### 2020-01-03 main.tex
Template

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Conference Paper Title*\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes,
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins,
column widths, line spaces, and text fonts are prescribed; please do not
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement
and others are deliberate, using specifications that anticipate your paper
as one part of the entire proceedings, and not as an independent document.
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a
separate text file. Complete all content and organizational editing before
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been
formatted and styled. Do not number text heads---{\LaTeX} will do that
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text,
even after they have been defined in the abstract. Abbreviations such as
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your
equations more compact, you may use the solidus (~/~), the exp function, or
appropriate exponents. Italicize Roman symbols for quantities and variables,
but not Greek symbols. Use a long dash rather than a hyphen for a minus
sign. Punctuate equations with commas or periods when they are part of a
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the
symbols in your equation have been defined before or immediately following
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A
minimum of one author is required for all conference articles. Author names
should be listed starting from left to right and then moving down to the
next line. This is the author sequence that will be used in future citations
and by indexing services. Names should not be listed in columns nor group by
affiliation. Please keep your affiliations as succinct as possible (for
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not
topically subordinate to each other. Examples include Acknowledgments and
References and, for these, the correct style to use is ``Heading 5''. Use
``figure caption'' for your Figure captions, and ``table head'' for your
table title. Run-in heads, such as ``Abstract'', will require you to apply a
style (in this case, italic) in addition to the style provided by the drop
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For
example, the paper title is the primary text head because all subsequent
material relates and elaborates on this one topic. If there are two or more
sub-topics, the next level head (uppercase Roman numerals) should be used
and, conversely, if there are not at least two sub-topics, then no subheads
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and
bottom of columns. Avoid placing them in the middle of columns. Large
figures and tables may span across both columns. Figure captions should be
below the figures; table heads should appear above the tables. Insert
figures and tables after they are cited in the text. Use the abbreviation
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4}
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
% \centerline{\includegraphics{fig1.png}}
foo
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2020-01-06 main.tex
Outline

@@
 \documentclass[conference]{IEEEtran}
+\newif\ifdraft \drafttrue
+\newif\iftext \draftfalse
+
 \IEEEoverridecommandlockouts
@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\newcommand{\bcp}[1]{\ifdraft\textcolor{violet}{{[BCP:~#1]}}\fi}
+
 \begin{document}
 
-\title{Conference Paper Title*\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
+\title{Interfering with Noninterference
+\iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF
+    and...?)}}\fi
 }
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
+\iftext
+\author{\IEEEauthorblockN{Roberto Blanco}
+\IEEEauthorblockA{\textit{\bcp{affiliation?}}
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID
+}
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
+\IEEEauthorblockN{Leonidas Lampropoulos}
+\IEEEauthorblockA{\textit{University of Maryland \bcp{right?}}
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID
+  }
 \and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
+\IEEEauthorblockN{Benjamin C. Pierce}
+\IEEEauthorblockA{\textit{University of Pennsylvania}
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID
+  }
 \and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
+\IEEEauthorblockN{Andrew Tolmach}
+\IEEEauthorblockA{\textit{Portland State University}
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID
 }
+}
+\fi
 
@@
 
+\iftext
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes,
-or Math in Paper Title or Abstract.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+% component, formatting, style, styling, insert
 \end{IEEEkeywords}
+\fi
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+\section{Preliminaries}
 
-\section{Ease of Use}
+\begin{itemize}
+\item (A sketch of) the RISC-V machine model and operational semantics.
+\item Reminder of the standard definition of NI.
 
-\subsection{Maintaining the Integrity of the Specifications}
+\bcp{Write it!}
 
-The IEEEtran class file is used to format your paper and style the text. All margins,
-column widths, line spaces, and text fonts are prescribed; please do not
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement
-and others are deliberate, using specifications that anticipate your paper
-as one part of the entire proceedings, and not as an independent document.
-Please do not revise any of the current designations.
+Designate sets of private and public addresses.
 
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a
-separate text file. Complete all content and organizational editing before
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on
-proofreading, spelling and grammar.
+Begin by defining what it means for two traces to witness interference?
 
-Keep your text and graphic files separate until after the text has been
-formatted and styled. Do not number text heads---{\LaTeX} will do that
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text,
-even after they have been defined in the abstract. Abbreviations such as
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
+\item Fact: The standard RISC-V semantics does not guarantee NI.  (Give a
+little example of a program that reads from a secret location and writes to
+a public one, just to illustrate the definition of NI.)
 \end{itemize}
 
-\subsection{Equations}
-Number equations consecutively. To make your
-equations more compact, you may use the solidus (~/~), the exp function, or
-appropriate exponents. Italicize Roman symbols for quantities and variables,
-but not Greek symbols. Use a long dash rather than a hyphen for a minus
-sign. Punctuate equations with commas or periods when they are part of a
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
+\section{Tainting is (a bit less than) noninterference}
 
-Be sure that the
-symbols in your equation have been defined before or immediately following
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+\begin{itemize}
+\item First intuition: A program is taint-NI if every pair of runs that don’t
+branch on high data are noninterfering in the standard sense.  (i.e., they
+don't write secret data to public locations and vice versa)
+\item Better alternative: … for every pair of runs that take the same control path (even if they branch on tainted data)...
+\item \bcp{We could even go a step farther: {\em truncate} the runs at the
+  first point where their PCs differ, then demand NI.}
+\end{itemize}
 
-\subsection{\LaTeX-Specific Advice}
+\section{Stack protection is (a bit more than) noninterference}
 
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
+Formal definition.
 
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
+\iftext
+\section{Stack + heap protection too}
 
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
+POTENTIALLY: Memory safety (before stack) + Conjunction of memory + stack
+safety (after).
 
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
+(Memory safety by itself is just arthur’s paper, but it’s cool to see how
+they are combined, especially what bits can be factored out as common
+structure, e.g. the notion of accessibility)
+\fi
 
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
+\section{Implementation}
 
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
+Sketch of the micropolicies that claim to implement these properties on
+PIPE...
 
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
+\subsection{A taint checking monitor}
 
-\subsection{Some Common Mistakes}\label{SCM}
+\subsection{A stack safety monitor}
+
+\section{Testing the implementations}
+
 \begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
+\item Experimental results of testing
 \end{itemize}
-An excellent style manual for science writers is \cite{b7}.
 
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A
-minimum of one author is required for all conference articles. Author names
-should be listed starting from left to right and then moving down to the
-next line. This is the author sequence that will be used in future citations
-and by indexing services. Names should not be listed in columns nor group by
-affiliation. Please keep your affiliations as succinct as possible (for
-example, do not differentiate among departments of the same organization).
+\section{Related work}
 
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through
-your paper. There are two types: component heads and text heads.
+\begin{itemize}
+\item Daniel's tainting paper (and some refs cited there)
+\item Nick and Andre's papers (and other micropolicy papers)
+\item StkTokens: Skorstengaard, Devriese, Birkedal
+\begin{itemize}
+\item {https://arxiv.org/pdf/1811.02787.pdf}
+\item {https://arxiv.org/pdf/1902.05283.pdf}
+\item {http://cs.au.dk/~lask/papers/esop2018.pdf}
+\end{itemize}
 
-Component heads identify the different components of your paper and are not
-topically subordinate to each other. Examples include Acknowledgments and
-References and, for these, the correct style to use is ``Heading 5''. Use
-``figure caption'' for your Figure captions, and ``table head'' for your
-table title. Run-in heads, such as ``Abstract'', will require you to apply a
-style (in this case, italic) in addition to the style provided by the drop
-down menu to differentiate the head from the text.
+\item Mechanized Reasoning about a Capability Machine: [Linn] Georges,
+Trieu, Birkedal (PriSC 2020).
+%
+Described at the Iris Workshop as formalizing the security properties of
+capabilities. This appears to involve a fairly complex and very large
+mechanization in Iris (also in the eyes of the Iris people).
+\end{itemize}
 
-Text heads organize the topics on a relational, hierarchical basis. For
-example, the paper title is the primary text head because all subsequent
-material relates and elaborates on this one topic. If there are two or more
-sub-topics, the next level head (uppercase Roman numerals) should be used
-and, conversely, if there are not at least two sub-topics, then no subheads
-should be introduced.
+\iftext
+\section{Future work}
 
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and
-bottom of columns. Avoid placing them in the middle of columns. Large
-figures and tables may span across both columns. Figure captions should be
-below the figures; table heads should appear above the tables. Insert
-figures and tables after they are cited in the text. Use the abbreviation
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
+\begin{itemize}
+\item Variations / extensions (OS, multiple processes, coroutines, generalizing depths to stacks of static call points, etc.)
+\item Extending the policy / property to more realistic ones
+\begin{itemize}
+\item Argument passing
+\item Variable-size stack frames
+\item Stack variables
+\item Relation to Nick and Andre’s policy
+\end{itemize}
+\item Could we also implement all this using Cheri capabilities?
+\end{itemize}
 
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4}
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
+More:
+\begin{itemize}
+\item More realistic calling conventions.
+(We need to eventually match this to the particular calling convention of whatever compiler we choose.)  E.g., the above will not quite work for passing arguments on the stack: callee won’t be able to read them…So perhaps need to support a special kind of write instruction that stores D.PCdepth+1 .
+[RB: Or implement this distinction through tags?]
+\item Combination with heap policy (colors, etc.)
+\item For coroutines:
+\begin{itemize}
+\item Generalize depth to depth + PID
+\item Private stuff not having changed as an idea should still work
+\item Maybe not so hard: The scheduler + other coroutines are just a sort of generalized “callee” as far as this process is concerned.
+Just need to fiddle th step relation so that it recognizes which process is getting returned to and updates its description appropriately.
+\end{itemize}
+\item If we had a heap policy, maybe we could talk about dynamic code generation, loading, as allocating a heap region, storing instructions into it, and then calling into them.
+For now, we just make all the instructions read only and assume they exist in the initial state.
+\item Question: how efficiently can consistency be tested?
+How much scrambling is needed, and does it really need to occur before every instruction?
+How expensive is it to check for agreement between the (in)accessible parts of two memories?
+Would spot checking for agreement be adequate?]
+\item What are the effects of compartmentalization on the ability of each compartment to change its own code/stack?
+\item Tail calls (maybe by not marking them as call sites?)
+\end{itemize}
 
-\begin{figure}[htbp]
-% \centerline{\includegraphics{fig1.png}}
-foo
-\caption{Example of a figure caption.}
-\label{fig}
-\end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
+\section*{Acknowledgments}
+\fi
 
 \section*{References}
+\bibliographystyle{IEEEabrv}
+\bibliography{bcp}
+\bibliography{local}
 
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
+\newpage
 
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
+\section*{Things to think about}
 
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
+\begin{itemize}
+\item
+How does Nick’s stack policy actually work these days?
+\item Andrew remembers that Catalin looked at something like this a long time ago and they decided that they needed unique tags for each dynamic call, not just levels.  We should check.
+\end{itemize}
 

### 2020-01-07 main.tex
Tweak
wip

@@
 
-Formal definition.
+\begin{itemize}
+\item
+  Idea of accessibility: set of data memory addresses that can be read or written at a particular execution point.
+  (Assume instruction memory is distinct.)
+\item
+  Stack protection intuition in a nutshell: caller's memory data is not accessible to callee, although callee's memory data is accessible to caller.
+  This gives both an integrity property (callee cannot affect caller's memory) and a secrecy property
+  (caller's memory cannot affect callee).
+  Can generalize to other control situations, such as coroutines, threads, compartments,...
+\item
+  In NI terms, protected programs are oblivious to variations in memory outside of their accessible set.
+  Unlike standard NI, the accessible set changes dynamically, so we cannot just frame the property in terms of
+  variant start states. [Is this true? Is it a significant difference? Are there other significant differences?]
+\item
+  We start with a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls:
+  \begin {itemize}
+  \item All memory (except for instructions) forms a single contiguous stack, divided into frames, each associated with a procedure activation.
+  \item Each frame consists of a saved return address and a single word of local data.
+These form the entirety of the procedure's memory data, and are the only thing we need to track when considering accessibility.
+  \item Registers other than RA, SP are not affected by calls or returns and hence are the (only) way to pass values between caller and callee.
+  \item Instructions serving as calls are statically marked as such in the code.
+  \item The code for each procedure is bracketed by fixed entry and exit sequences (details to follow), which save/restore the RA and
+    adjust the SP.  
+  \item During the body of the procedure, the single word of local data is accessed via SP, which does not change (except during a call, of course).
+The stored RA in the frame never changes.
+[I'm confused about these.
+Are they assumptions, or do they have a different status?
+We use tags to check them, but should we?]
+  \end {itemize}
+\item
+  Describe conjecture 2 property in terms of this model: informally and formally.
+\item 
+  Define abstract state descriptions.
+\item
+  Define step consistency.
+\item
+  Discuss relationship between state consistency and conjecture 2 property.  (Former is: presumably stronger; hopefully not
+  too strong that it produces lots of undesired test failures; surely easier to test).
+
+\end{itemize}
 

### 2020-01-07 main.tex
CP

@@
 \newif\ifdraft \drafttrue
-\newif\iftext \draftfalse
+\newif\iftext \textfalse
 
@@
 
-\section{Preliminaries}
+\section{Machine model}
+
+(A sketch of) the RISC-V machine model and operational semantics.
+
+Conventions:
+%
+\newcommand{\word}{W}
+\newcommand{\addr}{A}
+\newcommand{\WORDS}{{\mathcal W}}
+\newcommand{\reg}{R}
+\newcommand{\REGS}{{\mathcal R}}
+\newcommand{\mach}{M}
+\newcommand{\MACHS}{{\mathcal M}}
+\newcommand{\PC}[1]{\mathit{PC}(#1)}
+%
+\begin{itemize}
+\item Words / addresses: $\word,\addr \in \WORDS$
+\item Registers: $\reg \in \REGS$
+\item Machine states:
+$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
+\WORDS)$
+\item Abbreviations: \\
+$\mach(\addr) = \mach.1(\addr)$ \\
+$\mach(\reg) = \mach.2(\reg)$ \\
+$\mach(\textsc{pc}) = \mach.3$ \\
+\end{itemize}
+
+\section{Noninterference}
 
 \begin{itemize}
-\item (A sketch of) the RISC-V machine model and operational semantics.
+\item We use the word {\em policy} for any kind of dynamic enforcement
+mechanism (e.g., a runtime taint tracker, heap coloring rules, capability
+hardware, etc.). We use {\em property} for a mathematical predicate on
+programs (or machine states) capturing some useful security goal, e.g. noninterference.
+
 \item Reminder of the standard definition of NI.
@@
 
-Designate sets of private and public addresses.
+Designate sets of private, internal, and public addresses.
 
@@
 
-\section{Tainting is (a bit less than) noninterference}
+\section{Tainting is noninterference$-$}
 
@@
 \item \bcp{We could even go a step farther: {\em truncate} the runs at the
-  first point where their PCs differ, then demand NI.}
+  first point where their PCs differ, then demand NI.  This is a less
+  demanding property and probably a less desirable one, as it doesn't
+  predict the behavior of actual taint policies.  Discuss.}
+\item (Are there any other variations of tainting that we could present in
+this framework?)
 \end{itemize}
 
-\section{Stack protection is (a bit more than) noninterference}
+\section{Separation is noninterference$+$}
+
+\section{Stack protection is noninterference$++$}
 
@@
 \section*{References}
-\bibliographystyle{IEEEabrv}
-\bibliography{bcp}
-\bibliography{local}
+\bibliographystyle{IEEEtran}
+% \bibliography{bcp,local}
 

### 2020-01-07 main.tex
CP

@@
 \newcommand{\PC}[1]{\mathit{PC}(#1)}
+\newcommand{\pol}{P}
+\newcommand{\POLS}{{\mathcal P}}
+\newcommand{\trace}[1]{\overline{#1}}
+\newcommand{\TRACES}{{\mathcal T}}
+\newcommand{\stepsto}{\Rightarrow}
 %
@@
 \WORDS)$
-\item Abbreviations: \\
-$\mach(\addr) = \mach.1(\addr)$ \\
-$\mach(\reg) = \mach.2(\reg)$ \\
-$\mach(\textsc{pc}) = \mach.3$ \\
+\item Abbreviations:
+\\ $\mach(\addr) = \mach.1(\addr)$
+\\ $\mach(\reg) = \mach.2(\reg)$
+\\ $\mach(\textsc{pc}) = \mach.3$
+\item Step function\footnote{\bcp{I'm taking the step function to be total;
+    we could make it partial if we want, but I'm not sure this refinement is
+    important.}}: $\mathord{\stepsto} \in \MACHS \rightarrow \MACHS$
 \end{itemize}
+
+A {\em trace} $\trace{M} \in \TRACES$ is a nonempty, finite or infinite
+sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
+$\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.
+
+We use the word {\em policy} for any kind of dynamic enforcement
+mechanism (e.g., a runtime taint tracker, heap coloring rules, capability
+hardware, etc.). We use {\em property} for a mathematical predicate on
+programs (or machine states) capturing some useful security goal, e.g. noninterference.
+
+Formally, a {\em policy} $\pol \in \POLS$ is a prefix-closed set of traces.
+I.e., it is a restriction of the set of all traces generated by the standard
+step function, where some traces may be ``truncated early.''
+
+\iftext \bcp{Not sure we need names for $|TRACES$ or $\POLS$.} \fi
+
+\bcp{Now we need to define {\em property} formally.  But since we're talking
+about hyperproperties, I'm not certain how to go about it.  The standard way
+would be something like ``A policy $P$ has property $Q$ (i.e., $P$ is
+noninterfering) if, for every pair of traces $\trace{M}$ and $\trace{M'}$
+in $P \times P$, if the initial states $M_0$ and $M'_0$ differ only in
+high parts then every pair ...'' (blah: I'm not sure whether to talk about
+lockstep-corresponding states, which seems to restrictive, or final states,
+which doesn't work if the traces are infinite!  do we really need some kind
+of reactive noninterference here?? Urg.)}
 
@@
 \begin{itemize}
-\item We use the word {\em policy} for any kind of dynamic enforcement
-mechanism (e.g., a runtime taint tracker, heap coloring rules, capability
-hardware, etc.). We use {\em property} for a mathematical predicate on
-programs (or machine states) capturing some useful security goal, e.g. noninterference.
-
 \item Reminder of the standard definition of NI.
@@
 \section{Separation is noninterference$+$}
+
+Presentation idea: Start here with a super-simple separation policy, which
+just makes some part of the address space inaccessible, truncating any
+traces that read or write it.  Formulate the property as a flavor of NI.
+(The $+$ part in the title is a hint that one thing that gets added is that
+the PCs of corresponding traces always stay in lock step.)
 

### 2020-01-07 main.tex
wip

@@
 \item
-  Idea of accessibility: set of data memory addresses that can be read or written at a particular execution point.
+  Idea of accessibility: set of data memory addresses that can legally be read or written at a particular execution point.
   (Assume instruction memory is distinct.)
@@
   \begin {itemize}
-  \item All memory (except for instructions) forms a single contiguous stack, divided into frames, each associated with a procedure activation.
+  \item Memory is divided into an instruction region and a data region.
+The data region contains a single contiguous stack, divided into frames, each associated with a procedure activation.
+(The stack grows in the direction of increasing addresses.)
   \item Each frame consists of a saved return address and a single word of local data.
-These form the entirety of the procedure's memory data, and are the only thing we need to track when considering accessibility.
+These [both??] form the entirety of the procedure's memory data, and are the only thing we need to track when considering accessibility.
   \item Registers other than RA, SP are not affected by calls or returns and hence are the (only) way to pass values between caller and callee.
-  \item Instructions serving as calls are statically marked as such in the code.
+  \item Instructions serving as calls are statically marked as such in the code. [The entry and exit sequences must also be marked???]
   \item The code for each procedure is bracketed by fixed entry and exit sequences (details to follow), which save/restore the RA and
-    adjust the SP.
+    adjust the SP. [Are these assumptions or checked facts?] 
   \item During the body of the procedure, the single word of local data is accessed via SP, which does not change (except during a call, of course).
@@
 [I'm confused about these.
-Are they assumptions, or do they have a different status?
-We use tags to check them, but should we?]
+They are checked facts, not assumptions.]
   \end {itemize}
+  In this model, the definition of accessible memory is simply all (data) memory at addresses $\geq pc$. 
 \item

### 2020-01-07 main.tex
CP

--- 2020-01-07 main.tex
@@
 \usepackage{textcomp}
+\usepackage{cleveref}
 \usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \newcommand{\bcp}[1]{\ifdraft\textcolor{violet}{{[BCP:~#1]}}\fi}
+
+% Notational conventions
+\newcommand{\HIGH}{\textsc{h}}
+\newcommand{\LOW}{\textsc{l}}
+\newcommand{\word}{W}
+\newcommand{\addr}{A}
+\newcommand{\WORDS}{{\mathcal W}}
+\newcommand{\reg}{R}
+\newcommand{\REGS}{{\mathcal R}}
+\newcommand{\mach}{M}
+\newcommand{\MACHS}{{\mathcal M}}
+\newcommand{\PC}[1]{\PCname(#1)}
+\newcommand{\PCname}{\textsc{pc}}
+\newcommand{\polname}{P}
+\newcommand{\pol}[1]{\polname{#1}}
+\newcommand{\trace}[1]{\overline{#1}}
+\newcommand{\topo}{T}
+\newcommand{\stepsto}{\Rightarrow}
+\newcommand{\manystepsto}{\stepsto^\star}
 
@@
 
-\section{Machine model}
+\section{Preliminaries}
 
-(A sketch of) the RISC-V machine model and operational semantics.
+\subsection{Machine model}
+
+Concretely, we have in mind a conventional machine architecture with a
+RISC-V processor, perhaps extended with some kind of hardware-enforced
+security policy, but we won't need any RISC-V-specific details till we get
+to \cref{impl}.
 
 Conventions:
-%
-\newcommand{\word}{W}
-\newcommand{\addr}{A}
-\newcommand{\WORDS}{{\mathcal W}}
-\newcommand{\reg}{R}
-\newcommand{\REGS}{{\mathcal R}}
-\newcommand{\mach}{M}
-\newcommand{\MACHS}{{\mathcal M}}
-\newcommand{\PC}[1]{\mathit{PC}(#1)}
 %
@@
 \WORDS)$
-\item Abbreviations: \\
-$\mach(\addr) = \mach.1(\addr)$ \\
-$\mach(\reg) = \mach.2(\reg)$ \\
-$\mach(\textsc{pc}) = \mach.3$ \\
+\item Abbreviations:
+\\ $\mach(\addr) = \mach.1(\addr)$
+\\ $\mach(\reg) = \mach.2(\reg)$
+\\ $\mach(\textsc{pc}) = \mach.3$
+\item Step function\footnote{\bcp{I'm taking the step function to be total;
+    we could make it partial if we want, but I'm not sure this refinement is
+    important.}}: $\mathord{\stepsto} \in \MACHS \rightarrow \MACHS$
 \end{itemize}
 
-\section{Noninterference}
+A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
+sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
+$\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.
+
+% \bcp{Now we need to define {\em property} formally.  But since we're talking
+% about hyperproperties, I'm not certain how to go about it.  The standard way
+% would be something like ``A policy $P$ has property $Q$ (i.e., $P$ is
+% noninterfering) if, for every pair of traces $\trace{M}$ and $\trace{M'}$
+% in $P \times P$, if the initial states $M_0$ and $M'_0$ differ only in
+% high parts then every pair ...'' (blah: I'm not sure whether to talk about
+% lockstep-corresponding states, which seems too restrictive, or final states,
+% which doesn't work if the traces are infinite!  do we really need some kind
+% of reactive noninterference here?? Urg.)  (Well, no, we can just do
+% a flavor of termination-insensitive NI, which says that all infinite traces
+% are indistinguishable from anything...?)}
+
+% \bcp{Should we start from a notion of ``variant traces'' and then talk about
+% pairs of variant traces being ``distinguishable'' and then go to the full
+% thing?}
+
+The {\em result} of a machine state $\mach$ (if it exists) is a state
+$\mach_n$ such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step
+any further.
+
+\subsection{Noninterference}
+
+Intuitively, we use the word {\em policy} for any kind of dynamic
+enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
+capability hardware, etc.). We use {\em property} for a mathematical
+predicate on programs (i.e., machine states) articulating some useful
+security goal, e.g. noninterference.
+
+A {\em topography}\bcp{??}
+$\topo \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
+map from machine components to ``security levels,'' where $\HIGH$ marks
+locations containing secret or low-integrity data and $\LOW$ marks those
+with public or high-integrity data.
+
+Given a machine state $\mach$ and a topography $\topo$, a {\em variant}
+of $\mach$ is any machine state $\mach'$ whose components agree with $\mach$
+on all components marked $\LOW$.  \iftext\bcp{Spell it out formally.}\fi
+
+Formally, a {\em policy} $\polname$\iftext\bcp{show its type? then we need
+  some more metavariables... :-(}\fi is a function from topographies to
+prefix-closed set of traces.  I.e., given a topgraphy, it yields a
+restriction of the set of all traces generated by the standard step
+function, where some traces may be ``truncated early'' because they violate
+the conditions imposed by this topography.
+
+\bcp{Needs to be wrt a policy!!}
+Variant machine states $\mach$ and $\mach'$ are said to be {\em
+  indistinguishable} with respect to topography $\topo$ and policy
+$\pol{\topo}$ if either (1)
+one or both of $\mach$ and $\mach'$ fails to terminate, or (2) both
+terminate and their results are variants with respect to $\topo$.
+
+A policy $P$
 
 \begin{itemize}
-\item We use the word {\em policy} for any kind of dynamic enforcement
-mechanism (e.g., a runtime taint tracker, heap coloring rules, capability
-hardware, etc.). We use {\em property} for a mathematical predicate on
-programs (or machine states) capturing some useful security goal, e.g. noninterference.
+\item Our definition.
 
-\item Reminder of the standard definition of NI.
+\item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
+  have to begin with EENI and then get to SSNI as a (stronger) refinement.}
 
-\bcp{Write it!}
-
-Designate sets of private, internal, and public addresses.
-
-Begin by defining what it means for two traces to witness interference?
+\item Relation to the standard definition of NI.
 
@@
 
-\section{Tainting is noninterference$-$}
+\section{Tainting is noninterference($-$)}
 
@@
 
-\section{Separation is noninterference$+$}
+\section{Separation is noninterference($+$)}
 
-\section{Stack protection is noninterference$++$}
+Presentation idea: Start here with a super-simple separation policy, which
+just makes some part of the address space inaccessible, truncating any
+traces that read or write it.  Formulate the property as a flavor of NI.
+(The $+$ part in the title is a hint that one thing that gets added is that
+the PCs of corresponding traces always stay in lock step.)
+
+\section{Stack protection is noninterference($++$)}
 
@@
 \section{Implementation}
+\label{impl}
 
@@
 \item {http://cs.au.dk/~lask/papers/esop2018.pdf}
+\item {https://dl.acm.org/doi/fullHtml/10.1145/3363519}
 \end{itemize}

### 2020-01-07 main.tex
Strong noninterference ref from chalmers/PLAS

@@
 \iftext
-\author{\IEEEauthorblockN{Roberto Blanco}
-\IEEEauthorblockA{\textit{\bcp{affiliation?}}
+\IEEEauthorblockN{Leonidas Lampropoulos}
+\IEEEauthorblockA{\textit{University of Maryland \bcp{right?}}
 % \textit{name of organization (of Aff.)}\\
@@
 \and
-\IEEEauthorblockN{Leonidas Lampropoulos}
-\IEEEauthorblockA{\textit{University of Maryland \bcp{right?}}
+\author{\IEEEauthorblockN{Roberto Blanco}
+\IEEEauthorblockA{\textit{\bcp{affiliation?}}
 % \textit{name of organization (of Aff.)}\\
@@
 
-A {\em topography}\bcp{??}
+A {\em topography}\bcp{contour!}
 $\topo \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
@@
 
+\item Simple Noninterference by Normalization {http://nachivpn.me/nibnbe.pdf}
+
 \item Mechanized Reasoning about a Capability Machine: [Linn] Georges,

### 2020-01-07 main.tex
CP
First complete draft of section I

@@
 \newcommand{\polname}{P}
-\newcommand{\pol}[1]{\polname{#1}}
-\newcommand{\trace}[1]{\overline{#1}}
-\newcommand{\topo}{T}
-\newcommand{\stepsto}{\Rightarrow}
+\newcommand{\pol}[1]{\polname(#1)}
+% \newcommand{\trace}[1]{\overline{#1}}
+\newcommand{\contour}{C}
+\newcommand{\CONTOURS}{{\mathcal C}}
+\newcommand{\stepsto}{\Longrightarrow}
+\newcommand{\stepstounder}[1]{\stackrel{\mbox{\tiny{$#1$}}}{\Longrightarrow}}
+\newcommand{\stepstounderfull}{\stepstounder{\textsc{full}}}
 \newcommand{\manystepsto}{\stepsto^\star}
@@
 \\ $\mach(\textsc{pc}) = \mach.3$
-\item Step function\footnote{\bcp{I'm taking the step function to be total;
-    we could make it partial if we want, but I'm not sure this refinement is
-    important.}}: $\mathord{\stepsto} \in \MACHS \rightarrow \MACHS$
 \end{itemize}
-
-A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
-sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
-$\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.
 
@@
 
-The {\em result} of a machine state $\mach$ (if it exists) is a state
-$\mach_n$ such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step
-any further.
+The {\em result} of a machine state $\mach$ with respect to a step function
+$\mathord{\stepsto} \in \MACHS \rightharpoonup \MACHS$ is a state $\mach_n$
+such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step any
+further.
+
+The {\em full} step function, $\mathord{\stepstounderfull} \in \MACHS
+\rightharpoonup \MACHS$, embodies the standard RISC-V ISA semantics.
+
+% \bcp{Needed?}A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
+% sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
+% $\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.
 
@@
 
-A {\em topography}\bcp{contour!}
-$\topo \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
+A {\em contour}
+$\contour \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
 map from machine components to ``security levels,'' where $\HIGH$ marks
@@
 
-Given a machine state $\mach$ and a topography $\topo$, a {\em variant}
-of $\mach$ is any machine state $\mach'$ whose components agree with $\mach$
-on all components marked $\LOW$.  \iftext\bcp{Spell it out formally.}\fi
+Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
+$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
+components agree with $\mach$ on all components marked $\LOW$ by $\contour$.
+\iftext\bcp{Spell it out formally.}\fi
 
-Formally, a {\em policy} $\polname$\iftext\bcp{show its type? then we need
-  some more metavariables... :-(}\fi is a function from topographies to
-prefix-closed set of traces.  I.e., given a topgraphy, it yields a
-restriction of the set of all traces generated by the standard step
-function, where some traces may be ``truncated early'' because they violate
-the conditions imposed by this topography.
+Formally, a {\em policy}
+$\polname \in \CONTOURS \rightarrow (\MACHS \rightarrow \MACHS)$ is a
+function from contours to step functions (partial functions on $\MACHS$)
+that, wherever they are defined, agree with $\stepstounderfull$.  That is,
+it represents an enforcement mechanism that sometimes halts programs earlier
+than the bare step relation does, but that agrees with the bare step
+relation when it allows evaluation to proceed.  We write
+$\mach_1 \stepstounder{\pol{\contour}} \mach_2$ for
+$\pol{\contour}(\mach_1) = \mach_2$.
 
-\bcp{Needs to be wrt a policy!!}
-Variant machine states $\mach$ and $\mach'$ are said to be {\em
-  indistinguishable} with respect to topography $\topo$ and policy
-$\pol{\topo}$ if either (1)
-one or both of $\mach$ and $\mach'$ fails to terminate, or (2) both
-terminate and their results are variants with respect to $\topo$.
+If $\mach$ and $\mach'$ are variant machine states with respect to contour
+$\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
+under policy $\polname$ if either (1) one or both of $\mach$ and $\mach'$
+fail to terminate when executed with $\stepstounder{\pol{\contour}}$, or (2)
+both terminate and their results are variants with respect to $\contour$.
 
-A policy $P$
+Policy $\polname$ {\em guarantees noninterference} if, for every $\contour$,
+every pair $(\mach,\mach')$ of variant machine states with respect to
+$\contour$ are indistinguishable under $\polname$.
 
+\iftext
 \begin{itemize}
-\item Our definition.
-
 \item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
@@
 \end{itemize}
+\fi
 

### 2020-01-07 main.tex
Nit
Test
Thoughts

@@
 \item Registers: $\reg \in \REGS$
-\item Machine states:
+\item Machine states: \bcp{Simplify: define {\em machine components} and
+  then a machine state is a function from these to words.}
 $\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
@@
 \begin{itemize}
+\item Should we have an intermediate classification level (``internal'' or
+something) and make the policy say that it's only bad for data to flow from
+$\HIGH$ to $\LOW$?  This would correspond to more realistic taint policies.
+
 \item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
@@
 
-\section{Tainting is noninterference($-$)}
+\section{Tainting is noninterference$-$}
+
+\bcp{I'm having some trouble here.  Two main problems so far:
+  \begin{itemize}
+  \item Since the ``high part'' of the world keeps getting extended as
+  tainted data gets written here and there, the only way for something
+  ``bad'' to happen is for high/tainted data to get written to someplace
+  that was designated as an ``output''.  We don't currently have any way of
+  making such a designation.
+  \item Even after adding a separate IFC-like label for ``internal'' cells
+  (that are not scrambled but also not outputs), I'm not seeing clearly how
+  to formulate taint checking as an NI-like property.  Need to think about
+  it some more.
+  \end{itemize}
+}
 
@@
 \item Better alternative: … for every pair of runs that take the same control path (even if they branch on tainted data)...
-\item \bcp{We could even go a step farther: {\em truncate} the runs at the
+\item We could even go a step farther: {\em truncate} the runs at the
   first point where their PCs differ, then demand NI.  This is a less
   demanding property and probably a less desirable one, as it doesn't
-  predict the behavior of actual taint policies.  Discuss.}
+  predict the behavior of actual taint policies.  Discuss.
+\item Note that this is a super-simple property that doesn't really track
+taint at all.  
 \item (Are there any other variations of tainting that we could present in
@@
 
-\section{Separation is noninterference($+$)}
+\section{Separation is noninterference$+$}
 
@@
 
-\section{Stack protection is noninterference($++$)}
+\section{Stack protection is noninterference$++$}
 

### 2020-01-07 main.tex
Another thought

@@
   making such a designation.
-  \item Even after adding a separate IFC-like label for ``internal'' cells
-  (that are not scrambled but also not outputs), I'm not seeing clearly how
-  to formulate taint checking as an NI-like property.  Need to think about
-  it some more.
+
+  I guess we can add a ``middle'' classification level, and doing that here
+  seems to make pretty good sense, but it's hard to justify adding it to the
+  general definition of NI.  :-(
   \end{itemize}

### 2020-01-07 main.tex
Test
Comment

@@
 
-\bcp{I'm having some trouble here.  Two main problems so far:
-  \begin{itemize}
-  \item Since the ``high part'' of the world keeps getting extended as
+\bcp{I'm having some trouble here.
+%
+  Since the ``high part'' of the world keeps getting extended as
   tainted data gets written here and there, the only way for something
@@
   making such a designation.
-
+%
   I guess we can add a ``middle'' classification level, and doing that here
@@
   general definition of NI.  :-(
-  \end{itemize}
 }

### 2020-01-08 main.tex
wip

@@
 A {\em contour}
-$\contour \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
+$\contour \in \CONTOURS = (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
 map from machine components to ``security levels,'' where $\HIGH$ marks
@@
   variant start states. [Is this true? Is it a significant difference? Are there other significant differences?]
-\item
-  We start with a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls:
+\end{itemize}
+
+We start with a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls.
+We assume that memory is divided into two disjoint regions, one containing the code, the other the stack.
+Code begins at address 0; the stack begins at address $B$ and grows in the direction of increasing addresses.
+
+The part of the stack between $B$ and the stack pointer \textsc{sp} is structured into a sequence of consecutive
+\emph{frames}. Each frame consists of a single return address and a single word of local data.
+
   \begin {itemize}
@@
   In this model, the definition of accessible memory is simply all (data) memory at addresses $\geq pc$.
+\begin{itemize}
 \item

### 2020-01-09 main.tex
Slight simplication

@@
 \newcommand{\CONTOURS}{{\mathcal C}}
+\newcommand{\component}{K}
+\newcommand{\COMPONENTS}{{\mathcal K}}
 \newcommand{\stepsto}{\Longrightarrow}
@@
 \item Registers: $\reg \in \REGS$
-\item Machine states: \bcp{Simplify: define {\em machine components} and
-  then a machine state is a function from these to words.}
-$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
-\WORDS)$
-\item Abbreviations:
-\\ $\mach(\addr) = \mach.1(\addr)$
-\\ $\mach(\reg) = \mach.2(\reg)$
-\\ $\mach(\textsc{pc}) = \mach.3$
+\item Machine components: $\component \in \WORDS \cup \REGS \cup \{
+\textsc{pc} \}$
+\item Machine states: $\mach \in \MACHS = \COMPONENTS\rightarrow\WORDS$
 \end{itemize}
@@
 A {\em contour}
-$\contour \in \CONTOURS = (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
+$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}$ is a
 map from machine components to ``security levels,'' where $\HIGH$ marks

### 2020-01-10 main.tex
Notes

@@
 
+\iftext
+\section{Introduction}
+
+\bcp{How about coining the term ``lockstep noninterference'' and then saying
+that we study several variants...?}
+
+\fi
+
 \section{Preliminaries}
@@
 such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step any
-further.
+further.  \bcp{No halting!}
 
@@
 
-\subsection{Noninterference}
+\subsection{Policies}
+
+\bcp{Try this:
+  \begin{itemize}
+  \item two separate axes: high/low and observable/hidden
+  \item both part of contours
+  \end{itemize}
+}
 
@@
 $\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
-components agree with $\mach$ on all components marked $\LOW$ by $\contour$.
-\iftext\bcp{Spell it out formally.}\fi
+components agree with $\mach$ on all components marked $\LOW$ by
+$\contour$---i.e., if $\contour(\component) = \LOW$, then $\mach(\component)
+= \mach'(\component)$.
 
@@
 $\mach_1 \stepstounder{\pol{\contour}} \mach_2$ for
-$\pol{\contour}(\mach_1) = \mach_2$.
+$\pol{\contour}(\mach_1) = \mach_2$.  \bcp{It should also return a contour!}
+
+\subsection{Noninterference}
 

### 2020-01-10 main.tex
Updates based on yesterday's conversation
and some more questions!

--- 2020-01-09 main.tex
@@
 \newcommand{\LOW}{\textsc{l}}
+\newcommand{\OBS}{\textsc{obs}}
+\newcommand{\HIDDEN}{\textsc{hidden}}
 \newcommand{\word}{W}
@@
 \newcommand{\PCname}{\textsc{pc}}
-\newcommand{\polname}{P}
-\newcommand{\pol}[1]{\polname(#1)}
-% \newcommand{\trace}[1]{\overline{#1}}
+\newcommand{\pol}{P}
 \newcommand{\contour}{C}
@@
 \newcommand{\COMPONENTS}{{\mathcal K}}
+
 \newcommand{\stepsto}{\Longrightarrow}
 \newcommand{\stepstounder}[1]{\stackrel{\mbox{\tiny{$#1$}}}{\Longrightarrow}}
-\newcommand{\stepstounderfull}{\stepstounder{\textsc{full}}}
+\newcommand{\stepstounderfull}{\stepstounder{\textsc{RISCV}}}
 \newcommand{\manystepsto}{\stepsto^\star}
+\newcommand{\obstrace}{\mathit{obstrace}}
+
+\newcommand{\underscore}{\mbox{\_}}
 
@@
 
-The {\em result} of a machine state $\mach$ with respect to a step function
-$\mathord{\stepsto} \in \MACHS \rightharpoonup \MACHS$ is a state $\mach_n$
-such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step any
-further.
-
-The {\em full} step function, $\mathord{\stepstounderfull} \in \MACHS
+The {\em underlying} step function, $\mathord{\stepstounderfull} \in \MACHS
 \rightharpoonup \MACHS$, embodies the standard RISC-V ISA semantics.
@@
 
-\subsection{Noninterference}
+\subsection{Policies}
 
@@
 A {\em contour}
-$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}$ is a
-map from machine components to ``security levels,'' where $\HIGH$ marks
-locations containing secret or low-integrity data and $\LOW$ marks those
-with public or high-integrity data.
+$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}\times
+\{\OBS,\HIDDEN\}$ is a map from machine components to pairs of ``security
+levels,'' where $\HIGH$ marks locations containing secret or low-integrity
+data and $\LOW$ marks those with public or high-integrity data, and
+``observabilities,'' where $\OBS$ marks locations whose contents are
+observable from the outside (by the ``attacker'') and $\HIDDEN$ marks
+internal ``hidden'' state that cannot be observed directly.
 
@@
 $\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
-components agree with $\mach$ on all components marked $\LOW$ by $\contour$.
-\iftext\bcp{Spell it out formally.}\fi
+components agree with $\mach$ on all components marked $\LOW$ by
+$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
+$\mach(\component) = \mach'(\component)$.
 
 Formally, a {\em policy}
-$\polname \in \CONTOURS \rightarrow (\MACHS \rightarrow \MACHS)$ is a
-function from contours to step functions (partial functions on $\MACHS$)
-that, wherever they are defined, agree with $\stepstounderfull$.  That is,
+$\pol \in (\CONTOURS \times \MACHS) \rightarrow (\CONTOURS \times
+\MACHS)$ is an enriched
+step function that works on pairs of contours and machines---i.e., it is a
+partial function taking a contour and a machine state to an updated contour
+and machine state that, wherever is is defined, agrees with
+$\stepstounderfull$.  That is,
 it represents an enforcement mechanism that sometimes halts programs earlier
@@
 relation when it allows evaluation to proceed.  We write
-$\mach_1 \stepstounder{\pol{\contour}} \mach_2$ for
-$\pol{\contour}(\mach_1) = \mach_2$.
+$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
+$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
+
+The {\em observation trace} arising from an infinite sequence of contours
+and machine states $(\contour_0,mach_0),\, (\contour_1,\mach_1),\ ldots$,
+written $\obstrace((\contour_0,mach_0),\, (\contour_1,\mach_1),\ ldots)$
+ is
+the (finite or infinite) sequence of {\em distinct} values taken on by the
+{\em observable} components of each state.  \iftext\bcp{Give an example and
+  write it out formally.}\fi \bcp{Question: This could be defined in two
+  ways, and I'm not sure which is better or if it really matters: the
+  observation trace is a sequence of partial maps from components to their
+  values, or the observation trace is a partial maps from components to
+  their sequences of observable values.  This starts getting a bit
+  technically tangled, but I don't see a simpler alternative right now.}
+
+\subsection{Noninterference}
+
+\bcp{Proposal: sequences of values of the observable locations are what gets
+  compared.  Fix the following:}
 
@@
 $\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
-under policy $\polname$ if either (1) one or both of $\mach$ and $\mach'$
-fail to terminate when executed with $\stepstounder{\pol{\contour}}$, or (2)
-both terminate and their results are variants with respect to $\contour$.
+under policy $\pol$ if either (1) one or both of $(\contour,\mach)$ and
+$(\contour,\mach')$ fail to terminate when executed with
+$\stepstounder{\pol}$, or (2) both terminate and the machine-state
+components of their results are variants with respect to the original
+$\contour$.
 
-Policy $\polname$ {\em guarantees noninterference} if, for every $\contour$,
+Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
 every pair $(\mach,\mach')$ of variant machine states with respect to
-$\contour$ are indistinguishable under $\polname$.
+$\contour$ are indistinguishable under $\pol$.
 
@@
 
-\bcp{I'm having some trouble here.
-%
-  Since the ``high part'' of the world keeps getting extended as
-  tainted data gets written here and there, the only way for something
-  ``bad'' to happen is for high/tainted data to get written to someplace
-  that was designated as an ``output''.  We don't currently have any way of
-  making such a designation.
-%
-  I guess we can add a ``middle'' classification level, and doing that here
-  seems to make pretty good sense, but it's hard to justify adding it to the
-  general definition of NI.  :-(
-}
+\bcp{This needs to be written, now that we have added the hidden/observable stuff.}
 

### 2020-01-10 main.tex
typo

@@
 The {\em observation trace} arising from an infinite sequence of contours
-and machine states $(\contour_0,mach_0),\, (\contour_1,\mach_1),\ ldots$,
-written $\obstrace((\contour_0,mach_0),\, (\contour_1,\mach_1),\ ldots)$
+and machine states $(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$,
+written $\obstrace((\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots)$
  is

### 2020-01-10 main.tex
Fixes

@@
 The {\em underlying} step function, $\mathord{\stepstounderfull} \in \MACHS
-\rightharpoonup \MACHS$, embodies the standard RISC-V ISA semantics.
+\rightarrow \MACHS$, embodies the standard RISC-V ISA semantics.
 
@@
   their sequences of observable values.  This starts getting a bit
-  technically tangled, but I don't see a simpler alternative right now.}
+  technically tangled, but I don't see a simpler alternative right now.
+  Answer: map from components to sequences of words.}
 

### 2020-01-10 main.tex
Test
Macros file, plus some progress on NI definition

@@
 \documentclass[conference]{IEEEtran}
-\newif\ifdraft \drafttrue
-\newif\iftext \textfalse
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{cleveref}
-\usepackage{xcolor}
-\newcommand{\bcp}[1]{\ifdraft\textcolor{violet}{{[BCP:~#1]}}\fi}
-
-% Notational conventions
-\newcommand{\HIGH}{\textsc{h}}
-\newcommand{\LOW}{\textsc{l}}
-\newcommand{\OBS}{\textsc{obs}}
-\newcommand{\HIDDEN}{\textsc{hidden}}
-\newcommand{\word}{W}
-\newcommand{\addr}{A}
-\newcommand{\WORDS}{{\mathcal W}}
-\newcommand{\reg}{R}
-\newcommand{\REGS}{{\mathcal R}}
-\newcommand{\mach}{M}
-\newcommand{\MACHS}{{\mathcal M}}
-\newcommand{\PC}[1]{\PCname(#1)}
-\newcommand{\PCname}{\textsc{pc}}
-\newcommand{\pol}{P}
-\newcommand{\contour}{C}
-\newcommand{\CONTOURS}{{\mathcal C}}
-\newcommand{\component}{K}
-\newcommand{\COMPONENTS}{{\mathcal K}}
-
-\newcommand{\stepsto}{\Longrightarrow}
-\newcommand{\stepstounder}[1]{\stackrel{\mbox{\tiny{$#1$}}}{\Longrightarrow}}
-\newcommand{\stepstounderfull}{\stepstounder{\textsc{RISCV}}}
-\newcommand{\manystepsto}{\stepsto^\star}
-\newcommand{\obstrace}{\mathit{obstrace}}
-
-\newcommand{\underscore}{\mbox{\_}}
+\input{macros}
 
@@
 
-The {\em observation trace} arising from an infinite sequence of contours
-and machine states $(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$,
-written $\obstrace((\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots)$
- is
-the (finite or infinite) sequence of {\em distinct} values taken on by the
-{\em observable} components of each state.  \iftext\bcp{Give an example and
-  write it out formally.}\fi \bcp{Question: This could be defined in two
-  ways, and I'm not sure which is better or if it really matters: the
-  observation trace is a sequence of partial maps from components to their
-  values, or the observation trace is a partial maps from components to
-  their sequences of observable values.  This starts getting a bit
-  technically tangled, but I don't see a simpler alternative right now.
-  Answer: map from components to sequences of words.}
+The {\em observation trace} arising from a finite or infinite sequence of
+contours and machine states
+$(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$, written
+$\obstrace((\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots)$ is a
+partial map taking each machine component to the (finite or infinite)
+sequence of {\em distinct} values taken on by this component at moments when
+it is {\em observable}.  Formally, this map can be constructed as
+follows. For each component $\component$,
+%
+(1) if
+$\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
+$\obstrace(\component) = \bot$;
+%
+and otherwise (2), $\obstrace(\component)$ is a sequence whose first element
+is $\mach_i(\component)$ for the smallest $i$ such that
+$\contour_i(\component) = (\underscore,\OBS)$ and whose $(m+1)^{th}$ element
+(if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
+than the index $i$ of the $m^{th}$ element and such that
+$\contour_j(\component) = (\underscore,\OBS)$ and
+$\mach_j(\component) \neq \mach_i(\component)$.
+%
+\iftext
+\bcp{Unfinished and horrible...  But maybe the textual description is OK?}
+\newcommand{\aux}{\mathit{aux}}
+\bcp{Don't like the ``rest'' argument and the fact that we don't
+  have a metavariable for sequences of pairs of contours and machines.
+  Consider defining one.}
+%
+\[
+\begin{array}{lcl}
+\aux \; ( [\,] ) \; \trace& = & \trace
+\\
+\aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
+  & = &
+        \aux \;( \mathit{rest} )
+        \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
+\\
+\\
+\obstrace \;( (\contour_0,\mach_0),\, \ldots )
+  &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
+      \component.\, \bot)
+\end{array}
+\]
+\fi
 
@@
 
-\bcp{Proposal: sequences of values of the observable locations are what gets
-  compared.  Fix the following:}
+\bcp{I'm confused here about policy vs. property -- e.g., which contours are
+  requirements and which are chosen by the micropolicy.}
 
@@
 $\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
-under policy $\pol$ if either (1) one or both of $(\contour,\mach)$ and
-$(\contour,\mach')$ fail to terminate when executed with
-$\stepstounder{\pol}$, or (2) both terminate and the machine-state
-components of their results are variants with respect to the original
-$\contour$.
+under policy $\pol$ if their observable
+%
+$\obstrace(
+                      (\contour,\mach)
+  \stepstounder{\pol} (\contour_1,\mach_1)
+  \stepstounder{\pol} (\contour_2,\mach_2)
+  \stepstounder{\pol} \ldots)$
+%
+yields the same trace as
+$\obstrace(
+                      (\contour,\mach')
+  \stepstounder{\pol} (\contour_1,\mach_1')
+  \stepstounder{\pol} (\contour_2,\mach_2')
+  \stepstounder{\pol} \ldots)$.
+
+% either (1) one or both of $(\contour,\mach)$ and
+% $(\contour,\mach')$ fail to terminate when executed with
+% $\stepstounder{\pol}$, or (2) both terminate and the machine-state
+% components of their results are variants with respect to the original
+% $\contour$.
 

### 2020-01-10 main.tex
Notes.

@@
 
+\bcp{
+  Following is not quite right.  Better intuition:
+  \begin{itemize}
+  \item A (micro-){\em policy} is a function that takes a contour to
+  \begin{enumerate}
+  \item a private type $X$
+  \item an initial value of $X$
+  \item a partial step function of type $(\MACHS,X) \rightharpoonup 
+  (\MACHS,X)$ that, when defined, agrees with $\stepstounderfull$ on
+  machine states.
+  \end{enumerate}
+  That is, it takes a starting contour and produces a runtime monitor, which
+  carries along some information that it uses to decide when to terminate
+  execution early.
+  \item A {\em property} is a function that takes a contour to a
+  prefix-closed set of (finite and/or infinite) sequences of machine states.
+  That is, it takes a starting contour and produces a checker that tells
+  whether a given trace is OK to observe.
+  \item A policy $p$ {\em guarantees} a property $P$ if, for each contour
+  $\contour$, the sequence of machine states obtained by projecting the
+  sequence of ..... ARGH WHAT ABOUT VARIANTS!?!
+  \end{itemize}
+}
+
 Formally, a {\em policy}
@@
 
-\bcp{I'm confused here about policy vs. property -- e.g., which contours are
-  requirements and which are chosen by the micropolicy.}
-
 If $\mach$ and $\mach'$ are variant machine states with respect to contour
@@
 \begin{itemize}
-\item Should we have an intermediate classification level (``internal'' or
-something) and make the policy say that it's only bad for data to flow from
-$\HIGH$ to $\LOW$?  This would correspond to more realistic taint policies.
-
 \item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we

### 2020-01-12 main.tex
Checkpoint before messing with definitions

@@
   sequence of ..... ARGH WHAT ABOUT VARIANTS!?!
+  \item Proposal: Forget about full-blown NI and make the whole paper be
+  about ``lockstep NI'' where the PC is the same in every pair of
+  related runs
+  \begin{itemize}
+  \item One issue that I am not sure about: Could there be a sneaky way of
+  ``putting secret information into the control state'' where the PCs of
+  related runs are always in lockstep but, because the instruction memory is
+  writeable, different instructions can get dispatched.  E.g., XXXX
+  \item So maybe we should be asking that the instruction stream be
+  identical, not (just?) the PCs
+  \item One nice observation is that, if we are only interested in pairs of
+  runs with identical PCs, there are two different ways to proceed: (1) say
+  that we {\em ignore} runs where the PCs (or instruction streams) diverge
+  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
+  instruction streams) diverge, i.e., demand that this never happens (e.g.,
+  stack and heap policies).
+  \end{itemize}
+  \item A remaining question is what to do about the possibility that the
+  contour changes differently on two related runs.  We could disallow this,
+  but not certain it's necessary---maybe everything works out OK.
   \end{itemize}

### 2020-01-12 main.tex
Thoughts

@@
 
+Informally, a (micro-){\em policy}
+is an enriched
+step function that works on pairs of contours and machines---i.e., it is a
+partial function taking a contour and a machine state to an updated contour
+and machine state that, wherever is is defined, agrees with
+$\stepstounderfull$.  That is,
+it represents an enforcement mechanism that sometimes halts programs earlier
+than the bare step relation does, but that agrees with the bare step
+relation when it allows evaluation to proceed.  We write
+$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
+$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
+
+Formally, a policy
+%
+$\pol \in \CONTOURS \rightarrow \exists X.\; X \times (\MACHS \times X
+\rightharpoonup \MACHS \times X)$
+%
+ is a function that takes a contour to
+\begin{enumerate*}
+  \item a private type $X$,
+  \item an initial value of $X$, and
+  \item a partial step function of type
+  $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
+  agrees with $\stepstounderfull$ on machine states.
+\end{enumerate*}
+That is, a policy takes a starting contour and produces a runtime monitor
+that carries along some private information and uses it to decide whether to
+terminate execution early.
+
 Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
@@
 \bcp{
-  Following is not quite right.  Better intuition:
   \begin{itemize}
-  \item A (micro-){\em policy} is a function that takes a contour to
-  \begin{enumerate}
-  \item a private type $X$
-  \item an initial value of $X$
-  \item a partial step function of type $(\MACHS,X) \rightharpoonup
-  (\MACHS,X)$ that, when defined, agrees with $\stepstounderfull$ on
-  machine states.
-  \end{enumerate}
-  That is, it takes a starting contour and produces a runtime monitor, which
-  carries along some information that it uses to decide when to terminate
-  execution early.
   \item A {\em property} is a function that takes a contour to a
@@
 }
-
-Formally, a {\em policy}
-$\pol \in (\CONTOURS \times \MACHS) \rightarrow (\CONTOURS \times
-\MACHS)$ is an enriched
-step function that works on pairs of contours and machines---i.e., it is a
-partial function taking a contour and a machine state to an updated contour
-and machine state that, wherever is is defined, agrees with
-$\stepstounderfull$.  That is,
-it represents an enforcement mechanism that sometimes halts programs earlier
-than the bare step relation does, but that agrees with the bare step
-relation when it allows evaluation to proceed.  We write
-$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
-$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
 

### 2020-01-13 main.tex
Some more thoughts

@@
 
-\bcp{
+\bcp{Sketchy / out of date from here on...}
+
   \begin{itemize}
@@
   whether a given trace is OK to observe.
-  \item A policy $p$ {\em guarantees} a property $P$ if, for each contour
-  $\contour$, the sequence of machine states obtained by projecting the
-  sequence of ..... ARGH WHAT ABOUT VARIANTS!?!
-  \item Proposal: Forget about full-blown NI and make the whole paper be
-  about ``lockstep NI'' where the PC is the same in every pair of
-  related runs
+  \begin{itemize}
+  \item I am not sure whether we want to be talking in terms of {\em
+    properties} or {\em hyperproperties} here (and all the way through).
+  \end{itemize}
+  \item Proposal: Forget about full-blown NI and focus the whole paper on
+  variants of ``lockstep NI'' where the PC is the same in every pair of
+  related runs.
   \begin{itemize}
@@
   related runs are always in lockstep but, because the instruction memory is
-  writeable, different instructions can get dispatched.  E.g., XXXX
+  writeable, different instructions can get dispatched.
+
+  E.g., suppose the program looks like
+\begin{verbatim}
+0:   load from address 1000
+     write to address 100
+     jump to address 100
+100: <initial contents don't matter>
+     write R3 into address 1004
+     jump to address 100
+1000: <marked HIGH so that initial
+       contents vary on different runs>
+1004: <marked OBSERVABLE>
+\end{verbatim}
+  If the initial contents of address 1000 in one run is the encoding of the
+  instruction ``load immediate 0 into R3'' and the initial contents in the
+  other run is the encoding of the instruction ``load immediate 1 into R3,''
+  then the two programs execute different instructions (leading them to
+  write different things to an observable location) even though their PCs
+  are always in lock step.
+
   \item So maybe we should be asking that the instruction stream be
   identical, not (just?) the PCs
+  \end{itemize}
   \item One nice observation is that, if we are only interested in pairs of
@@
   stack and heap policies).
-  \end{itemize}
   \item A remaining question is what to do about the possibility that the
@@
   \end{itemize}
-}
 

### 2020-01-13 main.tex
Cp experiments with definitions

@@
 
-\subsection{Policies}
+\subsection{Contours and Variants}
 
@@
 predicate on programs (i.e., machine states) articulating some useful
-security goal, e.g. noninterference.
+security goal, e.g. noninterference. \iftext\bcp{Not quite true, if we stick
+  with hyperproperties.} \fi
 
@@
 \{\OBS,\HIDDEN\}$ is a map from machine components to pairs of ``security
-levels,'' where $\HIGH$ marks locations containing secret or low-integrity
-data and $\LOW$ marks those with public or high-integrity data, and
-``observabilities,'' where $\OBS$ marks locations whose contents are
-observable from the outside (by the ``attacker'') and $\HIDDEN$ marks
-internal ``hidden'' state that cannot be observed directly.
+levels,'' where $\HIGH$ marks locations that may contain secret or
+low-integrity data and $\LOW$ marks those that must only ever contain
+publicly readable or high-integrity data, and ``observabilities,'' where
+$\OBS$ marks locations whose contents are observable from the outside (by
+the ``attacker,'' e.g., by DMA) and $\HIDDEN$ marks internal ``hidden''
+state that cannot be observed directly (e.g., processor registers and memory
+addresses for temporary variables will usually be hidden).
+
+Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
+$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
+components agree with $\mach$ on all components marked $\LOW$ by
+$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
+$\mach(\component) = \mach'(\component)$.
+
+\subsection{Policies}
 
@@
 
-Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
-$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
-components agree with $\mach$ on all components marked $\LOW$ by
-$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
-$\mach(\component) = \mach'(\component)$.
+\subsection{Properties}
 
-\bcp{Sketchy / out of date from here on...}
+\bcp{Here is where it starts getting iffy...}
 
-  \begin{itemize}
-  \item A {\em property} is a function that takes a contour to a
-  prefix-closed set of (finite and/or infinite) sequences of machine states.
-  That is, it takes a starting contour and produces a checker that tells
-  whether a given trace is OK to observe.
-  \begin{itemize}
-  \item I am not sure whether we want to be talking in terms of {\em
-    properties} or {\em hyperproperties} here (and all the way through).
-  \end{itemize}
-  \item Proposal: Forget about full-blown NI and focus the whole paper on
-  variants of ``lockstep NI'' where the PC is the same in every pair of
-  related runs.
-  \begin{itemize}
-  \item One issue that I am not sure about: Could there be a sneaky way of
-  ``putting secret information into the control state'' where the PCs of
-  related runs are always in lockstep but, because the instruction memory is
-  writeable, different instructions can get dispatched.
-
-  E.g., suppose the program looks like
-\begin{verbatim}
-0:   load from address 1000
-     write to address 100
-     jump to address 100
-100: <initial contents don't matter>
-     write R3 into address 1004
-     jump to address 100
-1000: <marked HIGH so that initial
-       contents vary on different runs>
-1004: <marked OBSERVABLE>
-\end{verbatim}
-  If the initial contents of address 1000 in one run is the encoding of the
-  instruction ``load immediate 0 into R3'' and the initial contents in the
-  other run is the encoding of the instruction ``load immediate 1 into R3,''
-  then the two programs execute different instructions (leading them to
-  write different things to an observable location) even though their PCs
-  are always in lock step.
-
-  \item So maybe we should be asking that the instruction stream be
-  identical, not (just?) the PCs
-  \end{itemize}
-  \item One nice observation is that, if we are only interested in pairs of
-  runs with identical PCs, there are two different ways to proceed: (1) say
-  that we {\em ignore} runs where the PCs (or instruction streams) diverge
-  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
-  instruction streams) diverge, i.e., demand that this never happens (e.g.,
-  stack and heap policies).
-  \item A remaining question is what to do about the possibility that the
-  contour changes differently on two related runs.  We could disallow this,
-  but not certain it's necessary---maybe everything works out OK.
-  \end{itemize}
+A {\em contoured (hyper)property} is a function that takes a contour to a
+hyperproperty.  That is, it takes a starting contour and produces a
+\bcp{prefix-closed?} set of pairs of sequences of machine states.
 
@@
 
-Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
-every pair $(\mach,\mach')$ of variant machine states with respect to
-$\contour$ are indistinguishable under $\pol$.
-
 \iftext
@@
 \item Relation to the standard definition of NI.
-
-\item Fact: The standard RISC-V semantics does not guarantee NI.  (Give a
-little example of a program that reads from a secret location and writes to
-a public one, just to illustrate the definition of NI.)
 \end{itemize}
 \fi
+
+\bcp{Recheck.}
+Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
+every pair $(\mach,\mach')$ of variant machine states with respect to
+$\contour$ are indistinguishable under $\pol$.
+
+\iftext
+\bcp{
+Fact: The standard RISC-V semantics (i.e., the null policy) does not
+guarantee NI.  (Give a little example of a program that reads from a secret
+location and writes to a public one, just to illustrate the definition of
+NI.)}
+\fi
+
+\subsection{More...}
+  \begin{itemize}
+  \item Proposal: Maybe forget about full-blown NI and focus the whole paper
+  on variants of ``lockstep NI'' where the PC is the same in every pair of
+  related runs?
+  \item One issue that I am not sure about: Could there be a sneaky way of
+  ``putting secret information into the control state'' where the PCs of
+  related runs are always in lockstep but, because the instruction memory is
+  writeable, different instructions can get dispatched.
+
+  E.g., suppose the program looks like
+\begin{verbatim}
+0:   load from address 1000
+     write to address 100
+     jump to address 100
+100: <initial contents don't matter>
+     write R3 into address 1004
+     jump to address 100
+1000: <marked HIGH so that initial
+       contents vary on different runs>
+1004: <marked OBSERVABLE>
+\end{verbatim}
+  If the initial contents of address 1000 in one run is the encoding of the
+  instruction ``load immediate 0 into R3'' and the initial contents in the
+  other run is the encoding of the instruction ``load immediate 1 into R3,''
+  then the two programs execute different instructions (leading them to
+  write different things to an observable location) even though their PCs
+  are always in lock step.
+
+  So maybe we should be asking that the instruction stream be
+  identical, or maybe instructions paired with PCs?
+
+  \item One nice observation is that, if we are only interested in pairs of
+  runs with identical PCs, there are two different ways to proceed: (1) say
+  that we {\em ignore} runs where the PCs (or instruction streams) diverge
+  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
+  instruction streams) diverge, i.e., demand that this never happens (e.g.,
+  stack and heap policies).
+  \item A remaining question is what to do about the possibility that the
+  contour changes differently on two related runs.  We could disallow this,
+  but not certain it's necessary---maybe everything works out OK.
+  \end{itemize}
 

### 2020-01-14 main.tex
Notes
Thoughts

@@
 
-\subsection{Policies}
-
-Informally, a (micro-){\em policy}
-is an enriched
-step function that works on pairs of contours and machines---i.e., it is a
-partial function taking a contour and a machine state to an updated contour
-and machine state that, wherever is is defined, agrees with
-$\stepstounderfull$.  That is,
-it represents an enforcement mechanism that sometimes halts programs earlier
-than the bare step relation does, but that agrees with the bare step
-relation when it allows evaluation to proceed.  We write
-$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
-$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
-
-Formally, a policy
-%
-$\pol \in \CONTOURS \rightarrow \exists X.\; X \times (\MACHS \times X
-\rightharpoonup \MACHS \times X)$
-%
- is a function that takes a contour to
-\begin{enumerate*}
-  \item a private type $X$,
-  \item an initial value of $X$, and
-  \item a partial step function of type
-  $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
-  agrees with $\stepstounderfull$ on machine states.
-\end{enumerate*}
-That is, a policy takes a starting contour and produces a runtime monitor
-that carries along some private information and uses it to decide whether to
-terminate execution early.
-
-\subsection{Properties}
-
-\bcp{Here is where it starts getting iffy...}
-
-A {\em contoured (hyper)property} is a function that takes a contour to a
-hyperproperty.  That is, it takes a starting contour and produces a
-\bcp{prefix-closed?} set of pairs of sequences of machine states.
-
 The {\em observation trace} arising from a finite or infinite sequence of
-contours and machine states
+pairs of a contour and a machine state
 $(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$, written
@@
 partial map taking each machine component to the (finite or infinite)
-sequence of {\em distinct} values taken on by this component at moments when
-it is {\em observable}.  Formally, this map can be constructed as
+sequence of the {\em distinct} values taken on by this component at moments
+when it is {\em observable}.  Formally, this map can be constructed as
 follows. For each component $\component$,
@@
 \fi
+
+\subsection{Policies}
+
+Informally, a (micro-){\em policy}
+is an enriched
+step function that works on pairs of contours and machines---i.e., it is a
+partial function taking a contour and a machine state to an updated contour
+and machine state that, wherever is is defined, agrees with
+$\stepstounderfull$.  That is,
+it represents an enforcement mechanism that sometimes halts programs earlier
+than the bare step relation does, but that agrees with the bare step
+relation when it allows evaluation to proceed.  We write
+$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
+$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
+
+\bcp{Fix: even simpler and cleaner, just run the step function to completion
+  and forget about $X$---i.e., a property is just a function from a pair of
+  a contour and a machine state to a finite or infinite sequence of machine
+  states!}  Formally, a policy
+%
+$\pol \in \exists X.\; (\CONTOURS \rightarrow X) \times (\MACHS \times X
+\rightharpoonup \MACHS \times X)$
+%
+comprises
+\begin{enumerate*}
+  \item a private type $X$,
+  \item a function taking a contour to an initial value of $X$, and
+  \item a partial step function of type
+  $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
+  agrees with $\stepstounderfull$ on machine states. \bcp{We could actually
+    just make it return an X and calculate the machine state from the
+    ordinary step function.  This is perhaps a bit cleaner.}
+\end{enumerate*}
+That is, a policy takes a starting contour and produces a runtime monitor
+that carries along some private information and uses it to decide whether to
+terminate execution early.
+
+\subsection{Properties}
+
+% A {\em contoured (hyper)property} is a function that takes a contour to a
+% hyperproperty.  That is, it takes a starting contour and produces a
+% \bcp{prefix-closed?} set of pairs of sequences of machine states.
+
+\bcp{It gets sketchy from here on!  The observer needs to be simplified as
+  above, at a minimum.}
+
+An {\em observer} $\observer \in \exists X.\; (\CONTOURS \rightarrow X) \times
+(X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
+\times X) $
+comprises
+\begin{enumerate*}
+  \item a private type $X$,
+  \item a function taking a contour to an initial value of $X$, and
+  \item a function that allows ``reading off'' a contour from an $X$, and
+  \item a partial step function of type
+  $(\MACHS \times X) \rightharpoonup X$.
+\end{enumerate*}
 
@@
   identical, or maybe instructions paired with PCs?
+
+  \bcp{Answer: Just make instructions read-only.}
 

### 2020-01-14 main.tex
Progress

--- 2020-01-13 main.tex
@@
 
+A {\em contoured trace} is a finite or infinite sequence
+$\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
+(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs.
+
 Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
@@
 
-\subsection{Policies}
+\subsection{Observations}
 
-Informally, a (micro-){\em policy}
-is an enriched
-step function that works on pairs of contours and machines---i.e., it is a
-partial function taking a contour and a machine state to an updated contour
-and machine state that, wherever is is defined, agrees with
-$\stepstounderfull$.  That is,
-it represents an enforcement mechanism that sometimes halts programs earlier
-than the bare step relation does, but that agrees with the bare step
-relation when it allows evaluation to proceed.  We write
-$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
-$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
+A {\em component observation} is a finite or infinite sequence of words in
+which adjacent elements are always distinct (``no stuttering'').  An {\em
+  observation} is a map from machine components to component observations.
 
-Formally, a policy
-%
-$\pol \in \CONTOURS \rightarrow \exists X.\; X \times (\MACHS \times X
-\rightharpoonup \MACHS \times X)$
-%
- is a function that takes a contour to
-\begin{enumerate*}
-  \item a private type $X$,
-  \item an initial value of $X$, and
-  \item a partial step function of type
-  $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
-  agrees with $\stepstounderfull$ on machine states.
-\end{enumerate*}
-That is, a policy takes a starting contour and produces a runtime monitor
-that carries along some private information and uses it to decide whether to
-terminate execution early.
+The observation arising from a contoured trace $\seq{(\contour,\mach)}$,
+written $\obstrace(\seq{(\contour,\mach)})$, is a map taking each machine
+component to the (finite or infinite) sequence of the {\em distinct} values
+taken on by this component at moments when it is {\em observable}.
 
-\subsection{Properties}
-
-\bcp{Here is where it starts getting iffy...}
-
-A {\em contoured (hyper)property} is a function that takes a contour to a
-hyperproperty.  That is, it takes a starting contour and produces a
-\bcp{prefix-closed?} set of pairs of sequences of machine states.
-
-The {\em observation trace} arising from a finite or infinite sequence of
-contours and machine states
-$(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$, written
-$\obstrace((\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots)$ is a
-partial map taking each machine component to the (finite or infinite)
-sequence of {\em distinct} values taken on by this component at moments when
-it is {\em observable}.  Formally, this map can be constructed as
+Formally, $\obstrace(\seq{(\contour,\mach)})$ can be constructed as
 follows. For each component $\component$,
 %
-(1) if
+if
 $\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
@@
 %
-and otherwise (2), $\obstrace(\component)$ is a sequence whose first element
+otherwise, $\obstrace(\component)$ is a sequence whose first element
 is $\mach_i(\component)$ for the smallest $i$ such that
@@
 \bcp{Don't like the ``rest'' argument and the fact that we don't
-  have a metavariable for sequences of pairs of contours and machines.
+  have a metavariable for sequences of (contour, machine state) pairs.
   Consider defining one.}
@@
 \fi
+
+\subsection{Properties}
+
+% A {\em contoured (hyper)property} is a function that takes a contour to a
+% hyperproperty.  That is, it takes a starting contour and produces a
+% \bcp{prefix-closed?} set of pairs of sequences of machine states.
+
+% An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
+% (X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
+% \times X) $
+% comprises
+% \begin{enumerate*}
+%   \item a private type $X$,
+%   \item a function taking a contour to an initial value of $X$, and
+%   \item a function that allows ``reading off'' a contour from an $X$, and
+%   \item a partial step function of type
+%   $(\MACHS \times X) \rightharpoonup X$.
+% \end{enumerate*}
+
+A {\em contoured hyperproperty} is a set of pairs of contoured
+traces\iftext{} \bcp{Maybe prefix closed?}\fi.
+\bcp{Should we require that it contain all possible starting states and
+  contours (though of course the sequences can be singletons in many
+  cases!)?  What does it mean if it doesn't?}
+
+\subsection{Policies}
+
+Informally, a (micro-){\em policy} is an enriched step function that,
+whenever is is defined, agrees with $\stepstounderfull$.  It represents an
+enforcement mechanism that sometimes halts programs earlier than the bare
+step relation does but that agrees with the bare step relation as long as it
+allows evaluation to proceed.  We write
+$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
+$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
+
+Formally, a policy
+%
+$\pol \in \CONTOURS \rightarrow \MACHS \rightarrow \SEQ{\MACHS}$
+%
+takes a contour $\contour$ and a starting machine state $\mach_0$ and produces
+a finite or infinite sequence of states $\mach_0,\,\mach_1 ,\, \ldots$ such
+that $\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.
+
+% \bcp{Fix: even simpler and cleaner, just run the step function to completion
+%   and forget about $X$---i.e., a property is just a function from a pair of
+%   a contour and a machine state to a finite or infinite sequence of machine
+%   states!}
+% Formally, a policy
+% %
+% $\pol \in \exists X.\; (\CONTOURS \rightarrow X) \times (\MACHS \times X
+% \rightharpoonup \MACHS \times X)$
+% %
+% comprises
+% \begin{enumerate*}
+%   \item a private type $X$,
+%   \item a function taking a contour to an initial value of $X$, and
+%   \item a partial step function of type
+%   $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
+%   agrees with $\stepstounderfull$ on machine states. \bcp{We could actually
+%     just make it return an X and calculate the machine state from the
+%     ordinary step function.  This is perhaps a bit cleaner.}
+% \end{enumerate*}
+% That is, a policy takes a starting contour and produces a runtime monitor
+% that carries along some private information and uses it to decide whether to
+% terminate execution early.
+
+Policy $\pol$ {\em guarantees} property $\prop$ if \bcp{......}
+
+% every $\contour$,
+% every pair $(\mach,\mach')$ of variant machine states with respect to
+% $\contour$ are indistinguishable under $\pol$.
 
@@
 
-\bcp{Recheck.}
-Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
-every pair $(\mach,\mach')$ of variant machine states with respect to
-$\contour$ are indistinguishable under $\pol$.
+% Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
+% every pair $(\mach,\mach')$ of variant machine states with respect to
+% $\contour$ are indistinguishable under $\pol$.
 
@@
 
+\iftext
 \subsection{More...}
@@
 
+  \bcp{Answer: Just make instructions read-only.}
+
   \item One nice observation is that, if we are only interested in pairs of
@@
   \end{itemize}
+\fi
 

### 2020-01-14 main.tex
New try

@@
 $\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
-(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs.
+(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs \bcp{such
+  that the machine parts agree with the full step}.
 
@@
 $\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
-$\mach(\component) = \mach'(\component)$.
+$\mach(\component) = \mach'(\component)$.  \bcp{Move this to the NI section.}
 
@@
 $\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
-$\obstrace(\component) = \bot$;
+$\obstrace(\component) = [\,]$;
 %
@@
 %
+\bcp{Can be defined more simply}
 \iftext
@@
 
+\bcp{Maybe make it a sequence of maps from components to values?  This is
+  simpler, but it means we can ``observe the observbility of locations...''
+But is this too unrealistic?}
+
 \subsection{Properties}
@@
   cases!)?  What does it mean if it doesn't?}
+
+\subsection{Noninterference}
+
+\bcp{NI is the property where:
+  \begin{itemize}
+  \item In each top-level pair (each element of the hyperproperty), all the
+  contours coincide.
+  \item If the initial machine states plus contour are variants, then the
+  observervations of the two traces coincide.
+  \end{itemize}
+}
+
+% If $\mach$ and $\mach'$ are variant machine states with respect to contour
+% $\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
+% under policy $\pol$ if their observable
+% %
+% $\obstrace(
+%                       (\contour,\mach)
+%   \stepstounder{\pol} (\contour_1,\mach_1)
+%   \stepstounder{\pol} (\contour_2,\mach_2)
+%   \stepstounder{\pol} \ldots)$
+% %
+% yields the same trace as
+% $\obstrace(
+%                       (\contour,\mach')
+%   \stepstounder{\pol} (\contour_1,\mach_1')
+%   \stepstounder{\pol} (\contour_2,\mach_2')
+%   \stepstounder{\pol} \ldots)$.
+
+% % either (1) one or both of $(\contour,\mach)$ and
+% % $(\contour,\mach')$ fail to terminate when executed with
+% % $\stepstounder{\pol}$, or (2) both terminate and the machine-state
+% % components of their results are variants with respect to the original
+% % $\contour$.
+
+\iftext
+\begin{itemize}
+\item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
+  have to begin with EENI and then get to SSNI as a (stronger) refinement.}
+
+\item Relation to the standard definition of NI.
+\end{itemize}
+\fi
+
+% Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
+% every pair $(\mach,\mach')$ of variant machine states with respect to
+% $\contour$ are indistinguishable under $\pol$.
+
+\iftext
+\bcp{
+Fact: The standard RISC-V semantics (i.e., the null policy) does not
+guarantee NI.  (Give a little example of a program that reads from a secret
+location and writes to a public one, just to illustrate the definition of
+NI.)}
+\fi
 
@@
 
-Policy $\pol$ {\em guarantees} property $\prop$ if \bcp{......}
+Policy $\pol$ {\em guarantees} property $\prop$ if \bcp{for every pair of
+  starting states and every starting contour, the pair of traces that you
+  get from the policy is in the property for {\em some} pair of sequences of
+  contours that agree with the given contour on the first element.}
 
@@
 % $\contour$ are indistinguishable under $\pol$.
-
-\subsection{Noninterference}
-
-If $\mach$ and $\mach'$ are variant machine states with respect to contour
-$\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
-under policy $\pol$ if their observable
-%
-$\obstrace(
-                      (\contour,\mach)
-  \stepstounder{\pol} (\contour_1,\mach_1)
-  \stepstounder{\pol} (\contour_2,\mach_2)
-  \stepstounder{\pol} \ldots)$
-%
-yields the same trace as
-$\obstrace(
-                      (\contour,\mach')
-  \stepstounder{\pol} (\contour_1,\mach_1')
-  \stepstounder{\pol} (\contour_2,\mach_2')
-  \stepstounder{\pol} \ldots)$.
-
-% either (1) one or both of $(\contour,\mach)$ and
-% $(\contour,\mach')$ fail to terminate when executed with
-% $\stepstounder{\pol}$, or (2) both terminate and the machine-state
-% components of their results are variants with respect to the original
-% $\contour$.
-
-\iftext
-\begin{itemize}
-\item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
-  have to begin with EENI and then get to SSNI as a (stronger) refinement.}
-
-\item Relation to the standard definition of NI.
-\end{itemize}
-\fi
-
-% Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
-% every pair $(\mach,\mach')$ of variant machine states with respect to
-% $\contour$ are indistinguishable under $\pol$.
-
-\iftext
-\bcp{
-Fact: The standard RISC-V semantics (i.e., the null policy) does not
-guarantee NI.  (Give a little example of a program that reads from a secret
-location and writes to a public one, just to illustrate the definition of
-NI.)}
-\fi
 
@@
 
+\newpage
 \section{Tainting is noninterference$-$}

### 2020-01-14 main.tex
More rearrangement

@@
 
-\subsection{Contours and Variants}
+\subsection{Contours}
 
+\iftext
 Intuitively, we use the word {\em policy} for any kind of dynamic
@@
   with hyperproperties.} \fi
+\fi
+
+\bcp{Idea: Remove the HIGH/LOW part from contours and instead make it part
+  of the $X$ parameter to both properties and policies.  (I.e., we will have
+  ``X properties'' and ``X policies'' for each X.)}
 
@@
 
-Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
-$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
-components agree with $\mach$ on all components marked $\LOW$ by
-$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
-$\mach(\component) = \mach'(\component)$.  \bcp{Move this to the NI section.}
+\subsection{Observations}
 
-\subsection{Observations}
+\bcp{Try it the other way round: a sequence of maps from components to
+  values.  Need to extend values with $\bot$ for always-hidden elements.
+  Emphasize that ``observability is not observable.''}
+\bcp{Also, change OBS to VIS}
 
@@
 
-\bcp{Maybe make it a sequence of maps from components to values?  This is
-  simpler, but it means we can ``observe the observbility of locations...''
-But is this too unrealistic?}
+% \bcp{Maybe make it a sequence of maps from components to values?  This is
+%   simpler, but it means we can ``observe the observbility of locations...''
+%   But is this too unrealistic?}
 
@@
 traces\iftext{} \bcp{Maybe prefix closed?}\fi.
-\bcp{Should we require that it contain all possible starting states and
-  contours (though of course the sequences can be singletons in many
-  cases!)?  What does it mean if it doesn't?}
+% \bcp{Should we require that it contain all possible starting states and
+%   contours (though of course the sequences can be singletons in many
+%   cases!)?  What does it mean if it doesn't?}
+
+\bcp{Not quite enough: Policies and properties needs to know where are the
+  call and return sites.  So what we should do is to talk instead about ``$X$
+  policies'' (i.e., make ``policy'' a type operator) and replace the initial
+  $\CONTOURS$ argument to both policies and properties by an element of
+  $X$.}
 
 \subsection{Noninterference}
+
+Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
+$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
+components agree with $\mach$ on all components marked $\LOW$ by
+$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
+$\mach(\component) = \mach'(\component)$.
 

### 2020-01-15 main.tex
wip

@@
 
+Here we consider stack protection as an instance of generalized interference \apt{or whatever it's called}.
+Our notion of stack protection includes aspects both of control integrity and data security.
+In the terminology of Skorstengaard et al.~\cite{Skorstengaard+19} we enforce:
+
+\begin{itemize}
+\item \emph{Well-bracketed control flow (WBCF)}, which says that ``invoked functions must either return to their callers,
+  invoke other functions themselves, or diverge.''
+\item \emph{Local state encapsulation (LSE)}, which says that ``when a function invokes another function, its local variables
+  (saved on its stack frame) will not be read or modified until the invoked function returns.''
+\end{itemize}
+\apt{Should rephrase (and improve!) these in our own words.}
+
+These properties are essentially orthogonal: we could enforce WBCF without LSE, and variants of
+LSE could apply to control constructs beyond simple procedure calls (e.g. exceptions, coroutines, threads, monitors, etc.).
+
+In terms of NI....
+\apt{Perhaps the NI-like property is just LSE, but it holds only if WBCF holds?}
+
+To make these ideas concrete in this paper, we adopt a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls.
+Each procedure stack frame contains just one local variable (plus possible control information), which is pointed to by
+the SP whenever the PC is within the procedure body (excluding entry and exit sequences). Our WBCF property says that
+when a procedure is called from a given PC at a given SP, it eventually returns (if at all) to the next
+address following the PC with the same SP. Our LSE-like property says that addresses between the base of the stack
+up to (but not including) the current SP cannot be read or written.
+\apt{Not clear why we should bother allowing addresses beyond the current top of stack to be accessible.}
+In particular, this implies that upon return from a call,
+the value of the local variable will be unchanged; moreover, the values of local variables cannot influence the
+behavior of callees. Parameters and return values are passed in registers, which (except for SP and RA) do not change
+during call or return.
+
+{\bf Getting more precise about the property.}
+
+Assumptions on code and environment\apt{trying to be as minimal as possible}:
 \begin{itemize}
 \item 
-  Idea of accessibility: set of data memory addresses that can legally be read or written at a particular execution point.
-  (Assume instruction memory is distinct.)
+  Memory is divided into two disjoint regions, one containing the code, the other the stack.
+  Code begins at address 0; the stack begins at address B and grows up, in the direction of increasing addresses.
 \item
-  Stack protection intuition in a nutshell: caller's memory data is not accessible to callee, although callee's memory data is accessible to caller.
-  This gives both an integrity property (callee cannot affect caller's memory) and a secrecy property
-  (caller's memory cannot affect callee).
-  Can generalize to other control situations, such as coroutines, threads, compartments,...
+  The part of the stack between $B$ and the stack pointer \textsc{sp} (inclusive) is structured into a contiguous sequence of
+  \emph{frames}. Each frame contains a single word of local data, possibly preceded by control information. \apt{TMI?}
+  The initial value of \textsc{sp} is $B-1$.
 \item
-  In NI terms, protected programs are oblivious to variations in memory outside of their accessible set.
-  Unlike standard NI, the accessible set changes dynamically, so we cannot just frame the property in terms of
-  variant start states. [Is this true? Is it a significant difference? Are there other significant differences?]
+  Code is assumed to be be annotated to indicate (a) instructions representing calls, (b) instructions forming the \emph{entry sequence} of
+  a procedure, and (c) instructions forming the \emph{return sequence} of a procedure.
+\item
+  Execution of the entry sequence pushes a new frame on the stack, adjusting \textsc{sp} to point to that
+  frame's data word.  Execution of the return sequence pops the frame on top of the stack by readjusting \textsc{sp}.
+  \apt{TMI!?}
+  Outside of the entry and return sequences, no instruction alters \textsc{sp}.
+\item
+  The call instruction and entry and return sequences do not alter any registers except \textsc{sp} and one
+  particular register named \textsc{ra}.
 \end{itemize}
 
-We start with a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls.
-We assume that memory is divided into two disjoint regions, one containing the code, the other the stack.
-Code begins at address 0; the stack begins at address $B$ and grows in the direction of increasing addresses.
+DO HERE: Formal description of conjecture 2 property.  Or maybe we no longer really care about this particular property if we have something for conjecture 3?
 
-The part of the stack between $B$ and the stack pointer \textsc{sp} is structured into a sequence of consecutive
-\emph{frames}. Each frame consists of a single return address and a single word of local data.
-
-  \begin {itemize}
-  \item Memory is divided into an instruction region and a data region.
-The data region contains a single contiguous stack, divided into frames, each associated with a procedure activation.
-(The stack grows in the direction of increasing addresses.)
-  \item Each frame consists of a saved return address and a single word of local data.
-These [both??] form the entirety of the procedure's memory data, and are the only thing we need to track when considering accessibility.
-  \item Registers other than RA, SP are not affected by calls or returns and hence are the (only) way to pass values between caller and callee.
-  \item Instructions serving as calls are statically marked as such in the code. [The entry and exit sequences must also be marked???]
-  \item The code for each procedure is bracketed by fixed entry and exit sequences (details to follow), which save/restore the RA and
-    adjust the SP. [Are these assumptions or checked facts?]
-  \item During the body of the procedure, the single word of local data is accessed via SP, which does not change (except during a call, of course).
-The stored RA in the frame never changes.
-[I'm confused about these.
-They are checked facts, not assumptions.]
-  \end {itemize}
-  In this model, the definition of accessible memory is simply all (data) memory at addresses $\geq pc$.
+An (abstract) \emph{state description} $D$ consists of
 \begin{itemize}
 \item
-  Describe conjecture 2 property in terms of this model: informally and formally.
+  A stack $\dstk{D}$ of addresses, representing the return addresses of the current procedure and its callers. 
+  \apt{No need to remember the sp's, since they can be computed given that we know the fixed frame size.}
+  \apt{No need for $\dpcd{D}$ since it is just the size of $\dstk{D}$.}
 \item
-  Define abstract state descriptions.
+  For each memory location $a \geq B$, a natural number $\ddep{D}{a}$, representing the depth
+  of the frame to which $a$ belongs ($\infty$ if none).
+\end{itemize}
+
+The \emph{initial state description} $\dinit$ has $\dstk{\dinit} = \empstack$ and $\ddep{\dinit}{a} = \infty$ for all $a \geq B$.
+
+An address $a$ is \emph{accessible} with respect to a state $D$, written $\access{D}{a}$, if
+$a \geq B$ and $\ddep{D}{a} \geq \norm{\dstk{D}}$.
+
+TODO:
+\begin{itemize}
+\item
+  Define step relation.
 \item
@@
 \item   
-  Discuss relationship between state consistency and conjecture 2 property.  (Former is: presumably stronger; hopefully not
+  Discuss relationship between state consistency and top-level property.  (Former is: presumably stronger; hopefully not
   too strong that it produces lots of undesired test failures; surely easier to test).

### 2020-01-15 main.tex
typo
CP progress on basic definition

@@
 
-\bcp{Idea: Remove the HIGH/LOW part from contours and instead make it part
-  of the $X$ parameter to both properties and policies.  (I.e., we will have
-  ``X properties'' and ``X policies'' for each X.)}
-
 A {\em contour}
-$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}\times
-\{\OBS,\HIDDEN\}$ is a map from machine components to pairs of ``security
-levels,'' where $\HIGH$ marks locations that may contain secret or
-low-integrity data and $\LOW$ marks those that must only ever contain
-publicly readable or high-integrity data, and ``observabilities,'' where
-$\OBS$ marks locations whose contents are observable from the outside (by
-the ``attacker,'' e.g., by DMA) and $\HIDDEN$ marks internal ``hidden''
+$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\VIS,\HID\}$ is a map
+from machine components to ``observabilities,'' where
+$\VIS$ marks locations whose contents are observable from the outside (by
+the ``attacker,'' e.g., by DMA) and $\HID$ marks internal ``hidden''
 state that cannot be observed directly (e.g., processor registers and memory
@@
 $\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
-(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs \bcp{such
-  that the machine parts agree with the full step}.
+(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs in which
+adjacent machine states are related by $\stepstounderfull$---i.e., in which
+$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.
 
@@
 
-\bcp{Try it the other way round: a sequence of maps from components to
-  values.  Need to extend values with $\bot$ for always-hidden elements.
-  Emphasize that ``observability is not observable.''}
-\bcp{Also, change OBS to VIS}
+A {\em component observation} is either a word or the special value $\bot$.
+A {\em state observation} is a map from machine components to component
+observations.
 
-A {\em component observation} is a finite or infinite sequence of words in
-which adjacent elements are always distinct (``no stuttering'').  An {\em
-  observation} is a map from machine components to component observations.
+An {\em observation} is a nonempty, finite or infinite sequence of state
+observations in which (1) adjacent elements are always distinct (``no
+stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
+the the $j^{th}$ state observation in the sequence must also be mapped to
+$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
+component's value changes from $\bot$ to anything else, it never changes
+back to $\bot$ again).
 
-The observation arising from a contoured trace $\seq{(\contour,\mach)}$,
-written $\obstrace(\seq{(\contour,\mach)})$, is a map taking each machine
-component to the (finite or infinite) sequence of the {\em distinct} values
-taken on by this component at moments when it is {\em observable}.
+In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
+from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
+observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
+follows:
+\begin{itemize}
+\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
+to its value in $\mach_0$ and every hidden
+component to $\bot$; 
+\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
+visible component (where $\contour_{i+1} = \VIS$), copying its
+value from $\mach_{i+1}$, and (2) for each
+hidden component, where $\contour_i = \VIS$, copying its
+value from $\stateobs_{i+1}$, and (3) if the resulting $\stateobs_{i+1}$ is
+identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
+\iftext\bcp{Too informal!!}\fi
 
-Formally, $\obstrace(\seq{(\contour,\mach)})$ can be constructed as
-follows. For each component $\component$,
-%
-if
-$\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
-$\obstrace(\component) = [\,]$;
-%
-otherwise, $\obstrace(\component)$ is a sequence whose first element
-is $\mach_i(\component)$ for the smallest $i$ such that
-$\contour_i(\component) = (\underscore,\OBS)$ and whose $(m+1)^{th}$ element
-(if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
-than the index $i$ of the $m^{th}$ element and such that
-$\contour_j(\component) = (\underscore,\OBS)$ and
-$\mach_j(\component) \neq \mach_i(\component)$.
-%
-\bcp{Can be defined more simply}
-\iftext
-\bcp{Unfinished and horrible...  But maybe the textual description is OK?}
-\newcommand{\aux}{\mathit{aux}}
-\bcp{Don't like the ``rest'' argument and the fact that we don't
-  have a metavariable for sequences of (contour, machine state) pairs.
-  Consider defining one.}
-%
-\[
-\begin{array}{lcl}
-\aux \; ( [\,] ) \; \trace& = & \trace
-\\
-\aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
-  & = &
-        \aux \;( \mathit{rest} )
-        \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
-\\
-\\
-\obstrace \;( (\contour_0,\mach_0),\, \ldots )
-  &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
-      \component.\, \bot)
-\end{array}
-\]
-\fi
+% For each component $\component$,
+% %
+% if
+% $\contour_i(\component) = (\underscore,\HID)$ for all $i$, then
+% $\obstrace(\component) = [\,]$;
+% %
+% otherwise, $\obstrace(\component)$ is a sequence whose first element
+% is $\mach_i(\component)$ for the smallest $i$ such that
+% $\contour_i(\component) = (\underscore,\VIS)$ and whose $(m+1)^{th}$ element
+% (if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
+% than the index $i$ of the $m^{th}$ element and such that
+% $\contour_j(\component) = (\underscore,\VIS)$ and
+% $\mach_j(\component) \neq \mach_i(\component)$.
+% %
+% \bcp{Can be defined more simply}
+% \iftext
+% \bcp{Unfinished and horrible...  But maybe the textual description is OK?}
+% \newcommand{\aux}{\mathit{aux}}
+% \bcp{Don't like the ``rest'' argument and the fact that we don't
+%   have a metavariable for sequences of (contour, machine state) pairs.
+%   Consider defining one.}
+% %
+% \[
+% \begin{array}{lcl}
+% \aux \; ( [\,] ) \; \trace& = & \trace
+% \\
+% \aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
+%   & = &
+%         \aux \;( \mathit{rest} )
+%         \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
+% \\
+% \\
+% \obstrace \;( (\contour_0,\mach_0),\, \ldots )
+%   &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
+%       \component.\, \bot)
+% \end{array}
+% \]
+% \fi
 
@@
 
+\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
+
 \subsection{Properties}
+
+\bcp{Needs a parameter now.}
 
@@
 
-Here we consider stack protection as an instance of generalized interference \apt{or whatever it's called}.
+Here we consider stack protection as an instance of generalized noninterference \apt{or whatever it's called}.
 Our notion of stack protection includes aspects both of control integrity and data security.

### 2020-01-17 main.tex
CP progress on basic definition
Notes for meeting

@@
 \iftext\bcp{Too informal!!}\fi
+\end{itemize}
 
@@
 
+\leo{Low should be VIS here right?}
 Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
@@
 }
+
+\leo{Some notes for the meeting}
+
+In traditional NI viewed as a pair of contoured traces $\overline{(C, M_i)}$ and
+$\overline{(C, N_i)}$, we can have the following formulations:
+\begin{enumerate}
+\item $M_0, N_0 \in \mathit{Initial State}, ~\text{s.t.} ~\mathit{variant}(C, M_0, N_0)$
+\item $M_0 \in \mathit{Initial State}, N_0 \in \mathit{Variant}(C, M_0)$
+\item $M_0, N_0 \in \mathit{Reachable State}, ~\text{s.t.} ~\mathit{variant}(C, M_0, N_0)$
+\item $M_0 \in \mathit{Reachable State}, N_0 \in \mathit{Variant}(C, M_0)$
+\end{enumerate}
+%
+The (intuitive) top level property we care about is (1) - end to end
+noninterference.
+%
+Property (2) is easier to test - but not always equivalent (if the set
+of variants of an initial state can take you outside the set of
+initial states).
+%
+Property (3) is also easier to test than - you start from an arbitrary
+(reachable) state instead of an initial one, allowing to get to the
+interesting part sooner.
+%
+Property (4) corresponds to what we actually want to test - start from
+an arbitrary (reachable) state, vary it, test the trace.
+
+As far as facilitating testing goes, the assumption of SSNI is that
+you don't need to test full traces - you can test all traces of length
+1 and that somehow gives you confidence for the full property.
+%
+More importantly, note that going from (1,2) to (3,4) doesn't really
+change the property itself {\em as long as the contour stays identical
+  throughout the execution}.
+
+Now, when the contour does change the expressive power of (3,4) and (1,2)
+is slightly different:
+\begin{itemize}
+\item If the contour C makes more things hidden at step $i$, then
+  starting from the (reachable) state $M_i$ and taking its variant can
+  lead to a pair $M_i, N_i$ that is not reachable from a pair of variant
+  initial states.
+\item If the contour C declassifies things at step $i$, then traces
+  that started from variants $M_0, N_0$ based on a more restrictive
+  contour C, can reach that point and no longer remain variant based
+  on the new contour.
+\end{itemize}
+Therefore an additional property of the hyperproperties when
+classification/declassification is involved, is that the traces
+in the property start at some contour level C0 and are truncated
+after C becomes more permissive than C0.
+
+The stack property we test is just a way of testing such a
+hyperproperty more efficiently.
 

### 2020-01-17 main.tex
First stab at the (closure) properties of contoured hyperproperties. Too tedious?

@@
 
-A {\em contoured trace} is a finite or infinite sequence
-$\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
-(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs in which
-adjacent machine states are related by $\stepstounderfull$---i.e., in which
-$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.
+A {\em contoured trace} of length $n$, where $n$ can be finite or
+infinite, is a length-$n$ sequence $\seq{(\contour,\mach)}_n =
+(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$ of (contour,
+machine state) pairs in which adjacent machine states are related by
+$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
+\mach_i$ for each $i > 0$ in the sequence.
 
-\subsection{Observations}
-
-A {\em component observation} is either a word or the special value $\bot$.
-A {\em state observation} is a map from machine components to component
-observations.
-
-An {\em observation} is a nonempty, finite or infinite sequence of state
-observations in which (1) adjacent elements are always distinct (``no
-stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
-the the $j^{th}$ state observation in the sequence must also be mapped to
-$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
-component's value changes from $\bot$ to anything else, it never changes
-back to $\bot$ again).
-
-In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
-from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
-observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
-follows:
-\begin{itemize}
-\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
-to its value in $\mach_0$ and every hidden
-component to $\bot$; 
-\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
-visible component (where $\contour_{i+1} = \VIS$), copying its
-value from $\mach_{i+1}$, and (2) for each
-hidden component, where $\contour_i = \VIS$, copying its
-value from $\stateobs_{i+1}$, and (3) if the resulting $\stateobs_{i+1}$ is
-identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
-\iftext\bcp{Too informal!!}\fi
-\end{itemize}
+% 
+%\subsection{Observations}
+% 
+%A {\em component observation} is either a word or the special value $\bot$.
+%A {\em state observation} is a map from machine components to component
+%observations.
+% 
+%An {\em observation} is a nonempty, finite or infinite sequence of state
+%observations in which (1) adjacent elements are always distinct (``no
+%stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
+%the the $j^{th}$ state observation in the sequence must also be mapped to
+%$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
+%component's value changes from $\bot$ to anything else, it never changes
+%back to $\bot$ again).
+% 
+%In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
+%from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
+%observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
+%follows:
+%\begin{itemize}
+%\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
+%to its value in $\mach_0$ and every hidden
+%component to $\bot$; 
+%\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
+%visible component (where $\contour_{i+1} = \VIS$), copying its
+%value from $\mach_{i+1}$, and (2) for each
+%hidden component, where $\contour_i = \VIS$, copying its
+%value from $\stateobs_{i+1}$, and (3) if the resulting $\stateobs_{i+1}$ is
+%identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
+%\iftext\bcp{Too informal!!}\fi
+%\end{itemize}
 
@@
 % \end{enumerate*}
+\leo{The following all assumes lockstep-ness:}
+
+\leo{I think we need variants from here... Should we call it
+  indistinguishability?}
+
+Given a machine state $\mach$ and a contour $\contour$, a {\em
+  variant} of $\mach$ with respect to $\contour$ is any machine state
+$\mach'$ whose components agree with $\mach$ on all components marked
+$\VIS$ by $\contour$---i.e., if $\contour(\component) = \VIS$, then
+$\mach(\component) = \mach'(\component)$.
 
 A {\em contoured hyperproperty} is a set of pairs of contoured
-traces\iftext{} \bcp{Maybe prefix closed?}\fi.
+traces such that:
+\begin{itemize}
+\item {\em same-length:}
+
+  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
+    M')}_m$ in the set, $n = m$.
+
+\item {\em contour-lockstep:}
+
+  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
+    M')}_n$ in the set, $C_i = C_i'$ for all $i<n$.
+
+\item {\em prefix-closed:}
+
+  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
+    M')}_n$ in the set, the pair $\overline{(C, M)}_m$ and
+  $\overline{(C, M')}_m$ is also in the set if $m < n$.
+
+\item {\em suffix-closed:}
+
+  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
+    M')}_n$ in the set, and for every $0 \leq i < n$, there exists in
+  the set a pair of traces $\overline{(C', N)}_m$ and $\overline{(C',
+    N')}_m$ such that $C_0' = C_i$, $N_0 = M_i$, and $N_0' = M_0'$.
+
+\item {\em variant-closed:}
+
+  For every pair of traces $\overline{(C, M)}$ and $\overline{(C,
+    M')}$ in the set, and for every machine state $N_0$ such that
+  $N_0$ is a variant of $M_0$, there exists in the set a pair of
+  traces $\overline{(C, M)}_m$ and $\overline{(C, N)}_m$.
+
+\item {\em contour-wellformedness:}
+
+  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
+    M')}_n$ in the set, $C_0 \leq C_i$ for all $i<n$, where $C \leq
+  C'$ if $\forall \component, C(\component) \leq C'(\component)$ and
+  $\VIS < \HID$.
+
+\end{itemize}
+\iftext{} \bcp{Maybe prefix closed?}\fi.
 % \bcp{Should we require that it contain all possible starting states and
@@
 
-\leo{Low should be VIS here right?}
 Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of

### 2020-01-17 main.tex
What was wrong with obstraces?

--- 2020-01-15 main.tex
@@
 
-A {\em contoured trace} is a finite or infinite sequence
-$\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
-(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs in which
-adjacent machine states are related by $\stepstounderfull$---i.e., in which
-$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.
+A {\em contoured trace} of length $n$, where $n$ can be finite or
+infinite, is a length-$n$ sequence $\seq{(\contour,\mach)}_n =
+(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$ of (contour,
+machine state) pairs in which adjacent machine states are related by
+$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
+\mach_i$ for each $i > 0$ in the sequence.
 
@@
 
-Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
-$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
-components agree with $\mach$ on all components marked $\LOW$ by
-$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
-$\mach(\component) = \mach'(\component)$.
+NI is the contoured hyperproperty where:
+\begin{itemize}
+\item For each pair of traces in the hyperproperty, all the contours
+  coincide.
+\item For each pair of traces in the hyperproperty, If the initial
+  machine states plus contour are variants, then the observervations
+  of the two traces coincide.
+\end{itemize}
 
-\bcp{NI is the property where:
-  \begin{itemize}
-  \item In each top-level pair (each element of the hyperproperty), all the
-  contours coincide.
-  \item If the initial machine states plus contour are variants, then the
-  observervations of the two traces coincide.
-  \end{itemize}
-}
+%  
+% \leo{Some notes for the meeting}
+%  
+% In traditional NI viewed as a pair of contoured traces $\overline{(C, M_i)}$ and
+% $\overline{(C, N_i)}$, we can have the following formulations:
+% \begin{enumerate}
+% \item $M_0, N_0 \in \mathit{Initial State}, ~\text{s.t.} ~\mathit{variant}(C, M_0, N_0)$
+% \item $M_0 \in \mathit{Initial State}, N_0 \in \mathit{Variant}(C, M_0)$
+% \item $M_0, N_0 \in \mathit{Reachable State}, ~\text{s.t.} ~\mathit{variant}(C, M_0, N_0)$
+% \item $M_0 \in \mathit{Reachable State}, N_0 \in \mathit{Variant}(C, M_0)$
+% \end{enumerate}
+% %
+% The (intuitive) top level property we care about is (1) - end to end
+% noninterference.
+% %
+% Property (2) is easier to test - but not always equivalent (if the set
+% of variants of an initial state can take you outside the set of
+% initial states).
+% %
+% Property (3) is also easier to test than - you start from an arbitrary
+% (reachable) state instead of an initial one, allowing to get to the
+% interesting part sooner.
+% %
+% Property (4) corresponds to what we actually want to test - start from
+% an arbitrary (reachable) state, vary it, test the trace.
+%  
+% As far as facilitating testing goes, the assumption of SSNI is that
+% you don't need to test full traces - you can test all traces of length
+% 1 and that somehow gives you confidence for the full property.
+% %
+% More importantly, note that going from (1,2) to (3,4) doesn't really
+% change the property itself {\em as long as the contour stays identical
+%   throughout the execution}.
+%  
+% Now, when the contour does change the expressive power of (3,4) and (1,2)
+% is slightly different:
+% \begin{itemize}
+% \item If the contour C makes more things hidden at step $i$, then
+%   starting from the (reachable) state $M_i$ and taking its variant can
+%   lead to a pair $M_i, N_i$ that is not reachable from a pair of variant
+%   initial states.
+% \item If the contour C declassifies things at step $i$, then traces
+%   that started from variants $M_0, N_0$ based on a more restrictive
+%   contour C, can reach that point and no longer remain variant based
+%   on the new contour.
+% \end{itemize}
+% Therefore an additional property of the hyperproperties when
+% classification/declassification is involved, is that the traces
+% in the property start at some contour level C0 and are truncated
+% after C becomes more permissive than C0.
+%  
+% The stack property we test is just a way of testing such a
+% hyperproperty more efficiently.
 

### 2020-01-17 main.tex
Reorganize to make things better

@@
 
-\bcp{Needs a parameter now.}
-
 % A {\em contoured (hyper)property} is a function that takes a contour to a
@@
 % \end{enumerate*}
-\leo{The following all assumes lockstep-ness:}
-
-\leo{I think we need variants from here... Should we call it
-  indistinguishability?}
-
-Given a machine state $\mach$ and a contour $\contour$, a {\em
-  variant} of $\mach$ with respect to $\contour$ is any machine state
-$\mach'$ whose components agree with $\mach$ on all components marked
-$\VIS$ by $\contour$---i.e., if $\contour(\component) = \VIS$, then
-$\mach(\component) = \mach'(\component)$.
 
 A {\em contoured hyperproperty} is a set of pairs of contoured
-traces such that:
+traces.
+
+A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
+such that:
+
+\leo{Informally first, I don't like it too much:}
+\begin{itemize}
+\item each trace in every pair of a contoured trace in the set have the
+  same length, and
+\item each pair of elements in every pair of a contoured trace in the
+  set have the same contour.
+\end{itemize}
+
+\leo{Formally:}
 \begin{itemize}
@@
     M')}_n$ in the set, $C_i = C_i'$ for all $i<n$.
+\end{itemize}
 
+\subsection{Noninterference}
+
+\leo{Indistinguishability/variations:}
+
+Given a machine state $\mach$ and a contour $\contour$, a {\em
+  variant} of $\mach$ with respect to $\contour$ is any machine state
+$\mach'$ whose components agree with $\mach$ on all components marked
+$\VIS$ by $\contour$---i.e., if $\contour(\component) = \VIS$, then
+$\mach(\component) = \mach'(\component)$.
+
+\leo{Normal lockstep noninterference:}
+
+(Lockstep-)Noninterference is the contoured lockstep hyperproperty
+where:
+\begin{itemize}
+\item For each pair of traces in the hyperproperty, all the contours
+  coincide.
+\item For each pair of traces in the hyperproperty, If the initial
+  machine states plus contour are variants, then the observervations
+  of the two traces coincide.
+\end{itemize}
+
+\leo{Nested lockstep noninterference:}
+
+Generalizing to allow for contours to change throughout the execution,
+nested lockstep noninterference is the contoured lockstep
+hyperproperty where:
+
+\begin{itemize}
 \item {\em prefix-closed:}
@@
   For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
-    M')}_n$ in the set, the pair $\overline{(C, M)}_m$ and
-  $\overline{(C, M')}_m$ is also in the set if $m < n$.
+    M')}_n$ in the set, the pair of truncated traces $\overline{(C,
+    M)}_m$ and $\overline{(C, M')}_m$ is also in the set if $m < n$.
 
@@
   the set a pair of traces $\overline{(C', N)}_m$ and $\overline{(C',
-    N')}_m$ such that $C_0' = C_i$, $N_0 = M_i$, and $N_0' = M_0'$.
+    N')}_m$ such that $C_0' = C_i$, $N_0 = M_i$, and $N_0' = M_i'$.
 
@@
 
-  For every pair of traces $\overline{(C, M)}$ and $\overline{(C,
-    M')}$ in the set, and for every machine state $N_0$ such that
+  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
+    M')}_n$ in the set, and for every machine state $N_0$ such that
   $N_0$ is a variant of $M_0$, there exists in the set a pair of
@@
   $\VIS < \HID$.
+\end{itemize}
 
-\end{itemize}
-\iftext{} \bcp{Maybe prefix closed?}\fi.
+\leo{..and the observations of each trace coincide, right?}
+
 % \bcp{Should we require that it contain all possible starting states and
@@
   $X$.}
-
-\subsection{Noninterference}
-
-NI is the contoured hyperproperty where:
-\begin{itemize}
-\item For each pair of traces in the hyperproperty, all the contours
-  coincide.
-\item For each pair of traces in the hyperproperty, If the initial
-  machine states plus contour are variants, then the observervations
-  of the two traces coincide.
-\end{itemize}
 
@@
 
+\leo{Shouldn't this take a pair (C,M) to a pair (C,M)?}
 Formally, a policy

### 2020-01-18 main.tex
Policies

@@
 
+\leo{I think with the property being just a (special) hyperproperty
+  again, the policy can be ``more detailed'' - including low-level
+  enforcement details like tags, etc. But you should always be able to
+  reconstruct the contoured trace given an initial contour and an
+  initial state. I'm taking that approach here:}
+
+Formally, a policy
+%
+$\pol \in \exists X.\; (\CONTOURS \times \MACHS \rightarrow X)
+                       \times (X \rightarrow \CONTOURS)
+                       \times (\MACHS \times X \rightharpoonup \MACHS \times X)$
+%
+comprises
+\begin{enumerate*}
+  \item a private type $X$ corresponding to an ``implementation'' of the contour,
+  \item an initializer: a function taking a contour and a machine state to an initial value of $X$,
+  \item a projection: a way of reading a contour out of the private type,
+  \item a partial step function of type $(\MACHS \times X) \rightharpoonup X$
+    that, when defined, defines the private state after a step of the machine
+    (based on the $\stepstounderfull$ relation)
+\end{enumerate*}
+That is, a policy takes a starting contour and produces a runtime monitor
+that carries along some private information and uses it to decide whether to
+terminate execution early.
+
+\leo{Example? how detailed?}  For example, for standard
+lockstep-noninterference with only two secrecy levels, the private
+state {\em is} the contour.
+
+A policy $\pol$ {\em guarantees} property $\prop$ if for every pair of
+states $M$ and $M'$ and for every starting contour $C$, the pair of
+contoured traces that you get from the policy is in the property.
+
+\leo{New attempt ends here}
 Informally, a (micro-){\em policy} is an enriched step function that,

### 2020-01-18 main.tex
Comment

@@
 %
-$\pol \in \exists X.\; (\CONTOURS \times \MACHS \rightarrow X)
+$\pol \in \exists X.\; (\CONTOURS \rightarrow X)
                        \times (X \rightarrow \CONTOURS)
@@
   \item a private type $X$ corresponding to an ``implementation'' of the contour,
-  \item an initializer: a function taking a contour and a machine state to an initial value of $X$,
+  \item an initializer: a function taking a contour \leo{and a machine state?} to an initial value of $X$,
   \item a projection: a way of reading a contour out of the private type,
@@
     that, when defined, defines the private state after a step of the machine
-    (based on the $\stepstounderfull$ relation)
+    (based on the $\stepstounderfull$ relation).
+    \leo{Not sure this should be taking $\MACHS$ as input. It should take $X$
+      plus just the instruction being executed? The rest are the tags in $X$}
 \end{enumerate*}

### 2020-01-18 main.tex
Questions

@@
 Generalizing to allow for contours to change throughout the execution,
-nested lockstep noninterference is the contoured lockstep
-hyperproperty where:
+nested lockstep noninterference is the contoured lockstep hyperproperty
+where: \bcp{Which parts of this are common to {\em all} contoured lockstep
+  hyperproperties (i.e., which should be part of the definition of contoured
+  lockstep hyperproperties) and which are specific to (nested) NI?}
 
@@
   For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
-    M')}_n$ in the set, the pair of truncated traces $\overline{(C,
-    M)}_m$ and $\overline{(C, M')}_m$ is also in the set if $m < n$.
+    M')}_n$ in the set, \bcp{where $n$ is the length of the traces,} the
+  pair of truncated traces $\overline{(C, 
+    M)}_m$ and $\overline{(C, M')}_m$ is also in the set if $m < n$ \bcp{or
+    $n = \infty$}.
 
@@
 
-\leo{..and the observations of each trace coincide, right?}
+\leo{..and the observations of each trace coincide, right?}\bcp{Yes, I guess
+  so.}
+
+\bcp{It feels like something is missing here: This definition seems to be
+  requiring that secret data never affect observations, but our tester is
+  ensuring something more: that inaccessible locations are not {\em changed}
+(at least without also changing their tag to make them accessible).  Where
+is that condition being imposed?}
 
@@
 
-\bcp{Not quite enough: Policies and properties needs to know where are the
-  call and return sites.  So what we should do is to talk instead about ``$X$
-  policies'' (i.e., make ``policy'' a type operator) and replace the initial
-  $\CONTOURS$ argument to both policies and properties by an element of
-  $X$.}
+\bcp{Policies and properties need to know where are the call and return
+  sites, and for this they need to be initialized with some
+  application-specific information.  So it seems what we should do is to
+  talk instead about ``$X$ policies'' (i.e., make ``policy'' a type
+  operator) and replace the initial $\CONTOURS$ argument to both policies
+  and properties by an element of $X$.}
 

### 2020-01-18 main.tex
Some responses - still thinking about the X-policy part

@@
   lockstep hyperproperties) and which are specific to (nested) NI?}
+\leo{I'd say contoured lockstep hyperproperties are as above - and the following are
+  the generalization for nested ones.}
 
@@
   For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
-    M')}_n$ in the set, \bcp{where $n$ is the length of the traces,} the
+    M')}_n$ in the set, \bcp{where $n$ is the length of the traces,}\leo{Yes I put that in the definition of traces above, but we could repeat it here} the
   pair of truncated traces $\overline{(C, 
@@
   $\VIS < \HID$.
+
+\item {\em as in noninterfence:}
+
+  For each pair of traces in the hyperproperty, if the initial machine
+  states plus contour are variants, then the observervations of the
+  two traces coincide.
+
 \end{itemize}
-
-\leo{..and the observations of each trace coincide, right?}\bcp{Yes, I guess
-  so.}
 
@@
   is that condition being imposed?}
+\leo{It's not - I just went by the definition of (traditional) noninterference
+  you had and repeated that. I'm thinking about that now}
 

### 2020-01-19 main.tex
a few notational quibbles

@@
 visible component (where $\contour_{i+1} = \VIS$), copying its
-value from $\mach_{i+1}$, and (2) for each
+value from $\mach_{i+1}$\apt{$\mach_i$ ?}, and (2) for each
 hidden component, where $\contour_i = \VIS$, copying its
@@
 \item For each pair of traces in the hyperproperty, all the contours
-  coincide.
-\item For each pair of traces in the hyperproperty, If the initial
-  machine states plus contour are variants, then the observervations
+  coincide.\apt{This is still a bit ambiguous. How about: For each pair of
+    contoured traces in the hyperproperty, the contour is invariant 
+    along each trace.}
+  \item For each pair of contoured traces in the hyperproperty, if the
+    initial elements of the traces are variants, then the observervations
     of the two traces coincide.
@@
     N')}_m$ such that $C_0' = C_i$, $N_0 = M_i$, and $N_0' = M_i'$.
+  \apt{I don't see how we can leave $m$ unspecified: doesn't it need to be as long as possible subject to well-formedness?
+Also, the evolution of $C$ is not determined by the machine, so don't we need $C_j' = C_{i+j}$ for all $0 \leq j < m$ ?
+}
 
@@
   $N_0$ is a variant of $M_0$, there exists in the set a pair of
-  traces $\overline{(C, M)}_m$ and $\overline{(C, N)}_m$.
+  traces $\overline{(C, M)}_m$ and $\overline{(C, N)}_m$.\apt{Same complaint about $n$. And again, $N_0$ defines $\overline{N}$ but it doesn't define $\overline{(C,N)}$. }
 

### 2020-01-19 main.tex
concerns

@@
   $\VIS < \HID$.
+  \apt{Isn't this effectively still a negation in the functor?
+I thought for strategy for describing execution after a Return was to apply suffix closure to the original full trace -- but doesn't this criterion rule out the original full trace?}
 
@@
   For each pair of traces in the hyperproperty, if the initial machine
-  states plus contour are variants, then the observervations of the
+  states plus contour \apt{``elements''} are variants, then the observervations of the
   two traces coincide.
@@
   you had and repeated that. I'm thinking about that now}
+
+\apt{Even if it works for subroutines,
+  the basic idea of truncating traces when visibility increases doesn't 
+  seem to extend to things like coroutines, where we swap between two sets of visible locations on each transfer.}
 

### 2020-01-19 main.tex
fix a previous question

@@
 visible component (where $\contour_{i+1} = \VIS$), copying its
-value from $\mach_{i+1}$\apt{$\mach_i$ ?}, and (2) for each
+value from $\mach_{i+1}$, and (2) for each
 hidden component, where $\contour_i = \VIS$, copying its
-value from $\stateobs_{i+1}$, and (3) if the resulting $\stateobs_{i+1}$ is
+value from $\stateobs_{i+1}$\apt{$\stateobs_i$ ?}, and (3) if the resulting $\stateobs_{i+1}$ is
 identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.

### 2020-01-19 main.tex
splice-based alternative

@@
 
+\subsection{Alternative Definition of Contoured Hyperproperty}
+
+\apt{Just for the record, here's an attempt to write down the splice-based
+  alternative I suggested. I \emph{think} it might work for stack safety
+  (and coroutine safety).
+But there are some problems: (a) it doesn't account for the caller being able to write into the callee's local data (although perhaps there's no good reason to allow this anyway); (b) two calls that just happen to use the same memory space for their frames will share the same contour, which seems wrong.
+I'm wondering if it would be better to use a more
+  explicit and nominal refinement of contours (maybe called ``context''); two
+  contexts might have the same contour but the trace partitioning would be done based on context name. I haven't thought about relating all this to testing...}
+
+  Let $T = \overline{(C,M)}_n$ be a contoured trace of length $n \leq \infty$.
+  Write $T\vert_C$ for the result of
+  filtering $T$ to contain only elements with contour $C$ and 
+  then projecting the trace components.
+  The \emph{contour partition} of $T$ is the set
+  $P(T) = \{(C_{i_1},T\vert_{C_{i_1}}),
+  (C_{i_2},T\vert_{C_{i_2}}),
+  \ldots,(C_{i_d},T\vert_{C_{i_d}})\}$ where
+  $C_{i_1},C_{i_2},\ldots,C_{i_d}$ are the distinct contours appearing in $\overline{C}$.
+  In other words, $P(T)$ is the result of dividing the trace into sections based on contour and then splicing together the sections.
+Note that these traces no longer obey the underlying step function; e.g., there might be a step directly from a call Jal to the return address.
+(We may need to pin down exactly what is legal in these ``gappy'' sequences.)
+
+  Observations are a bit simpler than before. 
+  A \emph{state observation} with respect to a contour $C$ is a map from
+  the machine components visible in $C$ to words.
+  An \emph{observation} with respect to a contour $C$ is 
+  is a nonempty, finite or infinite sequence of
+  state observations with respect to $C$, in which adjacent elements
+  are distinct (``no stuttering''). 
+  In particular, the \emph{observation trace} of
+  a machine state sequence $\overline{M}$ with respect to a contour $C$,
+  written $\obstrace_C(\overline{M})$ is the observation 
+  obtained by removing duplicates from the sequence 
+  $\stateobs_0,\, \stateobs_1,\, \ldots$ where
+  $\stateobs_i(\component) = M_i(\component)$ for all $\component$ such that $C(\component) = \VIS$.
+  We extend observation traces to contour partitions piecewise:
+  if $P(T) = \{(C_1,\overline{M}_1),(C_2,\overline{M}_2),\ldots,(C_d,\overline{M}_d)\}$, then $\obstrace(P(T))$ is $\{(C_1,\obstrace(\overline{M}_1)),
+  (C_2,\obstrace(\overline{M}_2)),\ldots,(C_d,\obstrace(\overline{M}_d))\}$
+
+  Variants are as before.
+  Noninterference (no need for lockstep) with respect to a (single)
+  contour $C$, written $NI_C$, is defined as as a (standard) hyperproperty: it is the set of pairs of (ordinary, but possibly gappy) machine-state traces such that if the initial elements of the traces are variants with respect to $C$ then their observation traces with respect to $C$ are identical.
+
+  A \emph{contoured hyperproperty} (no restriction to lockstep) is still a set of pairs of contoured traces.
+
+  \emph{Contoured noninterference} is the contoured hyperproperty defined as follows: the pair of contoured traces $(T,T')$ is in the hyperproperty if,
+  for each $(C,\overline{M}) \in P(T)$ and $(C',\overline{M'}  ) \in P(T')$,
+  if $C = C'$ then $(\overline{M},\overline{M'}) \in NI_{C}$. (NB: This may
+  be too liberal; perhaps the sets of $C$'s should be the same?)
+
+Standard noninterference falls out in the obvious way when trace contours are invariant. (In this case, the machine-state traces are not gappy.)
+
 \iftext

### 2020-01-22 main.tex
Some nits

--- 2020-01-19 main.tex
@@
 
+Finally, we can parameterize the traces by some initial
+application-specific state to obtain a {\em contoured lockstep
+  $X$-hyperproperty}, where every trace comes with more information.
+\leo{Example, call sites in stack safety}
+
 \subsection{Noninterference}
@@
 
-(Lockstep-)Noninterference is the contoured lockstep hyperproperty
+\leo{For this the X- component is empty right?}
+Lockstep-Noninterference is the contoured lockstep hyperproperty
 where:
 \begin{itemize}
-\item For each pair of traces in the hyperproperty, all the contours
-  coincide.\apt{This is still a bit ambiguous. How about: For each pair of
-    contoured traces in the hyperproperty, the contour is invariant 
-    along each trace.}
+\item For each pair of contoured traces in the hyperproperty, the
+  contour is invariant along each trace.
 \item For each pair of contoured traces in the hyperproperty, if the
@@
 
-Generalizing to allow for contours to change throughout the execution,
-nested lockstep noninterference is the contoured lockstep hyperproperty
-where: \bcp{Which parts of this are common to {\em all} contoured lockstep
-  hyperproperties (i.e., which should be part of the definition of contoured
-  lockstep hyperproperties) and which are specific to (nested) NI?}
-\leo{I'd say contoured lockstep hyperproperties are as above - and the following are
-  the generalization for nested ones.}
+Lockstep-Noninterference is a restricted form of traditional
+noninterference, that is nevertheless useful (e.g. to encode taint
+tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
+At the same time, lockstepness makes it possible to generalize the
+setting and allow for the contours themselves to change throughout the
+execution.
+
+%Generalizing to allow for contours to change throughout the execution,
+%nested lockstep noninterference is the contoured lockstep hyperproperty
+%where: \bcp{Which parts of this are common to {\em all} contoured lockstep
+%  hyperproperties (i.e., which should be part of the definition of contoured
+%  lockstep hyperproperties) and which are specific to (nested) NI?}
+%\leo{I'd say contoured lockstep hyperproperties are as above - and the following are
+%  the generalization for nested ones.}
 
@@
   the set a pair of traces $\overline{(C', N)}_m$ and $\overline{(C',
-    N')}_m$ such that $C_0' = C_i$, $N_0 = M_i$, and $N_0' = M_i'$.
-  \apt{I don't see how we can leave $m$ unspecified: doesn't it need to be as long as possible subject to well-formedness?
-Also, the evolution of $C$ is not determined by the machine, so don't we need $C_j' = C_{i+j}$ for all $0 \leq j < m$ ?
-}
+    N')}_m$ such that for all $0 \leq j < m$, $C_j' = C_{i+j}$, $N_0 =
+  M_i$, and $N_0' = M_i'$, and where $m$ is the maximum index for
+  which $C_i \leq C_{i+j}$ for all $0 \leq j < m$. Here $C \leq C'$ if
+  $\forall \component, C(\component) \leq C'(\component)$ and $\VIS <
+  \HID$.
 
@@
   $N_0$ is a variant of $M_0$, there exists in the set a pair of
-  traces $\overline{(C, M)}_m$ and $\overline{(C, N)}_m$.\apt{Same complaint about $n$. And again, $N_0$ defines $\overline{N}$ but it doesn't define $\overline{(C,N)}$. }
-
-\item {\em contour-wellformedness:}
-
-  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
-    M')}_n$ in the set, $C_0 \leq C_i$ for all $i<n$, where $C \leq
-  C'$ if $\forall \component, C(\component) \leq C'(\component)$ and
-  $\VIS < \HID$.
-  \apt{Isn't this effectively still a negation in the functor?
-I thought for strategy for describing execution after a Return was to apply suffix closure to the original full trace -- but doesn't this criterion rule out the original full trace?}
+  traces $\overline{(C, M)}_n$ and $\overline{(C, N)}_n$.\apt{Same
+    complaint about $n$. And again, $N_0$ defines $\overline{N}$ but
+    it doesn't define $\overline{(C,N)}$. }\leo{Cs should be the same
+    right?}
 

### 2020-01-24 main.tex
Some progress in taint

@@
 
+\leo{Rough draft/take on this section}
+\begin{itemize}
+\item Taint tracking is an interesting property.
+\item It's weaker than traditional noninterference, but easier to enforce so people use it.
+\item \leo{Some of the intuition from below}
+\end{itemize}
+
+\paragraph{Taint Tracking Property}
+
+Given a contour C and pair of traces $\overline{M}$ and $\overline{N}$,
+the pair of contoured trace $\overline(C, M)_n$ and $\overline(C,N)_n$ is in the
+taint tracking contour-lockstep hyperproperty if:
+\begin{itemize}
+\item For all $0 \leq i < m$ the restriction of the trace to its
+  length $i$ prefix is in the standard noninterference property
+  \leo{Defined in the previous section}, where $m$ is the minimum
+  index (or infinity) for which $PC(M_i) \not = PC(N_i)$ or
+  $instr(M_i) \not = instr(N_i)$.
+\end{itemize}
+
+\paragraph{Taint Tracking Policy}
+
+\newcommand*{\clean}{{\em clean}}
+\newcommand*{\tainted}{{\em tainted}}
+\newcommand*{\boring}{{\em boring}}
+\begin{itemize}
+\item We define three {\em tags} \leo{have we defined what these
+  are?}: \clean, \tainted, and \boring. Machine components tagged
+  \clean by the policy represent protected locations that should never
+  be tainted. Components tagged \tainted are either sources of taint
+  (e.g. network inputs) or components that have been influenced by
+  other \tainted ones. Finally \boring components are locations that
+  are allowed to become tainted (perhaps temporarily) but have not
+  yet.
+\end{itemize}
+
 \bcp{This needs to be written, now that we have added the hidden/observable stuff.}

### 2020-01-25 main.tex
Some rough intro story

@@
   that we study several variants...?}
+\leo{Rough story of the paper based on the discussion at POPL:}
+\begin{itemize}
+\item Noninterference is a desirable security property: secret information
+  does not affect pulbic outputs of systems.
+\item Traditionally, it is expressed as a {\em hyperproperty}, a
+  property over pairs of execution traces. \leo{Two machines, differ
+    secret data, run, indist}
+\item The two execution traces are allowed differ arbitrarily in their
+  control flow: branching on secret data can cause the machines to follow
+  different execution paths.
+\item \leo{Why is that bad? Hard to test/visualize? Even though interesting and permissive?}
+\item \leo{Where do we mention contours, if at all?}
+\item 
+\item In this paper we describe a restricted form of hyperproperties:
+  lockstep-hyperproperties, where the two traces execute the same
+  sequence of instructions.
+\item This kind of hyperproperty is still of practical interest: we
+  can use it to encode both taint tracking, heap safety, and stack
+  safety as noninterference properties.
+\item Intuitively, taint tracking enforces noninterference until the
+  program branches on tainted data. Then all bets are off.
+\item On the other hand, stack safety is a monitoring policy that can
+  enforce this lockstepness.
+\item By restricting the class of hyperproperties we care about, we
+  are also able to give a pretty extensional characterization of
+  well-bracketed classification and declassification: what is
+  considered secret can change throughout the execution.
+\item For example, stack safety caller/callee accessible data.  
+\end{itemize}
+
+Contributions:
+\begin{itemize}
+\item \leo{Something about lockstep-hyperproperties as an abstraction}
+\item An extensional characterization of taint tracking.
+\item A novel stack safety policy and property characterized as noninterference.
+\item A testing framework for testing lockstep-hyperproperties (merge with 1?)  
+\end{itemize}
 
@@
   yet.
+\item To implement the taint tracking property as a policy, we
+  initialize a map from components to tags. From this map, we can read
+  a contour in a straightforward manner: \clean is \VIS, the other
+  tags are \HID.
 \end{itemize}
+
+\leo{How do we want to write these rules? Explain .dpl?}
+%//TODO: This is a bug opportunity, right?
+%   // explicit failure for write tainted val to clean addr
+%   storeGrp(   mem == [+Clean], addr == _, val == [+Taint], env == _
+%              -> fail "Tainted data written to clean memory")
+% 
+%   // Propogate taint
+%#ifndef ADD_NO_TAINT_1
+%   ^ arithGrp(env == _, op1 == [+Taint], op2 == _ -> env = env, res = op1)
+%#else
+%//   ^ arithGrp(env == _, op1 == [+Taint], op2 == _ -> env = env, res = {})
+%#endif
+%#ifndef ADD_NO_TAINT_2
+%   ^ arithGrp(env == _, op1 == _, op2 == [+Taint] -> env = env, res = op2)
+%#else
+%//   ^ arithGrp(env == _, op1 == _, op2 == [+Taint] -> env = env, res = {})
+%#endif
+%   ^ mulDivRemGrp(code == _, env == _, op1 == [+Taint], op2 == _ -> env = env, res = op1)
+%   ^ mulDivRemGrp(code == _, env == _, op1 == _, op2 == [+Taint] -> env = env, res = op2)
+%   ^ immArithGrp(env == _, op1 == [+Taint] -> env = env, res = op1)
+%#ifndef NO_MEM_TAINT_CHECK_LOAD
+%   ^ loadGrp(mem == [+Taint], addr == _, env == _ -> env = env, res = mem )
+%#else
+%// No mem check
+%#endif
+%#ifndef NO_ADDR_TAINT_CHECK_LOAD
+%   ^ loadGrp(mem == _, addr == [+Taint], env == _ -> env = env, res = addr )
+%#else
+%// THIS WAS AN ACTUAL BUG! Addres check ommitted
+%#endif
+%#ifndef NO_VAL_TAINT_CHECK_STORE
+%//NOT A BUG! We never write tainted data to clean memory because of the first rule
+%   ^ storeGrp(mem == _, addr == _, val == [+Taint], env == _ -> env = env, mem = val)
+%#else
+%// NO CHECK
+%#endif
+%#ifndef NO_MEM_TAINT_CHECK_STORE
+%   ^ storeGrp(mem == [+Taint], addr == _, val == _, env == _ -> env = env, mem = mem )
+%#else
+%// NO CHECK
+%#endif
+%#ifndef NO_ADDR_TAINT_CHECK_STORE
+%   ^ storeGrp(mem == _, addr == [+Taint], val == _, env == _ ->
+%     fail "Can't write through a tainted pointer")
+%#else
+%// Could also taint mem with addr. Also an actual bug (but one that I made).
+%// THIS WAS AN ACTUAL BUG! Addres check ommitted
+%#endif
+%   // keep clean mem stays clean always
+%   ^ storeGrp(   mem == [+Clean], addr == _, val == _, env == _ -> env = env, mem = mem)
 

### 2020-01-29 main.tex
Rough small intro

@@
 
-\iftext
-\IEEEauthorblockN{Leonidas Lampropoulos}
-\IEEEauthorblockA{\textit{University of Maryland \bcp{right?}}
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID
-  }
-\and
-\author{\IEEEauthorblockN{Roberto Blanco}
-\IEEEauthorblockA{\textit{\bcp{affiliation?}}
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID
-}
-\and
-\IEEEauthorblockN{Benjamin C. Pierce}
-\IEEEauthorblockA{\textit{University of Pennsylvania}
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID
-  }
-\and
-\IEEEauthorblockN{Andrew Tolmach}
-\IEEEauthorblockA{\textit{Portland State University}
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID
-}
-}
-\fi
+% I get a weird error, so I commented this out for now
+%\iftext
+%\IEEEauthorblockN{Leonidas Lampropoulos}
+%\IEEEauthorblockA{\textit{University of Maryland}}
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID
+%\and
+%\author{\IEEEauthorblockN{Roberto Blanco}
+%\IEEEauthorblockA{\textit{\bcp{affiliation?}}
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID
+%}
+%\and
+%\IEEEauthorblockN{Benjamin C. Pierce}
+%\IEEEauthorblockA{\textit{University of Pennsylvania}
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID
+%  }
+%\and
+%\IEEEauthorblockN{Andrew Tolmach}
+%\IEEEauthorblockA{\textit{Portland State University}
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID
+%}
+%}
+%\fi
 
@@
 
-\bcp{How about coining the term ``lockstep noninterference'' and then saying
-  that we study several variants...?}
-\leo{Rough story of the paper based on the discussion at POPL:}
+{\em Hyperproperties}, i.e. properties of pairs\leo{sets?} of execution traces,
+are the standard way of expressing security properties\cn. In
+particular, {\em noninterference} is commonly phrased as such a
+property: given two system states that seem identical to an observer
+with low clearance, their execution traces should also be
+indistinguishable\cn.
+
+In this paper we focus on a well-behaved closed subset of
+hyperproperties, where the execution traces remain in {\em lockstep}:
+the traces execute identical sequences of instructions. While such
+lockstep hyperproperties appear restrictive at first, they are still
+of practical interest, as they will allow us to give an extensional
+characterization of security policies such as taint tracking or stack
+safety, which have been traditionally approached
+intentionally. Moreover, this setting admits a principled treatment of
+well-bracketed classification and declassification. Finally, a
+fortunate side-effect is that because of the lockstep nature of the
+execution traces, lockstep-hyperproperties are far easier to test than
+their more general counterparts.
+
+Our contributions are:
 \begin{itemize}
-\item Noninterference is a desirable security property: secret information
-  does not affect pulbic outputs of systems.
-\item Traditionally, it is expressed as a {\em hyperproperty}, a
-  property over pairs of execution traces. \leo{Two machines, differ
-    secret data, run, indist}
-\item The two execution traces are allowed differ arbitrarily in their
-  control flow: branching on secret data can cause the machines to follow
-  different execution paths.
-\item \leo{Why is that bad? Hard to test/visualize? Even though interesting and permissive?}
-\item \leo{Where do we mention contours, if at all?}
-\item 
-\item In this paper we describe a restricted form of hyperproperties:
-  lockstep-hyperproperties, where the two traces execute the same
-  sequence of instructions.
-\item This kind of hyperproperty is still of practical interest: we
-  can use it to encode both taint tracking, heap safety, and stack
-  safety as noninterference properties.
-\item Intuitively, taint tracking enforces noninterference until the
-  program branches on tainted data. Then all bets are off.
-\item On the other hand, stack safety is a monitoring policy that can
-  enforce this lockstepness.
-\item By restricting the class of hyperproperties we care about, we
-  are also able to give a pretty extensional characterization of
-  well-bracketed classification and declassification: what is
-  considered secret can change throughout the execution.
-\item For example, stack safety caller/callee accessible data.  
-\end{itemize}
-
-Contributions:
-\begin{itemize}
-\item \leo{Something about lockstep-hyperproperties as an abstraction}
-\item An extensional characterization of taint tracking.
-\item A novel stack safety policy and property characterized as noninterference.
-\item A testing framework for testing lockstep-hyperproperties (merge with 1?)  
+\item We present a mathematical abstraction of a class of properties
+  called {\em lockstep hyperproperties}, that allow for intuitive
+  characterizations of interesting security policies.
+\item We give an extensional characterization of taint tracking as
+  lockstep-noninterference.
+\item We introduce a novel stack safety property that allows for
+  the notion of observability to change dynamically throughout
+  execution.
+\item We develop a testing framework for testing such
+  lockstep-hyperproperties, implemented as a parametric co-processor
+  for RISC-V.
 \end{itemize}

### 2020-01-29 main.tex
Little things

@@
 
-In this paper we focus on a well-behaved closed subset of
+In this paper we focus on a well-behaved closed\bcp{???} subset of
 hyperproperties, where the execution traces remain in {\em lockstep}:
@@
 safety, which have been traditionally approached
-intentionally. Moreover, this setting admits a principled treatment of
+intentionally. \bcp{reword, maybe using ``nested'':}Moreover, this setting
+admits a principled treatment of
 well-bracketed classification and declassification. Finally, a
@@
   characterizations of interesting security policies.
-\item We give an extensional characterization of taint tracking as
-  lockstep-noninterference.
+\item We give a novel {\em extensional} characterization of taint tracking
+  as a form of lockstep-noninterference.
 \item We introduce a novel stack safety property that allows for
@@
 \item Daniel's tainting paper (and some refs cited there)
+      \cite{DBLP:conf/eurosp/SchoepeBPS16} 
 \item Nick and Andre's papers (and other micropolicy papers)
@@
 \end{itemize}
+
+\item Stuff about dynamically changing labels (see footnote in Testing
+Noninterference Quickly paper)
 

### 2020-01-30 main.tex
stab at stack protection as L-S NI; don't see story for integrity

@@
 $\mach(\component) = \mach'(\component)$.
+\apt{Um..this seems to conflate visibility (\VIS,\HID) with
+  security/privacy level (\HIGH,\LOW).  Don't we want to distinguish these?}
 
@@
 
-These properties are essentially orthogonal: we could enforce WBCF without LSE, and variants of
+These properties are essentially orthogonal: one can certainly have WBCF without LSE, and variants of
 LSE could apply to control constructs beyond simple procedure calls (e.g. exceptions, coroutines, threads, monitors, etc.).
 
-In terms of NI....
-\apt{Perhaps the NI-like property is just LSE, but it holds only if WBCF holds?}
+Under the assumption that WBCF holds, we can describe LSE as an instance of Lockstep-Noninterference, where the contour at
+each step maps the current stack frame to \VIS\ and the remainder of the stack to \HID\. 
+The observations on which noninterference is defined will (necessarily)  be made on the values in the current stack frame.
+\apt{Again, we're conflating \VIS\ vs. \HID\ with \LOW\ vs \HIGH\; should probably be teased apart.}
+LSE has two aspects: secrecy (callees cannot read caller variables) and integrity (callees cannot alter caller variables).
+Secrecy is enforced by the variant-closure clause of L-N: since caller data is \HID, changing it cannot alter
+the callee's observable behavior.  Integrity is enforced why??? \apt{I've lost confidence that it is.}
 

### 2020-01-30 main.tex
Updated story

--- 2020-01-29 main.tex
@@
 
-{\em Hyperproperties}, i.e. properties of pairs\leo{sets?} of execution traces,
-are the standard way of expressing security properties\cn. In
-particular, {\em noninterference} is commonly phrased as such a
-property: given two system states that seem identical to an observer
-with low clearance, their execution traces should also be
-indistinguishable\cn.
+{\em Hyperproperties}, i.e. properties of pairs\leo{sets?} of
+execution traces, are the standard way of expressing security
+properties\cn. In particular, {\em noninterference} is commonly
+phrased as such a property: given two system states that seem
+identical to an observer with low clearance, their execution traces
+should also be indistinguishable\cn.
 
@@
 hyperproperties, where the execution traces remain in {\em lockstep}:
-the traces execute identical sequences of instructions. While such
-lockstep hyperproperties appear restrictive at first, they are still
-of practical interest, as they will allow us to give an extensional
-characterization of security policies such as taint tracking or stack
-safety, which have been traditionally approached
-intentionally. Moreover, this setting admits a principled treatment of
-well-bracketed classification and declassification. Finally, a
-fortunate side-effect is that because of the lockstep nature of the
-execution traces, lockstep-hyperproperties are far easier to test than
-their more general counterparts.
+the traces execute identical sequences of instructions.
+%
+This allows us to give a clean extensional characterization of taint
+tracking as a noninterference property: intuitively, taint tracking is
+a weak lockstep-hyperproperty that guarantees noninterference as long
+as the execution traces remain in sync, but then allows for arbitrary
+behavior after branching on a secret.
+%
+We also describe a strong version of lockstep hyperproperties, where
+the security policy enforces the lockstepness. \leo{Something about
+  full separation}
+%
+Finally, we generalize strong lockstep hyperproperties to allow for
+\leo{what's the correct adjective here? principled? nested?
+  restricted?}  changes to the observer clearance, and use this
+generalization to characterize a novel stack safety property as
+a variation of noninterference.
+
+As it happens, lockstep hyperproperties are also significantly more
+amenable to testing compared to traditional hyperproperties. Since
+executions remain in lockstep, there is no possibility of \leo{missing
+  bugs by never coming back down or a need to find an invariant strong
+  enough to prove things}. We have implemented enforcement mechanisms
+for all three policies for RISC-V and tested them.
 

### 2020-01-30 main.tex

--- 2020-01-30 main.tex
@@
 hyperproperties, where the execution traces remain in {\em lockstep}:
-the traces execute identical sequences of instructions. While such
-lockstep hyperproperties appear restrictive at first, they are still
-of practical interest, as they will allow us to give an extensional
-characterization of security policies such as taint tracking or stack
-safety, which have been traditionally approached
-intentionally. \bcp{reword, maybe using ``nested'':}Moreover, this setting
-admits a principled treatment of
-well-bracketed classification and declassification. Finally, a
-fortunate side-effect is that because of the lockstep nature of the
-execution traces, lockstep-hyperproperties are far easier to test than
-their more general counterparts.
+the traces execute identical sequences of instructions.
+%
+This allows us to give a clean extensional characterization of taint
+tracking as a noninterference property: intuitively, taint tracking is
+a weak lockstep-hyperproperty that guarantees noninterference as long
+as the execution traces remain in sync, but then allows for arbitrary
+behavior after branching on a secret.
+%
+We also describe a strong version of lockstep hyperproperties, where
+the security policy enforces the lockstepness. \leo{Something about
+  full separation}
+%
+Finally, we generalize strong lockstep hyperproperties to allow for
+\leo{what's the correct adjective here? principled? nested?
+  restricted?}  changes to the observer clearance, and use this
+generalization to characterize a novel stack safety property as
+a variation of noninterference.
+
+As it happens, lockstep hyperproperties are also significantly more
+amenable to testing compared to traditional hyperproperties. Since
+executions remain in lockstep, there is no possibility of \leo{missing
+  bugs by never coming back down or a need to find an invariant strong
+  enough to prove things}. We have implemented enforcement mechanisms
+for all three policies for RISC-V and tested them.
 

### 2020-02-01 main.tex
Writing down the stack safety formalization based on yesterdays talk

@@
 
+\begin{itemize}
+\item \leo{Not even sure this is true anymore}
+\item Noninterference is a security property combining {\em
+  confidentiality} and {\em integrity}: secret data should influence
+  public outputs and low integrity data should \leo{finish in a pretty
+    way}.
+\item In this paper we present a stack safety property as a variation
+  of noninterference.
+\end{itemize}
+
 {\em Hyperproperties}, i.e. properties of pairs\leo{sets?} of
@@
 
-A {\em contour}
-$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\VIS,\HID\}$ is a map
-from machine components to ``observabilities,'' where
-$\VIS$ marks locations whose contents are observable from the outside (by
-the ``attacker,'' e.g., by DMA) and $\HID$ marks internal ``hidden''
-state that cannot be observed directly (e.g., processor registers and memory
-addresses for temporary variables will usually be hidden).
-
-A {\em contoured trace} of length $n$, where $n$ can be finite or
-infinite, is a length-$n$ sequence $\seq{(\contour,\mach)}_n =
-(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$ of (contour,
-machine state) pairs in which adjacent machine states are related by
-$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
-\mach_i$ for each $i > 0$ in the sequence.
-
-\subsection{Observations}
-
-A {\em component observation} is either a word or the special value $\bot$.
-A {\em state observation} is a map from machine components to component
-observations.
-
-An {\em observation} is a nonempty, finite or infinite sequence of state
-observations in which (1) adjacent elements are always distinct (``no
-stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
-the the $j^{th}$ state observation in the sequence must also be mapped to
-$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
-component's value changes from $\bot$ to anything else, it never changes
-back to $\bot$ again).
-
-In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
-from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
-observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
-follows:
-\begin{itemize}
-\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
-to its value in $\mach_0$ and every hidden
-component to $\bot$; 
-\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
-visible component (where $\contour_{i+1} = \VIS$), copying its
-value from $\mach_{i+1}$, and (2) for each
-hidden component, where $\contour_i = \VIS$, copying its
-value from $\stateobs_{i+1}$\apt{$\stateobs_i$ ?}, and (3) if the resulting $\stateobs_{i+1}$ is
-identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
-\iftext\bcp{Too informal!!}\fi
-\end{itemize}
+\leo{We might need to change the intuition behind some of these things.}
+A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
+\{\HIGH,\LOW\} \times \{\HI, \LO\}$ is a map from machine components
+to confidentiality and integrity levels. $\LOW$ marks locations whose
+contents are observable from the outside (by the ``attacker,'' e.g.,
+by DMA) and $\HIGH$ marks internal ``hidden'' state that cannot be
+observed directly (e.g., processor registers and memory addresses for
+temporary variables will usually be hidden). Similarly, $\LO$
+locations correspond to low-integrity components (e.g. locations that
+can be written to by the network), while $\HI$ locations should be
+protected from such flows.
+
+% 
+%\subsection{Observations}
+% 
+%A {\em component observation} is either a word or the special value $\bot$.
+%A {\em state observation} is a map from machine components to component
+%observations.
+% 
+%An {\em observation} is a nonempty, finite or infinite sequence of state
+%observations in which (1) adjacent elements are always distinct (``no
+%stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
+%the the $j^{th}$ state observation in the sequence must also be mapped to
+%$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
+%component's value changes from $\bot$ to anything else, it never changes
+%back to $\bot$ again).
+% 
+%In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
+%from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
+%observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
+%follows:
+%\begin{itemize}
+%\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
+%to its value in $\mach_0$ and every hidden
+%component to $\bot$; 
+%\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
+%visible component (where $\contour_{i+1} = \VIS$), copying its
+%value from $\mach_{i+1}$, and (2) for each
+%hidden component, where $\contour_i = \VIS$, copying its
+%value from $\stateobs_{i+1}$\apt{$\stateobs_i$ ?}, and (3) if the resulting $\stateobs_{i+1}$ is
+%identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
+%\iftext\bcp{Too informal!!}\fi
+%\end{itemize}
 
@@
 
-\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
+%\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
 
@@
 
-A {\em contoured hyperproperty} is a set of pairs of contoured
-traces.
-
+A {\em trace} of length $n$, where $n$ can be finite or infinite, is a
+length-$n$ sequence $\seq{\mach}_n = \mach_0,\, \mach_1,\ldots$ of
+machine states in which adjacent machine states are related by
+$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
+\mach_i$ for each $i > 0$ in the sequence.
+
+A {\em contoured hyperproperty} is a set of triples: a contour along
+with a pair of traces.
+
+\leo{No longer needed:?}
 A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
-such that:
-
-\leo{Informally first, I don't like it too much:}
-\begin{itemize}
-\item each trace in every pair of a contoured trace in the set have the
-  same length, and
-\item each pair of elements in every pair of a contoured trace in the
-  set have the same contour.
-\end{itemize}
-
-\leo{Formally:}
-\begin{itemize}
-\item {\em same-length:}
-
-  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
-    M')}_m$ in the set, $n = m$.
-
-\item {\em contour-lockstep:}
-
-  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
-    M')}_n$ in the set, $C_i = C_i'$ for all $i<n$.
-\end{itemize}
+such that each trace in every pair of traces in the set executes the
+same instruction sequence.
 
@@
 application-specific state to obtain a {\em contoured lockstep
-  $X$-hyperproperty}, where every trace comes with more information.
-\leo{Example, call sites in stack safety}
+  $X$-hyperproperty}, where every trace comes with more information,
+for example call sites in stack safety.
 
@@
 $\mach'$ whose components agree with $\mach$ on all components marked
-$\VIS$ by $\contour$---i.e., if $\contour(\component) = \VIS$, then
+$\LOW$ by $\contour$---i.e., if $\contour(\component) = \LOW$, then
 $\mach(\component) = \mach'(\component)$.
-\apt{Um..this seems to conflate visibility (\VIS,\HID) with
-  security/privacy level (\HIGH,\LOW).  Don't we want to distinguish these?}
-
-\leo{Normal lockstep noninterference:}
-
-\leo{For this the X- component is empty right?}
-Lockstep-Noninterference is the contoured lockstep hyperproperty
-where:
-\begin{itemize}
-\item For each pair of contoured traces in the hyperproperty, the
-  contour is invariant along each trace.
-\item For each pair of contoured traces in the hyperproperty, if the
-  initial elements of the traces are variants, then the observervations
-  of the two traces coincide.
-\end{itemize}
-
-\leo{Nested lockstep noninterference:}
-
-Lockstep-Noninterference is a restricted form of traditional
-noninterference, that is nevertheless useful (e.g. to encode taint
-tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
-At the same time, lockstepness makes it possible to generalize the
-setting and allow for the contours themselves to change throughout the
-execution.
+
+\leo{Baseline noninterference:}
+
+\leo{Is this noninterference or is it noninterference + integrity.
+  I'm confused about terminology}
+
+Our baseline security property is noninterference.
+\begin{itemize}
+\item {\em Confidentiality:}
+
+  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$, if
+  the initial elements of the traces are variants, then the final
+  states of the traces are also variants.
+
+\item {\em Integrity: }
+
+  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$,
+  and for every high-integrity component $\component$ for which
+  $\contour(\component) = \HI$, $M_0(\component) = M_n(\component)$
+  and $M'_0(\component) = M'_{n'}(\component)$.
+\end{itemize}
+\leo{In the terminology of the original hyperproperty paper,
+  confidentiality is a hyperproperty, integrity is a trace property
+  that holds for every trace in a hyperproperty. This is a stronger
+  version of integrity than DLM for example}.
+% 
+%\leo{Nested lockstep noninterference:}
+% 
+%Lockstep-Noninterference is a restricted form of traditional
+%noninterference, that is nevertheless useful (e.g. to encode taint
+%tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
+%At the same time, lockstepness makes it possible to generalize the
+%setting and allow for the contours themselves to change throughout the
+%execution.
 
@@
 %  the generalization for nested ones.}
+
+\section{Separation is noninterference$+$}
+
+\leo{To warm up, a super simple separation policy.}
+\begin{itemize}
+\item Any instruction can serve as a final instruction.\leo{What about
+  instructions the are halted by the monitor?}
+\item For forward compatibility we could treat the first unmatched return as
+  a halt instruction.
+\item The instruction memory is marked $\LOW \times \HI$. That means
+  it is observable and can't change.
+\item The memory space is split in two. One part is marked $\LOW
+  \times \LO$: this is the public part (or the one that can be
+  accessed by the current process e.g.). The other part is marked
+  $\HIGH \times \HI$: it should be completely inaccessible to the
+  current process.
+\item Separation is noninterference:
+  \leo{Exactly the definition above}
+\item Depending on what we do with final instructions, the property
+  could be parameterized by the ``call/return sites''.
+\end{itemize}
+\leo{(The $+$ part in the title is a hint that one thing that gets added is that
+the PCs of corresponding traces always stay in lock step because of integrity?)}
+
+\section{Stack Safety is noninterference$+$}
+
+\leo{Build up to the interesting thing:}
+\begin{itemize}
+\item The first unmatched return serves as a halt.
+\item The instruction memory is again marked $\LOW \times \HI$. That
+  means it is observable (both traces have the same instructions!) and
+  can't change.
+\item At the beginning of time, the entire data memory is marked $\LOW
+  \times \LO$. That makes for a very unintersting property as all
+  variants are necessarily identical.
+\item BUT, we close the hyperproperty as follows:
+  \begin{itemize}
+  \item We need the state description stuff to enforce WBCF and be
+    able to talk about ``first unmatched returns''. That is also a
+    trace property.
+  \item After every call, we keep the part of the (left-hand, wlog
+    due to symmetry) trace until its first unmatched return.
+  \item We construct a new contour at that point: after every call,
+    the local state of the caller becomes inaccessible (marked
+    $\HIGH \times \HI$) until a subsequent return.
+  \item For every variant of the new initial state with respect to the
+    new contour, the contour plus pair of traces is in the
+    hyperproperty.
+  \end{itemize}  
+\end{itemize}
+
+\section{Old Stuff}
 

### 2020-02-02 main.tex
Comments

@@
 Concretely, we have in mind a conventional machine architecture with a
-RISC-V processor, perhaps extended with some kind of hardware-enforced
-security policy, but we won't need any RISC-V-specific details till we get
-to \cref{impl}.
+standard (RISC-V) processor extended with some kind of (static or dynamic,
+software- or hardware-enforced) security monitor, but we won't need to make
+any specific choices about details until we get to \cref{impl}.
 
@@
 A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
-\{\HIGH,\LOW\} \times \{\HI, \LO\}$ is a map from machine components
+\{\HIGH,\LOW\} \times \{\HI, \LO\}$\bcp{We need more mnemonic names for
+  these (that help us keep straight when we're talking about confidentiality
+  and when integrity), both in the tex and in the way they are displayed.}
+is a map from machine components 
 to confidentiality and integrity levels. $\LOW$ marks locations whose
@@
 locations correspond to low-integrity components (e.g. locations that
-can be written to by the network), while $\HI$ locations should be
+can be written to \bcp{directly or indirectly} by the network), while $\HI$ locations should be
 protected from such flows.
+
+\bcp{A random thought about integrity: One thing that may be helping confuse
+  us about integrity is that we've gone back and forth between ``eager'' and
+  ``lazy'' versions of various micropolicies, where the ``lazy'' ones do
+  allow high-integrity data to be overwritten, but only in a way that it
+  cannot be read again later (e.g., the lazy heap policy guarantees that, if
+  inaccessible memory is overwritten, then its color will change).  The lazy
+point of view is much more subtle as an integrity property.  So, to keep our
+heads straight here, we should be thinking only about eager properties in
+the first instance.}
 
@@
 
-A {\em trace} of length $n$, where $n$ can be finite or infinite, is a
+A {\em trace} of length $n$, where $n$ can be finite or infinite\bcp{awkward}, is a
 length-$n$ sequence $\seq{\mach}_n = \mach_0,\, \mach_1,\ldots$ of
@@
 
-A {\em contoured hyperproperty} is a set of triples: a contour along
-with a pair of traces.
+A {\em contoured hyperproperty} is a set of triples, each consisting of a
+contour along with a pair of traces.
 
@@
 
-Finally, we can parameterize the traces by some initial
-application-specific state to obtain a {\em contoured lockstep
-  $X$-hyperproperty}, where every trace comes with more information,
-for example call sites in stack safety.
+Finally, we can parameterize the traces by some initial application-specific
+state to obtain a {\em contoured lockstep $X$-hyperproperty}, where every
+trace comes with more information, for example call sites in stack safety.
+\bcp{If we use this later, we need to say it more formally to be clear about
+  what we mean---e.g., I had thought both the traces and the contour would
+  be functions of $X$!}
 
@@
   the initial elements of the traces are variants, then the final
-  states of the traces are also variants.
+  states of the traces are also variants.  \bcp{What does ``final''mean for
+    infinite traces?  What about pairs where one is finite and the other
+    infinite?}
 
@@
 \begin{itemize}
-\item Any instruction can serve as a final instruction.\leo{What about
-  instructions the are halted by the monitor?}
+\item Any instruction can serve as a final instruction.  \bcp{My picture
+  here was that, for this property, the $X$ parameter would be, say, the
+  address where the machine is considered to halt. (I guess the picture is
+  that we attach probes to the PC and to the rest of memory, and we read out
+  the memory when the PC reaches this address.  But when I write it down
+  that way it seems pretty awkward.  Maybe we should just assume that the
+  machine has a HALT instruction or some other way of signaling that it is
+  done---i.e., abstractly, simply take the step function as a partial
+  function.  We will need to put a footnote somewhere acknowledging that
+  this may be a bit of an abstraction from real modern processors, but the
+  simple separation policy is an abstraction anyway!)}
+
+\leo{What about
+  instructions the are halted by the monitor?}  \bcp{Not sure what the
+  question is here, but note that an instruction halted by the monitor will
+  not appear in the trace.}
 \item For forward compatibility we could treat the first unmatched return as
@@
 \item The instruction memory is marked $\LOW \times \HI$. That means
-  it is observable and can't change.
+  it is observable and can't change.  \bcp{What does this assumption mean,
+    precisely?  (I.e., what is ``the instruction memory''?)}
 \item The memory space is split in two. One part is marked $\LOW
@@
 the PCs of corresponding traces always stay in lock step because of integrity?)}
+\bcp{The section titles are not working so well, now that we're being a bit
+  less abstract about our definitions.}
 
@@
 \begin{itemize}
-\item The first unmatched return serves as a halt.
+\item The first unmatched return serves as a halt.  \bcp{What does that mean
+formally?}
 \item The instruction memory is again marked $\LOW \times \HI$. That
@@
   can't change.
-\item At the beginning of time, the entire data memory is marked $\LOW
+\item At the beginning of time\bcp{Not clear what ``beginning of time''
+  means when we're talking about hyperproperties---the intuition of time is
+  relevant to the traces within the hyperproperty, but not to the
+  hyperproperty itself.  Simpler, I think, to view the hyperproperty as a
+  starting set of traces (where the whole machine state [registers and PC too,
+  right?] is marked $\LOW$), closed under an operator that takes a
+  contour+trace+trace and snips out all the subtraces between a call and its
+return.  (But: eek---see below.)}, the entire data memory is marked $\LOW
   \times \LO$. That makes for a very unintersting property as all
@@
 \end{itemize}
+
+\bcp{Eeek.  Just realized why we absolutely need the ``lockstep'' part!
+  The closure operator that snips out a ``subtrace'' between a call and the
+  corresponding return has to do the snipping on {\em both} traces at once,
+  and this only makes sense if they are executing in lockstep.}
 

### 2020-02-02 main.tex
Comments

@@
 \begin{itemize}
-\item {https://arxiv.org/pdf/1811.02787.pdf}
-\item {https://arxiv.org/pdf/1902.05283.pdf}
-\item {http://cs.au.dk/~lask/papers/esop2018.pdf}
-\item {https://dl.acm.org/doi/fullHtml/10.1145/3363519}
+\item \url{https://arxiv.org/pdf/1811.02787.pdf}
+\item \url{https://arxiv.org/pdf/1902.05283.pdf}
+\item \url{http://cs.au.dk/~lask/papers/esop2018.pdf}
+\item \url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}
 \end{itemize}
@@
 mechanization in Iris (also in the eyes of the Iris people).
+
+\item \url{https://www.cis.upenn.edu/~stevez/papers/LZ05a.pdf}
+
+\item We'd better say something about the relation to the original
+hyperproperties paper, no?
 \end{itemize}
@@
 \item Could we also implement all this using Cheri capabilities?
+\item A more abstract presentation, perhaps unifying our treatments of
+confidentiality and integrity?
 \end{itemize}

### 2020-02-02 main.tex
Last comments for now

@@
 {\em Hyperproperties}, i.e. properties of pairs\leo{sets?} of
-execution traces, are the standard way of expressing security
+execution traces\footnote{\bcp{I'm not certain we're using the word
+    correctly.  E.g., ``A hyperproperty is a set of properties [Clarkson and Schneider 2008, 2010]''}}, are the standard way of expressing security
 properties\cn. In particular, {\em noninterference} is commonly
@@
 
-The {\em underlying} step function, $\mathord{\stepstounderfull} \in \MACHS
-\rightarrow \MACHS$, embodies the standard RISC-V ISA semantics.
+The {\em underlying} step function,
+$\mathord{\stepstounderfull} \in \MACHS \rightarrow \MACHS$, embodies the
+standard RISC-V ISA semantics.  \bcp{It's probably clearer not to put RISCV
+  in the notation---for most of the paper, the fact that it's a RISCV really
+  doesn't matter.  Moreover, I think we want to make it a partial function,
+  for simplicity of definitions in this part, and if we insist on RISCV here
+  this will confuse people that know it doesn't have a HALT instruction.}
 

### 2020-02-02 main.tex
Note

@@
 execution traces\footnote{\bcp{I'm not certain we're using the word
-    correctly.  E.g., ``A hyperproperty is a set of properties [Clarkson and Schneider 2008, 2010]''}}, are the standard way of expressing security
+    correctly.  E.g., ``A hyperproperty is a set of properties [Clarkson and
+    Schneider 2008, 2010]''} \bcp{After more reading, I am still wondering
+    if we're using the word the right way.  I think we mean something much
+    less general---just a relation on traces.  So I think we should just say
+    ``contoured relation'' (and maybe ``lockstep relation'').}}, are the
+standard way of expressing security 
 properties\cn. In particular, {\em noninterference} is commonly

### 2020-02-02 main.tex
Few notes about related work

@@
 \begin{itemize}
-\item Daniel's tainting paper (and some refs cited there)
-      \cite{DBLP:conf/eurosp/SchoepeBPS16}
 \item Nick and Andre's papers (and other micropolicy papers)
+\cite{DBLP:conf/sp/RoesslerD18}
+
+This is the basis of our run-time monitor in Section{XXX}.  Specifically,
+our micropolicy is a simplified variant of their ... (I think it is the
+Return-Address Protection + Static Authorities + Depth Isolation policy, but
+need to double check---be careful about laziness.)  Essentially, our work
+arose as an attempt to formalize the security property enforced by something
+like one of their stack policies.
+
 \item StkTokens: Skorstengaard, Devriese, Birkedal
@@
 
-\item Simple Noninterference by Normalization {http://nachivpn.me/nibnbe.pdf}
+\item Simple Noninterference by Normalization \url{http://nachivpn.me/nibnbe.pdf}
 
@@
 
-\item \url{https://www.cis.upenn.edu/~stevez/papers/LZ05a.pdf}
-
+\item \url{https://www.cis.upenn.edu/~stevez/papers/LZ05a.pdf} ??
+
+\item Not so closely related?
+\begin{itemize}
+\item Daniel's tainting paper (and some refs cited there)
+      \cite{DBLP:conf/eurosp/SchoepeBPS16} -- not needed if we postpone
+      dealing with tainting till another paper!
 \item We'd better say something about the relation to the original
-hyperproperties paper, no?
+hyperproperties paper, no?  (No, not if we drop the word hyperproperty!) 
+\end{itemize}
 \end{itemize}

### 2020-02-03 main.tex
Related work

@@
 
+\item Hritcu et al's work on ``Testing Noninterference
+Quickly''~\cite{DBLP:journals/jfp/HritcuLSADHPV16} laid the
+foundation of the random testing strategy that we use here.  In particular,
+that paper explored testing strategies based on both standard ``End-to-end
+noninterference''---a pleasantly extensional and easy-to-define property
+that leads to somewhat inefficient testing because process that misbehaves
+while its PC is labeled ``high'' must return to a low PC before the
+misbehavior can be identified as such---and ``Single-step
+noninterference''---a more intricate property related to Goguen and
+Meseguer's ``unwinding conditions''~\cite{GoguenMeseguer85} that ``fails
+faster'' and is thus easier to test.  \bcp{Explain the relation of the
+  latter property to our stack testing strategy.}
+
 \item StkTokens: Skorstengaard, Devriese, Birkedal

### 2020-02-03 main.tex
More related work comments

@@
 
+\item Azevedo de Amorim's presentation of heap safety as a noninterference
+property~\cite{DBLP:conf/post/AmorimHP18} was a key inspiration for our
+definition of stack safety.  (Details...)
+
 \item StkTokens: Skorstengaard, Devriese, Birkedal
 \begin{itemize}
-\item \url{https://arxiv.org/pdf/1811.02787.pdf}
-\item \url{https://arxiv.org/pdf/1902.05283.pdf}
+\item POPL \cite{skorstengaard2019stktokens}
+\item tech report with proofs: \cite{skorstengaard2018stktokens}
+\url{https://arxiv.org/pdf/1811.02787.pdf}
+\item TR with proofs for ``Reasoning about a machine...'', Feb 2019:
+\url{https://arxiv.org/pdf/1902.05283.pdf} 
 \item \url{http://cs.au.dk/~lask/papers/esop2018.pdf}
-\item \url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}
+\item Also: {\em Reasoning about a Machine with Local Capabilities: Provably
+  Safe Stack and Return Pointer Management} \cite{10.1145/3363519}
+\url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}
+
+\bcp{I'm nervous that these papers may be viewed as subsuming most of
+  ours---in particular, their security property is formulated as a logical
+  relation, i.e. a relational property!  We really need to dig into the
+  details (which are unfortunately rather technical).}
 \end{itemize}

### 2020-02-03 main.tex
Some work before other meetings

@@
 
-The {\em underlying} step function,
-$\mathord{\stepstounderfull} \in \MACHS \rightarrow \MACHS$, embodies the
-standard RISC-V ISA semantics.  \bcp{It's probably clearer not to put RISCV
-  in the notation---for most of the paper, the fact that it's a RISCV really
-  doesn't matter.  Moreover, I think we want to make it a partial function,
-  for simplicity of definitions in this part, and if we insist on RISCV here
-  this will confuse people that know it doesn't have a HALT instruction.}
+We assume we have an underlying step function that takes machine
+states to machine states, $\MACHS \rightharpoonup \MACHS$; we also
+assume the only possible partiality is because of a $\HALT$
+instruction.
+
+%The
+%{\em underlying} step function, $\mathord{\stepstounderfull} \in
+%\MACHS \rightharpoonup \MACHS$, embodies the standard RISC-V ISA
+%semantics.  \bcp{It's probably clearer not to put RISCV in the
+%  notation---for most of the paper, the fact that it's a RISCV really
+%  doesn't matter.  Moreover, I think we want to make it a partial
+%  function, for simplicity of definitions in this part, and if we
+%  insist on RISCV here this will confuse people that know it doesn't
+%  have a HALT instruction.}
 
@@
 A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
-\{\HIGH,\LOW\} \times \{\HI, \LO\}$\bcp{We need more mnemonic names for
+\{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT, \LOWINT\}$\bcp{We need more mnemonic names for
   these (that help us keep straight when we're talking about confidentiality
-  and when integrity), both in the tex and in the way they are displayed.}
+  and when integrity) in the way they are displayed.}
 is a map from machine components
-to confidentiality and integrity levels. $\LOW$ marks locations whose
+to confidentiality and integrity levels. $\LOWSEC$ marks locations whose
 contents are observable from the outside (by the ``attacker,'' e.g.,
-by DMA) and $\HIGH$ marks internal ``hidden'' state that cannot be
+by DMA) and $\HIGHSEC$ marks internal ``hidden'' state that cannot be
 observed directly (e.g., processor registers and memory addresses for
-temporary variables will usually be hidden). Similarly, $\LO$
+temporary variables will usually be hidden). Similarly, $\LOWINT$
 locations correspond to low-integrity components (e.g. locations that
-can be written to \bcp{directly or indirectly} by the network), while $\HI$ locations should be
+can be written to \bcp{directly or indirectly} by the network), while $\HIGHINT$ locations should be
 protected from such flows.
@@
 $\mach'$ whose components agree with $\mach$ on all components marked
-$\LOW$ by $\contour$---i.e., if $\contour(\component) = \LOW$, then
+$\LOWSEC$ by $\contour$---i.e., if $\contour(\component) = \LOWSEC$, then
 $\mach(\component) = \mach'(\component)$.
@@
   and for every high-integrity component $\component$ for which
-  $\contour(\component) = \HI$, $M_0(\component) = M_n(\component)$
+  $\contour(\component) = \HIGHINT$, $M_0(\component) = M_n(\component)$
   and $M'_0(\component) = M'_{n'}(\component)$.
@@
   a halt instruction.
-\item The instruction memory is marked $\LOW \times \HI$. That means
+\item The instruction memory is marked $\LOWSEC \times \HIGHINT$. That means
   it is observable and can't change.  \bcp{What does this assumption mean,
     precisely?  (I.e., what is ``the instruction memory''?)}
-\item The memory space is split in two. One part is marked $\LOW
-  \times \LO$: this is the public part (or the one that can be
+\item The memory space is split in two. One part is marked $\LOWSEC
+  \times \LOWINT$: this is the public part (or the one that can be
   accessed by the current process e.g.). The other part is marked
-  $\HIGH \times \HI$: it should be completely inaccessible to the
+  $\HIGHSEC \times \HIGHINT$: it should be completely inaccessible to the
   current process.
@@
 formally?}
-\item The instruction memory is again marked $\LOW \times \HI$. That
+\item The instruction memory is again marked $\LOWSEC \times \HIGHINT$. That
   means it is observable (both traces have the same instructions!) and
@@
   starting set of traces (where the whole machine state [registers and PC too,
-  right?] is marked $\LOW$), closed under an operator that takes a
+  right?] is marked $\LOWSEC$), closed under an operator that takes a
   contour+trace+trace and snips out all the subtraces between a call and its
-return.  (But: eek---see below.)}, the entire data memory is marked $\LOW
-  \times \LO$. That makes for a very unintersting property as all
+return.  (But: eek---see below.)}, the entire data memory is marked $\LOWSEC
+  \times \LOWINT$. That makes for a very unintersting property as all
   variants are necessarily identical.
@@
     the local state of the caller becomes inaccessible (marked
-    $\HIGH \times \HI$) until a subsequent return.
+    $\HIGHSEC \times \HIGHINT$) until a subsequent return.
   \item For every variant of the new initial state with respect to the
@@
 The observations on which noninterference is defined will (necessarily)  be made on the values in the current stack frame.
-\apt{Again, we're conflating \VIS\ vs. \HID\ with \LOW\ vs \HIGH\; should probably be teased apart.}
+\apt{Again, we're conflating \VIS\ vs. \HID\ with \LOWSEC\ vs \HIGHSEC\; should probably be teased apart.}
 LSE has two aspects: secrecy (callees cannot read caller variables) and integrity (callees cannot alter caller variables).

### 2020-02-04 main.tex
Notes
Formulation of stack safety based on the positive/negative functor. Hyperhyperproperties?

@@
 
-We assume we have an underlying step function that takes machine
-states to machine states, $\MACHS \rightharpoonup \MACHS$; we also
-assume the only possible partiality is because of a $\HALT$
-instruction.
-
-%The
-%{\em underlying} step function, $\mathord{\stepstounderfull} \in
-%\MACHS \rightharpoonup \MACHS$, embodies the standard RISC-V ISA
-%semantics.  \bcp{It's probably clearer not to put RISCV in the
+The {\em underlying} step function, $\mathord{\stepstounderfull} \in
+\MACHS \rightarrow \MACHS$, embodies the standard RISC-V ISA
+semantics.
+%\bcp{It's probably clearer not to put RISCV in the
 %  notation---for most of the paper, the fact that it's a RISCV really
@@
 
-% \bcp{Needed?}A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
-% sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
-% $\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.
+A {\em trace} is a finite or infinite sequence of machine states,
+$\seq{\mach} = \mach_0,\, \mach_1,\ldots$, in which adjacent machine
+states are related by $\stepstounderfull$---i.e., in which
+$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the
+sequence.
 
@@
 enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
-capability hardware, etc.). We use {\em property} for a mathematical
+capability hardware, etc.).
+\leo{FIX: terminology}
+We use {\em property} for a mathematical
 predicate on programs (i.e., machine states) articulating some useful
@@
 
+We denote this (reflexive) flows-to relation with $\flows$. Then
+$\LOWSEC \flowsstrict \HIGHSEC$ and $\HIGHINT \flowsstrict
+\LOWINT$. We can easily lift this relation to pairs of secrecy and
+integrity labels, and from there to contours pointwise.
+
 \bcp{A random thought about integrity: One thing that may be helping confuse
@@
 
-%
-%\subsection{Observations}
-%
-%A {\em component observation} is either a word or the special value $\bot$.
-%A {\em state observation} is a map from machine components to component
-%observations.
-%
-%An {\em observation} is a nonempty, finite or infinite sequence of state
-%observations in which (1) adjacent elements are always distinct (``no
-%stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
-%the the $j^{th}$ state observation in the sequence must also be mapped to
-%$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
-%component's value changes from $\bot$ to anything else, it never changes
-%back to $\bot$ again).
-%
-%In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
-%from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
-%observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
-%follows:
-%\begin{itemize}
-%\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
-%to its value in $\mach_0$ and every hidden
-%component to $\bot$;
-%\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
-%visible component (where $\contour_{i+1} = \VIS$), copying its
-%value from $\mach_{i+1}$, and (2) for each
-%hidden component, where $\contour_i = \VIS$, copying its
-%value from $\stateobs_{i+1}$\apt{$\stateobs_i$ ?}, and (3) if the resulting $\stateobs_{i+1}$ is
-%identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
-%\iftext\bcp{Too informal!!}\fi
-%\end{itemize}
+\subsection{Observations}
+
+A {\em component observation} is either a word or the special value $\bot$.
+A {\em state observation} is a map from machine components to component
+observations.
+
+An {\em observation} is a nonempty, finite or infinite sequence of state
+observations in which (1) adjacent elements are always distinct (``no
+stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
+the the $j^{th}$ state observation in the sequence must also be mapped to
+$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
+component's value changes from $\bot$ to anything else, it never changes
+back to $\bot$ again).
+
+In particular, the observation $\obstrace(\contour, \seq{\mach})$
+arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
+is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
+\ldots$ defined informally as follows:
+\begin{itemize}
+\item $\stateobs_0$ maps every visible component $\component$ (where
+  $\contour(\component) = (\LOWSEC, \_)$) to its value in $\mach_0$
+  and every hidden component to $\bot$;
+\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
+  each visible component $\component$ (where $\contour(\component) =
+  (\LOWSEC, \_)$), copying its value from $\mach_{i+1}$, and (2) for
+  each hidden component $\component$, where $\contour (\component) =
+  (\HIGHSEC,\_)$, copying its value from $\stateobs_{i}$, and (3) if
+  the resulting $\stateobs_{i+1}$ is identical to $\stateobs_{i}$,
+  then skipping $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
+\end{itemize}
 
@@
 
-%\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
+\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
 
@@
 % \bcp{prefix-closed?} set of pairs of sequences of machine states.
-
+% 
 % An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
@@
 
-A {\em trace} of length $n$, where $n$ can be finite or infinite\bcp{awkward}, is a
-length-$n$ sequence $\seq{\mach}_n = \mach_0,\, \mach_1,\ldots$ of
-machine states in which adjacent machine states are related by
-$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
-\mach_i$ for each $i > 0$ in the sequence.
-
-A {\em contoured hyperproperty} is a set of triples, each consisting of a
-contour along with a pair of traces.
-
-\leo{No longer needed:?}
-A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
-such that each trace in every pair of traces in the set executes the
-same instruction sequence.
-
-Finally, we can parameterize the traces by some initial application-specific
-state to obtain a {\em contoured lockstep $X$-hyperproperty}, where every
-trace comes with more information, for example call sites in stack safety.
-\bcp{If we use this later, we need to say it more formally to be clear about
-  what we mean---e.g., I had thought both the traces and the contour would
-  be functions of $X$!}
-
-\subsection{Noninterference}
-
-\leo{Indistinguishability/variations:}
+Let $(\stepplus) :: (\CONTOURS \times \MACHS) \rightarrow (\CONTOURS
+\times \MACHS)$ be an extension of the underlying step function, such
+that $(C,M) \stepplus (C',M')$ implies that $M \stepstounderfull M'$.
+Given a contour $C$ and a machine state $M$, the {\em contoured trace}
+induced by $C$ and $M$, written $\induced{C}{M}$ is the largest
+(finite or infinite) sequence of pairs of (contour, machine state)
+pairs, in which adjacent pairs are related by $\stepstounderfull$ and
+for which $C \flows C_i$ for all $i > 0$ in the sequence.
+\leo{That is, the induced contoured trace by a (C,M) pair incorporates
+  the ``return'' snipping}
 
@@
 $\mach(\component) = \mach'(\component)$.
+
+Let $\initmach$ be the set of initial states of the machine and let
+$\initcontour(M)$ be the initial contour assigned to each $M \in
+\initmach$.\leo{This could probably be generalized to a set}
+
+For each $M \in \initmach$ and $C = \initcontour(M)$, we define the
+{\em closure} of $(C,M)$ as the set of triples of (contour, trace,
+trace) as follows:
+\[
+\begin{array}{rcl}
+  \closure{(C,M)} = \{ ( C_i, \seq{N}, \seq{N'})
+  & | & C_{i-1} \not = C_i \\
+  & , & \seq{(C_i, M_i)} = \induced{C}{M}\\
+  & , & \seq{(\_, N_j)} = \induced{C_i}{M_i}\\
+  & , & N'_0 \in \variant{C_i}{M_i}\\
+  & , & \seq{(\_, N'_j)} = \induced{C_i}{N_0'} \}\\
+\end{array}  
+\]
+\leo{Alternative: We could do this for all $i$, but lose the
+  trivial instantiation of noninterference later on}
+
+A trace $\seq{M}$ satisfies the {\em integrity} safety property with
+respect to a contour $\contour$, if for all components $\component$
+where $\contour(\component) = (\_, \HIGHINT)$, $M_i(K) = M_0(K)$ for
+all $i$ in the trace.
+
+A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
+  confidentiality} safety hyperproperty with respect to a contour
+$\contour$, if, if the initials states are variants $N_0 \in
+\variant{C}{M_0}$, then their observations coincide
+$\obstrace(\contour, \seq{M}) = \obstrace(\contour, \seq{N})$.
+\leo{Should we cut at minimum length for termination-insensitiveness?}
+
+\leo{From now on I need better terminology, but hopefully you get the
+  point} An initial machine state \leo{or contour machine state pair}
+is \leo{good} if for every triple $(C, \seq{N}, \seq{N'})$ in
+$\closure{(\initcontour(M),M)}$ we have:
+\begin{itemize}
+\item The traces $\seq{N}$ and $\seq{N'}$ satisfy confidentiality
+  with respect to $C$.
+\item The traces $\seq{N}$ and $\seq{N'}$ each satisfy integrity
+  with respect to $C$.
+\end{itemize}
+
+\leo{Combined result}
+Noninterference holds for the set of \leo{good} initial machine
+states.
+
+\leo{NOTE:
+  \begin{itemize}
+    \item When the contours don't change this is trivially
+      noninterference: only a variation of the initial state is
+      added to the trace.
+    \item This almost directly coincides with the LSE part
+      of the stack-safety property.
+  \end{itemize}
+}
+
+%A {\em contoured hyperproperty} is a set of triples,
+%each consisting of a contour along with a pair of traces.
+% 
+%\leo{No longer needed:?}
+%A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
+%such that each trace in every pair of traces in the set executes the
+%same instruction sequence.
+% 
+%Finally, we can parameterize the traces by some initial application-specific
+%state to obtain a {\em contoured lockstep $X$-hyperproperty}, where every
+%trace comes with more information, for example call sites in stack safety.
+%\bcp{If we use this later, we need to say it more formally to be clear about
+%  what we mean---e.g., I had thought both the traces and the contour would
+%  be functions of $X$!}
+
+\subsection{Noninterference}
+
+\leo{Indistinguishability/variations:}
 

### 2020-02-04 main.tex
Notes / polish

@@
 capability hardware, etc.).
-\leo{FIX: terminology}
 We use {\em property} for a mathematical
-predicate on programs (i.e., machine states) articulating some useful
-security goal, e.g. noninterference. \iftext\bcp{Not quite true, if we stick
-  with hyperproperties.} \fi
+condition on programs (i.e., machine states) articulating some useful
+security goal, e.g. noninterference.
 \fi
@@
 locations correspond to low-integrity components (e.g. locations that
-can be written to \bcp{directly or indirectly} by the network), while $\HIGHINT$ locations should be
+can be written to \iftext \bcp{directly or indirectly} \fi by the network), while $\HIGHINT$ locations should be
 protected from such flows.
@@
 integrity labels, and from there to contours pointwise.
-
-\bcp{A random thought about integrity: One thing that may be helping confuse
-  us about integrity is that we've gone back and forth between ``eager'' and
-  ``lazy'' versions of various micropolicies, where the ``lazy'' ones do
-  allow high-integrity data to be overwritten, but only in a way that it
-  cannot be read again later (e.g., the lazy heap policy guarantees that, if
-  inaccessible memory is overwritten, then its color will change).  The lazy
-point of view is much more subtle as an integrity property.  So, to keep our
-heads straight here, we should be thinking only about eager properties in
-the first instance.}
 
@@
 Let $\initmach$ be the set of initial states of the machine and let
-$\initcontour(M)$ be the initial contour assigned to each $M \in
-\initmach$.\leo{This could probably be generalized to a set}
+$\initcontour(M)$ be the\bcp{an?} initial contour assigned to each $M \in
+\initmach$. \iftext\leo{This could probably be generalized to a set}\fi
 
@@
   trivial instantiation of noninterference later on}
+\bcp{Don't follow the notation here.  :-(}
 
@@
 $\contour$, if, if the initials states are variants $N_0 \in
-\variant{C}{M_0}$, then their observations coincide
+\variant{C}{M_0}$, then their observations coincide:
 $\obstrace(\contour, \seq{M}) = \obstrace(\contour, \seq{N})$.
 \leo{Should we cut at minimum length for termination-insensitiveness?}
+\bcp{Or just ask that one be a prefix of the other?}
 
@@
   and this only makes sense if they are executing in lockstep.}
+
+\section{Implementation}
+\label{impl}
+
+Sketch of the micropolicies that claim to implement these properties on
+PIPE...
+
+\subsection{A taint checking monitor}
+
+\subsection{A stack safety monitor}
+
+\section{Testing the implementations}
+
+\begin{itemize}
+\item Experimental results of testing
+\end{itemize}
+
+\section{Related work}
+
+\begin{itemize}
+\item Nick and Andre's papers (and other micropolicy papers)
+\cite{DBLP:conf/sp/RoesslerD18}
+
+This is the basis of our run-time monitor in Section{XXX}.  Specifically,
+our micropolicy is a simplified variant of their ... (I think it is the
+Return-Address Protection + Static Authorities + Depth Isolation policy, but
+need to double check---be careful about laziness.)  Essentially, our work
+arose as an attempt to formalize the security property enforced by something
+like one of their stack policies.
+
+\item Hritcu et al's work on ``Testing Noninterference
+Quickly''~\cite{DBLP:journals/jfp/HritcuLSADHPV16} laid the
+foundation of the random testing strategy that we use here.  In particular,
+that paper explored testing strategies based on both standard ``End-to-end
+noninterference''---a pleasantly extensional and easy-to-define property
+that leads to somewhat inefficient testing because process that misbehaves
+while its PC is labeled ``high'' must return to a low PC before the
+misbehavior can be identified as such---and ``Single-step
+noninterference''---a more intricate property related to Goguen and
+Meseguer's ``unwinding conditions''~\cite{GoguenMeseguer85} that ``fails
+faster'' and is thus easier to test.  \bcp{Explain the relation of the
+  latter property to our stack testing strategy.}
+
+\item Azevedo de Amorim's presentation of heap safety as a noninterference
+property~\cite{DBLP:conf/post/AmorimHP18} was a key inspiration for our
+definition of stack safety.  (Details...)
+
+\item StkTokens: Skorstengaard, Devriese, Birkedal
+\begin{itemize}
+\item POPL \cite{skorstengaard2019stktokens}
+\item tech report with proofs: \cite{skorstengaard2018stktokens}
+\url{https://arxiv.org/pdf/1811.02787.pdf}
+\item TR with proofs for ``Reasoning about a machine...'', Feb 2019:
+\url{https://arxiv.org/pdf/1902.05283.pdf}
+\item \url{http://cs.au.dk/~lask/papers/esop2018.pdf}
+\item Also: {\em Reasoning about a Machine with Local Capabilities: Provably
+  Safe Stack and Return Pointer Management} \cite{10.1145/3363519}
+\url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}
+
+\bcp{I'm nervous that these papers may be viewed as subsuming most of
+  ours---in particular, their security property is formulated as a logical
+  relation, i.e. a relational property!  We really need to dig into the
+  details (which are unfortunately rather technical).}
+\end{itemize}
+
+\item Stuff about dynamically changing labels (see footnote in Testing
+Noninterference Quickly paper)
+
+\item Simple Noninterference by Normalization \url{http://nachivpn.me/nibnbe.pdf}
+
+\item Mechanized Reasoning about a Capability Machine: [Linn] Georges,
+Trieu, Birkedal (PriSC 2020).
+%
+Described at the Iris Workshop as formalizing the security properties of
+capabilities. This appears to involve a fairly complex and very large
+mechanization in Iris (also in the eyes of the Iris people).
+
+\item \url{https://www.cis.upenn.edu/~stevez/papers/LZ05a.pdf} ??
+
+\item Not so closely related?
+\begin{itemize}
+\item Daniel's tainting paper (and some refs cited there)
+      \cite{DBLP:conf/eurosp/SchoepeBPS16} -- not needed if we postpone
+      dealing with tainting till another paper!
+\item We'd better say something about the relation to the original
+hyperproperties paper, no?  (No, not if we drop the word hyperproperty!)
+\end{itemize}
+\end{itemize}
+
+\iftext
+\section{Future work}
+
+\begin{itemize}
+\item Variations / extensions (OS, multiple processes, coroutines, generalizing depths to stacks of static call points, etc.)
+\item Extending the policy / property to more realistic ones
+\begin{itemize}
+\item Argument passing
+\item Variable-size stack frames
+\item Stack variables
+\item Relation to Nick and Andre’s policy
+\end{itemize}
+\item Could we also implement all this using Cheri capabilities?
+\item A more abstract presentation, perhaps unifying our treatments of
+confidentiality and integrity?
+\end{itemize}
+
+More:
+\begin{itemize}
+\item More realistic calling conventions.
+(We need to eventually match this to the particular calling convention of whatever compiler we choose.)  E.g., the above will not quite work for passing arguments on the stack: callee won’t be able to read them…So perhaps need to support a special kind of write instruction that stores D.PCdepth+1 .
+[RB: Or implement this distinction through tags?]
+\item Combination with heap policy (colors, etc.)
+\item For coroutines:
+\begin{itemize}
+\item Generalize depth to depth + PID
+\item Private stuff not having changed as an idea should still work
+\item Maybe not so hard: The scheduler + other coroutines are just a sort of generalized “callee” as far as this process is concerned.
+Just need to fiddle th step relation so that it recognizes which process is getting returned to and updates its description appropriately.
+\end{itemize}
+\item If we had a heap policy, maybe we could talk about dynamic code generation, loading, as allocating a heap region, storing instructions into it, and then calling into them.
+For now, we just make all the instructions read only and assume they exist in the initial state.
+\item Question: how efficiently can consistency be tested?
+How much scrambling is needed, and does it really need to occur before every instruction?
+How expensive is it to check for agreement between the (in)accessible parts of two memories?
+Would spot checking for agreement be adequate?]
+\item What are the effects of compartmentalization on the ability of each compartment to change its own code/stack?
+\item Tail calls (maybe by not marking them as call sites?)
+\end{itemize}
+
+\section*{Acknowledgments}
+\fi
+
+\section*{References}
+\bibliographystyle{IEEEtran}
+% \bibliography{bcp,local}
+
+\newpage
+
+\section*{Things to think about}
+
+\begin{itemize}
+\item
+How does Nick’s stack policy actually work these days?
+\item Andrew remembers that Catalin looked at something like this a long time ago and they decided that they needed unique tags for each dynamic call, not just levels.  We should check.
+\item \bcp{A random thought about integrity: One thing that may be helping confuse
+  us about integrity is that we've gone back and forth between ``eager'' and
+  ``lazy'' versions of various micropolicies, where the ``lazy'' ones do
+  allow high-integrity data to be overwritten, but only in a way that it
+  cannot be read again later (e.g., the lazy heap policy guarantees that, if
+  inaccessible memory is overwritten, then its color will change).  The lazy
+point of view is much more subtle as an integrity property.  So, to keep our
+heads straight here, we should be thinking only about eager properties in
+the first instance.}
+\end{itemize}
+
+\newpage
 
@@
 
-\section{Implementation}
-\label{impl}
-
-Sketch of the micropolicies that claim to implement these properties on
-PIPE...
-
-\subsection{A taint checking monitor}
-
-\subsection{A stack safety monitor}
-
-\section{Testing the implementations}
-
-\begin{itemize}
-\item Experimental results of testing
-\end{itemize}
-
-\section{Related work}
-
-\begin{itemize}
-\item Nick and Andre's papers (and other micropolicy papers)
-\cite{DBLP:conf/sp/RoesslerD18}
-
-This is the basis of our run-time monitor in Section{XXX}.  Specifically,
-our micropolicy is a simplified variant of their ... (I think it is the
-Return-Address Protection + Static Authorities + Depth Isolation policy, but
-need to double check---be careful about laziness.)  Essentially, our work
-arose as an attempt to formalize the security property enforced by something
-like one of their stack policies.
-
-\item Hritcu et al's work on ``Testing Noninterference
-Quickly''~\cite{DBLP:journals/jfp/HritcuLSADHPV16} laid the
-foundation of the random testing strategy that we use here.  In particular,
-that paper explored testing strategies based on both standard ``End-to-end
-noninterference''---a pleasantly extensional and easy-to-define property
-that leads to somewhat inefficient testing because process that misbehaves
-while its PC is labeled ``high'' must return to a low PC before the
-misbehavior can be identified as such---and ``Single-step
-noninterference''---a more intricate property related to Goguen and
-Meseguer's ``unwinding conditions''~\cite{GoguenMeseguer85} that ``fails
-faster'' and is thus easier to test.  \bcp{Explain the relation of the
-  latter property to our stack testing strategy.}
-
-\item Azevedo de Amorim's presentation of heap safety as a noninterference
-property~\cite{DBLP:conf/post/AmorimHP18} was a key inspiration for our
-definition of stack safety.  (Details...)
-
-\item StkTokens: Skorstengaard, Devriese, Birkedal
-\begin{itemize}
-\item POPL \cite{skorstengaard2019stktokens}
-\item tech report with proofs: \cite{skorstengaard2018stktokens}
-\url{https://arxiv.org/pdf/1811.02787.pdf}
-\item TR with proofs for ``Reasoning about a machine...'', Feb 2019:
-\url{https://arxiv.org/pdf/1902.05283.pdf}
-\item \url{http://cs.au.dk/~lask/papers/esop2018.pdf}
-\item Also: {\em Reasoning about a Machine with Local Capabilities: Provably
-  Safe Stack and Return Pointer Management} \cite{10.1145/3363519}
-\url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}
-
-\bcp{I'm nervous that these papers may be viewed as subsuming most of
-  ours---in particular, their security property is formulated as a logical
-  relation, i.e. a relational property!  We really need to dig into the
-  details (which are unfortunately rather technical).}
-\end{itemize}
-
-\item Stuff about dynamically changing labels (see footnote in Testing
-Noninterference Quickly paper)
-
-\item Simple Noninterference by Normalization \url{http://nachivpn.me/nibnbe.pdf}
-
-\item Mechanized Reasoning about a Capability Machine: [Linn] Georges,
-Trieu, Birkedal (PriSC 2020).
-%
-Described at the Iris Workshop as formalizing the security properties of
-capabilities. This appears to involve a fairly complex and very large
-mechanization in Iris (also in the eyes of the Iris people).
-
-\item \url{https://www.cis.upenn.edu/~stevez/papers/LZ05a.pdf} ??
-
-\item Not so closely related?
-\begin{itemize}
-\item Daniel's tainting paper (and some refs cited there)
-      \cite{DBLP:conf/eurosp/SchoepeBPS16} -- not needed if we postpone
-      dealing with tainting till another paper!
-\item We'd better say something about the relation to the original
-hyperproperties paper, no?  (No, not if we drop the word hyperproperty!) 
-\end{itemize}
-\end{itemize}
-
-\iftext
-\section{Future work}
-
-\begin{itemize}
-\item Variations / extensions (OS, multiple processes, coroutines, generalizing depths to stacks of static call points, etc.)
-\item Extending the policy / property to more realistic ones
-\begin{itemize}
-\item Argument passing
-\item Variable-size stack frames
-\item Stack variables
-\item Relation to Nick and Andre’s policy
-\end{itemize}
-\item Could we also implement all this using Cheri capabilities?
-\item A more abstract presentation, perhaps unifying our treatments of
-confidentiality and integrity?
-\end{itemize}
-
-More:
-\begin{itemize}
-\item More realistic calling conventions.
-(We need to eventually match this to the particular calling convention of whatever compiler we choose.)  E.g., the above will not quite work for passing arguments on the stack: callee won’t be able to read them…So perhaps need to support a special kind of write instruction that stores D.PCdepth+1 .
-[RB: Or implement this distinction through tags?]
-\item Combination with heap policy (colors, etc.)
-\item For coroutines:
-\begin{itemize}
-\item Generalize depth to depth + PID
-\item Private stuff not having changed as an idea should still work
-\item Maybe not so hard: The scheduler + other coroutines are just a sort of generalized “callee” as far as this process is concerned.
-Just need to fiddle th step relation so that it recognizes which process is getting returned to and updates its description appropriately.
-\end{itemize}
-\item If we had a heap policy, maybe we could talk about dynamic code generation, loading, as allocating a heap region, storing instructions into it, and then calling into them.
-For now, we just make all the instructions read only and assume they exist in the initial state.
-\item Question: how efficiently can consistency be tested?
-How much scrambling is needed, and does it really need to occur before every instruction?
-How expensive is it to check for agreement between the (in)accessible parts of two memories?
-Would spot checking for agreement be adequate?]
-\item What are the effects of compartmentalization on the ability of each compartment to change its own code/stack?
-\item Tail calls (maybe by not marking them as call sites?)
-\end{itemize}
-
-\section*{Acknowledgments}
-\fi
-
-\section*{References}
-\bibliographystyle{IEEEtran}
-% \bibliography{bcp,local}
-
-\newpage
-
-\section*{Things to think about}
-
-\begin{itemize}
-\item
-How does Nick’s stack policy actually work these days?
-\item Andrew remembers that Catalin looked at something like this a long time ago and they decided that they needed unique tags for each dynamic call, not just levels.  We should check.
-\end{itemize}
-
 \end{document}

### 2020-02-04 main.tex
Some notes no StkTokens
To be added: perspective w.r.t. ESOP paper.

@@
   details (which are unfortunately rather technical).}
+
+\rb{
+  Some notes on StkTokens.
+
+  A calling convention based on linear capabilities to enforce WBCF and LCE. The
+  program is divided into mutually distrustful components. Instead of using
+  separate stacks for each component and a shared control stack, it uses a
+  single shared stack. For this, it needs to pass stack and return capabilities
+  to possibly untrusted components temporarily, revoking them later. Together
+  with a small number of security checks, these measures suffice to guarantee
+  WBCF and LCE. StkTokens employs linear capabilities to obtain this kind of
+  result while limiting the amount of overhead needed to secure the transitions
+  between boundary crossings, modeled by a standard notion of components,
+  linking, disjoint interfaces, etc.
+
+  The capability machine that is modeled has two final configurations: halted
+  (for successful termination) and failed. It is a mostly standard capability
+  machine, modified by the additional requirements of linear capabilities
+  (moves, stores and loads, seals, splitting and splicing of ranges of
+  capability).
+
+  The formalization relies on the definition of a second operational semantics
+  for the target language, where WBCF and LCE are natively enforced, and a proof
+  of FA between that and the original operational semantics of the target. This
+  semantics executes in lockstep with the original target semantics; in
+  addition, it gives a special interpretation to the sequence of call
+  instructions on top of the regular step relation. It assumes that all
+  components meet some syntactic well-formedness requirements, and requires
+  trusted components to perform a number of simple checks in addition to
+  employing the calling convention proper (check of stack base address on call,
+  appropriate use and protection of return and closure seals), as well as some
+  conditions on initial system configurations. The formalization has a number
+  of advantages over previous work, notably the establishment of some temporal
+  safety guarantees.
+
+  The core of the proof of FA is the definition of when two components are
+  related. This is modeled by a step-indexed Kripke logical with recursive
+  worlds, which each collect assumptions about the system. Worlds are divided
+  into sub-worlds for heap, call stack and free stack, each mapping region names
+  to regions, which can be either spatial (all three) or shared (heap only,
+  encapsulating a system invariant, including an invariant on seals). Spatial
+  regions differ from shared regions in their linearity and (optionally)
+  ownership. ``Regular'' spatial regions can become owned or revoked without
+  while preserving the relation between memories in the two semantics. The proof
+  uses (shared) code regions, regions for heap and stack memory, and regions to
+  track immutable, inactive parts of the stack. With those Kripke worlds, the LR
+  defines relatedness of various entities. There are two value relations for
+  trusted and untrusted values (a subset of trusted values which are safe to
+  pass to an adversary, including stack pointers and their linear capabilities,
+  and sealed values, all under certain conditions on the ownership of the
+  various parts of the stack). The top-level FA proof is straightforward.
+
+  The extended operational semantics has support for tail calls. Because data
+  memory does not contain any seals or ways to get to those, it can always be
+  shared safely.
+
+  Like in our scheme, instructions must be marked as executable and a check on
+  the PC made at each step. Stack pointers are now linear stack capabilities
+  (and there should only be stack pointers for the stack memory). Return
+  pointers are pairs of sealed return capabilities, encapsulating the callers
+  data and return code; the callee can only jump to these. A caller will split
+  its own stack capability into two: one for its local stack frame (which will
+  be sealed as the data component of the return capability pair) and one for the
+  unused stack, which will be passed to the callee. This second capability is
+  also used as a token to prove that the callee is allowed to return to the
+  caller by adjacency with the sealed data capability (for this trick to work,
+  stack frames must be non-empty). Another requirement is that each call point
+  needs to be associated with a unique seal for returns, which should never be
+  leaked to an adversary.
+
+}
+
 \end{itemize}

### 2020-02-04 main.tex
Some notes on StkTokens' ESOP predecessor

@@
   leaked to an adversary.
+
+  In their ESOP 2018 paper, the same authors present a similar scheme that uses
+  local capabilities employing similar techniques. Their capability machine is
+  mostly standard. Operations other than those working on capabilities are
+  mostly standard, with small extensions to account for the presence
+  capabilities. Notably, jmp needs to account for enter capabilities, which can
+  only be used to jump to a location and then become read-execute; they
+  represent an opaque packing of some code together with local state. The store
+  operation is also complicated, this time by the presence of local
+  capabilities, which can only be stored in registers (and write-local memory)
+  and may be temporarily handed over to an adversary, say for the duration of a
+  call.
+
+  The work presents a concrete calling convention that relies on a single shared
+  stack and has moderate overhead with stack and return pointers passed through
+  registers as local capabilities. The scheme imposes some standard conditions
+  on initial state and memory allocation primitives, and safety measures a
+  component needs to perform to protect itself against an adversary. The most
+  demanding requirement is that not only non-return value registers and used
+  stack, but also the unused part of the stack (which may contain local
+  capabilities) needs to be cleared when returning to the adversary. Simple
+  measures are applied on calls and returns to and from an adversary, and on
+  calls from an adversary.
+
+  The logical relation that is presented is not specific to the particular
+  calling convention. It uses a variant of the public-private separation of
+  worlds to model the particular properties of local capabilities. In
+  particular, worlds have permanent, temporary and revoked regions, with
+  non-revoked regions comprising a state transition system with both public and
+  private transitions. Local and global capabilities can depend on protocols
+  from permanent regions, and local capabilities can also depend on protocols
+  from temporary regions (which may be revoked in the future, and may not
+  outlive those regions). A pair of future world relations describe the changes
+  of memory over time. In the public FW relation, temporary and permanent
+  regions can evolve according to the public part of their transition system,
+  and revoked regions can be reinstated as temporary. In the private FW
+  relation, temporary and revoked regions can be replaced by arbitrary regions,
+  and permanent regions can evolve according to the private part of their
+  transition system. The relation defines an observation relation for safe
+  configurations (i.e., the memory of a halting configuration is satisfied by
+  some private FW). The relation on values covers conditions for the safe use of
+  write and write-local capabilities, and similarly for enter capabilities.
+
+  All this is used to prove WBCF and LCE for concrete example programs.
 

### 2020-02-04 main.tex
Fix problem with line breaks inside comments

--- 2020-02-04 main.tex
@@
   details (which are unfortunately rather technical).}
+
+\rb{
+  Some notes on StkTokens.
+}
+\rb{
+  A calling convention based on linear capabilities to enforce WBCF and LCE. The
+  program is divided into mutually distrustful components. Instead of using
+  separate stacks for each component and a shared control stack, it uses a
+  single shared stack. For this, it needs to pass stack and return capabilities
+  to possibly untrusted components temporarily, revoking them later. Together
+  with a small number of security checks, these measures suffice to guarantee
+  WBCF and LCE. StkTokens employs linear capabilities to obtain this kind of
+  result while limiting the amount of overhead needed to secure the transitions
+  between boundary crossings, modeled by a standard notion of components,
+  linking, disjoint interfaces, etc.
+}
+\rb{
+  The capability machine that is modeled has two final configurations: halted
+  (for successful termination) and failed. It is a mostly standard capability
+  machine, modified by the additional requirements of linear capabilities
+  (moves, stores and loads, seals, splitting and splicing of ranges of
+  capability).
+}
+\rb{
+  The formalization relies on the definition of a second operational semantics
+  for the target language, where WBCF and LCE are natively enforced, and a proof
+  of FA between that and the original operational semantics of the target. This
+  semantics executes in lockstep with the original target semantics; in
+  addition, it gives a special interpretation to the sequence of call
+  instructions on top of the regular step relation. It assumes that all
+  components meet some syntactic well-formedness requirements, and requires
+  trusted components to perform a number of simple checks in addition to
+  employing the calling convention proper (check of stack base address on call,
+  appropriate use and protection of return and closure seals), as well as some
+  conditions on initial system configurations. The formalization has a number
+  of advantages over previous work, notably the establishment of some temporal
+  safety guarantees.
+}
+\rb{
+  The core of the proof of FA is the definition of when two components are
+  related. This is modeled by a step-indexed Kripke logical with recursive
+  worlds, which each collect assumptions about the system. Worlds are divided
+  into sub-worlds for heap, call stack and free stack, each mapping region names
+  to regions, which can be either spatial (all three) or shared (heap only,
+  encapsulating a system invariant, including an invariant on seals). Spatial
+  regions differ from shared regions in their linearity and (optionally)
+  ownership. ``Regular'' spatial regions can become owned or revoked without
+  while preserving the relation between memories in the two semantics. The proof
+  uses (shared) code regions, regions for heap and stack memory, and regions to
+  track immutable, inactive parts of the stack. With those Kripke worlds, the LR
+  defines relatedness of various entities. There are two value relations for
+  trusted and untrusted values (a subset of trusted values which are safe to
+  pass to an adversary, including stack pointers and their linear capabilities,
+  and sealed values, all under certain conditions on the ownership of the
+  various parts of the stack). The top-level FA proof is straightforward.
+}
+\rb{
+  The extended operational semantics has support for tail calls. Because data
+  memory does not contain any seals or ways to get to those, it can always be
+  shared safely.
+}
+\rb{
+  Like in our scheme, instructions must be marked as executable and a check on
+  the PC made at each step. Stack pointers are now linear stack capabilities
+  (and there should only be stack pointers for the stack memory). Return
+  pointers are pairs of sealed return capabilities, encapsulating the callers
+  data and return code; the callee can only jump to these. A caller will split
+  its own stack capability into two: one for its local stack frame (which will
+  be sealed as the data component of the return capability pair) and one for the
+  unused stack, which will be passed to the callee. This second capability is
+  also used as a token to prove that the callee is allowed to return to the
+  caller by adjacency with the sealed data capability (for this trick to work,
+  stack frames must be non-empty). Another requirement is that each call point
+  needs to be associated with a unique seal for returns, which should never be
+  leaked to an adversary.
+}
+\rb{
+  In their ESOP 2018 paper, the same authors present a similar scheme that uses
+  local capabilities employing similar techniques. Their capability machine is
+  mostly standard. Operations other than those working on capabilities are
+  mostly standard, with small extensions to account for the presence
+  capabilities. Notably, jmp needs to account for enter capabilities, which can
+  only be used to jump to a location and then become read-execute; they
+  represent an opaque packing of some code together with local state. The store
+  operation is also complicated, this time by the presence of local
+  capabilities, which can only be stored in registers (and write-local memory)
+  and may be temporarily handed over to an adversary, say for the duration of a
+  call.
+}
+\rb{
+  The work presents a concrete calling convention that relies on a single shared
+  stack and has moderate overhead with stack and return pointers passed through
+  registers as local capabilities. The scheme imposes some standard conditions
+  on initial state and memory allocation primitives, and safety measures a
+  component needs to perform to protect itself against an adversary. The most
+  demanding requirement is that not only non-return value registers and used
+  stack, but also the unused part of the stack (which may contain local
+  capabilities) needs to be cleared when returning to the adversary. Simple
+  measures are applied on calls and returns to and from an adversary, and on
+  calls from an adversary.
+}
+\rb{
+  The logical relation that is presented is not specific to the particular
+  calling convention. It uses a variant of the public-private separation of
+  worlds to model the particular properties of local capabilities. In
+  particular, worlds have permanent, temporary and revoked regions, with
+  non-revoked regions comprising a state transition system with both public and
+  private transitions. Local and global capabilities can depend on protocols
+  from permanent regions, and local capabilities can also depend on protocols
+  from temporary regions (which may be revoked in the future, and may not
+  outlive those regions). A pair of future world relations describe the changes
+  of memory over time. In the public FW relation, temporary and permanent
+  regions can evolve according to the public part of their transition system,
+  and revoked regions can be reinstated as temporary. In the private FW
+  relation, temporary and revoked regions can be replaced by arbitrary regions,
+  and permanent regions can evolve according to the private part of their
+  transition system. The relation defines an observation relation for safe
+  configurations (i.e., the memory of a halting configuration is satisfied by
+  some private FW). The relation on values covers conditions for the safe use of
+  write and write-local capabilities, and similarly for enter capabilities.
+}
+\rb{
+  All this is used to prove WBCF and LCE for concrete example programs.
+}
+
 \end{itemize}

### 2020-02-04 main.tex
Fix build

@@
 software- or hardware-enforced) security monitor, but we won't need to make
-any specific choices about details until we get to \cref{impl}.
+any specific choices about details until we get to \bcp{the implementation
+  section}.
+% \cref{impl}.
 

### 2020-02-17 main.tex
Fix ESOP'18 broken link
From Lars' homepage, as it is no longer available in Lau's.
Checkpoint: adjust existing terminology

@@
 
+A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\ACCYES,
+\ACCNO, \ACCCODE\}$ is a map from machine components to accessibility levels.
+$\ACCYES$ marks locations whose contents represent data observable from the
+outside (by the ``attacker,'' e.g., by DMA), which moreover correspond to
+low-integrity components (e.g., locations that can be written to \iftext
+\bcp{directly or indirectly} \fi by the network). $\ACCNO$ marks internal
+``hidden'' state that cannot be observed directly (e.g., processor registers and
+memory addresses for temporary variables will usually be hidden), which is
+moreover high-integrity (i.e., locations which should be protected from flows
+from the network)
+%
+\rb{Tried to merge discussion on integrity and secrecy, but maybe it is clearer
+  to separate both concerns?}
+%
+In addition to those two accessibility levels, $\ACCCODE$ marks locations that
+hold executable instructions, which are considered to be low-secrecy (i.e.,
+observable by the attacker) and high-integrity (i.e., the attacker cannot write
+to them).
+
+\rb{OLD:
 \leo{We might need to change the intuition behind some of these things.}
@@
 protected from such flows.
-
+}
+
+\rb{The flows-to relation was quite nice before, now it looks like it will get
+  messier\ldots}
+%
 We denote this (reflexive) flows-to relation with $\flows$. Then
@@
 \subsection{Observations}
+
+\rb{We are not using observations at the moment, but a few adjustments are made
+  here.}
 
@@
 \item $\stateobs_0$ maps every visible component $\component$ (where
-  $\contour(\component) = (\LOWSEC, \_)$) to its value in $\mach_0$
+  $\contour(\component) \in \{\ACCYES, \ACCCODE\}$) to its value in $\mach_0$
   and every hidden component to $\bot$;
 \item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
-  each visible component $\component$ (where $\contour(\component) =
-  (\LOWSEC, \_)$), copying its value from $\mach_{i+1}$, and (2) for
+  each visible component $\component$ (where $\contour(\component) \in
+  \{\ACCYES, \ACCCODE\}$), copying its value from $\mach_{i+1}$, and (2) for
   each hidden component $\component$, where $\contour (\component) =
-  (\HIGHSEC,\_)$, copying its value from $\stateobs_{i}$, and (3) if
+  \ACCNO$, copying its value from $\stateobs_{i}$, and (3) if
   the resulting $\stateobs_{i+1}$ is identical to $\stateobs_{i}$,
@@
 $\mach'$ whose components agree with $\mach$ on all components marked
-$\LOWSEC$ by $\contour$---i.e., if $\contour(\component) = \LOWSEC$, then
+$\ACCYES$ or $\ACCCODE$ by $\contour$---i.e., if $\contour(\component) \in \{\ACCYES, \ACCCODE\}$, then
 $\mach(\component) = \mach'(\component)$.
+%
+\rb{When we talk about observations, it is visibility that we are concerned with
+  and moving from $\HIGHSEC$ to $\ACCNO$ and $\LOWSEC$ to $\ACCYES$ and
+  $\ACCCODE$ is simple. Here, the case of code is slightly more
+  interesting (are $\ACCCODE$-variants ever useful?).}
 
@@
 all $i$ in the trace.
+%
+\rb{Here, $\HIGHINT$ would be replaced by $\in \{\ACCNO, \ACCCODE\}$.}
 
@@
 \url{https://arxiv.org/pdf/1902.05283.pdf}
-\item \url{http://cs.au.dk/~lask/papers/esop2018.pdf}
+\item \url{https://cs.au.dk/~birke/papers/local-capabilities-conf.pdf}
 \item Also: {\em Reasoning about a Machine with Local Capabilities: Provably

### 2020-02-17 main.tex
First iteration of a cleaner notion of safety

@@
 %  be functions of $X$!}
+
+\rb{New try. Contours are as above.}
+
+A {\em call/return map} $\kappa \in \WORDS \rightarrow \{\CRCALL, \CRRET,
+\CRBOT\}$ is a map from addresses to tags (not to be confused with tags at the
+micropolicy level \rb{reference if needed}) that denote whether a memory
+location is to be treated as (the first word of) a call instruction, a return
+instruction, or neither \rb{(and we should compare these with call and return
+{\em instructions})}.
+
+The {\em step relation} $(\stepkappa) :: \MACHS \rightharpoonup \MACHS$ is a
+partial function from machine states to machine states: a restriction of the
+native machine's step relation $\stepstounderfull$ indexed by a call/return
+map $\kappa$, such that \rb{(How much to build into this, and/or the machine
+states themselves? Also describe and notate the concept of ``running until
+the first unmatched return of two runs'') \ldots}
+
+Given a machine state $\mach$ and associated contour $\contour$, an
+$\ACCNO$-variant is any state $\mach'$ that varies machine components tagged as
+$\ACCNO$ in $\mach$: that is, for any component $\component$ of $\mach$, if
+$\contour(\component) \neq \ACCNO$, then
+$\mach(\component) = \mach'(\component)$; otherwise, if
+$\contour(\component) = \ACCNO$, then $\mach'(\component)$ holds an
+arbitrary value, unrelated to $\mach(\component)$. Similarly, $\ACCYES$-variant
+states are defined by scrambling those machine components tagged as $\ACCYES$
+and leaving all others unchanged.
+
+A machine state $\mach$ is {\em safe} with respect to a contour $\contour$
+\rb{and call/return map $\kappa$} if it satisfies the following conditions:
+%
+\begin{enumerate}
+
+\item {\em Integrity:} if we run $\mach$ until its first unmatched return (if
+  any), all components other than those marked as accessible must remain
+  unchanged with respect to $\mach$.
+
+\item {\em Confidentiality:} for every $\ACCNO$-variant $\mach'$ of $\mach$, if
+  we run $\mach$ and $\mach'$ until their first unmatched return, the resulting
+  final states are themselves $\ACCNO$-variant of one another. \rb{(Homogenize
+  notations, see also above.)}
+
+\item Let $\mach_1, \mach_2, \ldots$ be a machine execution starting from
+  $\mach$, i.e., $\mach_1 = \mach$. Let also $\contour_1, \contour_2, \ldots$
+  be the corresponding sequence of contours, such that $\contour_1 = \contour$
+  and at each step $\mach_i$ the corresponding $\contour_i$ \rb{\ldots better
+  define the contour-updating step in the style of $\stepplus$.} For all $i$,
+  if $M_i$ is a top-level call in the trace, then (the instruction immediately
+  after it is safe with respect to $C_i$.
+
+\item All instructions executed throughout any run starting from $\mach$ are
+marked as such in the contour $\contour$. \rb{Need notation to dereference the
+PC and have the memory be marked as $\ACCCODE$.}
+
+\end{enumerate}
+
+The step relation $\stepkappa$ is itself safe if every machine state $\mach$ is
+safe with respect to the special contour where all memory locations containing
+code are marked $\ACCCODE$ and all other components are marked $\ACCYES$.
 

### 2020-02-27 main.tex
Tweaks on safety

--- 2020-02-17 main.tex
@@
 %  be functions of $X$!}
+
+\rb{New try. Contours are as above.}
+
+A {\em call/return map} $\kappa \in \WORDS \rightarrow \{\CRCALL, \CRRET,
+\CRBOT\}$ is a map from addresses to tags (not to be confused with tags at the
+micropolicy level \rb{reference if needed}) that denote whether a memory
+location is to be treated as (the first word of) a call instruction, a return
+instruction, or neither \rb{(and we should compare these with call and return
+machine instructions ``proper'')}.
+
+The {\em step relation} $(\stepkappa) :: \MACHS \rightharpoonup \MACHS$ is a
+partial function from machine states to machine states: a restriction of the
+native machine's step relation $\stepstounderfull$ indexed by a call/return
+map $\kappa$, such that \rb{(How much to build into this, and/or the machine
+states themselves? Also describe and notate the concept of ``running until
+the first unmatched return (cf. observable halting states)'') \ldots}
+
+Given a machine state $\mach$ and associated contour $\contour$, an
+$\ACCNO$-variant is any state $\mach'$ that varies machine components tagged as
+$\ACCNO$ in $\mach$: that is, for any component $\component$ of $\mach$, if
+$\contour(\component) \neq \ACCNO$, then
+$\mach(\component) = \mach'(\component)$; otherwise (if
+$\contour(\component) = \ACCNO$), then $\mach'(\component)$ holds an
+arbitrary value, unrelated to $\mach(\component)$. Similarly, $\ACCYES$-variant
+states are defined by scrambling those machine components tagged as $\ACCYES$
+and leaving all others unchanged.
+
+A machine state $\mach$ is {\em safe} with respect to a contour $\contour$
+\rb{and call/return map $\kappa$} if it satisfies the following conditions:
+%
+\begin{enumerate}
+
+\item {\em Integrity:} if we run $\mach$ until its first unmatched return (if
+  any), all components other than those marked as accessible must remain
+  unchanged with respect to $\mach$. \rb{That is, the final state is an
+  $\ACCYES$-variant of $\mach$.}
+
+\item {\em Confidentiality:} for every $\ACCNO$-variant $\mach'$ of $\mach$, if
+  we run $\mach$ and $\mach'$ until their first unmatched returns, the resulting
+  final states are themselves $\ACCNO$-variant of one another. \rb{(Homogenize
+  notations, see also above.)}
+
+\item Let $\mach_1, \mach_2, \ldots$ be a machine execution starting from
+  $\mach$, i.e., $\mach_1 = \mach$. Let also $\contour_1, \contour_2, \ldots$
+  be the corresponding sequence of contours, such that $\contour_1 = \contour$
+  and at each step $\mach_i$ the corresponding $\contour_i$ \rb{\ldots better
+  define the contour-updating step in the style of $\stepplus$.} For all $i$,
+  if $M_i$ is a top-level call in the trace, then (the instruction immediately
+  after) it \rb{$\mach_{i+1}$?} is safe with respect to $C_i$.
+
+\item All instructions executed throughout any run starting from $\mach$ are
+  marked as such in the contour $\contour$. \rb{Need notation to dereference the
+  PC and have the memory be marked as $\ACCCODE$:
+  $\contour_i(\PC{\mach_i}) = \ACCCODE$ (or noting invariance in $\ACCCODE$
+  parts of $\contour$).}
+
+\end{enumerate}
+
+The step relation $\stepkappa$ is itself safe if every machine state $\mach$ is
+safe with respect to the special contour where all memory locations containing
+code are marked $\ACCCODE$ and all other components are marked $\ACCYES$.
 

### 2020-03-31 main.tex
Nits

--- 2020-02-17 main.tex
@@
 software- or hardware-enforced) security monitor, but we won't need to make
-any specific choices about details until we get to \bcp{the implementation
-  section}.
-% \cref{impl}.
+any specific choices about details until we get to \cref{impl}.
 
@@
 
-A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\ACCYES,
-\ACCNO, \ACCCODE\}$ is a map from machine components to accessibility levels.
-$\ACCYES$ marks locations whose contents represent data observable from the
-outside (by the ``attacker,'' e.g., by DMA), which moreover correspond to
-low-integrity components (e.g., locations that can be written to \iftext
-\bcp{directly or indirectly} \fi by the network). $\ACCNO$ marks internal
-``hidden'' state that cannot be observed directly (e.g., processor registers and
-memory addresses for temporary variables will usually be hidden), which is
-moreover high-integrity (i.e., locations which should be protected from flows
-from the network)
-%
-\rb{Tried to merge discussion on integrity and secrecy, but maybe it is clearer
-  to separate both concerns?}
-%
-In addition to those two accessibility levels, $\ACCCODE$ marks locations that
-hold executable instructions, which are considered to be low-secrecy (i.e.,
-observable by the attacker) and high-integrity (i.e., the attacker cannot write
-to them).
-
-\rb{OLD:
 \leo{We might need to change the intuition behind some of these things.}
@@
 protected from such flows.
-}
-
-\rb{The flows-to relation was quite nice before, now it looks like it will get
-  messier\ldots}
-%
+
 We denote this (reflexive) flows-to relation with $\flows$. Then
@@
 \subsection{Observations}
-
-\rb{We are not using observations at the moment, but a few adjustments are made
-  here.}
 
@@
 \item $\stateobs_0$ maps every visible component $\component$ (where
-  $\contour(\component) \in \{\ACCYES, \ACCCODE\}$) to its value in $\mach_0$
+  $\contour(\component) = (\LOWSEC, \_)$) to its value in $\mach_0$
   and every hidden component to $\bot$;
 \item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
-  each visible component $\component$ (where $\contour(\component) \in
-  \{\ACCYES, \ACCCODE\}$), copying its value from $\mach_{i+1}$, and (2) for
+  each visible component $\component$ (where $\contour(\component) =
+  (\LOWSEC, \_)$), copying its value from $\mach_{i+1}$, and (2) for
   each hidden component $\component$, where $\contour (\component) =
-  \ACCNO$, copying its value from $\stateobs_{i}$, and (3) if
+  (\HIGHSEC,\_)$, copying its value from $\stateobs_{i}$, and (3) if
   the resulting $\stateobs_{i+1}$ is identical to $\stateobs_{i}$,
@@
 $\mach'$ whose components agree with $\mach$ on all components marked
-$\ACCYES$ or $\ACCCODE$ by $\contour$---i.e., if $\contour(\component) \in \{\ACCYES, \ACCCODE\}$, then
+$\LOWSEC$ by $\contour$---i.e., if $\contour(\component) = \LOWSEC$, then
 $\mach(\component) = \mach'(\component)$.
-%
-\rb{When we talk about observations, it is visibility that we are concerned with
-  and moving from $\HIGHSEC$ to $\ACCNO$ and $\LOWSEC$ to $\ACCYES$ and
-  $\ACCCODE$ is simple. Here, the case of code is slightly more
-  interesting (are $\ACCCODE$-variants ever useful?).}
 
@@
 all $i$ in the trace.
-%
-\rb{Here, $\HIGHINT$ would be replaced by $\in \{\ACCNO, \ACCCODE\}$.}
 
@@
 \bcp{Or just ask that one be a prefix of the other?}
+\bcp{Or go back to halting states.  (Which I guess really means go back to
+  Halting + Failing states.  We never quite worked out how that would go.)}
 
@@
 
+\iftext
 \leo{Indistinguishability/variations:}
@@
   I'm confused about terminology}
+\fi
 
@@
 \end{itemize}
+
+\iftext
 \leo{In the terminology of the original hyperproperty paper,
@@
   version of integrity than DLM for example}.
+\fi
 %
@@
 \url{https://arxiv.org/pdf/1902.05283.pdf}
-\item \url{https://cs.au.dk/~birke/papers/local-capabilities-conf.pdf}
+\item \url{http://cs.au.dk/~lask/papers/esop2018.pdf}
 \item Also: {\em Reasoning about a Machine with Local Capabilities: Provably
@@
   details (which are unfortunately rather technical).}
-
-\rb{
-  Some notes on StkTokens.
-}
-\rb{
-  A calling convention based on linear capabilities to enforce WBCF and LCE. The
-  program is divided into mutually distrustful components. Instead of using
-  separate stacks for each component and a shared control stack, it uses a
-  single shared stack. For this, it needs to pass stack and return capabilities
-  to possibly untrusted components temporarily, revoking them later. Together
-  with a small number of security checks, these measures suffice to guarantee
-  WBCF and LCE. StkTokens employs linear capabilities to obtain this kind of
-  result while limiting the amount of overhead needed to secure the transitions
-  between boundary crossings, modeled by a standard notion of components,
-  linking, disjoint interfaces, etc.
-}
-\rb{
-  The capability machine that is modeled has two final configurations: halted
-  (for successful termination) and failed. It is a mostly standard capability
-  machine, modified by the additional requirements of linear capabilities
-  (moves, stores and loads, seals, splitting and splicing of ranges of
-  capability).
-}
-\rb{
-  The formalization relies on the definition of a second operational semantics
-  for the target language, where WBCF and LCE are natively enforced, and a proof
-  of FA between that and the original operational semantics of the target. This
-  semantics executes in lockstep with the original target semantics; in
-  addition, it gives a special interpretation to the sequence of call
-  instructions on top of the regular step relation. It assumes that all
-  components meet some syntactic well-formedness requirements, and requires
-  trusted components to perform a number of simple checks in addition to
-  employing the calling convention proper (check of stack base address on call,
-  appropriate use and protection of return and closure seals), as well as some
-  conditions on initial system configurations. The formalization has a number
-  of advantages over previous work, notably the establishment of some temporal
-  safety guarantees.
-}
-\rb{
-  The core of the proof of FA is the definition of when two components are
-  related. This is modeled by a step-indexed Kripke logical with recursive
-  worlds, which each collect assumptions about the system. Worlds are divided
-  into sub-worlds for heap, call stack and free stack, each mapping region names
-  to regions, which can be either spatial (all three) or shared (heap only,
-  encapsulating a system invariant, including an invariant on seals). Spatial
-  regions differ from shared regions in their linearity and (optionally)
-  ownership. ``Regular'' spatial regions can become owned or revoked without
-  while preserving the relation between memories in the two semantics. The proof
-  uses (shared) code regions, regions for heap and stack memory, and regions to
-  track immutable, inactive parts of the stack. With those Kripke worlds, the LR
-  defines relatedness of various entities. There are two value relations for
-  trusted and untrusted values (a subset of trusted values which are safe to
-  pass to an adversary, including stack pointers and their linear capabilities,
-  and sealed values, all under certain conditions on the ownership of the
-  various parts of the stack). The top-level FA proof is straightforward.
-}
-\rb{
-  The extended operational semantics has support for tail calls. Because data
-  memory does not contain any seals or ways to get to those, it can always be
-  shared safely.
-}
-\rb{
-  Like in our scheme, instructions must be marked as executable and a check on
-  the PC made at each step. Stack pointers are now linear stack capabilities
-  (and there should only be stack pointers for the stack memory). Return
-  pointers are pairs of sealed return capabilities, encapsulating the callers
-  data and return code; the callee can only jump to these. A caller will split
-  its own stack capability into two: one for its local stack frame (which will
-  be sealed as the data component of the return capability pair) and one for the
-  unused stack, which will be passed to the callee. This second capability is
-  also used as a token to prove that the callee is allowed to return to the
-  caller by adjacency with the sealed data capability (for this trick to work,
-  stack frames must be non-empty). Another requirement is that each call point
-  needs to be associated with a unique seal for returns, which should never be
-  leaked to an adversary.
-}
-\rb{
-  In their ESOP 2018 paper, the same authors present a similar scheme that uses
-  local capabilities employing similar techniques. Their capability machine is
-  mostly standard. Operations other than those working on capabilities are
-  mostly standard, with small extensions to account for the presence
-  capabilities. Notably, jmp needs to account for enter capabilities, which can
-  only be used to jump to a location and then become read-execute; they
-  represent an opaque packing of some code together with local state. The store
-  operation is also complicated, this time by the presence of local
-  capabilities, which can only be stored in registers (and write-local memory)
-  and may be temporarily handed over to an adversary, say for the duration of a
-  call.
-}
-\rb{
-  The work presents a concrete calling convention that relies on a single shared
-  stack and has moderate overhead with stack and return pointers passed through
-  registers as local capabilities. The scheme imposes some standard conditions
-  on initial state and memory allocation primitives, and safety measures a
-  component needs to perform to protect itself against an adversary. The most
-  demanding requirement is that not only non-return value registers and used
-  stack, but also the unused part of the stack (which may contain local
-  capabilities) needs to be cleared when returning to the adversary. Simple
-  measures are applied on calls and returns to and from an adversary, and on
-  calls from an adversary.
-}
-\rb{
-  The logical relation that is presented is not specific to the particular
-  calling convention. It uses a variant of the public-private separation of
-  worlds to model the particular properties of local capabilities. In
-  particular, worlds have permanent, temporary and revoked regions, with
-  non-revoked regions comprising a state transition system with both public and
-  private transitions. Local and global capabilities can depend on protocols
-  from permanent regions, and local capabilities can also depend on protocols
-  from temporary regions (which may be revoked in the future, and may not
-  outlive those regions). A pair of future world relations describe the changes
-  of memory over time. In the public FW relation, temporary and permanent
-  regions can evolve according to the public part of their transition system,
-  and revoked regions can be reinstated as temporary. In the private FW
-  relation, temporary and revoked regions can be replaced by arbitrary regions,
-  and permanent regions can evolve according to the private part of their
-  transition system. The relation defines an observation relation for safe
-  configurations (i.e., the memory of a halting configuration is satisfied by
-  some private FW). The relation on values covers conditions for the safe use of
-  write and write-local capabilities, and similarly for enter capabilities.
-}
-\rb{
-  All this is used to prove WBCF and LCE for concrete example programs.
-}
-
 \end{itemize}

### 2020-04-09 main.tex
Writeup of eager stuff safety

--- 2020-02-27 main.tex
@@
 $\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the
-sequence.
+sequence. We write $\seq{\mach}^i_j$ for the subtrace between $i$ and
+$j$.
+
+\subsection{Contours}
+
+A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
+\{\HIGHSEC, \LOWSEC\} \times \{\HIGHINT, \LOWINT\}$ is a map from
+machine components to a pair of confidentiality and integrity labels.
+$\LOWSEC$ marks locations whose contents represent data observable
+{\em by the current process}; $\HIGHSEC$ corresponds to non-observable
+data (e.g. local state of the process caller). $\HIGHINT$ marks
+locations that the current process does not have write access to;
+$\LOWINT$ is freely accessible memory. As examples of different
+combinations consider the following:
+\begin{itemize}
+\item Locations that belong to the current process such as
+  available registers, local stack, memory accessible through
+  available pointers will be tagged $\LOWSEC \times \LOWINT$:
+  they can be both read and written.
+\item The local state of the caller that needs to be protected from a
+  potentially untrusted callee will be marked $\HIGHSEC \times
+  \HIGHINT$. The callee should be able to neither read nor write
+  these locations.
+\item The instruction memory is read only for every process: $\LOWSEC
+  \times \HIGHINT$.
+\item We won't have examples of secret data that can be overwritten.
+  \leo{Maybe in the lazy setting?}
+\end{itemize}
+
+\subsection{Identity colorings}
+\leo{Feel free to coin another name}
+
+An {\em identity coloring} of a trace $\seq{\mach}$ is an equal-length
+sequence $\seq{\ID}$ of booleans $\IDS = \{\ME, \NOTME\}$, where
+$\ID_j = \ME$ iff the machine state $\mach_j$ is executing the same
+process as $\mach_0$.
+
+A {\em call map} is a set of instructions for a machine state that are
+(1) $JAL$s and (2) correspond to actual calls.
+
+Given a machine trace $\seq{\mach}$ and a call map this can be
+computed as follows:
+\begin{enumerate}
+\item $mach_0$ is colored $\ME$.
+\item Find the next machine state $\mach_c$ (if any) that appears in
+  the trace $\seq{mach}$ such that the instruction executed at
+  $\mach_c$ is a $JAL$ that is marked as a call in the call map.
+\item Color all states between $\mach_0$ and $\mach_c$ inclusive as
+  $\ME$. If no such state exists, color the remainder of the trace
+  $\ME$.
+\item Find the next state $\mach_r$ after $\mach_c$ in the trace (if
+  any) such that the pc of $\mach_r$ is equal to the pc of $\mach_c +
+  4$, and whose stack pointer is identical to that of $\mach_c$.
+\item Color all states between $\mach_{c+1}$ and $\mach_r$ as
+  $\NOTME$.  If no such state exists, color the remainder of the trace
+  $\NOTME$.
+\item Color $\mach_{r+1}$ as $\ME$. Go to step (2).
+\end{enumerate}
+
+\subsection{Contoured Trace}
+
+Given a contour $C$ and a machine state $M$, the {\em contoured trace}
+induced by $C$ and $M$, written $\induced{C}{M}$, is the (finite or
+infinite) sequence of pairs of (contour, machine state) pairs in which
+states in adjacent pairs are related by $\stepstounderfull$, and in which
+contours in adjacent pairs are constructed as follows:
+\begin{itemize}
+\item If in the identity coloring of $\seq{\mach}$ $\ID_i$ = $\ID_{i+1}$ then
+  $C_{i+1} = C_i$.
+\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
+  constructed from $C_i$ by setting the local state of $\mach_i$ (the
+  caller) to $\HIGHSEC \times \HIGHINT$.
+\item If $\ID_{i} = \NOTME$ and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
+  constructed from $C_i$ by setting the local state of the caller
+  to $\LOWSEC \times LOWINT$.
+\end{itemize}
+\leo{To make this more precise we need to accurately describe what
+  local state {\em is}}
+
+\subsection{Integrity}
+
+A trace $\seq{\mach}$ satisfies the {\em integrity} safety property
+with respect to a contour $\contour$, if for all components
+$\component$ where $\contour(\component) = (\_, \HIGHINT)$,
+$\mach_i(\component) = \mach_0(\component)$ for all $i$ in the trace.
+
+\leo{For confidentiality we need observations first}
+
+\subsection{Observations}
+
+A {\em component observation} is either a word or the special value
+$\bot$.  \leo{Fix notation, clashes with $\LOWINT$}  A {\em state
+  observation} is a map from machine components to component
+observations.
+
+An {\em observation} is a nonempty, finite or infinite sequence of state
+observations in which (1) adjacent elements are always distinct (``no
+stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
+the the $j^{th}$ state observation in the sequence must also be mapped to
+$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
+component's value changes from $\bot$ to anything else, it never changes
+back to $\bot$ again).
+
+In particular, the observation $\obstrace(\contour, \seq{\mach})$
+arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
+is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
+\ldots$ defined informally as follows:
+\begin{itemize}
+\item $\stateobs_0$ maps every visible component $\component$ (where
+  $\contour(\component) \in \{\LOWSEC \times \_\}$) to its value in
+  $\mach_0$ and every hidden component to $\bot$;
+\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
+  each visible component $\component$ (where
+  $\contour(\component) \in \{\LOWSEC \times \_ \}$) its value
+  from $\mach_{i+1}$, and (2) for each hidden component $\component$,
+  where $\contour (\component) = \{\HIGHSEC \times \_\}$,
+  copying its value from $\stateobs_{i}$, and (3) if the resulting
+  $\stateobs_{i+1}$ is identical to $\stateobs_{i}$, then skipping
+  $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
+\end{itemize}
+\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
+
+\subsection{Confidentiality}
+
+Given a machine state $\mach$ and a contour $\contour$, a {\em
+  variant} of $\mach$ with respect to $\contour$ is any machine state
+$\mach'$ whose components agree with $\mach$ on all components marked
+$(\LOWSEC, \_)$ by $\contour$---i.e., if $\contour(\component) =
+(\LOWSEC, \_)$, then $\mach(\component) = \mach'(\component)$.
+
+A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
+  confidentiality} safety hyperproperty with respect to a contour
+$\contour$, if, if the initials states are variants of each other $N_0
+\in \variant{C}{M_0}$, then the observations of one is a prefix of the
+observations of the other: $\obstrace(\contour, \seq{M}) \subseteq
+\obstrace(\contour, \seq{N})$.
+
+\subsection{Stack Safety}
+
+A trace $\seq{\mach}$ satisfies {\em stack safety} with respect to a
+contour $\contour$ if:
+\begin{itemize}
+\item $\seq{\mach}$ satisfies integrity with respect to $\contour$.
+\item For every variant $N \in \variant{C}{\mach_0}$, $\seq{\mach}$
+  and the the induced trace $\seq{N}$ satisfy confidentiality with
+  respect to $\contour$.
+\item Every subtrace $\seq{\mach}^r_c$ colored $\NOTME$ in the
+  identity coloring of $\seq{\mach}$ satisfies stack safety with
+  respect to the contour $\contour_c$.
+\end{itemize}
+
+\leo{OLD STUFF BEGINS HERE}
 

### 2020-04-10 main.tex
tightened up statement of identity coloring; fixed a few typos

@@
 locations that the current process does not have write access to;
-$\LOWINT$ is freely accessible memory. As examples of different
+$\LOWINT$ is freely accessible \apt{``accessible''?} memory. As examples of different
 combinations consider the following:
@@
 
-Given a machine trace $\seq{\mach}$ and a call map this can be
+Given a machine trace $\seq{\mach}$ and a call map the identity coloring can be
 computed as follows:
 \begin{enumerate}
-\item $mach_0$ is colored $\ME$.
+\item Set $i=0$.
+\item Color $\mach_i$ as $\ME$.
 \item Find the next machine state $\mach_c$ (if any) that appears in
-  the trace $\seq{mach}$ such that the instruction executed at
+  the trace $\seq{\mach}$ such that the instruction executed at
   $\mach_c$ is a $JAL$ that is marked as a call in the call map.
-\item Color all states between $\mach_0$ and $\mach_c$ inclusive as
+\item Color all states between $\mach_i$ and $\mach_c$ inclusive as
   $\ME$. If no such state exists, color the remainder of the trace
@@
   4$, and whose stack pointer is identical to that of $\mach_c$.
-\item Color all states between $\mach_{c+1}$ and $\mach_r$ as
-  $\NOTME$.  If no such state exists, color the remainder of the trace
-  $\NOTME$.
-\item Color $\mach_{r+1}$ as $\ME$. Go to step (2).
+\item If $M_r$ exists, color all states between $\mach_{c+1}$ and $\mach_{r-1}$ inclusive as
+  $\NOTME$. Otherwise, color the remainder of the trace starting from $\mach_{c+1}$ as $\NOTME$.  
+\item If $M_r$ exists, set $i=r$ and go to step (2). Otherwise, halt.
 \end{enumerate}
@@
   constructed from $C_i$ by setting the local state of the caller
-  to $\LOWSEC \times LOWINT$.
+  to $\LOWSEC \times \LOWINT$.
 \end{itemize}
@@
   local state {\em is}}
+
+\apt{How is this definition of contoured traces connected to the various properties
+  below (which seem to talk about an (uncontoured) trace plus a single contour) ?}
 

### 2020-04-10 main.tex
Edit/response to APT

@@
 locations that the current process does not have write access to;
-$\LOWINT$ is freely accessible \apt{``accessible''?} memory. As examples of different
+$\LOWINT$ is writeable. As examples of different
 combinations consider the following:
@@
   below (which seem to talk about an (uncontoured) trace plus a single contour) ?}
+\leo{It only appears in the $C_i$ part of the stack safety property}
 

### 2020-04-10 main.tex
clear/add comments

@@
 
+\apt{This definition may need refinement to reflect the fact that entry and exit sequences
+  are multiple instructions; in the middle of these sequences the state might be neither $\ME$ nor $\NOTME$.}
+
 \subsection{Contoured Trace}
@@
   local state {\em is}}
-
-\apt{How is this definition of contoured traces connected to the various properties
-  below (which seem to talk about an (uncontoured) trace plus a single contour) ?}
-\leo{It only appears in the $C_i$ part of the stack safety property}
 
@@
 
+\apt{This definition seems to have vestiges of being based on contoured
+  traces; for single contours, there is no possibility of changing from
+  $\bot$ to anything else.}
+
 \subsection{Confidentiality}
@@
 \obstrace(\contour, \seq{N})$.
+
+\apt{Does this differ from lockstep equivalence?}
 

### 2020-04-19 main.tex
one more question
Notes
Updated for the calling convention/eager setting

--- 2020-04-10 main.tex
@@
 
+\subsection{Calling Convention}
+
+\leo{Mainly @Andrew, check if this makes sense for the eager setting.}
+
+We adopt the following calling convention:
+\begin{itemize}
+\item First the caller increases the stack pointer to allocate stack space
+  for arguments and result and zeroes out anything not written.
+\item Then a $\JAL$ happens.
+\item Then the entry sequence of the callee follows which increases
+  the stack pointer again to allocate as much local space as needed
+  (e.g. for the return address), and it is all zeroed out if not
+  explicitly written. \leo{This can be lifted by dealing with
+    stack allocation in the contours}
+\end{itemize}
+
+A {\em call map} is a map from a sequence of instructions (read: {\em
+  a call sequence}) to a natural number (the number of arguments and
+results of the call plus the size of the local state of the callee,
+which corresponds to stack locations that will be made available to
+the callee).
+
 \subsection{Contours}
@@
   \times \HIGHINT$.
-\item We won't have examples of secret data that can be overwritten.
-  \leo{Maybe in the lazy setting?}
+\item Uninitialized memory can be written, but never read. This is only
+  applicable in a more lazy setting where we don't zero out memory.
 \end{itemize}
@@
 
+\leo{Went with S for ``self'' and O for ``other''. M/NM was confusing
+  with the machine states also being M. T/F didn't really work
+  with transitions. Open to other suggestions.}
 An {\em identity coloring} of a trace $\seq{\mach}$ is an equal-length
-sequence $\seq{\ID}$ of booleans $\IDS = \{\ME, \NOTME\}$, where
-$\ID_j = \ME$ iff the machine state $\mach_j$ is executing the same
-process as $\mach_0$.
-
-A {\em call map} is a set of instructions for a machine state that are
-(1) $JAL$s and (2) correspond to actual calls.
+sequence $\seq{\ID}$ of ids ($\IDS = \{\ME, \NOTME, \TRANS\}$, where
+$\ID_j = \ME$ if the machine state $\mach_j$ is executing the same
+process as $\mach_0$, $ID_j = \NOTME$ if it is executing a different
+process, and $ID_j = \TRANS$ during a transition (e.g. during a call
+or return sequence).
 
 Given a machine trace $\seq{\mach}$ and a call map the identity coloring can be
-computed as follows:
+computed by iterating through the trace:
 \begin{enumerate}
-\item Set $i=0$.
-\item Color $\mach_i$ as $\ME$.
-\item Find the next machine state $\mach_c$ (if any) that appears in
-  the trace $\seq{\mach}$ such that the instruction executed at
-  $\mach_c$ is a $JAL$ that is marked as a call in the call map.
-\item Color all states between $\mach_i$ and $\mach_c$ inclusive as
-  $\ME$. If no such state exists, color the remainder of the trace
-  $\ME$.
-\item Find the next state $\mach_r$ after $\mach_c$ in the trace (if
-  any) such that the pc of $\mach_r$ is equal to the pc of $\mach_c +
-  4$, and whose stack pointer is identical to that of $\mach_c$.
-\item If $M_r$ exists, color all states between $\mach_{c+1}$ and $\mach_{r-1}$ inclusive as
-  $\NOTME$. Otherwise, color the remainder of the trace starting from $\mach_{c+1}$ as $\NOTME$.  
-\item If $M_r$ exists, set $i=r$ and go to step (2). Otherwise, halt.
+\item Start by coloring $\mach_0$ as $\ME$.
+\item Color every subsequent state as $\ME$, until the first machine
+  state (if any) at which the instruction executed is a
+  the beginning of a call sequence.
+\item Color the entire call sequence $\TRANS$. This necessarily
+  contains a machine state $\mach_c$ which is a $\JAL$.
+\item Color every subsequence state as $\NOTME$, until the first state
+  $\mach_r$ after $\mach_c$ in the trace (if any) such that the pc of
+  $\PC{\mach_r}$ is equal to $\PC{\mach_c} + 4$, and whose stack
+  pointer is identical to that of $\mach_c$. Then mark $\mach_r$ as
+  $\ME$ and got to step (2).
 \end{enumerate}
@@
 \begin{itemize}
-\item If in the identity coloring of $\seq{\mach}$ $\ID_i$ = $\ID_{i+1}$ then
+\item If in the identity coloring of $\seq{\mach}$ the identity of
+  consecutive states is preserved ($\ID_i$ = $\ID_{i+1}$) then
   $C_{i+1} = C_i$.
-\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
+\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \TRANS$, then $C_{i+1}$ is
   constructed from $C_i$ by setting the local state of $\mach_i$ (the
-  caller) to $\HIGHSEC \times \HIGHINT$.
-\item If $\ID_{i} = \NOTME$ and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
+  caller) to $\HIGHSEC \times \HIGHINT$. The local state is comprised
+  of everything below the stack pointer at $\mach_i$.
+\item If $\ID_{i} = \TRANS$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
+  constructed from $C_i$ by making the top $n$ locations in the stack,
+  where $n$ is the mapping of the transition call sequence in the call map,
+  readable and writeable $\LOWSEC \times \LOWINT$.
+\item If $\ID_{i} = \TRANS$ and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
   constructed from $C_i$ by setting the local state of the caller
-  to $\LOWSEC \times \LOWINT$.
-\end{itemize}
-\leo{To make this more precise we need to accurately describe what
-  local state {\em is}}
-
-\apt{How is this definition of contoured traces connected to the various properties
-  below (which seem to talk about an (uncontoured) trace plus a single contour) ?}
-\leo{It only appears in the $C_i$ part of the stack safety property}
+  to $\LOWSEC \times \LOWINT$, and making everything above the stack
+  pointer $\HIGHSEC \times \HIGHINT$ again.
+\end{itemize}
 
@@
 In particular, the observation $\obstrace(\contour, \seq{\mach})$
-arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
+arising from a contoured trace $\seq{(\contour, \mach)}$
 is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
@@
 \item $\stateobs_0$ maps every visible component $\component$ (where
-  $\contour(\component) \in \{\LOWSEC \times \_\}$) to its value in
+  $\contour_0(\component) \in \{\LOWSEC \times \_\}$) to its value in
   $\mach_0$ and every hidden component to $\bot$;
@@
   each visible component $\component$ (where
-  $\contour(\component) \in \{\LOWSEC \times \_ \}$) its value
+  $\contour_{i+1}(\component) \in \{\LOWSEC \times \_ \}$) its value
   from $\mach_{i+1}$, and (2) for each hidden component $\component$,
-  where $\contour (\component) = \{\HIGHSEC \times \_\}$,
+  where $\contour_{i+1} (\component) = \{\HIGHSEC \times \_\}$,
   copying its value from $\stateobs_{i}$, and (3) if the resulting
@@
 \in \variant{C}{M_0}$, then the observations of one is a prefix of the
-observations of the other: $\obstrace(\contour, \seq{M}) \subseteq
-\obstrace(\contour, \seq{N})$.
+observations of the other: $\obstrace(\seq{\contour, \mach}) \subseteq
+\obstrace(\seq{\contour, N})$.
+
+\apt{Does this differ from lockstep equivalence? And why prefix?}
 
@@
 
+\bcp{Needs to be formulated coinductively?!}
+
+\bcp{What about unitialized memory (written by previous calls)}
+
+\section{Related work}
+
+\bcp{Make sure to mention that we are NOT enforcing well-bracketed control
+  flow a la Birkedal et al!}
+
+\cleardoublepage
 \leo{OLD STUFF BEGINS HERE}

### 2020-04-21 main.tex
a few comments/queries

@@
   contains a machine state $\mach_c$ which is a $\JAL$.
-\item Color every subsequence state as $\NOTME$, until the first state
+\item Color every subsequent state as $\NOTME$, until the first state
   $\mach_r$ after $\mach_c$ in the trace (if any) such that the pc of
@@
   pointer is identical to that of $\mach_c$. Then mark $\mach_r$ as
-  $\ME$ and got to step (2).
+  $\ME$ and go to step (2).
 \end{enumerate}
@@
   readable and writeable $\LOWSEC \times \LOWINT$.
-\item If $\ID_{i} = \TRANS$ and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
+\item If $\ID_{i} = \TRANS$ \apt {$\NOTME$ ? } and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
   constructed from $C_i$ by setting the local state of the caller

### 2020-04-26 LazyStackProperties.tex
Add tex about lazy properties

\documentclass{article}

\usepackage{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{xcolor}

\newcommand{\apt}[1]{\textcolor{blue}{{[APT:~#1]}}}
\newcommand{\sna}[1]{\textcolor{green}{{[SNA:~#1]}}}

\title{Lazy Stack Properties}
\author{Sean Anderson, Andrew Tolmach}

\begin{document}
  \maketitle

  \section{Introduction}

    Subroutine and coroutine models suggest logical encapsulation of stack frames that is not typically
    enforced in low level code. With an additional enforcement mechanism such as a tag-based reference monitor,
    we can prove integrity and confidentiality properties of subroutines and coroutines.
    The most obvious integrity and confidentiality properties reflect the behavior of eager enforcement mechanisms
    that prevent illegal writes and writes immediately. But lazy tag policies (see Roessler and DeHon, Section IV.A
    \cite{Roessler2018ProtectingHardware}) do not enforce these eager properties.
    In this paper we present properties that capture the inability of unauthorized reads and writes
    to interfere with observable behavior.

    In Section \ref{sec:model} we model a machine as a transition system with built-in coroutine and subroutine
    operations. Section \ref{sec:spec} gives the specification of integrity in terms of accessible and inaccessible
    memory, then Sections \ref{sec:propint} and \ref{sec:propcon} translates this specification into a property
    that captures when a violation would effect observable behavior.

    Our lazy integrity property, {\it observable integrity}, defines an idealized ``safe'' variant machine that fulfills
    eager integrity properties, but otherwise behaves identically to the original machine. The safe machine rolls
    back writes to inaccessible locations when they become accessible. Our lazy property is that the observable
    behavior of the idealized machine extends that of the reference machine.

    \paragraph{Eager vs. Lazy Enforcement}

      For illustrative purposes, consider the following program fragments in a simple language with subroutines.
      Suppose that addresses 20-29 are the stack frame of our main routine:

      \vspace*{0.2in}
      \begin{minipage}{.45\textwidth}
        \noindent
        {\tt
          0: call 10 \\
          1: if *20 goto 4 \\
          2: print 0 \\
          3: halt \\
          4: print 1 \\
          5: halt \\
          \vdots \\
          10: *21 := 0 \\
          11: return
          }
      \end{minipage}
      \begin{minipage}{.45\textwidth}
        \noindent
        {\tt
          0: call 10 \\
          1: if *20 goto 4 \\
          2: print 0 \\
          3: halt \\
          4: print 1 \\
          5: halt \\
          \vdots \\
          10: *20 := 0 \\
          11: return
          }
     \end{minipage}
      \vspace*{0.2in}

      The subroutine at 10 writes into its caller's stack, which is erroneous. An eager enforcement mechanism might
      be to tag 20-29 with a unique {\it color} matching a tag on the program counter (the {\it PC Tag}), change the
      tag on the program counter during the call out, and cause the system to fail-stop anytime a write occurs to an
      address not matching the program counter's tag. Then both programs will fail-stop. But this would require writing
      the color to all 10 addresses, which is expensive if most of those addresses are not actually used.

      In the lazy variation we construct our system to initially give all memory a tag of {\it init} and cause
      writes to tag the written location with a color matching that of the program counter. We further cause
      reads to fail-stop if the value being read does not match the program counter's. So failure occurs only when
      a location is read by a routine other than the one that wrote it. Note that the left program will not fail-stop
      under this policy, as address 21 is overwritten by the callee but never read. The right program, on the other
      hand, still fails.

      Both of these policies will be described in more detail in Section \ref{sec:pols}.
      The lazy mechanism enforces a weaker property than that of the eager mechanism since it fails fewer programs.
      Intuitively, it seems clear that this property still reflects a desirable notion of integrity;
      while the left program erred, it did so invisibly.

  \section{The Model}
    \label{sec:model}

    First, some types. \(\mathtt{PC}\) is the program counter.
    We have words and addresses:
    \[w,a \in \mathcal{W}\]
    Registers:
    \[r \in \mathcal{R}\]
    Stack names:
    \[S \in \mathcal{S}\]
    A stack is specified by a 4-tuple of addresses: base, frame, pointer, and top. Stack names are mapped to stacks
    by a stack configuration:
    \[\sigma \in \Sigma = \mathcal{S} \rightarrow \mathcal{W} \times \mathcal{W} \times \mathcal{W} \times \mathcal{W}\]
    Machine Components:
    \[K \in \mathcal{W} \cup \mathcal{R} \cup \{\mathtt{PC}\} \cup \{\mathtt{FP}\} \cup \{\mathtt{SP}\}\]
    States:
    \[M \in \mathcal{M} = \mathcal{K} \rightarrow \mathcal{W}\]
    Machine configurations:
    \[T \in \mathcal{T} = \mathcal{M} \times \Sigma \times \mathcal{S}\]

    If \(T = M,\sigma,S\) and \(\sigma(S) = b,f,p,t\), then \(M(\mathtt{FP})\) is \(f\) and \(M(\mathtt{SP})\) is \(p\).
    These special registers only change when \(\sigma\) or \(S\) changes.

    The step relation \(\longrightarrow\) represent the execution of a single instruction or pseudo-instruction,
    \(\cdot \xrightarrow{\cdot} \cdot \subseteq \mathcal{T} \times (\mathcal{W} \cup \{\tau\}) \times \mathcal{T}\).
    It may be labeled \(\xrightarrow{w}\) or silent \(\xrightarrow{\tau}\), which we will write \(\longrightarrow\).
    For generality we leave the set of instructions open except to specify the existence of pseudo-instructions for
    coroutine yields, subroutine calls and returns, and output.

    A yield instruction {\tt Yield \(S\)} saves the {\tt PC} to the current stack, changes the current
    stack to \(S\), and loads the {\tt PC} from the new stack.

    The call instruction {\tt Call \(w\) \(a\)} pushes a frame of size \(w\) onto the stack and sets the {\tt PC}
    to address \(a\). The return instruction {\tt Return} pops the top frame of the current stack and sets
    the {\tt PC} to the next instruction from a matched call -- see below.

    An output instruction, {\tt Output \(r\)}, takes a register \(r\) and steps by a transition labeled with the
    contents of \(r\). It is the only instruction that can produce a labeled transition.

    An execution is a sequence of configurations and labels, possibly infinite, of the form
    \(T_0 w_0 T_1 w_1 \dots\), such that \(T_i \xrightarrow{w_i} T_{i+1}\). We will omit \(w_i\) when
    it is \(\tau\). The behavior of an execution, or its trace, written \(\mathit{behav}(\overline{T})\), is the
    projection of the labels of \(\overline{T}\). A {\it property} is a set of executions, and a system
    enjoys those properties for which all of its possible executions are in the property.

  \paragraph{Pseudoinstruction Semantics} Here we describe the behavior of the call and return pseudoinstructions.
    Stacks grow upward; initially each stack's state is \(b,b,b+2,t\) where \(b\) is its base and \(t\) its top,
    and no stacks overlap. A separate code section is non-overlapping with the stacks. The program counter is initially
    in the code section.

    \vspace{.5cm}

    \begin{tabular}{c c}
      \(M_1(M_1(\mathtt{PC})) = \mathtt{Call\ w\ a}\) & \(\sigma(S) = b,f_1,p_1,t\) \\
      \(f_2 = p_1\) & \(p_2 = f_2 + \mathtt{w} + 2\) \\
      \(M_2 = M_1[f_2 \mapsto M_1(\mathtt{PC})][f_2+1 \mapsto f_1][\mathtt{PC} \mapsto a]\) &
        \(\sigma_2 = \sigma_1[S \mapsto b,f_2,p_2,t]\) \\
      \hline
      \multicolumn{2}{c}{\(M_1,\sigma_1,S \longrightarrow M_2,\sigma_2,S\)}
    \end{tabular}

    \vspace{.5cm}

    \begin{tabular}{c c c}
      \(M_1(M_1(\mathtt{PC})) = \mathtt{Return}\) & \(\sigma(S) = b,f_1,p_1,t\) \\
      \(M_2 = M_1[\mathtt{PC} \mapsto M_1(f_1)]\) & \(f_2 = M_1(f_1+1)\) & \(\sigma_2 = \sigma_1[S \mapsto b,f_2,f_1,t]\) \\
      \hline
      \multicolumn{3}{c}{\(M_1,\sigma_1,S \longrightarrow M_2,\sigma_2,S\)}
    \end{tabular}

    \vspace{.5cm}

    In an execution \(\overline{T} = \dots T_1 \dots T_2 \dots\), if \(T_1 = M_1,\sigma_1,S\)
    and \(T_2 = M_2,\sigma_2,S\), where \(\sigma_1(S) = \sigma_2(S)\) and \(T_1(\mathtt{PC}) = T_2(\mathtt{PC})\)
    --- and if for all states \(T' = M',\sigma',S\) between \(T_1\) and \(T_2\), \(\sigma'(S) \not = \sigma_1(S)\) ---
    \(T_1\) and \(T_2\) are said to be {\it call brackets}. We will use this concept to define an eager notion
    of integrity.

    Now we describe the behavior of the yield pseudoinstruction.

    \vspace{.5cm}

    \begin{tabular}{c c}
      \(M_1(M_1(\mathtt{PC})) = \mathtt{Yield} S_2\) & \(\sigma(S_1) = b,f_1,p_1,t\) \\
      \(\sigma(S_2) = b_2,f_2,s_2,t_2\) & \(M_2 = M_1[s_2 \mapsto M_1(\mathtt{PC})][\mathtt{PC} \mapsto M_1(s_2)]\) \\
      \hline
      \multicolumn{2}{c}{\(M_1,\sigma,S_1 \longrightarrow M_2,\sigma,S_2\)}
    \end{tabular}

    \vspace{.5cm}

    In an execution \(\overline{T} = \dots T_1 \dots T_2 \dots\), the \(T_1\) and \(T_2\) are considered
    {\it yield brackets} if the have the same active stack, and if every state between them has a different
    stack. Again this will be used in eager notions of integrity.

    Finally, output instructions generate a labeled transition.

    \vspace{.5cm}

    \begin{tabular}{c}
      \(M(M(\mathtt{PC})) = \mathtt{Output}\ r\) \\
      \hline
      \(M,\sigma,S \xrightarrow{M(r)} M[\mathtt{PC} \mapsto \mathtt{PC} + 1],\sigma,S\)
    \end{tabular}

    \vspace{.5cm}

\section{Security Levels and Contours}
    \label{sec:spec}

    At any given state we assign each component a security level. It is either accessible to the active
    subroutine within the active coroutine ({\tt A}), accessible but unallocated ({\tt AL}),
    inaccessible due to belonging to a caller in the local stack ({\tt IL}), inaccessible due to belonging to
    a foreign coroutine({\tt IF}), or control/code ({\tt K}).
    A contour \(C \in \mathcal{C} = \mathcal{K} \rightarrow \{\mathtt{A},\mathtt{AL},\mathtt{IL},\mathtt{IF},
    \mathtt{K}\}\) maps components to security levels, and within our model there is a clear mapping from pairs
    of stack configurations and active stack ids to contours.
    Here \(\mathit{control(T)}\) is the set of addresses reserved for control information: the first two addresses
    of each frame and the top of each inactive stack area.

    \[\begin{split}
      \mathit{contour}(\sigma,S) = \text{Let } \sigma(S) = b,f,p,t \text{ in} \\
      \lambda K . \begin{cases}
        \mathtt{A} & \text{if } K \not \in \mathcal{W} \\
        \mathtt{K} & \text{else if } K \in \mathit{control(T)} \\
        \mathtt{A} & \text{else if } f \leq K < p \\
        \mathtt{AL} & \text{else if } p \leq K < t \\
        \mathtt{IL} & \text{else if } b \leq K < f \\
        \mathtt{IF} & \text{else} \\
      \end{cases}
    \end{split}\]

    In this context we are concerned with protecting data, and will assume control is protected by some lower
    level policy such that only calls, returns, and yields may update {\tt K} components.

  \section{Integrity Properties}
    \label{sec:propint}

    \paragraph{Local, Frame, and Stack Integrity}

      From a contour as defined above, we have an obvious notion of integrity: a transition system
      enjoys {\it local integrity} when, for all \(T_1 = M_1,\sigma_1,S_1\) and \(T_2 = M_2,\sigma_2,S_2\), if
      \(T_1 \xrightarrow{w} T_2\) and \(M_1(K) \not = M_2(K)\), \(\mathit{contour}(\sigma_1,S_1)(K) =
      \{\mathtt{A},\mathit{AL}\}\). In other words, individual state transitions respect the contour of the
      first state, only modifying memory locations within the active stack above the frame pointer. This
      property is very strong and convenient to test for eager enforcement mechanisms as it only talks about
      a single step.

      Local integrity implies a pair of marginally weaker properties that we call frame integrity and stack integrity.
      A system enjoys {\it frame integrity} if for all \(K\) and \(\overline{T} = \dots T_1 \dots T_2 \dots\) where
      \(T_1\) and \(T_2\) are a call bracket, if \(\mathit{contour}(T_1)(K) = \mathtt{A}\) then
      \(T_1(K) = T_2(K)\). In other words, when we make a call, our accessible state is unchanged on return.

      A system has \emph{stack integrity} if, for all \(K\) and all \(\overline{T} = \dots T_1 \dots T_2 \dots\) where
      \(T_1\) and \(T_2\) are a yield bracket for all \(K\), if \(\mathit{contour}(\sigma_1,S_1)(K) \in
      \{\mathtt{IL},\mathtt{A},\mathtt{AL}\}\), then \(T_1(K) = T_2(K)\). In short, when a we yield control,
      our entire stack is unchanged when control is yielded back.

      Frame integrity weakens local integrity by focusing on changes visible to a caller, and stack
      integrity on changes visible to an entire coroutine. A system that fulfills both properties may write outside
      of accessible memory, but only if that memory will somehow be restored by the time it becomes accessible.

    \paragraph{Observable Integrity}

      In order for a lazy policy to fulfill an integrity property, the property must be weaker still --- specifically,
      it must allow for writes that violate the contour and remain after control changes, but only in ways that are
      still invisible. We construct this property by comparing the behavior of a system to the behavior of an
      idealized system that fulfills frame  and stack integrity by construction and that otherwise behaves like
      \(\longrightarrow\). We call this system \(\xrightarrow{\text{\sc Safe}}\), and it enforces both forms of
      integrity by rolling back illegal writes to an inactive frame when it becomes active again.

      Let \(\cdot,\cdot \xrightarrow{\text{\sc Safe}} \cdot,\cdot \subseteq
      \mathcal{T} \times \mathcal{M} \times \mathcal{T} \times \mathcal{M}\). We carry an
      extra memory configuration remembering the last known value for each component outside the active
      frame. If \(M_1,\sigma_1,S_1 \longrightarrow M_2,\sigma_2,S_2\), then for all \(\hat{M_1}\),
      \((M_1,\sigma_1,S_1),\hat{M_1} \xrightarrow{\text{\sc Safe}} (M_2',\sigma_2,S_2),\hat{M_2}\), where:
      \[\hat{M_2}(K) =
        \begin{cases}
          M_2(K) & \text{if } \mathit{contour}(\sigma_1,S_1)(K) \in \{\mathtt{A},\mathtt{AL}\} \\
          \hat{M_1}(K) & \text{else} \\
        \end{cases}\]
      \[M_2'(K) =
        \begin{cases}
          \hat{M_1}(K) & \text{if } \mathit{contour}(\sigma_1,S_1)(K) \in \{\mathtt{IL},\mathtt{IF}\} \text{ and }
            \mathit{contour}(\sigma_2,S_2)(K) \in \{\mathtt{A},\mathtt{AF}\} \\
          M_2(K) & \text{else} \\
        \end{cases}\]

      A series of safe states \(\overline{U} = (T_0,\hat{M_0}), \ldots\) with adjacent states related by
      \(\xrightarrow{\text{\sc Safe}}\) is a \emph{safe execution}, and its behavior is defined as the behavior
      of the execution formed by projecting the state components of the series:
      \[\mathit{behav}(\overline{U}) = \mathit{behav(map(fst,\overline{U}))}\]

      A system enjoys {\it observable integrity} if, for any normal execution \(\overline{T} = T_0,\ldots\) with
      \(T_0 = M_0,\sigma_0,S_0\), there exists a safe execution \(\overline{U} = (T_0,M_0),\ldots\) such that
      \(\mathit{behave}(\overline{T})\) is a prefix of \(\mathit{behave}(\overline{U})\). That is, the safe execution
      may not get stuck when the normal execution does, but up to that point the behaviors are identical.

  \subsection{Example}

    Here we examine a more detailed example. We will have two coroutines, with stacks \(S_1\) and \(S_2\), and
    three routines \(A\), \(B\), and \(C\). Coroutine 1 will start in \(A\), call \(B\), and yield to coroutine
    2, which starts in routine \(C\). Let \(\sigma_{init}(S_1) = 0,0,0,10\) and \(\sigma_{init}(S_2) = 11,11,11,21\).
    Subroutine \(A\) will start at address 30, \(B\) at 40, and \(C\) at 50. Here is our code, in a RISC-like style
    that folds some arithmetic into the instructions for simplicity:

    {\tt
    \begin{tabular}{l l | l}
      \vdots \\
      30 & add SP 3 SP & {\it allocate some space for local storage} \\
      31 & store 1 (FP+2) & {\it FP+2 is 2} \\
      32 & store 1 (FP+3) & {\it FP+3 is 3} \\
      33 & call 2 40 & {\it call B} \\
      34 & load (FP+2) r1 & \\
      35 & load (FP+3) r2 & \\
      36 & brne r1 r2 38 & {\it if our first two locals are equal,} \\
      37 & output 1 & {\it we output 1,} \\
      38 & halt & {\it and either way we halt} \\
      \vdots \\
      40 & store 5 (FP-1) & {\it FP-1 is 4} \\
      41 & load (FP-2) r3 & {\it FP-2 is 3} \\
      42 & mov FP r4 & {\it save frame pointer for second coroutine} \\
      42 & yield \(\mathtt{S_1}\) & {\it yield to second coroutine} \\
      43 & return \\
      \vdots \\
      50 & store r3 (r4-2) & {\it store the value read from 3 back to 3} \\
      51 & yield \(\mathtt{S_1}\) & \\
      \vdots
    \end{tabular}}

    This program makes multiple illegal writes. \(B\) writes into a location in \(A\)'s stack frame,
    and so does \(C\), but in the former case the location is never read, and in the latter the value written
    will be the same value that was there before. Under our \(\xrightarrow{\text{\sc Safe}}\) system,
    the values at addresses 3 and 4 are rolled back to their initial values, so 3 maps to 1 and 4 maps
    to 0. Under \(\longrightarrow\), 3 maps to 1 and 4 maps to 5. The value at 3 still matches the value
    at 2 in either run, so we output 1 and then halt. The programs have the same behavior, and observational
    integrity is maintained.

  \section{Confidentiality Properties}
    \label{sec:propcon}

    Confidentiality in general is commonly modeled as a {\it noninterference} property: if some components
    are considered secret, a system enjoys noninterference when it produces the same behavior regardless
    of the initial values of the secret components. We will present confidentiality analogs of each of our
    integrity properties.

    \paragraph{Local Noninterference}

      The strongest form of confidentiality is the dual of local integrity: {\it local noninterference}.
      For any transition \(T_1 \xrightarrow{w} T_2\), we use the contour of \(T_1\) to determine the accessibility
      of each component. Local noninterference holds if any inaccessible component can be varied in \(T_1\) without
      changing the values of any accessible component in \(T_2\). Formally, we define view-preserving
      equivalence relations between states:
      \[\begin{split}
        M_1,\sigma_1,S_1 \sim_L M_2,\sigma_2,S_2 \triangleq & S_1 = S_2 \text{ and } \sigma_1(S_1) = \sigma_2(S_2) \\
        & \text{and for all } K \text{, if } \mathit{contour}(\sigma_1,S_1)(K) \not = \mathtt{IL} \\
        & \text{then } M_1(K) = M_2(K) \\
      \end{split}\]
      \[\begin{split}
        M_1,\sigma_1,S_1 \sim_F M_2,\sigma_2,S_2 \triangleq & S_1 = S_2 \text{ and } \sigma_1(S_1) = \sigma_2(S_2) \\
        & \text{and for all } K \text{, if } \mathit{contour}(\sigma_1,S_1)(K) \not = \mathtt{IF} \\
        & \text{then } M_1(K) = M_2(K) \\
      \end{split}\]

      \[T_1 \sim T_2 \triangleq T_1 \sim_L T_2 \text{ and } T_1 \sim_F T_2\]

      Then a system enjoys local noninterference if for any transition \(T_1 \xrightarrow{w} T_2\), for all
      \(T_1'\) such that \(T_1 \sim T_1'\), there exists some \(T_2'\) such that \(T_2 \sim T_2'\) and
      \(T_1' \xrightarrow{w} T_2'\).

    \paragraph{Frame and Stack Interference}

      Local noninterference is very strong. Suppose a function writes outside of its
      accessible memory, then immediately reads from that location. We would not consider that read to
      be a failure of confidentiality in any reasonable attacker model. We can weaken our notion of
      confidentiality to reflect that information originating in the active routine need not be secret
      from it, even if stored in a secret location. We do this by varying the state only once
      for each activation. Then we need to check that the variation has not impacted the behavior of
      the execution. Conservatively, we will do so by comparing states at the end of the activation.

      The analog of frame integrity is {\it frame noninterference}. A system enjoys frame noninterference when
      for any \(\overline{T} = \dots T_1 T_2 \overline{T}' T_3 T_4 \dots\) where \(T_1\) and \(T_4\) bracket a
      subroutine, and any \(T_2'\) such that \(T_2 \sim_L T_2'\), there is some \(T_3'\) such that
      \(T_3 \sim_L T_3'\), \(T_2' \xrightarrow{\overline{w}}^* T_3'\), and \(\overline{W} =
      \mathit{behav}(T_2 \overline{T}' T_3)\). Note that we use \(\sim_L\), because a subroutine containing a yield
      might well have its behavior depend on a foreign stack, and protecting from that is a separate property.

      Then the analog to stack integrity is {\it stack noninterference}, which a system enjoys if for any
      execution \(\overline{T} = T_1 T_2 \overline{T}' T_3 T_4\) such that every state from \(T_2\) to \(T_3\)
      shares the same stack and neither \(T_1\) nor \(T_4\) do, for any \(T_2'\), if \(T_2 \sim_F T_2'\)
      there exists some \(T_3'\) such that \(T_3 \sim_F T_3'\), \(T_2' \xrightarrow{\overline{w}} T_3'\), and
      \(\overline{w} = \mathit{behav}(T_2 \overline{T}' T_3)\).

      Local noninterference implies both frame and stack noninterference. Naturally lazy enforcement
      does not enjoy local noninterference, but what about frame and stack noninterference? Stack noninterference
      would be violated by a coroutine that wrote outside the active stack, yielded, and then
      copied the value from that location into its active stack. But if no other coroutine tried to overwrite
      that location, it would not violate the policy. So stack noninterference would still be too strong.
      Frame noninterference does not have this issue, provided that the lazy policy distinguishes between
      distinct activations of the same frame depth. This is our final policy in the micropolicy section, below.
      We are still developing a property that works for coroutines.

    \paragraph{Observable Frame Noninterference}

      Suppose a function reads a value from outside of its accessible memory into
      a register, then immediately overwrites the register without influencing observations. This is benign,
      but our frame and stack noninterference properties do not allow it. They take a shortcut
      by checking the entire accessible state at the end of the activation. Our final policy enforces the
      frame noninterference property, but in principle we would like to present a looser property
      that captures the intuition that hidden information should not interfere with observable behavior.

      Here we attempt to define a looser noninterference property. For simplicity now let us assume a system with
      only one stack, and we will deal only with calls and returns.
      A system enjoys {\it observable frame noninterference} if, for all
      \(\overline{T} = \overline{T}_0 (M_1,\sigma_1,S) (M_2,\sigma_2,S) \dots (M_3,\sigma_3,S) \overline{T}_4\)
      such that \(M_1,\sigma_1,S\) and \(M_3,\sigma_3,S\) bracket a call, the following holds for any
      \(T_2' = M_2',\sigma_2,S\) such that \(M_2,\sigma_2,S \sim_L T_2'\).

      \begin{itemize}
        \item There is some \(T_3' = M_3',\sigma_3,S\) such that \(T_2' \xrightarrow{\overline{w}}^* T_3'\)
        \item \(T_3'\) has the same frame pointer as \(T_1\)
        \item Let \(T_3'' = M_3'', \sigma_3, S\) where
            \[M_3''(K) =
            \begin{cases}
              M_1(K) & \text{if } M_2(K) \not = M_2'(K) \\
              M_3'(K) & \text{else} \\
            \end{cases}\]
        \item \(\mathit{behav}(\overline{T}) = \mathit{behav}(\overline{T}_0 +\!\!\!+
          \overline{w} +\!\!\!+ \mathit{behav}(T_3'' \overline{T}_4')\) where \(T_3'' \overline{T}_4'\)
          is an execution.
      \end{itemize}

      So, \(T_3''\) preserves those values from \(T_3'\) that were not changed in the variation.
      Note that this is a rollback as seen in \(\xrightarrow{\text{\sc Safe}}\), so this property
      subsumes observable integrity. (We could narrow it to handle only confidentiality by not rolling
      back values that have been written to after being varied.)

%    \paragraph{Dynamic Noninterference}
%      
%      Neither local noninterference nor stack noninterference works well for lazy policies. The problem
%      with coroutines having a sustained state is that a write that makes future (apparently invalid) reads
%      turn out to be valid never goes away. The entire execution history of a given coroutine is always
%      relevant, and can't be forgotten across coroutine yields. So to use the noninterference model, we need
%      to think about variant states for only those components that have changed since the last time a given
%      subroutine or coroutine was active.

%     We define a function \(\mathit{lastvis} : \mathcal{T}^* \rightarrow \Sigma \rightarrow \mathcal{S}
%      \rightarrow \mathcal{K} \rightarrow \mathcal{W} + \bot\) that gives us the last value of a component
%      as seen by a given stack configuration, or \(\bot\) if no such value has been visible.

%      \[\mathit{lastvis}(\varepsilon, \sigma, S, K)  = \bot\]
%      \[\mathit{lastvis}(\overline{T} (M_i,\sigma_i,S_i), \sigma, S, K) =
%        \begin{cases}
%          M_i(K) & \text{if } \mathit{contour}(\sigma,S)(K) \in \{\mathtt{A},\mathtt{AL}\} \\
%          \mathit{lastvis}(\overline{T}, \sigma, S, K) & \text{else} \\
%        \end{cases}\]

%      A system enjoys {\it Dynamic Stack Noninterference} if for all executions \(\overline{T} = \overline{T}_0
%      (M_1, \sigma_1, S_1)(M_2, \sigma_2, S_2) \dots\) such that \(S_1 \not = S_2\), the following holds.
%      Let \(M_2'\) be a state where for all \(K\), if \(M_2(K) = \mathit{lastvis}(\overline{T}_0,\sigma_2,S_2,K)\)
%      or if \(\mathit{contour}(\sigma_2,S_2)(K) \in \{\mathtt{A},\mathtt{AL}\}\), then \(M_2'(K) = M_2(K)\).
%      Then there is some \(\overline{T}' = M_2' \dots\) such that \(\mathit{behav}(\overline{T}) = 
%      \mathit{behav}(\overline{T}_0) +\!\!\!+ \mathit{behav}(\overline{T}')\).

%      Dynamic stack noninterference is similar to observation integrity, in that it creates variant
%      states and then insists that the observable behavior of the system is identical in the execution
%      proceeding from either state.

    \paragraph{Safe Initialization}

      Not strictly a type of confidentiality, {\it safe initialization} -- the requirement that data be
      written before it is read -- is nevertheless important. In particular, frame noninterference avoids
      the flaw of stack noninterference by virtue of subroutine activations that share a stack frame not
      being treated as the same entity. Nevertheless, an activation using an overlapping frame to a prior
      activation could well be influenced by it. For that matter a caller could influence its callee
      improperly by writing to the callee's accessible memory, if the callee reads uninitialized memory.
      So we give a safe initialization property.

      For an execution \(\overline{T} = \dots (M_1,\sigma_1,S_1) (M_2,\sigma_2,S_2) \overline{T}_2\), let \(M_1'\)
      be a variant of \(M_1\) defined as follows. For all \(K\), if \(\mathit{contour}(\sigma_1,S_1)(K) =
      \mathtt{AL}\) and \(\mathit{contour}(\sigma_2,S_2) = \mathtt{A}\), \(M_1'(K)\) may have any value;
      otherwise \(M_1'(K) = M_1(K)\). In other words, \(M_1'\) varies from \(M_1\) at only those components
      that have just been allocated. Then there is an execution \((M_1',\sigma_1,S_2) T_2' \overline{T}_2'\)
      such that \(\mathit{behav}((M_2,\sigma_2,S_2) \overline{T}_2) = \mathit{behav}(T_2' \overline{T}_2')\).

%  \section{Testing}

%    This section focuses on how these properties would behave in a randomized testing framework, such as
%    QuickChick. Some of them present challenges. First we will describe the mechanism for testing local
%    properties, which is quite straightforward. The general framework allows us generate states and step
%    through their execution, and to make modified variations and step through their executions as well.

%    \paragraph{Local Integrity and Noninterference}

  \section{Micropolicies}
    \label{sec:pols}

    Here we describe some tag-based policies that enforce the integrity properties above. First we give the eager
    version, then the lazy version. In each case we modify the type of a memory configuration to map each component to a
    word and a tag of type \(\mathbb{T}\).
    \[t \in \mathbb{T} = \mathcal{S} \times \mathbb{N} + ()\]
    \[M \in \mathcal{M} = \mathcal{K} \rightarrow \mathcal{W} \times \mathbb{T}\]
    When \(M(K) = w,t\) we say that \(t\) is the tag on \(K\), or \(K\) is tagged with \(t\).
    The tags on special registers are referred to by name: PC tag for the tag on \(\mathtt{PC}\), FP tag, etc.
    In our first two policies a tag is a pair representing the current stack and the depth of the active call within
    the stack, or else the default tag, (). We call setting a tag to () clearing it.

    \paragraph{Eager Depth Tracking}

      Initially, the PC tag is \(S_{init},0\), where \(S_{init}\) is the initial stack. The value at the top of
      a stack \(S\) holding its coroutine's initial PC is tagged \(S,0\).
      On a function call, the return address (at the current \(\mathtt{SP}\) and new \(\mathtt{FP}\)) is tagged with
      the PC tag, \(S,d\). The new PC tag becomes \(S,d+1\), and every address of the new stack frame is tagged
      \(S,d+1\). On return, the tag on the return address is restored to the PC tag, and the tags on the rest of the
      frame are cleared. Yields write the PC tag along with the PC, and restore the new ones from memory.
      When the PC tag is \(S,d\), only addresses tagged \(S,d\) or () may be read or written.

      This policy enforces local integrity, and therefore frame and stack integrity. It has a major weakness: it
      does not permit the dynamic manipulation of the current stack frame by altering the stack pointer. Doing
      so would leave addresses in the current frame tagged () and therefore accessible by higher stack frames.
      The most likely solution would be to introduce yet more pseudo-instructions, this time for tagging a range
      of memory.

      On the confidentiality side, this enforces local noninterference, and therefore everything that follows
      from it. Because the entire new stack frame must have its tags updated on a call and cleared on return, if we
      use these writes to zero out the values in memory as well, then this policy also enforces temporal safety.

    \paragraph{Lazy Depth Tracking}

      The PC tag starts at \(S_{init},0\) and the values holding coroutine PCs are initialized as above. The policy is
      identical except that stack frames are not tagged or cleared as a stack is pushed and popped. Rather, a write
      tags the written location with the current PC tag. A read is permitted if the location read is tagged
      () or exactly matches the PC tag.

      This policy enforces nothing. Clearly it doesn't enforce local integrity. Consider observable integrity:
      suppose that a function \(A\) calls \(B\), which writes to an unused portion of \(A\)'s frame and returns.
      Then \(A\) calls another function \(B'\), and \(B'\) reads from that portion and outputs its value.
      \(B\) has successfully interfered with the behavior of the trace by writing to \(A\)'s frame. Worse,
      suppose that before calling \(B'\), \(A\) deallocates some stack memory, moving its frame pointer
      below the tainted memory. Then a \(B'\) would not even need to violate confidentiality to output
      \(B\)'s interference.

      The same scenario, in which \(B'\) reads outside its accessible memory, violates all of our
      noninterference properties as well. These issues could occur whether or not \(B = B'\). Roessler
      and DeHon add to the tag a function identifier to catch the case of different functions, but not
      the case in which they're the same.

      This policy also  does not enforce safe initialization. If data is left behind by \(B\) in its own frame,
      \(B'\) can read it if they are at the same depth.

    \paragraph{Instance Tracking}

      This variation does enforce observable integrity and frame noninterference. Tags are now a pair of
      a stack id and a natural number we will call the {\it instance}. Each time a function call is made,
      we generate a fresh instance \(i\) and set the new PC tag to \(S,i\). The rest of the policy works
      the same as the Lazy Depth Tracking policy. Sadly, the requirement that we constantly generate fresh
      instances makes this variation costly, as the underlying tag mechanism relies on caching for good
      performance, and each new instance will cause compulsory misses.

      Performance issues aside, the case above no longer violates. If \(B\) has the PC tag \(S,i\)
      then \(B'\) will have the tag \(S,i+1\), so \(B\)'s external writes will not be accessible to
      \(B'\).

  \bibliographystyle{acm}
  \bibliography{references}

\end{document}

### 2020-04-28 LazyStackProperties.tex
Added version of integrity that corresponds better to the coq development

@@
 
+    \paragraph{Observable Integrity Closer to the Development}
+
+      In the interest of matching the coq development, here is an alternate formulation of observable integrity
+      that talks just about subroutines and is just phrased as operating on traces. Assume that there is just
+      one stack, \(S\). In this version, a system enjoys {\it observable stack integrity} if for all
+      \(\overline{T} = \dots T_1 T_2 \dots T_3 \overline{T}_4\) such that \(T_1 = M_1,\sigma_1,S\), 
+      \(T_2 = M_2,\sigma_2,S\), \(T_3 = M_3,\sigma_3,S\), and \(T_1\) and \(T_3\) bracket a call,
+      there is a safe state \(T_3' = M_3',\sigma_3,S\) where:
+      \[M_3'(K) =
+      \begin{cases}
+        M_1(K) & \text{if } \mathit{contour}(\sigma_1,S)(K) = \mathtt{A} \text{ and }
+                 \mathit{contour}(\sigma_2,S)(K) = \mathtt{IL} \\
+        M_3(K) & \text{else} \\
+      \end{cases}\]
+
+      And the observable trace \(\mathit{behav}(T_3' \overline{T}_4')\) is prefixed by
+      \(\mathit{behave}(T_3 \overline{T}_4)\).
+
+      This variant of the property ought to be combined with a coroutine equivalent, and it captures the
+      same intuition as the {\sc Safe} execution version, though it would need to be proven that the two are
+      equivalent.
+
   \subsection{Example}

### 2020-05-22 props.tex
Started writing up properties in tex to match the coq (props.tex)

\documentclass{article}

\usepackage{geometry}
\usepackage{amssymb,amsmath}

\begin{document}

  \section{Preliminaries}

    Values are words:
    \[w \in \mathcal{W}\]

    Components are words (addresses) and registers, including special registers:
    \[r \in \mathcal{R}\]
    \[k \in \mathcal{K} ::= \mathcal{W} | \mathcal{R} | \mathtt{PC} | \mathtt{SP}\]

    A machine state maps components to values:
    \[m \in \mathcal{M} ::= \mathcal{K} \rightarrow \mathcal{W}\]

    Our properties are defined in terms of observations, which may be silent or a word of data:
    \[o \in \mathit{Obs} ::= \tau | w \in \mathcal{W}\]

    Execution is by a step function \(\cdot \xrightarrow{\cdot} \cdot : \mathcal{M} \rightarrow
    \mathit{Obs} \times \mathcal{M}\). For every machine state \(m\) there is exactly one \(o\) and \(m'\)
    such that \(m \xrightarrow{o} m'\).

    Meanwhile a policy tracks a mix of static and dynamic enforcement. A policy consists of
    a set of policy states, \(p \in \mathcal{P}\), a policy initializer
    \(pinit : \mathcal{M} \rightharpoonup \mathcal{P}\), and a policy-parameterized step relation
    \(\cdot, \cdot \xrightarrow{\cdot}_P \cdot, \cdot : \mathcal{M} \times \mathcal{P} \rightharpoonup
    \mathit{Obs} \times \mathcal{M} \times \mathcal{P}\) such that if \(m,p \xrightarrow{o}_P m',p'\),
    then \(m \xrightarrow{o} m'\). The initializer is a partial function from machine states to policy
    states, reflecting that a policy may apply static constraints on the initial state. The parameterized
    step relation is likewise partial to allow for fail-stop behavior.

  \paragraph{Machine, Policy, and Observation Traces}

    A machine trace \(M\) is a sequence of machine states \(m_0 m_1 m_2 \dots\) such that each
    pair of adjacent states are related by \(\longrightarrow\) with some observation. An
    {\it mptrace} is similarly a sequence of machine state-policy state pairs,
    \((m_0,\mathit{pinit}(m_0))(m_1,p_1) \dots\) such that adjacent pairs of pairs are related
    by \(\longrightarrow_P\). Steps being deterministic, we can relate an initial state to its trace
    via \(\cdot \hookrightarrow \cdot : \mathcal{M} \rightarrow \mathit{trace}\ \mathcal{M}\):

    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} m_1 \hookrightarrow M}
           {m_0 \hookrightarrow m_0 M}\]

    Since the step function is total, if \(m \hookrightarrow M\), then \(M\) must be infinite.
    We will also wish to take a prefix of the trace from an initial state up to when some condition
    holds. Let \(f\) be a predicate on machine states. Then we define \(m \hookrightarrow M | f\)
    (read ``\(M\) is the prefix from \(m\) until \(f\)'') by:

    \begin{minipage}[b]{.45\textwidth}
      \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1 \hookrightarrow M | f }
             {m_0 \hookrightarrow m_0 M | f}\]
    \end{minipage}
    \begin{minipage}[b]{.45\textwidth}
      \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} f\ m_1}
             {m_0 \hookrightarrow m_0 m_1}\]
    \end{minipage}
    \vspace{\belowdisplayskip}

    We similarly define \(\hookrightarrow_P\), writing \(m_0,\mathit{pinit}(m_0) \longrightarrow \bot\) when
    there is no \(m_1,p_1\) such that \(m_0,\mathit{pinit}(m_0) \xrightarrow{o} m_1,p_1\) for any \(o\).

    \begin{minipage}[b]{.45\textwidth}
      \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} m_1,p_1 \hookrightarrow_P MP}
             {m_0,p_0 \hookrightarrow_P m_0,p_0 MP}\]
    \end{minipage}
    \begin{minipage}[b]{.45\textwidth}
      \[\frac{m_0,p_0 \longrightarrow \bot}
             {m_0,p_0 \hookrightarrow_P m_0,p_0}\]
    \end{minipage}
    \vspace{\belowdisplayskip}

    With a policy we may see \(m,p \hookrightarrow MP\) where \(MP\) is finite due to a policy
    failure. We will identify the set of infinite traces as \(\top\), so \(MP \in \top\) will
    indicate an infinite trace and \(MP \not \in \top\) a finite one.
    We define a similar notion of taking the prefix until a condition:

    \begin{minipage}[b]{.45\textwidth}
      \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1,p_1
                \hookrightarrow_P M | f }
             {m_0,p_0 \hookrightarrow_P m_0,p_0 M | f}\]
    \end{minipage}
    \begin{minipage}[b]{.45\textwidth}
      \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} f\ m_1}
             {m_0,p_0 \hookrightarrow_P (m_0,p_0) (m_1,p_1)}\]
    \end{minipage}
    \vspace{\belowdisplayskip}

    An observation trace \(O\) is the trace of observations from a machine or {\it mptrace},
    as defined by \(\mathit{ObsOf}\):

    \begin{minipage}[b]{.45\textwidth}
      \[\frac{M = mM' \hspace{.5cm} m \xrightarrow{o} m' \hspace{.5cm} \mathit{ObsOf}(M') = O}
             {\mathit{ObsOf}(M) = oO}\]
    \end{minipage}
    \begin{minipage}[b]{.45\textwidth}
      \[\frac{M = m}
             {\mathit{ObsOf}(M) = \tau}\]
    \end{minipage}
    \vspace{\belowdisplayskip}

    \begin{minipage}[b]{.45\textwidth}
      \[\frac{MP = (m,p)MP' \hspace{.5cm} m,p \xrightarrow{o} m',p' \hspace{.5cm} \mathit{ObsOf}(MP') = O}
             {\mathit{ObsOf}(MP) = oO}\]
    \end{minipage}
    \begin{minipage}[b]{.45\textwidth}
      \[\frac{MP = m,p}
             {\mathit{ObsOf}(M) = \tau}\]
    \end{minipage}
    \vspace{\belowdisplayskip}

    We abbreviate the observation trace of the execution trace from a state as:
    \[m \rightsquigarrow O \triangleq m \hookrightarrow M \land \mathit{ObsOf}(M) = O\]
    \[m,p \rightsquigarrow O \triangleq m,p \hookrightarrow_P MP \land \mathit{ObsOf}(MP) = O\]

    An observational prefix \(O_1 \lesssim O_2\) means that \(O_1\) is a prefix of \(O_2\) up to the deletion
    of \(\tau\) observations:

    \[O_1 \sim O_2 \triangleq
      \begin{cases}
        O_1' \sim O_2 & \text{if } O_1 = \tau O_1' \\
        O_1 \sim O_2' & \text{if } O_2 = \tau O_2' \\
        O_1' \sim O_2' & \text{if } O_1 = w O_1' \text{ and } O_2 = w O_2' \\
        \text{True} & \text{if for all } o \text{ in } O_1, o = \tau \\
      \end{cases}\]

    Observational equivalence is defined as a symmetric prefix:
    \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]

  \section{Observable Stack Safety}

    {\it Observable stack safety} is a property of a transition system \(\longrightarrow_p\),
    which is decomposed into three subproperties:
    {\it observable stack integrity}, {\it observable stack confidentiality}, and
    {\it observable safe initialization}.

  %  We unify these properties with a structure
  %  called a {\it contour} that maps components to levels of confidentiality and integrity.
  %  Specifically, we have confidentiality labels \(\mathit{labelC} ::= \{\mathtt{HC}, \mathtt{LC}\}\)
  %  and integrity labels \(\mathit{labelI} ::= \{\mathtt{HI}, \mathtt{LI}\}\), and a
  %  type \(\mathit{contour}\)

    \paragraph{Observable Safe Initialization}

      In the initial state of a program, nothing has yet been allocated on the stack, and
      there is no guarantee of the initial value of memory in the stack. So there is no
      need to protect a caller's stack data, but reading uninitialized data is unsafe.

      Let \(C : \mathcal{K} \rightarrow \{\mathtt{HC},\mathtt{LC}\}\) map all addresses
      to \(\mathtt{HC}\) (high confidentiality) and all registers to \(\mathtt{LC}\)
      (low confidentiality). In the style of non-interference, a pair of states \(m\)
      and \(m'\) are variants with regard to \(C\), \(m \approx_C m'\), if:
      \[\forall k . C(k) = \mathtt{LC} \rightarrow m(k) = m'(k)\]

      A system enjoys {\it observable safe initialization} if for any initial states \(m_0\)
      and \(m_0'\) such that \(m_0 \approx_C m_0'\):

      \begin{itemize}
        \item Let \(m_0,\mathit{pinit}(m_0) \rightsquigarrow O\) and \(m_0' \rightsquigarrow O'\)
        \item If \(m_0,\mathit{pinit}(m_0) \hookrightarrow_P MP \in \top\) then \(O \simeq O'\)
        \item Otherwise \(O \lesssim O'\)
      \end{itemize}

    \paragraph{Call States, Returns, and Contours}

      Our remaining properties concern the behavior of function calls and require a dynamic treatment
      of confidentiality and integrity. We generalize \(C\) above to a {\it contour} mapping each component
      to a label:

      \[\mathit{label} ::= \{\mathtt{HC},\mathtt{LC}\} \times \{\mathtt{HI},\mathtt{LI}\}\]

      \[C \in \mathcal{C} ::= \mathcal{K} \rightarrow \mathit{label}\]

      [TODO: be more explicit about what the call map does]
      A state \(m_c\) is a {\it call state} if [call map says it is], and a state \(m_p\) is
      considered to have returned to \(m_c\) if [etc.] From a call state with \(n\) arguments to
      its call, we can construct a contour that should apply inside the call:
      \[Cof(m)(k) =
      \begin{cases}
        (\mathtt{HC},\mathtt{HI}) & \text{if } k \in \mathcal{W}
                                    \text{ and } k \leq m(\mathtt{SP}) - n \\
        (\mathtt{HC},\mathtt{LI}) & \text{if } k \in \mathcal{W}
                                    \text{ and } k \leq m(\mathtt{SP}) \\
        (\mathtt{LC},\mathtt{LI}) & \text{else} \\
      \end{cases}\]

      Above the stack pointer, memory is high integrity in an extension of safe initialization;
      from the perspective of the callee, that memory is not initialized. Below the old stack
      pointer, memory is high integrity and high confidentiality. Registers and the callee's
      frame are low integrity and low confidentiality. These categories inform our integrity
      and confidentiality properties.

    \paragraph{Observable Integrity}

      Observable integrity is concerned with the writes a call makes to high integrity components,
      and the effects those writes have on the observable behavior of its caller
      {\it after its return}. Given a contour, a call state, and a returned state, we define
      a ``rollback'' function \(\mathit{roll}_I : \mathcal{C} \times \mathcal{M} \times \mathcal{M}
      \rightarrow \mathcal{M}\). The rollback creates a state that matches the call
      state on those components that are high integrity in \(C\), and matches the return state
      on low integrity components:

      \[\mathit{roll}_I(C,m_c,m_r)(k) =
      \begin{cases}
        m_c(k) & C(k) = (\mathtt{HI},\_) \\
        m_r(k) & C(k) = (\mathtt{LI},\_) \\
      \end{cases}\]

      Intuitively, the result of a rollback is an idealized state that reflects changes to accessible
      components, but ignores the callee's interference with the caller's data. Then the observable
      behavior of the resulting trace is the standard to which we expect any real trace to conform.

      A system enjoys observable integrity if for any initial state \(m_0\), call state \(m_c\),
      policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0)
      \hookrightarrow \dots (m_c,p_c) \dots\):

      \begin{itemize}
        \item If \((m_c,p_c) \hookrightarrow \dots (m_r,p_r) | \mathit{returned}\ m_c\ m_r\):
          \begin{itemize}
            \item Let \(m_r' = \mathit{roll}_I(C,m_c,m_r)\), \((m_r,p_r) \rightsquigarrow O\), and
              \(m_r' \rightsquigarrow O'\)
            \item If \(m_r,p_r \hookrightarrow MP \in \top\) then \(O \simeq O'\)
            \item Otherwise \(O \lesssim O'\)
          \end{itemize}
      \end{itemize}

    \paragraph{Eager Integrity}

      An eager enforcement mechanism implements a stronger property, which we term {\it eager integrity}.
      Eager integrity simplifies observable integrity by requiring that high integrity data of \(m_r\) be
      identical to that of \(m_c\). Formally, a systemctl enjoys eager integrity if for any initial state
      \(m_0\), call state \(m_c\), policy state \(p_c\), and contour \(C = Cof(m_c)\) such that
      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots (m_r,p_r) \dots\), where \(\mathit{returned}\ m_c\ m_r\):

      \[\forall k . C(k) = \mathtt{HI} \rightarrow m_c(k) = m_r(k)\]

      This implies lazy integrity, as the rollback function will produce \(m_r' = m_r\).

    \paragraph{Observable Confidentiality}

      Observable confidentiality is concerned with whether high confidentiality data influences
      observable behavior during and after a call. We define variants given the generalized notion
      of contours:
      \[m \approx_C m' \triangleq \forall k . C(k) = \mathtt{LC} \rightarrow m(k) = m'(k)\]

      Data that is high confidentiality within a callee due either to being in its caller's frame
      will become low confidentiality if the callee returns. We face a subtle distinction: 
      a callee treats the contents of its caller's frame as secrets, and they should not influence
      its behavior. Nor should the callee be able to influence behavior after its return based
      on the caller's secrets, for instance by copying them to registers that will later be output.
      But the caller itself may still adjust its behavior based on its own secrets. So our property
      relies on another rollback function to preserve changes made by the caller but remove other
      differences between variant states.

      In this case, a confidentiality rollback takes a contour and four states: two call states
      that will be variants, and two return states. It is not symmetric, regarding \(m_c\) and \(m_r\)
      as the ``real'' call and return, \(m_c'\) as a variant and \(m_r'\) as a return that follows it.

      \[\mathit{roll}_C(C,m_c,m_c',m_r,m_r')(k) =
        \begin{cases}
          m_c(k) & \text{if } C(k) = \mathtt{HC} \text{ and } m_c(k) = m_r(k) \text{ and } m_c'(k) = m_r'(k) \\
          m_r'(k) & \text{else} \\
        \end{cases}\]

      A system enjoys observable confidentiality if for any initial state \(m_0\), call states \(m_c\)
      and \(m_c'\), policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0)
      \hookrightarrow \dots (m_c,p_c) \dots\) and \(m_c \approx_C m_c'\):

      \begin{itemize}
        \item If \((m_c,p_c) \hookrightarrow MP (m_r,p_r) | \mathit{returned}\ m_c\ m_r\)
          \begin{itemize}
            \item Then there is some \(m_r'\) where
              \(m_c' \hookrightarrow M m_r' | \mathit{returned}\ m_c'\ m_r'\)
            \item Let \(m_r'' = \mathit{roll}_C(C,m_c,m_r,m_c',m_r')\),
              \((m_r,p_r) \rightsquigarrow O\), and \(m_r'' \rightsquigarrow O'\)
            \item If \(m_r,p_r \hookrightarrow_P MP \in \top\) then \(O \simeq \mathit{ObsOf}(M) +\!\!\!+ O'\)
            \item Otherwise \(O \lesssim \mathit{ObsOf}(M) +\!\!\!+ O'\)
          \end{itemize}
        \item Else, let \(m_c,p_c \rightsquigarrow O\) and \(m_c' \rightsquigarrow O'\)
          \begin{itemize}
            \item If \(m_c,p_c \hookrightarrow_P MP \in \top\) then \(O \simeq O'\)
            \item Otherwise \(O \lesssim O'\)
          \end{itemize}
      \end{itemize}

    \paragraph{Eager Confidentiality}

      Once again an eager policy enforces a stronger property that only concerns the behavior of
      the original trace and its variant within each call. A system enjoys {\it eager stack
      confidentiality} if, for any initial state \(m_0\), call states \(m_c\) and \(m_c'\),
      policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow
      \dots (m_c,p_c) \dots\) and \(m_c \approx_C m_c'\):

      \begin{itemize}
        \item If \((m_c,p_c) \hookrightarrow_P MP (m_r,p_r) | \mathit{returned}\ m_c\ m_r\)
          \begin{itemize}
            \item Then there is some \(m_r'\) where
              \(m_c' \hookrightarrow M m_r' | \mathit{returned}\ m_c'\ m_r'\)
            \item \(\mathit{ObsOf}(MP) \simeq \mathit{ObsOf}(M)\)
            \item And for all \(k\), if \(m_c(k) \not = m_r(k)\) or \(m_c'(k) \not = m_r'(k)\),
              \(m_r(k) = m_r'(k)\)
          \end{itemize}
        \item Else, let \(m_c,p_c \rightsquigarrow O\) and \(m_c' \rightsquigarrow O'\)
          \begin{itemize}
            \item If \(m_c,p_c \hookrightarrow_P MP \in \top\) then \(O \simeq O'\)
            \item Otherwise \(O \lesssim O'\)
          \end{itemize}
      \end{itemize}

      Eager confidentiality implies observable confidentiality, because if \(m_r\) and \(m_r'\) agree
      on all \(k\) that changed during the call, then an \(m_r'' = \mathit{roll}_C(C,m_c,m_r,m_c',m_r')\)
      will be identical to \(m_r\), and future observations will be the same up to a potential policy
      stop.

\end{document}

### 2020-05-24 props.tex
Quick update of eager integrity to match the others
Made properties explicit about calls, though now returns don't line up right. (In .tex)

@@
 
+    \paragraph{Calling Conventions}
+
+      A subroutine call with \(n\) arguments follows the following convention:
+
+      \begin{itemize}
+        \item The stack pointer is increased by \(n+1\)
+        \item A {\tt Jal} occurs
+        \item The return address is stored at \(\mathtt{SP} - (n+1)\)
+      \end{itemize}
+
+      The final state of a call sequence state is a {\it call state}. We relate a call sequence \(M\) with
+      \(n\) arguments to its final state \(m\) with \(\mathit{isCall}\ M\ m\ n \subset \mathit{trace}\ \mathcal{M}
+      \times \mathcal{M} \times \mathbb{N}\).
+
+      A state is considered to have returned after a call if the stack pointer is the same as before the call
+      and the program counter is incremented. So:
+      \[\mathit{returned}\ m\ m' \triangleq \exists M,n . \mathit{isCall\ M\ m\ n} \land
+        m'(\mathtt{SP}) = m(\mathtt{SP}) - (n+1) \land m(\mathtt{PC}) = m'(\mathtt{PC}) + 1\]
+
   \section{Observable Stack Safety}
@@
     {\it Observable stack safety} is a property of a transition system \(\longrightarrow_p\),
-    which is decomposed into three subproperties:
-    {\it observable stack integrity}, {\it observable stack confidentiality}, and
-    {\it observable safe initialization}.
+    which is decomposed into three subproperties: {\it observable stack integrity},
+    {\it observable stack confidentiality}, and {\it observable safe initialization}.
 
@@
 
-    \paragraph{Call States, Returns, and Contours}
+    \paragraph{Contours}
 
@@
 
-      [TODO: be more explicit about what the call map does]
-      A state \(m_c\) is a {\it call state} if [call map says it is], and a state \(m_p\) is
-      considered to have returned to \(m_c\) if [etc.] From a call state with \(n\) arguments to
-      its call, we can construct a contour that should apply inside the call:
+      When \(m\) is a call state such that \(\mathit{isCall}\ M\ m\ n\) for some \(M\) and \(n\),
+      we can construct a contour based on the size of \(m\)'s stack frame, as follows:
       \[Cof(m)(k) =
@@
       Above the stack pointer, memory is high integrity in an extension of safe initialization;
-      from the perspective of the callee, that memory is not initialized. Below the old stack
-      pointer, memory is high integrity and high confidentiality. Registers and the callee's
+      from the perspective of the callee, that memory is not initialized. At and below the old stack
+      pointer, memory is high integrity and high confidentiality -- this includes the return address,
+      protecting control flow, as well as the caller's data. Registers and the callee's
       frame are low integrity and low confidentiality. These categories inform our integrity
@@
 
-      A system enjoys observable integrity if for any initial state \(m_0\), call state \(m_c\),
-      policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0)
-      \hookrightarrow \dots (m_c,p_c) \dots\):
+      A system enjoys observable integrity if for any initial state \(m_0\) and call state \(m_c\)
+      such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots MP_c (m_c,p_c) \dots\)
+      and \(\mathit{isCall}\ MP_c\ m_c\ n\):
 
@@
           \begin{itemize}
+            \item Let \(C = \mathit{Cof(m_c,n)}\)
             \item Let \(m_r' = \mathit{roll}_I(C,m_c,m_r)\), \((m_r,p_r) \rightsquigarrow O\), and
@@
       Eager integrity simplifies observable integrity by requiring that high integrity data of \(m_r\) be
-      identical to that of \(m_c\). Formally, a systemctl enjoys eager integrity if for any initial state
-      \(m_0\), call state \(m_c\), policy state \(p_c\), and contour \(C = Cof(m_c)\) such that
-      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots (m_r,p_r) \dots\), where \(\mathit{returned}\ m_c\ m_r\):
+      identical to that of \(m_c\). Formally, a system enjoys eager integrity if for any initial state
+      \(m_0\) and call state \(m_c\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots MP_c (m_c,p_c)
+      \dots (m_r,p_r) | \mathit{returned}\ m_c\ m_r\) and \(\mathit{isCall}\ MP_c (m_c,p_c)\):
 
@@
 
-      A system enjoys observable confidentiality if for any initial state \(m_0\), call states \(m_c\)
-      and \(m_c'\), policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0)
-      \hookrightarrow \dots (m_c,p_c) \dots\) and \(m_c \approx_C m_c'\):
+      A system enjoys observable confidentiality if for any initial state \(m_0\) and call state \(m_c\)
+      such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots MP_c (m_c,p_c) \dots\) and
+      \(\mathit{isCall}\ MP_c\ m_c\ n\):
+      and \(m_c'\),
 
       \begin{itemize}
+        \item Let \(C = \mathit{Cof}(m_c,n)\)
+        \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
         \item If \((m_c,p_c) \hookrightarrow MP (m_r,p_r) | \mathit{returned}\ m_c\ m_r\)
@@
       the original trace and its variant within each call. A system enjoys {\it eager stack
-      confidentiality} if, for any initial state \(m_0\), call states \(m_c\) and \(m_c'\),
-      policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow
-      \dots (m_c,p_c) \dots\) and \(m_c \approx_C m_c'\):
+      confidentiality} if, for any initial state \(m_0\) and call state \(m_c\) such that
+      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots MP_c (m_c,p_c) \dots\) and
+      \(\mathit{isCall}\ MP_c\ m_c\ n\):
 
       \begin{itemize}
+        \item Let \(C = \mathit{Cof}(m_c)\)
+        \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
         \item If \((m_c,p_c) \hookrightarrow_P MP (m_r,p_r) | \mathit{returned}\ m_c\ m_r\)

### 2020-05-24 props.tex
props.tex contains a slightly more honest treatment of calls and returns, and I think it works properly now.

--- 2020-05-22 props.tex
@@
 
-    Values are words:
-    \[w \in \mathcal{W}\]
+    Values and addresses are words:
+    \[a,w \in \mathcal{W}\]
 
@@
 
-    Execution is by a step function \(\cdot \xrightarrow{\cdot} \cdot : \mathcal{M} \rightarrow
-    \mathit{Obs} \times \mathcal{M}\). For every machine state \(m\) there is exactly one \(o\) and \(m'\)
-    such that \(m \xrightarrow{o} m'\).
+    Execution is by a step function \(m \xrightarrow{o} m' : m \in \mathcal{M} \rightarrow
+    o \in \mathit{Obs} \times m' \in \mathcal{M}\).
 
-    Meanwhile a policy tracks a mix of static and dynamic enforcement. A policy consists of
+    A policy tracks a mix of static and dynamic enforcement. A policy consists of
     a set of policy states, \(p \in \mathcal{P}\), a policy initializer
-    \(pinit : \mathcal{M} \rightharpoonup \mathcal{P}\), and a policy-parameterized step relation
-    \(\cdot, \cdot \xrightarrow{\cdot}_P \cdot, \cdot : \mathcal{M} \times \mathcal{P} \rightharpoonup
-    \mathit{Obs} \times \mathcal{M} \times \mathcal{P}\) such that if \(m,p \xrightarrow{o}_P m',p'\),
-    then \(m \xrightarrow{o} m'\). The initializer is a partial function from machine states to policy
+    \(pinit : \mathcal{M} \rightharpoonup \mathcal{P}\), and a policy-parameterized step function:
+    \[m, p \xrightarrow{o}_P m', p' : m \in \mathcal{M} \times p \in \mathcal{P} \rightharpoonup
+    o \in \mathit{Obs} \times m' \in \mathcal{M} \times p' \in \mathcal{P}\]
+
+    Axiomatically, if \(m,p \xrightarrow{o}_P m',p'\), then \(m \xrightarrow{o} m'\).
+    The initializer is a partial function from machine states to policy
     states, reflecting that a policy may apply static constraints on the initial state. The parameterized
@@
 
+  \paragraph{Calling Conventions and Contours}
+
+      A subroutine call with \(n\) arguments follows the following convention:
+
+      \begin{itemize}
+        \item The stack pointer is increased by \(n+1\)
+        \item A {\tt Jal} occurs
+        \item The return address is stored at \(\mathtt{SP} - 1\)
+      \end{itemize}
+
+      The final state of a call sequence is a {\it call state}. If \(M\) is a call sequence with
+      we map it to its call state, number of arguments, and the PC and SP values of its eventual
+      return by the function \(\mathit{call} : \mathit{trace}\ \mathcal{M} \rightharpoonup \mathcal{M}
+      \times \mathbb{N} \times \mathcal{W} \times \mathcal{W}\). We can also track a call sequence in the
+      policy state, defining \(\mathit{call}_P : \mathcal{M} \times \mathcal{P} \rightharpoonup \mathbb{N}
+      \times \mathcal{W} \times \mathcal{W}\) such that:
+
+      \[\frac{m_0,\mathit{pinit}(m_0) \hookrightarrow_P \dots m,p \dots \hspace{.5cm}
+              m_0 \hookrightarrow \dots M \dots \hspace{.5cm} \mathit{call}(M) = m, n, a_{pc}, a_{sp}}
+             {\mathit{call}_P(m,p) = n, a_{pc}, a_{sp}}\]
+
+      A state is considered to have returned after a call if the stack pointer is the same as before the call
+      (\(a_{sp}\)) and the program counter is the next address after (\(a_{pc}\)):
+      \[\mathit{returned}\ m\ a_{pc}\ a_{sp} \triangleq m'(\mathtt{SP}) = a_{sp} \land m(\mathtt{PC}) = a_{pc}\]
+
   \section{Observable Stack Safety}
@@
 
-      [TODO: be more explicit about what the call map does]
-      A state \(m_c\) is a {\it call state} if [call map says it is], and a state \(m_p\) is
-      considered to have returned to \(m_c\) if [etc.] From a call state with \(n\) arguments to
-      its call, we can construct a contour that should apply inside the call:
-      \[Cof(m)(k) =
+      When \(m,p\) is a call state \(\mathit{call}_P(m,p) = n,\_,\_\) for some \(n\),
+      we can construct a contour based on the size of \(m\)'s stack frame, as follows:
+      \[Cof(m,n)(k) =
       \begin{cases}
@@
         (\mathtt{HC},\mathtt{LI}) & \text{if } k \in \mathcal{W}
-                                    \text{ and } k \leq m(\mathtt{SP}) \\
+                                    \text{ and } k \geq m(\mathtt{SP}) \\
+        (\mathtt{LC},\mathtt{HI}) & \text{if } k \in \mathcal{W}
+                                    \text{ and } k = m(\mathtt{SP}) - 1 \\
         (\mathtt{LC},\mathtt{LI}) & \text{else} \\
@@
       from the perspective of the callee, that memory is not initialized. Below the old stack
-      pointer, memory is high integrity and high confidentiality. Registers and the callee's
-      frame are low integrity and low confidentiality. These categories inform our integrity
-      and confidentiality properties.
+      pointer, memory is high integrity and high confidentiality. The top of the callee's frame
+      is its return address, which it must be able to read in order to return but which is
+      high integrity [does this make sense?] Registers and the callee's frame are low integrity
+      and low confidentiality. These categories inform our integrity and confidentiality properties.
 
@@
       A system enjoys observable integrity if for any initial state \(m_0\), call state \(m_c\),
-      policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0)
-      \hookrightarrow \dots (m_c,p_c) \dots\):
+      and policy state \(p_c\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots\)
+      and \(\mathit{call}_P(m_c,p_c) = n,a_{pc},a_{sp}\):
 
       \begin{itemize}
-        \item If \((m_c,p_c) \hookrightarrow \dots (m_r,p_r) | \mathit{returned}\ m_c\ m_r\):
+        \item If \((m_c,p_c) \hookrightarrow \dots (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\):
           \begin{itemize}
+            \item Let \(C = \mathit{Cof}(m_c,n)\)
             \item Let \(m_r' = \mathit{roll}_I(C,m_c,m_r)\), \((m_r,p_r) \rightsquigarrow O\), and
@@
       Eager integrity simplifies observable integrity by requiring that high integrity data of \(m_r\) be
-      identical to that of \(m_c\). Formally, a systemctl enjoys eager integrity if for any initial state
-      \(m_0\), call state \(m_c\), policy state \(p_c\), and contour \(C = Cof(m_c)\) such that
-      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots (m_r,p_r) \dots\), where \(\mathit{returned}\ m_c\ m_r\):
+      identical to that of \(m_c\). Formally, a system enjoys eager integrity if for any initial state
+      \(m_0\), call state \(m_c\), and policy state \(p_c\) such that \(\mathit{call}(m_c,p_c) =
+      n, a_{pc}, a_{sp}\), and some \(m_r\) and \(p_r\) such that
+      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\):
 
@@
 
-      A system enjoys observable confidentiality if for any initial state \(m_0\), call states \(m_c\)
-      and \(m_c'\), policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0)
-      \hookrightarrow \dots (m_c,p_c) \dots\) and \(m_c \approx_C m_c'\):
+      A system enjoys observable confidentiality if for any initial state \(m_0\), call state \(m_c\),
+      and policy state \(p_c\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots\) and
+      \(\mathit{call}(m_c,p_c) = n, a_{pc}, a_{sp}\):
 
       \begin{itemize}
-        \item If \((m_c,p_c) \hookrightarrow MP (m_r,p_r) | \mathit{returned}\ m_c\ m_r\)
+        \item Let \(C = \mathit{Cof}(m_c,n)\)
+        \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
+        \item If \((m_c,p_c) \hookrightarrow MP (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\)
           \begin{itemize}
             \item Then there is some \(m_r'\) where
-              \(m_c' \hookrightarrow M m_r' | \mathit{returned}\ m_c'\ m_r'\)
+              \(m_c' \hookrightarrow M m_r' | \mathit{returned}\ m_r'\ a_{pc}\ a_{sp}\)
             \item Let \(m_r'' = \mathit{roll}_C(C,m_c,m_r,m_c',m_r')\),
@@
       the original trace and its variant within each call. A system enjoys {\it eager stack
-      confidentiality} if, for any initial state \(m_0\), call states \(m_c\) and \(m_c'\),
-      policy state \(p_c\), and contour \(C = Cof(m_c)\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow
-      \dots (m_c,p_c) \dots\) and \(m_c \approx_C m_c'\):
+      confidentiality} if, for any initial state \(m_0\) and call state \(m_c\) such that
+      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots\) and
+      \(\mathit{call}(m_c,p_c) = n, a_{pc}, a_{sp}\):
 
       \begin{itemize}
-        \item If \((m_c,p_c) \hookrightarrow_P MP (m_r,p_r) | \mathit{returned}\ m_c\ m_r\)
+        \item Let \(C = \mathit{Cof}(m_c,n)\)
+        \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
+        \item If \((m_c,p_c) \hookrightarrow_P MP (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\)
           \begin{itemize}

### 2020-05-25 props.tex
Cleaned up notation on props.tex and put it in a two-column form

@@
-\documentclass{article}
+\documentclass[conference]{IEEEtran}
 
@@
 \usepackage{amssymb,amsmath}
+
+\newcommand{\MP}{\mathit{MP}}
 
@@
     \(pinit : \mathcal{M} \rightharpoonup \mathcal{P}\), and a policy-parameterized step function:
-    \[m, p \xrightarrow{o}_P m', p' : m \in \mathcal{M} \times p \in \mathcal{P} \rightharpoonup
-    o \in \mathit{Obs} \times m' \in \mathcal{M} \times p' \in \mathcal{P}\]
+    \[\begin{split}
+      m, p \xrightarrow{o}_P m', p' : m \in \mathcal{M} \times p \in \mathcal{P} & \rightharpoonup \\
+      o \in \mathit{Obs} \times m' \in \mathcal{M} \times p' \in \mathcal{P} &
+    \end{split}\]
 
@@
     by \(\longrightarrow_P\). Steps being deterministic, we can relate an initial state to its trace
-    via \(\cdot \hookrightarrow \cdot : \mathcal{M} \rightarrow \mathit{trace}\ \mathcal{M}\):
+    via \(m \hookrightarrow M : m \in \mathcal{M} \rightarrow M \in \mathit{trace}\ \mathcal{M}\):
 
@@
     holds. Let \(f\) be a predicate on machine states. Then we define \(m \hookrightarrow M | f\)
-    (read ``\(M\) is the prefix from \(m\) until \(f\)'') by:
-
-    \begin{minipage}[b]{.45\textwidth}
+    (read ``\(M\) is the prefix from \(m\) until \(f\)'') where \(f\) may bind states in \(M\):
     \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1 \hookrightarrow M | f }
            {m_0 \hookrightarrow m_0 M | f}\]
-    \end{minipage}
-    \begin{minipage}[b]{.45\textwidth}
     \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} f\ m_1}
            {m_0 \hookrightarrow m_0 m_1}\]
-    \end{minipage}
-    \vspace{\belowdisplayskip}
 
@@
     there is no \(m_1,p_1\) such that \(m_0,\mathit{pinit}(m_0) \xrightarrow{o} m_1,p_1\) for any \(o\).
-
-    \begin{minipage}[b]{.45\textwidth}
-      \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} m_1,p_1 \hookrightarrow_P MP}
-             {m_0,p_0 \hookrightarrow_P m_0,p_0 MP}\]
-    \end{minipage}
-    \begin{minipage}[b]{.45\textwidth}
+    \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} m_1,p_1 \hookrightarrow_P \MP}
+           {m_0,p_0 \hookrightarrow_P m_0,p_0 \MP}\]
     \[\frac{m_0,p_0 \longrightarrow \bot}
            {m_0,p_0 \hookrightarrow_P m_0,p_0}\]
-    \end{minipage}
-    \vspace{\belowdisplayskip}
 
-    With a policy we may see \(m,p \hookrightarrow MP\) where \(MP\) is finite due to a policy
-    failure. We will identify the set of infinite traces as \(\top\), so \(MP \in \top\) will
-    indicate an infinite trace and \(MP \not \in \top\) a finite one.
+    With a policy we may see \(m,p \hookrightarrow \MP\) where \(\MP\) is finite due to a policy
+    failure. We will identify the set of infinite traces as \(\top\), so \(\MP \in \top\) will
+    indicate an infinite trace and \(\MP \not \in \top\) a finite one.
     We define a similar notion of taking the prefix until a condition:
 
-    \begin{minipage}[b]{.45\textwidth}
     \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1,p_1
@@
            {m_0,p_0 \hookrightarrow_P m_0,p_0 M | f}\]
-    \end{minipage}
-    \begin{minipage}[b]{.45\textwidth}
     \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} f\ m_1}
            {m_0,p_0 \hookrightarrow_P (m_0,p_0) (m_1,p_1)}\]
-    \end{minipage}
-    \vspace{\belowdisplayskip}
 
     An observation trace \(O\) is the trace of observations from a machine or {\it mptrace},
-    as defined by \(\mathit{ObsOf}\):
-
-    \begin{minipage}[b]{.45\textwidth}
-      \[\frac{M = mM' \hspace{.5cm} m \xrightarrow{o} m' \hspace{.5cm} \mathit{ObsOf}(M') = O}
-             {\mathit{ObsOf}(M) = oO}\]
-    \end{minipage}
-    \begin{minipage}[b]{.45\textwidth}
+    as defined by \(\mathit{Obs}\):
+    \[\frac{M = mM' \hspace{.5cm} m \xrightarrow{o} m'}
+           {\mathit{Obs}(M) = o\mathit{Obs}(M')}\]
     \[\frac{M = m}
-             {\mathit{ObsOf}(M) = \tau}\]
-    \end{minipage}
-    \vspace{\belowdisplayskip}
-
-    \begin{minipage}[b]{.45\textwidth}
-      \[\frac{MP = (m,p)MP' \hspace{.5cm} m,p \xrightarrow{o} m',p' \hspace{.5cm} \mathit{ObsOf}(MP') = O}
-             {\mathit{ObsOf}(MP) = oO}\]
-    \end{minipage}
-    \begin{minipage}[b]{.45\textwidth}
-      \[\frac{MP = m,p}
-             {\mathit{ObsOf}(M) = \tau}\]
-    \end{minipage}
-    \vspace{\belowdisplayskip}
-
-    We abbreviate the observation trace of the execution trace from a state as:
-    \[m \rightsquigarrow O \triangleq m \hookrightarrow M \land \mathit{ObsOf}(M) = O\]
-    \[m,p \rightsquigarrow O \triangleq m,p \hookrightarrow_P MP \land \mathit{ObsOf}(MP) = O\]
+           {\mathit{Obs}(M) = \tau}\]
+    \[\frac{\MP = (m,p)\MP' \hspace{.5cm} m,p \xrightarrow{o} m',p'}
+           {\mathit{Obs}(\MP) = o \mathit{Obs}(\MP')}\]
+    \[\frac{\MP = m,p}
+           {\mathit{Obs}(M) = \tau}\]
 
@@
     of \(\tau\) observations:
+    \[\begin{split}
+      O_1 & \sim O_2 \triangleq \forall o . o \in O_1 \Rightarrow o = \tau \lor \\
+      & \exists O_1' . O_1 = \tau O_1' \land O_1' \sim O_2 \lor \\
+      & \exists O_2' .  O_2 = \tau O_2' \land O_1 \sim O_2' \lor \\
+      & \exists O_1' O_2' . O_1 = w O_1' \land O_2 = w O_2' \land O_1' \sim O_2' \\
+    \end{split}\]
 
-    \[O_1 \sim O_2 \triangleq
-      \begin{cases}
-        O_1' \sim O_2 & \text{if } O_1 = \tau O_1' \\
-        O_1 \sim O_2' & \text{if } O_2 = \tau O_2' \\
-        O_1' \sim O_2' & \text{if } O_1 = w O_1' \text{ and } O_2 = w O_2' \\
-        \text{True} & \text{if for all } o \text{ in } O_1, o = \tau \\
-      \end{cases}\]
-
-    Observational equivalence is defined as a symmetric prefix:
+    Observational equivalence is defined as traces prefixing each other:
     \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
@@
 
-      \[\frac{m_0,\mathit{pinit}(m_0) \hookrightarrow_P \dots m,p \dots \hspace{.5cm}
-              m_0 \hookrightarrow \dots M \dots \hspace{.5cm} \mathit{call}(M) = m, n, a_{pc}, a_{sp}}
+      \[\frac{\begin{split}
+        m_0,\mathit{pinit}(m_0) & \hookrightarrow_P \dots m,p \dots \\ % \hspace{.5cm}
+        m_0 & \hookrightarrow \dots M \dots \\
+        \mathit{call}(M) & = m, n, a_{pc}, a_{sp}\end{split}}
              {\mathit{call}_P(m,p) = n, a_{pc}, a_{sp}}\]
@@
       (\(a_{sp}\)) and the program counter is the next address after (\(a_{pc}\)):
-      \[\mathit{returned}\ m\ a_{pc}\ a_{sp} \triangleq m'(\mathtt{SP}) = a_{sp} \land m(\mathtt{PC}) = a_{pc}\]
+      \[\mathit{ret}\ m\ a_{pc}\ a_{sp} \triangleq m(\mathtt{SP}) = a_{sp} \land m(\mathtt{PC}) = a_{pc}\]
 
@@
       \begin{itemize}
-        \item Let \(m_0,\mathit{pinit}(m_0) \rightsquigarrow O\) and \(m_0' \rightsquigarrow O'\)
-        \item If \(m_0,\mathit{pinit}(m_0) \hookrightarrow_P MP \in \top\) then \(O \simeq O'\)
-        \item Otherwise \(O \lesssim O'\)
+        \item Let \(m_0,\mathit{pinit}(m_0) \hookrightarrow \MP\) and \(m_0' \hookrightarrow M'\)
+        \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
+        \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
       \end{itemize}
@@
       we can construct a contour based on the size of \(m\)'s stack frame, as follows:
-      \[Cof(m,n)(k) =
+      \[\mathit{Cof}(m,n)(k) =
       \begin{cases}
-        (\mathtt{HC},\mathtt{HI}) & \text{if } k \in \mathcal{W}
-                                    \text{ and } k \leq m(\mathtt{SP}) - n \\
-        (\mathtt{HC},\mathtt{LI}) & \text{if } k \in \mathcal{W}
-                                    \text{ and } k \geq m(\mathtt{SP}) \\
-        (\mathtt{LC},\mathtt{HI}) & \text{if } k \in \mathcal{W}
-                                    \text{ and } k = m(\mathtt{SP}) - 1 \\
+        (\mathtt{HC},\mathtt{HI}) & \text{if } k \leq m(\mathtt{SP}) - n \\
+        (\mathtt{HC},\mathtt{LI}) & \text{if } k \geq m(\mathtt{SP}) \\
+        (\mathtt{LC},\mathtt{HI}) & \text{if } k = m(\mathtt{SP}) - 1 \\
         (\mathtt{LC},\mathtt{LI}) & \text{else} \\
@@
 
-      A system enjoys observable integrity if for any initial state \(m_0\), call state \(m_c\),
-      and policy state \(p_c\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots\)
-      and \(\mathit{call}_P(m_c,p_c) = n,a_{pc},a_{sp}\):
+      A system enjoys observable integrity as follows:
+
+      For any initial state \(m_0\), call state \(m_c\), and policy state \(p_c\) such that
+      \((m_0,\mathit{pinit}(m_0)) \hookrightarrow ... (m_c,p_c) ...\)
+      and \(\mathit{call}_P(m_c,p_c) = n,a_{pc},a_{sp}\), if 
+      \((m_c,p_c) \hookrightarrow ... (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\):
 
           \begin{itemize}
-        \item If \((m_c,p_c) \hookrightarrow \dots (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\):
-          \begin{itemize}
             \item Let \(C = \mathit{Cof}(m_c,n)\)
-            \item Let \(m_r' = \mathit{roll}_I(C,m_c,m_r)\), \((m_r,p_r) \rightsquigarrow O\), and
-              \(m_r' \rightsquigarrow O'\)
-            \item If \(m_r,p_r \hookrightarrow MP \in \top\) then \(O \simeq O'\)
-            \item Otherwise \(O \lesssim O'\)
-          \end{itemize}
+            \item Let \(m_r' = \mathit{roll}_I(C,m_c,m_r)\), \((m_r,p_r) \hookrightarrow \MP\), and
+              \(m_r' \hookrightarrow M'\)
+            \item If \(m_r,p_r \hookrightarrow \MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
+            \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
           \end{itemize}
@@
       n, a_{pc}, a_{sp}\), and some \(m_r\) and \(p_r\) such that
-      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\):
+      \(m_0,\mathit{pinit}(m_0) \hookrightarrow ... (m_c,p_c) ... (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\):
 
@@
       as the ``real'' call and return, \(m_c'\) as a variant and \(m_r'\) as a return that follows it.
+      \[\begin{split}
+        & \mathit{roll}_C(C,m_c,m_c',m_r,m_r')(k) = \\
+        & \begin{cases}
+          m_c(k) & \text{if } C(k) = \mathtt{HC} \text{ and } \\
+                  & m_c(k) = m_r(k) \text{ and } m_c'(k) = m_r'(k) \\
+          m_r'(k) & \text{else} \\
+        \end{cases}
+      \end{split}\]
 
-      \[\mathit{roll}_C(C,m_c,m_c',m_r,m_r')(k) =
-        \begin{cases}
-          m_c(k) & \text{if } C(k) = \mathtt{HC} \text{ and } m_c(k) = m_r(k) \text{ and } m_c'(k) = m_r'(k) \\
-          m_r'(k) & \text{else} \\
-        \end{cases}\]
+      A system enjoys observable confidentiality as follows:
 
-      A system enjoys observable confidentiality if for any initial state \(m_0\), call state \(m_c\),
-      and policy state \(p_c\) such that \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots\) and
+      For any initial state \(m_0\), call state \(m_c\), and policy state \(p_c\) such that
+      \(m_0,\mathit{pinit}(m_0) \hookrightarrow ... (m_c,p_c) ...\) and
       \(\mathit{call}(m_c,p_c) = n, a_{pc}, a_{sp}\):
@@
         \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
-        \item If \((m_c,p_c) \hookrightarrow MP (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\)
+        \item If \((m_c,p_c) \hookrightarrow \MP_1 (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\)
           \begin{itemize}
             \item Then there is some \(m_r'\) where
-              \(m_c' \hookrightarrow M m_r' | \mathit{returned}\ m_r'\ a_{pc}\ a_{sp}\)
+              \(m_c' \hookrightarrow M_1' m_r' | \mathit{ret}\ m_r'\ a_{pc}\ a_{sp}\)
             \item Let \(m_r'' = \mathit{roll}_C(C,m_c,m_r,m_c',m_r')\),
-              \((m_r,p_r) \rightsquigarrow O\), and \(m_r'' \rightsquigarrow O'\)
-            \item If \(m_r,p_r \hookrightarrow_P MP \in \top\) then \(O \simeq \mathit{ObsOf}(M) +\!\!\!+ O'\)
-            \item Otherwise \(O \lesssim \mathit{ObsOf}(M) +\!\!\!+ O'\)
+              \((m_r,p_r) \hookrightarrow \MP_2\), and \(m_r'' \hookrightarrow M_2'\)
+            \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP_1) \cdot \mathit{Obs}(\MP_2) \simeq
+              \mathit{Obs}(M_1') \cdot \mathit{Obs}(M_2')\)
+            \item Otherwise \(\mathit{Obs}(\MP_1) \cdot \mathit{Obs}(\MP_2) \lesssim
+              \mathit{Obs}(M_1') \cdot \mathit{Obs}(M_2')\)
           \end{itemize}
-        \item Else, let \(m_c,p_c \rightsquigarrow O\) and \(m_c' \rightsquigarrow O'\)
+        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\):
           \begin{itemize}
-            \item If \(m_c,p_c \hookrightarrow_P MP \in \top\) then \(O \simeq O'\)
-            \item Otherwise \(O \lesssim O'\)
+            \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
+            \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
           \end{itemize}
@@
       confidentiality} if, for any initial state \(m_0\) and call state \(m_c\) such that
-      \(m_0,\mathit{pinit}(m_0) \hookrightarrow \dots (m_c,p_c) \dots\) and
+      \(m_0,\mathit{pinit}(m_0) \hookrightarrow ... (m_c,p_c) ...\) and
       \(\mathit{call}(m_c,p_c) = n, a_{pc}, a_{sp}\):
@@
         \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
-        \item If \((m_c,p_c) \hookrightarrow_P MP (m_r,p_r) | \mathit{returned}\ m_r\ a_{pc}\ a_{sp}\)
+        \item If \((m_c,p_c) \hookrightarrow_P \MP (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\)
           \begin{itemize}
             \item Then there is some \(m_r'\) where
-              \(m_c' \hookrightarrow M m_r' | \mathit{returned}\ m_c'\ m_r'\)
-            \item \(\mathit{ObsOf}(MP) \simeq \mathit{ObsOf}(M)\)
+              \(m_c' \hookrightarrow M m_r' | \mathit{ret}\ m_c'\ m_r'\)
+            \item \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M)\)
             \item And for all \(k\), if \(m_c(k) \not = m_r(k)\) or \(m_c'(k) \not = m_r'(k)\),
@@
           \end{itemize}
-        \item Else, let \(m_c,p_c \rightsquigarrow O\) and \(m_c' \rightsquigarrow O'\)
+        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\)
           \begin{itemize}
-            \item If \(m_c,p_c \hookrightarrow_P MP \in \top\) then \(O \simeq O'\)
-            \item Otherwise \(O \lesssim O'\)
+            \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
+            \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
           \end{itemize}

### 2020-05-26 props.tex
Slightly cleaner, factored out policy failure case

@@
     \[\begin{split}
-      O_1 & \sim O_2 \triangleq \forall o . o \in O_1 \Rightarrow o = \tau \lor \\
-      & \exists O_1' . O_1 = \tau O_1' \land O_1' \sim O_2 \lor \\
-      & \exists O_2' .  O_2 = \tau O_2' \land O_1 \sim O_2' \lor \\
-      & \exists O_1' O_2' . O_1 = w O_1' \land O_2 = w O_2' \land O_1' \sim O_2' \\
+      O_1 & \lesssim O_2 \triangleq \forall o . o \in O_1 \Rightarrow o = \tau \lor \\
+      & \exists O_1' . O_1 = \tau O_1' \land O_1' \lesssim O_2 \lor \\
+      & \exists O_2' .  O_2 = \tau O_2' \land O_1 \lesssim O_2' \lor \\
+      & \exists O_1' O_2' . O_1 = w O_1' \land O_2 = w O_2' \land O_1' \lesssim O_2' \\
     \end{split}\]
 
-    Observational equivalence is defined as traces prefixing each other:
+    Equivalence of observation traces is defined as traces prefixing each other:
     \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
+
+    We define policy-sensitive observational equivalence, which relates machine traces and
+    {\it mptraces} whose observable behavior match up to a possible halt due to policy failure.
+    It relates lists of parameters as long as only the last is infinite.
+
+    \[\frac{\MP \not \in \top \hspace{.5cm} \mathit{Obs}(\MP) \lesssim \mathit{Obs}(M)}
+           {\MP \sim_P M}\]
+
+    \[\frac{\MP \in \top \hspace{.5cm} \mathit{Obs}(\MP) \simeq \mathit{Obs}(M)}
+           {\MP \sim_P M}\]
+
+    \[\frac{\begin{split}
+              \MP \not \in \top & \hspace{.5cm} M \not \in \top \\
+              \mathit{Obs}(\MP) \simeq \mathit{Obs}(M) & \hspace{.5cm} \overline{\MP} \sim_P \overline{M} \\
+            \end{split}}
+           {\MP \overline{\MP} \sim_P M \overline{M}}\]
 
@@
         \item Let \(m_0,\mathit{pinit}(m_0) \hookrightarrow \MP\) and \(m_0' \hookrightarrow M'\)
-        \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
-        \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
+        \item \(\MP \sim M'\)
       \end{itemize}
@@
               \(m_r' \hookrightarrow M'\)
-            \item If \(m_r,p_r \hookrightarrow \MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
-            \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
+            \item \(\MP \sim_P M'\)
           \end{itemize}
@@
             \item Then there is some \(m_r'\) where
+
               \(m_c' \hookrightarrow M_1' m_r' | \mathit{ret}\ m_r'\ a_{pc}\ a_{sp}\)
             \item Let \(m_r'' = \mathit{roll}_C(C,m_c,m_r,m_c',m_r')\),
+
               \((m_r,p_r) \hookrightarrow \MP_2\), and \(m_r'' \hookrightarrow M_2'\)
-            \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP_1) \cdot \mathit{Obs}(\MP_2) \simeq
-              \mathit{Obs}(M_1') \cdot \mathit{Obs}(M_2')\)
-            \item Otherwise \(\mathit{Obs}(\MP_1) \cdot \mathit{Obs}(\MP_2) \lesssim
-              \mathit{Obs}(M_1') \cdot \mathit{Obs}(M_2')\)
+            \item \(\MP_1 \MP_2 \sim_P M_1' M_2'\)
           \end{itemize}
-        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\):
-          \begin{itemize}
-            \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
-            \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
-          \end{itemize}
+        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\),
+          \(\MP \sim_P M'\)
       \end{itemize}
@@
             \item Then there is some \(m_r'\) where
+
               \(m_c' \hookrightarrow M m_r' | \mathit{ret}\ m_c'\ m_r'\)
@@
           \end{itemize}
-        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\)
-          \begin{itemize}
-            \item If \(\MP \in \top\) then \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M')\)
-            \item Otherwise \(\mathit{Obs}(\MP) \lesssim \mathit{Obs}(M')\)
-          \end{itemize}
+        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\),
+          \(\MP \sim_P M'\)
       \end{itemize}

### 2020-05-26 props.tex
Back to the "join" version.

@@
 
+    When we've taken a prefix and want to combine it with another trace by replacing its final
+    state and continuing execution, we use the join operator on any type of trace \(T_1 \cdot T_2 :
+    \mathit{trace}\ X \rightarrow \mathit{trace}\ X \rightharpoonup \mathit{trace}\ X\), defined
+    on \(T_1 \not \in \top\) and any \(T_2\):
+
+    \[T_1 \cdot T_2 \triangleq
+      \begin{cases}
+        T_2 & \text{if } T_1 = t \\
+        t T_1' \cdot T_2 & \text{if } T_1 = t T_1' \\
+      \end{cases}\]
+
     An observation trace \(O\) is the trace of observations from a machine or {\it mptrace},
@@
 
-    We define policy-sensitive observational equivalence, which relates machine traces and
-    {\it mptraces} whose observable behavior match up to a possible halt due to policy failure.
-    It relates lists of parameters as long as only the last is infinite.
+    We define policy-sensitive observational equivalence, which relates a machine trace and
+    an {\it mptrace} whose observable behavior match, up to a possible halt due to policy failure.
 
@@
            {\MP \sim_P M}\]
-
-    \[\frac{\begin{split}
-              \MP \not \in \top & \hspace{.5cm} M \not \in \top \\
-              \mathit{Obs}(\MP) \simeq \mathit{Obs}(M) & \hspace{.5cm} \overline{\MP} \sim_P \overline{M} \\
-            \end{split}}
-           {\MP \overline{\MP} \sim_P M \overline{M}}\]
 
@@
               \((m_r,p_r) \hookrightarrow \MP_2\), and \(m_r'' \hookrightarrow M_2'\)
-            \item \(\MP_1 \MP_2 \sim_P M_1' M_2'\)
+            \item \(\MP_1 \cdot \MP_2 \sim_P M_1' \cdot M_2'\)
           \end{itemize}

### 2020-05-26 props.tex
comments on section I

@@
 \documentclass[conference]{IEEEtran}
+\input{macros}
 
@@
 
+\apt{important tex note: adopt/adapt macros from macros.tex to make it easier to switch notations if desire later.}
+
     Values and addresses are words:
@@
     \[k \in \mathcal{K} ::= \mathcal{W} | \mathcal{R} | \mathtt{PC} | \mathtt{SP}\]
+
+\apt{Use set notation consistently rather than mixing in grammar notation ($::=$ and $|$).}
 
@@
     Execution is by a step function \(m \xrightarrow{o} m' : m \in \mathcal{M} \rightarrow
-    o \in \mathit{Obs} \times m' \in \mathcal{M}\).
+    o \in \mathit{Obs} \times m' \in \mathcal{M}\).\apt{Very awkward: don't repeat notation conventions in type.}
 
-    A policy tracks a mix of static and dynamic enforcement. A policy consists of
+    A policy tracks a mix of static and dynamic enforcement.\apt{??} A policy consists of
     a set of policy states, \(p \in \mathcal{P}\), a policy initializer
@@
     \end{split}\]
+    \apt{I would suggest parenthesizing the pairs $(m,p)$ unless that looks too busy. (This sort of experiment is exactly why macros are a good idea.)}
 
     Axiomatically, if \(m,p \xrightarrow{o}_P m',p'\), then \(m \xrightarrow{o} m'\).
+    \apt{Why not follow coq development by defining policy step function $\mathcal{M} \times \mathcal{P} \rightarrow \mathcal{P}$ separately and then \emph{defining}
+      combined function $\mathcal{M} \times \mathcal{P} \rightarrow \mathcal{M} \times \mathcal{P}$.}
     The initializer is a partial function from machine states to policy
@@
     {\it mptrace} is similarly a sequence of machine state-policy state pairs,
-    \((m_0,\mathit{pinit}(m_0))(m_1,p_1) \dots\) such that adjacent pairs of pairs are related
-    by \(\longrightarrow_P\). Steps being deterministic, we can relate an initial state to its trace
+    \((m_0,\mathit{pinit}(m_0))(m_1,p_1) \dots\) \apt{why does it need to start with $\mathit{pinit}$?}
+    such that adjacent pairs of pairs are related
+    by \(\longrightarrow_P\). Steps being deterministic\apt{awkward}, we can relate an initial state to its trace
     via \(m \hookrightarrow M : m \in \mathcal{M} \rightarrow M \in \mathit{trace}\ \mathcal{M}\):
@@
 
+    \apt{Note that this is a coinductive relation.}
+
     Since the step function is total, if \(m \hookrightarrow M\), then \(M\) must be infinite.
     We will also wish to take a prefix of the trace from an initial state up to when some condition
-    holds. Let \(f\) be a predicate on machine states. Then we define \(m \hookrightarrow M | f\)
-    (read ``\(M\) is the prefix from \(m\) until \(f\)'') where \(f\) may bind states in \(M\):
+    holds\apt{``up through the state where some first condition first holds'' (?)}.
+    Let \(f\) be a predicate on machine states. Then we define \(m \hookrightarrow M | f\)
+    (read ``\(M\) is the prefix from \(m\) until \(f\)'') where \(f\) may bind\apt{wrong word, and this phrase is redundant anyhow} states in \(M\):
     \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1 \hookrightarrow M | f }
@@
 
+           \apt{What if $f$ holds on $m_0$ ? }
+
     We similarly define \(\hookrightarrow_P\), writing \(m_0,\mathit{pinit}(m_0) \longrightarrow \bot\) when
     there is no \(m_1,p_1\) such that \(m_0,\mathit{pinit}(m_0) \xrightarrow{o} m_1,p_1\) for any \(o\).
+    \apt{What does $\mathit{pinit}$ have to do with it?. In any case, suggest using $\nrightarrow$ or similar.}
     \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} m_1,p_1 \hookrightarrow_P \MP}
@@
     With a policy we may see \(m,p \hookrightarrow \MP\) where \(\MP\) is finite due to a policy
-    failure. We will identify the set of infinite traces as \(\top\), so \(\MP \in \top\) will
+    failure. We will identify the set of infinite traces \apt{over what? Don't think set is the right notion here.}
+    as \(\top\), so \(\MP \in \top\) will
     indicate an infinite trace and \(\MP \not \in \top\) a finite one.
@@
 
+    \apt{If we're defining $\mathit{trace}$ generically, why not do so earlier to avoid multiple definitions of $\hookrightarrow$ etc.?}
     \[T_1 \cdot T_2 \triangleq
       \begin{cases}
-        T_2 & \text{if } T_1 = t \\
+        t T_2 & \text{if } T_1 = t \\
         t T_1' \cdot T_2 & \text{if } T_1 = t T_1' \\
@@
 
+      \apt{But actually, the coq definition is defined for all $T_1$ (even if that isn't very useful). And isn't it important to let $T_2$ be empty?}
     An observation trace \(O\) is the trace of observations from a machine or {\it mptrace},
@@
 
+    \apt{This is not quite the same as the coq definition. Is it really equivalent?}
+
     Equivalence of observation traces is defined as traces prefixing each other:
     \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
+
+    \apt{Ditto.}
 
@@
            {\MP \sim_P M}\]
+
+           \apt{Does this notion appear in Coq? It seems potentially useful, but it seems to require
+             knowing whether we are in the finite or infinite case.}
 
@@
 
-      The final state of a call sequence is a {\it call state}. If \(M\) is a call sequence with
-      we map it to its call state, number of arguments, and the PC and SP values of its eventual
+      \apt{But the \emph{properties} don't need these notions!}
+
+      The final state of a call sequence \apt{what's that ??} is a {\it call state}. If \(M\) is a call sequence with
+      we map it \apt{??} to its call state, number of arguments, and the PC and SP values of its eventual
       return by the function \(\mathit{call} : \mathit{trace}\ \mathcal{M} \rightharpoonup \mathcal{M}

### 2020-05-27 main.tex
Paper on formal CHERI security properties
Outline

@@
 
-\title{Interfering with Noninterference
+\title{Stack Safety as Security Property}
+%\title{The Meaning of/What is Stack Safety}
 \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF
@@
 
+\section{Introduction}
+
+Everybody likes stack safety. Nobody can say what it is precisely.
+
+The intention is to protect callers from callees.
+
+Examples of what it isn't in the literature.
+Small bad examples of things that can go wrong.
+
+Turn this into a positive security property: integrity + confidentiality.
+
+Challenges?
+Shifting notion of who the attacker is throughout execution.
+
+Roadmap: 
+Preliminaries before all these.
+
+Contributions:
+\begin{itemize}
+\item A formal property capturing the informal understanding of what ``stack safety'' is.
+\item Evidence that this property really does capture the folk intuition: we show that the stack
+  unsafe examples from the literature also fail our property.
+\item Evidence that the property is not too strong. It's realizible by
+  a previously proposed  micropolicy enforcement mechanism.
+\item A (inductive? step-by-step/stepwise) variant of this property
+  tuned for testing/verification together with a proof that it implies
+  the original property.  We test the whole thing with a cool
+  framework.
+\item A lazy version of the property that allows for more efficient/dynamic enforcement.
+  Maybe testing.
+\item TODO: Logic? Reasoning Principles?
+\end{itemize}
+
+Related work pointer + conclusions.
+
+\section{Preliminaries/Background}
+%\input{props}
+
+\begin{itemize}
+\item machine model
+\item Calling conventions
+\end{itemize}
+
+\section{Related Work}
+
 \iftext
@@
   flow a la Birkedal et al!}
+
+\section{Future Work}
+
+Is there a knowledge-based presentation of all this?
 
@@
 
+\item Rigorous engineering for hardware security: Formal modelling and proof in
+the CHERI design and implementation process
+\url{https://www.cl.cam.ac.uk/research/security/ctsrd/pdfs/202005oakland-cheri-formal.pdf}
+
 \item \url{https://www.cis.upenn.edu/~stevez/papers/LZ05a.pdf} ??

### 2020-05-28 props.tex
Updated props to match coq better and use macros

--- 2020-05-26 props.tex
@@
 \usepackage{amssymb,amsmath}
-
-\newcommand{\MP}{\mathit{MP}}
 
@@
     Values and addresses are words:
-    \[a,w \in \mathcal{W}\]
+    \[\word,\addr \in \WORDS\]
 
     Components are words (addresses) and registers, including special registers:
-    \[r \in \mathcal{R}\]
-    \[k \in \mathcal{K} ::= \mathcal{W} | \mathcal{R} | \mathtt{PC} | \mathtt{SP}\]
+    \[\reg \in \REGS\]
+    \[\component \in \COMPONENTS = \WORDS + \REGS + \SP + \PCname\]
 
     A machine state maps components to values:
-    \[m \in \mathcal{M} ::= \mathcal{K} \rightarrow \mathcal{W}\]
+    \[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]
 
     Our properties are defined in terms of observations, which may be silent or a word of data:
-    \[o \in \mathit{Obs} ::= \tau | w \in \mathcal{W}\]
+    \[\obs \in \OBSS = \tau | \word \in \WORDS\]
 
-    Execution is by a step function \(m \xrightarrow{o} m' : m \in \mathcal{M} \rightarrow
-    o \in \mathit{Obs} \times m' \in \mathcal{M}\).
+    Execution is by a step function \(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow
+    \OBSS \times \MACHS\).
 
-    A policy tracks a mix of static and dynamic enforcement. A policy consists of
-    a set of policy states, \(p \in \mathcal{P}\), a policy initializer
-    \(pinit : \mathcal{M} \rightharpoonup \mathcal{P}\), and a policy-parameterized step function:
-    \[\begin{split}
-      m, p \xrightarrow{o}_P m', p' : m \in \mathcal{M} \times p \in \mathcal{P} & \rightharpoonup \\
-      o \in \mathit{Obs} \times m' \in \mathcal{M} \times p' \in \mathcal{P} &
-    \end{split}\]
+    A policy is a general model of an enforcement mechanism, consisting of
+    a set of policy states, \(\pol \in \POLS\), a policy initializer
+    \(pinit : \MACHS \rightharpoonup \POLS\), and a policy step function \(\pol \stepstopol \pol' :
+    \POLS \rightharpoonup \POLS\). The initializer and step function are both partial functions,
+    the former to reflect static guarantees (some initial states may not be valid at all) and
+    the latter to reflect dynamic fail-stop behavior. We then define machine-policy pairs and
+    a step function on them:
 
-    Axiomatically, if \(m,p \xrightarrow{o}_P m',p'\), then \(m \xrightarrow{o} m'\).
-    The initializer is a partial function from machine states to policy
-    states, reflecting that a policy may apply static constraints on the initial state. The parameterized
-    step relation is likewise partial to allow for fail-stop behavior.
+    \[\MPS = \MACHS \times \POLS\]
+    \[\stepstoobs{}_P \subseteq \MPS \times \MPS\]
 
-  \paragraph{Machine, Policy, and Observation Traces}
+    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} \pol_1 \stepstopol \pol_2}
+           {\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}}\]
 
-    A machine trace \(M\) is a sequence of machine states \(m_0 m_1 m_2 \dots\) such that each
-    pair of adjacent states are related by \(\longrightarrow\) with some observation. An
-    {\it mptrace} is similarly a sequence of machine state-policy state pairs,
-    \((m_0,\mathit{pinit}(m_0))(m_1,p_1) \dots\) such that adjacent pairs of pairs are related
-    by \(\longrightarrow_P\). Steps being deterministic, we can relate an initial state to its trace
-    via \(m \hookrightarrow M : m \in \mathcal{M} \rightarrow M \in \mathit{trace}\ \mathcal{M}\):
+  \paragraph{Traces}
 
-    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} m_1 \hookrightarrow M}
-           {m_0 \hookrightarrow m_0 M}\]
+    A trace is a potentially infinite, nonempty sequence. We define the append operation \(T \cdot T'\)
+    where \(T'\) may be a trace or empty (\(\varepsilon\)):
+    \[T_1 \cdot T_2 \triangleq
+      \begin{cases}
+        T_1 & \text{if } T_2 = \varepsilon \\
+        t T_2 & \text{if } T_1 = t \\
+        t T_1' \cdot T_2 & \text{if } T_1 = t T_1' \\
+      \end{cases}\]
 
-    Since the step function is total, if \(m \hookrightarrow M\), then \(M\) must be infinite.
-    We will also wish to take a prefix of the trace from an initial state up to when some condition
-    holds. Let \(f\) be a predicate on machine states. Then we define \(m \hookrightarrow M | f\)
-    (read ``\(M\) is the prefix from \(m\) until \(f\)'') where \(f\) may bind states in \(M\):
-    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1 \hookrightarrow M | f }
-           {m_0 \hookrightarrow m_0 M | f}\]
-    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} f\ m_1}
-           {m_0 \hookrightarrow m_0 m_1}\]
+    We identify infinite traces with \(\isinf(T\). We define a ``trace-of'' operator \(\hookrightarrow\)
+    that coinductively relates a machine state with the trace of machine states paired with observations
+    produced by repeated application of step:
 
-    We similarly define \(\hookrightarrow_P\), writing \(m_0,\mathit{pinit}(m_0) \longrightarrow \bot\) when
-    there is no \(m_1,p_1\) such that \(m_0,\mathit{pinit}(m_0) \xrightarrow{o} m_1,p_1\) for any \(o\).
-    \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} m_1,p_1 \hookrightarrow_P \MP}
-           {m_0,p_0 \hookrightarrow_P m_0,p_0 \MP}\]
-    \[\frac{m_0,p_0 \longrightarrow \bot}
-           {m_0,p_0 \hookrightarrow_P m_0,p_0}\]
+    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
+           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]
 
-    With a policy we may see \(m,p \hookrightarrow \MP\) where \(\MP\) is finite due to a policy
-    failure. We will identify the set of infinite traces as \(\top\), so \(\MP \in \top\) will
-    indicate an infinite trace and \(\MP \not \in \top\) a finite one.
-    We define a similar notion of taking the prefix until a condition:
+    Since the step function is total, if \(\mach \hookrightarrow \machT\), then \(\machT\) must be infinite.
+    We will also wish to take a prefix of the trace from an initial state up through 
+    the first state where some condition holds. Let \(f\) be a predicate on machine states.
+    Then we define \(\mach \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix from
+    \(\mach\) until \(f\)''):
+    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm} \mach_1
+              \hookrightarrow \machT | f}
+           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT | f}\]
+    \[\frac{f\ m}
+           {m \hookrightarrow (m,\tau)}\]
 
-    \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1,p_1
-              \hookrightarrow_P M | f }
-           {m_0,p_0 \hookrightarrow_P m_0,p_0 M | f}\]
-    \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} f\ m_1}
-           {m_0,p_0 \hookrightarrow_P (m_0,p_0) (m_1,p_1)}\]
+    We define \(\hookrightarrow_P\) to relate a machine-policy pair to the trace of MP pairs and observations
+    created by \(\stepstoobs{}_P\):
 
-    An observation trace \(O\) is the trace of observations from a machine or {\it mptrace},
-    as defined by \(\mathit{Obs}\):
-    \[\frac{M = mM' \hspace{.5cm} m \xrightarrow{o} m'}
-           {\mathit{Obs}(M) = o\mathit{Obs}(M')}\]
-    \[\frac{M = m}
-           {\mathit{Obs}(M) = \tau}\]
-    \[\frac{\MP = (m,p)\MP' \hspace{.5cm} m,p \xrightarrow{o} m',p'}
-           {\mathit{Obs}(\MP) = o \mathit{Obs}(\MP')}\]
-    \[\frac{\MP = m,p}
-           {\mathit{Obs}(M) = \tau}\]
+    \[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
+           {(\mach_0,\pol_0) \hookrightarrow (\mach_0,\pol_0)}\]
+    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
+              (\mach_1,\pol_1) \hookrightarrow \machT}
+           {(\mach_0,\pol_0) \hookrightarrow_P (\mach_0,\pol_0,\obs) \machT}\]
 
-    An observational prefix \(O_1 \lesssim O_2\) means that \(O_1\) is a prefix of \(O_2\) up to the deletion
-    of \(\tau\) observations:
-    \[\begin{split}
-      O_1 & \lesssim O_2 \triangleq \forall o . o \in O_1 \Rightarrow o = \tau \lor \\
-      & \exists O_1' . O_1 = \tau O_1' \land O_1' \lesssim O_2 \lor \\
-      & \exists O_2' .  O_2 = \tau O_2' \land O_1 \lesssim O_2' \lor \\
-      & \exists O_1' O_2' . O_1 = w O_1' \land O_2 = w O_2' \land O_1' \lesssim O_2' \\
-    \end{split}\]
+    \[\frac{\mach_0,\pol_0 \xrightarrow{\obs} \mach_1,\pol_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm}
+              \mach_1,\pol_1 \hookrightarrow_P \machT | f }
+           {\mach_0,\pol_0 \hookrightarrow_P \mach_0,\pol_0 \machT | f}\]
+    \[\frac{\mach_0,\pol_0 \xrightarrow{\obs} \mach_1,\pol_1 \hspace{.5cm} f\ \mach_1}
+           {\mach_0,\pol_0 \hookrightarrow_P (\mach_0,\pol_0) (\mach_1,\pol_1)}\]
+
+    [Coq difference alert: this operation steps until it reaches a state on which \(f\) holds,
+    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
+    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
+    Still thinking about this.]
+
+    We project the machine states from either machine trace with \(\pi_m\) and the observations with \(\pi_o\).
+    An observational prefix \(\obsT_1 \lesssim \obsT_2\) means that the trace \(\obsT_1\) of observations is
+    a prefix of \(\obsT_2\) up to the deletion of \(\tau\) observations:
+
+    \[\frac{}
+           {\obsT \lesssim \obsT}\]
+
+    \[\frac{}
+           {\tau \lesssim \obsT}\]
+
+    \[\frac{\obsT_1 = \tau \obsT_1' \hspace{.5cm} \obsT_1' \lesssim \obsT_2}
+           {\obsT_1 \lesssim \obsT_2}\]
+
+    \[\frac{\obsT_2 = \tau \obsT_2' \hspace{.5cm} \obsT_1 \lesssim \obsT_2'}
+           {\obsT_1 \lesssim \obsT_2}\]
+
+    \[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2'}
+           {\obsT_1 \lesssim \obsT_2}\]
+
+    [This is the same as the coq version minus a redundant case.]
 
@@
 
-    We define policy-sensitive observational equivalence, which relates machine traces and
-    {\it mptraces} whose observable behavior match up to a possible halt due to policy failure.
-    It relates lists of parameters as long as only the last is infinite.
+    [Coq difference alert: this should be equivalent to the coq version, and we should adopt it.]
 
-    \[\frac{\MP \not \in \top \hspace{.5cm} \mathit{Obs}(\MP) \lesssim \mathit{Obs}(M)}
-           {\MP \sim_P M}\]
+    We define policy-sensitive observational equivalence, which relates a machine trace and
+    an {\it mptrace} whose observable behavior match it, up to a possible halt due to policy failure.
 
-    \[\frac{\MP \in \top \hspace{.5cm} \mathit{Obs}(\MP) \simeq \mathit{Obs}(M)}
-           {\MP \sim_P M}\]
+    \[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
+           {\MPT \sim_P \machT}\]
 
-    \[\frac{\begin{split}
-              \MP \not \in \top & \hspace{.5cm} M \not \in \top \\
-              \mathit{Obs}(\MP) \simeq \mathit{Obs}(M) & \hspace{.5cm} \overline{\MP} \sim_P \overline{M} \\
-            \end{split}}
-           {\MP \overline{\MP} \sim_P M \overline{M}}\]
+    \[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
+           {\MPT \sim_P \machT}\]
 
-  \paragraph{Calling Conventions and Contours}
+    [Todo: update coq to use this.]
 
-      A subroutine call with \(n\) arguments follows the following convention:
+  \paragraph{Call Maps and Returns}
 
-      \begin{itemize}
-        \item The stack pointer is increased by \(n+1\)
-        \item A {\tt Jal} occurs
-        \item The return address is stored at \(\mathtt{SP} - 1\)
-      \end{itemize}
+    A {\it call map} identifies machine states that represent a call, abstracting away the explicit
+    calling convention. Formally it maps some machine states to a number of arguments:
+    \[\callmap \in \CALLMAPS = \mach \rightharpoonup \mathbb{N}\]
 
-      The final state of a call sequence is a {\it call state}. If \(M\) is a call sequence with
-      we map it to its call state, number of arguments, and the PC and SP values of its eventual
-      return by the function \(\mathit{call} : \mathit{trace}\ \mathcal{M} \rightharpoonup \mathcal{M}
-      \times \mathbb{N} \times \mathcal{W} \times \mathcal{W}\). We can also track a call sequence in the
-      policy state, defining \(\mathit{call}_P : \mathcal{M} \times \mathcal{P} \rightharpoonup \mathbb{N}
-      \times \mathcal{W} \times \mathcal{W}\) such that:
+    As it pertains to a typical calling convention, for purposes of understanding the properties,
+    a call map may be considered to identify the state just before the \(\JAL\), in which space for
+    arguments is already allocated. But the mechanism is more general.
 
-      \[\frac{\begin{split}
-        m_0,\mathit{pinit}(m_0) & \hookrightarrow_P \dots m,p \dots \\ % \hspace{.5cm}
-        m_0 & \hookrightarrow \dots M \dots \\
-        \mathit{call}(M) & = m, n, a_{pc}, a_{sp}\end{split}}
-             {\mathit{call}_P(m,p) = n, a_{pc}, a_{sp}}\]
+    A state on which a call map is defined is referred to as a {\it call state}, and we consider
+    the call to have returned the first time control returns to the instruction following the
+    \(\JAL\) with the stack restored. Formally this is a relation between the states:
 
-      A state is considered to have returned after a call if the stack pointer is the same as before the call
-      (\(a_{sp}\)) and the program counter is the next address after (\(a_{pc}\)):
-      \[\mathit{ret}\ m\ a_{pc}\ a_{sp} \triangleq m(\mathtt{SP}) = a_{sp} \land m(\mathtt{PC}) = a_{pc}\]
+    \[\begin{split}
+      \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) = \mach_c(\PCname)+1 \land \\
+                                        & \mach_r(\SP) = \mach_c(\SP)
+    \end{split}\]
 
-  \section{Observable Stack Safety}
+  \section{Observable and Eager Stack Safety}
 
-    {\it Observable stack safety} is a property of a transition system \(\longrightarrow_p\),
-    which is decomposed into three subproperties: {\it observable stack integrity},
-    {\it observable stack confidentiality}, and {\it observable safe initialization}.
+    A stack safety property is a property of a transition system, a policy, and a callmap
+    that is decomposed into an integrity property, a confidentiality property, and a safe
+    initialization property. We will present eager and observational variants of stack safety.
+
+    {\it Eager stack integrity} states that if a function call returns, the caller's stack data
+    will be identical to what it was before the call. This implies the weaker {\it observable
+    stack integrity}, which states that any changes to the caller's frame will produce the same
+    observable behavior.
+
+    {\it Eager stack confidentiality} states that data outside a callee's stack frame is secret,
+    and does not interfere with its observable behavior or with the machine state at return. Its
+    weaker variant {\it observable stack safety} lifts the latter restriction and states that
+    the callee may not leak these secrets directly or by changing the behavior of its caller.
+
+    Finally, {\it safe initialization} deals with the top level outside of a function call, in
+    which uninitiaized data should not interfere with any observable behavior.
+
+    [Coq difference alert: safe initialization is implemented differently in each property, since
+     it's really just confidentiality. But because there's no return it's functionally the same.]
+
+    Our top level properties are {\it eager stack safety} and {\it observable stack safety}:
+    \[\begin{split}
+      & \textit{observable stack safety} \triangleq \\
+      & \textit{observable stack integrity} \land \\
+      & \textit{observable stack confidentiality} \land \\
+      & \textit{safe initialization}
+    \end{split}\]
+    \[\begin{split}
+      & \textit{eager stack safety} \triangleq \\
+      & \textit{eager stack integrity} \land \\
+      & \textit{eager stack confidentiality} \land \\
+      & \textit{safe initialization}
+    \end{split}\]
 
@@
   %  called a {\it contour} that maps components to levels of confidentiality and integrity.
-  %  Specifically, we have confidentiality labels \(\mathit{labelC} ::= \{\mathtt{HC}, \mathtt{LC}\}\)
-  %  and integrity labels \(\mathit{labelI} ::= \{\mathtt{HI}, \mathtt{LI}\}\), and a
+  %  Specifically, we have confidentiality labels \(\mathit{labelC} ::= \{\HIGHSEC, \LOWSEC\}\)
+  %  and integrity labels \(\mathit{labelI} ::= \{\HIGHINT, \LOWINT\}\), and a
   %  type \(\mathit{contour}\)
 
-    \paragraph{Observable Safe Initialization}
+    \paragraph{Safe Initialization}
 
@@
 
-      Let \(C : \mathcal{K} \rightarrow \{\mathtt{HC},\mathtt{LC}\}\) map all addresses
-      to \(\mathtt{HC}\) (high confidentiality) and all registers to \(\mathtt{LC}\)
-      (low confidentiality). In the style of non-interference, a pair of states \(m\)
-      and \(m'\) are variants with regard to \(C\), \(m \approx_C m'\), if:
-      \[\forall k . C(k) = \mathtt{LC} \rightarrow m(k) = m'(k)\]
+      Let \(\contour : \COMPONENTS \rightarrow \{\HIGHSEC,\LOWSEC\}\) map all addresses
+      to \(\HIGHSEC\) (high confidentiality) and all registers to \(\LOWSEC\)
+      (low confidentiality). In the style of non-interference, a pair of states \(\mach\)
+      and \(\mach'\) are variants with regard to \(\contour\), \(\mach \approx_\contour \mach'\), if:
+      \[\forall \component . \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
 
-      A system enjoys {\it observable safe initialization} if for any initial states \(m_0\)
-      and \(m_0'\) such that \(m_0 \approx_C m_0'\):
+      A system enjoys {\it safe initialization} if for any initial states \(\mach_0\)
+      and \(\mach_0'\) such that \(\mach_0 \approx_\contour \mach_0'\):
 
       \begin{itemize}
-        \item Let \(m_0,\mathit{pinit}(m_0) \hookrightarrow \MP\) and \(m_0' \hookrightarrow M'\)
-        \item \(\MP \sim M'\)
+        \item Let \(\mach_0,\mathit{pinit}(\mach_0) \hookrightarrow \MPT\) and \(\mach_0' \hookrightarrow \machT'\)
+        \item \(\MPT \sim_P \machT'\)
       \end{itemize}
@@
 
-      \[\mathit{label} ::= \{\mathtt{HC},\mathtt{LC}\} \times \{\mathtt{HI},\mathtt{LI}\}\]
+      \[\mathit{label} ::= \{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT,\LOWINT\}\]
 
-      \[C \in \mathcal{C} ::= \mathcal{K} \rightarrow \mathit{label}\]
+      \[\contour \in \CONTOURS ::= \COMPONENTS \rightarrow \mathit{label}\]
 
-      When \(m,p\) is a call state \(\mathit{call}_P(m,p) = n,\_,\_\) for some \(n\),
-      we can construct a contour based on the size of \(m\)'s stack frame, as follows:
-      \[\mathit{Cof}(m,n)(k) =
+      When \(\mpstate{\mach}{\pol}\) is a call state \(\callmap(\mach) = n\) for some \(n\),
+      we can construct a contour based on the bounds of the callee's stack frame, as follows:
+      \[\mathit{Cof}(\mach,n)(\component) =
       \begin{cases}
-        (\mathtt{HC},\mathtt{HI}) & \text{if } k \leq m(\mathtt{SP}) - n \\
-        (\mathtt{HC},\mathtt{LI}) & \text{if } k \geq m(\mathtt{SP}) \\
-        (\mathtt{LC},\mathtt{HI}) & \text{if } k = m(\mathtt{SP}) - 1 \\
-        (\mathtt{LC},\mathtt{LI}) & \text{else} \\
+        (\HIGHSEC,\HIGHINT) & \text{if } \component \leq \mach(\SP) - n \\
+        (\HIGHSEC,\LOWINT) & \text{if } \component \geq \mach(\SP) \\
+        (\LOWSEC,\LOWINT) & \text{else} \\
       \end{cases}\]
@@
       from the perspective of the callee, that memory is not initialized. Below the old stack
-      pointer, memory is high integrity and high confidentiality. The top of the callee's frame
-      is its return address, which it must be able to read in order to return but which is
-      high integrity [does this make sense?] Registers and the callee's frame are low integrity
-      and low confidentiality. These categories inform our integrity and confidentiality properties.
+      pointer, memory is high integrity and high confidentiality. Registers and the callee's frame
+      are low integrity and low confidentiality. These categories inform our integrity and
+      confidentiality properties.
+
+    \paragraph{Eager Integrity}
+
+      A system enjoys eager integrity with regard to \(\callmap\) if for any initial state \(\mach_0\),
+      call state \(\mach_c\) where \(\callmap(\mach_c) = n\), and policy state \(\pol_c\), and
+      return state \(\mpstate{\mach_r,\pol_r}\), where:
+      \[\mach_0,\mathit{pinit}(\mach_0) \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...
+        \mpstate{\mach_r}{\pol_r,\_} | \ret{\mach_c}\]
+
+      it is always the case that:
+      \[\forall \component . \contour(\component) = \HIGHINT \rightarrow \mach_c(\component) = \mach_r(\component)\]
 
@@
 
-      Observable integrity is concerned with the writes a call makes to high integrity components,
+      Observable integrity is concerned with the changes a callee makes to high integrity components,
       and the effects those writes have on the observable behavior of its caller
       {\it after its return}. Given a contour, a call state, and a returned state, we define
-      a ``rollback'' function \(\mathit{roll}_I : \mathcal{C} \times \mathcal{M} \times \mathcal{M}
-      \rightarrow \mathcal{M}\). The rollback creates a state that matches the call
-      state on those components that are high integrity in \(C\), and matches the return state
+      a ``rollback'' function \(\mathit{roll}_I : \CONTOURS \times \MACHS \times \MACHS
+      \rightarrow \MACHS\). The rollback creates a state that matches the call
+      state on those components that are high integrity in its contour, and matches the return state
       on low integrity components:
 
-      \[\mathit{roll}_I(C,m_c,m_r)(k) =
+      \[\mathit{roll}_I(\contour,\mach_c,\mach_r)(\component) =
       \begin{cases}
-        m_c(k) & C(k) = (\mathtt{HI},\_) \\
-        m_r(k) & C(k) = (\mathtt{LI},\_) \\
+        \mach_c(\component) & \contour(\component) = (\HIGHINT,\_) \\
+        \mach_r(\component) & \contour(\component) = (\LOWINT,\_) \\
       \end{cases}\]
@@
 
-      A system enjoys observable integrity as follows:
+      A system enjoys observable integrity with regard to \(\callmap\) as follows:
 
-      For any initial state \(m_0\), call state \(m_c\), and policy state \(p_c\) such that
-      \((m_0,\mathit{pinit}(m_0)) \hookrightarrow ... (m_c,p_c) ...\)
-      and \(\mathit{call}_P(m_c,p_c) = n,a_{pc},a_{sp}\), if 
-      \((m_c,p_c) \hookrightarrow ... (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\):
+      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...\)
+      and \(\callmap(\mach_c) = n\), if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow ...
+        \mpstate{\mach_r}{\pol_r,\_} | \ret{\mach_c}\):
 
           \begin{itemize}
-            \item Let \(C = \mathit{Cof}(m_c,n)\)
-            \item Let \(m_r' = \mathit{roll}_I(C,m_c,m_r)\), \((m_r,p_r) \hookrightarrow \MP\), and
-              \(m_r' \hookrightarrow M'\)
-            \item \(\MP \sim_P M'\)
+            \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
+            \item Let \(\mach_r' = \mathit{roll}_I(\contour,\mach_c,\mach_r)\), \(\mpstate{\mach_r}{\pol_r}
+              \hookrightarrow \MPT\), and \(\mach_r' \hookrightarrow \machT'\)
+            \item \(\MPT \sim_P \machT'\)
           \end{itemize}
 
-    \paragraph{Eager Integrity}
+    \paragraph{Eager Confidentiality}
 
-      An eager enforcement mechanism implements a stronger property, which we term {\it eager integrity}.
-      Eager integrity simplifies observable integrity by requiring that high integrity data of \(m_r\) be
-      identical to that of \(m_c\). Formally, a system enjoys eager integrity if for any initial state
-      \(m_0\), call state \(m_c\), and policy state \(p_c\) such that \(\mathit{call}(m_c,p_c) =
-      n, a_{pc}, a_{sp}\), and some \(m_r\) and \(p_r\) such that
-      \(m_0,\mathit{pinit}(m_0) \hookrightarrow ... (m_c,p_c) ... (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\):
+      A system enjoys eager stack confidentiality with regard to \(\callmap\) if, for any initial state
+      \(\mach_0\) and call state \(\mach_c\) such that
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...\) and
+      \(\callmap(\mach_c) = n\):
 
-      \[\forall k . C(k) = \mathtt{HI} \rightarrow m_c(k) = m_r(k)\]
+      \begin{itemize}
+        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
+        \item For all \(\mach_c'\) such that \(\mach_c \approx_\contour \mach_c'\):
+        \item If \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT (\mach_r,p_r,\_) | \ret{\mach_c}\)
+          \begin{itemize}
+            \item Then there is some \(\mach_r'\) where
 
-      This implies lazy integrity, as the rollback function will produce \(m_r' = m_r\).
+              \(\mach_c' \hookrightarrow \machT' (\mach_r',_) | \ret{\mach_c'}\)
+            \item \(\mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT')\)
+            \item And for all \(\component\), if \(\mach_c(\component) \not = \mach_r(\component)\) or
+              \(\mach_c'(\component) \not = \mach_r'(\component)\), \(\mach_r(\component) = \mach_r'(\component)\)
+          \end{itemize}
+        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
+          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
+      \end{itemize}
 
@@
       Observable confidentiality is concerned with whether high confidentiality data influences
-      observable behavior during and after a call. We define variants given the generalized notion
+      observable both behavior during and after a call. We define variants given the generalized notion
       of contours:
-      \[m \approx_C m' \triangleq \forall k . C(k) = \mathtt{LC} \rightarrow m(k) = m'(k)\]
+      \[\mach \approx_\contour \mach' \triangleq \forall \component .
+        \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
 
@@
       In this case, a confidentiality rollback takes a contour and four states: two call states
-      that will be variants, and two return states. It is not symmetric, regarding \(m_c\) and \(m_r\)
-      as the ``real'' call and return, \(m_c'\) as a variant and \(m_r'\) as a return that follows it.
+      that will be variants, and two return states. It is not symmetric, regarding \(\mach_c\) and \(\mach_r\)
+      as the ``real'' call and return, \(\mach_c'\) as a variant and \(\mach_r'\) as a return that follows it.
       \[\begin{split}
-        & \mathit{roll}_C(C,m_c,m_c',m_r,m_r')(k) = \\
+        & \mathit{roll}_C(\contour,\mach_c,\mach_c',\mach_r,\mach_r')(\component) = \\
         & \begin{cases}
-          m_c(k) & \text{if } C(k) = \mathtt{HC} \text{ and } \\
-                  & m_c(k) = m_r(k) \text{ and } m_c'(k) = m_r'(k) \\
-          m_r'(k) & \text{else} \\
+          \mach_c(\component) & \text{if } C(\component) = \HIGHSEC \text{ and } \\
+                  & \mach_c(\component) = \mach_r(\component) \text{ and }
+                    \mach_c'(\component) = \mach_r'(\component) \\
+          \mach_r'(\component) & \text{else} \\
         \end{cases}
@@
 
-      A system enjoys observable confidentiality as follows:
+      A system enjoys observable confidentiality with respect to \(\callmap\) as follows:
 
-      For any initial state \(m_0\), call state \(m_c\), and policy state \(p_c\) such that
-      \(m_0,\mathit{pinit}(m_0) \hookrightarrow ... (m_c,p_c) ...\) and
-      \(\mathit{call}(m_c,p_c) = n, a_{pc}, a_{sp}\):
+      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... (\mach_c,\pol_c,\_) ...\) and
+      \(\callmap(\mach_c) = n\):
 
       \begin{itemize}
-        \item Let \(C = \mathit{Cof}(m_c,n)\)
-        \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
-        \item If \((m_c,p_c) \hookrightarrow \MP_1 (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\)
+        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
+        \item For all \(\mach_c'\) such that \(\mach_c \approx_C \mach_c'\):
+        \item If \(\mpstate{\mach_c}{p_c} \hookrightarrow \MPT_1 (\mach_r,p_r,\_) | \ret{\mach_c}\)
           \begin{itemize}
-            \item Then there is some \(m_r'\) where
+            \item Then there is some \(\mach_r'\) where
 
-              \(m_c' \hookrightarrow M_1' m_r' | \mathit{ret}\ m_r'\ a_{pc}\ a_{sp}\)
-            \item Let \(m_r'' = \mathit{roll}_C(C,m_c,m_r,m_c',m_r')\),
+              \(\mach_c' \hookrightarrow \machT_1' (\mach_r',\_) | \ret{\mach_c'}\)
+            \item Let \(\mach_r'' = \mathit{roll}_C(\contour,\mach_c,\mach_r,\mach_c',\mach_r')\),
 
-              \((m_r,p_r) \hookrightarrow \MP_2\), and \(m_r'' \hookrightarrow M_2'\)
-            \item \(\MP_1 \MP_2 \sim_P M_1' M_2'\)
+              \(\mpstate{\mach_r}{p_r} \hookrightarrow \MPT_2\), and \(\mach_r'' \hookrightarrow \machT_2'\)
+            \item \(\MPT_1 \cdot \MPT_2 \sim_P \MPT_1' \cdot \MPT_2'\)
           \end{itemize}
-        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\),
-          \(\MP \sim_P M'\)
+        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
+          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
       \end{itemize}
 
-    \paragraph{Eager Confidentiality}
-
-      Once again an eager policy enforces a stronger property that only concerns the behavior of
-      the original trace and its variant within each call. A system enjoys {\it eager stack
-      confidentiality} if, for any initial state \(m_0\) and call state \(m_c\) such that
-      \(m_0,\mathit{pinit}(m_0) \hookrightarrow ... (m_c,p_c) ...\) and
-      \(\mathit{call}(m_c,p_c) = n, a_{pc}, a_{sp}\):
-
-      \begin{itemize}
-        \item Let \(C = \mathit{Cof}(m_c,n)\)
-        \item For all \(m_c'\) such that \(m_c \approx_C m_c'\):
-        \item If \((m_c,p_c) \hookrightarrow_P \MP (m_r,p_r) | \mathit{ret}\ m_r\ a_{pc}\ a_{sp}\)
-          \begin{itemize}
-            \item Then there is some \(m_r'\) where
-
-              \(m_c' \hookrightarrow M m_r' | \mathit{ret}\ m_c'\ m_r'\)
-            \item \(\mathit{Obs}(\MP) \simeq \mathit{Obs}(M)\)
-            \item And for all \(k\), if \(m_c(k) \not = m_r(k)\) or \(m_c'(k) \not = m_r'(k)\),
-              \(m_r(k) = m_r'(k)\)
-          \end{itemize}
-        \item Else, for \(\MP\) and \(M'\) such that \(m_c,p_c \hookrightarrow \MP\) and \(m_c' \hookrightarrow M'\),
-          \(\MP \sim_P M'\)
-      \end{itemize}
-
-      Eager confidentiality implies observable confidentiality, because if \(m_r\) and \(m_r'\) agree
-      on all \(k\) that changed during the call, then an \(m_r'' = \mathit{roll}_C(C,m_c,m_r,m_c',m_r')\)
-      will be identical to \(m_r\), and future observations will be the same up to a potential policy
-      stop.
-
 \end{document}

### 2020-05-28 props.tex
Some comments for Sean

@@
   \section{Preliminaries}
+
+  \leo{Low level presentation point that slightly bugs me: }
 
@@
     Our properties are defined in terms of observations, which may be silent or a word of data:
-    \[\obs \in \OBSS = \tau | \word \in \WORDS\]
+    \[\obs \in \OBSS = \tau ~ | ~ \word \in \WORDS\]
 
-    Execution is by a step function \(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow
+    Execution is by\leo{is by?} a step function \(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow
     \OBSS \times \MACHS\).
@@
     a set of policy states, \(\pol \in \POLS\), a policy initializer
-    \(pinit : \MACHS \rightharpoonup \POLS\), and a policy step function \(\pol \stepstopol \pol' :
+    \(\pinit : \MACHS \rightharpoonup \POLS\), and a policy step function \(\pol \stepstopol \pol' :
     \POLS \rightharpoonup \POLS\). The initializer and step function are both partial functions,
@@
 
-    We identify infinite traces with \(\isinf(T\). We define a ``trace-of'' operator \(\hookrightarrow\)
+    We identify infinite traces with \(\isinf(T)\). We define a ``trace-of'' operator \(\hookrightarrow\)
     that coinductively relates a machine state with the trace of machine states paired with observations
@@
 
-    [Coq difference alert: this operation steps until it reaches a state on which \(f\) holds,
+    \COQ{this operation steps until it reaches a state on which \(f\) holds,
     and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
     which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
-    Still thinking about this.]
+    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
+      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
+    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}
 
@@
            {\obsT_1 \lesssim \obsT_2}\]
+    \leo{Doesn't this case miss the recursive $\obsT_1 \lesssim \obsT_2$ condition?}
 
-    [This is the same as the coq version minus a redundant case.]
+    \COQ{This is the same as the coq version minus a redundant case.}
 
@@
 
-    [Coq difference alert: this should be equivalent to the coq version, and we should adopt it.]
+    \COQ{this should be equivalent to the coq version, and we should adopt it.}
+    \leo{We had this, but it was much harder to work with for my proofs. We should still
+      prove them equivalent if we want to present it this way (which is probably cleaner for
+      presentation purposes)}
 
@@
     [Todo: update coq to use this.]
+    \leo{Oh, I like this version. Should make things cleaner in the top part of my proofs
+      while leaving the hard low-level invariants unaffected.}
 
@@
 
-    [Coq difference alert: safe initialization is implemented differently in each property, since
-     it's really just confidentiality. But because there's no return it's functionally the same.]
+    \COQ{safe initialization is implemented differently in each property, since
+     it's really just confidentiality. But because there's no return it's functionally the same.}
 
@@
     \paragraph{Safe Initialization}
+
+    \leo{Do we really need safe initialization? Couldn't we just use the definitions of
+      integrity and confidentiality for the toplevel, pointing out that since it never
+      returns only confidentiality is interesting? Or would we miss something?}
 

### 2020-05-29 main.tex
Fix typo in main
Nit
From meeting

@@
 \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF
-    and...?)}}\fi
-}
+    and...?)}}
+\fi
 
@@
 \item Evidence that the property is not too strong. It's realizible by
-  a previously proposed  micropolicy enforcement mechanism.
+  a previously proposed  micropolicy enforcement mechanism \leo{And it
+    says yes on a substantial body of code}
 \item A (inductive? step-by-step/stepwise) variant of this property

### 2020-05-29 props.tex
From meeting

@@
 
-    A policy is a general model of an enforcement mechanism, consisting of
-    a set of policy states, \(\pol \in \POLS\), a policy initializer
-    \(\pinit : \MACHS \rightharpoonup \POLS\), and a policy step function \(\pol \stepstopol \pol' :
-    \POLS \rightharpoonup \POLS\). The initializer and step function are both partial functions,
+    \paragraph*{Policy}
+
+    A policy is a general model of an enforcement mechanism,
+    consisting of a set of policy states, \(\pol \in \POLS\) and a
+    policy step function \((\mach, \pol) \stepstopol \pol' : \MACHS
+    \times \POLS \rightharpoonup \POLS\).
+
+    The initializer and step function are both partial functions,
     the former to reflect static guarantees (some initial states may not be valid at all) and
@@
 
-    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} \pol_1 \stepstopol \pol_2}
+    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} (\mach_1, \pol_1) \stepstopol \pol_2}
            {\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}}\]
@@
     a prefix of \(\obsT_2\) up to the deletion of \(\tau\) observations:
+    \leo{Point out this is coinductive}
 
@@
 
-    \[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
+    \[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
            {\MPT \sim_P \machT}\]
 
-    \[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
+    \[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
            {\MPT \sim_P \machT}\]
@@
 
-  \paragraph{Call Maps and Returns}
+  \section{Observable and Eager Stack Safety}
+
+    \paragraph*{Call Maps and Returns}
 
@@
     calling convention. Formally it maps some machine states to a number of arguments:
-    \[\callmap \in \CALLMAPS = \mach \rightharpoonup \mathbb{N}\]
+    \[\callmap \in \CALLMAPS = \MACHS \rightharpoonup \mathbb{N}\]
 
@@
     \[\begin{split}
-      \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) = \mach_c(\PCname)+1 \land \\
+      \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) = \next(\mach_c(\PCname)) \land \\
                                         & \mach_r(\SP) = \mach_c(\SP)
@@
 
-  \section{Observable and Eager Stack Safety}
+    \leo{TODO: Initial state depends on call map}
+
+    \paragraph*{Stack safety}
 
@@
 
+      \leo{Think more about whether we vary mc or step mc}
+
     \paragraph{Observable Confidentiality}
@@
       \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
+      \leo{Move earlier}
 
       Data that is high confidentiality within a callee due either to being in its caller's frame
-      will become low confidentiality if the callee returns. We face a subtle distinction: 
+      will become low confidentiality if the callee returns.\leo{??} We face a subtle distinction: 
       a callee treats the contents of its caller's frame as secrets, and they should not influence
@@
       \end{split}\]
+      \leo{Maybe this should be called unscramble}
 
@@
 
+      \leo{Need to fix append. Maybe call it join. It should remove the last
+        state of left arg}
+
 \end{document}

### 2020-06-02 main.tex

@@
-\documentclass[conference]{IEEEtran}
+%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For double-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
+%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For single-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%% For final camera-ready submission, w/ required CCS and ACM Reference
+%\documentclass[acmsmall]{acmart}\settopmatter{}
+
+%% Journal information
+%% Supplied to authors by publisher for camera-ready submission;
+%% use defaults for review submission.
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
+\acmYear{2021}
+\acmMonth{1}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+
+%% Copyright information
+%% Supplied to authors (based on authors' rights management selection;
+%% see authors.acm.org) by publisher for camera-ready submission;
+%% use 'none' for review submission.
+\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\copyrightyear{2018}           %% If different from \acmYear
+
+%% Bibliography style
+\bibliographystyle{ACM-Reference-Format}
+%% Citation style
+%% Note: author/year citations are required for papers published as an
+%% issue of PACMPL.
+\citestyle{acmauthoryear}   %% For author/year citations
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Note: Authors migrating a paper from PACMPL format to traditional
+%% SIGPLAN proceedings format must update the '\documentclass' and
+%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Some recommended packages.
+\usepackage{booktabs}   %% For formal tables:
+                        %% http://ctan.org/pkg/booktabs
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+                        %% http://ctan.org/pkg/subcaption
+
 \input{macros}
@@
 
-\title{Stack Safety as Security Property}
-%\title{The Meaning of/What is Stack Safety}
-\iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF
-    and...?)}}
-\fi
-
-% I get a weird error, so I commented this out for now
-%\iftext
-%\IEEEauthorblockN{Leonidas Lampropoulos}
-%\IEEEauthorblockA{\textit{University of Maryland}}
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID
-%\and
-%\author{\IEEEauthorblockN{Roberto Blanco}
-%\IEEEauthorblockA{\textit{\bcp{affiliation?}}
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID
+%% Title information
+\title{Stack Safety as a Security Property}         %% [Short Title] is optional;
+                                        %% when present, will be used in
+                                        %% header instead of Full Title.
+%\titlenote{with title note}             %% \titlenote is optional;
+%                                        %% can be repeated if necessary;
+%                                        %% contents suppressed with 'anonymous'
+%\subtitle{Subtitle}                     %% \subtitle is optional
+%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
+%                                        %% can be repeated if necessary;
+%                                        %% contents suppressed with 'anonymous'
+
+%% Author information
+%% Contents and number of authors suppressed with 'anonymous'.
+%% Each author should be introduced by \author, followed by
+%% \authornote (optional), \orcid (optional), \affiliation, and
+%% \email.
+%% An author may have multiple affiliations and/or emails; repeat the
+%% appropriate command.
+%% Many elements are not rendered, but should be provided for metadata
+%% extraction tools.
+
+%% Author with single affiliation.
+%\author{First1 Last1}
+%\authornote{with author1 note}          %% \authornote is optional;
+%                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+%\affiliation{
+%  \position{Position1}
+%  \department{Department1}              %% \department is recommended
+%  \institution{Institution1}            %% \institution is required
+%  \streetaddress{Street1 Address1}
+%  \city{City1}
+%  \state{State1}
+%  \postcode{Post-Code1}
+%  \country{Country1}                    %% \country is recommended
 %}
-%\and
-%\IEEEauthorblockN{Benjamin C. Pierce}
-%\IEEEauthorblockA{\textit{University of Pennsylvania}
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID
-%  }
-%\and
-%\IEEEauthorblockN{Andrew Tolmach}
-%\IEEEauthorblockA{\textit{Portland State University}
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID
+%\email{first1.last1@inst1.edu}          %% \email is recommended
+% 
+%%% Author with two affiliations and emails.
+%\author{First2 Last2}
+%\authornote{with author2 note}          %% \authornote is optional;
+%                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+%\affiliation{
+%  \position{Position2a}
+%  \department{Department2a}             %% \department is recommended
+%  \institution{Institution2a}           %% \institution is required
+%  \streetaddress{Street2a Address2a}
+%  \city{City2a}
+%  \state{State2a}
+%  \postcode{Post-Code2a}
+%  \country{Country2a}                   %% \country is recommended
 %}
+%\email{first2.last2@inst2a.com}         %% \email is recommended
+%\affiliation{
+%  \position{Position2b}
+%  \department{Department2b}             %% \department is recommended
+%  \institution{Institution2b}           %% \institution is required
+%  \streetaddress{Street3b Address2b}
+%  \city{City2b}
+%  \state{State2b}
+%  \postcode{Post-Code2b}
+%  \country{Country2b}                   %% \country is recommended
 %}
-%\fi
-
+%\email{first2.last2@inst2b.org}         %% \email is recommended
+
+%% Abstract
+%% Note: \begin{abstract}...\end{abstract} environment must come
+%% before \maketitle command
+\begin{abstract}
+Text of abstract \ldots.
+\end{abstract}
+
+%% 2012 ACM Computing Classification System (CSS) concepts
+%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011008</concept_id>
+<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003456.10003457.10003521.10003525</concept_id>
+<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+%\ccsdesc[500]{Software and its engineering~General programming languages}
+%\ccsdesc[300]{Social and professional topics~History of programming languages}
+%% End of generated code
+
+%% Keywords
+%% comma separated list
+\keywords{Stack Safety, Micropolicy, Random Testing, RISC-V}  %% \keywords are mandatory in final camera-ready submission
+
+%% \maketitle
+%% Note: \maketitle command must come after title commands, author
+%% commands, abstract environment, Computing Classification System
+%% environment and commands, and keywords command.
 \maketitle
-
-\iftext
-\begin{abstract}
-\end{abstract}
-
-\begin{IEEEkeywords}
-% component, formatting, style, styling, insert
-\end{IEEEkeywords}
-\fi
 
@@
 Turn this into a positive security property: integrity + confidentiality.
+
+\leo{Variations of these should probably go here:}
+    {\it Eager stack integrity} states that if a function call returns, the caller's stack data
+    will be identical to what it was before the call. This implies the weaker {\it observable
+    stack integrity}, which states that any changes to the caller's frame will produce the same
+    observable behavior.
+
+    {\it Eager stack confidentiality} states that data outside a callee's stack frame is secret,
+    and does not interfere with its observable behavior or with the machine state at return. Its
+    weaker variant {\it observable stack safety} lifts the latter restriction and states that
+    the callee may not leak these secrets directly or by changing the behavior of its caller.
 
@@
 
-\section{Preliminaries/Background}
-%\input{props}
+\section{Background}
+
+We begin by describing our machine model. Concretely, we have in mind a
+conventional machine architecture with a standard (RISC-V) processor
+extended with some kind of (static or dynamic, software- or
+hardware-enforced) security monitor, but we won't need to make any
+specific choices about details until we get to \bcp{the policy one, probably}.
+%In this section we will describe an abstract interface to our machine model.
+
+Values and addresses are words drawn from some set of all words $\WORDS$. We will
+denote words with $\word$.\leo{TODO: If we ever need ``A'' come back to this.}
+
+Components are words (denoting addresses) and registers.
+%
+    \[\reg \in \REGS\]
+    \[\component \in \COMPONENTS = \WORDS + \REGS \]
+%
+The registers must include two special purpose registers: the program
+counter \PCname and the stack pointer \SP.
+
+A machine state is just a map from components to values:
+%
+\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]
+
+Our stack safety property will be defined in terms of {\em observations}, which may be
+silent (written $\tau$) or a word of data:
+%
+\[\obs \in \OBSS = \tau ~ | ~ \word \in \WORDS\]
+
+These observations are a byproduct of the step function of the
+machine, which takes a machine state and produces a pair of a
+machine state and an observation.
+%
+\(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \).
+
+\paragraph*{Policy}
+
+A policy is a general model of an enforcement mechanism, consisting of
+a set of policy states, \(\pol \in \POLS\) and a policy step function
+\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
+\rightharpoonup \POLS\).
+
+%    The initializer and step function are both partial functions,
+%    the former to reflect static guarantees (some initial states may not be valid at all) and
+%    the latter to reflect dynamic fail-stop behavior. We then define machine-policy pairs and
+%    a step function on them:
+
+Let $\MPS$ be the set of all pairs of machine states and policy states.
+%
+\[\MPS = \MACHS \times \POLS\]
+%
+Then we can lift this policy step function to operate on such pairs,
+by combining it with the regular step function for machine states.
+
+%    \[\stepstoobs{}_P \subseteq \MPS \times \MPS\]
+
+    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} (\mach_1, \pol_1) \stepstopol \pol_2}
+           {\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}}\]
+\leo{Notation: Should $\stepstoobs{}$ be partial?}
+
+\paragraph{Traces}
+
+A trace over some type $A$ is a potentially infinite, nonempty
+sequence of $A$s. We will mostly be concerned with sequences of
+machines states, usually denoted $\machT$, and with sequences of
+triples of (machine states, policy states, and observations), usually
+denoted $\MPT$.
+
+\leo{I tried the ? notation for the potentially empty trace, which reveals that
+  we're slightly informal in the second case. Expand?}
+We define the append operation \(T \cdot T?\), where \(T?\) may be a trace or empty (\(\varepsilon\)):
+\[T \cdot T? \triangleq
+  \begin{cases}
+    T & \text{if } T? = \varepsilon \\
+    t T? & \text{if } T = t \\
+    t T' \cdot T? & \text{if } T = t T' \\
+  \end{cases}\]
+
+We define the ``trace-of'' operator \(\hookrightarrow\) that coinductively
+relates a machine state with the trace of machine states paired with
+observations produced by repeated application of step:
+%
+\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
+       {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]
+%
+Since the step function is total, if \(\mach \hookrightarrow \machT\),
+then \(\machT\) must be infinite.
+
+We will also wish to take a prefix of the trace from an initial state
+up through the first state where some condition holds. Let \(f\) be
+such a predicate on machine states.  Then we define \(\mach
+\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix from
+\(\mach\) until \(f\)''):\leo{Abrupt. Explain in words?}
+    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm} \mach_1
+              \hookrightarrow \machT | f}
+           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT | f}\]
+    \[\frac{f\ m}
+           {m \hookrightarrow (m,\tau)}\]
+
+Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
+state pair to the trace of triples of machine states, policy states,
+and observations induced by by \(\stepstoobs{}_P\):
+%
+\leo{I think this would read better as not exists. Thoughts?}
+\[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
+       {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
+\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
+  \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
+       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]
+
+We also lift the prefix-up-to relation to such traces.
+\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm}
+  \neg f\ \mach_1 \hspace{.5cm} \mpstate{\mach_1}{\pol_1} \hookrightarrow_P \machT | f }
+       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT | f}\]
+
+\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm} f\ \mach_1}
+       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs}
+         \mpostate{\mach_1}{\pol_1}{\tau} | f}\]
+
+Finally, we will use \(\pi_m\) to project the trace of machine states
+out of a trace, \(\pi_p\) to project the trace of policy states, and
+\(\pi_o\) to project the observations.
+
+%    \COQ{this operation steps until it reaches a state on which \(f\) holds,
+%    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
+%    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
+%    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
+%      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
+%    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}
+
+\paragraph*{Observations and Equivalence}
+
+We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
+\leo{Benjamin suggested ``as far as an attacker is conerned in an
+  e-mail, but we haven't said anything about attackers yet.}, if the
+sequence of word-observations of $\obsT_1$ is a prefix of those of
+$\obsT_2$, that is we operate up to deletion of \(\tau\) observations, coinductively:
+
+\[\frac{} {\obsT \lesssim \obsT}\]
+
+\[\frac{} {\tau \lesssim \obsT}\]
+
+\[\frac{\obsT_1 = \tau \obsT_1' \hspace{.5cm} \obsT_1' \lesssim \obsT_2}
+       {\obsT_1 \lesssim \obsT_2}\]
+
+\[\frac{\obsT_2 = \tau \obsT_2' \hspace{.5cm} \obsT_1 \lesssim \obsT_2'}
+       {\obsT_1 \lesssim \obsT_2}\]
+
+\[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
+       {\obsT_1 \lesssim \obsT_2}\]
+
+% Fixed by Andrew:
+%\COQ{This is the same as the coq version minus a redundant case.}
+
+We then define equivalence of observation traces as traces prefixing each other:
+\[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
+
+\leo{Should we point out here the edge cases of all-tau traces?}
+
+% Fixed by Andrew:
+%    \COQ{this should be equivalent to the coq version, and we should adopt it.}
+%    \leo{We had this, but it was much harder to work with for my proofs. We should still
+%      prove them equivalent if we want to present it this way (which is probably cleaner for
+%      presentation purposes)}
+
+Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
+We then define {\em policy-sensitive observational equivalence}: a
+machine-policy-observation\leo{Call these MPO traces?} is equivalent
+to a machine-observation trace, if the MPO trace is infinite and the
+observations of MPO and M coincide, or if the MPO trace is finite
+(e.g. due to a policy fault) and its observations are a prefix of
+those of M.
+
+\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
+\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
+           {\MPT \sim_P \machT}\]
+
+\[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
+           {\MPT \sim_P \machT}\]
+
+\section{Stack Safety, Formally}
+
+In this section we will describe stack safety as a formal
+property.
+\leo{Should call maps and returns go here or in preliminaries?}
+
+\paragraph*{Call Maps and Returns}
+
+A {\it call map} $\callmap$ identifies machine states that represent a
+call, abstracting away from the calling convention of any concrete
+machine. Formally $\callmap$ is a partial map from machine states to a
+natural number, the number of arguments of that call.
+%
+\[\callmap \in \CALLMAPS = \MACHS \rightharpoonup \mathbb{N}\]
+
+For concreteness, in a typical calling convention, a call map may be
+considered to identify the state just before the \(\JAL\), in which
+space for arguments is already allocated, but the mechanism is more
+general.
+
+A state on which a call map is defined is referred to as a {\it call
+  state}, and we consider the call to have returned the first time
+control returns to the instruction following the \(\JAL\) with the
+stack restored. Formally this is a relation between the states:
+
+\[\begin{split}
+\ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) = \nextPC(\mach_c(\PCname)) \land \\
+& \mach_r(\SP) = \mach_c(\SP)
+\end{split}\]
+
+\leo{How do we initialize the first policy state based on a call map? It's not
+  part of the policy. Do we parameterize policies?}
+
+\paragraph*{Stack safety}
+
+A stack safety property is a property of a transition system, a
+policy, and a callmap that is decomposed into an integrity property and a
+confidentiality property.
+
+    \[\begin{split}
+      \textit{stack safety} \triangleq \textit{stack integrity} \land \textit{stack confidentiality} \\
+    \end{split}\]
+
+\leo{TODO: Rephrase from introduction} Informally, {\it Stack
+  integrity} states that if a function call returns, the caller's
+private stack data will be identical to what it was before the
+call. {\it Stack confidentiality} states that data outside a callee's
+stack frame is secret, and do not interfere with its observable
+behavior or with the machine state at return.
+
+As is standard with security properties, we associate with each
+component of the machine a {\em label}: a pair of security levels
+denoting high or low confidentiality and high or low integrity.
+%
+\[\mathit{label} ::= \{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT,\LOWINT\}\]
+%
+Since we need to enforce integrity and confidentiality at {\em every}
+function call in a trace, we must treat the security level of data
+dynamically. To that end, we introduce the notion of a {\em contour},
+a map from components to labels:
+%
+\[\contour \in \CONTOURS ::= \COMPONENTS \rightarrow \mathit{label}\]
+
+\leo{What about instructions? Shouldn't they be LC/HI?}
+When \(\mpstate{\mach}{\pol}\) is a call state \(\callmap(\mach) = n\)
+for some \(n\), we can construct a contour based on the bounds of the
+callee's stack frame, as follows:
+\leo{Should the first two cases have a $k \not \in \REGS$?}
+\[\mathit{Cof}(\mach,n)(\component) =
+\begin{cases}
+  (\HIGHSEC,\HIGHINT) & \text{if } \component \leq \mach(\SP) - n \\
+  (\HIGHSEC,\LOWINT) & \text{if } \component \geq \mach(\SP) \\
+  (\LOWSEC,\LOWINT) & \text{else} \\
+\end{cases}\]
+%
+The private state of the callee is the entire area of the memory below
+the stack pointer, with the exception of the $n$ locations
+corresponding to the arguments of a call. These components are tagged
+high confidentiality and high integrity: the callee should be able to
+neither read, nor write them.
+%
+Everything above the stack pointer at the point of the call is
+``uninitialized'' memory that can be written into but should not be
+read. To protect the callee itself from such reads, these components
+are marked high confidentiality and low integrity.
+%
+Registers and the frame of arguments to the callee are free for the
+callee to use, and are therefore tagged with low integrity and low
+confidentiality.
+
+\leo{Shouldn't this be a definition instead of a paragraph?}
+%\paragraph{Integrity}
+
+\definition{Stack Integrity}
+
+A system satisfies integrity with respect to a callmap \(\callmap\) if
+for any initial state \(\mach_0\), call state \(\mach_c\) where
+\(\callmap(\mach_c) = n\), and policy state \(\pol_c\), and return
+state \(\mpstate{\mach_r,\pol_r}\), where: \leo{words: induced trace, call in the trace, take the prefix}
+%
+\[\mach_0,\mathit{pinit}(\mach_0) \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ... \mpstate{\mach_r}{\pol_r,\_} | \ret{\mach_c}\]
+%
+it is always the case that: \leo{words here too: the value of confidential components is unchanged}
+\[\forall \component . \contour(\component) = \HIGHINT \rightarrow \mach_c(\component) = \mach_r(\component)\]
+
+\definition{Stack Confidentiality}
+
+A system enjoys eager stack confidentiality with respect to a callmap
+\(\callmap\) if, for any initial state \(\mach_0\) and call state
+\(\mach_c\) such that
+\(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}
+\hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...\) and
+\(\callmap(\mach_c) = n\):
 
       \begin{itemize}
-\item machine model
-\item Calling conventions
+        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
+        \item For all \(\mach_c'\) such that \(\mach_c \approx_\contour \mach_c'\):
+        \item If \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT (\mach_r,p_r,\_) | \ret{\mach_c}\)
+          \begin{itemize}
+            \item Then there is some \(\mach_r'\) where
+
+              \(\mach_c' \hookrightarrow \machT' (\mach_r',_) | \ret{\mach_c'}\)
+            \item \(\mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT')\)
+            \item And for all \(\component\), if \(\mach_c(\component) \not = \mach_r(\component)\) or
+              \(\mach_c'(\component) \not = \mach_r'(\component)\), \(\mach_r(\component) = \mach_r'(\component)\)
           \end{itemize}
-
-\section{Related Work}
-
-\iftext
-\section{Introduction}
+        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
+          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
+      \end{itemize}
+
+      \leo{Think more about whether we vary mc or step mc}
+      \leo{Also, words}
+
+\section{Lazy Stack Safety}
+
+    \paragraph{Observable Integrity}
+
+      Observable integrity is concerned with the changes a callee makes to high integrity components,
+      and the effects those writes have on the observable behavior of its caller
+      {\it after its return}. Given a contour, a call state, and a returned state, we define
+      a ``rollback'' function \(\mathit{roll}_I : \CONTOURS \times \MACHS \times \MACHS
+      \rightarrow \MACHS\). The rollback creates a state that matches the call
+      state on those components that are high integrity in its contour, and matches the return state
+      on low integrity components:
+
+      \[\mathit{roll}_I(\contour,\mach_c,\mach_r)(\component) =
+      \begin{cases}
+        \mach_c(\component) & \contour(\component) = (\HIGHINT,\_) \\
+        \mach_r(\component) & \contour(\component) = (\LOWINT,\_) \\
+      \end{cases}\]
+
+      Intuitively, the result of a rollback is an idealized state that reflects changes to accessible
+      components, but ignores the callee's interference with the caller's data. Then the observable
+      behavior of the resulting trace is the standard to which we expect any real trace to conform.
+
+      A system enjoys observable integrity with regard to \(\callmap\) as follows:
+
+      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\)
+      and \(\callmap(\mach_c) = n\), if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow ...
+        \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\):
 
           \begin{itemize}
-\item \leo{Not even sure this is true anymore}
-\item Noninterference is a security property combining {\em
-  confidentiality} and {\em integrity}: secret data should influence
-  public outputs and low integrity data should \leo{finish in a pretty
-    way}.
-\item In this paper we present a stack safety property as a variation
-  of noninterference.
+            \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
+            \item Let \(\mach_r' = \mathit{roll}_I(\contour,\mach_c,\mach_r)\), \(\mpstate{\mach_r}{\pol_r}
+              \hookrightarrow \MPT\), and \(\mach_r' \hookrightarrow \machT'\)
+            \item \(\MPT \sim_P \machT'\)
           \end{itemize}
 
-{\em Hyperproperties}, i.e. properties of pairs\leo{sets?} of
-execution traces\footnote{\bcp{I'm not certain we're using the word
-    correctly.  E.g., ``A hyperproperty is a set of properties [Clarkson and
-    Schneider 2008, 2010]''} \bcp{After more reading, I am still wondering
-    if we're using the word the right way.  I think we mean something much
-    less general---just a relation on traces.  So I think we should just say
-    ``contoured relation'' (and maybe ``lockstep relation'').}}, are the
-standard way of expressing security
-properties\cn. In particular, {\em noninterference} is commonly
-phrased as such a property: given two system states that seem
-identical to an observer with low clearance, their execution traces
-should also be indistinguishable\cn.
-
-In this paper we focus on a well-behaved closed\bcp{???} subset of
-hyperproperties, where the execution traces remain in {\em lockstep}:
-the traces execute identical sequences of instructions.
-%
-This allows us to give a clean extensional characterization of taint
-tracking as a noninterference property: intuitively, taint tracking is
-a weak lockstep-hyperproperty that guarantees noninterference as long
-as the execution traces remain in sync, but then allows for arbitrary
-behavior after branching on a secret.
-%
-We also describe a strong version of lockstep hyperproperties, where
-the security policy enforces the lockstepness. \leo{Something about
-  full separation}
-%
-Finally, we generalize strong lockstep hyperproperties to allow for
-\leo{what's the correct adjective here? principled? nested?
-  restricted?}  changes to the observer clearance, and use this
-generalization to characterize a novel stack safety property as
-a variation of noninterference.
-
-As it happens, lockstep hyperproperties are also significantly more
-amenable to testing compared to traditional hyperproperties. Since
-executions remain in lockstep, there is no possibility of \leo{missing
-  bugs by never coming back down or a need to find an invariant strong
-  enough to prove things}. We have implemented enforcement mechanisms
-for all three policies for RISC-V and tested them.
-
-Our contributions are:
+    \paragraph{Observable Confidentiality}
+
+      Observable confidentiality is concerned with whether high confidentiality data influences
+      observable both behavior during and after a call. We define variants given the generalized notion
+      of contours:
+      \[\mach \approx_\contour \mach' \triangleq \forall \component .
+      \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
+      \leo{Move earlier}
+
+      Data that is high confidentiality within a callee due either to being in its caller's frame
+      will become low confidentiality if the callee returns.\leo{??} We face a subtle distinction: 
+      a callee treats the contents of its caller's frame as secrets, and they should not influence
+      its behavior. Nor should the callee be able to influence behavior after its return based
+      on the caller's secrets, for instance by copying them to registers that will later be output.
+      But the caller itself may still adjust its behavior based on its own secrets. So our property
+      relies on another rollback function to preserve changes made by the caller but remove other
+      differences between variant states.
+
+      In this case, a confidentiality rollback takes a contour and four states: two call states
+      that will be variants, and two return states. It is not symmetric, regarding \(\mach_c\) and \(\mach_r\)
+      as the ``real'' call and return, \(\mach_c'\) as a variant and \(\mach_r'\) as a return that follows it.
+      \[\begin{split}
+        & \mathit{roll}_C(\contour,\mach_c,\mach_c',\mach_r,\mach_r')(\component) = \\
+        & \begin{cases}
+          \mach_c(\component) & \text{if } C(\component) = \HIGHSEC \text{ and } \\
+                  & \mach_c(\component) = \mach_r(\component) \text{ and }
+                    \mach_c'(\component) = \mach_r'(\component) \\
+          \mach_r'(\component) & \text{else} \\
+        \end{cases}
+      \end{split}\]
+      \leo{Maybe this should be called unscramble}
+
+      A system enjoys observable confidentiality with respect to \(\callmap\) as follows:
+
+      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\) and
+      \(\callmap(\mach_c) = n\):
+
       \begin{itemize}
-\item We present a mathematical abstraction of a class of properties
-  called {\em lockstep hyperproperties}, that allow for intuitive
-  characterizations of interesting security policies.
-\item We give a novel {\em extensional} characterization of taint tracking
-  as a form of lockstep-noninterference.
-\item We introduce a novel stack safety property that allows for
-  the notion of observability to change dynamically throughout
-  execution.
-\item We develop a testing framework for testing such
-  lockstep-hyperproperties, implemented as a parametric co-processor
-  for RISC-V.
+        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
+        \item For all \(\mach_c'\) such that \(\mach_c \approx_C \mach_c'\):
+        \item If \(\mpstate{\mach_c}{p_c} \hookrightarrow \MPT_1 \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\)
+          \begin{itemize}
+            \item Then there is some \(\mach_r'\) where
+
+              \(\mach_c' \hookrightarrow \machT_1' (\mach_r',\_) | \ret{\mach_c'}\)
+            \item Let \(\mach_r'' = \mathit{roll}_C(\contour,\mach_c,\mach_r,\mach_c',\mach_r')\),
+
+              \(\mpstate{\mach_r}{p_r} \hookrightarrow \MPT_2\), and \(\mach_r'' \hookrightarrow \machT_2'\)
+            \item \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1' \cdot \machT_2'\)
           \end{itemize}
-
-\fi
-
-\section{Preliminaries}
-
-\subsection{Machine model}
-
-Concretely, we have in mind a conventional machine architecture with a
-standard (RISC-V) processor extended with some kind of (static or dynamic,
-software- or hardware-enforced) security monitor, but we won't need to make
-any specific choices about details until we get to \bcp{the implementation
-  section}.
-% \cref{impl}.
-
-Conventions:
-%
-\begin{itemize}
-\item Words / addresses: $\word,\addr \in \WORDS$
-\item Registers: $\reg \in \REGS$
-\item Machine components: $\component \in \WORDS \cup \REGS \cup \{
-\textsc{pc} \}$
-\item Machine states: $\mach \in \MACHS = \COMPONENTS\rightarrow\WORDS$
+        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
+          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
       \end{itemize}
 
-% \bcp{Now we need to define {\em property} formally.  But since we're talking
-% about hyperproperties, I'm not certain how to go about it.  The standard way
-% would be something like ``A policy $P$ has property $Q$ (i.e., $P$ is
-% noninterfering) if, for every pair of traces $\trace{M}$ and $\trace{M'}$
-% in $P \times P$, if the initial states $M_0$ and $M'_0$ differ only in
-% high parts then every pair ...'' (blah: I'm not sure whether to talk about
-% lockstep-corresponding states, which seems too restrictive, or final states,
-% which doesn't work if the traces are infinite!  do we really need some kind
-% of reactive noninterference here?? Urg.)  (Well, no, we can just do
-% a flavor of termination-insensitive NI, which says that all infinite traces
-% are indistinguishable from anything...?)}
-
-% \bcp{Should we start from a notion of ``variant traces'' and then talk about
-% pairs of variant traces being ``distinguishable'' and then go to the full
-% thing?}
-
-The {\em underlying} step function, $\mathord{\stepstounderfull} \in
-\MACHS \rightarrow \MACHS$, embodies the standard RISC-V ISA
-semantics.
-%\bcp{It's probably clearer not to put RISCV in the
-%  notation---for most of the paper, the fact that it's a RISCV really
-%  doesn't matter.  Moreover, I think we want to make it a partial
-%  function, for simplicity of definitions in this part, and if we
-%  insist on RISCV here this will confuse people that know it doesn't
-%  have a HALT instruction.}
-
-A {\em trace} is a finite or infinite sequence of machine states,
-$\seq{\mach} = \mach_0,\, \mach_1,\ldots$, in which adjacent machine
-states are related by $\stepstounderfull$---i.e., in which
-$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the
-sequence. We write $\seq{\mach}^i_j$ for the subtrace between $i$ and
-$j$.
-
-\subsection{Calling Convention}
-
-\leo{Mainly @Andrew, check if this makes sense for the eager setting.}
-
-We adopt the following calling convention:
-\begin{itemize}
-\item First the caller increases the stack pointer to allocate stack space
-  for arguments and result and zeroes out anything not written.
-\item Then a $\JAL$ happens.
-\item Then the entry sequence of the callee follows which increases
-  the stack pointer again to allocate as much local space as needed
-  (e.g. for the return address), and it is all zeroed out if not
-  explicitly written. \leo{This can be lifted by dealing with
-    stack allocation in the contours}
-\end{itemize}
-
-A {\em call map} is a map from a sequence of instructions (read: {\em
-  a call sequence}) to a natural number (the number of arguments and
-results of the call plus the size of the local state of the callee,
-which corresponds to stack locations that will be made available to
-the callee).
-
-\subsection{Contours}
-
-A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
-\{\HIGHSEC, \LOWSEC\} \times \{\HIGHINT, \LOWINT\}$ is a map from
-machine components to a pair of confidentiality and integrity labels.
-$\LOWSEC$ marks locations whose contents represent data observable
-{\em by the current process}; $\HIGHSEC$ corresponds to non-observable
-data (e.g. local state of the process caller). $\HIGHINT$ marks
-locations that the current process does not have write access to;
-$\LOWINT$ is writeable. As examples of different
-combinations consider the following:
-\begin{itemize}
-\item Locations that belong to the current process such as
-  available registers, local stack, memory accessible through
-  available pointers will be tagged $\LOWSEC \times \LOWINT$:
-  they can be both read and written.
-\item The local state of the caller that needs to be protected from a
-  potentially untrusted callee will be marked $\HIGHSEC \times
-  \HIGHINT$. The callee should be able to neither read nor write
-  these locations.
-\item The instruction memory is read only for every process: $\LOWSEC
-  \times \HIGHINT$.
-\item Uninitialized memory can be written, but never read. This is only
-  applicable in a more lazy setting where we don't zero out memory.
-\end{itemize}
-
-\subsection{Identity colorings}
-\leo{Feel free to coin another name}
-
-\leo{Went with S for ``self'' and O for ``other''. M/NM was confusing
-  with the machine states also being M. T/F didn't really work
-  with transitions. Open to other suggestions.}
-An {\em identity coloring} of a trace $\seq{\mach}$ is an equal-length
-sequence $\seq{\ID}$ of ids ($\IDS = \{\ME, \NOTME, \TRANS\}$, where
-$\ID_j = \ME$ if the machine state $\mach_j$ is executing the same
-process as $\mach_0$, $ID_j = \NOTME$ if it is executing a different
-process, and $ID_j = \TRANS$ during a transition (e.g. during a call
-or return sequence).
-
-Given a machine trace $\seq{\mach}$ and a call map the identity coloring can be
-computed by iterating through the trace:
-\begin{enumerate}
-\item Start by coloring $\mach_0$ as $\ME$.
-\item Color every subsequent state as $\ME$, until the first machine
-  state (if any) at which the instruction executed is a
-  the beginning of a call sequence.
-\item Color the entire call sequence $\TRANS$. This necessarily
-  contains a machine state $\mach_c$ which is a $\JAL$.
-\item Color every subsequent state as $\NOTME$, until the first state
-  $\mach_r$ after $\mach_c$ in the trace (if any) such that the pc of
-  $\PC{\mach_r}$ is equal to $\PC{\mach_c} + 4$, and whose stack
-  pointer is identical to that of $\mach_c$. Then mark $\mach_r$ as
-  $\ME$ and go to step (2).
-\end{enumerate}
-
-\subsection{Contoured Trace}
-
-Given a contour $C$ and a machine state $M$, the {\em contoured trace}
-induced by $C$ and $M$, written $\induced{C}{M}$, is the (finite or
-infinite) sequence of pairs of (contour, machine state) pairs in which
-states in adjacent pairs are related by $\stepstounderfull$, and in which
-contours in adjacent pairs are constructed as follows:
-\begin{itemize}
-\item If in the identity coloring of $\seq{\mach}$ the identity of
-  consecutive states is preserved ($\ID_i$ = $\ID_{i+1}$) then
-  $C_{i+1} = C_i$.
-\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \TRANS$, then $C_{i+1}$ is
-  constructed from $C_i$ by setting the local state of $\mach_i$ (the
-  caller) to $\HIGHSEC \times \HIGHINT$. The local state is comprised
-  of everything below the stack pointer at $\mach_i$.
-\item If $\ID_{i} = \TRANS$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
-  constructed from $C_i$ by making the top $n$ locations in the stack,
-  where $n$ is the mapping of the transition call sequence in the call map,
-  readable and writeable $\LOWSEC \times \LOWINT$.
-\item If $\ID_{i} = \TRANS$ \apt {$\NOTME$ ? } and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
-  constructed from $C_i$ by setting the local state of the caller
-  to $\LOWSEC \times \LOWINT$, and making everything above the stack
-  pointer $\HIGHSEC \times \HIGHINT$ again.
-\end{itemize}
-
-\subsection{Integrity}
-
-A trace $\seq{\mach}$ satisfies the {\em integrity} safety property
-with respect to a contour $\contour$, if for all components
-$\component$ where $\contour(\component) = (\_, \HIGHINT)$,
-$\mach_i(\component) = \mach_0(\component)$ for all $i$ in the trace.
-
-\leo{For confidentiality we need observations first}
-
-\subsection{Observations}
-
-A {\em component observation} is either a word or the special value
-$\bot$.  \leo{Fix notation, clashes with $\LOWINT$}  A {\em state
-  observation} is a map from machine components to component
-observations.
-
-An {\em observation} is a nonempty, finite or infinite sequence of state
-observations in which (1) adjacent elements are always distinct (``no
-stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
-the the $j^{th}$ state observation in the sequence must also be mapped to
-$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
-component's value changes from $\bot$ to anything else, it never changes
-back to $\bot$ again).
-
-In particular, the observation $\obstrace(\contour, \seq{\mach})$
-arising from a contoured trace $\seq{(\contour, \mach)}$
-is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
-\ldots$ defined informally as follows:
-\begin{itemize}
-\item $\stateobs_0$ maps every visible component $\component$ (where
-  $\contour_0(\component) \in \{\LOWSEC \times \_\}$) to its value in
-  $\mach_0$ and every hidden component to $\bot$;
-\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
-  each visible component $\component$ (where
-  $\contour_{i+1}(\component) \in \{\LOWSEC \times \_ \}$) its value
-  from $\mach_{i+1}$, and (2) for each hidden component $\component$,
-  where $\contour_{i+1} (\component) = \{\HIGHSEC \times \_\}$,
-  copying its value from $\stateobs_{i}$, and (3) if the resulting
-  $\stateobs_{i+1}$ is identical to $\stateobs_{i}$, then skipping
-  $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
-\end{itemize}
-\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
-
-\subsection{Confidentiality}
-
-Given a machine state $\mach$ and a contour $\contour$, a {\em
-  variant} of $\mach$ with respect to $\contour$ is any machine state
-$\mach'$ whose components agree with $\mach$ on all components marked
-$(\LOWSEC, \_)$ by $\contour$---i.e., if $\contour(\component) =
-(\LOWSEC, \_)$, then $\mach(\component) = \mach'(\component)$.
-
-A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
-  confidentiality} safety hyperproperty with respect to a contour
-$\contour$, if, if the initials states are variants of each other $N_0
-\in \variant{C}{M_0}$, then the observations of one is a prefix of the
-observations of the other: $\obstrace(\seq{\contour, \mach}) \subseteq
-\obstrace(\seq{\contour, N})$.
-
-\apt{Does this differ from lockstep equivalence? And why prefix?}
-
-\subsection{Stack Safety}
-
-A trace $\seq{\mach}$ satisfies {\em stack safety} with respect to a
-contour $\contour$ if:
-\begin{itemize}
-\item $\seq{\mach}$ satisfies integrity with respect to $\contour$.
-\item For every variant $N \in \variant{C}{\mach_0}$, $\seq{\mach}$
-  and the the induced trace $\seq{N}$ satisfy confidentiality with
-  respect to $\contour$.
-\item Every subtrace $\seq{\mach}^r_c$ colored $\NOTME$ in the
-  identity coloring of $\seq{\mach}$ satisfies stack safety with
-  respect to the contour $\contour_c$.
-\end{itemize}
-
-\bcp{Needs to be formulated coinductively?!}
-
-\bcp{What about unitialized memory (written by previous calls)}
-
-\section{Related work}
-
-\bcp{Make sure to mention that we are NOT enforcing well-bracketed control
-  flow a la Birkedal et al!}
-
-\section{Future Work}
-
-Is there a knowledge-based presentation of all this?
-
-\cleardoublepage
-\leo{OLD STUFF BEGINS HERE}
-
-\subsection{Contours}
-
-\iftext
-Intuitively, we use the word {\em policy} for any kind of dynamic
-enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
-capability hardware, etc.).
-We use {\em property} for a mathematical
-condition on programs (i.e., machine states) articulating some useful
-security goal, e.g. noninterference.
-\fi
-
-A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\ACCYES,
-\ACCNO, \ACCCODE\}$ is a map from machine components to accessibility levels.
-$\ACCYES$ marks locations whose contents represent data observable from the
-outside (by the ``attacker,'' e.g., by DMA), which moreover correspond to
-low-integrity components (e.g., locations that can be written to \iftext
-\bcp{directly or indirectly} \fi by the network). $\ACCNO$ marks internal
-``hidden'' state that cannot be observed directly (e.g., processor registers and
-memory addresses for temporary variables will usually be hidden), which is
-moreover high-integrity (i.e., locations which should be protected from flows
-from the network)
-%
-\rb{Tried to merge discussion on integrity and secrecy, but maybe it is clearer
-  to separate both concerns?}
-%
-In addition to those two accessibility levels, $\ACCCODE$ marks locations that
-hold executable instructions, which are considered to be low-secrecy (i.e.,
-observable by the attacker) and high-integrity (i.e., the attacker cannot write
-to them).
-
-\rb{OLD:
-\leo{We might need to change the intuition behind some of these things.}
-A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
-\{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT, \LOWINT\}$\bcp{We need more mnemonic names for
-  these (that help us keep straight when we're talking about confidentiality
-  and when integrity) in the way they are displayed.}
-is a map from machine components
-to confidentiality and integrity levels. $\LOWSEC$ marks locations whose
-contents are observable from the outside (by the ``attacker,'' e.g.,
-by DMA) and $\HIGHSEC$ marks internal ``hidden'' state that cannot be
-observed directly (e.g., processor registers and memory addresses for
-temporary variables will usually be hidden). Similarly, $\LOWINT$
-locations correspond to low-integrity components (e.g. locations that
-can be written to \iftext \bcp{directly or indirectly} \fi by the network), while $\HIGHINT$ locations should be
-protected from such flows.
-}
-
-\rb{The flows-to relation was quite nice before, now it looks like it will get
-  messier\ldots}
-%
-We denote this (reflexive) flows-to relation with $\flows$. Then
-$\LOWSEC \flowsstrict \HIGHSEC$ and $\HIGHINT \flowsstrict
-\LOWINT$. We can easily lift this relation to pairs of secrecy and
-integrity labels, and from there to contours pointwise.
-
-\subsection{Observations}
-
-\rb{We are not using observations at the moment, but a few adjustments are made
-  here.}
-
-A {\em component observation} is either a word or the special value $\bot$.
-A {\em state observation} is a map from machine components to component
-observations.
-
-An {\em observation} is a nonempty, finite or infinite sequence of state
-observations in which (1) adjacent elements are always distinct (``no
-stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
-the the $j^{th}$ state observation in the sequence must also be mapped to
-$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
-component's value changes from $\bot$ to anything else, it never changes
-back to $\bot$ again).
-
-In particular, the observation $\obstrace(\contour, \seq{\mach})$
-arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
-is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
-\ldots$ defined informally as follows:
-\begin{itemize}
-\item $\stateobs_0$ maps every visible component $\component$ (where
-  $\contour(\component) \in \{\ACCYES, \ACCCODE\}$) to its value in $\mach_0$
-  and every hidden component to $\bot$;
-\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
-  each visible component $\component$ (where $\contour(\component) \in
-  \{\ACCYES, \ACCCODE\}$), copying its value from $\mach_{i+1}$, and (2) for
-  each hidden component $\component$, where $\contour (\component) =
-  \ACCNO$, copying its value from $\stateobs_{i}$, and (3) if
-  the resulting $\stateobs_{i+1}$ is identical to $\stateobs_{i}$,
-  then skipping $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
-\end{itemize}
-
-% For each component $\component$,
-% %
-% if
-% $\contour_i(\component) = (\underscore,\HID)$ for all $i$, then
-% $\obstrace(\component) = [\,]$;
-% %
-% otherwise, $\obstrace(\component)$ is a sequence whose first element
-% is $\mach_i(\component)$ for the smallest $i$ such that
-% $\contour_i(\component) = (\underscore,\VIS)$ and whose $(m+1)^{th}$ element
-% (if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
-% than the index $i$ of the $m^{th}$ element and such that
-% $\contour_j(\component) = (\underscore,\VIS)$ and
-% $\mach_j(\component) \neq \mach_i(\component)$.
-% %
-% \bcp{Can be defined more simply}
-% \iftext
-% \bcp{Unfinished and horrible...  But maybe the textual description is OK?}
-% \newcommand{\aux}{\mathit{aux}}
-% \bcp{Don't like the ``rest'' argument and the fact that we don't
-%   have a metavariable for sequences of (contour, machine state) pairs.
-%   Consider defining one.}
-% %
-% \[
-% \begin{array}{lcl}
-% \aux \; ( [\,] ) \; \trace& = & \trace
-% \\
-% \aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
-%   & = &
-%         \aux \;( \mathit{rest} )
-%         \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
-% \\
-% \\
-% \obstrace \;( (\contour_0,\mach_0),\, \ldots )
-%   &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
-%       \component.\, \bot)
-% \end{array}
-% \]
-% \fi
-
-% \bcp{Maybe make it a sequence of maps from components to values?  This is
-%   simpler, but it means we can ``observe the observbility of locations...''
-%   But is this too unrealistic?}
-
-\iftext\bcp{Emphasize that ``observability is not observable.''}\fi
-
-\subsection{Properties}
-
-% A {\em contoured (hyper)property} is a function that takes a contour to a
-% hyperproperty.  That is, it takes a starting contour and produces a
-% \bcp{prefix-closed?} set of pairs of sequences of machine states.
-%
-% An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
-% (X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
-% \times X) $
-% comprises
-% \begin{enumerate*}
-%   \item a private type $X$,
-%   \item a function taking a contour to an initial value of $X$, and
-%   \item a function that allows ``reading off'' a contour from an $X$, and
-%   \item a partial step function of type
-%   $(\MACHS \times X) \rightharpoonup X$.
-% \end{enumerate*}
-
-Let $(\stepplus) :: (\CONTOURS \times \MACHS) \rightarrow (\CONTOURS
-\times \MACHS)$ be an extension of the underlying step function, such
-that $(C,M) \stepplus (C',M')$ implies that $M \stepstounderfull M'$.
-Given a contour $C$ and a machine state $M$, the {\em contoured trace}
-induced by $C$ and $M$, written $\induced{C}{M}$ is the largest
-(finite or infinite) sequence of pairs of (contour, machine state)
-pairs, in which adjacent pairs are related by $\stepstounderfull$ and
-for which $C \flows C_i$ for all $i > 0$ in the sequence.
-\leo{That is, the induced contoured trace by a (C,M) pair incorporates
-  the ``return'' snipping}
-
-Given a machine state $\mach$ and a contour $\contour$, a {\em
-  variant} of $\mach$ with respect to $\contour$ is any machine state
-$\mach'$ whose components agree with $\mach$ on all components marked
-$\ACCYES$ or $\ACCCODE$ by $\contour$---i.e., if $\contour(\component) \in \{\ACCYES, \ACCCODE\}$, then
-$\mach(\component) = \mach'(\component)$.
-%
-\rb{When we talk about observations, it is visibility that we are concerned with
-  and moving from $\HIGHSEC$ to $\ACCNO$ and $\LOWSEC$ to $\ACCYES$ and
-  $\ACCCODE$ is simple. Here, the case of code is slightly more
-  interesting (are $\ACCCODE$-variants ever useful?).}
-
-Let $\initmach$ be the set of initial states of the machine and let
-$\initcontour(M)$ be the\bcp{an?} initial contour assigned to each $M \in
-\initmach$. \iftext\leo{This could probably be generalized to a set}\fi
-
-For each $M \in \initmach$ and $C = \initcontour(M)$, we define the
-{\em closure} of $(C,M)$ as the set of triples of (contour, trace,
-trace) as follows:
-\[
-\begin{array}{rcl}
-  \closure{(C,M)} = \{ ( C_i, \seq{N}, \seq{N'})
-  & | & C_{i-1} \not = C_i \\
-  & , & \seq{(C_i, M_i)} = \induced{C}{M}\\
-  & , & \seq{(\_, N_j)} = \induced{C_i}{M_i}\\
-  & , & N'_0 \in \variant{C_i}{M_i}\\
-  & , & \seq{(\_, N'_j)} = \induced{C_i}{N_0'} \}\\
-\end{array}
-\]
-\leo{Alternative: We could do this for all $i$, but lose the
-  trivial instantiation of noninterference later on}
-\bcp{Don't follow the notation here.  :-(}
-
-A trace $\seq{M}$ satisfies the {\em integrity} safety property with
-respect to a contour $\contour$, if for all components $\component$
-where $\contour(\component) = (\_, \HIGHINT)$, $M_i(K) = M_0(K)$ for
-all $i$ in the trace.
-%
-\rb{Here, $\HIGHINT$ would be replaced by $\in \{\ACCNO, \ACCCODE\}$.}
-
-A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
-  confidentiality} safety hyperproperty with respect to a contour
-$\contour$, if, if the initials states are variants $N_0 \in
-\variant{C}{M_0}$, then their observations coincide:
-$\obstrace(\contour, \seq{M}) = \obstrace(\contour, \seq{N})$.
-\leo{Should we cut at minimum length for termination-insensitiveness?}
-\bcp{Or just ask that one be a prefix of the other?}
-\bcp{Or go back to halting states.  (Which I guess really means go back to
-  Halting + Failing states.  We never quite worked out how that would go.)}
-
-\leo{From now on I need better terminology, but hopefully you get the
-  point} An initial machine state \leo{or contour machine state pair}
-is \leo{good} if for every triple $(C, \seq{N}, \seq{N'})$ in
-$\closure{(\initcontour(M),M)}$ we have:
-\begin{itemize}
-\item The traces $\seq{N}$ and $\seq{N'}$ satisfy confidentiality
-  with respect to $C$.
-\item The traces $\seq{N}$ and $\seq{N'}$ each satisfy integrity
-  with respect to $C$.
-\end{itemize}
-
-\leo{Combined result}
-Noninterference holds for the set of \leo{good} initial machine
-states.
-
-\leo{NOTE:
-  \begin{itemize}
-    \item When the contours don't change this is trivially
-      noninterference: only a variation of the initial state is
-      added to the trace.
-    \item This almost directly coincides with the LSE part
-      of the stack-safety property.
-  \end{itemize}
-}
-
-%A {\em contoured hyperproperty} is a set of triples,
-%each consisting of a contour along with a pair of traces.
-%
-%\leo{No longer needed:?}
-%A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
-%such that each trace in every pair of traces in the set executes the
-%same instruction sequence.
-%
-%Finally, we can parameterize the traces by some initial application-specific
-%state to obtain a {\em contoured lockstep $X$-hyperproperty}, where every
-%trace comes with more information, for example call sites in stack safety.
-%\bcp{If we use this later, we need to say it more formally to be clear about
-%  what we mean---e.g., I had thought both the traces and the contour would
-%  be functions of $X$!}
-
-\rb{New try. Contours are as above.}
-
-A {\em call/return map} $\kappa \in \WORDS \rightarrow \{\CRCALL, \CRRET,
-\CRBOT\}$ is a map from addresses to tags (not to be confused with tags at the
-micropolicy level \rb{reference if needed}) that denote whether a memory
-location is to be treated as (the first word of) a call instruction, a return
-instruction, or neither \rb{(and we should compare these with call and return
-machine instructions ``proper'')}.
-
-The {\em step relation} $(\stepkappa) :: \MACHS \rightharpoonup \MACHS$ is a
-partial function from machine states to machine states: a restriction of the
-native machine's step relation $\stepstounderfull$ indexed by a call/return
-map $\kappa$, such that \rb{(How much to build into this, and/or the machine
-states themselves? Also describe and notate the concept of ``running until
-the first unmatched return (cf. observable halting states)'') \ldots}
-
-Given a machine state $\mach$ and associated contour $\contour$, an
-$\ACCNO$-variant is any state $\mach'$ that varies machine components tagged as
-$\ACCNO$ in $\mach$: that is, for any component $\component$ of $\mach$, if
-$\contour(\component) \neq \ACCNO$, then
-$\mach(\component) = \mach'(\component)$; otherwise (if
-$\contour(\component) = \ACCNO$), then $\mach'(\component)$ holds an
-arbitrary value, unrelated to $\mach(\component)$. Similarly, $\ACCYES$-variant
-states are defined by scrambling those machine components tagged as $\ACCYES$
-and leaving all others unchanged.
-
-A machine state $\mach$ is {\em safe} with respect to a contour $\contour$
-\rb{and call/return map $\kappa$} if it satisfies the following conditions:
-%
-\begin{enumerate}
-
-\item {\em Integrity:} if we run $\mach$ until its first unmatched return (if
-  any), all components other than those marked as accessible must remain
-  unchanged with respect to $\mach$. \rb{That is, the final state is an
-  $\ACCYES$-variant of $\mach$.}
-
-\item {\em Confidentiality:} for every $\ACCNO$-variant $\mach'$ of $\mach$, if
-  we run $\mach$ and $\mach'$ until their first unmatched returns, the resulting
-  final states are themselves $\ACCNO$-variant of one another. \rb{(Homogenize
-  notations, see also above.)}
-
-\item Let $\mach_1, \mach_2, \ldots$ be a machine execution starting from
-  $\mach$, i.e., $\mach_1 = \mach$. Let also $\contour_1, \contour_2, \ldots$
-  be the corresponding sequence of contours, such that $\contour_1 = \contour$
-  and at each step $\mach_i$ the corresponding $\contour_i$ \rb{\ldots better
-  define the contour-updating step in the style of $\stepplus$.} For all $i$,
-  if $M_i$ is a top-level call in the trace, then (the instruction immediately
-  after) it \rb{$\mach_{i+1}$?} is safe with respect to $C_i$.
-
-\item All instructions executed throughout any run starting from $\mach$ are
-  marked as such in the contour $\contour$. \rb{Need notation to dereference the
-  PC and have the memory be marked as $\ACCCODE$:
-  $\contour_i(\PC{\mach_i}) = \ACCCODE$ (or noting invariance in $\ACCCODE$
-  parts of $\contour$).}
-
-\end{enumerate}
-
-The step relation $\stepkappa$ is itself safe if every machine state $\mach$ is
-safe with respect to the special contour where all memory locations containing
-code are marked $\ACCCODE$ and all other components are marked $\ACCYES$.
-
-\subsection{Noninterference}
-
-\iftext
-\leo{Indistinguishability/variations:}
-
-\leo{Baseline noninterference:}
-
-\leo{Is this noninterference or is it noninterference + integrity.
-  I'm confused about terminology}
-\fi
-
-Our baseline security property is noninterference.
-\begin{itemize}
-\item {\em Confidentiality:}
-
-  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$, if
-  the initial elements of the traces are variants, then the final
-  states of the traces are also variants.  \bcp{What does ``final''mean for
-    infinite traces?  What about pairs where one is finite and the other
-    infinite?}
-
-\item {\em Integrity: }
-
-  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$,
-  and for every high-integrity component $\component$ for which
-  $\contour(\component) = \HIGHINT$, $M_0(\component) = M_n(\component)$
-  and $M'_0(\component) = M'_{n'}(\component)$.
-\end{itemize}
-
-\iftext
-\leo{In the terminology of the original hyperproperty paper,
-  confidentiality is a hyperproperty, integrity is a trace property
-  that holds for every trace in a hyperproperty. This is a stronger
-  version of integrity than DLM for example}.
-\fi
-%
-%\leo{Nested lockstep noninterference:}
-%
-%Lockstep-Noninterference is a restricted form of traditional
-%noninterference, that is nevertheless useful (e.g. to encode taint
-%tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
-%At the same time, lockstepness makes it possible to generalize the
-%setting and allow for the contours themselves to change throughout the
-%execution.
-
-%Generalizing to allow for contours to change throughout the execution,
-%nested lockstep noninterference is the contoured lockstep hyperproperty
-%where: \bcp{Which parts of this are common to {\em all} contoured lockstep
-%  hyperproperties (i.e., which should be part of the definition of contoured
-%  lockstep hyperproperties) and which are specific to (nested) NI?}
-%\leo{I'd say contoured lockstep hyperproperties are as above - and the following are
-%  the generalization for nested ones.}
-
-\section{Separation is noninterference$+$}
-
-\leo{To warm up, a super simple separation policy.}
-\begin{itemize}
-\item Any instruction can serve as a final instruction.  \bcp{My picture
-  here was that, for this property, the $X$ parameter would be, say, the
-  address where the machine is considered to halt. (I guess the picture is
-  that we attach probes to the PC and to the rest of memory, and we read out
-  the memory when the PC reaches this address.  But when I write it down
-  that way it seems pretty awkward.  Maybe we should just assume that the
-  machine has a HALT instruction or some other way of signaling that it is
-  done---i.e., abstractly, simply take the step function as a partial
-  function.  We will need to put a footnote somewhere acknowledging that
-  this may be a bit of an abstraction from real modern processors, but the
-  simple separation policy is an abstraction anyway!)}
-
-\leo{What about
-  instructions the are halted by the monitor?}  \bcp{Not sure what the
-  question is here, but note that an instruction halted by the monitor will
-  not appear in the trace.}
-\item For forward compatibility we could treat the first unmatched return as
-  a halt instruction.
-\item The instruction memory is marked $\LOWSEC \times \HIGHINT$. That means
-  it is observable and can't change.  \bcp{What does this assumption mean,
-    precisely?  (I.e., what is ``the instruction memory''?)}
-\item The memory space is split in two. One part is marked $\LOWSEC
-  \times \LOWINT$: this is the public part (or the one that can be
-  accessed by the current process e.g.). The other part is marked
-  $\HIGHSEC \times \HIGHINT$: it should be completely inaccessible to the
-  current process.
-\item Separation is noninterference:
-  \leo{Exactly the definition above}
-\item Depending on what we do with final instructions, the property
-  could be parameterized by the ``call/return sites''.
-\end{itemize}
-\leo{(The $+$ part in the title is a hint that one thing that gets added is that
-the PCs of corresponding traces always stay in lock step because of integrity?)}
-\bcp{The section titles are not working so well, now that we're being a bit
-  less abstract about our definitions.}
-
-\section{Stack Safety is noninterference$+$}
-
-\leo{Build up to the interesting thing:}
-\begin{itemize}
-\item The first unmatched return serves as a halt.  \bcp{What does that mean
-formally?}
-\item The instruction memory is again marked $\LOWSEC \times \HIGHINT$. That
-  means it is observable (both traces have the same instructions!) and
-  can't change.
-\item At the beginning of time\bcp{Not clear what ``beginning of time''
-  means when we're talking about hyperproperties---the intuition of time is
-  relevant to the traces within the hyperproperty, but not to the
-  hyperproperty itself.  Simpler, I think, to view the hyperproperty as a
-  starting set of traces (where the whole machine state [registers and PC too,
-  right?] is marked $\LOWSEC$), closed under an operator that takes a
-  contour+trace+trace and snips out all the subtraces between a call and its
-return.  (But: eek---see below.)}, the entire data memory is marked $\LOWSEC
-  \times \LOWINT$. That makes for a very unintersting property as all
-  variants are necessarily identical.
-\item BUT, we close the hyperproperty as follows:
-  \begin{itemize}
-  \item We need the state description stuff to enforce WBCF and be
-    able to talk about ``first unmatched returns''. That is also a
-    trace property.
-  \item After every call, we keep the part of the (left-hand, wlog
-    due to symmetry) trace until its first unmatched return.
-  \item We construct a new contour at that point: after every call,
-    the local state of the caller becomes inaccessible (marked
-    $\HIGHSEC \times \HIGHINT$) until a subsequent return.
-  \item For every variant of the new initial state with respect to the
-    new contour, the contour plus pair of traces is in the
-    hyperproperty.
-  \end{itemize}
-\end{itemize}
-
-\bcp{Eeek.  Just realized why we absolutely need the ``lockstep'' part!
-  The closure operator that snips out a ``subtrace'' between a call and the
-  corresponding return has to do the snipping on {\em both} traces at once,
-  and this only makes sense if they are executing in lockstep.}
-
-\section{Implementation}
-\label{impl}
-
-Sketch of the micropolicies that claim to implement these properties on
-PIPE...
-
-\subsection{A taint checking monitor}
-
-\subsection{A stack safety monitor}
-
-\section{Testing the implementations}
-
-\begin{itemize}
-\item Experimental results of testing
-\end{itemize}
-
-\section{Related work}
-
-\begin{itemize}
-\item Nick and Andre's papers (and other micropolicy papers)
-\cite{DBLP:conf/sp/RoesslerD18}
-
-This is the basis of our run-time monitor in Section{XXX}.  Specifically,
-our micropolicy is a simplified variant of their ... (I think it is the
-Return-Address Protection + Static Authorities + Depth Isolation policy, but
-need to double check---be careful about laziness.)  Essentially, our work
-arose as an attempt to formalize the security property enforced by something
-like one of their stack policies.
-
-\item Hritcu et al's work on ``Testing Noninterference
-Quickly''~\cite{DBLP:journals/jfp/HritcuLSADHPV16} laid the
-foundation of the random testing strategy that we use here.  In particular,
-that paper explored testing strategies based on both standard ``End-to-end
-noninterference''---a pleasantly extensional and easy-to-define property
-that leads to somewhat inefficient testing because process that misbehaves
-while its PC is labeled ``high'' must return to a low PC before the
-misbehavior can be identified as such---and ``Single-step
-noninterference''---a more intricate property related to Goguen and
-Meseguer's ``unwinding conditions''~\cite{GoguenMeseguer85} that ``fails
-faster'' and is thus easier to test.  \bcp{Explain the relation of the
-  latter property to our stack testing strategy.}
-
-\item Azevedo de Amorim's presentation of heap safety as a noninterference
-property~\cite{DBLP:conf/post/AmorimHP18} was a key inspiration for our
-definition of stack safety.  (Details...)
-
-\item StkTokens: Skorstengaard, Devriese, Birkedal
-\begin{itemize}
-\item POPL \cite{skorstengaard2019stktokens}
-\item tech report with proofs: \cite{skorstengaard2018stktokens}
-\url{https://arxiv.org/pdf/1811.02787.pdf}
-\item TR with proofs for ``Reasoning about a machine...'', Feb 2019:
-\url{https://arxiv.org/pdf/1902.05283.pdf}
-\item \url{https://cs.au.dk/~birke/papers/local-capabilities-conf.pdf}
-\item Also: {\em Reasoning about a Machine with Local Capabilities: Provably
-  Safe Stack and Return Pointer Management} \cite{10.1145/3363519}
-\url{https://dl.acm.org/doi/fullHtml/10.1145/3363519}
-
-\bcp{I'm nervous that these papers may be viewed as subsuming most of
-  ours---in particular, their security property is formulated as a logical
-  relation, i.e. a relational property!  We really need to dig into the
-  details (which are unfortunately rather technical).}
-
-\rb{
-  Some notes on StkTokens.
-}
-\rb{
-  A calling convention based on linear capabilities to enforce WBCF and LCE. The
-  program is divided into mutually distrustful components. Instead of using
-  separate stacks for each component and a shared control stack, it uses a
-  single shared stack. For this, it needs to pass stack and return capabilities
-  to possibly untrusted components temporarily, revoking them later. Together
-  with a small number of security checks, these measures suffice to guarantee
-  WBCF and LCE. StkTokens employs linear capabilities to obtain this kind of
-  result while limiting the amount of overhead needed to secure the transitions
-  between boundary crossings, modeled by a standard notion of components,
-  linking, disjoint interfaces, etc.
-}
-\rb{
-  The capability machine that is modeled has two final configurations: halted
-  (for successful termination) and failed. It is a mostly standard capability
-  machine, modified by the additional requirements of linear capabilities
-  (moves, stores and loads, seals, splitting and splicing of ranges of
-  capability).
-}
-\rb{
-  The formalization relies on the definition of a second operational semantics
-  for the target language, where WBCF and LCE are natively enforced, and a proof
-  of FA between that and the original operational semantics of the target. This
-  semantics executes in lockstep with the original target semantics; in
-  addition, it gives a special interpretation to the sequence of call
-  instructions on top of the regular step relation. It assumes that all
-  components meet some syntactic well-formedness requirements, and requires
-  trusted components to perform a number of simple checks in addition to
-  employing the calling convention proper (check of stack base address on call,
-  appropriate use and protection of return and closure seals), as well as some
-  conditions on initial system configurations. The formalization has a number
-  of advantages over previous work, notably the establishment of some temporal
-  safety guarantees.
-}
-\rb{
-  The core of the proof of FA is the definition of when two components are
-  related. This is modeled by a step-indexed Kripke logical with recursive
-  worlds, which each collect assumptions about the system. Worlds are divided
-  into sub-worlds for heap, call stack and free stack, each mapping region names
-  to regions, which can be either spatial (all three) or shared (heap only,
-  encapsulating a system invariant, including an invariant on seals). Spatial
-  regions differ from shared regions in their linearity and (optionally)
-  ownership. ``Regular'' spatial regions can become owned or revoked without
-  while preserving the relation between memories in the two semantics. The proof
-  uses (shared) code regions, regions for heap and stack memory, and regions to
-  track immutable, inactive parts of the stack. With those Kripke worlds, the LR
-  defines relatedness of various entities. There are two value relations for
-  trusted and untrusted values (a subset of trusted values which are safe to
-  pass to an adversary, including stack pointers and their linear capabilities,
-  and sealed values, all under certain conditions on the ownership of the
-  various parts of the stack). The top-level FA proof is straightforward.
-}
-\rb{
-  The extended operational semantics has support for tail calls. Because data
-  memory does not contain any seals or ways to get to those, it can always be
-  shared safely.
-}
-\rb{
-  Like in our scheme, instructions must be marked as executable and a check on
-  the PC made at each step. Stack pointers are now linear stack capabilities
-  (and there should only be stack pointers for the stack memory). Return
-  pointers are pairs of sealed return capabilities, encapsulating the callers
-  data and return code; the callee can only jump to these. A caller will split
-  its own stack capability into two: one for its local stack frame (which will
-  be sealed as the data component of the return capability pair) and one for the
-  unused stack, which will be passed to the callee. This second capability is
-  also used as a token to prove that the callee is allowed to return to the
-  caller by adjacency with the sealed data capability (for this trick to work,
-  stack frames must be non-empty). Another requirement is that each call point
-  needs to be associated with a unique seal for returns, which should never be
-  leaked to an adversary.
-}
-\rb{
-  In their ESOP 2018 paper, the same authors present a similar scheme that uses
-  local capabilities employing similar techniques. Their capability machine is
-  mostly standard. Operations other than those working on capabilities are
-  mostly standard, with small extensions to account for the presence
-  capabilities. Notably, jmp needs to account for enter capabilities, which can
-  only be used to jump to a location and then become read-execute; they
-  represent an opaque packing of some code together with local state. The store
-  operation is also complicated, this time by the presence of local
-  capabilities, which can only be stored in registers (and write-local memory)
-  and may be temporarily handed over to an adversary, say for the duration of a
-  call.
-}
-\rb{
-  The work presents a concrete calling convention that relies on a single shared
-  stack and has moderate overhead with stack and return pointers passed through
-  registers as local capabilities. The scheme imposes some standard conditions
-  on initial state and memory allocation primitives, and safety measures a
-  component needs to perform to protect itself against an adversary. The most
-  demanding requirement is that not only non-return value registers and used
-  stack, but also the unused part of the stack (which may contain local
-  capabilities) needs to be cleared when returning to the adversary. Simple
-  measures are applied on calls and returns to and from an adversary, and on
-  calls from an adversary.
-}
-\rb{
-  The logical relation that is presented is not specific to the particular
-  calling convention. It uses a variant of the public-private separation of
-  worlds to model the particular properties of local capabilities. In
-  particular, worlds have permanent, temporary and revoked regions, with
-  non-revoked regions comprising a state transition system with both public and
-  private transitions. Local and global capabilities can depend on protocols
-  from permanent regions, and local capabilities can also depend on protocols
-  from temporary regions (which may be revoked in the future, and may not
-  outlive those regions). A pair of future world relations describe the changes
-  of memory over time. In the public FW relation, temporary and permanent
-  regions can evolve according to the public part of their transition system,
-  and revoked regions can be reinstated as temporary. In the private FW
-  relation, temporary and revoked regions can be replaced by arbitrary regions,
-  and permanent regions can evolve according to the private part of their
-  transition system. The relation defines an observation relation for safe
-  configurations (i.e., the memory of a halting configuration is satisfied by
-  some private FW). The relation on values covers conditions for the safe use of
-  write and write-local capabilities, and similarly for enter capabilities.
-}
-\rb{
-  All this is used to prove WBCF and LCE for concrete example programs.
-}
-
-\end{itemize}
-
-\item Stuff about dynamically changing labels (see footnote in Testing
-Noninterference Quickly paper)
-
-\item Simple Noninterference by Normalization \url{http://nachivpn.me/nibnbe.pdf}
-
-\item Mechanized Reasoning about a Capability Machine: [Linn] Georges,
-Trieu, Birkedal (PriSC 2020).
-%
-Described at the Iris Workshop as formalizing the security properties of
-capabilities. This appears to involve a fairly complex and very large
-mechanization in Iris (also in the eyes of the Iris people).
-
-\item Rigorous engineering for hardware security: Formal modelling and proof in
-the CHERI design and implementation process
-\url{https://www.cl.cam.ac.uk/research/security/ctsrd/pdfs/202005oakland-cheri-formal.pdf}
-
-\item \url{https://www.cis.upenn.edu/~stevez/papers/LZ05a.pdf} ??
-
-\item Not so closely related?
-\begin{itemize}
-\item Daniel's tainting paper (and some refs cited there)
-      \cite{DBLP:conf/eurosp/SchoepeBPS16} -- not needed if we postpone
-      dealing with tainting till another paper!
-\item We'd better say something about the relation to the original
-hyperproperties paper, no?  (No, not if we drop the word hyperproperty!)
-\end{itemize}
-\end{itemize}
-
-\iftext
-\section{Future work}
-
-\begin{itemize}
-\item Variations / extensions (OS, multiple processes, coroutines, generalizing depths to stacks of static call points, etc.)
-\item Extending the policy / property to more realistic ones
-\begin{itemize}
-\item Argument passing
-\item Variable-size stack frames
-\item Stack variables
-\item Relation to Nick and Andre’s policy
-\end{itemize}
-\item Could we also implement all this using Cheri capabilities?
-\item A more abstract presentation, perhaps unifying our treatments of
-confidentiality and integrity?
-\end{itemize}
-
-More:
-\begin{itemize}
-\item More realistic calling conventions.
-(We need to eventually match this to the particular calling convention of whatever compiler we choose.)  E.g., the above will not quite work for passing arguments on the stack: callee won’t be able to read them…So perhaps need to support a special kind of write instruction that stores D.PCdepth+1 .
-[RB: Or implement this distinction through tags?]
-\item Combination with heap policy (colors, etc.)
-\item For coroutines:
-\begin{itemize}
-\item Generalize depth to depth + PID
-\item Private stuff not having changed as an idea should still work
-\item Maybe not so hard: The scheduler + other coroutines are just a sort of generalized “callee” as far as this process is concerned.
-Just need to fiddle th step relation so that it recognizes which process is getting returned to and updates its description appropriately.
-\end{itemize}
-\item If we had a heap policy, maybe we could talk about dynamic code generation, loading, as allocating a heap region, storing instructions into it, and then calling into them.
-For now, we just make all the instructions read only and assume they exist in the initial state.
-\item Question: how efficiently can consistency be tested?
-How much scrambling is needed, and does it really need to occur before every instruction?
-How expensive is it to check for agreement between the (in)accessible parts of two memories?
-Would spot checking for agreement be adequate?]
-\item What are the effects of compartmentalization on the ability of each compartment to change its own code/stack?
-\item Tail calls (maybe by not marking them as call sites?)
-\end{itemize}
-
-\section*{Acknowledgments}
-\fi
-
-\section*{References}
-\bibliographystyle{IEEEtran}
-% \bibliography{bcp,local}
-
-\newpage
-
-\section*{Things to think about}
-
-\begin{itemize}
-\item
-How does Nick’s stack policy actually work these days?
-\item Andrew remembers that Catalin looked at something like this a long time ago and they decided that they needed unique tags for each dynamic call, not just levels.  We should check.
-\item \bcp{A random thought about integrity: One thing that may be helping confuse
-  us about integrity is that we've gone back and forth between ``eager'' and
-  ``lazy'' versions of various micropolicies, where the ``lazy'' ones do
-  allow high-integrity data to be overwritten, but only in a way that it
-  cannot be read again later (e.g., the lazy heap policy guarantees that, if
-  inaccessible memory is overwritten, then its color will change).  The lazy
-point of view is much more subtle as an integrity property.  So, to keep our
-heads straight here, we should be thinking only about eager properties in
-the first instance.}
-\end{itemize}
-
-\newpage
-
-\section{Old Stuff}
-
-\begin{itemize}
-\item {\em prefix-closed:}
-
-  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
-    M')}_n$ in the set, \bcp{where $n$ is the length of the traces,}\leo{Yes I put that in the definition of traces above, but we could repeat it here} the
-  pair of truncated traces $\overline{(C,
-    M)}_m$ and $\overline{(C, M')}_m$ is also in the set if $m < n$ \bcp{or
-    $n = \infty$}.
-
-\item {\em suffix-closed:}
-
-  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
-    M')}_n$ in the set, and for every $0 \leq i < n$, there exists in
-  the set a pair of traces $\overline{(C', N)}_m$ and $\overline{(C',
-    N')}_m$ such that for all $0 \leq j < m$, $C_j' = C_{i+j}$, $N_0 =
-  M_i$, and $N_0' = M_i'$, and where $m$ is the maximum index for
-  which $C_i \leq C_{i+j}$ for all $0 \leq j < m$. Here $C \leq C'$ if
-  $\forall \component, C(\component) \leq C'(\component)$ and $\VIS <
-  \HID$.
-
-\item {\em variant-closed:}
-
-  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C,
-    M')}_n$ in the set, and for every machine state $N_0$ such that
-  $N_0$ is a variant of $M_0$, there exists in the set a pair of
-  traces $\overline{(C, M)}_n$ and $\overline{(C, N)}_n$.\apt{Same
-    complaint about $n$. And again, $N_0$ defines $\overline{N}$ but
-    it doesn't define $\overline{(C,N)}$. }\leo{Cs should be the same
-    right?}
-
-\item {\em as in noninterfence:}
-
-  For each pair of traces in the hyperproperty, if the initial machine
-  states plus contour \apt{``elements''} are variants, then the observervations of the
-  two traces coincide.
-
-\end{itemize}
-
-\bcp{It feels like something is missing here: This definition seems to be
-  requiring that secret data never affect observations, but our tester is
-  ensuring something more: that inaccessible locations are not {\em changed}
-(at least without also changing their tag to make them accessible).  Where
-  is that condition being imposed?}
-\leo{It's not - I just went by the definition of (traditional) noninterference
-  you had and repeated that. I'm thinking about that now}
-
-\apt{Even if it works for subroutines,
-  the basic idea of truncating traces when visibility increases doesn't
-  seem to extend to things like coroutines, where we swap between two sets of visible locations on each transfer.}
-
-% \bcp{Should we require that it contain all possible starting states and
-%   contours (though of course the sequences can be singletons in many
-%   cases!)?  What does it mean if it doesn't?}
-
-\bcp{Policies and properties need to know where are the call and return
-  sites, and for this they need to be initialized with some
-  application-specific information.  So it seems what we should do is to
-  talk instead about ``$X$ policies'' (i.e., make ``policy'' a type
-  operator) and replace the initial $\CONTOURS$ argument to both policies
-  and properties by an element of $X$.}
-
-%
-% \leo{Some notes for the meeting}
-%
-% In traditional NI viewed as a pair of contoured traces $\overline{(C, M_i)}$ and
-% $\overline{(C, N_i)}$, we can have the following formulations:
-% \begin{enumerate}
-% \item $M_0, N_0 \in \mathit{Initial State}, ~\text{s.t.} ~\mathit{variant}(C, M_0, N_0)$
-% \item $M_0 \in \mathit{Initial State}, N_0 \in \mathit{Variant}(C, M_0)$
-% \item $M_0, N_0 \in \mathit{Reachable State}, ~\text{s.t.} ~\mathit{variant}(C, M_0, N_0)$
-% \item $M_0 \in \mathit{Reachable State}, N_0 \in \mathit{Variant}(C, M_0)$
-% \end{enumerate}
-% %
-% The (intuitive) top level property we care about is (1) - end to end
-% noninterference.
-% %
-% Property (2) is easier to test - but not always equivalent (if the set
-% of variants of an initial state can take you outside the set of
-% initial states).
-% %
-% Property (3) is also easier to test than - you start from an arbitrary
-% (reachable) state instead of an initial one, allowing to get to the
-% interesting part sooner.
-% %
-% Property (4) corresponds to what we actually want to test - start from
-% an arbitrary (reachable) state, vary it, test the trace.
-%
-% As far as facilitating testing goes, the assumption of SSNI is that
-% you don't need to test full traces - you can test all traces of length
-% 1 and that somehow gives you confidence for the full property.
-% %
-% More importantly, note that going from (1,2) to (3,4) doesn't really
-% change the property itself {\em as long as the contour stays identical
-%   throughout the execution}.
-%
-% Now, when the contour does change the expressive power of (3,4) and (1,2)
-% is slightly different:
-% \begin{itemize}
-% \item If the contour C makes more things hidden at step $i$, then
-%   starting from the (reachable) state $M_i$ and taking its variant can
-%   lead to a pair $M_i, N_i$ that is not reachable from a pair of variant
-%   initial states.
-% \item If the contour C declassifies things at step $i$, then traces
-%   that started from variants $M_0, N_0$ based on a more restrictive
-%   contour C, can reach that point and no longer remain variant based
-%   on the new contour.
-% \end{itemize}
-% Therefore an additional property of the hyperproperties when
-% classification/declassification is involved, is that the traces
-% in the property start at some contour level C0 and are truncated
-% after C becomes more permissive than C0.
-%
-% The stack property we test is just a way of testing such a
-% hyperproperty more efficiently.
-
-% If $\mach$ and $\mach'$ are variant machine states with respect to contour
-% $\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
-% under policy $\pol$ if their observable
-% %
-% $\obstrace(
-%                       (\contour,\mach)
-%   \stepstounder{\pol} (\contour_1,\mach_1)
-%   \stepstounder{\pol} (\contour_2,\mach_2)
-%   \stepstounder{\pol} \ldots)$
-% %
-% yields the same trace as
-% $\obstrace(
-%                       (\contour,\mach')
-%   \stepstounder{\pol} (\contour_1,\mach_1')
-%   \stepstounder{\pol} (\contour_2,\mach_2')
-%   \stepstounder{\pol} \ldots)$.
-
-% % either (1) one or both of $(\contour,\mach)$ and
-% % $(\contour,\mach')$ fail to terminate when executed with
-% % $\stepstounder{\pol}$, or (2) both terminate and the machine-state
-% % components of their results are variants with respect to the original
-% % $\contour$.
-
-\iftext
-\begin{itemize}
-\item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
-  have to begin with EENI and then get to SSNI as a (stronger) refinement.}
-
-\item Relation to the standard definition of NI.
-\end{itemize}
-\fi
-
-% Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
-% every pair $(\mach,\mach')$ of variant machine states with respect to
-% $\contour$ are indistinguishable under $\pol$.
-
-\iftext
-\bcp{
-Fact: The standard RISC-V semantics (i.e., the null policy) does not
-guarantee NI.  (Give a little example of a program that reads from a secret
-location and writes to a public one, just to illustrate the definition of
-NI.)}
-\fi
-
-\subsection{Policies}
-
-\leo{I think with the property being just a (special) hyperproperty
-  again, the policy can be ``more detailed'' - including low-level
-  enforcement details like tags, etc. But you should always be able to
-  reconstruct the contoured trace given an initial contour and an
-  initial state. I'm taking that approach here:}
-
-Formally, a policy
-%
-$\pol \in \exists X.\; (\CONTOURS \rightarrow X)
-                       \times (X \rightarrow \CONTOURS)
-                       \times (\MACHS \times X \rightharpoonup \MACHS \times X)$
-%
-comprises
-\begin{enumerate*}
-  \item a private type $X$ corresponding to an ``implementation'' of the contour,
-  \item an initializer: a function taking a contour \leo{and a machine state?} to an initial value of $X$,
-  \item a projection: a way of reading a contour out of the private type,
-  \item a partial step function of type $(\MACHS \times X) \rightharpoonup X$
-    that, when defined, defines the private state after a step of the machine
-    (based on the $\stepstounderfull$ relation).
-    \leo{Not sure this should be taking $\MACHS$ as input. It should take $X$
-      plus just the instruction being executed? The rest are the tags in $X$}
-\end{enumerate*}
-That is, a policy takes a starting contour and produces a runtime monitor
-that carries along some private information and uses it to decide whether to
-terminate execution early.
-
-\leo{Example? how detailed?}  For example, for standard
-lockstep-noninterference with only two secrecy levels, the private
-state {\em is} the contour.
-
-A policy $\pol$ {\em guarantees} property $\prop$ if for every pair of
-states $M$ and $M'$ and for every starting contour $C$, the pair of
-contoured traces that you get from the policy is in the property.
-
-\leo{New attempt ends here}
-Informally, a (micro-){\em policy} is an enriched step function that,
-whenever is is defined, agrees with $\stepstounderfull$.  It represents an
-enforcement mechanism that sometimes halts programs earlier than the bare
-step relation does but that agrees with the bare step relation as long as it
-allows evaluation to proceed.  We write
-$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
-$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.
-
-\leo{Shouldn't this take a pair (C,M) to a pair (C,M)?}
-Formally, a policy
-%
-$\pol \in \CONTOURS \rightarrow \MACHS \rightarrow \SEQ{\MACHS}$
-%
-takes a contour $\contour$ and a starting machine state $\mach_0$ and produces
-a finite or infinite sequence of states $\mach_0,\,\mach_1 ,\, \ldots$ such
-that $\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.
-
-% \bcp{Fix: even simpler and cleaner, just run the step function to completion
-%   and forget about $X$---i.e., a property is just a function from a pair of
-%   a contour and a machine state to a finite or infinite sequence of machine
-%   states!}
-% Formally, a policy
-% %
-% $\pol \in \exists X.\; (\CONTOURS \rightarrow X) \times (\MACHS \times X
-% \rightharpoonup \MACHS \times X)$
-% %
-% comprises
-% \begin{enumerate*}
-%   \item a private type $X$,
-%   \item a function taking a contour to an initial value of $X$, and
-%   \item a partial step function of type
-%   $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
-%   agrees with $\stepstounderfull$ on machine states. \bcp{We could actually
-%     just make it return an X and calculate the machine state from the
-%     ordinary step function.  This is perhaps a bit cleaner.}
-% \end{enumerate*}
-% That is, a policy takes a starting contour and produces a runtime monitor
-% that carries along some private information and uses it to decide whether to
-% terminate execution early.
-
-Policy $\pol$ {\em guarantees} property $\prop$ if \bcp{for every pair of
-  starting states and every starting contour, the pair of traces that you
-  get from the policy is in the property for {\em some} pair of sequences of
-  contours that agree with the given contour on the first element.}
-
-% every $\contour$,
-% every pair $(\mach,\mach')$ of variant machine states with respect to
-% $\contour$ are indistinguishable under $\pol$.
-
-\subsection{Alternative Definition of Contoured Hyperproperty}
-
-\apt{Just for the record, here's an attempt to write down the splice-based
-  alternative I suggested. I \emph{think} it might work for stack safety
-  (and coroutine safety).
-But there are some problems: (a) it doesn't account for the caller being able to write into the callee's local data (although perhaps there's no good reason to allow this anyway); (b) two calls that just happen to use the same memory space for their frames will share the same contour, which seems wrong.
-I'm wondering if it would be better to use a more
-  explicit and nominal refinement of contours (maybe called ``context''); two
-  contexts might have the same contour but the trace partitioning would be done based on context name. I haven't thought about relating all this to testing...}
-
-  Let $T = \overline{(C,M)}_n$ be a contoured trace of length $n \leq \infty$.
-  Write $T\vert_C$ for the result of
-  filtering $T$ to contain only elements with contour $C$ and
-  then projecting the trace components.
-  The \emph{contour partition} of $T$ is the set
-  $P(T) = \{(C_{i_1},T\vert_{C_{i_1}}),
-  (C_{i_2},T\vert_{C_{i_2}}),
-  \ldots,(C_{i_d},T\vert_{C_{i_d}})\}$ where
-  $C_{i_1},C_{i_2},\ldots,C_{i_d}$ are the distinct contours appearing in $\overline{C}$.
-  In other words, $P(T)$ is the result of dividing the trace into sections based on contour and then splicing together the sections.
-Note that these traces no longer obey the underlying step function; e.g., there might be a step directly from a call Jal to the return address.
-(We may need to pin down exactly what is legal in these ``gappy'' sequences.)
-
-  Observations are a bit simpler than before.
-  A \emph{state observation} with respect to a contour $C$ is a map from
-  the machine components visible in $C$ to words.
-  An \emph{observation} with respect to a contour $C$ is
-  is a nonempty, finite or infinite sequence of
-  state observations with respect to $C$, in which adjacent elements
-  are distinct (``no stuttering'').
-  In particular, the \emph{observation trace} of
-  a machine state sequence $\overline{M}$ with respect to a contour $C$,
-  written $\obstrace_C(\overline{M})$ is the observation
-  obtained by removing duplicates from the sequence
-  $\stateobs_0,\, \stateobs_1,\, \ldots$ where
-  $\stateobs_i(\component) = M_i(\component)$ for all $\component$ such that $C(\component) = \VIS$.
-  We extend observation traces to contour partitions piecewise:
-  if $P(T) = \{(C_1,\overline{M}_1),(C_2,\overline{M}_2),\ldots,(C_d,\overline{M}_d)\}$, then $\obstrace(P(T))$ is $\{(C_1,\obstrace(\overline{M}_1)),
-  (C_2,\obstrace(\overline{M}_2)),\ldots,(C_d,\obstrace(\overline{M}_d))\}$
-
-  Variants are as before.
-  Noninterference (no need for lockstep) with respect to a (single)
-  contour $C$, written $NI_C$, is defined as as a (standard) hyperproperty: it is the set of pairs of (ordinary, but possibly gappy) machine-state traces such that if the initial elements of the traces are variants with respect to $C$ then their observation traces with respect to $C$ are identical.
-
-  A \emph{contoured hyperproperty} (no restriction to lockstep) is still a set of pairs of contoured traces.
-
-  \emph{Contoured noninterference} is the contoured hyperproperty defined as follows: the pair of contoured traces $(T,T')$ is in the hyperproperty if,
-  for each $(C,\overline{M}) \in P(T)$ and $(C',\overline{M'}  ) \in P(T')$,
-  if $C = C'$ then $(\overline{M},\overline{M'}) \in NI_{C}$. (NB: This may
-  be too liberal; perhaps the sets of $C$'s should be the same?)
-
-Standard noninterference falls out in the obvious way when trace contours are invariant. (In this case, the machine-state traces are not gappy.)
-
-\iftext
-\subsection{More...}
-  \begin{itemize}
-  \item Proposal: Maybe forget about full-blown NI and focus the whole paper
-  on variants of ``lockstep NI'' where the PC is the same in every pair of
-  related runs?
-  \item One issue that I am not sure about: Could there be a sneaky way of
-  ``putting secret information into the control state'' where the PCs of
-  related runs are always in lockstep but, because the instruction memory is
-  writeable, different instructions can get dispatched.
-
-  E.g., suppose the program looks like
-\begin{verbatim}
-0:   load from address 1000
-     write to address 100
-     jump to address 100
-100: <initial contents don't matter>
-     write R3 into address 1004
-     jump to address 100
-1000: <marked HIGH so that initial
-       contents vary on different runs>
-1004: <marked OBSERVABLE>
-\end{verbatim}
-  If the initial contents of address 1000 in one run is the encoding of the
-  instruction ``load immediate 0 into R3'' and the initial contents in the
-  other run is the encoding of the instruction ``load immediate 1 into R3,''
-  then the two programs execute different instructions (leading them to
-  write different things to an observable location) even though their PCs
-  are always in lock step.
-
-  So maybe we should be asking that the instruction stream be
-  identical, or maybe instructions paired with PCs?
-
-  \bcp{Answer: Just make instructions read-only.}
-
-  \item One nice observation is that, if we are only interested in pairs of
-  runs with identical PCs, there are two different ways to proceed: (1) say
-  that we {\em ignore} runs where the PCs (or instruction streams) diverge
-  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
-  instruction streams) diverge, i.e., demand that this never happens (e.g.,
-  stack and heap policies).
-  \item A remaining question is what to do about the possibility that the
-  contour changes differently on two related runs.  We could disallow this,
-  but not certain it's necessary---maybe everything works out OK.
-  \end{itemize}
-\fi
-
-\newpage
-\section{Tainting is noninterference$-$}
-
-\leo{Rough draft/take on this section}
-\begin{itemize}
-\item Taint tracking is an interesting property.
-\item It's weaker than traditional noninterference, but easier to enforce so people use it.
-\item \leo{Some of the intuition from below}
-\end{itemize}
-
-\paragraph{Taint Tracking Property}
-
-Given a contour C and pair of traces $\overline{M}$ and $\overline{N}$,
-the pair of contoured trace $\overline(C, M)_n$ and $\overline(C,N)_n$ is in the
-taint tracking contour-lockstep hyperproperty if:
-\begin{itemize}
-\item For all $0 \leq i < m$ the restriction of the trace to its
-  length $i$ prefix is in the standard noninterference property
-  \leo{Defined in the previous section}, where $m$ is the minimum
-  index (or infinity) for which $PC(M_i) \not = PC(N_i)$ or
-  $instr(M_i) \not = instr(N_i)$.
-\end{itemize}
-
-\paragraph{Taint Tracking Policy}
-
-\newcommand*{\clean}{{\em clean}}
-\newcommand*{\tainted}{{\em tainted}}
-\newcommand*{\boring}{{\em boring}}
-\begin{itemize}
-\item We define three {\em tags} \leo{have we defined what these
-  are?}: \clean, \tainted, and \boring. Machine components tagged
-  \clean by the policy represent protected locations that should never
-  be tainted. Components tagged \tainted are either sources of taint
-  (e.g. network inputs) or components that have been influenced by
-  other \tainted ones. Finally \boring components are locations that
-  are allowed to become tainted (perhaps temporarily) but have not
-  yet.
-\item To implement the taint tracking property as a policy, we
-  initialize a map from components to tags. From this map, we can read
-  a contour in a straightforward manner: \clean is \VIS, the other
-  tags are \HID.
-\end{itemize}
-
-\leo{How do we want to write these rules? Explain .dpl?}
-%//TODO: This is a bug opportunity, right?
-%   // explicit failure for write tainted val to clean addr
-%   storeGrp(   mem == [+Clean], addr == _, val == [+Taint], env == _
-%              -> fail "Tainted data written to clean memory")
-%
-%   // Propogate taint
-%#ifndef ADD_NO_TAINT_1
-%   ^ arithGrp(env == _, op1 == [+Taint], op2 == _ -> env = env, res = op1)
-%#else
-%//   ^ arithGrp(env == _, op1 == [+Taint], op2 == _ -> env = env, res = {})
-%#endif
-%#ifndef ADD_NO_TAINT_2
-%   ^ arithGrp(env == _, op1 == _, op2 == [+Taint] -> env = env, res = op2)
-%#else
-%//   ^ arithGrp(env == _, op1 == _, op2 == [+Taint] -> env = env, res = {})
-%#endif
-%   ^ mulDivRemGrp(code == _, env == _, op1 == [+Taint], op2 == _ -> env = env, res = op1)
-%   ^ mulDivRemGrp(code == _, env == _, op1 == _, op2 == [+Taint] -> env = env, res = op2)
-%   ^ immArithGrp(env == _, op1 == [+Taint] -> env = env, res = op1)
-%#ifndef NO_MEM_TAINT_CHECK_LOAD
-%   ^ loadGrp(mem == [+Taint], addr == _, env == _ -> env = env, res = mem )
-%#else
-%// No mem check
-%#endif
-%#ifndef NO_ADDR_TAINT_CHECK_LOAD
-%   ^ loadGrp(mem == _, addr == [+Taint], env == _ -> env = env, res = addr )
-%#else
-%// THIS WAS AN ACTUAL BUG! Addres check ommitted
-%#endif
-%#ifndef NO_VAL_TAINT_CHECK_STORE
-%//NOT A BUG! We never write tainted data to clean memory because of the first rule
-%   ^ storeGrp(mem == _, addr == _, val == [+Taint], env == _ -> env = env, mem = val)
-%#else
-%// NO CHECK
-%#endif
-%#ifndef NO_MEM_TAINT_CHECK_STORE
-%   ^ storeGrp(mem == [+Taint], addr == _, val == _, env == _ -> env = env, mem = mem )
-%#else
-%// NO CHECK
-%#endif
-%#ifndef NO_ADDR_TAINT_CHECK_STORE
-%   ^ storeGrp(mem == _, addr == [+Taint], val == _, env == _ ->
-%     fail "Can't write through a tainted pointer")
-%#else
-%// Could also taint mem with addr. Also an actual bug (but one that I made).
-%// THIS WAS AN ACTUAL BUG! Addres check ommitted
-%#endif
-%   // keep clean mem stays clean always
-%   ^ storeGrp(   mem == [+Clean], addr == _, val == _, env == _ -> env = env, mem = mem)
-
-\bcp{This needs to be written, now that we have added the hidden/observable stuff.}
-
-\begin{itemize}
-\item First intuition: A program is taint-NI if every pair of runs that don’t
-branch on high data are noninterfering in the standard sense.  (i.e., they
-don't write secret data to public locations and vice versa)
-\item Better alternative: … for every pair of runs that take the same control path (even if they branch on tainted data)...
-\item We could even go a step farther: {\em truncate} the runs at the
-  first point where their PCs differ, then demand NI.  This is a less
-  demanding property and probably a less desirable one, as it doesn't
-  predict the behavior of actual taint policies.  Discuss.
-\item Note that this is a super-simple property that doesn't really track
-taint at all.
-\item (Are there any other variations of tainting that we could present in
-this framework?)
-\end{itemize}
-
-\section{Separation is noninterference$+$}
-
-Presentation idea: Start here with a super-simple separation policy, which
-just makes some part of the address space inaccessible, truncating any
-traces that read or write it.  Formulate the property as a flavor of NI.
-(The $+$ part in the title is a hint that one thing that gets added is that
-the PCs of corresponding traces always stay in lock step.)
-
-\section{Stack protection is noninterference$++$}
-
-Here we consider stack protection as an instance of generalized noninterference \apt{or whatever it's called}.
-Our notion of stack protection includes aspects both of control integrity and data security.
-In the terminology of Skorstengaard et al.~\cite{Skorstengaard+19} we enforce:
-
-\begin{itemize}
-\item \emph{Well-bracketed control flow (WBCF)}, which says that ``invoked functions must either return to their callers,
-  invoke other functions themselves, or diverge.''
-\item \emph{Local state encapsulation (LSE)}, which says that ``when a function invokes another function, its local variables
-  (saved on its stack frame) will not be read or modified until the invoked function returns.''
-\end{itemize}
-\apt{Should rephrase (and improve!) these in our own words.}
-
-These properties are essentially orthogonal: one can certainly have WBCF without LSE, and variants of
-LSE could apply to control constructs beyond simple procedure calls (e.g. exceptions, coroutines, threads, monitors, etc.).
-
-Under the assumption that WBCF holds, we can describe LSE as an instance of Lockstep-Noninterference, where the contour at
-each step maps the current stack frame to \VIS\ and the remainder of the stack to \HID\.
-The observations on which noninterference is defined will (necessarily)  be made on the values in the current stack frame.
-\apt{Again, we're conflating \VIS\ vs. \HID\ with \LOWSEC\ vs \HIGHSEC\; should probably be teased apart.}
-LSE has two aspects: secrecy (callees cannot read caller variables) and integrity (callees cannot alter caller variables).
-Secrecy is enforced by the variant-closure clause of L-N: since caller data is \HID, changing it cannot alter
-the callee's observable behavior.  Integrity is enforced why??? \apt{I've lost confidence that it is.}
-
-To make these ideas concrete in this paper, we adopt a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls.
-Each procedure stack frame contains just one local variable (plus possible control information), which is pointed to by
-the SP whenever the PC is within the procedure body (excluding entry and exit sequences). Our WBCF property says that
-when a procedure is called from a given PC at a given SP, it eventually returns (if at all) to the next
-address following the PC with the same SP. Our LSE-like property says that addresses between the base of the stack
-up to (but not including) the current SP cannot be read or written.
-\apt{Not clear why we should bother allowing addresses beyond the current top of stack to be accessible.}
-In particular, this implies that upon return from a call,
-the value of the local variable will be unchanged; moreover, the values of local variables cannot influence the
-behavior of callees. Parameters and return values are passed in registers, which (except for SP and RA) do not change
-during call or return.
-
-{\bf Getting more precise about the property.}
-
-Assumptions on code and environment\apt{trying to be as minimal as possible}:
-\begin{itemize}
-\item
-  Memory is divided into two disjoint regions, one containing the code, the other the stack.
-  Code begins at address 0; the stack begins at address B and grows up, in the direction of increasing addresses.
-\item
-  The part of the stack between $B$ and the stack pointer \textsc{sp} (inclusive) is structured into a contiguous sequence of
-  \emph{frames}. Each frame contains a single word of local data, possibly preceded by control information. \apt{TMI?}
-  The initial value of \textsc{sp} is $B-1$.
-\item
-  Code is assumed to be be annotated to indicate (a) instructions representing calls, (b) instructions forming the \emph{entry sequence} of
-  a procedure, and (c) instructions forming the \emph{return sequence} of a procedure.
-\item
-  Execution of the entry sequence pushes a new frame on the stack, adjusting \textsc{sp} to point to that
-  frame's data word.  Execution of the return sequence pops the frame on top of the stack by readjusting \textsc{sp}.
-  \apt{TMI!?}
-  Outside of the entry and return sequences, no instruction alters \textsc{sp}.
-\item
-  The call instruction and entry and return sequences do not alter any registers except \textsc{sp} and one
-  particular register named \textsc{ra}.
-\end{itemize}
-
-DO HERE: Formal description of conjecture 2 property.  Or maybe we no longer really care about this particular property if we have something for conjecture 3?
-
-An (abstract) \emph{state description} $D$ consists of
-\begin{itemize}
-\item
-  A stack $\dstk{D}$ of addresses, representing the return addresses of the current procedure and its callers.
-  \apt{No need to remember the sp's, since they can be computed given that we know the fixed frame size.}
-  \apt{No need for $\dpcd{D}$ since it is just the size of $\dstk{D}$.}
-\item
-  For each memory location $a \geq B$, a natural number $\ddep{D}{a}$, representing the depth
-  of the frame to which $a$ belongs ($\infty$ if none).
-\end{itemize}
-
-The \emph{initial state description} $\dinit$ has $\dstk{\dinit} = \empstack$ and $\ddep{\dinit}{a} = \infty$ for all $a \geq B$.
-
-An address $a$ is \emph{accessible} with respect to a state $D$, written $\access{D}{a}$, if
-$a \geq B$ and $\ddep{D}{a} \geq \norm{\dstk{D}}$.
-
-TODO:
-\begin{itemize}
-\item
-  Define step relation.
-\item
-  Define step consistency.
-\item
-  Discuss relationship between state consistency and top-level property.  (Former is: presumably stronger; hopefully not
-  too strong that it produces lots of undesired test failures; surely easier to test).
-
-\end{itemize}
-
-\iftext
-\section{Stack + heap protection too}
-
-POTENTIALLY: Memory safety (before stack) + Conjunction of memory + stack
-safety (after).
-
-(Memory safety by itself is just arthur’s paper, but it’s cool to see how
-they are combined, especially what bits can be factored out as common
-structure, e.g. the notion of accessibility)
-\fi
+      \leo{Need to fix append. Maybe call it join. It should remove the last
+        state of left arg}
+
+%% Acknowledgments
+\begin{acks}                            %% acks environment is optional
+                                        %% contents suppressed with 'anonymous'
+  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+  %% acknowledge financial support and will be used by metadata
+  %% extraction tools.
+  This material is based upon work supported by the
+  \grantsponsor{GS100000001}{National Science
+    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+  conclusions or recommendations expressed in this material are those
+  of the author and do not necessarily reflect the views of the
+  National Science Foundation.
+\end{acks}
+
+%% Bibliography
+%\bibliography{bibfile}
+
+%% Appendix
+\appendix
+\section{Appendix}
+
+Text of appendix \ldots
 

### 2020-06-02 main.tex
Notes

@@
 \item A formal property capturing the informal understanding of what ``stack safety'' is.
-\item Evidence that this property really does capture the folk intuition: we show that the stack
+\item Evidence that this property is strong enough to capture the folk
+intuition: we show that the stack
   unsafe examples from the literature also fail our property.
@@
 
+\section{Discussion}
+
+\bcp{Illustrate the definition by using it to discuss a range of ``stack
+  unsafety'' examples from the literature. }
+
 \section{Lazy Stack Safety}
@@
 
+\section{Future Work}
+
+We've assumed that callees always have less access than their callers, but
+in real systems things are more complicated (because of objects, static
+variables in C, software compartmentalization, coroutines, ...).  What we've
+defined is a very ``pure'' variant of stack safety.  In future work, we
+should go further. :-)
+
 %% Acknowledgments

### 2020-06-03 main.tex
Moved updates into main.tex, with a few other changes to preliminaries. Focusing on properties next.

@@
 
-    {\it Eager stack confidentiality} states that data outside a callee's stack frame is secret,
-    and does not interfere with its observable behavior or with the machine state at return. Its
-    weaker variant {\it observable stack safety} lifts the latter restriction and states that
-    the callee may not leak these secrets directly or by changing the behavior of its caller.
+    Confidentiality is in the style of noninterference, defined on a per-call basis. For each call,
+    the caller's data and the uninitialized stack do not interfere with observable behavior until return.
+    In {\it eager stack confidentiality} this means that when a function is called, the memory outside of
+    its stack frame could be scrambled without changing its observable behavior or the changes it makes
+    to the program state. This implies the weaker {\it observable stack confidentialty} in which
+    inaccessible memory could be scrambled at each call and restored on return without changing the
+    observable behavior of the remaining program.
 
@@
 
-\leo{I tried the ? notation for the potentially empty trace, which reveals that
-  we're slightly informal in the second case. Expand?}
-We define the append operation \(T \cdot T?\), where \(T?\) may be a trace or empty (\(\varepsilon\)):
-\[T \cdot T? \triangleq
-  \begin{cases}
-    T & \text{if } T? = \varepsilon \\
-    t T? & \text{if } T = t \\
-    t T' \cdot T? & \text{if } T = t T' \\
-  \end{cases}\]
+We define a few operations on traces generically.
+The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
+separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
+  \[\frac{f\ t}
+         {\mathit{split}(f,tT) = t, tT}\]
+  \[\frac{\neg f\ t \hspace{.5cm} \mathit{split}(f,T) = T_1,T_2}
+         {\mathit{split}(f,tT) = tT_1,T_2}\]
+
+The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
+of the first (if it has one) with the first of the second:
+  \[\frac{}
+         {T_1t \cdot T_2 = T_1 T_2}\]
 
@@
 observations produced by repeated application of step:
+\[\machT \in \mathit{MOTrace} = \mathit{Trace}\ (\MACHS \times \OBSS)\]
+\[\hookrightarrow : \machT \rightarrow \mathit{MOTrace}\]
 %
@@
 
-We will also wish to take a prefix of the trace from an initial state
-up through the first state where some condition holds. Let \(f\) be
-such a predicate on machine states.  Then we define \(\mach
-\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix from
-\(\mach\) until \(f\)''):\leo{Abrupt. Explain in words?}
-    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm} \mach_1
-              \hookrightarrow \machT | f}
-           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT | f}\]
-    \[\frac{f\ m}
-           {m \hookrightarrow (m,\tau)}\]
+\sna{This is a bit off, because \(f\) is a predicate on machine states, not the pairs (and later triples).
+     Will think about what the best way is to present this.}
+We will also wish to take a prefix of the trace from an initial state up through the first state where
+some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
+\(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
+  \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
+         {\mach \hookrightarrow \machT_1 | f}\]
 
@@
 \leo{I think this would read better as not exists. Thoughts?}
-\[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
+\sna{Looks good.}
+  \[\MPT \in \mathit{MPOTrace} = \mathit{Trace}\ (\MACHS \times \POLS \times \OBSS)\]
+  \[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \mathit{MPOTrace}\]
+  \[\frac{\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1}
          {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
-\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
+  \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \rightharpoonup \pol_1 \hspace{.5cm}
             \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
@@
 We also lift the prefix-up-to relation to such traces.
-\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm}
-  \neg f\ \mach_1 \hspace{.5cm} \mpstate{\mach_1}{\pol_1} \hookrightarrow_P \machT | f }
-       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT | f}\]
-
-\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm} f\ \mach_1}
-       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs}
-         \mpostate{\mach_1}{\pol_1}{\tau} | f}\]
+  \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
+         {\mpstatename \hookrightarrow \MPT_1 | f}\]
 
@@
 Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
-We then define {\em policy-sensitive observational equivalence}: a
-machine-policy-observation\leo{Call these MPO traces?} is equivalent
-to a machine-observation trace, if the MPO trace is infinite and the
-observations of MPO and M coincide, or if the MPO trace is finite
-(e.g. due to a policy fault) and its observations are a prefix of
-those of M.
+We then define {\em policy-sensitive observational equivalence}: an
+{\it MPOTrace} \(\MPT\) is equivalent to an {\it MOTrace} trace \(\machT\) if \(\MPT\)
+is infinite and the observations of \(\MPT\) and \(\machT\) coincide, or if the \(\MPT\)
+is finite (due to a policy fault) and its observations are a prefix of those of M.
 
-\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
-\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
+% Fixed by Sean
+%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
+\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \lesssim \pi_o(\machT)}
            {\MPT \sim_P \machT}\]
 
-\[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
+\[\frac{\isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \simeq \pi_o(\machT)}
            {\MPT \sim_P \machT}\]

### 2020-06-03 main.tex
More words in eager properties

@@
 
-A system satisfies integrity with respect to a callmap \(\callmap\) if
-for any initial state \(\mach_0\), call state \(\mach_c\) where
-\(\callmap(\mach_c) = n\), and policy state \(\pol_c\), and return
-state \(\mpstate{\mach_r,\pol_r}\), where: \leo{words: induced trace, call in the trace, take the prefix}
-%
-\[\mach_0,\mathit{pinit}(\mach_0) \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ... \mpstate{\mach_r}{\pol_r,\_} | \ret{\mach_c}\]
-%
-it is always the case that: \leo{words here too: the value of confidential components is unchanged}
-\[\forall \component . \contour(\component) = \HIGHINT \rightarrow \mach_c(\component) = \mach_r(\component)\]
+A system satisfies integrity with respect to a callmap \(\callmap\)
+if for any initial mpstate \(\mpstatename =
+\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each call state
+\(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from \(\mpstatename\)
+where \(\callmap(\mach_c) = n\), if we take the prefix from
+\(\mpstate{\mach_c}{\pol_c}\) until \(\ret{\mach_c}\) holds on some
+\(\mach_p\), \(\mach_p\) matches \(\mach_c\) on each component whose integrity
+in \(\mathit{Cof}(\mach_c,n)\) is \(\HIGHINT\).
 
@@
 A system enjoys eager stack confidentiality with respect to a callmap
-\(\callmap\) if, for any initial state \(\mach_0\) and call state
-\(\mach_c\) such that
-\(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}
-\hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...\) and
-\(\callmap(\mach_c) = n\):
+\(\callmap\) if, for any initial state \(\mpstatename =
+\mpstate{\mach_0}{\mathit{pinit}(\mach_0,cm)}\), for each call state
+\(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from \(\mpstatename\)
+where \(\callmap(\mach_c) = n\) and each variant \(\mach_c'\) such that
+\(\mach_c \approx_\contour \mach_c'\) for \(\contour = Cof(\mach_c, n)\),
+one of two cases holds.
 
-      \begin{itemize}
-        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
-        \item For all \(\mach_c'\) such that \(\mach_c \approx_\contour \mach_c'\):
-        \item If \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT (\mach_r,p_r,\_) | \ret{\mach_c}\)
-          \begin{itemize}
-            \item Then there is some \(\mach_r'\) where
+If the trace from the call reaches a return -- \(\mpstate{\mach_c}{\pol_c}
+\hookrightarrow_P \MPT \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\) --
+then the trace from the variant \(\mach_c'\) reaches a return --
+\(\mach_c' \hookrightarrow \machT (\mach_r',\_) | \ret{\mach_c}\).
+\(\MPT\) and \(\machT\) are observationally equivalent, \(\pi_o(\MPT) \eqsim
+\pi_o(\machT)\). For any component \(\component\) that changed between
+\(\mach_c\) and \(\mach_r\), or between \(\mach_c'\) and \(\mach_r'\),
+\(\mach_r\) agrees with \(\mach_r'\) on \(\component\).
 
-              \(\mach_c' \hookrightarrow \machT' (\mach_r',_) | \ret{\mach_c'}\)
-            \item \(\mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT')\)
-            \item And for all \(\component\), if \(\mach_c(\component) \not = \mach_r(\component)\) or
-              \(\mach_c'(\component) \not = \mach_r'(\component)\), \(\mach_r(\component) = \mach_r'(\component)\)
-          \end{itemize}
-        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
-          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
-      \end{itemize}
+If the trace from the call never reaches a return, neither does the
+trace from the variant. Then if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P
+\MPT\) and \(\mach_c' \hookrightarrow \machT\), \(\MPT \sim_P \machT\).
 

### 2020-06-05 main.tex
Words and an example of the ROP thing.

@@
 
+Confidentiality is in the style of non-interference, comparing the behavior of
+a program with an arbitrary variant that preserves low confidentiality data but
+may have any configuration of high confidentiality data. We define two machine
+states as variants given a contour if they agree on low confidentiality data:
+  \[\mach \approx_\contour \mach' \triangleq \forall \component .
+    \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
+
 A system enjoys eager stack confidentiality with respect to a callmap
@@
 
+\sna{Trying to get at why ROP is potentially a problem in non-returning calls
+     with an example.}
+  \paragraph{Non-stack Resources}
+
+    Stack integrity carries an implicit assumption that memory that has
+    become inaccessible should remain so until the correct return state, and
+    therefore that a call that never returns fulfills the requirements of
+    the property. The caller's integrity is not violated until it is able
+    to execute again and finds that its state has changed.
+
+    This assumption might be broken in any number of extended models, such as
+    the addition of heap memory or global variables. Suppose a function \(f\)
+    has some private state associated with it, with a static guarantee that
+    only its instructions access that state. A return oriented programming
+    attack from a caller could modify the private state using the caller's own
+    instructions, and even without returning, if the callee calls \(f\) the
+    modifications will be visible.
+
+    A safety property for such an extended model would extend the contour to
+    determine when the security labels of different classes of components
+    change, and integrity would mean that when a component is low
+    confidentiality, it must have the same value as the last time it was low
+    integrity. Stack integrity is a special case in which a component only
+    moves from high to low integrity on a return.
+
 \section{Lazy Stack Safety}
 
-    \paragraph{Observable Integrity}
+    \definition{Observable Integrity}
 
-      Observable integrity is concerned with the changes a callee makes to high integrity components,
-      and the effects those writes have on the observable behavior of its caller
-      {\it after its return}. Given a contour, a call state, and a returned state, we define
-      a ``rollback'' function \(\mathit{roll}_I : \CONTOURS \times \MACHS \times \MACHS
-      \rightarrow \MACHS\). The rollback creates a state that matches the call
-      state on those components that are high integrity in its contour, and matches the return state
-      on low integrity components:
+      Observable integrity weakens eager integrity by allowing a returned state
+      to differ from its call state at high integrity components, provided the
+      observable behavior of the induced trace from the return matches that of
+      an idealized state whose high integrity components do match the call
+      state.
 
-      \[\mathit{roll}_I(\contour,\mach_c,\mach_r)(\component) =
+      Given a contour, a call state, and a returned state, we define a
+      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
+      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
+      the call state on those components that are high integrity in its
+      contour, and matches the return state on low integrity components:
+      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
       \begin{cases}
@@
 
-      Intuitively, the result of a rollback is an idealized state that reflects changes to accessible
-      components, but ignores the callee's interference with the caller's data. Then the observable
-      behavior of the resulting trace is the standard to which we expect any real trace to conform.
+      Intuitively, the result of a rollback is an idealized state that
+      reflects changes to accessible components, but ignores the callee's
+      interference with the caller's data. Then the observable behavior of the
+      resulting trace is the standard to which we expect the real trace to
+      conform.
 
-      A system enjoys observable integrity with regard to \(\callmap\) as follows:
+      A system enjoys observable integrity with regard to \(\callmap\) if
+      for any initial state \(\mach_0\) and each call state \(\mpstatename =
+      \mpstate{\mach_c}{\pol_c}\) in the trace induced by
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where \(\callmap(\mach_c)
+      = n\), if there is a return state \(\mpstatename \hookrightarrow \MPT
+      \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\), the following holds.
+      Let \(\contour = \mathit{Cof}(\mach_c,n)\) describe the security levels
+      of the call and \(\mach_r' = \mathit{roll}(\contour,\mach_c,\mach_r)\).
+      The induced traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\)
+      and \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
+      possible policy failure: \(\MPT \sim_P \machT'\).
 
-      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\)
-      and \(\callmap(\mach_c) = n\), if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow ...
-        \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\):
+    \definition{Observable Confidentiality}
 
-          \begin{itemize}
-            \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
-            \item Let \(\mach_r' = \mathit{roll}_I(\contour,\mach_c,\mach_r)\), \(\mpstate{\mach_r}{\pol_r}
-              \hookrightarrow \MPT\), and \(\mach_r' \hookrightarrow \machT'\)
-            \item \(\MPT \sim_P \machT'\)
-          \end{itemize}
+      Observable confidentiality is concerned with whether high confidentiality
+      data influences observable both behavior during and after a call. As with
+      eager confidentiality, we will compare the behavior of the trace induced
+      by a state to another induced by its variant. But if the original returns,
+      before continuing, we must reset the variant to conform with the original
+      state.
 
-    \paragraph{Observable Confidentiality}
-
-      Observable confidentiality is concerned with whether high confidentiality data influences
-      observable both behavior during and after a call. We define variants given the generalized notion
-      of contours:
-      \[\mach \approx_\contour \mach' \triangleq \forall \component .
-      \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
-      \leo{Move earlier}
-
-      Data that is high confidentiality within a callee due either to being in its caller's frame
-      will become low confidentiality if the callee returns.\leo{??} We face a subtle distinction: 
-      a callee treats the contents of its caller's frame as secrets, and they should not influence
-      its behavior. Nor should the callee be able to influence behavior after its return based
-      on the caller's secrets, for instance by copying them to registers that will later be output.
-      But the caller itself may still adjust its behavior based on its own secrets. So our property
-      relies on another rollback function to preserve changes made by the caller but remove other
-      differences between variant states.
-
-      In this case, a confidentiality rollback takes a contour and four states: two call states
-      that will be variants, and two return states. It is not symmetric, regarding \(\mach_c\) and \(\mach_r\)
-      as the ``real'' call and return, \(\mach_c'\) as a variant and \(\mach_r'\) as a return that follows it.
-      \[\begin{split}
-        & \mathit{roll}_C(\contour,\mach_c,\mach_c',\mach_r,\mach_r')(\component) = \\
-        & \begin{cases}
-          \mach_c(\component) & \text{if } C(\component) = \HIGHSEC \text{ and } \\
+      The \(\mathit{restore}\) function takes a contour and four states: the
+      original call state, a variant call state, and corresponding return
+      states, and returns another state. The resulting state restores those
+      components that are high confidentiality to match the original call,
+      unless they were changed during execution:
+      \[\mathit{restore}(\contour,\mach_c,\mach_c',\mach_r,\mach_r')(\component) =
+        \begin{cases}
+          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_) \text{ and } \\
                   & \mach_c(\component) = \mach_r(\component) \text{ and }
@@
           \mach_r'(\component) & \text{else} \\
-        \end{cases}
-      \end{split}\]
-      \leo{Maybe this should be called unscramble}
+        \end{cases}\]
 
-      A system enjoys observable confidentiality with respect to \(\callmap\) as follows:
+      A system enjoys observable confidentiality with respect to \(\callmap\)
+      if for any initial mpstate \(\mpstatename =
+      \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), each call state
+      \(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from
+      \(\mpstatename\) where \(\callmap(\mach_c) = n\), and each variant
+      \(\mach_c'\) where \(\mach_c \approx_\contour \mach_c'\) and
+      \(\contour = \mathit{Cof}(\mach_c, n)\), one of two cases holds on the
+      induced traces from \(\mach_c\) and \(\mach_c'\).
 
-      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\) and
-      \(\callmap(\mach_c) = n\):
+      If \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT_1 | \ret{\mach_c}\)
+      with \(\MPT_1\) the trace within the callee ending with the machine state
+      \(\mach_r\) and policy state \(\pol_r\), the first state to return to the
+      caller, then the trace from the variant until return 
+      \(\mach_c' \hookrightarrow \machT_1 | \ret{\mach_c}\) ends with a return
+      state \(\mach_r'\). Let \(\mach_r'' = \mathit{restore}(\contour,\mach_c,
+      \mach_c',\mach_r,\mach_r')\), and let the induced traces from the
+      returned states be \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT_2\)
+      and \(\mach_r'' \hookrightarrow \machT_2\). With the callee and caller
+      traces joined, the traces are observationally equivalent up to policy
+      failures: \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1 \cdot \machT_2\).
 
-      \begin{itemize}
-        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
-        \item For all \(\mach_c'\) such that \(\mach_c \approx_C \mach_c'\):
-        \item If \(\mpstate{\mach_c}{p_c} \hookrightarrow \MPT_1 \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\)
-          \begin{itemize}
-            \item Then there is some \(\mach_r'\) where
-
-              \(\mach_c' \hookrightarrow \machT_1' (\mach_r',\_) | \ret{\mach_c'}\)
-            \item Let \(\mach_r'' = \mathit{roll}_C(\contour,\mach_c,\mach_r,\mach_c',\mach_r')\),
-
-              \(\mpstate{\mach_r}{p_r} \hookrightarrow \MPT_2\), and \(\mach_r'' \hookrightarrow \machT_2'\)
-            \item \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1' \cdot \machT_2'\)
-          \end{itemize}
-        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
-          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
-      \end{itemize}
-
-      \leo{Need to fix append. Maybe call it join. It should remove the last
-        state of left arg}
+      The case in which the induced trace does not return is the same as in the
+      eager property: the variant trace also does not return, and the induced
+      traces are observationally equivalent up to policy failure.
 

### 2020-06-05 main.tex
That memory layout took me more than I care to admit
Tweak to "prefix up to"

@@
 
+\begin{figure}
+  \begin{minipage}{\textwidth}
+  \begin{center}
+$\cdots$
+\memory{4}{gray}%
+\memory{8}{green}[{\makebox[0pt]{Caller}}]%
+\memory{4}{yellow}[{\makebox[0pt]{Args}}]%
+\memory{8}{red}[{\makebox[0pt]{Callee}}]%
+\memory{4}{gray}%
+~$\cdots$
+  \end{center}
+  \end{minipage}\\
+~\\
+~\\
+  \begin{minipage}{0.45\textwidth}
+    \begin{center}
+      Code 1
+    \end{center}
+  \end{minipage}~
+  \begin{minipage}{0.45\textwidth}
+    \begin{center}
+      Code 2
+    \end{center}
+  \end{minipage}
+  \caption{Examples of Unsafe Programs}
+  \label{fig:stackunsafety}
+\end{figure}
+
 Everybody likes stack safety. Nobody can say what it is precisely.
@@
 
-\sna{This is a bit off, because \(f\) is a predicate on machine states, not the pairs (and later triples).
-     Will think about what the best way is to present this.}
+\sna{Lambdas in the split are a bit busy, not sure anything will look better.}
 We will also wish to take a prefix of the trace from an initial state up through the first state where
@@
 \(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
-  \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
+  \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT)
+            = \machT_1,\machT_2}
          {\mach \hookrightarrow \machT_1 | f}\]
@@
 We also lift the prefix-up-to relation to such traces.
-  \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
+  \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm}
+            \mathit{split}(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT_1,\MPT_2}
          {\mpstatename \hookrightarrow \MPT_1 | f}\]
@@
 \leo{What about instructions? Shouldn't they be LC/HI?}
+\sna{We should consider whether applying stack safety to them is right, as
+     opposed to a stronger property. Code could change instructions as long as
+     they're restored before a callee returns. Shenanigans like with ROP would
+     be expected.}
 When \(\mpstate{\mach}{\pol}\) is a call state \(\callmap(\mach) = n\)
 for some \(n\), we can construct a contour based on the bounds of the
-callee's stack frame, as follows:
-\leo{Should the first two cases have a $k \not \in \REGS$?}
+callee's stack frame. Let \(\base \in \WORDS\) be the base of the stack,
+and memory below \(\base\) set aside for code.
+% Fixed. \leo{Should the first two cases have a $k \not \in \REGS$?}
 \[\mathit{Cof}(\mach,n)(\component) =
 \begin{cases}
-  (\HIGHSEC,\HIGHINT) & \text{if } \component \leq \mach(\SP) - n \\
-  (\HIGHSEC,\LOWINT) & \text{if } \component \geq \mach(\SP) \\
+  (\LOWSEC,\HIGHINT) & \text{if } \component \not \in \REGS
+                       \text{ and } \component < \base \\
+  (\HIGHSEC,\HIGHINT) & \text{if } \component \not \in \REGS
+                        \text{ and } \base \leq \component \leq \mach(\SP) - n \\
+  (\HIGHSEC,\LOWINT) & \text{if } \component \not \in \REGS
+                       \text{ and } \component \geq \mach(\SP) \\
   (\LOWSEC,\LOWINT) & \text{else} \\
 \end{cases}\]
+The code section below the stack base is low confidentiality and high
+integrity, as code should be readable but modifiable.
 %
-The private state of the callee is the entire area of the memory below
-the stack pointer, with the exception of the $n$ locations
+The private state of the callee is the entire area of the memory from stack
+base to stack pointer, with the exception of the $n$ locations
 corresponding to the arguments of a call. These components are tagged
@@
 
-\leo{Shouldn't this be a definition instead of a paragraph?}
+%\leo{Shouldn't this be a definition instead of a paragraph?}
 %\paragraph{Integrity}

### 2020-06-05 main.tex
CP - intro words

@@
 
+\newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}
+
 \begin{figure}
@@
     \begin{center}
+\MemoryLabel{10em}{1.5em}{40}
+\MemoryLabel{16em}{1.5em}{60}
+\MemoryLabel{22em}{1.5em}{80}      
 $\cdots$
 \memory{4}{gray}%
-\memory{8}{green}[{\makebox[0pt]{Caller}}]%
-\memory{4}{yellow}[{\makebox[0pt]{Args}}]%
+\memory{5}{green}[{\makebox[0pt]{Caller}}]%
+\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
 \memory{8}{red}[{\makebox[0pt]{Callee}}]%
 \memory{4}{gray}%
-~$\cdots$
+~$\cdots$\\
   \end{center}
@@
     \begin{center}
-      Code 1
+\begin{verbatim}
+jal ...
+lw r1 r0 40 
+\end{verbatim}
     \end{center}
@@
     \begin{center}
-      Code 2
+\begin{verbatim}
+jal ...
+sw r0 r0 40 
+\end{verbatim}
     \end{center}
@@
 
-Everybody likes stack safety. Nobody can say what it is precisely.
+\leo{I hate the writing here, but I wanted to make a start.}  The call
+stack is an increasingly popular target for attacks exploiting memory
+safety vulnerabilities, with potential consequences ranging from
+attackers reading or corrupting private stack data, to highjacking a
+programs control flow. To prevent such attacks, many techniques have
+been proposed in the literature\leo{cite, point out? stack canaries,
+  cherry, nicks...} aiming to achieve {\em stack safety}.
 
-The intention is to protect callers from callees.
+But what does stack safety even mean? A formal definition of stack
+safety is missing from the literature, with the notion only being
+defined negatively: through examples of stack unsafety. Two classical
+examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
+\leo{Goal: on the left, a bad read; on the right, a bad right.}
 
-Examples of what it isn't in the literature.
-Small bad examples of things that can go wrong.
+A common undertone in such examples is the intention to protect
+callers from callees: a callee should be not be able to access or
+overwrite a caller's private data. In this paper we turn this
+intuition into a formal security property, with an integrity and
+a confidentiality component.
 
-Turn this into a positive security property: integrity + confidentiality.
+Informally, {\em stack integrity} guarantees that if a function call
+returns, the caller's stack data will be identical to what it was
+before the call. {\em Stack confidentiality} ensures that a callee's
+observable behavior is not affected by memory outside its stack frame
+(e.g. a caller's private data or uninitialized memory).
 
-\leo{Variations of these should probably go here:}
-    {\it Eager stack integrity} states that if a function call returns, the caller's stack data
-    will be identical to what it was before the call. This implies the weaker {\it observable
-    stack integrity}, which states that any changes to the caller's frame will produce the same
-    observable behavior.
-
-    Confidentiality is in the style of noninterference, defined on a per-call basis. For each call,
-    the caller's data and the uninitialized stack do not interfere with observable behavior until return.
-    In {\it eager stack confidentiality} this means that when a function is called, the memory outside of
-    its stack frame could be scrambled without changing its observable behavior or the changes it makes
-    to the program state. This implies the weaker {\it observable stack confidentialty} in which
-    inaccessible memory could be scrambled at each call and restored on return without changing the
-    observable behavior of the remaining program.
-
-Challenges?
-Shifting notion of who the attacker is throughout execution.
+While integrity and confidentiality are relatively standard security
+properties, this setting provides an interesting additional challenge:
+the notion of who the attacker shifts throught execution. Not only
+does a caller need to be protected from its callee, but the callee
+itself needs to be potentially proteced from {\em its} callees.
 

### 2020-06-05 main.tex
CP

@@
 
-\leo{I hate the writing here, but I wanted to make a start.}  The call
-stack is an increasingly popular target for attacks exploiting memory
-safety vulnerabilities, with potential consequences ranging from
-attackers reading or corrupting private stack data, to highjacking a
-programs control flow. To prevent such attacks, many techniques have
-been proposed in the literature\leo{cite, point out? stack canaries,
-  cherry, nicks...} aiming to achieve {\em stack safety}.
+\leo{I hate the writing here, but I wanted to make a start.}
 
-But what does stack safety even mean? A formal definition of stack
-safety is missing from the literature, with the notion only being
-defined negatively: through examples of stack unsafety. Two classical
-examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
-\leo{Goal: on the left, a bad read; on the right, a bad right.}
+The call stack is an increasingly popular target for attacks
+exploiting memory safety vulnerabilities, with potential consequences
+ranging from attackers reading or corrupting private stack data, to
+highjacking a program's control flow. To prevent such attacks, many
+techniques have been proposed in the literature\leo{cite, point out?
+  stack canaries, cherry, nicks...} aiming to achieve {\em stack
+  safety}.
+
+But what {\em is} stack safety? A formal definition of stack safety is
+lacking from the literature, with the notion only being defined
+negatively: through examples of stack unsafety. Two classical examples
+of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
+\leo{Goal: on the left, a bad read; on the right, a bad write.}
 
@@
 does a caller need to be protected from its callee, but the callee
-itself needs to be potentially proteced from {\em its} callees.
+itself needs to be potentially proteced from {\em its} callees.  In
+this paper we address this challenge by varying the observation power
+of the attacker at call points, and demonstrate that our policy is
+implementable by pre-existing monitoring mechanisms.
 
-Roadmap: 
-Preliminaries before all these.
-
-Contributions:
+We begin by giving some background in Section~\ref{sec:prelim}.
+Our contributions are:
 \begin{itemize}
@@
   Maybe testing.
-\item TODO: Logic? Reasoning Principles?
 \end{itemize}
-
-Related work pointer + conclusions.
+Finally, in Section~\ref{sec:relwork} we discuss related work and
+sketch directions for future work in Section~\ref{sec:future}.
 
 \section{Background}
+\label{sec:prelim}
 
@@
 
+\section{Related Work}
+\label{sec:relwork}
+
 \section{Future Work}
+\label{sec:future}
 

### 2020-06-10 main.tex
Notes on things to write about
Add Well-bracketed control flow

@@
 
+\section{Well Bracketed Control Flow}
+
+  Our integrity and confidentiality properties concern themselves with data,
+  and may hold on a wide range of programs that violate expected control flow.
+  But some forms of security rely on protected control flow. In a classic
+  example, a system call that launches a missile might be gated behind code
+  that checks a password. A piece of malicious code might jump into the middle
+  of the sequence and execute the system call without the appropriate password,
+  and still obey integrity if the malicious code never returns to its caller.
+  Even with basic control flow protections, a return oriented programming
+  attack could overwrite the return address to produce the same effect.
+
+  Well bracketed control flow prevents this style of attack by guaranteeing
+  that every jump between functions is either a call or a return, and each
+  return jumps to the instruction after a corresponding call.
+
+  \paragraph*{Separating Code}
+
+    We extend our model with an explicit notion of functions and their code.
+    Each function has an identifier, and the partial function \(\mathit{own}\)
+    maps an address to the function in which it is an instruction.
+    %
+    \[\funid \in \FUNIDS\]
+    %
+    \[\own:\WORDS \rightharpoonup \FUNIDS\]
+
+  \paragraph*{Return Maps}
+
+    A {\em return map} is analogous to a call map, abstracting the recognition
+    of a state representing the end of a return sequence. In this case
+    \(\retmap \in \RETMAPS\) is a predicate on machine states.
+
+  \definition{Control Separation}
+
+    A system enjoys {\em control separation} with regard to an ownership
+    function \(\own\), a call map \(\callmap\), and a return map \(\retmap\)
+    if for any initial state \(\mach_0\) and any adjacent pair of states
+    \(\mach_1\) and \(\mach_2\) in the induced trace from
+    \(\mpstate{\mach_0}{\pinit(\callmap,\retmap,\mach_0)}\), if \(\own(\mach_1)
+    \not = \own(\mach_2)\), then either \(\callmap(\mach_1) = n\) for some
+    \(n\), or \(\retmap(\mach_1)\).
+
+  \definition{Bracketing}
+
+    A system enjoys {\em bracketing} with regard to a call map \(\callmap\) and
+    a return map \(\retmap\) if for any initial state \(\mach_0\) and any
+    call state and policy state \(\mach_c\) and \(\pol_c\) in the induced trace
+    from \(\mpstate{\mach_0}{\pinit(\callmap,\retmap,\mach_0)}\), if
+    \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT \mpostate{\mach'}
+    {\pol'}{\obs'} | \retmap\) and \(\mpstate{\mach'}{\pol'} \stepstoobs{\obs'}
+    \mpstate{\mach_r}{\pol_r}\), then \(\ret{\mach_c,\mach_r}\).
+
+  \definition{Well-bracketed Control Flow}
+
+    A system that enjoys control separation and bracketing with regard to some
+    call map, return map, and ownership function enjoys {\em well-bracketed
+    control flow} with respect to the three.
+
 \section{Related Work}

### 2020-06-10 main.tex
Forgot to commit yesterday

--- 2020-06-05 main.tex
@@
 \leo{Goal: on the left, a bad read; on the right, a bad write.}
+\leo{Do we also need an example of ROP/control flow for the new part?}
 
@@
 overwrite a caller's private data. In this paper we turn this
-intuition into a formal security property, with an integrity and
-a confidentiality component.
-
-Informally, {\em stack integrity} guarantees that if a function call
-returns, the caller's stack data will be identical to what it was
-before the call. {\em Stack confidentiality} ensures that a callee's
+intuition into a formal security property, with a confidentiality and
+an integrity component.
+%
+Informally, {\em stack confidentiality} ensures that a callee's
 observable behavior is not affected by memory outside its stack frame
-(e.g. a caller's private data or uninitialized memory).
+(e.g. a caller's private data or uninitialized memory).  {\em Stack
+  integrity} guarantees that if a function call returns, the caller's
+stack data will be identical to what it was before the call.
 
@@
 
+\leo{This will probably need to be fixed depending on what we end up
+  doing}
+%
+A second common concern is to protect the call structure itself: a
+callee should not be allowed to jump to the middle of an arbitrary
+block of code either directly (via a jump instruction) or indirectly
+(by corrupting stack data and returning). In this work we enforce
+well-bracketed control flow by ensuring that code transitions are
+valid based on a statically-known call graph of the
+program. \leo{e.g. provided by the compiler?}
+
 We begin by giving some background in Section~\ref{sec:prelim}.
@@
 \begin{itemize}
-\item A formal property capturing the informal understanding of what ``stack safety'' is.
+\item A formal property capturing the local state encaptulation
+  notion of stack safety.
 \item Evidence that this property is strong enough to capture the folk
@@
   Maybe testing.
+\item An extension to this formal property that enforces well-bracketed
+  control flow and an accompanying policy.
 \end{itemize}
@@
 
+\paragraph{Examples}
+
 \bcp{Illustrate the definition by using it to discuss a range of ``stack
@@
      with an example.}
+
+\paragraph{What about tail recursion and other non-stack-like control flow}
+
+\paragraph{What about well-bracketed control flow?}
+
+Andrew thinks it's mostly orthogonal...?
+
+\paragraph{Adding Heap Safety}
+
+Discuss how we would do it, referring to Arthur's paper.  Probably not very long or formal?
+
 \paragraph{Non-stack Resources}

### 2020-06-12 main.tex
Tex for WBCF and treatment of code in confidentiality

--- 2020-06-10 main.tex
@@
 
-\paragraph*{Call Maps and Returns}
+\paragraph*{Program Structure}
 
-A {\it call map} $\callmap$ identifies machine states that represent a
-call, abstracting away from the calling convention of any concrete
-machine. Formally $\callmap$ is a partial map from machine states to a
-natural number, the number of arguments of that call.
+  Our properties and policies are parameterized by several types of annotations
+  describing features of a program. A {\em code map} identifies some addresses
+  as code and assigns each to a function. A {\em call map} identifies addresses
+  that complete a call sequence, abstracting away from any concrete calling
+  convention. Likewise a {\em return map} identifies addresses that complete a
+  return sequence. An {\em entry map} identifies valid entry points to a
+  function.
+
+  Formally a code map is a partial function from addresses to a function
+  identifier drawn from the set \(\FUNIDS\).
   %
-\[\callmap \in \CALLMAPS = \MACHS \rightharpoonup \mathbb{N}\]
+  \[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]
+
+  A call map is a partial function from addresses to a natural number,
+  the number of arguments of the call.
+  %
+  \[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]
 
@@
   space for arguments is already allocated, but the mechanism is more
-general.
+  general. A state in which a call map is defined on the program counter is
+  referred to as a {\it call state}.
 
-A state on which a call map is defined is referred to as a {\it call
-  state}, and we consider the call to have returned the first time
-control returns to the instruction following the \(\JAL\) with the
-stack restored. Formally this is a relation between the states:
+  A return map is simply a predicate on addresses and may be considered to
+  identify the final jump of the return sequence. An entry map is likewise a
+  predicate on addresses.
+  %
+  \[\retmap \in \RETMAPS \subseteq \WORDS\]
+  \[\entmap \in \ENTMAPS \subseteq \WORDS\]
+
+\paragraph*{State Returns}
+
+  We separately identify, for any call state, what is means for a state to
+  have correctly returned from a call. We consider the call to have returned
+  the first time control returns to the instruction following the \(\JAL\) with
+  the stack restored. Formally this is a relation between the states:
 
   \[\begin{split}
-\ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) = \nextPC(\mach_c(\PCname)) \land \\
+    \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) =
+    \nextPC(\mach_c(\PCname)) \land \\
     & \mach_r(\SP) = \mach_c(\SP)
@@
 
-A stack safety property is a property of a transition system, a
-policy, and a callmap that is decomposed into an integrity property and a
+  A stack safety property is a property of a transition system, a policy, a
+  call map and a code map, that is decomposed into an integrity property and a
   confidentiality property.
@@
     \[\begin{split}
-      \textit{stack safety} \triangleq \textit{stack integrity} \land \textit{stack confidentiality} \\
+      \textit{stack safety} \triangleq \textit{stack integrity} \land
+      \textit{stack confidentiality} \\
     \end{split}\]
@@
   \leo{What about instructions? Shouldn't they be LC/HI?}
-\sna{We should consider whether applying stack safety to them is right, as
-     opposed to a stronger property. Code could change instructions as long as
-     they're restored before a callee returns. Shenanigans like with ROP would
-     be expected.}
-When \(\mpstate{\mach}{\pol}\) is a call state \(\callmap(\mach) = n\)
-for some \(n\), we can construct a contour based on the bounds of the
-callee's stack frame. Let \(\base \in \WORDS\) be the base of the stack,
-and memory below \(\base\) set aside for code.
-% Fixed. \leo{Should the first two cases have a $k \not \in \REGS$?}
-\[\mathit{Cof}(\mach,n)(\component) =
-\begin{cases}
-  (\LOWSEC,\HIGHINT) & \text{if } \component \not \in \REGS
-                       \text{ and } \component < \base \\
-  (\HIGHSEC,\HIGHINT) & \text{if } \component \not \in \REGS
-                        \text{ and } \base \leq \component \leq \mach(\SP) - n \\
-  (\HIGHSEC,\LOWINT) & \text{if } \component \not \in \REGS
-                       \text{ and } \component \geq \mach(\SP) \\
-  (\LOWSEC,\LOWINT) & \text{else} \\
-\end{cases}\]
-The code section below the stack base is low confidentiality and high
-integrity, as code should be readable but modifiable.
-%
+  \sna{Now we are dealing with this!}
+
+  When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) for some
+  \(n\) and \(\mach_c \stepstoobs{\obs} \mach_e\), we can construct a contour
+  based on the bounds of the callee's stack frame. Instructions are low
+  confidentiality if they belong to the callee and high confidentiality
+  otherwise; regardless they are always high integrity.
+
   The private state of the callee is the entire area of the memory from stack
   base to stack pointer, with the exception of the $n$ locations
-corresponding to the arguments of a call. These components are tagged
+  corresponding to the arguments of a call. Private components are marked
   high confidentiality and high integrity: the callee should be able to
   neither read, nor write them.
-%
+
   Everything above the stack pointer at the point of the call is
@@
   are marked high confidentiality and low integrity.
-%
+
   Registers and the frame of arguments to the callee are free for the
@@
 
-%\leo{Shouldn't this be a definition instead of a paragraph?}
-%\paragraph{Integrity}
+ \[\mathit{Cof}(\mach,\funid,n)(\component) =
+  \begin{cases}
+    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
+                         \text{ and } \codemap(\component) = \codemap(\mach(\PCname)) \\
+    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS
+                          \text{ and } \codemap(\component) = \funid \not = \codemap(\mach(\PCname)) \\
+    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
+                          \text{ and } \component \leq \mach(\SP) - n \\
+    (\HIGHSEC,\LOWINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
+                         \text{ and } \component \geq \mach(\SP) \\
+    (\LOWSEC,\LOWINT) & \text{else} \\
+  \end{cases}\]
 
@@
 
-A system satisfies integrity with respect to a callmap \(\callmap\)
-if for any initial mpstate \(\mpstatename =
-\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each call state
-\(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from \(\mpstatename\)
-where \(\callmap(\mach_c) = n\), if we take the prefix from
-\(\mpstate{\mach_c}{\pol_c}\) until \(\ret{\mach_c}\) holds on some
-\(\mach_p\), \(\mach_p\) matches \(\mach_c\) on each component whose integrity
-in \(\mathit{Cof}(\mach_c,n)\) is \(\HIGHINT\).
+  A system satisfies eager stack integrity with respect to a call map
+  \(\callmap\) and code map \(\codemap\) if for any initial mpstate
+  \(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each
+  adjacent pair of a call state \(\mach_c\) and state \(\mach_e\) with policy
+  state \(\pol_e\) in the induced trace from \(\mpstatename\), where
+  \(\callmap(\mach_c(\PCname)) = n\) and \(\codemap(\mach_e(\PCname)) =
+  \funid\), if we take the prefix until a return
+  \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_c}\) and
+  \(\MPT\) ends with some \(\mach_p\), \(\mach_p\) matches \(\mach_c\) on each
+  component whose integrity in \(\mathit{Cof}(\mach_c,\funid,n)\) is
+  \(\HIGHINT\).
 
@@
 
-Confidentiality is in the style of non-interference, comparing the behavior of
-a program with an arbitrary variant that preserves low confidentiality data but
-may have any configuration of high confidentiality data. We define two machine
-states as variants given a contour if they agree on low confidentiality data:
+  Confidentiality is in the style of non-interference, comparing the behavior
+  of a program with an arbitrary variant that preserves low confidentiality data
+  but may have any configuration of high confidentiality data. We define two
+  machine states as variants given a contour if they agree on low
+  confidentiality data:
   \[\mach \approx_\contour \mach' \triangleq \forall \component .
-    \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
+    \contour(\component) = \LOWSEC \rightarrow \mach(\component) =
+    \mach'(\component)\]
 
-A system enjoys eager stack confidentiality with respect to a callmap
-\(\callmap\) if, for any initial state \(\mpstatename =
-\mpstate{\mach_0}{\mathit{pinit}(\mach_0,cm)}\), for each call state
-\(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from \(\mpstatename\)
-where \(\callmap(\mach_c) = n\) and each variant \(\mach_c'\) such that
-\(\mach_c \approx_\contour \mach_c'\) for \(\contour = Cof(\mach_c, n)\),
-one of two cases holds.
+  A system satisfies eager stack confidentiality with respect to a call map
+  \(\callmap\) and code map \(\codemap\) if for any initial mpstate
+  \(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each
+  adjacent pair of a call state \(\mach_c\) and state \(\mach_e\) with policy
+  state \(\pol_e\) in the induced trace from \(\mpstatename\), where
+  \(\callmap(\mach_c(\PCname)) = n\) and \(\codemap(\mach_e(\PCname)) =
+  \funid\), one of two cases holds on the induced traces from
+  \(\mpstate{\mach_e}{\pol_e}\) and any variation \(\mach_e'\) with respect
+  to the contour \(\mathit{Cof}(\mach_c,\funid,n)\).
 
-If the trace from the call reaches a return -- \(\mpstate{\mach_c}{\pol_c}
-\hookrightarrow_P \MPT \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\) --
-then the trace from the variant \(\mach_c'\) reaches a return --
-\(\mach_c' \hookrightarrow \machT (\mach_r',\_) | \ret{\mach_c}\).
-\(\MPT\) and \(\machT\) are observationally equivalent, \(\pi_o(\MPT) \eqsim
-\pi_o(\machT)\). For any component \(\component\) that changed between
-\(\mach_c\) and \(\mach_r\), or between \(\mach_c'\) and \(\mach_r'\),
+  If the trace from the call reaches a return -- \(\mpstate{\mach_e}{\pol_e}
+  \hookrightarrow_P \MPT | \ret{\mach_c}\) -- then the trace from the variant
+  \(\mach_e'\) reaches a return -- \(\mach_e' \hookrightarrow \machT |
+  \ret{\mach_c}\). Let \(\mach_r\) and \(\mach_r'\) be the final states of
+  \(\pi_m(\MPT)\) and \(\machT\), respectively. Until their final states,
+  \(\MPT\) and \(\machT\) must be observationally equivalent, \(\pi_o(\MPT)
+  \cdot \tau \eqsim \pi_o(\machT) \cdot \tau\). For any component
+  \(\component\) where \(\mach_e(\component) \not = \mach_r(\component)\) or
+  \(\mach_e'(\component) \not = \mach_r'(\component)\),
   \(\mach_r\) agrees with \(\mach_r'\) on \(\component\).
@@
   \leo{Think more about whether we vary mc or step mc}
-      \leo{Also, words}
+  \sna{With code, it must be step mc}
 
@@
 
-\sna{Trying to get at why ROP is potentially a problem in non-returning calls
-     with an example.}
-
   \paragraph{What about tail recursion and other non-stack-like control flow}
 
-\paragraph{What about well-bracketed control flow?}
+  \paragraph{Control Hijacking and Well-bracketed Control Flow}
 
-Andrew thinks it's mostly orthogonal...?
+    Some forms of security rely on protected control flow. In a classic
+    example, a system call that launches a missile might be gated behind code
+    that checks a password. A piece of malicious code might jump into the middle
+    of the sequence and execute the system call without the appropriate password,
+    and still obey integrity if the malicious code never returns to its caller.
+    Integrity here protects data, not privileges. But such a jump does violate
+    eager stack confidentiality if the attacker jumps beyond its own function's
+    instructions without performing a designated call or returning to an
+    appropriate site.
+
+    A jump to a function's own code does not violate either, even if done using
+    nefarious means such as return-oriented programming (ROP). Our model would
+    permit such a jump within the function, so privilege is not elevated by
+    permitting ROP in this case.
+
+    But, our policies do enforce orthogonal control flow properties that
+    protect from calls that enter outside of valid entry points and from all
+    invalid returns, even if they would not escalate privilege. Most notably
+    {\em well-bracketed control flow}, below.
 
@@
     Discuss how we would do it, referring to Arthur's paper.  Probably not very long or formal?
-
-\paragraph{Non-stack Resources}
-
-Stack integrity carries an implicit assumption that memory that has
-become inaccessible should remain so until the correct return state, and
-therefore that a call that never returns fulfills the requirements of
-the property. The caller's integrity is not violated until it is able
-to execute again and finds that its state has changed.
-
-This assumption might be broken in any number of extended models, such as
-the addition of heap memory or global variables. Suppose a function \(f\)
-has some private state associated with it, with a static guarantee that
-only its instructions access that state. A return oriented programming
-attack from a caller could modify the private state using the caller's own
-instructions, and even without returning, if the callee calls \(f\) the
-modifications will be visible.
-
-A safety property for such an extended model would extend the contour to
-determine when the security labels of different classes of components
-change, and integrity would mean that when a component is low
-confidentiality, it must have the same value as the last time it was low
-integrity. Stack integrity is a special case in which a component only
-moves from high to low integrity on a return.
 
@@
 
-      A system enjoys observable integrity with regard to \(\callmap\) if
-      for any initial state \(\mach_0\) and each call state \(\mpstatename =
-      \mpstate{\mach_c}{\pol_c}\) in the trace induced by
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where \(\callmap(\mach_c)
-      = n\), if there is a return state \(\mpstatename \hookrightarrow \MPT
-      \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\), the following holds.
-      Let \(\contour = \mathit{Cof}(\mach_c,n)\) describe the security levels
-      of the call and \(\mach_r' = \mathit{roll}(\contour,\mach_c,\mach_r)\).
-      The induced traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\)
-      and \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
+      A system enjoys observable integrity with regard to \(\callmap\) and 
+      \(\codemap\) if for any initial state \(\mach_0\) and each adjacent pair
+      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
+      \(\pol_e\) in the trace induced by
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where
+      \(\callmap(\mach_c) = n\) and \(\codemap(\mach_e) = \funid\),
+      if \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
+      with the return state \(\mach_r\) at the end of \(\MPT\), then the
+      following holds.
+
+      Let \(\contour = \mathit{Cof}(\mach_c,\funid,n)\) describe the security
+      levels of the call and
+      \(\mach_r' = \mathit{roll}(\contour,\mach_e,\mach_r)\). The induced
+      traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\) and
+      \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
       possible policy failure: \(\MPT \sim_P \machT'\).
@@
       A system enjoys observable confidentiality with respect to \(\callmap\)
-      if for any initial mpstate \(\mpstatename =
-      \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), each call state
-      \(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from
-      \(\mpstatename\) where \(\callmap(\mach_c) = n\), and each variant
-      \(\mach_c'\) where \(\mach_c \approx_\contour \mach_c'\) and
-      \(\contour = \mathit{Cof}(\mach_c, n)\), one of two cases holds on the
-      induced traces from \(\mach_c\) and \(\mach_c'\).
+      and \(\codemap\) if for any initial state \(\mach_0\), each adjacent pair
+      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
+      \(\pol_e\) in the inducted trace from
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), and each variant
+      \(\mach_e' \approx_\contour \mach_e'\) where
+      \(\contour = \mathit{Cof}(\mach_c, \funid, n)\), one of two cases holds
+      on the induced traces from \(\mach_e\) and \(\mach_e'\).
 
-      If \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT_1 | \ret{\mach_c}\)
+      If \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_1 | \ret{\mach_c}\)
       with \(\MPT_1\) the trace within the callee ending with the machine state
-      \(\mach_r\) and policy state \(\pol_r\), the first state to return to the
-      caller, then the trace from the variant until return
-      \(\mach_c' \hookrightarrow \machT_1 | \ret{\mach_c}\) ends with a return
-      state \(\mach_r'\). Let \(\mach_r'' = \mathit{restore}(\contour,\mach_c,
-      \mach_c',\mach_r,\mach_r')\), and let the induced traces from the
+      \(\mach_r\), the first state to return to the caller, and policy state
+      \(\pol_r\), then the there is a return state \(\mach_r'\) reached by the
+      trace from the variant, \(\mach_c' \hookrightarrow \machT_1 |
+      \ret{\mach_c}\). Let \(\mach_r'' = \mathit{restore}(\contour,\mach_e,
+      \mach_e',\mach_r,\mach_r')\), and let the induced traces from the
       returned states be \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT_2\)
@@
       traces are observationally equivalent up to policy failure.
+
+    \definition{Observable Stack Safety}
+
+      A system enjoys {\em observable stack safety} with regard to a call map
+      and a code map if it enjoys both observable integrity and observable
+      confidentiality with regard to them.
+
+\section{Well-bracketed Control Flow}
+
+  Our integrity and confidentiality properties concern themselves with data,
+  and may hold on a wide range of programs that violate expected control flow.
+  But well-bracketed control flow prevents this style of attack by guaranteeing
+  that every jump between functions is either a call or a return, and each
+  return jumps to the instruction after a corresponding call.
+
+  \definition{Control Separation}
+
+    A system enjoys {\em control separation} with regard to a code map
+    \(\codemap\), a call map \(\callmap\), and a return map \(\retmap\)
+    if for any initial state \(\mach_0\) and any adjacent pair of states
+    \(\mach_1\) and \(\mach_2\) in the induced trace from
+    \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if \(\codemap(\mach_1(\PCname))
+    \not = \codemap(\mach_2)\), then either \(\callmap(\mach_1(\PCname)) = n\)
+    for some \(n\), or \(\retmap(\mach_1)\).
+
+  \definition{Entry Integrity}
+
+    A system enjoys {\em entry integrity} with regard to a call map
+    \(\callmap\) and entry map \(\entmap\) if for any initial state
+    \(\mach_0\) and any adjacent pair of states \(\mach_c\) and \(\mach_e\)
+    in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
+    \(\callmap(\mach_c(\PCname)) = n\) then \(\entmap(\mach_e)\).
+
+  \definition{Well-bracketed Control Flow}
+
+    Well-bracketed control flow captures the expected behavior of marked
+    returns with regard to their call points. The coinductive {\em bracketing}
+    predicate, \(\mathit{BR}(\machT)\), is parameterized by a call map
+    \(\callmap\) and a return map \(\retmap\):
+
+    \[\frac{\mathit{BR}(\machT) \hspace{.5cm} \callmap(\mach_c(\PCname)) = n
+              \hspace{.5cm} \retmap(\mach_c(\PCname))}
+           {\mathit{BR}(\mach_c \machT \mach_r)}\]
+
+    \[\frac{\mathit{BR}(\machT) \hspace{.5cm} \callmap(\mach(\PCname)) = \bot}
+           {\mathit{BR}(\mach \machT)}\]
+
+    A system enjoys {\em well-bracketed control flow} with regard to a call map
+    \(\callmap\) and return map \(\retmap\) if for any initial state
+    \(\mach_0\) and any call state and policy state \(\mach_c\) and \(\pol_c\)
+    in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
+    \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT \mpostate{\mach'}
+    {\pol'}{\obs'} ...\) and \(\MPT\) is bracketed with regard to \(\callmap\)
+    and \(\retmap\), then \(\ret{\mach_c,\mach'}\).
+
+  \definition{Standard Stack Control Flow}
+
+    The above properties are somewhat orthogonal, but together they capture a
+    standard notion of control flow for a stack: functions keep their code
+    separate outside of calls and returns, calls respect explicit entry points,
+    and returns respect implicit entry points corresponding to their call.
+    A system that enjoys them with regard to a code, a call, a return, and an
+    entry map is said to enjoy {\em standard stack control flow} with regard to
+    those maps.
 

### 2020-06-13 main.tex
Updated WBCF in tex and coq

@@
 
-  \definition{Well-bracketed Control Flow}
+  \definition{Return Integrity}
 
-    Well-bracketed control flow captures the expected behavior of marked
-    returns with regard to their call points. The coinductive {\em bracketing}
-    predicate, \(\mathit{BR}(\machT)\), is parameterized by a call map
-    \(\callmap\) and a return map \(\retmap\):
+    Return integrity captures the expected behavior of marked returns with
+    regard to their call points. We begin by defining a coinductive relation
+    \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1 \MPT_2\). Given a call map
+    \(\callmap\) and return map \(\retmap\), \(\MPT\) is related to
+    \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\) whose head is the first
+    unmatched return in \(\MPT\). So if \(\MPT\) starts with a return,
+    it is related to itself:
+%
+    \[\frac{\callmap(\mach(\PCname)) = \bot
+              \hspace{.5cm} \retmap(\mach(\PCname))}
+           {\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\ 
+              \mpstate{\mach}{\pol}}\]
+    \[\frac{\callmap(\mach(\PCname)) = \bot
+              \hspace{.5cm} \retmap(\mach(\PCname))}
+           {\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\ 
+              \mpstate{\mach}{\pol}\MPT}\]
+%
+    If \(\MPT\) starts with a non-call, non-return state, its unmatched
+    return must come later:
+%
+    \[\frac{\begin{split}
+            & \callmap(\mach(\PCname)) = \bot
+              \hspace{.5cm} \neg \retmap(\mach(\PCname)) \\
+            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'
+            \end{split}}
+           {\mathit{Unmatched}\ \callmap\ \retmap\ 
+            \mpstate{\mach}{\pol}\MPT\ \MPT'}\]
+%   
+    And most intricately, if \(\MPT\) starts with a call, we must find the
+    return that matches it, which will be the unmatched return once we step
+    past the call. Then the first unmatched return beyond it is the true
+    unmatched return.
+    \[\frac{\begin{split}
+            & \callmap(\mach(\PCname)) = n
+              \hspace{.5cm} \neg \retmap(\mach(\PCname)) \\
+            & \MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}} \\
+            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}} \\
+            & \MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}} \\
+            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}} \\
+            \end{split}}
+            {\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}}\]
 
-    \[\frac{\mathit{BR}(\machT) \hspace{.5cm} \callmap(\mach_c(\PCname)) = n
-              \hspace{.5cm} \retmap(\mach_c(\PCname))}
-           {\mathit{BR}(\mach_c \machT \mach_r)}\]
-
-    \[\frac{\mathit{BR}(\machT) \hspace{.5cm} \callmap(\mach(\PCname)) = \bot}
-           {\mathit{BR}(\mach \machT)}\]
-
-    A system enjoys {\em well-bracketed control flow} with regard to a call map
+    A system enjoys {\em return integrity} with regard to a call map
     \(\callmap\) and return map \(\retmap\) if for any initial state
@@
     in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
-    \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT \mpostate{\mach'}
-    {\pol'}{\obs'} ...\) and \(\MPT\) is bracketed with regard to \(\callmap\)
-    and \(\retmap\), then \(\ret{\mach_c,\mach'}\).
+    \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\) is the
+    first unmatched return, with \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT
+    \ \MPT'\) and \(\mach_r\) the first machine state of \(\MPT'\),
+    then \(\ret{\mach_c}\ \mach_r\).
 
-  \definition{Standard Stack Control Flow}
+  \definition{Well Bracketed Control Flow}
 
@@
     A system that enjoys them with regard to a code, a call, a return, and an
-    entry map is said to enjoy {\em standard stack control flow} with regard to
+    entry map is said to enjoy {\em well bracketed control flow} with regard to
     those maps.

### 2020-06-15 main.tex
Cleaned up judgments and wrote very briefly about heap safety

@@
 
-    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} (\mach_1, \pol_1) \stepstopol \pol_2}
-           {\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}}\]
+    \judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1) \stepstopol \pol_2\)}
+           {\(\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}\)}
+
 \leo{Notation: Should $\stepstoobs{}$ be partial?}
@@
 separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
-  \[\frac{f\ t}
-         {\mathit{split}(f,tT) = t, tT}\]
-  \[\frac{\neg f\ t \hspace{.5cm} \mathit{split}(f,T) = T_1,T_2}
-         {\mathit{split}(f,tT) = tT_1,T_2}\]
+
+  \judgment{\(f\ t\)}
+           {\(\mathit{split}(f,tT) = t, tT\)}
+
+  \judgmenttwo{\(\neg f\ t\)}{\(\mathit{split}(f,T) = T_1,T_2\)}
+              {\(\mathit{split}(f,tT) = tT_1,T_2\)}
 
@@
 of the first (if it has one) with the first of the second:
-  \[\frac{}
-         {T_1t \cdot T_2 = T_1 T_2}\]
+
+  \judgment{}
+           {\(T_1t \cdot T_2 = T_1 T_2\)}
 
@@
 \[\hookrightarrow : \machT \rightarrow \mathit{MOTrace}\]
-%
-\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
-       {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]
-%
+
+\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
+            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}
+
 Since the step function is total, if \(\mach \hookrightarrow \machT\),
@@
 \(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
-  \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT)
-            = \machT_1,\machT_2}
-         {\mach \hookrightarrow \machT_1 | f}\]
+
+  \judgmenttwo{\(\mach \hookrightarrow \machT\)}
+              {\(\mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT) = \machT_1,\machT_2\)}
+              {\(\mach \hookrightarrow \machT_1 | f\)}
 
@@
 and observations induced by by \(\stepstoobs{}_P\):
-%
+
 \leo{I think this would read better as not exists. Thoughts?}
@@
   \[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \mathit{MPOTrace}\]
-  \[\frac{\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1}
-         {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
-  \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \rightharpoonup \pol_1 \hspace{.5cm}
-            \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
-         {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]
+
+  \judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
+           {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
+
+  \judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
+                {\(\pol_0 \rightharpoonup \pol_1\)}
+                {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
+      {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
 
 We also lift the prefix-up-to relation to such traces.
-  \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm}
-            \mathit{split}(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT_1,\MPT_2}
-         {\mpstatename \hookrightarrow \MPT_1 | f}\]
+
+  \judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
+              {\(\mathit{split}(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT_1,\MPT_2\)}
+         {\(\mpstatename \hookrightarrow \MPT_1 | f\)}
 
@@
 
-\[\frac{} {\obsT \lesssim \obsT}\]
+\judgment{}{\(\obsT \lesssim \obsT\)}
 
-\[\frac{} {\tau \lesssim \obsT}\]
+\judgment{}{\(\tau \lesssim \obsT\)}
 
-\[\frac{\obsT_1 = \tau \obsT_1' \hspace{.5cm} \obsT_1' \lesssim \obsT_2}
-       {\obsT_1 \lesssim \obsT_2}\]
+\judgmenttwo{\(\obsT_1 = \tau \obsT_1'\)}{\(\obsT_1' \lesssim \obsT_2\)}
+            {\(\obsT_1 \lesssim \obsT_2\)}
 
-\[\frac{\obsT_2 = \tau \obsT_2' \hspace{.5cm} \obsT_1 \lesssim \obsT_2'}
-       {\obsT_1 \lesssim \obsT_2}\]
+\judgmenttwo{\(\obsT_2 = \tau \obsT_2'\)}{\(\obsT_1 \lesssim \obsT_2'\)}
+            {\(\obsT_1 \lesssim \obsT_2\)}
 
-\[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
-       {\obsT_1 \lesssim \obsT_2}\]
+\judgmentthree{\(\obsT_1 = w \obsT_1'\)}
+              {\(\obsT_2 = w \obsT_2'\)}
+              {\(\obsT_1' \lesssim \obsT_2'\)}
+              {\(\obsT_1 \lesssim \obsT_2\)}
 
@@
 %\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
-\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \lesssim \pi_o(\machT)}
-           {\MPT \sim_P \machT}\]
+\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
+            {\(\MPT \sim_P \machT\)}
 
-\[\frac{\isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \simeq \pi_o(\machT)}
-           {\MPT \sim_P \machT}\]
+\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
+            {\(\MPT \sim_P \machT\)}
 
@@
     (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
-                         \text{ and } \codemap(\component) = \codemap(\mach(\PCname)) \\
-    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS
-                          \text{ and } \codemap(\component) = \funid \not = \codemap(\mach(\PCname)) \\
-    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
+                         \text{ and for some } \funid, \codemap(\component) = \funid \\
+    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                           \text{ and } \component \leq \mach(\SP) - n \\
-    (\HIGHSEC,\LOWINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
+    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \geq \mach(\SP) \\
@@
   \leo{Think more about whether we vary mc or step mc}
-  \sna{With code, it must be step mc}
+  \sna{I've left it at step mc for now}
 
@@
     example, a system call that launches a missile might be gated behind code
-    that checks a password. A piece of malicious code might jump into the middle
+    that checks a password. A piece of malicious code can jump into the middle
     of the sequence and execute the system call without the appropriate password,
@@
 
-    Discuss how we would do it, referring to Arthur's paper.  Probably not very long or formal?
+    While these ideas are here presented in a stack-centric style, they could
+    be extended to heap safety in a system extended with a heap. Heap safety,
+    like stack safety, may be framed as a form of noninterference
+    \cite{AzevedoDeAmorim17}. As in stack safety, the mapping from components
+    to security labels would be dynamic, with any heap data unreachable from
+    the stack frame in a given state treated as high confidentiality and high
+    integrity. This would require updating the contour at each step.
 
@@
     it is related to itself:
-%
-    \[\frac{\callmap(\mach(\PCname)) = \bot
-              \hspace{.5cm} \retmap(\mach(\PCname))}
-           {\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\ 
-              \mpstate{\mach}{\pol}}\]
-    \[\frac{\callmap(\mach(\PCname)) = \bot
-              \hspace{.5cm} \retmap(\mach(\PCname))}
-           {\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\ 
-              \mpstate{\mach}{\pol}\MPT}\]
-%
+
+  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
+           {\(\retmap(\mach(\PCname))\)}
+      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\ 
+        \mpstate{\mach}{\pol}\)}
+  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
+              {\(\retmap(\mach(\PCname))\)}
+      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\ 
+              \mpstate{\mach}{\pol}\MPT\)}
+
     If \(\MPT\) starts with a non-call, non-return state, its unmatched
     return must come later:
-%
-    \[\frac{\begin{split}
-            & \callmap(\mach(\PCname)) = \bot
-              \hspace{.5cm} \neg \retmap(\mach(\PCname)) \\
-            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'
-            \end{split}}
-           {\mathit{Unmatched}\ \callmap\ \retmap\ 
-            \mpstate{\mach}{\pol}\MPT\ \MPT'}\]
-%   
+
+  \judgment{\(\callmap(\mach(\PCname)) = \bot
+              \hspace{.5cm} \neg \retmap(\mach(\PCname))
+
+              \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'\)}
+           {\(\mathit{Unmatched}\ \callmap\ \retmap\ 
+            \mpstate{\mach}{\pol}\MPT\ \MPT'\)}
+
     And most intricately, if \(\MPT\) starts with a call, we must find the
@@
     unmatched return.
-    \[\frac{\begin{split}
-            & \callmap(\mach(\PCname)) = n
-              \hspace{.5cm} \neg \retmap(\mach(\PCname)) \\
-            & \MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}} \\
-            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}} \\
-            & \MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}} \\
-            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}} \\
-            \end{split}}
-            {\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}}\]
+
+    \judgment{\(\callmap(\mach(\PCname)) = n
+              \hspace{.5cm} \neg \retmap(\mach(\PCname))\)
+
+              \(\MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}}\)
+
+              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}}\)
+
+              \(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\)
+
+              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}}\)}
+            {\(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}\)}
 

### 2020-06-22 main.tex
Skipping broken citation for now

@@
 \leo{Do we also need an example of ROP/control flow for the new part?}
+\leo{Stack tokens introduce two properties: lse + wbcf, high level
+  semantics, refinement to a machine with ``builtin'' ``by construction''
+  (pretty weak :)) We make them formal and extensional}
 
@@
 \begin{itemize}
-\item A formal property capturing the local state encaptulation
-  notion of stack safety.
-\item Evidence that this property is strong enough to capture the folk
-intuition: we show that the stack
-  unsafe examples from the literature also fail our property.
+\item \leo{More? Extensional} formal characterizations of the intuitive concepts
+  of local state encapsualtion \leo{ref} and well-bracketed control flow \leo{ref}.
+  Evidence that this property is strong enough to capture the folk
+  intuition: we show that the stack unsafe examples from the literature also fail
+  our property.
 \item Evidence that the property is not too strong. It's realizible by
@@
     says yes on a substantial body of code}
-\item A (inductive? step-by-step/stepwise) variant of this property
-  tuned for testing/verification together with a proof that it implies
-  the original property.  We test the whole thing with a cool
-  framework.
-\item A lazy version of the property that allows for more efficient/dynamic enforcement.
-  Maybe testing.
-\item An extension to this formal property that enforces well-bracketed
-  control flow and an accompanying policy.
+  We then present a variation of the property + policy that is lazy,
+  easier/more efficient to implement, and provably weaker.
+\item A provably stronger (inductive? step-by-step/stepwise) variant
+  of this property tuned for testing/verification.  We test the whole
+  thing with a cool framework.
 \end{itemize}
@@
     like stack safety, may be framed as a form of noninterference
-    \cite{AzevedoDeAmorim17}. As in stack safety, the mapping from components
-    to security labels would be dynamic, with any heap data unreachable from
-    the stack frame in a given state treated as high confidentiality and high
-    integrity. This would require updating the contour at each step.
+    [cite Meaning of Memory Safety]. As in stack safety, the mapping from
+    components to security labels would be dynamic, with any heap data
+    unreachable from the stack frame in a given state treated as high
+    confidentiality and high integrity. This would require updating the contour
+    at each step.
 
@@
 %% Bibliography
-%\bibliography{bibfile}
+%\bibliography{bcp.bib}
 

### 2020-06-24 main.tex
Better explanation (still brief) of how heap safety should work

@@
     like stack safety, may be framed as a form of noninterference
-    [cite Meaning of Memory Safety]. As in stack safety, the mapping from
-    components to security labels would be dynamic, with any heap data
-    unreachable from the stack frame in a given state treated as high
-    confidentiality and high integrity. This would require updating the contour
-    at each step.
+    [cite Meaning of Memory Safety]. For each certain component marked as
+    containing a valid pointer to a heap object, we can propagate that
+    validity when the pointer moves or is used in pointer arithmetic. Heap
+    objects to which a valid pointer is in a register are reachable, and
+    treated as low confidentiality and low integrity.
 

### 2020-06-24 main.tex
Revamp intro and background

@@
 
-But what {\em is} stack safety? A formal definition of stack safety is
-lacking from the literature, with the notion only being defined
-negatively: through examples of stack unsafety. Two classical examples
-of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
+But what {\em is} stack safety? A precise formal definition of stack
+safety is lacking from the literature, with the notion only being
+defined negatively: through examples of stack unsafety. Two classical
+examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
 \leo{Goal: on the left, a bad read; on the right, a bad write.}
 \leo{Do we also need an example of ROP/control flow for the new part?}
-\leo{Stack tokens introduce two properties: lse + wbcf, high level
-  semantics, refinement to a machine with ``builtin'' ``by construction''
-  (pretty weak :)) We make them formal and extensional}
+Recently, a \leo{inspiring/awesome/something good} work reasoned
+about a stack safety enforcement mechanism using capabilities\leo{cite};
+while impressive, stack safety was defined with respect to an abstract
+machine that ``exhibits stack safety by construction''. Which begs
+the question, what {\em is} the property that holds by construction?
 
-A common undertone in such examples is the intention to protect
-callers from callees: a callee should be not be able to access or
-overwrite a caller's private data. In this paper we turn this
-intuition into a formal security property, with a confidentiality and
-an integrity component.
+\leo{StackTokens} provides some intuition towards answering that question,
+by informally describing two components that comprise stack safety:
+local state encapsulation and well-bracketed control flow. In this paper
+we turn this intuition into two formal \leo{extensional} security properties,
+that together form a formal definition of stack safety.
+
+The underlying notion behind local state encapsulation is the
+intention to protect callers from callees: a callee should be not be
+able to access or overwrite a caller's private data. We formalize this
+intuition as a security property, with a confidentiality and an
+integrity component.
 %
-Informally, {\em stack confidentiality} ensures that a callee's
+At a high level, {\em stack confidentiality} ensures that a callee's
 observable behavior is not affected by memory outside its stack frame
@@
 
-\leo{This will probably need to be fixed depending on what we end up
-  doing}
-%
-A second common concern is to protect the call structure itself: a
-callee should not be allowed to jump to the middle of an arbitrary
-block of code either directly (via a jump instruction) or indirectly
-(by corrupting stack data and returning). In this work we enforce
-well-bracketed control flow by ensuring that code transitions are
-valid based on a statically-known call graph of the
-program. \leo{e.g. provided by the compiler?}
+Similarly, the underlying notion behind well-bracketed control flow is
+the need to protect the call structure itself: a callee should not be
+allowed to jump to the middle of an arbitrary block of code either
+directly (via a jump instruction) or indirectly (by corrupting stack
+data and returning). \leo{There needs to be some informal description
+  of the wbcf part here. Sean?}
 
@@
 \begin{itemize}
-\item \leo{More? Extensional} formal characterizations of the intuitive concepts
-  of local state encapsualtion \leo{ref} and well-bracketed control flow \leo{ref}.
-  Evidence that this property is strong enough to capture the folk
-  intuition: we show that the stack unsafe examples from the literature also fail
-  our property.
+\item \leo{More? Extensional} formal characterizations of the
+  intuitive concepts of local state encapsulation~\ref{sec:lse} and
+  well-bracketed control flow~\ref{sec:wbcf}.  Evidence that this
+  property is strong enough to capture the folk intuition: we show
+  that the stack unsafe examples from the literature also fail our
+  property.
 \item Evidence that the property is not too strong. It's realizible by
-  a previously proposed  micropolicy enforcement mechanism \leo{And it
-    says yes on a substantial body of code}
-  We then present a variation of the property + policy that is lazy,
-  easier/more efficient to implement, and provably weaker.
+  a previously proposed micropolicy enforcement
+  mechanism~\ref{sec:micropolicy}. \leo{And it says yes on a
+    substantial body of code} We then present a variation of the
+  property + policy that is lazy, easier/more efficient to implement,
+  and provably weaker.
 \item A provably stronger (inductive? step-by-step/stepwise) variant
@@
 
-We begin by describing our machine model. Concretely, we have in mind a
+Before we define stack safety, we need to first set the stage by
+describing our machine model. Concretely, we have in mind a
 conventional machine architecture with a standard (RISC-V) processor
 extended with some kind of (static or dynamic, software- or
-hardware-enforced) security monitor, but we won't need to make any
-specific choices about details until we get to \bcp{the policy one, probably}.
-%In this section we will describe an abstract interface to our machine model.
+hardware-enforced) security monitor. However, our definition of stack
+safety is generic: we will begin with an abstract interface to our
+machine model and only make specific choices about details when we get
+to the policy enforcement section~\ref{sec:micropolicy}.
 
-Values and addresses are words drawn from some set of all words $\WORDS$. We will
-denote words with $\word$.\leo{TODO: If we ever need ``A'' come back to this.}
+\paragraph*{Values, States, and Observations}
 
-Components are words (denoting addresses) and registers.
+The basic building blocks of the machine are {\em values} and {\em
+  addresses}. Both are drawn from some set of all possible words
+$\WORDS$. We will denote words with $\word$.
+%
+Our machine states are comprised of {\em components}, that can be word
+(denoting memory addresses) or registers (drawn from some set $\REGS$.
 %
@@
 counter \PCname and the stack pointer \SP.
-
-A machine state is just a map from components to values:
+%
+Formally, a machine state is just a map from components to values:
 %
@@
 
-Our stack safety property will be defined in terms of {\em observations}, which may be
-silent (written $\tau$) or a word of data:
+Our stack safety property will be defined in terms of {\em
+  observations}, which may be silent (written $\tau$) or a word of
+data:\leo{Needs more exposition}
 %
 \[\obs \in \OBSS = \tau ~ | ~ \word \in \WORDS\]
-
+%
 These observations are a byproduct of the step function of the
@@
 
-\paragraph*{Policy}
+\paragraph*{Policies}
 
 A policy is a general model of an enforcement mechanism, consisting of
-a set of policy states, \(\pol \in \POLS\) and a policy step function
+a set of policy states \(\pol \in \POLS\), and a policy step function
 \((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
@@
 
-%    The initializer and step function are both partial functions,
-%    the former to reflect static guarantees (some initial states may not be valid at all) and
-%    the latter to reflect dynamic fail-stop behavior. We then define machine-policy pairs and
-%    a step function on them:
-
-Let $\MPS$ be the set of all pairs of machine states and policy states.
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
 %
@@
 %
-Then we can lift this policy step function to operate on such pairs,
+We then lift this policy step function to operate on such pairs,
 by combining it with the regular step function for machine states.
-
-%    \[\stepstoobs{}_P \subseteq \MPS \times \MPS\]
 
@@
 
-\leo{Notation: Should $\stepstoobs{}$ be partial?}
+\leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted step?}
 
@@
 triples of (machine states, policy states, and observations), usually
-denoted $\MPT$.
+denoted $\MPT$. To continue, we will require three generic operations
+on traces: a ``trace-of'' operation to construct the potentially
+infinite traced induced by a step relation; a ``split'', to break a
+trace into two components based on a predicate; a ``join'', the
+reverse of split; and a ``prefix'', as a shorthand to only keep the
+first component of a split.
 
-We define a few operations on traces generically.
-The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
-separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
-
-  \judgment{\(f\ t\)}
-           {\(\mathit{split}(f,tT) = t, tT\)}
-
-  \judgmenttwo{\(\neg f\ t\)}{\(\mathit{split}(f,T) = T_1,T_2\)}
-              {\(\mathit{split}(f,tT) = tT_1,T_2\)}
-
-The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
-of the first (if it has one) with the first of the second:
-
-  \judgment{}
-           {\(T_1t \cdot T_2 = T_1 T_2\)}
-
-We define the ``trace-of'' operator \(\hookrightarrow\) that coinductively
+The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
 relates a machine state with the trace of machine states paired with
 observations produced by repeated application of step:
-\[\machT \in \mathit{MOTrace} = \mathit{Trace}\ (\MACHS \times \OBSS)\]
-\[\hookrightarrow : \machT \rightarrow \mathit{MOTrace}\]
-
+%
+\[\hookrightarrow : \machT \rightarrow \MOTRACE\]%
+%
 \judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
-            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}
-
+            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
+%
+\noindent
 Since the step function is total, if \(\mach \hookrightarrow \machT\),
 then \(\machT\) must be infinite.
-
-\sna{Lambdas in the split are a bit busy, not sure anything will look better.}
-We will also wish to take a prefix of the trace from an initial state up through the first state where
-some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
-\(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
-
-  \judgmenttwo{\(\mach \hookrightarrow \machT\)}
-              {\(\mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT) = \machT_1,\machT_2\)}
-              {\(\mach \hookrightarrow \machT_1 | f\)}
 
@@
 and observations induced by by \(\stepstoobs{}_P\):
-
-\leo{I think this would read better as not exists. Thoughts?}
-\sna{Looks good.}
-  \[\MPT \in \mathit{MPOTrace} = \mathit{Trace}\ (\MACHS \times \POLS \times \OBSS)\]
-  \[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \mathit{MPOTrace}\]
-
+%
+\[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \MPOTRACE \]
+%
 \judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
          {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
-
+%  
 \judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
@@
               {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
+We will also use \(\pi_m\) to project the trace of machine states
+out of a trace, \(\pi_p\) to project the trace of policy states, and
+\(\pi_o\) to project the observations.
 
-We also lift the prefix-up-to relation to such traces.
+The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\)
+and a predicate on elements \(f\) and separates \(T\) into a prefix
+and suffix at the first element on which \(f\) holds:
+
+\judgment{\(f\ t\)}
+         {\(\mathit{split}(f,tT) = t, tT\)}
+%  
+\judgmenttwo{\(\neg f\ t\)}{\(\mathit{split}(f,T) = T_1,T_2\)}
+            {\(\mathit{split}(f,tT) = tT_1,T_2\)}
+%
+Note that the last element of the prefix and the first element of the
+suffix coincide. As we will mostly be using $\mathit{split}$ to
+partition a trace to a callee and a caller portion, this choice will
+allow us to preserve local state encapsulation invariants during
+the transition between the two.
+
+The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
+$\mathit{split}$: it takes two traces and links them by replacing the
+last element of the first (if it has one) with the first element of
+the second
+
+\judgment{} {\(t \cdot T = T\)}
+\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
+\leo{Lemma/theorem: split/join = id?}
+\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
+
+Finally, we will also want to take a prefix of the trace from an
+initial state up through the first state where some condition holds.
+This can easily be implemented using $\mathit{split}$, but as we're
+going to be making heavy use of it we're going to provide special
+notation for convenience.
+%
+Let \(f\) be a predicate on machine states. Then we define \(\mach
+\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
+\(\machT\) up to \(f\)''):
+
+\judgmenttwo{\(\mach \hookrightarrow \machT\)}
+            {\(\mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT)
+              = (\machT_1,\machT_2)\)}
+            {\(\mach \hookrightarrow \machT_1 | f\)}
+\noindent
+And similarly for traces with policies:
 
@@
          {\(\mpstatename \hookrightarrow \MPT_1 | f\)}
-
-Finally, we will use \(\pi_m\) to project the trace of machine states
-out of a trace, \(\pi_p\) to project the trace of policy states, and
-\(\pi_o\) to project the observations.
 
@@
 
-\paragraph*{Observations and Equivalence}
+\paragraph*{Observational Equivalence}
 
@@
   e-mail, but we haven't said anything about attackers yet.}, if the
-sequence of word-observations of $\obsT_1$ is a prefix of those of
-$\obsT_2$, that is we operate up to deletion of \(\tau\) observations, coinductively:
+sequence of non-silent observations of $\obsT_1$ is a prefix of those
+of $\obsT_2$; that is we operate up to deletion of \(\tau\)
+observations, coinductively:
 
@@
 Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
-We then define {\em policy-sensitive observational equivalence}: an
-{\it MPOTrace} \(\MPT\) is equivalent to an {\it MOTrace} trace \(\machT\) if \(\MPT\)
-is infinite and the observations of \(\MPT\) and \(\machT\) coincide, or if the \(\MPT\)
-is finite (due to a policy fault) and its observations are a prefix of those of M.
+We then define {\em policy-sensitive observational equivalence}: a
+trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
+\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
+and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
+policy fault) and its observations are a prefix of those of M.
 

### 2020-06-25 main.tex
CP - conf

@@
 
-In this section we will describe stack safety as a formal
-property.
-\leo{Should call maps and returns go here or in preliminaries?}
+In this section we will describe stack safety as a formal property. We
+begin with our model of knowledge about the program structure\leo{I
+  hate this}, and then dive in to local state encapsulation and
+well-bracketed control flow.
 
-\paragraph*{Program Structure}
+\subsection{Program Structre}
 
-  Our properties and policies are parameterized by several types of annotations
-  describing features of a program. A {\em code map} identifies some addresses
-  as code and assigns each to a function. A {\em call map} identifies addresses
-  that complete a call sequence, abstracting away from any concrete calling
-  convention. Likewise a {\em return map} identifies addresses that complete a
-  return sequence. An {\em entry map} identifies valid entry points to a
-  function.
+Our properties and policies are parameterized by several types of
+annotations describing features of a program. A {\em code map}
+identifies some addresses as code and assigns each to a function. A
+{\em call map} identifies addresses that complete a call sequence,
+abstracting away from any concrete calling convention. Likewise a {\em
+  return map} identifies addresses that complete a return sequence. An
+{\em entry map} identifies valid entry points to a function.
 
@@
 space for arguments is already allocated, but the mechanism is more
-  general. A state in which a call map is defined on the program counter is
-  referred to as a {\it call state}.
+general. A state in which a call map is defined on the program counter
+is referred to as a {\it call state}.
 
-  A return map is simply a predicate on addresses and may be considered to
-  identify the final jump of the return sequence. An entry map is likewise a
-  predicate on addresses.
+A return map is simply a predicate on addresses and may be considered
+to identify the final jump of the return sequence. An entry map is
+likewise a predicate on addresses.
   %
@@
 
-\paragraph*{State Returns}
+\paragraph*{What is a return?}
 
-  We separately identify, for any call state, what is means for a state to
-  have correctly returned from a call. We consider the call to have returned
-  the first time control returns to the instruction following the \(\JAL\) with
-  the stack restored. Formally this is a relation between the states:
+We separately identify, for any call state, what is means for a state
+to have correctly returned from a call. We consider the call to have
+returned the first time control returns to the instruction following
+the \(\JAL\) with the stack restored. Formally this is a relation
+between the states:
 
@@
     \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) =
-    \nextPC(\mach_c(\PCname)) \land \\
-    & \mach_r(\SP) = \mach_c(\SP)
+    \nextPC(\mach_c(\PCname)) \land \\ & \mach_r(\SP) = \mach_c(\SP)
 \end{split}\]
 
-\leo{How do we initialize the first policy state based on a call map? It's not
-  part of the policy. Do we parameterize policies?}
+\subsection{Local State Encapsulation}
 
-\paragraph*{Stack safety}
-
-  A stack safety property is a property of a transition system, a policy, a
-  call map and a code map, that is decomposed into an integrity property and a
-  confidentiality property.
-
-    \[\begin{split}
-      \textit{stack safety} \triangleq \textit{stack integrity} \land
-      \textit{stack confidentiality} \\
-    \end{split}\]
-
-  \leo{TODO: Rephrase from introduction} Informally, {\it Stack
-  integrity} states that if a function call returns, the caller's
-  private stack data will be identical to what it was before the
-  call. {\it Stack confidentiality} states that data outside a callee's
-  stack frame is secret, and do not interfere with its observable
-  behavior or with the machine state at return.
+The first part of stack safety is {\em local state encapsulation}.
+Informally, it captures the notion that a caller's data is
+protected from a callee's actions. We formalize this notion
+as a security property with a confidentiality and an integrity
+component, where the former protects the caller's data from being
+read and the latter from being written.
 
@@
 
-  \leo{What about instructions? Shouldn't they be LC/HI?}
-  \sna{Now we are dealing with this!}
-
-  When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) for some
-  \(n\) and \(\mach_c \stepstoobs{\obs} \mach_e\), we can construct a contour
-  based on the bounds of the callee's stack frame. Instructions are low
-  confidentiality if they belong to the callee and high confidentiality
-  otherwise; regardless they are always high integrity.
-
-  The private state of the callee is the entire area of the memory from stack
-  base to stack pointer, with the exception of the $n$ locations
-  corresponding to the arguments of a call. Private components are marked
-  high confidentiality and high integrity: the callee should be able to
-  neither read, nor write them.
-
+When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) for
+some \(n\) and \(\mach_c \stepstoobs{\obs} \mach_e\), we can construct
+a contour based on the bounds of the callee's stack frame.
+%
+The private state of the callee is the entire area of the memory from
+stack base to stack pointer, with the exception of the $n$ locations
+corresponding to the arguments of a call. Private components are
+marked high confidentiality and high integrity: the callee should be
+able to neither read, nor write them.
+%  
 Everything above the stack pointer at the point of the call is
@@
 are marked high confidentiality and low integrity.
-
-  Registers and the frame of arguments to the callee are free for the
-  callee to use, and are therefore tagged with low integrity and low
-  confidentiality.
+%
+Special care is needed for the code memory corresponding to the
+callee: they are marked low confidentiality if they belong to the
+callee and high confidentiality otherwise; regardless they are always
+high integrity.\leo{That would mean we scramble some instructions, I
+  guess that makes sense.}
+%
+Finally, registers and the frame of arguments to the callee are free
+for the callee to use, and are therefore tagged with low integrity and
+low confidentiality.
 
@@
   \end{cases}\]
+
+\paragraph*{Stack Confidentiality}
+
+Confidentiality is a noninterference-style property, comparing the
+behavior of a program with an arbitrary variant that preserves low
+confidentiality data but may have any configuration of high
+confidentiality data.
+
+\definition{Variations}
+
+We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
+another given a contour $\contour$ if they agree on low
+confidentiality data:
+  \[\mach \approx_\contour \mach' \triangleq \forall \component .
+    \contour(\component) = \LOWSEC \rightarrow \mach(\component) =
+    \mach'(\component)\]
+
+\definition{Stack Confidentiality}
+
+A system satisfies eager {\em stack confidentiality} with respect to a
+call map \(\callmap\) and a code map \(\codemap\) if for any initial
+mpstate $\mpstate{\mach_0}{\pol_0}$
+%\(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\)
+and each adjacent pair of a call state \(\mach_c\) and state
+\(\mach_e\) with policy state \(\pol_e\) in the induced trace from
+\(\mpstatename\), where \(\callmap(\mach_c(\PCname)) = n\) and
+\(\codemap(\mach_e(\PCname)) = \funid\), one of two cases holds on the
+induced traces from \(\mpstate{\mach_e}{\pol_e}\) and any variation
+\(\mach_e'\) with respect to the contour
+\(\mathit{Cof}(\mach_c,\funid,n)\):
+
+\begin{itemize}
+\item
+If the trace from the call reaches a return -- \(\mpstate{\mach_e}{\pol_e}
+  \hookrightarrow_P \MPT | \ret{\mach_c}\) -- then the trace from the variant
+  \(\mach_e'\) reaches a return -- \(\mach_e' \hookrightarrow \machT |
+  \ret{\mach_c}\). Let \(\mach_r\) and \(\mach_r'\) be the final states of
+  \(\pi_m(\MPT)\) and \(\machT\), respectively. Until their final states,
+  \(\MPT\) and \(\machT\) must be observationally equivalent, \(\pi_o(\MPT)
+  \cdot \tau \eqsim \pi_o(\machT) \cdot \tau\). For any component
+  \(\component\) where \(\mach_e(\component) \not = \mach_r(\component)\) or
+  \(\mach_e'(\component) \not = \mach_r'(\component)\),
+  \(\mach_r\) agrees with \(\mach_r'\) on \(\component\).
+\item
+  If the trace from the call never reaches a return, neither does the
+  trace from the variant. Then if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P
+  \MPT\) and \(\mach_c' \hookrightarrow \machT\), \(\MPT \sim_P \machT\).
+\end{itemize}
+
+  \leo{Think more about whether we vary mc or step mc}
+  \sna{I've left it at step mc for now}
 
@@
   \(\HIGHINT\).
-
-\definition{Stack Confidentiality}
-
-  Confidentiality is in the style of non-interference, comparing the behavior
-  of a program with an arbitrary variant that preserves low confidentiality data
-  but may have any configuration of high confidentiality data. We define two
-  machine states as variants given a contour if they agree on low
-  confidentiality data:
-  \[\mach \approx_\contour \mach' \triangleq \forall \component .
-    \contour(\component) = \LOWSEC \rightarrow \mach(\component) =
-    \mach'(\component)\]
-
-  A system satisfies eager stack confidentiality with respect to a call map
-  \(\callmap\) and code map \(\codemap\) if for any initial mpstate
-  \(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each
-  adjacent pair of a call state \(\mach_c\) and state \(\mach_e\) with policy
-  state \(\pol_e\) in the induced trace from \(\mpstatename\), where
-  \(\callmap(\mach_c(\PCname)) = n\) and \(\codemap(\mach_e(\PCname)) =
-  \funid\), one of two cases holds on the induced traces from
-  \(\mpstate{\mach_e}{\pol_e}\) and any variation \(\mach_e'\) with respect
-  to the contour \(\mathit{Cof}(\mach_c,\funid,n)\).
-
-  If the trace from the call reaches a return -- \(\mpstate{\mach_e}{\pol_e}
-  \hookrightarrow_P \MPT | \ret{\mach_c}\) -- then the trace from the variant
-  \(\mach_e'\) reaches a return -- \(\mach_e' \hookrightarrow \machT |
-  \ret{\mach_c}\). Let \(\mach_r\) and \(\mach_r'\) be the final states of
-  \(\pi_m(\MPT)\) and \(\machT\), respectively. Until their final states,
-  \(\MPT\) and \(\machT\) must be observationally equivalent, \(\pi_o(\MPT)
-  \cdot \tau \eqsim \pi_o(\machT) \cdot \tau\). For any component
-  \(\component\) where \(\mach_e(\component) \not = \mach_r(\component)\) or
-  \(\mach_e'(\component) \not = \mach_r'(\component)\),
-  \(\mach_r\) agrees with \(\mach_r'\) on \(\component\).
-
-  If the trace from the call never reaches a return, neither does the
-  trace from the variant. Then if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P
-  \MPT\) and \(\mach_c' \hookrightarrow \machT\), \(\MPT \sim_P \machT\).
-
-  \leo{Think more about whether we vary mc or step mc}
-  \sna{I've left it at step mc for now}
 

### 2020-06-26 main.tex
NOtE
Cleaned up conf

@@
 
+\leo{TODO: We really need a better layout than centered paragraphs.}
 \judgment{}{\(\obsT \lesssim \obsT\)}
@@
 \judgment{}{\(\tau \lesssim \obsT\)}
+\judgment{}{\(w \lesssim w\obsT\)}
 
-\judgmenttwo{\(\obsT_1 = \tau \obsT_1'\)}{\(\obsT_1' \lesssim \obsT_2\)}
-            {\(\obsT_1 \lesssim \obsT_2\)}
+\judgment{\(\obsT_1 \lesssim \obsT_2\)}
+         {\(\tau \obsT_1 \lesssim \obsT_2\)}
 
-\judgmenttwo{\(\obsT_2 = \tau \obsT_2'\)}{\(\obsT_1 \lesssim \obsT_2'\)}
-            {\(\obsT_1 \lesssim \obsT_2\)}
+\judgment{\(\obsT_1 \lesssim \obsT_2\)}
+         {\(\obsT_1 \lesssim \tau \obsT_2\)}
 
-\judgmentthree{\(\obsT_1 = w \obsT_1'\)}
-              {\(\obsT_2 = w \obsT_2'\)}
-              {\(\obsT_1' \lesssim \obsT_2'\)}
-              {\(\obsT_1 \lesssim \obsT_2\)}
+\judgment{\(\obsT_1 \lesssim \obsT_2\)}
+         {\(w\obsT_1 \lesssim w\obsT_2\)}
 
@@
 
-\leo{Should we point out here the edge cases of all-tau traces?}
+\leo{Rephrase transition. I hate it:} There are two interesting things
+to note.  First, observation trace equivalence could alternatively be
+defined as the prefix relation but dropping the only assymetric rules
+\leo{2 + 3. TODO: name rules?}
 
-% Fixed by Andrew:
-%    \COQ{this should be equivalent to the coq version, and we should adopt it.}
-%    \leo{We had this, but it was much harder to work with for my proofs. We should still
-%      prove them equivalent if we want to present it this way (which is probably cleaner for
-%      presentation purposes)}
+\leo{Also rephrase this.}
+Second, an infinite silent trace is a prefix of (and equivalent to)
+any other trace. While this might seem surprising at first, it makes
+sense in a timing-insensitive context: an external observer looking
+at two machine runs cannot distinguish between a machine that
+steps forever or a machine that steps for a long time before producing
+some output, before that time has passed.
 
-Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
-We then define {\em policy-sensitive observational equivalence}: a
-trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
-\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
-and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
-policy fault) and its observations are a prefix of those of M.
-
-% Fixed by Sean
-%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
-\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
-            {\(\MPT \sim_P \machT\)}
-
-\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
-            {\(\MPT \sim_P \machT\)}
+%Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
+%We then define {\em policy-sensitive observational equivalence}: a
+%trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
+%\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
+%and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
+%policy fault) and its observations are a prefix of those of M.
+% 
+%% Fixed by Sean
+%%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
+%\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
+%            {\(\MPT \sim_P \machT\)}
+% 
+%\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
+%            {\(\MPT \sim_P \machT\)}
 
@@
 Confidentiality is a noninterference-style property, comparing the
-behavior of a program with an arbitrary variant that preserves low
-confidentiality data but may have any configuration of high
+behavior of a machine with an arbitrary variant that preserves low
 confidentiality data.
 
+\leo{Variations or variants?}
 \definition{Variations}
@@
 confidentiality data:
-  \[\mach \approx_\contour \mach' \triangleq \forall \component .
+%
+\[
+\mach \approx_\contour \mach' \triangleq \forall \component .
 \contour(\component) = \LOWSEC \rightarrow \mach(\component) =
-    \mach'(\component)\]
+\mach'(\component)
+\]
+
+Let $\callmap$ be a call map and $\codemap$ a code map.
+Confidentiality with respect to $\callmap$ and $\codemap$ should be
+enforced at {\em every} transition between a caller $f$ and a callee
+$f'$.
+%
+Suppose that an execution trace starting from some initial
+machine-policy state pair contains such a transition. That is, it
+contains some step between state state pairs $(\mach_c,\pol_c)
+\stepstoobs{\obs}_P (\mach_e, \pol_e)$ for some observation $\obs$
+such that $\mach_c$ is a valid call with $n$ arguments
+($\callmap(\mach_c(\PCname)) = n$) and $\mach_e$ is an entry point for
+$f'$ ($\codemap(\mach_e(\PCname)) = f'$).
+
+Now let $MP$ be the (potentially infinite) prefix of the trace
+starting at the entry $(\mach_e, \pol_e)$ until the corresponding
+return: $(\mach_e, \pol_e) \hookrightarrow MP ~ | ~ \ret{\mach_c}$ and
+$\contour$ be the contour at the call: $\contour =
+\mathit{Cof}(\mach_c, f', n)$.
+
+\definition{Trace Confidentiality}
+
+We will say that $MP$ satisfies {\em trace confidentiality} with
+respect to $\callmap$, $\codemap$, and $\contour$ if, for every
+$\contour$-variation $\mach_e'$ of the entry state, and the prefix $M$
+of its induced traced until the corresponding return $\mach_e'
+\hookrightarrow M ~ | ~ \ret{\mach_c}$, one of three cases holds:
+\begin{itemize}
+\item If the call successfully returns, that is $MP$ is finite and its
+  final state $\mach_r$ is a valid return, then $M$ must also be
+  finite with some final state $\mach_r'$ that has changed in the same
+  manner:
+$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
+  \vee \mach_e'(\component) \not = \mach_r'(\component) \rightarrow
+  \mach_r(\component) = \mach_r'(\component)$$
+
+  In addittion their observations must be equivalent:
+  $$\pi_o(MP) \eqsim \pi_o(M)$$
+\item If the call trace never returns and is infinite, then the
+  variant trace must also never return and their observations are
+  equivalent:
+  $$\pi_o(MP) \eqsim \pi_o(M)$$  
+\item Finally, if the call trace ends prematurely with a policy fault,
+  then we require that the observations of $MP$ are a prefix of those
+  of the variation:
+  $$\pi_o(MP) \lesssim \pi_o(M)$$
+\end{itemize}
 
@@
 
-A system satisfies eager {\em stack confidentiality} with respect to a
-call map \(\callmap\) and a code map \(\codemap\) if for any initial
-mpstate $\mpstate{\mach_0}{\pol_0}$
-%\(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\)
-and each adjacent pair of a call state \(\mach_c\) and state
-\(\mach_e\) with policy state \(\pol_e\) in the induced trace from
-\(\mpstatename\), where \(\callmap(\mach_c(\PCname)) = n\) and
-\(\codemap(\mach_e(\PCname)) = \funid\), one of two cases holds on the
-induced traces from \(\mpstate{\mach_e}{\pol_e}\) and any variation
-\(\mach_e'\) with respect to the contour
-\(\mathit{Cof}(\mach_c,\funid,n)\):
-
-\begin{itemize}
-\item
-If the trace from the call reaches a return -- \(\mpstate{\mach_e}{\pol_e}
-  \hookrightarrow_P \MPT | \ret{\mach_c}\) -- then the trace from the variant
-  \(\mach_e'\) reaches a return -- \(\mach_e' \hookrightarrow \machT |
-  \ret{\mach_c}\). Let \(\mach_r\) and \(\mach_r'\) be the final states of
-  \(\pi_m(\MPT)\) and \(\machT\), respectively. Until their final states,
-  \(\MPT\) and \(\machT\) must be observationally equivalent, \(\pi_o(\MPT)
-  \cdot \tau \eqsim \pi_o(\machT) \cdot \tau\). For any component
-  \(\component\) where \(\mach_e(\component) \not = \mach_r(\component)\) or
-  \(\mach_e'(\component) \not = \mach_r'(\component)\),
-  \(\mach_r\) agrees with \(\mach_r'\) on \(\component\).
-\item
-  If the trace from the call never reaches a return, neither does the
-  trace from the variant. Then if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P
-  \MPT\) and \(\mach_c' \hookrightarrow \machT\), \(\MPT \sim_P \machT\).
-\end{itemize}
-
-  \leo{Think more about whether we vary mc or step mc}
-  \sna{I've left it at step mc for now}
+We say that a system satisfies {\em stack confidentiality} with
+respect to a call map $\callmap$ and a code map $\codemap$ if for
+every initial machine-policy pair $(\mach_0, \pol_0)$ and its induced
+trace, every call subtrace\leo{Can i reuse stuff above?} in it
+satisfies trace confidentiality.
 

### 2020-06-27 main.tex
Stack safety section with explanation that actually flows and makes some sense - at least to me

@@
 protected from a callee's actions. We formalize this notion
-as a security property with a confidentiality and an integrity
+as a security property with an integrity and a confidentiality
 component, where the former protects the caller's data from being
-read and the latter from being written.
+written and the latter from being read.
 
@@
 %
-Since we need to enforce integrity and confidentiality at {\em every}
-function call in a trace, we must treat the security level of data
-dynamically. To that end, we introduce the notion of a {\em contour},
-a map from components to labels:
+In a traditional setting, integrity would enforce that the contents of
+$\HIGHINT$ locations remain unchanged, while confidentiality would
+enforce that the contents of $\HIGHSEC$ locations don't influence the
+observable behavior. However, when dealing with nested calls and
+returns, stack safety needs to enforce integrity and confidentiality
+for {\em every} function call in a trace; that is, we must treat the
+security level of data dynamically.
+%
+To that end, we introduce the notion of a {\em contour}, a map from
+components to labels:
 %
@@
 
+\paragraph*{Stack Integrity}
+
+The first component of local state encapsulation is an integrity
+property: it ensures that the caller's data is never overwritten by a
+callee. We will formally define integrity in two steps: first, we will
+define integrity as a trace property over the trace corresponding to a
+single call (from its entry point to its corresponding return), and
+then we will lift that to a property for every such subtrace of a
+callee.
+
+\definition{Trace Integrity}
+\label{def:trace-integrity}
+
+Let $\contour$ be a contour and $MP$ be a (potentially infinite)
+$\MPOTRACE$. We will say $MP$ satisfies {\em trace integrity} with
+respect to $\contour$ if its final state (if any) agrees with its
+initial one on all components marked $\HIGHSEC$ in $\contour$:
+%
+\[
+\mathit{int} ~ \contour ~ MP \triangleq
+\left \{ \begin{array}{ll}
+  \mathit{True}, & \text{if $MP$ is infinite}\\
+  \forall \component. ~ \contour(\component) = \HIGHSEC \rightarrow m_0(\component) = m_f(\component), & \text{otherwise} \\
+\end{array}
+\right .
+\]  
+\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
+
+Intuitively, the caller expects to find their private data untouched
+after the callee returns. That means that if the callee never returns,
+integrity is trivially satisfied. Similarly, the callee is allowed to
+temporarily overwrite these locations, as long as they are restored
+once control returns. We will see variations of this later on: a
+stronger, inductive integrity property that ensures that the caller's
+private data is never overwritten, even during the caller's execution
+(a variation that leads to easier reasoning and more effective testing);
+and a weaker, lazy integrity property that only ensures that the
+locations that a caller accesses are preserved (a variation that allows
+for a more efficient enforcement mechanism).
+
+Trace integrity characterizes the trace corresponding to a single
+callee. We now lift this definition to an integrity property for a
+complete execution trace. We do this by requiring that trace integrity
+holds for every transition between a caller and a callee $f$ for
+the entirety of the trace of $f$.
+
+\definition{Transition}
+
+Given a trace $MP$, a call map $\callmap$ and a code map $\codemap$,
+an $n-${\em transition} to a callee $f$ is a pair of machine-policy
+state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ in the trace
+such that:
+\begin{itemize}
+\item one state pair steps to the other: 
+  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstoobs{\obs}_P (\mach_e, \pol_e),$$
+\item $\mach_c$ is a valid call with $n$ arguments:
+  $$\callmap(\mach_c(\PCname)) = n,$$
+\item and $\mach_e$ is an entry point for $f$:
+  $$\codemap(\mach_e(\PCname)) = f.$$
+\end{itemize}
+
+\definition{Stack Integrity}
+
+Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
+system satisfies {\em stack integrity} with respect to $\callmap$ and
+$\codemap$ if, for every initial machine-policy state pair $(\mach_0,
+\pol_0)$ and its induced trace $MP$, and every $n$-transition
+$(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee $f$ in $MP$,
+the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$ until the first
+return corresponding to $\mach_c$ (if any) satisfies trace integrity
+with respect to the contour at the call:
+%
+$$\forall MP_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
+MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow
+\mathit{int}~\mathit{Cof}(\mach_c, f, n)~MP_{pre}$$
+
 \paragraph*{Stack Confidentiality}
 
-Confidentiality is a noninterference-style property, comparing the
-behavior of a machine with an arbitrary variant that preserves low
-confidentiality data.
+The second component of local state encapsulation is a confidentiality
+property: just like integrity ensures that a caller's data is never
+overwritten by a callee, confidentiality ensures that it never
+influences its observable behavior. We will formally define
+confidentiality in two similar steps: a trace version and a system
+version.
 
-\leo{Variations or variants?}
+Unlike integrity, confidentiality is a noninterference-style property,
+comparing the behavior of a machine with an arbitrary variant that
+preserves low confidentiality data.
+
 \definition{Variations}
@@
 \]
-
-Let $\callmap$ be a call map and $\codemap$ a code map.
-Confidentiality with respect to $\callmap$ and $\codemap$ should be
-enforced at {\em every} transition between a caller $f$ and a callee
-$f'$.
 %
-Suppose that an execution trace starting from some initial
-machine-policy state pair contains such a transition. That is, it
-contains some step between state state pairs $(\mach_c,\pol_c)
-\stepstoobs{\obs}_P (\mach_e, \pol_e)$ for some observation $\obs$
-such that $\mach_c$ is a valid call with $n$ arguments
-($\callmap(\mach_c(\PCname)) = n$) and $\mach_e$ is an entry point for
-$f'$ ($\codemap(\mach_e(\PCname)) = f'$).
-
-Now let $MP$ be the (potentially infinite) prefix of the trace
-starting at the entry $(\mach_e, \pol_e)$ until the corresponding
-return: $(\mach_e, \pol_e) \hookrightarrow MP ~ | ~ \ret{\mach_c}$ and
-$\contour$ be the contour at the call: $\contour =
-\mathit{Cof}(\mach_c, f', n)$.
+Then, the trace version of confidentiality operates on two traces: the
+``original'' one $MP \in \MPOTRACE$ that corresponds to a callee $f$
+and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
+if it started from a variant initial machine state.
 
@@
 
-We will say that $MP$ satisfies {\em trace confidentiality} with
-respect to $\callmap$, $\codemap$, and $\contour$ if, for every
-$\contour$-variation $\mach_e'$ of the entry state, and the prefix $M$
-of its induced traced until the corresponding return $\mach_e'
-\hookrightarrow M ~ | ~ \ret{\mach_c}$, one of three cases holds:
+\leo{Add info for ``valid return'':}
+We will say two traces $MP$ and $M$ satisfy {\em trace
+  confidentiality} with respect to $\contour$ if:
 \begin{itemize}
-\item If the call successfully returns, that is $MP$ is finite and its
-  final state $\mach_r$ is a valid return, then $M$ must also be
-  finite with some final state $\mach_r'$ that has changed in the same
-  manner:
+\item If $MP$ is finite and its final state $\mach_r$ is a valid
+  return, then $M$ is also finite with some final state $\mach_r'$ and
+  any changes from the initial states have changed in the same way:
 $$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
@@
   \mach_r(\component) = \mach_r'(\component)$$
-
   In addittion their observations must be equivalent:
   $$\pi_o(MP) \eqsim \pi_o(M)$$
-\item If the call trace never returns and is infinite, then the
-  variant trace must also never return and their observations are
-  equivalent:
+\item If $MP$ is infinite, then $M$ is also infinite and their
+  observations are equivalent:
   $$\pi_o(MP) \eqsim \pi_o(M)$$  
-\item Finally, if the call trace ends prematurely with a policy fault,
-  then we require that the observations of $MP$ are a prefix of those
-  of the variation:
+\item If the $MP$ ends prematurely with a policy fault, then
+  we require that its observations are a prefix of those of $M$:
   $$\pi_o(MP) \lesssim \pi_o(M)$$
@@
 
+\leo{Discuss variations in a similar manner}
+
+To lift trace confidentiality to a system property, we follow the same
+pattern as in stack integrity:
+
 \definition{Stack Confidentiality}
 
-We say that a system satisfies {\em stack confidentiality} with
-respect to a call map $\callmap$ and a code map $\codemap$ if for
-every initial machine-policy pair $(\mach_0, \pol_0)$ and its induced
-trace, every call subtrace\leo{Can i reuse stuff above?} in it
-satisfies trace confidentiality.
-
-\definition{Stack Integrity}
-
-  A system satisfies eager stack integrity with respect to a call map
-  \(\callmap\) and code map \(\codemap\) if for any initial mpstate
-  \(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each
-  adjacent pair of a call state \(\mach_c\) and state \(\mach_e\) with policy
-  state \(\pol_e\) in the induced trace from \(\mpstatename\), where
-  \(\callmap(\mach_c(\PCname)) = n\) and \(\codemap(\mach_e(\PCname)) =
-  \funid\), if we take the prefix until a return
-  \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_c}\) and
-  \(\MPT\) ends with some \(\mach_p\), \(\mach_p\) matches \(\mach_c\) on each
-  component whose integrity in \(\mathit{Cof}(\mach_c,\funid,n)\) is
-  \(\HIGHINT\).
+Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
+system satisfies {\em stack confidentiality} with respect to
+$\callmap$ and $\codemap$ if, for every initial machine-policy state
+pair $(\mach_0, \pol_0)$ and its induced trace $MP$, and every
+$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
+$f$ in $MP$, the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$
+until the first return corresponding to $\mach_c$ (if any) satisfies
+trace integrity with respect to every similarly prefixed machine trace
+of every variation $\mach_e'$ with the contour at the call:
+$$\forall MP_{pre} ~ M_{pre}. ~
+\mpstate{\mach_e}{\pol_e} \hookrightarrow_P MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
+$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
+\mach_e' \hookrightarrow M_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ MP_{pre} ~ M_{pre}$$
 

### 2020-06-27 main.tex
Reorg

@@
 \subsection{Local State Encapsulation}
+\label{sec:lse}
 
@@
 
-\section{Discussion}
-
-  \paragraph{Examples}
-
-  \bcp{Illustrate the definition by using it to discuss a range of ``stack
-    unsafety'' examples from the literature. }
-
-  \paragraph{What about tail recursion and other non-stack-like control flow}
-
-  \paragraph{Control Hijacking and Well-bracketed Control Flow}
-
+\subsection{Well-bracketed Control Flow}
+\label{sec:wbcf}
+
+Our integrity and confidentiality properties concern themselves with
+data, and may hold on a wide range of programs that violate expected
+control flow.  But well-bracketed control flow prevents this style of
+attack by guaranteeing that every jump between functions is either a
+call or a return, and each return jumps to the instruction after a
+corresponding call.
+
+\leo{Also dumping in relevant wbcf discussion part. TODO: merge}  
     Some forms of security rely on protected control flow. In a classic
@@
     {\em well-bracketed control flow}, below.
+
+\definition{Control Separation}
+
+A system enjoys {\em control separation} with regard to a code map
+\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\) if
+for any initial state \(\mach_0\) and any adjacent pair of states
+\(\mach_1\) and \(\mach_2\) in the induced trace from
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
+\(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2)\), then either
+\(\callmap(\mach_1(\PCname)) = n\) for some \(n\), or
+\(\retmap(\mach_1)\).
+
+\definition{Entry Integrity}
+
+A system enjoys {\em entry integrity} with regard to a call map
+\(\callmap\) and entry map \(\entmap\) if for any initial state
+\(\mach_0\) and any adjacent pair of states \(\mach_c\) and
+\(\mach_e\) in the induced trace from
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
+\(\callmap(\mach_c(\PCname)) = n\) then \(\entmap(\mach_e)\).
+
+\definition{Return Integrity}
+
+Return integrity captures the expected behavior of marked returns with
+regard to their call points. We begin by defining a coinductive
+relation \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1
+\MPT_2\). Given a call map \(\callmap\) and return map \(\retmap\),
+\(\MPT\) is related to \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\)
+whose head is the first unmatched return in \(\MPT\). So if \(\MPT\)
+starts with a return, it is related to itself:
+
+  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
+           {\(\retmap(\mach(\PCname))\)}
+      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\ 
+        \mpstate{\mach}{\pol}\)}
+  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
+              {\(\retmap(\mach(\PCname))\)}
+      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\ 
+              \mpstate{\mach}{\pol}\MPT\)}
+
+If \(\MPT\) starts with a non-call, non-return state, its unmatched
+return must come later:
+
+  \judgment{\(\callmap(\mach(\PCname)) = \bot
+              \hspace{.5cm} \neg \retmap(\mach(\PCname))
+
+              \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'\)}
+           {\(\mathit{Unmatched}\ \callmap\ \retmap\ 
+            \mpstate{\mach}{\pol}\MPT\ \MPT'\)}
+
+And most intricately, if \(\MPT\) starts with a call, we must find the
+return that matches it, which will be the unmatched return once we
+step past the call. Then the first unmatched return beyond it is the
+true unmatched return.
+
+    \judgment{\(\callmap(\mach(\PCname)) = n
+              \hspace{.5cm} \neg \retmap(\mach(\PCname))\)
+
+              \(\MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}}\)
+
+              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}}\)
+
+              \(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\)
+
+              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}}\)}
+            {\(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}\)}
+
+A system enjoys {\em return integrity} with regard to a call map
+\(\callmap\) and return map \(\retmap\) if for any initial state
+\(\mach_0\) and any call state and policy state \(\mach_c\) and
+\(\pol_c\) in the induced trace from
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if \(\mpstate{\mach_c}{\pol_c}
+\hookrightarrow_P \MPT\) and \(\mach_r\) is the first unmatched
+return, with \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT \ \MPT'\)
+and \(\mach_r\) the first machine state of \(\MPT'\), then
+\(\ret{\mach_c}\ \mach_r\).
+
+\definition{Well Bracketed Control Flow}
+
+The above properties are somewhat orthogonal, but together they
+capture a standard notion of control flow for a stack: functions keep
+their code separate outside of calls and returns, calls respect
+explicit entry points, and returns respect implicit entry points
+corresponding to their call.  A system that enjoys them with regard to
+a code, a call, a return, and an entry map is said to enjoy {\em well
+  bracketed control flow} with regard to those maps.
+
+\section{Discussion}
+
+  \paragraph{Examples}
+
+  \bcp{Illustrate the definition by using it to discuss a range of ``stack
+    unsafety'' examples from the literature. }
+
+  \paragraph{What about tail recursion and other non-stack-like control flow}
 
@@
       confidentiality with regard to them.
-
-\section{Well-bracketed Control Flow}
-
-  Our integrity and confidentiality properties concern themselves with data,
-  and may hold on a wide range of programs that violate expected control flow.
-  But well-bracketed control flow prevents this style of attack by guaranteeing
-  that every jump between functions is either a call or a return, and each
-  return jumps to the instruction after a corresponding call.
-
-  \definition{Control Separation}
-
-    A system enjoys {\em control separation} with regard to a code map
-    \(\codemap\), a call map \(\callmap\), and a return map \(\retmap\)
-    if for any initial state \(\mach_0\) and any adjacent pair of states
-    \(\mach_1\) and \(\mach_2\) in the induced trace from
-    \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if \(\codemap(\mach_1(\PCname))
-    \not = \codemap(\mach_2)\), then either \(\callmap(\mach_1(\PCname)) = n\)
-    for some \(n\), or \(\retmap(\mach_1)\).
-
-  \definition{Entry Integrity}
-
-    A system enjoys {\em entry integrity} with regard to a call map
-    \(\callmap\) and entry map \(\entmap\) if for any initial state
-    \(\mach_0\) and any adjacent pair of states \(\mach_c\) and \(\mach_e\)
-    in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
-    \(\callmap(\mach_c(\PCname)) = n\) then \(\entmap(\mach_e)\).
-
-  \definition{Return Integrity}
-
-    Return integrity captures the expected behavior of marked returns with
-    regard to their call points. We begin by defining a coinductive relation
-    \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1 \MPT_2\). Given a call map
-    \(\callmap\) and return map \(\retmap\), \(\MPT\) is related to
-    \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\) whose head is the first
-    unmatched return in \(\MPT\). So if \(\MPT\) starts with a return,
-    it is related to itself:
-
-  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
-           {\(\retmap(\mach(\PCname))\)}
-      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\ 
-        \mpstate{\mach}{\pol}\)}
-  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
-              {\(\retmap(\mach(\PCname))\)}
-      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\ 
-              \mpstate{\mach}{\pol}\MPT\)}
-
-    If \(\MPT\) starts with a non-call, non-return state, its unmatched
-    return must come later:
-
-  \judgment{\(\callmap(\mach(\PCname)) = \bot
-              \hspace{.5cm} \neg \retmap(\mach(\PCname))
-
-              \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'\)}
-           {\(\mathit{Unmatched}\ \callmap\ \retmap\ 
-            \mpstate{\mach}{\pol}\MPT\ \MPT'\)}
-
-    And most intricately, if \(\MPT\) starts with a call, we must find the
-    return that matches it, which will be the unmatched return once we step
-    past the call. Then the first unmatched return beyond it is the true
-    unmatched return.
-
-    \judgment{\(\callmap(\mach(\PCname)) = n
-              \hspace{.5cm} \neg \retmap(\mach(\PCname))\)
-
-              \(\MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}}\)
-
-              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}}\)
-
-              \(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\)
-
-              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}}\)}
-            {\(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}\)}
-
-    A system enjoys {\em return integrity} with regard to a call map
-    \(\callmap\) and return map \(\retmap\) if for any initial state
-    \(\mach_0\) and any call state and policy state \(\mach_c\) and \(\pol_c\)
-    in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
-    \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\) is the
-    first unmatched return, with \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT
-    \ \MPT'\) and \(\mach_r\) the first machine state of \(\MPT'\),
-    then \(\ret{\mach_c}\ \mach_r\).
-
-  \definition{Well Bracketed Control Flow}
-
-    The above properties are somewhat orthogonal, but together they capture a
-    standard notion of control flow for a stack: functions keep their code
-    separate outside of calls and returns, calls respect explicit entry points,
-    and returns respect implicit entry points corresponding to their call.
-    A system that enjoys them with regard to a code, a call, a return, and an
-    entry map is said to enjoy {\em well bracketed control flow} with regard to
-    those maps.
 

### 2020-06-29 main.tex
Write-up about the run-from-state version

@@
 
-The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\)
-and a predicate on elements \(f\) and separates \(T\) into a prefix
-and suffix at the first element on which \(f\) holds:
-
-\judgment{\(f\ t\)}
-         {\(\mathit{split}(f,tT) = t, tT\)}
-%  
-\judgmenttwo{\(\neg f\ t\)}{\(\mathit{split}(f,T) = T_1,T_2\)}
-            {\(\mathit{split}(f,tT) = tT_1,T_2\)}
-%
-Note that the last element of the prefix and the first element of the
-suffix coincide. As we will mostly be using $\mathit{split}$ to
-partition a trace to a callee and a caller portion, this choice will
-allow us to preserve local state encapsulation invariants during
-the transition between the two.
-
-The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
-$\mathit{split}$: it takes two traces and links them by replacing the
-last element of the first (if it has one) with the first element of
-the second
-
-\judgment{} {\(t \cdot T = T\)}
-\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
-\leo{Lemma/theorem: split/join = id?}
-\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
-
-Finally, we will also want to take a prefix of the trace from an
-initial state up through the first state where some condition holds.
-This can easily be implemented using $\mathit{split}$, but as we're
-going to be making heavy use of it we're going to provide special
-notation for convenience.
+The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
+and a predicate on elements \(f\) and gives the prefix of \(T\)
+ending with the first element on which \(f\) holds, coinductively:
+
+\judgment{\(f\ \traceelem\)}
+         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
+%
+\judgment{}
+         {\(\PUT(f,\traceelem) = \traceelem\)}
+%
+\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
+            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
+%
+Note that if the resulting trace is a strict prefix of the parameter, then the
+predicate must hold on its final element. Conversely, if the predicate never
+holds, \(\PUT\) is the identity function.
+
+% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
+%$\mathit{split}$: it takes two traces and links them by replacing the
+%last element of the first (if it has one) with the first element of
+%the second
+
+%\judgment{} {\(t \cdot T = T\)}
+%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
+%\leo{Lemma/theorem: split/join = id?}
+%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
+
+We will frequently take a prefix of the trace from an initial state up
+through the first state where some condition holds on the machine state.
+This can easily be implemented using \(\PUT\), and we provide special notation
+for convenience.
 %
@@
 \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
-\(\machT\) up to \(f\)''):
+running \(\mach\) up to \(f\)''):
 
 \judgmenttwo{\(\mach \hookrightarrow \machT\)}
-            {\(\mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT)
-              = (\machT_1,\machT_2)\)}
-            {\(\mach \hookrightarrow \machT_1 | f\)}
+            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT)
+              = \machT'\)}
+            {\(\mach \hookrightarrow \machT' | f\)}
 \noindent
@@
 \judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
-              {\(\mathit{split}(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT_1,\MPT_2\)}
-         {\(\mpstatename \hookrightarrow \MPT_1 | f\)}
+            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT'\)}
+         {\(\mpstatename \hookrightarrow \MPT' | f\)}
 
@@
 
+\sna{Version for stepping from a state}
+\definition{Confidential Execution}
+
+  Stack confidentiality, below, is a property of a system that, from a call to
+  a return, each trace obeys a local confidentiality property,
+  {\it confidential execution}.
+
+  We will say that a state \(\mpstate{\mach}{\pol}\) satisfies confidential
+  execution with regard to a contour \(\contour\) and an ending condition
+  \(R\) that is a predicate on machine states if, for all variant states
+  \(\mach' \approx_\contour \mach\), the induced traces from \(\mach\) and
+  \(\mach'\) behave the same across three cases. Let
+  \(\mpstate{\mach}{\pol} \hookrightarrow \MPT | R\) and
+  \(\mach' \hookrightarrow \machT | R\).
+
+  If \(\MPT\) ends with a state \(\mach_r\) and \(R\ \mach_r\), then \(\machT\)
+  ends with a state \(\mach_r'\) such that \(R\ \mach_r'\). \(\MPT\) and
+  \(\machT\) are observationally equivalent, and for each component
+  \(\component\) such that \(\contour(\component) = \HIGHSEC\),
+  \(\mach_r(\component) = \mach_r'(\component)\).
+
+  If \(\MPT\) is infinite, then \(\machT\) is infinite, and they are
+  observationally equivalent.
+
+  If \(\MPT\) ends before \(R\) ever holds, meaning that it halted due to a
+  policy violation, then it is a strict observational prefix of \(\machT\),
+  which must be infinite.
+
 \definition{Stack Confidentiality}

### 2020-06-29 main.tex
Quick read of first two sections, couple of notes

@@
 safety is lacking from the literature, with the notion only being
-defined negatively: through examples of stack unsafety. Two classical
+defined negatively: through examples of stack unsafety
+\rb{usually though not always; as noted below there are other approaches}. Two classical
 examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
@@
 that together form a formal definition of stack safety.
+\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
+rigor in it, even if some notions are implicit or not fully developed.}
 
@@
 properties, this setting provides an interesting additional challenge:
-the notion of who the attacker shifts throught execution. Not only
+the notion of who the attacker is shifts throught execution. Not only
 does a caller need to be protected from its callee, but the callee
@@
 extended with some kind of (static or dynamic, software- or
-hardware-enforced) security monitor. However, our definition of stack
+hardware-enforced) security monitor
+\rb{Put stress on genericity first, advance this concrete model after?}.
+However, our definition of stack
 safety is generic: we will begin with an abstract interface to our
@@
 Our machine states are comprised of {\em components}, that can be word
-(denoting memory addresses) or registers (drawn from some set $\REGS$.
+(denoting memory addresses) or registers (drawn from some set $\REGS$).
 %
@@
 The registers must include two special purpose registers: the program
-counter \PCname and the stack pointer \SP.
+counter {\PCname} and the stack pointer \SP.
 %
@@
 triples of (machine states, policy states, and observations), usually
-denoted $\MPT$. To continue, we will require three generic operations
+denoted $\MPT$.
+\rb{The way the definitions are set up, one may expect MPO, but\ldots}
+To continue, we will require three \{four?} generic operations
 on traces: a ``trace-of'' operation to construct the potentially
-infinite traced induced by a step relation; a ``split'', to break a
+infinite traced induced by a step relation
+\rb{not an op on traces per se?};
+a ``split'', to break a
 trace into two components based on a predicate; a ``join'', the
@@
 
-Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
+Let $\callmap$ be a call map and $\codemap$ a code map.
+\rb{Could we find better symbols for the maps?}
+We say that a
 system satisfies {\em stack confidentiality} with respect to

### 2020-06-29 main.tex
fix build

@@
 \rb{The way the definitions are set up, one may expect MPO, but\ldots}
-To continue, we will require three \{four?} generic operations
+To continue, we will require three \rb{four?} generic operations
 on traces: a ``trace-of'' operation to construct the potentially

### 2020-06-30 main.tex
nit
Started lazy integrity in new style

@@
 
-\subsection{Program Structre}
+\subsection{Program Structure}
 
@@
 
-    \definition{Observable Integrity}
-
-      Observable integrity weakens eager integrity by allowing a returned state
-      to differ from its call state at high integrity components, provided the
-      observable behavior of the induced trace from the return matches that of
-      an idealized state whose high integrity components do match the call
-      state.
-
-      Given a contour, a call state, and a returned state, we define a
-      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
-      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
-      the call state on those components that are high integrity in its
-      contour, and matches the return state on low integrity components:
-      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
+    \definition{Observable Trace Integrity}
+
+      Suppose a function \(f\) calls another function \(g\) with contour
+      \(\contour\). If \(g\) alters an address below its stack frame, where
+      \(\contour\) is \(\HIGHSEC\), the trace from the entry point to the
+      return will not enjoy Trace Integrity. But if \(f\) immediately
+      overwrites the overwritten location, the illegal write can be safely
+      ignored, as it will never affect the behavior of the program.
+      {\em Observable trace integrity} captures this intuition by guaranteeing
+      that the final state, instead of being identical to the initial one,
+      has the same behavior as it would if the writes had not occurred.
+
+      Let \(\contour\) be a contour and \(\MPT\) be a finite \(\MPOTRACE\)
+      whose initial machine state is \(\mach_e\). \(\MPT\) satisfies {\em trace
+      integrity} with respect to \(\contour\) if the trace induced by its final
+      state, \(\mpostate{\mach_r}{\pol}{\obs}\), is an observational prefix of
+      that induced by an idealized state \(\mach_r'\) that agrees with
+      \(\mach_r\) on all components marked \(\HIGHSEC\) in \(\contour\).
+      Formally we define a rollback function, \(\mathit{roll}\), that takes
+      two machine states and produces a third.
+      %
+      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
       \begin{cases}
@@
 
-      Intuitively, the result of a rollback is an idealized state that
-      reflects changes to accessible components, but ignores the callee's
-      interference with the caller's data. Then the observable behavior of the
-      resulting trace is the standard to which we expect the real trace to
-      conform.
-
-      A system enjoys observable integrity with regard to \(\callmap\) and 
-      \(\codemap\) if for any initial state \(\mach_0\) and each adjacent pair
-      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
-      \(\pol_e\) in the trace induced by
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where
-      \(\callmap(\mach_c) = n\) and \(\codemap(\mach_e) = \funid\),
-      if \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
-      with the return state \(\mach_r\) at the end of \(\MPT\), then the
-      following holds.
-
-      Let \(\contour = \mathit{Cof}(\mach_c,\funid,n)\) describe the security
-      levels of the call and
-      \(\mach_r' = \mathit{roll}(\contour,\mach_e,\mach_r)\). The induced
-      traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\) and
-      \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
-      possible policy failure: \(\MPT \sim_P \machT'\).
+      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
+      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
+      agrees with \(\mach_e\) on every component the callee should not change,
+      but with \(\mach_r\) on those components that it is permitted to change.
+      So the observable behavior that follows is the idealized behavior to
+      which we compare the trace from \(\mach_r\), which must prefix it (to
+      account for premature halts due to policy violations.)
+      %
+      \[\begin{split}
+        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
+          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
+            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
+        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' \Rightarrow \\
+        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
+      \end{split}\]
+
+    \definition{Observable Stack Integrity}
+
+      A system enjoys {\em observable integrity} with regard to \(\callmap\)
+      and  \(\codemap\) if for any initial state \(\mach_0\) and each
+      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
+      contour \(\contour\) in the induced trace from
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
+      then \(\text{\sc ObsInt}_\contour ~ \MPT\).
 

### 2020-07-01 main.tex
Program and initial memory for example

@@
 Before we define stack safety, we need to first set the stage by
-describing our machine model. Concretely, we have in mind a
+describing our machine model. Our definition of stack safety is
+generic and this sections discusses an abstract interface to our
+machine model and some kind of static or dynamic, software- or
+hardware-enforced security monitor. At the same time, for
+concreteness, we will build up our running example using a
 conventional machine architecture with a standard (RISC-V) processor
-extended with some kind of (static or dynamic, software- or
-hardware-enforced) security monitor
-\rb{Put stress on genericity first, advance this concrete model after?}.
-However, our definition of stack
-safety is generic: we will begin with an abstract interface to our
-machine model and only make specific choices about details when we get
-to the policy enforcement section~\ref{sec:micropolicy}.
+and later extend it with a dynamic policy mechanism.
 
@@
 \(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \).
+
+\paragraph*{Running Example: RISC-V}
+
+Mapping a conventional architecture to this abstract interface is
+straightforward. To build up our running example, we are going to use
+a standard 64-bit RISC machine, just like the one we are using in our
+experiments later on \leo{Mention FORVIS?}. In this setting, $\WORDS$
+is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
+r_{31}\} \cup {\PCname}$, and $\SP = r_2$. \leo{Mention We use the
+  standard encoding of instructions to words?}
+
+The first interesting choice is that of observations, as there are
+multiple reasonable options. At one extreme, we could choose to define
+observations as the value written at any component. However, that
+would pose problems later on when we treat some components as ``secret''
+when they correspond to a caller's private state. \leo{On the other extreme,
+we could pick a particular component and treat it as the ``output'' of
+our system. }
 
@@
 \leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted step?}
+
+\leo{TODO: Forward pointer to policies right?}
 
@@
 \rb{The way the definitions are set up, one may expect MPO, but\ldots}
-To continue, we will require three \rb{four?} generic operations
+To continue, we will require four generic operations
 on traces: a ``trace-of'' operation to construct the potentially
@@
 
+\leo{TODO: Add an example of a trace once I figure out the notations}
+
 Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
@@
 
+\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
+  seems better}
 The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
@@
 
+\subsection{Running Example}
+
+\leo{TODO: Figure out presentation}
+
+\newcommand*{\add}{\text{add}}
+\newcommand*{\addi}{\text{addi}}
+\newcommand*{\sw}{\text{sw}}
+\newcommand*{\lw}{\text{lw}}
+\newcommand*{\jal}{\text{jal}}
+\newcommand*{\jalr}{\text{jalr}}
+%\begin{lstlisting}
+%START:
+%  addi r2 r2 12 ; allocate 3 words in the stack
+%  sw r2 42 -8; store 42 in a private location
+%  sw r2 0 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 F ; call F and store return address
+% 
+%F:
+%  addi r2 r2 4 ; allocate for ra 
+%  sw r2 r1 -4 ; store the return address to sp-4
+%  addi r2 r2 8 ; allocate for arg + ret
+%  sw r2 17 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 G ; call G and store return address
+% 
+%G:  
+%  addi r2 r2 4 ; allocate for ra 
+%  sw r2 r1 -4 ; store the return address to sp-4
+%  
+%\end{lstlisting}
+
+Starting State:
+
+Instruction memory in locations 0 - ..
+PC at 100.
+Stack memory starting at 1000.
+Assume we're in a function. 
+Stack Pointer at 1040.
+1040: 42.
+
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
+\MemoryLabel{35em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{4}{gray}[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-18.7em}{0.6em}{42}
+\MemoryLabel{-16.4em}{0.6em}{0}
+\\
+\end{center}
+
+Program:
+
+\begin{verbatim}
+top{
+  return f(0);
+}
+
+f(x){
+  return (x + g(17));
+}
+
+g(x){
+  return (x + 1);
+}
+\end{verbatim}
+
+\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
+\newcommand*{\clabel}[1]{#1: & & &\\}
+\newcommand*{\rsp}{sp}
+\newcommand*{\rra}{ra}
+\[
+\begin{array}{llll}
+  \clabel{TOP}  
+  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
+  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
+  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
+  & & & \\
+  & & & \\
+  \clabel{F}
+  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
+  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
+  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
+  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
+  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
+  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
+  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
+  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
+  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
+  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
+  & & & \\
+  & & & \\  
+  \clabel{G}
+  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
+  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
+  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
+  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
+
+%  addi r2 r2 8 ; allocate for arg + ret
+%  sw r2 17 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 G ; call G and store return address
+
+\end{array}
+\]
+
+%  sw r2 42 -8; store 42 in a private location
+%  sw r2 0 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 F ; call F and store return address
+
+%  0: & \addi ~ r_4 ~ r_0 ~ 40 &
+%     ; ~ \text{\em put $40$ in register $r_4$}\\
+%  4: & \sw ~ r_4 ~ r_4 ~ 1000 &
+%     ; ~ \text{\em store $40$ in location $1040$}\\
+%  8: & \jal ~ r_1 ~ 20        &
+%     ; ~ \text{\em jump to $20$, store return addres in $r_1$}\\
+%  12: & & \\
+%  16: & & \\
+%  20: & \sw ~ r_2 ~ r_1 ~ 4 &
+%     ; ~ \text{\em store the return address to the $\SP + 4$}\\
+%  24: & \addi ~ r_2 ~ r_2 ~ 8 &
+%     ; ~ \text{\em increase the $\SP$ by 8}\\
+%\end{array}
+%\]
+% 
+%  , (SW sp ra 4  , tagH1)
+%  , (ADDI sp sp 8, tagH2)
+%  , (LW 3 3 1000, instrTag)
+%  ]
+
 \subsection{Program Structure}
@@
 
+\leo{Visible from assembly}
+
 A call map is a partial function from addresses to a natural number,
@@
 \[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]
+
+\leo{The two JAL instructions map to 2 (arg + return)}
 
@@
 
+\leo{retmap is the jal + 1}
+\leo{entrymap is the ones after the labels}
+
 \paragraph*{What is a return?}
@@
 \end{split}\]
+
+\leo{Point to the trace. two return points}
 
@@
   \end{cases}\]
+
+  \leo{Snapshot state at the point of the call from TOP to F, collor
+    the different things}
 

### 2020-07-01 main.tex
Started on observational conf section

@@
 
-\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
-  seems better}
+% SNA: fixed
+%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
+%  seems better}
 The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
@@
 
-\sna{Version for stepping from a state}
-\definition{Confidential Execution}
-
-  Stack confidentiality, below, is a property of a system that, from a call to
-  a return, each trace obeys a local confidentiality property,
-  {\it confidential execution}.
-
-  We will say that a state \(\mpstate{\mach}{\pol}\) satisfies confidential
-  execution with regard to a contour \(\contour\) and an ending condition
-  \(R\) that is a predicate on machine states if, for all variant states
-  \(\mach' \approx_\contour \mach\), the induced traces from \(\mach\) and
-  \(\mach'\) behave the same across three cases. Let
-  \(\mpstate{\mach}{\pol} \hookrightarrow \MPT | R\) and
-  \(\mach' \hookrightarrow \machT | R\).
-
-  If \(\MPT\) ends with a state \(\mach_r\) and \(R\ \mach_r\), then \(\machT\)
-  ends with a state \(\mach_r'\) such that \(R\ \mach_r'\). \(\MPT\) and
-  \(\machT\) are observationally equivalent, and for each component
-  \(\component\) such that \(\contour(\component) = \HIGHSEC\),
-  \(\mach_r(\component) = \mach_r'(\component)\).
-
-  If \(\MPT\) is infinite, then \(\machT\) is infinite, and they are
-  observationally equivalent.
-
-  If \(\MPT\) ends before \(R\) ever holds, meaning that it halted due to a
-  policy violation, then it is a strict observational prefix of \(\machT\),
-  which must be infinite.
-
 \definition{Stack Confidentiality}
@@
 
-    \definition{Observable Confidentiality}
-
-      Observable confidentiality is concerned with whether high confidentiality
-      data influences observable both behavior during and after a call. As with
-      eager confidentiality, we will compare the behavior of the trace induced
-      by a state to another induced by its variant. But if the original returns,
-      before continuing, we must reset the variant to conform with the original
-      state.
+    \definition{Observable Trace Confidentiality}
+
+      Trace confidentiality is a property of two traces whose initial machine
+      states are variations, with two sub-properties. First, the behavioral
+      condition -- that their observable behaviors prior to the end of the
+      trace are equivalent -- reflects the intuition that secret values should
+      not influence visible behavior. Second, we have the end condition that
+      their final states (if any) are equal up to the original variation. The
+      end condition over-approximates the extension of the behavioral condition
+      to future execution.
+
+      We introduce {\em observable trace confidentiality}, which extends the
+      behavioral condition to apply to execution from the final state on.
+      This requires restoring varied components to their original values,
+      since after return they will once again be accessible legitimately.
+
+      [Example goes here, showing how, when we return from a call, execution
+      becomes incoherent unless the state is restored.]
 
@@
 
-      A system enjoys observable confidentiality with respect to \(\callmap\)
-      and \(\codemap\) if for any initial state \(\mach_0\), each adjacent pair
-      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
-      \(\pol_e\) in the inducted trace from
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), and each variant
-      \(\mach_e' \approx_\contour \mach_e'\) where
-      \(\contour = \mathit{Cof}(\mach_c, \funid, n)\), one of two cases holds
-      on the induced traces from \(\mach_e\) and \(\mach_e'\).
-
-      If \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_1 | \ret{\mach_c}\)
-      with \(\MPT_1\) the trace within the callee ending with the machine state
-      \(\mach_r\), the first state to return to the caller, and policy state
-      \(\pol_r\), then the there is a return state \(\mach_r'\) reached by the
-      trace from the variant, \(\mach_c' \hookrightarrow \machT_1 |
-      \ret{\mach_c}\). Let \(\mach_r'' = \mathit{restore}(\contour,\mach_e,
-      \mach_e',\mach_r,\mach_r')\), and let the induced traces from the
-      returned states be \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT_2\)
-      and \(\mach_r'' \hookrightarrow \machT_2\). With the callee and caller
-      traces joined, the traces are observationally equivalent up to policy
-      failures: \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1 \cdot \machT_2\).
-
-      The case in which the induced trace does not return is the same as in the
-      eager property: the variant trace also does not return, and the induced
-      traces are observationally equivalent up to policy failure.
+      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_c\) and
+      \(\mach_c'\), respectively, with \(\mach_c \approx_\contour \mach_c'\),
+      enjoy observable trace confidentiality with respect to \(\contour\) if
+      two conditions hold:
+
+      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim_O
+      \pi_o(\machT)\).
+
+      If \(\MPT\) is finite and ends with the state
+      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
+      \(\mach_r'\) be the final state of \(\machT\). Then the induced traces
+      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
+      \(\mathit{restore}(\contour,\mach_c,\mach_c',\mach_r,\mach_r')
+      \hookrightarrow \machT'\) are observationally prefixed,
+      \(\pi_o(\MPT') \lesssim_O \pi_o(\machT')\).
+
+      \sna{Structured like this, we need to start with the call state, not
+           the entry. That's because we need to know what to restore to.
+           But we could also define restore in the stack confidentiality
+           section, maybe?}
+
+    \definition{Observable Stack Confidentiality}
 

### 2020-07-01 main.tex
Minor (maybe just spacing)

--- 2020-06-30 main.tex
@@
 safety is lacking from the literature, with the notion only being
-defined negatively: through examples of stack unsafety
-\rb{usually though not always; as noted below there are other approaches}. Two classical
+defined negatively: through examples of stack unsafety. Two classical
 examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
@@
 that together form a formal definition of stack safety.
-\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
-rigor in it, even if some notions are implicit or not fully developed.}
 
@@
 properties, this setting provides an interesting additional challenge:
-the notion of who the attacker is shifts throught execution. Not only
+the notion of who the attacker shifts throught execution. Not only
 does a caller need to be protected from its callee, but the callee
@@
 extended with some kind of (static or dynamic, software- or
-hardware-enforced) security monitor
-\rb{Put stress on genericity first, advance this concrete model after?}.
-However, our definition of stack
+hardware-enforced) security monitor. However, our definition of stack
 safety is generic: we will begin with an abstract interface to our
@@
 Our machine states are comprised of {\em components}, that can be word
-(denoting memory addresses) or registers (drawn from some set $\REGS$).
+(denoting memory addresses) or registers (drawn from some set $\REGS$.
 %
@@
 The registers must include two special purpose registers: the program
-counter {\PCname} and the stack pointer \SP.
+counter \PCname and the stack pointer \SP.
 %
@@
 triples of (machine states, policy states, and observations), usually
-denoted $\MPT$.
-\rb{The way the definitions are set up, one may expect MPO, but\ldots}
-To continue, we will require three \rb{four?} generic operations
+denoted $\MPT$. To continue, we will require three generic operations
 on traces: a ``trace-of'' operation to construct the potentially
-infinite traced induced by a step relation
-\rb{not an op on traces per se?};
-a ``split'', to break a
+infinite traced induced by a step relation; a ``split'', to break a
 trace into two components based on a predicate; a ``join'', the
@@
 
-\subsection{Program Structure}
+\subsection{Program Structre}
 
@@
 \label{def:trace-integrity}
-
 Let $\contour$ be a contour and $MP$ be a (potentially infinite)
@@
 \definition{Transition}
-
 Given a trace $MP$, a call map $\callmap$ and a code map $\codemap$,
-an $n-${\em transition} to a callee $f$ is a pair of machine-policy
+an $n$-{\em transition} to a callee $f$ is a pair of machine-policy
 state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ in the trace
@@
 \definition{Stack Integrity}
-
 Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
@@
 \definition{Variations}
-
 We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
@@
 \definition{Trace Confidentiality}
-
 \leo{Add info for ``valid return'':}
@@
 \definition{Confidential Execution}
-
   Stack confidentiality, below, is a property of a system that, from a call to
@@
 \definition{Stack Confidentiality}
-
-Let $\callmap$ be a call map and $\codemap$ a code map.
-\rb{Could we find better symbols for the maps?}
-We say that a
+Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
 system satisfies {\em stack confidentiality} with respect to
@@
 \definition{Control Separation}
-
 A system enjoys {\em control separation} with regard to a code map
@@
 \definition{Entry Integrity}
-
 A system enjoys {\em entry integrity} with regard to a call map
@@
 \definition{Return Integrity}
-
 Return integrity captures the expected behavior of marked returns with
@@
 \definition{Well Bracketed Control Flow}
-
 The above properties are somewhat orthogonal, but together they
@@
 
-    \definition{Observable Trace Integrity}
-
-      Suppose a function \(f\) calls another function \(g\) with contour
-      \(\contour\). If \(g\) alters an address below its stack frame, where
-      \(\contour\) is \(\HIGHSEC\), the trace from the entry point to the
-      return will not enjoy Trace Integrity. But if \(f\) immediately
-      overwrites the overwritten location, the illegal write can be safely
-      ignored, as it will never affect the behavior of the program.
-      {\em Observable trace integrity} captures this intuition by guaranteeing
-      that the final state, instead of being identical to the initial one,
-      has the same behavior as it would if the writes had not occurred.
-
-      Let \(\contour\) be a contour and \(\MPT\) be a finite \(\MPOTRACE\)
-      whose initial machine state is \(\mach_e\). \(\MPT\) satisfies {\em trace
-      integrity} with respect to \(\contour\) if the trace induced by its final
-      state, \(\mpostate{\mach_r}{\pol}{\obs}\), is an observational prefix of
-      that induced by an idealized state \(\mach_r'\) that agrees with
-      \(\mach_r\) on all components marked \(\HIGHSEC\) in \(\contour\).
-      Formally we define a rollback function, \(\mathit{roll}\), that takes
-      two machine states and produces a third.
-      %
-      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
+    \definition{Observable Integrity}
+      Observable integrity weakens eager integrity by allowing a returned state
+      to differ from its call state at high integrity components, provided the
+      observable behavior of the induced trace from the return matches that of
+      an idealized state whose high integrity components do match the call
+      state.
+
+      Given a contour, a call state, and a returned state, we define a
+      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
+      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
+      the call state on those components that are high integrity in its
+      contour, and matches the return state on low integrity components:
+      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
       \begin{cases}
@@
 
-      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
-      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
-      agrees with \(\mach_e\) on every component the callee should not change,
-      but with \(\mach_r\) on those components that it is permitted to change.
-      So the observable behavior that follows is the idealized behavior to
-      which we compare the trace from \(\mach_r\), which must prefix it (to
-      account for premature halts due to policy violations.)
-      %
-      \[\begin{split}
-        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
-          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
-            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
-        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' \Rightarrow \\
-        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
-      \end{split}\]
-
-    \definition{Observable Stack Integrity}
-
-      A system enjoys {\em observable integrity} with regard to \(\callmap\)
-      and  \(\codemap\) if for any initial state \(\mach_0\) and each
-      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
-      contour \(\contour\) in the induced trace from
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
-      then \(\text{\sc ObsInt}_\contour ~ \MPT\).
+      Intuitively, the result of a rollback is an idealized state that
+      reflects changes to accessible components, but ignores the callee's
+      interference with the caller's data. Then the observable behavior of the
+      resulting trace is the standard to which we expect the real trace to
+      conform.
+
+      A system enjoys observable integrity with regard to \(\callmap\) and 
+      \(\codemap\) if for any initial state \(\mach_0\) and each adjacent pair
+      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
+      \(\pol_e\) in the trace induced by
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where
+      \(\callmap(\mach_c) = n\) and \(\codemap(\mach_e) = \funid\),
+      if \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
+      with the return state \(\mach_r\) at the end of \(\MPT\), then the
+      following holds.
+
+      Let \(\contour = \mathit{Cof}(\mach_c,\funid,n)\) describe the security
+      levels of the call and
+      \(\mach_r' = \mathit{roll}(\contour,\mach_e,\mach_r)\). The induced
+      traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\) and
+      \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
+      possible policy failure: \(\MPT \sim_P \machT'\).
 
     \definition{Observable Confidentiality}
-
       Observable confidentiality is concerned with whether high confidentiality
@@
     \definition{Observable Stack Safety}
-
       A system enjoys {\em observable stack safety} with regard to a call map

### 2020-07-01 main.tex

@@
 Before we define stack safety, we need to first set the stage by
-describing our machine model. Concretely, we have in mind a
+describing our machine model. Our definition of stack safety is
+generic and this sections discusses an abstract interface to our
+machine model and some kind of static or dynamic, software- or
+hardware-enforced security monitor. At the same time, for
+concreteness, we will build up our running example using a
 conventional machine architecture with a standard (RISC-V) processor
-extended with some kind of (static or dynamic, software- or
-hardware-enforced) security monitor. However, our definition of stack
-safety is generic: we will begin with an abstract interface to our
-machine model and only make specific choices about details when we get
-to the policy enforcement section~\ref{sec:micropolicy}.
+and later extend it with a dynamic policy mechanism.
 
@@
 
+\leo{TODO: Forward pointer to policies right?}
+
 \paragraph{Traces}
@@
 triples of (machine states, policy states, and observations), usually
-denoted $\MPT$. To continue, we will require three generic operations
+denoted $\MPT$.
+\rb{The way the definitions are set up, one may expect MPO, but\ldots}
+To continue, we will require four generic operations
 on traces: a ``trace-of'' operation to construct the potentially
-infinite traced induced by a step relation; a ``split'', to break a
+infinite traced induced by a step relation
+\rb{not an op on traces per se?};
+a ``split'', to break a
 trace into two components based on a predicate; a ``join'', the
@@
 then \(\machT\) must be infinite.
+
+\leo{TODO: Add an example of a trace once I figure out the notations}
 
@@
 
+% SNA: fixed
+%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
+%  seems better}
 The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
@@
 
-\subsection{Program Structre}
+\subsection{Running Example}
+
+\leo{TODO: Figure out presentation}
+
+\newcommand*{\add}{\text{add}}
+\newcommand*{\addi}{\text{addi}}
+\newcommand*{\sw}{\text{sw}}
+\newcommand*{\lw}{\text{lw}}
+\newcommand*{\jal}{\text{jal}}
+\newcommand*{\jalr}{\text{jalr}}
+%\begin{lstlisting}
+%START:
+%  addi r2 r2 12 ; allocate 3 words in the stack
+%  sw r2 42 -8; store 42 in a private location
+%  sw r2 0 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 F ; call F and store return address
+% 
+%F:
+%  addi r2 r2 4 ; allocate for ra 
+%  sw r2 r1 -4 ; store the return address to sp-4
+%  addi r2 r2 8 ; allocate for arg + ret
+%  sw r2 17 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 G ; call G and store return address
+% 
+%G:  
+%  addi r2 r2 4 ; allocate for ra 
+%  sw r2 r1 -4 ; store the return address to sp-4
+%  
+%\end{lstlisting}
+
+Starting State:
+
+Instruction memory in locations 0 - ..
+PC at 100.
+Stack memory starting at 1000.
+Assume we're in a function. 
+Stack Pointer at 1040.
+1040: 42.
+
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
+\MemoryLabel{35em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{4}{gray}[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-18.7em}{0.6em}{42}
+\MemoryLabel{-16.4em}{0.6em}{0}
+\\
+\end{center}
+
+Program:
+
+\begin{verbatim}
+top{
+  return f(0);
+}
+
+f(x){
+  return (x + g(17));
+}
+
+g(x){
+  return (x + 1);
+}
+\end{verbatim}
+
+\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
+\newcommand*{\clabel}[1]{#1: & & &\\}
+\newcommand*{\rsp}{sp}
+\newcommand*{\rra}{ra}
+\[
+\begin{array}{llll}
+  \clabel{TOP}  
+  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
+  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
+  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
+  & & & \\
+  & & & \\
+  \clabel{F}
+  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
+  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
+  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
+  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
+  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
+  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
+  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
+  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
+  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
+  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
+  & & & \\
+  & & & \\  
+  \clabel{G}
+  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
+  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
+  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
+  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
+
+%  addi r2 r2 8 ; allocate for arg + ret
+%  sw r2 17 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 G ; call G and store return address
+
+\end{array}
+\]
+
+%  sw r2 42 -8; store 42 in a private location
+%  sw r2 0 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 F ; call F and store return address
+
+%  0: & \addi ~ r_4 ~ r_0 ~ 40 &
+%     ; ~ \text{\em put $40$ in register $r_4$}\\
+%  4: & \sw ~ r_4 ~ r_4 ~ 1000 &
+%     ; ~ \text{\em store $40$ in location $1040$}\\
+%  8: & \jal ~ r_1 ~ 20        &
+%     ; ~ \text{\em jump to $20$, store return addres in $r_1$}\\
+%  12: & & \\
+%  16: & & \\
+%  20: & \sw ~ r_2 ~ r_1 ~ 4 &
+%     ; ~ \text{\em store the return address to the $\SP + 4$}\\
+%  24: & \addi ~ r_2 ~ r_2 ~ 8 &
+%     ; ~ \text{\em increase the $\SP$ by 8}\\
+%\end{array}
+%\]
+% 
+%  , (SW sp ra 4  , tagH1)
+%  , (ADDI sp sp 8, tagH2)
+%  , (LW 3 3 1000, instrTag)
+%  ]
+
+\subsection{Program Structure}
 

### 2020-07-01 main.tex
Comments / polishing

@@
   \end{minipage}
-  \caption{Examples of Unsafe Programs}
+  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
+      the boxes; the program fragments need comments explaining what each
+      line does.}}
   \label{fig:stackunsafety}
@@
 techniques have been proposed in the literature\leo{cite, point out?
-  stack canaries, cherry, nicks...} aiming to achieve {\em stack
-  safety}.
-
-But what {\em is} stack safety? A precise formal definition of stack
-safety is lacking from the literature, with the notion only being
+  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
+  stronger point.} aiming to achieve ``stack
+  safety''.
+
+But what {\em is} this stack safety of which we speak? A precise formal
+definition is lacking in the literature, with the notion usually being
 defined negatively: through examples of stack unsafety
-\rb{usually though not always; as noted below there are other approaches}. Two classical
+\rb{usually though not always; as noted below there are other
+  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
 examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
 \leo{Goal: on the left, a bad read; on the right, a bad write.}
-\leo{Do we also need an example of ROP/control flow for the new part?}
-Recently, a \leo{inspiring/awesome/something good} work reasoned
+\leo{Do we also need an example of ROP/control flow for the new
+  part?}\bcp{If it is not too complicated to explain, we should definitely
+  put it
+  here.  If it is complicated, we can just say in text that there are other
+  sorts of attacks too, and that we deal with them, and give a forward
+  pointer.}
+Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
+  we are only aware of one, quite recent, attempt to do something better} work reasoned
 about a stack safety enforcement mechanism using capabilities\leo{cite};
-while impressive, stack safety was defined with respect to an abstract
-machine that ``exhibits stack safety by construction''. Which begs
-the question, what {\em is} the property that holds by construction?
+but even here, stack safety is defined with respect to an abstract
+machine that ``exhibits stack safety by construction,'' begging
+the question of what {\em is} the property that holds by construction?
 
@@
 by informally describing two components that comprise stack safety:
-local state encapsulation and well-bracketed control flow. In this paper
-we turn this intuition into two formal \leo{extensional} security properties,
+local state encapsulation and well-bracketed control flow. The first
+contribution of the present paper is to
+turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
+  claim this} security properties,
 that together form a formal definition of stack safety.
 \rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
-rigor in it, even if some notions are implicit or not fully developed.}
+rigor in it, even if some notions are implicit or not fully
+developed.}\bcp{Yes, and let's remember that they are highly likely to be
+reviewing this paper!  Let's be as generous as we can be without
+under-selling our delta.}
 
@@
 able to access or overwrite a caller's private data. We formalize this
-intuition as a security property, with a confidentiality and an
-integrity component.
-%
-At a high level, {\em stack confidentiality} ensures that a callee's
+intuition as a security property with both confidentiality and
+integrity components.
+%
+Intuitively, {\em stack confidentiality} promises that a callee's
 observable behavior is not affected by memory outside its stack frame
 (e.g. a caller's private data or uninitialized memory).  {\em Stack
-  integrity} guarantees that if a function call returns, the caller's
-stack data will be identical to what it was before the call.
+  integrity} guarantees that, if a function call ever returns, the caller's
+stack data at the point of return will be identical to what it was before
+the call.  \bcp{We need to point out somewhere that we are only dealing with
+stack safety, though we discuss (somewhere) how our property can be combined
+with the heap safety property of [Azevedo et al]}
 
 While integrity and confidentiality are relatively standard security
-properties, this setting provides an interesting additional challenge:
+properties, the stack-safety setting raises an interesting additional
+challenge:
 the notion of who the attacker is shifts throught execution. Not only
 does a caller need to be protected from its callee, but the callee
-itself needs to be potentially proteced from {\em its} callees.  In
-this paper we address this challenge by varying the observation power
+itself needs to be proteced from {\em its} callees.
+We address this challenge by varying the observation power
 of the attacker at call points, and demonstrate that our policy is
-implementable by pre-existing monitoring mechanisms.
+implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
+  the sentence are about two very different things.  The second deserves its
+  own paragraph, IMO.}
 
@@
 the need to protect the call structure itself: a callee should not be
-allowed to jump to the middle of an arbitrary block of code either
-directly (via a jump instruction) or indirectly (by corrupting stack
-data and returning). \leo{There needs to be some informal description
-  of the wbcf part here. Sean?}
-
-We begin by giving some background in Section~\ref{sec:prelim}.
-Our contributions are:
+allowed to jump to the middle of an arbitrary block of code, either
+directly (via a jump instruction) or indirectly (by corrupting the return
+address on the stack and then returning). \leo{There needs to be some
+  informal description of the wbcf part here. Sean?}
+
+In summary, we offer the following contributions:
 \begin{itemize}
@@
 \end{itemize}
-Finally, in Section~\ref{sec:relwork} we discuss related work and
-sketch directions for future work in Section~\ref{sec:future}.
+We begin with some technical preliminaries in
+Section~\ref{sec:prelim}. Section~\ref{sec:relwork} discusses related work, and
+Section~\ref{sec:future} sketches directions for future work.
 

### 2020-07-01 main.tex
More polishing on the contributions (but they need even more!)

@@
 \begin{itemize}
-\item \leo{More? Extensional} formal characterizations of the
-  intuitive concepts of local state encapsulation~\ref{sec:lse} and
-  well-bracketed control flow~\ref{sec:wbcf}.  Evidence that this
-  property is strong enough to capture the folk intuition: we show
+\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
+  intuitive concepts of local state encapsulation and
+  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
+  property is strong enough to capture the folk intuition, we show
   that the stack unsafe examples from the literature also fail our
-  property.
-\item Evidence that the property is not too strong. It's realizible by
+  property~(\Cref{sec:disc}).
+\item As evidence that the property is not too strong, we show that it is
+  realizible by
   a previously proposed micropolicy enforcement
-  mechanism~\ref{sec:micropolicy}. \leo{And it says yes on a
-    substantial body of code} We then present a variation of the
+  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
+    substantial body of code}. We also present a variation of the
   property + policy that is lazy, easier/more efficient to implement,
-  and provably weaker.
+  and provably weaker (\Cref{sec:lazy}).
 \item A provably stronger (inductive? step-by-step/stepwise) variant
   of this property tuned for testing/verification.  We test the whole
-  thing with a cool framework.
+  thing with a cool framework. \bcp{Which section?}
 \end{itemize}
+Section~\ref{sec:relwork} discusses related work, and
+Section~\ref{sec:future} sketches directions for future work.
+%
 We begin with some technical preliminaries in
-Section~\ref{sec:prelim}. Section~\ref{sec:relwork} discusses related work, and
-Section~\ref{sec:future} sketches directions for future work.
+Section~\ref{sec:prelim}.
 
@@
 \section{Stack Safety, Formally}
+\label{sec:lse-and-wbcf}
 
@@
 \section{Discussion}
+\bcp{Make this a subsection?}
+\label{sec:disc}
 
@@
 \section{Lazy Stack Safety}
+\label{sec:lazy}
 

### 2020-07-01 main.tex
Comments and polishing on 2

@@
 
-Before we define stack safety, we need to first set the stage by
-describing our machine model. Our definition of stack safety is
-generic and this sections discusses an abstract interface to our
+To set the stage, we first
+describe our machine model. Our definition of stack safety is
+generic and this section discusses an abstract interface to our
 machine model and some kind of static or dynamic, software- or
-hardware-enforced security monitor. At the same time, for
-concreteness, we will build up our running example using a
-conventional machine architecture with a standard (RISC-V) processor
-and later extend it with a dynamic policy mechanism.
+hardware-enforced security monitor.  For
+concreteness, our running examples will use a
+conventional machine architecture and a standard (RISC-V) instruction set;
+in~\Cref{sec:micropolicies} we will
+extend this model it with a dynamic policy enforcement mechanism based
+on \bcp{NickPaper,AndMaybeSomeOthers}.
 
@@
 The basic building blocks of the machine are {\em values} and {\em
-  addresses}. Both are drawn from some set of all possible words
-$\WORDS$. We will denote words with $\word$.
-%
-Our machine states are comprised of {\em components}, that can be word
-(denoting memory addresses) or registers (drawn from some set $\REGS$).
-%
-    \[\reg \in \REGS\]
+  addresses}. Both are drawn from some set of {\em words}
+$\WORDS$, ranged over by $\word$.
+%
+Our machine states are composed of {\em components} $k$, addressed either by
+words (memory addresses) or by by register names ($r$) drawn from some set
+$\REGS$.
+%
     \[\component \in \COMPONENTS = \WORDS + \REGS \]
 %
-The registers must include two special purpose registers: the program
-counter {\PCname} and the stack pointer \SP.
-%
-Formally, a machine state is just a map from components to values:
+The register names are assumed to include two special purpose registers: the
+program counter {\PCname} and the stack pointer \SP.
+
+A machine state is a map from components to values:
 %
@@
 
-Our stack safety property will be defined in terms of {\em
-  observations}, which may be silent (written $\tau$) or a word of
-data:\leo{Needs more exposition}
-%
-\[\obs \in \OBSS = \tau ~ | ~ \word \in \WORDS\]
-%
-These observations are a byproduct of the step function of the
-machine, which takes a machine state and produces a pair of a
-machine state and an observation.
-%
-\(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \).
+The step function of the
+machine takes a machine state and produces a pair of a
+machine state and an {\em observation}, which may be either
+silent (written $\tau$) or a word of data that the machine communicates to
+the outside world (e.g., by writing to a DMA address):\leo{Needs more
+  exposition}\bcp{Like that?}
+%
+\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
+\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
+%
 
@@
 is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
-r_{31}\} \cup {\PCname}$, and $\SP = r_2$. \leo{Mention We use the
-  standard encoding of instructions to words?}
+r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$.
+% \leo{Mention We use the standard encoding of instructions to words?}
+% BCP: Nah.
 
@@
 we could pick a particular component and treat it as the ``output'' of
-our system. }
+our system. }  \bcp{But in this paragraph we are describing the particular
+choise that we make for our running example, no?  But we certainly want to
+discuss this fully someplace!}
 
@@
 
-A policy is a general model of an enforcement mechanism, consisting of
+\bcp{More explanation needed, plus examples.}A policy is a general model of an enforcement mechanism, consisting of
 a set of policy states \(\pol \in \POLS\), and a policy step function
@@
 %
-We then lift this policy step function to operate on such pairs,
+We lift the policy step function to operate on such pairs
 by combining it with the regular step function for machine states.
@@
 
-\leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted step?}
-
-\leo{TODO: Forward pointer to policies right?}
+\leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted
+  step?}\bcp{Yes.}
+
+\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
+in more detail that the intention is to capture any kind of dynamic
+enforcement mechanism and give as many names and citations as we can.}
 
@@
 
-A trace over some type $A$ is a potentially infinite, nonempty
+A trace over a type $A$ is a nonempty, finite or infinite
 sequence of $A$s. We will mostly be concerned with sequences of
-machines states, usually denoted $\machT$, and with sequences of
+machines states, usually\bcp{not always?} denoted $\machT$, and with sequences of
 triples of (machine states, policy states, and observations), usually
@@
 \rb{The way the definitions are set up, one may expect MPO, but\ldots}
-To continue, we will require four generic operations
-on traces: a ``trace-of'' operation to construct the potentially
-infinite traced induced by a step relation
-\rb{not an op on traces per se?};
+To continue\bcp{??}, we will require four generic operations
+on traces: a ``trace-of'' operation to construct the
+trace induced by a step relation and a machine state
+\rb{not an op on traces per se?}\bcp{+1};
 a ``split'', to break a
@@
 reverse of split; and a ``prefix'', as a shorthand to only keep the
-first component of a split.
+first component of a split.  \bcp{Not sure it's worth enumerating them
+  rather than just getting down to the actual definitions below.}
 
@@
 relates a machine state with the trace of machine states paired with
-observations produced by repeated application of step:
+observations produced by repeated application of step:  \bcp{above, we
+  showed operators applied to metavariables, not just the bare operator
+  symbol.}
 %
@@
 \leo{TODO: Add an example of a trace once I figure out the notations}
+\bcp{Let's also try to clean up the spacing a little -- many of these
+  sequences of rules can be typeset on fewer lines}
 
@@
 state pair to the trace of triples of machine states, policy states,
-and observations induced by by \(\stepstoobs{}_P\):
+and observations induced by by \(\stepstoobs{}_P\): \bcp{Observe that this
+  time the trace can be either finite or infinite.}
 %
@@
 out of a trace, \(\pi_p\) to project the trace of policy states, and
-\(\pi_o\) to project the observations.
+\(\pi_o\) to project the observations.  \bcp{I'm getting a little lost in
+  all these.  E.g., do these traces include the final state of a finite
+  execution?  If so, what observation is it paired with?}
 
@@
 %  seems better}
-The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
+The {\it prefUpTo} operation \(\PUT(\Trace,f)\) \bcp{Why does it have two
+  different abbreviated names?  And wouldn't ``While'' be more mnemonic?}
+takes a trace \(T\) 
 and a predicate on elements \(f\) and gives the prefix of \(T\)
@@
 Note that if the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. Conversely, if the predicate never
+predicate must hold on its final element. Conversely\bcp{why conversely?}, if the predicate never
 holds, \(\PUT\) is the identity function.

### 2020-07-01 main.tex
Comments / polishing on section 2.

@@
 \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
-running \(\mach\) up to \(f\)''):
+running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I think.}
 
@@
 We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-\leo{Benjamin suggested ``as far as an attacker is conerned in an
-  e-mail, but we haven't said anything about attackers yet.}, if the
+\leo{Benjamin suggested ``as far as an attacker is concerned'' in an
+  e-mail, but we haven't said anything about attackers yet.}\bcp{We don't
+  have to talk about attackers, just external observers.}, if the
 sequence of non-silent observations of $\obsT_1$ is a prefix of those
@@
 
-We then define equivalence of observation traces as traces prefixing each other:
+We then define equivalence\bcp{Can we please call it something else?  An
+  ``equivalence'' should always be transitive.} of observation traces as traces prefixing each other:
 \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
@@
 sense in a timing-insensitive context: an external observer looking
-at two machine runs cannot distinguish between a machine that
-steps forever or a machine that steps for a long time before producing
-some output, before that time has passed.
+at two machine runs cannot (computably) distinguish between a machine that
+steps forever and a machine that steps for a long time before producing
+some output.
 

### 2020-07-01 main.tex
CP - 3

--- 2020-07-01 main.tex
@@
   \end{minipage}
-  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
-      the boxes; the program fragments need comments explaining what each
-      line does.}}
+  \caption{Examples of Unsafe Programs}
   \label{fig:stackunsafety}
@@
 techniques have been proposed in the literature\leo{cite, point out?
-  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
-  stronger point.} aiming to achieve ``stack
-  safety''.
-
-But what {\em is} this stack safety of which we speak? A precise formal
-definition is lacking in the literature, with the notion usually being
+  stack canaries, cherry, nicks...} aiming to achieve {\em stack
+  safety}.
+
+But what {\em is} stack safety? A precise formal definition of stack
+safety is lacking from the literature, with the notion only being
 defined negatively: through examples of stack unsafety
-\rb{usually though not always; as noted below there are other
-  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
+\rb{usually though not always; as noted below there are other approaches}. Two classical
 examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
 \leo{Goal: on the left, a bad read; on the right, a bad write.}
-\leo{Do we also need an example of ROP/control flow for the new
-  part?}\bcp{If it is not too complicated to explain, we should definitely
-  put it
-  here.  If it is complicated, we can just say in text that there are other
-  sorts of attacks too, and that we deal with them, and give a forward
-  pointer.}
-Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
-  we are only aware of one, quite recent, attempt to do something better} work reasoned
+\leo{Do we also need an example of ROP/control flow for the new part?}
+Recently, a \leo{inspiring/awesome/something good} work reasoned
 about a stack safety enforcement mechanism using capabilities\leo{cite};
-but even here, stack safety is defined with respect to an abstract
-machine that ``exhibits stack safety by construction,'' begging
-the question of what {\em is} the property that holds by construction?
+while impressive, stack safety was defined with respect to an abstract
+machine that ``exhibits stack safety by construction''. Which begs
+the question, what {\em is} the property that holds by construction?
 
@@
 by informally describing two components that comprise stack safety:
-local state encapsulation and well-bracketed control flow. The first
-contribution of the present paper is to
-turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
-  claim this} security properties,
+local state encapsulation and well-bracketed control flow. In this paper
+we turn this intuition into two formal \leo{extensional} security properties,
 that together form a formal definition of stack safety.
 \rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
-rigor in it, even if some notions are implicit or not fully
-developed.}\bcp{Yes, and let's remember that they are highly likely to be
-reviewing this paper!  Let's be as generous as we can be without
-under-selling our delta.}
+rigor in it, even if some notions are implicit or not fully developed.}
 
@@
 able to access or overwrite a caller's private data. We formalize this
-intuition as a security property with both confidentiality and
-integrity components.
-%
-Intuitively, {\em stack confidentiality} promises that a callee's
+intuition as a security property, with a confidentiality and an
+integrity component.
+%
+At a high level, {\em stack confidentiality} ensures that a callee's
 observable behavior is not affected by memory outside its stack frame
 (e.g. a caller's private data or uninitialized memory).  {\em Stack
-  integrity} guarantees that, if a function call ever returns, the caller's
-stack data at the point of return will be identical to what it was before
-the call.  \bcp{We need to point out somewhere that we are only dealing with
-stack safety, though we discuss (somewhere) how our property can be combined
-with the heap safety property of [Azevedo et al]}
+  integrity} guarantees that if a function call returns, the caller's
+stack data will be identical to what it was before the call.
 
 While integrity and confidentiality are relatively standard security
-properties, the stack-safety setting raises an interesting additional
-challenge:
+properties, this setting provides an interesting additional challenge:
 the notion of who the attacker is shifts throught execution. Not only
 does a caller need to be protected from its callee, but the callee
-itself needs to be proteced from {\em its} callees.
-We address this challenge by varying the observation power
+itself needs to be potentially proteced from {\em its} callees.  In
+this paper we address this challenge by varying the observation power
 of the attacker at call points, and demonstrate that our policy is
-implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
-  the sentence are about two very different things.  The second deserves its
-  own paragraph, IMO.}
+implementable by pre-existing monitoring mechanisms.
 
@@
 the need to protect the call structure itself: a callee should not be
-allowed to jump to the middle of an arbitrary block of code, either
-directly (via a jump instruction) or indirectly (by corrupting the return
-address on the stack and then returning). \leo{There needs to be some
-  informal description of the wbcf part here. Sean?}
-
-In summary, we offer the following contributions:
+allowed to jump to the middle of an arbitrary block of code either
+directly (via a jump instruction) or indirectly (by corrupting stack
+data and returning). \leo{There needs to be some informal description
+  of the wbcf part here. Sean?}
+
+We begin by giving some background in Section~\ref{sec:prelim}.
+Our contributions are:
 \begin{itemize}
-\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
-  intuitive concepts of local state encapsulation and
-  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
-  property is strong enough to capture the folk intuition, we show
+\item \leo{More? Extensional} formal characterizations of the
+  intuitive concepts of local state encapsulation~\ref{sec:lse} and
+  well-bracketed control flow~\ref{sec:wbcf}.  Evidence that this
+  property is strong enough to capture the folk intuition: we show
   that the stack unsafe examples from the literature also fail our
-  property~(\Cref{sec:disc}).
-\item As evidence that the property is not too strong, we show that it is
-  realizible by
+  property.
+\item Evidence that the property is not too strong. It's realizible by
   a previously proposed micropolicy enforcement
-  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
-    substantial body of code}. We also present a variation of the
+  mechanism~\ref{sec:micropolicy}. \leo{And it says yes on a
+    substantial body of code} We then present a variation of the
   property + policy that is lazy, easier/more efficient to implement,
-  and provably weaker (\Cref{sec:lazy}).
+  and provably weaker.
 \item A provably stronger (inductive? step-by-step/stepwise) variant
   of this property tuned for testing/verification.  We test the whole
-  thing with a cool framework. \bcp{Which section?}
+  thing with a cool framework.
 \end{itemize}
-Section~\ref{sec:relwork} discusses related work, and
-Section~\ref{sec:future} sketches directions for future work.
-%
-We begin with some technical preliminaries in
-Section~\ref{sec:prelim}.
+Finally, in Section~\ref{sec:relwork} we discuss related work and
+sketch directions for future work in Section~\ref{sec:future}.
 
@@
 
-% SNA: fixed
-%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
-%  seems better}
+\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
+  seems better}
 The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
@@
 \section{Stack Safety, Formally}
-\label{sec:lse-and-wbcf}
 
 In this section we will describe stack safety as a formal property. We
-begin with our model of knowledge about the program structure\leo{I
-  hate this}, and then dive in to local state encapsulation and
+begin with introducing a running example that will help keep things
+concrete throughout this section, continue with our model of
+\leo{compiler-provided?} knowledge about the program structure, and
+then dive in to formal definitions local state encapsulation and
 well-bracketed control flow.
 
+\begin{figure}[b]
+  \begin{centering}
+    \begin{minipage}{0.5\textwidth}
+\begin{verbatim}
+f{
+  int x, y;
+  ...
+  y = g(0);
+  ...
+}
+
+g(x){
+  return (x + h(17));
+}
+
+h(x){
+  return (x + 1);
+}
+\end{verbatim}
+\end{minipage}
+\end{centering}
+\caption{High-level Program with Simple Nested Calls}
+\label{fig:running-program}
+\end{figure}
+
 \subsection{Running Example}
 
-\leo{TODO: Figure out presentation}
+Our running example consists of three functions $f$, $g$, and $h$ that
+perform a nested sequence of calls. A high level C-like view of the
+excerpt program can be found in Figure~\ref{fig:running-program}.
+Suppose that the starting state for our running example is right
+before $f$'s call to $g$. A rough example layout of the memory at that
+point can be found in the top of Figure~\ref{fig:running-trace-b}:
+\begin{itemize}
+\item The low part of the memory (0-1000) is dedicated to instructions.
+\item The stack starts at location 1000.
+\item The stack of $f$ starts at 1040.
+\item The location of $x$ is 1040 and stores $42$.
+\item The location of $y$ is 1044 and stores $0$.
+\end{itemize}
+Directly below the layout, we summarize the parts of the register file
+that are relevant to the example.
+
+The rest of Figure~\ref{fig:running-trace-b} detail the execution of the
+program, showing the machine state at call and return points, as well
+as describing the full effect of each instruction in detail.
+\leo{Unoptimized, etc.}
 
@@
 \newcommand*{\jalr}{\text{jalr}}
-%\begin{lstlisting}
-%START:
-%  addi r2 r2 12 ; allocate 3 words in the stack
-%  sw r2 42 -8; store 42 in a private location
-%  sw r2 0 -4; store 0 as an argument
-%  sw r2 0 0; zero out the return location
-%  jal r1 F ; call F and store return address
-%
-%F:
-%  addi r2 r2 4 ; allocate for ra
-%  sw r2 r1 -4 ; store the return address to sp-4
-%  addi r2 r2 8 ; allocate for arg + ret
-%  sw r2 17 -4; store 0 as an argument
-%  sw r2 0 0; zero out the return location
-%  jal r1 G ; call G and store return address
-%
-%G:
-%  addi r2 r2 4 ; allocate for ra
-%  sw r2 r1 -4 ; store the return address to sp-4
-%
-%\end{lstlisting}
-
-Starting State:
-
-Instruction memory in locations 0 - ..
-PC at 100.
-Stack memory starting at 1000.
-Assume we're in a function.
-Stack Pointer at 1040.
-1040: 42.
-
+\newcommand*{\rsp}{\mathit{sp}}
+\newcommand*{\rra}{\mathit{ra}}
+
+\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}
+
+\begin{figure}
+
+%Starting Memory Layout
 \begin{center}
@@
 \memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{4}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-18.7em}{0.6em}{42}
-\MemoryLabel{-16.4em}{0.6em}{0}
+\memory{8}{gray}[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-29.6em}{0.6em}{42}
+\MemoryLabel{-27.4em}{0.6em}{0}
 \\
 \end{center}
-
-Program:
-
-\begin{verbatim}
-top{
-  return f(0);
-}
-
-f(x){
-  return (x + g(17));
-}
-
-g(x){
-  return (x + 1);
-}
-\end{verbatim}
-
-\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
-\newcommand*{\clabel}[1]{#1: & & &\\}
-\newcommand*{\rsp}{sp}
-\newcommand*{\rra}{ra}
+% Starting Register File
 \[
-\begin{array}{llll}
-  \clabel{TOP}
-  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
-  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
-  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
-  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
-  & & & \\
-  & & & \\
-  \clabel{F}
-  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
-  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
-  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
-  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
-  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
-  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
-  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
-  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
-  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
-  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
-  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
-  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
-  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
-  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
-  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
-  & & & \\
-  & & & \\
-  \clabel{G}
-  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
-  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
-  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
-  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
-  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
-  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
-  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
-  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
-
-%  addi r2 r2 8 ; allocate for arg + ret
-%  sw r2 17 -4; store 0 as an argument
-%  sw r2 0 0; zero out the return location
-%  jal r1 G ; call G and store return address
-
+\begin{array}{cccccc}
+  PC: 100 & r_0 : 0 & \rra : ? & \rsp : 1048 & r_4 : ? & r_5 : ? 
 \end{array}
 \]
+\vspace*{0.2em}
+\[
+  \begin{array}{r|c|c|c|c}
+    & PC & \mathit{Instruction} & \mathit{Effect} &  \\
+    \hline
+    \tracerow{1}{100}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1056}{Allocate call frame}
+    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ -8}{[1048] \leftarrow 0}{Store 0 as argument}
+    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ -4}{[1052] \leftarrow 0}{Zero out result}
+    \tracerow{4}{112}{\jal ~ \rra ~ 140}{PC \leftarrow 140; ~ \rra \leftarrow 116}{Make the call}
+  \end{array}
+  \]
+~ \\
+% At Call to G
+% Memory Layout
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{41em}{1.8em}{$\downarrow$}
+\MemoryLabel{40.8em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
+\memory{6}{gray}[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-29.4em}{0.6em}{42}
+\MemoryLabel{-27.2em}{0.6em}{0}
+\MemoryLabel{-25.0em}{0.6em}{0}
+\MemoryLabel{-22.8em}{0.6em}{0}
+\\
+\end{center}
+% Register File
+\[
+\begin{array}{cccccc}
+  PC: 140 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : ? & r_5 : ? 
+\end{array}
+\]
+% G - Body - 1
+\vspace*{0.2em}
+\[
+  \begin{array}{r|c|c|c|c}
+    \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
+    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ -4}{[1056] \leftarrow 116}{Store return address}
+    \tracerow{7}{148}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1068}{Allocate for call frame}
+    \tracerow{8}{152}{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow 17}{Store 17 to $r_4$}
+    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ -8}{[1060] \leftarrow 17}{Store 17 as the argument}
+    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ -4}{[1064] \leftarrow 0 }{Zero out return}
+    \tracerow{11}{164}{\jal ~ \rra ~ 200}{PC \leftarrow 200; ~ \rra \leftarrow 168}{Make the call to $h$}
+  \end{array}
+  \]
+  ~ \\
+% At Call to H
+% Memory Layout
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{49em}{1.8em}{$\downarrow$}
+\MemoryLabel{48.8em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
+\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
+\memory{3}{gray}[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-29.4em}{0.6em}{42}
+\MemoryLabel{-27.2em}{0.6em}{0}
+\MemoryLabel{-25.0em}{0.6em}{0}
+\MemoryLabel{-22.8em}{0.6em}{0}
+\MemoryLabel{-20.6em}{0.6em}{116}
+\MemoryLabel{-18.4em}{0.6em}{17}
+\MemoryLabel{-16.2em}{0.6em}{0}
+\\
+\end{center}
+% Register File
+\[
+\begin{array}{cccccc}
+  PC: 200 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 17 & r_5 : ? 
+\end{array}
+\]
+% H - Body
+\vspace*{0.2em}
+\[
+\begin{array}{r|c|c|c|c}
+  \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
+  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ -4}{[1068] \leftarrow 168}{Store return address}
+  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ -12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
+  \tracerow{15}{212}{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{Increment $r_4$ by 1}\end{array}
+\]
+% At Middle of H
+% Memory Layout
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{52em}{1.8em}{$\downarrow$}
+\MemoryLabel{51.8em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
+\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
+\memory{1}{cyan}%
+\memory{2}{gray}[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-29.4em}{0.6em}{42}
+\MemoryLabel{-27.2em}{0.6em}{0}
+\MemoryLabel{-25.0em}{0.6em}{0}
+\MemoryLabel{-22.8em}{0.6em}{0}
+\MemoryLabel{-20.6em}{0.6em}{116}
+\MemoryLabel{-18.4em}{0.6em}{17}
+\MemoryLabel{-16.2em}{0.6em}{0}
+\MemoryLabel{-14.0em}{0.6em}{168}
+\\
+\end{center}
+% Register File
+\[
+\begin{array}{cccccc}
+  PC: 212 & r_0 : 0 & \rra : 168 & \rsp : 1072 & r_4 : 18 & r_5 : ? 
+\end{array}
+\]
+\vspace*{0.2em}
+\[
+\begin{array}{r|c|c|c|c}
+  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ -8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
+  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 168}{Load return address}
+  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}
+  \tracerow{19}{228}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 168}{Return to $g$}
+  \end{array}
+\]
+\label{fig:running-trace-a}
+\end{figure}
+
+\begin{figure}
+% After Return from H
+% Memory Layout
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{49em}{1.8em}{$\downarrow$}
+\MemoryLabel{48.8em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
+\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
+\memory{3}{gray}
+~$\cdots$
+\MemoryLabel{-29.4em}{0.6em}{42}
+\MemoryLabel{-27.2em}{0.6em}{0}
+\MemoryLabel{-25.0em}{0.6em}{0}
+\MemoryLabel{-22.8em}{0.6em}{0}
+\MemoryLabel{-20.6em}{0.6em}{116}
+\MemoryLabel{-18.4em}{0.6em}{17}
+\MemoryLabel{-16.2em}{0.6em}{18}
+\\
+\end{center}
+% Register File
+\[
+\begin{array}{cccccc}
+  PC: 168 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 18 & r_5 : ? 
+\end{array}
+\]
+% H - Body
+\vspace*{0.2em}
+\[
+\begin{array}{r|c|c|c|c}
+  \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
+  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ -20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
+  \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
+  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ -8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
+  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1060}{Deallocate local state}
+  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 116}{Load return address}
+  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
+  \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
+\end{array}
+\]
+% At Call to G
+% Memory Layout
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{41em}{1.8em}{$\downarrow$}
+\MemoryLabel{40.8em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
+\memory{6}{gray}
+~$\cdots$
+\MemoryLabel{-29.4em}{0.6em}{42}
+\MemoryLabel{-27.2em}{0.6em}{0}
+\MemoryLabel{-25.0em}{0.6em}{0}
+\MemoryLabel{-22.8em}{0.6em}{18}
+\\
+\end{center}
+% Register File
+\[
+\begin{array}{cccccc}
+  PC: 116 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : 18 & r_5 : 0
+\end{array}
+\]
+\vspace*{0.2em}
+\[
+\begin{array}{r|c|c|c|c}
+  \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
+  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ -12}{[1048] \leftarrow 18}{Set $y$ to the return value}
+  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
+\end{array}
+\]
+\vspace*{0.2em}  
+% Ending Memory Layout
+\begin{center}
+\MemoryLabel{1.7em}{1.8em}{0}
+\MemoryLabel{21.2em}{1.8em}{1000}
+\MemoryLabel{29.7em}{1.8em}{1040}
+\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
+\MemoryLabel{35em}{3em}{SP}
+\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+~$\cdots$
+\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+~$\cdots$
+\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{8}{gray}
+~$\cdots$
+\MemoryLabel{-29.6em}{0.6em}{42}
+\MemoryLabel{-27.4em}{0.6em}{0}
+\\
+\end{center}
+% Starting Register File
+\[
+\begin{array}{cccccc}
+  PC: 128 & r_0 : 0 & \rra : 116 & \rsp : 1048 & r_4 : 18 & r_5 : 0
+\end{array}
+\]
+\caption{Detailed Execution Trace}
+\label{fig:running-trace-b}
+\end{figure}
+
+%\end{minipage}
+%\begin{minipage}{0.7\textwidth}
+%\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
+%\newcommand*{\clabel}[1]{#1: & & &\\}
+%\[
+%\begin{array}{llll}
+%  \clabel{TOP}  
+%  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
+%  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
+%  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+%  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
+%  & & & \\
+%  & & & \\
+%  \clabel{F}
+%  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+%  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+%  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
+%  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
+%  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
+%  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+%  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
+%  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
+%  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
+%  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
+%  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
+%  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
+%  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+%  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+%  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
+%  & & & \\
+%  & & & \\  
+%  \clabel{G}
+%  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+%  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+%  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
+%  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
+%  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
+%  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+%  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+%  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
+%\end{array}
+%\]
+%\end{minipage}
+%\label{fig:running-program}
+%\end{figure}
 
@@
 \[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]
-
-\leo{Visible from assembly}
+%
+In our running example, instructions 100-124 correspond to $f$,
+instructions 140-196 correspond to $g$, and instructions 200-228
+correspond to $h$.
 
 A call map is a partial function from addresses to a natural number,
-the number of arguments of the call.
+the number of arguments and return values of the call.
   %
 \[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]
-
-\leo{The two JAL instructions map to 2 (arg + return)}
+%
+In our running example, the callmap would contain (at least) two
+entries: the word 112 corresponding to the first $\jal$
+instruction to $g$, which would map to $2$ as the call has one
+argument; and the word 164 also mapping to $2$ corresponding
+to the call to $h$.
 
@@
 \[\entmap \in \ENTMAPS \subseteq \WORDS\]
-
-\leo{retmap is the jal + 1}
-\leo{entrymap is the ones after the labels}
+%
+In our running example, the return map identifies the $\jalr$
+instructions 196 and 224, while the entry map identifies
+instructions 140 and 200.
 
@@
 \label{def:trace-integrity}
+
 Let $\contour$ be a contour and $MP$ be a (potentially infinite)
@@
 \definition{Transition}
+
 Given a trace $MP$, a call map $\callmap$ and a code map $\codemap$,
-an $n$-{\em transition} to a callee $f$ is a pair of machine-policy
+an $n-${\em transition} to a callee $f$ is a pair of machine-policy
 state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ in the trace
@@
 \definition{Stack Integrity}
+
 Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
@@
 \definition{Variations}
+
 We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
@@
 \definition{Trace Confidentiality}
+
 \leo{Add info for ``valid return'':}
@@
 \definition{Confidential Execution}
+
   Stack confidentiality, below, is a property of a system that, from a call to
@@
 \definition{Stack Confidentiality}
-Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
+
+Let $\callmap$ be a call map and $\codemap$ a code map.
+\rb{Could we find better symbols for the maps?}
+We say that a
 system satisfies {\em stack confidentiality} with respect to
@@
 \definition{Control Separation}
+
 A system enjoys {\em control separation} with regard to a code map
@@
 \definition{Entry Integrity}
+
 A system enjoys {\em entry integrity} with regard to a call map
@@
 \definition{Return Integrity}
+
 Return integrity captures the expected behavior of marked returns with
@@
 \definition{Well Bracketed Control Flow}
+
 The above properties are somewhat orthogonal, but together they
@@
 \section{Discussion}
-\bcp{Make this a subsection?}
-\label{sec:disc}
 
@@
 \section{Lazy Stack Safety}
-\label{sec:lazy}
-
-    \definition{Observable Integrity}
-      Observable integrity weakens eager integrity by allowing a returned state
-      to differ from its call state at high integrity components, provided the
-      observable behavior of the induced trace from the return matches that of
-      an idealized state whose high integrity components do match the call
-      state.
-
-      Given a contour, a call state, and a returned state, we define a
-      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
-      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
-      the call state on those components that are high integrity in its
-      contour, and matches the return state on low integrity components:
-      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
+
+    \definition{Observable Trace Integrity}
+
+      Suppose a function \(f\) calls another function \(g\) with contour
+      \(\contour\). If \(g\) alters an address below its stack frame, where
+      \(\contour\) is \(\HIGHSEC\), the trace from the entry point to the
+      return will not enjoy Trace Integrity. But if \(f\) immediately
+      overwrites the overwritten location, the illegal write can be safely
+      ignored, as it will never affect the behavior of the program.
+      {\em Observable trace integrity} captures this intuition by guaranteeing
+      that the final state, instead of being identical to the initial one,
+      has the same behavior as it would if the writes had not occurred.
+
+      Let \(\contour\) be a contour and \(\MPT\) be a finite \(\MPOTRACE\)
+      whose initial machine state is \(\mach_e\). \(\MPT\) satisfies {\em trace
+      integrity} with respect to \(\contour\) if the trace induced by its final
+      state, \(\mpostate{\mach_r}{\pol}{\obs}\), is an observational prefix of
+      that induced by an idealized state \(\mach_r'\) that agrees with
+      \(\mach_r\) on all components marked \(\HIGHSEC\) in \(\contour\).
+      Formally we define a rollback function, \(\mathit{roll}\), that takes
+      two machine states and produces a third.
+      %
+      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
       \begin{cases}
@@
 
-      Intuitively, the result of a rollback is an idealized state that
-      reflects changes to accessible components, but ignores the callee's
-      interference with the caller's data. Then the observable behavior of the
-      resulting trace is the standard to which we expect the real trace to
-      conform.
-
-      A system enjoys observable integrity with regard to \(\callmap\) and
-      \(\codemap\) if for any initial state \(\mach_0\) and each adjacent pair
-      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
-      \(\pol_e\) in the trace induced by
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where
-      \(\callmap(\mach_c) = n\) and \(\codemap(\mach_e) = \funid\),
-      if \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
-      with the return state \(\mach_r\) at the end of \(\MPT\), then the
-      following holds.
-
-      Let \(\contour = \mathit{Cof}(\mach_c,\funid,n)\) describe the security
-      levels of the call and
-      \(\mach_r' = \mathit{roll}(\contour,\mach_e,\mach_r)\). The induced
-      traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\) and
-      \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
-      possible policy failure: \(\MPT \sim_P \machT'\).
+      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
+      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
+      agrees with \(\mach_e\) on every component the callee should not change,
+      but with \(\mach_r\) on those components that it is permitted to change.
+      So the observable behavior that follows is the idealized behavior to
+      which we compare the trace from \(\mach_r\), which must prefix it (to
+      account for premature halts due to policy violations.)
+      %
+      \[\begin{split}
+        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
+          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
+            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
+        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' \Rightarrow \\
+        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
+      \end{split}\]
+
+    \definition{Observable Stack Integrity}
+
+      A system enjoys {\em observable integrity} with regard to \(\callmap\)
+      and  \(\codemap\) if for any initial state \(\mach_0\) and each
+      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
+      contour \(\contour\) in the induced trace from
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
+      then \(\text{\sc ObsInt}_\contour ~ \MPT\).
 
     \definition{Observable Confidentiality}
+
       Observable confidentiality is concerned with whether high confidentiality
@@
 
-      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_c\) and
-      \(\mach_c'\), respectively, with \(\mach_c \approx_\contour \mach_c'\),
-      enjoy observable trace confidentiality with respect to \(\contour\) if
-      two conditions hold:
-
-      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim_O
-      \pi_o(\machT)\).
-
-      If \(\MPT\) is finite and ends with the state
-      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
-      \(\mach_r'\) be the final state of \(\machT\). Then the induced traces
-      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
-      \(\mathit{restore}(\contour,\mach_c,\mach_c',\mach_r,\mach_r')
-      \hookrightarrow \machT'\) are observationally prefixed,
-      \(\pi_o(\MPT') \lesssim_O \pi_o(\machT')\).
-
-      \sna{Structured like this, we need to start with the call state, not
-           the entry. That's because we need to know what to restore to.
-           But we could also define restore in the stack confidentiality
-           section, maybe?}
-
-    \definition{Observable Stack Confidentiality}
+      A system enjoys observable confidentiality with respect to \(\callmap\)
+      and \(\codemap\) if for any initial state \(\mach_0\), each adjacent pair
+      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
+      \(\pol_e\) in the inducted trace from
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), and each variant
+      \(\mach_e' \approx_\contour \mach_e'\) where
+      \(\contour = \mathit{Cof}(\mach_c, \funid, n)\), one of two cases holds
+      on the induced traces from \(\mach_e\) and \(\mach_e'\).
+
+      If \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_1 | \ret{\mach_c}\)
+      with \(\MPT_1\) the trace within the callee ending with the machine state
+      \(\mach_r\), the first state to return to the caller, and policy state
+      \(\pol_r\), then the there is a return state \(\mach_r'\) reached by the
+      trace from the variant, \(\mach_c' \hookrightarrow \machT_1 |
+      \ret{\mach_c}\). Let \(\mach_r'' = \mathit{restore}(\contour,\mach_e,
+      \mach_e',\mach_r,\mach_r')\), and let the induced traces from the
+      returned states be \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT_2\)
+      and \(\mach_r'' \hookrightarrow \machT_2\). With the callee and caller
+      traces joined, the traces are observationally equivalent up to policy
+      failures: \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1 \cdot \machT_2\).
+
+      The case in which the induced trace does not return is the same as in the
+      eager property: the variant trace also does not return, and the induced
+      traces are observationally equivalent up to policy failure.
 
     \definition{Observable Stack Safety}
+
       A system enjoys {\em observable stack safety} with regard to a call map

### 2020-07-01 main.tex

--- 2020-07-01 main.tex
@@
 In this section we will describe stack safety as a formal property. We
-begin with our model of knowledge about the program structure\leo{I
-  hate this}, and then dive in to local state encapsulation and
+begin with introducing a running example that will help keep things
+concrete throughout this section, continue with our model of
+\leo{compiler-provided?} knowledge about the program structure, and
+then dive in to formal definitions local state encapsulation and
 well-bracketed control flow.
 
+\begin{figure}[b]
+  \begin{centering}
+    \begin{minipage}{0.5\textwidth}
+\begin{verbatim}
+f{
+  int x, y;
+  ...
+  y = g(0);
+  ...
+}
+
+g(x){
+  return (x + h(17));
+}
+
+h(x){
+  return (x + 1);
+}
+\end{verbatim}
+\end{minipage}
+\end{centering}
+\caption{High-level Program with Simple Nested Calls}
+\label{fig:running-program}
+\end{figure}
+
 \subsection{Running Example}
 
-\leo{TODO: Figure out presentation}
+Our running example consists of three functions $f$, $g$, and $h$ that
+perform a nested sequence of calls. A high level C-like view of the
+excerpt program can be found in Figure~\ref{fig:running-program}.
+Suppose that the starting state for our running example is right
+before $f$'s call to $g$. A rough example layout of the memory at that
+point can be found in the top of Figure~\ref{fig:running-trace-b}:
+\begin{itemize}
+\item The low part of the memory (0-1000) is dedicated to instructions.
+\item The stack starts at location 1000.
+\item The stack of $f$ starts at 1040.
+\item The location of $x$ is 1040 and stores $42$.
+\item The location of $y$ is 1044 and stores $0$.
+\end{itemize}
+Directly below the layout, we summarize the parts of the register file
+that are relevant to the example.
+
+The rest of Figure~\ref{fig:running-trace-b} detail the execution of the
+program, showing the machine state at call and return points, as well
+as describing the full effect of each instruction in detail.
+\leo{Unoptimized, etc.}
 

### 2020-07-01 main.tex
Example in trace confidentiality.

--- 2020-07-01 main.tex
@@
 directly (via a jump instruction) or indirectly (by corrupting stack
-data and returning). \leo{There needs to be some informal description
-  of the wbcf part here. Sean?}
+data and returning). Well-bracketed control flow establshes safety between
+functions: all code is owned by a function, and all jumps between functions
+are either calls that enter at a marked entry point or returns that enter
+just after their associated call in the standard nested call structure.
+
+%\leo{There needs to be some informal description
+%  of the wbcf part here. Sean?}
 
@@
 In this section we will describe stack safety as a formal property. We
-begin with introducing a running example that will help keep things
-concrete throughout this section, continue with our model of
-\leo{compiler-provided?} knowledge about the program structure, and
-then dive in to formal definitions local state encapsulation and
+begin with our model of knowledge about the program structure\leo{I
+  hate this}, and then dive in to local state encapsulation and
 well-bracketed control flow.
 
-\begin{figure}[b]
-  \begin{centering}
-    \begin{minipage}{0.5\textwidth}
-\begin{verbatim}
-f{
-  int x, y;
-  ...
-  y = g(0);
-  ...
-}
-
-g(x){
-  return (x + h(17));
-}
-
-h(x){
-  return (x + 1);
-}
-\end{verbatim}
-\end{minipage}
-\end{centering}
-\caption{High-level Program with Simple Nested Calls}
-\label{fig:running-program}
-\end{figure}
-
 \subsection{Running Example}
 
-Our running example consists of three functions $f$, $g$, and $h$ that
-perform a nested sequence of calls. A high level C-like view of the
-excerpt program can be found in Figure~\ref{fig:running-program}.
-Suppose that the starting state for our running example is right
-before $f$'s call to $g$. A rough example layout of the memory at that
-point can be found in the top of Figure~\ref{fig:running-trace-b}:
-\begin{itemize}
-\item The low part of the memory (0-1000) is dedicated to instructions.
-\item The stack starts at location 1000.
-\item The stack of $f$ starts at 1040.
-\item The location of $x$ is 1040 and stores $42$.
-\item The location of $y$ is 1044 and stores $0$.
-\end{itemize}
-Directly below the layout, we summarize the parts of the register file
-that are relevant to the example.
-
-The rest of Figure~\ref{fig:running-trace-b} detail the execution of the
-program, showing the machine state at call and return points, as well
-as describing the full effect of each instruction in detail.
-\leo{Unoptimized, etc.}
+\leo{TODO: Figure out presentation}
 
@@
 \newcommand*{\jalr}{\text{jalr}}
-\newcommand*{\rsp}{\mathit{sp}}
-\newcommand*{\rra}{\mathit{ra}}
-
-\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}
-
-\begin{figure}
-
-%Starting Memory Layout
+%\begin{lstlisting}
+%START:
+%  addi r2 r2 12 ; allocate 3 words in the stack
+%  sw r2 42 -8; store 42 in a private location
+%  sw r2 0 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 F ; call F and store return address
+% 
+%F:
+%  addi r2 r2 4 ; allocate for ra 
+%  sw r2 r1 -4 ; store the return address to sp-4
+%  addi r2 r2 8 ; allocate for arg + ret
+%  sw r2 17 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 G ; call G and store return address
+% 
+%G:  
+%  addi r2 r2 4 ; allocate for ra 
+%  sw r2 r1 -4 ; store the return address to sp-4
+%  
+%\end{lstlisting}
+
+Starting State:
+
+Instruction memory in locations 0 - ..
+PC at 100.
+Stack memory starting at 1000.
+Assume we're in a function. 
+Stack Pointer at 1040.
+1040: 42.
+
 \begin{center}
@@
 \memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{8}{gray}[{\makebox[0pt]{Uninitialized}}]
+\memory{4}{gray}[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
-\MemoryLabel{-29.6em}{0.6em}{42}
-\MemoryLabel{-27.4em}{0.6em}{0}
+\MemoryLabel{-18.7em}{0.6em}{42}
+\MemoryLabel{-16.4em}{0.6em}{0}
 \\
 \end{center}
-% Starting Register File
+
+Program:
+
+\begin{verbatim}
+top{
+  return f(0);
+}
+
+f(x){
+  return (x + g(17));
+}
+
+g(x){
+  return (x + 1);
+}
+\end{verbatim}
+
+\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
+\newcommand*{\clabel}[1]{#1: & & &\\}
+\newcommand*{\rsp}{sp}
+\newcommand*{\rra}{ra}
 \[
-\begin{array}{cccccc}
-  PC: 100 & r_0 : 0 & \rra : ? & \rsp : 1048 & r_4 : ? & r_5 : ? 
+\begin{array}{llll}
+  \clabel{TOP}  
+  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
+  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
+  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
+  & & & \\
+  & & & \\
+  \clabel{F}
+  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
+  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
+  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
+  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
+  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
+  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
+  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
+  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
+  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
+  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
+  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
+  & & & \\
+  & & & \\  
+  \clabel{G}
+  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
+  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
+  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
+  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
+  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
+  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
+  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
+  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
+
+%  addi r2 r2 8 ; allocate for arg + ret
+%  sw r2 17 -4; store 0 as an argument
+%  sw r2 0 0; zero out the return location
+%  jal r1 G ; call G and store return address
+
 \end{array}
 \]
-\vspace*{0.2em}
-\[
-  \begin{array}{r|c|c|c|c}
-    & PC & \mathit{Instruction} & \mathit{Effect} &  \\
-    \hline
-    \tracerow{1}{100}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1056}{Allocate call frame}
-    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ -8}{[1048] \leftarrow 0}{Store 0 as argument}
-    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ -4}{[1052] \leftarrow 0}{Zero out result}
-    \tracerow{4}{112}{\jal ~ \rra ~ 140}{PC \leftarrow 140; ~ \rra \leftarrow 116}{Make the call}
-  \end{array}
-  \]
-~ \\
-% At Call to G
-% Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{41em}{1.8em}{$\downarrow$}
-\MemoryLabel{40.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
-\memory{6}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
-\\
-\end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 140 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : ? & r_5 : ? 
-\end{array}
-\]
-% G - Body - 1
-\vspace*{0.2em}
-\[
-  \begin{array}{r|c|c|c|c}
-    \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
-    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ -4}{[1056] \leftarrow 116}{Store return address}
-    \tracerow{7}{148}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1068}{Allocate for call frame}
-    \tracerow{8}{152}{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow 17}{Store 17 to $r_4$}
-    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ -8}{[1060] \leftarrow 17}{Store 17 as the argument}
-    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ -4}{[1064] \leftarrow 0 }{Zero out return}
-    \tracerow{11}{164}{\jal ~ \rra ~ 200}{PC \leftarrow 200; ~ \rra \leftarrow 168}{Make the call to $h$}
-  \end{array}
-  \]
-  ~ \\
-% At Call to H
-% Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{49em}{1.8em}{$\downarrow$}
-\MemoryLabel{48.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
-\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
-\memory{3}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
-\MemoryLabel{-20.6em}{0.6em}{116}
-\MemoryLabel{-18.4em}{0.6em}{17}
-\MemoryLabel{-16.2em}{0.6em}{0}
-\\
-\end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 200 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 17 & r_5 : ? 
-\end{array}
-\]
-% H - Body
-\vspace*{0.2em}
-\[
-\begin{array}{r|c|c|c|c}
-  \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
-  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ -4}{[1068] \leftarrow 168}{Store return address}
-  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ -12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
-  \tracerow{15}{212}{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{Increment $r_4$ by 1}\end{array}
-\]
-% At Middle of H
-% Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{52em}{1.8em}{$\downarrow$}
-\MemoryLabel{51.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
-\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
-\memory{1}{cyan}%
-\memory{2}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
-\MemoryLabel{-20.6em}{0.6em}{116}
-\MemoryLabel{-18.4em}{0.6em}{17}
-\MemoryLabel{-16.2em}{0.6em}{0}
-\MemoryLabel{-14.0em}{0.6em}{168}
-\\
-\end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 212 & r_0 : 0 & \rra : 168 & \rsp : 1072 & r_4 : 18 & r_5 : ? 
-\end{array}
-\]
-\vspace*{0.2em}
-\[
-\begin{array}{r|c|c|c|c}
-  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ -8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
-  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 168}{Load return address}
-  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}
-  \tracerow{19}{228}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 168}{Return to $g$}
-  \end{array}
-\]
-\label{fig:running-trace-a}
-\end{figure}
-
-\begin{figure}
-% After Return from H
-% Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{49em}{1.8em}{$\downarrow$}
-\MemoryLabel{48.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
-\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
-\memory{3}{gray}
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
-\MemoryLabel{-20.6em}{0.6em}{116}
-\MemoryLabel{-18.4em}{0.6em}{17}
-\MemoryLabel{-16.2em}{0.6em}{18}
-\\
-\end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 168 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 18 & r_5 : ? 
-\end{array}
-\]
-% H - Body
-\vspace*{0.2em}
-\[
-\begin{array}{r|c|c|c|c}
-  \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
-  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ -20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
-  \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
-  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ -8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
-  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1060}{Deallocate local state}
-  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 116}{Load return address}
-  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
-  \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
-\end{array}
-\]
-% At Call to G
-% Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{41em}{1.8em}{$\downarrow$}
-\MemoryLabel{40.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
-\memory{6}{gray}
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{18}
-\\
-\end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 116 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : 18 & r_5 : 0
-\end{array}
-\]
-\vspace*{0.2em}
-\[
-\begin{array}{r|c|c|c|c}
-  \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
-  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ -12}{[1048] \leftarrow 18}{Set $y$ to the return value}
-  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
-\end{array}
-\]
-\vspace*{0.2em}  
-% Ending Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
-\MemoryLabel{35em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{8}{gray}
-~$\cdots$
-\MemoryLabel{-29.6em}{0.6em}{42}
-\MemoryLabel{-27.4em}{0.6em}{0}
-\\
-\end{center}
-% Starting Register File
-\[
-\begin{array}{cccccc}
-  PC: 128 & r_0 : 0 & \rra : 116 & \rsp : 1048 & r_4 : 18 & r_5 : 0
-\end{array}
-\]
-\caption{Detailed Execution Trace}
-\label{fig:running-trace-b}
-\end{figure}
-
-%\end{minipage}
-%\begin{minipage}{0.7\textwidth}
-%\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
-%\newcommand*{\clabel}[1]{#1: & & &\\}
-%\[
-%\begin{array}{llll}
-%  \clabel{TOP}  
-%  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
-%  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
-%  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
-%  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
-%  & & & \\
-%  & & & \\
-%  \clabel{F}
-%  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
-%  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
-%  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
-%  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
-%  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
-%  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
-%  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
-%  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
-%  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
-%  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
-%  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
-%  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
-%  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
-%  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
-%  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
-%  & & & \\
-%  & & & \\  
-%  \clabel{G}
-%  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
-%  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
-%  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
-%  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
-%  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
-%  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
-%  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
-%  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
-%\end{array}
-%\]
-%\end{minipage}
-%\label{fig:running-program}
-%\end{figure}
 
@@
 \[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]
-%
-In our running example, instructions 100-124 correspond to $f$,
-instructions 140-196 correspond to $g$, and instructions 200-228
-correspond to $h$.
+
+\leo{Visible from assembly}
 
 A call map is a partial function from addresses to a natural number,
-the number of arguments and return values of the call.
+the number of arguments of the call.
   %
 \[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]
-%
-In our running example, the callmap would contain (at least) two
-entries: the word 112 corresponding to the first $\jal$
-instruction to $g$, which would map to $2$ as the call has one
-argument; and the word 164 also mapping to $2$ corresponding
-to the call to $h$.
+
+\leo{The two JAL instructions map to 2 (arg + return)}
 
@@
 \[\entmap \in \ENTMAPS \subseteq \WORDS\]
-%
-In our running example, the return map identifies the $\jalr$
-instructions 196 and 224, while the entry map identifies
-instructions 140 and 200.
+
+\leo{retmap is the jal + 1}
+\leo{entrymap is the ones after the labels}
 
@@
 
-    \definition{Observable Confidentiality}
-
-      Observable confidentiality is concerned with whether high confidentiality
-      data influences observable both behavior during and after a call. As with
-      eager confidentiality, we will compare the behavior of the trace induced
-      by a state to another induced by its variant. But if the original returns,
-      before continuing, we must reset the variant to conform with the original
-      state.
+    \definition{Observable Trace Confidentiality}
+
+      Trace confidentiality is a property of two traces whose initial machine
+      states are variations, with two sub-properties. First, the behavioral
+      condition -- that their observable behaviors prior to the end of the
+      trace are equivalent -- reflects the intuition that secret values should
+      not influence visible behavior. Second, we have the end condition that
+      their final states (if any) are equal up to the original variation. The
+      end condition over-approximates the extension of the behavioral condition
+      to future execution.
+
+      We introduce {\em observable trace confidentiality}, which extends the
+      behavioral condition to apply to execution from the final state on.
+      This requires restoring varied components to their original values,
+      since after return they will once again be accessible legitimately.
+
+\begin{figure}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \MemoryLabel{10em}{1.5em}{40}
+      \MemoryLabel{15em}{1.5em}{42}
+      \memory{2}{gray}%
+      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \MemoryLabel{10em}{1.5em}{40}
+      \MemoryLabel{15em}{1.5em}{42}
+      \memory{2}{gray}%
+      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage} \\
+  ~\\
+  ~\\
+  \begin{verbatim}
+    // read HC data and write based on it
+    lw r1 r0 40
+    bnz r1 2
+    sw 0 r0 42
+    jmp r2 // return
+    sw 1 r0 42
+    jmp r2 // return
+  \end{verbatim}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \memory{2}{gray}%
+      \memory{3}{red}
+      \memory{1}{black}
+      \memory{2}{green}
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \memory{2}{gray}%
+      \memory{3}{yellow}%
+      \memory{1}{white}
+      \memory{2}{green}
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage} \\
+
+  \caption{Failing Trace Confidentiality}
+  \label{fig:lazyex}
+\end{figure}
+
+      In Figure \ref{fig:lazyex}, the final states of the original and variant
+      differ, because a written value (in black or white) depended on the value
+      of a high confidentiality location. So the resulting trace fails Trace
+      Confidentiality. For observable trace confidentiality, we ask whether the
+      differing values at address 42 cause any observable distinctions in
+      future behavior. To do so, we must undo the variation of formerly high
+      confidentiality locations, so that those in yellow match those in red,
+      while preserving the updated values where data may have been leaked.
 

### 2020-07-02 main.tex

--- 2020-07-01 main.tex
@@
 the need to protect the call structure itself: a callee should not be
-allowed to jump to the middle of an arbitrary block of code, either
-directly (via a jump instruction) or indirectly (by corrupting the return
-address on the stack and then returning). \leo{There needs to be some
-  informal description of the wbcf part here. Sean?}
+allowed to jump to the middle of an arbitrary block of code either
+directly (via a jump instruction) or indirectly (by corrupting stack
+data and returning). Well-bracketed control flow establshes safety between
+functions: all code is owned by a function, and all jumps between functions
+are either calls that enter at a marked entry point or returns that enter
+just after their associated call in the standard nested call structure.
+
+%\leo{There needs to be some informal description
+%  of the wbcf part here. Sean?}
 
@@
 Note that if the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. Conversely\bcp{why conversely?}, if the predicate never
-holds, \(\PUT\) is the identity function.
+predicate must hold on its final element. If the predicate never holds,
+\(\PUT\) is the identity function.
 
@@
 
-\judgmenttwo{\(\mach \hookrightarrow \machT\)}
-            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT)
-              = \machT'\)}
-            {\(\mach \hookrightarrow \machT' | f\)}
+\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
+            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT')
+              = \machT\)}
+            {\(\mach \hookrightarrow \machT | f\)}
 \noindent
@@
 
-\leo{TODO: We really need a better layout than centered paragraphs.}
+%\leo{TODO: We really need a better layout than centered paragraphs.}
+
+\begin{minipage}{.3\textwidth}
   \judgment{}{\(\obsT \lesssim \obsT\)}
-
+\end{minipage}
+\begin{minipage}{.3\textwidth}
 \judgment{}{\(\tau \lesssim \obsT\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
 \judgment{}{\(w \lesssim w\obsT\)}
-
+\end{minipage}
+
+\begin{minipage}{.3\textwidth}
 \judgment{\(\obsT_1 \lesssim \obsT_2\)}
          {\(\tau \obsT_1 \lesssim \obsT_2\)}
-
+\end{minipage}
+\begin{minipage}{.3\textwidth}
 \judgment{\(\obsT_1 \lesssim \obsT_2\)}
          {\(\obsT_1 \lesssim \tau \obsT_2\)}
-
+\end{minipage}
+\begin{minipage}{.3\textwidth}
 \judgment{\(\obsT_1 \lesssim \obsT_2\)}
          {\(w\obsT_1 \lesssim w\obsT_2\)}
+\end{minipage}
 
@@
 
-We then define equivalence\bcp{Can we please call it something else?  An
-  ``equivalence'' should always be transitive.} of observation traces as traces prefixing each other:
+We then define similarity of observation traces as traces prefixing each other:
+
+\bcp{Can we please call it something else?  An
+  ``equivalence'' should always be transitive.}
+\sna{Similarity seems less loaded?}
 \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
@@
 and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
-if it started from a variant initial machine state.
+if it started from a variant initial machine state. An example in
 
@@
 
+Figure \ref{fig:lazyex1} shows the initial states \(\mach_e\), left,
+and \(\mach_e'\), right. Low confidentiality data in green is shared
+between them, while high confidentiality data in red and yellow may differ.
+After executing the code shown, the final states below have written different
+values to address 42, represented by the black and white components.
+It is allowed for address 40 to differ between the states due to the initial
+variation, but address 42 changed differently between the traces, representing
+a leak of data. Meanwhile \ref{fig:lazyex2}, while high confidentiality data
+is read, it is overwritten before return, and the write at step 3 always writes
+the same value. Yellow and red locations differ between variations only because
+of the initial variation, and the rest are identical. So this example passes.
+
+\sna{TODO: it would be nice for this to correspond at least roughly to
+     the layout in the running example.}
+
+\begin{figure}
+  % H - Body
+  \[
+  \begin{array}{r|c|c|c|c}
+    \tracerow{1}{100}{\lw ~ r_0 ~ 40}{}{Load value at 40 to $r_0$}
+    \tracerow{2}{104}{\sw ~ 42 ~ r_0}{}{Store that value to 42}
+    \tracerow{3}{108}{\jalr ~ \rra ~ \rra ~ 0}{}{Return}
+  \end{array}
+  \]
+
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \MemoryLabel{10em}{1.5em}{40}
+      \MemoryLabel{15em}{1.5em}{42}
+      \memory{2}{gray}%
+      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \MemoryLabel{10em}{1.5em}{40}
+      \MemoryLabel{15em}{1.5em}{42}
+      \memory{2}{gray}%
+      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage} \\
+  ~\\
+  ~\\
+ \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \memory{2}{gray}%
+      \memory{3}{red}
+      \memory{1}{black}
+      \memory{2}{green}
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \memory{2}{gray}%
+      \memory{3}{yellow}%
+      \memory{1}{white}
+      \memory{2}{green}
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage} \\
+
+  \caption{Failing Trace Confidentiality}
+  \label{fig:lazyex1}
+\end{figure}
 \leo{Discuss variations in a similar manner}
 
+\begin{figure}
+  % H - Body
+  \[
+  \begin{array}{r|c|c|c|c}
+    \tracerow{1}{100}{\lw ~ r_0 ~ 40}{}{Load value at 40 to $r_0$}
+    \tracerow{2}{104}{\sw ~ 42 ~ 0}{}{Store a constant value to 42}
+    \tracerow{3}{108}{\lw ~ r_0 ~ 45}{}{Overwrite $r_0$ with accessible value}
+    \tracerow{4}{112}{\jalr ~ \rra ~ \rra ~ 0}{}{Return}
+  \end{array}
+  \]
+
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \MemoryLabel{10em}{1.5em}{40}
+      \MemoryLabel{15em}{1.5em}{42}
+      \memory{2}{gray}%
+      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \MemoryLabel{10em}{1.5em}{40}
+      \MemoryLabel{15em}{1.5em}{42}
+      \memory{2}{gray}%
+      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage} \\
+  ~\\
+  ~\\
+ \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \memory{2}{gray}%
+      \memory{3}{red}
+      \memory{1}{black}
+      \memory{2}{green}
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage}
+  \begin{minipage}{.49\textwidth}
+    \begin{center}
+      \memory{2}{gray}%
+      \memory{3}{yellow}%
+      \memory{1}{black}
+      \memory{2}{green}
+      \memory{2}{gray}%
+      \\
+    \end{center}
+  \end{minipage} \\
+
+  \caption{Passing Confidentiality Despite Failing Integrity}
+  \label{fig:lazyex2}
+\end{figure}
 To lift trace confidentiality to a system property, we follow the same
 pattern as in stack integrity:
-
-\sna{Version for stepping from a state}
-\definition{Confidential Execution}
-  Stack confidentiality, below, is a property of a system that, from a call to
-  a return, each trace obeys a local confidentiality property,
-  {\it confidential execution}.
-
-  We will say that a state \(\mpstate{\mach}{\pol}\) satisfies confidential
-  execution with regard to a contour \(\contour\) and an ending condition
-  \(R\) that is a predicate on machine states if, for all variant states
-  \(\mach' \approx_\contour \mach\), the induced traces from \(\mach\) and
-  \(\mach'\) behave the same across three cases. Let
-  \(\mpstate{\mach}{\pol} \hookrightarrow \MPT | R\) and
-  \(\mach' \hookrightarrow \machT | R\).
-
-  If \(\MPT\) ends with a state \(\mach_r\) and \(R\ \mach_r\), then \(\machT\)
-  ends with a state \(\mach_r'\) such that \(R\ \mach_r'\). \(\MPT\) and
-  \(\machT\) are observationally equivalent, and for each component
-  \(\component\) such that \(\contour(\component) = \HIGHSEC\),
-  \(\mach_r(\component) = \mach_r'(\component)\).
-
-  If \(\MPT\) is infinite, then \(\machT\) is infinite, and they are
-  observationally equivalent.
-
-  If \(\MPT\) ends before \(R\) ever holds, meaning that it halted due to a
-  policy violation, then it is a strict observational prefix of \(\machT\),
-  which must be infinite.
 
@@
 
-      Intuitively, the result of a rollback is an idealized state that
-      reflects changes to accessible components, but ignores the callee's
-      interference with the caller's data. Then the observable behavior of the
-      resulting trace is the standard to which we expect the real trace to
-      conform.
-
-      A system enjoys observable integrity with regard to \(\callmap\) and
-      \(\codemap\) if for any initial state \(\mach_0\) and each adjacent pair
-      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
-      \(\pol_e\) in the trace induced by
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where
-      \(\callmap(\mach_c) = n\) and \(\codemap(\mach_e) = \funid\),
-      if \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
-      with the return state \(\mach_r\) at the end of \(\MPT\), then the
-      following holds.
-
-      Let \(\contour = \mathit{Cof}(\mach_c,\funid,n)\) describe the security
-      levels of the call and
-      \(\mach_r' = \mathit{roll}(\contour,\mach_e,\mach_r)\). The induced
-      traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\) and
-      \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
-      possible policy failure: \(\MPT \sim_P \machT'\).
-
-    \definition{Observable Confidentiality}
-      Observable confidentiality is concerned with whether high confidentiality
-      data influences observable both behavior during and after a call. As with
-      eager confidentiality, we will compare the behavior of the trace induced
-      by a state to another induced by its variant. But if the original returns,
-      before continuing, we must reset the variant to conform with the original
-      state.
+      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
+      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
+      agrees with \(\mach_e\) on every component the callee should not change,
+      but with \(\mach_r\) on those components that it is permitted to change.
+      So the observable behavior that follows is the idealized behavior to
+      which we compare the trace from \(\mach_r\), which must prefix it (to
+      account for premature halts due to policy violations.)
+      %
+      \[\begin{split}
+        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
+          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
+            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
+        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' \Rightarrow \\
+        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
+      \end{split}\]
+
+    \definition{Observable Stack Integrity}
+
+      A system enjoys {\em observable integrity} with regard to \(\callmap\)
+      and  \(\codemap\) if for any initial state \(\mach_0\) and each
+      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
+      contour \(\contour\) in the induced trace from
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
+      then \(\text{\sc ObsInt}_\contour ~ \MPT\).
+
+    \definition{Observable Trace Confidentiality}
+
+      Trace confidentiality is a property of two traces whose initial machine
+      states are variations, with two sub-properties. First, the behavioral
+      condition -- that their observable behaviors prior to the end of the
+      trace are equivalent -- reflects the intuition that secret values should
+      not influence visible behavior. Second, we have the end condition that
+      their final states (if any) are equal up to the original variation. The
+      end condition over-approximates the extension of the behavioral condition
+      to future execution.
+
+      We introduce {\em observable trace confidentiality}, which extends the
+      behavioral condition to apply to execution from the final state on.
+      This requires restoring varied components to their original values,
+      since after return they will once again be accessible legitimately.
+
+      In Figure \ref{fig:lazyex1}, the final states of the original and variant
+      differ, because a written value (in black or white) depended on the value
+      of a high confidentiality location. So the resulting trace fails trace
+      confidentiality. For observable trace confidentiality, we ask whether the
+      differing values at address 42 cause any observable distinctions in
+      future behavior. To do so, we must undo the variation of formerly high
+      confidentiality locations, so that those in yellow match those in red,
+      while preserving the updated values where data may have been leaked.
 
@@
       unless they were changed during execution:
-      \[\mathit{restore}(\contour,\mach_c,\mach_c',\mach_r,\mach_r')(\component) =
+      \[\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')(\component) =
         \begin{cases}
           \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_) \text{ and } \\
-                  & \mach_c(\component) = \mach_r(\component) \text{ and }
-                    \mach_c'(\component) = \mach_r'(\component) \\
+                  & \mach_e(\component) = \mach_r(\component) \text{ and }
+                    \mach_e'(\component) = \mach_r'(\component) \\
           \mach_r'(\component) & \text{else} \\
@@
 
-      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_c\) and
-      \(\mach_c'\), respectively, with \(\mach_c \approx_\contour \mach_c'\),
+      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_e\) and
+      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\),
       enjoy observable trace confidentiality with respect to \(\contour\) if
@@
 
-      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim_O
+      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
       \pi_o(\machT)\).
@@
       \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
-      \(\mathit{restore}(\contour,\mach_c,\mach_c',\mach_r,\mach_r')
+      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
       \hookrightarrow \machT'\) are observationally prefixed,
-      \(\pi_o(\MPT') \lesssim_O \pi_o(\machT')\).
-
-      \sna{Structured like this, we need to start with the call state, not
-           the entry. That's because we need to know what to restore to.
-           But we could also define restore in the stack confidentiality
-           section, maybe?}
+      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).
+
+      We write this \(\text{\sc ObsConf}_\contour ~ \MPT\).
 
     \definition{Observable Stack Confidentiality}
+
+      From observable trace confidentiality we construct {\em observable stack
+      confidentiality}. A system enjoys {\em observable stack confidentiality}
+      with regard to \(\callmap\) and \(\codemap\) if for any initial state
+      \(\mach_0\) and each transition with entry state
+      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
+      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
+      then \(\text{\sc ObsConf}_\contour ~ \MPT\).
 

### 2020-07-02 main.tex
Comments on the new running example

@@
   \begin{centering}
-    \begin{minipage}{0.5\textwidth}
+    \begin{minipage}{.8\textwidth}
 \begin{verbatim}
-f{
-  int x, y;
-  ...
+f = {                 g(x) = {                      h(x) = {
+   int x, y;                    return (x + h(17));                  return (x + 1);
+  ...                 }                             }
   y = g(0);
   ...
-}
-
-g(x){
-  return (x + h(17));
-}
-
-h(x){
-  return (x + 1);
 }
@@
 \end{centering}
-\caption{High-level Program with Simple Nested Calls}
+\caption{High-level Program with Simple Nested Calls \bcp{Rename parameters
+    of g and h to z and w}}
 \label{fig:running-program}
@@
 \subsection{Running Example}
+
+\bcp{
+Comments on the example:
+\begin{itemize}
+\item We can simplify a bit by renaming $f$ to $\mathit{main}$ and
+relocating its stack to location 1000.  And maybe move its instructions to
+start at 0.  (We can comment that on a real RISC-V this is where interrupt
+handlers go, but POPL readers will not care too much.)  We should streamline
+the example as much as possible.
+\item Do we need the ``...''s in $f$?
+%
+\item We should also emphasize, again, that there is no heap (if there were,
+we might put it above the stack or between the instructions and the stack).
+%
+\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
+capitalization of register names (sp vs SP), etc. Maybe typeset instructions
+in small caps?  Maybe typeset everything that is part of a concrete machine
+in small caps??
+%
+\item The colors are helpful in making the big page of instruction traces easier
+on the eye; maybe they would be even more helpful if
+corresponding colors appeared elsewhere in the example (on the instructions
+themselves, e.g.).  But we need to be careful to keep it readable -- maybe
+they should be background colors, like you've done, rather than text colors.
+\item The first column of numbers is redundant with the second column (PC).
+Suggest deleting the first.
+\item Asking readers to wade through all this is asking quite a bit!  We
+need to make sure we are going to get sufficient value out of the example
+later (I think we can, but we should really convince ourselves) to make it
+worthwhile.
+\item Use ``='' instead of ``:'' in the ``summaries'' below the pictures.
+And put a bit more space between the components.
+\item
+\end{itemize}
+}
 
@@
 Suppose that the starting state for our running example is right
-before $f$'s call to $g$. A rough example layout of the memory at that
+before $f$'s call to $g$. A sketch of the memory layout at that
 point can be found in the top of Figure~\ref{fig:running-trace-b}:
@@
 
-The rest of Figure~\ref{fig:running-trace-b} detail the execution of the
+The rest of Figure~\ref{fig:running-trace-b} details the execution of the
 program, showing the machine state at call and return points, as well
 as describing the full effect of each instruction in detail.
-\leo{Unoptimized, etc.}
+\leo{Unoptimized, etc.}\bcp{??}
 
@@
 ~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
+\memory{2}{green}[{\makebox[0pt]{$F_{\textit{priv}}$}}]%
 \memory{8}{gray}[{\makebox[0pt]{Uninitialized}}]
@@
 \[
-  \begin{array}{r|c|c|c|c}
+  \begin{array}{l|l|l|l|l}
     & PC & \mathit{Instruction} & \mathit{Effect} &  \\
@@
 \[
-  \begin{array}{r|c|c|c|c}
+  \begin{array}{l|l|l|l|l}
     \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
@@
 \[
-\begin{array}{r|c|c|c|c}
+\begin{array}{l|l|l|l|l}
   \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
@@
 \[
-\begin{array}{r|c|c|c|c}
+\begin{array}{l|l|l|l|l}
   \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ -8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
@@
 \[
-\begin{array}{r|c|c|c|c}
+\begin{array}{l|l|l|l|l}
   \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
@@
 \[
-\begin{array}{r|c|c|c|c}
+\begin{array}{l|l|l|l|l}
   \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
@@
   \[
-  \begin{array}{r|c|c|c|c}
+  \begin{array}{l|l|l|l|l}
     \tracerow{1}{100}{\lw ~ r_0 ~ 40}{}{Load value at 40 to $r_0$}

### 2020-07-02 main.tex
More comments

@@
 And put a bit more space between the components.
-\item
+\item Create a new column that just literally pronounces each instruction in
+English, so people don't have to read RISC-V opcodes.
+\item As discussed, we should move this to before section 2, we should
+extend it with a bit more at the end so we can talk about the lazy policy
+with the same example, we should add
+in some bad variants, and we should consider discussing the attacker model
+here (we need to discuss it someplace in any case!)
 \end{itemize}
@@
 \newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}
+\newcommand{\negate}{\textrm{-}}
 
@@
     \tracerow{1}{100}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1056}{Allocate call frame}
-    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ -8}{[1048] \leftarrow 0}{Store 0 as argument}
-    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ -4}{[1052] \leftarrow 0}{Zero out result}
+    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ \negate 8}{[1048] \leftarrow 0}{Store 0 as argument}
+    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1052] \leftarrow 0}{Zero out result}
     \tracerow{4}{112}{\jal ~ \rra ~ 140}{PC \leftarrow 140; ~ \rra \leftarrow 116}{Make the call}
@@
     \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
-    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ -4}{[1056] \leftarrow 116}{Store return address}
+    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1056] \leftarrow 116}{Store return address}
     \tracerow{7}{148}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1068}{Allocate for call frame}
     \tracerow{8}{152}{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow 17}{Store 17 to $r_4$}
-    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ -8}{[1060] \leftarrow 17}{Store 17 as the argument}
-    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ -4}{[1064] \leftarrow 0 }{Zero out return}
+    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1060] \leftarrow 17}{Store 17 as the argument}
+    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1064] \leftarrow 0 }{Zero out return}
     \tracerow{11}{164}{\jal ~ \rra ~ 200}{PC \leftarrow 200; ~ \rra \leftarrow 168}{Make the call to $h$}
@@
   \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
-  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ -4}{[1068] \leftarrow 168}{Store return address}
-  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ -12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
+  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1068] \leftarrow 168}{Store return address}
+  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ \negate 12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
   \tracerow{15}{212}{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{Increment $r_4$ by 1}\end{array}
@@
 \begin{array}{l|l|l|l|l}
-  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ -8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
-  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 168}{Load return address}
-  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}
+  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
+  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ \negate 4}{\rra \leftarrow 168}{Load return address}
+  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}
   \tracerow{19}{228}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 168}{Return to $g$}
@@
   \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
-  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ -20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
+  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ \negate 20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
   \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
-  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ -8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
-  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1060}{Deallocate local state}
-  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 116}{Load return address}
-  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
+  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
+  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ \negate 8}{\rsp \leftarrow 1060}{Deallocate local state}
+  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ \negate 4}{\rra \leftarrow 116}{Load return address}
+  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
   \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
@@
   \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
-  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ -12}{[1048] \leftarrow 18}{Set $y$ to the return value}
-  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
+  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ \negate 12}{[1048] \leftarrow 18}{Set $y$ to the return value}
+  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ \negate 8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
 \end{array}
@@
 
-Formally a code map is a partial function from addresses to a function
+Formally a code map\bcp{I'd love to change ``cdm'' to ``km''.  Should we
+  also call it a ``kode map'' in English?  Maybe that's a bit too far.} is a
+partial function from addresses to a function
 identifier drawn from the set \(\FUNIDS\).
@@
     (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
-                          \text{ and } \component \leq \mach(\SP) - n \\
+                          \text{ and } \component \leq \mach(\SP) \negate  n \\
     (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS

### 2020-07-02 main.tex
Notes

@@
 with the same example, we should add
-in some bad variants, and we should consider discussing the attacker model
+in some bad variants (doing it here replaces the ``Examples'' from the
+Discussion section), and we should consider discussing the attacker model 
 here (we need to discuss it someplace in any case!)
@@
 
+\bcp{Belongs with the running example:}
   \bcp{Illustrate the definition by using it to discuss a range of ``stack
@@
   \paragraph{What about tail recursion and other non-stack-like control flow}
+
+\bcp{And this and the next belong in Future Work...}
 

### 2020-07-02 main.tex
Updated running example

@@
 
-\begin{figure}
-  \begin{minipage}{\textwidth}
-    \begin{center}
-\MemoryLabel{10em}{1.5em}{40}
-\MemoryLabel{16em}{1.5em}{60}
-\MemoryLabel{22em}{1.5em}{80}
-$\cdots$
-\memory{4}{gray}%
-\memory{5}{green}[{\makebox[0pt]{Caller}}]%
-\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
-\memory{8}{red}[{\makebox[0pt]{Callee}}]%
-\memory{4}{gray}%
-~$\cdots$\\
-  \end{center}
-  \end{minipage}\\
-~\\
-~\\
-  \begin{minipage}{0.45\textwidth}
-    \begin{center}
-\begin{verbatim}
-jal ...
-lw r1 r0 40
-\end{verbatim}
-    \end{center}
-  \end{minipage}~
-  \begin{minipage}{0.45\textwidth}
-    \begin{center}
-\begin{verbatim}
-jal ...
-sw r0 r0 40
-\end{verbatim}
-    \end{center}
-  \end{minipage}
-  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
-      the boxes; the program fragments need comments explaining what each
-      line does.}}
-  \label{fig:stackunsafety}
-\end{figure}
-
-\leo{I hate the writing here, but I wanted to make a start.}
+%\begin{figure}
+%  \begin{minipage}{\textwidth}
+%    \begin{center}
+%\MemoryLabel{10em}{1.5em}{40}
+%\MemoryLabel{16em}{1.5em}{60}
+%\MemoryLabel{22em}{1.5em}{80}
+%$\cdots$
+%\memory{4}{gray}%
+%\memory{5}{green}[{\makebox[0pt]{Caller}}]%
+%\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
+%\memory{8}{red}[{\makebox[0pt]{Callee}}]%
+%\memory{4}{gray}%
+%~$\cdots$\\
+%  \end{center}
+%  \end{minipage}\\
+%~\\
+%~\\
+%  \begin{minipage}{0.45\textwidth}
+%    \begin{center}
+%\begin{verbatim}
+%jal ...
+%lw r1 r0 40
+%\end{verbatim}
+%    \end{center}
+%  \end{minipage}~
+%  \begin{minipage}{0.45\textwidth}
+%    \begin{center}
+%\begin{verbatim}
+%jal ...
+%sw r0 r0 40
+%\end{verbatim}
+%    \end{center}
+%  \end{minipage}
+%  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
+%      the boxes; the program fragments need comments explaining what each
+%      line does.}}
+%  \label{fig:stackunsafety}
+%\end{figure}
+% 
+%\leo{I hate the writing here, but I wanted to make a start.}
 
@@
 %
-We begin with some technical preliminaries in
-Section~\ref{sec:prelim}.
-
-\section{Background}
-\label{sec:prelim}
-
-To set the stage, we first
-describe our machine model. Our definition of stack safety is
-generic and this section discusses an abstract interface to our
-machine model and some kind of static or dynamic, software- or
-hardware-enforced security monitor.  For
-concreteness, our running examples will use a
-conventional machine architecture and a standard (RISC-V) instruction set;
-in~\Cref{sec:micropolicies} we will
-extend this model it with a dynamic policy enforcement mechanism based
-on \bcp{NickPaper,AndMaybeSomeOthers}.
-
-\paragraph*{Values, States, and Observations}
-
-The basic building blocks of the machine are {\em values} and {\em
-  addresses}. Both are drawn from some set of {\em words}
-$\WORDS$, ranged over by $\word$.
-%
-Our machine states are composed of {\em components} $k$, addressed either by
-words (memory addresses) or by by register names ($r$) drawn from some set
-$\REGS$.
-%
-    \[\component \in \COMPONENTS = \WORDS + \REGS \]
-%
-The register names are assumed to include two special purpose registers: the
-program counter {\PCname} and the stack pointer \SP.
-
-A machine state is a map from components to values:
-%
-\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]
-
-The step function of the
-machine takes a machine state and produces a pair of a
-machine state and an {\em observation}, which may be either
-silent (written $\tau$) or a word of data that the machine communicates to
-the outside world (e.g., by writing to a DMA address):\leo{Needs more
-  exposition}\bcp{Like that?}
-%
-\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
-\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
-%
-
-\paragraph*{Running Example: RISC-V}
-
-Mapping a conventional architecture to this abstract interface is
-straightforward. To build up our running example, we are going to use
-a standard 64-bit RISC machine, just like the one we are using in our
-experiments later on \leo{Mention FORVIS?}. In this setting, $\WORDS$
-is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
-r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$.
-% \leo{Mention We use the standard encoding of instructions to words?}
-% BCP: Nah.
-
-The first interesting choice is that of observations, as there are
-multiple reasonable options. At one extreme, we could choose to define
-observations as the value written at any component. However, that
-would pose problems later on when we treat some components as ``secret''
-when they correspond to a caller's private state. \leo{On the other extreme,
-we could pick a particular component and treat it as the ``output'' of
-our system. }  \bcp{But in this paragraph we are describing the particular
-choise that we make for our running example, no?  But we certainly want to
-discuss this fully someplace!}
-
-\paragraph*{Policies}
-
-\bcp{More explanation needed, plus examples.}A policy is a general model of an enforcement mechanism, consisting of
-a set of policy states \(\pol \in \POLS\), and a policy step function
-\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
-\rightharpoonup \POLS\).
-
-We write $\MPS$ for the set of all pairs of machine states and policy
-states.
-%
-\[\MPS = \MACHS \times \POLS\]
-%
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1) \stepstopol \pol_2\)}
-            {\(\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}\)}
-
-\leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted
-  step?}\bcp{Yes.}
-
-\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
-in more detail that the intention is to capture any kind of dynamic
-enforcement mechanism and give as many names and citations as we can.}
-
-\paragraph{Traces}
-
-A trace over a type $A$ is a nonempty, finite or infinite
-sequence of $A$s. We will mostly be concerned with sequences of
-machines states, usually\bcp{not always?} denoted $\machT$, and with sequences of
-triples of (machine states, policy states, and observations), usually
-denoted $\MPT$.
-\rb{The way the definitions are set up, one may expect MPO, but\ldots}
-To continue\bcp{??}, we will require four generic operations
-on traces: a ``trace-of'' operation to construct the
-trace induced by a step relation and a machine state
-\rb{not an op on traces per se?}\bcp{+1};
-a ``split'', to break a
-trace into two components based on a predicate; a ``join'', the
-reverse of split; and a ``prefix'', as a shorthand to only keep the
-first component of a split.  \bcp{Not sure it's worth enumerating them
-  rather than just getting down to the actual definitions below.}
-
-The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
-relates a machine state with the trace of machine states paired with
-observations produced by repeated application of step:  \bcp{above, we
-  showed operators applied to metavariables, not just the bare operator
-  symbol.}
-%
-\[\hookrightarrow : \machT \rightarrow \MOTRACE\]%
-%
-\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
-            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
-%
-\noindent
-Since the step function is total, if \(\mach \hookrightarrow \machT\),
-then \(\machT\) must be infinite.
-
-\leo{TODO: Add an example of a trace once I figure out the notations}
-\bcp{Let's also try to clean up the spacing a little -- many of these
-  sequences of rules can be typeset on fewer lines}
-
-Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
-state pair to the trace of triples of machine states, policy states,
-and observations induced by by \(\stepstoobs{}_P\): \bcp{Observe that this
-  time the trace can be either finite or infinite.}
-%
-\[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \MPOTRACE \]
-%
-\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
-         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
-%
-\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
-              {\(\pol_0 \rightharpoonup \pol_1\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
-We will also use \(\pi_m\) to project the trace of machine states
-out of a trace, \(\pi_p\) to project the trace of policy states, and
-\(\pi_o\) to project the observations.  \bcp{I'm getting a little lost in
-  all these.  E.g., do these traces include the final state of a finite
-  execution?  If so, what observation is it paired with?}
-
-% SNA: fixed
-%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
-%  seems better}
-The {\it prefUpTo} operation \(\PUT(\Trace,f)\) \bcp{Why does it have two
-  different abbreviated names?  And wouldn't ``While'' be more mnemonic?}
-takes a trace \(T\)
-and a predicate on elements \(f\) and gives the prefix of \(T\)
-ending with the first element on which \(f\) holds, coinductively:
-
-\judgment{\(f\ \traceelem\)}
-         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
-%
-\judgment{}
-         {\(\PUT(f,\traceelem) = \traceelem\)}
-%
-\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
-            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
-%
-Note that if the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate never holds,
-\(\PUT\) is the identity function.
-
-% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
-%$\mathit{split}$: it takes two traces and links them by replacing the
-%last element of the first (if it has one) with the first element of
-%the second
-
-%\judgment{} {\(t \cdot T = T\)}
-%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
-%\leo{Lemma/theorem: split/join = id?}
-%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
-
-We will frequently take a prefix of the trace from an initial state up
-through the first state where some condition holds on the machine state.
-This can easily be implemented using \(\PUT\), and we provide special notation
-for convenience.
-%
-Let \(f\) be a predicate on machine states. Then we define \(\mach
-\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
-running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I think.}
-
-\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
-            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT')
-              = \machT\)}
-            {\(\mach \hookrightarrow \machT | f\)}
-\noindent
-And similarly for traces with policies:
-
-\judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
-            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT'\)}
-         {\(\mpstatename \hookrightarrow \MPT' | f\)}
-
-%    \COQ{this operation steps until it reaches a state on which \(f\) holds,
-%    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
-%    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
-%    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
-%      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
-%    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}
-
-\paragraph*{Observational Equivalence}
-
-We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-\leo{Benjamin suggested ``as far as an attacker is concerned'' in an
-  e-mail, but we haven't said anything about attackers yet.}\bcp{We don't
-  have to talk about attackers, just external observers.}, if the
-sequence of non-silent observations of $\obsT_1$ is a prefix of those
-of $\obsT_2$; that is we operate up to deletion of \(\tau\)
-observations, coinductively:
-
-%\leo{TODO: We really need a better layout than centered paragraphs.}
-
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(\obsT \lesssim \obsT\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-\judgment{}{\(\tau \lesssim \obsT\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-\judgment{}{\(w \lesssim w\obsT\)}
-\end{minipage}
-
-\begin{minipage}{.3\textwidth}
-\judgment{\(\obsT_1 \lesssim \obsT_2\)}
-         {\(\tau \obsT_1 \lesssim \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-\judgment{\(\obsT_1 \lesssim \obsT_2\)}
-         {\(\obsT_1 \lesssim \tau \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-\judgment{\(\obsT_1 \lesssim \obsT_2\)}
-         {\(w\obsT_1 \lesssim w\obsT_2\)}
-\end{minipage}
-
-% Fixed by Andrew:
-%\COQ{This is the same as the coq version minus a redundant case.}
-
-We then define similarity of observation traces as traces prefixing each other:
-
-\bcp{Can we please call it something else?  An
-  ``equivalence'' should always be transitive.}
-\sna{Similarity seems less loaded?}
-\[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
-
-\leo{Rephrase transition. I hate it:} There are two interesting things
-to note.  First, observation trace equivalence could alternatively be
-defined as the prefix relation but dropping the only assymetric rules
-\leo{2 + 3. TODO: name rules?}
-
-\leo{Also rephrase this.}
-Second, an infinite silent trace is a prefix of (and equivalent to)
-any other trace. While this might seem surprising at first, it makes
-sense in a timing-insensitive context: an external observer looking
-at two machine runs cannot (computably) distinguish between a machine that
-steps forever and a machine that steps for a long time before producing
-some output.
-
-%Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
-%We then define {\em policy-sensitive observational equivalence}: a
-%trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
-%\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
-%and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
-%policy fault) and its observations are a prefix of those of M.
-%
-%% Fixed by Sean
-%%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
-%\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
-%            {\(\MPT \sim_P \machT\)}
-%
-%\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
-%            {\(\MPT \sim_P \machT\)}
-
-\section{Stack Safety, Formally}
-\label{sec:lse-and-wbcf}
-
-In this section we will describe stack safety as a formal property. We
-begin with introducing a running example that will help keep things
-concrete throughout this section, continue with our model of
-\leo{compiler-provided?} knowledge about the program structure, and
-then dive in to formal definitions local state encapsulation and
-well-bracketed control flow.
+We begin with a running example that we're going to keep coming back
+to throughout the paper (Section~\ref{sec:running-example}) and some
+technical preliminaries (Section~\ref{sec:prelim}).
+
+\section{Running Example}
+\label{sec:running-example}
+
+To help keep things concrete throughout the more technical sections
+that follow, we are first going to introduce a running example. We are
+going to use it to demonstrate various potentially problematic
+behaviours in this section, and then show how our properties and
+enforcement mechanisms preclude them later on in the paper.
+
+The program appears in Figure~\ref{fig:running-program}. It consists
+of a {\tt main} function that performs a call to some function {\tt
+  f}, which in turns makes a nested call to some function {\tt g}
+before returning, all while performing some simple arithmetic
+in the process.
 
@@
   \begin{centering}
-    \begin{minipage}{.8\textwidth}
+    \begin{minipage}{.3\textwidth}
 \begin{verbatim}
-f = {                 g(x) = {                      h(x) = {
-   int x, y;                    return (x + h(17));                  return (x + 1);
-  ...                 }                             }
+main = {
+  int x -42, y = 0;
   y = g(0);
-  ...
+  print (x + y);
 }
@@
     \end{minipage}
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+g(z) = {
+  int w = h(17);
+  print z;
+  return (z + w);
+}
+\end{verbatim}
+    \end{minipage}
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+h(v) = {
+  return (v + 1);
+}
+
+\end{verbatim}
+    \end{minipage}
 \end{centering}
-\caption{High-level Program with Simple Nested Calls \bcp{Rename parameters
-    of g and h to z and w}}
+\caption{High-level Program with Simple Nested Calls}
 \label{fig:running-program}
@@
 
-\subsection{Running Example}
-
-\bcp{
-Comments on the example:
+To show a trace of the execution of this program, we are going to need
+a low-level view of both the code and the machine it runs on. While
+our concrete implementation (Section~\leo{cite}) operates on a full
+conventional RISC-V architecture, we will make a few simplifying
+assumptions here to ease the exposition.
+
+First of all, we are going to ignore the heap. We are mainly concerned
+with stack interactions so that will be the main focus. \leo{We're
+  going to come back to that at the end. Will we?} We are then going
+to abstract away low-level details of memory layout and present a
+machine that is word-addressed, with instructions starting from
+location 0, a dedicated output at location 99, and lay the stack from
+location 100 upwards. \leo{Explain more? How to connect to realistic
+  one? Observations? DMA? Interrupt handlers?}
+%
+A sketch of the memory layout at that point can be found in the top of
+Figure~\ref{fig:running-trace-b}: instructions for {\tt main} begin at
+0, the dedicated output is empty, while the (uninitialized) stack
+begins at 100. Directly below the layout, we summarize the parts of
+the register file that are relevant to the example.
+%
+The rest of Figure~\ref{fig:running-trace-b} details the execution of the
+program, showing the machine state at call and return points, as well
+as describing the full effect of each instruction in detail.
+\leo{Expand upon this?}
+
+\bcp{ Comments on the example:
 \begin{itemize}
@@
 the example as much as possible.
-\item Do we need the ``...''s in $f$?
 %
@@
 
-Our running example consists of three functions $f$, $g$, and $h$ that
-perform a nested sequence of calls. A high level C-like view of the
-excerpt program can be found in Figure~\ref{fig:running-program}.
-Suppose that the starting state for our running example is right
-before $f$'s call to $g$. A sketch of the memory layout at that
-point can be found in the top of Figure~\ref{fig:running-trace-b}:
-\begin{itemize}
-\item The low part of the memory (0-1000) is dedicated to instructions.
-\item The stack starts at location 1000.
-\item The stack of $f$ starts at 1040.
-\item The location of $x$ is 1040 and stores $42$.
-\item The location of $y$ is 1044 and stores $0$.
-\end{itemize}
-Directly below the layout, we summarize the parts of the register file
-that are relevant to the example.
-
-The rest of Figure~\ref{fig:running-trace-b} details the execution of the
-program, showing the machine state at call and return points, as well
-as describing the full effect of each instruction in detail.
-\leo{Unoptimized, etc.}\bcp{??}
-
-\newcommand*{\add}{\text{add}}
-\newcommand*{\addi}{\text{addi}}
-\newcommand*{\sw}{\text{sw}}
-\newcommand*{\lw}{\text{lw}}
-\newcommand*{\jal}{\text{jal}}
-\newcommand*{\jalr}{\text{jalr}}
-\newcommand*{\rsp}{\mathit{sp}}
-\newcommand*{\rra}{\mathit{ra}}
-
-\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}
+\newcommand*{\add}{\textsc{add}}
+\newcommand*{\addi}{\textsc{addi}}
+\newcommand*{\sw}{\textsc{sw}}
+\newcommand*{\lw}{\textsc{lw}}
+\newcommand*{\jal}{\textsc{jal}}
+\newcommand*{\jalr}{\textsc{jalr}}
+\newcommand*{\rsp}{\textsc{sp}}
+\newcommand*{\rra}{\textsc{ra}}
+\newcommand*{\rout}{\textsc{out}}
+
+\newcounter{pcctr}
+\setcounter{pcctr}{0}
+\newcommand*{\row}[4]{
+  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
+}
+\newcommand*{\tracerow}[5]{
+  #1 & #2 & #3 & #4 & \text{#5}\\
+}
+\newcommand*{\summary}[6]{
+\[
+\begin{array}{cccccc}
+  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
+\end{array}
+\]
+}    
 \newcommand{\negate}{\textrm{-}}
@@
 \begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
-\MemoryLabel{35em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{25em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{10}{gray}[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{\textit{priv}}$}}]%
-\memory{8}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-29.6em}{0.6em}{42}
-\MemoryLabel{-27.4em}{0.6em}{0}
 \\
@@
 % Starting Register File
-\[
-\begin{array}{cccccc}
-  PC: 100 & r_0 : 0 & \rra : ? & \rsp : 1048 & r_4 : ? & r_5 : ?
-\end{array}
-\]
+\summary{0}{0}{?}{100}{?}{?}
 \vspace*{0.2em}
 \[
-  \begin{array}{l|l|l|l|l}
-    & PC & \mathit{Instruction} & \mathit{Effect} &  \\
+  \begin{array}{c|c|c|c|c}
+    PC & \mathit{Instruction} & \mathit{Description} & \mathit{Effect} & \\
     \hline
-    \tracerow{1}{100}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1056}{Allocate call frame}
-    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ \negate 8}{[1048] \leftarrow 0}{Store 0 as argument}
-    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1052] \leftarrow 0}{Zero out result}
-    \tracerow{4}{112}{\jal ~ \rra ~ 140}{PC \leftarrow 140; ~ \rra \leftarrow 116}{Make the call}
+    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 102}
+        {Allocate private {\tt main} data}
+    \row{\addi ~ r_4 ~ r_0 ~ 42}{r_4 \leftarrow r_0 + 42}{r_4 \leftarrow 42}
+        {Put 42 in $r_4$}
+    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp-2] \leftarrow r_4}{[100] \leftarrow 42}
+        {Set {\tt x} to 42}
+    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[101] \leftarrow 0}
+        {Set {\tt y} to 0}
+    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 104}
+        {Allocate {\tt g}'s call frame}
+    \row{\sw ~ \rsp ~ r_0 ~ \negate 2}{[\rsp-2] \leftarrow r_0}{[102] \leftarrow 0}
+        {Set argument to 0}
+    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[103] \leftarrow 0}
+        {Zero out result}
+    \row{\jal ~ \rra ~ 20}{??}{\PCname \leftarrow 20; ~ \rra \leftarrow \thepcctr}
+        {Make the call}
   \end{array}
@@
   ~ \\
+  ~\\
 % At Call to G
@@
 \begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{41em}{1.8em}{$\downarrow$}
-\MemoryLabel{40.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{43.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{6}{gray}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
-\memory{6}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
 \\
 \end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 140 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : ? & r_5 : ?
-\end{array}
-\]
+% Starting Register File
+\summary{20}{0}{\thepcctr}{104}{42}{?}
+%\vspace*{0.2em}
+\setcounter{pcctr}{20}%
 % G - Body - 1
@@
 \[
-  \begin{array}{l|l|l|l|l}
-    \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
-    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1056] \leftarrow 116}{Store return address}
-    \tracerow{7}{148}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1068}{Allocate for call frame}
-    \tracerow{8}{152}{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow 17}{Store 17 to $r_4$}
-    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1060] \leftarrow 17}{Store 17 as the argument}
-    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1064] \leftarrow 0 }{Zero out return}
-    \tracerow{11}{164}{\jal ~ \rra ~ 200}{PC \leftarrow 200; ~ \rra \leftarrow 168}{Make the call to $h$}
+  \begin{array}{c|c|c|c|c}
+    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 106}
+        {Allocate local {\tt g} state}
+    \row{\sw ~ \rsp ~ \rra ~ \negate 2}{[\rsp - 2] \leftarrow \rra}{[104] \leftarrow 8}
+        {Store return address}
+    % Zero out z?
+    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 108}
+        {Allocate for call frame}
+    \row{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow r_0 + 17}{r_4 \leftarrow 17}
+        {Store 17 to $r_4$}
+    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[106] \leftarrow 17}
+        {Store 17 as the argument {\tt v}}
+    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp - 1] \leftarrow r_0}{[107] \leftarrow 0}
+        {Zero out result}
+    \row{\jal ~ \rra ~ 40}{??}{PC \leftarrow 40; ~ \rra \leftarrow \thepcctr}
+        {Make the call to {\tt h}}
   \end{array}
   \]
+  ~ \\
   ~ \\
@@
 \begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{49em}{1.8em}{$\downarrow$}
-\MemoryLabel{48.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\\
+\end{center}
+% Starting Register File
+\summary{40}{0}{\thepcctr}{108}{17}{?}
+\setcounter{pcctr}{40}
+\vspace*{0.2em}
+% H - Body
+\[
+\begin{array}{c|c|c|c|c}
+  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
+      {Allocate for return address}
+  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
+      {Store return address}
+  \row{\lw ~ r_4 ~ \rsp ~ \negate 3}{r_4 \leftarrow [\rsp - 3]}{r_4 \leftarrow 17}
+      {Load the argument {\tt v} to $r_4$}
+  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 18}
+      {Increment $r_4$ by 1}\end{array}
+\]
+~\\
+~\\
+% At Middle of H
+% Memory Layout
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{65.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{1}{red}%[{\makebox[0pt]{{\texttt h} ra}}]
+\memory{1}{gray}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
-\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
-\memory{3}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
-\MemoryLabel{-20.6em}{0.6em}{116}
-\MemoryLabel{-18.4em}{0.6em}{17}
-\MemoryLabel{-16.2em}{0.6em}{0}
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\MemoryLabel{-14.5em}{0.75em}{27}
 \\
@@
 % Register File
+\summary{40}{0}{\thepcctr}{109}{18}{?}
+\vspace*{0.2em}
+\label{fig:running-trace-a}
+\end{figure}
+
+\begin{figure}
 \[
-\begin{array}{cccccc}
-  PC: 200 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 17 & r_5 : ?
+\begin{array}{c|c|c|c|c}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
+      {Store $r_4$ as the return value}
+  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
+      {Load return address}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
+      {Restore the stack pointer of {\tt g}}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
+      {Return to {\tt g}}
   \end{array}
 \]
+
+% After Return from H
+% Memory Layout
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{18}
+\\
+\end{center}
+% Starting Register File
+\summary{27}{0}{\thepcctr}{108}{18}{?}
+\vspace*{0.2em}
+
+\setcounter{pcctr}{27}
 % H - Body
@@
 \[
-\begin{array}{l|l|l|l|l}
-  \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
-  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1068] \leftarrow 168}{Store return address}
-  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ \negate 12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
-  \tracerow{15}{212}{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{Increment $r_4$ by 1}\end{array}
+\begin{array}{c|c|c|c|c}
+  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
+      {Load return value to $r_4$}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
+      {Set {\tt w} to 18}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
+      {Deallocate {\tt h}'s call frame}
+  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
+      {Load argument {\tt z} $r_5$}
+  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
+      {Print {\tt z}}      
+  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
+      {Add $r_4$ and $r_5$}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
+      {Store $r_4$ as ther return of {\tt g}}
+  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
+      {Load return address}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
+      {Deallocate {\tt g}'s local state}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 8}{Return to {\tt main}}
+\end{array}
 \]
-% At Middle of H
+% Return to main
 % Memory Layout
 \begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{52em}{1.8em}{$\downarrow$}
-\MemoryLabel{51.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{43.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{6}{gray}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
-\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
-\memory{1}{cyan}%
-\memory{2}{gray}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
-\MemoryLabel{-20.6em}{0.6em}{116}
-\MemoryLabel{-18.4em}{0.6em}{17}
-\MemoryLabel{-16.2em}{0.6em}{0}
-\MemoryLabel{-14.0em}{0.6em}{168}
+\MemoryLabel{-51em}{0.75em}{0}
+\MemoryLabel{-47em}{0.75em}{42}
+\MemoryLabel{-43em}{0.75em}{0}
+\MemoryLabel{-39em}{0.75em}{0}
+\MemoryLabel{-35em}{0.75em}{18}
 \\
@@
 % Register File
+\summary{8}{0}{\thepcctr}{104}{18}{0}
+%\vspace*{0.2em}
+\setcounter{pcctr}{8}
 \[
-\begin{array}{cccccc}
-  PC: 212 & r_0 : 0 & \rra : 168 & \rsp : 1072 & r_4 : 18 & r_5 : ?
+\begin{array}{c|c|c|c|c}
+  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
+      {Load return value to $r_4$}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
+      {Set {\tt y} to the result}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
+      {Deallocate {\tt g}'s call frame}
+  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
+      {Load {\tt x} to $r_5$}
+  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
+      {Add {\tt x} and {\tt y}}
+  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
+      {Print {\tt x + y}}
 \end{array}
@@
 \vspace*{0.2em}
-\[
-\begin{array}{l|l|l|l|l}
-  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
-  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ \negate 4}{\rra \leftarrow 168}{Load return address}
-  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}
-  \tracerow{19}{228}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 168}{Return to $g$}
-  \end{array}
-\]
-\label{fig:running-trace-a}
-\end{figure}
-
-\begin{figure}
-% After Return from H
+% Return to main
 % Memory Layout
 \begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{49em}{1.8em}{$\downarrow$}
-\MemoryLabel{48.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{31.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{8}{gray}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
-\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
-\memory{3}{gray}
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{0}
-\MemoryLabel{-20.6em}{0.6em}{116}
-\MemoryLabel{-18.4em}{0.6em}{17}
-\MemoryLabel{-16.2em}{0.6em}{18}
-\\
-\end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 168 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 18 & r_5 : ?
-\end{array}
-\]
-% H - Body
-\vspace*{0.2em}
-\[
-\begin{array}{l|l|l|l|l}
-  \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
-  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ \negate 20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
-  \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
-  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
-  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ \negate 8}{\rsp \leftarrow 1060}{Deallocate local state}
-  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ \negate 4}{\rra \leftarrow 116}{Load return address}
-  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
-  \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
-\end{array}
-\]
-% At Call to G
-% Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{41em}{1.8em}{$\downarrow$}
-\MemoryLabel{40.8em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
-\memory{6}{gray}
-~$\cdots$
-\MemoryLabel{-29.4em}{0.6em}{42}
-\MemoryLabel{-27.2em}{0.6em}{0}
-\MemoryLabel{-25.0em}{0.6em}{0}
-\MemoryLabel{-22.8em}{0.6em}{18}
-\\
-\end{center}
-% Register File
-\[
-\begin{array}{cccccc}
-  PC: 116 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : 18 & r_5 : 0
-\end{array}
-\]
-\vspace*{0.2em}
-\[
-\begin{array}{l|l|l|l|l}
-  \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
-  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ \negate 12}{[1048] \leftarrow 18}{Set $y$ to the return value}
-  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ \negate 8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
-\end{array}
-\]
-\vspace*{0.2em}
-% Ending Memory Layout
-\begin{center}
-\MemoryLabel{1.7em}{1.8em}{0}
-\MemoryLabel{21.2em}{1.8em}{1000}
-\MemoryLabel{29.7em}{1.8em}{1040}
-\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
-\MemoryLabel{35em}{3em}{SP}
-\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
-~$\cdots$
-\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
-~$\cdots$
-\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
-\memory{8}{gray}
-~$\cdots$
-\MemoryLabel{-29.6em}{0.6em}{42}
-\MemoryLabel{-27.4em}{0.6em}{0}
+\MemoryLabel{-51em}{0.75em}{60}
+\MemoryLabel{-47em}{0.75em}{60}
+\MemoryLabel{-43em}{0.75em}{42}
 \\
@@
 % Starting Register File
-\[
-\begin{array}{cccccc}
-  PC: 128 & r_0 : 0 & \rra : 116 & \rsp : 1048 & r_4 : 18 & r_5 : 0
-\end{array}
-\]
+\summary{\thepcctr}{0}{37}{102}{60}{42}
 \caption{Detailed Execution Trace}
@@
 
+\section{Background}
+\label{sec:prelim}
+
+To set the stage, we first
+describe our machine model. Our definition of stack safety is
+generic and this section discusses an abstract interface to our
+machine model and some kind of static or dynamic, software- or
+hardware-enforced security monitor.  For
+concreteness, our running examples will use a
+conventional machine architecture and a standard (RISC-V) instruction set;
+in~\Cref{sec:micropolicies} we will
+extend this model it with a dynamic policy enforcement mechanism based
+on \bcp{NickPaper,AndMaybeSomeOthers}.
+
+\paragraph*{Values, States, and Observations}
+
+The basic building blocks of the machine are {\em values} and {\em
+  addresses}. Both are drawn from some set of {\em words}
+$\WORDS$, ranged over by $\word$.
+%
+Our machine states are composed of {\em components} $k$, addressed either by
+words (memory addresses) or by by register names ($r$) drawn from some set
+$\REGS$.
+%
+    \[\component \in \COMPONENTS = \WORDS + \REGS \]
+%
+The register names are assumed to include two special purpose registers: the
+program counter {\PCname} and the stack pointer \SP.
+
+A machine state is a map from components to values:
+%
+\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]
+
+The step function of the
+machine takes a machine state and produces a pair of a
+machine state and an {\em observation}, which may be either
+silent (written $\tau$) or a word of data that the machine communicates to
+the outside world (e.g., by writing to a DMA address):\leo{Needs more
+  exposition}\bcp{Like that?}
+%
+\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
+\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
+%
+
+\paragraph*{Running Example: RISC-V}
+
+Mapping a conventional architecture to this abstract interface is
+straightforward. To build up our running example, we are going to use
+a standard 64-bit RISC machine, just like the one we are using in our
+experiments later on \leo{Mention FORVIS?}. In this setting, $\WORDS$
+is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
+r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$.
+% \leo{Mention We use the standard encoding of instructions to words?}
+% BCP: Nah.
+
+The first interesting choice is that of observations, as there are
+multiple reasonable options. At one extreme, we could choose to define
+observations as the value written at any component. However, that
+would pose problems later on when we treat some components as ``secret''
+when they correspond to a caller's private state. \leo{On the other extreme,
+we could pick a particular component and treat it as the ``output'' of
+our system. }  \bcp{But in this paragraph we are describing the particular
+choise that we make for our running example, no?  But we certainly want to
+discuss this fully someplace!}
+
+\paragraph*{Policies}
+
+\bcp{More explanation needed, plus examples.}A policy is a general model of an enforcement mechanism, consisting of
+a set of policy states \(\pol \in \POLS\), and a policy step function
+\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
+\rightharpoonup \POLS\).
+
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
+%
+\[\MPS = \MACHS \times \POLS\]
+%
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1) \stepstopol \pol_2\)}
+            {\(\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}\)}
+
+\leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted
+  step?}\bcp{Yes.}
+
+\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
+in more detail that the intention is to capture any kind of dynamic
+enforcement mechanism and give as many names and citations as we can.}
+
+\paragraph{Traces}
+
+A trace over a type $A$ is a nonempty, finite or infinite
+sequence of $A$s. We will mostly be concerned with sequences of
+machines states, usually\bcp{not always?} denoted $\machT$, and with sequences of
+triples of (machine states, policy states, and observations), usually
+denoted $\MPT$.
+\rb{The way the definitions are set up, one may expect MPO, but\ldots}
+To continue\bcp{??}, we will require four generic operations
+on traces: a ``trace-of'' operation to construct the
+trace induced by a step relation and a machine state
+\rb{not an op on traces per se?}\bcp{+1};
+a ``split'', to break a
+trace into two components based on a predicate; a ``join'', the
+reverse of split; and a ``prefix'', as a shorthand to only keep the
+first component of a split.  \bcp{Not sure it's worth enumerating them
+  rather than just getting down to the actual definitions below.}
+
+The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
+relates a machine state with the trace of machine states paired with
+observations produced by repeated application of step:  \bcp{above, we
+  showed operators applied to metavariables, not just the bare operator
+  symbol.}
+%
+\[\hookrightarrow : \machT \rightarrow \MOTRACE\]%
+%
+\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
+            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
+%
+\noindent
+Since the step function is total, if \(\mach \hookrightarrow \machT\),
+then \(\machT\) must be infinite.
+
+\leo{TODO: Add an example of a trace once I figure out the notations}
+\bcp{Let's also try to clean up the spacing a little -- many of these
+  sequences of rules can be typeset on fewer lines}
+
+Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
+state pair to the trace of triples of machine states, policy states,
+and observations induced by by \(\stepstoobs{}_P\): \bcp{Observe that this
+  time the trace can be either finite or infinite.}
+%
+\[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \MPOTRACE \]
+%
+\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
+         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
+%
+\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
+              {\(\pol_0 \rightharpoonup \pol_1\)}
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
+We will also use \(\pi_m\) to project the trace of machine states
+out of a trace, \(\pi_p\) to project the trace of policy states, and
+\(\pi_o\) to project the observations.  \bcp{I'm getting a little lost in
+  all these.  E.g., do these traces include the final state of a finite
+  execution?  If so, what observation is it paired with?}
+
+% SNA: fixed
+%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
+%  seems better}
+The {\it prefUpTo} operation \(\PUT(\Trace,f)\) \bcp{Why does it have two
+  different abbreviated names?  And wouldn't ``While'' be more mnemonic?}
+takes a trace \(T\)
+and a predicate on elements \(f\) and gives the prefix of \(T\)
+ending with the first element on which \(f\) holds, coinductively:
+
+\judgment{\(f\ \traceelem\)}
+         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
+%
+\judgment{}
+         {\(\PUT(f,\traceelem) = \traceelem\)}
+%
+\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
+            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
+%
+Note that if the resulting trace is a strict prefix of the parameter, then the
+predicate must hold on its final element. If the predicate never holds,
+\(\PUT\) is the identity function.
+
+% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
+%$\mathit{split}$: it takes two traces and links them by replacing the
+%last element of the first (if it has one) with the first element of
+%the second
+
+%\judgment{} {\(t \cdot T = T\)}
+%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
+%\leo{Lemma/theorem: split/join = id?}
+%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
+
+We will frequently take a prefix of the trace from an initial state up
+through the first state where some condition holds on the machine state.
+This can easily be implemented using \(\PUT\), and we provide special notation
+for convenience.
+%
+Let \(f\) be a predicate on machine states. Then we define \(\mach
+\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
+running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I think.}
+
+\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
+            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT')
+              = \machT\)}
+            {\(\mach \hookrightarrow \machT | f\)}
+\noindent
+And similarly for traces with policies:
+
+\judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
+            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT'\)}
+         {\(\mpstatename \hookrightarrow \MPT' | f\)}
+
+%    \COQ{this operation steps until it reaches a state on which \(f\) holds,
+%    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
+%    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
+%    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
+%      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
+%    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}
+
+\paragraph*{Observational Equivalence}
+
+We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
+\leo{Benjamin suggested ``as far as an attacker is concerned'' in an
+  e-mail, but we haven't said anything about attackers yet.}\bcp{We don't
+  have to talk about attackers, just external observers.}, if the
+sequence of non-silent observations of $\obsT_1$ is a prefix of those
+of $\obsT_2$; that is we operate up to deletion of \(\tau\)
+observations, coinductively:
+
+%\leo{TODO: We really need a better layout than centered paragraphs.}
+
+\begin{minipage}{.3\textwidth}
+  \judgment{}{\(\obsT \lesssim \obsT\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+\judgment{}{\(\tau \lesssim \obsT\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+\judgment{}{\(w \lesssim w\obsT\)}
+\end{minipage}
+
+\begin{minipage}{.3\textwidth}
+\judgment{\(\obsT_1 \lesssim \obsT_2\)}
+         {\(\tau \obsT_1 \lesssim \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+\judgment{\(\obsT_1 \lesssim \obsT_2\)}
+         {\(\obsT_1 \lesssim \tau \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+\judgment{\(\obsT_1 \lesssim \obsT_2\)}
+         {\(w\obsT_1 \lesssim w\obsT_2\)}
+\end{minipage}
+
+% Fixed by Andrew:
+%\COQ{This is the same as the coq version minus a redundant case.}
+
+We then define similarity of observation traces as traces prefixing each other:
+
+\bcp{Can we please call it something else?  An
+  ``equivalence'' should always be transitive.}
+\sna{Similarity seems less loaded?}
+\[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
+
+\leo{Rephrase transition. I hate it:} There are two interesting things
+to note.  First, observation trace equivalence could alternatively be
+defined as the prefix relation but dropping the only assymetric rules
+\leo{2 + 3. TODO: name rules?}
+
+\leo{Also rephrase this.}
+Second, an infinite silent trace is a prefix of (and equivalent to)
+any other trace. While this might seem surprising at first, it makes
+sense in a timing-insensitive context: an external observer looking
+at two machine runs cannot (computably) distinguish between a machine that
+steps forever and a machine that steps for a long time before producing
+some output.
+
+%Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
+%We then define {\em policy-sensitive observational equivalence}: a
+%trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
+%\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
+%and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
+%policy fault) and its observations are a prefix of those of M.
+%
+%% Fixed by Sean
+%%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
+%\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
+%            {\(\MPT \sim_P \machT\)}
+%
+%\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
+%            {\(\MPT \sim_P \machT\)}
+
+\section{Stack Safety, Formally}
+\label{sec:lse-and-wbcf}
+
+In this section we will describe stack safety as a formal property. We
+begin with introducing a running example that will help keep things
+concrete throughout this section, continue with our model of
+\leo{compiler-provided?} knowledge about the program structure, and
+then dive in to formal definitions local state encapsulation and
+well-bracketed control flow.
+
 \subsection{Program Structure}

### 2020-07-03 main.tex
Added better examples for integrity and confidentiality. Second conf example coming.

@@
 
+For an example see Figure \ref{fig:intex}, in which {\tt h} writes its return
+value to {\tt g}'s return location the stack rather than its own. That address
+is high integrity (red), and so writing to it causes the property to fail.
+
 \definition{Stack Integrity}
@@
 
-Figure \ref{fig:lazyex1} shows the initial states \(\mach_e\), left,
+Figure \ref{fig:confex2} shows the initial states \(\mach_e\), left,
 and \(\mach_e'\), right. Low confidentiality data in green is shared
@@
 variation, but address 42 changed differently between the traces, representing
-a leak of data. Meanwhile \ref{fig:lazyex2}, while high confidentiality data
-is read, it is overwritten before return, and the write at step 3 always writes
-the same value. Yellow and red locations differ between variations only because
-of the initial variation, and the rest are identical. So this example passes.
-
-\sna{TODO: it would be nice for this to correspond at least roughly to
-     the layout in the running example.}
+a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
+%is read, it is overwritten before return, and the write at step 3 always writes
+%the same value. Yellow and red locations differ between variations only because
+%of the initial variation, and the rest are identical. So this example passes.
 
 \begin{figure}
+% At Call to H
+% Memory Layout
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+  \memory{6}{red}[{\makebox[0pt]{\(\HIGHINT\)}}]%
+  \memory{2}{green}[{\makebox[0pt]{\(\LOWINT\)}}]%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\\
+\end{center}
+% Starting Register File
+\summary{40}{0}{\thepcctr}{108}{17}{?}
+\setcounter{pcctr}{40}
+\vspace*{0.2em}
 % H - Body
 \[
-  \begin{array}{l|l|l|l|l}
-    \tracerow{1}{100}{\lw ~ r_0 ~ 40}{}{Load value at 40 to $r_0$}
-    \tracerow{2}{104}{\sw ~ 42 ~ r_0}{}{Store that value to 42}
-    \tracerow{3}{108}{\jalr ~ \rra ~ \rra ~ 0}{}{Return}
+\begin{array}{c|c|c|c|c}
+  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
+      {Allocate for return address}
+  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
+      {Store return address}
+  \row{\lw ~ r_4 ~ \rsp ~ \negate 3}{r_4 \leftarrow [\rsp - 3]}{r_4 \leftarrow 17}
+      {Load the argument {\tt v} to $r_4$}
+  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 18}
+      {Increment $r_4$ by 1}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 5}{[\rsp - 5] \leftarrow r_4}{[103] \leftarrow 18}
+      {Store $r_4$ as {\tt g}'s return value}
+  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
+      {Load return address}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
+      {Restore the stack pointer of {\tt g}}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
+      {Return to {\tt g}}
   \end{array}
 \]
-
-  \begin{minipage}{.49\textwidth}
 \begin{center}
-      \MemoryLabel{10em}{1.5em}{40}
-      \MemoryLabel{15em}{1.5em}{42}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{44em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+  \memory{3}{red}%
+  \memory{1}{green}%
+  \memory{2}{red}%
   \memory{2}{gray}%
-      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
-      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
-      \memory{2}{gray}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{18}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\MemoryLabel{-16em}{0.75em}{108}
 \\
 \end{center}
-  \end{minipage}
-  \begin{minipage}{.49\textwidth}
+%
+  \caption{Failing Trace Integrity}
+  \label{fig:intex}
+\end{figure}
+
+\begin{figure}
+% At Call to H
+% Memory Layout
 \begin{center}
-      \MemoryLabel{10em}{1.5em}{40}
-      \MemoryLabel{15em}{1.5em}{42}
-      \memory{2}{gray}%
-      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+  \memory{6}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
   \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
-      \memory{2}{gray}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
 \\
 \end{center}
-  \end{minipage} \\
-  ~\\
-  ~\\
- \begin{minipage}{.49\textwidth}
+% Variant
 \begin{center}
-      \memory{2}{gray}%
-      \memory{3}{red}
-      \memory{1}{black}
-      \memory{2}{green}
-      \memory{2}{gray}%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}%
+  \memory{6}{yellow}[{\makebox[0pt]{Varied}}]%
+  \memory{2}{green}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{?}
+\MemoryLabel{-42.5em}{0.75em}{?}
+\MemoryLabel{-38.5em}{0.75em}{?}
+\MemoryLabel{-34.5em}{0.75em}{?}
+\MemoryLabel{-30.5em}{0.75em}{?}
+\MemoryLabel{-26.5em}{0.75em}{?}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
 \\
 \end{center}
-  \end{minipage}
-  \begin{minipage}{.49\textwidth}
+
+% Starting Register File
+\summary{40}{0}{\thepcctr}{108}{17}{?}
+\setcounter{pcctr}{40}
+\vspace*{0.2em}
+% H - Body
+\[
+\begin{array}{c|c|c|c|c}
+  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
+      {Allocate for return address}
+  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
+      {Store return address}
+  \row{\lw ~ r_4 ~ 100}{r_4 \leftarrow 100}{r_4 \leftarrow 42/?}
+      {Load {\tt main}'s secret {\tt x} to $r_4$}
+  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 43/(?+1)}
+      {Increment $r_4$ by 1}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 43/(?+1)}
+      {Store $r_4$ as the return value}
+  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
+      {Load return address}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
+      {Restore the stack pointer of {\tt g}}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
+      {Return to {\tt g}}
+  \end{array}
+\]
 \begin{center}
-      \memory{2}{gray}%
-      \memory{3}{yellow}%
-      \memory{1}{white}
-      \memory{2}{green}
-      \memory{2}{gray}%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{44em}{2em}{\SP}
+\memory{3}{gray}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}%
+  \memory{6}{red}%
+  \memory{1}{green}%
+  \memory{1}{pink}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-19em}{0.75em}{43}
+\MemoryLabel{-16em}{0.75em}{108}
 \\
 \end{center}
-  \end{minipage} \\
-
-  \caption{Failing Trace Confidentiality}
-  \label{fig:lazyex1}
-\end{figure}
-\leo{Discuss variations in a similar manner}
-
-\begin{figure}
-  % H - Body
-  \[
-  \begin{array}{l|l|l|l|l}
-    \tracerow{1}{100}{\lw ~ r_0 ~ 40}{}{Load value at 40 to $r_0$}
-    \tracerow{2}{104}{\sw ~ 42 ~ 0}{}{Store a constant value to 42}
-    \tracerow{3}{108}{\lw ~ r_0 ~ 45}{}{Overwrite $r_0$ with accessible value}
-    \tracerow{4}{112}{\jalr ~ \rra ~ \rra ~ 0}{}{Return}
-  \end{array}
-  \]
-
-  \begin{minipage}{.49\textwidth}
+%
 \begin{center}
-      \MemoryLabel{10em}{1.5em}{40}
-      \MemoryLabel{15em}{1.5em}{42}
-      \memory{2}{gray}%
-      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
-      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
-      \memory{2}{gray}%
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{44em}{2em}{\SP}
+\memory{3}{gray}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}%
+  \memory{6}{yellow}%
+  \memory{1}{green}%
+  \memory{1}{pink}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{?}
+\MemoryLabel{-42.5em}{0.75em}{?}
+\MemoryLabel{-38.5em}{0.75em}{?}
+\MemoryLabel{-34.5em}{0.75em}{?}
+\MemoryLabel{-30.5em}{0.75em}{?}
+\MemoryLabel{-26.5em}{0.75em}{?}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-19em}{0.75em}{?}
+\MemoryLabel{-16em}{0.75em}{108}
 \\
 \end{center}
-  \end{minipage}
-  \begin{minipage}{.49\textwidth}
-    \begin{center}
-      \MemoryLabel{10em}{1.5em}{40}
-      \MemoryLabel{15em}{1.5em}{42}
-      \memory{2}{gray}%
-      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
-      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
-      \memory{2}{gray}%
-      \\
-    \end{center}
-  \end{minipage} \\
-  ~\\
-  ~\\
- \begin{minipage}{.49\textwidth}
-    \begin{center}
-      \memory{2}{gray}%
-      \memory{3}{red}
-      \memory{1}{black}
-      \memory{2}{green}
-      \memory{2}{gray}%
-      \\
-    \end{center}
-  \end{minipage}
-  \begin{minipage}{.49\textwidth}
-    \begin{center}
-      \memory{2}{gray}%
-      \memory{3}{yellow}%
-      \memory{1}{black}
-      \memory{2}{green}
-      \memory{2}{gray}%
-      \\
-    \end{center}
-  \end{minipage} \\
-
-  \caption{Passing Confidentiality Despite Failing Integrity}
-  \label{fig:lazyex2}
+%
+  \caption{Failing Trace Confidentiality}
+  \label{fig:confex2}
 \end{figure}
+
+Figure \ref{confex2} gives an example of a trace that fails trace
+confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
+Since {\tt x}, at address 100, is high confidentiality (red) the variant
+may have a different value (yellow). So both \(r_4\) and address 107
+differ in the final state between the original trace and the variant one,
+failing the ending condition of trace confidentiality.
+
 To lift trace confidentiality to a system property, we follow the same

### 2020-07-03 main.tex
Rough draft of a section on "enforcement"

@@
 
+\section{Enforcement}
+\label{sec:enforcement}
+
+\paragraph*{Micropolicies}
+
+\paragraph*{A conservative policy}
+
+This conservative policy enforces our stack safety policy. Moreover, it enforces
+a provably stronger \emph{testing property}.
+
+The properties we give are indeed enforceable.
+
+Less conservative policies can be more efficient, but also harder to test.
+
+What is the strongest property enforced by the conservative policy?
+%
+\rb{Not too much focus on this}
+
+\rb{We have not proved that this policy guarantees our property, but we have
+tested it against a stronger property.}
+
+\paragraph*{Testing the conservative policy}
+
+\paragraph*{End-to-end stack safety property}
+
+\paragraph*{Eager vs lazy testing}
+
+\rb{This part should probably be somewhere in the lazy section.}
+%
+In practice, this \emph{eager policy} is too slow to be of practical interest:
+what we really want in practice is a \emph{lazy policy} that can be efficiently
+implemented.
+
+We can also test the lazy policy, even if less efficiently, and it should
+enforce an inductive property, which may be more amenable to proving.
+%
+\rb{Would need careful phrasing.}
+%
+\rb{Implied by the property we started with?}
+
+\rb{Explain and compare what each policy enforces.}
+
 \section{Lazy Stack Safety}

### 2020-07-03 main.tex
Quick intro to micropolicies

@@
 
-The call stack is an increasingly popular target for attacks
+The call stack is an increasingly\apt{``a perennially'' ?} popular target for attacks
 exploiting memory safety vulnerabilities, with potential consequences
@@
 examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
+\apt{is this gone?}
 \leo{Goal: on the left, a bad read; on the right, a bad write.}
@@
 but even here, stack safety is defined with respect to an abstract
-machine that ``exhibits stack safety by construction,'' begging
+machine that ``satisfies stack safety by construction,'' begging
 the question of what {\em is} the property that holds by construction?
+\apt{Suggest to move first sentence of next para up to this one.}
 
@@
 itself needs to be proteced from {\em its} callees.
-We address this challenge by varying the observation power
+We address this challenge by varying the observation power\apt{is this
+  the right word? doesn't it conflict with our later use of ``observation''?}
 of the attacker at call points, and demonstrate that our policy is
@@
   the sentence are about two very different things.  The second deserves its
-  own paragraph, IMO.}
+  own paragraph, IMO.}\apt{+1}
 
@@
 directly (via a jump instruction) or indirectly (by corrupting stack
-data and returning). Well-bracketed control flow establshes safety between
-functions: all code is owned by a function, and all jumps between functions
+data and returning). Well-bracketed control flow establishes safety between
+functions: all code is uniquely owned by some function, and all jumps between functions
 are either calls that enter at a marked entry point or returns that enter
@@
 
-In summary, we offer the following contributions:
+In summary, we offer the following contributions\apt{is this up to date with current paper organization?}:
 \begin{itemize}
@@
   that the stack unsafe examples from the literature also fail our
-  property~(\Cref{sec:disc}).
-\item As evidence that the property is not too strong, we show that it is
+  property~(\Cref{sec:disc}).\apt{Something about relation to StkTokens formalization?}
+\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
   realizible by
@@
 
+\rb{Better section name and headings, references.}
+
 \paragraph*{Micropolicies}
+%
+In order to enforce a given security policy on a system dynamically, we resort
+to a programmable, tag-based reference monitor that runs alongside the program.
+To control this monitor, we use a programming model that allows fine-grained
+manipulation of metadata tags to encode so-called \emph{micropolicies}
+\rb{cite}.
+
+In such a system, all memory addresses, registers and the program counter are
+enriched with an abstract metadata tag, which can point to the representation of
+an arbitrary data structure. A micropolicy is thus defined by a collection of
+rules to be checked as each instruction executes and together define the
+transfer function of the security policy. Each rule inspects the tags associated
+to the inputs of an instruction (PC, registers, memory) as well as the
+instruction opcode itself, and produces one of two outcomes: either the
+instruction is allowed to execute (and tags for the new PC and its result are
+produced and assigned to those), or the machine fail-stops with a policy
+violation.
+
+Efficient execution of these micropolicies relies on hardware implementations
+such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
+maps tag inputs to outputs and a mechanism to trap misses to software handlers
+and thereby computes the transfer function of the software-defined policy.
+Previous work has shown that a wide range of micropolicies can be defined and
+enforced efficiently \rb{cite}.
+%
+\rb{Maybe go into more detail, examples later\ldots or try to segue into stack
+policies}
 

### 2020-07-03 main.tex
a few comments on intro

@@
 
-\rb{Better section name and headings, references.}
-
 \paragraph*{Micropolicies}
-%
-In order to enforce a given security policy on a system dynamically, we resort
-to a programmable, tag-based reference monitor that runs alongside the program.
-To control this monitor, we use a programming model that allows fine-grained
-manipulation of metadata tags to encode so-called \emph{micropolicies}
-\rb{cite}.
-
-In such a system, all memory addresses, registers and the program counter are
-enriched with an abstract metadata tag, which can point to the representation of
-an arbitrary data structure. A micropolicy is thus defined by a collection of
-rules to be checked as each instruction executes and together define the
-transfer function of the security policy. Each rule inspects the tags associated
-to the inputs of an instruction (PC, registers, memory) as well as the
-instruction opcode itself, and produces one of two outcomes: either the
-instruction is allowed to execute (and tags for the new PC and its result are
-produced and assigned to those), or the machine fail-stops with a policy
-violation.
-
-Efficient execution of these micropolicies relies on hardware implementations
-such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
-maps tag inputs to outputs and a mechanism to trap misses to software handlers
-and thereby computes the transfer function of the software-defined policy.
-Previous work has shown that a wide range of micropolicies can be defined and
-enforced efficiently \rb{cite}.
-%
-\rb{Maybe go into more detail, examples later\ldots or try to segue into stack
-policies}
 

### 2020-07-03 main.tex
Fixes, nits, and variations

--- 2020-07-03 main.tex
@@
 \end{figure}
-
-To show a trace of the execution of this program, we are going to need
-a low-level view of both the code and the machine it runs on. While
-our concrete implementation (Section~\leo{cite}) operates on a full
-conventional RISC-V architecture, we will make a few simplifying
-assumptions here to ease the exposition.
-
-First of all, we are going to ignore the heap. We are mainly concerned
-with stack interactions so that will be the main focus. \leo{We're
-  going to come back to that at the end. Will we?} We are then going
-to abstract away low-level details of memory layout and present a
-machine that is word-addressed, with instructions starting from
-location 0, a dedicated output at location 99, and lay the stack from
-location 100 upwards. \leo{Explain more? How to connect to realistic
-  one? Observations? DMA? Interrupt handlers?}
-%
-A sketch of the memory layout at that point can be found in the top of
-Figure~\ref{fig:running-trace-b}: instructions for {\tt main} begin at
-0, the dedicated output is empty, while the (uninitialized) stack
-begins at 100. Directly below the layout, we summarize the parts of
-the register file that are relevant to the example.
-%
-The rest of Figure~\ref{fig:running-trace-b} details the execution of the
-program, showing the machine state at call and return points, as well
-as describing the full effect of each instruction in detail.
-\leo{Expand upon this?}
-
-\bcp{ Comments on the example:
-\begin{itemize}
-\item We can simplify a bit by renaming $f$ to $\mathit{main}$ and
-relocating its stack to location 1000.  And maybe move its instructions to
-start at 0.  (We can comment that on a real RISC-V this is where interrupt
-handlers go, but POPL readers will not care too much.)  We should streamline
-the example as much as possible.
-%
-\item We should also emphasize, again, that there is no heap (if there were,
-we might put it above the stack or between the instructions and the stack).
-%
-\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
-capitalization of register names (sp vs SP), etc. Maybe typeset instructions
-in small caps?  Maybe typeset everything that is part of a concrete machine
-in small caps??
-%
-\item The colors are helpful in making the big page of instruction traces easier
-on the eye; maybe they would be even more helpful if
-corresponding colors appeared elsewhere in the example (on the instructions
-themselves, e.g.).  But we need to be careful to keep it readable -- maybe
-they should be background colors, like you've done, rather than text colors.
-\item The first column of numbers is redundant with the second column (PC).
-Suggest deleting the first.
-\item Asking readers to wade through all this is asking quite a bit!  We
-need to make sure we are going to get sufficient value out of the example
-later (I think we can, but we should really convince ourselves) to make it
-worthwhile.
-\item Use ``='' instead of ``:'' in the ``summaries'' below the pictures.
-And put a bit more space between the components.
-\item Create a new column that just literally pronounces each instruction in
-English, so people don't have to read RISC-V opcodes.
-\item As discussed, we should move this to before section 2, we should
-extend it with a bit more at the end so we can talk about the lazy policy
-with the same example, we should add
-in some bad variants (doing it here replaces the ``Examples'' from the
-Discussion section), and we should consider discussing the attacker model 
-here (we need to discuss it someplace in any case!)
-\end{itemize}
-}
 
@@
       {Deallocate {\tt h}'s call frame}
-  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
+  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 6]}{r_5 \leftarrow 0}
       {Load argument {\tt z} $r_5$}
@@
 
-%\end{minipage}
-%\begin{minipage}{0.7\textwidth}
-%\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
-%\newcommand*{\clabel}[1]{#1: & & &\\}
-%\[
-%\begin{array}{llll}
-%  \clabel{TOP}
-%  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
-%  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
-%  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
-%  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
-%  & & & \\
-%  & & & \\
-%  \clabel{F}
-%  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
-%  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
-%  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
-%  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
-%  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
-%  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
-%  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
-%  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
-%  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
-%  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
-%  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
-%  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
-%  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
-%  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
-%  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
-%  & & & \\
-%  & & & \\
-%  \clabel{G}
-%  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
-%  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
-%  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
-%  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
-%  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
-%  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
-%  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
-%  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}
-%\end{array}
-%\]
-%\end{minipage}
-%\label{fig:running-program}
-%\end{figure}
-
-%  sw r2 42 -8; store 42 in a private location
-%  sw r2 0 -4; store 0 as an argument
-%  sw r2 0 0; zero out the return location
-%  jal r1 F ; call F and store return address
-
-%  0: & \addi ~ r_4 ~ r_0 ~ 40 &
-%     ; ~ \text{\em put $40$ in register $r_4$}\\
-%  4: & \sw ~ r_4 ~ r_4 ~ 1000 &
-%     ; ~ \text{\em store $40$ in location $1040$}\\
-%  8: & \jal ~ r_1 ~ 20        &
-%     ; ~ \text{\em jump to $20$, store return addres in $r_1$}\\
-%  12: & & \\
-%  16: & & \\
-%  20: & \sw ~ r_2 ~ r_1 ~ 4 &
-%     ; ~ \text{\em store the return address to the $\SP + 4$}\\
-%  24: & \addi ~ r_2 ~ r_2 ~ 8 &
-%     ; ~ \text{\em increase the $\SP$ by 8}\\
-%\end{array}
-%\]
-%
-%  , (SW sp ra 4  , tagH1)
-%  , (ADDI sp sp 8, tagH2)
-%  , (LW 3 3 1000, instrTag)
-%  ]
+To show a trace of the execution of this program, we are going to need
+a low-level view of both the code and the machine it runs on. While
+our concrete implementation (Section~\leo{cite}) operates on a full
+conventional RISC-V architecture, we will make a few simplifying
+assumptions here to ease the exposition.
+
+\bcp{We should also emphasize, again, that there is no heap (if there were,
+we might put it above the stack or between the instructions and the stack)}
+First of all, we are going to ignore the heap. We are mainly concerned
+with stack interactions so that will be the main focus. \leo{We're
+  going to come back to that at the end. Will we?} We are then going
+to abstract away low-level details of memory layout and present a
+machine that is word-addressed, with instructions starting from
+location 0, a dedicated output at location 99, and lay the stack from
+location 100 upwards. \leo{Explain more? How to connect to realistic
+  one? Observations? DMA? Interrupt handlers?}
+%
+A sketch of the memory layout at that point can be found in the top of
+Figure~\ref{fig:running-trace-b}: instructions for {\tt main} begin at
+0, the dedicated output is empty, while the (uninitialized) stack
+begins at 100. Directly below the layout, we summarize the parts of
+the register file that are relevant to the example.
+%
+The rest of Figure~\ref{fig:running-trace-b} details the execution of the
+program, showing the machine state at call and return points, as well
+as describing the full effect of each instruction in detail.
+\leo{Expand upon this?}
+
+\bcp{ Comments on the example:
+\begin{itemize}
+\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
+capitalization of register names (sp vs SP), etc. Maybe typeset instructions
+in small caps?  Maybe typeset everything that is part of a concrete machine
+in small caps??
+%
+\item The colors are helpful in making the big page of instruction traces easier
+on the eye; maybe they would be even more helpful if
+corresponding colors appeared elsewhere in the example (on the instructions
+themselves, e.g.).  But we need to be careful to keep it readable -- maybe
+they should be background colors, like you've done, rather than text colors.
+\item Asking readers to wade through all this is asking quite a bit!  We
+need to make sure we are going to get sufficient value out of the example
+later (I think we can, but we should really convince ourselves) to make it
+worthwhile.
+\item Create a new column that just literally pronounces each instruction in
+English, so people don't have to read RISC-V opcodes.
+\item As discussed, we should move this to before section 2, we should
+extend it with a bit more at the end so we can talk about the lazy policy
+with the same example, we should add
+in some bad variants (doing it here replaces the ``Examples'' from the
+Discussion section), and we should consider discussing the attacker model 
+here (we need to discuss it someplace in any case!)
+\end{itemize}
+}
+
+\subsection{Attacker Model + Unsafe Variations}
+
+\leo{TODO: Make attacker model more explicit} The execution trace
+shown in Figure~\ref{fig:running-trace-b} is an example of a
+``stack-safe'' execution: each process interacts only with the
+registers, its local stack space and its call frame. But what could go
+wrong if an attacker could hijack an arbitrary call and execute
+arbitrary code? In the rest of this section we'll go over examples of
+unsafe behaviors, before formally capturing what stack-safe means
+formally and discussing enforcement techniques from the literature
+in later sections.
+
+\paragraph*{Bad write}
+
+\[
+44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
+\]
+
+Instruction 44 currently is $\sw ~ \rsp ~ r_4 ~ \negate 2$, with the
+effect of storing the contents of $r_4$ as the result of {\tt h}. If
+instead it wrote to location $[\rsp - 5]$ then it would be overriding
+private data (the return address) belonging to {\tt g}.
+
+\paragraph*{Bad read}
+
+\[
+30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 8
+\]
+
+Instruction 30 currently is $\lw ~ r_5 ~ \rsp ~ \negate 6$, with the
+effect of loading the argument of {\tt g} to $r_4$. If instead it
+loaded the contents of location $[\rsp - 8]$, it would obtain access
+to {\tt main}'s private information. Worse, since the next instruction
+outputs the contents of $r_5$ that would immediately be made public.
+
+\paragraph*{Control flow}
+
+\leo{What should we do here? Overwrite our own ra and return?}
 
@@
 
-To set the stage, we first
-describe our machine model. Our definition of stack safety is
-generic and this section discusses an abstract interface to our
-machine model and some kind of static or dynamic, software- or
-hardware-enforced security monitor.  For
-concreteness, our running examples will use a
-conventional machine architecture and a standard (RISC-V) instruction set;
-in~\Cref{sec:micropolicies} we will
-extend this model it with a dynamic policy enforcement mechanism based
-on \bcp{NickPaper,AndMaybeSomeOthers}.
+To set the stage, we first describe our machine model. Our definition
+of stack safety is generic and this section discusses an abstract
+interface to our machine model and some kind of static or dynamic,
+software- or hardware-enforced security monitor.  For concreteness,
+our running examples will use a conventional machine architecture and
+a standard (RISC-V) instruction set; in~\Cref{sec:micropolicies} we
+will extend this model it with a dynamic policy enforcement mechanism
+based on \bcp{NickPaper,AndMaybeSomeOthers}.
 

### 2020-07-03 main.tex
wibbles, comments

@@
 To help keep things concrete throughout the more technical sections
-that follow, we are first going to introduce a running example. We are
-going to use it to demonstrate various potentially problematic
+that follow, we first introduce a running example. We will
+use it to demonstrate various potentially problematic
 behaviours in this section, and then show how our properties and
@@
     \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[103] \leftarrow 0}
-        {Zero out result}
+        {Zero out result\apt{why?}}
     \row{\jal ~ \rra ~ 20}{??}{\PCname \leftarrow 20; ~ \rra \leftarrow \thepcctr}
@@
 
-To show a trace of the execution of this program, we are going to need
+To show a trace of the execution of this program, we need
 a low-level view of both the code and the machine it runs on. While
 our concrete implementation (Section~\leo{cite}) operates on a full
-conventional RISC-V architecture, we will make a few simplifying
+conventional RISC-V architecture, we make a few simplifying
 assumptions here to ease the exposition.
@@
 we might put it above the stack or between the instructions and the stack)}
-First of all, we are going to ignore the heap. We are mainly concerned
+First of all, we ignore the heap. We are mainly concerned
 with stack interactions so that will be the main focus. \leo{We're
-  going to come back to that at the end. Will we?} We are then going
-to abstract away low-level details of memory layout and present a
+  going to come back to that at the end. Will we?} We also
+abstract away low-level details of memory layout and present a
 machine that is word-addressed, with instructions starting from
@@
 \leo{Expand upon this?}
+
+\apt{Need to clarify to what extent our properties depend on the
+particular compilation style, calling sequences, etc. illustrated here.}
 

### 2020-07-03 main.tex
typo
wip

@@
 
+\bcp{``alternative framework that uses the conceptual framework of language-based security''}
+\bcp{idea:intrinsic vs. extrinsic: stktokens only talks}
+
+Another dimension: spatial vs temporal safety
+
 While integrity and confidentiality are relatively standard security
@@
 itself needs to be proteced from {\em its} callees.
-We address this challenge by varying the observation power\apt{is this
-  the right word? doesn't it conflict with our later use of ``observation''?}
+We address this challenge by varying the access powers
 of the attacker at call points, and demonstrate that our policy is
@@
 main = {
-  int x -42, y = 0;
+  int x = 42, y = 0;
   y = g(0);

### 2020-07-03 main.tex

@@
 \rb{usually though not always; as noted below there are other
-  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
-examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
-\apt{is this gone?}
-\leo{Goal: on the left, a bad read; on the right, a bad write.}
-\leo{Do we also need an example of ROP/control flow for the new
-  part?}\bcp{If it is not too complicated to explain, we should definitely
-  put it
-  here.  If it is complicated, we can just say in text that there are other
-  sorts of attacks too, and that we deal with them, and give a forward
-  pointer.}
+  approaches}\bcp{Yes, we have to be very careful about this}.
+\leo{Examples at a high level}
 Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
@@
 We address this challenge by varying the access powers
-of the attacker at call points, and demonstrate that our policy is
-implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
-  the sentence are about two very different things.  The second deserves its
-  own paragraph, IMO.}\apt{+1}
+of the attacker at call points.
 
@@
 
+ and demonstrate that our policy is
+implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
+  the sentence are about two very different things.  The second deserves its
+  own paragraph, IMO.}\apt{+1}
 %\leo{There needs to be some informal description
@@
 \begin{itemize}
-\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
-  intuitive concepts of local state encapsulation and
-  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
-  property is strong enough to capture the folk intuition, we show
-  that the stack unsafe examples from the literature also fail our
-  property~(\Cref{sec:disc}).\apt{Something about relation to StkTokens formalization?}
-\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
-  realizible by
-  a previously proposed micropolicy enforcement
-  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
-    substantial body of code}. We also present a variation of the
-  property + policy that is lazy, easier/more efficient to implement,
-  and provably weaker (\Cref{sec:lazy}).
-\item A provably stronger (inductive? step-by-step/stepwise) variant
-  of this property tuned for testing/verification.  We test the whole
-  thing with a cool framework. \bcp{Which section?}
+\item \leo{More? Extensional}\bcp{We might need to invent a word for
+  this---''quasi-extensional''?} formal characterizations of the
+  intuitive concepts of local state encapsulation and well-bracketed
+  control flow~(\Cref{sec:lse-and-wbcf}). \leo{Make one sentence} examples from the literature also fail our
+  property.\apt{Something about relation to StkTokens formalization?}
+\item Case Study:
+  \leo{Existing enforcement (Nick-eager). Satisfies Property.
+  Satisfies even stronger property (provably).}
+\item Pragmatically motivated refinement:
+  \leo{Lazy enforcement mechanism (Nick-lazy).
+    Plus a lazy variant of the property.}
+%\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
+%  realizible by
+%  a previously proposed micropolicy enforcement
+%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
+%    substantial body of code}. We also present a variation of the
+%  property + policy that is lazy, easier/more efficient to implement,
+%  and provably weaker (\Cref{sec:lazy}).
+%\item A provably stronger (inductive? step-by-step/stepwise) variant
+%  of this property tuned for testing/verification.  We test the whole
+%  thing with a cool framework. \bcp{Which section?}
 \end{itemize}
@@
 
+\leo{TODO: Add citations of mitigation techniques}
+
 \paragraph*{Bad write}
@@
 
-\section{Background}
+\paragraph*{Reusing call frames}
+
+\leo{TODO}
+
+\section{Machines, Traces, and Observations}
 \label{sec:prelim}
@@
 enforcement mechanism and give as many names and citations as we can.}
+
+\leo{Policies of interest: null, micro, cherri, software-only}
+\leo{TODO: Write down carefully}
 
@@
 annotations describing features of a program. A {\em code map}
+\leo{TODO: function map}
 identifies some addresses as code and assigns each to a function. A
@@
 
-\section{Discussion}
-\bcp{Make this a subsection?}
-\label{sec:disc}
-
-  \paragraph{Examples}
-
-\bcp{Belongs with the running example:}
-  \bcp{Illustrate the definition by using it to discuss a range of ``stack
-    unsafety'' examples from the literature. }
-
-  \paragraph{What about tail recursion and other non-stack-like control flow}
-
-\bcp{And this and the next belong in Future Work...}
-
-  \paragraph{Adding Heap Safety}
-
-    While these ideas are here presented in a stack-centric style, they could
-    be extended to heap safety in a system extended with a heap. Heap safety,
-    like stack safety, may be framed as a form of noninterference
-    [cite Meaning of Memory Safety]. For each certain component marked as
-    containing a valid pointer to a heap object, we can propagate that
-    validity when the pointer moves or is used in pointer arithmetic. Heap
-    objects to which a valid pointer is in a register are reachable, and
-    treated as low confidentiality and low integrity.
-
 \section{Enforcement}
@@
 \label{sec:future}
+
+\section{Discussion}
+
+\bcp{Make this a subsection?}
+\label{sec:disc}
+
+  \paragraph{What about tail recursion and other non-stack-like control flow}
+
+\bcp{And this and the next belong in Future Work...}
+
+  \paragraph{Adding Heap Safety}
+
+    While these ideas are here presented in a stack-centric style, they could
+    be extended to heap safety in a system extended with a heap. Heap safety,
+    like stack safety, may be framed as a form of noninterference
+    [cite Meaning of Memory Safety]. For each certain component marked as
+    containing a valid pointer to a heap object, we can propagate that
+    validity when the pointer moves or is used in pointer arithmetic. Heap
+    objects to which a valid pointer is in a register are reachable, and
+    treated as low confidentiality and low integrity.
 

### 2020-07-03 main.tex
Quick proposal to describe JAL(R) in examples

@@
 
-\bcp{``alternative framework that uses the conceptual framework of language-based security''}
-\bcp{idea:intrinsic vs. extrinsic: stktokens only talks}
-
-Another dimension: spatial vs temporal safety
-
 While integrity and confidentiality are relatively standard security
@@
 itself needs to be proteced from {\em its} callees.
-We address this challenge by varying the access powers
+We address this challenge by varying the observation power\apt{is this
+  the right word? doesn't it conflict with our later use of ``observation''?}
 of the attacker at call points.
@@
 \summary{\thepcctr}{0}{37}{102}{60}{42}
-\caption{Detailed Execution Trace}
+\caption{Detailed Execution Trace
+  \rb{A suggestion for the description of $\jal$ and $\jalr$:
+    call(function label) [if labels are added next to the entrypoints]
+    or call(address of entrypoint) and return()?}
+}
 \label{fig:running-trace-b}

### 2020-07-04 main.tex
From call
done with intro for now; releasing lock

--- 2020-07-03 main.tex
@@
                                         %% when present, will be used in
-                                        %% header instead of Full Title.
+v                                        %% header instead of Full Title.
 %\titlenote{with title note}             %% \titlenote is optional;
@@
 
-The call stack is an increasingly\apt{``a perennially'' ?} popular target for attacks
+The call stack is a perennially popular target for attacks
 exploiting memory safety vulnerabilities, with potential consequences
-ranging from attackers reading or corrupting private stack data, to
-highjacking a program's control flow. To prevent such attacks, many
-techniques have been proposed in the literature\leo{cite, point out?
-  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
-  stronger point.} aiming to achieve ``stack
-  safety''.
-
-But what {\em is} this stack safety of which we speak? A precise formal
-definition is lacking in the literature, with the notion usually being
-defined negatively: through examples of stack unsafety
-\rb{usually though not always; as noted below there are other
-  approaches}\bcp{Yes, we have to be very careful about this}.
-\leo{Examples at a high level}
-Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
-  we are only aware of one, quite recent, attempt to do something better} work reasoned
-about a stack safety enforcement mechanism using capabilities\leo{cite};
-but even here, stack safety is defined with respect to an abstract
-machine that ``satisfies stack safety by construction,'' begging
-the question of what {\em is} the property that holds by construction?
-\apt{Suggest to move first sentence of next para up to this one.}
-
-\leo{StackTokens} provides some intuition towards answering that question,
-by informally describing two components that comprise stack safety:
-local state encapsulation and well-bracketed control flow. The first
-contribution of the present paper is to
-turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
-  claim this} security properties,
-that together form a formal definition of stack safety.
-\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
-rigor in it, even if some notions are implicit or not fully
-developed.}\bcp{Yes, and let's remember that they are highly likely to be
-reviewing this paper!  Let's be as generous as we can be without
-under-selling our delta.}
-
+ranging from leakage or corruption of private stack data to
+hijacking of the program's control flow. To prevent such attacks, many
+software and hardware techniques for stack safety have been proposed, % in the literature and implemented in production systems,
+including stack canaries~\citep{Cowan+98},
+shadow stacks~\citep{Dang+15,Shanbhogue+19},
+split stacks~\citep{Kuznetsov+14},
+bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
+capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from nick;there could be more}
+
+But what {\em is} this ``stack safety'' of which we speak? Precise formal
+definitions are lacking in the literature; instead, the notion of safety
+is usually defined negatively, by enumerating the bad things that a safe system
+prevents, such as corruption of return addresses, buffer overflows, or
+use of uninitialized variables.
+%with the notion usually being
+%defined negatively: through examples of stack-based exploits 
+%\rb{usually though not always; as noted below there are other
+%  approaches}\bcp{Yes, we have to be very careful about this}.
+%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ... 
+We are only aware of one, quite recent, attempt to frame a positive definition.
+\citet{Skorstengaard+19} define stack safety as
+the conjunction of two properties, local state encapsulation and well-bracketed control flow.
+They formalize these properties by defining an abstract capability-based machine   
+that ``satisfies [stack safety] by construction.'' We believe there is still room
+to craft a more abstract and less intensional characterization of
+stack safety, and that is the goal of the present paper. \apt{still needs tuning}
+
+Our approach is to give alternative formal
+definitions of local state encapsulation and well-bracketed control flow
+using the conceptual framework of language-based security.
+%
 The underlying notion behind local state encapsulation is the
@@
 
-\bcp{``alternative framework that uses the conceptual framework of language-based security''}
-\bcp{idea:intrinsic vs. extrinsic: stktokens only talks}
-
-Another dimension: spatial vs temporal safety
-
-While integrity and confidentiality are relatively standard security
-properties, the stack-safety setting raises an interesting additional
-challenge:
-the notion of who the attacker is shifts throught execution. Not only
+% BCP idea: contrast intrinsic vs. extrinsic property definition
+% Remember to talk about spatial vs temporal safety
+
+While integrity and confidentiality are standard security
+properties, applying them in the stack-safety setting raises a novel challenge:
+the notion of ``attacker'' changes during execution. Not only
 does a caller need to be protected from its callee, but the callee
-itself needs to be proteced from {\em its} callees.
-We address this challenge by varying the access powers
-of the attacker at call points.
-
-Similarly, the underlying notion behind well-bracketed control flow is
+itself needs to be protected from {\em its} callees.
+We address this challenge by adjusting the integrity and confidentiality
+of different stack frames at each call or return transition between functions.\apt{connection to ``attacker'' still vague.}
+
+The underlying notion behind well-bracketed control flow is
 the need to protect the call structure itself: a callee should not be
@@
 data and returning). Well-bracketed control flow establishes safety between
-functions: all code is uniquely owned by some function, and all jumps between functions
-are either calls that enter at a marked entry point or returns that enter
+functions: each instruction is uniquely owned by some function, and all jumps between functions
+are either calls that enter at a marked entry point or returns that re-enter
 just after their associated call in the standard nested call structure.
-
- and demonstrate that our policy is
-implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
-  the sentence are about two very different things.  The second deserves its
-  own paragraph, IMO.}\apt{+1}
-%\leo{There needs to be some informal description
-%  of the wbcf part here. Sean?}
-
-In summary, we offer the following contributions\apt{is this up to date with current paper organization?}:
+\apt{clarify what we are contributing here.}
+
+To provide evidence that our definitions capture the inituitive notion of stack safety,
+we show that they rule out a wide \apt{really?} range of exploitable bad behaviors.\apt{``from the literature''??}
+To assess whether they are implementable, 
+we compare them against the proposed capability-based enforcement
+mechanism of \citet{Skorstengaard+19} \apt{we really should} and the existing tag-based enforcement mechanism
+of \citet{DBLP:conf/sp/RoesslerD18}.
+
+In summary, we offer the following contributions:
 \begin{itemize}
-\item \leo{More? Extensional}\bcp{We might need to invent a word for
-  this---''quasi-extensional''?} formal characterizations of the
+\item
+  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
+  Abstract formal properties capturing the 
   intuitive concepts of local state encapsulation and well-bracketed
-  control flow~(\Cref{sec:lse-and-wbcf}). \leo{Make one sentence} examples from the literature also fail our
-  property.\apt{Something about relation to StkTokens formalization?}
-\item Case Study:
-  \leo{Existing enforcement (Nick-eager). Satisfies Property.
-  Satisfies even stronger property (provably).}
-\item Pragmatically motivated refinement:
-  \leo{Lazy enforcement mechanism (Nick-lazy).
-    Plus a lazy variant of the property.}
+  control flow~(\Cref{sec:lse-and-wbcf}) in a very general setting. These properties
+  rule out common examples of stack memory abuse.
+\item
+  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\Cref{sec:enforcement}).
+  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
+  \apt{If we have no testing, do we even want to talk about the testing property?}
+\item
+  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy 
+  of \citet{DBLP:conf/sp/RoesslerD18}~(\Cref{sec:lazy}).  This policy does not enforce our property, but we give a
+  variant of the policy that does enforce a (provably) weaker variant of our property. \apt{Awkward, but the truth I fear.}
+
   %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
@@
 %
-We begin with a running example that we're going to keep coming back
+We begin with a running example that we refer
 to throughout the paper (Section~\ref{sec:running-example}) and some
@@
 %% Bibliography
-%\bibliography{bcp.bib}
+\bibliography{bcp.bib,local.bib}
 

### 2020-07-04 main.tex
Committing to be able to pull; working on examples

@@
 \input{macros}
+\input{examples}
 
@@
       {Deallocate {\tt h}'s call frame}
-  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 6]}{r_5 \leftarrow 0}
+  \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
       {Load argument {\tt z} $r_5$}
@@
 \begin{figure}
-% At Call to H
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
-  \memory{6}{red}[{\makebox[0pt]{\(\HIGHINT\)}}]%
-  \memory{2}{green}[{\makebox[0pt]{\(\LOWINT\)}}]%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\MemoryLabel{-30.5em}{0.75em}{8}
-\MemoryLabel{-26.5em}{0.75em}{}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-18.5em}{0.75em}{0}
-\\
-\end{center}
-% Starting Register File
-\summary{40}{0}{\thepcctr}{108}{17}{?}
-\setcounter{pcctr}{40}
-\vspace*{0.2em}
-% H - Body
-\[
-\begin{array}{c|c|c|c|c}
-  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
-      {Allocate for return address}
-  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
-      {Store return address}
-  \row{\lw ~ r_4 ~ \rsp ~ \negate 3}{r_4 \leftarrow [\rsp - 3]}{r_4 \leftarrow 17}
-      {Load the argument {\tt v} to $r_4$}
-  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 18}
-      {Increment $r_4$ by 1}
-  \row{\sw ~ \rsp ~ r_4 ~ \negate 5}{[\rsp - 5] \leftarrow r_4}{[103] \leftarrow 18}
-      {Store $r_4$ as {\tt g}'s return value}
-  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
-      {Load return address}
-  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
-      {Restore the stack pointer of {\tt g}}
-  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
-      {Return to {\tt g}}
-  \end{array}
-\]
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{44em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
-  \memory{3}{red}%
-  \memory{1}{green}%
-  \memory{2}{red}%
-  \memory{2}{gray}%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{18}
-\MemoryLabel{-30.5em}{0.75em}{8}
-\MemoryLabel{-26.5em}{0.75em}{}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-18.5em}{0.75em}{0}
-\MemoryLabel{-16em}{0.75em}{108}
-\\
-\end{center}
-%
+  \integrityexample
   \caption{Failing Trace Integrity}
@@
 \begin{figure}
-% At Call to H
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
-  \memory{6}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
-  \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\MemoryLabel{-30.5em}{0.75em}{8}
-\MemoryLabel{-26.5em}{0.75em}{}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-18.5em}{0.75em}{0}
-\\
-\end{center}
-% Variant
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{gray}%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{white}%
-  \memory{6}{yellow}[{\makebox[0pt]{Varied}}]%
-  \memory{2}{green}%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{?}
-\MemoryLabel{-42.5em}{0.75em}{?}
-\MemoryLabel{-38.5em}{0.75em}{?}
-\MemoryLabel{-34.5em}{0.75em}{?}
-\MemoryLabel{-30.5em}{0.75em}{?}
-\MemoryLabel{-26.5em}{0.75em}{?}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-18.5em}{0.75em}{0}
-\\
-\end{center}
-
-% Starting Register File
-\summary{40}{0}{\thepcctr}{108}{17}{?}
-\setcounter{pcctr}{40}
-\vspace*{0.2em}
-% H - Body
-\[
-\begin{array}{c|c|c|c|c}
-  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
-      {Allocate for return address}
-  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
-      {Store return address}
-  \row{\lw ~ r_4 ~ 100}{r_4 \leftarrow 100}{r_4 \leftarrow 42/?}
-      {Load {\tt main}'s secret {\tt x} to $r_4$}
-  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 43/(?+1)}
-      {Increment $r_4$ by 1}
-  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 43/(?+1)}
-      {Store $r_4$ as the return value}
-  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
-      {Load return address}
-  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
-      {Restore the stack pointer of {\tt g}}
-  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
-      {Return to {\tt g}}
-  \end{array}
-\]
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{44em}{2em}{\SP}
-\memory{3}{gray}%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{white}%
-  \memory{6}{red}%
-  \memory{1}{green}%
-  \memory{1}{pink}%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\MemoryLabel{-30.5em}{0.75em}{8}
-\MemoryLabel{-26.5em}{0.75em}{}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-19em}{0.75em}{43}
-\MemoryLabel{-16em}{0.75em}{108}
-\\
-\end{center}
-%
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{44em}{2em}{\SP}
-\memory{3}{gray}%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{white}%
-  \memory{6}{yellow}%
-  \memory{1}{green}%
-  \memory{1}{pink}%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{?}
-\MemoryLabel{-42.5em}{0.75em}{?}
-\MemoryLabel{-38.5em}{0.75em}{?}
-\MemoryLabel{-34.5em}{0.75em}{?}
-\MemoryLabel{-30.5em}{0.75em}{?}
-\MemoryLabel{-26.5em}{0.75em}{?}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-19em}{0.75em}{?}
-\MemoryLabel{-16em}{0.75em}{108}
-\\
-\end{center}
-%
-  \caption{Failing Trace Confidentiality}
+  \integritylazyexample
+  \caption{Rolling back from Figure \ref{fig:intex}}
+  \label{fig:intlex}
+\end{figure}
+
+\begin{figure}
+  \confidentialityduringexample
+  \caption{Failing Trace Confidentiality by Varying Observations}
+  \label{fig:confex1}
+\end{figure}
+
+\begin{figure}
+  \confidentialityendexample
+  \caption{Failing Trace Confidentiality at End State}
   \label{fig:confex2}

### 2020-07-04 main.tex
consistent use of Cref

--- 2020-07-04 main.tex
@@
 \end{itemize}
-Section~\ref{sec:relwork} discusses related work, and
-Section~\ref{sec:future} sketches directions for future work.
+\Cref{sec:relwork} discusses related work, and
+\Cref{sec:future} sketches directions for future work.
 %
 We begin with a running example that we refer
-to throughout the paper (Section~\ref{sec:running-example}) and some
-technical preliminaries (Section~\ref{sec:prelim}).
+to throughout the paper (\Cref{sec:running-example}) and some
+technical preliminaries (\Cref{sec:prelim}).
 
@@
 
-The program appears in Figure~\ref{fig:running-program}. It consists
+The program appears in \Cref{fig:running-program}. It consists
 of a {\tt main} function that performs a call to some function {\tt
@@
 A sketch of the memory layout at that point can be found in the top of
-Figure~\ref{fig:running-trace-b}: instructions for {\tt main} begin at
+\Cref{fig:running-trace-b}: instructions for {\tt main} begin at
 0, the dedicated output is empty, while the (uninitialized) stack
@@
 %
-The rest of Figure~\ref{fig:running-trace-b} details the execution of the
+The rest of \Cref{fig:running-trace-b} details the execution of the
 program, showing the machine state at call and return points, as well
@@
 \leo{TODO: Make attacker model more explicit} The execution trace
-shown in Figure~\ref{fig:running-trace-b} is an example of a
+shown in \Cref{fig:running-trace-b} is an example of a
 ``stack-safe'' execution: each process interacts only with the
@@
 our running examples will use a conventional machine architecture and
-a standard (RISC-V) instruction set; in~\Cref{sec:micropolicies} we
+a standard (RISC-V) instruction set; in \Cref{sec:micropolicies} we
 will extend this model it with a dynamic policy enforcement mechanism
@@
 
-For an example see Figure \ref{fig:intex}, in which {\tt h} writes its return
+For an example see \Cref{fig:intex}, in which {\tt h} writes its return
 value to {\tt g}'s return location the stack rather than its own. That address
@@
 
-Figure \ref{fig:confex2} shows the initial states \(\mach_e\), left,
+\Cref{fig:confex2} shows the initial states \(\mach_e\), left,
 and \(\mach_e'\), right. Low confidentiality data in green is shared
@@
 
-Figure \ref{confex2} gives an example of a trace that fails trace
+\Cref{confex2} gives an example of a trace that fails trace
 confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
@@
 
-      In Figure \ref{fig:lazyex1}, the final states of the original and variant
+      In \Cref{fig:lazyex1}, the final states of the original and variant
       differ, because a written value (in black or white) depended on the value

### 2020-07-04 main.tex
rationalize cref vs Cref using [capitalize] option to cleverf

@@
                                         %% when present, will be used in
-v                                        %% header instead of Full Title.
+                                        %% header instead of Full Title.
 %\titlenote{with title note}             %% \titlenote is optional;
@@
   intuitive concepts of local state encapsulation and well-bracketed
-  control flow~(\Cref{sec:lse-and-wbcf}) in a very general setting. These properties
+  control flow~(\cref{sec:lse-and-wbcf}) in a very general setting. These properties
   rule out common examples of stack memory abuse.
 \item
-  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\Cref{sec:enforcement}).
+  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
   This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
@@
   A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy 
-  of \citet{DBLP:conf/sp/RoesslerD18}~(\Cref{sec:lazy}).  This policy does not enforce our property, but we give a
+  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
   variant of the policy that does enforce a (provably) weaker variant of our property. \apt{Awkward, but the truth I fear.}
@@
 \Cref{sec:relwork} discusses related work, and
-\Cref{sec:future} sketches directions for future work.
+\cref{sec:future} sketches directions for future work.
 %
 We begin with a running example that we refer
-to throughout the paper (\Cref{sec:running-example}) and some
-technical preliminaries (\Cref{sec:prelim}).
+to throughout the paper (\cref{sec:running-example}) and some
+technical preliminaries (\cref{sec:prelim}).
 
@@
 
-The program appears in \Cref{fig:running-program}. It consists
+The program appears in \cref{fig:running-program}. It consists
 of a {\tt main} function that performs a call to some function {\tt
@@
 A sketch of the memory layout at that point can be found in the top of
-\Cref{fig:running-trace-b}: instructions for {\tt main} begin at
+\cref{fig:running-trace-b}: instructions for {\tt main} begin at
 0, the dedicated output is empty, while the (uninitialized) stack
@@
 %
-The rest of \Cref{fig:running-trace-b} details the execution of the
+The rest of \cref{fig:running-trace-b} details the execution of the
 program, showing the machine state at call and return points, as well
@@
 \leo{TODO: Make attacker model more explicit} The execution trace
-shown in \Cref{fig:running-trace-b} is an example of a
+shown in \cref{fig:running-trace-b} is an example of a
 ``stack-safe'' execution: each process interacts only with the
@@
 our running examples will use a conventional machine architecture and
-a standard (RISC-V) instruction set; in \Cref{sec:micropolicies} we
+a standard (RISC-V) instruction set; in \cref{sec:micropolicies} we
 will extend this model it with a dynamic policy enforcement mechanism
@@
 
-For an example see \Cref{fig:intex}, in which {\tt h} writes its return
+For an example see \cref{fig:intex}, in which {\tt h} writes its return
 value to {\tt g}'s return location the stack rather than its own. That address
@@
 
-\Cref{confex2} gives an example of a trace that fails trace
+\Cref{fig:confex2} gives an example of a trace that fails trace
 confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
@@
 
-      In \Cref{fig:lazyex1}, the final states of the original and variant
+      In \cref{fig:lazyex1}, the final states of the original and variant
       differ, because a written value (in black or white) depended on the value

### 2020-07-04 main.tex
Revamp 2 and 3.1

--- 2020-07-04 main.tex
@@
 \input{macros}
-\input{examples}
 
@@
       {Deallocate {\tt h}'s call frame}
-  \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
+  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 6]}{r_5 \leftarrow 0}
       {Load argument {\tt z} $r_5$}
@@
 
-To set the stage, we first describe our machine model. Our definition
-of stack safety is generic and this section discusses an abstract
-interface to our machine model and some kind of static or dynamic,
-software- or hardware-enforced security monitor.  For concreteness,
-our running examples will use a conventional machine architecture and
-a standard (RISC-V) instruction set; in~\Cref{sec:micropolicies} we
-will extend this model it with a dynamic policy enforcement mechanism
-based on \bcp{NickPaper,AndMaybeSomeOthers}.
-
-\paragraph*{Values, States, and Observations}
+In this section we describe our machine model. Our definition of stack
+safety is generic and this section discusses an abstract interface to
+our machine model and some kind of static or dynamic, software- or
+hardware-enforced security monitor. We will refer back to the running
+example of the previous section for concreteness, and then,
+in~\Cref{sec:micropolicies}, extend this model with a dynamic policy
+enforcement mechanism in~\Cref{sec:micropolicies} based on
+\bcp{NickPaper,AndMaybeSomeOthers}.
+
+\subsection{Values, States, and Observations}
 
@@
 
-The step function of the
-machine takes a machine state and produces a pair of a
-machine state and an {\em observation}, which may be either
-silent (written $\tau$) or a word of data that the machine communicates to
-the outside world (e.g., by writing to a DMA address):\leo{Needs more
-  exposition}\bcp{Like that?}
+The step function of the machine takes a machine state and produces a
+pair of a machine state and an {\em observation}, which may be either
+silent (written $\tau$) or a word of data that the machine
+communicates to the outside world (e.g., by writing to a DMA
+address):
 %
@@
 
-\paragraph*{Running Example: RISC-V}
-
 Mapping a conventional architecture to this abstract interface is
-straightforward. To build up our running example, we are going to use
-a standard 64-bit RISC machine, just like the one we are using in our
-experiments later on \leo{Mention FORVIS?}. In this setting, $\WORDS$
-is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
-r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$.
-% \leo{Mention We use the standard encoding of instructions to words?}
-% BCP: Nah.
-
-The first interesting choice is that of observations, as there are
-multiple reasonable options. At one extreme, we could choose to define
-observations as the value written at any component. However, that
-would pose problems later on when we treat some components as ``secret''
-when they correspond to a caller's private state. \leo{On the other extreme,
-we could pick a particular component and treat it as the ``output'' of
-our system. }  \bcp{But in this paragraph we are describing the particular
+straightforward. For the running example of the previous section,
+$\WORDS$ is the set of 64-bit integers, $\REGS$ is the set $\{r_0,
+\cdots, r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$. Observations are
+the writes to the special $\rout$ location: if during an execution
+step a word $w$ is written to $\rout$ then $\obs = w$, otherwise $\obs
+= \tau$.
+
+\leo{Other choices for observations? Where should that be discussed?}
+\bcp{But in this paragraph we are describing the particular
 choise that we make for our running example, no?  But we certainly want to
@@
 
-\paragraph*{Policies}
+\subsection{Policies}
 
@@
 
-\paragraph{Traces}
-
-A trace over a type $A$ is a nonempty, finite or infinite
-sequence of $A$s. We will mostly be concerned with sequences of
-machines states, usually\bcp{not always?} denoted $\machT$, and with sequences of
-triples of (machine states, policy states, and observations), usually
-denoted $\MPT$.
-\rb{The way the definitions are set up, one may expect MPO, but\ldots}
-To continue\bcp{??}, we will require four generic operations
-on traces: a ``trace-of'' operation to construct the
-trace induced by a step relation and a machine state
-\rb{not an op on traces per se?}\bcp{+1};
-a ``split'', to break a
-trace into two components based on a predicate; a ``join'', the
-reverse of split; and a ``prefix'', as a shorthand to only keep the
-first component of a split.  \bcp{Not sure it's worth enumerating them
-  rather than just getting down to the actual definitions below.}
+\subsection{Traces}
+
+\leo{Which of the machine, machine-observation, machine-policy,
+  and machine-policy-observation traces should we distinguish
+  in metavariables?}
+
+A trace over a type $A$ is a nonempty, finite or infinite sequence of
+$A$s. We will mostly be concerned with sequences of machines states,
+denoted $\machT$, and with sequences of triples of (machine states,
+policy states, and observations), usually denoted $\MPT$.  \rb{The way
+  the definitions are set up, one may expect MPO, but\ldots} We
+will require four generic operations on traces.
+
+\bcp{TODO: Let's also try to clean up the spacing a little -- many of these
+  sequences of rules can be typeset on fewer lines}
+
+\paragraph*{Trace-Of}
 
@@
 relates a machine state with the trace of machine states paired with
-observations produced by repeated application of step:  \bcp{above, we
-  showed operators applied to metavariables, not just the bare operator
-  symbol.}
-%
-\[\hookrightarrow : \machT \rightarrow \MOTRACE\]%
+observations produced by repeated application of step:  
+%
+\[\mach \hookrightarrow \machOT : \MACHS \rightarrow \MOTRACE\]%
 %
@@
 
-\leo{TODO: Add an example of a trace once I figure out the notations}
-\bcp{Let's also try to clean up the spacing a little -- many of these
-  sequences of rules can be typeset on fewer lines}
-
-Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
-state pair to the trace of triples of machine states, policy states,
-and observations induced by by \(\stepstoobs{}_P\): \bcp{Observe that this
-  time the trace can be either finite or infinite.}
-%
-\[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \MPOTRACE \]
+In our running example in Figure~\Cref{fig:running-example-b} we can
+see a depiction of such a trace: starting from an initial machine
+state at the very top, each instruction constructs a new machine state
+(the diff of which compared to the previous one appears in the
+``Effect'' column) and the majority of observations are $\tau$ with
+the exception of instructions 31 and 13 which are writes to $\rout$
+and where the observations are $0$ and $60$ respectively.
+
+In a similar fashion, we define \(\hookrightarrow_P\) to relate a
+machine-policy state pair to the trace of triples of machine states,
+policy states, and observations induced by by \(\stepstoobs{}_P\).
+Unlike the $\MOTRACE$ above, this trace can be finite as the
+policy can cause a fault:
+%
+\[(m,p) \hookrightarrow_P MP : \MACHS \times \POLS \rightarrow \MPOTRACE \]
 %
@@
   all these.  E.g., do these traces include the final state of a finite
-  execution?  If so, what observation is it paired with?}
-
-% SNA: fixed
-%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
-%  seems better}
-The {\it prefUpTo} operation \(\PUT(\Trace,f)\) \bcp{Why does it have two
-  different abbreviated names?  And wouldn't ``While'' be more mnemonic?}
-takes a trace \(T\)
-and a predicate on elements \(f\) and gives the prefix of \(T\)
-ending with the first element on which \(f\) holds, coinductively:
-
+  execution?  If so, what observation is it paired with?}\leo{The final
+  observation is a $\tau$ based on the rule in the inlined figure, so
+  it is essentially ignored. Point that out somehow?}
+
+\paragraph*{Prefix-Up-To}
+The ``prefix-up-to'' operation \(\PUT(\Trace,f)\) \bcp{Wouldn't
+  ``While'' be more mnemonic?}  takes a trace \(T\) and a predicate on
+elements \(f\) and gives the prefix of \(T\) ending with the first
+element on which \(f\) holds, coinductively:
+
+\begin{center}
+\begin{minipage}{.3\textwidth}
 \judgment{\(f\ \traceelem\)}
          {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
-%
+\end{minipage}
+%
+\begin{minipage}{.3\textwidth}
 \judgment{}
@@
          %
+\end{minipage}                  
+\begin{minipage}{.3\textwidth}
 \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
             {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
+\end{minipage}
+\end{center}
 %
@@
 Let \(f\) be a predicate on machine states. Then we define \(\mach
-\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
-running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I think.}
+\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
+running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I
+  think.}  \leo{There were inconsistencies which i fixed. I wanted the
+  metavariables defined to be consistent and the ``temporary''
+  variable which is the full trace to be primed. Should I swap those?}
 
@@
 
-\judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
-            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT'\)}
-         {\(\mpstatename \hookrightarrow \MPT' | f\)}
+\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
+            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT') = \MPT\)}
+         {\(\mpstatename \hookrightarrow \MPT | f\)}
 
@@
 We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-\leo{Benjamin suggested ``as far as an attacker is concerned'' in an
-  e-mail, but we haven't said anything about attackers yet.}\bcp{We don't
-  have to talk about attackers, just external observers.}, if the
-sequence of non-silent observations of $\obsT_1$ is a prefix of those
-of $\obsT_2$; that is we operate up to deletion of \(\tau\)
-observations, coinductively:
+as far as an external observer is concerned, if the sequence of
+non-silent observations of $\obsT_1$ is a prefix of those of
+$\obsT_2$; that is we operate up to deletion of \(\tau\) observations,
+coinductively:
 
@@
 In this section we will describe stack safety as a formal property. We
-begin with introducing a running example that will help keep things
-concrete throughout this section, continue with our model of
-\leo{compiler-provided?} knowledge about the program structure, and
-then dive in to formal definitions local state encapsulation and
-well-bracketed control flow.
+begin with our model of \leo{compiler-provided?} knowledge about the
+program structure, and then dive in to formal definitions local state
+encapsulation and well-bracketed control flow.
 
@@
 Our properties and policies are parameterized by several types of
-annotations describing features of a program. A {\em code map}
-\leo{TODO: function map}
-identifies some addresses as code and assigns each to a function. A
-{\em call map} identifies addresses that complete a call sequence,
-abstracting away from any concrete calling convention. Likewise a {\em
-  return map} identifies addresses that complete a return sequence. An
-{\em entry map} identifies valid entry points to a function.
-
-Formally a code map\bcp{I'd love to change ``cdm'' to ``km''.  Should we
-  also call it a ``kode map'' in English?  Maybe that's a bit too far.} is a
+annotations describing features of a program.
+
+\paragraph*{Function Map}
+
+A {\em function map} identifies some addresses as code and assigns
+each to a function. Formally, a function map is a
 partial function from addresses to a function
@@
 %
-In our running example, instructions 100-124 correspond to $f$,
-instructions 140-196 correspond to $g$, and instructions 200-228
-correspond to $h$.
-
-A call map is a partial function from addresses to a natural number,
-the number of arguments and return values of the call.
+In our running example, instructions 1-13 correspond to {\tt main},
+instructions 20-36 correspond to {\tt g}, and instructions 40-47
+correspond to {\tt h}, which means that the function map would be:
+\[ \codemap = \{\texttt{main} \mapsto \{1\ldots13\}, \texttt{g} \mapsto \{20\ldots36\}, \texttt{h} \mapsto \{40\ldots47\}\}\]
+
+\paragraph*{Call Map}
+
+A {\em call map} identifies addresses that complete a call sequence,
+abstracting away from any concrete calling convention. Formally, it is
+a partial function from addresses to a natural number, the number of
+arguments and return values of the call.
   %
@@
 %
-In our running example, the callmap would contain (at least) two
-entries: the word 112 corresponding to the first $\jal$
-instruction to $g$, which would map to $2$ as the call has one
-argument; and the word 164 also mapping to $2$ corresponding
-to the call to $h$.
+In our running example, the callmap contains two entries: the word 7
+corresponding to the call to {\tt g} which has one argument and one
+return value; and the word 26 corresponding to the call to {\tt h},
+also with a single argument and return value:
+\[ \callmap = \{ 7 \mapsto 2, 26 \mapsto 2\}\]
 
 For concreteness, in a typical calling convention, a call map may be
-considered to identify the state just before the \(\JAL\), in which
+considered to identify the state just before the \jal, in which
 space for arguments is already allocated, but the mechanism is more
@@
 
-A return map is simply a predicate on addresses and may be considered
-to identify the final jump of the return sequence. An entry map is
-likewise a predicate on addresses.
+\paragraph*{Return Map}
+Similarly, a {\em return map} identifies addresses that complete a
+return sequence. Formally, it is simply a set of components; each one
+is the final jump corresponding to a function:
 %
 \[\retmap \in \RETMAPS \subseteq \WORDS\]
+%
+In our running example, the return map identifies the $\jalr$
+instructions 36 and 47:
+\[ \retmap = \{36, 47\} \]
+
+\paragraph*{Entry Map}
+Finally, the entry map identifies valid entry points to a
+function and is also a set:
 \[\entmap \in \ENTMAPS \subseteq \WORDS\]
 %
-In our running example, the return map identifies the $\jalr$
-instructions 196 and 224, while the entry map identifies
-instructions 140 and 200.
+In our running example, the entry map contains the entry
+points of {\tt g} and {\tt h}:
+\[ \entmap = \{20, 40\} \]
 
@@
 
-We separately identify, for any call state, what is means for a state
+We separately identify, for any call state, what it means for a state
 to have correctly returned from a call. We consider the call to have
 returned the first time control returns to the instruction following
-the \(\JAL\) with the stack restored. Formally this is a relation
+the \jal with the stack restored. Formally this is a relation
 between the states:
@@
 
-\leo{Point to the trace. two return points}
+There are two such return points in our running example, both of which
+are depicted in the figure: the state before the execution of
+instruction 27 corresponding to the return from {\tt h} and the state
+before instruction 8 corresponding to the return from {\tt g}.
 
@@
 \begin{figure}
-  \integrityexample
+% At Call to H
+% Memory Layout
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+  \memory{6}{red}[{\makebox[0pt]{\(\HIGHINT\)}}]%
+  \memory{2}{green}[{\makebox[0pt]{\(\LOWINT\)}}]%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\\
+\end{center}
+% Starting Register File
+\summary{40}{0}{\thepcctr}{108}{17}{?}
+\setcounter{pcctr}{40}
+\vspace*{0.2em}
+% H - Body
+\[
+\begin{array}{c|c|c|c|c}
+  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
+      {Allocate for return address}
+  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
+      {Store return address}
+  \row{\lw ~ r_4 ~ \rsp ~ \negate 3}{r_4 \leftarrow [\rsp - 3]}{r_4 \leftarrow 17}
+      {Load the argument {\tt v} to $r_4$}
+  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 18}
+      {Increment $r_4$ by 1}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 5}{[\rsp - 5] \leftarrow r_4}{[103] \leftarrow 18}
+      {Store $r_4$ as {\tt g}'s return value}
+  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
+      {Load return address}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
+      {Restore the stack pointer of {\tt g}}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
+      {Return to {\tt g}}
+  \end{array}
+\]
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{44em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+  \memory{3}{red}%
+  \memory{1}{green}%
+  \memory{2}{red}%
+  \memory{2}{gray}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{18}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\MemoryLabel{-16em}{0.75em}{108}
+\\
+\end{center}
+%
   \caption{Failing Trace Integrity}
@@
 \begin{figure}
-  \integritylazyexample
-  \caption{Rolling back from Figure \ref{fig:intex}}
-  \label{fig:intlex}
-\end{figure}
-
-\begin{figure}
-  \confidentialityduringexample
-  \caption{Failing Trace Confidentiality by Varying Observations}
-  \label{fig:confex1}
-\end{figure}
-
-\begin{figure}
-  \confidentialityendexample
-  \caption{Failing Trace Confidentiality at End State}
+% At Call to H
+% Memory Layout
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+  \memory{6}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
+  \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\\
+\end{center}
+% Variant
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{61.5em}{2em}{\SP}
+\memory{3}{gray}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}%
+  \memory{6}{yellow}[{\makebox[0pt]{Varied}}]%
+  \memory{2}{green}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{?}
+\MemoryLabel{-42.5em}{0.75em}{?}
+\MemoryLabel{-38.5em}{0.75em}{?}
+\MemoryLabel{-34.5em}{0.75em}{?}
+\MemoryLabel{-30.5em}{0.75em}{?}
+\MemoryLabel{-26.5em}{0.75em}{?}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-18.5em}{0.75em}{0}
+\\
+\end{center}
+
+% Starting Register File
+\summary{40}{0}{\thepcctr}{108}{17}{?}
+\setcounter{pcctr}{40}
+\vspace*{0.2em}
+% H - Body
+\[
+\begin{array}{c|c|c|c|c}
+  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
+      {Allocate for return address}
+  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
+      {Store return address}
+  \row{\lw ~ r_4 ~ 100}{r_4 \leftarrow 100}{r_4 \leftarrow 42/?}
+      {Load {\tt main}'s secret {\tt x} to $r_4$}
+  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 43/(?+1)}
+      {Increment $r_4$ by 1}
+  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 43/(?+1)}
+      {Store $r_4$ as the return value}
+  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
+      {Load return address}
+  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
+      {Restore the stack pointer of {\tt g}}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
+      {Return to {\tt g}}
+  \end{array}
+\]
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{44em}{2em}{\SP}
+\memory{3}{gray}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}%
+  \memory{6}{red}%
+  \memory{1}{green}%
+  \memory{1}{pink}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\MemoryLabel{-30.5em}{0.75em}{8}
+\MemoryLabel{-26.5em}{0.75em}{}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-19em}{0.75em}{43}
+\MemoryLabel{-16em}{0.75em}{108}
+\\
+\end{center}
+%
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{44em}{2em}{\SP}
+\memory{3}{gray}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}%
+  \memory{6}{yellow}%
+  \memory{1}{green}%
+  \memory{1}{pink}%
+\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{?}
+\MemoryLabel{-42.5em}{0.75em}{?}
+\MemoryLabel{-38.5em}{0.75em}{?}
+\MemoryLabel{-34.5em}{0.75em}{?}
+\MemoryLabel{-30.5em}{0.75em}{?}
+\MemoryLabel{-26.5em}{0.75em}{?}
+\MemoryLabel{-22.5em}{0.75em}{17}
+\MemoryLabel{-19em}{0.75em}{?}
+\MemoryLabel{-16em}{0.75em}{108}
+\\
+\end{center}
+%
+  \caption{Failing Trace Confidentiality}
   \label{fig:confex2}

### 2020-07-04 main.tex

@@
 \end{itemize}
-Section~\ref{sec:relwork} discusses related work, and
-Section~\ref{sec:future} sketches directions for future work.
+\Cref{sec:relwork} discusses related work, and
+\cref{sec:future} sketches directions for future work.
 %
 We begin with a running example that we refer
-to throughout the paper (Section~\ref{sec:running-example}) and some
-technical preliminaries (Section~\ref{sec:prelim}).
+to throughout the paper (\cref{sec:running-example}) and some
+technical preliminaries (\cref{sec:prelim}).
 
@@
 
-The program appears in Figure~\ref{fig:running-program}. It consists
+The program appears in \cref{fig:running-program}. It consists
 of a {\tt main} function that performs a call to some function {\tt
@@
 A sketch of the memory layout at that point can be found in the top of
-Figure~\ref{fig:running-trace-b}: instructions for {\tt main} begin at
+\cref{fig:running-trace-b}: instructions for {\tt main} begin at
 0, the dedicated output is empty, while the (uninitialized) stack
@@
 %
-The rest of Figure~\ref{fig:running-trace-b} details the execution of the
+The rest of \cref{fig:running-trace-b} details the execution of the
 program, showing the machine state at call and return points, as well
@@
 \leo{TODO: Make attacker model more explicit} The execution trace
-shown in Figure~\ref{fig:running-trace-b} is an example of a
+shown in \cref{fig:running-trace-b} is an example of a
 ``stack-safe'' execution: each process interacts only with the
@@
 
-For an example see Figure \ref{fig:intex}, in which {\tt h} writes its return
+For an example see \cref{fig:intex}, in which {\tt h} writes its return
 value to {\tt g}'s return location the stack rather than its own. That address
@@
 
-Figure \ref{confex2} gives an example of a trace that fails trace
+\Cref{fig:confex2} gives an example of a trace that fails trace
 confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
@@
 
-      In Figure \ref{fig:lazyex1}, the final states of the original and variant
+      In \cref{fig:lazyex1}, the final states of the original and variant
       differ, because a written value (in black or white) depended on the value

### 2020-07-04 main.tex
Some more notes derived from yesterday's meeting

@@
 
-% BCP idea: contrast intrinsic vs. extrinsic property definition
-% Remember to talk about spatial vs temporal safety
+\bcp{BCP idea: contrast intrinsic vs. extrinsic property definition
+     Remember to talk about spatial vs temporal safety}
 
@@
 of different stack frames at each call or return transition between functions.\apt{connection to ``attacker'' still vague.}
+\rb{Mention to varying ``accessing power'' of the attacker?
+  The notion of attacker changes on certain actions that are not externally
+  visible.}
 
@@
 \leo{TODO: Add citations of mitigation techniques}
+\rb{And for those techniques, explain how they reflect their motivating
+examples. What does each technique do for those examples? How are they
+connected to this work?}
+
+\rb{Intuition that the callee can do anything in principle, but the machine will
+halt if it does not follow the calling convention, etc.?}
 
@@
 
+\rb{Notes from meeting:
+  \begin{itemize}
+  \item Explain how the theory applies to production compilers like GCC and
+    LLVM, which aspects are directly applicable, what are the effects of
+    alternative calling conventions. Other less charted waters: tail calls,
+    coroutines, jumps into supervisor mode.
+  \item CHERI and its operation, documentation and practical configurations and
+    operation.
+  \item Compare and contrast with StkTokens.
+  \item Under what circumstances can stack safety exist without an enforcement
+    mechanism?
+  \item Zeroing out of allocated space inside blessed sequences, eager vs lazy.
+    Possibility of blessed alloc and dealloc sequences, which also manipulate SP
+    and would not be legal without blessing. In such an extended setting, the
+    return sequence would check for matching addresses (it would be nice to have
+    testing for this!).
+  \end{itemize}
+}
+
 %% Acknowledgments

### 2020-07-05 main.tex
Tweaks

--- 2020-07-04 main.tex
@@
 
-The call stack is a perennially popular target for attacks
+The call stack is an increasingly\apt{``a perennially'' ?} popular target for attacks
 exploiting memory safety vulnerabilities, with potential consequences
-ranging from leakage or corruption of private stack data to
-hijacking of the program's control flow. To prevent such attacks, many
-software and hardware techniques for stack safety have been proposed, % in the literature and implemented in production systems,
-including stack canaries~\citep{Cowan+98},
-shadow stacks~\citep{Dang+15,Shanbhogue+19},
-split stacks~\citep{Kuznetsov+14},
-bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
-capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from nick;there could be more}
-
-But what {\em is} this ``stack safety'' of which we speak? Precise formal
-definitions are lacking in the literature; instead, the notion of safety
-is usually defined negatively, by enumerating the bad things that a safe system
-prevents, such as corruption of return addresses, buffer overflows, or
-use of uninitialized variables.
-%with the notion usually being
-%defined negatively: through examples of stack-based exploits 
-%\rb{usually though not always; as noted below there are other
-%  approaches}\bcp{Yes, we have to be very careful about this}.
-%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ... 
-We are only aware of one, quite recent, attempt to frame a positive definition.
-\citet{Skorstengaard+19} define stack safety as
-the conjunction of two properties, local state encapsulation and well-bracketed control flow.
-They formalize these properties by defining an abstract capability-based machine   
-that ``satisfies [stack safety] by construction.'' We believe there is still room
-to craft a more abstract and less intensional characterization of
-stack safety, and that is the goal of the present paper. \apt{still needs tuning}
-
-Our approach is to give alternative formal
-definitions of local state encapsulation and well-bracketed control flow
-using the conceptual framework of language-based security.
-%
+ranging from attackers reading or corrupting private stack data, to
+highjacking a program's control flow. To prevent such attacks, many
+techniques have been proposed in the literature\leo{cite, point out?
+  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
+  stronger point.} aiming to achieve ``stack
+  safety''.
+
+But what {\em is} this stack safety of which we speak? A precise formal
+definition is lacking in the literature, with the notion usually being
+defined negatively: through examples of stack unsafety
+\rb{usually though not always; as noted below there are other
+  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
+examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
+\apt{is this gone?}
+\leo{Goal: on the left, a bad read; on the right, a bad write.}
+\leo{Do we also need an example of ROP/control flow for the new
+  part?}\bcp{If it is not too complicated to explain, we should definitely
+  put it
+  here.  If it is complicated, we can just say in text that there are other
+  sorts of attacks too, and that we deal with them, and give a forward
+  pointer.}
+Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
+  we are only aware of one, quite recent, attempt to do something better} work reasoned
+about a stack safety enforcement mechanism using capabilities\leo{cite};
+but even here, stack safety is defined with respect to an abstract
+machine that ``satisfies stack safety by construction,'' begging
+the question of what {\em is} the property that holds by construction?
+\apt{Suggest to move first sentence of next para up to this one.}
+
+\leo{StackTokens} provides some intuition towards answering that question,
+by informally describing two components that comprise stack safety:
+local state encapsulation and well-bracketed control flow. The first
+contribution of the present paper is to
+turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
+  claim this} security properties,
+that together form a formal definition of stack safety.
+\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
+rigor in it, even if some notions are implicit or not fully
+developed.}\bcp{Yes, and let's remember that they are highly likely to be
+reviewing this paper!  Let's be as generous as we can be without
+under-selling our delta.}
+
 The underlying notion behind local state encapsulation is the
@@
 
-% BCP idea: contrast intrinsic vs. extrinsic property definition
-% Remember to talk about spatial vs temporal safety
-
-While integrity and confidentiality are standard security
-properties, applying them in the stack-safety setting raises a novel challenge:
-the notion of ``attacker'' changes during execution. Not only
+While integrity and confidentiality are relatively standard security
+properties, the stack-safety setting raises an interesting additional
+challenge:
+the notion of who the attacker is shifts throught execution. Not only
 does a caller need to be protected from its callee, but the callee
-itself needs to be protected from {\em its} callees.
-We address this challenge by adjusting the integrity and confidentiality
-of different stack frames at each call or return transition between functions.\apt{connection to ``attacker'' still vague.}
-
-The underlying notion behind well-bracketed control flow is
+itself needs to be proteced from {\em its} callees.
+We address this challenge by varying the observation power\apt{is this
+  the right word? doesn't it conflict with our later use of ``observation''?}
+of the attacker at call points, and demonstrate that our policy is
+implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
+  the sentence are about two very different things.  The second deserves its
+  own paragraph, IMO.}\apt{+1}
+
+Similarly, the underlying notion behind well-bracketed control flow is
 the need to protect the call structure itself: a callee should not be
@@
 data and returning). Well-bracketed control flow establishes safety between
-functions: each instruction is uniquely owned by some function, and all jumps between functions
-are either calls that enter at a marked entry point or returns that re-enter
+functions: all code is uniquely owned by some function, and all jumps between functions
+are either calls that enter at a marked entry point or returns that enter
 just after their associated call in the standard nested call structure.
-\apt{clarify what we are contributing here.}
-
-To provide evidence that our definitions capture the inituitive notion of stack safety,
-we show that they rule out a wide \apt{really?} range of exploitable bad behaviors.\apt{``from the literature''??}
-To assess whether they are implementable, 
-we compare them against the proposed capability-based enforcement
-mechanism of \citet{Skorstengaard+19} \apt{we really should} and the existing tag-based enforcement mechanism
-of \citet{DBLP:conf/sp/RoesslerD18}.
-
-In summary, we offer the following contributions:
+
+%\leo{There needs to be some informal description
+%  of the wbcf part here. Sean?}
+
+In summary, we offer the following contributions\apt{is this up to date with current paper organization?}:
 \begin{itemize}
-\item
-  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
-  Abstract formal properties capturing the 
-  intuitive concepts of local state encapsulation and well-bracketed
-  control flow~(\Cref{sec:lse-and-wbcf}) in a very general setting. These properties
-  rule out common examples of stack memory abuse.
-\item
-  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\Cref{sec:enforcement}).
-  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
-  \apt{If we have no testing, do we even want to talk about the testing property?}
-\item
-  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy 
-  of \citet{DBLP:conf/sp/RoesslerD18}~(\Cref{sec:lazy}).  This policy does not enforce our property, but we give a
-  variant of the policy that does enforce a (provably) weaker variant of our property. \apt{Awkward, but the truth I fear.}
-
-  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
-%  realizible by
-%  a previously proposed micropolicy enforcement
-%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
-%    substantial body of code}. We also present a variation of the
-%  property + policy that is lazy, easier/more efficient to implement,
-%  and provably weaker (\Cref{sec:lazy}).
-%\item A provably stronger (inductive? step-by-step/stepwise) variant
-%  of this property tuned for testing/verification.  We test the whole
-%  thing with a cool framework. \bcp{Which section?}
+\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
+  intuitive concepts of local state encapsulation and
+  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
+  property is strong enough to capture the folk intuition, we show
+  that the stack unsafe examples from the literature also fail our
+  property~(\Cref{sec:disc}).\apt{Something about relation to StkTokens formalization?}
+\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
+  realizible by
+  a previously proposed micropolicy enforcement
+  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
+    substantial body of code}. We also present a variation of the
+  property + policy that is lazy, easier/more efficient to implement,
+  and provably weaker (\Cref{sec:lazy}).
+\item A provably stronger (inductive? step-by-step/stepwise) variant
+  of this property tuned for testing/verification.  We test the whole
+  thing with a cool framework. \bcp{Which section?}
 \end{itemize}
@@
 %
-We begin with a running example that we refer
+We begin with a running example that we're going to keep coming back
 to throughout the paper (Section~\ref{sec:running-example}) and some
@@
 
-\section{Running Example}
+\section{Stack Safety and Unsafety by Example}
 \label{sec:running-example}
@@
 enforcement mechanisms preclude them later on in the paper.
+
+\subsection{Running Example}
 
@@
 \vspace*{0.2em}
+\bcp{Notation nit: in many assembly notations, square brackets mean
+  ``{\em indirect} through this location...''  And a very small nit: Change
+  $\rsp-2$ to $\rsp\mathord{-}2$}
 \[
@@
 \summary{\thepcctr}{0}{37}{102}{60}{42}
-\caption{Detailed Execution Trace
-  \rb{A suggestion for the description of $\jal$ and $\jalr$:
-    call(function label) [if labels are added next to the entrypoints]
-    or call(address of entrypoint) and return()?}
-}
+\caption{Detailed Execution Trace}
 \label{fig:running-trace-b}
@@
 
-\leo{TODO: Add citations of mitigation techniques}
-
 \paragraph*{Bad write}
@@
 
-\paragraph*{Reusing call frames}
-
-\leo{TODO}
-
-\section{Machines, Traces, and Observations}
+\section{Background}
 \label{sec:prelim}
 
-In this section we describe our machine model. Our definition of stack
-safety is generic and this section discusses an abstract interface to
-our machine model and some kind of static or dynamic, software- or
-hardware-enforced security monitor. We will refer back to the running
-example of the previous section for concreteness, and then,
-in~\Cref{sec:micropolicies}, extend this model with a dynamic policy
-enforcement mechanism in~\Cref{sec:micropolicies} based on
-\bcp{NickPaper,AndMaybeSomeOthers}.
-
-\subsection{Values, States, and Observations}
+To set the stage, we first describe our machine model. Our definition
+of stack safety is generic and this section discusses an abstract
+interface to our machine model and some kind of static or dynamic,
+software- or hardware-enforced security monitor.  For concreteness,
+our running examples will use a conventional machine architecture and
+a standard (RISC-V) instruction set; in~\Cref{sec:micropolicies} we
+will extend this model it with a dynamic policy enforcement mechanism
+based on \bcp{NickPaper,AndMaybeSomeOthers}.
+
+\paragraph*{Values, States, and Observations}
 
@@
 
-The step function of the machine takes a machine state and produces a
-pair of a machine state and an {\em observation}, which may be either
-silent (written $\tau$) or a word of data that the machine
-communicates to the outside world (e.g., by writing to a DMA
-address):
+The step function of the
+machine takes a machine state and produces a pair of a
+machine state and an {\em observation}, which may be either
+silent (written $\tau$) or a word of data that the machine communicates to
+the outside world (e.g., by writing to a DMA address):\leo{Needs more
+  exposition}\bcp{Like that?}
 %
@@
 
+\paragraph*{Running Example: RISC-V}
+
 Mapping a conventional architecture to this abstract interface is
-straightforward. For the running example of the previous section,
-$\WORDS$ is the set of 64-bit integers, $\REGS$ is the set $\{r_0,
-\cdots, r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$. Observations are
-the writes to the special $\rout$ location: if during an execution
-step a word $w$ is written to $\rout$ then $\obs = w$, otherwise $\obs
-= \tau$.
-
-\leo{Other choices for observations? Where should that be discussed?}
-\bcp{But in this paragraph we are describing the particular
+straightforward. To build up our running example, we are going to use
+a standard 64-bit RISC machine, just like the one we are using in our
+experiments later on \leo{Mention FORVIS?}. In this setting, $\WORDS$
+is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
+r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$.
+% \leo{Mention We use the standard encoding of instructions to words?}
+% BCP: Nah.
+
+The first interesting choice is that of observations, as there are
+multiple reasonable options. At one extreme, we could choose to define
+observations as the value written at any component. However, that
+would pose problems later on when we treat some components as ``secret''
+when they correspond to a caller's private state. \leo{On the other extreme,
+we could pick a particular component and treat it as the ``output'' of
+our system. }  \bcp{But in this paragraph we are describing the particular
 choise that we make for our running example, no?  But we certainly want to
@@
 
-\subsection{Policies}
+\paragraph*{Policies}
 
@@
 
-\leo{Policies of interest: null, micro, cherri, software-only}
-\leo{TODO: Write down carefully}
-
-\subsection{Traces}
-
-\leo{Which of the machine, machine-observation, machine-policy,
-  and machine-policy-observation traces should we distinguish
-  in metavariables?}
-
-A trace over a type $A$ is a nonempty, finite or infinite sequence of
-$A$s. We will mostly be concerned with sequences of machines states,
-denoted $\machT$, and with sequences of triples of (machine states,
-policy states, and observations), usually denoted $\MPT$.  \rb{The way
-  the definitions are set up, one may expect MPO, but\ldots} We
-will require four generic operations on traces.
-
-\bcp{TODO: Let's also try to clean up the spacing a little -- many of these
-  sequences of rules can be typeset on fewer lines}
-
-\paragraph*{Trace-Of}
+\paragraph{Traces}
+
+A trace over a type $A$ is a nonempty, finite or infinite
+sequence of $A$s. We will mostly be concerned with sequences of
+machines states, usually\bcp{not always?} denoted $\machT$, and with sequences of
+triples of (machine states, policy states, and observations), usually
+denoted $\MPT$.
+\rb{The way the definitions are set up, one may expect MPO, but\ldots}
+To continue\bcp{??}, we will require four generic operations
+on traces: a ``trace-of'' operation to construct the
+trace induced by a step relation and a machine state
+\rb{not an op on traces per se?}\bcp{+1};
+a ``split'', to break a
+trace into two components based on a predicate; a ``join'', the
+reverse of split; and a ``prefix'', as a shorthand to only keep the
+first component of a split.  \bcp{Not sure it's worth enumerating them
+  rather than just getting down to the actual definitions below.}
 
@@
 relates a machine state with the trace of machine states paired with
-observations produced by repeated application of step:  
-%
-\[\mach \hookrightarrow \machOT : \MACHS \rightarrow \MOTRACE\]%
+observations produced by repeated application of step:  \bcp{above, we
+  showed operators applied to metavariables, not just the bare operator
+  symbol.}
+%
+\[\hookrightarrow : \machT \rightarrow \MOTRACE\]%
 %
@@
 
-In our running example in Figure~\Cref{fig:running-example-b} we can
-see a depiction of such a trace: starting from an initial machine
-state at the very top, each instruction constructs a new machine state
-(the diff of which compared to the previous one appears in the
-``Effect'' column) and the majority of observations are $\tau$ with
-the exception of instructions 31 and 13 which are writes to $\rout$
-and where the observations are $0$ and $60$ respectively.
-
-In a similar fashion, we define \(\hookrightarrow_P\) to relate a
-machine-policy state pair to the trace of triples of machine states,
-policy states, and observations induced by by \(\stepstoobs{}_P\).
-Unlike the $\MOTRACE$ above, this trace can be finite as the
-policy can cause a fault:
-%
-\[(m,p) \hookrightarrow_P MP : \MACHS \times \POLS \rightarrow \MPOTRACE \]
+\leo{TODO: Add an example of a trace once I figure out the notations}
+\bcp{Let's also try to clean up the spacing a little -- many of these
+  sequences of rules can be typeset on fewer lines}
+
+Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
+state pair to the trace of triples of machine states, policy states,
+and observations induced by by \(\stepstoobs{}_P\): \bcp{Observe that this
+  time the trace can be either finite or infinite.}
+%
+\[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \MPOTRACE \]
 %
@@
   all these.  E.g., do these traces include the final state of a finite
-  execution?  If so, what observation is it paired with?}\leo{The final
-  observation is a $\tau$ based on the rule in the inlined figure, so
-  it is essentially ignored. Point that out somehow?}
-
-\paragraph*{Prefix-Up-To}
-The ``prefix-up-to'' operation \(\PUT(\Trace,f)\) \bcp{Wouldn't
-  ``While'' be more mnemonic?}  takes a trace \(T\) and a predicate on
-elements \(f\) and gives the prefix of \(T\) ending with the first
-element on which \(f\) holds, coinductively:
-
-\begin{center}
-\begin{minipage}{.3\textwidth}
+  execution?  If so, what observation is it paired with?}
+
+% SNA: fixed
+%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
+%  seems better}
+The {\it prefUpTo} operation \(\PUT(\Trace,f)\) \bcp{Why does it have two
+  different abbreviated names?  And wouldn't ``While'' be more mnemonic?}
+takes a trace \(T\)
+and a predicate on elements \(f\) and gives the prefix of \(T\)
+ending with the first element on which \(f\) holds, coinductively:
+
 \judgment{\(f\ \traceelem\)}
          {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
-\end{minipage}
-%
-\begin{minipage}{.3\textwidth}
+%
 \judgment{}
@@
 %
-\end{minipage}                  
-\begin{minipage}{.3\textwidth}
 \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
             {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
-\end{minipage}
-\end{center}
 %
@@
 Let \(f\) be a predicate on machine states. Then we define \(\mach
-\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
-running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I
-  think.}  \leo{There were inconsistencies which i fixed. I wanted the
-  metavariables defined to be consistent and the ``temporary''
-  variable which is the full trace to be primed. Should I swap those?}
+\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
+running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I think.}
 
@@
 
-\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
-            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT') = \MPT\)}
-         {\(\mpstatename \hookrightarrow \MPT | f\)}
+\judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
+            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT'\)}
+         {\(\mpstatename \hookrightarrow \MPT' | f\)}
 
@@
 We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-as far as an external observer is concerned, if the sequence of
-non-silent observations of $\obsT_1$ is a prefix of those of
-$\obsT_2$; that is we operate up to deletion of \(\tau\) observations,
-coinductively:
+\leo{Benjamin suggested ``as far as an attacker is concerned'' in an
+  e-mail, but we haven't said anything about attackers yet.}\bcp{We don't
+  have to talk about attackers, just external observers.}, if the
+sequence of non-silent observations of $\obsT_1$ is a prefix of those
+of $\obsT_2$; that is we operate up to deletion of \(\tau\)
+observations, coinductively:
 
@@
 In this section we will describe stack safety as a formal property. We
-begin with our model of \leo{compiler-provided?} knowledge about the
-program structure, and then dive in to formal definitions local state
-encapsulation and well-bracketed control flow.
+begin with introducing a running example that will help keep things
+concrete throughout this section, continue with our model of
+\leo{compiler-provided?} knowledge about the program structure, and
+then dive in to formal definitions local state encapsulation and
+well-bracketed control flow.
 
@@
 Our properties and policies are parameterized by several types of
-annotations describing features of a program.
-
-\paragraph*{Function Map}
-
-A {\em function map} identifies some addresses as code and assigns
-each to a function. Formally, a function map is a
+annotations describing features of a program. A {\em code map}
+identifies some addresses as code and assigns each to a function. A
+{\em call map} identifies addresses that complete a call sequence,
+abstracting away from any concrete calling convention. Likewise a {\em
+  return map} identifies addresses that complete a return sequence. An
+{\em entry map} identifies valid entry points to a function.
+
+Formally a code map\bcp{I'd love to change ``cdm'' to ``km''.  Should we
+  also call it a ``kode map'' in English?  Maybe that's a bit too far.} is a
 partial function from addresses to a function
@@
 %
-In our running example, instructions 1-13 correspond to {\tt main},
-instructions 20-36 correspond to {\tt g}, and instructions 40-47
-correspond to {\tt h}, which means that the function map would be:
-\[ \codemap = \{\texttt{main} \mapsto \{1\ldots13\}, \texttt{g} \mapsto \{20\ldots36\}, \texttt{h} \mapsto \{40\ldots47\}\}\]
-
-\paragraph*{Call Map}
-
-A {\em call map} identifies addresses that complete a call sequence,
-abstracting away from any concrete calling convention. Formally, it is
-a partial function from addresses to a natural number, the number of
-arguments and return values of the call.
+In our running example, instructions 100-124 correspond to $f$,
+instructions 140-196 correspond to $g$, and instructions 200-228
+correspond to $h$.
+
+A call map is a partial function from addresses to a natural number,
+the number of arguments and return values of the call.
   %
@@
 %
-In our running example, the callmap contains two entries: the word 7
-corresponding to the call to {\tt g} which has one argument and one
-return value; and the word 26 corresponding to the call to {\tt h},
-also with a single argument and return value:
-\[ \callmap = \{ 7 \mapsto 2, 26 \mapsto 2\}\]
+In our running example, the callmap would contain (at least) two
+entries: the word 112 corresponding to the first $\jal$
+instruction to $g$, which would map to $2$ as the call has one
+argument; and the word 164 also mapping to $2$ corresponding
+to the call to $h$.
 
 For concreteness, in a typical calling convention, a call map may be
-considered to identify the state just before the \jal, in which
+considered to identify the state just before the \(\JAL\), in which
 space for arguments is already allocated, but the mechanism is more
@@
 
-\paragraph*{Return Map}
-Similarly, a {\em return map} identifies addresses that complete a
-return sequence. Formally, it is simply a set of components; each one
-is the final jump corresponding to a function:
+A return map is simply a predicate on addresses and may be considered
+to identify the final jump of the return sequence. An entry map is
+likewise a predicate on addresses.
   %
 \[\retmap \in \RETMAPS \subseteq \WORDS\]
+\[\entmap \in \ENTMAPS \subseteq \WORDS\]
 %
 In our running example, the return map identifies the $\jalr$
-instructions 36 and 47:
-\[ \retmap = \{36, 47\} \]
-
-\paragraph*{Entry Map}
-Finally, the entry map identifies valid entry points to a
-function and is also a set:
-\[\entmap \in \ENTMAPS \subseteq \WORDS\]
-%
-In our running example, the entry map contains the entry
-points of {\tt g} and {\tt h}:
-\[ \entmap = \{20, 40\} \]
+instructions 196 and 224, while the entry map identifies
+instructions 140 and 200.
 
@@
 
-We separately identify, for any call state, what it means for a state
+We separately identify, for any call state, what is means for a state
 to have correctly returned from a call. We consider the call to have
 returned the first time control returns to the instruction following
-the \jal with the stack restored. Formally this is a relation
+the \(\JAL\) with the stack restored. Formally this is a relation
 between the states:
@@
 
-There are two such return points in our running example, both of which
-are depicted in the figure: the state before the execution of
-instruction 27 corresponding to the return from {\tt h} and the state
-before instruction 8 corresponding to the return from {\tt g}.
+\leo{Point to the trace. two return points}
 
@@
 
+\section{Discussion}
+\bcp{Make this a subsection?}
+\label{sec:disc}
+
+  \paragraph{Examples}
+
+\bcp{Belongs with the running example:}
+  \bcp{Illustrate the definition by using it to discuss a range of ``stack
+    unsafety'' examples from the literature. }
+
+  \paragraph{What about tail recursion and other non-stack-like control flow}
+
+\bcp{And this and the next belong in Future Work...}
+
+  \paragraph{Adding Heap Safety}
+
+    While these ideas are here presented in a stack-centric style, they could
+    be extended to heap safety in a system extended with a heap. Heap safety,
+    like stack safety, may be framed as a form of noninterference
+    [cite Meaning of Memory Safety]. For each certain component marked as
+    containing a valid pointer to a heap object, we can propagate that
+    validity when the pointer moves or is used in pointer arithmetic. Heap
+    objects to which a valid pointer is in a register are reachable, and
+    treated as low confidentiality and low integrity.
+
 \section{Enforcement}
@@
 \label{sec:future}
-
-\section{Discussion}
-
-\bcp{Make this a subsection?}
-\label{sec:disc}
-
-  \paragraph{What about tail recursion and other non-stack-like control flow}
-
-\bcp{And this and the next belong in Future Work...}
-
-  \paragraph{Adding Heap Safety}
-
-    While these ideas are here presented in a stack-centric style, they could
-    be extended to heap safety in a system extended with a heap. Heap safety,
-    like stack safety, may be framed as a form of noninterference
-    [cite Meaning of Memory Safety]. For each certain component marked as
-    containing a valid pointer to a heap object, we can propagate that
-    validity when the pointer moves or is used in pointer arithmetic. Heap
-    objects to which a valid pointer is in a register are reachable, and
-    treated as low confidentiality and low integrity.
 
@@
 %% Bibliography
-\bibliography{bcp.bib,local.bib}
+%\bibliography{bcp.bib}
 

### 2020-07-05 main.tex

--- 2020-07-04 main.tex
@@
 ranging from leakage or corruption of private stack data to
-hijacking of the program's control flow. To prevent such attacks, many
-software and hardware techniques for stack safety have been proposed, % in the literature and implemented in production systems,
+control-flow hijacking. To detect or prevent such attacks, many
+software and hardware techniques for stack safety have been proposed,
+%
 including stack canaries~\citep{Cowan+98},
+bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
+split stacks~\citep{Kuznetsov+14},
 shadow stacks~\citep{Dang+15,Shanbhogue+19},
-split stacks~\citep{Kuznetsov+14},
-bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
 capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from nick;there could be more}
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from
+  nick;there could be more}\bcp{How about static techniques??  We decided
+  these were also in-scope.}
 
@@
 \citet{Skorstengaard+19} define stack safety as
-the conjunction of two properties, local state encapsulation and well-bracketed control flow.
+the conjunction of two properties, {\em local state encapsulation} and {\em
+  well-bracketed control flow}.
 They formalize these properties by defining an abstract capability-based machine
@@
 to craft a more abstract and less intensional characterization of
-stack safety, and that is the goal of the present paper. \apt{still needs tuning}
-
-Our approach is to give alternative formal
-definitions of local state encapsulation and well-bracketed control flow
-using the conceptual framework of language-based security.
-%
-The underlying notion behind local state encapsulation is the
-intention to protect callers from callees: a callee should be not be
-able to access or overwrite a caller's private data. We formalize this
-intuition as a security property with both confidentiality and
-integrity components.
+stack safety, and that is the goal of the present paper. \apt{still needs
+  tuning}\bcp{Yes, I think we need to go into a little more detail on what
+  their definition looks like in order to justify why an alternative
+  definition might be valuable.  What questions does their presentation not
+  answer? What uses does it not support? E.g., can it be enforced
+  efficiently? (I guess so, by capabilities, though I think their
+  capabilities are more powerful than what Cheri actually provides, while
+  our policies are exactly what Nick needs.)  Why would it be good to look
+  at stack safety through the lens of security?  (Because it is
+  fundamentally a security mechanism?)}
+\rb{BCP idea: contrast intrinsic vs. extrinsic property definition.
+     Remember to talk about spatial vs temporal safety}
+
+Our main contribution in this paper is an alternative formal presentation of
+local state encapsulation and well-bracketed control flow using the
+conceptual and technical framework of language-based security.
+%
+The underlying notion behind local state encapsulation is the intention to
+protect callers from callees: a callee should be not be able to access or
+overwrite private data in the caller's stack frame. We formalize this
+intuition as a security property with both confidentiality and integrity
+components.
 %
@@
 with the heap safety property of [Azevedo et al]}
-
-\bcp{BCP idea: contrast intrinsic vs. extrinsic property definition
-     Remember to talk about spatial vs temporal safety}
 
@@
   The notion of attacker changes on certain actions that are not externally
-  visible.}
+  visible.} \bcp{Maybe the best way is to grab the bull by the horns: ``If
+  we're going to phrase stack safety as a security property, we need to talk
+  about an attacker model, and this is where the technicalities get really
+  interesting...''}
 
@@
   This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
-  \apt{If we have no testing, do we even want to talk about the testing property?}
+  \apt{If we have no testing, do we even want to talk about the testing
+    property?} \bcp{Not here; just in future work.  I guess we need to see
+    whether we wind up with any results at all about testing.}
 \item
@@
   of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
-  variant of the policy that does enforce a (provably) weaker variant of our property. \apt{Awkward, but the truth I fear.}
+  variant of the policy that does enforce a (provably) weaker variant of our
+  property.
+%
+  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
+    but the point we can make is still strong: To make the property easier
+    to enforce, we can slightly weaken its guarantees; the weaker version is
+    a bit more complicated to state but still retains the pragmatic
+    protections of the original.}
 
@@
 
-\section{Running Example}
+\section{Stack Safety and Unsafety by Example}
 \label{sec:running-example}
 
-To help keep things concrete throughout the more technical sections
-that follow, we first introduce a running example. We will
-use it to demonstrate various potentially problematic
-behaviours in this section, and then show how our properties and
-enforcement mechanisms preclude them later on in the paper.
-
-The program appears in \cref{fig:running-program}. It consists
-of a {\tt main} function that performs a call to some function {\tt
-  f}, which in turns makes a nested call to some function {\tt g}
-before returning, all while performing some simple arithmetic
+We begin with a running example that will serve both to concretely
+demonstrate the ``stack-unsafe'' behaviours we intend to preclude and to
+ground and illustrate the formal properties and enforcement mechanisms in
+later sections.
+
+\subsection{Running Example}
+
+The example program appears in \Cref{fig:running-program}. It consists of a
+{\tt main} function, written in some C-like high-level language, that
+performs a call to some function {\tt f}, which in turns makes a nested call
+to some function {\tt g} before returning, performing some simple arithmetic
 in the process.
@@
 \begin{figure}[b]
-  \begin{centering}
+\hfill
     \begin{minipage}{.3\textwidth}
@@
     \end{minipage}
-\end{centering}
-\caption{High-level Program with Simple Nested Calls}
+\caption{A High-level Program With Nested Calls}
 \label{fig:running-program}
@@
 \vspace*{0.2em}
+\bcp{A very small nit: Change
+  $\rsp-2$ to $\rsp\mathord{-}2$}
 \[
@@
         {Zero out result}
-    \row{\jal ~ \rra ~ 40}{??}{PC \leftarrow 40; ~ \rra \leftarrow \thepcctr}
+    \row{\jal ~ \rra ~ 40}{??}{\PCname \leftarrow 40; ~ \rra \leftarrow \thepcctr}
         {Make the call to {\tt h}}
@@
       {Restore the stack pointer of {\tt g}}
-  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 31}
       {Return to {\tt g}}

### 2020-07-05 main.tex
clarify comment

@@
     a bit more complicated to state but still retains the pragmatic
-    protections of the original.}
+    protections of the original.}\apt{To clarify: Nick's LT&C policy uses
+    depth-based tags, which do not enforce even our lazy policy. So we need to
+    introduce a version of LT&C that uses unique activation tags.}
 

### 2020-07-05 main.tex
Wrote some future work, spun out examples into a separate file for readability

--- 2020-07-05 main.tex
@@
 
-The call stack is an increasingly\apt{``a perennially'' ?} popular target for attacks
+The call stack is a perennially popular target for attacks
 exploiting memory safety vulnerabilities, with potential consequences
-ranging from attackers reading or corrupting private stack data, to
-highjacking a program's control flow. To prevent such attacks, many
-techniques have been proposed in the literature\leo{cite, point out?
-  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
-  stronger point.} aiming to achieve ``stack
-  safety''.
-
-But what {\em is} this stack safety of which we speak? A precise formal
-definition is lacking in the literature, with the notion usually being
-defined negatively: through examples of stack unsafety
-\rb{usually though not always; as noted below there are other
-  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
-examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
-\apt{is this gone?}
-\leo{Goal: on the left, a bad read; on the right, a bad write.}
-\leo{Do we also need an example of ROP/control flow for the new
-  part?}\bcp{If it is not too complicated to explain, we should definitely
-  put it
-  here.  If it is complicated, we can just say in text that there are other
-  sorts of attacks too, and that we deal with them, and give a forward
-  pointer.}
-Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
-  we are only aware of one, quite recent, attempt to do something better} work reasoned
-about a stack safety enforcement mechanism using capabilities\leo{cite};
-but even here, stack safety is defined with respect to an abstract
-machine that ``satisfies stack safety by construction,'' begging
-the question of what {\em is} the property that holds by construction?
-\apt{Suggest to move first sentence of next para up to this one.}
-
-\leo{StackTokens} provides some intuition towards answering that question,
-by informally describing two components that comprise stack safety:
-local state encapsulation and well-bracketed control flow. The first
-contribution of the present paper is to
-turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
-  claim this} security properties,
-that together form a formal definition of stack safety.
-\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
-rigor in it, even if some notions are implicit or not fully
-developed.}\bcp{Yes, and let's remember that they are highly likely to be
-reviewing this paper!  Let's be as generous as we can be without
-under-selling our delta.}
-
+ranging from leakage or corruption of private stack data to
+hijacking of the program's control flow. To prevent such attacks, many
+software and hardware techniques for stack safety have been proposed, % in the literature and implemented in production systems,
+including stack canaries~\citep{Cowan+98},
+shadow stacks~\citep{Dang+15,Shanbhogue+19},
+split stacks~\citep{Kuznetsov+14},
+bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
+capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from nick;there could be more}
+
+But what {\em is} this ``stack safety'' of which we speak? Precise formal
+definitions are lacking in the literature; instead, the notion of safety
+is usually defined negatively, by enumerating the bad things that a safe system
+prevents, such as corruption of return addresses, buffer overflows, or
+use of uninitialized variables.
+%with the notion usually being
+%defined negatively: through examples of stack-based exploits 
+%\rb{usually though not always; as noted below there are other
+%  approaches}\bcp{Yes, we have to be very careful about this}.
+%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ... 
+We are only aware of one, quite recent, attempt to frame a positive definition.
+\citet{Skorstengaard+19} define stack safety as
+the conjunction of two properties, local state encapsulation and well-bracketed control flow.
+They formalize these properties by defining an abstract capability-based machine   
+that ``satisfies [stack safety] by construction.'' We believe there is still room
+to craft a more abstract and less intensional characterization of
+stack safety, and that is the goal of the present paper. \apt{still needs tuning}
+
+Our approach is to give alternative formal
+definitions of local state encapsulation and well-bracketed control flow
+using the conceptual framework of language-based security.
+%
 The underlying notion behind local state encapsulation is the
@@
 
-While integrity and confidentiality are relatively standard security
-properties, the stack-safety setting raises an interesting additional
-challenge:
-the notion of who the attacker is shifts throught execution. Not only
+% BCP idea: contrast intrinsic vs. extrinsic property definition
+% Remember to talk about spatial vs temporal safety
+
+While integrity and confidentiality are standard security
+properties, applying them in the stack-safety setting raises a novel challenge:
+the notion of ``attacker'' changes during execution. Not only
 does a caller need to be protected from its callee, but the callee
-itself needs to be proteced from {\em its} callees.
-We address this challenge by varying the observation power\apt{is this
-  the right word? doesn't it conflict with our later use of ``observation''?}
-of the attacker at call points, and demonstrate that our policy is
-implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
-  the sentence are about two very different things.  The second deserves its
-  own paragraph, IMO.}\apt{+1}
-
-Similarly, the underlying notion behind well-bracketed control flow is
+itself needs to be protected from {\em its} callees.
+We address this challenge by adjusting the integrity and confidentiality
+of different stack frames at each call or return transition between functions.\apt{connection to ``attacker'' still vague.}
+
+The underlying notion behind well-bracketed control flow is
 the need to protect the call structure itself: a callee should not be
@@
 data and returning). Well-bracketed control flow establishes safety between
-functions: all code is uniquely owned by some function, and all jumps between functions
-are either calls that enter at a marked entry point or returns that enter
+functions: each instruction is uniquely owned by some function, and all jumps between functions
+are either calls that enter at a marked entry point or returns that re-enter
 just after their associated call in the standard nested call structure.
-
-%\leo{There needs to be some informal description
-%  of the wbcf part here. Sean?}
-
-In summary, we offer the following contributions\apt{is this up to date with current paper organization?}:
+\apt{clarify what we are contributing here.}
+
+To provide evidence that our definitions capture the inituitive notion of stack safety,
+we show that they rule out a wide \apt{really?} range of exploitable bad behaviors.\apt{``from the literature''??}
+To assess whether they are implementable, 
+we compare them against the proposed capability-based enforcement
+mechanism of \citet{Skorstengaard+19} \apt{we really should} and the existing tag-based enforcement mechanism
+of \citet{DBLP:conf/sp/RoesslerD18}.
+
+In summary, we offer the following contributions:
 \begin{itemize}
-\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
-  intuitive concepts of local state encapsulation and
-  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
-  property is strong enough to capture the folk intuition, we show
-  that the stack unsafe examples from the literature also fail our
-  property~(\Cref{sec:disc}).\apt{Something about relation to StkTokens formalization?}
-\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
-  realizible by
-  a previously proposed micropolicy enforcement
-  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
-    substantial body of code}. We also present a variation of the
-  property + policy that is lazy, easier/more efficient to implement,
-  and provably weaker (\Cref{sec:lazy}).
-\item A provably stronger (inductive? step-by-step/stepwise) variant
-  of this property tuned for testing/verification.  We test the whole
-  thing with a cool framework. \bcp{Which section?}
+\item
+  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
+  Abstract formal properties capturing the 
+  intuitive concepts of local state encapsulation and well-bracketed
+  control flow~(\Cref{sec:lse-and-wbcf}) in a very general setting. These properties
+  rule out common examples of stack memory abuse.
+\item
+  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\Cref{sec:enforcement}).
+  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
+  \apt{If we have no testing, do we even want to talk about the testing property?}
+\item
+  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy 
+  of \citet{DBLP:conf/sp/RoesslerD18}~(\Cref{sec:lazy}).  This policy does not enforce our property, but we give a
+  variant of the policy that does enforce a (provably) weaker variant of our property. \apt{Awkward, but the truth I fear.}
+
+  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
+%  realizible by
+%  a previously proposed micropolicy enforcement
+%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
+%    substantial body of code}. We also present a variation of the
+%  property + policy that is lazy, easier/more efficient to implement,
+%  and provably weaker (\Cref{sec:lazy}).
+%\item A provably stronger (inductive? step-by-step/stepwise) variant
+%  of this property tuned for testing/verification.  We test the whole
+%  thing with a cool framework. \bcp{Which section?}
 \end{itemize}
@@
 
-\section{Stack Safety and Unsafety by Example}
+\section{Running Example}
 \label{sec:running-example}
@@
 enforcement mechanisms preclude them later on in the paper.
-
-\subsection{Running Example}
 
@@
 \vspace*{0.2em}
-\bcp{Notation nit: in many assembly notations, square brackets mean
-  ``{\em indirect} through this location...''  And a very small nit: Change
-  $\rsp-2$ to $\rsp\mathord{-}2$}
 \[
@@
 
-We've assumed that callees always have less access than their callers, but
-in real systems things are more complicated (because of objects, static
-variables in C, software compartmentalization, coroutines, ...).  What we've
-defined is a very ``pure'' variant of stack safety.  In future work, we
-should go further. :-)
+  We have presented a pure notion of stack safety that omits many complicating
+  factors common to real system. In particular, our model as presented exploits
+  the simplifying assumption that privilege never increases during a function
+  call, and that therefore a contour computed once on the entry to a function
+  describes its privilege throughout. Our future plans are to extend the model
+  with common language features, some of which violate this assumption. Here we
+  describe the basic principles by which our model can be extended to some of
+  these cases.
+
+  \paragraph{Stack-derived Pointers}
+
+    Common programming idioms involve a caller passing a pointer to its local
+    data into a callee. Using standard notions of pointer provenance [cite],
+    we can extend our model to distinguish safe use of this idiom from true
+    stack safety violations. We require additional annotation to identify
+    which addresses within a function's stack frame correspond to distinct
+    objects. Then a valid pointer to an object is one derived through legal
+    arithmetic from its base address, and at each call point, an object's
+    addresses are marked low confidentiality and integrity if a valid pointer
+    exists in a register or in the transitive closure of the callee's
+    accessible memory. All pointers to an object cease to be valid when it is
+    deallocated along with its stack frame.
+
+  \paragraph{Adding Heap Safety}
+
+    Heap safety, like stack safety, may be framed as a form of noninterference
+    [cite Meaning of Memory Safety]. The above model describing safety of
+    shared stack-allocated objects extends naturally to heap safety,
+    complicated by the fact that these objects are not deallocated by returns,
+    and therefore a function's privilege may increase or decrease after its
+    entry point. For instance, if a callee allocates a heap object and returns
+    the pointer to its caller, that object's addresses become accessible in
+    contradiction to a contour computed at the caller's entry. So such an
+    extension must enforce trace confidentiality and integrity properties
+    separately on continuous segments of each call, with contours computed at
+    each crossing between caller and callee.
+
+  \paragraph{Non-stack control flow}
+
+    Control structures beyond well-bracketed calls and returns require
+    modifications to the model. Tail calls, for instance, reuse the caller's
+    stack frame for its callee, and every nested tail call returns
+    simultaneously to the top non-tail-calling function. Under normal stack
+    integrity, the first function in a chain of tail calls returns and violates
+    integrity, while subsequent tail calls never actually return at all.
+    Instead we need a variant that treats the entry of a tail call as the
+    return of its caller for purposes.
+
+    More complex is the addition of a coroutine model, in which multiple stacks
+    respect stack safety internally, but might also yield to one another.
+    Here we must distinguish three levels of integrity: accessible, in the same
+    stack, and in another stack. Stack integrity requires that same-stack data
+    be unchanged when a function returns, but other-stack data may change. A
+    similar {\em yield integrity} property requires that, from a yield out of
+    a coroutine to the next yield back into it, all accessible and same-stack
+    data are unchanged. Confidentiality properties are split similarly.
+
+\section{Discussion}
+
+\bcp{Make this a subsection?}
+\label{sec:disc}
 

### 2020-07-05 main.tex

--- 2020-07-05 main.tex
@@
 
+  We have presented a pure notion of stack safety that omits many complicating
+  factors common to real system. In particular, our model as presented exploits
+  the simplifying assumption that privilege never increases during a function
+  call, and that therefore a contour computed once on the entry to a function
+  describes its privilege throughout. Our future plans are to extend the model
+  with common language features, some of which violate this assumption. Here we
+  describe the basic principles by which our model can be extended to some of
+  these cases.
+
+  \paragraph{Stack-derived Pointers}
+
+    Common programming idioms involve a caller passing a pointer to its local
+    data into a callee. Using standard notions of pointer provenance [cite],
+    we can extend our model to distinguish safe use of this idiom from true
+    stack safety violations. We require additional annotation to identify
+    which addresses within a function's stack frame correspond to distinct
+    objects. Then a valid pointer to an object is one derived through legal
+    arithmetic from its base address, and at each call point, an object's
+    addresses are marked low confidentiality and integrity if a valid pointer
+    exists in a register or in the transitive closure of the callee's
+    accessible memory. All pointers to an object cease to be valid when it is
+    deallocated along with its stack frame.
+
+  \paragraph{Adding Heap Safety}
+
+    Heap safety, like stack safety, may be framed as a form of noninterference
+    [cite Meaning of Memory Safety]. The above model describing safety of
+    shared stack-allocated objects extends naturally to heap safety,
+    complicated by the fact that these objects are not deallocated by returns,
+    and therefore a function's privilege may increase or decrease after its
+    entry point. For instance, if a callee allocates a heap object and returns
+    the pointer to its caller, that object's addresses become accessible in
+    contradiction to a contour computed at the caller's entry. So such an
+    extension must enforce trace confidentiality and integrity properties
+    separately on continuous segments of each call, with contours computed at
+    each crossing between caller and callee.
+
+  \paragraph{Non-stack control flow}
+
+    Control structures beyond well-bracketed calls and returns require
+    modifications to the model. Tail calls, for instance, reuse the caller's
+    stack frame for its callee, and every nested tail call returns
+    simultaneously to the top non-tail-calling function. Under normal stack
+    integrity, the first function in a chain of tail calls returns and violates
+    integrity, while subsequent tail calls never actually return at all.
+    Instead we need a variant that treats the entry of a tail call as the
+    return of its caller for purposes.
+
+    More complex is the addition of a coroutine model, in which multiple stacks
+    respect stack safety internally, but might also yield to one another.
+    Here we must distinguish three levels of integrity: accessible, in the same
+    stack, and in another stack. Stack integrity requires that same-stack data
+    be unchanged when a function returns, but other-stack data may change. A
+    similar {\em yield integrity} property requires that, from a yield out of
+    a coroutine to the next yield back into it, all accessible and same-stack
+    data are unchanged. Confidentiality properties are split similarly.
+
 \section{Discussion}
-
-\bcp{Make this a subsection?}
-\label{sec:disc}
-
-  \paragraph{What about tail recursion and other non-stack-like control flow}
-
-\bcp{And this and the next belong in Future Work...}
-
-  \paragraph{Adding Heap Safety}
-
-    While these ideas are here presented in a stack-centric style, they could
-    be extended to heap safety in a system extended with a heap. Heap safety,
-    like stack safety, may be framed as a form of noninterference
-    [cite Meaning of Memory Safety]. For each certain component marked as
-    containing a valid pointer to a heap object, we can propagate that
-    validity when the pointer moves or is used in pointer arithmetic. Heap
-    objects to which a valid pointer is in a register are reachable, and
-    treated as low confidentiality and low integrity.
-
-We've assumed that callees always have less access than their callers, but
-in real systems things are more complicated (because of objects, static
-variables in C, software compartmentalization, coroutines, ...).  What we've
-defined is a very ``pure'' variant of stack safety.  In future work, we
-should go further. :-)
 

### 2020-07-05 main.tex
Maybe call things MO and MPO

@@
     a bit more complicated to state but still retains the pragmatic
-    protections of the original.}\apt{To clarify: Nick's LT&C policy uses
+    protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
     depth-based tags, which do not enforce even our lazy policy. So we need to
-    introduce a version of LT&C that uses unique activation tags.}
+    introduce a version of LT\&C that uses unique activation tags.}
 
@@
 A trace over a type $A$ is a nonempty, finite or infinite sequence of
-$A$s. We will mostly be concerned with sequences of machines states,
-denoted $\machT$, and with sequences of triples of (machine states,
-policy states, and observations), usually denoted $\MPT$.  \rb{The way
-  the definitions are set up, one may expect MPO, but\ldots} We
-will require four generic operations on traces.
+$A$s. We will mostly be concerned with sequences of pairs of machine states
+and observations, called machine traces and denoted $\machOT \in \MOTRACE$,
+and with sequences of triples of a machine state, a policy state, and an
+observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces. 
+\rb{The way the definitions are set up, one may expect MPO, but\ldots}
 
@@
 %
-\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
-            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
+\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machOT\)}
+            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machOT\)}%
 %
 \noindent
-Since the step function is total, if \(\mach \hookrightarrow \machT\),
-then \(\machT\) must be infinite.
+Since the step function is total, if \(\mach \hookrightarrow \machOT\),
+then \(\machOT\) must be infinite.
 
@@
 
-In a similar fashion, we define \(\hookrightarrow_P\) to relate a
-machine-policy state pair to the trace of triples of machine states,
-policy states, and observations induced by by \(\stepstoobs{}_P\).
-Unlike the $\MOTRACE$ above, this trace can be finite as the
-policy can cause a fault:
-%
-\[(m,p) \hookrightarrow_P MP : \MACHS \times \POLS \rightarrow \MPOTRACE \]
+In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
+to relate a machine-policy state pair to the MP-trace
+induced by by \(\stepstoobs{}_P\). Unlike the $\MOTRACE$ above, this trace can
+be finite as the policy can cause a fault, in which case the final state is
+a \(\tau\):
 %
@@
               {\(\pol_0 \rightharpoonup \pol_1\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machOT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
 We will also use \(\pi_m\) to project the trace of machine states
 out of a trace, \(\pi_p\) to project the trace of policy states, and
-\(\pi_o\) to project the observations.  \bcp{I'm getting a little lost in
+\(\pi_o\) to project the observations, with \(\pi_m\) and \(\pi_o\) overloaded
+for machine traces and MP-traces.
+
+\bcp{I'm getting a little lost in
   all these.  E.g., do these traces include the final state of a finite
@@
   observation is a $\tau$ based on the rule in the inlined figure, so
-  it is essentially ignored. Point that out somehow?}
+  it is essentially ignored. Point that out somehow?}\sna{It is now pointed
+  out, but it also reminds me as we go back to prefix-up-to that the final
+  observation of a prefix needs to be tau...}
 
@@
 Let \(f\) be a predicate on machine states. Then we define \(\mach
-\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
+\hookrightarrow \machOT | f\) (read ``\(\machOT\) is the prefix of
 running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I
@@
 
-\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
-            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT')
-              = \machT\)}
-            {\(\mach \hookrightarrow \machT | f\)}
+\judgmenttwo{\(\mach \hookrightarrow \machOT'\)}
+            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machOT')
+              = \machOT\)}
+            {\(\mach \hookrightarrow \machOT | f\)}
 \noindent
@@
         & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
-        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' \Rightarrow \\
-        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
+        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machOT' \Rightarrow \\
+        & \pi_o(\MPT') \lesssim_O \pi_o(\machOT')
       \end{split}\]
@@
 
-      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_e\) and
+      A pair of traces \(\MPT\) and \(\machOT\) starting from \(\mach_e\) and
       \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\),
@@
       The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
-      \pi_o(\machT)\).
+      \pi_o(\machOT)\).
 
       If \(\MPT\) is finite and ends with the state
-      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
-      \(\mach_r'\) be the final state of \(\machT\). Then the induced traces
+      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machOT\) is finite. Let
+      \(\mach_r'\) be the final state of \(\machOT\). Then the induced traces
       \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
       \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
-      \hookrightarrow \machT'\) are observationally prefixed,
-      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).
+      \hookrightarrow \machOT'\) are observationally prefixed,
+      \(\pi_o(\MPT') \lesssim \pi_o(\machOT')\).
 

### 2020-07-05 main.tex
Fixed some spots where we weren't using the macros (so MPO and MO wasn't happening)

@@
 \label{def:trace-integrity}
-Let $\contour$ be a contour and $MP$ be a (potentially infinite)
-$\MPOTRACE$. We will say $MP$ satisfies {\em trace integrity} with
+Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
+$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
 respect to $\contour$ if its final state (if any) agrees with its
@@
 \[
-\mathit{int} ~ \contour ~ MP \triangleq
+\mathit{int} ~ \contour ~ \MPT \triangleq
 \left \{ \begin{array}{ll}
-  \mathit{True}, & \text{if $MP$ is infinite}\\
+  \mathit{True}, & \text{if $\MPT$ is infinite}\\
   \forall \component. ~ \contour(\component) = \HIGHSEC \rightarrow m_0(\component) = m_f(\component), & \text{otherwise} \\
@@
 $\codemap$ if, for every initial machine-policy state pair $(\mach_0,
-\pol_0)$ and its induced trace $MP$, and every $n$-transition
+\pol_0)$ and its induced trace $\MPT$, and every $n$-transition
 $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee $f$ in $MP$,
-the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$ until the first
+the prefix $\MPT_{pre}$ of $\MPT$ from $(\mach_e, \pol_e)$ until the first
 return corresponding to $\mach_c$ (if any) satisfies trace integrity
@@
 %
-$$\forall MP_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
-MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow
-\mathit{int}~\mathit{Cof}(\mach_c, f, n)~MP_{pre}$$
+$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
+\MPT_{pre} ~ | ~ \ret{\mach_c} \rightarrow
+\mathit{int}~\mathit{Cof}(\mach_c, f, n)~\MPT_{pre}$$
 
@@
 $\callmap$ and $\codemap$ if, for every initial machine-policy state
-pair $(\mach_0, \pol_0)$ and its induced trace $MP$, and every
+pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
 $n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
-$f$ in $MP$, the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$
+$f$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from $(\mach_e, \pol_e)$
 until the first return corresponding to $\mach_c$ (if any) satisfies
@@
 of every variation $\mach_e'$ with the contour at the call:
-$$\forall MP_{pre} ~ M_{pre}. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
+$$\forall \MPT_{pre} ~ \machOT_{pre}. ~
+\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
 $$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
-\mach_e' \hookrightarrow M_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ MP_{pre} ~ M_{pre}$$
+\mach_e' \hookrightarrow \machOT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ \MPT_{pre} ~ \machOT_{pre}$$
 

### 2020-07-05 main.tex
Description of the comservative policy

@@
 \newcommand*{\rout}{\textsc{out}}
+
+\newcommand*{\tagInstr}{\textsc{instr}}
+\newcommand*{\tagCall}{\textsc{call}}
+\newcommand*{\tagHa}{\textsc{h1}}
+\newcommand*{\tagHb}{\textsc{h2}}
+\newcommand*{\tagRa}{\textsc{r1}}
+\newcommand*{\tagRb}{\textsc{r2}}
+\newcommand*{\tagRc}{\textsc{r3}}
+\newcommand*{\tagNoDepth}{\textsc{unused}}
+\newcommand*{\tagStackDepth}[1]{\textsc{stack} ~ #1}
+\newcommand*{\tagPCDepth}[1]{\textsc{pc} ~ #1}
+\newcommand*{\tagSP}{\textsc{sp}}
 
@@
 \paragraph*{A conservative policy}
+%
+The first micropolicy we present for eager stack safety is very similar to the
+Depth Isolation policy presented by Roessler and DeHon \rb{cite}. Their policies
+tag the stack memory with ownership information associated to each stack frame
+(and to each separate object inside that frames, a more finely grained access
+control that we do not need to consider), and registers holding stack pointers
+with matching access permission information. Beyond those capabilities, the
+compiler generates designated sequences of instructions with special tags that
+grant them permission to manage the stack.
+
+Our micropolicy works as follows. We assume for ease of exposition that no
+arguments are passed on the stack \rb{Come back to this later?}. Initially, the
+entire stack is tagged as unused with $\tagNoDepth$ \rb{actually, the ``initial
+  frame'' should be tagged with its proper depth}, and the {\PCname} carries in
+its tag at all times the depth of the current activation, starting with
+$\tagPCDepth{0}$. The current activation depth needs to change exactly at those
+points in the program where the contour of the stack safety property changes:
+
+\begin{itemize}
+
+\item From caller to callee, when the machine executes an instruction tagged as
+  belonging to the call map with $\tagCall$. At this point, the {\PCname} depth
+  is incremented to, say $\tagPCDepth{n}$; on function entry, the locations of
+  the new frame is marked $\tagStackDepth{n}$.
+
+\item From callee back to caller, when execution reaches the instruction
+  immediately following the original call instruction and the $\SP$ is the same
+  as at the time of that call. At this point, the {\PCname} is decremented;
+  previously, the locations of the old frame are restored to $\tagNoDepth$.
+
+\end{itemize}
+
+To enforce this discipline, we define blessed header and return instruction
+sequences. These are specific sequences of tagged instructions that are intended
+to appear at the beginning and end of each function:
+
+\[
+  \begin{array}{c|c|c}
+    {\sw ~ \rsp ~ \rra ~ 1} & {\tagHa, \tagInstr} & \text{Store RA in stack} \\
+    {\addi ~ \rsp ~ \rsp ~ 2} & {\tagHb, \tagInstr} & \text{Increment SP by frame size}
+  \end{array}
+  \]
+
+\[
+  \begin{array}{c|c|c}
+    {\lw ~ \rra ~ \rsp ~ -1} & {\tagRa, \tagInstr} & \text{Load RA} \\
+    {\addi ~ \rsp ~ \rsp ~ -2} & {\tagRb, \tagInstr} & \text{Decrement SP by frame size} \\
+    {\jalr ~ \rra ~ \rra} & {\tagRc, \tagInstr} & \text{Jump back}
+  \end{array}
+\]
+
+\rb{Positioning of the above sequences, relation to running example. The tagging
+  of stack frames would be considered part of these sequences.}
+
+These sequences also enforce the $\SP$ discipline. When stack frames are of
+fixed size, it suffices to increment $\SP$ and give it with a special tag in the
+header sequence, which is later checked after decrementing $\SP$ during the exit
+sequence to authorize the return. The well-formedness of the blessed sequences
+can be checked statically. Only programs whose sequences are well-formed are
+protected by the micropolicy, although programs are still protected (by
+fail-stopping) even if the intended sequences are missing. By tracking the
+execution of the blessed sequences in the {\PCname} tag state, the micropolicy
+ensures that the entry sequence is executed correctly a declared call.
+
+% If anything overwrites the SP in between, the exit sequence will fail, which is safe.
+
+During normal execution, the {\PCname} uses the current activation depth in its
+tag $\tagPCDepth{n}$ to gate load and store operations, which are allowed to
+operate only on the corresponding stack frame, tagged $\tagStackDepth{n}$.
+\rb{Note that we have also considered and implemented more flexible schemes
+  based on allowing operations above the current activation depth. How do we put
+  these together?}
+\rb{Do we want to discuss variations involving frame pointers or more implicit
+  return addresses?}
+\rb{Where to talk about the role of the compiler?}
+
+\paragraph*{Testing the conservative policy}
 
@@
 
-\paragraph*{Testing the conservative policy}
+\rb{The Static Authorities differs from the Depth Isolation policy in that frame
+  ids aren't true activation ids, but unique ids associated to all activations
+  of the same function. Discuss unique activation ids and compare with depth
+  later, possibly when discussing lazy policies, as the difference becomes
+  tangible there.}
 

### 2020-07-05 main.tex
Remove the last element of prefix-up-to

--- 2020-07-05 main.tex
@@
 
-\paragraph*{Prefix-Up-To}
-The ``prefix-up-to'' operation \(\PUT(\Trace,f)\) \bcp{Wouldn't
-  ``While'' be more mnemonic?}  takes a trace \(T\) and a predicate on
-elements \(f\) and gives the prefix of \(T\) ending with the first
-element on which \(f\) holds, coinductively:
+\paragraph*{Until}
+The ``Until'' operation \(\PUT(\Trace,f)\) \bcp{Wouldn't
+``While'' be more mnemonic?} \sna{Appealing, but it's more of a
+``while-not''? Trying out ``until''. In any case... moving away
+from doing joins means this needs to get some work to eliminate
+the trailing observation. Here's an attempt.} takes a trace
+\(\Trace : \mathit{trace} \TRACEELEMS \times \OBSS\) and a predicate
+on elements \(f \subset \TRACEELEMS\) and gives the prefix of
+\(\Trace\) ending with the first element on which \(f\) holds.
+Subtly, because the observation associated with a trace represents its
+next step, when we take a prefix from a trace containing observations
+we must replace the final observation with a \(\tau\).
 
@@
     \judgment{\(f\ \traceelem\)}
-         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
+             {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
   \end{minipage}
@@
   \judgment{}
-         {\(\PUT(f,\traceelem) = \traceelem\)}
+           {\(\PUT(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
 %
@@
   \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
-            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
+              {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
   \end{minipage}
@@
 \sna{Similarity seems less loaded?}
-\[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
+\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
 

### 2020-07-06 main.tex
Work on 3

--- 2020-07-05 main.tex
@@
 %\memory{8}{red}[{\makebox[0pt]{Callee}}]%
-%\memory{4}{gray}%
+%\memory{4}{\unusedc}%
 %~$\cdots$\\
@@
 
+\newcommand{\instrc}{magenta}
+\newcommand{\outc}{white}
+\newcommand{\fprivc}{cyan}
+\newcommand{\fcallc}{green}
+\newcommand{\gprivc}{yellow}
+\newcommand{\gcallc}{orange}
+\newcommand{\hprivc}{red}
+\newcommand{\unusedc}{gray}
+
 \begin{figure}
@@
 \MemoryLabel{25em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
 \hspace*{3pt}
 $\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{10}{gray}[{\makebox[0pt]{Uninitialized}}]
+\memory{1}{\outc}[{\makebox[0pt]{Output}}]%
+\memory{10}{\unusedc}[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
@@
 \MemoryLabel{43.5em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
 \hspace*{3pt}
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
-\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{6}{gray}%[{\makebox[0pt]{Uninitialized}}]
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
@@
 \MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
 \hspace*{3pt}
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
-\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{orange}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{2}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
@@
 \MemoryLabel{65.5em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
 \hspace*{3pt}
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
-\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
-\memory{1}{red}%[{\makebox[0pt]{{\texttt h} ra}}]
-\memory{1}{gray}%[{\makebox[0pt]{Uninitialized}}]
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{orange}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{1}{\hprivc}%[{\makebox[0pt]{{\texttt h} ra}}]
+\memory{1}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
@@
 \MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\outc}[{\makebox[0pt]{Instructions}}]%
 \hspace*{3pt}
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
-\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
-\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
-\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{orange}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{2}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
@@
 \MemoryLabel{31.5em}{2em}{\SP}
-\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
 \hspace*{3pt}
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
-\memory{8}{gray}%[{\makebox[0pt]{Uninitialized}}]
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{8}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
 ~$\cdots$
@@
 example of the previous section for concreteness, and then,
-in~\Cref{sec:micropolicies}, extend this model with a dynamic policy
-enforcement mechanism in~\Cref{sec:micropolicies} based on
+in~\cref{sec:micropolicies}, extend this model with a dynamic policy
+enforcement mechanism in~\cref{sec:micropolicies} based on
 \bcp{NickPaper,AndMaybeSomeOthers}.
@@
 
-In our running example in Figure~\Cref{fig:running-example-b} we can
+In our running example in Figure~\cref{fig:running-example-b} we can
 see a depiction of such a trace: starting from an initial machine
@@
 enforce that the contents of $\HIGHSEC$ locations don't influence the
-observable behavior. However, when dealing with nested calls and
-returns, stack safety needs to enforce integrity and confidentiality
-for {\em every} function call in a trace; that is, we must treat the
-security level of data dynamically.
+observable behavior of the machine. However, when dealing with nested
+calls and returns, stack safety needs to enforce integrity and
+confidentiality for {\em every} function call in a trace; that is, we
+must treat the security level of data dynamically.
 %
@@
 
-When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) for
-some \(n\) and \(\mach_c \stepstoobs{\obs} \mach_e\), we can construct
-a contour based on the bounds of the callee's stack frame.
+When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) and
+\(\mach_c \stepstoobs{\obs} \mach_e\) for some $n$, $\obs$, and
+$\mach_e$, we can construct a contour based on the bounds of the
+callee's stack frame.
 %
@@
 stack base to stack pointer, with the exception of the $n$ locations
-corresponding to the arguments of a call. Private components are
-marked high confidentiality and high integrity: the callee should be
-able to neither read, nor write them.
+corresponding to the arguments and result locations of the
+call. Private components are marked high confidentiality and high
+integrity: the callee should be able to neither read, nor write them.
 %
@@
 are marked high confidentiality and low integrity.
+%
+The same is true for the designated output location: it can be
+freely written into but should not be read.
 %
@@
 
- \[\mathit{Cof}(\mach,\funid,n)(\component) =
+ \[\mathit{Cof}(\mach,n)(\component) =
   \begin{cases}
@@
     (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
-                         \text{ and } \component \geq \mach(\SP) \\
+                          \text{ and } \component \geq \mach(\SP)
+                          \text{ or if } \component = \rout\\
     (\LOWSEC,\LOWINT) & \text{else} \\
@@
 
-  \leo{Snapshot state at the point of the call from TOP to F, collor
-    the different things}
+Going back to our running example, recall the state of the machine
+at the call from {\tt main} to {\tt g}:
+\vspace*{0.2em}
+% At Call to G
+% Memory Layout
+\begin{center}
+\MemoryLabel{4em}{2em}{0}
+\MemoryLabel{21em}{2em}{99}
+\MemoryLabel{43.5em}{2em}{\SP}
+\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{6}{\unusedc}[{\makebox[0pt]{Uninitialized}}]
+~$\cdots$
+\MemoryLabel{-46.5em}{0.75em}{42}
+\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-38.5em}{0.75em}{0}
+\MemoryLabel{-34.5em}{0.75em}{0}
+\\
+\end{center}
+
+If $\mach$ is the state at this call, the corresponding contour would
+be:
+\[
+\mathit{Cof}(\mach,2) = \left \{
+\begin{array}{l}
+  0\ldots98\} \mapsto (\LOWSEC, \HIGHINT)\\
+  99 \mapsto (\HIGHSEC, \LOWINT)\\
+  100, 101 \mapsto (\HIGHSEC, \HIGHINT)\\
+  102, 103 \mapsto (\LOWSEC, \LOWINT)\\
+  104\ldots \mapsto (\HIGHSEC, \LOWINT)\\
+  r \mapsto (\LOWSEC, \LOWINT)\\
+\end{array}
+\right .
+\]
 
@@
 \leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
+
+\begin{figure}
+  \integrityexample
+  \caption{Failing Trace Integrity}
+  \label{fig:intex}
+\end{figure}
+
+\begin{figure}
+  \integritylazyexample
+  \caption{Rolling back from Figure \ref{fig:intex}\leo{Where is this referenced?
+    Shouldn't this be in the lazy part?}}
+  \label{fig:intlex}
+\end{figure}
 
@@
 Trace integrity characterizes the trace corresponding to a single
-callee. We now lift this definition to an integrity property for a
-complete execution trace. We do this by requiring that trace integrity
-holds for every transition between a caller and a callee $f$ for
-the entirety of the trace of $f$.
-
-\definition{Transition}
-Given a trace $MP$, a call map $\callmap$ and a code map $\codemap$,
-an $n$-{\em transition} to a callee $f$ is a pair of machine-policy
-state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ in the trace
-such that:
+callee. Going back to our running example, such a trace would be, for
+instance, the entire trace of {\tt h} from instruction 40 to
+instruction 47. This trace is repeated in \Cref{fig:intex}, and
+recolored: \leo{Explain colors here - not sure I agree with the
+  current choice}
+\leo{Old relevant text:
+For an example see \cref{fig:intex}, in which {\tt h} writes its return
+value to {\tt g}'s return location the stack rather than its own. That address
+is high integrity (red), and so writing to it causes the property to fail.
+}
+
+We now lift this definition to an integrity property for a complete
+execution trace. We do this by requiring that trace integrity holds
+for every transition between a caller and a callee for the entirety of
+the trace of the callee.
+
+\definition{Transition} Given a trace $MP$, a call map $\callmap$ and
+a code map $\codemap$, an $n$-{\em transition} a pair of
+machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
+in the trace such that:
 \begin{itemize}
@@
 
-For an example see \cref{fig:intex}, in which {\tt h} writes its return
-value to {\tt g}'s return location the stack rather than its own. That address
-is high integrity (red), and so writing to it causes the property to fail.
-
-\definition{Stack Integrity}
-Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
-system satisfies {\em stack integrity} with respect to $\callmap$ and
-$\codemap$ if, for every initial machine-policy state pair $(\mach_0,
-\pol_0)$ and its induced trace $\MPT$, and every $n$-transition
-$(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee $f$ in $MP$,
-the prefix $\MPT_{pre}$ of $\MPT$ from $(\mach_e, \pol_e)$ until the first
-return corresponding to $\mach_c$ (if any) satisfies trace integrity
-with respect to the contour at the call:
+\definition{Stack Integrity} Let $\callmap$ be a call map and
+$\codemap$ a code map. We say that a system satisfies {\em stack
+  integrity} with respect to $\callmap$ and $\codemap$ if, for every
+initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
+trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
+$(\mach_e, \pol_e)$ in $MP$, the prefix $\MPT_{pre}$ of $\MPT$ from
+$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
+(if any) satisfies trace integrity with respect to the contour at the
+call:
 %
@@
 \MPT_{pre} ~ | ~ \ret{\mach_c} \rightarrow
-\mathit{int}~\mathit{Cof}(\mach_c, f, n)~\MPT_{pre}$$
+\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$
 
@@
 and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
-if it started from a variant initial machine state. An example in
-
-\definition{Trace Confidentiality}
-\leo{Add info for ``valid return'':}
-We will say two traces $MP$ and $M$ satisfy {\em trace
-  confidentiality} with respect to $\contour$ if:
+if it started from a variant initial machine state.
+%
+As an example of variations, consider the top of \Cref{fig:confex1}
+which shows a snapshot of the state as in the running example (top
+memory layout) and a variation of it (bottom). \leo{Finalize
+  contents, describe}
+
+\begin{figure}
+  \confidentialityduringexample
+  \caption{Failing Trace Confidentiality by Varying Observations}
+  \label{fig:confex1}
+\end{figure}
+
+\begin{figure}
+  \confidentialityendexample
+  \caption{Failing Trace Confidentiality at End State}
+  \label{fig:confex2}
+\end{figure}
+
+\definition{Trace Confidentiality} We will say two traces $MP$ and $M$
+satisfy {\em trace confidentiality} with respect to $\contour$ if:
 \begin{itemize}
 \item If $MP$ is finite and its final state $\mach_r$ is a valid
-  return, then $M$ is also finite with some final state $\mach_r'$ and
+  return\leo{Is this ill-scoped?}, then $M$ is also finite with some final state $\mach_r'$ and
   any changes from the initial states have changed in the same way:
@@
 
-\Cref{fig:confex2} shows the initial states \(\mach_e\), left,
-and \(\mach_e'\), right. Low confidentiality data in green is shared
-between them, while high confidentiality data in red and yellow may differ.
-After executing the code shown, the final states below have written different
-values to address 42, represented by the black and white components.
-It is allowed for address 40 to differ between the states due to the initial
-variation, but address 42 changed differently between the traces, representing
-a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
+The rest of \Cref{fig:confex1} shows an example that fails \leo{number
+  items?  observations of trace?}  \Cref{fig:confex2} shows the
+initial states \(\mach_e\), left, and \(\mach_e'\), right. Low
+confidentiality data in \fcallc is shared between them, while high
+confidentiality data in red and \gprivc may differ.  After executing
+the code shown, the final states below have written different values
+to address 42, represented by the black and white components.  It is
+allowed for address 40 to differ between the states due to the initial
+variation, but address 42 changed differently between the traces,
+representing a leak of data. %Meanwhile \ref{fig:lazyex2}, while high
+confidentiality data
 %is read, it is overwritten before return, and the write at step 3 always writes
-%the same value. Yellow and red locations differ between variations only because
+%the same value. \Gprivc and red locations differ between variations only because
 %of the initial variation, and the rest are identical. So this example passes.
-
-\begin{figure}
-  \integrityexample
-  \caption{Failing Trace Integrity}
-  \label{fig:intex}
-\end{figure}
-
-\begin{figure}
-  \integritylazyexample
-  \caption{Rolling back from Figure \ref{fig:intex}}
-  \label{fig:intlex}
-\end{figure}
-
-\begin{figure}
-  \confidentialityduringexample
-  \caption{Failing Trace Confidentiality by Varying Observations}
-  \label{fig:confex1}
-\end{figure}
-
-\begin{figure}
-  \confidentialityendexample
-  \caption{Failing Trace Confidentiality at End State}
-  \label{fig:confex2}
-\end{figure}
-
 \Cref{fig:confex2} gives an example of a trace that fails trace
@@
 Since {\tt x}, at address 100, is high confidentiality (red) the variant
-may have a different value (yellow). So both \(r_4\) and address 107
+may have a different value (\gprivc). So both \(r_4\) and address 107
 differ in the final state between the original trace and the variant one,

### 2020-07-06 main.tex
Finished examples for lazy properties, suggested slightly different coq for prefix

--- 2020-07-05 main.tex
@@
       {Store $r_4$ as the return value}
-  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
+  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 27}
       {Load return address}
@@
       {Restore the stack pointer of {\tt g}}
-  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 31}
+  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 27}
       {Return to {\tt g}}
@@
 
+\sna{Variations here collides with state variations. I propose substitutions.}
+
 \leo{TODO: Make attacker model more explicit} The execution trace
@@
 
-A trace over a type $A$ is a nonempty, finite or infinite sequence of
-$A$s. We will mostly be concerned with sequences of pairs of machine states
+A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
+$\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
 and observations, called machine traces and denoted $\machOT \in \MOTRACE$,
@@
 
-The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
-relates a machine state with the trace of machine states paired with
-observations produced by repeated application of step:
-%
-\[\mach \hookrightarrow \machOT : \MACHS \rightarrow \MOTRACE\]%
-%
+The ``trace-of'' operator, written \(\mach \hookrightarrow \machOT\),
+coinductively relates a machine state with the trace of machine states
+paired with observations produced by repeated application of step:
+
 \judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machOT\)}
@@
 a \(\tau\):
-%
+
+\begin{minipage}{.4\textwidth}
 \judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
          {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
-%
+\end{minipage}
+\begin{minipage}{.55\textwidth}
 \judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
@@
               {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
+\end{minipage}
+
 We will also use \(\pi_m\) to project the trace of machine states
@@
 \judgmenttwo{\(\mach \hookrightarrow \machOT'\)}
-            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machOT')
+            {\(\PUT(f,\machOT')
               = \machOT\)}
@@
 \judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
-            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT') = \MPT\)}
+            {\(\PUT(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
          {\(\mpstatename \hookrightarrow \MPT | f\)}
@@
 
+\begin{figure}
+  \integrityexample
+  \caption{Failing Trace Integrity}
+  \label{fig:intex}
+\end{figure}
+
+For an example see \cref{fig:intex}, representing the unsafe substitution
+\[
+  44: ~ \sw ~ \rsp ~ r_4 ~ -5
+\]
+This instruction is in {\tt h}, and instead of storing data to the return,
+it overwrites {\tt g}'s variable, {\tt z}. At the top of the figure high
+integrity memory is in red, while low integrity is in green; at the end of the
+trace address 103 should match the initial state, and does not. So trace
+integrity is violated.
+
 Trace integrity characterizes the trace corresponding to a single
@@
 \definition{Transition}
-Given a trace $MP$, a call map $\callmap$ and a code map $\codemap$,
+Given a trace $\MPT$, a call map $\callmap$ and a code map $\codemap$,
 an $n$-{\em transition} to a callee $f$ is a pair of machine-policy
@@
 \end{itemize}
-
-For an example see \cref{fig:intex}, in which {\tt h} writes its return
-value to {\tt g}'s return location the stack rather than its own. That address
-is high integrity (red), and so writing to it causes the property to fail.
 
@@
 
-\Cref{fig:confex2} shows the initial states \(\mach_e\), left,
-and \(\mach_e'\), right. Low confidentiality data in green is shared
-between them, while high confidentiality data in red and yellow may differ.
-After executing the code shown, the final states below have written different
-values to address 42, represented by the black and white components.
-It is allowed for address 40 to differ between the states due to the initial
-variation, but address 42 changed differently between the traces, representing
-a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
+For example, consider the unsafe substitution:
+\[
+42 : ~ \sw ~ r_4 ~ \rsp 100
+\]
+Instead of reading the proper argument, {\tt h} reads directly from the
+secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
+shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
+in red and low confidentiality in green. Below it, the variant state
+\(\mach_e'\) agrees on low confidentiality memory, but the rest might have
+any value, in yellow. After the last step, the return states \(\mach_r\) and
+\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in pink, and while 109
+is consistent, 108 will not hold 43 in all variants, so the trace does not obey
+trace confidentiality.
+
+Meanwhile, the following unsafe substitution outputs the secret
+directly, violating the internal observation clause of confidentiality:
+\[
+30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+\]
+%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
+%and \(\mach_e'\), below. Low confidentiality data in green is shared
+%between them, while high confidentiality data in red and yellow may differ.
+%After executing the code shown, the final states below have written different
+%values to address 42, represented by the black and white components.
+%It is allowed for address 40 to differ between the states due to the initial
+%variation, but address 42 changed differently between the traces, representing
+%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
 %is read, it is overwritten before return, and the write at step 3 always writes
@@
 
-\begin{figure}
-  \integrityexample
-  \caption{Failing Trace Integrity}
-  \label{fig:intex}
-\end{figure}
-
-\begin{figure}
-  \integritylazyexample
-  \caption{Rolling back from Figure \ref{fig:intex}}
-  \label{fig:intlex}
-\end{figure}
-
-\begin{figure}
-  \confidentialityduringexample
-  \caption{Failing Trace Confidentiality by Varying Observations}
-  \label{fig:confex1}
-\end{figure}
+%\begin{figure}
+%  \confidentialityduringexample
+%  \caption{Failing Trace Confidentiality by Varying Observations}
+%  \label{fig:confex1}
+%\end{figure}
 
@@
   \confidentialityendexample
-  \caption{Failing Trace Confidentiality at End State}
-  \label{fig:confex2}
+  \caption{Failing Trace Confidentiality}
+  \label{fig:confex}
 \end{figure}
 
-\Cref{fig:confex2} gives an example of a trace that fails trace
-confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
-Since {\tt x}, at address 100, is high confidentiality (red) the variant
-may have a different value (yellow). So both \(r_4\) and address 107
-differ in the final state between the original trace and the variant one,
-failing the ending condition of trace confidentiality.
+%\Cref{fig:confex2} gives an example of a trace that fails trace
+%confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
+%Since {\tt x}, at address 100, is high confidentiality (red) the variant
+%may have a different value (yellow). So both \(r_4\) and address 107
+%differ in the final state between the original trace and the variant one,
+%failing the ending condition of trace confidentiality.
 
@@
 
+    \begin{figure}
+      \integritylazyexample
+      \caption{Rolling back from Figure \ref{fig:intex}}
+      \label{fig:intlex}
+    \end{figure}
+
+    For out example, we return to the violation of trace integrity in
+    \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
+    Whether observable integrity holds depends on the continuing trace, shown
+    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
+    \(\mach_r\), with high integrity components in red and low integrity in
+    green. Next is the rolled-back state, \(\mach_r' =
+    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
+    components are restored to their value as of \(\mach_e\). Specifically,
+    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
+    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
+    it prints 0. So observable trace integrity does not hold.
+
+    However, if {\tt g} had an additional instruction to overwrite {\tt z}
+    before printing, observable trace integrity would hold.
+
     \definition{Observable Stack Integrity}
@@
 
-      In \cref{fig:lazyex1}, the final states of the original and variant
-      differ, because a written value (in black or white) depended on the value
-      of a high confidentiality location. So the resulting trace fails trace
-      confidentiality. For observable trace confidentiality, we ask whether the
-      differing values at address 42 cause any observable distinctions in
-      future behavior. To do so, we must undo the variation of formerly high
-      confidentiality locations, so that those in yellow match those in red,
-      while preserving the updated values where data may have been leaked.
+      To return to the example in \cref{fig:confex}, the varied final state
+      \(\mach_r'\) might do anything depending on the values of the variation,
+      so we cannot immediately compare traces from \(\mach_r\) and
+      \(\mach_r'\). We must first restore the varied values to their original
+      state, while preserving differences that have occurred during
+      execution.
+
+%      In \cref{fig:lazyex1}, the final states of the original and variant
+%      differ, because a written value (in black or white) depended on the value
+%      of a high confidentiality location. So the resulting trace fails trace
+%      confidentiality. For observable trace confidentiality, we ask whether the
+%      differing values at address 42 cause any observable distinctions in
+%      future behavior. To do so, we must undo the variation of formerly high
+%      confidentiality locations, so that those in yellow match those in red,
+%      while preserving the updated values where data may have been leaked.
 
@@
       We write this \(\text{\sc ObsConf}_\contour ~ \MPT\).
+
+      \begin{figure}
+        \confidentialitylazyexample
+        \caption{Continuing from \cref{fig:confex}}
+        \label{fig:conflex}
+      \end{figure}
+
+      In our example, \cref{fig:conflex} shows how execution continues from
+      \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
+      \mach_e,\mach_e',\mach_r,\mach_r'\). Addresses 108 and 109, in pink, are
+      not restored; every address that was varied and not changed in one of the
+      executions is. Since 108 contains a value that originates in the
+      variation, values that depend on it might be anything; after returning
+      from {\tt g}, {\tt main} prints 95 in the trace from \(\mach_r\) and
+      some undetermined value in the trace from \(\mach_r''\). So observable
+      trace confidentiality does not hold.
 

### 2020-07-06 main.tex

--- 2020-07-06 main.tex
@@
 
-\definition{Transition} Given a trace $MP$, a call map $\callmap$ and
+\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
 a code map $\codemap$, an $n$-{\em transition} a pair of
@@
 trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
-$(\mach_e, \pol_e)$ in $MP$, the prefix $\MPT_{pre}$ of $\MPT$ from
+$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
 $(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
@@
 
-The rest of \Cref{fig:confex1} shows an example that fails \leo{number
-  items?  observations of trace?}  \Cref{fig:confex2} shows the
-initial states \(\mach_e\), left, and \(\mach_e'\), right. Low
-confidentiality data in \fcallc is shared between them, while high
-confidentiality data in red and \gprivc may differ.  After executing
-the code shown, the final states below have written different values
-to address 42, represented by the black and white components.  It is
-allowed for address 40 to differ between the states due to the initial
-variation, but address 42 changed differently between the traces,
-representing a leak of data. %Meanwhile \ref{fig:lazyex2}, while high
-confidentiality data
+For example, consider the unsafe substitution:
+\[
+42 : ~ \sw ~ r_4 ~ \rsp 100
+\]
+Instead of reading the proper argument, {\tt h} reads directly from the
+secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
+shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
+in red and low confidentiality in green. Below it, the variant state
+\(\mach_e'\) agrees on low confidentiality memory, but the rest might have
+any value, in yellow. After the last step, the return states \(\mach_r\) and
+\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in pink, and while 109
+is consistent, 108 will not hold 43 in all variants, so the trace does not obey
+trace confidentiality.
+
+Meanwhile, the following unsafe substitution outputs the secret
+directly, violating the internal observation clause of confidentiality:
+\[
+30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+\]
+%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
+%and \(\mach_e'\), below. Low confidentiality data in green is shared
+%between them, while high confidentiality data in red and yellow may differ.
+%After executing the code shown, the final states below have written different
+%values to address 42, represented by the black and white components.
+%It is allowed for address 40 to differ between the states due to the initial
+%variation, but address 42 changed differently between the traces, representing
+%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
 %is read, it is overwritten before return, and the write at step 3 always writes
@@
 %of the initial variation, and the rest are identical. So this example passes.
-\Cref{fig:confex2} gives an example of a trace that fails trace
-confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
-Since {\tt x}, at address 100, is high confidentiality (red) the variant
-may have a different value (\gprivc). So both \(r_4\) and address 107
-differ in the final state between the original trace and the variant one,
-failing the ending condition of trace confidentiality.
+
+%\begin{figure}
+%  \confidentialityduringexample
+%  \caption{Failing Trace Confidentiality by Varying Observations}
+%  \label{fig:confex1}
+%\end{figure}
+
+\begin{figure}
+  \confidentialityendexample
+  \caption{Failing Trace Confidentiality}
+  \label{fig:confex}
+\end{figure}
+
+%\Cref{fig:confex2} gives an example of a trace that fails trace
+%confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
+%Since {\tt x}, at address 100, is high confidentiality (red) the variant
+%may have a different value (yellow). So both \(r_4\) and address 107
+%differ in the final state between the original trace and the variant one,
+%failing the ending condition of trace confidentiality.
 

### 2020-07-06 main.tex
Description of eager testing property

@@
   in metavariables?}
+\rb{Should we have notation for traces somewhere here?}
 
@@
 
+%% Because the access rights of callers subsume those of callees, it is essential
+%% that the micropolicy change tags when the property thinks that a call occurs,
+%% but it is safe for the micropolicy to fail to reset to the caller's tag when the
+%% property thinks a return occurs. (Dually, it is safe for the micropolicy to
+%% change tags even without executing a call point, but not for it to reset tags to
+%% the caller when the property thinks execution is still in the caller).
+
+%% Because the property does not identify return sites, we have no way to require
+%% that ``intended returns'' are tagged correspondingly. Some unintended results
+%% might also be so tagged, but as long as the jump has the effect of a legal
+%% return, it is OK.
+
 \paragraph*{Testing the conservative policy}
-
-This conservative policy enforces our stack safety policy. Moreover, it enforces
-a provably stronger \emph{testing property}.
-
-The properties we give are indeed enforceable.
+%
+This conservative policy is can be shown to satisfy an eager \emph{testing
+property}, which enforces lockstep and performs integrity and confidentiality
+tests for a stack of running variants at each step.
+
+\definition{Eager integrity test}
+  Given two machines $\mach$ and $\mach'$ and a contour $\contour$, these
+  satisfy the \emph{eager integrity test} if their high-integrity components
+  match:
+  \[
+    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
+    \forall \component . ~
+    \contour(\component) = \HIGHINT \rightarrow
+    \mach(\contour) = \mach'(\contour)
+    \]
+
+\definition{Eager confidentiality test}
+  Given two machine states $\mach_1$ and $\mach_2$ with stepping states and
+  observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
+  \stepstoobs{\obs_2} \mach_2'$ these satisfy the \emph{eager confidentiality
+  test} if their observations and changed components coincide:
+  \[
+    \mathit{conf}_T ~ \mach_1 ~ \obs_1 ~ \mach_1' ~ \mach_2 ~ \obs_2 ~ \mach_2' ~ \triangleq
+    \obs_1 = \obs_2 \wedge \forall \component . ~
+   (     \mach_1(\component) \neq \mach_1'(\component)
+    \vee \mach_2(\component) \neq \mach_2'(\component))
+    \rightarrow \mach_1' = \mach_2'
+  \]
+
+\rb{Fix notations, including for properties below.}
+\newcommand{\vse}{\mathit{vse}}
+\newcommand{\VSES}{\mathit{VSE}}
+\newcommand{\vs}{\mathit{vs}}
+\newcommand{\VSS}{\mathit{VS}}
+
+\definition{Variant stacks}
+  The testing property keeps track of execution information on a stack whose
+  elements record an initial machine, an initial and current variant of the
+  machine, a contour and a return map:
+  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS (\MACHS \rightarrow \mathcal{P})$$
+  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
+  $$\vs \in \VSS$$
+  Given the current machine state $\mach$, a stack $\vs$ is \emph{well-formed}
+  with respect to it if:
+  \begin{itemize}
+  \item The current variant is indeed a variant of the current state.
+  \item The stack is nonempty and the bottom return map is the false property.
+  \item The current state is in the trace of every initial machine.
+  \item The variant state is in the trace of its own initial variant.
+  \item The return function is in the return map if not the last element. \rb{TODO}
+  \item All contours have public registers.
+  \end{itemize}
+
+\definition{Eager testing property}
+  Let $\callmap$ a call map and $\codemap$ a code map. Given a state
+  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
+  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
+  testing property} at a given call depth if:
+
+  \begin{itemize}
+
+  \item If the machine fail-stops,
+    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
+    the property is safisfied for the singleton trace $(\mach, \pol)$ at any
+    depth up to the size of $\vs$.
+
+  \item If the top variant of the stack (and only the top variant) returns,
+    the property is safisfied for the singleton trace $(\mach, \pol)$ at depth
+    equal to the size of the stack. \rb{Avoid confusion with return below}
+
+  \item If the machine takes a step
+    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    that is neither a call nor a return, all
+    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
+    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
+    (these tests are required also for calls and returns, and omitted in those cases),
+    and the stepped $(\mach', \pol')$ and $\vs'$ satisfy the eager testing property
+    for a given depth $d$ \rb{omitting stack size condition?} and MP-trace $\MPT$,
+    then the property is also satisfied for the extension of the trace \rb{notation}
+    at depth $d$.
+
+  \item If the machine takes a call step
+    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    the contour $\contour$ is updated based on the top of the current machine.
+    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    (and in addition to the general step conditions),
+    the eager integrity test must hold for $\contour$ at the call,
+    the eager confidentiality test must hold for all variants at the call,
+    including the new variant that is added to the top of the stack,
+    and the property recurses on an arbitrary variant at the new contour
+    at depth $d$ with trace $\MPT$.
+
+  \item If the machine takes a return step at the top of the variant stack
+    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    if (in addition to the general step conditions)
+    the property recurses at depth $d$ with trace $\MPT$
+    after removing the returning top variant from the stack.
+
+  \item If the machine takes a return step that is also a call, the last two
+    cases are combined.
+
+  \end{itemize}
+
+By using property-based testing techniques, we can determine with a high degree
+of confidence that an implementation of the conservative micropolicy satisfies
+this eager testing property. Although we have not proved that the conservative
+policy guarantees the stack safety property, the testing property is provably
+stronger than the stack safety property \rb{theorem?} and, by this implication,
+we can conclude that the conservative policy also enforces stack safety.
+
+% The properties we give are indeed enforceable.
+
+\rb{What is the strongest property enforced by the conservative policy?}
+
+\paragraph*{End-to-end stack safety property}
+
+\paragraph*{Eager vs lazy testing}
+
+\rb{This part should probably be somewhere in the lazy section.}
+%
+In practice, this \emph{eager policy} is too slow to be of practical interest:
+what we really want in practice is a \emph{lazy policy} that can be efficiently
+implemented.
 
@@
 
-What is the strongest property enforced by the conservative policy?
-%
-\rb{Not too much focus on this}
-
-\rb{We have not proved that this policy guarantees our property, but we have
-tested it against a stronger property.}
+We can also test the lazy policy, even if less efficiently, and it should
+enforce an inductive property, which may be more amenable to proving.
+%
+\rb{Would need careful phrasing.}
+%
+\rb{Implied by the property we started with?}
+
+\rb{Explain and compare what each policy enforces.}
 
@@
   tangible there.}
-
-\paragraph*{End-to-end stack safety property}
-
-\paragraph*{Eager vs lazy testing}
-
-\rb{This part should probably be somewhere in the lazy section.}
-%
-In practice, this \emph{eager policy} is too slow to be of practical interest:
-what we really want in practice is a \emph{lazy policy} that can be efficiently
-implemented.
-
-We can also test the lazy policy, even if less efficiently, and it should
-enforce an inductive property, which may be more amenable to proving.
-%
-\rb{Would need careful phrasing.}
-%
-\rb{Implied by the property we started with?}
-
-\rb{Explain and compare what each policy enforces.}
 

### 2020-07-06 main.tex
Font formatting in running example
High-level overview of StkTokens

@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
 \memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
 \memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
 \memory{2}{orange}[{\makebox[0pt]{{\tt h}'s call frame }}]%
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
 \memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{orange}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
+\memory{2}{orange}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
 \memory{2}{\gprivc}[{\makebox[0pt]{{\tt h}'s call frame }}]%
@@
 \memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
+\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
 \memory{8}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
@@
 
+StkTokens \citep{Skorstengaard+19} protects the stack by enforcing local state
+encapsulation and well-bracketed control flow. It does so by defining a calling
+convention that makes use of linear capabilities for stack and return pointers
+operating on a single shared stack, and by requiring that protected components
+avoid compromising their own security by following certain simple rules.
+Informally, local state encapsulation is defined by restricting accesses to the
+range of memory allocated to the current stack frame, and well-bracketed control
+flow by only allowing returns from the topmost frame to the immediately adjacent
+frame below. Formally, these properties are built into a capability machine
+defined by an overlay semantics, which is proven fully abstract with respect to
+a capability machine that replaces those abstract operations with their calling
+convention.
+
+\rb{There are quite a few parallels at various levels, although the definitions
+  are overall more implicit and not given formally. A limiting factor is the
+  lack of linear capabilities in existing models like CHERI---although these
+  could be implemented by a micropolicy.}
+
 \section{Future Work}

### 2020-07-06 main.tex
Comment squashing

@@
 
-\bcp{ Comments on the example:
-\begin{itemize}
-\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
-capitalization of register names (sp vs SP), etc. Maybe typeset instructions
-in small caps?  Maybe typeset everything that is part of a concrete machine
-in small caps??
-%
-\item The colors are helpful in making the big page of instruction traces easier
-on the eye; maybe they would be even more helpful if
-corresponding colors appeared elsewhere in the example (on the instructions
-themselves, e.g.).  But we need to be careful to keep it readable -- maybe
-they should be background colors, like you've done, rather than text colors.
-\item Asking readers to wade through all this is asking quite a bit!  We
-need to make sure we are going to get sufficient value out of the example
-later (I think we can, but we should really convince ourselves) to make it
-worthwhile.
-\item Create a new column that just literally pronounces each instruction in
-English, so people don't have to read RISC-V opcodes.
-\item As discussed, we should move this to before section 2, we should
-extend it with a bit more at the end so we can talk about the lazy policy
-with the same example, we should add
-in some bad variants (doing it here replaces the ``Examples'' from the
-Discussion section), and we should consider discussing the attacker model
-here (we need to discuss it someplace in any case!)
-\end{itemize}
-}
-
 \subsection{Attacker Model + Unsafe Variations}
@@
 
+\leo{Other choices for observations? Where should that be discussed?}
+\bcp{But in this paragraph we are describing the particular
+choise that we make for our running example, no?  But we certainly want to
+discuss this fully someplace!}
+\leo{Here: Many observation choices. Next: Simplify:}
 Mapping a conventional architecture to this abstract interface is
@@
 = \tau$.
-
-\leo{Other choices for observations? Where should that be discussed?}
-\bcp{But in this paragraph we are describing the particular
-choise that we make for our running example, no?  But we certainly want to
-discuss this fully someplace!}
 
@@
 
-\bcp{I'm getting a little lost in
-  all these.  E.g., do these traces include the final state of a finite
-  execution?  If so, what observation is it paired with?}\leo{The final
-  observation is a $\tau$ based on the rule in the inlined figure, so
-  it is essentially ignored. Point that out somehow?}\sna{It is now pointed
-  out, but it also reminds me as we go back to prefix-up-to that the final
-  observation of a prefix needs to be tau...}
-
 \paragraph*{Until}
-The ``Until'' operation \(\PUT(\Trace,f)\) \bcp{Wouldn't
-``While'' be more mnemonic?} \sna{Appealing, but it's more of a
-``while-not''? Trying out ``until''. In any case... moving away
-from doing joins means this needs to get some work to eliminate
-the trailing observation. Here's an attempt.} takes a trace
+The ``Until'' operation \(\PUT(\Trace,f)\)
+%\bcp{Wouldn't
+%``While'' be more mnemonic?} \sna{Appealing, but it's more of a
+%``while-not''? Trying out ``until''. In any case... moving away
+%from doing joins means this needs to get some work to eliminate
+%the trailing observation. Here's an attempt.}
+takes a trace
 \(\Trace : \mathit{trace} \TRACEELEMS \times \OBSS\) and a predicate
@@
 \hookrightarrow \machOT | f\) (read ``\(\machOT\) is the prefix of
-running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I
-  think.}  \leo{There were inconsistencies which i fixed. I wanted the
-  metavariables defined to be consistent and the ``temporary''
-  variable which is the full trace to be primed. Should I swap those?}
+running \(\mach\) up to \(f\)''):
 

### 2020-07-06 main.tex
query in 5
another query in 5

@@
 \definition{Eager integrity test}
-  Given two machines $\mach$ and $\mach'$ and a contour $\contour$, these
+  Given two machines \apt{machine states?} $\mach$ and $\mach'$ and a contour $\contour$, these
   satisfy the \emph{eager integrity test} if their high-integrity components
@@
     \]
+\apt{Shouldn't conclusion be $m(k) = m'(k)$ ? }
 
@@
   \]
-
+\apt{scoping of $\forall$ is ambiguous. Anyhow, shouldn't this be
+$
+\forall \component . ~
+   (     \mach_1(\component) \neq \mach_1'(\component)
+    \vee \mach_2(\component) \neq \mach_2'(\component))
+    \rightarrow \mach_1'(\component) = \mach_2'(\component)
+ $ ?}
 \rb{Fix notations, including for properties below.}

### 2020-07-06 main.tex
Some notes on notation in lazy properties

@@
 \end{minipage}
+\rb{Shouldn't the conclusion of the first judgment use $\hookrightarrow_P$?
+  And in the second judgment, wouldn't $MO$ be $MPO$, and the third premise $\hookrightarrow_P$?}
 
@@
 
+    \rb{Suggestion: name this ``observable \emph{trace? something else?}
+      integrity'' to avoid confusion with observable stack integrity (in fact,
+      it is called exactly this when discussing the example below). The rollback
+      may also be a definition of its own to match the granularity of
+      definitions in previous sections.}
+
     \definition{Observable Integrity}
@@
       account for premature halts due to policy violations.)
+      \rb{Here, should the second premise use $\hookrightarrow_P$?}
       %
@@
     \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
-    components are restored to their value as of \(\mach_e\). Specifically,
+    components are restored to their values as of \(\mach_e\). Specifically,
     address 103 is restored to 0. When instruction 31 prints {\tt z} in the
@@
 
-      A system enjoys {\em observable integrity} with regard to \(\callmap\)
+      A system enjoys {\em observable \rb{stack?} integrity} with regard to \(\callmap\)
       and  \(\codemap\) if for any initial state \(\mach_0\) and each
@@
       \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\) \rb{$\hookrightarrow_P$?},
       then \(\text{\sc ObsInt}_\contour ~ \MPT\).
@@
       execution.
+      \rb{To me, this reads like a prologue to the definitions properly
+        speaking---maybe the definition header above should be a paragraph and
+        a definition header appear next? I would also split it into two
+        definitions.}
 
@@
 
-      A pair of traces \(\MPT\) and \(\machOT\) starting from \(\mach_e\) and
-      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\),
+      A pair of traces \(\MPT\) and \(\machOT\) starting from \(\mach_e\) \rb{with what $\pol_e$?} and
+      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\) \rb{$\mach_e'$?},
       enjoy observable trace confidentiality with respect to \(\contour\) if
@@
       \(\mach_r'\) be the final state of \(\machOT\). Then the induced traces
-      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
+      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) \rb{$\hookrightarrow_P$?} and
       \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
@@
       We write this \(\text{\sc ObsConf}_\contour ~ \MPT\).
+      \rb{What about $\machOT$? Ditto for use in the definition of observable stack confidentiality below.}
 
@@
       \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
-      \mach_e,\mach_e',\mach_r,\mach_r'\). Addresses 108 and 109, in pink, are
+      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109, in pink, are
       not restored; every address that was varied and not changed in one of the
@@
       A system enjoys {\em observable stack safety} with regard to a call map
-      and a code map if it enjoys both observable integrity and observable
-      confidentiality with regard to them.
+      and a code map if it enjoys both observable \rb{stack?} integrity and observable
+      \rb{stack?} confidentiality with regard to them.
 

### 2020-07-06 main.tex
WBCF cleanup

--- 2020-07-06 main.tex
@@
 
-Our integrity and confidentiality properties concern themselves with
-data, and may hold on a wide range of programs that violate expected
-control flow.  But well-bracketed control flow prevents this style of
-attack by guaranteeing that every jump between functions is either a
-call or a return, and each return jumps to the instruction after a
+Both the integrity and confidentiality components of local state
+encapsulation concern themselves with the accessibility of data, and
+may hold on a wide range of programs that violate expected control
+flow. For example, if {\tt g} managed to jump in the middle of the
+code of {\tt h} with its call, confidentiality and integrity would
+still be satisfied. 
+%
+However, that is still unwanted behavior: for example, a caller (or a
+callee) should not be able to bypass a password check by jumping (or
+returning) in the middle of a block of code - even if such an execution
+would ostensibly satisfy local state encapsulation.
+%
+Stack safety is traditionally associated with preventing such unwanted
+flows: every jump between functions should either be a call or a
+return, and each return should to the instruction after its
 corresponding call.
 
-\leo{Also dumping in relevant wbcf discussion part. TODO: merge}
-    Some forms of security rely on protected control flow. In a classic
-    example, a system call that launches a missile might be gated behind code
-    that checks a password. A piece of malicious code can jump into the middle
-    of the sequence and execute the system call without the appropriate password,
-    and still obey integrity if the malicious code never returns to its caller.
-    Integrity here protects data, not privileges. But such a jump does violate
-    eager stack confidentiality if the attacker jumps beyond its own function's
-    instructions without performing a designated call or returning to an
-    appropriate site.
-
-    A jump to a function's own code does not violate either, even if done using
-    nefarious means such as return-oriented programming (ROP). Our model would
-    permit such a jump within the function, so privilege is not elevated by
-    permitting ROP in this case.
-
-    But, our policies do enforce orthogonal control flow properties that
-    protect from calls that enter outside of valid entry points and from all
-    invalid returns, even if they would not escalate privilege. Most notably
-    {\em well-bracketed control flow}, below.
+To formally capture this aspect of stack safety, we introduce three
+security properties that rely on the underlying knowledge of the
+program structure (code maps, function maps, return maps, and entry
+maps): {\em control separation} ensures the validity of transition
+sources between functions, {\em entry integrity} ensures the validity
+of call targets, and {\tt return integrity} ensures the validity of
+return targets.
 
 \definition{Control Separation}
-A system enjoys {\em control separation} with regard to a code map
-\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\) if
-for any initial state \(\mach_0\) and any adjacent pair of states
-\(\mach_1\) and \(\mach_2\) in the induced trace from
-\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
-\(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2)\), then either
-\(\callmap(\mach_1(\PCname)) = n\) for some \(n\), or
-\(\retmap(\mach_1)\).
+
+A system enjoys {\em control separation} with regard to a function map
+\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
+if for any initial state \((\mach_0, \pol_0)\) and any adjacent pair
+of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
+trace from \(\mpstate{\mach_0}{\pol_0})\) whose current instruction
+belongs to different functions according to the function map,
+either $\mach_1$ is a valid call or a valid return:
+
+\[ \codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname)) \rightarrow
+\exists n. ~ \callmap(\mach_1(\PCname)) = n \vee
+\retmap(\mach_1) \]
 
 \definition{Entry Integrity}
+
 A system enjoys {\em entry integrity} with regard to a call map
 \(\callmap\) and entry map \(\entmap\) if for any initial state
-\(\mach_0\) and any adjacent pair of states \(\mach_c\) and
-\(\mach_e\) in the induced trace from
-\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
-\(\callmap(\mach_c(\PCname)) = n\) then \(\entmap(\mach_e)\).
+\((\mach_0, \pol_0)\) and any n-transition \((\mach_c,\pol_c)\) and
+\((\mach_e, \pol_e)\) in the induced trace from
+\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
+\[\callmap(\mach_c(\PCname)) = n \rightarrow \entmap(\mach_e)\]
 
 \definition{Return Integrity}
+
+\newcommand*{\Unmatched}{\mathit{Unmatched}}
+
 Return integrity captures the expected behavior of marked returns with
-regard to their call points. We begin by defining a coinductive
-relation \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1
-\MPT_2\). Given a call map \(\callmap\) and return map \(\retmap\),
-\(\MPT\) is related to \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\)
-whose head is the first unmatched return in \(\MPT\). So if \(\MPT\)
-starts with a return, it is related to itself:
+regard to their call points. Given a call map $\callmap$ and a return
+map $\retmap$, we begin by defining a coinductive relation
+$\Unmatched$ over two $\MPOTRACE$ $\MPT$ and $\MPT'$, if $\MPT'$ is a
+suffix of $\MPT$ whose head is the first unmatched return in $\MPT$.
+
+Formally:
+\begin{itemize}
+\item A trace that starts with a return is related to itself:
 
@@
               {\(\retmap(\mach(\PCname))\)}
-      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\
-        \mpstate{\mach}{\pol}\)}
+              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
   \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
               {\(\retmap(\mach(\PCname))\)}
-      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\
-              \mpstate{\mach}{\pol}\MPT\)}
-
-If \(\MPT\) starts with a non-call, non-return state, its unmatched
-return must come later:
+              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}
+
+\item If a trace \(\MPT\) starts with a non-call, non-return state,
+  its unmatched return must come later:
 
@@
               \hspace{.5cm} \neg \retmap(\mach(\PCname))
-
-              \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'\)}
-           {\(\mathit{Unmatched}\ \callmap\ \retmap\
-            \mpstate{\mach}{\pol}\MPT\ \MPT'\)}
-
-And most intricately, if \(\MPT\) starts with a call, we must find the
-return that matches it, which will be the unmatched return once we
-step past the call. Then the first unmatched return beyond it is the
-true unmatched return.
-
-    \judgment{\(\callmap(\mach(\PCname)) = n
-              \hspace{.5cm} \neg \retmap(\mach(\PCname))\)
-
-              \(\MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}}\)
-
-              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}}\)
-
-              \(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\)
-
-              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}}\)}
-            {\(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}\)}
-
+              \Unmatched ~ \MPT ~ \MPT'\)}
+           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}
+
+\item Finally, if a trace starts with the call, then we must discard the first
+  unmatched return to get to the real one:
+
+  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
+             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
+             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
+              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
+            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
+\end{itemize}
 A system enjoys {\em return integrity} with regard to a call map
-\(\callmap\) and return map \(\retmap\) if for any initial state
-\(\mach_0\) and any call state and policy state \(\mach_c\) and
-\(\pol_c\) in the induced trace from
-\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if \(\mpstate{\mach_c}{\pol_c}
-\hookrightarrow_P \MPT\) and \(\mach_r\) is the first unmatched
-return, with \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT \ \MPT'\)
-and \(\mach_r\) the first machine state of \(\MPT'\), then
-\(\ret{\mach_c}\ \mach_r\).
+\(\callmap\) and return map \(\retmap\) if for any initial state pair
+\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
+corresponding to a call in its induced trace, if
+\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
+is the first unmatched return such that for some $\pol_r$
+\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
+then $\mach_r$ corresponds to a return state for $\mach_c$:
+\[\ret{\mach_c}\ \mach_r\]
 

### 2020-07-07 main.tex
queries in 2

@@
         {Zero out result\apt{why?}}
-    \row{\jal ~ \rra ~ 20}{??}{\PCname \leftarrow 20; ~ \rra \leftarrow \thepcctr}
+    \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
         {Make the call}
@@
   ~\\
+  \apt{How's that description for last line?}
+  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
+    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
+  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}
 % At Call to G
@@
 \cref{fig:running-trace-b}: instructions for {\tt main} begin at
-0, the dedicated output is empty, while the (uninitialized) stack
+0, the dedicated output is empty\apt{what does that mean? why not just uninitialized?}, while the (uninitialized) stack
 begins at 100. Directly below the layout, we summarize the parts of

### 2020-07-07 main.tex
Improved examples (haven't figured out cross-hatching though) and updated lazy definitions

--- 2020-07-06 main.tex
@@
 $\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
-and observations, called machine traces and denoted $\machOT \in \MOTRACE$,
+and observations, called machine traces and denoted $\machT \in \MOTRACE$,
 and with sequences of triples of a machine state, a policy state, and an
@@
 
-The ``trace-of'' operator, written \(\mach \hookrightarrow \machOT\),
+The ``trace-of'' operator, written \(\mach \hookrightarrow \machT\),
 coinductively relates a machine state with the trace of machine states
@@
 
-\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machOT\)}
-            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machOT\)}%
+\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
+            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
 %
 \noindent
-Since the step function is total, if \(\mach \hookrightarrow \machOT\),
-then \(\machOT\) must be infinite.
+Since the step function is total, if \(\mach \hookrightarrow \machT\),
+then \(\machT\) must be infinite.
 
@@
               {\(\pol_0 \rightharpoonup \pol_1\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machOT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
 \end{minipage}
@@
 Let \(f\) be a predicate on machine states. Then we define \(\mach
-\hookrightarrow \machOT | f\) (read ``\(\machOT\) is the prefix of
+\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
 running \(\mach\) up to \(f\)''):
 
-\judgmenttwo{\(\mach \hookrightarrow \machOT'\)}
-            {\(\PUT(f,\machOT')
-              = \machOT\)}
-            {\(\mach \hookrightarrow \machOT | f\)}
+\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
+            {\(\PUT(f,\machT')
+              = \machT\)}
+            {\(\mach \hookrightarrow \machT | f\)}
 \noindent
@@
 
-\begin{figure}
-  \integrityexample
-  \caption{Failing Trace Integrity}
-  \label{fig:intex}
-\end{figure}
-
-\begin{figure}
-  \integritylazyexample
-  \caption{Rolling back from Figure \ref{fig:intex}\leo{Where is this referenced?
-    Shouldn't this be in the lazy part?}}
-  \label{fig:intlex}
-\end{figure}
-
 Intuitively, the caller expects to find their private data untouched
@@
 
-\begin{figure}
-  \confidentialityduringexample
-  \caption{Failing Trace Confidentiality by Varying Observations}
-  \label{fig:confex1}
-\end{figure}
-
-\begin{figure}
-  \confidentialityendexample
-  \caption{Failing Trace Confidentiality at End State}
-  \label{fig:confex2}
-\end{figure}
-
 \definition{Trace Confidentiality} We will say two traces $MP$ and $M$
@@
 \begin{itemize}
-\item If $MP$ is finite and its final state $\mach_r$ is a valid
-  return\leo{Is this ill-scoped?}, then $M$ is also finite with some final state $\mach_r'$ and
+\item If $\MPT$ is finite and its final state $\mach_r$ is a valid
+  return\leo{Is this ill-scoped?}, then $\machT$ is also finite with some final state $\mach_r'$ and
   any changes from the initial states have changed in the same way:
@@
   In addittion their observations must be equivalent:
-  $$\pi_o(MP) \eqsim \pi_o(M)$$
-\item If $MP$ is infinite, then $M$ is also infinite and their
+  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
+\item If $\MPT$ is infinite, then $\machT$ is also infinite and their
   observations are equivalent:
-  $$\pi_o(MP) \eqsim \pi_o(M)$$
-\item If the $MP$ ends prematurely with a policy fault, then
-  we require that its observations are a prefix of those of $M$:
-  $$\pi_o(MP) \lesssim \pi_o(M)$$
+  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
+\item If the $\MPT$ ends prematurely with a policy fault, then
+  we require that its observations are a prefix of those of $\machT$:
+  $$\pi_o(\MPT) \lesssim \pi_o(\machT)$$
 \end{itemize}
@@
 
-%\begin{figure}
-%  \confidentialityduringexample
-%  \caption{Failing Trace Confidentiality by Varying Observations}
-%  \label{fig:confex1}
-%\end{figure}
-
 \begin{figure}
@@
 of every variation $\mach_e'$ with the contour at the call:
-$$\forall \MPT_{pre} ~ \machOT_{pre}. ~
+$$\forall \MPT_{pre} ~ \machT_{pre}. ~
 \mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
 $$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
-\mach_e' \hookrightarrow \machOT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ \MPT_{pre} ~ \machOT_{pre}$$
+\mach_e' \hookrightarrow \machT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ \MPT_{pre} ~ \machT_{pre}$$
 
@@
 
-Both the integrity and confidentiality components of local state
-encapsulation concern themselves with the accessibility of data, and
-may hold on a wide range of programs that violate expected control
-flow. For example, if {\tt g} managed to jump in the middle of the
-code of {\tt h} with its call, confidentiality and integrity would
-still be satisfied. 
-%
-However, that is still unwanted behavior: for example, a caller (or a
-callee) should not be able to bypass a password check by jumping (or
-returning) in the middle of a block of code - even if such an execution
-would ostensibly satisfy local state encapsulation.
-%
-Stack safety is traditionally associated with preventing such unwanted
-flows: every jump between functions should either be a call or a
-return, and each return should to the instruction after its
+Our integrity and confidentiality properties concern themselves with
+data, and may hold on a wide range of programs that violate expected
+control flow.  But well-bracketed control flow prevents this style of
+attack by guaranteeing that every jump between functions is either a
+call or a return, and each return jumps to the instruction after a
 corresponding call.
 
-To formally capture this aspect of stack safety, we introduce three
-security properties that rely on the underlying knowledge of the
-program structure (code maps, function maps, return maps, and entry
-maps): {\em control separation} ensures the validity of transition
-sources between functions, {\em entry integrity} ensures the validity
-of call targets, and {\tt return integrity} ensures the validity of
-return targets.
+\leo{Also dumping in relevant wbcf discussion part. TODO: merge}
+    Some forms of security rely on protected control flow. In a classic
+    example, a system call that launches a missile might be gated behind code
+    that checks a password. A piece of malicious code can jump into the middle
+    of the sequence and execute the system call without the appropriate password,
+    and still obey integrity if the malicious code never returns to its caller.
+    Integrity here protects data, not privileges. But such a jump does violate
+    eager stack confidentiality if the attacker jumps beyond its own function's
+    instructions without performing a designated call or returning to an
+    appropriate site.
+
+    A jump to a function's own code does not violate either, even if done using
+    nefarious means such as return-oriented programming (ROP). Our model would
+    permit such a jump within the function, so privilege is not elevated by
+    permitting ROP in this case.
+
+    But, our policies do enforce orthogonal control flow properties that
+    protect from calls that enter outside of valid entry points and from all
+    invalid returns, even if they would not escalate privilege. Most notably
+    {\em well-bracketed control flow}, below.
 
 \definition{Control Separation}
-
-A system enjoys {\em control separation} with regard to a function map
-\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
-if for any initial state \((\mach_0, \pol_0)\) and any adjacent pair
-of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
-trace from \(\mpstate{\mach_0}{\pol_0})\) whose current instruction
-belongs to different functions according to the function map,
-either $\mach_1$ is a valid call or a valid return:
-
-\[ \codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname)) \rightarrow
-\exists n. ~ \callmap(\mach_1(\PCname)) = n \vee
-\retmap(\mach_1) \]
+A system enjoys {\em control separation} with regard to a code map
+\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\) if
+for any initial state \(\mach_0\) and any adjacent pair of states
+\(\mach_1\) and \(\mach_2\) in the induced trace from
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
+\(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2)\), then either
+\(\callmap(\mach_1(\PCname)) = n\) for some \(n\), or
+\(\retmap(\mach_1)\).
 
 \definition{Entry Integrity}
-
 A system enjoys {\em entry integrity} with regard to a call map
 \(\callmap\) and entry map \(\entmap\) if for any initial state
-\((\mach_0, \pol_0)\) and any n-transition \((\mach_c,\pol_c)\) and
-\((\mach_e, \pol_e)\) in the induced trace from
-\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
-\[\callmap(\mach_c(\PCname)) = n \rightarrow \entmap(\mach_e)\]
+\(\mach_0\) and any adjacent pair of states \(\mach_c\) and
+\(\mach_e\) in the induced trace from
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
+\(\callmap(\mach_c(\PCname)) = n\) then \(\entmap(\mach_e)\).
 
 \definition{Return Integrity}
-
-\newcommand*{\Unmatched}{\mathit{Unmatched}}
-
 Return integrity captures the expected behavior of marked returns with
-regard to their call points. Given a call map $\callmap$ and a return
-map $\retmap$, we begin by defining a coinductive relation
-$\Unmatched$ over two $\MPOTRACE$ $\MPT$ and $\MPT'$, if $\MPT'$ is a
-suffix of $\MPT$ whose head is the first unmatched return in $\MPT$.
-
-Formally:
-\begin{itemize}
-\item A trace that starts with a return is related to itself:
+regard to their call points. We begin by defining a coinductive
+relation \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1
+\MPT_2\). Given a call map \(\callmap\) and return map \(\retmap\),
+\(\MPT\) is related to \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\)
+whose head is the first unmatched return in \(\MPT\). So if \(\MPT\)
+starts with a return, it is related to itself:
 
@@
            {\(\retmap(\mach(\PCname))\)}
-              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
+      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\
+        \mpstate{\mach}{\pol}\)}
   \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
               {\(\retmap(\mach(\PCname))\)}
-              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}
-
-\item If a trace \(\MPT\) starts with a non-call, non-return state,
-  its unmatched return must come later:
+      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\
+              \mpstate{\mach}{\pol}\MPT\)}
+
+If \(\MPT\) starts with a non-call, non-return state, its unmatched
+return must come later:
 
@@
               \hspace{.5cm} \neg \retmap(\mach(\PCname))
-              \Unmatched ~ \MPT ~ \MPT'\)}
-           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}
-
-\item Finally, if a trace starts with the call, then we must discard the first
-  unmatched return to get to the real one:
-
-  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
-             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
-             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
-              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
-            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
-\end{itemize}
+
+              \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'\)}
+           {\(\mathit{Unmatched}\ \callmap\ \retmap\
+            \mpstate{\mach}{\pol}\MPT\ \MPT'\)}
+
+And most intricately, if \(\MPT\) starts with a call, we must find the
+return that matches it, which will be the unmatched return once we
+step past the call. Then the first unmatched return beyond it is the
+true unmatched return.
+
+    \judgment{\(\callmap(\mach(\PCname)) = n
+              \hspace{.5cm} \neg \retmap(\mach(\PCname))\)
+
+              \(\MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}}\)
+
+              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}}\)
+
+              \(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\)
+
+              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}}\)}
+            {\(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}\)}
+
 A system enjoys {\em return integrity} with regard to a call map
-\(\callmap\) and return map \(\retmap\) if for any initial state pair
-\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
-corresponding to a call in its induced trace, if
-\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
-is the first unmatched return such that for some $\pol_r$
-\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
-then $\mach_r$ corresponds to a return state for $\mach_c$:
-\[\ret{\mach_c}\ \mach_r\]
+\(\callmap\) and return map \(\retmap\) if for any initial state
+\(\mach_0\) and any call state and policy state \(\mach_c\) and
+\(\pol_c\) in the induced trace from
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if \(\mpstate{\mach_c}{\pol_c}
+\hookrightarrow_P \MPT\) and \(\mach_r\) is the first unmatched
+return, with \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT \ \MPT'\)
+and \(\mach_r\) the first machine state of \(\MPT'\), then
+\(\ret{\mach_c}\ \mach_r\).
 
@@
 
-    \definition{Observable Integrity}
-      Observable integrity weakens eager integrity by allowing a returned state
-      to differ from its call state at high integrity components, provided the
-      observable behavior of the induced trace from the return matches that of
-      an idealized state whose high integrity components do match the call
-      state.
+    \paragraph{Rolling Back to Idealized State}
+      Observable integrity will weaken eager integrity by allowing a returned
+      state to differ from its call state at high integrity components,
+      provided the observable behavior of the induced trace from the return
+      matches that of an idealized state whose high integrity components do
+      match the call state.
 
@@
       account for premature halts due to policy violations.)
-      \rb{Here, should the second premise use $\hookrightarrow_P$?}
-      %
+
+      Now we can define a version of trace integrity that compares the trace
+      from a returned state to that from its idealized counterpart.
+
+    \definition{Observable Trace Integrity}
+
+      An MP-Trace enjoys {\em observable trace integrity} with respect to
+      some contour \(\contour\) if its final state (if any) has observable
+      behavior that prefixes that of its idealized version, constructed by
+      \(\mathit{roll}\):
+
       \[\begin{split}
         \text{\sc ObsInt}_\contour ~ \MPT \triangleq
-          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
-            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
-        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machOT' \Rightarrow \\
-        & \pi_o(\MPT') \lesssim_O \pi_o(\machOT')
+        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \\
+        & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT' \Rightarrow \\
+        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' 
+          \Rightarrow \\
+        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
       \end{split}\]
@@
 
-      A system enjoys {\em observable \rb{stack?} integrity} with regard to \(\callmap\)
-      and  \(\codemap\) if for any initial state \(\mach_0\) and each
-      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
-      contour \(\contour\) in the induced trace from
+      A system enjoys {\em observable stack integrity} with regard to
+      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
+      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
+      and contour \(\contour\) in the induced trace from
       \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\) \rb{$\hookrightarrow_P$?},
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_e}\),
       then \(\text{\sc ObsInt}_\contour ~ \MPT\).
 
-    \definition{Observable Trace Confidentiality}
+    \paragraph{Restoring Varied States}
 
@@
         definitions.}
+      \sna{Definitely better as a prologue. Where would you split it?}
 
@@
       unless they were changed during execution:
-      \[\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')(\component) =
+      \[\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
+        (\component) =
         \begin{cases}
-          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_) \text{ and } \\
+          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_)
+            \text{ and } \\
           & \mach_e(\component) = \mach_r(\component) \text{ and }
@@
 
-      A pair of traces \(\MPT\) and \(\machOT\) starting from \(\mach_e\) \rb{with what $\pol_e$?} and
-      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\) \rb{$\mach_e'$?},
-      enjoy observable trace confidentiality with respect to \(\contour\) if
-      two conditions hold:
+      Now we can describe a trace property that extends trace confidentiality.
+
+    \definition{Observable Trace Confidentiality}
+
+      A pair of traces \(\MPT\) and \(\machT\) starting from
+      \(\mpstate{\mach_e}{\pol_e}\) and \(\mach_e'\), respectively, with
+      \(\mach_e \approx_\contour \mach_e'\), enjoy observable trace
+      confidentiality with respect to \(\contour\) if two conditions hold:
 
       The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
-      \pi_o(\machOT)\).
+      \pi_o(\machT)\).
 
       If \(\MPT\) is finite and ends with the state
-      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machOT\) is finite. Let
-      \(\mach_r'\) be the final state of \(\machOT\). Then the induced traces
-      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) \rb{$\hookrightarrow_P$?} and
+      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
+      \(\mach_r' = \last(\machT)\). Then the induced traces
+      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT'\) and
       \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
-      \hookrightarrow \machOT'\) are observationally prefixed,
-      \(\pi_o(\MPT') \lesssim \pi_o(\machOT')\).
-
-      We write this \(\text{\sc ObsConf}_\contour ~ \MPT\).
-      \rb{What about $\machOT$? Ditto for use in the definition of observable stack confidentiality below.}
+      \hookrightarrow \machT'\) are observationally prefixed,
+      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).
+
+      We write this \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).
 
@@
       \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
-      then \(\text{\sc ObsConf}_\contour ~ \MPT\).
+      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).
 
@@
       A system enjoys {\em observable stack safety} with regard to a call map
-      and a code map if it enjoys both observable \rb{stack?} integrity and observable
-      \rb{stack?} confidentiality with regard to them.
+      and a code map if it enjoys both observable stack integrity and
+      observable stack confidentiality with regard to them.
 

### 2020-07-07 main.tex
Instantiate variant in example

@@
 
+\paragraph*{Head and Last}
+
+We take the first element of a trace with \(\head(\Trace)\), which is a total
+function since traces are non-empty, and the final element (if one exists) with
+\(\last(\Trace)\), which is partial.
+
 \paragraph*{Until}
@@
 
-%\leo{TODO: We really need a better layout than centered paragraphs.}
-
 \begin{minipage}{.3\textwidth}
@@
 \end{minipage}
-
-% Fixed by Andrew:
-%\COQ{This is the same as the coq version minus a redundant case.}
 
@@
 %
-\[
-\mathit{int} ~ \contour ~ \MPT \triangleq
-\left \{ \begin{array}{ll}
-  \mathit{True}, & \text{if $\MPT$ is infinite}\\
-  \forall \component. ~ \contour(\component) = \HIGHSEC \rightarrow m_0(\component) = m_f(\component), & \text{otherwise} \\
-\end{array}
-\right .
-\]
+\[\begin{split}
+  \mathit{int} ~ \contour ~ \MPT \triangleq &
+    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
+  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
+  & \forall \component. ~ \contour(\component) = \HIGHSEC \rightarrow
+    \mach_e(\component) = \mach_r(\component) \\
+\end{split}\]
 \leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
+\sna{We could do it like this, with finiteness implicit in the existence of
+     a final state.}
 
@@
 instance, the entire trace of {\tt h} from instruction 40 to
-instruction 47. This trace is repeated in \Cref{fig:intex}, and
-recolored: \leo{Explain colors here - not sure I agree with the
-  current choice}
-\leo{Old relevant text:
-For an example see \cref{fig:intex}, in which {\tt h} writes its return
-value to {\tt g}'s return location the stack rather than its own. That address
-is high integrity (red), and so writing to it causes the property to fail.
-}
+instruction 47. This trace is repeated in \Cref{fig:intex}, with the initial
+and final states colored to show the contour; high confidentiality
+components in magenta, low in teal. At the end of the trace addresses 103
+and 108 have both changed; since 103 was high confidentiality, the trace
+does not enjoy trace integrity.
 
@@
 shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
-in red and low confidentiality in green. Below it, the variant state
-\(\mach_e'\) agrees on low confidentiality memory, but the rest might have
-any value, in yellow. After the last step, the return states \(\mach_r\) and
+in magenta and low confidentiality in teal. Below it, the variant state
+\(\mach_e'\) agrees on low confidentiality memory, but all other addresses
+hold -5, in orange. After the last step, the return states \(\mach_r\) and
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
-\(\mach_e'\). Addresses 108 and 109 have changed, shown in pink, and while 109
-is consistent, 108 will not hold 43 in all variants, so the trace does not obey
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in gray, and while 109
+is consistent, 108 holds either 43 or -4, so the trace does not obey
 trace confidentiality.
+
+\sna{I like instantiating the variant, tomorrow I will extend it to the
+     lazy example as well.}
 
@@
 
+    \sna{Use the \(n\)-transition language above in all of these.}
+
     \paragraph{Rolling Back to Idealized State}

### 2020-07-07 main.tex
Found some mismatched color names
Eager vs lazy policies

--- 2020-07-07 main.tex
@@
 integrity is violated.
+\rb{There are some color name macros, but none for integrity yet? Use instead
+of hardcoded color names.}
 
@@
 trace confidentiality.
+\rb{Color name macros}
 
@@
 
+Integrity
+
+Confidentiality as a form of noninterference
+
 \paragraph*{Eager vs lazy testing}
 
-\rb{This part should probably be somewhere in the lazy section.}
-%
-In practice, this \emph{eager policy} is too slow to be of practical interest:
-what we really want in practice is a \emph{lazy policy} that can be efficiently
-implemented.
-
-Less conservative policies can be more efficient, but also harder to test.
-
-We can also test the lazy policy, even if less efficiently, and it should
-enforce an inductive property, which may be more amenable to proving.
-%
-\rb{Would need careful phrasing.}
-%
+The conservative policy we just described, while testable and enforceable, is
+too slow to be of practical use. What we want, instead, is a somewhat more
+permissive policy that moreover can be efficiently implemented. Most of the
+performance overhead incurred by this kind of \emph{eager policy} steps from the
+need to set stack activation tags as a frame is created and reset them as it is
+destructed, as \citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
+mitigate those costs, they propose optimizations like Lazy Tagging and Lazy
+Clearing that relax and defer some of the checks of an eager policy, being more
+permissive in their allowed operations and catching violations if and when their
+effect becomes manifest.
+\rb{How much detail to go into? Refer to examples if we have the space?}
+
+These \emph{lazy policies} admit more efficient implementations while being
+harder to test, given that potential violations may or may not materialize,
+requiring in turn additional bookkeeping. However, a critical observation is
+that the depth-based tagging scheme discussed so far with fail when combined
+with a lazy policy. The reason for this is that because tags are not set and
+cleared on function entry and exit points, stale tags may persist after a given
+call, and a subsequent call sharing the same depth as the old call may access
+memory contents tagged with the stale tags of a previous activation.
+\rb{Again, example?}
+
+The lazy policy can be repaired if, instead of tagging each function activation
+with its depth in the stack, we generate a fresh activation identifier on each
+call, which allows the policy to identify stale tags from previous activations.
+Under this stricted tagging discipline, the eager policy behaves almost exactly
+like its lazy counterpart. A related mechanism was explored in the Static
+Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which associates a
+unique activation identifier to each function, which is however shared by all
+activations of the same function in order to obtain good cacheability of tag
+rules.
+
+\rb{Parts of this discussion feel somewhat out of order...}
+
+\rb{We can also test the lazy policy, even if less efficiently, and it should
+  enforce an inductive property, which may be more amenable to proving --- this
+  should be the connection to the following section.}
+
 \rb{Implied by the property we started with?}
@@
 \rb{Explain and compare what each policy enforces.}
-
-\rb{The Static Authorities differs from the Depth Isolation policy in that frame
-  ids aren't true activation ids, but unique ids associated to all activations
-  of the same function. Discuss unique activation ids and compare with depth
-  later, possibly when discussing lazy policies, as the difference becomes
-  tangible there.}
 
@@
     it prints 0. So observable trace integrity does not hold.
+    \rb{Color name macros}
 
@@
       trace confidentiality does not hold.
+      \rb{Color name macro}
 

### 2020-07-07 main.tex
Wondering how to phrase the various policies

@@
 
-\leo{Policies of interest: null, micro, cherri, software-only}
-\leo{TODO: Write down carefully}
+This model is flexible enough to capture any kind of mechanism used for dynamic
+enforcement of properties of interest, including the following representative
+families of policies.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+\paragraph*{Null policy}
+%
+The simplest possible policy is the neutral policy that carries no information
+and does nothing. Its set of policy states $\POLS$ is the unit type, and its
+policy step function is the constant function that produces the unit, $\lambda
+\_ . \top$, thereby allowing all calls the machine step function. This
+corresponds to the lifting of our machine model to the policy setting.
+
+\paragraph*{Software-only policies}
+%
+\rb{TODO In a way, in these schemes the policy state is built in as part of the
+  program state and transparent to the hardware, although it could be simulated
+  by an overlay where enforcement is represented at the policy level.}
+
+\paragraph*{Capability machines}
+%
+\rb{TODO Here the policy is built into the hardware, although the extended ISA
+  could be split into a machine without capabilities and capabilities as the
+  policy.}
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
+
+\paragraph*{Micropolicies}
+%
+\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+  policy? }
 

### 2020-07-07 main.tex
Andrew's corrections on testing definitions

@@
 \definition{Eager integrity test}
-  Given two machines \apt{machine states?} $\mach$ and $\mach'$ and a contour $\contour$, these
+  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
   satisfy the \emph{eager integrity test} if their high-integrity components
@@
     \contour(\component) = \HIGHINT \rightarrow
-    \mach(\contour) = \mach'(\contour)
+    \mach(\component) = \mach'(\component)
     \]
-\apt{Shouldn't conclusion be $m(k) = m'(k)$ ? }
 
@@
     \vee \mach_2(\component) \neq \mach_2'(\component))
-    \rightarrow \mach_1' = \mach_2'
+    \rightarrow \mach_1'(\component) = \mach_2'(\component)
   \]
-\apt{scoping of $\forall$ is ambiguous. Anyhow, shouldn't this be
-$
-\forall \component . ~
-   (     \mach_1(\component) \neq \mach_1'(\component)
-    \vee \mach_2(\component) \neq \mach_2'(\component))
-    \rightarrow \mach_1'(\component) = \mach_2'(\component)
- $ ?}
+\apt{scoping of $\forall$ is ambiguous.}
 \rb{Fix notations, including for properties below.}
@@
   machine, a contour and a return map:
-  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS (\MACHS \rightarrow \mathcal{P})$$
+  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times 2^\MACHS)$$
   $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$

### 2020-07-07 main.tex
More on capabilities and StkTokens

@@
 
+\rb{Maybe provide a full definition of (eager) stack safety, here or at the end
+  of the section.}
+
 \subsection{Program Structure}
@@
 
+CHERI \rb{cite}
+
+\citet{Skorstengaard+19b} introduce a calling convention to preserve local state
+encapsulation and well-bracketed control flow by the use of local capabilities.
+Although the required hardware support is readily available in capability
+machines like CHERI, this technique incurs potentially prohibitive performance
+costs, because it requires the unused part of the stack to be cleared whenever a
+security boundary is crossed. In the discussion, ``while [the authors] claim
+that [their] calling convention enforces control-flow correctness, [they] do not
+prove a general theorem that shows this, because it is not clear what such a
+theorem should look like,'' noting that the correctness property enjoyed by
+their technique ``is not made very explicit.''
+
 StkTokens \citep{Skorstengaard+19} protects the stack by enforcing local state
@@
 flow by only allowing returns from the topmost frame to the immediately adjacent
-frame below. Formally, these properties are built into a capability machine
-defined by an overlay semantics, which is proven fully abstract with respect to
-a capability machine that replaces those abstract operations with their calling
-convention.
+frame below. Formally, it improves on \citet{Skorstengaard+19b} by building
+those properties into the semantics of a capability machine with a built-in call
+stack and call and return instructions, which is proven fully abstract with
+respect to a more concrete capability machine that replaces those
+pseudo-instructions with their calling convention.
+
+We are confident \rb{how strong a claim to make, with substantiation} that
+StkTokens satisfies the eager stack safety property, and even the stronger eager
+testing property \rb{refs}. The main limitation of StkTokens is its reliance on
+linear capabilities, as it is unclear now they could be added to practical
+capability machines, especially in terms of efficiency.
+%
+\rb{On the other hand, we know that these can be implemented as micropolicies,
+  cf. linear return capabilities in yannis-report (what about performance?)}
 
 \rb{There are quite a few parallels at various levels, although the definitions
-  are overall more implicit and not given formally. A limiting factor is the
-  lack of linear capabilities in existing models like CHERI---although these
-  could be implemented by a micropolicy.}
+  are overall more implicit and not given formally.}
 

### 2020-07-07 main.tex

@@
 shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
-in red and low confidentiality in green. Below it, the variant state
-\(\mach_e'\) agrees on low confidentiality memory, but the rest might have
-any value, in yellow. After the last step, the return states \(\mach_r\) and
+in magenta and low confidentiality in teal. Below it, the variant state
+\(\mach_e'\) agrees on low confidentiality memory, but all other addresses
+hold -5, in orange. After the last step, the return states \(\mach_r\) and
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
-\(\mach_e'\). Addresses 108 and 109 have changed, shown in pink, and while 109
-is consistent, 108 will not hold 43 in all variants, so the trace does not obey
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in gray, and while 109
+is consistent, 108 holds either 43 or -4, so the trace does not obey
 trace confidentiality.
 \rb{Color name macros}
+
+\sna{I like instantiating the variant, tomorrow I will extend it to the
+     lazy example as well.}
 

### 2020-07-07 main.tex
Fixed confidentiality example

@@
 \definition{Trace Confidentiality} We will say two traces $MP$ and $M$
-satisfy {\em trace confidentiality} with respect to $\contour$ if:
+satisfy {\em trace confidentiality} with respect to $\contour$ and
+a return predicate \(R \subset \MACHS\) if:
 \begin{itemize}
-\item If $\MPT$ is finite and its final state $\mach_r$ is a valid
-  return\leo{Is this ill-scoped?}, then $\machT$ is also finite with some final state $\mach_r'$ and
+\item If $\MPT$ is finite and its final state $R\ \mach_r$,
+  then $\machT$ is also finite with some final state $\mach_r'$ and
   any changes from the initial states have changed in the same way:
@@
   \mach_r(\component) = \mach_r'(\component)$$
-  In addittion their observations must be equivalent:
+  In addition their observations must be equivalent:
   $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
@@
   $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
-\item If the $\MPT$ ends prematurely with a policy fault, then
+\item If the $\MPT$ ends prematurely, without fulfilling \(R\), then
   we require that its observations are a prefix of those of $\machT$:
@@
       \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
-      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109, in pink, are
-      not restored; every address that was varied and not changed in one of the
-      executions is. Since 108 contains a value that originates in the
+      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109, in \tainted,
+      are not restored; every address that was varied and not changed in one
+      of the executions is. Since 108 contains a value that originates in the
       variation, values that depend on it might be anything; after returning

### 2020-07-07 main.tex
wbcf examples

--- 2020-07-07 main.tex
@@
 
-\subsection{Attacker Model + Unsafe Variations}
+\subsection{Attacker Model}
 
@@
 
+\subsection{Unsafe Variations}
+
 \leo{TODO: Add citations of mitigation techniques}
@@
 
-\paragraph*{Bad write}
+\paragraph*{Bad Write}
 
@@
 
-\paragraph*{Bad read}
+\paragraph*{Bad Read}
 
@@
 
-\paragraph*{Control flow}
-
-\leo{What should we do here? Overwrite our own ra and return?}
-
-\paragraph*{Reusing call frames}
-
-\leo{TODO}
+\paragraph*{Bad Control Flow}
+
+\[
+5 : \jalr ~ r_4 ~ \rra ~ 0 
+\]
+If a jump instruction was added to {\tt main} with its target being
+the middle of the code block of {\tt h}.
 
@@
 
+For example, consider the unsafe substitution:
+\[
+5 : ~ \jalr ~ r_4 ~ \rra ~ 0
+\]
+This violates control separation as the instruction that causes a
+function switch is not a valid call (or a return) according to the
+call map.
+
 \definition{Entry Integrity}
@@
 \[\callmap(\mach_c(\PCname)) = n \rightarrow \entmap(\mach_e)\]
+
+The same unsafe substitution violates entry integrity
+\[
+5 : ~ \jalr ~ r_4 ~ \rra ~ 0
+\]
+as its target (instruction 42) is not an established entry point for
+{\tt h} according to the entry map.
+
+\leo{Note something about not enforcing control flow integrity
+  with allowing a call from {\tt main} to {\tt h}?}
 
@@
 
+For example, if we added an instruction between 45 and 46:
+\[
+45.5: ~ \addi ~ \rra ~ r_0 ~ \negate 19
+\]
+This would cause the return from {\tt h} to go to instruction 8 and
+return to {\tt main} instead of {\tt g}. While the instruction that
+causes the return would still be a valid return (satisfying control
+separation), the state after the return would not correspond to an
+$\Unmatched$ return for {\tt g}'s call to {\tt h}, violating return
+integrity.
+
 \definition{Well Bracketed Control Flow}
+
 The above properties are somewhat orthogonal, but together they

### 2020-07-07 main.tex
Fix scoping in confidentiality

--- 2020-07-07 main.tex
@@
 $n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
-$f$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from $(\mach_e, \pol_e)$
-until the first return corresponding to $\mach_c$ (if any) satisfies
-trace integrity with respect to every similarly prefixed machine trace
-of every variation $\mach_e'$ with the contour at the call:
-$$\forall \MPT_{pre} ~ \machT_{pre}. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
+$f$ in $\MPT$, the following holds. Let $\MPT_{call}$ be the prefix where
+\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} | \ret{\mach_c}\).
+For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
+\mach_e\), let \(\machT\) be the similarly prefixed trace such that
+\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
+\(\MPT_{call}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
+and \(R = \ret{\mach_c}\):
+
+$$\forall \MPT_{call} ~ \machT. ~
+\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} ~ | ~ \ret{\mach_c} \rightarrow$$
 $$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
-\mach_e' \hookrightarrow \machT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ \MPT_{pre} ~ \machT_{pre}$$
+\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \rightarrow$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{pre} ~ \machT_{pre}$$
 

### 2020-07-07 main.tex
Fixed a few names

--- 2020-07-07 main.tex
@@
 $n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
-$f$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from $(\mach_e, \pol_e)$
-until the first return corresponding to $\mach_c$ (if any) satisfies
-trace integrity with respect to every similarly prefixed machine trace
-of every variation $\mach_e'$ with the contour at the call:
-$$\forall \MPT_{pre} ~ \machT_{pre}. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
+$f$ in $\MPT$, the following holds. Let $\MPT_{call}$ be the prefix where
+\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} | \ret{\mach_c}\).
+For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
+\mach_e\), let \(\machT\) be the similarly prefixed trace such that
+\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
+\(\MPT_{call}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
+and \(R = \ret{\mach_c}\):
+
+$$\forall \MPT_{call} ~ \machT. ~
+\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} ~ | ~ \ret{\mach_c} \rightarrow$$
 $$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
-\mach_e' \hookrightarrow \machT_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ \MPT_{pre} ~ \machT_{pre}$$
+\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \rightarrow$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{call} ~ \machT_{call}$$
 

### 2020-07-07 main.tex
Quick Intro Pass

@@
 They formalize these properties by defining an abstract capability-based machine
-that ``satisfies [stack safety] by construction.'' We believe there is still room
-to craft a more abstract and less intensional characterization of
-stack safety, and that is the goal of the present paper. \apt{still needs
-  tuning}\bcp{Yes, I think we need to go into a little more detail on what
-  their definition looks like in order to justify why an alternative
-  definition might be valuable.  What questions does their presentation not
-  answer? What uses does it not support? E.g., can it be enforced
-  efficiently? (I guess so, by capabilities, though I think their
-  capabilities are more powerful than what Cheri actually provides, while
-  our policies are exactly what Nick needs.)  Why would it be good to look
-  at stack safety through the lens of security?  (Because it is
-  fundamentally a security mechanism?)}
-\rb{BCP idea: contrast intrinsic vs. extrinsic property definition.
-     Remember to talk about spatial vs temporal safety}
+that ``satisfies [stack safety] by construction.''
+%
+While this is an important step in the right direction, we believe
+there is still room to craft a more abstract and less intensional
+characterization of stack safety, so that we are able to answer
+questions about other enforcement mechansims. That is the goal of this
+paper.
+\leo{I took a stab. I think it's still to dense, but captures the
+  feeling of today's discussion I think}
+\apt{still needs tuning}\bcp{Yes, I think we need to go into a
+  little more detail on what their definition looks like in order to
+  justify why an alternative definition might be valuable.  What
+  questions does their presentation not answer? What uses does it not
+  support? E.g., can it be enforced efficiently? (I guess so, by
+  capabilities, though I think their capabilities are more powerful
+  than what Cheri actually provides, while our policies are exactly
+  what Nick needs.)  Why would it be good to look at stack safety
+  through the lens of security?  (Because it is fundamentally a
+  security mechanism?)}  \rb{BCP idea: contrast intrinsic
+  vs. extrinsic property definition.  Remember to talk about spatial
+  vs temporal safety}
 
@@
 stack safety, though we discuss (somewhere) how our property can be combined
-with the heap safety property of [Azevedo et al]}
-
-While integrity and confidentiality are standard security
-properties, applying them in the stack-safety setting raises a novel challenge:
-the notion of ``attacker'' changes during execution. Not only
-does a caller need to be protected from its callee, but the callee
-itself needs to be protected from {\em its} callees.
-We address this challenge by adjusting the integrity and confidentiality
-of different stack frames at each call or return transition between functions.\apt{connection to ``attacker'' still vague.}
-\rb{Mention to varying ``accessing power'' of the attacker?
-  The notion of attacker changes on certain actions that are not externally
-  visible.} \bcp{Maybe the best way is to grab the bull by the horns: ``If
-  we're going to phrase stack safety as a security property, we need to talk
-  about an attacker model, and this is where the technicalities get really
-  interesting...''}
+with the heap safety property of [Azevedo et al]}\leo{Should this be
+  here or in the assumptions as presented in the running example?}
 
@@
 directly (via a jump instruction) or indirectly (by corrupting stack
-data and returning). Well-bracketed control flow establishes safety between
-functions: each instruction is uniquely owned by some function, and all jumps between functions
-are either calls that enter at a marked entry point or returns that re-enter
-just after their associated call in the standard nested call structure.
-\apt{clarify what we are contributing here.}
-
-To provide evidence that our definitions capture the inituitive notion of stack safety,
-we show that they rule out a wide \apt{really?} range of exploitable bad behaviors.\apt{``from the literature''??}
-To assess whether they are implementable,
-we compare them against the proposed capability-based enforcement
-mechanism of \citet{Skorstengaard+19} \apt{we really should} and the existing tag-based enforcement mechanism
-of \citet{DBLP:conf/sp/RoesslerD18}.
+data and returning). Well-bracketed control flow establishes safety
+between functions: each instruction is uniquely owned by some
+function, and all jumps between functions are either calls that enter
+at a marked entry point or returns that re-enter just after their
+associated call in the standard nested call structure.  \apt{clarify
+  what we are contributing here.}
+
+To phrase both aspects of stack safety as a security property requires
+an explicit attacker model which, for this paper, is quite strong: a
+callee can execute arbitrary code, and the enforcement mechanism
+should prevent any stack safety violations. Moreover, even though
+integrity and confidentiality are standard security properties,
+applying them in the stack-safety setting raises a novel challenge:
+the notion of ``attacker'' changes during execution. Not only does a
+caller need to be protected from its callee, but the callee itself
+needs to be protected from {\em its} callees. We address this
+challenge by adjusting the integrity and confidentiality of different
+stack frames at each call or return transition between
+functions.
+%\apt{connection to ``attacker'' still vague.}  \rb{Mention
+%  to varying ``accessing power'' of the attacker?  The notion of
+%  attacker changes on certain actions that are not externally
+%  visible.} \bcp{Maybe the best way is to grab the bull by the horns:
+%  ``If we're going to phrase stack safety as a security property, we
+%  need to talk about an attacker model, and this is where the
+%  technicalities get really interesting...''}
+
+To provide evidence that our definitions capture the inituitive notion
+of stack safety, we show that they rule out a range of exploitable bad
+behaviors.\apt{``from the literature''??}  To assess whether they are
+implementable, we compare them against the proposed capability-based
+enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
+  should} and the existing tag-based enforcement mechanism of
+\citet{DBLP:conf/sp/RoesslerD18}.\leo{Is this where we would rephrase this
+  because we would only talk about it in related work?}
 
@@
   intuitive concepts of local state encapsulation and well-bracketed
-  control flow~(\cref{sec:lse-and-wbcf}) in a very general setting. These properties
+  control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
   rule out common examples of stack memory abuse.

### 2020-07-07 main.tex
another pass on intro

@@
 While this is an important step in the right direction, we believe
-there is still room to craft a more abstract and less intensional
-characterization of stack safety, so that we are able to answer
-questions about other enforcement mechansims. That is the goal of this
-paper.
+there is still room to craft a more abstract, generic characterization
+of stack safety that is independent of any particular enforcement
+mechanism.  In other words, we seek formal criteria that can be applied
+against \emph{any} \apt{stack-based?} system to determine whether or not
+the system is stack-safe.  
+\apt{another stab.}
 \leo{I took a stab. I think it's still to dense, but captures the
@@
 
-Our main contribution in this paper is an alternative formal presentation of
+The main contribution of this paper is a novel formal presentation of
 local state encapsulation and well-bracketed control flow using the
@@
 behaviors.\apt{``from the literature''??}  To assess whether they are
-implementable, we compare them against the proposed capability-based
-enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
-  should} and the existing tag-based enforcement mechanism of
-\citet{DBLP:conf/sp/RoesslerD18}.\leo{Is this where we would rephrase this
-  because we would only talk about it in related work?}
+implementable, we compare them against 
+%proposed capability-based
+%enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
+%  should} and
+the existing tag-based enforcement mechanism of
+\citet{DBLP:conf/sp/RoesslerD18}.
+%\leo{Is this where we would rephrase this because we would only talk about it in related work?}
 
@@
   Abstract formal properties capturing the
-  intuitive concepts of local state encapsulation and well-bracketed
+  intuitive concepts of strict local state encapsulation and well-bracketed
   control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
@@
   A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
-  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
+  This policy enforces our strict stack safety property, and also a (provably) stronger property that is easier to test.
   \apt{If we have no testing, do we even want to talk about the testing
@@
   A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
-  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
-  variant of the policy that does enforce a (provably) weaker variant of our
-  property.
-%
-  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
-    but the point we can make is still strong: To make the property easier
-    to enforce, we can slightly weaken its guarantees; the weaker version is
-    a bit more complicated to state but still retains the pragmatic
-    protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
-    depth-based tags, which do not enforce even our lazy policy. So we need to
-    introduce a version of LT\&C that uses unique activation tags.}
+  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our strict
+  local state encapsulation property,
+  but a modified version of it using unique per-activation tags enforces 
+  a (provably) weaker lazy variant of the property. 
+%
+% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
+%   but the point we can make is still strong: To make the property easier
+%   to enforce, we can slightly weaken its guarantees; the weaker version is
+%   a bit more complicated to state but still retains the pragmatic
+%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
+%   depth-based tags, which do not enforce even our lazy policy. So we need to
+%   introduce a version of LT\&C that uses unique activation tags.}
 

### 2020-07-07 main.tex
More example fixes

--- 2020-07-07 main.tex
@@
 shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
-in magenta and low confidentiality in teal. Below it, the variant state
+in \high and low confidentiality in \low. Below it, the variant state
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
@@
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
-\(\mach_e'\). Addresses 108 and 109 have changed, shown in gray, and while 109
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
 is consistent, 108 holds either 43 or -4, so the trace does not obey
@@
 \rb{Color name macros}
-
-\sna{I like instantiating the variant, tomorrow I will extend it to the
-     lazy example as well.}
 
@@
     in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
-    \(\mach_r\), with high integrity components in red and low integrity in
-    green. Next is the rolled-back state, \(\mach_r' =
+    \(\mach_r\), and the rolled-back state, \(\mach_r' =
     \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
@@
     it prints 0. So observable trace integrity does not hold.
-    \rb{Color name macros}
 
@@
       are not restored; every address that was varied and not changed in one
-      of the executions is. Since 108 contains a value that originates in the
-      variation, values that depend on it might be anything; after returning
-      from {\tt g}, {\tt main} prints 95 in the trace from \(\mach_r\) and
-      some undetermined value in the trace from \(\mach_r''\). So observable
-      trace confidentiality does not hold.
-      \rb{Color name macro}
+      of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
+      and after returning from {\tt g}, {\tt main} prints 85. But in the trace
+      from \(\mach_r''\), 108 contains -4, and ultimately 38 is printed.
+      So observable trace confidentiality does not hold. This example
+      illustrates the weakness of observable confidentiality as a testing
+      property; it takes until the very end of the program to determine whether
+      a tainted value actually influences observable behavior.
 

### 2020-07-07 main.tex
Section 1 polishing / comments

@@
 
-The call stack is a perennially popular target for attacks
-exploiting memory safety vulnerabilities, with potential consequences
+The call stack is a perennial target for attacks
+based on memory safety vulnerabilities, with potential consequences
 ranging from leakage or corruption of private stack data to
-control-flow hijacking. To detect or prevent such attacks, many
+control-flow hijacking. To detect or prevent such attacks, a menagerie of
 software and hardware techniques for stack safety have been proposed,
@@
   nick;there could be more}\bcp{How about static techniques??  We decided
-  these were also in-scope.}
+  these were also in-scope.}\bcp{... or are they?  I guess in any case we should
+list any that we know about.}
 
 But what {\em is} this ``stack safety'' of which we speak? Precise formal
-definitions are lacking in the literature; instead, the notion of safety
-is usually defined negatively, by enumerating the bad things that a safe system
+definitions are lacking in the literature; instead, the notion
+is usually defined negatively, by enumerating the bad things that a given
+enforcement mechanism
 prevents, such as corruption of return addresses, buffer overflows, or
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-We are only aware of one, quite recent, attempt to frame a positive definition.
+Indeed, we are only aware of one---quite recent---attempt to frame a
+positive definition:
 \citet{Skorstengaard+19} define stack safety as
@@
 %
-While this is an important step in the right direction, we believe
-there is still room to craft a more abstract, generic characterization
-of stack safety that is independent of any particular enforcement
-mechanism.  In other words, we seek formal criteria that can be applied
-against \emph{any} \apt{stack-based?} system to determine whether or not
-the system is stack-safe.  
+Although this represents an important step in the right direction, we believe
+there is still room to craft a more abstract, more generic characterization
+of stack safety that is independent of capabilities or any other specific
+enforcement
+technique.  In other words, we seek formal criteria that can be applied
+against \emph{any} stack-based system + enforcement mechanism to judge
+whether or not the combination is actually stack-safe.
+
+\iflater
 \apt{another stab.}
@@
   vs temporal safety}
-
-The main contribution of this paper is a novel formal presentation of
+\fi
+
+The key contribution of this paper is a novel\bcp{generic?} presentation of
 local state encapsulation and well-bracketed control flow using the
-conceptual and technical framework of language-based security.
-%
-The underlying notion behind local state encapsulation is the intention to
-protect callers from callees: a callee should be not be able to access or
-overwrite private data in the caller's stack frame. We formalize this
-intuition as a security property with both confidentiality and integrity
-components.
-%
-Intuitively, {\em stack confidentiality} promises that a callee's
+technical framework of language-based security.
+
+The intuition behind \citet{Skorstengaard+19}'s\iflater\bcp{I forget how to
+  remove the year and just have the author names here...}\fi use of the term {\em
+  local state encapsulation} is that callers should be protected from
+callees: a callee should be not be able to access or overwrite private data
+in the caller's stack frame. We formalize this intuition as a security
+property with both confidentiality and integrity components.
+%
+We reflect this intuition in a formal
+{\em stack confidentiality} property, which promises that a callee's
 observable behavior is not affected by memory outside its stack frame
-(e.g. a caller's private data or uninitialized memory).  {\em Stack
-  integrity} guarantees that, if a function call ever returns, the caller's
+(e.g. a caller's private data or uninitialized memory).  Conversely, {\em stack
+  integrity} promises that, if a function call ever returns, the caller's
 stack data at the point of return will be identical to what it was before
-the call.  \bcp{We need to point out somewhere that we are only dealing with
+the call.
+
+\iflater
+\bcp{We need to point out somewhere that we are only dealing with
 stack safety, though we discuss (somewhere) how our property can be combined
 with the heap safety property of [Azevedo et al]}\leo{Should this be
-  here or in the assumptions as presented in the running example?}
-
-The underlying notion behind well-bracketed control flow is
-the need to protect the call structure itself: a callee should not be
-allowed to jump to the middle of an arbitrary block of code either
+  here or in the assumptions as presented in the running
+  example?}\bcp{Probably better there.}
+\fi
+
+The intuition behind \citet{Skorstengaard+19}'s {\em well-bracketed control
+  flow} is 
+the desire to protect the call graph itself: a callee should pass control
+back to the caller only by returning in the standard way (transferring
+control to the instruction following the call); it should 
+not be
+allowed to jump back into the caller's code at an arbitrary point, either
 directly (via a jump instruction) or indirectly (by corrupting stack
@@
 
-To phrase both aspects of stack safety as a security property requires
-an explicit attacker model which, for this paper, is quite strong: a
-callee can execute arbitrary code, and the enforcement mechanism
-should prevent any stack safety violations. Moreover, even though
-integrity and confidentiality are standard security properties,
-applying them in the stack-safety setting raises a novel challenge:
+To phrase these two aspects of stack safety in terms of security properties
+requires an explicit attacker model. \bcp{This paragraph goes too fast: We
+  need to start by asking, informally, who is being protected and who is the
+attacker. Then we observe that real-world stack protection mechanisms tend
+to imagine attack scenarios where the callee gets confused by something it
+receives from the outside; to avoid getting into the details of how this
+happens, we simply assume that the callee can execute arbitrary code.}
+For simplicity, we choose one that is
+fairly strong: when stating the protections that a function can rely on when
+it makes a call, we assume that the callee can execute arbitrary
+instructions until it returns.
+
+\bcp{Again, here, a bigger transition, more setup, and a longer discussion
+  are needed.  This is really the crux of what we are doing!}
+Moreover, applying the standard language-based-security notions of
+integrity and confidentiality in the stack-safety setting raises a novel
+challenge:
 the notion of ``attacker'' changes during execution. Not only does a
@@
 stack frames at each call or return transition between
-functions.
+functions.  \bcp{that's way too terse!}
 %\apt{connection to ``attacker'' still vague.}  \rb{Mention
@@
 
+\bcp{This little paragraph is our entire explanation of the second and third
+contributions!  Needs to be expanded considerably...}%
 To provide evidence that our definitions capture the inituitive notion
@@
 
-In summary, we offer the following contributions:
+In summary, we offer the following contributions:\BCP{I would still like to
+  see these points phrased in ``contribution normal form'': say exactly what
+  we did for each one (it's a bit vague---''considers,'' etc.), and
+  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
 \begin{itemize}
@@
 \end{itemize}
-\Cref{sec:relwork} discusses related work, and
+\Cref{sec:relwork} discusses related work;
 \cref{sec:future} sketches directions for future work.

### 2020-07-07 main.tex
Pass though 2-4

--- 2020-07-07 main.tex
@@
 While this is an important step in the right direction, we believe
-there is still room to craft a more abstract, generic characterization
-of stack safety that is independent of any particular enforcement
-mechanism.  In other words, we seek formal criteria that can be applied
-against \emph{any} \apt{stack-based?} system to determine whether or not
-the system is stack-safe.  
-\apt{another stab.}
+there is still room to craft a more abstract and less intensional
+characterization of stack safety, so that we are able to answer
+questions about other enforcement mechansims. That is the goal of this
+paper.
 \leo{I took a stab. I think it's still to dense, but captures the
@@
 
-The main contribution of this paper is a novel formal presentation of
+Our main contribution in this paper is an alternative formal presentation of
 local state encapsulation and well-bracketed control flow using the
@@
 behaviors.\apt{``from the literature''??}  To assess whether they are
-implementable, we compare them against 
-%proposed capability-based
-%enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
-%  should} and
-the existing tag-based enforcement mechanism of
-\citet{DBLP:conf/sp/RoesslerD18}.
-%\leo{Is this where we would rephrase this because we would only talk about it in related work?}
+implementable, we compare them against the proposed capability-based
+enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
+  should} and the existing tag-based enforcement mechanism of
+\citet{DBLP:conf/sp/RoesslerD18}.\leo{Is this where we would rephrase this
+  because we would only talk about it in related work?}
 
@@
   Abstract formal properties capturing the
-  intuitive concepts of strict local state encapsulation and well-bracketed
+  intuitive concepts of local state encapsulation and well-bracketed
   control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
@@
   A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
-  This policy enforces our strict stack safety property, and also a (provably) stronger property that is easier to test.
+  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
   \apt{If we have no testing, do we even want to talk about the testing
@@
   A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
-  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our strict
-  local state encapsulation property,
-  but a modified version of it using unique per-activation tags enforces 
-  a (provably) weaker lazy variant of the property. 
-%
-% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
-%   but the point we can make is still strong: To make the property easier
-%   to enforce, we can slightly weaken its guarantees; the weaker version is
-%   a bit more complicated to state but still retains the pragmatic
-%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
-%   depth-based tags, which do not enforce even our lazy policy. So we need to
-%   introduce a version of LT\&C that uses unique activation tags.}
+  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
+  variant of the policy that does enforce a (provably) weaker variant of our
+  property.
+%
+  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
+    but the point we can make is still strong: To make the property easier
+    to enforce, we can slightly weaken its guarantees; the weaker version is
+    a bit more complicated to state but still retains the pragmatic
+    protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
+    depth-based tags, which do not enforce even our lazy policy. So we need to
+    introduce a version of LT\&C that uses unique activation tags.}
 
@@
 
-The example program appears in \Cref{fig:running-program}. It consists of a
-{\tt main} function, written in some C-like high-level language, that
-performs a call to some function {\tt f}, which in turns makes a nested call
-to some function {\tt g} before returning, performing some simple arithmetic
-in the process.
+The example program appears in \Cref{fig:running-program}. It consists
+of a {\tt main} function, written in some C-like high-level language,
+that performs a call to some function {\tt f}, which in turns makes a
+nested call to some function {\tt g} before returning, while
+performing some simple arithmetic operations in the process.
 
@@
 \setcounter{pcctr}{0}
+%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
 \newcommand*{\row}[4]{
-  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
+  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\  
 }
@@
 
-\newcommand{\instrc}{magenta}
+\colorlet{lgray}{gray!40}
+
+\newcommand{\instrc}{lgray}
 \newcommand{\outc}{white}
@@
 \[
-  \begin{array}{c|c|c|c|c}
-    PC & \mathit{Instruction} & \mathit{Description} & \mathit{Effect} & \\
+  \begin{array}{c|c|c|c}
+    PC & \mathit{Instruction} & \mathit{Effect} & \\
     \hline
@@
     \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
-        {Make the call}
+        {Make the call (Jump and link)}
   \end{array}
@@
   ~\\
-  \apt{How's that description for last line?}
-  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
-    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
-  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}
+%  \apt{How's that description for last line?}
+%  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
+%    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
+%  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}
 % At Call to G
@@
   \[
-  \begin{array}{c|c|c|c|c}
+  \begin{array}{c|c|c|c}
     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 106}
@@
         {Zero out result}
-    \row{\jal ~ \rra ~ 40}{??}{\PCname \leftarrow 40; ~ \rra \leftarrow \thepcctr}
-        {Make the call to {\tt h}}
+    \row{\jal ~ \rra ~ 40}{\jal ~ \rra ~ 40}{\PCname \leftarrow 40; ~ \rra \leftarrow \thepcctr}
+        {Make the call to {\tt h} (jump-and-link)}
   \end{array}
@@
 \[
-\begin{array}{c|c|c|c|c}
+\begin{array}{c|c|c|c}
   \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
@@
 \[
-\begin{array}{c|c|c|c|c}
+\begin{array}{c|c|c|c}
   \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
@@
 \[
-\begin{array}{c|c|c|c|c}
+\begin{array}{c|c|c|c}
   \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
@@
 \[
-\begin{array}{c|c|c|c|c}
+\begin{array}{c|c|c|c}
   \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
@@
 
-To show a trace of the execution of this program, we need
-a low-level view of both the code and the machine it runs on. While
-our concrete implementation (Section~\leo{cite}) operates on a full
-conventional RISC-V architecture, we make a few simplifying
-assumptions here to ease the exposition.
-
-\bcp{We should also emphasize, again, that there is no heap (if there were,
-we might put it above the stack or between the instructions and the stack)}
-First of all, we ignore the heap. We are mainly concerned
-with stack interactions so that will be the main focus. \leo{We're
-  going to come back to that at the end. Will we?} We also
-abstract away low-level details of memory layout and present a
+To show a trace of the execution of this program, we need a low-level
+view of both the code and the machine it runs on. While the concrete
+implementation of our enforcement mechanism (\Cref{sec:enforcement})
+operates on a full conventional RISC-V architecture (using the FORVIS
+simulator\leo{cite}), we make a few simplifying assumptions here to
+ease the exposition.
+
+First of all, we ignore the heap. We are mainly concerned with stack
+interactions so that will be the main focus; integrating a heap in the
+style of \citet{AzevedoDeAmorim17} is a natural extension we discuss
+in~\Cref{sec:future}.
+%
+We also abstract away low-level details of memory layout and present a
 machine that is word-addressed, with instructions starting from
@@
 A sketch of the memory layout at that point can be found in the top of
-\cref{fig:running-trace-b}: instructions for {\tt main} begin at
-0, the dedicated output is empty\apt{what does that mean? why not just uninitialized?}, while the (uninitialized) stack
-begins at 100. Directly below the layout, we summarize the parts of
-the register file that are relevant to the example.
+\cref{fig:running-trace-b}: instructions for {\tt main} begin at 0,
+the dedicated output is uninitialized, while the (also uninitialized)
+stack begins at 100. Directly below the layout, we summarize the parts
+of the register file that are relevant to the example.
 %
@@
 
+Stack safety is meant to protect a caller from a malicious callee's
+attempts to access private data or highjack its control flow. Our
+attacker model reflects this intention: we assume that the attacker
+can run arbitrary code as a caller and the enforcement mechanism
+should protect against that. Hardware and timing attacks are out of
+scope. \leo{Anything else to add here?}
+\leo{Maybe a good place to discuss calling conventions and that they
+  are essentially enforced by the policy or sth?}
+\rb{Intuition that the callee can do anything in principle, but the machine will
+halt if it does not follow the calling convention, etc.?}
+
+\subsection{Unsafe Variations}
 \sna{Variations here collides with state variations. I propose substitutions.}
-
-\leo{TODO: Make attacker model more explicit} The execution trace
-shown in \cref{fig:running-trace-b} is an example of a
-``stack-safe'' execution: each process interacts only with the
-registers, its local stack space and its call frame. But what could go
-wrong if an attacker could hijack an arbitrary call and execute
-arbitrary code? In the rest of this section we'll go over examples of
-unsafe behaviors, before formally capturing what stack-safe means
-formally and discussing enforcement techniques from the literature
-in later sections.
-
-\subsection{Unsafe Variations}
+\leo{I agree there is a collision, not sure about ``substitutions''}
+
+The execution trace shown in \cref{fig:running-trace-b} is an example
+of a ``stack-safe'' execution: each process interacts only with the
+registers, its local stack space and its call frame. In the rest of
+this section we'll go over examples of unsafe behaviors, before
+formally capturing what stack-safe means formally and discussing
+enforcement techniques from the literature in later sections.
 
@@
 connected to this work?}
-
-\rb{Intuition that the callee can do anything in principle, but the machine will
-halt if it does not follow the calling convention, etc.?}
 
@@
 example of the previous section for concreteness, and then,
-in~\cref{sec:micropolicies}, extend this model with a dynamic policy
-enforcement mechanism in~\cref{sec:micropolicies} based on
-\bcp{NickPaper,AndMaybeSomeOthers}.
+in~\cref{sec:enforcement}, extend this model with a dynamic policy
+enforcement mechanism based on
+\citet{DBLP:conf/sp/RoesslerD18}.
 
@@
 
-\leo{Other choices for observations? Where should that be discussed?}
-\bcp{But in this paragraph we are describing the particular
-choise that we make for our running example, no?  But we certainly want to
-discuss this fully someplace!}
-\leo{Here: Many observation choices. Next: Simplify:}
-Mapping a conventional architecture to this abstract interface is
-straightforward. For the running example of the previous section,
-$\WORDS$ is the set of 64-bit integers, $\REGS$ is the set $\{r_0,
-\cdots, r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$. Observations are
-the writes to the special $\rout$ location: if during an execution
-step a word $w$ is written to $\rout$ then $\obs = w$, otherwise $\obs
-= \tau$.
+The choice of observations is an interesting one. A first approach
+would be to consider any change to the registers or memory observable.
+However, as some parts of the stack will be considered ``private''
+this choice won't work. A relaxation of this to consider all changes
+to ``non-private'' data observable could work in principle, but
+doesn't allow for an interesting weaker-but-more-efficient enforcement
+mechanism. We'll come back to this point in \Cref{sec:lazy}. On the
+other hand, we could use a particular location as an abstract output
+of our system (e.g. through DMA). This is the approach we chose for
+our running example to keep things simple. For that, $\WORDS$ is the
+set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
+\cup \{\PCname\}$, and $\SP = r_2$. Observations are the writes to the
+special $\rout$ location: if during an execution step a word $w$ is
+written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.
 
@@
 
-\leo{Which of the machine, machine-observation, machine-policy,
-  and machine-policy-observation traces should we distinguish
-  in metavariables?}
-\rb{Should we have notation for traces somewhere here?}
-
 A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
@@
 observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.
-\rb{The way the definitions are set up, one may expect MPO, but\ldots}
 
@@
 
-In our running example in Figure~\cref{fig:running-example-b} we can
+In our running example in \Cref{fig:running-trace-b} we can
 see a depiction of such a trace: starting from an initial machine
@@
 \judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
-         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
+         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
 \end{minipage}
@@
               {\(\pol_0 \rightharpoonup \pol_1\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
 \end{minipage}
-\rb{Shouldn't the conclusion of the first judgment use $\hookrightarrow_P$?
-  And in the second judgment, wouldn't $MO$ be $MPO$, and the third premise $\hookrightarrow_P$?}
 
@@
 
-\paragraph*{Observational Equivalence}
+\paragraph*{Observational Similarity}
 
@@
 
-\leo{Rephrase transition. I hate it:} There are two interesting things
-to note.  First, observation trace equivalence could alternatively be
-defined as the prefix relation but dropping the only assymetric rules
-\leo{2 + 3. TODO: name rules?}
-
-\leo{Also rephrase this.}
-Second, an infinite silent trace is a prefix of (and equivalent to)
-any other trace. While this might seem surprising at first, it makes
-sense in a timing-insensitive context: an external observer looking
-at two machine runs cannot (computably) distinguish between a machine that
-steps forever and a machine that steps for a long time before producing
-some output.
+\leo{I still don't know how to do this transition. Maybe contrast
+with the ``perhaps surprisingly'' later?} Note that
+similarity of observation traces could alternatively be defined coinductively,
+just like the prefix relation, but by dropping the assymetric rules that allow
+for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$)
+
+In addition, perhaps surprisingly, an infinite silent trace is a
+prefix of (and equivalent to) any other trace. While this might seem
+surprising at first, it makes sense in a timing-insensitive context:
+an external observer looking at two machine runs cannot (computably)
+distinguish between a machine that steps forever and a machine that
+steps for a long time before producing some output.
 
@@
 program structure, and then dive in to formal definitions local state
-encapsulation and well-bracketed control flow.
-
-\rb{Maybe provide a full definition of (eager) stack safety, here or at the end
-  of the section.}
+encapsulation and well-bracketed control flow that together constitute
+stack safety.
 
@@
 before instruction 8 corresponding to the return from {\tt g}.
+
+\medskip
+
+In the rest of this section we will implictly parameterize our
+definitions by a function map $\codemap$, a call map $\callmap$, a return
+map $\retmap$ and an entry map $\entmap$.
 
@@
 callee and high confidentiality otherwise; regardless they are always
-high integrity.\leo{That would mean we scramble some instructions, I
-  guess that makes sense.}
+high integrity.
 %
@@
                          \text{ and for some } \funid, \codemap(\component) = \funid \\
-    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
-                          \text{ and } \component \leq \mach(\SP) \negate  n \\
     (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
@@
                           \text{ or if } \component = \rout\\
+    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
+                          \text{ and } \component \leq \mach(\SP) \negate  n \\
     (\LOWSEC,\LOWINT) & \text{else} \\
@@
 
-Going back to our running example, recall the state of the machine
-at the call from {\tt main} to {\tt g}:
+Going back to our running example, recall the state of the machine at
+the call from {\tt main} to {\tt g}, annotated with their
+corresponding labels (not shown: the labels of the instructions
+corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
+$(\HIGHSEC, \HIGHINT)$):
 \vspace*{0.2em}
@@
 \MemoryLabel{43.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\instrc}[{\makebox[0pt]{$(\HIGHSEC,\HIGHINT)$}}]%
 \hspace*{3pt}
 $\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{6}{\unusedc}[{\makebox[0pt]{Uninitialized}}]
+\memory{1}{white}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
+\memory{2}{\fprivc}[{\makebox[0pt]{$(\HIGHSEC, \HIGHINT)$}}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{$(\LOWSEC, \LOWINT)$}}]%
+\memory{6}{\unusedc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]
 ~$\cdots$
@@
 
-If $\mach$ is the state at this call, the corresponding contour would
-be:
-\[
-\mathit{Cof}(\mach,2) = \left \{
-\begin{array}{l}
-  0\ldots98\} \mapsto (\LOWSEC, \HIGHINT)\\
-  99 \mapsto (\HIGHSEC, \LOWINT)\\
-  100, 101 \mapsto (\HIGHSEC, \HIGHINT)\\
-  102, 103 \mapsto (\LOWSEC, \LOWINT)\\
-  104\ldots \mapsto (\HIGHSEC, \LOWINT)\\
-  r \mapsto (\LOWSEC, \LOWINT)\\
-\end{array}
-\right .
-\]
+%If $\mach$ is the state at this call, the corresponding contour would
+%be:
+%\[
+%\mathit{Cof}(\mach,2) = \left \{
+%\begin{array}{l}
+%  0\ldots98\} \mapsto (\LOWSEC, \HIGHINT)\\
+%  99 \mapsto (\HIGHSEC, \LOWINT)\\
+%  100, 101 \mapsto (\HIGHSEC, \HIGHINT)\\
+%  102, 103 \mapsto (\LOWSEC, \LOWINT)\\
+%  104\ldots \mapsto (\HIGHSEC, \LOWINT)\\
+%  r \mapsto (\LOWSEC, \LOWINT)\\
+%\end{array}
+%\right .
+%\]
 
@@
 
-\definition{Trace Integrity}
-\label{def:trace-integrity}
+\definition
+
 Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
@@
   & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
-  & \forall \component. ~ \contour(\component) = \HIGHSEC \rightarrow
+  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
     \mach_e(\component) = \mach_r(\component) \\
 \end{split}\]
-\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
-\sna{We could do it like this, with finiteness implicit in the existence of
-     a final state.}
+%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
+%\sna{We could do it like this, with finiteness implicit in the existence of
+%     a final state.}+1
 
@@
 $$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
-\MPT_{pre} ~ | ~ \ret{\mach_c} \rightarrow
+\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
 \mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$
@@
 \mach \approx_\contour \mach' \triangleq \forall \component .
-\contour(\component) = \LOWSEC \rightarrow \mach(\component) =
+\contour(\component) = \LOWSEC \Rightarrow \mach(\component) =
 \mach'(\component)
@@
 $$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
-  \vee \mach_e'(\component) \not = \mach_r'(\component) \rightarrow
+  \vee \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
   \mach_r(\component) = \mach_r'(\component)$$
@@
 shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
-in \high and low confidentiality in \low. Below it, the variant state
+in magenta and low confidentiality in teal. Below it, the variant state
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
@@
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
-\(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in gray, and while 109
 is consistent, 108 holds either 43 or -4, so the trace does not obey
@@
 \rb{Color name macros}
+
+\sna{I like instantiating the variant, tomorrow I will extend it to the
+     lazy example as well.}
 
@@
 $$\forall \MPT_{call} ~ \machT. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} ~ | ~ \ret{\mach_c} \rightarrow$$
+\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} ~ | ~ \ret{\mach_c} \Rightarrow$$
 $$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
-\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \rightarrow$$
+\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
 $$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{call} ~ \machT_{call}$$
@@
 
-\[ \codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname)) \rightarrow
+\[ \codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname)) \Rightarrow
 \exists n. ~ \callmap(\mach_1(\PCname)) = n \vee
@@
 \(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
-\[\callmap(\mach_c(\PCname)) = n \rightarrow \entmap(\mach_e)\]
+\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]
 
@@
     in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
-    \(\mach_r\), and the rolled-back state, \(\mach_r' =
+    \(\mach_r\), with high integrity components in red and low integrity in
+    green. Next is the rolled-back state, \(\mach_r' =
     \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
@@
       are not restored; every address that was varied and not changed in one
-      of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
-      and after returning from {\tt g}, {\tt main} prints 85. But in the trace
-      from \(\mach_r''\), 108 contains -4, and ultimately 38 is printed.
-      So observable trace confidentiality does not hold. This example
-      illustrates the weakness of observable confidentiality as a testing
-      property; it takes until the very end of the program to determine whether
-      a tainted value actually influences observable behavior.
+      of the executions is. Since 108 contains a value that originates in the
+      variation, values that depend on it might be anything; after returning
+      from {\tt g}, {\tt main} prints 95 in the trace from \(\mach_r\) and
+      some undetermined value in the trace from \(\mach_r''\). So observable
+      trace confidentiality does not hold.
+      \rb{Color name macro}
 

### 2020-07-07 main.tex

@@
 %
-While this is an important step in the right direction, we believe
-there is still room to craft a more abstract and less intensional
-characterization of stack safety, so that we are able to answer
-questions about other enforcement mechansims. That is the goal of this
-paper.
+Although this represents an important step in the right direction, we believe
+there is still room to craft a more abstract, more generic characterization
+of stack safety that is independent of capabilities or any other specific
+enforcement
+technique.  In other words, we seek formal criteria that can be applied
+against \emph{any} stack-based system + enforcement mechanism to judge
+whether or not the combination is actually stack-safe.
+
+\iflater
+\apt{another stab.}
 \leo{I took a stab. I think it's still to dense, but captures the
@@
   vs temporal safety}
-
-Our main contribution in this paper is an alternative formal presentation of
+\fi
+
+The key contribution of this paper is a novel\bcp{generic?} presentation of
 local state encapsulation and well-bracketed control flow using the
-conceptual and technical framework of language-based security.
-%
-The underlying notion behind local state encapsulation is the intention to
-protect callers from callees: a callee should be not be able to access or
-overwrite private data in the caller's stack frame. We formalize this
-intuition as a security property with both confidentiality and integrity
-components.
-%
-Intuitively, {\em stack confidentiality} promises that a callee's
+technical framework of language-based security.
+
+The intuition behind \citet{Skorstengaard+19}'s\iflater\bcp{I forget how to
+  remove the year and just have the author names here...}\fi use of the term {\em
+  local state encapsulation} is that callers should be protected from
+callees: a callee should be not be able to access or overwrite private data
+in the caller's stack frame. We formalize this intuition as a security
+property with both confidentiality and integrity components.
+%
+We reflect this intuition in a formal
+{\em stack confidentiality} property, which promises that a callee's
 observable behavior is not affected by memory outside its stack frame
-(e.g. a caller's private data or uninitialized memory).  {\em Stack
-  integrity} guarantees that, if a function call ever returns, the caller's
+(e.g. a caller's private data or uninitialized memory).  Conversely, {\em stack
+  integrity} promises that, if a function call ever returns, the caller's
 stack data at the point of return will be identical to what it was before
-the call.  \bcp{We need to point out somewhere that we are only dealing with
+the call.
+
+\iflater
+\bcp{We need to point out somewhere that we are only dealing with
 stack safety, though we discuss (somewhere) how our property can be combined
 with the heap safety property of [Azevedo et al]}\leo{Should this be
-  here or in the assumptions as presented in the running example?}
-
-The underlying notion behind well-bracketed control flow is
-the need to protect the call structure itself: a callee should not be
-allowed to jump to the middle of an arbitrary block of code either
+  here or in the assumptions as presented in the running
+  example?}\bcp{Probably better there.}
+\fi
+
+The intuition behind \citet{Skorstengaard+19}'s {\em well-bracketed control
+  flow} is 
+the desire to protect the call graph itself: a callee should pass control
+back to the caller only by returning in the standard way (transferring
+control to the instruction following the call); it should 
+not be
+allowed to jump back into the caller's code at an arbitrary point, either
 directly (via a jump instruction) or indirectly (by corrupting stack
@@
 behaviors.\apt{``from the literature''??}  To assess whether they are
-implementable, we compare them against the proposed capability-based
-enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
-  should} and the existing tag-based enforcement mechanism of
-\citet{DBLP:conf/sp/RoesslerD18}.\leo{Is this where we would rephrase this
-  because we would only talk about it in related work?}
-
-In summary, we offer the following contributions:
+implementable, we compare them against
+%proposed capability-based
+%enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
+%  should} and
+the existing tag-based enforcement mechanism of
+\citet{DBLP:conf/sp/RoesslerD18}.
+%\leo{Is this where we would rephrase this because we would only talk about it in related work?}
+
+In summary, we offer the following contributions:\bcp{I would still like to
+  see these points phrased in ``contribution normal form'': say exactly what
+  we did for each one (it's a bit vague---''considers,'' etc.), and
+  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
 \begin{itemize}
@@
   A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
-  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
-  variant of the policy that does enforce a (provably) weaker variant of our
-  property.
-%
-  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
-    but the point we can make is still strong: To make the property easier
-    to enforce, we can slightly weaken its guarantees; the weaker version is
-    a bit more complicated to state but still retains the pragmatic
-    protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
-    depth-based tags, which do not enforce even our lazy policy. So we need to
-    introduce a version of LT\&C that uses unique activation tags.}
+  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our strict
+  local state encapsulation property,
+  but a modified version of it using unique per-activation tags enforces
+  a (provably) weaker lazy variant of the property.
+%
+% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
+%   but the point we can make is still strong: To make the property easier
+%   to enforce, we can slightly weaken its guarantees; the weaker version is
+%   a bit more complicated to state but still retains the pragmatic
+%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
+%   depth-based tags, which do not enforce even our lazy policy. So we need to
+%   introduce a version of LT\&C that uses unique activation tags.}
 
@@
 observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces. 
+observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.
 

### 2020-07-08 main.tex
some fixes in 2

@@
 \newcommand{\instrc}{lgray}
-\newcommand{\outc}{white}
 \newcommand{\fprivc}{cyan}
@@
 \newcommand{\unusedc}{gray}
+\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ? 
+\newcommand{\fulloutc}{white}
 
@@
 $\cdots$
-\memory{1}{\outc}[{\makebox[0pt]{Output}}]%
+\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
 \memory{10}{\unusedc}[{\makebox[0pt]{Uninitialized}}]
@@
 \bcp{A very small nit: Change
-  $\rsp-2$ to $\rsp\mathord{-}2$}
+  $\rsp-2$ to $\rsp\mathord{-}2$}\apt{??}
 \[
@@
     \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
-        {Make the call (Jump and link)}
+        {Call to {\tt g} ($\jal$)}
   \end{array}
@@
 $\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
 \memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
@@
     \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp - 1] \leftarrow r_0}{[107] \leftarrow 0}
-        {Zero out result}
-    \row{\jal ~ \rra ~ 40}{\jal ~ \rra ~ 40}{\PCname \leftarrow 40; ~ \rra \leftarrow \thepcctr}
-        {Make the call to {\tt h} (jump-and-link)}
+        {Zero out result \apt{why?}}
+    \row{\jal ~ \rra ~ 40}{\rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow 40}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 40}
+        {Call to {\tt h} ($\jal$)}
   \end{array}
@@
       {Restore the stack pointer of {\tt g}}
-  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 27}
-      {Return to {\tt g}}
+  \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 27}
+      {Return to {\tt g} ($\jalr$)}
   \end{array}
@@
 \MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{\outc}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\emptyoutc}[{\makebox[0pt]{Instructions}}]%
 \hspace*{3pt}
 $\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
 \memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
@@
   \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
-      {Load argument {\tt z} $r_5$}
+      {Load argument {\tt z} to $r_5$}
   \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
@@
   \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
-      {Store $r_4$ as ther return of {\tt g}}
+      {Store $r_4$ as the return value}
   \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
@@
       {Deallocate {\tt g}'s local state}
-  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 8}{Return to {\tt main}}
+  \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 8}{Return to {\tt main} ($\jalr$)}
 \end{array}
@@
 $\cdots$
-\memory{1}{white}[{\makebox[0pt]{Output}}]%
+\memory{1}{\fulloutc}[{\makebox[0pt]{Output}}]%
 \memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
@@
 \caption{Detailed Execution Trace
-  \rb{A suggestion for the description of $\jal$ and $\jalr$:
-    call(function label) [if labels are added next to the entrypoints]
-    or call(address of entrypoint) and return()?}
+%  \rb{A suggestion for the description of $\jal$ and $\jalr$:
+%    call(function label) [if labels are added next to the entrypoints]
+%    or call(address of entrypoint) and return()?}
 }
@@
 view of both the code and the machine it runs on. While the concrete
-implementation of our enforcement mechanism (\Cref{sec:enforcement})
+implementation of our enforcement mechanism (\Cref{sec:enforcement})\apt{Is it ``ours'' in this context?}
 operates on a full conventional RISC-V architecture (using the FORVIS
-simulator\leo{cite}), we make a few simplifying assumptions here to
+simulator\leo{cite}\apt{why mention this here?}), we make a few simplifying assumptions here to
 ease the exposition.
@@
 First of all, we ignore the heap. We are mainly concerned with stack
-interactions so that will be the main focus; integrating a heap in the
+interactions so that will be the main focus; integrating a safety
+propety for the heap in the 
 style of \citet{AzevedoDeAmorim17} is a natural extension we discuss
@@
 machine that is word-addressed, with instructions starting from
-location 0, a dedicated output at location 99, and lay the stack from
-location 100 upwards. \leo{Explain more? How to connect to realistic
+location 0, a dedicated memory-mapped output port at location 99, and
+the stack beginning at location 100 and growing upwards with increasing
+addresses.  \leo{Explain more? How to connect to realistic
   one? Observations? DMA? Interrupt handlers?}
 %
-A sketch of the memory layout at that point can be found in the top of
+A sketch of the initial memory layout can be found in the top of
 \cref{fig:running-trace-b}: instructions for {\tt main} begin at 0,
@@
 %
-The rest of \cref{fig:running-trace-b} details the execution of the
-program, showing the machine state at call and return points, as well
-as describing the full effect of each instruction in detail.
-\leo{Expand upon this?}
+The rest of \cref{fig:running-trace-b} traces the execution of the
+program, showing the machine state at call and return points.
+Instructions are intended to
+be similar to RISC-V, but are described in an ad-hoc notation
+rather than using assembler syntax. The $\jal$ (jump-and-link)
+instruction used for calls atomically saves the return address in a register
+and jumps to a specified new $\PCname$; the $\jalr$ (jump-and-link-via-register)
+instruction used for returns jumps to the address stored in a register.
+Register $r_0$ is hardwired to zero. \leo{Expand upon this?}
 
@@
 \[
-44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
+44 : [\rsp - 5] \leftarrow r_4
+%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
 \]
 
-Instruction 44 currently is $\sw ~ \rsp ~ r_4 ~ \negate 2$, with the
+Instruction 44 currently is
+$ [\rsp - 2] \leftarrow r_4$,
+%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
+with the
 effect of storing the contents of $r_4$ as the result of {\tt h}. If
-instead it wrote to location $[\rsp - 5]$ then it would be overriding
+instead it wrote to location $\rsp - 5$ then it would be overwriting
 private data (the return address) belonging to {\tt g}.
@@
 \[
-30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 8
+30 : r_5  \leftarrow [\rsp - 6]
+%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
 \]
 
-Instruction 30 currently is $\lw ~ r_5 ~ \rsp ~ \negate 6$, with the
-effect of loading the argument of {\tt g} to $r_4$. If instead it
-loaded the contents of location $[\rsp - 8]$, it would obtain access
-to {\tt main}'s private information. Worse, since the next instruction
-outputs the contents of $r_5$ that would immediately be made public.
+Instruction 30 currently is
+$ r_5 \leftarrow [\rsp - 4]$,
+%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
+with the
+effect of loading the argument of {\tt g} to $r_5$. If instead it
+loaded the contents of location $\rsp - 6$, it would obtain access
+to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
+outputs the contents of $r_5$, that value would immediately be made public.
 
@@
 \[
-5 : \jalr ~ r_4 ~ \rra ~ 0
+5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
+%5 : \jalr ~ r_4 ~ \rra ~ 0
 \]
-If a jump instruction was added to {\tt main} with its target being
-the middle of the code block of {\tt h}.
+If a {\jalr} instruction was added to {\tt main} with its target being
+the middle of the code block of {\tt h}.\apt{describe result of this?}
 
@@
 $\cdots$
-\memory{1}{white}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
+\memory{1}{\emptyoutc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
 \memory{2}{\fprivc}[{\makebox[0pt]{$(\HIGHSEC, \HIGHINT)$}}]%

### 2020-07-08 main.tex
Filling in policy discussion and more tweaks where needed

@@
                         %% http://ctan.org/pkg/subcaption
-
 \input{macros}
@@
 \leo{I agree there is a collision, not sure about ``substitutions''}
+\sna{Mutations, drawing from the randomized testing literature?}
 
@@
 formally capturing what stack-safe means formally and discussing
-enforcement techniques from the literature in later sections.
+enforcement techniques from the literature in later sections. Throughout
+the paper examples will be localize modifications of the example, by
+substituting an instruction at a given PC, written:
+\[\PCname :\mathit{inst}\]
 
@@
 
-\bcp{More explanation needed, plus examples.}A policy is a general model of an enforcement mechanism, consisting of
-a set of policy states \(\pol \in \POLS\), and a policy step function
+\bcp{More explanation needed, plus examples.}
+A policy is a general model of an enforcement mechanism, consisting of a
+set of policy states \(\pol \in \POLS\), a policy step function
 \((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
-\rightharpoonup \POLS\).
+\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
+\POLS\). The initializer establishes the initial policy state for a given
+machine state, but is partial to reflect that some policies are only defined
+on states that are well-formed in some way.
 
@@
 
-\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1) \stepstopol \pol_2\)}
-            {\(\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}\)}
-
-\leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted
-  step?}\bcp{Yes.}
+\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
+               \polstep \pol_2\)}
+            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
+               \mpstate{\mach_2}{\pol_2}\)}
 
@@
 policy step function is the constant function that produces the unit, $\lambda
-\_ . \top$, thereby allowing all calls the machine step function. This
-corresponds to the lifting of our machine model to the policy setting.
+\_ . \top$, thereby allowing all calls the machine step function. Its
+initialization function is defined on all states. This corresponds to the
+lifting of our machine model to the policy setting. Machines that fulfill a
+property by construction will also fulfill that property with the null
+policy..
 
@@
 %
+Some policies are implemented in software, by modifying code in the compiler
+to (e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
+canaries \citep{Cowan+98}. These approaches are represented as policies that
+accept only initial machine states that are well-formed for the policy (due
+to being produced by the compiler.) The set of policy states \(\POLS\) is again
+the unit type and the step function constant. Property enforcement comes from
+the fact that only ill-formed states can induce a trace that violates the
+property.
+
 \rb{TODO In a way, in these schemes the policy state is built in as part of the
@@
 %
+Micropolicies are another major family of policies \citep{pump_oakland2015},
+which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
+Here the machine state is extended with tags paired with all components of the
+system, and a monitor checks for each operation that the tags on the operands
+obey a set of rules. One could treat tags directly as part of the machine
+state, and build the rules into the step function. But local state
+encapsulation properties as we frame them are not intuitive in their
+interaction with tags or other additional enforcement state, so it is helpful
+to separate tags for clarity. This also distinguishes a hardware fail-stop from
+a software fail state.
 \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
   policy? }
+
+\paragraph*{Machine Termination}
+
+In our model, the machine step function is total -- a program does not halt
+on its own, but may instead enter a silent terminal loop. Runs are only cut
+short by the policy step function. This means that software-only policies that
+can catch errors must be modeled as reaching an error state, not outright
+halting.
 
@@
 to relate a machine-policy state pair to the MP-trace
-induced by by \(\stepstoobs{}_P\). Unlike the $\MOTRACE$ above, this trace can
+induced by \(\stepstopol{}\). Unlike the $\MOTRACE$ above, this trace can
 be finite as the policy can cause a fault, in which case the final state is
@@
 \begin{minipage}{.55\textwidth}
-\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
-              {\(\pol_0 \rightharpoonup \pol_1\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
+  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstoobs{\obs}
+                 \mpstate{\mach_1}{\pol_1}\)}
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow_P \MPT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P
+                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
 \end{minipage}
@@
 \item one state pair steps to the other:
-  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstoobs{\obs}_P (\mach_e, \pol_e),$$
+  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
 \item $\mach_c$ is a valid call with $n$ arguments:

### 2020-07-08 main.tex
Nit

@@
 
-\section{Stack Safety and Unsafety by Example}
+\section{Stack {\underline{Un}}safety by Example}
 \label{sec:running-example}
 
-We begin with a running example that will serve both to concretely
-demonstrate the ``stack-unsafe'' behaviours we intend to preclude and to
-ground and illustrate the formal properties and enforcement mechanisms in
-later sections.
+We begin with a running example that will serve both to demonstrate several
+``stack-unsafe'' behaviours that our stack safety property should prohibit
+and to ground the technical presentation in later sections.
 

### 2020-07-08 main.tex
A pass on sections 1-4
Small comments and nits, harmonized notation and spelling.

--- 2020-07-08 main.tex
@@
 technique.  In other words, we seek formal criteria that can be applied
-against \emph{any} stack-based system + enforcement mechanism to judge
+against \emph{any} stack-based system + \rb{in words or symbols? :-)} enforcement mechanism to judge
 whether or not the combination is actually stack-safe.
@@
 
-The intuition behind \citet{Skorstengaard+19}'s\iflater\bcp{I forget how to
+The intuition behind \citet{Skorstengaard+19}'s \iflater\bcp{I forget how to
   remove the year and just have the author names here...}\fi use of the term {\em
@@
 observable behavior is not affected by memory outside its stack frame
-(e.g. a caller's private data or uninitialized memory).  Conversely, {\em stack
+(e.g., a caller's private data or uninitialized memory).  Conversely, {\em stack
   integrity} promises that, if a function call ever returns, the caller's
@@
 We begin with a running example that will serve both to concretely
-demonstrate the ``stack-unsafe'' behaviours we intend to preclude and to
+demonstrate the ``stack-unsafe'' behaviors we intend to preclude and to
 ground and illustrate the formal properties and enforcement mechanisms in
@@
 %
-A sketch of the initial memory layout can be found in the top of
+A sketch of the initial memory layout can be found at the top of
 \cref{fig:running-trace-b}: instructions for {\tt main} begin at 0,
@@
 Stack safety is meant to protect a caller from a malicious callee's
-attempts to access private data or highjack its control flow. Our
+attempts to access private data or hijack its control flow. Our
 attacker model reflects this intention: we assume that the attacker
@@
 registers, its local stack space and its call frame. In the rest of
-this section we'll go over examples of unsafe behaviors, before
+this section we will go over examples of unsafe behaviors, before
 formally capturing what stack-safe means formally and discussing
 enforcement techniques from the literature in later sections. Throughout
-the paper examples will be localize modifications of the example, by
+the paper examples will be localized modifications of the example, by
 substituting an instruction at a given PC, written:
@@
 The choice of observations is an interesting one. A first approach
-would be to consider any change to the registers or memory observable.
+would be to consider any change to the registers or memory as observable.
 However, as some parts of the stack will be considered ``private''
 this choice won't work. A relaxation of this to consider all changes
-to ``non-private'' data observable could work in principle, but
-doesn't allow for an interesting weaker-but-more-efficient enforcement
-mechanism. We'll come back to this point in \Cref{sec:lazy}. On the
+to ``non-private'' data as observable could work in principle, but
+does not allow for an interesting weaker-but-more-efficient enforcement
+mechanism. We will come back to this point in \Cref{sec:lazy}. On the
 other hand, we could use a particular location as an abstract output
-of our system (e.g. through DMA). This is the approach we chose for
+of our system (e.g., through DMA). This is the approach we chose for
 our running example to keep things simple. For that, $\WORDS$ is the
@@
 property by construction will also fulfill that property with the null
-policy..
+policy.
 
@@
 observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.
-observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.
 
@@
 a \(\tau\):
+\rb{This $\hookrightarrow_P$ is often enough written just $\hookrightarrow$.
+Overload everywhere?}
 
@@
 takes a trace
-\(\Trace : \mathit{trace} \TRACEELEMS \times \OBSS\) and a predicate
+\(\Trace : \mathit{trace} ~ \TRACEELEMS \times \OBSS\) and a predicate
 on elements \(f \subset \TRACEELEMS\) and gives the prefix of
@@
 begin with our model of \leo{compiler-provided?} knowledge about the
-program structure, and then dive in to formal definitions local state
+program structure, and then dive into the formal definitions of local state
 encapsulation and well-bracketed control flow that together constitute
@@
 $\HIGHINT$ locations remain unchanged, while confidentiality would
-enforce that the contents of $\HIGHSEC$ locations don't influence the
+enforce that the contents of $\HIGHSEC$ locations do not influence the
 observable behavior of the machine. However, when dealing with nested
@@
 \definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
-a code map $\codemap$, an $n$-{\em transition} a pair of
+a code map $\codemap$, an $n$-{\em transition} is a pair of
 machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
@@
 shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
-in \high and low confidentiality in \low. Below it, the variant state
+in {\high} and low confidentiality in \low. Below it, the variant state
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
@@
 $n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
-$f$ in $\MPT$, the following holds. Let $\MPT_{call}$ be the prefix where
-\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} | \ret{\mach_c}\).
+$f$ in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
+\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
 For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
@@
 \(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
-\(\MPT_{call}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
+\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
 and \(R = \ret{\mach_c}\):
 
-$$\forall \MPT_{call} ~ \machT. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{call} ~ | ~ \ret{\mach_c} \Rightarrow$$
+$$\forall \MPT_{\mathit{call}} ~ \machT. ~
+\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
 $$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
 \mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{call} ~ \machT_{call}$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$
 
@@
 may hold on a wide range of programs that violate expected control
-flow. For example, if {\tt g} managed to jump in the middle of the
+flow. Say, if {\tt g} managed to jump in the middle of the
 code of {\tt h} with its call, confidentiality and integrity would
@@
 callee) should not be able to bypass a password check by jumping (or
-returning) in the middle of a block of code - even if such an execution
+returning) in the middle of a block of code --- even if such an execution
 would ostensibly satisfy local state encapsulation.
@@
 flows: every jump between functions should either be a call or a
-return, and each return should to the instruction after its
+return, and each return should be to the instruction after its
 corresponding call.
@@
 sources between functions, {\em entry integrity} ensures the validity
-of call targets, and {\tt return integrity} ensures the validity of
+of call targets, and {\em return integrity} ensures the validity of
 return targets.
@@
 
-\definition{Well Bracketed Control Flow}
+\definition{Well-Bracketed Control Flow}
 
@@
 corresponding to their call.  A system that enjoys them with regard to
-a code, a call, a return, and an entry map is said to enjoy {\em well
-  bracketed control flow} with regard to those maps.
+a code, a call, a return, and an entry map is said to enjoy
+{\em well-bracketed control flow} with regard to those maps.
 
@@
   machine, a contour and a return map:
-  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times 2^\MACHS)$$
+  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
   $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$

### 2020-07-08 main.tex

@@
 
-The intuition behind \citet{Skorstengaard+19}'s \iflater\bcp{I forget how to
+The intuition behind \citet{Skorstengaard+19}'s\iflater\bcp{I forget how to
   remove the year and just have the author names here...}\fi use of the term {\em
@@
 
-\section{Stack Safety and Unsafety by Example}
+\section{Stack {\underline{Un}}safety by Example}
 \label{sec:running-example}
 
-We begin with a running example that will serve both to concretely
-demonstrate the ``stack-unsafe'' behaviors we intend to preclude and to
-ground and illustrate the formal properties and enforcement mechanisms in
-later sections.
+We begin with a running example that will serve both to demonstrate several
+``stack-unsafe'' behaviors that our stack safety property should prohibit
+and to ground the technical presentation in later sections.
 
@@
 Stack safety is meant to protect a caller from a malicious callee's
-attempts to access private data or hijack its control flow. Our
+attempts to access private data or highjack its control flow. Our
 attacker model reflects this intention: we assume that the attacker

### 2020-07-08 main.tex
CP - intro

@@
 technique.  In other words, we seek formal criteria that can be applied
-against \emph{any} stack-based system + \rb{in words or symbols? :-)} enforcement mechanism to judge
+against \emph{any} stack-based system + enforcement mechanism to judge
 whether or not the combination is actually stack-safe.
@@
 \iflater
-\apt{another stab.}
-\leo{I took a stab. I think it's still to dense, but captures the
-  feeling of today's discussion I think}
-\apt{still needs tuning}\bcp{Yes, I think we need to go into a
-  little more detail on what their definition looks like in order to
-  justify why an alternative definition might be valuable.  What
-  questions does their presentation not answer? What uses does it not
-  support? E.g., can it be enforced efficiently? (I guess so, by
-  capabilities, though I think their capabilities are more powerful
-  than what Cheri actually provides, while our policies are exactly
-  what Nick needs.)  Why would it be good to look at stack safety
-  through the lens of security?  (Because it is fundamentally a
-  security mechanism?)}  \rb{BCP idea: contrast intrinsic
-  vs. extrinsic property definition.  Remember to talk about spatial
-  vs temporal safety}
+\leo{Should this be its own paragraph?}
 \fi
-
-The key contribution of this paper is a novel\bcp{generic?} presentation of
+The key contribution of this paper is a novel, generic presentation of
 local state encapsulation and well-bracketed control flow using the
@@
 
-The intuition behind \citet{Skorstengaard+19}'s\iflater\bcp{I forget how to
-  remove the year and just have the author names here...}\fi use of the term {\em
+The intuition behind \citeauthor{Skorstengaard+19}'s use of the term {\em
   local state encapsulation} is that callers should be protected from
@@
 
-The intuition behind \citet{Skorstengaard+19}'s {\em well-bracketed control
+The intuition behind \citeauthor{Skorstengaard+19}'s {\em well-bracketed control
   flow} is
@@
 
-To phrase these two aspects of stack safety in terms of security properties
-requires an explicit attacker model. \bcp{This paragraph goes too fast: We
-  need to start by asking, informally, who is being protected and who is the
-attacker. Then we observe that real-world stack protection mechanisms tend
-to imagine attack scenarios where the callee gets confused by something it
-receives from the outside; to avoid getting into the details of how this
-happens, we simply assume that the callee can execute arbitrary code.}
-For simplicity, we choose one that is
-fairly strong: when stating the protections that a function can rely on when
-it makes a call, we assume that the callee can execute arbitrary
-instructions until it returns.
+To phrase these two aspects of stack safety in terms of security
+properties \leo{in terms of security properties or ``using security
+  language''?}  requires an explicit attacker model: an clear
+understanding of who is being protected and from what. Most existing
+stack protection mechanisms deal with scenarios where a callee gets
+influenced by external inputs and ends up executing potentially
+malicious code. To avoid the details of exactly how this happens and
+keep the presentation clear, we pick an even stronger model: when
+stating the protections that a function can rely on when it makes a
+call, we assume that the callee can execute arbitrary instructions
+until it returns.
 

### 2020-07-08 main.tex
Tackle problematic paragraph in intro

@@
 
-\bcp{Again, here, a bigger transition, more setup, and a longer discussion
-  are needed.  This is really the crux of what we are doing!}
+Adopting that model, however, poses another novel challege: not only
+does a caller need to be protected from its callee, but the callee
+itself needs to be protected from {\em its} callees: the notion of
+``attacker'' changes throughout the execution of a program!
+
+\bcp{Again, here, a bigger transition, more setup, and a longer
+  discussion are needed.  This is really the crux of what we are
+  doing!} \leo{Attempt}
+%
 Moreover, applying the standard language-based-security notions of
-integrity and confidentiality in the stack-safety setting raises a novel
-challenge:
-the notion of ``attacker'' changes during execution. Not only does a
-caller need to be protected from its callee, but the callee itself
-needs to be protected from {\em its} callees. We address this
-challenge by adjusting the integrity and confidentiality of different
-stack frames at each call or return transition between
-functions.  \bcp{that's way too terse!}
+integrity and confidentiality in the stack-safety setting raises a
+novel challenge: the notion of ``attacker'' changes during execution.
+Not only does a caller need to be protected from its callee, but the
+callee itself needs to be protected from {\em its} callees. We address
+this problem by using a similarly changing notion of accessibility for
+the different components of the machine: we adjust the integrity and
+confidentiality level of different stack locations and registers at
+every transition (call or return) between functions.  \bcp{that's way too
+  terse!}\leo{Slightly better, still needs a bit probably}
 %\apt{connection to ``attacker'' still vague.}  \rb{Mention

### 2020-07-08 main.tex
Actually add the changes

@@
 
-Adopting that model, however, poses another novel challege: not only
-does a caller need to be protected from its callee, but the callee
-itself needs to be protected from {\em its} callees: the notion of
-``attacker'' changes throughout the execution of a program!
-
-\bcp{Again, here, a bigger transition, more setup, and a longer
-  discussion are needed.  This is really the crux of what we are
-  doing!} \leo{Attempt}
-%
-Moreover, applying the standard language-based-security notions of
-integrity and confidentiality in the stack-safety setting raises a
-novel challenge: the notion of ``attacker'' changes during execution.
-Not only does a caller need to be protected from its callee, but the
-callee itself needs to be protected from {\em its} callees. We address
-this problem by using a similarly changing notion of accessibility for
-the different components of the machine: we adjust the integrity and
-confidentiality level of different stack locations and registers at
-every transition (call or return) between functions.  \bcp{that's way too
-  terse!}\leo{Slightly better, still needs a bit probably}
+Adopting that model in our setting, however, poses another novel
+challege: not only does a caller need to be protected from its callee,
+but the callee itself needs to be protected from {\em its}
+callees. That is, the notion of ``attacker'' changes throughout the
+execution of a program!  We address this problem by using a similarly
+changing notion of accessibility for the different components of the
+machine: we adjust the integrity and confidentiality level of
+different stack locations and registers at every transition (call or
+return) between functions.  \bcp{that's way too terse!}\leo{Slightly
+  better, still needs a bit probably}
 %\apt{connection to ``attacker'' still vague.}  \rb{Mention

### 2020-07-08 main.tex
Better presentation of the testing property

@@
 
-In such a system, all memory addresses, registers and the program counter are
-enriched with an abstract metadata tag, which can point to the representation of
-an arbitrary data structure. A micropolicy is thus defined by a collection of
-rules to be checked as each instruction executes and together define the
-transfer function of the security policy. Each rule inspects the tags associated
-to the inputs of an instruction (PC, registers, memory) as well as the
-instruction opcode itself, and produces one of two outcomes: either the
-instruction is allowed to execute (and tags for the new PC and its result are
-produced and assigned to those), or the machine fail-stops with a policy
+In such a system, all memory addresses and registers (including the program
+counter) are enriched with an abstract metadata tag, which can point to the
+representation of an arbitrary data structure. A micropolicy is thus defined by
+a collection of rules to be checked as each instruction executes and together
+define the transfer function of the security policy. Each rule inspects the tags
+associated to the inputs of an instruction (PC, registers, memory) as well as
+the instruction opcode itself, and produces one of two outcomes: either the
+instruction is allowed to execute (and generates and assigns tags for the new PC
+and the result of the operation), or the machine fail-stops with a policy
 violation.
@@
 maps tag inputs to outputs and a mechanism to trap misses to software handlers
-and thereby computes the transfer function of the software-defined policy.
+and thereby compute the transfer function of the software-defined policy.
 Previous work has shown that a wide range of micropolicies can be defined and
@@
 %
-The first micropolicy we present for eager stack safety is very similar to the
-Depth Isolation policy presented by Roessler and DeHon \rb{cite}. Their policies
-tag the stack memory with ownership information associated to each stack frame
-(and to each separate object inside that frames, a more finely grained access
-control that we do not need to consider), and registers holding stack pointers
-with matching access permission information. Beyond those capabilities, the
-compiler generates designated sequences of instructions with special tags that
-grant them permission to manage the stack.
+The first micropolicy we present to enforce the definition of stack safety
+developed in \cref{sec:lse-and-wbcf} is very similar to the Depth Isolation
+policy presented by \citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the
+stack memory with ownership information associated to each stack frame (and to
+each separate object inside that frames, a more finely grained access control
+that we do not need to consider here), and registers holding stack pointers with
+matching access permission information. Beyond those capabilities, the compiler
+generates designated sequences of instructions with special tags that grant them
+permission to manage the stack, say during calls and returns. \rb{Conceivably on
+alloc operations, etc., which we have not yet mentioned.}
 
@@
 arguments are passed on the stack \rb{Come back to this later?}. Initially, the
-entire stack is tagged as unused with $\tagNoDepth$ \rb{actually, the ``initial
-  frame'' should be tagged with its proper depth}, and the {\PCname} carries in
-its tag at all times the depth of the current activation, starting with
-$\tagPCDepth{0}$. The current activation depth needs to change exactly at those
-points in the program where the contour of the stack safety property changes:
+entire stack is tagged as unused with an $\tagNoDepth$ tag \rb{actually, the
+``initial frame'' should be tagged with its proper depth}, and the {\PCname}
+carries in its tag at all times the depth of the current activation, starting
+with a $\tagPCDepth{0}$ tag. The current activation depth needs to change
+exactly at those points in the program where the contour of the stack safety
+property changes:
 
@@
 
-\item From caller to callee, when the machine executes an instruction tagged as
-  belonging to the call map with $\tagCall$. At this point, the {\PCname} depth
-  is incremented to, say $\tagPCDepth{n}$; on function entry, the locations of
-  the new frame is marked $\tagStackDepth{n}$.
-
-\item From callee back to caller, when execution reaches the instruction
-  immediately following the original call instruction and the $\SP$ is the same
-  as at the time of that call. At this point, the {\PCname} is decremented;
-  previously, the locations of the old frame are restored to $\tagNoDepth$.
+\item From caller to callee, when the machine executes an instruction tagged
+  with $\tagCall$ as belonging to the call map. At this point, the current
+  {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
+  the locations of the new frame is marked $\tagStackDepth{n + 1}$.
+
+\item From callee back to caller, when execution reaches the corresponding
+  return instruction for the call instruction, as defined by
+  $\ret{\mach_c}\ \mach_r$. At this point, the {\PCname} tag is decremented.
+  Prior to this point, the locations of the old frame have been restored to
+  $\tagNoDepth$.
 
@@
 These sequences also enforce the $\SP$ discipline. When stack frames are of
-fixed size, it suffices to increment $\SP$ and give it with a special tag in the
-header sequence, which is later checked after decrementing $\SP$ during the exit
-sequence to authorize the return. The well-formedness of the blessed sequences
-can be checked statically. Only programs whose sequences are well-formed are
-protected by the micropolicy, although programs are still protected (by
-fail-stopping) even if the intended sequences are missing. By tracking the
-execution of the blessed sequences in the {\PCname} tag state, the micropolicy
-ensures that the entry sequence is executed correctly a declared call.
+fixed size, it suffices to increment $\SP$ and tag it with a dedicated tag in
+the header sequence, which is later checked after decrementing $\SP$ during the
+exit sequence to authorize the return. The well-formedness of the blessed
+sequences can be checked statically. Only programs whose sequences are
+well-formed are protected by the micropolicy, although programs are still
+protected (by fail-stopping) even if the intended sequences are missing. By
+tracking the execution of the blessed sequences in the tag state of the
+{\PCname} register, the micropolicy ensures that the entry sequences of calls
+declared in the call map are executed correctly.
 
@@
 %
-This conservative policy is can be shown to satisfy an eager \emph{testing
-property}, which enforces lockstep and performs integrity and confidentiality
-tests for a stack of running variants at each step.
-
-\definition{Eager integrity test}
+This conservative policy can be shown to satisfy an eager \emph{testing
+property}, which at each step of the computation performs integrity and
+confidentiality tests for a stack of running variants, which are kept in
+lockstep with the original machine.
+
+\definition{Stack integrity test}
   Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
-  satisfy the \emph{eager integrity test} if their high-integrity components
+  satisfy the \emph{stack integrity test} if their high-integrity components
   match:
@@
 
-\definition{Eager confidentiality test}
-  Given two machine states $\mach_1$ and $\mach_2$ with stepping states and
-  observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
-  \stepstoobs{\obs_2} \mach_2'$ these satisfy the \emph{eager confidentiality
-  test} if their observations and changed components coincide:
+\definition{Stack confidentiality test}
+  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
+  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
+  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
+    confidentiality test} if their observations and changed components coincide:
   \[
-    \mathit{conf}_T ~ \mach_1 ~ \obs_1 ~ \mach_1' ~ \mach_2 ~ \obs_2 ~ \mach_2' ~ \triangleq
+    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
     \obs_1 = \obs_2 \wedge \forall \component . ~
@@
 \apt{scoping of $\forall$ is ambiguous.}
-\rb{Fix notations, including for properties below.}
+\rb{Fix notations, including for properties below.
+  In particular, note that the last four arguments could be computed from the
+  first two and it may be clearer to elide them.}
 \newcommand{\vse}{\mathit{vse}}
@@
 \definition{Variant stacks}
-  The testing property keeps track of execution information on a stack whose
-  elements record an initial machine, an initial and current variant of the
-  machine, a contour and a return map:
+  The testing property keeps track of relevant information on a stack whose
+  elements represent variants of the machine by recording an initial machine
+  state, an initial and current variant of the machine, a contour and a return
+  map:
+%
   $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
@@
   $$\vs \in \VSS$$
-  Given the current machine state $\mach$, a stack $\vs$ is \emph{well-formed}
-  with respect to it if:
+%
+  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
+  respect to it if the stack is nonempty, the return map of the bottom element
+  is the false property, and for each element $\vse \in \vs$:
   \begin{itemize}
-  \item The current variant is indeed a variant of the current state.
-  \item The stack is nonempty and the bottom return map is the false property.
-  \item The current state is in the trace of every initial machine.
-  \item The variant state is in the trace of its own initial variant.
+  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
+  \item $\mach$ is in the trace $\machT$ of its initial machine,
+    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
+  \item Its variant state $\mathit{curr\_variant}(\vse)$
+    is in the trace of its initial variant,
+    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
   \item The return function is in the return map if not the last element. \rb{TODO}
-  \item All contours have public registers.
+  \item The registers of its contour $\contour(\vse)$ are public.
   \end{itemize}
+
+\definition{Variant stack test}
+%
+  Our testing property extends the basic tests above to whole variant stacks.
+  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
+  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
+  $\mach$ safisfies the \emph{variant stack test} if for each variant element
+  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
+  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
+  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
+  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
+  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
 
@@
 
-  \item If the machine fail-stops,
+  \item If the machine fail-stops, i.e.,
     $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
-    the property is safisfied for the singleton trace $(\mach, \pol)$ at any
+    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at any
     depth up to the size of $\vs$.
@@
   \item If the top variant of the stack (and only the top variant) returns,
-    the property is safisfied for the singleton trace $(\mach, \pol)$ at depth
+    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at depth
     equal to the size of the stack. \rb{Avoid confusion with return below}
 
-  \item If the machine takes a step
+  \item If (1) the machine takes a step,
     $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
-    that is neither a call nor a return, all
+    that is neither a call nor a return, (2) all
     variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
@@
     (these tests are required also for calls and returns, and omitted in those cases),
-    and the stepped $(\mach', \pol')$ and $\vs'$ satisfy the eager testing property
-    for a given depth $d$ \rb{omitting stack size condition?} and MP-trace $\MPT$,
-    then the property is also satisfied for the extension of the trace \rb{notation}
+    and (3) the post-step state $(\mach', \pol')$ and stack $\vs'$ satisfy the eager testing property
+    for a given depth $d$ \rb{omitting stack size condition} and MP-trace $\MPT$,
+    then the property is also satisfied for trace $(\mach, \pol) : \MPT$
     at depth $d$.

### 2020-07-08 main.tex
call map -> function map

@@
 \definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
-a code map $\codemap$, an $n$-{\em transition} is a pair of
+a function map $\codemap$, an $n$-{\em transition} is a pair of
 machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
@@
 \definition{Stack Integrity} Let $\callmap$ be a call map and
-$\codemap$ a code map. We say that a system satisfies {\em stack
+$\codemap$ a function map. We say that a system satisfies {\em stack
   integrity} with respect to $\callmap$ and $\codemap$ if, for every
@@
 \definition{Stack Confidentiality}
-Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
+Let $\callmap$ be a call map and $\codemap$ a function map. We say that a
 system satisfies {\em stack confidentiality} with respect to
@@
 security properties that rely on the underlying knowledge of the
-program structure (code maps, function maps, return maps, and entry
+program structure (call maps, function maps, return maps, and entry
 maps): {\em control separation} ensures the validity of transition
@@
 \definition{Eager testing property}
-  Let $\callmap$ a call map and $\codemap$ a code map. Given a state
+  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
   $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
@@
       A system enjoys {\em observable stack safety} with regard to a call map
-      and a code map if it enjoys both observable stack integrity and
+      and a function map if it enjoys both observable stack integrity and
       observable stack confidentiality with regard to them.

### 2020-07-08 main.tex
Reorg done

@@
 %
-We begin with a running example that we refer
-to throughout the paper (\cref{sec:running-example}) and some
-technical preliminaries (\cref{sec:prelim}).
-
-\section{Stack {\underline{Un}}safety by Example}
+We begin with a running example that we refer to throughout the paper
+(\cref{sec:running-example}), a description of our threat model
+(\cref{sec:threat}) and some technical preliminaries
+(\cref{sec:prelim}).
+
+\section{Running Example}
 \label{sec:running-example}
 
-We begin with a running example that will serve both to demonstrate several
-``stack-unsafe'' behaviors that our stack safety property should prohibit
-and to ground the technical presentation in later sections.
-
-\subsection{Running Example}
+\leo{We've said this literally in the previous sentence. What should this intro
+  be in the new reorg?}
+%We begin with a running example that will serve both to demonstrate several
+%``stack-unsafe'' behaviors that our stack safety property should prohibit
+%and to ground the technical presentation in later sections.
 
@@
 
-\subsection{Attacker Model}
+\section{Threat Model}
+
+\leo{Needs high level signposting, transitions}
 
@@
 
-\subsection{Unsafe Variations}
-\sna{Variations here collides with state variations. I propose substitutions.}
-\leo{I agree there is a collision, not sure about ``substitutions''}
-\sna{Mutations, drawing from the randomized testing literature?}
-
-The execution trace shown in \cref{fig:running-trace-b} is an example
-of a ``stack-safe'' execution: each process interacts only with the
-registers, its local stack space and its call frame. In the rest of
-this section we will go over examples of unsafe behaviors, before
-formally capturing what stack-safe means formally and discussing
-enforcement techniques from the literature in later sections. Throughout
-the paper examples will be localized modifications of the example, by
-substituting an instruction at a given PC, written:
-\[\PCname :\mathit{inst}\]
-
-\leo{TODO: Add citations of mitigation techniques}
-\rb{And for those techniques, explain how they reflect their motivating
-examples. What does each technique do for those examples? How are they
-connected to this work?}
-
-\paragraph*{Bad Write}
-
-\[
-44 : [\rsp - 5] \leftarrow r_4
-%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
-\]
-
-Instruction 44 currently is
-$ [\rsp - 2] \leftarrow r_4$,
-%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
-with the
-effect of storing the contents of $r_4$ as the result of {\tt h}. If
-instead it wrote to location $\rsp - 5$ then it would be overwriting
-private data (the return address) belonging to {\tt g}.
-
-\paragraph*{Bad Read}
-
-\[
-30 : r_5  \leftarrow [\rsp - 6]
-%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-\]
-
-Instruction 30 currently is
-$ r_5 \leftarrow [\rsp - 4]$,
-%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
-with the
-effect of loading the argument of {\tt g} to $r_5$. If instead it
-loaded the contents of location $\rsp - 6$, it would obtain access
-to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
-outputs the contents of $r_5$, that value would immediately be made public.
-
-\paragraph*{Bad Control Flow}
-
-\[
-5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
-%5 : \jalr ~ r_4 ~ \rra ~ 0
-\]
-If a {\jalr} instruction was added to {\tt main} with its target being
-the middle of the code block of {\tt h}.\apt{describe result of this?}
+\bcp{More explanation needed, plus examples.}
+A policy is a general model of an enforcement mechanism, consisting of a
+set of policy states \(\pol \in \POLS\), a policy step function
+\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
+\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
+\POLS\). The initializer establishes the initial policy state for a given
+machine state, but is partial to reflect that some policies are only defined
+on states that are well-formed in some way.
+
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
+%
+\[\MPS = \MACHS \times \POLS\]
+%
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
+               \polstep \pol_2\)}
+            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
+               \mpstate{\mach_2}{\pol_2}\)}
+
+\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
+in more detail that the intention is to capture any kind of dynamic
+enforcement mechanism and give as many names and citations as we can.}
+
+This model is flexible enough to capture any kind of mechanism used for dynamic
+enforcement of properties of interest, including the following representative
+families of policies.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+\paragraph*{Null policy}
+%
+The simplest possible policy is the neutral policy that carries no information
+and does nothing. Its set of policy states $\POLS$ is the unit type, and its
+policy step function is the constant function that produces the unit, $\lambda
+\_ . \top$, thereby allowing all calls the machine step function. Its
+initialization function is defined on all states. This corresponds to the
+lifting of our machine model to the policy setting. Machines that fulfill a
+property by construction will also fulfill that property with the null
+policy.
+
+\paragraph*{Software-only policies}
+%
+Some policies are implemented in software, by modifying code in the compiler
+to (e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
+canaries \citep{Cowan+98}. These approaches are represented as policies that
+accept only initial machine states that are well-formed for the policy (due
+to being produced by the compiler.) The set of policy states \(\POLS\) is again
+the unit type and the step function constant. Property enforcement comes from
+the fact that only ill-formed states can induce a trace that violates the
+property.
+
+\rb{TODO In a way, in these schemes the policy state is built in as part of the
+  program state and transparent to the hardware, although it could be simulated
+  by an overlay where enforcement is represented at the policy level.}
+
+\paragraph*{Capability machines}
+%
+\rb{TODO Here the policy is built into the hardware, although the extended ISA
+  could be split into a machine without capabilities and capabilities as the
+  policy.}
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
+
+\paragraph*{Micropolicies}
+%
+Micropolicies are another major family of policies \citep{pump_oakland2015},
+which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
+Here the machine state is extended with tags paired with all components of the
+system, and a monitor checks for each operation that the tags on the operands
+obey a set of rules. One could treat tags directly as part of the machine
+state, and build the rules into the step function. But local state
+encapsulation properties as we frame them are not intuitive in their
+interaction with tags or other additional enforcement state, so it is helpful
+to separate tags for clarity. This also distinguishes a hardware fail-stop from
+a software fail state.
+\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+  policy? }
+
+% 
+%\leo{TODO: Add citations of mitigation techniques}
+%\rb{And for those techniques, explain how they reflect their motivating
+%examples. What does each technique do for those examples? How are they
+%connected to this work?}
+% 
+%\paragraph*{Bad Write}
+% 
+%\[
+%44 : [\rsp - 5] \leftarrow r_4
+%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
+%\]
+% 
+%Instruction 44 currently is
+%$ [\rsp - 2] \leftarrow r_4$,
+%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
+%with the
+%effect of storing the contents of $r_4$ as the result of {\tt h}. If
+%instead it wrote to location $\rsp - 5$ then it would be overwriting
+%private data (the return address) belonging to {\tt g}.
+% 
+%\paragraph*{Bad Read}
+% 
+%\[
+%30 : r_5  \leftarrow [\rsp - 6]
+%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+%\]
+% 
+%Instruction 30 currently is
+%$ r_5 \leftarrow [\rsp - 4]$,
+%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
+%with the
+%effect of loading the argument of {\tt g} to $r_5$. If instead it
+%loaded the contents of location $\rsp - 6$, it would obtain access
+%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
+%outputs the contents of $r_5$, that value would immediately be made public.
+% 
+%\paragraph*{Bad Control Flow}
+% 
+%\[
+%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
+%%5 : \jalr ~ r_4 ~ \rra ~ 0
+%\]
+%If a {\jalr} instruction was added to {\tt main} with its target being
+%the middle of the code block of {\tt h}.\apt{describe result of this?}
 
@@
 written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.
-
-\subsection{Policies}
-
-\bcp{More explanation needed, plus examples.}
-A policy is a general model of an enforcement mechanism, consisting of a
-set of policy states \(\pol \in \POLS\), a policy step function
-\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
-\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
-\POLS\). The initializer establishes the initial policy state for a given
-machine state, but is partial to reflect that some policies are only defined
-on states that are well-formed in some way.
-
-We write $\MPS$ for the set of all pairs of machine states and policy
-states.
-%
-\[\MPS = \MACHS \times \POLS\]
-%
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
-               \polstep \pol_2\)}
-            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
-               \mpstate{\mach_2}{\pol_2}\)}
-
-\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
-in more detail that the intention is to capture any kind of dynamic
-enforcement mechanism and give as many names and citations as we can.}
-
-This model is flexible enough to capture any kind of mechanism used for dynamic
-enforcement of properties of interest, including the following representative
-families of policies.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-\paragraph*{Null policy}
-%
-The simplest possible policy is the neutral policy that carries no information
-and does nothing. Its set of policy states $\POLS$ is the unit type, and its
-policy step function is the constant function that produces the unit, $\lambda
-\_ . \top$, thereby allowing all calls the machine step function. Its
-initialization function is defined on all states. This corresponds to the
-lifting of our machine model to the policy setting. Machines that fulfill a
-property by construction will also fulfill that property with the null
-policy.
-
-\paragraph*{Software-only policies}
-%
-Some policies are implemented in software, by modifying code in the compiler
-to (e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
-canaries \citep{Cowan+98}. These approaches are represented as policies that
-accept only initial machine states that are well-formed for the policy (due
-to being produced by the compiler.) The set of policy states \(\POLS\) is again
-the unit type and the step function constant. Property enforcement comes from
-the fact that only ill-formed states can induce a trace that violates the
-property.
-
-\rb{TODO In a way, in these schemes the policy state is built in as part of the
-  program state and transparent to the hardware, although it could be simulated
-  by an overlay where enforcement is represented at the policy level.}
-
-\paragraph*{Capability machines}
-%
-\rb{TODO Here the policy is built into the hardware, although the extended ISA
-  could be split into a machine without capabilities and capabilities as the
-  policy.}
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
-
-\paragraph*{Micropolicies}
-%
-Micropolicies are another major family of policies \citep{pump_oakland2015},
-which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
-Here the machine state is extended with tags paired with all components of the
-system, and a monitor checks for each operation that the tags on the operands
-obey a set of rules. One could treat tags directly as part of the machine
-state, and build the rules into the step function. But local state
-encapsulation properties as we frame them are not intuitive in their
-interaction with tags or other additional enforcement state, so it is helpful
-to separate tags for clarity. This also distinguishes a hardware fail-stop from
-a software fail state.
-\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-  policy? }
 

### 2020-07-08 main.tex
More formal stack testing, checkpoint

--- 2020-07-08 main.tex
@@
   \end{itemize}
+%
+  A variant stack steps if \rb{\ldots, explain, fix notation}, $\vs
+  \hookrightarrow \vs'$.
 
@@
 
-\definition{Eager testing property}
-  Let $\callmap$ a call map and $\codemap$ a code map. Given a state
-  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
-  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
-  testing property} at a given call depth if:
+\definition{Coinductive stack testing property}
+%
+  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
+  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
+  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
+  testing property} is satisfied at call depth $d$, and write
+  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:
 
@@
 
-  \item If the machine fail-stops, i.e.,
-    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
-    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at any
-    depth up to the size of $\vs$.
+  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
+    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
+    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
+    $\vs$. (That is, a violation was detected $d$ levels into the stack.
+    Elements are created and pushed to the variant stack on calls and popped
+    from it on returns, so its size always reflects the depth of the program
+    stack.)
 
   \item If the top variant of the stack (and only the top variant) returns,
-    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at depth
-    equal to the size of the stack. \rb{Avoid confusion with return below}
-
-  \item If (1) the machine takes a step,
-    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
-    that is neither a call nor a return, (2) all
-    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
-    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
-    (these tests are required also for calls and returns, and omitted in those cases),
-    and (3) the post-step state $(\mach', \pol')$ and stack $\vs'$ satisfy the eager testing property
-    for a given depth $d$ \rb{omitting stack size condition} and MP-trace $\MPT$,
-    then the property is also satisfied for trace $(\mach, \pol) : \MPT$
-    at depth $d$.
-
-  \item If the machine takes a call step
-    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
-    the contour $\contour$ is updated based on the top of the current machine.
-    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
-    (and in addition to the general step conditions),
-    the eager integrity test must hold for $\contour$ at the call,
-    the eager confidentiality test must hold for all variants at the call,
-    including the new variant that is added to the top of the stack,
-    and the property recurses on an arbitrary variant at the new contour
-    at depth $d$ with trace $\MPT$.
-
-  \item If the machine takes a return step at the top of the variant stack
+    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
+    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
+    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
+    use of this vs the return case below.}
+
+  \item If
+    %
+    (1) the machine takes a \emph{non-call, non-return step} $(\mach,
+    \pol) \hookrightarrow_P (\mach', \pol', \obs)$ and the variant stack steps
+    to $\vs \hookrightarrow \vs'$,
+    %
+    (2) the variant stack test $\mathit{\vs}_T ~
+    (\mach, \pol) ~ \vs$ is satisfied, and
+    %
+    (3) the stack safety test
+    $\mathit{stack}_T ~ d ~ \MPT ~ \vs'$ is coinductively satisfied at depth $d$
+    and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
+    \pol')$,
+    %
+    {\em then} we conclude $\mathit{stack}_T d ~ ((\mach, \pol) : \MPT) ~ vs$.
+    \rb{omitting stack size condition}
+
+  \item
+%
+    If (1) the machine takes a \emph{call step} $(\mach, \pol) \hookrightarrow_P
+    (\mach', \pol', \obs)$ and the variant stack steps to $\vs \hookrightarrow
+    \vs'$,
+%
+    (2) $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$,
+%
+    (3) for the new contour at the call, $\contour$, \rb{definition?} the
+    machine satisfies the stack integrity test $\mathit{int}_T ~ \contour ~
+    \mach ~ \mach'$ and the stack confidentiality test with regard to any
+    variant $\mathit{variant} ~ C ~ \mach ~ \mach_{\mathit{var}}$,
+    $\mathit{conf}_T ~ \mach ~ \mach_{\mathit{var}}$ \rb{definition?},
+%
+    and (4) the stack safety test is coinductively satisfied at depth $d$ for a
+    continuing trace $\MPT$ where $\head(\MPT) = (\mach', \pol')$ after an
+    arbitrary variant machine derived from $\contour$ has been added to the top
+    of the stack, $\mathit{stack}_T ~ d ~ \MPT ~ (\vse : \vs')$, \rb{define
+      $\vse$}
+%
+    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : \MPT) ~ vs$.
+
+    %% the contour $\contour$ is updated based on the top of the current machine.
+    %% For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    %% the eager integrity test must hold for $\contour$ at the call,
+    %% the eager confidentiality test must hold for all variants at the call,
+    %% including the new variant that is added to the top of the stack,
+    %% and the property recurses on an arbitrary variant at the new contour
+    %% at depth $d$ with trace $\MPT$.
+
+  \item If the machine takes a \emph{return step} at the top of the variant stack
     $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
@@
 
-  \item If the machine takes a return step that is also a call, the last two
-    cases are combined.
+  \item If the machine takes \emph{simultaneous return-call step}, that is, the
+    instruction to which execution returns immediately after completing a call
+    is itself a call instruction, the last two cases are combined.
 
   \end{itemize}
+
+\definition{Top-level stack testing property}
+%
+  \ldots wrap the coinductive stack testing property
 

### 2020-07-08 main.tex

--- 2020-07-08 main.tex
@@
 
+%  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
+%  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
+%  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
+%  testing property} is satisfied at call depth $d$, and write
+%  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:
+
   \begin{itemize}
 
-  \item If the machine fail-stops, i.e.,
-    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
-    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at any
-    depth up to the size of $\vs$.
+  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
+    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
+    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
+    $\vs$. (That is, a violation was detected $d$ levels into the stack.
+    Elements are created and pushed to the variant stack on calls and popped
+    from it on returns, so its size always reflects the depth of the program
+    stack.)
 
   \item If the top variant of the stack (and only the top variant) returns,
-    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at depth
-    equal to the size of the stack. \rb{Avoid confusion with return below}
-
-  \item If (1) the machine takes a step,
-    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
-    that is neither a call nor a return, (2) all
-    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
-    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
-    (these tests are required also for calls and returns, and omitted in those cases),
-    and (3) the post-step state $(\mach', \pol')$ and stack $\vs'$ satisfy the eager testing property
-    for a given depth $d$ \rb{omitting stack size condition} and MP-trace $\MPT$,
-    then the property is also satisfied for trace $(\mach, \pol) : \MPT$
-    at depth $d$.
-
-  \item If the machine takes a call step
-    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
-    the contour $\contour$ is updated based on the top of the current machine.
-    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
-    (and in addition to the general step conditions),
-    the eager integrity test must hold for $\contour$ at the call,
-    the eager confidentiality test must hold for all variants at the call,
-    including the new variant that is added to the top of the stack,
-    and the property recurses on an arbitrary variant at the new contour
-    at depth $d$ with trace $\MPT$.
-
-  \item If the machine takes a return step at the top of the variant stack
+    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
+    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
+    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
+    use of this vs the return case below.}
+
+  \item If
+    %
+    (1) the machine takes a \emph{non-call, non-return step} $(\mach,
+    \pol) \hookrightarrow_P (\mach', \pol', \obs)$ and the variant stack steps
+    to $\vs \hookrightarrow \vs'$,
+    %
+    (2) the variant stack test $\mathit{\vs}_T ~
+    (\mach, \pol) ~ \vs$ is satisfied, and
+    %
+    (3) the stack safety test
+    $\mathit{stack}_T ~ d ~ \MPT ~ \vs'$ is coinductively satisfied at depth $d$
+    and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
+    \pol')$,
+    %
+    {\em then} we conclude $\mathit{stack}_T d ~ ((\mach, \pol) : \MPT) ~ vs$.
+    \rb{omitting stack size condition}
+
+  \item
+%
+    If (1) the machine takes a \emph{call step} $(\mach, \pol) \hookrightarrow_P
+    (\mach', \pol', \obs)$ and the variant stack steps to $\vs \hookrightarrow
+    \vs'$,
+%
+    (2) $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$,
+%
+    (3) for the new contour at the call, $\contour$, \rb{definition?} the
+    machine satisfies the stack integrity test $\mathit{int}_T ~ \contour ~
+    \mach ~ \mach'$ and the stack confidentiality test with regard to any
+    variant $\mathit{variant} ~ C ~ \mach ~ \mach_{\mathit{var}}$,
+    $\mathit{conf}_T ~ \mach ~ \mach_{\mathit{var}}$ \rb{definition?},
+%
+    and (4) the stack safety test is coinductively satisfied at depth $d$ for a
+    continuing trace $\MPT$ where $\head(\MPT) = (\mach', \pol')$ after an
+    arbitrary variant machine derived from $\contour$ has been added to the top
+    of the stack, $\mathit{stack}_T ~ d ~ \MPT ~ (\vse : \vs')$, \rb{define
+      $\vse$}
+%
+    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : \MPT) ~ vs$.
+
+    %% the contour $\contour$ is updated based on the top of the current machine.
+    %% For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    %% the eager integrity test must hold for $\contour$ at the call,
+    %% the eager confidentiality test must hold for all variants at the call,
+    %% including the new variant that is added to the top of the stack,
+    %% and the property recurses on an arbitrary variant at the new contour
+    %% at depth $d$ with trace $\MPT$.
+
+  \item If the machine takes a \emph{return step} at the top of the variant stack
     $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,

### 2020-07-08 main.tex
pass through 4; releasing lock on 4

--- 2020-07-08 main.tex
@@
 \section{Threat Model}
+\label{sec:threat}
 
@@
 
-\bcp{More explanation needed, plus examples.}
-A policy is a general model of an enforcement mechanism, consisting of a
-set of policy states \(\pol \in \POLS\), a policy step function
-\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
-\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
-\POLS\). The initializer establishes the initial policy state for a given
-machine state, but is partial to reflect that some policies are only defined
-on states that are well-formed in some way.
-
-We write $\MPS$ for the set of all pairs of machine states and policy
-states.
-%
-\[\MPS = \MACHS \times \POLS\]
-%
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
-               \polstep \pol_2\)}
-            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
-               \mpstate{\mach_2}{\pol_2}\)}
-
-\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
-in more detail that the intention is to capture any kind of dynamic
-enforcement mechanism and give as many names and citations as we can.}
+\apt{Informal definition of policies as auxiliary state here.}
 
@@
 enforcement mechanism based on
-\citet{DBLP:conf/sp/RoesslerD18}.
+\citet{DBLP:conf/sp/RoesslerD18}.\apt{needs a relook}
 
@@
 Our machine states are composed of {\em components} $k$, addressed either by
-words (memory addresses) or by by register names ($r$) drawn from some set
+words (memory addresses) or by register names ($r$) drawn from some set
 $\REGS$.
@@
 silent (written $\tau$) or a word of data that the machine
-communicates to the outside world (e.g., by writing to a DMA
-address):
+communicates to the outside world:
 %
@@
 
-The choice of observations is an interesting one. A first approach
-would be to consider any change to the registers or memory as observable.
-However, as some parts of the stack will be considered ``private''
-this choice won't work. A relaxation of this to consider all changes
-to ``non-private'' data as observable could work in principle, but
-does not allow for an interesting weaker-but-more-efficient enforcement
-mechanism. We will come back to this point in \Cref{sec:lazy}. On the
-other hand, we could use a particular location as an abstract output
-of our system (e.g., through DMA). This is the approach we chose for
-our running example to keep things simple. For that, $\WORDS$ is the
-set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
-\cup \{\PCname\}$, and $\SP = r_2$. Observations are the writes to the
-special $\rout$ location: if during an execution step a word $w$ is
+Notice that we choose to model observations explicitly, rather than simply treating
+a subset of registers or memory as being observable.  This choice gives
+us the flexibility to treat programs as observationally equivalent even when
+their internal states may differ.
+%The choice of modelling observations is an interesting one. A first approach
+%would be to consider any change to the registers or memory as observable.
+%However, as some parts of the stack will be considered ``private''
+%this choice won't work. A relaxation of this to consider all changes
+%to ``non-private'' data as observable could work in principle, but
+%does not allow for an interesting weaker-but-more-efficient enforcement
+%mechanism. We will come back to this point in \Cref{sec:lazy}.
+
+In the running example of \Cref{sec:running-example}, 
+$\WORDS$ is the
+set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
+\cup \{\PCname,\SP\}$. Observations are the writes to the
+  special $\rout$ location, which can be thought of as a memory-mapped
+  output port: if during an execution step a word $w$ is
 written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.
+
+\subsection{Policies}
+
+A policy is a general model of an enforcement mechanism, consisting of a
+set of policy states \(\pol \in \POLS\), a policy step function
+\((\mach, \pol) \polstep \pol' : \MACHS \times \POLS
+\rightharpoonup \POLS\), and an initializer \(\pinit : \MACHS \rightharpoonup
+\POLS\). The initializer establishes the initial policy state for a given
+machine state, but is partial to reflect that some policies are only defined
+on states that are well-formed in some way.\apt{Is this our current thinking?}
+
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
+%
+\[\MPS = \MACHS \times \POLS\]
+%
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
+               \polstep \pol_2\)}
+            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
+               \mpstate{\mach_2}{\pol_2}\)}
+
+\apt{TODO: relate back to \Cref{sec:threat} and forward to \Cref{sec:enforcement}.}
+%\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
+%in more detail that the intention is to capture any kind of dynamic
+%enforcement mechanism and give as many names and citations as we can.}
 
@@
 short by the policy step function. This means that software-only policies that
-can catch errors must be modeled as reaching an error state, not outright
-halting.
+can catch errors must be modeled as reaching an error state\apt{??}, not 
+halting outright.\apt{Is this our current thinking?}
 
@@
 
-In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
+In a similar fashion, we define \(\mpstatename \hookrightarrow \MPT\)
 to relate a machine-policy state pair to the MP-trace
-induced by \(\stepstopol{}\). Unlike the $\MOTRACE$ above, this trace can
-be finite as the policy can cause a fault, in which case the final state is
+induced by the step function on $\MPS$. Unlike the $\MOTRACE$ above, this trace can
+be finite as the policy can cause a fault, in which case the final observation is
 a \(\tau\):
 \rb{This $\hookrightarrow_P$ is often enough written just $\hookrightarrow$.
-Overload everywhere?}
+  Overload everywhere?}\apt{Done, and for $\rightharpoonup$ too.}
 
 \begin{minipage}{.4\textwidth}
-\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
-         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
+\judgment{\(\neg \exists \pol_1 . \mpstate{\mach_0}{\pol_0} \polstep \pol_1\)}
+         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
 \end{minipage}
 \begin{minipage}{.55\textwidth}
-  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstoobs{\obs}
+  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstopol{\obs}
                  \mpstate{\mach_1}{\pol_1}\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow_P \MPT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow
                  \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
@@
 \paragraph*{Until}
-The ``Until'' operation \(\PUT(\Trace,f)\)
+The operation \(\PUT~(\Trace,f)\)
 %\bcp{Wouldn't
@@
     \judgment{\(f\ \traceelem\)}
-             {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
+             {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
   \end{minipage}
@@
   \judgment{}
-           {\(\PUT(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
+           {\(\PUT~(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
 %
@@
   \begin{minipage}{.3\textwidth}
-  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
-              {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
+  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT~(f,\Trace) = \Trace'\)}
+              {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
   \end{minipage}
@@
 \judgmenttwo{\(\mach \hookrightarrow \machT'\)}
-            {\(\PUT(f,\machT')
+            {\(\PUT~(f,\machT')
               = \machT\)}
@@
 \judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
-            {\(\PUT(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
+            {\(\PUT~(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
          {\(\mpstatename \hookrightarrow \MPT | f\)}
@@
 non-silent observations of $\obsT_1$ is a prefix of those of
-$\obsT_2$; that is we operate up to deletion of \(\tau\) observations,
+$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
 coinductively:
@@
 similarity of observation traces could alternatively be defined coinductively,
-just like the prefix relation, but by dropping the assymetric rules that allow
-for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$)
-
-In addition, perhaps surprisingly, an infinite silent trace is a
+just like the prefix relation, but by dropping the asymmetric rules that allow
+for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}
+
+Note that an infinite silent trace is a
 prefix of (and equivalent to) any other trace. While this might seem

### 2020-07-08 main.tex
Overloading hookrightarrow

--- 2020-07-08 main.tex
@@
 %
-We begin with a running example that we refer to throughout the paper
-(\cref{sec:running-example}), a description of our threat model
-(\cref{sec:threat}) and some technical preliminaries
-(\cref{sec:prelim}).
-
-\section{Running Example}
+We begin with a running example that we refer
+to throughout the paper (\cref{sec:running-example}) and some
+technical preliminaries (\cref{sec:prelim}).
+
+\section{Stack {\underline{Un}}safety by Example}
 \label{sec:running-example}
 
-\leo{We've said this literally in the previous sentence. What should this intro
-  be in the new reorg?}
-%We begin with a running example that will serve both to demonstrate several
-%``stack-unsafe'' behaviors that our stack safety property should prohibit
-%and to ground the technical presentation in later sections.
+We begin with a running example that will serve both to demonstrate several
+``stack-unsafe'' behaviors that our stack safety property should prohibit
+and to ground the technical presentation in later sections.
+
+\subsection{Running Example}
 
@@
 
-\section{Threat Model}
-
-\leo{Needs high level signposting, transitions}
+\subsection{Attacker Model}
 
@@
 
-\bcp{More explanation needed, plus examples.}
-A policy is a general model of an enforcement mechanism, consisting of a
-set of policy states \(\pol \in \POLS\), a policy step function
-\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
-\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
-\POLS\). The initializer establishes the initial policy state for a given
-machine state, but is partial to reflect that some policies are only defined
-on states that are well-formed in some way.
-
-We write $\MPS$ for the set of all pairs of machine states and policy
-states.
-%
-\[\MPS = \MACHS \times \POLS\]
-%
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
-               \polstep \pol_2\)}
-            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
-               \mpstate{\mach_2}{\pol_2}\)}
-
-\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
-in more detail that the intention is to capture any kind of dynamic
-enforcement mechanism and give as many names and citations as we can.}
-
-This model is flexible enough to capture any kind of mechanism used for dynamic
-enforcement of properties of interest, including the following representative
-families of policies.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-\paragraph*{Null policy}
-%
-The simplest possible policy is the neutral policy that carries no information
-and does nothing. Its set of policy states $\POLS$ is the unit type, and its
-policy step function is the constant function that produces the unit, $\lambda
-\_ . \top$, thereby allowing all calls the machine step function. Its
-initialization function is defined on all states. This corresponds to the
-lifting of our machine model to the policy setting. Machines that fulfill a
-property by construction will also fulfill that property with the null
-policy.
-
-\paragraph*{Software-only policies}
-%
-Some policies are implemented in software, by modifying code in the compiler
-to (e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
-canaries \citep{Cowan+98}. These approaches are represented as policies that
-accept only initial machine states that are well-formed for the policy (due
-to being produced by the compiler.) The set of policy states \(\POLS\) is again
-the unit type and the step function constant. Property enforcement comes from
-the fact that only ill-formed states can induce a trace that violates the
-property.
-
-\rb{TODO In a way, in these schemes the policy state is built in as part of the
-  program state and transparent to the hardware, although it could be simulated
-  by an overlay where enforcement is represented at the policy level.}
-
-\paragraph*{Capability machines}
-%
-\rb{TODO Here the policy is built into the hardware, although the extended ISA
-  could be split into a machine without capabilities and capabilities as the
-  policy.}
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
-
-\paragraph*{Micropolicies}
-%
-Micropolicies are another major family of policies \citep{pump_oakland2015},
-which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
-Here the machine state is extended with tags paired with all components of the
-system, and a monitor checks for each operation that the tags on the operands
-obey a set of rules. One could treat tags directly as part of the machine
-state, and build the rules into the step function. But local state
-encapsulation properties as we frame them are not intuitive in their
-interaction with tags or other additional enforcement state, so it is helpful
-to separate tags for clarity. This also distinguishes a hardware fail-stop from
-a software fail state.
-\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-  policy? }
-
-% 
-%\leo{TODO: Add citations of mitigation techniques}
-%\rb{And for those techniques, explain how they reflect their motivating
-%examples. What does each technique do for those examples? How are they
-%connected to this work?}
-% 
-%\paragraph*{Bad Write}
-% 
-%\[
-%44 : [\rsp - 5] \leftarrow r_4
-%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
-%\]
-% 
-%Instruction 44 currently is
-%$ [\rsp - 2] \leftarrow r_4$,
-%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
-%with the
-%effect of storing the contents of $r_4$ as the result of {\tt h}. If
-%instead it wrote to location $\rsp - 5$ then it would be overwriting
-%private data (the return address) belonging to {\tt g}.
-% 
-%\paragraph*{Bad Read}
-% 
-%\[
-%30 : r_5  \leftarrow [\rsp - 6]
-%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-%\]
-% 
-%Instruction 30 currently is
-%$ r_5 \leftarrow [\rsp - 4]$,
-%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
-%with the
-%effect of loading the argument of {\tt g} to $r_5$. If instead it
-%loaded the contents of location $\rsp - 6$, it would obtain access
-%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
-%outputs the contents of $r_5$, that value would immediately be made public.
-% 
-%\paragraph*{Bad Control Flow}
-% 
-%\[
-%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
-%%5 : \jalr ~ r_4 ~ \rra ~ 0
-%\]
-%If a {\jalr} instruction was added to {\tt main} with its target being
-%the middle of the code block of {\tt h}.\apt{describe result of this?}
+\subsection{Unsafe Variations}
+\sna{Variations here collides with state variations. I propose substitutions.}
+\leo{I agree there is a collision, not sure about ``substitutions''}
+\sna{Mutations, drawing from the randomized testing literature?}
+
+The execution trace shown in \cref{fig:running-trace-b} is an example
+of a ``stack-safe'' execution: each process interacts only with the
+registers, its local stack space and its call frame. In the rest of
+this section we will go over examples of unsafe behaviors, before
+formally capturing what stack-safe means formally and discussing
+enforcement techniques from the literature in later sections. Throughout
+the paper examples will be localized modifications of the example, by
+substituting an instruction at a given PC, written:
+\[\PCname :\mathit{inst}\]
+
+\leo{TODO: Add citations of mitigation techniques}
+\rb{And for those techniques, explain how they reflect their motivating
+examples. What does each technique do for those examples? How are they
+connected to this work?}
+
+\paragraph*{Bad Write}
+
+\[
+44 : [\rsp - 5] \leftarrow r_4
+%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
+\]
+
+Instruction 44 currently is
+$ [\rsp - 2] \leftarrow r_4$,
+%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
+with the
+effect of storing the contents of $r_4$ as the result of {\tt h}. If
+instead it wrote to location $\rsp - 5$ then it would be overwriting
+private data (the return address) belonging to {\tt g}.
+
+\paragraph*{Bad Read}
+
+\[
+30 : r_5  \leftarrow [\rsp - 6]
+%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+\]
+
+Instruction 30 currently is
+$ r_5 \leftarrow [\rsp - 4]$,
+%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
+with the
+effect of loading the argument of {\tt g} to $r_5$. If instead it
+loaded the contents of location $\rsp - 6$, it would obtain access
+to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
+outputs the contents of $r_5$, that value would immediately be made public.
+
+\paragraph*{Bad Control Flow}
+
+\[
+5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
+%5 : \jalr ~ r_4 ~ \rra ~ 0
+\]
+If a {\jalr} instruction was added to {\tt main} with its target being
+the middle of the code block of {\tt h}.\apt{describe result of this?}
 
@@
 
+\subsection{Policies}
+
+\bcp{More explanation needed, plus examples.}
+A policy is a general model of an enforcement mechanism, consisting of a
+set of policy states \(\pol \in \POLS\), a policy step function
+\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
+\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
+\POLS\). The initializer establishes the initial policy state for a given
+machine state, but is partial to reflect that some policies are only defined
+on states that are well-formed in some way.
+
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
+%
+\[\MPS = \MACHS \times \POLS\]
+%
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
+               \polstep \pol_2\)}
+            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
+               \mpstate{\mach_2}{\pol_2}\)}
+
+\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
+in more detail that the intention is to capture any kind of dynamic
+enforcement mechanism and give as many names and citations as we can.}
+
+This model is flexible enough to capture any kind of mechanism used for dynamic
+enforcement of properties of interest, including the following representative
+families of policies.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+\paragraph*{Null policy}
+%
+The simplest possible policy is the neutral policy that carries no information
+and does nothing. Its set of policy states $\POLS$ is the unit type, and its
+policy step function is the constant function that produces the unit, $\lambda
+\_ . \top$, thereby allowing all calls the machine step function. Its
+initialization function is defined on all states. This corresponds to the
+lifting of our machine model to the policy setting. Machines that fulfill a
+property by construction will also fulfill that property with the null
+policy.
+
+\paragraph*{Software-only policies}
+%
+Some policies are implemented in software, by modifying code in the compiler
+to (e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
+canaries \citep{Cowan+98}. These approaches are represented as policies that
+accept only initial machine states that are well-formed for the policy (due
+to being produced by the compiler.) The set of policy states \(\POLS\) is again
+the unit type and the step function constant. Property enforcement comes from
+the fact that only ill-formed states can induce a trace that violates the
+property.
+
+\rb{TODO In a way, in these schemes the policy state is built in as part of the
+  program state and transparent to the hardware, although it could be simulated
+  by an overlay where enforcement is represented at the policy level.}
+
+\paragraph*{Capability machines}
+%
+\rb{TODO Here the policy is built into the hardware, although the extended ISA
+  could be split into a machine without capabilities and capabilities as the
+  policy.}
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
+
+\paragraph*{Micropolicies}
+%
+Micropolicies are another major family of policies \citep{pump_oakland2015},
+which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
+Here the machine state is extended with tags paired with all components of the
+system, and a monitor checks for each operation that the tags on the operands
+obey a set of rules. One could treat tags directly as part of the machine
+state, and build the rules into the step function. But local state
+encapsulation properties as we frame them are not intuitive in their
+interaction with tags or other additional enforcement state, so it is helpful
+to separate tags for clarity. This also distinguishes a hardware fail-stop from
+a software fail state.
+\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+  policy? }
+
 \paragraph*{Machine Termination}
@@
 
-In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
+We similarly overload  \(\mpstatename \hookrightarrow \MPT\),
 to relate a machine-policy state pair to the MP-trace
@@
 a \(\tau\):
-\rb{This $\hookrightarrow_P$ is often enough written just $\hookrightarrow$.
-Overload everywhere?}
+%\rb{This $\hookrightarrow$ is often enough written just $\hookrightarrow$.
+%Overload everywhere?}
 
@@
 \judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
-         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
+         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
 \end{minipage}
@@
                  \mpstate{\mach_1}{\pol_1}\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow_P \MPT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow
                  \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
@@
 \definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
-a function map $\codemap$, an $n$-{\em transition} is a pair of
+a code map $\codemap$, an $n$-{\em transition} is a pair of
 machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
@@
 \definition{Stack Integrity} Let $\callmap$ be a call map and
-$\codemap$ a function map. We say that a system satisfies {\em stack
+$\codemap$ a code map. We say that a system satisfies {\em stack
   integrity} with respect to $\callmap$ and $\codemap$ if, for every
@@
 %
-$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
+$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow
 \MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
@@
 \definition{Stack Confidentiality}
-Let $\callmap$ be a call map and $\codemap$ a function map. We say that a
+Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
 system satisfies {\em stack confidentiality} with respect to
@@
 $f$ in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
-\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
+\(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} | \ret{\mach_c}\).
 For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
@@
 $$\forall \MPT_{\mathit{call}} ~ \machT. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
+\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
 $$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
@@
 security properties that rely on the underlying knowledge of the
-program structure (call maps, function maps, return maps, and entry
+program structure (code maps, function maps, return maps, and entry
 maps): {\em control separation} ensures the validity of transition
@@
 corresponding to a call in its induced trace, if
-\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
+\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
 is the first unmatched return such that for some $\pol_r$
@@
 
-In such a system, all memory addresses and registers (including the program
-counter) are enriched with an abstract metadata tag, which can point to the
-representation of an arbitrary data structure. A micropolicy is thus defined by
-a collection of rules to be checked as each instruction executes and together
-define the transfer function of the security policy. Each rule inspects the tags
-associated to the inputs of an instruction (PC, registers, memory) as well as
-the instruction opcode itself, and produces one of two outcomes: either the
-instruction is allowed to execute (and generates and assigns tags for the new PC
-and the result of the operation), or the machine fail-stops with a policy
+In such a system, all memory addresses, registers and the program counter are
+enriched with an abstract metadata tag, which can point to the representation of
+an arbitrary data structure. A micropolicy is thus defined by a collection of
+rules to be checked as each instruction executes and together define the
+transfer function of the security policy. Each rule inspects the tags associated
+to the inputs of an instruction (PC, registers, memory) as well as the
+instruction opcode itself, and produces one of two outcomes: either the
+instruction is allowed to execute (and tags for the new PC and its result are
+produced and assigned to those), or the machine fail-stops with a policy
 violation.
@@
 maps tag inputs to outputs and a mechanism to trap misses to software handlers
-and thereby compute the transfer function of the software-defined policy.
+and thereby computes the transfer function of the software-defined policy.
 Previous work has shown that a wide range of micropolicies can be defined and
@@
 %
-The first micropolicy we present to enforce the definition of stack safety
-developed in \cref{sec:lse-and-wbcf} is very similar to the Depth Isolation
-policy presented by \citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the
-stack memory with ownership information associated to each stack frame (and to
-each separate object inside that frames, a more finely grained access control
-that we do not need to consider here), and registers holding stack pointers with
-matching access permission information. Beyond those capabilities, the compiler
-generates designated sequences of instructions with special tags that grant them
-permission to manage the stack, say during calls and returns. \rb{Conceivably on
-alloc operations, etc., which we have not yet mentioned.}
+The first micropolicy we present for eager stack safety is very similar to the
+Depth Isolation policy presented by Roessler and DeHon \rb{cite}. Their policies
+tag the stack memory with ownership information associated to each stack frame
+(and to each separate object inside that frames, a more finely grained access
+control that we do not need to consider), and registers holding stack pointers
+with matching access permission information. Beyond those capabilities, the
+compiler generates designated sequences of instructions with special tags that
+grant them permission to manage the stack.
 
@@
 arguments are passed on the stack \rb{Come back to this later?}. Initially, the
-entire stack is tagged as unused with an $\tagNoDepth$ tag \rb{actually, the
-``initial frame'' should be tagged with its proper depth}, and the {\PCname}
-carries in its tag at all times the depth of the current activation, starting
-with a $\tagPCDepth{0}$ tag. The current activation depth needs to change
-exactly at those points in the program where the contour of the stack safety
-property changes:
+entire stack is tagged as unused with $\tagNoDepth$ \rb{actually, the ``initial
+  frame'' should be tagged with its proper depth}, and the {\PCname} carries in
+its tag at all times the depth of the current activation, starting with
+$\tagPCDepth{0}$. The current activation depth needs to change exactly at those
+points in the program where the contour of the stack safety property changes:
 
@@
 
-\item From caller to callee, when the machine executes an instruction tagged
-  with $\tagCall$ as belonging to the call map. At this point, the current
-  {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
-  the locations of the new frame is marked $\tagStackDepth{n + 1}$.
-
-\item From callee back to caller, when execution reaches the corresponding
-  return instruction for the call instruction, as defined by
-  $\ret{\mach_c}\ \mach_r$. At this point, the {\PCname} tag is decremented.
-  Prior to this point, the locations of the old frame have been restored to
-  $\tagNoDepth$.
+\item From caller to callee, when the machine executes an instruction tagged as
+  belonging to the call map with $\tagCall$. At this point, the {\PCname} depth
+  is incremented to, say $\tagPCDepth{n}$; on function entry, the locations of
+  the new frame is marked $\tagStackDepth{n}$.
+
+\item From callee back to caller, when execution reaches the instruction
+  immediately following the original call instruction and the $\SP$ is the same
+  as at the time of that call. At this point, the {\PCname} is decremented;
+  previously, the locations of the old frame are restored to $\tagNoDepth$.
 
@@
 These sequences also enforce the $\SP$ discipline. When stack frames are of
-fixed size, it suffices to increment $\SP$ and tag it with a dedicated tag in
-the header sequence, which is later checked after decrementing $\SP$ during the
-exit sequence to authorize the return. The well-formedness of the blessed
-sequences can be checked statically. Only programs whose sequences are
-well-formed are protected by the micropolicy, although programs are still
-protected (by fail-stopping) even if the intended sequences are missing. By
-tracking the execution of the blessed sequences in the tag state of the
-{\PCname} register, the micropolicy ensures that the entry sequences of calls
-declared in the call map are executed correctly.
+fixed size, it suffices to increment $\SP$ and give it with a special tag in the
+header sequence, which is later checked after decrementing $\SP$ during the exit
+sequence to authorize the return. The well-formedness of the blessed sequences
+can be checked statically. Only programs whose sequences are well-formed are
+protected by the micropolicy, although programs are still protected (by
+fail-stopping) even if the intended sequences are missing. By tracking the
+execution of the blessed sequences in the {\PCname} tag state, the micropolicy
+ensures that the entry sequence is executed correctly a declared call.
 
@@
 %
-This conservative policy can be shown to satisfy an eager \emph{testing
-property}, which at each step of the computation performs integrity and
-confidentiality tests for a stack of running variants, which are kept in
-lockstep with the original machine.
-
-\definition{Stack integrity test}
+This conservative policy is can be shown to satisfy an eager \emph{testing
+property}, which enforces lockstep and performs integrity and confidentiality
+tests for a stack of running variants at each step.
+
+\definition{Eager integrity test}
   Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
-  satisfy the \emph{stack integrity test} if their high-integrity components
+  satisfy the \emph{eager integrity test} if their high-integrity components
   match:
@@
 
-\definition{Stack confidentiality test}
-  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
-  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
-  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
-    confidentiality test} if their observations and changed components coincide:
+\definition{Eager confidentiality test}
+  Given two machine states $\mach_1$ and $\mach_2$ with stepping states and
+  observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
+  \stepstoobs{\obs_2} \mach_2'$ these satisfy the \emph{eager confidentiality
+  test} if their observations and changed components coincide:
   \[
-    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
+    \mathit{conf}_T ~ \mach_1 ~ \obs_1 ~ \mach_1' ~ \mach_2 ~ \obs_2 ~ \mach_2' ~ \triangleq
     \obs_1 = \obs_2 \wedge \forall \component . ~
@@
 \apt{scoping of $\forall$ is ambiguous.}
-\rb{Fix notations, including for properties below.
-  In particular, note that the last four arguments could be computed from the
-  first two and it may be clearer to elide them.}
+\rb{Fix notations, including for properties below.}
 \newcommand{\vse}{\mathit{vse}}
@@
 \definition{Variant stacks}
-  The testing property keeps track of relevant information on a stack whose
-  elements represent variants of the machine by recording an initial machine
-  state, an initial and current variant of the machine, a contour and a return
-  map:
-%
+  The testing property keeps track of execution information on a stack whose
+  elements record an initial machine, an initial and current variant of the
+  machine, a contour and a return map:
   $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
@@
   $$\vs \in \VSS$$
-%
-  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
-  respect to it if the stack is nonempty, the return map of the bottom element
-  is the false property, and for each element $\vse \in \vs$:
+  Given the current machine state $\mach$, a stack $\vs$ is \emph{well-formed}
+  with respect to it if:
   \begin{itemize}
-  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
-  \item $\mach$ is in the trace $\machT$ of its initial machine,
-    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
-  \item Its variant state $\mathit{curr\_variant}(\vse)$
-    is in the trace of its initial variant,
-    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
+  \item The current variant is indeed a variant of the current state.
+  \item The stack is nonempty and the bottom return map is the false property.
+  \item The current state is in the trace of every initial machine.
+  \item The variant state is in the trace of its own initial variant.
   \item The return function is in the return map if not the last element. \rb{TODO}
-  \item The registers of its contour $\contour(\vse)$ are public.
+  \item All contours have public registers.
   \end{itemize}
 
-\definition{Variant stack test}
-%
-  Our testing property extends the basic tests above to whole variant stacks.
-  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
-  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
-  $\mach$ safisfies the \emph{variant stack test} if for each variant element
-  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
-  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
-  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
-  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
-  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
-
 \definition{Eager testing property}
-  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
+  Let $\callmap$ a call map and $\codemap$ a code map. Given a state
   $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
@@
 
-%  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
-%  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
-%  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
-%  testing property} is satisfied at call depth $d$, and write
-%  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:
-
   \begin{itemize}
 
-  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
-    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
-    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
-    $\vs$. (That is, a violation was detected $d$ levels into the stack.
-    Elements are created and pushed to the variant stack on calls and popped
-    from it on returns, so its size always reflects the depth of the program
-    stack.)
+  \item If the machine fail-stops,
+    $(\mach, \pol) \hookrightarrow (\mach, \pol, \tau)$,
+    the property is safisfied for the singleton trace $(\mach, \pol)$ at any
+    depth up to the size of $\vs$.
 
   \item If the top variant of the stack (and only the top variant) returns,
-    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
-    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
-    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
-    use of this vs the return case below.}
-
-  \item If
-    %
-    (1) the machine takes a \emph{non-call, non-return step} $(\mach,
-    \pol) \hookrightarrow_P (\mach', \pol', \obs)$ and the variant stack steps
-    to $\vs \hookrightarrow \vs'$,
-    %
-    (2) the variant stack test $\mathit{\vs}_T ~
-    (\mach, \pol) ~ \vs$ is satisfied, and
-    %
-    (3) the stack safety test
-    $\mathit{stack}_T ~ d ~ \MPT ~ \vs'$ is coinductively satisfied at depth $d$
-    and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
-    \pol')$,
-    %
-    {\em then} we conclude $\mathit{stack}_T d ~ ((\mach, \pol) : \MPT) ~ vs$.
-    \rb{omitting stack size condition}
-
-  \item
-%
-    If (1) the machine takes a \emph{call step} $(\mach, \pol) \hookrightarrow_P
-    (\mach', \pol', \obs)$ and the variant stack steps to $\vs \hookrightarrow
-    \vs'$,
-%
-    (2) $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$,
-%
-    (3) for the new contour at the call, $\contour$, \rb{definition?} the
-    machine satisfies the stack integrity test $\mathit{int}_T ~ \contour ~
-    \mach ~ \mach'$ and the stack confidentiality test with regard to any
-    variant $\mathit{variant} ~ C ~ \mach ~ \mach_{\mathit{var}}$,
-    $\mathit{conf}_T ~ \mach ~ \mach_{\mathit{var}}$ \rb{definition?},
-%
-    and (4) the stack safety test is coinductively satisfied at depth $d$ for a
-    continuing trace $\MPT$ where $\head(\MPT) = (\mach', \pol')$ after an
-    arbitrary variant machine derived from $\contour$ has been added to the top
-    of the stack, $\mathit{stack}_T ~ d ~ \MPT ~ (\vse : \vs')$, \rb{define
-      $\vse$}
-%
-    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : \MPT) ~ vs$.
-
-    %% the contour $\contour$ is updated based on the top of the current machine.
-    %% For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
-    %% the eager integrity test must hold for $\contour$ at the call,
-    %% the eager confidentiality test must hold for all variants at the call,
-    %% including the new variant that is added to the top of the stack,
-    %% and the property recurses on an arbitrary variant at the new contour
-    %% at depth $d$ with trace $\MPT$.
-
-  \item If the machine takes a \emph{return step} at the top of the variant stack
-    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    the property is safisfied for the singleton trace $(\mach, \pol)$ at depth
+    equal to the size of the stack. \rb{Avoid confusion with return below}
+
+  \item If the machine takes a step
+    $(\mach, \pol) \hookrightarrow (\mach', \pol', \obs)$,
+    that is neither a call nor a return, all
+    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
+    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
+    (these tests are required also for calls and returns, and omitted in those cases),
+    and the stepped $(\mach', \pol')$ and $\vs'$ satisfy the eager testing property
+    for a given depth $d$ \rb{omitting stack size condition?} and MP-trace $\MPT$,
+    then the property is also satisfied for the extension of the trace \rb{notation}
+    at depth $d$.
+
+  \item If the machine takes a call step
+    $(\mach, \pol) \hookrightarrow (\mach', \pol', \obs)$,
+    the contour $\contour$ is updated based on the top of the current machine.
+    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    (and in addition to the general step conditions),
+    the eager integrity test must hold for $\contour$ at the call,
+    the eager confidentiality test must hold for all variants at the call,
+    including the new variant that is added to the top of the stack,
+    and the property recurses on an arbitrary variant at the new contour
+    at depth $d$ with trace $\MPT$.
+
+  \item If the machine takes a return step at the top of the variant stack
+    $(\mach, \pol) \hookrightarrow (\mach', \pol', \obs)$,
     the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
@@
 
-  \item If the machine takes \emph{simultaneous return-call step}, that is, the
-    instruction to which execution returns immediately after completing a call
-    is itself a call instruction, the last two cases are combined.
+  \item If the machine takes a return step that is also a call, the last two
+    cases are combined.
 
   \end{itemize}
-
-\definition{Top-level stack testing property}
-%
-  \ldots wrap the coinductive stack testing property
 
@@
         & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT' \Rightarrow \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
         & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
@@
       \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_e}\),
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
       then \(\text{\sc ObsInt}_\contour ~ \MPT\).
@@
       \(\mach_r' = \last(\machT)\). Then the induced traces
-      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT'\) and
+      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
       \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
@@
       A system enjoys {\em observable stack safety} with regard to a call map
-      and a function map if it enjoys both observable stack integrity and
+      and a code map if it enjoys both observable stack integrity and
       observable stack confidentiality with regard to them.

### 2020-07-08 main.tex
CP - 6

--- 2020-07-08 main.tex
@@
 
-\rb{Better section name and headings, references.}
+In this section we are going to see how an existing enforcement mechanism,
+Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
+the formal stack safety property described in the previous section: it in
+fact enforces a stronger, inductive version of it that strengthens the
+local state encapsulation component by checking for integrity and
+confidentiality violations at every step of the callee's execution trace.
 
@@
 %
-In order to enforce a given security policy on a system dynamically, we resort
-to a programmable, tag-based reference monitor that runs alongside the program.
-To control this monitor, we use a programming model that allows fine-grained
-manipulation of metadata tags to encode so-called \emph{micropolicies}
-\rb{cite}.
-
-In such a system, all memory addresses and registers (including the program
-counter) are enriched with an abstract metadata tag, which can point to the
-representation of an arbitrary data structure. A micropolicy is thus defined by
-a collection of rules to be checked as each instruction executes and together
-define the transfer function of the security policy. Each rule inspects the tags
-associated to the inputs of an instruction (PC, registers, memory) as well as
-the instruction opcode itself, and produces one of two outcomes: either the
-instruction is allowed to execute (and generates and assigns tags for the new PC
-and the result of the operation), or the machine fail-stops with a policy
-violation.
-
-Efficient execution of these micropolicies relies on hardware implementations
-such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
-maps tag inputs to outputs and a mechanism to trap misses to software handlers
-and thereby compute the transfer function of the software-defined policy.
-Previous work has shown that a wide range of micropolicies can be defined and
+The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
+on programmable, tag-based reference monitor that runs alongside the
+program.  To control this monitor, they use a programming model that
+allows fine-grained manipulation of metadata tags to encode so-called
+\emph{micropolicies}~\citep{pump_oakland15}.
+
+In such a system, all memory addresses and registers (including the
+program counter) are enriched with an abstract metadata tag, which can
+point to the representation of an arbitrary data structure. A
+micropolicy is thus defined by a collection of rules to be checked as
+each instruction executes and together define the transfer function of
+the security policy. Each rule inspects the tags associated to the
+inputs of an instruction (PC, registers, memory) as well as the
+instruction opcode itself, and produces one of two outcomes: either
+the instruction is allowed to execute (and generates and assigns tags
+for the new PC and the result of the operation), or the machine
+fail-stops with a policy violation.
+
+Efficient execution of these micropolicies relies on hardware
+implementations, such as the PUMP architecture~\citep{pump_oakland15},
+which include a monitor rule cache that maps tag inputs to outputs and
+a mechanism to trap misses to software handlers and thereby compute
+the transfer function of the software-defined policy.  Previous work
+has shown that a wide range of micropolicies can be defined and
 enforced efficiently \rb{cite}.
@@
 %
-The first micropolicy we present to enforce the definition of stack safety
-developed in \cref{sec:lse-and-wbcf} is very similar to the Depth Isolation
-policy presented by \citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the
-stack memory with ownership information associated to each stack frame (and to
-each separate object inside that frames, a more finely grained access control
-that we do not need to consider here), and registers holding stack pointers with
-matching access permission information. Beyond those capabilities, the compiler
-generates designated sequences of instructions with special tags that grant them
-permission to manage the stack, say during calls and returns. \rb{Conceivably on
-alloc operations, etc., which we have not yet mentioned.}
+The first micropolicy we present to enforce the definition of stack
+safety developed in \cref{sec:lse-and-wbcf} is \leo{a small variation
+  of?}  the Depth Isolation policy presented by
+\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
+with ownership information associated to each stack frame (and to each
+separate object inside that frames, a more finely grained access
+control that we do not need to consider here), and registers holding
+stack pointers with matching access permission information. Beyond
+those capabilities, the compiler generates designated sequences of
+instructions with special tags that grant them permission to manage
+the stack, say during calls and returns. \rb{Conceivably on alloc
+  operations, etc., which we have not yet mentioned.}
 
@@
 
-These sequences also enforce the $\SP$ discipline. When stack frames are of
-fixed size, it suffices to increment $\SP$ and tag it with a dedicated tag in
-the header sequence, which is later checked after decrementing $\SP$ during the
-exit sequence to authorize the return. The well-formedness of the blessed
-sequences can be checked statically. Only programs whose sequences are
-well-formed are protected by the micropolicy, although programs are still
-protected (by fail-stopping) even if the intended sequences are missing. By
-tracking the execution of the blessed sequences in the tag state of the
-{\PCname} register, the micropolicy ensures that the entry sequences of calls
-declared in the call map are executed correctly.
+These sequences also enforce the $\SP$ discipline. When stack frames
+are of fixed size, it suffices to increment $\SP$ and tag it with a
+dedicated tag in the header sequence, which is later checked after
+decrementing $\SP$ during the exit sequence to authorize the
+return. The well-formedness of the blessed sequences can be checked
+statically. Only programs whose sequences are well-formed are
+protected by the micropolicy, although programs are still protected
+(by fail-stopping) even if the intended sequences are missing. By
+tracking the execution of the blessed sequences in the tag state of
+the {\PCname} register, the micropolicy ensures that the entry
+sequences of calls declared in the call map are executed correctly.
 
@@
 
-During normal execution, the {\PCname} uses the current activation depth in its
-tag $\tagPCDepth{n}$ to gate load and store operations, which are allowed to
-operate only on the corresponding stack frame, tagged $\tagStackDepth{n}$.
-\rb{Note that we have also considered and implemented more flexible schemes
-  based on allowing operations above the current activation depth. How do we put
-  these together?}
-\rb{Do we want to discuss variations involving frame pointers or more implicit
-  return addresses?}
-\rb{Where to talk about the role of the compiler?}
+During normal execution, the {\PCname} uses the current activation
+depth in its tag $\tagPCDepth{n}$ to gate load and store operations,
+which are allowed to operate only on the corresponding stack frame,
+tagged $\tagStackDepth{n}$.  \rb{Note that we have also considered and
+  implemented more flexible schemes based on allowing operations above
+  the current activation depth. How do we put these together?}  \rb{Do
+  we want to discuss variations involving frame pointers or more
+  implicit return addresses?}  \rb{Where to talk about the role of the
+  compiler?}
 

### 2020-07-08 main.tex
pass on 5.1 and 5.2; releasing lock on them; keeping lock on 5.3

--- 2020-07-08 main.tex
@@
 
-In this section we will describe stack safety as a formal property. We
+In this section we formalize our definition of stack safety.We
 begin with our model of \leo{compiler-provided?} knowledge about the
@@
 Similarly, a {\em return map} identifies addresses that complete a
-return sequence. Formally, it is simply a set of components; each one
-is the final jump corresponding to a function:
+return sequence. Formally, it is simply a set of words; each one
+is the address of the final jump corresponding to a function return:
 %
@@
 to have correctly returned from a call. We consider the call to have
-returned the first time control returns to the instruction following
-the \jal with the stack restored. Formally this is a relation
-between the states:
-
-\[\begin{split}
-    \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) =
-    \nextPC(\mach_c(\PCname)) \land \\ & \mach_r(\SP) = \mach_c(\SP)
-\end{split}\]
-
+returned when control returns to the instruction following
+the \jal with the stack restored. Formally, this is a predicate on
+states that says when the machine has just returned from a specified call state
+(where $\nextPC$ gives the address of the next instruction after a given address):
+\[%\begin{split}
+    \ret{\mach_c} \triangleq  \lambda \mach_r . \mach_r(\PCname) =
+    \nextPC(\mach_c(\PCname)) \land  \mach_r(\SP) = \mach_c(\SP)
+%\end{split}
+\]
 There are two such return points in our running example, both of which
@@
 
-In the rest of this section we will implictly parameterize our
+In the rest of this section we implictly parameterize all 
 definitions by a function map $\codemap$, a call map $\callmap$, a return
@@
 
-When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) and
-\(\mach_c \stepstoobs{\obs} \mach_e\) for some $n$, $\obs$, and
-$\mach_e$, we can construct a contour based on the bounds of the
+When \(\mach_c\) is a call state (\(\callmap(\mach_c(\PCname)) = n\) for some $n$) and
+\(\mach_c \stepstoobs{\obs} \mach_e\) for some 
+$\mach_e$ and $\obs$, we can construct a contour based on the bounds of the
 callee's stack frame.
@@
 call. Private components are marked high confidentiality and high
-integrity: the callee should be able to neither read, nor write them.
+integrity: the callee should be able to neither read nor write them.
 %
@@
 callee and high confidentiality otherwise; regardless they are always
-high integrity.
-%
-Finally, registers and the frame of arguments to the callee are free
+high integrity.\apt{Sean:I think this is vestigial from an aborted version
+  in which you were modeling WBCF using confidentiality. In any casen
+  it is not reflected in the definition below.}
+%
+Finally, registers and the frame of arguments and results are free
 for the callee to use, and are therefore tagged with low integrity and
@@
     (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
-                         \text{ and for some } \funid, \codemap(\component) = \funid \\
+                         \text{ and } \codemap(\component) = \funid \text{ for some }\funid\\
     (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
@@
 
-Going back to our running example, recall the state of the machine at
-the call from {\tt main} to {\tt g}, annotated with their
+Going back to our running example, here is the state of the machine at
+the call from {\tt main} to {\tt g}, showing addresses annotated with their 
 corresponding labels (not shown: the labels of the instructions
 corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
-$(\HIGHSEC, \HIGHINT)$):
+$(\HIGHSEC, \HIGHINT)$)\apt{see previous note}:
 \vspace*{0.2em}
@@
 
-\paragraph*{Stack Integrity}
+\subsubsection{Stack Integrity}
 
@@
 \definition
-
 Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
@@
   \mathit{int} ~ \contour ~ \MPT \triangleq &
-    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
-  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
+    \head(\MPT) = \mpostate{\mach_e}{\_}{\_}  \land 
+    \last(\MPT) = \mpostate{\mach_r}{\_}{\_} \Rightarrow \\
   & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
@@
 private data is never overwritten, even during the caller's execution
-(a variation that leads to easier reasoning and more effective testing);
-and a weaker, lazy integrity property that only ensures that the
-locations that a caller accesses are preserved (a variation that allows
-for a more efficient enforcement mechanism).
+(which leads to easier reasoning and more effective testing);
+and a weaker, lazy integrity property that only ensures the preservation
+of locations that subsequently affect the caller's observable actions 
+(which allows for a more efficient enforcement mechanism).
 
@@
 
-For an example see \cref{fig:intex}, representing the unsafe substitution
-\[
-  44: ~ \sw ~ \rsp ~ r_4 ~ -5
-\]
-This instruction is in {\tt h}, and instead of storing data to the return,
-it overwrites {\tt g}'s variable, {\tt z}. At the top of the figure high
-integrity memory is in red, while low integrity is in green; at the end of the
-trace address 103 should match the initial state, and does not. So trace
-integrity is violated.
+Trace integrity characterizes the trace corresponding to a single
+callee.
+Going back to our running example, such a trace would be, for
+instance, the entire trace of {\tt h} from instruction 40 to
+instruction 47. 
+As an example of an integrity violation affecting this trace, \cref{fig:intex} shows what happens
+if we change line 44 of the example program to $[\rsp - 5] \leftarrow r_4$.
+Instead of storing data to the return value slot,
+the changed version overwrites {\tt g}'s local variable {\tt z}.
+The figure shows the modified trace, with snapshots of the initial
+and final states colored to show the contour; high 
+integrity memory is in red\apt{NO?}, while low integrity is in green\apt{NO?}.
+At the end of the trace high integrity address 103 should match the initial state, and does not.
+So trace integrity is violated.
 \rb{There are some color name macros, but none for integrity yet? Use instead
-of hardcoded color names.}
-
-Trace integrity characterizes the trace corresponding to a single
-callee. Going back to our running example, such a trace would be, for
-instance, the entire trace of {\tt h} from instruction 40 to
-instruction 47. This trace is repeated in \Cref{fig:intex}, with the initial
-and final states colored to show the contour; high confidentiality
-components in magenta, low in teal. At the end of the trace addresses 103
-and 108 have both changed; since 103 was high confidentiality, the trace
-does not enjoy trace integrity.
+  of hardcoded color names.}\apt{I don't understand what's going on with the colors.
+  I thought we agreed to use the same coloring scheme throughout?}
 
@@
 execution trace. We do this by requiring that trace integrity holds
-for every transition between a caller and a callee for the entirety of
-the trace of the callee.
-
-\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
-a function map $\codemap$, an $n$-{\em transition} is a pair of
+for every subtrace delimited by a call and its corresponding return.
+%for every transition between a caller and a callee for the entirety of
+%the trace of the callee.\apt{?????}
+
+\apt{Some more words needed here to smooth the technicalities.}
+
+\definition%{Transition}
+Given a trace $\MPT$, an $n$-{\em transition} is a pair of
 machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
@@
   $$\callmap(\mach_c(\PCname)) = n,$$
-\item and $\mach_e$ is an entry point for $f$:
-  $$\codemap(\mach_e(\PCname)) = f.$$
+\item and $\mach_e$ is an entry point\apt{or is in some function?? and why do we need either??}:
+  $$\exists f. \codemap(\mach_e(\PCname)) = f.$$
 \end{itemize}
 
-\definition{Stack Integrity} Let $\callmap$ be a call map and
-$\codemap$ a function map. We say that a system satisfies {\em stack
-  integrity} with respect to $\callmap$ and $\codemap$ if, for every
+\definition%{Stack Integrity}
+A system satisfies {\em stack
+  integrity} if, for every
 initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
@@
 
-\paragraph*{Stack Confidentiality}
+\subsubsection{Stack Confidentiality}
 
@@
 property: just like integrity ensures that a caller's data is never
-overwritten by a callee, confidentiality ensures that it never
-influences its observable behavior. We will formally define
+overwritten by a callee, confidentiality ensures that the caller's data never
+influences the callee's observable behavior. We will formally define
 confidentiality in two similar steps: a trace version and a system
@@
 
-\definition{Variations}
+\definition%{Variations}
 We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
-another given a contour $\contour$ if they agree on low
+another with respect to a contour $\contour$ if they agree on low
 confidentiality data:
@@
 Then, the trace version of confidentiality operates on two traces: the
-``original'' one $MP \in \MPOTRACE$ that corresponds to a callee $f$
-and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
-if it started from a variant initial machine state.
+``original'' tracce $MP \in \MPOTRACE$ that corresponds to a callee $f$
+and a ``variant'' trace $M \in \MTRACE$ that corresponds to how $f$
+would behaves if it were started from a variant initial machine state.
 %
@@
 
-\definition{Trace Confidentiality} We will say two traces $MP$ and $M$
-satisfy {\em trace confidentiality} with respect to $\contour$ and
-a return predicate \(R \subset \MACHS\) if:
+\definition%{Trace Confidentiality}
+%% Suppose $m_e \approx_\contour m_e'$.
+%% Two traces $MP$ with $\head(MP) = m_e$ and $M$ with $head(M) = m_e'$
+%% satisfy {\em trace confidentiality} with respect to $\contour$ and
+%% a return predicate \(R \subset \MACHS\) if:
+Two traces $MP$, with head $m_e$, and $M$, with head $m_e'$,
+satisfy {\em trace confidentiality} with respect to contour $\contour$ and
+a return predicate \(R \subset \MACHS\) if, whenever $m_e \approx_\contour m_e'$ then:
 \begin{itemize}
-\item If $\MPT$ is finite and its final state $R\ \mach_r$,
+\item If $\MPT$ is finite with final state $\mach_r$ and $\mach_r \in R$, 
   then $\machT$ is also finite with some final state $\mach_r'$ and
@@
   \mach_r(\component) = \mach_r'(\component)$$
-  In addition their observations must be equivalent:
+  In addition their observations must be similar:
   $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
 \item If $\MPT$ is infinite, then $\machT$ is also infinite and their
-  observations are equivalent:
+  observations are similar:
   $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
@@
 \end{itemize}
-
-For example, consider the unsafe substitution:
-\[
-42 : ~ \sw ~ r_4 ~ \rsp 100
-\]
-Instead of reading the proper argument, {\tt h} reads directly from the
+\apt{Needs words about why we need a return predicate abstraction.}
+
+Consider the result of modifying line 42 of our running example to $r_4 \leftarrow [\rsp - 9]$.
+Instead of reading the proper argument {\tt v}, {\tt h} now reads directly from the
 secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
@@
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
-hold -5, in orange. After the last step, the return states \(\mach_r\) and
+hold -5, in orange.\apt{NO?}  After the last step, the return states \(\mach_r\) and
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
 \(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
-is consistent, 108 holds either 43 or -4, so the trace does not obey
-trace confidentiality.
+had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
 \rb{Color name macros}
-
-Meanwhile, the following unsafe substitution outputs the secret
-directly, violating the internal observation clause of confidentiality:
-\[
-30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-\]
+\apt{Explain slash notation in effects column.}
+
+As another example, 
+changing line 30 of the running example to $r_5  \leftarrow [\rsp - 6]$ would cause
+{\tt g} to read the value of {\tt main}'s local variable {\tt x} and then immediately
+output it, violating the internal observation clause of the confidentiality definition.
 %\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
@@
 
-\definition{Stack Confidentiality}
-Let $\callmap$ be a call map and $\codemap$ a function map. We say that a
-system satisfies {\em stack confidentiality} with respect to
-$\callmap$ and $\codemap$ if, for every initial machine-policy state
+\definition%{Stack Confidentiality}
+%Let $\callmap$ be a call map and $\codemap$ a function map.
+We say that a
+system satisfies {\em stack confidentiality}
+% with respect to $\callmap$ and $\codemap$
+if, for every initial machine-policy state
 pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
-$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
-$f$ in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
+$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
+in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
 \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
-For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
+For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
 \mach_e\), let \(\machT\) be the similarly prefixed trace such that
 \(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
-\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
+\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
 and \(R = \ret{\mach_c}\):
@@
 \mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
-$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
+$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, n)} \mach_e' \wedge
 \mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$
+
+\apt{oy. still needs work to make this intelligible}
 

### 2020-07-08 main.tex
CP for rob

--- 2020-07-08 main.tex
@@
 
-\paragraph*{Testing the conservative policy}
-%
-This conservative policy can be shown to satisfy an eager \emph{testing
-property}, which at each step of the computation performs integrity and
-confidentiality tests for a stack of running variants, which are kept in
-lockstep with the original machine.
-
-\definition{Stack integrity test}
+\paragraph*{Strong Stack Safety}
+%
+This conservative policy does not only satisfy the stack safety above,
+but also a provably stronger one, which at each step of the
+computation performs integrity and confidentiality tests for a stack
+of running variants, which are kept in lockstep with the original
+machine.
+
+\definition
   Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
@@
 
-\definition{Stack confidentiality test}
+\definition
   Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
@@
 
-\definition{Variant stacks}
-  The testing property keeps track of relevant information on a stack whose
-  elements represent variants of the machine by recording an initial machine
-  state, an initial and current variant of the machine, a contour and a return
-  map:
-%
-  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
-  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
-  $$\vs \in \VSS$$
-%
+\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
+
+\definition The testing property keeps track of relevant information
+on a {\em variant stack}: a list of machine states, each element of
+which (denoted $\vm{C}$) of which is a variation of the actual machine
+according to a contour $C$.
+%
+\footnote{The full formal definition in Coq also records additional
+  information (the initial machine state at the call, its initial
+  variation, the contour, and a return predicate) which we elide
+  here for simplicity of exposition.}
+%%
+%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
+%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
+%  $$\vs \in \VSS$$
+%%
+\iflater
+\leo{This can be elided}
   Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
@@
   \end{itemize}
-
-\definition{Variant stack test}
-%
-  Our testing property extends the basic tests above to whole variant stacks.
-  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
-  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
-  $\mach$ safisfies the \emph{variant stack test} if for each variant element
-  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
-  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
-  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
-  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
-  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
+  \fi
+
+\definition
+%
+  Our testing property extends the basic tests above to whole variant
+  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
+  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
+  is well-formed with respect to $\mach$\fi safisfies the
+  \emph{variant stack test} if for each variant element $\vm{C}$
+  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step
+  $\mathit{curr\_variant}(\vse) \stepstoobs{\obs} \mach'$ while
+  satisfying the stack integrity test $\mathit{int}_T ~ \contour(\vse)
+  ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~
+  \mach ~ \mach'$ with regard to the original machine state. We write
+  this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
 

### 2020-07-08 main.tex

--- 2020-07-08 main.tex
@@
 a \(\tau\):
-\rb{This $\hookrightarrow_P$ is often enough written just $\hookrightarrow$.
-  Overload everywhere?}\apt{Done, and for $\rightharpoonup$ too.}
 
@@
 
-Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
-$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
-respect to $\contour$ if its final state (if any) agrees with its
-initial one on all components marked $\HIGHSEC$ in $\contour$:
+Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
+\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
+respect to \(\contour\) if its final state (if any) agrees with its
+initial one on all components marked \(\HIGHSEC\) in \(\contour\):
 %
 \[\begin{split}
-  \mathit{int} ~ \contour ~ \MPT \triangleq &
+  \propdef{TraceIntegrity} ~ \contour ~ \MPT \triangleq &
     \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
@@
 \end{split}\]
-%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
-%\sna{We could do it like this, with finiteness implicit in the existence of
-%     a final state.}+1
 
@@
 for every transition between a caller and a callee for the entirety of
-the trace of the callee.
-
-\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
-a function map $\codemap$, an $n$-{\em transition} is a pair of
-machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
-in the trace such that:
-\begin{itemize}
-\item one state pair steps to the other:
-  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
-\item $\mach_c$ is a valid call with $n$ arguments:
-  $$\callmap(\mach_c(\PCname)) = n,$$
-\item and $\mach_e$ is an entry point for $f$:
-  $$\codemap(\mach_e(\PCname)) = f.$$
-\end{itemize}
-
-\definition{Stack Integrity} Let $\callmap$ be a call map and
-$\codemap$ a function map. We say that a system satisfies {\em stack
-  integrity} with respect to $\callmap$ and $\codemap$ if, for every
-initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
-trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
-$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
-$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
-(if any) satisfies trace integrity with respect to the contour at the
+the trace of the callee, with a transition defined:
+
+\definition Given a trace \(\MPT\), a call map \(\callmap\) and
+a function map \(\codemap\), an \(n\)-{\em transition} is a pair of
+machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
+\(\mpstate{\mach_e}{\pol_e}\) in the trace such that
+\(\mpstate{\mach_c}{\pol_c}\) steps to \(\mpstate{\mach_e}{\pol_e}\),
+\(\mach_c\) is a valid call with \(n\) arguments, and
+and \(\mach_e\) is an entry point for some function \(f\):
+ % 
+  \[\begin{split}
+    \mathit{Tr}_n ~ \mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}
+    \triangleq & \exists \obs, f.
+    ~ \mpstate{\mach_c}{\pol_c} \stepstopol{\obs} \mpstate{\mach_e}{\pol_e} \\
+    & \land \callmap(\mach_c(\PCname)) = n
+      \land \codemap(\mach_e(\PCname)) = f
+  \end{split}\]
+
+\definition Let \(\callmap\) be a call map and \(\codemap\) a function map.
+We say that a system satisfies {\em stack integrity} with respect to
+\(\callmap\) and \(\codemap\) if, for every initial machine-policy state pair
+\(\mpstate{\mach_0}{\pinit(\mach_0)})\) and its induced trace \(\MPT\), and
+every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\) and
+\(\mpstate{\mach_e}{\pol_e}\) in \(\MPT\), the prefix \(\MPT_{call}\) of \(\MPT\)
+from \(\mpstate{\mach_e}{\pol_e}\) until the first return corresponding to
+\(\mach_c\) satisfies trace integrity with respect to the contour at the
 call:
 %
-$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
-\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
-\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$
+  \[\begin{split}
+    \propdef{StackIntegrity} \triangleq & \forall \MPT_{call}. ~
+      \mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} ~ | ~ \ret{\mach_c}
+      \Rightarrow \\
+    & \propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c, n) ~ \MPT_{call}
+  \end{split}\]
 
@@
 Unlike integrity, confidentiality is a noninterference-style property,
-comparing the behavior of a machine with an arbitrary variant that
+comparing the behavior of a machine with an arbitrary variation that
 preserves low confidentiality data.
 
-\definition{Variations}
-We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
-another given a contour $\contour$ if they agree on low
+\definition
+We say two machine states \(\mach\) and \(\mach'\) are {\em variations}
+of one another given a contour \(\contour\) if they agree on low
 confidentiality data:
@@
 
-\definition{Trace Confidentiality} We will say two traces $MP$ and $M$
-satisfy {\em trace confidentiality} with respect to $\contour$ and
-a return predicate \(R \subset \MACHS\) if:
+\definition
+
+We will say two traces \(\MPT\) and \(\machT\) satisfy {\em trace
+confidentiality} with respect to a return predicate
+\(R \subset \MACHS\) if they behave the same in their observations and the ways
+that they change their states, captured across the following three cases:
+\sna{Interestingly, split out like this, we don't need to talk about
+     contour here -- they are varied in the next property.}
 \begin{itemize}
-\item If $\MPT$ is finite and its final state $R\ \mach_r$,
-  then $\machT$ is also finite with some final state $\mach_r'$ and
-  any changes from the initial states have changed in the same way:
-$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
-  \vee \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
-  \mach_r(\component) = \mach_r'(\component)$$
-  In addition their observations must be equivalent:
-  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
-\item If $\MPT$ is infinite, then $\machT$ is also infinite and their
-  observations are equivalent:
-  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
-\item If the $\MPT$ ends prematurely, without fulfilling \(R\), then
-  we require that its observations are a prefix of those of $\machT$:
-  $$\pi_o(\MPT) \lesssim \pi_o(\machT)$$
+  \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
+    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
+    some final state \(\mach_r'\) and any changes from the initial states are
+    matched between \(\mach_r\) and \(\mach_r'\). In addition their
+    observations are similar.
+    \[\begin{split}
+      \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
+      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
+        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
+        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
+      & ~ (\mach_e(\component) \not = \mach_r(\component)
+        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
+      \mach_r(\component) = \mach_r'(\component))
+    \end{split}\]
+  \item If \(\MPT\) is infinite, then \(\machT\) is also infinite and their
+  observations are similar.
+    \[\begin{split}
+      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
+      & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
+      & \neg (\exists mo . \last(\machT) = mo)
+        \land \pi_o(\MPT) \eqsim \pi_o(\machT)
+    \end{split}\]
+\item If \(\MPT\) ends prematurely without fulfilling \(R\), then
+  its observations are a prefix of those of \(\machT\).
+    \[\begin{split}
+      \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
+      & \forall \mach_r ~ \pol_r ~ \obs_r .
+        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+      & \neg ~ R ~ \mach_r \Rightarrow \\
+      & \pi_o(\MPT) \lesssim \pi_o(\machT)
+    \end{split}\]
 \end{itemize}
 
-For example, consider the unsafe substitution:
+Putting it all together:
+
+\[\begin{split}
+  \propdef{TraceConfidentiality} ~ R ~ \MPT ~ \machT \triangleq
+  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
+  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
+    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
+\end{split}\]
+
+In the non-return cases, we only care that the traces exhibit the same
+observable behavior (or a prefix if the trace halts prematurely), as changes
+in the state will not affect the caller absent a return.  In the return case
+we must additionally guarantee that no changes were made based on secret data
+that could corrupt the caller upon return.
+
+For example, consider if we replaced instruction 42 with the unsafe instruction:
 \[
@@
 \]
-%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
-%and \(\mach_e'\), below. Low confidentiality data in green is shared
-%between them, while high confidentiality data in red and yellow may differ.
-%After executing the code shown, the final states below have written different
-%values to address 42, represented by the black and white components.
-%It is allowed for address 40 to differ between the states due to the initial
-%variation, but address 42 changed differently between the traces, representing
-%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
-%is read, it is overwritten before return, and the write at step 3 always writes
-%the same value. \Gprivc and red locations differ between variations only because
-%of the initial variation, and the rest are identical. So this example passes.
 
@@
 
-In this section we are going to see how an existing enforcement mechanism,
-Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
-the formal stack safety property described in the previous section: it in
-fact enforces a stronger, inductive version of it that strengthens the
-local state encapsulation component by checking for integrity and
-confidentiality violations at every step of the callee's execution trace.
+\rb{Better section name and headings, references.}
 
@@
 %
-The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
-on programmable, tag-based reference monitor that runs alongside the
-program.  To control this monitor, they use a programming model that
-allows fine-grained manipulation of metadata tags to encode so-called
-\emph{micropolicies}~\citep{pump_oakland15}.
-
-In such a system, all memory addresses and registers (including the
-program counter) are enriched with an abstract metadata tag, which can
-point to the representation of an arbitrary data structure. A
-micropolicy is thus defined by a collection of rules to be checked as
-each instruction executes and together define the transfer function of
-the security policy. Each rule inspects the tags associated to the
-inputs of an instruction (PC, registers, memory) as well as the
-instruction opcode itself, and produces one of two outcomes: either
-the instruction is allowed to execute (and generates and assigns tags
-for the new PC and the result of the operation), or the machine
-fail-stops with a policy violation.
-
-Efficient execution of these micropolicies relies on hardware
-implementations, such as the PUMP architecture~\citep{pump_oakland15},
-which include a monitor rule cache that maps tag inputs to outputs and
-a mechanism to trap misses to software handlers and thereby compute
-the transfer function of the software-defined policy.  Previous work
-has shown that a wide range of micropolicies can be defined and
+In order to enforce a given security policy on a system dynamically, we resort
+to a programmable, tag-based reference monitor that runs alongside the program.
+To control this monitor, we use a programming model that allows fine-grained
+manipulation of metadata tags to encode so-called \emph{micropolicies}
+\rb{cite}.
+
+In such a system, all memory addresses and registers (including the program
+counter) are enriched with an abstract metadata tag, which can point to the
+representation of an arbitrary data structure. A micropolicy is thus defined by
+a collection of rules to be checked as each instruction executes and together
+define the transfer function of the security policy. Each rule inspects the tags
+associated to the inputs of an instruction (PC, registers, memory) as well as
+the instruction opcode itself, and produces one of two outcomes: either the
+instruction is allowed to execute (and generates and assigns tags for the new PC
+and the result of the operation), or the machine fail-stops with a policy
+violation.
+
+Efficient execution of these micropolicies relies on hardware implementations
+such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
+maps tag inputs to outputs and a mechanism to trap misses to software handlers
+and thereby compute the transfer function of the software-defined policy.
+Previous work has shown that a wide range of micropolicies can be defined and
 enforced efficiently \rb{cite}.
@@
 %
-The first micropolicy we present to enforce the definition of stack
-safety developed in \cref{sec:lse-and-wbcf} is \leo{a small variation
-  of?}  the Depth Isolation policy presented by
-\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
-with ownership information associated to each stack frame (and to each
-separate object inside that frames, a more finely grained access
-control that we do not need to consider here), and registers holding
-stack pointers with matching access permission information. Beyond
-those capabilities, the compiler generates designated sequences of
-instructions with special tags that grant them permission to manage
-the stack, say during calls and returns. \rb{Conceivably on alloc
-  operations, etc., which we have not yet mentioned.}
+The first micropolicy we present to enforce the definition of stack safety
+developed in \cref{sec:lse-and-wbcf} is very similar to the Depth Isolation
+policy presented by \citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the
+stack memory with ownership information associated to each stack frame (and to
+each separate object inside that frames, a more finely grained access control
+that we do not need to consider here), and registers holding stack pointers with
+matching access permission information. Beyond those capabilities, the compiler
+generates designated sequences of instructions with special tags that grant them
+permission to manage the stack, say during calls and returns. \rb{Conceivably on
+alloc operations, etc., which we have not yet mentioned.}
 
@@
 
-These sequences also enforce the $\SP$ discipline. When stack frames
-are of fixed size, it suffices to increment $\SP$ and tag it with a
-dedicated tag in the header sequence, which is later checked after
-decrementing $\SP$ during the exit sequence to authorize the
-return. The well-formedness of the blessed sequences can be checked
-statically. Only programs whose sequences are well-formed are
-protected by the micropolicy, although programs are still protected
-(by fail-stopping) even if the intended sequences are missing. By
-tracking the execution of the blessed sequences in the tag state of
-the {\PCname} register, the micropolicy ensures that the entry
-sequences of calls declared in the call map are executed correctly.
+These sequences also enforce the $\SP$ discipline. When stack frames are of
+fixed size, it suffices to increment $\SP$ and tag it with a dedicated tag in
+the header sequence, which is later checked after decrementing $\SP$ during the
+exit sequence to authorize the return. The well-formedness of the blessed
+sequences can be checked statically. Only programs whose sequences are
+well-formed are protected by the micropolicy, although programs are still
+protected (by fail-stopping) even if the intended sequences are missing. By
+tracking the execution of the blessed sequences in the tag state of the
+{\PCname} register, the micropolicy ensures that the entry sequences of calls
+declared in the call map are executed correctly.
 
@@
 
-During normal execution, the {\PCname} uses the current activation
-depth in its tag $\tagPCDepth{n}$ to gate load and store operations,
-which are allowed to operate only on the corresponding stack frame,
-tagged $\tagStackDepth{n}$.  \rb{Note that we have also considered and
-  implemented more flexible schemes based on allowing operations above
-  the current activation depth. How do we put these together?}  \rb{Do
-  we want to discuss variations involving frame pointers or more
-  implicit return addresses?}  \rb{Where to talk about the role of the
-  compiler?}
+During normal execution, the {\PCname} uses the current activation depth in its
+tag $\tagPCDepth{n}$ to gate load and store operations, which are allowed to
+operate only on the corresponding stack frame, tagged $\tagStackDepth{n}$.
+\rb{Note that we have also considered and implemented more flexible schemes
+  based on allowing operations above the current activation depth. How do we put
+  these together?}
+\rb{Do we want to discuss variations involving frame pointers or more implicit
+  return addresses?}
+\rb{Where to talk about the role of the compiler?}
 
@@
 
-\paragraph*{Strong Stack Safety}
-%
-This conservative policy does not only satisfy the stack safety above,
-but also a provably stronger one, which at each step of the
-computation performs integrity and confidentiality tests for a stack
-of running variants, which are kept in lockstep with the original
-machine.
-
-\definition
+\paragraph*{Testing the conservative policy}
+%
+This conservative policy can be shown to satisfy an eager \emph{testing
+property}, which at each step of the computation performs integrity and
+confidentiality tests for a stack of running variants, which are kept in
+lockstep with the original machine.
+
+\definition{Stack integrity test}
   Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
@@
 
-\definition
+\definition{Stack confidentiality test}
   Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
@@
 
-\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
-
-\definition The testing property keeps track of relevant information
-on a {\em variant stack}: a list of machine states, each element of
-which (denoted $\vm{C}$) of which is a variation of the actual machine
-according to a contour $C$.
-%
-\footnote{The full formal definition in Coq also records additional
-  information (the initial machine state at the call, its initial
-  variation, the contour, and a return predicate) which we elide
-  here for simplicity of exposition.}
-%%
-%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
-%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
-%  $$\vs \in \VSS$$
-%%
-\iflater
-\leo{This can be elided}
+\definition{Variant stacks}
+  The testing property keeps track of relevant information on a stack whose
+  elements represent variants of the machine by recording an initial machine
+  state, an initial and current variant of the machine, a contour and a return
+  map:
+%
+  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
+  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
+  $$\vs \in \VSS$$
+%
   Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
@@
   \end{itemize}
-  \fi
-
-\definition
-%
-  Our testing property extends the basic tests above to whole variant
-  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
-  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
-  is well-formed with respect to $\mach$\fi safisfies the
-  \emph{variant stack test} if for each variant element $\vm{C}$
-  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step
-  $\mathit{curr\_variant}(\vse) \stepstoobs{\obs} \mach'$ while
-  satisfying the stack integrity test $\mathit{int}_T ~ \contour(\vse)
-  ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~
-  \mach ~ \mach'$ with regard to the original machine state. We write
-  this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
+
+\definition{Variant stack test}
+%
+  Our testing property extends the basic tests above to whole variant stacks.
+  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
+  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
+  $\mach$ safisfies the \emph{variant stack test} if for each variant element
+  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
+  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
+  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
+  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
+  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
 
@@
 
-  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
+  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow
     (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
@@
   \item If the machine takes a \emph{return step} at the top of the variant stack
-    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    $(\mach, \pol) \hookrightarrow (\mach', \pol', \obs)$,
     the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$

### 2020-07-08 main.tex
Fix definition 6.4
Remove vestigial notation and use the parts that are implicitly
defined by the new notation, with some trivial lifting of steps.

--- 2020-07-08 main.tex
@@
   \emph{variant stack test} if for each variant element $\vm{C}$
-  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step
-  $\mathit{curr\_variant}(\vse) \stepstoobs{\obs} \mach'$ while
-  satisfying the stack integrity test $\mathit{int}_T ~ \contour(\vse)
+  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
+  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
+  satisfying the stack integrity test $\mathit{int}_T ~ C
   ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~

### 2020-07-08 main.tex
More compact/formal phrasing of Definition 6.5

@@
   say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
-  testing property} at a given call depth if:
+  testing property} at a given call depth as follows.
+% Refactor repetitive notation before case split
+  In all cases that assume a machine step, $(\mach, \pol) \hookrightarrow_P
+  (\mach', \pol', \obs)$, assume also (unless the step is a return) that the
+  variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
+  machines can step.
 
@@
 %
-    (1) the machine takes a \emph{non-call, non-return step} $(\mach,
-    \pol) \hookrightarrow_P (\mach', \pol', \obs)$ and the variant stack steps
-    to $\vs \hookrightarrow \vs'$,
-    %
-    (2) the variant stack test $\mathit{\vs}_T ~
-    (\mach, \pol) ~ \vs$ is satisfied, and
-    %
-    (3) the stack safety test
-    $\mathit{stack}_T ~ d ~ \MPT ~ \vs'$ is coinductively satisfied at depth $d$
-    and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
-    \pol')$,
-    %
-    {\em then} we conclude $\mathit{stack}_T d ~ ((\mach, \pol) : \MPT) ~ vs$.
-    \rb{omitting stack size condition}
+    (1) the machine takes a \emph{non-call, non-return step},
+%
+    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
+    satisfied, and
+%
+    (3) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') : \MPT) ~
+    \vs'$ is coinductively satisfied % at depth $d$
+    %% and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
+    %% \pol')$,
+%
+    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
+    \pol') : \MPT) ~ vs$.
+    %% \rb{omitting stack size condition}
 
@@
 %
-    If (1) the machine takes a \emph{call step} $(\mach, \pol) \hookrightarrow_P
-    (\mach', \pol', \obs)$ and the variant stack steps to $\vs \hookrightarrow
-    \vs'$,
-%
-    (2) $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$,
-%
-    (3) for the new contour at the call, $\contour$, \rb{definition?} the
-    machine satisfies the stack integrity test $\mathit{int}_T ~ \contour ~
-    \mach ~ \mach'$ and the stack confidentiality test with regard to any
-    variant $\mathit{variant} ~ C ~ \mach ~ \mach_{\mathit{var}}$,
-    $\mathit{conf}_T ~ \mach ~ \mach_{\mathit{var}}$ \rb{definition?},
-%
-    and (4) the stack safety test is coinductively satisfied at depth $d$ for a
-    continuing trace $\MPT$ where $\head(\MPT) = (\mach', \pol')$ after an
-    arbitrary variant machine derived from $\contour$ has been added to the top
-    of the stack, $\mathit{stack}_T ~ d ~ \MPT ~ (\vse : \vs')$, \rb{define
-      $\vse$}
-%
-    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : \MPT) ~ vs$.
+    If (1) the machine takes a \emph{call step} and $\contour$ is the contour at
+    the call,
+%
+    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
+    satisfied,
+%
+    (3) the stack integrity test $\mathit{int}_T ~ \contour ~ \mach ~ \mach'$ is
+    satisfied and the stack confidentiality test $\mathit{conf}_T ~ \mach ~
+    \vm{\contour}$ is satisfied with regard to an arbitrary $\contour$-variant
+    of $\mach$,
+%
+    and (4) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') :
+    \MPT) ~ (\vm{\contour} : \vs')$ is coinductively satisfied, for an arbitrary
+    $\contour$-variant of the machine,
+%
+    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
+    \pol') : \MPT) ~ vs$.
 
@@
 
-  \item If the machine takes a \emph{return step} at the top of the variant stack
-    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
-    the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
-    if (in addition to the general step conditions)
-    the property recurses at depth $d$ with trace $\MPT$
-    after removing the returning top variant from the stack.
+  \item
+%
+    If (1) the machine takes a \emph{return step} at the top of the variant
+    stack (and only the top variant),
+%
+    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
+    satisfied,
+%
+    (3) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') : \MPT) ~
+    \mathit{tail}(\vs)$ is coinductively satisfied
+%
+    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
+    \pol') : \MPT) ~ vs$.
 
@@
     instruction to which execution returns immediately after completing a call
-    is itself a call instruction, the last two cases are combined.
+    is itself a call instruction, the \emph{call step} and \emph{return step}
+    cases are combined.
 

### 2020-07-08 main.tex
Definition 6.6

@@
 
-\definition{Top-level stack testing property}
-%
-  \ldots wrap the coinductive stack testing property
+\definition
+%
+  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
+  satisfies the eager testing property at depth 1 with regard to the initial
+  variant, $\mathit{stack}_T ~ 1 ~ \MPT ~ \vm{\contour_0}$.
+%
+  \rb{I don't think the initial variant has been defined, but it's
+    straightforward.}
 

### 2020-07-08 main.tex
Some notation for Section 6

--- 2020-07-08 main.tex
@@
 
+\newcommand{\vse}{\mathit{vse}}
+\newcommand{\VSES}{\mathit{VSE}}
+\newcommand{\vs}{\mathit{vs}}
+\newcommand{\VSS}{\mathit{VS}}
+
+\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
+\newcommand*{\testInt}[3]{\ensuremath{\mathit{int}_T ~ #1 ~ #2 ~ #3}}
+\newcommand*{\testConf}[2]{\ensuremath{\mathit{conf}_T ~ #1 ~ #2}}
+\newcommand*{\testVS}[2]{\ensuremath{\mathit{\vs}_T ~ #1 ~ #2}}
+\newcommand*{\testProp}[3]{\ensuremath{\mathit{stack}_T ~ #1 ~ #2 ~ #3}}
+
 \definition
@@
   \[
-    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
+    \testInt{\contour}{\mach}{\mach'} \triangleq
     \forall \component . ~
@@
   first two and it may be clearer to elide them.}
-\newcommand{\vse}{\mathit{vse}}
-\newcommand{\VSES}{\mathit{VSE}}
-\newcommand{\vs}{\mathit{vs}}
-\newcommand{\VSS}{\mathit{VS}}
-
-\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
 
@@
   denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
-  satisfying the stack integrity test $\mathit{int}_T ~ C
-  ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~
-  \mach ~ \mach'$ with regard to the original machine state. We write
-  this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
+  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
+  and stack confidentiality test $\testConf{\mach}{\mach'}$
+  with regard to the original machine state. We write
+  this $\testVS{(\mach, \pol)}{\vs}$.
 
@@
     (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
-    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
+    $\testProp{d}{[(\mach, \pol)]}{\vs}$, for any $d$ up to the size of
     $\vs$. (That is, a violation was detected $d$ levels into the stack.
@@
   \item If the top variant of the stack (and only the top variant) returns,
-    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
+    %i.e., $\mathit{ret}(\head(\vs), \mach)$ holds,
+    the property is safisfied for
     the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
-    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
+    stack: $\testProp{|\vs|}{[(\mach, \pol)]}{\vs}$. \rb{Explain the
     use of this vs the return case below.}
@@
 %
-    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
+    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
     satisfied, and
 %
-    (3) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') : \MPT) ~
-    \vs'$ is coinductively satisfied % at depth $d$
+    (3) the stack safety test $\testProp{d}{((\mach', \pol') : \MPT)}{\vs'}$
+    is coinductively satisfied % at depth $d$
     %% and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
@@
 %
-    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
-    \pol') : \MPT) ~ vs$.
+    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach', \pol') :
+    \MPT)}{vs}$.
     %% \rb{omitting stack size condition}
@@
 %
-    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
+    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
     satisfied,
 %
-    (3) the stack integrity test $\mathit{int}_T ~ \contour ~ \mach ~ \mach'$ is
-    satisfied and the stack confidentiality test $\mathit{conf}_T ~ \mach ~
-    \vm{\contour}$ is satisfied with regard to an arbitrary $\contour$-variant
+    (3) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
+    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
+    is satisfied with regard to an arbitrary $\contour$-variant
     of $\mach$,
 %
-    and (4) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') :
-    \MPT) ~ (\vm{\contour} : \vs')$ is coinductively satisfied, for an arbitrary
+    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
+    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
     $\contour$-variant of the machine,
 %
-    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
-    \pol') : \MPT) ~ vs$.
+    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
+    \pol') : \MPT)}{vs}$.
 
@@
 %
-    If (1) the machine takes a \emph{return step} at the top of the variant
-    stack (and only the top variant),
-%
-    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
+    If (1) the machine takes a \emph{return step}, so the top of the variant
+    stack (and only the top variant) returns,
+%
+    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
     satisfied,
 %
-    (3) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') : \MPT) ~
-    \mathit{tail}(\vs)$ is coinductively satisfied
-%
-    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
-    \pol') : \MPT) ~ vs$.
+    (3) the stack safety test $\testProp{d}{((\mach', \pol') :
+    \MPT)}{\mathit{tail}(\vs)}$ is coinductively satisfied
+%
+    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
+    \pol') : \MPT)}{vs}$.
 
@@
 
-\definition{Top-level stack testing property}
-%
-  \ldots wrap the coinductive stack testing property
+\definition
+%
+  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
+  satisfies the eager testing property at depth 1 with regard to the initial
+  variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
+%
+  \rb{I don't think the initial variant has been defined, but it's
+    straightforward.}
 

### 2020-07-08 main.tex
Typo

@@
 
-  \item If the machine takes \emph{simultaneous return-call step}, that is, the
-    instruction to which execution returns immediately after completing a call
-    is itself a call instruction, the \emph{call step} and \emph{return step}
-    cases are combined.
+  \item If the machine takes a \emph{simultaneous return-call step}, that is,
+    the instruction to which execution returns immediately after completing a
+    call is itself a call instruction, the \emph{call step} and \emph{return
+    step} cases are combined.
 

### 2020-07-08 main.tex
Note on the StkTokens proof technique

@@
 
+\rb{As far as I can tell, while their notion of components and interfaces seems
+  standard, observable behavior appears to be limited to cotermination. The
+  previous paper makes explicit that (at least at that point) they are not
+  interested in externally observable side effects like console output or memory
+  access traces.}
+
 We are confident \rb{how strong a claim to make, with substantiation} that

### 2020-07-08 main.tex
CP - 6. keeping lock

@@
 
-\definition
-  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
-  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
-  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
-    confidentiality test} if their observations and changed components coincide:
+\definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
+\emph{stack confidentiality test} if the machines they stepped to
+have their changed components coincide and the observations during
+these steps are equal:
 \[
-    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
-    \obs_1 = \obs_2 \wedge \forall \component . ~
+\begin{array}{rl}
+  \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \triangleq &
+  \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
+  \mach_1 \stepstoobs{\obs_1} \mach_1' \wedge \mach_2 \stepstoobs{\obs_2} \mach_2' \Rightarrow \\
+&    \obs_1 = \obs_2 \wedge \forall \component . ~
    (     \mach_1(\component) \neq \mach_1'(\component)
@@
     \rightarrow \mach_1'(\component) = \mach_2'(\component)
+\end{array}
   \]
-\apt{scoping of $\forall$ is ambiguous.}
-\rb{Fix notations, including for properties below.
-  In particular, note that the last four arguments could be computed from the
-  first two and it may be clearer to elide them.}
-
-\definition The testing property keeps track of relevant information
-on a {\em variant stack}: a list of machine states, each element of
-which (denoted $\vm{C}$) of which is a variation of the actual machine
+
+\definition The testing property keeps track of a variation machine
+for every nested call performed so far on a {\em variant stack}: a
+list of machine states annotated with a contour, each element of which
+(denoted $\vm{C}$) of which is a variation of the actual machine
 according to a contour $C$.
@@
 
-\definition{Eager testing property}
-  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
-  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
-  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
-  testing property} at a given call depth as follows.
+\definition
+%  Let $\callmap$ a call map and $\codemap$ a function map.
+  Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
+  variant stack $\vs$, we say that an MP-trace induced from $(\mach,
+  \pol)$ satisfies the \emph{strong stack safety} at a given call
+  depth as follows.
 % Refactor repetitive notation before case split
-  In all cases that assume a machine step, $(\mach, \pol) \hookrightarrow_P
-  (\mach', \pol', \obs)$, assume also (unless the step is a return) that the
-  variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
-  machines can step.
+  In all cases that let $(\mach, \pol) \hookrightarrow_P
+  (\mach', \pol', \obs)$.
+  %and assume (unless the step is a return) that the
+  %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
+  %machines can step.
 
@@
   \begin{itemize}
+
+  \item
+%
+    If the machine takes a \emph{call step} and $\contour$ is the contour at
+    the call then,
+%
+    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
+    satisfied,
+%
+    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
+    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
+    is satisfied with regard to an arbitrary $\contour$-variant
+    of $\mach$,
+%
+    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
+    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
+    $\contour$-variant of the machine,
+%
+    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
+    \pol') : \MPT)}{vs}$.
 
@@
     %% \rb{omitting stack size condition}
-
-  \item
-%
-    If (1) the machine takes a \emph{call step} and $\contour$ is the contour at
-    the call,
-%
-    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
-    satisfied,
-%
-    (3) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
-    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-    is satisfied with regard to an arbitrary $\contour$-variant
-    of $\mach$,
-%
-    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
-    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
-    $\contour$-variant of the machine,
-%
-    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
-    \pol') : \MPT)}{vs}$.
 

### 2020-07-08 main.tex
Intro to capability machines in threat model

@@
 %
-\rb{TODO Here the policy is built into the hardware, although the extended ISA
-  could be split into a machine without capabilities and capabilities as the
-  policy.}
+Other policies are implemented in hardware or built on such hardware mechanisms.
+Capability machines like CHERI \rb{Woodruff+14} extend conventional
+architectures to support efficient and fine-grained control over memory
+accesses, which can be used to implement security policies involving, for
+example, compartmentalization or the stack protection policies of
+\citet{Skorstengaard+19,Skorstengaard+19b}. Mapping the underlying protection
+mechanisms to policy states and step function would be a delicate operation, as
+those are tightly coupled to the hardware design. However, separating the
+enforcement aspects of a concrete policy (such as StkTokens, where \(\POLS\)
+would encompass details like the abstract call stack, the capability sealing
+scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
+the machine would be relatively straightforward.
 

### 2020-07-09 main.tex
done with pass through 5.3; releasing lock

--- 2020-07-08 main.tex
@@
 %
-\[\MPS = \MACHS \times \POLS\]
+\[\mpstatename \in \MPS = \MACHS \times \POLS\]
 %
@@
 \(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
-\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
+\(\MPT_{\mathit{call}}\) and \(\machT\) with respect to \(\mathit{Cof}(\mach_c, n)\)
 and \(R = \ret{\mach_c}\):
@@
 Both the integrity and confidentiality components of local state
-encapsulation concern themselves with the accessibility of data, and
-may hold on a wide range of programs that violate expected control
+encapsulation concern themselves only with the accessibility of data, and
+may still hold on a wide range of programs that violate expected control
 flow. Say, if {\tt g} managed to jump in the middle of the
@@
 %
-Stack safety is traditionally associated with preventing such unwanted
+Informal definitions of stack safety usually \apt{evidence?} aim to prevent such unwanted
 flows: every jump between functions should either be a call or a
@@
 security properties that rely on the underlying knowledge of the
-program structure (call maps, function maps, return maps, and entry
-maps): {\em control separation} ensures the validity of transition
+program structure (call map $\callmap$, function map $\codemap$, return map $\retmap$, and entry
+map $\entmap$): {\em control separation} ensures the validity of transition
 sources between functions, {\em entry integrity} ensures the validity
@@
 
-\definition{Control Separation}
-
-A system enjoys {\em control separation} with regard to a function map
-\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
-if for any initial state \((\mach_0, \pol_0)\) and any adjacent pair
+\definition%{Control Separation}
+A system enjoys {\em control separation}
+% with respect to a function map
+%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
+if, for any initial state $\mpstatename$ and any adjacent pair
 of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
-trace from \(\mpstate{\mach_0}{\pol_0})\) whose current instruction
+trace from $\mpstatename$ whose current instruction
 belongs to different functions according to the function map,
-either $\mach_1$ is a valid call or a valid return:
+$\mach_1$ is either a valid call or a valid return:
 
@@
 
-For example, consider the unsafe substitution:
-\[
-5 : ~ \jalr ~ r_4 ~ \rra ~ 0
-\]
-This violates control separation as the instruction that causes a
+For example, consider changing line 5 of the running example to a $\jalr$,
+$\PCname \leftarrow r_4$, with effect $\PCname \leftarrow 42$, causing {\tt main}
+to jump into the middle of {\tt h}. 
+This violates control separation because the instruction that causes a
 function switch is not a valid call (or a return) according to the
@@
 
-\definition{Entry Integrity}
-
-A system enjoys {\em entry integrity} with regard to a call map
-\(\callmap\) and entry map \(\entmap\) if for any initial state
-\((\mach_0, \pol_0)\) and any n-transition \((\mach_c,\pol_c)\) and
+\definition%{Entry Integrity}
+A system enjoys {\em entry integrity}
+%with respect to a call map
+%\(\callmap\) and entry map \(\entmap\)
+if, for any initial state $\mpstatename$
+and any n-transition \((\mach_c,\pol_c)\) and
 \((\mach_e, \pol_e)\) in the induced trace from
-\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
+$\mpstatename$, $\mach_e$ is a member of the entry map:
 \[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]
 
-The same unsafe substitution violates entry integrity
-\[
-5 : ~ \jalr ~ r_4 ~ \rra ~ 0
-\]
-as its target (instruction 42) is not an established entry point for
-{\tt h} according to the entry map.
+The changed program described just above also violates entry integrity,
+because the target (instruction 42) of the new $\jalr$ is not an established entry point
+according to the entry map.
 
@@
 
-\definition{Return Integrity}
-
-\newcommand*{\Unmatched}{\mathit{Unmatched}}
+\newcommand*{\Unmatched}{\mathit{um}}
 
 Return integrity captures the expected behavior of marked returns with
-regard to their call points. Given a call map $\callmap$ and a return
-map $\retmap$, we begin by defining a coinductive relation
-$\Unmatched$ over two $\MPOTRACE$ $\MPT$ and $\MPT'$, if $\MPT'$ is a
-suffix of $\MPT$ whose head is the first unmatched return in $\MPT$.
-
-Formally:
+respect to their call points. Intuitively, it says that the marked return
+matching a call $\mach_c$ should satisfy $\ret{\mach_c}$.
+Observe that, given an MP-trace $\mpstate{\mach_c}{\pol}\MPT$, if $\mach_c$ is a call state then
+the matching return will be the first \emph{unmatched} return in $\MPT$. So
+we begin by defining an auxiliary ``unmatched'' relation between two MP-traces.
+
+\definition
+We write $\Unmatched~\MPT~\MPT'$ if $\MPT'$ is
+obtained from $\MPT$ by dropping elements until we reach the first unmatched return
+(according to the return map).  Formally, $\Unmatched$ is defined coinductively as follows:
 \begin{itemize}
@@
 
+\begin{minipage}{.3\textwidth}
   \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
@@
               {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
+\end{minipage}
+\hspace*{0.1\textwidth}
+\begin{minipage}{.3\textwidth}
   \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
@@
               {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}
+\end{minipage}
 
@@
               \hspace{.5cm} \neg \retmap(\mach(\PCname))
-              \Unmatched ~ \MPT ~ \MPT'\)}
+              \hspace{.5cm}\Unmatched ~ \MPT ~ \MPT'\)}
            {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}
 
-\item Finally, if a trace starts with the call, then we must discard the first
-  unmatched return to get to the real one:
+\item Finally, if a trace starts with a call, then we must discard the first
+  unmatched return from its tail to get to the real one:
 
@@
 \end{itemize}
-A system enjoys {\em return integrity} with regard to a call map
-\(\callmap\) and return map \(\retmap\) if for any initial state pair
-\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
+\apt{Could cut formal definition for space.}
+
+\definition%{Return Integrity}
+A system enjoys {\em return integrity}
+% with respect to a call map \(\callmap\) and return map \(\retmap\)
+if, for any initial state pair 
+$\mpstatename$ and any state pair \((\mach_c, \pol_c)\)
 corresponding to a call in its induced trace, if
-\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
-is the first unmatched return such that for some $\pol_r$
-\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
-then $\mach_r$ corresponds to a return state for $\mach_c$:
-\[\ret{\mach_c}\ \mach_r\]
-
-For example, if we added an instruction between 45 and 46:
+\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
+is the first unmatched return in $\MPT$, then $\mach_r$ corresponds
+to a return state for $\mach_c$:
+
 \[
-45.5: ~ \addi ~ \rra ~ r_0 ~ \negate 19
+\exists p_r.\Unmatched ~ \MPT ~ (m_r,p_r)\MPT' \Rightarrow \ret{\mach_c}\ \mach_r.
 \]
-This would cause the return from {\tt h} to go to instruction 8 and
+\apt{I've tried to tidy this, but I don't think it corresponds to the Coq.
+  Shouldn't we apply $\Unmatched$ to the tail of $\MPT$?
+  And what about the no-unmatched-return-at-top-level clause?}
+
+For example, if we modified the running example by adding
+an instruction $ \rra \rightarrow \rra - 19$
+between 45 and 46, this would cause the return from {\tt h} to go to instruction 8 and
 return to {\tt main} instead of {\tt g}. While the instruction that
@@
 separation), the state after the return would not correspond to an
-$\Unmatched$ return for {\tt g}'s call to {\tt h}, violating return
+unmatched return for {\tt g}'s call to {\tt h}, violating return
 integrity.
@@
 explicit entry points, and returns respect implicit entry points
-corresponding to their call.  A system that enjoys them with regard to
-a code, a call, a return, and an entry map is said to enjoy
-{\em well-bracketed control flow} with regard to those maps.
+corresponding to their call.  A system that enjoys all three
+properties (with respect to a function map, call map, return map,
+and entry map) is said to enjoy
+{\em well-bracketed control flow} (with respect to those maps).
 
@@
   and stack confidentiality test $\testConf{\mach}{\mach'}$
-  with regard to the original machine state. We write
+  with respect to the original machine state. We write
   this $\testVS{(\mach, \pol)}{\vs}$.
@@
   A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
-  satisfies the eager testing property at depth 1 with regard to the initial
+  satisfies the eager testing property at depth 1 with respect to the initial
   variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
@@
 
-      A system enjoys {\em observable stack integrity} with regard to
+      A system enjoys {\em observable stack integrity} with respect to
       \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
@@
       confidentiality}. A system enjoys {\em observable stack confidentiality}
-      with regard to \(\callmap\) and \(\codemap\) if for any initial state
+      with respect to \(\callmap\) and \(\codemap\) if for any initial state
       \(\mach_0\) and each transition with entry state
@@
     \definition{Observable Stack Safety}
-      A system enjoys {\em observable stack safety} with regard to a call map
+      A system enjoys {\em observable stack safety} with respect to a call map
       and a function map if it enjoys both observable stack integrity and
-      observable stack confidentiality with regard to them.
+      observable stack confidentiality with respect to them.
 

### 2020-07-09 main.tex

--- 2020-07-08 main.tex
@@
 %
-Other policies are implemented in hardware or built on such hardware mechanisms.
-Capability machines like CHERI \rb{Woodruff+14} extend conventional
-architectures to support efficient and fine-grained control over memory
-accesses, which can be used to implement security policies involving, for
-example, compartmentalization or the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}. Mapping the underlying protection
-mechanisms to policy states and step function would be a delicate operation, as
-those are tightly coupled to the hardware design. However, separating the
-enforcement aspects of a concrete policy (such as StkTokens, where \(\POLS\)
-would encompass details like the abstract call stack, the capability sealing
-scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-the machine would be relatively straightforward.
+\rb{TODO Here the policy is built into the hardware, although the extended ISA
+  could be split into a machine without capabilities and capabilities as the
+  policy.}
 
@@
 steps for a long time before producing some output.
-
-%Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
-%We then define {\em policy-sensitive observational equivalence}: a
-%trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
-%\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
-%and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
-%policy fault) and its observations are a prefix of those of M.
-%
-%% Fixed by Sean
-%%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
-%\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
-%            {\(\MPT \sim_P \machT\)}
-%
-%\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
-%            {\(\MPT \sim_P \machT\)}
 
@@
 %
-Special care is needed for the code memory corresponding to the
-callee: they are marked low confidentiality if they belong to the
-callee and high confidentiality otherwise; regardless they are always
-high integrity.\apt{Sean:I think this is vestigial from an aborted version
-  in which you were modeling WBCF using confidentiality. In any casen
-  it is not reflected in the definition below.}
+Code is always high integrity and low confidentiality.
 %
@@
 \definition
-Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
-$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
-respect to $\contour$ if its final state (if any) agrees with its
-initial one on all components marked $\HIGHSEC$ in $\contour$:
+
+Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
+\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
+respect to \(\contour\) if its final state (if any) agrees with its
+initial one on all components marked \(\HIGHSEC\) in \(\contour\):
 %
 \[\begin{split}
-  \mathit{int} ~ \contour ~ \MPT \triangleq &
-    \head(\MPT) = \mpostate{\mach_e}{\_}{\_}  \land 
-    \last(\MPT) = \mpostate{\mach_r}{\_}{\_} \Rightarrow \\
+  \propdef{TraceIntegrity} ~ \contour ~ \MPT \triangleq &
+    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
+  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
   & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
@@
 \end{split}\]
-%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
-%\sna{We could do it like this, with finiteness implicit in the existence of
-%     a final state.}+1
+\sna{Experimenting with font on the mathy version of the property.}
 
@@
 and final states colored to show the contour; high 
-integrity memory is in red\apt{NO?}, while low integrity is in green\apt{NO?}.
-At the end of the trace high integrity address 103 should match the initial state, and does not.
-So trace integrity is violated.
+integrity memory is in \high, while low integrity is in \low.
+At the end of the trace high integrity address 103 should match the initial
+state, and does not. So trace integrity is violated.
 \rb{There are some color name macros, but none for integrity yet? Use instead
@@
 for every subtrace delimited by a call and its corresponding return.
-%for every transition between a caller and a callee for the entirety of
-%the trace of the callee.\apt{?????}
+
+\definition Given a trace \(\MPT\) and a call map \(\callmap\)
+%and a function map \(\codemap\),
+an \(n\)-{\em transition} is a pair of
+machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
+\(\mpstate{\mach_e}{\pol_e}\) in the trace such that
+\(\mach_c\) is a valid call with \(n\) arguments, and
+\(\mpstate{\mach_c}{\pol_c}\) steps to \(\mpstate{\mach_e}{\pol_e}\).
+
+\apt{or is in some function?? and why do we need either??}:
+\sna{I think Andrew is right that we don't need to know the function
+     here, or that the entry point is valid; that's part of WBCF.}
+%and \(\mach_e\) is an entry point for some function \(f\):
+
+  \[\begin{split}
+    \mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}
+    \triangleq & \exists \obs, f.
+    ~ \mpstate{\mach_c}{\pol_c} \stepstopol{\obs} \mpstate{\mach_e}{\pol_e} \\
+    & \land \callmap(\mach_c(\PCname)) = n
+      %\land \codemap(\mach_e(\PCname)) = f
+  \end{split}\]
+
+\definition
+
+{\em Stack integrity} is a property of a system with regard to a call map
+\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
+and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
+on the prefix up to its return, if any. For any initial state \(\mach_0\)
+we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
+\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
+\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
+the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
+\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
+\(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
+system as a whole enjoys stack integrity with regard to \(\callmap\).
 
 \apt{Some more words needed here to smooth the technicalities.}
-
-\definition%{Transition}
-Given a trace $\MPT$, an $n$-{\em transition} is a pair of
-machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
-in the trace such that:
-\begin{itemize}
-\item one state pair steps to the other:
-  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
-\item $\mach_c$ is a valid call with $n$ arguments:
-  $$\callmap(\mach_c(\PCname)) = n,$$
-\item and $\mach_e$ is an entry point\apt{or is in some function?? and why do we need either??}:
-  $$\exists f. \codemap(\mach_e(\PCname)) = f.$$
-\end{itemize}
-
-\definition%{Stack Integrity}
-A system satisfies {\em stack
-  integrity} if, for every
-initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
-trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
-$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
-$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
-(if any) satisfies trace integrity with respect to the contour at the
-call:
-%
-$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
-\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
-\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$
+\sna{Attempt...}
 
@@
 Unlike integrity, confidentiality is a noninterference-style property,
-comparing the behavior of a machine with an arbitrary variant that
+comparing the behavior of a machine with an arbitrary variation that
 preserves low confidentiality data.
@@
 \definition%{Variations}
-We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
-another with respect to a contour $\contour$ if they agree on low
+We say two machine states \(\mach\) and \(\mach'\) are {\em variations}
+of one another given a contour \(\contour\) if they agree on low
 confidentiality data:
@@
 Then, the trace version of confidentiality operates on two traces: the
-``original'' tracce $MP \in \MPOTRACE$ that corresponds to a callee $f$
-and a ``variant'' trace $M \in \MTRACE$ that corresponds to how $f$
+``original'' trace $\MPT \in \MPOTRACE$ that corresponds to a callee $f$
+and a ``variant'' trace $\machT \in \MTRACE$ that corresponds to how $f$
 would behaves if it were started from a variant initial machine state.
@@
 
-\definition%{Trace Confidentiality}
-%% Suppose $m_e \approx_\contour m_e'$.
-%% Two traces $MP$ with $\head(MP) = m_e$ and $M$ with $head(M) = m_e'$
-%% satisfy {\em trace confidentiality} with respect to $\contour$ and
-%% a return predicate \(R \subset \MACHS\) if:
-Two traces $MP$, with head $m_e$, and $M$, with head $m_e'$,
-satisfy {\em trace confidentiality} with respect to contour $\contour$ and
-a return predicate \(R \subset \MACHS\) if, whenever $m_e \approx_\contour m_e'$ then:
+\definition
+
+{\em Trace confidentiality} will be a property of two traces, \(\MPT\) and
+\(\machT\), with regard to a return predicate \(R \subset \MACHS\) that
+reflects the states that are considered to have returned. We abstract over
+\(R\) because what it means to return depends on the state from which
+\(f\) was called. There are three cases in which \(\MPT\) and \(\machT\)
+enjoy trace confidentiality with respect to \(R\).
+
+\sna{Interestingly, split out like this, we don't need to talk about
+     contour here -- they are varied in the next property.}
 \begin{itemize}
-\item If $\MPT$ is finite with final state $\mach_r$ and $\mach_r \in R$, 
-  then $\machT$ is also finite with some final state $\mach_r'$ and
-  any changes from the initial states have changed in the same way:
-$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
-  \vee \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
-  \mach_r(\component) = \mach_r'(\component)$$
-  In addition their observations must be similar:
-  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
-\item If $\MPT$ is infinite, then $\machT$ is also infinite and their
-  observations are similar:
-  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
-\item If the $\MPT$ ends prematurely, without fulfilling \(R\), then
-  we require that its observations are a prefix of those of $\machT$:
-  $$\pi_o(\MPT) \lesssim \pi_o(\machT)$$
+  \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
+    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
+    some final state \(\mach_r'\) and any changes from the initial states are
+    matched between \(\mach_r\) and \(\mach_r'\). In addition their
+    observations are similar.
+    \[\begin{split}
+      \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
+      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
+        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
+        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
+      & ~ (\mach_e(\component) \not = \mach_r(\component)
+        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
+      \mach_r(\component) = \mach_r'(\component))
+    \end{split}\]
+  \item If \(\MPT\) is infinite, then \(\machT\) is also infinite and their
+  observations are similar.
+    \[\begin{split}
+      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
+      & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
+      & \neg (\exists mo . \last(\machT) = mo)
+        \land \pi_o(\MPT) \eqsim \pi_o(\machT)
+    \end{split}\]
+\item If \(\MPT\) ends prematurely without fulfilling \(R\), then
+  its observations are a prefix of those of \(\machT\).
+    \[\begin{split}
+      \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
+      & \forall \mach_r ~ \pol_r ~ \obs_r .
+        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+      & \neg ~ R ~ \mach_r \Rightarrow \\
+      & \pi_o(\MPT) \lesssim \pi_o(\machT)
+    \end{split}\]
 \end{itemize}
-\apt{Needs words about why we need a return predicate abstraction.}
+
+Putting it all together:
+
+\[\begin{split}
+  \propdef{TraceConfidentiality} ~ R ~ \MPT ~ \machT \triangleq
+  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
+  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
+    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
+\end{split}\]
+
+In the non-return cases, we only care that the traces exhibit the same
+observable behavior (or a prefix if the trace halts prematurely), as changes
+in the state will not affect the caller absent a return.  In the return case
+we must additionally guarantee that no changes were made based on secret data
+that could corrupt the caller upon return.
 
@@
 shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
-in {\high} and low confidentiality in \low. Below it, the variant state
+in {\high} and low confidentiality in {\low}. Below it, the variant state
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
-hold -5, in orange.\apt{NO?}  After the last step, the return states \(\mach_r\) and
+hold -5, in {\varied}.  After the last step, the return states \(\mach_r\) and
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
-\(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in light {\low}, and while 109
 had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
@@
 output it, violating the internal observation clause of the confidentiality definition.
-%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
-%and \(\mach_e'\), below. Low confidentiality data in green is shared
-%between them, while high confidentiality data in red and yellow may differ.
-%After executing the code shown, the final states below have written different
-%values to address 42, represented by the black and white components.
-%It is allowed for address 40 to differ between the states due to the initial
-%variation, but address 42 changed differently between the traces, representing
-%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
-%is read, it is overwritten before return, and the write at step 3 always writes
-%the same value. \Gprivc and red locations differ between variations only because
-%of the initial variation, and the rest are identical. So this example passes.
 
@@
 
-%\Cref{fig:confex2} gives an example of a trace that fails trace
-%confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
-%Since {\tt x}, at address 100, is high confidentiality (red) the variant
-%may have a different value (yellow). So both \(r_4\) and address 107
-%differ in the final state between the original trace and the variant one,
-%failing the ending condition of trace confidentiality.
-
 To lift trace confidentiality to a system property, we follow the same
@@
 %Let $\callmap$ be a call map and $\codemap$ a function map.
-We say that a
-system satisfies {\em stack confidentiality}
-% with respect to $\callmap$ and $\codemap$
-if, for every initial machine-policy state
-pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
-$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
-in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
-\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
-For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
-\mach_e\), let \(\machT\) be the similarly prefixed trace such that
-\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
-\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
-and \(R = \ret{\mach_c}\):
-
-$$\forall \MPT_{\mathit{call}} ~ \machT. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
-$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, n)} \mach_e' \wedge
-\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$
+We will say that a system satisfies {\em stack confidentiality} with respect to
+\(\callmap\) in a similar way to stack integrity. For every initial state
+\(\mach_0\), we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)}
+\hookrightarrow \MPT\). Then for any \(n\)-transition
+\(\mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~
+\mpstate{\mach_e}{\pol_e}\), we take the prefix trace until return:
+%
+\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\]
+
+We also take the corresponding trace from any variant \(\mach_e' 
+\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
+%
+\[\mach_e' \hookrightarrow \machT_{call} | \ret{\mach_c}\]
+
+If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
+\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
+stack confidentiality with regard to \(\callmap\).
 
 \apt{oy. still needs work to make this intelligible}
+\sna{Attempted. There's a tension between introducing all the variables
+     and keeping a good flow that I'm not sure I'm managing here though.}
 
@@
 
-\newcommand{\vse}{\mathit{vse}}
-\newcommand{\VSES}{\mathit{VSE}}
-\newcommand{\vs}{\mathit{vs}}
-\newcommand{\VSS}{\mathit{VS}}
-
-\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
-\newcommand*{\testInt}[3]{\ensuremath{\mathit{int}_T ~ #1 ~ #2 ~ #3}}
-\newcommand*{\testConf}[2]{\ensuremath{\mathit{conf}_T ~ #1 ~ #2}}
-\newcommand*{\testVS}[2]{\ensuremath{\mathit{\vs}_T ~ #1 ~ #2}}
-\newcommand*{\testProp}[3]{\ensuremath{\mathit{stack}_T ~ #1 ~ #2 ~ #3}}
-
 \definition
@@
   \[
-    \testInt{\contour}{\mach}{\mach'} \triangleq
+    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
     \forall \component . ~
@@
 
-\definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
-\emph{stack confidentiality test} if the machines they stepped to
-have their changed components coincide and the observations during
-these steps are equal:
+\definition
+  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
+  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
+  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
+    confidentiality test} if their observations and changed components coincide:
   \[
-\begin{array}{rl}
-  \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \triangleq &
-  \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
-  \mach_1 \stepstoobs{\obs_1} \mach_1' \wedge \mach_2 \stepstoobs{\obs_2} \mach_2' \Rightarrow \\
-&    \obs_1 = \obs_2 \wedge \forall \component . ~
+    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
+    \obs_1 = \obs_2 \wedge \forall \component . ~
    (     \mach_1(\component) \neq \mach_1'(\component)
@@
     \rightarrow \mach_1'(\component) = \mach_2'(\component)
-\end{array}
   \]
-
-\definition The testing property keeps track of a variation machine
-for every nested call performed so far on a {\em variant stack}: a
-list of machine states annotated with a contour, each element of which
-(denoted $\vm{C}$) of which is a variation of the actual machine
+\apt{scoping of $\forall$ is ambiguous.}
+\rb{Fix notations, including for properties below.
+  In particular, note that the last four arguments could be computed from the
+  first two and it may be clearer to elide them.}
+\newcommand{\vse}{\mathit{vse}}
+\newcommand{\VSES}{\mathit{VSE}}
+\newcommand{\vs}{\mathit{vs}}
+\newcommand{\VSS}{\mathit{VS}}
+
+\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
+
+\definition The testing property keeps track of relevant information
+on a {\em variant stack}: a list of machine states, each element of
+which (denoted $\vm{C}$) of which is a variation of the actual machine
 according to a contour $C$.
@@
   \emph{variant stack test} if for each variant element $\vm{C}$
-  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
-  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
-  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
-  and stack confidentiality test $\testConf{\mach}{\mach'}$
-  with regard to the original machine state. We write
-  this $\testVS{(\mach, \pol)}{\vs}$.
-
-\definition
-%  Let $\callmap$ a call map and $\codemap$ a function map.
-  Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
-  variant stack $\vs$, we say that an MP-trace induced from $(\mach,
-  \pol)$ satisfies the \emph{strong stack safety} at a given call
-  depth as follows.
-% Refactor repetitive notation before case split
-  In all cases that let $(\mach, \pol) \hookrightarrow_P
-  (\mach', \pol', \obs)$.
-  %and assume (unless the step is a return) that the
-  %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
-  %machines can step.
+  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step
+  $\mathit{curr\_variant}(\vse) \stepstoobs{\obs} \mach'$ while
+  satisfying the stack integrity test $\mathit{int}_T ~ \contour(\vse)
+  ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~
+  \mach ~ \mach'$ with regard to the original machine state. We write
+  this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
+
+\definition{Eager testing property}
+  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
+  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
+  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
+  testing property} at a given call depth if:
 
@@
 
-  \item
-%
-    If the machine takes a \emph{call step} and $\contour$ is the contour at
-    the call then,
-%
-    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
-    satisfied,
-%
-    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
-    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-    is satisfied with regard to an arbitrary $\contour$-variant
-    of $\mach$,
-%
-    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
-    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
-    $\contour$-variant of the machine,
-%
-    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
-    \pol') : \MPT)}{vs}$.
-
-  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
+  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow
     (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
-    $\testProp{d}{[(\mach, \pol)]}{\vs}$, for any $d$ up to the size of
+    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
     $\vs$. (That is, a violation was detected $d$ levels into the stack.
@@
   \item If the top variant of the stack (and only the top variant) returns,
-    %i.e., $\mathit{ret}(\head(\vs), \mach)$ holds,
-    the property is safisfied for
+    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
     the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
-    stack: $\testProp{|\vs|}{[(\mach, \pol)]}{\vs}$. \rb{Explain the
+    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
     use of this vs the return case below.}
@@
     %
-    (1) the machine takes a \emph{non-call, non-return step},
-%
-    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
-    satisfied, and
-%
-    (3) the stack safety test $\testProp{d}{((\mach', \pol') : \MPT)}{\vs'}$
-    is coinductively satisfied % at depth $d$
-    %% and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
-    %% \pol')$,
-%
-    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach', \pol') :
-    \MPT)}{vs}$.
-    %% \rb{omitting stack size condition}
+    (1) the machine takes a \emph{non-call, non-return step} $(\mach,
+    \pol) \hookrightarrow_P (\mach', \pol', \obs)$ and the variant stack steps
+    to $\vs \hookrightarrow \vs'$,
+    %
+    (2) the variant stack test $\mathit{\vs}_T ~
+    (\mach, \pol) ~ \vs$ is satisfied, and
+    %
+    (3) the stack safety test
+    $\mathit{stack}_T ~ d ~ \MPT ~ \vs'$ is coinductively satisfied at depth $d$
+    and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
+    \pol')$,
+    %
+    {\em then} we conclude $\mathit{stack}_T d ~ ((\mach, \pol) : \MPT) ~ vs$.
+    \rb{omitting stack size condition}
+
+  \item
+%
+    If (1) the machine takes a \emph{call step} $(\mach, \pol) \hookrightarrow_P
+    (\mach', \pol', \obs)$ and the variant stack steps to $\vs \hookrightarrow
+    \vs'$,
+%
+    (2) $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$,
+%
+    (3) for the new contour at the call, $\contour$, \rb{definition?} the
+    machine satisfies the stack integrity test $\mathit{int}_T ~ \contour ~
+    \mach ~ \mach'$ and the stack confidentiality test with regard to any
+    variant $\mathit{variant} ~ C ~ \mach ~ \mach_{\mathit{var}}$,
+    $\mathit{conf}_T ~ \mach ~ \mach_{\mathit{var}}$ \rb{definition?},
+%
+    and (4) the stack safety test is coinductively satisfied at depth $d$ for a
+    continuing trace $\MPT$ where $\head(\MPT) = (\mach', \pol')$ after an
+    arbitrary variant machine derived from $\contour$ has been added to the top
+    of the stack, $\mathit{stack}_T ~ d ~ \MPT ~ (\vse : \vs')$, \rb{define
+      $\vse$}
+%
+    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : \MPT) ~ vs$.
 
@@
 
-  \item
-%
-    If (1) the machine takes a \emph{return step}, so the top of the variant
-    stack (and only the top variant) returns,
-%
-    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
-    satisfied,
-%
-    (3) the stack safety test $\testProp{d}{((\mach', \pol') :
-    \MPT)}{\mathit{tail}(\vs)}$ is coinductively satisfied
-%
-    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
-    \pol') : \MPT)}{vs}$.
-
-  \item If the machine takes a \emph{simultaneous return-call step}, that is,
-    the instruction to which execution returns immediately after completing a
-    call is itself a call instruction, the \emph{call step} and \emph{return
-    step} cases are combined.
+  \item If the machine takes a \emph{return step} at the top of the variant stack
+    $(\mach, \pol) \hookrightarrow (\mach', \pol', \obs)$,
+    the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    if (in addition to the general step conditions)
+    the property recurses at depth $d$ with trace $\MPT$
+    after removing the returning top variant from the stack.
+
+  \item If the machine takes \emph{simultaneous return-call step}, that is, the
+    instruction to which execution returns immediately after completing a call
+    is itself a call instruction, the last two cases are combined.
 
@@
 
-\definition
-%
-  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
-  satisfies the eager testing property at depth 1 with regard to the initial
-  variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
-%
-  \rb{I don't think the initial variant has been defined, but it's
-    straightforward.}
+\definition{Top-level stack testing property}
+%
+  \ldots wrap the coinductive stack testing property
 
@@
       \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
-      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109, in \tainted,
+      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109
       are not restored; every address that was varied and not changed in one
@@
 pseudo-instructions with their calling convention.
-
-\rb{As far as I can tell, while their notion of components and interfaces seems
-  standard, observable behavior appears to be limited to cotermination. The
-  previous paper makes explicit that (at least at that point) they are not
-  interested in externally observable side effects like console output or memory
-  access traces.}
 

### 2020-07-09 main.tex

--- 2020-07-09 main.tex
@@
 %Let $\callmap$ be a call map and $\codemap$ a function map.
-We say that a
-system satisfies {\em stack confidentiality}
-% with respect to $\callmap$ and $\codemap$
-if, for every initial machine-policy state
-pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
-$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
-in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
-\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
-For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
-\mach_e\), let \(\machT\) be the similarly prefixed trace such that
-\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
-\(\MPT_{\mathit{call}}\) and \(\machT\) with respect to \(\mathit{Cof}(\mach_c, n)\)
-and \(R = \ret{\mach_c}\):
-
-$$\forall \MPT_{\mathit{call}} ~ \machT. ~
-\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
-$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, n)} \mach_e' \wedge
-\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
-$$\mathit{conf}~\mathit{Cof}(\mach_c, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$
+We will say that a system satisfies {\em stack confidentiality} with respect to
+\(\callmap\) in a similar way to stack integrity. For every initial state
+\(\mach_0\), we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)}
+\hookrightarrow \MPT\). Then for any \(n\)-transition
+\(\mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~
+\mpstate{\mach_e}{\pol_e}\), we take the prefix trace until return:
+%
+\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\]
+
+We also take the corresponding trace from any variant \(\mach_e' 
+\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
+%
+\[\mach_e' \hookrightarrow \machT_{call} | \ret{\mach_c}\]
+
+If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
+\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
+stack confidentiality with regard to \(\callmap\).
 
 \apt{oy. still needs work to make this intelligible}
+\sna{Attempted. There's a tension between introducing all the variables
+     and keeping a good flow that I'm not sure I'm managing here though.}
 
@@
 \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
-is the first unmatched return in $\MPT$, then $\mach_r$ corresponds
-to a return state for $\mach_c$:
+is the first unmatched return such that for some $\pol_r$
+\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
+then $\mach_r$ corresponds to a return state for $\mach_c$:
+\[\ret{\mach_c}\ \mach_r\]
 

### 2020-07-09 main.tex
Revamp of 6

--- 2020-07-09 main.tex
@@
 %
-\[\mpstatename \in \MPS = \MACHS \times \POLS\]
+\[\MPS = \MACHS \times \POLS\]
 %
@@
 \(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
-\(\MPT_{\mathit{call}}\) and \(\machT\) with respect to \(\mathit{Cof}(\mach_c, n)\)
+\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
 and \(R = \ret{\mach_c}\):
@@
 Both the integrity and confidentiality components of local state
-encapsulation concern themselves only with the accessibility of data, and
-may still hold on a wide range of programs that violate expected control
+encapsulation concern themselves with the accessibility of data, and
+may hold on a wide range of programs that violate expected control
 flow. Say, if {\tt g} managed to jump in the middle of the
@@
 %
-Informal definitions of stack safety usually \apt{evidence?} aim to prevent such unwanted
+Stack safety is traditionally associated with preventing such unwanted
 flows: every jump between functions should either be a call or a
@@
 security properties that rely on the underlying knowledge of the
-program structure (call map $\callmap$, function map $\codemap$, return map $\retmap$, and entry
-map $\entmap$): {\em control separation} ensures the validity of transition
+program structure (call maps, function maps, return maps, and entry
+maps): {\em control separation} ensures the validity of transition
 sources between functions, {\em entry integrity} ensures the validity
@@
 
-\definition%{Control Separation}
-A system enjoys {\em control separation}
-% with respect to a function map
-%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
-if, for any initial state $\mpstatename$ and any adjacent pair
+\definition{Control Separation}
+
+A system enjoys {\em control separation} with regard to a function map
+\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
+if for any initial state \((\mach_0, \pol_0)\) and any adjacent pair
 of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
-trace from $\mpstatename$ whose current instruction
+trace from \(\mpstate{\mach_0}{\pol_0})\) whose current instruction
 belongs to different functions according to the function map,
-$\mach_1$ is either a valid call or a valid return:
+either $\mach_1$ is a valid call or a valid return:
 
@@
 
-For example, consider changing line 5 of the running example to a $\jalr$,
-$\PCname \leftarrow r_4$, with effect $\PCname \leftarrow 42$, causing {\tt main}
-to jump into the middle of {\tt h}. 
-This violates control separation because the instruction that causes a
+For example, consider the unsafe substitution:
+\[
+5 : ~ \jalr ~ r_4 ~ \rra ~ 0
+\]
+This violates control separation as the instruction that causes a
 function switch is not a valid call (or a return) according to the
@@
 
-\definition%{Entry Integrity}
-A system enjoys {\em entry integrity}
-%with respect to a call map
-%\(\callmap\) and entry map \(\entmap\)
-if, for any initial state $\mpstatename$
-and any n-transition \((\mach_c,\pol_c)\) and
+\definition{Entry Integrity}
+
+A system enjoys {\em entry integrity} with regard to a call map
+\(\callmap\) and entry map \(\entmap\) if for any initial state
+\((\mach_0, \pol_0)\) and any n-transition \((\mach_c,\pol_c)\) and
 \((\mach_e, \pol_e)\) in the induced trace from
-$\mpstatename$, $\mach_e$ is a member of the entry map:
+\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
 \[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]
 
-The changed program described just above also violates entry integrity,
-because the target (instruction 42) of the new $\jalr$ is not an established entry point
-according to the entry map.
+The same unsafe substitution violates entry integrity
+\[
+5 : ~ \jalr ~ r_4 ~ \rra ~ 0
+\]
+as its target (instruction 42) is not an established entry point for
+{\tt h} according to the entry map.
 
@@
 
-\newcommand*{\Unmatched}{\mathit{um}}
+\definition{Return Integrity}
+
+\newcommand*{\Unmatched}{\mathit{Unmatched}}
 
 Return integrity captures the expected behavior of marked returns with
-respect to their call points. Intuitively, it says that the marked return
-matching a call $\mach_c$ should satisfy $\ret{\mach_c}$.
-Observe that, given an MP-trace $\mpstate{\mach_c}{\pol}\MPT$, if $\mach_c$ is a call state then
-the matching return will be the first \emph{unmatched} return in $\MPT$. So
-we begin by defining an auxiliary ``unmatched'' relation between two MP-traces.
-
-\definition
-We write $\Unmatched~\MPT~\MPT'$ if $\MPT'$ is
-obtained from $\MPT$ by dropping elements until we reach the first unmatched return
-(according to the return map).  Formally, $\Unmatched$ is defined coinductively as follows:
+regard to their call points. Given a call map $\callmap$ and a return
+map $\retmap$, we begin by defining a coinductive relation
+$\Unmatched$ over two $\MPOTRACE$ $\MPT$ and $\MPT'$, if $\MPT'$ is a
+suffix of $\MPT$ whose head is the first unmatched return in $\MPT$.
+
+Formally:
 \begin{itemize}
@@
 
-\begin{minipage}{.3\textwidth}
   \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
@@
               {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
-\end{minipage}
-\hspace*{0.1\textwidth}
-\begin{minipage}{.3\textwidth}
   \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
@@
               {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}
-\end{minipage}
 
@@
               \hspace{.5cm} \neg \retmap(\mach(\PCname))
-              \hspace{.5cm}\Unmatched ~ \MPT ~ \MPT'\)}
+              \Unmatched ~ \MPT ~ \MPT'\)}
            {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}
 
-\item Finally, if a trace starts with a call, then we must discard the first
-  unmatched return from its tail to get to the real one:
+\item Finally, if a trace starts with the call, then we must discard the first
+  unmatched return to get to the real one:
 
@@
 \end{itemize}
-\apt{Could cut formal definition for space.}
-
-\definition%{Return Integrity}
-A system enjoys {\em return integrity}
-% with respect to a call map \(\callmap\) and return map \(\retmap\)
-if, for any initial state pair 
-$\mpstatename$ and any state pair \((\mach_c, \pol_c)\)
+A system enjoys {\em return integrity} with regard to a call map
+\(\callmap\) and return map \(\retmap\) if for any initial state pair
+\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
 corresponding to a call in its induced trace, if
-\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
-is the first unmatched return in $\MPT$, then $\mach_r$ corresponds
-to a return state for $\mach_c$:
-
+\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
+is the first unmatched return such that for some $\pol_r$
+\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
+then $\mach_r$ corresponds to a return state for $\mach_c$:
+\[\ret{\mach_c}\ \mach_r\]
+
+For example, if we added an instruction between 45 and 46:
 \[
-\exists p_r.\Unmatched ~ \MPT ~ (m_r,p_r)\MPT' \Rightarrow \ret{\mach_c}\ \mach_r.
+45.5: ~ \addi ~ \rra ~ r_0 ~ \negate 19
 \]
-\apt{I've tried to tidy this, but I don't think it corresponds to the Coq.
-  Shouldn't we apply $\Unmatched$ to the tail of $\MPT$?
-  And what about the no-unmatched-return-at-top-level clause?}
-
-For example, if we modified the running example by adding
-an instruction $ \rra \rightarrow \rra - 19$
-between 45 and 46, this would cause the return from {\tt h} to go to instruction 8 and
+This would cause the return from {\tt h} to go to instruction 8 and
 return to {\tt main} instead of {\tt g}. While the instruction that
@@
 separation), the state after the return would not correspond to an
-unmatched return for {\tt g}'s call to {\tt h}, violating return
+$\Unmatched$ return for {\tt g}'s call to {\tt h}, violating return
 integrity.
@@
 explicit entry points, and returns respect implicit entry points
-corresponding to their call.  A system that enjoys all three
-properties (with respect to a function map, call map, return map,
-and entry map) is said to enjoy
-{\em well-bracketed control flow} (with respect to those maps).
+corresponding to their call.  A system that enjoys them with regard to
+a code, a call, a return, and an entry map is said to enjoy
+{\em well-bracketed control flow} with regard to those maps.
 
@@
 \paragraph*{Strong Stack Safety}
-%
+
+\leo{Maybe lockstep?}
+
 This conservative policy does not only satisfy the stack safety above,
-but also a provably stronger one, which at each step of the
-computation performs integrity and confidentiality tests for a stack
-of running variants, which are kept in lockstep with the original
-machine.
+but also a provably (in Coq) stronger one that enforces local state
+encapsulation throughout every calls execution. To encode that
+property, we not only keep track of the execution of one machine
+(whose trace we're trying to characterize as safe), but of an entire
+stack of machine variations corresponding to nested calls of the
+original machine. For each step of the original machine, we take a
+step in every variant machine (thus, keeping them in lockstep
+execution) and enforce a single-step variant of integrity and
+confidentiality tests. In addition, every time a call is made we push
+a variation of the original machine state at the time of that call on
+top of the variant stack (just like the stack confidentiality property
+enforced trace confidentiality for every transition), and every time
+the machine returns from a call we pop the top variation (to account
+for stack confidentiality taking the prefix of the transition trace
+until its return).
+
+We start by defining the integrity and confidentiality tests, as well
+as the variant stack, before giving the full definition of strong
+stack safety.
 
@@
 
-\definition
-%
-  Our testing property extends the basic tests above to whole variant
-  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
-  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
-  is well-formed with respect to $\mach$\fi safisfies the
-  \emph{variant stack test} if for each variant element $\vm{C}$
-  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
-  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
-  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
-  and stack confidentiality test $\testConf{\mach}{\mach'}$
-  with respect to the original machine state. We write
-  this $\testVS{(\mach, \pol)}{\vs}$.
-
-\definition
-%  Let $\callmap$ a call map and $\codemap$ a function map.
-  Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
-  variant stack $\vs$, we say that an MP-trace induced from $(\mach,
-  \pol)$ satisfies the \emph{strong stack safety} at a given call
-  depth as follows.
-% Refactor repetitive notation before case split
-  In all cases that let $(\mach, \pol) \hookrightarrow_P
-  (\mach', \pol', \obs)$.
-  %and assume (unless the step is a return) that the
-  %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
-  %machines can step.
-
-%  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
-%  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
-%  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
-%  testing property} is satisfied at call depth $d$, and write
-%  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:
+We also need two auxilliary notations that relate to variant stacks.
+% 
+First, strong safety will check that the basic tests above hold for every
+machine in the variant stack.  Given a machine that steps $(\mach,
+\pol) \hookrightarrow_P (\mach', \pol', \obs)$, say that a variant
+stack $\vs$ safisfies the \emph{variant stack test} if for each
+variant element $\vm{C}$ (which is a variation of $\mach$ for $C$),
+$\vm{C}$ can take a step, denoted $\vm{C} \stepstoobs{\obs} \mach'$,
+while satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
+and stack confidentiality test $\testConf{\mach}{\mach'}$ with regard
+to the original machine state. We will write $\testVS{(\mach,
+  \pol)}{\vs}$ for this compound check.
+%
+Second, we will write $\vs \stepsto \vs'$ if every variant machine
+in $\vs$ steps to the corresponding machine in $\vs'$.
+%
+
+We are now ready to define strong stack safety coinductively.
+
+\definition Given a state $(\mach, \pol) \in \MPS$ and a variant stack
+$\vs$ in which every element $\vm{\contour}$ is a variation of $\mach$
+with respect to its contour, we say that an MP-trace induced from
+$(\mach, \pol)$ satisfies the \emph{trace strong stack safety}, and write
+$\testProp{}{\MPT}{\vs}$, if:
 
@@
 \item
-%
+  If the machine takes a \emph{non-call, non-return step}
+  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)\MPT,\]
+  then the variant stack test must be satisfied:
+  \[\testVS{(\mach, \pol)}{\vs},\]
+  and trace strong stack safety must be coinductively satisfied for
+  the remainder of the trace and the variant stack $vs'$ that
+  steps to ($\vs \stepsto \vs'$):
+  \[ \testProp{}{((\mach',\pol',\obs)\MPT)}{\vs'}\]
+
+\item
   If the machine takes a \emph{call step} and $\contour$ is the contour at
-    the call then,
-%
-    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
-    satisfied,
-%
-    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
-    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-    is satisfied with regard to an arbitrary $\contour$-variant
-    of $\mach$,
-%
-    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
-    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
-    $\contour$-variant of the machine,
-%
-    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
-    \pol') : \MPT)}{vs}$.
+  the call
+  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \callmap(\mach) = n,\]
+%
+  then the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
+  satisfied:
+  \[\testVS{(\mach, \pol)}{\vs},\]
+%
+%    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
+%    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
+%    is satisfied with regard to an arbitrary $\contour$-variant
+%    of $\mach$,
+%
+  and trace strong stack safety must be coinductively satisfied
+  for the remainder of the trace and for every variant stack
+  that is the result of adding a variation of $\mach'$ with
+  respect to $C$ at the top of the stack $\vs'$ that $\vs$ steps
+  to:
+  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\vm{\contour} : \vs')}\]
+
+\item
+%
+  If the machine $\mach$ takes a \emph{return step},
+  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \retmap(\mach),\]
+  %
+  the top of the variant stack also returns,
+  the tail of $\vs$ satisfies the stack safety test
+  \[\testVS{(\mach, \pol)}{\mathit{tail}(\vs)},\]
+  %
+  and trace strong stack safety must be coinductively satisfied for
+  the remainder of the trace and for the tail of the varant stack
+  that $\vs$ steps to:
+  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\mathit{tail}(\vs'))}\]
 
@@
     (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
-    $\testProp{d}{[(\mach, \pol)]}{\vs}$, for any $d$ up to the size of
-    $\vs$. (That is, a violation was detected $d$ levels into the stack.
-    Elements are created and pushed to the variant stack on calls and popped
-    from it on returns, so its size always reflects the depth of the program
-    stack.)
-
-  \item If the top variant of the stack (and only the top variant) returns,
-    %i.e., $\mathit{ret}(\head(\vs), \mach)$ holds,
-    the property is safisfied for
-    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
-    stack: $\testProp{|\vs|}{[(\mach, \pol)]}{\vs}$. \rb{Explain the
-    use of this vs the return case below.}
-
-  \item If
-%
-    (1) the machine takes a \emph{non-call, non-return step},
-%
-    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
-    satisfied, and
-%
-    (3) the stack safety test $\testProp{d}{((\mach', \pol') : \MPT)}{\vs'}$
-    is coinductively satisfied % at depth $d$
-    %% and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
-    %% \pol')$,
-%
-    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach', \pol') :
-    \MPT)}{vs}$.
-    %% \rb{omitting stack size condition}
-
-    %% the contour $\contour$ is updated based on the top of the current machine.
-    %% For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
-    %% the eager integrity test must hold for $\contour$ at the call,
-    %% the eager confidentiality test must hold for all variants at the call,
-    %% including the new variant that is added to the top of the stack,
-    %% and the property recurses on an arbitrary variant at the new contour
-    %% at depth $d$ with trace $\MPT$.
-
-  \item
-%
-    If (1) the machine takes a \emph{return step}, so the top of the variant
-    stack (and only the top variant) returns,
-%
-    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
-    satisfied,
-%
-    (3) the stack safety test $\testProp{d}{((\mach', \pol') :
-    \MPT)}{\mathit{tail}(\vs)}$ is coinductively satisfied
-%
-    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
-    \pol') : \MPT)}{vs}$.
-
-  \item If the machine takes a \emph{simultaneous return-call step}, that is,
-    the instruction to which execution returns immediately after completing a
-    call is itself a call instruction, the \emph{call step} and \emph{return
-    step} cases are combined.
+    $\testProp{}{[(\mach, \pol)]}{\vs}$.
+
+%  \item If the machine takes a \emph{simultaneous return-call step}, that is,
+%    the instruction to which execution returns immediately after completing a
+%    call is itself a call instruction, the \emph{call step} and \emph{return
+%    step} cases are combined.
 
@@
 %
-  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
-  satisfies the eager testing property at depth 1 with respect to the initial
-  variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
-%
-  \rb{I don't think the initial variant has been defined, but it's
-    straightforward.}
-
-By using property-based testing techniques, we can determine with a high degree
-of confidence that an implementation of the conservative micropolicy satisfies
-this eager testing property. Although we have not proved that the conservative
-policy guarantees the stack safety property, the testing property is provably
-stronger than the stack safety property \rb{theorem?} and, by this implication,
-we can conclude that the conservative policy also enforces stack safety.
-
-% The properties we give are indeed enforceable.
-
-\rb{What is the strongest property enforced by the conservative policy?}
-
-\paragraph*{End-to-end stack safety property}
-
-Integrity
-
-Confidentiality as a form of noninterference
-
-\paragraph*{Eager vs lazy testing}
+A system satisfies {\em strong stack safety} if for every initial machine $\mach$,
+the trace induced from $(\mach, \pinit{\mach})$ satisfies trace strong stack safety with
+respect to the empty stack.
+
+\theorem
+Strong stack safety implies stack safety.
+
+\proof
+By coinduction on trace strong stack safety.
+\leo{3000 lines Coq?}
+\leo{I can write a TON about this proof, but I'm not sure how much is interesting/how much space we have}.
+
+\paragraph*{Eager vs Lazy Stack Safety}
 
@@
 effect becomes manifest.
-\rb{How much detail to go into? Refer to examples if we have the space?}
-
+
+\leo{Not sure how much of the following could go after the lazy stack safety property}
 These \emph{lazy policies} admit more efficient implementations while being
@@
 
-      A system enjoys {\em observable stack integrity} with respect to
+      A system enjoys {\em observable stack integrity} with regard to
       \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
@@
       confidentiality}. A system enjoys {\em observable stack confidentiality}
-      with respect to \(\callmap\) and \(\codemap\) if for any initial state
+      with regard to \(\callmap\) and \(\codemap\) if for any initial state
       \(\mach_0\) and each transition with entry state
@@
     \definition{Observable Stack Safety}
-      A system enjoys {\em observable stack safety} with respect to a call map
+      A system enjoys {\em observable stack safety} with regard to a call map
       and a function map if it enjoys both observable stack integrity and
-      observable stack confidentiality with respect to them.
+      observable stack confidentiality with regard to them.
 

### 2020-07-09 main.tex
light pass over 7; releasing lock

--- 2020-07-09 main.tex
@@
 
-{\em Stack integrity} is a property of a system with regard to a call map
+{\em Stack integrity} is a property of a system with respect to a call map
 \(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
@@
 \(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
-system as a whole enjoys stack integrity with regard to \(\callmap\).
+system as a whole enjoys stack integrity with respect to \(\callmap\).
 
@@
 {\em Trace confidentiality} will be a property of two traces, \(\MPT\) and
-\(\machT\), with regard to a return predicate \(R \subset \MACHS\) that
+\(\machT\), with respect to a return predicate \(R \subset \MACHS\) that
 reflects the states that are considered to have returned. We abstract over
@@
 \ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
-stack confidentiality with regard to \(\callmap\).
+stack confidentiality with respect to \(\callmap\).
 
@@
     satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-    is satisfied with regard to an arbitrary $\contour$-variant
+    is satisfied with resepct to an arbitrary $\contour$-variant
     of $\mach$,
@@
 
+\apt{Need some intro text here.}
+
+\subsection{Observable Integrity}
+
     \paragraph{Rolling Back to Idealized State}
-      Observable integrity will weaken eager integrity by allowing a returned
+      Observable integrity weakens eager integrity by allowing a returned
       state to differ from its call state at high integrity components,
-      provided the observable behavior of the induced trace from the return
-      matches that of an idealized state whose high integrity components do
-      match the call state.
+      provided that the differences do not change the observable behavior
+      of the rest of the program.  
+%      provided the observable behavior of the induced trace from the return
+%      matches that of an idealized state whose high integrity components do
+%      match the call state.
 
@@
       \begin{cases}
-        \mach_c(\component) & \contour(\component) = (\HIGHINT,\_) \\
-        \mach_r(\component) & \contour(\component) = (\LOWINT,\_) \\
+        \mach_c(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
+        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
       \end{cases}\]
@@
 
+      \apt{Much confusion here and below about whether $\mathit{roll}$ applies to call state
+        or entry state just after.}
+
       Now we can define a version of trace integrity that compares the trace
@@
 
-    \definition{Observable Trace Integrity}
+    \definition%{Observable Trace Integrity}
 
@@
       some contour \(\contour\) if its final state (if any) has observable
-      behavior that prefixes that of its idealized version, constructed by
+      behavior \apt{states don't have behavior} that prefixes that of its idealized version, constructed by
       \(\mathit{roll}\):
@@
 
-    For out example, we return to the violation of trace integrity in
+    For our example, we return to the violation of trace integrity in
     \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
@@
     before printing, observable trace integrity would hold.
-
-    \definition{Observable Stack Integrity}
-
-      A system enjoys {\em observable stack integrity} with respect to
-      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
+    \apt{Better to give an example in which {\tt h} overwrites something that
+      is not needed to compute an output, so OTI does not fail.
+      E.g., {\tt h} could overwrite {\tt w} with nonsense before returning -- but
+      this doesn't matter, because the return value will replace the nonsense
+      before this figures in the final print. 
+      Also could be good to give an example in which we 
+      overwrite {\tt x} instead of {\tt z} to show that we still fail OTI -- in other words,
+      the entire remainder of the program matters, not just the caller.
+    }
+
+    \definition%{Observable Stack Integrity}
+
+    A system enjoys {\em observable stack integrity}
+    % with respect to    \(\callmap\) and \(\codemap\)
+    if, for any initial state \(\mach_0\) and
       each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
@@
       then \(\text{\sc ObsInt}_\contour ~ \MPT\).
+\apt{more words please}
+
+\subsection{Observable Confidentiality}
 
@@
       end condition over-approximates the extension of the behavioral condition
-      to future execution.
+      to future execution.\apt{That's quite mysterious.}
 
@@
       This requires restoring varied components to their original values,
-      since after return they will once again be accessible legitimately.
+      since after return they will once again be legitimately accessible.
 
@@
 
-    \definition{Observable Trace Confidentiality}
-
-      A pair of traces \(\MPT\) and \(\machT\) starting from
+    \definition{%Observable Trace Confidentiality}
+
+      Let \(\MPT\) and \(\machT\) be a pair of traces starting from
       \(\mpstate{\mach_e}{\pol_e}\) and \(\mach_e'\), respectively, with
-      \(\mach_e \approx_\contour \mach_e'\), enjoy observable trace
-      confidentiality with respect to \(\contour\) if two conditions hold:
-
-      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
-      \pi_o(\machT)\).
-
-      If \(\MPT\) is finite and ends with the state
-      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
-      \(\mach_r' = \last(\machT)\). Then the induced traces
-      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
-      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
-      \hookrightarrow \machT'\) are observationally prefixed,
-      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).
-
-      We write this \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).
+      \(\mach_e \approx_\contour \mach_e'\). \apt{Should some of this move into the defining clauses below?} Then
+      $\MPT$ and $\machT$ enjoy \emph{observable trace
+        confidentiality} with respect to \(\contour\), written
+            \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\), if:
+
+      \begin{itemize}
+      \item
+      the two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
+      \pi_o(\machT)\); and
+
+      \item
+      if \(\MPT\) is finite and ends with the state
+      \(\mpostate{\mach_r}{\pol_r}{\_}\), then \(\machT\) is finite, and
+      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\), where
+      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\),
+      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \hookrightarrow \machT'\),
+      and \(\mach_r' = \last(\machT)\). 
+      \end{itemize}
 
@@
       In our example, \cref{fig:conflex} shows how execution continues from
-      \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
-      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109
+      \(\mach_r'\) and the restored
+      \[\mach_r'' = \mathit{restore}(\contour,
+      \mach_e,\mach_e',\mach_r,\mach_r').
+      \]
+      Addresses 108 and 109
       are not restored; every address that was varied and not changed in one
@@
 
-    \definition{Observable Stack Confidentiality}
-
-      From observable trace confidentiality we construct {\em observable stack
-      confidentiality}. A system enjoys {\em observable stack confidentiality}
-      with respect to \(\callmap\) and \(\codemap\) if for any initial state
+    \definition%{Observable Stack Confidentiality}
+
+      A system enjoys {\em observable stack confidentiality}
+      %with respect to \(\callmap\) and \(\codemap\)
+      if, for any initial state
       \(\mach_0\) and each transition with entry state
@@
 
-    \definition{Observable Stack Safety}
-      A system enjoys {\em observable stack safety} with respect to a call map
-      and a function map if it enjoys both observable stack integrity and
+    \definition%{Observable Stack Safety}
+    A system enjoys {\em observable stack safety}
+    % with respect to a call map       and a function map
+    if it enjoys both observable stack integrity and
       observable stack confidentiality with respect to them.
+
+\apt{Words needed to connect these ideas back to the lazy policies which motivated them in the first place!}
 

### 2020-07-09 main.tex

--- 2020-07-09 main.tex
@@
 
+%% <<<<<<< HEAD
+%% \definition
+%% %  Let $\callmap$ a call map and $\codemap$ a function map.
+%%   Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
+%%   variant stack $\vs$, we say that an MP-trace induced from $(\mach,
+%%   \pol)$ satisfies the \emph{strong stack safety} at a given call
+%%   depth as follows.
+%% % Refactor repetitive notation before case split
+%%   In all cases that let $(\mach, \pol) \hookrightarrow_P
+%%   (\mach', \pol', \obs)$.
+%%   %and assume (unless the step is a return) that the
+%%   %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
+%%   %machines can step.
+
+%% %  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
+%% %  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
+%% %  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
+%% %  testing property} is satisfied at call depth $d$, and write
+%% %  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:
+
+%%   \begin{itemize}
+
+%%   \item
+%% %
+%%     If the machine takes a \emph{call step} and $\contour$ is the contour at
+%%     the call then,
+%% %
+%%     (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
+%%     satisfied,
+%% %
+%%     (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
+%%     satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
+%%     is satisfied with resepct to an arbitrary $\contour$-variant
+%%     of $\mach$,
+%% =======
 We also need two auxilliary notations that relate to variant stacks.
@@
   \pol)}{\vs}$ for this compound check.
+%% >>>>>>> 1886f7a7bf858ce0c9518e407d9481518a4f028d
 %

### 2020-07-09 main.tex
Overhaul of section 7, discussion of lazy policies/properties, a few example fixes

@@
   \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
-      {Set {\tt w} to 18}
+      {Set {\tt w} to return value}
   \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
@@
 \[\begin{split}
-  \propdef{TraceIntegrity} ~ \contour ~ \MPT \triangleq &
+  \propdef{TraceInt} ~ \contour ~ \MPT \triangleq &
     \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
@@
 \MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
-\(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
+\(\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
 system as a whole enjoys stack integrity with respect to \(\callmap\).
@@
 \[\begin{split}
-  \propdef{TraceConfidentiality} ~ R ~ \MPT ~ \machT \triangleq
+  \propdef{TraceConf} ~ R ~ \MPT ~ \machT \triangleq
   & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
@@
 
-If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
+If, for all such pairs of traces, \(\propdef{TraceConf} ~
 \ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
@@
 
-\section{Lazy Stack Safety}
+\section{Stack Safety Properties for Lazy Enforcement}
 \label{sec:lazy}
 
-    \rb{Suggestion: name this ``observable \emph{trace? something else?}
-      integrity'' to avoid confusion with observable stack integrity (in fact,
-      it is called exactly this when discussing the example below). The rollback
-      may also be a definition of its own to match the granularity of
-      definitions in previous sections.}
+  A lazy policy permits some violations of integrity or confidentiality,
+  justified because a truly harmful violation will be caught later.
+  In a typical lazy version of the tag policy, writes are always permitted,
+  but tagged to identify the context in which they were written. But reads
+  check that the value read was written by the current context and fail-stop
+  if it wasn't. So, if a callee makes an illegal write to its caller, when the
+  caller tries to read it the policy will catch the error.
+
+  The natural question is, how do we characterize the protections provided
+  by these policies once stack safety is broken? How does a harmful violation
+  that must be caught later differ from a harmless one? The properties in this
+  section use the observation model to draw this distinction.
 
@@
       provided that the differences do not change the observable behavior
-      of the rest of the program.  
-%      provided the observable behavior of the induced trace from the return
-%      matches that of an idealized state whose high integrity components do
-%      match the call state.
+      of the rest of the program. It is not immediately clear what constitutes
+      changing behavior; we construct an idealized return state that contains
+      only the changes that the caller was allowed to make, and use the trace
+      from that state as a baseline against which the real trace will be
+      compared.
+
+      \definition
 
       Given a contour, a call state, and a returned state, we define a
-      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
+      {\em rollback} function \(\mathit{roll} : \CONTOURS \times \MACHS \times
       \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
@@
       contour, and matches the return state on low integrity components:
-      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
+      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
       \begin{cases}
-        \mach_c(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
+        \mach_e(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
         \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
@@
 
-      \apt{Much confusion here and below about whether $\mathit{roll}$ applies to call state
-        or entry state just after.}
-
       Now we can define a version of trace integrity that compares the trace
@@
 
-      An MP-Trace enjoys {\em observable trace integrity} with respect to
-      some contour \(\contour\) if its final state (if any) has observable
-      behavior \apt{states don't have behavior} that prefixes that of its idealized version, constructed by
-      \(\mathit{roll}\):
+      A trace enjoys {\em observable trace integrity} with respect to
+      some contour \(\contour\) if the trace induced by its final state
+      (if any) has observable behavior that prefixes that of its idealized
+      version, constructed by \(\mathit{roll}\):
 
       \[\begin{split}
-        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
+        \propdef{ObsTraceInt} ~ \contour ~ \MPT \triangleq
         & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \\
@@
 
-    However, if {\tt g} had an additional instruction to overwrite {\tt z}
-    before printing, observable trace integrity would hold.
-    \apt{Better to give an example in which {\tt h} overwrites something that
-      is not needed to compute an output, so OTI does not fail.
-      E.g., {\tt h} could overwrite {\tt w} with nonsense before returning -- but
-      this doesn't matter, because the return value will replace the nonsense
-      before this figures in the final print. 
-      Also could be good to give an example in which we 
-      overwrite {\tt x} instead of {\tt z} to show that we still fail OTI -- in other words,
-      the entire remainder of the program matters, not just the caller.
-    }
+    However, consider instead if we insert after instruction 44 a new
+    instruction \(45: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
+    {\tt g}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
+    rolled-back state \(\mach_r'\) maps it to 0. But in the trace from either
+    state, instructions 27 and 28 load the return value (18) and store it to
+    105, eliminating the only difference. The remainder of the observation
+    traces must be identical, so this trace does obey observable trace
+    integrity.
 
@@
 
-    A system enjoys {\em observable stack integrity}
-    % with respect to    \(\callmap\) and \(\codemap\)
-    if, for any initial state \(\mach_0\) and
-      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
-      and contour \(\contour\) in the induced trace from
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
-      then \(\text{\sc ObsInt}_\contour ~ \MPT\).
-\apt{more words please}
+      A system enjoys {\em observable stack integrity}, like stack integrity,
+      if, for any initial state and each \(n\)-transition from
+      \(\mpstate{\mach_c}{\pol_c}\) to \(\mpstate{\mach_e}{\pol_e}\) in its
+      induced trace, we take the prefix of the call
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\)
+      and observable trace integrity holds on \(\MPT_{call}\) with
+      contour \(\mathit{Cof}(\mach_c,n)\).
 
@@
 
-    \paragraph{Restoring Varied States}
-
-      Trace confidentiality is a property of two traces whose initial machine
-      states are variations, with two sub-properties. First, the behavioral
-      condition -- that their observable behaviors prior to the end of the
-      trace are equivalent -- reflects the intuition that secret values should
-      not influence visible behavior. Second, we have the end condition that
-      their final states (if any) are equal up to the original variation. The
-      end condition over-approximates the extension of the behavioral condition
-      to future execution.\apt{That's quite mysterious.}
-
-      We introduce {\em observable trace confidentiality}, which extends the
-      behavioral condition to apply to execution from the final state on.
-      This requires restoring varied components to their original values,
-      since after return they will once again be legitimately accessible.
-
-      To return to the example in \cref{fig:confex}, the varied final state
-      \(\mach_r'\) might do anything depending on the values of the variation,
-      so we cannot immediately compare traces from \(\mach_r\) and
-      \(\mach_r'\). We must first restore the varied values to their original
-      state, while preserving differences that have occurred during
-      execution.
-      \rb{To me, this reads like a prologue to the definitions properly
-        speaking---maybe the definition header above should be a paragraph and
-        a definition header appear next? I would also split it into two
-        definitions.}
-      \sna{Definitely better as a prologue. Where would you split it?}
-
-%      In \cref{fig:lazyex1}, the final states of the original and variant
-%      differ, because a written value (in black or white) depended on the value
-%      of a high confidentiality location. So the resulting trace fails trace
-%      confidentiality. For observable trace confidentiality, we ask whether the
-%      differing values at address 42 cause any observable distinctions in
-%      future behavior. To do so, we must undo the variation of formerly high
-%      confidentiality locations, so that those in yellow match those in red,
-%      while preserving the updated values where data may have been leaked.
+      Two of the cases of trace confidentiality, \(\mathit{InfCase}\) and
+      \(\mathit{FailCase}\), are purely properties of the observations produced
+      by the traces they are properties of. If the traces are infinite, they
+      should be similar; if one fail-stops, it should be a prefix. Only
+      the \(\mathit{ReturnCase}\) checks machine states, because in that case
+      the traces could continue, but have been cut off by a return. The key
+      to making confidentiality lazy is to consider the traces after the
+      return, considering only whether they are observationally similar
+      or prefixes.
+
+      This requires restoring varied components to their original values
+      in the trace from the variant state, since after return they will
+      once again be legitimately accessible. To return to the example in
+      \cref{fig:confex}, the varied final state \(\mach_r'\) still has
+      -5 as a return address for {\tt g}, so the trace after that return will
+      behave oddly in a way that doesn't reflect a confidentiality violation.
+      So we cannot immediately compare traces from \(\mach_r\) and
+      \(\mach_r'\). We must restore the varied values while preserving
+      differences that have occurred during execution.
+
+      \definition
 
@@
 
-      Now we can describe a trace property that extends trace confidentiality.
+      Now we can describe a trace property akin to trace confidentiality, but
+      continuing execution after the end of the callee trace.
 
@@
 
-      Let \(\MPT\) and \(\machT\) be a pair of traces starting from
-      \(\mpstate{\mach_e}{\pol_e}\) and \(\mach_e'\), respectively, with
-      \(\mach_e \approx_\contour \mach_e'\). \apt{Should some of this move into the defining clauses below?} Then
-      $\MPT$ and $\machT$ enjoy \emph{observable trace
-        confidentiality} with respect to \(\contour\), written
-            \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\), if:
-
-      \begin{itemize}
-      \item
-      the two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
-      \pi_o(\machT)\); and
-
-      \item
-      if \(\MPT\) is finite and ends with the state
-      \(\mpostate{\mach_r}{\pol_r}{\_}\), then \(\machT\) is finite, and
-      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\), where
-      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\),
-      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \hookrightarrow \machT'\),
-      and \(\mach_r' = \last(\machT)\). 
-      \end{itemize}
+      {\em Observable trace confidentiality} is the lazy analog to trace
+      confidentiality, and reuses the cases in which the traces don't return.
+      For a contout \(\contour\), return predicate \(R\), and a pair of traces
+      \(\MPT\) and \(\machT\), we consider the return case, in which
+      \(\last(\MPT)\) is some \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and
+      \(R ~ \mach_r\). Then \(\last(\machT)\) is some \((\mach_r',\obs_r')\)
+      and \(R ~ \mach_r'\). As above, the observations of \(\MPT\) are similar
+      to those of \(\machT\). In place of the requirement that \(\mach_r\)
+      and \(\mach_r'\) agree on components, we have the following requirement.
+
+      Finally, let \(\mach_e\) be the machine state of \(\head(\MPT)\),
+      and \(\mach_e'\) be that of \(\head(\machT)\). We restore \(\mach_r'\)
+      based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
+      \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')\).
+      The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
+      observational prefix of the induced trace from \(\mach_r'\),
+      reflecting the possibility of a fail-stop after return.
+
+      \[\begin{split}
+        \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \triangleq
+        & \head(\MPT) = \mpostate{\mach_e}{\pol_e}{\obs_e} \Rightarrow \\
+        & \head(\machT) = (\mach_e',\obs_e') \Rightarrow \\
+        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+        & \last(\machT) = (\mach_r',\obs_r') \Rightarrow \\
+        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
+        & \mach_r'' \hookrightarrow \machT' \Rightarrow \\
+        & \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
+        & \pi_o(\MPT') \lesssim \pi_o(\machT') \\
+      \end{split}\]
+
+      We combine this case with the other two cases from ordinary trace
+      confidentiality:
+
+      \[\begin{split}
+        \propdef{ObsTraceConf} ~ \contour ~ R ~ \MPT ~ \machT \triangleq
+        & \mathit{ObsReturnCase} ~ \contour ~ R ~ \MPT ~ \machT \land \\
+        & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
+          \MPT ~ \machT \\
+      \end{split}\]
 
@@
 
+      On the other hand, suppose that we inserted an instruction at 43:
+      \(r_6 \leftarrow [\rsp-8]\). \(r_6\) would hold 42 in \(\mach_r\)
+      and -5 in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
+      never used otherwise, the traces of {\tt h} will enjoy observable trace
+      confidentiality.
+
     \definition%{Observable Stack Confidentiality}
 
-      A system enjoys {\em observable stack confidentiality}
-      %with respect to \(\callmap\) and \(\codemap\)
-      if, for any initial state
-      \(\mach_0\) and each transition with entry state
-      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
-      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
-      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).
+      {\em Observable stack confidentiality} is structured like stack
+      confidentiality. Let \(\callmap\) be a callmap; for each initial
+      state and \(n\)-transition consisting of \(\mpstate{\mach_c}{\pol_c}\)
+      and \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
+      \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
+      and \(\mach_e' \hookrightarrow \machT | \ret{\mach_c}\). A system
+      enjoys observable stack confidentiality if, for all such
+      \(\MPT\) and \(\machT\), \(\propdef{ObsTraceConf} ~
+      \mathit{Cof}(\mach_c,n) ~ \ret{\mach_c} ~ \MPT ~ \machT\).
 

### 2020-07-09 main.tex
Overall pass, some reworking of policy discussions

@@
   local state encapsulation} is that callers should be protected from
-callees: a callee should be not be able to access or overwrite private data
+callees: a callee should not be able to access or overwrite private data
 in the caller's stack frame. We formalize this intuition as a security
@@
 properties \leo{in terms of security properties or ``using security
-  language''?}  requires an explicit attacker model: an clear
+  language''?}  requires an explicit attacker model: a clear
 understanding of who is being protected and from what. Most existing
@@
 
-This model is flexible enough to capture any kind of mechanism used for dynamic
-enforcement of properties of interest, including the following representative
-families of policies.
+Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
+policy state to the machine state and a partial policy step function that
+refines the machine step function by allowing or disallowing a step to be taken
+based on its auxiliary state. This model will be made more precise in
+\cref{sec:prelim}. It is flexible enough to capture any kind of mechanism used
+for dynamic enforcement of properties of interest, including the following
+representative families of policies.
 
@@
 The simplest possible policy is the neutral policy that carries no information
-and does nothing. Its set of policy states $\POLS$ is the unit type, and its
-policy step function is the constant function that produces the unit, $\lambda
-\_ . \top$, thereby allowing all calls the machine step function. Its
-initialization function is defined on all states. This corresponds to the
-lifting of our machine model to the policy setting. Machines that fulfill a
-property by construction will also fulfill that property with the null
-policy.
+and does nothing. Its set of policy states is the unit type, and its policy step
+function is the constant unit function, thereby allowing all calls to the machine
+step function. Its initialization function is defined on all states. This
+corresponds to the lifting of our machine model to the policy setting. Machines
+that fulfill a property by construction will also fulfill that property with the
+null policy.
 
@@
 %
-Some policies are implemented in software, by modifying code in the compiler
-to (e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
+Some policies are implemented in software, by modifying code in the compiler to
+(e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
 canaries \citep{Cowan+98}. These approaches are represented as policies that
-accept only initial machine states that are well-formed for the policy (due
-to being produced by the compiler.) The set of policy states \(\POLS\) is again
-the unit type and the step function constant. Property enforcement comes from
-the fact that only ill-formed states can induce a trace that violates the
-property.
+accept only initial machine states that are well-formed for the policy (due to
+being produced by the compiler.) The set of policy states is again the unit type
+and the step function constant. Property enforcement comes from the fact that
+only ill-formed states can induce a trace that violates the property.
 
@@
 Other policies are implemented in hardware or built on such hardware mechanisms.
-Capability machines like CHERI \rb{Woodruff+14} extend conventional
+Capability machines like CHERI \citep{Woodruff+14} extend conventional
 architectures to support efficient and fine-grained control over memory
@@
 those are tightly coupled to the hardware design. However, separating the
-enforcement aspects of a concrete policy (such as StkTokens, where \(\POLS\)
+enforcement aspects of a concrete policy (such as StkTokens, where policy state
 would encompass details like the abstract call stack, the capability sealing
@@
 hardware-enforced security monitor. We will refer back to the running
-example of the previous section for concreteness, and then,
+example of \cref{sec:running-example} for concreteness, and then,
 in~\cref{sec:enforcement}, extend this model with a dynamic policy
@@
 
-In our model, the machine step function is total -- a program does not halt
+In our model, the machine step function is total---a program does not halt
 on its own, but may instead enter a silent terminal loop. Runs are only cut
@@
 halting outright.\apt{Is this our current thinking?}
+\rb{And if so, should the running example loop at the end, or do we assume
+that is exactly what happens after main is done?}
 
@@
 
-In this section we formalize our definition of stack safety.We
+In this section we formalize our definition of stack safety. We
 begin with our model of \leo{compiler-provided?} knowledge about the
@@
 %
-\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\]
+\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} | \ret{\mach_c}\]
 
@@
 %
-\[\mach_e' \hookrightarrow \machT_{call} | \ret{\mach_c}\]
+\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]
 
@@
 
+\rb{Wondering if we should use sub-sections here, too.}
+
 In this section we are going to see how an existing enforcement mechanism,
@@
 The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
-on programmable, tag-based reference monitor that runs alongside the
+on a programmable, tag-based reference monitor that runs alongside the
 program.  To control this monitor, they use a programming model that
@@
 
-In such a system, all memory addresses and registers (including the
-program counter) are enriched with an abstract metadata tag, which can
-point to the representation of an arbitrary data structure. A
-micropolicy is thus defined by a collection of rules to be checked as
-each instruction executes and together define the transfer function of
-the security policy. Each rule inspects the tags associated to the
-inputs of an instruction (PC, registers, memory) as well as the
-instruction opcode itself, and produces one of two outcomes: either
-the instruction is allowed to execute (and generates and assigns tags
-for the new PC and the result of the operation), or the machine
-fail-stops with a policy violation.
-
-Efficient execution of these micropolicies relies on hardware
-implementations, such as the PUMP architecture~\citep{pump_oakland15},
-which include a monitor rule cache that maps tag inputs to outputs and
-a mechanism to trap misses to software handlers and thereby compute
-the transfer function of the software-defined policy.  Previous work
-has shown that a wide range of micropolicies can be defined and
-enforced efficiently \rb{cite}.
+In such a system, all memory addresses and registers (including the program
+counter) are enriched with an abstract metadata tag, which can represent an
+arbitrary data structure. A micropolicy is defined as a set of tags, and a
+collection of software-defined rules that represent the transfer function of the
+policy. Each rule inspects the tags associated to the inputs of an instruction
+(PC, registers, memory) as well as the instruction opcode itself, and produces
+one of two outcomes: either the instruction is allowed to execute (and generates
+and assigns tags for the new PC and the result of the operation), or the machine
+fail-stops with a policy violation. During program execution, micropolicy rules
+are checked before each machine step.
+
+Efficient execution of these micropolicies relies on hardware implementations,
+such as the PUMP architecture~\citep{pump_oakland15}. This hardware incorporates
+a monitor rule cache to allow quick retrieval of transfer function outputs for
+mapped inputs. Cache misses are detected and trapped to software handlers that
+use the software rules of the policy to compute the relevant parts of the
+transfer function. Previous work has shown that a wide range of micropolicies
+can be defined and enforced efficiently in these machines
+\citep{pump_oakland2015}.
 %
@@
 with ownership information associated to each stack frame (and to each
-separate object inside that frames, a more finely grained access
+separate object inside that frame, a more finely grained access
 control that we do not need to consider here), and registers holding
@@
 ``initial frame'' should be tagged with its proper depth}, and the {\PCname}
-carries in its tag at all times the depth of the current activation, starting
+carries in its tag the depth of the current activation at all times, starting
 with a $\tagPCDepth{0}$ tag. The current activation depth needs to change
@@
   {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
-  the locations of the new frame is marked $\tagStackDepth{n + 1}$.
+  the locations of the newly allocated frame are marked $\tagStackDepth{n + 1}$.
 
@@
 
-This conservative policy does not only satisfy the stack safety above,
+This conservative policy satisfies not only the stack safety above,
 but also a provably (in Coq) stronger one that enforces local state
-encapsulation throughout every calls execution. To encode that
+encapsulation throughout every call's execution. To encode that
 property, we not only keep track of the execution of one machine
-(whose trace we're trying to characterize as safe), but of an entire
+(whose trace we are trying to characterize as safe), but of an entire
 stack of machine variations corresponding to nested calls of the
@@
 machine in the variant stack.  Given a machine that steps $(\mach,
-\pol) \hookrightarrow_P (\mach', \pol', \obs)$, say that a variant
+\pol) \stepstopol{\obs} (\mach', \pol')$, say that a variant
 stack $\vs$ safisfies the \emph{variant stack test} if for each
@@
   If the machine takes a \emph{call step} and $\contour$ is the contour at
-  the call
-  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \callmap(\mach) = n,\]
+  the call \rb{Corrected the step function here, or should it be
+  $\hookrightarrow$ and give the rest of the induced trace?}
+  \[(\mach, \pol) \stepstopol{\obs} (\mach', \pol') \wedge \callmap(\mach) = n,\]
 %
@@
   If the machine $\mach$ takes a \emph{return step},
-  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \retmap(\mach),\]
+  \[(\mach, \pol) \stepstopol{\obs} (\mach', \pol') \wedge \retmap(\mach),\]
   %

### 2020-07-09 main.tex
Work on intro

--- 2020-07-09 main.tex
@@
 \begin{abstract}
-Text of abstract \ldots.
+\bcp{Hey, we should have an abstract \ldots}
 \end{abstract}
@@
 
-To phrase these two aspects of stack safety in terms of security
-properties \leo{in terms of security properties or ``using security
-  language''?}  requires an explicit attacker model: an clear
-understanding of who is being protected and from what. Most existing
-stack protection mechanisms deal with scenarios where a callee gets
-influenced by external inputs and ends up executing potentially
-malicious code. To avoid the details of exactly how this happens and
-keep the presentation clear, we pick an even stronger model: when
-stating the protections that a function can rely on when it makes a
-call, we assume that the callee can execute arbitrary instructions
+To phrase these two aspects of stack safety in terms of security properties
+\leo{in terms of security properties or ``using security language''?}
+requires an explicit attacker model: an clear understanding of who is being
+protected and from what. Most existing stack protection mechanisms deal with
+scenarios where a callee is influenced by external inputs and ends up
+corrupting the stack in some way. To avoid getting into the details of
+exactly how this happens in our formal properties, we use an even stronger
+model: when stating the protections that a function can rely on when it
+makes a call, we assume that the callee can execute arbitrary instructions
 until it returns.
 
-Adopting that model in our setting, however, poses another novel
-challege: not only does a caller need to be protected from its callee,
-but the callee itself needs to be protected from {\em its}
-callees. That is, the notion of ``attacker'' changes throughout the
-execution of a program!  We address this problem by using a similarly
-changing notion of accessibility for the different components of the
-machine: we adjust the integrity and confidentiality level of
-different stack locations and registers at every transition (call or
-return) between functions.  \bcp{that's way too terse!}\leo{Slightly
-  better, still needs a bit probably}
-%\apt{connection to ``attacker'' still vague.}  \rb{Mention
-%  to varying ``accessing power'' of the attacker?  The notion of
-%  attacker changes on certain actions that are not externally
-%  visible.} \bcp{Maybe the best way is to grab the bull by the horns:
-%  ``If we're going to phrase stack safety as a security property, we
-%  need to talk about an attacker model, and this is where the
-%  technicalities get really interesting...''}
+A key challenge in adapting the concepts of language-based security to this
+setting is that stack safety is a ``nested'' property: not only does a given
+caller need to be protected from its callee, but the callee also needs to be
+protected from {\em its} callees, and so on. That is, the notion of
+``attacker'' changes throughout the execution of the program.  We address
+this issue using a dynamically changing notion of {\em accessibility} for
+the different components of the machine state and adjusting the integrity
+and confidentiality levels of different stack locations and registers at
+every transition (call or return) between functions.
 
@@
 \section{Threat Model}
-\label{sec:threat}
 
@@
 
-\apt{Informal definition of policies as auxiliary state here.}
+\bcp{More explanation needed, plus examples.}
+A policy is a general model of an enforcement mechanism, consisting of a
+set of policy states \(\pol \in \POLS\), a policy step function
+\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
+\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
+\POLS\). The initializer establishes the initial policy state for a given
+machine state, but is partial to reflect that some policies are only defined
+on states that are well-formed in some way.
+
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
+%
+\[\MPS = \MACHS \times \POLS\]
+%
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
+               \polstep \pol_2\)}
+            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
+               \mpstate{\mach_2}{\pol_2}\)}
+
+\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
+in more detail that the intention is to capture any kind of dynamic
+enforcement mechanism and give as many names and citations as we can.}
 
@@
 enforcement mechanism based on
-\citet{DBLP:conf/sp/RoesslerD18}.\apt{needs a relook}
+\citet{DBLP:conf/sp/RoesslerD18}.
 
@@
 Our machine states are composed of {\em components} $k$, addressed either by
-words (memory addresses) or by register names ($r$) drawn from some set
+words (memory addresses) or by by register names ($r$) drawn from some set
 $\REGS$.
@@
 silent (written $\tau$) or a word of data that the machine
-communicates to the outside world:
+communicates to the outside world (e.g., by writing to a DMA
+address):
 %
@@
 
-Notice that we choose to model observations explicitly, rather than simply treating
-a subset of registers or memory as being observable.  This choice gives
-us the flexibility to treat programs as observationally equivalent even when
-their internal states may differ.
-%The choice of modelling observations is an interesting one. A first approach
-%would be to consider any change to the registers or memory as observable.
-%However, as some parts of the stack will be considered ``private''
-%this choice won't work. A relaxation of this to consider all changes
-%to ``non-private'' data as observable could work in principle, but
-%does not allow for an interesting weaker-but-more-efficient enforcement
-%mechanism. We will come back to this point in \Cref{sec:lazy}.
-
-In the running example of \Cref{sec:running-example}, 
-$\WORDS$ is the
-set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
-\cup \{\PCname,\SP\}$. Observations are the writes to the
-  special $\rout$ location, which can be thought of as a memory-mapped
-  output port: if during an execution step a word $w$ is
+The choice of observations is an interesting one. A first approach
+would be to consider any change to the registers or memory as observable.
+However, as some parts of the stack will be considered ``private''
+this choice won't work. A relaxation of this to consider all changes
+to ``non-private'' data as observable could work in principle, but
+does not allow for an interesting weaker-but-more-efficient enforcement
+mechanism. We will come back to this point in \Cref{sec:lazy}. On the
+other hand, we could use a particular location as an abstract output
+of our system (e.g., through DMA). This is the approach we chose for
+our running example to keep things simple. For that, $\WORDS$ is the
+set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
+\cup \{\PCname\}$, and $\SP = r_2$. Observations are the writes to the
+special $\rout$ location: if during an execution step a word $w$ is
 written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.
-
-\subsection{Policies}
-
-A policy is a general model of an enforcement mechanism, consisting of a
-set of policy states \(\pol \in \POLS\), a policy step function
-\((\mach, \pol) \polstep \pol' : \MACHS \times \POLS
-\rightharpoonup \POLS\), and an initializer \(\pinit : \MACHS \rightharpoonup
-\POLS\). The initializer establishes the initial policy state for a given
-machine state, but is partial to reflect that some policies are only defined
-on states that are well-formed in some way.\apt{Is this our current thinking?}
-
-We write $\MPS$ for the set of all pairs of machine states and policy
-states.
-%
-\[\mpstatename \in \MPS = \MACHS \times \POLS\]
-%
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
-               \polstep \pol_2\)}
-            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
-               \mpstate{\mach_2}{\pol_2}\)}
-
-\apt{TODO: relate back to \Cref{sec:threat} and forward to \Cref{sec:enforcement}.}
-%\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
-%in more detail that the intention is to capture any kind of dynamic
-%enforcement mechanism and give as many names and citations as we can.}
 
@@
 short by the policy step function. This means that software-only policies that
-can catch errors must be modeled as reaching an error state\apt{??}, not 
-halting outright.\apt{Is this our current thinking?}
+can catch errors must be modeled as reaching an error state, not outright
+halting.
 
@@
 
-In a similar fashion, we define \(\mpstatename \hookrightarrow \MPT\)
+In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
 to relate a machine-policy state pair to the MP-trace
-induced by the step function on $\MPS$. Unlike the $\MOTRACE$ above, this trace can
-be finite as the policy can cause a fault, in which case the final observation is
+induced by \(\stepstopol{}\). Unlike the $\MOTRACE$ above, this trace can
+be finite as the policy can cause a fault, in which case the final state is
 a \(\tau\):
+\rb{This $\hookrightarrow_P$ is often enough written just $\hookrightarrow$.
+Overload everywhere?}
 
 \begin{minipage}{.4\textwidth}
-\judgment{\(\neg \exists \pol_1 . \mpstate{\mach_0}{\pol_0} \polstep \pol_1\)}
-         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
+\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
+         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
 \end{minipage}
 \begin{minipage}{.55\textwidth}
-  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstopol{\obs}
+  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstoobs{\obs}
                  \mpstate{\mach_1}{\pol_1}\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow
+              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow_P \MPT\)}
+              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P
                  \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
@@
 \paragraph*{Until}
-The operation \(\PUT~(\Trace,f)\)
+The ``Until'' operation \(\PUT(\Trace,f)\)
 %\bcp{Wouldn't
@@
     \judgment{\(f\ \traceelem\)}
-             {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
+             {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
   \end{minipage}
@@
   \judgment{}
-           {\(\PUT~(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
+           {\(\PUT(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
 %
@@
   \begin{minipage}{.3\textwidth}
-  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT~(f,\Trace) = \Trace'\)}
-              {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
+  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
+              {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
   \end{minipage}
@@
 \judgmenttwo{\(\mach \hookrightarrow \machT'\)}
-            {\(\PUT~(f,\machT')
+            {\(\PUT(f,\machT')
               = \machT\)}
@@
 \judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
-            {\(\PUT~(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
+            {\(\PUT(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
          {\(\mpstatename \hookrightarrow \MPT | f\)}
@@
 non-silent observations of $\obsT_1$ is a prefix of those of
-$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
+$\obsT_2$; that is we operate up to deletion of \(\tau\) observations,
 coinductively:
@@
 similarity of observation traces could alternatively be defined coinductively,
-just like the prefix relation, but by dropping the asymmetric rules that allow
-for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}
-
-Note that an infinite silent trace is a
+just like the prefix relation, but by dropping the assymetric rules that allow
+for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$)
+
+In addition, perhaps surprisingly, an infinite silent trace is a
 prefix of (and equivalent to) any other trace. While this might seem
@@
 
+%Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
+%We then define {\em policy-sensitive observational equivalence}: a
+%trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
+%\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
+%and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
+%policy fault) and its observations are a prefix of those of M.
+%
+%% Fixed by Sean
+%%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
+%\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
+%            {\(\MPT \sim_P \machT\)}
+%
+%\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
+%            {\(\MPT \sim_P \machT\)}
+
 \section{Stack Safety, Formally}
@@
 
-In this section we formalize our definition of stack safety.We
+In this section we will describe stack safety as a formal property. We
 begin with our model of \leo{compiler-provided?} knowledge about the
@@
 Similarly, a {\em return map} identifies addresses that complete a
-return sequence. Formally, it is simply a set of words; each one
-is the address of the final jump corresponding to a function return:
+return sequence. Formally, it is simply a set of components; each one
+is the final jump corresponding to a function:
 %
@@
 to have correctly returned from a call. We consider the call to have
-returned when control returns to the instruction following
-the \jal with the stack restored. Formally, this is a predicate on
-states that says when the machine has just returned from a specified call state
-(where $\nextPC$ gives the address of the next instruction after a given address):
-\[%\begin{split}
-    \ret{\mach_c} \triangleq  \lambda \mach_r . \mach_r(\PCname) =
-    \nextPC(\mach_c(\PCname)) \land  \mach_r(\SP) = \mach_c(\SP)
-%\end{split}
-\]
+returned the first time control returns to the instruction following
+the \jal with the stack restored. Formally this is a relation
+between the states:
+
+\[\begin{split}
+    \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) =
+    \nextPC(\mach_c(\PCname)) \land \\ & \mach_r(\SP) = \mach_c(\SP)
+\end{split}\]
+
 There are two such return points in our running example, both of which
@@
 
-In the rest of this section we implictly parameterize all 
+In the rest of this section we will implictly parameterize our
 definitions by a function map $\codemap$, a call map $\callmap$, a return
@@
 
-When \(\mach_c\) is a call state (\(\callmap(\mach_c(\PCname)) = n\) for some $n$) and
-\(\mach_c \stepstoobs{\obs} \mach_e\) for some 
-$\mach_e$ and $\obs$, we can construct a contour based on the bounds of the
+When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) and
+\(\mach_c \stepstoobs{\obs} \mach_e\) for some $n$, $\obs$, and
+$\mach_e$, we can construct a contour based on the bounds of the
 callee's stack frame.
@@
 call. Private components are marked high confidentiality and high
-integrity: the callee should be able to neither read nor write them.
+integrity: the callee should be able to neither read, nor write them.
 %
@@
 %
-Code is always high integrity and low confidentiality.
-%
-Finally, registers and the frame of arguments and results are free
+Special care is needed for the code memory corresponding to the
+callee: they are marked low confidentiality if they belong to the
+callee and high confidentiality otherwise; regardless they are always
+high integrity.
+%
+Finally, registers and the frame of arguments to the callee are free
 for the callee to use, and are therefore tagged with low integrity and
@@
     (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
-                         \text{ and } \codemap(\component) = \funid \text{ for some }\funid\\
+                         \text{ and for some } \funid, \codemap(\component) = \funid \\
     (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
@@
 
-Going back to our running example, here is the state of the machine at
-the call from {\tt main} to {\tt g}, showing addresses annotated with their 
+Going back to our running example, recall the state of the machine at
+the call from {\tt main} to {\tt g}, annotated with their
 corresponding labels (not shown: the labels of the instructions
 corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
-$(\HIGHSEC, \HIGHINT)$)\apt{see previous note}:
+$(\HIGHSEC, \HIGHINT)$):
 \vspace*{0.2em}
@@
 
-\subsubsection{Stack Integrity}
+\paragraph*{Stack Integrity}
 
@@
 
-Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
-\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
-respect to \(\contour\) if its final state (if any) agrees with its
-initial one on all components marked \(\HIGHSEC\) in \(\contour\):
+Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
+$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
+respect to $\contour$ if its final state (if any) agrees with its
+initial one on all components marked $\HIGHSEC$ in $\contour$:
 %
 \[\begin{split}
-  \propdef{TraceIntegrity} ~ \contour ~ \MPT \triangleq &
+  \mathit{int} ~ \contour ~ \MPT \triangleq &
     \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
@@
 \end{split}\]
-\sna{Experimenting with font on the mathy version of the property.}
+%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
+%\sna{We could do it like this, with finiteness implicit in the existence of
+%     a final state.}+1
 
@@
 private data is never overwritten, even during the caller's execution
-(which leads to easier reasoning and more effective testing);
-and a weaker, lazy integrity property that only ensures the preservation
-of locations that subsequently affect the caller's observable actions 
-(which allows for a more efficient enforcement mechanism).
+(a variation that leads to easier reasoning and more effective testing);
+and a weaker, lazy integrity property that only ensures that the
+locations that a caller accesses are preserved (a variation that allows
+for a more efficient enforcement mechanism).
 
@@
 
+For an example see \cref{fig:intex}, representing the unsafe substitution
+\[
+  44: ~ \sw ~ \rsp ~ r_4 ~ -5
+\]
+This instruction is in {\tt h}, and instead of storing data to the return,
+it overwrites {\tt g}'s variable, {\tt z}. At the top of the figure high
+integrity memory is in red, while low integrity is in green; at the end of the
+trace address 103 should match the initial state, and does not. So trace
+integrity is violated.
+\rb{There are some color name macros, but none for integrity yet? Use instead
+of hardcoded color names.}
+
 Trace integrity characterizes the trace corresponding to a single
-callee.
-Going back to our running example, such a trace would be, for
+callee. Going back to our running example, such a trace would be, for
 instance, the entire trace of {\tt h} from instruction 40 to
-instruction 47. 
-As an example of an integrity violation affecting this trace, \cref{fig:intex} shows what happens
-if we change line 44 of the example program to $[\rsp - 5] \leftarrow r_4$.
-Instead of storing data to the return value slot,
-the changed version overwrites {\tt g}'s local variable {\tt z}.
-The figure shows the modified trace, with snapshots of the initial
-and final states colored to show the contour; high 
-integrity memory is in \high, while low integrity is in \low.
-At the end of the trace high integrity address 103 should match the initial
-state, and does not. So trace integrity is violated.
-\rb{There are some color name macros, but none for integrity yet? Use instead
-  of hardcoded color names.}\apt{I don't understand what's going on with the colors.
-  I thought we agreed to use the same coloring scheme throughout?}
+instruction 47. This trace is repeated in \Cref{fig:intex}, with the initial
+and final states colored to show the contour; high confidentiality
+components in magenta, low in teal. At the end of the trace addresses 103
+and 108 have both changed; since 103 was high confidentiality, the trace
+does not enjoy trace integrity.
 
@@
 execution trace. We do this by requiring that trace integrity holds
-for every subtrace delimited by a call and its corresponding return.
-
-\definition Given a trace \(\MPT\) and a call map \(\callmap\)
-%and a function map \(\codemap\),
-an \(n\)-{\em transition} is a pair of
-machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
-\(\mpstate{\mach_e}{\pol_e}\) in the trace such that
-\(\mach_c\) is a valid call with \(n\) arguments, and
-\(\mpstate{\mach_c}{\pol_c}\) steps to \(\mpstate{\mach_e}{\pol_e}\).
-
-\apt{or is in some function?? and why do we need either??}:
-\sna{I think Andrew is right that we don't need to know the function
-     here, or that the entry point is valid; that's part of WBCF.}
-%and \(\mach_e\) is an entry point for some function \(f\):
-
-  \[\begin{split}
-    \mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}
-    \triangleq & \exists \obs, f.
-    ~ \mpstate{\mach_c}{\pol_c} \stepstopol{\obs} \mpstate{\mach_e}{\pol_e} \\
-    & \land \callmap(\mach_c(\PCname)) = n
-      %\land \codemap(\mach_e(\PCname)) = f
-  \end{split}\]
-
-\definition
-
-{\em Stack integrity} is a property of a system with respect to a call map
-\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
-and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
-on the prefix up to its return, if any. For any initial state \(\mach_0\)
-we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
-\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
-\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
-the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
-\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
-\(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
-system as a whole enjoys stack integrity with respect to \(\callmap\).
-
-\apt{Some more words needed here to smooth the technicalities.}
-\sna{Attempt...}
-
-\subsubsection{Stack Confidentiality}
+for every transition between a caller and a callee for the entirety of
+the trace of the callee.
+
+\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
+a function map $\codemap$, an $n$-{\em transition} is a pair of
+machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
+in the trace such that:
+\begin{itemize}
+\item one state pair steps to the other:
+  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
+\item $\mach_c$ is a valid call with $n$ arguments:
+  $$\callmap(\mach_c(\PCname)) = n,$$
+\item and $\mach_e$ is an entry point for $f$:
+  $$\codemap(\mach_e(\PCname)) = f.$$
+\end{itemize}
+
+\definition{Stack Integrity} Let $\callmap$ be a call map and
+$\codemap$ a function map. We say that a system satisfies {\em stack
+  integrity} with respect to $\callmap$ and $\codemap$ if, for every
+initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
+trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
+$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
+$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
+(if any) satisfies trace integrity with respect to the contour at the
+call:
+%
+$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
+\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
+\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$
+
+\paragraph*{Stack Confidentiality}
 
@@
 property: just like integrity ensures that a caller's data is never
-overwritten by a callee, confidentiality ensures that the caller's data never
-influences the callee's observable behavior. We will formally define
+overwritten by a callee, confidentiality ensures that it never
+influences its observable behavior. We will formally define
 confidentiality in two similar steps: a trace version and a system
@@
 Unlike integrity, confidentiality is a noninterference-style property,
-comparing the behavior of a machine with an arbitrary variation that
+comparing the behavior of a machine with an arbitrary variant that
 preserves low confidentiality data.
 
-\definition%{Variations}
-We say two machine states \(\mach\) and \(\mach'\) are {\em variations}
-of one another given a contour \(\contour\) if they agree on low
+\definition{Variations}
+We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
+another given a contour $\contour$ if they agree on low
 confidentiality data:
@@
 Then, the trace version of confidentiality operates on two traces: the
-``original'' trace $\MPT \in \MPOTRACE$ that corresponds to a callee $f$
-and a ``variant'' trace $\machT \in \MTRACE$ that corresponds to how $f$
-would behaves if it were started from a variant initial machine state.
+``original'' one $MP \in \MPOTRACE$ that corresponds to a callee $f$
+and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
+if it started from a variant initial machine state.
 %
@@
 
-\definition
-
-{\em Trace confidentiality} will be a property of two traces, \(\MPT\) and
-\(\machT\), with respect to a return predicate \(R \subset \MACHS\) that
-reflects the states that are considered to have returned. We abstract over
-\(R\) because what it means to return depends on the state from which
-\(f\) was called. There are three cases in which \(\MPT\) and \(\machT\)
-enjoy trace confidentiality with respect to \(R\).
-
-\sna{Interestingly, split out like this, we don't need to talk about
-     contour here -- they are varied in the next property.}
+\definition{Trace Confidentiality} We will say two traces $MP$ and $M$
+satisfy {\em trace confidentiality} with respect to $\contour$ and
+a return predicate \(R \subset \MACHS\) if:
 \begin{itemize}
-  \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
-    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
-    some final state \(\mach_r'\) and any changes from the initial states are
-    matched between \(\mach_r\) and \(\mach_r'\). In addition their
-    observations are similar.
-    \[\begin{split}
-      \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
-      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
-        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
-      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
-        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
-      & ~ (\mach_e(\component) \not = \mach_r(\component)
-        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
-      \mach_r(\component) = \mach_r'(\component))
-    \end{split}\]
-  \item If \(\MPT\) is infinite, then \(\machT\) is also infinite and their
-  observations are similar.
-    \[\begin{split}
-      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
-      & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
-      & \neg (\exists mo . \last(\machT) = mo)
-        \land \pi_o(\MPT) \eqsim \pi_o(\machT)
-    \end{split}\]
-\item If \(\MPT\) ends prematurely without fulfilling \(R\), then
-  its observations are a prefix of those of \(\machT\).
-    \[\begin{split}
-      \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
-      & \forall \mach_r ~ \pol_r ~ \obs_r .
-        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
-      & \neg ~ R ~ \mach_r \Rightarrow \\
-      & \pi_o(\MPT) \lesssim \pi_o(\machT)
-    \end{split}\]
+\item If $\MPT$ is finite and its final state $R\ \mach_r$,
+  then $\machT$ is also finite with some final state $\mach_r'$ and
+  any changes from the initial states have changed in the same way:
+$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
+  \vee \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
+  \mach_r(\component) = \mach_r'(\component)$$
+  In addition their observations must be equivalent:
+  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
+\item If $\MPT$ is infinite, then $\machT$ is also infinite and their
+  observations are equivalent:
+  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
+\item If the $\MPT$ ends prematurely, without fulfilling \(R\), then
+  we require that its observations are a prefix of those of $\machT$:
+  $$\pi_o(\MPT) \lesssim \pi_o(\machT)$$
 \end{itemize}
 
-Putting it all together:
-
-\[\begin{split}
-  \propdef{TraceConfidentiality} ~ R ~ \MPT ~ \machT \triangleq
-  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
-  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
-    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
-\end{split}\]
-
-In the non-return cases, we only care that the traces exhibit the same
-observable behavior (or a prefix if the trace halts prematurely), as changes
-in the state will not affect the caller absent a return.  In the return case
-we must additionally guarantee that no changes were made based on secret data
-that could corrupt the caller upon return.
-
-Consider the result of modifying line 42 of our running example to $r_4 \leftarrow [\rsp - 9]$.
-Instead of reading the proper argument {\tt v}, {\tt h} now reads directly from the
+For example, consider the unsafe substitution:
+\[
+42 : ~ \sw ~ r_4 ~ \rsp 100
+\]
+Instead of reading the proper argument, {\tt h} reads directly from the
 secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
 shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
-in {\high} and low confidentiality in {\low}. Below it, the variant state
+in {\high} and low confidentiality in \low. Below it, the variant state
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
-hold -5, in {\varied}.  After the last step, the return states \(\mach_r\) and
+hold -5, in orange. After the last step, the return states \(\mach_r\) and
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
-\(\mach_e'\). Addresses 108 and 109 have changed, shown in light {\low}, and while 109
-had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
+\(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
+is consistent, 108 holds either 43 or -4, so the trace does not obey
+trace confidentiality.
 \rb{Color name macros}
-\apt{Explain slash notation in effects column.}
-
-As another example, 
-changing line 30 of the running example to $r_5  \leftarrow [\rsp - 6]$ would cause
-{\tt g} to read the value of {\tt main}'s local variable {\tt x} and then immediately
-output it, violating the internal observation clause of the confidentiality definition.
+
+Meanwhile, the following unsafe substitution outputs the secret
+directly, violating the internal observation clause of confidentiality:
+\[
+30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+\]
+%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
+%and \(\mach_e'\), below. Low confidentiality data in green is shared
+%between them, while high confidentiality data in red and yellow may differ.
+%After executing the code shown, the final states below have written different
+%values to address 42, represented by the black and white components.
+%It is allowed for address 40 to differ between the states due to the initial
+%variation, but address 42 changed differently between the traces, representing
+%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
+%is read, it is overwritten before return, and the write at step 3 always writes
+%the same value. \Gprivc and red locations differ between variations only because
+%of the initial variation, and the rest are identical. So this example passes.
 
@@
 
+%\Cref{fig:confex2} gives an example of a trace that fails trace
+%confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
+%Since {\tt x}, at address 100, is high confidentiality (red) the variant
+%may have a different value (yellow). So both \(r_4\) and address 107
+%differ in the final state between the original trace and the variant one,
+%failing the ending condition of trace confidentiality.
+
 To lift trace confidentiality to a system property, we follow the same
@@
 
-\definition%{Stack Confidentiality}
-%Let $\callmap$ be a call map and $\codemap$ a function map.
-We will say that a system satisfies {\em stack confidentiality} with respect to
-\(\callmap\) in a similar way to stack integrity. For every initial state
-\(\mach_0\), we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)}
-\hookrightarrow \MPT\). Then for any \(n\)-transition
-\(\mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~
-\mpstate{\mach_e}{\pol_e}\), we take the prefix trace until return:
-%
-\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\]
-
-We also take the corresponding trace from any variant \(\mach_e' 
-\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
-%
-\[\mach_e' \hookrightarrow \machT_{call} | \ret{\mach_c}\]
-
-If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
-\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
-stack confidentiality with respect to \(\callmap\).
-
-\apt{oy. still needs work to make this intelligible}
-\sna{Attempted. There's a tension between introducing all the variables
-     and keeping a good flow that I'm not sure I'm managing here though.}
+\definition{Stack Confidentiality}
+Let $\callmap$ be a call map and $\codemap$ a function map. We say that a
+system satisfies {\em stack confidentiality} with respect to
+$\callmap$ and $\codemap$ if, for every initial machine-policy state
+pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
+$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
+$f$ in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
+\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
+For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
+\mach_e\), let \(\machT\) be the similarly prefixed trace such that
+\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
+\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
+and \(R = \ret{\mach_c}\):
+
+$$\forall \MPT_{\mathit{call}} ~ \machT. ~
+\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
+$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
+\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
+$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$
 
@@
 \end{itemize}
-\apt{Could cut formal definition for space.}
-
-\definition%{Return Integrity}
-A system enjoys {\em return integrity}
-% with respect to a call map \(\callmap\) and return map \(\retmap\)
-if, for any initial state pair 
-$\mpstatename$ and any state pair \((\mach_c, \pol_c)\)
+A system enjoys {\em return integrity} with regard to a call map
+\(\callmap\) and return map \(\retmap\) if for any initial state pair
+\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
 corresponding to a call in its induced trace, if
-\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
+\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
 is the first unmatched return such that for some $\pol_r$
@@
 
+For example, if we added an instruction between 45 and 46:
 \[
-\exists p_r.\Unmatched ~ \MPT ~ (m_r,p_r)\MPT' \Rightarrow \ret{\mach_c}\ \mach_r.
+45.5: ~ \addi ~ \rra ~ r_0 ~ \negate 19
 \]
-\apt{I've tried to tidy this, but I don't think it corresponds to the Coq.
-  Shouldn't we apply $\Unmatched$ to the tail of $\MPT$?
-  And what about the no-unmatched-return-at-top-level clause?}
-
-For example, if we modified the running example by adding
-an instruction $ \rra \rightarrow \rra - 19$
-between 45 and 46, this would cause the return from {\tt h} to go to instruction 8 and
+This would cause the return from {\tt h} to go to instruction 8 and
 return to {\tt main} instead of {\tt g}. While the instruction that
@@
 
-\paragraph*{Strong Stack Safety}
-
-\leo{Maybe lockstep?}
-
-This conservative policy does not only satisfy the stack safety above,
-but also a provably (in Coq) stronger one that enforces local state
-encapsulation throughout every calls execution. To encode that
-property, we not only keep track of the execution of one machine
-(whose trace we're trying to characterize as safe), but of an entire
-stack of machine variations corresponding to nested calls of the
-original machine. For each step of the original machine, we take a
-step in every variant machine (thus, keeping them in lockstep
-execution) and enforce a single-step variant of integrity and
-confidentiality tests. In addition, every time a call is made we push
-a variation of the original machine state at the time of that call on
-top of the variant stack (just like the stack confidentiality property
-enforced trace confidentiality for every transition), and every time
-the machine returns from a call we pop the top variation (to account
-for stack confidentiality taking the prefix of the transition trace
-until its return).
-
-We start by defining the integrity and confidentiality tests, as well
-as the variant stack, before giving the full definition of strong
-stack safety.
-
+\paragraph*{Testing the conservative policy}
+%
+This conservative policy can be shown to satisfy an eager \emph{testing
+property}, which at each step of the computation performs integrity and
+confidentiality tests for a stack of running variants, which are kept in
+lockstep with the original machine.
+
+\definition{Stack integrity test}
+  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
+  satisfy the \emph{stack integrity test} if their high-integrity components
+  match:
+  \[
+    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
+    \forall \component . ~
+    \contour(\component) = \HIGHINT \rightarrow
+    \mach(\component) = \mach'(\component)
+  \]
+
+\definition{Stack confidentiality test}
+  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
+  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
+  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
+    confidentiality test} if their observations and changed components coincide:
+  \[
+    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
+    \obs_1 = \obs_2 \wedge \forall \component . ~
+   (     \mach_1(\component) \neq \mach_1'(\component)
+    \vee \mach_2(\component) \neq \mach_2'(\component))
+    \rightarrow \mach_1'(\component) = \mach_2'(\component)
+  \]
+\apt{scoping of $\forall$ is ambiguous.}
+\rb{Fix notations, including for properties below.
+  In particular, note that the last four arguments could be computed from the
+  first two and it may be clearer to elide them.}
 \newcommand{\vse}{\mathit{vse}}
@@
 
-\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
-\newcommand*{\testInt}[3]{\ensuremath{\mathit{int}_T ~ #1 ~ #2 ~ #3}}
-\newcommand*{\testConf}[2]{\ensuremath{\mathit{conf}_T ~ #1 ~ #2}}
-\newcommand*{\testVS}[2]{\ensuremath{\mathit{\vs}_T ~ #1 ~ #2}}
-\newcommand*{\testProp}[3]{\ensuremath{\mathit{stack}_T ~ #1 ~ #2 ~ #3}}
-
-\definition
-  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
-  satisfy the \emph{stack integrity test} if their high-integrity components
-  match:
-  \[
-    \testInt{\contour}{\mach}{\mach'} \triangleq
-    \forall \component . ~
-    \contour(\component) = \HIGHINT \rightarrow
-    \mach(\component) = \mach'(\component)
-  \]
-
-\definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
-\emph{stack confidentiality test} if the machines they stepped to
-have their changed components coincide and the observations during
-these steps are equal:
-\[
-\begin{array}{rl}
-  \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \triangleq &
-  \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
-  \mach_1 \stepstoobs{\obs_1} \mach_1' \wedge \mach_2 \stepstoobs{\obs_2} \mach_2' \Rightarrow \\
-&    \obs_1 = \obs_2 \wedge \forall \component . ~
-   (     \mach_1(\component) \neq \mach_1'(\component)
-    \vee \mach_2(\component) \neq \mach_2'(\component))
-    \rightarrow \mach_1'(\component) = \mach_2'(\component)
-\end{array}
-  \]
-
-\definition The testing property keeps track of a variation machine
-for every nested call performed so far on a {\em variant stack}: a
-list of machine states annotated with a contour, each element of which
-(denoted $\vm{C}$) of which is a variation of the actual machine
-according to a contour $C$.
-%
-\footnote{The full formal definition in Coq also records additional
-  information (the initial machine state at the call, its initial
-  variation, the contour, and a return predicate) which we elide
-  here for simplicity of exposition.}
-%%
-%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
-%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
-%  $$\vs \in \VSS$$
-%%
-\iflater
-\leo{This can be elided}
+\definition{Variant stacks}
+  The testing property keeps track of relevant information on a stack whose
+  elements represent variants of the machine by recording an initial machine
+  state, an initial and current variant of the machine, a contour and a return
+  map:
+%
+  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
+  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
+  $$\vs \in \VSS$$
+%
   Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
@@
   \end{itemize}
-  \fi
-
-%% <<<<<<< HEAD
-%% \definition
-%% %  Let $\callmap$ a call map and $\codemap$ a function map.
-%%   Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
-%%   variant stack $\vs$, we say that an MP-trace induced from $(\mach,
-%%   \pol)$ satisfies the \emph{strong stack safety} at a given call
-%%   depth as follows.
-%% % Refactor repetitive notation before case split
-%%   In all cases that let $(\mach, \pol) \hookrightarrow_P
-%%   (\mach', \pol', \obs)$.
-%%   %and assume (unless the step is a return) that the
-%%   %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
-%%   %machines can step.
-
-%% %  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
-%% %  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
-%% %  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
-%% %  testing property} is satisfied at call depth $d$, and write
-%% %  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:
-
-%%   \begin{itemize}
-
-%%   \item
-%% %
-%%     If the machine takes a \emph{call step} and $\contour$ is the contour at
-%%     the call then,
-%% %
-%%     (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
-%%     satisfied,
-%% %
-%%     (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
-%%     satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-%%     is satisfied with resepct to an arbitrary $\contour$-variant
-%%     of $\mach$,
-%% =======
-We also need two auxilliary notations that relate to variant stacks.
-% 
-First, strong safety will check that the basic tests above hold for every
-machine in the variant stack.  Given a machine that steps $(\mach,
-\pol) \hookrightarrow_P (\mach', \pol', \obs)$, say that a variant
-stack $\vs$ safisfies the \emph{variant stack test} if for each
-variant element $\vm{C}$ (which is a variation of $\mach$ for $C$),
-$\vm{C}$ can take a step, denoted $\vm{C} \stepstoobs{\obs} \mach'$,
-while satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
-and stack confidentiality test $\testConf{\mach}{\mach'}$ with regard
-to the original machine state. We will write $\testVS{(\mach,
-  \pol)}{\vs}$ for this compound check.
-%% >>>>>>> 1886f7a7bf858ce0c9518e407d9481518a4f028d
-%
-Second, we will write $\vs \stepsto \vs'$ if every variant machine
-in $\vs$ steps to the corresponding machine in $\vs'$.
-%
-
-We are now ready to define strong stack safety coinductively.
-
-\definition Given a state $(\mach, \pol) \in \MPS$ and a variant stack
-$\vs$ in which every element $\vm{\contour}$ is a variation of $\mach$
-with respect to its contour, we say that an MP-trace induced from
-$(\mach, \pol)$ satisfies the \emph{trace strong stack safety}, and write
-$\testProp{}{\MPT}{\vs}$, if:
+
+\definition{Variant stack test}
+%
+  Our testing property extends the basic tests above to whole variant stacks.
+  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
+  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
+  $\mach$ safisfies the \emph{variant stack test} if for each variant element
+  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
+  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
+  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
+  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
+  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
+
+\definition{Eager testing property}
+  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
+  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
+  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
+  testing property} at a given call depth if:
 
@@
 
-\item
-  If the machine takes a \emph{non-call, non-return step}
-  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)\MPT,\]
-  then the variant stack test must be satisfied:
-  \[\testVS{(\mach, \pol)}{\vs},\]
-  and trace strong stack safety must be coinductively satisfied for
-  the remainder of the trace and the variant stack $vs'$ that
-  steps to ($\vs \stepsto \vs'$):
-  \[ \testProp{}{((\mach',\pol',\obs)\MPT)}{\vs'}\]
-
-\item
-  If the machine takes a \emph{call step} and $\contour$ is the contour at
-  the call
-  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \callmap(\mach) = n,\]
-%
-  then the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
-  satisfied:
-  \[\testVS{(\mach, \pol)}{\vs},\]
-%
-%    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
-%    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-%    is satisfied with regard to an arbitrary $\contour$-variant
-%    of $\mach$,
-%
-  and trace strong stack safety must be coinductively satisfied
-  for the remainder of the trace and for every variant stack
-  that is the result of adding a variation of $\mach'$ with
-  respect to $C$ at the top of the stack $\vs'$ that $\vs$ steps
-  to:
-  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\vm{\contour} : \vs')}\]
-
-\item
-%
-  If the machine $\mach$ takes a \emph{return step},
-  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \retmap(\mach),\]
-  %
-  the top of the variant stack also returns,
-  the tail of $\vs$ satisfies the stack safety test
-  \[\testVS{(\mach, \pol)}{\mathit{tail}(\vs)},\]
-  %
-  and trace strong stack safety must be coinductively satisfied for
-  the remainder of the trace and for the tail of the varant stack
-  that $\vs$ steps to:
-  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\mathit{tail}(\vs'))}\]
-
-\item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
-    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
-    $\testProp{}{[(\mach, \pol)]}{\vs}$.
-
-%  \item If the machine takes a \emph{simultaneous return-call step}, that is,
-%    the instruction to which execution returns immediately after completing a
-%    call is itself a call instruction, the \emph{call step} and \emph{return
-%    step} cases are combined.
+  \item If the machine fail-stops, i.e.,
+    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
+    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at any
+    depth up to the size of $\vs$.
+
+  \item If the top variant of the stack (and only the top variant) returns,
+    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at depth
+    equal to the size of the stack. \rb{Avoid confusion with return below}
+
+  \item If (1) the machine takes a step,
+    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    that is neither a call nor a return, (2) all
+    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
+    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
+    (these tests are required also for calls and returns, and omitted in those cases),
+    and (3) the post-step state $(\mach', \pol')$ and stack $\vs'$ satisfy the eager testing property
+    for a given depth $d$ \rb{omitting stack size condition} and MP-trace $\MPT$,
+    then the property is also satisfied for trace $(\mach, \pol) : \MPT$
+    at depth $d$.
+
+  \item If the machine takes a call step
+    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    the contour $\contour$ is updated based on the top of the current machine.
+    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    (and in addition to the general step conditions),
+    the eager integrity test must hold for $\contour$ at the call,
+    the eager confidentiality test must hold for all variants at the call,
+    including the new variant that is added to the top of the stack,
+    and the property recurses on an arbitrary variant at the new contour
+    at depth $d$ with trace $\MPT$.
+
+  \item If the machine takes a return step at the top of the variant stack
+    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
+    the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
+    if (in addition to the general step conditions)
+    the property recurses at depth $d$ with trace $\MPT$
+    after removing the returning top variant from the stack.
+
+  \item If the machine takes a return step that is also a call, the last two
+    cases are combined.
 
@@
 
-\definition 
-%
-A system satisfies {\em strong stack safety} if for every initial machine $\mach$,
-the trace induced from $(\mach, \pinit{\mach})$ satisfies trace strong stack safety with
-respect to the empty stack.
-
-\theorem
-Strong stack safety implies stack safety.
-
-\proof
-By coinduction on trace strong stack safety.
-\leo{3000 lines Coq?}
-\leo{I can write a TON about this proof, but I'm not sure how much is interesting/how much space we have}.
-
-\paragraph*{Eager vs Lazy Stack Safety}
+By using property-based testing techniques, we can determine with a high degree
+of confidence that an implementation of the conservative micropolicy satisfies
+this eager testing property. Although we have not proved that the conservative
+policy guarantees the stack safety property, the testing property is provably
+stronger than the stack safety property \rb{theorem?} and, by this implication,
+we can conclude that the conservative policy also enforces stack safety.
+
+% The properties we give are indeed enforceable.
+
+\rb{What is the strongest property enforced by the conservative policy?}
+
+\paragraph*{End-to-end stack safety property}
+
+Integrity
+
+Confidentiality as a form of noninterference
+
+\paragraph*{Eager vs lazy testing}
 
@@
 effect becomes manifest.
-
-\leo{Not sure how much of the following could go after the lazy stack safety property}
+\rb{How much detail to go into? Refer to examples if we have the space?}
+
 These \emph{lazy policies} admit more efficient implementations while being
@@
 
-\apt{Need some intro text here.}
-
-\subsection{Observable Integrity}
-
     \paragraph{Rolling Back to Idealized State}
-      Observable integrity weakens eager integrity by allowing a returned
+      Observable integrity will weaken eager integrity by allowing a returned
       state to differ from its call state at high integrity components,
-      provided that the differences do not change the observable behavior
-      of the rest of the program.  
-%      provided the observable behavior of the induced trace from the return
-%      matches that of an idealized state whose high integrity components do
-%      match the call state.
+      provided the observable behavior of the induced trace from the return
+      matches that of an idealized state whose high integrity components do
+      match the call state.
 
@@
       \begin{cases}
-        \mach_c(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
-        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
+        \mach_c(\component) & \contour(\component) = (\HIGHINT,\_) \\
+        \mach_r(\component) & \contour(\component) = (\LOWINT,\_) \\
       \end{cases}\]
@@
 
-    \definition%{Observable Trace Integrity}
+    \definition{Observable Trace Integrity}
 
@@
       some contour \(\contour\) if its final state (if any) has observable
-      behavior \apt{states don't have behavior} that prefixes that of its idealized version, constructed by
+      behavior that prefixes that of its idealized version, constructed by
       \(\mathit{roll}\):
@@
         & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT' \Rightarrow \\
         & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
@@
 
-    For our example, we return to the violation of trace integrity in
+    For out example, we return to the violation of trace integrity in
     \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
@@
     before printing, observable trace integrity would hold.
-    \apt{Better to give an example in which {\tt h} overwrites something that
-      is not needed to compute an output, so OTI does not fail.
-      E.g., {\tt h} could overwrite {\tt w} with nonsense before returning -- but
-      this doesn't matter, because the return value will replace the nonsense
-      before this figures in the final print. 
-      Also could be good to give an example in which we 
-      overwrite {\tt x} instead of {\tt z} to show that we still fail OTI -- in other words,
-      the entire remainder of the program matters, not just the caller.
-    }
-
-    \definition%{Observable Stack Integrity}
-
-    A system enjoys {\em observable stack integrity}
-    % with respect to    \(\callmap\) and \(\codemap\)
-    if, for any initial state \(\mach_0\) and
+
+    \definition{Observable Stack Integrity}
+
+      A system enjoys {\em observable stack integrity} with regard to
+      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
       each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
@@
       \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_e}\),
       then \(\text{\sc ObsInt}_\contour ~ \MPT\).
-\apt{more words please}
-
-\subsection{Observable Confidentiality}
 
@@
       end condition over-approximates the extension of the behavioral condition
-      to future execution.\apt{That's quite mysterious.}
+      to future execution.
 
@@
       This requires restoring varied components to their original values,
-      since after return they will once again be legitimately accessible.
+      since after return they will once again be accessible legitimately.
 
@@
 
-    \definition{%Observable Trace Confidentiality}
-
-      Let \(\MPT\) and \(\machT\) be a pair of traces starting from
+    \definition{Observable Trace Confidentiality}
+
+      A pair of traces \(\MPT\) and \(\machT\) starting from
       \(\mpstate{\mach_e}{\pol_e}\) and \(\mach_e'\), respectively, with
-      \(\mach_e \approx_\contour \mach_e'\). \apt{Should some of this move into the defining clauses below?} Then
-      $\MPT$ and $\machT$ enjoy \emph{observable trace
-        confidentiality} with respect to \(\contour\), written
-            \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\), if:
-
-      \begin{itemize}
-      \item
-      the two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
-      \pi_o(\machT)\); and
-
-      \item
-      if \(\MPT\) is finite and ends with the state
-      \(\mpostate{\mach_r}{\pol_r}{\_}\), then \(\machT\) is finite, and
-      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\), where
-      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\),
-      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \hookrightarrow \machT'\),
-      and \(\mach_r' = \last(\machT)\). 
-      \end{itemize}
+      \(\mach_e \approx_\contour \mach_e'\), enjoy observable trace
+      confidentiality with respect to \(\contour\) if two conditions hold:
+
+      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
+      \pi_o(\machT)\).
+
+      If \(\MPT\) is finite and ends with the state
+      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
+      \(\mach_r' = \last(\machT)\). Then the induced traces
+      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT'\) and
+      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
+      \hookrightarrow \machT'\) are observationally prefixed,
+      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).
+
+      We write this \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).
 
@@
       In our example, \cref{fig:conflex} shows how execution continues from
-      \(\mach_r'\) and the restored
-      \[\mach_r'' = \mathit{restore}(\contour,
-      \mach_e,\mach_e',\mach_r,\mach_r').
-      \]
-      Addresses 108 and 109
+      \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
+      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109, in \tainted,
       are not restored; every address that was varied and not changed in one
@@
 
-    \definition%{Observable Stack Confidentiality}
-
-      A system enjoys {\em observable stack confidentiality}
-      %with respect to \(\callmap\) and \(\codemap\)
-      if, for any initial state
+    \definition{Observable Stack Confidentiality}
+
+      From observable trace confidentiality we construct {\em observable stack
+      confidentiality}. A system enjoys {\em observable stack confidentiality}
+      with regard to \(\callmap\) and \(\codemap\) if for any initial state
       \(\mach_0\) and each transition with entry state
@@
 
-    \definition%{Observable Stack Safety}
-    A system enjoys {\em observable stack safety}
-    % with respect to a call map       and a function map
-    if it enjoys both observable stack integrity and
-      observable stack confidentiality with respect to them.
-
-\apt{Words needed to connect these ideas back to the lazy policies which motivated them in the first place!}
+    \definition{Observable Stack Safety}
+      A system enjoys {\em observable stack safety} with regard to a call map
+      and a function map if it enjoys both observable stack integrity and
+      observable stack confidentiality with regard to them.
 

### 2020-07-09 main.tex

--- 2020-07-09 main.tex
@@
 
-To phrase these two aspects of stack safety in terms of security
-properties \leo{in terms of security properties or ``using security
-  language''?}  requires an explicit attacker model: a clear
-understanding of who is being protected and from what. Most existing
-stack protection mechanisms deal with scenarios where a callee gets
-influenced by external inputs and ends up executing potentially
-malicious code. To avoid the details of exactly how this happens and
-keep the presentation clear, we pick an even stronger model: when
-stating the protections that a function can rely on when it makes a
-call, we assume that the callee can execute arbitrary instructions
-until it returns.
-
-Adopting that model in our setting, however, poses another novel
-challege: not only does a caller need to be protected from its callee,
-but the callee itself needs to be protected from {\em its}
-callees. That is, the notion of ``attacker'' changes throughout the
-execution of a program!  We address this problem by using a similarly
-changing notion of accessibility for the different components of the
-machine: we adjust the integrity and confidentiality level of
-different stack locations and registers at every transition (call or
-return) between functions.  \bcp{that's way too terse!}\leo{Slightly
-  better, still needs a bit probably}
-%\apt{connection to ``attacker'' still vague.}  \rb{Mention
-%  to varying ``accessing power'' of the attacker?  The notion of
-%  attacker changes on certain actions that are not externally
-%  visible.} \bcp{Maybe the best way is to grab the bull by the horns:
-%  ``If we're going to phrase stack safety as a security property, we
-%  need to talk about an attacker model, and this is where the
-%  technicalities get really interesting...''}
+To phrase these two aspects of stack safety as security properties requires
+an explicit attacker model: a clear understanding of who is being protected
+and from what. Most existing stack protection mechanisms envisage scenarios
+where a callee is influenced by external inputs to corrupt the stack in some
+way. To avoid getting into the details of exactly how this happens when
+formalizing our properties, we use an even stronger model: when we state the
+protections that a function can rely on when it makes a call, we assume that
+the callee can execute arbitrary instructions until it returns.
+
+A key challenge in adapting the concepts of language-based security to this
+setting is that stack safety is a ``nested'' property: not only does a given
+caller need to be protected from its callee, but the callee also needs to be
+protected from {\em its} callees, and so on. That is, the notion of
+``attacker'' changes throughout the execution of the program.  We address
+this issue using a dynamically changing notion of {\em accessibility} for
+the different components of the machine state and adjusting the integrity
+and confidentiality levels of different stack locations and registers at
+every transition (call or return) between functions.
 
@@
 
-%% <<<<<<< HEAD
 %% \definition

### 2020-07-09 main.tex
Polishing

@@
 
+\bcp{We don't actually say very clearly anywhere that we are interested in
+  low-level enforcement mechanisms for machine-code programs (because this
+  is where stack attacks occur, and because at higher levels of abstraction
+  the stack is a built-in concept, not something that can be attacked even
+  in principle).  This could cause confusion.}
+
 The call stack is a perennial target for attacks
@@
 capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \iflater\apt{Mostly from
   nick;there could be more}\bcp{How about static techniques??  We decided
   these were also in-scope.}\bcp{... or are they?  I guess in any case we should
-list any that we know about.}
-
-But what {\em is} this ``stack safety'' of which we speak? Precise formal
-definitions are lacking in the literature; instead, the notion
-is usually defined negatively, by enumerating the bad things that a given
-enforcement mechanism
+list any that we know about.}\fi
+
+But what {\em is} this stack safety of which people speak? Precise formal
+definitions are mostly lacking: rather, the notion
+is commonly defined in a negative way, by enumerating the bad things that a
+given enforcement mechanism
 prevents, such as corruption of return addresses, buffer overflows, or
-use of uninitialized variables.
+use of uninitialized variables\bcp{Does use of unitialized variables lead to
+stack safety attacks?}.
 %with the notion usually being
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-Indeed, we are only aware of one---quite recent---attempt to frame a
-positive definition:
-\citet{Skorstengaard+19} define stack safety as
-the conjunction of two properties, {\em local state encapsulation} and {\em
+Indeed, we are only aware of one quite recent attempt to frame a
+positive definition, by
+\citet{Skorstengaard+19}, who define stack safety as
+the conjunction of two properties: {\em local state encapsulation} and {\em
   well-bracketed control flow}.
@@
 %
-Although this represents an important step in the right direction, we believe
+While this represents an important step in the right direction, we believe
 there is still room to craft a more abstract, more generic characterization
@@
 technique.  In other words, we seek formal criteria that can be applied
-against \emph{any} stack-based system + enforcement mechanism to judge
-whether or not the combination is actually stack-safe.
-
-\iflater
-\leo{Should this be its own paragraph?}
-\fi
-The key contribution of this paper is a novel, generic presentation of
+to \emph{any} mechanism for protecting stack-based low-level programs to judge
+whether or not it actually ``guarantees stack safety.''
+%
+Our key contribution is a novel, generic presentation of
 local state encapsulation and well-bracketed control flow using the
@@
 callees: a callee should not be able to access or overwrite private data
-in the caller's stack frame. We formalize this intuition as a security
-property with both confidentiality and integrity components.
-%
-We reflect this intuition in a formal
-{\em stack confidentiality} property, which promises that a callee's
-observable behavior is not affected by memory outside its stack frame
-(e.g., a caller's private data or uninitialized memory).  Conversely, {\em stack
-  integrity} promises that, if a function call ever returns, the caller's
-stack data at the point of return will be identical to what it was before
-the call.
+in the caller's stack frame.
+%
+We formalize this intuition in (1) a {\em stack confidentiality}
+property, which promises that a callee's observable behavior is not affected
+by memory outside its stack frame (e.g., a caller's private data or
+uninitialized memory), plus (2) a {\em stack integrity} property,
+which promises that, if a function call ever returns, the caller's stack
+data at the point of return will be identical to what it was before the
+call.
 
@@
   flow} is
-the desire to protect the call graph itself: a callee should pass control
+the need to protect the call graph itself: a callee should pass control
 back to the caller only by returning in the standard way (transferring
@@
 \label{sec:running-example}
-
-\leo{We've said this literally in the previous sentence. What should this intro
-  be in the new reorg?}
-%We begin with a running example that will serve both to demonstrate several
-%``stack-unsafe'' behaviors that our stack safety property should prohibit
-%and to ground the technical presentation in later sections.
-
-The example program appears in \Cref{fig:running-program}. It consists
-of a {\tt main} function, written in some C-like high-level language,
-that performs a call to some function {\tt f}, which in turns makes a
-nested call to some function {\tt g} before returning, while
-performing some simple arithmetic operations in the process.
 
@@
 
-To show a trace of the execution of this program, we need a low-level
-view of both the code and the machine it runs on. While the concrete
-implementation of our enforcement mechanism (\Cref{sec:enforcement})\apt{Is it ``ours'' in this context?}
-operates on a full conventional RISC-V architecture (using the FORVIS
-simulator\leo{cite}\apt{why mention this here?}), we make a few simplifying assumptions here to
-ease the exposition.
-
-First of all, we ignore the heap. We are mainly concerned with stack
-interactions so that will be the main focus; integrating a safety
-propety for the heap in the
-style of \citet{AzevedoDeAmorim17} is a natural extension we discuss
-in~\Cref{sec:future}.
-%
-We also abstract away low-level details of memory layout and present a
-machine that is word-addressed, with instructions starting from
+\Cref{fig:running-program} shows a high-level view of our running example,
+written in C-like pseudo-code.  It consists of a {\tt main} function that
+performs a call to some function {\tt f}, which in turns makes a nested call
+to some function {\tt g} before returning, while performing some simple
+arithmetic operations in the process. \cref{fig:running-trace-b} shows an
+execution trace of corresponding machine instructions. (These might have
+been generated by compiling the high-level code, but it is only the behavior
+of the machine instructions that we care about
+here---\Cref{fig:running-program} is our actual running example.)
+
+To streamline the exposition of the example, we use a small instruction set
+reminiscent of RISC-V and make few simplifying assumptions.
+%
+First, since our goal is to formalize stack safety, we ignore the heap and
+assume that the memory is divided between the instructions and the stack.
+(Integrating a memory safety property for the heap in the style of
+\citet{AzevedoDeAmorim17} is a natural extension that we discuss
+in~\Cref{sec:future}.)
+%
+We also abstract away low-level details of memory layout and assume a
+machine that is word- (not byte-) addressed, with instructions starting from
 location 0, a dedicated memory-mapped output port at location 99, and
@@
 addresses.  \leo{Explain more? How to connect to realistic
-  one? Observations? DMA? Interrupt handlers?}
-%
-A sketch of the initial memory layout can be found at the top of
-\cref{fig:running-trace-b}: instructions for {\tt main} begin at 0,
+  one? Observations? DMA? Interrupt handlers?}\bcp{Yes.}
+
+A sketch of the initial memory layout appears at the top of
+\cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
 the dedicated output is uninitialized, while the (also uninitialized)
@@
 instruction used for returns jumps to the address stored in a register.
-Register $r_0$ is hardwired to zero. \leo{Expand upon this?}
+Register $r_0$ is hardwired to zero. \leo{Expand upon this?}\bcp{Yes, if it
+  matters to understanding the example---point to where.}
 

### 2020-07-09 main.tex
Polishing section 2

@@
 
-To streamline the exposition of the example, we use a small instruction set
-reminiscent of RISC-V and make few simplifying assumptions.
+To streamline the example, we use a small instruction set reminiscent of
+RISC-V and make a few simplifying assumptions about the organization of
+memory.
 %
@@
 %
-We also abstract away low-level details of memory layout and assume a
-machine that is word- (not byte-) addressed, with instructions starting from
-location 0, a dedicated memory-mapped output port at location 99, and
-the stack beginning at location 100 and growing upwards with increasing
-addresses.  \leo{Explain more? How to connect to realistic
-  one? Observations? DMA? Interrupt handlers?}\bcp{Yes.}
+We also assume that the machine is word- (not byte-) addressed, that program
+instructions start from location 0, that the machine communicates with the
+outside world through a single memory-mapped output port at address 99, and that the stack begins at location 100 and grows upwards.
+% \leo{Explain more?  How to connect to realistic one? Observations? DMA?
+%   Interrupt handlers?}\bcp{Yes.}
 
@@
 
-Stack safety is meant to protect a caller from a malicious callee's
+Stack safety is meant to protect a caller from a confused or malicious callee's
 attempts to access private data or highjack its control flow. Our
-attacker model reflects this intention: we assume that the attacker
-can run arbitrary code as a caller and the enforcement mechanism
-should protect against that. Hardware and timing attacks are out of
+attacker model reflects this intention by assuming that the attacker
+can run arbitrary code to try to access or overwrite its caller's stack data
+or disrupt its control state. Hardware and timing attacks are out of
 scope. \leo{Anything else to add here?}

### 2020-07-09 main.tex
wibbles

--- 2020-07-09 main.tex
@@
                           \text{ and } \component \geq \mach(\SP)
-                          \text{ or if } \component = \rout\\
+                          \text{ or } \component = \rout\\
     (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
@@
 the call from {\tt main} to {\tt g}, showing addresses annotated with their
-corresponding labels (not shown: the labels of the instructions
-corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
-$(\HIGHSEC, \HIGHINT)$)\apt{see previous note}:
+corresponding labels:
 \vspace*{0.2em}
@@
 Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
-\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
+\(\MPOTRACE\). We say \(\MPT\) satisfies {\em trace integrity} with
 respect to \(\contour\) if its final state (if any) agrees with its

### 2020-07-09 main.tex
More polishing

--- 2020-07-09 main.tex
@@
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \iflater\apt{Mostly from
-  nick;there could be more}\bcp{How about static techniques??  We decided
-  these were also in-scope.}\bcp{... or are they?  I guess in any case we should
-list any that we know about.}\fi
+  nick; there could be more}\fi
 
@@
 \vspace*{0.2em}
-\bcp{A very small nit: Change
-  $\rsp-2$ to $\rsp\mathord{-}2$}\apt{??}
+\iflater\bcp{A very small nit: Change
+  $\rsp-2$ to $\rsp\mathord{-}2$}\apt{??}\bcp{I meant remove spaces around +
+and $-$ when then are used inside brackets.  I think it will look better.
+But not critical.}\fi
 \[
-  \begin{array}{c|c|c|c}
+  \begin{array}{l|l|l|l}
     PC & \mathit{Instruction} & \mathit{Effect} & \\
@@
     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 102}
-        {Allocate private {\tt main} data}
+        {Allocate private data for {\tt main}}
     \row{\addi ~ r_4 ~ r_0 ~ 42}{r_4 \leftarrow r_0 + 42}{r_4 \leftarrow 42}
-        {Put 42 in $r_4$}
+        {Load 42 into $r_4$}
     \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp-2] \leftarrow r_4}{[100] \leftarrow 42}
@@
     \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[103] \leftarrow 0}
-        {Zero out result\apt{why?}}
+        {Zero out result\apt{why?}\bcp{+1}}
     \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
-        {Call to {\tt g} ($\jal$)}
+        {Call ($\jal$ to) {\tt g}}
   \end{array}
@@
   \[
-  \begin{array}{c|c|c|c}
+  \begin{array}{l|l|l|l}
     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 106}
-        {Allocate local {\tt g} state}
+        {Allocate stack space for {\tt w}}
     \row{\sw ~ \rsp ~ \rra ~ \negate 2}{[\rsp - 2] \leftarrow \rra}{[104] \leftarrow 8}
-        {Store return address}
+        {Save return address}
     % Zero out z?
     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 108}
-        {Allocate for call frame}
+        {Allocate space for call frame}
     \row{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow r_0 + 17}{r_4 \leftarrow 17}
-        {Store 17 to $r_4$}
+        {Store 17 into $r_4$}
     \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[106] \leftarrow 17}
@@
     \row{\jal ~ \rra ~ 40}{\rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow 40}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 40}
-        {Call to {\tt h} ($\jal$)}
+        {Call {\tt h}}
   \end{array}
@@
 \[
-\begin{array}{c|c|c|c}
+\begin{array}{l|l|l|l}
   \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
-      {Allocate for return address}
+      {Allocate space for return address}
   \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
-      {Store return address}
+      {Save return address}
   \row{\lw ~ r_4 ~ \rsp ~ \negate 3}{r_4 \leftarrow [\rsp - 3]}{r_4 \leftarrow 17}
-      {Load the argument {\tt v} to $r_4$}
+      {Load the argument ({\tt v}) into $r_4$}
   \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 18}
-      {Increment $r_4$ by 1}\end{array}
+      {Increment $r_4$}\end{array}
 \]
@@
 \[
-\begin{array}{c|c|c|c}
+\begin{array}{l|l|l|l}
   \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
@@
 \summary{27}{0}{\thepcctr}{108}{18}{?}
-\vspace*{0.2em}
+% \vspace*{0.2em}
 
@@
 % H - Body
-\vspace*{0.2em}
+% \vspace*{0.2em}
 \[
-\begin{array}{c|c|c|c}
+\begin{array}{l|l|l|l}
   \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
-      {Load return value to $r_4$}
+      {Load return value into $r_4$}
   \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
@@
 \[
-\begin{array}{c|c|c|c}
+\begin{array}{l|l|l|l}
   \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
@@
 
+\bcp{We really want readers to understand this trace.  I think we should
+  spend a few more words walking through it in detail.}
+
 \Cref{fig:running-program} shows a high-level view of our running example,
@@
 First, since our goal is to formalize stack safety, we ignore the heap and
-assume that the memory is divided between the instructions and the stack.
+assume that all variables are stack allocated.
 (Integrating a memory safety property for the heap in the style of
@@
 instructions start from location 0, that the machine communicates with the
-outside world through a single memory-mapped output port at address 99, and that the stack begins at location 100 and grows upwards.
+outside world through a single memory-mapped output port at address 99 (so
+a {\tt print} statement will be compiled to a store to this
+location), and that the stack begins at location 100 and grows upwards.
 % \leo{Explain more?  How to connect to realistic one? Observations? DMA?
@@
 instruction used for returns jumps to the address stored in a register.
-Register $r_0$ is hardwired to zero. \leo{Expand upon this?}\bcp{Yes, if it
-  matters to understanding the example---point to where.}
+Register $r_0$ is hardwired to zero (so that, e.g., instruction 3 stores $0$
+into address 101.).
 
@@
 Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
-policy state to the machine state and a partial policy step function that
+policy state to the ordinary machine state\bcp{This may be surprising to
+  people who are not thinking about micropolicies.} and a partial policy
+step function that
 refines the machine step function by allowing or disallowing a step to be taken
@@
 canaries \citep{Cowan+98}. These approaches are represented as policies that
-accept only initial machine states that are well-formed for the policy (due to
+accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
 being produced by the compiler.) The set of policy states is again the unit type
@@
 %
-Other policies are implemented in hardware or built on such hardware mechanisms.
+Other policies are implemented in hardware or built on hardware mechanisms.
 Capability machines like CHERI \citep{Woodruff+14} extend conventional
@@
 scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-the machine would be relatively straightforward.
+the machine would be relatively straightforward.  \bcp{Not sure that any of
+  this is comprehensible... :-(}
 
@@
 observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.
-
-\bcp{TODO: Let's also try to clean up the spacing a little -- many of these
-  sequences of rules can be typeset on fewer lines}
 

### 2020-07-09 main.tex
subsections in 6
Misc

@@
 %% comma separated list
-\keywords{Stack Safety, Micropolicy, Random Testing, RISC-V}  %% \keywords are mandatory in final camera-ready submission
+\keywords{Stack Safety, Micropolicies, Random Testing\bcp{Really??}, RISC-V}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
+  \paragraph{Heap Safety as a Security Property}
+\bcp{Talk about the relation.}
+    Heap safety, like stack safety, may be framed as a form of
+    noninterference\citet{DBLP:conf/post/AmorimHP18}. The above model
+    describing safety of
+    shared stack-allocated objects extends naturally to heap safety,
+    complicated by the fact that these objects are not deallocated by returns,
+    and therefore a function's privilege may increase or decrease after its
+    entry point. For instance, if a callee allocates a heap object and returns
+    the pointer to its caller, that object's addresses become accessible in
+    contradiction to a contour computed at the caller's entry. So such an
+    extension must enforce trace confidentiality and integrity properties
+    separately on continuous segments of each call, with contours computed at
+    each crossing between caller and callee.
+
 \section{Future Work}
@@
     deallocated along with its stack frame.
-
-  \paragraph{Adding Heap Safety}
-
-    Heap safety, like stack safety, may be framed as a form of noninterference
-    [cite Meaning of Memory Safety]. The above model describing safety of
-    shared stack-allocated objects extends naturally to heap safety,
-    complicated by the fact that these objects are not deallocated by returns,
-    and therefore a function's privilege may increase or decrease after its
-    entry point. For instance, if a callee allocates a heap object and returns
-    the pointer to its caller, that object's addresses become accessible in
-    contradiction to a contour computed at the caller's entry. So such an
-    extension must enforce trace confidentiality and integrity properties
-    separately on continuous segments of each call, with contours computed at
-    each crossing between caller and callee.
 

### 2020-07-09 main.tex
Tweaks

@@
 of the machine instructions that we care about
-here---\Cref{fig:running-program} is our actual running example.)
+here---\Cref{fig:running-trace-b} is our actual running example.)
 
@@
   \integrityexample
-  \caption{Failing Trace Integrity}
+  \caption{Failing Trace Integrity\bcp{the ``108'' is not placed right}}
   \label{fig:intex}
@@
 
-\paragraph*{A conservative policy}
+\subsection{A Conservative Policy}
 %
@@
 
-\paragraph*{Strong Stack Safety}
+\subsection{A Stronger Stack Safety Property}
 
@@
 
-\section{Discussion}
-
-\rb{Notes from meeting:
-  \begin{itemize}
-  \item Explain how the theory applies to production compilers like GCC and
-    LLVM, which aspects are directly applicable, what are the effects of
-    alternative calling conventions. Other less charted waters: tail calls,
-    coroutines, jumps into supervisor mode.
-  \item CHERI and its operation, documentation and practical configurations and
-    operation.
-  \item Compare and contrast with StkTokens.
-  \item Under what circumstances can stack safety exist without an enforcement
-    mechanism?
-  \item Zeroing out of allocated space inside blessed sequences, eager vs lazy.
-    Possibility of blessed alloc and dealloc sequences, which also manipulate SP
-    and would not be legal without blessing. In such an extended setting, the
-    return sequence would check for matching addresses (it would be nice to have
-    testing for this!).
-  \end{itemize}
-}
+% \rb{Notes from meeting:
+%   \begin{itemize}
+%   \item Explain how the theory applies to production compilers like GCC and
+%     LLVM, which aspects are directly applicable, what are the effects of
+%     alternative calling conventions. Other less charted waters: tail calls,
+%     coroutines, jumps into supervisor mode.
+%   \item CHERI and its operation, documentation and practical configurations and
+%     operation.
+%   \item Compare and contrast with StkTokens.
+%   \item Under what circumstances can stack safety exist without an enforcement
+%     mechanism?
+%   \item Zeroing out of allocated space inside blessed sequences, eager vs lazy.
+%     Possibility of blessed alloc and dealloc sequences, which also manipulate SP
+%     and would not be legal without blessing. In such an extended setting, the
+%     return sequence would check for matching addresses (it would be nice to have
+%     testing for this!).
+%   \end{itemize}
+% }
 

### 2020-07-09 main.tex
Intro done - unlocking. Locking abstract

@@
 
-\bcp{This little paragraph is our entire explanation of the second and third
-contributions!  Needs to be expanded considerably...}%
-To provide evidence that our definitions capture the inituitive notion
-of stack safety, we show that they rule out a range of exploitable bad
-behaviors.\apt{``from the literature''??}  To assess whether they are
-implementable, we compare them against
-%proposed capability-based
-%enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
-%  should} and
-the existing tag-based enforcement mechanism of
-\citet{DBLP:conf/sp/RoesslerD18}.
-%\leo{Is this where we would rephrase this because we would only talk about it in related work?}
+\leo{Took a stab. Needs rewriting}%
+
+After formally defining stack safety, we show that our definition is
+enforceable using an existing tag-based enforcement
+mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
+conservative Depth Isolation policy, which not only enforces the
+our formal definition of stack safety, but also a much stronger
+property that requires integrity and confidentiality at every step of
+an execution trace. This property is provably (in Coq)\leo{How to
+  mention this?} stronger and stemmed out of the need to more
+efficiently test stack safety. However, at the same time the
+conservative enforcement mechanism is expensive to implement. In the
+same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different
+mechanism that is more efficient by virtue of being lazy in its
+enforcement. At a high level, instead of guaranteeing integrity and
+confidentiality when a callee returns, they check it when the caller
+actually makes use of (potentially compromised) data. Thus, they
+enforce a (also provably in Coq) weaker property that still protects
+from unsafe callees, but is much less intuitive.
 
@@
   A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
-  This policy enforces our strict stack safety property, and also a (provably) stronger property that is easier to test.
+  This policy enforces our strict stack safety property, and also a (provably in Coq) stronger property that is easier to test.
   \apt{If we have no testing, do we even want to talk about the testing
@@
   but a modified version of it using unique per-activation tags enforces
-  a (provably) weaker lazy variant of the property.
+  an (also provably) weaker lazy variant of the property, which still prohibits
+  the same intuitively stack unsafe behaviors.
 %

### 2020-07-09 main.tex
More on observable properties

--- 2020-07-09 main.tex
@@
 
-\paragraph*{Eager vs Lazy Stack Safety}
+\paragraph*{Eager vs Lazy Policies}
 
@@
 
-\leo{Not sure how much of the following could go after the lazy stack safety property}
-These \emph{lazy policies} admit more efficient implementations while being
-harder to test, given that potential violations may or may not materialize,
-requiring in turn additional bookkeeping. However, a critical observation is
-that the depth-based tagging scheme discussed so far with fail when combined
-with a lazy policy. The reason for this is that because tags are not set and
-cleared on function entry and exit points, stale tags may persist after a given
-call, and a subsequent call sharing the same depth as the old call may access
-memory contents tagged with the stale tags of a previous activation.
-\rb{Again, example?}
-
-The lazy policy can be repaired if, instead of tagging each function activation
-with its depth in the stack, we generate a fresh activation identifier on each
-call, which allows the policy to identify stale tags from previous activations.
-Under this stricted tagging discipline, the eager policy behaves almost exactly
-like its lazy counterpart. A related mechanism was explored in the Static
-Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which associates a
-unique activation identifier to each function, which is however shared by all
-activations of the same function in order to obtain good cacheability of tag
-rules.
-
-\rb{Parts of this discussion feel somewhat out of order...}
-
-\rb{We can also test the lazy policy, even if less efficiently, and it should
-  enforce an inductive property, which may be more amenable to proving --- this
-  should be the connection to the following section.}
-
-\rb{Implied by the property we started with?}
-
-\rb{Explain and compare what each policy enforces.}
+These \emph{lazy policies} admit more efficient implementations, but they
+deliberately allow violations of stack integrity. So it becomes important to
+characterize what alternative property these policies ought to implement.
+
+\sna{moved a lot of discussion to after property, along with Rob's
+     comments on it}
 
@@
   justified because a truly harmful violation will be caught later.
-  In a typical lazy version of the tag policy, writes are always permitted,
-  but tagged to identify the context in which they were written. But reads
-  check that the value read was written by the current context and fail-stop
-  if it wasn't. So, if a callee makes an illegal write to its caller, when the
-  caller tries to read it the policy will catch the error.
-
   The natural question is, how do we characterize the protections provided
@@
   section use the observation model to draw this distinction.
-
-    \sna{Use the \(n\)-transition language above in all of these.}
-
-\apt{Need some intro text here.}
 
@@
 
-\apt{Words needed to connect these ideas back to the lazy policies which motivated them in the first place!}
+    \paragraph{Connection to Policies}
+
+    Observable properties allow us defer enforcement until a property
+    violation would become visible, as in lazy policies. But by formalizing
+    the connection to the eager policy, we can identify when lazy policies
+    miss violations that can become visible. The basic lazy tag policy from
+    \citet{DBLP:conf/sp/RoesslerD18} tags written memory with the depth in the
+    stack at which it was written, and enforces that it must be read from the
+    same depth. This allows violations of observable stack safety!
+    Say we have functions {\tt foo}, {\tt bar} and {\tt baz}, where {\tt foo}
+    calls {\tt bar}, and then after {\tt bar} returns {\tt foo} calls
+    {\tt baz}. Since {\tt bar} and {\tt baz} are at the same stack depth, if
+    {\tt bar} writes to {\tt foo}'s frame, {\tt baz} can then read the tainted
+    write and use it to create an observable difference that violates both
+    observable integrity and confidentiality without violating the policy.
+    This can be fixed by generating distinct tags per-activation.
+
+    The lazy policy can be repaired if, instead of tagging each function 
+    activation with its depth in the stack, we generate a fresh activation
+    identifier on each call, which prevents a program from exploiting stale
+    tags from previous activations. A related mechanism was explored in the
+    Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
+    associates a unique activation identifier to each function, which is
+    however shared by all activations of the same function in order to obtain
+    good cacheability of tag rules. Static Authorities comes closer to
+    observable stack safety, though the example above would still exist
+    if {\tt bar} and {\tt baz} were merged into a single function that tracked
+    whether it had been called previously and changed its behavior accordingly.
+
+  \paragraph{Relation to Stack Safety}
+
+\rb{Implied by the property we started with?}
+
+\rb{We can also test the lazy policy, even if less efficiently, and it should
+  enforce an inductive property, which may be more amenable to proving --- this
+  should be the connection to the following section.}
+
+\rb{Explain and compare what each policy enforces.}
 

### 2020-07-09 main.tex
wip

--- 2020-07-09 main.tex
@@
 
-\leo{Took a stab. Needs rewriting}%
-
-After formally defining stack safety, we show that our definition is
-enforceable using an existing tag-based enforcement
-mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
-conservative Depth Isolation policy, which not only enforces the
-our formal definition of stack safety, but also a much stronger
-property that requires integrity and confidentiality at every step of
-an execution trace. This property is provably (in Coq)\leo{How to
-  mention this?} stronger and stemmed out of the need to more
-efficiently test stack safety. However, at the same time the
-conservative enforcement mechanism is expensive to implement. In the
-same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different
-mechanism that is more efficient by virtue of being lazy in its
-enforcement. At a high level, instead of guaranteeing integrity and
-confidentiality when a callee returns, they check it when the caller
-actually makes use of (potentially compromised) data. Thus, they
-enforce a (also provably in Coq) weaker property that still protects
-from unsafe callees, but is much less intuitive.
+\bcp{This little paragraph is our entire explanation of the second and third
+contributions!  Needs to be expanded considerably...}%
+To provide evidence that our definitions capture the inituitive notion
+of stack safety, we show that they rule out a range of exploitable bad
+behaviors.\apt{``from the literature''??}  To assess whether they are
+implementable, we compare them against
+%proposed capability-based
+%enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
+%  should} and
+the existing tag-based enforcement mechanism of
+\citet{DBLP:conf/sp/RoesslerD18}.
+%\leo{Is this where we would rephrase this because we would only talk about it in related work?}
 
@@
   A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
-  This policy enforces our strict stack safety property, and also a (provably in Coq) stronger property that is easier to test.
+  This policy enforces our strict stack safety property, and also a (provably) stronger property that is easier to test.
   \apt{If we have no testing, do we even want to talk about the testing
@@
   but a modified version of it using unique per-activation tags enforces
-  an (also provably) weaker lazy variant of the property, which still prohibits
-  the same intuitively stack unsafe behaviors.
+  a (provably) weaker lazy variant of the property.
 %
@@
 policy state to the ordinary machine state\bcp{This may be surprising to
-  people who are not thinking about micropolicies.} and a partial policy
+  people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
 step function that
@@
 representative families of policies.
+
+We model ordinary program state and policy enforcement state separately to make it 
+easier to use well-established concepts from language-based security.  
+In particular, confidentiality has an elegant extensional characterization in terms of
+\emph{noninterference}: a program preserves the confidentiality of certain data if
+\emph{varying} that data would not change the program's observable behavior~\cite{??}.
+In applying this idea to a system that incorporates policy enforcement, only
+ordinary program data should be varied, whereas any associated policy enforcement
+state should be left alone; our model makes this distinction obvious.
+%That is, we compare the actual execution of the program with a \emph{hypothetical}
+%execution where data has been varied. 
+We also introduce a notion of observational integrity that relies on the idea of
+\emph{rolling back} program state; again, our model makes it easy to describe
+rollback of ordinary program data while leaving policy enforcement data unchanged.
 

### 2020-07-09 main.tex
typo
A pass on related work

--- 2020-07-09 main.tex
@@
 
-CHERI \rb{cite}
-
-\citet{Skorstengaard+19b} introduce a calling convention to preserve local state
-encapsulation and well-bracketed control flow by the use of local capabilities.
-Although the required hardware support is readily available in capability
-machines like CHERI, this technique incurs potentially prohibitive performance
-costs, because it requires the unused part of the stack to be cleared whenever a
-security boundary is crossed. In the discussion, ``while [the authors] claim
-that [their] calling convention enforces control-flow correctness, [they] do not
-prove a general theorem that shows this, because it is not clear what such a
-theorem should look like,'' noting that the correctness property enjoyed by
-their technique ``is not made very explicit.''
-
-StkTokens \citep{Skorstengaard+19} protects the stack by enforcing local state
-encapsulation and well-bracketed control flow. It does so by defining a calling
-convention that makes use of linear capabilities for stack and return pointers
-operating on a single shared stack, and by requiring that protected components
-avoid compromising their own security by following certain simple rules.
-Informally, local state encapsulation is defined by restricting accesses to the
-range of memory allocated to the current stack frame, and well-bracketed control
-flow by only allowing returns from the topmost frame to the immediately adjacent
-frame below. Formally, it improves on \citet{Skorstengaard+19b} by building
-those properties into the semantics of a capability machine with a built-in call
-stack and call and return instructions, which is proven fully abstract with
-respect to a more concrete capability machine that replaces those
-pseudo-instructions with their calling convention.
-
-\rb{As far as I can tell, while their notion of components and interfaces seems
-  standard, observable behavior appears to be limited to cotermination. The
-  previous paper makes explicit that (at least at that point) they are not
-  interested in externally observable side effects like console output or memory
-  access traces.}
-
-We are confident \rb{how strong a claim to make, with substantiation} that
-StkTokens satisfies the eager stack safety property, and even the stronger eager
-testing property \rb{refs}. The main limitation of StkTokens is its reliance on
-linear capabilities, as it is unclear now they could be added to practical
-capability machines, especially in terms of efficiency.
-%
-\rb{On the other hand, we know that these can be implemented as micropolicies,
-  cf. linear return capabilities in yannis-report (what about performance?)}
-
-\rb{There are quite a few parallels at various levels, although the definitions
-  are overall more implicit and not given formally.}
+%% CHERI \rb{cite}
+
+\paragraph{Formal Stack Safety on Capability Machines}
+%
+To our knowledge, the first and only line of work that has attempted to furnish
+a positive and formal characterization of the meaning of stack safety was
+inaugurated by \citet{Skorstengaard+19b}, who introduce a calling convention
+that uses local capabilities to preserve local state encapsulation and
+well-bracketed control flow, using a logical relation to reason about the stack
+safety of concrete programs. Although the required hardware support is readily
+available in capability machines like CHERI, this technique incurs potentially
+prohibitive performance costs, because it requires the entire unused part of the
+stack to be cleared whenever a security boundary is crossed. Their logical
+relation captures capability safety without ``externally observable side-effects
+(like console output or memory access traces).'' In the discussion, ``while [the
+authors] claim that [their] calling convention enforces control-flow
+correctness, [they] do not prove a general theorem that shows this, because it
+is not clear what such a theorem should look like,'' noting that the correctness
+property enjoyed by their technique ``is not made very explicit.''
+
+StkTokens \citep{Skorstengaard+19} is a natural continuation of that work. Like
+its predecessor, it aims to protect the stack by enforcing local state
+encapsulation and well-bracketed control flow. It does so by defining a new
+calling convention that makes use of linear capabilities for stack and return
+pointers. The convention operates on a single shared stack and requires that
+protected components avoid compromising their own security by following certain
+simple rules---like avoiding to leak their private capability seals. More
+precisely but still informally, local state encapsulation is defined as
+restricting accesses to the range of memory allocated to the current stack
+frame, and well-bracketed control flow as only allowing returns from the topmost
+frame to the immediately adjacent frame below. Formally, it improves on
+\citet{Skorstengaard+19b} by building those properties into the semantics of a
+capability machine with a built-in call stack and call and return instructions,
+which is proven fully abstract with respect to a more concrete capability
+machine that replaces those pseudo-instructions with their calling convention.
+Their proof of full abstraction uses a standard notion of components, which
+import and export functions through their interfaces; their model of
+observations is limited to cotermination.
+
+We are confident that the linear capability machines introduced by
+\citep{Skorstengaard+19}, in combination with the StkTokens calling convention,
+can be modeled in our framework and satisfy our definition of stack safety in
+\cref{sec:lse-and-wbcf}. Further, we expect them to satisfy the even stronger
+property developed in \cref{sec:enforcement}. Proving these conjectures would
+involve materializing their definitions of local state encapsulation and
+well-bracketed control flow, which exist only in enmeshed form as part of the
+semantics of the capability machines. The main practical limitation of StkTokens
+is its reliance on linear capabilities, as it is unclear now they could be added
+to practical capability machines, especially in terms of efficiency---previous
+work on micropolicies \citep{yannis-report} has shown how to use that framework
+to implement linear return capabilities.
+
+\paragraph{Protecting the Stack with Micropolicies}
+%
+\citet{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
+attacks against stack data are in scope (but not side channels or hardware
+attacks), and study the protection of stack data through three families of
+micropolicies that tag stack objects with a pair of frame and object identifiers
+used to validate accesses to the stack: Return Address Protection (which
+prevents an adversary from overwriting designated return addresses), Static
+Authorities (which only allows the code of a function to access the stack frames
+of its own dynamic instances), and Depth Isolation (described in
+\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
+temporal locality of stack memory and local call graphs, as well as information
+generated by the compilation toolchain, to strike various balances between
+precision and cacheability, evaluated through benchmarks that demonstrate
+limited performance overhead. In addition to the baseline ``eager'' policies,
+they propose a number of lazy optimizations, notably Lazy Tagging and Lazy
+Clearing, discussed in \cref{sec:lazy}, with associated improvements in
+performance.
 
@@
   We have presented a pure notion of stack safety that omits many complicating
-  factors common to real system. In particular, our model as presented exploits
+  factors common to real systems. In particular, our model as presented exploits
   the simplifying assumption that privilege never increases during a function

### 2020-07-09 main.tex
Abstract

--- 2020-07-09 main.tex
@@
 \begin{abstract}
-\bcp{Hey, we should have an abstract \ldots}
+  What is stack safety? While many mechanisms have been proposed to
+  protect against attacks that influence the call stack of a program,
+  and many others rely on a protected stack as an assumption to
+  provide higher level guarantees, a clear stack safety definition
+  is missing from the literature.
+%
+  In this paper we give such a formal definition of stack safety as a
+  security property.
+  %
+  We also explore how this formal definition can be enforced by
+  previously proposed security mechanisms and discuss provably
+  stronger and weaker variations of interest.
 \end{abstract}
@@
 policy state to the ordinary machine state\bcp{This may be surprising to
-  people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
+  people who are not thinking about micropolicies.} and a partial policy
 step function that
@@
 representative families of policies.
-
-We model ordinary program state and policy enforcement state separately to make it 
-easier to use well-established concepts from language-based security.  
-In particular, confidentiality has an elegant extensional characterization in terms of
-\emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable behavior~\cite{??}.
-In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, whereas any associated policy enforcement
-state should be left alone; our model makes this distinction obvious.
-%That is, we compare the actual execution of the program with a \emph{hypothetical}
-%execution where data has been varied. 
-We also introduce a notion of observational integrity that relies on the idea of
-\emph{rolling back} program state; again, our model makes it easy to describe
-rollback of ordinary program data while leaving policy enforcement data unchanged.
 

### 2020-07-09 main.tex
Abstract

--- 2020-07-09 main.tex
@@
 \begin{abstract}
-\bcp{Hey, we should have an abstract \ldots}
+What is stack safety? Numerous mechanisms have been proposed to guard
+against attacks that corrupt the call stacks of running programs, and many
+others rely on a protected stack as an assumption to provide higher level
+guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
+  paper, do we?}, a clear and precise definition of stack safety is missing
+from the literature.
+
+We propose a formal definition of stack safety phrased in the technical
+terminology of language-based security---as a combination of a
+confidentiality property (callees should not be able to read their caller's
+stack frames) and an integrity property (callees should not corrupt their
+caller's stack frames or control state).
+  %
+We show how these properties are enforced or approximated by a range of
+existing stack protection mechanisms\bcp{that is not exactly what we do,
+  though.  Let's (re)write the contributions first.} and discuss some useful
+stronger and weaker variants.  \bcp{should we mention stktokens here?}
 \end{abstract}

### 2020-07-09 main.tex
wip on 3

--- 2020-07-09 main.tex
@@
 \begin{abstract}
-  What is stack safety? While many mechanisms have been proposed to
-  protect against attacks that influence the call stack of a program,
-  and many others rely on a protected stack as an assumption to
-  provide higher level guarantees, a clear stack safety definition
-  is missing from the literature.
-%
-  In this paper we give such a formal definition of stack safety as a
-  security property.
-  %
-  We also explore how this formal definition can be enforced by
-  previously proposed security mechanisms and discuss provably
-  stronger and weaker variations of interest.
+\bcp{Hey, we should have an abstract \ldots}
 \end{abstract}
@@
 
-\apt{Need to clarify to what extent our properties depend on the
-particular compilation style, calling sequences, etc. illustrated here.}
-
-\section{Threat Model}
+%\apt{Need to clarify to what extent our properties depend on the
+%particular compilation style, calling sequences, etc. illustrated here.}
+
+\section{Threat Model and Assumptions}
 \label{sec:threat}
@@
 Stack safety is meant to protect a caller from a confused or malicious callee's
-attempts to access private data or highjack its control flow. Our
+attempts to access private data or hijack its control flow. Our
 attacker model reflects this intention by assuming that the attacker
-can run arbitrary code to try to access or overwrite its caller's stack data
+can run arbitrary machine code to try to access or overwrite its caller's stack data
 or disrupt its control state.  Hardware and timing attacks are out of
-scope. \leo{Anything else to add here?}
-\leo{Maybe a good place to discuss calling conventions and that they
-  are essentially enforced by the policy or sth?}
-\rb{Intuition that the callee can do anything in principle, but the machine will
-halt if it does not follow the calling convention, etc.?}
-
-\apt{Informal definition of policies as auxiliary state here.}
+scope.
+%\leo{Anything else to add here?}
+%\leo{Maybe a good place to discuss calling conventions and that they
+%  are essentially enforced by the policy or sth?}
+%\rb{Intuition that the callee can do anything in principle, but the machine will
+%halt if it does not follow the calling convention, etc.?}
+
+We assume we are given a machine language program, together with certain auxiliary declarations
+about its structure: the code footprint of each function and the locations of instructions
+representing calls, returns, and function entry points.
+Although this information would typically come from a compiler, we do not
+assume anything about the provenance of the machine
+code. We assume that the machine has a program counter $\PCname$
+and stack pointer $\rsp$ and can perform the usual
+loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
+
+To make our stack safety definitions precise, we make some simple assumptions about
+memory layout; our definitions could easily be changed to be parametric over
+these assumptions, at the cost of slightly greater complexity.  Memory 
+consists of an instruction region and a stack that grows upwards; for simplicity,
+we omit a heap. (Integrating a memory safety property for the heap in the style of
+\citet{AzevedoDeAmorim17} is a natural extension that we discuss
+in~\Cref{sec:future}.)  The stack has a conventional structure, with
+contiguously allocated per-function frames.
+The boundary between the frames of a caller and callee is given by the $\rsp$ register
+at the time of the call.  The stack can be used to
+pass arguments or return results in the usual way, by placing them
+at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
+definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
+
+Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
+which are not fully explicit in machine code.  We define a call to be
+any instruction declared as such in the program.
+If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
+  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
+of well-bracketed control flow relates this definition of return to the program's
+declared set of return points.)
+Our property definitions make no other detailed assumptions about call or return sequences, although 
+some of the enforcement mechanisms we consider do.
 
@@
 policy state to the ordinary machine state\bcp{This may be surprising to
-  people who are not thinking about micropolicies.} and a partial policy
+people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
 step function that 
 refines the machine step function by allowing or disallowing a step to be taken
-based on its auxiliary state. This model will be made more precise in
-\cref{sec:prelim}. It is flexible enough to capture any kind of mechanism used
+based on its auxiliary state.   Enforcement works by halting the program before
+it can perform an action that would violate the security property; our properties
+are naturally therefore \emph{termination insensitive.}
+This model will be made more precise in \cref{sec:prelim}.
+
+We model ordinary program state and policy enforcement state separately to make it 
+easier to use well-established concepts from language-based security.  
+In particular, confidentiality has an elegant extensional characterization in terms of
+\emph{noninterference}: a program preserves the confidentiality of certain data if
+\emph{varying} that data would not change the program's observable behavior~\citep{??}.
+In applying this idea to a system that incorporates policy enforcement, only
+ordinary program data should be varied, whereas any associated policy enforcement
+state should be left alone; our model makes this distinction obvious.
+%That is, we compare the actual execution of the program with a \emph{hypothetical}
+%execution where data has been varied. 
+We also introduce a notion of observational integrity that relies on the idea of
+\emph{rolling back} program state; again, our model makes it easy to describe
+rollback of ordinary program data while leaving policy enforcement data unchanged.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+This enforcement model is flexible enough to capture any kind of mechanism used
 for dynamic enforcement of properties of interest, including the following
@@
 
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
+\apt{rest of this section still needs rethink}
 \paragraph*{Null policy}

### 2020-07-09 main.tex
Clean up blessed sequences

--- 2020-07-09 main.tex
@@
 sequences. These are specific sequences of tagged instructions that are intended
-to appear at the beginning and end of each function:
-
+to appear at the beginning and end of each function. This is the header
+sequence:
+%
+\setcounter{pcctr}{1}
 \[
-  \begin{array}{c|c|c}
-    {\sw ~ \rsp ~ \rra ~ 1} & {\tagHa, \tagInstr} & \text{Store RA in stack} \\
-    {\addi ~ \rsp ~ \rsp ~ 2} & {\tagHb, \tagInstr} & \text{Increment SP by frame size}
+  \begin{array}{l|l|l|l}
+     & \mathit{Instruction} & \mathit{Tags} & \\
+    \hline
+    \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
+        {Store return address in stack}
+    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
+        {Increment stack pointer by frame size}
   \end{array}
 \]
-
+%
+And the return sequence:
+%
+\setcounter{pcctr}{1}
 \[
-  \begin{array}{c|c|c}
-    {\lw ~ \rra ~ \rsp ~ -1} & {\tagRa, \tagInstr} & \text{Load RA} \\
-    {\addi ~ \rsp ~ \rsp ~ -2} & {\tagRb, \tagInstr} & \text{Decrement SP by frame size} \\
-    {\jalr ~ \rra ~ \rra} & {\tagRc, \tagInstr} & \text{Jump back}
+  \begin{array}{l|l|l|l}
+     & \mathit{Instruction} & \mathit{Tags} & \\
+    \hline
+    \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
+        {Load return address}
+    \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
+        {Decrement stack pointer by frame size}
+    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
+        {Return to caller ($\jalr$)}
   \end{array}

### 2020-07-09 main.tex
Cleaned up 5.3 (WBCF)

--- 2020-07-09 main.tex
@@
 
-%\apt{Need to clarify to what extent our properties depend on the
-%particular compilation style, calling sequences, etc. illustrated here.}
-
-\section{Threat Model and Assumptions}
+\apt{Need to clarify to what extent our properties depend on the
+particular compilation style, calling sequences, etc. illustrated here.}
+
+\section{Threat Model}
 \label{sec:threat}
@@
 Stack safety is meant to protect a caller from a confused or malicious callee's
-attempts to access private data or hijack its control flow. Our
+attempts to access private data or highjack its control flow. Our
 attacker model reflects this intention by assuming that the attacker
-can run arbitrary machine code to try to access or overwrite its caller's stack data
+can run arbitrary code to try to access or overwrite its caller's stack data
 or disrupt its control state. Hardware and timing attacks are out of
-scope.
-%\leo{Anything else to add here?}
-%\leo{Maybe a good place to discuss calling conventions and that they
-%  are essentially enforced by the policy or sth?}
-%\rb{Intuition that the callee can do anything in principle, but the machine will
-%halt if it does not follow the calling convention, etc.?}
-
-We assume we are given a machine language program, together with certain auxiliary declarations
-about its structure: the code footprint of each function and the locations of instructions
-representing calls, returns, and function entry points.
-Although this information would typically come from a compiler, we do not
-assume anything about the provenance of the machine
-code. We assume that the machine has a program counter $\PCname$
-and stack pointer $\rsp$ and can perform the usual
-loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
-
-To make our stack safety definitions precise, we make some simple assumptions about
-memory layout; our definitions could easily be changed to be parametric over
-these assumptions, at the cost of slightly greater complexity.  Memory 
-consists of an instruction region and a stack that grows upwards; for simplicity,
-we omit a heap. (Integrating a memory safety property for the heap in the style of
-\citet{AzevedoDeAmorim17} is a natural extension that we discuss
-in~\Cref{sec:future}.)  The stack has a conventional structure, with
-contiguously allocated per-function frames.
-The boundary between the frames of a caller and callee is given by the $\rsp$ register
-at the time of the call.  The stack can be used to
-pass arguments or return results in the usual way, by placing them
-at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
-definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
-
-Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code.  We define a call to be
-any instruction declared as such in the program.
-If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
-  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
-of well-bracketed control flow relates this definition of return to the program's
-declared set of return points.)
-Our property definitions make no other detailed assumptions about call or return sequences, although 
-some of the enforcement mechanisms we consider do.
+scope. \leo{Anything else to add here?}
+\leo{Maybe a good place to discuss calling conventions and that they
+  are essentially enforced by the policy or sth?}
+\rb{Intuition that the callee can do anything in principle, but the machine will
+halt if it does not follow the calling convention, etc.?}
+
+\apt{Informal definition of policies as auxiliary state here.}
 
@@
 policy state to the ordinary machine state\bcp{This may be surprising to
-people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
+  people who are not thinking about micropolicies.} and a partial policy
 step function that
 refines the machine step function by allowing or disallowing a step to be taken
-based on its auxiliary state.   Enforcement works by halting the program before
-it can perform an action that would violate the security property; our properties
-are naturally therefore \emph{termination insensitive.}
-This model will be made more precise in \cref{sec:prelim}.
-
-We model ordinary program state and policy enforcement state separately to make it 
-easier to use well-established concepts from language-based security.  
-In particular, confidentiality has an elegant extensional characterization in terms of
-\emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable behavior~\citep{??}.
-In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, whereas any associated policy enforcement
-state should be left alone; our model makes this distinction obvious.
-%That is, we compare the actual execution of the program with a \emph{hypothetical}
-%execution where data has been varied. 
-We also introduce a notion of observational integrity that relies on the idea of
-\emph{rolling back} program state; again, our model makes it easy to describe
-rollback of ordinary program data while leaving policy enforcement data unchanged.
+based on its auxiliary state. This model will be made more precise in
+\cref{sec:prelim}. It is flexible enough to capture any kind of mechanism used
+for dynamic enforcement of properties of interest, including the following
+representative families of policies.
 
@@
 
-This enforcement model is flexible enough to capture any kind of mechanism used
-for dynamic enforcement of properties of interest, including the following
-representative families of policies.
-
-\apt{rest of this section still needs rethink}
 \paragraph*{Null policy}
@@
 %\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
-if, for any initial state $\mpstatename$ and any adjacent pair
-of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
-trace from $\mpstatename$ whose current instruction
-belongs to different functions according to the function map,
-$\mach_1$ is either a valid call or a valid return:
-
-\[ \codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname)) \Rightarrow
-\exists n. ~ \callmap(\mach_1(\PCname)) = n \vee
-\retmap(\mach_1) \]
+if, for any initial state \(\mach_0\) and any adjacent pair of states
+\(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
+in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\),
+if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
+then \(\mpstate{\mach_1}{\pol_1}\) is either a call \(\callmap(\mach_1(\PCname))\)
+or a return \(\retmap(\mach_1(\PCname))\).
 
@@
 function switch is not a valid call (or a return) according to the
-call map.
+call map. On the other hand we do permit functions to jump freely internally.
 
@@
 %\(\callmap\) and entry map \(\entmap\)
-if, for any initial state $\mpstatename$
+if, for any initial state \(\mach_0\)
 and any n-transition \((\mach_c,\pol_c)\) and
 \((\mach_e, \pol_e)\) in the induced trace from
-$\mpstatename$, $\mach_e$ is a member of the entry map:
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), $\mach_e$ is a valid entry:
 \[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]
@@
 The changed program described just above also violates entry integrity,
-because the target (instruction 42) of the new $\jalr$ is not an established entry point
-according to the entry map.
-
-\leo{Note something about not enforcing control flow integrity
-  with allowing a call from {\tt main} to {\tt h}?}
+because the target (instruction 42) of the new $\jalr$ is not an established
+entry point according to the entry map. Our properties to not enforce which
+call points jump to which entry points, however; that is left for a stronger
+notion of control flow integrity.
 
@@
 matching a call $\mach_c$ should satisfy $\ret{\mach_c}$.
-Observe that, given an MP-trace $\mpstate{\mach_c}{\pol}\MPT$, if $\mach_c$ is a call state then
-the matching return will be the first \emph{unmatched} return in $\MPT$. So
-we begin by defining an auxiliary ``unmatched'' relation between two MP-traces.
+Observe that, given an MP-trace $\mpstate{\mach_c}{\pol}\MPT$, if $\mach_c$ is
+a call state then the matching return will be the first \emph{unmatched} return
+in $\MPT$. So we begin by defining an auxiliary ``unmatched'' relation between
+two traces.
 
@@
 \definition%{Return Integrity}
-A system enjoys {\em return integrity}
-% with respect to a call map \(\callmap\) and return map \(\retmap\)
-if, for any initial state pair
-$\mpstatename$ and any state pair \((\mach_c, \pol_c)\)
-corresponding to a call in its induced trace, if
-\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
-is the first unmatched return such that for some $\pol_r$
-\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
-then $\mach_r$ corresponds to a return state for $\mach_c$:
-\[\ret{\mach_c}\ \mach_r\]
-
-\[
-\exists p_r.\Unmatched ~ \MPT ~ (m_r,p_r)\MPT' \Rightarrow \ret{\mach_c}\ \mach_r.
-\]
-\apt{I've tried to tidy this, but I don't think it corresponds to the Coq.
-  Shouldn't we apply $\Unmatched$ to the tail of $\MPT$?
-  And what about the no-unmatched-return-at-top-level clause?}
+A system enjoys {\em return integrity} if all of its returns match prior calls
+and return to the appropriate location. For any initial state \(\mach_0\), let
+\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pinit(\mach_0)}\).
+\(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
+may not hold for any \(\MPT_{tail}\). For any state pair \((\mach_c, \pol_c)\)
+corresponding to a call \(\MPT\), if 
+\(\mpstatename = \mpostate{\mach_r}{\pol_r}{\obs_r}\)
+is the first unmatched return such that
+\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{\MPT_{tail}} = \mpstatename\),
+then $\mach_r$ must correspond to a return state for $\mach_c$:
+\(\ret{\mach_c}\ \mach_r\)
 

### 2020-07-09 main.tex
done with pass on 3

--- 2020-07-09 main.tex
@@
 policy state to the ordinary machine state\bcp{This may be surprising to
-  people who are not thinking about micropolicies.} and a partial policy
+people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
 step function that 
 refines the machine step function by allowing or disallowing a step to be taken
-based on its auxiliary state. This model will be made more precise in
-\cref{sec:prelim}. It is flexible enough to capture any kind of mechanism used
+based on its auxiliary state.   Enforcement works by halting the program before
+it can perform an action that would violate the security property; our properties
+are naturally therefore \emph{termination insensitive.}
+This model will be made more precise in \cref{sec:prelim}.
+
+We model ordinary program state and policy enforcement state separately to make it 
+easier to use well-established concepts from language-based security.  
+In particular, confidentiality has an elegant extensional characterization in terms of
+\emph{noninterference}: a program preserves the confidentiality of certain data if
+\emph{varying} that data would not change the program's observable behavior~\citep{??}.
+In applying this idea to a system that incorporates policy enforcement, only
+ordinary program data should be varied, whereas any associated policy enforcement
+state should be left alone; our model makes this distinction obvious.
+%That is, we compare the actual execution of the program with a \emph{hypothetical}
+%execution where data has been varied. 
+We also introduce a notion of observational integrity that relies on the idea of
+\emph{rolling back} program state; again, our model makes it easy to describe
+rollback of ordinary program data while leaving policy enforcement data unchanged.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+This enforcement model is flexible enough to capture any kind of mechanism used
 for dynamic enforcement of properties of interest, including the following
 representative families of policies.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-\paragraph*{Null policy}
-%
-The simplest possible policy is the neutral policy that carries no information
-and does nothing. Its set of policy states is the unit type, and its policy step
-function is the constant unit function, thereby allowing all calls to the machine
-step function. Its initialization function is defined on all states. This
-corresponds to the lifting of our machine model to the policy setting. Machines
-that fulfill a property by construction will also fulfill that property with the
-null policy.
+\apt{I don't really buy this, because we really depend on the ability to define
+  variation and rollback. I think we should back off and say: this
+  works well for micropolicies, and we conjecture that we can make it work
+  for code-altering appraoches (pure software or capability).}
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
+
+\paragraph*{Micropolicies}
+%
+Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
+reference monitoring mechanism, which have been applied to stack safety by 
+\citet{DBLP:conf/sp/RoesslerD18}.
+Here the policy state consists of metadata tags (e.g., identifying stack frames)
+attached to each value in memory or registers, and the policy step function
+checks that each machine operation obeys a set of rules on tags
+(e.g., that the current $\PCname$ tag matches the tag on the stack
+location being accessed), and halts the machine if not.  Distinguishing
+ordinary machine state and policy state is very natural in this setting.
+%% Here the machine state is extended with tags paired with all components of the
+%% system, and a monitor checks for each operation that the tags on the operands
+%% obey a set of rules. One could treat tags directly as part of the machine
+%% state, and build the rules into the step function. But local state
+%% encapsulation properties as we frame them are not intuitive in their
+%% interaction with tags or other additional enforcement state, so it is helpful
+%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
+%% a software fail state.
+%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+%%   policy? }
+
+\paragraph*{Capability machines}
+%
+%Other policies are implemented in hardware or built on hardware mechanisms.
+Capability machines like CHERI \citep{Woodruff+14} extend conventional
+architectures to support efficient and fine-grained control over memory
+accesses, which can be used to implement security policies, including 
+the stack protection policies of
+\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
+(ordinary data) with base and bounds information (policy data),
+and their use requires rewriting source code (e.g. to use capabilities in place
+of ordinary pointers), so separating ordinary machine state from policy
+state is delicate. \apt{But we conjecture it can be done...?}
+
+%% Mapping the underlying protection
+%% mechanisms to policy states and step function would be a delicate operation, as
+%% those are tightly coupled to the hardware design. However, separating the
+%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
+%% would encompass details like the abstract call stack, the capability sealing
+%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
+%% the machine would be relatively straightforward.  \bcp{Not sure that any of
+%%   this is comprehensible... :-(}
 
@@
 Some policies are implemented in software, by modifying code in the compiler to
-(e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
-canaries \citep{Cowan+98}. These approaches are represented as policies that
-accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
-being produced by the compiler.) The set of policy states is again the unit type
-and the step function constant. Property enforcement comes from the fact that
-only ill-formed states can induce a trace that violates the property.
+(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
+canaries~\citep{Cowan+98}. \apt{not sure what to say here}
+%These approaches are represented as policies that
+%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
+%being produced by the compiler.) The set of policy states is again the unit type
+%and the step function constant. Property enforcement comes from the fact that
+%only ill-formed states can induce a trace that violates the property.
 
@@
 
-\paragraph*{Capability machines}
-%
-Other policies are implemented in hardware or built on hardware mechanisms.
-Capability machines like CHERI \citep{Woodruff+14} extend conventional
-architectures to support efficient and fine-grained control over memory
-accesses, which can be used to implement security policies involving, for
-example, compartmentalization or the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}. Mapping the underlying protection
-mechanisms to policy states and step function would be a delicate operation, as
-those are tightly coupled to the hardware design. However, separating the
-enforcement aspects of a concrete policy (such as StkTokens, where policy state
-would encompass details like the abstract call stack, the capability sealing
-scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-the machine would be relatively straightforward.  \bcp{Not sure that any of
-  this is comprehensible... :-(}
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
-
-\paragraph*{Micropolicies}
-%
-Micropolicies are another major family of policies \citep{pump_oakland2015},
-which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
-Here the machine state is extended with tags paired with all components of the
-system, and a monitor checks for each operation that the tags on the operands
-obey a set of rules. One could treat tags directly as part of the machine
-state, and build the rules into the step function. But local state
-encapsulation properties as we frame them are not intuitive in their
-interaction with tags or other additional enforcement state, so it is helpful
-to separate tags for clarity. This also distinguishes a hardware fail-stop from
-a software fail state.
-\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-  policy? }
+%% \Paragraph*{Null policy}
+%% %
+%% The simplest possible policy is the neutral policy that carries no information
+%% and does nothing. Its set of policy states is the unit type, and its policy step
+%% function is the constant unit function, thereby allowing all calls to the machine
+%% step function. Its initialization function is defined on all states. This
+%% corresponds to the lifting of our machine model to the policy setting. Machines
+%% that fulfill a property by construction will also fulfill that property with the
+%% null policy.
 
@@
 sequences. These are specific sequences of tagged instructions that are intended
-to appear at the beginning and end of each function. This is the header
-sequence:
-%
-\setcounter{pcctr}{1}
+to appear at the beginning and end of each function:
+
 \[
-  \begin{array}{l|l|l|l}
-     & \mathit{Instruction} & \mathit{Tags} & \\
-    \hline
-    \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
-        {Store return address in stack}
-    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
-        {Increment stack pointer by frame size}
+  \begin{array}{c|c|c}
+    {\sw ~ \rsp ~ \rra ~ 1} & {\tagHa, \tagInstr} & \text{Store RA in stack} \\
+    {\addi ~ \rsp ~ \rsp ~ 2} & {\tagHb, \tagInstr} & \text{Increment SP by frame size}
   \end{array}
   \]
-%
-And the return sequence:
-%
-\setcounter{pcctr}{1}
+
 \[
-  \begin{array}{l|l|l|l}
-     & \mathit{Instruction} & \mathit{Tags} & \\
-    \hline
-    \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
-        {Load return address}
-    \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
-        {Decrement stack pointer by frame size}
-    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
-        {Return to caller ($\jalr$)}
+  \begin{array}{c|c|c}
+    {\lw ~ \rra ~ \rsp ~ -1} & {\tagRa, \tagInstr} & \text{Load RA} \\
+    {\addi ~ \rsp ~ \rsp ~ -2} & {\tagRb, \tagInstr} & \text{Decrement SP by frame size} \\
+    {\jalr ~ \rra ~ \rra} & {\tagRc, \tagInstr} & \text{Jump back}
   \end{array}

### 2020-07-09 main.tex
Cleanup on section 6

@@
 
-\rb{Wondering if we should use sub-sections here, too.}
-
 In this section we are going to see how an existing enforcement mechanism,
@@
 allows fine-grained manipulation of metadata tags to encode so-called
-\emph{micropolicies}~\citep{pump_oakland15}.
+\emph{micropolicies}~\citep{pump_oakland2015}.
 
@@
 Efficient execution of these micropolicies relies on hardware implementations,
-such as the PUMP architecture~\citep{pump_oakland15}. This hardware incorporates
-a monitor rule cache to allow quick retrieval of transfer function outputs for
-mapped inputs. Cache misses are detected and trapped to software handlers that
-use the software rules of the policy to compute the relevant parts of the
-transfer function. Previous work has shown that a wide range of micropolicies
-can be defined and enforced efficiently in these machines
-\citep{pump_oakland2015}.
-%
-\rb{Maybe go into more detail, examples later\ldots or try to segue into stack
-policies}
+such as the PUMP architecture. This hardware incorporates a monitor rule cache
+to allow quick retrieval of transfer function outputs for mapped inputs. Cache
+misses are detected and trapped to software handlers that use the software rules
+of the policy to compute the relevant parts of the transfer function.
+\citeauthor{pump_oakland2015} have shown that a wide range of micropolicies can
+be defined and enforced efficiently in these machines.
+%% %
+%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
+%% policies}
 
@@
 The first micropolicy we present to enforce the definition of stack
-safety developed in \cref{sec:lse-and-wbcf} is \leo{a small variation
-  of?}  the Depth Isolation policy presented by
+safety developed in \cref{sec:lse-and-wbcf} is a small variation
+on the Depth Isolation policy presented by
 \citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
@@
 separate object inside that frame, a more finely grained access
-control that we do not need to consider here), and registers holding
-stack pointers with matching access permission information. Beyond
+control that we do not need to consider here), and tags registers containing
+stack pointers with access permission information. Beyond
 those capabilities, the compiler generates designated sequences of
 instructions with special tags that grant them permission to manage
-the stack, say during calls and returns. \rb{Conceivably on alloc
-  operations, etc., which we have not yet mentioned.}
+the stack, say during calls and returns.
+%% \rb{Conceivably on alloc
+%%   operations, etc., which we have not yet mentioned.}
 
 Our micropolicy works as follows. We assume for ease of exposition that no
-arguments are passed on the stack \rb{Come back to this later?}. Initially, the
-entire stack is tagged as unused with an $\tagNoDepth$ tag \rb{actually, the
-``initial frame'' should be tagged with its proper depth}, and the {\PCname}
-carries in its tag the depth of the current activation at all times, starting
-with a $\tagPCDepth{0}$ tag. The current activation depth needs to change
-exactly at those points in the program where the contour of the stack safety
-property changes:
+arguments are passed on the stack.
+%% \rb{Come back to this later?}.
+Initially, the entire stack is tagged as unused with an $\tagNoDepth$ tag
+\rb{actually, the ``initial frame'' should be tagged with its proper depth}, and
+the {\PCname} carries in its tag the depth of the current activation at all
+times, starting with a $\tagPCDepth{0}$ tag. The current activation depth needs
+to change exactly at those points in the program where the contour of the stack
+safety property changes:
 
@@
 
-\rb{Positioning of the above sequences, relation to running example. The tagging
-  of stack frames would be considered part of these sequences.}
+%% \rb{Positioning of the above sequences, relation to running example. The tagging
+%%   of stack frames would be considered part of these sequences.}
 
@@
 which are allowed to operate only on the corresponding stack frame,
-tagged $\tagStackDepth{n}$.  \rb{Note that we have also considered and
-  implemented more flexible schemes based on allowing operations above
-  the current activation depth. How do we put these together?}  \rb{Do
-  we want to discuss variations involving frame pointers or more
-  implicit return addresses?}  \rb{Where to talk about the role of the
-  compiler?}
+tagged $\tagStackDepth{n}$.
+\rb{Note that we have also considered and implemented more flexible schemes
+  based on allowing operations above the current activation depth. How do we put
+  these together?}
+\rb{Do we want to discuss variations involving frame pointers or more implicit
+  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
+\rb{Where to talk about the role of the compiler?}
 
@@
 
-\subsection{A Stronger Stack Safety Property}
-
-\leo{Maybe lockstep?}
+\subsection{A Stronger Lockstep Stack Safety Property}
+
+%% \leo{Maybe lockstep?}
 
@@
 We start by defining the integrity and confidentiality tests, as well
-as the variant stack, before giving the full definition of strong
+as the variant stack, before giving the full definition of lockstep
 stack safety.
@@
 %
-First, strong safety will check that the basic tests above hold for every
+First, lockstep stack safety will check that the basic tests above hold for every
 machine in the variant stack.  Given a machine that steps $(\mach,
@@
 
-We are now ready to define strong stack safety coinductively.
+We are now ready to define lockstep stack safety coinductively.
 
@@
 with respect to its contour, we say that an MP-trace induced from
-$(\mach, \pol)$ satisfies the \emph{trace strong stack safety}, and write
+$(\mach, \pol)$ satisfies the \emph{trace lockstep stack safety}, and write
 $\testProp{}{\MPT}{\vs}$, if:
@@
   \[\testVS{(\mach, \pol)}{\vs},\]
-  and trace strong stack safety must be coinductively satisfied for
+  and trace lockstep stack safety must be coinductively satisfied for
   the remainder of the trace and the variant stack $vs'$ that
@@
   If the machine takes a \emph{call step} and $\contour$ is the contour at
-  the call \rb{Corrected the step function here, or should it be
-  $\hookrightarrow$ and give the rest of the induced trace?}
+  the call
+  %% \rb{Corrected the step function here, or should it be
+  %% $\hookrightarrow$ and give the rest of the induced trace?}
   \[(\mach, \pol) \stepstopol{\obs} (\mach', \pol') \wedge \callmap(\mach) = n,\]
@@
 %
-  and trace strong stack safety must be coinductively satisfied
+  and trace lockstep stack safety must be coinductively satisfied
   for the remainder of the trace and for every variant stack
@@
   %
-  and trace strong stack safety must be coinductively satisfied for
+  and trace lockstep stack safety must be coinductively satisfied for
   the remainder of the trace and for the tail of the varant stack
@@
 %
-A system satisfies {\em strong stack safety} if for every initial machine $\mach$,
-the trace induced from $(\mach, \pinit{\mach})$ satisfies trace strong stack safety with
+A system satisfies {\em lockstep stack safety} if for every initial machine $\mach$,
+the trace induced from $(\mach, \pinit{\mach})$ satisfies trace lockstep stack safety with
 respect to the empty stack.
@@
 \theorem
-Strong stack safety implies stack safety.
+Lockstep stack safety implies stack safety.
 
 \proof
-By coinduction on trace strong stack safety.
+By coinduction on trace lockstep stack safety.
 \leo{3000 lines Coq?}

### 2020-07-09 main.tex
Pass on 7

--- 2020-07-09 main.tex
@@
 
-This enforcement model is flexible enough to capture any kind of mechanism used
-for dynamic enforcement of properties of interest, including the following
-representative families of policies.
-\apt{I don't really buy this, because we really depend on the ability to define
-  variation and rollback. I think we should back off and say: this
-  works well for micropolicies, and we conjecture that we can make it work
-  for code-altering appraoches (pure software or capability).}
+\paragraph*{Null policy}
+%
+The simplest possible policy is the neutral policy that carries no information
+and does nothing. Its set of policy states is the unit type, and its policy step
+function is the constant unit function, thereby allowing all calls to the machine
+step function. Its initialization function is defined on all states. This
+corresponds to the lifting of our machine model to the policy setting. Machines
+that fulfill a property by construction will also fulfill that property with the
+null policy.
+
+\paragraph*{Software-only policies}
+%
+Some policies are implemented in software, by modifying code in the compiler to
+(e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
+canaries \citep{Cowan+98}. These approaches are represented as policies that
+accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
+being produced by the compiler.) The set of policy states is again the unit type
+and the step function constant. Property enforcement comes from the fact that
+only ill-formed states can induce a trace that violates the property.
+
+\rb{TODO In a way, in these schemes the policy state is built in as part of the
+  program state and transparent to the hardware, although it could be simulated
+  by an overlay where enforcement is represented at the policy level.}
+
+\paragraph*{Capability machines}
+%
+Other policies are implemented in hardware or built on hardware mechanisms.
+Capability machines like CHERI \citep{Woodruff+14} extend conventional
+architectures to support efficient and fine-grained control over memory
+accesses, which can be used to implement security policies involving, for
+example, compartmentalization or the stack protection policies of
+\citet{Skorstengaard+19,Skorstengaard+19b}. Mapping the underlying protection
+mechanisms to policy states and step function would be a delicate operation, as
+those are tightly coupled to the hardware design. However, separating the
+enforcement aspects of a concrete policy (such as StkTokens, where policy state
+would encompass details like the abstract call stack, the capability sealing
+scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
+the machine would be relatively straightforward.  \bcp{Not sure that any of
+  this is comprehensible... :-(}
 
@@
 %
-Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
-reference monitoring mechanism, which have been applied to stack safety by 
-\citet{DBLP:conf/sp/RoesslerD18}.
-Here the policy state consists of metadata tags (e.g., identifying stack frames)
-attached to each value in memory or registers, and the policy step function
-checks that each machine operation obeys a set of rules on tags
-(e.g., that the current $\PCname$ tag matches the tag on the stack
-location being accessed), and halts the machine if not.  Distinguishing
-ordinary machine state and policy state is very natural in this setting.
-%% Here the machine state is extended with tags paired with all components of the
-%% system, and a monitor checks for each operation that the tags on the operands
-%% obey a set of rules. One could treat tags directly as part of the machine
-%% state, and build the rules into the step function. But local state
-%% encapsulation properties as we frame them are not intuitive in their
-%% interaction with tags or other additional enforcement state, so it is helpful
-%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
-%% a software fail state.
-%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-%%   policy? }
-
-\paragraph*{Capability machines}
-%
-%Other policies are implemented in hardware or built on hardware mechanisms.
-Capability machines like CHERI \citep{Woodruff+14} extend conventional
-architectures to support efficient and fine-grained control over memory
-accesses, which can be used to implement security policies, including 
-the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
-(ordinary data) with base and bounds information (policy data),
-and their use requires rewriting source code (e.g. to use capabilities in place
-of ordinary pointers), so separating ordinary machine state from policy
-state is delicate. \apt{But we conjecture it can be done...?}
-
-%% Mapping the underlying protection
-%% mechanisms to policy states and step function would be a delicate operation, as
-%% those are tightly coupled to the hardware design. However, separating the
-%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
-%% would encompass details like the abstract call stack, the capability sealing
-%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-%% the machine would be relatively straightforward.  \bcp{Not sure that any of
-%%   this is comprehensible... :-(}
-
-\paragraph*{Software-only policies}
-%
-Some policies are implemented in software, by modifying code in the compiler to
-(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
-canaries~\citep{Cowan+98}. \apt{not sure what to say here}
-%These approaches are represented as policies that
-%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
-%being produced by the compiler.) The set of policy states is again the unit type
-%and the step function constant. Property enforcement comes from the fact that
-%only ill-formed states can induce a trace that violates the property.
-
-\rb{TODO In a way, in these schemes the policy state is built in as part of the
-  program state and transparent to the hardware, although it could be simulated
-  by an overlay where enforcement is represented at the policy level.}
-
-%% \Paragraph*{Null policy}
-%% %
-%% The simplest possible policy is the neutral policy that carries no information
-%% and does nothing. Its set of policy states is the unit type, and its policy step
-%% function is the constant unit function, thereby allowing all calls to the machine
-%% step function. Its initialization function is defined on all states. This
-%% corresponds to the lifting of our machine model to the policy setting. Machines
-%% that fulfill a property by construction will also fulfill that property with the
-%% null policy.
+Micropolicies are another major family of policies \citep{pump_oakland2015},
+which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
+Here the machine state is extended with tags paired with all components of the
+system, and a monitor checks for each operation that the tags on the operands
+obey a set of rules. One could treat tags directly as part of the machine
+state, and build the rules into the step function. But local state
+encapsulation properties as we frame them are not intuitive in their
+interaction with tags or other additional enforcement state, so it is helpful
+to separate tags for clarity. This also distinguishes a hardware fail-stop from
+a software fail state.
+\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+  policy? }
 
@@
 
-\paragraph*{Eager vs Lazy Policies}
-
-The conservative policy we just described, while testable and enforceable, is
-too slow to be of practical use. What we want, instead, is a somewhat more
-permissive policy that moreover can be efficiently implemented. Most of the
-performance overhead incurred by this kind of \emph{eager policy} steps from the
-need to set stack activation tags as a frame is created and reset them as it is
-destructed, as \citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
-mitigate those costs, they propose optimizations like Lazy Tagging and Lazy
-Clearing that relax and defer some of the checks of an eager policy, being more
-permissive in their allowed operations and catching violations if and when their
-effect becomes manifest.
-
-These \emph{lazy policies} admit more efficient implementations, but they
-deliberately allow violations of stack integrity. So it becomes important to
-characterize what alternative property these policies ought to implement.
-
-\sna{moved a lot of discussion to after property, along with Rob's
-     comments on it}
-
 \section{Stack Safety Properties for Lazy Enforcement}
@@
 
-  A lazy policy permits some violations of integrity or confidentiality,
-  justified because a truly harmful violation will be caught later.
-  The natural question is, how do we characterize the protections provided
-  by these policies once stack safety is broken? How does a harmful violation
-  that must be caught later differ from a harmless one? The properties in this
-  section use the observation model to draw this distinction.
+The conservative policy described in the previous section, while
+testable and enforceable, is rather slow to be of practical use. What
+we want, instead, is a somewhat more permissive policy that can also
+be efficiently implemented. Most of the performance overhead incurred
+stems from the need to set stack activation tags as a frame is created
+and reset them as it is destructed, as
+\citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
+mitigate those costs, they propose optimizations like Lazy Tagging and
+Lazy Clearing that relax and defer some of the checks of an eager
+policy, being more permissive in their allowed operations and catching
+violations if and when their effect becomes manifest.
+%
+These \emph{lazy policies} admit more efficient implementations, but
+they deliberately allow violations of stack integrity temporarily,
+with the checks deferred until the point a violation truly becomes
+harmful. The natural question then is, how do we characterize the
+protections provided by these policies once stack safety is broken?
+How does a harmful violation that must be caught later differ from a
+harmless one? The properties in this section rely on our notion of
+observations to draw this distinction.
 
@@
 \paragraph{Rolling Back to Idealized State}
+
 Observable integrity weakens eager integrity by allowing a returned
 state to differ from its call state at high integrity components,
-      provided that the differences do not change the observable behavior
-      of the rest of the program. It is not immediately clear what constitutes
-      changing behavior; we construct an idealized return state that contains
-      only the changes that the caller was allowed to make, and use the trace
-      from that state as a baseline against which the real trace will be
-      compared.
+provided that the differences do not change the observable behavior of
+the rest of the program. It is not immediately clear what constitutes
+changing behavior: we construct an idealized return state that
+contains only the changes that the caller was allowed to make, and use
+the trace from that state as a baseline against which the real trace
+will be compared.
 
@@
         \propdef{ObsTraceInt} ~ \contour ~ \MPT \triangleq
-        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \\
+        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
         & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
@@
 
-      This requires restoring varied components to their original values
-      in the trace from the variant state, since after return they will
-      once again be legitimately accessible. To return to the example in
-      \cref{fig:confex}, the varied final state \(\mach_r'\) still has
-      -5 as a return address for {\tt g}, so the trace after that return will
-      behave oddly in a way that doesn't reflect a confidentiality violation.
-      So we cannot immediately compare traces from \(\mach_r\) and
-      \(\mach_r'\). We must restore the varied values while preserving
-      differences that have occurred during execution.
+      This requires restoring varied components to their original
+      values in the trace from the variant state, since after return
+      they will once again be legitimately accessible. To return to
+      the example in \cref{fig:confex}, the varied final state
+      \(\mach_r'\) still has -5 as a return address for {\tt g}, so
+      the trace after that return will behave oddly in a way that
+      doesn't reflect a confidentiality violation.  That means we
+      cannot immediately compare traces from \(\mach_r\) and
+      \(\mach_r'\), but rather that we must restore the varied values
+      while preserving differences that have occurred during
+      execution.
 
@@
 
+      \leo{This is starting to be too spread out as a definition. Pack a bit?}
       \[\begin{split}
@@
         & \last(\machT) = (\mach_r',\obs_r') \Rightarrow \\
-        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \\
+        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \Rightarrow\\
         & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
@@
       confidentiality. Let \(\callmap\) be a callmap; for each initial
-      state and \(n\)-transition consisting of \(\mpstate{\mach_c}{\pol_c}\)
-      and \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
+      state and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
+      to \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
       \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
@@
 
+      \theorem Stack safety implies observable stack safety.
+
+      \proof \leo{TODO: Whatever we say about the previous proof
+        we have to say similar things here}
+
     \paragraph{Connection to Policies}
@@
     stack at which it was written, and enforces that it must be read from the
-    same depth. This allows violations of observable stack safety!
+    same depth. However, this allows violations of observable stack safety!
     Say we have functions {\tt foo}, {\tt bar} and {\tt baz}, where {\tt foo}
@@
     observable integrity and confidentiality without violating the policy.
-    This can be fixed by generating distinct tags per-activation.
 
@@
     whether it had been called previously and changed its behavior accordingly.
-
-  \paragraph{Relation to Stack Safety}
-
-\rb{Implied by the property we started with?}
-
-\rb{We can also test the lazy policy, even if less efficiently, and it should
-  enforce an inductive property, which may be more amenable to proving --- this
-  should be the connection to the following section.}
-
-\rb{Explain and compare what each policy enforces.}
 

### 2020-07-09 main.tex
Clean up keywords
Reorg 3-2
\iflater
Massage 2+3

@@
 %% comma separated list
-\keywords{Stack Safety, Micropolicies, Random Testing\bcp{Really??}, RISC-V}  %% \keywords are mandatory in final camera-ready submission
+\keywords{Stack Safety, Micropolicies}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
+\section{Threat Model and Assumptions}
+\label{sec:threat}
+
+\leo{Needs high level signposting, transitions}
+
+Stack safety is meant to protect a caller from a confused or malicious callee's
+attempts to access private data or hijack its control flow. Our
+attacker model reflects this intention by assuming that the attacker
+can run arbitrary machine code to try to access or overwrite its caller's stack data
+or disrupt its control state.  Hardware and timing attacks are out of
+scope.
+%\leo{Anything else to add here?}
+%\leo{Maybe a good place to discuss calling conventions and that they
+%  are essentially enforced by the policy or sth?}
+%\rb{Intuition that the callee can do anything in principle, but the machine will
+%halt if it does not follow the calling convention, etc.?}
+
+\newcommand*{\rsp}{\textsc{sp}}
+
+We assume we are given a machine language program, together with certain auxiliary declarations
+about its structure: the code footprint of each function and the locations of instructions
+representing calls, returns, and function entry points.
+Although this information would typically come from a compiler, we do not
+assume anything about the provenance of the machine
+code. We assume that the machine has a program counter $\PCname$
+and stack pointer $\rsp$ and can perform the usual
+loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
+
+To make our stack safety definitions precise, we make some simple assumptions about
+memory layout; our definitions could easily be changed to be parametric over
+these assumptions, at the cost of slightly greater complexity.  Memory 
+consists of an instruction region and a stack that grows upwards; for simplicity,
+we omit a heap. (Integrating a memory safety property for the heap in the style of
+\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
+in~\Cref{sec:related}.)  The stack has a conventional structure, with
+contiguously allocated per-function frames.
+The boundary between the frames of a caller and callee is given by the $\rsp$ register
+at the time of the call.  The stack can be used to
+pass arguments or return results in the usual way, by placing them
+at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
+definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
+
+Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
+which are not fully explicit in machine code.  We define a call to be
+any instruction declared as such in the program.
+If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
+  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
+of well-bracketed control flow relates this definition of return to the program's
+declared set of return points.)
+Our property definitions make no other detailed assumptions about call or return sequences, although 
+some of the enforcement mechanisms we consider do.
+
+Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
+policy state to the ordinary machine state
+%\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
+and a partial policy step function that refines the machine step
+function by allowing or disallowing a step to be taken based on its
+auxiliary state.  Enforcement works by halting the program before it
+can perform an action that would violate the security property; our
+properties are naturally therefore \emph{termination insensitive.}
+This model will be made more precise in \cref{sec:prelim}.
+
+We model ordinary program state and policy enforcement state separately to make it 
+easier to use well-established concepts from language-based security.  
+In particular, confidentiality has an elegant extensional characterization in terms of
+\emph{noninterference}: a program preserves the confidentiality of certain data if
+\emph{varying} that data would not change the program's observable behavior~\citep{??}.
+In applying this idea to a system that incorporates policy enforcement, only
+ordinary program data should be varied, whereas any associated policy enforcement
+state should be left alone; our model makes this distinction obvious.
+%That is, we compare the actual execution of the program with a \emph{hypothetical}
+%execution where data has been varied. 
+We also introduce a notion of observational integrity that relies on the idea of
+\emph{rolling back} program state; again, our model makes it easy to describe
+rollback of ordinary program data while leaving policy enforcement data unchanged.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+This enforcement model is flexible enough to capture
+micropolicies~\citep{pump_oakland2015}, and we believe that it can
+also capture code-altering approaches (such as software-based
+or capability-based techniques).
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
+
+\paragraph*{Micropolicies}
+%
+Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
+reference monitoring mechanism, which have been applied to stack safety by 
+\citet{DBLP:conf/sp/RoesslerD18}.
+Here the policy state consists of metadata tags (e.g., identifying stack frames)
+attached to each value in memory or registers, and the policy step function
+checks that each machine operation obeys a set of rules on tags
+(e.g., that the current $\PCname$ tag matches the tag on the stack
+location being accessed), and halts the machine if not.  Distinguishing
+ordinary machine state and policy state is very natural in this setting.
+%% Here the machine state is extended with tags paired with all components of the
+%% system, and a monitor checks for each operation that the tags on the operands
+%% obey a set of rules. One could treat tags directly as part of the machine
+%% state, and build the rules into the step function. But local state
+%% encapsulation properties as we frame them are not intuitive in their
+%% interaction with tags or other additional enforcement state, so it is helpful
+%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
+%% a software fail state.
+%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+%%   policy? }
+
+\paragraph*{Capability machines}
+%
+%Other policies are implemented in hardware or built on hardware mechanisms.
+Capability machines like CHERI \citep{Woodruff+14} extend conventional
+architectures to support efficient and fine-grained control over memory
+accesses, which can be used to implement security policies, including 
+the stack protection policies of
+\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
+(ordinary data) with base and bounds information (policy data),
+and their use requires rewriting source code (e.g. to use capabilities in place
+of ordinary pointers), so separating ordinary machine state from policy
+state is delicate. \apt{But we conjecture it can be done...?}\leo{Just say that?}
+
+%% Mapping the underlying protection
+%% mechanisms to policy states and step function would be a delicate operation, as
+%% those are tightly coupled to the hardware design. However, separating the
+%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
+%% would encompass details like the abstract call stack, the capability sealing
+%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
+%% the machine would be relatively straightforward.  \bcp{Not sure that any of
+%%   this is comprehensible... :-(}
+
+\paragraph*{Software-only policies}
+%
+Some policies are implemented in software, by modifying code in the
+compiler to (e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or
+to insert stack canaries~\citep{Cowan+98}. Such approaches should
+still satisfy our formal notion of stack safety with a trivial null
+policy, but only for programs produced by the compiler.
+\apt{not sure what to say here}
+\leo{Something like this?}
+%These approaches are represented as policies that
+%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
+%being produced by the compiler.) The set of policy states is again the unit type
+%and the step function constant. Property enforcement comes from the fact that
+%only ill-formed states can induce a trace that violates the property.
+
+%% \Paragraph*{Null policy}
+%% %
+%% The simplest possible policy is the neutral policy that carries no information
+%% and does nothing. Its set of policy states is the unit type, and its policy step
+%% function is the constant unit function, thereby allowing all calls to the machine
+%% step function. Its initialization function is defined on all states. This
+%% corresponds to the lifting of our machine model to the policy setting. Machines
+%% that fulfill a property by construction will also fulfill that property with the
+%% null policy.
+
+%
+%\leo{TODO: Add citations of mitigation techniques}
+%\rb{And for those techniques, explain how they reflect their motivating
+%examples. What does each technique do for those examples? How are they
+%connected to this work?}
+%
+%\paragraph*{Bad Write}
+%
+%\[
+%44 : [\rsp - 5] \leftarrow r_4
+%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
+%\]
+%
+%Instruction 44 currently is
+%$ [\rsp - 2] \leftarrow r_4$,
+%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
+%with the
+%effect of storing the contents of $r_4$ as the result of {\tt h}. If
+%instead it wrote to location $\rsp - 5$ then it would be overwriting
+%private data (the return address) belonging to {\tt g}.
+%
+%\paragraph*{Bad Read}
+%
+%\[
+%30 : r_5  \leftarrow [\rsp - 6]
+%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+%\]
+%
+%Instruction 30 currently is
+%$ r_5 \leftarrow [\rsp - 4]$,
+%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
+%with the
+%effect of loading the argument of {\tt g} to $r_5$. If instead it
+%loaded the contents of location $\rsp - 6$, it would obtain access
+%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
+%outputs the contents of $r_5$, that value would immediately be made public.
+%
+%\paragraph*{Bad Control Flow}
+%
+%\[
+%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
+%%5 : \jalr ~ r_4 ~ \rra ~ 0
+%\]
+%If a {\jalr} instruction was added to {\tt main} with its target being
+%the middle of the code block of {\tt h}.\apt{describe result of this?}
+
 \section{Running Example}
@@
 \newcommand*{\jalr}{\textsc{jalr}}
-\newcommand*{\rsp}{\textsc{sp}}
+%\newcommand*{\rsp}{\textsc{sp}}
 \newcommand*{\rra}{\textsc{ra}}
@@
 \vspace*{0.2em}
-\iflater\bcp{A very small nit: Change
+\ifaftersubmission\bcp{A very small nit: Change
   $\rsp-2$ to $\rsp\mathord{-}2$}\apt{??}\bcp{I meant remove spaces around +
@@
 %
-First, since our goal is to formalize stack safety, we ignore the heap and
-assume that all variables are stack allocated.
-(Integrating a memory safety property for the heap in the style of
-\citet{AzevedoDeAmorim17} is a natural extension that we discuss
-in~\Cref{sec:future}.)
+First, since our goal is to formalize stack safety, we ignore the heap
+and assume that all variables are stack allocated. As noted in the
+previous section, the extension to account for memory safety is
+discussed in~\Cref{sec:related}.
 %
@@
 %particular compilation style, calling sequences, etc. illustrated here.}
-
-\section{Threat Model and Assumptions}
-\label{sec:threat}
-
-\leo{Needs high level signposting, transitions}
-
-Stack safety is meant to protect a caller from a confused or malicious callee's
-attempts to access private data or hijack its control flow. Our
-attacker model reflects this intention by assuming that the attacker
-can run arbitrary machine code to try to access or overwrite its caller's stack data
-or disrupt its control state.  Hardware and timing attacks are out of
-scope.
-%\leo{Anything else to add here?}
-%\leo{Maybe a good place to discuss calling conventions and that they
-%  are essentially enforced by the policy or sth?}
-%\rb{Intuition that the callee can do anything in principle, but the machine will
-%halt if it does not follow the calling convention, etc.?}
-
-We assume we are given a machine language program, together with certain auxiliary declarations
-about its structure: the code footprint of each function and the locations of instructions
-representing calls, returns, and function entry points.
-Although this information would typically come from a compiler, we do not
-assume anything about the provenance of the machine
-code. We assume that the machine has a program counter $\PCname$
-and stack pointer $\rsp$ and can perform the usual
-loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
-
-To make our stack safety definitions precise, we make some simple assumptions about
-memory layout; our definitions could easily be changed to be parametric over
-these assumptions, at the cost of slightly greater complexity.  Memory 
-consists of an instruction region and a stack that grows upwards; for simplicity,
-we omit a heap. (Integrating a memory safety property for the heap in the style of
-\citet{AzevedoDeAmorim17} is a natural extension that we discuss
-in~\Cref{sec:future}.)  The stack has a conventional structure, with
-contiguously allocated per-function frames.
-The boundary between the frames of a caller and callee is given by the $\rsp$ register
-at the time of the call.  The stack can be used to
-pass arguments or return results in the usual way, by placing them
-at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
-definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
-
-Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code.  We define a call to be
-any instruction declared as such in the program.
-If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
-  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
-of well-bracketed control flow relates this definition of return to the program's
-declared set of return points.)
-Our property definitions make no other detailed assumptions about call or return sequences, although 
-some of the enforcement mechanisms we consider do.
-
-Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
-policy state to the ordinary machine state\bcp{This may be surprising to
-people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
-step function that 
-refines the machine step function by allowing or disallowing a step to be taken
-based on its auxiliary state.   Enforcement works by halting the program before
-it can perform an action that would violate the security property; our properties
-are naturally therefore \emph{termination insensitive.}
-This model will be made more precise in \cref{sec:prelim}.
-
-We model ordinary program state and policy enforcement state separately to make it 
-easier to use well-established concepts from language-based security.  
-In particular, confidentiality has an elegant extensional characterization in terms of
-\emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable behavior~\citep{??}.
-In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, whereas any associated policy enforcement
-state should be left alone; our model makes this distinction obvious.
-%That is, we compare the actual execution of the program with a \emph{hypothetical}
-%execution where data has been varied. 
-We also introduce a notion of observational integrity that relies on the idea of
-\emph{rolling back} program state; again, our model makes it easy to describe
-rollback of ordinary program data while leaving policy enforcement data unchanged.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-This enforcement model is flexible enough to capture any kind of mechanism used
-for dynamic enforcement of properties of interest, including the following
-representative families of policies.
-\apt{I don't really buy this, because we really depend on the ability to define
-  variation and rollback. I think we should back off and say: this
-  works well for micropolicies, and we conjecture that we can make it work
-  for code-altering appraoches (pure software or capability).}
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
-
-\paragraph*{Micropolicies}
-%
-Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
-reference monitoring mechanism, which have been applied to stack safety by 
-\citet{DBLP:conf/sp/RoesslerD18}.
-Here the policy state consists of metadata tags (e.g., identifying stack frames)
-attached to each value in memory or registers, and the policy step function
-checks that each machine operation obeys a set of rules on tags
-(e.g., that the current $\PCname$ tag matches the tag on the stack
-location being accessed), and halts the machine if not.  Distinguishing
-ordinary machine state and policy state is very natural in this setting.
-%% Here the machine state is extended with tags paired with all components of the
-%% system, and a monitor checks for each operation that the tags on the operands
-%% obey a set of rules. One could treat tags directly as part of the machine
-%% state, and build the rules into the step function. But local state
-%% encapsulation properties as we frame them are not intuitive in their
-%% interaction with tags or other additional enforcement state, so it is helpful
-%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
-%% a software fail state.
-%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-%%   policy? }
-
-\paragraph*{Capability machines}
-%
-%Other policies are implemented in hardware or built on hardware mechanisms.
-Capability machines like CHERI \citep{Woodruff+14} extend conventional
-architectures to support efficient and fine-grained control over memory
-accesses, which can be used to implement security policies, including 
-the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
-(ordinary data) with base and bounds information (policy data),
-and their use requires rewriting source code (e.g. to use capabilities in place
-of ordinary pointers), so separating ordinary machine state from policy
-state is delicate. \apt{But we conjecture it can be done...?}
-
-%% Mapping the underlying protection
-%% mechanisms to policy states and step function would be a delicate operation, as
-%% those are tightly coupled to the hardware design. However, separating the
-%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
-%% would encompass details like the abstract call stack, the capability sealing
-%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-%% the machine would be relatively straightforward.  \bcp{Not sure that any of
-%%   this is comprehensible... :-(}
-
-\paragraph*{Software-only policies}
-%
-Some policies are implemented in software, by modifying code in the compiler to
-(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
-canaries~\citep{Cowan+98}. \apt{not sure what to say here}
-%These approaches are represented as policies that
-%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
-%being produced by the compiler.) The set of policy states is again the unit type
-%and the step function constant. Property enforcement comes from the fact that
-%only ill-formed states can induce a trace that violates the property.
-
-\rb{TODO In a way, in these schemes the policy state is built in as part of the
-  program state and transparent to the hardware, although it could be simulated
-  by an overlay where enforcement is represented at the policy level.}
-
-%% \Paragraph*{Null policy}
-%% %
-%% The simplest possible policy is the neutral policy that carries no information
-%% and does nothing. Its set of policy states is the unit type, and its policy step
-%% function is the constant unit function, thereby allowing all calls to the machine
-%% step function. Its initialization function is defined on all states. This
-%% corresponds to the lifting of our machine model to the policy setting. Machines
-%% that fulfill a property by construction will also fulfill that property with the
-%% null policy.
-
-%
-%\leo{TODO: Add citations of mitigation techniques}
-%\rb{And for those techniques, explain how they reflect their motivating
-%examples. What does each technique do for those examples? How are they
-%connected to this work?}
-%
-%\paragraph*{Bad Write}
-%
-%\[
-%44 : [\rsp - 5] \leftarrow r_4
-%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
-%\]
-%
-%Instruction 44 currently is
-%$ [\rsp - 2] \leftarrow r_4$,
-%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
-%with the
-%effect of storing the contents of $r_4$ as the result of {\tt h}. If
-%instead it wrote to location $\rsp - 5$ then it would be overwriting
-%private data (the return address) belonging to {\tt g}.
-%
-%\paragraph*{Bad Read}
-%
-%\[
-%30 : r_5  \leftarrow [\rsp - 6]
-%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-%\]
-%
-%Instruction 30 currently is
-%$ r_5 \leftarrow [\rsp - 4]$,
-%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
-%with the
-%effect of loading the argument of {\tt g} to $r_5$. If instead it
-%loaded the contents of location $\rsp - 6$, it would obtain access
-%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
-%outputs the contents of $r_5$, that value would immediately be made public.
-%
-%\paragraph*{Bad Control Flow}
-%
-%\[
-%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
-%%5 : \jalr ~ r_4 ~ \rra ~ 0
-%\]
-%If a {\jalr} instruction was added to {\tt main} with its target being
-%the middle of the code block of {\tt h}.\apt{describe result of this?}
 

### 2020-07-09 main.tex
Abstract nit

@@
   %
-We show how these properties are enforced or approximated by a range of
-existing stack protection mechanisms\bcp{that is not exactly what we do,
-  though.  Let's (re)write the contributions first.} and discuss some useful
-stronger and weaker variants.  \bcp{should we mention stktokens here?}
+We discuss how these properties are enforced or approximated by  
+existing stack protection mechanisms and describe some useful
+provably stronger and weaker variants.  \bcp{should we mention stktokens here?}
 \end{abstract}

### 2020-07-09 main.tex
pass on contributions

--- 2020-07-09 main.tex
@@
   the stack is a built-in concept, not something that can be attacked even
-  in principle).  This could cause confusion.}
+  in principle).  This could cause confusion.\apt{not so worried}}
 
@@
 use of uninitialized variables\bcp{Does use of unitialized variables lead to
-stack safety attacks?}.
+stack safety attacks?\apt{sure, via buffer overflow}}.
 %with the notion usually being
@@
 associated call in the standard nested call structure.  \apt{clarify
-  what we are contributing here.}
-
-To phrase these two aspects of stack safety as security properties requires
+  what we are contributing here. We are not trying to phrase this as a security property.}
+
+To phrase these two \apt{which two? Not WBCF!} aspects of stack safety as security properties requires
 an explicit attacker model: a clear understanding of who is being protected
@@
 mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
-conservative Depth Isolation policy, which not only enforces the
-our formal definition of stack safety, but also a much stronger
-property that requires integrity and confidentiality at every step of
-an execution trace. This property is provably (in Coq)\leo{How to
-  mention this?} stronger and stemmed out of the need to more
-efficiently test stack safety. However, at the same time the
+conservative Depth Isolation policy, which enforces the
+our basic stack safety property. It also enforces an alternative
+lockstep property that requires integrity and confidentiality at every step of
+an execution trace, which fails faster and thus is more efficient to test;
+we prove (in Coq) that the lockstep property is stronger 
+than our basic property,
+However, at the same time the
 conservative enforcement mechanism is expensive to implement. In the
@@
 enforcement. At a high level, instead of guaranteeing integrity and
-confidentiality when a callee returns, they check it when the caller
-actually makes use of (potentially compromised) data. Thus, they
-enforce a (also provably in Coq) weaker property that still protects
-from unsafe callees, but is much less intuitive.
+confidentiality when a callee returns, this policy checks it when the caller
+actually makes use of (potentially compromised) data. The lazy policy
+does not enforce our basic propety, but an enhanced version of it does
+enforce a proveably weaker property that still protects
+from unsafe callees, but is less intuitive.
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
-  Abstract formal properties capturing the
+  A novel formal stack safety property capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
-  control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
-  rule out common examples of stack memory abuse.
+  control flow~(\cref{sec:lse-and-wbcf}) in terms of security concepts.
+  % These properties   rule out common examples of stack memory abuse.
+  This property is enforced by 
+  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:conservative}).
 \item
-  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
-  This policy enforces our strict stack safety property, and also a (provably in Coq) stronger property that is easier to test.
-  \apt{If we have no testing, do we even want to talk about the testing
-    property?} \bcp{Not here; just in future work.  I guess we need to see
-    whether we wind up with any results at all about testing.}
+  A lockstep stack safety property that is also enforced by the Depth Isolation policy,
+  and is more efficient to test, which we prove (in Coq) is stronger than the basic property(~\cref{sec:lockstep}).
 \item
-  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
-  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our strict
-  local state encapsulation property,
-  but a modified version of it using unique per-activation tags enforces
-  an (also provably) weaker lazy variant of the property, which still prohibits
-  the same intuitively stack unsafe behaviors.
+  A novel lazy stack safety property that is enforced by the more efficient
+  Lazy Tagging and Clearing policy
+  of \citet{DBLP:conf/sp/RoesslerD18}, which we prove (in Coq) is weaker than the basic property~(\cref{sec:lazy}).
+%  This policy does not enforce our strict
+%  local state encapsulation property,
+%  but a modified version of it using unique per-activation tags enforces
+%  an (also provably) weaker lazy variant of the property, which still prohibits
+%  the same intuitively stack unsafe behaviors.
 %
@@
 %
-We begin with a running example that we refer to throughout the paper
-(\cref{sec:running-example}), a description of our threat model
-(\cref{sec:threat}) and some technical preliminaries
+We begin with a description of our threat model
+(\cref{sec:threat}), a running example that we refer to throughout the paper
+(\cref{sec:running-example}),  and some technical preliminaries
 (\cref{sec:prelim}).
+
+\section{Threat Model and Assumptions}
+\label{sec:threat}
+
+\leo{Needs high level signposting, transitions}
+
+Stack safety is meant to protect a caller from a confused or malicious callee's
+attempts to access private data or hijack its control flow. Our
+attacker model reflects this intention by assuming that the attacker
+can run arbitrary machine code to try to access or overwrite its caller's stack data
+or disrupt its control state.  Hardware and timing attacks are out of
+scope.
+%\leo{Anything else to add here?}
+%\leo{Maybe a good place to discuss calling conventions and that they
+%  are essentially enforced by the policy or sth?}
+%\rb{Intuition that the callee can do anything in principle, but the machine will
+%halt if it does not follow the calling convention, etc.?}
+
+\newcommand*{\rsp}{\textsc{sp}}
+
+We assume we are given a machine language program, together with certain auxiliary declarations
+about its structure: the code footprint of each function and the locations of instructions
+representing calls, returns, and function entry points.
+Although this information would typically come from a compiler, we do not
+assume anything about the provenance of the machine
+code. We assume that the machine has a program counter $\PCname$
+and stack pointer $\rsp$ and can perform the usual
+loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
+
+To make our stack safety definitions precise, we make some simple assumptions about
+memory layout; our definitions could easily be changed to be parametric over
+these assumptions, at the cost of slightly greater complexity.  Memory 
+consists of an instruction region and a stack that grows upwards; for simplicity,
+we omit a heap. (Integrating a memory safety property for the heap in the style of
+\citet{AzevedoDeAmorim17} is a natural extension that we discuss
+in~\Cref{sec:future}.)  The stack has a conventional structure, with
+contiguously allocated per-function frames.
+The boundary between the frames of a caller and callee is given by the $\rsp$ register
+at the time of the call.  The stack can be used to
+pass arguments or return results in the usual way, by placing them
+at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
+definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
+
+Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
+which are not fully explicit in machine code.  We define a call to be
+any instruction declared as such in the program.
+If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
+  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
+of well-bracketed control flow relates this definition of return to the program's
+declared set of return points.)
+Our property definitions make no other detailed assumptions about call or return sequences, although 
+some of the enforcement mechanisms we consider do.
+
+Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
+policy state to the ordinary machine state\bcp{This may be surprising to
+people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
+step function that 
+refines the machine step function by allowing or disallowing a step to be taken
+based on its auxiliary state.   Enforcement works by halting the program before
+it can perform an action that would violate the security property; our properties
+are naturally therefore \emph{termination insensitive.}
+This model will be made more precise in \cref{sec:prelim}.
+
+We model ordinary program state and policy enforcement state separately to make it 
+easier to use well-established concepts from language-based security.  
+In particular, confidentiality has an elegant extensional characterization in terms of
+\emph{noninterference}: a program preserves the confidentiality of certain data if
+\emph{varying} that data would not change the program's observable behavior~\citep{??}.
+In applying this idea to a system that incorporates policy enforcement, only
+ordinary program data should be varied, whereas any associated policy enforcement
+state should be left alone; our model makes this distinction obvious.
+%That is, we compare the actual execution of the program with a \emph{hypothetical}
+%execution where data has been varied. 
+We also introduce a notion of observational integrity that relies on the idea of
+\emph{rolling back} program state; again, our model makes it easy to describe
+rollback of ordinary program data while leaving policy enforcement data unchanged.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+This enforcement model is flexible enough to capture any kind of mechanism used
+for dynamic enforcement of properties of interest, including the following
+representative families of policies.
+\apt{I don't really buy this, because we really depend on the ability to define
+  variation and rollback. I think we should back off and say: this
+  works well for micropolicies, and we conjecture that we can make it work
+  for code-altering appraoches (pure software or capability).}
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
+
+\paragraph*{Micropolicies}
+%
+Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
+reference monitoring mechanism, which have been applied to stack safety by 
+\citet{DBLP:conf/sp/RoesslerD18}.
+Here the policy state consists of metadata tags (e.g., identifying stack frames)
+attached to each value in memory or registers, and the policy step function
+checks that each machine operation obeys a set of rules on tags
+(e.g., that the current $\PCname$ tag matches the tag on the stack
+location being accessed), and halts the machine if not.  Distinguishing
+ordinary machine state and policy state is very natural in this setting.
+%% Here the machine state is extended with tags paired with all components of the
+%% system, and a monitor checks for each operation that the tags on the operands
+%% obey a set of rules. One could treat tags directly as part of the machine
+%% state, and build the rules into the step function. But local state
+%% encapsulation properties as we frame them are not intuitive in their
+%% interaction with tags or other additional enforcement state, so it is helpful
+%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
+%% a software fail state.
+%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+%%   policy? }
+
+\paragraph*{Capability machines}
+%
+%Other policies are implemented in hardware or built on hardware mechanisms.
+Capability machines like CHERI \citep{Woodruff+14} extend conventional
+architectures to support efficient and fine-grained control over memory
+accesses, which can be used to implement security policies, including 
+the stack protection policies of
+\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
+(ordinary data) with base and bounds information (policy data),
+and their use requires rewriting source code (e.g. to use capabilities in place
+of ordinary pointers), so separating ordinary machine state from policy
+state is delicate. \apt{But we conjecture it can be done...?}
+
+%% Mapping the underlying protection
+%% mechanisms to policy states and step function would be a delicate operation, as
+%% those are tightly coupled to the hardware design. However, separating the
+%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
+%% would encompass details like the abstract call stack, the capability sealing
+%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
+%% the machine would be relatively straightforward.  \bcp{Not sure that any of
+%%   this is comprehensible... :-(}
+
+\paragraph*{Software-only policies}
+%
+Some policies are implemented in software, by modifying code in the compiler to
+(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
+canaries~\citep{Cowan+98}. \apt{not sure what to say here}
+%These approaches are represented as policies that
+%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
+%being produced by the compiler.) The set of policy states is again the unit type
+%and the step function constant. Property enforcement comes from the fact that
+%only ill-formed states can induce a trace that violates the property.
+
+\rb{TODO In a way, in these schemes the policy state is built in as part of the
+  program state and transparent to the hardware, although it could be simulated
+  by an overlay where enforcement is represented at the policy level.}
+
+%% \Paragraph*{Null policy}
+%% %
+%% The simplest possible policy is the neutral policy that carries no information
+%% and does nothing. Its set of policy states is the unit type, and its policy step
+%% function is the constant unit function, thereby allowing all calls to the machine
+%% step function. Its initialization function is defined on all states. This
+%% corresponds to the lifting of our machine model to the policy setting. Machines
+%% that fulfill a property by construction will also fulfill that property with the
+%% null policy.
+
+%
+%\leo{TODO: Add citations of mitigation techniques}
+%\rb{And for those techniques, explain how they reflect their motivating
+%examples. What does each technique do for those examples? How are they
+%connected to this work?}
+%
+%\paragraph*{Bad Write}
+%
+%\[
+%44 : [\rsp - 5] \leftarrow r_4
+%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
+%\]
+%
+%Instruction 44 currently is
+%$ [\rsp - 2] \leftarrow r_4$,
+%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
+%with the
+%effect of storing the contents of $r_4$ as the result of {\tt h}. If
+%instead it wrote to location $\rsp - 5$ then it would be overwriting
+%private data (the return address) belonging to {\tt g}.
+%
+%\paragraph*{Bad Read}
+%
+%\[
+%30 : r_5  \leftarrow [\rsp - 6]
+%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+%\]
+%
+%Instruction 30 currently is
+%$ r_5 \leftarrow [\rsp - 4]$,
+%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
+%with the
+%effect of loading the argument of {\tt g} to $r_5$. If instead it
+%loaded the contents of location $\rsp - 6$, it would obtain access
+%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
+%outputs the contents of $r_5$, that value would immediately be made public.
+%
+%\paragraph*{Bad Control Flow}
+%
+%\[
+%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
+%%5 : \jalr ~ r_4 ~ \rra ~ 0
+%\]
+%If a {\jalr} instruction was added to {\tt main} with its target being
+%the middle of the code block of {\tt h}.\apt{describe result of this?}
 
@@
 
-\section{Threat Model and Assumptions}
-\label{sec:threat}
-
-\leo{Needs high level signposting, transitions}
-
-Stack safety is meant to protect a caller from a confused or malicious callee's
-attempts to access private data or hijack its control flow. Our
-attacker model reflects this intention by assuming that the attacker
-can run arbitrary machine code to try to access or overwrite its caller's stack data
-or disrupt its control state.  Hardware and timing attacks are out of
-scope.
-%\leo{Anything else to add here?}
-%\leo{Maybe a good place to discuss calling conventions and that they
-%  are essentially enforced by the policy or sth?}
-%\rb{Intuition that the callee can do anything in principle, but the machine will
-%halt if it does not follow the calling convention, etc.?}
-
-We assume we are given a machine language program, together with certain auxiliary declarations
-about its structure: the code footprint of each function and the locations of instructions
-representing calls, returns, and function entry points.
-Although this information would typically come from a compiler, we do not
-assume anything about the provenance of the machine
-code. We assume that the machine has a program counter $\PCname$
-and stack pointer $\rsp$ and can perform the usual
-loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
-
-To make our stack safety definitions precise, we make some simple assumptions about
-memory layout; our definitions could easily be changed to be parametric over
-these assumptions, at the cost of slightly greater complexity.  Memory 
-consists of an instruction region and a stack that grows upwards; for simplicity,
-we omit a heap. (Integrating a memory safety property for the heap in the style of
-\citet{AzevedoDeAmorim17} is a natural extension that we discuss
-in~\Cref{sec:future}.)  The stack has a conventional structure, with
-contiguously allocated per-function frames.
-The boundary between the frames of a caller and callee is given by the $\rsp$ register
-at the time of the call.  The stack can be used to
-pass arguments or return results in the usual way, by placing them
-at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
-definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
-
-Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code.  We define a call to be
-any instruction declared as such in the program.
-If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
-  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
-of well-bracketed control flow relates this definition of return to the program's
-declared set of return points.)
-Our property definitions make no other detailed assumptions about call or return sequences, although 
-some of the enforcement mechanisms we consider do.
-
-Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
-policy state to the ordinary machine state\bcp{This may be surprising to
-people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
-step function that 
-refines the machine step function by allowing or disallowing a step to be taken
-based on its auxiliary state.   Enforcement works by halting the program before
-it can perform an action that would violate the security property; our properties
-are naturally therefore \emph{termination insensitive.}
-This model will be made more precise in \cref{sec:prelim}.
-
-We model ordinary program state and policy enforcement state separately to make it 
-easier to use well-established concepts from language-based security.  
-In particular, confidentiality has an elegant extensional characterization in terms of
-\emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable behavior~\citep{??}.
-In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, whereas any associated policy enforcement
-state should be left alone; our model makes this distinction obvious.
-%That is, we compare the actual execution of the program with a \emph{hypothetical}
-%execution where data has been varied. 
-We also introduce a notion of observational integrity that relies on the idea of
-\emph{rolling back} program state; again, our model makes it easy to describe
-rollback of ordinary program data while leaving policy enforcement data unchanged.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-This enforcement model is flexible enough to capture any kind of mechanism used
-for dynamic enforcement of properties of interest, including the following
-representative families of policies.
-\apt{I don't really buy this, because we really depend on the ability to define
-  variation and rollback. I think we should back off and say: this
-  works well for micropolicies, and we conjecture that we can make it work
-  for code-altering appraoches (pure software or capability).}
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
-
-\paragraph*{Micropolicies}
-%
-Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
-reference monitoring mechanism, which have been applied to stack safety by 
-\citet{DBLP:conf/sp/RoesslerD18}.
-Here the policy state consists of metadata tags (e.g., identifying stack frames)
-attached to each value in memory or registers, and the policy step function
-checks that each machine operation obeys a set of rules on tags
-(e.g., that the current $\PCname$ tag matches the tag on the stack
-location being accessed), and halts the machine if not.  Distinguishing
-ordinary machine state and policy state is very natural in this setting.
-%% Here the machine state is extended with tags paired with all components of the
-%% system, and a monitor checks for each operation that the tags on the operands
-%% obey a set of rules. One could treat tags directly as part of the machine
-%% state, and build the rules into the step function. But local state
-%% encapsulation properties as we frame them are not intuitive in their
-%% interaction with tags or other additional enforcement state, so it is helpful
-%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
-%% a software fail state.
-%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-%%   policy? }
-
-\paragraph*{Capability machines}
-%
-%Other policies are implemented in hardware or built on hardware mechanisms.
-Capability machines like CHERI \citep{Woodruff+14} extend conventional
-architectures to support efficient and fine-grained control over memory
-accesses, which can be used to implement security policies, including 
-the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
-(ordinary data) with base and bounds information (policy data),
-and their use requires rewriting source code (e.g. to use capabilities in place
-of ordinary pointers), so separating ordinary machine state from policy
-state is delicate. \apt{But we conjecture it can be done...?}
-
-%% Mapping the underlying protection
-%% mechanisms to policy states and step function would be a delicate operation, as
-%% those are tightly coupled to the hardware design. However, separating the
-%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
-%% would encompass details like the abstract call stack, the capability sealing
-%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-%% the machine would be relatively straightforward.  \bcp{Not sure that any of
-%%   this is comprehensible... :-(}
-
-\paragraph*{Software-only policies}
-%
-Some policies are implemented in software, by modifying code in the compiler to
-(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
-canaries~\citep{Cowan+98}. \apt{not sure what to say here}
-%These approaches are represented as policies that
-%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
-%being produced by the compiler.) The set of policy states is again the unit type
-%and the step function constant. Property enforcement comes from the fact that
-%only ill-formed states can induce a trace that violates the property.
-
-\rb{TODO In a way, in these schemes the policy state is built in as part of the
-  program state and transparent to the hardware, although it could be simulated
-  by an overlay where enforcement is represented at the policy level.}
-
-%% \Paragraph*{Null policy}
-%% %
-%% The simplest possible policy is the neutral policy that carries no information
-%% and does nothing. Its set of policy states is the unit type, and its policy step
-%% function is the constant unit function, thereby allowing all calls to the machine
-%% step function. Its initialization function is defined on all states. This
-%% corresponds to the lifting of our machine model to the policy setting. Machines
-%% that fulfill a property by construction will also fulfill that property with the
-%% null policy.
-
-%
-%\leo{TODO: Add citations of mitigation techniques}
-%\rb{And for those techniques, explain how they reflect their motivating
-%examples. What does each technique do for those examples? How are they
-%connected to this work?}
-%
-%\paragraph*{Bad Write}
-%
-%\[
-%44 : [\rsp - 5] \leftarrow r_4
-%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
-%\]
-%
-%Instruction 44 currently is
-%$ [\rsp - 2] \leftarrow r_4$,
-%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
-%with the
-%effect of storing the contents of $r_4$ as the result of {\tt h}. If
-%instead it wrote to location $\rsp - 5$ then it would be overwriting
-%private data (the return address) belonging to {\tt g}.
-%
-%\paragraph*{Bad Read}
-%
-%\[
-%30 : r_5  \leftarrow [\rsp - 6]
-%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-%\]
-%
-%Instruction 30 currently is
-%$ r_5 \leftarrow [\rsp - 4]$,
-%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
-%with the
-%effect of loading the argument of {\tt g} to $r_5$. If instead it
-%loaded the contents of location $\rsp - 6$, it would obtain access
-%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
-%outputs the contents of $r_5$, that value would immediately be made public.
-%
-%\paragraph*{Bad Control Flow}
-%
-%\[
-%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
-%%5 : \jalr ~ r_4 ~ \rra ~ 0
-%\]
-%If a {\jalr} instruction was added to {\tt main} with its target being
-%the middle of the code block of {\tt h}.\apt{describe result of this?}
-
 \section{Machines, Traces, and Observations}
@@
 \subsection{A Conservative Policy}
+\label{sec:conservative}
 %
@@
 \subsection{A Stronger Lockstep Stack Safety Property}
+\label{sec:lockstep}
 

### 2020-07-09 main.tex

@@
 we omit a heap. (Integrating a memory safety property for the heap in the style of
-\citet{AzevedoDeAmorim17} is a natural extension that we discuss
-in~\Cref{sec:future}.)  The stack has a conventional structure, with
+\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
+in~\Cref{sec:related}.)  The stack has a conventional structure, with
 contiguously allocated per-function frames.
@@
 Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
-policy state to the ordinary machine state\bcp{This may be surprising to
-people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
-step function that 
-refines the machine step function by allowing or disallowing a step to be taken
-based on its auxiliary state.   Enforcement works by halting the program before
-it can perform an action that would violate the security property; our properties
-are naturally therefore \emph{termination insensitive.}
+policy state to the ordinary machine state
+%\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
+and a partial policy step function that refines the machine step
+function by allowing or disallowing a step to be taken based on its
+auxiliary state.  Enforcement works by halting the program before it
+can perform an action that would violate the security property; our
+properties are naturally therefore \emph{termination insensitive.}
 This model will be made more precise in \cref{sec:prelim}.
@@
 
-This enforcement model is flexible enough to capture any kind of mechanism used
-for dynamic enforcement of properties of interest, including the following
-representative families of policies.
-\apt{I don't really buy this, because we really depend on the ability to define
-  variation and rollback. I think we should back off and say: this
-  works well for micropolicies, and we conjecture that we can make it work
-  for code-altering appraoches (pure software or capability).}
+This enforcement model is flexible enough to capture
+micropolicies~\citep{pump_oakland2015}, and we believe that it can
+also capture code-altering approaches (such as software-based
+or capability-based techniques).
 
@@
 of ordinary pointers), so separating ordinary machine state from policy
-state is delicate. \apt{But we conjecture it can be done...?}
+state is delicate. \apt{But we conjecture it can be done...?}\leo{Just say that?}
 
@@
 %
-Some policies are implemented in software, by modifying code in the compiler to
-(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
-canaries~\citep{Cowan+98}. \apt{not sure what to say here}
+Some policies are implemented in software, by modifying code in the
+compiler to (e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or
+to insert stack canaries~\citep{Cowan+98}. Such approaches should
+still satisfy our formal notion of stack safety with a trivial null
+policy, but only for programs produced by the compiler.
+\apt{not sure what to say here}
+\leo{Something like this?}
 %These approaches are represented as policies that
@@
 %only ill-formed states can induce a trace that violates the property.
-
-\rb{TODO In a way, in these schemes the policy state is built in as part of the
-  program state and transparent to the hardware, although it could be simulated
-  by an overlay where enforcement is represented at the policy level.}
 

### 2020-07-09 main.tex
Read 4, added pointers to policy intro and example

@@
 
+We now present our model of enforcement, introduced informally in \Cref{sec:threat}.
 A policy is a general model of an enforcement mechanism, consisting of a
@@
 
+A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be given in
+\Cref{sec:enforcement}.
+
 \apt{TODO: relate back to \Cref{sec:threat} and forward to \Cref{sec:enforcement}.}
+\rb{Added some quick pointers, although one might expect to see Nick's policy
+as a more formal instance of this framework, which at the moment it is not.}
+
 %\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say

### 2020-07-09 main.tex
Submit draft

@@
   A lockstep stack safety property that is also enforced by the Depth Isolation policy,
-  and is more efficient to test, which we prove (in Coq) is stronger than the basic property(~\cref{sec:lockstep}).
+  and is more efficient to test, which we prove (in Coq) is stronger than the basic property~(\cref{sec:lockstep}).
 \item
@@
 \label{sec:threat}
-
-\leo{Needs high level signposting, transitions}
 
@@
 
-\theorem
+\paragraph*{THEOREM}
 Lockstep stack safety implies stack safety.
 
-\proof
+\paragraph*{PROOF}
 By coinduction on trace lockstep stack safety.
@@
 
-      \theorem Stack safety implies observable stack safety.
-
-      \proof \leo{TODO: Whatever we say about the previous proof
-        we have to say similar things here}
+      \paragraph*{THEOREM} Stack safety implies observable stack safety.
+
+%      \paragraph*{PROOF} \leo{TODO: Whatever we say about the previous proof
+ %       we have to say similar things here}
 

### 2020-07-09 main.tex
Quick pass on 5

@@
      here, or that the entry point is valid; that's part of WBCF.}
+\rb{Agreed, removed $f$}
 %and \(\mach_e\) is an entry point for some function \(f\):
@@
     \mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}
-    \triangleq & \exists \obs, f.
+    \triangleq & \exists \obs.
     ~ \mpstate{\mach_c}{\pol_c} \stepstopol{\obs} \mpstate{\mach_e}{\pol_e} \\
@@
 and a ``variant'' trace $\machT \in \MTRACE$ that corresponds to how $f$
-would behaves if it were started from a variant initial machine state.
-%
-As an example of variations, consider the top of \Cref{fig:confex1}
+would behave if it were started from a variant initial machine state.
+%
+As an example of variations, consider the top of \Cref{fig:confex}
 which shows a snapshot of the state as in the running example (top
@@
   contents, describe}
+%
+The ``Effect'' column shows differences between the two runs by writing values
+in the original run in {\high} and values in the variant run in {\varied}
+separated by a slash.
 
@@
 \apt{Explain slash notation in effects column.}
+\rb{Wrote some words to explain after def. 5.4 (variations).}
 

### 2020-07-09 main.tex
1+2

@@
 guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
-  paper, do we?}, a clear and precise definition of stack safety is missing
+  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}, a clear and precise definition of stack safety is missing
 from the literature.
@@
 provably stronger and weaker variants.  \bcp{should we mention stktokens here?}
+\leo{Not sure. We almost never cite stuff in the abstract, do we?}
 \end{abstract}
@@
 
-\bcp{We don't actually say very clearly anywhere that we are interested in
-  low-level enforcement mechanisms for machine-code programs (because this
-  is where stack attacks occur, and because at higher levels of abstraction
-  the stack is a built-in concept, not something that can be attacked even
-  in principle).  This could cause confusion.\apt{not so worried}}
+%\bcp{We don't actually say very clearly anywhere that we are interested in
+%  low-level enforcement mechanisms for machine-code programs (because this
+%  is where stack attacks occur, and because at higher levels of abstraction
+%  the stack is a built-in concept, not something that can be attacked even
+%  in principle).  This could cause confusion.\apt{not so worried}}
+%\leo{I agree. I've clarified the writing in the top of the running example
+%  a bit to make sure though}
 
@@
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \iflater\apt{Mostly from
-  nick; there could be more}\fi
+  nick; there could be more}\leo{Do we want to hunt for more refs here?}\fi
 
@@
 prevents, such as corruption of return addresses, buffer overflows, or
-use of uninitialized variables\bcp{Does use of unitialized variables lead to
-stack safety attacks?\apt{sure, via buffer overflow}}.
+use of uninitialized variables.
+%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
 %with the notion usually being
@@
 %
-Our key contribution is a novel, generic presentation of
+Our key contribution is such a novel and generic presentation of
 local state encapsulation and well-bracketed control flow using the
@@
 
-\iflater
-\bcp{We need to point out somewhere that we are only dealing with
-stack safety, though we discuss (somewhere) how our property can be combined
-with the heap safety property of [Azevedo et al]}\leo{Should this be
-  here or in the assumptions as presented in the running
-  example?}\bcp{Probably better there.}
-\fi
-
-The intuition behind \citeauthor{Skorstengaard+19}'s {\em well-bracketed control
-  flow} is
-the need to protect the call graph itself: a callee should pass control
-back to the caller only by returning in the standard way (transferring
-control to the instruction following the call); it should
-not be
-allowed to jump back into the caller's code at an arbitrary point, either
-directly (via a jump instruction) or indirectly (by corrupting stack
-data and returning). Well-bracketed control flow establishes safety
-between functions: each instruction is uniquely owned by some
-function, and all jumps between functions are either calls that enter
-at a marked entry point or returns that re-enter just after their
-associated call in the standard nested call structure.  \apt{clarify
-  what we are contributing here. We are not trying to phrase this as a security property.}
-
-To phrase these two \apt{which two? Not WBCF!} aspects of stack safety as security properties requires
-an explicit attacker model: a clear understanding of who is being protected
-and from what. Most existing stack protection mechanisms envisage scenarios
-where a callee is influenced by external inputs to corrupt the stack in some
-way. To avoid getting into the details of exactly how this happens when
-formalizing our properties, we use an even stronger model: when we state the
-protections that a function can rely on when it makes a call, we assume that
-the callee can execute arbitrary instructions until it returns.
+The intuition behind \citeauthor{Skorstengaard+19}'s {\em
+  well-bracketed control flow} is the need to protect the call graph
+itself: a callee should pass control back to the caller only by
+returning in the standard way (transferring control to the instruction
+following the call); it should not be allowed to jump back into the
+caller's code at an arbitrary point, either directly (via a jump
+instruction) or indirectly (by corrupting stack data and
+returning). Well-bracketed control flow establishes safety between
+functions: each instruction is uniquely owned by some function, and
+all jumps between functions are either calls that enter at a marked
+entry point or returns that re-enter just after their associated call
+in the standard nested call structure.  \apt{clarify what we are
+  contributing here. We are not trying to phrase this as a security
+  property.}\leo{Better now? Should this be fixed two paragraphs above too?}
+
+To phrase stack safety, and especially local state encapsulation, as a
+security property requires an explicit attacker model: a clear
+understanding of who is being protected and from what. Most existing
+stack protection mechanisms envisage scenarios where a callee is
+influenced by external inputs to corrupt the stack in some way. To
+avoid getting into the details of exactly how this happens when
+formalizing our properties, we use an even stronger model: when we
+state the protections that a function can rely on when it makes a
+call, we assume that the callee can execute arbitrary instructions
+until it returns.
 
@@
 
-\leo{Took a stab. Needs rewriting}%
-
 After formally defining stack safety, we show that our definition is
@@
 mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
-conservative Depth Isolation policy, which enforces the
-our basic stack safety property. It also enforces an alternative
-lockstep property that requires integrity and confidentiality at every step of
-an execution trace, which fails faster and thus is more efficient to test;
-we prove (in Coq) that the lockstep property is stronger 
-than our basic property,
-However, at the same time the
-conservative enforcement mechanism is expensive to implement. In the
-same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different
-mechanism that is more efficient by virtue of being lazy in its
-enforcement. At a high level, instead of guaranteeing integrity and
-confidentiality when a callee returns, this policy checks it when the caller
-actually makes use of (potentially compromised) data. The lazy policy
-does not enforce our basic propety, but an enhanced version of it does
-enforce a proveably weaker property that still protects
-from unsafe callees, but is less intuitive.
-
-In summary, we offer the following contributions:\bcp{I would still like to
-  see these points phrased in ``contribution normal form'': say exactly what
-  we did for each one (it's a bit vague---''considers,'' etc.), and
-  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
+conservative Depth Isolation policy, which not only enforces our basic
+stack safety property, but also an alternative lockstep property that
+checks for integrity and confidentiality violations at every step of
+an execution trace. Thus, this property fails faster and can lead to
+more efficient testing. We prove (in Coq) that this lockstep property
+implies our basic definition of stack safety.
+
+However, at the same time the conservative enforcement mechanism is
+expensive to implement. In the same work,
+\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
+that is more efficient by virtue of being lazy in its enforcement. At
+a high level, instead of guaranteeing integrity and confidentiality
+when a callee returns, this policy checks for violations when the
+caller actually makes use of (potentially compromised) data. The lazy
+policy does not enforce our basic propety, but an enhanced version of
+it does enforce a proveably weaker property that still protects from
+unsafe callees, but is less intuitive.
+
+In summary, we offer the following contributions:
+%\bcp{I would still like to
+%  see these points phrased in ``contribution normal form'': say exactly what
+%  we did for each one (it's a bit vague---''considers,'' etc.), and
+%  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
+% +1 on Andrew's rewrite
 \begin{itemize}
@@
 memory layout; our definitions could easily be changed to be parametric over
-these assumptions, at the cost of slightly greater complexity.  Memory 
+these assumptions, at the cost of greater complexity.  Memory 
 consists of an instruction region and a stack that grows upwards; for simplicity,
@@
 \citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
-in~\Cref{sec:related}.)  The stack has a conventional structure, with
+in~\Cref{sec:relwork}.)  The stack has a conventional structure, with
 contiguously allocated per-function frames.
@@
 \emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable behavior~\citep{??}.
+\emph{varying} that data would not change the program's observable behavior.\leo{What citation should go here? Not the hyperproperties one?}
 In applying this idea to a system that incorporates policy enforcement, only
@@
 of ordinary pointers), so separating ordinary machine state from policy
-state is delicate. \apt{But we conjecture it can be done...?}\leo{Just say that?}
+state is delicate, but we conjecture it can be done with some effort.
+%\apt{But we conjecture it can be done...?}\leo{Just say that?}
 
@@
 \bcp{We really want readers to understand this trace.  I think we should
-  spend a few more words walking through it in detail.}
+  spend a few more words walking through it in detail.}\leo{Not sure
+  exactly what you mean here. It would take roughly a page to go through
+  each instruction in detail - that's what the last column is for in my mind.
+  Or do you have something else in mind?}
 
@@
 previous section, the extension to account for memory safety is
-discussed in~\Cref{sec:related}.
+discussed in~\Cref{sec:relwork}.
 %

### 2020-07-09 main.tex
Nit: missing word in abstract

--- 2020-07-09 main.tex
@@
 guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
-  paper, do we?}, a clear and precise definition of stack safety is missing
+  paper, do we?}, but a clear and precise definition of stack safety is missing
 from the literature.

### 2020-07-09 main.tex
Polishing / squishing comments

--- 2020-07-09 main.tex
@@
 What is stack safety? Numerous mechanisms have been proposed to guard
-against attacks that corrupt the call stacks of running programs, and many
+against attacks that corrupt the call stacks of running programs,
+\ifaftersubmission
+and many
 others rely on a protected stack as an assumption to provide higher level
 guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
-  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}, a clear and precise definition of stack safety is missing
+  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
+\bcp{Let's delete for now; later, we can try to incorporate it in the body
+  of the paper.}
+\fi,
+but a clear and precise definition of stack safety is missing
 from the literature.
-
-We propose a formal definition of stack safety phrased in the technical
+%
+We propose a formal definition phrased in the technical
 terminology of language-based security---as a combination of a
@@
 existing stack protection mechanisms and describe some useful
-provably stronger and weaker variants.  \bcp{should we mention stktokens here?}
-\leo{Not sure. We almost never cite stuff in the abstract, do we?}
+provably stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
+\leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi
 \end{abstract}
@@
 capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \iflater\apt{Mostly from
-  nick; there could be more}\leo{Do we want to hunt for more refs here?}\fi
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
+  nick; there could be more}\leo{Do we want to hunt for more refs here?}
+\bcp{Sure, but later!}
+\fi
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
-  A novel formal stack safety property capturing the
+  A new {\em formal stack-safety property} capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
   % These properties   rule out common examples of stack memory abuse.
-  This property is enforced by 
-  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:conservative}).
+  This property can be enforced by
+  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
+  based on hardware tagging~(\cref{sec:conservative}).
 \item
-  A lockstep stack safety property that is also enforced by the Depth Isolation policy,
-  and is more efficient to test, which we prove (in Coq) is stronger than the basic property~(\cref{sec:lockstep}).
+  A {\em lockstep} variant of the stack-safety property that is also
+  enforced by the Depth Isolation policy
+  but is more efficient to test, plus a proof (in Coq) that this variant is
+  stronger than the original~(\cref{sec:lockstep}).
 \item
-  A novel lazy stack safety property that is enforced by the more efficient
-  Lazy Tagging and Clearing policy
-  of \citet{DBLP:conf/sp/RoesslerD18}, which we prove (in Coq) is weaker than the basic property~(\cref{sec:lazy}).
+  A {\em lazy} variant of the stack-safety property that can be enforced by the
+  more efficient Lazy Tagging and Clearing policy
+  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
+  variant is weaker than the original~(\cref{sec:lazy}).
 %  This policy does not enforce our strict
@@
 \cref{sec:future} sketches directions for future work.
-%
+
 We begin with a description of our threat model

### 2020-07-09 main.tex
CP - 3,4 done. Mid 5

--- 2020-07-09 main.tex
@@
 in~\cref{sec:enforcement}, extend this model with a dynamic policy
-enforcement mechanism based on
-\citet{DBLP:conf/sp/RoesslerD18}.\apt{needs a relook}
+enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
 
@@
 %
+This step function is total---execution traces are infinite unless cut
+short externally (e.g. by a policy as we'll see below).
 
@@
 machine state, but is partial to reflect that some policies are only defined
-on states that are well-formed in some way.\apt{Is this our current thinking?}
+on states that are well-formed in some way.
+%
+\apt{Is this our current thinking?}
+\leo{I thought this was what we settled on in the last two days' discussions}
+%
+A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be given in
+\Cref{sec:enforcement}.
 
@@
                \mpstate{\mach_2}{\pol_2}\)}
-
-A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be given in
-\Cref{sec:enforcement}.
-
-\apt{TODO: relate back to \Cref{sec:threat} and forward to \Cref{sec:enforcement}.}
-\rb{Added some quick pointers, although one might expect to see Nick's policy
-as a more formal instance of this framework, which at the moment it is not.}
-
-%\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
-%in more detail that the intention is to capture any kind of dynamic
-%enforcement mechanism and give as many names and citations as we can.}
-
-\paragraph*{Machine Termination}
-
-In our model, the machine step function is total---a program does not halt
-on its own, but may instead enter a silent terminal loop. Runs are only cut
-short by the policy step function. This means that software-only policies that
-can catch errors must be modeled as reaching an error state\apt{??}, not
-halting outright.\apt{Is this our current thinking?}
-\rb{And if so, should the running example loop at the end, or do we assume
-that is exactly what happens after main is done?}
 
@@
 \paragraph*{Until}
-The operation \(\PUT~(\Trace,f)\)
+The operation \(\PUT~(f,\Trace)\)
 %\bcp{Wouldn't
@@
 %
-Note that if the resulting trace is a strict prefix of the parameter, then the
+If the resulting trace is a strict prefix of the parameter, then the
 predicate must hold on its final element. If the predicate never holds,
@@
 
-\bcp{Can we please call it something else?  An
-  ``equivalence'' should always be transitive.}
-\sna{Similarity seems less loaded?}
 \[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
-
-\leo{I still don't know how to do this transition. Maybe contrast
-with the ``perhaps surprisingly'' later?} Note that
-similarity of observation traces could alternatively be defined coinductively,
-just like the prefix relation, but by dropping the asymmetric rules that allow
-for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}
+% 
+%\leo{I still don't know how to do this transition. Maybe contrast
+%with the ``perhaps surprisingly'' later?} Note that
+%similarity of observation traces could alternatively be defined coinductively,
+%just like the prefix relation, but by dropping the asymmetric rules that allow
+%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}
 
@@
 
-In this section we formalize our definition of stack safety. We
-begin with our model of \leo{compiler-provided?} knowledge about the
-program structure, and then dive into the formal definitions of local state
-encapsulation and well-bracketed control flow that together constitute
-stack safety.
+In this section we formalize our definition of stack safety. We begin
+by describing our model of knowledge about the program structure, and
+then dive into the formal definitions of local state encapsulation and
+well-bracketed control flow that together constitute stack safety.
 
@@
   \integrityexample
-  \caption{Failing Trace Integrity\bcp{the ``108'' is not placed right}}
+  \caption{Failing Trace Integrity}
   \label{fig:intex}
@@
 state, and does not. So trace integrity is violated.
-\rb{There are some color name macros, but none for integrity yet? Use instead
-  of hardcoded color names.}\apt{I don't understand what's going on with the colors.
-  I thought we agreed to use the same coloring scheme throughout?}
 
@@
 
-\apt{or is in some function?? and why do we need either??}:
-\sna{I think Andrew is right that we don't need to know the function
-     here, or that the entry point is valid; that's part of WBCF.}
-\rb{Agreed, removed $f$}
 %and \(\mach_e\) is an entry point for some function \(f\):
@@
 
-{\em Stack integrity} is a property of a system with respect to a call map
-\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
-and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
-on the prefix up to its return, if any. For any initial state \(\mach_0\)
-we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
-\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
-\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
-the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
-\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\),
-\(\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
-system as a whole enjoys stack integrity with respect to \(\callmap\).
-
-\apt{Some more words needed here to smooth the technicalities.}
-\sna{Attempt...}
+We say that a system enjoys {\em stack integrity} if for every initial
+state $\mach_0$ and its corresponding induced trace $\MPT$ (with
+$\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT$), for every
+$n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
+$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, if we take the prefix of the
+call until its corresponding return $\mpstate{\mach_e}{\pol_e}
+\hookrightarrow \MPT_{call} ~ | ~ \ret{\mach_c}$, then $\MPT_{call}$ must
+satisfy trace integrity:
+\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]
+
+%{\em Stack integrity} is a property of a system with respect to a call map
+%\(\callmap\) that holds when every \(n\)-transition  in every legal trace exhibits trace integrity
+%on the prefix up to its return, if any. For any initial state \(\mach_0\)
+%we take the induced trace. For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
+%\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
+%the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
+%\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\),
+%\(\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
+%system as a whole enjoys stack integrity with respect to \(\callmap\).
+% 
+%\apt{Some more words needed here to smooth the technicalities.}
+%\sna{Attempt...}
+%Rewrote for clarity
 
@@
 which shows a snapshot of the state as in the running example (top
-memory layout) and a variation of it (bottom). \leo{Finalize
-  contents, describe}
+memory layout) and a variation of it (bottom). 
 %
@@
 separated by a slash.
+%
+This example will end up a violation of confidentiality, as a
+difference in a confidential location (\textcolor{cyan}{42} vs
+\textcolor{magenta}{5} in location 100) leads to a visible
+difference in return value.
 
@@
 
-\sna{Interestingly, split out like this, we don't need to talk about
-     contour here -- they are varied in the next property.}
+%\sna{Interestingly, split out like this, we don't need to talk about
+%     contour here -- they are varied in the next property.}+1
 \begin{itemize}
@@
 had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
-\rb{Color name macros}
-\apt{Explain slash notation in effects column.}
-\rb{Wrote some words to explain after def. 5.4 (variations).}
-
-As another example,
-changing line 30 of the running example to $r_5  \leftarrow [\rsp - 6]$ would cause
-{\tt g} to read the value of {\tt main}'s local variable {\tt x} and then immediately
-output it, violating the internal observation clause of the confidentiality definition.
+%\rb{Color name macros}
+%\apt{Explain slash notation in effects column.}
+%\rb{Wrote some words to explain after def. 5.4 (variations).}
+
+As another example, changing instruction 30 of the running example to
+$r_5 \leftarrow [\rsp - 6]$ would cause {\tt g} to read the value of
+{\tt main}'s local variable {\tt x} and then immediately output it,
+violating the internal observation clause of the confidentiality
+definition.
 
@@
 %Let $\callmap$ be a call map and $\codemap$ a function map.
-We will say that a system satisfies {\em stack confidentiality} with respect to
-\(\callmap\) in a similar way to stack integrity. For every initial state
-\(\mach_0\), we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)}
-\hookrightarrow \MPT\). Then for any \(n\)-transition
-\(\mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~
-\mpstate{\mach_e}{\pol_e}\), we take the prefix trace until return:
-%
-\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} | \ret{\mach_c}\]
-
+For every initial state \(\mach_0\), we take the induced trace
+\(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\). Then for
+any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
+$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, we take the prefix of the trace
+from the entry point until its corresponding return:
+%
+\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
+%
 We also take the corresponding trace from any variant \(\mach_e'
@@
 %
-\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]
-
-If, for all such pairs of traces, \(\propdef{TraceConf} ~
-\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
-stack confidentiality with respect to \(\callmap\).
-
-\apt{oy. still needs work to make this intelligible}
-\sna{Attempted. There's a tension between introducing all the variables
-     and keeping a good flow that I'm not sure I'm managing here though.}
+\[\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
+%
+If trace confidentiality holds for all such pairs of traces,
+\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT ~ \machT\), then the
+system as a whole enjoys {\em stack confidentiality} with respect to
+\(\callmap\).
+
+%\apt{oy. still needs work to make this intelligible}
+%\sna{Attempted. There's a tension between introducing all the variables
+%     and keeping a good flow that I'm not sure I'm managing here though.}
+%Added even more
 
@@
 if, for any initial state \(\mach_0\)
-and any n-transition \((\mach_c,\pol_c)\) and
+and any n-transition from \((\mach_c,\pol_c)\) to
 \((\mach_e, \pol_e)\) in the induced trace from

### 2020-07-09 main.tex
Nit

--- 2020-07-09 main.tex
@@
 existing stack protection mechanisms and describe some useful
-provably stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
+stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
 \leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi

### 2020-07-09 main.tex
Fixes to properties I noticed while doing coq

--- 2020-07-09 main.tex
@@
 What is stack safety? Numerous mechanisms have been proposed to guard
-against attacks that corrupt the call stacks of running programs,
-\ifaftersubmission
-and many
+against attacks that corrupt the call stacks of running programs, and many
 others rely on a protected stack as an assumption to provide higher level
 guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
-  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
-\bcp{Let's delete for now; later, we can try to incorporate it in the body
-  of the paper.}
-\fi,
-but a clear and precise definition of stack safety is missing
+  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}, but a clear and precise definition of stack safety is missing
 from the literature.
-%
-We propose a formal definition phrased in the technical
+
+We propose a formal definition of stack safety phrased in the technical
 terminology of language-based security---as a combination of a
@@
 existing stack protection mechanisms and describe some useful
-stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
-\leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi
+provably stronger and weaker variants.  \bcp{should we mention stktokens here?}
+\leo{Not sure. We almost never cite stuff in the abstract, do we?}
 \end{abstract}
@@
 capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\leo{Do we want to hunt for more refs here?}
-\bcp{Sure, but later!}
-\fi
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \iflater\apt{Mostly from
+  nick; there could be more}\leo{Do we want to hunt for more refs here?}\fi
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
-  A new {\em formal stack-safety property} capturing the
+  A novel formal stack safety property capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
   % These properties   rule out common examples of stack memory abuse.
-  This property can be enforced by
-  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
-  based on hardware tagging~(\cref{sec:conservative}).
+  This property is enforced by 
+  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:conservative}).
 \item
-  A {\em lockstep} variant of the stack-safety property that is also
-  enforced by the Depth Isolation policy
-  but is more efficient to test, plus a proof (in Coq) that this variant is
-  stronger than the original~(\cref{sec:lockstep}).
+  A lockstep stack safety property that is also enforced by the Depth Isolation policy,
+  and is more efficient to test, which we prove (in Coq) is stronger than the basic property~(\cref{sec:lockstep}).
 \item
-  A {\em lazy} variant of the stack-safety property that can be enforced by the
-  more efficient Lazy Tagging and Clearing policy
-  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
-  variant is weaker than the original~(\cref{sec:lazy}).
+  A novel lazy stack safety property that is enforced by the more efficient
+  Lazy Tagging and Clearing policy
+  of \citet{DBLP:conf/sp/RoesslerD18}, which we prove (in Coq) is weaker than the basic property~(\cref{sec:lazy}).
 %  This policy does not enforce our strict
@@
 \cref{sec:future} sketches directions for future work.
-
+%
 We begin with a description of our threat model
@@
 
-\apt{Some more words needed here to smooth the technicalities.}
-\sna{Attempt...}
-
 \subsubsection{Stack Confidentiality}
@@
 
-\sna{Interestingly, split out like this, we don't need to talk about
-     contour here -- they are varied in the next property.}
 \begin{itemize}
@@
 
-If, for all such pairs of traces, \(\propdef{TraceConf} ~
-\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
+For all such pairs of traces, \(\propdef{TraceConf} ~
+\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
+Trace confidentiality must also hold on the trace as a whole, so that
+unallocated memory is still secret. Let \(\mach_0'
+\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
+\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
+
+If these conditions hold, then the system as a whole enjoys
 stack confidentiality with respect to \(\callmap\).
-
-\apt{oy. still needs work to make this intelligible}
-\sna{Attempted. There's a tension between introducing all the variables
-     and keeping a good flow that I'm not sure I'm managing here though.}
 
@@
 
-      The \(\mathit{restore}\) function takes a contour and four states: the
-      original call state, a variant call state, and corresponding return
-      states, and returns another state. The resulting state restores those
-      components that are high confidentiality to match the original call,
+      The \(\mathit{restore}\) function takes four states: the
+      original entry state, a variant entry state, and corresponding return
+      states. It returns another state. The resulting state restores those
+      components that varied between entries to match the original entry,
       unless they were changed during execution:
-      \[\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
+      \[\mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')
         (\component) =
         \begin{cases}
-          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_)
-            \text{ and } \\
+          \mach_e(\component) & \text{if } \mach_e(\component) \not =
+                                \mach_e'(\component) \text { and } \\
           & \mach_e(\component) = \mach_r(\component) \text{ and }
@@
       confidentiality, and reuses the cases in which the traces don't return.
-      For a contout \(\contour\), return predicate \(R\), and a pair of traces
+      For a return predicate \(R\) and a pair of traces
       \(\MPT\) and \(\machT\), we consider the return case, in which
@@
       based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
-      \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')\).
+      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).
       The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
@@
       \[\begin{split}
-        \propdef{ObsTraceConf} ~ \contour ~ R ~ \MPT ~ \machT \triangleq
-        & \mathit{ObsReturnCase} ~ \contour ~ R ~ \MPT ~ \machT \land \\
+        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \triangleq
+        & \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \land \\
         & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
@@
       confidentiality. Let \(\callmap\) be a callmap; for each initial
-      state and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
-      to \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
-      \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
-      and \(\mach_e' \hookrightarrow \machT | \ret{\mach_c}\). A system
-      enjoys observable stack confidentiality if, for all such
-      \(\MPT\) and \(\machT\), \(\propdef{ObsTraceConf} ~
-      \mathit{Cof}(\mach_c,n) ~ \ret{\mach_c} ~ \MPT ~ \machT\).
+      state \(\mach_0\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
+      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
+      \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\), take any
+      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
+      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{call} |
+      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
+      \ret{\mach_c} ~ \MPT ~ \machT\).
+
+      Additional, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
+      \mach_0\), taking the induced trace  \(\mach_0' \hookrightarrow
+      \machT\), we must have \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\).
+
+      A system enjoys observable stack confidentiality if observable
+      confidentiality holds in all such cases.
 

### 2020-07-10 main.tex
Rest of 5 + 6

--- 2020-07-09 main.tex
@@
 existing stack protection mechanisms and describe some useful
-stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
+provably stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
 \leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi
@@
 Both the integrity and confidentiality components of local state
-encapsulation concern themselves only with the accessibility of data, and
-may still hold on a wide range of programs that violate expected control
-flow. Say, if {\tt g} managed to jump in the middle of the
-code of {\tt h} with its call, confidentiality and integrity would
+encapsulation concern themselves only with the accessibility of data,
+and may still hold on a wide range of programs that violate expected
+control flow. For instance, if {\tt g} managed to jump in the middle of
+the code of {\tt h} with its call, confidentiality and integrity would
 still be satisfied.
@@
 %
-Informal definitions of stack safety usually \apt{evidence?} aim to prevent such unwanted
-flows: every jump between functions should either be a call or a
-return, and each return should be to the instruction after its
-corresponding call.
+Stack safety enforcement mechanisms like \citet{Skorstengaard+19}
+usually aim to prevent such unwanted flows: every jump between
+functions should either be a call or a return, and each return should
+be to the instruction after its corresponding call.
 
@@
 sources between functions, {\em entry integrity} ensures the validity
-of call targets, and {\em return integrity} ensures the validity of
+of call targets, and {\em return integrity} ensures that of
 return targets.
@@
 or a return \(\retmap(\mach_1(\PCname))\).
+
+\smallskip
 
@@
 \end{itemize}
+\ifspace
 \apt{Could cut formal definition for space.}
+\fi
 
@@
 %% \rb{Come back to this later?}.
-Initially, the entire stack is tagged as unused with an $\tagNoDepth$ tag
-\rb{actually, the ``initial frame'' should be tagged with its proper depth}, and
-the {\PCname} carries in its tag the depth of the current activation at all
-times, starting with a $\tagPCDepth{0}$ tag. The current activation depth needs
-to change exactly at those points in the program where the contour of the stack
-safety property changes:
+Initially, the entire stack is tagged as unused with an $\tagNoDepth$ tag,
+%\rb{actually, the ``initial frame'' should be tagged with its proper depth}
+and the {\PCname} carries in its tag the depth of the current
+activation at all times, starting with a $\tagPCDepth{0}$ tag. The
+current activation depth needs to change exactly at those points in
+the program where the contour of the stack safety property changes:
 
@@
 sequences of calls declared in the call map are executed correctly.
-
-% If anything overwrites the SP in between, the exit sequence will fail, which is safe.
-
+%
 During normal execution, the {\PCname} uses the current activation
@@
 tagged $\tagStackDepth{n}$.
-\rb{Note that we have also considered and implemented more flexible schemes
-  based on allowing operations above the current activation depth. How do we put
-  these together?}
-\rb{Do we want to discuss variations involving frame pointers or more implicit
-  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
-\rb{Where to talk about the role of the compiler?}
+%\rb{Note that we have also considered and implemented more flexible schemes
+%  based on allowing operations above the current activation depth. How do we put
+%  these together?}
+%\rb{Do we want to discuss variations involving frame pointers or more implicit
+%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
+%\rb{Where to talk about the role of the compiler?}
 
@@
 but also a provably (in Coq) stronger one that enforces local state
-encapsulation throughout every call's execution. To encode that
+encapsulation throughout the execution. To encode that
 property, we not only keep track of the execution of one machine
@@
 %%
-\iflater
-\leo{This can be elided}
-  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
-  respect to it if the stack is nonempty, the return map of the bottom element
-  is the false property, and for each element $\vse \in \vs$:
-  \begin{itemize}
-  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
-  \item $\mach$ is in the trace $\machT$ of its initial machine,
-    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
-  \item Its variant state $\mathit{curr\_variant}(\vse)$
-    is in the trace of its initial variant,
-    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
-  \item The return function is in the return map if not the last element. \rb{TODO}
-  \item The registers of its contour $\contour(\vse)$ are public.
-  \end{itemize}
-  \fi
+%\iflater
+%\leo{This can be elided}
+%  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
+%  respect to it if the stack is nonempty, the return map of the bottom element
+%  is the false property, and for each element $\vse \in \vs$:
+%  \begin{itemize}
+%  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
+%  \item $\mach$ is in the trace $\machT$ of its initial machine,
+%    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
+%  \item Its variant state $\mathit{curr\_variant}(\vse)$
+%    is in the trace of its initial variant,
+%    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
+%  \item The return function is in the return map if not the last element. \rb{TODO}
+%  \item The registers of its contour $\contour(\vse)$ are public.
+%  \end{itemize}
+%  \fi
 
@@
 
-\paragraph*{THEOREM}
+\medskip
+\textsc{Theorem.}
 Lockstep stack safety implies stack safety.
-
-\paragraph*{PROOF}
-By coinduction on trace lockstep stack safety.
-\leo{3000 lines Coq?}
-\leo{I can write a TON about this proof, but I'm not sure how much is interesting/how much space we have}.
+\smallskip
+
+The proof of this theorem is a fairly involved coinductive proof
+(around 3K LoC in Coq) on the definition trace lockstep stack safety,
+using an auxilliary depth-index to ensure that proper aligmnent of
+calls and returns.
 

### 2020-07-10 main.tex

--- 2020-07-09 main.tex
@@
 
-%\sna{Interestingly, split out like this, we don't need to talk about
-%     contour here -- they are varied in the next property.}+1
 \begin{itemize}
@@
 %
-\[\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
-%
-If trace confidentiality holds for all such pairs of traces,
-\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT ~ \machT\), then the
+\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]
+
+For all such pairs of traces, \(\propdef{TraceConf} ~
+\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
+Trace confidentiality must also hold on the trace as a whole, so that
+unallocated memory is still secret. Let \(\mach_0'
+\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
+\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
+
+If trace confidentiality holds for all such traces,
+\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
+\machT_{\mathit{call}}\) and
+\(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
 system as a whole enjoys {\em stack confidentiality} with respect to
 \(\callmap\).
-
-%\apt{oy. still needs work to make this intelligible}
-%\sna{Attempted. There's a tension between introducing all the variables
-%     and keeping a good flow that I'm not sure I'm managing here though.}
-%Added even more
 

### 2020-07-10 main.tex
Nits

@@
 others rely on a protected stack as an assumption to provide higher level
-guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
+guarantees,\bcp{this is a good point, but we don't pick it up anywhere in the
   paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
@@
   of the paper.}
-\fi,
+\fi
 but a clear and precise definition of stack safety is missing
@@
 positive definition, by
-\citet{Skorstengaard+19}, who define stack safety as
+\citet{Skorstengaard+19}, which defines stack safety as
 the conjunction of two properties: {\em local state encapsulation} and {\em
   well-bracketed control flow}.
-They formalize these properties by defining an abstract capability-based machine
+These properties are formalized by defining an abstract capability-based machine
 that ``satisfies [stack safety] by construction.''
 %
-While this represents an important step in the right direction, we believe
+While this work represents an important step in the right direction, we believe
 there is still room to craft a more abstract, more generic characterization
@@
 %
-Our key contribution is such a novel and generic presentation of
-local state encapsulation and well-bracketed control flow using the
+Our key contribution is such a generic presentation of
+local state encapsulation and well-bracketed control\bcp{we need to either
+  weaken the previous sentence or else add ``first steps toward'' in this sentence...} flow using the
 technical framework of language-based security.
@@
 policy does not enforce our basic propety, but an enhanced version of
-it does enforce a proveably weaker property that still protects from
+it does enforce a provably weaker property that still protects from
 unsafe callees, but is less intuitive.

### 2020-07-10 main.tex
Issue

--- 2020-07-10 main.tex
@@
 
+%\sna{Interestingly, split out like this, we don't need to talk about
+%     contour here -- they are varied in the next property.}+1
 \begin{itemize}
@@
 %
-\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]
-
-For all such pairs of traces, \(\propdef{TraceConf} ~
-\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
-Trace confidentiality must also hold on the trace as a whole, so that
-unallocated memory is still secret. Let \(\mach_0'
-\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
-\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
-
-If trace confidentiality holds for all such traces,
-\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
-\machT_{\mathit{call}}\) and
-\(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
+\[\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
+%
+If trace confidentiality holds for all such pairs of traces,
+\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT ~ \machT\), then the
 system as a whole enjoys {\em stack confidentiality} with respect to
 \(\callmap\).
+
+%\apt{oy. still needs work to make this intelligible}
+%\sna{Attempted. There's a tension between introducing all the variables
+%     and keeping a good flow that I'm not sure I'm managing here though.}
+%Added even more
 
@@
 
-      The \(\mathit{restore}\) function takes four states: the
-      original entry state, a variant entry state, and corresponding return
-      states. It returns another state. The resulting state restores those
-      components that varied between entries to match the original entry,
+      The \(\mathit{restore}\) function takes a contour and four states: the
+      original call state, a variant call state, and corresponding return
+      states, and returns another state. The resulting state restores those
+      components that are high confidentiality to match the original call,
       unless they were changed during execution:
-      \[\mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')
+      \[\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
         (\component) =
         \begin{cases}
-          \mach_e(\component) & \text{if } \mach_e(\component) \not =
-                                \mach_e'(\component) \text { and } \\
+          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_)
+            \text{ and } \\
           & \mach_e(\component) = \mach_r(\component) \text{ and }
@@
       confidentiality, and reuses the cases in which the traces don't return.
-      For a return predicate \(R\) and a pair of traces
+      For a contout \(\contour\), return predicate \(R\), and a pair of traces
       \(\MPT\) and \(\machT\), we consider the return case, in which
@@
       based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
-      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).
+      \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')\).
       The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
@@
       \[\begin{split}
-        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \triangleq
-        & \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \land \\
+        \propdef{ObsTraceConf} ~ \contour ~ R ~ \MPT ~ \machT \triangleq
+        & \mathit{ObsReturnCase} ~ \contour ~ R ~ \MPT ~ \machT \land \\
         & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
@@
       confidentiality. Let \(\callmap\) be a callmap; for each initial
-      state \(\mach_0\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
-      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
-      \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\), take any
-      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
-      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{call} |
-      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
-      \ret{\mach_c} ~ \MPT ~ \machT\).
-
-      Additional, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
-      \mach_0\), taking the induced trace  \(\mach_0' \hookrightarrow
-      \machT\), we must have \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\).
-
-      A system enjoys observable stack confidentiality if observable
-      confidentiality holds in all such cases.
+      state and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
+      to \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
+      \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
+      and \(\mach_e' \hookrightarrow \machT | \ret{\mach_c}\). A system
+      enjoys observable stack confidentiality if, for all such
+      \(\MPT\) and \(\machT\), \(\propdef{ObsTraceConf} ~
+      \mathit{Cof}(\mach_c,n) ~ \ret{\mach_c} ~ \MPT ~ \machT\).
 

### 2020-07-10 main.tex

--- 2020-07-10 main.tex
@@
 others rely on a protected stack as an assumption to provide higher level
-guarantees,\bcp{this is a good point, but we don't pick it up anywhere in the
+guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
   paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
@@
   of the paper.}
-\fi
+\fi,
 but a clear and precise definition of stack safety is missing
@@
 policy does not enforce our basic propety, but an enhanced version of
-it does enforce a provably weaker property that still protects from
+it does enforce a proveably weaker property that still protects from
 unsafe callees, but is less intuitive.

### 2020-07-10 main.tex
Nits on 2

@@
 can perform an action that would violate the security property; our
-properties are naturally therefore \emph{termination insensitive.}
+properties are therefore naturally \emph{termination insensitive.}
 This model will be made more precise in \cref{sec:prelim}.

### 2020-07-10 main.tex
CP

--- 2020-07-10 main.tex
@@
 
-The intuition behind \citeauthor{Skorstengaard+19}'s use of the term {\em
+The intuition behind \citeauthor{Skorstengaard+19}'s {\em
   local state encapsulation} is that callers should be protected from
@@
 itself: a callee should pass control back to the caller only by
-returning in the standard way (transferring control to the instruction
-following the call); it should not be allowed to jump back into the
+transferring control to the instruction
+following the call; it should not be allowed to jump back into the
 caller's code at an arbitrary point, either directly (via a jump
@@
   contributing here. We are not trying to phrase this as a security
-  property.}\leo{Better now? Should this be fixed two paragraphs above too?}
+  property.}\leo{Better now? Should this be fixed two paragraphs above
+  too?}\bcp{Hmmm this seems like a fundamental issue with the way we are
+  framing the story...}
 
@@
 protected from {\em its} callees, and so on. That is, the notion of
-``attacker'' changes throughout the execution of the program.  We address
-this issue using a dynamically changing notion of {\em accessibility} for
+``attacker'' changes at many points throughout the execution of the program.
+We address
+this issue by introducing a dynamically changing notion of {\em accessibility} for
 the different components of the machine state and adjusting the integrity
@@
 
-After formally defining stack safety, we show that our definition is
+After formally defining stack safety, we show that the definition is
 enforceable using an existing tag-based enforcement
-mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
+mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
+their\ifaftersubmission\bcp{Ugh}\fi{}
 conservative Depth Isolation policy, which not only enforces our basic
@@
 checks for integrity and confidentiality violations at every step of
-an execution trace. Thus, this property fails faster and can lead to
-more efficient testing. We prove (in Coq) that this lockstep property
-implies our basic definition of stack safety.
+an execution trace. The advantage of this property is that it ``fails
+faster'' and os therefore more suitable for efficient testing. We prove (in
+Coq) that this ``lockstep'' property implies our original definition of stack
+safety.
 

### 2020-07-10 main.tex
Nits in 3

@@
 (ordinary data) with base and bounds information (policy data),
-and their use requires rewriting source code (e.g. to use capabilities in place
+and their use requires rewriting source code (e.g., to use capabilities in place
 of ordinary pointers), so separating ordinary machine state from policy
@@
 written in C-like pseudo-code.  It consists of a {\tt main} function that
-performs a call to some function {\tt f}, which in turns makes a nested call
+performs a call to some function {\tt f}, which in turn makes a nested call
 to some function {\tt g} before returning, while performing some simple
-arithmetic operations in the process. \cref{fig:running-trace-b} shows an
+arithmetic operations in the process. \Cref{fig:running-trace-b} shows an
 execution trace of corresponding machine instructions. (These might have
@@
 First, since our goal is to formalize stack safety, we ignore the heap
-and assume that all variables are stack allocated. As noted in the
+and assume that all variables are stack-allocated. As noted in the
 previous section, the extension to account for memory safety is
@@
 A sketch of the initial memory layout appears at the top of
-\cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
+\Cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
 the dedicated output is uninitialized, while the (also uninitialized)
@@
 %
-The rest of \cref{fig:running-trace-b} traces the execution of the
+The rest of \Cref{fig:running-trace-b} traces the execution of the
 program, showing the machine state at call and return points.
@@
 Register $r_0$ is hardwired to zero (so that, e.g., instruction 3 stores $0$
-into address 101.).
+into address 101).
 

### 2020-07-10 main.tex
CP

--- 2020-07-10 main.tex
@@
 mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
-their\ifaftersubmission\bcp{Ugh}\fi{}
+\citeauthor{DBLP:conf/sp/RoesslerD18}'s
 conservative Depth Isolation policy, which not only enforces our basic
@@
 
-However, at the same time the conservative enforcement mechanism is
+The downside of the Depth Isolation policy is that it is rather
 expensive to implement. In the same work,
 \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
-that is more efficient by virtue of being lazy in its enforcement. At
-a high level, instead of guaranteeing integrity and confidentiality
-when a callee returns, this policy checks for violations when the
-caller actually makes use of (potentially compromised) data. The lazy
-policy does not enforce our basic propety, but an enhanced version of
-it does enforce a provably weaker property that still protects from
-unsafe callees, but is less intuitive.
+that is more efficient by virtue of being lazy in its enforcement.
+Instead of enforcing the integrity and confidentiality of the stack at the
+exact moment
+when a callee returns, this policy signals violations when the
+caller actually uses data that the callee has written into  its stack
+frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
+  property inthe same way...}\fi.
+The lazy
+policy does not enforce our original stack-safety property, but (a slightly
+enhanced variant of the policy\bcp{I worry that we are being a little
+  dishonest here: the stated advantage of the lazy policy is that it is more
+efficient, but our variant is {\em not} efficient!}) of it does enforce a slightly weaker
+lazy version of the property that still protects from unsafe callees but is
+less intuitive to state.
 
 In summary, we offer the following contributions:
-%\bcp{I would still like to
-%  see these points phrased in ``contribution normal form'': say exactly what
-%  we did for each one (it's a bit vague---''considers,'' etc.), and
-%  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
-% +1 on Andrew's rewrite
 \begin{itemize}
@@
   intuitive concepts of strict local state encapsulation and well-bracketed
-  control flow~(\cref{sec:lse-and-wbcf}) in terms of security concepts.
+  control flow~(\cref{sec:lse-and-wbcf}) in terms of familiar security concepts.
   % These properties   rule out common examples of stack memory abuse.

### 2020-07-10 main.tex
Pass complete

--- 2020-07-10 main.tex
@@
 positive definition, by
-\citet{Skorstengaard+19}, which defines stack safety as
+\citet{Skorstengaard+19}, who define stack safety as
 the conjunction of two properties: {\em local state encapsulation} and {\em
   well-bracketed control flow}.
-These properties are formalized by defining an abstract capability-based machine
+They formalize these properties by defining an abstract capability-based machine
 that ``satisfies [stack safety] by construction.''
 %
-While this work represents an important step in the right direction, we believe
+While this represents an important step in the right direction, we believe
 there is still room to craft a more abstract, more generic characterization
@@
 %
-Our key contribution is such a generic presentation of
-local state encapsulation and well-bracketed control\bcp{we need to either
-  weaken the previous sentence or else add ``first steps toward'' in this sentence...} flow using the
+Our key contribution is such a novel and generic presentation of
+local state encapsulation and well-bracketed control flow using the
 technical framework of language-based security.
 
-The intuition behind \citeauthor{Skorstengaard+19}'s {\em
+The intuition behind \citeauthor{Skorstengaard+19}'s use of the term {\em
   local state encapsulation} is that callers should be protected from
@@
 itself: a callee should pass control back to the caller only by
-transferring control to the instruction
-following the call; it should not be allowed to jump back into the
+returning in the standard way (transferring control to the instruction
+following the call); it should not be allowed to jump back into the
 caller's code at an arbitrary point, either directly (via a jump
@@
   contributing here. We are not trying to phrase this as a security
-  property.}\leo{Better now? Should this be fixed two paragraphs above
-  too?}\bcp{Hmmm this seems like a fundamental issue with the way we are
-  framing the story...}
+  property.}\leo{Better now? Should this be fixed two paragraphs above too?}
 
@@
 protected from {\em its} callees, and so on. That is, the notion of
-``attacker'' changes at many points throughout the execution of the program.
-We address
-this issue by introducing a dynamically changing notion of {\em accessibility} for
+``attacker'' changes throughout the execution of the program.  We address
+this issue using a dynamically changing notion of {\em accessibility} for
 the different components of the machine state and adjusting the integrity
@@
 
-After formally defining stack safety, we show that the definition is
+After formally defining stack safety, we show that our definition is
 enforceable using an existing tag-based enforcement
-mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
-\citeauthor{DBLP:conf/sp/RoesslerD18}'s
+mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
 conservative Depth Isolation policy, which not only enforces our basic
@@
 checks for integrity and confidentiality violations at every step of
-an execution trace. The advantage of this property is that it ``fails
-faster'' and os therefore more suitable for efficient testing. We prove (in
-Coq) that this ``lockstep'' property implies our original definition of stack
-safety.
-
-The downside of the Depth Isolation policy is that it is rather
+an execution trace. Thus, this property fails faster and can lead to
+more efficient testing. We prove (in Coq) that this lockstep property
+implies our basic definition of stack safety.
+
+However, at the same time the conservative enforcement mechanism is
 expensive to implement. In the same work,
 \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
-that is more efficient by virtue of being lazy in its enforcement.
-Instead of enforcing the integrity and confidentiality of the stack at the
-exact moment
-when a callee returns, this policy signals violations when the
-caller actually uses data that the callee has written into  its stack
-frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
-  property inthe same way...}\fi.
-The lazy
-policy does not enforce our original stack-safety property, but (a slightly
-enhanced variant of the policy\bcp{I worry that we are being a little
-  dishonest here: the stated advantage of the lazy policy is that it is more
-efficient, but our variant is {\em not} efficient!}) of it does enforce a slightly weaker
-lazy version of the property that still protects from unsafe callees but is
-less intuitive to state.
+that is more efficient by virtue of being lazy in its enforcement. At
+a high level, instead of guaranteeing integrity and confidentiality
+when a callee returns, this policy checks for violations when the
+caller actually makes use of (potentially compromised) data. The lazy
+policy does not enforce our basic propety, but an enhanced version of
+it does enforce a proveably weaker property that still protects from
+unsafe callees, but is less intuitive.
 
 In summary, we offer the following contributions:
+%\bcp{I would still like to
+%  see these points phrased in ``contribution normal form'': say exactly what
+%  we did for each one (it's a bit vague---''considers,'' etc.), and
+%  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
+% +1 on Andrew's rewrite
 \begin{itemize}
@@
   intuitive concepts of strict local state encapsulation and well-bracketed
-  control flow~(\cref{sec:lse-and-wbcf}) in terms of familiar security concepts.
+  control flow~(\cref{sec:lse-and-wbcf}) in terms of security concepts.
   % These properties   rule out common examples of stack memory abuse.
@@
 can perform an action that would violate the security property; our
-properties are therefore naturally \emph{termination insensitive.}
+properties are naturally therefore \emph{termination insensitive.}
 This model will be made more precise in \cref{sec:prelim}.
@@
 
-      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
-      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
-      agrees with \(\mach_e\) on every component the callee should not change,
-      but with \(\mach_r\) on those components that it is permitted to change.
-      So the observable behavior that follows is the idealized behavior to
-      which we compare the trace from \(\mach_r\), which must prefix it (to
-      account for premature halts due to policy violations.)
+      If \(\mach_e\) is an entry to a call with contour \(\contour\)
+      and \(\mach_r\) its return, the result of
+      \(\mathit{roll}(\contour,\mach_e,\mach_r)\) agrees with the
+      entry \(\mach_e\) on every component the callee should not
+      change, and with the actual return \(\mach_r\) on those
+      components that it is permitted to change.  So the observable
+      behavior that follows is the idealized behavior to which we
+      compare the trace from \(\mach_r\).
 
@@
       A trace enjoys {\em observable trace integrity} with respect to
-      some contour \(\contour\) if the trace induced by its final state
-      (if any) has observable behavior that prefixes that of its idealized
-      version, constructed by \(\mathit{roll}\):
+      some contour \(\contour\) if the trace induced by its final
+      state (if any) has observable behavior that prefixes that of its
+      idealized version (to account for premature halts due to policy
+      violations):
 
@@
     trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
-    it prints 0. So observable trace integrity does not hold.
+    it prints 0. Only then is observable trace integrity violated.
 
@@
 
-      Two of the cases of trace confidentiality, \(\mathit{InfCase}\) and
-      \(\mathit{FailCase}\), are purely properties of the observations produced
-      by the traces they are properties of. If the traces are infinite, they
-      should be similar; if one fail-stops, it should be a prefix. Only
-      the \(\mathit{ReturnCase}\) checks machine states, because in that case
-      the traces could continue, but have been cut off by a return. The key
-      to making confidentiality lazy is to consider the traces after the
-      return, considering only whether they are observationally similar
-      or prefixes.
+      Two of the cases of trace confidentiality, \(\mathit{InfCase}\)
+      and \(\mathit{FailCase}\), are purely properties of the
+      observations produced by the traces they are properties of. If
+      the traces are infinite, they should be similar; if one
+      fail-stops, it should be a prefix. Only the
+      \(\mathit{ReturnCase}\) checks machine states, because in that
+      case the traces could continue, but have been cut off by a
+      return. The key to making confidentiality lazy not perform that
+      check but rather to consider the traces after that return,
+      checking only whether they are observationally similar.
 
@@
 
-      \paragraph*{THEOREM} Stack safety implies observable stack safety.
-
-%      \paragraph*{PROOF} \leo{TODO: Whatever we say about the previous proof
- %       we have to say similar things here}
+      \medskip
+      \textsc{Theorem.}
+      Stack safety implies observable stack safety.
+      \smallskip
+
+      This is more straightforward (1500 LoC) as the recursive
+      structure of the two properties is similar.
 
@@
 %
-To our knowledge, the first and only line of work that has attempted to furnish
-a positive and formal characterization of the meaning of stack safety was
-inaugurated by \citet{Skorstengaard+19b}, who introduce a calling convention
-that uses local capabilities to preserve local state encapsulation and
-well-bracketed control flow, using a logical relation to reason about the stack
-safety of concrete programs. Although the required hardware support is readily
-available in capability machines like CHERI, this technique incurs potentially
-prohibitive performance costs, because it requires the entire unused part of the
-stack to be cleared whenever a security boundary is crossed. Their logical
-relation captures capability safety without ``externally observable side-effects
-(like console output or memory access traces).'' In the discussion, ``while [the
-authors] claim that [their] calling convention enforces control-flow
-correctness, [they] do not prove a general theorem that shows this, because it
-is not clear what such a theorem should look like,'' noting that the correctness
-property enjoyed by their technique ``is not made very explicit.''
+To our knowledge, the first and only line of work that has attempted
+to furnish a positive and formal characterization of the meaning of
+stack safety is \citet{Skorstengaard+19b}, who introduce a calling
+convention that uses local capabilities to preserve local state
+encapsulation and well-bracketed control flow, using a logical
+relation to reason about the stack safety of concrete
+programs. Although the required hardware support is readily available
+in capability machines like CHERI, this technique incurs significant
+costs, because it requires the entire unused part of the stack to be
+cleared whenever a security boundary is crossed. Their logical
+relation captures capability safety without ``externally observable
+side-effects (like console output or memory access traces).'' In the
+discussion, ``while [the authors] claim that [their] calling
+convention enforces control-flow correctness, [they] do not prove a
+general theorem that shows this, because it is not clear what such a
+theorem should look like,'' noting that the correctness property
+enjoyed by their technique ``is not made very explicit.''
 
@@
 \paragraph{Heap Safety as a Security Property}
-\bcp{Talk about the relation.}
-    Heap safety, like stack safety, may be framed as a form of
-    noninterference\citet{DBLP:conf/post/AmorimHP18}. The above model
-    describing safety of
-    shared stack-allocated objects extends naturally to heap safety,
-    complicated by the fact that these objects are not deallocated by returns,
-    and therefore a function's privilege may increase or decrease after its
-    entry point. For instance, if a callee allocates a heap object and returns
-    the pointer to its caller, that object's addresses become accessible in
-    contradiction to a contour computed at the caller's entry. So such an
-    extension must enforce trace confidentiality and integrity properties
-    separately on continuous segments of each call, with contours computed at
-    each crossing between caller and callee.
+
+Heap safety, like stack safety, may be framed as a security property
+in the form of noninterference~\citet{DBLP:conf/post/AmorimHP18}. Just
+like~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
+characterization of the meaning of (heap) memory safety, in this paper
+we aim to do the same for the stack. 
+%
+Their model describing safety of shared stack-allocated
+objects extends naturally to heap safety, complicated by the fact that
+these objects are not deallocated by returns, and therefore a
+function's privilege may increase or decrease after its entry
+point. For instance, if a callee allocates a heap object and returns
+the pointer to its caller, that object's addresses become accessible
+in contradiction to a contour computed at the caller's entry. So such
+an extension must enforce trace confidentiality and integrity
+properties separately on continuous segments of each call, with
+contours computed at each crossing between caller and callee.
 
@@
   describe the basic principles by which our model can be extended to some of
-  these cases.
+  these cases. We also plan to develop a testing framework for quickly
+  checking the validity of different enforcement mechanisms.
 
@@
 
-    Common programming idioms involve a caller passing a pointer to its local
-    data into a callee. Using standard notions of pointer provenance [cite],
-    we can extend our model to distinguish safe use of this idiom from true
-    stack safety violations. We require additional annotation to identify
-    which addresses within a function's stack frame correspond to distinct
-    objects. Then a valid pointer to an object is one derived through legal
-    arithmetic from its base address, and at each call point, an object's
-    addresses are marked low confidentiality and integrity if a valid pointer
-    exists in a register or in the transitive closure of the callee's
-    accessible memory. All pointers to an object cease to be valid when it is
-    deallocated along with its stack frame.
+    Common programming idioms involve a caller passing a pointer to
+    its local data into a callee. Using standard notions of pointer
+    provenance~\citep{provenance}, we can extend our model to
+    distinguish safe use of this idiom from true stack safety
+    violations. We require additional annotation to identify which
+    addresses within a function's stack frame correspond to distinct
+    objects. Then a valid pointer to an object is one derived through
+    legal arithmetic from its base address, and at each call point, an
+    object's addresses are marked low confidentiality and integrity if
+    a valid pointer exists in a register or in the transitive closure
+    of the callee's accessible memory. All pointers to an object cease
+    to be valid when it is deallocated along with its stack frame.
 
@@
     data are unchanged. Confidentiality properties are split similarly.
+
+  \paragraph{Random Testing}
+
+  Part of the motivation for the strong lockstep stack safety property
+  is that it uncovers errors quickly, and therefore should be quicker
+  to test just like the smarter testing properties
+  of~\citet{TestingNI:ICFP}. We have already developed a testing
+  framework for debugging our policy enforcement for this property, and
+  it would be interesting to explore whether a testing-amenable variant
+  of the lazy property also exists.
 

### 2020-07-10 main.tex
CP

@@
 \item
-  A {\em lazy} variant of the stack-safety property that can be enforced by the
+  A {\em lazy} variant of the stack-safety property that can be enforced by
+  (a variant of\bcp{same concern}) the
   more efficient Lazy Tagging and Clearing policy
@@
 
-We assume we are given a machine language program, together with certain auxiliary declarations
+We assume we are given a machine language program, together with some auxiliary information
 about its structure: the code footprint of each function and the locations of instructions
 representing calls, returns, and function entry points.
-Although this information would typically come from a compiler, we do not
-assume anything about the provenance of the machine
-code. We assume that the machine has a program counter $\PCname$
-and stack pointer $\rsp$ and can perform the usual
-loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
-
-To make our stack safety definitions precise, we make some simple assumptions about
+Although this information would typically come from a compiler together with
+the machine code, we do not
+assume anything about their provenance.  We assume that the machine has a
+program counter $\PCname$ 
+and stack pointer $\rsp$\bcp{but not a frame pointer?!}; other details of the machine's ISA are unimportant.
+
+To make our stack safety definitions precise, we make some straightforward
+assumptions about 
 memory layout; our definitions could easily be changed to be parametric over
-these assumptions, at the cost of greater complexity.  Memory
-consists of an instruction region and a stack that grows upwards; for simplicity,
-we omit a heap. (Integrating a memory safety property for the heap in the style of
+these assumptions, at the cost of some complexity in their statements.  Memory
+consists of an instruction region and a stack that grows upwards. (For simplicity,
+we do not consider the heap. Integrating a memory safety property for the
+heap in the style of 
 \citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
@@
 contiguously allocated per-function frames.
-The boundary between the frames of a caller and callee is given by the $\rsp$ register
+The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
 at the time of the call.  The stack can be used to
 pass arguments or return results in the usual way, by placing them
-at the top of the caller's frame. Functions can also communicate via registers; indeed, our
-definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
+at the top of the caller's frame. Callers and callees can also communicate
+via registers if they wish; indeed, our 
+definitions say nothing at all about registers (except $\PCname$ and
+$\rsp$). \bcp{This implicitly amounts to choosing a caller-saves register
+  convention, though, doesn't it?}
 
 Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code.  We define a call to be
-any instruction declared as such in the program.
+which are not fully explicit in machine code (the same instruction may be
+used in some places as part of a call or return sequence and in other places
+for other purposes).  We define a call to be
+any instruction declared as such in the program\bcp{in the call map, not the
+program, right?}.
 If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding

### 2020-07-10 main.tex
Pass on 7

@@
 (around 3K LoC in Coq) on the definition trace lockstep stack safety,
-using an auxilliary depth-index to ensure that proper aligmnent of
+using an auxiliary depth-index to ensure the proper aligmnent of
 calls and returns.
@@
       induced trace, we take the prefix of the call
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\)
-      and observable trace integrity holds on \(\MPT_{call}\) with
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} | \ret{\mach_c}\)
+      and observable trace integrity holds on \(\MPT_{\mathit{call}}\) with
       contour \(\mathit{Cof}(\mach_c,n)\).
@@
       case the traces could continue, but have been cut off by a
-      return. The key to making confidentiality lazy not perform that
+      return. The key to making confidentiality lazy is not to perform that
       check but rather to consider the traces after that return,
@@
       the example in \cref{fig:confex}, the varied final state
-      \(\mach_r'\) still has -5 as a return address for {\tt g}, so
+      \(\mach_r'\) still has $-5$ as a return address for {\tt g}, so
       the trace after that return will behave oddly in a way that
-      doesn't reflect a confidentiality violation.  That means we
+      does not reflect a confidentiality violation.  That means we
       cannot immediately compare traces from \(\mach_r\) and
@@
       {\em Observable trace confidentiality} is the lazy analog to trace
-      confidentiality, and reuses the cases in which the traces don't return.
+      confidentiality, and reuses the cases in which the traces do not return.
       For a return predicate \(R\) and a pair of traces
@@
       and \(\mach_r'\) agree on components, we have the following requirement.
+      \rb{This reads a little funny, ending this paragraph with the promise of a
+      requirement and starting the next one introducing a conclusion.}
 
@@
       and after returning from {\tt g}, {\tt main} prints 85. But in the trace
-      from \(\mach_r''\), 108 contains -4, and ultimately 38 is printed.
+      from \(\mach_r''\), 108 contains $-4$, and ultimately 38 is printed.
       So observable trace confidentiality does not hold. This example
       illustrates the weakness of observable confidentiality as a testing
-      property; it takes until the very end of the program to determine whether
+      property: it takes until the very end of the program to determine whether
       a tainted value actually influences observable behavior.
@@
       \(r_6 \leftarrow [\rsp-8]\). \(r_6\) would hold 42 in \(\mach_r\)
-      and -5 in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
+      and $-5$ in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
       never used otherwise, the traces of {\tt h} will enjoy observable trace
@@
       \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
-      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{call} |
+      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} |
       \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
@@
 
-      This is more straightforward (1500 LoC) as the recursive
+      This is more straightforward \rb{than what?} (1,5K LoC in Coq) as the recursive
       structure of the two properties is similar.

### 2020-07-10 main.tex
Nits

@@
 %% =======
-We also need two auxilliary notations that relate to variant stacks.
+We also need two auxiliary notations that relate to variant stacks.
 %
@@
 A system satisfies {\em lockstep stack safety} if for every initial machine $\mach$,
-the trace induced from $(\mach, \pinit{\mach})$ satisfies trace lockstep stack safety with
+the trace induced from $(\mach, \pinit(\mach))$ satisfies trace lockstep stack safety with
 respect to the empty stack.

### 2020-07-10 main.tex
Remove token appendix
Nit in 5

@@
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
-hold -5, in {\varied}.  After the last step, the return states \(\mach_r\) and
+hold $-5$, in {\varied}.  After the last step, the return states \(\mach_r\) and
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
 \(\mach_e'\). Addresses 108 and 109 have changed, shown in light {\low}, and while 109
-had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
+had changed in a consistent way (independent of the variation), 108 holds either 43 or $-4$, so the trace does not obey trace confidentiality.
 %\rb{Color name macros}
@@
 %% Appendix
-\appendix
-\section{Appendix}
-
-Text of appendix \ldots
+%\appendix
+%\section{Appendix}
+%Text of appendix \ldots
 

### 2020-07-10 main.tex
Address comment in intro

@@
 %
-Our key contribution is such a generic presentation of
-local state encapsulation and well-bracketed control\bcp{we need to either
-  weaken the previous sentence or else add ``first steps toward'' in this sentence...} flow using the
-technical framework of language-based security.
+Our key contribution is another important step towards this goal: a generic
+presentation of local state encapsulation and well-bracketed
+control\bcp{we need to either weaken the previous sentence or else add
+  ``first steps toward'' in this sentence...}\leo{Weakened enough?} flow using the technical
+framework of language-based security.
 

### 2020-07-10 main.tex
Moving on to section 3

--- 2020-07-10 main.tex
@@
 
-Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
+Enforcement mechanisms are modeled as {\em policies}\bcp{this is a somewhat
+  special use of the word ``policy'' that comes from the micropolicies
+  world.  We should include a comment about this to avoid confusing
+  readers.} that add some auxiliary
 policy state to the ordinary machine state
@@
 
-We model ordinary program state and policy enforcement state separately to make it
+We model ordinary program state and policy-enforcement state separately to make it
 easier to use well-established concepts from language-based security.
@@
 \emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable behavior.\leo{What citation should go here? Not the hyperproperties one?}
+\emph{varying} that data would not change the program's observable
+behavior.\leo{What citation should go here? Not the hyperproperties one?
+  Goguen and Meseguer?}
 In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, whereas any associated policy enforcement
+ordinary program data should be varied, while any associated policy
 state should be left alone; our model makes this distinction obvious.
@@
 %execution where data has been varied.
-We also introduce a notion of observational integrity that relies on the idea of
+We will also introduce a notion of {\em observational integrity} that relies on the idea of
 \emph{rolling back} program state; again, our model makes it easy to describe
-rollback of ordinary program data while leaving policy enforcement data unchanged.
+rollback of ordinary program data while leaving policy state alone.
 
@@
 
-This enforcement model is flexible enough to capture
-micropolicies~\citep{pump_oakland2015}, and we believe that it can
-also capture code-altering approaches (such as software-based
+This enforcement model directly describes
+micropolicies~\citep{pump_oakland2015}, and we believe that it is
+flexible enough to capture code-altering approaches (such as code rewriting
 or capability-based techniques).
@@
 %
-Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
-reference monitoring mechanism, which have been applied to stack safety by
-\citet{DBLP:conf/sp/RoesslerD18}.
+Micropolicies are a flexible tag-based, hardware-accelerated
+reference monitoring mechanism; they been applied to stack safety
+enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
 Here the policy state consists of metadata tags (e.g., identifying stack frames)
-attached to each value in memory or registers, and the policy step function
+attached to each value in memory or registers, and the policy step
+function\bcp{undefined!}
 checks that each machine operation obeys a set of rules on tags
@@
 %Other policies are implemented in hardware or built on hardware mechanisms.
-Capability machines like CHERI \citep{Woodruff+14} extend conventional
+Capability machines such as CHERI \citep{Woodruff+14} extend conventional
 architectures to support efficient and fine-grained control over memory
@@
 the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
+\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
+linear capabilities?}.  Capabilities package up pointers
 (ordinary data) with base and bounds information (policy data),
-and their use requires rewriting source code (e.g., to use capabilities in place
+and they require rewriting source code (e.g., to use capabilities in place
 of ordinary pointers), so separating ordinary machine state from policy
-state is delicate, but we conjecture it can be done with some effort.
-%\apt{But we conjecture it can be done...?}\leo{Just say that?}
+state is delicate; we conjecture it can be done with some effort.
 
@@
 Some policies are implemented in software, by modifying code in the
-compiler to (e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or
+compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
 to insert stack canaries~\citep{Cowan+98}. Such approaches should
@@
 \leo{Something like this?}
+\bcp{OK}
 %These approaches are represented as policies that
@@
 %% =======
-We also need two auxiliary notations that relate to variant stacks.
+We also need two auxilliary notations that relate to variant stacks.
 %
@@
 A system satisfies {\em lockstep stack safety} if for every initial machine $\mach$,
-the trace induced from $(\mach, \pinit(\mach))$ satisfies trace lockstep stack safety with
+the trace induced from $(\mach, \pinit{\mach})$ satisfies trace lockstep stack safety with
 respect to the empty stack.
@@
 (around 3K LoC in Coq) on the definition trace lockstep stack safety,
-using an auxiliary depth-index to ensure the proper aligmnent of
+using an auxilliary depth-index to ensure that proper aligmnent of
 calls and returns.
@@
       induced trace, we take the prefix of the call
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} | \ret{\mach_c}\)
-      and observable trace integrity holds on \(\MPT_{\mathit{call}}\) with
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\)
+      and observable trace integrity holds on \(\MPT_{call}\) with
       contour \(\mathit{Cof}(\mach_c,n)\).
@@
       case the traces could continue, but have been cut off by a
-      return. The key to making confidentiality lazy is not to perform that
+      return. The key to making confidentiality lazy not perform that
       check but rather to consider the traces after that return,
@@
       the example in \cref{fig:confex}, the varied final state
-      \(\mach_r'\) still has $-5$ as a return address for {\tt g}, so
+      \(\mach_r'\) still has -5 as a return address for {\tt g}, so
       the trace after that return will behave oddly in a way that
-      does not reflect a confidentiality violation.  That means we
+      doesn't reflect a confidentiality violation.  That means we
       cannot immediately compare traces from \(\mach_r\) and
@@
       {\em Observable trace confidentiality} is the lazy analog to trace
-      confidentiality, and reuses the cases in which the traces do not return.
+      confidentiality, and reuses the cases in which the traces don't return.
       For a return predicate \(R\) and a pair of traces
@@
       and \(\mach_r'\) agree on components, we have the following requirement.
-      \rb{This reads a little funny, ending this paragraph with the promise of a
-      requirement and starting the next one introducing a conclusion.}
 
@@
       and after returning from {\tt g}, {\tt main} prints 85. But in the trace
-      from \(\mach_r''\), 108 contains $-4$, and ultimately 38 is printed.
+      from \(\mach_r''\), 108 contains -4, and ultimately 38 is printed.
       So observable trace confidentiality does not hold. This example
       illustrates the weakness of observable confidentiality as a testing
-      property: it takes until the very end of the program to determine whether
+      property; it takes until the very end of the program to determine whether
       a tainted value actually influences observable behavior.
@@
       \(r_6 \leftarrow [\rsp-8]\). \(r_6\) would hold 42 in \(\mach_r\)
-      and $-5$ in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
+      and -5 in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
       never used otherwise, the traces of {\tt h} will enjoy observable trace
@@
       \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
-      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} |
+      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{call} |
       \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
@@
 
-      This is more straightforward \rb{than what?} (1,5K LoC in Coq) as the recursive
+      This is more straightforward (1500 LoC) as the recursive
       structure of the two properties is similar.

### 2020-07-10 main.tex
Squash another comment
More comment answers/squashing

@@
 for other purposes).  We define a call to be
-any instruction declared as such in the program\bcp{in the call map, not the
-program, right?}.
+any instruction declared as such in the compiler-provided call structure.
 If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
@@
 \emph{varying} that data would not change the program's observable
-behavior.\leo{What citation should go here? Not the hyperproperties one?
-  Goguen and Meseguer?}
+behavior~\citep{6234468}
 In applying this idea to a system that incorporates policy enforcement, only
@@
 attached to each value in memory or registers, and the policy step
-function\bcp{undefined!}
+function\bcp{undefined!}\leo{``partial policy step'' is explained informally above}
 checks that each machine operation obeys a set of rules on tags
@@
 \citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
-linear capabilities?}.  Capabilities package up pointers
+  linear capabilities?}\leo{Only the second one AFAIK. What would you
+  change to characterize it more accurately?}.  Capabilities package up pointers
 (ordinary data) with base and bounds information (policy data),
@@
 policy, but only for programs produced by the compiler.
-\apt{not sure what to say here}
-\leo{Something like this?}
-\bcp{OK}
 %These approaches are represented as policies that

### 2020-07-10 main.tex
Moving on from 3 to 4

--- 2020-07-10 main.tex
@@
 
+\ifaftersubmission
 \bcp{We really want readers to understand this trace.  I think we should
@@
   Or do you have something else in mind?}
+\bcp{I was thinking about just a few words, not a page.  But let's do this
+  later.}
+\fi
 
@@
 arithmetic operations in the process. \Cref{fig:running-trace-b} shows an
-execution trace of corresponding machine instructions. (These might have
+execution trace of the corresponding machine instructions. These might have
 been generated by compiling the high-level code, but it is only the behavior
 of the machine instructions that we care about
-here---\Cref{fig:running-trace-b} is our actual running example.)
+here---\Cref{fig:running-trace-b} is our actual example.
 
@@
 %
-First, since our goal is to formalize stack safety, we ignore the heap
-and assume that all variables are stack-allocated. As noted in the
-previous section, the extension to account for memory safety is
-discussed in~\Cref{sec:relwork}.
+First, as noted above, since our goal is to formalize stack safety, we
+ignore the heap
+and assume that all variables are stack allocated.
 %
@@
 \Cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
-the dedicated output is uninitialized, while the (also uninitialized)
-stack begins at 100. Directly below the layout, we summarize the parts
-of the register file that are relevant to the example.
+and the (uninitialized)
+stack begins at 100. Directly below the memory sketch, we summarize the parts
+of the register file that are relevant at this point in the example.
 %
@@
 Instructions are intended to
-be similar to RISC-V, but are described in an ad-hoc notation
+be similar to RISC-V, but are described in a (hopefully more readable) an
+ad-hoc notation
 rather than using assembler syntax. The $\jal$ (jump-and-link)
-instruction used for calls atomically saves the return address in a register
+instruction used for calls atomically simultaneously the return address in a
+register
 and jumps to a specified new $\PCname$; the $\jalr$ (jump-and-link-via-register)
 instruction used for returns jumps to the address stored in a register.
-Register $r_0$ is hardwired to zero (so that, e.g., instruction 3 stores $0$
-into address 101).
+Register $r_0$ is hardwired to zero, so that, e.g., instruction 3 stores $0$
+into address 101.
 
@@
 
-In this section we describe our machine model. Our definition of stack
-safety is generic and this section discusses an abstract interface to
-our machine model and some kind of static or dynamic, software- or
-hardware-enforced security monitor. We will refer back to the running
-example of \cref{sec:running-example} for concreteness, and then,
-in~\cref{sec:enforcement}, extend this model with a dynamic policy
-enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
+We next describe our machine model. To make our definition of stack safety
+as generic as possible, this section proposes an abstract interface to a
+machine model plus some kind of (static or dynamic, software- or
+hardware-enforced) policy monitor. In~\cref{sec:enforcement}, we will
+instantiate this model with a concrete dynamic policy enforcement mechanism
+based on \citet{DBLP:conf/sp/RoesslerD18}.
+
+We will often refer to the running example of \cref{sec:running-example} for
+concreteness.
 
@@
 
-A machine state is a map from components to values:
+A {\em machine state} is a map from components to values:
 %

### 2020-07-10 main.tex
Comments, polishing

@@
 
-We will often refer to the running example of \cref{sec:running-example} for
-concreteness.
-
 \subsection{Values, States, and Observations}
@@
 pair of a machine state and an {\em observation}, which may be either
-silent (written $\tau$) or a word of data that the machine
+silent (written $\tau$) or a word of data\bcp{This is a rather concrete
+  choice!  Why not leave the set of observations abstract?} that the machine
 communicates to the outside world:
@@
 %
-This step function is total---execution traces are infinite unless cut
-short externally (e.g. by a policy as we'll see below).
-
-Notice that we choose to model observations explicitly, rather than simply treating
+This step function is total: execution traces are infinite unless
+cut short externally by a policy fault (as explained below).
+
+We choose to model observations explicitly, rather than simply
+treating
 a subset of registers or memory as being observable.  This choice gives
@@
 set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
-\cup \{\PCname,\SP\}$. Observations are the writes to the
-  special $\rout$ location, which can be thought of as a memory-mapped
+\cup \{\PCname,\SP\}$. Observations are writes to the
+special $\rout$ location\bcp{Which is 100! (do we need a name for it? seems
+  not...)}, which can be thought of as a memory-mapped
   output port: if during an execution step a word $w$ is
@@
 
-We now present our model of enforcement, introduced informally in \Cref{sec:threat}.
+We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
 A policy is a general model of an enforcement mechanism, consisting of a
@@
 \apt{Is this our current thinking?}
-\leo{I thought this was what we settled on in the last two days' discussions}
-%
-A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be given in
+\leo{I thought this was what we settled on in the last two days'
+  discussions}\bcp{That was not my impression!  I don't see how the policy
+  state can be recovered from just the machine state---e.g., the machine
+  state doesn't say where the calls and returns are, does it?!}
+%
+A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
 \Cref{sec:enforcement}.

### 2020-07-10 main.tex
Add comment
work through 6.1

--- 2020-07-10 main.tex
@@
 %
-Our key contribution is another important step towards this goal: a generic
-presentation of local state encapsulation and well-bracketed
-control\bcp{we need to either weaken the previous sentence or else add
-  ``first steps toward'' in this sentence...}\leo{Weakened enough?} flow using the technical
-framework of language-based security.
-
-The intuition behind \citeauthor{Skorstengaard+19}'s {\em
+Our key contribution is such a generic presentation of
+local state encapsulation and well-bracketed control\bcp{we need to either
+  weaken the previous sentence or else add ``first steps toward'' in this sentence...} flow using the
+technical framework of language-based security.
+
+The intuition behind \citeauthor{Skorstengaard+19}'s use of the term {\em
   local state encapsulation} is that callers should be protected from
@@
 \item
-  A {\em lazy} variant of the stack-safety property that can be enforced by
-  (a variant of\bcp{same concern}) the
+  A {\em lazy} variant of the stack-safety property that can be enforced by the
   more efficient Lazy Tagging and Clearing policy
@@
 
-We assume we are given a machine language program, together with some auxiliary information
+We assume we are given a machine language program, together with certain auxiliary declarations
 about its structure: the code footprint of each function and the locations of instructions
 representing calls, returns, and function entry points.
-Although this information would typically come from a compiler together with
-the machine code, we do not
-assume anything about their provenance.  We assume that the machine has a
-program counter $\PCname$
-and stack pointer $\rsp$\bcp{but not a frame pointer?!}; other details of the machine's ISA are unimportant.
-
-To make our stack safety definitions precise, we make some straightforward
-assumptions about
+Although this information would typically come from a compiler, we do not
+assume anything about the provenance of the machine
+code. We assume that the machine has a program counter $\PCname$
+and stack pointer $\rsp$ and can perform the usual
+loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.
+
+To make our stack safety definitions precise, we make some simple assumptions about
 memory layout; our definitions could easily be changed to be parametric over
-these assumptions, at the cost of some complexity in their statements.  Memory
-consists of an instruction region and a stack that grows upwards. (For simplicity,
-we do not consider the heap. Integrating a memory safety property for the
-heap in the style of
+these assumptions, at the cost of greater complexity.  Memory
+consists of an instruction region and a stack that grows upwards; for simplicity,
+we omit a heap. (Integrating a memory safety property for the heap in the style of
 \citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
@@
 contiguously allocated per-function frames.
-The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
+The boundary between the frames of a caller and callee is given by the $\rsp$ register
 at the time of the call.  The stack can be used to
 pass arguments or return results in the usual way, by placing them
-at the top of the caller's frame. Callers and callees can also communicate
-via registers if they wish; indeed, our
-definitions say nothing at all about registers (except $\PCname$ and
-$\rsp$). \bcp{This implicitly amounts to choosing a caller-saves register
-  convention, though, doesn't it?}
+at the top of the caller's frame. Functions can also communicate via registers; indeed, our
+definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
 
 Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code (the same instruction may be
-used in some places as part of a call or return sequence and in other places
-for other purposes).  We define a call to be
-any instruction declared as such in the program\bcp{in the call map, not the
-program, right?}.
+which are not fully explicit in machine code.  We define a call to be
+any instruction declared as such in the program.
 If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
@@
 
-Enforcement mechanisms are modeled as {\em policies}\bcp{this is a somewhat
-  special use of the word ``policy'' that comes from the micropolicies
-  world.  We should include a comment about this to avoid confusing
-  readers.} that add some auxiliary
+Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
 policy state to the ordinary machine state
@@
 can perform an action that would violate the security property; our
-properties are therefore naturally \emph{termination insensitive.}
+properties are naturally therefore \emph{termination insensitive.}
 This model will be made more precise in \cref{sec:prelim}.
 
-We model ordinary program state and policy-enforcement state separately to make it
+We model ordinary program state and policy enforcement state separately to make it
 easier to use well-established concepts from language-based security.
@@
 \emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable
-behavior.\leo{What citation should go here? Not the hyperproperties one?
-  Goguen and Meseguer?}
+\emph{varying} that data would not change the program's observable behavior.\leo{What citation should go here? Not the hyperproperties one?}
 In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, while any associated policy
+ordinary program data should be varied, whereas any associated policy enforcement
 state should be left alone; our model makes this distinction obvious.
@@
 %execution where data has been varied.
-We will also introduce a notion of {\em observational integrity} that relies on the idea of
+We also introduce a notion of observational integrity that relies on the idea of
 \emph{rolling back} program state; again, our model makes it easy to describe
-rollback of ordinary program data while leaving policy state alone.
+rollback of ordinary program data while leaving policy enforcement data unchanged.
 
@@
 
-This enforcement model directly describes
-micropolicies~\citep{pump_oakland2015}, and we believe that it is
-flexible enough to capture code-altering approaches (such as code rewriting
+This enforcement model is flexible enough to capture
+micropolicies~\citep{pump_oakland2015}, and we believe that it can
+also capture code-altering approaches (such as software-based
 or capability-based techniques).
@@
 %
-Micropolicies are a flexible tag-based, hardware-accelerated
-reference monitoring mechanism; they been applied to stack safety
-enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
+Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
+reference monitoring mechanism, which have been applied to stack safety by
+\citet{DBLP:conf/sp/RoesslerD18}.
 Here the policy state consists of metadata tags (e.g., identifying stack frames)
-attached to each value in memory or registers, and the policy step
-function\bcp{undefined!}
+attached to each value in memory or registers, and the policy step function
 checks that each machine operation obeys a set of rules on tags
@@
 %Other policies are implemented in hardware or built on hardware mechanisms.
-Capability machines such as CHERI \citep{Woodruff+14} extend conventional
+Capability machines like CHERI \citep{Woodruff+14} extend conventional
 architectures to support efficient and fine-grained control over memory
@@
 the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
-linear capabilities?}.  Capabilities package up pointers
+\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
 (ordinary data) with base and bounds information (policy data),
-and they require rewriting source code (e.g., to use capabilities in place
+and their use requires rewriting source code (e.g. to use capabilities in place
 of ordinary pointers), so separating ordinary machine state from policy
-state is delicate; we conjecture it can be done with some effort.
+state is delicate, but we conjecture it can be done with some effort.
+%\apt{But we conjecture it can be done...?}\leo{Just say that?}
 
@@
 Some policies are implemented in software, by modifying code in the
-compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
+compiler to (e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or
 to insert stack canaries~\citep{Cowan+98}. Such approaches should
@@
 \leo{Something like this?}
-\bcp{OK}
 %These approaches are represented as policies that
@@
 
-\ifaftersubmission
 \bcp{We really want readers to understand this trace.  I think we should
@@
   Or do you have something else in mind?}
-\bcp{I was thinking about just a few words, not a page.  But let's do this
-  later.}
-\fi
 
@@
 written in C-like pseudo-code.  It consists of a {\tt main} function that
-performs a call to some function {\tt f}, which in turn makes a nested call
+performs a call to some function {\tt f}, which in turns makes a nested call
 to some function {\tt g} before returning, while performing some simple
-arithmetic operations in the process. \Cref{fig:running-trace-b} shows an
-execution trace of the corresponding machine instructions. These might have
+arithmetic operations in the process. \cref{fig:running-trace-b} shows an
+execution trace of corresponding machine instructions. (These might have
 been generated by compiling the high-level code, but it is only the behavior
 of the machine instructions that we care about
-here---\Cref{fig:running-trace-b} is our actual example.
+here---\Cref{fig:running-trace-b} is our actual running example.)
 
@@
 %
-First, as noted above, since our goal is to formalize stack safety, we
-ignore the heap
-and assume that all variables are stack allocated.
+First, since our goal is to formalize stack safety, we ignore the heap
+and assume that all variables are stack allocated. As noted in the
+previous section, the extension to account for memory safety is
+discussed in~\Cref{sec:relwork}.
 %
@@
 A sketch of the initial memory layout appears at the top of
-\Cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
-and the (uninitialized)
-stack begins at 100. Directly below the memory sketch, we summarize the parts
-of the register file that are relevant at this point in the example.
-%
-The rest of \Cref{fig:running-trace-b} traces the execution of the
+\cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
+the dedicated output is uninitialized, while the (also uninitialized)
+stack begins at 100. Directly below the layout, we summarize the parts
+of the register file that are relevant to the example.
+%
+The rest of \cref{fig:running-trace-b} traces the execution of the
 program, showing the machine state at call and return points.
 Instructions are intended to
-be similar to RISC-V, but are described in a (hopefully more readable) an
-ad-hoc notation
+be similar to RISC-V, but are described in an ad-hoc notation
 rather than using assembler syntax. The $\jal$ (jump-and-link)
-instruction used for calls atomically simultaneously the return address in a
-register
+instruction used for calls atomically saves the return address in a register
 and jumps to a specified new $\PCname$; the $\jalr$ (jump-and-link-via-register)
 instruction used for returns jumps to the address stored in a register.
-Register $r_0$ is hardwired to zero, so that, e.g., instruction 3 stores $0$
-into address 101.
+Register $r_0$ is hardwired to zero (so that, e.g., instruction 3 stores $0$
+into address 101.).
 
@@
 
-We next describe our machine model. To make our definition of stack safety
-as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (static or dynamic, software- or
-hardware-enforced) policy monitor. In~\cref{sec:enforcement}, we will
-instantiate this model with a concrete dynamic policy enforcement mechanism
-based on \citet{DBLP:conf/sp/RoesslerD18}.
-
-We will often refer to the running example of \cref{sec:running-example} for
-concreteness.
+In this section we describe our machine model. Our definition of stack
+safety is generic and this section discusses an abstract interface to
+our machine model and some kind of static or dynamic, software- or
+hardware-enforced security monitor. We will refer back to the running
+example of \cref{sec:running-example} for concreteness, and then,
+in~\cref{sec:enforcement}, extend this model with a dynamic policy
+enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
 
@@
 
-A {\em machine state} is a map from components to values:
+A machine state is a map from components to values:
 %
@@
 \(\mach_e'\) agrees on low confidentiality memory, but all other addresses
-hold $-5$, in {\varied}.  After the last step, the return states \(\mach_r\) and
+hold -5, in {\varied}.  After the last step, the return states \(\mach_r\) and
 \(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
 \(\mach_e'\). Addresses 108 and 109 have changed, shown in light {\low}, and while 109
-had changed in a consistent way (independent of the variation), 108 holds either 43 or $-4$, so the trace does not obey trace confidentiality.
+had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
 %\rb{Color name macros}
@@
 
-In this section we are going to see how an existing enforcement mechanism,
+In this section we examine how an existing enforcement mechanism,
 Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
 the formal stack safety property described in the previous section: it in
-fact enforces a stronger, inductive version of it that strengthens the
+fact enforces a stronger, inductive version of that property which strengthens the
 local state encapsulation component by checking for integrity and
@@
 
-\paragraph*{Micropolicies}
+%\paragraph*{Micropolicies}
 %
@@
 \emph{micropolicies}~\citep{pump_oakland2015}.
-
-In such a system, all memory addresses and registers (including the program
-counter) are enriched with an abstract metadata tag, which can represent an
-arbitrary data structure. A micropolicy is defined as a set of tags, and a
-collection of software-defined rules that represent the transfer function of the
-policy. Each rule inspects the tags associated to the inputs of an instruction
-(PC, registers, memory) as well as the instruction opcode itself, and produces
+In such a system, all values in
+memory addresses and registers (including the $\PCname$)
+are enriched with an abstract metadata tag, which can represent 
+arbitrary information about the value. A micropolicy is defined as a set of tags and a
+collection of software-defined rules, indexed by machine opcode. 
+At each step of the machine, the relevant rule is applied to the tags on 
+the instructions inputs (PC, registers, memory) and on the instruction itself, and produces
 one of two outcomes: either the instruction is allowed to execute (and generates
-and assigns tags for the new PC and the result of the operation), or the machine
-fail-stops with a policy violation. During program execution, micropolicy rules
-are checked before each machine step.
+tags for the result of the operation and the new PC), or the machine
+fail-stops with a policy violation. 
+\citeauthor{pump_oakland2015} have shown that a wide range of micropolicies can
+be defined using this scheme.
 
 Efficient execution of these micropolicies relies on hardware implementations,
-such as the PUMP architecture. This hardware incorporates a monitor rule cache
-to allow quick retrieval of transfer function outputs for mapped inputs. Cache
-misses are detected and trapped to software handlers that use the software rules
-of the policy to compute the relevant parts of the transfer function.
-\citeauthor{pump_oakland2015} have shown that a wide range of micropolicies can
-be defined and enforced efficiently in these machines.
+such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
+as word-size bit vectors, stored separately from the regular memory and registers.
+The hardware incorporates a rule cache
+to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
+the hardware traps to a software handler (running in a privileged context or
+on a co-processor) to compute the rule result. To obtain adequate performance,
+it is important to design micropolicies so that they hit in the cache as
+much as possible. Thus, practical policies maintain a small working set of
+distinct tags.
 %% %
@@
 control that we do not need to consider here), and tags registers containing
-stack pointers with access permission information. Beyond
-those capabilities, the compiler generates designated sequences of
-instructions with special tags that grant them permission to manage
-the stack, say during calls and returns.
+stack pointers with access permission information. The compiler also
+uses tags on instructions to identify the code 
+sequences that have permission to manage the stack, say during calls and returns.
 %% \rb{Conceivably on alloc
@@
 
-Our micropolicy works as follows. We assume for ease of exposition that no
-arguments are passed on the stack.
+Our micropolicy works as follows. (We assume for ease of exposition that no
+arguments are passed on the stack.)
 %% \rb{Come back to this later?}.
-Initially, the entire stack is tagged as unused with an $\tagNoDepth$ tag,
+The micropolicy maintains tags on values in memory of the form $\tagStackDepth{n}$,
+indicating locations that belong to the stack frame at activation depth $n$,
+or $\tagNoDepth$; it tags the $\PCname$ with tags of the form $\tagPCDepth{n}$.
+During normal execution, the micropolicy rules only permit load and store operations
+when the target memory is tagged with the same depth as the current {\PCname} tag.
+Initially, the entire stack is tagged $\tagNoDepth$,
 %\rb{actually, the ``initial frame'' should be tagged with its proper depth}
-and the {\PCname} carries in its tag the depth of the current
-activation at all times, starting with a $\tagPCDepth{0}$ tag. The
-current activation depth needs to change exactly at those points in
+and the {\PCname} has tag $\tagPCDepth{0}$. These tags are altered
+at exactly those points in
 the program where the contour of the stack safety property changes:
@@
 
-\item From caller to callee, when the machine executes an instruction tagged
-  with $\tagCall$ as belonging to the call map. At this point, the current
-  {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
-  the locations of the newly allocated frame are marked $\tagStackDepth{n + 1}$.
-
-\item From callee back to caller, when execution reaches the corresponding
-  return instruction for the call instruction, as defined by
-  $\ret{\mach_c}\ \mach_r$. At this point, the {\PCname} tag is decremented.
-  Prior to this point, the locations of the old frame have been restored to
-  $\tagNoDepth$.
+\item From caller to callee, when the machine executes an instruction marked as a call in
+  the call map. 
+  At this point, the current
+  $\tagPCDepth{n}$ tag is incremented to $\tagPCDepth{(n + 1)}$, and the function
+  entry sequence initializes  all locations of the new function frame with
+  tag $\tagStackDepth{(n + 1)}$.
+
+\item From callee back to caller, when execution is about to reach the return point
+  $\ret{c}$
+  corresponding to the call at $c$.  The exit sequence retags all locations in the function
+  frame with $\tagNoDepth$. At the actual return instruction, the
+  $\PCname$ tag is decremented.
 
@@
 
-To enforce this discipline, we define blessed header and return instruction
-sequences. These are specific sequences of tagged instructions that are intended
-to appear at the beginning and end of each function. This is the header
-sequence:
-%
-\setcounter{pcctr}{1}
-\[
-  \begin{array}{l|l|l|l}
-     & \mathit{Instruction} & \mathit{Tags} & \\
-    \hline
-    \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
-        {Store return address in stack}
-    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
-        {Increment stack pointer by frame size}
-  \end{array}
-\]
-%
-And the return sequence:
-%
-\setcounter{pcctr}{1}
-\[
-  \begin{array}{l|l|l|l}
-     & \mathit{Instruction} & \mathit{Tags} & \\
-    \hline
-    \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
-        {Load return address}
-    \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
-        {Decrement stack pointer by frame size}
-    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
-        {Return to caller ($\jalr$)}
-  \end{array}
-\]
+This discipline suffices to enforce local stack encapsulation.
+To implement it,  we define ``blessed'' instruction sequences
+intended to appear at the entry and exit of each function,
+which manipulate tags as just described in addition to performing the
+usual calling convention tasks of saving/restoring the return address to/from
+the stack and adjusting the stack pointer. With the aid of an additional tag on
+the $\rsp$ register, these sequences also serve to enforce
+well-bracketed control flow.
+The micropolicy guarantees atomic execution of these sequences
+using a combination of tags on the instructions
+and an additional tag on the $\PCname$; we omit the details here.
+
+There remains the question of how to ensure that the sequences are
+invoked at the right places. For a program to enjoy the stack safety property
+(with respect to a particular call map), all we need require is that 
+the entry sequence be initiated at any instruction marked as a call
+in the call map. This is achieved by giving these instructions (another) special tag,
+and it is easy to check statically that this has been done correctly.
+If the code fails to initiate an exit
+sequence at a point where the stack safety property expects a return,
+the micropolicy will incorrectly behave as if execution
+is continuing in the callee, but since the callee never has stronger access
+rights than the caller, this is harmless.
+
+%% APT: removed because the tag part is too mysterious and the instruction part
+%% is standard. 
+%% This is the entry
+%% sequence:
+%% %
+%% \setcounter{pcctr}{1}
+%% \[
+%%   \begin{array}{l|l|l|l}
+%%      & \mathit{Instruction} & \mathit{Tags} & \\
+%%     \hline
+%%     \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
+%%         {Store return address in stack}
+%%     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
+%%         {Increment stack pointer by frame size}
+%%   \end{array}
+%% \]
+%% %
+%% And the exit sequence:
+%% %
+%% \setcounter{pcctr}{1}
+%% \[
+%%   \begin{array}{l|l|l|l}
+%%      & \mathit{Instruction} & \mathit{Tags} & \\
+%%     \hline
+%%     \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
+%%         {Load return address}
+%%     \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
+%%         {Decrement stack pointer by frame size}
+%%     \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
+%%         {Return to caller ($\jalr$)}
+%%   \end{array}
+%% \]
 
@@
 
-These sequences also enforce the $\SP$ discipline. When stack frames
-are of fixed size, it suffices to increment $\SP$ and tag it with a
-dedicated tag in the header sequence, which is later checked after
-decrementing $\SP$ during the exit sequence to authorize the
-return. The well-formedness of the blessed sequences can be checked
-statically. Only programs whose sequences are well-formed are
-protected by the micropolicy, although programs are still protected
-(by fail-stopping) even if the intended sequences are missing. By
-tracking the execution of the blessed sequences in the tag state of
-the {\PCname} register, the micropolicy ensures that the entry
-sequences of calls declared in the call map are executed correctly.
-%
-During normal execution, the {\PCname} uses the current activation
-depth in its tag $\tagPCDepth{n}$ to gate load and store operations,
-which are allowed to operate only on the corresponding stack frame,
-tagged $\tagStackDepth{n}$.
+%% These sequences also enforce the $\SP$ discipline. When stack frames
+%% are of fixed size, it suffices to increment $\SP$ and tag it with a
+%% dedicated tag in the header sequence, which is later checked after
+%% decrementing $\SP$ during the exit sequence to authorize the
+%% return. The well-formedness of the blessed sequences can be checked
+%% statically.
+
+%% Only programs whose sequences are well-formed are
+%% protected by the micropolicy, although programs are still protected
+%% (by fail-stopping) even if the intended sequences are missing. By
+%% tracking the execution of the blessed sequences in the tag state of
+%% the {\PCname} register, the micropolicy ensures that the entry
+%% sequences of calls declared in the call map are executed correctly.
+
+Note that in this policy both the entry and exit sequences
+must write to each element of
+the frame, which can be quite expensive, especially for programs that
+allocate large but sparsely populated frames; we return to this point
+in \cref{sec:lazy}.
+%
 %\rb{Note that we have also considered and implemented more flexible schemes
@@
 
-      If \(\mach_e\) is an entry to a call with contour \(\contour\)
-      and \(\mach_r\) its return, the result of
-      \(\mathit{roll}(\contour,\mach_e,\mach_r)\) agrees with the
-      entry \(\mach_e\) on every component the callee should not
-      change, and with the actual return \(\mach_r\) on those
-      components that it is permitted to change.  So the observable
-      behavior that follows is the idealized behavior to which we
-      compare the trace from \(\mach_r\).
+      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
+      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
+      agrees with \(\mach_e\) on every component the callee should not change,
+      but with \(\mach_r\) on those components that it is permitted to change.
+      So the observable behavior that follows is the idealized behavior to
+      which we compare the trace from \(\mach_r\), which must prefix it (to
+      account for premature halts due to policy violations.)
 
@@
       A trace enjoys {\em observable trace integrity} with respect to
-      some contour \(\contour\) if the trace induced by its final
-      state (if any) has observable behavior that prefixes that of its
-      idealized version (to account for premature halts due to policy
-      violations):
+      some contour \(\contour\) if the trace induced by its final state
+      (if any) has observable behavior that prefixes that of its idealized
+      version, constructed by \(\mathit{roll}\):
 
@@
     trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
-    it prints 0. Only then is observable trace integrity violated.
+    it prints 0. So observable trace integrity does not hold.
 
@@
 
-      Two of the cases of trace confidentiality, \(\mathit{InfCase}\)
-      and \(\mathit{FailCase}\), are purely properties of the
-      observations produced by the traces they are properties of. If
-      the traces are infinite, they should be similar; if one
-      fail-stops, it should be a prefix. Only the
-      \(\mathit{ReturnCase}\) checks machine states, because in that
-      case the traces could continue, but have been cut off by a
-      return. The key to making confidentiality lazy is not to perform that
-      check but rather to consider the traces after that return,
-      checking only whether they are observationally similar.
+      Two of the cases of trace confidentiality, \(\mathit{InfCase}\) and
+      \(\mathit{FailCase}\), are purely properties of the observations produced
+      by the traces they are properties of. If the traces are infinite, they
+      should be similar; if one fail-stops, it should be a prefix. Only
+      the \(\mathit{ReturnCase}\) checks machine states, because in that case
+      the traces could continue, but have been cut off by a return. The key
+      to making confidentiality lazy is to consider the traces after the
+      return, considering only whether they are observationally similar
+      or prefixes.
 
@@
 
-      \medskip
-      \textsc{Theorem.}
-      Stack safety implies observable stack safety.
-      \smallskip
-
-      This is more straightforward \rb{than what?} (1,5K LoC in Coq) as the recursive
-      structure of the two properties is similar.
+      \paragraph*{THEOREM} Stack safety implies observable stack safety.
+
+%      \paragraph*{PROOF} \leo{TODO: Whatever we say about the previous proof
+ %       we have to say similar things here}
 
@@
 %
-To our knowledge, the first and only line of work that has attempted
-to furnish a positive and formal characterization of the meaning of
-stack safety is \citet{Skorstengaard+19b}, who introduce a calling
-convention that uses local capabilities to preserve local state
-encapsulation and well-bracketed control flow, using a logical
-relation to reason about the stack safety of concrete
-programs. Although the required hardware support is readily available
-in capability machines like CHERI, this technique incurs significant
-costs, because it requires the entire unused part of the stack to be
-cleared whenever a security boundary is crossed. Their logical
-relation captures capability safety without ``externally observable
-side-effects (like console output or memory access traces).'' In the
-discussion, ``while [the authors] claim that [their] calling
-convention enforces control-flow correctness, [they] do not prove a
-general theorem that shows this, because it is not clear what such a
-theorem should look like,'' noting that the correctness property
-enjoyed by their technique ``is not made very explicit.''
+To our knowledge, the first and only line of work that has attempted to furnish
+a positive and formal characterization of the meaning of stack safety was
+inaugurated by \citet{Skorstengaard+19b}, who introduce a calling convention
+that uses local capabilities to preserve local state encapsulation and
+well-bracketed control flow, using a logical relation to reason about the stack
+safety of concrete programs. Although the required hardware support is readily
+available in capability machines like CHERI, this technique incurs potentially
+prohibitive performance costs, because it requires the entire unused part of the
+stack to be cleared whenever a security boundary is crossed. Their logical
+relation captures capability safety without ``externally observable side-effects
+(like console output or memory access traces).'' In the discussion, ``while [the
+authors] claim that [their] calling convention enforces control-flow
+correctness, [they] do not prove a general theorem that shows this, because it
+is not clear what such a theorem should look like,'' noting that the correctness
+property enjoyed by their technique ``is not made very explicit.''
 
@@
   \paragraph{Heap Safety as a Security Property}
-
-Heap safety, like stack safety, may be framed as a security property
-in the form of noninterference~\citet{DBLP:conf/post/AmorimHP18}. Just
-like~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
-characterization of the meaning of (heap) memory safety, in this paper
-we aim to do the same for the stack.
-%
-Their model describing safety of shared stack-allocated
-objects extends naturally to heap safety, complicated by the fact that
-these objects are not deallocated by returns, and therefore a
-function's privilege may increase or decrease after its entry
-point. For instance, if a callee allocates a heap object and returns
-the pointer to its caller, that object's addresses become accessible
-in contradiction to a contour computed at the caller's entry. So such
-an extension must enforce trace confidentiality and integrity
-properties separately on continuous segments of each call, with
-contours computed at each crossing between caller and callee.
+\bcp{Talk about the relation.}
+    Heap safety, like stack safety, may be framed as a form of
+    noninterference\citet{DBLP:conf/post/AmorimHP18}. The above model
+    describing safety of
+    shared stack-allocated objects extends naturally to heap safety,
+    complicated by the fact that these objects are not deallocated by returns,
+    and therefore a function's privilege may increase or decrease after its
+    entry point. For instance, if a callee allocates a heap object and returns
+    the pointer to its caller, that object's addresses become accessible in
+    contradiction to a contour computed at the caller's entry. So such an
+    extension must enforce trace confidentiality and integrity properties
+    separately on continuous segments of each call, with contours computed at
+    each crossing between caller and callee.
 
@@
   describe the basic principles by which our model can be extended to some of
-  these cases. We also plan to develop a testing framework for quickly
-  checking the validity of different enforcement mechanisms.
+  these cases.
 
@@
 
-    Common programming idioms involve a caller passing a pointer to
-    its local data into a callee. Using standard notions of pointer
-    provenance~\citep{provenance}, we can extend our model to
-    distinguish safe use of this idiom from true stack safety
-    violations. We require additional annotation to identify which
-    addresses within a function's stack frame correspond to distinct
-    objects. Then a valid pointer to an object is one derived through
-    legal arithmetic from its base address, and at each call point, an
-    object's addresses are marked low confidentiality and integrity if
-    a valid pointer exists in a register or in the transitive closure
-    of the callee's accessible memory. All pointers to an object cease
-    to be valid when it is deallocated along with its stack frame.
+    Common programming idioms involve a caller passing a pointer to its local
+    data into a callee. Using standard notions of pointer provenance [cite],
+    we can extend our model to distinguish safe use of this idiom from true
+    stack safety violations. We require additional annotation to identify
+    which addresses within a function's stack frame correspond to distinct
+    objects. Then a valid pointer to an object is one derived through legal
+    arithmetic from its base address, and at each call point, an object's
+    addresses are marked low confidentiality and integrity if a valid pointer
+    exists in a register or in the transitive closure of the callee's
+    accessible memory. All pointers to an object cease to be valid when it is
+    deallocated along with its stack frame.
 
@@
     data are unchanged. Confidentiality properties are split similarly.
-
-  \paragraph{Random Testing}
-
-  Part of the motivation for the strong lockstep stack safety property
-  is that it uncovers errors quickly, and therefore should be quicker
-  to test just like the smarter testing properties
-  of~\citet{TestingNI:ICFP}. We have already developed a testing
-  framework for debugging our policy enforcement for this property, and
-  it would be interesting to explore whether a testing-amenable variant
-  of the lazy property also exists.
 
@@
 %% Appendix
-%\appendix
-%\section{Appendix}
-%Text of appendix \ldots
+\appendix
+\section{Appendix}
+
+Text of appendix \ldots
 

### 2020-07-10 main.tex

--- 2020-07-10 main.tex
@@
   linear capabilities?}\leo{Only the second one AFAIK. What would you
-  change to characterize it more accurately?}.  Capabilities package up pointers
+  change to characterize it more accurately?}\rb{StkPointers uses linear
+  capabilities, its predecessor users local capabilities (and unlike
+  StkPointers could be implemented in CHERI)}.
+Capabilities package up pointers
 (ordinary data) with base and bounds information (policy data),

### 2020-07-10 main.tex
Introduction work

--- 2020-07-10 main.tex
@@
 %
-Our key contribution is another important step towards this goal: a generic
-presentation of local state encapsulation and well-bracketed
-control\bcp{we need to either weaken the previous sentence or else add
-  ``first steps toward'' in this sentence...}\leo{Weakened enough?} flow using the technical
-framework of language-based security.
+Our key contribution is such a generic presentation of
+local state encapsulation and well-bracketed control\bcp{we need to either
+  weaken the previous sentence or else add ``first steps toward'' in this sentence...} flow using the
+technical framework of language-based security.
 
@@
 A key challenge in adapting the concepts of language-based security to this
-setting is that stack safety is a ``nested'' property: not only does a given
-caller need to be protected from its callee, but the callee also needs to be
-protected from {\em its} callees, and so on. That is, the notion of
-``attacker'' changes at many points throughout the execution of the program.
+setting is that integrity and confidentiality are ``nested'' properties: not
+only does a given caller need to be protected from its callee, but the callee
+also needs to be protected from {\em its} callees, and so on. That is, the notion
+of ``attacker'' changes at many points throughout the execution of the program.
 We address
@@
 when a callee returns, this policy signals violations when the
-caller actually uses data that the callee has written into  its stack
+caller actually accesses data that the callee has written into its stack
 frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
   property inthe same way...}\fi.
-The lazy
-policy does not enforce our original stack-safety property, but (a slightly
-enhanced variant of the policy\bcp{I worry that we are being a little
-  dishonest here: the stated advantage of the lazy policy is that it is more
-efficient, but our variant is {\em not} efficient!}) of it does enforce a slightly weaker
-lazy version of the property that still protects from unsafe callees but is
-less intuitive to state.
+The lazy policy does not enforce our original integrity and confidentiality
+properties, and to our knowledge there is no formal statement of the properties that
+it ought to enforce. We propose ``observable'' versions of integrity and
+confidentiality property, which capture the desired behavior of a policy that
+detects violations between when they occur and when they make a visible impact.
+Unfortunately the existing Lazy Tagging and Clearing policy of
+\citeauthor{DBLP:conf/sp/RoesslerD18} does not quite enforce these properties,
+but would if enhanced (at some performance cost).
+%\bcp{I worry that we are being a little
+%  dishonest here: the stated advantage of the lazy policy is that it is more
+%efficient, but our variant is {\em not} efficient!}
+
+\sna{There's been some inconsistency in ``stack safety'' vs.
+     ``stack-safety,'' I'm fixing on the former.}
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A new {\em formal stack-safety property} capturing the
+  A new {\em formal stack safety property} capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
 \item
-  A {\em lockstep} variant of the stack-safety property that is also
+  A {\em lockstep} variant of the stack safety property that is also
   enforced by the Depth Isolation policy
@@
 \item
-  A {\em lazy} variant of the stack-safety property that can be enforced by
-  (a variant of\bcp{same concern}) the
-  more efficient Lazy Tagging and Clearing policy
-  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
+  A {\em lazy} variant of the stack safety property that can be enforced if the
+  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
+  with unique tags on each activation, plus a proof (in Coq) that this
   variant is weaker than the original~(\cref{sec:lazy}).
@@
 contiguously allocated per-function frames.
-The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
+The boundary between the frames of a caller and callee is given by the $\rsp$ register
 at the time of the call.  The stack can be used to
 pass arguments or return results in the usual way, by placing them
-at the top of the caller's frame. Callers and callees can also communicate
-via registers if they wish; indeed, our
-definitions say nothing at all about registers (except $\PCname$ and
-$\rsp$). \bcp{This implicitly amounts to choosing a caller-saves register
-  convention, though, doesn't it?}
+at the top of the caller's frame. Functions can also communicate via registers; indeed, our
+definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
 
 Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code (the same instruction may be
-used in some places as part of a call or return sequence and in other places
-for other purposes).  We define a call to be
-any instruction declared as such in the compiler-provided call structure.
+which are not fully explicit in machine code.  We define a call to be
+any instruction declared as such in the program.
 If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
@@
 
-We model ordinary program state and policy-enforcement state separately to make it
+We model ordinary program state and policy enforcement state separately to make it
 easier to use well-established concepts from language-based security.
@@
 \emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable
-behavior~\citep{6234468}
+\emph{varying} that data would not change the program's observable behavior.\leo{What citation should go here? Not the hyperproperties one?}
 In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, while any associated policy
+ordinary program data should be varied, whereas any associated policy enforcement
 state should be left alone; our model makes this distinction obvious.
@@
 %
-Micropolicies are a flexible tag-based, hardware-accelerated
-reference monitoring mechanism; they been applied to stack safety
-enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
+Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
+reference monitoring mechanism, which have been applied to stack safety by
+\citet{DBLP:conf/sp/RoesslerD18}.
 Here the policy state consists of metadata tags (e.g., identifying stack frames)
-attached to each value in memory or registers, and the policy step
-function\bcp{undefined!}\leo{``partial policy step'' is explained informally above}
+attached to each value in memory or registers, and the policy step function
 checks that each machine operation obeys a set of rules on tags
@@
 the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
-  linear capabilities?}\leo{Only the second one AFAIK. What would you
-  change to characterize it more accurately?}.  Capabilities package up pointers
+\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
 (ordinary data) with base and bounds information (policy data),
-and they require rewriting source code (e.g., to use capabilities in place
+and their use requires rewriting source code (e.g., to use capabilities in place
 of ordinary pointers), so separating ordinary machine state from policy
-state is delicate; we conjecture it can be done with some effort.
+state is delicate, but we conjecture it can be done with some effort.
+%\apt{But we conjecture it can be done...?}\leo{Just say that?}
 
@@
 policy, but only for programs produced by the compiler.
+\apt{not sure what to say here}
+\leo{Something like this?}
 %These approaches are represented as policies that
@@
 arithmetic operations in the process. \Cref{fig:running-trace-b} shows an
-execution trace of the corresponding machine instructions. These might have
+execution trace of corresponding machine instructions. (These might have
 been generated by compiling the high-level code, but it is only the behavior
 of the machine instructions that we care about
-here---\Cref{fig:running-trace-b} is our actual example.
+here---\Cref{fig:running-trace-b} is our actual running example.)
 
@@
 %
-First, as noted above, since our goal is to formalize stack safety, we
-ignore the heap
-and assume that all variables are stack allocated.
+First, since our goal is to formalize stack safety, we ignore the heap
+and assume that all variables are stack-allocated. As noted in the
+previous section, the extension to account for memory safety is
+discussed in~\Cref{sec:relwork}.
 %
@@
 \Cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
-and the (uninitialized)
-stack begins at 100. Directly below the memory sketch, we summarize the parts
-of the register file that are relevant at this point in the example.
+the dedicated output is uninitialized, while the (also uninitialized)
+stack begins at 100. Directly below the layout, we summarize the parts
+of the register file that are relevant to the example.
 %
@@
 Instructions are intended to
-be similar to RISC-V, but are described in a (hopefully more readable) an
-ad-hoc notation
+be similar to RISC-V, but are described in an ad-hoc notation
 rather than using assembler syntax. The $\jal$ (jump-and-link)
-instruction used for calls atomically simultaneously the return address in a
-register
+instruction used for calls atomically saves the return address in a register
 and jumps to a specified new $\PCname$; the $\jalr$ (jump-and-link-via-register)
 instruction used for returns jumps to the address stored in a register.
-Register $r_0$ is hardwired to zero, so that, e.g., instruction 3 stores $0$
-into address 101.
+Register $r_0$ is hardwired to zero (so that, e.g., instruction 3 stores $0$
+into address 101).
 
@@
 
-We next describe our machine model. To make our definition of stack safety
-as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (static or dynamic, software- or
-hardware-enforced) policy monitor. In~\cref{sec:enforcement}, we will
-instantiate this model with a concrete dynamic policy enforcement mechanism
-based on \citet{DBLP:conf/sp/RoesslerD18}.
+In this section we describe our machine model. Our definition of stack
+safety is generic and this section discusses an abstract interface to
+our machine model and some kind of static or dynamic, software- or
+hardware-enforced security monitor. We will refer back to the running
+example of \cref{sec:running-example} for concreteness, and then,
+in~\cref{sec:enforcement}, extend this model with a dynamic policy
+enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
 
@@
 pair of a machine state and an {\em observation}, which may be either
-silent (written $\tau$) or a word of data\bcp{This is a rather concrete
-  choice!  Why not leave the set of observations abstract?} that the machine
+silent (written $\tau$) or a word of data that the machine
 communicates to the outside world:
@@
 %
-This step function is total: execution traces are infinite unless
-cut short externally by a policy fault (as explained below).
-
-We choose to model observations explicitly, rather than simply
-treating
+This step function is total---execution traces are infinite unless cut
+short externally (e.g. by a policy as we'll see below).
+
+Notice that we choose to model observations explicitly, rather than simply treating
 a subset of registers or memory as being observable.  This choice gives
@@
 set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
-\cup \{\PCname,\SP\}$. Observations are writes to the
-special $\rout$ location\bcp{Which is 100! (do we need a name for it? seems
-  not...)}, which can be thought of as a memory-mapped
+\cup \{\PCname,\SP\}$. Observations are the writes to the
+  special $\rout$ location, which can be thought of as a memory-mapped
   output port: if during an execution step a word $w$ is
@@
 
-We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
+We now present our model of enforcement, introduced informally in \Cref{sec:threat}.
 A policy is a general model of an enforcement mechanism, consisting of a
@@
 \apt{Is this our current thinking?}
-\leo{I thought this was what we settled on in the last two days'
-  discussions}\bcp{That was not my impression!  I don't see how the policy
-  state can be recovered from just the machine state---e.g., the machine
-  state doesn't say where the calls and returns are, does it?!}
-%
-A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
+\leo{I thought this was what we settled on in the last two days' discussions}
+%
+A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be given in
 \Cref{sec:enforcement}.

### 2020-07-10 main.tex

--- 2020-07-10 main.tex
@@
 
-In this section we examine how an existing enforcement mechanism,
+In this section we are going to see how an existing enforcement mechanism,
 Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
 the formal stack safety property described in the previous section: it in
-fact enforces a stronger, inductive version of that property which strengthens the
+fact enforces a stronger, inductive version of it that strengthens the
 local state encapsulation component by checking for integrity and
@@
 
-%\paragraph*{Micropolicies}
+\paragraph*{Micropolicies}
 %
@@
 \emph{micropolicies}~\citep{pump_oakland2015}.
-In such a system, all values in
-memory addresses and registers (including the $\PCname$)
-are enriched with an abstract metadata tag, which can represent 
-arbitrary information about the value. A micropolicy is defined as a set of tags and a
-collection of software-defined rules, indexed by machine opcode. 
-At each step of the machine, the relevant rule is applied to the tags on 
-the instructions inputs (PC, registers, memory) and on the instruction itself, and produces
+
+In such a system, all memory addresses and registers (including the program
+counter) are enriched with an abstract metadata tag, which can represent an
+arbitrary data structure. A micropolicy is defined as a set of tags, and a
+collection of software-defined rules that represent the transfer function of the
+policy. Each rule inspects the tags associated to the inputs of an instruction
+(PC, registers, memory) as well as the instruction opcode itself, and produces
 one of two outcomes: either the instruction is allowed to execute (and generates
-tags for the result of the operation and the new PC), or the machine
-fail-stops with a policy violation. 
+and assigns tags for the new PC and the result of the operation), or the machine
+fail-stops with a policy violation. During program execution, micropolicy rules
+are checked before each machine step.
+
+Efficient execution of these micropolicies relies on hardware implementations,
+such as the PUMP architecture. This hardware incorporates a monitor rule cache
+to allow quick retrieval of transfer function outputs for mapped inputs. Cache
+misses are detected and trapped to software handlers that use the software rules
+of the policy to compute the relevant parts of the transfer function.
 \citeauthor{pump_oakland2015} have shown that a wide range of micropolicies can
-be defined using this scheme.
-
-Efficient execution of these micropolicies relies on hardware implementations,
-such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
-as word-size bit vectors, stored separately from the regular memory and registers.
-The hardware incorporates a rule cache
-to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
-the hardware traps to a software handler (running in a privileged context or
-on a co-processor) to compute the rule result. To obtain adequate performance,
-it is important to design micropolicies so that they hit in the cache as
-much as possible. Thus, practical policies maintain a small working set of
-distinct tags.
+be defined and enforced efficiently in these machines.
 %% %
@@
 control that we do not need to consider here), and tags registers containing
-stack pointers with access permission information. The compiler also
-uses tags on instructions to identify the code 
-sequences that have permission to manage the stack, say during calls and returns.
+stack pointers with access permission information. Beyond
+those capabilities, the compiler generates designated sequences of
+instructions with special tags that grant them permission to manage
+the stack, say during calls and returns.
 %% \rb{Conceivably on alloc
@@
 
-Our micropolicy works as follows. (We assume for ease of exposition that no
-arguments are passed on the stack.)
+Our micropolicy works as follows. We assume for ease of exposition that no
+arguments are passed on the stack.
 %% \rb{Come back to this later?}.
-The micropolicy maintains tags on values in memory of the form $\tagStackDepth{n}$,
-indicating locations that belong to the stack frame at activation depth $n$,
-or $\tagNoDepth$; it tags the $\PCname$ with tags of the form $\tagPCDepth{n}$.
-During normal execution, the micropolicy rules only permit load and store operations
-when the target memory is tagged with the same depth as the current {\PCname} tag.
-Initially, the entire stack is tagged $\tagNoDepth$,
+Initially, the entire stack is tagged as unused with an $\tagNoDepth$ tag,
 %\rb{actually, the ``initial frame'' should be tagged with its proper depth}
-and the {\PCname} has tag $\tagPCDepth{0}$. These tags are altered
-at exactly those points in
+and the {\PCname} carries in its tag the depth of the current
+activation at all times, starting with a $\tagPCDepth{0}$ tag. The
+current activation depth needs to change exactly at those points in
 the program where the contour of the stack safety property changes:
@@
 
-\item From caller to callee, when the machine executes an instruction marked as a call in
-  the call map. 
-  At this point, the current
-  $\tagPCDepth{n}$ tag is incremented to $\tagPCDepth{(n + 1)}$, and the function
-  entry sequence initializes  all locations of the new function frame with
-  tag $\tagStackDepth{(n + 1)}$.
-
-\item From callee back to caller, when execution is about to reach the return point
-  $\ret{c}$
-  corresponding to the call at $c$.  The exit sequence retags all locations in the function
-  frame with $\tagNoDepth$. At the actual return instruction, the
-  $\PCname$ tag is decremented.
+\item From caller to callee, when the machine executes an instruction tagged
+  with $\tagCall$ as belonging to the call map. At this point, the current
+  {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
+  the locations of the newly allocated frame are marked $\tagStackDepth{n + 1}$.
+
+\item From callee back to caller, when execution reaches the corresponding
+  return instruction for the call instruction, as defined by
+  $\ret{\mach_c}\ \mach_r$. At this point, the {\PCname} tag is decremented.
+  Prior to this point, the locations of the old frame have been restored to
+  $\tagNoDepth$.
 
@@
 
-This discipline suffices to enforce local stack encapsulation.
-To implement it,  we define ``blessed'' instruction sequences
-intended to appear at the entry and exit of each function,
-which manipulate tags as just described in addition to performing the
-usual calling convention tasks of saving/restoring the return address to/from
-the stack and adjusting the stack pointer. With the aid of an additional tag on
-the $\rsp$ register, these sequences also serve to enforce
-well-bracketed control flow.
-The micropolicy guarantees atomic execution of these sequences
-using a combination of tags on the instructions
-and an additional tag on the $\PCname$; we omit the details here.
-
-There remains the question of how to ensure that the sequences are
-invoked at the right places. For a program to enjoy the stack safety property
-(with respect to a particular call map), all we need require is that 
-the entry sequence be initiated at any instruction marked as a call
-in the call map. This is achieved by giving these instructions (another) special tag,
-and it is easy to check statically that this has been done correctly.
-If the code fails to initiate an exit
-sequence at a point where the stack safety property expects a return,
-the micropolicy will incorrectly behave as if execution
-is continuing in the callee, but since the callee never has stronger access
-rights than the caller, this is harmless.
-
-%% APT: removed because the tag part is too mysterious and the instruction part
-%% is standard. 
-%% This is the entry
-%% sequence:
-%% %
-%% \setcounter{pcctr}{1}
-%% \[
-%%   \begin{array}{l|l|l|l}
-%%      & \mathit{Instruction} & \mathit{Tags} & \\
-%%     \hline
-%%     \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
-%%         {Store return address in stack}
-%%     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
-%%         {Increment stack pointer by frame size}
-%%   \end{array}
-%% \]
-%% %
-%% And the exit sequence:
-%% %
-%% \setcounter{pcctr}{1}
-%% \[
-%%   \begin{array}{l|l|l|l}
-%%      & \mathit{Instruction} & \mathit{Tags} & \\
-%%     \hline
-%%     \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
-%%         {Load return address}
-%%     \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
-%%         {Decrement stack pointer by frame size}
-%%     \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
-%%         {Return to caller ($\jalr$)}
-%%   \end{array}
-%% \]
+To enforce this discipline, we define blessed header and return instruction
+sequences. These are specific sequences of tagged instructions that are intended
+to appear at the beginning and end of each function. This is the header
+sequence:
+%
+\setcounter{pcctr}{1}
+\[
+  \begin{array}{l|l|l|l}
+     & \mathit{Instruction} & \mathit{Tags} & \\
+    \hline
+    \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
+        {Store return address in stack}
+    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
+        {Increment stack pointer by frame size}
+  \end{array}
+\]
+%
+And the return sequence:
+%
+\setcounter{pcctr}{1}
+\[
+  \begin{array}{l|l|l|l}
+     & \mathit{Instruction} & \mathit{Tags} & \\
+    \hline
+    \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
+        {Load return address}
+    \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
+        {Decrement stack pointer by frame size}
+    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
+        {Return to caller ($\jalr$)}
+  \end{array}
+\]
 
@@
 
-%% These sequences also enforce the $\SP$ discipline. When stack frames
-%% are of fixed size, it suffices to increment $\SP$ and tag it with a
-%% dedicated tag in the header sequence, which is later checked after
-%% decrementing $\SP$ during the exit sequence to authorize the
-%% return. The well-formedness of the blessed sequences can be checked
-%% statically.
-
-%% Only programs whose sequences are well-formed are
-%% protected by the micropolicy, although programs are still protected
-%% (by fail-stopping) even if the intended sequences are missing. By
-%% tracking the execution of the blessed sequences in the tag state of
-%% the {\PCname} register, the micropolicy ensures that the entry
-%% sequences of calls declared in the call map are executed correctly.
-
-Note that in this policy both the entry and exit sequences
-must write to each element of
-the frame, which can be quite expensive, especially for programs that
-allocate large but sparsely populated frames; we return to this point
-in \cref{sec:lazy}.
-%
+These sequences also enforce the $\SP$ discipline. When stack frames
+are of fixed size, it suffices to increment $\SP$ and tag it with a
+dedicated tag in the header sequence, which is later checked after
+decrementing $\SP$ during the exit sequence to authorize the
+return. The well-formedness of the blessed sequences can be checked
+statically. Only programs whose sequences are well-formed are
+protected by the micropolicy, although programs are still protected
+(by fail-stopping) even if the intended sequences are missing. By
+tracking the execution of the blessed sequences in the tag state of
+the {\PCname} register, the micropolicy ensures that the entry
+sequences of calls declared in the call map are executed correctly.
+%
+During normal execution, the {\PCname} uses the current activation
+depth in its tag $\tagPCDepth{n}$ to gate load and store operations,
+which are allowed to operate only on the corresponding stack frame,
+tagged $\tagStackDepth{n}$.
 %\rb{Note that we have also considered and implemented more flexible schemes

### 2020-07-10 main.tex
CP 4

@@
 A key challenge in adapting the concepts of language-based security to this
-setting is that integrity and confidentiality are ``nested'' properties: not
-only does a given caller need to be protected from its callee, but the callee
-also needs to be protected from {\em its} callees, and so on. That is, the notion
-of ``attacker'' changes at many points throughout the execution of the program.
+setting is that stack safety is a ``nested'' property: not only does a given
+caller need to be protected from its callee, but the callee also needs to be
+protected from {\em its} callees, and so on. That is, the notion of
+``attacker'' changes at many points throughout the execution of the program.
 We address
@@
 when a callee returns, this policy signals violations when the
-caller actually accesses data that the callee has written into its stack
+caller actually uses data that the callee has written into  its stack
 frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
   property inthe same way...}\fi.
-The lazy policy does not enforce our original integrity and confidentiality
-properties, and to our knowledge there is no formal statement of the properties that
-it ought to enforce. We propose ``observable'' versions of integrity and
-confidentiality property, which capture the desired behavior of a policy that
-detects violations between when they occur and when they make a visible impact.
-Unfortunately the existing Lazy Tagging and Clearing policy of
-\citeauthor{DBLP:conf/sp/RoesslerD18} does not quite enforce these properties,
-but would if enhanced (at some performance cost).
-%\bcp{I worry that we are being a little
-%  dishonest here: the stated advantage of the lazy policy is that it is more
-%efficient, but our variant is {\em not} efficient!}
-
-\sna{There's been some inconsistency in ``stack safety'' vs.
-     ``stack-safety,'' I'm fixing on the former.}
+The lazy
+policy does not enforce our original stack-safety property, but (a slightly
+enhanced variant of the policy\bcp{I worry that we are being a little
+  dishonest here: the stated advantage of the lazy policy is that it is more
+efficient, but our variant is {\em not} efficient!}) of it does enforce a slightly weaker
+lazy version of the property that still protects from unsafe callees but is
+less intuitive to state.
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
-  A new {\em formal stack safety property} capturing the
+  A new {\em formal stack-safety property} capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
 \item
-  A {\em lockstep} variant of the stack safety property that is also
+  A {\em lockstep} variant of the stack-safety property that is also
   enforced by the Depth Isolation policy
@@
 \item
-  A {\em lazy} variant of the stack safety property that can be enforced if the
-  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
-  with unique tags on each activation, plus a proof (in Coq) that this
+  A {\em lazy} variant of the stack-safety property that can be enforced by
+  (a variant of\bcp{same concern}) the
+  more efficient Lazy Tagging and Clearing policy
+  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
   variant is weaker than the original~(\cref{sec:lazy}).
@@
   linear capabilities?}\leo{Only the second one AFAIK. What would you
-  change to characterize it more accurately?}\rb{StkPointers uses linear
-  capabilities, its predecessor users local capabilities (and unlike
-  StkPointers could be implemented in CHERI)}.
-Capabilities package up pointers
+  change to characterize it more accurately?}.  Capabilities package up pointers
 (ordinary data) with base and bounds information (policy data),
@@
 A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
-$\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
+$\TRACEELEMS$s; we write $\TRACE{\TRACEELEMS}$ for the set of traces over
+$\TRACEELEMS$. We will mostly be concerned with sequences of pairs of
+machine states
 and observations, called machine traces and denoted $\machT \in \MOTRACE$,
@@
 coinductively relates a machine state with the trace of machine states
-paired with observations produced by repeated application of step:
+and observations produced by repeated application of step:
 
@@
 \noindent
+(We use juxtaposition of an element and a sequence to represent ``cons.'')
 Since the step function is total, if \(\mach \hookrightarrow \machT\),
@@
 see a depiction of such a trace: starting from an initial machine
-state at the very top, each instruction constructs a new machine state
-(the diff of which compared to the previous one appears in the
-``Effect'' column) and the majority of observations are $\tau$ with
-the exception of instructions 31 and 13 which are writes to $\rout$
-and where the observations are $0$ and $60$ respectively.
+state at the very top, each instruction leads to a new machine state
+(the diff of which, compared to the previous one, appears in the
+``{\textit{Effect}'' column), and the majority of observations are $\tau$, except
+for instructions 31 and 13, which are writes to $\rout$
+where the observations are $0$ and $60$, respectively.
 
@@
 %the trailing observation. Here's an attempt.}
-takes a trace
-\(\Trace : \mathit{trace} ~ \TRACEELEMS \times \OBSS\) and a predicate
+takes a trace\bcp{I changed a : here to $\in$.  We should check if there are
+any more to fix.}
+\(\Trace \in \TRACE{\TRACEELEMS} \times \OBSS\) and a predicate
 on elements \(f \subset \TRACEELEMS\) and gives the prefix of
 \(\Trace\) ending with the first element on which \(f\) holds.
-Subtly, because the observation associated with a trace represents its
-next step, when we take a prefix from a trace containing observations
+One subtlety is that, because the observation associated with some element
+in a trace
+represents what can be observed when this element is executed (i.e., the
+observation on that step), when we take a prefix from a trace containing
+observations
 we must replace the final observation with a \(\tau\).
@@
 If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate never holds,
+predicate must hold on its final element. If the predicate never holds, then
 \(\PUT\) is the identity function.
@@
 %
-Let \(f\) be a predicate on machine states. Then we define \(\mach
+If \(f\) be a predicate on machine states, we define \(\mach
 \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
-running \(\mach\) up to \(f\)''):
+running \(\mach\) up to \(f\)'')
 
@@
 \noindent
-And similarly for traces with policies:
+and similarly for traces with policies:
 
@@
 We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-as far as an external observer is concerned, if the sequence of
+as far as an external observer is concerned, written \(\obsT_1 \lesssim
+\obsT_2\), if the sequence of
 non-silent observations of $\obsT_1$ is a prefix of those of
@@
 
+\smallskip
 We then define similarity of observation traces as traces prefixing each other:
-
 \[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
@@
 Note that an infinite silent trace is a
-prefix of (and equivalent to) any other trace. While this might seem
+prefix of (and equivalent\bcp{similar?} to) any other trace. While this might seem
 surprising at first, it makes sense in a timing-insensitive context:

### 2020-07-10 main.tex
Remove pinit

--- 2020-07-10 main.tex
@@
 
-We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
-A policy is a general model of an enforcement mechanism, consisting of a
-set of policy states \(\pol \in \POLS\), a policy step function
-\((\mach, \pol) \polstep \pol' : \MACHS \times \POLS
-\rightharpoonup \POLS\), and an initializer \(\pinit : \MACHS \rightharpoonup
-\POLS\). The initializer establishes the initial policy state for a given
-machine state, but is partial to reflect that some policies are only defined
-on states that are well-formed in some way.
-%
-\apt{Is this our current thinking?}
-\leo{I thought this was what we settled on in the last two days'
-  discussions}\bcp{That was not my impression!  I don't see how the policy
-  state can be recovered from just the machine state---e.g., the machine
-  state doesn't say where the calls and returns are, does it?!}
+We now present our model of enforcement, which we introduced
+informally in \Cref{sec:threat}.  A policy is a general model of an
+enforcement mechanism, consisting of a set of policy states \(\pol \in
+\POLS\), a policy step function \((\mach, \pol) \polstep \pol' :
+\MACHS \times \POLS \rightharpoonup \POLS\). We assume that the
+compiler provides an initial policy state $\pol_0$ to accompany 
+the initial machine state.
 %
@@
 We say that a system enjoys {\em stack integrity} if for every initial
-state $\mach_0$ and its corresponding induced trace $\MPT$ (with
-$\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT$), for every
+state $(\mach_0,\pol_0)$ and its corresponding induced trace $\MPT$ (with
+$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), for every
 $n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
@@
 %Let $\callmap$ be a call map and $\codemap$ a function map.
-For every initial state \(\mach_0\), we take the induced trace
-\(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\). Then for
+For every initial state \((\mach_0,\pol_0)\), we take the induced trace
+\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\). Then for
 any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
@@
 \(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
-\machT_{\mathit{call}}\) and
-\(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
+\machT_{\mathit{call}}\) and it also holds for the toplevel execution
+trace \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
 system as a whole enjoys {\em stack confidentiality} with respect to
@@
 %\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
-if, for any initial state \(\mach_0\) and any adjacent pair of states
+if, for any initial state \((\mach_0, \pol_0)\) and any adjacent pair of states
 \(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
-in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\),
+in the induced trace from \(\mpstate{\mach_0}{\pol_0}\),
 if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
@@
 %\(\callmap\) and entry map \(\entmap\)
-if, for any initial state \(\mach_0\)
+if, for any initial state \((\mach_0, \pol_0)\)
 and any n-transition from \((\mach_c,\pol_c)\) to
 \((\mach_e, \pol_e)\) in the induced trace from
-\(\mpstate{\mach_0}{\pinit(\mach_0)}\), $\mach_e$ is a valid entry:
+\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a valid entry:
 \[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]
@@
 A system enjoys {\em return integrity} if all of its returns match prior calls
-and return to the appropriate location. For any initial state \(\mach_0\), let
-\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pinit(\mach_0)}\).
+and return to the appropriate location. For any initial state \((\mach_0\, \pol_0)\), let
+\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\).
 \(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
@@
 %
-A system satisfies {\em lockstep stack safety} if for every initial machine $\mach$,
-the trace induced from $(\mach, \pinit(\mach))$ satisfies trace lockstep stack safety with
+A system satisfies {\em lockstep stack safety} if for every initial machine $(\mach,\pol)$,
+the trace induced from $(\mach, \pol)$ satisfies trace lockstep stack safety with
 respect to the empty stack.
@@
       confidentiality. Let \(\callmap\) be a callmap; for each initial
-      state \(\mach_0\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
+      state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
       to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
-      \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\), take any
+      \(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\), take any
       variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let

### 2020-07-10 main.tex
Fixed duplicate in 5.6

@@
 %
-\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]
-
-For all such pairs of traces, \(\propdef{TraceConf} ~
-\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
-Trace confidentiality must also hold on the trace as a whole, so that
-unallocated memory is still secret. Let \(\mach_0'
-\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
-\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
-
+\[\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
+%
+%For all such pairs of traces, \(\propdef{TraceConf} ~
+%\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
+%Trace confidentiality must also hold on the trace as a whole, so that
+%unallocated memory is still secret. Let \(\mach_0'
+%\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
+%\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
+%
 If trace confidentiality holds for all such traces,
@@
 \machT_{\mathit{call}}\) and it also holds for the toplevel execution
-trace \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
-system as a whole enjoys {\em stack confidentiality} with respect to
-\(\callmap\).
+trace to account for unallocated memory \(\propdef{TraceConf} ~
+\emptyset ~ \MPT ~ \machT\) , then the system as a whole enjoys {\em
+  stack confidentiality} with respect to \(\callmap\).
 

### 2020-07-10 main.tex
CP

--- 2020-07-10 main.tex
@@
 
-We now present our model of enforcement, which we introduced
-informally in \Cref{sec:threat}.  A policy is a general model of an
-enforcement mechanism, consisting of a set of policy states \(\pol \in
-\POLS\), a policy step function \((\mach, \pol) \polstep \pol' :
-\MACHS \times \POLS \rightharpoonup \POLS\). We assume that the
-compiler provides an initial policy state $\pol_0$ to accompany 
-the initial machine state.
+We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
+A policy is a general model of an enforcement mechanism, consisting of a
+set of policy states \(\pol \in \POLS\), a policy step function
+\((\mach, \pol) \polstep \pol' : \MACHS \times \POLS
+\rightharpoonup \POLS\), and an initializer \(\pinit : \MACHS \rightharpoonup
+\POLS\). The initializer establishes the initial policy state for a given
+machine state, but is partial to reflect that some policies are only defined
+on states that are well-formed in some way.
+%
+\apt{Is this our current thinking?}
+\leo{I thought this was what we settled on in the last two days'
+  discussions}\bcp{That was not my impression!  I don't see how the policy
+  state can be recovered from just the machine state---e.g., the machine
+  state doesn't say where the calls and returns are, does it?!}
 %
@@
 
-In this section we formalize our definition of stack safety. We begin
-by describing our model of knowledge about the program structure, and
+We are finally ready for our definition of stack safety. We begin
+by describing how we model knowledge about the program structure,
 then dive into the formal definitions of local state encapsulation and
-well-bracketed control flow that together constitute stack safety.
+well-bracketed control flow, which together constitute stack safety.
 
@@
 We say that a system enjoys {\em stack integrity} if for every initial
-state $(\mach_0,\pol_0)$ and its corresponding induced trace $\MPT$ (with
-$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), for every
+state $\mach_0$ and its corresponding induced trace $\MPT$ (with
+$\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT$), for every
 $n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
@@
 %Let $\callmap$ be a call map and $\codemap$ a function map.
-For every initial state \((\mach_0,\pol_0)\), we take the induced trace
-\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\). Then for
+For every initial state \(\mach_0\), we take the induced trace
+\(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\). Then for
 any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
@@
 \(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
-\machT_{\mathit{call}}\) and it also holds for the toplevel execution
-trace \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
+\machT_{\mathit{call}}\) and
+\(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
 system as a whole enjoys {\em stack confidentiality} with respect to
@@
 %\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
-if, for any initial state \((\mach_0, \pol_0)\) and any adjacent pair of states
+if, for any initial state \(\mach_0\) and any adjacent pair of states
 \(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
-in the induced trace from \(\mpstate{\mach_0}{\pol_0}\),
+in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\),
 if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
@@
 %\(\callmap\) and entry map \(\entmap\)
-if, for any initial state \((\mach_0, \pol_0)\)
+if, for any initial state \(\mach_0\)
 and any n-transition from \((\mach_c,\pol_c)\) to
 \((\mach_e, \pol_e)\) in the induced trace from
-\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a valid entry:
+\(\mpstate{\mach_0}{\pinit(\mach_0)}\), $\mach_e$ is a valid entry:
 \[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]
@@
 A system enjoys {\em return integrity} if all of its returns match prior calls
-and return to the appropriate location. For any initial state \((\mach_0\, \pol_0)\), let
-\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\).
+and return to the appropriate location. For any initial state \(\mach_0\), let
+\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pinit(\mach_0)}\).
 \(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
@@
 %
-A system satisfies {\em lockstep stack safety} if for every initial machine $(\mach,\pol)$,
-the trace induced from $(\mach, \pol)$ satisfies trace lockstep stack safety with
+A system satisfies {\em lockstep stack safety} if for every initial machine $\mach$,
+the trace induced from $(\mach, \pinit(\mach))$ satisfies trace lockstep stack safety with
 respect to the empty stack.
@@
       confidentiality. Let \(\callmap\) be a callmap; for each initial
-      state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
+      state \(\mach_0\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
       to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
-      \(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\), take any
+      \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\), take any
       variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let

### 2020-07-10 main.tex
Clean up 7.5

--- 2020-07-10 main.tex
@@
 
-      {\em Observable trace confidentiality} is the lazy analog to trace
-      confidentiality, and reuses the cases in which the traces do not return.
-      For a return predicate \(R\) and a pair of traces
-      \(\MPT\) and \(\machT\), we consider the return case, in which
-      \(\last(\MPT)\) is some \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and
-      \(R ~ \mach_r\). Then \(\last(\machT)\) is some \((\mach_r',\obs_r')\)
-      and \(R ~ \mach_r'\). As above, the observations of \(\MPT\) are similar
-      to those of \(\machT\). In place of the requirement that \(\mach_r\)
-      and \(\mach_r'\) agree on components, we have the following requirement.
-      \rb{This reads a little funny, ending this paragraph with the promise of a
-      requirement and starting the next one introducing a conclusion.}
-
-      Finally, let \(\mach_e\) be the machine state of \(\head(\MPT)\),
-      and \(\mach_e'\) be that of \(\head(\machT)\). We restore \(\mach_r'\)
-      based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
-      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).
-      The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
+      {\em Observable trace confidentiality} is the lazy analog to
+      trace confidentiality, and reuses the cases in which the traces
+      do not return.  For a return predicate \(R\) and a pair of
+      traces \(\MPT\) and \(\machT\), we consider the return case, in
+      which $\MPT$ has a last element
+      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and that machine state is
+      a return: \(R ~ \mach_r\).
+%
+      Observable trace confidentiality then requires that $\machT$
+      also has a final state \((\mach_r',\obs_r')\) that is a return:
+      \(R ~ \mach_r'\).
+%
+      In addition, just like in the eager case, 
+      observations of \(\MPT\) are similar to those of \(\machT\).
+
+      The only thing that changes is the requirement that \(\mach_r\)
+      and \(\mach_r'\) agree on components. Instead, let \(\mach_e\)
+      be the machine state of \(\head(\MPT)\), and \(\mach_e'\) be
+      that of \(\head(\machT)\). We restore \(\mach_r'\) based on
+      \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
+      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).  Then
+      observable trace confidentiality requires that the
+      induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
       observational prefix of the induced trace from \(\mach_r'\),
@@
 
-      \leo{This is starting to be too spread out as a definition. Pack a bit?}
       \[\begin{split}
-        \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \triangleq
-        & \head(\MPT) = \mpostate{\mach_e}{\pol_e}{\obs_e} \Rightarrow \\
-        & \head(\machT) = (\mach_e',\obs_e') \Rightarrow \\
+        \mathit{ObsReturnCase} ~ R ~ (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~ ((\mach_e',\obs_e')\machT) \triangleq
         & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\

### 2020-07-10 main.tex
wip on 6.2

--- 2020-07-10 main.tex
@@
 A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
-$\TRACEELEMS$s; we write $\TRACE{\TRACEELEMS}$ for the set of traces over
-$\TRACEELEMS$. We will mostly be concerned with sequences of pairs of
-machine states
+$\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
 and observations, called machine traces and denoted $\machT \in \MOTRACE$,
@@
 coinductively relates a machine state with the trace of machine states
-and observations produced by repeated application of step:
+paired with observations produced by repeated application of step:
 
@@
 \noindent
-(We use juxtaposition of an element and a sequence to represent ``cons.'')
 Since the step function is total, if \(\mach \hookrightarrow \machT\),
@@
 see a depiction of such a trace: starting from an initial machine
-state at the very top, each instruction leads to a new machine state
-(the diff of which, compared to the previous one, appears in the
-``{\textit{Effect}'' column), and the majority of observations are $\tau$, except
-for instructions 31 and 13, which are writes to $\rout$
-where the observations are $0$ and $60$, respectively.
+state at the very top, each instruction constructs a new machine state
+(the diff of which compared to the previous one appears in the
+``Effect'' column) and the majority of observations are $\tau$ with
+the exception of instructions 31 and 13 which are writes to $\rout$
+and where the observations are $0$ and $60$ respectively.
 
@@
 %the trailing observation. Here's an attempt.}
-takes a trace\bcp{I changed a : here to $\in$.  We should check if there are
-any more to fix.}
-\(\Trace \in \TRACE{\TRACEELEMS} \times \OBSS\) and a predicate
+takes a trace
+\(\Trace : \mathit{trace} ~ \TRACEELEMS \times \OBSS\) and a predicate
 on elements \(f \subset \TRACEELEMS\) and gives the prefix of
 \(\Trace\) ending with the first element on which \(f\) holds.
-One subtlety is that, because the observation associated with some element
-in a trace
-represents what can be observed when this element is executed (i.e., the
-observation on that step), when we take a prefix from a trace containing
-observations
+Subtly, because the observation associated with a trace represents its
+next step, when we take a prefix from a trace containing observations
 we must replace the final observation with a \(\tau\).
@@
 If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate never holds, then
+predicate must hold on its final element. If the predicate never holds,
 \(\PUT\) is the identity function.
@@
 %
-If \(f\) be a predicate on machine states, we define \(\mach
+Let \(f\) be a predicate on machine states. Then we define \(\mach
 \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
-running \(\mach\) up to \(f\)'')
+running \(\mach\) up to \(f\)''):
 
@@
 \noindent
-and similarly for traces with policies:
+And similarly for traces with policies:
 
@@
 We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-as far as an external observer is concerned, written \(\obsT_1 \lesssim
-\obsT_2\), if the sequence of
+as far as an external observer is concerned, if the sequence of
 non-silent observations of $\obsT_1$ is a prefix of those of
@@
 
-\smallskip
 We then define similarity of observation traces as traces prefixing each other:
+
 \[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
@@
 Note that an infinite silent trace is a
-prefix of (and equivalent\bcp{similar?} to) any other trace. While this might seem
+prefix of (and equivalent to) any other trace. While this might seem
 surprising at first, it makes sense in a timing-insensitive context:
@@
 
-We are finally ready for our definition of stack safety. We begin
-by describing how we model knowledge about the program structure,
+In this section we formalize our definition of stack safety. We begin
+by describing our model of knowledge about the program structure, and
 then dive into the formal definitions of local state encapsulation and
-well-bracketed control flow, which together constitute stack safety.
+well-bracketed control flow that together constitute stack safety.
 
@@
 %
-\[\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
-%
-%For all such pairs of traces, \(\propdef{TraceConf} ~
-%\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
-%Trace confidentiality must also hold on the trace as a whole, so that
-%unallocated memory is still secret. Let \(\mach_0'
-%\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
-%\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
-%
+\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]
+
+For all such pairs of traces, \(\propdef{TraceConf} ~
+\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
+Trace confidentiality must also hold on the trace as a whole, so that
+unallocated memory is still secret. Let \(\mach_0'
+\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
+\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
+
 If trace confidentiality holds for all such traces,
 \(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
-\machT_{\mathit{call}}\) and it also holds for the toplevel execution
-trace to account for unallocated memory \(\propdef{TraceConf} ~
-\emptyset ~ \MPT ~ \machT\) , then the system as a whole enjoys {\em
-  stack confidentiality} with respect to \(\callmap\).
+\machT_{\mathit{call}}\) and
+\(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
+system as a whole enjoys {\em stack confidentiality} with respect to
+\(\callmap\).
 
@@
 
-This conservative policy satisfies not only the stack safety above,
+This conservative policy satisfies not only the stack safety policy of
+\cref{sec:lse-and-wbcf},
 but also a provably (in Coq) stronger one that enforces local state
-encapsulation throughout the execution. To encode that
+encapsulation at every point during execution. To encode that
 property, we not only keep track of the execution of one machine
@@
 top of the variant stack (just like the stack confidentiality property
-enforced trace confidentiality for every transition), and every time
+enforced trace confidentiality for every transition\apt{can we change ``transition'' to ``call transition'' at least sometimes? -- I keep thinking it means ``step''}), and every time
 the machine returns from a call we pop the top variation (to account
@@
   \]
+\apt{We could have used this to define trace integrity in the first place...}
 
@@
   \]
+\apt{And the last clause here could have been used for trace confidentiality...}
 
@@
 for every nested call performed so far on a {\em variant stack}: a
-list of machine states annotated with a contour, each element of which
-(denoted $\vm{C}$) of which is a variation of the actual machine
+list of machine states annotated with a contour\apt{one contour per list or one contour per stack??}, each element of which
+(denoted $\vm{C}$) is a variation of the actual machine
 according to a contour $C$.

### 2020-07-10 main.tex
CP

--- 2020-07-10 main.tex
@@
 
-\definition{Well-Bracketed Control Flow}
-
+\definition% {Well-Bracketed Control Flow}
 The above properties are somewhat orthogonal, but together they
@@
 explicit entry points, and returns respect implicit entry points
-corresponding to their call.  A system that enjoys all three
+corresponding to their call.  A system with all three
 properties (with respect to a function map, call map, return map,
@@
 $(\mach, \pol)$ satisfies the \emph{trace lockstep stack safety}, and write
-$\testProp{}{\MPT}{\vs}$, if:
+$\testProp{}{\MPT}{\vs}$, if the following conditions are satisfied:
 
@@
 
-The proof of this theorem is a fairly involved coinductive proof
+The proof is a fairly involved coinductive argument
 (around 3K LoC in Coq) on the definition trace lockstep stack safety,
@@
       \[\begin{split}
-        \propdef{ObsTraceInt} ~ \contour ~ \MPT \triangleq
+        \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
         & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
@@
       \integritylazyexample
-      \caption{Rolling back from Figure \ref{fig:intex}}
+      \caption{Rolling back from Figure \ref{fig:intex}\bcp{too much space!}}
       \label{fig:intlex}
@@
 
-      {\em Observable trace confidentiality} is the lazy analog to
-      trace confidentiality, and reuses the cases in which the traces
-      do not return.  For a return predicate \(R\) and a pair of
-      traces \(\MPT\) and \(\machT\), we consider the return case, in
-      which $\MPT$ has a last element
-      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and that machine state is
-      a return: \(R ~ \mach_r\).
-%
-      Observable trace confidentiality then requires that $\machT$
-      also has a final state \((\mach_r',\obs_r')\) that is a return:
-      \(R ~ \mach_r'\).
-%
-      In addition, just like in the eager case, 
-      observations of \(\MPT\) are similar to those of \(\machT\).
-
-      The only thing that changes is the requirement that \(\mach_r\)
-      and \(\mach_r'\) agree on components. Instead, let \(\mach_e\)
-      be the machine state of \(\head(\MPT)\), and \(\mach_e'\) be
-      that of \(\head(\machT)\). We restore \(\mach_r'\) based on
-      \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
-      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).  Then
-      observable trace confidentiality requires that the
-      induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
+      {\em Observable trace confidentiality} is the lazy analog to trace
+      confidentiality, and reuses the cases in which the traces do not return.
+      For a return predicate \(R\) and a pair of traces
+      \(\MPT\) and \(\machT\), we consider the return case, in which
+      \(\last(\MPT)\) is some \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and
+      \(R ~ \mach_r\). Then \(\last(\machT)\) is some \((\mach_r',\obs_r')\)
+      and \(R ~ \mach_r'\). As above, the observations of \(\MPT\) are similar
+      to those of \(\machT\). In place of the requirement that \(\mach_r\)
+      and \(\mach_r'\) agree on components, we have the following requirement.
+      \rb{This reads a little funny, ending this paragraph with the promise of a
+      requirement and starting the next one introducing a conclusion.}
+
+      Finally, let \(\mach_e\) be the machine state of \(\head(\MPT)\),
+      and \(\mach_e'\) be that of \(\head(\machT)\). We restore \(\mach_r'\)
+      based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
+      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).
+      The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
       observational prefix of the induced trace from \(\mach_r'\),
@@
 
+      \leo{This is starting to be too spread out as a definition. Pack a bit?}
       \[\begin{split}
-        \mathit{ObsReturnCase} ~ R ~ (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~ ((\mach_e',\obs_e')\machT) \triangleq
+        \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \ \ \triangleq\ \
+        & \head(\MPT) = \mpostate{\mach_e}{\pol_e}{\obs_e} \Rightarrow \\
+        & \head(\machT) = (\mach_e',\obs_e') \Rightarrow \\
         & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
@@
       \[\begin{split}
-        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \triangleq
+        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
         & \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \land \\
@@
         \confidentialitylazyexample
-        \caption{Continuing from \cref{fig:confex}}
+        \caption{Continuing from \cref{fig:confex}\bcp{Spacing!}}
         \label{fig:conflex}

### 2020-07-10 main.tex
Comment squash

@@
 
-\definition% {Well-Bracketed Control Flow}
+\definition{Well-Bracketed Control Flow}
+
 The above properties are somewhat orthogonal, but together they
@@
 explicit entry points, and returns respect implicit entry points
-corresponding to their call.  A system with all three
+corresponding to their call.  A system that enjoys all three
 properties (with respect to a function map, call map, return map,
@@
 $(\mach, \pol)$ satisfies the \emph{trace lockstep stack safety}, and write
-$\testProp{}{\MPT}{\vs}$, if the following conditions are satisfied:
+$\testProp{}{\MPT}{\vs}$, if:
 
@@
 
-The proof is a fairly involved coinductive argument
+The proof of this theorem is a fairly involved coinductive proof
 (around 3K LoC in Coq) on the definition trace lockstep stack safety,
@@
       \[\begin{split}
-        \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
+        \propdef{ObsTraceInt} ~ \contour ~ \MPT \triangleq
         & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
@@
       \integritylazyexample
-      \caption{Rolling back from Figure \ref{fig:intex}\bcp{too much space!}}
+      \caption{Rolling back from Figure \ref{fig:intex}}
       \label{fig:intlex}
@@
 
-      \leo{This is starting to be too spread out as a definition. Pack a bit?}
       \[\begin{split}
-        \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \ \ \triangleq\ \
-        & \head(\MPT) = \mpostate{\mach_e}{\pol_e}{\obs_e} \Rightarrow \\
-        & \head(\machT) = (\mach_e',\obs_e') \Rightarrow \\
+        \mathit{ObsReturnCase} ~ R ~ (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~ ((\mach_e',\obs_e')\machT) \triangleq
         & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
@@
       \[\begin{split}
-        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
+        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \triangleq
         & \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \land \\
@@
         \confidentialitylazyexample
-        \caption{Continuing from \cref{fig:confex}\bcp{Spacing!}}
+        \caption{Continuing from \cref{fig:confex}}
         \label{fig:conflex}
@@
 
-      This is more straightforward \rb{than what?} (1,5K LoC in Coq) as the recursive
+      This is more straightforward than the implication in the previous
+      section (1,5K LoC in Coq) as the recursive
       structure of the two properties is similar.

### 2020-07-10 main.tex

--- 2020-07-10 main.tex
@@
 control\bcp{we need to either weaken the previous sentence or else add
-  ``first steps toward'' in this sentence...}\leo{Weakened enough?} flow using the technical
+  ``first steps toward'' in this sentence...}\leo{Weakened enough?}\bcp{Not
+  sure: it still says ``generic,'' and it's not clear why we're saying it's
+  a step rather than that we achieved the goal...} flow using the technical
 framework of language-based security.
@@
      ``stack-safety,'' I'm fixing on the former.}
+\bcp{Be careful: When a compound phrase is used as an adjective to modify a
+  following noun, it gets hyphenated; otherwise it does not.  This is
+  correct grammar, not inconsistency!  :-)}
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A new {\em formal stack safety property} capturing the
+  A novel {\em formal stack safety property} formalizing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
       \[\begin{split}
-        \mathit{ObsReturnCase} ~ R ~ (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~ ((\mach_e',\obs_e')\machT) \triangleq
+        \mathit{ObsReturnCase} ~ R ~
+        (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~
+        ((\mach_e',\obs_e')\machT) \ \ \triangleq \ \ 
         & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\

### 2020-07-10 main.tex
wip

@@
 control\bcp{we need to either weaken the previous sentence or else add
-  ``first steps toward'' in this sentence...}\leo{Weakened enough?}\bcp{Not
-  sure: it still says ``generic,'' and it's not clear why we're saying it's
-  a step rather than that we achieved the goal...} flow using the technical
+  ``first steps toward'' in this sentence...}\leo{Weakened enough?} flow using the technical
 framework of language-based security.
@@
      ``stack-safety,'' I'm fixing on the former.}
-\bcp{Be careful: When a compound phrase is used as an adjective to modify a
-  following noun, it gets hyphenated; otherwise it does not.  This is
-  correct grammar, not inconsistency!  :-)}
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A novel {\em formal stack safety property} formalizing the
+  A new {\em formal stack safety property} capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
 for every nested call performed so far on a {\em variant stack}: a
-list of machine states annotated with a contour\apt{one contour per list or one contour per stack??}, each element of which
+list of machine states annotated with a contour\apt{one contour per list or one contour per state??}, each element of which
 (denoted $\vm{C}$) is a variation of the actual machine
@@
       \[\begin{split}
-        \mathit{ObsReturnCase} ~ R ~
-        (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~
-        ((\mach_e',\obs_e')\machT) \ \ \triangleq \ \ 
+        \mathit{ObsReturnCase} ~ R ~ (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~ ((\mach_e',\obs_e')\machT) \triangleq
         & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\

### 2020-07-10 main.tex
done with 6

@@
 for every nested call performed so far on a {\em variant stack}: a
-list of machine states annotated with a contour\apt{one contour per list or one contour per state??}, each element of which
+list of machine states annotated with a contour, each element of which
 (denoted $\vm{C}$) is a variation of the actual machine
@@
 to the original machine state. We will write $\testVS{(\mach,
-  \pol)}{\vs}$ for this compound check.
+  \pol)}{\vs}$ for this compound check. \apt{I'm having trouble here and below with
+  the scoping of $C$. Each variant stack is for one particular $C$, right?
+  So why not write $vs_C$ rather than $vm_C$ ?}
 %% >>>>>>> 1886f7a7bf858ce0c9518e407d9481518a4f028d
@@
 $\vs$ in which every element $\vm{\contour}$ is a variation of $\mach$
-with respect to its contour, we say that an MP-trace induced from
+with respect to its \apt{whose?} contour, we say that an MP-trace induced from
 $(\mach, \pol)$ satisfies the \emph{trace lockstep stack safety}, and write

### 2020-07-10 main.tex
Nit

--- 2020-07-10 main.tex
@@
 for every nested call performed so far on a {\em variant stack}: a
-list of machine states annotated with a contour\apt{one contour per list or one contour per state??}, each element of which
+list of machine states annotated with a contour\apt{one contour per list or one contour per stack??}, each element of which
 (denoted $\vm{C}$) is a variation of the actual machine
@@
 
-We are confident that the linear capability machines introduced by
+We believe that the linear capability machines introduced by
 \citep{Skorstengaard+19}, in combination with the StkTokens calling convention,

### 2020-07-10 main.tex
CP

--- 2020-07-10 main.tex
@@
 stack frames) and an integrity property (callees should not corrupt their
-caller's stack frames or control state).
+caller's stack frames or control state), together with a straightforward
+well-bracketedness condition on the program's control flow.
   %
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A novel {\em formal stack safety property} formalizing the
+  A novel {\em stack safety property} abstractly capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
-  control flow~(\cref{sec:lse-and-wbcf}) in terms of familiar security concepts.
+  control flow~(\cref{sec:lse-and-wbcf}).
   % These properties rule out common examples of stack memory abuse.

### 2020-07-10 main.tex
CP

@@
   well-bracketed control flow}.
-These properties are formalized by defining an abstract capability-based machine
+These properties are formalized by defining a capability-based machine
 that ``satisfies [stack safety] by construction.''
@@
 While this work represents an important step in the right direction, we believe
-there is still room to craft a more abstract, more generic characterization
+there is still room to craft a more abstract characterization
 of stack safety that is independent of capabilities or any other specific
@@
 %
-Our key contribution is another important step towards this goal: a generic
-presentation of local state encapsulation and well-bracketed
-control\bcp{we need to either weaken the previous sentence or else add
-  ``first steps toward'' in this sentence...}\leo{Weakened enough?}\bcp{Not
-  sure: it still says ``generic,'' and it's not clear why we're saying it's
-  a step rather than that we achieved the goal...} flow using the technical
-framework of language-based security.
+Our key contribution is an abstract
+presentation of local state encapsulation using the technical
+framework of language-based security, together with a simple
+characterization of well-bracketed control flow.
 

### 2020-07-10 main.tex
CP

@@
 technique.  In other words, we seek formal criteria that can be applied
-to \emph{any} mechanism for protecting stack-based low-level programs to judge
+to a mechanism for protecting stack-based low-level programs to judge
 whether or not it actually ``guarantees stack safety.''
@@
 entry point or returns that re-enter just after their associated call
-in the standard nested call structure.  \apt{clarify what we are
-  contributing here. We are not trying to phrase this as a security
-  property.}\leo{Better now? Should this be fixed two paragraphs above
-  too?}\bcp{Hmmm this seems like a fundamental issue with the way we are
-  framing the story...}
-
-To phrase stack safety, and especially local state encapsulation, as a
+in the standard nested call structure.
+
+To phrase local state encapsulation as a
 security property requires an explicit attacker model: a clear

### 2020-07-10 main.tex

@@
 We discuss how these properties are enforced or approximated by
-existing stack protection mechanisms and describe some useful
+existing stack protection mechanisms and describe useful
 stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
@@
 an execution trace. The advantage of this property is that it ``fails
-faster'' and os therefore more suitable for efficient testing. We prove (in
+faster'' and is therefore more suitable for efficient testing. We prove (in
 Coq) that this ``lockstep'' property implies our original definition of stack
@@
 frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
-  property inthe same way...}\fi.
+  property in the same way...}\fi.
 The lazy policy does not enforce our original integrity and confidentiality
@@
 
-Enforcement mechanisms are modeled as {\em policies}\bcp{this is a somewhat
-  special use of the word ``policy'' that comes from the micropolicies
-  world.  We should include a comment about this to avoid confusing
-  readers.} that add some auxiliary
-policy state to the ordinary machine state
+Enforcement mechanisms are modeled as {\em policies} (borrowing the
+terminology from tag-based mechanisms) that add some auxiliary policy
+state to the ordinary machine state
 %\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
@@
 
-In the rest of this section we implictly parameterize all
+In the rest of this section we implicitly parameterize all
 definitions by a function map $\codemap$, a call map $\callmap$, a return
@@
 top of the variant stack (just like the stack confidentiality property
-enforced trace confidentiality for every transition\apt{can we change ``transition'' to ``call transition'' at least sometimes? -- I keep thinking it means ``step''}), and every time
+enforced trace confidentiality for every call transition), and every time
 the machine returns from a call we pop the top variation (to account
@@
     \]
+    \ifaftersubmission
 \apt{We could have used this to define trace integrity in the first place...}
-
+\fi
 \definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
@@
 \]
+\ifaftersubmission
 \apt{And the last clause here could have been used for trace confidentiality...}
+\fi
 
@@
 for every nested call performed so far on a {\em variant stack}: a
-list of machine states annotated with a contour, each element of which
-(denoted $\vm{C}$) is a variation of the actual machine
-according to a contour $C$.
+list of machine states, each annotated with its own contour written
+as $\vm{C}$), and each being a variation of the actual machine
+according to $C$.
 %
@@
 \pol) \stepstopol{\obs} (\mach', \pol')$, say that a variant
-stack $\vs$ safisfies the \emph{variant stack test} if for each
+stack $\vs$ satisfies the \emph{variant stack test} if for each
 variant element $\vm{C}$ (which is a variation of $\mach$ for $C$),
@@
 to the original machine state. We will write $\testVS{(\mach,
-  \pol)}{\vs}$ for this compound check. \apt{I'm having trouble here and below with
-  the scoping of $C$. Each variant stack is for one particular $C$, right?
-  So why not write $vs_C$ rather than $vm_C$ ?}
+  \pol)}{\vs}$ for this compound check.
+%\apt{I'm having trouble here and below with
+%  the scoping of $C$. Each variant stack is for one particular $C$, right?
+%  So why not write $vs_C$ rather than $vm_C$ ?}
+% LEO: Each machine has its own contour. Clarified above, hopefully
 %% >>>>>>> 1886f7a7bf858ce0c9518e407d9481518a4f028d
@@
 $\vs$ in which every element $\vm{\contour}$ is a variation of $\mach$
-with respect to its \apt{whose?} contour, we say that an MP-trace induced from
-$(\mach, \pol)$ satisfies the \emph{trace lockstep stack safety}, and write
-$\testProp{}{\MPT}{\vs}$, if the following conditions are satisfied:
+with respect to its annotated contour $\contour$, we say that an
+MP-trace induced from $(\mach, \pol)$ satisfies the \emph{trace
+  lockstep stack safety}, and write $\testProp{}{\MPT}{\vs}$, if the
+following conditions are satisfied:
 
@@
   and trace lockstep stack safety must be coinductively satisfied for
-  the remainder of the trace and for the tail of the varant stack
+  the remainder of the trace and for the tail of the variant stack
   that $\vs$ steps to:
@@
 \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
-    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
+    (\mach, \pol, \tau)$, the property is satisfied for the singleton trace:
     $\testProp{}{[(\mach, \pol)]}{\vs}$.
@@
 (around 3K LoC in Coq) on the definition trace lockstep stack safety,
-using an auxiliary depth-index to ensure the proper aligmnent of
+using an auxiliary depth-index to ensure the proper alignment of
 calls and returns.
@@
       induced trace, we take the prefix of the call
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} | \ret{\mach_c}\)
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\)
       and observable trace integrity holds on \(\MPT_{\mathit{call}}\) with
@@
       variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
-      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} |
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~
+      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~
       \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
@@
 
+    \leo{Should we clearpage this once we're sure of the layout?}
 \section{Related Work}
@@
   is that it uncovers errors quickly, and therefore should be quicker
-  to test just like the smarter testing properties
+  to test, just like the smarter testing properties
   of~\citet{TestingNI:ICFP}. We have already developed a testing

### 2020-07-10 main.tex
read through 3

@@
 program counter $\PCname$
-and stack pointer $\rsp$\bcp{but not a frame pointer?!}; other details of the machine's ISA are unimportant.
+and stack pointer $\rsp$\bcp{but not a frame pointer?!\apt{right; we're agnostic about that}}; other details of the machine's ISA are unimportant.
 
@@
 $\rsp$). \bcp{This implicitly amounts to choosing a caller-saves register
-  convention, though, doesn't it?}
+  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}
 
 Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code (the same instruction may be
+which are not fully explicit in machine code (the same instruction opcode may be
 used in some places as part of a call or return sequence and in other places
@@
 If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
-  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
+  return} has occurred when the $\PCname$ points to the next instruction after $a$ and $\rsp = s$. (Our definition
 of well-bracketed control flow relates this definition of return to the program's
@@
 \emph{varying} that data would not change the program's observable
-behavior~\citep{6234468}
+behavior~\citep{6234468}.
 In applying this idea to a system that incorporates policy enforcement, only
@@
 attached to each value in memory or registers, and the policy step
-function\bcp{undefined!}\leo{``partial policy step'' is explained informally above}
-checks that each machine operation obeys a set of rules on tags
+function checks that each machine operation obeys a set of rules on tags
 (e.g., that the current $\PCname$ tag matches the tag on the stack
@@
 the stack protection policies of
-\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
-  linear capabilities?}\leo{Only the second one AFAIK. What would you
-  change to characterize it more accurately?}\rb{StkPointers uses linear
-  capabilities, its predecessor users local capabilities (and unlike
-  StkPointers could be implemented in CHERI)}.
+\citet{Skorstengaard+19b} and (with an extension to support linear capabilities) \citet{Skorstengaard+19}.
+%\bcp{I thought these policies used
+%  linear capabilities?}\leo{Only the second one AFAIK. What would you
+%  change to characterize it more accurately?}\rb{StkPointers uses linear
+%  capabilities, its predecessor users local capabilities (and unlike
+%  StkPointers could be implemented in CHERI)}.
 Capabilities package up pointers
@@
 to insert stack canaries~\citep{Cowan+98}. Such approaches should
-still satisfy our formal notion of stack safety with a trivial null
-policy, but only for programs produced by the compiler.
+still satisfy our formal notion of stack safety \apt{Stack canaries certainly would not!} with a trivial null
+policy, but only for programs produced by the compiler.\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}
 %These approaches are represented as policies that
@@
 
-\Cref{fig:running-program} shows a high-level view of our running example,
+\Cref{fig:running-program} shows a high-level view of a running example,
 written in C-like pseudo-code.  It consists of a {\tt main} function that
@@
 execution trace of the corresponding machine instructions. These might have
-been generated by compiling the high-level code, but it is only the behavior
+been generated from the high-level code by a (rather simple-minded) compiler,
+but it is only the behavior 
 of the machine instructions that we care about
@@
 Instructions are intended to
-be similar to RISC-V, but are described in a (hopefully more readable) an
+be similar to RISC-V, but are described in a (hopefully more readable)
 ad-hoc notation
@@
 into address 101.
+
+The code itself is straightforward, and should be largely self-explanatory, but
+a few points deserve clarification. Functions return their result value on the
+stack rather than in a register.  Stack frames are built in two parts:
+space for local variables and the return address is allocated and initialized at function entry, and deallocated at function exit; space for arguments and return values is allocated and initialized immediately prior to each function call and deallocated on return.
+The return value slot is always initialized
+because some of our enforcement mechanisms will rely on this.
+\apt{this is bogus, but oh well}
+The return address is spilled to the stack even by leaf functions such as
+{\tt h} that could just leave it in $\rra$.
 
@@
 
+\apt{It would be good to relate this back to the running example. This is
+  complicated because the example passes arguments/returns, and it builds
+  and initializes this part of the frame separately.}
+
 %% APT: removed because the tag part is too mysterious and the instruction part

### 2020-07-10 main.tex
cp through 5.1

@@
 as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (static or dynamic, software- or
+machine model plus some kind of 
+% static or dynamic,
+(software- or
 hardware-enforced) policy monitor. In~\cref{sec:enforcement}, we will
@@
 %
-This step function is total: execution traces are infinite unless
-cut short externally by a policy fault (as explained below).
+This step function is total, so the machine's execution traces
+(defined in \cref{sec:traces}) are infinite.
 
@@
 \POLS\), a policy step function \((\mach, \pol) \polstep \pol' :
-\MACHS \times \POLS \rightharpoonup \POLS\). We assume that the
+\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial; 
+it is undefined on input configurations that correspond to a policy fault.
+We assume that the
 compiler provides an initial policy state $\pol_0$ to accompany
@@
 
+This combined step function is also partial, and traces of the machine-with-policy
+will be finite iff a policy fault occurs.
+
 \subsection{Traces}
+\label{sec:traces}
 
@@
 machine states
-and observations, called machine traces and denoted $\machT \in \MOTRACE$,
+and observations, called \emph{machine traces} and denoted $\machT \in \MOTRACE$,
 and with sequences of triples of a machine state, a policy state, and an
-observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.
+observation, called \emph{MP-traces} and denoted $\MPT \in \MPOTRACE$.
 
@@
 In our running example in \Cref{fig:running-trace-b} we can
-see a depiction of such a trace: starting from an initial machine
+see a depiction of (a prefix of) such a trace: starting from an initial machine
 state at the very top, each instruction leads to a new machine state
@@
 \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
-running \(\mach\) up to \(f\)'')
-
+running \(\mach\) up to \(f\)'') and similarly for traces with policies:
+
+\begin{center}
+\begin{minipage}{.3\textwidth}
 \judgmenttwo{\(\mach \hookrightarrow \machT'\)}
@@
             {\(\mach \hookrightarrow \machT | f\)}
-\noindent
-and similarly for traces with policies:
-
+\end{minipage}
+\begin{minipage}{.6\textwidth}
 \judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
@@
          {\(\mpstatename \hookrightarrow \MPT | f\)}
+\end{minipage}
+\end{center}
 
@@
 Note that an infinite silent trace is a
-prefix of (and equivalent\bcp{similar?} to) any other trace. While this might seem
+prefix of (and similar to) any other trace. While this might seem
 surprising at first, it makes sense in a timing-insensitive context:
@@
 
-Our properties and policies are parameterized by several types of
-annotations describing features of a program.
+Our properties and policies assume that the program is annotated
+with various information about its structure.
 
@@
 correspond to {\tt h}, which means that the function map would be:
-\[ \codemap = \{\texttt{main} \mapsto \{1\ldots13\}, \texttt{g} \mapsto \{20\ldots36\}, \texttt{h} \mapsto \{40\ldots47\}\}\]
+
+\[ \codemap(w) = 
+  \begin{cases}
+    \texttt{main} & \text{if } 1 \leq w \leq 13 \\
+    \texttt{g} & \text{if } 20 \leq w \leq 36 \\
+    \texttt{h} & \text{if } 40 \leq w \leq 47 \\
+    \mathit{undefined} & \text {otherwise}\\
+  \end{cases}
+\]
 
@@
 
-For concreteness, in a typical calling convention, a call map may be
-considered to identify the state just before the \jal, in which
-space for arguments is already allocated, but the mechanism is more
-general. A state in which a call map is defined on the program counter
+For concreteness, in a typical calling convention, the call map will
+identify the locations of the $\jal$ instructions representing calls.
+% APT: I found this more confusing than helpful...
+%, in which
+%space for arguments is already allocated, but the mechanism is more
+%general.
+A state in which a call map is defined on the program counter
 is referred to as a {\it call state}.
@@
 Similarly, a {\em return map} identifies addresses that complete a
-return sequence. Formally, it is simply a set of words; each one
+return sequence. Formally, it is a set of words; each one
 is the address of the final jump corresponding to a function return:
@@
 
-We separately identify, for any call state, what it means for a state
-to have correctly returned from a call. We consider the call to have
+We separately identify, for any call state, what it means to have 
+correctly returned from that call. We consider the call to have
 returned when control returns to the instruction following
-the \jal with the stack restored. Formally, this is a predicate on
+the $\jal$ with the stack restored. Formally, this is a predicate on
 states that says when the machine has just returned from a specified call state

### 2020-07-10 main.tex
Sections 1-2 pass.

@@
 The call stack is a perennial target for attacks
-based on memory safety vulnerabilities, with potential consequences
+based on memory-safety vulnerabilities, with potential consequences
 ranging from leakage or corruption of private stack data to
@@
 %
-We formalize this intuition in (1) a {\em stack confidentiality}
+We formalize this intuition in (1) a {\em stack-confidentiality}
 property, which promises that a callee's observable behavior is not affected
 by memory outside its stack frame (e.g., a caller's private data or
-uninitialized memory), plus (2) a {\em stack integrity} property,
+uninitialized memory), plus (2) a {\em stack-integrity} property,
 which promises that, if a function call ever returns, the caller's stack
@@
 conservative Depth Isolation policy, which not only enforces our basic
-stack safety property, but also an alternative lockstep property that
+stack-safety property, but also an alternative lockstep property that
 checks for integrity and confidentiality violations at every step of
@@
 detects violations between when they occur and when they make a visible impact.
-Unfortunately the existing Lazy Tagging and Clearing policy of
-\citeauthor{DBLP:conf/sp/RoesslerD18} does not quite enforce these properties,
+Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing ``Lazy Tagging
+and Clearing'' policy does not quite enforce these properties,
 but would if enhanced (at some performance cost).
@@
   correct grammar, not inconsistency!  :-)}
+\sna{Aha, I forgot that. Going hard the other way.}
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A novel {\em stack safety property} abstractly capturing the
+  A novel {\em stack-safety property} abstractly capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
 \item
-  A {\em lazy} variant of the stack safety property that can be enforced if the
+  A {\em lazy} variant of the stack-safety property that can be enforced if the
   Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
@@
 consists of an instruction region and a stack that grows upwards. (For simplicity,
-we do not consider the heap. Integrating a memory safety property for the
+we do not consider the heap. Integrating a memory-safety property for the
 heap in the style of
@@
 
-Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
+Our stack-safety concepts depend critically on the notions of ``call'' and ``return,''
 which are not fully explicit in machine code (the same instruction opcode may be
@@
 compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
-to insert stack canaries~\citep{Cowan+98}. Such approaches should
-still satisfy our formal notion of stack safety \apt{Stack canaries certainly would not!} with a trivial null
-policy, but only for programs produced by the compiler.\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}
+to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
+be evaluated in terms of our formal notion of stack safety with a trivial null
+policy. But any state maintained by the enforcement mechanism must be in normal
+memory and subject to the same confidentiality and integrity concerns (to be
+varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
+out as policy state provided that it can be proven to protect itself.
+
+\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}
 %These approaches are represented as policies that
@@
 Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
-the formal stack safety property described in the previous section: it in
+the formal stack-safety property described in the previous section: it in
 fact enforces a stronger, inductive version of that property which strengthens the
@@
 at exactly those points in
-the program where the contour of the stack safety property changes:
+the program where the contour of the stack-safety property changes:
 
@@
 There remains the question of how to ensure that the sequences are
-invoked at the right places. For a program to enjoy the stack safety property
+invoked at the right places. For a program to enjoy the stack-safety property
 (with respect to a particular call map), all we need require is that
@@
 If the code fails to initiate an exit
-sequence at a point where the stack safety property expects a return,
+sequence at a point where the stack-safety property expects a return,
 the micropolicy will incorrectly behave as if execution
@@
 a variation of the original machine state at the time of that call on
-top of the variant stack (just like the stack confidentiality property
+top of the variant stack (just like the stack-confidentiality property
 enforced trace confidentiality for every call transition), and every time
@@
 
-  Part of the motivation for the strong lockstep stack safety property
+  Part of the motivation for the strong lockstep stack-safety property
   is that it uncovers errors quickly, and therefore should be quicker

### 2020-07-10 main.tex
done through 5

--- 2020-07-10 main.tex
@@
     \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[103] \leftarrow 0}
-        {Zero out result\apt{why?}\bcp{+1}}
+        {Zero out result}
     \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
@@
     \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp - 1] \leftarrow r_0}{[107] \leftarrow 0}
-        {Zero out result \apt{why?}}
+        {Zero out result}
     \row{\jal ~ \rra ~ 40}{\rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow 40}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 40}
@@
 Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
-\(\MPOTRACE\). We say \(\MPT\) satisfies {\em trace integrity} with
+MP-trace. say \(\MPT\) satisfies {\em trace integrity} with
 respect to \(\contour\) if its final state (if any) agrees with its
@@
   \propdef{TraceInt} ~ \contour ~ \MPT \triangleq &
-    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
-  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
+    \head(\MPT) = \mpostate{\mach_e}{\_}{\_} \land
+    \last(\MPT) = \mpostate{\mach_r}{\_}{\_} \Rightarrow \\
   & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
@@
 \end{split}\]
-\sna{Experimenting with font on the mathy version of the property.}
+\sna{Experimenting with font on the mathy version of the property.\apt{looks fine to me}}
 
@@
 integrity is trivially satisfied. Similarly, the callee is allowed to
-temporarily overwrite these locations, as long as they are restored
-once control returns. We will see variations of this later on: a
+temporarily overwrite the caller's private locations,
+as long as their values are restored
+by the time the callee returns. We will see variations of this later on: a
 stronger, inductive integrity property that ensures that the caller's
-private data is never overwritten, even during the caller's execution
+private data is never overwritten, even during the callee's execution
 (which leads to easier reasoning and more effective testing);
 and a weaker, lazy integrity property that only ensures the preservation
-of locations that subsequently affect the caller's observable actions
+of locations that subsequently affect the program's observable actions
 (which allows for a more efficient enforcement mechanism).
@@
 
-\definition Given a trace \(\MPT\) and a call map \(\callmap\)
+\definition Given a trace \(\MPT\), % and a call map \(\callmap\)
 %and a function map \(\codemap\),
@@
 machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
-\(\mpstate{\mach_e}{\pol_e}\) in the trace such that
+\(\mpstate{\mach_e}{\pol_e}\) in \(\MPT\) such that
 \(\mach_c\) is a valid call with \(n\) arguments, and
@@
 state $(\mach_0,\pol_0)$ and its corresponding induced trace $\MPT$ (with
-$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), for every
+$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), and for every
 $n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
@@
 \[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]
+\apt{This could still be clearer. Interleaving informal and mathy in the
+same sentence doesn't work so well.}
 
@@
 overwritten by a callee, confidentiality ensures that the caller's data never
-influences the callee's observable behavior. We will formally define
-confidentiality in two similar steps: a trace version and a system
+influences the callee's observable behavior. Again, we will formally define
+confidentiality in two steps: a trace version and a system
 version.
@@
 \definition%{Variations}
-We say two machine states \(\mach\) and \(\mach'\) are {\em variations}
+Two machine states \(\mach\) and \(\mach'\) are {\em variations}
 of one another given a contour \(\contour\) if they agree on low
@@
 ``original'' trace $\MPT \in \MPOTRACE$ that corresponds to a callee $f$
-and a ``variant'' trace $\machT \in \MTRACE$ that corresponds to how $f$
+and a ``variant'' trace $\machT \in \MOTRACE$ that corresponds to how $f$
 would behave if it were started from a variant initial machine state.
@@
 %
-This example will end up a violation of confidentiality, as a
-difference in a confidential location (\textcolor{cyan}{42} vs
-\textcolor{magenta}{5} in location 100) leads to a visible
+This example will end up with a violation of confidentiality, as a
+variation in a confidential location (\textcolor{cyan}{42} vs
+\textcolor{magenta}{-5} in location 100) leads to a visible
 difference in return value.
@@
   \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
-    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
+    {\_}{\_}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
     some final state \(\mach_r'\) and any changes from the initial states are
@@
       \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
-      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
-        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
-      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
-        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
-      & ~ (\mach_e(\component) \not = \mach_r(\component)
-        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
-      \mach_r(\component) = \mach_r'(\component))
+      \last(\MPT) =
+        \mpostate{\mach_r}{\_}{\_} ~ \land ~ R ~ \mach_r \Rightarrow \\
+      & \exists \mach_r' . \last(\machT) = (\mach_r',\_)
+        \land \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land \\
+      & \forall ~ \component . (\mach_e(\component) \not = \mach_r(\component)
+        \lor \mach_e'(\component) \not = \mach_r'(\component)) \Rightarrow
+      \mach_r(\component) = \mach_r'(\component)
     \end{split}\]
@@
     \[\begin{split}
-      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
+      \mathit{InfCase} ~ \MPT ~ \machT \triangleq
       & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
@@
       \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
-      & \forall \mach_r ~ \pol_r ~ \obs_r .
-        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
-      & \neg ~ R ~ \mach_r \Rightarrow \\
+      & \forall \mach_r .
+        \last(\MPT) = \mpostate{\mach_r}{\_}{\_} ~ \land ~ \neg ~ R ~ \mach_r \Rightarrow \\
       & \pi_o(\MPT) \lesssim \pi_o(\machT)
@@
 
-\[\begin{split}
-  \propdef{TraceConf} ~ R ~ \MPT ~ \machT \triangleq
-  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
-  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
-    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
-\end{split}\]
+\[  \propdef{TraceConf} ~ R ~ \MPT ~ \machT \triangleq
+   \mathit{ReturnCase} ~ R ~ \MPT ~ \machT ~ \land 
+   \mathit{InfCase} ~ \MPT ~ \machT \land
+    \mathit{FailCase} ~ R ~ \MPT ~ \machT 
+\]
 
@@
 \(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
-\machT_{\mathit{call}}\) and it also holds for the toplevel execution
-trace to account for unallocated memory \(\propdef{TraceConf} ~
+\machT_{\mathit{call}}\), and it also holds for the toplevel execution
+trace to account for unallocated memory, \(\propdef{TraceConf} ~
 \emptyset ~ \MPT ~ \machT\) , then the system as a whole enjoys {\em
@@
 because the target (instruction 42) of the new $\jalr$ is not an established
-entry point according to the entry map. Our properties to not enforce which
-call points jump to which entry points, however; that is left for a stronger
+entry point according to the entry map. Our properties do not enforce which
+call points jump to which entry points, however; that would require a stronger
 notion of control flow integrity.
@@
 
+\begin{center}
 \begin{minipage}{.3\textwidth}
@@
 \end{minipage}
+\end{center}
 
@@
 and return to the appropriate location. For any initial state \((\mach_0\, \pol_0)\), let
-\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\).
-\(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
-may not hold for any \(\MPT_{tail}\). For any state pair \((\mach_c, \pol_c)\)
+\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\). Then two
+conditions must hold. 
+\emph{(i)} \(\MPT\) must not have an unmatched return, i.e., \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
+must not hold for any \(\MPT_{tail}\). \emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
 corresponding to a call \(\MPT\), if
-\(\mpstatename = \mpostate{\mach_r}{\pol_r}{\obs_r}\)
+\(\mpstatename = \mpostate{\mach_r}{\_}{\_}\)
 is the first unmatched return such that
-\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{\MPT_{tail}} = \mpstatename\),
-then $\mach_r$ must correspond to a return state for $\mach_c$:
-\(\ret{\mach_c}\ \mach_r\)
+\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{(\MPT_{tail})} = \mpstatename\),
+then $\mach_r$ must correspond to a return state for $\mach_c$, i.e.,
+\(\ret{\mach_c}\ \mach_r\) holds.\apt{This sentence still doesn't make sense to me.}
 

### 2020-07-10 main.tex
Pass

--- 2020-07-10 main.tex
@@
 The call stack is a perennial target for attacks
-based on memory-safety vulnerabilities, with potential consequences
+based on memory safety vulnerabilities, with potential consequences
 ranging from leakage or corruption of private stack data to
@@
 %
-We formalize this intuition in (1) a {\em stack-confidentiality}
+We formalize this intuition in (1) a {\em stack confidentiality}
 property, which promises that a callee's observable behavior is not affected
 by memory outside its stack frame (e.g., a caller's private data or
-uninitialized memory), plus (2) a {\em stack-integrity} property,
+uninitialized memory), plus (2) a {\em stack integrity} property,
 which promises that, if a function call ever returns, the caller's stack
@@
 conservative Depth Isolation policy, which not only enforces our basic
-stack-safety property, but also an alternative lockstep property that
+stack safety property, but also an alternative lockstep property that
 checks for integrity and confidentiality violations at every step of
@@
 detects violations between when they occur and when they make a visible impact.
-Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing ``Lazy Tagging
-and Clearing'' policy does not quite enforce these properties,
+Unfortunately the existing Lazy Tagging and Clearing policy of
+\citeauthor{DBLP:conf/sp/RoesslerD18} does not quite enforce these properties,
 but would if enhanced (at some performance cost).
@@
   correct grammar, not inconsistency!  :-)}
-\sna{Aha, I forgot that. Going hard the other way.}
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A novel {\em stack-safety property} abstractly capturing the
+  A novel {\em stack safety property} abstractly capturing the
   intuitive concepts of strict local state encapsulation and well-bracketed
@@
 \item
-  A {\em lazy} variant of the stack-safety property that can be enforced if the
+  A {\em lazy} variant of the stack safety property that can be enforced if the
   Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
@@
 program counter $\PCname$
-and stack pointer $\rsp$\bcp{but not a frame pointer?!\apt{right; we're agnostic about that}}; other details of the machine's ISA are unimportant.
+and stack pointer $\rsp$\bcp{but not a frame pointer?!}; other details of the machine's ISA are unimportant.
 
@@
 consists of an instruction region and a stack that grows upwards. (For simplicity,
-we do not consider the heap. Integrating a memory-safety property for the
+we do not consider the heap. Integrating a memory safety property for the
 heap in the style of
@@
 $\rsp$). \bcp{This implicitly amounts to choosing a caller-saves register
-  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}
-
-Our stack-safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code (the same instruction opcode may be
+  convention, though, doesn't it?}
+
+Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
+which are not fully explicit in machine code (the same instruction may be
 used in some places as part of a call or return sequence and in other places
@@
 If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
-  return} has occurred when the $\PCname$ points to the next instruction after $a$ and $\rsp = s$. (Our definition
+  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
 of well-bracketed control flow relates this definition of return to the program's
@@
 
-Enforcement mechanisms are modeled as {\em policies} (borrowing the
-terminology from tag-based mechanisms) that add some auxiliary policy
-state to the ordinary machine state
+Enforcement mechanisms are modeled as {\em policies}\bcp{this is a somewhat
+  special use of the word ``policy'' that comes from the micropolicies
+  world.  We should include a comment about this to avoid confusing
+  readers.} that add some auxiliary
+policy state to the ordinary machine state
 %\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
@@
 \emph{varying} that data would not change the program's observable
-behavior~\citep{6234468}.
+behavior~\citep{6234468}
 In applying this idea to a system that incorporates policy enforcement, only
@@
 attached to each value in memory or registers, and the policy step
-function checks that each machine operation obeys a set of rules on tags
+function\bcp{undefined!}\leo{``partial policy step'' is explained informally above}
+checks that each machine operation obeys a set of rules on tags
 (e.g., that the current $\PCname$ tag matches the tag on the stack
@@
 the stack protection policies of
-\citet{Skorstengaard+19b} and (with an extension to support linear capabilities) \citet{Skorstengaard+19}.
-%\bcp{I thought these policies used
-%  linear capabilities?}\leo{Only the second one AFAIK. What would you
-%  change to characterize it more accurately?}\rb{StkPointers uses linear
-%  capabilities, its predecessor users local capabilities (and unlike
-%  StkPointers could be implemented in CHERI)}.
+\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
+  linear capabilities?}\leo{Only the second one AFAIK. What would you
+  change to characterize it more accurately?}\rb{StkPointers uses linear
+  capabilities, its predecessor users local capabilities (and unlike
+  StkPointers could be implemented in CHERI)}.
 Capabilities package up pointers
@@
 compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
-to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
-be evaluated in terms of our formal notion of stack safety with a trivial null
-policy. But any state maintained by the enforcement mechanism must be in normal
-memory and subject to the same confidentiality and integrity concerns (to be
-varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
-out as policy state provided that it can be proven to protect itself.
-
-\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}
+to insert stack canaries~\citep{Cowan+98}. Such approaches should
+still satisfy our formal notion of stack safety with a trivial null
+policy, but only for programs produced by the compiler.
 %These approaches are represented as policies that
@@
 
-\Cref{fig:running-program} shows a high-level view of a running example,
+\Cref{fig:running-program} shows a high-level view of our running example,
 written in C-like pseudo-code.  It consists of a {\tt main} function that
@@
 execution trace of the corresponding machine instructions. These might have
-been generated from the high-level code by a (rather simple-minded) compiler,
-but it is only the behavior 
+been generated by compiling the high-level code, but it is only the behavior
 of the machine instructions that we care about
@@
 Instructions are intended to
-be similar to RISC-V, but are described in a (hopefully more readable)
+be similar to RISC-V, but are described in a (hopefully more readable) an
 ad-hoc notation
@@
 
-The code itself is straightforward, and should be largely self-explanatory, but
-a few points deserve clarification. Functions return their result value on the
-stack rather than in a register.  Stack frames are built in two parts:
-space for local variables and the return address is allocated and initialized at function entry, and deallocated at function exit; space for arguments and return values is allocated and initialized immediately prior to each function call and deallocated on return.
-The return value slot is always initialized
-because some of our enforcement mechanisms will rely on this.
-\apt{this is bogus, but oh well}
-The return address is spilled to the stack even by leaf functions such as
-{\tt h} that could just leave it in $\rra$.
-
 %\apt{Need to clarify to what extent our properties depend on the
@@
 as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of 
-% static or dynamic,
-(software- or
+machine model plus some kind of (static or dynamic, software- or
 hardware-enforced) policy monitor. In~\cref{sec:enforcement}, we will
@@
 %
-This step function is total, so the machine's execution traces
-(defined in \cref{sec:traces}) are infinite.
+This step function is total: execution traces are infinite unless
+cut short externally by a policy fault (as explained below).
 
@@
 \POLS\), a policy step function \((\mach, \pol) \polstep \pol' :
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial; 
-it is undefined on input configurations that correspond to a policy fault.
-We assume that the
+\MACHS \times \POLS \rightharpoonup \POLS\). We assume that the
 compiler provides an initial policy state $\pol_0$ to accompany
@@
 
-This combined step function is also partial, and traces of the machine-with-policy
-will be finite iff a policy fault occurs.
-
 \subsection{Traces}
-\label{sec:traces}
 
@@
 machine states
-and observations, called \emph{machine traces} and denoted $\machT \in \MOTRACE$,
+and observations, called machine traces and denoted $\machT \in \MOTRACE$,
 and with sequences of triples of a machine state, a policy state, and an
-observation, called \emph{MP-traces} and denoted $\MPT \in \MPOTRACE$.
+observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.
 
@@
 In our running example in \Cref{fig:running-trace-b} we can
-see a depiction of (a prefix of) such a trace: starting from an initial machine
+see a depiction of such a trace: starting from an initial machine
 state at the very top, each instruction leads to a new machine state
@@
 \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
-running \(\mach\) up to \(f\)'') and similarly for traces with policies:
-
-\begin{center}
-\begin{minipage}{.3\textwidth}
+running \(\mach\) up to \(f\)'')
+
 \judgmenttwo{\(\mach \hookrightarrow \machT'\)}
@@
             {\(\mach \hookrightarrow \machT | f\)}
-\end{minipage}
-\begin{minipage}{.6\textwidth}
+\noindent
+and similarly for traces with policies:
+
 \judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
@@
          {\(\mpstatename \hookrightarrow \MPT | f\)}
-\end{minipage}
-\end{center}
 
@@
 Note that an infinite silent trace is a
-prefix of (and similar to) any other trace. While this might seem
+prefix of (and equivalent\bcp{similar?} to) any other trace. While this might seem
 surprising at first, it makes sense in a timing-insensitive context:
@@
 
-Our properties and policies assume that the program is annotated
-with various information about its structure.
+Our properties and policies are parameterized by several types of
+annotations describing features of a program.
 
@@
 correspond to {\tt h}, which means that the function map would be:
-
-\[ \codemap(w) = 
-  \begin{cases}
-    \texttt{main} & \text{if } 1 \leq w \leq 13 \\
-    \texttt{g} & \text{if } 20 \leq w \leq 36 \\
-    \texttt{h} & \text{if } 40 \leq w \leq 47 \\
-    \mathit{undefined} & \text {otherwise}\\
-  \end{cases}
-\]
+\[ \codemap = \{\texttt{main} \mapsto \{1\ldots13\}, \texttt{g} \mapsto \{20\ldots36\}, \texttt{h} \mapsto \{40\ldots47\}\}\]
 
@@
 
-For concreteness, in a typical calling convention, the call map will
-identify the locations of the $\jal$ instructions representing calls.
-% APT: I found this more confusing than helpful...
-%, in which
-%space for arguments is already allocated, but the mechanism is more
-%general.
-A state in which a call map is defined on the program counter
+For concreteness, in a typical calling convention, a call map may be
+considered to identify the state just before the \jal, in which
+space for arguments is already allocated, but the mechanism is more
+general. A state in which a call map is defined on the program counter
 is referred to as a {\it call state}.
@@
 Similarly, a {\em return map} identifies addresses that complete a
-return sequence. Formally, it is a set of words; each one
+return sequence. Formally, it is simply a set of words; each one
 is the address of the final jump corresponding to a function return:
@@
 
-We separately identify, for any call state, what it means to have 
-correctly returned from that call. We consider the call to have
+We separately identify, for any call state, what it means for a state
+to have correctly returned from a call. We consider the call to have
 returned when control returns to the instruction following
-the $\jal$ with the stack restored. Formally, this is a predicate on
+the \jal with the stack restored. Formally, this is a predicate on
 states that says when the machine has just returned from a specified call state
@@
 Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
-the formal stack-safety property described in the previous section: it in
+the formal stack safety property described in the previous section: it in
 fact enforces a stronger, inductive version of that property which strengthens the
@@
 at exactly those points in
-the program where the contour of the stack-safety property changes:
+the program where the contour of the stack safety property changes:
 
@@
 There remains the question of how to ensure that the sequences are
-invoked at the right places. For a program to enjoy the stack-safety property
+invoked at the right places. For a program to enjoy the stack safety property
 (with respect to a particular call map), all we need require is that
@@
 If the code fails to initiate an exit
-sequence at a point where the stack-safety property expects a return,
+sequence at a point where the stack safety property expects a return,
 the micropolicy will incorrectly behave as if execution
@@
 rights than the caller, this is harmless.
-
-\apt{It would be good to relate this back to the running example. This is
-  complicated because the example passes arguments/returns, and it builds
-  and initializes this part of the frame separately.}
 
@@
 a variation of the original machine state at the time of that call on
-top of the variant stack (just like the stack-confidentiality property
+top of the variant stack (just like the stack confidentiality property
 enforced trace confidentiality for every call transition), and every time
@@
 
-  Part of the motivation for the strong lockstep stack-safety property
+  Part of the motivation for the strong lockstep stack safety property
   is that it uncovers errors quickly, and therefore should be quicker

### 2020-07-10 main.tex
done through 7.1

--- 2020-07-10 main.tex
@@
 stack rather than in a register.  Stack frames are built in two parts:
-space for local variables and the return address is allocated and initialized at function entry, and deallocated at function exit; space for arguments and return values is allocated and initialized immediately prior to each function call and deallocated on return.
+space for local variables (which are all stack-allocated) and the return address is allocated and initialized at function entry, and deallocated at function exit; space for arguments and return values is allocated and initialized immediately prior to each function call and deallocated on return.
 The return value slot is always initialized
@@
 At each step of the machine, the relevant rule is applied to the tags on
-the instructions inputs (PC, registers, memory) and on the instruction itself, and produces
+the instruction's inputs (PC, registers, memory) and on the instruction itself, and produces
 one of two outcomes: either the instruction is allowed to execute (and generates
@@
 control that we do not need to consider here), and tags registers containing
-stack pointers with access permission information. The compiler also
+stack pointers with access permission information. The policy also
 uses tags on instructions to identify the code
@@
 for every nested call performed so far on a {\em variant stack}: a
-list of machine states, each annotated with its own contour written
+list of machine states, each annotated with its own contour (written
 as $\vm{C}$), and each being a variation of the actual machine
-according to $C$.
+according to $C$.%
 %
@@
 MP-trace induced from $(\mach, \pol)$ satisfies the \emph{trace
-  lockstep stack safety}, and write $\testProp{}{\MPT}{\vs}$, if the
+  lockstep stack safety}, written $\testProp{}{\MPT}{\vs}$, if the
 following conditions are satisfied:
@@
   and trace lockstep stack safety must be coinductively satisfied for
-  the remainder of the trace and the variant stack $vs'$ that
+  the remainder of the trace and the variant stack $vs'$ that $vs$
   steps to ($\vs \stepsto \vs'$):
@@
 %
-  then the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
+  then the variant stack test must be
   satisfied:
@@
   %
-  the top of the variant stack also returns,
-  the tail of $\vs$ satisfies the stack safety test
+  then the top of the variant stack also returns,
+  the tail of $\vs$ satisfies the stack safety test:
   \[\testVS{(\mach, \pol)}{\mathit{tail}(\vs)},\]
@@
 %
-A system satisfies {\em lockstep stack safety} if for every initial machine $(\mach,\pol)$,
+A system satisfies {\em lockstep stack safety} if, for every initial machine $(\mach,\pol)$,
 the trace induced from $(\mach, \pol)$ satisfies trace lockstep stack safety with
@@
 The proof is a fairly involved coinductive argument
-(around 3K LoC in Coq) on the definition trace lockstep stack safety,
+(around 3K LoC in Coq) on the definition of trace lockstep stack safety,
 using an auxiliary depth-index to ensure the proper alignment of
@@
 \citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
-mitigate those costs, they propose optimizations like Lazy Tagging and
-Lazy Clearing that relax and defer some of the checks of an eager
-policy, being more permissive in their allowed operations and catching
-violations if and when their effect becomes manifest.
+mitigate those costs, they propose optimizations called Lazy Tagging and
+Lazy Clearing that relax and defer some of the checks of the Depth Isolation
+policy. Under these optimizations, the policy does not initialize stack
+frames on entry or clear them on exit (thus speeding up execution),
+and it permits all writes to the stack, even when the $\PCname$ tag
+does not match the memory tag. 
+However, reads from the stack do still require $\PCname$ tag and memory tag
+to match. Thus, even if a callee illicitly writes to a private location
+in its caller, the caller will eventually detect this if it ever tries to read from
+that location.
 %
@@
 provided that the differences do not change the observable behavior of
-the rest of the program. It is not immediately clear what constitutes
-changing behavior: we construct an idealized return state that
+the rest of the program. The challenge here is to define what constitutes
+a change in behavior. We construct an idealized return state that
 contains only the changes that the caller was allowed to make, and use
@@
         \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
-        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
-        & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
-        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
+        & \head(\MPT) = \mpostate{\mach_e}{\_}{\_} ~ \land ~
+          \last(\MPT) = \mpostate{\mach_r}{\_}{\_} ~ \land \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
+          \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
           \Rightarrow \\
@@
       \integritylazyexample
-      \caption{Rolling back from Figure \ref{fig:intex}\bcp{too much space!}}
+      \caption{Rolling back from Figure \ref{fig:intex}.\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
       \label{fig:intlex}
@@
 
-    However, consider instead if we insert after instruction 44 a new
-    instruction \(45: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
+    However, consider instead if we modify the original code of \cref{fig:running-trace-b} to insert after instruction 44 a new
+    instruction \(44.5: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
     {\tt g}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
@@
         \confidentialitylazyexample
-        \caption{Continuing from \cref{fig:confex}\bcp{Spacing!}}
+        \caption{Continuing from \cref{fig:confex}.}
         \label{fig:conflex}

### 2020-07-10 main.tex
Updated a number of things in text, coq

--- 2020-07-10 main.tex
@@
   correct grammar, not inconsistency!  :-)}
-\sna{Aha, I forgot that. Going hard the other way.}
+\sna{Aha, I forgot that. Corrected, and I've added it in more places.}
 
@@
     \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[103] \leftarrow 0}
-        {Zero out result}
+        {Zero out result\apt{why?}\bcp{+1}}
     \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
-        {Call ($\jal$ to) {\tt g}}
+        {Call ($\jal$ to) {\tt f}}
   \end{array}
@@
     \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp - 1] \leftarrow r_0}{[107] \leftarrow 0}
-        {Zero out result}
+        {Zero out result \apt{why?}}
     \row{\jal ~ \rra ~ 40}{\rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow 40}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 40}
-        {Call {\tt h}}
+        {Call {\tt g}}
   \end{array}
@@
 \memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
 \memory{2}{\gprivc}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
-\memory{2}{orange}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
 \memory{2}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
@@
 \memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
-\memory{2}{orange}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
+\memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
 \memory{1}{\hprivc}%[{\makebox[0pt]{{\texttt h} ra}}]
@@
 \memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{orange}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{{\tt h}'s call frame }}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
+\memory{2}{orange}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
 \memory{2}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
@@
   \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
-      {Deallocate {\tt h}'s call frame}
+      {Deallocate {\tt g}'s call frame}
   \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
@@
 The return address is spilled to the stack even by leaf functions such as
-{\tt h} that could just leave it in $\rra$.
+{\tt g} that could just leave it in $\rra$.
 
@@
 In our running example, instructions 1-13 correspond to {\tt main},
-instructions 20-36 correspond to {\tt g}, and instructions 40-47
-correspond to {\tt h}, which means that the function map would be:
+instructions 20-36 correspond to {\tt f}, and instructions 40-47
+correspond to {\tt g}, which means that the function map would be:
 
@@
 In our running example, the callmap contains two entries: the word 7
-corresponding to the call to {\tt g} which has one argument and one
-return value; and the word 26 corresponding to the call to {\tt h},
+corresponding to the call to {\tt f} which has one argument and one
+return value; and the word 26 corresponding to the call to {\tt g},
 also with a single argument and return value:
@@
 In our running example, the entry map contains the entry
-points of {\tt g} and {\tt h}:
+points of {\tt f} and {\tt g}:
 \[ \entmap = \{20, 40\} \]
@@
 are depicted in the figure: the state before the execution of
-instruction 27 corresponding to the return from {\tt h} and the state
+instruction 27 corresponding to the return from {\tt f} and the state
 before instruction 8 corresponding to the return from {\tt g}.
@@
 Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
-MP-trace. say \(\MPT\) satisfies {\em trace integrity} with
+\(\MPOTRACE\). We say \(\MPT\) satisfies {\em trace integrity} with
 respect to \(\contour\) if its final state (if any) agrees with its
@@
   \propdef{TraceInt} ~ \contour ~ \MPT \triangleq &
-    \head(\MPT) = \mpostate{\mach_e}{\_}{\_} \land
-    \last(\MPT) = \mpostate{\mach_r}{\_}{\_} \Rightarrow \\
+    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
+  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
   & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
@@
 \end{split}\]
-\sna{Experimenting with font on the mathy version of the property.\apt{looks fine to me}}
+\sna{Experimenting with font on the mathy version of the property.}
 
@@
 integrity is trivially satisfied. Similarly, the callee is allowed to
-temporarily overwrite the caller's private locations,
-as long as their values are restored
-by the time the callee returns. We will see variations of this later on: a
+temporarily overwrite these locations, as long as they are restored
+once control returns. We will see variations of this later on: a
 stronger, inductive integrity property that ensures that the caller's
-private data is never overwritten, even during the callee's execution
+private data is never overwritten, even during the caller's execution
 (which leads to easier reasoning and more effective testing);
 and a weaker, lazy integrity property that only ensures the preservation
-of locations that subsequently affect the program's observable actions
+of locations that subsequently affect the caller's observable actions
 (which allows for a more efficient enforcement mechanism).
@@
 Going back to our running example, such a trace would be, for
-instance, the entire trace of {\tt h} from instruction 40 to
+instance, the entire trace of {\tt g} from instruction 40 to
 instruction 47.
@@
 Instead of storing data to the return value slot,
-the changed version overwrites {\tt g}'s local variable {\tt z}.
+the changed version overwrites {\tt f}'s local variable {\tt z}.
 The figure shows the modified trace, with snapshots of the initial
@@
 
-\definition Given a trace \(\MPT\), % and a call map \(\callmap\)
+\definition Given a trace \(\MPT\) and a call map \(\callmap\)
 %and a function map \(\codemap\),
@@
 machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
-\(\mpstate{\mach_e}{\pol_e}\) in \(\MPT\) such that
+\(\mpstate{\mach_e}{\pol_e}\) in the trace such that
 \(\mach_c\) is a valid call with \(n\) arguments, and
@@
 state $(\mach_0,\pol_0)$ and its corresponding induced trace $\MPT$ (with
-$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), and for every
+$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), for every
 $n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
@@
 \[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]
-\apt{This could still be clearer. Interleaving informal and mathy in the
-same sentence doesn't work so well.}
 
@@
 overwritten by a callee, confidentiality ensures that the caller's data never
-influences the callee's observable behavior. Again, we will formally define
-confidentiality in two steps: a trace version and a system
+influences the callee's observable behavior. We will formally define
+confidentiality in two similar steps: a trace version and a system
 version.
@@
 \definition%{Variations}
-Two machine states \(\mach\) and \(\mach'\) are {\em variations}
+We say two machine states \(\mach\) and \(\mach'\) are {\em variations}
 of one another given a contour \(\contour\) if they agree on low
@@
 ``original'' trace $\MPT \in \MPOTRACE$ that corresponds to a callee $f$
-and a ``variant'' trace $\machT \in \MOTRACE$ that corresponds to how $f$
+and a ``variant'' trace $\machT \in \MTRACE$ that corresponds to how $f$
 would behave if it were started from a variant initial machine state.
@@
 %
-This example will end up with a violation of confidentiality, as a
-variation in a confidential location (\textcolor{cyan}{42} vs
-\textcolor{magenta}{-5} in location 100) leads to a visible
+This example will end up a violation of confidentiality, as a
+difference in a confidential location (\textcolor{cyan}{42} vs
+\textcolor{magenta}{5} in location 100) leads to a visible
 difference in return value.
@@
   \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
-    {\_}{\_}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
+    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
     some final state \(\mach_r'\) and any changes from the initial states are
@@
       \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
-      \last(\MPT) =
-        \mpostate{\mach_r}{\_}{\_} ~ \land ~ R ~ \mach_r \Rightarrow \\
-      & \exists \mach_r' . \last(\machT) = (\mach_r',\_)
-        \land \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land \\
-      & \forall ~ \component . (\mach_e(\component) \not = \mach_r(\component)
-        \lor \mach_e'(\component) \not = \mach_r'(\component)) \Rightarrow
-      \mach_r(\component) = \mach_r'(\component)
+      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
+        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
+        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
+      & ~ (\mach_e(\component) \not = \mach_r(\component)
+        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
+      \mach_r(\component) = \mach_r'(\component))
     \end{split}\]
@@
     \[\begin{split}
-      \mathit{InfCase} ~ \MPT ~ \machT \triangleq
+      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
       & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
@@
       \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
-      & \forall \mach_r .
-        \last(\MPT) = \mpostate{\mach_r}{\_}{\_} ~ \land ~ \neg ~ R ~ \mach_r \Rightarrow \\
+      & \forall \mach_r ~ \pol_r ~ \obs_r .
+        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
+      & \neg ~ R ~ \mach_r \Rightarrow \\
       & \pi_o(\MPT) \lesssim \pi_o(\machT)
@@
 
-\[  \propdef{TraceConf} ~ R ~ \MPT ~ \machT \triangleq
-   \mathit{ReturnCase} ~ R ~ \MPT ~ \machT ~ \land 
-   \mathit{InfCase} ~ \MPT ~ \machT \land
-    \mathit{FailCase} ~ R ~ \MPT ~ \machT 
-\]
+\[\begin{split}
+  \propdef{TraceConf} ~ R ~ \MPT ~ \machT \triangleq
+  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
+  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
+    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
+\end{split}\]
 
@@
 Consider the result of modifying line 42 of our running example to $r_4 \leftarrow [\rsp - 9]$.
-Instead of reading the proper argument {\tt v}, {\tt h} now reads directly from the
+Instead of reading the proper argument {\tt v}, {\tt g} now reads directly from the
 secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
-shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
+shows \(\mach_e\), the entry state of {\tt g}, with high confidentiality memory
 in {\high} and low confidentiality in {\low}. Below it, the variant state
@@
 \(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
-\machT_{\mathit{call}}\), and it also holds for the toplevel execution
-trace to account for unallocated memory, \(\propdef{TraceConf} ~
+\machT_{\mathit{call}}\) and it also holds for the toplevel execution
+trace to account for unallocated memory \(\propdef{TraceConf} ~
 \emptyset ~ \MPT ~ \machT\) , then the system as a whole enjoys {\em
@@
 and may still hold on a wide range of programs that violate expected
-control flow. For instance, if {\tt g} managed to jump in the middle of
-the code of {\tt h} with its call, confidentiality and integrity would
+control flow. For instance, if {\tt f} managed to jump in the middle of
+the code of {\tt g} with its call, confidentiality and integrity would
 still be satisfied.
@@
 $\PCname \leftarrow r_4$, with effect $\PCname \leftarrow 42$, causing {\tt main}
-to jump into the middle of {\tt h}.
+to jump into the middle of {\tt g}.
 This violates control separation because the instruction that causes a
@@
 because the target (instruction 42) of the new $\jalr$ is not an established
-entry point according to the entry map. Our properties do not enforce which
-call points jump to which entry points, however; that would require a stronger
+entry point according to the entry map. Our properties to not enforce which
+call points jump to which entry points, however; that is left for a stronger
 notion of control flow integrity.
@@
 
-\begin{center}
 \begin{minipage}{.3\textwidth}
@@
 \end{minipage}
-\end{center}
 
@@
 and return to the appropriate location. For any initial state \((\mach_0\, \pol_0)\), let
-\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\). Then two
-conditions must hold. 
-\emph{(i)} \(\MPT\) must not have an unmatched return, i.e., \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
-must not hold for any \(\MPT_{tail}\). \emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
+\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\).
+\(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
+may not hold for any \(\MPT_{tail}\). For any state pair \((\mach_c, \pol_c)\)
 corresponding to a call \(\MPT\), if
-\(\mpstatename = \mpostate{\mach_r}{\_}{\_}\)
+\(\mpstatename = \mpostate{\mach_r}{\pol_r}{\obs_r}\)
 is the first unmatched return such that
-\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{(\MPT_{tail})} = \mpstatename\),
-then $\mach_r$ must correspond to a return state for $\mach_c$, i.e.,
-\(\ret{\mach_c}\ \mach_r\) holds.\apt{This sentence still doesn't make sense to me.}
+\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{\MPT_{tail}} = \mpstatename\),
+then $\mach_r$ must correspond to a return state for $\mach_c$:
+\(\ret{\mach_c}\ \mach_r\)
 
@@
 an instruction $ \rra \rightarrow \rra - 19$
-between 45 and 46, this would cause the return from {\tt h} to go to instruction 8 and
-return to {\tt main} instead of {\tt g}. While the instruction that
+between 45 and 46, this would cause the return from {\tt g} to go to instruction 8 and
+return to {\tt main} instead of {\tt f}. While the instruction that
 causes the return would still be a valid return (satisfying control
 separation), the state after the return would not correspond to an
-unmatched return for {\tt g}'s call to {\tt h}, violating return
+unmatched return for {\tt f}'s call to {\tt g}, violating return
 integrity.
@@
     For our example, we return to the violation of trace integrity in
-    \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
+    \cref{fig:intex}, in which {\tt g} has written to {\tt z} and returned.
     Whether observable integrity holds depends on the continuing trace, shown
@@
       and $-5$ in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
-      never used otherwise, the traces of {\tt h} will enjoy observable trace
+      never used otherwise, the traces of {\tt g} will enjoy observable trace
       confidentiality.

### 2020-07-10 main.tex

--- 2020-07-10 main.tex
@@
     \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
-        {Call ($\jal$ to) {\tt g}}
+        {Call ($\jal$ to) {\tt f}}
   \end{array}
@@
     \row{\jal ~ \rra ~ 40}{\rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow 40}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 40}
-        {Call {\tt h}}
+        {Call {\tt g}}
   \end{array}

### 2020-07-10 main.tex
through Fig. 6

--- 2020-07-10 main.tex
@@
 \definition
+\label{def:trace-conf}
 
@@
         & \head(\MPT) = \mpostate{\mach_e}{\_}{\_} ~ \land ~
-          \last(\MPT) = \mpostate{\mach_r}{\_}{\_} ~ \land \\
+          \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land \\
         & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
@@
 
-      Two of the cases of trace confidentiality, \(\mathit{InfCase}\)
-      and \(\mathit{FailCase}\), are purely properties of the
-      observations produced by the traces they are properties of. If
+      Two of the cases of trace confidentiality (\cref{def:trace-conf}), \(\mathit{InfCase}\)
+      and \(\mathit{FailCase}\), are defined solely in terms of
+      observation traces. If 
       the traces are infinite, they should be similar; if one
@@
       which $\MPT$ has a last element
-      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and that machine state is
-      a return: \(R ~ \mach_r\).
+      \(\mpostate{\mach_r}{\_}{\_}\) and that machine state is
+      a return, i.e. \(R ~ \mach_r\) holds.
 %
       Observable trace confidentiality then requires that $\machT$
-      also has a final state \((\mach_r',\obs_r')\) that is a return:
-      \(R ~ \mach_r'\).
+      also has a final state \((\mach_r',\_)\) that is a return, i.e.
+      \(R ~ \mach_r'\) holds.
 %
@@
         \mathit{ObsReturnCase} ~ R ~
-        (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~
-        ((\mach_e',\obs_e')\machT) \ \ \triangleq \ \
-        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
-        & \last(\machT) = (\mach_r',\obs_r') \Rightarrow \\
-        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \Rightarrow\\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
-        & \mach_r'' \hookrightarrow \machT' \Rightarrow \\
-        & \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
-        & \pi_o(\MPT') \lesssim \pi_o(\machT') \\
+        (\mpostate{\mach_e}{\_}{\_}\MPT) ~
+        ((\mach_e',\_)\machT) \ \ \triangleq \ \
+        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land ~
+         \last(\machT) = (\mach_r',\_) ~ \land \\
+        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') ~ \land \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
+        \mach_r'' \hookrightarrow \machT' \Rightarrow \\
+        & \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land ~
+        \pi_o(\MPT') \lesssim \pi_o(\machT') \\
       \end{split}\]
@@
       confidentiality:
-
-      \[\begin{split}
+\[
         \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
-        & \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \land \\
-        & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
-          \MPT ~ \machT \\
-      \end{split}\]
+         \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT ~ \land 
+         \mathit{InfCase} ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
+          \MPT ~ \machT 
+\]
 
@@
       \]
-      Addresses 108 and 109
+      Addresses 108\apt{107??} and 109\apt{108??}
       are not restored; every address that was varied and not changed in one

### 2020-07-10 main.tex
done with pass

@@
       {\em Observable stack confidentiality} is structured like stack
-      confidentiality. Let \(\callmap\) be a callmap; for each initial
+      confidentiality. Let \(\callmap\) be a callmap\apt{should be implicit and is not used anyway}; for each initial
       state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
@@
       \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
-      \ret{\mach_c} ~ \MPT ~ \machT\).
+      \ret{\mach_c} ~ \MPT ~ \machT\).\apt{What is $M$ ??}.
 
@@
 
-    Observable properties allow us defer enforcement until a property
+    Observable properties allow us to defer enforcement until a property
     violation would become visible, as in lazy policies. But by formalizing
     the connection to the eager policy, we can identify when lazy policies
-    miss violations that can become visible. The basic lazy tag policy from
-    \citet{DBLP:conf/sp/RoesslerD18} tags written memory with the depth in the
-    stack at which it was written, and enforces that it must be read from the
+    miss violations that can become visible. The optimized lazy tag policy from
+    \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth 
+    at which it was written, and enforces that it must be read from the
     same depth. However, this allows violations of observable stack safety!
@@
     whether it had been called previously and changed its behavior accordingly.
+    \apt{More blatently, static authorities doesn't work for recursive calls.}
 
@@
 protected components avoid compromising their own security by following certain
-simple rules---like avoiding to leak their private capability seals. More
+simple rules---like not leaking their private capability seals. More
 precisely but still informally, local state encapsulation is defined as
@@
 Heap safety, like stack safety, may be framed as a security property
-in the form of noninterference~\citet{DBLP:conf/post/AmorimHP18}. Just
+in the form of noninterference~\citep{DBLP:conf/post/AmorimHP18}. Just
 like~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
@@
     Instead we need a variant that treats the entry of a tail call as the
-    return of its caller for purposes.
+    return of its caller for accounting\apt{better word?} purposes.
 

### 2020-07-10 main.tex
Changed dangling colons, some mathit

@@
 \[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
-\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
+\[\mach \stepstoobs{\obs} \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
 %
@@
 enforcement mechanism, consisting of a set of policy states \(\pol \in
-\POLS\), a policy step function \((\mach, \pol) \polstep \pol' :
+\POLS\), a policy step function \((\mach, \pol) \polstep \pol' \in
 \MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial; 
@@
 %the trailing observation. Here's an attempt.}
-takes a trace\bcp{I changed a : here to $\in$.  We should check if there are
-any more to fix.}
+takes a trace
+%% \bcp{I changed a : here to $\in$.  We should check if there are
+%% any more to fix.}
 \(\Trace \in \TRACE{\TRACEELEMS} \times \OBSS\) and a predicate
@@
 variation in a confidential location (\textcolor{cyan}{42} vs
-\textcolor{magenta}{-5} in location 100) leads to a visible
+\textcolor{magenta}{$-5$} in location 100) leads to a visible
 difference in return value.
@@
 conditions must hold. 
-\emph{(i)} \(\MPT\) must not have an unmatched return, i.e., \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
-must not hold for any \(\MPT_{tail}\). \emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
+\emph{(i)} \(\MPT\) must not have an unmatched return, i.e., \(\Unmatched ~ \MPT ~ \MPT_{\mathit{tail}}\)
+must not hold for any \(\MPT_{\mathit{tail}}\). \emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
 corresponding to a call \(\MPT\), if
@@
 is the first unmatched return such that
-\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{(\MPT_{tail})} = \mpstatename\),
+\( \Unmatched ~ \MPT ~ \MPT_{\mathit{tail}}\) and \(\head{(\MPT_{\mathit{tail}})} = \mpstatename\),
 then $\mach_r$ must correspond to a return state for $\mach_c$, i.e.,
@@
       Given a contour, a call state, and a returned state, we define a
-      {\em rollback} function \(\mathit{roll} : \CONTOURS \times \MACHS \times
+      {\em rollback} function \(\mathit{roll} \in \CONTOURS \times \MACHS \times
       \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
@@
       \integritylazyexample
-      \caption{Rolling back from Figure \ref{fig:intex}.\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
+      \caption{Rolling back from Figure \ref{fig:intex}.}
+        %\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
       \label{fig:intlex}
@@
       \]
-      Addresses 108\apt{107??} and 109\apt{108??}
+      Addresses 107 and 108
       are not restored; every address that was varied and not changed in one

### 2020-07-10 main.tex
Addressing Andrew's comment on 6

--- 2020-07-10 main.tex
@@
 \begin{verbatim}
-g(z) = {
-  int w = h(17);
+f(z) = {
+  int w = g(17);
   print z;
@@
 \begin{verbatim}
-h(v) = {
+g(v) = {
   return (v + 1);
@@
     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 104}
-        {Allocate {\tt g}'s call frame}
+        {Allocate {\tt f}'s call frame}
     \row{\sw ~ \rsp ~ r_0 ~ \negate 2}{[\rsp-2] \leftarrow r_0}{[102] \leftarrow 0}
@@
 \memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
 \memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
@@
     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 108}
-        {Allocate space for call frame}
+        {Allocate {\tt g}'s call frame}
     \row{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow r_0 + 17}{r_4 \leftarrow 17}
@@
 \memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
 \memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
@@
   \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
-      {Restore the stack pointer of {\tt g}}
+      {Deallocate {\tt g}'s local state\rb{Yes, still {\tt g}}}
   \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 27}
-      {Return to {\tt g} ($\jalr$)}
+      {Return to {\tt f} ($\jalr$)}
   \end{array}
@@
   \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
-      {Deallocate {\tt g}'s local state}
+      {Deallocate {\tt f}'s local state}
   \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 8}{Return to {\tt main} ($\jalr$)}
@@
   \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
-      {Deallocate {\tt g}'s call frame}
+      {Deallocate {\tt f}'s call frame}
   \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
@@
 call until its corresponding return $\mpstate{\mach_e}{\pol_e}
-\hookrightarrow \MPT_{call} ~ | ~ \ret{\mach_c}$, then $\MPT_{call}$ must
+\hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}$, then $\MPT_{\mathit{call}}$ must
 satisfy trace integrity:
@@
   and initializes this part of the frame separately.}
+\rb{Added a few words at the end of this sub-section.}
 
@@
 %
+The running example in \cref{fig:running-trace-a} manages the stack in this way.
+For example, {\tt f}'s ``blessed'' sequences allocate space to save its return
+address (and private variables) in instructions 20--21, and later restores that
+return address and deallocates its private frame space before returning in
+34--36. Additionally, unlike in the simplified presentation of the policy, the
+caller passes information to the callee through the stack, and for this creates
+and initializes (23 and 24--25) and destructs (29) the argument part of the
+callee's stack frame.\rb{The policy would also need to control this, I think it
+  is clear}
+%
 %\rb{Note that we have also considered and implemented more flexible schemes

### 2020-07-10 main.tex
Proposal: label gpriv in Fig.2
Nits

@@
 \memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{1}{\hprivc}%[{\makebox[0pt]{{\texttt h} ra}}]
+\memory{1}{\hprivc}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
 \memory{1}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
@@
 respect to \(\contour\) if its final state (if any) agrees with its
-initial one on all components marked \(\HIGHSEC\) in \(\contour\):
+initial one on all components marked \(\HIGHSEC\) in \(\contour\):\bcp{I've
+  fixed some problems with not enough space around $\triangleq$, but I see
+  another here --- should check carefully for more.}
 %
@@
   \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)\MPT,\]
-  then the variant stack test must be satisfied:
+  then the variant stack test must be satisfied
   \[\testVS{(\mach, \pol)}{\vs},\]
@@
   then the variant stack test must be
-  satisfied:
+  satisfied
   \[\testVS{(\mach, \pol)}{\vs},\]
@@
   then the top of the variant stack also returns,
-  the tail of $\vs$ satisfies the stack safety test:
+  the tail of $\vs$ satisfies the stack safety test
   \[\testVS{(\mach, \pol)}{\mathit{tail}(\vs)},\]

### 2020-07-10 main.tex
Quick fix on def. 7.6

@@
       {\em Observable stack confidentiality} is structured like stack
-      confidentiality. Let \(\callmap\) be a callmap\apt{should be implicit and is not used anyway}; for each initial
+      confidentiality. For each initial
       state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
@@
       \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
-      \ret{\mach_c} ~ \MPT ~ \machT\).\apt{What is $M$ ??}.
+      \ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}\).\apt{What is $M$ ??}\rb{I think both of these are the call traces, cf. ObservableStackConfPaper in Coq }.
 

### 2020-07-10 main.tex
One error in example-policy connection, one typo

@@
 caller passes information to the callee through the stack, and for this creates
-and initializes (23 and 24--25) and destructs (29) the argument part of the
+and initializes (22 and 24--25) and destructs (29) the argument part of the
 callee's stack frame.\rb{The policy would also need to control this, I think it
@@
 
-      Additional, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
+      Additionally, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
       \mach_0\), taking the induced trace  \(\mach_0' \hookrightarrow

### 2020-07-10 main.tex
Cleanup with Leo

@@
 program counter $\PCname$
-and stack pointer $\rsp$\bcp{but not a frame pointer?!\apt{right; we're agnostic about that}}; other details of the machine's ISA are unimportant.
+and stack pointer $\rsp$; other details of the machine's ISA are unimportant.
 
@@
 definitions say nothing at all about registers (except $\PCname$ and
-$\rsp$). \bcp{This implicitly amounts to choosing a caller-saves register
-  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}
+$\rsp$). \ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
+  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
 
@@
 
-\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}
+\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
 %These approaches are represented as policies that
@@
   int x = 42, y = 0;
-  y = g(0);
+  y = f(0);
   print (x + y);
@@
 \memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
-\memory{2}{orange}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
+\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
+\memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
 \memory{2}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
@@
 because some of our enforcement mechanisms will rely on this.
-\apt{this is bogus, but oh well}
+\ifaftersubmission\apt{this is bogus, but oh well}\fi
 The return address is spilled to the stack even by leaf functions such as
@@
 pair of a machine state and an {\em observation}, which may be either
-silent (written $\tau$) or a word of data\bcp{This is a rather concrete
-  choice!  Why not leave the set of observations abstract?} that the machine
+silent (written $\tau$) or a word of data\ifaftersubmission\bcp{This is a rather concrete
+  choice!  Why not leave the set of observations abstract?}\fi that the machine
 communicates to the outside world:
@@
 \cup \{\PCname,\SP\}$. Observations are writes to the
-special $\rout$ location\bcp{Which is 100! (do we need a name for it? seems
-  not...)}, which can be thought of as a memory-mapped
+special $\rout$ location\ifaftersubmission\bcp{Which is 100! (do we need a name for it? seems
+  not...)}\fi, which can be thought of as a memory-mapped
   output port: if during an execution step a word $w$ is
@@
 \end{split}\]
-\sna{Experimenting with font on the mathy version of the property.\apt{looks fine to me}}
 
@@
 satisfy trace integrity:
-\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]
+\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{\mathit{call}}\]
 \apt{This could still be clearer. Interleaving informal and mathy in the
@@
 
-\apt{It would be good to relate this back to the running example. This is
-  complicated because the example passes arguments/returns, and it builds
-  and initializes this part of the frame separately.}
-\rb{Added a few words at the end of this sub-section.}
-
 %% APT: removed because the tag part is too mysterious and the instruction part
@@
       \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
-      \ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}\).\apt{What is $M$ ??}\rb{I think both of these are the call traces, cf. ObservableStackConfPaper in Coq }.
+      \ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}\).
 
@@
     whether it had been called previously and changed its behavior accordingly.
-    \apt{More blatently, static authorities doesn't work for recursive calls.}
-
-    \leo{Should we clearpage this once we're sure of the layout?}
+    \ifaftersubmission\apt{More blatently, static authorities doesn't work for recursive calls.}\fi
+
 \section{Related Work}
@@
     Instead we need a variant that treats the entry of a tail call as the
-    return of its caller for accounting\apt{better word?} purposes.
+    return of its caller.
 

### 2020-07-10 main.tex
Clearer definition of return integrity

@@
 A system enjoys {\em return integrity} if all of its returns match prior calls
-and return to the appropriate location. For any initial state \((\mach_0\, \pol_0)\), let
-\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\). Then two
+and return to the appropriate location.
+For any initial state and its induced trace \((\mach_0\, \pol_0) \hookrightarrow \MPT\), two
 conditions must hold.
 \emph{(i)} \(\MPT\) must not have an unmatched return, i.e., \(\Unmatched ~ \MPT ~ \MPT_{\mathit{tail}}\)
-must not hold for any \(\MPT_{\mathit{tail}}\). \emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
-corresponding to a call \(\MPT\), if
+must not hold for any \(\MPT_{\mathit{tail}}\).
+%
+\emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
+corresponding to a call in \(\MPT\), if
+\(\MPT_{\mathit{call}}\) is the suffix of the trace starting at the call,
 \(\mpstatename = \mpostate{\mach_r}{\_}{\_}\)
 is the first unmatched return such that
-\( \Unmatched ~ \MPT ~ \MPT_{\mathit{tail}}\) and \(\head{(\MPT_{\mathit{tail}})} = \mpstatename\),
+\( \Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{tail}}\) and \(\head{(\MPT_{\mathit{tail}})} = \mpstatename\),
 then $\mach_r$ must correspond to a return state for $\mach_c$, i.e.,
-\(\ret{\mach_c}\ \mach_r\) holds.\apt{This sentence still doesn't make sense to me.}
+\(\ret{\mach_c}\ \mach_r\) holds.
 

### 2020-07-10 main.tex
Checking function names

@@
 %with the
-%effect of storing the contents of $r_4$ as the result of {\tt h}. If
+%effect of storing the contents of $r_4$ as the result of {\tt g}. If
 %instead it wrote to location $\rsp - 5$ then it would be overwriting
-%private data (the return address) belonging to {\tt g}.
+%private data (the return address) belonging to {\tt f}.
 %
@@
 %with the
-%effect of loading the argument of {\tt g} to $r_5$. If instead it
+%effect of loading the argument of {\tt f} to $r_5$. If instead it
 %loaded the contents of location $\rsp - 6$, it would obtain access
@@
 %If a {\jalr} instruction was added to {\tt main} with its target being
-%the middle of the code block of {\tt h}.\apt{describe result of this?}
+%the middle of the code block of {\tt g}.\apt{describe result of this?}
 
@@
     \texttt{main} & \text{if } 1 \leq w \leq 13 \\
-    \texttt{g} & \text{if } 20 \leq w \leq 36 \\
-    \texttt{h} & \text{if } 40 \leq w \leq 47 \\
+    \texttt{f} & \text{if } 20 \leq w \leq 36 \\
+    \texttt{g} & \text{if } 40 \leq w \leq 47 \\
     \mathit{undefined} & \text {otherwise}\\
@@
 Going back to our running example, here is the state of the machine at
-the call from {\tt main} to {\tt g}, showing addresses annotated with their
+the call from {\tt main} to {\tt f}, showing addresses annotated with their
 corresponding labels:
@@
 As another example, changing instruction 30 of the running example to
-$r_5 \leftarrow [\rsp - 6]$ would cause {\tt g} to read the value of
+$r_5 \leftarrow [\rsp - 6]$ would cause {\tt f} to read the value of
 {\tt main}'s local variable {\tt x} and then immediately output it,
@@
     For our example, we return to the violation of trace integrity in
-    \cref{fig:intex}, in which {\tt g} has written to {\tt z} and returned.
+    \cref{fig:intex}, in which {\tt g} \rb{??} has written to {\tt z} and returned.
     Whether observable integrity holds depends on the continuing trace, shown
@@
     instruction \(44.5: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
-    {\tt g}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
+    {\tt f}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
     rolled-back state \(\mach_r'\) maps it to 0. But in the trace from either

### 2020-07-10 main.tex
Nits
Last few instances

@@
 corresponding to a call in \(\MPT\), if
-\(\MPT_{\mathit{call}}\) is the suffix of the trace starting at the call,
+\(\MPT_{\mathit{call}}\) is the suffix of the trace starting at the call, and
 \(\mpstatename = \mpostate{\mach_r}{\_}{\_}\)
@@
       the example in \cref{fig:confex}, the varied final state
-      \(\mach_r'\) still has $-5$ as a return address for {\tt g}, so
+      \(\mach_r'\) still has $-5$ as a return address for {\tt f}, so
       the trace after that return will behave oddly in a way that
@@
       of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
-      and after returning from {\tt g}, {\tt main} prints 85. But in the trace
+      and after returning from {\tt f}, {\tt main} prints 85. But in the trace
       from \(\mach_r''\), 108 contains $-4$, and ultimately 38 is printed.
@@
       and $-5$ in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
-      never used otherwise, the traces of {\tt g} will enjoy observable trace
+      never used otherwise, the traces of {\tt f} will enjoy observable trace
       confidentiality.

### 2020-07-10 main.tex
A few last comments and polishings (not critical for submission)

@@
 %
-To our knowledge, the first and only line of work that has attempted
-to furnish a positive and formal characterization of the meaning of
+To our knowledge, the only line of work to date that has attempted
+a positive and formal characterization of
 stack safety is \citet{Skorstengaard+19b}, who introduce a calling
@@
 programs. Although the required hardware support is readily available
-in capability machines like CHERI, this technique incurs significant
-costs, because it requires the entire unused part of the stack to be
+in capability machines like CHERI\bcp{including linear capabilities?? or
+  don't they use them in this paper?}, this technique incurs significant
+costs\bcp{This is the pot calling the kettle black}, because it requires the entire unused part of the stack to be
 cleared whenever a security boundary is crossed. Their logical
 relation captures capability safety without ``externally observable
-side-effects (like console output or memory access traces).'' In the
+side-effects (like console output or memory access traces).'' \bcp{Not very clear:}In the
 discussion, ``while [the authors] claim that [their] calling
@@
 
-StkTokens \citep{Skorstengaard+19} is a natural continuation of that work. Like
-its predecessor, it aims to protect the stack by enforcing local state
+StkTokens \citep{Skorstengaard+19} continues this line of work. Like
+the earlier paper, it aims to protect the stack by enforcing local state
 encapsulation and well-bracketed control flow. It does so by defining a new
@@
 work on micropolicies \citep{yannis-report} has shown how to use that framework
-to implement linear return capabilities.
+to implement linear return capabilities.\bcp{But IIUC the Dover realization
+  of micropolicies can't do this...?}
 
@@
 characterization of the meaning of (heap) memory safety, in this paper
-we aim to do the same for the stack.
+we aim to do the same for the stack.\bcp{Let's describe their work in a bit
+  more detail.}
 %

### 2020-09-08 rebuttal.txt
Beginning of rebuttal

Motivation - Applicability:

Our original goal in this work wasn't to formalize a stack safety
notion that is specific to micropolicies, but in hindsight we agree
that the result appears to be tailored to that particular enforcement
mechanism---in large part because of our presentation choices and
exposition-helping simplifications. In particular, we believe that the
enforcement model can be used (with some effort) to describe
capability-based enforcement mechanisms (as we state in Section 2).
However, given that this effort didn't make it into the final product,
we agree that the material present would be more accurately described
as an abstract property capturing micropolicy enforcement rather than
the all-encompassing stack safety notion we set out to state.
LEO: ...and we plan to do so for the full version?

Attacker Model:

The attacker model for the properties is discussed in the paper
(Section 2). For example, lines 104-106 state that "the attacker can
run arbitrary machine code" and that "hardware and timing attacks are
out of scope". The attacker model of the micropolicy enforcement
mechanism is described in [Roessler and Dehon 18].

5 Criteria - StkTokens:
LEO: Not really sure what to say here. 
(1) intuitive: this is subjective but we were aiming for a more extensional property than full abstraction from a "clearly stack safe" machine
(2) useful for reasoning: LEO: do we have anything to say here other than you're right?
(3) reusable in compiler chains: not really applicable?
(4) complete: ??
(5) scalable: what can we extend this to? the maps appear particularly restrictive here...

> Review #27A
> ===========================================================================
>  
> Overall merit
> -------------
> B. Weak Accept
>  
> Reviewer expertise
> ------------------
> Y. Knowledgeable
>  
> Paper summary
> -------------
> This paper defines an abstraction of stack safety in a positive way, meaning defining what operations are allowed. This work improves upon a previous work, which also defines stack safety in a positive way, but whose definitions are tied to capability machines. this work generalizes the stack safety definitions to be applicable to a wide class of  enforcement mechanisms.
>  
> Strengths
> ---------
> + The authors addresses an important problem in software security with a solid methodology. Indeed, stack safety defined in a negative way leads to arm races between defenses and attacks.
>  
>  + The definitions and theory in this work seem sound
>  
>  + The authors apply the proposed the stack safety definitions to characterize two variations of enforcement policies that uses tag-based hardware. The characterization enables us to precisely understand what is the safety achieved by the enforcement policy and what safety is lost to achieve a faster implementation.
>  
> Weaknesses
> ----------
> - The authors do not explicitly discuss how multiple concurrent threads fit in the definition. It is not clear to me how to apply the definitions in this work to multiple threads. Since the stacks of all threads are allocated in the same linear virtual address space, it is possible for one thread’s stack to overlap the stack of another thread. How should stack safety be defined in this case? I think this is an very important question as recent work points out cross-thread stack smashing attack [1].
>  
>  - In the future work section, the authors mentions several types of non-stack control flow that are not addressed by the current model, including tail calls. Other common non-stack control flow includes exceptions (such as c++ exceptions) and setjmp/longjmp. I understand that these non-stack control flow is more difficult to model, but for the exact reasons, how to formally characterize stack safety for these difficult cases becomes critical.
>  
>  [1] Xiaoyang Xu, Masoud Ghaffarinia, Wenhao Wang, Kevin W. Hamlen, and Zhiqiang Lin. 2019. CONFIRM: evaluating compatibility and relevance of control-flow integrity protections for modern software. In Proceedings of the 28th USENIX Conference on Security Symposium (SEC’19). USENIX Association, USA, 18051821.
>  
> Comments for author
> -------------------
> Please see the weaknesses section.
>  
>  
>  
> Review #27B
> ===========================================================================
>  
> Overall merit
> -------------
> D. Strong Reject
>  
> Reviewer expertise
> ------------------
> Y. Knowledgeable
>  
> Paper summary
> -------------
> This paper:
> - Defines stack safety as confidentiality, integrity, and well-bracketedness.
> - Evaluates existing stack protection mechanisms w.r.t. the defined notion of stack safety.
> - Proposes other protection mechanisms.
>  
> Strengths
> ---------
> - This paper is clearly defining a notion of stack safety.
>  
> Weaknesses
> ----------
> - The proposed notion of stack safety is too strong to be used for real-world programs.
> - The proposed notion of stack safety is too strong to be enforced by existing stack protection mechanisms.
> - This paper is lacking proper discussion on related work.
>  
> Comments for author
> -------------------
> This paper is not clearly motivated, and as a result, has the following problems:
>  
> - The proposed notion of stack safety is too strong to be used for real-world programs.
>  
>   As discussed in Section 9, stack-derived pointers are not supported (L1199), but they are widely used in real-world programs.
>   Furthermore, without stack-derived pointers, stack safety is more-or-less obvious to define.
>  
> - The proposed notion of stack safety is too strong to be enforced by existing stack protection mechanisms.
>  
>   [Roessler and DeHon 2018]'s "lazy tagging and clearing" is itself not stack-safe according to this paper.
>   To understand the mechanism, this paper additionally defines the "observable" versions of stack safety.
>   But then why you didn't use the observable version as the authoritative definition?
>
> - This paper is lacking proper discussion on related work.
>  
>    On the contrary to this paper's observation (L25), there're a few formalizations of stack safety (and rely/guarantee conditions for function invocation).
>       For example, see CompCert's assumption on function calls: https://github.com/AbsInt/CompCert/blob/94558ecb3e48261f12c644045d40c7d0784415e0/common/Events.v#L619
>  
> Questions for the response period
> ---------------------------------
> Please address the three weaknesses I discussed above.
>  
>  
>  
> Review #27C
> ===========================================================================
>  
> Overall merit
> -------------
> C. Weak Reject
>  
> Reviewer expertise
> ------------------
> X. Expert
>  
> Paper summary
> -------------
> The authors propose a novel characterization of stack safety in terms of a confidentiality policy, an integrity policy and a well-bracketedness condition.
> The proposal assumes a static code layout.
> In terms of this, it is defined which stack frame the currently running code and the different memory locations belong to.
> In the next step, the policies define which memory locations may be modified and which memory locations may influence writes to memory and output events, and which code may be returned to.
> In addition to this base policy, the authors describe a strong and lazy variant which can be quite directly implemented on a system with micropolicy-based security primitives.
>  
> Strengths
> ---------
> * Novel proposal for characterizing stack safety.
>  
> * The characterization is quite precise about which memory locations may be modified at which time.
>   As such, it can be used directly to inform and evaluate micropolicy-based control flow enforcement.
>  
> Weaknesses
> ----------
> * very poor evaluation of the proposed characterization of stack safety
>   no clear statement of the advantages of the proposed technique.
>   no evaluation in terms of previously proposed informal criteria
>  
> * no discussion of how the proposed technique can be used for reasoning about programs.
>  
> * the proposed policies appear specific to certain enforcement techniques.
>   This is true particularly for the proposed strong and lazy variants of stack safety (which are designed to express exactly what is enforced by a specific enforcement technique).
>   However, even the normal variant assumes, for example, call, return, entry maps which are, of course, very convenient for defining micropolicies, but fundamentally in conflict with certain features like dynamic code generation, dynamic linking, JIT-compilation, variable-arity functions etc.
>  
> * proposed characterization seems fundamentally in conflict with features like dynamic code generation, dynamic linking, varargs, closures that have access to their parent stack frame etc.
>  
> Comments for author
> -------------------
> Summary
>  
> This paper does an interesting proposal for characterizing stack safety rather operationally in terms of memory locations that are allowed to be modified or depended on and code locations which may be returned to at each step of execution.
> However, the proposal is very poorly evaluated against existing proposals, with only some vague and misleading comments.
> This is unfortunate especially because Skorstengaard et al 2019, explicitizes an informal list of desirable qualities from a characterization of stack safety which the authors could easily use to compare the approaches.
>  
> The two variant properties appears specifically designed for capturing the guarantees of very specific micropolicy-based enforcement policies.
> However, also the more general stack safety characterization makes some assumptions that are undesirable in a general characterization of stack safety, like a static code layout and the complete unmodifiability of inactive stack frames (i.e. no closures with access to their parents' stack frames) and these are very fundamental: the proposed characterization strongly depends on these assumptions.
>  
> It is my impression that the authors are not really looking for a characterization of stack safety that is general and widely applicable but rather for one that is concrete and operational enough to inform and support the design of micropolicy-based stack safety enforcement.
> This goal is in my opinion valid and useful, and I believe it would strengthen the paper if the authors were more upfront about this goal.
>  
> In summary, I think this is a promising paper, but it requires a significant rework, particularly a motivation that is more upfront about the goals of the work and a better evaluation of the policy with respect to alternative proposals.
>  
> Discussion
>  
> The weakest point of this paper is the evaluation of the proposed characterization.
> Particularly, the authors don't actually compare their approach to existing approaches, specifically Skorstengaard 2019b.
> In the paragraph on l22-35, the authors suggest that Skorstengaard 2019b's characterization of stack safety is specific to capability machines and they are after "a more abstract characterization [...] that is independent of [a] specific enforcement technique".
> I found no other place in the paper where the authors discuss the advantage of their approach with respect to other approaches, specifically Skorstengaard 2019b.
> There are a number of problems with this:
>  
> - The claim about Skorstengaard et al. 2019b is left completely unsubstantiated and it is in my opinion wrong.
>   It is true that Skorstengaard et al. only instantiate their approach for a capability machine.
>   However, similar to this paper, Skorstengaard et al. describe their characterization from a more general perspective.
>   Their approach is to characterize stack safety as full abstraction of the embedding function from a language with a natively stack safe semantics to the same language with a regular, non-natively stack safe semantics.
>   I don't see why this approach could not be applied in the setting of this paper as well, or in other settings.
>  
> - To the contrary even, my impression is that the current paper is *much more* specific to micropolicy-based enforcement, than Skorstengaard et al.'s approach is specific to capabilities.
>   For example, the use of function, call, return and entry maps are big assumptions that rule out dynamic code generation, dynamic linking, JIT-compilation, variable-arity functions (e.g. C varargs) and possibly other features.
>   At the same time, the maps align suspiciously well with the requirements of micropolicy-based enforcement.
>   Other aspects of the approach appear to rule out the use of per-component stacks or separate control/data stacks.
>  
> - Skorstengaard et al. actually provide a list of informal criteria for judging a characterization of full abstraction:
>   "(1)intuitive
>    (2)useful for reasoning: we should be able to use WBCF and LSE when reasoning about correctnessand security of programs using StkTokens.
>    (3)reusable in secure compiler chains: for compilers using StkTokens, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
>    (4)arguably "complete": the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
>    (5)potentially scalable: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings."
>   In a paper claiming to improve on their work, I think you are morally obliged to at least mention these criteria and explain why you believe your approach scores better on these criteria.
>   Of course, if you don't agree with the criteria, you are free to criticize them and/or propose your own.
>   But leaving them entirely unmentioned strikes me as odd.
>  
> - Taking a quick look at the criteria above, (1) seems subjective, but the others can clearly be used to compare the two appraoches:
>   (2) useful for reasoning: Skorstengaard et al. 2019a seems to win here, since it offers a logical relation and use it to prove soundness of an assembly implementation of the "awkward example" (whose validity depends on stack safety).
>       It is useful to think about how one would prove this awkward example safe using the formalizations of stack safety of Skorstengaard et al. 2019b and the current paper.
>  
>       Skorstengaard et al. 2019b unfortunately leave the awkward example undiscussed, but it seems like one could prove the example safe in the natively stack-safe semantics and then use full abstraction to extrapolate the results to the real semantics.
>       This seems like it would work, although perhaps only for formulations of safety that can be expressed as program equivalences.
>  
>       In the current approach, it seems like one would need to reason about a trace of execution, the different calls and returns and then deduce something from that.
>       However, encapsulation of heap locations like the one used in the awkward example seems out of scope, so one would either need to combine this with a formulation of memory safety (perhaps Azevedo de Amorim 2018?) or allocate the location on the stack instead.
>       Unfortunately, the latter wouldn't work because this paper's approach doesn't support closures that have access to the stack frame in which they were defined. 
>  
>       In summary, on this criterion, I would expect to score the contenders as Skorstengaard 2019a > Skorstengaard 2019b > current paper
>  
>   (3) This secure compilation criterion seems to be something that Skorstengaard 2019b was specifically designed for, so perhaps it's a bit biased.
>       But still, it seems potentially desirable and I don't immediately see how this could be achieved with Skorstengaard 2019a or the current paper.
>  
>   (4) This criterion asks for an argument that any property of a program that relies on stack safety can be proven with the characterization of stack safety.
>       For Skorstengaard 2019b, this can be argued because any property that holds in the natively stack safe source language can be carried over to the original assembly.
>       On the other hand, full abstraction only applies to program equivalences, not safety properties or other.
>       For Skorstengaard 2019a, there's no argument and I don't see an argument for the current paper.
>  
>   (5) dynamic code generation seems like it might be supported in Skorstengaard 2019a and 2019b (with some non-trivial extra effort), but it seems quite contradictory with the use of function, call, return and entry maps in the current paper.
>       concurrency seems fine.
>       The current paper also seems like it would have a lot of trouble with closures that have access to the stack frame in which they were defined, because the stack safety properties requires the original stack frames to be unmodified and not influence the influence of higher stack frames.
>  
>   In summary, this quick exercise to evaluate the current approach on the intuitive criteria proposed by Skorstengaard et al., does not strike me as very favorable for the current approach, to the contrary even.
>   Perhaps this is because Skorstengaard et al. 2019b chose criteria on which they knew their approach would do well, but then why don't the current authors don't propose any alternative criteria themselves?
>  
> I believe the authors should do at least the following:
> * remove the claim about Skorstengaard et al., 2019b or substantiate it with a clear argument about why it is specific to capability machine-based enforcement.
> * discuss the criteria proposed by Skorstengaard et al. and investigate how their approach "scores" on them.
>   For example, they should demonstrate a specific program whose validity depends on stack safety and demonstrate how they would reason about it using their criterion?
>   Also, they should discuss language features (more than the ones currently in section 9, like closures and dynamic code generation and discuss how their approach could potentially scale to it in the long term.
> * discuss in more detail the fundamental limitations of their approach, particularly the assumption of a static code layout.
> * provide a clear, maximally objective/neutral statement of the benefits of their approach w.r.t. Skorstengaard et al. 2019b
>  
> The authors do not discuss any attacker model, both for their properties and for their enforcement technique.
> Who can modify the micropolicy tables?
> Do you have to trust the operating system or just the hardware?
> Is the attacker capable of running arbitrary assembly code (like in Skorstengaard et al, 2019ab) or just inject stack modifications like in Abadi et al., 2005?
>  
> Another odd aspect of this paper is the fact that it doesn't propose one but three increasingly strong properties that characterize stack safety (see l86-96).
> What does this mean: why do we want stack safety to mean different things and which meaning is appropriate in which setting?
> Do the three variants score equally well on the above criteria?
> - For the "lockstep" variant, I get the impression that this is a proof-technical definition that is only used for proving the regularly stack-safety property for the micropolicy enforcement and not really important/useful in itself: is this also how you see it?
> - The lazy variant of stack safety strikes me as problematic because of several reasons:
>   * It seems very specific to exactly one enforcement technique.
>     There isn't a single argument in favour of this property except that it is enforced by this one enforcement technique.
>   * The property is quite complex to formalize and understand. I found it hard to motivate myself to study it in detail, because it is rather complicated and so specific to this one type of enforcement.
>   * The lazy enforcement seems to create new side-channels in the presence of an attacker being invoked by an adversary.
>     Can't the lazy enforcement be exploited by such an adversary to cause the trusted code to halt in the middle of its execution, possibly in an inconsistent state?
>     This reminds me of page-table based attacks in the context of SGX, see, for example, Xu et al. (https://ieeexplore.ieee.org/abstract/document/7163052?casa_token=0ycUzVYD0FoAAAAA:bi8eLFwzqvCnRjtcAWD-y9xG473kJeLQdZaLWEbU43cpAwvGldBu840sXWtcC2mnOnOqGsUPruc).
>  
> I'm also a bit unconvinced about the paper's motivation.
> Based on comments throughout the paper, they seem to be looking for a property that describes in detail how the stack looks in memory and how the different memory locations are allowed to be modified or how it is allowed to influence the running program.
> I get the impression they are motivated to find such a property in order to be able to derive a policy that can be enforced on a micropolicy machine.
> In my opinion, this is not an unreasonable goal, but it is a different goal than finding an abstract definition of stack safety that is intuitive, abstract, and maximally applicable to diverse enforcement mechanisms and complicated features like dynamic code generation, varargs etc.
> Properties from related work (Skorstengaard et al., 2019b) seem to strive for a more abstract statement along the lines of "the program should behave the same as when it were running on a machine with a native stack, even in the presence of an adversary" and avoids going into the matter of precisely which memory cell is allowed to be modified at which moment.
> I have the impression the paper would be improved by being more clear about the real goal of the work: finding a more concrete specification of stack safety that is closer to implementable micro-policies.
> To this end, the authors should be open about this and remove statements that suggest their goal is different, for example l29-31 "there is still room to craft a more abstract characterization of stack safety that is independent of capabilities or any other specific enforcement technique".
>  
>  
> General comments:
> * There is some important missing related work on formalizing stack safety:
>   - Martín Abadi, Mihai Budiu, Úlfar Erlingsson, and Jay Ligatti. 2005. A Theory of Secure Control Flow. In Formal Methods and Software Engineering. Springer Berlin Heidelberg.
>   - Martín Abadi, Mihai Budiu, Úlfar Erlingsson, and Jay Ligatti. 2005. Control-flow Integrity. In Conference on Computer and Communications Security. ACM. https://doi.org/10.1145/1102120.1102165
>   This probably will not compromise novelty of the paper's contributions, but I think it must be cited and compared against.
>  
> * l17-21: strange that you don't mention CFI here or on l149-150 where you mention "[stack safety enforcement using] code rewriting".
>  
> * l23: "defined in a negative way": it's quite unclear what you mean exactly with this phrase.
>   For example, does Abadi2005 (see above) fall under this phrasing?
>   Skorstengaard 2019b can also be described negatively as "Full abstraction should not be broken." and the same applies to this paper.
>   Can you more precisely describe what you mean?
>  
> * l197-244: these figures are very clear and helpful, nice...
>   perhaps you could additionally annotate l203-211 as "main", l219-225 as "f" and l233-237 as "g", to clearly point out the link to Fig. 1
>  
> * l299-300 and Fig.2: I find it confusing that you use JALR (jump-and-link-register) when you really mean JR (jump to register).
>   IIUC, this is because RISC-V doesn't actually offer JR because it can be emulated as JALR with r0 as the target for storing the return pointer.
>   However, I would prefer if you simply pretend that a direct JR instruction exists. 
>  
> * l318: "components K": odd name. why not locations as in compiler terminology?
>  
> * l426: "and a machine that steps for a long time before producing some output": actually, it's even more strange: the observer cannot distinguish between a machine that steps forever without producing output and a machine that *immediately* produces output
>  
> * l483: err.. shouldn't you also require that the stacks below SP are the same in m_c and m_r and that there have been no returns beyond SP between m_c and m_r?
>  
> * l515-516: here, I found myself confused because I was interpreting your policy as an information flow policy, so I thought you were allowing to write data from uninitialized stack locations to out.
>  
> * l563: "Intuitively, the caller expects to find their private data untouched after the callee returns": what about settings where closures can access the stack frame they were defined in?
>  
> * l688 "it also holds for the toplevel execution trace to account for unallocated memory": I don't understand this, please elaborate.
>  
> * l1110-1126: I don't understand what you're saying here: Roessler and Dehon's lazy tag policies proposals do not actually give you lazy stack safety?
>   Is there one enforcement technique that satisfies this lazy version of stack safety?
>  
> Minor comments:
>  
> * l153: "they been applied"
>  
> * l298: "[jal] atomically simultaneously the return address in a register": verb missing
>  
> * l368: "coinductively": I would say "corecursively" here
>  
> * l393: "T in Traces(A) x Obs": do you mean "Traces(T x Obs)"?
>  
> * l405: "if f be a predicate": if f is a predicate?
>  
> * l537: "MP-trace. say MP": capitalize Say?
>  
> * l559: "marked HC in C": HI?
>  
> * l687: "TraceConf justret m_c MP_call M_call": parentheses around justret m_c
>  
> * l872: "with regard to the original machine state": what does this mean?
>  
> Questions for the response period
> ---------------------------------
> * You seem to say/suggest that the approach of Skorstengaard 2019b to defining stack safety is specific to capability machines.
>   Do you really think so and if so, why do you think their approach could not be applied to a different enforcement mechanism?
>  
> * What exactly is the benefit of your approach to defining stack safety w.r.t. the approach of Skorstengaard 2019b?
>   Do you think it can be applied to more or less languages?
>   Do you think it is better suited or less well suited for reasoning about programs which depend on stack safety?
>   Do you think it is etc.
>  
> * Do you think you could show safety of a specific example program that depends on stack safety (like the awkward example) and if so, how would such a proof work? 
>  
> * Do you think your approach to defining stack safety could be extended to languages with the following features (and if so, how?):
>   - varargs?  (seems problematic for call map)
>   - dynamic code generation (seems problematic for the function map, call map and return map)
>   - dynamic linking (seems problematic for the function map, call map and return map)
>   - self-editing code (seems problematic for call map and return map)
>   - tail calls (seems potentially okay, but I'd like to be sure) 
>   - exceptions and stack unwinding
>   - closures (seems potentially problematic for precision of the entry map)
>   How would you extend your approach to define stack safety in the presence of these features?
>   
> * Do you think your approach implies the following properties?
>   - temporal safety in the presence of closures (i.e. closures should only be invokable by the adversary after the adversary first receives a reference to them)
>   - encapsulation of closures: it should not be possible for the adversary to invoke a closure with a different environment than the one they received? 
>   If not, do you think it should imply these properties and if so, how do you think it could be extended?
>  
> * Are the lazy and strong variant of your properties specifically intended for the corresponding micropolicy-based enforcement techniques or do you believe they are more broadly applicable?

### 2020-09-08 main.tex
Oops, the RA's and final main_priv seem to be off
Typo in figure: code is low confidentiality

@@
 % Register File
-\summary{40}{0}{\thepcctr}{109}{18}{?}
+\summary{40\rb{44}}{0}{\thepcctr\rb{27}}{109}{18}{?}
 \vspace*{0.2em}
@@
   \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
-      {Deallocate {\tt g}'s local state\rb{Yes, still {\tt g}}}
+      {Deallocate {\tt g}'s local state}
   \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 27}
@@
 % Starting Register File
-\summary{27}{0}{\thepcctr}{108}{18}{?}
+\summary{27}{0}{\thepcctr\rb{27}}{108}{18}{?}
 % \vspace*{0.2em}
@@
 \begin{array}{l|l|l|l}
-  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
+  \row{\lw ~ r_4 ~ \rsp ~ 0\rb{\negate 1}}{r_4 \leftarrow [\rsp \rb{- 1}]}{r_4 \leftarrow 18}
       {Load return value into $r_4$}
@@
 % Register File
-\summary{8}{0}{\thepcctr}{104}{18}{0}
+\summary{8}{0}{\thepcctr\rb{8}}{104}{18}{0}
 %\vspace*{0.2em}
@@
 \MemoryLabel{-51em}{0.75em}{60}
-\MemoryLabel{-47em}{0.75em}{60}
-\MemoryLabel{-43em}{0.75em}{42}
+\MemoryLabel{-47em}{0.75em}{60\rb{42}}
+\MemoryLabel{-43em}{0.75em}{42\rb{18}}
 \\
@@
 % Starting Register File
-\summary{\thepcctr}{0}{37}{102}{60}{42}
+\summary{\thepcctr}{0}{37\rb{8}}{102}{60}{42}
 \caption{Detailed Execution Trace
@@
 \MemoryLabel{43.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{$(\HIGHSEC,\HIGHINT)$}}]%
+\memory{3}{\instrc}[{\makebox[0pt]{$(\LOWSEC,\HIGHINT)$}}]%
 \hspace*{3pt}
@@
     For our example, we return to the violation of trace integrity in
-    \cref{fig:intex}, in which {\tt g} \rb{??} has written to {\tt z} and returned.
+    \cref{fig:intex}, in which {\tt g} has written to {\tt z} and returned.
     Whether observable integrity holds depends on the continuing trace, shown
@@
       \(\mach_r'\) and the restored
-      \[\mach_r'' = \mathit{restore}(\contour,
+      \[\mach_r'' = \mathit{restore}(%\contour,
       \mach_e,\mach_e',\mach_r,\mach_r').
       \]
-      Addresses 107 and 108
+      Addresses 106 and 107
       are not restored; every address that was varied and not changed in one
-      of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
+      of the executions is. So 107 contains 43 in the trace from \(\mach_r\),
       and after returning from {\tt f}, {\tt main} prints 85. But in the trace
-      from \(\mach_r''\), 108 contains $-4$, and ultimately 38 is printed.
+      from \(\mach_r''\), 107 contains $-4$, and ultimately 38 is printed.
       So observable trace confidentiality does not hold. This example

### 2020-09-08 rebuttal.txt
Some comments on the rebuttal for the 5 criteria

@@
 (5) scalable: what can we extend this to? the maps appear particularly restrictive here...
+
+SNA: Thoughts on the criteria.
+(1) The LSE property can be conceptually divided into the contour function, which divides the stack into
+    HC/LC and HI/LI components, and the actual definitions of confidentiality and integrity w.r.t. that
+    division. Insofar as confidentiality and integrity are believable, I believe the contour is very
+    intuitive. Confidentiality is not especially intuitive, but is based on non-interference which
+    is well-known. Eager integrity is pretty intuitive. The lazy properties are highly non-intuitive,
+    which makes sense as the policies they aim to characterize permit violations of the standard intuition.
+
+    The WBCF property simply describes the standard behavior of stack-based control flow, it seems very intuitive.
+
+    Relating to Skorstengaard, they rely on different prior intuitions. Skorstengaard is intuitive to exactly the
+    degree that one understands the behavior of a pure stack in general and the built-in stack of oLCM in
+    particular. As the behavior of stacks is generally understood, Skorstengaard has an edge, but there is
+    value in offering an intuition that is less circularly dependent on understanding stacks.
+(2) We probably don't have anything to say here. A "Meaning of Memory Safety"-style frame rule would be just the
+    ticket, but we haven't done it.
+(3) Not really applicable, since this is a criterion tailored to what Skorstengaard is aiming to do. We might consider
+    what the equivalently tailored criterion would be for us -- usable in "dynamic enforcement environments" with
+    potentially unsafe commodity compilers? Where we score well but not perfectly, because while we do need some compiler 
+    support it's in the form of very basic control flow information.
+(4) Subject to the limitations of the formalization, our properties should apply to any reasonable program.
+    But we would need to make some argument that we're not missing something obvious.
+(5) It would be very challenging to mess with dynamic code, unfortunately. In the tag world we would handle
+    generated code by tagging the instructions that generate it, I suppose. Which would correspond to
+    a map that allows some instructions to update the maps, including potentially itself, etc...
+
+    On the other hand, we described how we would handle coroutines, and threads would be similar. Basically
+    when we lose the property of the contour that exiting a context means either shrinking access monotonically
+    or never returning to that context, we need to change the LSE property to vary at each entry to a context
+    and exit at each return. But once that is done the space of threads and coroutines should be wide open.
+    As for WBCF, we would need to update Control Separation and Entry Integrity to account for changes in
+    acceptable modes of control transfer. Return integrity applies to calls and returns even if separated by
+    yields.
 

### 2020-09-08 rebuttal.txt
Above the fold

--- 2020-09-08 rebuttal.txt
@@
-Motivation - Applicability:
+We thank the reviewers and in particular reviewer C for their
+thoughtful comments.
 
-Our original goal in this work wasn't to formalize a stack safety
-notion that is specific to micropolicies, but in hindsight we agree
-that the result appears to be tailored to that particular enforcement
-mechanism---in large part because of our presentation choices and
-exposition-helping simplifications. In particular, we believe that the
-enforcement model can be used (with some effort) to describe
-capability-based enforcement mechanisms (as we state in Section 2).
-However, given that this effort didn't make it into the final product,
-we agree that the material present would be more accurately described
-as an abstract property capturing micropolicy enforcement rather than
-the all-encompassing stack safety notion we set out to state.
-LEO: ...and we plan to do so for the full version?
+We first discuss the most important points raised in reviews. Then, in
+an optional appendix, we respond point-by-point to each individual
+piece of reviewer feedback.
 
-Attacker Model:
+Reviewer C is particularly concerned with the way this work is
+motivated. Our original goal in this work wasn't to formalize a stack
+safety notion that is specific to micropolicies, but in hindsight we
+agree that the result appears to be tailored to that particular
+enforcement mechanism---in large part because of our presentation
+choices and exposition-helping simplifications. In particular, we
+believe that the enforcement model can be used (with some effort) to
+describe capability-based enforcement mechanisms (as we state in
+Section 2).  However, given that this effort didn't make it into the
+final product, we agree that the material present would be more
+accurately described as an abstract property capturing micropolicy
+enforcement rather than the all-encompassing stack safety notion we
+set out to state, and intend to do so for a final version.
 
-The attacker model for the properties is discussed in the paper
+All reviewers also wanted additional discussion of related work.  We
+agree that there should be a clearer comparison to both Skorstengaard
+papers, using the informal criteria layed out there. We plan to
+include a discussion like the following:
+
+(1) On intuition: The LSE part of property we present aims to
+characterize stack safety in security terms. For readers familiar with
+integrity/confidentiality/noninterference, the eager version should be
+fairly intuitive. The lazy/observable properties, are less intuitive,
+but more interesting in the violations they permit. The WBCF part simply
+formalizes standard behavior of stack-based control flow.
+
+On the other hand, Skorstengaard relies on different prior intuitions:
+for readers familiar with the behavior of pure stacks (or rather the
+built-in oLCM stack), their property is more intuitive. For readers
+versed in security, we believe there is value in a different
+perspective.
+
+(2) useful for reasoning: In order to reason about actual programs we
+would indeed need to incorporate a "Meaning of Memory Safety"-style
+frame rule. Our goal with this paper was to distill the essence of
+stack safety without adding the complex (but solved) overhead of
+dealing with heap locations. But the properties (and the testing
+framework behind it) were designed with [Azevedo et al] in mind.
+
+(3) reusable in compiler chains: As you point out, this is not really
+applicable, since this is a criterion tailored to what Skorstengaard
+is aiming to do. We might consider what the equivalently tailored
+criterion would be for us -- usable in "dynamic enforcement
+environments" with potentially unsafe commodity compilers.
+
+LEO: I'm REALLY not sure about this:
+(4) complete: Subject to the limitations of the formalization, our
+properties should apply to any reasonable program.
+
+(5) scalable: Dynamic code would pose a significant challenge in this
+setup. In the tag world, we could handle generated code by tagging the
+instructions that generate it, which would correspond to a map that
+allows some instructions to update the maps, including potentially
+itself.  On the other hand, we described how we would handle
+coroutines, and threads would be similar. Basically when we lose the
+monotonicity of contours throughout an execution trace, and we would
+need to change the LSE property to vary at each context entry and and
+to exit at each return.
+
+Finally, reviewer C is also concerned with the attacker model. The
+attacker model concerning the properties is discussed in the paper
 (Section 2). For example, lines 104-106 state that "the attacker can
@@
 mechanism is described in [Roessler and Dehon 18].
-
-5 Criteria - StkTokens:
-LEO: Not really sure what to say here. 
-(1) intuitive: this is subjective but we were aiming for a more extensional property than full abstraction from a "clearly stack safe" machine
-(2) useful for reasoning: LEO: do we have anything to say here other than you're right?
-(3) reusable in compiler chains: not really applicable?
-(4) complete: ??
-(5) scalable: what can we extend this to? the maps appear particularly restrictive here...
 

### 2020-09-08 rebuttal.txt
Quick comments
a few notes on C's questions

@@
 LEO: I'm REALLY not sure about this:
+APT: Reviewer C interprets (4) to mean: 
+"This criterion asks for an argument that any property of a program
+that relies on stack safety can be proven with the characterization of stack safety."
+Since we're in the business of _defining_ stack safety, this seems pretty circular to me!
+Maybe the point is that we should not omit any characteristic of
+everyone's informal notion of stack safety (for the set of language features
+we support -- which is the business of (5)).  Can we claim that (informally, of course)?
 (4) complete: Subject to the limitations of the formalization, our
@@
 >   Do you really think so and if so, why do you think their approach could not be applied to a different enforcement mechanism?
->  
+APT:  Sk2019b _define_ stack safety via a capability machine semantics that is supposed to embody stack
+safety "by construction," and relating it to a standard semantics for the same machine via a full abstraction result.
+We did not mean to imply that the same idea could not work for other machine models or enforcement mechanisms (although
+this generalization is not actually in Sk2019b, even as suggested future work). Rather, we wished to distinguish our
+approach from theirs by emphasizing that we define stack safety based not on any "correct by construction" 
+machine tied to some flavor of enforcement,  but rather on more abstract notions of observable behavior. 
 > * What exactly is the benefit of your approach to defining stack safety w.r.t. the approach of Skorstengaard 2019b?

### 2020-09-09 rebuttal.txt

@@
 
+BCP: Basically good, but I think we can be a little less apologetic, and
+stronger on what we'd propose to do to fix it.
+
 All reviewers also wanted additional discussion of related work.  We
@@
 
-On the other hand, Skorstengaard relies on different prior intuitions:
+By contrast, Skorstengaard relies on different prior intuitions:
 for readers familiar with the behavior of pure stacks (or rather the
@@
 >   Do you really think so and if so, why do you think their approach could not be applied to a different enforcement mechanism?
+
 APT:  Sk2019b _define_ stack safety via a capability machine semantics that is supposed to embody stack
@@
 machine tied to some flavor of enforcement,  but rather on more abstract notions of observable behavior.
+
+BCP: Nice.
+
 > * What exactly is the benefit of your approach to defining stack safety w.r.t. the approach of Skorstengaard 2019b?

### 2020-09-09 rebuttal.txt
More comments

@@
-We thank the reviewers and in particular reviewer C for their
-thoughtful comments.
+We thank the reviewers and in particular reviewer C for their thoughtful
+comments.  We first discuss the most important points raised in
+reviews. Then, in an optional appendix, we respond point-by-point to each
+individual piece of reviewer feedback.
 
-We first discuss the most important points raised in reviews. Then, in
-an optional appendix, we respond point-by-point to each individual
-piece of reviewer feedback.
+MOTIVATION
 
 Reviewer C is particularly concerned with the way this work is
-motivated. Our original goal in this work wasn't to formalize a stack
-safety notion that is specific to micropolicies, but in hindsight we
-agree that the result appears to be tailored to that particular
-enforcement mechanism---in large part because of our presentation
-choices and exposition-helping simplifications. In particular, we
-believe that the enforcement model can be used (with some effort) to
-describe capability-based enforcement mechanisms (as we state in
-Section 2).  However, given that this effort didn't make it into the
-final product, we agree that the material present would be more
-accurately described as an abstract property capturing micropolicy
-enforcement rather than the all-encompassing stack safety notion we
-set out to state, and intend to do so for a final version.
+motivated. Our original goal in this work wasn't to formalize a stack safety
+notion that is specific to micropolicies, but in hindsight we agree that the
+result may appear to be tailored to that particular enforcement mechanism,
+though we believe this is more a matter of presentation choices and
+exposition-helping simplifications than a fundamental limitation. In
+particular, we believe that the enforcement model can be used (with some
+effort) to describe capability-based enforcement mechanisms, as we state in
+Section 2.  However, given that this effort didn't make it into the final
+product, we agree that the paper as it stands would be more accurately
+described as an abstract property capturing micropolicy enforcement rather
+than the all-encompassing stack safety notion we set out to state, and
+intend to do so for a final version.
 
-BCP: Basically good, but I think we can be a little less apologetic, and
-stronger on what we'd propose to do to fix it.
+BCP: I.e., we plan to limit our claims rather than fixing the exposition?
+Why not promise the latter (and lay out what we would do)?
 
-All reviewers also wanted additional discussion of related work.  We
-agree that there should be a clearer comparison to both Skorstengaard
-papers, using the informal criteria layed out there. We plan to
-include a discussion like the following:
+RELATED WORK
 
-(1) On intuition: The LSE part of property we present aims to
+All reviewers also wanted additional discussion of related work.  We agree
+that there should be a clearer comparison to both Skorstengaard papers,
+using the informal criteria proposed there. We plan to include a discussion
+like the following:
+
+(1) On intuition: The LSE part of the property we present aims to
 characterize stack safety in security terms. For readers familiar with
@@
 formalizes standard behavior of stack-based control flow.
+
+BCP: "more interesting in the violations they permit"???
 
@@
 
-Finally, reviewer C is also concerned with the attacker model. The
-attacker model concerning the properties is discussed in the paper
-(Section 2). For example, lines 104-106 state that "the attacker can
-run arbitrary machine code" and that "hardware and timing attacks are
-out of scope". The attacker model of the micropolicy enforcement
-mechanism is described in [Roessler and Dehon 18].
+ATTACKER MODEL
+
+Finally, reviewer C is also concerned with the attacker model. The attacker
+model we consider is discussed in Section 2.  In particular, lines 104-106
+state that "the attacker can run arbitrary machine code" and that "hardware
+and timing attacks are out of scope". The attacker model of the micropolicy
+enforcement mechanism is described in [Roessler and Dehon 18].
+
+BCP: Can we sketch it here?
 
@@
 >   But then why you didn't use the observable version as the authoritative definition?
->
+
+BCP: This is a silly objection, but it should be addressed clearly and
+strongly.  E.g.:
+
+We propose two definitions of stack safety: one that is easier to understand
+because it corresponds to a more naive view of what stack safety intuitively
+means, and a slightly more complex and permissive one that (as Roessler and
+Dehon showed) can be implemented more efficiently.  We believe both are of
+interest.
+
 > - This paper is lacking proper discussion on related work.
@@
 >       For example, see CompCert's assumption on function calls: https://github.com/AbsInt/CompCert/blob/94558ecb3e48261f12c644045d40c7d0784415e0/common/Events.v#L619
->
+
+BCP: Obviously, this has to be addressed strongly and concretely.
+
 > Questions for the response period

### 2020-09-09 rebuttal.txt
Closure notes

--- 2020-09-08 rebuttal.txt
@@
-We thank the reviewers and in particular reviewer C for their
-thoughtful comments.
+Motivation - Applicability:
 
-We first discuss the most important points raised in reviews. Then, in
-an optional appendix, we respond point-by-point to each individual
-piece of reviewer feedback.
+Our original goal in this work wasn't to formalize a stack safety
+notion that is specific to micropolicies, but in hindsight we agree
+that the result appears to be tailored to that particular enforcement
+mechanism---in large part because of our presentation choices and
+exposition-helping simplifications. In particular, we believe that the
+enforcement model can be used (with some effort) to describe
+capability-based enforcement mechanisms (as we state in Section 2).
+However, given that this effort didn't make it into the final product,
+we agree that the material present would be more accurately described
+as an abstract property capturing micropolicy enforcement rather than
+the all-encompassing stack safety notion we set out to state.
+LEO: ...and we plan to do so for the full version?
 
-Reviewer C is particularly concerned with the way this work is
-motivated. Our original goal in this work wasn't to formalize a stack
-safety notion that is specific to micropolicies, but in hindsight we
-agree that the result appears to be tailored to that particular
-enforcement mechanism---in large part because of our presentation
-choices and exposition-helping simplifications. In particular, we
-believe that the enforcement model can be used (with some effort) to
-describe capability-based enforcement mechanisms (as we state in
-Section 2).  However, given that this effort didn't make it into the
-final product, we agree that the material present would be more
-accurately described as an abstract property capturing micropolicy
-enforcement rather than the all-encompassing stack safety notion we
-set out to state, and intend to do so for a final version.
+Attacker Model:
 
-All reviewers also wanted additional discussion of related work.  We
-agree that there should be a clearer comparison to both Skorstengaard
-papers, using the informal criteria layed out there. We plan to
-include a discussion like the following:
-
-(1) On intuition: The LSE part of property we present aims to
-characterize stack safety in security terms. For readers familiar with
-integrity/confidentiality/noninterference, the eager version should be
-fairly intuitive. The lazy/observable properties, are less intuitive,
-but more interesting in the violations they permit. The WBCF part simply
-formalizes standard behavior of stack-based control flow.
-
-On the other hand, Skorstengaard relies on different prior intuitions:
-for readers familiar with the behavior of pure stacks (or rather the
-built-in oLCM stack), their property is more intuitive. For readers
-versed in security, we believe there is value in a different
-perspective.
-
-(2) useful for reasoning: In order to reason about actual programs we
-would indeed need to incorporate a "Meaning of Memory Safety"-style
-frame rule. Our goal with this paper was to distill the essence of
-stack safety without adding the complex (but solved) overhead of
-dealing with heap locations. But the properties (and the testing
-framework behind it) were designed with [Azevedo et al] in mind.
-
-(3) reusable in compiler chains: As you point out, this is not really
-applicable, since this is a criterion tailored to what Skorstengaard
-is aiming to do. We might consider what the equivalently tailored
-criterion would be for us -- usable in "dynamic enforcement
-environments" with potentially unsafe commodity compilers.
-
-LEO: I'm REALLY not sure about this:
-APT: Reviewer C interprets (4) to mean: 
-"This criterion asks for an argument that any property of a program
-that relies on stack safety can be proven with the characterization of stack safety."
-Since we're in the business of _defining_ stack safety, this seems pretty circular to me!
-Maybe the point is that we should not omit any characteristic of
-everyone's informal notion of stack safety (for the set of language features
-we support -- which is the business of (5)).  Can we claim that (informally, of course)?
-(4) complete: Subject to the limitations of the formalization, our
-properties should apply to any reasonable program.
-
-(5) scalable: Dynamic code would pose a significant challenge in this
-setup. In the tag world, we could handle generated code by tagging the
-instructions that generate it, which would correspond to a map that
-allows some instructions to update the maps, including potentially
-itself.  On the other hand, we described how we would handle
-coroutines, and threads would be similar. Basically when we lose the
-monotonicity of contours throughout an execution trace, and we would
-need to change the LSE property to vary at each context entry and and
-to exit at each return.
-
-Finally, reviewer C is also concerned with the attacker model. The
-attacker model concerning the properties is discussed in the paper
+The attacker model for the properties is discussed in the paper
 (Section 2). For example, lines 104-106 state that "the attacker can
@@
 mechanism is described in [Roessler and Dehon 18].
+
+5 Criteria - StkTokens:
+LEO: Not really sure what to say here. 
+(1) intuitive: this is subjective but we were aiming for a more extensional property than full abstraction from a "clearly stack safe" machine
+(2) useful for reasoning: LEO: do we have anything to say here other than you're right?
+(3) reusable in compiler chains: not really applicable?
+(4) complete: ??
+(5) scalable: what can we extend this to? the maps appear particularly restrictive here...
+
+SNA: Thoughts on the criteria.
+(1) The LSE property can be conceptually divided into the contour function, which divides the stack into
+    HC/LC and HI/LI components, and the actual definitions of confidentiality and integrity w.r.t. that
+    division. Insofar as confidentiality and integrity are believable, I believe the contour is very
+    intuitive. Confidentiality is not especially intuitive, but is based on non-interference which
+    is well-known. Eager integrity is pretty intuitive. The lazy properties are highly non-intuitive,
+    which makes sense as the policies they aim to characterize permit violations of the standard intuition.
+
+    The WBCF property simply describes the standard behavior of stack-based control flow, it seems very intuitive.
+
+    Relating to Skorstengaard, they rely on different prior intuitions. Skorstengaard is intuitive to exactly the
+    degree that one understands the behavior of a pure stack in general and the built-in stack of oLCM in
+    particular. As the behavior of stacks is generally understood, Skorstengaard has an edge, but there is
+    value in offering an intuition that is less circularly dependent on understanding stacks.
+(2) We probably don't have anything to say here. A "Meaning of Memory Safety"-style frame rule would be just the
+    ticket, but we haven't done it.
+(3) Not really applicable, since this is a criterion tailored to what Skorstengaard is aiming to do. We might consider
+    what the equivalently tailored criterion would be for us -- usable in "dynamic enforcement environments" with
+    potentially unsafe commodity compilers? Where we score well but not perfectly, because while we do need some compiler 
+    support it's in the form of very basic control flow information.
+(4) Subject to the limitations of the formalization, our properties should apply to any reasonable program.
+    But we would need to make some argument that we're not missing something obvious.
+(5) It would be very challenging to mess with dynamic code, unfortunately. In the tag world we would handle
+    generated code by tagging the instructions that generate it, I suppose. Which would correspond to
+    a map that allows some instructions to update the maps, including potentially itself, etc...
+
+    On the other hand, we described how we would handle coroutines, and threads would be similar. Basically
+    when we lose the property of the contour that exiting a context means either shrinking access monotonically
+    or never returning to that context, we need to change the LSE property to vary at each entry to a context
+    and test at each exit from it. But once that is done the space of threads and coroutines should be wide open.
+    As for WBCF, we would need to update Control Separation and Entry Integrity to account for changes in
+    acceptable modes of control transfer. Return integrity applies to calls and returns even if separated by
+    yields.
+
+SNA: Thoughts on closures (very sketchy).
+
+  Suppose we added yet another partial map (sigh) representing the sealing of a closure. Suppose as well that our
+  model is enhanced to track the provenance of values as discussed in the future work section on stack-derived
+  pointers. A closure map would take a code address as usual and yield: a register r that holds the pointer, an
+  entry point e, a unique Seal identifier i, and a set of components ks. We extend the tuple of a security state a
+  contour can give with (forall i, Seal i). When we take a step from an instruction where the closure map is defined,
+  we update the contour C so that for each k in ks, C(k) is (*,*,Seal i). We also need to add state to track the
+  association between i, e, and r, and carry that association to any k with pointer provenance from r.
+  Finally, we modify call maps to give an optional component whose contents are being used as a function pointer.
+
+  Then, when a call map says that we are making a call using a closure associated with (Seal i), the contour
+  update sets all (Seal i) components (LC,LI) and LSE works just the same, treating sealed components as (HC/HI).
+  WBCF will need an extra clause stating that we really did jump to e. (I'm a little unclear as to whether a
+  closure should be re-sealed on return, but we can do that if we need to.)
+
+  (Notes: arg, maybe we actually need a list of closures.)
+
+  To answer the questions: yes, we believe this can be extended to closures, and with these additions stack safety
+  would indeed imply both temporal safety and encapsulation of closures.
 
@@
 >   Do you really think so and if so, why do you think their approach could not be applied to a different enforcement mechanism?
-APT:  Sk2019b _define_ stack safety via a capability machine semantics that is supposed to embody stack
-safety "by construction," and relating it to a standard semantics for the same machine via a full abstraction result.
-We did not mean to imply that the same idea could not work for other machine models or enforcement mechanisms (although
-this generalization is not actually in Sk2019b, even as suggested future work). Rather, we wished to distinguish our
-approach from theirs by emphasizing that we define stack safety based not on any "correct by construction" 
-machine tied to some flavor of enforcement,  but rather on more abstract notions of observable behavior. 
+>  
 > * What exactly is the benefit of your approach to defining stack safety w.r.t. the approach of Skorstengaard 2019b?

### 2020-09-09 rebuttal.txt

--- 2020-09-09 rebuttal.txt
@@
 coroutines, and threads would be similar. Basically when we lose the
-monotonicity of contours throughout an execution trace, and we would
-need to change the LSE property to vary at each context entry and and
-to exit at each return.
+monotonicity of contours relative to stack depth, we will
+need to change the LSE property to check integrity and confidentiality
+each time we cross between security contexts - nested calls are not
+strictly more limited.
 
@@
 >   If not, do you think it should imply these properties and if so, how do you think it could be extended?
->
+
+  Adding closures requires:
+  - identifying code points where closures are created
+  - adding machinery to track association of closure identities, entry points, and components
+    containing valid closure pointers (including pointer provenance, as with stack-derived pointers)
+  - extending WBCF to require that closure invocations respect the closure's designated entry point
+    and use a valid closure pointer (ensuring temporal safety and encapsulation)
+  - extending contours to treat some components as sealed, associated with closure identities
+  - on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
+    (allowing access to closure)
+  - otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
+    nested calls (monotonicity is lost)
+
+  The answers are yes, we believe that with addition of closures, natural extensions to WBCF and LSE would imply these
+  properties.
+
 > * Are the lazy and strong variant of your properties specifically intended for the corresponding micropolicy-based enforcement techniques or do you believe they are more broadly applicable?

### 2020-09-09 rebuttal.txt
Added some thoughts on closures

@@
 
+SNA:
   Adding closures requires:

### 2020-09-09 rebuttal.txt
CP/comments

@@
 I think this is an very important question as recent work points out cross-thread stack smashing attack [1].
->
+
+We sketch how multiple threads/coroutines would work in the future
+work section (l 1213-1219). In particular, we would need to further
+distinguish stack locations as either "accessible" (current frame),
+"belonging to the same call stack", or "belonging in another
+stack". And then integrity would require that when a function changes
+the locations in the "same" stack should be unchanged. Similarly for
+cofnidentiality.
+
 >  - In the future work section, the authors mentions several types of non-stack control flow that are not addressed by the current model, including tail calls.
@@
 I understand that these non-stack control flow is more difficult to model, but for the exact reasons, how to formally characterize stack safety for these difficult cases becomes critical.
->
+
+LEO: Not sure what the take should be here. Something like the following?
+Tail calls were left out as it's only the characterization of the return that poses
+a problem - once the point of context switch is determined, the statement of
+stack safety remains largely the same.
+
 >  [1] Xiaoyang Xu, Masoud Ghaffarinia, Wenhao Wang, Kevin W.
@@
 >   Furthermore, without stack-derived pointers, stack safety is more-or-less obvious to define.
->
+
+Leo: The sketch of stack-derived pointers in 1199 seems fairly intuitive to me. Double down on it?
+
 > - The proposed notion of stack safety is too strong to be enforced by existing stack protection mechanisms.
@@
 >   But then why you didn't use the observable version as the authoritative definition?
-
-BCP: This is a silly objection, but it should be addressed clearly and
-strongly.  E.g.:
 
@@
 
-APT:  Sk2019b _define_ stack safety via a capability machine semantics that is supposed to embody stack
+Sk2019b _define_ stack safety via a capability machine semantics that is supposed to embody stack
 safety "by construction," and relating it to a standard semantics for the same machine via a full abstraction result.
@@
 machine tied to some flavor of enforcement,  but rather on more abstract notions of observable behavior.
-
-BCP: Nice.
 

### 2020-09-09 rebuttal.txt
More responses

@@
 Why not promise the latter (and lay out what we would do)?
+LEO: Going with "we'll limit our claims and add the related work comparison"
+seems more doable (and fair) in a revision round than promising to do
+new work (or work that we started doing and then it was clear we couldn't
+work for the deadline).
 
@@
 fairly intuitive. The lazy/observable properties, are less intuitive,
-but more interesting in the violations they permit. The WBCF part simply
-formalizes standard behavior of stack-based control flow.
-
-BCP: "more interesting in the violations they permit"???
+but more interesting in that they precisely characterize violations to
+the more intuitive view of stack safety that can still be considered
+safe to an external observer. The WBCF part simply formalizes standard
+behavior of stack-based control flow.
 
@@
 
-LEO: I'm REALLY not sure about this:
 APT: Reviewer C interprets (4) to mean:
@@
 we support -- which is the business of (5)).  Can we claim that (informally, of course)?
-(4) complete: Subject to the limitations of the formalization, our
-properties should apply to any reasonable program.
+
+(4) complete: This criterion seems a bit circular. It asks for being
+able to prove any property of a program that relies on stack safety
+with a given characterization of stack safety - but we're in the
+buisness of _defining_ stack safety in the first place. If the point
+is that our definition should capture characteristics of established
+informal notions of stack safety (subject to the limitations of the
+language features our formalization supports - see 5), our properties
+should apply to any such reasonable program.
 
@@
 BCP: Obviously, this has to be addressed strongly and concretely.
+LEO: I'm not really following the CompCert definition. Has anybody taking a crack at this?
 
@@
 > * provide a clear, maximally objective/neutral statement of the benefits of their approach w.r.t. Skorstengaard et al. 2019b
->
+
+We agree that there should be a much more thorough comparison with Sk2019a/b. See the summary above for more details on what we plan to include.
+
 > The authors do not discuss any attacker model, both for their properties and for their enforcement technique.
@@
 > Is the attacker capable of running arbitrary assembly code (like in Skorstengaard et al, 2019ab) or just inject stack modifications like in Abadi et al., 2005?
->
+
+See ATTACKER MODEL in the summary above.
+
 > Another odd aspect of this paper is the fact that it doesn't propose one but three increasingly strong properties that characterize stack safety (see l86-96).
 > What does this mean: why do we want stack safety to mean different things and which meaning is appropriate in which setting?
+
+We view these stack safety properties as a spectrum. The main property
+strives to capture an intuitive notion of stack safety, by relying on
+abstract notions of observable behavior.  The stronger, lockstep
+enforces a more conservative notion of stack safety, where almost any
+digression from a canonical execution is considered unsafe. On the
+other hand, the weaker lazy variant pushes the limit of the standard
+intuition by allowing executions that would otherwise be unsafe, as
+long as their effects are not actually observable. We think all
+three are interesting in their own right and reveal interesting trade
+offs in terms of enforcement or reasoning efficiency.
+
 > Do the three variants score equally well on the above criteria?
 > - For the "lockstep" variant, I get the impression that this is a proof-technical definition that is only used for proving the regularly stack-safety property for the micropolicy enforcement and not really important/useful in itself: is this also how you see it?
+
+The lockstep variant is a slightly more technical artifact that is (1)
+easy to enforce and (2) easy to test, but that is interesting in its
+own right because it captures a very conservative notion of stack
+safety---while still allowing the majority of reasonable stack-safe
+programs.
+
 > - The lazy variant of stack safety strikes me as problematic because of several reasons:
@@
 (https://ieeexplore.ieee.org/abstract/document/7163052?casa_token=0ycUzVYD0FoAAAAA:bi8eLFwzqvCnRjtcAWD-y9xG473kJeLQdZaLWEbU43cpAwvGldBu840sXWtcC2mnOnOqGsUPruc).
->
+
+Similarly, the lazy variant is interesting because it is a property
+that is as liberal as possible in the behaviors it allows. Compared to
+the other properties, it's much harder to reason about or test (as we
+have to consider the entire execution trace), but it's simpler to enforce
+and more permissive.
+
 > I'm also a bit unconvinced about the paper's motivation.
@@
 > To this end, the authors should be open about this and remove statements that suggest their goal is different, for example l29-31 "there is still room to craft a more abstract characterization of stack safety that is independent of capabilities or any other specific enforcement technique".
->
->
+
+See MOTIVATION in the summary above.
+
 > General comments:

### 2020-09-09 rebuttal.txt
More detailed answers to C

@@
 
-LEO: Not sure what the take should be here. Something like the following?
-Tail calls were left out as it's only the characterization of the return that poses
-a problem - once the point of context switch is determined, the statement of
-stack safety remains largely the same.
+Tail calls were left out as it's only the characterization of the
+return that poses a problem - once the point of context switch is
+determined, the statement of stack safety remains largely the same.
 
@@
 >   Do you think it is etc.
->
+
+See the full comparison in terms of the criteria in the summary.
+
 > * Do you think you could show safety of a specific example program that depends on stack safety (like the awkward example) and if so, how would such a proof work?
->
+
+LEO: This is the (4) criterion I guess again. I'm not sure exactly what this means.
+
 > * Do you think your approach to defining stack safety could be extended to languages with the following features (and if so, how?):
 >   - varargs?  (seems problematic for call map)
+
+Yes. The call map serves two purposes: identifying call points and
+allowing for an easy calculation of the stack pointer at the time of
+the call. We could keep the call map essentially as a set and use a
+frame pointer (or a similar abstraction) for the latter functionality.
+Keeping the number of arguments in the map is certainly convenient,
+but not necessary.
+
 >   - dynamic code generation (seems problematic for the function map, call map and return map)
@@
 >   - self-editing code (seems problematic for call map and return map)
+
+LEO: Polish needed here.
+These three extensions would indeed be challenging. We could
+potentially handle generated code by using special contours for the
+instructions that generate it, which would correspond to a map that
+allows some instructions to update the maps, including potentially
+the map itself.
+
 >   - tail calls (seems potentially okay, but I'd like to be sure)
+
+Yes. The only minor problem with tail calls is characterizing the
+return - once the point of context switch is determined, the statement
+of stack safety remains largely the sam.
+
 >   - exceptions and stack unwinding
+
+LEO: Hmmm. I don't remember ever discussing this.
+
 >   - closures (seems potentially problematic for precision of the entry map)
+
+See below.
+
 >   How would you extend your approach to define stack safety in the presence of these features?
@@
 
-SNA:
-  Adding closures requires:
-  - identifying code points where closures are created
-  - adding machinery to track association of closure identities, entry points, and components
+Adding closures would require: 
+(1) identifying code points where closures are created
+(2) adding machinery to track association of closure identities, entry points, and components
     containing valid closure pointers (including pointer provenance, as with stack-derived pointers)
-  - extending WBCF to require that closure invocations respect the closure's designated entry point
+(3) extending WBCF to require that closure invocations respect the closure's designated entry point
     and use a valid closure pointer (ensuring temporal safety and encapsulation)
-  - extending contours to treat some components as sealed, associated with closure identities
-  - on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
+(4) extending contours to treat some components as sealed, associated with closure identities
+(5) on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
     (allowing access to closure)
-  - otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
+(6) otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
     nested calls (monotonicity is lost)
 
-  The answers are yes, we believe that with addition of closures, natural extensions to WBCF and LSE would imply these
-  properties.
+So to sum up, the answers are yes, we believe that to add closures,
+natural extensions to WBCF and LSE would imply the properties
+mentioned.
 
 > * Are the lazy and strong variant of your properties specifically intended for the corresponding micropolicy-based enforcement techniques or do you believe they are more broadly applicable?
+
+As discussed in the summary above and in your detailed comments, we
+did not intend for the properties to be micropolicy specific, but a
+spectrum of stack safety properties presenting an interesting
+trade-off in terms of enforcement or reasoning efficiency. But we
+agree that the current exposition does not convey what we set out to
+do, and that the overall motivation of the included work would be
+better portrayed as a formal characterization of micropolicy-enforced
+stack safety.

### 2020-09-09 rebuttal.txt
Some thoughts on simpler modifications, probably about to get overwritten

--- 2020-09-09 rebuttal.txt
@@
 >   - varargs?  (seems problematic for call map)
+SNA:
+
+  Varargs can easily be implemented by establishing a more concrete calling convention in which
+  the number of arguments for a call is apparent by examining the call state, at the cost of the
+  convention-agnosticism of the call map. Alternatively the call map can take the entire state as
+  an input and once again abstract over calling conventions. Once we can deduce the number of
+  arguments of each call, our properties would work without alteration.
+
 >   - dynamic code generation (seems problematic for the function map, call map and return map)
@@
 >   - tail calls (seems potentially okay, but I'd like to be sure)
+SNA:
+
+  A tail call in terms of our properties would be a call for which confidentiality and integrity hold,
+  but which is selectively exempted from Return Integrity. From the (tail) callees' perspective, it is
+  permitted to return anywhere. But the return state for the callee is also that of its caller, and so
+  the nested nature of the property means that if the caller obeys Return Integrity, then the callee
+  returns to the correct place.
+
 >   - exceptions and stack unwinding
@@
   Adding closures requires:
-  - identifying code points where closures are created
+  - identifying code points where closures are created (i.e., adding another map)
   - adding machinery to track association of closure identities, entry points, and components
-    containing valid closure pointers (including pointer provenance, as with stack-derived pointers)
+    containing valid closure pointers (via pointer provenance, as with both stack-derived and heao pointers)
   - extending WBCF to require that closure invocations respect the closure's designated entry point
-    and use a valid closure pointer (ensuring temporal safety and encapsulation)
+    and use a valid closure pointer (ensuring temporal safety)
   - extending contours to treat some components as sealed, associated with closure identities
   - on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
-    (allowing access to closure)
+    (together with above, ensuring encapsulation)
   - otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
@@
   The answers are yes, we believe that with addition of closures, natural extensions to WBCF and LSE would imply these
-  properties.
+  properties. The addition of closures is very non-trivial but plausible.
 

### 2020-09-09 rebuttal.txt
Comments / polish

--- 2020-09-09 rebuttal.txt
@@
-We thank the reviewers and in particular reviewer C for their thoughtful
-comments.  We first discuss the most important points raised in
-reviews. Then, in an optional appendix, we respond point-by-point to each
+Thank you for the thoughtful comments.  We first discuss the most important
+points raised, then, in an optional appendix, respond point-by-point to each
 individual piece of reviewer feedback.
@@
 described as an abstract property capturing micropolicy enforcement rather
-than the all-encompassing stack safety notion we set out to state, and
-intend to do so for a final version.
-
-BCP: I.e., we plan to limit our claims rather than fixing the exposition?
-Why not promise the latter (and lay out what we would do)?
-LEO: Going with "we'll limit our claims and add the related work comparison"
-seems more doable (and fair) in a revision round than promising to do
-new work (or work that we started doing and then it was clear we couldn't
-work for the deadline).
+than the all-encompassing stack safety notion we set out to state; we will
+clarify this for a final version.
 
@@
 
-All reviewers also wanted additional discussion of related work.  We agree
-that there should be a clearer comparison to both Skorstengaard papers,
-using the informal criteria proposed there. We plan to include a discussion
-like the following:
+All reviewers also wanted additional discussion of related work.  We agree,
+in particular, that there should be a clearer comparison to the two
+Skorstengaard et al papers, using the five informal criteria proposed
+there. We plan to include a discussion like the following:
+
+BCP: Can we format these so that it is clear which bits we're
+getting (verbatim, even) from Skorstengaard and which are us?
+
+BCP: Spell out abbreviations like LSE.  In general, we should not assume
+reviewers remember all the details of the paper (much less the details of
+Skorstengaard)!
 
@@
 integrity/confidentiality/noninterference, the eager version should be
-fairly intuitive. The lazy/observable properties, are less intuitive,
-but more interesting in that they precisely characterize violations to
-the more intuitive view of stack safety that can still be considered
-safe to an external observer. The WBCF part simply formalizes standard
-behavior of stack-based control flow.
+fairly intuitive. The lazy/observable properties, though less intuitive, are
+more interesting in that they precisely characterize behaviors that would be
+considered violations to the more intuitive view of stack safety but that
+can still be considered safe by an external observer. The WBCF part simply
+formalizes standard behavior of stack-based control flow.
 
@@
 
+BCP: I'm not sure this is hitting the right point.  Clearly, to reason about
+programs with both stack and heap, we need something like that frame rule.
+But the reviewer might wonder about how our property can be used for
+_reasoning_ about programs with just a stack...
+
 (3) reusable in compiler chains: As you point out, this is not really
@@
 environments" with potentially unsafe commodity compilers.
+BCP: "you" is just one reviewer, but this is to be read by all three...
 
@@
 
+BCP: Don't understand.  Are we saying tail calls are easy?  (Then why didn't
+we just add them?)
+
 >  [1] Xiaoyang Xu, Masoud Ghaffarinia, Wenhao Wang, Kevin W.
@@
 
+BCP: Yes.
+
 > - The proposed notion of stack safety is too strong to be enforced by existing stack protection mechanisms.
@@
 LEO: I'm not really following the CompCert definition. Has anybody taking a crack at this?
+
+There is some relation.  E.g., the ec_readonly condition
+
+  https://github.com/AbsInt/CompCert/blob/94558ecb3e48261f12c644045d40c7d0784415e0/common/Events.v#L651
+
+looks related.  But CompCert's definition (a) doesn't protect the stack at
+all from _reading_, and (b) isn't "nested" the way ours is.  This makes
+sense, since C doesn't protect the stack!
 
@@
 
-The lockstep variant is a slightly more technical artifact that is (1)
-easy to enforce and (2) easy to test, but that is interesting in its
-own right because it captures a very conservative notion of stack
-safety---while still allowing the majority of reasonable stack-safe
-programs.
+The lockstep variant is a slightly more technical artifact that is (1) easy
+to enforce and (2) easy to test, but that is also interesting in its own
+right because it captures a quite conservative notion of stack safety (while
+still allowing the majority of reasonable stack-safe programs).
 
@@
 
+BCP: I think this is a more focused question: Suppose you had a program that
+you believed was correct according to this criterion and you wanted to prove
+that it was.  How would the proof look?
+
 > * Do you think your approach to defining stack safety could be extended to languages with the following features (and if so, how?):
@@
 
-Yes. The call map serves two purposes: identifying call points and
-allowing for an easy calculation of the stack pointer at the time of
-the call. We could keep the call map essentially as a set and use a
-frame pointer (or a similar abstraction) for the latter functionality.
-Keeping the number of arguments in the map is certainly convenient,
-but not necessary.
+We believe this extension is possible.
+
+The call map serves two purposes: identifying call points and allowing for
+an easy calculation of the stack pointer at the time of the call. We could
+keep the call map essentially as a set and use a frame pointer (or a similar
+abstraction) for the latter functionality.  Keeping the number of arguments
+in the map is certainly convenient, but not necessary.
 
@@
 
+BCP: I don't actually believe the second sentence.
+
 >   - tail calls (seems potentially okay, but I'd like to be sure)
 
-Yes. The only minor problem with tail calls is characterizing the
-return - once the point of context switch is determined, the statement
-of stack safety remains largely the sam.
+Yes. The only minor problem with tail calls is characterizing the return;
+once the point of context switch is determined, the statement of stack
+safety remains largely the sam.
 
@@
 
-As discussed in the summary above and in your detailed comments, we
-did not intend for the properties to be micropolicy specific, but a
-spectrum of stack safety properties presenting an interesting
-trade-off in terms of enforcement or reasoning efficiency. But we
-agree that the current exposition does not convey what we set out to
-do, and that the overall motivation of the included work would be
-better portrayed as a formal characterization of micropolicy-enforced
-stack safety.
+As discussed in the summary above and in your detailed comments, we did not
+intend for the properties to be micropolicy specific, but rather to
+represent a spectrum of stack safety properties presenting an interesting
+trade-off in terms of enforcement or reasoning efficiency.
+
+But we agree that the current exposition does not convey this generality,
+and that the overall motivation of the work would be better portrayed as a
+formal characterization of micropolicy-enforced stack safety.

### 2020-09-09 rebuttal.txt

--- 2020-09-09 rebuttal.txt
@@
 return - once the point of context switch is determined, the statement
-of stack safety remains largely the sam.
+of stack safety remains largely the same. Notably a tail call's return
+point is shared with its caller, so if the (non-tail-call) caller obeys
+Return Integrity, the expected control flow of the tail call is already
+captured.
 
@@
 (3) extending WBCF to require that closure invocations respect the closure's designated entry point
-    and use a valid closure pointer (ensuring temporal safety and encapsulation)
+    and use a valid closure pointer (ensuring temporal safety)
 (4) extending contours to treat some components as sealed, associated with closure identities
 (5) on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
-    (allowing access to closure)
+    (together with the above, ensuring encapsulation)
 (6) otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during

### 2020-09-09 rebuttal.txt
attempt at answer about CompCert

@@
 Leo: The sketch of stack-derived pointers in 1199 seems fairly intuitive to me. Double down on it?
+APT: Does Sk2019b support stack-derived pointers? I don't think so, although I'm not certain.
 
@@
 LEO: I'm not really following the CompCert definition. Has anybody taking a crack at this?
+APT: The CompCert specification cited here characterizes the assumptions that the comiler
+correctness proof makes about the behavior of external functions.  
+When taken together with other parts of the CompCert spec that
+describe the block memory model and the treatment of control stacks,
+this does imply some facts about the stack safety conditions that CompCert
+assumes are provided by the external environment. But it is certainly not intended
+as a general characterization of stack safety, and it does not seem suitable
+as a basis for checking validity of an enforcement mechanism.
 

### 2020-09-10 rebuttal.txt
note on criterion 2

--- 2020-09-09 rebuttal.txt
@@
 LEO: This is the (4) criterion I guess again. I'm not sure exactly what this means.
+APT: I agree with SNA that this is the (2) criterion. Maybe we could doctor up the example program to
+include some external call to an adversary and sketch how it we would use it to prove 
+that it still prints 0, 60 (or whatever).
 
@@
 
-Yes. The only minor problem with tail calls is characterizing the return;
-once the point of context switch is determined, the statement of stack
-safety remains largely the sam.
+Yes. The only (minor) problem with tail calls is characterizing the
+return - once the point of context switch is determined, the statement
+of stack safety remains largely the same. Notably a tail call's return
+point is shared with its caller, so if the (non-tail-call) caller obeys
+Return Integrity, the expected control flow of the tail call is already
+captured.
 

### 2020-09-10 rebuttal.txt
Response/polish

@@
 
-BCP: Can we format these so that it is clear which bits we're
-getting (verbatim, even) from Skorstengaard and which are us?
+(1) Intuitiveness:
 
-BCP: Spell out abbreviations like LSE.  In general, we should not assume
-reviewers remember all the details of the paper (much less the details of
-Skorstengaard)!
+The local state encapsulation (LSE) part of the
+property we present aims to characterize stack safety in security
+terms. For readers familiar with integrity / confidentiality /
+noninterference, the eager version should be fairly intuitive. The
+lazy/observable properties, though less intuitive, are more
+interesting in that they precisely characterize behaviors that would
+be considered violations to the more intuitive view of stack safety
+but that can still be considered safe by an external observer. The
+WBCF part simply formalizes standard behavior of stack-based control
+flow.
 
-(1) On intuition: The LSE part of the property we present aims to
-characterize stack safety in security terms. For readers familiar with
-integrity/confidentiality/noninterference, the eager version should be
-fairly intuitive. The lazy/observable properties, though less intuitive, are
-more interesting in that they precisely characterize behaviors that would be
-considered violations to the more intuitive view of stack safety but that
-can still be considered safe by an external observer. The WBCF part simply
-formalizes standard behavior of stack-based control flow.
-
-By contrast, Skorstengaard relies on different prior intuitions:
-for readers familiar with the behavior of pure stacks (or rather the
+By contrast, Skorstengaard relies on different prior intuitions: for
+readers familiar with the behavior of pure stacks (or rather the
 built-in oLCM stack), their property is more intuitive. For readers
@@
 
-(2) useful for reasoning: In order to reason about actual programs we
+(2) Useful for reasoning:
+
+In order to reason about actual programs we
 would indeed need to incorporate a "Meaning of Memory Safety"-style
@@
 _reasoning_ about programs with just a stack...
+LEO: Isn't that the point of (4)? These criteria confuse me.
 
-(3) reusable in compiler chains: As you point out, this is not really
-applicable, since this is a criterion tailored to what Skorstengaard
-is aiming to do. We might consider what the equivalently tailored
-criterion would be for us -- usable in "dynamic enforcement
-environments" with potentially unsafe commodity compilers.
-BCP: "you" is just one reviewer, but this is to be read by all three...
+(3) reusable in compiler chains:
 
-APT: Reviewer C interprets (4) to mean:
-"This criterion asks for an argument that any property of a program
-that relies on stack safety can be proven with the characterization of stack safety."
-Since we're in the business of _defining_ stack safety, this seems pretty circular to me!
-Maybe the point is that we should not omit any characteristic of
-everyone's informal notion of stack safety (for the set of language features
-we support -- which is the business of (5)).  Can we claim that (informally, of course)?
+As reviewer C points out, this is not really applicable, since this is
+a criterion tailored to what Skorstengaard is aiming to do. We might
+consider what the equivalently tailored criterion would be for us --
+usable in "dynamic enforcement environments" with potentially unsafe
+commodity compilers.
 
-(4) complete: This criterion seems a bit circular. It asks for being
+(4) complete:
+
+This criterion seems a bit circular. It asks for being
 able to prove any property of a program that relies on stack safety
@@
 
-(5) scalable: Dynamic code would pose a significant challenge in this
-setup. In the tag world, we could handle generated code by tagging the
-instructions that generate it, which would correspond to a map that
-allows some instructions to update the maps, including potentially
-itself.  On the other hand, we described how we would handle
-coroutines, and threads would be similar. Basically when we lose the
-monotonicity of contours relative to stack depth, we will
-need to change the LSE property to check integrity and confidentiality
-each time we cross between security contexts - nested calls are not
-strictly more limited.
+(5) scalable:
+
+Dynamic code would pose a significant challenge in this setup. In the
+tag world, we could handle generated code by tagging the instructions
+that generate it, which would correspond to a map that allows some
+instructions to update the maps, including potentially itself.  On the
+other hand, we described how we would handle coroutines, and threads
+would be similar. Basically when we lose the monotonicity of contours
+relative to stack depth, we will need to change the LSE property to
+check integrity and confidentiality each time we cross between
+security contexts - nested calls are not strictly more limited.
 
@@
 we just add them?)
+LEO: It's much less clean in terms of the formalization and I thought
+we had decided it wasn't worth the trouble - a sketch should be enough.
 
@@
 
-BCP: Obviously, this has to be addressed strongly and concretely.
-LEO: I'm not really following the CompCert definition. Has anybody taking a crack at this?
-APT: The CompCert specification cited here characterizes the assumptions that the comiler
-correctness proof makes about the behavior of external functions.  
-When taken together with other parts of the CompCert spec that
-describe the block memory model and the treatment of control stacks,
-this does imply some facts about the stack safety conditions that CompCert
-assumes are provided by the external environment. But it is certainly not intended
-as a general characterization of stack safety, and it does not seem suitable
-as a basis for checking validity of an enforcement mechanism.
+The CompCert specification cited here characterizes the assumptions
+that the comiler correctness proof makes about the behavior of
+external functions.  When taken together with other parts of the
+CompCert spec that describe the block memory model and the treatment
+of control stacks, this does imply some facts about the stack safety
+conditions that CompCert assumes are provided by the external
+environment. But it is certainly not intended as a general
+characterization of stack safety, and it does not seem suitable as a
+basis for checking validity of an enforcement mechanism.
 
@@
 sense, since C doesn't protect the stack!
+
+Thanks for the pointer, we will make sure to include a discussion like this
+in the revised version.
 

### 2020-09-10 rebuttal.txt
Polish

@@
 motivated. Our original goal in this work wasn't to formalize a stack safety
-notion that is specific to micropolicies, but in hindsight we agree that the
-result may appear to be tailored to that particular enforcement mechanism,
-though we believe this is more a matter of presentation choices and
-exposition-helping simplifications than a fundamental limitation. In
-particular, we believe that the enforcement model can be used (with some
-effort) to describe capability-based enforcement mechanisms, as we state in
-Section 2.  However, given that this effort didn't make it into the final
-product, we agree that the paper as it stands would be more accurately
-described as an abstract property capturing micropolicy enforcement rather
-than the all-encompassing stack safety notion we set out to state; we will
-clarify this for a final version.
+notion that is specific to micropolicies. In particular, we believe that the
+enforcement model can be used (with some effort) to describe
+capability-based enforcement mechanisms, as we state in Section 2.  In
+retrospect we agree that our results may appear to be tailored to that
+particular enforcement mechanism.
+
+We believe this is more a matter of presentation choices and
+exposition-helping simplifications than a fundamental limitation.  However,
+given that a satisfying treatment of capability-based enforcement didn't
+make it into the final product, we agree that the paper as it stands would
+be more accurately described as an abstract property capturing micropolicy
+enforcement rather than the all-encompassing stack safety notion we set out
+to state; we will clarify this for a final version.
 
@@
 
-The local state encapsulation (LSE) part of the
-property we present aims to characterize stack safety in security
-terms. For readers familiar with integrity / confidentiality /
-noninterference, the eager version should be fairly intuitive. The
-lazy/observable properties, though less intuitive, are more
-interesting in that they precisely characterize behaviors that would
-be considered violations to the more intuitive view of stack safety
-but that can still be considered safe by an external observer. The
-WBCF part simply formalizes standard behavior of stack-based control
-flow.
+The local state encapsulation (LSE) part of the property we present aims to
+characterize stack safety in security terms. For readers familiar with
+integrity / confidentiality / noninterference, the eager version should be
+fairly intuitive. The lazy/observable properties, though less intuitive, are
+more interesting in that they precisely characterize behaviors that would be
+considered violations to the more intuitive view of stack safety but that
+can still be considered safe by an external observer. The WBCF part simply
+formalizes standard behavior of stack-based control flow.
 
@@
 
-(2) Useful for reasoning:
+(2) Usefulness for reasoning:
 
@@
 LEO: Isn't that the point of (4)? These criteria confuse me.
+BCP: Maybe 2 and 4 go together?  2 is about whether you can use it for
+reasoning; 4 is about whether this reasoning is complete in the sense that
+the proof principle is strong enough to prove safety of any safe program.
 
-(3) reusable in compiler chains:
+(3) Reusability in compiler chains:
 
-As reviewer C points out, this is not really applicable, since this is
-a criterion tailored to what Skorstengaard is aiming to do. We might
-consider what the equivalently tailored criterion would be for us --
-usable in "dynamic enforcement environments" with potentially unsafe
-commodity compilers.
+As reviewer C points out, this is not really applicable, since this is a
+criterion tailored to what Skorstengaard is aiming to do. We might consider
+what the equivalently tailored criterion would be for us -- usable in
+"dynamic enforcement environments" with potentially unsafe commodity
+compilers.
 
-(4) complete:
+(4) Completeness:
 
-This criterion seems a bit circular. It asks for being
-able to prove any property of a program that relies on stack safety
-with a given characterization of stack safety - but we're in the
-buisness of _defining_ stack safety in the first place. If the point
-is that our definition should capture characteristics of established
-informal notions of stack safety (subject to the limitations of the
-language features our formalization supports - see 5), our properties
-should apply to any such reasonable program.
+This criterion seems a bit circular. It asks for being able to prove any
+property of a program that relies on stack safety with a given
+characterization of stack safety - but we're in the buisness of _defining_
+stack safety in the first place. If the point is that our definition should
+capture characteristics of established informal notions of stack
+safety (subject to the limitations of the language features our
+formalization supports - see 5), our properties should apply to any such
+reasonable program.
 
-(5) scalable:
+(5) Scalability:
 
-Dynamic code would pose a significant challenge in this setup. In the
-tag world, we could handle generated code by tagging the instructions
-that generate it, which would correspond to a map that allows some
-instructions to update the maps, including potentially itself.  On the
-other hand, we described how we would handle coroutines, and threads
-would be similar. Basically when we lose the monotonicity of contours
-relative to stack depth, we will need to change the LSE property to
+Dynamic code would pose a significant challenge in this setup. In the tag
+world, we could handle generated code by tagging the instructions that
+generate it, which would correspond to a map that allows some instructions
+to update the maps, including potentially itself.  On the other hand, we
+described how we would handle coroutines, and threads would be
+similar. Basically when we lose the monotonicity of contours relative to
+stack depth, we will need to change the LSE property to
 check integrity and confidentiality each time we cross between
@@
 the locations in the "same" stack should be unchanged. Similarly for
-cofnidentiality.
+confidentiality.
 
@@
 
-Tail calls were left out as it's only the characterization of the
+We omitted tail calls as it's really only the characterization of the
 return that poses a problem - once the point of context switch is
-determined, the statement of stack safety remains largely the same.
-
-BCP: Don't understand.  Are we saying tail calls are easy?  (Then why didn't
-we just add them?)
-LEO: It's much less clean in terms of the formalization and I thought
-we had decided it wasn't worth the trouble - a sketch should be enough.
+determined, the statement of stack safety remains largely the same.  We will
+add a more detailed discussion.
 
@@
 APT: Does Sk2019b support stack-derived pointers? I don't think so, although I'm not certain.
-
-BCP: Yes.
 

### 2020-09-10 rebuttal.txt
another round

--- 2020-09-10 rebuttal.txt
@@
 but that can still be considered safe by an external observer. The
-WBCF part simply formalizes standard behavior of stack-based control
-flow.
+well-bracketed control flow (WBCF) part simply formalizes standard
+behavior of stack-based control flow.
 
-By contrast, Skorstengaard relies on different prior intuitions: for
-readers familiar with the behavior of pure stacks (or rather the
+By contrast, [Skorstengaard 2019b] relies on different prior intuitions: for
+readers familiar with operational models of stacks (here the
 built-in oLCM stack), their property is more intuitive. For readers
-versed in security, we believe there is value in a different
-perspective.
+versed in security theory, we believe that making an explicit connection
+with integrity and confidentiality gives useful insight.
 
@@
 In order to reason about actual programs we
-would indeed need to incorporate a "Meaning of Memory Safety"-style
+would indeed need to incorporate a frame ruleMeaning of Memory Safety"-style
 frame rule. Our goal with this paper was to distill the essence of
@@
 dealing with heap locations. But the properties (and the testing
-framework behind it) were designed with [Azevedo et al] in mind.
+framework behind it) were designed with [Azevedo et al 2018] in mind.
 
@@
 LEO: Isn't that the point of (4)? These criteria confuse me.
+APT: My take is that (2) is about what kind of programming logic
+we can support, whereas (4) is about arguing that our definition
+captures all the standard intuitions one might want to use in proofs.
+I think the heap is a red herring here.
 
@@
 
+APT: I don't understand the last sentence.
+
 (4) complete:
@@
 informal notions of stack safety (subject to the limitations of the
-language features our formalization supports - see 5), our properties
-should apply to any such reasonable program.
+language features our formalization supports - see 5), then we believe
+it does.
 
@@
 security contexts - nested calls are not strictly more limited.
+
+APT: shouldn't we mention stack-derived pointers too?
 

### 2020-09-10 rebuttal.txt

--- 2020-09-10 rebuttal.txt
@@
 
-By contrast, Skorstengaard relies on different prior intuitions: for
-readers familiar with the behavior of pure stacks (or rather the
+By contrast, [Skorstengaard 2019b] relies on different prior intuitions: for
+readers familiar with operational models of stacks (here the
 built-in oLCM stack), their property is more intuitive. For readers
-versed in security, we believe there is value in a different
-perspective.
+versed in security theory, we believe that making an explicit connection
+with integrity and confidentiality gives useful insight.
 
@@
 LEO: Isn't that the point of (4)? These criteria confuse me.
+APT: My take is that (2) is about what kind of programming logic
+we can support, whereas (4) is about arguing that our definition
+captures all the standard intuitions one might want to use in proofs.
+I think the heap is a red herring here.
 BCP: Maybe 2 and 4 go together?  2 is about whether you can use it for
@@
 
-(3) Reusability in compiler chains:
+(3) reusable in compiler chains:
 
-As reviewer C points out, this is not really applicable, since this is a
-criterion tailored to what Skorstengaard is aiming to do. We might consider
-what the equivalently tailored criterion would be for us -- usable in
-"dynamic enforcement environments" with potentially unsafe commodity
-compilers.
+As reviewer C points out, this is not really applicable, since this is
+a criterion tailored to what Skorstengaard is aiming to do. We might
+consider what the equivalently tailored criterion would be for us --
+usable in "dynamic enforcement environments" with potentially unsafe
+commodity compilers.
 
-(4) Completeness:
+APT: I don't understand the last sentence.
 
-This criterion seems a bit circular. It asks for being able to prove any
-property of a program that relies on stack safety with a given
-characterization of stack safety - but we're in the buisness of _defining_
-stack safety in the first place. If the point is that our definition should
-capture characteristics of established informal notions of stack
-safety (subject to the limitations of the language features our
-formalization supports - see 5), our properties should apply to any such
-reasonable program.
+(4) complete:
 
-(5) Scalability:
+This criterion seems a bit circular. It asks for being
+able to prove any property of a program that relies on stack safety
+with a given characterization of stack safety - but we're in the
+buisness of _defining_ stack safety in the first place. If the point
+is that our definition should capture characteristics of established
+informal notions of stack safety (subject to the limitations of the
+language features our formalization supports - see 5), then we believe
+it does.
 
-Dynamic code would pose a significant challenge in this setup. In the tag
-world, we could handle generated code by tagging the instructions that
-generate it, which would correspond to a map that allows some instructions
-to update the maps, including potentially itself.  On the other hand, we
-described how we would handle coroutines, and threads would be
-similar. Basically when we lose the monotonicity of contours relative to
-stack depth, we will need to change the LSE property to
+(5) scalable:
+
+Dynamic code would pose a significant challenge in this setup. In the
+tag world, we could handle generated code by tagging the instructions
+that generate it, which would correspond to a map that allows some
+instructions to update the maps, including potentially itself.  On the
+other hand, we described how we would handle coroutines, and threads
+would be similar. Basically when we lose the monotonicity of contours
+relative to stack depth, we will need to change the LSE property to
 check integrity and confidentiality each time we cross between
 security contexts - nested calls are not strictly more limited.
+
+APT: shouldn't we mention stack-derived pointers too?
 

### 2020-09-10 rebuttal.txt
Tiny comment
Another crack.

@@
 
-Leo: The sketch of stack-derived pointers in 1199 seems fairly intuitive to me. Double down on it?
+Stack-derived pointers are not supported yet, but based on the sketch in future work (L1199+)
+we feel it is relatively straightforward combination of known techniques (on top of the
+definition of the properties we introduced).
+
 APT: Does Sk2019b support stack-derived pointers? I don't think so, although I'm not certain.
+LEO: This reviewer doesn't complain about Sk2019b - maybe not make reviewer C look at this even more closely by invoking that?
 
@@
 
-LEO: Polish needed here.
 These three extensions would indeed be challenging. We could
@@
 instructions that generate it, which would correspond to a map that
-allows some instructions to update the maps, including potentially
-the map itself.
+allows some instructions to update the maps, including potentially the
+map itself.
 
@@
 >   - tail calls (seems potentially okay, but I'd like to be sure)
-
-Yes. The only (minor) problem with tail calls is characterizing the
-return - once the point of context switch is determined, the statement
-of stack safety remains largely the same. Notably a tail call's return
-point is shared with its caller, so if the (non-tail-call) caller obeys
-Return Integrity, the expected control flow of the tail call is already
-captured.
-
 >   - exceptions and stack unwinding
 
-LEO: Hmmm. I don't remember ever discussing this.
+We believe so. The (minor) problem with tail calls is characterizing
+the return - once the point of context switch is determined, the
+statement of stack safety remains largely the same. Notably a tail
+call's return point is shared with its caller, so if the
+(non-tail-call) caller obeys Return Integrity, the expected control
+flow of the tail call is already captured. Exceptions would be
+slightly trickier, but could potentially be handled by multiple
+allowable return points (e.g. one for the call and one for the
+exception handler).
 

### 2020-09-10 rebuttal.txt
Some updates to model extension

@@
 I think the heap is a red herring here.
-BCP: Maybe 2 and 4 go together?  2 is about whether you can use it for
+BCP: Agree
+
+BCP (OLD): Maybe 2 and 4 go together?  2 is about whether you can use it for
 reasoning; 4 is about whether this reasoning is complete in the sense that
@@
 
-APT: I don't understand the last sentence.
+SNA: trying to phrase this better.
+A more general version of this criterion might be, "Is the model suitable
+for a practical application in the engineering of secure systems?"
+In our case, such an application would be use in "dynamic enforcement
+environments" with code generated by non-secure compilation, ideally
+with commodity compilers.
 
@@
 that generate it, which would correspond to a map that allows some
-instructions to update the maps, including potentially itself.  On the
-other hand, we described how we would handle coroutines, and threads
+instructions to update the maps, including potentially itself.
+
+SNA: how about...
+
+Dynamic code would pose a significant challenge in this setting.
+Our various maps would need to be made dynamic, which would require
+either very restricted code modification or very complex machinery.
+It would also raise questions about how to handle all of our trace
+properties when (for instance) a return map might change mid-call.
+Substantial rethinking of both the machinery and the property statements
+themselves make this an unlikely feature to add.
+
+On the other hand, we described how we would handle coroutines, and threads
 would be similar. Basically when we lose the monotonicity of contours
@@
 check integrity and confidentiality each time we cross between
-security contexts - nested calls are not strictly more limited.
+security contexts - nested calls are not strictly more limited than
+their callers.
 
-APT: shouldn't we mention stack-derived pointers too?
+Loss of monotonicity, and the attendant changes in the LSE property,
+are a consistent theme among potential expansions to the model.
+Stack-derived pointers, as described in Section 9, are a natural
+extension of the model that will give up monotonicity.
 

### 2020-09-10 rebuttal.txt
wip

--- 2020-09-10 rebuttal.txt
@@
 Reviewer C is particularly concerned with the way this work is
-motivated. Our original goal in this work wasn't to formalize a stack safety
-notion that is specific to micropolicies. In particular, we believe that the
-enforcement model can be used (with some effort) to describe
-capability-based enforcement mechanisms, as we state in Section 2.  In
-retrospect we agree that our results may appear to be tailored to that
-particular enforcement mechanism.
+motivated. Our original goal in this work was to formalize a stack
+safety notion that applies to many enforcement mechanisms, not just to
+micropolicies.  In particular, we believe that the enforcement model
+can be used (with some effort) to describe capability-based
+enforcement mechanisms, as we state in Section 2.  But in retrospect
+we agree that our results may appear to be tailored to micropolicies.
 
 We believe this is more a matter of presentation choices and
-exposition-helping simplifications than a fundamental limitation.  However,
+simplifications intended to help exposition than a fundamental limitation.  However,
 given that a satisfying treatment of capability-based enforcement didn't
@@
 to state; we will clarify this for a final version.
+
+APT: Not to beat a dead horse, but this seems to completely throw away
+our claim that security theory ideas are useful for characterizing
+stack safety.  Could we incorporate that point somewhere in the above?
 
@@
 
-The local state encapsulation (LSE) part of the property we present aims to
-characterize stack safety in security terms. For readers familiar with
-integrity / confidentiality / noninterference, the eager version should be
-fairly intuitive. The lazy/observable properties, though less intuitive, are
-more interesting in that they precisely characterize behaviors that would be
-considered violations to the more intuitive view of stack safety but that
-can still be considered safe by an external observer. The WBCF part simply
-formalizes standard behavior of stack-based control flow.
+The local state encapsulation (LSE) part of the property we present
+aims to characterize stack safety in security terms. For readers
+familiar with integrity / confidentiality / noninterference, the eager
+version should be fairly intuitive. The lazy/observable properties,
+though less intuitive, are more interesting in that they precisely
+characterize behaviors that would be considered violations to the more
+intuitive view of stack safety but that can still be considered safe
+by an external observer. The well-bracketec control flow (WBCF) part
+simply formalizes standard behavior of stack-based control flow.
 
@@
 In order to reason about actual programs we
-would indeed need to incorporate a frame ruleMeaning of Memory Safety"-style
+would indeed need to incorporate a "Meaning of Memory Safety"-style
 frame rule. Our goal with this paper was to distill the essence of
@@
 As reviewer C points out, this is not really applicable, since this is
-a criterion tailored to what Skorstengaard is aiming to do. We might
+a criterion tailored to what Skorstengaard et al. are aiming to do. We might
 consider what the equivalently tailored criterion would be for us --

### 2020-09-10 rebuttal.txt
accept some SNA changes

@@
 
-SNA: trying to phrase this better.
 A more general version of this criterion might be, "Is the model suitable
@@
 
-Dynamic code would pose a significant challenge in this setup. In the
-tag world, we could handle generated code by tagging the instructions
-that generate it, which would correspond to a map that allows some
-instructions to update the maps, including potentially itself.
-
-SNA: how about...
-
 Dynamic code would pose a significant challenge in this setting.
@@
 check integrity and confidentiality each time we cross between
-security contexts - nested calls are not strictly more limited than
+security contexts - nested calls are no longer strictly more limited than
 their callers.

### 2020-09-10 rebuttal.txt
Fix pointers

--- 2020-09-10 rebuttal.txt
@@
 Reviewer C is particularly concerned with the way this work is
-motivated. Our original goal in this work was to formalize a stack
-safety notion that applies to many enforcement mechanisms, not just to
-micropolicies.  In particular, we believe that the enforcement model
-can be used (with some effort) to describe capability-based
-enforcement mechanisms, as we state in Section 2.  But in retrospect
-we agree that our results may appear to be tailored to micropolicies.
+motivated. Our original goal in this work wasn't to formalize a stack safety
+notion that is specific to micropolicies. In particular, we believe that the
+enforcement model can be used (with some effort) to describe
+capability-based enforcement mechanisms, as we state in Section 2.  In
+retrospect we agree that our results may appear to be tailored to that
+particular enforcement mechanism.
 
 We believe this is more a matter of presentation choices and
-simplifications intended to help exposition than a fundamental limitation.  However,
+exposition-helping simplifications than a fundamental limitation.  However,
 given that a satisfying treatment of capability-based enforcement didn't
@@
 to state; we will clarify this for a final version.
-
-APT: Not to beat a dead horse, but this seems to completely throw away
-our claim that security theory ideas are useful for characterizing
-stack safety.  Could we incorporate that point somewhere in the above?
 
@@
 
-The local state encapsulation (LSE) part of the property we present
-aims to characterize stack safety in security terms. For readers
-familiar with integrity / confidentiality / noninterference, the eager
-version should be fairly intuitive. The lazy/observable properties,
-though less intuitive, are more interesting in that they precisely
-characterize behaviors that would be considered violations to the more
-intuitive view of stack safety but that can still be considered safe
-by an external observer. The well-bracketec control flow (WBCF) part
-simply formalizes standard behavior of stack-based control flow.
+The local state encapsulation (LSE) part of the property we present aims to
+characterize stack safety in security terms. For readers familiar with
+integrity / confidentiality / noninterference, the eager version should be
+fairly intuitive. The lazy/observable properties, though less intuitive, are
+more interesting in that they precisely characterize behaviors that would be
+considered violations to the more intuitive view of stack safety but that
+can still be considered safe by an external observer. The WBCF part simply
+formalizes standard behavior of stack-based control flow.
 
@@
 As reviewer C points out, this is not really applicable, since this is
-a criterion tailored to what Skorstengaard et al. are aiming to do. We might
+a criterion tailored to what Skorstengaard is aiming to do. We might
 consider what the equivalently tailored criterion would be for us --
@@
 (5) scalable:
-
-Dynamic code would pose a significant challenge in this setup. In the
-tag world, we could handle generated code by tagging the instructions
-that generate it, which would correspond to a map that allows some
-instructions to update the maps, including potentially itself.
-
-SNA: how about...
 
@@
 
-LEO: This is the (4) criterion I guess again. I'm not sure exactly what this means.
-APT: I agree with SNA that this is the (2) criterion. Maybe we could doctor up the example program to
-include some external call to an adversary and sketch how it we would use it to prove
-that it still prints 0, 60 (or whatever).
-
-BCP: I think this is a more focused question: Suppose you had a program that
-you believed was correct according to this criterion and you wanted to prove
-that it was.  How would the proof look?
+See discussion of criterion (2) in the summary.
 
@@
 
-These three extensions would indeed be challenging. We could
-potentially handle generated code by using special contours for the
-instructions that generate it, which would correspond to a map that
-allows some instructions to update the maps, including potentially the
-map itself.
-
-BCP: I don't actually believe the second sentence.
+As discussed in the toplevel summary, for criterion (5), these three
+extensions would indeed be challenging.
 

### 2020-09-10 rebuttal.txt
Comment
one last contribution to the (2) story

@@
 framework behind it) were designed with [Azevedo et al 2018] in mind.
+
+APT: Leo asks: "what kind of logic do we want to claim to support?"
+I dunno, a Hoare logic with some simple procedure rule?
+Integrity says that local variables do not change over calls,
+which basically says that procedure calls are a no-op (except
+for the return value).  Confidentiality says that lemmas proved about
+procedure return values can be applied at call sites.
 
@@
 definition of the properties we introduced).
-
-APT: Does Sk2019b support stack-derived pointers? I don't think so, although I'm not certain.
-LEO: This reviewer doesn't complain about Sk2019b - maybe not make reviewer C look at this even more closely by invoking that?
 

### 2020-09-10 rebuttal.txt
Some work on 2

--- 2020-09-10 rebuttal.txt
@@
 (2) Usefulness for reasoning:
+
+Local state encapsulation enables reasoning about the behavior of a function
+in a vacuum, without needing to account for either its caller's state or
+for changes to its state during its callees. Without confidentiality,
+proving a property of a function requires a quantification over all possible
+configurations of memory. With confidentiality, only all possible configurations
+of the stack frame itself need be considered.
+
+As for integrity, our threat model permits a callee to execute arbitrary code,
+severely limiting our ability to reason about the behavior of its caller after
+return. But we can now prove properties of the caller by quantifying only over
+configurations of registers and memory above the stack. In security terms, we
+make no guarantees about the callee's behavior, but we can guarantee that it
+communicates with the caller only through the approved interface.
+
+Meanwhile, WBCF enables reasoning about the whole structure of the program
+in terms of predictable control flows between its components.
+
+SNA: I feel like it would be good to be less informal and put this in programming
+     logic terms, but I'm not familiar enough with the language. Also we didn't
+     formalize a programming logic, so I'm not sure how much we can say formally.
+
+------- old -------
 
@@
 (5) scalable:
+
+Dynamic code would pose a significant challenge in this setup. In the
+tag world, we could handle generated code by tagging the instructions
+that generate it, which would correspond to a map that allows some
+instructions to update the maps, including potentially itself.
+
+SNA: how about...
 
@@
 
-See discussion of criterion (2) in the summary.
+LEO: This is the (4) criterion I guess again. I'm not sure exactly what this means.
+APT: I agree with SNA that this is the (2) criterion. Maybe we could doctor up the example program to
+include some external call to an adversary and sketch how it we would use it to prove
+that it still prints 0, 60 (or whatever).
+
+BCP: I think this is a more focused question: Suppose you had a program that
+you believed was correct according to this criterion and you wanted to prove
+that it was.  How would the proof look?
 
@@
 
-As discussed in the toplevel summary, for criterion (5), these three
-extensions would indeed be challenging.
+These three extensions would indeed be challenging. Our static maps would need
+to be replaced with more dynamic machinery, likely becoming either very complex
+or much more concrete about what constitutes a call, return, etc.
 

### 2020-09-10 rebuttal.txt

--- 2020-09-10 rebuttal.txt
@@
 
+Local state encapsulation enables reasoning about the behavior of a function
+in a vacuum, without needing to account for either its caller's state or
+for changes to its state during its callees. Without confidentiality,
+proving a property of a function requires a quantification over all possible
+configurations of memory. With confidentiality, only all possible configurations
+of the stack frame itself need be considered.
+
+As for integrity, our threat model permits a callee to execute arbitrary code,
+severely limiting our ability to reason about the behavior of its caller after
+return. But we can now prove properties of the caller by quantifying only over
+configurations of registers and memory above the stack. In security terms, we
+make no guarantees about the callee's behavior, but we can guarantee that it
+communicates with the caller only through the approved interface.
+
+Meanwhile, WBCF enables reasoning about the whole structure of the program
+in terms of predictable control flows between its components.
+
+SNA: I feel like it would be good to be less informal and put this in programming
+     logic terms, but I'm not familiar enough with the language. Also we didn't
+     formalize a programming logic, so I'm not sure how much we can say formally.
+
+------- old -------
+
 In order to reason about actual programs we
@@
 BCP: Agree
+
+BCP: I think we should rewrite the point to simply speculate that something
+analogous to the frame rule from the "meaning of memory safety" paper could
+be developed here for the stack.
 

### 2020-09-10 rebuttal.txt
Rephrase 2 and incorporate Andrew's

@@
 
-Local state encapsulation enables reasoning about the behavior of a function
-in a vacuum, without needing to account for either its caller's state or
-for changes to its state during its callees. Without confidentiality,
-proving a property of a function requires a quantification over all possible
-configurations of memory. With confidentiality, only all possible configurations
-of the stack frame itself need be considered.
+Local state encapsulation enables reasoning about the behavior of a
+function in a vacuum, without needing to account for either its
+caller's state or for changes to its state during its callees. When
+proving a property of a function, confidentiality allows us to only
+reason about configurations of the stack frame itself (instead of all
+possible _memory_ configurations).  As for integrity, our threat model
+permits a callee to execute arbitrary code, severely limiting our
+ability to reason about the behavior of its caller after return. But
+we can now prove properties of the caller by quantifying only over
+configurations of registers and memory above the stack.
 
-As for integrity, our threat model permits a callee to execute arbitrary code,
-severely limiting our ability to reason about the behavior of its caller after
-return. But we can now prove properties of the caller by quantifying only over
-configurations of registers and memory above the stack. In security terms, we
-make no guarantees about the callee's behavior, but we can guarantee that it
-communicates with the caller only through the approved interface.
+In security terms, we make no guarantees about the callee's behavior,
+but we can guarantee that it communicates with the caller only through
+the approved interface.  In terms of a logic, integrity and
+confidentiality would allow a procedure rule that states that local
+variables do not change over calls, (which basically says that
+procedure calls are a no-op except for the return value) and that
+lemmas proved about procedure return values can be applied at call
+sites.
 
@@
 in terms of predictable control flows between its components.
-
-SNA: I feel like it would be good to be less informal and put this in programming
-     logic terms, but I'm not familiar enough with the language. Also we didn't
-     formalize a programming logic, so I'm not sure how much we can say formally.
-
-------- old -------
-
-In order to reason about actual programs we
-would indeed need to incorporate a "Meaning of Memory Safety"-style
-frame rule. Our goal with this paper was to distill the essence of
-stack safety without adding the complex (but solved) overhead of
-dealing with heap locations. But the properties (and the testing
-framework behind it) were designed with [Azevedo et al 2018] in mind.
-
-APT: Leo asks: "what kind of logic do we want to claim to support?"
-I dunno, a Hoare logic with some simple procedure rule?
-Integrity says that local variables do not change over calls,
-which basically says that procedure calls are a no-op (except
-for the return value).  Confidentiality says that lemmas proved about
-procedure return values can be applied at call sites.
-
-BCP: I'm not sure this is hitting the right point.  Clearly, to reason about
-programs with both stack and heap, we need something like that frame rule.
-But the reviewer might wonder about how our property can be used for
-_reasoning_ about programs with just a stack...
-LEO: Isn't that the point of (4)? These criteria confuse me.
-APT: My take is that (2) is about what kind of programming logic
-we can support, whereas (4) is about arguing that our definition
-captures all the standard intuitions one might want to use in proofs.
-I think the heap is a red herring here.
-BCP: Agree
-
-BCP: I think we should rewrite the point to simply speculate that something
-analogous to the frame rule from the "meaning of memory safety" paper could
-be developed here for the stack.
-
-BCP (OLD): Maybe 2 and 4 go together?  2 is about whether you can use it for
-reasoning; 4 is about whether this reasoning is complete in the sense that
-the proof principle is strong enough to prove safety of any safe program.
 

### 2020-09-10 rebuttal.txt
Add APT in motivation

@@
 We believe this is more a matter of presentation choices and
-simplifications intended to help exposition than a fundamental limitation.  However,
-given that a satisfying treatment of capability-based enforcement didn't
-make it into the final product, we agree that the paper as it stands would
-be more accurately described as an abstract property capturing micropolicy
-enforcement rather than the all-encompassing stack safety notion we set out
-to state; we will clarify this for a final version.
-
-APT: Not to beat a dead horse, but this seems to completely throw away
-our claim that security theory ideas are useful for characterizing
-stack safety.  Could we incorporate that point somewhere in the above?
+simplifications intended to help exposition than a fundamental
+limitation.  However, given that a satisfying treatment of
+capability-based enforcement didn't make it into the final product, we
+agree that the paper as it stands would be more accurately described
+as an abstract property capturing micropolicy enforcement rather than
+the all-encompassing stack safety notion we set out to state; we will
+clarify this for a final version. We still maintain that security
+theory ideas are are useful and interesting for characterizing stack
+safety, but we will restrict our claims to the micropolicy setting.
 

### 2020-09-10 rebuttal.txt
Squash
typo
Added example of the use of a procedure rule.

--- 2020-09-10 rebuttal.txt
@@
 variables do not change over calls, (which basically says that
-procedure calls are a no-op except for the return value) and that
-lemmas proved about procedure return values can be applied at call
-sites.
+procedure calls are a no-op except for observations and the return value)
+and that lemmas proved about procedure return values and observations can
+be applied at call sites.
+
+For a security-oriented example, suppose that a procedure f takes two
+arguments, x and y, passes y on to a procedure g, and outputs the result. We
+may wish to prove that x is not leaked - a noninterference proof that
+for and x, x', and y, the output trace of f(x,y) is identical to that of
+f(x',y). We would prove this easily for each step of f before the call to g;
+at the call point, we would use confidentiality to claim that since the states
+are variants of one another under the contour of g, they have the same
+observable behavior and state changes if there is a return. Integrity guarantees
+that at a return point, f's code has not changed. So, whatever value was returned
+from g, it is output in both traces.
 

### 2020-09-10 rebuttal.txt

@@
 intuitive view of stack safety but that can still be considered safe
-by an external observer. The well-bracketec control flow (WBCF) part
+by an external observer. The well-bracketed control flow (WBCF) part
 simply formalizes standard behavior of stack-based control flow.
@@
 enforcement mechanism is described in [Roessler and Dehon 18].
-
-BCP: Can we sketch it here?
 

### 2021-02-14 main.tex
Contexts in tex

@@
 
-\section{Stack Safety, Formally}
-\label{sec:lse-and-wbcf}
+\section{Simple Stack Safety, Formally}
+\label{sec:sss}
 
@@
 well-bracketed control flow, which together constitute stack safety.
+Initially we will focus on a simple stack model, with a single stack
+that doesn't share between caller and callee. All arguments are passed
+in registers.
 
@@
 Our properties and policies assume that the program is annotated
-with various information about its structure.
-
-\paragraph*{Function Map}
-
-A {\em function map} identifies some addresses as code and assigns
-each to a function. Formally, a function map is a
-partial function from addresses to a function
-identifier drawn from the set \(\FUNIDS\).
-  %
-\[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]
-%
-In our running example, instructions 1-13 correspond to {\tt main},
-instructions 20-36 correspond to {\tt f}, and instructions 40-47
-correspond to {\tt g}, which means that the function map would be:
-
-\[ \codemap(w) =
+by the compiler with information about its structure. At minimum,
+we need to know which addresses contain instructions, which instructions
+constitute calls and returns, and to which function an instruction
+belongs.
+
+This information is represented by a {\em code map}, a partial function from
+addresses to pairs consisting of a function identifier and an {\em annotation}.
+An annotation may be {\em call} or {\em return}.
+
+\[\ann \in \ANNS ::= \callmap + \retmap\]
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option{\ann}\]
+
+\paragraph*{Call Depth and Domains}
+
+At any given step of a program, some number of callers have pushed their frames
+onto the stack, ``sealing'' that memory for later use; this number is the current
+{\em call depth}. The call depth, as well as the association between stack addresses
+and sealed frames, are dynamic and cannot be deduced from the state alone, so we will
+track them over the course of an execution as additional context that informs the property.
+
+Each component will be assigned to a {\em domain}, which may be {\em Outside} of the
+stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed} by a particular
+depth.
+
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{n}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+  \times (n \in \mathbb{N}) \]
+
+An initial context \(\context_0 = \domMap_0 \times 0\) where \(\domMap_0\) maps
+the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
+Then at each step we update the context according to the context-step function,
+\(\stepsto_C\), which we will define in terms of the operations
+\(\push{\domMap}{n}{\components}\), in which \(\components \subseteq \COMPONENTS\),
+and \(\pop{\domMap}{n}\), pronounced ``push'' and ``pop'' and defined as follows.
+
+\[
+(\push{\domMap}{n}{\components}) ~ \component =
 \begin{cases}
-    \texttt{main} & \text{if } 1 \leq w \leq 13 \\
-    \texttt{f} & \text{if } 20 \leq w \leq 36 \\
-    \texttt{g} & \text{if } 40 \leq w \leq 47 \\
-    \mathit{undefined} & \text {otherwise}\\
+  \sealed{n} & \text{if } \domMap ~ \component = \unsealed
+               \text{ and } \component \in \components \\
+  \domMap ~ \component & \text{else} \\
 \end{cases}
 \]
-
-\paragraph*{Call Map}
-
-A {\em call map} identifies addresses that complete a call sequence,
-abstracting away from any concrete calling convention. Formally, it is
-a partial function from addresses to a natural number, the number of
-arguments and return values of the call.
-  %
-\[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]
-%
-In our running example, the callmap contains two entries: the word 7
-corresponding to the call to {\tt f} which has one argument and one
-return value; and the word 26 corresponding to the call to {\tt g},
-also with a single argument and return value:
-\[ \callmap = \{ 7 \mapsto 2, 26 \mapsto 2\}\]
-
-For concreteness, in a typical calling convention, the call map will
-identify the locations of the $\jal$ instructions representing calls.
-% APT: I found this more confusing than helpful...
-%, in which
-%space for arguments is already allocated, but the mechanism is more
-%general.
-A state in which a call map is defined on the program counter
-is referred to as a {\it call state}.
-
-\paragraph*{Return Map}
-Similarly, a {\em return map} identifies addresses that complete a
-return sequence. Formally, it is a set of words; each one
-is the address of the final jump corresponding to a function return:
-%
-\[\retmap \in \RETMAPS \subseteq \WORDS\]
-%
-In our running example, the return map identifies the $\jalr$
-instructions 36 and 47:
-\[ \retmap = \{36, 47\} \]
-
-\paragraph*{Entry Map}
-Finally, the entry map identifies valid entry points to a
-function and is also a set:
-\[\entmap \in \ENTMAPS \subseteq \WORDS\]
-%
-In our running example, the entry map contains the entry
-points of {\tt f} and {\tt g}:
-\[ \entmap = \{20, 40\} \]
-
-\paragraph*{What is a return?}
-
-We separately identify, for any call state, what it means to have
-correctly returned from that call. We consider the call to have
-returned when control returns to the instruction following
-the $\jal$ with the stack restored. Formally, this is a predicate on
-states that says when the machine has just returned from a specified call state
-(where $\nextPC$ gives the address of the next instruction after a given address):
-\[%\begin{split}
-    \ret{\mach_c} \triangleq  \lambda \mach_r . \mach_r(\PCname) =
-    \nextPC(\mach_c(\PCname)) \land  \mach_r(\SP) = \mach_c(\SP)
-%\end{split}
+\[
+(\pop{\domMap}{n}) ~ \component =
+\begin{cases}
+  \unsealed & \text{if } \domMap ~ \component = \sealed{n-1} \\
+  \domMap ~ \component & \text{else} \\
+\end{cases}
 \]
-There are two such return points in our running example, both of which
-are depicted in the figure: the state before the execution of
-instruction 27 corresponding to the return from {\tt f} and the state
-before instruction 8 corresponding to the return from {\tt g}.
-
-\medskip
-
-In the rest of this section we implicitly parameterize all
-definitions by a function map $\codemap$, a call map $\callmap$, a return
-map $\retmap$ and an entry map $\entmap$.
+
+Then we can define the context update relation. Let \(\sealcon\) be the sealing convention
+that maps a machine state to a set of components, typically the set of addresses below
+the stack pointer. The {\sc Default} rule applies only if neither of the others do.
+
+\begin{center}
+\begin{minipage}{.4\textwidth}
+  \judgmenttwo{\(\mach \stepsto \mach'\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
+              {\(\mach,(\domMap,n) \stepsto_C \mach',(\push{\domMap}{n}{\sealcon ~ \mach},n+1)\)}
+\end{minipage}
+\hspace*{0.1\textwidth}
+\begin{minipage}{.4\textwidth}
+  \judgmenttwo{\(\mach \stepsto \mach'\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
+              {\(\mach,(\domMap,n) \stepsto_C \mach',(\pop{\domMap}{n},n-1)\)}
+\end{minipage}
+\end{center}
+
+\begin{center}
+  \judgment[Default]
+           {\(\mach \stepsto \mach'\)}
+           {\(\mach,\context \stepsto_C \mach',\context\)}
+\end{center}
+
+\subsection{Ultra-eager Integrity}
+
+\subsection{Eager Integrity and the Call Rule}
+
+\subsection{Eager Confidentiality and the Transparency Rule}
+
+\section{Stack Safety with Sharing}
+
+\subsection{Trace Properties}
+
+\section{Coroutine Safety}
 

### 2021-02-14 rebuttal.txt

@@
 reason about configurations of the stack frame itself (instead of all
-possible _memory_ configurations).  As for integrity, our threat model
+possible memory configurations).  Moreover, while our threat model
 permits a callee to execute arbitrary code, severely limiting our
-ability to reason about the behavior of its caller after return. But
-we can now prove properties of the caller by quantifying only over
-configurations of registers and memory above the stack.
+ability to reason about the behavior of its caller after return,
+integrity allows us to prove properties of the caller by quantifying
+only over configurations of registers and memory above the stack.
 
@@
 confidentiality would allow a procedure rule that states that local
-variables do not change over calls, (which basically says that
-procedure calls are a no-op except for observations and the return value)
-and that lemmas proved about procedure return values and observations can
-be applied at call sites.
+variables do not change over calls (which basically says that
+procedure calls are a no-op except for observations and the return
+value), and that lemmas proved about procedure return values and
+observations can be applied at call sites.
 
@@
 from g, it is output in both traces.
-
-Meanwhile, WBCF enables reasoning about the whole structure of the program
-in terms of predictable control flows between its components.
 

### 2021-02-14 main.tex
Progress on properties

@@
 
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{n}\]
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
 \[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-  \times (n \in \mathbb{N}) \]
-
-An initial context \(\context_0 = \domMap_0 \times 0\) where \(\domMap_0\) maps
+  \times (\depth \in \mathbb{N}) \]
+
+An initial context \(\context_0 = \domMap_0, 0\) where \(\domMap_0\) maps
 the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
@@
 \begin{cases}
-  \sealed{n} & \text{if } \domMap ~ \component = \unsealed
+  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
                     \text{ and } \component \in \components \\
@@
 \begin{cases}
-  \unsealed & \text{if } \domMap ~ \component = \sealed{n-1} \\
+  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1} \\
   \domMap ~ \component & \text{else} \\
@@
               {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
-              {\(\mach,(\domMap,n) \stepsto_C \mach',(\push{\domMap}{n}{\sealcon ~ \mach},n+1)\)}
+              {\(\mach,(\domMap,\depth) \stepsto_C \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
 \end{minipage}
@@
               {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
-              {\(\mach,(\domMap,n) \stepsto_C \mach',(\pop{\domMap}{n},n-1)\)}
+              {\(\mach,(\domMap,\depth) \stepsto_C \mach',(\pop{\domMap}{\depth},\depth-1)\)}
 \end{minipage}
@@
 
+Our first property will be {\em ultra eager}, meaning that it will be quantified over each step
+of execution from any initial state. The criterion for integrity is simple: sealed components do not
+change.
+
+\definition A system enjoys {\em ultra-eager integrity} if, for any initial state \(\mach_0\),
+its initial context \(\context_0\), and its initial policy state \(\pol_0\), and each
+pair of adjacent states \(\mach,(\domMap,\depth),\pol \stepsto_C \mach',\context',\pol'\) in
+the induced trace from \(\mach_0,\context_0,\pol_0\), if \(\domMap ~ \component = \sealed{\depth'}\)
+for any \(\depth'\), then \(\mach(\component) = \mach'(\component)\).
+
 \subsection{Eager Integrity and the Call Rule}
 
+\definition Call Rule goes here.
+
+\definition A system enjoys {\em eager integrity} if, for any initial state \(\mach_0\), initial
+context \(\context_0\), and initial policy state \(\pol_0\) and any call depth \(\depth\), each segment
+\(\MCPTrace\) taken from the induced trace from \(\mach_0,\context_0,\pol_0\) by
+\(\prop (\domMap,\depth') = \depth' \geq \depth\) satisfies \(\pi_m(\head ~ \MCPTrace) \component =
+\pi_m(\last ~ \MCPTrace) \component\) for all \(\component\) where \(\domMap ~ \component = \sealed{\depth''}\)
+for any \(\depth''\).
+
 \subsection{Eager Confidentiality and the Transparency Rule}
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if, for
+all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\). We write
+this \(\mach \approx_\components \nach\).
+
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) have the {\em same differences}
+if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
+\(\nach ~ \component \not = \nach' ~ \component\), \(\mach' ~ \component = \nach' ~ \component\).
+This is written \(\mach,\mach' \diamond \nach,\nach'\).
+
+\definition Transparency Rule
+
+\definition A system enjoys {\em eager confidentiality} if, for any initial state \(\mach_0\), initial
+context \(\context_0\), and initial policy state \(\pol_0\) and any call depth \(\depth\), each segment
+\(\MCPTrace\) taken from the induced trace from \(\mach_0,\context_0,\pol_0\) by
+\(\prop (\domMap,\depth') = \depth' \geq \depth\) behaves the same as any \(\components\)-variant
+as follows.
+
+\begin{itemized}
+\item
+\item
+\item
+\end{itemized}
 

### 2021-02-15 main.tex
Add context to states

@@
 as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of
-% static or dynamic,
-(software- or
-hardware-enforced) policy monitor. In~\cref{sec:enforcement}, we will
-instantiate this model with a concrete dynamic policy enforcement mechanism
-based on \citet{DBLP:conf/sp/RoesslerD18}.
+machine model plus some kind of (software- or hardware-enforced) policy monitor.
+In~\cref{sec:enforcement}, we will instantiate this model with a concrete dynamic
+policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
 
@@
   addresses}. Both are drawn from some set of {\em words}
-$\WORDS$, ranged over by $\word$.
-%
-Our machine states are composed of {\em components} $k$, addressed either by
-words (memory addresses) or by register names ($r$) drawn from some set
-$\REGS$.
+\(\WORDS\), ranged over by \(\word\).
+%
+Our machine states are composed of {\em components} \(\component\), addressed either by
+words (memory addresses) or by register names (\(\reg\)) drawn from some set
+\(\REGS\).
 %
@@
 The register names are assumed to include two special purpose registers: the
-program counter {\PCname} and the stack pointer \SP.
+program counter {\PCname} and the stack pointer \(\SP\).
 
@@
 pair of a machine state and an {\em observation}, which may be either
-silent (written $\tau$) or a word of data\ifaftersubmission\bcp{This is a rather concrete
+silent (written \(\tau\)) or a word of data\ifaftersubmission\bcp{This is a rather concrete
   choice!  Why not leave the set of observations abstract?}\fi that the machine
@@
 
-We choose to model observations explicitly, rather than simply
-treating
+We choose to model observations explicitly, rather than simply treating
 a subset of registers or memory as being observable.  This choice gives
@@
 
-We now present our model of enforcement, which we introduced
-informally in \Cref{sec:threat}.  A policy is a general model of an
-enforcement mechanism, consisting of a set of policy states \(\pol \in
-\POLS\), a policy step function \((\mach, \pol) \polstep \pol' \in
+We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
+A policy is a general model of an enforcement mechanism, consisting of a set of policy
+states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
 \MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault.
-We assume that the
-compiler provides an initial policy state $\pol_0$ to accompany
-the initial machine state.
+it is undefined on input configurations that correspond to a policy fault. The policy
+can see the machine state, but cannot modify it. We assume that the
+compiler provides an initial policy state \(\pol_0\) to accompany the initial machine state.
 %
@@
 
+\subsection{Contexts}
+
+Separate from the machine state, our formal properties will keep track of additional context
+information representing the history of the run. The following definitions are parameterized
+over an arbitrary set of context states, \(\CONTEXTS\), an initial context state \(\context_0
+\in \CONTEXTS\), and a context step function \(\mach, \context \constep \context' \in
+\MACHS \times \CONTEXTS \rightarrow \CONTEXTS\).
+
+We lift contexts into an MP trace in the same way as we lifted policy states to get
+a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).
+
+\judgmenttwo{\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs} \mpstate{\mach_2}{\pol_2}\)}
+            {\(\mach_1, \context_1 \constep \context_2\)}
+            {\(\mpcstate{\mach_1}{\pol_1}{\context_1} \stepstocon{\obs}
+               \mpcstate{\mach_2}{\pol_2}{\context_2}\)}
+
 \subsection{Traces}
@@
 
-A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
-$\TRACEELEMS$s; we write $\TRACE{\TRACEELEMS}$ for the set of traces over
-$\TRACEELEMS$. We will mostly be concerned with sequences of pairs of
-machine states
-and observations, called \emph{machine traces} and denoted $\machT \in \MOTRACE$,
-and with sequences of triples of a machine state, a policy state, and an
-observation, called \emph{MP-traces} and denoted $\MPT \in \MPOTRACE$.
+A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
+\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
+\(\TRACEELEMS\). We will mostly be concerned with sequences of pairs of machine states
+and observations, called \emph{machine traces} and denoted \(\machT \in \MOTRACE\),
+and with sequences of 4-tuples of a machine state, a policy state, a context, and an
+observation, called \emph{MPC-traces} and denoted \(\MPCT \in \MPCOTRACE\).
 
@@
 %
-If \(f\) be a predicate on machine states, we define \(\mach
+If \(f\) a predicate on machine states, we define \(\mach
 \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
@@
 \judgmenttwo{\(\mach \hookrightarrow \machT'\)}
-            {\(\PUT~(f,\machT')
-              = \machT\)}
+            {\(\PUT~(f,\machT') = \machT\)}
             {\(\mach \hookrightarrow \machT | f\)}

### 2021-02-17 main.tex
Redid a bunch of trace stuff. Needs to be clearer on whether observations are in the state, but intro is next on my plate

@@
 
-\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
-               \polstep \pol_2\)}
-            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
-               \mpstate{\mach_2}{\pol_2}\)}
+\judgmenttwo{\(\mach \stepstoobs{\obs} \mach'\)}{\((\mach, \pol)
+               \polstep \pol'\)}
+            {\(\mpstate{\mach}{\pol} \stepstopol{\obs}
+               \mpstate{\mach'}{\pol'}\)}
 
@@
 
-\judgmenttwo{\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs} \mpstate{\mach_2}{\pol_2}\)}
-            {\(\mach_1, \context_1 \constep \context_2\)}
-            {\(\mpcstate{\mach_1}{\pol_1}{\context_1} \stepstocon{\obs}
-               \mpcstate{\mach_2}{\pol_2}{\context_2}\)}
+\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol{\obs} \mpstate{\mach'}{\pol'}\)}
+            {\(\mach, \context \constep \context'\)}
+            {\(\mpcstate{\mach}{\pol}{\context} \stepstocon{\obs}
+               \mpcstate{\mach'}{\pol'}{\context'}\)}
 
@@
 coinductively relates a machine state with the trace of machine states
-and observations produced by repeated application of step:
-
-\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
-            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
-%
+and observations produced by repeated application of step. It's defined in
+terms of a coinductive helper, \(\mathit{obs-forward}\):
+
+\judgmenttwo{\(\mach \stepstoobs{\obs'} \mach'\)}
+            {\(\mathit{obs-forward} ~ \mach' ~ \obs' = \machT\)}
+            {\(\mathit{obs-forward} ~ \mach ~ \obs = (\mach,\obs) \machT\)}%
+%
+\judgment{\(\mathit{obs-forward} ~ \mach ~ \tau = \machT\)}
+         {\(\mach \hookrightarrow \machT\)}
 \noindent
@@
 
-In a similar fashion, we define \(\mpstatename \hookrightarrow \MPT\)
-to relate a machine-policy state pair to the MP-trace
-induced by the step function on $\MPS$. Unlike the $\MOTRACE$ above, this trace can
-be finite as the policy can cause a fault, in which case the final observation is
-a \(\tau\):
-
-\begin{minipage}{.4\textwidth}
-\judgment{\(\neg \exists \pol_1 . \mpstate{\mach_0}{\pol_0} \polstep \pol_1\)}
-         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
-\end{minipage}
-\begin{minipage}{.55\textwidth}
-  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstopol{\obs}
-                 \mpstate{\mach_1}{\pol_1}\)}
-              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
-              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow
-                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
-\end{minipage}
-
-We will also use \(\pi_m\) to project the trace of machine states
-out of a trace, \(\pi_p\) to project the trace of policy states, and
-\(\pi_o\) to project the observations, with \(\pi_m\) and \(\pi_o\) overloaded
-for machine traces and MP-traces.
+In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
+to relate a machine-policy-context state triple to the MPCO-trace
+induced by the step function.
+
+\judgment{\(\neg \exists \pol' . \mpcstate{\mach}{\pol} \polstep \pol'\)}
+         {\(\mathit{obs-forward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
+           \mpcstate{\mach}{\pol}{\context},\obs\)}
+%
+\judgmenttwo{\(\mpcstate{\mach}{\pol}{\context} \stepstoobs{\obs'} \mpcstate{\mach'}{\pol'}{\context'}\)}
+            {\(\mathit{obs-forward} ~ \mpcstate{\mach'}{\pol'}{\context'} ~ \obs' = \machT\)}
+            {\(\mathit{obs-forward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
+              (\mpcstate{\mach}{\pol}{\context},\obs) \machT\)}%
+%
+\judgment{\(\mathit{obs-forward} ~ \mpcstatename ~ \tau = \machT\)}
+         {\(\mpcstatename \hookrightarrow \machT\)}
+
+We will also use \(\pi_m\), \(\pi_p\), \(\pi_c\), and \(\pi_o\) to project the components out of an
+MPC state, and overload them to project a trace of the components out of an MPC trace.
 
@@
 \paragraph*{Until}
-The operation \(\PUT~(f,\Trace)\)
-%\bcp{Wouldn't
-%``While'' be more mnemonic?} \sna{Appealing, but it's more of a
-%``while-not''? Trying out ``until''. In any case... moving away
-%from doing joins means this needs to get some work to eliminate
-%the trailing observation. Here's an attempt.}
-takes a trace
-%% \bcp{I changed a : here to $\in$.  We should check if there are
-%% any more to fix.}
-\(\Trace \in \TRACE{\TRACEELEMS} \times \OBSS\) and a predicate
+The operation \(\PUT ~ f ~ \Trace\) takes a trace
+\(\Trace \in \TRACE{\TRACEELEMS \times \OBSS}\) and a predicate
 on elements \(f \subset \TRACEELEMS\) and gives the prefix of
 \(\Trace\) ending with the first element on which \(f\) holds.
-One subtlety is that, because the observation associated with some element
-in a trace
-represents what can be observed when this element is executed (i.e., the
-observation on that step), when we take a prefix from a trace containing
-observations
-we must replace the final observation with a \(\tau\).
 
@@
   \begin{minipage}{.3\textwidth}
-    \judgment{\(f\ \traceelem\)}
-             {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
+    \judgment{\(f ~ \traceelem\)}
+             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
   \end{minipage}
@@
     \judgment{}
-           {\(\PUT~(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
+             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
 %
@@
   \begin{minipage}{.3\textwidth}
-  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT~(f,\Trace) = \Trace'\)}
-              {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
+    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
+                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
   \end{minipage}
@@
 If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate never holds, then
+predicate must hold on its final element. If the predicate always holds, then
 \(\PUT\) is the identity function.
@@
 %
-If \(f\) a predicate on machine states, we define \(\mach
-\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
-running \(\mach\) up to \(f\)'') and similarly for traces with policies:
+If \(f\) is a predicate on machine states, we define \(\mpcstatename
+\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
+running \(\mpcstatename\) up to \(f\)'') and similarly for traces with policies:
 
 \begin{center}
-\begin{minipage}{.3\textwidth}
-\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
-            {\(\PUT~(f,\machT') = \machT\)}
-            {\(\mach \hookrightarrow \machT | f\)}
-\end{minipage}
-\begin{minipage}{.6\textwidth}
-\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
-            {\(\PUT~(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
-         {\(\mpstatename \hookrightarrow \MPT | f\)}
-\end{minipage}
+\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
+            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
 \end{center}
 
-%    \COQ{this operation steps until it reaches a state on which \(f\) holds,
-%    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
-%    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
-%    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
-%      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
-%    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}
+\paragraph*{Context Segments}
+
+In particular, we wish to formulate properties around consecutive subtraces
+in which the context obeys a given predicate. This will commonly be used to extract
+individual calls from a trace by taking subtraces in which the stack is
+at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
+we have \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
+``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''
+
+\begin{center}
+  \judgmenttwo[Now]
+              {\(f ~ \mach ~ \context\)}
+              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
+              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
+%
+  \judgmentthree[Skip]
+                {\(f ~ \mach ~ \context\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
+%
+  \judgmentthree[Search]
+                {\(\neg f ~ \mach ~ \context\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
+\end{center}
+
+\paragraph*{Reachability}
+
+We often want to quantify over all traces from any initial state so that our
+properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
+is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
+and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
+such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
 
@@
 \[
-(\push{\domMap}{n}{\components}) ~ \component =
+(\push{\domMap}{\depth}{\components}) ~ \component =
 \begin{cases}
@@
 \[
-(\pop{\domMap}{n}) ~ \component =
+(\pop{\domMap}{\depth}) ~ \component =
 \begin{cases}
@@
               {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
-              {\(\mach,(\domMap,\depth) \stepsto_C \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
+              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
 \end{minipage}
@@
               {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
-              {\(\mach,(\domMap,\depth) \stepsto_C \mach',(\pop{\domMap}{\depth},\depth-1)\)}
+              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
 \end{minipage}
@@
            {\(\mach \stepsto \mach'\)}
-           {\(\mach,\context \stepsto_C \mach',\context\)}
+           {\(\mach,\context \stepstocon \mach',\context\)}
 \end{center}
@@
 
-\definition A system enjoys {\em ultra-eager integrity} if, for any initial state \(\mach_0\),
-its initial context \(\context_0\), and its initial policy state \(\pol_0\), and each
-pair of adjacent states \(\mach,(\domMap,\depth),\pol \stepsto_C \mach',\context',\pol'\) in
-the induced trace from \(\mach_0,\context_0,\pol_0\), if \(\domMap ~ \component = \sealed{\depth'}\)
-for any \(\depth'\), then \(\mach(\component) = \mach'(\component)\).
+\definition A system enjoys {\em ultra-eager integrity} if for any reachable state
+\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon{} \mpcstatename'\),
+and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
+\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
+\pi_\mach(\mpcstatename') ~ \component\).
 
@@
 
-\definition Call Rule goes here.
-
-\definition A system enjoys {\em eager integrity} if, for any initial state \(\mach_0\), initial
-context \(\context_0\), and initial policy state \(\pol_0\) and any call depth \(\depth\), each segment
-\(\MCPTrace\) taken from the induced trace from \(\mach_0,\context_0,\pol_0\) by
-\(\prop (\domMap,\depth') = \depth' \geq \depth\) satisfies \(\pi_m(\head ~ \MCPTrace) \component =
-\pi_m(\last ~ \MCPTrace) \component\) for all \(\component\) where \(\domMap ~ \component = \sealed{\depth''}\)
-for any \(\depth''\).
+\definition Let \(\mpcstatename\) be a reachable state whose next instruction is annotated
+as a call and whose call depth is \(\depth\). Let \(f ~ \mach ~ \context\) hold when \(\context = \_,\depth\),
+and \(\components\) be the set of components \(\component\) such that
+\(\sealcon ~ \pi_\mach(\mpcstatename) ~ \component\) holds. Then the {\em Call Rule} says that if
+\(\mpcstatename \stepstocon \mpcstatename'\), it is sound to assume that
+\(\mpcstatename' \hookrightarrow \MPCT | f\) and if \(\last(\MPCT) = \mach'\), then
+\(\pi_\mach(\mpcstatename) ~ \component = \mach' ~ \component\).
+
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold when
+\(\depth' \geq \depth\). A system enjoys {\em eager integrity} if each reachable segment
+\(\MPCT \langle f_\depth \rangle\) satisfies \(\pi_m(\head(\MPCT)) ~ \component =
+\pi_m(\last(\MPCT)) ~ \component\) for all \(\component\) where \(\domMap ~ \component =
+\sealed{\_}\).
 
@@
 
-\definition A system enjoys {\em eager confidentiality} if, for any initial state \(\mach_0\), initial
-context \(\context_0\), and initial policy state \(\pol_0\) and any call depth \(\depth\), each segment
-\(\MCPTrace\) taken from the induced trace from \(\mach_0,\context_0,\pol_0\) by
-\(\prop (\domMap,\depth') = \depth' \geq \depth\) behaves the same as any \(\components\)-variant
-as follows.
-
-\begin{itemized}
-\item
-\item
-\item
-\end{itemized}
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold for
+\(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth),\_\) and
+\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\). \(\MPCT\) corresponds to the
+trace of a subroutine from its call to its return.
+
+If we take the induced trace from this variant until a return,
+\(\nach,\pol,\context \hookrightarrow \NPCT | \neg f_\depth\), we get a subtrace corresponding
+to the same subrouting called with the same (register) arguments but with a potentially different
+state at the call-site. A system enjoys {\em eager confidentiality} if for all such pairs of traces,
+they correspond according to one of four cases:
+
+\begin{enumerate}
+\item \(\MPCT\) may be finite and \(\pi_\context(\last(\MPCT)) = (\_,\depth')\) with \(\depth' < \depth\),
+  and \(\pi_\context(\last(\NPCT)) = (\_,\depth')\) with \(\depth' < \depth\). That is, both traces end
+  having left the call of depth \(\depth\). Further:
+  \begin{itemize}
+  \item They have the same observable behavior, \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
+  \item Their final states have the same differences, \(\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
+  \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\)
+  \end{itemize}
+\item \(\MPCT\) and \(\NPCT\) may both be infinite, with \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
+\item \(\MPCT\) may be finite but end without returning -- which can only have happened if it fail-stopped --
+  and its observable behavior prefixes that of \(\NPCT\), \(\pi_\obs(\MPCT) \lesssim \pi_\obs(\NPCT)\)
+\item \(\NPCT\) may end with a fail-stop as in case 3, with \(\pi_\obs(\NPCT) \lesssim \pi_\obs(\MPCT)\)
+\end{enumerate}
 

### 2021-02-17 main.tex
Added introduction examples

--- 2021-02-15 main.tex
@@
 While this work represents an important step in the right direction, we believe
-there is still room to craft a more abstract characterization
+there is still room to craft an operational characterization
 of stack safety that is independent of capabilities or any other specific
-enforcement
-technique.  In other words, we seek formal criteria that can be applied
+enforcement technique.  In other words, we seek formal criteria that can be applied
 to a mechanism for protecting stack-based low-level programs to judge
@@
 %
-Our key contribution is an abstract
-presentation of local state encapsulation using the technical
-framework of language-based security, together with a simple
+Our key contribution is an abstract presentation of local state encapsulation
+using the technical framework of language-based security, together with a simple
 characterization of well-bracketed control flow.
 
-The intuition behind \citeauthor{Skorstengaard+19}'s {\em
-  local state encapsulation} is that callers should be protected from
-callees: a callee should not be able to access or overwrite private data
-in the caller's stack frame.
-%
-We formalize this intuition in (1) a {\em stack-confidentiality}
-property, which promises that a callee's observable behavior is not affected
-by memory outside its stack frame (e.g., a caller's private data or
-uninitialized memory), plus (2) a {\em stack-integrity} property,
-which promises that, if a function call ever returns, the caller's stack
-data at the point of return will be identical to what it was before the
-call.
-
-The intuition behind \citeauthor{Skorstengaard+19}'s {\em
-  well-bracketed control flow} is the need to protect the call graph
-itself: a callee should pass control back to the caller only by
-transferring control to the instruction
-following the call; it should not be allowed to jump back into the
-caller's code at an arbitrary point, either directly (via a jump
-instruction) or indirectly (by corrupting stack data and
-returning). Well-bracketed control flow establishes safety between
-functions: each instruction is uniquely owned by some function, and
-all jumps between functions are either calls that enter at a marked
-entry point or returns that re-enter just after their associated call
-in the standard nested call structure.
-
-To phrase local state encapsulation as a
-security property requires an explicit attacker model: a clear
-understanding of who is being protected and from what. Most existing
+\begin{figure}[b]
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+main = {
+  int x = 42;
+  f();
+  print(x);
+}
+\end{verbatim}
+    \end{minipage}
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+main = {
+  int x = 1;
+  f();
+  if(x)
+    g();
+  else
+    h();
+}
+\end{verbatim}
+    \end{minipage}
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+main = {
+  int x = 1;
+  f();
+  x = -x;
+  f();
+  print(x);
+}
+\end{verbatim}
+  \end{minipage}
+  \caption{Integrity Examples}
+  \label{fig:intexamples}
+\end{figure}
+
+\begin{figure}[b]
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+main = {
+  int key = get_key();
+  f();
+  use_key(key);
+}
+\end{verbatim}
+    \end{minipage}
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+main = {
+  int x = get_secret();
+  if(x)
+    g();
+  else
+    h();
+  f();
+}
+\end{verbatim}
+    \end{minipage}
+    \begin{minipage}{.3\textwidth}
+\begin{verbatim}
+main = {
+  int x = 0;
+  int y = get_secret();
+  f();
+  print(x);
+}
+\end{verbatim}
+  \end{minipage}
+  \caption{Confidentiality Examples}
+  \label{fig:confexamples}
+\end{figure}
+
+\paragraph*{The Use of Stack Safety}
+
+When does the programmer rely on stack safety? The most natural example is the first
+in figure \ref{fig:intexample}, in which {\tt main} calls stores 42 to a local variable {\tt x},
+calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
+{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
+overwrite it's caller's stack data. To this end, our properties support a
+{\em call rule} that can be used to reason that on return from {\tt f}, {\tt x} still
+contains the same value. This is not only useful for the programmer: consider the second
+example, where the call to {\tt h} should be unreachable and therefore subject to dead code
+elimination, if only the compiler can soundly reason that {\tt x} does not change.
+The property that enables the call rule is, in security terms, {\em integrity}: callers'
+data is protected from modification by callees.
+
+Alongside integrity, we have {\em confidentiality}, in which callers' data is protected
+from leakage by their callees. Consider the first example in figure \ref{fig:confexamples},
+in which {\tt main} recieves a key, then calls {\tt f}. If {\tt f} is an untrusted function,
+it would be bad if it could just read {\tt key}. But there are more subtle examples,
+such as the second one, where a secret value controls which function of {\tt g} and
+{\tt h} are called prior to calling {\tt f}. Here, if {\tt g} or {\tt h} leaves behind
+identifying information in unused stack space, {\tt f} could still deduce information
+about the secret. So our confidentiality properties also support a {\em transparency rule},
+that each function's behavior is independent of the state of its caller, other than arguments.
+This also makes it easier, though not trivial, to prove idempotence: the channels through
+which a callee can remember its own history are limited to those outside the stack.
+
+Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
+{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
+callers from their callees. Consider the last example in figure \ref{fig:intexample},
+a variant of the famous ``awkward example,'' in which {\tt f} is called twice and may not
+be idempotent. It should be the case that {\tt main} prints -1, but if
+{\tt f}'s first activation stores its return address somewhere accessible, the second
+activation can return to before {\tt x} is negated, cause {\tt x} to be set to
+1 again before a third activation that would return normally. Then {\tt main} would print 1.
+
+Preventing this form of reentrancy attack is the intuition behind
+\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
+the call graph itself. A callee should pass control back to the caller only by
+transferring control to the instruction following the call; it should not be allowed
+to jump back into the caller's code at an arbitrary point, either directly (via a jump
+instruction) or indirectly (by corrupting stack data and returning). It also maintains
+the integrity of the stack structure; otherwise a callee could leak data in the third
+example of \ref{fig:conf-example} by failing to properly restore the stack pointer.
+As {\tt x} is accessed by its offset from the stack pointer, an adversary could leave
+the stack pointer so that {\tt x}'s offset now points to {\tt y}, and trick {\tt main}
+into printing its secret.
+
+To phrase LSE and WBCF as security properties requires an explicit attacker model:
+a clear understanding of who is being protected and from what. Most existing
 stack protection mechanisms envisage scenarios where a callee is
@@
 of ``attacker'' changes at many points throughout the execution of the program.
-We address
-this issue by introducing a dynamically changing notion of {\em accessibility} for
-the different components of the machine state and adjusting the integrity
-and confidentiality levels of different stack locations and registers at
-every transition (call or return) between functions.
+We address this issue by introducing a dynamically model of {\em domains} for
+the different components of the machine state, in which some stack locations are
+sealed by the caller that needs to preserve them while the rest are unsealed and
+available for use.
+
+First we introduce this model for a single simple stack with no sharing between callers and
+callee. Then we expand it to allow components to be passed, granting a callee one time
+access to the caller's data, or shared, allowing components to be outright
+exempted from the stack safety properties. Shared components could model objects whose
+addresses are passed and whose security must be modeled by a separate property, such as
+memory safety enforced by capabilities.
+
+\sna{Rewriting done up to here, then skips down to property details (sections 4/5).}
 
@@
 %efficient, but our variant is {\em not} efficient!}
-
-\sna{There's been some inconsistency in ``stack safety'' vs.
-     ``stack-safety,'' I'm fixing on the former.}
-\bcp{Be careful: When a compound phrase is used as an adjective to modify a
-  following noun, it gets hyphenated; otherwise it does not.  This is
-  correct grammar, not inconsistency!  :-)}
-\sna{Aha, I forgot that. Corrected, and I've added it in more places.}
 

### 2021-02-18 CheriApproaches.tex
Introduction additions in main.tex and new discussion of approaches to cheri (really, to everything)

\documentclass{article}

\usepackage{geometry}

\begin{document}

This document deals with questions of the division between machine state and policy state.
In existing models of our security properties, the state is divided into a machine state,
which is a mapping from addresses and registers to data, and a policy state, which may be
any arbitrary data. A machine state can step on its own, or in conjunction with a policy
state; when states step together the machine state component always steps as it would on
its own. A policy state may only cause the step to fail outright, never change the result.

We are faced with the task of deciding 1) which parts of an architecture should be treated
as machine, and which parts as policy; 2) what is to be done with the policy state in the
context of confidentiality properties; and 3) whether there are parts of the machine state
that ought to be treated differently in such properties.

We will discuss these ideas in terms of three enforcement mechanisms:

\begin{itemize}
\item HOPE
\item Cheri
\item A software-based mechanism such as SFI
\end{itemize}

\section{What if we eliminate Policy States?}

Let's start by considering what a system would look like that doesn't use policy state. The
simplest such model simply makes the machine transition function partial, with failure to step
indicating a fail-stop. Here the enforcement is built directly into the machine. This applies to
our three enforcement mechanisms as follows.

\paragraph*{HOPE} The machine contains both ordinary registers and memory, and tag registers and memory.
The transition function is parameterized by a tag rule table, and we must be careful to construct the model
such that tags only be read and updated as described by the table. Fail-stops should occur only on tag
rule failures.

\paragraph*{Cheri} The machine contains ordinary registers and memory, with some effort needed to model
the extra space that capabilities take up. (Either words need to be larger to fit capabilities, or we
need to model them being spread across multiple words.) It also contains the map of 1-bit tags marking
capabilities as valid. The ISA is extended with the Cheri operations and the transition function handles
the valid-capability tags appropriately. Here a fail-stop occurs whenever the Cheri architecture would
trap to a failstate, mostly when reading or writing through an invalid capability or outside of a
valid capability's authority.

\paragraph*{Software} Here the machine gets to be completely unchanged, and never failstops. Instead
we restrict the set of initial states to those whose code comes from the compiler with the appropriate
modifications, and it will simply never violate the policy - any errors will take it to some error
state where it silently diverges, which is perfectly allowed.

\paragraph*{Implications for confidentiality} If everything lives in the machine state, then
we need to think about what it means to vary it to create a confidentiality property. Take the HOPE
example: does a variant include states with different tags, such that either the primary
trace failstops and the variant does not, or the variant does and the primary does not? Similarly,
in Cheri, do we vary the valid-capability bit? What about bounds information?

We can pretty firmly say that we do not need to vary tags, including the capability bit. That's
because these can never be read, and therefore it is meaningless to talk about their confidentiality.
But the bounds information in Cheri presents a bigger problem. We will revisit it as we start moving
policy data out of the machine.

\section{Separate Policy States}

We have identified some kinds of data that can never be read and serve exclusively to inform
the enforcement mechanism. We can think of data as having two meanings, its ``application content'' and its
``security content.'' In HOPE, the distinction seems clear: tags have only security content, never application.
So we can justify separating out security-only data into a separate part of the state, from which it cannot
escape and therefore does not need to be varied to support confidentiality. Let's discuss how that
works in HOPE and Cheri (in software methods, of course, everything has application content.)

\paragraph*{HOPE} In HOPE, we put all tags in the policy state, along with any policy-specific machinery such as
counters that give us the fresh tags. The machine step function can now be total, with failure only
coming from the policy step function. But it turns out that some machine data still has security content.
Consider a load instruction: its tag rule must take into account the tag at the target address, which
means that the policy step function must be aware of the machine state.

This gives us the same question of what we do in confidentiality with variant states that fail-stop
when the original didn't.

\paragraph*{Cheri} In Cheri, the valid-capability bit has no application content, so these can live in
the policy state. But the application content of a capability is its bitwise value. We might argue
that actually the base and offset are application content, while permissions and bounds information
are security-only, but it is not feasible to separate the latter from the former. Even if we did, the
base and bounds would give us the same issues as HOPE pointers.

\paragraph*{Implications} The primary implication is that we no longer need to worry about what happens
if we vary tags, whether that's the full HOPE tag or the 1-bit Cheri tag. But that doesn't save us
from sometimes varying security content, because pointer values still have security content. In both systems,
if we read a valid pointer that is supposed to be secret and then dereference it, there is a variant run
in which the pointer is invalid and the variant fail-stops. This will always be the case no matter how we
separate machine and policy.

The advantages that we do get by separating them as much as possible are: it's easier to model the separate
state than if we built it into the machine architecture, which matters more for HOPE but is a small concern for
Cheri. And depending on how we handle these variant fail-stops, it may be more efficient to test if fewer of
them occur.

\section{Handling Variations in Security Content}

Either way the question becomes: should the variant trace be one with a policy active, or no? Let's
look at a concrete example.

\begin{verbatim}
main() {
  int[2] x = {0,0};
  int* y = &x+secret();
  f(&x);
}

f() {
  int** z = guess_y_addr();
  print(z**);
}
\end{verbatim}

Here {\tt main} has an array {\tt x} and saves a pointer to one of its elements, but which one is secret.
Then it passes the address to  {\tt f}, so that {\tt f} can access {\tt x} legally according to sharing.
{\tt x} has a tag that marks it permissible for anyone to access it, but {\tt f} has actually accessed {\tt x}
via reading the hidden information in {\tt y}. Suppose that our policy contains an error such that the access
of {\tt y} is successful - then {\tt f} always ends up printing 0 regardless, so we haven't violated
confidentiality. But we would if we printed {\tt y} outright or if the values in {\tt x[0]} and
{\tt x[1]} differed.

Consider the variant states that we can have. {\tt x} does not vary, but {\tt y} can point to {\tt x[0]},
{\tt x[1]}, or a location beyond {\tt x} entirely. In the first two cases, our output is 0,
so confidentiality is maintained. What do we do with that case? We can:

\begin{enumerate}
\item Recognize that the variant should fail-stop, and treat that as a violation
\item Recognize that the variant should fail-stop, and ignore this variant
\item Do not check for a variant fail-stop, and therefore likely get a violation
\end{enumerate}

I argue for option 2, because either some other variation that doesn't create an illegal pointer
will be a violation, or the particular trace has not leaked any information. In this case,
the error will be detected if we ever run the comparable case where {\tt x = {0,1}}, but if the compiler
somehow only produced traces where this issue is invisible, we need not consider it. Still, option 1
could be fine, since our policies are meant to prevent {\tt y} from being read in the first place.

Option 3 is clearly bad because if the contents of {\tt y} are used in some other way, its behavior
could be radically different from the primary trace. Whether it fails or not is virtually unrelated to
the initial potential violation. This seems very bad.

Considering the same situation in Cheri, we also have the possibility to get additional tag failures
by varying not just the address pointed to, but the permissions bits on a capability. But the fundamental
question remains the same: what to do with fail-stop variants.

\section{Cheri Implementation}

In order to implement Cheri we'll need to make a couple of changes. First of all, we need to extend the model
with capability instructions and preferably add some capability-sized registers. It will be tedious aligning
capabilities properly in memory; could we instead take a shortcut by just making words capability-sized, and
only reading half of them for non-capabilities? We need at the very least cap-load, cap-store, arithmetic,
and the instructions to derive capabilities from the ambient capabilities (esp. stack capability). Currently
looking into just how much we can strip out and still have it work.

The policy state will be the mapping from addresses/registers to their capability bits. Capability instructions
will maintain the mapping as appropriate. Certain operations will fail-stop based on violating that mapping.

\end{document}

### 2021-02-18 main.tex
Introduction additions in main.tex and new discussion of approaches to cheri (really, to everything)

@@
 
-\sna{Rewriting done up to here, then skips down to property details (sections 4/5).}
-
 After formally defining stack safety, we show that the definition is
@@
 mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
-\citeauthor{DBLP:conf/sp/RoesslerD18}'s
-conservative Depth Isolation policy, which not only enforces our basic
-stack-safety property, but also an alternative lockstep property that
-checks for integrity and confidentiality violations at every step of
-an execution trace. The advantage of this property is that it ``fails
+\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
+We give {\em eager} properties that this policy enforces, in which each
+callee, at the point of its return, can be seen to respect the integrity
+and confidentiality of its caller. Additionally we have an
+{\em ultra-eager} integrity property which guarantees that integrity
+is respected at each step. The advantage of this property is that it ``fails
 faster'' and is therefore more suitable for efficient testing. We prove (in
-Coq) that this ``lockstep'' property implies our original definition of stack
+Coq) that the ultra-eager property implies our original definition of stack
 safety.
@@
 Instead of enforcing the integrity and confidentiality of the stack at the
-exact moment
-when a callee returns, this policy signals violations when the
+exact moment when a callee returns, this policy signals violations when the
 caller actually accesses data that the callee has written into its stack
@@
 properties, and to our knowledge there is no formal statement of the properties that
-it ought to enforce. We propose ``observable'' versions of integrity and
-confidentiality property, which capture the desired behavior of a policy that
-detects violations between when they occur and when they make a visible impact.
-Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing ``Lazy Tagging
-and Clearing'' policy does not quite enforce these properties,
-but would if enhanced (at some performance cost).
+it ought to enforce. We propose a {\em lazy} integrity property, which capture the
+desired behavior of a policy that detects violations between when they occur and when
+they make a visible impact. Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s
+existing ``Lazy Tagging and Clearing'' policy does not quite enforce these properties,
+but formalizing the ideal property allows us to see how it falls short.
+We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
+at some performance cost.
+
+We extend our properties to allow sharing, both in the form of arguments passed
+on the stack and of objects shared widely beyond their owning function. In the latter
+case, restricting access the shared objects is beyond the scope of stack safety,
+but our model delimits them from other objects in the same frame.
+
 %\bcp{I worry that we are being a little
@@
 
-The first part of stack safety is {\em local state encapsulation}.
-Informally, it captures the notion that a caller's data is
-protected from a callee's actions. We formalize this notion
-as a security property with an integrity and a confidentiality
-component, where the former protects the caller's data from being
-written and the latter from being read.
-
-As is standard with security properties, we associate with each
-component of the machine a {\em label}: a pair of security levels
-denoting high or low confidentiality and high or low integrity.
-%
-\[\mathit{label} ::= \{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT,\LOWINT\}\]
-%
-In a traditional setting, integrity would enforce that the contents of
-$\HIGHINT$ locations remain unchanged, while confidentiality would
-enforce that the contents of $\HIGHSEC$ locations do not influence the
-observable behavior of the machine. However, when dealing with nested
-calls and returns, stack safety needs to enforce integrity and
-confidentiality for {\em every} function call in a trace; that is, we
-must treat the security level of data dynamically.
-%
-To that end, we introduce the notion of a {\em contour}, a map from
-components to labels:
-%
-\[\contour \in \CONTOURS ::= \COMPONENTS \rightarrow \mathit{label}\]
-
-When \(\mach_c\) is a call state (\(\callmap(\mach_c(\PCname)) = n\) for some $n$) and
-\(\mach_c \stepstoobs{\obs} \mach_e\) for some
-$\mach_e$ and $\obs$, we can construct a contour based on the bounds of the
-callee's stack frame.
-%
-The private state of the callee is the entire area of the memory from
-stack base to stack pointer, with the exception of the $n$ locations
-corresponding to the arguments and result locations of the
-call. Private components are marked high confidentiality and high
-integrity: the callee should be able to neither read nor write them.
-%
-Everything above the stack pointer at the point of the call is
-``uninitialized'' memory that can be written into but should not be
-read. To protect the callee itself from such reads, these components
-are marked high confidentiality and low integrity.
-%
-The same is true for the designated output location: it can be
-freely written into but should not be read.
-%
-Code is always high integrity and low confidentiality.
-%
-Finally, registers and the frame of arguments and results are free
-for the callee to use, and are therefore tagged with low integrity and
-low confidentiality.
-
- \[\mathit{Cof}(\mach,n)(\component) =
-  \begin{cases}
-    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
-                         \text{ and } \codemap(\component) = \funid \text{ for some }\funid\\
-    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
-                          \text{ and } \component \geq \mach(\SP)
-                          \text{ or } \component = \rout\\
-    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
-                          \text{ and } \component \leq \mach(\SP) \negate  n \\
-    (\LOWSEC,\LOWINT) & \text{else} \\
-  \end{cases}\]
-
-Going back to our running example, here is the state of the machine at
-the call from {\tt main} to {\tt f}, showing addresses annotated with their
-corresponding labels:
-\vspace*{0.2em}
-% At Call to G
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{43.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{$(\LOWSEC,\HIGHINT)$}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{$(\HIGHSEC, \HIGHINT)$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{$(\LOWSEC, \LOWINT)$}}]%
-\memory{6}{\unusedc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\\
-\end{center}
-
-%If $\mach$ is the state at this call, the corresponding contour would
-%be:
-%\[
-%\mathit{Cof}(\mach,2) = \left \{
-%\begin{array}{l}
-%  0\ldots98\} \mapsto (\LOWSEC, \HIGHINT)\\
-%  99 \mapsto (\HIGHSEC, \LOWINT)\\
-%  100, 101 \mapsto (\HIGHSEC, \HIGHINT)\\
-%  102, 103 \mapsto (\LOWSEC, \LOWINT)\\
-%  104\ldots \mapsto (\HIGHSEC, \LOWINT)\\
-%  r \mapsto (\LOWSEC, \LOWINT)\\
-%\end{array}
-%\right .
-%\]
-
-\subsubsection{Stack Integrity}
-
-The first component of local state encapsulation is an integrity
-property: it ensures that the caller's data is never overwritten by a
-callee. We will formally define integrity in two steps: first, we will
-define integrity as a trace property over the trace corresponding to a
-single call (from its entry point to its corresponding return), and
-then we will lift that to a property for every such subtrace of a
-callee.
-
-\definition
-
-Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
-MP-trace. say \(\MPT\) satisfies {\em trace integrity} with
-respect to \(\contour\) if its final state (if any) agrees with its
-initial one on all components marked \(\HIGHSEC\) in \(\contour\):\bcp{I've
-  fixed some problems with not enough space around $\triangleq$, but I see
-  another here --- should check carefully for more.}
-%
-\[\begin{split}
-  \propdef{TraceInt} ~ \contour ~ \MPT \triangleq &
-    \head(\MPT) = \mpostate{\mach_e}{\_}{\_} \land
-    \last(\MPT) = \mpostate{\mach_r}{\_}{\_} \Rightarrow \\
-  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
-    \mach_e(\component) = \mach_r(\component) \\
-\end{split}\]
-
-Intuitively, the caller expects to find their private data untouched
-after the callee returns. That means that if the callee never returns,
-integrity is trivially satisfied. Similarly, the callee is allowed to
-temporarily overwrite the caller's private locations,
-as long as their values are restored
-by the time the callee returns. We will see variations of this later on: a
-stronger, inductive integrity property that ensures that the caller's
-private data is never overwritten, even during the callee's execution
-(which leads to easier reasoning and more effective testing);
-and a weaker, lazy integrity property that only ensures the preservation
-of locations that subsequently affect the program's observable actions
-(which allows for a more efficient enforcement mechanism).
-
-\begin{figure}
-  \integrityexample
-  \caption{Failing Trace Integrity}
-  \label{fig:intex}
-\end{figure}
-
-Trace integrity characterizes the trace corresponding to a single
-callee.
-Going back to our running example, such a trace would be, for
-instance, the entire trace of {\tt g} from instruction 40 to
-instruction 47.
-As an example of an integrity violation affecting this trace, \cref{fig:intex} shows what happens
-if we change line 44 of the example program to $[\rsp - 5] \leftarrow r_4$.
-Instead of storing data to the return value slot,
-the changed version overwrites {\tt f}'s local variable {\tt z}.
-The figure shows the modified trace, with snapshots of the initial
-and final states colored to show the contour; high
-integrity memory is in \high, while low integrity is in \low.
-At the end of the trace high integrity address 103 should match the initial
-state, and does not. So trace integrity is violated.
-
-We now lift this definition to an integrity property for a complete
-execution trace. We do this by requiring that trace integrity holds
-for every subtrace delimited by a call and its corresponding return.
-
-\definition Given a trace \(\MPT\), % and a call map \(\callmap\)
-%and a function map \(\codemap\),
-an \(n\)-{\em transition} is a pair of
-machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
-\(\mpstate{\mach_e}{\pol_e}\) in \(\MPT\) such that
-\(\mach_c\) is a valid call with \(n\) arguments, and
-\(\mpstate{\mach_c}{\pol_c}\) steps to \(\mpstate{\mach_e}{\pol_e}\).
-
-%and \(\mach_e\) is an entry point for some function \(f\):
-
-  \[\begin{split}
-    \mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}
-    \triangleq & \exists \obs.
-    ~ \mpstate{\mach_c}{\pol_c} \stepstopol{\obs} \mpstate{\mach_e}{\pol_e} \\
-    & \land \callmap(\mach_c(\PCname)) = n
-      %\land \codemap(\mach_e(\PCname)) = f
-  \end{split}\]
-
-\definition
-
-We say that a system enjoys {\em stack integrity} if for every initial
-state $(\mach_0,\pol_0)$ and its corresponding induced trace $\MPT$ (with
-$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), and for every
-$n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
-$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, if we take the prefix of the
-call until its corresponding return $\mpstate{\mach_e}{\pol_e}
-\hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}$, then $\MPT_{\mathit{call}}$ must
-satisfy trace integrity:
-\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{\mathit{call}}\]
-\apt{This could still be clearer. Interleaving informal and mathy in the
-same sentence doesn't work so well.}
-
-%{\em Stack integrity} is a property of a system with respect to a call map
-%\(\callmap\) that holds when every \(n\)-transition  in every legal trace exhibits trace integrity
-%on the prefix up to its return, if any. For any initial state \(\mach_0\)
-%we take the induced trace. For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
-%\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
-%the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
-%\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\),
-%\(\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
-%system as a whole enjoys stack integrity with respect to \(\callmap\).
-%
-%\apt{Some more words needed here to smooth the technicalities.}
-%\sna{Attempt...}
-%Rewrote for clarity
-
-\subsubsection{Stack Confidentiality}
-
-The second component of local state encapsulation is a confidentiality
-property: just like integrity ensures that a caller's data is never
-overwritten by a callee, confidentiality ensures that the caller's data never
-influences the callee's observable behavior. Again, we will formally define
-confidentiality in two steps: a trace version and a system
-version.
-
-Unlike integrity, confidentiality is a noninterference-style property,
-comparing the behavior of a machine with an arbitrary variation that
-preserves low confidentiality data.
-
-\definition%{Variations}
-Two machine states \(\mach\) and \(\mach'\) are {\em variations}
-of one another given a contour \(\contour\) if they agree on low
-confidentiality data:
-%
-\[
-\mach \approx_\contour \mach' \triangleq \forall \component .
-\contour(\component) = \LOWSEC \Rightarrow \mach(\component) =
-\mach'(\component)
-\]
-%
-Then, the trace version of confidentiality operates on two traces: the
-``original'' trace $\MPT \in \MPOTRACE$ that corresponds to a callee $f$
-and a ``variant'' trace $\machT \in \MOTRACE$ that corresponds to how $f$
-would behave if it were started from a variant initial machine state.
-%
-As an example of variations, consider the top of \Cref{fig:confex}
-which shows a snapshot of the state as in the running example (top
-memory layout) and a variation of it (bottom).
-%
-The ``Effect'' column shows differences between the two runs by writing values
-in the original run in {\high} and values in the variant run in {\varied}
-separated by a slash.
-%
-This example will end up with a violation of confidentiality, as a
-variation in a confidential location (\textcolor{cyan}{42} vs
-\textcolor{magenta}{$-5$} in location 100) leads to a visible
-difference in return value.
-
-\definition
-\label{def:trace-conf}
-
-{\em Trace confidentiality} will be a property of two traces, \(\MPT\) and
-\(\machT\), with respect to a return predicate \(R \subset \MACHS\) that
-reflects the states that are considered to have returned. We abstract over
-\(R\) because what it means to return depends on the state from which
-\(f\) was called. There are three cases in which \(\MPT\) and \(\machT\)
-enjoy trace confidentiality with respect to \(R\).
-
-\begin{itemize}
-  \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
-    {\_}{\_}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
-    some final state \(\mach_r'\) and any changes from the initial states are
-    matched between \(\mach_r\) and \(\mach_r'\). In addition their
-    observations are similar.
-    \[\begin{split}
-      \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
-      \last(\MPT) =
-        \mpostate{\mach_r}{\_}{\_} ~ \land ~ R ~ \mach_r \Rightarrow \\
-      & \exists \mach_r' . \last(\machT) = (\mach_r',\_)
-        \land \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land \\
-      & \forall ~ \component . (\mach_e(\component) \not = \mach_r(\component)
-        \lor \mach_e'(\component) \not = \mach_r'(\component)) \Rightarrow
-      \mach_r(\component) = \mach_r'(\component)
-    \end{split}\]
-  \item If \(\MPT\) is infinite, then \(\machT\) is also infinite and their
-  observations are similar.
-    \[\begin{split}
-      \mathit{InfCase} ~ \MPT ~ \machT \triangleq
-      & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
-      & \neg (\exists mo . \last(\machT) = mo)
-        \land \pi_o(\MPT) \eqsim \pi_o(\machT)
-    \end{split}\]
-\item If \(\MPT\) ends prematurely without fulfilling \(R\), then
-  its observations are a prefix of those of \(\machT\).
-    \[\begin{split}
-      \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
-      & \forall \mach_r .
-        \last(\MPT) = \mpostate{\mach_r}{\_}{\_} ~ \land ~ \neg ~ R ~ \mach_r \Rightarrow \\
-      & \pi_o(\MPT) \lesssim \pi_o(\machT)
-    \end{split}\]
-\end{itemize}
-
-Putting it all together:
-
-\[  \propdef{TraceConf} ~ R ~ \MPT ~ \machT \triangleq
-   \mathit{ReturnCase} ~ R ~ \MPT ~ \machT ~ \land
-   \mathit{InfCase} ~ \MPT ~ \machT \land
-    \mathit{FailCase} ~ R ~ \MPT ~ \machT
-\]
-
-In the non-return cases, we only care that the traces exhibit the same
-observable behavior (or a prefix if the trace halts prematurely), as changes
-in the state will not affect the caller absent a return.  In the return case
-we must additionally guarantee that no changes were made based on secret data
-that could corrupt the caller upon return.
-
-Consider the result of modifying line 42 of our running example to $r_4 \leftarrow [\rsp - 9]$.
-Instead of reading the proper argument {\tt v}, {\tt g} now reads directly from the
-secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
-shows \(\mach_e\), the entry state of {\tt g}, with high confidentiality memory
-in {\high} and low confidentiality in {\low}. Below it, the variant state
-\(\mach_e'\) agrees on low confidentiality memory, but all other addresses
-hold $-5$, in {\varied}.  After the last step, the return states \(\mach_r\) and
-\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
-\(\mach_e'\). Addresses 108 and 109 have changed, shown in light {\low}, and while 109
-had changed in a consistent way (independent of the variation), 108 holds either 43 or $-4$, so the trace does not obey trace confidentiality.
-%\rb{Color name macros}
-%\apt{Explain slash notation in effects column.}
-%\rb{Wrote some words to explain after def. 5.4 (variations).}
-
-As another example, changing instruction 30 of the running example to
-$r_5 \leftarrow [\rsp - 6]$ would cause {\tt f} to read the value of
-{\tt main}'s local variable {\tt x} and then immediately output it,
-violating the internal observation clause of the confidentiality
-definition.
-
-\begin{figure}
-  \confidentialityendexample
-  \caption{Failing Trace Confidentiality}
-  \label{fig:confex}
-\end{figure}
-
-To lift trace confidentiality to a system property, we follow the same
-pattern as in stack integrity:
-
-\definition%{Stack Confidentiality}
-%Let $\callmap$ be a call map and $\codemap$ a function map.
-For every initial state \((\mach_0,\pol_0)\), we take the induced trace
-\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\). Then for
-any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
-$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, we take the prefix of the trace
-from the entry point until its corresponding return:
-%
-\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
-%
-We also take the corresponding trace from any variant \(\mach_e'
-\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
-%
-\[\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
-%
-%For all such pairs of traces, \(\propdef{TraceConf} ~
-%\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
-%Trace confidentiality must also hold on the trace as a whole, so that
-%unallocated memory is still secret. Let \(\mach_0'
-%\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
-%\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.
-%
-If trace confidentiality holds for all such traces,
-\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
-\machT_{\mathit{call}}\), and it also holds for the toplevel execution
-trace to account for unallocated memory, \(\propdef{TraceConf} ~
-\emptyset ~ \MPT ~ \machT\) , then the system as a whole enjoys {\em
-  stack confidentiality} with respect to \(\callmap\).
-
 \subsection{Well-bracketed Control Flow}

### 2021-02-18 CheriApproaches.tex
Updated example to be much simpler

@@
 
-Either way the question becomes: should the variant trace be one with a policy active, or no? Let's
-look at a concrete example.
+Lets consider a very simple machine with three addresses and an external feed of instructions
+(instructions are not in memory.) An address can be marked either secret or public. Let's start with
+the following configuration (top rows are addresses marked S or P, bottom are contents):
 
-\begin{verbatim}
-main() {
-  int[2] x = {0,0};
-  int* y = &x+secret();
-  f(&x);
-}
+\begin{tabular}{| c | c | c |}
+  \hline
+  0(S) & 1(P) & 2(P) \\
+  \hline
+  0 & 0 & 0 \\
+  \hline
+\end{tabular}
 
-f() {
-  int** z = guess_y_addr();
-  print(z**);
-}
-\end{verbatim}
+We have a register \(r\), initialized to 0, and an output {\sc Out} that can be loaded to like a register.
+We want to apply our confidentiality model by varying secret locations and determining if programs
+differ in output between the original state and a variant. I'm using the simplified format from our
+examples in the paper. Assume that loads use the address modulo three so that the step operation is
+total. Consider the program:
 
-Here {\tt main} has an array {\tt x} and saves a pointer to one of its elements, but which one is secret.
-Then it passes the address to  {\tt f}, so that {\tt f} can access {\tt x} legally according to sharing.
-{\tt x} has a tag that marks it permissible for anyone to access it, but {\tt f} has actually accessed {\tt x}
-via reading the hidden information in {\tt y}. Suppose that our policy contains an error such that the access
-of {\tt y} is successful - then {\tt f} always ends up printing 0 regardless, so we haven't violated
-confidentiality. But we would if we printed {\tt y} outright or if the values in {\tt x[0]} and
-{\tt x[1]} differed.
+\vspace{\abovedisplayskip}
 
-Consider the variant states that we can have. {\tt x} does not vary, but {\tt y} can point to {\tt x[0]},
-{\tt x[1]}, or a location beyond {\tt x} entirely. In the first two cases, our output is 0,
-so confidentiality is maintained. What do we do with that case? We can:
+\begin{tabular}{l l l}
+  Line & Instr & Description \\
+  \hline
+  1 & \(r \leftarrow [0]\) & Load from address 0 \\
+  2 & {\sc Out} \(\leftarrow [r]\) & Load from the resulting address and output it \\
+  3 & \(r \leftarrow 0\) & Reset \(r\) to 0 \\
+\end{tabular}
+
+\vspace{\belowdisplayskip}
+
+At the end of this program, 0 has been printed. In any variant when address 0 mods to a 1 or a 2, it
+will also print 0. And since the final instruction clears \(r\), the final state is identical as well.
+So we would not say that this program violates confidentiality of address 0, even though it was read.
+
+Now suppose that we also make address 1 secret. The variation might vary both 0 and 1, so that the
+program reads address 1 and outputs a non-zero result. Now, confidentiality is violated, because this
+variation produces a different output. Suppose we consider some possible security properties to
+prevent this from occurring. The property we come up with attaches to the load instruction and,
+if the source register contains an address that is marked secret, fail-stops the machine. But we forget
+to implement a similar rule for loading through immediate values, oops! So the load on line 1 never failstops.
+Let's look at that second start state and the same program:
+
+\begin{tabular}{| c | c | c |}
+  \hline
+  0(S) & 1(S) & 2(P) \\
+  \hline
+  0 & 0 & 0 \\
+  \hline
+\end{tabular}
+
+Now the primary execution will load from address 0, then attempt to do so again and fail-stop. The variant
+might instead try to read from a different location, but we are termination insensitive, so even if it reads
+from 2 (successfully) the property holds. But what about this start state?
+
+\begin{tabular}{| c | c | c |}
+  \hline
+  0(S) & 1(S) & 2(P) \\
+  \hline
+  2 & 0 & 0 \\
+  \hline
+\end{tabular}
+
+Here the primary execution loads from 2 and prints 0. But a variant in which address 0 holds 0 or 1 will
+load from those addresses, which would violate the policy and cause a failstop. So the question is, how do
+we handle those variants? We can:
 
 \begin{enumerate}
-\item Recognize that the variant should fail-stop, and treat that as a violation
-\item Recognize that the variant should fail-stop, and ignore this variant
-\item Do not check for a variant fail-stop, and therefore likely get a violation
+\item Recognize that the variant should fail-stop, and since the primary execution does not,
+  treat that as a violation.
+\item Recognize that the variant should fail-stop, and since we are termination insensitive,
+  treat that as acceptable and look for another variant that actually produces different output.
+\item Do not check whether the variant fail-stops. So when we vary address 0 to 1 and address 1 to 1,
+  the program prints 1. Thus the property is violated.
 \end{enumerate}
 
-I argue for option 2, because either some other variation that doesn't create an illegal pointer
-will be a violation, or the particular trace has not leaked any information. In this case,
-the error will be detected if we ever run the comparable case where {\tt x = {0,1}}, but if the compiler
-somehow only produced traces where this issue is invisible, we need not consider it. Still, option 1
-could be fine, since our policies are meant to prevent {\tt y} from being read in the first place.
+Options 1 and 2, in testing, will let us stop short and either move on to a different test or
+declare a failure. Option 1 seems at odds with the notion of termination-insensitivity; so does option 3
+in its own way, with the added downside that it's hard to predict whether a given case might fail. Option
+2 is similar to only doing variants that maintain the legality of pointers. I've been leaning toward option
+2, and I tend not to like option 3 because our variant run can show behavior that is infeasible in the
+target machine.
 
-Option 3 is clearly bad because if the contents of {\tt y} are used in some other way, its behavior
-could be radically different from the primary trace. Whether it fails or not is virtually unrelated to
-the initial potential violation. This seems very bad.
-
-Considering the same situation in Cheri, we also have the possibility to get additional tag failures
-by varying not just the address pointed to, but the permissions bits on a capability. But the fundamental
-question remains the same: what to do with fail-stop variants.
+This small example illustrates a scenario that can come up in HOPE and Cheri, when pointers and capabilities
+being varied introduces fail-stop behavior in the variant trace when none was present in the primary. This
+is an edge case, of course, because in any such scenario the program could instead print the pointer and thus
+violate the property immediately. But we do need to decide how to approach it. We could imagine a policy that
+prevents printing the pointer directly (via taint-tracking, say) but does not prevent dereferencing it.
 

### 2021-02-22 example1.tex
Filled in first example

\label{sec:running-example}

\begin{figure}[b]
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int x = 42, y = 0;
  y = f(10);
  print (x + y);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
f(z) = {
  print z;
  return (2*z);
}
\end{verbatim}
    \end{minipage}
\caption{A Simple Call}
\label{fig:simple-program}
\end{figure}

\newcommand*{\add}{\textsc{add}}
\newcommand*{\addi}{\textsc{addi}}
\newcommand*{\sw}{\textsc{sw}}
\newcommand*{\lw}{\textsc{lw}}
\newcommand*{\jal}{\textsc{jal}}
\newcommand*{\jalr}{\textsc{jalr}}
%\newcommand*{\rsp}{\textsc{sp}}
\newcommand*{\rra}{\textsc{ra}}
\newcommand*{\rout}{\textsc{out}}

\newcommand*{\tagInstr}{\textsc{instr}}
\newcommand*{\tagCall}{\textsc{call}}
\newcommand*{\tagHa}{\textsc{h1}}
\newcommand*{\tagHb}{\textsc{h2}}
\newcommand*{\tagRa}{\textsc{r1}}
\newcommand*{\tagRb}{\textsc{r2}}
\newcommand*{\tagRc}{\textsc{r3}}
\newcommand*{\tagNoDepth}{\textsc{unused}}
\newcommand*{\tagStackDepth}[1]{\textsc{stack} ~ #1}
\newcommand*{\tagPCDepth}[1]{\textsc{pc} ~ #1}
\newcommand*{\tagSP}{\textsc{sp}}

\newcounter{pcctr}
\setcounter{pcctr}{0}
%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
\newcommand*{\row}[4]{
  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\
}
\newcommand*{\tracerow}[5]{
  #1 & #2 & #3 & #4 & \text{#5}\\
}
\newcommand*{\summary}[6]{
\[
\begin{array}{cccccc}
  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
\end{array}
\]
}
\newcommand{\negate}{\textrm{-}}

\colorlet{lgray}{gray!40}

\newcommand{\instrc}{lgray}
\newcommand{\mainsealc}{cyan}
\newcommand{\fsealc}{green}
\newcommand{\unsealc}{gray}
\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
\newcommand{\fulloutc}{white}

\begin{figure}

%Starting Memory Layout
\begin{center}
\memoryaddrs{25em}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
\memory{10}{\unsealc}[{\makebox[0pt]{Unsealed}}]
~$\cdots$
\\
\end{center}
% Starting Register File
\summary{0}{0}{?}{100}{?}{?}
\vspace*{0.2em}
\ifaftersubmission\bcp{A very small nit: Change
  $\rsp-2$ to $\rsp\mathord{-}2$}\apt{??}\bcp{I meant remove spaces around +
and $-$ when then are used inside brackets.  I think it will look better.
But not critical.}\fi
\[
  \begin{array}{l|l|l|l}
    PC & \mathit{Instruction} & \mathit{Effect} & \\
    \hline
    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 102}
        {Allocate private data for {\tt main}}
    \row{\addi ~ r_4 ~ r_0 ~ 42}{r_4 \leftarrow r_0 + 42}{r_4 \leftarrow 42}
        {Load 42 into $r_4$}
    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp-2] \leftarrow r_4}{[100] \leftarrow 42}
        {Set {\tt x} to 42}
    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[101] \leftarrow 0}
        {Set {\tt y} to 0}
    \row{\addi ~ r_4 ~ r_0 ~ 10}{r_4 \leftarrow r_0 + 10}{r_4 \leftarrow 10}
        {Load 10 in \(r_4\) as argument}
    \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
        {Call ($\jal$ to) {\tt f}}
  \end{array}
  \]
  ~ \\
  ~\\
%  \apt{How's that description for last line?}
%  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
%    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
%  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}
% At Call to G
% Memory Layout
\begin{center}
\memoryaddrs{35em}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed 0}}]%
\memory{8}{\unsealc}[{\makebox[0pt]{Unsealed}}]
~$\cdots$
\\
\end{center}
% Starting Register File
\summary{20}{0}{\thepcctr}{102}{10}{?}
%\vspace*{0.2em}
\setcounter{pcctr}{20}%
% G - Body - 1
\vspace*{0.2em}
  \[
  \begin{array}{l|l|l|l}
    \row{\addi ~ \rsp ~ \rsp ~ + 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow \rra}
        {Allocate space for return address}
    \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[103] \leftarrow 6}
        {Save return address}
    \row{\sw ~ \rout ~ r_4}{[\rout] \leftarrow r_4}{[\rout] \leftarrow 10}
        {Print argument}
        % Zero out z?
    \row{\addi ~ r_4 ~ r_4 ~ r_4}{r_4 \leftarrow r_4 + r_4}{r_4 \leftarrow 20}
        {Double the argument}
    \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 8}
        {Load return address}
    \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 102}
        {Deallocate return address}
    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 6}{Return to {\tt main} ($\jalr$)}
  \end{array}
  \]
  ~ \\
  ~ \\
\begin{center}
\memoryaddrs{25em}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
\memory{10}{\unsealc}[{\makebox[0pt]{Unsealed}}]
~$\cdots$
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\\
\end{center}

\setcounter{pcctr}{6}

% Register File
\summary{\thepcctr}{0}{\thepcctr}{102}{20}{?}
%\vspace*{0.2em}
\[
\begin{array}{l|l|l|l}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 1}{[\rsp - 1] \leftarrow r_4}{[101] \leftarrow 20}
      {Store the result to {\tt y}}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 2}{r_4 \leftarrow [\rsp - 2]}{r_4 \leftarrow 42}
      {Load {\tt x} to \(r_4\)}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 1}{r_5 \leftarrow [\rsp - 1]}{r_5 \leftarrow 20}
      {Load {\tt y} to \(r_5\)}
  \row{\addi ~ r_5 ~ r_4 ~ r_5}{r_5 \leftarrow r_4 + r_5}{r_5 \leftarrow 62}
      {Add {\tt x} to {\tt y}}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 62}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
% Starting Register File
\caption{Detailed Execution Trace
%  \rb{A suggestion for the description of $\jal$ and $\jalr$:
%    call(function label) [if labels are added next to the entrypoints]
%    or call(address of entrypoint) and return()?}
}
\label{fig:simple-trace}
\end{figure}

\Cref{fig:simple-program} shows a high-level view of a running example,
written in C-like pseudo-code.  It consists of a {\tt main} function that
performs a call to a function {\tt f}, which prints its argument, performs
arithmetic, and returns a result. \Cref{fig:simple-trace} shows an
execution trace of the corresponding machine instructions. These might have
been generated from the high-level code by a (rather simple-minded) compiler,
but we focus on the behavior of the machine instructions. Note that for this
simple example, arguments and return values are passed on the stack.

To streamline the example, we use a small instruction set reminiscent of
RISC-V and make a few simplifying assumptions about the organization of
memory.
%
First, as noted above, since our goal is to formalize stack safety, we
ignore the heap and assume that all variables are stack allocated.
%
We also assume that the machine is word- (not byte-) addressed, that program
instructions start from location 0, that the machine communicates with the
outside world through a single memory-mapped output port at address 99 (so
a {\tt print} statement will be compiled to a store to this
location), and that the stack begins at location 100 and grows upwards.
% \leo{Explain more?  How to connect to realistic one? Observations? DMA?
%   Interrupt handlers?}\bcp{Yes.}

A sketch of the initial memory layout appears at the top of
\Cref{fig:simple-trace}: the instructions for {\tt main} begin at 0,
and the (uninitialized) stack begins at 100. The entire stack is currently
{\em unsealed}, meaning that it is eligible to be written to.
Directly below the memory sketch, we summarize the parts
of the register file that are relevant at this point in the example.

The rest of \Cref{fig:simple-trace} traces the execution of the
program, showing the machine state at call and return points.
Instructions are intended to be similar to RISC-V, but are described
in a (hopefully more readable) ad-hoc notation rather than assembler
syntax. The $\jal$ (jump-and-link) instruction used for calls atomically
stores the return address in a register and jumps to a specified new $\PCname$;
the $\jalr$ (jump-and-link-via-register) instruction used for returns jumps
to the address stored in a register. Register $r_0$ is hardwired to zero,
so that, e.g., instruction 3 stores $0$ into address 101.

At the call from {\tt main} to {\tt f}, {\tt main}'s locals (as
designated by the stack pointer) are {\em sealed} and labeled with the depth
of their owner, which is 0. Sealing is the fundamental integrity contract
between the caller and callee: {\tt main} expects that {\tt x} and {\tt y}
will be unchanged when it gets control back. This gives us our simple,
intuitive statements of integrity, in their {\em eager} and {\em ultra-eager}
forms, either of which supports a {\em call rule}.

\begin{itemize}
\item {\bf Eager Integrity}: on entry to a subroutine, whichever components are
  sealed will have the same values when the subroutine returns
\item {\bf Ultra-eager Integrity}: at any step, whichever components are sealed
  will be invariant
\item {\bf Call Rule}: from the \(\jal\) instruction of a call, if the step is
  successful and there is eventually a corresponding return, that return is to
  the appropriate return address and all sealed components are unchanged
\end{itemize}

Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
of the context of its call (with the exception, of course, of its arguments).
We state this in terms of noninterference: if we say ``components \(\components\)
interfere with the call to {\tt f},'' we mean that at some point up to when it returns,
{\tt f} would display different behavior if called from a state with a different
assignment of values to \(\components\). Here behavior encompasses uses of {\tt print},
return values, and changes to state that will persist beyond the return.
For instance, if {\tt f} computed the offset to access {\tt x} and branched on its value
to decide whether to double or triple its argument, a change in {\tt x} could translate
to a change in {\tt f}'s return value, making {\tt x} interfere with {\tt f}.

Note that it is not sufficient to prevent sealed memory from interfering: unsealed
memory that is, from the callee's perspective, uninitialized is also part of the calling
context. Our model therefore does not support an ultra-eager confidentiality property.
We have only eager confidentiality, which supports a {\em transparency rule}. But the
notion of non-interference also lets us weaken integrity to make a {\em lazy} property
for use with lazy tag policies.

\begin{itemize}
\item {\bf Eager Confidentiality}: the state of the stack on entry to a subroutine
  does not interfere with its behavior up to the point that it returns
\item {\bf Transparency Rule}: from the \(\jal\) instruction of a call, if the step is
  successful, the entry state can be arbitrarily scrambled without changing the callee's
  behavior up to return
\item {\bf Lazy Integrity}:  on entry to a subroutine, whichever components are
  sealed will either have the same values when the subroutine returns, {\em or} they will
  not interfere with the caller's behavior up to its own return
\end{itemize}

### 2021-02-22 main.tex
Filled in first example

@@
 
-\section{Running Example}
-\label{sec:running-example}
-
-\begin{figure}[b]
-\hfill
-    \begin{minipage}{.3\textwidth}
-\begin{verbatim}
-main = {
-  int x = 42, y = 0;
-  y = f(0);
-  print (x + y);
-}
-\end{verbatim}
-    \end{minipage}
-    \begin{minipage}{.3\textwidth}
-\begin{verbatim}
-f(z) = {
-  int w = g(17);
-  print z;
-  return (z + w);
-}
-\end{verbatim}
-    \end{minipage}
-    \begin{minipage}{.3\textwidth}
-\begin{verbatim}
-g(v) = {
-  return (v + 1);
-}
-
-\end{verbatim}
-    \end{minipage}
-\caption{A High-level Program With Nested Calls}
-\label{fig:running-program}
-\end{figure}
-
-\newcommand*{\add}{\textsc{add}}
-\newcommand*{\addi}{\textsc{addi}}
-\newcommand*{\sw}{\textsc{sw}}
-\newcommand*{\lw}{\textsc{lw}}
-\newcommand*{\jal}{\textsc{jal}}
-\newcommand*{\jalr}{\textsc{jalr}}
-%\newcommand*{\rsp}{\textsc{sp}}
-\newcommand*{\rra}{\textsc{ra}}
-\newcommand*{\rout}{\textsc{out}}
-
-\newcommand*{\tagInstr}{\textsc{instr}}
-\newcommand*{\tagCall}{\textsc{call}}
-\newcommand*{\tagHa}{\textsc{h1}}
-\newcommand*{\tagHb}{\textsc{h2}}
-\newcommand*{\tagRa}{\textsc{r1}}
-\newcommand*{\tagRb}{\textsc{r2}}
-\newcommand*{\tagRc}{\textsc{r3}}
-\newcommand*{\tagNoDepth}{\textsc{unused}}
-\newcommand*{\tagStackDepth}[1]{\textsc{stack} ~ #1}
-\newcommand*{\tagPCDepth}[1]{\textsc{pc} ~ #1}
-\newcommand*{\tagSP}{\textsc{sp}}
-
-\newcounter{pcctr}
-\setcounter{pcctr}{0}
-%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
-\newcommand*{\row}[4]{
-  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\
-}
-\newcommand*{\tracerow}[5]{
-  #1 & #2 & #3 & #4 & \text{#5}\\
-}
-\newcommand*{\summary}[6]{
-\[
-\begin{array}{cccccc}
-  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
-\end{array}
-\]
-}
-\newcommand{\negate}{\textrm{-}}
-
-\colorlet{lgray}{gray!40}
-
-\newcommand{\instrc}{lgray}
-\newcommand{\fprivc}{cyan}
-\newcommand{\fcallc}{green}
-\newcommand{\gprivc}{yellow}
-\newcommand{\gcallc}{orange}
-\newcommand{\hprivc}{red}
-\newcommand{\unusedc}{gray}
-\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
-\newcommand{\fulloutc}{white}
-
-\begin{figure}
-
-%Starting Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{25em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{10}{\unusedc}[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\\
-\end{center}
-% Starting Register File
-\summary{0}{0}{?}{100}{?}{?}
-\vspace*{0.2em}
-\ifaftersubmission\bcp{A very small nit: Change
-  $\rsp-2$ to $\rsp\mathord{-}2$}\apt{??}\bcp{I meant remove spaces around +
-and $-$ when then are used inside brackets.  I think it will look better.
-But not critical.}\fi
-\[
-  \begin{array}{l|l|l|l}
-    PC & \mathit{Instruction} & \mathit{Effect} & \\
-    \hline
-    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 102}
-        {Allocate private data for {\tt main}}
-    \row{\addi ~ r_4 ~ r_0 ~ 42}{r_4 \leftarrow r_0 + 42}{r_4 \leftarrow 42}
-        {Load 42 into $r_4$}
-    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp-2] \leftarrow r_4}{[100] \leftarrow 42}
-        {Set {\tt x} to 42}
-    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[101] \leftarrow 0}
-        {Set {\tt y} to 0}
-    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 104}
-        {Allocate {\tt f}'s call frame}
-    \row{\sw ~ \rsp ~ r_0 ~ \negate 2}{[\rsp-2] \leftarrow r_0}{[102] \leftarrow 0}
-        {Set argument to 0}
-    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[103] \leftarrow 0}
-        {Zero out result}
-    \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
-        {Call ($\jal$ to) {\tt f}}
-  \end{array}
-  \]
-  ~ \\
-  ~\\
-%  \apt{How's that description for last line?}
-%  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
-%    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
-%  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}
-% At Call to G
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{43.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
-\memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\\
-\end{center}
-% Starting Register File
-\summary{20}{0}{\thepcctr}{104}{42}{?}
-%\vspace*{0.2em}
-\setcounter{pcctr}{20}%
-% G - Body - 1
-\vspace*{0.2em}
-  \[
-  \begin{array}{l|l|l|l}
-    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 106}
-        {Allocate stack space for {\tt w}}
-    \row{\sw ~ \rsp ~ \rra ~ \negate 2}{[\rsp - 2] \leftarrow \rra}{[104] \leftarrow 8}
-        {Save return address}
-    % Zero out z?
-    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 108}
-        {Allocate {\tt g}'s call frame}
-    \row{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow r_0 + 17}{r_4 \leftarrow 17}
-        {Store 17 into $r_4$}
-    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[106] \leftarrow 17}
-        {Store 17 as the argument {\tt v}}
-    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp - 1] \leftarrow r_0}{[107] \leftarrow 0}
-        {Zero out result}
-    \row{\jal ~ \rra ~ 40}{\rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow 40}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 40}
-        {Call {\tt g}}
-  \end{array}
-  \]
-  ~ \\
-  ~ \\
-% At Call to H
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
-\memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\MemoryLabel{-30.5em}{0.75em}{8}
-\MemoryLabel{-26.5em}{0.75em}{}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-18.5em}{0.75em}{0}
-\\
-\end{center}
-% Starting Register File
-\summary{40}{0}{\thepcctr}{108}{17}{?}
-\setcounter{pcctr}{40}
-\vspace*{0.2em}
-% H - Body
-\[
-\begin{array}{l|l|l|l}
-  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
-      {Allocate space for return address}
-  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
-      {Save return address}
-  \row{\lw ~ r_4 ~ \rsp ~ \negate 3}{r_4 \leftarrow [\rsp - 3]}{r_4 \leftarrow 17}
-      {Load the argument ({\tt v}) into $r_4$}
-  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 18}
-      {Increment $r_4$}\end{array}
-\]
-~\\
-~\\
-% At Middle of H
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{65.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
-\memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{1}{\hprivc}[{\makebox[0pt]{${\texttt g}_{\mathit{priv}}$}}]%
-\memory{1}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\MemoryLabel{-30.5em}{0.75em}{8}
-\MemoryLabel{-26.5em}{0.75em}{}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-18.5em}{0.75em}{0}
-\MemoryLabel{-14.5em}{0.75em}{27}
-\\
-\end{center}
-% Register File
-\summary{40\rb{44}}{0}{\thepcctr\rb{27}}{109}{18}{?}
-\vspace*{0.2em}
-\label{fig:running-trace-a}
-\end{figure}
-
-\begin{figure}
-\[
-\begin{array}{l|l|l|l}
-  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
-      {Store $r_4$ as the return value}
-  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 27}
-      {Load return address}
-  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
-      {Deallocate {\tt g}'s local state}
-  \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 27}
-      {Return to {\tt f} ($\jalr$)}
-  \end{array}
-\]
-
-% After Return from H
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{61.5em}{2em}{\SP}
-\memory{3}{\emptyoutc}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
-\memory{2}{\gprivc}[{\makebox[0pt]{${\texttt f}_{\mathit{priv}}$}}]%
-\memory{2}{orange}[{\makebox[0pt]{{\tt g}'s call frame }}]%
-\memory{2}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
-\MemoryLabel{-38.5em}{0.75em}{0}
-\MemoryLabel{-34.5em}{0.75em}{0}
-\MemoryLabel{-30.5em}{0.75em}{8}
-\MemoryLabel{-26.5em}{0.75em}{}
-\MemoryLabel{-22.5em}{0.75em}{17}
-\MemoryLabel{-18.5em}{0.75em}{18}
-\\
-\end{center}
-% Starting Register File
-\summary{27}{0}{\thepcctr\rb{27}}{108}{18}{?}
-% \vspace*{0.2em}
-
-\setcounter{pcctr}{27}
-% H - Body
-% \vspace*{0.2em}
-\[
-\begin{array}{l|l|l|l}
-  \row{\lw ~ r_4 ~ \rsp ~ 0\rb{\negate 1}}{r_4 \leftarrow [\rsp \rb{- 1}]}{r_4 \leftarrow 18}
-      {Load return value into $r_4$}
-  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
-      {Set {\tt w} to return value}
-  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
-      {Deallocate {\tt g}'s call frame}
-  \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
-      {Load argument {\tt z} to $r_5$}
-  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
-      {Print {\tt z}}
-  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
-      {Add $r_4$ and $r_5$}
-  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
-      {Store $r_4$ as the return value}
-  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
-      {Load return address}
-  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
-      {Deallocate {\tt f}'s local state}
-  \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 8}{Return to {\tt main} ($\jalr$)}
-\end{array}
-\]
-% Return to main
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{43.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\fulloutc}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{2}{\fcallc}[{\makebox[0pt]{{\tt f}'s call frame }}]%
-\memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-51em}{0.75em}{0}
-\MemoryLabel{-47em}{0.75em}{42}
-\MemoryLabel{-43em}{0.75em}{0}
-\MemoryLabel{-39em}{0.75em}{0}
-\MemoryLabel{-35em}{0.75em}{18}
-\\
-\end{center}
-% Register File
-\summary{8}{0}{\thepcctr\rb{8}}{104}{18}{0}
-%\vspace*{0.2em}
-\setcounter{pcctr}{8}
-\[
-\begin{array}{l|l|l|l}
-  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
-      {Load return value to $r_4$}
-  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
-      {Set {\tt y} to the result}
-  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
-      {Deallocate {\tt f}'s call frame}
-  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
-      {Load {\tt x} to $r_5$}
-  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
-      {Add {\tt x} and {\tt y}}
-  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
-      {Print {\tt x + y}}
-\end{array}
-\]
-\vspace*{0.2em}
-% Return to main
-% Memory Layout
-\begin{center}
-\MemoryLabel{4em}{2em}{0}
-\MemoryLabel{21em}{2em}{99}
-\MemoryLabel{31.5em}{2em}{\SP}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
-\hspace*{3pt}
-$\cdots$
-\memory{1}{\fulloutc}[{\makebox[0pt]{Output}}]%
-\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
-\memory{8}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
-~$\cdots$
-\MemoryLabel{-51em}{0.75em}{60}
-\MemoryLabel{-47em}{0.75em}{60\rb{42}}
-\MemoryLabel{-43em}{0.75em}{42\rb{18}}
-\\
-\end{center}
-% Starting Register File
-\summary{\thepcctr}{0}{37\rb{8}}{102}{60}{42}
-\caption{Detailed Execution Trace
-%  \rb{A suggestion for the description of $\jal$ and $\jalr$:
-%    call(function label) [if labels are added next to the entrypoints]
-%    or call(address of entrypoint) and return()?}
-}
-\label{fig:running-trace-b}
-\end{figure}
-
-\ifaftersubmission
-\bcp{We really want readers to understand this trace.  I think we should
-  spend a few more words walking through it in detail.}\leo{Not sure
-  exactly what you mean here. It would take roughly a page to go through
-  each instruction in detail - that's what the last column is for in my mind.
-  Or do you have something else in mind?}
-\bcp{I was thinking about just a few words, not a page.  But let's do this
-  later.}
-\fi
-
-\Cref{fig:running-program} shows a high-level view of a running example,
-written in C-like pseudo-code.  It consists of a {\tt main} function that
-performs a call to some function {\tt f}, which in turn makes a nested call
-to some function {\tt g} before returning, while performing some simple
-arithmetic operations in the process. \Cref{fig:running-trace-b} shows an
-execution trace of the corresponding machine instructions. These might have
-been generated from the high-level code by a (rather simple-minded) compiler,
-but it is only the behavior
-of the machine instructions that we care about
-here---\Cref{fig:running-trace-b} is our actual example.
-
-To streamline the example, we use a small instruction set reminiscent of
-RISC-V and make a few simplifying assumptions about the organization of
-memory.
-%
-First, as noted above, since our goal is to formalize stack safety, we
-ignore the heap
-and assume that all variables are stack allocated.
-%
-We also assume that the machine is word- (not byte-) addressed, that program
-instructions start from location 0, that the machine communicates with the
-outside world through a single memory-mapped output port at address 99 (so
-a {\tt print} statement will be compiled to a store to this
-location), and that the stack begins at location 100 and grows upwards.
-% \leo{Explain more?  How to connect to realistic one? Observations? DMA?
-%   Interrupt handlers?}\bcp{Yes.}
-
-A sketch of the initial memory layout appears at the top of
-\Cref{fig:running-trace-b}: the instructions for {\tt main} begin at 0,
-and the (uninitialized)
-stack begins at 100. Directly below the memory sketch, we summarize the parts
-of the register file that are relevant at this point in the example.
-%
-The rest of \Cref{fig:running-trace-b} traces the execution of the
-program, showing the machine state at call and return points.
-Instructions are intended to
-be similar to RISC-V, but are described in a (hopefully more readable)
-ad-hoc notation
-rather than using assembler syntax. The $\jal$ (jump-and-link)
-instruction used for calls atomically simultaneously the return address in a
-register
-and jumps to a specified new $\PCname$; the $\jalr$ (jump-and-link-via-register)
-instruction used for returns jumps to the address stored in a register.
-Register $r_0$ is hardwired to zero, so that, e.g., instruction 3 stores $0$
-into address 101.
-
-The code itself is straightforward, and should be largely self-explanatory, but
-a few points deserve clarification. Functions return their result value on the
-stack rather than in a register.  Stack frames are built in two parts:
-space for local variables (which are all stack-allocated) and the return address is allocated and initialized at function entry, and deallocated at function exit; space for arguments and return values is allocated and initialized immediately prior to each function call and deallocated on return.
-The return value slot is always initialized
-because some of our enforcement mechanisms will rely on this.
-\ifaftersubmission\apt{this is bogus, but oh well}\fi
-The return address is spilled to the stack even by leaf functions such as
-{\tt g} that could just leave it in $\rra$.
-
-%\apt{Need to clarify to what extent our properties depend on the
-%particular compilation style, calling sequences, etc. illustrated here.}
+\section{Examples and Intuition}
+\input{example1}
 

### 2021-02-24 example1.tex
Narrowed example diagram
Example description to merge

@@
 We state this in terms of noninterference: if we say ``components \(\components\)
-interfere with the call to {\tt f},'' we mean that at some point up to when it returns,
+interfere with {\tt f} up to its return,'' we mean that at some point up to when it returns,
 {\tt f} would display different behavior if called from a state with a different

### 2021-03-01 main.tex
Add lazy and other updates

@@
 identifying information in unused stack space, {\tt f} could still deduce information
-about the secret. So our confidentiality properties also support a {\em transparency rule},
+about the secret. So our confidentiality properties also support a {\em insensitivity rule},
 that each function's behavior is independent of the state of its caller, other than arguments.
@@
 \[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
-\[\mach \stepstoobs{\obs} \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
+\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
 %
@@
 
-\judgmenttwo{\(\mach \stepstoobs{\obs} \mach'\)}{\((\mach, \pol)
+\judgmenttwo{\(\mach \stepstoobs[\obs] \mach'\)}{\((\mach, \pol)
                \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol{\obs}
+            {\(\mpstate{\mach}{\pol} \stepstopol[\obs]
                \mpstate{\mach'}{\pol'}\)}
@@
 and observations produced by repeated application of step. It's defined in
-terms of a coinductive helper, \(\mathit{obs-forward}\):
-
-\judgmenttwo{\(\mach \stepstoobs{\obs'} \mach'\)}
-            {\(\mathit{obs-forward} ~ \mach' ~ \obs' = \machT\)}
-            {\(\mathit{obs-forward} ~ \mach ~ \obs = (\mach,\obs) \machT\)}%
-%
-\judgment{\(\mathit{obs-forward} ~ \mach ~ \tau = \machT\)}
+terms of a coinductive helper, \(\mathit{obsforward}\):
+
+\judgmenttwo{\(\mach \stepstoobs[\obs'] \mach'\)}
+            {\(\mathit{obsforward} ~ \mach' ~ \obs' = \machT\)}
+            {\(\mathit{obsforward} ~ \mach ~ \obs = (\mach,\obs) \machT\)}%
+%
+\judgment{\(\mathit{obsforward} ~ \mach ~ \tau = \machT\)}
          {\(\mach \hookrightarrow \machT\)}
@@
 \judgment{\(\neg \exists \pol' . \mpcstate{\mach}{\pol} \polstep \pol'\)}
-         {\(\mathit{obs-forward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
+         {\(\mathit{obsforward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
            \mpcstate{\mach}{\pol}{\context},\obs\)}
 %
-\judgmenttwo{\(\mpcstate{\mach}{\pol}{\context} \stepstoobs{\obs'} \mpcstate{\mach'}{\pol'}{\context'}\)}
-            {\(\mathit{obs-forward} ~ \mpcstate{\mach'}{\pol'}{\context'} ~ \obs' = \machT\)}
-            {\(\mathit{obs-forward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
+\judgmenttwo{\(\mpcstate{\mach}{\pol}{\context} \stepstoobs[\obs'] \mpcstate{\mach'}{\pol'}{\context'}\)}
+            {\(\mathit{obsforward} ~ \mpcstate{\mach'}{\pol'}{\context'} ~ \obs' = \machT\)}
+            {\(\mathit{obsforward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
               (\mpcstate{\mach}{\pol}{\context},\obs) \machT\)}%
 %
-\judgment{\(\mathit{obs-forward} ~ \mpcstatename ~ \tau = \machT\)}
+\judgment{\(\mathit{obsforward} ~ \mpcstatename ~ \tau = \machT\)}
          {\(\mpcstatename \hookrightarrow \machT\)}
@@
 
-\section{Simple Stack Safety, Formally}
-\label{sec:sss}
+\section{Local State Encapsulation, Formally}
+\label{sec:lse}
 
@@
 
-\[\ann \in \ANNS ::= \callmap + \retmap\]
+\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap\})\]
 \[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option{\ann}\]
@@
 
-\definition Let \(\mpcstatename\) be a reachable state whose next instruction is annotated
-as a call and whose call depth is \(\depth\). Let \(f ~ \mach ~ \context\) hold when \(\context = \_,\depth\),
-and \(\components\) be the set of components \(\component\) such that
-\(\sealcon ~ \pi_\mach(\mpcstatename) ~ \component\) holds. Then the {\em Call Rule} says that if
-\(\mpcstatename \stepstocon \mpcstatename'\), it is sound to assume that
-\(\mpcstatename' \hookrightarrow \MPCT | f\) and if \(\last(\MPCT) = \mach'\), then
+\definition Let \(\mpcstatename\) be a reachable state whose PC value is annotated
+as a call and whose call depth is \(\depth\). Assume \(\mpcstatename \stepstocon \mpcstatename'\).
+Let \(\MPCT\) be the trace ending after the corresponding return, if any, defined by
+\(\mpcstatename' \hookrightarrow \MPCT | f\) with \(f ~ \mach ~ \context\) holding when
+\(\context = \_,\depth\). Finally let \(\components\) be the set of components \(\component\)
+such that \(\sealcon ~ \pi_\mach(\mpcstatename) ~ \component\) holds. Then by the {\em Call Rule},
+it is sound to assume that if \(\last(\pi_\mach(\MPCT)) = \mach'\), then
 \(\pi_\mach(\mpcstatename) ~ \component = \mach' ~ \component\).
@@
 
-\subsection{Eager Confidentiality and the Transparency Rule}
+\subsection{Eager Confidentiality and the Insensitivity Rule}
 
@@
 
-\definition Transparency Rule
+\definition Let \(\mpcstatename\) be a reachable call state. Assume the that the call is successful
+and \(\mpcstatename \stepstocon \mpcstatename'\). Then let \(\npcstatename'\) be a \(\components\)-variant
+where \(\components\) is the set of all addresses in the stack. Then by the {\em Insensitivity Rule},
+if the trace from \(\mpcstatename'\) eventually returns, so does that from \(\npcstatename'\),
+their states after the return have the same differences, and up to return their traces are observationally
+equivalent.
 
@@
 
-\section{Stack Safety with Sharing}
+\subsection{Stack Safety with Sharing}
+\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap\} \cup
+\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
+\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
 
@@
 
+\subsection{Local State Encapsulation}
+
+\definition A system enjoys {\em local state encapsulation} if it enjoys eager integrity and
+eager confidentiality.
+
+\definition A system enjoys {\em lazy local state encapsulation} if it enjoys lazy integrity
+and eager confidentiality.
+
+We do not give a confidentiality equivalent to lazy integrity, because the family of lazy
+enforcement mechanisms that we consider under the HOPE system in fact enforce eager confidentiality.
+A policy based on taint-tracking might require such a lazy confidentiality, and it would be constructed
+by replacing the ``same-difference'' condition with one that requires any differences to produce
+observationally equivalent states. But for our purposes it is irrelevant.
+
 \section{Coroutine Safety}
-
-\subsection{Local State Encapsulation}
-\label{sec:lse}
+\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap\} \cup
+\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
+\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
 
@@
 and may still hold on a wide range of programs that violate expected
-control flow. For instance, if {\tt f} managed to jump in the middle of
-the code of {\tt g} with its call, confidentiality and integrity would
-still be satisfied.
-%
-However, that is still unwanted behavior: for example, a caller (or a
-callee) should not be able to bypass a password check by jumping (or
-returning) in the middle of a block of code --- even if such an execution
-would ostensibly satisfy local state encapsulation.
+control flow. The archetypal example of this is the {\em awkward example},
+in which an adversarial callee stores a return address and returns to it
+on a subsequent call, thus violating the expected control flow of the
+caller without actually violating the caller's integrity. But it also
+encompasses simpler violations, such as a function bypassing a password
+check by jumping (or returning) in the middle of a block of code.
 %
@@
 
-To formally capture this aspect of stack safety, we introduce three
-security properties that rely on the underlying knowledge of the
-program structure (call map $\callmap$, function map $\codemap$, return map $\retmap$, and entry
-map $\entmap$): {\em control separation} ensures the validity of transition
-sources between functions, {\em entry integrity} ensures the validity
-of call targets, and {\em return integrity} ensures that of
-return targets.
+To formally capture this aspect of stack safety, we introduce some more
+security properties. {\em Control separation} restricts transitions between
+functions to follow the interfaces set out by call, return, and yield, forming
+a foundation for all of the other properties that rely on those interfaces for
+meaning. {\em Entry integrity} ensures the validity of call targets,
+{\em return integrity} ensures that of return targets, and
+{\em yield-back integrity} ensures that of yield targets.
+
+Recall that our properties use a compiler-derived code-map determine
+which addresses are instructions and which functions they belong to, as well
+as annotating them as calls, returns, and yields, as well as sharing information.
+We now extend it with an annotation for valid entry points:
+\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap, \entrymap\} \cup
+\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
+\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
 
 \definition%{Control Separation}
-A system enjoys {\em control separation}
+A system enjoys {\em control separation} when each reachable step \(\mpcstatename\)
+where 
 % with respect to a function map
@@
 notion of control flow integrity.
-
-\newcommand*{\Unmatched}{\mathit{um}}
-
-Return integrity captures the expected behavior of marked returns with
-respect to their call points. Intuitively, it says that the marked return
-matching a call $\mach_c$ should satisfy $\ret{\mach_c}$.
-Observe that, given an MP-trace $\mpstate{\mach_c}{\pol}\MPT$, if $\mach_c$ is
-a call state then the matching return will be the first \emph{unmatched} return
-in $\MPT$. So we begin by defining an auxiliary ``unmatched'' relation between
-two traces.
-
-\definition
-We write $\Unmatched~\MPT~\MPT'$ if $\MPT'$ is
-obtained from $\MPT$ by dropping elements until we reach the first unmatched return
-(according to the return map).  Formally, $\Unmatched$ is defined coinductively as follows:
-\begin{itemize}
-\item A trace that starts with a return is related to itself:
-
-\begin{center}
-\begin{minipage}{.3\textwidth}
-  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
-              {\(\retmap(\mach(\PCname))\)}
-              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
-\end{minipage}
-\hspace*{0.1\textwidth}
-\begin{minipage}{.3\textwidth}
-  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
-              {\(\retmap(\mach(\PCname))\)}
-              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}
-\end{minipage}
-\end{center}
-
-\item If a trace \(\MPT\) starts with a non-call, non-return state,
-  its unmatched return must come later:
-
-  \judgment{\(\callmap(\mach(\PCname)) = \bot
-              \hspace{.5cm} \neg \retmap(\mach(\PCname))
-              \hspace{.5cm}\Unmatched ~ \MPT ~ \MPT'\)}
-           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}
-
-\item Finally, if a trace starts with a call, then we must discard the first
-  unmatched return from its tail to get to the real one:
-
-  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
-             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
-             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
-              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
-            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
-\end{itemize}
-\ifspace
-\apt{Could cut formal definition for space.}
-\fi
 
@@
   \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
-  \mach_1 \stepstoobs{\obs_1} \mach_1' \wedge \mach_2 \stepstoobs{\obs_2} \mach_2' \Rightarrow \\
+  \mach_1 \stepstoobs[\obs_1] \mach_1' \wedge \mach_2 \stepstoobs[\obs_2] \mach_2' \Rightarrow \\
 &    \obs_1 = \obs_2 \wedge \forall \component . ~
@@
 machine in the variant stack.  Given a machine that steps $(\mach,
-\pol) \stepstopol{\obs} (\mach', \pol')$, say that a variant
+\pol) \stepstopol[\obs] (\mach', \pol')$, say that a variant
 stack $\vs$ satisfies the \emph{variant stack test} if for each
 variant element $\vm{C}$ (which is a variation of $\mach$ for $C$),
-$\vm{C}$ can take a step, denoted $\vm{C} \stepstoobs{\obs} \mach'$,
-while satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
+$\vm{C}$ can take a step, denoted \(\vm{C} \stepstoobs[\obs] \mach'\),
+while satisfying the stack integrity test \(\testInt{C}{\mach}{\mach'}\)
 and stack confidentiality test $\testConf{\mach}{\mach'}$ with regard
@@
   %% $\hookrightarrow$ and give the rest of the induced trace?}
-  \[(\mach, \pol) \stepstopol{\obs} (\mach', \pol') \wedge \callmap(\mach) = n,\]
+  \[(\mach, \pol) \stepstopol[\obs] (\mach', \pol') \wedge \callmap(\mach) = n,\]
 %
@@
   If the machine $\mach$ takes a \emph{return step},
-  \[(\mach, \pol) \stepstopol{\obs} (\mach', \pol') \wedge \retmap(\mach),\]
+  \[(\mach, \pol) \stepstopol[\obs] (\mach', \pol') \wedge \retmap(\mach),\]
   %

### 2021-03-01 main.tex
Cleaned up formalism

@@
 \begin{figure}[b]
-    \begin{minipage}{.3\textwidth}
+    \begin{minipage}{.4\textwidth}
 \begin{verbatim}
-main = {
+int main() {
   int x = 42;
@@
     \end{minipage}
-    \begin{minipage}{.3\textwidth}
+    \begin{minipage}{.4\textwidth}
 \begin{verbatim}
-main = {
+int main() {
   int x = 1;
@@
     \end{minipage}
-    \begin{minipage}{.3\textwidth}
+  \caption{Simple Examples}
+  \label{fig:examples1}
+\end{figure}
+
+\begin{figure}[b]
+    \begin{minipage}{.4\textwidth}
 \begin{verbatim}
-main = {
+int main(int arg) {
+  if(arg)
+    g();
+  else
+    h();
+  f();
+}
+\end{verbatim}
+    \end{minipage}
+    \begin{minipage}{.4\textwidth}
+\begin{verbatim}
+int main() {
   int x = 1;
@@
   \end{minipage}
-  \caption{Integrity Examples}
-  \label{fig:intexamples}
+
+  \caption{More Examples}
+  \label{fig:examples2}
 \end{figure}
 
-\begin{figure}[b]
-    \begin{minipage}{.3\textwidth}
-\begin{verbatim}
-main = {
-  int key = get_key();
-  f();
-  use_key(key);
-}
-\end{verbatim}
-    \end{minipage}
-    \begin{minipage}{.3\textwidth}
-\begin{verbatim}
-main = {
-  int x = get_secret();
-  if(x)
-    g();
-  else
-    h();
-  f();
-}
-\end{verbatim}
-    \end{minipage}
-    \begin{minipage}{.3\textwidth}
-\begin{verbatim}
-main = {
-  int x = 0;
-  int y = get_secret();
-  f();
-  print(x);
-}
-\end{verbatim}
-  \end{minipage}
-  \caption{Confidentiality Examples}
-  \label{fig:confexamples}
-\end{figure}
-
 \paragraph*{The Use of Stack Safety}
@@
 When does the programmer rely on stack safety? The most natural example is the first
-in figure \ref{fig:intexample}, in which {\tt main} calls stores 42 to a local variable {\tt x},
+in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
 calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
@@
 Alongside integrity, we have {\em confidentiality}, in which callers' data is protected
-from leakage by their callees. Consider the first example in figure \ref{fig:confexamples},
-in which {\tt main} recieves a key, then calls {\tt f}. If {\tt f} is an untrusted function,
-it would be bad if it could just read {\tt key}. But there are more subtle examples,
-such as the second one, where a secret value controls which function of {\tt g} and
-{\tt h} are called prior to calling {\tt f}. Here, if {\tt g} or {\tt h} leaves behind
-identifying information in unused stack space, {\tt f} could still deduce information
-about the secret. So our confidentiality properties also support a {\em insensitivity rule},
+from leakage by their callees. Consider again the first example in figure \ref{fig:examples1},
+in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
+programmer had initialized {\tt x} after calling {\tt f} rather than before; such a change
+should not alter the behavior of {\tt main}, but if {\tt f} can directly read {\tt x}, it
+may. But there are more subtle examples, such as the first in figure \ref{fig:examples2},
+where {\tt main} may call {\tt g} or {\tt h} are called prior to calling {\tt f}, depending on
+its argument. Here, if {\tt g} or {\tt h} leaves behind identifying information in unused stack
+space, {\tt f} could still deduce information about the argument and behave differently, violating
+the abstraction of its call. So our confidentiality properties also support a {\em insensitivity rule},
 that each function's behavior is independent of the state of its caller, other than arguments.
-This also makes it easier, though not trivial, to prove idempotence: the channels through
+This also makes it easier, though not trivial, to prove a function's idempotence: the channels through
 which a callee can remember its own history are limited to those outside the stack.
@@
 {\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
-callers from their callees. Consider the last example in figure \ref{fig:intexample},
+callers from their callees. Consider the last example in figure \ref{fig:examples2},
 a variant of the famous ``awkward example,'' in which {\tt f} is called twice and may not
@@
 {\tt f}'s first activation stores its return address somewhere accessible, the second
-activation can return to before {\tt x} is negated, cause {\tt x} to be set to
+activation can return to before {\tt x} is negated, causing {\tt x} to be set to
 1 again before a third activation that would return normally. Then {\tt main} would print 1.
+Once again this is rather unexpected behavior.
 
@@
 to jump back into the caller's code at an arbitrary point, either directly (via a jump
-instruction) or indirectly (by corrupting stack data and returning). It also maintains
-the integrity of the stack structure; otherwise a callee could leak data in the third
-example of \ref{fig:conf-example} by failing to properly restore the stack pointer.
-As {\tt x} is accessed by its offset from the stack pointer, an adversary could leave
-the stack pointer so that {\tt x}'s offset now points to {\tt y}, and trick {\tt main}
-into printing its secret.
+instruction) or indirectly (by corrupting stack data and returning).
 
@@
 \(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
-\(\TRACEELEMS\). We will mostly be concerned with sequences of pairs of machine states
-and observations, called \emph{machine traces} and denoted \(\machT \in \MOTRACE\),
-and with sequences of 4-tuples of a machine state, a policy state, a context, and an
-observation, called \emph{MPC-traces} and denoted \(\MPCT \in \MPCOTRACE\).
+\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
+called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\), and of observations.
 
@@
 
-The ``trace-of'' operator, written \(\mach \hookrightarrow \machT\),
-coinductively relates a machine state with the trace of machine states
-and observations produced by repeated application of step. It's defined in
-terms of a coinductive helper, \(\mathit{obsforward}\):
-
-\judgmenttwo{\(\mach \stepstoobs[\obs'] \mach'\)}
-            {\(\mathit{obsforward} ~ \mach' ~ \obs' = \machT\)}
-            {\(\mathit{obsforward} ~ \mach ~ \obs = (\mach,\obs) \machT\)}%
-%
-\judgment{\(\mathit{obsforward} ~ \mach ~ \tau = \machT\)}
-         {\(\mach \hookrightarrow \machT\)}
-\noindent
-(We use juxtaposition of an element and a sequence to represent ``cons.'')
-Since the step function is total, if \(\mach \hookrightarrow \machT\),
-then \(\machT\) must be infinite.
-
-In our running example in \Cref{fig:running-trace-b} we can
-see a depiction of (a prefix of) such a trace: starting from an initial machine
-state at the very top, each instruction leads to a new machine state
-(the diff of which, compared to the previous one, appears in the
-``{\textit{Effect}'' column), and the majority of observations are $\tau$, except
-for instructions 31 and 13, which are writes to $\rout$
-where the observations are $0$ and $60$, respectively.
-
-In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
-to relate a machine-policy-context state triple to the MPCO-trace
-induced by the step function.
+The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
+coinductively relates an initial state with the trace of states
+produced by repeated application of step, and potentially cut short by
+the policy component.
 
@@
 
-We will also use \(\pi_m\), \(\pi_p\), \(\pi_c\), and \(\pi_o\) to project the components out of an
-MPC state, and overload them to project a trace of the components out of an MPC trace.
+\paragraph*{Projections}
+
+We project out the machine state of an MPC-state with \(\pi_\mach\), the
+policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
+these operators to project out of an machine-trace a trace of machine states,
+policy states, and contexts, respectively.
+
+\paragraph*{Observations-Of}
+
+The ``observations-of'' operator, written \(\pi_\obs\), takes a machine trace and
+creates a trace of observations corresponding to the observations of each step in
+the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
+initial observation, a machine-trace, and an observation trace:
+
+\judgmentthree{\(\pi_\mach ~ \mpcstatename \stepstoobs[\obs'] \pi_\mach ~ \mpcstatename'\)}
+              {\(\mpcstatename' \hookrightarrow \MPCT\)}
+              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
+              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
+\judgment{}
+         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}
+
+Then to take the observations from any particular trace, we start with a \(\tau\) observation:
+
+\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
+         {\(\pi_\obs ~ \MPCT = \obsT\)}
+\noindent
+(We use juxtaposition of an element and a sequence to represent ``cons.'')
+
+In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
+to relate a machine-policy-context state triple to the MPCO-trace
+induced by the step function.
 
@@
 We now extend it with an annotation for valid entry points:
-\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap, \entrymap\} \cup
+\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
 \{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
@@
 \definition%{Control Separation}
-A system enjoys {\em control separation} when each reachable step \(\mpcstatename\)
-where 
-% with respect to a function map
-%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
-if, for any initial state \((\mach_0, \pol_0)\) and any adjacent pair of states
-\(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
-in the induced trace from \(\mpstate{\mach_0}{\pol_0}\),
-if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
-then \(\mpstate{\mach_1}{\pol_1}\) is either a call \(\callmap(\mach_1(\PCname))\)
-or a return \(\retmap(\mach_1(\PCname))\).
-
-\smallskip
-
-For example, consider changing line 5 of the running example to a $\jalr$,
-$\PCname \leftarrow r_4$, with effect $\PCname \leftarrow 42$, causing {\tt main}
-to jump into the middle of {\tt g}.
-This violates control separation because the instruction that causes a
-function switch is not a valid call (or a return) according to the
-call map. On the other hand we do permit functions to jump freely internally.
+
+Let \(\mpcstatename\) be a reachable state whose program counter points to code in
+function \(\funid\), and \(\mpcstatename \stepstopol \mpcstatename'\) where
+\(\mpcstatename'\) is state whose program counter points to a different function
+\(\funid'\). A system enjoys {\em control separation} when all such \(\mpcstatename\)
+are annotated \(\callmap\), \(\retmap\), or \(\yieldmap\).
 
 \definition%{Entry Integrity}
-A system enjoys {\em entry integrity}
-%with respect to a call map
-%\(\callmap\) and entry map \(\entmap\)
-if, for any initial state \((\mach_0, \pol_0)\)
-and any n-transition from \((\mach_c,\pol_c)\) to
-\((\mach_e, \pol_e)\) in the induced trace from
-\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a valid entry:
-\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]
-
-The changed program described just above also violates entry integrity,
-because the target (instruction 42) of the new $\jalr$ is not an established
-entry point according to the entry map. Our properties do not enforce which
-call points jump to which entry points, however; that would require a stronger
-notion of control flow integrity.
+A system enjoys {\em entry integrity} when, for any reachable \(\mpcstatename\) and
+\(\mpcstatename'\) such that \(\mpcstatename\) is annotated \(\callmap\) and
+\(\mpcstatename \stepstopol \mpcstatename'\), \(\mpcstatename'\) is annotated \(\entmap\).
 
 \definition%{Return Integrity}
-A system enjoys {\em return integrity} if all of its returns match prior calls
-and return to the appropriate location.
-For any initial state and its induced trace \((\mach_0\, \pol_0) \hookrightarrow \MPT\), two
-conditions must hold.
-\emph{(i)} \(\MPT\) must not have an unmatched return, i.e., \(\Unmatched ~ \MPT ~ \MPT_{\mathit{tail}}\)
-must not hold for any \(\MPT_{\mathit{tail}}\).
-%
-\emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
-corresponding to a call in \(\MPT\), if
-\(\MPT_{\mathit{call}}\) is the suffix of the trace starting at the call, and
-\(\mpstatename = \mpostate{\mach_r}{\_}{\_}\)
-is the first unmatched return such that
-\( \Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{tail}}\) and \(\head{(\MPT_{\mathit{tail}})} = \mpstatename\),
-then $\mach_r$ must correspond to a return state for $\mach_c$, i.e.,
-\(\ret{\mach_c}\ \mach_r\) holds.
-
-For example, if we modified the running example by adding
-an instruction $ \rra \rightarrow \rra - 19$
-between 45 and 46, this would cause the return from {\tt g} to go to instruction 8 and
-return to {\tt main} instead of {\tt f}. While the instruction that
-causes the return would still be a valid return (satisfying control
-separation), the state after the return would not correspond to an
-unmatched return for {\tt f}'s call to {\tt g}, violating return
-integrity.
+A system enjoys {\em return integrity}... [this one needs to talk about context.]
+
+\definition%{Yield-back Integrity}
+A system enjoys {\em yield-back integrity}... [same.]
 
@@
 
-\subsection{A Conservative Policy}
+\subsection{A Conservative Tag Policy}
 \label{sec:conservative}
@@
 
-\subsection{A Stronger Lockstep Stack Safety Property}
-\label{sec:lockstep}
-
-%% \leo{Maybe lockstep?}
-
-This conservative policy satisfies not only the stack safety policy of
-\cref{sec:lse-and-wbcf},
-but also a provably (in Coq) stronger one that enforces local state
-encapsulation at every point during execution. To encode that
-property, we not only keep track of the execution of one machine
-(whose trace we are trying to characterize as safe), but of an entire
-stack of machine variations corresponding to nested calls of the
-original machine. For each step of the original machine, we take a
-step in every variant machine (thus, keeping them in lockstep
-execution) and enforce a single-step variant of integrity and
-confidentiality tests. In addition, every time a call is made we push
-a variation of the original machine state at the time of that call on
-top of the variant stack (just like the stack-confidentiality property
-enforced trace confidentiality for every call transition), and every time
-the machine returns from a call we pop the top variation (to account
-for stack confidentiality taking the prefix of the transition trace
-until its return).
-
-We start by defining the integrity and confidentiality tests, as well
-as the variant stack, before giving the full definition of lockstep
-stack safety.
-
-\newcommand{\vse}{\mathit{vse}}
-\newcommand{\VSES}{\mathit{VSE}}
-\newcommand{\vs}{\mathit{vs}}
-\newcommand{\VSS}{\mathit{VS}}
-
-\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
-\newcommand*{\testInt}[3]{\ensuremath{\mathit{int}_T ~ #1 ~ #2 ~ #3}}
-\newcommand*{\testConf}[2]{\ensuremath{\mathit{conf}_T ~ #1 ~ #2}}
-\newcommand*{\testVS}[2]{\ensuremath{\mathit{\vs}_T ~ #1 ~ #2}}
-\newcommand*{\testProp}[3]{\ensuremath{\mathit{stack}_T ~ #1 ~ #2 ~ #3}}
-
-\definition
-  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
-  satisfy the \emph{stack integrity test} if their high-integrity components
-  match:
-  \[
-    \testInt{\contour}{\mach}{\mach'} \triangleq
-    \forall \component . ~
-    \contour(\component) = \HIGHINT \rightarrow
-    \mach(\component) = \mach'(\component)
-    \]
-    \ifaftersubmission
-\apt{We could have used this to define trace integrity in the first place...}
-\fi
-\definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
-\emph{stack confidentiality test} if the machines they stepped to
-have their changed components coincide and the observations during
-these steps are equal:
-\[
-\begin{array}{rl}
-  \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \triangleq &
-  \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
-  \mach_1 \stepstoobs[\obs_1] \mach_1' \wedge \mach_2 \stepstoobs[\obs_2] \mach_2' \Rightarrow \\
-&    \obs_1 = \obs_2 \wedge \forall \component . ~
-   (     \mach_1(\component) \neq \mach_1'(\component)
-    \vee \mach_2(\component) \neq \mach_2'(\component))
-    \rightarrow \mach_1'(\component) = \mach_2'(\component)
-\end{array}
-\]
-\ifaftersubmission
-\apt{And the last clause here could have been used for trace confidentiality...}
-\fi
-
-\definition The testing property keeps track of a variation machine
-for every nested call performed so far on a {\em variant stack}: a
-list of machine states, each annotated with its own contour (written
-as $\vm{C}$), and each being a variation of the actual machine
-according to $C$.%
-%
-\footnote{The full formal definition in Coq also records additional
-  information (the initial machine state at the call, its initial
-  variation, the contour, and a return predicate) which we elide
-  here for simplicity of exposition.}
-%%
-%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
-%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
-%  $$\vs \in \VSS$$
-%%
-%\iflater
-%\leo{This can be elided}
-%  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
-%  respect to it if the stack is nonempty, the return map of the bottom element
-%  is the false property, and for each element $\vse \in \vs$:
-%  \begin{itemize}
-%  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
-%  \item $\mach$ is in the trace $\machT$ of its initial machine,
-%    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
-%  \item Its variant state $\mathit{curr\_variant}(\vse)$
-%    is in the trace of its initial variant,
-%    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
-%  \item The return function is in the return map if not the last element. \rb{TODO}
-%  \item The registers of its contour $\contour(\vse)$ are public.
-%  \end{itemize}
-%  \fi
-
-%% \definition
-%% %  Let $\callmap$ a call map and $\codemap$ a function map.
-%%   Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
-%%   variant stack $\vs$, we say that an MP-trace induced from $(\mach,
-%%   \pol)$ satisfies the \emph{strong stack safety} at a given call
-%%   depth as follows.
-%% % Refactor repetitive notation before case split
-%%   In all cases that let $(\mach, \pol) \hookrightarrow_P
-%%   (\mach', \pol', \obs)$.
-%%   %and assume (unless the step is a return) that the
-%%   %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
-%%   %machines can step.
-
-%% %  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
-%% %  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
-%% %  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
-%% %  testing property} is satisfied at call depth $d$, and write
-%% %  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:
-
-%%   \begin{itemize}
-
-%%   \item
-%% %
-%%     If the machine takes a \emph{call step} and $\contour$ is the contour at
-%%     the call then,
-%% %
-%%     (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
-%%     satisfied,
-%% %
-%%     (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
-%%     satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-%%     is satisfied with resepct to an arbitrary $\contour$-variant
-%%     of $\mach$,
-%% =======
-We also need two auxiliary notations that relate to variant stacks.
-%
-First, lockstep stack safety will check that the basic tests above hold for every
-machine in the variant stack.  Given a machine that steps $(\mach,
-\pol) \stepstopol[\obs] (\mach', \pol')$, say that a variant
-stack $\vs$ satisfies the \emph{variant stack test} if for each
-variant element $\vm{C}$ (which is a variation of $\mach$ for $C$),
-$\vm{C}$ can take a step, denoted \(\vm{C} \stepstoobs[\obs] \mach'\),
-while satisfying the stack integrity test \(\testInt{C}{\mach}{\mach'}\)
-and stack confidentiality test $\testConf{\mach}{\mach'}$ with regard
-to the original machine state. We will write $\testVS{(\mach,
-  \pol)}{\vs}$ for this compound check.
-%\apt{I'm having trouble here and below with
-%  the scoping of $C$. Each variant stack is for one particular $C$, right?
-%  So why not write $vs_C$ rather than $vm_C$ ?}
-% LEO: Each machine has its own contour. Clarified above, hopefully
-%% >>>>>>> 1886f7a7bf858ce0c9518e407d9481518a4f028d
-%
-Second, we will write $\vs \stepsto \vs'$ if every variant machine
-in $\vs$ steps to the corresponding machine in $\vs'$.
-%
-
-We are now ready to define lockstep stack safety coinductively.
-
-\definition Given a state $(\mach, \pol) \in \MPS$ and a variant stack
-$\vs$ in which every element $\vm{\contour}$ is a variation of $\mach$
-with respect to its annotated contour $\contour$, we say that an
-MP-trace induced from $(\mach, \pol)$ satisfies the \emph{trace
-  lockstep stack safety}, written $\testProp{}{\MPT}{\vs}$, if the
-following conditions are satisfied:
-
-\begin{itemize}
-
-\item
-  If the machine takes a \emph{non-call, non-return step}
-  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)\MPT,\]
-  then the variant stack test must be satisfied
-  \[\testVS{(\mach, \pol)}{\vs},\]
-  and trace lockstep stack safety must be coinductively satisfied for
-  the remainder of the trace and the variant stack $vs'$ that $vs$
-  steps to ($\vs \stepsto \vs'$):
-  \[ \testProp{}{((\mach',\pol',\obs)\MPT)}{\vs'}\]
-
-\item
-  If the machine takes a \emph{call step} and $\contour$ is the contour at
-  the call
-  %% \rb{Corrected the step function here, or should it be
-  %% $\hookrightarrow$ and give the rest of the induced trace?}
-  \[(\mach, \pol) \stepstopol[\obs] (\mach', \pol') \wedge \callmap(\mach) = n,\]
-%
-  then the variant stack test must be
-  satisfied
-  \[\testVS{(\mach, \pol)}{\vs},\]
-%
-%    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
-%    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
-%    is satisfied with regard to an arbitrary $\contour$-variant
-%    of $\mach$,
-%
-  and trace lockstep stack safety must be coinductively satisfied
-  for the remainder of the trace and for every variant stack
-  that is the result of adding a variation of $\mach'$ with
-  respect to $C$ at the top of the stack $\vs'$ that $\vs$ steps
-  to:
-  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\vm{\contour} : \vs')}\]
-
-\item
-%
-  If the machine $\mach$ takes a \emph{return step},
-  \[(\mach, \pol) \stepstopol[\obs] (\mach', \pol') \wedge \retmap(\mach),\]
-  %
-  then the top of the variant stack also returns,
-  the tail of $\vs$ satisfies the stack safety test
-  \[\testVS{(\mach, \pol)}{\mathit{tail}(\vs)},\]
-  %
-  and trace lockstep stack safety must be coinductively satisfied for
-  the remainder of the trace and for the tail of the variant stack
-  that $\vs$ steps to:
-  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\mathit{tail}(\vs'))}\]
-
-\item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
-    (\mach, \pol, \tau)$, the property is satisfied for the singleton trace:
-    $\testProp{}{[(\mach, \pol)]}{\vs}$.
-
-%  \item If the machine takes a \emph{simultaneous return-call step}, that is,
-%    the instruction to which execution returns immediately after completing a
-%    call is itself a call instruction, the \emph{call step} and \emph{return
-%    step} cases are combined.
-
-  \end{itemize}
-
-\definition
-%
-A system satisfies {\em lockstep stack safety} if, for every initial machine $(\mach,\pol)$,
-the trace induced from $(\mach, \pol)$ satisfies trace lockstep stack safety with
-respect to the empty stack.
-
-\medskip
-\textsc{Theorem.}
-Lockstep stack safety implies stack safety.
-\smallskip
-
-The proof is a fairly involved coinductive argument
-(around 3K LoC in Coq) on the definition of trace lockstep stack safety,
-using an auxiliary depth-index to ensure the proper alignment of
-calls and returns.
+\section{Extension to Cheri}
+
+\section{Extension to Software-based Enforcement}
 

### 2021-03-02 main.tex
Reorganized structure, fixed powerset
Edits to section 4

@@
 \input{examples}
+
+\newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}
 
@@
 pair of a machine state and an {\em observation}, which may be either
-silent (written \(\tau\)) or a word of data\ifaftersubmission\bcp{This is a rather concrete
-  choice!  Why not leave the set of observations abstract?}\fi that the machine
+silent (written \(\tau\)) or a word of data that the machine
 communicates to the outside world:
+\ifaftersubmission\bcp{This is a rather concrete
+  choice!  Why not leave the set of observations abstract?}\fi
 %
@@
 
-\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol{\obs} \mpstate{\mach'}{\pol'}\)}
+\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
             {\(\mach, \context \constep \context'\)}
-            {\(\mpcstate{\mach}{\pol}{\context} \stepstocon{\obs}
+            {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
                \mpcstate{\mach'}{\pol'}{\context'}\)}
@@
 produced by repeated application of step, and potentially cut short by
-the policy component.
-
-\judgment{\(\neg \exists \pol' . \mpcstate{\mach}{\pol} \polstep \pol'\)}
-         {\(\mathit{obsforward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
-           \mpcstate{\mach}{\pol}{\context},\obs\)}
-%
-\judgmenttwo{\(\mpcstate{\mach}{\pol}{\context} \stepstoobs[\obs'] \mpcstate{\mach'}{\pol'}{\context'}\)}
-            {\(\mathit{obsforward} ~ \mpcstate{\mach'}{\pol'}{\context'} ~ \obs' = \machT\)}
-            {\(\mathit{obsforward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
-              (\mpcstate{\mach}{\pol}{\context},\obs) \machT\)}%
-%
-\judgment{\(\mathit{obsforward} ~ \mpcstatename ~ \tau = \machT\)}
-         {\(\mpcstatename \hookrightarrow \machT\)}
+the policy component. The {\sc Default} rule applies only if the other does not.
+(We use juxtaposition of an element and a trace to represent ``cons.'')
+
+\begin{minipage}{.4\textwidth}
+\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
+            {\(\mpcstatename' \hookrightarrow \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
+\end{minipage}
+\begin{minipage}{.4\textwidth}
+\judgment[Default]
+         {}
+         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
+\end{minipage}
+%
 
@@
 
-\judgmentthree{\(\pi_\mach ~ \mpcstatename \stepstoobs[\obs'] \pi_\mach ~ \mpcstatename'\)}
+\judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
               {\(\mpcstatename' \hookrightarrow \MPCT\)}
@@
 \judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
-         {\(\pi_\obs ~ \MPCT = \obsT\)}
+         {\(\pi_\obs(\MPCT) = \obsT\)}
 \noindent
-(We use juxtaposition of an element and a sequence to represent ``cons.'')
 
@@
 
-\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap\})\]
+\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap\})\]
 \[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option{\ann}\]
@@
 \subsection{Stack Safety with Sharing}
-\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap\} \cup
+\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap\} \cup
 \{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
@@
 \section{Coroutine Safety}
-\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap\} \cup
+\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap\} \cup
 \{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
@@
 We now extend it with an annotation for valid entry points:
-\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
+\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
 \{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
@@
 {\em well-bracketed control flow} (with respect to those maps).
+
+\section{Programming Logic Perspective}
 
@@
 
-\section{Extension to Cheri}
-
-\section{Extension to Software-based Enforcement}
+\section{Randomized Testing}
 
@@
     \ifaftersubmission\apt{More blatently, static authorities doesn't work for recursive calls.}\fi
+
+\section{Extension to Cheri}
+
+\section{Extension to Software-based Enforcement}
 

### 2021-03-03 example2.tex
Working on example for sharing

--- 2021-02-22 example1.tex
@@
-\label{sec:running-example}
-
 \begin{figure}[b]
-    \begin{minipage}{.3\textwidth}
+    \begin{minipage}{.25\textwidth}
 \begin{verbatim}
-main = {
-  int x = 42, y = 0;
-  y = f(10);
-  print (x + y);
+int main() {
+  long int x = 42;
+  print f(x);
 }
@@
     \end{minipage}
-    \begin{minipage}{.3\textwidth}
+    \begin{minipage}{.4\textwidth}
 \begin{verbatim}
-f(z) = {
-  print z;
-  return (2*z);
+long int f(long int z) {
+  print g();
+  return (z+z);
+}
+\end{verbatim}
+    \end{minipage}
+    \begin{minipage}{.25\textwidth}
+\begin{verbatim}
+int g() {
+  return 5;
 }
@@
 
-\newcommand*{\add}{\textsc{add}}
-\newcommand*{\addi}{\textsc{addi}}
-\newcommand*{\sw}{\textsc{sw}}
-\newcommand*{\lw}{\textsc{lw}}
-\newcommand*{\jal}{\textsc{jal}}
-\newcommand*{\jalr}{\textsc{jalr}}
-%\newcommand*{\rsp}{\textsc{sp}}
-\newcommand*{\rra}{\textsc{ra}}
-\newcommand*{\rout}{\textsc{out}}
-
-\newcommand*{\tagInstr}{\textsc{instr}}
-\newcommand*{\tagCall}{\textsc{call}}
-\newcommand*{\tagHa}{\textsc{h1}}
-\newcommand*{\tagHb}{\textsc{h2}}
-\newcommand*{\tagRa}{\textsc{r1}}
-\newcommand*{\tagRb}{\textsc{r2}}
-\newcommand*{\tagRc}{\textsc{r3}}
-\newcommand*{\tagNoDepth}{\textsc{unused}}
-\newcommand*{\tagStackDepth}[1]{\textsc{stack} ~ #1}
-\newcommand*{\tagPCDepth}[1]{\textsc{pc} ~ #1}
-\newcommand*{\tagSP}{\textsc{sp}}
-
-\newcounter{pcctr}
-\setcounter{pcctr}{0}
-%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
-\newcommand*{\row}[4]{
-  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\
-}
-\newcommand*{\tracerow}[5]{
-  #1 & #2 & #3 & #4 & \text{#5}\\
-}
-\newcommand*{\summary}[6]{
-\[
-\begin{array}{cccccc}
-  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
-\end{array}
-\]
-}
-\newcommand{\negate}{\textrm{-}}
-
 \colorlet{lgray}{gray!40}
@@
 \newcommand{\mainsealc}{cyan}
+\newcommand{\mainpassc}{magenta}
 \newcommand{\fsealc}{green}
@@
 
-%Starting Memory Layout
 \begin{center}
+\begin{tabular}{l l}
+Entry to {\tt main}
+ &
 \memoryaddrs{25em}
@@
 \memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{10}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+\memory{8}{\unsealc}[{\makebox[0pt]{Unsealed}}]
 ~$\cdots$
-\\
-\end{center}
-% Starting Register File
-\summary{0}{0}{?}{100}{?}{?}
-\vspace*{0.2em}
-\ifaftersubmission\bcp{A very small nit: Change
-  $\rsp-2$ to $\rsp\mathord{-}2$}\apt{??}\bcp{I meant remove spaces around +
-and $-$ when then are used inside brackets.  I think it will look better.
-But not critical.}\fi
-\[
-  \begin{array}{l|l|l|l}
-    PC & \mathit{Instruction} & \mathit{Effect} & \\
-    \hline
-    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 102}
-        {Allocate private data for {\tt main}}
-    \row{\addi ~ r_4 ~ r_0 ~ 42}{r_4 \leftarrow r_0 + 42}{r_4 \leftarrow 42}
-        {Load 42 into $r_4$}
-    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp-2] \leftarrow r_4}{[100] \leftarrow 42}
-        {Set {\tt x} to 42}
-    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[101] \leftarrow 0}
-        {Set {\tt y} to 0}
-    \row{\addi ~ r_4 ~ r_0 ~ 10}{r_4 \leftarrow r_0 + 10}{r_4 \leftarrow 10}
-        {Load 10 in \(r_4\) as argument}
-    \row{\jal ~ \rra ~ 20}{\rra \leftarrow \PCname + 1; ~ \PCname \leftarrow 20}{\rra \leftarrow \thepcctr; ~ \PCname \leftarrow 20}
-        {Call ($\jal$ to) {\tt f}}
-  \end{array}
-  \]
-  ~ \\
-  ~\\
-%  \apt{How's that description for last line?}
-%  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
-%    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
-%  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}
-% At Call to G
-% Memory Layout
-\begin{center}
+\\ \\
+Call to {\tt f}
+ &
 \memoryaddrs{35em}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
+\memory{3}{\instrc}%
 \hspace*{3pt}
 $\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed 0}}]%
-\memory{8}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+\memory{1}{\emptyoutc}%
+\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]%
+\memory{2}{\mainpassc}[{\makebox[0pt]{Passed(0)}}]%
+\memory{4}{\unsealc}[{\makebox[0pt]{Unsealed}}]
 ~$\cdots$
-\\
-\end{center}
-% Starting Register File
-\summary{20}{0}{\thepcctr}{102}{10}{?}
-%\vspace*{0.2em}
-\setcounter{pcctr}{20}%
-% G - Body - 1
-\vspace*{0.2em}
-  \[
-  \begin{array}{l|l|l|l}
-    \row{\addi ~ \rsp ~ \rsp ~ + 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow \rra}
-        {Allocate space for return address}
-    \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[103] \leftarrow 6}
-        {Save return address}
-    \row{\sw ~ \rout ~ r_4}{[\rout] \leftarrow r_4}{[\rout] \leftarrow 10}
-        {Print argument}
-        % Zero out z?
-    \row{\addi ~ r_4 ~ r_4 ~ r_4}{r_4 \leftarrow r_4 + r_4}{r_4 \leftarrow 20}
-        {Double the argument}
-    \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 8}
-        {Load return address}
-    \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 102}
-        {Deallocate return address}
-    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 6}{Return to {\tt main} ($\jalr$)}
-  \end{array}
-  \]
-  ~ \\
-  ~ \\
-\begin{center}
-\memoryaddrs{25em}
-\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
+\MemoryLabel{-37.5em}{0.75em}{0\(\ldots\)}
+\MemoryLabel{-33.5em}{0.75em}{42}
+\MemoryLabel{-29.5em}{0.75em}{0\(\ldots\)}
+\MemoryLabel{-25.5em}{0.75em}{42}
+\\ \\
+Entry to {\tt f} &
+\memoryaddrs{35em}
+\memory{3}{\instrc}%
 \hspace*{3pt}
 $\cdots$
-\memory{1}{\emptyoutc}[{\makebox[0pt]{Output}}]%
-\memory{10}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+\memory{1}{\emptyoutc}%
+\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
+\memory{2}{\mainpassc}[{\makebox[0pt]{Passed(0)}}]%
+\memory{4}{\unsealc}[{\makebox[0pt]{Unsealed}}]
 ~$\cdots$
-\MemoryLabel{-46.5em}{0.75em}{42}
-\MemoryLabel{-42.5em}{0.75em}{0}
+\MemoryLabel{-29em}{0.75em}{0\(\ldots\)}
+\MemoryLabel{-24em}{0.75em}{42}
+\\ \\
+Return to {\tt main}
+ &
+\memoryaddrs{25em}
+\memory{3}{\instrc}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{\emptyoutc}%
+\memory{8}{\unsealc}
+~$\cdots$
+\MemoryLabel{-37.5em}{0.75em}{0\(\ldots\)}
+\MemoryLabel{-33.5em}{0.75em}{42}
+\MemoryLabel{-29.5em}{0.75em}{0\(\ldots\)}
+\MemoryLabel{-25.5em}{0.75em}{42}
 \\
+\end{tabular}
 \end{center}
 
-\setcounter{pcctr}{6}
+\vspace{\abovedisplayskip}
 
-% Register File
-\summary{\thepcctr}{0}{\thepcctr}{102}{20}{?}
-%\vspace*{0.2em}
-\[
-\begin{array}{l|l|l|l}
-  \row{\sw ~ \rsp ~ r_4 ~ \negate 1}{[\rsp - 1] \leftarrow r_4}{[101] \leftarrow 20}
-      {Store the result to {\tt y}}
-  \row{\lw ~ \rsp ~ r_4 ~ \negate 2}{r_4 \leftarrow [\rsp - 2]}{r_4 \leftarrow 42}
-      {Load {\tt x} to \(r_4\)}
-  \row{\lw ~ \rsp ~ r_5 ~ \negate 1}{r_5 \leftarrow [\rsp - 1]}{r_5 \leftarrow 20}
-      {Load {\tt y} to \(r_5\)}
-  \row{\addi ~ r_5 ~ r_4 ~ r_5}{r_5 \leftarrow r_4 + r_5}{r_5 \leftarrow 62}
-      {Add {\tt x} to {\tt y}}
-  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 62}
-      {Print {\tt x + y}}
-\end{array}
-\]
-\vspace*{0.2em}
-% Return to main
-% Memory Layout
-% Starting Register File
 \caption{Detailed Execution Trace
@@
 }
-\label{fig:simple-trace}
+\label{fig:share-trace}
 \end{figure}
-
-\Cref{fig:simple-program} shows a high-level view of a running example,
-written in C-like pseudo-code.  It consists of a {\tt main} function that
-performs a call to a function {\tt f}, which prints its argument, performs
-arithmetic, and returns a result. \Cref{fig:simple-trace} shows an
-execution trace of the corresponding machine instructions. These might have
-been generated from the high-level code by a (rather simple-minded) compiler,
-but we focus on the behavior of the machine instructions. Note that for this
-simple example, arguments and return values are passed on the stack.
-
-To streamline the example, we use a small instruction set reminiscent of
-RISC-V and make a few simplifying assumptions about the organization of
-memory.
-%
-First, as noted above, since our goal is to formalize stack safety, we
-ignore the heap and assume that all variables are stack allocated.
-%
-We also assume that the machine is word- (not byte-) addressed, that program
-instructions start from location 0, that the machine communicates with the
-outside world through a single memory-mapped output port at address 99 (so
-a {\tt print} statement will be compiled to a store to this
-location), and that the stack begins at location 100 and grows upwards.
-% \leo{Explain more?  How to connect to realistic one? Observations? DMA?
-%   Interrupt handlers?}\bcp{Yes.}
-
-A sketch of the initial memory layout appears at the top of
-\Cref{fig:simple-trace}: the instructions for {\tt main} begin at 0,
-and the (uninitialized) stack begins at 100. The entire stack is currently
-{\em unsealed}, meaning that it is eligible to be written to.
-Directly below the memory sketch, we summarize the parts
-of the register file that are relevant at this point in the example.
-
-The rest of \Cref{fig:simple-trace} traces the execution of the
-program, showing the machine state at call and return points.
-Instructions are intended to be similar to RISC-V, but are described
-in a (hopefully more readable) ad-hoc notation rather than assembler
-syntax. The $\jal$ (jump-and-link) instruction used for calls atomically
-stores the return address in a register and jumps to a specified new $\PCname$;
-the $\jalr$ (jump-and-link-via-register) instruction used for returns jumps
-to the address stored in a register. Register $r_0$ is hardwired to zero,
-so that, e.g., instruction 3 stores $0$ into address 101.
-
-At the call from {\tt main} to {\tt f}, {\tt main}'s locals (as
-designated by the stack pointer) are {\em sealed} and labeled with the depth
-of their owner, which is 0. Sealing is the fundamental integrity contract
-between the caller and callee: {\tt main} expects that {\tt x} and {\tt y}
-will be unchanged when it gets control back. This gives us our simple,
-intuitive statements of integrity, in their {\em eager} and {\em ultra-eager}
-forms, either of which supports a {\em call rule}.
-
-\begin{itemize}
-\item {\bf Eager Integrity}: on entry to a subroutine, whichever components are
-  sealed will have the same values when the subroutine returns
-\item {\bf Ultra-eager Integrity}: at any step, whichever components are sealed
-  will be invariant
-\item {\bf Call Rule}: from the \(\jal\) instruction of a call, if the step is
-  successful and there is eventually a corresponding return, that return is to
-  the appropriate return address and all sealed components are unchanged
-\end{itemize}
-
-Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
-of the context of its call (with the exception, of course, of its arguments).
-We state this in terms of noninterference: if we say ``components \(\components\)
-interfere with the call to {\tt f},'' we mean that at some point up to when it returns,
-{\tt f} would display different behavior if called from a state with a different
-assignment of values to \(\components\). Here behavior encompasses uses of {\tt print},
-return values, and changes to state that will persist beyond the return.
-For instance, if {\tt f} computed the offset to access {\tt x} and branched on its value
-to decide whether to double or triple its argument, a change in {\tt x} could translate
-to a change in {\tt f}'s return value, making {\tt x} interfere with {\tt f}.
-
-Note that it is not sufficient to prevent sealed memory from interfering: unsealed
-memory that is, from the callee's perspective, uninitialized is also part of the calling
-context. Our model therefore does not support an ultra-eager confidentiality property.
-We have only eager confidentiality, which supports a {\em transparency rule}. But the
-notion of non-interference also lets us weaken integrity to make a {\em lazy} property
-for use with lazy tag policies.
-
-\begin{itemize}
-\item {\bf Eager Confidentiality}: the state of the stack on entry to a subroutine
-  does not interfere with its behavior up to the point that it returns
-\item {\bf Transparency Rule}: from the \(\jal\) instruction of a call, if the step is
-  successful, the entry state can be arbitrarily scrambled without changing the callee's
-  behavior up to return
-\item {\bf Lazy Integrity}:  on entry to a subroutine, whichever components are
-  sealed will either have the same values when the subroutine returns, {\em or} they will
-  not interfere with the caller's behavior up to its own return
-\end{itemize}

### 2021-03-03 example2.tex
Added example

@@
 \begin{figure}[b]
-    \begin{minipage}{.25\textwidth}
+    \begin{minipage}{.35\textwidth}
 \begin{verbatim}
 int main() {
-  long int x = 42;
-  print f(x);
+  long int x = 42, y = 0;
+  print &y;
+  y = f(x);
+  return y;
 }
@@
     \end{minipage}
-    \begin{minipage}{.4\textwidth}
+    \begin{minipage}{.35\textwidth}
 \begin{verbatim}
@@
 \caption{A Simple Call}
-\label{fig:simple-program}
+\label{fig:share-program}
 \end{figure}
@@
 \newcommand{\fsealc}{green}
+\newcommand{\sharec}{lgray}
 \newcommand{\unsealc}{gray}
@@
  &
-\memoryaddrs{35em}
+\memoryaddrs{52em}
 \memory{3}{\instrc}%
@@
 \memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]%
+\memory{2}{\sharec}[{\makebox[0pt]{Shared}}]%
 \memory{2}{\mainpassc}[{\makebox[0pt]{Passed(0)}}]%
-\memory{4}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
 ~$\cdots$
-\MemoryLabel{-37.5em}{0.75em}{0\(\ldots\)}
+\MemoryLabel{-37.5em}{0.75em}{\(\ldots\)}
 \MemoryLabel{-33.5em}{0.75em}{42}
-\MemoryLabel{-29.5em}{0.75em}{0\(\ldots\)}
-\MemoryLabel{-25.5em}{0.75em}{42}
+\MemoryLabel{-29.5em}{0.75em}{\(\ldots\)}
+\MemoryLabel{-25.5em}{0.75em}{0}
+\MemoryLabel{-21.5em}{0.75em}{\(\ldots\)}
+\MemoryLabel{-17.5em}{0.75em}{42}
 \\ \\
 Entry to {\tt f} &
-\memoryaddrs{35em}
+\memoryaddrs{52em}
 \memory{3}{\instrc}%
@@
 \memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
+\memory{2}{\sharec}[{\makebox[0pt]{Shared}}]%
 \memory{2}{\mainpassc}[{\makebox[0pt]{Passed(0)}}]%
-\memory{4}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
 ~$\cdots$
-\MemoryLabel{-29em}{0.75em}{0\(\ldots\)}
-\MemoryLabel{-24em}{0.75em}{42}
+\MemoryLabel{-29.5em}{0.75em}{\(\ldots\)}
+\MemoryLabel{-25.5em}{0.75em}{0}
+\MemoryLabel{-21.5em}{0.75em}{\(\ldots\)}
+\MemoryLabel{-17.5em}{0.75em}{42}
 \\ \\
-Return to {\tt main}
- &
-\memoryaddrs{25em}
+Entry to {\tt g} &
+\memoryaddrs{61em}
 \memory{3}{\instrc}%
@@
 \memory{1}{\emptyoutc}%
-\memory{8}{\unsealc}
+\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
+\memory{2}{\sharec}[{\makebox[0pt]{Shared}}]%
+\memory{2}{\mainpassc}[{\makebox[0pt]{Passed(0)}}]%
+\memory{2}{\fsealc}[{\makebox[0pt]{Sealed(1)}}]%
+%\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
 ~$\cdots$
-\MemoryLabel{-37.5em}{0.75em}{0\(\ldots\)}
+\MemoryLabel{-29.5em}{0.75em}{\(\ldots\)}
+\MemoryLabel{-25.5em}{0.75em}{0}
+\\ \\
+Return to {\tt main}
+ &
+\memoryaddrs{52em}
+\memory{3}{\instrc}%
+\hspace*{3pt}
+$\cdots$
+\memory{1}{\emptyoutc}%
+\memory{2}{\unsealc}%
+\memory{2}{\sharec}[{\makebox[0pt]{Shared}}]%
+\memory{4}{\unsealc}
+~$\cdots$
+\MemoryLabel{-37.5em}{0.75em}{\(\ldots\)}
 \MemoryLabel{-33.5em}{0.75em}{42}
-\MemoryLabel{-29.5em}{0.75em}{0\(\ldots\)}
-\MemoryLabel{-25.5em}{0.75em}{42}
+\MemoryLabel{-29.5em}{0.75em}{\(\ldots\)}
+\MemoryLabel{-25.5em}{0.75em}{0}
+\MemoryLabel{-21.5em}{0.75em}{\(\ldots\)}
+\MemoryLabel{-17.5em}{0.75em}{84}
 \\
@@
 \end{figure}
+
+In our second example, we have a more complicated program with nested calls
+and shared stack memory. Assuming that a long int takes two words, {\tt main}
+can't pass {\tt x} by value via a register, but must pass it on the stack. This
+is one way that memory might be shared: {\tt z} is in memory that would be sealed
+by {\tt main} but should be accessible by {\tt f}. We would not expect that {\tt g}
+should be able to access {\tt z}, however.
+
+By contrast, {\tt main} has taken the address of {\tt y}. This is commonly a precursor
+to intentionally sharing access to {\tt y}, and unless we can determine statically that
+the address does not escape, we must treat {\tt y} as shared. This form of sharing is
+more permissive than passing, as once the pointer escapes, {\tt main} has no way of knowing
+how far it will be passed. We therefore make no restrictions on {\tt y} for the remainder
+of its life.
+
+There are many ways that we might want to restrict access to a shared address. For instance,
+since in principle our sharing is related to the passing of an address, it would be standard
+to restrict access to only those callees that actually recieve the address, as in a capability
+system. This model typically works by enforcing the memory safety of all objects, including
+addressed stack variables. Or, we could imagine a mandatory access control approach in which
+the code is annotated to specifically identify which functions may access the shared memory.
+Whatever restrictions we desire, they are orthogonal to stack safety.
+
+Note that both passing and sharing are already visible at the point that {\tt main} calls
+{\tt f}, before the remainder of its frame is sealed. The compiler annotations telling
+us that passing and sharing occur are tied, respectively, to the copying of {\tt x} into
+{\tt z}, and to the taking of {\tt y}'s address. Then as before, everything below the stack
+pointer that is not passed or shared is sealed. Later, when {\tt f} calls {\tt g}, it also
+seals its own frame with its depth (1). We skip ahead to when we return to {\tt main} to note
+that while {\tt z} is no longer passed, {\tt y} is still shared, and will be until {\tt main}
+returns.

### 2021-03-04 main.tex
Working on example for sharing
Small reorg
Updating annotation and context steps for sharing

@@
 \citet{Skorstengaard+19}, which defines stack safety as
-the conjunction of two properties: {\em local state encapsulation} and {\em
-  well-bracketed control flow}.
+the conjunction of two properties: {\em local state encapsulation} and
+{\em well-bracketed control flow}.
 These properties are formalized by defining a capability-based machine
@@
 \section{Examples and Intuition}
+\label{sec:running-example}
+
 \input{example1}
+
+\input{example2}
 
@@
 information representing the history of the run. The following definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), an initial context state \(\context_0
-\in \CONTEXTS\), and a context step function \(\mach, \context \constep \context' \in
-\MACHS \times \CONTEXTS \rightarrow \CONTEXTS\).
+over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
+\(\mach, \context, \ann \constep \context' \in \MACHS \times \CONTEXTS \times \ANNS \rightarrow \CONTEXTS\).
 
@@
 
-\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap\})\]
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option{\ann}\]
+\[\ann \in \ANNS ::= \callmap + \retmap\]
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option(\ann)\]
 
@@
 \definition A system enjoys {\em ultra-eager integrity} if for any reachable state
-\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon{} \mpcstatename'\),
+\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
 and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
@@
 \(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth),\_\) and
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth)\) and
 \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
@@
 
+\subsection{Trace Properties}
+
+Having seen instances of eager confidentiality and integrity, defined in terms of reachable trace
+segments, we can generalize them to properties of the trace segments themselves. Then as we expand
+our model, our further properties will be defined in terms of the trace properties.
+
+\definition Let \(\MPCT\) be a reachable segment and \(\components\) be any set of components.
+\(\MPCT\) enjoys {\em trace integrity} when \(\last(\MPCT) ~ \component = \head(\MPCT) ~ \component\)
+for all \(\component \in \components\), and when \(\last(\MPCT)\) is undefined.
+
+\definition Let \(\MPCT \langle f \rangle\) be a reachable segment delimited by the predicate \(f\)
+and the \(\components\) be any set of components. Let \(\head(\MPCT) = \mpcstate{\mach}{\pol}{\context}\)
+and \(\nach\) be a machine state such that \(\mach \approx_\components \nach\). Suppose we take the
+induced trace from \(\npcstatename = \mpcstate{\nach}{\pol}{\context}\) that corresponds to \(\MPCT\),
+\(\npcstatename \hookrightarrow \NPCT | \neg f\). The trace \(\MPCT\) enjoys {\em trace confidentiality}
+with respect to \(f\) and \(P\) if, for all such \(\NPCT\), they correspond according to the four cases
+above:
+
+\begin{enumerate}
+\item \(\MPCT\) may be finite and \(\pi_\context(\last(\MPCT)) = (\_,\depth')\) with \(\depth' < \depth\),
+  and \(\pi_\context(\last(\NPCT)) = (\_,\depth')\) with \(\depth' < \depth\). That is, both traces end
+  having left the call of depth \(\depth\). Further:
+  \begin{itemize}
+  \item They have the same observable behavior, \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
+  \item Their final states have the same differences, \(\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
+  \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\)
+  \end{itemize}
+\item \(\MPCT\) and \(\NPCT\) may both be infinite, with \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
+\item \(\MPCT\) may be finite but end without returning -- which can only have happened if it fail-stopped --
+  and its observable behavior prefixes that of \(\NPCT\), \(\pi_\obs(\MPCT) \lesssim \pi_\obs(\NPCT)\)
+\item \(\NPCT\) may end with a fail-stop as in case 3, with \(\pi_\obs(\NPCT) \lesssim \pi_\obs(\MPCT)\)
+\end{enumerate}
+
 \subsection{Stack Safety with Sharing}
-\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap\} \cup
-\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
-\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
-
-\subsection{Trace Properties}
+
+Next we introduce sharing. We will need to extend the set of annotations to include
+sharing and passing, and now the code map may map a single instruction to multiple
+annotations, because in some calling conventions a single instruction might carry
+meaning as both a call and a pass. But we will never have a single instruction annotated
+as both a call and a return.
+
+\[\ann \in \ANNS ::= \callmap, \retmap + (\sharemap ~ P \in \MACHS \times \COMPONENTS) +
+(\passmap ~ P \in \MACHS \times \COMPONENTS)\]
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ann)\]
+
+And we extend the set of domains to include shared and passed components, the latter of
+which are labeled with the depth of the caller that has passed them.
+
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \shared{\depth} + \sealed{\depth} + \passed{\depth}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+\times (\depth \in \mathbb{N}) \]
+
+What it means to push the stack is similar to above, but two details are worth noting.
+First, shared and passed components are not sealed by the sealing convention, and second, anything
+not touched by the sealing convention must be unsealed. Sharing and passing cannot apply to memory
+the callee could otherwise use.
+
+\[
+(\push{\domMap}{\depth}{\components}) ~ \component =
+\begin{cases}
+  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
+                    \text{ and } \component \in \components \\
+  \domMap ~ \component & \text{else if } \component \in \components \\
+  \unsealed & \text{else} \\                  
+\end{cases}
+\]
+
+Popping is simpler, but note that while passed components are unsealed on return to their
+passer, shared ones are unsealed on return {\em from} their sharer - on deallocation, in other words.
+
+\[
+(\pop{\domMap}{\depth}) ~ \component =
+\begin{cases}
+  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1}
+              \text{ or } \domMap ~ \component = \passed{\depth-1}
+              \text{ or } \domMap ~ \component = \shared{\depth} \\
+  \domMap ~ \component & \text{else} \\
+\end{cases}
+\]
+
+Now we can inductively define the context update function; as these could be ambiguous assume that
+they apply in the order presented.
+
+\begin{center}
+  \judgmenttwo{\(\mach \stepsto \mach'\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \callmap \cup \passmap \cup \_\)}
+              {\(\mach,(\domMap,\depth) \stepstocon (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
+\hspace*{0.1\textwidth}
+
+  \judgmenttwo{\(\mach \stepsto \mach'\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
+              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
+\end{center}
+
+\begin{center}
+  \judgment[Default]
+           {\(\mach \stepsto \mach'\)}
+           {\(\mach,\context \stepstocon \mach',\context\)}
+\end{center}
 
@@
 
-\section{Coroutine Safety}
+\subsection{Coroutine Safety}
 \[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap\} \cup
@@
 
-\subsection{Well-bracketed Control Flow}
+\section{Well-bracketed Control Flow}
 \label{sec:wbcf}

### 2021-03-04 main.tex
Added sharing

@@
 used in some places as part of a call or return sequence and in other places
-for other purposes).  We define a call to be
-any instruction declared as such in the compiler-provided call structure.
-If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
-  return} has occurred when the $\PCname$ points to the next instruction after $a$ and $\rsp = s$. (Our definition
-of well-bracketed control flow relates this definition of return to the program's
-declared set of return points.)
-Our property definitions make no other detailed assumptions about call or return sequences, although
-some of the enforcement mechanisms we consider do.
+for other purposes). We assume the compiler's cooperation in labeling instructions
+with a variety of annotations, and in fact in identifying them as instructions.
+A {\em call-map} is a partial map from addresses to pairs of a function identifier --
+every instruction belongs to a function -- and a set of annotations. This gives
+us flexibility to have instructions simultaneously play multiple roles.
 
@@
 
-The basic building blocks of the machine are {\em values} and {\em
-  addresses}. Both are drawn from some set of {\em words}
-\(\WORDS\), ranged over by \(\word\).
+The basic building blocks of the machine are {\em values} and {\em addresses}.
+Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\).
 %
@@
 an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context, \ann \constep \context' \in \MACHS \times \CONTEXTS \times \ANNS \rightarrow \CONTEXTS\).
-
-We lift contexts into an MP trace in the same way as we lifted policy states to get
+\(\mach, \context, \ann \in \powerset \ANNS \constep \context' \in \MACHS \times \CONTEXTS
+\times \ANNS \rightarrow \CONTEXTS\). We also assume for any set of annotations that we have
+code maps, with \(\FUNIDS\) a set of function identifiers:
+\[\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]
+
+We lift contexts into an MP trace in much the same way as we lifted policy states to get
 a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).
 
-\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
-            {\(\mach, \context \constep \context'\)}
+\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \ann\)}
+              {\(\mach, \context, \ann \constep \context'\)}
               {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
@@
 
-\subsection{Program Structure}
-
-Our properties and policies assume that the program is annotated
-by the compiler with information about its structure. At minimum,
-we need to know which addresses contain instructions, which instructions
-constitute calls and returns, and to which function an instruction
-belongs.
-
-This information is represented by a {\em code map}, a partial function from
-addresses to pairs consisting of a function identifier and an {\em annotation}.
-An annotation may be {\em call} or {\em return}.
+We begin with our only annotations being {\em call} or {\em return}, and
+define our code map accordingly.
 
 \[\ann \in \ANNS ::= \callmap + \retmap\]
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option(\ann)\]
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]
 
@@
 Then at each step we update the context according to the context-step function,
-\(\stepsto_C\), which we will define in terms of the operations
+\(\stepsto_C\). We will define this in terms of the operations
 \(\push{\domMap}{n}{\components}\), in which \(\components \subseteq \COMPONENTS\),
@@
 \begin{minipage}{.4\textwidth}
-  \judgmenttwo{\(\mach \stepsto \mach'\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
-              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
+  \judgment{}
+           {\(\mach,(\domMap,\depth),\{\callmap\} \stepstocon
+             \push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1\)}
 \end{minipage}
@@
 \begin{minipage}{.4\textwidth}
-  \judgmenttwo{\(\mach \stepsto \mach'\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
-              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
+  \judgment{}
+           {\(\mach,(\domMap,\depth),\{\retmap\} \stepstocon
+             \pop{\domMap}{\depth},\depth-1\)}
 \end{minipage}
@@
   \judgment[Default]
-           {\(\mach \stepsto \mach'\)}
-           {\(\mach,\context \stepstocon \mach',\context\)}
+           {}
+           {\(\_,\context,\_ \stepstocon \context\)}
 \end{center}
@@
 
-\[\ann \in \ANNS ::= \callmap, \retmap + (\sharemap ~ P \in \MACHS \times \COMPONENTS) +
-(\passmap ~ P \in \MACHS \times \COMPONENTS)\]
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ann)\]
+\[\ann \in \ANNS ::= \callmap + \retmap + \sharemap(P \subseteq \MACHS \times \COMPONENTS) +
+\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
 
@@
 What it means to push the stack is similar to above, but two details are worth noting.
-First, shared and passed components are not sealed by the sealing convention, and second, anything
-not touched by the sealing convention must be unsealed. Sharing and passing cannot apply to memory
-the callee could otherwise use.
+First, shared and passed components are not sealed by the sealing convention, and second,
+anything not touched by the sealing convention must be unsealed. Sharing and passing cannot
+apply to memory the callee could otherwise use.
 
@@
 Popping is simpler, but note that while passed components are unsealed on return to their
-passer, shared ones are unsealed on return {\em from} their sharer - on deallocation, in other words.
+passer, shared ones are unsealed on return {\em from} their sharer - on deallocation,
+in other words.
 
@@
 
+Finally we introduce a more direct update. If \(P\) is a relation between machine states and
+components, we write \(\domMap[P ~ \mach \mapsto \stackDom]\) to say \(\domMap\) updated
+so that each component \(\component\) where \(P ~ \mach ~ \component\) holds is mapped to
+\(\stackDom\).
+
 Now we can inductively define the context update function; as these could be ambiguous assume that
@@
 \begin{center}
-  \judgmenttwo{\(\mach \stepsto \mach'\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \callmap \cup \passmap \cup \_\)}
-              {\(\mach,(\domMap,\depth) \stepstocon (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
-\hspace*{0.1\textwidth}
-
-  \judgmenttwo{\(\mach \stepsto \mach'\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
-              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\callmap\} \cup \ann) \stepstocon
+             (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\retmap\} \cup \ann) \stepstocon
+             \mach',(\pop{\domMap}{\depth},\depth-1)\)}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\passmap ~ P\} \cup \ann) \stepstocon
+             \domMap[P ~ \mach \mapsto \passed{\depth}], \depth)\)}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\sharemap ~ P\} \cup \ann) \stepstocon
+             \domMap[P ~ \mach \mapsto \shared{\depth}], \depth)\)}
+  \judgment{}
+           {\(\mach,\context,\_ \stepstocon \context\)}
 \end{center}
 
-\begin{center}
-  \judgment[Default]
-           {\(\mach \stepsto \mach'\)}
-           {\(\mach,\context \stepstocon \mach',\context\)}
-\end{center}
+\definition Let \(\MPCT\) be a reachable trace segment of the form
+\(\mpcstatename ~ \mpcstatename'\), where \(\pi_\context(\mpcstatename) = (\domMap,\depth)\).
+Let \(\components\) be the set of components \(\component\) such that either
+\(\domMap ~ \component = \sealed{\_}\) or \(\domMap ~ \component = \passed{\depth'}\) for
+some depth \(\depth' < \depth-1\). If any such \(\MPCT\) obeys trace integrity with respect
+to \(\components\), then the system enjoys {\em ultra-eager confidentiality} in the sharing
+model.
+
+\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
+on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
+\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
+\(\component\) such that either \(\domMap ~ \component = \sealed{\_}\) or
+\(\domMap ~ \component = \passed{\depth''}\) for some depth \(\depth'' < \depth'-1\).
+Then the system enjoys {\em eager integrity} in the sharing model if \(\MPCT\) obeys
+trace integrity with respect to \(\components\).
+
+\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
+on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
+\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
+\(\component\) such that either \(\domMap ~ \component = \outside\) or
+\(\domMap ~ \component = \shared{\_}\). Then the system enjoys {\em eager confidentiality}
+in the sharing model if \(\MPCT\) obeys trace confidentiality with respect to \(\components\)
+and \(f\).
+
+\subsection{Coroutine Safety}
+\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap\} \cup
+\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
+\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
 
@@
 observationally equivalent states. But for our purposes it is irrelevant.
-
-\subsection{Coroutine Safety}
-\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap\} \cup
-\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
-\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
 

### 2021-03-07 main.tex
Fixed up some broken proofs in ObsTrace.v

@@
   \item They have the same observable behavior, \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
-  \item Their final states have the same differences, \(\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
-  \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\)
+  \item Their final states have the same differences:
+    \[\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
+    \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\]
   \end{itemize}
@@
 \subsection{Coroutine Safety}
-\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap\} \cup
-\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
-\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
+
+Our final extension to the model is to add coroutines. We assume a very simple model with
+a fixed number of coroutines, each with its own stack.
+Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
+\(\stackof\) from addresses to stack ids. The ``active stack'' -- the stack belonging to the
+currently executing coroutine -- in any given machine state is \(\stackof(\mach ~ \SP)\).
+We extend the set of possible annotations with a yield.
+
+\[\ann \in \ANNS ::= \callmap + \retmap + \yieldmap +
+\sharemap(P \subseteq \MACHS \times \COMPONENTS) +
+\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
+
+We use the same domain model as above, but instead of keeping track of a single depth
+we use a map from stacks to their depths.
+
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
+(\depthMap \in \STACKS \rightarrow \mathbb{N})\]
 

### 2021-03-18 CheriApproaches.tex
Expanded Cheri writeup

@@
 
-This document deals with questions of the division between machine state and policy state.
-In existing models of our security properties, the state is divided into a machine state,
-which is a mapping from addresses and registers to data, and a policy state, which may be
-any arbitrary data. A machine state can step on its own, or in conjunction with a policy
-state; when states step together the machine state component always steps as it would on
-its own. A policy state may only cause the step to fail outright, never change the result.
+\section{A Stack-safe Calling Convention for Capabilities}
+
+This is Skorstengaard's calling convention from the worse of their papers, with a
+difference in that we use a single piece of code to handle the ``landing pad'' of a return.
+We must combine this with descriptions of how to adjust out model to support them. At the
+high level, we need to accept that their threat model is slightly incompatible with ours:
+adversarial code, defined as code whose dynamic behavior does not obey simple reasonableness
+properties, is exempt from the LSE and WBCF protections in its own calls. For instance, such
+a function might make a call in such a way that the return skips over it and returns straight
+to its caller, violating WBCF. Or it might create capabilities on its own data and pass them
+along, violating LSE. This latter piece of unreasonableness is easy to handle: we add a clause
+to integrity to assume that the caller is reasonable. Equivalent adjustments to confidentiality
+might be trickier, in that we should acknowledge the possibility of unreasonableness in either
+the callee or the caller. But the violations of WBCF are much more problematic, because our
+model of counting depths assumes that callers are never ``skipped'' by a multi-level return.
+
+\subsection{Updating Stack Safety to Support Unreasonable Functions}
+
+The solution given in StkTokens is, essentially, that unreasonable functions don't get to
+make ``real'' (protected) calls. Rather, they perform all the operations involved in a call,
+then jump into another function at its entry point. That function may then jump back -- which
+is not a ``real'' return -- or do an actual return to the original caller. The equivalent to
+this is in our model is for us to, on a call-annotated instruction, check that the current
+function is reasonable. If it is not, we do not update the depth, and given the callee an
+unsealed piece of stack to work with. Similarly, on return, we only update the context
+if we are returning to a reasonable function. That requires examining the program counter
+that we're returning to. The problem with this whole approach is that it doesn't play nice
+with the initialization side of confidentiality. The adversary gets to initialize its
+callee's frame before fake-calling it. A reasonable callee will not read those initialized
+variables, anyway, but the property will not capture that the protection is needed. We
+could potentially treat this as a special kind of call, obeying confidentiality but not
+integrity or WBCF, which gets us closer to the next approach.
+
+The more complex approach is to accept that unreasonable callers might get skipped or
+accessed, but otherwise model their calls and returns normally. That means extending
+our model with support for dynamic returns, which are then forbidden by WBCF in the case
+of reasonable callers. Specifically, the {\sc Ret} annotation is now parameterized by
+a function that takes a machine state and a context and returns an integer depth: the
+return target. Instead of only popping the top depth, we pop the depth
+we're returning to and every one above it, by the same mechanisms as in a single return.
+In the context of PIPE, where we always guarantee single returns, the function can simply
+subtract 1 from the current depth, as before. We will discuss how to implement this
+function for Cheri once we cover the calling convention.
+
+Now most of our properties are unaffected by this, even Return Integrity, which after
+all talks about what happens {\it when} we return to a caller. Then we simply need a second
+property, which I propose to call Return Guarantee, that if a caller is reasonable it
+will see a return {\it before} any caller with a lower depth does.
+
+\subsection{Calling Convention Details}
+
+First let's discuss our assumptions about the capability system, starting with basic
+capabilities.
+
+A {\em capability} is composed of a {\em base}, {\em offset}, and {\em bound}
+Each word of memory carries a 2-bit tag, the first bit indicated whether it is a
+{\em valid} capability and the second whether that capability is {\em local}
+A local capability can never be stored in memory outside of the stack. A load or store
+failstops if the capability used as its pointer does not have its offset within its
+bounds. A load or store using a non-capability pointer uses the bounds of one of the
+following special capabilities.
+
+There are three special capability registers. The {\em default data capability} ({\tt \$ddc}),
+is applied to reads and writes through non-capability pointers. The
+{\em program counter capability} ({\tt \$pcc}) is applied to instruction fetches.
+The {\em stack capability} ({\tt spc}) takes the place of the stack pointer and is used by
+compiler generated code. Basic capabilities are supported by the following instructions:
+
+\begin{tabular}{l l l}
+  {\tt st}      & Data Store       & Stores data, clearing any capability bits \\
+  {\tt ld}      & Data Load        & Loads data, clearing any capability bits \\
+  {\tt cst}     & Capability Store & Stores a capability, but failstops rather than store
+                                     a local one outside the stack \\
+  {\tt cld}     & Capability Load  & Loads a capability from memory \\
+  {\tt incbase} & Increment Base   & Increases the base of a capability by one and
+                                     decrements the offset \\
+  {\tt decbnd}  & Decrement Bound  & Decreases the bound of a capability by one \\
+\end{tabular}
+
+We will need to use {\em object capabilities}. Object capabilities come in twos:
+a code pointer and a memory capability. They are sealed with a matching identifier and
+can only be unsealed by invoking them together, simultaneously jumping to the code pointer
+and unsealing the memory capability. This way we ensure that the memory capability is
+only used by trusted code. These are supported by the following instructions:
+
+\begin{tabular}{l l l}
+  {\tt so} & Seal Object   & Takes a pair of capabilities and
+                                  seals them with matching identifiers \\
+  {\tt io} & Invoke Object & Takes matching sealed capabilities \(c\) and
+                                  \(p\), jumps to \(p\), and unseals \(c\) \\
+\end{tabular}
+
+Finally, we will use monotone capabilities.
+
+\subsection{The calling convention}
+
+From here on out, assume all capabilities are local. They can live in registers and stack
+memory only. In the interest of corresponding to out model, the {\tt \$ddc} is permissive
+and allows access to everything that isn't stack or code, and the {\tt \$pcc} allows access
+only to our code.
+
+{\bf Initially:} The {\tt \$stc} encompasses the full stack in its bounds. There is a
+special bit of code that has no function identifier that serves as a landing pad for
+functions, that takes a capability on a two-word block, loads the first word to {\tt \$stc}
+and jumps to the second. The jump is annotated as a return by the compiler.
+
+{\bf Caller:} To make a call, the caller allocates two extra words on the stack. It copies
+the current {\tt \$stc} and the address of the instruction after the call into these. Then
+it creates a capability restricted to this object and links it as an object capability with
+the address of the landing pad. The object capability is stored in a register. Then the
+caller increments the base of the {\tt \$stc} until its offset is zero, which will prevent
+access to its frame. Finally, the caller jumps to the call target -- this instruction is
+annotated as a call by the compiler.
+
+After a return, the caller immediately clears all non-return-value registers and the stack
+above the {\tt \$stc}'s offset. This prevents a copy of the object capability from being
+stashed somewhere for a multiple return attack.
+
+{\bf Callee:} On entry, the callee allocates memory by increasing the {\tt \$stc} and stores
+the object capability it was passed. To return, it invokes that capability, executing the
+landing pad and unsealing the data portion of the object capability. At the end of the
+landing pad sequence, there is a jump instruction, which is annotated as a return. As
+discussed above, to dynamically determine the target of the return, we must look at the
+machine state and contet: the data portion of the return capability points to the activation
+record of the caller, which is labeled {\it Passed(d)} for some depth \(d\) in the context.
+(That is, it lies with the frame of the activation at depth \(d\).) The return target is
+therefore \(d\). An intuitive way of describing this is that on return, we unseal layers of
+calls until the activation record we're using to return has been unsealed.
+
+\subsection{Analysis}
+
+This calling convention prevents the callee from accessing the caller's state. The caller
+guarantees that there is no additional object capability floating around somewhere that
+the callee could access, and the {\tt \$stc} has its bounds shrunk to exclude the caller's
+frame. It can only be restored with the object invocation.
+
+What it doesn't necessarily do is prevent the callee from reading its frame. An adversarial
+caller could write above its stack pointer before the call and there would be no protection.
+So we have ``sealed confidentiality'' instead of normal confidentiality that includes all
+uninitialized memory.
+
+Luckily, if we extend our model with the ``uninitialized capabilities'' of Georges et. al.
+we can once again enforce the confidentiality of uninitializes memory, as well. Then
+we have full local state encapsulation.
+
+The generic return code being shared between all functions is a little unfortunate, as it
+puts a hole in control separation. We'll need to describe the correct behavior carefully
+and adjust the property.
+
+We may have a problem with returns, in that a callee could potentially pass along its
+return object to a further callee, which could then bypass the first and return straight
+to the original caller. This appears to be a hazard of our attacker model.
+
+The caller would be happy enough, as its state would have been protected,
+but it messes up our whole model of depths and counting
+of returns. And the callee might rely on getting returned to.
+
+\section{Well-formedness, Reasonableness, and PIPE}
+
+As discussed above, Skortengaard et. al. impose static well-formedness conditions on
+all of their code, as well as certain
+
+\section{The Question of Policy States}
+
+This half of the document deals with questions of the division between machine state and
+policy state. In existing models of our security properties, the state is divided into a
+machine state, which is a mapping from addresses and registers to data, and a policy state,
+which may be any arbitrary data. A machine state can step on its own, or in conjunction
+with a policy state; when states step together the machine state component always steps
+as it would on its own. A policy state may only cause the step to fail outright, never
+change the result.
 
@@
 
-\paragraph*{HOPE} The machine contains both ordinary registers and memory, and tag registers and memory.
-The transition function is parameterized by a tag rule table, and we must be careful to construct the model
-such that tags only be read and updated as described by the table. Fail-stops should occur only on tag
-rule failures.
+\paragraph*{HOPE} The machine contains both ordinary registers and memory, and tag registers
+and memory. The transition function is parameterized by a tag rule table, and we must be
+careful to construct the model such that tags only be read and updated as described by the
+table. Fail-stops should occur only on tag rule failures.
 
-\paragraph*{Cheri} The machine contains ordinary registers and memory, with some effort needed to model
-the extra space that capabilities take up. (Either words need to be larger to fit capabilities, or we
-need to model them being spread across multiple words.) It also contains the map of 1-bit tags marking
-capabilities as valid. The ISA is extended with the Cheri operations and the transition function handles
-the valid-capability tags appropriately. Here a fail-stop occurs whenever the Cheri architecture would
-trap to a failstate, mostly when reading or writing through an invalid capability or outside of a
-valid capability's authority.
+\paragraph*{Cheri} The machine contains ordinary registers and memory, with some effort needed
+to model the extra space that capabilities take up. (Either words need to be larger to fit
+capabilities, or we need to model them being spread across multiple words.) It also contains
+the map of 1-bit tags marking capabilities as valid. The ISA is extended with the Cheri
+operations and the transition function handles the valid-capability tags appropriately.
+Here a fail-stop occurs whenever the Cheri architecture would trap to a failstate, mostly
+when reading or writing through an invalid capability or outside of a valid capability's
+authority.
 
-\paragraph*{Software} Here the machine gets to be completely unchanged, and never failstops. Instead
-we restrict the set of initial states to those whose code comes from the compiler with the appropriate
-modifications, and it will simply never violate the policy - any errors will take it to some error
-state where it silently diverges, which is perfectly allowed.
+\paragraph*{Software} Here the machine gets to be completely unchanged, and never failstops.
+Instead we restrict the set of initial states to those whose code comes from the compiler
+with the appropriate modifications, and it will simply never violate the policy - any errors
+will take it to some error state where it silently diverges, which is perfectly allowed.
 
 \paragraph*{Implications for confidentiality} If everything lives in the machine state, then
-we need to think about what it means to vary it to create a confidentiality property. Take the HOPE
-example: does a variant include states with different tags, such that either the primary
-trace failstops and the variant does not, or the variant does and the primary does not? Similarly,
-in Cheri, do we vary the valid-capability bit? What about bounds information?
+we need to think about what it means to vary it to create a confidentiality property. Take
+the HOPE example: does a variant include states with different tags, such that either the
+primary trace failstops and the variant does not, or the variant does and the primary does
+not? Similarly, in Cheri, do we vary the valid-capability bit? What about bounds information?
 
 We can pretty firmly say that we do not need to vary tags, including the capability bit. That's
-because these can never be read, and therefore it is meaningless to talk about their confidentiality.
-But the bounds information in Cheri presents a bigger problem. We will revisit it as we start moving
-policy data out of the machine.
+because these can never be read, and therefore it is meaningless to talk about their
+confidentiality. But the bounds information in Cheri presents a bigger problem. We will
+revisit it as we start moving policy data out of the machine.
 
@@
 We have identified some kinds of data that can never be read and serve exclusively to inform
-the enforcement mechanism. We can think of data as having two meanings, its ``application content'' and its
-``security content.'' In HOPE, the distinction seems clear: tags have only security content, never application.
-So we can justify separating out security-only data into a separate part of the state, from which it cannot
-escape and therefore does not need to be varied to support confidentiality. Let's discuss how that
-works in HOPE and Cheri (in software methods, of course, everything has application content.)
+the enforcement mechanism. We can think of data as having two meanings, its ``application
+content'' and its ``security content.'' In HOPE, the distinction seems clear: tags have only
+security content, never application. So we can justify separating out security-only data
+into a separate part of the state, from which it cannot escape and therefore does not need
+to be varied to support confidentiality. Let's discuss how that works in HOPE and Cheri
+(in software methods, of course, everything has application content.)
 
-\paragraph*{HOPE} In HOPE, we put all tags in the policy state, along with any policy-specific machinery such as
-counters that give us the fresh tags. The machine step function can now be total, with failure only
-coming from the policy step function. But it turns out that some machine data still has security content.
-Consider a load instruction: its tag rule must take into account the tag at the target address, which
-means that the policy step function must be aware of the machine state.
+\paragraph*{HOPE} In HOPE, we put all tags in the policy state, along with any policy-specific
+machinery such as counters that give us the fresh tags. The machine step function can now be
+total, with failure only coming from the policy step function. But it turns out that some
+machine data still has security content. Consider a load instruction: its tag rule must
+take into account the tag at the target address, which means that the policy step function
+must be aware of the machine state.
 
-This gives us the same question of what we do in confidentiality with variant states that fail-stop
-when the original didn't.
+This gives us the same question of what we do in confidentiality with variant states that
+fail-stop when the original didn't.
 
-\paragraph*{Cheri} In Cheri, the valid-capability bit has no application content, so these can live in
-the policy state. But the application content of a capability is its bitwise value. We might argue
-that actually the base and offset are application content, while permissions and bounds information
-are security-only, but it is not feasible to separate the latter from the former. Even if we did, the
-base and bounds would give us the same issues as HOPE pointers.
+\paragraph*{Cheri} In Cheri, the valid-capability bit has no application content, so these can
+live in the policy state. But the application content of a capability is its bitwise value.
+We might argue that actually the base and offset are application content, while permissions
+and bounds information are security-only, but it is not feasible to separate the latter
+from the former. Even if we did, the base and bounds would give us the same issues as HOPE
+pointers.
 
-\paragraph*{Implications} The primary implication is that we no longer need to worry about what happens
-if we vary tags, whether that's the full HOPE tag or the 1-bit Cheri tag. But that doesn't save us
-from sometimes varying security content, because pointer values still have security content. In both systems,
-if we read a valid pointer that is supposed to be secret and then dereference it, there is a variant run
-in which the pointer is invalid and the variant fail-stops. This will always be the case no matter how we
-separate machine and policy.
+\paragraph*{Implications} The primary implication is that we no longer need to worry about
+what happens if we vary tags, whether that's the full HOPE tag or the 1-bit Cheri tag.
+But that doesn't save us from sometimes varying security content, because pointer values
+still have security content. In both systems, if we read a valid pointer that is supposed to
+be secret and then dereference it, there is a variant run in which the pointer is invalid and
+the variant fail-stops. This will always be the case no matter how we separate machine and
+policy.
 
-The advantages that we do get by separating them as much as possible are: it's easier to model the separate
-state than if we built it into the machine architecture, which matters more for HOPE but is a small concern for
-Cheri. And depending on how we handle these variant fail-stops, it may be more efficient to test if fewer of
-them occur.
+The advantages that we do get by separating them as much as possible are: it's easier to model
+the separate state than if we built it into the machine architecture, which matters more for
+HOPE but is a small concern for Cheri. And depending on how we handle these variant fail-stops,
+it may be more efficient to test if fewer of them occur.
 

### 2021-03-22 CheriApproaches.tex
Finished up some of the Cheri work I was doing

@@
 {\em program counter capability} ({\tt \$pcc}) is applied to instruction fetches.
-The {\em stack capability} ({\tt spc}) takes the place of the stack pointer and is used by
+The {\em stack capability} ({\tt \$stc}) takes the place of the stack pointer and is used by
 compiler generated code. Basic capabilities are supported by the following instructions:
@@
   {\tt ld}      & Data Load        & Loads data, clearing any capability bits \\
-  {\tt cst}     & Capability Store & Stores a capability, but failstops rather than store
+  {\tt stc}     & Capability Store & Stores a capability, but failstops rather than store
                                      a local one outside the stack \\
-  {\tt cld}     & Capability Load  & Loads a capability from memory \\
-  {\tt incbase} & Increment Base   & Increases the base of a capability by one and
-                                     decrements the offset \\
-  {\tt decbnd}  & Decrement Bound  & Decreases the bound of a capability by one \\
+  {\tt ldc}     & Capability Load  & Loads a capability from memory \\
+%  {\tt incbase} & Increment Base   & Increases the base of a capability by one and
+%                                     decrements the offset \\
+%  {\tt decbnd}  & Decrement Bound  & Decreases the bound of a capability by one \\
 \end{tabular}
@@
 \begin{tabular}{l l l}
-  {\tt so} & Seal Object   & Takes a pair of capabilities and
-                                  seals them with matching identifiers \\
-  {\tt io} & Invoke Object & Takes matching sealed capabilities \(c\) and
+  {\tt seal}   & Seal Object    & Takes a pair of capabilities and a seal identifier and
+                                  seals them with the identifier \\
+  {\tt invoke} & Invoke Object  & Takes matching sealed capabilities \(c\) and
                                   \(p\), jumps to \(p\), and unseals \(c\) \\
@@
 
-Finally, we will use monotone capabilities.
+Finally, we will use {\em uninitialized capabilities}. These have a special permission structure
+where their offset only moves when written to. The portion of the capability below the offset,
+which is initialized, can be promoted into a normal capability. They support the following
+instructions:
 
-\subsection{The calling convention}
+\begin{tabular}{l l l}
+  {\tt stu}  & Data Store (Uninitialized) & Stores data and increments offset of Ucap \\
+  {\tt stuc} & Capability Store           & As {\tt stU} but stores a capability \\
+  {\tt prom} & Promote                    & Promote initialized part of an uninitialized capability
+                                            into a normal capability \\
+  {\tt morp} & Co-promote                 & Remove the initialized part of an uninitialized capability \\
+\end{tabular}
 
-From here on out, assume all capabilities are local. They can live in registers and stack
-memory only. In the interest of corresponding to out model, the {\tt \$ddc} is permissive
-and allows access to everything that isn't stack or code, and the {\tt \$pcc} allows access
-only to our code.
+\section{A Capability Machine}
 
-{\bf Initially:} The {\tt \$stc} encompasses the full stack in its bounds. There is a
-special bit of code that has no function identifier that serves as a landing pad for
-functions, that takes a capability on a two-word block, loads the first word to {\tt \$stc}
-and jumps to the second. The jump is annotated as a return by the compiler.
+Here I'll describe a simple machine implementing the operations above. Eventually. For now
+I mocked it up in CheriModel.v. It's pretty close to what's described above. The most interesting
+thing is the semantics of invoking a sealed object. A sealed object is a pair of a code capability
+and a data capability (though these are distinguished only by convention.) There is a special register,
+{\tt \$invdst}, in which the unsealed data capability is stored (and of course, the unsealed code capability
+is copied to the {\tt \$pcc}. The idea is that on entry to an invoked capability, we always need to know
+which register has the data capability, so it can't be dynamic.
 
-{\bf Caller:} To make a call, the caller allocates two extra words on the stack. It copies
-the current {\tt \$stc} and the address of the instruction after the call into these. Then
-it creates a capability restricted to this object and links it as an object capability with
-the address of the landing pad. The object capability is stored in a register. Then the
-caller increments the base of the {\tt \$stc} until its offset is zero, which will prevent
-access to its frame. Finally, the caller jumps to the call target -- this instruction is
-annotated as a call by the compiler.
+\subsection{A call}
 
-After a return, the caller immediately clears all non-return-value registers and the stack
-above the {\tt \$stc}'s offset. This prevents a copy of the object capability from being
-stashed somewhere for a multiple return attack.
+Here we will give an example of a call to a function at address {\tt 20} that takes an
+argument in register {\tt ra}. Let our call start at address {\tt 0}. Suppose that we have a piece
+of code at {\tt 50} that executes a ``landing pad'' for a return. Assume that all capabilities referenced
+here are local, and the stack capability is an uninitialized capability that can write locally. There
+could be nonlocal capabilities as well, for use in the heap, but to match our model we can also just assume
+that such capabilities are derived from a permissive {\tt $ddc}.
 
-{\bf Callee:} On entry, the callee allocates memory by increasing the {\tt \$stc} and stores
-the object capability it was passed. To return, it invokes that capability, executing the
-landing pad and unsealing the data portion of the object capability. At the end of the
-landing pad sequence, there is a jump instruction, which is annotated as a return. As
-discussed above, to dynamically determine the target of the return, we must look at the
-machine state and contet: the data portion of the return capability points to the activation
-record of the caller, which is labeled {\it Passed(d)} for some depth \(d\) in the context.
-(That is, it lies with the frame of the activation at depth \(d\).) The return target is
-therefore \(d\). An intuitive way of describing this is that on return, we unseal layers of
-calls until the activation record we're using to return has been unsealed.
+\begin{verbatim}
+<0: [clear all registers]
+0: mov 42 ra
+1: mov $stc r1
+2: morp $stc
+3: stuc r1 $stc               [PASS $stc]
+4: stuc [$pcc + 7] $stc       [PASS $stc]
+5: mov $stc r1
+6: prom r1
+7: morp $stc
+8: mov [$pcc + 42] r2
+9: seal s r1 r2
+10: jmp +10                   [CALL]
+11: ...
+\end{verbatim}
 
-\subsection{Analysis}
+We can walk through what this does step by step. First, to make sure that we don't have any extra
+capabilities laying around, we clear all of our registers. Then we set the argument in {\tt ra}.
+Now things get interesting: instructions 1-2 save the current stack capability and cut off
+the readable portion, leaving only the uninitialized portion. Instructions 3-4 construct an activation
+record containing the previous stack capability and the appropriate return address. Note that each {\tt stuc}
+instruction increments the stack pointer, so we are implicitly allocating these two variables.
+They will need to be accessible to the callee, so we annotate them as passed. Next, we create a capability
+that can only access them, in instructions 5-6: the promoted stack capability is a capability on
+only the written portion. Finally instruction 7 ensures that the stack capability cannot access
+them. Instructions 8 and 9 now seal together the data capability and a code capability on address
+50, where our ``landing pad'' code lives, and address 10 jumps us ahead to our function entry.
 
-This calling convention prevents the callee from accessing the caller's state. The caller
-guarantees that there is no additional object capability floating around somewhere that
-the callee could access, and the {\tt \$stc} has its bounds shrunk to exclude the caller's
-frame. It can only be restored with the object invocation.
+Now, what happens when we return? Let's look at the landing pad. Recall that we have a special
+register {\tt \$invdst} that is where the unsealed data capability lives.
 
-What it doesn't necessarily do is prevent the callee from reading its frame. An adversarial
-caller could write above its stack pointer before the call and there would be no protection.
-So we have ``sealed confidentiality'' instead of normal confidentiality that includes all
-uninitialized memory.
+\begin{verbatim}
+50: ldc $invdst $stk
+51: jalr [$invdst+1]          [RETURN]
+\end{verbatim}
 
-Luckily, if we extend our model with the ``uninitialized capabilities'' of Georges et. al.
-we can once again enforce the confidentiality of uninitializes memory, as well. Then
-we have full local state encapsulation.
+Pretty simple! Notice something interesting: if we invoke a (properly constructed) return capability,
+we {\it always} hit a {\tt jalr} labeled as a return two steps later. In other words, an adversary can't do
+something that looks like a return without it being labeled a return.
 
-The generic return code being shared between all functions is a little unfortunate, as it
-puts a hole in control separation. We'll need to describe the correct behavior carefully
-and adjust the property.
+Now it's pretty easy here to argue some of our properties. We should have that no unsealed capabilities are
+present unless they were passed as arguments, except the stack capability, which is completely uninitialized.
+So we should not be able to read anything without writing it or access a caller's frame at all.
 
-We may have a problem with returns, in that a callee could potentially pass along its
-return object to a further callee, which could then bypass the first and return straight
-to the original caller. This appears to be a hazard of our attacker model.
+When we've returned, we've restored the old stack capability, so the callee's frame is again uninitialized.
+But! We may have a capability on it that the callee returned. That's a potential problem, because the capability
+could then be passed to a future call in a way that allows it to persist. There's a solution to that, unpleasant
+though it is: clear every register on return except for the value being returned, and manually check that it
+isn't a capability that overlaps the uninitialize portion of the stack. Or indeed, any stack capability at all.
 
-The caller would be happy enough, as its state would have been protected,
-but it messes up our whole model of depths and counting
-of returns. And the callee might rely on getting returned to.
+\subsection{Friction points}
 
-\section{Well-formedness, Reasonableness, and PIPE}
+In this model we don't really have control separation. And if we did, the landing pad wouldn't
+actually belong to any one function, so it'd be a bit odd. We could get separation back by making
+the call an invocation of an object, with its own launchpad that handles all of the rest of the
+construction.
 
-As discussed above, Skortengaard et. al. impose static well-formedness conditions on
-all of their code, as well as certain
+We still have the problem with returns, in that a callee could potentially pass along its
+return object as an argument to a further callee, which could then bypass the first and return
+straight to the original caller. We can model what happens in this case by checking, at the return point,
+where we are in the stack - that is the depth we are about to return to, and it may be more than one depth
+down (or zero, as the callee could always manufacture its own return object to ``return'' to itself).
 

### 2021-03-24 cheri/CheriApproaches.tex
Merged Leo's risc-v work with my most recent and organized the directory

--- 2021-03-22 CheriApproaches.tex
@@
   {\tt ldc}     & Capability Load   & Loads a capability from memory \\
+  {\tt cca}     & Adjust capability & Adds argument to the capability offset \\ 
 %  {\tt incbase} & Increment Base   & Increases the base of a capability by one and
@@
 could be nonlocal capabilities as well, for use in the heap, but to match our model we can also just assume
-that such capabilities are derived from a permissive {\tt $ddc}.
+that such capabilities are derived from a permissive {\tt \$ddc}.
 

### 2021-03-24 tex/props.tex
Merged Leo's risc-v work with my most recent and organized the directory

--- 2020-05-29 props.tex
@@
 
-    A trace is a potentially infinite, nonempty sequence. We define the append operation \(T \cdot T'\)
-    where \(T'\) may be a trace or empty (\(\varepsilon\)):
-    \[T_1 \cdot T_2 \triangleq
-      \begin{cases}
-        T_1 & \text{if } T_2 = \varepsilon \\
-        t T_2 & \text{if } T_1 = t \\
-        t T_1' \cdot T_2 & \text{if } T_1 = t T_1' \\
-      \end{cases}\]
+    A trace is a potentially infinite, nonempty sequence. We define a few operations on traces generically.
+    The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
+    separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
+    \[\frac{f\ t}
+           {\mathit{split}(f,tT) = t, tT}\]
+    \[\frac{\neg f\ t \hspace{.5cm} \mathit{split}(f,T) = T_1,T_2}
+           {\mathit{split}(f,tT) = tT_1,T_2}\]
 
-    We identify infinite traces with \(\isinf(T)\). We define a ``trace-of'' operator \(\hookrightarrow\)
-    that coinductively relates a machine state with the trace of machine states paired with observations
-    produced by repeated application of step:
+    The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
+    of the first (if it has one) with the first of the second:
+    \[\frac{}
+           {T_1t \cdot T_2 = T_1 T_2}\]
+
+    Note that both operations are defined coinductively. We identify infinite traces with \(\isinf(T)\).
+    We define a ``trace-of'' operator \(\hookrightarrow\) that coinductively relates a machine state with the
+    trace of machine states paired with observations produced by repeated application of step:
 
@@
     Since the step function is total, if \(\mach \hookrightarrow \machT\), then \(\machT\) must be infinite.
-    We will also wish to take a prefix of the trace from an initial state up through 
-    the first state where some condition holds. Let \(f\) be a predicate on machine states.
-    Then we define \(\mach \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix from
-    \(\mach\) until \(f\)''):
-    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm} \mach_1
-              \hookrightarrow \machT | f}
-           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT | f}\]
-    \[\frac{f\ m}
-           {m \hookrightarrow (m,\tau)}\]
+    We will also wish to take a prefix of the trace from an initial state up through the first state where
+    some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
+    \(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
+    \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
+           {\mach \hookrightarrow \machT_1 | f}\]
+
+    %\[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
+    %       {\mach \hookrightarrow f | \machT_2}\]
 
@@
     \[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
-           {(\mach_0,\pol_0) \hookrightarrow (\mach_0,\pol_0)}\]
+           {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
     \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
-              (\mach_1,\pol_1) \hookrightarrow \machT}
-           {(\mach_0,\pol_0) \hookrightarrow_P (\mach_0,\pol_0,\obs) \machT}\]
-
-    \[\frac{\mach_0,\pol_0 \xrightarrow{\obs} \mach_1,\pol_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm}
-              \mach_1,\pol_1 \hookrightarrow_P \machT | f }
-           {\mach_0,\pol_0 \hookrightarrow_P \mach_0,\pol_0 \machT | f}\]
-    \[\frac{\mach_0,\pol_0 \xrightarrow{\obs} \mach_1,\pol_1 \hspace{.5cm} f\ \mach_1}
-           {\mach_0,\pol_0 \hookrightarrow_P (\mach_0,\pol_0) (\mach_1,\pol_1)}\]
-
-    \COQ{this operation steps until it reaches a state on which \(f\) holds,
-    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
-    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
-    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
-      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
-    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}
+              \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
+           {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]
+    \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
+           {\mpstatename \hookrightarrow \MPT_1 | f}\]
+    %\[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
+    %       {\mpstatename \hookrightarrow f | \MPT_2}\]
 
@@
     An observational prefix \(\obsT_1 \lesssim \obsT_2\) means that the trace \(\obsT_1\) of observations is
-    a prefix of \(\obsT_2\) up to the deletion of \(\tau\) observations:
-    \leo{Point out this is coinductive}
+    a prefix of \(\obsT_2\) up to the deletion of \(\tau\) observations (once again, coinductively):
 
@@
 
-    \[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2'}
+    \[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
            {\obsT_1 \lesssim \obsT_2}\]
-    \leo{Doesn't this case miss the recursive $\obsT_1 \lesssim \obsT_2$ condition?}
-
-    \COQ{This is the same as the coq version minus a redundant case.}
 
@@
     \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]
-
-    \COQ{this should be equivalent to the coq version, and we should adopt it.}
-    \leo{We had this, but it was much harder to work with for my proofs. We should still
-      prove them equivalent if we want to present it this way (which is probably cleaner for
-      presentation purposes)}
 
@@
            {\MPT \sim_P \machT}\]
-
-    [Todo: update coq to use this.]
-    \leo{Oh, I like this version. Should make things cleaner in the top part of my proofs
-      while leaving the hard low-level invariants unaffected.}
 
@@
     A stack safety property is a property of a transition system, a policy, and a callmap
-    that is decomposed into an integrity property, a confidentiality property, and a safe
-    initialization property. We will present eager and observational variants of stack safety.
+    that is decomposed into an integrity property and a confidentiality property.
+    We will present eager and observational variants of stack safety.
 
@@
 
-    {\it Eager stack confidentiality} states that data outside a callee's stack frame is secret,
-    and does not interfere with its observable behavior or with the machine state at return. Its
-    weaker variant {\it observable stack safety} lifts the latter restriction and states that
-    the callee may not leak these secrets directly or by changing the behavior of its caller.
-
-    Finally, {\it safe initialization} deals with the top level outside of a function call, in
-    which uninitiaized data should not interfere with any observable behavior.
-
-    \COQ{safe initialization is implemented differently in each property, since
-     it's really just confidentiality. But because there's no return it's functionally the same.}
+    \sna{Attempt 2...}
+    {\it Observable stack confidentiality} states that data outside a callee's stack frame are secret,
+    and may not interfere with the observable behavior of the program. That is, the callee's behavior must
+    be identical regardless of the configuration of memory outside its frame, and it must not change the
+    program state based on that memory in a way that causes future changes in behavior.
+    {\it Eager stack confidentiality} additionally states that changes the callee makes to the program state
+    must be identical regardless of the configuration of memory.
 
@@
       & \textit{observable stack integrity} \land \\
-      & \textit{observable stack confidentiality} \land \\
-      & \textit{safe initialization}
+      & \textit{observable stack confidentiality} \\
     \end{split}\]
@@
       & \textit{eager stack integrity} \land \\
-      & \textit{eager stack confidentiality} \land \\
-      & \textit{safe initialization}
+      & \textit{eager stack confidentiality} \\
     \end{split}\]
+
+    We will separately introduce confidentiality at the top level, to illustrate it simply before
+    moving into calls. Integrity at the top level is vacuous because there is nothing to protect.
 
@@
 
-    \paragraph{Safe Initialization}
+    \paragraph{Top-level Confidentiality}
 
@@
       returns only confidentiality is interesting? Or would we miss something?}
+    \sna{Changed to make it confidentiality, but still introducing contours after. Maybe
+         contours should be before?}
 
@@
       there is no guarantee of the initial value of memory in the stack. So there is no
-      need to protect a caller's stack data, but reading uninitialized data is unsafe.
+      need to protect a caller's stack data, but data above the stack frame
+      (all of memory) are out of scope and uninitialized, and therefore unsafe to read.
 
@@
 
-      A system enjoys {\it safe initialization} if for any initial states \(\mach_0\)
+      A system enjoys confidentiality in this case if for any initial states \(\mach_0\)
       and \(\mach_0'\) such that \(\mach_0 \approx_\contour \mach_0'\):
@@
 
-      Above the stack pointer, memory is high integrity in an extension of safe initialization;
-      from the perspective of the callee, that memory is not initialized. Below the old stack
-      pointer, memory is high integrity and high confidentiality. Registers and the callee's frame
-      are low integrity and low confidentiality. These categories inform our integrity and
-      confidentiality properties.
+      Above the stack pointer, memory is high integrity; from the perspective of the callee, that
+      memory is not initialized. Below the old stack pointer, memory is high integrity and high
+      confidentiality. Registers and the callee's frame are low integrity and low confidentiality.
+      These categories inform our integrity and confidentiality properties.
 
@@
       For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...\)
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\)
       and \(\callmap(\mach_c) = n\), if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow ...
-        \mpstate{\mach_r}{\pol_r,\_} | \ret{\mach_c}\):
+        \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\):
 
@@
       For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
-      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... (\mach_c,\pol_c,\_) ...\) and
+      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\) and
       \(\callmap(\mach_c) = n\):
@@
         \item For all \(\mach_c'\) such that \(\mach_c \approx_C \mach_c'\):
-        \item If \(\mpstate{\mach_c}{p_c} \hookrightarrow \MPT_1 (\mach_r,p_r,\_) | \ret{\mach_c}\)
+        \item If \(\mpstate{\mach_c}{p_c} \hookrightarrow \MPT_1 \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\)
           \begin{itemize}
@@
               \(\mpstate{\mach_r}{p_r} \hookrightarrow \MPT_2\), and \(\mach_r'' \hookrightarrow \machT_2'\)
-            \item \(\MPT_1 \cdot \MPT_2 \sim_P \MPT_1' \cdot \MPT_2'\)
+            \item \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1' \cdot \machT_2'\)
           \end{itemize}

### 2021-03-30 tex/main.tex
Work on introduction and threat model

--- 2021-03-07 main.tex
@@
 \begin{abstract}
-What is stack safety? Numerous mechanisms have been proposed to guard
-against attacks that corrupt the call stacks of running programs,
-\ifaftersubmission
-and many
-others rely on a protected stack as an assumption to provide higher level
-guarantees,\bcp{this is a good point, but we don't pick it up anywhere in the
-  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
-\bcp{Let's delete for now; later, we can try to incorporate it in the body
-  of the paper.}
+  What is stack safety? Numerous mechanisms have been proposed to guard against
+  attacks that corrupt the call stacks of running programs, including micro-policies
+  implemented using metadata tagging mechanisms such as PIPE. But for specification
+  these implementations largely appeal either to intuitive notions of the behavior
+  of a stack, or to specific examples of bad behavior.
+  %
+  We propose a formal definition phrased in the technical terminology of language-based
+  security---as a combination of an integrity property (a caller's stack frame is
+  invariant over calls) and a confidentiality property (callees lack knowledge of the
+  contexts of their calls). These combine with a straightforward well-bracketedness
+  condition on the program's control flow that ensures that callers and callees remain
+  meaningfully distinct.
+  %
+  While our formalism is informed by the PIPE implementation, the primary properties reflect
+  general principles of stack safety. We define additional properties tailored for specific
+  contexts: a stronger property amenable to efficient testing and a weaker one that is
+  approximated by more efficient, lazy policies.
+
+  \ifaftersubmission\bcp{should we mention stktokens here?}
+  \leo{Not sure. We almost never cite stuff in the abstract, do we?}
+  \sna{On the same note, cite Nick and Andre?}
   \fi
-but a clear and precise definition of stack safety is missing
-from the literature.
-%
-We propose a formal definition phrased in the technical
-terminology of language-based security---as a combination of a
-confidentiality property (callees should not be able to read their caller's
-stack frames) and an integrity property (callees should not corrupt their
-caller's stack frames or control state), together with a straightforward
-well-bracketedness condition on the program's control flow.
-  %
-We discuss how these properties are enforced or approximated by
-existing stack protection mechanisms and describe useful
-stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
-\leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi
 \end{abstract}
@@
 \fi
-
-But what {\em is} this stack safety of which people speak? Precise formal
-definitions are mostly lacking: rather, the notion
-is commonly defined in a negative way, by enumerating the bad things that a
-given enforcement mechanism
+This last provides an array of micro-policies offering protection at various granularities,
+which offer a trade-off between protection and efficiently.
+
+But what {\em is} this stack safety of which they speak? Precise formal
+definitions are mostly lacking: rather, the notion is commonly defined in a
+negative way, by enumerating the bad things that a given enforcement mechanism
 prevents, such as corruption of return addresses, buffer overflows, or
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-Indeed, we are only aware of one quite recent attempt to frame a
-positive definition, by
-\citet{Skorstengaard+19}, which defines stack safety as
-the conjunction of two properties: {\em local state encapsulation} and
-{\em well-bracketed control flow}.
+Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
+\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
+{\em local state encapsulation} and {\em well-bracketed control flow}.
 These properties are formalized by defining a capability-based machine
@@
 there is still room to craft an operational characterization
-of stack safety that is independent of capabilities or any other specific
-enforcement technique.  In other words, we seek formal criteria that can be applied
+of stack safety. In other words, we seek formal criteria that can be applied
 to a mechanism for protecting stack-based low-level programs to judge
-whether or not it actually ``guarantees stack safety.''
+whether or not it actually ``guarantees stack safety.'' While our criteria are
+designed with PIPE micro-policies in mind, they are agnostic about many implementation
+details and should be applicable to other systems with minor adjustment.
 %
@@
 {\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
-overwrite it's caller's stack data. To this end, our properties support a
-{\em call rule} that can be used to reason that on return from {\tt f}, {\tt x} still
-contains the same value. This is not only useful for the programmer: consider the second
+overwrite {\tt main}'s stack data. This expectation is for {\em integrity}: that a
+caller's data is protected from modification by its callee.
+This is not only useful for the programmer: consider the second
 example, where the call to {\tt h} should be unreachable and therefore subject to dead code
 elimination, if only the compiler can soundly reason that {\tt x} does not change.
-The property that enables the call rule is, in security terms, {\em integrity}: callers'
-data is protected from modification by callees.
-
-Alongside integrity, we have {\em confidentiality}, in which callers' data is protected
-from leakage by their callees. Consider again the first example in figure \ref{fig:examples1},
+
+Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
+from leakage by its callee. Consider again the first example in figure \ref{fig:examples1},
 in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
@@
 space, {\tt f} could still deduce information about the argument and behave differently, violating
-the abstraction of its call. So our confidentiality properties also support a {\em insensitivity rule},
-that each function's behavior is independent of the state of its caller, other than arguments.
-This also makes it easier, though not trivial, to prove a function's idempotence: the channels through
-which a callee can remember its own history are limited to those outside the stack.
+the abstraction of its call.
 
@@
 callers from their callees. Consider the last example in figure \ref{fig:examples2},
-a variant of the famous ``awkward example,'' in which {\tt f} is called twice and may not
-be idempotent. It should be the case that {\tt main} prints -1, but if
+a variant of the famous ``awkward example,'' in which {\tt f} is called twice and might store
+data outside the stack between calls. It should be the case that {\tt main} prints -1, but if
 {\tt f}'s first activation stores its return address somewhere accessible, the second
@@
 1 again before a third activation that would return normally. Then {\tt main} would print 1.
-Once again this is rather unexpected behavior.
-
-Preventing this form of reentrancy attack is the intuition behind
+This is rather unexpected behavior.
+
+Preventing this form of control-flow attack is the intuition behind
 \citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
 the call graph itself. A callee should pass control back to the caller only by
-transferring control to the instruction following the call; it should not be allowed
-to jump back into the caller's code at an arbitrary point, either directly (via a jump
-instruction) or indirectly (by corrupting stack data and returning).
+transferring control to the instruction following the call; otherwise it could
+subvert LSE by appearing to return -- thus restoring access to the caller's data --
+while causing the execution of arbitrary instructions on that data.
 
@@
 call, we assume that the callee can execute arbitrary instructions
-until it returns.
-
-A key challenge in adapting the concepts of language-based security to this
-setting is that integrity and confidentiality are ``nested'' properties: not
-only does a given caller need to be protected from its callee, but the callee
-also needs to be protected from {\em its} callees, and so on. That is, the notion
-of ``attacker'' changes at many points throughout the execution of the program.
-We address this issue by introducing a dynamically model of {\em domains} for
-the different components of the machine state, in which some stack locations are
-sealed by the caller that needs to preserve them while the rest are unsealed and
-available for use.
+until it returns. Thus we have no need to differentiate safe code
+from adversarial - all functions, and indeed all activations of the same function,
+are mutually distrusting.
+%
+Even more fundamentally, we need to know when a call or a return is taking
+place. We assume that calls and returns are annotated in the code, so that
+we can track the point at which control transfers from the protected caller
+to the attacking callee -- and when that callee in turn becomes a protected
+caller in its own right.
+
+This ``nesting'' of properties is a key challenge in adapting the concepts
+of language-based security to the stack setting. We address this issue by
+turning any set of annotations into a dynamic model of {\em domains},
+in which some stack locations are sealed by the caller that needs to preserve
+them while the rest are unsealed and available for use by the active, attacking
+callee.
 
@@
 addresses are passed and whose security must be modeled by a separate property, such as
-memory safety enforced by capabilities.
+memory safety enforced by capabilities. Finally, we extend the model to a coroutine
+system featuring finitely-bounded stacks.
 
@@
 \citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
-We give {\em eager} properties that this policy enforces, in which each
+This policy testably enforces our {\em eager} properties, in which each
 callee, at the point of its return, can be seen to respect the integrity
@@
 faster'' and is therefore more suitable for efficient testing. We prove (in
-Coq) that the ultra-eager property implies our original definition of stack
-safety.
+Coq) that the ultra-eager property implies our original definition of integrity.
 
@@
 caller actually accesses data that the callee has written into its stack
-frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
-  property in the same way...}\fi.
-The lazy policy does not enforce our original integrity and confidentiality
-properties, and to our knowledge there is no formal statement of the properties that
-it ought to enforce. We propose a {\em lazy} integrity property, which capture the
-desired behavior of a policy that detects violations between when they occur and when
-they make a visible impact. Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s
-existing ``Lazy Tagging and Clearing'' policy does not quite enforce these properties,
-but formalizing the ideal property allows us to see how it falls short.
+frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
+  property in the same way...}\fi
+The lazy policy does not enforce our original integrity property, and to our
+knowledge there is no formal statement of the properties that it ought to enforce.
+We propose a {\em lazy} integrity property, which capture the desired behavior of a
+policy that detects integrity violations between when they occur and when they make
+a visible impact. Intuitively this can be phrased, ``No visible reads after dangerous
+writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
+``Lazy Tagging and Clearing'' policy does not quite enforce this property, as shown in
+testing, but formalizing the ideal property allows us to see how it falls short.
 We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
 at some performance cost.
-
-We extend our properties to allow sharing, both in the form of arguments passed
-on the stack and of objects shared widely beyond their owning function. In the latter
-case, restricting access the shared objects is beyond the scope of stack safety,
-but our model delimits them from other objects in the same frame.
 
@@
   % These properties rule out common examples of stack memory abuse.
-  This property can be enforced by
-  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
-  based on hardware tagging~(\cref{sec:conservative}).
 \item
-  A {\em lockstep} variant of the stack-safety property that is also
-  enforced by the Depth Isolation policy
-  but is more efficient to test, plus a proof (in Coq) that this variant is
-  stronger than the original~(\cref{sec:lockstep}).
+  A property-based randomized testing framework that demonstrates that this
+  property can be enforced by the Depth Isolation policy of
+  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware tagging~(\cref{sec:conservative}).
+\item
+  An {\em ultra-eager} variant of the stack-integrity property that is also
+  enforced by the Depth Isolation policy but is more efficient to test, plus a proof
+  (in Coq) that this variant is stronger than the original~(\cref{sec:lockstep}).
 \item
@@
   variant is weaker than the original~(\cref{sec:lazy}).
+\item
+  Standardized trace-based formulations of eager confidentiality and of ultra-eager,
+  eager, and lazy integrity, parameterizeable by arbitrary control-flow models and assignments
+  of machine components to {\em LSE domains}.
+\item
+  An LSE-domain model tracking the passing of function arguments on the stack and the sharing
+  of arbitrary stack data.
+\item
+  An extension of stack safety to a simple coroutine model with corresponding LSE
+  properties and a well-structured control flow property that subsumes WBCF for each
+  stack as well as guaranteeing legitimate coroutine crossings.
 %  This policy does not enforce our strict
@@
 Stack safety is meant to protect a caller from a confused or malicious callee's
-attempts to access private data or hijack its control flow. Our
-attacker model reflects this intention by assuming that the attacker
-can run arbitrary machine code to try to access or overwrite its caller's stack data
-or disrupt its control state.  Hardware and timing attacks are out of
-scope.
+attempts to access private data or hijack its control flow. And in fact the caller,
+while presumably not malicious, may itself be confused or erroneous. Our attacker
+model reflects this intention by assuming that both the attacker and the defender
+may run arbitrary machine code. Hardware and timing attacks are out of scope.
+
 %\leo{Anything else to add here?}
@@
 
-We assume we are given a machine language program, together with some auxiliary information
-about its structure: the code footprint of each function and the locations of instructions
-representing calls, returns, and function entry points.
-Although this information would typically come from a compiler together with
-the machine code, we do not
-assume anything about their provenance.  We assume that the machine has a
-program counter $\PCname$
-and stack pointer $\rsp$; other details of the machine's ISA are unimportant.
-
-To make our stack safety definitions precise, we make some straightforward
-assumptions about
-memory layout; our definitions could easily be changed to be parametric over
-these assumptions, at the cost of some complexity in their statements.  Memory
-consists of an instruction region and a stack that grows upwards. (For simplicity,
+In this very strong attacker model, there is little obvious distinction between
+the attacker and the defender. They are separated only by the control transfer
+inherent in their calls and returns. Our stack-safety concepts therefore depend
+critically on the notions of ``call'' and ``return,'' which are not fully explicit
+in machine code (the same instruction opcode may be used in some places as part of
+a call or return sequence and in other places for other purposes).
+%
+We assume we are given a machine language program annotated with the locations of
+instructions representing calls, returns, and function entry points. Although this
+information would typically come from a compiler together with the machine code,
+we do not assume anything about their provenance. We do treat the execution of an
+annotated instruction as in some sense intentional on the part of the current
+function. No accidental calls or returns, and attackers control their own returns
+but cannot change the annotations on the caller's code.
+
+\paragraph{Well-formedness and Bad Annotations}
+
+Since we rely so heavily on code annotations, we must ask: what if the annotations
+are wrong? What if a call or return occurs that is not annotated, or an annotation
+fails to correspond to an actual return? In the former case, the answer is simple --
+there has been no call or return, only instruction resembling such an operation, and
+therefore no associated change in permissions. This is particularly interesting in the
+case of a false return, where control may appear to have returned to the caller, but
+the enforcement mechanism must still protect the caller's memory.
+
+The latter case depends on the enforcement mechanism and the nature of the false
+annotation. A given enforcement mechanism might place a static well-formedness criterion
+on the code it applies to. For instance, our micro-policies expect that call and return
+annotations occur on jump instructions. On the other hand, not all jumps are proper
+calls or returns, only those preceded by appropriate sequences: in these cases the
+bad annotation might appear well formed but the system should either respect the change
+in permissions or, if doing so would compromise security, fail-stop rather than
+execute the annotated instruction.
+
+\paragraph{System Assumptions}
+
+Our primary target is a RISC-V architecture extended with the PIPE architecture, but the
+model makes very few assumptions about the machine it applies to. We assume a register
+machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other details of
+the machine's ISA are unimportant.
+
+Our definitions are largely parametric over the layout of the stack in memory, but for
+purposes of explanation we will describe a standard layout. Memory consists of an
+instruction region and a stack that grows upwards. (For simplicity,
 we do not consider the heap. Integrating a memory-safety property for the
-heap in the style of
-\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
-in~\Cref{sec:relwork}.)  The stack has a conventional structure, with
+heap in the style of \citet{DBLP:conf/post/AmorimHP18} is a natural extension
+that we discuss in~\Cref{sec:relwork}.) The stack has a conventional structure, with
 contiguously allocated per-function frames.
+
 The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
-at the time of the call.  The stack can be used to
-pass arguments or return results in the usual way, by placing them
-at the top of the caller's frame. Callers and callees can also communicate
-via registers if they wish; indeed, our
-definitions say nothing at all about registers (except $\PCname$ and
-$\rsp$). \ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
+at the time of the call. The stack can be used to pass arguments or return results
+in the usual way, by placing them at the top of the caller's frame. This will be modeled
+as a form of sharing in section \sna{cite the sharing section}.
+Callers and callees can also communicate via registers if they wish; indeed, our
+definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
+
+\ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
   convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
 
-Our stack-safety concepts depend critically on the notions of ``call'' and ``return,''
-which are not fully explicit in machine code (the same instruction opcode may be
-used in some places as part of a call or return sequence and in other places
-for other purposes). We assume the compiler's cooperation in labeling instructions
-with a variety of annotations, and in fact in identifying them as instructions.
-A {\em call-map} is a partial map from addresses to pairs of a function identifier --
-every instruction belongs to a function -- and a set of annotations. This gives
-us flexibility to have instructions simultaneously play multiple roles.
-
-Enforcement mechanisms are modeled as {\em policies} (borrowing the
-terminology from tag-based mechanisms) that add some auxiliary policy
-state to the ordinary machine state
-%\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
-and a partial policy step function that refines the machine step
-function by allowing or disallowing a step to be taken based on its
-auxiliary state.  Enforcement works by halting the program before it
-can perform an action that would violate the security property; our
+\paragraph{Policies}
+
+We model the base machine as a step function over machine states, and a machine enhanced with
+an enforcement mechanism (in the language of tag-based enforcement, a {\em policy}) as 
+a partial step function over machine states extended with some auxiliary state, refining
+the base machine. When the policy-enhanced machine fails to step, this is a {\em failstop},
+halting the program before it can perform an action that would violate the security property; our
 properties are therefore naturally \emph{termination insensitive.}
@@
 
-We model ordinary program state and policy-enforcement state separately to make it
-easier to use well-established concepts from language-based security.
-In particular, confidentiality has an elegant extensional characterization in terms of
-\emph{noninterference}: a program preserves the confidentiality of certain data if
-\emph{varying} that data would not change the program's observable
-behavior~\citep{6234468}.
-In applying this idea to a system that incorporates policy enforcement, only
-ordinary program data should be varied, while any associated policy
-state should be left alone; our model makes this distinction obvious.
+%We model ordinary program state and policy-enforcement state separately to make it
+%easier to use well-established concepts from language-based security.
+%In particular, confidentiality has an elegant extensional characterization in terms of
+%\emph{noninterference}: a program preserves the confidentiality of certain data if
+%\emph{varying} that data would not change the program's observable
+%behavior~\citep{6234468}.
+%In applying this idea to a system that incorporates policy enforcement, only
+%ordinary program data should be varied, while any associated policy
+%state should be left alone; our model makes this distinction obvious.
 %That is, we compare the actual execution of the program with a \emph{hypothetical}
 %execution where data has been varied.
-We will also introduce a notion of {\em observational integrity} that relies on the idea of
-\emph{rolling back} program state; again, our model makes it easy to describe
-rollback of ordinary program data while leaving policy state alone.
+%We will also introduce a notion of {\em observational integrity} that relies on the idea of
+%\emph{rolling back} program state; again, our model makes it easy to describe
+%rollback of ordinary program data while leaving policy state alone.
 
@@
 
-This enforcement model directly describes
-micropolicies~\citep{pump_oakland2015}, and we believe that it is
-flexible enough to capture code-altering approaches (such as code rewriting
-or capability-based techniques).
+This enforcement model directly describes micropolicies~\citep{pump_oakland2015}.
+We believe that it is flexible enough to capture code-altering approaches, such as code
+rewriting or capability-based techniques, perhaps with less elegance.
 
@@
 
-\paragraph*{Capability machines}
-%
-%Other policies are implemented in hardware or built on hardware mechanisms.
-Capability machines such as CHERI \citep{Woodruff+14} extend conventional
-architectures to support efficient and fine-grained control over memory
-accesses, which can be used to implement security policies, including
-the stack protection policies of
-\citet{Skorstengaard+19b} and (with an extension to support linear capabilities) \citet{Skorstengaard+19}.
-%\bcp{I thought these policies used
-%  linear capabilities?}\leo{Only the second one AFAIK. What would you
-%  change to characterize it more accurately?}\rb{StkPointers uses linear
-%  capabilities, its predecessor users local capabilities (and unlike
-%  StkPointers could be implemented in CHERI)}.
-Capabilities package up pointers
-(ordinary data) with base and bounds information (policy data),
-and they require rewriting source code (e.g., to use capabilities in place
-of ordinary pointers), so separating ordinary machine state from policy
-state is delicate; we conjecture it can be done with some effort.
-
-%% Mapping the underlying protection
-%% mechanisms to policy states and step function would be a delicate operation, as
-%% those are tightly coupled to the hardware design. However, separating the
-%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
-%% would encompass details like the abstract call stack, the capability sealing
-%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-%% the machine would be relatively straightforward.  \bcp{Not sure that any of
-%%   this is comprehensible... :-(}
-
-\paragraph*{Software-only policies}
-%
-Some policies are implemented in software, by modifying code in the
-compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
-to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
-be evaluated in terms of our formal notion of stack safety with a trivial null
-policy. But any state maintained by the enforcement mechanism must be in normal
-memory and subject to the same confidentiality and integrity concerns (to be
-varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
-out as policy state provided that it can be proven to protect itself.
-
 \ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
@@
 
-\subsection{Eager Integrity and the Call Rule}
-
-\definition Let \(\mpcstatename\) be a reachable state whose PC value is annotated
-as a call and whose call depth is \(\depth\). Assume \(\mpcstatename \stepstocon \mpcstatename'\).
-Let \(\MPCT\) be the trace ending after the corresponding return, if any, defined by
-\(\mpcstatename' \hookrightarrow \MPCT | f\) with \(f ~ \mach ~ \context\) holding when
-\(\context = \_,\depth\). Finally let \(\components\) be the set of components \(\component\)
-such that \(\sealcon ~ \pi_\mach(\mpcstatename) ~ \component\) holds. Then by the {\em Call Rule},
-it is sound to assume that if \(\last(\pi_\mach(\MPCT)) = \mach'\), then
-\(\pi_\mach(\mpcstatename) ~ \component = \mach' ~ \component\).
+\subsection{Eager Integrity}
 
@@
 
-\subsection{Eager Confidentiality and the Insensitivity Rule}
+\subsection{Eager Confidentiality}
 
@@
 This is written \(\mach,\mach' \diamond \nach,\nach'\).
-
-\definition Let \(\mpcstatename\) be a reachable call state. Assume the that the call is successful
-and \(\mpcstatename \stepstocon \mpcstatename'\). Then let \(\npcstatename'\) be a \(\components\)-variant
-where \(\components\) is the set of all addresses in the stack. Then by the {\em Insensitivity Rule},
-if the trace from \(\mpcstatename'\) eventually returns, so does that from \(\npcstatename'\),
-their states after the return have the same differences, and up to return their traces are observationally
-equivalent.
 
@@
 {\em well-bracketed control flow} (with respect to those maps).
-
-\section{Programming Logic Perspective}
 
@@
 
-\section{Stack Safety Properties for Lazy Enforcement}
+\section{Stack Safety for Lazy Enforcement}
 \label{sec:lazy}
@@
 
-  \paragraph{Stack-derived Pointers}
-
-    Common programming idioms involve a caller passing a pointer to
-    its local data into a callee. Using standard notions of pointer
-    provenance~\citep{provenance}, we can extend our model to
-    distinguish safe use of this idiom from true stack safety
-    violations. We require additional annotation to identify which
-    addresses within a function's stack frame correspond to distinct
-    objects. Then a valid pointer to an object is one derived through
-    legal arithmetic from its base address, and at each call point, an
-    object's addresses are marked low confidentiality and integrity if
-    a valid pointer exists in a register or in the transitive closure
-    of the callee's accessible memory. All pointers to an object cease
-    to be valid when it is deallocated along with its stack frame.
-
-  \paragraph{Non-stack control flow}
-
-    Control structures beyond well-bracketed calls and returns require
-    modifications to the model. Tail calls, for instance, reuse the caller's
-    stack frame for its callee, and every nested tail call returns
-    simultaneously to the top non-tail-calling function. Under normal stack
-    integrity, the first function in a chain of tail calls returns and violates
-    integrity, while subsequent tail calls never actually return at all.
-    Instead we need a variant that treats the entry of a tail call as the
-    return of its caller.
-
-    More complex is the addition of a coroutine model, in which multiple stacks
-    respect stack safety internally, but might also yield to one another.
-    Here we must distinguish three levels of integrity: accessible, in the same
-    stack, and in another stack. Stack integrity requires that same-stack data
-    be unchanged when a function returns, but other-stack data may change. A
-    similar {\em yield integrity} property requires that, from a yield out of
-    a coroutine to the next yield back into it, all accessible and same-stack
-    data are unchanged. Confidentiality properties are split similarly.
-
-  \paragraph{Random Testing}
-
-  Part of the motivation for the strong lockstep stack-safety property
-  is that it uncovers errors quickly, and therefore should be quicker
-  to test, just like the smarter testing properties
-  of~\citet{TestingNI:ICFP}. We have already developed a testing
-  framework for debugging our policy enforcement for this property, and
-  it would be interesting to explore whether a testing-amenable variant
-  of the lazy property also exists.
+  \paragraph*{Capability machines}
+  %
+  %Other policies are implemented in hardware or built on hardware mechanisms.
+  Capability machines such as CHERI \citep{Woodruff+14} extend conventional
+  architectures to support efficient and fine-grained control over memory
+  accesses, which can be used to implement security policies, including
+  the stack protection policies of
+  \citet{Skorstengaard+19b} and (with an extension to support linear capabilities) \citet{Skorstengaard+19}.
+  %\bcp{I thought these policies used
+  %  linear capabilities?}\leo{Only the second one AFAIK. What would you
+  %  change to characterize it more accurately?}\rb{StkPointers uses linear
+  %  capabilities, its predecessor users local capabilities (and unlike
+  %  StkPointers could be implemented in CHERI)}.
+  Capabilities package up pointers
+  (ordinary data) with base and bounds information (policy data),
+  and they require rewriting source code (e.g., to use capabilities in place
+  of ordinary pointers), so separating ordinary machine state from policy
+  state is delicate; we conjecture it can be done with some effort.
+
+  %% Mapping the underlying protection
+  %% mechanisms to policy states and step function would be a delicate operation, as
+  %% those are tightly coupled to the hardware design. However, separating the
+  %% enforcement aspects of a concrete policy (such as StkTokens, where policy state
+  %% would encompass details like the abstract call stack, the capability sealing
+  %% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
+  %% the machine would be relatively straightforward.  \bcp{Not sure that any of
+  %%   this is comprehensible... :-(}
+
+  %\paragraph*{Software-only policies}
+  %
+  %Some policies are implemented in software, by modifying code in the
+  %compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
+  %to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
+  %be evaluated in terms of our formal notion of stack safety with a trivial null
+  %policy. But any state maintained by the enforcement mechanism must be in normal
+  %memory and subject to the same confidentiality and integrity concerns (to be
+  %varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
+  %out as policy state provided that it can be proven to protect itself.
+
+%  \paragraph{Stack-derived Pointers}
+
+%    Common programming idioms involve a caller passing a pointer to
+%    its local data into a callee. Using standard notions of pointer
+%    provenance~\citep{provenance}, we can extend our model to
+%    distinguish safe use of this idiom from true stack safety
+%    violations. We require additional annotation to identify which
+%    addresses within a function's stack frame correspond to distinct
+%    objects. Then a valid pointer to an object is one derived through
+%    legal arithmetic from its base address, and at each call point, an
+%    object's addresses are marked low confidentiality and integrity if
+%    a valid pointer exists in a register or in the transitive closure
+%    of the callee's accessible memory. All pointers to an object cease
+%    to be valid when it is deallocated along with its stack frame.
+
+%  \paragraph{Non-stack control flow}
+
+%    Control structures beyond well-bracketed calls and returns require
+%    modifications to the model. Tail calls, for instance, reuse the caller's
+%    stack frame for its callee, and every nested tail call returns
+%    simultaneously to the top non-tail-calling function. Under normal stack
+%    integrity, the first function in a chain of tail calls returns and violates
+%    integrity, while subsequent tail calls never actually return at all.
+%    Instead we need a variant that treats the entry of a tail call as the
+%    return of its caller.
+
+%    More complex is the addition of a coroutine model, in which multiple stacks
+%    respect stack safety internally, but might also yield to one another.
+%    Here we must distinguish three levels of integrity: accessible, in the same
+%    stack, and in another stack. Stack integrity requires that same-stack data
+%    be unchanged when a function returns, but other-stack data may change. A
+%    similar {\em yield integrity} property requires that, from a yield out of
+%    a coroutine to the next yield back into it, all accessible and same-stack
+%    data are unchanged. Confidentiality properties are split similarly.
+
+%    \paragraph{Random Testing}
+
+%    Part of the motivation for the strong lockstep stack-safety property
+%    is that it uncovers errors quickly, and therefore should be quicker
+%    to test, just like the smarter testing properties
+%    of~\citet{TestingNI:ICFP}. We have already developed a testing
+%    framework for debugging our policy enforcement for this property, and
+%    it would be interesting to explore whether a testing-amenable variant
+%    of the lazy property also exists.
 

### 2021-03-31 tex/example1.tex
Work on examples

\newcommand*{\add}{\textsc{add}}
\newcommand*{\addi}{\textsc{addi}}
\newcommand*{\sw}{\textsc{sw}}
\newcommand*{\lw}{\textsc{lw}}
\newcommand*{\jal}{\textsc{jal}}
\newcommand*{\jalr}{\textsc{jalr}}
%\newcommand*{\rsp}{\textsc{sp}}
\newcommand*{\rra}{\textsc{ra}}
\newcommand*{\rout}{\textsc{out}}

\newcommand*{\tagInstr}{\textsc{instr}}
\newcommand*{\tagCall}{\textsc{call}}
\newcommand*{\tagHa}{\textsc{h1}}
\newcommand*{\tagHb}{\textsc{h2}}
\newcommand*{\tagRa}{\textsc{r1}}
\newcommand*{\tagRb}{\textsc{r2}}
\newcommand*{\tagRc}{\textsc{r3}}
\newcommand*{\tagNoDepth}{\textsc{unused}}
\newcommand*{\tagStackDepth}[1]{\textsc{stack} ~ #1}
\newcommand*{\tagPCDepth}[1]{\textsc{pc} ~ #1}
\newcommand*{\tagSP}{\textsc{sp}}

%\newcounter{pcctr}
%\setcounter{pcctr}{0}
%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
%\newcommand*{\row}[4]{
%  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\
%}
%\newcommand*{\tracerow}[5]{
%  #1 & #2 & #3 & #4 & \text{#5}\\
%}
%\newcommand*{\summary}[6]{
%\[
%\begin{array}{cccccc}
%  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
%\end{array}
%\]
%}
\newcommand{\negate}{\textrm{-}}

\colorlet{lgray}{gray!40}

\newcommand{\instrc}{lgray}
\newcommand{\mainsealc}{cyan}
\newcommand{\fsealc}{green}
\newcommand{\unsealc}{gray}
\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
\newcommand{\fulloutc}{white}

\begin{figure}

\begin{subfigure}{.4\textwidth}
\begin{verbatim}
int main() {
0:     int x = 42, y = 0;
1[C]:  y = f(10);
2:     print (x + y);
}

int f(int z) {
10:     print z;
11[R]:  return (z+z);
}
\end{verbatim}
\caption{A Simple Call}
\label{fig:simple-program}
\end{subfigure}
\begin{subfigure}{.4\textwidth}  
\begin{center}
\begin{tabular}{l l}
{\tt 0} &
\memoryaddrs{4em}
\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
~$\cdots$
\\
{\tt 1} &
\memoryaddrs{12em}
\memory{3}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{42}
\MemoryLabel{-10em}{0.75em}{0}
\vspace{.5em}
\\
{\tt 10} &
\memoryaddrs{12em}
\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
\memory{1}{\unsealc}
~$\cdots$
\\
{\tt 2} &
\memoryaddrs{12em}
\memory{3}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{42}
\MemoryLabel{-10em}{0.75em}{0}
\\
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}

\caption{Trace of Stack Domains}

\label{fig:simple-trace}
\end{subfigure}
\end{figure}

This section will elucidate the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will insert our model between the enforcement and
the examples, arguing that:

\begin{itemize}
\item Examples that appear to violate stack safety, violate at least one of our properties
\item We do not find desireable behaviors that appear stack-safe yet violate our properties
\item Our properties are testably implemented by stack safety micro-policies from the literature,
  and the micro-policies are not significantly more conservative (i.e., policy failstops tend to
  correspond to imminent property violations)
\end{itemize}

\Cref{fig:simple-program} shows a simple example, written in C-like pseudo-code.
It consists of a {\tt main} function that performs a call to a function {\tt f},
which prints its argument, performs arithmetic, and returns a result.
\Cref{fig:simple-trace} shows the stack as it is divided into domains throughout
the program execution. We use high-level code for compactness, but reason about
a standard compilation of each function. Line numbers correspond to the code
addresses just before the code that implements the line, and annotations
{\tt C} and {\tt R} indicate that an instruction is annotated as a call or
return, respectively.

We also assume that the machine is word- (not byte-) addressed and that the
machine communicates with the outside world through a single memory-mapped
output port whose address is elided (so a {\tt print} statement will be
compiled to a store to this location), and that the stack grows upwards.

In the initial memory layout the entire stack is currently {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals (as designated by the stack pointer) are
{\em sealed} and labeled with the depth of their owner, in this case 0.
Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. Later we will
loosen this condition in the interest of extensionality -- the caller only
cares about changes it can see after the return, not changes during execution.

Suppose we insert an instruction after line 10, that stores 5 to the stack pointer - 1.
Then we would be changing {\tt y}, which is sealed, in violation of stepwise integrity.

\begin{figure}

\begin{subfigure}{.35\textwidth}
\begin{verbatim}
int main() {
0:     int x = 42, y = 0;
1[C]:  y = f();
2:     print (x + y);
}

int f() {
10:     int z = *($SP-1);
11[R]:  return (z+z);
}
\end{verbatim}
\caption{A Confidentiality Violation}
\label{fig:conf1-program}
\end{subfigure}
\begin{subfigure}{.5\textwidth}  
\begin{center}
\begin{tabular}{l l l l}
{\tt 0} &
\memoryaddrs{4em}
\memory{3}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{a}
\MemoryLabel{-10em}{0.75em}{b}
\MemoryLabel{-7em}{0.75em}{c}
\vspace{.5em}
& &
\\
{\tt 1} &
\memoryaddrs{12em}
\memory{3}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{42}
\MemoryLabel{-10em}{0.75em}{0}
\MemoryLabel{-7em}{0.75em}{c}
\vspace{.5em} & &
\\
{\tt 10} &
\memoryaddrs{12em}
\memory{2}{\mainsealc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{42}
\MemoryLabel{-10em}{0.75em}{0}
\MemoryLabel{-7em}{0.75em}{c} &
\memoryaddrs{12em}
\memory{2}{\mainsealc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{a'}
\MemoryLabel{-10em}{0.75em}{b'}
\MemoryLabel{-7em}{0.75em}{c'}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{42}
\MemoryLabel{-10em}{0.75em}{0}
\MemoryLabel{-7em}{0.75em}{0} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{a'}
\MemoryLabel{-10em}{0.75em}{b'}
\MemoryLabel{-7em}{0.75em}{b'}
\\
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}

\caption{Trace of Stack Domains}

\label{fig:conf1-trace}
\end{subfigure}
\end{figure}

Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
of the context of its call (with the exception, of course, of its arguments).
We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
security. Noninterference describes knowledge: it states that for some component(s)
that we do not know, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the components by observing those results, violating
confidentiality.

In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt y}, which was not passed to it. The trace
in figure \ref{fig:conf1-trace} shows how noninterference works. First, let \(a\),
\(b\), and \(c\) be the unknown initial values at their locations. On entry to {\tt f}
we again have 42 and 0 in {\tt x} and {\tt y}, and {\tt z} contains \(c\). We consider a
{\em variant state} that replaces the values of components that should be secret with
arbitrary values \(a'\), \(b'\), and \(c'\). After a step, we've copied 0 and \(c'\),
respectively, into {\tt z}. Since {\tt z} changed value, for confidentiality to hold,
its value should be the same in all variant states. But when \(c' \not = 0\), the
states changes differently, which violates confidentiality. It would also be a violation
to print the secret value.

This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. Any components
whose values update and any printed values will be identical between the primary and variant
traces.

\begin{figure}

\begin{subfigure}{.35\textwidth}
\begin{verbatim}
int main() {
0[C]:  int t = f();
1[R]:  return;
}

int f() {
10:     int x, y = 5, z = 6;
11:     print y+z;
12:     $SP-=2;
13[C]:  x = g();
14[R]:  return x;
}

int g() {
20:     int z = *$SP;
21[R]:  return (z+z);
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.5\textwidth}  
\begin{center}
\begin{tabular}{l l l l}
{\tt 13} &
\memoryaddrs{12em}
\memory{1}{\mainsealc}
\memory{3}{\unsealc}
~$\cdots$
\MemoryLabel{-18em}{0.75em}{a}
\MemoryLabel{-14em}{0.75em}{b}
\MemoryLabel{-10em}{0.75em}{5}
\MemoryLabel{-7em}{0.75em}{6}
\vspace{.5em}
& &
\\
{\tt 20} &
\memoryaddrs{12em}
\memory{1}{\mainsealc}
\memory{1}{\fsealc}
\memory{2}{\unsealc}
~$\cdots$
\MemoryLabel{-18em}{0.75em}{a}
\MemoryLabel{-14em}{0.75em}{b}
\MemoryLabel{-10em}{0.75em}{5}
\MemoryLabel{-7em}{0.75em}{6}
\vspace{.5em} &
\memoryaddrs{12em}
\memory{1}{\mainsealc}
\memory{1}{\fsealc}
\memory{2}{\unsealc}
~$\cdots$
\MemoryLabel{-18em}{0.75em}{a'}
\MemoryLabel{-14em}{0.75em}{b'}
\MemoryLabel{-10em}{0.75em}{c'}
\MemoryLabel{-7em}{0.75em}{d'}
\vspace{.5em}
&
\\
{\tt 21} &
\memoryaddrs{12em}
\memory{1}{\mainsealc}
\memory{1}{\fsealc}
\memory{2}{\unsealc}
~$\cdots$
\MemoryLabel{-18em}{0.75em}{a}
\MemoryLabel{-14em}{0.75em}{b}
\MemoryLabel{-10em}{0.75em}{5}
\MemoryLabel{-7em}{0.75em}{6}
\vspace{.5em} &
\memoryaddrs{12em}
\memory{1}{\mainsealc}
\memory{1}{\fsealc}
\memory{2}{\unsealc}
~$\cdots$
\MemoryLabel{-18em}{0.75em}{a'}
\MemoryLabel{-14em}{0.75em}{b'}
\MemoryLabel{-10em}{0.75em}{c'}
\MemoryLabel{-7em}{0.75em}{d'}
\vspace{.5em}
&
\\

\end{tabular}
\end{center}

\vspace{\abovedisplayskip}

\end{subfigure}

\caption{Another Confidentiality Violation}
\label{fig:conf2}
\end{figure}

Two important details of confidentiality are reflected in figure \ref{fig:conf2}.

### 2021-04-01 tex/example1.tex
Extended examples a bit more

@@
 
-%\newcounter{pcctr}
-%\setcounter{pcctr}{0}
-%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
-%\newcommand*{\row}[4]{
-%  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\
-%}
-%\newcommand*{\tracerow}[5]{
-%  #1 & #2 & #3 & #4 & \text{#5}\\
-%}
-%\newcommand*{\summary}[6]{
-%\[
-%\begin{array}{cccccc}
-%  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
-%\end{array}
-%\]
-%}
 \newcommand{\negate}{\textrm{-}}
@@
 
-\newcommand{\instrc}{lgray}
 \newcommand{\mainsealc}{cyan}
 \newcommand{\fsealc}{green}
-\newcommand{\unsealc}{gray}
+\newcommand{\unsealc}{lgray}
 \newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
@@
 \end{verbatim}
+
 \caption{A Simple Call}
@@
 {\tt 10} &
-\memoryaddrs{12em}
+\memoryaddrs{17em}
 \memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
-\memory{1}{\unsealc}
+\memory{2}{\unsealc}
 ~$\cdots$
+\MemoryLabel{-10em}{0.75em}{2}
 \\
@@
 \MemoryLabel{-10em}{0.75em}{0}
+\MemoryLabel{-7em}{0.75em}{2}
 \\
@@
 
-\begin{subfigure}{.35\textwidth}
+\begin{subfigure}{.33\textwidth}
 \begin{verbatim}
@@
 \end{verbatim}
-\caption{A Confidentiality Violation}
 \label{fig:conf1-program}
 \end{subfigure}
-\begin{subfigure}{.5\textwidth}  
-\begin{center}
-\begin{tabular}{l l l l}
+\begin{subfigure}{.59\textwidth}
+\centering
+\begin{tabular}{l l l}
   {\tt 0} &
+  \multicolumn{2}{c}{
     \memoryaddrs{4em}
@@
     \vspace{.5em}
-& &
-\\
+  } \\
   {\tt 1} &
+  \multicolumn{2}{c}{
     \memoryaddrs{12em}
@@
     \MemoryLabel{-7em}{0.75em}{c}
-\vspace{.5em} & &
-\\
+    \vspace{.5em}
+  } \\
   {\tt 10} &
-\memoryaddrs{12em}
+  \memoryaddrs{20em}
   \memory{2}{\mainsealc}
-\memory{1}{\unsealc}
+  \memory{2}{\unsealc}
   ~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\MemoryLabel{-7em}{0.75em}{c} &
-\memoryaddrs{12em}
+  \MemoryLabel{-19em}{0.75em}{42}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf d}
+ &
+  \memoryaddrs{20em}
   \memory{2}{\mainsealc}
-\memory{1}{\unsealc}
+  \memory{2}{\unsealc}
   ~$\cdots$
-\MemoryLabel{-14em}{0.75em}{a'}
-\MemoryLabel{-10em}{0.75em}{b'}
-\MemoryLabel{-7em}{0.75em}{c'}
+  \MemoryLabel{-19em}{0.75em}{a'}
+  \MemoryLabel{-14em}{0.75em}{b'}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf d'}
   \\
   {\tt 11} &
-\memoryaddrs{16em}
+  \memoryaddrs{20em}
   \memory{2}{\mainsealc}
-\memory{1}{\unsealc}
+  \memory{2}{\unsealc}
   ~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\MemoryLabel{-7em}{0.75em}{0} &
-\memoryaddrs{16em}
+  \MemoryLabel{-19em}{0.75em}{42}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf 0}
+  &
+  \memoryaddrs{20em}
   \memory{2}{\mainsealc}
-\memory{1}{\unsealc}
+  \memory{2}{\unsealc}
   ~$\cdots$
-\MemoryLabel{-14em}{0.75em}{a'}
-\MemoryLabel{-10em}{0.75em}{b'}
-\MemoryLabel{-7em}{0.75em}{b'}
+  \MemoryLabel{-19em}{0.75em}{a'}
+  \MemoryLabel{-14em}{0.75em}{b'}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf b'}
   \\
 \end{tabular}
-\end{center}
 
@@
 
-\caption{Trace of Stack Domains}
-
 \label{fig:conf1-trace}
 \end{subfigure}
+\caption{A Confidentiality Violation}
 \end{figure}
@@
 
-\begin{subfigure}{.35\textwidth}
+\begin{subfigure}{.32\textwidth}
 \begin{verbatim}
@@
 int g() {
-20:     int z = *$SP;
+20:     int z;
 21[R]:  return (z+z);
@@
 Two important details of confidentiality are reflected in figure \ref{fig:conf2}.
+First of all, it has multiple calls, and while we focus on the stack at the call
+from line 13 to line 20, it should be clear that we also need to consider a similar
+variant at the call from 0 to 10. Confidentiality is in this sense recursive: each
+nested must obey confidentiality in its own right.
+
+More importantly, careful readers will have noticed that the informal definition
+of stack confidentiality refers to a variation of the entire stack, not just the
+sealed portion. In fact, we support temporal stack safety, in which values left behind
+by previous activations are also inaccessible. In this case, we connect temporal safety
+firmly to the caller-callee attacker model -- {\tt f} stored some data in the variables
+{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.
+After the stack pointer is written over {\tt y}, the value of {\tt z} is still
+left behind. Then {\tt g} can read it just by allocating and accessing an uninitialized
+variable!
+
+\paragraph*{Awkward Examples}

### 2021-04-02 tex/example1.tex
More work on examples and intro

--- 2021-03-31 tex/example1.tex
@@
 
-%\newcounter{pcctr}
-%\setcounter{pcctr}{0}
-%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
-%\newcommand*{\row}[4]{
-%  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\
-%}
-%\newcommand*{\tracerow}[5]{
-%  #1 & #2 & #3 & #4 & \text{#5}\\
-%}
-%\newcommand*{\summary}[6]{
-%\[
-%\begin{array}{cccccc}
-%  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
-%\end{array}
-%\]
-%}
 \newcommand{\negate}{\textrm{-}}
-
-\colorlet{lgray}{gray!40}
-
-\newcommand{\instrc}{lgray}
-\newcommand{\mainsealc}{cyan}
-\newcommand{\fsealc}{green}
-\newcommand{\unsealc}{gray}
-\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
-\newcommand{\fulloutc}{white}
-
-\begin{figure}
-
-\begin{subfigure}{.4\textwidth}
-\begin{verbatim}
-int main() {
-0:     int x = 42, y = 0;
-1[C]:  y = f(10);
-2:     print (x + y);
-}
-
-int f(int z) {
-10:     print z;
-11[R]:  return (z+z);
-}
-\end{verbatim}
-\caption{A Simple Call}
-\label{fig:simple-program}
-\end{subfigure}
-\begin{subfigure}{.4\textwidth}  
-\begin{center}
-\begin{tabular}{l l}
-{\tt 0} &
-\memoryaddrs{4em}
-\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
-~$\cdots$
-\\
-{\tt 1} &
-\memoryaddrs{12em}
-\memory{3}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\vspace{.5em}
-\\
-{\tt 10} &
-\memoryaddrs{12em}
-\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
-\memory{1}{\unsealc}
-~$\cdots$
-\\
-{\tt 2} &
-\memoryaddrs{12em}
-\memory{3}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\caption{Trace of Stack Domains}
-
-\label{fig:simple-trace}
-\end{subfigure}
-\end{figure}
-
-This section will elucidate the intuition behind the properties through a series
-of examples. Existing stack safety enforcement work mostly refers directly to examples
-of bad behavior that it prevents; we will insert our model between the enforcement and
-the examples, arguing that:
-
-\begin{itemize}
-\item Examples that appear to violate stack safety, violate at least one of our properties
-\item We do not find desireable behaviors that appear stack-safe yet violate our properties
-\item Our properties are testably implemented by stack safety micro-policies from the literature,
-  and the micro-policies are not significantly more conservative (i.e., policy failstops tend to
-  correspond to imminent property violations)
-\end{itemize}
-
-\Cref{fig:simple-program} shows a simple example, written in C-like pseudo-code.
-It consists of a {\tt main} function that performs a call to a function {\tt f},
-which prints its argument, performs arithmetic, and returns a result.
-\Cref{fig:simple-trace} shows the stack as it is divided into domains throughout
-the program execution. We use high-level code for compactness, but reason about
-a standard compilation of each function. Line numbers correspond to the code
-addresses just before the code that implements the line, and annotations
-{\tt C} and {\tt R} indicate that an instruction is annotated as a call or
-return, respectively.
-
-We also assume that the machine is word- (not byte-) addressed and that the
-machine communicates with the outside world through a single memory-mapped
-output port whose address is elided (so a {\tt print} statement will be
-compiled to a store to this location), and that the stack grows upwards.
-
-In the initial memory layout the entire stack is currently {\em unsealed},
-meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals (as designated by the stack pointer) are
-{\em sealed} and labeled with the depth of their owner, in this case 0.
-Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
-This gives us our simple, intuitive statement of {\em stepwise integrity}:
-if a component is sealed before a step, it will be unchanged afterward. Stepwise
-refers to the fact that the property must hold on every step. Later we will
-loosen this condition in the interest of extensionality -- the caller only
-cares about changes it can see after the return, not changes during execution.
-
-Suppose we insert an instruction after line 10, that stores 5 to the stack pointer - 1.
-Then we would be changing {\tt y}, which is sealed, in violation of stepwise integrity.
-
-\begin{figure}
-
-\begin{subfigure}{.35\textwidth}
-\begin{verbatim}
-int main() {
-0:     int x = 42, y = 0;
-1[C]:  y = f();
-2:     print (x + y);
-}
-
-int f() {
-10:     int z = *($SP-1);
-11[R]:  return (z+z);
-}
-\end{verbatim}
-\caption{A Confidentiality Violation}
-\label{fig:conf1-program}
-\end{subfigure}
-\begin{subfigure}{.5\textwidth}  
-\begin{center}
-\begin{tabular}{l l l l}
-{\tt 0} &
-\memoryaddrs{4em}
-\memory{3}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{a}
-\MemoryLabel{-10em}{0.75em}{b}
-\MemoryLabel{-7em}{0.75em}{c}
-\vspace{.5em}
-& &
-\\
-{\tt 1} &
-\memoryaddrs{12em}
-\memory{3}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\MemoryLabel{-7em}{0.75em}{c}
-\vspace{.5em} & &
-\\
-{\tt 10} &
-\memoryaddrs{12em}
-\memory{2}{\mainsealc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\MemoryLabel{-7em}{0.75em}{c} &
-\memoryaddrs{12em}
-\memory{2}{\mainsealc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{a'}
-\MemoryLabel{-10em}{0.75em}{b'}
-\MemoryLabel{-7em}{0.75em}{c'}
-\\
-{\tt 11} &
-\memoryaddrs{16em}
-\memory{2}{\mainsealc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\MemoryLabel{-7em}{0.75em}{0} &
-\memoryaddrs{16em}
-\memory{2}{\mainsealc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{a'}
-\MemoryLabel{-10em}{0.75em}{b'}
-\MemoryLabel{-7em}{0.75em}{b'}
-\\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\caption{Trace of Stack Domains}
-
-\label{fig:conf1-trace}
-\end{subfigure}
-\end{figure}
-
-Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
-of the context of its call (with the exception, of course, of its arguments).
-We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
-security. Noninterference describes knowledge: it states that for some component(s)
-that we do not know, we can substitute any values whatsoever without changing the results
-of our computation. If there exists some value for which we see different results, then
-in fact we can learn something about the components by observing those results, violating
-confidentiality.
-
-In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
-by reading {\tt main}'s variable {\tt y}, which was not passed to it. The trace
-in figure \ref{fig:conf1-trace} shows how noninterference works. First, let \(a\),
-\(b\), and \(c\) be the unknown initial values at their locations. On entry to {\tt f}
-we again have 42 and 0 in {\tt x} and {\tt y}, and {\tt z} contains \(c\). We consider a
-{\em variant state} that replaces the values of components that should be secret with
-arbitrary values \(a'\), \(b'\), and \(c'\). After a step, we've copied 0 and \(c'\),
-respectively, into {\tt z}. Since {\tt z} changed value, for confidentiality to hold,
-its value should be the same in all variant states. But when \(c' \not = 0\), the
-states changes differently, which violates confidentiality. It would also be a violation
-to print the secret value.
-
-This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
-as: for any state at the entry to a function, for any variant over the values in the stack,
-the primary state and its variant will step in lockstep until they return. Any components
-whose values update and any printed values will be identical between the primary and variant
-traces.
-
-\begin{figure}
-
-\begin{subfigure}{.35\textwidth}
-\begin{verbatim}
-int main() {
-0[C]:  int t = f();
-1[R]:  return;
-}
-
-int f() {
-10:     int x, y = 5, z = 6;
-11:     print y+z;
-12:     $SP-=2;
-13[C]:  x = g();
-14[R]:  return x;
-}
-
-int g() {
-20:     int z = *$SP;
-21[R]:  return (z+z);
-}
-\end{verbatim}
-\end{subfigure}
-\begin{subfigure}{.5\textwidth}  
-\begin{center}
-\begin{tabular}{l l l l}
-{\tt 13} &
-\memoryaddrs{12em}
-\memory{1}{\mainsealc}
-\memory{3}{\unsealc}
-~$\cdots$
-\MemoryLabel{-18em}{0.75em}{a}
-\MemoryLabel{-14em}{0.75em}{b}
-\MemoryLabel{-10em}{0.75em}{5}
-\MemoryLabel{-7em}{0.75em}{6}
-\vspace{.5em}
-& &
-\\
-{\tt 20} &
-\memoryaddrs{12em}
-\memory{1}{\mainsealc}
-\memory{1}{\fsealc}
-\memory{2}{\unsealc}
-~$\cdots$
-\MemoryLabel{-18em}{0.75em}{a}
-\MemoryLabel{-14em}{0.75em}{b}
-\MemoryLabel{-10em}{0.75em}{5}
-\MemoryLabel{-7em}{0.75em}{6}
-\vspace{.5em} &
-\memoryaddrs{12em}
-\memory{1}{\mainsealc}
-\memory{1}{\fsealc}
-\memory{2}{\unsealc}
-~$\cdots$
-\MemoryLabel{-18em}{0.75em}{a'}
-\MemoryLabel{-14em}{0.75em}{b'}
-\MemoryLabel{-10em}{0.75em}{c'}
-\MemoryLabel{-7em}{0.75em}{d'}
-\vspace{.5em}
-&
-\\
-{\tt 21} &
-\memoryaddrs{12em}
-\memory{1}{\mainsealc}
-\memory{1}{\fsealc}
-\memory{2}{\unsealc}
-~$\cdots$
-\MemoryLabel{-18em}{0.75em}{a}
-\MemoryLabel{-14em}{0.75em}{b}
-\MemoryLabel{-10em}{0.75em}{5}
-\MemoryLabel{-7em}{0.75em}{6}
-\vspace{.5em} &
-\memoryaddrs{12em}
-\memory{1}{\mainsealc}
-\memory{1}{\fsealc}
-\memory{2}{\unsealc}
-~$\cdots$
-\MemoryLabel{-18em}{0.75em}{a'}
-\MemoryLabel{-14em}{0.75em}{b'}
-\MemoryLabel{-10em}{0.75em}{c'}
-\MemoryLabel{-7em}{0.75em}{d'}
-\vspace{.5em}
-&
-\\
-
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-
-\caption{Another Confidentiality Violation}
-\label{fig:conf2}
-\end{figure}
-
-Two important details of confidentiality are reflected in figure \ref{fig:conf2}.

### 2021-04-02 tex/main.tex
Work on examples
Extended examples a bit more
More work on examples and intro

@@
 \input{examples}
+\input{example1}
 
@@
 \newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}
+\newcommand{\memoryaddrs}[1]
+  {
+    \MemoryLabel{#1}{2em}{\SP}
+  }
 
@@
 to a mechanism for protecting stack-based low-level programs to judge
-whether or not it actually ``guarantees stack safety.'' While our criteria are
-designed with PIPE micro-policies in mind, they are agnostic about many implementation
-details and should be applicable to other systems with minor adjustment.
-%
-Our key contribution is an abstract presentation of local state encapsulation
-using the technical framework of language-based security, together with a simple
-characterization of well-bracketed control flow.
-
-\begin{figure}[b]
-    \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-int main() {
-  int x = 42;
-  f();
-  print(x);
-}
-\end{verbatim}
-    \end{minipage}
-    \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-int main() {
-  int x = 1;
-  f();
-  if(x)
-    g();
-  else
-    h();
-}
-\end{verbatim}
-    \end{minipage}
-  \caption{Simple Examples}
-  \label{fig:examples1}
-\end{figure}
-
-\begin{figure}[b]
-    \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-int main(int arg) {
-  if(arg)
-    g();
-  else
-    h();
-  f();
-}
-\end{verbatim}
-    \end{minipage}
-    \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-int main() {
-  int x = 1;
-  f();
-  x = -x;
-  f();
-  print(x);
-}
-\end{verbatim}
-  \end{minipage}
-
-  \caption{More Examples}
-  \label{fig:examples2}
-\end{figure}
+whether or not it actually ``guarantees stack safety.'' We begin with a strong
+local state encapsulation property that closely models the behavior of eager
+micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
+is too strong to characterize stack safety; reasonable enforcement mechanisms
+exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
+the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
+to support this technique and argue that the new property, defined relative to an
+arbitrary notion of observation, is weak enough to characterize stack safety.
+
+While our criteria are designed with PIPE micro-policies in mind, they are
+agnostic about many implementation details and should be applicable to other
+systems with minor adjustment.
+%
+Our key contribution is an operational model of local state encapsulation
+using the technical framework of language-based security, with eager and lazy variants,
+together with a simple characterization of well-bracketed control flow.
+
+%\begin{figure}[b]
+%    \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%int main() {
+%  int x = 42;
+%  f();
+%  print(x);
+%}
+%\end{verbatim}
+%    \end{minipage}
+%    \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%int main() {
+%  int x = 1;
+%  f();
+%  if(x)
+%    g();
+%  else
+%    h();
+%}
+%\end{verbatim}
+%    \end{minipage}
+%  \caption{Simple Examples}
+%  \label{fig:examples1}
+%\end{figure}
+
+%\begin{figure}[b]
+%    \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%int main(int arg) {
+%  if(arg)
+%    g();
+%  else
+%    h();
+%  f();
+%}
+%\end{verbatim}
+%    \end{minipage}
+%    \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%int main() {
+%  int x = 1;
+%  f();
+%  x = -x;
+%  f();
+%  print(x);
+%}
+%\end{verbatim}
+%  \end{minipage}
+%
+%  \caption{More Examples}
+%  \label{fig:examples2}
+%\end{figure}
 
@@
 
-When does the programmer rely on stack safety? The most natural example is the first
-in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
-calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
-{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
-overwrite {\tt main}'s stack data. This expectation is for {\em integrity}: that a
-caller's data is protected from modification by its callee.
-This is not only useful for the programmer: consider the second
-example, where the call to {\tt h} should be unreachable and therefore subject to dead code
-elimination, if only the compiler can soundly reason that {\tt x} does not change.
+How does the programmer rely on stack safety? The stack makes implicit promises about
+the behavior of function calls. First, that when a call is finished, the caller can continue
+from where it left off with its local state unchanged. And second, that the call is
+self-contained, its behavior determined only by a predictable interface -- at minimum
+parameters and global variables, but never local state.
+
+%The most natural example is the first
+%in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
+%calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
+%{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
+%overwrite {\tt main}'s stack data.
+
+Implicit in both expectations are a predictable control flow, termed {\em well-bracketed
+control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls jump to a consistent entry
+point and return to the call site. Then the data portion of the first expectation is for
+{\em integrity}: that a caller's local data is protected from modification by its callee.
+This is vital for local reasoning about the behavior of the caller. The security implications
+are significant: a caller that branches on a local variable could be manipulated by its
+callee into unexpected behavior without integrity.
 
 Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
-from leakage by its callee. Consider again the first example in figure \ref{fig:examples1},
-in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
-programmer had initialized {\tt x} after calling {\tt f} rather than before; such a change
-should not alter the behavior of {\tt main}, but if {\tt f} can directly read {\tt x}, it
-may. But there are more subtle examples, such as the first in figure \ref{fig:examples2},
-where {\tt main} may call {\tt g} or {\tt h} are called prior to calling {\tt f}, depending on
-its argument. Here, if {\tt g} or {\tt h} leaves behind identifying information in unused stack
-space, {\tt f} could still deduce information about the argument and behave differently, violating
-the abstraction of its call.
-
+from leakage by its callee. As a reasoning principle, confidentiality tells us that to
+predict the behavior of a call we need only consider a subset of the machine state.
+Rearranging local data will have no impact. And in security terms, we can guarantee
+the secrecy of local data that we haven't passed even when calling a badly behaved function.
 Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
-{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
-callers from their callees. Consider the last example in figure \ref{fig:examples2},
-a variant of the famous ``awkward example,'' in which {\tt f} is called twice and might store
-data outside the stack between calls. It should be the case that {\tt main} prints -1, but if
-{\tt f}'s first activation stores its return address somewhere accessible, the second
-activation can return to before {\tt x} is negated, causing {\tt x} to be set to
-1 again before a third activation that would return normally. Then {\tt main} would print 1.
-This is rather unexpected behavior.
-
-Preventing this form of control-flow attack is the intuition behind
-\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
-the call graph itself. A callee should pass control back to the caller only by
-transferring control to the instruction following the call; otherwise it could
-subvert LSE by appearing to return -- thus restoring access to the caller's data --
-while causing the execution of arbitrary instructions on that data.
+{\em local state encapsulation} (LSE) property.
+
+It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
+could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
+accessing the caller's state, but the execution would remain under the callee's control.
 
@@
 callee. Then we expand it to allow components to be passed, granting a callee one time
-access to the caller's data, or shared, allowing components to be outright
-exempted from the stack safety properties. Shared components could model objects whose
-addresses are passed and whose security must be modeled by a separate property, such as
-memory safety enforced by capabilities. Finally, we extend the model to a coroutine
-system featuring finitely-bounded stacks.
+access to the caller's data. We do not model more sophisticated sharing, because to do
+so realistically would require a separate protection scheme, such as a capability
+model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
+a separate policy outside the scope of stack safety. (The combination of such policies is
+a strength of tag-based approaches.)
+
+Finally, we extend the model to a coroutine system featuring finitely-bounded stacks. Though
+simple, this extension is a step toward concurrency and demonstrates that the model is flexible
+enough to handle more sophisticated styles of control-flow.
 
@@
 \citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
-This policy testably enforces our {\em eager} properties, in which each
-callee, at the point of its return, can be seen to respect the integrity
-and confidentiality of its caller. Additionally we have an
-{\em ultra-eager} integrity property which guarantees that integrity
-is respected at each step. The advantage of this property is that it ``fails
-faster'' and is therefore more suitable for efficient testing. We prove (in
-Coq) that the ultra-eager property implies our original definition of integrity.
-
-The downside of the Depth Isolation policy is that it is rather
-expensive to implement. In the same work,
-\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
+This policy testably enforces our {\em eager} properties, in which integrity is respected
+at every step of a program, and each call respects integrity at each of its internal steps.
+These properties are very strong, and our policies enforce them; they also have the advantage
+that they ``fail faster'' and are therefore suitable for efficient testing.
+
+The downside of the Depth Isolation policy is that it is rather expensive to implement.
+In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
 that is more efficient by virtue of being lazy in its enforcement.
@@
   property in the same way...}\fi
-The lazy policy does not enforce our original integrity property, and to our
+The lazy policy does not enforce our eager integrity property, and to our
 knowledge there is no formal statement of the properties that it ought to enforce.
-We propose a {\em lazy} integrity property, which capture the desired behavior of a
+We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
 policy that detects integrity violations between when they occur and when they make
@@
 writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
-``Lazy Tagging and Clearing'' policy does not quite enforce this property, as shown in
+``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
 testing, but formalizing the ideal property allows us to see how it falls short.
 We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
-at some performance cost.
+at some performance cost. We prove (in Coq) that the eager property implies the lazy one.
+We argue that any system purporting to enforce stack safety ought to obey lazy stack safety.
 
@@
   % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A novel {\em stack-safety property} abstractly capturing the
-  intuitive concepts of strict local state encapsulation and well-bracketed
-  control flow~(\cref{sec:lse-and-wbcf}).
+  A novel {\em stack-safety property} capturing the intuitive concepts of strictly enforced
+  local state encapsulation and well-bracketed control flow~(\cref{sec:lse-and-wbcf}).
   % These properties rule out common examples of stack memory abuse.
@@
   property can be enforced by the Depth Isolation policy of
-  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware tagging~(\cref{sec:conservative}).
-\item
-  An {\em ultra-eager} variant of the stack-integrity property that is also
-  enforced by the Depth Isolation policy but is more efficient to test, plus a proof
-  (in Coq) that this variant is stronger than the original~(\cref{sec:lockstep}).
+  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware
+  tagging~(\cref{sec:conservative}).
 \item
@@
 \item
-  Standardized trace-based formulations of eager confidentiality and of ultra-eager,
-  eager, and lazy integrity, parameterizeable by arbitrary control-flow models and assignments
-  of machine components to {\em LSE domains}.
+  Standardized trace-based formulations of both eager and lazy confidentiality and integrity,
+  parameterizeable by arbitrary control-flow models and assignments of machine components
+  to {\em LSE domains}.
 \item
-  An LSE-domain model tracking the passing of function arguments on the stack and the sharing
-  of arbitrary stack data.
+  An LSE-domain model enhanced with passing of function arguments on the stack, with which our
+  standardized trace properties extend stack safety with local sharing.
 \item
-  An extension of stack safety to a simple coroutine model with corresponding LSE
-  properties and a well-structured control flow property that subsumes WBCF for each
+  An LSE-domain model extending the stack to a simple coroutine model, with corresponding LSE
+  properties and a {\em well-structured control flow} property that subsumes WBCF for each
   stack as well as guaranteeing legitimate coroutine crossings.
@@
 
-\paragraph{Well-formedness and Bad Annotations}
+\paragraph*{Well-formedness and Bad Annotations}
 
@@
 
-\paragraph{System Assumptions}
+\paragraph*{System Assumptions}
 
@@
 
-\paragraph{Policies}
+\paragraph*{Policies}
 
@@
 
-\section{Examples and Intuition}
+\section{Stack Safety, Intuitively}
 \label{sec:running-example}
-
-\input{example1}
-
-\input{example2}
+\colorlet{lgray}{gray!40}
+
+\newcommand{\mainsealc}{cyan}
+\newcommand{\fsealc}{green}
+\newcommand{\unsealc}{lgray}
+\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
+\newcommand{\fulloutc}{white}
+
+\begin{figure}
+
+\begin{subfigure}{.4\textwidth}
+\begin{verbatim}
+int main() {
+0:     int x = 42, y = 0;
+1[C]:  y = f(10);
+2:     print (x + y);
+}
+
+int f(int z) {
+10:     print z;
+11[R]:  return (z+z);
+}
+\end{verbatim}
+
+\caption{A Simple Call}
+\label{fig:simple-program}
+\end{subfigure}
+\begin{subfigure}{.4\textwidth}  
+\begin{center}
+\begin{tabular}{l l}
+{\tt 0} &
+\memoryaddrs{4em}
+\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+~$\cdots$
+\\
+{\tt 1} &
+\memoryaddrs{12em}
+\memory{3}{\unsealc}
+~$\cdots$
+\MemoryLabel{-14em}{0.75em}{42}
+\MemoryLabel{-10em}{0.75em}{0}
+\vspace{.5em}
+\\
+{\tt 10} &
+\memoryaddrs{17em}
+\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
+\memory{2}{\unsealc}
+~$\cdots$
+\MemoryLabel{-10em}{0.75em}{2}
+\\
+{\tt 2} &
+\memoryaddrs{12em}
+\memory{3}{\unsealc}
+~$\cdots$
+\MemoryLabel{-14em}{0.75em}{42}
+\MemoryLabel{-10em}{0.75em}{0}
+\MemoryLabel{-7em}{0.75em}{2}
+\\
+\end{tabular}
+\end{center}
+
+\vspace{\abovedisplayskip}
+
+\caption{Trace of Stack Domains}
+
+\label{fig:simple-trace}
+\end{subfigure}
+\end{figure}
+
+This section will elucidate the intuition behind the properties through a series
+of examples. Existing stack safety enforcement work mostly refers directly to examples
+of bad behavior that it prevents; we will insert our model between the enforcement and
+the examples, arguing that:
+
+\begin{itemize}
+\item Examples that appear to violate stack safety, violate at least one of our properties
+\item We do not find desireable behaviors that appear stack-safe yet violate our properties
+\item Our properties are testably implemented by stack safety micro-policies from the literature,
+  and the micro-policies are not significantly more conservative (i.e., policy failstops tend to
+  correspond to imminent property violations)
+\end{itemize}
+
+\Cref{fig:simple-program} shows a simple example, written in C-like pseudo-code.
+It consists of a {\tt main} function that performs a call to a function {\tt f},
+which prints its argument, performs arithmetic, and returns a result.
+\Cref{fig:simple-trace} shows the stack as it is divided into domains throughout
+the program execution. We use high-level code for compactness, but reason about
+a standard compilation of each function. Line numbers correspond to the code
+addresses just before the code that implements the line, and annotations
+{\tt C} and {\tt R} indicate that an instruction is annotated as a call or
+return, respectively.
+
+We also assume that the machine is word- (not byte-) addressed and that the
+machine communicates with the outside world through a single memory-mapped
+output port whose address is elided (so a {\tt print} statement will be
+compiled to a store to this location), and that the stack grows upwards.
+
+In the initial memory layout the entire stack is currently {\em unsealed},
+meaning that it is eligible to be written to. At the call from {\tt main} to
+{\tt f}, {\tt main}'s locals (as designated by the stack pointer) are
+{\em sealed} and labeled with the depth of their owner, in this case 0.
+Sealing is the fundamental integrity contract between the caller and callee:
+{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
+This gives us our simple, intuitive statement of {\em stepwise integrity}:
+if a component is sealed before a step, it will be unchanged afterward. Stepwise
+refers to the fact that the property must hold on every step. Later we will
+loosen this condition in the interest of extensionality -- the caller only
+cares about changes it can see after the return, not changes during execution.
+
+Suppose we insert an instruction after line 10, that stores 5 to the stack pointer - 1.
+Then we would be changing {\tt y}, which is sealed, in violation of stepwise integrity.
+
+\begin{figure}
+
+\begin{subfigure}{.33\textwidth}
+\begin{verbatim}
+int main() {
+0:     int x = 42, y = 0;
+1[C]:  y = f();
+2:     print (x + y);
+}
+
+int f() {
+10:     int z = *($SP-1);
+11[R]:  return (z+z);
+}
+\end{verbatim}
+\label{fig:conf1-program}
+\end{subfigure}
+\begin{subfigure}{.59\textwidth}
+\centering
+\begin{tabular}{l l l}
+  {\tt 0} &
+  \multicolumn{2}{c}{
+    \memoryaddrs{4em}
+    \memory{3}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-14em}{0.75em}{a}
+    \MemoryLabel{-10em}{0.75em}{b}
+    \MemoryLabel{-7em}{0.75em}{c}
+    \vspace{.5em}
+  } \\
+  {\tt 1} &
+  \multicolumn{2}{c}{
+    \memoryaddrs{12em}
+    \memory{3}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-14em}{0.75em}{42}
+    \MemoryLabel{-10em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{c}
+    \vspace{.5em}
+  } \\
+  {\tt 10} &
+  \memoryaddrs{20em}
+  \memory{2}{\mainsealc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{42}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf d}
+ &
+  \memoryaddrs{20em}
+  \memory{2}{\mainsealc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{a'}
+  \MemoryLabel{-14em}{0.75em}{b'}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf d'}
+  \\
+  {\tt 11} &
+  \memoryaddrs{20em}
+  \memory{2}{\mainsealc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{42}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf 0}
+  &
+  \memoryaddrs{20em}
+  \memory{2}{\mainsealc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{a'}
+  \MemoryLabel{-14em}{0.75em}{b'}
+  \MemoryLabel{-10em}{0.75em}{2}
+  \MemoryLabel{-7em}{0.75em}{\bf b'}
+  \\
+\end{tabular}
+
+\vspace{\abovedisplayskip}
+
+\label{fig:conf1-trace}
+\end{subfigure}
+\caption{A Confidentiality Violation}
+\end{figure}
+
+Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
+of the context of its call (with the exception, of course, of its arguments).
+We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
+security. Noninterference describes knowledge: it states that for some component(s)
+that we do not know, we can substitute any values whatsoever without changing the results
+of our computation. If there exists some value for which we see different results, then
+in fact we can learn something about the components by observing those results, violating
+confidentiality.
+
+In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
+by reading {\tt main}'s variable {\tt y}, which was not passed to it. The trace
+in figure \ref{fig:conf1-trace} shows how noninterference works. First, let \(a\),
+\(b\), and \(c\) be the unknown initial values at their locations. On entry to {\tt f}
+we again have 42 and 0 in {\tt x} and {\tt y}, and {\tt z} contains \(c\). We consider a
+{\em variant state} that replaces the values of components that should be secret with
+arbitrary values \(a'\), \(b'\), and \(c'\). After a step, we've copied 0 and \(c'\),
+respectively, into {\tt z}. Since {\tt z} changed value, for confidentiality to hold,
+its value should be the same in all variant states. But when \(c' \not = 0\), the
+states changes differently, which violates confidentiality. It would also be a violation
+to print the secret value.
+
+This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
+as: for any state at the entry to a function, for any variant over the values in the stack,
+the primary state and its variant will step in lockstep until they return. Any components
+whose values update and any printed values will be identical between the primary and variant
+traces.
+
+\begin{figure}
+
+\begin{subfigure}{.25\textwidth}
+\begin{verbatim}
+int main() {
+0[C]:  int t = f();
+1[R]:  return;
+}
+\end{verbatim}
+\end{subfigure}
+\begin{subfigure}{.37\textwidth}
+\begin{verbatim}
+int f() {
+10:     int x, y = 5, z = 6;
+11:     print y+z;
+12:     $SP-=2;
+13[C]:  x = g();
+14[R]:  return x;
+}
+\end{verbatim}
+\end{subfigure}
+\begin{subfigure}{.32\textwidth}
+\begin{verbatim}
+int g() {
+20:     int z;
+21[R]:  return (z+z);
+}
+\end{verbatim}
+\end{subfigure}
+
+\begin{subfigure}{.8\textwidth}  
+\begin{center}
+\begin{tabular}{l l l}
+  {\tt 13} &
+  \multicolumn{2}{c}{
+    \memoryaddrs{16em}
+    \memory{1}{\mainsealc}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-23em}{0.75em}{a}
+    \MemoryLabel{-19em}{0.75em}{b}
+    \MemoryLabel{-15em}{0.75em}{1}
+    \MemoryLabel{-11em}{0.75em}{5}
+    \MemoryLabel{-8em}{0.75em}{6}
+    \vspace{.5em}
+  }
+  &
+\\
+  {\tt 20} &
+  \memoryaddrs{25em}
+  \memory{1}{\mainsealc}
+  \memory{2}{\fsealc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-23em}{0.75em}{a}
+  \MemoryLabel{-19em}{0.75em}{b}
+  \MemoryLabel{-15em}{0.75em}{1}
+  \MemoryLabel{-11em}{0.75em}{14}
+  \MemoryLabel{-8em}{0.75em}{\bf 6}
+  \vspace{.5em}
+  &
+  \memoryaddrs{25em}
+  \memory{1}{\mainsealc}
+  \memory{2}{\fsealc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-23em}{0.75em}{a'}
+  \MemoryLabel{-19em}{0.75em}{b'}
+  \MemoryLabel{-15em}{0.75em}{c'}
+  \MemoryLabel{-11em}{0.75em}{14}
+  \MemoryLabel{-8em}{0.75em}{\bf d'}
+  \vspace{.5em}
+\\
+\end{tabular}
+\end{center}
+
+\vspace{\abovedisplayskip}
+
+\end{subfigure}
+
+\caption{Another Confidentiality Violation}
+\label{fig:conf2}
+\end{figure}
+
+Two important details of confidentiality are reflected in figure \ref{fig:conf2}.
+First of all, it has multiple calls, and while we focus on the stack at the call
+from line 13 to line 20, it should be clear that we also need to consider a similar
+variant at the call from 0 to 10. Confidentiality is in this sense recursive: each
+nested call must obey confidentiality in its own right.
+
+More importantly, careful readers will have noticed that the informal definition
+of stack confidentiality refers to a variation of the entire stack, not just the
+sealed portion. In fact, we support temporal stack safety, in which values left behind
+by previous activations are also inaccessible. In this case, we connect temporal safety
+firmly to the caller-callee attacker model -- {\tt f} stored some data in the variables
+{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.
+After the stack pointer is written over {\tt y}, the value of {\tt z} is still
+left behind. Then {\tt g} can read it just by allocating and accessing an uninitialized
+variable!
+
+\paragraph*{Control Flow and Awkward Examples}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.4\textwidth}
+\begin{verbatim}
+int stash = 0;
+
+int main() {
+0:     int x = 1;
+1[C]:  f();
+2:     x = -x;
+3[C]:  f();
+4[R]:  return x;
+}
+
+void f() {
+10:     if(!stash) {
+        // store return pointer
+11:       stash = $SP-1;
+12:     } else {
+13:       $SP-1 = stash;
+14:       stash = 0;
+15[R]:    return;
+        }
+}
+\end{verbatim}
+  \end{subfigure}
+\label{fig:WBCF1}
+  \begin{subfigure}[t]{.4\textwidth}
+\begin{verbatim}
+int stash = 0;
+
+int main() {
+0:     int x = 1;
+1[C]:  f();
+2:     x = -x;
+3[R]:  return x;
+}
+
+void f() {
+10:     if(!stash) {
+        // store return pointer
+11:       stash = $SP-1;
+12[C]     main();
+12:     } else {
+13:       $SP-1 = stash;
+14:       stash = 0;
+15[R]:    return;
+        }
+}
+\end{verbatim}
+\label{fig:WBCF2}
+  \end{subfigure}
+  \caption{Control Flow Attacks}
+  \label{fig:WBCF}
+\end{figure}
+
+Example \ref{fig:WBCF1} shows how control flow attacks can compromise integrity in principle
+even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
+global variable and retrieves it the second time it is called, returning to the first call
+point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
+line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.
+
+To prevent this, we define Well-bracketed Control Flow. Throughout the above examples, we
+have implicitly tracked the depth of the current call in order to seal memory at the correct
+depth. We also have a clear sense of where the entry points to functions are -- the top line,
+in our examples, and another instruction annotation in practice. Well-bracketed Control Flow
+holds if:
+
+\begin{itemize}
+\item All calls jump to entry points ({\em entry integrity})
+\item The depth counter decreases by exactly one on a return ({\em non-skipping})
+\item From a call at depth \(d\), the next state to have depth \(d\) matches the stack pointer,
+  and the program counter is at the instruction following the jal ({\em return integrity})
+\end{itemize}
+
+Non-skipping comes into play for attacks such as seen in \ref{WBCF2}, where the calls are
+nested rather than sequential. \sna{The example isn't quite right; need to look at the
+  Towards Complete... paper again.}
+
+\sna{Put sharing example here, or after the formalism?}
 
@@
 
-    \begin{figure}
-      \integritylazyexample
-      \caption{Rolling back from Figure \ref{fig:intex}.}
+%    \begin{figure}
+%      \integritylazyexample
+%      \caption{Rolling back from Figure \ref{fig:intex}.}
         %\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
-      \label{fig:intlex}
-    \end{figure}
+%      \label{fig:intlex}
+%    \end{figure}
 
@@
 
-      \begin{figure}
-        \confidentialitylazyexample
-        \caption{Continuing from \cref{fig:confex}.}
-        \label{fig:conflex}
-      \end{figure}
+%      \begin{figure}
+%        \confidentialitylazyexample
+%        \caption{Continuing from \cref{fig:confex}.}
+%        \label{fig:conflex}
+%      \end{figure}
 
@@
 
-\section{Extension to Software-based Enforcement}
+%\section{Extension to Software-based Enforcement}
 

### 2021-04-02 tex/main.tex
Finished up draft of examples section

@@
 
+We alternatively could imagine a stepwise form of confidentiality in which we explicitly
+track which addresses may be read (i.e., which have been initialized), and at each step
+guarantee that the step alters state identically to a step from a variant state. Then we
+wouldn't need to divide up the execution by calls. But this is further from the normal
+framing of noninterference, requires explicitly tracking initializations, and would behave
+no differently relative to our policies.
+
 \begin{figure}
@@
 0[C]:  int t = f();
-1[R]:  return;
+1[R]:  return t;
 }
@@
   }
-  &
 \\
@@
   \centering
-  \begin{subfigure}[t]{.4\textwidth}
+  \begin{minipage}{.4\textwidth}
 \begin{verbatim}
@@
 }
-
+\end{verbatim}
+  \end{minipage}
+  \begin{minipage}{.4\textwidth}
+\begin{verbatim}
 void f() {
@@
 \end{verbatim}
-  \end{subfigure}
+  \end{minipage}
 \label{fig:WBCF1}
-  \begin{subfigure}[t]{.4\textwidth}
-\begin{verbatim}
-int stash = 0;
-
-int main() {
-0:     int x = 1;
-1[C]:  f();
-2:     x = -x;
-3[R]:  return x;
-}
-
-void f() {
-10:     if(!stash) {
-        // store return pointer
-11:       stash = $SP-1;
-12[C]     main();
-12:     } else {
-13:       $SP-1 = stash;
-14:       stash = 0;
-15[R]:    return;
-        }
-}
-\end{verbatim}
-\label{fig:WBCF2}
-  \end{subfigure}
+%  \begin{subfigure}[t]{.4\textwidth}
+%\begin{verbatim}
+%int counter = 0;
+%int stash = 0;
+%
+%int main() {
+%0:     int x = 1;
+%1[C]:  f();
+%2:     x = -x;
+%3[C]:  f();
+%4[R]:  return x;
+%}
+%
+%void f() {
+%10:     switch(counter++) {
+%11:       case 0:
+%12[R]:      return;
+%13:       case 1:
+%14:         stash = $SP-1;
+%15[C]:      main();
+%16:         return;
+%17:       default:
+%18:         $SP-1 = stash;
+%19[R]:      return;
+%        }
+%}
+%\end{verbatim}
+%\label{fig:WBCF2}
+%  \end{subfigure}
   \caption{Control Flow Attacks}
@@
 
-Example \ref{fig:WBCF1} shows how control flow attacks can compromise integrity in principle
+Example \ref{fig:WBCF} shows how control flow attacks can compromise integrity in principle
 even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
@@
 
-Non-skipping comes into play for attacks such as seen in \ref{WBCF2}, where the calls are
-nested rather than sequential. \sna{The example isn't quite right; need to look at the
-  Towards Complete... paper again.}
-
-\sna{Put sharing example here, or after the formalism?}
+Non-skipping may seem strange -- surely the depth counter ought to decrease by one at a time
+as part of the property definition, and the problem comes when the return is to a shallower
+call's return pointer, violating return integrity? Yet some weaker enforcement mechanisms,
+such as those in which returns are implemented as invocation of a Cheri object capability,
+do allow skipping when adversarial code acts unreasonably by leaking its own return capability,
+reasoning that this is acceptable as long as only the unreasonable function is skipped
+(e.g. \citep{Skorstengaard+19b}). A variant of our model supports this weaker form of WBCF,
+but with tag-based policies we can enforce the non-skipping version entirely.
+\sna{Cite future work for details when the relevant section is finished.}
+
+\paragraph*{Pass by Reference and Stack Arguments}
+
+\newcommand{\mainpassc}{magenta}
+\newcommand{\sharec}{lgray}
+
+\begin{figure}[b]
+  \centering
+  \begin{subfigure}{.45\textwidth}
+\begin{verbatim}
+int stash;
+
+int main() {
+0:             int x = 42, y = 0;
+1[P(x)],2[C]:  f(&x);
+3:             x = 0;
+4[P(y)],5[C]:  f(&y);
+6[R]:          return 0;
+}
+
+void f(int* a) {
+10:  if(!stash)
+11:    stash = a;
+12:  *stash = 5;
+13:  return;
+}
+\end{verbatim}
+\end{subfigure}
+  \begin{subfigure}{.5\textwidth}
+    \begin{tabular}{l l}
+      {\tt 2} &
+      \memoryaddrs{12em}
+      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
+      \memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+      ~$\cdots$ \\
+      {\tt 10} &
+      \memoryaddrs{17em}
+      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{2}{\unsealc}
+      ~$\cdots$ \\
+      {\tt 3} &
+      \memoryaddrs{12em}
+      \memory{3}{\unsealc}
+      ~$\cdots$ \\
+      {\tt 5} &
+      \memoryaddrs{12em}
+      \memory{1}{\unsealc}
+      \memory{1}{\mainpassc}
+      \memory{1}{\unsealc}
+      ~$\cdots$ \\
+      {\tt 10} &
+      \memoryaddrs{17em}
+      \memory{1}{\mainsealc}
+      \memory{1}{\mainpassc}
+      \memory{2}{\unsealc}
+      ~$\cdots$ \\
+\end{tabular}
+\end{subfigure}
+
+\label{fig:passing}
+\end{figure}
+
+Up til now we have assumed that function arguments and returns are passed solely in registers.
+This is an oversimplication, which we will now correct. There are important stack safety concerns
+when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
+{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
+{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
+\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
+magenta. Then at the call, the caller's frame that is not passed is sealed, as usual. Passed
+memory also tracks which depth owns it, and it may only be accessed by the immediate callee
+(at its depth plus one).
+
+But because {\tt f} stashes the original pointer to {\tt x}, it will overwrite {\tt x} while
+{\tt x} is sealed, an integrity violation. This fits with the intuition of passing by reference,
+that the passed variable might be read or written during the call, not that it is permanently
+open to be read or written by all subsequent calls. This also applies to passing large objects
+by value in the stack: the caller allocates them, and the relevant instructions are annotated to
+indicate passing.
+
+We do not model cases where permanent sharing is desireable, because they are uninteresting
+and hard to motivate: we would expect to treat such sharing as completely unprotected, and
+perhaps attach a separate memory safety policy to add some protection to such variables.
 

### 2021-04-06 tex/main.tex
Couple little things (still getting started)

@@
 \begin{abstract}
-  What is stack safety? Numerous mechanisms have been proposed to guard against
+\bcp{How about ``Stack Safety {\em is} a Security Property''?}
+  What is stack safety? The phrase has been used to describe the protections
+  offered by a variety of mechanisms for guarding against
   attacks that corrupt the call stacks of running programs, including micro-policies

### 2021-04-06 tex/main.tex
a few comments and rewords on 1-2

@@
 
-But what {\em is} this stack safety of which they speak? Precise formal
+But what {\em is} the stack safety that these techniques enforce? Precise formal
 definitions are mostly lacking: rather, the notion is commonly defined in a
@@
 there is still room to craft an operational characterization
-of stack safety. In other words, we seek formal criteria that can be applied
+of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
+In other words, we seek formal criteria that can be applied
 to a mechanism for protecting stack-based low-level programs to judge
@@
 arbitrary notion of observation, is weak enough to characterize stack safety.
+\apt{Reads oddly. ``string enough'' is what we want, surely? (While still weak
+  enough to be enforced by the lazy policies.)}
 
@@
 Our key contribution is an operational model of local state encapsulation
-using the technical framework of language-based security, with eager and lazy variants,
+using the technical framework of language-based security\apt{this idea deserves more prominence}, with eager and lazy variants,
 together with a simple characterization of well-bracketed control flow.
@@
 self-contained, its behavior determined only by a predictable interface -- at minimum
-parameters and global variables, but never local state.
+parameters and global variables, but never local state.\apt{This last sentence is not so clear.}
 
@@
 
-Implicit in both expectations are a predictable control flow, termed {\em well-bracketed
-control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls jump to a consistent entry
+Both expectations imply predictable control flow, termed {\em well-bracketed
+control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls jump to a consistent \apt{??} entry
 point and return to the call site. Then the data portion of the first expectation is for
@@
 This is vital for local reasoning about the behavior of the caller. The security implications
-are significant: a caller that branches on a local variable could be manipulated by its
-callee into unexpected behavior without integrity.
+are significant: without integrity, a caller that branches on a local variable could be manipulated by its
+callee into unexpected behavior.
 
@@
 predict the behavior of a call we need only consider a subset of the machine state.
-Rearranging local data will have no impact. And in security terms, we can guarantee
+Rearranging local data will have no impact.\apt{??} And in security terms, we can guarantee
 the secrecy of local data that we haven't passed even when calling a badly behaved function.
@@
 %
-Even more fundamentally, we need to know when a call or a return is taking
+Even more fundamentally, \apt{``For this model to make sense,'' (??)} we need to know when a call or a return is taking
 place. We assume that calls and returns are annotated in the code, so that
@@
 First we introduce this model for a single simple stack with no sharing between callers and
-callee. Then we expand it to allow components to be passed, granting a callee one time
+callee. Then we expand it to allow components\apt{??} to be passed, granting a callee one time
 access to the caller's data. We do not model more sophisticated sharing, because to do
@@
 Finally, we extend the model to a coroutine system featuring finitely-bounded stacks. Though
-simple, this extension is a step toward concurrency and demonstrates that the model is flexible
+simple\apt{simplistic}, this extension is a step toward concurrency and demonstrates that the model is flexible
 enough to handle more sophisticated styles of control-flow.
@@
 These properties are very strong, and our policies enforce them; they also have the advantage
-that they ``fail faster'' and are therefore suitable for efficient testing.
+that they ``fail faster'' \apt{than what?} and are therefore suitable for efficient testing.
 
@@
 policy that detects integrity violations between when they occur and when they make
-a visible impact. Intuitively this can be phrased, ``No visible reads after dangerous
+a visible impact. Intuitively this can be phrased as ``no visible reads after dangerous
 writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
@@
 We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
-at some performance cost. We prove (in Coq) that the eager property implies the lazy one.
-We argue that any system purporting to enforce stack safety ought to obey lazy stack safety.
+at some performance cost. We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
+  and is it so interesting anyhow?}
+We argue that any system purporting to enforce stack safety ought to obey \apt{``at least'' ?} xslazy stack safety.
 
@@
 \item
-  An LSE-domain model extending the stack to a simple coroutine model, with corresponding LSE
+  An LSE-domain model that applies to simple coroutining, with corresponding LSE
   properties and a {\em well-structured control flow} property that subsumes WBCF for each
-  stack as well as guaranteeing legitimate coroutine crossings.
+  stack as well as guaranteeing legitimate coroutine control transfers.
 %  This policy does not enforce our strict
@@
 function. No accidental calls or returns, and attackers control their own returns
-but cannot change the annotations on the caller's code.
+but cannot change the annotations on the caller's code.\apt{cryptic}
 
@@
 Since we rely so heavily on code annotations, we must ask: what if the annotations
-are wrong? What if a call or return occurs that is not annotated, or an annotation
-fails to correspond to an actual return? In the former case, the answer is simple --
-there has been no call or return, only instruction resembling such an operation, and
+are wrong? First, suppose a call or return occurs that is not annotated.
+In this case, the answer is simple:
+there has been no call or return\apt{from whose perspective?}, only an instruction resembling such an operation, and
 therefore no associated change in permissions. This is particularly interesting in the
@@
 
-The latter case depends on the enforcement mechanism and the nature of the false
+Alternatively, suppose an annotated instruction is not really an actual \apt{``call or'' ?} return.
+What happens in this case depends on the enforcement mechanism and the nature of the false
 annotation. A given enforcement mechanism might place a static well-formedness criterion
-on the code it applies to. For instance, our micro-policies expect that call and return
+on the code it applies to. For instance, our\apt{or \citeauthor{DBLP:conf/sp/RoesslerD18}'s ? }
+micro-policies expect that call and return
 annotations occur on jump instructions. On the other hand, not all jumps are proper
@@
 in permissions or, if doing so would compromise security, fail-stop rather than
-execute the annotated instruction.
+execute the annotated instruction.\apt{hard to follow}
 
 \paragraph*{System Assumptions}
+
+\apt{I'm unclear about how many of these assumptions are relevant to the properties
+  as opposed to the (micro)policies.}
 
@@
 purposes of explanation we will describe a standard layout. Memory consists of an
-instruction region and a stack that grows upwards. (For simplicity,
+instruction region and a stack that grows upwards.\apt{if we want to be ``standard''
+  we really should have the stack grown downwards.}
+(For simplicity,
 we do not consider the heap. Integrating a memory-safety property for the
@@
 \item Our properties are testably implemented by stack safety micro-policies from the literature,
-  and the micro-policies are not significantly more conservative (i.e., policy failstops tend to
+  and the micro-policies are not significantly more conservative (i.e., policy failstops usually 
   correspond to imminent property violations)

### 2021-04-06 tex/main.tex
comments on 3

@@
 the program execution. We use high-level code for compactness, but reason about
-a standard compilation of each function. Line numbers correspond to the code
+a standard compilation of each function. \apt{Note assumption that arguments and return values are in registers.}
+Line numbers correspond to the code
 addresses just before the code that implements the line, and annotations
@@
 return, respectively.
+\apt{In trace, clarify difference between values (like 42,0) and line numbers (like 2).}
 
@@
 meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals (as designated by the stack pointer) are
+{\tt f}, {\tt main}'s locals (as designated \apt{``delimited''? in any case, how?} by the stack pointer) are
 {\em sealed} and labeled with the depth of their owner, in this case 0.
@@
 
-Suppose we insert an instruction after line 10, that stores 5 to the stack pointer - 1.
+Suppose we insert an instruction after line 10 that stores 5 to the stack pointer - 1.\apt{Actually, -2.
+  But this also raises the question of if/how return addresses are protected.}
 Then we would be changing {\tt y}, which is sealed, in violation of stepwise integrity.
@@
 Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
-of the context of its call (with the exception, of course, of its arguments).
+of the context of its call (except, of course, for its arguments).
 We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
-security. Noninterference describes knowledge: it states that for some component(s)
+security. Noninterference describes knowledge: it states that for some component(s)\apt{??}
 that we do not know, we can substitute any values whatsoever without changing the results
@@
 by reading {\tt main}'s variable {\tt y}, which was not passed to it. The trace
-in figure \ref{fig:conf1-trace} shows how noninterference works. First, let \(a\),
+in figure \ref{fig:conf1-trace} shows how noninterference works. \apt{Is this right? what is d?}
+First, let \(a\),
 \(b\), and \(c\) be the unknown initial values at their locations. On entry to {\tt f}
@@
 its value should be the same in all variant states. But when \(c' \not = 0\), the
-states changes differently, which violates confidentiality. It would also be a violation
-to print the secret value.
+states change differently, which violates confidentiality. It would also be a violation
+to print the secret value.\apt{all this is not too clear.}
 
@@
 framing of noninterference, requires explicitly tracking initializations, and would behave
-no differently relative to our policies.
+no differently relative to our policies.\apt{not very clear.}
 
@@
 firmly to the caller-callee attacker model -- {\tt f} stored some data in the variables
-{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.
+{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.\apt{Where is the storage for x ?.
+And need to explain what the colors mean.}
 After the stack pointer is written over {\tt y}, the value of {\tt z} is still
 left behind. Then {\tt g} can read it just by allocating and accessing an uninitialized
-variable!
+variable! \apt{Need to be more explicit about what is getting varied, when, and why.}
 
@@
 reasoning that this is acceptable as long as only the unreasonable function is skipped
-(e.g. \citep{Skorstengaard+19b}). A variant of our model supports this weaker form of WBCF,
-but with tag-based policies we can enforce the non-skipping version entirely.
+(e.g. \citep{Skorstengaard+19b}).\apt{???} A variant of our model supports this weaker form of WBCF,
+but with tag-based policies we can enforce the non-skipping version entirely.\apt{??}
 \sna{Cite future work for details when the relevant section is finished.}
@@
 \(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
-magenta. Then at the call, the caller's frame that is not passed is sealed, as usual. Passed
+magenta. \apt{Why are there two line numbers assoicated with the two calls?}
+Then at the call, the parts of the caller's frame that are not passed are sealed, as usual. Passed
 memory also tracks which depth owns it, and it may only be accessed by the immediate callee
@@
 
+\apt{This needs to be talked through in more detail.}
 But because {\tt f} stashes the original pointer to {\tt x}, it will overwrite {\tt x} while

### 2021-04-07 tex/main.tex
Update based on some low-hanging-fruit comments

@@
 %% Title information
-\title{Stack Safety as a Security Property}         %% [Short Title] is optional;
+\title{Stack Safety is a Security Property}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 \begin{abstract}
-\bcp{How about ``Stack Safety {\em is} a Security Property''?}
   What is stack safety? The phrase has been used to describe the protections
@@
 to support this technique and argue that the new property, defined relative to an
-arbitrary notion of observation, is weak enough to characterize stack safety.
-\apt{Reads oddly. ``string enough'' is what we want, surely? (While still weak
-  enough to be enforced by the lazy policies.)}
+arbitrary notion of observation. This property is still strong enough to prevent
+the bad behaviors we associate with unsafe stack behavior, but weak enough that
+other reasonable enforcement mechanisms can enforce it.
 
@@
 Our key contribution is an operational model of local state encapsulation
-using the technical framework of language-based security\apt{this idea deserves more prominence}, with eager and lazy variants,
+using the technical framework of language-based security
+\apt{this idea deserves more prominence}, with eager and lazy variants,
 together with a simple characterization of well-bracketed control flow.
@@
 the behavior of function calls. First, that when a call is finished, the caller can continue
-from where it left off with its local state unchanged. And second, that the call is
-self-contained, its behavior determined only by a predictable interface -- at minimum
-parameters and global variables, but never local state.\apt{This last sentence is not so clear.}
+from where it left off with its local state unchanged. And second, that the call's
+behavior is determined only by a predictable interface that does not include the caller's
+local state.
 
@@
 Both expectations imply predictable control flow, termed {\em well-bracketed
-control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls jump to a consistent \apt{??} entry
-point and return to the call site. Then the data portion of the first expectation is for
+control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls always jump to valid entry
+points and return to their call sites. Then the data portion of the first expectation is for
 {\em integrity}: that a caller's local data is protected from modification by its callee.
 This is vital for local reasoning about the behavior of the caller. The security implications
-are significant: without integrity, a caller that branches on a local variable could be manipulated by its
-callee into unexpected behavior.
+are significant: without integrity, a caller that branches on a local variable could be
+manipulated by its callee into unexpected behavior.
 
@@
 predict the behavior of a call we need only consider a subset of the machine state.
-Rearranging local data will have no impact.\apt{??} And in security terms, we can guarantee
-the secrecy of local data that we haven't passed even when calling a badly behaved function.
-Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
-{\em local state encapsulation} (LSE) property.
+And in security terms, we can guarantee the secrecy of local data that we haven't passed
+even when calling a badly behaved function. Together, integrity and confidentiality
+comprise \citeauthor{Skorstengaard+19}'s {\em local state encapsulation} (LSE) property.
 
@@
 %
-Even more fundamentally, \apt{``For this model to make sense,'' (??)} we need to know when a call or a return is taking
+For this model to make sense, we need to know when a call or a return is taking
 place. We assume that calls and returns are annotated in the code, so that
@@
 First we introduce this model for a single simple stack with no sharing between callers and
-callee. Then we expand it to allow components\apt{??} to be passed, granting a callee one time
+callee. Then we expand it to allow passing of stack data, granting a callee one time
 access to the caller's data. We do not model more sophisticated sharing, because to do
@@
 model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
-a separate policy outside the scope of stack safety. (The combination of such policies is
-a strength of tag-based approaches.)
-
-Finally, we extend the model to a coroutine system featuring finitely-bounded stacks. Though
-simple\apt{simplistic}, this extension is a step toward concurrency and demonstrates that the model is flexible
-enough to handle more sophisticated styles of control-flow.
+a separate policy outside the scope of stack safety. (A tag-based approach can easily
+compose multiple such policies.)
+
+Finally, we extend the model to a coroutine system featuring finitely-bounded stacks.
+Though simplistic, this extension is a step toward concurrency and demonstrates that
+the model is flexible enough to handle more sophisticated styles of control-flow.
 
@@
 at every step of a program, and each call respects integrity at each of its internal steps.
-These properties are very strong, and our policies enforce them; they also have the advantage
-that they ``fail faster'' \apt{than what?} and are therefore suitable for efficient testing.
+These properties are very strong, and our policies enforce them; this also makes them very
+intensional, in that they do not distinguish internal behavior from external.
+They also have the advantage that they are suitable for efficient testing, as they
+fail immediately when a violation occurs.
 
@@
 We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
-at some performance cost. We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
-  and is it so interesting anyhow?}
-We argue that any system purporting to enforce stack safety ought to obey \apt{``at least'' ?} xslazy stack safety.
+at some performance cost.
+%We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
+%  and is it so interesting anyhow?}
+We argue that any system purporting to enforce stack safety ought to obey at least lazy
+stack safety, making it a plausible extensional characterization.
 
@@
 information would typically come from a compiler together with the machine code,
-we do not assume anything about their provenance. We do treat the execution of an
-annotated instruction as in some sense intentional on the part of the current
-function. No accidental calls or returns, and attackers control their own returns
-but cannot change the annotations on the caller's code.\apt{cryptic}
+we do not assume anything about their provenance. This means that attackers (or
+foolish defenders) are under no obligation to have reasonable annotations.
 
@@
 
-Alternatively, suppose an annotated instruction is not really an actual \apt{``call or'' ?} return.
+Alternatively, suppose an annotated instruction is not really an actual call or return.
 What happens in this case depends on the enforcement mechanism and the nature of the false
@@
 the base machine. When the policy-enhanced machine fails to step, this is a {\em failstop},
-halting the program before it can perform an action that would violate the security property; our
-properties are therefore naturally \emph{termination insensitive.}
+halting the program before it can perform an action that would violate the security property;
+our properties are therefore naturally \emph{termination insensitive.}
 This model will be made more precise in \cref{sec:prelim}.

### 2021-04-07 tex/main.tex
Further edits based on Andrew's feedback

@@
 are wrong? First, suppose a call or return occurs that is not annotated.
-In this case, the answer is simple:
-there has been no call or return\apt{from whose perspective?}, only an instruction resembling such an operation, and
+In this case, the answer is simple: as far as the properties are concerned, there has
+been no call or return, only an instruction sequence resembling such an operation, and
 therefore no associated change in permissions. This is particularly interesting in the
@@
 
-Alternatively, suppose an annotated instruction is not really an actual call or return.
-What happens in this case depends on the enforcement mechanism and the nature of the false
-annotation. A given enforcement mechanism might place a static well-formedness criterion
-on the code it applies to. For instance, our\apt{or \citeauthor{DBLP:conf/sp/RoesslerD18}'s ? }
-micro-policies expect that call and return
-annotations occur on jump instructions. On the other hand, not all jumps are proper
-calls or returns, only those preceded by appropriate sequences: in these cases the
-bad annotation might appear well formed but the system should either respect the change
-in permissions or, if doing so would compromise security, fail-stop rather than
-execute the annotated instruction.\apt{hard to follow}
+Alternatively, an annotated instruction might not really be a call or return.
+In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
+and returns are identified with ``blessed sequences'' of instructions, which the tag policy
+will force to be executed from start to finish. The annotated instruction should always
+be the final instruction of the sequence. What if an annotated instruction exists outside
+such a sequence? We will statically rule out some scenarios with a well-formedness criterion
+on initial states: annotated instructions will initially be tagged apropriately, as the
+final instruction in the call or return sequence. This is all we need for the policy to
+protect itself: it can then enforce that the call or return only occurs if that instruction
+is a jump that follows the blessed sequence. So, even in our strong attacker model, we need
+very little in the way of extra guarantees.
+
+In general, regardless of enforcement mechanism, we take this approach: 
+a given enforcement mechanism may place a static well-formedness criterion
+on the code it applies to. Weaker enforcement mechanisms might also need dynamic
+reasonableness conditions to prevent the defender from sabotaging enforcement, as
+discussed in section \ref{sec:cheri}.
 
@@
 
-\apt{I'm unclear about how many of these assumptions are relevant to the properties
-  as opposed to the (micro)policies.}
-
-Our primary target is a RISC-V architecture extended with the PIPE architecture, but the
-model makes very few assumptions about the machine it applies to. We assume a register
-machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other details of
-the machine's ISA are unimportant.
-
-Our definitions are largely parametric over the layout of the stack in memory, but for
-purposes of explanation we will describe a standard layout. Memory consists of an
-instruction region and a stack that grows upwards.\apt{if we want to be ``standard''
-  we really should have the stack grown downwards.}
-(For simplicity,
-we do not consider the heap. Integrating a memory-safety property for the
-heap in the style of \citet{DBLP:conf/post/AmorimHP18} is a natural extension
-that we discuss in~\Cref{sec:relwork}.) The stack has a conventional structure, with
-contiguously allocated per-function frames.
-
-The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
-at the time of the call. The stack can be used to pass arguments or return results
-in the usual way, by placing them at the top of the caller's frame. This will be modeled
-as a form of sharing in section \sna{cite the sharing section}.
-Callers and callees can also communicate via registers if they wish; indeed, our
-definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
-
-\ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
-  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
+Our properties are defined in terms of an abstract machine model that makes very few
+assumptions about the machine that instantiates it.  We assume a register
+machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other
+details of the machine's ISA are unimportant. We assume that we know ahead of time
+the layout of memory, including code (with annotations) and which addresses are within
+the bounds of the stack.
+%
+We further parameterize over the structure of the stack, particularly the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention} and model it as a relation between machine states
+and addresses. Then from any call, those the sealing convention relates to the
+state will be protected until the return.
 
@@
 1[C]:  y = f(10);
-2:     print (x + y);
+2[R]:  return (x + y);
 }
@@
 int f(int z) {
-10:     print z;
+10:     $(1,0) = 0;
 11[R]:  return (z+z);
@@
 \end{verbatim}
-
-\caption{A Simple Call}
-\label{fig:simple-program}
 \end{subfigure}
@@
 {\tt 0} &
-\memoryaddrs{4em}
 \memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
-~$\cdots$
+\vspace{.5em}
 \\
 {\tt 1} &
-\memoryaddrs{12em}
-\memory{3}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
+\memory{2}{\unsealc}
+\MemoryLabel{-6em}{0.75em}{42}
+\MemoryLabel{-4em}{0.75em}{0}
 \vspace{.5em}
@@
 {\tt 10} &
-\memoryaddrs{17em}
+\memory[1em]{1}{\emptyoutc}
+\hspace{.5em}
 \memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
-\memory{2}{\unsealc}
-~$\cdots$
-\MemoryLabel{-10em}{0.75em}{2}
 \\
-{\tt 2} &
-\memoryaddrs{12em}
-\memory{3}{\unsealc}
-~$\cdots$
-\MemoryLabel{-14em}{0.75em}{42}
-\MemoryLabel{-10em}{0.75em}{0}
-\MemoryLabel{-7em}{0.75em}{2}
-\\
+{\tt 11} &
+\memory[1em]{1}{\emptyoutc}
+\hspace{.5em}
+\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
+\MemoryLabel{-6em}{0.75em}{\bf 0}
 \end{tabular}
@@
 \vspace{\abovedisplayskip}
-
-\caption{Trace of Stack Domains}
-
-\label{fig:simple-trace}
 \end{subfigure}
+\label{fig:int1}
+\caption{Integrity Violation}
 \end{figure}
@@
 \item We do not find desireable behaviors that appear stack-safe yet violate our properties
-\item Our properties are testably implemented by stack safety micro-policies from the literature,
-  and the micro-policies are not significantly more conservative (i.e., policy failstops usually 
-  correspond to imminent property violations)
+\item Our properties are testably implemented by stack safety micro-policies from the
+  literature, and the micro-policies are not significantly more conservative
+  (i.e., policy failstops usually correspond to imminent property violations)
 \end{itemize}
 
-\Cref{fig:simple-program} shows a simple example, written in C-like pseudo-code.
-It consists of a {\tt main} function that performs a call to a function {\tt f},
-which prints its argument, performs arithmetic, and returns a result.
-\Cref{fig:simple-trace} shows the stack as it is divided into domains throughout
-the program execution. We use high-level code for compactness, but reason about
-a standard compilation of each function. \apt{Note assumption that arguments and return values are in registers.}
-Line numbers correspond to the code
-addresses just before the code that implements the line, and annotations
-{\tt C} and {\tt R} indicate that an instruction is annotated as a call or
-return, respectively.
-\apt{In trace, clarify difference between values (like 42,0) and line numbers (like 2).}
-
-We also assume that the machine is word- (not byte-) addressed and that the
-machine communicates with the outside world through a single memory-mapped
-output port whose address is elided (so a {\tt print} statement will be
-compiled to a store to this location), and that the stack grows upwards.
-
-In the initial memory layout the entire stack is currently {\em unsealed},
+Our definitions are largely parametric over the layout of the stack in memory, so we
+will give our initial examples in a simplified layout. Treat the stack as a non-empty
+list of frames, each a contiguous block of memory (that may be of length zero). When
+the active function allocates local variables, it does so by increasing the size of the
+head of the stack; variables declared in nested scopes increase the size further when
+entering the scope, and decrease it when leaving. During a call, the sealing convention
+designates the entire head frame to be sealed, and a new frame is pushed onto the stack.
+At a return, the stack is popped. We omit return addresses from these examples, as they
+focus on local state encapsulation.
+
+We have described an idealized stack, so we need to add a way for the attacker to violate
+stack safety. We define the expression {\tt \$(i,j)} as accessing the \(i\)th frame in the
+list at the offset \(j\). We will show how uses of this expression produce violations
+of both integrity and confidentiality. Besides this operator, examples are in C-like
+pseudocode. Line numbers represent the point in execution before executing all instructions
+compiled from that line, with annotations in brackets. (Initially, only {\tt C} for call
+and {\tt R} for return.) Variables are already allocated at the first line of a block.
+All arguments and return values are passed in registers.
+
+%The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
+%at the time of the call. The stack can be used to pass arguments or return results
+%in the usual way, by placing them at the top of the caller's frame. This will be modeled
+%as a form of sharing in section \sna{cite the sharing section}.
+%Callers and callees can also communicate via registers if they wish; indeed, our
+%definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
+
+%\ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
+%  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
+
+\Cref{fig:int1} shows an integrity violation. It consists of a {\tt main}
+function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
+variable {\tt x}, performs arithmetic, and returns a result.
+In the initial memory layout the only frame is currently {\em unsealed},
 meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals (as designated \apt{``delimited''? in any case, how?} by the stack pointer) are
-{\em sealed} and labeled with the depth of their owner, in this case 0.
-Sealing is the fundamental integrity contract between the caller and callee:
+{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
+in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
 {\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
@@
 if a component is sealed before a step, it will be unchanged afterward. Stepwise
-refers to the fact that the property must hold on every step. Later we will
-loosen this condition in the interest of extensionality -- the caller only
-cares about changes it can see after the return, not changes during execution.
-
-Suppose we insert an instruction after line 10 that stores 5 to the stack pointer - 1.\apt{Actually, -2.
-  But this also raises the question of if/how return addresses are protected.}
-Then we would be changing {\tt y}, which is sealed, in violation of stepwise integrity.
+refers to the fact that the property must hold on every step. %Later we will
+%loosen this condition in the interest of extensionality -- the caller only
+%cares about changes it can see after the return, not changes during execution.
+In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.
+
+\sna{I'm experimenting with the example style, and it ends here currently, so below
+  this is the old version in which the stack is contiguous. Current plan is to switch
+  to contiguous stack when adding stack passing.}
 
@@
   \multicolumn{2}{c}{
-    \memoryaddrs{4em}
     \memory{3}{\unsealc}
@@
 \section{Extension to Cheri}
+\label{sec:cheri}
 

### 2021-04-08 tex/main.tex
Further progress on examples

@@
 \colorlet{lgray}{gray!40}
+\colorlet{lred}{red!40}
+\colorlet{lblue}{blue!20}
 
@@
 \newcommand{\fulloutc}{white}
+\newcommand{\badc}{lred}
+\newcommand{\goodc}{lblue}
 
@@
 
-\Cref{fig:int1} shows an integrity violation. It consists of a {\tt main}
+Example \ref{fig:int1} shows an integrity violation. It consists of a {\tt main}
 function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
@@
 
-\sna{I'm experimenting with the example style, and it ends here currently, so below
-  this is the old version in which the stack is contiguous. Current plan is to switch
-  to contiguous stack when adding stack passing.}
-
 \begin{figure}
@@
 int main() {
-0:     int x = 42, y = 0;
-1[C]:  y = f();
-2:     print (x + y);
+0:     int x,y;
+1:     x = 42;
+2:     y = 0;
+3[C]:  y = f();
+4[R]:  return x+y;
 }
@@
 int f() {
-10:     int z = *($SP-1);
-11[R]:  return (z+z);
+10:     int w,z;
+11:     w = 5;
+12:     z = $(1,0);
+13[R]:  return w+z;
 }
@@
 \centering
-\begin{tabular}{l l l}
-  {\tt 0} &
+\begin{tabular}{l l | l}
+  {\tt 2} &
   \multicolumn{2}{c}{
-    \memory{3}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-14em}{0.75em}{a}
-    \MemoryLabel{-10em}{0.75em}{b}
-    \MemoryLabel{-7em}{0.75em}{c}
+    \memory{2}{\unsealc}
+    \MemoryLabel{-6em}{0.75em}{42}
+    \MemoryLabel{-3em}{0.75em}{0}
     \vspace{.5em}
   } \\
-  {\tt 1} &
-  \multicolumn{2}{c}{
-    \memoryaddrs{12em}
-    \memory{3}{\unsealc}
-    ~$\cdots$
+  {\tt 11} &
+  \memory{2}{\unsealc}
+  \hspace{1em}
+  \memory{2}{\unsealc}
+  \MemoryLabel{-17em}{0.75em}{0}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-7em}{0.75em}{42}
+  \MemoryLabel{-3em}{0.75em}{0}
+  &
+  \memory{2}{\unsealc}
+  \hspace{1em}
+  \memory{2}{\unsealc}
+  \MemoryLabel{-17em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{\(v_1\)}
+  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
+  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
+  \\
+  {\tt 12} &
+  \memory{1}{\goodc}
+  \memory{1}{\unsealc}
+  \hspace{1em}
+  \memory{2}{\unsealc}
+  \MemoryLabel{-17em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-7em}{0.75em}{42}
+  \MemoryLabel{-3em}{0.75em}{0}
+  &
+  \memory{1}{\goodc}
+  \memory{1}{\unsealc}
+  \hspace{1em}
+  \memory{2}{\unsealc}
+  \MemoryLabel{-17em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{\(v_1\)}
+  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
+  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
+  \\
+    {\tt 13} &
+  \memory{1}{\unsealc}
+  \memory{1}{\badc}
+  \hspace{1em}
+  \memory{2}{\unsealc}
+  \MemoryLabel{-17em}{0.75em}{5}
   \MemoryLabel{-14em}{0.75em}{42}
-    \MemoryLabel{-10em}{0.75em}{0}
-    \MemoryLabel{-7em}{0.75em}{c}
-    \vspace{.5em}
-  } \\
-  {\tt 10} &
-  \memoryaddrs{20em}
-  \memory{2}{\mainsealc}
+  \MemoryLabel{-7em}{0.75em}{42}
+  \MemoryLabel{-3em}{0.75em}{0}
+  &
+  \memory{1}{\unsealc}
+  \memory{1}{\badc}
+  \hspace{1em}
   \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-10em}{0.75em}{2}
-  \MemoryLabel{-7em}{0.75em}{\bf d}
- &
-  \memoryaddrs{20em}
-  \memory{2}{\mainsealc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{a'}
-  \MemoryLabel{-14em}{0.75em}{b'}
-  \MemoryLabel{-10em}{0.75em}{2}
-  \MemoryLabel{-7em}{0.75em}{\bf d'}
-  \\
-  {\tt 11} &
-  \memoryaddrs{20em}
-  \memory{2}{\mainsealc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-10em}{0.75em}{2}
-  \MemoryLabel{-7em}{0.75em}{\bf 0}
-  &
-  \memoryaddrs{20em}
-  \memory{2}{\mainsealc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{a'}
-  \MemoryLabel{-14em}{0.75em}{b'}
-  \MemoryLabel{-10em}{0.75em}{2}
-  \MemoryLabel{-7em}{0.75em}{\bf b'}
-  \\
+  \MemoryLabel{-17em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{\(v_2\)}
+  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
+  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
 \end{tabular}
@@
 
-\label{fig:conf1-trace}
+\label{fig:conf1}
 \end{subfigure}
@@
 We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
-security. Noninterference describes knowledge: it states that for some component(s)\apt{??}
-that we do not know, we can substitute any values whatsoever without changing the results
+security. Noninterference describes knowledge: it states that for some system components
+containing secret values, we can substitute any values whatsoever without changing the results
 of our computation. If there exists some value for which we see different results, then
@@
 In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
-by reading {\tt main}'s variable {\tt y}, which was not passed to it. The trace
-in figure \ref{fig:conf1-trace} shows how noninterference works. \apt{Is this right? what is d?}
-First, let \(a\),
-\(b\), and \(c\) be the unknown initial values at their locations. On entry to {\tt f}
-we again have 42 and 0 in {\tt x} and {\tt y}, and {\tt z} contains \(c\). We consider a
-{\em variant state} that replaces the values of components that should be secret with
-arbitrary values \(a'\), \(b'\), and \(c'\). After a step, we've copied 0 and \(c'\),
-respectively, into {\tt z}. Since {\tt z} changed value, for confidentiality to hold,
-its value should be the same in all variant states. But when \(c' \not = 0\), the
-states change differently, which violates confidentiality. It would also be a violation
-to print the secret value.\apt{all this is not too clear.}
+by reading {\tt main}'s variable {\tt x}, which was not passed to it. The trace
+shows how noninterference applies to this scenario. At line 11, we create a {\em variant}
+state with arbitrary values \(v_0\), etc., in place of the values that should
+be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt w}, has
+changed to 5 in both the primary trace and the variant. The primary and variant agree on
+the location that changed, so confidentiality was not violated. But after line 13,
+the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
+in the variant trace it might have any value \(v_2\). Say \(v_2 = 43\) -- then the variant
+has behaved differently than the primary, violating confidentiality.
 
@@
 as: for any state at the entry to a function, for any variant over the values in the stack,
-the primary state and its variant will step in lockstep until they return. Any components
-whose values update and any printed values will be identical between the primary and variant
-traces.
+the primary state and its variant will step in lockstep until they return. Any system components
+whose values update will be identical between the primary and variant traces. This includes not
+just memory, but registers as well, so that returning a secret is a violation. And when we extend
+the model with a notion of external output, it will include any values that are outputted.
 
@@
 guarantee that the step alters state identically to a step from a variant state. Then we
-wouldn't need to divide up the execution by calls. But this is further from the normal
-framing of noninterference, requires explicitly tracking initializations, and would behave
-no differently relative to our policies.\apt{not very clear.}
+wouldn't need to divide up the execution by calls. But we can see in this example that taking
+a single variant at the start of the call effectively captures initialization: when {\tt w} is
+initialized, the primary and the variant agree on its value, so it can now be safely read.
+This treatment is roughly in line with more standard uses of noninterference, which typically
+vary states at the starts of whole program runs.
+
+Together, stepwise integrity and lockstep confidentiality form the eager version of
+local state encapsulation.
 
@@
 
+It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt w}
+and {\tt z}. We will demonstrate by amending our toy model to make it closer to a real stack.
+In figure \ref{fig:conf2} the topmost stack frame is no longer allocated as needed, but is an
+infinite array of addresses. Variables are allocated and deallocated by moving the stack pointer,
+and on a call, addresses to the left of the stack pointer are split off and pushed onto the
+stack as sealed frames. On a return, the topmost sealed frame is prepended to the unsealed
+frame. Now we can see how secrets might be left behind even in the unsealed portion of the stack.
+
+\sna{Updated up to here.}
+
 Two important details of confidentiality are reflected in figure \ref{fig:conf2}.

### 2021-04-08 tex/main.tex
A comment and a few nits

@@
 This last provides an array of micro-policies offering protection at various granularities,
-which offer a trade-off between protection and efficiently.
+which offer a trade-off between protection and efficiency.
 
@@
 to a mechanism for protecting stack-based low-level programs to judge
-whether or not it actually ``guarantees stack safety.'' We begin with a strong
+whether or not it actually ``guarantees stack safety.''
+\rb{The remainder of this paragraph may be diving too quickly, would a reader be
+confused by the references to these types of properties, yet to be introduced?}
+We begin with a strong
 local state encapsulation property that closely models the behavior of eager
@@
 until it returns. Thus we have no need to differentiate safe code
-from adversarial - all functions, and indeed all activations of the same function,
+from adversarial --- all functions, and indeed all activations of the same function,
 are mutually distrusting.
@@
 we can track the point at which control transfers from the protected caller
-to the attacking callee -- and when that callee in turn becomes a protected
+to the attacking callee --- and when that callee in turn becomes a protected
 caller in its own right.
@@
 First we introduce this model for a single simple stack with no sharing between callers and
-callee. Then we expand it to allow passing of stack data, granting a callee one time
+callees. Then we expand it to allow passing of stack data, granting a callee one-time
 access to the caller's data. We do not model more sophisticated sharing, because to do

### 2021-04-10 tex/main.tex
Some changes to text that will be useful.
Mostly done with examples

@@
                         %% http://ctan.org/pkg/subcaption
+\usepackage{tikz}
+\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
+    \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
+
 \input{macros}
@@
 \newcommand{\goodc}{lblue}
+\newcommand{\retptrc}{black}
 
@@
 int main() {
-0:     int x = 42, y = 0;
-1[C]:  y = f(10);
-2[R]:  return (x + y);
+0:     int x, y;
+1:     y = 42;
+2[C]:  x = f();
+3[R]:  return x+y;
 }
 
-int f(int z) {
-10:     $(1,0) = 0;
-11[R]:  return (z+z);
+int f() {
+10:     $SP-2 = 0;
+11[R]:  return 5;
 }
@@
 \begin{tabular}{l l}
-{\tt 0} &
-\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
+{\tt 1} &
+\memoryaddrs{12em}
+\memory{4}{\unsealc}[Unseal]
+~$\cdots$
 \vspace{.5em}
 \\
-{\tt 1} &
-\memory{2}{\unsealc}
-\MemoryLabel{-6em}{0.75em}{42}
-\MemoryLabel{-4em}{0.75em}{0}
+{\tt 2} &
+\memoryaddrs{12em}
+\memory{4}{\unsealc}
+~$\cdots$
+    \MemoryLabel{-15em}{0.75em}{42}
     \vspace{.5em}
@@
 {\tt 10} &
-\memory[1em]{1}{\emptyoutc}
-\hspace{.5em}
-\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
+\memoryaddrs{16em}
+\memory{2}{\mainsealc}[Seal(0)]%
+\memory{1}{\retptrc}[RetPtr]%
+\memory{1}{\unsealc}[Unseal]
+~$\cdots$
+\MemoryLabel{-15em}{0.75em}{42}
+\vspace{.5em}
 \\
 {\tt 11} &
-\memory[1em]{1}{\emptyoutc}
-\hspace{.5em}
-\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
-\MemoryLabel{-6em}{0.75em}{\bf 0}
+\memoryaddrs{16em}
+\memory{2}{\mainsealc}
+\memory{1}{\retptrc}
+\memory{1}{\unsealc}
+~$\cdots$
+\MemoryLabel{-15em}{0.75em}{\bf 0}
+\vspace{.5em}
 \end{tabular}
@@
 
-Our definitions are largely parametric over the layout of the stack in memory, so we
-will give our initial examples in a simplified layout. Treat the stack as a non-empty
-list of frames, each a contiguous block of memory (that may be of length zero). When
-the active function allocates local variables, it does so by increasing the size of the
-head of the stack; variables declared in nested scopes increase the size further when
-entering the scope, and decrease it when leaving. During a call, the sealing convention
-designates the entire head frame to be sealed, and a new frame is pushed onto the stack.
-At a return, the stack is popped. We omit return addresses from these examples, as they
-focus on local state encapsulation.
-
-We have described an idealized stack, so we need to add a way for the attacker to violate
-stack safety. We define the expression {\tt \$(i,j)} as accessing the \(i\)th frame in the
-list at the offset \(j\). We will show how uses of this expression produce violations
-of both integrity and confidentiality. Besides this operator, examples are in C-like
-pseudocode. Line numbers represent the point in execution before executing all instructions
-compiled from that line, with annotations in brackets. (Initially, only {\tt C} for call
-and {\tt R} for return.) Variables are already allocated at the first line of a block.
-All arguments and return values are passed in registers.
+Our examples are written in C-like syntax, labeled with line numbers for reference,
+but to reflect the increased power of the underlying assembly code, we allow code
+to access the stack pointer directly, written {\tt \$SP}. The layout of the stack
+is given to the right, with a box representing a word of memory, and variables stored
+in declaration order from left to right. The stack grows upward in memory and to the
+right on paper.
 
@@
 function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
-variable {\tt x}, performs arithmetic, and returns a result.
-In the initial memory layout the only frame is currently {\em unsealed},
+variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
 meaning that it is eligible to be written to. At the call from {\tt main} to
@@
 
+The black rectangle represents the return address. In a typical calling convention
+this will have been passed in a register during the call itself, and {\tt f} stores
+it in the stack during its preamble, so it is not itself sealed.
+
 \begin{figure}
-
 \begin{subfigure}{.33\textwidth}
@@
 int main() {
-0:     int x,y;
+0:     int x;
 1:     x = 42;
-2:     y = 0;
-3[C]:  y = f();
-4[R]:  return x+y;
+2[C]:  x = f();
+3[R]:  return x+y;
 }
@@
 int f() {
-10:     int w,z;
-11:     w = 5;
-12:     z = $(1,0);
-13[R]:  return w+z;
+10:     int y,z;
+11:     y = 5;
+12:     z = $SP-2;
+13[R]:  return y+z;
 }
 \end{verbatim}
-\label{fig:conf1-program}
 \end{subfigure}
@@
 \begin{tabular}{l l | l}
-  {\tt 2} &
+  {\tt 1} &
   \multicolumn{2}{c}{
-    \memory{2}{\unsealc}
-    \MemoryLabel{-6em}{0.75em}{42}
-    \MemoryLabel{-3em}{0.75em}{0}
+    \memoryaddrs{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-18em}{0.75em}{0}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
   } \\
+  {\tt 10} &
+  \memoryaddrs{20em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{42}
+  \MemoryLabel{-10em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs{20em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
+  \\
   {\tt 11} &
-  \memory{2}{\unsealc}
-  \hspace{1em}
-  \memory{2}{\unsealc}
-  \MemoryLabel{-17em}{0.75em}{0}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-7em}{0.75em}{42}
-  \MemoryLabel{-3em}{0.75em}{0}
+  \memoryaddrs{20em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{1}{\goodc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{42}
+  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{0}
   &
-  \memory{2}{\unsealc}
-  \hspace{1em}
-  \memory{2}{\unsealc}
-  \MemoryLabel{-17em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
-  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
+  \memoryaddrs{20em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{1}{\goodc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
   \\
   {\tt 12} &
-  \memory{1}{\goodc}
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
-  \hspace{1em}
-  \memory{2}{\unsealc}
-  \MemoryLabel{-17em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-7em}{0.75em}{42}
-  \MemoryLabel{-3em}{0.75em}{0}
-  &
-  \memory{1}{\goodc}
-  \memory{1}{\unsealc}
-  \hspace{1em}
-  \memory{2}{\unsealc}
-  \MemoryLabel{-17em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
-  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
-  \\
-    {\tt 13} &
+  \memory{1}{\retptrc}
   \memory{1}{\unsealc}
   \memory{1}{\badc}
-  \hspace{1em}
-  \memory{2}{\unsealc}
-  \MemoryLabel{-17em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{42}
-  \MemoryLabel{-7em}{0.75em}{42}
-  \MemoryLabel{-3em}{0.75em}{0}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{42}
+  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{42}
+  \vspace{.5em}
   &
+  \memoryaddrs{20em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
   \memory{1}{\unsealc}
   \memory{1}{\badc}
-  \hspace{1em}
-  \memory{2}{\unsealc}
-  \MemoryLabel{-17em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{\(v_2\)}
-  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
-  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
+  \vspace{.5em}
 \end{tabular}
@@
 
-Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
-of the context of its call (except, of course, for its arguments).
+Meanwhile confidentiality intuitively states that a callee is unaware
+of the context of its call (except for its arguments and any global variables).
 We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
@@
 of our computation. If there exists some value for which we see different results, then
-in fact we can learn something about the components by observing those results, violating
+in fact we can learn something about the initial state by observing those results, violating
 confidentiality.
@@
 In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
-by reading {\tt main}'s variable {\tt x}, which was not passed to it. The trace
+by reading {\tt main}'s variable {\tt x}. The trace
 shows how noninterference applies to this scenario. At line 11, we create a {\em variant}
 state with arbitrary values \(v_0\), etc., in place of the values that should
-be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt w}, has
+be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt y}, has
 changed to 5 in both the primary trace and the variant. The primary and variant agree on
@@
 whose values update will be identical between the primary and variant traces. This includes not
-just memory, but registers as well, so that returning a secret is a violation. And when we extend
-the model with a notion of external output, it will include any values that are outputted.
+just memory, but registers as well, so that returning a secret is a violation. And when we
+extend the model with a notion of external output, it will include any values that are
+outputted.
 
@@
 wouldn't need to divide up the execution by calls. But we can see in this example that taking
-a single variant at the start of the call effectively captures initialization: when {\tt w} is
+a single variant at the start of the call effectively captures initialization: when {\tt y} is
 initialized, the primary and the variant agree on its value, so it can now be safely read.
@@
 
-Together, stepwise integrity and lockstep confidentiality form the eager version of
+Together, stepwise integrity and lockstep confidentiality form the stepwise version of
 local state encapsulation.
@@
 
-\begin{subfigure}{.25\textwidth}
+\begin{subfigure}{.3\textwidth}
 \begin{verbatim}
 int main() {
-0[C]:  int t = f();
-1[R]:  return t;
+0:     int x;
+1[C]:  x = f();
+2[CR]: return g();
+}
+
+int f() {
+10:     int y;
+11:     y = 5;
+12[R]:  return y;
+}
+
+int g() {
+20:     int z;
+21:     if(z == 5)
+22[R]:    return 100;
+23:     else
+24[R]:    return 10;
 }
@@
 \end{subfigure}
-\begin{subfigure}{.37\textwidth}
-\begin{verbatim}
-int f() {
-10:     int x, y = 5, z = 6;
-11:     print y+z;
-12:     $SP-=2;
-13[C]:  x = g();
-14[R]:  return x;
-}
-\end{verbatim}
-\end{subfigure}
-\begin{subfigure}{.32\textwidth}
-\begin{verbatim}
-int g() {
-20:     int z;
-21[R]:  return (z+z);
-}
-\end{verbatim}
-\end{subfigure}
-
-\begin{subfigure}{.8\textwidth}
+\begin{subfigure}{.65\textwidth}
 \begin{center}
-\begin{tabular}{l l l}
-  {\tt 13} &
+\begin{tabular}{l r | l}
+  \circled{\tt 01} &
   \multicolumn{2}{c}{
-    \memoryaddrs{16em}
-    \memory{1}{\mainsealc}
+    \memoryaddrs{8em}
     \memory{4}{\unsealc}
     ~$\cdots$
-    \MemoryLabel{-23em}{0.75em}{a}
-    \MemoryLabel{-19em}{0.75em}{b}
-    \MemoryLabel{-15em}{0.75em}{1}
-    \MemoryLabel{-11em}{0.75em}{5}
-    \MemoryLabel{-8em}{0.75em}{6}
+    \MemoryLabel{-18em}{0.75em}{0}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{0}
+    \vspace{.5em}
+  } \\
+  \circled{\tt 11} &
+  \memoryaddrs{16em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs{16em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
+  \\
+  \circled{\tt 12} &
+  \memoryaddrs{16em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs{16em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-10em}{0.75em}{\(5\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
+  \\
+  \circled{\tt 02} &
+  \multicolumn{2}{c}{
+    \memoryaddrs{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-18em}{0.75em}{5}
+    \MemoryLabel{-14em}{0.75em}{2}
+    \MemoryLabel{-10em}{0.75em}{5}
+    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
@@
   \\
-  {\tt 20} &
-  \memoryaddrs{25em}
-  \memory{1}{\mainsealc}
-  \memory{2}{\fsealc}
+  \circled{\tt 21} &
+  \memoryaddrs{16em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
   \memory{2}{\unsealc}
   ~$\cdots$
-  \MemoryLabel{-23em}{0.75em}{a}
-  \MemoryLabel{-19em}{0.75em}{b}
-  \MemoryLabel{-15em}{0.75em}{1}
-  \MemoryLabel{-11em}{0.75em}{14}
-  \MemoryLabel{-8em}{0.75em}{\bf 6}
-  \vspace{.5em}
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{0}
   &
-  \memoryaddrs{25em}
-  \memory{1}{\mainsealc}
-  \memory{2}{\fsealc}
+  \memoryaddrs{16em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
   \memory{2}{\unsealc}
   ~$\cdots$
-  \MemoryLabel{-23em}{0.75em}{a'}
-  \MemoryLabel{-19em}{0.75em}{b'}
-  \MemoryLabel{-15em}{0.75em}{c'}
-  \MemoryLabel{-11em}{0.75em}{14}
-  \MemoryLabel{-8em}{0.75em}{\bf d'}
-  \vspace{.5em}
+  \MemoryLabel{-19em}{0.75em}{\(v_3\)}
+  \MemoryLabel{-10em}{0.75em}{\(v_4\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_5\)}
+  \\ &
+  \circled{\bf 22}
+  ~$\cdots$
+  &
+  \circled{\bf 24}
+  ~$\cdots$
   \\
@@
 
-It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt w}
-and {\tt z}. We will demonstrate by amending our toy model to make it closer to a real stack.
-In figure \ref{fig:conf2} the topmost stack frame is no longer allocated as needed, but is an
-infinite array of addresses. Variables are allocated and deallocated by moving the stack pointer,
-and on a call, addresses to the left of the stack pointer are split off and pushed onto the
-stack as sealed frames. On a return, the topmost sealed frame is prepended to the unsealed
-frame. Now we can see how secrets might be left behind even in the unsealed portion of the stack.
-
-\sna{Updated up to here.}
-
-Two important details of confidentiality are reflected in figure \ref{fig:conf2}.
-First of all, it has multiple calls, and while we focus on the stack at the call
-from line 13 to line 20, it should be clear that we also need to consider a similar
-variant at the call from 0 to 10. Confidentiality is in this sense recursive: each
-nested call must obey confidentiality in its own right.
-
-More importantly, careful readers will have noticed that the informal definition
-of stack confidentiality refers to a variation of the entire stack, not just the
-sealed portion. In fact, we support temporal stack safety, in which values left behind
-by previous activations are also inaccessible. In this case, we connect temporal safety
-firmly to the caller-callee attacker model -- {\tt f} stored some data in the variables
-{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.\apt{Where is the storage for x ?.
-And need to explain what the colors mean.}
-After the stack pointer is written over {\tt y}, the value of {\tt z} is still
-left behind. Then {\tt g} can read it just by allocating and accessing an uninitialized
-variable! \apt{Need to be more explicit about what is getting varied, when, and why.}
-
-\paragraph*{Control Flow and Awkward Examples}
-
-\begin{figure}
-  \centering
-  \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-int stash = 0;
-
-int main() {
-0:     int x = 1;
-1[C]:  f();
-2:     x = -x;
-3[C]:  f();
-4[R]:  return x;
-}
-\end{verbatim}
-  \end{minipage}
-  \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-void f() {
-10:     if(!stash) {
-        // store return pointer
-11:       stash = $SP-1;
-12:     } else {
-13:       $SP-1 = stash;
-14:       stash = 0;
-15[R]:    return;
-        }
-}
-\end{verbatim}
-  \end{minipage}
-\label{fig:WBCF1}
+It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt y}
+and {\tt z}. In figure \ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
+that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
+{\tt y} without even directly using its ability to read the stack pointer. In
+this case, {\tt g} branches on that value, which is a different way of
+violating lockstep confidentiality. Technically, this violation
+occurs because the program counter is also a component of the state that must change
+identically between the primary and variant traces. In a case where \(v_1 \not = 5\),
+the program counters will differ, breaking the lockstep.
+
+%\paragraph*{Control Flow and Awkward Examples}
+
+%\begin{figure}
+%  \centering
+%  \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%int stash = 0;
+
+%int main() {
+%0:     int x = 1;
+%1[C]:  f();
+%2:     x = -x;
+%3[C]:  f();
+%4[R]:  return x;
+%}
+%\end{verbatim}
+%  \end{minipage}
+%  \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%void f() {
+%10:     if(!stash) {
+%        // store return pointer
+%11:       stash = $SP-1;
+%12:     } else {
+%13:       $SP-1 = stash;
+%14:       stash = 0;
+%15[R]:    return;
+%        }
+%}
+%\end{verbatim}
+%  \end{minipage}
+%\label{fig:WBCF1}
 %  \begin{subfigure}[t]{.4\textwidth}
@@
 %  \end{subfigure}
-  \caption{Control Flow Attacks}
-  \label{fig:WBCF}
-\end{figure}
-
-Example \ref{fig:WBCF} shows how control flow attacks can compromise integrity in principle
-even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
-global variable and retrieves it the second time it is called, returning to the first call
-point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
-line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.
-
-To prevent this, we define Well-bracketed Control Flow. Throughout the above examples, we
-have implicitly tracked the depth of the current call in order to seal memory at the correct
-depth. We also have a clear sense of where the entry points to functions are -- the top line,
-in our examples, and another instruction annotation in practice. Well-bracketed Control Flow
-holds if:
-
-\begin{itemize}
-\item All calls jump to entry points ({\em entry integrity})
-\item The depth counter decreases by exactly one on a return ({\em non-skipping})
-\item From a call at depth \(d\), the next state to have depth \(d\) matches the stack pointer,
-  and the program counter is at the instruction following the jal ({\em return integrity})
-\end{itemize}
-
-Non-skipping may seem strange -- surely the depth counter ought to decrease by one at a time
-as part of the property definition, and the problem comes when the return is to a shallower
-call's return pointer, violating return integrity? Yet some weaker enforcement mechanisms,
-such as those in which returns are implemented as invocation of a Cheri object capability,
-do allow skipping when adversarial code acts unreasonably by leaking its own return capability,
-reasoning that this is acceptable as long as only the unreasonable function is skipped
-(e.g. \citep{Skorstengaard+19b}).\apt{???} A variant of our model supports this weaker form of WBCF,
-but with tag-based policies we can enforce the non-skipping version entirely.\apt{??}
-\sna{Cite future work for details when the relevant section is finished.}
+%  \caption{Control Flow Attacks}
+%  \label{fig:WBCF}
+%\end{figure}
+
+%Example \ref{fig:WBCF} shows how control flow attacks can compromise integrity in principle
+%even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
+%global variable and retrieves it the second time it is called, returning to the first call
+%point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
+%line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.
+
+%To prevent this, we define Well-bracketed Control Flow. Throughout the above examples, we
+%have implicitly tracked the depth of the current call in order to seal memory at the correct
+%depth. We also have a clear sense of where the entry points to functions are -- the top line,
+%in our examples, and another instruction annotation in practice. Well-bracketed Control Flow
+%holds if:
+
+%\begin{itemize}
+%\item All calls jump to entry points ({\em entry integrity})
+%\item The depth counter decreases by exactly one on a return ({\em non-skipping})
+%\item From a call at depth \(d\), the next state to have depth \(d\) matches the stack pointer,
+%  and the program counter is at the instruction following the jal ({\em return integrity})
+%\end{itemize}
+
+%Non-skipping may seem strange -- surely the depth counter ought to decrease by one at a time
+%as part of the property definition, and the problem comes when the return is to a shallower
+%call's return pointer, violating return integrity? Yet some weaker enforcement mechanisms,
+%such as those in which returns are implemented as invocation of a Cheri object capability,
+%do allow skipping when adversarial code acts unreasonably by leaking its own return capability,
+%reasoning that this is acceptable as long as only the unreasonable function is skipped
+%(e.g. \citep{Skorstengaard+19b}).\apt{???} A variant of our model supports this weaker form of WBCF,
+%but with tag-based policies we can enforce the non-skipping version entirely.\apt{??}
+%\sna{Cite future work for details when the relevant section is finished.}
 
@@
 \newcommand{\mainpassc}{magenta}
-\newcommand{\sharec}{lgray}
 

### 2021-04-10 tex/main.tex
major pass on intro

--- 2021-04-08 tex/main.tex
@@
 \begin{abstract}
+\apt{maybe: ``Stack safety is a testable security property'' ?}a
   What is stack safety? The phrase has been used to describe the protections
@@
 %% comma separated list
-\keywords{Stack Safety, Micropolicies}  %% \keywords are mandatory in final camera-ready submission
+\keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission
 
@@
 shadow stacks~\citep{Dang+15,Shanbhogue+19},
-capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
+capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
@@
 \fi
-This last provides an array of micro-policies offering protection at various granularities,
-which offer a trade-off between protection and efficiently.
-
-But what {\em is} the stack safety that these techniques enforce? Precise formal
-definitions are mostly lacking: rather, the notion is commonly defined in a
-negative way, by enumerating the bad things that a given enforcement mechanism
+% 
+The implementation and configuration of these mechanisms can be complex, 
+and often involves making compromises that trade off protection for efficiency.  
+Thus it would be very useful to have precise formal specifications for stack safety
+that give us a basis for comparing the security guarantees claimed by different enforcement
+techniques, and for verifying that these claims are met by particular implementations.
+
+For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
+using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
+In essence, memory locations within each stack frame are tagged with the identity of
+the function that owns the frame, and access to those locations is then gated
+according to the function that is currently executing. Many alternatives can
+be implemented on top of this basic idea, by choosing how to identify functions
+and deciding just which accesses are checked.  These alternatives vary in their
+efficiency, but also in the spatial and temporal precision of their security guarantees.
+For example, the \emph{Depth Isolation} policy strictly requires that a function owns any
+frame into which it writes; this seems simple and intuitively correct, but can be expensive to implement.
+On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by 
+allowing illicit writes to the frame so long as the written locations are not
+subsequently read. The intuition justifying laziness is that any genuinely dangerous writes will ultimately be
+detected before they can affect program behavior.  But we clearly need a more formal foundation for
+expressing, comparing and verifying these micro-policy security claims, as well as those of
+other enforcement technologies.
+
+So how should we formally define stack safety? Definitions in the literature are commonly informal
+and proceed by enumerating the specific bad things that a given enforcement mechanism
 prevents, such as corruption of return addresses, buffer overflows, or
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
-\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
-{\em local state encapsulation} and {\em well-bracketed control flow}.
-These properties are formalized by defining a capability-based machine
-that ``satisfies [stack safety] by construction.''
-%
-While this work represents an important step in the right direction, we believe
-there is still room to craft an operational characterization
-of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
-In other words, we seek formal criteria that can be applied
-to a mechanism for protecting stack-based low-level programs to judge
-whether or not it actually ``guarantees stack safety.'' We begin with a strong
-local state encapsulation property that closely models the behavior of eager
-micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
-is too strong to characterize stack safety; reasonable enforcement mechanisms
-exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
-the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
-to support this technique and argue that the new property, defined relative to an
-arbitrary notion of observation. This property is still strong enough to prevent
-the bad behaviors we associate with unsafe stack behavior, but weak enough that
-other reasonable enforcement mechanisms can enforce it.
-
-While our criteria are designed with PIPE micro-policies in mind, they are
+The most abstract characterization of which we are aware is in 
+recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
+{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
+Informally, LSE says that the contents of
+a caller's stack frame are not read and written while its callees (and their callees, etc.) are
+executing, and WBCF says that callees always
+return to the point just after their call (or don't return at all). 
+\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
+in which they visibly hold by construction; an enforcement mechanism (in their case, hardware capabilities)
+can then be shown to provide stack safety by giving a fully abstract embedding function
+from the idealized machine to an ordinary machine equipped with the enforcement.
+
+In this work, we take the (informal) definitions of WBCF and LSE as a starting point,
+but choose a different approach to formalizing them, based on the
+technical framework of language-based security~\citep{??}.  We define
+trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
+In applying these security concepts, we treat each caller-callee pair as mutually distrustful
+and view the caller as needing protection from the callee. \apt{Need more explanation of
+  ``nesting'' and our novelty factor here? Also about domains??}
+{\em Integrity} says that a caller's local data is protected from modification by its callee.
+This is vital for local reasoning about the behavior of the caller. 
+{\em Confidentiality} states that a caller's data is protected
+from being leaked into its callee, and also that the callee cannot see uninitialized data.
+As a reasoning principle, confidentiality guarantees that the callee's
+behavior is determined only by a predictable interface that does not include the caller's
+local state or the uninitialized portion of the stack.
+Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
+notion that data outside of live frames should be inaccessible (which LSE does not address).
+We also give a simple trace-based formalization of WBCF, which is essential for keeping the
+notions of ``caller' and ``callee'' sensible.
+
+To illustrate the applicability of these properties, we perform randomized testing to
+validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces them.
+\apt{more to say?}
+
+However, their Lazy Tagging and Clearing micro-policy does not enforce stepwise integrity.
+To address this, we develop an \emph{observational stack integrity} property that 
+is intended to specify enforcement mechanisms that 
+detect integrity violations between the point when they occur and the point when they make
+an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous
+writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.
+
+Unfortunately the existing Lazy Tagging and Clearing 
+micro-policy does not quite enforce this property either, as we show in
+testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
+We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce observational integrity,
+at some performance cost.\apt{but we're not testing that, right?}
+
+Observational integrity is attractive in being significantly more \emph{extensional} than
+stepwise integrity.  We argue that it may therefore be a better minimal specification for
+stack safety enforcement mechanisms in general.
+
+Our initial description of the properties assumes single simple stack with no sharing between callers and
+callees.  We expand it to allow passing of scalar stack data, granting a callee 
+access to the caller's data. We do not model more sophisticated sharing, because to do
+so realistically would require a separate protection scheme, such as a capability
+model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
+a separate policy outside the scope of stack safety. (A tag-based approach can easily
+compose multiple such policies.)
+
+Finally, we define integrity and confidentiality properties for a coroutine system featuring finitely-bounded stacks.
+Though simplistic, this extension is a step toward concurrency and demonstrates that
+our approach is flexible enough to handle more sophisticated styles of control-flow.
+
+While our property definitions were developed with micro-policy enforcement in mind, they
 agnostic about many implementation details and should be applicable to other
 systems with minor adjustment.
-%
-Our key contribution is an operational model of local state encapsulation
-using the technical framework of language-based security
-\apt{this idea deserves more prominence}, with eager and lazy variants,
-together with a simple characterization of well-bracketed control flow.
 
@@
 
-\paragraph*{The Use of Stack Safety}
-
-How does the programmer rely on stack safety? The stack makes implicit promises about
-the behavior of function calls. First, that when a call is finished, the caller can continue
-from where it left off with its local state unchanged. And second, that the call's
-behavior is determined only by a predictable interface that does not include the caller's
-local state.
-
-%The most natural example is the first
-%in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
-%calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
-%{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
-%overwrite {\tt main}'s stack data.
-
-Both expectations imply predictable control flow, termed {\em well-bracketed
-control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls always jump to valid entry
-points and return to their call sites. Then the data portion of the first expectation is for
-{\em integrity}: that a caller's local data is protected from modification by its callee.
-This is vital for local reasoning about the behavior of the caller. The security implications
-are significant: without integrity, a caller that branches on a local variable could be
-manipulated by its callee into unexpected behavior.
-
-Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
-from leakage by its callee. As a reasoning principle, confidentiality tells us that to
-predict the behavior of a call we need only consider a subset of the machine state.
-And in security terms, we can guarantee the secrecy of local data that we haven't passed
-even when calling a badly behaved function. Together, integrity and confidentiality
-comprise \citeauthor{Skorstengaard+19}'s {\em local state encapsulation} (LSE) property.
-
-It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
-could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
-accessing the caller's state, but the execution would remain under the callee's control.
-
-To phrase LSE and WBCF as security properties requires an explicit attacker model:
-a clear understanding of who is being protected and from what. Most existing
-stack protection mechanisms envisage scenarios where a callee is
-influenced by external inputs to corrupt the stack in some way. To
-avoid getting into the details of exactly how this happens when
-formalizing our properties, we use an even stronger model: when we
-state the protections that a function can rely on when it makes a
-call, we assume that the callee can execute arbitrary instructions
-until it returns. Thus we have no need to differentiate safe code
-from adversarial - all functions, and indeed all activations of the same function,
-are mutually distrusting.
-%
-For this model to make sense, we need to know when a call or a return is taking
-place. We assume that calls and returns are annotated in the code, so that
-we can track the point at which control transfers from the protected caller
-to the attacking callee -- and when that callee in turn becomes a protected
-caller in its own right.
-
-This ``nesting'' of properties is a key challenge in adapting the concepts
-of language-based security to the stack setting. We address this issue by
-turning any set of annotations into a dynamic model of {\em domains},
-in which some stack locations are sealed by the caller that needs to preserve
-them while the rest are unsealed and available for use by the active, attacking
-callee.
-
-First we introduce this model for a single simple stack with no sharing between callers and
-callee. Then we expand it to allow passing of stack data, granting a callee one time
-access to the caller's data. We do not model more sophisticated sharing, because to do
-so realistically would require a separate protection scheme, such as a capability
-model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
-a separate policy outside the scope of stack safety. (A tag-based approach can easily
-compose multiple such policies.)
-
-Finally, we extend the model to a coroutine system featuring finitely-bounded stacks.
-Though simplistic, this extension is a step toward concurrency and demonstrates that
-the model is flexible enough to handle more sophisticated styles of control-flow.
-
-After formally defining stack safety, we show that the definition is
-enforceable using an existing tag-based enforcement
-mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
-\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy.
-This policy testably enforces our {\em eager} properties, in which integrity is respected
-at every step of a program, and each call respects integrity at each of its internal steps.
-These properties are very strong, and our policies enforce them; this also makes them very
-intensional, in that they do not distinguish internal behavior from external.
-They also have the advantage that they are suitable for efficient testing, as they
-fail immediately when a violation occurs.
-
-The downside of the Depth Isolation policy is that it is rather expensive to implement.
-In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
-that is more efficient by virtue of being lazy in its enforcement.
-Instead of enforcing the integrity and confidentiality of the stack at the
-exact moment when a callee returns, this policy signals violations when the
-caller actually accesses data that the callee has written into its stack
-frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
-  property in the same way...}\fi
-The lazy policy does not enforce our eager integrity property, and to our
-knowledge there is no formal statement of the properties that it ought to enforce.
-We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
-policy that detects integrity violations between when they occur and when they make
-a visible impact. Intuitively this can be phrased as ``no visible reads after dangerous
-writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
-``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
-testing, but formalizing the ideal property allows us to see how it falls short.
-We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
-at some performance cost.
-%We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
-%  and is it so interesting anyhow?}
-We argue that any system purporting to enforce stack safety ought to obey at least lazy
-stack safety, making it a plausible extensional characterization.
-
-%\bcp{I worry that we are being a little
-%  dishonest here: the stated advantage of the lazy policy is that it is more
-%efficient, but our variant is {\em not} efficient!}
-
 In summary, we offer the following contributions:
@@
 \item 
-  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
-  A novel {\em stack-safety property} capturing the intuitive concepts of strictly enforced
-  local state encapsulation and well-bracketed control flow~(\cref{sec:lse-and-wbcf}).
-  % These properties rule out common examples of stack memory abuse.
+  We harness the language-based security concepts of integrity and confidentiality 
+  to formalize a parameterized {\em stepwise stack-safety property}, and argue that
+  it captures the intuitive meaning of local stack encapsulation, well-bracketed control
+  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
 \item
-  A property-based randomized testing framework that demonstrates that this
-  property can be enforced by the Depth Isolation policy of
-  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware
-  tagging~(\cref{sec:conservative}).
+  We apply a property-based randomized testing framework to provide evidence that 
+  this stepwise property is enforced by the Depth Isolation policy of
+  \citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:conservative}).
 \item 
-  A {\em lazy} variant of the stack-safety property that can be enforced if the
-  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
-  with unique tags on each activation, plus a proof (in Coq) that this
-  variant is weaker than the original~(\cref{sec:lazy}).
+  We formalize a weaker but more extensional \emph{observational} variant of the
+  stack safety property, and argue that it accurately characterizes 
+  the Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18},
+  when enhanced with unique tags on each activation~(\cref{sec:lazy}).
 \item 
-  Standardized trace-based formulations of both eager and lazy confidentiality and integrity,
-  parameterizeable by arbitrary control-flow models and assignments of machine components
-  to {\em LSE domains}.
-\item
-  An LSE-domain model enhanced with passing of function arguments on the stack, with which our
-  standardized trace properties extend stack safety with local sharing.
-\item
-  An LSE-domain model that applies to simple coroutining, with corresponding LSE
-  properties and a {\em well-structured control flow} property that subsumes WBCF for each
-  stack as well as guaranteeing legitimate coroutine control transfers.
+  We show how our models can be instantiated and extended to handle passing
+  arguments on the stack~(\cref{sec:??}) and to support simple coroutining~(\cref{sec:??}).
 %  This policy does not enforce our strict
@@
 %  realizible by
-%  a previously proposed micropolicy enforcement
-%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
+%  a previously proposed micro-policy enforcement
+%  mechanism~(\Cref{sec:micro-policy}). \leo{And it says yes on a
 %    substantial body of code}. We also present a variation of the
@@
 (\cref{sec:running-example}),  and some technical preliminaries
-(\cref{sec:prelim}).
+(\cref{sec:prelim}).\apt{this para needs updating}
 
@@
 
-This enforcement model directly describes micropolicies~\citep{pump_oakland2015}.
+This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
 We believe that it is flexible enough to capture code-altering approaches, such as code
@@
 
-\paragraph*{Micropolicies}
-%
-Micropolicies are a flexible tag-based, hardware-accelerated
+\paragraph*{Micro-Policies}
+%
+Micro-Policies are a flexible tag-based, hardware-accelerated
 reference monitoring mechanism; they been applied to stack safety
@@
 
-Up til now we have assumed that function arguments and returns are passed solely in registers.
+Until now we have assumed that function arguments and returns are passed solely in registers.
 This is an oversimplication, which we will now correct. There are important stack safety concerns
@@
 
-%\paragraph*{Micropolicies}
+%\paragraph*{Micro-Policies}
 %
@@
 allows fine-grained manipulation of metadata tags to encode so-called
-\emph{micropolicies}~\citep{pump_oakland2015}.
+\emph{micro-policies}~\citep{pump_oakland2015}.
 In such a system, all values in
@@
 are enriched with an abstract metadata tag, which can represent
-arbitrary information about the value. A micropolicy is defined as a set of tags and a
+arbitrary information about the value. A micro-policy is defined as a set of tags and a
 collection of software-defined rules, indexed by machine opcode.
@@
 fail-stops with a policy violation.
-\citeauthor{pump_oakland2015} have shown that a wide range of micropolicies can
+\citeauthor{pump_oakland2015} have shown that a wide range of micro-policies can
 be defined using this scheme.
 
-Efficient execution of these micropolicies relies on hardware implementations,
+Efficient execution of these micro-policies relies on hardware implementations,
 such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
@@
 on a co-processor) to compute the rule result. To obtain adequate performance,
-it is important to design micropolicies so that they hit in the cache as
+it is important to design micro-policies so that they hit in the cache as
 much as possible. Thus, practical policies maintain a small working set of
@@
 %
-The first micropolicy we present to enforce the definition of stack
+The first micro-policy we present to enforce the definition of stack
 safety developed in \cref{sec:lse-and-wbcf} is a small variation
@@
 
-Our micropolicy works as follows. (We assume for ease of exposition that no
+Our micro-policy works as follows. (We assume for ease of exposition that no
 arguments are passed on the stack.)
 %% \rb{Come back to this later?}.
-The micropolicy maintains tags on values in memory of the form $\tagStackDepth{n}$,
+The micro-policy maintains tags on values in memory of the form $\tagStackDepth{n}$,
 indicating locations that belong to the stack frame at activation depth $n$,
 or $\tagNoDepth$; it tags the $\PCname$ with tags of the form $\tagPCDepth{n}$.
-During normal execution, the micropolicy rules only permit load and store operations
+During normal execution, the micro-policy rules only permit load and store operations
 when the target memory is tagged with the same depth as the current {\PCname} tag.
@@
 well-bracketed control flow.
-The micropolicy guarantees atomic execution of these sequences
+The micro-policy guarantees atomic execution of these sequences
 using a combination of tags on the instructions
@@
 sequence at a point where the stack-safety property expects a return,
-the micropolicy will incorrectly behave as if execution
+the micro-policy will incorrectly behave as if execution
 is continuing in the callee, but since the callee never has stronger access
@@
 %% Only programs whose sequences are well-formed are
-%% protected by the micropolicy, although programs are still protected
+%% protected by the micro-policy, although programs are still protected
 %% (by fail-stopping) even if the intended sequences are missing. By
 %% tracking the execution of the blessed sequences in the tag state of
-%% the {\PCname} register, the micropolicy ensures that the entry
+%% the {\PCname} register, the micro-policy ensures that the entry
 %% sequences of calls declared in the call map are executed correctly.
@@
 %% Because the access rights of callers subsume those of callees, it is essential
-%% that the micropolicy change tags when the property thinks that a call occurs,
-%% but it is safe for the micropolicy to fail to reset to the caller's tag when the
-%% property thinks a return occurs. (Dually, it is safe for the micropolicy to
+%% that the micro-policy change tags when the property thinks that a call occurs,
+%% but it is safe for the micro-policy to fail to reset to the caller's tag when the
+%% property thinks a return occurs. (Dually, it is safe for the micro-policy to
 %% change tags even without executing a call point, but not for it to reset tags to
@@
 to practical capability machines, especially in terms of efficiency---previous
-work on micropolicies \citep{yannis-report} has shown how to use that framework
+work on micro-policies \citep{yannis-report} has shown how to use that framework
 to implement linear return capabilities.\bcp{But IIUC the Dover realization
-  of micropolicies can't do this...?}
-
-\paragraph{Protecting the Stack with Micropolicies}
+  of micro-policies can't do this...?}
+
+\paragraph{Protecting the Stack with Micro-Policies}
 %
@@
 attacks), and study the protection of stack data through three families of
-micropolicies that tag stack objects with a pair of frame and object identifiers
+micro-policies that tag stack objects with a pair of frame and object identifiers
 used to validate accesses to the stack: Return Address Protection (which

### 2021-04-10 tex/main.tex
recover most of old intro as comments

@@
 from the idealized machine to an ordinary machine equipped with the enforcement.
+%%
+%% While this work represents an important step in the right direction, we believe
+%% there is still room to craft an operational characterization
+%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
+%% In other words, we seek formal criteria that can be applied
+%% to a mechanism for protecting stack-based low-level programs to judge
+%% whether or not it actually ``guarantees stack safety.''
+%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
+%% confused by the references to these types of properties, yet to be introduced?}
+%% We begin with a strong
+%% local state encapsulation property that closely models the behavior of eager
+%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
+%% is too strong to characterize stack safety; reasonable enforcement mechanisms
+%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
+%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
+%% to support this technique and argue that the new property, defined relative to an
+%% arbitrary notion of observation. This property is still strong enough to prevent
+%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
+%% other reasonable enforcement mechanisms can enforce it.
 
@@
 agnostic about many implementation details and should be applicable to other
-systems with minor adjustment.
-
-%\begin{figure}[b]
-%    \begin{minipage}{.4\textwidth}
-%\begin{verbatim}
-%int main() {
-%  int x = 42;
-%  f();
-%  print(x);
-%}
-%\end{verbatim}
-%    \end{minipage}
-%    \begin{minipage}{.4\textwidth}
-%\begin{verbatim}
-%int main() {
-%  int x = 1;
-%  f();
-%  if(x)
-%    g();
-%  else
-%    h();
-%}
-%\end{verbatim}
-%    \end{minipage}
-%  \caption{Simple Examples}
-%  \label{fig:examples1}
-%\end{figure}
-
-%\begin{figure}[b]
-%    \begin{minipage}{.4\textwidth}
-%\begin{verbatim}
-%int main(int arg) {
-%  if(arg)
-%    g();
-%  else
-%    h();
-%  f();
-%}
-%\end{verbatim}
-%    \end{minipage}
-%    \begin{minipage}{.4\textwidth}
-%\begin{verbatim}
-%int main() {
-%  int x = 1;
-%  f();
-%  x = -x;
-%  f();
-%  print(x);
-%}
-%\end{verbatim}
-%  \end{minipage}
-%
-%  \caption{More Examples}
-%  \label{fig:examples2}
-%\end{figure}
+systems with minor adjustment.\apt{move this up?}
+
+%% %\begin{figure}[b]
+%% %    \begin{minipage}{.4\textwidth}
+%% %\begin{verbatim}
+%% %int main() {
+%% %  int x = 42;
+%% %  f();
+%% %  print(x);
+%% %}
+%% %\end{verbatim}
+%% %    \end{minipage}
+%% %    \begin{minipage}{.4\textwidth}
+%% %\begin{verbatim}
+%% %int main() {
+%% %  int x = 1;
+%% %  f();
+%% %  if(x)
+%% %    g();
+%% %  else
+%% %    h();
+%% %}
+%% %\end{verbatim}
+%% %    \end{minipage}
+%% %  \caption{Simple Examples}
+%% %  \label{fig:examples1}
+%% %\end{figure}
+
+%% %\begin{figure}[b]
+%% %    \begin{minipage}{.4\textwidth}
+%% %\begin{verbatim}
+%% %int main(int arg) {
+%% %  if(arg)
+%% %    g();
+%% %  else
+%% %    h();
+%% %  f();
+%% %}
+%% %\end{verbatim}
+%% %    \end{minipage}
+%% %    \begin{minipage}{.4\textwidth}
+%% %\begin{verbatim}
+%% %int main() {
+%% %  int x = 1;
+%% %  f();
+%% %  x = -x;
+%% %  f();
+%% %  print(x);
+%% %}
+%% %\end{verbatim}
+%% %  \end{minipage}
+%% %
+%% %  \caption{More Examples}
+%% %  \label{fig:examples2}
+%% %\end{figure}
+
+%% \paragraph*{The Use of Stack Safety}
+
+%% How does the programmer rely on stack safety? The stack makes implicit promises about
+%% the behavior of function calls. First, that when a call is finished, the caller can continue
+%% from where it left off with its local state unchanged. And second, that the call's
+%% behavior is determined only by a predictable interface that does not include the caller's
+%% local state.
+
+%% %The most natural example is the first
+%% %in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
+%% %calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
+%% %{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
+%% %overwrite {\tt main}'s stack data.
+
+%% Both expectations imply predictable control flow, termed {\em well-bracketed
+%% control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls always jump to valid entry
+%% points and return to their call sites. Then the data portion of the first expectation is for
+%% {\em integrity}: that a caller's local data is protected from modification by its callee.
+%% This is vital for local reasoning about the behavior of the caller. The security implications
+%% are significant: without integrity, a caller that branches on a local variable could be
+%% manipulated by its callee into unexpected behavior.
+
+%% Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
+%% from leakage by its callee. As a reasoning principle, confidentiality tells us that to
+%% predict the behavior of a call we need only consider a subset of the machine state.
+%% And in security terms, we can guarantee the secrecy of local data that we haven't passed
+%% even when calling a badly behaved function. Together, integrity and confidentiality
+%% comprise \citeauthor{Skorstengaard+19}'s {\em local state encapsulation} (LSE) property.
+
+%% It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
+%% could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
+%% accessing the caller's state, but the execution would remain under the callee's control.
+
+%% To phrase LSE and WBCF as security properties requires an explicit attacker model:
+%% a clear understanding of who is being protected and from what. Most existing
+%% stack protection mechanisms envisage scenarios where a callee is
+%% influenced by external inputs to corrupt the stack in some way. To
+%% avoid getting into the details of exactly how this happens when
+%% formalizing our properties, we use an even stronger model: when we
+%% state the protections that a function can rely on when it makes a
+%% call, we assume that the callee can execute arbitrary instructions
+%% until it returns. Thus we have no need to differentiate safe code
+%% from adversarial --- all functions, and indeed all activations of the same function,
+%% are mutually distrusting.
+%% %
+%% For this model to make sense, we need to know when a call or a return is taking
+%% place. We assume that calls and returns are annotated in the code, so that
+%% we can track the point at which control transfers from the protected caller
+%% to the attacking callee --- and when that callee in turn becomes a protected
+%% caller in its own right.
+
+%% This ``nesting'' of properties is a key challenge in adapting the concepts
+%% of language-based security to the stack setting. We address this issue by
+%% turning any set of annotations into a dynamic model of {\em domains},
+%% in which some stack locations are sealed by the caller that needs to preserve
+%% them while the rest are unsealed and available for use by the active, attacking
+%% callee.
+
+%% First we introduce this model for a single simple stack with no sharing between callers and
+%% callees. Then we expand it to allow passing of stack data, granting a callee one-time
+%% access to the caller's data. We do not model more sophisticated sharing, because to do
+%% so realistically would require a separate protection scheme, such as a capability
+%% model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
+%% a separate policy outside the scope of stack safety. (A tag-based approach can easily
+%% compose multiple such policies.)
+
+%% Finally, we extend the model to a coroutine system featuring finitely-bounded stacks.
+%% Though simplistic, this extension is a step toward concurrency and demonstrates that
+%% the model is flexible enough to handle more sophisticated styles of control-flow.
+
+%% After formally defining stack safety, we show that the definition is
+%% enforceable using an existing tag-based enforcement
+%% mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
+%% \citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy.
+%% This policy testably enforces our {\em eager} properties, in which integrity is respected
+%% at every step of a program, and each call respects integrity at each of its internal steps.
+%% These properties are very strong, and our policies enforce them; this also makes them very
+%% intensional, in that they do not distinguish internal behavior from external.
+%% They also have the advantage that they are suitable for efficient testing, as they
+%% fail immediately when a violation occurs.
+
+%% The downside of the Depth Isolation policy is that it is rather expensive to implement.
+%% In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
+%% that is more efficient by virtue of being lazy in its enforcement.
+%% Instead of enforcing the integrity and confidentiality of the stack at the
+%% exact moment when a callee returns, this policy signals violations when the
+%% caller actually accesses data that the callee has written into its stack
+%% frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
+%%   property in the same way...}\fi
+%% The lazy policy does not enforce our eager integrity property, and to our
+%% knowledge there is no formal statement of the properties that it ought to enforce.
+%% We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
+%% policy that detects integrity violations between when they occur and when they make
+%% a visible impact. Intuitively this can be phrased as ``no visible reads after dangerous
+%% writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
+%% ``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
+%% testing, but formalizing the ideal property allows us to see how it falls short.
+%% We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
+%% at some performance cost.
+%% %We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
+%% %  and is it so interesting anyhow?}
+%% We argue that any system purporting to enforce stack safety ought to obey at least lazy
+%% stack safety, making it a plausible extensional characterization.
+
+%% %\bcp{I worry that we are being a little
+%% %  dishonest here: the stated advantage of the lazy policy is that it is more
+%% %efficient, but our variant is {\em not} efficient!}
 
@@
   We apply a property-based randomized testing framework to provide evidence that 
-  this stepwise property is enforced by the Depth Isolation policy of
-  \citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:conservative}).
+  this stepwise property is enforced by the Depth Isolation policy
+  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
 \item 
@@
   stack safety property, and argue that it accurately characterizes 
-  the Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18},
+  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18},
   when enhanced with unique tags on each activation~(\cref{sec:lazy}).

### 2021-04-10 tex/main.tex
Redid section 3

--- 2021-04-10 tex/main.tex
@@
 
-\begin{figure}[b]
+\begin{figure}
   \centering
@@
 \begin{verbatim}
-int stash;
+int stash = 0;
 
 int main() {
-0:             int x = 42, y = 0;
-1[P(x)],2[C]:  f(&x);
-3:             x = 0;
-4[P(y)],5[C]:  f(&y);
-6[R]:          return 0;
+0:         int x, y;
+1[P(x)C]:  f(x);
+2:         x = 0;
+3[P(y)C]:  f(y);
+4[R]:      return x;
 }
 
-void f(int* a) {
+void f(int &a) {
 10:    if(!stash)
-11:    stash = a;
+11:      stash = &a;
 12:    *stash = 5;
-13:  return;
+13[R]:   return;
 }
@@
     \begin{tabular}{l l}
-      {\tt 2} &
-      \memoryaddrs{12em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
-      \memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
-      ~$\cdots$ \\
       {\tt 10} &
@@
       \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{2}{\unsealc}
+      \memory{1}{\retptrc}
+      \memory{1}{\unsealc}
       ~$\cdots$ \\
+      {\tt 11} &
+      \memoryaddrs{17em}
+      \memory{1}{\mainpassc}
+      \memory{1}{\mainsealc}
+      \memory{1}{\retptrc}
+      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{5}
+      \\
       {\tt 3} &
       \memoryaddrs{12em}
-      \memory{3}{\unsealc}
-      ~$\cdots$ \\
-      {\tt 5} &
-      \memoryaddrs{12em}
+      \memory{4}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 10} &
+      \memoryaddrs{17em}
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
+      \memory{1}{\retptrc}
       \memory{1}{\unsealc}
-      \memory{1}{\mainpassc}
-      \memory{1}{\unsealc}
-      ~$\cdots$ \\
-      {\tt 10} &
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 11} &
       \memoryaddrs{17em}
@@
       \memory{1}{\mainpassc}
-      \memory{2}{\unsealc}
-      ~$\cdots$ \\
+      \memory{1}{\retptrc}
+      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{\bf 5}
 \end{tabular}
@@
 
-Up til now we have assumed that function arguments and returns are passed solely in registers.
-This is an oversimplication, which we will now correct. There are important stack safety concerns
-when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
-{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
-{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
-\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
-magenta. \apt{Why are there two line numbers assoicated with the two calls?}
-Then at the call, the parts of the caller's frame that are not passed are sealed, as usual. Passed
-memory also tracks which depth owns it, and it may only be accessed by the immediate callee
-(at its depth plus one).
-
-\apt{This needs to be talked through in more detail.}
-But because {\tt f} stashes the original pointer to {\tt x}, it will overwrite {\tt x} while
-{\tt x} is sealed, an integrity violation. This fits with the intuition of passing by reference,
-that the passed variable might be read or written during the call, not that it is permanently
-open to be read or written by all subsequent calls. This also applies to passing large objects
-by value in the stack: the caller allocates them, and the relevant instructions are annotated to
-indicate passing.
-
-We do not model cases where permanent sharing is desireable, because they are uninteresting
-and hard to motivate: we would expect to treat such sharing as completely unprotected, and
-perhaps attach a separate memory safety policy to add some protection to such variables.
+Up til now we have ignored function arguments, since our simplest model places them in
+registers. This is an oversimplication, which we will now correct, starting with the more
+interesting case: call-by-reference. There are important stack safety concerns
+when passing arguments by reference, as seen in example \ref{fig:passing}. In this example,
+{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
+passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
+{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
+{\tt x} in the second call.
+
+At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying the memory to be
+passed, and our stack model marks it in magenta. Then the caller's frame that are not
+passed are sealed, as usual. Just as we marked sealed memory with the depth of its owner,
+we do the same to passed memory. We keep track of the current depth of the active function
+next to the stack pointer in the diagram. A passed address may only be accessed when the
+current depth is one higher than the one it is sealed with.
+
+So at the initial call, {\tt x} is passed, and {\tt f} is free to write to it. Then upon
+return, {\tt main}'s frame is unsealed. At the second call it is {\tt y} that is passed,
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
+a future call with no arguments passed by reference would also violate integrity if it wrote
+to either variable.
+
+Passed addresses are also excluded from confidentiality: at the start of each call, we vary
+only stack addresses that are not passed. This is important, as we use the passing label
+for those arguments that are passed by value on the stack, as well. We might instead
+treat them as unsealed, but then we would treat them as confidential like the rest of the stack.
+
+We do not model permanent sharing, in which a variable's address is taken and it might be
+passed far afield before use, because this situation is uninteresting and hard to motivate
+as an instance of stack safety. We would expect to treat such sharing as completely
+unprotected, and perhaps attach a separate memory safety policy to add some protection
+to such variables.
 

### 2021-04-10 tex/main.tex

--- 2021-04-10 tex/main.tex
@@
 
-It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt w}
-and {\tt z}. We will demonstrate by amending our toy model to make it closer to a real stack.
-In figure \ref{fig:conf2} the topmost stack frame is no longer allocated as needed, but is an
-infinite array of addresses. Variables are allocated and deallocated by moving the stack pointer,
-and on a call, addresses to the left of the stack pointer are split off and pushed onto the
-stack as sealed frames. On a return, the topmost sealed frame is prepended to the unsealed
-frame. Now we can see how secrets might be left behind even in the unsealed portion of the stack.
-
-\sna{Updated up to here.}
-
-Two important details of confidentiality are reflected in figure \ref{fig:conf2}.
-First of all, it has multiple calls, and while we focus on the stack at the call
-from line 13 to line 20, it should be clear that we also need to consider a similar
-variant at the call from 0 to 10. Confidentiality is in this sense recursive: each
-nested call must obey confidentiality in its own right.
-
-More importantly, careful readers will have noticed that the informal definition
-of stack confidentiality refers to a variation of the entire stack, not just the
-sealed portion. In fact, we support temporal stack safety, in which values left behind
-by previous activations are also inaccessible. In this case, we connect temporal safety
-firmly to the caller-callee attacker model -- {\tt f} stored some data in the variables
-{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.\apt{Where is the storage for x ?.
-And need to explain what the colors mean.}
-After the stack pointer is written over {\tt y}, the value of {\tt z} is still
-left behind. Then {\tt g} can read it just by allocating and accessing an uninitialized
-variable! \apt{Need to be more explicit about what is getting varied, when, and why.}
-
-\paragraph*{Control Flow and Awkward Examples}
-
-\begin{figure}
-  \centering
-  \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-int stash = 0;
-
-int main() {
-0:     int x = 1;
-1[C]:  f();
-2:     x = -x;
-3[C]:  f();
-4[R]:  return x;
-}
-\end{verbatim}
-  \end{minipage}
-  \begin{minipage}{.4\textwidth}
-\begin{verbatim}
-void f() {
-10:     if(!stash) {
-        // store return pointer
-11:       stash = $SP-1;
-12:     } else {
-13:       $SP-1 = stash;
-14:       stash = 0;
-15[R]:    return;
-        }
-}
-\end{verbatim}
-  \end{minipage}
-\label{fig:WBCF1}
+It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt y}
+and {\tt z}. In figure \ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
+that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
+{\tt y} without even directly using its ability to read the stack pointer. In
+this case, {\tt g} branches on that value, which is a different way of
+violating lockstep confidentiality. Technically, this violation
+occurs because the program counter is also a component of the state that must change
+identically between the primary and variant traces. In a case where \(v_1 \not = 5\),
+the program counters will differ, breaking the lockstep.
+
+\sna{Not terribly happy with the circles yet; the idea is that instead of a meta-level
+  label for the line numbers we could treat the PC as a part of the system we're showing.}
+
+%\paragraph*{Control Flow and Awkward Examples}
+
+%\begin{figure}
+%  \centering
+%  \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%int stash = 0;
+
+%int main() {
+%0:     int x = 1;
+%1[C]:  f();
+%2:     x = -x;
+%3[C]:  f();
+%4[R]:  return x;
+%}
+%\end{verbatim}
+%  \end{minipage}
+%  \begin{minipage}{.4\textwidth}
+%\begin{verbatim}
+%void f() {
+%10:     if(!stash) {
+%        // store return pointer
+%11:       stash = $SP-1;
+%12:     } else {
+%13:       $SP-1 = stash;
+%14:       stash = 0;
+%15[R]:    return;
+%        }
+%}
+%\end{verbatim}
+%  \end{minipage}
+%\label{fig:WBCF1}
 %  \begin{subfigure}[t]{.4\textwidth}
@@
 \newcommand{\mainpassc}{magenta}
-\newcommand{\sharec}{lgray}
-
-\begin{figure}[b]
+
+\begin{figure}
   \centering
@@
 
-Until now we have assumed that function arguments and returns are passed solely in registers.
-This is an oversimplication, which we will now correct. There are important stack safety concerns
-when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
-{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
-{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
-\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
-magenta. \apt{Why are there two line numbers assoicated with the two calls?}
-Then at the call, the parts of the caller's frame that are not passed are sealed, as usual. Passed
-memory also tracks which depth owns it, and it may only be accessed by the immediate callee
-(at its depth plus one).
-
-\apt{This needs to be talked through in more detail.}
-But because {\tt f} stashes the original pointer to {\tt x}, it will overwrite {\tt x} while
-{\tt x} is sealed, an integrity violation. This fits with the intuition of passing by reference,
-that the passed variable might be read or written during the call, not that it is permanently
-open to be read or written by all subsequent calls. This also applies to passing large objects
-by value in the stack: the caller allocates them, and the relevant instructions are annotated to
-indicate passing.
-
-We do not model cases where permanent sharing is desireable, because they are uninteresting
-and hard to motivate: we would expect to treat such sharing as completely unprotected, and
-perhaps attach a separate memory safety policy to add some protection to such variables.
+Up til now we have ignored function arguments, since our simplest model places them in
+registers. This is an oversimplication, which we will now correct, starting with the more
+interesting case: call-by-reference. There are important stack safety concerns
+when passing arguments by reference, as seen in example \ref{fig:passing}. In this example,
+{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
+passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
+{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
+{\tt x} in the second call.
+
+At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying the memory to be
+passed, and our stack model marks it in magenta. Then the caller's frame that are not
+passed are sealed, as usual. Just as we marked sealed memory with the depth of its owner,
+we do the same to passed memory. We keep track of the current depth of the active function
+next to the stack pointer in the diagram. A passed address may only be accessed when the
+current depth is one higher than the one it is sealed with.
+
+So at the initial call, {\tt x} is passed, and {\tt f} is free to write to it. Then upon
+return, {\tt main}'s frame is unsealed. At the second call it is {\tt y} that is passed,
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
+a future call with no arguments passed by reference would also violate integrity if it wrote
+to either variable.
+
+Passed addresses are also excluded from confidentiality: at the start of each call, we vary
+only stack addresses that are not passed. This is important, as we use the passing label
+for those arguments that are passed by value on the stack, as well. We might instead
+treat them as unsealed, but then we would treat them as confidential like the rest of the stack.
+
+We do not model permanent sharing, in which a variable's address is taken and it might be
+passed far afield before use, because this situation is uninteresting and hard to motivate
+as an instance of stack safety. We would expect to treat such sharing as completely
+unprotected, and perhaps attach a separate memory safety policy to add some protection
+to such variables.
 

### 2021-04-10 tex/main.tex
Rewrite abstract; work on intro

@@
 \begin{abstract}
-\apt{maybe: ``Stack safety is a testable security property'' ?}a
-  What is stack safety? The phrase has been used to describe the protections
-  offered by a variety of mechanisms for guarding against
-  attacks that corrupt the call stacks of running programs, including micro-policies
-  implemented using metadata tagging mechanisms such as PIPE. But for specification
-  these implementations largely appeal either to intuitive notions of the behavior
-  of a stack, or to specific examples of bad behavior.
-  %
-  We propose a formal definition phrased in the technical terminology of language-based
-  security---as a combination of an integrity property (a caller's stack frame is
-  invariant over calls) and a confidentiality property (callees lack knowledge of the
-  contexts of their calls). These combine with a straightforward well-bracketedness
-  condition on the program's control flow that ensures that callers and callees remain
-  meaningfully distinct.
-  %
-  While our formalism is informed by the PIPE implementation, the primary properties reflect
-  general principles of stack safety. We define additional properties tailored for specific
-  contexts: a stronger property amenable to efficient testing and a weaker one that is
-  approximated by more efficient, lazy policies.
+\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
+  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
+  to leave it out of the title but hit it strongly in the abstract.}%
+What is stack safety? The phrase is associated with a variety of compiler,
+run-time, and hardware mechanisms for guarding against corruption of stack
+memory.  But these mechanisms mostly lack rigorous specifications, appealing
+either to intuitive expectations of how good stack-using programs should
+behave or to specific examples of bad behavior.
+
+We propose a formal criterion for stack safety, phrased in terms of
+language-based security: a combination of an integrity property (``each
+caller's stack frame is invariant by the callee'') and a confidentiality
+property (``the callee's behavior is insensitive to differences in the
+caller's stack frame''). These combine with a straightforward
+well-bracketedness condition on the program's control flow that ensures that
+callers and callees remain meaningfully distinct.\bcp{I am still not clear
+  how well-bracketedness fits into the story.  (For me, it is the bracketing
+  that {\em defines} where calls and returns happen, so there is something
+  tautological about saying that `` well-bracketedness condition ensures
+  that callers and callees remain meaningfully distinct''!)}
+
+Our criterion is inspired by hardware tagging mechanisms for stack
+protection, in particular the stack-safety ``micro-policies''
+of~\citet{DBLP:conf/sp/RoesslerD18}, but its core reflects general
+principles of stack safety\bcp{I wish we did not have to be so apologetic
+  about being tied to micro-policies (and it is still not clear to me that
+  we have to be... do we have a clear statement someplace of why our
+  definition would not apply to CFI, for example? This seems like a critical
+  discussion...)}. We define additional properties tailored for specific
+contexts: a stronger property amenable to efficient testing and a weaker one
+that is approximated by more efficient, lazy policies.
+
+\bcp{... and now a sentence on testing!}
 
@@
   \leo{Not sure. We almost never cite stuff in the abstract, do we?}
-  \sna{On the same note, cite Nick and Andre?}
+  \bcp{Some people say not to, but I think their argument is weak: I've
+    added Nick/Andre above.  But I'm not sure we want to get into stkTokens
+    in the abstract.}
   \fi
@@
 
-The call stack is a perennial target for attacks
-based on memory-safety vulnerabilities, with potential consequences
+The call stack is a perennial target for low-level attacks
+based on memory-safety vulnerabilities, with consequences
 ranging from leakage or corruption of private stack data to
 control-flow hijacking. To detect or prevent such attacks, a menagerie of
-software and hardware techniques for stack safety have been proposed,
+software and hardware protections have been proposed,
 %
@@
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\leo{Do we want to hunt for more refs here?}
-\bcp{Sure, but later!}
+  nick; there could be more}
 \fi
 %
-The implementation and configuration of these mechanisms can be complex, 
-and often involves making compromises that trade off protection for efficiency.  
-Thus it would be very useful to have precise formal specifications for stack safety
-that give us a basis for comparing the security guarantees claimed by different enforcement
-techniques, and for verifying that these claims are met by particular implementations.
+The implementation and configuration of these  stack-safety mechanisms can be complex,
+and often involves compromises trading protection for efficiency.
+Thus it would be useful to have a precise formal specification for stack
+safety, both as a basis for comparing the security guarantees claimed by different enforcement
+techniques and for verifying that these claims are met by particular implementations.
 
@@
 validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces them.
-\apt{more to say?}
+\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
+  significant contribution, and partly to give the reader some context for
+  why this is part of our story.  E.g., ``One nice thing we can do with our
+  spec is to use it to evaluate the claims of existing or future proposals
+  for enforcement mechanisms.  One relatively inexpensive way to carry out
+  such an evaluation is to see whether property-based testing against this
+  spec can distinguish correct implementations of the mechanism from buggy
+  ones...''}
 
@@
 \item
-  We harness the language-based security concepts of integrity and confidentiality 
-  to formalize a parameterized {\em stepwise stack-safety property}, and argue that
-  it captures the intuitive meaning of local stack encapsulation, well-bracketed control
+  We adapt formulations of integrity and confidentiality from
+  language-based security
+  to formalize a parameterized {\em stepwise stack-safety property} and argue that
+  it captures the intuitive concepts of local stack encapsulation, well-bracketed control
   flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
@@
   We formalize a weaker but more extensional \emph{observational} variant of the
-  stack safety property, and argue that it accurately characterizes 
-  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18},
-  when enhanced with unique tags on each activation~(\cref{sec:lazy}).
+  stack safety property and argue that it accurately characterizes
+  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18}
+  enhanced with unique tags on each activation~(\cref{sec:lazy}).
 \item
   We show how our models can be instantiated and extended to handle passing
-  arguments on the stack~(\cref{sec:??}) and to support simple coroutining~(\cref{sec:??}).
+  arguments on the stack~(\cref{sec:??}) and simple
+  coroutines~(\cref{sec:??}).
 %  This policy does not enforce our strict

### 2021-04-10 tex/main.tex
Couple more things

@@
 
-In summary, we offer the following contributions:
+In summary, we offer the following contributions: \bcp{Section numbers are
+  wonky---why just 7.1, where is 8, etc.}
 \begin{itemize}
@@
 \item
-  We apply a property-based randomized testing framework to provide evidence that
-  this stepwise property is enforced by the Depth Isolation policy
+  We use a property-based randomized testing framework to support the claim
+  that this stepwise property is actually enforced by the Depth Isolation policy
   of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
@@
   arguments on the stack~(\cref{sec:??}) and simple
-  coroutines~(\cref{sec:??}).
+  coroutines~(\cref{sec:??}).\bcp{``instantiated and extended'' is awkward.
+  Maybe ``We show how our models can be instantiated to handle passing
+  arguments on the stack~(\cref{sec:??}) and extended with simple
+  coroutines~(\cref{sec:??}).''  If that's what we mean.}
 %  This policy does not enforce our strict
@@
 \end{itemize}
-\Cref{sec:relwork} discusses related work;
+\Cref{sec:relwork} discusses related work, and 
 \cref{sec:future} sketches directions for future work.

### 2021-04-10 tex/main.tex
Couple of comments
Responded to some abstract comments

@@
 property (``the callee's behavior is insensitive to differences in the
-caller's stack frame''). These combine with a straightforward
-well-bracketedness condition on the program's control flow that ensures that
-callers and callees remain meaningfully distinct.\bcp{I am still not clear
+caller's state''). These are defined in terms of a control flow model
+in which callees are meaningfully distinct from their callers due to being
+bracketed by calls and returns. We further apply a straightforward notion of
+``well-bracketed control flow'' that ensures that callers' control flow
+can continue predictably after the return.
+
+  \bcp{I am still not clear
   how well-bracketedness fits into the story.  (For me, it is the bracketing
@@
   that callers and callees remain meaningfully distinct''!)}
+  \sna{Attempted to resolve this, it may need some work.}
 
@@
   definition would not apply to CFI, for example? This seems like a critical
-  discussion...)}. We define additional properties tailored for specific
-contexts: a stronger property amenable to efficient testing and a weaker one
-that is approximated by more efficient, lazy policies.
-
-\bcp{... and now a sentence on testing!}
+  discussion...)}. Finally, we show that micro-policies do indeed enforce
+our properties using property-based randomized testing. An eager micropolicy
+that catches violations immediately enforces our stronger property, and we
+define a weaker, more extensional form of the property that is enforced
+by a ``lazy'' micropolicy -- one that, for performance reasons, may permit
+local violations as long as they do not influence future behavior.
 
@@
 systems with minor adjustment.\apt{move this up?}
-
-%% %\begin{figure}[b]
-%% %    \begin{minipage}{.4\textwidth}
-%% %\begin{verbatim}
-%% %int main() {
-%% %  int x = 42;
-%% %  f();
-%% %  print(x);
-%% %}
-%% %\end{verbatim}
-%% %    \end{minipage}
-%% %    \begin{minipage}{.4\textwidth}
-%% %\begin{verbatim}
-%% %int main() {
-%% %  int x = 1;
-%% %  f();
-%% %  if(x)
-%% %    g();
-%% %  else
-%% %    h();
-%% %}
-%% %\end{verbatim}
-%% %    \end{minipage}
-%% %  \caption{Simple Examples}
-%% %  \label{fig:examples1}
-%% %\end{figure}
-
-%% %\begin{figure}[b]
-%% %    \begin{minipage}{.4\textwidth}
-%% %\begin{verbatim}
-%% %int main(int arg) {
-%% %  if(arg)
-%% %    g();
-%% %  else
-%% %    h();
-%% %  f();
-%% %}
-%% %\end{verbatim}
-%% %    \end{minipage}
-%% %    \begin{minipage}{.4\textwidth}
-%% %\begin{verbatim}
-%% %int main() {
-%% %  int x = 1;
-%% %  f();
-%% %  x = -x;
-%% %  f();
-%% %  print(x);
-%% %}
-%% %\end{verbatim}
-%% %  \end{minipage}
-%% %
-%% %  \caption{More Examples}
-%% %  \label{fig:examples2}
-%% %\end{figure}
 

### 2021-04-10 tex/main.tex
Polishing and comments

@@
 The call stack is a perennial target for low-level attacks
-based on memory-safety vulnerabilities, with consequences
-ranging from leakage or corruption of private stack data to
-control-flow hijacking. To detect or prevent such attacks, a menagerie of
+based on memory-safety vulnerabilities, leading to consequences
+from leakage or corruption of private stack data to
+control-flow hijacking. To detect or prevent such attacks, a profusion of
 software and hardware protections have been proposed,
@@
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
-  nick; there could be more}
+  nick; there could be more}\bcp{Yes, going back to MIT days---we should
+  include several more of these, if only to give readers the impression that
+this is a well-studied mechanism (so formalizing its protections is
+useful).} 
 \fi
 %
-The implementation and configuration of these  stack-safety mechanisms can be complex,
-and often involves compromises trading protection for efficiency.
-Thus it would be useful to have a precise formal specification for stack
-safety, both as a basis for comparing the security guarantees claimed by different enforcement
-techniques and for verifying that these claims are met by particular implementations.
+The protections offered by such mechanisms are generally described in terms
+of a set of concrete examples of attacks that they can prevent, leaving a
+more general characterization to the reader's intuition.
+Moreover, the details of these mechanisms can be quite intricate
+and often involve trading off protection and efficiency.
+Thus, it would be useful to have a precise, generic, and formal
+specification for stack
+safety, both as a basis for comparing the security claims of different
+enforcement techniques and for validating more rigorously that these claims
+are met by particular implementations.
 
@@
 using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
-In essence, memory locations within each stack frame are tagged with the identity of
+Intuitively, memory cells within each stack frame are tagged with the
+identity of
 the function that owns the frame, and access to those locations is then gated
-according to the function that is currently executing. Many alternatives can
-be implemented on top of this basic idea, by choosing how to identify functions
-and deciding just which accesses are checked.  These alternatives vary in their
-efficiency, but also in the spatial and temporal precision of their security guarantees.
-For example, the \emph{Depth Isolation} policy strictly requires that a function owns any
+according to the function that is currently executing. Roessler and Dehon
+explored a number of versions
+of this basic idea, with different ways of identifying
+functions and deciding exactly which accesses are allowed.  These alternatives
+vary in their
+efficiency and in the spatial and temporal precision of their security guarantees.
+For example, their \emph{Depth Isolation} policy demands that a function
+owns any
 frame into which it writes; this seems simple and intuitively correct, but can be expensive to implement.
-On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by
-allowing illicit writes to the frame so long as the written locations are not
-subsequently read. The intuition justifying laziness is that any genuinely dangerous writes will ultimately be
-detected before they can affect program behavior.  But we clearly need a more formal foundation for
+Conversely, their \emph{Lazy Tagging and Clearing} policy improves
+performance by
+allowing ``illicit'' writes to out-of-scope stack frames so long as the
+written locations are not
+subsequently read. The justification for this form of laziness is that any
+genuinely dangerous writes will ultimately be
+detected before they can affect program behavior. \bcp{Awkward transition here.} But we clearly need a
+more formal foundation for
 expressing, comparing and verifying these micro-policy security claims, as well as those of
@@
 at some performance cost.\apt{but we're not testing that, right?}
+\rb{Not sure about the property part right now, the machine side of this is
+pretty much done}
 
@@
 compose multiple such policies.)
+\rb{So no pointer passing? Pass by reference is now discussed at the end of
+Section 3, however it looks like things get tricky rather quickly.}
 

### 2021-04-10 tex/main.tex
Comments and polishing

@@
 
-For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
+\bcp{this paragraph seems a bit in the way at the moment: the previous
+  paragraph has already motivated the need for a formal characterization, so
+  why not just go straight on and talk about ours?  I propose moving it to,
+  say, just before ``To illustrate...''  We can also then drop the awkward
+  last sentence.}
+}For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
 using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
@@
 according to the function that is currently executing. Roessler and Dehon
-explored a number of versions
+explore alternative versions
 of this basic idea, with different ways of identifying
-functions and deciding exactly which accesses are allowed.  These alternatives
-vary in their
-efficiency and in the spatial and temporal precision of their security guarantees.
+functions and deciding exactly which accesses are allowed, varying in their
+efficiency and in the spatial and temporal precision of their security
+guarantees.
 For example, their \emph{Depth Isolation} policy demands that a function
@@
 performance by
-allowing ``illicit'' writes to out-of-scope stack frames so long as the
+allowing ``illicit'' writes to inaccessible stack frames so long as the
 written locations are not
@@
 
-In this work, we take the (informal) definitions of WBCF and LSE as a starting point,
-but choose a different approach to formalizing them, based on the
+\bcp{If we end up adopting my recent proposal for defining subtraces and
+  separating WBCF, I think we can be a bit stronger here: We define stack
+  safety in IFC terminology based on the same intuitions as their LSE, but
+  we consider (and define) WBCF to be a separate thing---desirable, but
+  distinct from safety.  (If we make this argument, we'll need to show an
+  example of a program that is stack safe but doesn't have WBCF.)}In this
+work, we take the (informal) definitions of WBCF and LSE as a starting
+point, but choose a different approach to formalizing them, based on the
 technical framework of language-based security~\citep{??}.  We define

### 2021-04-10 tex/main.tex
One morec

@@
 trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
-In applying these security concepts, we treat each caller-callee pair as mutually distrustful
-and view the caller as needing protection from the callee. \apt{Need more explanation of
+In applying these security concepts, we treat each caller-callee pair as
+mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
   ``nesting'' and our novelty factor here? Also about domains??}
-{\em Integrity} says that a caller's local data is protected from modification by its callee.
+{\em Integrity} says that a caller's local data is protected from
+modification by its callee.\bcp{What about the ``stepwise'' part?}
 This is vital for local reasoning about the behavior of the caller.

### 2021-04-11 tex/main.tex
Passed over intro

--- 2021-04-10 tex/main.tex
@@
 The call stack is a perennial target for low-level attacks
-based on memory-safety vulnerabilities, leading to consequences
-from leakage or corruption of private stack data to
-control-flow hijacking. To detect or prevent such attacks, a profusion of
+based on memory-safety vulnerabilities, with consequences
+ranging from leakage or corruption of private stack data to
+control-flow hijacking. To detect or prevent such attacks, a menagerie of
 software and hardware protections have been proposed,
@@
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\bcp{Yes, going back to MIT days---we should
-  include several more of these, if only to give readers the impression that
-this is a well-studied mechanism (so formalizing its protections is
-useful).} 
+  nick; there could be more}
 \fi
 %
-The protections offered by such mechanisms are generally described in terms
-of a set of concrete examples of attacks that they can prevent, leaving a
-more general characterization to the reader's intuition.
-Moreover, the details of these mechanisms can be quite intricate
-and often involve trading off protection and efficiency.
-Thus, it would be useful to have a precise, generic, and formal
-specification for stack
-safety, both as a basis for comparing the security claims of different
-enforcement techniques and for validating more rigorously that these claims
-are met by particular implementations.
+The implementation and configuration of these  stack-safety mechanisms can be complex,
+and often involves compromises trading protection for efficiency.
+Thus it would be useful to have a precise formal specification for stack
+safety, both as a basis for comparing the security guarantees claimed by different enforcement
+techniques and for verifying that these claims are met by particular implementations.
 
@@
 using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
-Intuitively, memory cells within each stack frame are tagged with the
-identity of
-the function that owns the frame, and access to those locations is then gated
-according to the function that is currently executing. Roessler and Dehon
-explored a number of versions
-of this basic idea, with different ways of identifying
-functions and deciding exactly which accesses are allowed.  These alternatives
-vary in their
-efficiency and in the spatial and temporal precision of their security guarantees.
-For example, their \emph{Depth Isolation} policy demands that a function
-owns any
-frame into which it writes; this seems simple and intuitively correct, but can be expensive to implement.
-Conversely, their \emph{Lazy Tagging and Clearing} policy improves
-performance by
-allowing ``illicit'' writes to out-of-scope stack frames so long as the
-written locations are not
-subsequently read. The justification for this form of laziness is that any
-genuinely dangerous writes will ultimately be
-detected before they can affect program behavior. \bcp{Awkward transition here.} But we clearly need a
-more formal foundation for
+In essence, memory locations within each stack frame are tagged with the identity of
+the function activation that owns the frame, and access to those locations is then gated
+according to the activation that is currently executing. Many alternatives can
+be implemented on top of this basic idea, by choosing how to identify activations
+and how to maintain the connection between an activation identity and the memory it owns.
+These alternatives vary in their
+efficiency, but also in the spatial and temporal precision of their security guarantees.
+For example, the \emph{Depth Isolation} policy requires a strict mapping from a function
+to its frame from the start of the call to its end, and permits the function to write
+into that frame. This seems simple and intuitively correct, but it can be expensive to
+implement that strict mapping.
+On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by
+assigning locations to the active function when they are written, which permits
+illicit writes outside the function's frame so long as the proper owner of the written
+locations does not subsequently read them. The intuition justifying laziness is that any
+genuinely dangerous writes will ultimately be detected before they can affect program
+behavior.  But we clearly need a more formal foundation for
 expressing, comparing and verifying these micro-policy security claims, as well as those of
@@
 
-So how should we formally define stack safety? Definitions in the literature are commonly informal
-and proceed by enumerating the specific bad things that a given enforcement mechanism
+So how should we formally define stack safety? Definitions in the literature are commonly
+informal and proceed by enumerating the specific bad things that a given enforcement mechanism
 prevents, such as corruption of return addresses, buffer overflows, or
@@
 The most abstract characterization of which we are aware is in
-recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
+recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
+properties:
 {\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
@@
 \citet{Skorstengaard+19} formalize these properties by defining an idealized machine
-in which they visibly hold by construction; an enforcement mechanism (in their case, hardware capabilities)
+in which they visibly hold by construction; an enforcement mechanism (in their case, hardware
+capabilities)
 can then be shown to provide stack safety by giving a fully abstract embedding function
@@
 but choose a different approach to formalizing them, based on the
-technical framework of language-based security~\citep{??}.  We define
+technical framework of language-based security~\citep{??}.
+While our property definitions were developed with micro-policy enforcement in mind, they
+agnostic about many implementation details and should be applicable to other
+systems with minor adjustment. We define
 trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
@@
   ``nesting'' and our novelty factor here? Also about domains??}
+This differs from common instances in language-based security in that the
+confidentiality property applies to many nested subtraces within a single execution.
+\sna{Thinking about how to express domains, or whether to. I'm not sure it's
+  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
 {\em Integrity} says that a caller's local data is protected from modification by its callee.
@@
 We also give a simple trace-based formalization of WBCF, which is essential for keeping the
-notions of ``caller' and ``callee'' sensible.
+notions of ``caller'' and ``callee'' sensible.
 
 To illustrate the applicability of these properties, we perform randomized testing to
-validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces them.
-\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
-  significant contribution, and partly to give the reader some context for
-  why this is part of our story.  E.g., ``One nice thing we can do with our
-  spec is to use it to evaluate the claims of existing or future proposals
-  for enforcement mechanisms.  One relatively inexpensive way to carry out
-  such an evaluation is to see whether property-based testing against this
-  spec can distinguish correct implementations of the mechanism from buggy
-  ones...''}
-
-However, their Lazy Tagging and Clearing micro-policy does not enforce stepwise integrity.
+validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
+them. Randomized testing is a relatively low-cost way to validate a potential enforcement
+mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
+of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
+covers potentially erroneous programs, we test intentionally buggy policies alongside
+the real one. If the testing can differentiate correct implementations from buggy ones,
+it is an argument in favor of their correctness.
+%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
+%  significant contribution, and partly to give the reader some context for
+%  why this is part of our story.  E.g., ``One nice thing we can do with our
+%  spec is to use it to evaluate the claims of existing or future proposals
+%  for enforcement mechanisms.  One relatively inexpensive way to carry out
+%  such an evaluation is to see whether property-based testing against this
+%  spec can distinguish correct implementations of the mechanism from buggy
+%  ones...''}
+
+However, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
+Clearing micro-policy does not enforce stepwise integrity.
 To address this, we develop an \emph{observational stack integrity} property that
@@
 detect integrity violations between the point when they occur and the point when they make
-an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous
-writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.
+an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.
+\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
+  established that this was important.}
 
@@
 testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
-We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce observational integrity,
-at some performance cost.\apt{but we're not testing that, right?}
+We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
+observational integrity, at some performance cost.\apt{but we're not testing that, right?}
 \rb{Not sure about the property part right now, the machine side of this is
   pretty much done}
+\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
+  need a caveat about how it's harder to test?}
 
@@
 
-Our initial description of the properties assumes single simple stack with no sharing between callers and
-callees.  We expand it to allow passing of scalar stack data, granting a callee
+Our initial description of the properties assumes single simple stack with no sharing between
+callers and callees.  We expand it to allow passing of scalar stack data, granting a callee
 access to the caller's data. We do not model more sophisticated sharing, because to do
@@
   Section 3, however it looks like things get tricky rather quickly.}
-
-Finally, we define integrity and confidentiality properties for a coroutine system featuring finitely-bounded stacks.
+\sna{The trick is that if a programmer is passing pointers, we can't presume their
+  intent -- how long should we permit memory to be shared? But if the compiler is passing
+  pointers, as in pass by reference, we have a clear sense of where sharing should stop.}
+
+Finally, we define integrity and confidentiality properties for a coroutine system featuring
+finitely-bounded stacks.
 Though simplistic, this extension is a step toward concurrency and demonstrates that
 our approach is flexible enough to handle more sophisticated styles of control-flow.
-
-While our property definitions were developed with micro-policy enforcement in mind, they
-agnostic about many implementation details and should be applicable to other
-systems with minor adjustment.\apt{move this up?}
 
@@
 
-We begin with a description of our threat model
-(\cref{sec:threat}), a running example that we refer to throughout the paper
-(\cref{sec:running-example}),  and some technical preliminaries
-(\cref{sec:prelim}).\apt{this para needs updating}
+We begin with a description of our threat model (\cref{sec:threat}),
+then show how our properties apply to examples of stack unsafety 
+(\cref{sec:running-example}). Then, after some technical preliminaries
+(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
+and describe testing (\cref{sec:testing}).
 

### 2021-04-11 tex/main.tex
Last set of comments for today

--- 2021-04-10 tex/main.tex
@@
 prevents, such as corruption of return addresses, buffer overflows, or
-use of uninitialized variables.
+use of uninitialized variables.\bcp{Most of that sentence should be merged 
+  into the similar sentence above.  Doesn't belong here.}
 %\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
@@
   ``nesting'' and our novelty factor here? Also about domains??}
-{\em Integrity} says that a caller's local data is protected from
-modification by its callee.\bcp{What about the ``stepwise'' part?}
-This is vital for local reasoning about the behavior of the caller.
-{\em Confidentiality} states that a caller's data is protected
+{\em Integrity} says informally that a caller's local data is protected from
+modification by its callees.\bcp{What about the ``stepwise'' part?}
+This is vital for local reasoning about the behavior of the
+caller.\bcp{Either delete that sentence or explain it more.}
+       {\em Confidentiality} says that a caller's data is protected
 from being leaked into its callee, and also that the callee cannot see uninitialized data.
@@
 Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
-notion that data outside of live frames should be inaccessible (which LSE does not address).
+notion that data outside of live frames should be inaccessible (which LSE
+does not address\bcp{Which supports my feeling that we should not advertise
+  ourselves as reimplementing LSE but rather as defining a different
+  condition based on similar intuitions.}).
+\bcp{These are the intuitions, but we should also mention that we want to define
+  these (at least partly) in terms of a sort-of-standard notion of ``nested
+  noninterference''!  Then, at the end of the intro (just before the
+  contributions) we can summarize the whole thing like ``So what is stack
+  safety?  It is nested NI.''}
 We also give a simple trace-based formalization of WBCF, which is essential for keeping the
-notions of ``caller' and ``callee'' sensible.
+notions of ``caller' and ``callee'' sensible. \bcp{That last should probably
+be a separate discussion, if I've succeeded in convincing people that WBCF
+is not part of stack safety.}
 
@@
 Observational integrity is attractive in being significantly more \emph{extensional} than
-stepwise integrity.  We argue that it may therefore be a better minimal specification for
+stepwise integrity.  We therefore argue that, despite its somewhat greater complexity,
+it may be a better minimal specification for
 stack safety enforcement mechanisms in general.
 
-Our initial description of the properties assumes single simple stack with no sharing between callers and
-callees.  We expand it to allow passing of scalar stack data, granting a callee
-access to the caller's data. We do not model more sophisticated sharing, because to do
+Our initial description of these properties assumes a single simple stack with no sharing between callers and
+callees: all parameters and return values are passed in registers.  However,
+we can refine it to allow passing of scalar stack data, granting a callee
+access to the caller's data. (We do not model more sophisticated sharing, because to do
 so realistically would require a separate protection scheme, such as a capability
-model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
-a separate policy outside the scope of stack safety. (A tag-based approach can easily
+model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
+a separate policy outside the scope of stack safety. A tag-based approach can easily
 compose multiple such policies.)
@@
 Though simplistic, this extension is a step toward concurrency and demonstrates that
-our approach is flexible enough to handle more sophisticated styles of control-flow.
-
-While our property definitions were developed with micro-policy enforcement in mind, they
-agnostic about many implementation details and should be applicable to other
-systems with minor adjustment.\apt{move this up?}
+our approach is flexible enough to handle more sophisticated styles of
+control flow.
+
+While our property definitions were developed with micro-policy enforcement
+in mind, they agnostic about many implementation details and should be
+applicable to other systems with minor adjustment.\apt{move this up?} \bcp{I
+  am still hoping to get a clearer understanding for myself of just what is
+  micropolicy-specific about our definitions...}
 
@@
 
-Stack safety is meant to protect a caller from a confused or malicious callee's
-attempts to access private data or hijack its control flow. And in fact the caller,
-while presumably not malicious, may itself be confused or erroneous. Our attacker
-model reflects this intention by assuming that both the attacker and the defender
+\bcp{Again, I think we need to put ``Assumptions'' (i.e., our basic picture
+  of the world) first, as context.  E.g., none of the threat model
+  discussion makes sense until the reader understands that we {\em define}
+  function calls and returns by means of annotations (and that we MUST do
+  something like that because we are talking about machine code).}
+
+Intuititively, stack safety means that a caller's stack frame is protected
+from a confused or malicious callee's
+attempts to read or write its private data or hijack its control flow. And in fact the caller,
+while presumably\bcp{not ``presumably'': {\em by definition}!} not malicious, may itself be confused or erroneous. Our attacker
+model reflects this intention\bcp{I don't understand how it reflects this
+  intuition about the caller---indeed, I don't understand the point about
+  the caller.} by assuming that both the attacker and the defender
 may run arbitrary machine code. Hardware and timing attacks are out of scope.

### 2021-04-11 tex/main.tex
Oops

@@
   last sentence.}
-}For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
+For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
 using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.

### 2021-04-11 tex/main.tex

--- 2021-04-11 tex/main.tex
@@
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
-  nick; there could be more}
+  nick; there could be more}\bcp{Yes, going back to MIT days---we should
+  include several more of these, if only to give readers the impression that
+this is a well-studied mechanism (so formalizing its protections is
+useful).}
 \fi
 %
-The implementation and configuration of these  stack-safety mechanisms can be complex,
-and often involves compromises trading protection for efficiency.
-Thus it would be useful to have a precise formal specification for stack
-safety, both as a basis for comparing the security guarantees claimed by different enforcement
-techniques and for verifying that these claims are met by particular implementations.
-
-For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
-using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
-In essence, memory locations within each stack frame are tagged with the identity of
-the function activation that owns the frame, and access to those locations is then gated
-according to the activation that is currently executing. Many alternatives can
-be implemented on top of this basic idea, by choosing how to identify activations
-and how to maintain the connection between an activation identity and the memory it owns.
-These alternatives vary in their
-efficiency, but also in the spatial and temporal precision of their security guarantees.
-For example, the \emph{Depth Isolation} policy requires a strict mapping from a function
-to its frame from the start of the call to its end, and permits the function to write
-into that frame. This seems simple and intuitively correct, but it can be expensive to
-implement that strict mapping.
-On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by
-assigning locations to the active function when they are written, which permits
-illicit writes outside the function's frame so long as the proper owner of the written
-locations does not subsequently read them. The intuition justifying laziness is that any
-genuinely dangerous writes will ultimately be detected before they can affect program
-behavior.  But we clearly need a more formal foundation for
-expressing, comparing and verifying these micro-policy security claims, as well as those of
-other enforcement technologies.
+The protections offered by such mechanisms are generally described in terms
+of a set of concrete examples of attacks that they can prevent, leaving a
+more general characterization to the reader's intuition.
+Moreover, the details of these mechanisms can be quite intricate
+and often involve trading off protection and efficiency.
+Thus, it would be useful to have a precise, generic, and formal
+specification for stack
+safety, both as a basis for comparing the security claims of different
+enforcement techniques and for validating more rigorously that these claims
+are met by particular implementations.
 
@@
 
-In this work, we take the (informal) definitions of WBCF and LSE as a starting point,
-but choose a different approach to formalizing them, based on the
-technical framework of language-based security~\citep{??}.
-While our property definitions were developed with micro-policy enforcement in mind, they
-agnostic about many implementation details and should be applicable to other
-systems with minor adjustment. We define
+\bcp{If we end up adopting my recent proposal for defining subtraces and
+  separating WBCF, I think we can be a bit stronger here: We define stack
+  safety in IFC terminology based on the same intuitions as their LSE, but
+  we consider (and define) WBCF to be a separate thing---desirable, but
+  distinct from safety.  (If we make this argument, we'll need to show an
+  example of a program that is stack safe but doesn't have WBCF.)}In this
+work, we take the (informal) definitions of WBCF and LSE as a starting
+point, but choose a different approach to formalizing them, based on the
+technical framework of language-based security~\citep{??}. We define
 trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
-In applying these security concepts, we treat each caller-callee pair as mutually distrustful
-and view the caller as needing protection from the callee. \apt{Need more explanation of
+In applying these security concepts, we treat each caller-callee pair as
+mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
   ``nesting'' and our novelty factor here? Also about domains??}
@@
 {\em Integrity} says that a caller's local data is protected from modification by its callee.
+\bcp{What about the ``stepwise'' part?}
 This is vital for local reasoning about the behavior of the caller.
@@
 notions of ``caller'' and ``callee'' sensible.
+
+For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
+using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
+Intuitively, memory cells within each stack frame are tagged with the
+identity of the function activation
+that owns the frame, and access to those locations is then gated
+according to the activation that is currently executing. Roessler and Dehon
+explore alternative versions of this basic idea, with different ways of identifying
+functions and deciding exactly which accesses are allowed, to enable different ways
+of maintaining the connection between an activation identity and the memory it owns.
+These vary in their efficiency and in the spatial and temporal precision of their security
+guarantees.
+For example, their \emph{Depth Isolation} policy  requires a strict mapping from a function
+to its frame from the start of the call to its end, and permits the function to write
+only into that frame.
+This seems simple and intuitively correct, but it can be expensive to implement the strict
+mapping. Conversely, their \emph{Lazy Tagging and Clearing} policy improves
+performance by assigning locations to the active function when they are written, which
+permits  ``illicit'' writes to inaccessible stack frames so long as the
+written locations are not subsequently read by their ``real'' owners. The justification
+for this form of laziness is that any genuinely dangerous writes will ultimately be
+detected before they can affect program behavior.
 

### 2021-04-11 tex/main.tex

@@
 prevents, such as corruption of return addresses, buffer overflows, or
-use of uninitialized variables.
+use of uninitialized variables.\bcp{Most of that sentence should be merged
+  into the similar sentence above.  Doesn't belong here.}
 %\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
@@
 \bcp{What about the ``stepwise'' part?}
-This is vital for local reasoning about the behavior of the caller.
-{\em Confidentiality} states that a caller's data is protected
+This is vital for local reasoning about the behavior of the
+caller.\bcp{Either delete that sentence or explain it more.}
+{\em Confidentiality} says that a caller's data is protected
 from being leaked into its callee, and also that the callee cannot see uninitialized data.
@@
 Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
-notion that data outside of live frames should be inaccessible (which LSE does not address).
+notion that data outside of live frames should be inaccessible (which LSE
+does not address\bcp{Which supports my feeling that we should not advertise
+  ourselves as reimplementing LSE but rather as defining a different
+  condition based on similar intuitions.}).
+\bcp{These are the intuitions, but we should also mention that we want to define
+  these (at least partly) in terms of a sort-of-standard notion of ``nested
+  noninterference''!  Then, at the end of the intro (just before the
+  contributions) we can summarize the whole thing like ``So what is stack
+  safety?  It is nested NI.''}
 We also give a simple trace-based formalization of WBCF, which is essential for keeping the
-notions of ``caller'' and ``callee'' sensible.
+notions of ``caller'' and ``callee'' sensible. \bcp{That last should probably
+be a separate discussion, if I've succeeded in convincing people that WBCF
+is not part of stack safety.}
 
@@
 Observational integrity is attractive in being significantly more \emph{extensional} than
-stepwise integrity.  We argue that it may therefore be a better minimal specification for
+stepwise integrity.  We therefore argue that, despite its somewhat greater complexity,
+it may be a better minimal specification for
 stack safety enforcement mechanisms in general.
 
-Our initial description of the properties assumes single simple stack with no sharing between
-callers and callees.  We expand it to allow passing of scalar stack data, granting a callee
-access to the caller's data. We do not model more sophisticated sharing, because to do
+Our initial description of these properties assumes a single simple stack with no sharing between callers and
+callees: all parameters and return values are passed in registers.  However,
+we can refine it to allow passing of scalar stack data, granting a callee
+access to the caller's data. (We do not model more sophisticated sharing, because to do
 so realistically would require a separate protection scheme, such as a capability
-model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
-a separate policy outside the scope of stack safety. (A tag-based approach can easily
+model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
+a separate policy outside the scope of stack safety. A tag-based approach can easily
 compose multiple such policies.)
@@
 Though simplistic, this extension is a step toward concurrency and demonstrates that
-our approach is flexible enough to handle more sophisticated styles of control-flow.
+our approach is flexible enough to handle more sophisticated styles of
+control flow.
+
+While our property definitions were developed with micro-policy enforcement
+in mind, they agnostic about many implementation details and should be
+applicable to other systems with minor adjustment.\apt{move this up?} \bcp{I
+  am still hoping to get a clearer understanding for myself of just what is
+  micropolicy-specific about our definitions...}
 

### 2021-04-11 tex/main.tex
More discussion of properties relating to the broader world

@@
   distinct from safety.  (If we make this argument, we'll need to show an
-  example of a program that is stack safe but doesn't have WBCF.)}In this
-work, we take the (informal) definitions of WBCF and LSE as a starting
+  example of a program that is stack safe but doesn't have WBCF.)}
+In this work, we take the (informal) definitions of WBCF and LSE as a starting
 point, but choose a different approach to formalizing them, based on the
 technical framework of language-based security~\citep{??}. We define
-trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
+trace-based properties of \emph{stepwise stack integrity} and
+\emph{lockstep stack confidentiality}.
 In applying these security concepts, we treat each caller-callee pair as
@@
   an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
-{\em Integrity} says that a caller's local data is protected from modification by its callee.
-\bcp{What about the ``stepwise'' part?}
-This is vital for local reasoning about the behavior of the
-caller.\bcp{Either delete that sentence or explain it more.}
-{\em Confidentiality} says that a caller's data is protected
+{\em Integrity} says that a caller's local data is protected from modification by its callee,
+and its {\em stepwise} interpretation formalizes this very naturally as a caller's local data
+never being modified during a call. {\em Confidentiality} says that a caller's data is protected
 from being leaked into its callee, and also that the callee cannot see uninitialized data.
@@
 behavior is determined only by a predictable interface that does not include the caller's
-local state or the uninitialized portion of the stack.
-Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
-notion that data outside of live frames should be inaccessible (which LSE
-does not address\bcp{Which supports my feeling that we should not advertise
-  ourselves as reimplementing LSE but rather as defining a different
-  condition based on similar intuitions.}).
-\bcp{These are the intuitions, but we should also mention that we want to define
-  these (at least partly) in terms of a sort-of-standard notion of ``nested
-  noninterference''!  Then, at the end of the intro (just before the
-  contributions) we can summarize the whole thing like ``So what is stack
-  safety?  It is nested NI.''}
-We also give a simple trace-based formalization of WBCF, which is essential for keeping the
-notions of ``caller'' and ``callee'' sensible. \bcp{That last should probably
-be a separate discussion, if I've succeeded in convincing people that WBCF
-is not part of stack safety.}
+local state or the uninitialized portion of the stack. This is formalized as a
+{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
+compares a pair of subtraces for each call, nested within other calls. The term
+{\em lockstep} describes the behavior of these pairs: they step together, and any
+deviation is recognized as a violation of confidentiality.
+Together, these properties capture the rough intuition of LSE, but are not identical --
+confidentiality contains temporal protections that LSE does not address.
+
+Since our stack safety occupies the same conceptual space as LSE, we should consider how it
+relates to WBCF. First, we observe that some degree of bracketed control flow is implicit
+in any notion stack safety, simply because calls and returns must be recognizeable; callers
+and callees, distinct. We treat control flow that creates an ambiguity between caller and
+callee as keeping the callee in control, and therefore such behavior can never give access
+to the caller's data. An example would be if the callee attempts to ``return'' to the
+wrong location, hoping to execute code on behalf of the caller: it may make the jump
+without violating stack safety, but it gains no privileges on the caller's data.
+Alternatively a callee might return past its immediate caller, to somewhere deeper in the
+stack, and stack safety does not inherently rule this out. We then define WBCF as an additional
+restriction, describing the standard control flow principle that a return pops exactly one frame
+from the stack.
+\sna{We got side-tracked a bit from talk of entry integrity, and Andrew made some good points
+  about it, so let's not rule that out just yet. But I focused on returns because there's a
+  clear story already.}
+
+Part of the challenge in fully characterizing stack safety is that at a low level,
+without built-in stacks, we are at the mercy of architectural choices that
+do not relate to the abstract notion of stack safety. We abstract over these choices
+as much as possible, so that our model can be applied to a wide range of architectures
+and enforcement mechanisms. But a second challenge is the diversity of enforcement
+mechanisms and their need for efficiency. In the name of efficiency, enforcement
+mechanisms may weaken their protections, arguing that the protections they do provide
+still capture stack safety in a meaningful sense. This calls for an {\em extensional}
+characterization of stack safety, one that describes the externally visible behavior
+of a stack safe system. Our stepwise and lockstep properties are {\em intensional},
+focusing on the internal behavior.
 
@@
 of maintaining the connection between an activation identity and the memory it owns.
-These vary in their efficiency and in the spatial and temporal precision of their security
-guarantees.
-For example, their \emph{Depth Isolation} policy  requires a strict mapping from a function
+These vary in their efficiency and in the spatial and temporal precision of their
+security guarantees.
+Their \emph{Depth Isolation} policy requires a strict mapping from a function
 to its frame from the start of the call to its end, and permits the function to write
@@
 detected before they can affect program behavior.
+
+To address this, we develop an \emph{observational integrity} property that
+is supports enforcement mechanisms that detect violations between the point when
+they occur and the point when they make an observable impact. Intuitively this property
+can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
+integrity implies observational integrity.
+\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
+  established that this was important.}
+We additionally sketch an associated {\em observational confidentiality} property
+makes confidentiality extensional in the same way, though we do not know of a real
+enforcement mechanism that weakens confidentiality in this way. Together,
+these form an {\em observational stack safety} that is highly extensional.
+We therefore argue that, despite its somewhat greater complexity,
+it may be a better minimal specification for
+stack safety enforcement mechanisms in general.
 
@@
 %  ones...''}
-
-However, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
+Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
 Clearing micro-policy does not enforce stepwise integrity.
-To address this, we develop an \emph{observational stack integrity} property that
-is intended to specify enforcement mechanisms that
-detect integrity violations between the point when they occur and the point when they make
-an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.
-\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
-  established that this was important.}
-
 Unfortunately the existing Lazy Tagging and Clearing
-micro-policy does not quite enforce this property either, as we show in
-testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
+micro-policy does not quite enforce observable confidentiality either. %, as we show in
+%testing~\apt{really??},
+but formalizing the ideal property allows us to see how it falls short.
 We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
@@
   need a caveat about how it's harder to test?}
-
-Observational integrity is attractive in being significantly more \emph{extensional} than
-stepwise integrity.  We therefore argue that, despite its somewhat greater complexity,
-it may be a better minimal specification for
-stack safety enforcement mechanisms in general.
 
@@
 attempts to read or write its private data or hijack its control flow. And in fact the caller,
-while presumably\bcp{not ``presumably'': {\em by definition}!} not malicious, may itself be confused or erroneous. Our attacker
+while not malicious, may itself be confused or erroneous. Our attacker
 model reflects this intention\bcp{I don't understand how it reflects this

### 2021-04-11 tex/main.tex
Tweak to lazy description
Redid Assumptions/Threat model section

@@
 %testing~\apt{really??},
-but formalizing the ideal property allows us to see how it falls short.
+But formalizing the ideal property allows us to see how it falls short -- a specific
+violation of the temporal aspect of confidentiality.
 We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
@@
 
-We begin with a description of our threat model (\cref{sec:threat}),
+We begin with a description of our assumptions and threat model (\cref{sec:threat}),
 then show how our properties apply to examples of stack unsafety 
@@
 
-\section{Threat Model and Assumptions}
+\section{Assumptions and Threat Model}
 \label{sec:threat}
@@
 
+\newcommand*{\rsp}{\textsc{sp}}
+
+Our properties are defined in terms of an abstract machine model that makes very few
+assumptions about the machine that instantiates it.  We assume a register
+machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other
+details of the machine's ISA are unimportant. We assume that we know ahead of time
+the layout of memory, including code and which addresses are within
+the bounds of the stack.
+
+Our stack-safety concepts depend
+critically on the notions of ``call'' and ``return,'' which are not fully explicit
+in machine code (the same instruction opcode may be used in some places as part of
+a call or return sequence and in other places for other purposes).
+%
+We assume we are given a machine language program annotated with the locations of
+instructions representing calls, returns, and function entry points. Although this
+information would typically come from a compiler together with the machine code,
+we do not assume anything about their provenance.
+
+We further parameterize over the structure of the stack, particularly the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention} and model it as a relation between machine states
+and addresses. Then from any call, those the sealing convention relates to the
+state will be protected until the return.
+
+We model the base machine as a step function over machine states, and a machine
+enhanced with an enforcement mechanism (in the language of tag-based enforcement,
+a {\em policy}) as a partial step function over machine states extended with some
+auxiliary state, refining the base machine. When the policy-enhanced machine fails
+to step, this is a {\em failstop}, halting the program before it can perform an
+action that would violate the security property;
+our properties are therefore naturally \emph{termination insensitive.}
+This model will be made more precise in \cref{sec:prelim}.
+
+This division of policy state from machine state is helpful when reasoning about
+enforcement mechanisms that keep a significant amount of data separate from the
+primary execution data to inform security. Obviously tag-based systems do this,
+with tags and extra policy state separated from and untouchable by application code.
+Another significant example is the shadow stack, where a second stack
+mirrors the real one, and a return may only occur if both stacks agree on the
+target address. Typically the shadow stack is isolated using a separate protection
+mechanism.
+
+The separation is less useful in protection mechanisms that interact more directly
+with application code, such as Cheri-style capabilities. In these instances
+the policy would not have auxiliary state, but would still refine the machine step
+function to create a distinct failstop condition. Other techniques such as software-based
+ones might never actually failstop, but fulfill the properties by never creating
+traces that violate them.
+
+\paragraph*{Threat Model}
+
 Intuititively, stack safety means that a caller's stack frame is protected
-from a confused or malicious callee's
-attempts to read or write its private data or hijack its control flow. And in fact the caller,
-while not malicious, may itself be confused or erroneous. Our attacker
-model reflects this intention\bcp{I don't understand how it reflects this
-  intuition about the caller---indeed, I don't understand the point about
-  the caller.} by assuming that both the attacker and the defender
-may run arbitrary machine code. Hardware and timing attacks are out of scope.
+from a confused or malicious callee's attempts to read or write its private data.
+The attacker model therefore treats every caller as a defender and its callee as
+an attacker, meaning that with nested calls, a single function may play both roles
+at varying points.
+
+We take a very strong threat model in which both caller and callee may execute
+arbitrary machine code. This means that even the defender might be ill-behaved
+to the point that it attempts to undermine its own protections, and we expect
+enforcement mechanisms to ensure that it fails.
+Hardware and timing attacks are out of scope, however.
 
@@
 
-\newcommand*{\rsp}{\textsc{sp}}
-
-In this very strong attacker model, there is little obvious distinction between
-the attacker and the defender. They are separated only by the control transfer
-inherent in their calls and returns. Our stack-safety concepts therefore depend
-critically on the notions of ``call'' and ``return,'' which are not fully explicit
-in machine code (the same instruction opcode may be used in some places as part of
-a call or return sequence and in other places for other purposes).
-%
-We assume we are given a machine language program annotated with the locations of
-instructions representing calls, returns, and function entry points. Although this
-information would typically come from a compiler together with the machine code,
-we do not assume anything about their provenance. This means that attackers (or
-foolish defenders) are under no obligation to have reasonable annotations.
-
 \paragraph*{Well-formedness and Bad Annotations}
 
-Since we rely so heavily on code annotations, we must ask: what if the annotations
+Since we rely heavily on code annotations, we must ask: what if the annotations
 are wrong? First, suppose a call or return occurs that is not annotated.
@@
 discussed in section \ref{sec:cheri}.
-
-\paragraph*{System Assumptions}
-
-Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the machine that instantiates it.  We assume a register
-machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other
-details of the machine's ISA are unimportant. We assume that we know ahead of time
-the layout of memory, including code (with annotations) and which addresses are within
-the bounds of the stack.
-%
-We further parameterize over the structure of the stack, particularly the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention} and model it as a relation between machine states
-and addresses. Then from any call, those the sealing convention relates to the
-state will be protected until the return.
-
-\paragraph*{Policies}
-
-We model the base machine as a step function over machine states, and a machine enhanced with
-an enforcement mechanism (in the language of tag-based enforcement, a {\em policy}) as
-a partial step function over machine states extended with some auxiliary state, refining
-the base machine. When the policy-enhanced machine fails to step, this is a {\em failstop},
-halting the program before it can perform an action that would violate the security property;
-our properties are therefore naturally \emph{termination insensitive.}
-This model will be made more precise in \cref{sec:prelim}.
 
@@
 
-\paragraph*{Micro-Policies}
-%
-Micro-Policies are a flexible tag-based, hardware-accelerated
-reference monitoring mechanism; they been applied to stack safety
-enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
-Here the policy state consists of metadata tags (e.g., identifying stack frames)
-attached to each value in memory or registers, and the policy step
-function checks that each machine operation obeys a set of rules on tags
-(e.g., that the current $\PCname$ tag matches the tag on the stack
-location being accessed), and halts the machine if not.  Distinguishing
-ordinary machine state and policy state is very natural in this setting.
+%\paragraph*{Micro-Policies}
+%
+%Micro-Policies are a flexible tag-based, hardware-accelerated
+%reference monitoring mechanism; they been applied to stack safety
+%enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
+%Here the policy state consists of metadata tags (e.g., identifying stack frames)
+%attached to each value in memory or registers, and the policy step
+%function checks that each machine operation obeys a set of rules on tags
+%(e.g., that the current $\PCname$ tag matches the tag on the stack
+%location being accessed), and halts the machine if not.  Distinguishing
+%ordinary machine state and policy state is very natural in this setting.
 %% Here the machine state is extended with tags paired with all components of the
@@
 
-\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
+%\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
 %These approaches are represented as policies that

### 2021-04-11 tex/main.tex
Indentation fix

@@
 \begin{abstract}
-\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
-  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
-  to leave it out of the title but hit it strongly in the abstract.}%
+%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
+%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
+%  to leave it out of the title but hit it strongly in the abstract.}%
 What is stack safety? The phrase is associated with a variety of compiler,
 run-time, and hardware mechanisms for guarding against corruption of stack
-memory.  But these mechanisms mostly lack rigorous specifications, appealing
-either to intuitive expectations of how good stack-using programs should
-behave or to specific examples of bad behavior.
-
-We propose a formal criterion for stack safety, phrased in terms of
+memory.  But these mechanisms typically lack rigorous specifications;
+they rely instead on informal descriptions of
+the safe stack manipulation behaviors that they allow,
+or specific examples of the bad behaviors that they prevent.
+
+We propose a formal criterion \apt{criteria?} for stack safety, phrased in terms of
 language-based security: a combination of an integrity property (``each
-caller's stack frame is invariant by the callee'') and a confidentiality
+caller's stack frame is held invariant by the callee'') and a confidentiality
 property (``the callee's behavior is insensitive to differences in the
-caller's state''). These are defined in terms of a control flow model
-in which callees are meaningfully distinct from their callers due to being
-bracketed by calls and returns. We further apply a straightforward notion of
-``well-bracketed control flow'' that ensures that callers' control flow
-can continue predictably after the return.
-
-  \bcp{I am still not clear
-  how well-bracketedness fits into the story.  (For me, it is the bracketing
-  that {\em defines} where calls and returns happen, so there is something
-  tautological about saying that `` well-bracketedness condition ensures
-  that callers and callees remain meaningfully distinct''!)}
-  \sna{Attempted to resolve this, it may need some work.}
-
-Our criterion is inspired by hardware tagging mechanisms for stack
-protection, in particular the stack-safety ``micro-policies''
-of~\citet{DBLP:conf/sp/RoesslerD18}, but its core reflects general
-principles of stack safety\bcp{I wish we did not have to be so apologetic
-  about being tied to micro-policies (and it is still not clear to me that
-  we have to be... do we have a clear statement someplace of why our
-  definition would not apply to CFI, for example? This seems like a critical
-  discussion...)}. Finally, we show that micro-policies do indeed enforce
-our properties using property-based randomized testing. An eager micropolicy
-that catches violations immediately enforces our stronger property, and we
-define a weaker, more extensional form of the property that is enforced
-by a ``lazy'' micropolicy -- one that, for performance reasons, may permit
-local violations as long as they do not influence future behavior.
+caller's state''). These properties are sensible even in the presence of
+abnormal control flow.\apt{Is this important enough to mention here?}
+We further apply\apt{what does this mean? we didn't
+  invent it} a straightforward notion of
+``well-bracketed control flow'' that restricts control flow to the
+typical pattern of matched calls and returns.
+
+We show that our properties describe the behavior of the stack-safety
+``micro-policies'' of~\citet{DBLP:conf/sp/RoesslerD18}
+using property-based randomized testing. An eager micro-policy
+that catches violations immediately enforces our primary ``stepwise''
+properties, and we define weaker, more extensional ``observational''
+properties that are enforced even by a ``lazy'' micro-policy -- one
+that, for performance reasons, may catch violations after they occur
+but before they become visible.
+
+We argue \apt{do we?} that our properties are suited to a range of other enforcement
+mechanisms. The strict stepwise properties can describe mechanisms
+that catch errors immediately, as is typical. While the observational
+properties are closely linked to lazy micro-policies they are weak enough
+to be a lower bound that captures stack safety for any reasonable enforcement mechanism.
+
+\apt{This is a good story, but does it describe the paper we are able to produce this week?
+  We were rejected before exactly because we claimed to have
+  a general criterion, but did not in fact demonstrate its generality (which is not
+  obvious). What has changed?}
 
@@
 based on memory-safety vulnerabilities, leading to consequences
-from leakage or corruption of private stack data to
+that range from leakage or corruption of private stack data to
 control-flow hijacking. To detect or prevent such attacks, a profusion of
@@
 The protections offered by such mechanisms are generally described in terms
-of a set of concrete examples of attacks that they can prevent, leaving a
-more general characterization to the reader's intuition.
+of a set of concrete examples of attacks that they can prevent, such as corruption
+of return addresses, buffer overflows, or use of uninitialized variables. This leaves the
+task of formulating a more general characterization to the reader's intuition.
 Moreover, the details of these mechanisms can be quite intricate
@@
 
-So how should we formally define stack safety? Definitions in the literature are commonly
-informal and proceed by enumerating the specific bad things that a given enforcement mechanism
-prevents, such as corruption of return addresses, buffer overflows, or
-use of uninitialized variables.\bcp{Most of that sentence should be merged
-  into the similar sentence above.  Doesn't belong here.}
 %\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
@@
 
-\bcp{If we end up adopting my recent proposal for defining subtraces and
-  separating WBCF, I think we can be a bit stronger here: We define stack
-  safety in IFC terminology based on the same intuitions as their LSE, but
-  we consider (and define) WBCF to be a separate thing---desirable, but
-  distinct from safety.  (If we make this argument, we'll need to show an
-  example of a program that is stack safe but doesn't have WBCF.)}
-In this work, we take the (informal) definitions of WBCF and LSE as a starting
+In this work, we take the (informal) definitions of WBCF\apt{or maybe not?}  and LSE as a starting
 point, but choose a different approach to formalizing them, based on the
@@
 trace-based properties of \emph{stepwise stack integrity} and
-\emph{lockstep stack confidentiality}.
-In applying these security concepts, we treat each caller-callee pair as
-mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
-  ``nesting'' and our novelty factor here? Also about domains??}
-This differs from common instances in language-based security in that the
+\emph{lockstep stack confidentiality}.\apt{Could we just say ``stepwise'' for both?}
+In applying these security concepts, we treat each caller as needing protection
+from its callee, even as that callee may need protection from its callee in turn.
+This differs from common practice in language-based security in that the
 confidentiality property applies to many nested subtraces within a single execution.
-\sna{Thinking about how to express domains, or whether to. I'm not sure it's
-  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
+\sna{How novel is this nesting?} \apt{Lost the reasoning principle that was here?}
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
-and its {\em stepwise} interpretation formalizes this very naturally as a caller's local data
-never being modified during a call. {\em Confidentiality} says that a caller's data is protected
+and its {\em stepwise} interpretation formalizes this very naturally: a caller's local data
+is never modified during a call. {\em Confidentiality} says that a caller's data is protected
 from being leaked into its callee, and also that the callee cannot see uninitialized data.
@@
 deviation is recognized as a violation of confidentiality.
-Together, these properties capture the rough intuition of LSE, but are not identical --
-confidentiality contains temporal protections that LSE does not address.
+Together, these properties capture the rough intuition of LSE, but are somewhat stronger:
+confidentiality contains temporal protections that LSE does not address.\apt{Maybe want to
+  say a little more about this here, since it is a novelty wrt/ existing work?}
 
 Since our stack safety occupies the same conceptual space as LSE, we should consider how it
-relates to WBCF. First, we observe that some degree of bracketed control flow is implicit
-in any notion stack safety, simply because calls and returns must be recognizeable; callers
-and callees, distinct. We treat control flow that creates an ambiguity between caller and
-callee as keeping the callee in control, and therefore such behavior can never give access
+relates to WBCF.\apt{Or maybe not. This disucssion seems too detailed for this spot in the intro.
+  While it is good to note that our integrity/confidentiality properties make sense without
+  WBCF-or-similar, LSE is also described independently from WBCF, so there's no novelty in that.}
+First, we observe that some degree of bracketed control flow is implicit
+in any notion of stack safety, simply because calls and returns must be recognizeable so
+that callers and callees are distinguished. We treat control flow that creates an ambiguity between caller and
+callee\apt{this is obscure}  as keeping the callee in control, and therefore such behavior can never give access
 to the caller's data. An example would be if the callee attempts to ``return'' to the
@@
 Alternatively a callee might return past its immediate caller, to somewhere deeper in the
-stack, and stack safety does not inherently rule this out. We then define WBCF as an additional
-restriction, describing the standard control flow principle that a return pops exactly one frame
+stack, and stack safety does not inherently rule this out. We therefore define WBCF as an additional
+restriction \apt{on what? in what way is it formalized differently than by Skorstengaard?},
+describing the standard control flow principle that a return pops exactly one frame
 from the stack.
@@
 
-Part of the challenge in fully characterizing stack safety is that at a low level,
-without built-in stacks, we are at the mercy of architectural choices that
-do not relate to the abstract notion of stack safety. We abstract over these choices
+Part of the challenge in fully\apt{omit?} characterizing stack safety is that in a low-level machine
+without built-in stacks, we are at the mercy of architectural
+choices that do not relate to the abstract notion of stack safety\apt{a bit vague. what kind of choices?}.
+We abstract over these choices
 as much as possible, so that our model can be applied to a wide range of architectures
-and enforcement mechanisms. But a second challenge is the diversity of enforcement
-mechanisms and their need for efficiency. In the name of efficiency, enforcement
-mechanisms may weaken their protections, arguing that the protections they do provide
-still capture stack safety in a meaningful sense. This calls for an {\em extensional}
-characterization of stack safety, one that describes the externally visible behavior
-of a stack safe system. Our stepwise and lockstep properties are {\em intensional},
+and enforcement mechanisms.
+
+\apt{Intra-paragraph flow seemed awkward.}
+A second challenge is
+% the diversity of enforcement mechanisms APT:already implied above
+enforcement mechanisms may weaken their protections that in the name of efficiency,
+while arguing that they still capture stack safety in a meaningful sense.
+\apt{Remainder of this paragraph might be better \emph{after} the R+DH para below.}
+This calls \apt{not
+  immediately obvious why. is this true for anything other than laziness?}
+for an {\em extensional}
+characterization of stack safety that describes the externally observable behavior
+of a stack-safe system. In contract, our stepwise and lockstep properties are {\em intensional},
 focusing on the internal behavior.
@@
 using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
-Intuitively, memory cells within each stack frame are tagged with the
+Memory cells within each stack frame are tagged with the
 identity of the function activation
 that owns the frame, and access to those locations is then gated
-according to the activation that is currently executing. Roessler and Dehon
+according to the activation that is currently executing. Roessler and DeHon
 explore alternative versions of this basic idea, with different ways of identifying
-functions and deciding exactly which accesses are allowed, to enable different ways
-of maintaining the connection between an activation identity and the memory it owns.
+functions and deciding exactly which accesses are allowed.
+% APT: seem repetitive:
+%, to enable different ways
+%of maintaining the connection between an activation identity and the memory it owns.
 These vary in their efficiency and in the spatial and temporal precision of their
 security guarantees.
-Their \emph{Depth Isolation} policy requires a strict mapping from a function
-to its frame from the start of the call to its end, and permits the function to write
+The \emph{Depth Isolation} policy requires a strict mapping from a function
+to its frame from the start of the call to its end\apt{unclear. why not simply say
+  what depth is?}, and permits the function to write
 only into that frame.
 This seems simple and intuitively correct, but it can be expensive to implement the strict
-mapping. Conversely, their \emph{Lazy Tagging and Clearing} policy improves
+mapping. Conversely, the \emph{Lazy Tagging and Clearing} policy improves
 performance by assigning locations to the active function when they are written, which
 permits  ``illicit'' writes to inaccessible stack frames so long as the
-written locations are not subsequently read by their ``real'' owners. The justification
+written locations are not subsequently read by their ``real'' owners. The intuitive justification
 for this form of laziness is that any genuinely dangerous writes will ultimately be
@@
 
-To address this, we develop an \emph{observational integrity} property that
-is supports enforcement mechanisms that detect violations between the point when
+To formalize this intuition, we develop an \emph{observational integrity} property which
+characterizes enforcement mechanisms that detect violations between the point when
 they occur and the point when they make an observable impact. Intuitively this property
@@
 \sna{There is a bit of massaging old proofs to do here, and I don't recall if we
-  established that this was important.}
+  established that this was important.\apt{medium importance. it is obvious that
+    observational integerity allows behaviors that stepwise does not, but
+    maybe not obvious that the converse is not true too.}}
 We additionally sketch an associated {\em observational confidentiality} property
-makes confidentiality extensional in the same way, though we do not know of a real
-enforcement mechanism that weakens confidentiality in this way. Together,
-these form an {\em observational stack safety} that is highly extensional.
+makes confidentiality extensional in the same way, although we do not know of a real
+enforcement mechanism that weakens confidentiality in this way.\apt{why bother mentioning it then?}
+Together, these form an {\em observational stack safety} that is highly extensional.
 We therefore argue that, despite its somewhat greater complexity,
@@
 \sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
-  need a caveat about how it's harder to test?}
+  need a caveat about how it's harder to test?}\apt{As per email discussion,
+  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and Clearing,
+  by loosening the temporal safety properties?}
 
@@
   intent -- how long should we permit memory to be shared? But if the compiler is passing
-  pointers, as in pass by reference, we have a clear sense of where sharing should stop.}
+  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
+  \apt{N.B.Doing call-by-reference properly requires that the passed variable is accessible
+    by all descendents, not just immediate callee, which is doable with depth tags but not
+    obviously with per-activation tags }}
 
@@
 control flow.
-
-While our property definitions were developed with micro-policy enforcement
-in mind, they agnostic about many implementation details and should be
-applicable to other systems with minor adjustment.\apt{move this up?} \bcp{I
-  am still hoping to get a clearer understanding for myself of just what is
-  micropolicy-specific about our definitions...}
 
@@
   language-based security
-  to formalize a parameterized {\em stepwise stack-safety property} and argue that
+  to formalize a parameterized {\em stepwise stack-safety property} \apt{lockstep?} and argue that
   it captures the intuitive concepts of local stack encapsulation, well-bracketed control
@@
 \item
-  We show how our models can be instantiated and extended to handle passing
-  arguments on the stack~(\cref{sec:??}) and simple
-  coroutines~(\cref{sec:??}).\bcp{``instantiated and extended'' is awkward.
-  Maybe ``We show how our models can be instantiated to handle passing
-  arguments on the stack~(\cref{sec:??}) and extended with simple
-  coroutines~(\cref{sec:??}).''  If that's what we mean.}
-%  This policy does not enforce our strict
-%  local state encapsulation property,
-%  but a modified version of it using unique per-activation tags enforces
-%  an (also provably) weaker lazy variant of the property, which still prohibits
-%  the same intuitively stack unsafe behaviors.
-%
-% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
-%   but the point we can make is still strong: To make the property easier
-%   to enforce, we can slightly weaken its guarantees; the weaker version is
-%   a bit more complicated to state but still retains the pragmatic
-%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
-%   depth-based tags, which do not enforce even our lazy policy. So we need to
-%   introduce a version of LT\&C that uses unique activation tags.}
-
-  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
-%  realizible by
-%  a previously proposed micro-policy enforcement
-%  mechanism~(\Cref{sec:micro-policy}). \leo{And it says yes on a
-%    substantial body of code}. We also present a variation of the
-%  property + policy that is lazy, easier/more efficient to implement,
-%  and provably weaker (\Cref{sec:lazy}).
-%\item A provably stronger (inductive? step-by-step/stepwise) variant
-%  of this property tuned for testing/verification.  We test the whole
-%  thing with a cool framework. \bcp{Which section?}
+  We extend our model so that it can be instantiated in systems that pass arguments
+  on the stack~(\cref{sec:??}) and in those with simple coroutines~(\cref{sec:??}).
 \end{itemize}
-\Cref{sec:relwork} discusses related work, and
-\cref{sec:future} sketches directions for future work.
 
@@
 and describe testing (\cref{sec:testing}).
+\Cref{sec:relwork} discusses related work, and
+\cref{sec:future} sketches directions for future work.
 
@@
 
-\bcp{Again, I think we need to put ``Assumptions'' (i.e., our basic picture
-  of the world) first, as context.  E.g., none of the threat model
-  discussion makes sense until the reader understands that we {\em define}
-  function calls and returns by means of annotations (and that we MUST do
-  something like that because we are talking about machine code).}
-
 \newcommand*{\rsp}{\textsc{sp}}
@@
 Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the machine that instantiates it.  We assume a register
-machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other
+assumptions about the concrete machine that instantiates it.  We assume a register
+machine with a program counter \(\PCname\) and stack pointer \(\rsp\)\apt{do we really
+  need these except for WBCF?}; other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
-the layout of memory, including code and which addresses are within
+the layout of memory, including where code lives and which addresses are within
 the bounds of the stack.
@@
 critically on the notions of ``call'' and ``return,'' which are not fully explicit
-in machine code (the same instruction opcode may be used in some places as part of
-a call or return sequence and in other places for other purposes).
-%
-We assume we are given a machine language program annotated with the locations of
-instructions representing calls, returns, and function entry points. Although this
-information would typically come from a compiler together with the machine code,
+in machine code---the same instruction opcode may be used in some places as part of
+a call or return sequence and in other places for other purposes.
+%
+So we assume we are given a machine language program annotated with the locations of
+instructions representing calls, returns, and function entry points. Although these
+annotations would typically come from a compiler together with the machine code,
 we do not assume anything about their provenance.
@@
 this the {\em sealing convention} and model it as a relation between machine states
-and addresses. Then from any call, those the sealing convention relates to the
-state will be protected until the return.
+and addresses. From any call, those addresses that the sealing convention relates to the
+state\apt{rephrased, but still obscure} will be protected until the return.
 
@@
 
+\sna{This is awkward.}
 This division of policy state from machine state is helpful when reasoning about
@@
 The separation is less useful in protection mechanisms that interact more directly
-with application code, such as Cheri-style capabilities. In these instances
+with application code, such as Cheri-style capabilities\citep{??}. Here
 the policy would not have auxiliary state, but would still refine the machine step
-function to create a distinct failstop condition. Other techniques such as software-based
-ones might never actually failstop, but fulfill the properties by never creating
-traces that violate them.
+function to create a distinct failstop condition. Other techniques such as
+software fault isolation~\citep{??} might never actually failstop, but fulfill
+the properties by rewriting code so that traces can never violate them.
 
@@
 
-Intuititively, stack safety means that a caller's stack frame is protected
+Intuitively, stack safety means that a caller's stack frame is protected
 from a confused or malicious callee's attempts to read or write its private data.
@@
 to the point that it attempts to undermine its own protections, and we expect
-enforcement mechanisms to ensure that it fails.
+enforcement mechanisms to ensure that it fails.\apt{obscure}
 Hardware and timing attacks are out of scope, however.
 
-%\leo{Anything else to add here?}
-%\leo{Maybe a good place to discuss calling conventions and that they
-%  are essentially enforced by the policy or sth?}
-%\rb{Intuition that the callee can do anything in principle, but the machine will
-%halt if it does not follow the calling convention, etc.?}
-
 \paragraph*{Well-formedness and Bad Annotations}
 
+\sna{TODO: Need to rewrite this for the new model of WBCF.}
 Since we rely heavily on code annotations, we must ask: what if the annotations
@@
 In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
-and returns are identified with ``blessed sequences'' of instructions, which the tag policy
-will force to be executed from start to finish. The annotated instruction should always
-be the final instruction of the sequence. What if an annotated instruction exists outside
+and returns are implemented by fixed sequences of instructions which the tag policy
+will force to be executed from start to finish. The annotated call or return instruction should always
+be the final instruction of the appropriate sequence. What if an annotated instruction exists outside
 such a sequence? We will statically rule out some scenarios with a well-formedness criterion
@@
 We believe that it is flexible enough to capture code-altering approaches, such as code
-rewriting or capability-based techniques, perhaps with less elegance.
+rewriting or capability-based techniques, though perhaps with less elegance.
 

### 2021-04-12 tex/main.tex
Addressed notes in abstract, introduction, some in section 2

--- 2021-04-11 tex/main.tex
@@
 property (``the callee's behavior is insensitive to differences in the
-caller's state''). These are defined in terms of a control flow model
-in which callees are meaningfully distinct from their callers due to being
-bracketed by calls and returns. We further apply a straightforward notion of
-``well-bracketed control flow'' that ensures that callers' control flow
-can continue predictably after the return.
-
-  \bcp{I am still not clear
-  how well-bracketedness fits into the story.  (For me, it is the bracketing
-  that {\em defines} where calls and returns happen, so there is something
-  tautological about saying that `` well-bracketedness condition ensures
-  that callers and callees remain meaningfully distinct''!)}
-  \sna{Attempted to resolve this, it may need some work.}
-
-Our criterion is inspired by hardware tagging mechanisms for stack
-protection, in particular the stack-safety ``micro-policies''
-of~\citet{DBLP:conf/sp/RoesslerD18}, but its core reflects general
-principles of stack safety\bcp{I wish we did not have to be so apologetic
-  about being tied to micro-policies (and it is still not clear to me that
-  we have to be... do we have a clear statement someplace of why our
-  definition would not apply to CFI, for example? This seems like a critical
-  discussion...)}. Finally, we show that micro-policies do indeed enforce
-our properties using property-based randomized testing. An eager micropolicy
-that catches violations immediately enforces our stronger property, and we
-define a weaker, more extensional form of the property that is enforced
-by a ``lazy'' micropolicy -- one that, for performance reasons, may permit
-local violations as long as they do not influence future behavior.
+caller's state''). These properties are sensible even in the presence of
+abnormal control flow. We further apply a straightforward notion of
+``well-bracketed control flow'' that restricts control flow to the
+typical pattern matched calls and returns.
+
+We show that our properties describe the behavior the stack-safety
+``micro-policies'' of~\citet{DBLP:conf/sp/RoesslerD18}
+using property-based randomized testing. An eager micropolicy
+that catches violations immediately enforces our primary ``stepwise''
+properties, and we define weaker, more extensional ``observational''
+properties that are enforced even by a ``lazy'' micropolicy -- one
+that, for performance reasons, may catch violations after they occur
+but before they become visible.
+
+We argue that our properties are suited to a range of other enforcement
+mechanisms. The strict stepwise properties can describe mechanisms
+that catch errors immediately, as is typical. While the observational
+properties are closely linked to lazy micropolicies they are weak enough
+to be a lower bound that captures stack safety for any reasonable enforcement.
 
@@
 The protections offered by such mechanisms are generally described in terms
-of a set of concrete examples of attacks that they can prevent, leaving a
+of a set of concrete examples of attacks that they can prevent, such as corruption
+of return addresses, buffer overflows, or use of uninitialized variables. This leaves a
 more general characterization to the reader's intuition.
@@
 
-\bcp{If we end up adopting my recent proposal for defining subtraces and
-  separating WBCF, I think we can be a bit stronger here: We define stack
-  safety in IFC terminology based on the same intuitions as their LSE, but
-  we consider (and define) WBCF to be a separate thing---desirable, but
-  distinct from safety.  (If we make this argument, we'll need to show an
-  example of a program that is stack safe but doesn't have WBCF.)}
 In this work, we take the (informal) definitions of WBCF and LSE as a starting
@@
 \emph{lockstep stack confidentiality}.
-In applying these security concepts, we treat each caller-callee pair as
-mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
-  ``nesting'' and our novelty factor here? Also about domains??}
+In applying these security concepts, we treat each caller as needing protection
+from its callee, even as that callee may need protection from its callee in turn.
 This differs from common instances in language-based security in that the
 confidentiality property applies to many nested subtraces within a single execution.
-\sna{Thinking about how to express domains, or whether to. I'm not sure it's
-  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
+\sna{How novel is this nesting?}
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
@@
 
-While our property definitions were developed with micro-policy enforcement
-in mind, they agnostic about many implementation details and should be
-applicable to other systems with minor adjustment.\apt{move this up?} \bcp{I
-  am still hoping to get a clearer understanding for myself of just what is
-  micropolicy-specific about our definitions...}
-
 %% \paragraph*{The Use of Stack Safety}
@@
 \label{sec:threat}
-
-\bcp{Again, I think we need to put ``Assumptions'' (i.e., our basic picture
-  of the world) first, as context.  E.g., none of the threat model
-  discussion makes sense until the reader understands that we {\em define}
-  function calls and returns by means of annotations (and that we MUST do
-  something like that because we are talking about machine code).}
 
@@
 
-%\leo{Anything else to add here?}
-%\leo{Maybe a good place to discuss calling conventions and that they
-%  are essentially enforced by the policy or sth?}
-%\rb{Intuition that the callee can do anything in principle, but the machine will
-%halt if it does not follow the calling convention, etc.?}
-
 \paragraph*{Well-formedness and Bad Annotations}
 
+\sna{TODO: Need to rewrite this for the new model of WBCF.}
 Since we rely heavily on code annotations, we must ask: what if the annotations

### 2021-04-12 tex/main.tex
pass on abstract

@@
 \begin{abstract}
-\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
-  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
-  to leave it out of the title but hit it strongly in the abstract.}%
+%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
+%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
+%  to leave it out of the title but hit it strongly in the abstract.}%
 What is stack safety? The phrase is associated with a variety of compiler,
 run-time, and hardware mechanisms for guarding against corruption of stack
-memory.  But these mechanisms mostly lack rigorous specifications, appealing
-either to intuitive expectations of how good stack-using programs should
-behave or to specific examples of bad behavior.
-
-We propose a formal criterion for stack safety, phrased in terms of
+memory.  But these mechanisms typically lack rigorous specifications;
+they rely instead on informal descriptions of
+the safe stack manipulation behaviors that they allow,
+or specific examples of the bad behaviors that they prevent.
+
+We propose a formal criterion \apt{criteria?} for stack safety, phrased in terms of
 language-based security: a combination of an integrity property (``each
-caller's stack frame is invariant by the callee'') and a confidentiality
+caller's stack frame is held invariant by the callee'') and a confidentiality
 property (``the callee's behavior is insensitive to differences in the
 caller's state''). These properties are sensible even in the presence of
-abnormal control flow. We further apply a straightforward notion of
+abnormal control flow.\apt{Is this important enough to mention here?}
+We further apply\apt{what does this mean? we didn't
+  invent it} a straightforward notion of
 ``well-bracketed control flow'' that restricts control flow to the
-typical pattern matched calls and returns.
-
-We show that our properties describe the behavior the stack-safety
+typical pattern of matched calls and returns.
+
+We show that our properties describe the behavior of the stack-safety
 ``micro-policies'' of~\citet{DBLP:conf/sp/RoesslerD18}
-using property-based randomized testing. An eager micropolicy
+using property-based randomized testing. An eager micro-policy
 that catches violations immediately enforces our primary ``stepwise''
 properties, and we define weaker, more extensional ``observational''
-properties that are enforced even by a ``lazy'' micropolicy -- one
+properties that are enforced even by a ``lazy'' micro-policy -- one
 that, for performance reasons, may catch violations after they occur
@@
 
-We argue that our properties are suited to a range of other enforcement
+We argue \apt{do we?} that our properties are suited to a range of other enforcement
 mechanisms. The strict stepwise properties can describe mechanisms
 that catch errors immediately, as is typical. While the observational
-properties are closely linked to lazy micropolicies they are weak enough
-to be a lower bound that captures stack safety for any reasonable enforcement.
+properties are closely linked to lazy micro-policies they are weak enough
+to be a lower bound that captures stack safety for any reasonable enforcement mechanism.
+
+\apt{This is a good story, but does it describe the paper we are able to produce this week?
+  We were rejected before exactly because we claimed to have
+  a general criterion, but did not in fact demonstrate its generality (which is not
+  obvious). What has changed?}
 

### 2021-04-12 tex/main.tex
comments and word-smithing on 2
comments and word-smithing and diagram-smithing on 3

--- 2021-04-11 tex/main.tex
@@
   pointers, as in pass by reference, we have a clear sense of where sharing should stop.
-  \apt{N.B.Doing call-by-reference properly requires that the passed variable is accessible
+  \apt{N.B.Doing pass-by-reference properly requires that the passed variable is accessible
     by all descendents, not just immediate callee, which is doable with depth tags but not
@@
 \section{Stack Safety, Intuitively}
+\apt{section title is a bit sad since we're trying to convert intuition to formal spec!}
 \label{sec:running-example}
@@
 int f() {
-10:     $SP-2 = 0;
+10:     *($SP-2) = 0;
 11[R]:  return 5;
@@
 \end{subfigure}
+\caption{An Integrity Violation}
 \label{fig:int1}
-\caption{Integrity Violation}
 \end{figure}
@@
 \begin{itemize}
-\item Examples that appear to violate stack safety, violate at least one of our properties
-\item We do not find desireable behaviors that appear stack-safe yet violate our properties
-\item Our properties are testably implemented by stack safety micro-policies from the
+\item examples that appear to violate stack safety violate at least one of our properties;
+\item we do not find desireable behaviors that appear stack-safe yet violate our properties; and
+\item our properties are testably implemented by stack safety micro-policies from the
   literature, and the micro-policies are not significantly more conservative
-  (i.e., policy failstops usually correspond to imminent property violations)
+  (i.e., policy failstops usually correspond to imminent property violations).
 \end{itemize}
@@
 but to reflect the increased power of the underlying assembly code, we allow code
-to access the stack pointer directly, written {\tt \$SP}. The layout of the stack
-is given to the right, with a box representing a word of memory, and variables stored
-in declaration order from left to right. The stack grows upward in memory and to the
-right on paper.
+to read the stack pointer ({\tt \$SP}) directly and use
+it to form pointers.\apt{I've added some strategic *'s}
+Call and return annotations are marked {\tt[C]} and {\tt[R]}
+respectively. Local variables are assumed to be stored in the stack frame,
+in declaration order. Unless otherwise stated, arguments and return values are passed in registers.
+The machine is word-addressed. The stack grows upward in
+memory and {\tt \$SP} points to the first unused word above the top of the stack. 
+At the right of each example, we show the layout of stack memory just before
+the specified line number is executed; each box represents a memory word and
+addresses increase to the right.
 
@@
 
-Example \ref{fig:int1} shows an integrity violation. It consists of a {\tt main}
+Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
 function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
@@
 
-The black rectangle represents the return address. In a typical calling convention
-this will have been passed in a register during the call itself, and {\tt f} stores
+The black rectangle represents the return address. In a typical RISC calling convention
+this will have been stored in a register by call itself, and {\tt f} stores
 it in the stack during its preamble, so it is not itself sealed.
@@
 11:     y = 5;
-12:     z = $SP-2;
+12:     z = *($SP-4);
 13[R]:  return y+z;
@@
 
-\label{fig:conf1}
 \end{subfigure}
 \caption{A Confidentiality Violation}
+\label{fig:conf1}
 \end{figure}
@@
 of the context of its call (except for its arguments and any global variables).
-We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
-security. Noninterference describes knowledge: it states that for some system components
+We state this in a style similar to {\em noninterference}~\citep{??}.
+Noninterference describes knowledge: it states that for some system components
 containing secret values, we can substitute any values whatsoever without changing the results
@@
 
-In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
+In Figure~\ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
 by reading {\tt main}'s variable {\tt x}. The trace
@@
 changed to 5 in both the primary trace and the variant. The primary and variant agree on
-the location that changed, so confidentiality was not violated. But after line 13,
+the value in the changed location, so confidentiality was not violated. But after line 13,
 the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
-in the variant trace it might have any value \(v_2\). Say \(v_2 = 43\) -- then the variant
+in the variant trace it might have any value \(v_2\). If \(v_2 \neq 42\) then the variant
 has behaved differently than the primary, violating confidentiality.
@@
 as: for any state at the entry to a function, for any variant over the values in the stack,
-the primary state and its variant will step in lockstep until they return. Any system components
-whose values update will be identical between the primary and variant traces. This includes not
+the primary state and its variant will step in lockstep until they return. Any system components\apt{first use of this?}
+whose values change will be identical between the primary and variant traces. This includes not
 just memory, but registers as well, so that returning a secret is a violation. And when we
-extend the model with a notion of external output, it will include any values that are
-outputted.
+extend the model with a notion of external output, it\apt{what?}  will include any values that
+emitted.
 
@@
 This treatment is roughly in line with more standard uses of noninterference, which typically
-vary states at the starts of whole program runs.
+vary states at the starts of whole program runs.\apt{This is all a bit cryptic.}
 
@@
     \memoryaddrs{8em}
-    \memory{4}{\unsealc}
+%    \memory{4}{\unsealc}
+    \memory{1}{\unsealc}
+    \memory{1}{\retptrc}
+    \memory{2}{\unsealc}
     ~$\cdots$
@@
 
-It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt y}
-and {\tt z}. In figure \ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
+It may be unclear in figure \ref{fig:conf1} why we vary the initial values of {\tt y}
+and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
+temporal stack safety.
+In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
 that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
-{\tt y} without even directly using its ability to read the stack pointer. In
+{\tt y} without even directly reading the stack pointer. In
 this case, {\tt g} branches on that value, which is a different way of
@@
 \begin{verbatim}
-int stash = 0;
+int *stash = 0;
 
@@
 10:    if(!stash)
-11:      stash = &a;
+11:      stash = a;
 12:    *stash = 5;
@@
       ~$\cdots$ \\
-      {\tt 11} &
+      {\tt 13} &
       \memoryaddrs{17em}
@@
       \\
-      {\tt 11} &
+      {\tt 13} &
       \memoryaddrs{17em}
@@
 \end{subfigure}
-
+\caption{A violation with pass-by-reference}
 \label{fig:passing}
@@
 
-Up til now we have ignored function arguments, since our simplest model places them in
-registers. This is an oversimplication, which we will now correct, starting with the more
-interesting case: call-by-reference. There are important stack safety concerns
-when passing arguments by reference, as seen in example \ref{fig:passing}. In this example,
+To this point we have ignored function arguments, since our simplest model places them in
+registers. This is an oversimplification, which we will now correct, starting with the more
+interesting case: pass-by-reference.
+\apt{This is a nice and clear example, but I think the ordering here is confusing. Pass-by-reference and passing arguments on the stack are
+  entirely separable issues. I think it would be easiest to explain the latter first, as the
+  simplest example of passed arguments.}
+There are important stack safety concerns
+when passing arguments by reference, as seen in Figure~\ref{fig:passing}. In this example,
 {\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
@@
 
-At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying the memory to be
-passed, and our stack model marks it in magenta. Then the caller's frame that are not
-passed are sealed, as usual. Just as we marked sealed memory with the depth of its owner,
-we do the same to passed memory. We keep track of the current depth of the active function
-next to the stack pointer in the diagram. A passed address may only be accessed when the
-current depth is one higher than the one it is sealed with.
+At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying a memory location to be
+passed, and our diagram displays it in magenta. The caller's frame locations that are not
+passed are sealed as usual. Just as we marked sealed memory with the depth of its owner,
+we do the same for passed memory. We keep track of the current depth of the active function
+next to the stack pointer in the diagram.\apt{Not seeing that}  A passed address may only be accessed when the
+current depth is one higher than the one it is sealed with. \apt{This is not good enough!
+  The callee can in turn pass the same argument by reference to its callee...}
 
@@
 Passed addresses are also excluded from confidentiality: at the start of each call, we vary
-only stack addresses that are not passed. This is important, as we use the passing label
-for those arguments that are passed by value on the stack, as well. We might instead
-treat them as unsealed, but then we would treat them as confidential like the rest of the stack.
+only stack addresses that are not passed. This is important, as we also use the passing label
+for those arguments that are passed by value on the stack. \apt{this is hidden like an afterthought...}
+We might instead
+treat them as unsealed, but then we would treat them as confidential like the rest of the stack.\apt{unclear.
+  would this work or not?}
 
@@
 unprotected, and perhaps attach a separate memory safety policy to add some protection
-to such variables.
+to such variables.\apt{Neither clear nor convincing! Taking explicit addresses of stack
+  variables is very interesting to many reviewers!}
 

### 2021-04-12 tex/main.tex
Section 5 redone somewhat to take return targets into account

--- 2021-04-12 tex/main.tex
@@
 
-\section{Local State Encapsulation, Formally}
+\section{Stack Safety, Formally}
 \label{sec:lse}
@@
 \[\ann \in \ANNS ::= \callmap + \retmap\]
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]
-
-\paragraph*{Call Depth and Domains}
-
-At any given step of a program, some number of callers have pushed their frames
-onto the stack, ``sealing'' that memory for later use; this number is the current
-{\em call depth}. The call depth, as well as the association between stack addresses
-and sealed frames, are dynamic and cannot be deduced from the state alone, so we will
-track them over the course of an execution as additional context that informs the property.
-
-Each component will be assigned to a {\em domain}, which may be {\em Outside} of the
-stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed} by a particular
-depth.
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \ANNS\]
+
+When a function makes a call, it establishes that its frame should be
+``sealed'' for later use. At any given step of a program, any number of
+activations may have claimed memory in this way, but which cells are sealed
+is dynamic information and cannot be deduced from the state alone. So it must
+be tracked separately. We must also track the conditions under which a caller
+considers itself to have been returned to.
+
+Each system component will be assigned to a {\em domain}, which may be {\em Outside}
+of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
+at a particular depth. Each caller will register a {\em target} condition -- a
+predicate on machine states that must hold for it to consider itself active.
+A context is a pair of a ``domain map'' from components to domains, and a
+``return target list,'' which is a stack of targets, one for each caller awaiting
+return.
 
@@
 \[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-  \times (\depth \in \mathbb{N}) \]
-
-An initial context \(\context_0 = \domMap_0, 0\) where \(\domMap_0\) maps
+  \times (\rts \in \listT (\target \subseteq \MACHS)) \]
+
+An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
 the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
@@
 \(\stepsto_C\). We will define this in terms of the operations
-\(\push{\domMap}{n}{\components}\), in which \(\components \subseteq \COMPONENTS\),
-and \(\pop{\domMap}{n}\), pronounced ``push'' and ``pop'' and defined as follows.
+\(\push{\domMap}{\depth}{\components}\), in which \(\depth \in \mathbb{Z}\) and
+\(\components \subseteq \COMPONENTS\), and \(\pop{\domMap}{\depth}\), pronounced
+``push'' and ``pop'' and defined as follows.
 
@@
 \begin{cases}
-  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1} \\
+  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth'} \text{ where }
+              \depth' >= \depth\\
   \domMap ~ \component & \text{else} \\
@@
 \]
-
-Then we can define the context update relation. Let \(\sealcon\) be the sealing convention
+%
+Push seals a set of components with its depth argument, and pop unseals all components
+that were as deep as its argument or deeper.
+
+Now we can define the context update relation. Let \(\sealcon\) be the sealing convention
 that maps a machine state to a set of components, typically the set of addresses below
-the stack pointer. The {\sc Default} rule applies only if neither of the others do.
-
-\begin{center}
-\begin{minipage}{.4\textwidth}
-  \judgment{}
-           {\(\mach,(\domMap,\depth),\{\callmap\} \stepstocon
-             \push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1\)}
-\end{minipage}
-\hspace*{0.1\textwidth}
-\begin{minipage}{.4\textwidth}
-  \judgment{}
-           {\(\mach,(\domMap,\depth),\{\retmap\} \stepstocon
-             \pop{\domMap}{\depth},\depth-1\)}
-\end{minipage}
-\end{center}
-
-\begin{center}
+the stack pointer. Similarly, \(\retcon ~ \mach ~ \mach'\) will be the return convention
+that relates the call state \(\mach\) to return states \(\mach'\) that are valid returns.
+The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
+and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
+
+The call rule pushes the components sealed by the sealing convention with the depth
+of the current caller, and determines the return target by partially applying the
+return convention to the machine state, and pushes that onto the targets stack.
+The return rule checks whether the attempted return actually manages to reach a
+return target, traversing down the stack until the first one it finds. Recall that
+we are not discussing well-bracketed control flow, so the return might skip over
+a waiting caller. When a target is reached, {\em all} sealed components at or above
+its depth are unsealed by the pop operation. If no target is reached, nothing is
+popped at all. These rules are given inductively, and apply in the listed order.
+
+\judgmenttwo[Call]
+            {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
+            {\(\depth = |\rts|\)}
+            {\(\mach,(\domMap,\rts) \stepsto_C
+              \push{\domMap}{\depth}{\sealcon ~ \mach},(\retcon ~ \mach)::\rts\)}
+
+\judgmentthree[ReturnFound]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+              {\(\mach \stepsto \mach' \in \target\)}
+              {\(\depth = |\rts|\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C
+                \pop{\domMap}{\depth},\rts\)}
+
+\judgmentthree[ReturnRec]
+              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \_,\rts'\)}
+              {\(\depth = |\rts'|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C
+                \pop{\domMap}{\depth},\rts'\)}
+
 \judgment[Default]
          {}
-           {\(\_,\context,\_ \stepstocon \context\)}
-\end{center}
-
-\subsection{Ultra-eager Integrity}
-
-Our first property will be {\em ultra eager}, meaning that it will be quantified over each step
-of execution from any initial state. The criterion for integrity is simple: sealed components do not
-change.
-
-\definition A system enjoys {\em ultra-eager integrity} if for any reachable state
+         {\(\_,\context \stepsto_C \context\)}
+
+When we build machine traces with this step function, we always know how deep in the
+call stack we are and which components are sealed. We may now proceed to state our properties.
+Our first property will be {\em stepwise stack integrity}, stepwise meaning that it will be
+quantified over each step of execution from any initial state. The criterion for integrity
+is simple: sealed components do not change.
+
+\definition A system enjoys {\em stepwise stack integrity} if for any reachable state
 \(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
@@
 
-\subsection{Eager Integrity}
-
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold when
-\(\depth' \geq \depth\). A system enjoys {\em eager integrity} if each reachable segment
-\(\MPCT \langle f_\depth \rangle\) satisfies \(\pi_m(\head(\MPCT)) ~ \component =
-\pi_m(\last(\MPCT)) ~ \component\) for all \(\component\) where \(\domMap ~ \component =
-\sealed{\_}\).
+Next we consider {\em lockstep stack confidentiality}. Here it behooves us to factor out
+the general notion of lockstep confidentiality independent of a particular context model.
 

### 2021-04-13 tex/main.tex
Another go at the abstract

@@
 run-time, and hardware mechanisms for guarding against corruption of stack
-memory.  But these mechanisms typically lack rigorous specifications;
-they rely instead on informal descriptions of
-the safe stack manipulation behaviors that they allow,
-or specific examples of the bad behaviors that they prevent.
-
-We propose a formal criterion \apt{criteria?} for stack safety, phrased in terms of
+memory.  But these mechanisms typically lack rigorous specifications,
+relying instead on informal descriptions of
+the safe stack manipulation behaviors that they allow
+and/or specific examples of the bad behaviors that they prevent.
+
+We propose formal criteria for stack safety, phrased in terms of
 language-based security: a combination of an integrity property (``each
@@
 property (``the callee's behavior is insensitive to differences in the
-caller's state''). These properties are sensible even in the presence of
-abnormal control flow.\apt{Is this important enough to mention here?}
-We further apply\apt{what does this mean? we didn't
-  invent it} a straightforward notion of
-``well-bracketed control flow'' that restricts control flow to the
-typical pattern of matched calls and returns.
-
-We show that our properties describe the behavior of the stack-safety
-``micro-policies'' of~\citet{DBLP:conf/sp/RoesslerD18}
-using property-based randomized testing. An eager micro-policy
-that catches violations immediately enforces our primary ``stepwise''
-properties, and we define weaker, more extensional ``observational''
-properties that are enforced even by a ``lazy'' micro-policy -- one
-that, for performance reasons, may catch violations after they occur
-but before they become visible.
-
-We argue \apt{do we?} that our properties are suited to a range of other enforcement
-mechanisms. The strict stepwise properties can describe mechanisms
-that catch errors immediately, as is typical. While the observational
-properties are closely linked to lazy micro-policies they are weak enough
-to be a lower bound that captures stack safety for any reasonable enforcement mechanism.
+caller's state'').
+% These properties are sensible even in the presence of
+% abnormal control flow.\apt{Is this important enough to mention here?}
+% BCP: I think not in the abstract, but maybe in the intro.
+% We further apply\apt{what does this mean? we didn't
+%  invent it} a straightforward notion of
+% ``well-bracketed control flow'' that restricts control flow to the
+% typical pattern of matched calls and returns.
+%
+We argue that these criteria match common intuitions (1) by showing that
+they are violated by a variety of intuitively unsafe examples and (2) by
+using property-based random testing to check that appropriate variants of
+the criteria are actually enforced by the stack-safety ``micro-policies''
+proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, Roessler and
+dehon's ``eager'' micro-policy, which catches violations as early as
+possible, enforces simple ``stepwise'' versions of our confidentiality and
+integrity properties, while their more performant ``lazy'' micro-policy
+enforces weaker, more extensional ``observational'' variants of the
+properties.
+
+\bcp{We should mention stack data and coroutines, I think?}
+
+% We argue \apt{do we?}\bcp{I guess we will not have time to.  Propose
+%   deleting this para.} that our properties are suited to a range of other
+% enforcement mechanisms. The strict stepwise properties can describe
+% mechanisms that catch errors immediately, as is typical. While the
+% observational properties are closely linked to lazy micro-policies they are
+% weak enough to be a lower bound that captures stack safety for any
+% reasonable enforcement mechanism.
 
@@
   a general criterion, but did not in fact demonstrate its generality (which is not
-  obvious). What has changed?}
-
-  \ifaftersubmission\bcp{should we mention stktokens here?}
-  \leo{Not sure. We almost never cite stuff in the abstract, do we?}
-  \bcp{Some people say not to, but I think their argument is weak: I've
-    added Nick/Andre above.  But I'm not sure we want to get into stkTokens
-    in the abstract.}
-  \fi
+  obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
+  para about the general story.  Does the rest now avoid making (or
+  implying) claims that we can't defend?}
+\bcp{Also, I think we should save
+  WBCF for the intro---I can't find a way to discuss it here that is brief
+  and not confusing/distracting.}
+
+  % \ifaftersubmission\bcp{should we mention stktokens here?}
+  % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
+  % \bcp{Some people say not to, but I think their argument is weak: I've
+  %   added Nick/Andre above.  But I'm not sure we want to get into stkTokens
+  %   in the abstract.}
+  % \fi
 \end{abstract}
@@
 %% comma separated list
+\ifcameraready
 \keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission
+\fi
 

### 2021-04-13 tex/main.tex
A few more things in the abstract

@@
 the criteria are actually enforced by the stack-safety ``micro-policies''
-proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, Roessler and
-dehon's ``eager'' micro-policy, which catches violations as early as
-possible, enforces simple ``stepwise'' versions of our confidentiality and
-integrity properties, while their more performant ``lazy'' micro-policy
+proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
+broken versions of these enforcement mechanisms).  Specifically, Roessler
+and Dehon's ``eager'' micro-policy, which catches violations as early as
+possible, enforces the basic ``stepwise'' versions of our confidentiality
+and integrity properties, while their more performant ``lazy'' micro-policy
 enforces weaker, more extensional ``observational'' variants of the
@@
 
-\bcp{We should mention stack data and coroutines, I think?}
+\bcp{We should mention stack data and coroutines, I think?  Maybe in the
+  context of an expectation-lowering hedge, like:}%
+For the sake of simplicity, most of our presentation assumes a RISC-V
+machine architecture with a basic subroutine calling convention, where
+parameters are passed in registers.  We also show how stack-allocated
+parameter passing and coroutines can be accommodated.
 

### 2021-04-13 tex/main.tex
INtro

@@
   context of an expectation-lowering hedge, like:}%
-For the sake of simplicity, most of our presentation assumes a RISC-V
+For the sake of concreteness, most of our presentation assumes a RISC-V
 machine architecture with a basic subroutine calling convention, where
 parameters are passed in registers.  We also show how stack-allocated
-parameter passing and coroutines can be accommodated.
+parameter passing and coroutines can be accommodated.  \bcp{this is also a
+  good place to admit that our definitions are tuned for micropolicies, if
+  we can find a succinct way to say it.}
 
@@
 that range from leakage or corruption of private stack data to
-control-flow hijacking. To detect or prevent such attacks, a profusion of
+control-flow hijacking. To deter or detect such attacks, a profusion of
 software and hardware protections have been proposed,
@@
 %
-The protections offered by such mechanisms are generally described in terms
-of a set of concrete examples of attacks that they can prevent, such as corruption
-of return addresses, buffer overflows, or use of uninitialized variables. This leaves the
-task of formulating a more general characterization to the reader's intuition.
-Moreover, the details of these mechanisms can be quite intricate
-and often involve trading off protection and efficiency.
+The protections offered by such mechanisms are commonly described in terms
+of concrete examples of attacks that they can prevent, such as corruption
+of return addresses, buffer overflows, or use of uninitialized variables,
+leaving a more abstract characterization to the reader's intuition.
+But these mechanisms can be quite intricate
+and often require trading off protection and efficiency.
 Thus, it would be useful to have a precise, generic, and formal
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-The most abstract characterization of which we are aware is in
+The most abstract characterization stack safety of which we are aware is in
 recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
@@
 executing, and WBCF says that callees always
-return to the point just after their call (or don't return at all).
-\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
-in which they visibly hold by construction; an enforcement mechanism (in their case, hardware
+return to the point just after their call (if they return at all).
+\citet{Skorstengaard+19} formalize these properties by defining an idealized
+machine in which they hold {\em by construction}; an enforcement mechanism (in their case, hardware
 capabilities)
-can then be shown to provide stack safety by giving a fully abstract embedding function
-from the idealized machine to an ordinary machine equipped with the enforcement.
+can then be shown to guarantee stack safety by exhibiting a fully abstract
+embedding function
+from the idealized machine to an ordinary machine equipped with this
+enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
+  understand that sentence.}
 %%
@@
 
-In this work, we take the (informal) definitions of WBCF\apt{or maybe not?}  and LSE as a starting
+In this work, we take the (informal) definitions of WBCF\apt{or maybe
+  not?}\bcp{Definitely not, if we're following the approach I proposed.
+  WBCF should be discussed second, and we should say that, while it is a
+  very reasonable property to want, we do not consider
+it as part of stack safety (an enforcement mechanism could be intuitively
+``safe'' in the information-flow sense while not enforcing it), but rather
+as a separate condition that can be stated straightforwardly using the same
+technical machinery.}  and LSE as a starting
 point, but choose a different approach to formalizing them, based on the
@@
 trace-based properties of \emph{stepwise stack integrity} and
-\emph{lockstep stack confidentiality}.\apt{Could we just say ``stepwise'' for both?}
+\emph{lockstep stack confidentiality}.\apt{Could we just say ``stepwise''
+  for both?}\bcp{+1}
 In applying these security concepts, we treat each caller as needing protection
@@
 deviation is recognized as a violation of confidentiality.
-Together, these properties capture the rough intuition of LSE, but are somewhat stronger:
+Together, these properties capture roughly the same intuition as LSE, but are somewhat stronger:
 confidentiality contains temporal protections that LSE does not address.\apt{Maybe want to
-  say a little more about this here, since it is a novelty wrt/ existing work?}
+  say a little more about this here, since it is a novelty wrt/ existing work?}\bcp{+1}
 
@@
 
-Part of the challenge in fully\apt{omit?} characterizing stack safety is that in a low-level machine
+One challenge in characterizing stack safety formally is that in a low-level machine
 without built-in stacks, we are at the mercy of architectural
@@
 as much as possible, so that our model can be applied to a wide range of architectures
-and enforcement mechanisms.
-
-\apt{Intra-paragraph flow seemed awkward.}
-A second challenge is
+and enforcement mechanisms.  \bcp{As APT has worried, this kind of claims of
+generality may be getting us onto thin ice.  Maybe remove the second
+sentence (and expand the first)?  }
+
+A second challenge is that
 % the diversity of enforcement mechanisms APT:already implied above
@@
 while arguing that they still capture stack safety in a meaningful sense.
-\apt{Remainder of this paragraph might be better \emph{after} the R+DH para below.}
+\apt{Remainder of this paragraph might be better \emph{after} the R+DH para below.}\bcp{+1}
 This calls \apt{not
-  immediately obvious why. is this true for anything other than laziness?}
+  immediately obvious why. is this true for anything other than
+  laziness?}\bcp{Agreed.  If we move this later, we can just say that we can
+capture the more performant policies with a more extensional variant of the
+criteria that is a bit more complex but usefully weaker.  }
 for an {\em extensional}
@@
 
+\bcp{This paragraph doesn't apply to the lazy properties, so maybe it should
+be moved before they are introduced?}%
 To illustrate the applicability of these properties, we perform randomized testing to
@@
 the real one. If the testing can differentiate correct implementations from buggy ones,
-it is an argument in favor of their correctness.
+it is an argument in favor of their correctness. \bcp{Hmm--don't understand
+  that sentence}
 %\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
@@
 Our initial description of these properties assumes a single simple stack with no sharing between callers and
-callees: all parameters and return values are passed in registers.  However,
-we can refine it to allow passing of scalar stack data, granting a callee
+callees: all parameters and return values are passed in registers.  Later,
+we show how to refine it to allow passing of scalar stack data, granting a
+callee
 access to the caller's data. (We do not model more sophisticated sharing, because to do
@@
 Finally, we define integrity and confidentiality properties for a coroutine system featuring
-finitely-bounded stacks.
+finitely-bounded\bcp{What's that?} stacks.
 Though simplistic, this extension is a step toward concurrency and demonstrates that

### 2021-04-13 tex/main.tex
Pretty sure I need to merge

--- 2021-04-13 tex/main.tex
@@
 run-time, and hardware mechanisms for guarding against corruption of stack
-memory.  But these mechanisms typically lack rigorous specifications,
-relying instead on informal descriptions of
-the safe stack manipulation behaviors that they allow
-and/or specific examples of the bad behaviors that they prevent.
+memory.  But these mechanisms typically lack rigorous specifications;
+they rely instead on informal descriptions of
+the safe stack manipulation behaviors that they allow,
+or specific examples of the bad behaviors that they prevent.
 
@@
 property (``the callee's behavior is insensitive to differences in the
-caller's state'').
-% These properties are sensible even in the presence of
-% abnormal control flow.\apt{Is this important enough to mention here?}
-% BCP: I think not in the abstract, but maybe in the intro.
-% We further apply\apt{what does this mean? we didn't
-%  invent it} a straightforward notion of
-% ``well-bracketed control flow'' that restricts control flow to the
-% typical pattern of matched calls and returns.
-%
-We argue that these criteria match common intuitions (1) by showing that
-they are violated by a variety of intuitively unsafe examples and (2) by
-using property-based random testing to check that appropriate variants of
-the criteria are actually enforced by the stack-safety ``micro-policies''
-proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, Roessler and
-dehon's ``eager'' micro-policy, which catches violations as early as
-possible, enforces simple ``stepwise'' versions of our confidentiality and
-integrity properties, while their more performant ``lazy'' micro-policy
-enforces weaker, more extensional ``observational'' variants of the
-properties.
-
-\bcp{We should mention stack data and coroutines, I think?}
-
-% We argue \apt{do we?}\bcp{I guess we will not have time to.  Propose
-%   deleting this para.} that our properties are suited to a range of other
-% enforcement mechanisms. The strict stepwise properties can describe
-% mechanisms that catch errors immediately, as is typical. While the
-% observational properties are closely linked to lazy micro-policies they are
-% weak enough to be a lower bound that captures stack safety for any
-% reasonable enforcement mechanism.
+caller's state''). We show that our properties describe the behavior of the stack-safety
+``micro-policies'' of~\citet{DBLP:conf/sp/RoesslerD18}
+using property-based randomized testing. An eager micro-policy
+that catches violations immediately enforces our primary ``stepwise''
+properties, and we define weaker, more extensional ``observational''
+properties that are enforced even by a ``lazy'' micro-policy -- one
+that, for performance reasons, may catch violations after they occur
+but before they become visible.
+
+These two classes of properties are interesting points along a spectrum
+of notions of integrity and confidentiality. The strict stepwise properties can
+describe mechanisms that catch errors immediately, the expected behavior of most
+enforcement techniques. While the observational properties are closely linked to
+lazy micro-policies, they also have the benefit of being highly extensional,
+and therefore approaching a plausible minimum specification for stack safety in this
+setting.
 
@@
   a general criterion, but did not in fact demonstrate its generality (which is not
-  obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
-  para about the general story.  Does the rest now avoid making (or
-  implying) claims that we can't defend?}
-\bcp{Also, I think we should save
-  WBCF for the intro---I can't find a way to discuss it here that is brief
-  and not confusing/distracting.}
-
-  % \ifaftersubmission\bcp{should we mention stktokens here?}
-  % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
-  % \bcp{Some people say not to, but I think their argument is weak: I've
-  %   added Nick/Andre above.  But I'm not sure we want to get into stkTokens
-  %   in the abstract.}
-  % \fi
+  obvious). What has changed?}
+\sna{It's a delicate line we have to walk, because while our generality may not be
+  easily demonstrable, {\em lack} of generality, as BCP pointed out, is also not
+  demonstrable. So now I've just taken it out, and I'll plan to talk in the
+  introduction about how we hope to be general later.}
+
+  \ifaftersubmission\bcp{should we mention stktokens here?}
+  \leo{Not sure. We almost never cite stuff in the abstract, do we?}
+  \bcp{Some people say not to, but I think their argument is weak: I've
+    added Nick/Andre above.  But I'm not sure we want to get into stkTokens
+    in the abstract.}
+  \fi
 \end{abstract}
@@
 %% comma separated list
-\ifcameraready
 \keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission
-\fi
 
@@
 
-In this work, we take the (informal) definitions of WBCF\apt{or maybe not?}  and LSE as a starting
-point, but choose a different approach to formalizing them, based on the
-technical framework of language-based security~\citep{??}. We define
-trace-based properties of \emph{stepwise stack integrity} and
-\emph{lockstep stack confidentiality}.\apt{Could we just say ``stepwise'' for both?}
+In this work, we take the (informal) definition of LSE as a starting
+point, but choose a different approach to formalizing it, based on the
+technical framework of language-based security~\citep{??}. 
+Our core properties are \emph{stepwise stack integrity} and
+\emph{stepwise stack confidentiality}.
 In applying these security concepts, we treat each caller as needing protection
@@
 confidentiality property applies to many nested subtraces within a single execution.
-\sna{How novel is this nesting?} \apt{Lost the reasoning principle that was here?}
+\sna{How novel is this nesting?}
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
-and its {\em stepwise} interpretation formalizes this very naturally: a caller's local data
+motivated by the reasoning principle that local data persists across a call. Its
+{\em stepwise} interpretation formalizes this very naturally: a caller's local data
 is never modified during a call. {\em Confidentiality} says that a caller's data is protected
@@
 {\em noninterference}-style property that, rather than comparing pairs of whole program traces,
-compares a pair of subtraces for each call, nested within other calls. The term
-{\em lockstep} describes the behavior of these pairs: they step together, and any
-deviation is recognized as a violation of confidentiality.
-Together, these properties capture the rough intuition of LSE, but are somewhat stronger:
-confidentiality contains temporal protections that LSE does not address.\apt{Maybe want to
-  say a little more about this here, since it is a novelty wrt/ existing work?}
-
-Since our stack safety occupies the same conceptual space as LSE, we should consider how it
-relates to WBCF.\apt{Or maybe not. This disucssion seems too detailed for this spot in the intro.
-  While it is good to note that our integrity/confidentiality properties make sense without
-  WBCF-or-similar, LSE is also described independently from WBCF, so there's no novelty in that.}
-First, we observe that some degree of bracketed control flow is implicit
-in any notion of stack safety, simply because calls and returns must be recognizeable so
-that callers and callees are distinguished. We treat control flow that creates an ambiguity between caller and
-callee\apt{this is obscure}  as keeping the callee in control, and therefore such behavior can never give access
-to the caller's data. An example would be if the callee attempts to ``return'' to the
-wrong location, hoping to execute code on behalf of the caller: it may make the jump
-without violating stack safety, but it gains no privileges on the caller's data.
-Alternatively a callee might return past its immediate caller, to somewhere deeper in the
-stack, and stack safety does not inherently rule this out. We therefore define WBCF as an additional
-restriction \apt{on what? in what way is it formalized differently than by Skorstengaard?},
-describing the standard control flow principle that a return pops exactly one frame
-from the stack.
-\sna{We got side-tracked a bit from talk of entry integrity, and Andrew made some good points
-  about it, so let's not rule that out just yet. But I focused on returns because there's a
-  clear story already.}
-
-Part of the challenge in fully\apt{omit?} characterizing stack safety is that in a low-level machine
+compares a pair of subtraces for each call, nested within other calls. Once again
+{\em stepwise} describes the granularity of enforcement: at every step of a call,
+changes in the state are independent of the caller's data.
+
+Together, these properties capture the rough intuition of LSE, but are somewhat stronger.
+First, we model returns as returns to a valid return target; an invalid return
+is treated as a mere jump within the callee, and does not grant access to the caller's
+data. So, even without WBCF, our stack safety properties rule out many control flow attacks.
+The attacker in such an instance is permitted to ``return'' so long as it avoids
+violating the integrity and confidentiality of the caller. Second, confidentiality
+contains temporal protections that LSE does not implicitly address.
+\apt{Maybe want to say a little more about this here, since it is a novelty wrt/ existing work?}
+\sna{TODO: Say a little more, but make sure it's accurate first.}
+
+One challenge in characterizing stack safety is that in a low-level machine
 without built-in stacks, we are at the mercy of architectural
-choices that do not relate to the abstract notion of stack safety\apt{a bit vague. what kind of choices?}.
-We abstract over these choices
+choices that do not relate to the abstract notion of stack safety. Some implementations
+may have only a stack pointer while others add a frame pointer, in some systems return
+addresses and data live in separate stacks, and so on. We abstract over these choices
 as much as possible, so that our model can be applied to a wide range of architectures
@@
 
-\apt{Intra-paragraph flow seemed awkward.}
-A second challenge is
-% the diversity of enforcement mechanisms APT:already implied above
+A second challenge is that
 enforcement mechanisms may weaken their protections that in the name of efficiency,
 while arguing that they still capture stack safety in a meaningful sense.
-\apt{Remainder of this paragraph might be better \emph{after} the R+DH para below.}
-This calls \apt{not
-  immediately obvious why. is this true for anything other than laziness?}
-for an {\em extensional}
-characterization of stack safety that describes the externally observable behavior
-of a stack-safe system. In contract, our stepwise and lockstep properties are {\em intensional},
-focusing on the internal behavior.
-
 For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
@@
 functions and deciding exactly which accesses are allowed.
-% APT: seem repetitive:
-%, to enable different ways
-%of maintaining the connection between an activation identity and the memory it owns.
 These vary in their efficiency and in the spatial and temporal precision of their
 security guarantees.
-The \emph{Depth Isolation} policy requires a strict mapping from a function
-to its frame from the start of the call to its end\apt{unclear. why not simply say
-  what depth is?}, and permits the function to write
-only into that frame.
-This seems simple and intuitively correct, but it can be expensive to implement the strict
-mapping. Conversely, the \emph{Lazy Tagging and Clearing} policy improves
+The \emph{Depth Isolation} policy prevents a callee from writing outside its frame
+at all, which is simple and intuitively correct, but it must manually clear the tags
+on the frame on each return, which is very expensive. Conversely, the
+\emph{Lazy Tagging and Clearing} policy improves
 performance by assigning locations to the active function when they are written, which
 permits  ``illicit'' writes to inaccessible stack frames so long as the
-written locations are not subsequently read by their ``real'' owners. The intuitive justification
-for this form of laziness is that any genuinely dangerous writes will ultimately be
-detected before they can affect program behavior.
+written locations are not subsequently read by their ``real'' owners. Intuitively,
+this form of laziness is justified if any dangerous writes will
+ultimately be detected before they can affect program behavior.
+
+But what does ``affecting program behavior'' really mean? This micro-policy
+calls for a second characterization of stack safety in terms of the externally
+observable behavior of a stack-safe system. Such a characterization will have the
+additional benefit of being more {\em extensional}, concerned primarily with the
+external behavior of the system. In contract, our stepwise properties are {\em intensional},
+focusing on internal details that may not actually matter to overall security.
 
@@
 makes confidentiality extensional in the same way, although we do not know of a real
-enforcement mechanism that weakens confidentiality in this way.\apt{why bother mentioning it then?}
+enforcement mechanism that weakens confidentiality in this way.
+\apt{why bother mentioning it then?}
+\sna{If a relatively extensional characterization is a worthwhile contribution
+  in its own right, then it's worth mentioning that we have one. If not, then indeed,
+  we shouldn't mention it -- or get too excited about extensionality in general, because
+  lockstep isn't extensional.}
 Together, these form an {\em observational stack safety} that is highly extensional.
@@
 \item
-  We adapt formulations of integrity and confidentiality from
-  language-based security
-  to formalize a parameterized {\em stepwise stack-safety property} \apt{lockstep?} and argue that
-  it captures the intuitive concepts of local stack encapsulation, well-bracketed control
-  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
+  We adapt formulations of integrity and confidentiality from language-based security
+  to formalize a parameterized {\em stepwise stack-safety} property and argue that it
+  captures the intuitive concepts of stack safety~(\cref{sec:lse}). Our model extends to
+  argument passing on the stack and to a simple coroutine system.
 \item
@@
   that this stepwise property is actually enforced by the Depth Isolation policy
-  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
+  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:testing}).
 \item
@@
   enhanced with unique tags on each activation~(\cref{sec:lazy}).
-\item
-  We extend our model so that it can be instantiated in systems that pass arguments
-  on the stack~(\cref{sec:??}) and in those with simple coroutines~(\cref{sec:??}).
 \end{itemize}
@@
 assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter \(\PCname\) and stack pointer \(\rsp\)\apt{do we really
-  need these except for WBCF?}; other
+machine with a program counter \(\PCname\); other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
@@
 which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention} and model it as a relation between machine states
-and addresses. From any call, those addresses that the sealing convention relates to the
-state\apt{rephrased, but still obscure} will be protected until the return.
+this the {\em sealing convention}, which maps a machine state to a set of addresses.
+If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
+which addresses must be protected until the return.
 
@@
 We take a very strong threat model in which both caller and callee may execute
-arbitrary machine code. This means that even the defender might be ill-behaved
-to the point that it attempts to undermine its own protections, and we expect
-enforcement mechanisms to ensure that it fails.\apt{obscure}
-Hardware and timing attacks are out of scope, however.
+arbitrary machine code. Hardware and timing attacks are out of scope, however.
 
@@
 
-\section{Stack Safety, Intuitively}
-\apt{section title is a bit sad since we're trying to convert intuition to formal spec!}
+\section{Demonstrating Stack Safety}
 \label{sec:running-example}
@@
 
-%\paragraph*{Control Flow and Awkward Examples}
-
-%\begin{figure}
-%  \centering
-%  \begin{minipage}{.4\textwidth}
-%\begin{verbatim}
-%int stash = 0;
-
-%int main() {
-%0:     int x = 1;
-%1[C]:  f();
-%2:     x = -x;
-%3[C]:  f();
-%4[R]:  return x;
-%}
-%\end{verbatim}
-%  \end{minipage}
-%  \begin{minipage}{.4\textwidth}
-%\begin{verbatim}
-%void f() {
-%10:     if(!stash) {
-%        // store return pointer
-%11:       stash = $SP-1;
-%12:     } else {
-%13:       $SP-1 = stash;
-%14:       stash = 0;
-%15[R]:    return;
-%        }
-%}
-%\end{verbatim}
-%  \end{minipage}
-%\label{fig:WBCF1}
+\paragraph*{Control Flow Attacks}
+\begin{figure}
+  \centering
+  \begin{minipage}{.4\textwidth}
+\begin{verbatim}
+int *stash = 0;
+
+int main() {
+0:     int x;
+1:     x = 1;
+2[C]:  f();
+3:     x = -x;
+4[C]:  f();
+5[R]:  return x;
+}
+
+void f() {
+10:     if(!stash) {
+        // store return pointer
+11:       stash = $SP-1;
+12:     } else {
+13:       $SP-1 = stash;
+14:       stash = 0;
+        }
+15[R]:    return;
+}
+\end{verbatim}
+  \end{minipage}
+\label{fig:controlflow}
  %  \begin{subfigure}[t]{.4\textwidth}
@@
  %int counter = 0;
-%int stash = 0;
-%
-%int main() {
-%0:     int x = 1;
-%1[C]:  f();
-%2:     x = -x;
-%3[C]:  f();
-%4[R]:  return x;
-%}
-%
-%void f() {
-%10:     switch(counter++) {
-%11:       case 0:
-%12[R]:      return;
-%13:       case 1:
-%14:         stash = $SP-1;
-%15[C]:      main();
-%16:         return;
-%17:       default:
-%18:         $SP-1 = stash;
-%19[R]:      return;
-%        }
-%}
+@@ -1194,42 +1090,43 @@ the program counters will differ, breaking the lockstep.
  %\end{verbatim}
@@
  %  \end{subfigure}
-%  \caption{Control Flow Attacks}
-%  \label{fig:WBCF}
-%\end{figure}
-
-%Example \ref{fig:WBCF} shows how control flow attacks can compromise integrity in principle
-%even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
-%global variable and retrieves it the second time it is called, returning to the first call
-%point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
-%line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.
-
-%To prevent this, we define Well-bracketed Control Flow. Throughout the above examples, we
-%have implicitly tracked the depth of the current call in order to seal memory at the correct
-%depth. We also have a clear sense of where the entry points to functions are -- the top line,
-%in our examples, and another instruction annotation in practice. Well-bracketed Control Flow
-%holds if:
-
-%\begin{itemize}
-%\item All calls jump to entry points ({\em entry integrity})
-%\item The depth counter decreases by exactly one on a return ({\em non-skipping})
-%\item From a call at depth \(d\), the next state to have depth \(d\) matches the stack pointer,
-%  and the program counter is at the instruction following the jal ({\em return integrity})
-%\end{itemize}
-
-%Non-skipping may seem strange -- surely the depth counter ought to decrease by one at a time
-%as part of the property definition, and the problem comes when the return is to a shallower
-%call's return pointer, violating return integrity? Yet some weaker enforcement mechanisms,
-%such as those in which returns are implemented as invocation of a Cheri object capability,
-%do allow skipping when adversarial code acts unreasonably by leaking its own return capability,
-%reasoning that this is acceptable as long as only the unreasonable function is skipped
-%(e.g. \citep{Skorstengaard+19b}).\apt{???} A variant of our model supports this weaker form of WBCF,
-%but with tag-based policies we can enforce the non-skipping version entirely.\apt{??}
-%\sna{Cite future work for details when the relevant section is finished.}
+-%  \caption{Control Flow Attacks}
+-%  \label{fig:WBCF}
+-%\end{figure}
++  \caption{Control Flow Attacks}
++  \label{fig:WBCF}
++\end{figure}
 
@@
 \section{Randomized Testing}
+\label{sec:testing}
 
@@
     \ifaftersubmission\apt{More blatently, static authorities doesn't work for recursive calls.}\fi
-
-\section{Extension to Cheri}
-\label{sec:cheri}
-
-%\section{Extension to Software-based Enforcement}
 

### 2021-04-13 tex/main.tex

@@
 run-time, and hardware mechanisms for guarding against corruption of stack
-memory.  But these mechanisms typically lack rigorous specifications;
-they rely instead on informal descriptions of
-the safe stack manipulation behaviors that they allow,
-or specific examples of the bad behaviors that they prevent.
+memory.  But these mechanisms typically lack rigorous specifications,
+relying instead on informal descriptions of
+the safe stack manipulation behaviors that they allow
+and/or specific examples of the bad behaviors that they prevent.
 
@@
 property (``the callee's behavior is insensitive to differences in the
-caller's state''). We show that our properties describe the behavior of the stack-safety
-``micro-policies'' of~\citet{DBLP:conf/sp/RoesslerD18}
-using property-based randomized testing. An eager micro-policy
-that catches violations immediately enforces our primary ``stepwise''
-properties, and we define weaker, more extensional ``observational''
-properties that are enforced even by a ``lazy'' micro-policy -- one
-that, for performance reasons, may catch violations after they occur
-but before they become visible.
-
-These two classes of properties are interesting points along a spectrum
-of notions of integrity and confidentiality. The strict stepwise properties can
-describe mechanisms that catch errors immediately, the expected behavior of most
-enforcement techniques. While the observational properties are closely linked to
-lazy micro-policies, they also have the benefit of being highly extensional,
-and therefore approaching a plausible minimum specification for stack safety in this
-setting.
+caller's state'').
+% These properties are sensible even in the presence of
+% abnormal control flow.\apt{Is this important enough to mention here?}
+% BCP: I think not in the abstract, but maybe in the intro.
+% We further apply\apt{what does this mean? we didn't
+%  invent it} a straightforward notion of
+% ``well-bracketed control flow'' that restricts control flow to the
+% typical pattern of matched calls and returns.
+%
+We argue that these criteria match common intuitions (1) by showing that
+they are violated by a variety of intuitively unsafe examples and (2) by
+using property-based random testing to check that appropriate variants of
+the criteria are actually enforced by the stack-safety ``micro-policies''
+proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
+broken versions of these enforcement mechanisms).  Specifically, Roessler
+and Dehon's ``eager'' micro-policy, which catches violations as early as
+possible, enforces the basic ``stepwise'' versions of our confidentiality
+and integrity properties. Their more performant ``lazy'' micro-policy
+does not, but inspires weaker, more extensional ``observational'' variants of the
+properties. While we do not rule out other enforcement mechanisms, our
+definitions are strongly informed by micropolicy systems, especially our
+notion of observational integrity.
+\sna{Seems like the best place to admit we're tied to micropolicies is
+  while discussing micropolicies.}
+
+\bcp{We should mention stack data and coroutines, I think?  Maybe in the
+  context of an expectation-lowering hedge, like:}%
+For the sake of concreteness, most of our presentation assumes a RISC-V
+machine architecture with a basic subroutine calling convention, where
+parameters are passed in registers. We also show how to accomodate
+stack-allocated parameter passing and coroutines.
+
+\bcp{this is also a
+  good place to admit that our definitions are tuned for micropolicies, if
+  we can find a succinct way to say it.}
+\sna{See above.}
+
+% We argue \apt{do we?}\bcp{I guess we will not have time to.  Propose
+%   deleting this para.} that our properties are suited to a range of other
+% enforcement mechanisms. The strict stepwise properties can describe
+% mechanisms that catch errors immediately, as is typical. While the
+% observational properties are closely linked to lazy micro-policies they are
+% weak enough to be a lower bound that captures stack safety for any
+% reasonable enforcement mechanism.
 
@@
   a general criterion, but did not in fact demonstrate its generality (which is not
-  obvious). What has changed?}
-\sna{It's a delicate line we have to walk, because while our generality may not be
-  easily demonstrable, {\em lack} of generality, as BCP pointed out, is also not
-  demonstrable. So now I've just taken it out, and I'll plan to talk in the
-  introduction about how we hope to be general later.}
-
-  \ifaftersubmission\bcp{should we mention stktokens here?}
-  \leo{Not sure. We almost never cite stuff in the abstract, do we?}
-  \bcp{Some people say not to, but I think their argument is weak: I've
-    added Nick/Andre above.  But I'm not sure we want to get into stkTokens
-    in the abstract.}
-  \fi
+  obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
+  para about the general story.  Does the rest now avoid making (or
+  implying) claims that we can't defend?}
+\bcp{Also, I think we should save
+  WBCF for the intro---I can't find a way to discuss it here that is brief
+  and not confusing/distracting.}
+
+  % \ifaftersubmission\bcp{should we mention stktokens here?}
+  % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
+  % \bcp{Some people say not to, but I think their argument is weak: I've
+  %   added Nick/Andre above.  But I'm not sure we want to get into stkTokens
+  %   in the abstract.}
+  % \fi
 \end{abstract}
@@
 \newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}
-\newcommand{\memoryaddrs}[1]
+\newcommand{\memoryaddrs}[2][]
   {
-    \MemoryLabel{#1}{2em}{\SP}
+    \MemoryLabel{#2}{2em}{\SP #1}
   }
@@
 
-In this work, we take the (informal) definition of LSE as a starting
-point, but choose a different approach to formalizing it, based on the
-technical framework of language-based security~\citep{??}. 
+In this work, we also distinguish between protecting data (like LSE) and
+control flow (with WBCF), and formalize a data protection property
+based on the technical framework of language-based security~\citep{??}.
 Our core properties are \emph{stepwise stack integrity} and
-\emph{stepwise stack confidentiality}.
+\emph{stepwise stack confidentiality}. WBCF is a desireable property, but
+we can express safety in an information-flow sense without reference to it,
+even safety from attacks that hijack control flow.
+
 In applying these security concepts, we treat each caller as needing protection
@@
 \apt{Maybe want to say a little more about this here, since it is a novelty wrt/ existing work?}
-\sna{TODO: Say a little more, but make sure it's accurate first.}
-
-One challenge in characterizing stack safety is that in a low-level machine
+\bcp{+1}\sna{TODO: Say a little more, but make sure it's accurate first.}
+
+One challenge in characterizing stack safety formally is that in a low-level machine
 without built-in stacks, we are at the mercy of architectural
@@
 addresses and data live in separate stacks, and so on. We abstract over these choices
-as much as possible, so that our model can be applied to a wide range of architectures
-and enforcement mechanisms.
+as much as possible.
 
@@
   need a caveat about how it's harder to test?}\apt{As per email discussion,
-  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and Clearing,
-  by loosening the temporal safety properties?}
-
-Our initial description of these properties assumes a single simple stack with no sharing between callers and
-callees: all parameters and return values are passed in registers.  However,
-we can refine it to allow passing of scalar stack data, granting a callee
-access to the caller's data. (We do not model more sophisticated sharing, because to do
-so realistically would require a separate protection scheme, such as a capability
-model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
-a separate policy outside the scope of stack safety. A tag-based approach can easily
-compose multiple such policies.)
+  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and
+  Clearing, by loosening the temporal safety properties?}
+\sna{I'm pretty sure for the main policy it's just lazy integrity, no confidentiality.
+  There was a variant Nick and Andre gave that would restrict the leaks to only be between
+  instances of the same function, so that's the same property under a weaker attacker model,
+  where we don't protect a function from itself.}
+
+Our initial description of these properties assumes a single simple stack with no sharing
+between callers and callees: all parameters and return values are passed in registers.  Later,
+we show how to refine it to allow passing of scalar stack data, granting a
+callee access to the caller's data. We do not model more sophisticated sharing,
+such as taking the address of a variable and treating it like a pointer. Doing so without
+protecting the variable is not interesting, and protecting would be better suited to
+a different kind of enforcement, such as a memory safety micro-policy or Cheri-style capability.
+\bcp{This is all a bit mysterious} \sna{Any less mysterious now?}
+
 \rb{So no pointer passing? Pass by reference is now discussed at the end of
@@
 Finally, we define integrity and confidentiality properties for a coroutine system featuring
-finitely-bounded stacks.
+a static layout, in which each coroutine's stack is constrained to a fixed region of memory.
 Though simplistic, this extension is a step toward concurrency and demonstrates that
@@
 
-In summary, we offer the following contributions: \bcp{Section numbers are
-  wonky---why just 7.1, where is 8, etc.}
+In summary, we offer the following contributions:
 \begin{itemize}
@@
 to read the stack pointer ({\tt \$SP}) directly and use
-it to form pointers.\apt{I've added some strategic *'s}
+it to form pointers.
 Call and return annotations are marked {\tt[C]} and {\tt[R]}
@@
 
-This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
+This form of confidentiality we will term {\em stepwise confidentiality}, defined informally
 as: for any state at the entry to a function, for any variant over the values in the stack,
-the primary state and its variant will step in lockstep until they return. Any system components\apt{first use of this?}
-whose values change will be identical between the primary and variant traces. This includes not
-just memory, but registers as well, so that returning a secret is a violation. And when we
-extend the model with a notion of external output, it\apt{what?}  will include any values that
-emitted.
-
-We alternatively could imagine a stepwise form of confidentiality in which we explicitly
+the primary state and its variant will step in lockstep until they return. To step in lockstep
+means that if both states step -- that is, neither the primary nor the variant failstops --
+any {\em system component} (such as memory cell, register, program counter) that changes
+in one step has a matching value at the end of the other step. And when we
+extend the model with a notion of external output, values that are outputted will need
+to match as well.
+
+We alternatively could imagine a version of stepwise confidentiality in which we explicitly
 track which addresses may be read (i.e., which have been initialized), and at each step
 guarantee that the step alters state identically to a step from a variant state. Then we
-wouldn't need to divide up the execution by calls. But we can see in this example that taking
-a single variant at the start of the call effectively captures initialization: when {\tt y} is
-initialized, the primary and the variant agree on its value, so it can now be safely read.
-This treatment is roughly in line with more standard uses of noninterference, which typically
-vary states at the starts of whole program runs.\apt{This is all a bit cryptic.}
-
-Together, stepwise integrity and lockstep confidentiality form the stepwise version of
-local state encapsulation.
+wouldn't need to divide up the execution by calls. But the approach we take is more
+consistent with a typical noninterference approach, in that it varies the memory once
+at the beginning and trace changes through the execution.
 
@@
 \begin{tabular}{l r | l}
-  \circled{\tt 01} &
+  {\tt 01} &
   \multicolumn{2}{c}{
@@
   } \\
-  \circled{\tt 11} &
+  {\tt 11} &
   \memoryaddrs{16em}
@@
   \\
-  \circled{\tt 12} &
+  {\tt 12} &
   \memoryaddrs{16em}
@@
   \\
-  \circled{\tt 02} &
+  {\tt 02} &
   \multicolumn{2}{c}{
@@
   \\
-  \circled{\tt 21} &
+  {\tt 21} &
   \memoryaddrs{16em}
@@
   \\ &
-  \circled{\bf 22}
+  {\bf 22}
   ~$\cdots$
   &
-  \circled{\bf 24}
+  {\bf 24}
   ~$\cdots$
@@
 
-\sna{Not terribly happy with the circles yet; the idea is that instead of a meta-level
-  label for the line numbers we could treat the PC as a part of the system we're showing.}
-
 \paragraph*{Control Flow Attacks}
@@
   \centering
-  \begin{minipage}{.4\textwidth}
+  \begin{subfigure}{.4\textwidth}
 \begin{verbatim}
@@
 12:     } else {
-13:       $SP-1 = stash;
+13:       *($SP-1) = stash;
 14:       stash = 0;
@@
 \end{verbatim}
-  \end{minipage}
+  \end{subfigure}
+  \begin{subfigure}{.5\textwidth}
+    \begin{tabular}{l l l}
+      {\tt 2} &
+      \memoryaddrs{8em}
+      \memory{4}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{1} \\
+      {\tt 15} &
+      \memoryaddrs{12em}
+      \memory{1}{\mainsealc}[Seal(0)]%
+      \memory{1}{\goodc}[RetPtr]%
+      \memory{2}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{1}
+      \MemoryLabel{-15em}{0.75em}{3}
+      \vspace{.5em} &
+      \memory[3em]{1}{\mainsealc}[Target]%
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 3}
+      \\
+      {\tt 3} &
+      \memoryaddrs{8em}
+      \memory{4}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{1} \\
+      {\tt 10} &
+      \memoryaddrs{12em}
+      \memory{1}{\mainsealc}
+      \memory{1}{\goodc}
+      \memory{2}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{-1}
+      \MemoryLabel{-15em}{0.75em}{5} &
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
+      \\
+      {\tt 15} &
+      \memoryaddrs{12em}
+      \memory{1}{\mainsealc}
+      \memory{1}{\badc}
+      \memory{2}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{-1}
+      \MemoryLabel{-15em}{0.75em}{3} &
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
+      \\
+      {\tt 3} &
+      \memoryaddrs{8em}
+      \memory{1}{\mainsealc}
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{-1} &
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
+      \\
+      {\tt 4} &
+      \memoryaddrs{8em}
+      \memory{1}{\mainsealc}
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{\bf 1} &
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5} \\
+    \end{tabular}
+  \end{subfigure}
   \label{fig:controlflow}
- %  \begin{subfigure}[t]{.4\textwidth}
- %\begin{verbatim}
- %int counter = 0;
-@@ -1194,42 +1090,43 @@ the program counters will differ, breaking the lockstep.
- %\end{verbatim}
- %\label{fig:WBCF2}
- %  \end{subfigure}
--%  \caption{Control Flow Attacks}
--%  \label{fig:WBCF}
--%\end{figure}
-+  \caption{Control Flow Attacks}
-+  \label{fig:WBCF}
-+\end{figure}
+  \caption{A Control Flow Attack}
+\end{figure}
+
+From the classic stack smashing attack to return oriented programming, 
+one of the most pernicious ways to attack the stack is by hijacking control flow.
+Our stack safety properties do not assume that control flow is well-bracketed; rather,
+they express what it means to protect data even in the presence of control-flow attacks.
+Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''
+in which {\tt f} is called twice, at lines 2 and 4. It should therefore return to lines
+3 and 5, respectively. But the first time, before returning, it stashes its
+return pointer away, and the second time it returns to line 3 instead of 5; then it ends
+up called a third time, at which point it returns to line 5.
+
+Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that we return
+-1. But because of {\tt f}'s interference, we negate {\tt x} twice, and return 1.
+This has the appearance of an integrity violation. But how can that be?
+
+The key intuition is that to return to {\tt main}, {\tt f} needs to do more than just
+execute a return-labeled instruction. It must then actually reach the appropriate
+{\em return target} in {\tt main}. In our context, we keep a stack of return targets,
+which correspond to the depths at which stack frames have been sealed. A callee is
+only considered to have returned when it executes a return-annotated instruction
+and then reaches a valid return target. Only then do the sealed frames become unsealed.
+
+In the example, on its first return, {\tt f} reaches its return target, because the
+\(\PCname\) is 3 after its return. But on the second, {\tt f} does not properly return,
+but rather changes its stack pointer and jumps into {\tt main}'s code. Then {\tt f}
+attempts to overwrite the sealed value of {\tt x}, violating integrity.
 
@@
 \newcommand{\mainpassc}{magenta}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.3\textwidth}
+\begin{verbatim}
+int main() {
+0:           int x;
+1:           x = 5;
+2[P($SP)C]:  x = f(x);
+3[R]:        return x;
+}
+
+int f(int a) {
+10:    int y;
+11:    y = 2*a;
+12:    a = 0;
+13[R]: return y;
+}
+\end{verbatim}
+  \end{subfigure}
+\begin{subfigure}{.65\textwidth}
+\begin{center}
+\begin{tabular}{l r | l}
+  {\tt 2} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{5}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{0}
+    \vspace{.5em}
+  } \\
+  {\tt 11} &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}[Seal(0)]%
+  \memory{1}{\mainpassc}[P(0-1)]%
+  \memory{1}{\retptrc}[RetPtr]%
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
+  \\
+  {\tt 12} &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  {\tt 13} &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  {\tt 3} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{10}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{3}
+    \MemoryLabel{-7em}{0.75em}{10}
+    \vspace{.5em}
+  } \\
+\end{tabular}
+\end{center}
+
+\vspace{\abovedisplayskip}
+
+\end{subfigure}
+
+\caption{Sharing on the Stack}
+\label{fig:passingsimple}
+\end{figure}
 
@@
       {\tt 10} &
-      \memoryaddrs{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
+      \memoryaddrs[(0)]{17em}
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
       \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
@@
       {\tt 13} &
-      \memoryaddrs{17em}
+      \memoryaddrs[(1)]{17em}
       \memory{1}{\mainpassc}
@@
       {\tt 3} &
-      \memoryaddrs{12em}
+      \memoryaddrs[(0)]{12em}
       \memory{4}{\unsealc}
@@
       {\tt 10} &
-      \memoryaddrs{17em}
+      \memoryaddrs[(0)]{17em}
       \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
       \memory{1}{\retptrc}
@@
       {\tt 13} &
-      \memoryaddrs{17em}
+      \memoryaddrs[(0)]{17em}
       \memory{1}{\mainsealc}
@@
 To this point we have ignored function arguments, since our simplest model places them in
-registers. This is an oversimplification, which we will now correct, starting with the more
-interesting case: pass-by-reference.
-\apt{This is a nice and clear example, but I think the ordering here is confusing. Pass-by-reference and passing arguments on the stack are
-  entirely separable issues. I think it would be easiest to explain the latter first, as the
-  simplest example of passed arguments.}
-There are important stack safety concerns
-when passing arguments by reference, as seen in Figure~\ref{fig:passing}. In this example,
+registers. This is an oversimplification, which we will now correct, starting with the simple
+case: passing by value on the stack. Example \ref{fig:passingsimple} shows a sample trace of
+a program that passes a value on the stack, without an attack.
+At the call, we add a new annotation \(P(\$SP)\) identifying a memory location to be
+passed, and our diagram displays it in magenta. The caller's frame locations that are not
+passed are sealed as usual. Just as we marked sealed memory with the depth of its owner,
+we do the same for passed memory -- but it has multiple owners. We keep track of the
+current depth of the active function next to the stack pointer in the diagram, for reference.
+The argument {\tt a} has
+its value copied from {\tt x}, and we note that {\tt f} can read it -- because it is marked
+as passed between depths 0 and 1, it is not varied. Then from line 12 to 13 it is even
+modified, and naturally this is permitted as well.
+
+Passing on the stack is thus straightforward. The same mechanism can be used to pass
+variables by reference, and here there are important stack safety concerns,
+as seen in Figure~\ref{fig:passing}. In this example,
 {\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
@@
 {\tt x} in the second call.
-
-At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying a memory location to be
-passed, and our diagram displays it in magenta. The caller's frame locations that are not
-passed are sealed as usual. Just as we marked sealed memory with the depth of its owner,
-we do the same for passed memory. We keep track of the current depth of the active function
-next to the stack pointer in the diagram.\apt{Not seeing that}  A passed address may only be accessed when the
-current depth is one higher than the one it is sealed with. \apt{This is not good enough!
-  The callee can in turn pass the same argument by reference to its callee...}
 
@@
 
-Passed addresses are also excluded from confidentiality: at the start of each call, we vary
-only stack addresses that are not passed. This is important, as we also use the passing label
-for those arguments that are passed by value on the stack. \apt{this is hidden like an afterthought...}
-We might instead
-treat them as unsealed, but then we would treat them as confidential like the rest of the stack.\apt{unclear.
-  would this work or not?}
+When passing by reference, it is possible to repeatedly pass the same reference further
+and further down the stack. Note, however, that this must be a continuous chain; it will
+not skip over any calls in a chain. So we continue to track two depths, for each passed
+cell: the original owner, and the highest depth it has been passed to. It is considered
+accessible anywhere between the two.
 
 We do not model permanent sharing, in which a variable's address is taken and it might be
-passed far afield before use, because this situation is uninteresting and hard to motivate
-as an instance of stack safety. We would expect to treat such sharing as completely
-unprotected, and perhaps attach a separate memory safety policy to add some protection
-to such variables.\apt{Neither clear nor convincing! Taking explicit addresses of stack
-  variables is very interesting to many reviewers!}
+passed far afield before use, in detail. The appropriate way to do so in our model is
+for such a variable to have no protections except for temporal safety -- it would be
+unprotected until its original owner returns and it is deallocated. Then we would expect
+to separately require memory safety of such data. This is
+a separate property that is often enforced using capabilities, whether tag-based or
+otherwise. In fact, memory safety is the interesting part of this scenario, but is out of
+scope.
 

### 2021-04-13 tex/main.tex
Attacker model tweaks

@@
 properties. While we do not rule out other enforcement mechanisms, our
-definitions are strongly informed by micropolicy systems, especially our
-notion of observational integrity.
+definitions are strongly informed by micropolicy systems, especially
+our attacker model.
 \sna{Seems like the best place to admit we're tied to micropolicies is
@@
 
-\sna{TODO: Need to rewrite this for the new model of WBCF.}
 Since we rely heavily on code annotations, we must ask: what if the annotations
@@
 on initial states: annotated instructions will initially be tagged apropriately, as the
-final instruction in the call or return sequence. This is all we need for the policy to
-protect itself: it can then enforce that the call or return only occurs if that instruction
-is a jump that follows the blessed sequence. So, even in our strong attacker model, we need
-very little in the way of extra guarantees.
+final instruction in the call or return sequence, and all blessed sequences will end with
+annotated instructions. This is all we need for the policy to
+protect itself: it can then failstop if it would execute the blessed sequence out of order.
+So, even in our strong attacker model, we need very little in the way of restrictions on
+the code when considering micropolicy-based enforcement.
 

### 2021-04-13 tex/main.tex
Comments and polishing on abstract / intro

@@
 our attacker model.
-\sna{Seems like the best place to admit we're tied to micropolicies is
-  while discussing micropolicies.}
-
-\bcp{We should mention stack data and coroutines, I think?  Maybe in the
-  context of an expectation-lowering hedge, like:}%
+
 For the sake of concreteness, most of our presentation assumes a RISC-V
-machine architecture with a basic subroutine calling convention, where
-parameters are passed in registers. We also show how to accomodate
-stack-allocated parameter passing and coroutines.
+machine architecture with a basic subroutine calling convention (passing
+parameters in registers) and a micropolicy-like enforcement mechanism. We
+also show how to accomodate stack-allocated parameter passing and
+coroutines.
 
@@
 \sna{See above.}
+\bcp{I added it to the last paragraph (which is pretty close to the preceding
+  paragraph :-)  We should follow this up with some discussion in the (new)
+  Assumptions section about how different enforcement mechanisms (us,
+  StkTokens, maybe others) adopt different perspectives on what the stack is
+for and how it is organized and so lead to somewhat different notions of
+low-level safety.}
 
@@
 
+\bcp{I would still advocate splitting this into two sections, just so the
+  transition from one to the other and the division of responsibility are
+  clear.  The first section could be called ``What is 'the Stack'?'' instead
+of ``Assumptions'' if we want.}
+
 \newcommand*{\rsp}{\textsc{sp}}

### 2021-04-13 tex/main.tex
a few things in 3

@@
 The black rectangle represents the return address. In a typical RISC calling convention
-this will have been stored in a register by call itself, and {\tt f} stores
+this will have been saved into a register by the call itself, and {\tt f} stores
 it in the stack during its preamble, so it is not itself sealed.
@@
       \MemoryLabel{-19em}{0.75em}{1}
-      \MemoryLabel{-15em}{0.75em}{3}
+      \MemoryLabel{-15em}{0.75em}{\#3}
       \vspace{.5em} &
       \memory[3em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 3}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#3}
       \\
@@
       \MemoryLabel{-19em}{0.75em}{-1}
-      \MemoryLabel{-15em}{0.75em}{5} &
+      \MemoryLabel{-15em}{0.75em}{\#5} &
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
       \\
@@
       \MemoryLabel{-19em}{0.75em}{-1}
-      \MemoryLabel{-15em}{0.75em}{3} &
+      \MemoryLabel{-15em}{0.75em}{\#3} &
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
       \\
@@
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
       \\
@@
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5} \\
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5} \\
     \end{tabular}
   \end{subfigure}
+  \caption{A Control Flow Attack.}
   \label{fig:controlflow}
-  \caption{A Control Flow Attack}
 \end{figure}
 
-From the classic stack smashing attack to return oriented programming, 
-one of the most pernicious ways to attack the stack is by hijacking control flow.
+%From the classic stack smashing attack to return oriented programming, 
+One of the most pernicious ways to attack the stack is by hijacking control flow.
 Our stack safety properties do not assume that control flow is well-bracketed; rather,
 they express what it means to protect data even in the presence of control-flow attacks.
-Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''
+Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}
 in which {\tt f} is called twice, at lines 2 and 4. It should therefore return to lines
@@
 return pointer away, and the second time it returns to line 3 instead of 5; then it ends
-up called a third time, at which point it returns to line 5.
+up being called a third time, at which point it returns to line 5.
+\apt{I added \# marks to the line numbers to avoid confusion; this should
+  probably be explained. What do the colors mean in this example?}
 
@@
 -1. But because of {\tt f}'s interference, we negate {\tt x} twice, and return 1.
-This has the appearance of an integrity violation. But how can that be?
-
-The key intuition is that to return to {\tt main}, {\tt f} needs to do more than just
-execute a return-labeled instruction. It must then actually reach the appropriate
-{\em return target} in {\tt main}. In our context, we keep a stack of return targets,
+This has the appearance of an integrity violation.  How do we capture this in our
+integrity property?
+
+The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs to do more than just
+execute a return-labeled instruction: it must actually reach the appropriate
+{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep a stack of return targets,
 which correspond to the depths at which stack frames have been sealed. A callee is

### 2021-04-13 tex/main.tex
More on 3

@@
 
+\apt{How about an example incorporating temporal safety violations due to
+   reading uninitialized values?}
+
 \paragraph*{Control Flow Attacks}
@@
 they express what it means to protect data even in the presence of control-flow attacks.
-Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}
-in which {\tt f} is called twice, at lines 2 and 4. It should therefore return to lines
+Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}.
+(Here we show return addresses explicitly as line numbers, in the form $\#n$.)
+\apt{what do the colors mean in this example?}
+Function {\tt f} is called twice, at lines 2 and 4; it should therefore return to lines
 3 and 5, respectively. But the first time, before returning, it stashes its
@@
 up being called a third time, at which point it returns to line 5.
-\apt{I added \# marks to the line numbers to avoid confusion; this should
-  probably be explained. What do the colors mean in this example?}
 
@@
 
-Passing on the stack is thus straightforward. The same mechanism can be used to pass
+The same annotation mechanism can be used to pass
 variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example,
+as seen in Figure~\ref{fig:passing}. In this example, where we revert to assuming
+that arguments are passed in registers,
 {\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
@@
 When passing by reference, it is possible to repeatedly pass the same reference further
-and further down the stack. Note, however, that this must be a continuous chain; it will
-not skip over any calls in a chain. So we continue to track two depths, for each passed
+and further down the stack. Note, however, that this sharing of the reference must be over
+an unbroken chain of calls. So it suffices to track two depths for each passed
 cell: the original owner, and the highest depth it has been passed to. It is considered
@@
 
-We do not model permanent sharing, in which a variable's address is taken and it might be
-passed far afield before use, in detail. The appropriate way to do so in our model is
-for such a variable to have no protections except for temporal safety -- it would be
-unprotected until its original owner returns and it is deallocated. Then we would expect
-to separately require memory safety of such data. This is
-a separate property that is often enforced using capabilities, whether tag-based or
-otherwise. In fact, memory safety is the interesting part of this scenario, but is out of
-scope.
+Our model provides only limited support for handling fully first-class pointers into
+stack frames, e.g., pointers obtained by taking the address of a local variable.
+Unlike the case for pass-by-reference, there is no easy way to model the intended sharing behavior
+of such pointers in terms of stack frame activation patterns.  The best approximation we can make is to
+treat a location whose address is taken as unsealed for the rest of its lifetime, i.e., until
+its original owner deallocates it (typically by returning).  This prevents violations
+of temporal safety (use after free), but still permits callees to access the location without having
+been explicitly given a pointer to it.  Preventing the latter kind of abuse
+is the provence of a fine-grained memory safety policy, e.g. based on capabilities
+(implemented by tags or some other mechanism); we believe it is beyond the scope of stack safety
+\emph{per se}.
 

### 2021-04-13 tex/main.tex
remove foolish comment

@@
 
-\apt{How about an example incorporating temporal safety violations due to
-   reading uninitialized values?}
-
 \paragraph*{Control Flow Attacks}
@@
 of such pointers in terms of stack frame activation patterns.  The best approximation we can make is to
-treat a location whose address is taken as unsealed for the rest of its lifetime, i.e., until
-its original owner deallocates it (typically by returning).  This prevents violations
-of temporal safety (use after free), but still permits callees to access the location without having
+treat an address-taken variable in the same way as a passed argument, with the additional property
+that it remains accessible at \emph{all} stack depths above the caller. 
+This prevents most violations of temporal safety, although stale pointers can still be
+used in certain corner cases. \apt{Trying not to spell out the details, which are complex.}
+More seriously, callees are permitted to access the location without having
 been explicitly given a pointer to it.  Preventing the latter kind of abuse
 is the provence of a fine-grained memory safety policy, e.g. based on capabilities
-(implemented by tags or some other mechanism); we believe it is beyond the scope of stack safety
+(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
 \emph{per se}.

### 2021-04-14 tex/main.tex
Simple properties stated

--- 2021-04-13 tex/main.tex
@@
 
-Next we consider {\em lockstep stack confidentiality}. Here it behooves us to factor out
-the general notion of lockstep confidentiality independent of a particular context model.
-
-\subsection{Eager Confidentiality}
+Next we consider {\em stepwise stack confidentiality}. There are a number of preliminary
+concepts we must discuss here -- confidentiality quantifies over {\em variants} of the
+initial state of a call, and requires that the traces from them move in {\em lockstep}.
 
@@
 
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) have the {\em same differences}
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are {\em synched}
 if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
@@
 This is written \(\mach,\mach' \diamond \nach,\nach'\).
+
+\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
+are synched, written as a coinductive relation \(\machT \doteq \nachT\):
+
+\begin{minipage}{.3\textwidth}
+\judgment{}
+         {\(\mach \doteq \nach\)}
+\end{minipage}
+\begin{minipage}{.6\textwidth}
+\judgmenttwo{\(\mach, \head(\machT) \diamond \nach, \head(\nachT)\)}
+            {\(\machT \doteq \nachT\)}         
+            {\(\mach \machT \doteq \nach \nachT\)}
+\end{minipage}
 
@@
 \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\). \(\MPCT\) corresponds to the
-trace of a subroutine from its call to its return.
-
-If we take the induced trace from this variant until a return,
-\(\nach,\pol,\context \hookrightarrow \NPCT | \neg f_\depth\), we get a subtrace corresponding
-to the same subrouting called with the same (register) arguments but with a potentially different
-state at the call-site. A system enjoys {\em eager confidentiality} if for all such pairs of traces,
-they correspond according to one of four cases:
-
-\begin{enumerate}
-\item \(\MPCT\) may be finite and \(\pi_\context(\last(\MPCT)) = (\_,\depth')\) with \(\depth' < \depth\),
-  and \(\pi_\context(\last(\NPCT)) = (\_,\depth')\) with \(\depth' < \depth\). That is, both traces end
-  having left the call of depth \(\depth\). Further:
-  \begin{itemize}
-  \item They have the same observable behavior, \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
-  \item Their final states have the same differences:
-    \[\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
-    \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\]
-  \end{itemize}
-\item \(\MPCT\) and \(\NPCT\) may both be infinite, with \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
-\item \(\MPCT\) may be finite but end without returning -- which can only have happened if it fail-stopped --
-  and its observable behavior prefixes that of \(\NPCT\), \(\pi_\obs(\MPCT) \lesssim \pi_\obs(\NPCT)\)
-\item \(\NPCT\) may end with a fail-stop as in case 3, with \(\pi_\obs(\NPCT) \lesssim \pi_\obs(\MPCT)\)
-\end{enumerate}
-
-\subsection{Trace Properties}
-
-Having seen instances of eager confidentiality and integrity, defined in terms of reachable trace
-segments, we can generalize them to properties of the trace segments themselves. Then as we expand
-our model, our further properties will be defined in terms of the trace properties.
-
-\definition Let \(\MPCT\) be a reachable segment and \(\components\) be any set of components.
-\(\MPCT\) enjoys {\em trace integrity} when \(\last(\MPCT) ~ \component = \head(\MPCT) ~ \component\)
-for all \(\component \in \components\), and when \(\last(\MPCT)\) is undefined.
-
-\definition Let \(\MPCT \langle f \rangle\) be a reachable segment delimited by the predicate \(f\)
-and the \(\components\) be any set of components. Let \(\head(\MPCT) = \mpcstate{\mach}{\pol}{\context}\)
-and \(\nach\) be a machine state such that \(\mach \approx_\components \nach\). Suppose we take the
-induced trace from \(\npcstatename = \mpcstate{\nach}{\pol}{\context}\) that corresponds to \(\MPCT\),
-\(\npcstatename \hookrightarrow \NPCT | \neg f\). The trace \(\MPCT\) enjoys {\em trace confidentiality}
-with respect to \(f\) and \(P\) if, for all such \(\NPCT\), they correspond according to the four cases
-above:
-
-\begin{enumerate}
-\item \(\MPCT\) may be finite and \(\pi_\context(\last(\MPCT)) = (\_,\depth')\) with \(\depth' < \depth\),
-  and \(\pi_\context(\last(\NPCT)) = (\_,\depth')\) with \(\depth' < \depth\). That is, both traces end
-  having left the call of depth \(\depth\). Further:
-  \begin{itemize}
-  \item They have the same observable behavior, \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
-  \item Their final states have the same differences:
-    \[\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
-    \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\]
-  \end{itemize}
-\item \(\MPCT\) and \(\NPCT\) may both be infinite, with \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
-\item \(\MPCT\) may be finite but end without returning -- which can only have happened if it fail-stopped --
-  and its observable behavior prefixes that of \(\NPCT\), \(\pi_\obs(\MPCT) \lesssim \pi_\obs(\NPCT)\)
-\item \(\NPCT\) may end with a fail-stop as in case 3, with \(\pi_\obs(\NPCT) \lesssim \pi_\obs(\MPCT)\)
-\end{enumerate}
+\(\MPCT\) corresponds to the trace of a subroutine from its call to its return.
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
+\(\MPCT \doteq \NPCT\).
 

### 2021-04-14 tex/main.tex
pass through 4

--- 2021-04-13 tex/main.tex
@@
 
+Observations are only essential for describing the extensional properties
+described in Section~\ref{sec:lazy}, but for simplicity, we include them
+in the machine model from the beginning.
 We choose to model observations explicitly, rather than simply treating
@@
 
-In the running example of \Cref{sec:running-example},
+\apt{This para is out of date. Do we want to say something equivalent
+  about new examples (hard, because they are C-based).}
 $\WORDS$ is the
@@
 
+\noindent
 This combined step function is also partial, and traces of the machine-with-policy
@@
 an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context, \ann \in \powerset \ANNS \constep \context' \in \MACHS \times \CONTEXTS
-\times \ANNS \rightarrow \CONTEXTS\). We also assume for any set of annotations that we have
-code maps, with \(\FUNIDS\) a set of function identifiers:
-\[\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]
-
+\(\mach, \context, \anns \constep \context' \in \MACHS \times \CONTEXTS
+\times \powerset(\ANNS) \rightarrow \CONTEXTS\). We also assume a set of function identifiers
+\(\FUNIDS\) and the existence of a \emph{code map} \(\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\)
+that relates each code address to its owning function and set of annotations.
 We lift contexts into an MP trace in much the same way as we lifted policy states to get
@@
 \judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \ann\)}
-              {\(\mach, \context, \ann \constep \context'\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
+              {\(\mach, \context, \anns \constep \context'\)}
               {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
@@
 \(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
-called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\), and of observations.
+called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
 
@@
 
-The ``observations-of'' operator, written \(\pi_\obs\), takes a machine trace and
+The ``observations-of'' operator, written \(\pi_\obs\), \apt{since this isn't
+  actually a projection, it is slightly confusing to use \(\pi\)} takes a machine trace and
 creates a trace of observations corresponding to the observations of each step in
@@
 to relate a machine-policy-context state triple to the MPCO-trace
-induced by the step function.
+induced by the step function.\apt{?? This looks indistinguishable
+  from the \(\hookrightarrow\) we've already defined.}
 
@@
 \hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)'') and similarly for traces with policies:
+running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:
 
@@
 
-In particular, we wish to formulate properties around consecutive subtraces
+In particular, we wish to formulate properties about consecutive subtraces
 in which the context obeys a given predicate. This will commonly be used to extract
@@
 at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
-we have \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
+we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
 ``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''
@@
 \begin{center}
-  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\callmap\} \cup \ann) \stepstocon
+  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\callmap\} \cup \anns) \stepstocon
              (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
-  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\retmap\} \cup \ann) \stepstocon
+  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\retmap\} \cup \anns) \stepstocon
              \mach',(\pop{\domMap}{\depth},\depth-1)\)}
-  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\passmap ~ P\} \cup \ann) \stepstocon
+  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\passmap ~ P\} \cup \anns) \stepstocon
              \domMap[P ~ \mach \mapsto \passed{\depth}], \depth)\)}
-  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\sharemap ~ P\} \cup \ann) \stepstocon
+  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\sharemap ~ P\} \cup \anns) \stepstocon
              \domMap[P ~ \mach \mapsto \shared{\depth}], \depth)\)}
@@
 We now extend it with an annotation for valid entry points:
-\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
-\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
-\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
+\[\ann \in \ANNS ::= \callmap + \retmap + \yieldmap +  \entmap + 
+\sharemap(P \subseteq \MACHS \times \COMPONENTS) +
+\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
 

### 2021-04-14 tex/main.tex
Filled in section 5

@@
                         %% http://ctan.org/pkg/subcaption
-\usepackage{tikz}
-\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
-    \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
 
@@
 the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
-Then at each step we update the context according to the context-step function,
-\(\stepsto_C\). We will define this in terms of the operations
-\(\push{\domMap}{\depth}{\components}\), in which \(\depth \in \mathbb{Z}\) and
-\(\components \subseteq \COMPONENTS\), and \(\pop{\domMap}{\depth}\), pronounced
-``push'' and ``pop'' and defined as follows.
-
-\[
-(\push{\domMap}{\depth}{\components}) ~ \component =
-\begin{cases}
-  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
-                    \text{ and } \component \in \components \\
-  \domMap ~ \component & \text{else} \\
-\end{cases}
-\]
-\[
-(\pop{\domMap}{\depth}) ~ \component =
-\begin{cases}
-  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth'} \text{ where }
-              \depth' >= \depth\\
-  \domMap ~ \component & \text{else} \\
-\end{cases}
-\]
-%
-Push seals a set of components with its depth argument, and pop unseals all components
-that were as deep as its argument or deeper.
 
@@
 and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-
-The call rule pushes the components sealed by the sealing convention with the depth
-of the current caller, and determines the return target by partially applying the
-return convention to the machine state, and pushes that onto the targets stack.
-The return rule checks whether the attempted return actually manages to reach a
-return target, traversing down the stack until the first one it finds. Recall that
-we are not discussing well-bracketed control flow, so the return might skip over
-a waiting caller. When a target is reached, {\em all} sealed components at or above
-its depth are unsealed by the pop operation. If no target is reached, nothing is
-popped at all. These rules are given inductively, and apply in the listed order.
-
-\judgmenttwo[Call]
+Then at each step we update the context according to the context-step function,
+\(\stepsto_C\).
+
+The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
+A call simply pushes the return target onto the return stack and seals all components
+with the depth of the previous return stack. A return
+recursively searches for a valid return target at the next step, and unseals
+every component sealed at a depth that it passes.
+
+\judgmentthree[Call]
               {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
-            {\(\depth = |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C
-              \push{\domMap}{\depth}{\sealcon ~ \mach},(\retcon ~ \mach)::\rts\)}
-
-\judgmentthree[ReturnFound]
+              {\(\components = \sealcon ~ \mach\)}
+              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
+                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\judgmenttwobr[ReturnFound]
               {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
               {\(\mach \stepsto \mach' \in \target\)}
-              {\(\depth = |\rts|\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C
-                \pop{\domMap}{\depth},\rts\)}
-
-\judgmentthree[ReturnRec]
-              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \_,\rts'\)}
-              {\(\depth = |\rts'|\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C
-                \pop{\domMap}{\depth},\rts'\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\judgmenttwobr[ReturnRec]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
 
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are {\em synched}
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change together
 if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
@@
 \definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
-are synched, written as a coinductive relation \(\machT \doteq \nachT\):
+change together, written as a coinductive relation \(\machT \doteq \nachT\):
 
@@
 
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold for
-\(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth)\) and
-\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
 \(\MPCT\) corresponds to the trace of a subroutine from its call to its return.
@@
 
-\subsection{Stack Safety with Sharing}
-
-Next we introduce sharing. We will need to extend the set of annotations to include
-sharing and passing, and now the code map may map a single instruction to multiple
-annotations, because in some calling conventions a single instruction might carry
-meaning as both a call and a pass. But we will never have a single instruction annotated
-as both a call and a return.
-
-\[\ann \in \ANNS ::= \callmap + \retmap + \sharemap(P \subseteq \MACHS \times \COMPONENTS) +
-\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
-
-And we extend the set of domains to include shared and passed components, the latter of
-which are labeled with the depth of the caller that has passed them.
-
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \shared{\depth} + \sealed{\depth} + \passed{\depth}\]
+\subsection{Stack Safety with Passing}
+
+Next we introduce variable passing. We will need to add extra information to our call
+annotations to reflect that they may now pass variables on the stack, and that the locations
+of those variables may be dynamic (as in the case of pass-by-reference).
+A call annotation now includes a predicate relating machine states to the components
+that should be passed -- when passing by value, this will typically give the memory cells
+at the appropriate offset from the stack pointer, while passing by reference gives the
+addresses of the appropriate variables wherever they happen to be.
+
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap\]
+
+And we extend the set of domains to include passed components,
+which are labeled with the range of depths of the caller that has passed them.
+
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
 \[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\depth \in \mathbb{N}) \]
-
-What it means to push the stack is similar to above, but two details are worth noting.
-First, shared and passed components are not sealed by the sealing convention, and second,
-anything not touched by the sealing convention must be unsealed. Sharing and passing cannot
-apply to memory the callee could otherwise use.
-
-\[
-(\push{\domMap}{\depth}{\components}) ~ \component =
+\times (\rts \in \listT (\target \subseteq \MACHS)) \]
+
+We straightforwardly describe how the sealing convention and passing predicate combine to
+update a domain map when pushing it, where \(\depth\) is an integer depth.
+
+\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
 \begin{cases}
-  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
-                    \text{ and } \component \in \components \\
-  \domMap ~ \component & \text{else if } \component \in \components \\
-  \unsealed & \text{else} \\
-\end{cases}
-\]
-
-Popping is simpler, but note that while passed components are unsealed on return to their
-passer, shared ones are unsealed on return {\em from} their sharer - on deallocation,
-in other words.
-
-\[
-(\pop{\domMap}{\depth}) ~ \component =
+  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
+  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
+  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
+                                 \text{ for any } \depth' \\
+  \domMap~\component & \text{else} \\
+\end{cases}\]
+
+And to pop we just unseal the components at depth \(\depth\):
+
+\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
 \begin{cases}
-  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1}
-              \text{ or } \domMap ~ \component = \passed{\depth-1}
-              \text{ or } \domMap ~ \component = \shared{\depth} \\
-  \domMap ~ \component & \text{else} \\
-\end{cases}
-\]
-
-Finally we introduce a more direct update. If \(P\) is a relation between machine states and
-components, we write \(\domMap[P ~ \mach \mapsto \stackDom]\) to say \(\domMap\) updated
-so that each component \(\component\) where \(P ~ \mach ~ \component\) holds is mapped to
-\(\stackDom\).
-
-Now we can inductively define the context update function; as these could be ambiguous assume that
-they apply in the order presented.
-
-\begin{center}
-  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\callmap\} \cup \anns) \stepstocon
-             (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
-  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\retmap\} \cup \anns) \stepstocon
-             \mach',(\pop{\domMap}{\depth},\depth-1)\)}
-  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\passmap ~ P\} \cup \anns) \stepstocon
-             \domMap[P ~ \mach \mapsto \passed{\depth}], \depth)\)}
-  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
-           {\(\mach,\context,(\{\sharemap ~ P\} \cup \anns) \stepstocon
-             \domMap[P ~ \mach \mapsto \shared{\depth}], \depth)\)}
-  \judgment{}
-           {\(\mach,\context,\_ \stepstocon \context\)}
-\end{center}
-
-\definition Let \(\MPCT\) be a reachable trace segment of the form
-\(\mpcstatename ~ \mpcstatename'\), where \(\pi_\context(\mpcstatename) = (\domMap,\depth)\).
-Let \(\components\) be the set of components \(\component\) such that either
-\(\domMap ~ \component = \sealed{\_}\) or \(\domMap ~ \component = \passed{\depth'}\) for
-some depth \(\depth' < \depth-1\). If any such \(\MPCT\) obeys trace integrity with respect
-to \(\components\), then the system enjoys {\em ultra-eager confidentiality} in the sharing
-model.
-
-\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
-on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
-\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
-\(\component\) such that either \(\domMap ~ \component = \sealed{\_}\) or
-\(\domMap ~ \component = \passed{\depth''}\) for some depth \(\depth'' < \depth'-1\).
-Then the system enjoys {\em eager integrity} in the sharing model if \(\MPCT\) obeys
-trace integrity with respect to \(\components\).
-
-\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
-on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
-\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
-\(\component\) such that either \(\domMap ~ \component = \outside\) or
-\(\domMap ~ \component = \shared{\_}\). Then the system enjoys {\em eager confidentiality}
-in the sharing model if \(\MPCT\) obeys trace confidentiality with respect to \(\components\)
-and \(f\).
+  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
+  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth,\depth} \\
+  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
+                              \text{ for any } \depth' < \depth \\
+  \domMap~\component        & \text{else} \\
+\end{cases}\]
+
+Now we can inductively define the context update function; these again apply in the order
+presented.
+
+\judgmenttwo[Call]
+            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
+            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
+            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\judgmentthree[ReturnFound]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+              {\(\mach \stepsto \mach' \in \target\)}
+              {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\judgmentthree[ReturnRec]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\domMap'' = \mathit{pop} ~ \domMap ~ |\rts|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\judgment[Default]
+         {}
+         {\(\_,\context \stepsto_C \context\)}
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
+Let \(\components\) be the set of components \(\component\) where 
+\(\domMap ~ \component = \sealed{\_}\) or \(\passed{\depth,\depth'}\) where \(|\rts| < \depth\)
+or \(|\rts| > \depth'\). A system enjoys {\em stepwise stack integrity with passing} if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \unsealed\) or \(\passed{\depth,\depth'}\) where
+\(\depth \leq |\rts| \leq \depth'\) .
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
+\(\MPCT \doteq \NPCT\).
 
@@
 Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids. The ``active stack'' -- the stack belonging to the
-currently executing coroutine -- in any given machine state is \(\stackof(\mach ~ \SP)\).
-We extend the set of possible annotations with a yield.
-
-\[\ann \in \ANNS ::= \callmap + \retmap + \yieldmap +
-\sharemap(P \subseteq \MACHS \times \COMPONENTS) +
-\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
-
-We use the same domain model as above, but instead of keeping track of a single depth
-we use a map from stacks to their depths.
-
+\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
+pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
+We extend the set of possible annotations with a yield annotation.
+
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap + \yieldmap\]
+
+The first major difference in this model is the domains. Stack domains are now nested inside
+top-level domains, and associated with stack identifiers.
+Our context now consists of a domain map, a map from stack identifiers to return target stacks,
+an additional map of targets for yields, and a stack identifier of the active stack.
+Our yield target map \(\yts\) is a map from stack identifiers to targets.
+In a standard system an initial yield target map will map each stack identifier to the
+set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
+\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
+
+\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
+\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
 \[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
-(\depthMap \in \STACKS \rightarrow \mathbb{N})\]
-
-\subsection{Local State Encapsulation}
-
-\definition A system enjoys {\em local state encapsulation} if it enjoys eager integrity and
-eager confidentiality.
-
-\definition A system enjoys {\em lazy local state encapsulation} if it enjoys lazy integrity
-and eager confidentiality.
-
-We do not give a confidentiality equivalent to lazy integrity, because the family of lazy
-enforcement mechanisms that we consider under the HOPE system in fact enforce eager confidentiality.
-A policy based on taint-tracking might require such a lazy confidentiality, and it would be constructed
-by replacing the ``same-difference'' condition with one that requires any differences to produce
-observationally equivalent states. But for our purposes it is irrelevant.
-
-\section{Well-bracketed Control Flow}
+(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
+\target) \times (\stackid \in \STACKS)\]
+
+In our context update rules, calls and returns work the same as in the sharing model,
+with domain updates restricted to the top-level domain corresponding to the active stack,
+and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
+
+\judgmentthreebr[Yield]
+                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
+                {\(\mach \stepsto \mach'\)}
+                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
+                {\(\stackid \not = \stackid'\)}
+                {\(\yts ~ \stackid'\)}
+                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
+                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
+
+This calls for some explanation. First, we determine which coroutine we are yielding to by
+checking which stack the stack pointer of the next state is pointing to. Then we can only yield
+if the next state is a valid yield target for that stack, and we update the yield target for
+the yielding stack. So it is once again possible for a coroutine to claim to yield but not
+actually change the active stack, with similar effects to when a callee fails to return properly.
+Now we can define both stack safety and coroutine safety under this model.
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where 
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
+{\em stepwise stack integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where 
+\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
+A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\_}\).
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise coroutine confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
+\subsection{Well-bracketed Control Flow}
 \label{sec:wbcf}
 
-Both the integrity and confidentiality components of local state
-encapsulation concern themselves only with the accessibility of data,
-and may still hold on a wide range of programs that violate expected
-control flow. The archetypal example of this is the {\em awkward example},
-in which an adversarial callee stores a return address and returns to it
-on a subsequent call, thus violating the expected control flow of the
-caller without actually violating the caller's integrity. But it also
-encompasses simpler violations, such as a function bypassing a password
-check by jumping (or returning) in the middle of a block of code.
-%
-Stack safety enforcement mechanisms like \citet{Skorstengaard+19}
-usually aim to prevent such unwanted flows: every jump between
-functions should either be a call or a return, and each return should
-be to the instruction after its corresponding call.
-
-To formally capture this aspect of stack safety, we introduce some more
-security properties. {\em Control separation} restricts transitions between
-functions to follow the interfaces set out by call, return, and yield, forming
-a foundation for all of the other properties that rely on those interfaces for
-meaning. {\em Entry integrity} ensures the validity of call targets,
-{\em return integrity} ensures that of return targets, and
-{\em yield-back integrity} ensures that of yield targets.
-
-Recall that our properties use a compiler-derived code-map determine
-which addresses are instructions and which functions they belong to, as well
-as annotating them as calls, returns, and yields, as well as sharing information.
-We now extend it with an annotation for valid entry points:
-\[\ann \in \ANNS ::= \callmap + \retmap + \yieldmap +  \entmap + 
-\sharemap(P \subseteq \MACHS \times \COMPONENTS) +
-\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
-
-\definition%{Control Separation}
-
-Let \(\mpcstatename\) be a reachable state whose program counter points to code in
-function \(\funid\), and \(\mpcstatename \stepstopol \mpcstatename'\) where
-\(\mpcstatename'\) is state whose program counter points to a different function
-\(\funid'\). A system enjoys {\em control separation} when all such \(\mpcstatename\)
-are annotated \(\callmap\), \(\retmap\), or \(\yieldmap\).
-
-\definition%{Entry Integrity}
-A system enjoys {\em entry integrity} when, for any reachable \(\mpcstatename\) and
-\(\mpcstatename'\) such that \(\mpcstatename\) is annotated \(\callmap\) and
-\(\mpcstatename \stepstopol \mpcstatename'\), \(\mpcstatename'\) is annotated \(\entmap\).
-
-\definition%{Return Integrity}
-A system enjoys {\em return integrity}... [this one needs to talk about context.]
-
-\definition%{Yield-back Integrity}
-A system enjoys {\em yield-back integrity}... [same.]
-
-\definition% {Well-Bracketed Control Flow}
-The above properties are somewhat orthogonal, but together they
-capture a standard notion of control flow for a stack: functions keep
-their code separate outside of calls and returns, calls respect
-explicit entry points, and returns respect implicit entry points
-corresponding to their call.  A system with all three
-properties (with respect to a function map, call map, return map,
-and entry map) is said to enjoy
-{\em well-bracketed control flow} (with respect to those maps).
+Both the integrity and confidentiality components of stack safety
+are reasonable even in programs that violate expected
+control flow. But even though stack safety prevents control-flow violations
+from being used to gain privileges, it is worth protecting control flow as well.
+Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
+control flow.
+
+When we focus on subroutines, the essence of well-bracketed control flow
+is that a caller can expect that if its callee ever returns, it returns
+to the caller at the appropriate return target. This is tantamount to saying
+that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
+the simple model:
+
+\definition A system enjoys {\em well-bracketed control flow} if for all
+reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
+\(\mpcstatename \stepstocon \mpcstatename'\), if
+\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
+is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
+
+Extending this idea to coroutines, we use the same notion that there are
+no ``false'' yields.
+
+\definition A system enjoys {\em well-structured control flow} if for all
+reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
+\(\mpcstatename \stepstocon \mpcstatename'\), if
+\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
+is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
+where \(\stackid \not = \stackid'\).
 
@@
 
-\paragraph{Rolling Back to Idealized State}
-
-Observable integrity weakens eager integrity by allowing a returned
-state to differ from its call state at high integrity components,
-provided that the differences do not change the observable behavior of
-the rest of the program. The challenge here is to define what constitutes
-a change in behavior. We construct an idealized return state that
-contains only the changes that the caller was allowed to make, and use
-the trace from that state as a baseline against which the real trace
-will be compared.
-
-      \definition
-
-      Given a contour, a call state, and a returned state, we define a
-      {\em rollback} function \(\mathit{roll} \in \CONTOURS \times \MACHS \times
-      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
-      the call state on those components that are high integrity in its
-      contour, and matches the return state on low integrity components:
-      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
-      \begin{cases}
-        \mach_e(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
-        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
-      \end{cases}\]
-
-      If \(\mach_e\) is an entry to a call with contour \(\contour\)
-      and \(\mach_r\) its return, the result of
-      \(\mathit{roll}(\contour,\mach_e,\mach_r)\) agrees with the
-      entry \(\mach_e\) on every component the callee should not
-      change, and with the actual return \(\mach_r\) on those
-      components that it is permitted to change.  So the observable
-      behavior that follows is the idealized behavior to which we
-      compare the trace from \(\mach_r\).
-
-      Now we can define a version of trace integrity that compares the trace
-      from a returned state to that from its idealized counterpart.
-
-    \definition%{Observable Trace Integrity}
-
-      A trace enjoys {\em observable trace integrity} with respect to
-      some contour \(\contour\) if the trace induced by its final
-      state (if any) has observable behavior that prefixes that of its
-      idealized version (to account for premature halts due to policy
-      violations):
-
-      \[\begin{split}
-        \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
-        & \head(\MPT) = \mpostate{\mach_e}{\_}{\_} ~ \land ~
-          \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land \\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
-          \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
-          \Rightarrow \\
-        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
-      \end{split}\]
-
-%    \begin{figure}
-%      \integritylazyexample
-%      \caption{Rolling back from Figure \ref{fig:intex}.}
-        %\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
-%      \label{fig:intlex}
-%    \end{figure}
-
-    For our example, we return to the violation of trace integrity in
-    \cref{fig:intex}, in which {\tt g} has written to {\tt z} and returned.
-    Whether observable integrity holds depends on the continuing trace, shown
-    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
-    \(\mach_r\), and the rolled-back state, \(\mach_r' =
-    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
-    components are restored to their values as of \(\mach_e\). Specifically,
-    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
-    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
-    it prints 0. Only then is observable trace integrity violated.
-
-    However, consider instead if we modify the original code of \cref{fig:running-trace-b} to insert after instruction 44 a new
-    instruction \(44.5: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
-    {\tt f}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
-    rolled-back state \(\mach_r'\) maps it to 0. But in the trace from either
-    state, instructions 27 and 28 load the return value (18) and store it to
-    105, eliminating the only difference. The remainder of the observation
-    traces must be identical, so this trace does obey observable trace
-    integrity.
-
-    \definition%{Observable Stack Integrity}
-
-      A system enjoys {\em observable stack integrity}, like stack integrity,
-      if, for any initial state and each \(n\)-transition from
-      \(\mpstate{\mach_c}{\pol_c}\) to \(\mpstate{\mach_e}{\pol_e}\) in its
-      induced trace, we take the prefix of the call
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\)
-      and observable trace integrity holds on \(\MPT_{\mathit{call}}\) with
-      contour \(\mathit{Cof}(\mach_c,n)\).
-
 \subsection{Observable Confidentiality}
-
-      Two of the cases of trace confidentiality (\cref{def:trace-conf}), \(\mathit{InfCase}\)
-      and \(\mathit{FailCase}\), are defined solely in terms of
-      observation traces. If
-      the traces are infinite, they should be similar; if one
-      fail-stops, it should be a prefix. Only the
-      \(\mathit{ReturnCase}\) checks machine states, because in that
-      case the traces could continue, but have been cut off by a
-      return. The key to making confidentiality lazy is not to perform that
-      check but rather to consider the traces after that return,
-      checking only whether they are observationally similar.
-
-      This requires restoring varied components to their original
-      values in the trace from the variant state, since after return
-      they will once again be legitimately accessible. To return to
-      the example in \cref{fig:confex}, the varied final state
-      \(\mach_r'\) still has $-5$ as a return address for {\tt f}, so
-      the trace after that return will behave oddly in a way that
-      does not reflect a confidentiality violation.  That means we
-      cannot immediately compare traces from \(\mach_r\) and
-      \(\mach_r'\), but rather that we must restore the varied values
-      while preserving differences that have occurred during
-      execution.
-
-      \definition
-
-      The \(\mathit{restore}\) function takes four states: the
-      original entry state, a variant entry state, and corresponding return
-      states. It returns another state. The resulting state restores those
-      components that varied between entries to match the original entry,
-      unless they were changed during execution:
-      \[\mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')
-        (\component) =
-        \begin{cases}
-          \mach_e(\component) & \text{if } \mach_e(\component) \not =
-                                \mach_e'(\component) \text { and } \\
-          & \mach_e(\component) = \mach_r(\component) \text{ and }
-            \mach_e'(\component) = \mach_r'(\component) \\
-          \mach_r'(\component) & \text{else} \\
-        \end{cases}\]
-
-      Now we can describe a trace property akin to trace confidentiality, but
-      continuing execution after the end of the callee trace.
-
-    \definition{%Observable Trace Confidentiality}
-
-      {\em Observable trace confidentiality} is the lazy analog to
-      trace confidentiality, and reuses the cases in which the traces
-      do not return.  For a return predicate \(R\) and a pair of
-      traces \(\MPT\) and \(\machT\), we consider the return case, in
-      which $\MPT$ has a last element
-      \(\mpostate{\mach_r}{\_}{\_}\) and that machine state is
-      a return, i.e. \(R ~ \mach_r\) holds.
-%
-      Observable trace confidentiality then requires that $\machT$
-      also has a final state \((\mach_r',\_)\) that is a return, i.e.
-      \(R ~ \mach_r'\) holds.
-%
-      In addition, just like in the eager case,
-      observations of \(\MPT\) are similar to those of \(\machT\).
-
-      The only thing that changes is the requirement that \(\mach_r\)
-      and \(\mach_r'\) agree on components. Instead, let \(\mach_e\)
-      be the machine state of \(\head(\MPT)\), and \(\mach_e'\) be
-      that of \(\head(\machT)\). We restore \(\mach_r'\) based on
-      \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
-      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).  Then
-      observable trace confidentiality requires that the
-      induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
-      observational prefix of the induced trace from \(\mach_r'\),
-      reflecting the possibility of a fail-stop after return.
-
-      \[\begin{split}
-        \mathit{ObsReturnCase} ~ R ~
-        (\mpostate{\mach_e}{\_}{\_}\MPT) ~
-        ((\mach_e',\_)\machT) \ \ \triangleq \ \
-        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land ~
-         \last(\machT) = (\mach_r',\_) ~ \land \\
-        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') ~ \land \\
-        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
-        \mach_r'' \hookrightarrow \machT' \Rightarrow \\
-        & \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land ~
-        \pi_o(\MPT') \lesssim \pi_o(\machT') \\
-      \end{split}\]
-
-      We combine this case with the other two cases from ordinary trace
-      confidentiality:
-\[
-        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
-         \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT ~ \land
-         \mathit{InfCase} ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
-          \MPT ~ \machT
-\]
-
-%      \begin{figure}
-%        \confidentialitylazyexample
-%        \caption{Continuing from \cref{fig:confex}.}
-%        \label{fig:conflex}
-%      \end{figure}
-
-      In our example, \cref{fig:conflex} shows how execution continues from
-      \(\mach_r'\) and the restored
-      \[\mach_r'' = \mathit{restore}(%\contour,
-      \mach_e,\mach_e',\mach_r,\mach_r').
-      \]
-      Addresses 106 and 107
-      are not restored; every address that was varied and not changed in one
-      of the executions is. So 107 contains 43 in the trace from \(\mach_r\),
-      and after returning from {\tt f}, {\tt main} prints 85. But in the trace
-      from \(\mach_r''\), 107 contains $-4$, and ultimately 38 is printed.
-      So observable trace confidentiality does not hold. This example
-      illustrates the weakness of observable confidentiality as a testing
-      property: it takes until the very end of the program to determine whether
-      a tainted value actually influences observable behavior.
-
-      On the other hand, suppose that we inserted an instruction at 43:
-      \(r_6 \leftarrow [\rsp-8]\). \(r_6\) would hold 42 in \(\mach_r\)
-      and $-5$ in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
-      never used otherwise, the traces of {\tt f} will enjoy observable trace
-      confidentiality.
-
-    \definition%{Observable Stack Confidentiality}
-
-      {\em Observable stack confidentiality} is structured like stack
-      confidentiality. For each initial
-      state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
-      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
-      \(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\), take any
-      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
-      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~
-      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~
-      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
-      \ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}\).
-
-      Additionally, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
-      \mach_0\), taking the induced trace  \(\mach_0' \hookrightarrow
-      \machT\), we must have \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\).
-
-      A system enjoys observable stack confidentiality if observable
-      confidentiality holds in all such cases.
-
-    \definition%{Observable Stack Safety}
-    A system enjoys {\em observable stack safety}
-    % with respect to a call map       and a function map
-    if it enjoys both observable stack integrity and
-      observable stack confidentiality with respect to them.
-
-      \medskip
-      \textsc{Theorem.}
-      Stack safety implies observable stack safety.
-      \smallskip
-
-      This is more straightforward than the implication in the previous
-      section (1,5K LoC in Coq) as the recursive
-      structure of the two properties is similar.
 

### 2021-04-14 tex/main.tex
Comments

@@
 
-For the sake of concreteness, most of our presentation assumes a RISC-V
-machine architecture with a basic subroutine calling convention (passing
-parameters in registers) and a micropolicy-like enforcement mechanism. We
+For the sake of concreteness, most of our presentation assumes a standard RISC
+machine architecture with a basic subroutine calling convention (parameters
+passed in registers) and a micropolicy-like enforcement mechanism. We
 also show how to accomodate stack-allocated parameter passing and
@@
 \sna{See above.}
-\bcp{I added it to the last paragraph (which is pretty close to the preceding
+\BCP{I added it to the last paragraph (which is pretty close to the preceding
   paragraph :-)  We should follow this up with some discussion in the (new)
@@
 
-\bcp{This paragraph doesn't apply to the lazy properties, so maybe it should
+\BCP{This paragraph doesn't apply to the lazy properties, so maybe it should
 be moved before they are introduced?}%
@@
 
-\bcp{I would still advocate splitting this into two sections, just so the
+\BCP{I would still advocate splitting this into two sections, just so the
   transition from one to the other and the division of responsibility are
-  clear.  The first section could be called ``What is 'the Stack'?'' instead
-of ``Assumptions'' if we want.}
+  clear.  The first section could be called ``Setup'' or something like
+  that.}
 
@@
 the layout of memory, including where code lives and which addresses are within
-the bounds of the stack.
+the bounds of the stack.  \BCP{What should we say about the heap?}
 
@@
 So, even in our strong attacker model, we need very little in the way of restrictions on
-the code when considering micropolicy-based enforcement.
+the code when considering micropolicy-based enforcement.  \BCP{Is this
+  discussion still needed?  I had hoped that switching to our ``new''
+  definition on subtraces would eliminate such complexities!}
 
@@
 
-\section{Demonstrating Stack Safety}
+\section{Key Ideas}
 \label{sec:running-example}

### 2021-04-14 tex/main.tex
TIny
CP

@@
 memory.  But these mechanisms typically lack rigorous specifications,
-relying instead on informal descriptions of
-the safe stack manipulation behaviors that they allow
-and/or specific examples of the bad behaviors that they prevent.
-
-We propose formal criteria for stack safety, phrased in terms of
+relying instead on informal descriptions and specific examples of the bad behaviors that they prevent.
+
+We propose formal criteria for stack safety, using the concepts of
 language-based security: a combination of an integrity property (``each
@@
 %
-We argue that these criteria match common intuitions (1) by showing that
-they are violated by a variety of intuitively unsafe examples and (2) by
+We argue that these criteria match common intuitions by (1) showing that
+they are violated by a variety of intuitively unsafe examples and (2)
 using property-based random testing to check that appropriate variants of
@@
 proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
-broken versions of these enforcement mechanisms).  Specifically, Roessler
+broken variants).  Specifically, Roessler
 and Dehon's ``eager'' micro-policy, which catches violations as early as
-possible, enforces the basic ``stepwise'' versions of our confidentiality
+possible, enforces a basic ``stepwise'' versions of our confidentiality
 and integrity properties. Their more performant ``lazy'' micro-policy
-does not, but inspires weaker, more extensional ``observational'' variants of the
+does not, but corresponds to weaker, more extensional ``observational'' variants
+of the
 properties.  While we do not rule out other enforcement mechanisms, our
 definitions are strongly informed by micropolicy systems, especially
-our attacker model.
-
-For the sake of concreteness, most of our presentation assumes a standard RISC
-machine architecture with a basic subroutine calling convention (parameters
-passed in registers) and a micropolicy-like enforcement mechanism. We
-also show how to accomodate stack-allocated parameter passing and
-coroutines.
-
-\bcp{this is also a
-  good place to admit that our definitions are tuned for micropolicies, if
-  we can find a succinct way to say it.}
-\sna{See above.}
-\BCP{I added it to the last paragraph (which is pretty close to the preceding
-  paragraph :-)  We should follow this up with some discussion in the (new)
-  Assumptions section about how different enforcement mechanisms (us,
-  StkTokens, maybe others) adopt different perspectives on what the stack is
-for and how it is organized and so lead to somewhat different notions of
-low-level safety.}
+our attacker model. \BCP{I am still not clear in what way, exactly.}
+
+% For the sake of concreteness, most of our presentation assumes a standard RISC
+% machine architecture with a basic subroutine calling convention (parameters
+% passed in registers) and a micropolicy-like enforcement mechanism. We
+% also show how to accomodate stack-allocated parameter passing and
+% coroutines.
+
+% \bcp{this is also a
+%   good place to admit that our definitions are tuned for micropolicies, if
+%   we can find a succinct way to say it.}
+% \sna{See above.}
+% \BCP{I added it to the last paragraph (which is pretty close to the preceding
+%   paragraph :-)  We should follow this up with some discussion in the (new)
+%   Assumptions section about how different enforcement mechanisms (us,
+%   StkTokens, maybe others) adopt different perspectives on what the stack is
+% for and how it is organized and so lead to somewhat different notions of
+% low-level safety.}
 
@@
 
-\apt{This is a good story, but does it describe the paper we are able to produce this week?
-  We were rejected before exactly because we claimed to have
-  a general criterion, but did not in fact demonstrate its generality (which is not
-  obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
-  para about the general story.  Does the rest now avoid making (or
-  implying) claims that we can't defend?}
-\bcp{Also, I think we should save
-  WBCF for the intro---I can't find a way to discuss it here that is brief
-  and not confusing/distracting.}
-
-  % \ifaftersubmission\bcp{should we mention stktokens here?}
+% \apt{This is a good story, but does it describe the paper we are able to produce this week?
+%   We were rejected before exactly because we claimed to have
+%   a general criterion, but did not in fact demonstrate its generality (which is not
+%   obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
+%   para about the general story.  Does the rest now avoid making (or
+%   implying) claims that we can't defend?}
+% \bcp{Also, I think we should save
+%   WBCF for the intro---I can't find a way to discuss it here that is brief
+%   and not confusing/distracting.}
+
+  % \iftrue\bcp{should we mention stktokens here?}
   % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
@@
 %% caller actually accesses data that the callee has written into its stack
-%% frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
+%% frame. \iftrue\bcp{We could also hint at the lazy confidentiality
 %%   property in the same way...}\fi
@@
 
-%\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
+%\iftrue\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
 %These approaches are represented as policies that
@@
 
-\section{Key Ideas}
+\section{Key Ideas, By Example}
 \label{sec:running-example}
@@
 
-%\ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
+%\iftrue\bcp{This implicitly amounts to choosing a caller-saves register
 %  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
@@
 \label{sec:conservative}
-%
+
+\bcp{Is this the only subsection in this section?}
+
 The first micro-policy we present to enforce the definition of stack
@@
 
+\BCP{This section seems incomplete?}
+
   %% Mapping the underlying protection

### 2021-04-14 tex/main.tex
CP

@@
 This differs from common practice in language-based security in that the
-confidentiality property applies to many nested subtraces within a single execution.
-\sna{How novel is this nesting?}
+confidentiality property applies to many nested subtraces within a single execution.\sna{How novel is this nesting?}\bcp{I believe completely novel?}
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
@@
 
-\section{Assumptions and Threat Model}
+\section{Set Up and Threat Model}
 \label{sec:threat}
-
-\BCP{I would still advocate splitting this into two sections, just so the
-  transition from one to the other and the division of responsibility are
-  clear.  The first section could be called ``Setup'' or something like
-  that.}
 
@@
 assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter \(\PCname\); other
+machine with a program counter (\(\PCname\)); other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
 the layout of memory, including where code lives and which addresses are within
-the bounds of the stack.  \BCP{What should we say about the heap?}
-
-Our stack-safety concepts depend
+the bounds of the stack.\bcp{This implies no run-time code generation,
+  right?  This should probably be called out.}
+
+\BCP{What should we say about the heap?}
+
+Our stack-safety criteria depend
 critically on the notions of ``call'' and ``return,'' which are not fully explicit
@@
 %
-So we assume we are given a machine language program annotated with the locations of
-instructions representing calls, returns, and function entry points. Although these
+So we assume we are given a machine language program {\em annotated} with the locations of
+instructions representing calls, returns, and function entry points\BCP{I
+  still want to talk about whether returns and entry points need to be
+  annotated.  To me, these annotations do not make intuitive sense: it is
+  the caller's responsibility to know whether the address that it is calling
+is really an entry point, and we can trust it to check this.  Similarly, it
+is not the callee's role to say when it is returning; it is the caller's
+prerogative to decide when it has been returned to!  I.e., from the caller's
+point of view, it has been returned to when control resumes at the
+instruction following the call with the SP reset to where it was at the
+point of the call. (This is the point where it wants to be able to access
+its own local state again and expected to find it unmolested.)}. Although these
 annotations would typically come from a compiler together with the machine code,
@@
 
-We further parameterize over the structure of the stack, particularly the means by
+\bcp{Can this be postponed to the formal sections?}We further parameterize over the structure of the stack, particularly the means by
 which a caller identifies the data that should be protected during a call. We term
@@
 
-The separation is less useful in protection mechanisms that interact more directly
-with application code, such as Cheri-style capabilities\citep{??}. Here
+The separation is less useful\bcp{but still sensible, right} in protection mechanisms that interact more directly
+with application code, such as Cheri-style capabilities\citep{??}\bcp{This
+  section would read better if we could cleanly separate the general
+  definitions from examples}. Here
 the policy would not have auxiliary state, but would still refine the machine step
@@
 
-Intuitively, stack safety means that a caller's stack frame is protected
-from a confused or malicious callee's attempts to read or write its private data.
+Intuitively, stack safety means that the stack frame of each caller (and its
+those of its ancestors) is protected
+from attempts by a confused or malicious callee (including all its descendants)
+to read or write its private data. 
 The attacker model therefore treats every caller as a defender and its callee as
-an attacker, meaning that with nested calls, a single function may play both roles
-at varying points.
+an attacker, meaning that with nested calls, a single function invocation
+may be viewed in both ways.
 
 We take a very strong threat model in which both caller and callee may execute
-arbitrary machine code. Hardware and timing attacks are out of scope, however.
+arbitrary machine code. Hardware and timing attacks are out of scope.
 
 \paragraph*{Well-formedness and Bad Annotations}
+
+\BCP{I believe we could delete this section---and never mention WBCF at all
+  (except to say that we leave it out, compared to STkTokens)---if we
+  eliminated return (and entry point) annotations.}
 
@@
 
-In general, regardless of enforcement mechanism, we take this approach:
+\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
 a given enforcement mechanism may place a static well-formedness criterion
@@
 reasonableness conditions to prevent the defender from sabotaging enforcement, as
-discussed in section \ref{sec:cheri}.
+discussed in section \ref{sec:cheri}.  \BCP{Weak.}
 
@@
 of bad behavior that it prevents; we will insert our model between the enforcement and
-the examples, arguing that:
+the examples\bcp{Didn't understand that}, arguing that:
 
@@
 Our examples are written in C-like syntax, labeled with line numbers for reference,
-but to reflect the increased power of the underlying assembly code, we allow code
+but to reflect the power of the underlying assembly code, we allow code
 to read the stack pointer ({\tt \$SP}) directly and use
@@
 respectively. Local variables are assumed to be stored in the stack frame,
-in declaration order. Unless otherwise stated, arguments and return values are passed in registers.
+in declaration order. \bcp{This sort of thing implies that readers will have
+to mentally compile the source programs we're showing them.}Unless otherwise stated, arguments and return values are passed in registers.
 The machine is word-addressed. The stack grows upward in
@@
 the specified line number is executed; each box represents a memory word and
-addresses increase to the right.
+addresses increase to the right.  \BCP{I am worried about the C-like syntax:
+I think the mental complexity and imprecision that it introduces may not be
+worth the concision that we get from it.  Couldn't we get most of the
+concision by just assuming an ISA with CISC-style single-instruction Call
+and Return?}
 
@@
 meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
+{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of
+their owner\bcp{are we still using depths?},
 in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
@@
 \end{subfigure}
-\caption{A Confidentiality Violation}
+\caption{A Confidentiality Violation \bcp{It would help if the SPs could
+    have a little arrow after them pointing to the cell that they are
+    associated with.}}
 \label{fig:conf1}
@@
 Meanwhile confidentiality intuitively states that a callee is unaware
-of the context of its call (except for its arguments and any global variables).
+of the context of its call (except for its arguments and any global
+variables\bcp{We haven't talked about global variables.  We should.}).
 We state this in a style similar to {\em noninterference}~\citep{??}.

### 2021-04-14 tex/main.tex
Notes on enforcement, a few nits

@@
 A second challenge is that
-enforcement mechanisms may weaken their protections that in the name of efficiency,
+enforcement mechanisms may weaken their protections in the name of efficiency,
 while arguing that they still capture stack safety in a meaningful sense.
@@
 At each step of the machine, the relevant rule is applied to the tags on
-the instruction's inputs (PC, registers, memory) and on the instruction itself, and produces
+the instruction's inputs ($\PCname$, registers, memory) and on the instruction itself, and produces
 one of two outcomes: either the instruction is allowed to execute (and generates
-tags for the result of the operation and the new PC), or the machine
+tags for the result of the operation and the new $\PCname$), or the machine
 fail-stops with a policy violation.
@@
 %\rb{actually, the ``initial frame'' should be tagged with its proper depth}
-and the {\PCname} has tag $\tagPCDepth{0}$. These tags are altered
+and the {\PCname} has tag $\tagPCDepth{0}$, where $n$ is the current activation depth.
+These tags are altered
 at exactly those points in
@@
   $\PCname$ tag is decremented.
+  \rb{Problem: this is currently the first mention of $\ret{\cdot}$}
 
@@
 This discipline suffices to enforce local stack encapsulation.
-To implement it,  we define ``blessed'' instruction sequences
+To implement it,  we define blessed instruction sequences
 intended to appear at the entry and exit of each function,
@@
   is clear}
+\rb{TODO: References to the running example are no longer valid! We could use
+figures 5 or 6, referencing the figures, not the invisible assembly code.}
 %

### 2021-04-14 tex/main.tex

@@
 \acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{SPLASH} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
@@
 on initial states: annotated instructions will initially be tagged apropriately, as the
-final instruction in the call or return sequence, and all blessed sequences will end with
+final instruction in the call or return sequence, and all blessed\bcp{undefined} sequences will end with
 annotated instructions. This is all we need for the policy to
@@
 variables\bcp{We haven't talked about global variables.  We should.}).
-We state this in a style similar to {\em noninterference}~\citep{??}.
+We state this in a style similar\bcp{??} to {\em noninterference}~\citep{??}.
 Noninterference describes knowledge: it states that for some system components
@@
 
-We alternatively could imagine a version of stepwise confidentiality in which we explicitly
+\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
 track which addresses may be read (i.e., which have been initialized), and at each step
@@
 One of the most pernicious ways to attack the stack is by hijacking control flow.
-Our stack safety properties do not assume that control flow is well-bracketed; rather,
+Our stack safety properties do not assume that control flow is
+well-bracketed\bcp{undefined term}; rather,
 they express what it means to protect data even in the presence of control-flow attacks.
-Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}.
+Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''\bcp{??}~\citep{??}.
 (Here we show return addresses explicitly as line numbers, in the form $\#n$.)
@@
 execute a return-labeled instruction: it must actually reach the appropriate
-{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep a stack of return targets,
+{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep
+a stack of return targets\bcp{do we still},
 which correspond to the depths at which stack frames have been sealed. A callee is
 only considered to have returned when it executes a return-annotated instruction
-and then reaches a valid return target. Only then do the sealed frames become unsealed.
+and then reaches a valid return target\bcp{what's that?}. Only then do the sealed frames become unsealed.
 
@@
 \paragraph*{Pass by Reference and Stack Arguments}
+\bcp{I wonder if this could be postponed so that we get to the formal
+  definitions for the basic property sooner...}
 
@@
 passed, and our diagram displays it in magenta. The caller's frame locations that are not
-passed are sealed as usual. Just as we marked sealed memory with the depth of its owner,
+passed are sealed as usual. Just as we marked sealed memory with the depth of its owner\bcp{??},
 we do the same for passed memory -- but it has multiple owners. We keep track of the
@@
 
-We next describe our machine model. To make our definition of stack safety
+We next define our machine model precisely. To make our definition of stack safety
 as generic as possible, this section proposes an abstract interface to a
 machine model plus some kind of (software- or hardware-enforced) policy monitor.
-In~\cref{sec:enforcement}, we will instantiate this model with a concrete dynamic
+In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
 policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
@@
 %
-Our machine states are composed of {\em components} \(\component\), addressed either by
+Our machine states are composed of {\em components} \(\component\),
+addressed\bcp{no, the components are not addressed, they are addresses!} either by
 words (memory addresses) or by register names (\(\reg\)) drawn from some set
@@
 We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
-A policy is a general model of an enforcement mechanism, consisting of a set of policy
+A {\em policy} consists of a set of policy
 states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
@@
 can see the machine state, but cannot modify it. We assume that the
-compiler provides an initial policy state \(\pol_0\) to accompany the initial machine state.
+compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.
 %
@@
 information representing the history of the run. The following definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+over an arbitrary\bcp{why?} set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
 an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
@@
 We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).
+a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
+\CONTEXTS\).  \bcp{I don't get the intuition for this whole para.}
 
@@
 
-A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
+A trace over a type \(\TRACEELEMS\)\bcp{``over a type A'' is a Coq-ism} is a nonempty, finite or infinite sequence of
 \(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
@@
 the policy component. The {\sc Default} rule applies only if the other does not.
-(We use juxtaposition of an element and a trace to represent ``cons.'')
+(We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
 
@@
 
+\bcp{Could any of this be postponed?}
+
 The ``observations-of'' operator, written \(\pi_\obs\), \apt{since this isn't
@@
 individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
+at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
@@
 then dive into the formal definitions of local state encapsulation and
-well-bracketed control flow, which together constitute stack safety.
+well-bracketed control flow, which together constitute stack safety\bcp{not
+  any more?}.
 Initially we will focus on a simple stack model, with a single stack
@@
 of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller will register a {\em target} condition -- a
+at a particular depth\bcp{What about code??}. Each caller will register a {\em target} condition -- a
 predicate on machine states that must hold for it to consider itself active.
@@
 ``return target list,'' which is a stack of targets, one for each caller awaiting
-return.
+return.  \ifaftersubmission\bcp{spacing is funny}\fi
 
@@
 \definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
-change together, written as a coinductive relation \(\machT \doteq \nachT\):
+change together, written as a coinductive relation \(\machT \doteq
+\nachT\):\bcp{People may not be familiar with the notation}
 
@@
 local state encapsulation component by checking for integrity and
-confidentiality violations at every step of the callee's execution trace.
+confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}
 
@@
 fail-stops with a policy violation.
-\citeauthor{pump_oakland2015} have shown that a wide range of micro-policies can
+\citeauthor{pump_oakland2015}\bcp{and PIPE, if there are any references to
+  cite} have shown that a wide range of micro-policies can
 be defined using this scheme.
@@
 This discipline suffices to enforce local stack encapsulation.
-To implement it,  we define blessed instruction sequences
+To implement it,  we define ``blessed'' instruction sequences
 intended to appear at the entry and exit of each function,
@@
 programs. Although the required hardware support is readily available
-in capability machines like CHERI\bcp{including linear capabilities?? or
-  don't they use them in this paper?}, this technique incurs significant
-costs\bcp{This is the pot calling the kettle black}, because it requires the entire unused part of the stack to be
+in capability machines like CHERI, this technique incurs significant
+costs, because it requires the entire unused part of the stack to be
 cleared whenever a security boundary is crossed. Their logical
@@
 
-We believe that the linear capability machines introduced by
+We conjecture that the linear capability machines introduced by
 \citep{Skorstengaard+19}, in combination with the StkTokens calling convention,
@@
 \cref{sec:lse-and-wbcf}. Further, we expect them to satisfy the even stronger
-property developed in \cref{sec:enforcement}. Proving these conjectures would
-involve materializing their definitions of local state encapsulation and
-well-bracketed control flow, which exist only in enmeshed form as part of the
+property developed in \cref{sec:enforcement}. Proving\bcp{or testing!} these conjectures would
+involve materializing\bcp{?} their definitions of local state encapsulation and
+well-bracketed control flow, which exist only in enmeshed\bcp{?} form as part of the
 semantics of the capability machines. The main practical limitation of StkTokens
 is its reliance on linear capabilities, as it is unclear now they could be added
-to practical capability machines, especially in terms of efficiency---previous
+to practical capability machines, especially in terms of efficiency (previous
 work on micro-policies \citep{yannis-report} has shown how to use that framework
 to implement linear return capabilities.\bcp{But IIUC the Dover realization
-  of micro-policies can't do this...?}
+  of micro-policies can't do this...?})
 
@@
 Heap safety, like stack safety, may be framed as a security property
-in the form of noninterference~\citep{DBLP:conf/post/AmorimHP18}. Just
-like~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
+in the form of
+noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
+be ``Azedevo et al.''}\fi. Just
+as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
 characterization of the meaning of (heap) memory safety, in this paper
@@
 Their model describing safety of shared stack-allocated
-objects extends naturally to heap safety, complicated by the fact that
-these objects are not deallocated by returns, and therefore a
+objects\bcp{??} extends naturally to heap safety, complicated by the fact that
+these objects are not deallocated by returns\bcp{??}, and therefore a
 function's privilege may increase or decrease after its entry
@@
 properties separately on continuous segments of each call, with
-contours computed at each crossing between caller and callee.
+contours computed at each crossing between caller and callee.  \bcp{Didn't
+  understand last sentence.}
 

### 2021-04-14 tex/main.tex
Pushed through section 3 (except the parts that are now section 4)

--- 2021-04-14 tex/main.tex
@@
 \newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}
+\newcommand*{\rsp}{\textsc{sp}}
 
@@
 memory.  But these mechanisms typically lack rigorous specifications,
-relying instead on informal descriptions and specific examples of the bad behaviors that they prevent.
-
-We propose formal criteria for stack safety, using the concepts of
+relying instead on informal descriptions of
+the safe stack manipulation behaviors that they allow
+and/or specific examples of the bad behaviors that they prevent.
+
+We propose formal criteria for stack safety, phrased in terms of
 language-based security: a combination of an integrity property (``each
@@
 %
-We argue that these criteria match common intuitions by (1) showing that
-they are violated by a variety of intuitively unsafe examples and (2)
+We argue that these criteria match common intuitions (1) by showing that
+they are violated by a variety of intuitively unsafe examples and (2) by
 using property-based random testing to check that appropriate variants of
@@
 proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
-broken variants).  Specifically, Roessler
+broken versions of these enforcement mechanisms).  Specifically, Roessler
 and Dehon's ``eager'' micro-policy, which catches violations as early as
-possible, enforces a basic ``stepwise'' versions of our confidentiality
+possible, enforces the basic ``stepwise'' versions of our confidentiality
 and integrity properties. Their more performant ``lazy'' micro-policy
-does not, but corresponds to weaker, more extensional ``observational'' variants
-of the
+does not, but inspires weaker, more extensional ``observational'' variants of the
 properties. While we do not rule out other enforcement mechanisms, our
 definitions are strongly informed by micropolicy systems, especially
-our attacker model. \BCP{I am still not clear in what way, exactly.}
-
-% For the sake of concreteness, most of our presentation assumes a standard RISC
-% machine architecture with a basic subroutine calling convention (parameters
-% passed in registers) and a micropolicy-like enforcement mechanism. We
-% also show how to accomodate stack-allocated parameter passing and
-% coroutines.
-
-% \bcp{this is also a
-%   good place to admit that our definitions are tuned for micropolicies, if
-%   we can find a succinct way to say it.}
-% \sna{See above.}
-% \BCP{I added it to the last paragraph (which is pretty close to the preceding
-%   paragraph :-)  We should follow this up with some discussion in the (new)
-%   Assumptions section about how different enforcement mechanisms (us,
-%   StkTokens, maybe others) adopt different perspectives on what the stack is
-% for and how it is organized and so lead to somewhat different notions of
-% low-level safety.}
-
-% We argue \apt{do we?}\bcp{I guess we will not have time to.  Propose
-%   deleting this para.} that our properties are suited to a range of other
-% enforcement mechanisms. The strict stepwise properties can describe
-% mechanisms that catch errors immediately, as is typical. While the
-% observational properties are closely linked to lazy micro-policies they are
-% weak enough to be a lower bound that captures stack safety for any
-% reasonable enforcement mechanism.
-
-% \apt{This is a good story, but does it describe the paper we are able to produce this week?
-%   We were rejected before exactly because we claimed to have
-%   a general criterion, but did not in fact demonstrate its generality (which is not
-%   obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
-%   para about the general story.  Does the rest now avoid making (or
-%   implying) claims that we can't defend?}
-% \bcp{Also, I think we should save
-%   WBCF for the intro---I can't find a way to discuss it here that is brief
-%   and not confusing/distracting.}
-
-  % \iftrue\bcp{should we mention stktokens here?}
-  % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
-  % \bcp{Some people say not to, but I think their argument is weak: I've
-  %   added Nick/Andre above.  But I'm not sure we want to get into stkTokens
-  %   in the abstract.}
-  % \fi
+our attacker model.
+
+For the sake of concreteness, most of our presentation assumes a standard RISC
+machine architecture with a basic subroutine calling convention (parameters
+passed in registers) and a micropolicy-like enforcement mechanism. We
+also show how to accomodate stack-allocated parameter passing and
+coroutines.
+
+\bcp{ We should follow this up with some discussion in the (new)
+  Assumptions section about how different enforcement mechanisms (us,
+  StkTokens, maybe others) adopt different perspectives on what the stack is
+for and how it is organized and so lead to somewhat different notions of
+low-level safety.}
+
 \end{abstract}
@@
 \citet{Skorstengaard+19} formalize these properties by defining an idealized
-machine in which they hold {\em by construction}; an enforcement mechanism (in their case, hardware
-capabilities)
-can then be shown to guarantee stack safety by exhibiting a fully abstract
-embedding function
-from the idealized machine to an ordinary machine equipped with this
-enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
-  understand that sentence.}
+machine in which they hold {\em by construction};
+they then show that an enforcement
+mechanism (in their case, hardware capabilities) guarantees stack safety by
+proving that an ordinary machine with enforcement can simulate the idealized machine.
+%an enforcement mechanism (in their case, hardware
+%capabilities)
+%can then be shown to guarantee stack safety by exhibiting a fully abstract
+%embedding function
+%from the idealized machine to an ordinary machine equipped with this
+%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
+%  understand that sentence.}
 %%
@@
 This differs from common practice in language-based security in that the
-confidentiality property applies to many nested subtraces within a single execution.\sna{How novel is this nesting?}\bcp{I believe completely novel?}
+confidentiality property applies to many nested subtraces within a single execution.
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
@@
 violating the integrity and confidentiality of the caller. Second, confidentiality
-contains temporal protections that LSE does not implicitly address.
-\apt{Maybe want to say a little more about this here, since it is a novelty wrt/ existing work?}
-\bcp{+1}\sna{TODO: Say a little more, but make sure it's accurate first.}
+contains temporal protections that LSE does not necessarily address. The intuition
+for LSE is that it protects the caller's stack frame, not necessarily memory that was
+previously allocated but now outside of the caller's frame. (Cheri-capability-based
+implementations of LSE {\em do} protect previously allocated data, typically by clearing
+it, but because it is a potential vector for leaking capabilities into the caller's frame,
+not as a goal of the property.)
+
+To illustrate the applicability of these properties, we perform randomized testing to
+validate that an existing enforcement mechanism does enforce them.
+\citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
+using tag-based \emph{micro-policies}~\citep{pump_oakland2015}, but offer no formal properties.
+We test against their {\em Depth Isolation} micro-policy, in which memory cells within each stack
+frame are tagged with the identity of the function activation
+that owns the frame, and access to those locations is then gated
+according to the activation that is currently executing. This is simple and corresponds
+to intuitive notions of stack safety, and we argue that by describing its behavior
+we approach a reasonable stack-safety property that other micro-policies can be
+validated against.
+
+Randomized testing is a relatively low-cost way to validate a potential enforcement
+mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
+of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
+covers potentially erroneous programs, we test intentionally buggy policies alongside
+the real one. If we can catch known bugs, we gain confidence that we can catch unknown ones,
+strengthening the validation of the enforcement mechanism.
 
@@
 
-A second challenge is that
-enforcement mechanisms may weaken their protections that in the name of efficiency,
-while arguing that they still capture stack safety in a meaningful sense.
-For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
-using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
-Memory cells within each stack frame are tagged with the
-identity of the function activation
-that owns the frame, and access to those locations is then gated
-according to the activation that is currently executing. Roessler and DeHon
-explore alternative versions of this basic idea, with different ways of identifying
-functions and deciding exactly which accesses are allowed.
+A second challenge is that enforcement mechanisms may weaken their protections
+in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
+\citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
+ways of identifying functions and deciding exactly which accesses are allowed.
 These vary in their efficiency and in the spatial and temporal precision of their
 security guarantees.
-The \emph{Depth Isolation} policy prevents a callee from writing outside its frame
+The Depth Isolation policy prevents a callee from writing outside its frame
 at all, which is simple and intuitively correct, but it must manually clear the tags
-on the frame on each return, which is very expensive. Conversely, the
+on the frame on each return, which is very expensive. Conversely, their
 \emph{Lazy Tagging and Clearing} policy improves
@@
 permits  ``illicit'' writes to inaccessible stack frames so long as the
-written locations are not subsequently read by their ``real'' owners. Intuitively,
+written locations are not subsequently read by their ``real'' owners. This violates our
+straightforward properties, but ntuitively,
 this form of laziness is justified if any dangerous writes will
@@
 
-But what does ``affecting program behavior'' really mean? This micro-policy
+What does ``affecting program behavior'' really mean? This micro-policy
 calls for a second characterization of stack safety in terms of the externally
@@
 can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
-integrity implies observational integrity.
-\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
-  established that this was important.\apt{medium importance. it is obvious that
-    observational integerity allows behaviors that stepwise does not, but
-    maybe not obvious that the converse is not true too.}}
+integrity implies observational integrity. \sna{TODO: check this.}
+
 We additionally sketch an associated {\em observational confidentiality} property
-makes confidentiality extensional in the same way, although we do not know of a real
-enforcement mechanism that weakens confidentiality in this way.
-\apt{why bother mentioning it then?}
-\sna{If a relatively extensional characterization is a worthwhile contribution
-  in its own right, then it's worth mentioning that we have one. If not, then indeed,
-  we shouldn't mention it -- or get too excited about extensionality in general, because
-  lockstep isn't extensional.}
-Together, these form an {\em observational stack safety} that is highly extensional.
+makes confidentiality extensional in the same way, in order to form an
+{\em observational stack safety} that is highly extensional.
 We therefore argue that, despite its somewhat greater complexity,
@@
 
-\BCP{This paragraph doesn't apply to the lazy properties, so maybe it should
-be moved before they are introduced?}%
-To illustrate the applicability of these properties, we perform randomized testing to
-validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
-them. Randomized testing is a relatively low-cost way to validate a potential enforcement
-mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
-of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
-covers potentially erroneous programs, we test intentionally buggy policies alongside
-the real one. If the testing can differentiate correct implementations from buggy ones,
-it is an argument in favor of their correctness. \bcp{Hmm--don't understand
-  that sentence}
 %\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
@@
 %  ones...''}
-Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
-Clearing micro-policy does not enforce stepwise integrity.
-Unfortunately the existing Lazy Tagging and Clearing
-micro-policy does not quite enforce observable confidentiality either. %, as we show in
-%testing~\apt{really??},
+As it turns out, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
+Clearing micro-policy enforces observational integrity, but not confidentiality,
+whether stepwise or observational.
 But formalizing the ideal property allows us to see how it falls short -- a specific
-violation of the temporal aspect of confidentiality.
+violation of the temporal aspect of confidentiality occurs in corner cases where callees
+can leak data between one another.
 We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
-observational integrity, at some performance cost.\apt{but we're not testing that, right?}
-\rb{Not sure about the property part right now, the machine side of this is
-  pretty much done}
-\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
-  need a caveat about how it's harder to test?}\apt{As per email discussion,
-  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and
-  Clearing, by loosening the temporal safety properties?}
-\sna{I'm pretty sure for the main policy it's just lazy integrity, no confidentiality.
-  There was a variant Nick and Andre gave that would restrict the leaks to only be between
-  instances of the same function, so that's the same property under a weaker attacker model,
-  where we don't protect a function from itself.}
+confidentiality, albeit at some performance cost.
 
@@
 callee access to the caller's data. We do not model more sophisticated sharing,
-such as taking the address of a variable and treating it like a pointer. Doing so without
-protecting the variable is not interesting, and protecting would be better suited to
-a different kind of enforcement, such as a memory safety micro-policy or Cheri-style capability.
-\bcp{This is all a bit mysterious} \sna{Any less mysterious now?}
-
-\rb{So no pointer passing? Pass by reference is now discussed at the end of
-  Section 3, however it looks like things get tricky rather quickly.}
-\sna{The trick is that if a programmer is passing pointers, we can't presume their
-  intent -- how long should we permit memory to be shared? But if the compiler is passing
-  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
-  \apt{N.B.Doing pass-by-reference properly requires that the passed variable is accessible
-    by all descendents, not just immediate callee, which is doable with depth tags but not
-    obviously with per-activation tags }}
+such as taking the address of a variable and treating it like a pointer.
 
@@
 %% caller actually accesses data that the callee has written into its stack
-%% frame. \iftrue\bcp{We could also hint at the lazy confidentiality
+%% frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
 %%   property in the same way...}\fi
@@
 
-\section{Set Up and Threat Model}
-\label{sec:threat}
-
-\newcommand*{\rsp}{\textsc{sp}}
+\section{Setup}
+\label{sec:setup}
 
@@
 assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter (\(\PCname\)); other
+machine with a program counter \(\PCname\); other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
 the layout of memory, including where code lives and which addresses are within
-the bounds of the stack.\bcp{This implies no run-time code generation,
-  right?  This should probably be called out.}
-
-\BCP{What should we say about the heap?}
-
-Our stack-safety criteria depend
+the bounds of the stack. We do not model a heap; programs might choose to use memory
+that is neither code nor stack to store global variables, as a heap, etc., but our
+properties do not protect it or assume any particular behavior.
+
+Our stack-safety concepts depend
 critically on the notions of ``call'' and ``return,'' which are not fully explicit
@@
 %
-So we assume we are given a machine language program {\em annotated} with the locations of
-instructions representing calls, returns, and function entry points\BCP{I
-  still want to talk about whether returns and entry points need to be
-  annotated.  To me, these annotations do not make intuitive sense: it is
-  the caller's responsibility to know whether the address that it is calling
-is really an entry point, and we can trust it to check this.  Similarly, it
-is not the callee's role to say when it is returning; it is the caller's
-prerogative to decide when it has been returned to!  I.e., from the caller's
-point of view, it has been returned to when control resumes at the
-instruction following the call with the SP reset to where it was at the
-point of the call. (This is the point where it wants to be able to access
-its own local state again and expected to find it unmolested.)}. Although these
+So we assume we are given a machine language program annotated with the locations of
+instructions representing calls, returns, and function entry points. Although these
 annotations would typically come from a compiler together with the machine code,
@@
 
-\bcp{Can this be postponed to the formal sections?}We further parameterize over the structure of the stack, particularly the means by
+We further parameterize over the structure of the stack, particularly the means by
 which a caller identifies the data that should be protected during a call. We term
@@
 
-\sna{This is awkward.}
-This division of policy state from machine state is helpful when reasoning about
+Dividing policy state from machine state helps us reason about
 enforcement mechanisms that keep a significant amount of data separate from the
@@
 
-The separation is less useful\bcp{but still sensible, right} in protection mechanisms that interact more directly
-with application code, such as Cheri-style capabilities\citep{??}\bcp{This
-  section would read better if we could cleanly separate the general
-  definitions from examples}. Here
+The separation is less useful in protection mechanisms that interact more directly
+with application code, such as Cheri-style capabilities\citep{??}. Here
 the policy would not have auxiliary state, but would still refine the machine step
@@
 
-\paragraph*{Threat Model}
-
-Intuitively, stack safety means that the stack frame of each caller (and its
-those of its ancestors) is protected
-from attempts by a confused or malicious callee (including all its descendants)
-to read or write its private data. 
+\section{Threat Model}
+\label{sec:threat}
+
+Intuitively, stack safety means that a caller's stack frame is protected
+from a confused or malicious callee's attempts to read or write its private data.
 The attacker model therefore treats every caller as a defender and its callee as
-an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways.
+an attacker, meaning that with nested calls, a single function may play both roles
+at varying points.
 
 We take a very strong threat model in which both caller and callee may execute
-arbitrary machine code. Hardware and timing attacks are out of scope.
+arbitrary machine code. Hardware and timing attacks are out of scope, however.
 
 \paragraph*{Well-formedness and Bad Annotations}
-
-\BCP{I believe we could delete this section---and never mention WBCF at all
-  (except to say that we leave it out, compared to STkTokens)---if we
-  eliminated return (and entry point) annotations.}
 
@@
 
-\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
+In general, regardless of enforcement mechanism, we take this approach:
 a given enforcement mechanism may place a static well-formedness criterion
@@
 reasonableness conditions to prevent the defender from sabotaging enforcement, as
-discussed in section \ref{sec:cheri}.  \BCP{Weak.}
+discussed in section \ref{sec:cheri}.
 
@@
 
-%\iftrue\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
+%\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
 %These approaches are represented as policies that
@@
 of bad behavior that it prevents; we will insert our model between the enforcement and
-the examples\bcp{Didn't understand that}, arguing that:
+the examples, arguing that:
 
@@
 Our examples are written in C-like syntax, labeled with line numbers for reference,
-but to reflect the power of the underlying assembly code, we allow code
+but to reflect the increased power of the underlying assembly code, we allow code
 to read the stack pointer ({\tt \$SP}) directly and use
@@
 respectively. Local variables are assumed to be stored in the stack frame,
-in declaration order. \bcp{This sort of thing implies that readers will have
-to mentally compile the source programs we're showing them.}Unless otherwise stated, arguments and return values are passed in registers.
+in declaration order. Unless otherwise stated, arguments and return values are passed in registers.
 The machine is word-addressed. The stack grows upward in
@@
 the specified line number is executed; each box represents a memory word and
-addresses increase to the right.  \BCP{I am worried about the C-like syntax:
-I think the mental complexity and imprecision that it introduces may not be
-worth the concision that we get from it.  Couldn't we get most of the
-concision by just assuming an ISA with CISC-style single-instruction Call
-and Return?}
+addresses increase to the right.
 
@@
 
-%\iftrue\bcp{This implicitly amounts to choosing a caller-saves register
+%\ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
 %  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
@@
 meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of
-their owner\bcp{are we still using depths?},
+{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
 in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
@@
 \end{subfigure}
-\caption{A Confidentiality Violation \bcp{It would help if the SPs could
-    have a little arrow after them pointing to the cell that they are
-    associated with.}}
+\caption{A Confidentiality Violation}
 \label{fig:conf1}
@@
 Meanwhile confidentiality intuitively states that a callee is unaware
-of the context of its call (except for its arguments and any global
-variables\bcp{We haven't talked about global variables.  We should.}).
+of the context of its call (except for its arguments and any global variables).
 We state this in a style similar to {\em noninterference}~\citep{??}.
@@
 
+We will describe observable integrity in our most detailed setting. In the
+presence
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\sealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(|\rtm ~ \stackid| < \depth\) or \(\depth' < |\rtm ~ \stackid|\) .
+Then a system enjoys observable
+
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\_}\).
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise coroutine confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
 \subsection{Observable Confidentiality}

### 2021-04-14 tex/main.tex

--- 2021-04-14 tex/main.tex
@@
 definitions are strongly informed by micropolicy systems, especially
-our attacker model. \BCP{I am still not clear in what way, exactly.}
+our attacker model, which is stronger than other enforcement mechanisms
+can enable.
 
@@
 from its callee, even as that callee may need protection from its callee in turn.
-This differs from common practice in language-based security in that the
-confidentiality property applies to many nested subtraces within a single execution.\sna{How novel is this nesting?}\bcp{I believe completely novel?}
+This is novel in that the
+confidentiality property applies to many nested subtraces within a single execution,
+as opposed to traces of entire executions.
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
@@
 
-A second challenge is that
-enforcement mechanisms may weaken their protections in the name of efficiency,
-while arguing that they still capture stack safety in a meaningful sense.
-For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
-using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
-Memory cells within each stack frame are tagged with the
-identity of the function activation
-that owns the frame, and access to those locations is then gated
-according to the activation that is currently executing. Roessler and DeHon
-explore alternative versions of this basic idea, with different ways of identifying
-functions and deciding exactly which accesses are allowed.
+A second challenge is that enforcement mechanisms may weaken their protections
+in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
+\citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
+ways of identifying functions and deciding exactly which accesses are allowed.
 These vary in their efficiency and in the spatial and temporal precision of their
 security guarantees.
-The \emph{Depth Isolation} policy prevents a callee from writing outside its frame
+The Depth Isolation policy prevents a callee from writing outside its frame
 at all, which is simple and intuitively correct, but it must manually clear the tags
-on the frame on each return, which is very expensive. Conversely, the
+on the frame on each return, which is very expensive. Conversely, their
 \emph{Lazy Tagging and Clearing} policy improves
@@
 the layout of memory, including where code lives and which addresses are within
-the bounds of the stack.\bcp{This implies no run-time code generation,
-  right?  This should probably be called out.}
-
-\BCP{What should we say about the heap?}
+the bounds of the stack. We do not model a heap; programs might choose to use memory
+that is neither code nor stack to store global variables, as a heap, etc., but our
+properties do not protect it or assume any particular behavior. We also do not
+model run-time code generation.
 
@@
 
-\paragraph*{Threat Model}
+\section{Threat Model}
+\label{sec:threat}
 

### 2021-04-14 tex/main.tex
machine code for first example

--- 2021-04-14 tex/main.tex
@@
   {
-    \MemoryLabel{#2}{2em}{\SP #1}
+    \MemoryLabel{#2}{2em}{$\downarrow$\SP #1}
   }
@@
 
-\begin{subfigure}{.4\textwidth}
+\begin{subfigure}{.2\textwidth}
+{\small
 \begin{verbatim}
 int main() {
-0:     int x, y;
-1:     y = 42;
-2[C]:  x = f();
-3[R]:  return x+y;
+  int x = 42;
+  return f()+x;
 }
@@
 int f() {
-10:     *($SP-2) = 0;
-11[R]:  return 5;
+  int y;
+  *(&y-3) = 0;
+  return 5;
 }
 \end{verbatim}
+}
 \end{subfigure}
-\begin{subfigure}{.4\textwidth}
+\begin{subfigure}{.5\textwidth}
+{\small
+\begin{verbatim}
+1 main: add $1,%sp      ; allocate frame
+2       mov $42,-1(%sp) ; initialize local
+3       call f
+4       add -1(%sp),%ra ; set return value
+5       sub $1,%sp      ; deallocate frame
+6       ret
+7 f:    add $1,%sp      ; allocate frame
+8       mov $0,-3(%sp)  ; violation!
+9       mov $5, %ra     ; set return value
+10      sub $1,%sp      ; deallocate frame
+11      ret
+\end{verbatim}
+}
+\end{subfigure}
+\begin{subfigure}{.25\textwidth}
 \begin{center}
 \begin{tabular}{l l}
-{\tt 1} &
-\memoryaddrs{12em}
-\memory{4}{\unsealc}[Unseal]
+{\tt 2} &
+\memoryaddrs{8em}
+\memory{3}{\unsealc}[Unseal]
 ~$\cdots$
@@
 \\
-{\tt 2} &
-\memoryaddrs{12em}
-\memory{4}{\unsealc}
+{\tt 3} &
+\memoryaddrs{8em}
+\memory{3}{\unsealc}
 ~$\cdots$
@@
 \\
-{\tt 10} &
+{\tt 8} &
 \memoryaddrs{16em}
-\memory{2}{\mainsealc}[Seal(0)]%
+\memory{1}{\mainsealc}[Seal(0)]%
 \memory{1}{\retptrc}[RetPtr]%
@@
 \\
-{\tt 11} &
+{\tt 9} &
 \memoryaddrs{16em}
-\memory{2}{\mainsealc}
+\memory{1}{\mainsealc}
 \memory{1}{\retptrc}
@@
 
-Our examples are written in C-like syntax, labeled with line numbers for reference,
-but to reflect the power of the underlying assembly code, we allow code
-to read the stack pointer ({\tt \$SP}) directly and use
-it to form pointers.
-Call and return annotations are marked {\tt[C]} and {\tt[R]}
-respectively. Local variables are assumed to be stored in the stack frame,
-in declaration order. \bcp{This sort of thing implies that readers will have
-to mentally compile the source programs we're showing them.}Unless otherwise stated, arguments and return values are passed in registers.
+We show our examples in C syntax and give the corresponding assembly code for a
+simple two-address machine, with {\tt call} and {\tt ret} instructions that
+push and pop the return address from the stack, implicitly adjusting the stack pointer.
+The assembly code reflects a simple compilation model in which  
+local variables are stored in the stack frame.
+Unless otherwise stated, arguments and return values are passed in registers.
 The machine is word-addressed. The stack grows upward in
@@
 At the right of each example, we show the layout of stack memory just before
-the specified line number is executed; each box represents a memory word and
-addresses increase to the right.  \BCP{I am worried about the C-like syntax:
-I think the mental complexity and imprecision that it introduces may not be
-worth the concision that we get from it.  Couldn't we get most of the
-concision by just assuming an ISA with CISC-style single-instruction Call
-and Return?}
+the specified instruction is executed; each box represents a memory word and
+addresses increase to the right.
 
@@
 in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
+{\tt main} expects that {\tt x} will be unchanged when it gets control back.
 This gives us our simple, intuitive statement of {\em stepwise integrity}:
@@
 %cares about changes it can see after the return, not changes during execution.
-In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.
-
-The black rectangle represents the return address. In a typical RISC calling convention
-this will have been saved into a register by the call itself, and {\tt f} stores
-it in the stack during its preamble, so it is not itself sealed.
+In this case, after instruction 8, {\tt f} has violated integrity by writing to {\tt x},
+so the final return value produced by {\tt main} is wrong.
+
+The black rectangle represents the return address, which is saved by the call
+instruction itself, so is not sealed.
 
@@
   {\tt 10} &
-  \memoryaddrs{20em}
+  \memoryaddrs{21em}
   \memory{1}{\unsealc}
@@
   &
-  \memoryaddrs{20em}
+  \memoryaddrs{21em}
   \memory{1}{\unsealc}
@@
   {\tt 11} &
-  \memoryaddrs{20em}
+  \memoryaddrs{21em}
   \memory{1}{\unsealc}
@@
   {\tt 12} &
-  \memoryaddrs{20em}
+  \memoryaddrs{21em}
   \memory{1}{\unsealc}

### 2021-04-14 tex/main.tex
Minor changes

--- 2021-04-14 tex/main.tex
@@
 
-The separation is less useful\bcp{but still sensible, right} in protection mechanisms that interact more directly
-with application code, such as Cheri-style capabilities\citep{??}\bcp{This
-  section would read better if we could cleanly separate the general
-  definitions from examples}. Here
-the policy would not have auxiliary state, but would still refine the machine step
-function to create a distinct failstop condition. Other techniques such as
-software fault isolation~\citep{??} might never actually failstop, but fulfill
-the properties by rewriting code so that traces can never violate them.
+The separation is smaller in protection mechanisms that interact more directly
+with application code. Some mechanisms might exhibit failstop behavior, but have
+no protected policy state; others might never even failstop. Then the policy-enhanced
+machine is simply identical to the original. This is the case for software-based
+techniques.
 

### 2021-04-14 tex/main.tex

--- 2021-04-14 tex/main.tex
@@
   int y;
-  *(&y-3) = 0;
+  *(&y-2) = 0;
   return 5;
@@
 6       ret
-7 f:    add $1,%sp      ; allocate frame
-8       mov $0,-3(%sp)  ; violation!
-9       mov $5, %ra     ; set return value
-10      sub $1,%sp      ; deallocate frame
-11      ret
+
+10 f:   add $1,%sp      ; allocate frame
+11      mov $0,-3(%sp)  ; violation!
+12      mov $5, %ra     ; set return value
+13      sub $1,%sp      ; deallocate frame
+14      ret
 \end{verbatim}
@@
 \\
-{\tt 8} &
+{\tt 11} &
 \memoryaddrs{16em}
@@
 \\
-{\tt 9} &
+{\tt 12} &
 \memoryaddrs{16em}
@@
 
-We show our examples in C syntax and give the corresponding assembly code for a
-simple two-address machine, with {\tt call} and {\tt ret} instructions that
-push and pop the return address from the stack, implicitly adjusting the stack pointer.
+We show our examples both in C syntax and in corresponding assembly code for a
+simple two-address machine. The address of each instruction is shown to its left.
 The assembly code reflects a simple compilation model in which  
@@
 memory and {\tt \$SP} points to the first unused word above the top of the stack.
+The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
+implicitly adjusting the stack pointer.
 At the right of each example, we show the layout of stack memory just before
@@
 \begin{figure}
-\begin{subfigure}{.33\textwidth}
+\begin{subfigure}{.2\textwidth}
+{\small
 \begin{verbatim}
 int main() {
-0:     int x;
-1:     x = 42;
-2[C]:  x = f();
-3[R]:  return x+y;
+  int x = 42;
+  return f()+x;
 }
@@
 int f() {
-10:     int y,z;
-11:     y = 5;
-12:     z = *($SP-4);
-13[R]:  return y+z;
+  int y,z;
+  y = 5;
+  z = *(&y-2);
+  return y+z;
 }
 \end{verbatim}
+}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+{\small
+\begin{verbatim}
+1 main: add $1,%sp      ; allocate frame
+2       mov $42,-1(%sp) ; initialize local
+3       call f
+4       add -1(%sp),%ra ; set return value
+5       sub $1,%sp      ; deallocate frame
+6       ret
+
+10 f:   add $2,%sp      ; allocate frame
+11      mov $5,-2(%sp)  ; assign into y
+12      mov -4(%sp),-1(%sp)  ; violation!
+13      mov -2(%sp),%ra ; calculate and
+14      add -1(%sp),%ra ; set return value
+15      sub $2,%sp      ; deallocate frame
+16      ret
+\end{verbatim}
+}
 \end{subfigure}
@@
 \begin{tabular}{l l | l}
-  {\tt 1} &
+  {\tt 2} &
   \multicolumn{2}{c}{
@@
     ~$\cdots$
-    \MemoryLabel{-18em}{0.75em}{0}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{0}
-    \MemoryLabel{-7em}{0.75em}{0}
+%    \MemoryLabel{-18em}{0.75em}{0}
+%    \MemoryLabel{-14em}{0.75em}{0}
+%    \MemoryLabel{-10em}{0.75em}{0}
+%    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
   } \\
-  {\tt 10} &
+  {\tt 11} &
   \memoryaddrs{21em}
@@
   \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-10em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{0}
+%  \MemoryLabel{-10em}{0.75em}{0}
+%  \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   \\
-  {\tt 11} &
+  {\tt 12} &
   \memoryaddrs{21em}
@@
   \MemoryLabel{-10em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{0}
+ % \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   \\
-  {\tt 12} &
+  {\tt 13} &
   \memoryaddrs{21em}
@@
 \end{subfigure}
-\caption{A Confidentiality Violation \bcp{It would help if the SPs could
-    have a little arrow after them pointing to the cell that they are
-    associated with.}}
+\caption{A Confidentiality Violation\apt{Could save space by not repeating C or assemly code for {\tt main}.}}
 \label{fig:conf1}
@@
 by reading {\tt main}'s variable {\tt x}. The trace
-shows how noninterference applies to this scenario. At line 11, we create a {\em variant}
+shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
 state with arbitrary values \(v_0\), etc., in place of the values that should
-be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt y}, has
+be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
 changed to 5 in both the primary trace and the variant. The primary and variant agree on
-the value in the changed location, so confidentiality was not violated. But after line 13,
+the value in the changed location, so confidentiality was not violated. But at instruction 13,
 the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
-in the variant trace it might have any value \(v_2\). If \(v_2 \neq 42\) then the variant
+in the variant trace it might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
 has behaved differently than the primary, violating confidentiality.
@@
 
-\begin{subfigure}{.3\textwidth}
+\begin{subfigure}{.2\textwidth}
+  {\small
 \begin{verbatim}
 int main() {
-0:     int x;
-1[C]:  x = f();
-2[CR]: return g();
+  int x;
+  x = f();
+  return g();
 }
@@
 int f() {
-10:     int y;
-11:     y = 5;
-12[R]:  return y;
+  int y = 5;
+  return y;
 }
@@
 int g() {
-20:     int z;
-21:     if(z == 5)
-22[R]:    return 100;
-23:     else
-24[R]:    return 10;
+  int z;
+  if (z == 5)
+    return 10;
+  else
+    return 20;
 }
 \end{verbatim}
+}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+  {\small
+\begin{verbatim}
+1 main:  add $1,%sp ; allocate frame
+2        call f
+3        mov %rax,-1(%sp) ; store x
+4        call g
+5        sub $1,%sp ; deallocate frame
+6        ret
+
+10 f:    add $1,%sp ; allocate frame
+11       mov $5,-1(%sp) ; initialize y
+12       mov -1(%sp),%rax ; set return value
+13       sub $1,%sp ; deallocate frame
+14       ret
+
+20 g:    add $1,%sp ; allocate frame
+21       cmp $5,-1(%sp) ; compare 
+22       brneq #25      ; branch if not equal
+23       mov $10,%rax ; set return value
+24       jmp #26
+25       mov $20,%rax  ; set return value
+26       sub $1,%sp ; deallocate frame
+27       ret 
+}
+\end{verbatim}
+}
 \end{subfigure}
@@
 \begin{tabular}{l r | l}
-  {\tt 01} &
+  {\tt 2} &
   \multicolumn{2}{c}{
@@
     ~$\cdots$
-    \MemoryLabel{-18em}{0.75em}{0}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{0}
-    \MemoryLabel{-7em}{0.75em}{0}
+%    \MemoryLabel{-18em}{0.75em}{0}
+%    \MemoryLabel{-14em}{0.75em}{0}
+%    \MemoryLabel{-10em}{0.75em}{0}
+%    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
@@
   ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{0}
-  \MemoryLabel{-10em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{0}
+ % \MemoryLabel{-19em}{0.75em}{0}
+%  \MemoryLabel{-10em}{0.75em}{0}
+%  \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{0}
+ % \MemoryLabel{-19em}{0.75em}{0}
   \MemoryLabel{-10em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{0}
+%  \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   \\
-  {\tt 02} &
+  {\tt 4} &
   \multicolumn{2}{c}{
@@
     \MemoryLabel{-10em}{0.75em}{5}
-    \MemoryLabel{-7em}{0.75em}{0}
+%    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
@@
 that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
-{\tt y} without even directly reading the stack pointer. In
+{\tt y} and hence {\tt x} without directly reading the stack pointer. In
 this case, {\tt g} branches on that value, which is a different way of
@@
 occurs because the program counter is also a component of the state that must change
-identically between the primary and variant traces. In a case where \(v_1 \not = 5\),
-the program counters will differ, breaking the lockstep.
+identically between the primary and variant traces. In a case where \(v_4 \not = 5\),
+the program counters will differ, breaking the lockstep.\apt{Is there any reason
+to complicate things like this by bringing in the PC?}
 

### 2021-04-15 tex/main.tex
Minor changes

--- 2021-04-14 tex/main.tex
@@
 So we assume we are given a machine language program {\em annotated} with the locations of
-instructions representing calls, returns, and function entry points\BCP{I
-  still want to talk about whether returns and entry points need to be
-  annotated.  To me, these annotations do not make intuitive sense: it is
-  the caller's responsibility to know whether the address that it is calling
-is really an entry point, and we can trust it to check this.  Similarly, it
-is not the callee's role to say when it is returning; it is the caller's
-prerogative to decide when it has been returned to!  I.e., from the caller's
-point of view, it has been returned to when control resumes at the
-instruction following the call with the SP reset to where it was at the
-point of the call. (This is the point where it wants to be able to access
-its own local state again and expected to find it unmolested.)}. Although these
+instructions representing calls.%, returns, and function entry points\BCP{I
+%  still want to talk about whether returns and entry points need to be
+%  annotated.  To me, these annotations do not make intuitive sense: it is
+%  the caller's responsibility to know whether the address that it is calling
+%is really an entry point, and we can trust it to check this.  Similarly, it
+%is not the callee's role to say when it is returning; it is the caller's
+%prerogative to decide when it has been returned to!  I.e., from the caller's
+%point of view, it has been returned to when control resumes at the
+%instruction following the call with the SP reset to where it was at the
+%point of the call. (This is the point where it wants to be able to access
+%its own local state again and expected to find it unmolested.)}.
+Although these
 annotations would typically come from a compiler together with the machine code,
 we do not assume anything about their provenance.
-
-\bcp{Can this be postponed to the formal sections?}We further parameterize over the structure of the stack, particularly the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, which maps a machine state to a set of addresses.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return.
 
@@
 of examples. Existing stack safety enforcement work mostly refers directly to examples
-of bad behavior that it prevents; we will insert our model between the enforcement and
-the examples\bcp{Didn't understand that}, arguing that:
+of bad behavior that it prevents; we will argue that:
 
@@
 the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
+
+We parameterize over the structure of the stack, particularly the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
+maps a machine state to a set of addresses.
+If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
+which addresses must be protected until the return. Similarly, we abstract over what
+it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
+which relates
 

### 2021-04-15 tex/main.tex
Reduced section 4, starting on 5

@@
 \newcommand*{\rsp}{\textsc{sp}}
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
 
@@
 So we assume we are given a machine language program {\em annotated} with the locations of
-instructions representing calls.%, returns, and function entry points\BCP{I
+instructions representing calls, and we determine when a return has occurred by tracking
+control flow and observing when we reach a {\em return target}. For a given state annotated
+as a call, in a standard calling convention, another state fulfills its return target if
+it has the same stack pointer and the program counter is at the instruction following
+the call, but we abstract over different conventions.
+%, returns, and function entry points\BCP{I
 %  still want to talk about whether returns and entry points need to be
@@
 %its own local state again and expected to find it unmolested.)}.
-Although these
-annotations would typically come from a compiler together with the machine code,
-we do not assume anything about their provenance.
+Although call annotations would typically come from a compiler together with
+the machine code, we do not assume anything about their provenance.
 
@@
   eliminated return (and entry point) annotations.}
-
-Since we rely heavily on code annotations, we must ask: what if the annotations
-are wrong? First, suppose a call or return occurs that is not annotated.
+\sna{I'm not sure we can escape questions about annotations, even if we have
+  just calls, but I tried to abbreviate it.}
+
+Since we rely heavily on call annotations, we must ask: what if the annotations
+are wrong? First, suppose a call occurs that is not annotated.
 In this case, the answer is simple: as far as the properties are concerned, there has
-been no call or return, only an instruction sequence resembling such an operation, and
-therefore no associated change in permissions. This is particularly interesting in the
-case of a false return, where control may appear to have returned to the caller, but
-the enforcement mechanism must still protect the caller's memory.
-
-Alternatively, an annotated instruction might not really be a call or return.
+been no call, only an instruction sequence resembling one, and
+therefore no associated change in permissions. Then the caller is guaranteed no
+protections.
+
+Alternatively, an annotated instruction might not obey the calling convention.
 In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
 and returns are implemented by fixed sequences of instructions which the tag policy
-will force to be executed from start to finish. The annotated call or return instruction should always
-be the final instruction of the appropriate sequence. What if an annotated instruction exists outside
-such a sequence? We will statically rule out some scenarios with a well-formedness criterion
-on initial states: annotated instructions will initially be tagged apropriately, as the
-final instruction in the call or return sequence, and all blessed\bcp{undefined} sequences will end with
-annotated instructions. This is all we need for the policy to
-protect itself: it can then failstop if it would execute the blessed sequence out of order.
-So, even in our strong attacker model, we need very little in the way of restrictions on
-the code when considering micropolicy-based enforcement.  \BCP{Is this
-  discussion still needed?  I had hoped that switching to our ``new''
-  definition on subtraces would eliminate such complexities!}
-
-\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
-a given enforcement mechanism may place a static well-formedness criterion
-on the code it applies to. Weaker enforcement mechanisms might also need dynamic
-reasonableness conditions to prevent the defender from sabotaging enforcement, as
-discussed in section \ref{sec:cheri}.  \BCP{Weak.}
+will force to be executed from start to finish, termed {\em blessed sequences}.
+When applying our properties to a micro-policy setting, we will statically rule out some
+scenarios with a well-formedness criterion on initial states: all annotated instructions will
+be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
+This is all we need for the policy to protect itself: it can then failstop if it would
+execute the blessed sequence out of order.
 
@@
 
-This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
-We believe that it is flexible enough to capture code-altering approaches, such as code
-rewriting or capability-based techniques, though perhaps with less elegance.
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
+%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
+%We believe that it is flexible enough to capture code-altering approaches, such as code
+%rewriting or capability-based techniques, though perhaps with less elegance.
 
@@
 
-%The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
-%at the time of the call. The stack can be used to pass arguments or return results
-%in the usual way, by placing them at the top of the caller's frame. This will be modeled
-%as a form of sharing in section \sna{cite the sharing section}.
-%Callers and callees can also communicate via registers if they wish; indeed, our
-%definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
-
-%\iftrue\bcp{This implicitly amounts to choosing a caller-saves register
-%  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
+Unless otherwise noted, colored boxes indicate an assignment of the address to a
+{\em domain} -- a label reflecting a particular set of access rights. Domains
+are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
+(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
+(outside the stack entirely.)
 

### 2021-04-15 tex/main.tex
Removed function id from annotation

--- 2021-04-15 tex/main.tex
@@
 \acmVolume{1}
-\acmNumber{SPLASH} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
@@
                         %% http://ctan.org/pkg/subcaption
+\usepackage{tikz}
+\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
+    \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
 
@@
 \newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}
-\newcommand*{\rsp}{\textsc{sp}}
 
@@
 proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
-broken variants).  Specifically, Roessler
+broken versions of these enforcement mechanisms).  Specifically, Roessler
 and Dehon's ``eager'' micro-policy, which catches violations as early as
-possible, enforces a basic ``stepwise'' versions of our confidentiality
+possible, enforces the basic ``stepwise'' versions of our confidentiality
 and integrity properties. Their more performant ``lazy'' micro-policy
-does not, but corresponds to weaker, more extensional ``observational'' variants
-of the
+does not, but inspires weaker, more extensional ``observational'' variants of the
 properties. While we do not rule out other enforcement mechanisms, our
 definitions are strongly informed by micropolicy systems, especially
-our attacker model, which is stronger than other enforcement mechanisms
-can enable.
-
-% For the sake of concreteness, most of our presentation assumes a standard RISC
-% machine architecture with a basic subroutine calling convention (parameters
-% passed in registers) and a micropolicy-like enforcement mechanism. We
-% also show how to accomodate stack-allocated parameter passing and
-% coroutines.
-
-% \bcp{this is also a
-%   good place to admit that our definitions are tuned for micropolicies, if
-%   we can find a succinct way to say it.}
-% \sna{See above.}
-% \BCP{I added it to the last paragraph (which is pretty close to the preceding
-%   paragraph :-)  We should follow this up with some discussion in the (new)
-%   Assumptions section about how different enforcement mechanisms (us,
-%   StkTokens, maybe others) adopt different perspectives on what the stack is
-% for and how it is organized and so lead to somewhat different notions of
-% low-level safety.}
+our attacker model.
+
+For the sake of concreteness, most of our presentation assumes a RISC-V
+machine architecture with a basic subroutine calling convention (passing
+parameters in registers) and a micropolicy-like enforcement mechanism. We
+also show how to accomodate stack-allocated parameter passing and
+coroutines.
+
+\bcp{this is also a
+  good place to admit that our definitions are tuned for micropolicies, if
+  we can find a succinct way to say it.}
+\sna{See above.}
+\bcp{I added it to the last paragraph (which is pretty close to the preceding
+  paragraph :-)  We should follow this up with some discussion in the (new)
+  Assumptions section about how different enforcement mechanisms (us,
+  StkTokens, maybe others) adopt different perspectives on what the stack is
+for and how it is organized and so lead to somewhat different notions of
+low-level safety.}
 
@@
 
-% \apt{This is a good story, but does it describe the paper we are able to produce this week?
-%   We were rejected before exactly because we claimed to have
-%   a general criterion, but did not in fact demonstrate its generality (which is not
-%   obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
-%   para about the general story.  Does the rest now avoid making (or
-%   implying) claims that we can't defend?}
-% \bcp{Also, I think we should save
-%   WBCF for the intro---I can't find a way to discuss it here that is brief
-%   and not confusing/distracting.}
-
-  % \iftrue\bcp{should we mention stktokens here?}
+\apt{This is a good story, but does it describe the paper we are able to produce this week?
+  We were rejected before exactly because we claimed to have
+  a general criterion, but did not in fact demonstrate its generality (which is not
+  obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
+  para about the general story.  Does the rest now avoid making (or
+  implying) claims that we can't defend?}
+\bcp{Also, I think we should save
+  WBCF for the intro---I can't find a way to discuss it here that is brief
+  and not confusing/distracting.}
+
+  % \ifaftersubmission\bcp{should we mention stktokens here?}
   % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
@@
   {
-    \MemoryLabel{#2}{2em}{$\downarrow$\SP #1}
+    \MemoryLabel{#2}{2em}{\SP #1}
   }
@@
 \citet{Skorstengaard+19} formalize these properties by defining an idealized
-machine in which they hold {\em by construction};
-they then show that an enforcement
-mechanism (in their case, hardware capabilities) guarantees stack safety by
-proving that an ordinary machine with enforcement can simulate the idealized machine.
-%an enforcement mechanism (in their case, hardware
-%capabilities)
-%can then be shown to guarantee stack safety by exhibiting a fully abstract
-%embedding function
-%from the idealized machine to an ordinary machine equipped with this
-%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
-%  understand that sentence.}
+machine in which they hold {\em by construction}; an enforcement mechanism (in their case, hardware
+capabilities)
+can then be shown to guarantee stack safety by exhibiting a fully abstract
+embedding function
+from the idealized machine to an ordinary machine equipped with this
+enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
+  understand that sentence.}
 %%
@@
 from its callee, even as that callee may need protection from its callee in turn.
-This is novel in that the
-confidentiality property applies to many nested subtraces within a single execution,
-as opposed to traces of entire executions.
+This differs from common practice in language-based security in that the
+confidentiality property applies to many nested subtraces within a single execution.
+\sna{How novel is this nesting?}
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
@@
 violating the integrity and confidentiality of the caller. Second, confidentiality
-contains temporal protections that LSE does not necessarily address. The intuition
-for LSE is that it protects the caller's stack frame, not necessarily memory that was
-previously allocated but now outside of the caller's frame. (Cheri-capability-based
-implementations of LSE {\em do} protect previously allocated data, typically by clearing
-it, but because it is a potential vector for leaking capabilities into the caller's frame,
-not as a goal of the property.)
-
-To illustrate the applicability of these properties, we perform randomized testing to
-validate that an existing enforcement mechanism does enforce them.
-\citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
-using tag-based \emph{micro-policies}~\citep{pump_oakland2015}, but offer no formal properties.
-We test against their {\em Depth Isolation} micro-policy, in which memory cells within each stack
-frame are tagged with the identity of the function activation
-that owns the frame, and access to those locations is then gated
-according to the activation that is currently executing. This is simple and corresponds
-to intuitive notions of stack safety, and we argue that by describing its behavior
-we approach a reasonable stack-safety property that other micro-policies can be
-validated against.
-
-Randomized testing is a relatively low-cost way to validate a potential enforcement
-mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
-of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
-covers potentially erroneous programs, we test intentionally buggy policies alongside
-the real one. If we can catch known bugs, we gain confidence that we can catch unknown ones,
-strengthening the validation of the enforcement mechanism.
+contains temporal protections that LSE does not implicitly address.
+\apt{Maybe want to say a little more about this here, since it is a novelty wrt/ existing work?}
+\bcp{+1}\sna{TODO: Say a little more, but make sure it's accurate first.}
 
@@
 
-A second challenge is that enforcement mechanisms may weaken their protections
-in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
-\citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
-ways of identifying functions and deciding exactly which accesses are allowed.
+A second challenge is that
+enforcement mechanisms may weaken their protections that in the name of efficiency,
+while arguing that they still capture stack safety in a meaningful sense.
+For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
+using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
+Memory cells within each stack frame are tagged with the
+identity of the function activation
+that owns the frame, and access to those locations is then gated
+according to the activation that is currently executing. Roessler and DeHon
+explore alternative versions of this basic idea, with different ways of identifying
+functions and deciding exactly which accesses are allowed.
 These vary in their efficiency and in the spatial and temporal precision of their
 security guarantees.
-The Depth Isolation policy prevents a callee from writing outside its frame
+The \emph{Depth Isolation} policy prevents a callee from writing outside its frame
 at all, which is simple and intuitively correct, but it must manually clear the tags
-on the frame on each return, which is very expensive. Conversely, their
+on the frame on each return, which is very expensive. Conversely, the
 \emph{Lazy Tagging and Clearing} policy improves
@@
 permits  ``illicit'' writes to inaccessible stack frames so long as the
-written locations are not subsequently read by their ``real'' owners. This violates our
-straightforward properties, but ntuitively,
+written locations are not subsequently read by their ``real'' owners. Intuitively,
 this form of laziness is justified if any dangerous writes will
@@
 
-What does ``affecting program behavior'' really mean? This micro-policy
+But what does ``affecting program behavior'' really mean? This micro-policy
 calls for a second characterization of stack safety in terms of the externally
@@
 can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
-integrity implies observational integrity. \sna{TODO: check this.}
-
+integrity implies observational integrity.
+\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
+  established that this was important.\apt{medium importance. it is obvious that
+    observational integerity allows behaviors that stepwise does not, but
+    maybe not obvious that the converse is not true too.}}
 We additionally sketch an associated {\em observational confidentiality} property
-makes confidentiality extensional in the same way, in order to form an
-{\em observational stack safety} that is highly extensional.
+makes confidentiality extensional in the same way, although we do not know of a real
+enforcement mechanism that weakens confidentiality in this way.
+\apt{why bother mentioning it then?}
+\sna{If a relatively extensional characterization is a worthwhile contribution
+  in its own right, then it's worth mentioning that we have one. If not, then indeed,
+  we shouldn't mention it -- or get too excited about extensionality in general, because
+  lockstep isn't extensional.}
+Together, these form an {\em observational stack safety} that is highly extensional.
 We therefore argue that, despite its somewhat greater complexity,
@@
 
+\bcp{This paragraph doesn't apply to the lazy properties, so maybe it should
+be moved before they are introduced?}%
+To illustrate the applicability of these properties, we perform randomized testing to
+validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
+them. Randomized testing is a relatively low-cost way to validate a potential enforcement
+mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
+of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
+covers potentially erroneous programs, we test intentionally buggy policies alongside
+the real one. If the testing can differentiate correct implementations from buggy ones,
+it is an argument in favor of their correctness. \bcp{Hmm--don't understand
+  that sentence}
 %\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
@@
 %  ones...''}
-As it turns out, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
-Clearing micro-policy enforces observational integrity, but not confidentiality,
-whether stepwise or observational.
+Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
+Clearing micro-policy does not enforce stepwise integrity.
+Unfortunately the existing Lazy Tagging and Clearing
+micro-policy does not quite enforce observable confidentiality either. %, as we show in
+%testing~\apt{really??},
 But formalizing the ideal property allows us to see how it falls short -- a specific
-violation of the temporal aspect of confidentiality occurs in corner cases where callees
-can leak data between one another.
+violation of the temporal aspect of confidentiality.
 We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
-confidentiality, albeit at some performance cost.
+observational integrity, at some performance cost.\apt{but we're not testing that, right?}
+\rb{Not sure about the property part right now, the machine side of this is
+  pretty much done}
+\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
+  need a caveat about how it's harder to test?}\apt{As per email discussion,
+  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and
+  Clearing, by loosening the temporal safety properties?}
+\sna{I'm pretty sure for the main policy it's just lazy integrity, no confidentiality.
+  There was a variant Nick and Andre gave that would restrict the leaks to only be between
+  instances of the same function, so that's the same property under a weaker attacker model,
+  where we don't protect a function from itself.}
 
@@
 callee access to the caller's data. We do not model more sophisticated sharing,
-such as taking the address of a variable and treating it like a pointer.
+such as taking the address of a variable and treating it like a pointer. Doing so without
+protecting the variable is not interesting, and protecting would be better suited to
+a different kind of enforcement, such as a memory safety micro-policy or Cheri-style capability.
+\bcp{This is all a bit mysterious} \sna{Any less mysterious now?}
+
+\rb{So no pointer passing? Pass by reference is now discussed at the end of
+  Section 3, however it looks like things get tricky rather quickly.}
+\sna{The trick is that if a programmer is passing pointers, we can't presume their
+  intent -- how long should we permit memory to be shared? But if the compiler is passing
+  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
+  \apt{N.B.Doing pass-by-reference properly requires that the passed variable is accessible
+    by all descendents, not just immediate callee, which is doable with depth tags but not
+    obviously with per-activation tags }}
 
@@
 
-\section{Setup}
-\label{sec:setup}
+\section{Assumptions and Threat Model}
+\label{sec:threat}
+
+\bcp{I would still advocate splitting this into two sections, just so the
+  transition from one to the other and the division of responsibility are
+  clear.  The first section could be called ``What is 'the Stack'?'' instead
+of ``Assumptions'' if we want.}
+
+\newcommand*{\rsp}{\textsc{sp}}
 
@@
 assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter (\(\PCname\)); other
+machine with a program counter \(\PCname\); other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
 the layout of memory, including where code lives and which addresses are within
-the bounds of the stack. We do not model a heap; programs might choose to use memory
-that is neither code nor stack to store global variables, as a heap, etc., but our
-properties do not protect it or assume any particular behavior. We also do not
-model run-time code generation.
-
-Our stack-safety criteria depend
+the bounds of the stack.
+
+Our stack-safety concepts depend
 critically on the notions of ``call'' and ``return,'' which are not fully explicit
@@
 %
-So we assume we are given a machine language program {\em annotated} with the locations of
-instructions representing calls.%, returns, and function entry points\BCP{I
-%  still want to talk about whether returns and entry points need to be
-%  annotated.  To me, these annotations do not make intuitive sense: it is
-%  the caller's responsibility to know whether the address that it is calling
-%is really an entry point, and we can trust it to check this.  Similarly, it
-%is not the callee's role to say when it is returning; it is the caller's
-%prerogative to decide when it has been returned to!  I.e., from the caller's
-%point of view, it has been returned to when control resumes at the
-%instruction following the call with the SP reset to where it was at the
-%point of the call. (This is the point where it wants to be able to access
-%its own local state again and expected to find it unmolested.)}.
-Although these
+So we assume we are given a machine language program annotated with the locations of
+instructions representing calls, returns, and function entry points. Although these
 annotations would typically come from a compiler together with the machine code,
 we do not assume anything about their provenance.
+
+We further parameterize over the structure of the stack, particularly the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention}, which maps a machine state to a set of addresses.
+If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
+which addresses must be protected until the return.
 
@@
 
-Dividing policy state from machine state helps us reason about
+\sna{This is awkward.}
+This division of policy state from machine state is helpful when reasoning about
 enforcement mechanisms that keep a significant amount of data separate from the
@@
 
-The separation is smaller in protection mechanisms that interact more directly
-with application code. Some mechanisms might exhibit failstop behavior, but have
-no protected policy state; others might never even failstop. Then the policy-enhanced
-machine is simply identical to the original. This is the case for software-based
-techniques.
-
-\section{Threat Model}
-\label{sec:threat}
-
-Intuitively, stack safety means that the stack frame of each caller (and its
-those of its ancestors) is protected
-from attempts by a confused or malicious callee (including all its descendants)
-to read or write its private data.
+The separation is less useful in protection mechanisms that interact more directly
+with application code, such as Cheri-style capabilities\citep{??}. Here
+the policy would not have auxiliary state, but would still refine the machine step
+function to create a distinct failstop condition. Other techniques such as
+software fault isolation~\citep{??} might never actually failstop, but fulfill
+the properties by rewriting code so that traces can never violate them.
+
+\paragraph*{Threat Model}
+
+Intuitively, stack safety means that a caller's stack frame is protected
+from a confused or malicious callee's attempts to read or write its private data.
 The attacker model therefore treats every caller as a defender and its callee as
-an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways.
+an attacker, meaning that with nested calls, a single function may play both roles
+at varying points.
 
 We take a very strong threat model in which both caller and callee may execute
-arbitrary machine code. Hardware and timing attacks are out of scope.
+arbitrary machine code. Hardware and timing attacks are out of scope, however.
 
 \paragraph*{Well-formedness and Bad Annotations}
-
-\BCP{I believe we could delete this section---and never mention WBCF at all
-  (except to say that we leave it out, compared to STkTokens)---if we
-  eliminated return (and entry point) annotations.}
 
@@
 on initial states: annotated instructions will initially be tagged apropriately, as the
-final instruction in the call or return sequence, and all blessed\bcp{undefined} sequences will end with
+final instruction in the call or return sequence, and all blessed sequences will end with
 annotated instructions. This is all we need for the policy to
@@
 So, even in our strong attacker model, we need very little in the way of restrictions on
-the code when considering micropolicy-based enforcement.  \BCP{Is this
-  discussion still needed?  I had hoped that switching to our ``new''
-  definition on subtraces would eliminate such complexities!}
-
-\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
+the code when considering micropolicy-based enforcement.
+
+In general, regardless of enforcement mechanism, we take this approach:
 a given enforcement mechanism may place a static well-formedness criterion
@@
 
-\section{Key Ideas, By Example}
+\section{Demonstrating Stack Safety}
 \label{sec:running-example}
@@
 
-\begin{subfigure}{.2\textwidth}
-{\small
+\begin{subfigure}{.4\textwidth}
 \begin{verbatim}
 int main() {
-  int x = 42;
-  return f()+x;
+0:     int x, y;
+1:     y = 42;
+2[C]:  x = f();
+3[R]:  return x+y;
 }
@@
 int f() {
-  int y;
-  *(&y-3) = 0;
-  return 5;
+10:     *($SP-2) = 0;
+11[R]:  return 5;
 }
 \end{verbatim}
-}
 \end{subfigure}
-\begin{subfigure}{.5\textwidth}
-{\small
-\begin{verbatim}
-1 main: add $1,%sp      ; allocate frame
-2       mov $42,-1(%sp) ; initialize local
-3       call f
-4       add -1(%sp),%ra ; set return value
-5       sub $1,%sp      ; deallocate frame
-6       ret
-7 f:    add $1,%sp      ; allocate frame
-8       mov $0,-3(%sp)  ; violation!
-9       mov $5, %ra     ; set return value
-10      sub $1,%sp      ; deallocate frame
-11      ret
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}{.25\textwidth}
+\begin{subfigure}{.4\textwidth}
 \begin{center}
 \begin{tabular}{l l}
-{\tt 2} &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}[Unseal]
+{\tt 1} &
+\memoryaddrs{12em}
+\memory{4}{\unsealc}[Unseal]
 ~$\cdots$
@@
 \\
-{\tt 3} &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}
+{\tt 2} &
+\memoryaddrs{12em}
+\memory{4}{\unsealc}
 ~$\cdots$
@@
 \\
-{\tt 8} &
+{\tt 10} &
 \memoryaddrs{16em}
-\memory{1}{\mainsealc}[Seal(0)]%
+\memory{2}{\mainsealc}[Seal(0)]%
 \memory{1}{\retptrc}[RetPtr]%
@@
 \\
-{\tt 9} &
+{\tt 11} &
 \memoryaddrs{16em}
-\memory{1}{\mainsealc}
+\memory{2}{\mainsealc}
 \memory{1}{\retptrc}
@@
 of examples. Existing stack safety enforcement work mostly refers directly to examples
-of bad behavior that it prevents; we will argue that:
+of bad behavior that it prevents; we will insert our model between the enforcement and
+the examples, arguing that:
 
@@
 
-We show our examples in C syntax and give the corresponding assembly code for a
-simple two-address machine, with {\tt call} and {\tt ret} instructions that
-push and pop the return address from the stack, implicitly adjusting the stack pointer.
-The assembly code reflects a simple compilation model in which  
-local variables are stored in the stack frame.
-Unless otherwise stated, arguments and return values are passed in registers.
+Our examples are written in C-like syntax, labeled with line numbers for reference,
+but to reflect the increased power of the underlying assembly code, we allow code
+to read the stack pointer ({\tt \$SP}) directly and use
+it to form pointers.
+Call and return annotations are marked {\tt[C]} and {\tt[R]}
+respectively. Local variables are assumed to be stored in the stack frame,
+in declaration order. Unless otherwise stated, arguments and return values are passed in registers.
 The machine is word-addressed. The stack grows upward in
@@
 At the right of each example, we show the layout of stack memory just before
-the specified instruction is executed; each box represents a memory word and
+the specified line number is executed; each box represents a memory word and
 addresses increase to the right.
@@
 meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of
-their owner\bcp{are we still using depths?},
+{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
 in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt x} will be unchanged when it gets control back.
+{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
 This gives us our simple, intuitive statement of {\em stepwise integrity}:
@@
 %cares about changes it can see after the return, not changes during execution.
-In this case, after instruction 8, {\tt f} has violated integrity by writing to {\tt x},
-so the final return value produced by {\tt main} is wrong.
-
-The black rectangle represents the return address, which is saved by the call
-instruction itself, so is not sealed.
+In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.
+
+The black rectangle represents the return address. In a typical RISC calling convention
+this will have been saved into a register by the call itself, and {\tt f} stores
+it in the stack during its preamble, so it is not itself sealed.
 
@@
   {\tt 10} &
-  \memoryaddrs{21em}
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
   &
-  \memoryaddrs{21em}
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
   {\tt 11} &
-  \memoryaddrs{21em}
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
   {\tt 12} &
-  \memoryaddrs{21em}
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
 Meanwhile confidentiality intuitively states that a callee is unaware
-of the context of its call (except for its arguments and any global
-variables\bcp{We haven't talked about global variables.  We should.}).
-We state this in a style similar\bcp{??} to {\em noninterference}~\citep{??}.
+of the context of its call (except for its arguments and any global variables).
+We state this in a style similar to {\em noninterference}~\citep{??}.
 Noninterference describes knowledge: it states that for some system components
@@
 
-\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
+We alternatively could imagine a version of stepwise confidentiality in which we explicitly
 track which addresses may be read (i.e., which have been initialized), and at each step
@@
 One of the most pernicious ways to attack the stack is by hijacking control flow.
-Our stack safety properties do not assume that control flow is
-well-bracketed\bcp{undefined term}; rather,
+Our stack safety properties do not assume that control flow is well-bracketed; rather,
 they express what it means to protect data even in the presence of control-flow attacks.
-Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''\bcp{??}~\citep{??}.
+Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}.
 (Here we show return addresses explicitly as line numbers, in the form $\#n$.)
@@
 execute a return-labeled instruction: it must actually reach the appropriate
-{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep
-a stack of return targets\bcp{do we still},
+{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep a stack of return targets,
 which correspond to the depths at which stack frames have been sealed. A callee is
 only considered to have returned when it executes a return-annotated instruction
-and then reaches a valid return target\bcp{what's that?}. Only then do the sealed frames become unsealed.
+and then reaches a valid return target. Only then do the sealed frames become unsealed.
 
@@
 \paragraph*{Pass by Reference and Stack Arguments}
-\bcp{I wonder if this could be postponed so that we get to the formal
-  definitions for the basic property sooner...}
 
@@
 passed, and our diagram displays it in magenta. The caller's frame locations that are not
-passed are sealed as usual. Just as we marked sealed memory with the depth of its owner\bcp{??},
+passed are sealed as usual. Just as we marked sealed memory with the depth of its owner,
 we do the same for passed memory -- but it has multiple owners. We keep track of the
@@
 
-We next define our machine model precisely. To make our definition of stack safety
+We next describe our machine model. To make our definition of stack safety
 as generic as possible, this section proposes an abstract interface to a
 machine model plus some kind of (software- or hardware-enforced) policy monitor.
-In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
+In~\cref{sec:enforcement}, we will instantiate this model with a concrete dynamic
 policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
@@
 %
-Our machine states are composed of {\em components} \(\component\),
-addressed\bcp{no, the components are not addressed, they are addresses!} either by
+Our machine states are composed of {\em components} \(\component\), addressed either by
 words (memory addresses) or by register names (\(\reg\)) drawn from some set
@@
 
-Observations are only essential for describing the extensional properties
-described in Section~\ref{sec:lazy}, but for simplicity, we include them
-in the machine model from the beginning.
 We choose to model observations explicitly, rather than simply treating
@@
 
-\apt{This para is out of date. Do we want to say something equivalent
-  about new examples (hard, because they are C-based).}
+In the running example of \Cref{sec:running-example},
 $\WORDS$ is the
@@
 We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
-A {\em policy} consists of a set of policy
+A policy is a general model of an enforcement mechanism, consisting of a set of policy
 states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
@@
 can see the machine state, but cannot modify it. We assume that the
-compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.
+compiler provides an initial policy state \(\pol_0\) to accompany the initial machine state.
 %
@@
 
-\noindent
 This combined step function is also partial, and traces of the machine-with-policy
@@
 information representing the history of the run. The following definitions are parameterized
-over an arbitrary\bcp{why?} set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
 an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context, \anns \constep \context' \in \MACHS \times \CONTEXTS
-\times \powerset(\ANNS) \rightarrow \CONTEXTS\). We also assume a set of function identifiers
-\(\FUNIDS\) and the existence of a \emph{code map} \(\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\)
-that relates each code address to its owning function and set of annotations.
+\(\mach, \context, \ann \in \powerset \ANNS \constep \context' \in \MACHS \times \CONTEXTS
+\times \ANNS \rightarrow \CONTEXTS\). We also assume for any set of annotations that we have
+code maps, with \(\FUNIDS\) a set of function identifiers:
+\[\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]
+
 We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).  \bcp{I don't get the intuition for this whole para.}
+a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).
 
 \judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
-              {\(\mach, \context, \anns \constep \context'\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \ann\)}
+              {\(\mach, \context, \ann \constep \context'\)}
               {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
@@
 
-A trace over a type \(\TRACEELEMS\)\bcp{``over a type A'' is a Coq-ism} is a nonempty, finite or infinite sequence of
+A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
 \(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
 \(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
-called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
+called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\), and of observations.
 
@@
 the policy component. The {\sc Default} rule applies only if the other does not.
-(We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
+(We use juxtaposition of an element and a trace to represent ``cons.'')
 
@@
 
-\bcp{Could any of this be postponed?}
-
-The ``observations-of'' operator, written \(\pi_\obs\), \apt{since this isn't
-  actually a projection, it is slightly confusing to use \(\pi\)} takes a machine trace and
+The ``observations-of'' operator, written \(\pi_\obs\), takes a machine trace and
 creates a trace of observations corresponding to the observations of each step in
@@
 to relate a machine-policy-context state triple to the MPCO-trace
-induced by the step function.\apt{?? This looks indistinguishable
-  from the \(\hookrightarrow\) we've already defined.}
+induced by the step function.
 
@@
 \hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:
+running \(\mpcstatename\) up to \(f\)'') and similarly for traces with policies:
 
@@
 
-In particular, we wish to formulate properties about consecutive subtraces
+In particular, we wish to formulate properties around consecutive subtraces
 in which the context obeys a given predicate. This will commonly be used to extract
 individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
+at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
+we have \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
 ``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''
@@
 then dive into the formal definitions of local state encapsulation and
-well-bracketed control flow, which together constitute stack safety\bcp{not
-  any more?}.
+well-bracketed control flow, which together constitute stack safety.
 Initially we will focus on a simple stack model, with a single stack
@@
 \[\ann \in \ANNS ::= \callmap + \retmap\]
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \ANNS\]
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \ANNS\]
 
@@
 of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth\bcp{What about code??}. Each caller will register a {\em target} condition -- a
+at a particular depth. Each caller will register a {\em target} condition -- a
 predicate on machine states that must hold for it to consider itself active.
@@
 ``return target list,'' which is a stack of targets, one for each caller awaiting
-return.  \ifaftersubmission\bcp{spacing is funny}\fi
+return.
 
@@
 the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
-
-We parameterize over the structure of the stack, particularly the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
-maps a machine state to a set of addresses.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
-which relates
+Then at each step we update the context according to the context-step function,
+\(\stepsto_C\). We will define this in terms of the operations
+\(\push{\domMap}{\depth}{\components}\), in which \(\depth \in \mathbb{Z}\) and
+\(\components \subseteq \COMPONENTS\), and \(\pop{\domMap}{\depth}\), pronounced
+``push'' and ``pop'' and defined as follows.
+
+\[
+(\push{\domMap}{\depth}{\components}) ~ \component =
+\begin{cases}
+  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
+                    \text{ and } \component \in \components \\
+  \domMap ~ \component & \text{else} \\
+\end{cases}
+\]
+\[
+(\pop{\domMap}{\depth}) ~ \component =
+\begin{cases}
+  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth'} \text{ where }
+              \depth' >= \depth\\
+  \domMap ~ \component & \text{else} \\
+\end{cases}
+\]
+%
+Push seals a set of components with its depth argument, and pop unseals all components
+that were as deep as its argument or deeper.
 
@@
 and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-Then at each step we update the context according to the context-step function,
-\(\stepsto_C\).
-
-The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
-A call simply pushes the return target onto the return stack and seals all components
-with the depth of the previous return stack. A return
-recursively searches for a valid return target at the next step, and unseals
-every component sealed at a depth that it passes.
-
-\judgmentthree[Call]
+
+The call rule pushes the components sealed by the sealing convention with the depth
+of the current caller, and determines the return target by partially applying the
+return convention to the machine state, and pushes that onto the targets stack.
+The return rule checks whether the attempted return actually manages to reach a
+return target, traversing down the stack until the first one it finds. Recall that
+we are not discussing well-bracketed control flow, so the return might skip over
+a waiting caller. When a target is reached, {\em all} sealed components at or above
+its depth are unsealed by the pop operation. If no target is reached, nothing is
+popped at all. These rules are given inductively, and apply in the listed order.
+
+\judgmenttwo[Call]
             {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
-              {\(\components = \sealcon ~ \mach\)}
-              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\judgmenttwobr[ReturnFound]
+            {\(\depth = |\rts|\)}
+            {\(\mach,(\domMap,\rts) \stepsto_C
+              \push{\domMap}{\depth}{\sealcon ~ \mach},(\retcon ~ \mach)::\rts\)}
+
+\judgmentthree[ReturnFound]
               {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
               {\(\mach \stepsto \mach' \in \target\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\judgmenttwobr[ReturnRec]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+              {\(\depth = |\rts|\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C
+                \pop{\domMap}{\depth},\rts\)}
+
+\judgmentthree[ReturnRec]
+              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \_,\rts'\)}
+              {\(\depth = |\rts'|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C
+                \pop{\domMap}{\depth},\rts'\)}
 
@@
 
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change together
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are {\em synched}
 if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
@@
 \definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
-change together, written as a coinductive relation \(\machT \doteq
-\nachT\):\bcp{People may not be familiar with the notation}
+are synched, written as a coinductive relation \(\machT \doteq \nachT\):
 
@@
 
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold for
+\(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth)\) and
+\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
 \(\MPCT\) corresponds to the trace of a subroutine from its call to its return.
@@
 
-\subsection{Stack Safety with Passing}
-
-Next we introduce variable passing. We will need to add extra information to our call
-annotations to reflect that they may now pass variables on the stack, and that the locations
-of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the components
-that should be passed -- when passing by value, this will typically give the memory cells
-at the appropriate offset from the stack pointer, while passing by reference gives the
-addresses of the appropriate variables wherever they happen to be.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap\]
-
-And we extend the set of domains to include passed components,
-which are labeled with the range of depths of the caller that has passed them.
-
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
+\subsection{Stack Safety with Sharing}
+
+Next we introduce sharing. We will need to extend the set of annotations to include
+sharing and passing, and now the code map may map a single instruction to multiple
+annotations, because in some calling conventions a single instruction might carry
+meaning as both a call and a pass. But we will never have a single instruction annotated
+as both a call and a return.
+
+\[\ann \in \ANNS ::= \callmap + \retmap + \sharemap(P \subseteq \MACHS \times \COMPONENTS) +
+\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
+
+And we extend the set of domains to include shared and passed components, the latter of
+which are labeled with the depth of the caller that has passed them.
+
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \shared{\depth} + \sealed{\depth} + \passed{\depth}\]
 \[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\rts \in \listT (\target \subseteq \MACHS)) \]
-
-We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map when pushing it, where \(\depth\) is an integer depth.
-
-\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
+\times (\depth \in \mathbb{N}) \]
+
+What it means to push the stack is similar to above, but two details are worth noting.
+First, shared and passed components are not sealed by the sealing convention, and second,
+anything not touched by the sealing convention must be unsealed. Sharing and passing cannot
+apply to memory the callee could otherwise use.
+
+\[
+(\push{\domMap}{\depth}{\components}) ~ \component =
 \begin{cases}
-  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
-  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
-  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
-                                 \text{ for any } \depth' \\
-  \domMap~\component & \text{else} \\
-\end{cases}\]
-
-And to pop we just unseal the components at depth \(\depth\):
-
-\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
+  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
+                    \text{ and } \component \in \components \\
+  \domMap ~ \component & \text{else if } \component \in \components \\
+  \unsealed & \text{else} \\
+\end{cases}
+\]
+
+Popping is simpler, but note that while passed components are unsealed on return to their
+passer, shared ones are unsealed on return {\em from} their sharer - on deallocation,
+in other words.
+
+\[
+(\pop{\domMap}{\depth}) ~ \component =
 \begin{cases}
-  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth,\depth} \\
-  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
-                              \text{ for any } \depth' < \depth \\
-  \domMap~\component        & \text{else} \\
-\end{cases}\]
-
-Now we can inductively define the context update function; these again apply in the order
-presented.
-
-\judgmenttwo[Call]
-            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
-            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\judgmentthree[ReturnFound]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
-              {\(\mach \stepsto \mach' \in \target\)}
-              {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\judgmentthree[ReturnRec]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\judgment[Default]
-         {}
-         {\(\_,\context \stepsto_C \context\)}
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \sealed{\_}\) or \(\passed{\depth,\depth'}\) where \(|\rts| < \depth\)
-or \(|\rts| > \depth'\). A system enjoys {\em stepwise stack integrity with passing} if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \unsealed\) or \(\passed{\depth,\depth'}\) where
-\(\depth \leq |\rts| \leq \depth'\) .
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
-\(\MPCT \doteq \NPCT\).
+  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1}
+              \text{ or } \domMap ~ \component = \passed{\depth-1}
+              \text{ or } \domMap ~ \component = \shared{\depth} \\
+  \domMap ~ \component & \text{else} \\
+\end{cases}
+\]
+
+Finally we introduce a more direct update. If \(P\) is a relation between machine states and
+components, we write \(\domMap[P ~ \mach \mapsto \stackDom]\) to say \(\domMap\) updated
+so that each component \(\component\) where \(P ~ \mach ~ \component\) holds is mapped to
+\(\stackDom\).
+
+Now we can inductively define the context update function; as these could be ambiguous assume that
+they apply in the order presented.
+
+\begin{center}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\callmap\} \cup \ann) \stepstocon
+             (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\retmap\} \cup \ann) \stepstocon
+             \mach',(\pop{\domMap}{\depth},\depth-1)\)}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\passmap ~ P\} \cup \ann) \stepstocon
+             \domMap[P ~ \mach \mapsto \passed{\depth}], \depth)\)}
+  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
+           {\(\mach,\context,(\{\sharemap ~ P\} \cup \ann) \stepstocon
+             \domMap[P ~ \mach \mapsto \shared{\depth}], \depth)\)}
+  \judgment{}
+           {\(\mach,\context,\_ \stepstocon \context\)}
+\end{center}
+
+\definition Let \(\MPCT\) be a reachable trace segment of the form
+\(\mpcstatename ~ \mpcstatename'\), where \(\pi_\context(\mpcstatename) = (\domMap,\depth)\).
+Let \(\components\) be the set of components \(\component\) such that either
+\(\domMap ~ \component = \sealed{\_}\) or \(\domMap ~ \component = \passed{\depth'}\) for
+some depth \(\depth' < \depth-1\). If any such \(\MPCT\) obeys trace integrity with respect
+to \(\components\), then the system enjoys {\em ultra-eager confidentiality} in the sharing
+model.
+
+\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
+on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
+\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
+\(\component\) such that either \(\domMap ~ \component = \sealed{\_}\) or
+\(\domMap ~ \component = \passed{\depth''}\) for some depth \(\depth'' < \depth'-1\).
+Then the system enjoys {\em eager integrity} in the sharing model if \(\MPCT\) obeys
+trace integrity with respect to \(\components\).
+
+\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
+on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
+\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
+\(\component\) such that either \(\domMap ~ \component = \outside\) or
+\(\domMap ~ \component = \shared{\_}\). Then the system enjoys {\em eager confidentiality}
+in the sharing model if \(\MPCT\) obeys trace confidentiality with respect to \(\components\)
+and \(f\).
 
@@
 Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
-pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We extend the set of possible annotations with a yield annotation.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap + \yieldmap\]
-
-The first major difference in this model is the domains. Stack domains are now nested inside
-top-level domains, and associated with stack identifiers.
-Our context now consists of a domain map, a map from stack identifiers to return target stacks,
-an additional map of targets for yields, and a stack identifier of the active stack.
-Our yield target map \(\yts\) is a map from stack identifiers to targets.
-In a standard system an initial yield target map will map each stack identifier to the
-set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
-\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
-
-\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
-\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
+\(\stackof\) from addresses to stack ids. The ``active stack'' -- the stack belonging to the
+currently executing coroutine -- in any given machine state is \(\stackof(\mach ~ \SP)\).
+We extend the set of possible annotations with a yield.
+
+\[\ann \in \ANNS ::= \callmap + \retmap + \yieldmap +
+\sharemap(P \subseteq \MACHS \times \COMPONENTS) +
+\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
+
+We use the same domain model as above, but instead of keeping track of a single depth
+we use a map from stacks to their depths.
+
 \[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
-(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
-\target) \times (\stackid \in \STACKS)\]
-
-In our context update rules, calls and returns work the same as in the sharing model,
-with domain updates restricted to the top-level domain corresponding to the active stack,
-and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
-
-\judgmentthreebr[Yield]
-                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
-                {\(\mach \stepsto \mach'\)}
-                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
-                {\(\stackid \not = \stackid'\)}
-                {\(\yts ~ \stackid'\)}
-                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
-                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
-
-This calls for some explanation. First, we determine which coroutine we are yielding to by
-checking which stack the stack pointer of the next state is pointing to. Then we can only yield
-if the next state is a valid yield target for that stack, and we update the yield target for
-the yielding stack. So it is once again possible for a coroutine to claim to yield but not
-actually change the active stack, with similar effects to when a callee fails to return properly.
-Now we can define both stack safety and coroutine safety under this model.
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
-{\em stepwise stack integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
-A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\subsection{Well-bracketed Control Flow}
+(\depthMap \in \STACKS \rightarrow \mathbb{N})\]
+
+\subsection{Local State Encapsulation}
+
+\definition A system enjoys {\em local state encapsulation} if it enjoys eager integrity and
+eager confidentiality.
+
+\definition A system enjoys {\em lazy local state encapsulation} if it enjoys lazy integrity
+and eager confidentiality.
+
+We do not give a confidentiality equivalent to lazy integrity, because the family of lazy
+enforcement mechanisms that we consider under the HOPE system in fact enforce eager confidentiality.
+A policy based on taint-tracking might require such a lazy confidentiality, and it would be constructed
+by replacing the ``same-difference'' condition with one that requires any differences to produce
+observationally equivalent states. But for our purposes it is irrelevant.
+
+\section{Well-bracketed Control Flow}
 \label{sec:wbcf}
 
-Both the integrity and confidentiality components of stack safety
-are reasonable even in programs that violate expected
-control flow. But even though stack safety prevents control-flow violations
-from being used to gain privileges, it is worth protecting control flow as well.
-Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
-control flow.
-
-When we focus on subroutines, the essence of well-bracketed control flow
-is that a caller can expect that if its callee ever returns, it returns
-to the caller at the appropriate return target. This is tantamount to saying
-that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
-the simple model:
-
-\definition A system enjoys {\em well-bracketed control flow} if for all
-reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-\(\mpcstatename \stepstocon \mpcstatename'\), if
-\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
-is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
-
-Extending this idea to coroutines, we use the same notion that there are
-no ``false'' yields.
-
-\definition A system enjoys {\em well-structured control flow} if for all
-reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-\(\mpcstatename \stepstocon \mpcstatename'\), if
-\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
-is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
-where \(\stackid \not = \stackid'\).
+Both the integrity and confidentiality components of local state
+encapsulation concern themselves only with the accessibility of data,
+and may still hold on a wide range of programs that violate expected
+control flow. The archetypal example of this is the {\em awkward example},
+in which an adversarial callee stores a return address and returns to it
+on a subsequent call, thus violating the expected control flow of the
+caller without actually violating the caller's integrity. But it also
+encompasses simpler violations, such as a function bypassing a password
+check by jumping (or returning) in the middle of a block of code.
+%
+Stack safety enforcement mechanisms like \citet{Skorstengaard+19}
+usually aim to prevent such unwanted flows: every jump between
+functions should either be a call or a return, and each return should
+be to the instruction after its corresponding call.
+
+To formally capture this aspect of stack safety, we introduce some more
+security properties. {\em Control separation} restricts transitions between
+functions to follow the interfaces set out by call, return, and yield, forming
+a foundation for all of the other properties that rely on those interfaces for
+meaning. {\em Entry integrity} ensures the validity of call targets,
+{\em return integrity} ensures that of return targets, and
+{\em yield-back integrity} ensures that of yield targets.
+
+Recall that our properties use a compiler-derived code-map determine
+which addresses are instructions and which functions they belong to, as well
+as annotating them as calls, returns, and yields, as well as sharing information.
+We now extend it with an annotation for valid entry points:
+\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
+\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
+\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]
+
+\definition%{Control Separation}
+
+Let \(\mpcstatename\) be a reachable state whose program counter points to code in
+function \(\funid\), and \(\mpcstatename \stepstopol \mpcstatename'\) where
+\(\mpcstatename'\) is state whose program counter points to a different function
+\(\funid'\). A system enjoys {\em control separation} when all such \(\mpcstatename\)
+are annotated \(\callmap\), \(\retmap\), or \(\yieldmap\).
+
+\definition%{Entry Integrity}
+A system enjoys {\em entry integrity} when, for any reachable \(\mpcstatename\) and
+\(\mpcstatename'\) such that \(\mpcstatename\) is annotated \(\callmap\) and
+\(\mpcstatename \stepstopol \mpcstatename'\), \(\mpcstatename'\) is annotated \(\entmap\).
+
+\definition%{Return Integrity}
+A system enjoys {\em return integrity}... [this one needs to talk about context.]
+
+\definition%{Yield-back Integrity}
+A system enjoys {\em yield-back integrity}... [same.]
+
+\definition% {Well-Bracketed Control Flow}
+The above properties are somewhat orthogonal, but together they
+capture a standard notion of control flow for a stack: functions keep
+their code separate outside of calls and returns, calls respect
+explicit entry points, and returns respect implicit entry points
+corresponding to their call.  A system with all three
+properties (with respect to a function map, call map, return map,
+and entry map) is said to enjoy
+{\em well-bracketed control flow} (with respect to those maps).
 
@@
 local state encapsulation component by checking for integrity and
-confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}
+confidentiality violations at every step of the callee's execution trace.
 
@@
 At each step of the machine, the relevant rule is applied to the tags on
-the instruction's inputs ($\PCname$, registers, memory) and on the instruction itself, and produces
+the instruction's inputs (PC, registers, memory) and on the instruction itself, and produces
 one of two outcomes: either the instruction is allowed to execute (and generates
-tags for the result of the operation and the new $\PCname$), or the machine
+tags for the result of the operation and the new PC), or the machine
 fail-stops with a policy violation.
-\citeauthor{pump_oakland2015}\bcp{and PIPE, if there are any references to
-  cite} have shown that a wide range of micro-policies can
+\citeauthor{pump_oakland2015} have shown that a wide range of micro-policies can
 be defined using this scheme.
@@
 \label{sec:conservative}
-
-\bcp{Is this the only subsection in this section?}
-
+%
 The first micro-policy we present to enforce the definition of stack
@@
 %\rb{actually, the ``initial frame'' should be tagged with its proper depth}
-and the {\PCname} has tag $\tagPCDepth{0}$, where $n$ is the current activation depth.
-These tags are altered
+and the {\PCname} has tag $\tagPCDepth{0}$. These tags are altered
 at exactly those points in
@@
   $\PCname$ tag is decremented.
-  \rb{Problem: this is currently the first mention of $\ret{\cdot}$}
 
@@
   is clear}
-\rb{TODO: References to the running example are no longer valid! We could use
-figures 5 or 6, referencing the figures, not the invisible assembly code.}
 %
@@
 
-We will describe observable integrity in our most detailed setting. In the
-presence
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\sealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(\depth' < |\rtm ~ \stackid|\) .
-Then a system enjoys observable
-
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+\paragraph{Rolling Back to Idealized State}
+
+Observable integrity weakens eager integrity by allowing a returned
+state to differ from its call state at high integrity components,
+provided that the differences do not change the observable behavior of
+the rest of the program. The challenge here is to define what constitutes
+a change in behavior. We construct an idealized return state that
+contains only the changes that the caller was allowed to make, and use
+the trace from that state as a baseline against which the real trace
+will be compared.
+
+      \definition
+
+      Given a contour, a call state, and a returned state, we define a
+      {\em rollback} function \(\mathit{roll} \in \CONTOURS \times \MACHS \times
+      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
+      the call state on those components that are high integrity in its
+      contour, and matches the return state on low integrity components:
+      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
+      \begin{cases}
+        \mach_e(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
+        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
+      \end{cases}\]
+
+      If \(\mach_e\) is an entry to a call with contour \(\contour\)
+      and \(\mach_r\) its return, the result of
+      \(\mathit{roll}(\contour,\mach_e,\mach_r)\) agrees with the
+      entry \(\mach_e\) on every component the callee should not
+      change, and with the actual return \(\mach_r\) on those
+      components that it is permitted to change.  So the observable
+      behavior that follows is the idealized behavior to which we
+      compare the trace from \(\mach_r\).
+
+      Now we can define a version of trace integrity that compares the trace
+      from a returned state to that from its idealized counterpart.
+
+    \definition%{Observable Trace Integrity}
+
+      A trace enjoys {\em observable trace integrity} with respect to
+      some contour \(\contour\) if the trace induced by its final
+      state (if any) has observable behavior that prefixes that of its
+      idealized version (to account for premature halts due to policy
+      violations):
+
+      \[\begin{split}
+        \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
+        & \head(\MPT) = \mpostate{\mach_e}{\_}{\_} ~ \land ~
+          \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
+          \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
+          \Rightarrow \\
+        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
+      \end{split}\]
+
+%    \begin{figure}
+%      \integritylazyexample
+%      \caption{Rolling back from Figure \ref{fig:intex}.}
+        %\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
+%      \label{fig:intlex}
+%    \end{figure}
+
+    For our example, we return to the violation of trace integrity in
+    \cref{fig:intex}, in which {\tt g} has written to {\tt z} and returned.
+    Whether observable integrity holds depends on the continuing trace, shown
+    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
+    \(\mach_r\), and the rolled-back state, \(\mach_r' =
+    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
+    components are restored to their values as of \(\mach_e\). Specifically,
+    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
+    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
+    it prints 0. Only then is observable trace integrity violated.
+
+    However, consider instead if we modify the original code of \cref{fig:running-trace-b} to insert after instruction 44 a new
+    instruction \(44.5: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
+    {\tt f}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
+    rolled-back state \(\mach_r'\) maps it to 0. But in the trace from either
+    state, instructions 27 and 28 load the return value (18) and store it to
+    105, eliminating the only difference. The remainder of the observation
+    traces must be identical, so this trace does obey observable trace
+    integrity.
+
+    \definition%{Observable Stack Integrity}
+
+      A system enjoys {\em observable stack integrity}, like stack integrity,
+      if, for any initial state and each \(n\)-transition from
+      \(\mpstate{\mach_c}{\pol_c}\) to \(\mpstate{\mach_e}{\pol_e}\) in its
+      induced trace, we take the prefix of the call
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\)
+      and observable trace integrity holds on \(\MPT_{\mathit{call}}\) with
+      contour \(\mathit{Cof}(\mach_c,n)\).
 
   \subsection{Observable Confidentiality}
+
+      Two of the cases of trace confidentiality (\cref{def:trace-conf}), \(\mathit{InfCase}\)
+      and \(\mathit{FailCase}\), are defined solely in terms of
+      observation traces. If
+      the traces are infinite, they should be similar; if one
+      fail-stops, it should be a prefix. Only the
+      \(\mathit{ReturnCase}\) checks machine states, because in that
+      case the traces could continue, but have been cut off by a
+      return. The key to making confidentiality lazy is not to perform that
+      check but rather to consider the traces after that return,
+      checking only whether they are observationally similar.
+
+      This requires restoring varied components to their original
+      values in the trace from the variant state, since after return
+      they will once again be legitimately accessible. To return to
+      the example in \cref{fig:confex}, the varied final state
+      \(\mach_r'\) still has $-5$ as a return address for {\tt f}, so
+      the trace after that return will behave oddly in a way that
+      does not reflect a confidentiality violation.  That means we
+      cannot immediately compare traces from \(\mach_r\) and
+      \(\mach_r'\), but rather that we must restore the varied values
+      while preserving differences that have occurred during
+      execution.
+
+      \definition
+
+      The \(\mathit{restore}\) function takes four states: the
+      original entry state, a variant entry state, and corresponding return
+      states. It returns another state. The resulting state restores those
+      components that varied between entries to match the original entry,
+      unless they were changed during execution:
+      \[\mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')
+        (\component) =
+        \begin{cases}
+          \mach_e(\component) & \text{if } \mach_e(\component) \not =
+                                \mach_e'(\component) \text { and } \\
+          & \mach_e(\component) = \mach_r(\component) \text{ and }
+            \mach_e'(\component) = \mach_r'(\component) \\
+          \mach_r'(\component) & \text{else} \\
+        \end{cases}\]
+
+      Now we can describe a trace property akin to trace confidentiality, but
+      continuing execution after the end of the callee trace.
+
+    \definition{%Observable Trace Confidentiality}
+
+      {\em Observable trace confidentiality} is the lazy analog to
+      trace confidentiality, and reuses the cases in which the traces
+      do not return.  For a return predicate \(R\) and a pair of
+      traces \(\MPT\) and \(\machT\), we consider the return case, in
+      which $\MPT$ has a last element
+      \(\mpostate{\mach_r}{\_}{\_}\) and that machine state is
+      a return, i.e. \(R ~ \mach_r\) holds.
+%
+      Observable trace confidentiality then requires that $\machT$
+      also has a final state \((\mach_r',\_)\) that is a return, i.e.
+      \(R ~ \mach_r'\) holds.
+%
+      In addition, just like in the eager case,
+      observations of \(\MPT\) are similar to those of \(\machT\).
+
+      The only thing that changes is the requirement that \(\mach_r\)
+      and \(\mach_r'\) agree on components. Instead, let \(\mach_e\)
+      be the machine state of \(\head(\MPT)\), and \(\mach_e'\) be
+      that of \(\head(\machT)\). We restore \(\mach_r'\) based on
+      \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
+      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).  Then
+      observable trace confidentiality requires that the
+      induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
+      observational prefix of the induced trace from \(\mach_r'\),
+      reflecting the possibility of a fail-stop after return.
+
+      \[\begin{split}
+        \mathit{ObsReturnCase} ~ R ~
+        (\mpostate{\mach_e}{\_}{\_}\MPT) ~
+        ((\mach_e',\_)\machT) \ \ \triangleq \ \
+        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land ~
+         \last(\machT) = (\mach_r',\_) ~ \land \\
+        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') ~ \land \\
+        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
+        \mach_r'' \hookrightarrow \machT' \Rightarrow \\
+        & \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land ~
+        \pi_o(\MPT') \lesssim \pi_o(\machT') \\
+      \end{split}\]
+
+      We combine this case with the other two cases from ordinary trace
+      confidentiality:
+\[
+        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
+         \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT ~ \land
+         \mathit{InfCase} ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
+          \MPT ~ \machT
+\]
+
+%      \begin{figure}
+%        \confidentialitylazyexample
+%        \caption{Continuing from \cref{fig:confex}.}
+%        \label{fig:conflex}
+%      \end{figure}
+
+      In our example, \cref{fig:conflex} shows how execution continues from
+      \(\mach_r'\) and the restored
+      \[\mach_r'' = \mathit{restore}(%\contour,
+      \mach_e,\mach_e',\mach_r,\mach_r').
+      \]
+      Addresses 106 and 107
+      are not restored; every address that was varied and not changed in one
+      of the executions is. So 107 contains 43 in the trace from \(\mach_r\),
+      and after returning from {\tt f}, {\tt main} prints 85. But in the trace
+      from \(\mach_r''\), 107 contains $-4$, and ultimately 38 is printed.
+      So observable trace confidentiality does not hold. This example
+      illustrates the weakness of observable confidentiality as a testing
+      property: it takes until the very end of the program to determine whether
+      a tainted value actually influences observable behavior.
+
+      On the other hand, suppose that we inserted an instruction at 43:
+      \(r_6 \leftarrow [\rsp-8]\). \(r_6\) would hold 42 in \(\mach_r\)
+      and $-5$ in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
+      never used otherwise, the traces of {\tt f} will enjoy observable trace
+      confidentiality.
+
+    \definition%{Observable Stack Confidentiality}
+
+      {\em Observable stack confidentiality} is structured like stack
+      confidentiality. For each initial
+      state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
+      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
+      \(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\), take any
+      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
+      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~
+      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~
+      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
+      \ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}\).
+
+      Additionally, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
+      \mach_0\), taking the induced trace  \(\mach_0' \hookrightarrow
+      \machT\), we must have \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\).
+
+      A system enjoys observable stack confidentiality if observable
+      confidentiality holds in all such cases.
+
+    \definition%{Observable Stack Safety}
+    A system enjoys {\em observable stack safety}
+    % with respect to a call map       and a function map
+    if it enjoys both observable stack integrity and
+      observable stack confidentiality with respect to them.
+
+      \medskip
+      \textsc{Theorem.}
+      Stack safety implies observable stack safety.
+      \smallskip
+
+      This is more straightforward than the implication in the previous
+      section (1,5K LoC in Coq) as the recursive
+      structure of the two properties is similar.
 
@@
 programs. Although the required hardware support is readily available
-in capability machines like CHERI, this technique incurs significant
-costs, because it requires the entire unused part of the stack to be
+in capability machines like CHERI\bcp{including linear capabilities?? or
+  don't they use them in this paper?}, this technique incurs significant
+costs\bcp{This is the pot calling the kettle black}, because it requires the entire unused part of the stack to be
 cleared whenever a security boundary is crossed. Their logical
@@
 
-We conjecture that the linear capability machines introduced by
+We believe that the linear capability machines introduced by
 \citep{Skorstengaard+19}, in combination with the StkTokens calling convention,
@@
 \cref{sec:lse-and-wbcf}. Further, we expect them to satisfy the even stronger
-property developed in \cref{sec:enforcement}. Proving\bcp{or testing!} these conjectures would
-involve materializing\bcp{?} their definitions of local state encapsulation and
-well-bracketed control flow, which exist only in enmeshed\bcp{?} form as part of the
+property developed in \cref{sec:enforcement}. Proving these conjectures would
+involve materializing their definitions of local state encapsulation and
+well-bracketed control flow, which exist only in enmeshed form as part of the
 semantics of the capability machines. The main practical limitation of StkTokens
 is its reliance on linear capabilities, as it is unclear now they could be added
-to practical capability machines, especially in terms of efficiency (previous
+to practical capability machines, especially in terms of efficiency---previous
 work on micro-policies \citep{yannis-report} has shown how to use that framework
 to implement linear return capabilities.\bcp{But IIUC the Dover realization
-  of micro-policies can't do this...?})
+  of micro-policies can't do this...?}
 
@@
 Heap safety, like stack safety, may be framed as a security property
-in the form of
-noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
-be ``Azedevo et al.''}\fi. Just
-as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
+in the form of noninterference~\citep{DBLP:conf/post/AmorimHP18}. Just
+like~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
 characterization of the meaning of (heap) memory safety, in this paper
@@
 Their model describing safety of shared stack-allocated
-objects\bcp{??} extends naturally to heap safety, complicated by the fact that
-these objects are not deallocated by returns\bcp{??}, and therefore a
+objects extends naturally to heap safety, complicated by the fact that
+these objects are not deallocated by returns, and therefore a
 function's privilege may increase or decrease after its entry
@@
 properties separately on continuous segments of each call, with
-contours computed at each crossing between caller and callee.  \bcp{Didn't
-  understand last sentence.}
+contours computed at each crossing between caller and callee.
 
@@
   state is delicate; we conjecture it can be done with some effort.
-
-\BCP{This section seems incomplete?}
 

### 2021-04-15 tex/main.tex
More explanations in examples

--- 2021-04-15 tex/main.tex
@@
 (in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
-(outside the stack entirely.)
+(outside the stack entirely.) For our purposes, \(\outside\) will apply to global
+variables. Domains are part of the {\em context} of a state.
+We will introduce the remaining context as needed.
 
@@
 meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of
-their owner\bcp{are we still using depths?},
+{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
 in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
@@
 of the context of its call (except for its arguments and any global
-variables\bcp{We haven't talked about global variables.  We should.}).
-We state this in a style similar\bcp{??} to {\em noninterference}~\citep{??}.
-Noninterference describes knowledge: it states that for some system components
+variables).
+We state this as a kind of {\em noninterference} property~\citep{??}.
+Noninterference describes knowledge: it states that for system components
 containing secret values, we can substitute any values whatsoever without changing the results
@@
 
-\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
-track which addresses may be read (i.e., which have been initialized), and at each step
-guarantee that the step alters state identically to a step from a variant state. Then we
-wouldn't need to divide up the execution by calls. But the approach we take is more
-consistent with a typical noninterference approach, in that it varies the memory once
-at the beginning and trace changes through the execution.
+%\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
+%track which addresses may be read (i.e., which have been initialized), and at each step
+%guarantee that the step alters state identically to a step from a variant state. Then we
+%wouldn't need to divide up the execution by calls. But the approach we take is more
+%consistent with a typical noninterference approach, in that it varies the memory once
+%at the beginning and trace changes through the execution.
 
@@
       \memory{1}{\mainsealc}[Seal(0)]%
-      \memory{1}{\goodc}[RetPtr]%
+      \memory{1}{\unsealc}[RetPtr]%
       \memory{2}{\unsealc}%
@@
       \memory{1}{\mainsealc}
-      \memory{1}{\goodc}
-      \memory{2}{\unsealc}%
+      \memory{3}{\unsealc}%
       ~$\cdots$
@@
 One of the most pernicious ways to attack the stack is by hijacking control flow.
-Our stack safety properties do not assume that control flow is
-well-bracketed\bcp{undefined term}; rather,
-they express what it means to protect data even in the presence of control-flow attacks.
-Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''\bcp{??}~\citep{??}.
-(Here we show return addresses explicitly as line numbers, in the form $\#n$.)
-\apt{what do the colors mean in this example?}
+Our stack safety properties do not assume that control flow follows a typical structure
+of matching sets calls and returns nested within on another.
+Rather, they express what it means to protect data even in the presence of control-flow attacks.
+Consider example \ref{fig:controlflow}.
+Here we show return addresses explicitly as line numbers, in the form $\#n$.
 Function {\tt f} is called twice, at lines 2 and 4; it should therefore return to lines
@@
 
-The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs to do more than just
-execute a return-labeled instruction: it must actually reach the appropriate
-{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep
-a stack of return targets\bcp{do we still},
-which correspond to the depths at which stack frames have been sealed. A callee is
-only considered to have returned when it executes a return-annotated instruction
-and then reaches a valid return target\bcp{what's that?}. Only then do the sealed frames become unsealed.
+The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
+to do more than just execute a return-labeled instruction: it must actually reach the
+appropriate {\em return target} in {\tt main}. A return target is a predicate on states,
+in this case one that holds on any state whose program counter points to the correct line.
+We introduce a new piece of context: we keep a stack of return targets, and a callee is
+only considered to have returned when it reaches a state that matches the caller's target.
+A return target is associated with a set of sealed locations that will become unsealed
+when it is reached, here represented by sharing a color with the sealed data.
 
@@
 \(\PCname\) is 3 after its return. But on the second, {\tt f} does not properly return,
-but rather changes its stack pointer and jumps into {\tt main}'s code. Then {\tt f}
-attempts to overwrite the sealed value of {\tt x}, violating integrity.
+but rather changes its stack pointer (shown in light red) and jumps into
+{\tt main}'s code. Then {\tt f}
+attempts to overwrite the (still sealed) value of {\tt x}, violating integrity.
 

### 2021-04-15 tex/main.tex
example 3 formatting
Moved passing examples to after some formalism

@@
 
-\paragraph*{Pass by Reference and Stack Arguments}
-\bcp{I wonder if this could be postponed so that we get to the formal
-  definitions for the basic property sooner...}
+\section{Machines, Traces, and Observations}
+\label{sec:prelim}
+
+We next define our machine model precisely. To make our definition of stack safety
+as generic as possible, this section proposes an abstract interface to a
+machine model plus some kind of (software- or hardware-enforced) policy monitor.
+In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
+policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
+
+\subsection{Values, States, and Observations}
+
+The basic building blocks of the machine are {\em values} and {\em addresses}.
+Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\).
+%
+Our machine states are composed of {\em components} \(\component\),
+addressed\bcp{no, the components are not addressed, they are addresses!} either by
+words (memory addresses) or by register names (\(\reg\)) drawn from some set
+\(\REGS\).
+%
+    \[\component \in \COMPONENTS = \WORDS + \REGS \]
+%
+The register names are assumed to include two special purpose registers: the
+program counter {\PCname} and the stack pointer \(\SP\).
+
+A {\em machine state} is a map from components to values:
+%
+\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]
+
+The step function of the machine takes a machine state and produces a
+pair of a machine state and an {\em observation}, which may be either
+silent (written \(\tau\)) or a word of data that the machine
+communicates to the outside world:
+\ifaftersubmission\bcp{This is a rather concrete
+  choice!  Why not leave the set of observations abstract?}\fi
+%
+\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
+\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
+%
+This step function is total, so the machine's execution traces
+(defined in \cref{sec:traces}) are infinite.
+
+Observations are only essential for describing the extensional properties
+described in Section~\ref{sec:lazy}, but for simplicity, we include them
+in the machine model from the beginning.
+We choose to model observations explicitly, rather than simply treating
+a subset of registers or memory as being observable.  This choice gives
+us the flexibility to treat programs as observationally equivalent even when
+their internal states may differ.
+%The choice of modelling observations is an interesting one. A first approach
+%would be to consider any change to the registers or memory as observable.
+%However, as some parts of the stack will be considered ``private''
+%this choice won't work. A relaxation of this to consider all changes
+%to ``non-private'' data as observable could work in principle, but
+%does not allow for an interesting weaker-but-more-efficient enforcement
+%mechanism. We will come back to this point in \Cref{sec:lazy}.
+
+\apt{This para is out of date. Do we want to say something equivalent
+  about new examples (hard, because they are C-based).}
+$\WORDS$ is the
+set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
+\cup \{\PCname,\SP\}$. Observations are writes to the
+special $\rout$ location\ifaftersubmission\bcp{Which is 100! (do we need a name for it? seems
+  not...)}\fi, which can be thought of as a memory-mapped
+  output port: if during an execution step a word $w$ is
+written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.
+
+\subsection{Policies}
+
+We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
+A {\em policy} consists of a set of policy
+states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
+\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
+it is undefined on input configurations that correspond to a policy fault. The policy
+can see the machine state, but cannot modify it. We assume that the
+compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.
+%
+A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
+\Cref{sec:enforcement}.
+
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
+%
+\[\mpstatename \in \MPS = \MACHS \times \POLS\]
+%
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach \stepstoobs[\obs] \mach'\)}{\((\mach, \pol)
+               \polstep \pol'\)}
+            {\(\mpstate{\mach}{\pol} \stepstopol[\obs]
+               \mpstate{\mach'}{\pol'}\)}
+
+\noindent
+This combined step function is also partial, and traces of the machine-with-policy
+will be finite iff a policy fault occurs.
+
+\subsection{Contexts}
+
+Separate from the machine state, our formal properties will keep track of additional context
+information representing the history of the run. The following definitions are parameterized
+over an arbitrary\bcp{why?} set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
+\(\mach, \context, \anns \constep \context' \in \MACHS \times \CONTEXTS
+\times \powerset(\ANNS) \rightarrow \CONTEXTS\). We also assume a set of function identifiers
+\(\FUNIDS\) and the existence of a \emph{code map} \(\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\)
+that relates each code address to its owning function and set of annotations.
+We lift contexts into an MP trace in much the same way as we lifted policy states to get
+a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
+\CONTEXTS\).  \bcp{I don't get the intuition for this whole para.}
+
+\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
+              {\(\mach, \context, \anns \constep \context'\)}
+              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
+                \mpcstate{\mach'}{\pol'}{\context'}\)}
+
+\subsection{Traces}
+\label{sec:traces}
+
+A trace over a type \(\TRACEELEMS\)\bcp{``over a type A'' is a Coq-ism} is a nonempty, finite or infinite sequence of
+\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
+\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
+called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
+
+\paragraph*{Trace-Of}
+
+The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
+coinductively relates an initial state with the trace of states
+produced by repeated application of step, and potentially cut short by
+the policy component. The {\sc Default} rule applies only if the other does not.
+(We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
+
+\begin{minipage}{.4\textwidth}
+\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
+            {\(\mpcstatename' \hookrightarrow \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
+\end{minipage}
+\begin{minipage}{.4\textwidth}
+\judgment[Default]
+         {}
+         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
+\end{minipage}
+%
+
+\paragraph*{Projections}
+
+We project out the machine state of an MPC-state with \(\pi_\mach\), the
+policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
+these operators to project out of an machine-trace a trace of machine states,
+policy states, and contexts, respectively.
+
+\paragraph*{Observations-Of}
+
+\bcp{Could any of this be postponed?}
+
+The ``observations-of'' operator, written \(\pi_\obs\), \apt{since this isn't
+  actually a projection, it is slightly confusing to use \(\pi\)} takes a machine trace and
+creates a trace of observations corresponding to the observations of each step in
+the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
+initial observation, a machine-trace, and an observation trace:
+
+\judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
+              {\(\mpcstatename' \hookrightarrow \MPCT\)}
+              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
+              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
+\judgment{}
+         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}
+
+Then to take the observations from any particular trace, we start with a \(\tau\) observation:
+
+\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
+         {\(\pi_\obs(\MPCT) = \obsT\)}
+\noindent
+
+In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
+to relate a machine-policy-context state triple to the MPCO-trace
+induced by the step function.\apt{?? This looks indistinguishable
+  from the \(\hookrightarrow\) we've already defined.}
+
+\paragraph*{Head and Last}
+
+We take the first element of a trace with \(\head(\Trace)\), which is a total
+function since traces are non-empty, and the final element (if one exists) with
+\(\last(\Trace)\), which is partial.
+
+\paragraph*{Until}
+The operation \(\PUT ~ f ~ \Trace\) takes a trace
+\(\Trace \in \TRACE{\TRACEELEMS \times \OBSS}\) and a predicate
+on elements \(f \subset \TRACEELEMS\) and gives the prefix of
+\(\Trace\) ending with the first element on which \(f\) holds.
+
+\begin{center}
+  \begin{minipage}{.3\textwidth}
+    \judgment{\(f ~ \traceelem\)}
+             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
+  \end{minipage}
+%
+  \begin{minipage}{.3\textwidth}
+    \judgment{}
+             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
+%
+  \end{minipage}
+  \begin{minipage}{.3\textwidth}
+    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
+                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
+  \end{minipage}
+\end{center}
+%
+If the resulting trace is a strict prefix of the parameter, then the
+predicate must hold on its final element. If the predicate always holds, then
+\(\PUT\) is the identity function.
+
+% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
+%$\mathit{split}$: it takes two traces and links them by replacing the
+%last element of the first (if it has one) with the first element of
+%the second
+
+%\judgment{} {\(t \cdot T = T\)}
+%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
+%\leo{Lemma/theorem: split/join = id?}
+%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
+
+We will frequently take a prefix of the trace from an initial state up
+through the first state where some condition holds on the machine state.
+This can easily be implemented using \(\PUT\), and we provide special notation
+for convenience.
+%
+If \(f\) is a predicate on machine states, we define \(\mpcstatename
+\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
+running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:
+
+\begin{center}
+\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
+            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
+\end{center}
+
+\paragraph*{Context Segments}
+
+In particular, we wish to formulate properties about consecutive subtraces
+in which the context obeys a given predicate. This will commonly be used to extract
+individual calls from a trace by taking subtraces in which the stack is
+at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
+we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
+``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''
+
+\begin{center}
+  \judgmenttwo[Now]
+              {\(f ~ \mach ~ \context\)}
+              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
+              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
+%
+  \judgmentthree[Skip]
+                {\(f ~ \mach ~ \context\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
+%
+  \judgmentthree[Search]
+                {\(\neg f ~ \mach ~ \context\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
+\end{center}
+
+\paragraph*{Reachability}
+
+We often want to quantify over all traces from any initial state so that our
+properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
+is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
+and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
+such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
+
+\paragraph*{Observational Similarity}
+
+We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
+as far as an external observer is concerned, written \(\obsT_1 \lesssim
+\obsT_2\), if the sequence of
+non-silent observations of $\obsT_1$ is a prefix of those of
+$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
+coinductively:
+
+\begin{minipage}{.3\textwidth}
+  \judgment{}{\(\obsT \lesssim \obsT\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{}{\(\tau \lesssim \obsT\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{}{\(w \lesssim w\obsT\)}
+\end{minipage}
+
+\begin{minipage}{.3\textwidth}
+  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
+           {\(\tau \obsT_1 \lesssim \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
+           {\(\obsT_1 \lesssim \tau \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
+           {\(w\obsT_1 \lesssim w\obsT_2\)}
+\end{minipage}
+
+\smallskip
+We then define similarity of observation traces as traces prefixing each other:
+\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
+%
+%\leo{I still don't know how to do this transition. Maybe contrast
+%with the ``perhaps surprisingly'' later?} Note that
+%similarity of observation traces could alternatively be defined coinductively,
+%just like the prefix relation, but by dropping the asymmetric rules that allow
+%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}
+
+Note that an infinite silent trace is a
+prefix of (and similar to) any other trace. While this might seem
+surprising at first, it makes sense in a timing-insensitive context:
+an external observer looking at two machine runs cannot (computably)
+distinguish between a machine that steps forever and a machine that
+steps for a long time before producing some output.
+
+\section{Stack Safety, Formally}
+\label{sec:lse}
+
+We are finally ready for our definition of stack safety. We begin
+by describing how we model knowledge about the program structure,
+then dive into the formal definitions of local state encapsulation and
+well-bracketed control flow, which together constitute stack safety\bcp{not
+  any more?}.
+Initially we will focus on a simple stack model, with a single stack
+that doesn't share between caller and callee. All arguments are passed
+in registers. Recall that we keep track of which code addresses are
+annotated as calls, with a ``code map'':
+\[\codemap \in \CODEMAPS \subseteq \WORDS\]
+When a function makes a call, it establishes that its frame should be
+``sealed'' for later use. At any given step of a program, any number of
+activations may have claimed memory in this way, but which cells are sealed
+is dynamic information and cannot be deduced from the state alone. So it must
+be tracked separately. We must also track the conditions under which a caller
+considers itself to have been returned to.
+
+Each system component will be assigned to a {\em domain}, which may be {\em Outside}
+of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
+at a particular depth. Each caller will register a {\em target}
+condition -- a predicate on machine states that must hold for it to consider itself active.
+A context is a pair of a ``domain map'' from components to domains, and a
+``return target list,'' which is a stack of targets, one for each caller awaiting
+return.  \ifaftersubmission\bcp{spacing is funny}\fi
+
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+  \times (\rts \in \listT (\target \subseteq \MACHS)) \]
+
+An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
+the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
+
+We parameterize over the structure of the stack, particularly the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
+maps a machine state to a set of addresses.
+If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
+which addresses must be protected until the return. Similarly, we abstract over what
+it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
+which relates call states to the states that can be considered valid returns to their caller.
+The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
+and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
+But whatever the return convention, at each step we update the context according
+to the context-step function, \(\stepsto_C\).
+
+The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
+A call annotation tells us to push the return target onto the return stack and
+seal all components with the depth of the previous return stack. If we aren't making a call,
+we identify a return by looking ahead and checking if we're about to step to a state
+that matches the topmost return target. If neither is true, the context remains unchanged.
+
+\judgmentthree[Call]
+              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
+              {\(\components = \sealcon ~ \mach\)}
+              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
+                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\judgmentthree[ReturnFound]
+              {\(\mach \stepsto \mach' \in \target\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\judgmenttwobr[ReturnRec]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\judgment[Default]
+         {}
+         {\(\_,\context \stepsto_C \context\)}
+
+When we build machine traces with this step function, we always know how deep in the
+call stack we are and which components are sealed. We may now proceed to state our properties.
+Our first property will be {\em stepwise stack integrity}, stepwise meaning that it will be
+quantified over each step of execution from any initial state. The criterion for integrity
+is simple: sealed components do not change.
+
+\definition A system enjoys {\em stepwise stack integrity} if for any reachable state
+\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
+and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
+\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
+\pi_\mach(\mpcstatename') ~ \component\).
+
+Next we consider {\em stepwise stack confidentiality}. There are a number of preliminary
+concepts we must discuss here -- confidentiality quantifies over {\em variants} of the
+initial state of a call, and requires that the traces from them move in {\em lockstep}.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if, for
+all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\). We write
+this \(\mach \approx_\components \nach\).
+
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change together
+if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
+\(\nach ~ \component \not = \nach' ~ \component\), \(\mach' ~ \component = \nach' ~ \component\).
+This is written \(\mach,\mach' \diamond \nach,\nach'\).
+
+\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
+change together, written as a coinductive relation \(\machT \doteq
+\nachT\):\bcp{People may not be familiar with the notation}
+
+\begin{minipage}{.3\textwidth}
+\judgment{}
+         {\(\mach \doteq \nach\)}
+\end{minipage}
+\begin{minipage}{.6\textwidth}
+\judgmenttwo{\(\mach, \head(\machT) \diamond \nach, \head(\nachT)\)}
+            {\(\machT \doteq \nachT\)}
+            {\(\mach \machT \doteq \nach \nachT\)}
+\end{minipage}
+
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
+\(\MPCT\) corresponds to the trace of a subroutine from its call to its return.
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
+\(\MPCT \doteq \NPCT\).
+
+\subsection{Stack Safety with Passing}
 
@@
 1:          x = 5;
-2[P($SP)C]:  x = f(x);
+2[C($SP)]:  x = f(x);
 3[R]:       return x;
@@
   \memory{1}{\mainsealc}[Seal(0)]%
-  \memory{1}{\mainpassc}[P(0-1)]%
+  \memory{1}{\mainpassc}[P(0,1)]%
   \memory{1}{\retptrc}[RetPtr]%
@@
 0:         int x, y;
-1[P(x)C]:  f(x);
+1[C(x)]:  f(x);
 2:         x = 0;
-3[P(y)C]:  f(y);
+3[C(y)]:  f(y);
 4[R]:      return x;
@@
       \memoryaddrs[(0)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
       \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
@@
 To this point we have ignored function arguments, since our simplest model places them in
-registers. This is an oversimplification, which we will now correct, starting with the simple
-case: passing by value on the stack. Example \ref{fig:passingsimple} shows a sample trace of
+registers. This is an oversimplification, which we will now correct, aiming at two kinds
+of sharing: passing by value on the stack, and passing by reference. First we will explore
+some examples at the same level of detail as our earlier ones, and then move on to
+extending our formalization.
+
+Example \ref{fig:passingsimple} shows a sample trace of
 a program that passes a value on the stack, without an attack.
-At the call, we add a new annotation \(P(\$SP)\) identifying a memory location to be
-passed, and our diagram displays it in magenta. The caller's frame locations that are not
-passed are sealed as usual. Just as we marked sealed memory with the depth of its owner\bcp{??},
-we do the same for passed memory -- but it has multiple owners. We keep track of the
+We extend our call annotation with an additional parameter identifying a memory location to be
+passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
+to include \(\passed{\depth,\depth'}\), recording two depths. We note the
 current depth of the active function next to the stack pointer in the diagram, for reference.
@@
 stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike the case for pass-by-reference, there is no easy way to model the intended sharing behavior
-of such pointers in terms of stack frame activation patterns.  The best approximation we can make is to
-treat an address-taken variable in the same way as a passed argument, with the additional property
+Unlike the case for pass-by-reference, there is no easy way to model the intended
+sharing behavior of such pointers in terms of stack frame activation patterns.
+The best approximation we can make is to treat an address-taken variable in the
+same way as a passed argument, with the additional property
 that it remains accessible at \emph{all} stack depths above the caller.
 This prevents most violations of temporal safety, although stale pointers can still be
-used in certain corner cases. \apt{Trying not to spell out the details, which are complex.}
+used in certain corner cases.
 More seriously, callees are permitted to access the location without having
@@
 \emph{per se}.
-
-\section{Machines, Traces, and Observations}
-\label{sec:prelim}
-
-We next define our machine model precisely. To make our definition of stack safety
-as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (software- or hardware-enforced) policy monitor.
-In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
-policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
-
-\subsection{Values, States, and Observations}
-
-The basic building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\).
-%
-Our machine states are composed of {\em components} \(\component\),
-addressed\bcp{no, the components are not addressed, they are addresses!} either by
-words (memory addresses) or by register names (\(\reg\)) drawn from some set
-\(\REGS\).
-%
-    \[\component \in \COMPONENTS = \WORDS + \REGS \]
-%
-The register names are assumed to include two special purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-
-A {\em machine state} is a map from components to values:
-%
-\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]
-
-The step function of the machine takes a machine state and produces a
-pair of a machine state and an {\em observation}, which may be either
-silent (written \(\tau\)) or a word of data that the machine
-communicates to the outside world:
-\ifaftersubmission\bcp{This is a rather concrete
-  choice!  Why not leave the set of observations abstract?}\fi
-%
-\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
-\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
-%
-This step function is total, so the machine's execution traces
-(defined in \cref{sec:traces}) are infinite.
-
-Observations are only essential for describing the extensional properties
-described in Section~\ref{sec:lazy}, but for simplicity, we include them
-in the machine model from the beginning.
-We choose to model observations explicitly, rather than simply treating
-a subset of registers or memory as being observable.  This choice gives
-us the flexibility to treat programs as observationally equivalent even when
-their internal states may differ.
-%The choice of modelling observations is an interesting one. A first approach
-%would be to consider any change to the registers or memory as observable.
-%However, as some parts of the stack will be considered ``private''
-%this choice won't work. A relaxation of this to consider all changes
-%to ``non-private'' data as observable could work in principle, but
-%does not allow for an interesting weaker-but-more-efficient enforcement
-%mechanism. We will come back to this point in \Cref{sec:lazy}.
-
-\apt{This para is out of date. Do we want to say something equivalent
-  about new examples (hard, because they are C-based).}
-$\WORDS$ is the
-set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
-\cup \{\PCname,\SP\}$. Observations are writes to the
-special $\rout$ location\ifaftersubmission\bcp{Which is 100! (do we need a name for it? seems
-  not...)}\fi, which can be thought of as a memory-mapped
-  output port: if during an execution step a word $w$ is
-written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.
-
-\subsection{Policies}
-
-We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
-A {\em policy} consists of a set of policy
-states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. We assume that the
-compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.
-%
-A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-\Cref{sec:enforcement}.
-
-We write $\MPS$ for the set of all pairs of machine states and policy
-states.
-%
-\[\mpstatename \in \MPS = \MACHS \times \POLS\]
-%
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepstoobs[\obs] \mach'\)}{\((\mach, \pol)
-               \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol[\obs]
-               \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-This combined step function is also partial, and traces of the machine-with-policy
-will be finite iff a policy fault occurs.
-
-\subsection{Contexts}
-
-Separate from the machine state, our formal properties will keep track of additional context
-information representing the history of the run. The following definitions are parameterized
-over an arbitrary\bcp{why?} set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
-an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context, \anns \constep \context' \in \MACHS \times \CONTEXTS
-\times \powerset(\ANNS) \rightarrow \CONTEXTS\). We also assume a set of function identifiers
-\(\FUNIDS\) and the existence of a \emph{code map} \(\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\)
-that relates each code address to its owning function and set of annotations.
-We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).  \bcp{I don't get the intuition for this whole para.}
-
-\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
-              {\(\mach, \context, \anns \constep \context'\)}
-              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
-                \mpcstate{\mach'}{\pol'}{\context'}\)}
-
-\subsection{Traces}
-\label{sec:traces}
-
-A trace over a type \(\TRACEELEMS\)\bcp{``over a type A'' is a Coq-ism} is a nonempty, finite or infinite sequence of
-\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
-\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
-called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
-
-\paragraph*{Trace-Of}
-
-The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
-coinductively relates an initial state with the trace of states
-produced by repeated application of step, and potentially cut short by
-the policy component. The {\sc Default} rule applies only if the other does not.
-(We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
-
-\begin{minipage}{.4\textwidth}
-\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
-            {\(\mpcstatename' \hookrightarrow \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
-\end{minipage}
-\begin{minipage}{.4\textwidth}
-\judgment[Default]
-         {}
-         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
-\end{minipage}
-%
-
-\paragraph*{Projections}
-
-We project out the machine state of an MPC-state with \(\pi_\mach\), the
-policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
-these operators to project out of an machine-trace a trace of machine states,
-policy states, and contexts, respectively.
-
-\paragraph*{Observations-Of}
-
-\bcp{Could any of this be postponed?}
-
-The ``observations-of'' operator, written \(\pi_\obs\), \apt{since this isn't
-  actually a projection, it is slightly confusing to use \(\pi\)} takes a machine trace and
-creates a trace of observations corresponding to the observations of each step in
-the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
-initial observation, a machine-trace, and an observation trace:
-
-\judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
-              {\(\mpcstatename' \hookrightarrow \MPCT\)}
-              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
-              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
-\judgment{}
-         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}
-
-Then to take the observations from any particular trace, we start with a \(\tau\) observation:
-
-\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
-         {\(\pi_\obs(\MPCT) = \obsT\)}
-\noindent
-
-In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
-to relate a machine-policy-context state triple to the MPCO-trace
-induced by the step function.\apt{?? This looks indistinguishable
-  from the \(\hookrightarrow\) we've already defined.}
-
-\paragraph*{Head and Last}
-
-We take the first element of a trace with \(\head(\Trace)\), which is a total
-function since traces are non-empty, and the final element (if one exists) with
-\(\last(\Trace)\), which is partial.
-
-\paragraph*{Until}
-The operation \(\PUT ~ f ~ \Trace\) takes a trace
-\(\Trace \in \TRACE{\TRACEELEMS \times \OBSS}\) and a predicate
-on elements \(f \subset \TRACEELEMS\) and gives the prefix of
-\(\Trace\) ending with the first element on which \(f\) holds.
-
-\begin{center}
-  \begin{minipage}{.3\textwidth}
-    \judgment{\(f ~ \traceelem\)}
-             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
-  \end{minipage}
-%
-  \begin{minipage}{.3\textwidth}
-    \judgment{}
-             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
-%
-  \end{minipage}
-  \begin{minipage}{.3\textwidth}
-    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
-                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
-  \end{minipage}
-\end{center}
-%
-If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate always holds, then
-\(\PUT\) is the identity function.
-
-% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
-%$\mathit{split}$: it takes two traces and links them by replacing the
-%last element of the first (if it has one) with the first element of
-%the second
-
-%\judgment{} {\(t \cdot T = T\)}
-%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
-%\leo{Lemma/theorem: split/join = id?}
-%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
-
-We will frequently take a prefix of the trace from an initial state up
-through the first state where some condition holds on the machine state.
-This can easily be implemented using \(\PUT\), and we provide special notation
-for convenience.
-%
-If \(f\) is a predicate on machine states, we define \(\mpcstatename
-\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:
-
-\begin{center}
-\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
-            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-\end{center}
-
-\paragraph*{Context Segments}
-
-In particular, we wish to formulate properties about consecutive subtraces
-in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''
-
-\begin{center}
-  \judgmenttwo[Now]
-              {\(f ~ \mach ~ \context\)}
-              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
-              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
-%
-  \judgmentthree[Skip]
-                {\(f ~ \mach ~ \context\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
-%
-  \judgmentthree[Search]
-                {\(\neg f ~ \mach ~ \context\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{center}
-
-\paragraph*{Reachability}
-
-We often want to quantify over all traces from any initial state so that our
-properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
-such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
-
-\paragraph*{Observational Similarity}
-
-We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-as far as an external observer is concerned, written \(\obsT_1 \lesssim
-\obsT_2\), if the sequence of
-non-silent observations of $\obsT_1$ is a prefix of those of
-$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
-coinductively:
-
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(\obsT \lesssim \obsT\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(\tau \lesssim \obsT\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(w \lesssim w\obsT\)}
-\end{minipage}
-
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(\tau \obsT_1 \lesssim \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(\obsT_1 \lesssim \tau \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(w\obsT_1 \lesssim w\obsT_2\)}
-\end{minipage}
-
-\smallskip
-We then define similarity of observation traces as traces prefixing each other:
-\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
-%
-%\leo{I still don't know how to do this transition. Maybe contrast
-%with the ``perhaps surprisingly'' later?} Note that
-%similarity of observation traces could alternatively be defined coinductively,
-%just like the prefix relation, but by dropping the asymmetric rules that allow
-%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}
-
-Note that an infinite silent trace is a
-prefix of (and similar to) any other trace. While this might seem
-surprising at first, it makes sense in a timing-insensitive context:
-an external observer looking at two machine runs cannot (computably)
-distinguish between a machine that steps forever and a machine that
-steps for a long time before producing some output.
-
-\section{Stack Safety, Formally}
-\label{sec:lse}
-
-We are finally ready for our definition of stack safety. We begin
-by describing how we model knowledge about the program structure,
-then dive into the formal definitions of local state encapsulation and
-well-bracketed control flow, which together constitute stack safety\bcp{not
-  any more?}.
-Initially we will focus on a simple stack model, with a single stack
-that doesn't share between caller and callee. All arguments are passed
-in registers.
-
-We begin with our only annotations being {\em call} or {\em return}, and
-define our code map accordingly.
-
-\[\ann \in \ANNS ::= \callmap + \retmap\]
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \ANNS\]
-
-When a function makes a call, it establishes that its frame should be
-``sealed'' for later use. At any given step of a program, any number of
-activations may have claimed memory in this way, but which cells are sealed
-is dynamic information and cannot be deduced from the state alone. So it must
-be tracked separately. We must also track the conditions under which a caller
-considers itself to have been returned to.
-
-Each system component will be assigned to a {\em domain}, which may be {\em Outside}
-of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth\bcp{What about code??}. Each caller will register a {\em target} condition -- a
-predicate on machine states that must hold for it to consider itself active.
-A context is a pair of a ``domain map'' from components to domains, and a
-``return target list,'' which is a stack of targets, one for each caller awaiting
-return.  \ifaftersubmission\bcp{spacing is funny}\fi
-
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-  \times (\rts \in \listT (\target \subseteq \MACHS)) \]
-
-An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
-the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
-
-We parameterize over the structure of the stack, particularly the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
-maps a machine state to a set of addresses.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
-which relates
-
-Now we can define the context update relation. Let \(\sealcon\) be the sealing convention
-that maps a machine state to a set of components, typically the set of addresses below
-the stack pointer. Similarly, \(\retcon ~ \mach ~ \mach'\) will be the return convention
-that relates the call state \(\mach\) to return states \(\mach'\) that are valid returns.
-The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
-and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-Then at each step we update the context according to the context-step function,
-\(\stepsto_C\).
-
-The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
-A call simply pushes the return target onto the return stack and seals all components
-with the depth of the previous return stack. A return
-recursively searches for a valid return target at the next step, and unseals
-every component sealed at a depth that it passes.
-
-\judgmentthree[Call]
-              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
-              {\(\components = \sealcon ~ \mach\)}
-              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\judgmenttwobr[ReturnFound]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
-              {\(\mach \stepsto \mach' \in \target\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\judgmenttwobr[ReturnRec]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\judgment[Default]
-         {}
-         {\(\_,\context \stepsto_C \context\)}
-
-When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We may now proceed to state our properties.
-Our first property will be {\em stepwise stack integrity}, stepwise meaning that it will be
-quantified over each step of execution from any initial state. The criterion for integrity
-is simple: sealed components do not change.
-
-\definition A system enjoys {\em stepwise stack integrity} if for any reachable state
-\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
-\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
-\pi_\mach(\mpcstatename') ~ \component\).
-
-Next we consider {\em stepwise stack confidentiality}. There are a number of preliminary
-concepts we must discuss here -- confidentiality quantifies over {\em variants} of the
-initial state of a call, and requires that the traces from them move in {\em lockstep}.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if, for
-all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\). We write
-this \(\mach \approx_\components \nach\).
-
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change together
-if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
-\(\nach ~ \component \not = \nach' ~ \component\), \(\mach' ~ \component = \nach' ~ \component\).
-This is written \(\mach,\mach' \diamond \nach,\nach'\).
-
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
-change together, written as a coinductive relation \(\machT \doteq
-\nachT\):\bcp{People may not be familiar with the notation}
-
-\begin{minipage}{.3\textwidth}
-\judgment{}
-         {\(\mach \doteq \nach\)}
-\end{minipage}
-\begin{minipage}{.6\textwidth}
-\judgmenttwo{\(\mach, \head(\machT) \diamond \nach, \head(\nachT)\)}
-            {\(\machT \doteq \nachT\)}
-            {\(\mach \machT \doteq \nach \nachT\)}
-\end{minipage}
-
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
-\(\MPCT\) corresponds to the trace of a subroutine from its call to its return.
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
-\(\MPCT \doteq \NPCT\).
-
-\subsection{Stack Safety with Passing}
 

### 2021-04-15 tex/main.tex

@@
 \end{subfigure}
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.7\textwidth}
   {\small
@@
 2        call f
-3        mov %rax,-1(%sp) ; store x
+3        mov %ra,-1(%sp) ; store x
 4        call g
@@
 11       mov $5,-1(%sp)  ; initialize y
-12       mov -1(%sp),%rax ; set return value
+12       mov -1(%sp),%ra ; set return value
 13       sub $1,%sp      ; deallocate frame
@@
 22       brneq #25       ; branch if not equal
-23       mov $10,%rax ; set return value
+23       mov $10,%ra     ; set return value
 24       jmp #26
-25       mov $20,%rax  ; set return value
+25       mov $20,%ra     ; set return value
 26       sub $1,%sp      ; deallocate frame
@@
   \\
-  {\tt 21} &
+  {\tt 22} &
   \memoryaddrs{16em}
@@
   \\ &
-  {\bf 22}
+  {\bf 23}
   ~$\cdots$
   &
-  {\bf 24}
+  {\bf 25}
   ~$\cdots$
@@
   \centering
-  \begin{subfigure}{.4\textwidth}
+  \begin{subfigure}{.3\textwidth}
+{\small
 \begin{verbatim}
@@
 int main() {
-0:     int x;
-1:     x = 1;
-2[C]:  f();
-3:     x = -x;
-4[C]:  f();
-5[R]:  return x;
+  int x = 1;
+  f();
+  x = -x;
+  f();
+  return x;
 }
@@
 void f() {
-10:     if(!stash) {
-        // store return pointer
-11:       stash = $SP-1;
-12:     } else {
-13:       *($SP-1) = stash;
-14:       stash = 0;
+  int y;
+  if (!stash) {
+    stash = *(&y-1);
+  } else {
+    *(&y-1) = stash;
+    stash = 0;
   }
-15[R]:    return;
+  return;
 }
 \end{verbatim}
+}
 \end{subfigure}
-  \begin{subfigure}{.5\textwidth}
+  \begin{subfigure}{.65\textwidth}
+{\small
+\begin{verbatim}
+   stash: .word 0
+1  main:  add $1,%sp        ; allocate frame
+2         mov $1,-1(%sp)    ; initialize x
+3         call f
+4         neg -1(%sp)       ; x = -x 
+5         call f           
+6         mov -1(%sp),%ra   ; set return value
+7         sub $1,%sp        ; deallocate frame
+6         ret
+
+10 f:     add $1,%sp        ; allocate frame
+11        cmp stash, $0     ; compare
+12        breq #15          ; branch if equal
+13        mov -2(%sp),stash ; update stash
+14        jmp #17
+15        mov stash,-2(%sp) ; retrieve stash
+16        mov $0,stash
+17        sub $1,%sp        ; deallocate frame
+18        ret 
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}{.4\textwidth}
     \begin{tabular}{l l l}
-      {\tt 2} &
+      {\tt 3} &
       \memoryaddrs{8em}
-      \memory{4}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-19em}{0.75em}{1} \\
-      {\tt 15} &
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-15em}{0.75em}{1} \\
+      {\tt 18} &
       \memoryaddrs{12em}
@@
       \memory{1}{\unsealc}[RetPtr]%
-      \memory{2}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-19em}{0.75em}{1}
-      \MemoryLabel{-15em}{0.75em}{\#3}
+      \memory{1}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-15em}{0.75em}{1}
+      \MemoryLabel{-11em}{0.75em}{\#4}
       \vspace{.5em} &
       \memory[3em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#3}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#4}
       \\
-      {\tt 3} &
+      {\tt 4} &
       \memoryaddrs{8em}
-      \memory{4}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-19em}{0.75em}{1} \\
-      {\tt 10} &
-      \memoryaddrs{12em}
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-15em}{0.75em}{1} \\
+      {\tt 11} &
+      \memoryaddrs{16em}
       \memory{1}{\mainsealc}
-      \memory{3}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-19em}{0.75em}{-1}
-      \MemoryLabel{-15em}{0.75em}{\#5} &
+      \memory{2}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-15em}{0.75em}{-1}
+      \MemoryLabel{-11em}{0.75em}{\#6} &
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
       \\
-      {\tt 15} &
+      {\tt 18} &
       \memoryaddrs{12em}
@@
       \memory{1}{\badc}
-      \memory{2}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-19em}{0.75em}{-1}
-      \MemoryLabel{-15em}{0.75em}{\#3} &
+      \memory{1}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-15em}{0.75em}{-1}
+      \MemoryLabel{-11em}{0.75em}{\#4} &
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
-      \\
-      {\tt 3} &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-19em}{0.75em}{-1} &
-      \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
       \\
@@
       \memory{1}{\mainsealc}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-19em}{0.75em}{\bf 1} &
+      \memory{2}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-15em}{0.75em}{-1} &
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5} \\
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
+      \\
+      {\tt 5} &
+      \memoryaddrs{8em}
+      \memory{1}{\mainsealc}
+      \memory{2}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-15em}{0.75em}{\bf 1} &
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6} \\
     \end{tabular}
@@
 Here we show return addresses explicitly as line numbers, in the form $\#n$.
-Function {\tt f} is called twice, at lines 2 and 4; it should therefore return to lines
+Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
 3 and 5, respectively. But the first time, before returning, it stashes its
-return pointer away, and the second time it returns to line 3 instead of 5; then it ends
-up being called a third time, at which point it returns to line 5.
-
-Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that we return
--1. But because of {\tt f}'s interference, we negate {\tt x} twice, and return 1.
+return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
+up being called a third time, at which point it finally returns to instruction 5.
+
+Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that {\tt main} returns
+-1. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
 This has the appearance of an integrity violation.  How do we capture this in our
@@
 appropriate {\em return target} in {\tt main}. A return target is a predicate on states,
-in this case one that holds on any state whose program counter points to the correct line.
+in this case one that holds on any state whose program counter points to the correct instruction.
 We introduce a new piece of context: we keep a stack of return targets, and a callee is
@@
 In the example, on its first return, {\tt f} reaches its return target, because the
-\(\PCname\) is 3 after its return. But on the second, {\tt f} does not properly return,
-but rather changes its stack pointer (shown in light red) and jumps into
-{\tt main}'s code. Then {\tt f}
-attempts to overwrite the (still sealed) value of {\tt x}, violating integrity.
+\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
+(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
+Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
+violates the integrity property.
 

### 2021-04-15 tex/main.tex
Cleaned up eager properties section

--- 2021-04-15 tex/main.tex
@@
 that doesn't share between caller and callee. All arguments are passed
-in registers. Recall that we keep track of which code addresses are
+in registers. Recall that we keep track of which addresses are
 annotated as calls, with a ``code map'':
 \[\codemap \in \CODEMAPS \subseteq \WORDS\]
-When a function makes a call, it establishes that its frame should be
-``sealed'' for later use. At any given step of a program, any number of
-activations may have claimed memory in this way, but which cells are sealed
-is dynamic information and cannot be deduced from the state alone. So it must
-be tracked separately. We must also track the conditions under which a caller
-considers itself to have been returned to.
-
-Each system component will be assigned to a {\em domain}, which may be {\em Outside}
-of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller will register a {\em target}
-condition -- a predicate on machine states that must hold for it to consider itself active.
-A context is a pair of a ``domain map'' from components to domains, and a
-``return target list,'' which is a stack of targets, one for each caller awaiting
-return.  \ifaftersubmission\bcp{spacing is funny}\fi
-
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-  \times (\rts \in \listT (\target \subseteq \MACHS)) \]
-
-An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
-the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
 
@@
 
+Each system component will be assigned to a {\em domain}, which may be {\em Outside}
+of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
+at a particular depth. Each caller will register a {\em target}
+condition -- a predicate on machine states that must hold for it to consider itself active.
+A context is a pair of a ``domain map'' from components to domains, and a
+``return target list,'' which is a stack of targets, one for each caller awaiting
+return.  \ifaftersubmission\bcp{spacing is funny}\fi
+
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+  \times (\rts \in \listT (\target \subseteq \MACHS)) \]
+
+An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
+the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
+
 The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
@@
 
-\judgmenttwobr[ReturnRec]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
-
 \judgment[Default]
@@
 
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change together
-if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
-\(\nach ~ \component \not = \nach' ~ \component\), \(\mach' ~ \component = \nach' ~ \component\).
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change
+together if for all components \(\component\) such that
+\(\mach ~ \component \not = \mach' ~ \component\) or
+\(\nach ~ \component \not = \nach' ~ \component\),
+\(\mach' ~ \component = \nach' ~ \component\).
 This is written \(\mach,\mach' \diamond \nach,\nach'\).
 
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
-change together, written as a coinductive relation \(\machT \doteq
-\nachT\):\bcp{People may not be familiar with the notation}
+\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
+if their adjacent pairs change together. We will represent this relation with
+\(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
+\sna{It's a new notation, we use this enough that it seems worth picking an
+  evocative symbol.} 
+We define it as a coinductive relation:
 
@@
 
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap\]
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
 
@@
   \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth,\depth} \\
+  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
   \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
@@
 
-\judgmentthree[ReturnFound]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+\judgmenttwo[ReturnFound]
             {\(\mach \stepsto \mach' \in \target\)}
@@
             {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\judgmentthree[ReturnRec]
-              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
 pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We extend the set of possible annotations with a yield annotation.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap + \yieldmap\]
+We add a second annotation indicating yields.
+
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
 
@@
 if the next state is a valid yield target for that stack, and we update the yield target for
-the yielding stack. So it is once again possible for a coroutine to claim to yield but not
-actually change the active stack, with similar effects to when a callee fails to return properly.
-Now we can define both stack safety and coroutine safety under this model.
+the current stack, ensuring that when the system yields back we expect the current coroutine
+to continue from where it left off. So, it is possible for a coroutine to claim to yield
+but not actually change the active stack, with similar effects to when a callee fails
+to return properly. Now we can define both stack safety and coroutine safety under this model.
 
@@
 Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
 where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
@@
 \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
+\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
 where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
@@
 
-\subsection{Well-bracketed Control Flow}
-\label{sec:wbcf}
-
-Both the integrity and confidentiality components of stack safety
-are reasonable even in programs that violate expected
-control flow. But even though stack safety prevents control-flow violations
-from being used to gain privileges, it is worth protecting control flow as well.
-Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
-control flow.
-
-When we focus on subroutines, the essence of well-bracketed control flow
-is that a caller can expect that if its callee ever returns, it returns
-to the caller at the appropriate return target. This is tantamount to saying
-that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
-the simple model:
-
-\definition A system enjoys {\em well-bracketed control flow} if for all
-reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-\(\mpcstatename \stepstocon \mpcstatename'\), if
-\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
-is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
-
-Extending this idea to coroutines, we use the same notion that there are
-no ``false'' yields.
-
-\definition A system enjoys {\em well-structured control flow} if for all
-reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-\(\mpcstatename \stepstocon \mpcstatename'\), if
-\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
-is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
-where \(\stackid \not = \stackid'\).
+%\subsection{Well-bracketed Control Flow}
+%\label{sec:wbcf}
+
+%Both the integrity and confidentiality components of stack safety
+%are reasonable even in programs that violate expected
+%control flow. But even though stack safety prevents control-flow violations
+%from being used to gain privileges, it is worth protecting control flow as well.
+%Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
+%control flow.
+
+%When we focus on subroutines, the essence of well-bracketed control flow
+%is that a caller can expect that if its callee ever returns, it returns
+%to the caller at the appropriate return target. This is tantamount to saying
+%that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
+%the simple model:
+
+%\definition A system enjoys {\em well-bracketed control flow} if for all
+%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
+%\(\mpcstatename \stepstocon \mpcstatename'\), if
+%\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
+%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
+
+%Extending this idea to coroutines, we use the same notion that there are
+%no ``false'' yields.
+
+%\definition A system enjoys {\em well-structured control flow} if for all
+%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
+%\(\mpcstatename \stepstocon \mpcstatename'\), if
+%\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
+%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
+%where \(\stackid \not = \stackid'\).
 
@@
 
-\subsection{Observable Integrity}
-
-We will describe observable integrity in our most detailed setting. In the
-presence
+\subsection{Lazy Stepwise Integrity}
 

### 2021-04-15 tex/main.tex
redo stack sharing example

--- 2021-04-15 tex/main.tex
@@
   \centering
-  \begin{subfigure}{.3\textwidth}
+  \begin{subfigure}{.2\textwidth}
 \begin{verbatim}
-int main() {
-0:          int x;
-1:          x = 5;
-2[C($SP)]:  x = f(x);
-3[R]:       return x;
+void main() {
+  int x = 5;
+  x = f(x);
+  return;
 }
 
-int f(int a) {
-10:    int y;
-11:    y = 2*a;
-12:    a = 0;
-13[R]: return y;
+void f(int a) {
+  int y = 2*a;
+  a = 0;
+  return;
 }
+\end{verbatim}
+  \end{subfigure}
+  \begin{subfigure}{.6\textwidth}
+\begin{verbatim}
+1  main: add $1,%sp          ; allocate frame
+2        mov $5,-1(%sp)      ; initalize x
+3        add $1,%sp          ; allocate arg
+4        mov -2(%sp),-1(%sp) ; set arg
+5        call f              ; SHARE: %sp-1
+6        sub $1,%sp          ; deallocate arg
+7        mov %ra,-1(%sp)     ; store x
+8        sub $1,%sp          ; deallocate frame
+9        ret
+
+10  f:   add $1,%sp          ; allocate frame
+11       mov -3(%sp),-1(%sp) ; y = a
+12       mul $2,-1(%sp)      ; y = a*2
+13       mov $0,-3(%sp)      ; a = 0
+14       sub $1,%sp          ; deallocate frame
+15       ret
 \end{verbatim}
@@
 \begin{tabular}{l r | l}
-  {\tt 2} &
+  {\tt 3} &
   \multicolumn{2}{c}{
@@
     \MemoryLabel{-19em}{0.75em}{5}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{0}
-    \MemoryLabel{-7em}{0.75em}{0}
+%    \MemoryLabel{-14em}{0.75em}{0}
+%    \MemoryLabel{-10em}{0.75em}{0}
+%    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
@@
   {\tt 11} &
-  \memoryaddrs[(1)]{20em}
+  \memoryaddrs[(1)]{21.5em}
   \memory{1}{\mainsealc}[Seal(0)]%
@@
   \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{0}
+%  \MemoryLabel{-6em}{0.75em}{0}
   &
-  \memoryaddrs[(1)]{20em}
+  \memoryaddrs[(1)]{21.5em}
   \memory{1}{\mainsealc}
@@
   \\
-  {\tt 12} &
-  \memoryaddrs[(1)]{20em}
+  {\tt 13} &
+  \memoryaddrs[(1)]{21.5em}
   \memory{1}{\mainsealc}
@@
   &
-  \memoryaddrs[(1)]{20em}
+  \memoryaddrs[(1)]{21.5em}
   \memory{1}{\mainsealc}
@@
   \\
-  {\tt 13} &
-  \memoryaddrs[(1)]{20em}
+  {\tt 14} &
+  \memoryaddrs[(1)]{21.5em}
   \memory{1}{\mainsealc}
@@
   \\
-  {\tt 3} &
+  \\
+  {\tt 8} &
   \multicolumn{2}{c}{
@@
     \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{3}
+    \MemoryLabel{-10em}{0.75em}{\#6}
     \MemoryLabel{-7em}{0.75em}{10}
@@
 its value copied from {\tt x}, and we note that {\tt f} can read it -- because it is marked
-as passed between depths 0 and 1, it is not varied. Then from line 12 to 13 it is even
+as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
 modified, and naturally this is permitted as well.

### 2021-04-15 tex/main.tex
redid pass-by-ref example

@@
   \centering
-  \begin{subfigure}{.45\textwidth}
+  \begin{subfigure}{.2\textwidth}
+{\small
 \begin{verbatim}
@@
 int main() {
-0:         int x, y;
-1[C(x)]:  f(x);
-2:         x = 0;
-3[C(y)]:  f(y);
-4[R]:      return x;
+  int x, y;
+  f(x);
+  x = 0;
+  f(y);
+  return x;
 }
@@
 void f(int &a) {
-10:    if(!stash)
-11:      stash = a;
-12:    *stash = 5;
-13[R]: return;
+  if(!stash)
+    stash = &a; 
+  *stash = 5;
+  return;
 }
 \end{verbatim}
+}
   \end{subfigure}
-  \begin{subfigure}{.5\textwidth}
+  \begin{subfigure}{.49\textwidth}
+{\small
+\begin{verbatim}
+   stash: .word 0
+1  main:  add $2,%sp      ; allocate frame
+2         mov %sp,%ra
+3         add -2,%ra      ; ra = &x
+4         call f          ; SHARE: %sp-2
+5         mov $0,-2(%sp)  ; x = 0
+6         mov %sp,%ra
+7         add -1,%ra      ; ra = &y
+8         call f          ; SHARE: %sp-1
+9         mov -2(%sp),%ra ; set return value
+10        ret
+
+20 f:     cmp stash, $0   ; compare
+21        breq #13        ; branch if equal
+22        mov %ra,stash   ; stash = &a
+23        jmp #15
+24        mov $5, (stash) ; *stash = 5
+25        ret 
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}{.29\textwidth}
     \begin{tabular}{l l}
-      {\tt 10} &
-      \memoryaddrs[(0)]{17em}
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
       \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
@@
       \memory{1}{\retptrc}
-      \memory{1}{\unsealc}
+%      \memory{1}{\unsealc}
       ~$\cdots$ \\
-      {\tt 13} &
+      {\tt 25} &
       \memoryaddrs[(1)]{17em}
@@
       \memory{1}{\retptrc}
-      \memory{1}{\unsealc}
+%      \memory{1}{\unsealc}
       ~$\cdots$
@@
       \\
-      {\tt 3} &
+      {\tt 8} &
       \memoryaddrs[(0)]{12em}
-      \memory{4}{\unsealc}
+      \memory{3}{\unsealc}
       ~$\cdots$
@@
       \\
-      {\tt 10} &
-      \memoryaddrs[(0)]{17em}
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
       \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
@@
       \\
-      {\tt 13} &
-      \memoryaddrs[(0)]{17em}
+      {\tt 25} &
+      \memoryaddrs[(1)]{17em}
       \memory{1}{\mainsealc}
@@
 \end{subfigure}
-\caption{A violation with pass-by-reference}
+\caption{A violation with pass-by-reference\apt{SP(n) depths looked wrong; SNA double-check that I fixed them right.}}
 \label{fig:passing}
@@
 variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example, where we revert to assuming
-that arguments are passed in registers,
-{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
-passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
+as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
+{\tt f} takes a single argument by reference.
+{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
+{\tt f} again passing a reference to {\tt y}.
+We would expect that the first call to {\tt f} might modify
 {\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call.
-
-So at the initial call, {\tt x} is passed, and {\tt f} is free to write to it. Then upon
-return, {\tt main}'s frame is unsealed. At the second call it is {\tt y} that is passed,
+{\tt x} in the second call. (Incidentally, this is perfectly compilable C++!)
+
+To avoid cluttering the picture at the assembly level, we revert to assuming that the
+argument {\tt a} is passed in a register, namely {\tt \%ra}.
+The question is how the local variables {\tt x} and {\tt y} themselves
+should be labelled in order for the policy to detect an integrity violation.
+At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
+return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
 and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
@@
 an unbroken chain of calls. So it suffices to track two depths for each passed
-cell: the original owner, and the highest depth it has been passed to. It is considered
-accessible anywhere between the two.
+cell: the original owner, and the highest depth it has been passed to. The policy considers
+it to be accessible anywhere between the two.
 

### 2021-04-15 tex/main.tex
add missing \small in stack sharing example
add missing deallocate frame instr in pass-by-reference example

@@
   \begin{subfigure}{.2\textwidth}
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
   \end{subfigure}
   \begin{subfigure}{.6\textwidth}
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
   \end{subfigure}
@@
 9         mov -2(%sp),%ra ; set return value
-10        ret
+10        sub $2,%sp      ; deallocate frame
+11        ret
 

### 2021-04-15 tex/main.tex
Separated out observations and moved them down to their own section, wrote up properties

--- 2021-04-15 tex/main.tex
@@
 
-\section{Machines, Traces, and Observations}
+\section{Machines and Traces}
 \label{sec:prelim}
@@
 
-\subsection{Values, States, and Observations}
+\subsection{Values and States}
 
@@
 Our machine states are composed of {\em components} \(\component\),
-addressed\bcp{no, the components are not addressed, they are addresses!} either by
-words (memory addresses) or by register names (\(\reg\)) drawn from some set
+which are either words (memory addresses) or by register names (\(\reg\)) drawn from some set
 \(\REGS\).
 %
-    \[\component \in \COMPONENTS = \WORDS + \REGS \]
+    \[\component \in \COMPONENTS ::= \WORDS + \REGS \]
 %
@@
 %
-\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]
-
-The step function of the machine takes a machine state and produces a
-pair of a machine state and an {\em observation}, which may be either
-silent (written \(\tau\)) or a word of data that the machine
-communicates to the outside world:
-\ifaftersubmission\bcp{This is a rather concrete
-  choice!  Why not leave the set of observations abstract?}\fi
-%
-\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
-\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
-%
-This step function is total, so the machine's execution traces
-(defined in \cref{sec:traces}) are infinite.
-
-Observations are only essential for describing the extensional properties
-described in Section~\ref{sec:lazy}, but for simplicity, we include them
-in the machine model from the beginning.
-We choose to model observations explicitly, rather than simply treating
-a subset of registers or memory as being observable.  This choice gives
-us the flexibility to treat programs as observationally equivalent even when
-their internal states may differ.
-%The choice of modelling observations is an interesting one. A first approach
-%would be to consider any change to the registers or memory as observable.
-%However, as some parts of the stack will be considered ``private''
-%this choice won't work. A relaxation of this to consider all changes
-%to ``non-private'' data as observable could work in principle, but
-%does not allow for an interesting weaker-but-more-efficient enforcement
-%mechanism. We will come back to this point in \Cref{sec:lazy}.
-
-\apt{This para is out of date. Do we want to say something equivalent
-  about new examples (hard, because they are C-based).}
-$\WORDS$ is the
-set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
-\cup \{\PCname,\SP\}$. Observations are writes to the
-special $\rout$ location\ifaftersubmission\bcp{Which is 100! (do we need a name for it? seems
-  not...)}\fi, which can be thought of as a memory-mapped
-  output port: if during an execution step a word $w$ is
-written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.
+\[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
+
+We have a total step function between machine states \(\mach \stepsto \mach'\).
 
@@
 
-\judgmenttwo{\(\mach \stepstoobs[\obs] \mach'\)}{\((\mach, \pol)
+\judgmenttwo{\(\mach \stepstoobs \mach'\)}{\((\mach, \pol)
                \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol[\obs]
+            {\(\mpstate{\mach}{\pol} \stepstopol
                \mpstate{\mach'}{\pol'}\)}
@@
 
-\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
+\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
               {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
               {\(\mach, \context, \anns \constep \context'\)}
-              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
+              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
                 \mpcstate{\mach'}{\pol'}{\context'}\)}
@@
 
-\paragraph*{Observations-Of}
-
-\bcp{Could any of this be postponed?}
-
-The ``observations-of'' operator, written \(\pi_\obs\), \apt{since this isn't
-  actually a projection, it is slightly confusing to use \(\pi\)} takes a machine trace and
-creates a trace of observations corresponding to the observations of each step in
-the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
-initial observation, a machine-trace, and an observation trace:
-
-\judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
-              {\(\mpcstatename' \hookrightarrow \MPCT\)}
-              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
-              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
-\judgment{}
-         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}
-
-Then to take the observations from any particular trace, we start with a \(\tau\) observation:
-
-\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
-         {\(\pi_\obs(\MPCT) = \obsT\)}
-\noindent
-
-In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
-to relate a machine-policy-context state triple to the MPCO-trace
-induced by the step function.\apt{?? This looks indistinguishable
-  from the \(\hookrightarrow\) we've already defined.}
-
 \paragraph*{Head and Last}
@@
 such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
-
-\paragraph*{Observational Similarity}
-
-We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-as far as an external observer is concerned, written \(\obsT_1 \lesssim
-\obsT_2\), if the sequence of
-non-silent observations of $\obsT_1$ is a prefix of those of
-$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
-coinductively:
-
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(\obsT \lesssim \obsT\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(\tau \lesssim \obsT\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(w \lesssim w\obsT\)}
-\end{minipage}
-
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(\tau \obsT_1 \lesssim \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(\obsT_1 \lesssim \tau \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(w\obsT_1 \lesssim w\obsT_2\)}
-\end{minipage}
-
-\smallskip
-We then define similarity of observation traces as traces prefixing each other:
-\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
-%
-%\leo{I still don't know how to do this transition. Maybe contrast
-%with the ``perhaps surprisingly'' later?} Note that
-%similarity of observation traces could alternatively be defined coinductively,
-%just like the prefix relation, but by dropping the asymmetric rules that allow
-%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}
-
-Note that an infinite silent trace is a
-prefix of (and similar to) any other trace. While this might seem
-surprising at first, it makes sense in a timing-insensitive context:
-an external observer looking at two machine runs cannot (computably)
-distinguish between a machine that steps forever and a machine that
-steps for a long time before producing some output.
 
@@
 \citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
-mitigate those costs, they propose optimizations called Lazy Tagging and
-Lazy Clearing that relax and defer some of the checks of the Depth Isolation
-policy. Under these optimizations, the policy does not initialize stack
-frames on entry or clear them on exit (thus speeding up execution),
+mitigate those costs, they propose a number of optimizations.
+
+First, they eliminate the initialization pass. In this {\em Lazy Tagging}
+system, the stack is initially tagged as ``clear.'' A function is permitted
+to overwrite clear memory, and when it does so its depth tag overwrites
+the clear tag, claiming the memory; the remainder of the policy functions
+as {\em Depth Isolation}. On return, the callee must clear its frame, which
+is still a heavy cost.
+
+To eliminate this cost, \citet{DBLP:conf/sp/RoesslerD18} introduce
+{\em Lazy Tagging and Lazy Clearing}. Under this optimizations,
+the policy does not initialize stack frames on entry or clear them on exit,
 and it permits all writes to the stack, even when the $\PCname$ tag
-does not match the memory tag.
-However, reads from the stack do still require $\PCname$ tag and memory tag
-to match. Thus, even if a callee illicitly writes to a private location
-in its caller, the caller will eventually detect this if it ever tries to read from
-that location.
+does not match the memory tag. Instead it propagates the \(\PCname\) tag
+to written memory. Then reads from the stack do still require $\PCname$
+tag and memory tag to match. Thus, even if a callee illicitly writes to
+a private location in its caller, the caller will eventually detect this
+if it ever tries to read from that location.
 %
@@
 How does a harmful violation that must be caught later differ from a
-harmless one? The properties in this section rely on our notion of
-observations to draw this distinction.
-
-\subsection{Lazy Stepwise Integrity}
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\sealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(\depth' < |\rtm ~ \stackid|\) .
-Then a system enjoys observable
-
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\subsection{Observable Confidentiality}
-
-    \paragraph{Connection to Policies}
+harmless one?
+
+\paragraph*{Deferred Stepwise Integrity}
+
+We first notice that, in principle, these properties appear to enforce
+confidentiality outright. So we wish in particular to find an integrity
+property. The intuition is that sealed memory can change, but it is
+then tainted, and its original owner cannot actually read it. This
+leads us to a hybrid of integrity and confidentiality.
+
+For simplicity, our definitions return to the simple stack without sharing,
+but all properties in this section can be extended to more complex settings
+very naturally.
+
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \sealed{\_}\).
+
+Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\) -- meaning that the
+call does return -- let \(\components'\) be the set of components that differ
+between \(\mach\) and \(\mach'\). Take any \(\nach\) such that
+\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
+\(\nach \hookrightarrow \NPCT\). A system enjoys
+{\em deferred stepwise integrity} if for all such \(\MPCT\) and \(\NPCT\),
+\(\MPCT \doteq \NPCT\).
+
+In short, if sealed components are modified during the call, the caller may not
+access their values. ``No reads after dangerous writes.''
+
+\subsection{Observability}
+
+We argue that while deferred stepwise integrity meaningfully captures the intuition
+behind lazy tagging, it stops short of the kind of extensional characterization
+that would be viable as a weakest reasonable definition. Consider a hypothetical
+taint-tracking policy, that allows dangerous writes to be copied as long as the
+taint remains hidden from outside observers. In order to characterize
+such a policy, we would need a model of observation.
+Similarly, for such a taint-based policy, lockstep confidentiality would be too
+strong, as we might permit secrets to move about as long as they were not
+leaked to the outside world.
+
+So we extend our model with a notion of observations. We abstract over an
+observation type \(\obs \in \OBSS\), containing at minimum the silent observation
+\(\tau\). Our machine step function now takes a state and returns a state and an observation,
+which is carried over into the policy and context step functions.
+%
+\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
+%
+
+\paragraph*{Observations-Of}
+
+The ``observations-of'' operator, written \(\obsof\), takes a machine trace and
+creates a trace of observations corresponding to the observations of each step in
+the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
+initial observation, a machine-trace, and an observation trace:
+
+\judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
+              {\(\mpcstatename' \hookrightarrow \MPCT\)}
+              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
+              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
+\judgment{}
+         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}
+
+Then to take the observations from any particular trace, we start with a \(\tau\) observation:
+
+\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
+         {\(\obsof(\MPCT) = \obsT\)}
+
+\paragraph*{Observational Similarity}
+
+We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
+as far as an external observer is concerned, written \(\obsT_1 \lesssim
+\obsT_2\), if the sequence of
+non-silent observations of $\obsT_1$ is a prefix of those of
+$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
+coinductively:
+
+\begin{minipage}{.3\textwidth}
+  \judgment{}{\(\obsT \lesssim \obsT\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{}{\(\tau \lesssim \obsT\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{}{\(w \lesssim w\obsT\)}
+\end{minipage}
+
+\begin{minipage}{.3\textwidth}
+  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
+           {\(\tau \obsT_1 \lesssim \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
+           {\(\obsT_1 \lesssim \tau \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
+           {\(w\obsT_1 \lesssim w\obsT_2\)}
+\end{minipage}
+
+\smallskip
+We then define similarity of observation traces as traces prefixing each other:
+\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
+%
+%\leo{I still don't know how to do this transition. Maybe contrast
+%with the ``perhaps surprisingly'' later?} Note that
+%similarity of observation traces could alternatively be defined coinductively,
+%just like the prefix relation, but by dropping the asymmetric rules that allow
+%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}
+
+Note that an infinite silent trace is a
+prefix of (and similar to) any other trace. While this might seem
+surprising at first, it makes sense in a timing-insensitive context:
+an external observer looking at two machine runs cannot (computably)
+distinguish between a machine that steps forever and a machine that
+steps for a long time before producing some output.
+
+\subsection{Observational Stack Safety}
+
+We now can sketch observational versions of both integrity and confidentiality.
+Integrity is already straightforward, simply weakening the condition that must hold
+after the return.
+
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \sealed{\_}\).
+
+Then, again supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
+let \(\components'\) be the set of components that differ
+between \(\mach\) and \(\mach'\). Take any \(\nach\) such that
+\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
+\(\nach \hookrightarrow \NPCT\). A system enjoys
+{\em observational integrity} if for all such \(\MPCT\) and \(\NPCT\),
+\(\obsof(\MPCT) \simeq \obsof(\NPCT)\).
+
+We amend our intuitive statement to: ``No visible reads after dangerous writes.''
+
+Confidentiality is more sophisticated. We must capture the intuition that
+secrets do not escape during a call {\em and} are not leaked during a return.
+This require us to be able to quantify over components that preserve our \(\diamond\)
+relation from earlier.
+
+\definition Let \(\components\) be a set of components and \(\mach,\mach'\) and \(\nach,\nach'\)
+be pairs of states. The {\em unsafe set} of \(\components\), written
+\(\bar{\Diamond}(\components,\mach,\mach',\nach,\nach')\), is the set of all components
+\(\component \in \components\) such that if
+\(\mach ~ \component \not = \mach' ~ \component\) or
+\(\nach ~ \component \not = \nach' ~ \component\), then
+\(\mach' ~ \component = \nach' ~ \component\).
+
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
+Let \(\nach\) be a variant such that \(\mach \approx_\components \nach\) and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\).
+
+A system enjoys {\em observational confidentiality} if for all such
+\(\MPCT\) and \(\NPCT\), the following conditions all hold:
+
+\begin{itemize}
+\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
+  meaning that no secrets leak during the call
+\item \(f_\depth ~ \last(\MPCT) \leftrightarrow f_\depth ~ \last(\NPCT)\) --
+  if one call returns, so does the other
+\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
+  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
+  Let \(\components' = \bar{\Diamond}(\components,\mach,\mach',\nach,\nach'\).
+  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
+  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
+  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
+  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
+\end{itemize}
+
+This last condition is the most interesting: we identify the components that
+contain leaked information on either side of the first pair of variant traces.
+This becomes the basis for the second variation, because all of these values
+must be protected from leaking for the remainder of the program.
+
+\sna{More to write here. The gist is: this is the extensional property.
+  The one that we would tentatively propose as a real characterization of stack
+  safety.}
+
+\subsection{Connection to Policies}
 

### 2021-04-15 tex/main.tex

@@
 void f(int &a) {
-10:    if(!stash)
-11:      stash = a;
-12:    *stash = 5;
-13[R]: return;
+  if(!stash)
+    stash = &a; 
+  *stash = 5;
+  return;
 }
 \end{verbatim}
+}
   \end{subfigure}
-  \begin{subfigure}{.5\textwidth}
+  \begin{subfigure}{.49\textwidth}
+{\small
+\begin{verbatim}
+   stash: .word 0
+1  main:  add $2,%sp      ; allocate frame
+2         mov %sp,%ra
+3         add -2,%ra      ; ra = &x
+4         call f          ; SHARE: %sp-2
+5         mov $0,-2(%sp)  ; x = 0
+6         mov %sp,%ra
+7         add -1,%ra      ; ra = &y
+8         call f          ; SHARE: %sp-1
+9         mov -2(%sp),%ra ; set return value
+10        sub $2,%sp      ; deallocate frame
+11        ret
+
+20 f:     cmp stash, $0   ; compare
+21        breq #13        ; branch if equal
+22        mov %ra,stash   ; stash = &a
+23        jmp #15
+24        mov $5, (stash) ; *stash = 5
+25        ret 
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}{.29\textwidth}
     \begin{tabular}{l l}
-      {\tt 10} &
-      \memoryaddrs[(0)]{17em}
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
       \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%

### 2021-04-15 tex/main.tex
Nits

--- 2021-04-15 tex/main.tex
@@
 \newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}
-\newcommand*{\rsp}{\textsc{sp}}
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
 
@@
 definitions are strongly informed by micropolicy systems, especially
-our attacker model, which is stronger than other enforcement mechanisms
-can enable.
+our attacker model. \BCP{I am still not clear in what way, exactly.}
 
@@
 from its callee, even as that callee may need protection from its callee in turn.
-This is novel in that the
-confidentiality property applies to many nested subtraces within a single execution,
-as opposed to traces of entire executions.
+This differs from common practice in language-based security in that the
+confidentiality property applies to many nested subtraces within a single execution.\sna{How novel is this nesting?}\bcp{I believe completely novel?}
 {\em Integrity} says that a caller's local data is protected from modification by its callee,
@@
 
-A second challenge is that enforcement mechanisms may weaken their protections
-in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
-\citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
-ways of identifying functions and deciding exactly which accesses are allowed.
+A second challenge is that
+enforcement mechanisms may weaken their protections in the name of efficiency,
+while arguing that they still capture stack safety in a meaningful sense.
+For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
+using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
+Memory cells within each stack frame are tagged with the
+identity of the function activation
+that owns the frame\BCP{Really??}, and access to those locations is then gated
+according to the activation that is currently executing. Roessler and DeHon
+explore alternative versions of this basic idea, with different ways of identifying
+functions and deciding exactly which accesses are allowed.
 These vary in their efficiency and in the spatial and temporal precision of their
 security guarantees.
-The Depth Isolation policy prevents a callee from writing outside its frame
+The \emph{Depth Isolation} policy prevents a callee from writing outside its frame
 at all, which is simple and intuitively correct, but it must manually clear the tags
-on the frame on each return, which is very expensive. Conversely, their
+on the frame on each return, which is very expensive. Conversely, the
 \emph{Lazy Tagging and Clearing} policy improves
@@
 
+\BCP{This paragraph doesn't apply to the lazy properties, so maybe it should
+be moved before they are introduced?}%
+To illustrate the applicability of these properties, we perform randomized testing to
+validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
+them. Randomized testing is a relatively low-cost way to validate a potential enforcement
+mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
+of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
+covers potentially erroneous programs, we test intentionally buggy policies alongside
+the real one. If the testing can differentiate correct implementations from buggy ones,
+it is an argument in favor of their correctness. \bcp{Hmm--don't understand
+  that sentence}
 %\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
@@
 
-\section{Setup}
-\label{sec:setup}
+\section{Set Up and Threat Model}
+\label{sec:threat}
+
+\newcommand*{\rsp}{\textsc{sp}}
 
@@
 the layout of memory, including where code lives and which addresses are within
-the bounds of the stack. We do not model a heap; programs might choose to use memory
-that is neither code nor stack to store global variables, as a heap, etc., but our
-properties do not protect it or assume any particular behavior. We also do not
-model run-time code generation.
+the bounds of the stack.\bcp{This implies no run-time code generation,
+  right?  This should probably be called out.}
+
+\BCP{What should we say about the heap?}
 
@@
 So we assume we are given a machine language program {\em annotated} with the locations of
-instructions representing calls, and we determine when a return has occurred by tracking
-control flow and observing when we reach a {\em return target}. For a given state annotated
-as a call, in a standard calling convention, another state fulfills its return target if
-it has the same stack pointer and the program counter is at the instruction following
-the call, but we abstract over different conventions.
-%, returns, and function entry points\BCP{I
-%  still want to talk about whether returns and entry points need to be
-%  annotated.  To me, these annotations do not make intuitive sense: it is
-%  the caller's responsibility to know whether the address that it is calling
-%is really an entry point, and we can trust it to check this.  Similarly, it
-%is not the callee's role to say when it is returning; it is the caller's
-%prerogative to decide when it has been returned to!  I.e., from the caller's
-%point of view, it has been returned to when control resumes at the
-%instruction following the call with the SP reset to where it was at the
-%point of the call. (This is the point where it wants to be able to access
-%its own local state again and expected to find it unmolested.)}.
-Although call annotations would typically come from a compiler together with
-the machine code, we do not assume anything about their provenance.
+instructions representing calls, returns, and function entry points\BCP{I
+  still want to talk about whether returns and entry points need to be
+  annotated.  To me, these annotations do not make intuitive sense: it is
+  the caller's responsibility to know whether the address that it is calling
+is really an entry point, and we can trust it to check this.  Similarly, it
+is not the callee's role to say when it is returning; it is the caller's
+prerogative to decide when it has been returned to!  I.e., from the caller's
+point of view, it has been returned to when control resumes at the
+instruction following the call with the SP reset to where it was at the
+point of the call. (This is the point where it wants to be able to access
+its own local state again and expected to find it unmolested.)}. Although these
+annotations would typically come from a compiler together with the machine code,
+we do not assume anything about their provenance.
+
+\bcp{Can this be postponed to the formal sections?}We further parameterize over the structure of the stack, particularly the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention}, which maps a machine state to a set of addresses.
+If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
+which addresses must be protected until the return.
 
@@
 
-The separation is smaller in protection mechanisms that interact more directly
-with application code. Some mechanisms might exhibit failstop behavior, but have
-no protected policy state; others might never even failstop. Then the policy-enhanced
-machine is simply identical to the original. This is the case for software-based
-techniques.
-
-\section{Threat Model}
-\label{sec:threat}
+The separation is less useful\bcp{but still sensible, right} in protection mechanisms that interact more directly
+with application code, such as Cheri-style capabilities\citep{??}\bcp{This
+  section would read better if we could cleanly separate the general
+  definitions from examples}. Here
+the policy would not have auxiliary state, but would still refine the machine step
+function to create a distinct failstop condition. Other techniques such as
+software fault isolation~\citep{??} might never actually failstop, but fulfill
+the properties by rewriting code so that traces can never violate them.
+
+\paragraph*{Threat Model}
 
@@
   eliminated return (and entry point) annotations.}
-\sna{I'm not sure we can escape questions about annotations, even if we have
-  just calls, but I tried to abbreviate it.}
-
-Since we rely heavily on call annotations, we must ask: what if the annotations
-are wrong? First, suppose a call occurs that is not annotated.
+
+Since we rely heavily on code annotations, we must ask: what if the annotations
+are wrong? First, suppose a call or return occurs that is not annotated.
 In this case, the answer is simple: as far as the properties are concerned, there has
-been no call, only an instruction sequence resembling one, and
-therefore no associated change in permissions. Then the caller is guaranteed no
-protections.
-
-Alternatively, an annotated instruction might not obey the calling convention.
+been no call or return, only an instruction sequence resembling such an operation, and
+therefore no associated change in permissions. This is particularly interesting in the
+case of a false return, where control may appear to have returned to the caller, but
+the enforcement mechanism must still protect the caller's memory.
+
+Alternatively, an annotated instruction might not really be a call or return.
 In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
 and returns are implemented by fixed sequences of instructions which the tag policy
-will force to be executed from start to finish, termed {\em blessed sequences}.
-When applying our properties to a micro-policy setting, we will statically rule out some
-scenarios with a well-formedness criterion on initial states: all annotated instructions will
-be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
-This is all we need for the policy to protect itself: it can then failstop if it would
-execute the blessed sequence out of order.
+will force to be executed from start to finish. The annotated call or return instruction should always
+be the final instruction of the appropriate sequence. What if an annotated instruction exists outside
+such a sequence? We will statically rule out some scenarios with a well-formedness criterion
+on initial states: annotated instructions will initially be tagged apropriately, as the
+final instruction in the call or return sequence, and all blessed\bcp{undefined} sequences will end with
+annotated instructions. This is all we need for the policy to
+protect itself: it can then failstop if it would execute the blessed sequence out of order.
+So, even in our strong attacker model, we need very little in the way of restrictions on
+the code when considering micropolicy-based enforcement.  \BCP{Is this
+  discussion still needed?  I had hoped that switching to our ``new''
+  definition on subtraces would eliminate such complexities!}
+
+\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
+a given enforcement mechanism may place a static well-formedness criterion
+on the code it applies to. Weaker enforcement mechanisms might also need dynamic
+reasonableness conditions to prevent the defender from sabotaging enforcement, as
+discussed in section \ref{sec:cheri}.  \BCP{Weak.}
 
@@
 
-%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
-%We believe that it is flexible enough to capture code-altering approaches, such as code
-%rewriting or capability-based techniques, though perhaps with less elegance.
+This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
+We believe that it is flexible enough to capture code-altering approaches, such as code
+rewriting or capability-based techniques, though perhaps with less elegance.
+
+\newcommand{\TAGS}{\mathcal{T}}
+\newcommand{\tagname}{t}
+\newcommand{\uP}{{\mu P}}
 
@@
 int f() {
-  int y;
-  *(&y-2) = 0;
-  return 5;
+10:     *($SP-2) = 0;
+11[R]:  return 5;
 }
 \end{verbatim}
-}
 \end{subfigure}
-\begin{subfigure}{.5\textwidth}
-{\small
-\begin{verbatim}
-1 main: add $1,%sp      ; allocate frame
-2       mov $42,-1(%sp) ; initialize local
-3       call f
-4       add -1(%sp),%ra ; set return value
-5       sub $1,%sp      ; deallocate frame
-6       ret
-
-10 f:   add $1,%sp      ; allocate frame
-11      mov $0,-3(%sp)  ; violation!
-12      mov $5, %ra     ; set return value
-13      sub $1,%sp      ; deallocate frame
-14      ret
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}{.25\textwidth}
+\begin{subfigure}{.4\textwidth}
 \begin{center}
 \begin{tabular}{l l}
-{\tt 2} &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}[Unseal]
+{\tt 1} &
+\memoryaddrs{12em}
+\memory{4}{\unsealc}[Unseal]
 ~$\cdots$
@@
 \\
-{\tt 3} &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}
+{\tt 2} &
+\memoryaddrs{12em}
+\memory{4}{\unsealc}
+~$\cdots$
+    \MemoryLabel{-15em}{0.75em}{42}
+    \vspace{.5em}
+\\
+{\tt 10} &
+\memoryaddrs{16em}
+\memory{2}{\mainsealc}[Seal(0)]%
+\memory{1}{\retptrc}[RetPtr]%
+\memory{1}{\unsealc}[Unseal]
 ~$\cdots$
@@
 \memoryaddrs{16em}
-\memory{1}{\mainsealc}[Seal(0)]%
-\memory{1}{\retptrc}[RetPtr]%
-\memory{1}{\unsealc}[Unseal]
-~$\cdots$
-\MemoryLabel{-15em}{0.75em}{42}
-\vspace{.5em}
-\\
-{\tt 12} &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}
+\memory{2}{\mainsealc}
 \memory{1}{\retptrc}
@@
 of examples. Existing stack safety enforcement work mostly refers directly to examples
-of bad behavior that it prevents; we will argue that:
+of bad behavior that it prevents; we will insert our model between the enforcement and
+the examples\bcp{Didn't understand that}, arguing that:
 
@@
 
-We show our examples both in C syntax and in corresponding assembly code for a
-simple two-address machine. The address of each instruction is shown to its left.
-The assembly code reflects a simple compilation model in which  
-local variables are stored in the stack frame.
-Unless otherwise stated, arguments and return values are passed in registers.
+Our examples are written in C-like syntax, labeled with line numbers for reference,
+but to reflect the power of the underlying assembly code, we allow code
+to read the stack pointer ({\tt \$SP}) directly and use
+it to form pointers.
+Call and return annotations are marked {\tt[C]} and {\tt[R]}
+respectively. Local variables are assumed to be stored in the stack frame,
+in declaration order. \bcp{This sort of thing implies that readers will have
+to mentally compile the source programs we're showing them.}Unless otherwise stated, arguments and return values are passed in registers.
 The machine is word-addressed. The stack grows upward in
 memory and {\tt \$SP} points to the first unused word above the top of the stack.
-The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
-implicitly adjusting the stack pointer.
 At the right of each example, we show the layout of stack memory just before
-the specified instruction is executed; each box represents a memory word and
-addresses increase to the right.
-
-Unless otherwise noted, colored boxes indicate an assignment of the address to a
-{\em domain} -- a label reflecting a particular set of access rights. Domains
-are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
-(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
-(outside the stack entirely.) For our purposes, \(\outside\) will apply to global
-variables. Domains are part of the {\em context} of a state.
-We will introduce the remaining context as needed.
+the specified line number is executed; each box represents a memory word and
+addresses increase to the right.  \BCP{I am worried about the C-like syntax:
+I think the mental complexity and imprecision that it introduces may not be
+worth the concision that we get from it.  Couldn't we get most of the
+concision by just assuming an ISA with CISC-style single-instruction Call
+and Return?}
+
+%The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
+%at the time of the call. The stack can be used to pass arguments or return results
+%in the usual way, by placing them at the top of the caller's frame. This will be modeled
+%as a form of sharing in section \sna{cite the sharing section}.
+%Callers and callees can also communicate via registers if they wish; indeed, our
+%definitions say nothing at all about registers (except $\PCname$ and $\rsp$).
+
+%\iftrue\bcp{This implicitly amounts to choosing a caller-saves register
+%  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi
 
@@
 meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
+{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of
+their owner\bcp{are we still using depths?},
 in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt x} will be unchanged when it gets control back.
+{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
 This gives us our simple, intuitive statement of {\em stepwise integrity}:
@@
 %cares about changes it can see after the return, not changes during execution.
-In this case, after instruction 8, {\tt f} has violated integrity by writing to {\tt x},
-so the final return value produced by {\tt main} is wrong.
-
-The black rectangle represents the return address, which is saved by the call
-instruction itself, so is not sealed.
+In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.
+
+The black rectangle represents the return address. In a typical RISC calling convention
+this will have been saved into a register by the call itself, and {\tt f} stores
+it in the stack during its preamble, so it is not itself sealed.
 
 \begin{figure}
-\begin{subfigure}{.2\textwidth}
-{\small
+\begin{subfigure}{.33\textwidth}
 \begin{verbatim}
 int main() {
-  int x = 42;
-  return f()+x;
+0:     int x;
+1:     x = 42;
+2[C]:  x = f();
+3[R]:  return x+y;
 }
@@
 int f() {
-  int y,z;
-  y = 5;
-  z = *(&y-2);
-  return y+z;
+10:     int y,z;
+11:     y = 5;
+12:     z = *($SP-4);
+13[R]:  return y+z;
 }
 \end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}{.5\textwidth}
-{\small
-\begin{verbatim}
-1 main: add $1,%sp      ; allocate frame
-2       mov $42,-1(%sp) ; initialize local
-3       call f
-4       add -1(%sp),%ra ; set return value
-5       sub $1,%sp      ; deallocate frame
-6       ret
-
-10 f:   add $2,%sp      ; allocate frame
-11      mov $5,-2(%sp)  ; assign into y
-12      mov -4(%sp),-1(%sp)  ; violation!
-13      mov -2(%sp),%ra ; calculate and
-14      add -1(%sp),%ra ; set return value
-15      sub $2,%sp      ; deallocate frame
-16      ret
-\end{verbatim}
-}
 \end{subfigure}
@@
 \begin{tabular}{l l | l}
-  {\tt 2} &
+  {\tt 1} &
   \multicolumn{2}{c}{
@@
     ~$\cdots$
-%    \MemoryLabel{-18em}{0.75em}{0}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
+    \MemoryLabel{-18em}{0.75em}{0}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
   } \\
-  {\tt 11} &
-  \memoryaddrs{21em}
+  {\tt 10} &
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
   \MemoryLabel{-19em}{0.75em}{42}
-%  \MemoryLabel{-10em}{0.75em}{0}
-%  \MemoryLabel{-6em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{0}
   &
-  \memoryaddrs{21em}
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
   \\
-  {\tt 12} &
-  \memoryaddrs{21em}
+  {\tt 11} &
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
   \MemoryLabel{-10em}{0.75em}{5}
- % \MemoryLabel{-6em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{0}
   &
-  \memoryaddrs{21em}
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
   \\
-  {\tt 13} &
-  \memoryaddrs{21em}
+  {\tt 12} &
+  \memoryaddrs{20em}
   \memory{1}{\unsealc}
@@
 \end{subfigure}
-\caption{A Confidentiality Violation\apt{Could save space by not repeating C or assemly code for {\tt main}.}}
+\caption{A Confidentiality Violation \bcp{It would help if the SPs could
+    have a little arrow after them pointing to the cell that they are
+    associated with.}}
 \label{fig:conf1}
@@
 of the context of its call (except for its arguments and any global
-variables).
-We state this as a kind of {\em noninterference} property~\citep{??}.
-Noninterference describes knowledge: it states that for system components
+variables\bcp{We haven't talked about global variables.  We should.}).
+We state this in a style similar\bcp{??} to {\em noninterference}~\citep{??}.
+Noninterference describes knowledge: it states that for some system components
 containing secret values, we can substitute any values whatsoever without changing the results
@@
 by reading {\tt main}'s variable {\tt x}. The trace
-shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
+shows how noninterference applies to this scenario. At line 11, we create a {\em variant}
 state with arbitrary values \(v_0\), etc., in place of the values that should
-be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
+be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt y}, has
 changed to 5 in both the primary trace and the variant. The primary and variant agree on
-the value in the changed location, so confidentiality was not violated. But at instruction 13,
+the value in the changed location, so confidentiality was not violated. But after line 13,
 the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
-in the variant trace it might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
+in the variant trace it might have any value \(v_2\). If \(v_2 \neq 42\) then the variant
 has behaved differently than the primary, violating confidentiality.
@@
 
-%\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
-%track which addresses may be read (i.e., which have been initialized), and at each step
-%guarantee that the step alters state identically to a step from a variant state. Then we
-%wouldn't need to divide up the execution by calls. But the approach we take is more
-%consistent with a typical noninterference approach, in that it varies the memory once
-%at the beginning and trace changes through the execution.
+\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
+track which addresses may be read (i.e., which have been initialized), and at each step
+guarantee that the step alters state identically to a step from a variant state. Then we
+wouldn't need to divide up the execution by calls. But the approach we take is more
+consistent with a typical noninterference approach, in that it varies the memory once
+at the beginning and trace changes through the execution.
 
@@
 
-\begin{subfigure}{.2\textwidth}
-  {\small
+\begin{subfigure}{.3\textwidth}
 \begin{verbatim}
 int main() {
-  int x;
-  x = f();
-  return g();
+0:     int x;
+1[C]:  x = f();
+2[CR]: return g();
 }
@@
 int f() {
-  int y = 5;
-  return y;
+10:     int y;
+11:     y = 5;
+12[R]:  return y;
 }
@@
 int g() {
-  int z;
-  if (z == 5)
-    return 10;
-  else
-    return 20;
+20:     int z;
+21:     if(z == 5)
+22[R]:    return 100;
+23:     else
+24[R]:    return 10;
 }
 \end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}{.7\textwidth}
-  {\small
-\begin{verbatim}
-1 main:  add $1,%sp      ; allocate frame
-2        call f
-3        mov %ra,-1(%sp) ; store x
-4        call g
-5        sub $1,%sp      ; deallocate frame
-6        ret
-
-10 f:    add $1,%sp      ; allocate frame
-11       mov $5,-1(%sp)  ; initialize y
-12       mov -1(%sp),%ra ; set return value
-13       sub $1,%sp      ; deallocate frame
-14       ret
-
-20 g:    add $1,%sp      ; allocate frame
-21       cmp $5,-1(%sp)  ; compare 
-22       brneq #25       ; branch if not equal
-23       mov $10,%ra     ; set return value
-24       jmp #26
-25       mov $20,%ra     ; set return value
-26       sub $1,%sp      ; deallocate frame
-27       ret 
-}
-\end{verbatim}
-}
 \end{subfigure}
@@
 \begin{tabular}{l r | l}
-  {\tt 2} &
+  {\tt 01} &
   \multicolumn{2}{c}{
@@
     ~$\cdots$
-%    \MemoryLabel{-18em}{0.75em}{0}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
+    \MemoryLabel{-18em}{0.75em}{0}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
@@
   ~$\cdots$
- % \MemoryLabel{-19em}{0.75em}{0}
-%  \MemoryLabel{-10em}{0.75em}{0}
-%  \MemoryLabel{-6em}{0.75em}{0}
+  \MemoryLabel{-19em}{0.75em}{0}
+  \MemoryLabel{-10em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   ~$\cdots$
- % \MemoryLabel{-19em}{0.75em}{0}
+  \MemoryLabel{-19em}{0.75em}{0}
   \MemoryLabel{-10em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   \\
-  {\tt 4} &
+  {\tt 02} &
   \multicolumn{2}{c}{
@@
     \MemoryLabel{-10em}{0.75em}{5}
-%    \MemoryLabel{-7em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
@@
   \\
-  {\tt 22} &
+  {\tt 21} &
   \memoryaddrs{16em}
@@
   \MemoryLabel{-10em}{0.75em}{5}
- % \MemoryLabel{-6em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   \\ &
-  {\bf 23}
+  {\bf 22}
   ~$\cdots$
   &
-  {\bf 25}
+  {\bf 24}
   ~$\cdots$
@@
 that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
-{\tt y} and hence {\tt x} without directly reading the stack pointer. In
+{\tt y} without even directly reading the stack pointer. In
 this case, {\tt g} branches on that value, which is a different way of
@@
 occurs because the program counter is also a component of the state that must change
-identically between the primary and variant traces. In a case where \(v_4 \not = 5\),
-the program counters will differ, breaking the lockstep.\apt{Is there any reason
-to complicate things like this by bringing in the PC?}
+identically between the primary and variant traces. In a case where \(v_1 \not = 5\),
+the program counters will differ, breaking the lockstep.
 
@@
   \centering
-  \begin{subfigure}{.3\textwidth}
-{\small
+  \begin{subfigure}{.4\textwidth}
 \begin{verbatim}
@@
 int main() {
-  int x = 1;
-  f();
-  x = -x;
-  f();
-  return x;
+0:     int x;
+1:     x = 1;
+2[C]:  f();
+3:     x = -x;
+4[C]:  f();
+5[R]:  return x;
 }
@@
 void f() {
-  int y;
-  if (!stash) {
-    stash = *(&y-1);
-  } else {
-    *(&y-1) = stash;
-    stash = 0;
+10:     if(!stash) {
+        // store return pointer
+11:       stash = $SP-1;
+12:     } else {
+13:       *($SP-1) = stash;
+14:       stash = 0;
         }
-  return;
+15[R]:    return;
 }
 \end{verbatim}
-}
   \end{subfigure}
-  \begin{subfigure}{.65\textwidth}
-{\small
-\begin{verbatim}
-   stash: .word 0
-1  main:  add $1,%sp        ; allocate frame
-2         mov $1,-1(%sp)    ; initialize x
-3         call f
-4         neg -1(%sp)       ; x = -x 
-5         call f           
-6         mov -1(%sp),%ra   ; set return value
-7         sub $1,%sp        ; deallocate frame
-6         ret
-
-10 f:     add $1,%sp        ; allocate frame
-11        cmp stash, $0     ; compare
-12        breq #15          ; branch if equal
-13        mov -2(%sp),stash ; update stash
-14        jmp #17
-15        mov stash,-2(%sp) ; retrieve stash
-16        mov $0,stash
-17        sub $1,%sp        ; deallocate frame
-18        ret 
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}{.4\textwidth}
+  \begin{subfigure}{.5\textwidth}
     \begin{tabular}{l l l}
+      {\tt 2} &
+      \memoryaddrs{8em}
+      \memory{4}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{1} \\
+      {\tt 15} &
+      \memoryaddrs{12em}
+      \memory{1}{\mainsealc}[Seal(0)]%
+      \memory{1}{\goodc}[RetPtr]%
+      \memory{2}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{1}
+      \MemoryLabel{-15em}{0.75em}{\#3}
+      \vspace{.5em} &
+      \memory[3em]{1}{\mainsealc}[Target]%
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#3}
+      \\
       {\tt 3} &
       \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      {\tt 18} &
+      \memory{4}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{1} \\
+      {\tt 10} &
       \memoryaddrs{12em}
-      \memory{1}{\mainsealc}[Seal(0)]%
-      \memory{1}{\unsealc}[RetPtr]%
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1}
-      \MemoryLabel{-11em}{0.75em}{\#4}
-      \vspace{.5em} &
-      \memory[3em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#4}
+      \memory{1}{\mainsealc}
+      \memory{1}{\goodc}
+      \memory{2}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{-1}
+      \MemoryLabel{-15em}{0.75em}{\#5} &
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
       \\
-      {\tt 4} &
-      \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      {\tt 11} &
-      \memoryaddrs{16em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#6} &
-      \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
-      \\
-      {\tt 18} &
+      {\tt 15} &
       \memoryaddrs{12em}
@@
       \memory{1}{\badc}
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#4} &
+      \memory{2}{\unsealc}%
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{-1}
+      \MemoryLabel{-15em}{0.75em}{\#3} &
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
+      \\
+      {\tt 3} &
+      \memoryaddrs{8em}
+      \memory{1}{\mainsealc}
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{-1} &
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
       \\
@@
       \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1} &
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-19em}{0.75em}{\bf 1} &
       \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
-      \\
-      {\tt 5} &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{\bf 1} &
-      \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6} \\
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5} \\
     \end{tabular}
@@
 One of the most pernicious ways to attack the stack is by hijacking control flow.
-Our stack safety properties do not assume that control flow follows a typical structure
-of matching sets calls and returns nested within on another.
-Rather, they express what it means to protect data even in the presence of control-flow attacks.
-Consider example \ref{fig:controlflow}.
-Here we show return addresses explicitly as line numbers, in the form $\#n$.
-Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
+Our stack safety properties do not assume that control flow is
+well-bracketed\bcp{undefined term}; rather,
+they express what it means to protect data even in the presence of control-flow attacks.
+Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''\bcp{??}~\citep{??}.
+(Here we show return addresses explicitly as line numbers, in the form $\#n$.)
+\apt{what do the colors mean in this example?}
+Function {\tt f} is called twice, at lines 2 and 4; it should therefore return to lines
 3 and 5, respectively. But the first time, before returning, it stashes its
-return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
-up being called a third time, at which point it finally returns to instruction 5.
-
-Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that {\tt main} returns
--1. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
+return pointer away, and the second time it returns to line 3 instead of 5; then it ends
+up being called a third time, at which point it returns to line 5.
+
+Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that we return
+-1. But because of {\tt f}'s interference, we negate {\tt x} twice, and return 1.
 This has the appearance of an integrity violation.  How do we capture this in our
@@
 
-The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
-to do more than just execute a return-labeled instruction: it must actually reach the
-appropriate {\em return target} in {\tt main}. A return target is a predicate on states,
-in this case one that holds on any state whose program counter points to the correct instruction.
-We introduce a new piece of context: we keep a stack of return targets, and a callee is
-only considered to have returned when it reaches a state that matches the caller's target.
-A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, here represented by sharing a color with the sealed data.
+The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs to do more than just
+execute a return-labeled instruction: it must actually reach the appropriate
+{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep
+a stack of return targets\bcp{do we still},
+which correspond to the depths at which stack frames have been sealed. A callee is
+only considered to have returned when it executes a return-annotated instruction
+and then reaches a valid return target\bcp{what's that?}. Only then do the sealed frames become unsealed.
 
 In the example, on its first return, {\tt f} reaches its return target, because the
-\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
-(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
-Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.
+\(\PCname\) is 3 after its return. But on the second, {\tt f} does not properly return,
+but rather changes its stack pointer and jumps into {\tt main}'s code. Then {\tt f}
+attempts to overwrite the sealed value of {\tt x}, violating integrity.
+
+\paragraph*{Pass by Reference and Stack Arguments}
+\bcp{I wonder if this could be postponed so that we get to the formal
+  definitions for the basic property sooner...}
+
+\newcommand{\mainpassc}{magenta}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.3\textwidth}
+\begin{verbatim}
+int main() {
+0:           int x;
+1:           x = 5;
+2[P($SP)C]:  x = f(x);
+3[R]:        return x;
+}
+
+int f(int a) {
+10:    int y;
+11:    y = 2*a;
+12:    a = 0;
+13[R]: return y;
+}
+\end{verbatim}
+  \end{subfigure}
+\begin{subfigure}{.65\textwidth}
+\begin{center}
+\begin{tabular}{l r | l}
+  {\tt 2} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{5}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{0}
+    \MemoryLabel{-7em}{0.75em}{0}
+    \vspace{.5em}
+  } \\
+  {\tt 11} &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}[Seal(0)]%
+  \memory{1}{\mainpassc}[P(0-1)]%
+  \memory{1}{\retptrc}[RetPtr]%
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
+  \\
+  {\tt 12} &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  {\tt 13} &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{20em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  {\tt 3} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{10}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{3}
+    \MemoryLabel{-7em}{0.75em}{10}
+    \vspace{.5em}
+  } \\
+\end{tabular}
+\end{center}
+
+\vspace{\abovedisplayskip}
+
+\end{subfigure}
+
+\caption{Sharing on the Stack}
+\label{fig:passingsimple}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.45\textwidth}
+\begin{verbatim}
+int *stash = 0;
+
+int main() {
+0:         int x, y;
+1[P(x)C]:  f(x);
+2:         x = 0;
+3[P(y)C]:  f(y);
+4[R]:      return x;
+}
+
+void f(int &a) {
+10:    if(!stash)
+11:      stash = a;
+12:    *stash = 5;
+13[R]: return;
+}
+\end{verbatim}
+  \end{subfigure}
+  \begin{subfigure}{.5\textwidth}
+    \begin{tabular}{l l}
+      {\tt 10} &
+      \memoryaddrs[(0)]{17em}
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\retptrc}
+      \memory{1}{\unsealc}
+      ~$\cdots$ \\
+      {\tt 13} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainpassc}
+      \memory{1}{\mainsealc}
+      \memory{1}{\retptrc}
+      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{5}
+      \\
+      {\tt 3} &
+      \memoryaddrs[(0)]{12em}
+      \memory{4}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 10} &
+      \memoryaddrs[(0)]{17em}
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
+      \memory{1}{\retptrc}
+      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 13} &
+      \memoryaddrs[(0)]{17em}
+      \memory{1}{\mainsealc}
+      \memory{1}{\mainpassc}
+      \memory{1}{\retptrc}
+      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{\bf 5}
+\end{tabular}
+\end{subfigure}
+\caption{A violation with pass-by-reference}
+\label{fig:passing}
+\end{figure}
+
+To this point we have ignored function arguments, since our simplest model places them in
+registers. This is an oversimplification, which we will now correct, starting with the simple
+case: passing by value on the stack. Example \ref{fig:passingsimple} shows a sample trace of
+a program that passes a value on the stack, without an attack.
+At the call, we add a new annotation \(P(\$SP)\) identifying a memory location to be
+passed, and our diagram displays it in magenta. The caller's frame locations that are not
+passed are sealed as usual. Just as we marked sealed memory with the depth of its owner\bcp{??},
+we do the same for passed memory -- but it has multiple owners. We keep track of the
+current depth of the active function next to the stack pointer in the diagram, for reference.
+The argument {\tt a} has
+its value copied from {\tt x}, and we note that {\tt f} can read it -- because it is marked
+as passed between depths 0 and 1, it is not varied. Then from line 12 to 13 it is even
+modified, and naturally this is permitted as well.
+
+The same annotation mechanism can be used to pass
+variables by reference, and here there are important stack safety concerns,
+as seen in Figure~\ref{fig:passing}. In this example, where we revert to assuming
+that arguments are passed in registers,
+{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
+passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
+{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
+{\tt x} in the second call.
+
+So at the initial call, {\tt x} is passed, and {\tt f} is free to write to it. Then upon
+return, {\tt main}'s frame is unsealed. At the second call it is {\tt y} that is passed,
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
+a future call with no arguments passed by reference would also violate integrity if it wrote
+to either variable.
+
+When passing by reference, it is possible to repeatedly pass the same reference further
+and further down the stack. Note, however, that this sharing of the reference must be over
+an unbroken chain of calls. So it suffices to track two depths for each passed
+cell: the original owner, and the highest depth it has been passed to. It is considered
+accessible anywhere between the two.
+
+Our model provides only limited support for handling fully first-class pointers into
+stack frames, e.g., pointers obtained by taking the address of a local variable.
+Unlike the case for pass-by-reference, there is no easy way to model the intended sharing behavior
+of such pointers in terms of stack frame activation patterns.  The best approximation we can make is to
+treat an address-taken variable in the same way as a passed argument, with the additional property
+that it remains accessible at \emph{all} stack depths above the caller.
+This prevents most violations of temporal safety, although stale pointers can still be
+used in certain corner cases. \apt{Trying not to spell out the details, which are complex.}
+More seriously, callees are permitted to access the location without having
+been explicitly given a pointer to it.  Preventing the latter kind of abuse
+is the provence of a fine-grained memory safety policy, e.g. based on capabilities
+(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
+\emph{per se}.
 
@@
 that doesn't share between caller and callee. All arguments are passed
-in registers. Recall that we keep track of which addresses are
-annotated as calls, with a ``code map'':
-\[\codemap \in \CODEMAPS \subseteq \WORDS\]
-
-We parameterize over the structure of the stack, particularly the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
-maps a machine state to a set of addresses.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
-which relates call states to the states that can be considered valid returns to their caller.
-The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
-and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-But whatever the return convention, at each step we update the context according
-to the context-step function, \(\stepsto_C\).
+in registers.
+
+We begin with our only annotations being {\em call} or {\em return}, and
+define our code map accordingly.
+
+\[\ann \in \ANNS ::= \callmap + \retmap\]
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \ANNS\]
+
+When a function makes a call, it establishes that its frame should be
+``sealed'' for later use. At any given step of a program, any number of
+activations may have claimed memory in this way, but which cells are sealed
+is dynamic information and cannot be deduced from the state alone. So it must
+be tracked separately. We must also track the conditions under which a caller
+considers itself to have been returned to.
 
@@
 of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller will register a {\em target}
-condition -- a predicate on machine states that must hold for it to consider itself active.
+at a particular depth\bcp{What about code??}. Each caller will register a {\em target} condition -- a
+predicate on machine states that must hold for it to consider itself active.
 A context is a pair of a ``domain map'' from components to domains, and a
@@
 
+Now we can define the context update relation. Let \(\sealcon\) be the sealing convention
+that maps a machine state to a set of components, typically the set of addresses below
+the stack pointer. Similarly, \(\retcon ~ \mach ~ \mach'\) will be the return convention
+that relates the call state \(\mach\) to return states \(\mach'\) that are valid returns.
+The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
+and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
+Then at each step we update the context according to the context-step function,
+\(\stepsto_C\).
+
 The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
-A call annotation tells us to push the return target onto the return stack and
-seal all components with the depth of the previous return stack. If we aren't making a call,
-we identify a return by looking ahead and checking if we're about to step to a state
-that matches the topmost return target. If neither is true, the context remains unchanged.
+A call simply pushes the return target onto the return stack and seals all components
+with the depth of the previous return stack. A return
+recursively searches for a valid return target at the next step, and unseals
+every component sealed at a depth that it passes.
 
@@
 
-\judgmentthree[ReturnFound]
+\judgmenttwobr[ReturnFound]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
               {\(\mach \stepsto \mach' \in \target\)}
@@
               {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\judgmenttwobr[ReturnRec]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
 
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change
-together if for all components \(\component\) such that
-\(\mach ~ \component \not = \mach' ~ \component\) or
-\(\nach ~ \component \not = \nach' ~ \component\),
-\(\mach' ~ \component = \nach' ~ \component\).
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change together
+if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
+\(\nach ~ \component \not = \nach' ~ \component\), \(\mach' ~ \component = \nach' ~ \component\).
 This is written \(\mach,\mach' \diamond \nach,\nach'\).
 
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
-if their adjacent pairs change together. We will represent this relation with
-\(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
-\sna{It's a new notation, we use this enough that it seems worth picking an
-  evocative symbol.} 
-We define it as a coinductive relation:
+\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep} if their adjacent pairs
+change together, written as a coinductive relation \(\machT \doteq
+\nachT\):\bcp{People may not be familiar with the notation}
 
@@
 \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
-\(\MPCT\) corresponds to the trace of a subroutine from its call to its return.
+\(\MPCT\) corresponds to the trace of a function from its call to its return.
 Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
@@
 \subsection{Stack Safety with Passing}
-
-\newcommand{\mainpassc}{magenta}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{.2\textwidth}
-{\small
-\begin{verbatim}
-void main() {
-  int x = 5;
-  x = f(x);
-  return;
-}
-
-void f(int a) {
-  int y = 2*a;
-  a = 0;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}{.6\textwidth}
-{\small
-\begin{verbatim}
-1  main: add $1,%sp          ; allocate frame
-2        mov $5,-1(%sp)      ; initalize x
-3        add $1,%sp          ; allocate arg
-4        mov -2(%sp),-1(%sp) ; set arg
-5        call f              ; SHARE: %sp-1
-6        sub $1,%sp          ; deallocate arg
-7        mov %ra,-1(%sp)     ; store x
-8        sub $1,%sp          ; deallocate frame
-9        ret
-
-10  f:   add $1,%sp          ; allocate frame
-11       mov -3(%sp),-1(%sp) ; y = a
-12       mul $2,-1(%sp)      ; y = a*2
-13       mov $0,-3(%sp)      ; a = 0
-14       sub $1,%sp          ; deallocate frame
-15       ret
-\end{verbatim}
-}
-  \end{subfigure}
-\begin{subfigure}{.65\textwidth}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 3} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{5}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[Seal(0)]%
-  \memory{1}{\mainpassc}[P(0,1)]%
-  \memory{1}{\retptrc}[RetPtr]%
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  {\tt 13} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  {\tt 14} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  \\
-  {\tt 8} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{10}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{\#6}
-    \MemoryLabel{-7em}{0.75em}{10}
-    \vspace{.5em}
-  } \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-
-\caption{Sharing on the Stack}
-\label{fig:passingsimple}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{.2\textwidth}
-{\small
-\begin{verbatim}
-int *stash = 0;
-
-int main() {
-  int x, y;
-  f(x);
-  x = 0;
-  f(y);
-  return x;
-}
-
-void f(int &a) {
-  if(!stash)
-    stash = &a; 
-  *stash = 5;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}{.49\textwidth}
-{\small
-\begin{verbatim}
-   stash: .word 0
-1  main:  add $2,%sp      ; allocate frame
-2         mov %sp,%ra
-3         add -2,%ra      ; ra = &x
-4         call f          ; SHARE: %sp-2
-5         mov $0,-2(%sp)  ; x = 0
-6         mov %sp,%ra
-7         add -1,%ra      ; ra = &y
-8         call f          ; SHARE: %sp-1
-9         mov -2(%sp),%ra ; set return value
-10        sub $2,%sp      ; deallocate frame
-11        ret
-
-20 f:     cmp stash, $0   ; compare
-21        breq #13        ; branch if equal
-22        mov %ra,stash   ; stash = &a
-23        jmp #15
-24        mov $5, (stash) ; *stash = 5
-25        ret 
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}{.29\textwidth}
-    \begin{tabular}{l l}
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$ \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}
-      \memory{1}{\mainsealc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{5}
-      \\
-      {\tt 8} &
-      \memoryaddrs[(0)]{12em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\mainpassc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{\bf 5}
-\end{tabular}
-\end{subfigure}
-\caption{A violation with pass-by-reference\apt{SP(n) depths looked wrong; SNA double-check that I fixed them right.}}
-\label{fig:passing}
-\end{figure}
-
-To this point we have ignored function arguments, since our simplest model places them in
-registers. This is an oversimplification, which we will now correct, aiming at two kinds
-of sharing: passing by value on the stack, and passing by reference. First we will explore
-some examples at the same level of detail as our earlier ones, and then move on to
-extending our formalization.
-
-Example \ref{fig:passingsimple} shows a sample trace of
-a program that passes a value on the stack, without an attack.
-We extend our call annotation with an additional parameter identifying a memory location to be
-passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
-to include \(\passed{\depth,\depth'}\), recording two depths. We note the
-current depth of the active function next to the stack pointer in the diagram, for reference.
-The argument {\tt a} has
-its value copied from {\tt x}, and we note that {\tt f} can read it -- because it is marked
-as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
-modified, and naturally this is permitted as well.
-
-The same annotation mechanism can be used to pass
-variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference.
-{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
-{\tt f} again passing a reference to {\tt y}.
-We would expect that the first call to {\tt f} might modify
-{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++!)
-
-To avoid cluttering the picture at the assembly level, we revert to assuming that the
-argument {\tt a} is passed in a register, namely {\tt \%ra}.
-The question is how the local variables {\tt x} and {\tt y} themselves
-should be labelled in order for the policy to detect an integrity violation.
-At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
-return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
-a future call with no arguments passed by reference would also violate integrity if it wrote
-to either variable.
-
-When passing by reference, it is possible to repeatedly pass the same reference further
-and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices to track two depths for each passed
-cell: the original owner, and the highest depth it has been passed to. The policy considers
-it to be accessible anywhere between the two.
-
-Our model provides only limited support for handling fully first-class pointers into
-stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike the case for pass-by-reference, there is no easy way to model the intended
-sharing behavior of such pointers in terms of stack frame activation patterns.
-The best approximation we can make is to treat an address-taken variable in the
-same way as a passed argument, with the additional property
-that it remains accessible at \emph{all} stack depths above the caller.
-This prevents most violations of temporal safety, although stale pointers can still be
-used in certain corner cases.
-More seriously, callees are permitted to access the location without having
-been explicitly given a pointer to it.  Preventing the latter kind of abuse
-is the provence of a fine-grained memory safety policy, e.g. based on capabilities
-(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
-\emph{per se}.
 
@@
 
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap\]
 
@@
   \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
+  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth,\depth} \\
   \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
@@
 
-\judgmenttwo[ReturnFound]
+\judgmentthree[ReturnFound]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
               {\(\mach \stepsto \mach' \in \target\)}
@@
               {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\judgmentthree[ReturnRec]
+              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\domMap'' = \mathit{pop} ~ \domMap ~ |\rts|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
 pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We add a second annotation indicating yields.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
+We extend the set of possible annotations with a yield annotation.
+
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \retmap + \yieldmap\]
 
@@
 if the next state is a valid yield target for that stack, and we update the yield target for
-the current stack, ensuring that when the system yields back we expect the current coroutine
-to continue from where it left off. So, it is possible for a coroutine to claim to yield
-but not actually change the active stack, with similar effects to when a callee fails
-to return properly. Now we can define both stack safety and coroutine safety under this model.
+the yielding stack. So it is once again possible for a coroutine to claim to yield but not
+actually change the active stack, with similar effects to when a callee fails to return properly.
+Now we can define both stack safety and coroutine safety under this model.
 
@@
 Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
 where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
@@
 \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
 where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
@@
 
-%\subsection{Well-bracketed Control Flow}
-%\label{sec:wbcf}
-
-%Both the integrity and confidentiality components of stack safety
-%are reasonable even in programs that violate expected
-%control flow. But even though stack safety prevents control-flow violations
-%from being used to gain privileges, it is worth protecting control flow as well.
-%Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
-%control flow.
-
-%When we focus on subroutines, the essence of well-bracketed control flow
-%is that a caller can expect that if its callee ever returns, it returns
-%to the caller at the appropriate return target. This is tantamount to saying
-%that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
-%the simple model:
-
-%\definition A system enjoys {\em well-bracketed control flow} if for all
-%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-%\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
-
-%Extending this idea to coroutines, we use the same notion that there are
-%no ``false'' yields.
-
-%\definition A system enjoys {\em well-structured control flow} if for all
-%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-%\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
-%where \(\stackid \not = \stackid'\).
+\subsection{Well-bracketed Control Flow}
+\label{sec:wbcf}
+
+Both the integrity and confidentiality components of stack safety
+are reasonable even in programs that violate expected
+control flow. But even though stack safety prevents control-flow violations
+from being used to gain privileges, it is worth protecting control flow as well.
+Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
+control flow.
+
+When we focus on functions, the essence of well-bracketed control flow
+is that a caller can expect that if its callee ever returns, it returns
+to the caller at the appropriate return target. This is tantamount to saying
+that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
+the simple model:
+
+\definition A system enjoys {\em well-bracketed control flow} if for all
+reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
+\(\mpcstatename \stepstocon \mpcstatename'\), if
+\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
+is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
+
+Extending this idea to coroutines, we use the same notion that there are
+no ``false'' yields.
+
+\definition A system enjoys {\em well-structured control flow} if for all
+reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
+\(\mpcstatename \stepstocon \mpcstatename'\), if
+\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
+is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
+where \(\stackid \not = \stackid'\).
 
@@
 
-\subsection{Lazy Stepwise Integrity}
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\sealed}\) or \(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(\depth' < |\rtm ~ \stackid|\) .
-Then a system enjoys observable
-
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+\subsection{Observable Integrity}
 

### 2021-04-15 tex/main.tex
Some polishing and comments.

--- 2021-04-15 tex/main.tex
@@
 %  to leave it out of the title but hit it strongly in the abstract.}%
-What is stack safety? The phrase is associated with a variety of compiler,
+What does ``stack safety'' mean, exactly? The phrase is associated with a
+variety of compiler,
 run-time, and hardware mechanisms for guarding against corruption of stack
-memory.  But these mechanisms typically lack rigorous specifications,
-relying instead on informal descriptions and specific examples of the bad behaviors that they prevent.
+memory.  But these mechanisms typically lack precise specifications,
+relying instead on informal descriptions and specific examples of bad
+behaviors that they prevent.
 
 We propose formal criteria for stack safety, using the concepts of
-language-based security: a combination of an integrity property (``each
-caller's stack frame is held invariant by the callee'') and a confidentiality
-property (``the callee's behavior is insensitive to differences in the
-caller's state'').
+language-based security: a combination of an integrity property (``the private
+state in each caller's stack frame is held invariant by the callee'')
+and a confidentiality property (``the callee's behavior is insensitive to the
+caller's private state'').
 % These properties are sensible even in the presence of
@@
 and integrity properties. Their more performant ``lazy'' micro-policy
-does not, but corresponds to weaker, more extensional ``observational'' variants
-of the
-properties.  While we do not rule out other enforcement mechanisms, our
+does not, but it does (almost) correspond to weaker, more extensional
+variants of the properties.
+%
+\BCP{I still don't think the next sentence really nails it.  And (sigh) I am
+  still not really clear yet on whether our definition applies / might apply
+  to other enforcement mechanisms or not.  Saying that our attacker model is
+  stronger does not help---that should make it easier for us to describe
+  their protections, not harder.  So I don't understand what we're saying
+  here.}%
+%
+While we do not rule out other enforcement mechanisms, our
 definitions are strongly informed by micropolicy systems, especially
@@
 can enable.
+
+\bcp{We should decide between ``criteria'' and ``properties''...}
 
@@
 
-The call stack is a perennial target for low-level attacks
-based on memory-safety vulnerabilities, leading to consequences
-that range from leakage or corruption of private stack data to
+The call stack is a perennial target for low-level attacks, leading to
+consequences
+from leakage or corruption of private stack data to
 control-flow hijacking. To deter or detect such attacks, a profusion of
@@
 The protections offered by such mechanisms are commonly described in terms
-of concrete examples of attacks that they can prevent, such as corruption
-of return addresses, buffer overflows, or use of uninitialized variables,
-leaving a more abstract characterization to the reader's intuition.
-But these mechanisms can be quite intricate
-and often require trading off protection and efficiency.
+of concrete examples of attacks that they can prevent---corruption
+of return addresses, buffer overflows, use of uninitialized variables,
+etc.---leaving a more abstract characterization to the reader's intuition.
+But these mechanisms can be quite intricate,
+and they often require trading off protection and efficiency.
 Thus, it would be useful to have a precise, generic, and formal
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-The most abstract characterization stack safety of which we are aware is in
+The most abstract characterization stack safety that we are aware of is in
 recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
@@
 return to the point just after their call (if they return at all).
-\citet{Skorstengaard+19} formalize these properties by defining an idealized
+Skorstengaard et al.{} formalize these properties by defining an idealized
 machine in which they hold {\em by construction};
 they then show that an enforcement
-mechanism (in their case, hardware capabilities) guarantees stack safety by
+mechanism (in their case, hardware capabilities) guarantees stack safety, by
 proving that an ordinary machine with enforcement can simulate the idealized machine.
@@
 
-In this work, we also distinguish between protecting data (like LSE) and
-control flow (with WBCF), and formalize a data protection property
-based on the technical framework of language-based security~\citep{??}.
+In the present work, we also distinguish between protecting stack data (like
+LSE) and
+control flow (with WBCF).  But rather than defining stack safety as ``what
+a certain idealized machine does,'' we formalize explicit criteria
+based on the technical framework of language-based
+security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
+  survey?}\fi.
 Our core properties are \emph{stepwise stack integrity} and
@@
 we can express safety in an information-flow sense without reference to it,
-even safety from attacks that hijack control flow.
+even safety from attacks that hijack control flow.  \bcp{On consideration
+  (especially in light of the weird ``C confuses A by returning directly to
+  A instead of returning to B so that B can finish up some stuff before
+  returning to A'' example\footnote{Actually, this example is not so weird.
+  If C, the attacker, can return a {\em result} directly to A, who is
+  expecting a result from B, then that is clearly bad!}), I
+agree now that we should discuss WBCF.  It should be kept as distinct as
+possible from the ``data safety'' part, since it is not really an
+information flow property.  But our treatment of it is interestingly
+different from StkTokens, because of our call annotations, and we should
+discuss that.}
 
@@
 
-\section{Key Ideas, By Example}
+\section{Key Ideas by Example}
 \label{sec:running-example}
@@
 \sna{It's a new notation, we use this enough that it seems worth picking an
-  evocative symbol.} 
+  evocative symbol.}\bcp{Sorry -- I meant the notational convention of representing a
+coinductive relation with inference rules in this way, not the symbol...}
 We define it as a coinductive relation:

### 2021-04-15 tex/main.tex
comments on 5

--- 2021-04-15 tex/main.tex
@@
 Our machine states are composed of {\em components} \(\component\),
-which are either words (memory addresses) or by register names (\(\reg\)) drawn from some set
-\(\REGS\).
+which are either memory addresses (words) or register names (\(\reg\)) drawn from some set
+\(\REGS\), which is assumed to include two special purpose registers: the
+program counter {\PCname} and the stack pointer \(\SP\).
 %
@@
 %
-The register names are assumed to include two special purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-
 A {\em machine state} is a map from components to values:
@@
 \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-
-We have a total step function between machine states \(\mach \stepsto \mach'\).
+%
+We have a total step function between machine states, written \(\mach \stepsto \mach'\).
+\apt{But just below we write $\mach \stepstoobs \mach'$. Which is it?}
+\apt{Say something about initial machine state?}
 
@@
 
-We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
+We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
 A {\em policy} consists of a set of policy
@@
 can see the machine state, but cannot modify it. We assume that the
-compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.
+compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.\apt{see comment above}
 %
@@
 
-A trace over a type \(\TRACEELEMS\)\bcp{``over a type A'' is a Coq-ism} is a nonempty, finite or infinite sequence of
-\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
-\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
+A trace \(T \in \TRACE{\TRACEELEMS}\) is a nonempty, finite or infinite sequence of elements
+of some type \(\TRACEELEMS\).  We will mostly be concerned with sequences of MPC-states,
 called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
@@
 coinductively relates an initial state with the trace of states
-produced by repeated application of step, and potentially cut short by
-the policy component. The {\sc Default} rule applies only if the other does not.
+produced by repeated application of \(\stepstocon\), and potentially cut short by
+the policy component. % The {\sc Default} rule applies only if the other does not.
 (We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
 
+\begin{center}
 \begin{minipage}{.4\textwidth}
@@
 \begin{minipage}{.4\textwidth}
-\judgment[Default]
-         {}
+\judgment%[Default]
+         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
          {\(\mpcstatename \hookrightarrow \mpcstatename\)}
 \end{minipage}
+\end{center}
 %
@@
 The operation \(\PUT ~ f ~ \Trace\) takes a trace
-\(\Trace \in \TRACE{\TRACEELEMS \times \OBSS}\) and a predicate
+\(\Trace \in \TRACE{\TRACEELEMS %\times \OBSS
+}  \) and a predicate
 on elements \(f \subset \TRACEELEMS\) and gives the prefix of
@@
 %
+\apt{If we are writing \(f \subset \TRACEELEMS\), perhaps we should also
+  be writing \(a \in f\), etc.  rather than \(f a\), etc.?}
 If the resulting trace is a strict prefix of the parameter, then the
@@
 such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
+\apt{Unclear how there can be multiple initial states.}
 

### 2021-04-15 tex/main.tex
Some clarifications on related work
Added a few words about Georges et al.'s follow-up work, tried to
address Benjamin's comments and expressing more clearly the fact that
the StkToken definitions are embedded in the semantics.

@@
 relation captures capability safety without ``externally observable
-side-effects (like console output or memory access traces).'' \bcp{Not very clear:}In the
+side-effects (like console output or memory access traces)''
+and can be used to reason about individual programs. \bcp{Not very clear:}In the
 discussion, ``while [the authors] claim that [their] calling
@@
 can be modeled in our framework and satisfy our definition of stack safety in
-\cref{sec:lse-and-wbcf}. Further, we expect them to satisfy the even stronger
-property developed in \cref{sec:enforcement}. Proving\bcp{or testing!} these conjectures would
-involve materializing\bcp{?} their definitions of local state encapsulation and
-well-bracketed control flow, which exist only in enmeshed\bcp{?} form as part of the
-semantics of the capability machines. The main practical limitation of StkTokens
+\cref{sec:lse}. Further, we expect them to satisfy the even stronger
+property developed in \cref{sec:enforcement}. Proving (or testing) these conjectures would
+involve exposing their formal definitions of local state encapsulation and
+well-bracketed control flow, which are built-in as part of the
+semantics of the capability machines and not given explicitly.
+The main practical limitation of StkTokens
 is its reliance on linear capabilities, as it is unclear now they could be added
@@
 to implement linear return capabilities.\bcp{But IIUC the Dover realization
-  of micro-policies can't do this...?})
+  of micro-policies can't do this...?})\rb{Correct}
+%
+In recent work, \citet{Georges+21} address these efficiency concerns
+with the introduction of a new type of uninitialized capability that
+avoids the need to clear the stack while being more amenable to
+practical hardware implementation, although like
+\citet{Skorstengaard+19b} it lacks a general stack safety theorem,
+instead using a similar logical relation to reason about the safety of
+individual programs.
 

### 2021-04-15 tex/main.tex
Update enforcement with current example

@@
 
-Example \ref{fig:passingsimple} shows a sample trace of
+Example~\ref{fig:passingsimple} shows a sample trace of
 a program that passes a value on the stack, without an attack.
@@
 %
-The running example in \cref{fig:running-trace-a} manages the stack in this way.
-For example, {\tt f}'s ``blessed'' sequences allocate space to save its return
-address (and private variables) in instructions 20--21, and later restores that
-return address and deallocates its private frame space before returning in
-34--36. Additionally, unlike in the simplified presentation of the policy, the
-caller passes information to the callee through the stack, and for this creates
-and initializes (22 and 24--25) and destructs (29) the argument part of the
-callee's stack frame.\rb{The policy would also need to control this, I think it
-  is clear}
-\rb{TODO: References to the running example are no longer valid! We could use
-figures 5 or 6, referencing the figures, not the invisible assembly code.}
+For example, the program in \cref{fig:passingsimple} manages the stack
+in this way. Here, the call pseudo-instruction to {\tt f} in line 5
+hides a blessed sequence of instructions that allocate space to save
+its return address to {\tt main} and seal the protected part of the
+stack before {\tt f}'s body starts executing at 10. Conversely, the
+return instruction in line 15 restores and jumps to the return address
+and deallocates its frame space, and unseals the corresponding parts
+of the stack. The function itself allocates and deallocates its own
+private frame space in instructions 10 and 14. Additionally, unlike in
+the simplified presentation of the policy, the caller shares the
+arguments passed to the callee through the stack, and for this creates
+and initializes (3--4) and destructs (6) the argument part of the
+callee's stack frame.
 %

### 2021-04-15 tex/main.tex
Comments / polishing

--- 2021-04-15 tex/main.tex
@@
 LSE) and
-control flow (with WBCF).  But rather than defining stack safety as ``what
-a certain idealized machine does,'' we formalize explicit criteria
+control flow (like WBCF).  But our formulations of both properties are
+somewhat different from those of~\cite{Skorstengaard+19}.  Rather than
+defining stack safety implicitly,
+as ``what a certain idealized machine does,'' we formalize explicit criteria
 based on the technical framework of language-based
@@
 we can express safety in an information-flow sense without reference to it,
-even safety from attacks that hijack control flow.  \bcp{On consideration
+even safety from attacks that hijack control flow.  \BCP{On consideration
   (especially in light of the weird ``C confuses A by returning directly to
@@
 Our machine states are composed of {\em components} \(\component\),
-which are either memory addresses (words) or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special purpose registers: the
+which are either words (memory addresses) or by register names (\(\reg\)) drawn from some set
+\(\REGS\).
+%
+    \[\component \in \COMPONENTS ::= \WORDS + \REGS \]
+%
+The register names are assumed to include two special purpose registers: the
 program counter {\PCname} and the stack pointer \(\SP\).
-%
-    \[\component \in \COMPONENTS ::= \WORDS + \REGS \]
-%
+
 A {\em machine state} is a map from components to values:
@@
 \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-%
-We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-\apt{But just below we write $\mach \stepstoobs \mach'$. Which is it?}
-\apt{Say something about initial machine state?}
+
+We have a total step function between machine states \(\mach \stepsto \mach'\).
 
@@
 
-We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
+We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
 A {\em policy} consists of a set of policy
@@
 can see the machine state, but cannot modify it. We assume that the
-compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.\apt{see comment above}
+compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.
 %
@@
 
-A trace \(T \in \TRACE{\TRACEELEMS}\) is a nonempty, finite or infinite sequence of elements
-of some type \(\TRACEELEMS\).  We will mostly be concerned with sequences of MPC-states,
+A trace over a type \(\TRACEELEMS\)\bcp{``over a type A'' is a Coq-ism} is a nonempty, finite or infinite sequence of
+\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
+\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
 called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
@@
 coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\), and potentially cut short by
-the policy component. % The {\sc Default} rule applies only if the other does not.
+produced by repeated application of step, and potentially cut short by
+the policy component. The {\sc Default} rule applies only if the other does not.
 (We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
 
-\begin{center}
 \begin{minipage}{.4\textwidth}
@@
 \begin{minipage}{.4\textwidth}
-\judgment%[Default]
-         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
+\judgment[Default]
+         {}
          {\(\mpcstatename \hookrightarrow \mpcstatename\)}
 \end{minipage}
-\end{center}
 %
@@
 The operation \(\PUT ~ f ~ \Trace\) takes a trace
-\(\Trace \in \TRACE{\TRACEELEMS %\times \OBSS
-}  \) and a predicate
+\(\Trace \in \TRACE{\TRACEELEMS \times \OBSS}\) and a predicate
 on elements \(f \subset \TRACEELEMS\) and gives the prefix of
@@
 %
-\apt{If we are writing \(f \subset \TRACEELEMS\), perhaps we should also
-  be writing \(a \in f\), etc.  rather than \(f a\), etc.?}
 If the resulting trace is a strict prefix of the parameter, then the
@@
 such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
-\apt{Unclear how there can be multiple initial states.}
 

### 2021-04-15 tex/main.tex
Oops
CP

@@
 control flow (like WBCF).  But our formulations of both properties are
-somewhat different from those of~\cite{Skorstengaard+19}.  Rather than
+somewhat different from those of Skorstengaard et al.  Rather than
 defining stack safety implicitly,
@@
 
-In applying these security concepts, we treat each caller as needing protection
-from its callee, even as that callee may need protection from its callee in turn.
+In formalizing these security concepts, we treat {\em each} caller as needing
+protection from its callee (which, in turn, needs protection from its
+callee, etc.).
 This is novel in that the

### 2021-04-15 tex/main.tex
subtrace -> sub-trace

@@
 This is novel in that the
-confidentiality property applies to many nested subtraces within a single execution,
+confidentiality property applies to many nested sub-traces within a single execution,
 as opposed to traces of entire executions.
@@
 {\em noninterference}-style property that, rather than comparing pairs of whole program traces,
-compares a pair of subtraces for each call, nested within other calls. Once again
+compares a pair of sub-traces for each call, nested within other calls. Once again
 {\em stepwise} describes the granularity of enforcement: at every step of a call,
@@
 
-In particular, we wish to formulate properties about consecutive subtraces
+In particular, we wish to formulate properties about consecutive sub-traces
 in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking subtraces in which the stack is
+individual calls from a trace by taking sub-traces in which the stack is
 at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''
+``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
 
@@
 is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
+and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a sub-trace
 such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).

### 2021-04-15 tex/main.tex
work on 6

--- 2021-04-15 tex/main.tex
@@
 LSE) and
-control flow (like WBCF).  But our formulations of both properties are
-somewhat different from those of~\cite{Skorstengaard+19}.  Rather than
-defining stack safety implicitly,
-as ``what a certain idealized machine does,'' we formalize explicit criteria
+control flow (with WBCF).  But rather than defining stack safety as ``what
+a certain idealized machine does,'' we formalize explicit criteria
 based on the technical framework of language-based
@@
 we can express safety in an information-flow sense without reference to it,
-even safety from attacks that hijack control flow.  \BCP{On consideration
+even safety from attacks that hijack control flow.  \bcp{On consideration
   (especially in light of the weird ``C confuses A by returning directly to
@@
 \[\codemap \in \CODEMAPS \subseteq \WORDS\]
+\apt{Doesn't match previous definition any more.}
 
@@
 this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
-maps a machine state to a set of addresses.
+maps a machine state to a set of addresses. \apt{Is it a relation or a function (seems to
+  be the latter)?}
 If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
@@
 it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
-which relates call states to the states that can be considered valid returns to their caller.
+which relates call states to the states that can be considered valid returns to that call.
 The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
@@
 But whatever the return convention, at each step we update the context according
-to the context-step function, \(\stepsto_C\).
-
-Each system component will be assigned to a {\em domain}, which may be {\em Outside}
+to the context-step function, \(\stepsto_C\).\apt{abrupt transition}
+
+Each system component is assigned to a {\em domain}, which may be {\em Outside}
 of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller will register a {\em target}
+at a particular depth. Each caller registers a {\em target}
 condition -- a predicate on machine states that must hold for it to consider itself active.
@@
 return.  \ifaftersubmission\bcp{spacing is funny}\fi
-
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-  \times (\rts \in \listT (\target \subseteq \MACHS)) \]
-
-An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
+%
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
+\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
+  \times (\listT (\target)) \]
+%
+The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
 the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
@@
 When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We may now proceed to state our properties.
-Our first property will be {\em stepwise stack integrity}, stepwise meaning that it will be
+call stack we are and which components are sealed. We now proceed to state our properties.
+The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is 
 quantified over each step of execution from any initial state. The criterion for integrity
@@
 
-\definition A system enjoys {\em stepwise stack integrity} if for any reachable state
-\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
-\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
-\pi_\mach(\mpcstatename') ~ \component\).
+\definition A system enjoys {\em stepwise stack integrity} if, for any reachable state
+\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
+and any component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
+\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename') ~ \component =
+\pi_\mach(\mpcstatename) ~ \component\).
 
@@
 
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if, for
-all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\). We write
-this \(\mach \approx_\components \nach\).
-
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change
-together if for all components \(\component\) such that
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+written \(\mach \approx_\components \nach\), if, for
+all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
+
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to
+\emph{change together}, written \(\mach,\mach' \diamond \nach,\nach'\), if,
+for all components \(\component\) such that
 \(\mach ~ \component \not = \mach' ~ \component\) or
@@
 \(\mach' ~ \component = \nach' ~ \component\).
-This is written \(\mach,\mach' \diamond \nach,\nach'\).
 
 \definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
-if their adjacent pairs change together. We will represent this relation with
+if their adjacent pairs of states change together. We will represent this relation with
 \(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
@@
 
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\definition For any depth \(\depth\), define \(f_\depth ~ \mach ~ (\_, \rts)\) to be 
+\(|\rts| \geq \depth\). Then each reachable segment \(\MPCT \langle f_\depth \rangle\) 
+corresponds to a trace of a function from its call to its return.
+Given such a segment \(S\), let \(\mach,\_,(\domMap,\_)\)  = \(\head(S)\) and let 
 \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
-\(\MPCT\) corresponds to the trace of a function from its call to its return.
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
-\(\MPCT \doteq \NPCT\).
+A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
+with associated \(m\) and \(\components\) as above, and 
+for all \(\nach\) such that \(\mach \approx_\components \nach\) and
+ \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\), 
+ \(\MPCT\langle f_\depth\rangle \doteq \NPCT\langle f_\depth\rangle \).\apt{I found the previous
+   version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
+%% \definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+%% \(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+%% let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+%%  \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
+%% \(\MPCT\) corresponds to the trace of a function from its call to its return.
+%% Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+%% \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+%% {\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
+%% \(\MPCT \doteq \NPCT\).
 
@@
 
-Example~\ref{fig:passingsimple} shows a sample trace of
+Example \ref{fig:passingsimple} shows a sample trace of
 a program that passes a value on the stack, without an attack.
@@
 %
-For example, the program in \cref{fig:passingsimple} manages the stack
-in this way. Here, the call pseudo-instruction to {\tt f} in line 5
-hides a blessed sequence of instructions that allocate space to save
-its return address to {\tt main} and seal the protected part of the
-stack before {\tt f}'s body starts executing at 10. Conversely, the
-return instruction in line 15 restores and jumps to the return address
-and deallocates its frame space, and unseals the corresponding parts
-of the stack. The function itself allocates and deallocates its own
-private frame space in instructions 10 and 14. Additionally, unlike in
-the simplified presentation of the policy, the caller shares the
-arguments passed to the callee through the stack, and for this creates
-and initializes (3--4) and destructs (6) the argument part of the
-callee's stack frame.
+The running example in \cref{fig:running-trace-a} manages the stack in this way.
+For example, {\tt f}'s ``blessed'' sequences allocate space to save its return
+address (and private variables) in instructions 20--21, and later restores that
+return address and deallocates its private frame space before returning in
+34--36. Additionally, unlike in the simplified presentation of the policy, the
+caller passes information to the callee through the stack, and for this creates
+and initializes (22 and 24--25) and destructs (29) the argument part of the
+callee's stack frame.\rb{The policy would also need to control this, I think it
+  is clear}
+\rb{TODO: References to the running example are no longer valid! We could use
+figures 5 or 6, referencing the figures, not the invisible assembly code.}
 %

### 2021-04-15 tex/main.tex

@@
 LSE) and
-control flow (with WBCF).  But rather than defining stack safety as ``what
-a certain idealized machine does,'' we formalize explicit criteria
+control flow (like WBCF).  But our formulations of both properties are
+somewhat different from those of Skorstengaard et al.  Rather than
+defining stack safety implicitly,
+as ``what a certain idealized machine does,'' we formalize explicit criteria
 based on the technical framework of language-based
@@
 
-In applying these security concepts, we treat each caller as needing protection
-from its callee, even as that callee may need protection from its callee in turn.
+In formalizing these security concepts, we treat {\em each} caller as needing
+protection from its callee (which, in turn, needs protection from its
+callee, etc.).
 This is novel in that the
-confidentiality property applies to many nested subtraces within a single execution,
+confidentiality property applies to many nested sub-traces within a single execution,
 as opposed to traces of entire executions.

### 2021-04-15 tex/main.tex
More polish / comments

--- 2021-04-15 tex/main.tex
@@
 
-In formalizing these security concepts, we treat {\em each} caller as needing
+\bcp{This para could be clearer.  (I can try to fix it later but I want to
+  keep reading now...)}In formalizing these security concepts, we treat {\em each} caller as needing
 protection from its callee (which, in turn, needs protection from its
@@
 as opposed to traces of entire executions.
-{\em Integrity} says that a caller's local data is protected from modification by its callee,
+{\em Integrity} says that each caller's local data is protected from
+modification by its callee(s),
 motivated by the reasoning principle that local data persists across a call. Its
-{\em stepwise} interpretation formalizes this very naturally: a caller's local data
-is never modified during a call. {\em Confidentiality} says that a caller's data is protected
-from being leaked into its callee, and also that the callee cannot see uninitialized data.
+{\em stepwise} interpretation\bcp{We're introducing ``stepwise'' in a pretty
+abrupt way---potentially confusing.} formalizes this very naturally: a caller's local data
+is never modified during a call. {\em Confidentiality} says that a caller's
+data is not leaked into its callee, and  that the callee cannot see uninitialized data.
 As a reasoning principle, confidentiality guarantees that the callee's
-behavior is determined only by a predictable interface that does not include the caller's
+behavior is determined only by a predictable interface (function arguments and
+register contents) that does not include the caller's
 local state or the uninitialized portion of the stack. This is formalized as a
 {\em noninterference}-style property that, rather than comparing pairs of whole program traces,
-compares a pair of sub-traces for each call, nested within other calls. Once again
-{\em stepwise} describes the granularity of enforcement: at every step of a call,
+compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
+again
+{\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
 changes in the state are independent of the caller's data.
 
-Together, these properties capture the rough intuition of LSE, but are somewhat stronger.
-First, we model returns as returns to a valid return target; an invalid return
+Together, these properties capture roughly the same intuition as LSE, but
+they are somewhat stronger.
+First, we model returns as returns to a valid return target\bcp{that will
+  make no sense to the reader at this point}; an invalid return
 is treated as a mere jump within the callee, and does not grant access to the caller's
@@
 The attacker in such an instance is permitted to ``return'' so long as it avoids
-violating the integrity and confidentiality of the caller. Second, confidentiality
-contains temporal protections that LSE does not necessarily address. The intuition
+violating the integrity and confidentiality of the caller.\bcp{ditto} Second, confidentiality
+contains\bcp{?} temporal protections that LSE does not necessarily address\bcp{?}. The intuition
 for LSE is that it protects the caller's stack frame, not necessarily memory that was
@@
 it, but because it is a potential vector for leaking capabilities into the caller's frame,
-not as a goal of the property.)
-
-To illustrate the applicability of these properties, we perform randomized testing to
+not as a goal of the property.\bcp{Oof.  We need to do better than that!})
+
+To illustrate\bcp{IMO, this is not just an illustration; validating and
+  debugging a published enforcement mechanism is a top-level contribution.}
+the applicability of these properties, we perform randomized testing to
 validate that an existing enforcement mechanism does enforce them.
@@
 
-One challenge in characterizing stack safety formally is that in a low-level machine
-without built-in stacks, we are at the mercy of architectural
+One challenge in characterizing stack safety formally is that, in a low-level machine
+without built-in stacks, we are at the mercy\bcp{?} of architectural\bcp{?}
 choices that do not relate to the abstract notion of stack safety. Some implementations
@@
 
-A second challenge is that enforcement mechanisms may weaken their protections
+A second challenge is that enforcement mechanisms may weaken their
+protections\bcp{``weaken their protections'' doesn't convey what we mean
+  very well}
 in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
@@
 
+\bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
 What does ``affecting program behavior'' really mean? This micro-policy
@@
 
+\bcp{Ditto, even moreso.}
 To formalize this intuition, we develop an \emph{observational integrity} property which
@@
 {\em observational stack safety} that is highly extensional.
-We therefore argue that, despite its somewhat greater complexity,
-it may be a better minimal specification for
+We argue that, despite its somewhat greater complexity,
+it may be a better specification for
 stack safety enforcement mechanisms in general.
@@
 We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
-confidentiality, albeit at some performance cost.
+confidentiality, albeit at some performance cost.  \bcp{Did Nick and Andre
+  know about this deficiency?  Did they mention it in their paper?  (If
+  neither, we should make a bigger deal out of the fact that we discovered it.)}
 
@@
 %% %efficient, but our variant is {\em not} efficient!}
-
 In summary, we offer the following contributions:
+
 \begin{itemize}
@@
 (\cref{sec:running-example}). Then, after some technical preliminaries
-(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
-and describe testing (\cref{sec:testing}).
+(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}\bcp{??}, \cref{sec:lazy})
+and describe testing (\cref{sec:testing})\bcp{The sections are out of order?}.
 \Cref{sec:relwork} discusses related work, and
-\cref{sec:future} sketches directions for future work.
-
-\section{Setup}
+\cref{sec:future} sketches directions for future work.  \bcp{Where is 7?}
+
+\section{Set Up}
 \label{sec:setup}
@@
 control flow and observing when we reach a {\em return target}. For a given state annotated
-as a call, in a standard calling convention, another state fulfills its return target if
+as a call, in a standard calling convention, another state fulfils its return target if
 it has the same stack pointer and the program counter is at the instruction following
@@
 
-Intuitively, stack safety means that the stack frame of each caller (and its
-those of its ancestors) is protected
-from attempts by a confused or malicious callee (including all its descendants)
+Intuitively, stack safety means that the stack frame of each caller (and 
+those of its ancestor callers) is protected
+from attempts by a confused or malicious callee (including all its
+descendant callees)
 to read or write its private data.
-The attacker model therefore treats every caller as a defender and its callee as
+That is, the attacker model treats every caller as a defender and its callee as
 an attacker, meaning that with nested calls, a single function invocation
@@
 
-We take a very strong threat model in which both caller and callee may execute
+We adopt a strong threat model in which both caller and callee may execute
 arbitrary machine code. Hardware and timing attacks are out of scope.
@@
 \paragraph*{Well-formedness and Bad Annotations}
+\bcp{What is ``well-formedness''??}
 
@@
 
-Since we rely heavily on call annotations, we must ask: what if the annotations
+Since we rely on call annotations\bcp{to...}, we must ask: what if the annotations
 are wrong? First, suppose a call occurs that is not annotated.
@@
 
-Alternatively, an annotated instruction might not obey the calling convention.
+\bcp{Confusing:}Alternatively, an annotated instruction might not obey the calling convention.
 In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
@@
 \end{subfigure}
-\caption{An Integrity Violation}
+\caption{An Integrity Violation\bcp{Can we align the C-like snippets with
+    the corresponding assembly?}  \ifaftersubmission\bcp{I wonder if we
+    could make the figures easier to read by putting very light, differently
+  colored backgrounds under the different parts}\fi}
 \label{fig:int1}
@@
 
-This section will elucidate the intuition behind the properties through a series
+This section introduces the intuition behind the properties through a series
 of examples. Existing stack safety enforcement work mostly refers directly to examples

### 2021-04-15 tex/main.tex
CP

@@
 
-We begin with a description of our assumptions and threat model (\cref{sec:threat}),
+We begin with a description of our setting (\cref{sec:setup}) and threat
+model (\cref{sec:threat}),
 then show how our properties apply to examples of stack unsafety
@@
 
-\section{Set Up}
+\section{Setting}
 \label{sec:setup}
@@
 of examples. Existing stack safety enforcement work mostly refers directly to examples
-of bad behavior that it prevents; we will argue that:
+of bad behavior that it prevents; we will argue that: \bcp{Not sure this
+  enumeration is useful.}
 
@@
 local variables are stored in the stack frame.
-Unless otherwise stated, arguments and return values are passed in registers.
+For the moment, arguments and return values are passed in registers.
 The machine is word-addressed. The stack grows upward in
@@
 At the right of each example, we show the layout of stack memory just before
-the specified instruction is executed; each box represents a memory word and
+the specified instruction is executed; each box represents a memory word, and
 addresses increase to the right.
 
-Unless otherwise noted, colored boxes indicate an assignment of the address to a
+Unless otherwise noted\bcp{where is it ``otherwise noted''??}, colored boxes
+indicate an assignment of the address to a
 {\em domain} -- a label reflecting a particular set of access rights. Domains
-are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
+are \(\unsealed\)\bcp{``unsealed'' would read better} (in the stack and accessible to writes), \(\sealed{\depth}\)
 (in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
@@
 
-Meanwhile confidentiality intuitively states that a callee is unaware
+Meanwhile, confidentiality intuitively states that a callee is unaware\bcp{??}
 of the context of its call (except for its arguments and any global
@@
 shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
-state with arbitrary values \(v_0\), etc., in place of the values that should
+state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
+understand if we chose concrete values for the examples}\fi \(v_0\), etc., in place of the values that should
 be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
@@
 the program counters will differ, breaking the lockstep.\apt{Is there any reason
-to complicate things like this by bringing in the PC?}
+to complicate things like this by bringing in the PC?}\bcp{+1}
 
@@
 of matching sets calls and returns nested within on another.
-Rather, they express what it means to protect data even in the presence of control-flow attacks.
+Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
 Consider example \ref{fig:controlflow}.
@@
 
+\vspace*{-1ex}
 \judgmentthree[ReturnFound]
@@
 
+\vspace*{-3.5ex}
 \judgment[Default]

### 2021-04-15 tex/main.tex
control-flow, and capitalization

@@
 is treated as a mere jump within the callee, and does not grant access to the caller's
-data. So, even without WBCF, our stack safety properties rule out many control flow attacks.
+data. So, even without WBCF, our stack safety properties rule out many control-flow attacks.
 The attacker in such an instance is permitted to ``return'' so long as it avoids
@@
 \end{subfigure}
-\caption{An Integrity Violation\bcp{Can we align the C-like snippets with
+\caption{An integrity violation\bcp{Can we align the C-like snippets with
     the corresponding assembly?}  \ifaftersubmission\bcp{I wonder if we
@@
 \end{subfigure}
-\caption{A Confidentiality Violation\apt{Could save space by not repeating C or assemly code for {\tt main}.}}
+\caption{A confidentiality violation\apt{Could save space by not repeating C or assemly code for {\tt main}.}}
 \label{fig:conf1}
@@
 
-\caption{Another Confidentiality Violation}
+\caption{Another confidentiality violation}
 \label{fig:conf2}
@@
 
-\paragraph*{Control Flow Attacks}
+\paragraph*{Control-Flow Attacks}
 \begin{figure}
@@
   \end{subfigure}
-  \caption{A Control Flow Attack.}
+  \caption{A control-flow attack}
   \label{fig:controlflow}
@@
 
-\caption{Sharing on the Stack}
+\caption{Sharing on the stack}
 \label{fig:passingsimple}

### 2021-04-15 tex/main.tex
More polishing / comments

@@
 
-\caption{Another confidentiality violation}
+\caption{Another confidentiality violation \bcp{Again, let's align the C
+    code with the assembly, etc.}}
 \label{fig:conf2}
@@
   \end{subfigure}
-  \caption{A control-flow attack}
+  \caption{A control-flow attack \bcp{wonder if we can cram the bottom part
+      onto the same row as the rest.  Might need to make things smaller,
+      spill into the margin, etc., but we really need the space.  (Another
+      option is to make it a separate figure and flow text around it, but
+      this will save less space.)}}
   \label{fig:controlflow}
@@
 
-We next define our machine model precisely. To make our definition of stack safety
+We now define our machine model precisely. To make our definition of stack safety
 as generic as possible, this section proposes an abstract interface to a
@@
 
-\caption{Sharing on the stack}
+\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
+    the figure (at least in the latex build that I'm looking at...)}\fi}
 \label{fig:passingsimple}

### 2021-04-15 tex/main.tex

@@
 %  to leave it out of the title but hit it strongly in the abstract.}%
-What does ``stack safety'' mean, exactly? The phrase is associated with a
+What does ``stack safety'' mean? The phrase is associated with a
 variety of compiler,
@@
 memory.  But these mechanisms typically lack precise specifications,
-relying instead on informal descriptions and specific examples of bad
+relying instead on informal descriptions and examples of bad
 behaviors that they prevent.
 
-We propose formal criteria for stack safety, using the concepts of
-language-based security: a combination of an integrity property (``the private
+We propose a formal characterization
+of stack safety, phrased using concepts from language-based security: a
+combination of an integrity property (``the private
 state in each caller's stack frame is held invariant by the callee'')
 and a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state'').
+caller's private state'')\bcp{and something about WBCF...}.
 % These properties are sensible even in the presence of
@@
 %
-We argue that these criteria match common intuitions by (1) showing that
-they are violated by a variety of intuitively unsafe examples and (2)
-using property-based random testing to check that appropriate variants of
-the criteria are actually enforced by the stack-safety ``micro-policies''
-proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
-broken variants).  Specifically, Roessler
+We use our characterization to validate the stack-safety ``micro-policies''
+proposed by~\citet{DBLP:conf/sp/RoesslerD18}.
+Specifically, we use property-based random testing to check that Roessler
 and Dehon's ``eager'' micro-policy, which catches violations as early as
-possible, enforces a basic ``stepwise'' versions of our confidentiality
-and integrity properties. Their more performant ``lazy'' micro-policy
-does not, but it does (almost) correspond to weaker, more extensional
-variants of the properties.
-%
-\BCP{I still don't think the next sentence really nails it.  And (sigh) I am
-  still not really clear yet on whether our definition applies / might apply
-  to other enforcement mechanisms or not.  Saying that our attacker model is
-  stronger does not help---that should make it easier for us to describe
-  their protections, not harder.  So I don't understand what we're saying
-  here.}%
-%
-While we do not rule out other enforcement mechanisms, our
-definitions are strongly informed by micropolicy systems, especially
-our attacker model, which is stronger than other enforcement mechanisms
-can enable.
-
-\bcp{We should decide between ``criteria'' and ``properties''...}
+possible, enforces a simple ``stepwise'' variant of our properties (and
+detects several broken variants), and that a slightly repaired version of
+their more performant ``lazy'' micro-policy corresponds to weaker and more
+extensional variants of the properties.
+%
+% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
+%   still not really clear yet on whether our definition applies / might apply
+%   to other enforcement mechanisms or not.  Saying that our attacker model is
+%   stronger does not help---that should make it easier for us to describe
+%   their protections, not harder.  So I don't understand what we're saying
+%   here.}%
+% %
+% While we do not rule out other enforcement mechanisms, our
+% definitions are strongly informed by micropolicy systems, especially
+% our attacker model, which is stronger than other enforcement mechanisms
+% can enable.
 
@@
 defining stack safety implicitly,
-as ``what a certain idealized machine does,'' we formalize explicit criteria
+as ``what a certain idealized machine does,'' we formalize explicit properties
 based on the technical framework of language-based
@@
 
-Our stack-safety criteria depend
+Our stack-safety properties depend
 critically on the notions of ``call'' and ``return,'' which are not fully explicit
@@
 
-\bcp{Confusing:}Alternatively, an annotated instruction might not obey the calling convention.
+\bcp{Confusing, and maybe out of date?}Alternatively, an annotated instruction might not obey the calling convention.
 In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls

### 2021-04-15 tex/main.tex
Rework of formal preliminaries

--- 2021-04-15 tex/main.tex
@@
 LSE) and
-control flow (like WBCF).  But our formulations of both properties are
-somewhat different from those of Skorstengaard et al.  Rather than
-defining stack safety implicitly,
-as ``what a certain idealized machine does,'' we formalize explicit criteria
+control flow (with WBCF).  But rather than defining stack safety as ``what
+a certain idealized machine does,'' we formalize explicit criteria
 based on the technical framework of language-based
@@
 
-\bcp{This para could be clearer.  (I can try to fix it later but I want to
-  keep reading now...)}In formalizing these security concepts, we treat {\em each} caller as needing
-protection from its callee (which, in turn, needs protection from its
-callee, etc.).
+In applying these security concepts, we treat each caller as needing protection
+from its callee, even as that callee may need protection from its callee in turn.
 This is novel in that the
-confidentiality property applies to many nested sub-traces within a single execution,
+confidentiality property applies to many nested subtraces within a single execution,
 as opposed to traces of entire executions.
-{\em Integrity} says that each caller's local data is protected from
-modification by its callee(s),
+{\em Integrity} says that a caller's local data is protected from modification by its callee,
 motivated by the reasoning principle that local data persists across a call. Its
-{\em stepwise} interpretation\bcp{We're introducing ``stepwise'' in a pretty
-abrupt way---potentially confusing.} formalizes this very naturally: a caller's local data
-is never modified during a call. {\em Confidentiality} says that a caller's
-data is not leaked into its callee, and  that the callee cannot see uninitialized data.
+{\em stepwise} interpretation formalizes this very naturally: a caller's local data
+is never modified during a call. {\em Confidentiality} says that a caller's data is protected
+from being leaked into its callee, and also that the callee cannot see uninitialized data.
 As a reasoning principle, confidentiality guarantees that the callee's
-behavior is determined only by a predictable interface (function arguments and
-register contents) that does not include the caller's
+behavior is determined only by a predictable interface that does not include the caller's
 local state or the uninitialized portion of the stack. This is formalized as a
 {\em noninterference}-style property that, rather than comparing pairs of whole program traces,
-compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
-again
-{\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
+compares a pair of subtraces for each call, nested within other calls. Once again
+{\em stepwise} describes the granularity of enforcement: at every step of a call,
 changes in the state are independent of the caller's data.
 
-Together, these properties capture roughly the same intuition as LSE, but
-they are somewhat stronger.
-First, we model returns as returns to a valid return target\bcp{that will
-  make no sense to the reader at this point}; an invalid return
+Together, these properties capture the rough intuition of LSE, but are somewhat stronger.
+First, we model returns as returns to a valid return target; an invalid return
 is treated as a mere jump within the callee, and does not grant access to the caller's
-data. So, even without WBCF, our stack safety properties rule out many control-flow attacks.
+data. So, even without WBCF, our stack safety properties rule out many control flow attacks.
 The attacker in such an instance is permitted to ``return'' so long as it avoids
-violating the integrity and confidentiality of the caller.\bcp{ditto} Second, confidentiality
-contains\bcp{?} temporal protections that LSE does not necessarily address\bcp{?}. The intuition
+violating the integrity and confidentiality of the caller. Second, confidentiality
+contains temporal protections that LSE does not necessarily address. The intuition
 for LSE is that it protects the caller's stack frame, not necessarily memory that was
@@
 it, but because it is a potential vector for leaking capabilities into the caller's frame,
-not as a goal of the property.\bcp{Oof.  We need to do better than that!})
-
-To illustrate\bcp{IMO, this is not just an illustration; validating and
-  debugging a published enforcement mechanism is a top-level contribution.}
-the applicability of these properties, we perform randomized testing to
+not as a goal of the property.)
+
+To illustrate the applicability of these properties, we perform randomized testing to
 validate that an existing enforcement mechanism does enforce them.
@@
 
-One challenge in characterizing stack safety formally is that, in a low-level machine
-without built-in stacks, we are at the mercy\bcp{?} of architectural\bcp{?}
+One challenge in characterizing stack safety formally is that in a low-level machine
+without built-in stacks, we are at the mercy of architectural
 choices that do not relate to the abstract notion of stack safety. Some implementations
@@
 
-A second challenge is that enforcement mechanisms may weaken their
-protections\bcp{``weaken their protections'' doesn't convey what we mean
-  very well}
+A second challenge is that enforcement mechanisms may weaken their protections
 in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
@@
 
-\bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
 What does ``affecting program behavior'' really mean? This micro-policy
@@
 
-\bcp{Ditto, even moreso.}
 To formalize this intuition, we develop an \emph{observational integrity} property which
@@
 {\em observational stack safety} that is highly extensional.
-We argue that, despite its somewhat greater complexity,
-it may be a better specification for
+We therefore argue that, despite its somewhat greater complexity,
+it may be a better minimal specification for
 stack safety enforcement mechanisms in general.
@@
 We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
-confidentiality, albeit at some performance cost.  \bcp{Did Nick and Andre
-  know about this deficiency?  Did they mention it in their paper?  (If
-  neither, we should make a bigger deal out of the fact that we discovered it.)}
+confidentiality, albeit at some performance cost.
 
@@
 %% %efficient, but our variant is {\em not} efficient!}
+
 In summary, we offer the following contributions:
-
 \begin{itemize}
@@
 
-We begin with a description of our setting (\cref{sec:setup}) and threat
-model (\cref{sec:threat}),
+We begin with a description of our assumptions and threat model (\cref{sec:threat}),
 then show how our properties apply to examples of stack unsafety
 (\cref{sec:running-example}). Then, after some technical preliminaries
-(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}\bcp{??}, \cref{sec:lazy})
-and describe testing (\cref{sec:testing})\bcp{The sections are out of order?}.
+(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
+and describe testing (\cref{sec:testing}).
 \Cref{sec:relwork} discusses related work, and
-\cref{sec:future} sketches directions for future work.  \bcp{Where is 7?}
-
-\section{Setting}
+\cref{sec:future} sketches directions for future work.
+
+\section{Setup}
 \label{sec:setup}
@@
 control flow and observing when we reach a {\em return target}. For a given state annotated
-as a call, in a standard calling convention, another state fulfils its return target if
+as a call, in a standard calling convention, another state fulfills its return target if
 it has the same stack pointer and the program counter is at the instruction following
@@
 
-Intuitively, stack safety means that the stack frame of each caller (and
-those of its ancestor callers) is protected
-from attempts by a confused or malicious callee (including all its
-descendant callees)
+Intuitively, stack safety means that the stack frame of each caller (and its
+those of its ancestors) is protected
+from attempts by a confused or malicious callee (including all its descendants)
 to read or write its private data.
-That is, the attacker model treats every caller as a defender and its callee as
+The attacker model therefore treats every caller as a defender and its callee as
 an attacker, meaning that with nested calls, a single function invocation
@@
 
-We adopt a strong threat model in which both caller and callee may execute
+We take a very strong threat model in which both caller and callee may execute
 arbitrary machine code. Hardware and timing attacks are out of scope.
@@
 \paragraph*{Well-formedness and Bad Annotations}
-\bcp{What is ``well-formedness''??}
 
@@
 
-Since we rely on call annotations\bcp{to...}, we must ask: what if the annotations
+Since we rely heavily on call annotations, we must ask: what if the annotations
 are wrong? First, suppose a call occurs that is not annotated.
@@
 
-\bcp{Confusing:}Alternatively, an annotated instruction might not obey the calling convention.
+Alternatively, an annotated instruction might not obey the calling convention.
 In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
@@
 \end{subfigure}
-\caption{An integrity violation\bcp{Can we align the C-like snippets with
-    the corresponding assembly?}  \ifaftersubmission\bcp{I wonder if we
-    could make the figures easier to read by putting very light, differently
-  colored backgrounds under the different parts}\fi}
+\caption{An Integrity Violation}
 \label{fig:int1}
@@
 
-This section introduces the intuition behind the properties through a series
+This section will elucidate the intuition behind the properties through a series
 of examples. Existing stack safety enforcement work mostly refers directly to examples
-of bad behavior that it prevents; we will argue that: \bcp{Not sure this
-  enumeration is useful.}
+of bad behavior that it prevents; we will argue that:
 
@@
 local variables are stored in the stack frame.
-For the moment, arguments and return values are passed in registers.
+Unless otherwise stated, arguments and return values are passed in registers.
 The machine is word-addressed. The stack grows upward in
@@
 At the right of each example, we show the layout of stack memory just before
-the specified instruction is executed; each box represents a memory word, and
+the specified instruction is executed; each box represents a memory word and
 addresses increase to the right.
 
-Unless otherwise noted\bcp{where is it ``otherwise noted''??}, colored boxes
-indicate an assignment of the address to a
+Unless otherwise noted, colored boxes indicate an assignment of the address to a
 {\em domain} -- a label reflecting a particular set of access rights. Domains
-are \(\unsealed\)\bcp{``unsealed'' would read better} (in the stack and accessible to writes), \(\sealed{\depth}\)
+are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
 (in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
@@
 \end{subfigure}
-\caption{A confidentiality violation\apt{Could save space by not repeating C or assemly code for {\tt main}.}}
+\caption{A Confidentiality Violation\apt{Could save space by not repeating C or assemly code for {\tt main}.}}
 \label{fig:conf1}
@@
 
-Meanwhile, confidentiality intuitively states that a callee is unaware\bcp{??}
+Meanwhile confidentiality intuitively states that a callee is unaware
 of the context of its call (except for its arguments and any global
@@
 shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
-state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
-understand if we chose concrete values for the examples}\fi \(v_0\), etc., in place of the values that should
+state with arbitrary values \(v_0\), etc., in place of the values that should
 be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
@@
 
-\caption{Another confidentiality violation}
+\caption{Another Confidentiality Violation}
 \label{fig:conf2}
@@
 the program counters will differ, breaking the lockstep.\apt{Is there any reason
-to complicate things like this by bringing in the PC?}\bcp{+1}
-
-\paragraph*{Control-Flow Attacks}
+to complicate things like this by bringing in the PC?}
+
+\paragraph*{Control Flow Attacks}
 \begin{figure}
@@
   \end{subfigure}
-  \caption{A control-flow attack}
+  \caption{A Control Flow Attack.}
   \label{fig:controlflow}
@@
 of matching sets calls and returns nested within on another.
-Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
+Rather, they express what it means to protect data even in the presence of control-flow attacks.
 Consider example \ref{fig:controlflow}.
@@
 
-The basic building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\).
+The building blocks of the machine are {\em values} and {\em addresses}.
+Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
+\(\addr\), respectively.
 %
 Our machine states are composed of {\em components} \(\component\),
-which are either memory addresses (words) or register names (\(\reg\)) drawn from some set
+which are either addresses or register names (\(\reg\)) drawn from some set
 \(\REGS\), which is assumed to include two special purpose registers: the
 program counter {\PCname} and the stack pointer \(\SP\).
+A {\em machine state} is a map from components to values.
 %
@@
 %
-A {\em machine state} is a map from components to values:
-%
     \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
@@
 We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-\apt{But just below we write $\mach \stepstoobs \mach'$. Which is it?}
-\apt{Say something about initial machine state?}
-
-\subsection{Policies}
 
@@
 it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. We assume that the
-compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.\apt{see comment above}
+can see the machine state, but cannot modify it. The policy also provides a
+well-formedness condition, \(\WF (\mpstatename \in \MACHS \times \POLS)\), describing
+initial states that are properly configured.
 %
@@
 
-We write $\MPS$ for the set of all pairs of machine states and policy
-states.
-%
-\[\mpstatename \in \MPS = \MACHS \times \POLS\]
-%
 We lift the policy step function to operate on such pairs
 by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepstoobs \mach'\)}{\((\mach, \pol)
+%
+\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
                \polstep \pol'\)}
@@
                \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-This combined step function is also partial, and traces of the machine-with-policy
-will be finite iff a policy fault occurs.
+%
+Note that the base machine has a total step function, and the policy-enhanced machine
+refines it as a partial step. The policy state never changes behavior, but can cause
+premature termination, which represents failstop behavior.
 
@@
 
-Separate from the machine state, our formal properties will keep track of additional context
-information representing the history of the run. The following definitions are parameterized
-over an arbitrary\bcp{why?} set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+Separate from the machine state, we will keep track of additional context
+information representing the history of the run. This has no influence on execution, it merely
+describes information relevant to the particular property. Different information is needed
+for different versions of the model, so the definitions are parameterized
+over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
 an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context, \anns \constep \context' \in \MACHS \times \CONTEXTS
-\times \powerset(\ANNS) \rightarrow \CONTEXTS\). We also assume a set of function identifiers
-\(\FUNIDS\) and the existence of a \emph{code map} \(\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\)
-that relates each code address to its owning function and set of annotations.
+\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
+\times \rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
+\(\codemap \in \WORDS \rightharpoonup \ANNS\)
+that relates each code address to its annotation, if any.
 We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).  \bcp{I don't get the intuition for this whole para.}
-
-\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
-              {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
-              {\(\mach, \context, \anns \constep \context'\)}
+a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
+\CONTEXTS\).
+
+\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
+              {\(\mach, \context \constep \context'\)}
               {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
@@
 
-\subsection{Traces}
+\subsection{Machine Traces}
 \label{sec:traces}
 
-A trace \(T \in \TRACE{\TRACEELEMS}\) is a nonempty, finite or infinite sequence of elements
-of some type \(\TRACEELEMS\).  We will mostly be concerned with sequences of MPC-states,
-called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
-
-\paragraph*{Trace-Of}
+A {\em machine trace} is a nonempty, finite or infinite sequence of elements
+of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
+We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
 
@@
 coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\), and potentially cut short by
-the policy component. % The {\sc Default} rule applies only if the other does not.
-(We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
+produced by repeated application of \(\stepstocon\)
 
@@
 
-\paragraph*{Projections}
-
 We project out the machine state of an MPC-state with \(\pi_\mach\), the
-policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
-these operators to project out of an machine-trace a trace of machine states,
-policy states, and contexts, respectively.
-
-\paragraph*{Head and Last}
-
-We take the first element of a trace with \(\head(\Trace)\), which is a total
+policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
+We take the first element of a trace with \(\head(\MPCT)\), which is a total
 function since traces are non-empty, and the final element (if one exists) with
-\(\last(\Trace)\), which is partial.
+\(\last(\MPCT)\), which is partial.
 
 \paragraph*{Until}
-The operation \(\PUT ~ f ~ \Trace\) takes a trace
-\(\Trace \in \TRACE{\TRACEELEMS %\times \OBSS
-}  \) and a predicate
-on elements \(f \subset \TRACEELEMS\) and gives the prefix of
-\(\Trace\) ending with the first element on which \(f\) holds.
+The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
+and a predicate on states \(f ~ \mpcstate \in \MPCS\) and gives the prefix of
+\(\MPCT\) ending with the first element on which \(f\) holds.
 
@@
   \begin{minipage}{.3\textwidth}
-    \judgment{\(f ~ \traceelem\)}
-             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
+    \judgment{\(f ~ \mpcstatename\)}
+             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
   \end{minipage}
@@
     \judgment{}
-             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
+             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
 %
@@
   \begin{minipage}{.3\textwidth}
-    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
-                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
+    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
+                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
   \end{minipage}
@@
 \(\PUT\) is the identity function.
-
-% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
-%$\mathit{split}$: it takes two traces and links them by replacing the
-%last element of the first (if it has one) with the first element of
-%the second
-
-%\judgment{} {\(t \cdot T = T\)}
-%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
-%\leo{Lemma/theorem: split/join = id?}
-%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
 
@@
 \hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:
-
+running \(\mpcstatename\) up to \(f\)''):
+%
 \begin{center}
@@
 
-In particular, we wish to formulate properties about consecutive sub-traces
+In particular, we wish to formulate properties about consecutive subtraces
 in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking sub-traces in which the stack is
-at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
+individual calls from a trace by taking subtraces in which the stack is
+at or below a given depth. If \(f\) is still a predicate on states, then
+we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\):
 
@@
   \judgmenttwo[Now]
-              {\(f ~ \mach ~ \context\)}
-              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
-              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
+              {\(f ~ \mpcstatename\)}
+              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
+              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
 %
   \judgmentthree[Skip]
-                {\(f ~ \mach ~ \context\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
+                {\(f ~ \mpcstatename\)}
+                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
                 {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
 %
   \judgmentthree[Search]
-                {\(\neg f ~ \mach ~ \context\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
+                {\(\neg f ~ \mpcstatename\)}
+                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
                 {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
 \end{center}
@@
 properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a sub-trace
-such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
-\apt{Unclear how there can be multiple initial states.}
+is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
+where \(\WF ~ \mach,\pol\), \(\mpcstatename \hookrightarrow \MPCT\)
+and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
+such that from some similarly well-formed initial state \(\mpcstatename\),
+\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
 
@@
 \[\codemap \in \CODEMAPS \subseteq \WORDS\]
-\apt{Doesn't match previous definition any more.}
 
@@
 this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
-maps a machine state to a set of addresses. \apt{Is it a relation or a function (seems to
-  be the latter)?}
+maps a machine state to a set of addresses.
 If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
@@
 it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
-which relates call states to the states that can be considered valid returns to that call.
+which relates call states to the states that can be considered valid returns to their caller.
 The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
@@
 But whatever the return convention, at each step we update the context according
-to the context-step function, \(\stepsto_C\).\apt{abrupt transition}
-
-Each system component is assigned to a {\em domain}, which may be {\em Outside}
+to the context-step function, \(\stepsto_C\).
+
+Each system component will be assigned to a {\em domain}, which may be {\em Outside}
 of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller registers a {\em target}
+at a particular depth. Each caller will register a {\em target}
 condition -- a predicate on machine states that must hold for it to consider itself active.
@@
 return.  \ifaftersubmission\bcp{spacing is funny}\fi
-%
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-  \times (\listT (\target)) \]
-%
-The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
+
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+  \times (\rts \in \listT (\target \subseteq \MACHS)) \]
+
+An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
 the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
@@
 
-\vspace*{-1ex}
 \judgmentthree[ReturnFound]
@@
 
-\vspace*{-3.5ex}
 \judgment[Default]
@@
 When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We now proceed to state our properties.
-The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
+call stack we are and which components are sealed. We may now proceed to state our properties.
+Our first property will be {\em stepwise stack integrity}, stepwise meaning that it will be
 quantified over each step of execution from any initial state. The criterion for integrity
@@
 
-\definition A system enjoys {\em stepwise stack integrity} if, for any reachable state
-\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
-\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename') ~ \component =
-\pi_\mach(\mpcstatename) ~ \component\).
+\definition A system enjoys {\em stepwise stack integrity} if for any reachable state
+\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
+and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
+\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
+\pi_\mach(\mpcstatename') ~ \component\).
 
@@
 
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
-
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to
-\emph{change together}, written \(\mach,\mach' \diamond \nach,\nach'\), if,
-for all components \(\component\) such that
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if, for
+all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\). We write
+this \(\mach \approx_\components \nach\).
+
+\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to change
+together if for all components \(\component\) such that
 \(\mach ~ \component \not = \mach' ~ \component\) or
@@
 \(\mach' ~ \component = \nach' ~ \component\).
+This is written \(\mach,\mach' \diamond \nach,\nach'\).
 
 \definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
-if their adjacent pairs of states change together. We will represent this relation with
+if their adjacent pairs change together. We will represent this relation with
 \(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
@@
 
-\definition For any depth \(\depth\), define \(f_\depth ~ \mach ~ (\_, \rts)\) to be
-\(|\rts| \geq \depth\). Then each reachable segment \(\MPCT \langle f_\depth \rangle\)
-corresponds to a trace of a function from its call to its return.
-Given such a segment \(S\), let \(\mach,\_,(\domMap,\_)\)  = \(\head(S)\) and let
+\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
 \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
-A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
-with associated \(m\) and \(\components\) as above, and
-for all \(\nach\) such that \(\mach \approx_\components \nach\) and
- \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\),
- \(\MPCT\langle f_\depth\rangle \doteq \NPCT\langle f_\depth\rangle \).\apt{I found the previous
-   version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
-%% \definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-%% \(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-%% let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-%%  \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
-%% \(\MPCT\) corresponds to the trace of a function from its call to its return.
-%% Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-%% \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-%% {\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
-%% \(\MPCT \doteq \NPCT\).
+\(\MPCT\) corresponds to the trace of a function from its call to its return.
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT | \neg f_\depth \). A system enjoys
+{\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
+\(\MPCT \doteq \NPCT\).
 
@@
 
-\caption{Sharing on the stack}
+\caption{Sharing on the Stack}
 \label{fig:passingsimple}
@@
 %
-\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
+\[\mach \stepsto[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
 %

### 2021-04-15 tex/main.tex

--- 2021-04-15 tex/main.tex
@@
 
-In particular, we wish to formulate properties about consecutive sub-traces
-in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking sub-traces in which the stack is
-at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
+individual calls from a trace by taking subtraces in which the stack is
+at or below a given depth. If \(f\) is still a predicate on states, then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
@@
 A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
-with associated \(m\) and \(\components\) as above, and
+with associated \(\mach\) and \(\components\) as above, and
 for all \(\nach\) such that \(\mach \approx_\components \nach\) and
- \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\),
- \(\MPCT\langle f_\depth\rangle \doteq \NPCT\langle f_\depth\rangle \).\apt{I found the previous
+ \(\nach \hookrightarrow \NPCT\),
+ \(\MPCT\langle f_\depth\rangle \doteq \NPCT | \neg f_\depth \).\apt{I found the previous
    version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
-%% \definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-%% \(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-%% let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-%%  \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
-%% \(\MPCT\) corresponds to the trace of a function from its call to its return.
-%% Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-%% \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-%% {\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
-%% \(\MPCT \doteq \NPCT\).
+ \sna{It's right, but I need to rethink how the M<f> notation works...}
 
@@
 Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
 {\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
@@
 Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
 {\em stepwise coroutine confidentiality} in the coroutine setting if for all
@@
 Let \(\nach\) be a variant such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\).
+\(\nach \hookrightarrow \NPCT | \neg f_\depth \).
 

### 2021-04-15 tex/main.tex
More work on abstract and intro

@@
 We propose a formal characterization
-of stack safety, phrased using concepts from language-based security: a
+of stack safety, formulated with concepts from language-based security: a
 combination of an integrity property (``the private
 state in each caller's stack frame is held invariant by the callee'')
-and a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state'')\bcp{and something about WBCF...}.
+a confidentiality property (``the callee's behavior is insensitive to the
+caller's private state''), and a well-bracketedness property (``each
+callee returns control to its immediate caller'').
 % These properties are sensible even in the presence of
@@
 %
-We use our characterization to validate the stack-safety ``micro-policies''
-proposed by~\citet{DBLP:conf/sp/RoesslerD18}.
-Specifically, we use property-based random testing to check that Roessler
-and Dehon's ``eager'' micro-policy, which catches violations as early as
-possible, enforces a simple ``stepwise'' variant of our properties (and
-detects several broken variants), and that a slightly repaired version of
-their more performant ``lazy'' micro-policy corresponds to weaker and more
-extensional variants of the properties.
+We use these properties to validate the stack-safety ``micro-policies''
+proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
+property-based random testing) that Roessler and Dehon's ``eager''
+micro-policy, which catches violations as early as possible, enforces a
+simple ``stepwise'' variant of our properties and correctly detects several
+broken variants, and that (a repaired version of) their more performant
+``lazy'' micro-policy corresponds to a slightly weaker and more extensional
+variant of our properties.
 %
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-The most abstract characterization stack safety that we are aware of is in
-recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
+The most abstract characterization stack safety that we are aware of is
+recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
 properties:
@@
 mechanism (in their case, hardware capabilities) guarantees stack safety, by
-proving that an ordinary machine with enforcement can simulate the idealized machine.
+proving that an ordinary machine with enforcement simulates the idealized
+machine.
 %an enforcement mechanism (in their case, hardware
@@
 
-In the present work, we also distinguish between protecting stack data (like
+In the present work, we also aim to protect both stack data (like
 LSE) and
-control flow (like WBCF).  But our formulations of both properties are
+control flow (like WBCF).  But our formulations of these properties are
 somewhat different from those of Skorstengaard et al.  Rather than
 defining stack safety implicitly,
-as ``what a certain idealized machine does,'' we formalize explicit properties
+as ``what a certain idealized machine does,'' our properties are
 based on the technical framework of language-based
@@
   survey?}\fi.
-Our core properties are \emph{stepwise stack integrity} and
-\emph{stepwise stack confidentiality}.  WBCF is a desireable property, but
-we can express safety in an information-flow sense without reference to it,
-even safety from attacks that hijack control flow.  \BCP{On consideration
-  (especially in light of the weird ``C confuses A by returning directly to
-  A instead of returning to B so that B can finish up some stuff before
-  returning to A'' example\footnote{Actually, this example is not so weird.
-  If C, the attacker, can return a {\em result} directly to A, who is
-  expecting a result from B, then that is clearly bad!}), I
-agree now that we should discuss WBCF.  It should be kept as distinct as
-possible from the ``data safety'' part, since it is not really an
-information flow property.  But our treatment of it is interestingly
-different from StkTokens, because of our call annotations, and we should
-discuss that.}
+Our core properties are \emph{stepwise stack integrity},
+\emph{stepwise stack confidentiality}, and \emph{return to caller}.
+% \BCP{On consideration
+%   (especially in light of the weird ``C confuses A by returning directly to
+%   A instead of returning to B so that B can finish up some stuff before
+%   returning to A'' example\footnote{Actually, this example is not so weird.
+%   If C, the attacker, can return a {\em result} directly to A, who is
+%   expecting a result from B, then that is clearly bad!}), I
+% agree now that we should discuss WBCF.  It should be kept as distinct as
+% possible from the ``data safety'' part, since it is not really an
+% information flow property.  But our treatment of it is interestingly
+% different from StkTokens, because of our call annotations, and we should
+% discuss that.}
 
@@
   keep reading now...)}In formalizing these security concepts, we treat {\em each} caller as needing
-protection from its callee (which, in turn, needs protection from its
-callee, etc.).
-This is novel in that the
-confidentiality property applies to many nested sub-traces within a single execution,
+protection from its immediate callees (which, in turn, need protection from
+their immediate callees, etc.).
+This approach is novel in that the
+confidentiality\bcp{only?} property applies to many nested sub-traces within a single execution,
 as opposed to traces of entire executions.
@@
 modification by its callee(s),
-motivated by the reasoning principle that local data persists across a call. Its
-{\em stepwise} interpretation\bcp{We're introducing ``stepwise'' in a pretty
-abrupt way---potentially confusing.} formalizes this very naturally: a caller's local data
+motivated by the reasoning principle that local data persists across a
+call.
+
+For the confidentiality and integrity properties, we consider two variants.
+The
+{\em stepwise} variants formalize the intuition that a caller's
+local data
 is never modified during a call. {\em Confidentiality} says that a caller's
@@
 
+% BCP: This will intentionally fail when someone adds the real intended
+% definition of \WF!
+\newcommand{\WF}{\mathit{WF}}
+
 We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
@@
 %
-Note that the base machine has a total step function, and the policy-enhanced machine
-refines it as a partial step. The policy state never changes behavior, but can cause
-premature termination, which represents failstop behavior.
+Note that the base machine has a total step function, while the policy-enhanced machine
+refines it to a partial function. The extra policy state never changes behavior,
+but it can cause
+premature termination, representing failstop behavior.
 

### 2021-04-16 tex/main.tex
Working on intro

@@
 Informally, LSE says that the contents of
-a caller's stack frame are not read and written while its callees (and their callees, etc.) are
+a caller's stack frame are not read or written while its callees (and their callees, etc.) are
 executing, and WBCF says that callees always
-return to the point just after their call (if they return at all).
+return to the instruction following their call (if they return at all).
 Skorstengaard et al.{} formalize these properties by defining an idealized
@@
 control flow (like WBCF).  But our formulations of these properties are
-somewhat different from those of Skorstengaard et al.  Rather than
+rather different from those of Skorstengaard et al.  Instead of
 defining stack safety implicitly,
-as ``what a certain idealized machine does,'' our properties are
-based on the technical framework of language-based
+as ``what a certain idealized machine does,'' we define it
+\emph{explicitly} using the technical framework of language-based
 security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
@@
 \emph{stepwise stack confidentiality}, and \emph{return to caller}.
-% \BCP{On consideration
-%   (especially in light of the weird ``C confuses A by returning directly to
-%   A instead of returning to B so that B can finish up some stuff before
-%   returning to A'' example\footnote{Actually, this example is not so weird.
-%   If C, the attacker, can return a {\em result} directly to A, who is
-%   expecting a result from B, then that is clearly bad!}), I
-% agree now that we should discuss WBCF.  It should be kept as distinct as
-% possible from the ``data safety'' part, since it is not really an
-% information flow property.  But our treatment of it is interestingly
-% different from StkTokens, because of our call annotations, and we should
-% discuss that.}
-
-\bcp{This para could be clearer.  (I can try to fix it later but I want to
-  keep reading now...)}In formalizing these security concepts, we treat {\em each} caller as needing
-protection from its immediate callees (which, in turn, need protection from
-their immediate callees, etc.).
-This approach is novel in that the
-confidentiality\bcp{only?} property applies to many nested sub-traces within a single execution,
-as opposed to traces of entire executions.
-{\em Integrity} says that each caller's local data is protected from
-modification by its callee(s),
-motivated by the reasoning principle that local data persists across a
-call.
-
-For the confidentiality and integrity properties, we consider two variants.
-The
-{\em stepwise} variants formalize the intuition that a caller's
-local data
-is never modified during a call. {\em Confidentiality} says that a caller's
-data is not leaked into its callee, and  that the callee cannot see uninitialized data.
-As a reasoning principle, confidentiality guarantees that the callee's
-behavior is determined only by a predictable interface (function arguments and
-register contents) that does not include the caller's
-local state or the uninitialized portion of the stack. This is formalized as a
-{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
-compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
-again
-{\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
-changes in the state are independent of the caller's data.
-
-Together, these properties capture roughly the same intuition as LSE, but
-they are somewhat stronger.
-First, we model returns as returns to a valid return target\bcp{that will
-  make no sense to the reader at this point}; an invalid return
-is treated as a mere jump within the callee, and does not grant access to the caller's
-data. So, even without WBCF, our stack safety properties rule out many control-flow attacks.
-The attacker in such an instance is permitted to ``return'' so long as it avoids
-violating the integrity and confidentiality of the caller.\bcp{ditto} Second, confidentiality
-contains\bcp{?} temporal protections that LSE does not necessarily address\bcp{?}. The intuition
-for LSE is that it protects the caller's stack frame, not necessarily memory that was
-previously allocated but now outside of the caller's frame. (Cheri-capability-based
-implementations of LSE {\em do} protect previously allocated data, typically by clearing
-it, but because it is a potential vector for leaking capabilities into the caller's frame,
-not as a goal of the property.\bcp{Oof.  We need to do better than that!})
-
-To illustrate\bcp{IMO, this is not just an illustration; validating and
-  debugging a published enforcement mechanism is a top-level contribution.}
-the applicability of these properties, we perform randomized testing to
-validate that an existing enforcement mechanism does enforce them.
-\citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
-using tag-based \emph{micro-policies}~\citep{pump_oakland2015}, but offer no formal properties.
+%
+A key technical novelty in the definitions of confidentiality and integrity,
+compared to standard formulations from the security literature, is that they
+are ``nested'': {\em each} caller is guaranteed protection from its
+immediate callees (which, in turn, need protection from their immediate
+callees, etc.).
+
+% This approach is novel in that the
+% confidentiality\bcp{only?} property applies to many nested sub-traces within a single execution,
+% as opposed to traces of entire executions.
+% {\em Integrity} says that each caller's local data is protected from
+% modification by callees,
+% motivated by the reasoning principle that local data persists across a
+% call.
+
+For the confidentiality and integrity properties, we consider both {\em
+  stepwise} and {\em extensional} variants.  The simpler {stepwise} variants
+formalize the intuition that a caller's local data is {\em never} read or
+modified during a call. The extensional variants allow a callee to read from
+from and write to their caller's stack frame, as long as nothing that they
+read affects their observable behavior and nothing they write affects the
+observable behavior of the caller after they return.  \bcp{Please check: is
+  that accurate?}
+
+% {\em Confidentiality} says that a caller's
+% data is not leaked into its callee, and  that the callee cannot see uninitialized data.
+% As a reasoning principle, confidentiality guarantees that the callee's
+% behavior is determined only by a predictable interface (function arguments and
+% register contents) that does not include the caller's
+% local state or the uninitialized portion of the stack. This is formalized as a
+% {\em noninterference}-style property that, rather than comparing pairs of whole program traces,
+% compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
+% again
+% {\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
+% changes in the state are independent of the caller's data.
+
+% Together, these properties capture roughly the same intuition as LSE, but
+% they are somewhat stronger.
+% First, we model returns as returns to a valid return target\bcp{that will
+%   make no sense to the reader at this point}; an invalid return
+% is treated as a mere jump within the callee, and does not grant access to the caller's
+% data. So, even without WBCF, our stack safety properties rule out many control-flow attacks.
+% The attacker in such an instance is permitted to ``return'' so long as it avoids
+% violating the integrity and confidentiality of the caller.\bcp{ditto} Second, confidentiality
+% contains\bcp{?} temporal protections that LSE does not necessarily address\bcp{?}. The intuition
+% for LSE is that it protects the caller's stack frame, not necessarily memory that was
+% previously allocated but now outside of the caller's frame. (Cheri-capability-based
+% implementations of LSE {\em do} protect previously allocated data, typically by clearing
+% it, but because it is a potential vector for leaking capabilities into the caller's frame,
+% not as a goal of the property.\bcp{Oof.  We need to do better than that!})
+
+\bcp{I trimmed away a ton of text here that I thought was getting too much
+  into the weeds for the introduction.  However, in the process I lost the
+  explanation of how our properties are a bit stronger than LSE.  One way to
+put it back would be simply to add to the contributions a claim that we are
+a bit stronger than them, with a pointer to where we discuss this in detail.}
+
+To demonstrate the utility of our formal characterization, we use these
+properties to validate and improve an existing enforcement mechanism, the
+{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
+use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for the Coq proof assistant, to generate many random programs and check
+that Roessler and Dehon's micro-policies correctly abort the ones that
+attempt to violate one of our properties; furthermore, we introduce several
+intentionally broken micro-policies and check that the testing framework is
+able to generate counterexamples that violate our properties but are not
+halted by the incorrect enforcement variants.  (Showing that the testing
+framework detects intentional bugs in enforcement mechanisms increases our
+confidence that it can also catch unknown ones, strengthening our confidence
+in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
+  sentence belongs in the testing section?}\fi)
+
 We test against their {\em Depth Isolation} micro-policy, in which memory cells within each stack
@@
 
-Randomized testing is a relatively low-cost way to validate a potential enforcement
-mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
-of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
-covers potentially erroneous programs, we test intentionally buggy policies alongside
-the real one. If we can catch known bugs, we gain confidence that we can catch unknown ones,
-strengthening the validation of the enforcement mechanism.
-
-One challenge in characterizing stack safety formally is that, in a low-level machine
-without built-in stacks, we are at the mercy\bcp{?} of architectural\bcp{?}
-choices that do not relate to the abstract notion of stack safety. Some implementations
-may have only a stack pointer while others add a frame pointer, in some systems return
-addresses and data live in separate stacks, and so on. We abstract over these choices
-as much as possible.
+% Randomized testing is a relatively low-cost way to validate a potential enforcement
+% mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
+% of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
+% covers potentially erroneous programs, we test intentionally buggy policies alongside
+% the real one.
 
@@
 critically on the notions of ``call'' and ``return,'' which are not fully explicit
-in machine code---the same instruction opcode may be used in some places as part of
+in machine code%
+%
+\footnote{\bcp{Some useful text: A major challenge in characterizing stack
+    safety formally is that, in a low-level machine without built-in stacks,
+    we are at the mercy\bcp{?} of architectural\bcp{?}  choices that do not
+    relate to the abstract notion of stack safety. Some implementations may
+    have only a stack pointer while others add a frame pointer, in some
+    systems return addresses and data live in separate stacks, and so on. We
+    abstract over these choices as much as possible.  }}%
+%
+---the same instruction opcode may be used in some places as part of
 a call or return sequence and in other places for other purposes.

### 2021-04-16 tex/main.tex
For real now
Major haircut for the intro -- feels much better now!

@@
 For the confidentiality and integrity properties, we consider both {\em
-  stepwise} and {\em extensional} variants.  The simpler {stepwise} variants
-formalize the intuition that a caller's local data is {\em never} read or
-modified during a call. The extensional variants allow a callee to read from
-from and write to their caller's stack frame, as long as nothing that they
-read affects their observable behavior and nothing they write affects the
-observable behavior of the caller after they return.  \bcp{Please check: is
-  that accurate?}
+  stepwise} and {\em observational} variants.  The simpler {stepwise}
+variants formalize the intuition that a caller's local data is {\em never}
+read or modified during a call. The more extensional observational variants
+allow a callee to read from from and write to their caller's stack frame, as
+long as nothing that they read affects their observable behavior and nothing
+they write affects the observable behavior of the caller after they return.
+\bcp{Please check: is that accurate?}
 
@@
   sentence belongs in the testing section?}\fi)
-
-We test against their {\em Depth Isolation} micro-policy, in which memory cells within each stack
-frame are tagged with the identity of the function activation
-that owns the frame, and access to those locations is then gated
-according to the activation that is currently executing. This is simple and corresponds
-to intuitive notions of stack safety, and we argue that by describing its behavior
-we approach a reasonable stack-safety property that other micro-policies can be
-validated against.
-
-% Randomized testing is a relatively low-cost way to validate a potential enforcement
-% mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
-% of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
-% covers potentially erroneous programs, we test intentionally buggy policies alongside
-% the real one.
-
-A second challenge is that enforcement mechanisms may weaken their
-protections\bcp{``weaken their protections'' doesn't convey what we mean
-  very well}
-in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
-\citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
-ways of identifying functions and deciding exactly which accesses are allowed.
-These vary in their efficiency and in the spatial and temporal precision of their
-security guarantees.
-The Depth Isolation policy prevents a callee from writing outside its frame
-at all, which is simple and intuitively correct, but it must manually clear the tags
-on the frame on each return, which is very expensive. Conversely, their
-\emph{Lazy Tagging and Clearing} policy improves
-performance by assigning locations to the active function when they are written, which
-permits  ``illicit'' writes to inaccessible stack frames so long as the
-written locations are not subsequently read by their ``real'' owners. This violates our
-straightforward properties, but ntuitively,
-this form of laziness is justified if any dangerous writes will
-ultimately be detected before they can affect program behavior.
-
-\bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
-What does ``affecting program behavior'' really mean? This micro-policy
-calls for a second characterization of stack safety in terms of the externally
-observable behavior of a stack-safe system. Such a characterization will have the
-additional benefit of being more {\em extensional}, concerned primarily with the
-external behavior of the system. In contract, our stepwise properties are {\em intensional},
-focusing on internal details that may not actually matter to overall security.
-
-\bcp{Ditto, even moreso.}
-To formalize this intuition, we develop an \emph{observational integrity} property which
-characterizes enforcement mechanisms that detect violations between the point when
-they occur and the point when they make an observable impact. Intuitively this property
-can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
-integrity implies observational integrity. \sna{TODO: check this.}
-
-We additionally sketch an associated {\em observational confidentiality} property
-makes confidentiality extensional in the same way, in order to form an
-{\em observational stack safety} that is highly extensional.
-We argue that, despite its somewhat greater complexity,
-it may be a better specification for
-stack safety enforcement mechanisms in general.
+%
+We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
+which memory cells within each stack frame are tagged with the identity of
+the function activation that owns the frame and access to those locations is
+then controlled by the activation that is currently executing, validates our
+stepwise properties.  On the other hand, their more efficient \emph{Lazy
+  Tagging and Clearing} policy enforces observational integrity, but testing
+reveals that it does not enforce confidentiality (either stepwise or
+observational): it violates of the temporal aspect of confidentiality in
+corner cases where data can leak across calls to the same callee.  We
+propose a variant of Lazy Tagging and Clearing that \emph{does} enforce
+confidentiality (as far as our testing can tell), albeit at some performance
+cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
+  it in their paper?  (If neither, we should make a bigger deal out of the
+  fact that we discovered it.)}
+
+% A second challenge is that enforcement mechanisms may weaken their
+% protections\bcp{``weaken their protections'' doesn't convey what we mean
+%   very well}
+% in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
+% \citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
+% ways of identifying functions and deciding exactly which accesses are allowed.
+% These vary in their efficiency and in the spatial and temporal precision of their
+% security guarantees.
+% The Depth Isolation policy prevents a callee from writing outside its frame
+% at all, which is simple and intuitively correct, but it must manually clear the tags
+% on the frame on each return, which is very expensive. Conversely, their
+% \emph{Lazy Tagging and Clearing} policy improves
+% performance by assigning locations to the active function when they are written, which
+% permits  ``illicit'' writes to inaccessible stack frames so long as the
+% written locations are not subsequently read by their ``real'' owners. This violates our
+% straightforward properties, but ntuitively,
+% this form of laziness is justified if any dangerous writes will
+% ultimately be detected before they can affect program behavior.
+
+% \bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
+% What does ``affecting program behavior'' really mean? This micro-policy
+% calls for a second characterization of stack safety in terms of the externally
+% observable behavior of a stack-safe system. Such a characterization will have the
+% additional benefit of being more {\em extensional}, concerned primarily with the
+% external behavior of the system. In contract, our stepwise properties are {\em intensional},
+% focusing on internal details that may not actually matter to overall security.
+
+% \bcp{Ditto, even moreso.}
+% To formalize this intuition, we develop an \emph{observational integrity} property which
+% characterizes enforcement mechanisms that detect violations between the point when
+% they occur and the point when they make an observable impact. Intuitively this property
+% can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
+% integrity implies observational integrity. \sna{TODO: check this.}
+
+% We additionally sketch an associated {\em observational confidentiality} property
+% makes confidentiality extensional in the same way, in order to form an
+% {\em observational stack safety} that is highly extensional.
+% We argue that, despite its somewhat greater complexity,
+% it may be a better specification for
+% stack safety enforcement mechanisms in general.
 
@@
 %  ones...''}
-As it turns out, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
-Clearing micro-policy enforces observational integrity, but not confidentiality,
-whether stepwise or observational.
-But formalizing the ideal property allows us to see how it falls short -- a specific
-violation of the temporal aspect of confidentiality occurs in corner cases where callees
-can leak data between one another.
-We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
-confidentiality, albeit at some performance cost.  \bcp{Did Nick and Andre
-  know about this deficiency?  Did they mention it in their paper?  (If
-  neither, we should make a bigger deal out of the fact that we discovered it.)}
-
-Our initial description of these properties assumes a single simple stack with no sharing
-between callers and callees: all parameters and return values are passed in registers.  Later,
-we show how to refine it to allow passing of scalar stack data, granting a
-callee access to the caller's data. We do not model more sophisticated sharing,
-such as taking the address of a variable and treating it like a pointer.
-
-Finally, we define integrity and confidentiality properties for a coroutine system featuring
-a static layout, in which each coroutine's stack is constrained to a fixed region of memory.
-Though simplistic, this extension is a step toward concurrency and demonstrates that
-our approach is flexible enough to handle more sophisticated styles of
-control flow.
+
+Our first presentation of these properties assumes a single simple stack
+with no sharing between callers and callees: all parameters and return
+values are passed in registers.  Later, we show how to refine it to allow
+passing of scalar stack data, granting a callee access to the caller's
+data. We do not model more sophisticated sharing, such as taking the address
+of a variable and treating it like a pointer.
+%
+Finally, we define integrity and confidentiality properties for a coroutine
+system featuring a static layout, in which each coroutine's stack is
+constrained to a fixed region of memory.  Though simplistic, this extension
+is a step toward concurrency and demonstrates that our approach is flexible
+enough to handle more sophisticated styles of control flow.
 

### 2021-04-16 tex/main.tex
7 + 8 unlocked

--- 2021-04-15 tex/main.tex
@@
 We propose a formal characterization
-of stack safety, formulated with concepts from language-based security: a
+of stack safety, phrased using concepts from language-based security: a
 combination of an integrity property (``the private
 state in each caller's stack frame is held invariant by the callee'')
-a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state''), and a well-bracketedness property (``each
-callee returns control to its immediate caller'').
+and a confidentiality property (``the callee's behavior is insensitive to the
+caller's private state'')\bcp{and something about WBCF...}.
 % These properties are sensible even in the presence of
@@
 %
-We use these properties to validate the stack-safety ``micro-policies''
-proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
-property-based random testing) that Roessler and Dehon's ``eager''
-micro-policy, which catches violations as early as possible, enforces a
-simple ``stepwise'' variant of our properties and correctly detects several
-broken variants, and that (a repaired version of) their more performant
-``lazy'' micro-policy corresponds to a slightly weaker and more extensional
-variant of our properties.
+We use our characterization to validate the stack-safety ``micro-policies''
+proposed by~\citet{DBLP:conf/sp/RoesslerD18}.
+Specifically, we use property-based random testing to check that Roessler
+and Dehon's ``eager'' micro-policy, which catches violations as early as
+possible, enforces a simple ``stepwise'' variant of our properties (and
+detects several broken variants), and that a slightly repaired version of
+their more performant ``lazy'' micro-policy corresponds to weaker and more
+extensional variants of the properties.
 %
@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-The most abstract characterization stack safety that we are aware of is
-recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
+The most abstract characterization stack safety that we are aware of is in
+recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
 properties:
@@
 mechanism (in their case, hardware capabilities) guarantees stack safety, by
-proving that an ordinary machine with enforcement simulates the idealized
-machine.
+proving that an ordinary machine with enforcement can simulate the idealized machine.
 %an enforcement mechanism (in their case, hardware
@@
 
-In the present work, we also aim to protect both stack data (like
+In the present work, we also distinguish between protecting stack data (like
 LSE) and
-control flow (like WBCF).  But our formulations of these properties are
+control flow (like WBCF).  But our formulations of both properties are
 somewhat different from those of Skorstengaard et al.  Rather than
 defining stack safety implicitly,
-as ``what a certain idealized machine does,'' our properties are
+as ``what a certain idealized machine does,'' we formalize explicit properties
 based on the technical framework of language-based
@@
   survey?}\fi.
-Our core properties are \emph{stepwise stack integrity},
-\emph{stepwise stack confidentiality}, and \emph{return to caller}.
-% \BCP{On consideration
-%   (especially in light of the weird ``C confuses A by returning directly to
-%   A instead of returning to B so that B can finish up some stuff before
-%   returning to A'' example\footnote{Actually, this example is not so weird.
-%   If C, the attacker, can return a {\em result} directly to A, who is
-%   expecting a result from B, then that is clearly bad!}), I
-% agree now that we should discuss WBCF.  It should be kept as distinct as
-% possible from the ``data safety'' part, since it is not really an
-% information flow property.  But our treatment of it is interestingly
-% different from StkTokens, because of our call annotations, and we should
-% discuss that.}
+Our core properties are \emph{stepwise stack integrity} and
+\emph{stepwise stack confidentiality}.  WBCF is a desireable property, but
+we can express safety in an information-flow sense without reference to it,
+even safety from attacks that hijack control flow.  \BCP{On consideration
+  (especially in light of the weird ``C confuses A by returning directly to
+  A instead of returning to B so that B can finish up some stuff before
+  returning to A'' example\footnote{Actually, this example is not so weird.
+  If C, the attacker, can return a {\em result} directly to A, who is
+  expecting a result from B, then that is clearly bad!}), I
+agree now that we should discuss WBCF.  It should be kept as distinct as
+possible from the ``data safety'' part, since it is not really an
+information flow property.  But our treatment of it is interestingly
+different from StkTokens, because of our call annotations, and we should
+discuss that.}
 
@@
   keep reading now...)}In formalizing these security concepts, we treat {\em each} caller as needing
-protection from its immediate callees (which, in turn, need protection from
-their immediate callees, etc.).
-This approach is novel in that the
-confidentiality\bcp{only?} property applies to many nested sub-traces within a single execution,
+protection from its callee (which, in turn, needs protection from its
+callee, etc.).
+This is novel in that the
+confidentiality property applies to many nested sub-traces within a single execution,
 as opposed to traces of entire executions.
@@
 modification by its callee(s),
-motivated by the reasoning principle that local data persists across a
-call.
-
-For the confidentiality and integrity properties, we consider two variants.
-The
-{\em stepwise} variants formalize the intuition that a caller's
-local data
+motivated by the reasoning principle that local data persists across a call. Its
+{\em stepwise} interpretation\bcp{We're introducing ``stepwise'' in a pretty
+abrupt way---potentially confusing.} formalizes this very naturally: a caller's local data
 is never modified during a call. {\em Confidentiality} says that a caller's
@@
 
-The building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively.
+The basic building blocks of the machine are {\em values} and {\em addresses}.
+Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\).
 %
 Our machine states are composed of {\em components} \(\component\),
-which are either addresses or register names (\(\reg\)) drawn from some set
+which are either memory addresses (words) or register names (\(\reg\)) drawn from some set
 \(\REGS\), which is assumed to include two special purpose registers: the
 program counter {\PCname} and the stack pointer \(\SP\).
-A {\em machine state} is a map from components to values.
 %
@@
 %
+A {\em machine state} is a map from components to values:
+%
 \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
@@
 We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-
-% BCP: This will intentionally fail when someone adds the real intended
-% definition of \WF!
-\newcommand{\WF}{\mathit{WF}}
+\apt{But just below we write $\mach \stepstoobs \mach'$. Which is it?}
+\apt{Say something about initial machine state?}
+
+\subsection{Policies}
 
@@
 it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. The policy also provides a
-well-formedness condition, \(\WF (\mpstatename \in \MACHS \times \POLS)\), describing
-initial states that are properly configured.
+can see the machine state, but cannot modify it. We assume that the
+compiler\bcp{why?} provides an initial policy state \(\pol_0\) to accompany the initial machine state.\apt{see comment above}
 %
@@
 
+We write $\MPS$ for the set of all pairs of machine states and policy
+states.
+%
+\[\mpstatename \in \MPS = \MACHS \times \POLS\]
+%
 We lift the policy step function to operate on such pairs
 by combining it with the regular step function for machine states.
-%
-\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
+
+\judgmenttwo{\(\mach \stepstoobs \mach'\)}{\((\mach, \pol)
                \polstep \pol'\)}
@@
                \mpstate{\mach'}{\pol'}\)}
-%
-Note that the base machine has a total step function, while the policy-enhanced machine
-refines it to a partial function. The extra policy state never changes behavior,
-but it can cause
-premature termination, representing failstop behavior.
+
+\noindent
+This combined step function is also partial, and traces of the machine-with-policy
+will be finite iff a policy fault occurs.
 
@@
 
-Separate from the machine state, we will keep track of additional context
-information representing the history of the run. This has no influence on execution, it merely
-describes information relevant to the particular property. Different information is needed
-for different versions of the model, so the definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+Separate from the machine state, our formal properties will keep track of additional context
+information representing the history of the run. The following definitions are parameterized
+over an arbitrary\bcp{why?} set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
 an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
-\times \rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
-\(\codemap \in \WORDS \rightharpoonup \ANNS\)
-that relates each code address to its annotation, if any.
+\(\mach, \context, \anns \constep \context' \in \MACHS \times \CONTEXTS
+\times \powerset(\ANNS) \rightarrow \CONTEXTS\). We also assume a set of function identifiers
+\(\FUNIDS\) and the existence of a \emph{code map} \(\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\)
+that relates each code address to its owning function and set of annotations.
 We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).
-
-\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
-              {\(\mach, \context \constep \context'\)}
+a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
+\CONTEXTS\).  \bcp{I don't get the intuition for this whole para.}
+
+\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
+              {\(\mach, \context, \anns \constep \context'\)}
               {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
@@
 
-\subsection{Machine Traces}
+\subsection{Traces}
 \label{sec:traces}
 
-A {\em machine trace} is a nonempty, finite or infinite sequence of elements
-of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
+A trace \(T \in \TRACE{\TRACEELEMS}\) is a nonempty, finite or infinite sequence of elements
+of some type \(\TRACEELEMS\).  We will mostly be concerned with sequences of MPC-states,
+called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.
+
+\paragraph*{Trace-Of}
 
@@
 coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\)
+produced by repeated application of \(\stepstocon\), and potentially cut short by
+the policy component. % The {\sc Default} rule applies only if the other does not.
+(We use juxtaposition of an element and a trace to represent ``cons.''\bcp{Ugh})
 
@@
 
+\paragraph*{Projections}
+
 We project out the machine state of an MPC-state with \(\pi_\mach\), the
-policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
-We take the first element of a trace with \(\head(\MPCT)\), which is a total
+policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
+these operators to project out of an machine-trace a trace of machine states,
+policy states, and contexts, respectively.
+
+\paragraph*{Head and Last}
+
+We take the first element of a trace with \(\head(\Trace)\), which is a total
 function since traces are non-empty, and the final element (if one exists) with
-\(\last(\MPCT)\), which is partial.
+\(\last(\Trace)\), which is partial.
 
 \paragraph*{Until}
-The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
-and a predicate on states \(f ~ \mpcstate \in \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first element on which \(f\) holds.
+The operation \(\PUT ~ f ~ \Trace\) takes a trace
+\(\Trace \in \TRACE{\TRACEELEMS %\times \OBSS
+}  \) and a predicate
+on elements \(f \subset \TRACEELEMS\) and gives the prefix of
+\(\Trace\) ending with the first element on which \(f\) holds.
 
@@
   \begin{minipage}{.3\textwidth}
-    \judgment{\(f ~ \mpcstatename\)}
-             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
+    \judgment{\(f ~ \traceelem\)}
+             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
   \end{minipage}
@@
     \judgment{}
-             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
+             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
 %
@@
   \begin{minipage}{.3\textwidth}
-    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
-                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
+    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
+                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
   \end{minipage}
@@
 \(\PUT\) is the identity function.
+
+% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
+%$\mathit{split}$: it takes two traces and links them by replacing the
+%last element of the first (if it has one) with the first element of
+%the second
+
+%\judgment{} {\(t \cdot T = T\)}
+%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
+%\leo{Lemma/theorem: split/join = id?}
+%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}
 
@@
 \hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)''):
-%
+running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:
+
 \begin{center}
@@
 
-individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth. If \(f\) is still a predicate on states, then
+In particular, we wish to formulate properties about consecutive sub-traces
+in which the context obeys a given predicate. This will commonly be used to extract
+individual calls from a trace by taking sub-traces in which the stack is
+at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
@@
   \judgmenttwo[Now]
-              {\(f ~ \mpcstatename\)}
-              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
+              {\(f ~ \mach ~ \context\)}
+              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
+              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
 %
   \judgmentthree[Skip]
-                {\(f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
+                {\(f ~ \mach ~ \context\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
                 {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
 %
   \judgmentthree[Search]
-                {\(\neg f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
+                {\(\neg f ~ \mach ~ \context\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
                 {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
 \end{center}
@@
 properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
-where \(\WF ~ \mach,\pol\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
-such that from some similarly well-formed initial state \(\mpcstatename\),
-\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
+is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
+and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a sub-trace
+such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
+\apt{Unclear how there can be multiple initial states.}
 
@@
 A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
-with associated \(\mach\) and \(\components\) as above, and
+with associated \(m\) and \(\components\) as above, and
 for all \(\nach\) such that \(\mach \approx_\components \nach\) and
- \(\nach \hookrightarrow \NPCT\),
- \(\MPCT\langle f_\depth\rangle \doteq \NPCT | \neg f_\depth \).\apt{I found the previous
+ \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\),
+ \(\MPCT\langle f_\depth\rangle \doteq \NPCT\langle f_\depth\rangle \).\apt{I found the previous
    version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
- \sna{It's right, but I need to rethink how the M<f> notation works...}
+%% \definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
+%% \(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+%% let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+%%  \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
+%% \(\MPCT\) corresponds to the trace of a function from its call to its return.
+%% Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+%% \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+%% {\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
+%% \(\MPCT \doteq \NPCT\).
 
@@
 Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
 {\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
@@
 Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
 {\em stepwise coroutine confidentiality} in the coroutine setting if for all
@@
 
-In this section we examine how an existing enforcement mechanism,
-Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
-the formal stack-safety property described in the previous section: it in
-fact enforces a stronger, inductive version of that property which strengthens the
+\leo{I am not sure we've hit the correct level of detail in this section.}
+
+In this section we detail an existing enforcement mechanism,
+Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
+the formal stack-safety property described in the previous section; in
+fact, it enforces a stronger, inductive version of that property which strengthens the
 local state encapsulation component by checking for integrity and
@@
 
-\subsection{A Conservative Tag Policy}
-\label{sec:conservative}
-
-\bcp{Is this the only subsection in this section?}
-
-The first micro-policy we present to enforce the definition of stack
-safety developed in \cref{sec:lse-and-wbcf} is a small variation
+%\paragraph{A Conservative Tag Policy}
+%\label{sec:conservative}
+
+The micro-policy we present here to enforce the definition of stack
+safety developed in \cref{sec:lse} is a small variation
 on the Depth Isolation policy presented by
@@
 sequences that have permission to manage the stack, say during calls and returns.
+In the following, for ease of exposition, we will assume that no
+arguments are passed on the stack.
 %% \rb{Conceivably on alloc
@@
 
-Our micro-policy works as follows. (We assume for ease of exposition that no
-arguments are passed on the stack.)
-%% \rb{Come back to this later?}.
-The micro-policy maintains tags on values in memory of the form $\tagStackDepth{n}$,
-indicating locations that belong to the stack frame at activation depth $n$,
-or $\tagNoDepth$; it tags the $\PCname$ with tags of the form $\tagPCDepth{n}$.
-During normal execution, the micro-policy rules only permit load and store operations
-when the target memory is tagged with the same depth as the current {\PCname} tag.
-Initially, the entire stack is tagged $\tagNoDepth$,
-%\rb{actually, the ``initial frame'' should be tagged with its proper depth}
-and the {\PCname} has tag $\tagPCDepth{0}$, where $n$ is the current activation depth.
-These tags are altered
-at exactly those points in
-the program where the contour of the stack-safety property changes:
+The micro-policy maintains tags of the form $\tagStackDepth{n}$ for
+each stack location belonging to the stack frame at activation depth
+$n$, or the tag $\tagNoDepth$ for stack locations that are currently
+unused (but could potentially be used in the future).
+%
+The $\PCname$ is tagged similarly as $\tagPCDepth{n}$.
+%
+Initially, the entire stack is considered unused (and therefore tagged
+$\tagNoDepth$), and the {\PCname} has tag $\tagPCDepth{0}$.
+%
+During normal execution, the micro-policy rules only permit load and
+store operations when the target memory is tagged {\em with the same
+  depth} as the current {\PCname} tag.
+
+More interesting (from a stack safety perspective) changes to the tags
+happen at exactly the points in the program where we switch from a
+caller to a callee or vice-versa:
 
@@
 
-\item From callee back to caller, when execution is about to reach the return point
-  $\ret{c}$
-  corresponding to the call at $c$.  The exit sequence retags all locations in the function
-  frame with $\tagNoDepth$. At the actual return instruction, the
-  $\PCname$ tag is decremented.
-  \rb{Problem: this is currently the first mention of $\ret{\cdot}$}
+\item From callee back to caller, when execution is about to reach a
+  return point.  The exit sequence retags all locations in the
+  function frame with $\tagNoDepth$. At the actual return instruction,
+  the $\PCname$ tag is decremented. \leo{Not sure the actual policy
+    does that now (retag unused). But I don't think it's a
+    confidentiality or integrity violation.}
 
@@
 
+\leo{I'm not a fan of the next paragraph.}
 Note that in this policy both the entry and exit sequences
@@
 
+We validate the micro-policy of the previous section by systematically
+testing that it satisfies eager stack integrity and
+confidentiality. We used a Coq specification of the RISC-V
+architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
+extended it with a runtime monitor implementing the stack-safety
+micro-policy. \leo{Footnote: code will be made public} We chose the
+Coq proof assistant as the setting for our implementation for two
+reasons: first, to ensure that our coinductive trace definitions are
+well-formed, and reason about them\leo{Do we prove anything
+  interesting anymore?}, and to leverage the power of the QuickChick
+property-based testing framework.\leo{cite}
+
+To use QuickChick, we built generators that produce:
+\begin{enumerate}
+\item An initial RISC-V machine state, including most notably
+  the program to be executed.
+\item An initial policy state, tagging instructions corresponding to
+  ``blessed'' call or return sequences appropriately, while marking
+  all potential stack locations as $\tagNoDepth$.
+\item A code map and a stack map, accurately characterizing the behavior
+  of callers and callees.
+\end{enumerate}
+%
+That is, our generators need to produce not only a low-level program
+that exhibits interesting control-flow patterns, but also policy
+meta-information and code annotations that would normally be produced
+by the compiler.
+
+\paragraph*{Generator Overview}
+
+To write such generators we build on the work of \leo{ICFP 2013},
+where they introduced {\em generation by execution} to produce
+interesting programs. Generation by execution receives as an input a
+partially instantiated machine state, and attempts to generate an
+instruction (or a sequence of instructions such as a ``blessed''
+sequence) that makes sense locally (e.g., jumps to a potentially valid
+code location or loads from a potentially valid stack location). Then,
+we step the machine and repeat the process until we generate or
+execute some target number of instructions (or if we reached a point
+where the machine can't step anymore).
+
+We extend this technique to keep track of the control flow behavior of
+the program being generated: each time a call or return sequence is
+generated, we ensure that the appropriate policy tags and code
+annotations are set for the entry or return points. At the same time,
+we sometimes allow the generation to relax those constraints,
+introducing potentially ill-formed flows: that causes our programs
+to failstop when executed in conjunction with the policy monitor,
+but also allows for revealing errors in our setup.
+
+\paragraph*{Findings}
+
+To evaluate the effectiveness of our testing, we turn to {\em mutation
+  testing}\leo{cite}. In mutation testing, we inject errors
+(mutations) in a program that should cause the property of interest
+(here, stack safety) to fail, and ensure that the testing framework
+can find them. The bugs we used for our evaluation were both
+artificially generated (by deliberately weakening the micro-policy in
+ways that should break enforcement), or actual bugs that we discovered
+through testing in our implementation. We elaborate on some such bugs
+below.
+
+The mean-time-to-failure (MTTF) for various bugs can be found in
+\leo{Table with numbers?}. All experiments were run in a \leo{Machine
+  spec}. \leo{Something like: There is really no baseline to compare
+  this against, but include it here for the future/reasons?}
+
+The artificially injected bugs focus on weakening the tag propagation
+and checking mechanism. \leo{Mention them all?}
+%
+For example, when loading from a stack location, the correct
+micro-policy needs to enforce that the tag on the location being read
+is $\tagStackDepth{n}$ for some number $n$ and that the one of the
+current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
+that restriction by not checking the depth equality, or not checking
+the tags at all, causing bugs of different difficulty.
+%
+Similarly, when storing to a stack location, the correct micro-policy
+needs to ensure that the tag on the memory location is either
+$\tagNoDepth$ or has again the same depth as the current $\PCname$
+tag. Relaxing these constraints also causes violations to the
+property.
+
+Using our testing framework, we also found various errors in our
+implementation, ranging from trivial typos to ones that required an
+intriguingly complex setup to appear. Most notably, the initial
+``blessed'' call sequence/policy combination did not initialize all
+locations for the newly allocated stack frame correctly, but left some
+of them as $\tagNoDepth$. This allows for a potential integrity
+violation, but only if a rather complicated sequence of events
+occured: \leo{call - write - call (seal) - return - call - write?}
+
 \section{Stack Safety for Lazy Enforcement}
@@
 %
-\[\mach \stepsto[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
+\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
 %
@@
 Let \(\nach\) be a variant such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth \).
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\).
 

### 2021-04-16 tex/main.tex

--- 2021-04-16 tex/main.tex
@@
 
-% BCP: This will intentionally fail when someone adds the real intended
-% definition of \WF!
-\newcommand{\WF}{\mathit{WF}}
-
 We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
@@
 \label{sec:testing}
+
+We validate the micro-policy of the previous section by systematically
+testing that it satisfies eager stack integrity and
+confidentiality. We used a Coq specification of the RISC-V
+architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
+extended it with a runtime monitor implementing the stack-safety
+micro-policy. \leo{Footnote: code will be made public} We chose the
+Coq proof assistant as the setting for our implementation for two
+reasons: first, to ensure that our coinductive trace definitions are
+well-formed, and reason about them\leo{Do we prove anything
+  interesting anymore?}, and to leverage the power of the QuickChick
+property-based testing framework.\leo{cite}
+
+To use QuickChick, we built generators that produce:
+\begin{enumerate}
+\item An initial RISC-V machine state, including most notably
+  the program to be executed.
+\item An initial policy state, tagging instructions corresponding to
+  ``blessed'' call or return sequences appropriately, while marking
+  all potential stack locations as $\tagNoDepth$.
+\item A code map and a stack map, accurately characterizing the behavior
+  of callers and callees.
+\end{enumerate}
+%
+That is, our generators need to produce not only a low-level program
+that exhibits interesting control-flow patterns, but also policy
+meta-information and code annotations that would normally be produced
+by the compiler.
+
+\paragraph*{Generator Overview}
+
+To write such generators we build on the work of \leo{ICFP 2013},
+where they introduced {\em generation by execution} to produce
+interesting programs. Generation by execution receives as an input a
+partially instantiated machine state, and attempts to generate an
+instruction (or a sequence of instructions such as a ``blessed''
+sequence) that makes sense locally (e.g., jumps to a potentially valid
+code location or loads from a potentially valid stack location). Then,
+we step the machine and repeat the process until we generate or
+execute some target number of instructions (or if we reached a point
+where the machine can't step anymore).
+
+We extend this technique to keep track of the control flow behavior of
+the program being generated: each time a call or return sequence is
+generated, we ensure that the appropriate policy tags and code
+annotations are set for the entry or return points. At the same time,
+we sometimes allow the generation to relax those constraints,
+introducing potentially ill-formed flows: that causes our programs
+to failstop when executed in conjunction with the policy monitor,
+but also allows for revealing errors in our setup.
+
+\paragraph*{Findings}
+
+To evaluate the effectiveness of our testing, we turn to {\em mutation
+  testing}\leo{cite}. In mutation testing, we inject errors
+(mutations) in a program that should cause the property of interest
+(here, stack safety) to fail, and ensure that the testing framework
+can find them. The bugs we used for our evaluation were both
+artificially generated (by deliberately weakening the micro-policy in
+ways that should break enforcement), or actual bugs that we discovered
+through testing in our implementation. We elaborate on some such bugs
+below.
+
+The mean-time-to-failure (MTTF) for various bugs can be found in
+\leo{Table with numbers?}. All experiments were run in a \leo{Machine
+  spec}. \leo{Something like: There is really no baseline to compare
+  this against, but include it here for the future/reasons?}
+
+The artificially injected bugs focus on weakening the tag propagation
+and checking mechanism. \leo{Mention them all?}
+%
+For example, when loading from a stack location, the correct
+micro-policy needs to enforce that the tag on the location being read
+is $\tagStackDepth{n}$ for some number $n$ and that the one of the
+current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
+that restriction by not checking the depth equality, or not checking
+the tags at all, causing bugs of different difficulty.
+%
+Similarly, when storing to a stack location, the correct micro-policy
+needs to ensure that the tag on the memory location is either
+$\tagNoDepth$ or has again the same depth as the current $\PCname$
+tag. Relaxing these constraints also causes violations to the
+property.
+
+Using our testing framework, we also found various errors in our
+implementation, ranging from trivial typos to ones that required an
+intriguingly complex setup to appear. Most notably, the initial
+``blessed'' call sequence/policy combination did not initialize all
+locations for the newly allocated stack frame correctly, but left some
+of them as $\tagNoDepth$. This allows for a potential integrity
+violation, but only if a rather complicated sequence of events
+occured: \leo{call - write - call (seal) - return - call - write?}
 

### 2021-04-16 tex/main.tex
Small

--- 2021-04-16 tex/main.tex
@@
 passing of scalar stack data, granting a callee access to the caller's
-data. We do not model more sophisticated sharing, such as taking the address
-of a variable and treating it like a pointer.
+data. (We do not consider more sophisticated forms of sharing, such as
+taking the address of a variable and treating it like a pointer.)
 %
 Finally, we define integrity and confidentiality properties for a coroutine
-system featuring a static layout, in which each coroutine's stack is
-constrained to a fixed region of memory.  Though simplistic, this extension
-is a step toward concurrency and demonstrates that our approach is flexible
+system with a static layout, in which each coroutine's stack is constrained
+to a fixed region of memory.  Though simplistic, this extension is a step
+toward full-blown concurrency and demonstrates that our approach is flexible
 enough to handle more sophisticated styles of control flow.
@@
 \item
-  We use a property-based randomized testing framework to support the claim
-  that this stepwise property is actually enforced by the Depth Isolation policy
+  We use a property-based randomized testing framework to validate the claim
+  that this stepwise property is actually enforced by the {\em Depth Isolation} policy
   of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:testing}).
@@
   We formalize a weaker but more extensional \emph{observational} variant of the
-  stack safety property and argue that it accurately characterizes
-  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18}
-  enhanced with unique tags on each activation~(\cref{sec:lazy}).
+  stack safety property and argue that it accurately characterizes a small
+  refinement of
+  the {\em Lazy Tagging and Clearing} policy
+  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).
 \end{itemize}
@@
 then dive into the formal definitions of local state encapsulation and
-well-bracketed control flow, which together constitute stack safety\bcp{not
-  any more?}.
-Initially we will focus on a simple stack model, with a single stack
-that doesn't share between caller and callee. All arguments are passed
-in registers. Recall that we keep track of which addresses are
-annotated as calls, with a ``code map'':
-\[\codemap \in \CODEMAPS \subseteq \WORDS\]
-\apt{Doesn't match previous definition any more.}
-
-We parameterize over the structure of the stack, particularly the means by
+well-bracketed control flow, which together constitute stack safety\BCP{need
+  to put back WBCF!}.
+
+\subsection{Parameters in Registers}
+
+To begin, we will focus on a simple stack model, with a single stack
+that doesn't share between caller and callee: arguments and results are passed
+in registers.
+
+To accommodate different machine architectures and different calling
+conventions, we parameterize over the structure of the stack, particularly
+the means by 
 which a caller identifies the data that should be protected during a call. We term
@@
 that matches the topmost return target. If neither is true, the context remains unchanged.
+%
+\bcp{(Moved this here, BTW.)}Recall that we keep track of which addresses are
+annotated as calls, with a ``code map'' $\codemap \in \CODEMAPS \subseteq
+\WORDS$.
+\apt{Doesn't match previous definition any more.}
 
@@
 
-\subsection{Stack Safety with Passing}
+\subsection{Parameters Passed on the Stack}
 
@@
 
-\subsection{Coroutine Safety}
+\subsection{Coroutines}
 

### 2021-04-16 tex/main.tex
Comments on 7 and 8

@@
   Tagging and Clearing} policy enforces observational integrity, but testing
-reveals that it does not enforce confidentiality (either stepwise or
+reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
+stepwise or
 observational): it violates of the temporal aspect of confidentiality in
@@
 
-\leo{I am not sure we've hit the correct level of detail in this section.}
-
-In this section we detail an existing enforcement mechanism,
+\bcp{``Enforcement'' seems like too broad a title.  What about calling it
+  ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
+  Safety?''? And what about folding the testing section into this one?  That
+would keep related things together, would avoid appearing to make overly
+grand claims in either section, and would give us an excuse to write less
+about the mechanism here---we only need to write enough about the mechanism
+itself that readers can understand our validation case study.}
+
+\leo{I am not sure we've hit the correct level of detail in this
+  section.}\bcp{Agreed!}
+
+In this section we sketch an existing enforcement mechanism,
 Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
@@
 
-\leo{I'm not a fan of the next paragraph.}
+\leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
+  the weeds.  I don't think this level of detail is needed, in the context
+  of a paper that is mostly about our definitions, not theirs.}
 Note that in this policy both the entry and exit sequences
@@
 
-We validate the micro-policy of the previous section by systematically
+We validated the micro-policy of the previous section by systematically
 testing that it satisfies eager stack integrity and
@@
 extended it with a runtime monitor implementing the stack-safety
-micro-policy. \leo{Footnote: code will be made public} We chose the
+micro-policy.\leo{Footnote: code will be made public}\bcp{For the reviewers?
+ Or later?} We chose the
 Coq proof assistant as the setting for our implementation for two
@@
   interesting anymore?}, and to leverage the power of the QuickChick
-property-based testing framework.\leo{cite}
-
-To use QuickChick, we built generators that produce:
+property-based testing framework~\citep{Pierce:SF4}.
+
+To use QuickChick, we built random test-case generators that produce:
 \begin{enumerate}
@@
 That is, our generators need to produce not only a low-level program
-that exhibits interesting control-flow patterns, but also policy
+that exhibits interesting control-flow patterns, but also the policy
 meta-information and code annotations that would normally be produced
@@
 
-\paragraph*{Generator Overview}
+\paragraph*{Generators}
 
@@
 interesting programs. Generation by execution receives as an input a
-partially instantiated machine state, and attempts to generate an
+partially instantiated machine state and attempts to generate an
 instruction (or a sequence of instructions such as a ``blessed''
-sequence) that makes sense locally (e.g., jumps to a potentially valid
-code location or loads from a potentially valid stack location). Then,
+sequence) that makes sense locally (e.g., jumps go to a potentially valid
+code location and loads read from from a potentially valid stack
+location). Then
 we step the machine and repeat the process until we generate or
-execute some target number of instructions (or if we reached a point
-where the machine can't step anymore).
-
-We extend this technique to keep track of the control flow behavior of
+execute some target number of instructions, or if we reach a point
+where the machine can't step anymore.
+
+We extended this technique to keep track of the control flow behavior of
 the program being generated: each time a call or return sequence is
@@
 
-To evaluate the effectiveness of our testing, we turn to {\em mutation
-  testing}\leo{cite}. In mutation testing, we inject errors
-(mutations) in a program that should cause the property of interest
-(here, stack safety) to fail, and ensure that the testing framework
-can find them. The bugs we used for our evaluation were both
-artificially generated (by deliberately weakening the micro-policy in
-ways that should break enforcement), or actual bugs that we discovered
-through testing in our implementation. We elaborate on some such bugs
-below.
-
-The mean-time-to-failure (MTTF) for various bugs can be found in
-\leo{Table with numbers?}. All experiments were run in a \leo{Machine
-  spec}. \leo{Something like: There is really no baseline to compare
-  this against, but include it here for the future/reasons?}
+To evaluate the effectiveness of testing against our formal properties, we
+used {\em mutation testing}\leo{cite}. In mutation testing, we inject errors
+(mutations) in a program that should cause the property of interest (here,
+stack safety) to fail, and ensure that the testing framework can find
+them. The bugs we used for our evaluation were either artificially generated
+by us (deliberately weakening the micro-policy in ways that we expected
+should break its guarantees), or actual bugs that we discovered through
+testing our implementation. We elaborate on some such bugs below.
+
+The mean-time-to-failure (MTTF) for various bugs can be found in \leo{Table
+  with numbers?}\bcp{Sure.}. All experiments were run in a \leo{Machine
+  spec}. \leo{Something like: There is really no baseline to compare this
+  against, but include it here for the future/reasons?}
 
 The artificially injected bugs focus on weakening the tag propagation
-and checking mechanism. \leo{Mention them all?}
+and checking mechanism. \leo{Mention them all?}\bcp{Nah, especially if they
+  are all listed in the table.}
 %
@@
 
-Using our testing framework, we also found various errors in our
-implementation, ranging from trivial typos to ones that required an
-intriguingly complex setup to appear. Most notably, the initial
+Naturally, testing also revealed a number of errors in our
+implementation of the enforcement mechanism (the original was written in C++
+and targeted ARM machine code\bcp{right?}; we re-implemented it in Coq targeting
+RISC-V).  These errors ranged from trivial typos to
+ones that required an 
+intriguingly complex setup to appear.  Notably, our first try at the 
 ``blessed'' call sequence/policy combination did not initialize all
 locations for the newly allocated stack frame correctly, but left some
-of them as $\tagNoDepth$. This allows for a potential integrity
+of them as $\tagNoDepth$. This allowed for a potential integrity
 violation, but only if a rather complicated sequence of events

### 2021-04-16 tex/main.tex
Lots of polishing here and there

@@
 %  to leave it out of the title but hit it strongly in the abstract.}%
-What does ``stack safety'' mean? The phrase is associated with a
+What exactly does ``stack safety'' mean? The phrase is associated with a
 variety of compiler,
-run-time, and hardware mechanisms for guarding against corruption of stack
+run-time, and hardware mechanisms for protecting stack
 memory.  But these mechanisms typically lack precise specifications,
@@
 Our core properties are \emph{stepwise stack integrity},
-\emph{stepwise stack confidentiality}, and \emph{return to caller}.
+\emph{stepwise stack confidentiality}, and \emph{return to caller}\BCP{I
+  gave it a name different from WBCF, but I could be talked out of that.  In
+any case, it now needs to be discussed / defined in several places below.}.
 %
@@
 \begin{itemize}
-\item
-  We adapt formulations of integrity and confidentiality from language-based security
-  to formalize a parameterized {\em stepwise stack-safety} property and argue that it
-  captures the intuitive concepts of stack safety~(\cref{sec:lse}). Our model extends to
-  argument passing on the stack and to a simple coroutine system.
-\item
-  We use a property-based randomized testing framework to validate the claim
-  that this stepwise property is actually enforced by the {\em Depth Isolation} policy
-  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:testing}).
-\item
-  We formalize a weaker but more extensional \emph{observational} variant of the
-  stack safety property and argue that it accurately characterizes a small
-  refinement of
-  the {\em Lazy Tagging and Clearing} policy
-  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).
+\item We adapt concepts of integrity and confidentiality from language-based
+security to formalize a parameterized\bcp{is that qualifier important?  what
+  are readers supposed to understand from it?} {\em stepwise stack-safety}
+property and argue that it captures the intuitive concept of stack
+safety~(\cref{sec:prelim,sec:lse}). We show that the definition can be
+extended to argument passing on the stack and to a simple coroutine system.
+\item We use property-based randomized testing to check that this stepwise
+property is actually enforced by the {\em Depth Isolation} policy
+of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}), and we formalize a
+weaker but more extensional \emph{observational} variant of the stack safety
+property that characterizes a refinement of their {\em Lazy Tagging and
+  Clearing} policy (\cref{sec:lazy}).
 \end{itemize}
@@
 We begin with a description of our setting (\cref{sec:setup}) and threat
-model (\cref{sec:threat}),
-then show how our properties apply to examples of stack unsafety
-(\cref{sec:running-example}). Then, after some technical preliminaries
-(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}\bcp{??}, \cref{sec:lazy})
-and describe testing (\cref{sec:testing})\bcp{The sections are out of order?}.
-\Cref{sec:relwork} discusses related work, and
-\cref{sec:future} sketches directions for future work.  \bcp{Where is 7?}
+model (\cref{sec:threat}), and an informal overview of our key ideas via
+examples (\cref{sec:running-example}). \Cref{sec:relwork,sec:future} discuss related and future work.
 
@@
 %
-\footnote{\bcp{Some useful text: A major challenge in characterizing stack
-    safety formally is that, in a low-level machine without built-in stacks,
-    we are at the mercy\bcp{?} of architectural\bcp{?}  choices that do not
-    relate to the abstract notion of stack safety. Some implementations may
-    have only a stack pointer while others add a frame pointer, in some
-    systems return addresses and data live in separate stacks, and so on. We
-    abstract over these choices as much as possible.  }}%
+\footnote{\bcp{Some possibly useful text: A major challenge in
+    characterizing stack safety formally is that, in a low-level machine
+    without built-in stacks, we are at the mercy\bcp{?} of
+    architectural\bcp{?}  choices that do not relate to the abstract notion
+    of stack safety. Some implementations may have only a stack pointer
+    while others add a frame pointer, in some systems return addresses and
+    data live in separate stacks, and so on. We abstract over these choices
+    as much as possible.  }}%
 %
@@
 
+\bcp{I think we are not explicit enough about the sorts of threat situations
+  we imagine we are addressing.  A canonical example (that we could spell
+  out) is a developer who wants to call out to potentially buggy or
+  malicious libraries.  The library code is going to be linked with the
+  developer's in the same address space, but the developer wants her private
+  data to be protected from both reading/exfiltrating and overwriting...}
+
 Intuitively, stack safety means that the stack frame of each caller (and
@@
 
-\paragraph*{Well-formedness and Bad Annotations}
-\bcp{What is ``well-formedness''??}
-
-\BCP{I believe we could delete this section---and never mention WBCF at all
-  (except to say that we leave it out, compared to STkTokens)---if we
-  eliminated return (and entry point) annotations.}
-\sna{I'm not sure we can escape questions about annotations, even if we have
-  just calls, but I tried to abbreviate it.}
+\bcp{We should say explicitly that we do not address attacks on the heap
+  (cite Arthur's paper and say we believe it can be combined smoothly with
+  this), direct attacks on code (we assume that code can't be over written),
+etc.}
+
+% \paragraph*{Well-formedness and Bad Annotations}
+% \bcp{What is ``well-formedness''??}
 
@@
 by combining it with the regular step function for machine states.
-%
+
 \judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
@@
                \mpstate{\mach'}{\pol'}\)}
-%
+
+\noindent
 Note that the base machine has a total step function, while the policy-enhanced machine
@@
 ``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
+\bcp{We could save space by putting two of these on a line.  And,
+  throughout, we could save space and improve beauty by removing some of the
+vertical space between inference rules.}
 
 \begin{center}
-  \judgmenttwo[Now]
+  \judgmenttwo[ Now]
               {\(f ~ \mpcstatename\)}
@@
 %
-  \judgmentthree[Skip]
+  \judgmentthree[ Skip]
                 {\(f ~ \mpcstatename\)}
@@
 %
-  \judgmentthree[Search]
+  \judgmentthree[ Search]
                 {\(\neg f ~ \mpcstatename\)}
@@
 
-\judgmentthree[Call]
+\judgmentthree[ Call]
               {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
@@
 \vspace*{-1ex}
-\judgmentthree[ReturnFound]
+\judgmentthree[ ReturnFound]
               {\(\mach \stepsto \mach' \in \target\)}
@@
 \vspace*{-3.5ex}
-\judgment[Default]
+\judgment[ Default]
          {}
@@
 addresses of the appropriate variables wherever they happen to be.
+\bcp{Several math displays in this section have too much space above...}
 
@@
 Now we can inductively define the context update function; these again apply in the order
-presented.
-
-\judgmenttwo[Call]
+presented. \bcp{Too much space between the second and third.}
+
+\judgmenttwo[ Call]
             {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
@@
 
-\judgmenttwo[ReturnFound]
+\judgmenttwo[ ReturnFound]
             {\(\mach \stepsto \mach' \in \target\)}
@@
 
-\judgment[Default]
+\judgment[ Default]
          {}
@@
 
-\judgmentthreebr[Yield]
+\judgmentthreebr[ Yield]
                 {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
@@
 
-To use QuickChick, we built random test-case generators that produce:
-\begin{enumerate}
-\item An initial RISC-V machine state, including most notably
-  the program to be executed.
-\item An initial policy state, tagging instructions corresponding to
+To use QuickChick, we built random test-case generators that produce
+% \begin{enumerate}
+% \item
+(1) an initial RISC-V machine state, including most notably
+  the program to be executed;
+  % \item
+(2)
+  an initial policy state, tagging instructions corresponding to
   ``blessed'' call or return sequences appropriately, while marking
-  all potential stack locations as $\tagNoDepth$.
-\item A code map and a stack map, accurately characterizing the behavior
+  all potential stack locations as $\tagNoDepth$, and
+% \item
+(3)
+ a code map and a stack map, accurately characterizing the behavior
   of callers and callees.
-\end{enumerate}
+% \end{enumerate}
 %

### 2021-04-16 tex/main.tex
pass on 1

@@
 The call stack is a perennial target for low-level attacks, leading to
-consequences
-from leakage or corruption of private stack data to
-control-flow hijacking. To deter or detect such attacks, a profusion of
+consequences ranging from leakage or corruption of private stack data to
+control-flow hijacking. To prevent or detect such attacks, a profusion of
 software and hardware protections have been proposed,
@@
 read or modified during a call. The more extensional observational variants
-allow a callee to read from from and write to their caller's stack frame, as
+allow a callee to read from and write to their caller's stack frame, as
 long as nothing that they read affects their observable behavior and nothing
@@
 the function activation that owns the frame and access to those locations is
-then controlled by the activation that is currently executing, validates our
+then permitted only when that activation is currently executing, validates our
 stepwise properties.  On the other hand, their more efficient \emph{Lazy
@@
 stepwise or
-observational): it violates of the temporal aspect of confidentiality in
-corner cases where data can leak across calls to the same callee.  We
+observational): it violates the temporal aspect of confidentiality in
+corner cases where data can leak across repeated calls to the same callee.  We
 propose a variant of Lazy Tagging and Clearing that \emph{does} enforce
@@
   it in their paper?  (If neither, we should make a bigger deal out of the
-  fact that we discovered it.)}
+  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
+    They refer to ``temporal security properties'' as something to do with dangling
+    pointers, but they don't define it. 
+    Then they say that Depth Isolation has weaker temporal security
+    guarantees than Static Authorities (tag=function name), but they don't explain why
+    (and I don't see why). They don't say anything explicit about the temporal
+    security properties of Lazy Tagging and Clearing, but they consider it to enforce
+    a completely different (and generally weaker) policy anyway (``data flow integrity'' vs. ``memory safety'').}}
 
@@
 with no sharing between callers and callees: all parameters and return
-values are passed in registers.  Later, we show how to refine it to allow
+values are passed in registers.  Later, we show how to refine the properties to allow
 passing of scalar stack data, granting a callee access to the caller's
@@
 \item We adapt concepts of integrity and confidentiality from language-based
-security to formalize a parameterized\bcp{is that qualifier important?  what
-  are readers supposed to understand from it?} {\em stepwise stack-safety}
+security to formalize a {\em stepwise stack-safety}
 property and argue that it captures the intuitive concept of stack
@@
 property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}), and we formalize a
+of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
+\item We formalize a
 weaker but more extensional \emph{observational} variant of the stack safety
-property that characterizes a refinement of their {\em Lazy Tagging and
-  Clearing} policy (\cref{sec:lazy}).
+property that characterizes a refinement of Roessler and DeHon's more
+performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
+\apt{I think this is worth counting as a separate contribution.}
 \end{itemize}

### 2021-04-16 tex/main.tex
Merge 7+8+comments

--- 2021-04-16 tex/main.tex
@@
 
-\section{Enforcement}
+\section{Do Micro-Policies Enforce Stack Safety?}
 \label{sec:enforcement}
@@
 itself that readers can understand our validation case study.}
+\leo{Agreed. Took a crack at it.}
 
@@
 
-In this section we sketch an existing enforcement mechanism,
-Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
-the formal stack-safety property described in the previous section; in
-fact, it enforces a stronger, inductive version of that property which strengthens the
-local state encapsulation component by checking for integrity and
-confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}
-
-%\paragraph*{Micro-Policies}
+In this section we sketch an existing enforcement mechanism, Depth
+Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
+formal stack-safety property described in the previous section; in
+fact, it enforces a stronger, inductive version of that property which
+strengthens the local state encapsulation component by checking for
+integrity and confidentiality violations at every step of the callee's
+execution trace. We validate that claim by developing a property-based
+testing framework that can quickly uncover inconsistencies between the
+enforcement mechanism and the stack-safety property.
+
+\subsection*{A Conservative Stack-Safety Enforcement Mechanism}
 %
@@
 
-\leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
-  the weeds.  I don't think this level of detail is needed, in the context
-  of a paper that is mostly about our definitions, not theirs.}
-Note that in this policy both the entry and exit sequences
-must write to each element of
-the frame, which can be quite expensive, especially for programs that
-allocate large but sparsely populated frames; we return to this point
-in \cref{sec:lazy}.
-%
-For example, the program in \cref{fig:passingsimple} manages the stack
-in this way. Here, the call pseudo-instruction to {\tt f} in line 5
-hides a blessed sequence of instructions that allocate space to save
-its return address to {\tt main} and seal the protected part of the
-stack before {\tt f}'s body starts executing at 10. Conversely, the
-return instruction in line 15 restores and jumps to the return address
-and deallocates its frame space, and unseals the corresponding parts
-of the stack. The function itself allocates and deallocates its own
-private frame space in instructions 10 and 14. Additionally, unlike in
-the simplified presentation of the policy, the caller shares the
-arguments passed to the callee through the stack, and for this creates
-and initializes (3--4) and destructs (6) the argument part of the
-callee's stack frame.
-%
+% \leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
+%   the weeds.  I don't think this level of detail is needed, in the context
+%   of a paper that is mostly about our definitions, not theirs.}
+% Note that in this policy both the entry and exit sequences
+% must write to each element of
+% the frame, which can be quite expensive, especially for programs that
+% allocate large but sparsely populated frames; we return to this point
+% in \cref{sec:lazy}.
+% %
+% For example, the program in \cref{fig:passingsimple} manages the stack
+% in this way. Here, the call pseudo-instruction to {\tt f} in line 5
+% hides a blessed sequence of instructions that allocate space to save
+% its return address to {\tt main} and seal the protected part of the
+% stack before {\tt f}'s body starts executing at 10. Conversely, the
+% return instruction in line 15 restores and jumps to the return address
+% and deallocates its frame space, and unseals the corresponding parts
+% of the stack. The function itself allocates and deallocates its own
+% private frame space in instructions 10 and 14. Additionally, unlike in
+% the simplified presentation of the policy, the caller shares the
+% arguments passed to the callee through the stack, and for this creates
+% and initializes (3--4) and destructs (6) the argument part of the
+% callee's stack frame.
+
 %\rb{Note that we have also considered and implemented more flexible schemes
@@
 
-\section{Randomized Testing}
+\subsection*{Validation through Random Testing}
 \label{sec:testing}
@@
 implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code\bcp{right?}; we re-implemented it in Coq targeting
+and targeted ARM machine code
+%\bcp{right?}\leo{yeah}
+; we re-implemented it in Coq targeting
 RISC-V).  These errors ranged from trivial typos to

### 2021-04-16 tex/main.tex
Tried to elucide confidentiality with more exampels

--- 2021-04-16 tex/main.tex
@@
 
-Next we consider {\em stepwise stack confidentiality}. There are a number of preliminary
-concepts we must discuss here -- confidentiality quantifies over {\em variants} of the
-initial state of a call, and requires that the traces from them move in {\em lockstep}.
+Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
+nested: we will quantify over subtraces that represent whole calls. For each such
+subtrace, we will quantify over {\em variants} of its initial state. Then we require
+that the original, {\em primary} traces move in {\em lockstep} with the variant.
+
+\definition \(\MPCT^*\) is the set of all traces of calls, defined as
+\[\begin{align*}
+\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
+\text{ is a reachable segment}\} \hspace{1em} &
+\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth
+\end{align*}\]
 
@@
 
-\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to
-\emph{change together}, written \(\mach,\mach' \diamond \nach,\nach'\), if,
-for all components \(\component\) such that
-\(\mach ~ \component \not = \mach' ~ \component\) or
-\(\nach ~ \component \not = \nach' ~ \component\),
-\(\mach' ~ \component = \nach' ~ \component\).
+\definition Let \(\Delta(\mach,\mach')\) be the set of set of components \(\component\)
+such that \(\mach ~ \component \not = \mach' ~ \component\).
+
+\definition Between two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\), we
+say that they {\em lock} if for all components \(\components\) in
+\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
+\(\mach' ~ \component = \nach' ~ \component\). We write this \((\mach,\mach') \diamond
+(\nach,\nach')\).
 
@@
 if their adjacent pairs of states change together. We will represent this relation with
-\(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
-\sna{It's a new notation, we use this enough that it seems worth picking an
-  evocative symbol.}\bcp{Sorry -- I meant the notational convention of representing a
-coinductive relation with inference rules in this way, not the symbol...}
-We define it as a coinductive relation:
+\(\machT \doteq \nachT\). We define it as a coinductive relation:
 
@@
 \judgment{}
-         {\(\mach \doteq \nach\)}
+         {\(\mpcstatename \doteq \npcstatename\)}
 \end{minipage}
 \begin{minipage}{.6\textwidth}
-\judgmenttwo{\(\mach, \head(\machT) \diamond \nach, \head(\nachT)\)}
+\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
             {\(\machT \doteq \nachT\)}
-            {\(\mach \machT \doteq \nach \nachT\)}
+            {\(\notfinished{(\mach,\_,\_)}{\machT} \doteq \notfinished{(\nach,\_,\_)}{\nachT}\)}
 \end{minipage}
 
-\definition For any depth \(\depth\), define \(f_\depth ~ \mach ~ (\_, \rts)\) to be
-\(|\rts| \geq \depth\). Then each reachable segment \(\MPCT \langle f_\depth \rangle\)
-corresponds to a trace of a function from its call to its return.
-Given such a segment \(S\), let \(\mach,\_,(\domMap,\_)\)  = \(\head(S)\) and let
-\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
-A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
-with associated \(\mach\) and \(\components\) as above, and
+\definition {\bf Stepwise Stack Confidentiality}
+
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
+let \(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \outside\).
+A system enjoys \emph{stepwise stack confidentiality} if,
 for all \(\nach\) such that \(\mach \approx_\components \nach\) and
- \(\nach \hookrightarrow \NPCT\),
- \(\MPCT\langle f_\depth\rangle \doteq \NPCT | \neg f_\depth \).\apt{I found the previous
-   version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
- \sna{It's right, but I need to rethink how the M<f> notation works...}
+\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\).
+
+\vspace{\afterdisplayskip}
+
+Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
+of the components in \(\components\), and in addition their subtraces are also in
+\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
+will contain these calls and returns as well, so its return stack pushes and pops
+in lockstep until it breaks at the final return.
 
@@
 
-\section{Do Micro-Policies Enforce Stack Safety?}
+\section{Enforcement}
 \label{sec:enforcement}
@@
 itself that readers can understand our validation case study.}
-\leo{Agreed. Took a crack at it.}
 
@@
 
-In this section we sketch an existing enforcement mechanism, Depth
-Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
-formal stack-safety property described in the previous section; in
-fact, it enforces a stronger, inductive version of that property which
-strengthens the local state encapsulation component by checking for
-integrity and confidentiality violations at every step of the callee's
-execution trace. We validate that claim by developing a property-based
-testing framework that can quickly uncover inconsistencies between the
-enforcement mechanism and the stack-safety property.
-
-\subsection*{A Conservative Stack-Safety Enforcement Mechanism}
+In this section we sketch an existing enforcement mechanism,
+Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
+the formal stack-safety property described in the previous section; in
+fact, it enforces a stronger, inductive version of that property which strengthens the
+local state encapsulation component by checking for integrity and
+confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}
+
+%\paragraph*{Micro-Policies}
 %
@@
 
-% \leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
-%   the weeds.  I don't think this level of detail is needed, in the context
-%   of a paper that is mostly about our definitions, not theirs.}
-% Note that in this policy both the entry and exit sequences
-% must write to each element of
-% the frame, which can be quite expensive, especially for programs that
-% allocate large but sparsely populated frames; we return to this point
-% in \cref{sec:lazy}.
-% %
-% For example, the program in \cref{fig:passingsimple} manages the stack
-% in this way. Here, the call pseudo-instruction to {\tt f} in line 5
-% hides a blessed sequence of instructions that allocate space to save
-% its return address to {\tt main} and seal the protected part of the
-% stack before {\tt f}'s body starts executing at 10. Conversely, the
-% return instruction in line 15 restores and jumps to the return address
-% and deallocates its frame space, and unseals the corresponding parts
-% of the stack. The function itself allocates and deallocates its own
-% private frame space in instructions 10 and 14. Additionally, unlike in
-% the simplified presentation of the policy, the caller shares the
-% arguments passed to the callee through the stack, and for this creates
-% and initializes (3--4) and destructs (6) the argument part of the
-% callee's stack frame.
-
+\leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
+  the weeds.  I don't think this level of detail is needed, in the context
+  of a paper that is mostly about our definitions, not theirs.}
+Note that in this policy both the entry and exit sequences
+must write to each element of
+the frame, which can be quite expensive, especially for programs that
+allocate large but sparsely populated frames; we return to this point
+in \cref{sec:lazy}.
+%
+For example, the program in \cref{fig:passingsimple} manages the stack
+in this way. Here, the call pseudo-instruction to {\tt f} in line 5
+hides a blessed sequence of instructions that allocate space to save
+its return address to {\tt main} and seal the protected part of the
+stack before {\tt f}'s body starts executing at 10. Conversely, the
+return instruction in line 15 restores and jumps to the return address
+and deallocates its frame space, and unseals the corresponding parts
+of the stack. The function itself allocates and deallocates its own
+private frame space in instructions 10 and 14. Additionally, unlike in
+the simplified presentation of the policy, the caller shares the
+arguments passed to the callee through the stack, and for this creates
+and initializes (3--4) and destructs (6) the argument part of the
+callee's stack frame.
+%
 %\rb{Note that we have also considered and implemented more flexible schemes
@@
 
-\subsection*{Validation through Random Testing}
+\section{Randomized Testing}
 \label{sec:testing}
@@
 implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code
-%\bcp{right?}\leo{yeah}
-; we re-implemented it in Coq targeting
+and targeted ARM machine code\bcp{right?}; we re-implemented it in Coq targeting
 RISC-V).  These errors ranged from trivial typos to

### 2021-04-16 tex/main.tex
Fixes some issues
pass up through 4

--- 2021-04-16 tex/main.tex
@@
 control flow and observing when we reach a {\em return target}. For a given state annotated
-as a call, in a standard calling convention, another state fulfils its return target if
+as a call, another state fulfils its return target under a standard calling convention if
 it has the same stack pointer and the program counter is at the instruction following
@@
 
+\apt{Suddenly we are talking about alternative enforcement mechanisms for the first time.
+  This surely needs \emph{some} foreshadowing in the abstract and intro?}
 Dividing policy state from machine state helps us reason about
@@
 \caption{An integrity violation\bcp{Can we align the C-like snippets with
-    the corresponding assembly?}  \ifaftersubmission\bcp{I wonder if we
+    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
+      at places in both the C and asssembly, making them harder to read
+  and eating more space.}}  \ifaftersubmission\bcp{I wonder if we
     could make the figures easier to read by putting very light, differently
@@
 The machine is word-addressed. The stack grows upward in
-memory and {\tt \$SP} points to the first unused word above the top of the stack.
+memory and {\tt \%sp} points to the first unused word above the top of the stack.
 The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
@@
 
-Meanwhile, confidentiality intuitively states that a callee is unaware\bcp{??}
-of the context of its call (except for its arguments and any global
+Meanwhile, confidentiality intuitively states that a callee is insensitive to 
+the context of its call (except for its arguments and any global
 variables).
@@
 the primary state and its variant will step in lockstep until they return. To step in lockstep
-means that if both states step -- that is, neither the primary nor the variant failstops --
-any {\em system component} (such as memory cell, register, program counter) that changes
-in one step has a matching value at the end of the other step. And when we
-extend the model with a notion of external output, values that are outputted will need
+means that if both states step---that is, neither the primary nor the variant failstops---any
+{\em system component} (such as memory cell, register, program counter) that changes
+in one step has a matching value at the end of the other step. When we later
+extend the model with a notion of observable output, values that are emitted will need
 to match as well.
@@
   \centering
-  \begin{subfigure}{.3\textwidth}
-{\small
+  \begin{subfigure}{.2\textwidth}
+{\footnotesize
 \begin{verbatim}
@@
 \end{subfigure}
-  \begin{subfigure}{.65\textwidth}
-{\small
+  \begin{subfigure}{.4\textwidth}
+{\footnotesize
 \begin{verbatim}
    stash: .word 0
-1  main:  add $1,%sp        ; allocate frame
-2         mov $1,-1(%sp)    ; initialize x
+1  main:  add $1,%sp     ; alloc frame
+2         mov $1,-1(%sp) ; init x
 3         call f
@@
 5         call f
-6         mov -1(%sp),%ra   ; set return value
-7         sub $1,%sp        ; deallocate frame
+6         mov -1(%sp),%ra ; set ret val
+7         sub $1,%sp     ; dealloc frame
 6         ret
 
-10 f:     add $1,%sp        ; allocate frame
-11        cmp stash, $0     ; compare
-12        breq #15          ; branch if equal
-13        mov -2(%sp),stash ; update stash
+10 f:     add $1,%sp     ; alloc frame
+11        cmp stash, $0  ; stash = 0? 
+12        breq #15       ; branch if so
+13        mov -2(%sp),stash ; set stash
 14        jmp #17
-15        mov stash,-2(%sp) ; retrieve stash
+15        mov stash,-2(%sp) ; get stash
 16        mov $0,stash
-17        sub $1,%sp        ; deallocate frame
+17        sub $1,%sp     ; dealloc frame
 18        ret
@@
   \end{subfigure}
-  \begin{subfigure}{.4\textwidth}
+  \begin{subfigure}{.25\textwidth}
     \begin{tabular}{l l l}
@@
       \vspace{.5em} &
-      \memory[3em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#4}
+      \hspace*{-1.5em}
+      \memory[1.2em]{1}{\mainsealc}[Target]%
+      \MemoryLabel{-2.2em}{0.75em}{\#4}
       \\
@@
       \MemoryLabel{-11em}{0.75em}{\#6} &
-      \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
+      \hspace*{-1em}
+      \memory[1.2em]{1}{\mainsealc}
+      \MemoryLabel{-1.2em}{0.75em}{\#6}
       \\
@@
       \MemoryLabel{-11em}{0.75em}{\#4} &
-      \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
+      \hspace*{-1em}
+      \memory[1.2em]{1}{\mainsealc}
+      \MemoryLabel{-1.2em}{0.75em}{\#6}
       \\
@@
       \MemoryLabel{-15em}{0.75em}{-1} &
-      \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
+      \hspace*{-1em}
+      \memory[1.2em]{1}{\mainsealc}
+      \MemoryLabel{-1.2em}{0.75em}{\#6}
       \\
@@
       \MemoryLabel{-15em}{0.75em}{\bf 1} &
-      \memory[3em]{1}{\mainsealc}
-      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6} \\
+      \hspace*{-1em}
+      \memory[1.2em]{1}{\mainsealc}
+      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
     \end{tabular}
   \end{subfigure}
-  \caption{A control-flow attack \bcp{wonder if we can cram the bottom part
-      onto the same row as the rest.  Might need to make things smaller,
-      spill into the margin, etc., but we really need the space.  (Another
-      option is to make it a separate figure and flow text around it, but
-      this will save less space.)}}
+  \caption{A control-flow attack}
   \label{fig:controlflow}
@@
 Our stack safety properties do not assume that control flow follows a typical structure
-of matching sets calls and returns nested within on another.
+of matching sets of calls and returns nested within on another.
 Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.

### 2021-04-16 tex/main.tex

--- 2021-04-16 tex/main.tex
@@
 \definition \(\MPCT^*\) is the set of all traces of calls, defined as
-\[\begin{align*}
+\[\begin{aligned}
 \bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
 \text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth
-\end{align*}\]
+\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
+\end{aligned}\]
 
@@
 
-\vspace{\afterdisplayskip}
+\vspace{\belowdisplayskip}
 

### 2021-04-16 tex/main.tex
pass on 5

@@
 %
-    \[\component \in \COMPONENTS ::= \WORDS + \REGS \]
-%
-    \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
+    \[\component \in \COMPONENTS ::= \WORDS + \REGS  \hspace{1in} 
+    \mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
 %
@@
 can see the machine state, but cannot modify it. The policy also provides a
-well-formedness condition, \(\WF (\mpstatename \in \MACHS \times \POLS)\), describing
+well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
 initial states that are properly configured.
@@
 
-Separate from the machine state, we will keep track of additional context
+Separate from the machine state, we keep track of additional context
 information representing the history of the run. This has no influence on execution, it merely
@@
 \(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
-\times \rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
+\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
 \(\codemap \in \WORDS \rightharpoonup \ANNS\)
@@
 
+\apt{Some words have been lost here.}
 individual calls from a trace by taking subtraces in which the stack is
@@
 is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
-where \(\WF ~ \mach,\pol\), \(\mpcstatename \hookrightarrow \MPCT\)
+where \((\mach,\pol) \in \WF\), \(\mpcstatename \hookrightarrow \MPCT\)
 and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace

### 2021-04-16 tex/main.tex
pass through 6.1

@@
 ``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
-\bcp{We could save space by putting two of these on a line.  And,
-  throughout, we could save space and improve beauty by removing some of the
+\bcp{Throughout, we could save space and improve beauty by removing some of the
 vertical space between inference rules.}
@@
 \begin{center}
+\begin{minipage}{.38\textwidth}
   \judgmenttwo[ Now]
@@
               {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
-%
+\end{minipage}
+%
+\begin{minipage}{.58\textwidth}
   \judgmentthree[ Skip]
@@
                 {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
+\end{minipage}
 %
@@
 
-\definition \(\MPCT^*\) is the set of all traces of calls, defined as
+\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
 \[\begin{aligned}
@@
 
-\definition Let \(\Delta(\mach,\mach')\) be the set of set of components \(\component\)
+\definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
+is the set of components \(\component\)
 such that \(\mach ~ \component \not = \mach' ~ \component\).
 
-\definition Between two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\), we
-say that they {\em lock} if for all components \(\components\) in
+\definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
+ {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all components \(\component\) in
 \(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
-\(\mach' ~ \component = \nach' ~ \component\). We write this \((\mach,\mach') \diamond
-(\nach,\nach')\).
-
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
-if their adjacent pairs of states change together. We will represent this relation with
-\(\machT \doteq \nachT\). We define it as a coinductive relation:
+\(\mach' ~ \component = \nach' ~ \component\).
+
+\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep},
+written \(\machT \doteq \nachT\), 
+if their adjacent pairs of states change together.  We define this as a coinductive relation:
 
@@
 
+\apt{This is way better up until the final definition, which still needs more unpacking in words.}
 Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
@@
 The argument {\tt a} has
-its value copied from {\tt x}, and we note that {\tt f} can read it -- because it is marked
+its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
 as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
@@
 {\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++!)
+{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
 
@@
 at the appropriate offset from the stack pointer, while passing by reference gives the
-addresses of the appropriate variables wherever they happen to be.
-\bcp{Several math displays in this section have too much space above...}
-
+addresses of the appropriate variables wherever they happen to be.\apt{Why can't
+  these also be given as offsets from stack pointer (as in the example)?}
 \[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-
 And we extend the set of domains to include passed components,
 which are labeled with the range of depths of the caller that has passed them.
-
 \[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
@@
 \times (\rts \in \listT (\target \subseteq \MACHS)) \]
-
 We straightforwardly describe how the sealing convention and passing predicate combine to
 update a domain map when pushing it, where \(\depth\) is an integer depth.
-
 \[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =

### 2021-04-16 tex/main.tex
Put control flow back a little, fixed up properties

--- 2021-04-16 tex/main.tex
@@
 
-\newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}
 \newcommand*{\rsp}{\textsc{sp}}
@@
 Our core properties are \emph{stepwise stack integrity},
-\emph{stepwise stack confidentiality}, and \emph{return to caller}\BCP{I
-  gave it a name different from WBCF, but I could be talked out of that.  In
-any case, it now needs to be discussed / defined in several places below.}.
+\emph{stepwise stack confidentiality}, and \emph{return to caller}.
 %
@@
 immediate callees (which, in turn, need protection from their immediate
-callees, etc.).
-
-% This approach is novel in that the
-% confidentiality\bcp{only?} property applies to many nested sub-traces within a single execution,
-% as opposed to traces of entire executions.
-% {\em Integrity} says that each caller's local data is protected from
-% modification by callees,
-% motivated by the reasoning principle that local data persists across a
-% call.
+callees, etc.). They are also stronger than LSE; even in the absence of the
+control flow guarantees of {\em return to caller}, they guarantee callers'
+data safety in the face of control-flow attacks (see \cref{fig:control} and discussion).
+This reflects a cleaner division between data protection and control-flow protection
+than offered by LSE and WBCF.
 
@@
 they write affects the observable behavior of the caller after they return.
-\bcp{Please check: is that accurate?}
 
@@
 % not as a goal of the property.\bcp{Oof.  We need to do better than that!})
-
-\bcp{I trimmed away a ton of text here that I thought was getting too much
-  into the weeds for the introduction.  However, in the process I lost the
-  explanation of how our properties are a bit stronger than LSE.  One way to
-put it back would be simply to add to the contributions a claim that we are
-a bit stronger than them, with a pointer to where we discuss this in detail.}
 
@@
 security to formalize a {\em stepwise stack-safety}
-property and argue that it captures the intuitive concept of stack
+property and argue that it captures the intuitive concept of strict stack
 safety~(\cref{sec:prelim,sec:lse}). We show that the definition can be
@@
 performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
-\apt{I think this is worth counting as a separate contribution.}
 \end{itemize}
@@
 control flow and observing when we reach a {\em return target}. For a given state annotated
-as a call, another state fulfils its return target under a standard calling convention if
+as a call, in a standard calling convention, another state fulfils its return target if
 it has the same stack pointer and the program counter is at the instruction following
@@
 
-\apt{Suddenly we are talking about alternative enforcement mechanisms for the first time.
-  This surely needs \emph{some} foreshadowing in the abstract and intro?}
 Dividing policy state from machine state helps us reason about
@@
 \caption{An integrity violation\bcp{Can we align the C-like snippets with
-    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
-      at places in both the C and asssembly, making them harder to read
-  and eating more space.}}  \ifaftersubmission\bcp{I wonder if we
+    the corresponding assembly?}  \ifaftersubmission\bcp{I wonder if we
     could make the figures easier to read by putting very light, differently
@@
 The machine is word-addressed. The stack grows upward in
-memory and {\tt \%sp} points to the first unused word above the top of the stack.
+memory and {\tt \$SP} points to the first unused word above the top of the stack.
 The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
@@
 
-Meanwhile, confidentiality intuitively states that a callee is insensitive to 
-the context of its call (except for its arguments and any global
+Meanwhile, confidentiality intuitively states that a callee is unaware\bcp{??}
+of the context of its call (except for its arguments and any global
 variables).
@@
 the primary state and its variant will step in lockstep until they return. To step in lockstep
-means that if both states step---that is, neither the primary nor the variant failstops---any
-{\em system component} (such as memory cell, register, program counter) that changes
-in one step has a matching value at the end of the other step. When we later
-extend the model with a notion of observable output, values that are emitted will need
+means that if both states step -- that is, neither the primary nor the variant failstops --
+any {\em system component} (such as memory cell, register, program counter) that changes
+in one step has a matching value at the end of the other step. And when we
+extend the model with a notion of external output, values that are outputted will need
 to match as well.
@@
   \centering
-  \begin{subfigure}{.2\textwidth}
-{\footnotesize
+  \begin{subfigure}{.3\textwidth}
+{\small
 \begin{verbatim}
@@
 \end{subfigure}
-  \begin{subfigure}{.4\textwidth}
-{\footnotesize
+  \begin{subfigure}{.65\textwidth}
+{\small
 \begin{verbatim}
    stash: .word 0
-1  main:  add $1,%sp     ; alloc frame
-2         mov $1,-1(%sp) ; init x
+1  main:  add $1,%sp        ; allocate frame
+2         mov $1,-1(%sp)    ; initialize x
 3         call f
@@
 5         call f
-6         mov -1(%sp),%ra ; set ret val
-7         sub $1,%sp     ; dealloc frame
+6         mov -1(%sp),%ra   ; set return value
+7         sub $1,%sp        ; deallocate frame
 6         ret
 
-10 f:     add $1,%sp     ; alloc frame
-11        cmp stash, $0  ; stash = 0? 
-12        breq #15       ; branch if so
-13        mov -2(%sp),stash ; set stash
+10 f:     add $1,%sp        ; allocate frame
+11        cmp stash, $0     ; compare
+12        breq #15          ; branch if equal
+13        mov -2(%sp),stash ; update stash
 14        jmp #17
-15        mov stash,-2(%sp) ; get stash
+15        mov stash,-2(%sp) ; retrieve stash
 16        mov $0,stash
-17        sub $1,%sp     ; dealloc frame
+17        sub $1,%sp        ; deallocate frame
 18        ret
@@
   \end{subfigure}
-  \begin{subfigure}{.25\textwidth}
+  \begin{subfigure}{.4\textwidth}
     \begin{tabular}{l l l}
@@
       \vspace{.5em} &
-      \hspace*{-1.5em}
-      \memory[1.2em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-2.2em}{0.75em}{\#4}
+      \memory[3em]{1}{\mainsealc}[Target]%
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#4}
       \\
@@
       \MemoryLabel{-11em}{0.75em}{\#6} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
       \\
@@
       \MemoryLabel{-11em}{0.75em}{\#4} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
       \\
@@
       \MemoryLabel{-15em}{0.75em}{-1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
       \\
@@
       \MemoryLabel{-15em}{0.75em}{\bf 1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
+      \memory[3em]{1}{\mainsealc}
+      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6} \\
     \end{tabular}
   \end{subfigure}
-  \caption{A control-flow attack}
+  \caption{A control-flow attack \bcp{wonder if we can cram the bottom part
+      onto the same row as the rest.  Might need to make things smaller,
+      spill into the margin, etc., but we really need the space.  (Another
+      option is to make it a separate figure and flow text around it, but
+      this will save less space.)}}
   \label{fig:controlflow}
@@
 Our stack safety properties do not assume that control flow follows a typical structure
-of matching sets of calls and returns nested within on another.
+of matching sets calls and returns nested within on another.
 Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
@@
 
+\vspace*{-1ex}
+\judgmentthree[ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
 \vspace*{-3.5ex}
@@
 in lockstep until it breaks at the final return.
+
+\subsection{Control Flow and Return-to-Caller}
+\label{sec:RTC}
+
+Together, integrity and confidentiality protect a caller's data until its callee returns
+to it. But it is possible that an intermediate caller is never returned to, but returned
+{\em past}, to a deeper caller. Fundamentally, this is a control-flow problem, not a
+data-protection problem. From that caller's perspective, it is no different from the
+callee diverging. But a lower level caller might have been relying on the intermediate
+caller to finish its computation, so it is worth protecting control flow even from
+attacks that do not violate data protections. We formalize this as {\em return-to-caller}.
+
+\definition {\bf Return-To-Caller}
+
+A system enjoys {\em return-to-caller} when, for every pair of reachable states
+\(\mpcstatename = (\_,\_,(\_,\rts))\) and
+\(\mpcstatename' = (\_,\_,(\_,\rts'))\) such that
+\(\mpcstatename \stepsto \mpcstatename'\), \(|\rts| - 1 \leq |\rts'| \leq |\rts| + 1\).
 
@@
 
-\judgmenttwo[ Call]
+\judgmenttwo[Call]
             {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
@@
 
+\vspace*{-1ex}
 \judgmenttwo[ ReturnFound]
@@
 
+\vspace*{-1ex}
+\judgmentthree[ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\vspace*{-3.5ex}            
 \judgment[ Default]
@@
 
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
+be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
+or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
+\(|\rts| > \depth'\).
+
+\definition {\bf Stepwise Stack Integrity (Passing)}
+
+Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
 \(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \sealed{\_}\) or \(\passed{\depth,\depth'}\) where \(|\rts| < \depth\)
-or \(|\rts| > \depth'\). A system enjoys {\em stepwise stack integrity with passing} if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \unsealed\) or \(\passed{\depth,\depth'}\) where
-\(\depth \leq |\rts| \leq \depth'\) .
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
-{\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
+A system enjoys {\em stepwise stack integrity with passing} if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
+for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
+
+\definition {\bf Stepwise Stack Confidentiality (Passing)}
+
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
+let \(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
+and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
+\emph{stepwise stack confidentiality} if,
+for all \(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
 \(\MPCT \doteq \NPCT\).

### 2021-04-16 tex/main.tex
Swept through in reverse, added future work, some related work

@@
 control flow guarantees of {\em return to caller}, they guarantee callers'
-data safety in the face of control-flow attacks (see \cref{fig:control} and discussion).
+data safety in the face of control-flow attacks (see \cref{sec:control}).
 This reflects a cleaner division between data protection and control-flow protection
@@
 
-Unless otherwise noted\bcp{where is it ``otherwise noted''??}, colored boxes
-indicate an assignment of the address to a
+Colored boxes indicate an assignment of the address to a
 {\em domain} -- a label reflecting a particular set of access rights. Domains
@@
 
-\begin{subfigure}{.2\textwidth}
+\begin{subfigure}[t]{.2\textwidth}
   {\small
@@
   int z;
-  if (z == 5)
-    return 10;
-  else
-    return 20;
+  return z;
 }
@@
 \end{subfigure}
-\begin{subfigure}{.7\textwidth}
+\begin{subfigure}[t]{.7\textwidth}
   {\small
@@
 20 g:    add $1,%sp      ; allocate frame
-21       cmp $5,-1(%sp)  ; compare
-22       brneq #25       ; branch if not equal
-23       mov $10,%ra     ; set return value
-24       jmp #26
-25       mov $20,%ra     ; set return value
-26       sub $1,%sp      ; deallocate frame
-27       ret
+21       mov -1(%sp),%ra ; violation!
+22       sub $1,%sp      ; deallocate frame
+23       ret
 }
@@
   ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
+  \MemoryLabel{-19em}{0.75em}{62}
+  \MemoryLabel{-10em}{0.75em}{105}
   \\
@@
   ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
   \MemoryLabel{-10em}{0.75em}{\(5\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
   \\
@@
   \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-10em}{0.75em}{\bf 5}
  % \MemoryLabel{-6em}{0.75em}{0}
@@
   ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_3\)}
-  \MemoryLabel{-10em}{0.75em}{\(v_4\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_5\)}
-  \\ &
-  {\bf 23}
-  ~$\cdots$
-  &
-  {\bf 25}
-  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{36}
+  \MemoryLabel{-10em}{0.75em}{\bf 8}
+  \MemoryLabel{-6em}{0.75em}{90}
   \\
@@
 
-\caption{Another confidentiality violation \bcp{Again, let's align the C
-    code with the assembly, etc.}}
+\caption{Another confidentiality violation}
 \label{fig:conf2}
@@
 
+\label{sec:controlflow}
+
 %From the classic stack smashing attack to return oriented programming,
@@
 to do more than just execute a return-labeled instruction: it must actually reach the
-appropriate {\em return target} in {\tt main}. A return target is a predicate on states,
-in this case one that holds on any state whose program counter points to the correct instruction.
-We introduce a new piece of context: we keep a stack of return targets, and a callee is
+appropriate return target in {\tt main}.
+We add a new piece of context to our diagram: we keep a stack of return targets, and a callee is
 only considered to have returned when it reaches a state that matches the caller's target.
@@
 that matches the topmost return target. If neither is true, the context remains unchanged.
-%
-\bcp{(Moved this here, BTW.)}Recall that we keep track of which addresses are
-annotated as calls, with a ``code map'' $\codemap \in \CODEMAPS \subseteq
-\WORDS$.
-\apt{Doesn't match previous definition any more.}
+Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
 
 \judgmentthree[ Call]
-              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
+              {\(\codemap ~ (\mach ~ \PCname)\)}
               {\(\components = \sealcon ~ \mach\)}
@@
 
-\definition A system enjoys {\em stepwise stack integrity} if, for any reachable state
+\definition {\bf Stepwise Stack Integrity}
+
+A system enjoys {\em stepwise stack integrity} if, for any reachable state
 \(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
@@
 let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\).
+\(\domMap ~ \component = \outside\). That is, at the start of some call, we take every
+component in the stack.
 A system enjoys \emph{stepwise stack confidentiality} if,
 for all \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\).
+\(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\) -- if every variant (with the same policy state)
+steps in lockstep until both traces have reached a point whose depth is
+below the starting one. This is a return.
 
@@
 
-\apt{This is way better up until the final definition, which still needs more unpacking in words.}
 Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
@@
 \(\mpcstatename' = (\_,\_,(\_,\rts'))\) such that
-\(\mpcstatename \stepsto \mpcstatename'\), \(|\rts| - 1 \leq |\rts'| \leq |\rts| + 1\).
+\(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
+\(|\rts'| = |\rts|\)
 
@@
 A call annotation now includes a predicate relating machine states to the components
-that should be passed -- when passing by value, this will typically give the memory cells
-at the appropriate offset from the stack pointer, while passing by reference gives the
-addresses of the appropriate variables wherever they happen to be.\apt{Why can't
-  these also be given as offsets from stack pointer (as in the example)?}
+that should be passed -- when passing by value, this will typically be freshly allocated
+cells at the top of the frame, while passing by reference just gives the
+addresses of the appropriate variables.
 \[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
@@
 
-\paragraph*{Deferred Stepwise Integrity}
+\paragraph*{Deferring Integrity Checks}
 
@@
 
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+\definition {\bf Deferred Stack Integrity}
+
+For any \(\MPCT \in \MPCT^*\),
 let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \sealed{\_}\).
-
-Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\) -- meaning that the
-call does return -- let \(\components'\) be the set of components that differ
-between \(\mach\) and \(\mach'\). Take any \(\nach\) such that
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \sealed{\_}\).
+Then if \(\last(\MPCT) = \mach',\pol',\context'\), let \(\components' = \Delta(\mach,\mach')\).
+ A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
 \(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
-\(\nach \hookrightarrow \NPCT\). A system enjoys
-{\em deferred stepwise integrity} if for all such \(\MPCT\) and \(\NPCT\),
-\(\MPCT \doteq \NPCT\).
+\(\nach \hookrightarrow \NPCT\), \(\MPCT \doteq \NPCT\).
+
+\vspace{\belowdisplayskip}
 
@@
 Integrity is already straightforward, simply weakening the condition that must hold
-after the return.
-
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+after the return -- instead of preventing dangerous writes from being read, we
+prevent them from being leaked.
+
+\definition {\bf Observational Stack Integrity}
+
+For any call trace \(\MPCT \in \MPCT^*\),
 let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \sealed{\_}\).
-
-Then, again supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
+\(\components\) be the set of components \(\component\)
+such that \(\domMap ~ \component = \sealed{\_}\).
+
+Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
 let \(\components'\) be the set of components that differ
-between \(\mach\) and \(\mach'\). Take any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
-\(\nach \hookrightarrow \NPCT\). A system enjoys
-{\em observational integrity} if for all such \(\MPCT\) and \(\NPCT\),
-\(\obsof(\MPCT) \simeq \obsof(\NPCT)\).
-
-We amend our intuitive statement to: ``No visible reads after dangerous writes.''
+between \(\mach\) and \(\mach'\). A system enjoys
+{\em observational integrity} if for any \(\nach\) such that
+\(\mach' \approx_{\components'} \nach\), \((\mach,\pol',\context') \hookrightarrow \MPCT\), and
+\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT) \simeq \obsof(\NPCT)\).
+
+\vspace{\belowdisplayskip}
 
@@
 \(\bar{\Diamond}(\components,\mach,\mach',\nach,\nach')\), is the set of all components
-\(\component \in \components\) such that if
-\(\mach ~ \component \not = \mach' ~ \component\) or
-\(\nach ~ \component \not = \nach' ~ \component\), then
-\(\mach' ~ \component = \nach' ~ \component\).
-
-\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
-\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
+\(\mach' ~ \component \not = \nach' ~ \component\).
+
+\definition {\bf Observational Stack Confidentiality}
+
+For any call trace \(\MPCT \in \MPCT^*\),
 let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
-Let \(\nach\) be a variant such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth \).
-
-A system enjoys {\em observational confidentiality} if for all such
+\(\components\) be the set of components \(\component\)
+such that \(\domMap ~ \component \not = \outside\).
+
+A system enjoys {\em observational confidentiality} if for all
+\(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\nach \hookrightarrow \NPCT | \neg f_\depth \),
 \(\MPCT\) and \(\NPCT\), the following conditions all hold:
@@
 
-\sna{More to write here. The gist is: this is the extensional property.
-  The one that we would tentatively propose as a real characterization of stack
-  safety.}
+Together with return-after-call, these two properties give us
+{\em Observable Stack Safety}. This is a weaker property
+than the stepwise one, and also much more extensional.
+We can think of stepwise, deferred stepwise, and observational as
+three points along a spectrum from eager to lazy. The first two
+are useful for understanding particular enforcement mechanisms,
+but if we had to choose a single property to represent stack safety,
+it would be observational state safety.
 
@@
 the connection to the eager policy, we can identify when lazy policies
-    miss violations that can become visible. The optimized lazy tag policy from
-    \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
+miss violations that can become visible. The Lazy Tagging and Clearing
+micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
 at which it was written, and enforces that it must be read from the
 same depth. However, this allows violations of observable stack safety!
-    Say we have functions {\tt foo}, {\tt bar} and {\tt baz}, where {\tt foo}
-    calls {\tt bar}, and then after {\tt bar} returns {\tt foo} calls
-    {\tt baz}. Since {\tt bar} and {\tt baz} are at the same stack depth, if
-    {\tt bar} writes to {\tt foo}'s frame, {\tt baz} can then read the tainted
-    write and use it to create an observable difference that violates both
-    observable integrity and confidentiality without violating the policy.
-
-    The lazy policy can be repaired if, instead of tagging each function
+Specifically, a callee might write into the stack and leave tags behind
+for a future callee at the same depth to read. This violates confidentiality and,
+if the writes were in the caller's frame, integrity.
+
+Lazy Tagging and Clearing can be repaired if, instead of tagging each function
 activation with its depth in the stack, we generate a fresh activation
@@
 however shared by all activations of the same function in order to obtain
-    good cacheability of tag rules. Static Authorities comes closer to
-    observable stack safety, though the example above would still exist
-    if {\tt bar} and {\tt baz} were merged into a single function that tracked
-    whether it had been called previously and changed its behavior accordingly.
-    \ifaftersubmission\apt{More blatently, static authorities doesn't work for recursive calls.}\fi
+good cacheability of tag rules, but that policy allows leaks between
+instances of the same function.
+
+Lazy Per-activation Tagging and Clearing ends up enforcing the combination
+of deferred stepwise stack integrity and stepwise confidentiality -- we
+don't actually need a deferred confidentiality property here. Therefore,
+it also enforces observational stack safety, so we declare it stack safe.
+Our properties have allowed us to reject an enforcement mechanism as
+not really stack safe, and describe a fix.
 
@@
 observations is limited to cotermination.
-
-We conjecture that the linear capability machines introduced by
-\citep{Skorstengaard+19}, in combination with the StkTokens calling convention,
-can be modeled in our framework and satisfy our definition of stack safety in
-\cref{sec:lse}. Further, we expect them to satisfy the even stronger
-property developed in \cref{sec:enforcement}. Proving (or testing) these conjectures would
-involve exposing their formal definitions of local state encapsulation and
-well-bracketed control flow, which are built-in as part of the
-semantics of the capability machines and not given explicitly.
-The main practical limitation of StkTokens
-is its reliance on linear capabilities, as it is unclear now they could be added
-to practical capability machines, especially in terms of efficiency (previous
-work on micro-policies \citep{yannis-report} has shown how to use that framework
-to implement linear return capabilities.\bcp{But IIUC the Dover realization
-  of micro-policies can't do this...?})\rb{Correct}
-%
-In recent work, \citet{Georges+21} address these efficiency concerns
-with the introduction of a new type of uninitialized capability that
-avoids the need to clear the stack while being more amenable to
-practical hardware implementation, although like
-\citet{Skorstengaard+19b} it lacks a general stack safety theorem,
-instead using a similar logical relation to reason about the safety of
-individual programs.
+\sna{Some of this seems like more detail than we need, and we need to
+  compare ourselves to their list of criteria.}
 
@@
 
+We have evaluated Depth Isolation, Lazy Tagging, and Lazy Clearing against
+our properties, finding that Depth Isolation does indeed enforce stepwise
+stack safety and therefore also observable stack safety. We also found that
+neither lazy micro-policy actually does enforce stack safety, and proposed
+a fix.
+
 \paragraph{Heap Safety as a Security Property}
@@
 
-  We have presented a pure notion of stack safety that omits many complicating
-  factors common to real systems. In particular, our model as presented exploits
-  the simplifying assumption that privilege never increases during a function
-  call, and that therefore a contour computed once on the entry to a function
-  describes its privilege throughout. Our future plans are to extend the model
-  with common language features, some of which violate this assumption. Here we
-  describe the basic principles by which our model can be extended to some of
-  these cases. We also plan to develop a testing framework for quickly
-  checking the validity of different enforcement mechanisms.
-
-  \paragraph*{Capability machines}
-  %
-  %Other policies are implemented in hardware or built on hardware mechanisms.
-  Capability machines such as CHERI \citep{Woodruff+14} extend conventional
-  architectures to support efficient and fine-grained control over memory
-  accesses, which can be used to implement security policies, including
-  the stack protection policies of
-  \citet{Skorstengaard+19b} and (with an extension to support linear capabilities) \citet{Skorstengaard+19}.
-  %\bcp{I thought these policies used
-  %  linear capabilities?}\leo{Only the second one AFAIK. What would you
-  %  change to characterize it more accurately?}\rb{StkPointers uses linear
-  %  capabilities, its predecessor users local capabilities (and unlike
-  %  StkPointers could be implemented in CHERI)}.
-  Capabilities package up pointers
-  (ordinary data) with base and bounds information (policy data),
-  and they require rewriting source code (e.g., to use capabilities in place
-  of ordinary pointers), so separating ordinary machine state from policy
-  state is delicate; we conjecture it can be done with some effort.
-
-\BCP{This section seems incomplete?}
-
-  %% Mapping the underlying protection
-  %% mechanisms to policy states and step function would be a delicate operation, as
-  %% those are tightly coupled to the hardware design. However, separating the
-  %% enforcement aspects of a concrete policy (such as StkTokens, where policy state
-  %% would encompass details like the abstract call stack, the capability sealing
-  %% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
-  %% the machine would be relatively straightforward.  \bcp{Not sure that any of
-  %%   this is comprehensible... :-(}
-
-  %\paragraph*{Software-only policies}
-  %
-  %Some policies are implemented in software, by modifying code in the
-  %compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
-  %to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
-  %be evaluated in terms of our formal notion of stack safety with a trivial null
-  %policy. But any state maintained by the enforcement mechanism must be in normal
-  %memory and subject to the same confidentiality and integrity concerns (to be
-  %varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
-  %out as policy state provided that it can be proven to protect itself.
-
-%  \paragraph{Stack-derived Pointers}
-
-%    Common programming idioms involve a caller passing a pointer to
-%    its local data into a callee. Using standard notions of pointer
-%    provenance~\citep{provenance}, we can extend our model to
-%    distinguish safe use of this idiom from true stack safety
-%    violations. We require additional annotation to identify which
-%    addresses within a function's stack frame correspond to distinct
-%    objects. Then a valid pointer to an object is one derived through
-%    legal arithmetic from its base address, and at each call point, an
-%    object's addresses are marked low confidentiality and integrity if
-%    a valid pointer exists in a register or in the transitive closure
-%    of the callee's accessible memory. All pointers to an object cease
-%    to be valid when it is deallocated along with its stack frame.
-
-%  \paragraph{Non-stack control flow}
-
-%    Control structures beyond well-bracketed calls and returns require
-%    modifications to the model. Tail calls, for instance, reuse the caller's
-%    stack frame for its callee, and every nested tail call returns
-%    simultaneously to the top non-tail-calling function. Under normal stack
-%    integrity, the first function in a chain of tail calls returns and violates
-%    integrity, while subsequent tail calls never actually return at all.
-%    Instead we need a variant that treats the entry of a tail call as the
-%    return of its caller.
-
-%    More complex is the addition of a coroutine model, in which multiple stacks
-%    respect stack safety internally, but might also yield to one another.
-%    Here we must distinguish three levels of integrity: accessible, in the same
-%    stack, and in another stack. Stack integrity requires that same-stack data
-%    be unchanged when a function returns, but other-stack data may change. A
-%    similar {\em yield integrity} property requires that, from a yield out of
-%    a coroutine to the next yield back into it, all accessible and same-stack
-%    data are unchanged. Confidentiality properties are split similarly.
-
-%    \paragraph{Random Testing}
-
-%    Part of the motivation for the strong lockstep stack-safety property
-%    is that it uncovers errors quickly, and therefore should be quicker
-%    to test, just like the smarter testing properties
-%    of~\citet{TestingNI:ICFP}. We have already developed a testing
-%    framework for debugging our policy enforcement for this property, and
-%    it would be interesting to explore whether a testing-amenable variant
-%    of the lazy property also exists.
-
-% \rb{Notes from meeting:
-%   \begin{itemize}
-%   \item Explain how the theory applies to production compilers like GCC and
-%     LLVM, which aspects are directly applicable, what are the effects of
-%     alternative calling conventions. Other less charted waters: tail calls,
-%     coroutines, jumps into supervisor mode.
-%   \item CHERI and its operation, documentation and practical configurations and
-%     operation.
-%   \item Compare and contrast with StkTokens.
-%   \item Under what circumstances can stack safety exist without an enforcement
-%     mechanism?
-%   \item Zeroing out of allocated space inside blessed sequences, eager vs lazy.
-%     Possibility of blessed alloc and dealloc sequences, which also manipulate SP
-%     and would not be legal without blessing. In such an extended setting, the
-%     return sequence would check for matching addresses (it would be nice to have
-%     testing for this!).
-%   \end{itemize}
-% }
+We have introduced a characterization against which stack safety enforcement
+mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
+From here there are two natural ways to extend the work. We should thoroughly evaluate
+other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
+the functionality of our definitions by showing that they can be used to prove
+the correctness of a programming logic with strengthened rules around calls and returns.
+
+\paragraph*{Stack Safety in the Cheri Zoo}
+
+There are several proposals around the use of Cheri capabilities to enforce stack safety,
+ranging from mechanisms that use the standard Cheri hardware \citep{Skorstengaard+19b},
+to those that propose entirely new types of capabilities, such as linear
+(\citep{Skorstengaard+19}), uninitialized (\citep{Georges+21}), and monotonic (\sna{cite}).
+Of these, uninitialized capabilities seem particularly promising -- they naturally enforce
+write-before-read behavior, and therefore will enable confidentiality without the need
+to clear large swaths of the stack. We propose to implement a model of the stack safety
+system described in (\citep{Georges+21}) and test it against our properties using
+QuickChick.
+
+Notably, we conjecture that while these systems may implement integrity and confidentiality,
+they fall just short of return-to-caller. This is because the standard way to protect returns
+in Cheri is to construct {\em return capability}, an object capability that must be invoked
+to simultaneously transfer control back to the caller and unseal a capability on an activation
+record. An ill-behaved intermediate caller could always pass its caller's return capability on
+to its callee instead of its own, ensuring that the callee will return past it. Instead,
+we believe that Cheri supports a slightly weaker control-flow property in which only callers
+that obey a dynamic reasonableness property are guaranteed not to be skipped.
+
+\paragraph*{Program Logic for Stacks}
+
+Stack integrity suggests a reasoning principle in which knowledge about a caller's
+local data can be carried across a call site. We propose to develop this notion
+into a low-level separation logic with a ``call rule'' that preserves propositions
+whose domains are sealed across a call, while allowing those propositions proven
+about the behavior of the callee to apply to the accessible stack.
+
+It is straightforward to imagine that such a rule could be proven sound in a
+system with lockstep stack safety. But there is a real novelty in considering
+what kind of soundness can be proven from a system that obeys only our observational
+properties.
 

### 2021-04-16 tex/main.tex
Added shortening

@@
 that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
-{\tt y} and hence {\tt x} without directly reading the stack pointer. In
-this case, {\tt g} branches on that value, which is a different way of
-violating lockstep confidentiality. Technically, this violation
-occurs because the program counter is also a component of the state that must change
-identically between the primary and variant traces. In a case where \(v_4 \not = 5\),
-the program counters will differ, breaking the lockstep.\apt{Is there any reason
-to complicate things like this by bringing in the PC?}\bcp{+1}
+{\tt y} and hence {\tt x} without even directly reading the stack pointer.
 
@@
 creates a trace of observations corresponding to the observations of each step in
-the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
+the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
 initial observation, a machine-trace, and an observation trace:
 
+\begin{minipage}[b]{.6\textwidth}
   \judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
                 {\(\mpcstatename' \hookrightarrow \MPCT\)}
-              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
-              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
+                {\(\mathit{obsfor} ~ \obs' ~ \MPCT = \obsT\)}
+                {\(\mathit{obsfor} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
+\end{minipage}
+\begin{minipage}[b]{.35\textwidth}
   \judgment{}
-         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}
+           {\(\mathit{obsfor} ~ \obs ~ \mpcstatename = \obs\)}
+\end{minipage}
 
@@
 
-\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
+\judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
          {\(\obsof(\MPCT) = \obsT\)}
@@
 
-We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
-as far as an external observer is concerned, written \(\obsT_1 \lesssim
-\obsT_2\), if the sequence of
-non-silent observations of $\obsT_1$ is a prefix of those of
-$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
-coinductively:
-
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(\obsT \lesssim \obsT\)}
+We say that a pair of observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
+as far as an external observer is concerned, written \(\obsT_1 \eqsim
+\obsT_2\), if the sequence of non-silent observations is the same. That is, we
+compare up to deletion of \(\tau\) observations, coinductively:
+
+\begin{minipage}{.2\textwidth}
+  \judgment{}{\(\obsT \eqsim \obsT\)}
 \end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(\tau \lesssim \obsT\)}
+\begin{minipage}{.2\textwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\tau \obsT_1 \eqsim \obsT_2\)}
 \end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{}{\(w \lesssim w\obsT\)}
+\begin{minipage}{.2\textwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\obsT_1 \eqsim \tau \obsT_2\)}
 \end{minipage}
-
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(\tau \obsT_1 \lesssim \obsT_2\)}
+\begin{minipage}{.2\textwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\obs\obsT_1 \eqsim \obs\obsT_2\)}
 \end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(\obsT_1 \lesssim \tau \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \judgment{\(\obsT_1 \lesssim \obsT_2\)}
-           {\(w\obsT_1 \lesssim w\obsT_2\)}
-\end{minipage}
-
-\smallskip
-We then define similarity of observation traces as traces prefixing each other:
-\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
-%
+
 %\leo{I still don't know how to do this transition. Maybe contrast
@@
 than the stepwise one, and also much more extensional.
-We can think of stepwise, deferred stepwise, and observational as
+We can think of stepwise, deferred, and observational as
 three points along a spectrum from eager to lazy. The first two

### 2021-04-16 tex/main.tex
Minor tweaks

@@
 which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), which
-maps a machine state to a set of addresses. \apt{Is it a relation or a function (seems to
-  be the latter)?}
+this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
+relation between machine states and the addresses that should be sealed.
 If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
@@
 and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-But whatever the return convention, at each step we update the context according
-to the context-step function, \(\stepsto_C\).\apt{abrupt transition}
 
@@
 \end{subfigure}
-\caption{A violation with pass-by-reference\apt{SP(n) depths looked wrong; SNA double-check that I fixed them right.}}
+\caption{A violation with pass-by-reference}
 \label{fig:passing}

### 2021-04-16 tex/main.tex
CP

@@
 corner cases where data can leak across repeated calls to the same callee.  We
-propose a variant of Lazy Tagging and Clearing that \emph{does} enforce
+propose a variant of {\em Lazy Tagging and Clearing} that \emph{does} enforce
 confidentiality (as far as our testing can tell), albeit at some performance
@@
     security properties of Lazy Tagging and Clearing, but they consider it to enforce
-    a completely different (and generally weaker) policy anyway (``data flow integrity'' vs. ``memory safety'').}}
+    a completely different (and generally weaker) policy anyway (``data flow
+    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
+  leave this issue to deal with later (in consultation with them, which we
+  don't have time to do a good job of today).  At any rate, by my reading we
+are not saying anything incorrect at the moment.}
 
@@
 \apt{Suddenly we are talking about alternative enforcement mechanisms for the first time.
-  This surely needs \emph{some} foreshadowing in the abstract and intro?}
+  This surely needs \emph{some} foreshadowing in the abstract and
+  intro?}\bcp{Yes!}
 Dividing policy state from machine state helps us reason about
@@
   developer's in the same address space, but the developer wants her private
-  data to be protected from both reading/exfiltrating and overwriting...}
+  data to be protected from both reading/exfiltrating and
+  overwriting... More generally, we can talk about the fact that programmers
+  use simple reasoning principles to think about their programs (e.g., local
+  variables don't change across function calls), and life is better when
+  these principles are actually true...}
 
@@
       at places in both the C and asssembly, making them harder to read
-  and eating more space.}}  \ifaftersubmission\bcp{I wonder if we
+  and eating more space.}\bcp{I'm just talking about removing the space
+  above "main" in the left column and moving f down a bit.}}
+\ifaftersubmission\bcp{I wonder if we
     could make the figures easier to read by putting very light, differently
@@
 so the final return value produced by {\tt main} is wrong.
-
-The black rectangle represents the return address, which is saved by the call
-instruction itself, so is not sealed.
+%
+(The black rectangle represents the return address, which is saved by the call
+instruction itself, so it is not sealed.)
 
@@
 
-\section{Do Micro-Policies Enforce Stack Safety?}
+\section{Case Study: Do Micro-Policies Enforce Stack Safety?}
 \label{sec:enforcement}
 
-\bcp{``Enforcement'' seems like too broad a title.  What about calling it
-  ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
-  Safety?''? And what about folding the testing section into this one?  That
-would keep related things together, would avoid appearing to make overly
-grand claims in either section, and would give us an excuse to write less
-about the mechanism here---we only need to write enough about the mechanism
-itself that readers can understand our validation case study.}
-\leo{Agreed. Took a crack at it.}
-
-\leo{I am not sure we've hit the correct level of detail in this
-  section.}\bcp{Agreed!}
-
-In this section we sketch an existing enforcement mechanism, Depth
-Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
-formal stack-safety property described in the previous section; in
-fact, it enforces a stronger, inductive version of that property which
+% \bcp{``Enforcement'' seems like too broad a title.  What about calling it
+%   ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
+%   Safety?''? And what about folding the testing section into this one?  That
+% would keep related things together, would avoid appearing to make overly
+% grand claims in either section, and would give us an excuse to write less
+% about the mechanism here---we only need to write enough about the mechanism
+% itself that readers can understand our validation case study.}
+% \leo{Agreed. Took a crack at it.}
+
+% \leo{I am not sure we've hit the correct level of detail in this
+%   section.}\bcp{Agreed!}
+
+In this section we sketch an existing enforcement mechanism, {\em Depth
+Isolation}, from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
+formal stack-safety property described in the previous section. (In
+fact, it enforces a slightly stronger, inductive version of that property which
 strengthens the local state encapsulation component by checking for
 integrity and confidentiality violations at every step of the callee's
-execution trace. We validate that claim by developing a property-based
+execution trace.) We validate this claim with a property-based
 testing framework that can quickly uncover inconsistencies between the
@@
 on a programmable, tag-based reference monitor that runs alongside the
-program.  To control this monitor, they use a programming model that
-allows fine-grained manipulation of metadata tags to encode so-called
+program,
+allowing fine-grained manipulation of metadata tags to encode so-called
 \emph{micro-policies}~\citep{pump_oakland2015}.
-In such a system, all values in
-memory addresses and registers (including the $\PCname$)
-are enriched with an abstract metadata tag, which can represent
-arbitrary information about the value. A micro-policy is defined as a set of tags and a
-collection of software-defined rules, indexed by machine opcode.
+Every value in
+memory and registers (including the $\PCname$)
+is enriched with an abstract metadata tag, which can represent
+arbitrary information about the value. A micro-policy is then defined as a
+set of tags and a
+collection of software-defined rules.
 At each step of the machine, the relevant rule is applied to the tags on
-the instruction's inputs ($\PCname$, registers, memory) and on the instruction itself, and produces
-one of two outcomes: either the instruction is allowed to execute (and generates
-tags for the result of the operation and the new $\PCname$), or the machine
-fail-stops with a policy violation.
-\citeauthor{pump_oakland2015}\bcp{and PIPE, if there are any references to
-  cite} have shown that a wide range of micro-policies can
-be defined using this scheme.
-
-Efficient execution of these micro-policies relies on hardware implementations,
-such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
-as word-size bit vectors, stored separately from the regular memory and registers.
+the instruction's inputs ($\PCname$, registers, memory) and on the
+instruction opcode itself, producing
+one of two outcomes: either the instruction is permitted to execute (and
+generates tags for the result of the operation and the new $\PCname$), or
+the machine fail-stops with a policy violation.
+\citeauthor{pump_oakland2015}\bcp{additional citations: the noninterference
+  paper, Nick and Andre's papers, ...} show that a wide range of
+micro-policies can be defined using this scheme.
+
+Efficient execution of these micro-policies relies on hardware acceleration,
+such as the PUMP architecture~\citep{pump:asplos2015}.
 The hardware incorporates a rule cache
@@
 it is important to design micro-policies so that they hit in the cache as
-much as possible. Thus, practical policies maintain a small working set of
-distinct tags.
+much as possible. Thus, practical policies need to maintain a small working
+set of distinct tags.
 %% %
@@
 safety developed in \cref{sec:lse} is a small variation
-on the Depth Isolation policy presented by
-\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
+on the {\em Depth Isolation} policy presented by
+\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
 with ownership information associated to each stack frame (and to each
@@
 the connection to the eager policy, we can identify when lazy policies
-miss violations that can become visible. The Lazy Tagging and Clearing
+miss violations that can become visible. The {\em Lazy Tagging and Clearing}
 micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
@@
 
-Lazy Tagging and Clearing can be repaired if, instead of tagging each function
+{\em Lazy Tagging and Clearing} can be repaired if, instead of tagging each function
 activation with its depth in the stack, we generate a fresh activation
@@
 micro-policies that tag stack objects with a pair of frame and object identifiers
-used to validate accesses to the stack: Return Address Protection (which
-prevents an adversary from overwriting designated return addresses), Static
-Authorities (which only allows the code of a function to access the stack frames
-of its own dynamic instances), and Depth Isolation (described in
+used to validate accesses to the stack: {\em Return Address Protection} (which
+prevents an adversary from overwriting designated return addresses), {\em Static
+Authorities} (which only allows the code of a function to access the stack frames
+of its own dynamic instances), and {\em Depth Isolation} (described in
 \cref{sec:enforcement}). All these policies exploit various kinds of spatial and
@@
 limited performance overhead. In addition to the baseline ``eager'' policies,
-they propose a number of lazy optimizations, notably Lazy Tagging and Lazy
-Clearing, discussed in \cref{sec:lazy}, with associated improvements in
+they propose a number of lazy optimizations, notably {\em Lazy Tagging} and
+{\em Lazy
+Clearing}, discussed in \cref{sec:lazy}, with associated improvements in
 performance.
 
-We have evaluated Depth Isolation, Lazy Tagging, and Lazy Clearing against
-our properties, finding that Depth Isolation does indeed enforce stepwise
-stack safety and therefore also observable stack safety. We also found that
-neither lazy micro-policy actually does enforce stack safety, and proposed
-a fix.
+We have evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
+  Clearing} against our properties, finding that {\em Depth Isolation} does
+indeed enforce stepwise stack safety and therefore also observable stack
+safety. We also found that neither lazy micro-policy actually does enforce
+stack safety, and proposed a fix.
 
@@
 noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
-be ``Azedevo et al.''}\fi. Just
+be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
+(or ask arthur how!).}\fi. Just
 as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
-characterization of the meaning of (heap) memory safety, in this paper
-we aim to do the same for the stack.\bcp{Let's describe their work in a bit
-  more detail.}
+characterization of the meaning of (heap) memory safety, in this paper,
+inspired by their work, we aim to do the same for the stack.
 %
 Their model describing safety of shared stack-allocated
-objects\bcp{??} extends naturally to heap safety, complicated by the fact that
+objects\BCP{huh?? they were not talking about stacks} extends naturally to heap safety, complicated by the fact that
 these objects are not deallocated by returns\bcp{??}, and therefore a
@@
 properties separately on continuous segments of each call, with
-contours computed at each crossing between caller and callee.  \bcp{Didn't
-  understand last sentence.}
+contours computed at each crossing between caller and callee.  \BCP{Didn't
+  understand very much of that paragraph.}
 

### 2021-04-16 tex/main.tex
Polishing

@@
 Colored boxes indicate an assignment of the address to a
-{\em domain} -- a label reflecting a particular set of access rights. Domains
+{\em domain}---a label reflecting a particular set of access rights. Domains
 are \(\unsealed\)\bcp{``unsealed'' would read better} (in the stack and accessible to writes), \(\sealed{\depth}\)
@@
 at a particular depth. Each caller registers a {\em target}
-condition -- a predicate on machine states that must hold for it to consider itself active.
+condition---a predicate on machine states that must hold for it to consider itself active.
 A context is a pair of a ``domain map'' from components to domains, and a
@@
 
-\definition {\bf Stepwise Stack Integrity}
-
+\definition
 A system enjoys {\em stepwise stack integrity} if, for any reachable state
@@
 
-\definition {\bf Stepwise Stack Confidentiality}
-
+\definition
 For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
@@
 \(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\) -- if every variant (with the same policy state)
+\(\MPCT \doteq \NPCT\)---if every variant (with the same policy state)
 steps in lockstep until both traces have reached a point whose depth is
@@
 
-\vspace{\belowdisplayskip}
+\medskip
 
@@
 
-\definition {\bf Return-To-Caller}
-
+\definition
 A system enjoys {\em return-to-caller} when, for every pair of reachable states
@@
 A call annotation now includes a predicate relating machine states to the components
-that should be passed -- when passing by value, this will typically be freshly allocated
+that should be passed---when passing by value, this will typically be freshly allocated
 cells at the top of the frame, while passing by reference just gives the
@@
 
-\definition {\bf Stepwise Stack Integrity (Passing)}
-
+\definition
 Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
@@
 
-\definition {\bf Stepwise Stack Confidentiality (Passing)}
-
+\definition
 For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
@@
 using a combination of tags on the instructions
-and an additional tag on the $\PCname$; we omit the details here.
+and an additional tag on the $\PCname$.
 
@@
 
-\definition {\bf Deferred Stack Integrity}
-
+\definition
 For any \(\MPCT \in \MPCT^*\),
@@
 
-\vspace{\belowdisplayskip}
+% \vspace{\belowdisplayskip}
 
@@
 \(\tau\). Our machine step function now takes a state and returns a state and an observation,
-which is carried over into the policy and context step functions.
-%
-\[\mach \stepsto[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
+which is carried over into the policy and context step functions:
+%
+$\mach \stepsto[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS $.
 %
@@
 Integrity is already straightforward, simply weakening the condition that must hold
-after the return -- instead of preventing dangerous writes from being read, we
+after the return---instead of preventing dangerous writes from being read, we
 prevent them from being leaked.
 
-\definition {\bf Observational Stack Integrity}
-
+\definition
 For any call trace \(\MPCT \in \MPCT^*\),
@@
 such that \(\domMap ~ \component = \sealed{\_}\).
-
+%
 Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
@@
 
-\definition {\bf Observational Stack Confidentiality}
-
+\definition
 For any call trace \(\MPCT \in \MPCT^*\),
@@
 such that \(\domMap ~ \component \not = \outside\).
-
+%
 A system enjoys {\em observational confidentiality} if for all
@@
   meaning that no secrets leak during the call
-\item \(f_\depth ~ \last(\MPCT) \leftrightarrow f_\depth ~ \last(\NPCT)\) --
-  if one call returns, so does the other
+\item \(f_\depth ~ \last(\MPCT) \leftrightarrow f_\depth ~
+\last(\NPCT)\)---if one call returns, so does the other
 \item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
@@
 
-Lazy Per-activation Tagging and Clearing ends up enforcing the combination
-of deferred stepwise stack integrity and stepwise confidentiality -- we
+{\em Lazy Per-activation Tagging and Clearing} ends up enforcing the combination
+of deferred stepwise stack integrity and stepwise confidentiality---we
 don't actually need a deferred confidentiality property here. Therefore,
@@
 (\citep{Skorstengaard+19}), uninitialized (\citep{Georges+21}), and monotonic (\sna{cite}).
-Of these, uninitialized capabilities seem particularly promising -- they naturally enforce
+Of these, uninitialized capabilities seem particularly promising---they naturally enforce
 write-before-read behavior, and therefore will enable confidentiality without the need

### 2021-04-16 tex/main.tex
Policy comments

@@
     confidentiality or integrity violation.}
+  \rb{Correct, and it doesn't. Some versions of the policy do clear
+    the stack, retagging would be easily added.}
 
@@
   an initial policy state, tagging instructions corresponding to
-  ``blessed'' call or return sequences appropriately, while marking
+  blessed call or return sequences appropriately, while marking
   all potential stack locations as $\tagNoDepth$, and
@@
 partially instantiated machine state and attempts to generate an
-instruction (or a sequence of instructions such as a ``blessed''
+instruction (or a sequence of instructions such as a blessed
 sequence) that makes sense locally (e.g., jumps go to a potentially valid
@@
 intriguingly complex setup to appear.  Notably, our first try at the
-``blessed'' call sequence/policy combination did not initialize all
+blessed call sequence/policy combination did not initialize all
 locations for the newly allocated stack frame correctly, but left some
@@
 First, they eliminate the initialization pass. In this {\em Lazy Tagging}
-system, the stack is initially tagged as ``clear.'' A function is permitted
+system, the stack is initially tagged as ``clear.''
+\rb{The previous section already presents ``unused'' as the default tag,
+though it doesn't say that it can be overwritten whenever}
+A function is permitted
 to overwrite clear memory, and when it does so its depth tag overwrites
@@
 if it ever tries to read from that location.
+\rb{This is in line with the level of detail in the last section.}
 %

### 2021-04-16 tex/main.tex
Fix reference to example 1
Lazy integrity example

@@
 %cares about changes it can see after the return, not changes during execution.
-In this case, after instruction 8, {\tt f} has violated integrity by writing to {\tt x},
+In this case, after instruction 11, {\tt f} has violated integrity by writing to {\tt x},
 so the final return value produced by {\tt main} is wrong.
@@
 Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that {\tt main} returns
--1. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
+$-1$. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
 This has the appearance of an integrity violation.  How do we capture this in our
@@
 access their values. ``No reads after dangerous writes.''
+%
+For example, consider again the integrity violation in
+Figure~\ref{fig:int1}. Under a deferred discipline, the write to {\tt
+  x} in instruction 11 taints the variable but does not in itself
+constitute a violation, which happens only when the tainted value is
+read in instruction 4. However, if this instruction added the constant
+that was originally stored in {\tt x} without reading the variable,
+i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
+\rb{Tried to give a quick example without being too disruptive.
+  However, having it follow the slogan seems to rob it of part of its
+  punch.}
 

### 2021-04-16 tex/main.tex
CP

--- 2021-04-16 tex/main.tex
@@
 Together, integrity and confidentiality protect a caller's data until its callee returns
-to it. But it is possible that an intermediate caller is never returned to, but returned
-{\em past}, to a deeper caller. Fundamentally, this is a control-flow problem, not a
-data-protection problem. From that caller's perspective, it is no different from the
-callee diverging. But a lower level caller might have been relying on the intermediate
+to it. However, it is possible that an intermediate caller is not returned
+{\em to}, but rather returned {\em past}, to some ancestor caller! For
+example, function $A$ calls function B, which in turn calls $C$, which
+returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
+and branching to the address after the call from $A$ to $B$.
+Strictly speaking, this is not an information-flow problem: the caller's XXXXXX.
+\BCP{Still working on this bit.}
+But a lower level caller might have been relying on the intermediate
 caller to finish its computation, so it is worth protecting control flow even from

### 2021-04-16 tex/main.tex
Reorg contributions

@@
 \acmVolume{1}
-\acmNumber{SPLASH} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
@@
 confidentiality (as far as our testing can tell), albeit at some performance
-cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
+cost.
+%
+\ifaftersubmission
+\bcp{Did Nick and Andre know about this deficiency?  Did they mention
   it in their paper?  (If neither, we should make a bigger deal out of the
@@
 are not saying anything incorrect at the moment.}
+\fi
 
@@
 \item We adapt concepts of integrity and confidentiality from language-based
-security to formalize a {\em stepwise stack-safety}
-property and argue that it captures the intuitive concept of strict stack
-safety~(\cref{sec:prelim,sec:lse}). We show that the definition can be
-extended to argument passing on the stack and to a simple coroutine system.
+security, combine these with an existing notion of well-bracketed
+control flow\bcp{need to mention this above} to formalize a {\em stepwise stack-safety}
+property, and argue that this property captures the intuitive concept of
+strict stack
+safety~(\cref{sec:prelim,sec:lse}).
 \item We use property-based randomized testing to check that this stepwise
@@
 of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
+\bcp{mention that we extend the ICFP gen-by-exec with nested stuff...}
 \item We formalize a
@@
 performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
+\item We show that the definition can be
+extended to argument passing on the stack and to a simple coroutine system.
+\bcp{fill in pointer to new section}
 \end{itemize}

### 2021-04-16 tex/main.tex
figure alignment

@@
 
-\begin{subfigure}{.2\textwidth}
+\begin{subfigure}[t]{.2\textwidth}
 {\small
@@
 \end{subfigure}
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}[t]{.5\textwidth}
 {\small
@@
 \end{subfigure}
-\begin{subfigure}{.25\textwidth}
+\begin{subfigure}[t]{.25\textwidth}
 \begin{center}
@@
 \end{subfigure}
-\caption{An integrity violation\bcp{Can we align the C-like snippets with
-    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
-      at places in both the C and asssembly, making them harder to read
-  and eating more space.}\bcp{I'm just talking about removing the space
-  above "main" in the left column and moving f down a bit.}}
+\caption{An integrity violation
 \ifaftersubmission\bcp{I wonder if we
@@
 of bad behavior that it prevents; we will argue that: \bcp{Not sure this
-  enumeration is useful.}
+  enumeration is useful.}\apt{+1}
 
@@
 \begin{figure}
-\begin{subfigure}{.2\textwidth}
+\begin{subfigure}[t]{.2\textwidth}
 {\small

### 2021-04-16 tex/main.tex
more alignment wip
More polishing in case-study

--- 2021-04-16 tex/main.tex
@@
 
-In this section we sketch an existing enforcement mechanism, {\em Depth
-Isolation}, from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
-formal stack-safety property described in the previous section. (In
-fact, it enforces a slightly stronger, inductive version of that property which
-strengthens the local state encapsulation component by checking for
-integrity and confidentiality violations at every step of the callee's
-execution trace.) We validate this claim with a property-based
-testing framework that can quickly uncover inconsistencies between the
-enforcement mechanism and the stack-safety property.
+In this section we explore how an existing enforcement mechanism, {\em
+  Depth Isolation}, from \citet{DBLP:conf/sp/RoesslerD18} implements
+the formal stack-safety property described in the previous
+section. (In fact, it enforces a slightly stronger, inductive version
+of that property which strengthens the local state encapsulation
+component by checking for integrity and confidentiality violations at
+every step of the callee's execution trace.) We validate this claim
+with a property-based testing framework that can quickly uncover
+inconsistencies between the enforcement mechanism and the stack-safety
+property.
 
@@
 sequences that have permission to manage the stack, say during calls and returns.
+\leo{If we have the simple exposition so far, no need to mention the following sentence:}
 In the following, for ease of exposition, we will assume that no
@@
 \item From callee back to caller, when execution is about to reach a
-  return point.  The exit sequence retags all locations in the
-  function frame with $\tagNoDepth$. At the actual return instruction,
-  the $\PCname$ tag is decremented. \leo{Not sure the actual policy
-    does that now (retag unused). But I don't think it's a
-    confidentiality or integrity violation.}
-  \rb{Correct, and it doesn't. Some versions of the policy do clear
-    the stack, retagging would be easily added.}
+  return point.
+  %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
+  At the actual return instruction,
+  the $\PCname$ tag is decremented.
+%  \leo{Not sure the actual policy
+%    does that now (retag unused). But I don't think it's a
+%    confidentiality or integrity violation.}
+%  \rb{Correct, and it doesn't. Some versions of the policy do clear
+%    the stack, retagging would be easily added.}
 
@@
 
-We validated the micro-policy of the previous section by systematically
-testing that it satisfies eager stack integrity and
+We validated the micro-policy of the previous section by
+systematically testing that it satisfies eager stack integrity and
 confidentiality. We used a Coq specification of the RISC-V
-architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
-extended it with a runtime monitor implementing the stack-safety
-micro-policy.\leo{Footnote: code will be made public}\bcp{For the reviewers?
- Or later?} We chose the
-Coq proof assistant as the setting for our implementation for two
-reasons: first, to ensure that our coinductive trace definitions are
-well-formed, and reason about them\leo{Do we prove anything
-  interesting anymore?}, and to leverage the power of the QuickChick
-property-based testing framework~\citep{Pierce:SF4}.
+architecture
+\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
+\leo{translated from Haskell via hs-2-coq?}  and extended it with a
+runtime monitor implementing the stack-safety
+micro-policy.\footnote{This code will be made public. \leo{More?}} We
+  chose the Coq proof assistant as the setting for our implementation
+  for two reasons: first, to ensure that our coinductive trace
+  definitions are well-formed, and reason about them\leo{Do we prove
+    anything interesting anymore?}, and to leverage the power of the
+  QuickChick property-based testing framework~\citep{Pierce:SF4}.
 
@@
 
-To write such generators we build on the work of \leo{ICFP 2013},
-where they introduced {\em generation by execution} to produce
-interesting programs. Generation by execution receives as an input a
-partially instantiated machine state and attempts to generate an
-instruction (or a sequence of instructions such as a blessed
-sequence) that makes sense locally (e.g., jumps go to a potentially valid
-code location and loads read from from a potentially valid stack
-location). Then
-we step the machine and repeat the process until we generate or
-execute some target number of instructions, or if we reach a point
-where the machine can't step anymore.
+To write such generators we build on the work of
+\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, where they
+introduced {\em generation by execution} to produce progams leading
+abstract machines towards interesting behaviors. Generation by
+execution receives as an input a partially instantiated machine state
+and attempts to generate an instruction (or a sequence of instructions
+such as a blessed sequence) that makes sense locally (e.g., jumps go
+to a potentially valid code location and loads read from from a
+potentially valid stack location). Then we step the machine and repeat
+the process until we generate or execute some target number of
+instructions, or if we reach a point where the machine can't step
+anymore.
 
@@
 
+In addition, we needed to extend this technique to handle the nested
+nature of our confidentiality property: rather than generating two
+initial machines that are variant of one another and letting them
+execute to test for noninterference, we need to generate a new variant
+{\em every time a call is made} and check confidentiality for the
+subtrace produced from that variant state until its corresponding
+return. As a result, a ``single'' confidentiality test compactly
+checks the confidentiality of multiple nested calls.
+
 \paragraph*{Findings}
 
-To evaluate the effectiveness of testing against our formal properties, we
-used {\em mutation testing}\leo{cite}. In mutation testing, we inject errors
+To ensure the effectiveness of testing against our formal properties, we
+used {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
 (mutations) in a program that should cause the property of interest (here,
@@
 
-The mean-time-to-failure (MTTF) for various bugs can be found in \leo{Table
-  with numbers?}\bcp{Sure.}. All experiments were run in a \leo{Machine
-  spec}. \leo{Something like: There is really no baseline to compare this
+\begin{table}[]
+\centering
+\begin{tabular}{c|c|c}
+  Bug & Property Violated & MTTF (ms) \\
+  \hline
+      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
+      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
+      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
+\end{tabular}
+\vspace*{1em}
+\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+\vspace*{-2em}
+\label{tab:bug-table}
+\end{table}
+
+The mean-time-to-failure (MTTF) for various bugs can be found in
+Table~\ref{tab:bug-table}. All experiments were run in a desktop
+machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
+RAM. \leo{Something like: There is really no baseline to compare this
   against, but include it here for the future/reasons?}
-
+%
 The artificially injected bugs focus on weakening the tag propagation
-and checking mechanism. \leo{Mention them all?}\bcp{Nah, especially if they
-  are all listed in the table.}
+and checking mechanism. 
 %
@@
 current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
-that restriction by not checking the depth equality, or not checking
-the tags at all, causing bugs of different difficulty.
+that restriction by not checking the depth equality (row {\em
+  LOAD\_NO\_CHECK}).
 %
@@
 $\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing these constraints also causes violations to the
-property.
+tag. Relaxing that constraints causes violations to the integrity
+property (row {\em STORE\_NO\_CHECK}).
 
@@
 %\bcp{right?}\leo{yeah}
-; we re-implemented it in Coq targeting
-RISC-V).  These errors ranged from trivial typos to
-ones that required an
-intriguingly complex setup to appear.  Notably, our first try at the
-blessed call sequence/policy combination did not initialize all
-locations for the newly allocated stack frame correctly, but left some
-of them as $\tagNoDepth$. This allowed for a potential integrity
-violation, but only if a rather complicated sequence of events
-occured: \leo{call - write - call (seal) - return - call - write?}
+; we re-implemented it in Coq targeting RISC-V).  These errors ranged
+from trivial typos to ones that required an intriguingly complex setup
+to appear.  The most interesting bug (included in the table as row
+{\em HEADER\_NO\_INIT}), in our first try the blessed call
+sequence/policy combination did not initialize all locations for the
+newly allocated stack frame correctly, but left some of them as
+$\tagNoDepth$. This allowed for a potential integrity violation, but
+only if a rather complicated sequence of events occured: in
+particular, the smallest counterexample requires calling a function,
+writing in the uninitialized location (which is always fine), calling
+another function (to seal what was written for integrity purposes),
+and then attempt to write in the same location again from the new
+callee, which is only allowed because the enforcement still thinks
+the location is unused, but violates the integrity of the caller's data.
 

### 2021-04-16 tex/main.tex

@@
   \centering
-  \begin{subfigure}{.2\textwidth}
+  \begin{subfigure}[t]{.2\textwidth}
 {\footnotesize
@@
 int *stash = 0;
-
 int main() {
@@
 \end{subfigure}
-  \begin{subfigure}{.4\textwidth}
+  \begin{subfigure}[t]{.4\textwidth}
 {\footnotesize
@@
   \end{subfigure}
-  \begin{subfigure}{.25\textwidth}
+  \begin{subfigure}[t]{.25\textwidth}
     \begin{tabular}{l l l}
@@
   \centering
-  \begin{subfigure}{.2\textwidth}
+  \begin{subfigure}[t]{.2\textwidth}
 {\small
@@
   \end{subfigure}
-  \begin{subfigure}{.6\textwidth}
+  \begin{subfigure}[t]{.6\textwidth}
 {\small
@@
 \begin{subfigure}{.65\textwidth}
+\vspace*{0.2in}
 \begin{center}

### 2021-04-16 tex/main.tex
Intro nits

@@
 %\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-The most abstract characterization stack safety that we are aware of is
+The most abstract characterization of stack safety that we are aware of is
 recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
@@
 data safety in the face of control-flow attacks (see \cref{sec:control}).
+\rb{\cref{sec:RTC}?}
 This reflects a cleaner division between data protection and control-flow protection
@@
 
-\label{sec:controlflow}
+\label{sec:controlflow} % RB: ???
 

### 2021-04-16 tex/main.tex
finish adjusting example alignments
No pass by reference?
A paragraph about WBCF (may need to be relocated to section 4)

@@
 
-Together, integrity and confidentiality protect a caller's data until its callee returns
-to it. However, it is possible that an intermediate caller is not returned
-{\em to}, but rather returned {\em past}, to some ancestor caller! For
-example, function $A$ calls function B, which in turn calls $C$, which
+\BCP{We agreed that most or all of this belongs in Section 4...}Together,
+integrity and confidentiality protect a caller's data until its callee
+returns to it. However, it is possible that an intermediate caller is not
+returned {\em to}, but rather returned {\em past}, to some ancestor caller!
+For example, function $A$ calls function $B$, which in turn calls $C$, which
 returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
-and branching to the address after the call from $A$ to $B$.
-Strictly speaking, this is not an information-flow problem: the caller's XXXXXX.
-\BCP{Still working on this bit.}
-But a lower level caller might have been relying on the intermediate
-caller to finish its computation, so it is worth protecting control flow even from
-attacks that do not violate data protections. We formalize this as {\em return-to-caller}.
+and branching to the address after the call from $A$ to $B$.  Strictly
+speaking, this is not an information-flow violation, since neither $A$'s nor
+$B$'s stack frames are improperly read or written by $C$.  But it is still
+arguably unsafe in general.  For example, suppose that $B$'s continuation
+after the call to $C$ takes $C$'s result, doubles it, and returns this to
+$A$.  If $A$'s correct behavior after the call to $B$ returns depends on the
+result it receives being a even number---i.e., $A$ is trusting that $B$'s
+continuation will execute after $C$ returns, perhaps because $A$ and $B$ are
+written by the same person while $C$ is an untrusted library---then $C$ can
+cause it to behave incorrectly, possibly even insecurely, by returning an
+odd result directly to $A$.
+%
+Thus, a formal definition of safe use of the stack should include the
+condition that calls and returns are {\em well bracketed}, e.g., $C$ should
+only return to $A$.
 
@@
 \(|\rts'| = |\rts|\)
+
+\bcp{we probably want to remark that this notion is pretty
+much the same as in StkTokens, just formalized a little differently.}
 

### 2021-04-16 tex/main.tex
Numbered subsections in case study
Moved sections around

--- 2021-04-16 tex/main.tex
@@
 \acmVolume{1}
-\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{SPLASH} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
@@
 confidentiality (as far as our testing can tell), albeit at some performance
-cost.
-%
-\ifaftersubmission
-\bcp{Did Nick and Andre know about this deficiency?  Did they mention
+cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
   it in their paper?  (If neither, we should make a bigger deal out of the
@@
 are not saying anything incorrect at the moment.}
-\fi
 
@@
 \item We adapt concepts of integrity and confidentiality from language-based
-security, combine these with an existing notion of well-bracketed
-control flow\bcp{need to mention this above} to formalize a {\em stepwise stack-safety}
-property, and argue that this property captures the intuitive concept of
-strict stack
+security to formalize a {\em stepwise stack-safety}
+property and argue that it captures the intuitive concept of strict stack
 safety~(\cref{sec:prelim,sec:lse}).
@@
 of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
-\bcp{mention that we extend the ICFP gen-by-exec with nested stuff...}
 \item We formalize a
@@
 performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
-\item We show that the definition can be
-extended to argument passing on the stack and to a simple coroutine system.
-\bcp{fill in pointer to new section}
+\item We show that these definitions can be
+  extended to argument passing on the stack and to a simple coroutine system
+  (\cref{sec:ext}).
 \end{itemize}
@@
 
-\begin{subfigure}[t]{.2\textwidth}
+\begin{subfigure}{.2\textwidth}
 {\small
@@
 \end{subfigure}
-\begin{subfigure}[t]{.5\textwidth}
+\begin{subfigure}{.5\textwidth}
 {\small
@@
 \end{subfigure}
-\begin{subfigure}[t]{.25\textwidth}
+\begin{subfigure}{.25\textwidth}
 \begin{center}
@@
 \end{subfigure}
-\caption{An integrity violation
+\caption{An integrity violation\bcp{Can we align the C-like snippets with
+    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
+      at places in both the C and asssembly, making them harder to read
+  and eating more space.}\bcp{I'm just talking about removing the space
+  above "main" in the left column and moving f down a bit.}}
 \ifaftersubmission\bcp{I wonder if we
@@
 of bad behavior that it prevents; we will argue that: \bcp{Not sure this
-  enumeration is useful.}\apt{+1}
+  enumeration is useful.}
 
@@
 \begin{figure}
-\begin{subfigure}[t]{.2\textwidth}
+\begin{subfigure}{.2\textwidth}
 {\small
@@
   \centering
-  \begin{subfigure}[t]{.2\textwidth}
+  \begin{subfigure}{.2\textwidth}
 {\footnotesize
@@
 int *stash = 0;
+
 int main() {
@@
 \end{subfigure}
-  \begin{subfigure}[t]{.4\textwidth}
+  \begin{subfigure}{.4\textwidth}
 {\footnotesize
@@
   \end{subfigure}
-  \begin{subfigure}[t]{.25\textwidth}
+  \begin{subfigure}{.25\textwidth}
     \begin{tabular}{l l l}
@@
 Together, integrity and confidentiality protect a caller's data until its callee returns
-to it. However, it is possible that an intermediate caller is not returned
-{\em to}, but rather returned {\em past}, to some ancestor caller! For
-example, function $A$ calls function B, which in turn calls $C$, which
-returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
-and branching to the address after the call from $A$ to $B$.
-Strictly speaking, this is not an information-flow problem: the caller's XXXXXX.
-\BCP{Still working on this bit.}
-But a lower level caller might have been relying on the intermediate
+to it. But it is possible that an intermediate caller is never returned to, but returned
+{\em past}, to a deeper caller. Fundamentally, this is a control-flow problem, not a
+data-protection problem. From that caller's perspective, it is no different from the
+callee diverging. But a lower level caller might have been relying on the intermediate
 caller to finish its computation, so it is worth protecting control flow even from
@@
 \(|\rts'| = |\rts|\)
-
-\subsection{Parameters Passed on the Stack}
-
-\newcommand{\mainpassc}{magenta}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-void main() {
-  int x = 5;
-  x = f(x);
-  return;
-}
-
-void f(int a) {
-  int y = 2*a;
-  a = 0;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.6\textwidth}
-{\small
-\begin{verbatim}
-1  main: add $1,%sp          ; allocate frame
-2        mov $5,-1(%sp)      ; initalize x
-3        add $1,%sp          ; allocate arg
-4        mov -2(%sp),-1(%sp) ; set arg
-5        call f              ; SHARE: %sp-1
-6        sub $1,%sp          ; deallocate arg
-7        mov %ra,-1(%sp)     ; store x
-8        sub $1,%sp          ; deallocate frame
-9        ret
-
-10  f:   add $1,%sp          ; allocate frame
-11       mov -3(%sp),-1(%sp) ; y = a
-12       mul $2,-1(%sp)      ; y = a*2
-13       mov $0,-3(%sp)      ; a = 0
-14       sub $1,%sp          ; deallocate frame
-15       ret
-\end{verbatim}
-}
-  \end{subfigure}
-\begin{subfigure}{.65\textwidth}
-\vspace*{0.2in}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 3} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{5}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[Seal(0)]%
-  \memory{1}{\mainpassc}[P(0,1)]%
-  \memory{1}{\retptrc}[RetPtr]%
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  {\tt 13} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  {\tt 14} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  \\
-  {\tt 8} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{10}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{\#6}
-    \MemoryLabel{-7em}{0.75em}{10}
-    \vspace{.5em}
-  } \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-
-\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
-    the figure (at least in the latex build that I'm looking at...)}\fi}
-\label{fig:passingsimple}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{.2\textwidth}
-{\small
-\begin{verbatim}
-int *stash = 0;
-
-int main() {
-  int x, y;
-  f(x);
-  x = 0;
-  f(y);
-  return x;
-}
-
-void f(int &a) {
-  if(!stash)
-    stash = &a;
-  *stash = 5;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}{.49\textwidth}
-{\small
-\begin{verbatim}
-   stash: .word 0
-1  main:  add $2,%sp      ; allocate frame
-2         mov %sp,%ra
-3         add -2,%ra      ; ra = &x
-4         call f          ; SHARE: %sp-2
-5         mov $0,-2(%sp)  ; x = 0
-6         mov %sp,%ra
-7         add -1,%ra      ; ra = &y
-8         call f          ; SHARE: %sp-1
-9         mov -2(%sp),%ra ; set return value
-10        sub $2,%sp      ; deallocate frame
-11        ret
-
-20 f:     cmp stash, $0   ; compare
-21        breq #13        ; branch if equal
-22        mov %ra,stash   ; stash = &a
-23        jmp #15
-24        mov $5, (stash) ; *stash = 5
-25        ret
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}{.29\textwidth}
-    \begin{tabular}{l l}
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$ \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}
-      \memory{1}{\mainsealc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{5}
-      \\
-      {\tt 8} &
-      \memoryaddrs[(0)]{12em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\mainpassc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{\bf 5}
-\end{tabular}
-\end{subfigure}
-\caption{A violation with pass-by-reference}
-\label{fig:passing}
-\end{figure}
-
-To this point we have ignored function arguments, since our simplest model places them in
-registers. This is an oversimplification, which we will now correct, aiming at two kinds
-of sharing: passing by value on the stack, and passing by reference. First we will explore
-some examples at the same level of detail as our earlier ones, and then move on to
-extending our formalization.
-
-Example~\ref{fig:passingsimple} shows a sample trace of
-a program that passes a value on the stack, without an attack.
-We extend our call annotation with an additional parameter identifying a memory location to be
-passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
-to include \(\passed{\depth,\depth'}\), recording two depths. We note the
-current depth of the active function next to the stack pointer in the diagram, for reference.
-The argument {\tt a} has
-its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
-as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
-modified, and naturally this is permitted as well.
-
-The same annotation mechanism can be used to pass
-variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference.
-{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
-{\tt f} again passing a reference to {\tt y}.
-We would expect that the first call to {\tt f} might modify
-{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
-
-To avoid cluttering the picture at the assembly level, we revert to assuming that the
-argument {\tt a} is passed in a register, namely {\tt \%ra}.
-The question is how the local variables {\tt x} and {\tt y} themselves
-should be labelled in order for the policy to detect an integrity violation.
-At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
-return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
-a future call with no arguments passed by reference would also violate integrity if it wrote
-to either variable.
-
-When passing by reference, it is possible to repeatedly pass the same reference further
-and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices to track two depths for each passed
-cell: the original owner, and the highest depth it has been passed to. The policy considers
-it to be accessible anywhere between the two.
-
-Our model provides only limited support for handling fully first-class pointers into
-stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike the case for pass-by-reference, there is no easy way to model the intended
-sharing behavior of such pointers in terms of stack frame activation patterns.
-The best approximation we can make is to treat an address-taken variable in the
-same way as a passed argument, with the additional property
-that it remains accessible at \emph{all} stack depths above the caller.
-This prevents most violations of temporal safety, although stale pointers can still be
-used in certain corner cases.
-More seriously, callees are permitted to access the location without having
-been explicitly given a pointer to it.  Preventing the latter kind of abuse
-is the provence of a fine-grained memory safety policy, e.g. based on capabilities
-(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
-\emph{per se}.
-
-Next we introduce variable passing. We will need to add extra information to our call
-annotations to reflect that they may now pass variables on the stack, and that the locations
-of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the components
-that should be passed---when passing by value, this will typically be freshly allocated
-cells at the top of the frame, while passing by reference just gives the
-addresses of the appropriate variables.
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-And we extend the set of domains to include passed components,
-which are labeled with the range of depths of the caller that has passed them.
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\rts \in \listT (\target \subseteq \MACHS)) \]
-We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map when pushing it, where \(\depth\) is an integer depth.
-\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
-\begin{cases}
-  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
-  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
-  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
-                                 \text{ for any } \depth' \\
-  \domMap~\component & \text{else} \\
-\end{cases}\]
-
-And to pop we just unseal the components at depth \(\depth\):
-
-\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
-\begin{cases}
-  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
-  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
-                              \text{ for any } \depth' < \depth \\
-  \domMap~\component        & \text{else} \\
-\end{cases}\]
-
-Now we can inductively define the context update function; these again apply in the order
-presented. \bcp{Too much space between the second and third.}
-
-\judgmenttwo[Call]
-            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
-            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmenttwo[ ReturnFound]
-            {\(\mach \stepsto \mach' \in \target\)}
-            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3.5ex}
-\judgment[ Default]
-         {}
-         {\(\_,\context \stepsto_C \context\)}
-
-\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
-or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
-\(|\rts| > \depth'\).
-
-\definition
-Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
-A system enjoys {\em stepwise stack integrity with passing} if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
-for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
-
-\definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
-let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
-and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
-\emph{stepwise stack confidentiality} if,
-for all \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\).
-
-\subsection{Coroutines}
-
-Our final extension to the model is to add coroutines. We assume a very simple model with
-a fixed number of coroutines, each with its own stack.
-Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
-pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We add a second annotation indicating yields.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
-
-The first major difference in this model is the domains. Stack domains are now nested inside
-top-level domains, and associated with stack identifiers.
-Our context now consists of a domain map, a map from stack identifiers to return target stacks,
-an additional map of targets for yields, and a stack identifier of the active stack.
-Our yield target map \(\yts\) is a map from stack identifiers to targets.
-In a standard system an initial yield target map will map each stack identifier to the
-set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
-\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
-
-\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
-\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
-(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
-\target) \times (\stackid \in \STACKS)\]
-
-In our context update rules, calls and returns work the same as in the sharing model,
-with domain updates restricted to the top-level domain corresponding to the active stack,
-and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
-
-\judgmentthreebr[ Yield]
-                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
-                {\(\mach \stepsto \mach'\)}
-                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
-                {\(\stackid \not = \stackid'\)}
-                {\(\yts ~ \stackid'\)}
-                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
-                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
-
-This calls for some explanation. First, we determine which coroutine we are yielding to by
-checking which stack the stack pointer of the next state is pointing to. Then we can only yield
-if the next state is a valid yield target for that stack, and we update the yield target for
-the current stack, ensuring that when the system yields back we expect the current coroutine
-to continue from where it left off. So, it is possible for a coroutine to claim to yield
-but not actually change the active stack, with similar effects to when a callee fails
-to return properly. Now we can define both stack safety and coroutine safety under this model.
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
-{\em stepwise stack integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
-A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
 
@@
 
-In this section we explore how an existing enforcement mechanism, {\em
-  Depth Isolation}, from \citet{DBLP:conf/sp/RoesslerD18} implements
-the formal stack-safety property described in the previous
-section. (In fact, it enforces a slightly stronger, inductive version
-of that property which strengthens the local state encapsulation
-component by checking for integrity and confidentiality violations at
-every step of the callee's execution trace.) We validate this claim
-with a property-based testing framework that can quickly uncover
-inconsistencies between the enforcement mechanism and the stack-safety
-property.
+In this section we sketch an existing enforcement mechanism, {\em Depth
+Isolation}, from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
+formal stack-safety property described in the previous section. (In
+fact, it enforces a slightly stronger, inductive version of that property which
+strengthens the local state encapsulation component by checking for
+integrity and confidentiality violations at every step of the callee's
+execution trace.) We validate this claim with a property-based
+testing framework that can quickly uncover inconsistencies between the
+enforcement mechanism and the stack-safety property.
 
@@
 sequences that have permission to manage the stack, say during calls and returns.
-\leo{If we have the simple exposition so far, no need to mention the following sentence:}
 In the following, for ease of exposition, we will assume that no
@@
 \item From callee back to caller, when execution is about to reach a
-  return point.
-  %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
-  At the actual return instruction,
-  the $\PCname$ tag is decremented.
-%  \leo{Not sure the actual policy
-%    does that now (retag unused). But I don't think it's a
-%    confidentiality or integrity violation.}
-%  \rb{Correct, and it doesn't. Some versions of the policy do clear
-%    the stack, retagging would be easily added.}
+  return point.  The exit sequence retags all locations in the
+  function frame with $\tagNoDepth$. At the actual return instruction,
+  the $\PCname$ tag is decremented. \leo{Not sure the actual policy
+    does that now (retag unused). But I don't think it's a
+    confidentiality or integrity violation.}
+  \rb{Correct, and it doesn't. Some versions of the policy do clear
+    the stack, retagging would be easily added.}
 
@@
 
-We validated the micro-policy of the previous section by
-systematically testing that it satisfies eager stack integrity and
+We validated the micro-policy of the previous section by systematically
+testing that it satisfies eager stack integrity and
 confidentiality. We used a Coq specification of the RISC-V
-architecture
-\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
-\leo{translated from Haskell via hs-2-coq?}  and extended it with a
-runtime monitor implementing the stack-safety
-micro-policy.\footnote{This code will be made public. \leo{More?}} We
-  chose the Coq proof assistant as the setting for our implementation
-  for two reasons: first, to ensure that our coinductive trace
-  definitions are well-formed, and reason about them\leo{Do we prove
-    anything interesting anymore?}, and to leverage the power of the
-  QuickChick property-based testing framework~\citep{Pierce:SF4}.
+architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
+extended it with a runtime monitor implementing the stack-safety
+micro-policy.\leo{Footnote: code will be made public}\bcp{For the reviewers?
+ Or later?} We chose the
+Coq proof assistant as the setting for our implementation for two
+reasons: first, to ensure that our coinductive trace definitions are
+well-formed, and reason about them\leo{Do we prove anything
+  interesting anymore?}, and to leverage the power of the QuickChick
+property-based testing framework~\citep{Pierce:SF4}.
 
@@
 
-To write such generators we build on the work of
-\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, where they
-introduced {\em generation by execution} to produce progams leading
-abstract machines towards interesting behaviors. Generation by
-execution receives as an input a partially instantiated machine state
-and attempts to generate an instruction (or a sequence of instructions
-such as a blessed sequence) that makes sense locally (e.g., jumps go
-to a potentially valid code location and loads read from from a
-potentially valid stack location). Then we step the machine and repeat
-the process until we generate or execute some target number of
-instructions, or if we reach a point where the machine can't step
-anymore.
+To write such generators we build on the work of \leo{ICFP 2013},
+where they introduced {\em generation by execution} to produce
+interesting programs. Generation by execution receives as an input a
+partially instantiated machine state and attempts to generate an
+instruction (or a sequence of instructions such as a blessed
+sequence) that makes sense locally (e.g., jumps go to a potentially valid
+code location and loads read from from a potentially valid stack
+location). Then
+we step the machine and repeat the process until we generate or
+execute some target number of instructions, or if we reach a point
+where the machine can't step anymore.
 
@@
 
-In addition, we needed to extend this technique to handle the nested
-nature of our confidentiality property: rather than generating two
-initial machines that are variant of one another and letting them
-execute to test for noninterference, we need to generate a new variant
-{\em every time a call is made} and check confidentiality for the
-subtrace produced from that variant state until its corresponding
-return. As a result, a ``single'' confidentiality test compactly
-checks the confidentiality of multiple nested calls.
-
 \paragraph*{Findings}
 
-To ensure the effectiveness of testing against our formal properties, we
-used {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
+To evaluate the effectiveness of testing against our formal properties, we
+used {\em mutation testing}\leo{cite}. In mutation testing, we inject errors
 (mutations) in a program that should cause the property of interest (here,
@@
 
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c}
-  Bug & Property Violated & MTTF (ms) \\
-  \hline
-      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
-      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
-      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:bug-table}
-\end{table}
-
-The mean-time-to-failure (MTTF) for various bugs can be found in
-Table~\ref{tab:bug-table}. All experiments were run in a desktop
-machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
-RAM. \leo{Something like: There is really no baseline to compare this
+The mean-time-to-failure (MTTF) for various bugs can be found in \leo{Table
+  with numbers?}\bcp{Sure.}. All experiments were run in a \leo{Machine
+  spec}. \leo{Something like: There is really no baseline to compare this
   against, but include it here for the future/reasons?}
-%
+
 The artificially injected bugs focus on weakening the tag propagation
-and checking mechanism. 
+and checking mechanism. \leo{Mention them all?}\bcp{Nah, especially if they
+  are all listed in the table.}
 %
@@
 current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
-that restriction by not checking the depth equality (row {\em
-  LOAD\_NO\_CHECK}).
+that restriction by not checking the depth equality, or not checking
+the tags at all, causing bugs of different difficulty.
 %
@@
 $\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing that constraints causes violations to the integrity
-property (row {\em STORE\_NO\_CHECK}).
+tag. Relaxing these constraints also causes violations to the
+property.
 
@@
 %\bcp{right?}\leo{yeah}
-; we re-implemented it in Coq targeting RISC-V).  These errors ranged
-from trivial typos to ones that required an intriguingly complex setup
-to appear.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}), in our first try the blessed call
-sequence/policy combination did not initialize all locations for the
-newly allocated stack frame correctly, but left some of them as
-$\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured: in
-particular, the smallest counterexample requires calling a function,
-writing in the uninitialized location (which is always fine), calling
-another function (to seal what was written for integrity purposes),
-and then attempt to write in the same location again from the new
-callee, which is only allowed because the enforcement still thinks
-the location is unused, but violates the integrity of the caller's data.
+; we re-implemented it in Coq targeting
+RISC-V).  These errors ranged from trivial typos to
+ones that required an
+intriguingly complex setup to appear.  Notably, our first try at the
+blessed call sequence/policy combination did not initialize all
+locations for the newly allocated stack frame correctly, but left some
+of them as $\tagNoDepth$. This allowed for a potential integrity
+violation, but only if a rather complicated sequence of events
+occured: \leo{call - write - call (seal) - return - call - write?}
 
@@
 not really stack safe, and describe a fix.
+
+\section{Extensions}
+
+The model above elucidates the meaning of our properties, but it does not
+correspond well to realistic machine. And our properties are designed to
+support a wide range of architectural choices. We will demonstrate the
+flexibility of the formalism by extending it to two interesting settings.
+First, a subrouting model in which arguments are passed on the stack,
+and second, a very simple coroutine model.
+
+\subsection{Parameters Passed on the Stack}
+
+\newcommand{\mainpassc}{magenta}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.2\textwidth}
+{\small
+\begin{verbatim}
+void main() {
+  int x = 5;
+  x = f(x);
+  return;
+}
+
+void f(int a) {
+  int y = 2*a;
+  a = 0;
+  return;
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}{.6\textwidth}
+{\small
+\begin{verbatim}
+1  main: add $1,%sp          ; allocate frame
+2        mov $5,-1(%sp)      ; initalize x
+3        add $1,%sp          ; allocate arg
+4        mov -2(%sp),-1(%sp) ; set arg
+5        call f              ; SHARE: %sp-1
+6        sub $1,%sp          ; deallocate arg
+7        mov %ra,-1(%sp)     ; store x
+8        sub $1,%sp          ; deallocate frame
+9        ret
+
+10  f:   add $1,%sp          ; allocate frame
+11       mov -3(%sp),-1(%sp) ; y = a
+12       mul $2,-1(%sp)      ; y = a*2
+13       mov $0,-3(%sp)      ; a = 0
+14       sub $1,%sp          ; deallocate frame
+15       ret
+\end{verbatim}
+}
+  \end{subfigure}
+\begin{subfigure}{.65\textwidth}
+\begin{center}
+\begin{tabular}{l r | l}
+  {\tt 3} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{5}
+%    \MemoryLabel{-14em}{0.75em}{0}
+%    \MemoryLabel{-10em}{0.75em}{0}
+%    \MemoryLabel{-7em}{0.75em}{0}
+    \vspace{.5em}
+  } \\
+  {\tt 11} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}[Seal(0)]%
+  \memory{1}{\mainpassc}[P(0,1)]%
+  \memory{1}{\retptrc}[RetPtr]%
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+%  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
+  \\
+  {\tt 13} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  {\tt 14} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  \\
+  {\tt 8} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{10}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{\#6}
+    \MemoryLabel{-7em}{0.75em}{10}
+    \vspace{.5em}
+  } \\
+\end{tabular}
+\end{center}
+
+\vspace{\abovedisplayskip}
+
+\end{subfigure}
+
+\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
+    the figure (at least in the latex build that I'm looking at...)}\fi}
+\label{fig:passingsimple}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{.2\textwidth}
+{\small
+\begin{verbatim}
+int *stash = 0;
+
+int main() {
+  int x, y;
+  f(x);
+  x = 0;
+  f(y);
+  return x;
+}
+
+void f(int &a) {
+  if(!stash)
+    stash = &a;
+  *stash = 5;
+  return;
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}{.49\textwidth}
+{\small
+\begin{verbatim}
+   stash: .word 0
+1  main:  add $2,%sp      ; allocate frame
+2         mov %sp,%ra
+3         add -2,%ra      ; ra = &x
+4         call f          ; SHARE: %sp-2
+5         mov $0,-2(%sp)  ; x = 0
+6         mov %sp,%ra
+7         add -1,%ra      ; ra = &y
+8         call f          ; SHARE: %sp-1
+9         mov -2(%sp),%ra ; set return value
+10        sub $2,%sp      ; deallocate frame
+11        ret
+
+20 f:     cmp stash, $0   ; compare
+21        breq #13        ; branch if equal
+22        mov %ra,stash   ; stash = &a
+23        jmp #15
+24        mov $5, (stash) ; *stash = 5
+25        ret
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}{.29\textwidth}
+    \begin{tabular}{l l}
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$ \\
+      {\tt 25} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainpassc}
+      \memory{1}{\mainsealc}
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{5}
+      \\
+      {\tt 8} &
+      \memoryaddrs[(0)]{12em}
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 25} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainsealc}
+      \memory{1}{\mainpassc}
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{\bf 5}
+\end{tabular}
+\end{subfigure}
+\caption{A violation with pass-by-reference}
+\label{fig:passing}
+\end{figure}
+
+\label{sec:ext}
+
+Our simple model placed function parameters in registers, which by definition need no protection.
+We add two kinds of protected sharing: passing by value on the stack, and passing by reference.
+First we will explore some examples at the same level of detail as our earlier ones,
+and then move on to extending our formalization.
+
+Example~\ref{fig:passingsimple} shows a sample trace of
+a program that passes a value on the stack, without an attack.
+We extend our call annotation with an additional parameter identifying a memory location to be
+passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
+to include \(\passed{\depth,\depth'}\), recording two depths. We note the
+current depth of the active function next to the stack pointer in the diagram, for reference.
+The argument {\tt a} has
+its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
+as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
+modified, and naturally this is permitted as well.
+
+The same annotation mechanism can be used to pass
+variables by reference, and here there are important stack safety concerns,
+as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
+{\tt f} takes a single argument by reference.
+{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
+{\tt f} again passing a reference to {\tt y}.
+We would expect that the first call to {\tt f} might modify
+{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
+{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
+
+To avoid cluttering the picture at the assembly level, we revert to assuming that the
+argument {\tt a} is passed in a register, namely {\tt \%ra}.
+The question is how the local variables {\tt x} and {\tt y} themselves
+should be labelled in order for the policy to detect an integrity violation.
+At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
+return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
+a future call with no arguments passed by reference would also violate integrity if it wrote
+to either variable.
+
+When passing by reference, it is possible to repeatedly pass the same reference further
+and further down the stack. Note, however, that this sharing of the reference must be over
+an unbroken chain of calls. So it suffices to track two depths for each passed
+cell: the original owner, and the highest depth it has been passed to. The policy considers
+it to be accessible anywhere between the two.
+
+Our model provides only limited support for handling fully first-class pointers into
+stack frames, e.g., pointers obtained by taking the address of a local variable.
+Unlike the case for pass-by-reference, there is no easy way to model the intended
+sharing behavior of such pointers in terms of stack frame activation patterns.
+The best approximation we can make is to treat an address-taken variable in the
+same way as a passed argument, with the additional property
+that it remains accessible at \emph{all} stack depths above the caller.
+This prevents most violations of temporal safety, although stale pointers can still be
+used in certain corner cases.
+More seriously, callees are permitted to access the location without having
+been explicitly given a pointer to it.  Preventing the latter kind of abuse
+is the provence of a fine-grained memory safety policy, e.g. based on capabilities
+(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
+\emph{per se}.
+
+Next we introduce variable passing. We will need to add extra information to our call
+annotations to reflect that they may now pass variables on the stack, and that the locations
+of those variables may be dynamic (as in the case of pass-by-reference).
+A call annotation now includes a predicate relating machine states to the components
+that should be passed---when passing by value, this will typically be freshly allocated
+cells at the top of the frame, while passing by reference just gives the
+addresses of the appropriate variables.
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
+And we extend the set of domains to include passed components,
+which are labeled with the range of depths of the caller that has passed them.
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+\times (\rts \in \listT (\target \subseteq \MACHS)) \]
+We straightforwardly describe how the sealing convention and passing predicate combine to
+update a domain map when pushing it, where \(\depth\) is an integer depth.
+\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
+\begin{cases}
+  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
+  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
+  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
+                                 \text{ for any } \depth' \\
+  \domMap~\component & \text{else} \\
+\end{cases}\]
+
+And to pop we just unseal the components at depth \(\depth\):
+
+\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
+\begin{cases}
+  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
+  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
+  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
+                              \text{ for any } \depth' < \depth \\
+  \domMap~\component        & \text{else} \\
+\end{cases}\]
+
+Now we can inductively define the context update function; these again apply in the order
+presented. \bcp{Too much space between the second and third.}
+
+\judgmenttwo[Call]
+            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
+            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
+            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\vspace*{-1ex}
+\judgmenttwo[ ReturnFound]
+            {\(\mach \stepsto \mach' \in \target\)}
+            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
+            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\vspace*{-1ex}
+\judgmentthree[ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\vspace*{-3.5ex}
+\judgment[ Default]
+         {}
+         {\(\_,\context \stepsto_C \context\)}
+
+\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
+be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
+or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
+\(|\rts| > \depth'\).
+
+\definition
+Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
+A system enjoys {\em stepwise stack integrity with passing} if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
+for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
+
+\definition
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
+let \(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
+and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
+\emph{stepwise stack confidentiality} if,
+for all \(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\).
+
+\subsection{Coroutines}
+
+Our final extension to the model is to add coroutines. We assume a very simple model with
+a fixed number of coroutines, each with its own stack.
+Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
+\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
+pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
+We add a second annotation indicating yields.
+
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
+
+The first major difference in this model is the domains. Stack domains are now nested inside
+top-level domains, and associated with stack identifiers.
+Our context now consists of a domain map, a map from stack identifiers to return target stacks,
+an additional map of targets for yields, and a stack identifier of the active stack.
+Our yield target map \(\yts\) is a map from stack identifiers to targets.
+In a standard system an initial yield target map will map each stack identifier to the
+set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
+\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
+
+\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
+\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
+(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
+\target) \times (\stackid \in \STACKS)\]
+
+In our context update rules, calls and returns work the same as in the sharing model,
+with domain updates restricted to the top-level domain corresponding to the active stack,
+and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
+
+\judgmentthreebr[ Yield]
+                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
+                {\(\mach \stepsto \mach'\)}
+                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
+                {\(\stackid \not = \stackid'\)}
+                {\(\yts ~ \stackid'\)}
+                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
+                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
+
+This calls for some explanation. First, we determine which coroutine we are yielding to by
+checking which stack the stack pointer of the next state is pointing to. Then we can only yield
+if the next state is a valid yield target for that stack, and we update the yield target for
+the current stack, ensuring that when the system yields back we expect the current coroutine
+to continue from where it left off. So, it is possible for a coroutine to claim to yield
+but not actually change the active stack, with similar effects to when a callee fails
+to return properly. Now we can define both stack safety and coroutine safety under this model.
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
+{\em stepwise stack integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where
+\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
+A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\_}\).
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
+{\em stepwise coroutine confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
 

### 2021-04-16 tex/main.tex
polishing example 3

--- 2021-04-16 tex/main.tex
@@
 \begin{figure}
-
 \begin{subfigure}[t]{.2\textwidth}
@@
     \memoryaddrs{8em}
-    \memory{4}{\unsealc}
+    \memory{3}{\unsealc}
     ~$\cdots$
@@
   \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
+  \memory{1}{\unsealc}
   ~$\cdots$
@@
   \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{62}
-  \MemoryLabel{-10em}{0.75em}{105}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
   \\
@@
   \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
+  \memory{1}{\unsealc}
   ~$\cdots$
  % \MemoryLabel{-19em}{0.75em}{0}
-  \MemoryLabel{-10em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{5}
 %  \MemoryLabel{-6em}{0.75em}{0}
@@
   \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-10em}{0.75em}{\(5\)}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
+  \MemoryLabel{-6em}{0.75em}{\(5\)}
   \\
@@
     \memory{1}{\retptrc}
-    \memory{2}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-18em}{0.75em}{5}
-    \MemoryLabel{-14em}{0.75em}{2}
-    \MemoryLabel{-10em}{0.75em}{5}
+    \memory{1}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-14em}{0.75em}{5}
+%    \MemoryLabel{-10em}{0.75em}{\(v_1\)}
+    \MemoryLabel{-6em}{0.75em}{5}
 %    \MemoryLabel{-7em}{0.75em}{0}
@@
   \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-10em}{0.75em}{\bf 5}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-15em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\bf 5}
  % \MemoryLabel{-6em}{0.75em}{0}
@@
   \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{36}
-  \MemoryLabel{-10em}{0.75em}{\bf 8}
-  \MemoryLabel{-6em}{0.75em}{90}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-15em}{0.75em}{\(v_3\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_4\)}
+%  \MemoryLabel{-2em}{0.75em}{90}
   \\
@@
 \end{subfigure}
-
 \caption{Another confidentiality violation}
@@
 In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
-that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
+that later becomes the value of {\tt main}'s variable {\tt x}. Then {\tt g} is able to access
 {\tt y} and hence {\tt x} without even directly reading the stack pointer.
+In the primary trace, {\tt z} will have the value 5, but in the variant it can have
+an arbitrary value $v_4$; since the variant can behave different from the primary,
+the policy again signals a confidentiality violation.
 
@@
   \end{subfigure}
-  \begin{subfigure}[t]{.25\textwidth}
+%
+  \begin{subfigure}[t]{.2\textwidth}
+    \begin{center}
     \begin{tabular}{l l l}
@@
     \end{tabular}
+    \end{center}
+    \vspace{\abovedisplayskip}
   \end{subfigure}
-  \caption{A control-flow attack}
+  \caption{A control-flow attack.}
   \label{fig:controlflow}
@@
   \end{subfigure}
-  \begin{subfigure}{.49\textwidth}
+  \begin{subfigure}[t]{.49\textwidth}
 {\small
@@
 
-20 f:     cmp stash, $0   ; compare
-21        breq #13        ; branch if equal
+20 f:     cmp stash, $0   ; stash = 0? 
+21        breq #13        ; branch if so
 22        mov %ra,stash   ; stash = &a
@@
   \end{subfigure}
-  \begin{subfigure}{.29\textwidth}
+  \begin{subfigure}[t]{.29\textwidth}
+    \begin{center}
     \begin{tabular}{l l}
@@
 \end{tabular}
+\end{center}
+\vspace{\abovedisplayskip}
 \end{subfigure}

### 2021-04-16 tex/main.tex

--- 2021-04-16 tex/main.tex
@@
 taking the address of a variable and treating it like a pointer.)
+\rb{But actually, this is discussed at some length is
+  \cref{sec:passing}, and it is the topic of \cref{fig:passing}.
+  Reaching that point after reading this disclaimer is confusing.}
 %
@@
 \subsection{Parameters Passed on the Stack}
+\label{sec:passing}
 
@@
 
-\subsection*{A Conservative Stack-Safety Enforcement Mechanism}
+\subsection{A Conservative Stack-Safety Enforcement Mechanism}
 %
@@
 
-\subsection*{Validation through Random Testing}
+\subsection{Validation through Random Testing}
 \label{sec:testing}

### 2021-04-16 tex/main.tex
Did the move

@@
 \item We show that the definition can be
-extended to argument passing on the stack and to a simple coroutine system.
-\bcp{fill in pointer to new section}
+extended to argument passing on the stack and to a simple coroutine system (\cref{sec:ext}).
 \end{itemize}
@@
 much the same as in StkTokens, just formalized a little differently.}
-
-\subsection{Parameters Passed on the Stack}
-\label{sec:passing}
-
-\newcommand{\mainpassc}{magenta}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-void main() {
-  int x = 5;
-  x = f(x);
-  return;
-}
-
-void f(int a) {
-  int y = 2*a;
-  a = 0;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.6\textwidth}
-{\small
-\begin{verbatim}
-1  main: add $1,%sp          ; allocate frame
-2        mov $5,-1(%sp)      ; initalize x
-3        add $1,%sp          ; allocate arg
-4        mov -2(%sp),-1(%sp) ; set arg
-5        call f              ; SHARE: %sp-1
-6        sub $1,%sp          ; deallocate arg
-7        mov %ra,-1(%sp)     ; store x
-8        sub $1,%sp          ; deallocate frame
-9        ret
-
-10  f:   add $1,%sp          ; allocate frame
-11       mov -3(%sp),-1(%sp) ; y = a
-12       mul $2,-1(%sp)      ; y = a*2
-13       mov $0,-3(%sp)      ; a = 0
-14       sub $1,%sp          ; deallocate frame
-15       ret
-\end{verbatim}
-}
-  \end{subfigure}
-\begin{subfigure}{.65\textwidth}
-\vspace*{0.2in}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 3} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{5}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[Seal(0)]%
-  \memory{1}{\mainpassc}[P(0,1)]%
-  \memory{1}{\retptrc}[RetPtr]%
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  {\tt 13} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  {\tt 14} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  \\
-  {\tt 8} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{10}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{\#6}
-    \MemoryLabel{-7em}{0.75em}{10}
-    \vspace{.5em}
-  } \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-
-\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
-    the figure (at least in the latex build that I'm looking at...)}\fi}
-\label{fig:passingsimple}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-int *stash = 0;
-int main() {
-  int x, y;
-  f(x);
-  x = 0;
-  f(y);
-  return x;
-}
-
-void f(int &a) {
-  if(!stash)
-    stash = &a;
-  *stash = 5;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.49\textwidth}
-{\small
-\begin{verbatim}
-   stash: .word 0
-1  main:  add $2,%sp      ; allocate frame
-2         mov %sp,%ra
-3         add -2,%ra      ; ra = &x
-4         call f          ; SHARE: %sp-2
-5         mov $0,-2(%sp)  ; x = 0
-6         mov %sp,%ra
-7         add -1,%ra      ; ra = &y
-8         call f          ; SHARE: %sp-1
-9         mov -2(%sp),%ra ; set return value
-10        sub $2,%sp      ; deallocate frame
-11        ret
-
-20 f:     cmp stash, $0   ; stash = 0? 
-21        breq #13        ; branch if so
-22        mov %ra,stash   ; stash = &a
-23        jmp #15
-24        mov $5, (stash) ; *stash = 5
-25        ret
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.29\textwidth}
-    \begin{center}
-    \begin{tabular}{l l}
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$ \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}
-      \memory{1}{\mainsealc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{5}
-      \\
-      {\tt 8} &
-      \memoryaddrs[(0)]{12em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\mainpassc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{\bf 5}
-\end{tabular}
-\end{center}
-\vspace{\abovedisplayskip}
-\end{subfigure}
-\caption{A violation with pass-by-reference}
-\label{fig:passing}
-\end{figure}
-
-To this point we have ignored function arguments, since our simplest model places them in
-registers. This is an oversimplification, which we will now correct, aiming at two kinds
-of sharing: passing by value on the stack, and passing by reference. First we will explore
-some examples at the same level of detail as our earlier ones, and then move on to
-extending our formalization.
-
-Example~\ref{fig:passingsimple} shows a sample trace of
-a program that passes a value on the stack, without an attack.
-We extend our call annotation with an additional parameter identifying a memory location to be
-passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
-to include \(\passed{\depth,\depth'}\), recording two depths. We note the
-current depth of the active function next to the stack pointer in the diagram, for reference.
-The argument {\tt a} has
-its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
-as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
-modified, and naturally this is permitted as well.
-
-The same annotation mechanism can be used to pass
-variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference.
-{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
-{\tt f} again passing a reference to {\tt y}.
-We would expect that the first call to {\tt f} might modify
-{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
-
-To avoid cluttering the picture at the assembly level, we revert to assuming that the
-argument {\tt a} is passed in a register, namely {\tt \%ra}.
-The question is how the local variables {\tt x} and {\tt y} themselves
-should be labelled in order for the policy to detect an integrity violation.
-At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
-return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
-a future call with no arguments passed by reference would also violate integrity if it wrote
-to either variable.
-
-When passing by reference, it is possible to repeatedly pass the same reference further
-and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices to track two depths for each passed
-cell: the original owner, and the highest depth it has been passed to. The policy considers
-it to be accessible anywhere between the two.
-
-Our model provides only limited support for handling fully first-class pointers into
-stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike the case for pass-by-reference, there is no easy way to model the intended
-sharing behavior of such pointers in terms of stack frame activation patterns.
-The best approximation we can make is to treat an address-taken variable in the
-same way as a passed argument, with the additional property
-that it remains accessible at \emph{all} stack depths above the caller.
-This prevents most violations of temporal safety, although stale pointers can still be
-used in certain corner cases.
-More seriously, callees are permitted to access the location without having
-been explicitly given a pointer to it.  Preventing the latter kind of abuse
-is the provence of a fine-grained memory safety policy, e.g. based on capabilities
-(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
-\emph{per se}.
-
-Next we introduce variable passing. We will need to add extra information to our call
-annotations to reflect that they may now pass variables on the stack, and that the locations
-of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the components
-that should be passed---when passing by value, this will typically be freshly allocated
-cells at the top of the frame, while passing by reference just gives the
-addresses of the appropriate variables.
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-And we extend the set of domains to include passed components,
-which are labeled with the range of depths of the caller that has passed them.
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\rts \in \listT (\target \subseteq \MACHS)) \]
-We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map when pushing it, where \(\depth\) is an integer depth.
-\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
-\begin{cases}
-  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
-  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
-  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
-                                 \text{ for any } \depth' \\
-  \domMap~\component & \text{else} \\
-\end{cases}\]
-
-And to pop we just unseal the components at depth \(\depth\):
-
-\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
-\begin{cases}
-  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
-  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
-                              \text{ for any } \depth' < \depth \\
-  \domMap~\component        & \text{else} \\
-\end{cases}\]
-
-Now we can inductively define the context update function; these again apply in the order
-presented. \bcp{Too much space between the second and third.}
-
-\judgmenttwo[Call]
-            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
-            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmenttwo[ ReturnFound]
-            {\(\mach \stepsto \mach' \in \target\)}
-            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3.5ex}
-\judgment[ Default]
-         {}
-         {\(\_,\context \stepsto_C \context\)}
-
-\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
-or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
-\(|\rts| > \depth'\).
-
-\definition
-Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
-A system enjoys {\em stepwise stack integrity with passing} if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
-for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
-
-\definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
-let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
-and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
-\emph{stepwise stack confidentiality} if,
-for all \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\).
-
-\subsection{Coroutines}
-
-Our final extension to the model is to add coroutines. We assume a very simple model with
-a fixed number of coroutines, each with its own stack.
-Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
-pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We add a second annotation indicating yields.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
-
-The first major difference in this model is the domains. Stack domains are now nested inside
-top-level domains, and associated with stack identifiers.
-Our context now consists of a domain map, a map from stack identifiers to return target stacks,
-an additional map of targets for yields, and a stack identifier of the active stack.
-Our yield target map \(\yts\) is a map from stack identifiers to targets.
-In a standard system an initial yield target map will map each stack identifier to the
-set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
-\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
-
-\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
-\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
-(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
-\target) \times (\stackid \in \STACKS)\]
-
-In our context update rules, calls and returns work the same as in the sharing model,
-with domain updates restricted to the top-level domain corresponding to the active stack,
-and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
-
-\judgmentthreebr[ Yield]
-                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
-                {\(\mach \stepsto \mach'\)}
-                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
-                {\(\stackid \not = \stackid'\)}
-                {\(\yts ~ \stackid'\)}
-                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
-                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
-
-This calls for some explanation. First, we determine which coroutine we are yielding to by
-checking which stack the stack pointer of the next state is pointing to. Then we can only yield
-if the next state is a valid yield target for that stack, and we update the yield target for
-the current stack, ensuring that when the system yields back we expect the current coroutine
-to continue from where it left off. So, it is possible for a coroutine to claim to yield
-but not actually change the active stack, with similar effects to when a callee fails
-to return properly. Now we can define both stack safety and coroutine safety under this model.
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
-{\em stepwise stack integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
-A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
 
@@
 
+\section{Extensions}
+\label{sec:ext}
+
+We have shown how our properties apply to an extremely simply
+subroutine model and validated them through randomized testing.
+But what about more realistic systems? We now demonstrate the
+flexibility of these core concepts as we extend
+the core concepts of stack safety to a system that models the passing of
+parameters on the stack, and then to a very simple coroutine model.
+
+\subsection{Parameters Passed on the Stack}
+\label{sec:passing}
+
+\newcommand{\mainpassc}{magenta}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.2\textwidth}
+{\small
+\begin{verbatim}
+void main() {
+  int x = 5;
+  x = f(x);
+  return;
+}
+
+void f(int a) {
+  int y = 2*a;
+  a = 0;
+  return;
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}[t]{.6\textwidth}
+{\small
+\begin{verbatim}
+1  main: add $1,%sp          ; allocate frame
+2        mov $5,-1(%sp)      ; initalize x
+3        add $1,%sp          ; allocate arg
+4        mov -2(%sp),-1(%sp) ; set arg
+5        call f              ; SHARE: %sp-1
+6        sub $1,%sp          ; deallocate arg
+7        mov %ra,-1(%sp)     ; store x
+8        sub $1,%sp          ; deallocate frame
+9        ret
+
+10  f:   add $1,%sp          ; allocate frame
+11       mov -3(%sp),-1(%sp) ; y = a
+12       mul $2,-1(%sp)      ; y = a*2
+13       mov $0,-3(%sp)      ; a = 0
+14       sub $1,%sp          ; deallocate frame
+15       ret
+\end{verbatim}
+}
+  \end{subfigure}
+\begin{subfigure}{.65\textwidth}
+\vspace*{0.2in}
+\begin{center}
+\begin{tabular}{l r | l}
+  {\tt 3} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{5}
+%    \MemoryLabel{-14em}{0.75em}{0}
+%    \MemoryLabel{-10em}{0.75em}{0}
+%    \MemoryLabel{-7em}{0.75em}{0}
+    \vspace{.5em}
+  } \\
+  {\tt 11} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}[Seal(0)]%
+  \memory{1}{\mainpassc}[P(0,1)]%
+  \memory{1}{\retptrc}[RetPtr]%
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+%  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
+  \\
+  {\tt 13} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  {\tt 14} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  \\
+  {\tt 8} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{10}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{\#6}
+    \MemoryLabel{-7em}{0.75em}{10}
+    \vspace{.5em}
+  } \\
+\end{tabular}
+\end{center}
+
+\vspace{\abovedisplayskip}
+
+\end{subfigure}
+
+\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
+    the figure (at least in the latex build that I'm looking at...)}\fi}
+\label{fig:passingsimple}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.2\textwidth}
+{\small
+\begin{verbatim}
+int *stash = 0;
+int main() {
+  int x, y;
+  f(x);
+  x = 0;
+  f(y);
+  return x;
+}
+
+void f(int &a) {
+  if(!stash)
+    stash = &a;
+  *stash = 5;
+  return;
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}[t]{.49\textwidth}
+{\small
+\begin{verbatim}
+   stash: .word 0
+1  main:  add $2,%sp      ; allocate frame
+2         mov %sp,%ra
+3         add -2,%ra      ; ra = &x
+4         call f          ; SHARE: %sp-2
+5         mov $0,-2(%sp)  ; x = 0
+6         mov %sp,%ra
+7         add -1,%ra      ; ra = &y
+8         call f          ; SHARE: %sp-1
+9         mov -2(%sp),%ra ; set return value
+10        sub $2,%sp      ; deallocate frame
+11        ret
+
+20 f:     cmp stash, $0   ; stash = 0? 
+21        breq #13        ; branch if so
+22        mov %ra,stash   ; stash = &a
+23        jmp #15
+24        mov $5, (stash) ; *stash = 5
+25        ret
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}[t]{.29\textwidth}
+    \begin{center}
+    \begin{tabular}{l l}
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$ \\
+      {\tt 25} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainpassc}
+      \memory{1}{\mainsealc}
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{5}
+      \\
+      {\tt 8} &
+      \memoryaddrs[(0)]{12em}
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 25} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainsealc}
+      \memory{1}{\mainpassc}
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{\bf 5}
+\end{tabular}
+\end{center}
+\vspace{\abovedisplayskip}
+\end{subfigure}
+\caption{A violation with pass-by-reference}
+\label{fig:passing}
+\end{figure}
+
+To this point we have ignored function arguments, since our simplest model places them in
+registers. We now implement two kinds of protected
+sharing: passing by value on the stack, and passing by reference. First we will explore
+some examples at the same level of detail as our earlier ones, and then move on to
+extending our formalization.
+
+Example~\ref{fig:passingsimple} shows a sample trace of
+a program that passes a value on the stack, without an attack.
+We extend our call annotation with an additional parameter identifying a memory location to be
+passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
+to include \(\passed{\depth,\depth'}\), recording two depths. We note the
+current depth of the active function next to the stack pointer in the diagram, for reference.
+The argument {\tt a} has
+its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
+as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
+modified, and naturally this is permitted as well.
+
+The same annotation mechanism can be used to pass
+variables by reference, and here there are important stack safety concerns,
+as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
+{\tt f} takes a single argument by reference.
+{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
+{\tt f} again passing a reference to {\tt y}.
+We would expect that the first call to {\tt f} might modify
+{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
+{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
+
+To avoid cluttering the picture at the assembly level, we revert to assuming that the
+argument {\tt a} is passed in a register, namely {\tt \%ra}.
+The question is how the local variables {\tt x} and {\tt y} themselves
+should be labelled in order for the policy to detect an integrity violation.
+At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
+return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
+a future call with no arguments passed by reference would also violate integrity if it wrote
+to either variable.
+
+When passing by reference, it is possible to repeatedly pass the same reference further
+and further down the stack. Note, however, that this sharing of the reference must be over
+an unbroken chain of calls. So it suffices to track two depths for each passed
+cell: the original owner, and the highest depth it has been passed to. The policy considers
+it to be accessible anywhere between the two.
+
+Our model provides only limited support for handling fully first-class pointers into
+stack frames, e.g., pointers obtained by taking the address of a local variable.
+Unlike the case for pass-by-reference, there is no easy way to model the intended
+sharing behavior of such pointers in terms of stack frame activation patterns.
+The best approximation we can make is to treat an address-taken variable in the
+same way as a passed argument, with the additional property
+that it remains accessible at \emph{all} stack depths above the caller.
+This prevents most violations of temporal safety, although stale pointers can still be
+used in certain corner cases.
+More seriously, callees are permitted to access the location without having
+been explicitly given a pointer to it.  Preventing the latter kind of abuse
+is the provence of a fine-grained memory safety policy, e.g. based on capabilities
+(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
+\emph{per se}.
+
+Next we introduce variable passing. We will need to add extra information to our call
+annotations to reflect that they may now pass variables on the stack, and that the locations
+of those variables may be dynamic (as in the case of pass-by-reference).
+A call annotation now includes a predicate relating machine states to the components
+that should be passed---when passing by value, this will typically be freshly allocated
+cells at the top of the frame, while passing by reference just gives the
+addresses of the appropriate variables.
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
+And we extend the set of domains to include passed components,
+which are labeled with the range of depths of the caller that has passed them.
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+\times (\rts \in \listT (\target \subseteq \MACHS)) \]
+We straightforwardly describe how the sealing convention and passing predicate combine to
+update a domain map when pushing it, where \(\depth\) is an integer depth.
+\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
+\begin{cases}
+  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
+  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
+  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
+                                 \text{ for any } \depth' \\
+  \domMap~\component & \text{else} \\
+\end{cases}\]
+
+And to pop we just unseal the components at depth \(\depth\):
+
+\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
+\begin{cases}
+  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
+  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
+  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
+                              \text{ for any } \depth' < \depth \\
+  \domMap~\component        & \text{else} \\
+\end{cases}\]
+
+Now we can inductively define the context update function; these again apply in the order
+presented. \bcp{Too much space between the second and third.}
+
+\judgmenttwo[Call]
+            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
+            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
+            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\vspace*{-1ex}
+\judgmenttwo[ ReturnFound]
+            {\(\mach \stepsto \mach' \in \target\)}
+            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
+            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\vspace*{-1ex}
+\judgmentthree[ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\vspace*{-3.5ex}
+\judgment[ Default]
+         {}
+         {\(\_,\context \stepsto_C \context\)}
+
+\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
+be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
+or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
+\(|\rts| > \depth'\).
+
+\definition
+Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
+A system enjoys {\em stepwise stack integrity with passing} if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
+for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
+
+\definition
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
+let \(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
+and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
+\emph{stepwise stack confidentiality} if,
+for all \(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\).
+
+\subsection{Coroutines}
+
+Our final extension to the model is to add coroutines. We assume a very simple model with
+a fixed number of coroutines, each with its own stack.
+Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
+\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
+pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
+We add a second annotation indicating yields.
+
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
+
+The first major difference in this model is the domains. Stack domains are now nested inside
+top-level domains, and associated with stack identifiers.
+Our context now consists of a domain map, a map from stack identifiers to return target stacks,
+an additional map of targets for yields, and a stack identifier of the active stack.
+Our yield target map \(\yts\) is a map from stack identifiers to targets.
+In a standard system an initial yield target map will map each stack identifier to the
+set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
+\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
+
+\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
+\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
+(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
+\target) \times (\stackid \in \STACKS)\]
+
+In our context update rules, calls and returns work the same as in the sharing model,
+with domain updates restricted to the top-level domain corresponding to the active stack,
+and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
+
+\judgmentthreebr[ Yield]
+                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
+                {\(\mach \stepsto \mach'\)}
+                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
+                {\(\stackid \not = \stackid'\)}
+                {\(\yts ~ \stackid'\)}
+                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
+                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
+
+This calls for some explanation. First, we determine which coroutine we are yielding to by
+checking which stack the stack pointer of the next state is pointing to. Then we can only yield
+if the next state is a valid yield target for that stack, and we update the yield target for
+the current stack, ensuring that when the system yields back we expect the current coroutine
+to continue from where it left off. So, it is possible for a coroutine to claim to yield
+but not actually change the active stack, with similar effects to when a callee fails
+to return properly. Now we can define both stack safety and coroutine safety under this model.
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
+{\em stepwise stack integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where
+\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
+A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\_}\).
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
+{\em stepwise coroutine confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
 \section{Related Work}

### 2021-04-16 tex/main.tex
polishing definition of stepwise confidentiality

--- 2021-04-16 tex/main.tex
@@
 let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\). That is, at the start of some call, we take every
-component in the stack.
+\(\domMap ~ \component = \outside\).
+%That is, at the start of some call, we take every component in the stack.
 A system enjoys \emph{stepwise stack confidentiality} if,
-for all \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\)---if every variant (with the same policy state)
-steps in lockstep until both traces have reached a point whose depth is
-below the starting one. This is a return.
-
-\medskip
+for all \(\nach\) and \(\NPCT\) such that \(\mach \approx_\components \nach\)
+and \(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\) holds.
+
+%\medskip
+
+In other words, stepwise stack confidentiality states that if we vary every component of the stack,
+the resulting trace (starting in the same policy state) steps in lockstep with the
+original trace until both have reached a point whose depth is below the starting
+depth (this is a return).
 

### 2021-04-16 tex/main.tex
Threat model discussion

--- 2021-04-16 tex/main.tex
@@
 
-\bcp{I think we are not explicit enough about the sorts of threat situations
-  we imagine we are addressing.  A canonical example (that we could spell
-  out) is a developer who wants to call out to potentially buggy or
-  malicious libraries.  The library code is going to be linked with the
-  developer's in the same address space, but the developer wants her private
-  data to be protected from both reading/exfiltrating and
-  overwriting... More generally, we can talk about the fact that programmers
-  use simple reasoning principles to think about their programs (e.g., local
-  variables don't change across function calls), and life is better when
-  these principles are actually true...}
+Consider a common scenario: a developer wants to link against an existing
+code-base, which will then occupy the same address space as their own code.
+But perhaps the external code is buggy, or even malicious, and contains
+exploits that can be used to cause it to modify the developer's data or
+exfiltrate secrets. For that matter, the developer may not be a particularly
+secure programmer, and may not be able to rely on their own defensive programming
+to keep their secrets safe.
+
+The developer likely relies instead on simple reasoning principles about how
+their code should work. Local variables do not change across function calls,
+and the callees generally do the same thing regardless of when and where they
+are called (up to a reasonable interface, such as their parameters and global
+variables.) If these intuitions prove false, then the programmer will reason
+incorrectly about their program and especially about its security, and
+be exposed to attack.
 

### 2021-04-16 tex/main.tex
Address heap

@@
 arbitrary machine code. Hardware and timing attacks are out of scope.
-
-\bcp{We should say explicitly that we do not address attacks on the heap
-  (cite Arthur's paper and say we believe it can be combined smoothly with
-  this), direct attacks on code (we assume that code can't be over written),
-etc.}
+So are attacks on the heap -- we believe that the work of
+\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with our
+to protect both stack and heap.
 

### 2021-04-16 tex/main.tex
Removed extraneous bit of examples section

@@
 This section introduces the intuition behind the properties through a series
-of examples. Existing stack safety enforcement work mostly refers directly to examples
-of bad behavior that it prevents; we will argue that: \bcp{Not sure this
-  enumeration is useful.}\apt{+1}
-
-\begin{itemize}
-\item examples that appear to violate stack safety violate at least one of our properties;
-\item we do not find desireable behaviors that appear stack-safe yet violate our properties; and
-\item our properties are testably implemented by stack safety micro-policies from the
-  literature, and the micro-policies are not significantly more conservative
-  (i.e., policy failstops usually correspond to imminent property violations).
-\end{itemize}
-
-We show our examples both in C syntax and in corresponding assembly code for a
+of examples. We show our examples both in C syntax and in corresponding assembly code for a
 simple two-address machine. The address of each instruction is shown to its left.

### 2021-04-16 tex/main.tex
Nits in sections 3-4

@@
 state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
-understand if we chose concrete values for the examples}\fi \(v_0\), etc., in place of the values that should
+understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
 be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
@@
 
-It may be unclear in figure \ref{fig:conf1} why we vary the initial values of {\tt y}
+It may be unclear in Figure~\ref{fig:conf1} why we vary the initial values of {\tt y}
 and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
@@
 Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
-Consider example \ref{fig:controlflow}.
+Consider Figure~\ref{fig:controlflow}.
 Here we show return addresses explicitly as line numbers, in the form $\#n$.

### 2021-04-16 tex/main.tex
Notation fixes in 5

@@
 which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special purpose registers: the
+\(\REGS\), which is assumed to include two special-purpose registers: the
 program counter {\PCname} and the stack pointer \(\SP\).
@@
 We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
+\rb{Could we use $\cdot$?}
 
@@
 The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
-and a predicate on states \(f ~ \mpcstate \in \MPCS\) and gives the prefix of
+and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
 \(\MPCT\) ending with the first element on which \(f\) holds.
@@
 
-%% CHERI \rb{cite}
-
 \paragraph{Formal Stack Safety on Capability Machines}
@@
 programs. Although the required hardware support is readily available
-in capability machines like CHERI, this technique incurs significant
+in capability machines like Cheri, this technique incurs significant
 costs, because it requires the entire unused part of the stack to be

### 2021-04-16 tex/main.tex
Restore missing text from dd43ba24

@@
 
-\apt{Some words have been lost here.}
+In particular, we wish to formulate properties around consecutive subtraces
+in which the context obeys a given predicate. This will commonly be used to extract
 individual calls from a trace by taking subtraces in which the stack is

### 2021-04-16 tex/main.tex
CP

--- 2021-04-16 tex/main.tex
@@
 LSE) and 
-control flow (like WBCF).  But our formulations of these properties are
-rather different from those of Skorstengaard et al.  Instead of
-defining stack safety implicitly,
-as ``what a certain idealized machine does,'' we define it
-\emph{explicitly} using the technical framework of language-based
+control flow (like WBCF).  Our technical formulation of WBCF is similar to
+that of Skorstengaard et al.{}, but our treatment of stack data protection 
+is rather different.  Instead of defining it implicitly,
+as ``what a certain idealized machine does,'' we give an
+\emph{explicit} definition using the technical framework of language-based
 security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
-  survey?}\fi.
-Our core properties are \emph{stepwise stack integrity},
-\emph{stepwise stack confidentiality}, and \emph{return to caller}.
-%
-A key technical novelty in the definitions of confidentiality and integrity,
-compared to standard formulations from the security literature, is that they
+  survey?}\fi, phrased as the conjunction of
+\emph{stack integrity} and \emph{stack confidentiality}.
+%
+A key technical novelty in these definitions,
+compared to standard formulations of confidentiality and integrity from the
+security literature, is that they
 are ``nested'': {\em each} caller is guaranteed protection from its
@@
 \rb{\cref{sec:RTC}?}
-This reflects a cleaner division between data protection and control-flow protection
-than offered by LSE and WBCF.
+This reflects an arguably cleaner division between data protection and
+control-flow protection than LSE / WBCF.
 
@@
 \item We show that the definition can be
-extended to argument passing on the stack and to a simple coroutine system (\cref{sec:ext}).
+extended to argument passing on the stack and to a simple coroutine system.
+\bcp{fill in pointer to new section}
 \end{itemize}
@@
 
-Consider a common scenario: a developer wants to link against an existing
-code-base, which will then occupy the same address space as their own code.
-But perhaps the external code is buggy, or even malicious, and contains
-exploits that can be used to cause it to modify the developer's data or
-exfiltrate secrets. For that matter, the developer may not be a particularly
-secure programmer, and may not be able to rely on their own defensive programming
-to keep their secrets safe.
-
-The developer likely relies instead on simple reasoning principles about how
-their code should work. Local variables do not change across function calls,
-and the callees generally do the same thing regardless of when and where they
-are called (up to a reasonable interface, such as their parameters and global
-variables.) If these intuitions prove false, then the programmer will reason
-incorrectly about their program and especially about its security, and
-be exposed to attack.
+\bcp{I think we are not explicit enough about the sorts of threat situations
+  we imagine we are addressing.  A canonical example (that we could spell
+  out) is a developer who wants to call out to potentially buggy or
+  malicious libraries.  The library code is going to be linked with the
+  developer's in the same address space, but the developer wants her private
+  data to be protected from both reading/exfiltrating and
+  overwriting... More generally, we can talk about the fact that programmers
+  use simple reasoning principles to think about their programs (e.g., local
+  variables don't change across function calls), and life is better when
+  these principles are actually true...}
 
@@
 arbitrary machine code. Hardware and timing attacks are out of scope.
-So are attacks on the heap -- we believe that the work of
-\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with our
-to protect both stack and heap.
+
+\bcp{We should say explicitly that we do not address attacks on the heap
+  (cite Arthur's paper and say we believe it can be combined smoothly with
+  this), direct attacks on code (we assume that code can't be over written),
+etc.}
 
@@
 \end{center}
-
-\vspace{\abovedisplayskip}
+\vspace{-4ex}
 \end{subfigure}
@@
 This section introduces the intuition behind the properties through a series
-of examples. We show our examples both in C syntax and in corresponding assembly code for a
+of examples. Existing stack safety enforcement work mostly refers directly to examples
+of bad behavior that it prevents; we will argue that: \bcp{Not sure this
+  enumeration is useful.}\apt{+1}
+
+\begin{itemize}
+\item examples that appear to violate stack safety violate at least one of our properties;
+\item we do not find desireable behaviors that appear stack-safe yet violate our properties; and
+\item our properties are testably implemented by stack safety micro-policies from the
+  literature, and the micro-policies are not significantly more conservative
+  (i.e., policy failstops usually correspond to imminent property violations).
+\end{itemize}
+
+We show our examples both in C syntax and in corresponding assembly code for a
 simple two-address machine. The address of each instruction is shown to its left.
@@
 state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
-understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
+understand if we chose concrete values for the examples}\fi \(v_0\), etc., in place of the values that should
 be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
@@
 \begin{figure}
+
 \begin{subfigure}[t]{.2\textwidth}
@@
     \memoryaddrs{8em}
-    \memory{3}{\unsealc}
+    \memory{4}{\unsealc}
     ~$\cdots$
@@
   \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
+  \memory{2}{\unsealc}
   ~$\cdots$
@@
   \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{62}
+  \MemoryLabel{-10em}{0.75em}{105}
   \\
@@
   \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
+  \memory{2}{\unsealc}
   ~$\cdots$
  % \MemoryLabel{-19em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{5}
+  \MemoryLabel{-10em}{0.75em}{5}
 %  \MemoryLabel{-6em}{0.75em}{0}
@@
   \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(5\)}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-10em}{0.75em}{\(5\)}
   \\
@@
     \memory{1}{\retptrc}
-    \memory{1}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-14em}{0.75em}{5}
-%    \MemoryLabel{-10em}{0.75em}{\(v_1\)}
-    \MemoryLabel{-6em}{0.75em}{5}
+    \memory{2}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-18em}{0.75em}{5}
+    \MemoryLabel{-14em}{0.75em}{2}
+    \MemoryLabel{-10em}{0.75em}{5}
 %    \MemoryLabel{-7em}{0.75em}{0}
@@
   \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\bf 5}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-10em}{0.75em}{\bf 5}
  % \MemoryLabel{-6em}{0.75em}{0}
@@
   \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_3\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_4\)}
-%  \MemoryLabel{-2em}{0.75em}{90}
+  \memory{2}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{36}
+  \MemoryLabel{-10em}{0.75em}{\bf 8}
+  \MemoryLabel{-6em}{0.75em}{90}
   \\
@@
 \end{subfigure}
+
 \caption{Another confidentiality violation}
@@
 
-It may be unclear in Figure~\ref{fig:conf1} why we vary the initial values of {\tt y}
+It may be unclear in figure \ref{fig:conf1} why we vary the initial values of {\tt y}
 and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
@@
 In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
-that later becomes the value of {\tt main}'s variable {\tt x}. Then {\tt g} is able to access
+that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
 {\tt y} and hence {\tt x} without even directly reading the stack pointer.
-In the primary trace, {\tt z} will have the value 5, but in the variant it can have
-an arbitrary value $v_4$; since the variant can behave different from the primary,
-the policy again signals a confidentiality violation.
 
@@
 Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
-Consider Figure~\ref{fig:controlflow}.
+Consider example \ref{fig:controlflow}.
 Here we show return addresses explicitly as line numbers, in the form $\#n$.
@@
 let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\).
-%That is, at the start of some call, we take every component in the stack.
+\(\domMap ~ \component = \outside\). That is, at the start of some call, we take every
+component in the stack.
 A system enjoys \emph{stepwise stack confidentiality} if,
-for all \(\nach\) and \(\NPCT\) such that \(\mach \approx_\components \nach\)
-and \(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\) holds.
-
-%\medskip
-
-In other words, stepwise stack confidentiality states that if we vary every component of the stack,
-the resulting trace (starting in the same policy state) steps in lockstep with the
-original trace until both have reached a point whose depth is below the starting
-depth (this is a return).
+for all \(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\)---if every variant (with the same policy state)
+steps in lockstep until both traces have reached a point whose depth is
+below the starting one. This is a return.
+
+\medskip
 
@@
 
+\subsection{Parameters Passed on the Stack}
+\label{sec:passing}
+
+\newcommand{\mainpassc}{magenta}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.2\textwidth}
+{\small
+\begin{verbatim}
+void main() {
+  int x = 5;
+  x = f(x);
+  return;
+}
+
+void f(int a) {
+  int y = 2*a;
+  a = 0;
+  return;
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}[t]{.6\textwidth}
+{\small
+\begin{verbatim}
+1  main: add $1,%sp          ; allocate frame
+2        mov $5,-1(%sp)      ; initalize x
+3        add $1,%sp          ; allocate arg
+4        mov -2(%sp),-1(%sp) ; set arg
+5        call f              ; SHARE: %sp-1
+6        sub $1,%sp          ; deallocate arg
+7        mov %ra,-1(%sp)     ; store x
+8        sub $1,%sp          ; deallocate frame
+9        ret
+
+10  f:   add $1,%sp          ; allocate frame
+11       mov -3(%sp),-1(%sp) ; y = a
+12       mul $2,-1(%sp)      ; y = a*2
+13       mov $0,-3(%sp)      ; a = 0
+14       sub $1,%sp          ; deallocate frame
+15       ret
+\end{verbatim}
+}
+  \end{subfigure}
+\begin{subfigure}{.65\textwidth}
+\vspace*{0.2in}
+\begin{center}
+\begin{tabular}{l r | l}
+  {\tt 3} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{5}
+%    \MemoryLabel{-14em}{0.75em}{0}
+%    \MemoryLabel{-10em}{0.75em}{0}
+%    \MemoryLabel{-7em}{0.75em}{0}
+    \vspace{.5em}
+  } \\
+  {\tt 11} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}[Seal(0)]%
+  \memory{1}{\mainpassc}[P(0,1)]%
+  \memory{1}{\retptrc}[RetPtr]%
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+%  \MemoryLabel{-6em}{0.75em}{0}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
+  \\
+  {\tt 13} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{5}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  {\tt 14} &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{5}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  &
+  \memoryaddrs[(1)]{21.5em}
+  \memory{1}{\mainsealc}
+  \memory{1}{\mainpassc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-14em}{0.75em}{0}
+  \MemoryLabel{-6em}{0.75em}{10}
+  \\
+  \\
+  {\tt 8} &
+  \multicolumn{2}{c}{
+    \memoryaddrs[(0)]{8em}
+    \memory{4}{\unsealc}
+    ~$\cdots$
+    \MemoryLabel{-19em}{0.75em}{10}
+    \MemoryLabel{-14em}{0.75em}{0}
+    \MemoryLabel{-10em}{0.75em}{\#6}
+    \MemoryLabel{-7em}{0.75em}{10}
+    \vspace{.5em}
+  } \\
+\end{tabular}
+\end{center}
+
+\vspace{\abovedisplayskip}
+
+\end{subfigure}
+
+\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
+    the figure (at least in the latex build that I'm looking at...)}\fi}
+\label{fig:passingsimple}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.2\textwidth}
+{\small
+\begin{verbatim}
+int *stash = 0;
+int main() {
+  int x, y;
+  f(x);
+  x = 0;
+  f(y);
+  return x;
+}
+
+void f(int &a) {
+  if(!stash)
+    stash = &a;
+  *stash = 5;
+  return;
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}[t]{.49\textwidth}
+{\small
+\begin{verbatim}
+   stash: .word 0
+1  main:  add $2,%sp      ; allocate frame
+2         mov %sp,%ra
+3         add -2,%ra      ; ra = &x
+4         call f          ; SHARE: %sp-2
+5         mov $0,-2(%sp)  ; x = 0
+6         mov %sp,%ra
+7         add -1,%ra      ; ra = &y
+8         call f          ; SHARE: %sp-1
+9         mov -2(%sp),%ra ; set return value
+10        sub $2,%sp      ; deallocate frame
+11        ret
+
+20 f:     cmp stash, $0   ; compare
+21        breq #13        ; branch if equal
+22        mov %ra,stash   ; stash = &a
+23        jmp #15
+24        mov $5, (stash) ; *stash = 5
+25        ret
+}
+\end{verbatim}
+}
+  \end{subfigure}
+  \begin{subfigure}[t]{.29\textwidth}
+    \begin{center}
+    \begin{tabular}{l l}
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$ \\
+      {\tt 25} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainpassc}
+      \memory{1}{\mainsealc}
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{5}
+      \\
+      {\tt 8} &
+      \memoryaddrs[(0)]{12em}
+      \memory{3}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 20} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{0}
+      \\
+      {\tt 25} &
+      \memoryaddrs[(1)]{17em}
+      \memory{1}{\mainsealc}
+      \memory{1}{\mainpassc}
+      \memory{1}{\retptrc}
+%      \memory{1}{\unsealc}
+      ~$\cdots$
+      \MemoryLabel{-18em}{0.75em}{\bf 5}
+\end{tabular}
+\end{center}
+\vspace{\abovedisplayskip}
+\end{subfigure}
+\caption{A violation with pass-by-reference}
+\label{fig:passing}
+\end{figure}
+
+To this point we have ignored function arguments, since our simplest model places them in
+registers. This is an oversimplification, which we will now correct, aiming at two kinds
+of sharing: passing by value on the stack, and passing by reference. First we will explore
+some examples at the same level of detail as our earlier ones, and then move on to
+extending our formalization.
+
+Example~\ref{fig:passingsimple} shows a sample trace of
+a program that passes a value on the stack, without an attack.
+We extend our call annotation with an additional parameter identifying a memory location to be
+passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
+to include \(\passed{\depth,\depth'}\), recording two depths. We note the
+current depth of the active function next to the stack pointer in the diagram, for reference.
+The argument {\tt a} has
+its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
+as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
+modified, and naturally this is permitted as well.
+
+The same annotation mechanism can be used to pass
+variables by reference, and here there are important stack safety concerns,
+as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
+{\tt f} takes a single argument by reference.
+{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
+{\tt f} again passing a reference to {\tt y}.
+We would expect that the first call to {\tt f} might modify
+{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
+{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
+
+To avoid cluttering the picture at the assembly level, we revert to assuming that the
+argument {\tt a} is passed in a register, namely {\tt \%ra}.
+The question is how the local variables {\tt x} and {\tt y} themselves
+should be labelled in order for the policy to detect an integrity violation.
+At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
+return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
+a future call with no arguments passed by reference would also violate integrity if it wrote
+to either variable.
+
+When passing by reference, it is possible to repeatedly pass the same reference further
+and further down the stack. Note, however, that this sharing of the reference must be over
+an unbroken chain of calls. So it suffices to track two depths for each passed
+cell: the original owner, and the highest depth it has been passed to. The policy considers
+it to be accessible anywhere between the two.
+
+Our model provides only limited support for handling fully first-class pointers into
+stack frames, e.g., pointers obtained by taking the address of a local variable.
+Unlike the case for pass-by-reference, there is no easy way to model the intended
+sharing behavior of such pointers in terms of stack frame activation patterns.
+The best approximation we can make is to treat an address-taken variable in the
+same way as a passed argument, with the additional property
+that it remains accessible at \emph{all} stack depths above the caller.
+This prevents most violations of temporal safety, although stale pointers can still be
+used in certain corner cases.
+More seriously, callees are permitted to access the location without having
+been explicitly given a pointer to it.  Preventing the latter kind of abuse
+is the provence of a fine-grained memory safety policy, e.g. based on capabilities
+(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
+\emph{per se}.
+
+Next we introduce variable passing. We will need to add extra information to our call
+annotations to reflect that they may now pass variables on the stack, and that the locations
+of those variables may be dynamic (as in the case of pass-by-reference).
+A call annotation now includes a predicate relating machine states to the components
+that should be passed---when passing by value, this will typically be freshly allocated
+cells at the top of the frame, while passing by reference just gives the
+addresses of the appropriate variables.
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
+And we extend the set of domains to include passed components,
+which are labeled with the range of depths of the caller that has passed them.
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+\times (\rts \in \listT (\target \subseteq \MACHS)) \]
+We straightforwardly describe how the sealing convention and passing predicate combine to
+update a domain map when pushing it, where \(\depth\) is an integer depth.
+\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
+\begin{cases}
+  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
+  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
+  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
+                                 \text{ for any } \depth' \\
+  \domMap~\component & \text{else} \\
+\end{cases}\]
+
+And to pop we just unseal the components at depth \(\depth\):
+
+\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
+\begin{cases}
+  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
+  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
+  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
+                              \text{ for any } \depth' < \depth \\
+  \domMap~\component        & \text{else} \\
+\end{cases}\]
+
+Now we can inductively define the context update function; these again apply in the order
+presented. \bcp{Too much space between the second and third.}
+
+\judgmenttwo[Call]
+            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
+            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
+            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\vspace*{-1ex}
+\judgmenttwo[ ReturnFound]
+            {\(\mach \stepsto \mach' \in \target\)}
+            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
+            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\vspace*{-1ex}
+\judgmentthree[ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\vspace*{-3.5ex}
+\judgment[ Default]
+         {}
+         {\(\_,\context \stepsto_C \context\)}
+
+\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
+be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
+or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
+\(|\rts| > \depth'\).
+
+\definition
+Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
+A system enjoys {\em stepwise stack integrity with passing} if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
+for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
+
+\definition
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
+let \(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
+and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
+\emph{stepwise stack confidentiality} if,
+for all \(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\).
+
+\subsection{Coroutines}
+
+Our final extension to the model is to add coroutines. We assume a very simple model with
+a fixed number of coroutines, each with its own stack.
+Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
+\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
+pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
+We add a second annotation indicating yields.
+
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
+
+The first major difference in this model is the domains. Stack domains are now nested inside
+top-level domains, and associated with stack identifiers.
+Our context now consists of a domain map, a map from stack identifiers to return target stacks,
+an additional map of targets for yields, and a stack identifier of the active stack.
+Our yield target map \(\yts\) is a map from stack identifiers to targets.
+In a standard system an initial yield target map will map each stack identifier to the
+set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
+\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
+
+\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
+\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
+(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
+\target) \times (\stackid \in \STACKS)\]
+
+In our context update rules, calls and returns work the same as in the sharing model,
+with domain updates restricted to the top-level domain corresponding to the active stack,
+and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
+
+\judgmentthreebr[ Yield]
+                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
+                {\(\mach \stepsto \mach'\)}
+                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
+                {\(\stackid \not = \stackid'\)}
+                {\(\yts ~ \stackid'\)}
+                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
+                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
+
+This calls for some explanation. First, we determine which coroutine we are yielding to by
+checking which stack the stack pointer of the next state is pointing to. Then we can only yield
+if the next state is a valid yield target for that stack, and we update the yield target for
+the current stack, ensuring that when the system yields back we expect the current coroutine
+to continue from where it left off. So, it is possible for a coroutine to claim to yield
+but not actually change the active stack, with similar effects to when a callee fails
+to return properly. Now we can define both stack safety and coroutine safety under this model.
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
+{\em stepwise stack integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
+Let \(\components\) be the set of components \(\component\) where
+\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
+A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\_}\).
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
+{\em stepwise coroutine confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
 %\subsection{Well-bracketed Control Flow}
@@
 
-\subsection{A Conservative Stack-Safety Enforcement Mechanism}
+\subsection*{A Conservative Stack-Safety Enforcement Mechanism}
 %
@@
 
-\subsection{Validation through Random Testing}
+\subsection*{Validation through Random Testing}
 \label{sec:testing}
@@
 not really stack safe, and describe a fix.
-
-\section{Extensions}
-\label{sec:ext}
-
-We have shown how our properties apply to an extremely simply
-subroutine model and validated them through randomized testing.
-But what about more realistic systems? We now demonstrate the
-flexibility of these core concepts as we extend
-the core concepts of stack safety to a system that models the passing of
-parameters on the stack, and then to a very simple coroutine model.
-
-\subsection{Parameters Passed on the Stack}
-\label{sec:passing}
-
-\newcommand{\mainpassc}{magenta}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-void main() {
-  int x = 5;
-  x = f(x);
-  return;
-}
-
-void f(int a) {
-  int y = 2*a;
-  a = 0;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.6\textwidth}
-{\small
-\begin{verbatim}
-1  main: add $1,%sp          ; allocate frame
-2        mov $5,-1(%sp)      ; initalize x
-3        add $1,%sp          ; allocate arg
-4        mov -2(%sp),-1(%sp) ; set arg
-5        call f              ; SHARE: %sp-1
-6        sub $1,%sp          ; deallocate arg
-7        mov %ra,-1(%sp)     ; store x
-8        sub $1,%sp          ; deallocate frame
-9        ret
-
-10  f:   add $1,%sp          ; allocate frame
-11       mov -3(%sp),-1(%sp) ; y = a
-12       mul $2,-1(%sp)      ; y = a*2
-13       mov $0,-3(%sp)      ; a = 0
-14       sub $1,%sp          ; deallocate frame
-15       ret
-\end{verbatim}
-}
-  \end{subfigure}
-\begin{subfigure}{.65\textwidth}
-\vspace*{0.2in}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 3} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{5}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[Seal(0)]%
-  \memory{1}{\mainpassc}[P(0,1)]%
-  \memory{1}{\retptrc}[RetPtr]%
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  {\tt 13} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  {\tt 14} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  \\
-  {\tt 8} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{10}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{\#6}
-    \MemoryLabel{-7em}{0.75em}{10}
-    \vspace{.5em}
-  } \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-
-\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
-    the figure (at least in the latex build that I'm looking at...)}\fi}
-\label{fig:passingsimple}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-int *stash = 0;
-int main() {
-  int x, y;
-  f(x);
-  x = 0;
-  f(y);
-  return x;
-}
-
-void f(int &a) {
-  if(!stash)
-    stash = &a;
-  *stash = 5;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.49\textwidth}
-{\small
-\begin{verbatim}
-   stash: .word 0
-1  main:  add $2,%sp      ; allocate frame
-2         mov %sp,%ra
-3         add -2,%ra      ; ra = &x
-4         call f          ; SHARE: %sp-2
-5         mov $0,-2(%sp)  ; x = 0
-6         mov %sp,%ra
-7         add -1,%ra      ; ra = &y
-8         call f          ; SHARE: %sp-1
-9         mov -2(%sp),%ra ; set return value
-10        sub $2,%sp      ; deallocate frame
-11        ret
-
-20 f:     cmp stash, $0   ; stash = 0? 
-21        breq #13        ; branch if so
-22        mov %ra,stash   ; stash = &a
-23        jmp #15
-24        mov $5, (stash) ; *stash = 5
-25        ret
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.29\textwidth}
-    \begin{center}
-    \begin{tabular}{l l}
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$ \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}
-      \memory{1}{\mainsealc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{5}
-      \\
-      {\tt 8} &
-      \memoryaddrs[(0)]{12em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\mainpassc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{\bf 5}
-\end{tabular}
-\end{center}
-\vspace{\abovedisplayskip}
-\end{subfigure}
-\caption{A violation with pass-by-reference}
-\label{fig:passing}
-\end{figure}
-
-To this point we have ignored function arguments, since our simplest model places them in
-registers. We now implement two kinds of protected
-sharing: passing by value on the stack, and passing by reference. First we will explore
-some examples at the same level of detail as our earlier ones, and then move on to
-extending our formalization.
-
-Example~\ref{fig:passingsimple} shows a sample trace of
-a program that passes a value on the stack, without an attack.
-We extend our call annotation with an additional parameter identifying a memory location to be
-passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
-to include \(\passed{\depth,\depth'}\), recording two depths. We note the
-current depth of the active function next to the stack pointer in the diagram, for reference.
-The argument {\tt a} has
-its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
-as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
-modified, and naturally this is permitted as well.
-
-The same annotation mechanism can be used to pass
-variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference.
-{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
-{\tt f} again passing a reference to {\tt y}.
-We would expect that the first call to {\tt f} might modify
-{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
-
-To avoid cluttering the picture at the assembly level, we revert to assuming that the
-argument {\tt a} is passed in a register, namely {\tt \%ra}.
-The question is how the local variables {\tt x} and {\tt y} themselves
-should be labelled in order for the policy to detect an integrity violation.
-At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
-return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
-a future call with no arguments passed by reference would also violate integrity if it wrote
-to either variable.
-
-When passing by reference, it is possible to repeatedly pass the same reference further
-and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices to track two depths for each passed
-cell: the original owner, and the highest depth it has been passed to. The policy considers
-it to be accessible anywhere between the two.
-
-Our model provides only limited support for handling fully first-class pointers into
-stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike the case for pass-by-reference, there is no easy way to model the intended
-sharing behavior of such pointers in terms of stack frame activation patterns.
-The best approximation we can make is to treat an address-taken variable in the
-same way as a passed argument, with the additional property
-that it remains accessible at \emph{all} stack depths above the caller.
-This prevents most violations of temporal safety, although stale pointers can still be
-used in certain corner cases.
-More seriously, callees are permitted to access the location without having
-been explicitly given a pointer to it.  Preventing the latter kind of abuse
-is the provence of a fine-grained memory safety policy, e.g. based on capabilities
-(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
-\emph{per se}.
-
-Next we introduce variable passing. We will need to add extra information to our call
-annotations to reflect that they may now pass variables on the stack, and that the locations
-of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the components
-that should be passed---when passing by value, this will typically be freshly allocated
-cells at the top of the frame, while passing by reference just gives the
-addresses of the appropriate variables.
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-And we extend the set of domains to include passed components,
-which are labeled with the range of depths of the caller that has passed them.
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\rts \in \listT (\target \subseteq \MACHS)) \]
-We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map when pushing it, where \(\depth\) is an integer depth.
-\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
-\begin{cases}
-  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
-  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
-  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
-                                 \text{ for any } \depth' \\
-  \domMap~\component & \text{else} \\
-\end{cases}\]
-
-And to pop we just unseal the components at depth \(\depth\):
-
-\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
-\begin{cases}
-  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
-  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
-                              \text{ for any } \depth' < \depth \\
-  \domMap~\component        & \text{else} \\
-\end{cases}\]
-
-Now we can inductively define the context update function; these again apply in the order
-presented. \bcp{Too much space between the second and third.}
-
-\judgmenttwo[Call]
-            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
-            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmenttwo[ ReturnFound]
-            {\(\mach \stepsto \mach' \in \target\)}
-            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3.5ex}
-\judgment[ Default]
-         {}
-         {\(\_,\context \stepsto_C \context\)}
-
-\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
-or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
-\(|\rts| > \depth'\).
-
-\definition
-Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
-A system enjoys {\em stepwise stack integrity with passing} if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
-for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
-
-\definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
-let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
-and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
-\emph{stepwise stack confidentiality} if,
-for all \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\).
-
-\subsection{Coroutines}
-
-Our final extension to the model is to add coroutines. We assume a very simple model with
-a fixed number of coroutines, each with its own stack.
-Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
-pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We add a second annotation indicating yields.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
-
-The first major difference in this model is the domains. Stack domains are now nested inside
-top-level domains, and associated with stack identifiers.
-Our context now consists of a domain map, a map from stack identifiers to return target stacks,
-an additional map of targets for yields, and a stack identifier of the active stack.
-Our yield target map \(\yts\) is a map from stack identifiers to targets.
-In a standard system an initial yield target map will map each stack identifier to the
-set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
-\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
-
-\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
-\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
-(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
-\target) \times (\stackid \in \STACKS)\]
-
-In our context update rules, calls and returns work the same as in the sharing model,
-with domain updates restricted to the top-level domain corresponding to the active stack,
-and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
-
-\judgmentthreebr[ Yield]
-                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
-                {\(\mach \stepsto \mach'\)}
-                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
-                {\(\stackid \not = \stackid'\)}
-                {\(\yts ~ \stackid'\)}
-                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
-                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
-
-This calls for some explanation. First, we determine which coroutine we are yielding to by
-checking which stack the stack pointer of the next state is pointing to. Then we can only yield
-if the next state is a valid yield target for that stack, and we update the yield target for
-the current stack, ensuring that when the system yields back we expect the current coroutine
-to continue from where it left off. So, it is possible for a coroutine to claim to yield
-but not actually change the active stack, with similar effects to when a callee fails
-to return properly. Now we can define both stack safety and coroutine safety under this model.
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
-{\em stepwise stack integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
-A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
 

### 2021-04-16 tex/main.tex
Factored out protected set from integrity props

--- 2021-04-16 tex/main.tex
@@
 which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special-purpose registers: the
+\(\REGS\), which is assumed to include two special purpose registers: the
 program counter {\PCname} and the stack pointer \(\SP\).
@@
 We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
-\rb{Could we use $\cdot$?}
 
@@
 The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
-and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
+and a predicate on states \(f ~ \mpcstate \in \MPCS\) and gives the prefix of
 \(\MPCT\) ending with the first element on which \(f\) holds.
@@
 
-\subsection{Parameters in Registers}
-
-To begin, we will focus on a simple stack model, with a single stack
+We will focus on a simple stack model, with a single stack
 that doesn't share between caller and callee: arguments and results are passed
-in registers.
-
-To accommodate different machine architectures and different calling
+in registers. To accommodate different machine architectures and different calling
 conventions, we parameterize over the structure of the stack, particularly
@@
 \definition
+Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\), be the
+set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
+for any depth \(\depth\).
+
+\definition
 A system enjoys {\em stepwise stack integrity} if, for any reachable state
 \(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
-\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename') ~ \component =
-\pi_\mach(\mpcstatename) ~ \component\).
+and any component \(\component \in \mathit{protected}(\domMap,\rts)\),
+\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
 
@@
 \definition
-For any \(\MPCT \in \MPCT^*\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \sealed{\_}\).
-Then if \(\last(\MPCT) = \mach',\pol',\context'\), let \(\components' = \Delta(\mach,\mach')\).
+For any \(\MPCT \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
+suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
+\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Find the set of components
+that changed during the call, \(\components = \mathit{protected}(\context) \cap \Delta(\mach,\mach')\).
 A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
-\(\nach \hookrightarrow \NPCT\), \(\MPCT \doteq \NPCT\).
-
-% \vspace{\belowdisplayskip}
+\(\mach' \approx_{\components'} \nach\) and its induced trace
+\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\MPCT' \doteq \NPCT\).
 
 In short, if sealed components are modified during the call, the caller may not
-access their values. ``No reads after dangerous writes.''
+access their values. 
 %
@@
 i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
-\rb{Tried to give a quick example without being too disruptive.
-  However, having it follow the slogan seems to rob it of part of its
-  punch.}
+Instead of forbidding such dangerous writes, we say,
+``No reads after dangerous writes.''
 
@@
 The ``observations-of'' operator, written \(\obsof\), takes a machine trace and
-creates a trace of observations corresponding to the observations of each step in
+gives the trace of observations of each step in
 the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
@@
 
-Then to take the observations from any particular trace, we start with a \(\tau\) observation:
-
+\begin{minipage}{.35\textwidth}
+  Then we start with a \(\tau\) observation:
+\end{minipage}
+\begin{minipage}{.6\textwidth}
 \judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
          {\(\obsof(\MPCT) = \obsT\)}
+\end{minipage}
 
@@
 For any call trace \(\MPCT \in \MPCT^*\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\)
-such that \(\domMap ~ \component = \sealed{\_}\).
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\).
 %
 Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
-let \(\components'\) be the set of components that differ
-between \(\mach\) and \(\mach'\). A system enjoys
-{\em observational integrity} if for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\), \((\mach,\pol',\context') \hookrightarrow \MPCT\), and
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT) \simeq \obsof(\NPCT)\).
+let \(\components\) be the set of protected components that changed during
+the call, \(\components = \mathit{protected}(\domMap,\rts) \cap \Delta(\mach, \mach')\),
+and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
+A system enjoys {\em observational integrity} if for any \(\nach\) such that
+\(\mach' \approx_{\components'} \nach\), and its induced trace
+\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
 
@@
 
+%% CHERI \rb{cite}
+
 \paragraph{Formal Stack Safety on Capability Machines}
@@
 programs. Although the required hardware support is readily available
-in capability machines like Cheri, this technique incurs significant
+in capability machines like CHERI, this technique incurs significant
 costs, because it requires the entire unused part of the stack to be

### 2021-04-16 tex/main.tex
Here's the payoff

--- 2021-04-16 tex/main.tex
@@
 \definition
-Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
-A system enjoys {\em stepwise stack integrity with passing} if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
-for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
+For each of the stepwise, deferred, and observational versions of
+stack integrity in the simple model, the equivalent property in the
+passing model substitutes the above definitions of \(\context\), initial contexts and update
+function, and \(\mathit{protected}\), and is otherwise identical.
 

### 2021-04-16 tex/main.tex
Cheri related work: lighter, a little shorter

@@
 programs. Although the required hardware support is readily available
-in capability machines like Cheri, this technique incurs significant
-costs, because it requires the entire unused part of the stack to be
+in capability machines like Cheri,
+%% this technique incurs significant
+%% costs, because
+it requires the entire unused part of the stack to be
 cleared whenever a security boundary is crossed. Their logical
@@
 side-effects (like console output or memory access traces)''
-and can be used to reason about individual programs. \bcp{Not very clear:}In the
-discussion, ``while [the authors] claim that [their] calling
+and can be used to reason about individual programs.
+\citeauthor{Skorstengaard+19b} acknowledge the difficulty of defining
+these concepts precisely:
+``while [they] claim that [their] calling
 convention enforces control-flow correctness, [they] do not prove a
@@
 
-StkTokens \citep{Skorstengaard+19} continues this line of work. Like
-the earlier paper, it aims to protect the stack by enforcing local state
-encapsulation and well-bracketed control flow. It does so by defining a new
+StkTokens \citep{Skorstengaard+19} continues this work by defining a more efficient
 calling convention that makes use of linear capabilities for stack and return
-pointers. The convention operates on a single shared stack and requires that
-protected components avoid compromising their own security by following certain
-simple rules---like not leaking their private capability seals. More
-precisely but still informally, local state encapsulation is defined as
+pointers.
+%% The convention operates on a single shared stack and requires that
+%% protected components avoid compromising their own security by following certain
+%% simple rules---like not leaking their private capability seals. More
+%% precisely but
+Still informally, local state encapsulation is defined as
 restricting accesses to the range of memory allocated to the current stack
@@
 which is proven fully abstract with respect to a more concrete capability
-machine that replaces those pseudo-instructions with their calling convention.
+machine.
+%% that replaces those pseudo-instructions with their calling convention.
 Their proof of full abstraction uses a standard notion of components, which
-import and export functions through their interfaces; their model of
-observations is limited to cotermination.
+import and export functions through their interfaces.
+%% ; their model of
+%% observations is limited to cotermination.
 \sna{Some of this seems like more detail than we need, and we need to
   compare ourselves to their list of criteria.}
+\rb{Trimmed this a bit. Are you referring to their list of attack
+  examples, or to something else?}
 

### 2021-04-16 tex/main.tex
polishing on 7

--- 2021-04-16 tex/main.tex
@@
 
-In particular, we wish to formulate properties around consecutive subtraces
-in which the context obeys a given predicate. This will commonly be used to extract
+\apt{Some words have been lost here.}
 individual calls from a trace by taking subtraces in which the stack is
@@
 by describing how we model knowledge about the program structure,
-then dive into the formal definitions of local state encapsulation and
+then dive into the formal definitions of local state encapsulation\apt{we don't want
+  to call it that here, do we?} and
 well-bracketed control flow, which together constitute stack safety\BCP{need
@@
 
-We will focus on a simple stack model, with a single stack
+\subsection{Parameters in Registers}
+
+To begin, we will focus on a simple stack model, with a single stack
 that doesn't share between caller and callee: arguments and results are passed
-in registers. To accommodate different machine architectures and different calling
+in registers.
+
+To accommodate different machine architectures and different calling
 conventions, we parameterize over the structure of the stack, particularly
@@
 \definition
-Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\), be the
-set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
-for any depth \(\depth\).
-
-\definition
 A system enjoys {\em stepwise stack integrity} if, for any reachable state
 \(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component \in \mathit{protected}(\domMap,\rts)\),
-\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
+and any component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
+\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename') ~ \component =
+\pi_\mach(\mpcstatename) ~ \component\).
 
@@
 In this section we explore how an existing enforcement mechanism, {\em
-  Depth Isolation}, from \citet{DBLP:conf/sp/RoesslerD18} implements
+  Depth Isolation}~\citep{DBLP:conf/sp/RoesslerD18}, implements
 the formal stack-safety property described in the previous
 section. (In fact, it enforces a slightly stronger, inductive version
-of that property which strengthens the local state encapsulation
-component by checking for integrity and confidentiality violations at
+of that property which checks for integrity and confidentiality violations at
 every step of the callee's execution trace.) We validate this claim
@@
 
-The micro-policy we present here to enforce the definition of stack
-safety developed in \cref{sec:lse} is a small variation
+The micro-policy we present here to illustrate enforcement of stack
+safety as defined in \cref{sec:lse} is a small variation
 on the {\em Depth Isolation} policy presented by
@@
 sequences that have permission to manage the stack, say during calls and returns.
-\leo{If we have the simple exposition so far, no need to mention the following sentence:}
-In the following, for ease of exposition, we will assume that no
-arguments are passed on the stack.
+%\leo{If we have the simple exposition so far, no need to mention the following sentence:}
+%In the following, for ease of exposition, we will assume that no
+%arguments are passed on the stack.
 %% \rb{Conceivably on alloc
@@
 There remains the question of how to ensure that the sequences are
-invoked at the right places. For a program to enjoy the stack-safety property
+invoked at the right places from the perspective of the stack-safety
+property. For a program to enjoy stack safety
 (with respect to a particular call map), all we need require is that
@@
 
-We validated the micro-policy of the previous section by
+We validated\apt{could we stay in present tense throughout?}  the micro-policy of the previous section by
 systematically testing that it satisfies eager stack integrity and
@@
 \footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
-\leo{translated from Haskell via hs-2-coq?}  and extended it with a
+\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since
+  the MIT people did it}}  and extended it with a
 runtime monitor implementing the stack-safety
@@
 To write such generators we build on the work of
-\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, where they
+\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
 introduced {\em generation by execution} to produce progams leading
@@
 the process until we generate or execute some target number of
-instructions, or if we reach a point where the machine can't step
+instructions, or we reach a point where the machine can't step
 anymore.
@@
 we sometimes allow the generation to relax those constraints,
-introducing potentially ill-formed flows: that causes our programs
+introducing potentially ill-formed flows: this causes our programs
 to failstop when executed in conjunction with the policy monitor,
@@
 In addition, we needed to extend this technique to handle the nested
-nature of our confidentiality property: rather than generating two
-initial machines that are variant of one another and letting them
-execute to test for noninterference, we need to generate a new variant
+nature of our confidentiality property: rather than just generating two
+initial machines that are variants of one another and letting them
+execute to test for noninterference, we generate a new variant
 {\em every time a call is made} and check confidentiality for the
@@
 to appear.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}), in our first try the blessed call
-sequence/policy combination did not initialize all locations for the
+{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
+sequence/policy combination\apt{??} did not initialize all locations for the
 newly allocated stack frame correctly, but left some of them as
 $\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured: in
-particular, the smallest counterexample requires calling a function,
-writing in the uninitialized location (which is always fine), calling
-another function (to seal what was written for integrity purposes),
-and then attempt to write in the same location again from the new
-callee, which is only allowed because the enforcement still thinks
-the location is unused, but violates the integrity of the caller's data.
-
+only if a rather complicated sequence of events occured. 
+The smallest counterexample requires calling a function {\tt f},
+which writes into an uninitialized location $l$ (which is always fine), and then
+calls a further function {\tt g} (which should have the effect of sealing $l$
+for integrity purposes). Then {\tt g} attempts to write to $l$; 
+this is allowed because the enforcement mechanism still has 
+$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
 \section{Stack Safety for Lazy Enforcement}
@@
 \definition
-For any \(\MPCT \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
-suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
-\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Find the set of components
-that changed during the call, \(\components = \mathit{protected}(\context) \cap \Delta(\mach,\mach')\).
+For any \(\MPCT \in \MPCT^*\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \sealed{\_}\).
+Then if \(\last(\MPCT) = \mach',\pol',\context'\), let \(\components' = \Delta(\mach,\mach')\).
  A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\) and its induced trace
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\MPCT' \doteq \NPCT\).
+\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
+\(\nach \hookrightarrow \NPCT\), \(\MPCT \doteq \NPCT\).
+
+% \vspace{\belowdisplayskip}
 
 In short, if sealed components are modified during the call, the caller may not
-access their values. 
+access their values. ``No reads after dangerous writes.''
 %
@@
 i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
-Instead of forbidding such dangerous writes, we say,
-``No reads after dangerous writes.''
+\rb{Tried to give a quick example without being too disruptive.
+  However, having it follow the slogan seems to rob it of part of its
+  punch.}
 
@@
 The ``observations-of'' operator, written \(\obsof\), takes a machine trace and
-gives the trace of observations of each step in
+creates a trace of observations corresponding to the observations of each step in
 the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
@@
 
-\begin{minipage}{.35\textwidth}
-  Then we start with a \(\tau\) observation:
-\end{minipage}
-\begin{minipage}{.6\textwidth}
+Then to take the observations from any particular trace, we start with a \(\tau\) observation:
+
 \judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
          {\(\obsof(\MPCT) = \obsT\)}
-\end{minipage}
 
@@
 For any call trace \(\MPCT \in \MPCT^*\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\).
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
+\(\components\) be the set of components \(\component\)
+such that \(\domMap ~ \component = \sealed{\_}\).
 %
 Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
-let \(\components\) be the set of protected components that changed during
-the call, \(\components = \mathit{protected}(\domMap,\rts) \cap \Delta(\mach, \mach')\),
-and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
-A system enjoys {\em observational integrity} if for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\), and its induced trace
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
+let \(\components'\) be the set of components that differ
+between \(\mach\) and \(\mach'\). A system enjoys
+{\em observational integrity} if for any \(\nach\) such that
+\(\mach' \approx_{\components'} \nach\), \((\mach,\pol',\context') \hookrightarrow \MPCT\), and
+\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT) \simeq \obsof(\NPCT)\).
 
@@
 \definition
-For each of the stepwise, deferred, and observational versions of
-stack integrity in the simple model, the equivalent property in the
-passing model substitutes the above definitions of \(\context\), initial contexts and update
-function, and \(\mathit{protected}\), and is otherwise identical.
+Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
+\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
+A system enjoys {\em stepwise stack integrity with passing} if for all
+such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
+for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)
 

### 2021-04-16 tex/main.tex

--- 2021-04-16 tex/main.tex
@@
 to appear.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}), in our first try the blessed call
-sequence/policy combination did not initialize all locations for the
+{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
+sequence %/policy combination\apt{??}
+did not initialize all locations for the
 newly allocated stack frame correctly, but left some of them as
 $\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured: in
-particular, the smallest counterexample requires calling a function,
-writing in the uninitialized location (which is always fine), calling
-another function (to seal what was written for integrity purposes),
-and then attempt to write in the same location again from the new
-callee, which is only allowed because the enforcement still thinks
-the location is unused, but violates the integrity of the caller's data.
-
+only if a rather complicated sequence of events occured. 
+The smallest counterexample requires calling a function {\tt f},
+which writes into an uninitialized location $l$ (which is always fine), and then
+calls a further function {\tt g} (which should have the effect of sealing $l$
+for integrity purposes). Then {\tt g} attempts to write to $l$; 
+this is allowed because the enforcement mechanism still has 
+$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
 \section{Stack Safety for Lazy Enforcement}

### 2021-04-16 tex/main.tex
modularized integrity

--- 2021-04-16 tex/main.tex
@@
 
-\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
+\definition Let the protected set of a context, \(\mathit{protected}_P(\domMap,\rts)\),
 be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
@@
 passing model substitutes the above definitions of \(\context\), initial contexts and update
-function, and \(\mathit{protected}\), and is otherwise identical.
+function, and \(\mathit{protected}_P\) for their simple equivalents, and is otherwise identical.
 
@@
 
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS) + \yieldmap\]
+\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
 \[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
@@
 In a standard system an initial yield target map will map each stack identifier to the
-set of all states with their stack pointer at the stack base. When yielding from state \(\mach\),
-\(\ycon ~ \mach \subseteq \MACHS\) tells us what yielding back looks like.
+set of all states with their stack pointer at the stack base, and include any restrictions
+on entry points. Just as a return target tells us that we have really returned, a yield
+target tells us that we have really yielded. When yielding from state \(\mach\),
+\(\ycon ~ \mach \in \MACHS\) gives us the yield target for coming back to the yielding
+coroutine.
 
@@
 
-\judgmentthreebr[ Yield]
-                {\(\codemap ~ (\mach ~ \PCname) = \yieldmap\)}
-                {\(\mach \stepsto \mach'\)}
-                {\(\stackid' = \stackof ~ (\mach'~ \rsp)\)}
-                {\(\stackid \not = \stackid'\)}
-                {\(\yts ~ \stackid'\)}
-                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
+\judgmentfour[ Yield]
+             {\(\mach \stepsto \mach'\) ~ ~}
+             {~ ~ \(\stackid' = \stackof ~ (\mach'~ \rsp) \not = \stackid\) ~ ~}
+             {~ ~ \(\yts ~ \stackid'\) ~ ~}
+             {~ ~ \(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
              {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
@@
 
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\). A system enjoys
-{\em stepwise stack integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
-
-\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
-\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rtm,\yts,\stackid\).
-Let \(\components\) be the set of components \(\component\) where
-\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid'\).
-A system enjoys {\em stepwise coroutine integrity} in the coroutine setting if for all
-such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).
+\definition Let the call-protected set of a context,
+\(\mathit{protected}_C(\domMap,\rtm,\yts,\stackid)\),
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\)
+or \(\domMap ~ \component = \instack{\stackid}{\passed{\depth,\depth'}}\) and \(|\rtm ~ \stackid| < \depth\) or
+\(|\rtm ~ \stackid| > \depth'\).
+
+\definition
+For each of the stepwise, deferred, and observational versions of
+stack integrity in the simple model, the equivalent property in the
+passing model substitutes the above definitions of \(\context\), initial contexts and update
+function, and \(\mathit{protected}_C\) for their simple equivalents, and is otherwise identical.
+
+\definition Let the yield-protected set of a context,
+\(\mathit{protected}_Y(\domMap,\rtm,\yts,\stackid)\),
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid`\).
+
+\definition
+We define {\em stepwise coroutine integrity} in the same way as stepwise stack integrity,
+substituting the above definitions of \(\context\), initial contexts and update
+function, and \(\mathit{protected}_Y\) for their equivalents. Likewise
+{\em deferred coroutine integrity} and {\em observational coroutine integrity} are defined
+in terms of their simple equivalents.
 
@@
 programs. Although the required hardware support is readily available
-in capability machines like Cheri,
-%% this technique incurs significant
-%% costs, because
-it requires the entire unused part of the stack to be
+in capability machines like Cheri, this technique incurs significant
+costs, because it requires the entire unused part of the stack to be
 cleared whenever a security boundary is crossed. Their logical
@@
 side-effects (like console output or memory access traces)''
-and can be used to reason about individual programs.
-\citeauthor{Skorstengaard+19b} acknowledge the difficulty of defining
-these concepts precisely:
-``while [they] claim that [their] calling
+and can be used to reason about individual programs. \bcp{Not very clear:}In the
+discussion, ``while [the authors] claim that [their] calling
 convention enforces control-flow correctness, [they] do not prove a
@@
 
-StkTokens \citep{Skorstengaard+19} continues this work by defining a more efficient
+StkTokens \citep{Skorstengaard+19} continues this line of work. Like
+the earlier paper, it aims to protect the stack by enforcing local state
+encapsulation and well-bracketed control flow. It does so by defining a new
 calling convention that makes use of linear capabilities for stack and return
-pointers.
-%% The convention operates on a single shared stack and requires that
-%% protected components avoid compromising their own security by following certain
-%% simple rules---like not leaking their private capability seals. More
-%% precisely but
-Still informally, local state encapsulation is defined as
+pointers. The convention operates on a single shared stack and requires that
+protected components avoid compromising their own security by following certain
+simple rules---like not leaking their private capability seals. More
+precisely but still informally, local state encapsulation is defined as
 restricting accesses to the range of memory allocated to the current stack
@@
 which is proven fully abstract with respect to a more concrete capability
-machine.
-%% that replaces those pseudo-instructions with their calling convention.
+machine that replaces those pseudo-instructions with their calling convention.
 Their proof of full abstraction uses a standard notion of components, which
-import and export functions through their interfaces.
-%% ; their model of
-%% observations is limited to cotermination.
+import and export functions through their interfaces; their model of
+observations is limited to cotermination.
 \sna{Some of this seems like more detail than we need, and we need to
   compare ourselves to their list of criteria.}
-\rb{Trimmed this a bit. Are you referring to their list of attack
-  examples, or to something else?}
 

### 2021-04-16 tex/main.tex
Deal with some comments in the intro

@@
   sentence belongs in the testing section?}\fi)
-%
+
 We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
@@
 stepwise properties.  On the other hand, their more efficient \emph{Lazy
-  Tagging and Clearing} policy enforces observational integrity, but testing
+  Tagging and Clearing} policy enforces observational integrity, but
+testing
 reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
@@
 propose a variant of {\em Lazy Tagging and Clearing} that \emph{does} enforce
-confidentiality (as far as our testing can tell), albeit at some performance
-cost.
+confidentiality (as far as our testing can tell\bcp{right, leo?}), albeit at
+some performance cost.
 %
@@
 
-Our first presentation of these properties assumes a single simple stack
-with no sharing between callers and callees: all parameters and return
-values are passed in registers.  Later, we show how to refine the properties to allow
-passing of scalar stack data, granting a callee access to the caller's
-data. (We do not consider more sophisticated forms of sharing, such as
-taking the address of a variable and treating it like a pointer.)
-\rb{But actually, this is discussed at some length is
-  \cref{sec:passing}, and it is the topic of \cref{fig:passing}.
-  Reaching that point after reading this disclaimer is confusing.}
-%
-Finally, we define integrity and confidentiality properties for a coroutine
+At first, for ease of exposition, our development assumes a single simple
+stack with no sharing between callers and callees: all parameters and return
+values are passed in registers.  Later, we show how to refine the properties
+to allow passing of scalar stack data, granting a callee access to the
+caller's data. (We do not consider more sophisticated forms of sharing, such
+as taking the address of a variable and treating it like a pointer.)
+\rb{But actually, this is discussed at some length is \cref{sec:passing},
+  and it is the topic of \cref{fig:passing}.  Reaching that point after
+  reading this disclaimer is confusing.}\bcp{So should we just remove that
+  sentence?}
+%
+Lastly, we define integrity and confidentiality properties for a coroutine
 system with a static layout, in which each coroutine's stack is constrained
@@
 security, combine these with an existing notion of well-bracketed
-control flow\bcp{need to mention this above} to formalize a {\em stepwise stack-safety}
+control flow to formalize a {\em stepwise stack-safety}
 property, and argue that this property captures the intuitive concept of

### 2021-04-16 tex/main.tex
CP work on section 2.  Breaking for dinner for a few minutes.

@@
 property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
-\bcp{mention that we extend the ICFP gen-by-exec with nested stuff...}
+of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement}).
+\bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
 \item We formalize a
@@
 
+\paragraph*{Machine model}
 Our properties are defined in terms of an abstract machine model that makes very few
@@
 
-Our stack-safety properties depend
-critically on the notions of ``call'' and ``return,'' which are not fully explicit
-in machine code%
-%
-\footnote{\bcp{Some possibly useful text: A major challenge in
-    characterizing stack safety formally is that, in a low-level machine
-    without built-in stacks, we are at the mercy\bcp{?} of
-    architectural\bcp{?}  choices that do not relate to the abstract notion
-    of stack safety. Some implementations may have only a stack pointer
-    while others add a frame pointer, in some systems return addresses and
-    data live in separate stacks, and so on. We abstract over these choices
-    as much as possible.  }}%
-%
----the same instruction opcode may be used in some places as part of
-a call or return sequence and in other places for other purposes.
-%
-So we assume we are given a machine language program {\em annotated} with the locations of
-instructions representing calls, and we determine when a return has occurred by tracking
-control flow and observing when we reach a {\em return target}. For a given state annotated
-as a call, another state fulfils its return target under a standard calling convention if
-it has the same stack pointer and the program counter is at the instruction following
-the call, but we abstract over different conventions.
+\bcp{What is a ``call''?}  One significant challenge in formulating our
+properties arises from the fact that we want to define stack safety at the
+machine-code level for a variety of ISAs, including RISCy ones.  In
+particular, there may be no single ``call'' instruction that does all the
+required manipulation of the \(\PCname\) and stack pointer. Instead, there
+might be just a simple branch instruction that stores the old \(\PCname\) in
+a register, with the rest of the call sequence performed by other
+instructions either before or after the jump.  Similarly, there may be no
+``return'' macro-instruction, but rather a ``return sequence'' involving an
+indirect branch at some point.  Thus, the same instruction opcode may be
+used in some places as part of a call or return sequence and in other places
+for other purposes.
+
+Since our stack-safety properties depend
+critically on being able to recognize when the machine is ``performing a
+call'' or a ``returning from a call'', we need some other way of
+
+%
+% \footnote{\bcp{Some possibly useful text: A major challenge in
+%     characterizing stack safety formally is that, in a low-level machine
+%     without built-in stacks, we are at the mercy\bcp{?} of
+%     architectural\bcp{?}  choices that do not relate to the abstract notion
+%     of stack safety. Some implementations may have only a stack pointer
+%     while others add a frame pointer, in some systems return addresses and
+%     data live in separate stacks, and so on. We abstract over these choices
+%     as much as possible.  }}%
+%
+%
+
+We assume we are given a machine language program {\em annotated} with the
+locations of instructions representing calls.  We declare that a call has
+occurred when the machine has just executed an instruction marked as a
+call. Conversely, we declare that a return
+has occurred when we see the state of the machine reach a {\em return
+  target}.
+\bcp{Still working from here:}
+For a given
+state annotated as a call, another state fulfils its return target under a
+standard calling convention if it has the same stack pointer and the program
+counter is at the instruction following the call, but we abstract over
+different conventions.
 %, returns, and function entry points\BCP{I
@@
 
+\paragraph*{Enforcement mechanisms}
 We model the base machine as a step function over machine states, and a machine

### 2021-04-16 tex/main.tex
Factored out "break condition"

--- 2021-04-16 tex/main.tex
@@
 
+\definition
+For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \_,\_,(\_,\rts)\), the
+{\em break condition} of \(\MPCT\) is a predicate on
+states \(\mathit{brk}(\MPCT) = \neg f_{|\rts|}\). It follows from our definitions
+that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
+
 \definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
@@
 \definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
-let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\).
-%That is, at the start of some call, we take every component in the stack.
+Let the secret set of a context, \(\mathit{secret}(\domMap,\rts)\), be the
+set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
+
+\definition
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
+let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
+\(\mach \approx_\components \nach\).
 A system enjoys \emph{stepwise stack confidentiality} if,
-for all \(\nach\) and \(\NPCT\) such that \(\mach \approx_\components \nach\)
-and \(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+for \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
 \(\MPCT \doteq \NPCT\) holds.
@@
 the resulting trace (starting in the same policy state) steps in lockstep with the
-original trace until both have reached a point whose depth is below the starting
-depth (this is a return).
+original trace until both have reached the break condition.
 
@@
 For any call trace \(\MPCT \in \MPCT^*\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\).
+let \(\head(\MPCT) = \mach,\pol,\context\).
 %
@@
 let \(\components\) be the set of protected components that changed during
-the call, \(\components = \mathit{protected}(\domMap,\rts) \cap \Delta(\mach, \mach')\),
+the call, \(\components = \mathit{protected}(\context) \cap \Delta(\mach, \mach')\),
 and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
@@
 secrets do not escape during a call {\em and} are not leaked during a return.
-This require us to be able to quantify over components that preserve our \(\diamond\)
-relation from earlier.
-
-\definition Let \(\components\) be a set of components and \(\mach,\mach'\) and \(\nach,\nach'\)
-be pairs of states. The {\em unsafe set} of \(\components\), written
-\(\bar{\Diamond}(\components,\mach,\mach',\nach,\nach')\), is the set of all components
+This require us to be able to quantify over components that fail to preserve
+our \(\diamond\) relation.
+
+\definition Let \(\mach,\mach'\) and \(\nach,\nach'\)
+be pairs of states. Their {\em unsafe set}, written
+\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all components
 \(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
@@
 For any call trace \(\MPCT \in \MPCT^*\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
-\(\components\) be the set of components \(\component\)
-such that \(\domMap ~ \component \not = \outside\).
+let \(\head(\MPCT) = \mach,\pol,\context\) and
+\(\components = \mathit{secret}(\context)\).
 %
@@
 \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth \),
+\(\nach \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
 \(\MPCT\) and \(\NPCT\), the following conditions all hold:
@@
   meaning that no secrets leak during the call
-\item \(f_\depth ~ \last(\MPCT) \leftrightarrow f_\depth ~
+\item \(\mathit{brk}(\MPCT) ~ \last(\MPCT) \leftrightarrow \mathit{brk}(\MPCT) ~
 \last(\NPCT)\)---if one call returns, so does the other
@@
   and \(\pi_\mach(\last(\NPCT)) = \nach'\).
-  Let \(\components' = \bar{\Diamond}(\components,\mach,\mach',\nach,\nach'\).
+  Let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\).
   And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).

### 2021-04-16 tex/main.tex
Finished reworking section 2

@@
 
-\bcp{What is a ``call''?}  One significant challenge in formulating our
+% \footnote{\bcp{Some possibly useful text: A major challenge in
+%     characterizing stack safety formally is that, in a low-level machine
+%     without built-in stacks, we are at the mercy\bcp{?} of
+%     architectural\bcp{?}  choices that do not relate to the abstract notion
+%     of stack safety. Some implementations may have only a stack pointer
+%     while others add a frame pointer, in some systems return addresses and
+%     data live in separate stacks, and so on. We abstract over these choices
+%     as much as possible.  }}%
+
+\paragraph*{What is a ``call''?}  One significant challenge in formulating our
 properties arises from the fact that we want to define stack safety at the
@@
 
-Since our stack-safety properties depend
-critically on being able to recognize when the machine is ``performing a
-call'' or a ``returning from a call'', we need some other way of
-
-%
-% \footnote{\bcp{Some possibly useful text: A major challenge in
-%     characterizing stack safety formally is that, in a low-level machine
-%     without built-in stacks, we are at the mercy\bcp{?} of
-%     architectural\bcp{?}  choices that do not relate to the abstract notion
-%     of stack safety. Some implementations may have only a stack pointer
-%     while others add a frame pointer, in some systems return addresses and
-%     data live in separate stacks, and so on. We abstract over these choices
-%     as much as possible.  }}%
-%
-%
-
-We assume we are given a machine language program {\em annotated} with the
-locations of instructions representing calls.  We declare that a call has
-occurred when the machine has just executed an instruction marked as a
-call. Conversely, we declare that a return
-has occurred when we see the state of the machine reach a {\em return
-  target}.
-\bcp{Still working from here:}
-For a given
-state annotated as a call, another state fulfils its return target under a
-standard calling convention if it has the same stack pointer and the program
-counter is at the instruction following the call, but we abstract over
-different conventions.
+Since our stack-safety properties are all about protecting ``callers'' from
+their ``callees,'' we need some clear-cut way to recognize when the program
+makes or returns from a call.  Accordingly, we assume we are given a machine
+language program {\em annotated} with the locations of instructions
+representing calls.  We declare that a call has occurred when the machine
+has just executed an instruction marked as a call.
+%
+These call annotations would typically come from a compiler together with
+the machine code, but we do to assume anything about their provenance.
+
+Returns, on the other hand, need not be annotated: we can tell when a callee
+has returned control to the caller when... we see control return to the
+caller!  For example, under a standard calling convention, we can see that a
+return has occurred occurred when state of the machine reaches a state where
+the stack pointer has the same value as it did at the point of some earlier
+call and the $\PCname$ points to the instruction following that call.  We
+abstract away from the details of specific calling conventions by assuming
+that we are given some way of telling when a given machine state is a
+{\em return target}\bcp{with respect to some set of pending calls?}.
+
 %, returns, and function entry points\BCP{I
@@
 %its own local state again and expected to find it unmolested.)}.
-Although call annotations would typically come from a compiler together with
-the machine code, we do not assume anything about their provenance.
-
-\paragraph*{Enforcement mechanisms}
-We model the base machine as a step function over machine states, and a machine
-enhanced with an enforcement mechanism (in the language of tag-based enforcement,
-a {\em policy}) as a partial step function over machine states extended with some
-auxiliary state, refining the base machine. When the policy-enhanced machine fails
-to step, this is a {\em failstop}, halting the program before it can perform an
+
+\paragraph*{Policies}
+We model the base machine's behavior as a step function over machine states,
+and a machine
+enhanced with an enforcement mechanism, or {\em policy} for short, as a
+partial step function over machine states extended with some
+auxiliary state, refining the base machine. When the policy-enhanced machine
+does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
 action that would violate the security property;
@@
 
-\apt{Suddenly we are talking about alternative enforcement mechanisms for the first time.
-  This surely needs \emph{some} foreshadowing in the abstract and
-  intro?}\bcp{Yes!}
+\apt{Suddenly we are talking about alternative enforcement mechanisms for
+  the first time.  This surely needs \emph{some} foreshadowing in the
+  abstract and intro?}\bcp{Yes!}
+%
 Dividing policy state from machine state helps us reason about

### 2021-04-16 tex/main.tex
Updated through the passing model

--- 2021-04-16 tex/main.tex
@@
               {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
          {}
-         {\(\_,\context \stepsto_C \context\)}
+         {\(\underscore,\context \stepsto_C \context\)}
 
@@
 \definition
-Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\), be the
+Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
 set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
@@
 \(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component \in \mathit{protected}(\domMap,\rts)\),
+and any component \(\component \in \prot(\domMap,\rts)\),
 \(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
@@
 \definition
-For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \_,\_,(\_,\rts)\), the
+For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \underscore,\underscore,(\underscore,\rts)\), the
 {\em break condition} of \(\MPCT\) is a predicate on
-states \(\mathit{brk}(\MPCT) = \neg f_{|\rts|}\). It follows from our definitions
+states \(\brk(\MPCT) = \neg f_{|\rts|}\). It follows from our definitions
 that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
@@
             {\(\machT \doteq \nachT\)}
-            {\(\notfinished{(\mach,\_,\_)}{\machT} \doteq \notfinished{(\nach,\_,\_)}{\nachT}\)}
+            {\(\notfinished{(\mach,\underscore,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore,\underscore)}{\nachT}\)}
 \end{minipage}
@@
 \definition
-Let the secret set of a context, \(\mathit{secret}(\domMap,\rts)\), be the
+Let the secret set of a context, \(\secret(\domMap,\rts)\), be the
 set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
@@
 For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
+let \(\components = \secret(\context)\) and \(\nach\) be any variant
 \(\mach \approx_\components \nach\).
 A system enjoys \emph{stepwise stack confidentiality} if,
-for \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
+for \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk(\MPCT)\),
 \(\MPCT \doteq \NPCT\) holds.
@@
 A system enjoys {\em return-to-caller} when, for every pair of reachable states
-\(\mpcstatename = (\_,\_,(\_,\rts))\) and
-\(\mpcstatename' = (\_,\_,(\_,\rts'))\) such that
+\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
+\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
 \(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
@@
 %\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
+%\(\pi_c(\mpcstatename) = (\underscore, \underscore::\rts)\) and \(\pi_m(\mpcstatename)\)
+%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore, \rts)\).
 
@@
 %\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
+%\(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid)\) and \(\pi_m(\mpcstatename)\)
+%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid')\)
 %where \(\stackid \not = \stackid'\).
@@
 \((\mach',\pol',\context') \hookrightarrow \MPCT'\). Find the set of components
-that changed during the call, \(\components = \mathit{protected}(\context) \cap \Delta(\mach,\mach')\).
+that changed during the call, \(\components = \prot(\context) \cap \Delta(\mach,\mach')\).
 A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
@@
 let \(\components\) be the set of protected components that changed during
-the call, \(\components = \mathit{protected}(\context) \cap \Delta(\mach, \mach')\),
+the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
 and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
@@
 let \(\head(\MPCT) = \mach,\pol,\context\) and
-\(\components = \mathit{secret}(\context)\).
+\(\components = \secret(\context)\).
 %
@@
 \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
+\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
 \(\MPCT\) and \(\NPCT\), the following conditions all hold:
@@
   meaning that no secrets leak during the call
-\item \(\mathit{brk}(\MPCT) ~ \last(\MPCT) \leftrightarrow \mathit{brk}(\MPCT) ~
+\item \(\brk(\MPCT) ~ \last(\MPCT) \leftrightarrow \brk(\MPCT) ~
 \last(\NPCT)\)---if one call returns, so does the other
@@
               {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
          {}
-         {\(\_,\context \stepsto_C \context\)}
-
-\definition Let the protected set of a context, \(\mathit{protected}_P(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
+         {\(\underscore,\context \stepsto_C \context\)}
+
+\definition Let the protected set of a context, \(\prot_P(\domMap,\rts)\),
+be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
 or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
@@
 passing model substitutes the above definitions of \(\context\), initial contexts and update
-function, and \(\mathit{protected}_P\) for their simple equivalents, and is otherwise identical.
+function, and \(\prot_P\) for their simple equivalents, and is otherwise identical.
+
+\definition Let the secret set of a context, \(\secret_P(\domMap,\rts)\),
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \sealed{\underscore}\) or
+\(\domMap ~ \component = \passed{\depth,\depth'}\)
+where \(\depth \leq |\rts| \leq \depth'\).
 
 \definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
-let \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
-and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
-\emph{stepwise stack confidentiality} if,
-for all \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
-\(\MPCT \doteq \NPCT\).
+For both the stepwise and observational versions of stack confidentiality in
+the original model, the equivalent property in the passing model is the same after
+substituting the above definitions of \(\context\), initial contexts and update
+function, and \(\prot_P\) for their simple equivalents.
 
@@
 
-\definition Let the call-protected set of a context,
-\(\mathit{protected}_C(\domMap,\rtm,\yts,\stackid)\),
+\definition
+Let the call-protected set of a context, \(\prot_C(\domMap,\rtm,\yts,\stackid)\),
 be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\)
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\underscore}}\)
 or \(\domMap ~ \component = \instack{\stackid}{\passed{\depth,\depth'}}\) and \(|\rtm ~ \stackid| < \depth\) or
@@
 passing model substitutes the above definitions of \(\context\), initial contexts and update
-function, and \(\mathit{protected}_C\) for their simple equivalents, and is otherwise identical.
-
-\definition Let the yield-protected set of a context,
-\(\mathit{protected}_Y(\domMap,\rtm,\yts,\stackid)\),
+function, and \(\prot_C\) for their simple equivalents, and is otherwise identical.
+
+\definition
+Let the yield-protected set of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
 be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid`\).
+\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
 
@@
 substituting the above definitions of \(\context\), initial contexts and update
-function, and \(\mathit{protected}_Y\) for their equivalents. Likewise
+function, and \(\prot_Y\) for their equivalents. Likewise
 {\em deferred coroutine integrity} and {\em observational coroutine integrity} are defined
@@
 
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+\definition
+Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\underscore, \rtm, \underscore, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
 For any reachable segment \(\MPCT \langle f_\depth \rangle\),
@@
 \definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+\(f_\depth ~ \mach ~ (\underscore, \rtm, \underscore, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
 For any reachable segment \(\MPCT \langle f_\depth \rangle\),
@@
 \(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\_}\).
+\(\domMap ~ \component = \instack{\stackid}{\underscore}\).
 Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and

### 2021-04-16 tex/main.tex
questions and polishing on 6

--- 2021-04-16 tex/main.tex
@@
 in fact we can learn something about the initial state by observing those results, violating
-confidentiality.
+confidentiality. \apt{Add some motivation for why we chose to view confidentiality in this way.}
 
@@
 
-\section{Machines and Traces}
-\label{sec:prelim}
-
-We now define our machine model precisely. To make our definition of stack safety
-as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (software- or hardware-enforced) policy monitor.
-In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
-policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
-
-\subsection{Values and States}
-
-The building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively.
-%
-Our machine states are composed of {\em components} \(\component\),
-which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special-purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-A {\em machine state} is a map from components to values.
-%
-    \[\component \in \COMPONENTS ::= \WORDS + \REGS  \hspace{1in}
-    \mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-%
-We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-
-We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
-A {\em policy} consists of a set of policy
-states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. The policy also provides a
-well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
-initial states that are properly configured.
-%
-A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-\Cref{sec:enforcement}.
-
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
-               \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol
-               \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-Note that the base machine has a total step function, while the policy-enhanced machine
-refines it to a partial function. The extra policy state never changes behavior,
-but it can cause
-premature termination, representing failstop behavior.
-
-\subsection{Contexts}
-
-Separate from the machine state, we keep track of additional context
-information representing the history of the run. This has no influence on execution, it merely
-describes information relevant to the particular property. Different information is needed
-for different versions of the model, so the definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
-an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
-\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
-\(\codemap \in \WORDS \rightharpoonup \ANNS\)
-that relates each code address to its annotation, if any.
-We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).
-
-\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
-              {\(\mach, \context \constep \context'\)}
-              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
-                \mpcstate{\mach'}{\pol'}{\context'}\)}
-
-\subsection{Machine Traces}
-\label{sec:traces}
-
-A {\em machine trace} is a nonempty, finite or infinite sequence of elements
-of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
-\rb{Could we use $\cdot$?}
-
-The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
-coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\)
-
-\begin{center}
-\begin{minipage}{.4\textwidth}
-\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
-            {\(\mpcstatename' \hookrightarrow \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
-\end{minipage}
-\begin{minipage}{.4\textwidth}
-\judgment%[Default]
-         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
-         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
-\end{minipage}
-\end{center}
-%
-
-We project out the machine state of an MPC-state with \(\pi_\mach\), the
-policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
-We take the first element of a trace with \(\head(\MPCT)\), which is a total
-function since traces are non-empty, and the final element (if one exists) with
-\(\last(\MPCT)\), which is partial.
-
-\paragraph*{Until}
-The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
-and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first element on which \(f\) holds.
-
-\begin{center}
-  \begin{minipage}{.3\textwidth}
-    \judgment{\(f ~ \mpcstatename\)}
-             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
-  \end{minipage}
-%
-  \begin{minipage}{.3\textwidth}
-    \judgment{}
-             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
-%
-  \end{minipage}
-  \begin{minipage}{.3\textwidth}
-    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
-                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
-  \end{minipage}
-\end{center}
-%
-If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate always holds, then
-\(\PUT\) is the identity function.
-
-We will frequently take a prefix of the trace from an initial state up
-through the first state where some condition holds on the machine state.
-This can easily be implemented using \(\PUT\), and we provide special notation
-for convenience.
-%
-If \(f\) is a predicate on machine states, we define \(\mpcstatename
-\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)''):
-%
-\begin{center}
-\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
-            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-\end{center}
-
-\paragraph*{Context Segments}
-
-In particular, we wish to formulate properties around consecutive subtraces
-in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth. If \(f\) is still a predicate on states, then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
-\bcp{Throughout, we could save space and improve beauty by removing some of the
-vertical space between inference rules.}
-
-\begin{center}
-\begin{minipage}{.38\textwidth}
-  \judgmenttwo[ Now]
-              {\(f ~ \mpcstatename\)}
-              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
-\end{minipage}
-%
-\begin{minipage}{.58\textwidth}
-  \judgmentthree[ Skip]
-                {\(f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{minipage}
-%
-  \judgmentthree[ Search]
-                {\(\neg f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{center}
-
-\paragraph*{Reachability}
-
-We often want to quantify over all traces from any initial state so that our
-properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
-where \((\mach,\pol) \in \WF\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
-such that from some similarly well-formed initial state \(\mpcstatename\),
-\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
-
-\section{Stack Safety, Formally}
-\label{sec:lse}
-
-We are finally ready for our definition of stack safety. We begin
-by describing how we model knowledge about the program structure,
-then dive into the formal definitions of local state encapsulation\apt{we don't want
-  to call it that here, do we?} and
-well-bracketed control flow, which together constitute stack safety\BCP{need
-  to put back WBCF!}.
-
-We will focus on a simple stack model, with a single stack
-that doesn't share between caller and callee: arguments and results are passed
-in registers. To accommodate different machine architectures and different calling
-conventions, we parameterize over the structure of the stack, particularly
-the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
-relation between machine states and the addresses that should be sealed.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
-which relates call states to the states that can be considered valid returns to that call.
-The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
-and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-
-Each system component is assigned to a {\em domain}, which may be {\em Outside}
-of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller registers a {\em target}
-condition---a predicate on machine states that must hold for it to consider itself active.
-A context is a pair of a ``domain map'' from components to domains, and a
-``return target list,'' which is a stack of targets, one for each caller awaiting
-return.  \ifaftersubmission\bcp{spacing is funny}\fi
-%
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-  \times (\listT (\target)) \]
-%
-The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
-the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
-
-The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
-A call annotation tells us to push the return target onto the return stack and
-seal all components with the depth of the previous return stack. If we aren't making a call,
-we identify a return by looking ahead and checking if we're about to step to a state
-that matches the topmost return target. If neither is true, the context remains unchanged.
-Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
-
-\judgmentthree[ Call]
-              {\(\codemap ~ (\mach ~ \PCname)\)}
-              {\(\components = \sealcon ~ \mach\)}
-              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ ReturnFound]
-              {\(\mach \stepsto \mach' \in \target\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3.5ex}
-\judgment[ Default]
-         {}
-         {\(\_,\context \stepsto_C \context\)}
-
-When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We now proceed to state our properties.
-The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
-quantified over each step of execution from any initial state. The criterion for integrity
-is simple: sealed components do not change.
-
-\definition
-Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\), be the
-set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
-for any depth \(\depth\).
-
-\definition
-A system enjoys {\em stepwise stack integrity} if, for any reachable state
-\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component \in \mathit{protected}(\domMap,\rts)\),
-\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
-
-Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
-nested: we will quantify over subtraces that represent whole calls. For each such
-subtrace, we will quantify over {\em variants} of its initial state. Then we require
-that the original, {\em primary} traces move in {\em lockstep} with the variant.
-
-\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
-\[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
-\end{aligned}\]
-
-\definition
-For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \_,\_,(\_,\rts)\), the
-{\em break condition} of \(\MPCT\) is a predicate on
-states \(\mathit{brk}(\MPCT) = \neg f_{|\rts|}\). It follows from our definitions
-that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
-
-\definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
-is the set of components \(\component\)
-such that \(\mach ~ \component \not = \mach' ~ \component\).
-
-\definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
- {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all components \(\component\) in
-\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
-\(\mach' ~ \component = \nach' ~ \component\).
-
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep},
-written \(\machT \doteq \nachT\),
-if their adjacent pairs of states change together.  We define this as a coinductive relation:
-
-\begin{minipage}{.3\textwidth}
-\judgment{}
-         {\(\mpcstatename \doteq \npcstatename\)}
-\end{minipage}
-\begin{minipage}{.6\textwidth}
-\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
-            {\(\machT \doteq \nachT\)}
-            {\(\notfinished{(\mach,\_,\_)}{\machT} \doteq \notfinished{(\nach,\_,\_)}{\nachT}\)}
-\end{minipage}
-
-\definition
-Let the secret set of a context, \(\mathit{secret}(\domMap,\rts)\), be the
-set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
-
-\definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
-\(\mach \approx_\components \nach\).
-A system enjoys \emph{stepwise stack confidentiality} if,
-for \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
-\(\MPCT \doteq \NPCT\) holds.
-
-%\medskip
-
-In other words, stepwise stack confidentiality states that if we vary every component of the stack,
-the resulting trace (starting in the same policy state) steps in lockstep with the
-original trace until both have reached the break condition.
-
-Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
-of the components in \(\components\), and in addition their subtraces are also in
-\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
-will contain these calls and returns as well, so its return stack pushes and pops
-in lockstep until it breaks at the final return.
-
-\subsection{Control Flow and Return-to-Caller}
-\label{sec:RTC}
-
-\BCP{We agreed that most or all of this belongs in Section 4...}Together,
+\paragraph{Well-bracketed control flow.} Together,
 integrity and confidentiality protect a caller's data until its callee
@@
 condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $A$.
+only return to $A$.  This is what \citet{Skorstengaard+19b} describe
+as \emph{well-bracketed control flow (WBCF)}.
+
+\section{Machines and Traces}
+\label{sec:prelim}
+
+We now define our machine model precisely. To make our definition of stack safety
+as generic as possible, this section proposes an abstract interface to a
+machine model plus some kind of (software- or hardware-enforced) policy monitor.
+In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
+policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
+
+\subsection{Values and States}
+
+The building blocks of the machine are {\em values} and {\em addresses}.
+Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
+\(\addr\), respectively.
+%
+Our machine states are composed of {\em components} \(\component\),
+which are either addresses or register names (\(\reg\)) drawn from some set
+\(\REGS\), which is assumed to include two special-purpose registers: the
+program counter {\PCname} and the stack pointer \(\SP\).
+A {\em machine state} is a map from components to values.
+%
+    \[\component \in \COMPONENTS ::= \WORDS + \REGS  \hspace{1in}
+    \mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
+%
+We have a total step function between machine states, written \(\mach \stepsto \mach'\).
+
+We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
+A {\em policy} consists of a set of policy
+states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
+\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
+it is undefined on input configurations that correspond to a policy fault. The policy
+can see the machine state, but cannot modify it. The policy also provides a
+well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
+initial states that are properly configured.
+%
+A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
+\Cref{sec:enforcement}.
+
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
+               \polstep \pol'\)}
+            {\(\mpstate{\mach}{\pol} \stepstopol
+               \mpstate{\mach'}{\pol'}\)}
+
+\noindent
+Note that the base machine has a total step function, while the policy-enhanced machine
+refines it to a partial function. The extra policy state never changes behavior,
+but it can cause
+premature termination, representing failstop behavior.
+
+\subsection{Contexts}
+
+Separate from the machine state, we keep track of additional context
+information representing the history of the run. This has no influence on execution, it merely
+describes information relevant to the particular property. Different information is needed
+for different versions of the model, so the definitions are parameterized
+over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
+\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
+\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
+\(\codemap \in \WORDS \rightharpoonup \ANNS\)
+that relates each code address to its annotation, if any.
+We lift contexts into an MP trace in much the same way as we lifted policy states to get
+a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
+\CONTEXTS\).
+
+\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
+              {\(\mach, \context \constep \context'\)}
+              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
+                \mpcstate{\mach'}{\pol'}{\context'}\)}
+
+\subsection{Machine Traces}
+\label{sec:traces}
+
+A {\em machine trace} is a nonempty, finite or infinite sequence of elements
+of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
+We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
+\rb{Could we use $\cdot$?}
+
+The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
+coinductively relates an initial state with the trace of states
+produced by repeated application of \(\stepstocon\)
+
+\begin{center}
+\begin{minipage}{.4\textwidth}
+\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
+            {\(\mpcstatename' \hookrightarrow \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
+\end{minipage}
+\begin{minipage}{.4\textwidth}
+\judgment%[Default]
+         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
+         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
+\end{minipage}
+\end{center}
+%
+
+We project out the machine state of an MPC-state with \(\pi_\mach\), the
+policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
+We take the first element of a trace with \(\head(\MPCT)\), which is a total
+function since traces are non-empty, and the final element (if one exists) with
+\(\last(\MPCT)\), which is partial.
+
+\paragraph*{Until}
+The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
+and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
+\(\MPCT\) ending with the first element on which \(f\) holds.
+
+\begin{center}
+  \begin{minipage}{.3\textwidth}
+    \judgment{\(f ~ \mpcstatename\)}
+             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
+  \end{minipage}
+%
+  \begin{minipage}{.3\textwidth}
+    \judgment{}
+             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
+%
+  \end{minipage}
+  \begin{minipage}{.3\textwidth}
+    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
+                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
+  \end{minipage}
+\end{center}
+%
+If the resulting trace is a strict prefix of the parameter, then the
+predicate must hold on its final element. If the predicate always holds, then
+\(\PUT\) is the identity function.
+
+We will frequently take a prefix of the trace from an initial state up
+through the first state where some condition holds on the machine state.
+This can easily be implemented using \(\PUT\), and we provide special notation
+for convenience.
+%
+If \(f\) is a predicate on machine states, we define \(\mpcstatename
+\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
+running \(\mpcstatename\) up to \(f\)''):
+%
+\begin{center}
+\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
+            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
+\end{center}
+
+\paragraph*{Context Segments}
+
+In particular, we wish to formulate properties around consecutive subtraces
+in which the context obeys a given predicate. This will commonly be used to extract
+individual calls from a trace by taking subtraces in which the stack is
+at or below a given depth. If \(f\) is still a predicate on states, then
+we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
+``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
+\bcp{Throughout, we could save space and improve beauty by removing some of the
+vertical space between inference rules.}
+
+\begin{center}
+\begin{minipage}{.38\textwidth}
+  \judgmenttwo[ Now]
+              {\(f ~ \mpcstatename\)}
+              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
+              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
+\end{minipage}
+%
+\begin{minipage}{.58\textwidth}
+  \judgmentthree[ Skip]
+                {\(f ~ \mpcstatename\)}
+                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
+\end{minipage}
+%
+  \judgmentthree[ Search]
+                {\(\neg f ~ \mpcstatename\)}
+                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
+\end{center}
+
+\paragraph*{Reachability}
+
+We often want to quantify over all traces from any initial state so that our
+properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
+is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
+where \((\mach,\pol) \in \WF\), \(\mpcstatename \hookrightarrow \MPCT\)
+and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
+such that from some similarly well-formed initial state \(\mpcstatename\),
+\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
+
+\section{Stack Safety, Formally}
+\label{sec:lse}
+
+We are finally ready for our definition of stack safety. We begin
+by describing how we model knowledge about the program structure,
+then dive into the formal definitions of local state encapsulation\apt{we don't want
+  to call it that here, do we?} and
+well-bracketed control flow, which together constitute stack safety.
+
+We will focus on a simple stack model, with a single stack
+that doesn't share between caller and callee: arguments and results are passed
+in registers. To accommodate different machine architectures and different calling
+conventions, we parameterize over the structure of the stack, particularly
+the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
+relation between machine states and the addresses that should be sealed.
+If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
+which addresses must be protected until the return. Similarly, we abstract over what
+it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
+which relates call states to the states that can be considered valid returns to that call.
+The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
+and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
+
+Each system component is assigned to a {\em domain}, which may be {\em Outside}
+of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
+at a particular depth. Each caller registers a {\em target}
+condition---a predicate on machine states that must hold for it to consider itself active.
+A context is a pair of a ``domain map'' from components to domains, and a
+``return target list,'' which is a stack of targets, one for each caller awaiting
+return.  \ifaftersubmission\bcp{spacing is funny}\fi
+%
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
+\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
+  \times (\listT (\target)) \]
+%
+The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
+the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
+
+The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
+A call annotation tells us to push the return target onto the return stack and
+seal all components with the depth of the previous return stack. If we aren't making a call,
+we identify a return by looking ahead and checking if we're about to step to a state
+that matches the topmost return target. If neither is true, the context remains unchanged.
+Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
+
+\judgmentthree[ Call]
+              {\(\codemap ~ (\mach ~ \PCname)\)}
+              {\(\components = \sealcon ~ \mach\)}
+              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
+                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\vspace*{-1ex}
+\judgmentthree[ ReturnFound]
+              {\(\mach \stepsto \mach' \in \target\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\vspace*{-1ex}
+\judgmentthree[ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\vspace*{-3.5ex}
+\judgment[ Default]
+         {}
+         {\(\_,\context \stepsto_C \context\)}
+
+When we build machine traces with this step function, we always know how deep in the
+call stack we are and which components are sealed. We now proceed to state our properties.
+The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
+quantified over each step of execution from any initial state. The criterion for integrity
+is simple: sealed components do not change.
 
 \definition
-A system enjoys {\em return-to-caller} when, for every pair of reachable states
+Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\), be the
+set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
+for some depth \(\depth\).
+
+\definition
+A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
+\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
+and any component \(\component \in \mathit{protected}(\domMap,\rts)\),
+\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
+\apt{but now $dm$ and $rts$ are unbound}.
+
+Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
+nested: we will quantify over subtraces that represent whole calls. For each such
+subtrace, we will quantify over {\em variants} of its initial state. Then we require
+that the original, {\em primary} traces move in {\em lockstep} with the variant.
+
+\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
+\[\begin{aligned}
+\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
+\text{ is a reachable segment}\} \hspace{1em} &
+\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
+\end{aligned}\]
+
+\definition
+For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \_,\_,(\_,\rts)\), the
+{\em break condition} of \(\MPCT\) is a predicate on
+states defined as \(\mathit{brk}(\MPCT) = \neg f_{|\rts|}(\MPCT)\). It follows from our definitions
+that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+written \(\mach \approx_\components \nach\), if, for
+all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
+
+\definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
+is the set of components \(\component\)
+such that \(\mach ~ \component \not = \mach' ~ \component\).
+
+\definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
+ {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all components \(\component\) in
+\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
+\(\mach' ~ \component = \nach' ~ \component\).
+
+\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep},
+written \(\machT \doteq \nachT\),
+if their adjacent pairs of states change together.  We define this as a coinductive relation:
+
+\begin{minipage}{.3\textwidth}
+\judgment{}
+         {\(\mpcstatename \doteq \npcstatename\)}
+\end{minipage}
+\begin{minipage}{.6\textwidth}
+\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
+            {\(\machT \doteq \nachT\)}
+            {\(\notfinished{(\mach,\_,\_)}{\machT} \doteq \notfinished{(\nach,\_,\_)}{\nachT}\)}
+\end{minipage}
+
+\definition
+Let the secret set of a context, \(\mathit{secret}(\domMap,\rts)\), be the
+set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
+
+\definition
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
+let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
+such that \(\mach \approx_\components \nach\).
+A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
+for all \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
+\(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
+universally quantified, and probably also \(\MPCT\) (?) }
+
+%\medskip
+
+In other words, stepwise stack confidentiality states that if we vary every component of the stack,
+the resulting trace (starting in the same policy state) steps in lockstep with the
+original trace until both have reached the break condition.
+%
+Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
+of the components in \(\components\), and in addition their subtraces are also in
+\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
+will contain these calls and returns as well, so its return stack pushes and pops
+in lockstep until it breaks at the final return.
+
+Finally, we formalize a version of WBCF, which capture the same informal idea
+as that of \citet{Skorstengaard+19b} using our notion of contexts.
+\definition
+A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
 \(\mpcstatename = (\_,\_,(\_,\rts))\) and
@@
 \(|\rts'| = |\rts|\)
-
-\bcp{we probably want to remark that this notion is pretty
-much the same as in StkTokens, just formalized a little differently.}
 

### 2021-04-16 tex/main.tex
Cutting the stack definitions from coroutines

--- 2021-04-16 tex/main.tex
@@
 
-\definition
-Let the call-protected set of a context, \(\prot_C(\domMap,\rtm,\yts,\stackid)\),
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\underscore}}\)
-or \(\domMap ~ \component = \instack{\stackid}{\passed{\depth,\depth'}}\) and \(|\rtm ~ \stackid| < \depth\) or
-\(|\rtm ~ \stackid| > \depth'\).
-
-\definition
-For each of the stepwise, deferred, and observational versions of
-stack integrity in the simple model, the equivalent property in the
-passing model substitutes the above definitions of \(\context\), initial contexts and update
-function, and \(\prot_C\) for their simple equivalents, and is otherwise identical.
+Stack integrity and confidentiality apply in the coroutine model, based on straightforward
+projections of each stack's domains, etc. and the machinery of the passing model. We will give
+the coroutine-specific properties.
+
+%\definition
+%The protected and secret sets of a context as used for calls,
+%\(\prot_C(\domMap,\rtm,\yts,\stackid)\) and \(\secret_C(\domMap,\rtm,\yts,\stackid)\),
+%are equivalent to their counterparts \(\prot_P\) and \(\secret_P\) applied to a domain map
+%that projects out the stack domains of components in stack \(\stackid\) and a
+%return target stack \(\rts = \rtm ~ \stackid\). Intuitively, within the active stack we protect
+%the same components we would have in the passing model.
+
+%\definition
+%For each of the stepwise, deferred, and observational versions of
+%stack integrity in the simple model, the equivalent property in the
+%passing model substitutes the above definitions of \(\context\), initial contexts and update
+%function, and \(\prot_C\) for their simple equivalents, and is otherwise identical.
+
+%\definition \(\MPCT^*\) is now defined as
+%\[\begin{aligned}
+%\bigcup_{\stackid \in \STACKS,\depth \in \mathbb{Z}} \hspace{-1em} \{\MPCT | \MPCT\langle f_\depth \rangle
+%\text{ is a reachable segment}\} \hspace{1em} &
+%\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rtm,\yts,\stackid')) \triangleq |\rtm ~ \stackid| \geq \depth \\
+%\end{aligned}\]
+%and \(\brk{\MPCT}\) similarly projects the depth of the active stack.
+
+%\definition
+%For either of the stepwise and observational versions of
+%stack confidentiality in the passing model, the equivalent property in the
+%coroutine model substitutes the above definitions.
 
@@
 in terms of their simple equivalents.
+
+\definition
+Let the call-secret set of a context, 
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\underscore}}\)
+or \(\domMap ~ \component = \instack{\stackid}{\passed{\depth,\depth'}}\) and
+\(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\).
+
+\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
+\[\begin{aligned}
+\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
+\text{ is a reachable segment}\} \hspace{1em} &
+\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
+\end{aligned}\]
+
+\definition
+Let the yield-secret set of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
 
@@
 
-Notably, we conjecture that while these systems may implement integrity and confidentiality,
-they fall just short of return-to-caller. This is because the standard way to protect returns
-in Cheri is to construct {\em return capability}, an object capability that must be invoked
-to simultaneously transfer control back to the caller and unseal a capability on an activation
-record. An ill-behaved intermediate caller could always pass its caller's return capability on
-to its callee instead of its own, ensuring that the callee will return past it. Instead,
-we believe that Cheri supports a slightly weaker control-flow property in which only callers
-that obey a dynamic reasonableness property are guaranteed not to be skipped.
-
 \paragraph*{Program Logic for Stacks}

### 2021-04-16 tex/main.tex

@@
 
-\section{Machines and Traces}
-\label{sec:prelim}
-
-We now define our machine model precisely. To make our definition of stack safety
-as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (software- or hardware-enforced) policy monitor.
-In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
-policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
-
-\subsection{Values and States}
-
-The building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively.
-%
-Our machine states are composed of {\em components} \(\component\),
-which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special-purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-A {\em machine state} is a map from components to values.
-%
-    \[\component \in \COMPONENTS ::= \WORDS + \REGS  \hspace{1in}
-    \mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-%
-We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-
-We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
-A {\em policy} consists of a set of policy
-states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. The policy also provides a
-well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
-initial states that are properly configured.
-%
-A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-\Cref{sec:enforcement}.
-
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
-               \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol
-               \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-Note that the base machine has a total step function, while the policy-enhanced machine
-refines it to a partial function. The extra policy state never changes behavior,
-but it can cause
-premature termination, representing failstop behavior.
-
-\subsection{Contexts}
-
-Separate from the machine state, we keep track of additional context
-information representing the history of the run. This has no influence on execution, it merely
-describes information relevant to the particular property. Different information is needed
-for different versions of the model, so the definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
-an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
-\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
-\(\codemap \in \WORDS \rightharpoonup \ANNS\)
-that relates each code address to its annotation, if any.
-We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).
-
-\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
-              {\(\mach, \context \constep \context'\)}
-              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
-                \mpcstate{\mach'}{\pol'}{\context'}\)}
-
-\subsection{Machine Traces}
-\label{sec:traces}
-
-A {\em machine trace} is a nonempty, finite or infinite sequence of elements
-of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
-\rb{Could we use $\cdot$?}
-
-The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
-coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\)
-
-\begin{center}
-\begin{minipage}{.4\textwidth}
-\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
-            {\(\mpcstatename' \hookrightarrow \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
-\end{minipage}
-\begin{minipage}{.4\textwidth}
-\judgment%[Default]
-         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
-         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
-\end{minipage}
-\end{center}
-%
-
-We project out the machine state of an MPC-state with \(\pi_\mach\), the
-policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
-We take the first element of a trace with \(\head(\MPCT)\), which is a total
-function since traces are non-empty, and the final element (if one exists) with
-\(\last(\MPCT)\), which is partial.
-
-\paragraph*{Until}
-The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
-and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first element on which \(f\) holds.
-
-\begin{center}
-  \begin{minipage}{.3\textwidth}
-    \judgment{\(f ~ \mpcstatename\)}
-             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
-  \end{minipage}
-%
-  \begin{minipage}{.3\textwidth}
-    \judgment{}
-             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
-%
-  \end{minipage}
-  \begin{minipage}{.3\textwidth}
-    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
-                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
-  \end{minipage}
-\end{center}
-%
-If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate always holds, then
-\(\PUT\) is the identity function.
-
-We will frequently take a prefix of the trace from an initial state up
-through the first state where some condition holds on the machine state.
-This can easily be implemented using \(\PUT\), and we provide special notation
-for convenience.
-%
-If \(f\) is a predicate on machine states, we define \(\mpcstatename
-\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)''):
-%
-\begin{center}
-\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
-            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-\end{center}
-
-\paragraph*{Context Segments}
-
-In particular, we wish to formulate properties around consecutive subtraces
-in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth. If \(f\) is still a predicate on states, then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
-\bcp{Throughout, we could save space and improve beauty by removing some of the
-vertical space between inference rules.}
-
-\begin{center}
-\begin{minipage}{.38\textwidth}
-  \judgmenttwo[ Now]
-              {\(f ~ \mpcstatename\)}
-              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
-\end{minipage}
-%
-\begin{minipage}{.58\textwidth}
-  \judgmentthree[ Skip]
-                {\(f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{minipage}
-%
-  \judgmentthree[ Search]
-                {\(\neg f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{center}
-
-\paragraph*{Reachability}
-
-We often want to quantify over all traces from any initial state so that our
-properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
-where \((\mach,\pol) \in \WF\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
-such that from some similarly well-formed initial state \(\mpcstatename\),
-\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
-
-\section{Stack Safety, Formally}
-\label{sec:lse}
-
-We are finally ready for our definition of stack safety. We begin
-by describing how we model knowledge about the program structure,
-then dive into the formal definitions of local state encapsulation\apt{we don't want
-  to call it that here, do we?} and
-well-bracketed control flow, which together constitute stack safety\BCP{need
-  to put back WBCF!}.
-
-We will focus on a simple stack model, with a single stack
-that doesn't share between caller and callee: arguments and results are passed
-in registers. To accommodate different machine architectures and different calling
-conventions, we parameterize over the structure of the stack, particularly
-the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
-relation between machine states and the addresses that should be sealed.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
-which relates call states to the states that can be considered valid returns to that call.
-The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
-and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-
-Each system component is assigned to a {\em domain}, which may be {\em Outside}
-of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller registers a {\em target}
-condition---a predicate on machine states that must hold for it to consider itself active.
-A context is a pair of a ``domain map'' from components to domains, and a
-``return target list,'' which is a stack of targets, one for each caller awaiting
-return.  \ifaftersubmission\bcp{spacing is funny}\fi
-%
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-  \times (\listT (\target)) \]
-%
-The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
-the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
-
-The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
-A call annotation tells us to push the return target onto the return stack and
-seal all components with the depth of the previous return stack. If we aren't making a call,
-we identify a return by looking ahead and checking if we're about to step to a state
-that matches the topmost return target. If neither is true, the context remains unchanged.
-Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
-
-\judgmentthree[ Call]
-              {\(\codemap ~ (\mach ~ \PCname)\)}
-              {\(\components = \sealcon ~ \mach\)}
-              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ ReturnFound]
-              {\(\mach \stepsto \mach' \in \target\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3.5ex}
-\judgment[ Default]
-         {}
-         {\(\underscore,\context \stepsto_C \context\)}
-
-When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We now proceed to state our properties.
-The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
-quantified over each step of execution from any initial state. The criterion for integrity
-is simple: sealed components do not change.
-
-\definition
-Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
-set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
-for any depth \(\depth\).
-
-\definition
-A system enjoys {\em stepwise stack integrity} if, for any reachable state
-\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component \in \prot(\domMap,\rts)\),
-\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
-
-Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
-nested: we will quantify over subtraces that represent whole calls. For each such
-subtrace, we will quantify over {\em variants} of its initial state. Then we require
-that the original, {\em primary} traces move in {\em lockstep} with the variant.
-
-\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
-\[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
-\end{aligned}\]
-
-\definition
-For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \underscore,\underscore,(\underscore,\rts)\), the
-{\em break condition} of \(\MPCT\) is a predicate on
-states \(\brk(\MPCT) = \neg f_{|\rts|}\). It follows from our definitions
-that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
-
-\definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
-is the set of components \(\component\)
-such that \(\mach ~ \component \not = \mach' ~ \component\).
-
-\definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
- {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all components \(\component\) in
-\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
-\(\mach' ~ \component = \nach' ~ \component\).
-
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep},
-written \(\machT \doteq \nachT\),
-if their adjacent pairs of states change together.  We define this as a coinductive relation:
-
-\begin{minipage}{.3\textwidth}
-\judgment{}
-         {\(\mpcstatename \doteq \npcstatename\)}
-\end{minipage}
-\begin{minipage}{.6\textwidth}
-\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
-            {\(\machT \doteq \nachT\)}
-            {\(\notfinished{(\mach,\underscore,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore,\underscore)}{\nachT}\)}
-\end{minipage}
-
-\definition
-Let the secret set of a context, \(\secret(\domMap,\rts)\), be the
-set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
-
-\definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-let \(\components = \secret(\context)\) and \(\nach\) be any variant
-\(\mach \approx_\components \nach\).
-A system enjoys \emph{stepwise stack confidentiality} if,
-for \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk(\MPCT)\),
-\(\MPCT \doteq \NPCT\) holds.
-
-%\medskip
-
-In other words, stepwise stack confidentiality states that if we vary every component of the stack,
-the resulting trace (starting in the same policy state) steps in lockstep with the
-original trace until both have reached the break condition.
-
-Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
-of the components in \(\components\), and in addition their subtraces are also in
-\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
-will contain these calls and returns as well, so its return stack pushes and pops
-in lockstep until it breaks at the final return.
-
-\subsection{Control Flow and Return-to-Caller}
-\label{sec:RTC}
-
-\BCP{We agreed that most or all of this belongs in Section 4...}Together,
+\paragraph{Well-bracketed control flow.} Together,
 integrity and confidentiality protect a caller's data until its callee
@@
 condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $A$.
+only return to $A$.  This is what \citet{Skorstengaard+19b} describe
+as \emph{well-bracketed control flow (WBCF)}.
+
+\section{Machines and Traces}
+\label{sec:prelim}
+
+We now define our machine model precisely. To make our definition of stack safety
+as generic as possible, this section proposes an abstract interface to a
+machine model plus some kind of (software- or hardware-enforced) policy monitor.
+In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
+policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
+
+\subsection{Values and States}
+
+The building blocks of the machine are {\em values} and {\em addresses}.
+Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
+\(\addr\), respectively.
+%
+Our machine states are composed of {\em components} \(\component\),
+which are either addresses or register names (\(\reg\)) drawn from some set
+\(\REGS\), which is assumed to include two special-purpose registers: the
+program counter {\PCname} and the stack pointer \(\SP\).
+A {\em machine state} is a map from components to values.
+%
+    \[\component \in \COMPONENTS ::= \WORDS + \REGS  \hspace{1in}
+    \mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
+%
+We have a total step function between machine states, written \(\mach \stepsto \mach'\).
+
+We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
+A {\em policy} consists of a set of policy
+states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
+\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
+it is undefined on input configurations that correspond to a policy fault. The policy
+can see the machine state, but cannot modify it. The policy also provides a
+well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
+initial states that are properly configured.
+%
+A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
+\Cref{sec:enforcement}.
+
+We lift the policy step function to operate on such pairs
+by combining it with the regular step function for machine states.
+
+\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
+               \polstep \pol'\)}
+            {\(\mpstate{\mach}{\pol} \stepstopol
+               \mpstate{\mach'}{\pol'}\)}
+
+\noindent
+Note that the base machine has a total step function, while the policy-enhanced machine
+refines it to a partial function. The extra policy state never changes behavior,
+but it can cause
+premature termination, representing failstop behavior.
+
+\subsection{Contexts}
+
+Separate from the machine state, we keep track of additional context
+information representing the history of the run. This has no influence on execution, it merely
+describes information relevant to the particular property. Different information is needed
+for different versions of the model, so the definitions are parameterized
+over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
+an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
+\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
+\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
+\(\codemap \in \WORDS \rightharpoonup \ANNS\)
+that relates each code address to its annotation, if any.
+We lift contexts into an MP trace in much the same way as we lifted policy states to get
+a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
+\CONTEXTS\).
+
+\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
+              {\(\mach, \context \constep \context'\)}
+              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
+                \mpcstate{\mach'}{\pol'}{\context'}\)}
+
+\subsection{Machine Traces}
+\label{sec:traces}
+
+A {\em machine trace} is a nonempty, finite or infinite sequence of elements
+of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
+We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
+\rb{Could we use $\cdot$?}
+
+The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
+coinductively relates an initial state with the trace of states
+produced by repeated application of \(\stepstocon\)
+
+\begin{center}
+\begin{minipage}{.4\textwidth}
+\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
+            {\(\mpcstatename' \hookrightarrow \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
+\end{minipage}
+\begin{minipage}{.4\textwidth}
+\judgment%[Default]
+         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
+         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
+\end{minipage}
+\end{center}
+%
+
+We project out the machine state of an MPC-state with \(\pi_\mach\), the
+policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
+We take the first element of a trace with \(\head(\MPCT)\), which is a total
+function since traces are non-empty, and the final element (if one exists) with
+\(\last(\MPCT)\), which is partial.
+
+\paragraph*{Until}
+The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
+and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
+\(\MPCT\) ending with the first element on which \(f\) holds.
+
+\begin{center}
+  \begin{minipage}{.3\textwidth}
+    \judgment{\(f ~ \mpcstatename\)}
+             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
+  \end{minipage}
+%
+  \begin{minipage}{.3\textwidth}
+    \judgment{}
+             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
+%
+  \end{minipage}
+  \begin{minipage}{.3\textwidth}
+    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
+                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
+  \end{minipage}
+\end{center}
+%
+If the resulting trace is a strict prefix of the parameter, then the
+predicate must hold on its final element. If the predicate always holds, then
+\(\PUT\) is the identity function.
+
+We will frequently take a prefix of the trace from an initial state up
+through the first state where some condition holds on the machine state.
+This can easily be implemented using \(\PUT\), and we provide special notation
+for convenience.
+%
+If \(f\) is a predicate on machine states, we define \(\mpcstatename
+\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
+running \(\mpcstatename\) up to \(f\)''):
+%
+\begin{center}
+\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
+            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
+            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
+\end{center}
+
+\paragraph*{Context Segments}
+
+In particular, we wish to formulate properties around consecutive subtraces
+in which the context obeys a given predicate. This will commonly be used to extract
+individual calls from a trace by taking subtraces in which the stack is
+at or below a given depth. If \(f\) is still a predicate on states, then
+we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
+``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
+\bcp{Throughout, we could save space and improve beauty by removing some of the
+vertical space between inference rules.}
+
+\begin{center}
+\begin{minipage}{.38\textwidth}
+  \judgmenttwo[ Now]
+              {\(f ~ \mpcstatename\)}
+              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
+              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
+\end{minipage}
+%
+\begin{minipage}{.58\textwidth}
+  \judgmentthree[ Skip]
+                {\(f ~ \mpcstatename\)}
+                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
+\end{minipage}
+%
+  \judgmentthree[ Search]
+                {\(\neg f ~ \mpcstatename\)}
+                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
+                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
+                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
+\end{center}
+
+\paragraph*{Reachability}
+
+We often want to quantify over all traces from any initial state so that our
+properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
+is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
+where \((\mach,\pol) \in \WF\), \(\mpcstatename \hookrightarrow \MPCT\)
+and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
+such that from some similarly well-formed initial state \(\mpcstatename\),
+\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
+
+\section{Stack Safety, Formally}
+\label{sec:lse}
+
+We are finally ready for our definition of stack safety. We begin
+by describing how we model knowledge about the program structure,
+then dive into the formal definitions of local state encapsulation\apt{we don't want
+  to call it that here, do we?} and
+well-bracketed control flow, which together constitute stack safety.
+
+We will focus on a simple stack model, with a single stack
+that doesn't share between caller and callee: arguments and results are passed
+in registers. To accommodate different machine architectures and different calling
+conventions, we parameterize over the structure of the stack, particularly
+the means by
+which a caller identifies the data that should be protected during a call. We term
+this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
+relation between machine states and the addresses that should be sealed.
+If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
+which addresses must be protected until the return. Similarly, we abstract over what
+it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
+which relates call states to the states that can be considered valid returns to that call.
+The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
+and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
+
+Each system component is assigned to a {\em domain}, which may be {\em Outside}
+of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
+at a particular depth. Each caller registers a {\em target}
+condition---a predicate on machine states that must hold for it to consider itself active.
+A context is a pair of a ``domain map'' from components to domains, and a
+``return target list,'' which is a stack of targets, one for each caller awaiting
+return.  \ifaftersubmission\bcp{spacing is funny}\fi
+%
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
+\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
+  \times (\listT (\target)) \]
+%
+The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
+the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
+
+The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
+A call annotation tells us to push the return target onto the return stack and
+seal all components with the depth of the previous return stack. If we aren't making a call,
+we identify a return by looking ahead and checking if we're about to step to a state
+that matches the topmost return target. If neither is true, the context remains unchanged.
+Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
+
+\judgmentthree[ Call]
+              {\(\codemap ~ (\mach ~ \PCname)\)}
+              {\(\components = \sealcon ~ \mach\)}
+              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
+                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\vspace*{-1ex}
+\judgmentthree[ ReturnFound]
+              {\(\mach \stepsto \mach' \in \target\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\vspace*{-1ex}
+\judgmentthree[ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
+              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
+              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\vspace*{-3.5ex}
+\judgment[ Default]
+         {}
+         {\(\underscore,\context \stepsto_C \context\)}
+
+When we build machine traces with this step function, we always know how deep in the
+call stack we are and which components are sealed. We now proceed to state our properties.
+The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
+quantified over each step of execution from any initial state. The criterion for integrity
+is simple: sealed components do not change.
 
 \definition
-A system enjoys {\em return-to-caller} when, for every pair of reachable states
+Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
+set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
+for some depth \(\depth\).
+
+\definition
+A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
+\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
+and any component \(\component \in \prot(\domMap,\rts)\),
+\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
+\apt{but now $dm$ and $rts$ are unbound}.
+
+Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
+nested: we will quantify over subtraces that represent whole calls. For each such
+subtrace, we will quantify over {\em variants} of its initial state. Then we require
+that the original, {\em primary} traces move in {\em lockstep} with the variant.
+
+\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
+\[\begin{aligned}
+\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
+\text{ is a reachable segment}\} \hspace{1em} &
+\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
+\end{aligned}\]
+
+\definition
+For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \underscore,\underscore,(\underscore,\rts)\), the
+{\em break condition} of \(\MPCT\) is a predicate on
+states defined as \(\brk(\MPCT) = \neg f_{|\rts|}(\MPCT)\). It follows from our definitions
+that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+written \(\mach \approx_\components \nach\), if, for
+all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
+
+\definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
+is the set of components \(\component\)
+such that \(\mach ~ \component \not = \mach' ~ \component\).
+
+\definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
+ {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all components \(\component\) in
+\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
+\(\mach' ~ \component = \nach' ~ \component\).
+
+\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep},
+written \(\machT \doteq \nachT\),
+if their adjacent pairs of states change together.  We define this as a coinductive relation:
+
+\begin{minipage}{.3\textwidth}
+\judgment{}
+         {\(\mpcstatename \doteq \npcstatename\)}
+\end{minipage}
+\begin{minipage}{.6\textwidth}
+\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
+            {\(\machT \doteq \nachT\)}
+            {\(\notfinished{(\mach,\underscore,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore,\underscore)}{\nachT}\)}
+\end{minipage}
+
+\definition
+Let the secret set of a context, \(\secret(\domMap,\rts)\), be the
+set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
+
+\definition
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
+let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
+such that \(\mach \approx_\components \nach\).
+A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
+for all \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk(\MPCT)\),
+\(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
+universally quantified, and probably also \(\MPCT\) (?) }
+
+%\medskip
+
+In other words, stepwise stack confidentiality states that if we vary every component of the stack,
+the resulting trace (starting in the same policy state) steps in lockstep with the
+original trace until both have reached the break condition.
+%
+Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
+of the components in \(\components\), and in addition their subtraces are also in
+\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
+will contain these calls and returns as well, so its return stack pushes and pops
+in lockstep until it breaks at the final return.
+
+Finally, we formalize a version of WBCF, which capture the same informal idea
+as that of \citet{Skorstengaard+19b} using our notion of contexts.
+\definition
+A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
 \(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and

### 2021-04-16 tex/main.tex
Cleaned up coroutines

@@
 
-\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
+\definition \(\MPCT^*\) is the \emph{call set}, defined as
 \[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
+\bigcup_{\depth \in \mathbb{Z}} \{\MPCT, \neg f_\depth | \MPCT\langle f_\depth \rangle
 \text{ is a reachable segment}\} \hspace{1em} &
@@
 
-\definition
-For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \underscore,\underscore,(\underscore,\rts)\), the
-{\em break condition} of \(\MPCT\) is a predicate on
-states defined as \(\brk(\MPCT) = \neg f_{|\rts|}(\MPCT)\). It follows from our definitions
-that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
+We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
+states that holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
 
@@
 \definition
-For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
-such that \(\mach \approx_\components \nach\).
+For any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
+let \(\components = \mathit{secret}(\context)\).
 A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for all \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk(\MPCT)\),
+for any \(\nach\) such that \(\mach \approx_\components \nach\) and the induced trace
+\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
 \(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
   universally quantified, and probably also \(\MPCT\) (?) }
+\sna{Better?}
 
@@
 \(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
-\(|\rts'| = |\rts|\)
+\(|\rts'| = |\rts|-1\)
 
@@
 \definition
-For any \(\MPCT \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
+For any \(\MPCT,\brk \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
 suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
@@
 \definition
-For any call trace \(\MPCT \in \MPCT^*\),
+For any \(\MPCT,\brk \in \MPCT^*\),
 let \(\head(\MPCT) = \mach,\pol,\context\).
@@
 \definition
-For any call trace \(\MPCT \in \MPCT^*\),
+For any \(\MPCT,\brk \in \MPCT^*\),
 let \(\head(\MPCT) = \mach,\pol,\context\) and
@@
   meaning that no secrets leak during the call
-\item \(\brk(\MPCT) ~ \last(\MPCT) \leftrightarrow \brk(\MPCT) ~
+\item \(\brk ~ \last(\MPCT) \leftrightarrow \brk ~
 \last(\NPCT)\)---if one call returns, so does the other
@@
 
+\definition \(\MPCT^\wedge\) is the \emph{yield set}, defined as
+\[\begin{aligned}
+\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{\MPCT, \neg f_\stackid | \MPCT\langle f_\stackid \rangle
+\text{ is a reachable segment}\} \hspace{1em} &
+\text{where } f_\stackid ~ (\underscore,\underscore,(\underscore,\underscore,\underscore,\stackid')) \triangleq \stackid \not = \stackid' \\
+\end{aligned}\]
+
+Here, we are capturing the entire span that the system spends {\em away} from a given
+coroutine.
+
 \definition
-Let the yield-protected set of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
+Let the protected set of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
+
+\definition
+Let the secret set of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
 be the set of components \(\component\) such that
@@
 {\em deferred coroutine integrity} and {\em observational coroutine integrity} are defined
-in terms of their simple equivalents.
+in terms of their simple equivalents using \(\MPCT^\wedge\) in place of \(\MPCT^*\).
 
 \definition
-Let the call-secret set of a context, 
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\sealed{\underscore}}\)
-or \(\domMap ~ \component = \instack{\stackid}{\passed{\depth,\depth'}}\) and
-\(|\rtm ~ \stackid| < \depth\) or \(|\rtm ~ \stackid| > \depth'\).
-
-\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
-\[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
-\end{aligned}\]
-
-\definition
-Let the yield-secret set of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
-
-\definition
-Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\underscore, \rtm, \underscore, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
-\(\instack{\stackid}{\passed{\depth,\depth'}}\)
-where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
-{\em stepwise stack confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
-
-\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
-\(f_\depth ~ \mach ~ (\underscore, \rtm, \underscore, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
-For any reachable segment \(\MPCT \langle f_\depth \rangle\),
-let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
-\(\components\) be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid}{\underscore}\).
-Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
-\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
-{\em stepwise coroutine confidentiality} in the coroutine setting if for all
-such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+We define {\em stepwise coroutine integrity} in the same way as stepwise stack
+integrity, substituting the above definitions. Likewise
+{\em observational coroutine confidentiality} is defined in terms of observational
+stack confidentiality.
 

### 2021-04-17 tex/main.tex
WBCF: C returns to B (not A) in A -> B -> C :-)

@@
 condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $A$.  This is what \citet{Skorstengaard+19b} describe
+only return to $B$.  This is what \citet{Skorstengaard+19b} describe
 as \emph{well-bracketed control flow (WBCF)}.

### 2021-04-17 tex/main.tex
Added a dot for cons
TIny fix

@@
 of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use juxtaposition of an element and a trace to represent ``cons.'' (\bcp{Ugh})
-\rb{Could we use $\cdot$?}
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces (ordinary lists will use
+``::'').
 
@@
             {\(\mpcstatename' \hookrightarrow \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \mpcstatename' \MPCT\)}%
+            {\(\mpcstatename \hookrightarrow \notfinished{\mpcstatename'}{\MPCT}\)}%
 \end{minipage}
@@
     \judgment{\(f ~ \mpcstatename\)}
-             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
+             {\(\PUT ~ f ~ \notfinished{\mpcstatename}{\MPCT} = \mpcstatename\)}
   \end{minipage}
@@
     \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
-                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
+                {\(\PUT ~ f ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\mpcstatename}{\MPCT'}\)}
   \end{minipage}
@@
 
-Finally, we formalize a version of WBCF, which capture the same informal idea
-as that of \citet{Skorstengaard+19b} using our notion of contexts.
+Finally, we formalize a version of WBCF, capturing the same informal idea
+as \citet{Skorstengaard+19b} using our notion of contexts.
 \definition
@@
                 {\(\mathit{obsfor} ~ \obs' ~ \MPCT = \obsT\)}
-                {\(\mathit{obsfor} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
+                {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}%
 \end{minipage}
@@
   \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\tau \obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
 \end{minipage}
@@
   \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obsT_1 \eqsim \tau \obsT_2\)}
+           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
 \end{minipage}
@@
   \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obs\obsT_1 \eqsim \obs\obsT_2\)}
+           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
 \end{minipage}

### 2021-04-17 tex/main.tex
comments and polishing on 6 and 8

--- 2021-04-16 tex/main.tex
@@
               {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
          {}
-         {\(\underscore,\context \stepsto_C \context\)}
+         {\(\_,\context \stepsto_C \context\)}
 
@@
 \definition
-Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
+Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\), be the
 set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
@@
 \(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component \in \prot(\domMap,\rts)\),
+and any component \(\component \in \mathit{protected}(\domMap,\rts)\),
 \(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
@@
 
-\definition \(\MPCT^*\) is the \emph{call set}, defined as
+\definition \(\MPCT^*\) is the set of all \emph{call traces}, defined as
 \[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{\MPCT, \neg f_\depth | \MPCT\langle f_\depth \rangle
+\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
 \text{ is a reachable segment}\} \hspace{1em} &
@@
 
-We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
-states that holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
+\definition
+For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \_,\_,(\_,\rts)\), the
+{\em break condition} of \(\MPCT\) is a predicate on
+states defined as \(\mathit{brk}(\MPCT) = \neg f_{|\rts|}(\MPCT)\). It follows from our definitions
+that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
 
@@
             {\(\machT \doteq \nachT\)}
-            {\(\notfinished{(\mach,\underscore,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore,\underscore)}{\nachT}\)}
+            {\(\notfinished{(\mach,\_,\_)}{\machT} \doteq \notfinished{(\nach,\_,\_)}{\nachT}\)}
 \end{minipage}
@@
 \definition
-Let the secret set of a context, \(\secret(\domMap,\rts)\), be the
+Let the secret set of a context, \(\mathit{secret}(\domMap,\rts)\), be the
 set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
@@
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-let \(\components = \mathit{secret}(\context)\).
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
+let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
+such that \(\mach \approx_\components \nach\).
 A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for any \(\nach\) such that \(\mach \approx_\components \nach\) and the induced trace
-\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
+for all \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
 \(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
 universally quantified, and probably also \(\MPCT\) (?) }
-\sna{Better?}
 
@@
 
-In other words, stepwise stack confidentiality states that if we vary every component of the stack,
-the resulting trace (starting in the same policy state) steps in lockstep with the
-original trace until both have reached the break condition.
+In other words, stepwise stack confidentiality states that if, at the start of the call,
+we vary all stack locations that are supposed to be secret from the callee
+(which in our simple setting is the entirety of the stack),
+the resulting trace remains in lockstep with the original trace
+until either both return together or one failstops. This implies that
+if the callee reads a secret location, it must failstop.\apt{right?}
 %
@@
 
-Finally, we formalize a version of WBCF, which capture the same informal idea
-as that of \citet{Skorstengaard+19b} using our notion of contexts.
+Finally, we formalize a version of WBCF, which captures the same informal idea
+as that of \citet{Skorstengaard+19b} using our context machinery.
 \definition
 A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
-\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
-\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
-\(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
-\(|\rts'| = |\rts|-1\)
+\(\mpcstatename = (\_,\_,(\_,\rts))\) and
+\(\mpcstatename' = (\_,\_,(\_,\rts'))\) such that
+\(\mpcstatename \stepsto \mpcstatename'\)\apt{doesn't reachability of $\mpcstatename$ imply reachability of $\mpcstatename'$?}, if \(|\rts'| < |\rts|\), then
+\(|\rts'| = |\rts|\)
 
@@
 %\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\underscore, \underscore::\rts)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore, \rts)\).
+%\(\pi_c(\mpcstatename) = (\_, \_::\rts)\) and \(\pi_m(\mpcstatename)\)
+%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_, \rts)\).
 
@@
 %\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid')\)
+%\(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid)\) and \(\pi_m(\mpcstatename)\)
+%is annotated as a return, then \(\pi_c(\mpcstatename) = (\_,\_,\_,\stackid')\)
 %where \(\stackid \not = \stackid'\).
@@
 The smallest counterexample requires calling a function {\tt f},
-which writes into an uninitialized location $l$ (which is always fine), and then
+which writes into an uninitialized location $l$ (which is always fine\apt{is it??? in the strict policy?}, and then
 calls a further function {\tt g} (which should have the effect of sealing $l$
@@
 \rb{The previous section already presents ``unused'' as the default tag,
-though it doesn't say that it can be overwritten whenever}
+  though it doesn't say that it can be overwritten whenever\apt{but see my
+query re the HEADER\_NO\_INIT bug}}
 A function is permitted
@@
 
-To eliminate this cost, \citet{DBLP:conf/sp/RoesslerD18} introduce
-{\em Lazy Tagging and Lazy Clearing}. Under this optimizations,
-the policy does not initialize stack frames on entry or clear them on exit,
+To eliminate this cost, \citet{DBLP:conf/sp/RoesslerD18} augment Lazy Tagging
+with {\em Lazy Clearing}. Under this combination of optimizations,
+the policy does not initialize stack frames on entry \emph{or} clear them on exit,
 and it permits all writes to the stack, even when the $\PCname$ tag
-does not match the memory tag. Instead it propagates the \(\PCname\) tag
-to written memory. Then reads from the stack do still require $\PCname$
+does not match the memory tag; the \(\PCname\) tag is always propagated
+to the written location. Reads from the stack do still require $\PCname$
 tag and memory tag to match. Thus, even if a callee illicitly writes to
@@
 These \emph{lazy policies} admit more efficient implementations, but
-they deliberately allow violations of stack integrity temporarily,
+they deliberately allow temporary violations of stack integrity,
 with the checks deferred until the point a violation truly becomes
-harmful. The natural question then is, how do we characterize the
-protections provided by these policies once stack safety is broken?
+harmful. The natural question then is: how do we characterize the
+protections provided by these policies once stepwise stack safety is broken?
 How does a harmful violation that must be caught later differ from a
@@
 
-We first notice that, in principle, these properties appear to enforce
+We first notice that, in principle\apt{??}, these properties appear to enforce
 confidentiality outright. So we wish in particular to find an integrity
@@
 
-For simplicity, our definitions return to the simple stack without sharing,
-but all properties in this section can be extended to more complex settings
-very naturally.
+%For simplicity, our definitions return to the simple stack without sharing,
+%but all properties in this section can be extended to more complex settings
+%very naturally.
 
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
+For any \(\MPCT \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
 suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
-\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Find the set of components
-that changed during the call, \(\components = \prot(\context) \cap \Delta(\mach,\mach')\).
-A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\) and its induced trace
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\MPCT' \doteq \NPCT\).
+\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Let 
+\(\components = \mathit{protected}(\context) \cap \Delta(\mach,\mach')\), the
+set of sealed components that changed during the call.
+A system enjoys \textbf{\textit{deferred stepwise integrity}} if, for any \(\nach'\) and \(\NPCT'\)
+such that \(\mach' \approx_{\components} \nach'\) and 
+\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.\apt{again, shouldn't
+we quantify over all $M$?}
 
 In short, if sealed components are modified during the call, the caller may not
-access their values.
+access their values.\apt{This is pretty cryptic.}
 %
@@
 i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
-Instead of forbidding such dangerous writes, we say,
+Instead of strictly forbidding such dangerous writes, we say,
 ``No reads after dangerous writes.''
@@
 \subsection{Observability}
+
+\apt{Given that the deferred property suffices to specifcy the lazy policies,
+  the case for observable properties is pretty weak.  Should we just cut them?}
 
@@
 that would be viable as a weakest reasonable definition. Consider a hypothetical
-taint-tracking policy, that allows dangerous writes to be copied as long as the
+taint-tracking policy that allows dangerous writes to be copied as long as the
 taint remains hidden from outside observers. In order to characterize
@@
 
-We say that a pair of observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
+We say that two observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
 as far as an external observer is concerned, written \(\obsT_1 \eqsim
@@
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\),
+For any call trace \(\MPCT \in \MPCT^*\),
 let \(\head(\MPCT) = \mach,\pol,\context\).
@@
 let \(\components\) be the set of protected components that changed during
-the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
+the call, \(\components = \mathit{protected}(\context) \cap \Delta(\mach, \mach')\),
 and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
@@
 \((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
+\apt{Adjust quantification to match final version of deferred intergrity.}
 
@@
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\),
+For any call trace \(\MPCT \in \MPCT^*\),
 let \(\head(\MPCT) = \mach,\pol,\context\) and
-\(\components = \secret(\context)\).
+\(\components = \mathit{secret}(\context)\).
 %
@@
 \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
+\(\nach \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
 \(\MPCT\) and \(\NPCT\), the following conditions all hold:
@@
   meaning that no secrets leak during the call
-\item \(\brk ~ \last(\MPCT) \leftrightarrow \brk ~
+\item \(\mathit{brk}(\MPCT) ~ \last(\MPCT) \leftrightarrow \mathit{brk}(\MPCT) ~
 \last(\NPCT)\)---if one call returns, so does the other
@@
   \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
-  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
+  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)\apt{adjust quantification if needed}
 \end{itemize}
@@
 violation would become visible, as in lazy policies. But by formalizing
-the connection to the eager policy, we can identify when lazy policies
+the connection to the eager policy\apt{???}, we can identify when lazy policies
 miss violations that can become visible. The {\em Lazy Tagging and Clearing}
@@
 however shared by all activations of the same function in order to obtain
-good cacheability of tag rules, but that policy allows leaks between
+good cacheability of tag rules---but that policy allows leaks between
 instances of the same function.
@@
               {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
+              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}
 
@@
          {}
-         {\(\underscore,\context \stepsto_C \context\)}
-
-\definition Let the protected set of a context, \(\prot_P(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+         {\(\_,\context \stepsto_C \context\)}
+
+\definition Let the protected set of a context, \(\mathit{protected}_P(\domMap,\rts)\),
+be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
 or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
@@
 passing model substitutes the above definitions of \(\context\), initial contexts and update
-function, and \(\prot_P\) for their simple equivalents, and is otherwise identical.
-
-\definition Let the secret set of a context, \(\secret_P(\domMap,\rts)\),
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\) or
-\(\domMap ~ \component = \passed{\depth,\depth'}\)
-where \(\depth \leq |\rts| \leq \depth'\).
+function, and \(\mathit{protected}_P\) for their simple equivalents, and is otherwise identical.
 
 \definition
-For both the stepwise and observational versions of stack confidentiality in
-the original model, the equivalent property in the passing model is the same after
-substituting the above definitions of \(\context\), initial contexts and update
-function, and \(\prot_P\) for their simple equivalents.
+For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
+let \(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
+and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
+\emph{stepwise stack confidentiality} if,
+for all \(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
+\(\MPCT \doteq \NPCT\).
 
@@
 
-Stack integrity and confidentiality apply in the coroutine model, based on straightforward
-projections of each stack's domains, etc. and the machinery of the passing model. We will give
-the coroutine-specific properties.
-
-%\definition
-%The protected and secret sets of a context as used for calls,
-%\(\prot_C(\domMap,\rtm,\yts,\stackid)\) and \(\secret_C(\domMap,\rtm,\yts,\stackid)\),
-%are equivalent to their counterparts \(\prot_P\) and \(\secret_P\) applied to a domain map
-%that projects out the stack domains of components in stack \(\stackid\) and a
-%return target stack \(\rts = \rtm ~ \stackid\). Intuitively, within the active stack we protect
-%the same components we would have in the passing model.
-
-%\definition
-%For each of the stepwise, deferred, and observational versions of
-%stack integrity in the simple model, the equivalent property in the
-%passing model substitutes the above definitions of \(\context\), initial contexts and update
-%function, and \(\prot_C\) for their simple equivalents, and is otherwise identical.
-
-%\definition \(\MPCT^*\) is now defined as
-%\[\begin{aligned}
-%\bigcup_{\stackid \in \STACKS,\depth \in \mathbb{Z}} \hspace{-1em} \{\MPCT | \MPCT\langle f_\depth \rangle
-%\text{ is a reachable segment}\} \hspace{1em} &
-%\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rtm,\yts,\stackid')) \triangleq |\rtm ~ \stackid| \geq \depth \\
-%\end{aligned}\]
-%and \(\brk{\MPCT}\) similarly projects the depth of the active stack.
-
-%\definition
-%For either of the stepwise and observational versions of
-%stack confidentiality in the passing model, the equivalent property in the
-%coroutine model substitutes the above definitions.
-
-\definition \(\MPCT^\wedge\) is the \emph{yield set}, defined as
-\[\begin{aligned}
-\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{\MPCT, \neg f_\stackid | \MPCT\langle f_\stackid \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\stackid ~ (\underscore,\underscore,(\underscore,\underscore,\underscore,\stackid')) \triangleq \stackid \not = \stackid' \\
-\end{aligned}\]
-
-Here, we are capturing the entire span that the system spends {\em away} from a given
-coroutine.
+\definition Let the call-protected set of a context,
+\(\mathit{protected}_C(\domMap,\rtm,\yts,\stackid)\),
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\sealed{\_}}\)
+or \(\domMap ~ \component = \instack{\stackid}{\passed{\depth,\depth'}}\) and \(|\rtm ~ \stackid| < \depth\) or
+\(|\rtm ~ \stackid| > \depth'\).
 
 \definition
-Let the protected set of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
+For each of the stepwise, deferred, and observational versions of
+stack integrity in the simple model, the equivalent property in the
+passing model substitutes the above definitions of \(\context\), initial contexts and update
+function, and \(\mathit{protected}_C\) for their simple equivalents, and is otherwise identical.
+
+\definition Let the yield-protected set of a context,
+\(\mathit{protected}_Y(\domMap,\rtm,\yts,\stackid)\),
 be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
-
-\definition
-Let the secret set of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
+\(\domMap ~ \component = \instack{\stackid'}{\_}\) where \(\stackid \not = \stackid`\).
 
@@
 substituting the above definitions of \(\context\), initial contexts and update
-function, and \(\prot_Y\) for their equivalents. Likewise
+function, and \(\mathit{protected}_Y\) for their equivalents. Likewise
 {\em deferred coroutine integrity} and {\em observational coroutine integrity} are defined
-in terms of their simple equivalents using \(\MPCT^\wedge\) in place of \(\MPCT^*\).
-
-\definition
-We define {\em stepwise coroutine integrity} in the same way as stepwise stack
-integrity, substituting the above definitions. Likewise
-{\em observational coroutine confidentiality} is defined in terms of observational
-stack confidentiality.
+in terms of their simple equivalents.
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\unsealed}\) or
+\(\instack{\stackid}{\passed{\depth,\depth'}}\)
+where \(\depth \leq |\rtm ~ \stackid| \leq \depth'\) .
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
+{\em stepwise stack confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
+
+\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
+\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
+For any reachable segment \(\MPCT \langle f_\depth \rangle\),
+let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
+\(\components\) be the set of components \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid}{\_}\).
+Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
+\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
+{\em stepwise coroutine confidentiality} in the coroutine setting if for all
+such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).
 
@@
 
+Notably, we conjecture that while these systems may implement integrity and confidentiality,
+they fall just short of return-to-caller. This is because the standard way to protect returns
+in Cheri is to construct {\em return capability}, an object capability that must be invoked
+to simultaneously transfer control back to the caller and unseal a capability on an activation
+record. An ill-behaved intermediate caller could always pass its caller's return capability on
+to its callee instead of its own, ensuring that the callee will return past it. Instead,
+we believe that Cheri supports a slightly weaker control-flow property in which only callers
+that obey a dynamic reasonableness property are guaranteed not to be skipped.
+
 \paragraph*{Program Logic for Stacks}

### 2021-04-17 tex/main.tex

--- 2021-04-17 tex/main.tex
@@
   universally quantified, and probably also \(\MPCT\) (?) }
-\sna{Better?}
+\sna{Better?\apt{only partly. Quantification of $M,\brk$ still seems wrong to me. }}
 
@@
 
-Finally, we formalize a version of WBCF, capturing the same informal idea
-as \citet{Skorstengaard+19b} using our notion of contexts.
+Finally, we formalize a version of WBCF, which captures the same informal idea
+as that of \citet{Skorstengaard+19b} using our context machinery.
 \definition
@@
 \(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
-\(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
+\(\mpcstatename \stepsto \mpcstatename'\)\apt{doesn't reachability of $\mpcstatename$ imply reachability of $\mpcstatename'$?}, if \(|\rts'| < |\rts|\), then
 \(|\rts'| = |\rts|-1\)
@@
 
-For simplicity, our definitions return to the simple stack without sharing,
-but all properties in this section can be extended to more complex settings
-very naturally.
+%For simplicity, our definitions return to the simple stack without sharing,
+%but all properties in this section can be extended to more complex settings
+%very naturally.
 
@@
 suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
-\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Find the set of components
-that changed during the call, \(\components = \prot(\context) \cap \Delta(\mach,\mach')\).
-A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\) and its induced trace
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\MPCT' \doteq \NPCT\).
+\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Let 
+\(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
+set of sealed components that changed during the call.
+A system enjoys \textbf{\textit{deferred stepwise integrity}} if, for any \(\nach'\) and \(\NPCT'\)
+such that \(\mach' \approx_{\components} \nach'\) and 
+\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.\apt{again, shouldn't
+we quantify over all $M$?}
 
 In short, if sealed components are modified during the call, the caller may not
-access their values.
+access their values.\apt{This is pretty cryptic.}
 %

### 2021-04-17 tex/main.tex
Code map predicate?
Present-tense 7.2 (except when referring to bugs)

@@
 Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
+\rb{Technically a partial function from addresses to annotations,
+  though its domain can function as a predicate.}
 
@@
 
-We validated\apt{could we stay in present tense throughout?}  the micro-policy of the previous section by
+We validate the micro-policy of the previous section by
 systematically testing that it satisfies eager stack integrity and
-confidentiality. We used a Coq specification of the RISC-V
+confidentiality. We use a Coq specification of the RISC-V
 architecture
@@
 \leo{translated from Haskell via hs-2-coq?\apt{not worth saying since
-  the MIT people did it}}  and extended it with a
+  the MIT people did it}}  and extend it with a
 runtime monitor implementing the stack-safety
 micro-policy.\footnote{This code will be made public. \leo{More?}} We
-  chose the Coq proof assistant as the setting for our implementation
+  choose the Coq proof assistant as the setting for our implementation
   for two reasons: first, to ensure that our coinductive trace
@@
 
-To use QuickChick, we built random test-case generators that produce
+To use QuickChick, we build random test-case generators that produce
 % \begin{enumerate}
@@
   blessed call or return sequences appropriately, while marking
-  all potential stack locations as $\tagNoDepth$, and
+  all potential stack locations as $\tagNoDepth$; and
 % \item
@@
 
-We extended this technique to keep track of the control flow behavior of
+We extend this technique to keep track of the control flow behavior of
 the program being generated: each time a call or return sequence is
@@
 
-In addition, we needed to extend this technique to handle the nested
+In addition, we need to extend this technique to handle the nested
 nature of our confidentiality property: rather than just generating two
@@
 To ensure the effectiveness of testing against our formal properties, we
-used {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
+use {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
 (mutations) in a program that should cause the property of interest (here,
 stack safety) to fail, and ensure that the testing framework can find
-them. The bugs we used for our evaluation were either artificially generated
-by us (deliberately weakening the micro-policy in ways that we expected
+them. The bugs we use for our evaluation are either artificially generated
+by us (deliberately weakening the micro-policy in ways that we expect
 should break its guarantees), or actual bugs that we discovered through
@@
 $\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing that constraints causes violations to the integrity
+tag. Relaxing that constraint causes violations to the integrity
 property (row {\em STORE\_NO\_CHECK}).
@@
 implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code
+and targeted ARM machine code;
 %\bcp{right?}\leo{yeah}
-; we re-implemented it in Coq targeting RISC-V).  These errors ranged
-from trivial typos to ones that required an intriguingly complex setup
+we re-implemented it in Coq targeting RISC-V).  These errors range
+from trivial typos to ones that require an intriguingly complex setup
 to appear.  The most interesting bug (included in the table as row

### 2021-04-17 tex/main.tex
pass through 9

--- 2021-04-17 tex/main.tex
@@
 
-We have shown how our properties apply to an extremely simply
+We have shown how our properties apply to an extremely simple
 subroutine model and validated them through randomized testing.
@@
 To this point we have ignored function arguments, since our simplest model places them in
-registers. We now implement two kinds of protected
-sharing: passing by value on the stack, and passing by reference. First we will explore
+registers. We now show how to implement two kinds of protected
+sharing: passing by value on the stack, and passing by reference. We first explore
 some examples at the same level of detail as our earlier ones, and then move on to
@@
 We extend our call annotation with an additional parameter identifying a memory location to be
-passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
+passed (the diagram displays it in magenta). Our notion of domains is extended as well,
 to include \(\passed{\depth,\depth'}\), recording two depths. We note the
@@
 return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
-a future call with no arguments passed by reference would also violate integrity if it wrote
-to either variable.
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity.
+%For that matter
+%a future call with no arguments passed by reference would also violate integrity if it wrote
+%to either variable.
 
@@
 and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices to track two depths for each passed
+an unbroken chain of calls. So it suffices for the domain to track two depths for each passed
 cell: the original owner, and the highest depth it has been passed to. The policy considers
@@
 used in certain corner cases.
-More seriously, callees are permitted to access the location without having
+But more seriously, callees are permitted to access the location without having
 been explicitly given a pointer to it.  Preventing the latter kind of abuse
@@
 
-Next we introduce variable passing. We will need to add extra information to our call
+\paragraph{Formalization of variable passing.}
+We add extra information to our call
 annotations to reflect that they may now pass variables on the stack, and that the locations
@@
 cells at the top of the frame, while passing by reference just gives the
-addresses of the appropriate variables.
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-And we extend the set of domains to include passed components,
+addresses of the appropriate variables.\apt{This still doesn't make sense to me. We cannot
+  give a static annotation that identifies the absolute address of a stack-allocated variable!
+  Only sp-relative addresses make sense.}
+
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P) \hspace*{1in} P \subseteq (\MACHS \times \COMPONENTS)\]
+We extend the set of domains to include passed components,
 which are labeled with the range of depths of the caller that has passed them.
 \[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\rts \in \listT (\target \subseteq \MACHS)) \]
+%\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+%\times (\rts \in \listT (\target \subseteq \MACHS)) \]
+The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
+\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
+ \times (\listT (\target)) \]
 We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map when pushing it, where \(\depth\) is an integer depth.
+update a domain map at a call from depth \(\depth\).
 \[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
@@
 
-And to pop we just unseal the components at depth \(\depth\):
+At the matching return we just unseal the components at depth \(\depth\):
 
@@
 
-Now we can inductively define the context update function; these again apply in the order
+Now we can inductively define the context update function; these rules again apply in the order
 presented. \bcp{Too much space between the second and third.}
@@
 stack integrity in the simple model, the equivalent property in the
-passing model substitutes the above definitions of \(\context\), initial contexts and update
+passing model substitutes the above definitions of \(\context\), initial contexts\apt{not defined above??}  and update
 function, and \(\prot_P\) for their simple equivalents, and is otherwise identical.
@@
 the original model, the equivalent property in the passing model is the same after
-substituting the above definitions of \(\context\), initial contexts and update
+substituting the above definitions of \(\context\), initial contexts\apt{ditto} and update
 function, and \(\prot_P\) for their simple equivalents.
@@
 
-Our final extension to the model is to add coroutines. We assume a very simple model with
+Our final extension to the model adds coroutines. We assume a very simple model with
 a fixed number of coroutines, each with its own stack.
@@
 pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We add a second annotation indicating yields.
+We add a second annotation indicating yields.\apt{seems unchanged from (old version of) previous section}
 
@@
 \[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
-(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
-\target) \times (\stackid \in \STACKS)\]
+\[(\domMap,\rtm,\yts,\stackid) = \context \in \CONTEXTS ::= 
+ (\COMPONENTS \rightarrow \STKDOMS) \times
+(\STACKS \rightarrow \listT(\target)) \times (\STACKS \rightarrow \target) \times \STACKS\]
 
@@
 We define {\em stepwise coroutine integrity} in the same way as stepwise stack integrity,
-substituting the above definitions of \(\context\), initial contexts and update
+substituting the above definitions of \(\context\), initial contexts\apt{??} and update
 function, and \(\prot_Y\) for their equivalents. Likewise

### 2021-04-17 tex/main.tex
pass on 10,11. Changed bib keys for Skors papers to avoid confusing myself even further

@@
 shadow stacks~\citep{Dang+15,Shanbhogue+19},
-capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
+capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
@@
 The most abstract characterization of stack safety that we are aware of is
-recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
+recent work by \citet{SkorstengaardSTK}, which defines stack safety as the conjunction of two
 properties:
@@
 condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $B$.  This is what \citet{Skorstengaard+19b} describe
+only return to $B$.  This is what \citet{SkorstengaardSTK} describe
 as \emph{well-bracketed control flow (WBCF)}.
@@
 Finally, we formalize a version of WBCF, which captures the same informal idea
-as that of \citet{Skorstengaard+19b} using our context machinery.
+as that of \citet{SkorstengaardSTK} using our context machinery.
 \definition
@@
 a positive and formal characterization of
-stack safety is \citet{Skorstengaard+19b}, who introduce a calling
+stack safety begins with \citet{SkorstengaardLocal}, who introduce a calling
 convention that uses local capabilities to preserve local state
@@
 and can be used to reason about individual programs.
-\citeauthor{Skorstengaard+19b} acknowledge the difficulty of defining
+\citeauthor{SkorstengaardSTK} acknowledge the difficulty of defining
 these concepts precisely:
@@
 
-StkTokens \citep{Skorstengaard+19} continues this work by defining a more efficient
+StkTokens \citep{SkorstengaardSTK} continues this work by defining a more efficient
 calling convention that makes use of linear capabilities for stack and return
@@
 frame to the immediately adjacent frame below. Formally, it improves on
-\citet{Skorstengaard+19b} by building those properties into the semantics of a
+\citet{SkorstengaardLocal} by building those properties into the semantics of a
 capability machine with a built-in call stack and call and return instructions,
@@
 \rb{Trimmed this a bit. Are you referring to their list of attack
-  examples, or to something else?}
+  examples, or to something else?\apt{Their list at start of section 4.}}
 
@@
 
-\paragraph*{Stack Safety in the Cheri Zoo}
+\paragraph*{Stack Safety in the Cheri Ecosystem}
 
 There are several proposals around the use of Cheri capabilities to enforce stack safety,
-ranging from mechanisms that use the standard Cheri hardware \citep{Skorstengaard+19b},
+ranging from mechanisms that use the standard Cheri hardware (which includes local
+capabilities) \citep{SkorstengaardLocal},
 to those that propose entirely new types of capabilities, such as linear
-(\citep{Skorstengaard+19}), uninitialized (\citep{Georges+21}), and monotonic (\sna{cite}).
+\citep{SkorstengaardSTK}, uninitialized \citep{Georges+21}, lifetime \citep{Tsampas+19}, and monotonic \citep{GeorgesComplete}.
 Of these, uninitialized capabilities seem particularly promising---they naturally enforce
@@
 to clear large swaths of the stack. We propose to implement a model of the stack safety
-system described in (\citep{Georges+21}) and test it against our properties using
+system described in \citet{Georges+21} and test it against our properties using
 QuickChick.

### 2021-04-17 tex/main.tex
Remove note about not dealing with "rich" sharing

@@
 to allow passing of scalar stack data, granting a callee access to the
-caller's data. (We do not consider more sophisticated forms of sharing, such
-as taking the address of a variable and treating it like a pointer.)
-\rb{But actually, this is discussed at some length is \cref{sec:passing},
-  and it is the topic of \cref{fig:passing}.  Reaching that point after
-  reading this disclaimer is confusing.}\bcp{So should we just remove that
-  sentence?}
+caller's data.
+%% (We do not consider more sophisticated forms of sharing, such
+%% as taking the address of a variable and treating it like a pointer.)
+%% \rb{But actually, this is discussed at some length is \cref{sec:passing},
+%%   and it is the topic of \cref{fig:passing}.  Reaching that point after
+%%   reading this disclaimer is confusing.}\bcp{So should we just remove that
+%%   sentence?}
 %

### 2021-04-17 tex/main.tex
Intro comments

@@
 data safety in the face of control-flow attacks (see \cref{sec:control}).
-\rb{\cref{sec:RTC}?}
+\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
 This reflects an arguably cleaner division between data protection and
@@
 property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement}).
-\bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
+of~\citet{DBLP:conf/sp/RoesslerD18}, and extend the generation by
+execution techniques of \citet{TestingNI:ICFP} to handle nested
+properties (\cref{sec:enforcement}).
+%% \bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
 \item We formalize a

### 2021-04-17 tex/main.tex
Finished pass

--- 2021-04-17 tex/main.tex
@@
 ``lazy'' micro-policy corresponds to a slightly weaker and more extensional
-variant of our properties.
+``observational'' variant of our properties.
 %
@@
 shadow stacks~\citep{Dang+15,Shanbhogue+19},
-capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
+capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
@@
 The most abstract characterization of stack safety that we are aware of is
-recent work by \citet{SkorstengaardSTK}, which defines stack safety as the conjunction of two
+recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
 properties:
@@
 they then show that an enforcement
-mechanism (in their case, hardware capabilities) guarantees stack safety, by
+mechanism (in their case, hardware capabilities) guarantees these properties, by
 proving that an ordinary machine with enforcement simulates the idealized
@@
 
-In the present work, we also aim to protect both stack data (like
-LSE) and
+In the present work, we also describe the protection of both stack data (like LSE) and
 control flow (like WBCF).  Our technical formulation of WBCF is similar to
-that of Skorstengaard et al.{}, but our treatment of stack data protection
-is rather different.  Instead of defining it implicitly,
+theirs, but our treatment of stack data protection is rather different,
+as is our approach to the definition itself. Instead of defining stack safety implicitly,
 as ``what a certain idealized machine does,'' we give an
@@
 callees, etc.). They are also stronger than LSE; even in the absence of the
-control flow guarantees of {\em return to caller}, they guarantee callers'
-data safety in the face of control-flow attacks (see \cref{sec:control}).
-\rb{\cref{sec:RTC}?}
+control flow guarantees of WBCF, they guarantee callers'
+data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
 This reflects an arguably cleaner division between data protection and
@@
 
-For the confidentiality and integrity properties, we consider both {\em
-  stepwise} and {\em observational} variants.  The simpler {stepwise}
+For confidentiality and integrity , we consider both {\em stepwise}
+and {\em observational} variants.  The simpler stepwise
 variants formalize the intuition that a caller's local data is {\em never}
-read or modified during a call. The more extensional observational variants
+read or modified during a call. The observational variants
 allow a callee to read from and write to their caller's stack frame, as
@@
 they write affects the observable behavior of the caller after they return.
+The observational properties are meant to capture the same sense of security,
+extensionally enough that any reasonable protection mechanism ought to enforce
+them, even if it does not prevent every single dangerous read or write.
 
@@
 then permitted only when that activation is currently executing, validates our
-stepwise properties.  On the other hand, their more efficient \emph{Lazy
-  Tagging and Clearing} policy enforces observational integrity, but
-testing
-reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
-stepwise or
-observational): it violates the temporal aspect of confidentiality in
-corner cases where data can leak across repeated calls to the same callee.  We
-propose a variant of {\em Lazy Tagging and Clearing} that \emph{does} enforce
-confidentiality (as far as our testing can tell\bcp{right, leo?}), albeit at
-some performance cost.
+stepwise properties.  On the other hand, testing reveals their more efficient
+\emph{Lazy Tagging} policy does not enforce step-wise integrity. In fact
+we reason that both of their lazy policies violate the temporal aspect of confidentiality in
+corner cases where data can leak across repeated calls to the same callee,
+and \emph{Lazy Tagging and Clearing} also violates integrity in these instances. We
+propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
+confidentiality, albeit at some performance cost.
 %
@@
 properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy ones.  In
+machine-code level for a variety of ISAs, including RISCy \sna{RISCy?} ones.  In
 particular, there may be no single ``call'' instruction that does all the
@@
 that we are given some way of telling when a given machine state is a
-{\em return target}\bcp{with respect to some set of pending calls?}.
+{\em return target} with respect to some set of pending calls.
 
@@
 
-Intuitively, stack safety means that the stack frame of each caller (and
-those of its ancestor callers) is protected
+Intuitively, stack safety means that the stack frame of each caller is protected
 from attempts by a confused or malicious callee (including all its
-descendant callees)
-to read or write its private data.
+descendant callees) to read or write its private data.
 That is, the attacker model treats every caller as a defender and its callee as
 an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways.
+may be viewed in both ways. We do rule out some
+scenarios with a ``well-formedness'' criterion on initial states, ensuring that
+they are set up properly for the enforcement mechanism in question.
+When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
+well-behaved calls and returns are implemented by fixed sequences of instructions
+which the tag policy will force to be executed from start to finish,
+termed {\em blessed sequences}. Well-formedness requires that all annotated
+instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
+This is all we need for the policy to protect itself: it can then failstop if it would
+execute the blessed sequence out of order.
 
@@
 to protect both stack and heap.
-
-% \paragraph*{Well-formedness and Bad Annotations}
-% \bcp{What is ``well-formedness''??}
-
-Since we rely on call annotations\bcp{to...}, we must ask: what if the annotations
-are wrong? First, suppose a call occurs that is not annotated.
-In this case, the answer is simple: as far as the properties are concerned, there has
-been no call, only an instruction sequence resembling one, and
-therefore no associated change in permissions. Then the caller is guaranteed no
-protections.
-
-\bcp{Confusing, and maybe out of date?}Alternatively, an annotated instruction might not obey the calling convention.
-In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
-and returns are implemented by fixed sequences of instructions which the tag policy
-will force to be executed from start to finish, termed {\em blessed sequences}.
-When applying our properties to a micro-policy setting, we will statically rule out some
-scenarios with a well-formedness criterion on initial states: all annotated instructions will
-be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
-This is all we need for the policy to protect itself: it can then failstop if it would
-execute the blessed sequence out of order.
 
@@
 \memoryaddrs{8em}
-\memory{3}{\unsealc}[Unseal]
+\memory{3}{\unsealc}[U]
 ~$\cdots$
@@
 \memoryaddrs{16em}
-\memory{1}{\mainsealc}[Seal(0)]%
+\memory{1}{\mainsealc}[S(0)]%
 \memory{1}{\retptrc}[RetPtr]%
-\memory{1}{\unsealc}[Unseal]
+\memory{1}{\unsealc}[U]
 ~$\cdots$
@@
 {\em domain}---a label reflecting a particular set of access rights. Domains
-are \(\unsealed\)\bcp{``unsealed'' would read better} (in the stack and accessible to writes), \(\sealed{\depth}\)
+are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
 (in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
-(outside the stack entirely.) For our purposes, \(\outside\) will apply to global
+(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
+figures. For our purposes, \(\outside\) will apply to global
 variables. Domains are part of the {\em context} of a state.
@@
 condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $B$.  This is what \citet{SkorstengaardSTK} describe
+only return to $B$.  This is what \citet{Skorstengaard+19b} describe
 as \emph{well-bracketed control flow (WBCF)}.
@@
 by describing how we model knowledge about the program structure,
-then dive into the formal definitions of local state encapsulation\apt{we don't want
-  to call it that here, do we?} and
+then dive into the formal definitions of integrity, confidentiality, and
 well-bracketed control flow, which together constitute stack safety.
@@
 which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
+this the {\em sealing convention},
+\(\sealcon ~ (\mach \in \MACHS) ~ (\component \in \COMPONENTS)\), a
 relation between machine states and the addresses that should be sealed.
@@
 which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
+it means to return with a {\em return convention},
+\(\retcon ~ (\mach \in \MACHS) ~ (\mach' \in \MACHS)\),
 which relates call states to the states that can be considered valid returns to that call.
@@
 at a particular depth. Each caller registers a {\em target}
-condition---a predicate on machine states that must hold for it to consider itself active.
+condition, \(\target ~ \mach\)---a predicate on machine states that means it
+has been returned to.
 A context is a pair of a ``domain map'' from components to domains, and a
 ``return target list,'' which is a stack of targets, one for each caller awaiting
-return.  \ifaftersubmission\bcp{spacing is funny}\fi
+return.
 %
@@
 Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
-\rb{Technically a partial function from addresses to annotations,
-  though its domain can function as a predicate.}
 
@@
 A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
-\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
-and any component \(\component \in \prot(\domMap,\rts)\),
-\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
-\apt{but now $dm$ and $rts$ are unbound}.
+\(\mach,\pol,\context\) such that \(\mach,\pol,\context \stepstocon \mpcstatename'\),
+and any component \(\component \in \prot(\context)\),
+\(\pi_\mach(\mpcstatename') ~ \component = \mach ~ \component\).
 
@@
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-let \(\components = \mathit{secret}(\context)\).
 A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for any \(\nach\) such that \(\mach \approx_\components \nach\) and the induced trace
+for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
+\(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\) and the induced trace
 \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
-\(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
-  universally quantified, and probably also \(\MPCT\) (?) }
-\sna{Better?\apt{only partly. Quantification of $M,\brk$ still seems wrong to me. }}
+\(\MPCT \doteq \NPCT\) holds.
 
@@
 Finally, we formalize a version of WBCF, which captures the same informal idea
-as that of \citet{SkorstengaardSTK} using our context machinery.
+as that of \citet{Skorstengaard+19b} using our context machinery.
 \definition
@@
 systematically testing that it satisfies eager stack integrity and
-confidentiality. We use a Coq specification of the RISC-V
+confidentiality. We used a Coq specification of the RISC-V
 architecture
@@
 \leo{translated from Haskell via hs-2-coq?\apt{not worth saying since
-  the MIT people did it}}  and extend it with a
+  the MIT people did it}}  and extended it with a
 runtime monitor implementing the stack-safety
 micro-policy.\footnote{This code will be made public. \leo{More?}} We
-  choose the Coq proof assistant as the setting for our implementation
+  chose the Coq proof assistant as the setting for our implementation
   for two reasons: first, to ensure that our coinductive trace
@@
 
-To use QuickChick, we build random test-case generators that produce
+To use QuickChick, we built random test-case generators that produce
 % \begin{enumerate}
@@
   blessed call or return sequences appropriately, while marking
-  all potential stack locations as $\tagNoDepth$; and
+  all potential stack locations as $\tagNoDepth$, and
 % \item
@@
 
-We extend this technique to keep track of the control flow behavior of
+We extended this technique to keep track of the control flow behavior of
 the program being generated: each time a call or return sequence is
@@
 
-In addition, we need to extend this technique to handle the nested
+In addition, we needed to extend this technique to handle the nested
 nature of our confidentiality property: rather than just generating two
@@
 To ensure the effectiveness of testing against our formal properties, we
-use {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
+used {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
 (mutations) in a program that should cause the property of interest (here,
 stack safety) to fail, and ensure that the testing framework can find
-them. The bugs we use for our evaluation are either artificially generated
-by us (deliberately weakening the micro-policy in ways that we expect
+them. The bugs we used for our evaluation were either artificially generated
+by us (deliberately weakening the micro-policy in ways that we expected
 should break its guarantees), or actual bugs that we discovered through
@@
 $\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing that constraint causes violations to the integrity
+tag. Relaxing that constraints causes violations to the integrity
 property (row {\em STORE\_NO\_CHECK}).
@@
 implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code;
+and targeted ARM machine code
 %\bcp{right?}\leo{yeah}
-we re-implemented it in Coq targeting RISC-V).  These errors range
-from trivial typos to ones that require an intriguingly complex setup
+; we re-implemented it in Coq targeting RISC-V).  These errors ranged
+from trivial typos to ones that required an intriguingly complex setup
 to appear.  The most interesting bug (included in the table as row
@@
 $l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
+\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
+  and the write from {\tt f} was already a violation.}
+
 \section{Stack Safety for Lazy Enforcement}
@@
 
-First, they eliminate the initialization pass. In this {\em Lazy Tagging}
-system, the stack is initially tagged as ``clear.''
+First, they eliminate the initialization pass. In the {\em Lazy Tagging}
+system, the stack is initially tagged as \(\tagNoDepth\),
 \rb{The previous section already presents ``unused'' as the default tag,
   though it doesn't say that it can be overwritten whenever\apt{but see my
-query re the HEADER\_NO\_INIT bug}}
+    query re the HEADER\_NO\_INIT bug}}\sna{Correct there, but now I'm talking
+  about laziness.}
 A function is permitted
@@
 the clear tag, claiming the memory; the remainder of the policy functions
-as {\em Depth Isolation}. On return, the callee must clear its frame, which
-is still a heavy cost.
+as {\em Depth Isolation}. On return, the callee must clear its frame, restoring
+it to \(\tagNoDepth\), which is still a heavy cost.
 
@@
 if it ever tries to read from that location.
-\rb{This is in line with the level of detail in the last section.}
 %
@@
 
-We first notice that, in principle\apt{??}, these properties appear to enforce
-confidentiality outright. So we wish in particular to find an integrity
-property. The intuition is that sealed memory can change, but it is
+We first notice that these properties prevent invalid reads, as well as reads
+that follow invalid writes, but the writes themselves are permitted.
+So we wish in particular to find an integrity property that characterizes them.
+The intuition is that sealed memory can change, but it is
 then tainted, and its original owner cannot actually read it. This
@@
 
-%For simplicity, our definitions return to the simple stack without sharing,
-%but all properties in this section can be extended to more complex settings
-%very naturally.
-
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
-suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
-\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Let 
+A system enjoys \textbf{\textit{deferred stepwise integrity}} if,
+for any \(\MPCT,\brk \in \MPCT^*\) where \(\head(\MPCT) = \mach,\pol,\context\) and
+\(\last(\MPCT) = \mach',\pol',\context'\), the following holds. Let
+\((\mach',\pol',\context') \hookrightarrow \MPCT'\) and
 \(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
 set of sealed components that changed during the call.
-A system enjoys \textbf{\textit{deferred stepwise integrity}} if, for any \(\nach'\) and \(\NPCT'\)
-such that \(\mach' \approx_{\components} \nach'\) and 
-\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.\apt{again, shouldn't
-we quantify over all $M$?}
-
-In short, if sealed components are modified during the call, the caller may not
-access their values.\apt{This is pretty cryptic.}
-%
+For any \(\nach'\) and \(\NPCT'\) such that \(\mach' \approx_{\components} \nach'\) and 
+\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.
+
 For example, consider again the integrity violation in
@@
 
-We argue that while deferred stepwise integrity meaningfully captures the intuition
-behind lazy tagging, it stops short of the kind of extensional characterization
-that would be viable as a weakest reasonable definition. Consider a hypothetical
-taint-tracking policy that allows dangerous writes to be copied as long as the
-taint remains hidden from outside observers. In order to characterize
-such a policy, we would need a model of observation.
-Similarly, for such a taint-based policy, lockstep confidentiality would be too
-strong, as we might permit secrets to move about as long as they were not
-leaked to the outside world.
+While deferred integrity meaningfully captures the intuition
+behind lazy tagging, we seek properties that can be used to evaluate any reasonable
+enforcement mechanism. Consider a hypothetical
+taint-tracking policy that allows high-security to be copied as long as it is
+tagged as such and not exfiltrated, or that allows dangerous writes to be
+copied as long as they don't affect critical operations. This calls for
+an abstract mechanism to characterize behaviors that are externally relevant.
 
@@
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\),
-let \(\head(\MPCT) = \mach,\pol,\context\).
-%
-Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
-let \(\components\) be the set of protected components that changed during
+{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
+where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
+the following holds. Let \(\components\) be the set of protected components that changed during
 the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
 and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
-A system enjoys {\em observational integrity} if for any \(\nach\) such that
+Then for any \(\nach\) such that
 \(\mach' \approx_{\components'} \nach\), and its induced trace
 \((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
-\apt{Adjust quantification to match final version of deferred intergrity.}
 
@@
 \definition
-For any \(\MPCT,\brk \in \MPCT^*\),
-let \(\head(\MPCT) = \mach,\pol,\context\) and
-\(\components = \secret(\context)\).
-%
 A system enjoys {\em observational confidentiality} if for all
-\(\nach\) such that \(\mach \approx_\components \nach\) and
+\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\pol,\context\) and
+\(\components = \secret(\context)\), the following holds.
+%
+For any \(\nach\) such that \(\mach \approx_\components \nach\) and
 \(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
-\(\MPCT\) and \(\NPCT\), the following conditions all hold:
+\(\MPCT\) and \(\NPCT\) respect three conditions:
 
@@
   \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
-  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)\apt{adjust quantification if needed}
+  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
 \end{itemize}
@@
 Observable properties allow us to defer enforcement until a property
-violation would become visible, as in lazy policies. But by formalizing
-the connection to the eager policy\apt{???}, we can identify when lazy policies
+violation would become visible. Now we can identify when lazy policies
 miss violations that can become visible. The {\em Lazy Tagging and Clearing}
@@
 
-We have shown how our properties apply to an extremely simple
+We have shown how our properties apply to an extremely simply
 subroutine model and validated them through randomized testing.
@@
 To this point we have ignored function arguments, since our simplest model places them in
-registers. We now show how to implement two kinds of protected
-sharing: passing by value on the stack, and passing by reference. We first explore
+registers. We now implement two kinds of protected
+sharing: passing by value on the stack, and passing by reference. First we will explore
 some examples at the same level of detail as our earlier ones, and then move on to
@@
 We extend our call annotation with an additional parameter identifying a memory location to be
-passed (the diagram displays it in magenta). Our notion of domains is extended as well,
+passed, and our diagram displays it in magenta. Our notion of domains is extended as well,
 to include \(\passed{\depth,\depth'}\), recording two depths. We note the
@@
 return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity.
-%For that matter
-%a future call with no arguments passed by reference would also violate integrity if it wrote
-%to either variable.
+and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
+a future call with no arguments passed by reference would also violate integrity if it wrote
+to either variable.
 
@@
 and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices for the domain to track two depths for each passed
+an unbroken chain of calls. So it suffices to track two depths for each passed
 cell: the original owner, and the highest depth it has been passed to. The policy considers
@@
 used in certain corner cases.
-But more seriously, callees are permitted to access the location without having
+More seriously, callees are permitted to access the location without having
 been explicitly given a pointer to it.  Preventing the latter kind of abuse
@@
 
-\paragraph{Formalization of variable passing.}
-We add extra information to our call
+Next we introduce variable passing. We will need to add extra information to our call
 annotations to reflect that they may now pass variables on the stack, and that the locations
@@
 cells at the top of the frame, while passing by reference just gives the
-addresses of the appropriate variables.\apt{This still doesn't make sense to me. We cannot
-  give a static annotation that identifies the absolute address of a stack-allocated variable!
-  Only sp-relative addresses make sense.}
-
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P) \hspace*{1in} P \subseteq (\MACHS \times \COMPONENTS)\]
-We extend the set of domains to include passed components,
+addresses of the appropriate variables.
+\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
+And we extend the set of domains to include passed components,
 which are labeled with the range of depths of the caller that has passed them.
 \[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
-%\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-%\times (\rts \in \listT (\target \subseteq \MACHS)) \]
-The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
- \times (\listT (\target)) \]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+\times (\rts \in \listT (\target \subseteq \MACHS)) \]
 We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map at a call from depth \(\depth\).
+update a domain map when pushing it, where \(\depth\) is an integer depth.
 \[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
@@
 
-At the matching return we just unseal the components at depth \(\depth\):
+And to pop we just unseal the components at depth \(\depth\):
 
@@
 
-Now we can inductively define the context update function; these rules again apply in the order
+Now we can inductively define the context update function; these again apply in the order
 presented. \bcp{Too much space between the second and third.}
@@
 stack integrity in the simple model, the equivalent property in the
-passing model substitutes the above definitions of \(\context\), initial contexts\apt{not defined above??}  and update
+passing model substitutes the above definitions of \(\context\), initial contexts and update
 function, and \(\prot_P\) for their simple equivalents, and is otherwise identical.
@@
 the original model, the equivalent property in the passing model is the same after
-substituting the above definitions of \(\context\), initial contexts\apt{ditto} and update
+substituting the above definitions of \(\context\), initial contexts and update
 function, and \(\prot_P\) for their simple equivalents.
@@
 
-Our final extension to the model adds coroutines. We assume a very simple model with
+Our final extension to the model is to add coroutines. We assume a very simple model with
 a fixed number of coroutines, each with its own stack.
@@
 pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We add a second annotation indicating yields.\apt{seems unchanged from (old version of) previous section}
+We add a second annotation indicating yields.
 
@@
 \[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
-\[(\domMap,\rtm,\yts,\stackid) = \context \in \CONTEXTS ::= 
- (\COMPONENTS \rightarrow \STKDOMS) \times
-(\STACKS \rightarrow \listT(\target)) \times (\STACKS \rightarrow \target) \times \STACKS\]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times
+(\rtm \in \STACKS \rightarrow \listT ~ \target) \times (\yts \in \STACKS \rightarrow
+\target) \times (\stackid \in \STACKS)\]
 
@@
 We define {\em stepwise coroutine integrity} in the same way as stepwise stack integrity,
-substituting the above definitions of \(\context\), initial contexts\apt{??} and update
+substituting the above definitions of \(\context\), initial contexts and update
 function, and \(\prot_Y\) for their equivalents. Likewise
@@
 a positive and formal characterization of
-stack safety begins with \citet{SkorstengaardLocal}, who introduce a calling
+stack safety is \citet{Skorstengaard+19b}, who introduce a calling
 convention that uses local capabilities to preserve local state
@@
 and can be used to reason about individual programs.
-\citeauthor{SkorstengaardSTK} acknowledge the difficulty of defining
+\citeauthor{Skorstengaard+19b} acknowledge the difficulty of defining
 these concepts precisely:
@@
 
-StkTokens \citep{SkorstengaardSTK} continues this work by defining a more efficient
+StkTokens \citep{Skorstengaard+19} continues this work by defining a more efficient
 calling convention that makes use of linear capabilities for stack and return
@@
 frame to the immediately adjacent frame below. Formally, it improves on
-\citet{SkorstengaardLocal} by building those properties into the semantics of a
+\citet{Skorstengaard+19b} by building those properties into the semantics of a
 capability machine with a built-in call stack and call and return instructions,
@@
 %% observations is limited to cotermination.
-\sna{Some of this seems like more detail than we need, and we need to
-  compare ourselves to their list of criteria.}
-\rb{Trimmed this a bit. Are you referring to their list of attack
-  examples, or to something else?\apt{Their list at start of section 4.}}
 
@@
 characterization of the meaning of (heap) memory safety, in this paper,
-inspired by their work, we aim to do the same for the stack.
-%
-Their model describing safety of shared stack-allocated
-objects\BCP{huh?? they were not talking about stacks} extends naturally to heap safety, complicated by the fact that
-these objects are not deallocated by returns\bcp{??}, and therefore a
-function's privilege may increase or decrease after its entry
-point. For instance, if a callee allocates a heap object and returns
-the pointer to its caller, that object's addresses become accessible
-in contradiction to a contour computed at the caller's entry. So such
-an extension must enforce trace confidentiality and integrity
-properties separately on continuous segments of each call, with
-contours computed at each crossing between caller and callee.  \BCP{Didn't
-  understand very much of that paragraph.}
+inspired by their work, we aim to do the same for the stack. They are a
+significant inspiration for this work.
 
@@
 
-\paragraph*{Stack Safety in the Cheri Ecosystem}
+\paragraph*{Stack Safety in the Cheri Zoo}
 
 There are several proposals around the use of Cheri capabilities to enforce stack safety,
-ranging from mechanisms that use the standard Cheri hardware (which includes local
-capabilities) \citep{SkorstengaardLocal},
+ranging from mechanisms that use the standard Cheri hardware \citep{Skorstengaard+19b},
 to those that propose entirely new types of capabilities, such as linear
-\citep{SkorstengaardSTK}, uninitialized \citep{Georges+21}, lifetime \citep{Tsampas+19}, and monotonic \citep{GeorgesComplete}.
+(\citep{Skorstengaard+19}), uninitialized (\citep{Georges+21}), and monotonic (\sna{cite}).
 Of these, uninitialized capabilities seem particularly promising---they naturally enforce
@@
 to clear large swaths of the stack. We propose to implement a model of the stack safety
-system described in \citet{Georges+21} and test it against our properties using
+system described in (\citep{Georges+21}) and test it against our properties using
 QuickChick.

### 2021-04-17 tex/main.tex

--- 2021-04-17 tex/main.tex
@@
 callees, etc.). They are also stronger than LSE; even in the absence of the
-control flow guarantees of {\em return to caller}, they guarantee callers'
-data safety in the face of control-flow attacks (see \cref{sec:control}).
+control flow guarantees of WBCF, they guarantee callers'
+data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
 \rb{Reference to \cref{sec:running-example}, where this is discussed now?}

### 2021-04-17 tex/main.tex
wibbles

--- 2021-04-17 tex/main.tex
@@
 combination of an integrity property (``the private
-state in each caller's stack frame is held invariant by the callee'')
+state in each caller's stack frame is held invariant by the callee''),
 a confidentiality property (``the callee's behavior is insensitive to the
@@
 data safety in the face of control-flow attacks (see \cref{sec:control}).
-\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
+\rb{\cref{sec:RTC}?}
 This reflects an arguably cleaner division between data protection and
@@
 property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18}, and extend the generation by
-execution techniques of \citet{TestingNI:ICFP} to handle nested
-properties (\cref{sec:enforcement}).
-%% \bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
+of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement}).
+\bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
 \item We formalize a
@@
 variables).
-We state this as a kind of {\em noninterference} property~\citep{??}.
+We state this as a kind of {\em noninterference} property~\citep{Goguen1982SecurityPA}.
 Noninterference describes knowledge: it states that for system components

### 2021-04-17 tex/main.tex
Rephrase PBT in intro + NITS

@@
 data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
-\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
+%\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
 This reflects an arguably cleaner division between data protection and
@@
 
-For confidentiality and integrity , we consider both {\em stepwise}
+For confidentiality and integrity, we consider both {\em stepwise}
 and {\em observational} variants.  The simpler stepwise
@@
 properties to validate and improve an existing enforcement mechanism, the
-{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
+{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
+in the Coq proof assistant on top of a RISC-V specification.  We
 use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for the Coq proof assistant, to generate many random programs and check
+tool for Coq, to generate many random programs and check
 that Roessler and Dehon's micro-policies correctly abort the ones that
-attempt to violate one of our properties; furthermore, we introduce several
-intentionally broken micro-policies and check that the testing framework is
-able to generate counterexamples that violate our properties but are not
-halted by the incorrect enforcement variants.  (Showing that the testing
-framework detects intentional bugs in enforcement mechanisms increases our
-confidence that it can also catch unknown ones, strengthening our confidence
-in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
-  sentence belongs in the testing section?}\fi)
+attempt to violate one of our properties; furthermore, we
+%
+check that the testing framework is able to generate counterexamples
+that violate our properties but are not halted by incorrect
+enforcement variants (both variants that we accidentally created
+during our re-implementation of the micro-policy and ones that we
+intentionally crafted to be broken in order to increase our confidence
+in testing and the enforcement mechanism itself).
+%introduce several
+%intentionally broken micro-policies and check that the testing framework is
+%able to generate counterexamples that violate our properties but are not
+%halted by the incorrect enforcement variants.  (Showing that the testing
+%framework detects intentional bugs in enforcement mechanisms increases our
+%confidence that it can also catch unknown ones, strengthening our confidence
+%in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
+%  sentence belongs in the testing section?}\fi)
 

### 2021-04-17 tex/main.tex
Pass up to 7

@@
 properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy \sna{RISCy?} ones.  In
+machine-code level for a variety of ISAs, including RISCy \sna{RISCy?}\leo{RISC-like maybe?} ones.  In
 particular, there may be no single ``call'' instruction that does all the
@@
 
-This form of confidentiality we will term {\em stepwise confidentiality}, defined informally
-as: for any state at the entry to a function, for any variant over the values in the stack,
+We will term this form of confidentiality as {\em stepwise confidentiality}. Informally,
+for any state at the entry to a function, for any variant over the values in the stack,
 the primary state and its variant will step in lockstep until they return. To step in lockstep
@@
 ``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
-\bcp{Throughout, we could save space and improve beauty by removing some of the
-vertical space between inference rules.}
+\ifspace{\bcp{Throughout, we could save space and improve beauty by removing some of the
+vertical space between inference rules.}\fi
 
@@
 that matches the topmost return target. If neither is true, the context remains unchanged.
-Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
-\rb{Technically a partial function from addresses to annotations,
-  though its domain can function as a predicate.}
+Recall that a ``code map'' \(\codemap\) is partial function, and we can use its domain
+as a predicate on addresses.
 
@@
 until either both return together or one failstops. This implies that
-if the callee reads a secret location, it must failstop.\apt{right?}
+if the callee reads a secret location, it must failstop.%\apt{right?}\leo{yes}
 %
@@
 the formal stack-safety property described in the previous
-section. (In fact, it enforces a slightly stronger, inductive version
+section. In fact, it enforces a slightly stronger, inductive version
 of that property which checks for integrity and confidentiality violations at
-every step of the callee's execution trace.) We validate this claim
+every step of the callee's execution trace. We validate this claim
 with a property-based testing framework that can quickly uncover
@@
 the machine fail-stops with a policy violation.
-\citeauthor{pump_oakland2015}\bcp{additional citations: the noninterference
-  paper, Nick and Andre's papers, ...} show that a wide range of
+Existing work \citep{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
+%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
+has shown that a wide range of
 micro-policies can be defined using this scheme.
@@
 store operations when the target memory is tagged {\em with the same
-  depth} as the current {\PCname} tag.
+  depth} as the current {\PCname} tag, and store operations also when the
+target memory is $\tagNoDepth$.
 
@@
 confidentiality. We use a Coq specification of the RISC-V
-architecture
-\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
-\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since
-  the MIT people did it}}  and extend it with a
+architecture\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
+%\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}
+and extend it with a
 runtime monitor implementing the stack-safety
-micro-policy.\footnote{This code will be made public. \leo{More?}} We
-  choose the Coq proof assistant as the setting for our implementation
+micro-policy.\footnote{This code is {\em redacted for review}.} We
+  chose the Coq proof assistant as the setting for our implementation
   for two reasons: first, to ensure that our coinductive trace
@@
 annotations are set for the entry or return points. At the same time,
-we sometimes allow the generation to relax those constraints,
+we allow the generation to sometimes relax those constraints,
 introducing potentially ill-formed flows: this causes our programs
@@
 The smallest counterexample requires calling a function {\tt f},
-which writes into an uninitialized location $l$ (which is always fine\apt{is it??? in the strict policy?}, and then
+which writes into an uninitialized location $l$,
+% (which is always fine\apt{is it??? in the strict policy?}
+and then
 calls a further function {\tt g} (which should have the effect of sealing $l$
@@
 $l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
-\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
-  and the write from {\tt f} was already a violation.}
+%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
+%  and the write from {\tt f} was already a violation.}
 

### 2021-04-17 tex/main.tex
Nits

@@
 First, they eliminate the initialization pass. In the {\em Lazy Tagging}
-system, the stack is initially tagged as \(\tagNoDepth\),
-\rb{The previous section already presents ``unused'' as the default tag,
-  though it doesn't say that it can be overwritten whenever\apt{but see my
-    query re the HEADER\_NO\_INIT bug}}\sna{Correct there, but now I'm talking
-  about laziness.}
+system, the stack is again initially tagged as \(\tagNoDepth\),
+%\rb{The previous section already presents ``unused'' as the default tag,
+%  though it doesn't say that it can be overwritten whenever\apt{but see my
+%    query re the HEADER\_NO\_INIT bug}}\sna{Correct there, but now I'm talking
+%  about laziness.}\leo{Should be better now}
 A function is permitted
@@
 cells at the top of the frame, while passing by reference just gives the
-addresses of the appropriate variables.\apt{This still doesn't make sense to me. We cannot
+addresses of the appropriate variables relative to the stack pointer.
+\apt{This still doesn't make sense to me. We cannot
   give a static annotation that identifies the absolute address of a stack-allocated variable!
-  Only sp-relative addresses make sense.}
+  Only sp-relative addresses make sense.}\leo{Better now? :)}
 
@@
 Now we can inductively define the context update function; these rules again apply in the order
-presented. \bcp{Too much space between the second and third.}
+presented. \ifspace\bcp{Too much space between the second and third.}\fi
 
@@
 write-before-read behavior, and therefore will enable confidentiality without the need
-to clear large swaths of the stack. We propose to implement a model of the stack safety
+to clear large swaths of the stack. We plan to implement a model of the stack safety
 system described in \citet{Georges+21} and test it against our properties using

### 2021-04-17 tex/main.tex
comments and queries 1-3

--- 2021-04-17 tex/main.tex
@@
 they write affects the observable behavior of the caller after they return.
-The observational properties are meant to capture the same sense of security,
-extensionally enough that any reasonable protection mechanism ought to enforce
-them, even if it does not prevent every single dangerous read or write.
+The observational properties are more extensional, and represent a baseline 
+% meant to capture the same sense of security, extensionally enough
+that any reasonable protection mechanism ought to enforce,
+even if it does not prevent every single dangerous read or write.
 
@@
 corner cases where data can leak across repeated calls to the same callee,
-and \emph{Lazy Tagging and Clearing} also violates integrity in these instances. We
+and \emph{Lazy Tagging and Clearing} also violates integrity in these instances.\apt{is this
+  right?? I though Lazy tagging was fully ok, and lazy tagging + clearing violated
+  temporal confidentialitty (only).} We
 propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
@@
 
-At first, for ease of exposition, our development assumes a single simple
+For ease of exposition, our development initially assumes a single simple
 stack with no sharing between callers and callees: all parameters and return
@@
 assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter (\(\PCname\)); other
+machine with a program counter (\(\PCname\))\apt{and a stack pointer? 5.2 says we need that too}; other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
@@
 {\em return target} with respect to some set of pending calls.
+\apt{Ideally, I'd still like to get across the idea that, in general, annotations are
+  also needed to connect the policy to
+  to the enforcement mechanism and provide a basis for the ``well-formedness''
+  criterion discussed in 3. We are only able to get away with annotating
+  just calls because our enforcement mechanism doesn't require any cooperation
+  on the callee side.}
 

### 2021-04-17 tex/main.tex
pass through 4

@@
 figures. For our purposes, \(\outside\) will apply to global
-variables. Domains are part of the {\em context} of a state.
+variables.\apt{and code?} Domains are part of the {\em context} of a state.
 We will introduce the remaining context as needed.
@@
 \begin{figure}
-\begin{subfigure}[t]{.2\textwidth}
+\begin{subfigure}[t]{.25\textwidth}
 {\small
@@
 in fact we can learn something about the initial state by observing those results, violating
-confidentiality. \apt{Add some motivation for why we chose to view confidentiality in this way.}
+confidentiality.
 
@@
 {\em system component} (such as memory cell, register, program counter) that changes
-in one step has a matching value at the end of the other step. When we later
-extend the model with a notion of observable output, values that are emitted will need
-to match as well.
+in one step has a matching value at the end of the other step.
+When we later extend the model with a notion of observable output (\cref{sec:lazy}),  we will instead apply the matching
+criteria to the outputs.
 
@@
 \begin{figure}
+\begin{center}
 \begin{subfigure}[t]{.2\textwidth}
@@
 \end{subfigure}
-\begin{subfigure}[t]{.7\textwidth}
+\begin{subfigure}[t]{.68\textwidth}
   {\small
@@
 \end{subfigure}
+\end{center}
 \begin{subfigure}{.65\textwidth}
@@
       \memoryaddrs{12em}
-      \memory{1}{\mainsealc}[Seal(0)]%
+      \memory{1}{\mainsealc}[S(0)]%
       \memory{1}{\unsealc}[RetPtr]%
@@
 Consider Figure~\ref{fig:controlflow}.
-Here we show return addresses explicitly as line numbers, in the form $\#n$.
+(Here the diagram shows return addresses explicitly as line numbers, in the form $\#n$.)
 Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions

### 2021-04-17 tex/main.tex
Working through comments

--- 2021-04-17 tex/main.tex
@@
   right?? I though Lazy tagging was fully ok, and lazy tagging + clearing violated
-  temporal confidentialitty (only).} We
+  temporal confidentialitty (only).}\sna{I was wrong about that. It's true of lazy tagging
+  (provided the caller initializes their frame) but not lt+c}We
 propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
@@
 assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter (\(\PCname\))\apt{and a stack pointer? 5.2 says we need that too}; other
+machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
@@
 properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy \sna{RISCy?}\leo{RISC-like maybe?} ones.  In
+machine-code level for a variety of ISAs, including RISCy
+\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
 particular, there may be no single ``call'' instruction that does all the

### 2021-04-17 tex/main.tex
Comparison with StkTokens property desiderata

@@
 %% observations is limited to cotermination.
+
+Although there are some important differences between
+\citeauthor{SkorstengaardSTK}'s properties and our own, the new
+properties arguably satisfy their list of desiderata for such
+formulations: notably, they reflect (and refine) programmer intuitions
+about stack safety, furnish a comprehensive picture of the meaning of
+stack safety, can be used to reason about program behavior, and are
+flexible enough to be allow extensions to richer settings.
+%
 \sna{Some of this seems like more detail than we need, and we need to
@@
 \rb{Trimmed this a bit. Are you referring to their list of attack
-  examples, or to something else?\apt{Their list at start of section 4.}}
+  examples, or to something else?\apt{Their list at start of section 4.
+  \rb{Had a go at this.}}}
 

### 2021-04-17 tex/main.tex
Annotation connection

--- 2021-04-17 tex/main.tex
@@
 These call annotations would typically come from a compiler together with
-the machine code, but we do to assume anything about their provenance.
+the machine code, but we do not assume anything about their provenance.
+Annotations connect the enforcement mechanism to the property: we don't
+protect unmarked calls, because a mechanism's starting conditions might
+depend on knowing where calls are.
 
@@
 has returned control to the caller when... we see control return to the
-caller!  For example, under a standard calling convention, we can see that a
+caller! We don't require the callee's cooperation at all.
+For example, under a standard calling convention, we can see that a
 return has occurred occurred when state of the machine reaches a state where
@@
 {\em return target} with respect to some set of pending calls.
-\apt{Ideally, I'd still like to get across the idea that, in general, annotations are
-  also needed to connect the policy to
-  to the enforcement mechanism and provide a basis for the ``well-formedness''
-  criterion discussed in 3. We are only able to get away with annotating
-  just calls because our enforcement mechanism doesn't require any cooperation
-  on the callee side.}
 

### 2021-04-17 tex/main.tex
pass through 6

--- 2021-04-17 tex/main.tex
@@
   right?? I though Lazy tagging was fully ok, and lazy tagging + clearing violated
-  temporal confidentialitty (only).}\sna{I was wrong about that. It's true of lazy tagging
-  (provided the caller initializes their frame) but not lt+c}We
+  temporal confidentialitty (only).} We
 propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
@@
 assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
+machine with a program counter (\(\PCname\))\apt{and a stack pointer? 5.2 says we need that too}; other
 details of the machine's ISA are unimportant. We assume that we know ahead of time
@@
 properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy
-\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
+machine-code level for a variety of ISAs, including RISCy \sna{RISCy?}\leo{RISC-like maybe?} ones.  In
 particular, there may be no single ``call'' instruction that does all the
@@
 %
-A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-\Cref{sec:enforcement}.
+%A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
+%\Cref{sec:enforcement}. -- APT: We just said this.
 
@@
     \judgment{\(f ~ \mpcstatename\)}
-             {\(\PUT ~ f ~ \notfinished{\mpcstatename}{\MPCT} = \mpcstatename\)}
+             {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \mpcstatename\)}
   \end{minipage}
@@
     \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
-                {\(\PUT ~ f ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\mpcstatename}{\MPCT'}\)}
+                {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \notfinished{\mpcstatename}{\MPCT'}\)}
   \end{minipage}
@@
 \judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
-            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
+%            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
+            {\(\PUT ~ f ~ \MPCT' = \MPCT\)}
             {\(\mpcstatename \hookrightarrow \MPCT | f\)}
 \end{center}
+\apt{above changed: SNA please check.}
 
@@
 this the {\em sealing convention},
-\(\sealcon ~ (\mach \in \MACHS) ~ (\component \in \COMPONENTS)\), a
-relation between machine states and the addresses that should be sealed.
+%\(\sealcon ~ (\mach \in \MACHS) ~ (\component \in \COMPONENTS)\),
+\(\sealcon \subseteq (\MACHS \times \COMPONENTS)\),
+a relation between machine states and the addresses that should be sealed.
 If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
@@
 it means to return with a {\em return convention},
-\(\retcon ~ (\mach \in \MACHS) ~ (\mach' \in \MACHS)\),
+%\(\retcon ~ (\mach \in \MACHS) ~ (\mach' \in \MACHS)\),
+\(\retcon \subseteq (\MACHS \times \MACHS)\),
 which relates call states to the states that can be considered valid returns to that call.
@@
 at a particular depth. Each caller registers a {\em target}
-condition, \(\target ~ \mach\)---a predicate on machine states that means it
+condition, \(\target \subseteq \MACHS\), which is a predicate on machine states that means it
 has been returned to.
@@
 %
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
 \[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
@@
 that matches the topmost return target. If neither is true, the context remains unchanged.
-Recall that a ``code map'' \(\codemap\) is partial function, and we can use its domain
+Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
 as a predicate on addresses.
@@
               {\(\codemap ~ (\mach ~ \PCname)\)}
-              {\(\components = \sealcon ~ \mach\)}
+%              {\(\components = \sealcon ~ \mach\)}
+              {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
               {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
@@
 for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-\(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\) and the induced trace
-\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
+any \(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\), and any trace
+%\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
+\(\NPCT\) such that \(\nach,\pol,\context \hookrightarrow \NPCT | \brk\),
 \(\MPCT \doteq \NPCT\) holds.
+\apt{changed $dm$ to $ctx$; SNA please check.}
 
@@
 until either both return together or one failstops. This implies that
-if the callee reads a secret location, it must failstop.%\apt{right?}\leo{yes}
+if the callee reads a secret location, it must failstop. %\apt{right?}\leo{yes}
 %
@@
 \(\mpcstatename \stepsto \mpcstatename'\)\apt{doesn't reachability of $\mpcstatename$ imply reachability of $\mpcstatename'$?}, if \(|\rts'| < |\rts|\), then
-\(|\rts'| = |\rts|-1\)
+\(|\rts'| = |\rts|-1\).
 
@@
 
-\definition Let the secret set of a context, \(\secret_P(\domMap,\rts)\),
+\definition Let the \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
 be the set of components \(\component\) such that
@@
 %% observations is limited to cotermination.
-
-Although there are some important differences between
-\citeauthor{SkorstengaardSTK}'s properties and our own, the new
-properties arguably satisfy their list of desiderata for such
-formulations: notably, they reflect (and refine) programmer intuitions
-about stack safety, furnish a comprehensive picture of the meaning of
-stack safety, can be used to reason about program behavior, and are
-flexible enough to be allow extensions to richer settings.
-%
 \sna{Some of this seems like more detail than we need, and we need to
@@
 \rb{Trimmed this a bit. Are you referring to their list of attack
-  examples, or to something else?\apt{Their list at start of section 4.
-  \rb{Had a go at this.}}}
+  examples, or to something else?\apt{Their list at start of section 4.}}
 

### 2021-04-17 tex/main.tex
More tweaks

--- 2021-04-17 tex/main.tex
@@
 to allow passing of scalar stack data, granting a callee access to the
-caller's data.
-%% (We do not consider more sophisticated forms of sharing, such
-%% as taking the address of a variable and treating it like a pointer.)
-%% \rb{But actually, this is discussed at some length is \cref{sec:passing},
-%%   and it is the topic of \cref{fig:passing}.  Reaching that point after
-%%   reading this disclaimer is confusing.}\bcp{So should we just remove that
-%%   sentence?}
-%
+caller's data. Throughout, we note how we expect elements of the formalism
+to interact with enforcement mechanisms beyond micro-policies.
+
 Lastly, we define integrity and confidentiality properties for a coroutine
@@
 
-\apt{Suddenly we are talking about alternative enforcement mechanisms for
-  the first time.  This surely needs \emph{some} foreshadowing in the
-  abstract and intro?}\bcp{Yes!}
-%
 Dividing policy state from machine state helps us reason about
@@
 (outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
-figures. For our purposes, \(\outside\) will apply to global
-variables.\apt{and code?} Domains are part of the {\em context} of a state.
+figures. \sna{TODO}. For our purposes, \(\outside\) will apply to global
+variables and code. Domains are part of the {\em context} of a state.
 We will introduce the remaining context as needed.
@@
 The smallest counterexample requires calling a function {\tt f},
-which writes into an uninitialized location $l$,
-% (which is always fine\apt{is it??? in the strict policy?}
-and then
-calls a further function {\tt g} (which should have the effect of sealing $l$
-for integrity purposes). Then {\tt g} attempts to write to $l$;
-this is allowed because the enforcement mechanism still has
+which fails to initialize some of its frame during the blessed sequence,
+but writes into an uninitialized location $l$ later, treating \(l\) as outside
+the stack. Then {\tt f} calls a further function {\tt g} (which should have
+the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
+which is allowed because the enforcement mechanism still has
 $l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
@@
 First, they eliminate the initialization pass. In the {\em Lazy Tagging}
-system, the stack is again initially tagged as \(\tagNoDepth\),
-%\rb{The previous section already presents ``unused'' as the default tag,
-%  though it doesn't say that it can be overwritten whenever\apt{but see my
-%    query re the HEADER\_NO\_INIT bug}}\sna{Correct there, but now I'm talking
-%  about laziness.}\leo{Should be better now}
+system, the stack is again initially tagged as \(\tagNoDepth\).
 A function is permitted
-to overwrite clear memory, and when it does so its depth tag overwrites
+to overwrite \(\tagNoDepth\) memory, and when it does so its depth tag overwrites
 the clear tag, claiming the memory; the remainder of the policy functions

### 2021-04-17 tex/main.tex
wibble

--- 2021-04-17 tex/main.tex
@@
 
-More interesting (from a stack safety perspective) changes to the tags
+More interesting changes to the tags (from a stack safety perspective) 
 happen at exactly the points in the program where we switch from a

### 2021-04-17 tex/main.tex
pass through 8

@@
 does not match the memory tag; the \(\PCname\) tag is always propagated
-to the written location. Reads from the stack do still require $\PCname$
+to the written location. Reads from the stack \emph{do} require $\PCname$
 tag and memory tag to match. Thus, even if a callee illicitly writes to
@@
 For any \(\nach'\) and \(\NPCT'\) such that \(\mach' \approx_{\components} \nach'\) and 
-\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.
+\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
 
@@
 
-\apt{Given that the deferred property suffices to specifcy the lazy policies,
-  the case for observable properties is pretty weak.  Should we just cut them?}
+%\apt{Given that the deferred property suffices to specifcy the lazy policies,
+%  the case for observable properties is pretty weak.  Should we just cut them?}
 
 While deferred integrity meaningfully captures the intuition
-behind lazy tagging, we seek properties that can be used to evaluate any reasonable
-enforcement mechanism. Consider a hypothetical
+behind lazy tagging, we seek properties that can be used to evaluate the broadest
+possible range of reasonable
+enforcement mechanisms. Consider a hypothetical
 taint-tracking policy that allows high-security to be copied as long as it is
@@
   meaning that no secrets leak during the call
-\item \(\brk ~ \last(\MPCT) \leftrightarrow \brk ~
-\last(\NPCT)\)---if one call returns, so does the other
+\item \(\brk ~ (\last(\MPCT)) \leftrightarrow \brk ~
+(\last(\NPCT))\), i.e. if one call returns, so does the other
 \item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
-  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
-  Let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\).
-  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
+  and \(\pi_\mach(\last(\NPCT)) = \nach'\),
+  let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\),
+  and let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
   Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
@@
 don't actually need a deferred confidentiality property here. Therefore,
-it also enforces observational stack safety, so we declare it stack safe.
+it also enforces observational stack safety, so we declare it stack safe.\apt{Ugh.
+  We haven't set up this definition clearly enough (nor should we try).}
 Our properties have allowed us to reject an enforcement mechanism as

### 2021-04-17 tex/main.tex
Answering notes and simplifying

--- 2021-04-17 tex/main.tex
@@
 \definition
-A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
-\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
-\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
-\(\mpcstatename \stepsto \mpcstatename'\)\apt{doesn't reachability of $\mpcstatename$ imply reachability of $\mpcstatename'$?}, if \(|\rts'| < |\rts|\), then
-\(|\rts'| = |\rts|-1\).
+A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
+\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts))\), and
+\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts'))\),
+if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).
 
@@
 
-More interesting changes to the tags (from a stack safety perspective) 
+More interesting (from a stack safety perspective) changes to the tags
 happen at exactly the points in the program where we switch from a
@@
 
+\ifaftersubmission
 \apt{Given that the deferred property suffices to specifcy the lazy policies,
   the case for observable properties is pretty weak.  Should we just cut them?}
+\fi
 
@@
 A call annotation now includes a predicate relating machine states to the components
-that should be passed---when passing by value, this will typically be freshly allocated
-cells at the top of the frame, while passing by reference just gives the
-addresses of the appropriate variables relative to the stack pointer.
-\apt{This still doesn't make sense to me. We cannot
-  give a static annotation that identifies the absolute address of a stack-allocated variable!
-  Only sp-relative addresses make sense.}\leo{Better now? :)}
+that should be passed---we can always determine where they should be in terms of their
+offsets from the stack pointer, and embed this knowledge in an annotation.
 
@@
 \[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
-%\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-%\times (\rts \in \listT (\target \subseteq \MACHS)) \]
 The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
@@
 \times (\listT (\target)) \]
+The initial context again maps all components to \(\unsealed\).
 We straightforwardly describe how the sealing convention and passing predicate combine to
@@
 
-\definition
-For each of the stepwise, deferred, and observational versions of
-stack integrity in the simple model, the equivalent property in the
-passing model substitutes the above definitions of \(\context\), initial contexts\apt{not defined above??}  and update
-function, and \(\prot_P\) for their simple equivalents, and is otherwise identical.
-
 \definition Let the \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
@@
 \definition
+For each of the stepwise, deferred, and observational versions of
+stack integrity in the simple model, the equivalent property in the
+passing model substitutes the above \(\CONTEXTS\) type and update
+function, and \(\prot_P\), for their equivalents in the simple model.
+
 For both the stepwise and observational versions of stack confidentiality in
-the original model, the equivalent property in the passing model is the same after
-substituting the above definitions of \(\context\), initial contexts\apt{ditto} and update
-function, and \(\prot_P\) for their simple equivalents.
+the original model, the passing model equivalents additionally substitute \(\secret_P\).
 

### 2021-04-17 tex/main.tex

@@
 
-\ifaftersubmission
-\apt{Given that the deferred property suffices to specifcy the lazy policies,
-  the case for observable properties is pretty weak.  Should we just cut them?}
-\fi
+%\apt{Given that the deferred property suffices to specifcy the lazy policies,
+%  the case for observable properties is pretty weak.  Should we just cut them?}
 
 While deferred integrity meaningfully captures the intuition
-behind lazy tagging, we seek properties that can be used to evaluate any reasonable
-enforcement mechanism. Consider a hypothetical
+behind lazy tagging, we seek properties that can be used to evaluate the broadest
+possible range of reasonable
+enforcement mechanisms. Consider a hypothetical
 taint-tracking policy that allows high-security to be copied as long as it is

### 2021-04-17 tex/main.tex
through 9

--- 2021-04-17 tex/main.tex
@@
 But what about more realistic systems? We now demonstrate the
-flexibility of these core concepts as we extend
-the core concepts of stack safety to a system that models the passing of
+flexibility of our core concepts by extending 
+them to a system that models the passing of
 parameters on the stack, and then to a very simple coroutine model.
@@
   give a static annotation that identifies the absolute address of a stack-allocated variable!
-  Only sp-relative addresses make sense.}\leo{Better now? :)}
+  Only sp-relative addresses make sense.}\leo{Better now? :)}\apt{no: the definition of $P$ is still
+  bogus. It needs to be something like a relation between  $M$ and a set of SP offsets.}
 

### 2021-04-17 tex/main.tex

--- 2021-04-17 tex/main.tex
@@
 A call annotation now includes a predicate relating machine states to the components
-that should be passed---we can always determine where they should be in terms of their
-offsets from the stack pointer, and embed this knowledge in an annotation.
+that should be passed, expressed as offsets from the stack pointer. \apt{Much better, but the
+actual definition of $P$ is still wrong.}
 

### 2021-04-17 tex/main.tex
Cleaning

--- 2021-04-17 tex/main.tex
@@
 Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids. Note that we now explicitly reference the stack
-pointer, using the \(\stackof\) of the stack pointer to determine the target of a yield.
-We add a second annotation indicating yields.\apt{seems unchanged from (old version of) previous section}
+\(\stackof\) from addresses to stack ids.
 
@@
 an additional map of targets for yields, and a stack identifier of the active stack.
-Our yield target map \(\yts\) is a map from stack identifiers to targets.
+A yield target map \(\yts\) is a map from stack identifiers to targets.
 In a standard system an initial yield target map will map each stack identifier to the
@@
 (\STACKS \rightarrow \listT(\target)) \times (\STACKS \rightarrow \target) \times \STACKS\]
+
+The initial domain map maps each stack address in stack \(\stackid\) to
+\(\instack{\stackid}{\unsealed}\), the initial return map maps all stack identifiers to empty
+return stacks, the initial yield targets are defined above, and the initial stack id matches the initial
+coroutine. Together these form the initial context.
 
@@
 \definition
-We define {\em stepwise coroutine integrity} in the same way as stepwise stack integrity,
-substituting the above definitions of \(\context\), initial contexts\apt{??} and update
+We define \textbf{\em stepwise coroutine integrity} in the same way as stepwise stack integrity,
+substituting the above \(\CONTEXTS\) type, initial context and update
 function, and \(\prot_Y\) for their equivalents. Likewise
-{\em deferred coroutine integrity} and {\em observational coroutine integrity} are defined
+\textbf{\em deferred coroutine integrity} and \textbf{\em observational coroutine integrity} are defined
 in terms of their simple equivalents using \(\MPCT^\wedge\) in place of \(\MPCT^*\).
@@
 \definition
-We define {\em stepwise coroutine integrity} in the same way as stepwise stack
+We define \textbf{\em stepwise coroutine confidentiality} in the same way as stepwise stack
 integrity, substituting the above definitions. Likewise
-{\em observational coroutine confidentiality} is defined in terms of observational
+\textbf{\em observational coroutine confidentiality} is defined in terms of observational
 stack confidentiality.

### 2021-04-17 tex/main.tex
10 wip

--- 2021-04-17 tex/main.tex
@@
 convention that uses local capabilities to preserve local state
-encapsulation and well-bracketed control flow, using a logical
-relation to reason about the stack safety of concrete
-programs. Although the required hardware support is readily available
-in capability machines like Cheri,
+encapsulation and well-bracketed control flow
+using a logical relation to reason about the stack safety of concrete
+programs.
+% Although the required hardware support is readily available
+%in capability machines like Cheri,
 %% this technique incurs significant
 %% costs, because
-it requires the entire unused part of the stack to be
-cleared whenever a security boundary is crossed. Their logical
+%it requires the entire unused part of the stack to be
+%cleared whenever a security boundary is crossed.
+Their logical
 relation captures capability safety without ``externally observable
@@
 and can be used to reason about individual programs.
-\citeauthor{SkorstengaardSTK} acknowledge the difficulty of defining
+\citet{SkorstengaardSTK} acknowledge the difficulty of defining
 these concepts precisely:

### 2021-04-17 tex/main.tex

--- 2021-04-17 tex/main.tex
@@
 Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids.
-
-\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
+\(\stackof\) from addresses to stack ids.  Annotations and code maps are unchanged from
+the subroutine model.
+%\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
+%\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
 
@@
 This calls for some explanation. First, we determine which coroutine we are yielding to by
-checking which stack the stack pointer of the next state is pointing to. Then we can only yield
+kchecking which stack the stack pointer of the next state is pointing to. Then we can only yield
 if the next state is a valid yield target for that stack, and we update the yield target for

### 2021-04-17 tex/main.tex
Fixed passing predicate

--- 2021-04-17 tex/main.tex
@@
   \memory{1}{\retptrc}
-  \memory{1}{\goodc}
-  \memory{1}{\unsealc}
+  \memory{2}{\unsealc}
   ~$\cdots$
@@
 A call annotation now includes a predicate relating machine states to the components
-that should be passed, expressed as offsets from the stack pointer. \apt{Much better, but the
-actual definition of $P$ is still wrong.}
-
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P) \hspace*{1in} P \subseteq (\MACHS \times \COMPONENTS)\]
+that should be passed, expressed as offsets from the stack pointer.
+%
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \WORDS)\]
+
 We extend the set of domains to include passed components,
@@
   \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
+                                 \sealcon ~ \mach ~ \component \text{ and }
+                                 \component - (\mach ~ \rsp) \not \in P \\
   \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
+                                 \sealcon ~ \mach ~ \component \text{ and }
+                                 \component - (\mach ~ \rsp) \in P \\
   \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
-                                 \text{ for any } \depth' \\
+                                 \sealcon ~ \mach ~ \component \\
+                               & \text{ and } (\mach ~ \rsp) \in P, \text{ for any } \depth' \\
   \domMap~\component & \text{else} \\
@@
 At the matching return we just unseal the components at depth \(\depth\):
-
+%
 \[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
@@
 mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
-From here there are two natural ways to extend the work. We should thoroughly evaluate
+Our immediate next step is to implement testing of observational properties,
+both to validate them and to explore how to effectively test properties that do not
+fail immediately.
+From there there are two natural ways to extend the work. We should thoroughly evaluate
 other mechanisms, especially Cheri-based capability systems. And we can further demonstrate

### 2021-04-17 tex/main.tex
Dealing with lazy policies

@@
 then permitted only when that activation is currently executing, validates our
-stepwise properties.  On the other hand, testing reveals their more efficient
-\emph{Lazy Tagging} policy does not enforce step-wise integrity. In fact
-we reason that both of their lazy policies violate the temporal aspect of confidentiality in
+stepwise properties. On the other hand, we reason that \emph{Lazy Tagging and Clearing}
+violates the temporal aspect of confidentiality in
 corner cases where data can leak across repeated calls to the same callee,
-and \emph{Lazy Tagging and Clearing} also violates integrity in these instances.\apt{is this
-  right?? I though Lazy tagging was fully ok, and lazy tagging + clearing violated
-  temporal confidentialitty (only).}\sna{I was wrong about that. It's true of lazy tagging
-  (provided the caller initializes their frame) but not lt+c}We
+and also violates integrity if the leak happens to use the caller's frame. We
 propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
@@
 
-{\em Lazy Per-activation Tagging and Clearing} ends up enforcing the combination
-of deferred stepwise stack integrity and stepwise confidentiality---we
-don't actually need a deferred confidentiality property here. Therefore,
-it also enforces observational stack safety, so we declare it stack safe.\apt{Ugh.
-  We haven't set up this definition clearly enough (nor should we try).}
+We believe that {\em Lazy Per-activation Tagging and Clearing}, in which
+each activation is tagged uniquely, ends up enforcing the combination
+of deferred stepwise stack integrity and stepwise confidentiality.
+This is because tags that persist between calls will not become active again.
 Our properties have allowed us to reject an enforcement mechanism as
-not really stack safe, and describe a fix.
+not really stack safe, and describe a potential fix that itself is a
+prime candidate for future testing.
 

### 2021-04-17 tex/main.tex
Checked
Todid
further stab at Stktokens criteria comparison

--- 2021-04-17 tex/main.tex
@@
   \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
+  \memory{1}{\goodc}
+  \memory{1}{\unsealc}
   ~$\cdots$
@@
 A call annotation now includes a predicate relating machine states to the components
-that should be passed, expressed as offsets from the stack pointer.
-%
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \WORDS)\]
-
+that should be passed, expressed as offsets from the stack pointer. \apt{Much better, but the
+actual definition of $P$ is still wrong.}
+
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P) \hspace*{1in} P \subseteq (\MACHS \times \COMPONENTS)\]
 We extend the set of domains to include passed components,
@@
   \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and }
-                                 \component - (\mach ~ \rsp) \not \in P \\
+                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
   \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and }
-                                 \component - (\mach ~ \rsp) \in P \\
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
   \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \\
-                               & \text{ and } (\mach ~ \rsp) \in P, \text{ for any } \depth' \\
+                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
+                                 \text{ for any } \depth' \\
   \domMap~\component & \text{else} \\
@@
 At the matching return we just unseal the components at depth \(\depth\):
-%
+
 \[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
@@
 
-Although there are some important differences between
-\citeauthor{SkorstengaardSTK}'s properties and our own, the new
-properties arguably satisfy their list of desiderata for such
-formulations: notably, they reflect (and refine) programmer intuitions
-about stack safety, furnish a comprehensive picture of the meaning of
-stack safety, can be used to reason about program behavior, and are
-flexible enough to be allow extensions to richer settings.
-%
-\sna{Some of this seems like more detail than we need, and we need to
-  compare ourselves to their list of criteria.}
-\rb{Trimmed this a bit. Are you referring to their list of attack
-  examples, or to something else?\apt{Their list at start of section 4.
-  \rb{Had a go at this.}}}
+\citet{SkorstengaardSTK} enumerate a useful list of criteria for judging
+proposed stack safety properties: they should be (1) intuitive; (2)
+useful for reasoning; (3) reusable in secure compiler chains; (4)
+arguably ``complete;'' and (5) potentially scalable.  We believe our
+approach competes well with theirs on criteria (1), (4), and (5).
+Their approach is specifically set up to address (3), while
+ours is not. Addressing (2) remains as future work for us.
+
+%% Although there are some important differences between
+%% \citeauthor{SkorstengaardSTK}'s properties and our own, the new
+%% properties arguably satisfy their list of desiderata for such
+%% formulations: notably, they reflect (and refine) programmer intuitions
+%% about stack safety, furnish a comprehensive picture of the meaning of
+%% stack safety, can be used to reason about program behavior, and are
+%% flexible enough to be allow extensions to richer settings.
+%% %
+%% \sna{Some of this seems like more detail than we need, and we need to
+%%   compare ourselves to their list of criteria.}
+%% \rb{Trimmed this a bit. Are you referring to their list of attack
+%%   examples, or to something else?\apt{Their list at start of section 4.
+%%   \rb{Had a go at this.}}}
 
@@
 mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
-Our immediate next step is to implement testing of observational properties,
-both to validate them and to explore how to effectively test properties that do not
-fail immediately.
-From there there are two natural ways to extend the work. We should thoroughly evaluate
+From here there are two natural ways to extend the work. We should thoroughly evaluate
 other mechanisms, especially Cheri-based capability systems. And we can further demonstrate

### 2021-04-17 tex/main.tex

--- 2021-04-17 tex/main.tex
@@
 \judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
-%            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
             {\(\PUT ~ f ~ \MPCT' = \MPCT\)}
@@
 \end{center}
-\apt{above changed: SNA please check.}
 
@@
 \(\MPCT \doteq \NPCT\) holds.
-\apt{changed $dm$ to $ctx$; SNA please check.}
 
@@
   \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[Seal(0)]%
+  \memory{1}{\mainsealc}[S(0)]%
   \memory{1}{\mainpassc}[P(0,1)]%
@@
       \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{S(0)}}]%
       \memory{1}{\retptrc}
@@
       \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{S(0)}}]%
       \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
@@
 
-Although there are some important differences between
-\citeauthor{SkorstengaardSTK}'s properties and our own, the new
-properties arguably satisfy their list of desiderata for such
-formulations: notably, they reflect (and refine) programmer intuitions
-about stack safety, furnish a comprehensive picture of the meaning of
-stack safety, can be used to reason about program behavior, and are
-flexible enough to be allow extensions to richer settings.
-%
-\sna{Some of this seems like more detail than we need, and we need to
-  compare ourselves to their list of criteria.}
-\rb{Trimmed this a bit. Are you referring to their list of attack
-  examples, or to something else?\apt{Their list at start of section 4.
-  \rb{Had a go at this.}}}
+\citet{SkorstengaardSTK} enumerate a useful list of criteria for judging
+proposed stack safety properties: they should be (1) intuitive; (2)
+useful for reasoning; (3) reusable in secure compiler chains; (4)
+arguably ``complete;'' and (5) potentially scalable.  We believe our
+approach competes well with theirs on criteria (1), (4), and (5).
+Their approach is specifically set up to address (3), while
+ours is not. Addressing (2) remains as future work for us.\apt{OK?}
+
+%% Although there are some important differences between
+%% \citeauthor{SkorstengaardSTK}'s properties and our own, the new
+%% properties arguably satisfy their list of desiderata for such
+%% formulations: notably, they reflect (and refine) programmer intuitions
+%% about stack safety, furnish a comprehensive picture of the meaning of
+%% stack safety, can be used to reason about program behavior, and are
+%% flexible enough to be allow extensions to richer settings.
+%% %
+%% \sna{Some of this seems like more detail than we need, and we need to
+%%   compare ourselves to their list of criteria.}
+%% \rb{Trimmed this a bit. Are you referring to their list of attack
+%%   examples, or to something else?\apt{Their list at start of section 4.
+%%   \rb{Had a go at this.}}}
 

### 2021-04-17 tex/main.tex
Squeeze a little
Final

@@
 (outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
-figures. \sna{TODO}. For our purposes, \(\outside\) will apply to global
+figures. For our purposes, \(\outside\) will apply to global
 variables and code. Domains are part of the {\em context} of a state.
@@
 \end{subfigure}
-\caption{A confidentiality violation\apt{Could save space by not repeating C or assemly code for {\tt main}.}}
+\caption{A confidentiality violation\ifspace\apt{Could save space by not repeating C or assemly code for {\tt main}.}\fi}
 \label{fig:conf1}
@@
 \vspace*{-1ex}
-\judgmentthree[ReturnRec]
+\judgmentthree[ ReturnRec]
               {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
@@
   for two reasons: first, to ensure that our coinductive trace
-  definitions are well-formed, and reason about them\leo{Do we prove
-    anything interesting anymore?}, and to leverage the power of the
+  definitions are well-formed, and reason about them, and to leverage the power of the
   QuickChick property-based testing framework~\citep{Pierce:SF4}.
@@
 machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
-RAM. \leo{Something like: There is really no baseline to compare this
-  against, but include it here for the future/reasons?}
+RAM.
+% \leo{Something like: There is really no baseline to compare this
+  %against, but include it here for the future/reasons?}
 %
@@
 
-\judgmenttwo[Call]
+\judgmenttwo[ Call]
             {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
@@
 
-\vspace*{-3.5ex}
+\vspace*{-5ex}
 \judgment[ Default]
@@
 coroutine.
-
+%
 \[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
@@
 Their approach is specifically set up to address (3), while
-ours is not. Addressing (2) remains as future work for us.\apt{OK?}
+ours is not. Addressing (2) remains as future work for us.
 

### 2021-04-23 submission2/example2.tex
Moved tex/ to submission2/

--- 2021-03-03 example2.tex
@@
 int main() {
-  long int x = 42, y = 0;
-  print &y;
-  y = f(x);
+  int x = 42, y = 0;
+  swap(&x,&y);
   return y;
@@
 \begin{verbatim}
-long int f(long int z) {
+void swap(int* a, int* b) {
+  int tmp = *a;
+  *a = *b;
+  *b = tmp;
   print g();
@@
 
-\colorlet{lgray}{gray!40}
-
-\newcommand{\instrc}{lgray}
-\newcommand{\mainsealc}{cyan}
 \newcommand{\mainpassc}{magenta}
-\newcommand{\fsealc}{green}
 \newcommand{\sharec}{lgray}
-\newcommand{\unsealc}{gray}
-\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
-\newcommand{\fulloutc}{white}
 

### 2021-05-10 short/main.tex
Added short paper draft

--- 2021-04-17 tex/main.tex
@@
 %% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
+%\acmJournal{PACMPL}
+%\acmVolume{1}
+%\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
+%\acmArticle{1}
+%\acmYear{2021}
+%\acmMonth{1}
+%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+%\startPage{1}
 
@@
 \input{macros}
-\input{examples}
-\input{example1}
-
-\newcommand*{\rsp}{\textsc{sp}}
-
-\newcommand{\TAGS}{\mathcal{T}}
-\newcommand{\tagname}{t}
-\newcommand{\uP}{{\mu P}}
 
@@
 %% Title information
-\title{Stack Safety is a Security Property}         %% [Short Title] is optional;
+\title{Security Properties for Stack Safety}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 %% Author with single affiliation.
-%\author{First1 Last1}
+%\author{Sean Noble Anderson}
 %\authornote{with author1 note}          %% \authornote is optional;
@@
 %  \position{Position1}
-%  \department{Department1}              %% \department is recommended
-%  \institution{Institution1}            %% \institution is required
+%  \department{Computer Science}              %% \department is recommended
+%  \institution{Portland State University}            %% \institution is required
 %  \streetaddress{Street1 Address1}
@@
 %}
-%\email{first1.last1@inst1.edu}          %% \email is recommended
+%\email{ander28@pdx.edu}          %% \email is recommended
+
+%\author{Leonidas Lampropoulos}
+%\affiliation{
+%  \institution{University of Maryland, College Park}
+%}
+%\email{leonidas@umd.edu}
+
+%\author{Roberto Blanco}
+%\affiliation{
+%  \institution{Max Planck Institute for Security and Privacy}
+%}
+%\email{roberto.blanco@mpi-sp.org}
+
+%\author{Benjamin C. Pierce}
+%\affiliation{
+%  \institution{University of Pennsylvania}
+%}
+%\email{bcpierce@cis.upenn.edu}
+
+%\author{Andrew Tolmach}
+%\affiliation{
+%  \institution{Portland State University}
+%}
+%\email{tolmach@pdx.edu}
+
 %
@@
 %% before \maketitle command
-\begin{abstract}
-%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
-%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
-%  to leave it out of the title but hit it strongly in the abstract.}%
-What exactly does ``stack safety'' mean? The phrase is associated with a
-variety of compiler,
-run-time, and hardware mechanisms for protecting stack
-memory.  But these mechanisms typically lack precise specifications,
-relying instead on informal descriptions and examples of bad
-behaviors that they prevent.
+%\begin{abstract}
+%What exactly does ``stack safety'' mean? The phrase is associated with a
+%variety of compiler,
+%run-time, and hardware mechanisms for protecting stack
+%memory.  But these mechanisms typically lack precise specifications,
+%relying instead on informal descriptions and examples of bad
+%behaviors that they prevent.
 
-We propose a formal characterization
-of stack safety, formulated with concepts from language-based security: a
-combination of an integrity property (``the private
-state in each caller's stack frame is held invariant by the callee''),
-a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state''), and a well-bracketedness property (``each
-callee returns control to its immediate caller'').
-% These properties are sensible even in the presence of
-% abnormal control flow.\apt{Is this important enough to mention here?}
-% BCP: I think not in the abstract, but maybe in the intro.
-% We further apply\apt{what does this mean? we didn't
-%  invent it} a straightforward notion of
-% ``well-bracketed control flow'' that restricts control flow to the
-% typical pattern of matched calls and returns.
-%
-We use these properties to validate the stack-safety ``micro-policies''
-proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
-property-based random testing) that Roessler and Dehon's ``eager''
-micro-policy, which catches violations as early as possible, enforces a
-simple ``stepwise'' variant of our properties and correctly detects several
-broken variants, and that (a repaired version of) their more performant
-``lazy'' micro-policy corresponds to a slightly weaker and more extensional
-``observational'' variant of our properties.
-%
-% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
-%   still not really clear yet on whether our definition applies / might apply
-%   to other enforcement mechanisms or not.  Saying that our attacker model is
-%   stronger does not help---that should make it easier for us to describe
-%   their protections, not harder.  So I don't understand what we're saying
-%   here.}%
-% %
-% While we do not rule out other enforcement mechanisms, our
-% definitions are strongly informed by micropolicy systems, especially
-% our attacker model, which is stronger than other enforcement mechanisms
-% can enable.
+%We propose a formal characterization
+%of stack safety, formulated with concepts from language-based security: a
+%combination of an integrity property (``the private
+%state in each caller's stack frame is held invariant by the callee''),
+%a confidentiality property (``the callee's behavior is insensitive to the
+%caller's private state''), and a well-bracketedness property (``each
+%callee returns control to its immediate caller'').
 
-% For the sake of concreteness, most of our presentation assumes a standard RISC
-% machine architecture with a basic subroutine calling convention (parameters
-% passed in registers) and a micropolicy-like enforcement mechanism. We
-% also show how to accomodate stack-allocated parameter passing and
-% coroutines.
+%We use these properties to validate the stack-safety ``micro-policies''
+%proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
+%property-based random testing) that Roessler and Dehon's ``eager''
+%micro-policy, which catches violations as early as possible, enforces a
+%simple ``stepwise'' variant of our properties and correctly detects several
+%broken variants, and that (a repaired version of) their more performant
+%``lazy'' micro-policy corresponds to a slightly weaker and more extensional
+%``observational'' variant of our properties.
 
-% \bcp{this is also a
-%   good place to admit that our definitions are tuned for micropolicies, if
-%   we can find a succinct way to say it.}
-% \sna{See above.}
-% \BCP{I added it to the last paragraph (which is pretty close to the preceding
-%   paragraph :-)  We should follow this up with some discussion in the (new)
-%   Assumptions section about how different enforcement mechanisms (us,
-%   StkTokens, maybe others) adopt different perspectives on what the stack is
-% for and how it is organized and so lead to somewhat different notions of
-% low-level safety.}
-
-% We argue \apt{do we?}\bcp{I guess we will not have time to.  Propose
-%   deleting this para.} that our properties are suited to a range of other
-% enforcement mechanisms. The strict stepwise properties can describe
-% mechanisms that catch errors immediately, as is typical. While the
-% observational properties are closely linked to lazy micro-policies they are
-% weak enough to be a lower bound that captures stack safety for any
-% reasonable enforcement mechanism.
-
-% \apt{This is a good story, but does it describe the paper we are able to produce this week?
-%   We were rejected before exactly because we claimed to have
-%   a general criterion, but did not in fact demonstrate its generality (which is not
-%   obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
-%   para about the general story.  Does the rest now avoid making (or
-%   implying) claims that we can't defend?}
-% \bcp{Also, I think we should save
-%   WBCF for the intro---I can't find a way to discuss it here that is brief
-%   and not confusing/distracting.}
-
-  % \iftrue\bcp{should we mention stktokens here?}
-  % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
-  % \bcp{Some people say not to, but I think their argument is weak: I've
-  %   added Nick/Andre above.  But I'm not sure we want to get into stkTokens
-  %   in the abstract.}
-  % \fi
-\end{abstract}
+%\end{abstract}
 
@@
 
-\section{Introduction}
-
-\newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}
-\newcommand{\memoryaddrs}[2][]
-  {
-    \MemoryLabel{#2}{2em}{$\downarrow$\SP #1}
-  }
-
-%\begin{figure}
-%  \begin{minipage}{\textwidth}
-%    \begin{center}
-%\MemoryLabel{10em}{1.5em}{40}
-%\MemoryLabel{16em}{1.5em}{60}
-%\MemoryLabel{22em}{1.5em}{80}
-%$\cdots$
-%\memory{4}{gray}%
-%\memory{5}{green}[{\makebox[0pt]{Caller}}]%
-%\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
-%\memory{8}{red}[{\makebox[0pt]{Callee}}]%
-%\memory{4}{\unusedc}%
-%~$\cdots$\\
-%  \end{center}
-%  \end{minipage}\\
-%~\\
-%~\\
-%  \begin{minipage}{0.45\textwidth}
-%    \begin{center}
-%\begin{verbatim}
-%jal ...
-%lw r1 r0 40
-%\end{verbatim}
-%    \end{center}
-%  \end{minipage}~
-%  \begin{minipage}{0.45\textwidth}
-%    \begin{center}
-%\begin{verbatim}
-%jal ...
-%sw r0 r0 40
-%\end{verbatim}
-%    \end{center}
-%  \end{minipage}
-%  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
-%      the boxes; the program fragments need comments explaining what each
-%      line does.}}
-%  \label{fig:stackunsafety}
-%\end{figure}
-%
-%\leo{I hate the writing here, but I wanted to make a start.}
-
-%\bcp{We don't actually say very clearly anywhere that we are interested in
-%  low-level enforcement mechanisms for machine-code programs (because this
-%  is where stack attacks occur, and because at higher levels of abstraction
-%  the stack is a built-in concept, not something that can be attacked even
-%  in principle).  This could cause confusion.\apt{not so worried}}
-%\leo{I agree. I've clarified the writing in the top of the running example
-%  a bit to make sure though}
+%\section{Introduction}
 
@@
 capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\bcp{Yes, going back to MIT days---we should
-  include several more of these, if only to give readers the impression that
-this is a well-studied mechanism (so formalizing its protections is
-useful).}
-\fi
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}.
 %
 The protections offered by such mechanisms are commonly described in terms
-of concrete examples of attacks that they can prevent---corruption
-of return addresses, buffer overflows, use of uninitialized variables,
-etc.---leaving a more abstract characterization to the reader's intuition.
-But these mechanisms can be quite intricate,
-and they often require trading off protection and efficiency.
-Thus, it would be useful to have a precise, generic, and formal
-specification for stack
-safety, both as a basis for comparing the security claims of different
-enforcement techniques and for validating more rigorously that these claims
-are met by particular implementations.
+of concrete examples of attacks that they can prevent.
+At best, they define stack safety by reference to an idealized machine that
+is arguably stack safe by construction \citep{SkorstengaardSTK}.
+But these mechanisms can be intricate, and it would be useful to have a precise, generic, and formal
+specification for stack safety, both to compare the security claims of different
+enforcement techniques and to rigorously validate such claims.
 
-%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
-%with the notion usually being
-%defined negatively: through examples of stack-based exploits
-%\rb{usually though not always; as noted below there are other
-%  approaches}\bcp{Yes, we have to be very careful about this}.
-%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
-The most abstract characterization of stack safety that we are aware of is
-recent work by \citet{SkorstengaardSTK}, which defines stack safety as the conjunction of two
-properties:
-{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
-Informally, LSE says that the contents of
-a caller's stack frame are not read or written while its callees (and their callees, etc.) are
-executing, and WBCF says that callees always
-return to the instruction following their call (if they return at all).
-Skorstengaard et al.{} formalize these properties by defining an idealized
-machine in which they hold {\em by construction};
-they then show that an enforcement
-mechanism (in their case, hardware capabilities) guarantees these properties, by
-proving that an ordinary machine with enforcement simulates the idealized
-machine.
-%an enforcement mechanism (in their case, hardware
-%capabilities)
-%can then be shown to guarantee stack safety by exhibiting a fully abstract
-%embedding function
-%from the idealized machine to an ordinary machine equipped with this
-%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
-%  understand that sentence.}
-%%
-%% While this work represents an important step in the right direction, we believe
-%% there is still room to craft an operational characterization
-%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
-%% In other words, we seek formal criteria that can be applied
-%% to a mechanism for protecting stack-based low-level programs to judge
-%% whether or not it actually ``guarantees stack safety.''
-%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
-%% confused by the references to these types of properties, yet to be introduced?}
-%% We begin with a strong
-%% local state encapsulation property that closely models the behavior of eager
-%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
-%% is too strong to characterize stack safety; reasonable enforcement mechanisms
-%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
-%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
-%% to support this technique and argue that the new property, defined relative to an
-%% arbitrary notion of observation. This property is still strong enough to prevent
-%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
-%% other reasonable enforcement mechanisms can enforce it.
+We propose such a characterization,
+using the technical framework of language-based security: a
+combination of an {\em integrity property},
+a {\em confidentiality property}, and a {\em well-bracketedness property}.
+%
+In a setting containing unstructured machine code, true
+control-flow attacks are only meaningful when they bypass a caller that was
+waiting for its return, so well-bracketness guarantees that callees return
+only to their callers. Other control-flow attacks, like
+return-oriented-programming (ROP), appear to return but keep
+the callee in control while accessing the caller's data.
+%
+Our properties treat this scenario as the callee not returning at all,
+so if the ROP attack touches the caller's data, it violates integrity
+or confidentiality.
 
-In the present work, we also describe the protection of both stack data (like LSE) and
-control flow (like WBCF).  Our technical formulation of WBCF is similar to
-theirs, but our treatment of stack data protection is rather different,
-as is our approach to the definition itself. Instead of defining stack safety implicitly,
-as ``what a certain idealized machine does,'' we give an
-\emph{explicit} definition using the technical framework of language-based
-security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
-  survey?}\fi, phrased as the conjunction of
-\emph{stack integrity} and \emph{stack confidentiality}.
-%
-A key technical novelty in these definitions,
-compared to standard formulations of confidentiality and integrity from the
-security literature, is that they
-are ``nested'': {\em each} caller is guaranteed protection from its
-immediate callees (which, in turn, need protection from their immediate
-callees, etc.). They are also stronger than LSE; even in the absence of the
-control flow guarantees of WBCF, they guarantee callers'
-data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
-%\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
-This reflects an arguably cleaner division between data protection and
-control-flow protection than LSE / WBCF.
-
-For confidentiality and integrity, we consider both {\em stepwise}
-and {\em observational} variants.  The simpler stepwise
-variants formalize the intuition that a caller's local data is {\em never}
-read or modified during a call. The observational variants
-allow a callee to read from and write to their caller's stack frame, as
-long as nothing that they read affects their observable behavior and nothing
-they write affects the observable behavior of the caller after they return.
-The observational properties are more extensional, and represent a baseline 
-% meant to capture the same sense of security, extensionally enough
-that any reasonable protection mechanism ought to enforce,
+Integrity and confidentiality are also novel in that they
+are ``nested'': {\em each} caller is protected from its
+callees. They come in {\em stepwise} variants, in
+which a caller's data is {\em never}
+read or modified during a call, and {\em observational} ones, in which
+callees may read from and write to their caller's stack frame, as
+long as these ``risky'' behaviors do not affect the system's observable
+behavior. The observational properties are more extensional, and
+any reasonable protection mechanism ought to enforce them,
 even if it does not prevent every single dangerous read or write.
 
-% {\em Confidentiality} says that a caller's
-% data is not leaked into its callee, and  that the callee cannot see uninitialized data.
-% As a reasoning principle, confidentiality guarantees that the callee's
-% behavior is determined only by a predictable interface (function arguments and
-% register contents) that does not include the caller's
-% local state or the uninitialized portion of the stack. This is formalized as a
-% {\em noninterference}-style property that, rather than comparing pairs of whole program traces,
-% compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
-% again
-% {\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
-% changes in the state are independent of the caller's data.
-
-% Together, these properties capture roughly the same intuition as LSE, but
-% they are somewhat stronger.
-% First, we model returns as returns to a valid return target\bcp{that will
-%   make no sense to the reader at this point}; an invalid return
-% is treated as a mere jump within the callee, and does not grant access to the caller's
-% data. So, even without WBCF, our stack safety properties rule out many control-flow attacks.
-% The attacker in such an instance is permitted to ``return'' so long as it avoids
-% violating the integrity and confidentiality of the caller.\bcp{ditto} Second, confidentiality
-% contains\bcp{?} temporal protections that LSE does not necessarily address\bcp{?}. The intuition
-% for LSE is that it protects the caller's stack frame, not necessarily memory that was
-% previously allocated but now outside of the caller's frame. (Cheri-capability-based
-% implementations of LSE {\em do} protect previously allocated data, typically by clearing
-% it, but because it is a potential vector for leaking capabilities into the caller's frame,
-% not as a goal of the property.\bcp{Oof.  We need to do better than that!})
-
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
+To demonstrate the utility of our properties, we use them
+to validate an existing mechanism, the
 {\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
@@
 use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate many random programs and check
-that Roessler and Dehon's micro-policies correctly abort the ones that
-attempt to violate one of our properties; furthermore, we
-%
-check that the testing framework is able to generate counterexamples
-that violate our properties but are not halted by incorrect
-enforcement variants (both variants that we accidentally created
-during our re-implementation of the micro-policy and ones that we
-intentionally crafted to be broken in order to increase our confidence
-in testing and the enforcement mechanism itself).
-%introduce several
-%intentionally broken micro-policies and check that the testing framework is
-%able to generate counterexamples that violate our properties but are not
-%halted by the incorrect enforcement variants.  (Showing that the testing
-%framework detects intentional bugs in enforcement mechanisms increases our
-%confidence that it can also catch unknown ones, strengthening our confidence
-%in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
-%  sentence belongs in the testing section?}\fi)
+tool for Coq, to generate random programs and check
+that these micro-policies correctly abort programs that
+would violate stack safety.
 
@@
 confidentiality, albeit at some performance cost.
-%
-\ifaftersubmission
-\bcp{Did Nick and Andre know about this deficiency?  Did they mention
-  it in their paper?  (If neither, we should make a bigger deal out of the
-  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
-    They refer to ``temporal security properties'' as something to do with dangling
-    pointers, but they don't define it.
-    Then they say that Depth Isolation has weaker temporal security
-    guarantees than Static Authorities (tag=function name), but they don't explain why
-    (and I don't see why). They don't say anything explicit about the temporal
-    security properties of Lazy Tagging and Clearing, but they consider it to enforce
-    a completely different (and generally weaker) policy anyway (``data flow
-    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
-  leave this issue to deal with later (in consultation with them, which we
-  don't have time to do a good job of today).  At any rate, by my reading we
-are not saying anything incorrect at the moment.}
-\fi
 
-% A second challenge is that enforcement mechanisms may weaken their
-% protections\bcp{``weaken their protections'' doesn't convey what we mean
-%   very well}
-% in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
-% \citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
-% ways of identifying functions and deciding exactly which accesses are allowed.
-% These vary in their efficiency and in the spatial and temporal precision of their
-% security guarantees.
-% The Depth Isolation policy prevents a callee from writing outside its frame
-% at all, which is simple and intuitively correct, but it must manually clear the tags
-% on the frame on each return, which is very expensive. Conversely, their
-% \emph{Lazy Tagging and Clearing} policy improves
-% performance by assigning locations to the active function when they are written, which
-% permits  ``illicit'' writes to inaccessible stack frames so long as the
-% written locations are not subsequently read by their ``real'' owners. This violates our
-% straightforward properties, but ntuitively,
-% this form of laziness is justified if any dangerous writes will
-% ultimately be detected before they can affect program behavior.
+Finally, we demonstrate our model's flexibility by extending it with the passing
+of variables on the stack and to a simple coroutine model.
 
-% \bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
-% What does ``affecting program behavior'' really mean? This micro-policy
-% calls for a second characterization of stack safety in terms of the externally
-% observable behavior of a stack-safe system. Such a characterization will have the
-% additional benefit of being more {\em extensional}, concerned primarily with the
-% external behavior of the system. In contract, our stepwise properties are {\em intensional},
-% focusing on internal details that may not actually matter to overall security.
-
-% \bcp{Ditto, even moreso.}
-% To formalize this intuition, we develop an \emph{observational integrity} property which
-% characterizes enforcement mechanisms that detect violations between the point when
-% they occur and the point when they make an observable impact. Intuitively this property
-% can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
-% integrity implies observational integrity. \sna{TODO: check this.}
-
-% We additionally sketch an associated {\em observational confidentiality} property
-% makes confidentiality extensional in the same way, in order to form an
-% {\em observational stack safety} that is highly extensional.
-% We argue that, despite its somewhat greater complexity,
-% it may be a better specification for
-% stack safety enforcement mechanisms in general.
-
-%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
-%  significant contribution, and partly to give the reader some context for
-%  why this is part of our story.  E.g., ``One nice thing we can do with our
-%  spec is to use it to evaluate the claims of existing or future proposals
-%  for enforcement mechanisms.  One relatively inexpensive way to carry out
-%  such an evaluation is to see whether property-based testing against this
-%  spec can distinguish correct implementations of the mechanism from buggy
-%  ones...''}
-
-For ease of exposition, our development initially assumes a single simple
-stack with no sharing between callers and callees: all parameters and return
-values are passed in registers.  Later, we show how to refine the properties
-to allow passing of scalar stack data, granting a callee access to the
-caller's data. Throughout, we note how we expect elements of the formalism
-to interact with enforcement mechanisms beyond micro-policies.
-
-Lastly, we define integrity and confidentiality properties for a coroutine
-system with a static layout, in which each coroutine's stack is constrained
-to a fixed region of memory.  Though simplistic, this extension is a step
-toward full-blown concurrency and demonstrates that our approach is flexible
-enough to handle more sophisticated styles of control flow.
-
-%% \paragraph*{The Use of Stack Safety}
-
-%% How does the programmer rely on stack safety? The stack makes implicit promises about
-%% the behavior of function calls. First, that when a call is finished, the caller can continue
-%% from where it left off with its local state unchanged. And second, that the call's
-%% behavior is determined only by a predictable interface that does not include the caller's
-%% local state.
-
-%% %The most natural example is the first
-%% %in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
-%% %calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
-%% %{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
-%% %overwrite {\tt main}'s stack data.
-
-%% Both expectations imply predictable control flow, termed {\em well-bracketed
-%% control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls always jump to valid entry
-%% points and return to their call sites. Then the data portion of the first expectation is for
-%% {\em integrity}: that a caller's local data is protected from modification by its callee.
-%% This is vital for local reasoning about the behavior of the caller. The security implications
-%% are significant: without integrity, a caller that branches on a local variable could be
-%% manipulated by its callee into unexpected behavior.
-
-%% Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
-%% from leakage by its callee. As a reasoning principle, confidentiality tells us that to
-%% predict the behavior of a call we need only consider a subset of the machine state.
-%% And in security terms, we can guarantee the secrecy of local data that we haven't passed
-%% even when calling a badly behaved function. Together, integrity and confidentiality
-%% comprise \citeauthor{Skorstengaard+19}'s {\em local state encapsulation} (LSE) property.
-
-%% It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
-%% could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
-%% accessing the caller's state, but the execution would remain under the callee's control.
-
-%% To phrase LSE and WBCF as security properties requires an explicit attacker model:
-%% a clear understanding of who is being protected and from what. Most existing
-%% stack protection mechanisms envisage scenarios where a callee is
-%% influenced by external inputs to corrupt the stack in some way. To
-%% avoid getting into the details of exactly how this happens when
-%% formalizing our properties, we use an even stronger model: when we
-%% state the protections that a function can rely on when it makes a
-%% call, we assume that the callee can execute arbitrary instructions
-%% until it returns. Thus we have no need to differentiate safe code
-%% from adversarial --- all functions, and indeed all activations of the same function,
-%% are mutually distrusting.
-%% %
-%% For this model to make sense, we need to know when a call or a return is taking
-%% place. We assume that calls and returns are annotated in the code, so that
-%% we can track the point at which control transfers from the protected caller
-%% to the attacking callee --- and when that callee in turn becomes a protected
-%% caller in its own right.
-
-%% This ``nesting'' of properties is a key challenge in adapting the concepts
-%% of language-based security to the stack setting. We address this issue by
-%% turning any set of annotations into a dynamic model of {\em domains},
-%% in which some stack locations are sealed by the caller that needs to preserve
-%% them while the rest are unsealed and available for use by the active, attacking
-%% callee.
-
-%% First we introduce this model for a single simple stack with no sharing between callers and
-%% callees. Then we expand it to allow passing of stack data, granting a callee one-time
-%% access to the caller's data. We do not model more sophisticated sharing, because to do
-%% so realistically would require a separate protection scheme, such as a capability
-%% model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
-%% a separate policy outside the scope of stack safety. (A tag-based approach can easily
-%% compose multiple such policies.)
-
-%% Finally, we extend the model to a coroutine system featuring finitely-bounded stacks.
-%% Though simplistic, this extension is a step toward concurrency and demonstrates that
-%% the model is flexible enough to handle more sophisticated styles of control-flow.
-
-%% After formally defining stack safety, we show that the definition is
-%% enforceable using an existing tag-based enforcement
-%% mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
-%% \citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy.
-%% This policy testably enforces our {\em eager} properties, in which integrity is respected
-%% at every step of a program, and each call respects integrity at each of its internal steps.
-%% These properties are very strong, and our policies enforce them; this also makes them very
-%% intensional, in that they do not distinguish internal behavior from external.
-%% They also have the advantage that they are suitable for efficient testing, as they
-%% fail immediately when a violation occurs.
-
-%% The downside of the Depth Isolation policy is that it is rather expensive to implement.
-%% In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
-%% that is more efficient by virtue of being lazy in its enforcement.
-%% Instead of enforcing the integrity and confidentiality of the stack at the
-%% exact moment when a callee returns, this policy signals violations when the
-%% caller actually accesses data that the callee has written into its stack
-%% frame. \iftrue\bcp{We could also hint at the lazy confidentiality
-%%   property in the same way...}\fi
-%% The lazy policy does not enforce our eager integrity property, and to our
-%% knowledge there is no formal statement of the properties that it ought to enforce.
-%% We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
-%% policy that detects integrity violations between when they occur and when they make
-%% a visible impact. Intuitively this can be phrased as ``no visible reads after dangerous
-%% writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
-%% ``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
-%% testing, but formalizing the ideal property allows us to see how it falls short.
-%% We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
-%% at some performance cost.
-%% %We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
-%% %  and is it so interesting anyhow?}
-%% We argue that any system purporting to enforce stack safety ought to obey at least lazy
-%% stack safety, making it a plausible extensional characterization.
-
-%% %\bcp{I worry that we are being a little
-%% %  dishonest here: the stated advantage of the lazy policy is that it is more
-%% %efficient, but our variant is {\em not} efficient!}
-In summary, we offer the following contributions:
-
-\begin{itemize}
-\item We adapt concepts of integrity and confidentiality from language-based
-security, combine these with an existing notion of well-bracketed
-control flow to formalize a {\em stepwise stack-safety}
-property, and argue that this property captures the intuitive concept of
-strict stack
-safety~(\cref{sec:prelim,sec:lse}).
-\item We use property-based randomized testing to check that this stepwise
-property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18}, and extend the generation by
-execution techniques of \citet{TestingNI:ICFP} to handle nested
-properties (\cref{sec:enforcement}).
-%% \bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
-\item We formalize a
-weaker but more extensional \emph{observational} variant of the stack safety
-property that characterizes a refinement of Roessler and DeHon's more
-performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
-\item We show that the definition can be
-extended to argument passing on the stack and to a simple coroutine system (\cref{sec:ext}).
-\end{itemize}
-
-We begin with a description of our setting (\cref{sec:setup}) and threat
-model (\cref{sec:threat}), and an informal overview of our key ideas via
-examples (\cref{sec:running-example}). \Cref{sec:relwork,sec:future} discuss related and future work.
-
-\section{Setting}
-\label{sec:setup}
-
-\paragraph*{Machine model}
-Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the concrete machine that instantiates it.  We assume a register
-machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
-details of the machine's ISA are unimportant. We assume that we know ahead of time
-the layout of memory, including where code lives and which addresses are within
-the bounds of the stack. We do not model a heap; programs might choose to use memory
-that is neither code nor stack to store global variables, as a heap, etc., but our
-properties do not protect it or assume any particular behavior. We also do not
-model run-time code generation.
-
-% \footnote{\bcp{Some possibly useful text: A major challenge in
-%     characterizing stack safety formally is that, in a low-level machine
-%     without built-in stacks, we are at the mercy\bcp{?} of
-%     architectural\bcp{?}  choices that do not relate to the abstract notion
-%     of stack safety. Some implementations may have only a stack pointer
-%     while others add a frame pointer, in some systems return addresses and
-%     data live in separate stacks, and so on. We abstract over these choices
-%     as much as possible.  }}%
-
-\paragraph*{What is a ``call''?}  One significant challenge in formulating our
-properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy
-\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
-particular, there may be no single ``call'' instruction that does all the
-required manipulation of the \(\PCname\) and stack pointer. Instead, there
-might be just a simple branch instruction that stores the old \(\PCname\) in
-a register, with the rest of the call sequence performed by other
-instructions either before or after the jump.  Similarly, there may be no
-``return'' macro-instruction, but rather a ``return sequence'' involving an
-indirect branch at some point.  Thus, the same instruction opcode may be
-used in some places as part of a call or return sequence and in other places
-for other purposes.
-
-Since our stack-safety properties are all about protecting ``callers'' from
-their ``callees,'' we need some clear-cut way to recognize when the program
-makes or returns from a call.  Accordingly, we assume we are given a machine
-language program {\em annotated} with the locations of instructions
-representing calls.  We declare that a call has occurred when the machine
-has just executed an instruction marked as a call.
-%
-These call annotations would typically come from a compiler together with
-the machine code, but we do not assume anything about their provenance.
-Annotations connect the enforcement mechanism to the property: we don't
-protect unmarked calls, because a mechanism's starting conditions might
-depend on knowing where calls are.
-
-Returns, on the other hand, need not be annotated: we can tell when a callee
-has returned control to the caller when... we see control return to the
-caller! We don't require the callee's cooperation at all.
-For example, under a standard calling convention, we can see that a
-return has occurred occurred when state of the machine reaches a state where
-the stack pointer has the same value as it did at the point of some earlier
-call and the $\PCname$ points to the instruction following that call.  We
-abstract away from the details of specific calling conventions by assuming
-that we are given some way of telling when a given machine state is a
-{\em return target} with respect to some set of pending calls.
-
-%, returns, and function entry points\BCP{I
-%  still want to talk about whether returns and entry points need to be
-%  annotated.  To me, these annotations do not make intuitive sense: it is
-%  the caller's responsibility to know whether the address that it is calling
-%is really an entry point, and we can trust it to check this.  Similarly, it
-%is not the callee's role to say when it is returning; it is the caller's
-%prerogative to decide when it has been returned to!  I.e., from the caller's
-%point of view, it has been returned to when control resumes at the
-%instruction following the call with the SP reset to where it was at the
-%point of the call. (This is the point where it wants to be able to access
-%its own local state again and expected to find it unmolested.)}.
-
-\paragraph*{Policies}
-We model the base machine's behavior as a step function over machine states,
-and a machine
-enhanced with an enforcement mechanism, or {\em policy} for short, as a
-partial step function over machine states extended with some
-auxiliary state, refining the base machine. When the policy-enhanced machine
-does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
-action that would violate the security property;
-our properties are therefore naturally \emph{termination insensitive.}
-This model will be made more precise in \cref{sec:prelim}.
-
-Dividing policy state from machine state helps us reason about
-enforcement mechanisms that keep a significant amount of data separate from the
-primary execution data to inform security. Obviously tag-based systems do this,
-with tags and extra policy state separated from and untouchable by application code.
-Another significant example is the shadow stack, where a second stack
-mirrors the real one, and a return may only occur if both stacks agree on the
-target address. Typically the shadow stack is isolated using a separate protection
-mechanism.
-
-The separation is smaller in protection mechanisms that interact more directly
-with application code. Some mechanisms might exhibit failstop behavior, but have
-no protected policy state; others might never even failstop. Then the policy-enhanced
-machine is simply identical to the original. This is the case for software-based
-techniques.
-
-\section{Threat Model}
-\label{sec:threat}
-
-Consider a common scenario: a developer wants to link against an existing
-code-base, which will then occupy the same address space as their own code.
-But perhaps the external code is buggy, or even malicious, and contains
-exploits that can be used to cause it to modify the developer's data or
-exfiltrate secrets. For that matter, the developer may not be a particularly
-secure programmer, and may not be able to rely on their own defensive programming
-to keep their secrets safe.
-
-The developer likely relies instead on simple reasoning principles about how
-their code should work. Local variables do not change across function calls,
-and the callees generally do the same thing regardless of when and where they
-are called (up to a reasonable interface, such as their parameters and global
-variables.) If these intuitions prove false, then the programmer will reason
-incorrectly about their program and especially about its security, and
-be exposed to attack.
-
-Intuitively, stack safety means that the stack frame of each caller is protected
-from attempts by a confused or malicious callee (including all its
-descendant callees) to read or write its private data.
-That is, the attacker model treats every caller as a defender and its callee as
-an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways. We do rule out some
-scenarios with a ``well-formedness'' criterion on initial states, ensuring that
-they are set up properly for the enforcement mechanism in question.
-When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
-well-behaved calls and returns are implemented by fixed sequences of instructions
-which the tag policy will force to be executed from start to finish,
-termed {\em blessed sequences}. Well-formedness requires that all annotated
-instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
-This is all we need for the policy to protect itself: it can then failstop if it would
-execute the blessed sequence out of order.
-
-We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. Hardware and timing attacks are out of scope.
-So are attacks on the heap -- we believe that the work of
-\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with our
-to protect both stack and heap.
-
-%We model ordinary program state and policy-enforcement state separately to make it
-%easier to use well-established concepts from language-based security.
-%In particular, confidentiality has an elegant extensional characterization in terms of
-%\emph{noninterference}: a program preserves the confidentiality of certain data if
-%\emph{varying} that data would not change the program's observable
-%behavior~\citep{6234468}.
-%In applying this idea to a system that incorporates policy enforcement, only
-%ordinary program data should be varied, while any associated policy
-%state should be left alone; our model makes this distinction obvious.
-%That is, we compare the actual execution of the program with a \emph{hypothetical}
-%execution where data has been varied.
-%We will also introduce a notion of {\em observational integrity} that relies on the idea of
-%\emph{rolling back} program state; again, our model makes it easy to describe
-%rollback of ordinary program data while leaving policy state alone.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
-%We believe that it is flexible enough to capture code-altering approaches, such as code
-%rewriting or capability-based techniques, though perhaps with less elegance.
-
-%\paragraph*{Micro-Policies}
-%
-%Micro-Policies are a flexible tag-based, hardware-accelerated
-%reference monitoring mechanism; they been applied to stack safety
-%enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
-%Here the policy state consists of metadata tags (e.g., identifying stack frames)
-%attached to each value in memory or registers, and the policy step
-%function checks that each machine operation obeys a set of rules on tags
-%(e.g., that the current $\PCname$ tag matches the tag on the stack
-%location being accessed), and halts the machine if not.  Distinguishing
-%ordinary machine state and policy state is very natural in this setting.
-%% Here the machine state is extended with tags paired with all components of the
-%% system, and a monitor checks for each operation that the tags on the operands
-%% obey a set of rules. One could treat tags directly as part of the machine
-%% state, and build the rules into the step function. But local state
-%% encapsulation properties as we frame them are not intuitive in their
-%% interaction with tags or other additional enforcement state, so it is helpful
-%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
-%% a software fail state.
-%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-%%   policy? }
-
-%\iftrue\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
-%These approaches are represented as policies that
-%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
-%being produced by the compiler.) The set of policy states is again the unit type
-%and the step function constant. Property enforcement comes from the fact that
-%only ill-formed states can induce a trace that violates the property.
-
-%% \Paragraph*{Null policy}
-%% %
-%% The simplest possible policy is the neutral policy that carries no information
-%% and does nothing. Its set of policy states is the unit type, and its policy step
-%% function is the constant unit function, thereby allowing all calls to the machine
-%% step function. Its initialization function is defined on all states. This
-%% corresponds to the lifting of our machine model to the policy setting. Machines
-%% that fulfill a property by construction will also fulfill that property with the
-%% null policy.
-
-%
-%\leo{TODO: Add citations of mitigation techniques}
-%\rb{And for those techniques, explain how they reflect their motivating
-%examples. What does each technique do for those examples? How are they
-%connected to this work?}
-%
-%\paragraph*{Bad Write}
-%
-%\[
-%44 : [\rsp - 5] \leftarrow r_4
-%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
-%\]
-%
-%Instruction 44 currently is
-%$ [\rsp - 2] \leftarrow r_4$,
-%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
-%with the
-%effect of storing the contents of $r_4$ as the result of {\tt g}. If
-%instead it wrote to location $\rsp - 5$ then it would be overwriting
-%private data (the return address) belonging to {\tt f}.
-%
-%\paragraph*{Bad Read}
-%
-%\[
-%30 : r_5  \leftarrow [\rsp - 6]
-%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-%\]
-%
-%Instruction 30 currently is
-%$ r_5 \leftarrow [\rsp - 4]$,
-%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
-%with the
-%effect of loading the argument of {\tt f} to $r_5$. If instead it
-%loaded the contents of location $\rsp - 6$, it would obtain access
-%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
-%outputs the contents of $r_5$, that value would immediately be made public.
-%
-%\paragraph*{Bad Control Flow}
-%
-%\[
-%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
-%%5 : \jalr ~ r_4 ~ \rra ~ 0
-%\]
-%If a {\jalr} instruction was added to {\tt main} with its target being
-%the middle of the code block of {\tt g}.\apt{describe result of this?}
-
-\section{Key Ideas by Example}
-\label{sec:running-example}
-\colorlet{lgray}{gray!40}
-\colorlet{lred}{red!40}
-\colorlet{lblue}{blue!20}
-
-\newcommand{\mainsealc}{cyan}
-\newcommand{\fsealc}{green}
-\newcommand{\unsealc}{lgray}
-\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
-\newcommand{\fulloutc}{white}
-\newcommand{\badc}{lred}
-\newcommand{\goodc}{lblue}
-\newcommand{\retptrc}{black}
-
-\begin{figure}
-
-\begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-int main() {
-  int x = 42;
-  return f()+x;
-}
-
-int f() {
-  int y;
-  *(&y-2) = 0;
-  return 5;
-}
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}[t]{.5\textwidth}
-{\small
-\begin{verbatim}
-1 main: add $1,%sp      ; allocate frame
-2       mov $42,-1(%sp) ; initialize local
-3       call f
-4       add -1(%sp),%ra ; set return value
-5       sub $1,%sp      ; deallocate frame
-6       ret
-
-10 f:   add $1,%sp      ; allocate frame
-11      mov $0,-3(%sp)  ; violation!
-12      mov $5, %ra     ; set return value
-13      sub $1,%sp      ; deallocate frame
-14      ret
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}[t]{.25\textwidth}
-\begin{center}
-\begin{tabular}{l l}
-{\tt 2} &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}[U]
-~$\cdots$
-\vspace{.5em}
-\\
-{\tt 3} &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}
-~$\cdots$
-    \MemoryLabel{-15em}{0.75em}{42}
-    \vspace{.5em}
-\\
-{\tt 11} &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}[S(0)]%
-\memory{1}{\retptrc}[RetPtr]%
-\memory{1}{\unsealc}[U]
-~$\cdots$
-\MemoryLabel{-15em}{0.75em}{42}
-\vspace{.5em}
-\\
-{\tt 12} &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}
-\memory{1}{\retptrc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-15em}{0.75em}{\bf 0}
-\vspace{.5em}
-\end{tabular}
-\end{center}
-\vspace{-4ex}
-\end{subfigure}
-\caption{An integrity violation
-\ifaftersubmission\bcp{I wonder if we
-    could make the figures easier to read by putting very light, differently
-  colored backgrounds under the different parts}\fi}
-\label{fig:int1}
-\end{figure}
-
-This section introduces the intuition behind the properties through a series
-of examples. We show our examples both in C syntax and in corresponding assembly code for a
-simple two-address machine. The address of each instruction is shown to its left.
-The assembly code reflects a simple compilation model in which
-local variables are stored in the stack frame.
-For the moment, arguments and return values are passed in registers.
-The machine is word-addressed. The stack grows upward in
-memory and {\tt \%sp} points to the first unused word above the top of the stack.
-The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
-implicitly adjusting the stack pointer.
-At the right of each example, we show the layout of stack memory just before
-the specified instruction is executed; each box represents a memory word, and
-addresses increase to the right.
-
-Colored boxes indicate an assignment of the address to a
-{\em domain}---a label reflecting a particular set of access rights. Domains
-are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
-(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
-(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
-figures. For our purposes, \(\outside\) will apply to global
-variables and code. Domains are part of the {\em context} of a state.
-We will introduce the remaining context as needed.
-
-Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
-function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
-variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
-meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
-in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt x} will be unchanged when it gets control back.
-This gives us our simple, intuitive statement of {\em stepwise integrity}:
-if a component is sealed before a step, it will be unchanged afterward. Stepwise
-refers to the fact that the property must hold on every step. %Later we will
-%loosen this condition in the interest of extensionality -- the caller only
-%cares about changes it can see after the return, not changes during execution.
-In this case, after instruction 11, {\tt f} has violated integrity by writing to {\tt x},
-so the final return value produced by {\tt main} is wrong.
-%
-(The black rectangle represents the return address, which is saved by the call
-instruction itself, so it is not sealed.)
-
-\begin{figure}
-\begin{subfigure}[t]{.25\textwidth}
-{\small
-\begin{verbatim}
-int main() {
-  int x = 42;
-  return f()+x;
-}
-
-int f() {
-  int y,z;
-  y = 5;
-  z = *(&y-2);
-  return y+z;
-}
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}[t]{.5\textwidth}
-{\small
-\begin{verbatim}
-1 main: add $1,%sp      ; allocate frame
-2       mov $42,-1(%sp) ; initialize local
-3       call f
-4       add -1(%sp),%ra ; set return value
-5       sub $1,%sp      ; deallocate frame
-6       ret
-
-10 f:   add $2,%sp      ; allocate frame
-11      mov $5,-2(%sp)  ; assign into y
-12      mov -4(%sp),-1(%sp)  ; violation!
-13      mov -2(%sp),%ra ; calculate and
-14      add -1(%sp),%ra ; set return value
-15      sub $2,%sp      ; deallocate frame
-16      ret
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}{.59\textwidth}
-\centering
-\begin{tabular}{l l | l}
-  {\tt 2} &
-  \multicolumn{2}{c}{
-    \memoryaddrs{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-%    \MemoryLabel{-18em}{0.75em}{0}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-%  \MemoryLabel{-10em}{0.75em}{0}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
-  \\
-  {\tt 12} &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-10em}{0.75em}{5}
- % \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-10em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
-  \\
-  {\tt 13} &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-10em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{42}
-  \vspace{.5em}
-  &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-10em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
-  \vspace{.5em}
-\end{tabular}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-\caption{A confidentiality violation\ifspace\apt{Could save space by not repeating C or assemly code for {\tt main}.}\fi}
-\label{fig:conf1}
-\end{figure}
-
-Meanwhile, confidentiality intuitively states that a callee is insensitive to
-the context of its call (except for its arguments and any global
-variables).
-We state this as a kind of {\em noninterference} property~\citep{Goguen1982SecurityPA}.
-Noninterference describes knowledge: it states that for system components
-containing secret values, we can substitute any values whatsoever without changing the results
-of our computation. If there exists some value for which we see different results, then
-in fact we can learn something about the initial state by observing those results, violating
-confidentiality.
-
-In Figure~\ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
-by reading {\tt main}'s variable {\tt x}. The trace
-shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
-state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
-understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
-be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
-changed to 5 in both the primary trace and the variant. The primary and variant agree on
-the value in the changed location, so confidentiality was not violated. But at instruction 13,
-the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
-in the variant trace it might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
-has behaved differently than the primary, violating confidentiality.
-
-We will term this form of confidentiality as {\em stepwise confidentiality}. Informally,
-for any state at the entry to a function, for any variant over the values in the stack,
-the primary state and its variant will step in lockstep until they return. To step in lockstep
-means that if both states step---that is, neither the primary nor the variant failstops---any
-{\em system component} (such as memory cell, register, program counter) that changes
-in one step has a matching value at the end of the other step.
-When we later extend the model with a notion of observable output (\cref{sec:lazy}),  we will instead apply the matching
-criteria to the outputs.
-
-%\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
-%track which addresses may be read (i.e., which have been initialized), and at each step
-%guarantee that the step alters state identically to a step from a variant state. Then we
-%wouldn't need to divide up the execution by calls. But the approach we take is more
-%consistent with a typical noninterference approach, in that it varies the memory once
-%at the beginning and trace changes through the execution.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}[t]{.2\textwidth}
-  {\small
-\begin{verbatim}
-int main() {
-  int x;
-  x = f();
-  return g();
-}
-
-int f() {
-  int y = 5;
-  return y;
-}
-
-int g() {
-  int z;
-  return z;
-}
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}[t]{.68\textwidth}
-  {\small
-\begin{verbatim}
-1 main:  add $1,%sp      ; allocate frame
-2        call f
-3        mov %ra,-1(%sp) ; store x
-4        call g
-5        sub $1,%sp      ; deallocate frame
-6        ret
-
-10 f:    add $1,%sp      ; allocate frame
-11       mov $5,-1(%sp)  ; initialize y
-12       mov -1(%sp),%ra ; set return value
-13       sub $1,%sp      ; deallocate frame
-14       ret
-
-20 g:    add $1,%sp      ; allocate frame
-21       mov -1(%sp),%ra ; violation!
-22       sub $1,%sp      ; deallocate frame
-23       ret
-}
-\end{verbatim}
-}
-\end{subfigure}
-\end{center}
-\begin{subfigure}{.65\textwidth}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 2} &
-  \multicolumn{2}{c}{
-    \memoryaddrs{8em}
-    \memory{3}{\unsealc}
-    ~$\cdots$
-%    \MemoryLabel{-18em}{0.75em}{0}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
- % \MemoryLabel{-19em}{0.75em}{0}
-%  \MemoryLabel{-10em}{0.75em}{0}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
-  \\
-  {\tt 12} &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
- % \MemoryLabel{-19em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(5\)}
-  \\
-  {\tt 4} &
-  \multicolumn{2}{c}{
-    \memoryaddrs{8em}
-%    \memory{4}{\unsealc}
-    \memory{1}{\unsealc}
-    \memory{1}{\retptrc}
-    \memory{1}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-14em}{0.75em}{5}
-%    \MemoryLabel{-10em}{0.75em}{\(v_1\)}
-    \MemoryLabel{-6em}{0.75em}{5}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  }
-  \\
-  {\tt 22} &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\bf 5}
- % \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_3\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_4\)}
-%  \MemoryLabel{-2em}{0.75em}{90}
-  \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-\caption{Another confidentiality violation}
-\label{fig:conf2}
-\end{figure}
-
-It may be unclear in Figure~\ref{fig:conf1} why we vary the initial values of {\tt y}
-and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
-temporal stack safety.
-In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
-that later becomes the value of {\tt main}'s variable {\tt x}. Then {\tt g} is able to access
-{\tt y} and hence {\tt x} without even directly reading the stack pointer.
-In the primary trace, {\tt z} will have the value 5, but in the variant it can have
-an arbitrary value $v_4$; since the variant can behave different from the primary,
-the policy again signals a confidentiality violation.
-
-\paragraph*{Control-Flow Attacks}
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\footnotesize
-\begin{verbatim}
-int *stash = 0;
-int main() {
-  int x = 1;
-  f();
-  x = -x;
-  f();
-  return x;
-}
-
-void f() {
-  int y;
-  if (!stash) {
-    stash = *(&y-1);
-  } else {
-    *(&y-1) = stash;
-    stash = 0;
-  }
-  return;
-}
-\end{verbatim}
-}
-\end{subfigure}
-  \begin{subfigure}[t]{.4\textwidth}
-{\footnotesize
-\begin{verbatim}
-   stash: .word 0
-1  main:  add $1,%sp     ; alloc frame
-2         mov $1,-1(%sp) ; init x
-3         call f
-4         neg -1(%sp)    ; x = -x
-5         call f
-6         mov -1(%sp),%ra ; set ret val
-7         sub $1,%sp     ; dealloc frame
-6         ret
-
-10 f:     add $1,%sp     ; alloc frame
-11        cmp stash, $0  ; stash = 0?
-12        breq #15       ; branch if so
-13        mov -2(%sp),stash ; set stash
-14        jmp #17
-15        mov stash,-2(%sp) ; get stash
-16        mov $0,stash
-17        sub $1,%sp     ; dealloc frame
-18        ret
-\end{verbatim}
-}
-  \end{subfigure}
-%
-  \begin{subfigure}[t]{.2\textwidth}
-    \begin{center}
-    \begin{tabular}{l l l}
-      {\tt 3} &
-      \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      {\tt 18} &
-      \memoryaddrs{12em}
-      \memory{1}{\mainsealc}[S(0)]%
-      \memory{1}{\unsealc}[RetPtr]%
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1}
-      \MemoryLabel{-11em}{0.75em}{\#4}
-      \vspace{.5em} &
-      \hspace*{-1.5em}
-      \memory[1.2em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-2.2em}{0.75em}{\#4}
-      \\
-      {\tt 4} &
-      \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      {\tt 11} &
-      \memoryaddrs{16em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#6} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      {\tt 18} &
-      \memoryaddrs{12em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\badc}
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#4} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      {\tt 4} &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      {\tt 5} &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{\bf 1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
-    \end{tabular}
-    \end{center}
-    \vspace{\abovedisplayskip}
-  \end{subfigure}
-  \caption{A control-flow attack.}
-  \label{fig:controlflow}
-\end{figure}
-
-\label{sec:controlflow} % RB: ???
-
-%From the classic stack smashing attack to return oriented programming,
-One of the most pernicious ways to attack the stack is by hijacking control flow.
-Our stack safety properties do not assume that control flow follows a typical structure
-of matching sets of calls and returns nested within on another.
-Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
-Consider Figure~\ref{fig:controlflow}.
-(Here the diagram shows return addresses explicitly as line numbers, in the form $\#n$.)
-Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
-3 and 5, respectively. But the first time, before returning, it stashes its
-return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
-up being called a third time, at which point it finally returns to instruction 5.
-
-Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that {\tt main} returns
-$-1$. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
-This has the appearance of an integrity violation.  How do we capture this in our
-integrity property?
-
-The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
-to do more than just execute a return-labeled instruction: it must actually reach the
-appropriate return target in {\tt main}.
-We add a new piece of context to our diagram: we keep a stack of return targets, and a callee is
-only considered to have returned when it reaches a state that matches the caller's target.
-A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, here represented by sharing a color with the sealed data.
-
-In the example, on its first return, {\tt f} reaches its return target, because the
-\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
-(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
-Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.
-
-\paragraph{Well-bracketed control flow.} Together,
-integrity and confidentiality protect a caller's data until its callee
-returns to it. However, it is possible that an intermediate caller is not
-returned {\em to}, but rather returned {\em past}, to some ancestor caller!
-For example, function $A$ calls function $B$, which in turn calls $C$, which
-returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
-and branching to the address after the call from $A$ to $B$.  Strictly
-speaking, this is not an information-flow violation, since neither $A$'s nor
-$B$'s stack frames are improperly read or written by $C$.  But it is still
-arguably unsafe in general.  For example, suppose that $B$'s continuation
-after the call to $C$ takes $C$'s result, doubles it, and returns this to
-$A$.  If $A$'s correct behavior after the call to $B$ returns depends on the
-result it receives being a even number---i.e., $A$ is trusting that $B$'s
-continuation will execute after $C$ returns, perhaps because $A$ and $B$ are
-written by the same person while $C$ is an untrusted library---then $C$ can
-cause it to behave incorrectly, possibly even insecurely, by returning an
-odd result directly to $A$.
-%
-Thus, a formal definition of safe use of the stack should include the
-condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $B$.  This is what \citet{SkorstengaardSTK} describe
-as \emph{well-bracketed control flow (WBCF)}.
-
-\section{Machines and Traces}
-\label{sec:prelim}
-
-We now define our machine model precisely. To make our definition of stack safety
-as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (software- or hardware-enforced) policy monitor.
-In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
-policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
-
-\subsection{Values and States}
-
-The building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively.
-%
-Our machine states are composed of {\em components} \(\component\),
-which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special-purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-A {\em machine state} is a map from components to values.
-%
-    \[\component \in \COMPONENTS ::= \WORDS + \REGS  \hspace{1in}
-    \mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-%
-We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-
-We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
-A {\em policy} consists of a set of policy
-states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. The policy also provides a
-well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
-initial states that are properly configured.
-%
-%A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-%\Cref{sec:enforcement}. -- APT: We just said this.
-
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
-               \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol
-               \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-Note that the base machine has a total step function, while the policy-enhanced machine
-refines it to a partial function. The extra policy state never changes behavior,
-but it can cause
-premature termination, representing failstop behavior.
-
-\subsection{Contexts}
-
-Separate from the machine state, we keep track of additional context
-information representing the history of the run. This has no influence on execution, it merely
-describes information relevant to the particular property. Different information is needed
-for different versions of the model, so the definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
-an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
-\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
-\(\codemap \in \WORDS \rightharpoonup \ANNS\)
-that relates each code address to its annotation, if any.
-We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).
-
-\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
-              {\(\mach, \context \constep \context'\)}
-              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
-                \mpcstate{\mach'}{\pol'}{\context'}\)}
-
-\subsection{Machine Traces}
-\label{sec:traces}
-
-A {\em machine trace} is a nonempty, finite or infinite sequence of elements
-of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces (ordinary lists will use
-``::'').
-
-The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
-coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\)
-
-\begin{center}
-\begin{minipage}{.4\textwidth}
-\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
-            {\(\mpcstatename' \hookrightarrow \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \notfinished{\mpcstatename'}{\MPCT}\)}%
-\end{minipage}
-\begin{minipage}{.4\textwidth}
-\judgment%[Default]
-         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
-         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
-\end{minipage}
-\end{center}
-%
-
-We project out the machine state of an MPC-state with \(\pi_\mach\), the
-policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
-We take the first element of a trace with \(\head(\MPCT)\), which is a total
-function since traces are non-empty, and the final element (if one exists) with
-\(\last(\MPCT)\), which is partial.
-
-\paragraph*{Until}
-The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
-and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first element on which \(f\) holds.
-
-\begin{center}
-  \begin{minipage}{.3\textwidth}
-    \judgment{\(f ~ \mpcstatename\)}
-             {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \mpcstatename\)}
-  \end{minipage}
-%
-  \begin{minipage}{.3\textwidth}
-    \judgment{}
-             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
-%
-  \end{minipage}
-  \begin{minipage}{.3\textwidth}
-    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
-                {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \notfinished{\mpcstatename}{\MPCT'}\)}
-  \end{minipage}
-\end{center}
-%
-If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate always holds, then
-\(\PUT\) is the identity function.
-
-We will frequently take a prefix of the trace from an initial state up
-through the first state where some condition holds on the machine state.
-This can easily be implemented using \(\PUT\), and we provide special notation
-for convenience.
-%
-If \(f\) is a predicate on machine states, we define \(\mpcstatename
-\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)''):
-%
-\begin{center}
-\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
-            {\(\PUT ~ f ~ \MPCT' = \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-\end{center}
-
-\paragraph*{Context Segments}
-
-In particular, we wish to formulate properties around consecutive subtraces
-in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth. If \(f\) is still a predicate on states, then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
-\ifspace{\bcp{Throughout, we could save space and improve beauty by removing some of the
-vertical space between inference rules.}\fi
-
-\begin{center}
-\begin{minipage}{.38\textwidth}
-  \judgmenttwo[ Now]
-              {\(f ~ \mpcstatename\)}
-              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
-\end{minipage}
-%
-\begin{minipage}{.58\textwidth}
-  \judgmentthree[ Skip]
-                {\(f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{minipage}
-%
-  \judgmentthree[ Search]
-                {\(\neg f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{center}
-
-\paragraph*{Reachability}
-
-We often want to quantify over all traces from any initial state so that our
-properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
-where \((\mach,\pol) \in \WF\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
-such that from some similarly well-formed initial state \(\mpcstatename\),
-\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
-
-\section{Stack Safety, Formally}
-\label{sec:lse}
-
-We are finally ready for our definition of stack safety. We begin
-by describing how we model knowledge about the program structure,
-then dive into the formal definitions of integrity, confidentiality, and
-well-bracketed control flow, which together constitute stack safety.
-
-We will focus on a simple stack model, with a single stack
-that doesn't share between caller and callee: arguments and results are passed
-in registers. To accommodate different machine architectures and different calling
-conventions, we parameterize over the structure of the stack, particularly
-the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention},
-%\(\sealcon ~ (\mach \in \MACHS) ~ (\component \in \COMPONENTS)\),
-\(\sealcon \subseteq (\MACHS \times \COMPONENTS)\),
-a relation between machine states and the addresses that should be sealed.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention},
-%\(\retcon ~ (\mach \in \MACHS) ~ (\mach' \in \MACHS)\),
-\(\retcon \subseteq (\MACHS \times \MACHS)\),
-which relates call states to the states that can be considered valid returns to that call.
-The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
-and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-
-Each system component is assigned to a {\em domain}, which may be {\em Outside}
-of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
-at a particular depth. Each caller registers a {\em target}
-condition, \(\target \subseteq \MACHS\), which is a predicate on machine states that means it
-has been returned to.
-A context is a pair of a ``domain map'' from components to domains, and a
-``return target list,'' which is a stack of targets, one for each caller awaiting
-return.
-%
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-  \times (\listT (\target)) \]
-%
-The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
-the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
-
-The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
-A call annotation tells us to push the return target onto the return stack and
-seal all components with the depth of the previous return stack. If we aren't making a call,
-we identify a return by looking ahead and checking if we're about to step to a state
-that matches the topmost return target. If neither is true, the context remains unchanged.
-Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
-as a predicate on addresses.
-
-\judgmentthree[ Call]
-              {\(\codemap ~ (\mach ~ \PCname)\)}
-%              {\(\components = \sealcon ~ \mach\)}
-              {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ ReturnFound]
-              {\(\mach \stepsto \mach' \in \target\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3.5ex}
-\judgment[ Default]
-         {}
-         {\(\underscore,\context \stepsto_C \context\)}
-
-When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We now proceed to state our properties.
-The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
-quantified over each step of execution from any initial state. The criterion for integrity
-is simple: sealed components do not change.
-
-\definition
-Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
-set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
-for some depth \(\depth\).
-
-\definition
-A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
-\(\mach,\pol,\context\) such that \(\mach,\pol,\context \stepstocon \mpcstatename'\),
-and any component \(\component \in \prot(\context)\),
-\(\pi_\mach(\mpcstatename') ~ \component = \mach ~ \component\).
-
-Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
-nested: we will quantify over subtraces that represent whole calls. For each such
-subtrace, we will quantify over {\em variants} of its initial state. Then we require
-that the original, {\em primary} traces move in {\em lockstep} with the variant.
-
-\definition \(\MPCT^*\) is the \emph{call set}, defined as
-\[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{\MPCT, \neg f_\depth | \MPCT\langle f_\depth \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
-\end{aligned}\]
-
-We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
-states that holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
-
-\definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
-is the set of components \(\component\)
-such that \(\mach ~ \component \not = \mach' ~ \component\).
-
-\definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
- {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all components \(\component\) in
-\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
-\(\mach' ~ \component = \nach' ~ \component\).
-
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep},
-written \(\machT \doteq \nachT\),
-if their adjacent pairs of states change together.  We define this as a coinductive relation:
-
-\begin{minipage}{.3\textwidth}
-\judgment{}
-         {\(\mpcstatename \doteq \npcstatename\)}
-\end{minipage}
-\begin{minipage}{.6\textwidth}
-\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
-            {\(\machT \doteq \nachT\)}
-            {\(\notfinished{(\mach,\underscore,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore,\underscore)}{\nachT}\)}
-\end{minipage}
-
-\definition
-Let the secret set of a context, \(\secret(\domMap,\rts)\), be the
-set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
-
-\definition
-A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
-any \(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\), and any trace
-%\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
-\(\NPCT\) such that \(\nach,\pol,\context \hookrightarrow \NPCT | \brk\),
-\(\MPCT \doteq \NPCT\) holds.
-
-%\medskip
-
-In other words, stepwise stack confidentiality states that if, at the start of the call,
-we vary all stack locations that are supposed to be secret from the callee
-(which in our simple setting is the entirety of the stack),
-the resulting trace remains in lockstep with the original trace
-until either both return together or one failstops. This implies that
-if the callee reads a secret location, it must failstop. %\apt{right?}\leo{yes}
-%
-Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
-of the components in \(\components\), and in addition their subtraces are also in
-\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
-will contain these calls and returns as well, so its return stack pushes and pops
-in lockstep until it breaks at the final return.
-
-Finally, we formalize a version of WBCF, which captures the same informal idea
-as that of \citet{SkorstengaardSTK} using our context machinery.
-\definition
-A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
-\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts))\), and
-\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts'))\),
-if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).
-
-%\subsection{Well-bracketed Control Flow}
-%\label{sec:wbcf}
-
-%Both the integrity and confidentiality components of stack safety
-%are reasonable even in programs that violate expected
-%control flow. But even though stack safety prevents control-flow violations
-%from being used to gain privileges, it is worth protecting control flow as well.
-%Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
-%control flow.
-
-%When we focus on subroutines, the essence of well-bracketed control flow
-%is that a caller can expect that if its callee ever returns, it returns
-%to the caller at the appropriate return target. This is tantamount to saying
-%that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
-%the simple model:
-
-%\definition A system enjoys {\em well-bracketed control flow} if for all
-%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-%\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\underscore, \underscore::\rts)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore, \rts)\).
-
-%Extending this idea to coroutines, we use the same notion that there are
-%no ``false'' yields.
-
-%\definition A system enjoys {\em well-structured control flow} if for all
-%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-%\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid')\)
-%where \(\stackid \not = \stackid'\).
-
-\section{Case Study: Do Micro-Policies Enforce Stack Safety?}
-\label{sec:enforcement}
-
-% \bcp{``Enforcement'' seems like too broad a title.  What about calling it
-%   ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
-%   Safety?''? And what about folding the testing section into this one?  That
-% would keep related things together, would avoid appearing to make overly
-% grand claims in either section, and would give us an excuse to write less
-% about the mechanism here---we only need to write enough about the mechanism
-% itself that readers can understand our validation case study.}
-% \leo{Agreed. Took a crack at it.}
-
-% \leo{I am not sure we've hit the correct level of detail in this
-%   section.}\bcp{Agreed!}
-
-In this section we explore how an existing enforcement mechanism, {\em
-  Depth Isolation}~\citep{DBLP:conf/sp/RoesslerD18}, implements
-the formal stack-safety property described in the previous
-section. In fact, it enforces a slightly stronger, inductive version
-of that property which checks for integrity and confidentiality violations at
-every step of the callee's execution trace. We validate this claim
-with a property-based testing framework that can quickly uncover
-inconsistencies between the enforcement mechanism and the stack-safety
-property.
-
-\subsection{A Conservative Stack-Safety Enforcement Mechanism}
-%
-The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
-on a programmable, tag-based reference monitor that runs alongside the
-program,
-allowing fine-grained manipulation of metadata tags to encode so-called
-\emph{micro-policies}~\citep{pump_oakland2015}.
-Every value in
-memory and registers (including the $\PCname$)
-is enriched with an abstract metadata tag, which can represent
-arbitrary information about the value. A micro-policy is then defined as a
-set of tags and a
-collection of software-defined rules.
-At each step of the machine, the relevant rule is applied to the tags on
-the instruction's inputs ($\PCname$, registers, memory) and on the
-instruction opcode itself, producing
-one of two outcomes: either the instruction is permitted to execute (and
-generates tags for the result of the operation and the new $\PCname$), or
-the machine fail-stops with a policy violation.
-Existing work \citep{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
-%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
-has shown that a wide range of
-micro-policies can be defined using this scheme.
-
-Efficient execution of these micro-policies relies on hardware acceleration,
-such as the PUMP architecture~\citep{pump:asplos2015}.
-The hardware incorporates a rule cache
-to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
-the hardware traps to a software handler (running in a privileged context or
-on a co-processor) to compute the rule result. To obtain adequate performance,
-it is important to design micro-policies so that they hit in the cache as
-much as possible. Thus, practical policies need to maintain a small working
-set of distinct tags.
-%% %
-%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
-%% policies}
-
-%\paragraph{A Conservative Tag Policy}
-%\label{sec:conservative}
-
-The micro-policy we present here to illustrate enforcement of stack
-safety as defined in \cref{sec:lse} is a small variation
-on the {\em Depth Isolation} policy presented by
-\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
-with ownership information associated to each stack frame (and to each
-separate object inside that frame, a more finely grained access
-control that we do not need to consider here), and tags registers containing
-stack pointers with access permission information. The policy also
-uses tags on instructions to identify the code
-sequences that have permission to manage the stack, say during calls and returns.
-%\leo{If we have the simple exposition so far, no need to mention the following sentence:}
-%In the following, for ease of exposition, we will assume that no
-%arguments are passed on the stack.
-%% \rb{Conceivably on alloc
-%%   operations, etc., which we have not yet mentioned.}
-
-The micro-policy maintains tags of the form $\tagStackDepth{n}$ for
-each stack location belonging to the stack frame at activation depth
-$n$, or the tag $\tagNoDepth$ for stack locations that are currently
-unused (but could potentially be used in the future).
-%
-The $\PCname$ is tagged similarly as $\tagPCDepth{n}$.
-%
-Initially, the entire stack is considered unused (and therefore tagged
-$\tagNoDepth$), and the {\PCname} has tag $\tagPCDepth{0}$.
-%
-During normal execution, the micro-policy rules only permit load and
-store operations when the target memory is tagged {\em with the same
-  depth} as the current {\PCname} tag, and store operations also when the
-target memory is $\tagNoDepth$.
-
-More interesting changes to the tags (from a stack safety perspective) 
-happen at exactly the points in the program where we switch from a
-caller to a callee or vice-versa:
-
-\begin{itemize}
-
-\item From caller to callee, when the machine executes an instruction marked as a call in
-  the call map.
-  At this point, the current
-  $\tagPCDepth{n}$ tag is incremented to $\tagPCDepth{(n + 1)}$, and the function
-  entry sequence initializes  all locations of the new function frame with
-  tag $\tagStackDepth{(n + 1)}$.
-
-\item From callee back to caller, when execution is about to reach a
-  return point.
-  %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
-  At the actual return instruction,
-  the $\PCname$ tag is decremented.
-%  \leo{Not sure the actual policy
-%    does that now (retag unused). But I don't think it's a
-%    confidentiality or integrity violation.}
-%  \rb{Correct, and it doesn't. Some versions of the policy do clear
-%    the stack, retagging would be easily added.}
-
-\end{itemize}
-
-This discipline suffices to enforce local stack encapsulation.
-To implement it,  we define blessed instruction sequences
-intended to appear at the entry and exit of each function,
-which manipulate tags as just described in addition to performing the
-usual calling convention tasks of saving/restoring the return address to/from
-the stack and adjusting the stack pointer. With the aid of an additional tag on
-the $\rsp$ register, these sequences also serve to enforce
-well-bracketed control flow.
-The micro-policy guarantees atomic execution of these sequences
-using a combination of tags on the instructions
-and an additional tag on the $\PCname$.
-
-There remains the question of how to ensure that the sequences are
-invoked at the right places from the perspective of the stack-safety
-property. For a program to enjoy stack safety
-(with respect to a particular call map), all we need require is that
-the entry sequence be initiated at any instruction marked as a call
-in the call map. This is achieved by giving these instructions (another) special tag,
-and it is easy to check statically that this has been done correctly.
-If the code fails to initiate an exit
-sequence at a point where the stack-safety property expects a return,
-the micro-policy will incorrectly behave as if execution
-is continuing in the callee, but since the callee never has stronger access
-rights than the caller, this is harmless.
-
-%% APT: removed because the tag part is too mysterious and the instruction part
-%% is standard.
-%% This is the entry
-%% sequence:
-%% %
-%% \setcounter{pcctr}{1}
-%% \[
-%%   \begin{array}{l|l|l|l}
-%%      & \mathit{Instruction} & \mathit{Tags} & \\
-%%     \hline
-%%     \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
-%%         {Store return address in stack}
-%%     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
-%%         {Increment stack pointer by frame size}
-%%   \end{array}
-%% \]
-%% %
-%% And the exit sequence:
-%% %
-%% \setcounter{pcctr}{1}
-%% \[
-%%   \begin{array}{l|l|l|l}
-%%      & \mathit{Instruction} & \mathit{Tags} & \\
-%%     \hline
-%%     \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
-%%         {Load return address}
-%%     \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
-%%         {Decrement stack pointer by frame size}
-%%     \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
-%%         {Return to caller ($\jalr$)}
-%%   \end{array}
-%% \]
-
-%% \rb{Positioning of the above sequences, relation to running example. The tagging
-%%   of stack frames would be considered part of these sequences.}
-
-%% These sequences also enforce the $\SP$ discipline. When stack frames
-%% are of fixed size, it suffices to increment $\SP$ and tag it with a
-%% dedicated tag in the header sequence, which is later checked after
-%% decrementing $\SP$ during the exit sequence to authorize the
-%% return. The well-formedness of the blessed sequences can be checked
-%% statically.
-
-%% Only programs whose sequences are well-formed are
-%% protected by the micro-policy, although programs are still protected
-%% (by fail-stopping) even if the intended sequences are missing. By
-%% tracking the execution of the blessed sequences in the tag state of
-%% the {\PCname} register, the micro-policy ensures that the entry
-%% sequences of calls declared in the call map are executed correctly.
-
-% \leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
-%   the weeds.  I don't think this level of detail is needed, in the context
-%   of a paper that is mostly about our definitions, not theirs.}
-% Note that in this policy both the entry and exit sequences
-% must write to each element of
-% the frame, which can be quite expensive, especially for programs that
-% allocate large but sparsely populated frames; we return to this point
-% in \cref{sec:lazy}.
-% %
-% For example, the program in \cref{fig:passingsimple} manages the stack
-% in this way. Here, the call pseudo-instruction to {\tt f} in line 5
-% hides a blessed sequence of instructions that allocate space to save
-% its return address to {\tt main} and seal the protected part of the
-% stack before {\tt f}'s body starts executing at 10. Conversely, the
-% return instruction in line 15 restores and jumps to the return address
-% and deallocates its frame space, and unseals the corresponding parts
-% of the stack. The function itself allocates and deallocates its own
-% private frame space in instructions 10 and 14. Additionally, unlike in
-% the simplified presentation of the policy, the caller shares the
-% arguments passed to the callee through the stack, and for this creates
-% and initializes (3--4) and destructs (6) the argument part of the
-% callee's stack frame.
-
-%\rb{Note that we have also considered and implemented more flexible schemes
-%  based on allowing operations above the current activation depth. How do we put
-%  these together?}
-%\rb{Do we want to discuss variations involving frame pointers or more implicit
-%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
-%\rb{Where to talk about the role of the compiler?}
-
-%% Because the access rights of callers subsume those of callees, it is essential
-%% that the micro-policy change tags when the property thinks that a call occurs,
-%% but it is safe for the micro-policy to fail to reset to the caller's tag when the
-%% property thinks a return occurs. (Dually, it is safe for the micro-policy to
-%% change tags even without executing a call point, but not for it to reset tags to
-%% the caller when the property thinks execution is still in the caller).
-
-%% Because the property does not identify return sites, we have no way to require
-%% that ``intended returns'' are tagged correspondingly. Some unintended results
-%% might also be so tagged, but as long as the jump has the effect of a legal
-%% return, it is OK.
-
-\subsection{Validation through Random Testing}
-\label{sec:testing}
-
-We validate the micro-policy of the previous section by
-systematically testing that it satisfies eager stack integrity and
-confidentiality. We use a Coq specification of the RISC-V
-architecture\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
-%\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}
-and extend it with a
-runtime monitor implementing the stack-safety
-micro-policy.\footnote{This code is {\em redacted for review}.} We
-  chose the Coq proof assistant as the setting for our implementation
-  for two reasons: first, to ensure that our coinductive trace
-  definitions are well-formed, and reason about them, and to leverage the power of the
-  QuickChick property-based testing framework~\citep{Pierce:SF4}.
-
-To use QuickChick, we build random test-case generators that produce
-% \begin{enumerate}
-% \item
-(1) an initial RISC-V machine state, including most notably
-  the program to be executed;
-  % \item
-(2)
-  an initial policy state, tagging instructions corresponding to
-  blessed call or return sequences appropriately, while marking
-  all potential stack locations as $\tagNoDepth$; and
-% \item
-(3)
- a code map and a stack map, accurately characterizing the behavior
-  of callers and callees.
-% \end{enumerate}
-%
-That is, our generators need to produce not only a low-level program
-that exhibits interesting control-flow patterns, but also the policy
-meta-information and code annotations that would normally be produced
-by the compiler.
-
-\paragraph*{Generators}
-
-To write such generators we build on the work of
-\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
-introduced {\em generation by execution} to produce progams leading
-abstract machines towards interesting behaviors. Generation by
-execution receives as an input a partially instantiated machine state
-and attempts to generate an instruction (or a sequence of instructions
-such as a blessed sequence) that makes sense locally (e.g., jumps go
-to a potentially valid code location and loads read from from a
-potentially valid stack location). Then we step the machine and repeat
-the process until we generate or execute some target number of
-instructions, or we reach a point where the machine can't step
-anymore.
-
-We extend this technique to keep track of the control flow behavior of
-the program being generated: each time a call or return sequence is
-generated, we ensure that the appropriate policy tags and code
-annotations are set for the entry or return points. At the same time,
-we allow the generation to sometimes relax those constraints,
-introducing potentially ill-formed flows: this causes our programs
-to failstop when executed in conjunction with the policy monitor,
-but also allows for revealing errors in our setup.
-
-In addition, we need to extend this technique to handle the nested
-nature of our confidentiality property: rather than just generating two
-initial machines that are variants of one another and letting them
-execute to test for noninterference, we generate a new variant
-{\em every time a call is made} and check confidentiality for the
-subtrace produced from that variant state until its corresponding
-return. As a result, a ``single'' confidentiality test compactly
-checks the confidentiality of multiple nested calls.
-
-\paragraph*{Findings}
-
-To ensure the effectiveness of testing against our formal properties, we
-use {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
-(mutations) in a program that should cause the property of interest (here,
-stack safety) to fail, and ensure that the testing framework can find
-them. The bugs we use for our evaluation are either artificially generated
-by us (deliberately weakening the micro-policy in ways that we expect
-should break its guarantees), or actual bugs that we discovered through
-testing our implementation. We elaborate on some such bugs below.
-
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c}
-  Bug & Property Violated & MTTF (ms) \\
-  \hline
-      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
-      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
-      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:bug-table}
-\end{table}
-
-The mean-time-to-failure (MTTF) for various bugs can be found in
-Table~\ref{tab:bug-table}. All experiments were run in a desktop
-machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
-RAM.
-% \leo{Something like: There is really no baseline to compare this
-  %against, but include it here for the future/reasons?}
-%
-The artificially injected bugs focus on weakening the tag propagation
-and checking mechanism.
-%
-For example, when loading from a stack location, the correct
-micro-policy needs to enforce that the tag on the location being read
-is $\tagStackDepth{n}$ for some number $n$ and that the one of the
-current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
-that restriction by not checking the depth equality (row {\em
-  LOAD\_NO\_CHECK}).
-%
-Similarly, when storing to a stack location, the correct micro-policy
-needs to ensure that the tag on the memory location is either
-$\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing that constraint causes violations to the integrity
-property (row {\em STORE\_NO\_CHECK}).
-
-Naturally, testing also revealed a number of errors in our
-implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code;
-%\bcp{right?}\leo{yeah}
-we re-implemented it in Coq targeting RISC-V).  These errors range
-from trivial typos to ones that require an intriguingly complex setup
-to appear.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
-sequence %/policy combination\apt{??}
-did not initialize all locations for the
-newly allocated stack frame correctly, but left some of them as
-$\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured.
-The smallest counterexample requires calling a function {\tt f},
-which fails to initialize some of its frame during the blessed sequence,
-but writes into an uninitialized location $l$ later, treating \(l\) as outside
-the stack. Then {\tt f} calls a further function {\tt g} (which should have
-the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
-which is allowed because the enforcement mechanism still has
-$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
-%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
-%  and the write from {\tt f} was already a violation.}
-
-\section{Stack Safety for Lazy Enforcement}
-\label{sec:lazy}
-
-The conservative policy described in the previous section, while
-testable and enforceable, is rather slow to be of practical use. What
-we want, instead, is a somewhat more permissive policy that can also
-be efficiently implemented. Most of the performance overhead incurred
-stems from the need to set stack activation tags as a frame is created
-and reset them as it is destructed, as
-\citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
-mitigate those costs, they propose a number of optimizations.
-
-First, they eliminate the initialization pass. In the {\em Lazy Tagging}
-system, the stack is again initially tagged as \(\tagNoDepth\).
-A function is permitted
-to overwrite \(\tagNoDepth\) memory, and when it does so its depth tag overwrites
-the clear tag, claiming the memory; the remainder of the policy functions
-as {\em Depth Isolation}. On return, the callee must clear its frame, restoring
-it to \(\tagNoDepth\), which is still a heavy cost.
-
-To eliminate this cost, \citet{DBLP:conf/sp/RoesslerD18} augment Lazy Tagging
-with {\em Lazy Clearing}. Under this combination of optimizations,
-the policy does not initialize stack frames on entry \emph{or} clear them on exit,
-and it permits all writes to the stack, even when the $\PCname$ tag
-does not match the memory tag; the \(\PCname\) tag is always propagated
-to the written location. Reads from the stack \emph{do} require $\PCname$
-tag and memory tag to match. Thus, even if a callee illicitly writes to
-a private location in its caller, the caller will eventually detect this
-if it ever tries to read from that location.
-%
-These \emph{lazy policies} admit more efficient implementations, but
-they deliberately allow temporary violations of stack integrity,
-with the checks deferred until the point a violation truly becomes
-harmful. The natural question then is: how do we characterize the
-protections provided by these policies once stepwise stack safety is broken?
-How does a harmful violation that must be caught later differ from a
-harmless one?
-
-\paragraph*{Deferring Integrity Checks}
-
-We first notice that these properties prevent invalid reads, as well as reads
-that follow invalid writes, but the writes themselves are permitted.
-So we wish in particular to find an integrity property that characterizes them.
-The intuition is that sealed memory can change, but it is
-then tainted, and its original owner cannot actually read it. This
-leads us to a hybrid of integrity and confidentiality.
-
-\definition
-A system enjoys \textbf{\textit{deferred stepwise integrity}} if,
-for any \(\MPCT,\brk \in \MPCT^*\) where \(\head(\MPCT) = \mach,\pol,\context\) and
-\(\last(\MPCT) = \mach',\pol',\context'\), the following holds. Let
-\((\mach',\pol',\context') \hookrightarrow \MPCT'\) and
-\(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
-set of sealed components that changed during the call.
-For any \(\nach'\) and \(\NPCT'\) such that \(\mach' \approx_{\components} \nach'\) and 
-\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
-
-For example, consider again the integrity violation in
-Figure~\ref{fig:int1}. Under a deferred discipline, the write to {\tt
-  x} in instruction 11 taints the variable but does not in itself
-constitute a violation, which happens only when the tainted value is
-read in instruction 4. However, if this instruction added the constant
-that was originally stored in {\tt x} without reading the variable,
-i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
-Instead of strictly forbidding such dangerous writes, we say,
-``No reads after dangerous writes.''
-
-\subsection{Observability}
-
-%\apt{Given that the deferred property suffices to specifcy the lazy policies,
-%  the case for observable properties is pretty weak.  Should we just cut them?}
-
-While deferred integrity meaningfully captures the intuition
-behind lazy tagging, we seek properties that can be used to evaluate the broadest
-possible range of reasonable
-enforcement mechanisms. Consider a hypothetical
-taint-tracking policy that allows high-security to be copied as long as it is
-tagged as such and not exfiltrated, or that allows dangerous writes to be
-copied as long as they don't affect critical operations. This calls for
-an abstract mechanism to characterize behaviors that are externally relevant.
-
-So we extend our model with a notion of observations. We abstract over an
-observation type \(\obs \in \OBSS\), containing at minimum the silent observation
-\(\tau\). Our machine step function now takes a state and returns a state and an observation,
-which is carried over into the policy and context step functions:
-%
-$\mach \stepsto[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS $.
-%
-
-\paragraph*{Observations-Of}
-
-The ``observations-of'' operator, written \(\obsof\), takes a machine trace and
-gives the trace of observations of each step in
-the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
-initial observation, a machine-trace, and an observation trace:
-
-\begin{minipage}[b]{.6\textwidth}
-  \judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
-                {\(\mpcstatename' \hookrightarrow \MPCT\)}
-                {\(\mathit{obsfor} ~ \obs' ~ \MPCT = \obsT\)}
-                {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}%
-\end{minipage}
-\begin{minipage}[b]{.35\textwidth}
-  \judgment{}
-           {\(\mathit{obsfor} ~ \obs ~ \mpcstatename = \obs\)}
-\end{minipage}
-
-\begin{minipage}{.35\textwidth}
-  Then we start with a \(\tau\) observation:
-\end{minipage}
-\begin{minipage}{.6\textwidth}
-\judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
-         {\(\obsof(\MPCT) = \obsT\)}
-\end{minipage}
-
-\paragraph*{Observational Similarity}
-
-We say that two observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
-as far as an external observer is concerned, written \(\obsT_1 \eqsim
-\obsT_2\), if the sequence of non-silent observations is the same. That is, we
-compare up to deletion of \(\tau\) observations, coinductively:
-
-\begin{minipage}{.2\textwidth}
-  \judgment{}{\(\obsT \eqsim \obsT\)}
-\end{minipage}
-\begin{minipage}{.2\textwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.2\textwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
-\end{minipage}
-\begin{minipage}{.2\textwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
-\end{minipage}
-
-%\leo{I still don't know how to do this transition. Maybe contrast
-%with the ``perhaps surprisingly'' later?} Note that
-%similarity of observation traces could alternatively be defined coinductively,
-%just like the prefix relation, but by dropping the asymmetric rules that allow
-%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}
-
-Note that an infinite silent trace is a
-prefix of (and similar to) any other trace. While this might seem
-surprising at first, it makes sense in a timing-insensitive context:
-an external observer looking at two machine runs cannot (computably)
-distinguish between a machine that steps forever and a machine that
-steps for a long time before producing some output.
-
-\subsection{Observational Stack Safety}
-
-We now can sketch observational versions of both integrity and confidentiality.
-Integrity is already straightforward, simply weakening the condition that must hold
-after the return---instead of preventing dangerous writes from being read, we
-prevent them from being leaked.
-
-\definition
-{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
-where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
-the following holds. Let \(\components\) be the set of protected components that changed during
-the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
-and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
-Then for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\), and its induced trace
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
-
-\medskip
-
-Confidentiality is more sophisticated. We must capture the intuition that
-secrets do not escape during a call {\em and} are not leaked during a return.
-This require us to be able to quantify over components that fail to preserve
-our \(\diamond\) relation.
-
-\definition Let \(\mach,\mach'\) and \(\nach,\nach'\)
-be pairs of states. Their {\em unsafe set}, written
-\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all components
-\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
-\(\mach' ~ \component \not = \nach' ~ \component\).
-
-\definition
-A system enjoys {\em observational confidentiality} if for all
-\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\pol,\context\) and
-\(\components = \secret(\context)\), the following holds.
-%
-For any \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
-\(\MPCT\) and \(\NPCT\) respect three conditions:
-
-\begin{itemize}
-\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
-  meaning that no secrets leak during the call
-\item \(\brk ~ (\last(\MPCT)) \leftrightarrow \brk ~
-(\last(\NPCT))\), i.e. if one call returns, so does the other
-\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
-  and \(\pi_\mach(\last(\NPCT)) = \nach'\),
-  let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\),
-  and let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
-  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
-  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
-  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
-\end{itemize}
-
-This last condition is the most interesting: we identify the components that
-contain leaked information on either side of the first pair of variant traces.
-This becomes the basis for the second variation, because all of these values
-must be protected from leaking for the remainder of the program.
-
-Together with return-after-call, these two properties give us
-{\em Observable Stack Safety}. This is a weaker property
-than the stepwise one, and also much more extensional.
-We can think of stepwise, deferred, and observational as
-three points along a spectrum from eager to lazy. The first two
-are useful for understanding particular enforcement mechanisms,
-but if we had to choose a single property to represent stack safety,
-it would be observational state safety.
-
-\subsection{Connection to Policies}
-
-Observable properties allow us to defer enforcement until a property
-violation would become visible. Now we can identify when lazy policies
-miss violations that can become visible. The {\em Lazy Tagging and Clearing}
-micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
-at which it was written, and enforces that it must be read from the
-same depth. However, this allows violations of observable stack safety!
-Specifically, a callee might write into the stack and leave tags behind
-for a future callee at the same depth to read. This violates confidentiality and,
-if the writes were in the caller's frame, integrity.
-
-{\em Lazy Tagging and Clearing} can be repaired if, instead of tagging each function
-activation with its depth in the stack, we generate a fresh activation
-identifier on each call, which prevents a program from exploiting stale
-tags from previous activations. A related mechanism was explored in the
-Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
-associates a unique activation identifier to each function, which is
-however shared by all activations of the same function in order to obtain
-good cacheability of tag rules---but that policy allows leaks between
-instances of the same function.
-
-We believe that {\em Lazy Per-activation Tagging and Clearing}, in which
-each activation is tagged uniquely, ends up enforcing the combination
-of deferred stepwise stack integrity and stepwise confidentiality.
-This is because tags that persist between calls will not become active again.
-Our properties have allowed us to reject an enforcement mechanism as
-not really stack safe, and describe a potential fix that itself is a
-prime candidate for future testing.
-
-\section{Extensions}
-\label{sec:ext}
-
-We have shown how our properties apply to an extremely simple
-subroutine model and validated them through randomized testing.
-But what about more realistic systems? We now demonstrate the
-flexibility of our core concepts by extending 
-them to a system that models the passing of
-parameters on the stack, and then to a very simple coroutine model.
-
-\subsection{Parameters Passed on the Stack}
-\label{sec:passing}
-
-\newcommand{\mainpassc}{magenta}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-void main() {
-  int x = 5;
-  x = f(x);
-  return;
-}
-
-void f(int a) {
-  int y = 2*a;
-  a = 0;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.6\textwidth}
-{\small
-\begin{verbatim}
-1  main: add $1,%sp          ; allocate frame
-2        mov $5,-1(%sp)      ; initalize x
-3        add $1,%sp          ; allocate arg
-4        mov -2(%sp),-1(%sp) ; set arg
-5        call f              ; SHARE: %sp-1
-6        sub $1,%sp          ; deallocate arg
-7        mov %ra,-1(%sp)     ; store x
-8        sub $1,%sp          ; deallocate frame
-9        ret
-
-10  f:   add $1,%sp          ; allocate frame
-11       mov -3(%sp),-1(%sp) ; y = a
-12       mul $2,-1(%sp)      ; y = a*2
-13       mov $0,-3(%sp)      ; a = 0
-14       sub $1,%sp          ; deallocate frame
-15       ret
-\end{verbatim}
-}
-  \end{subfigure}
-\begin{subfigure}{.65\textwidth}
-\vspace*{0.2in}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 3} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{5}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[S(0)]%
-  \memory{1}{\mainpassc}[P(0,1)]%
-  \memory{1}{\retptrc}[RetPtr]%
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  {\tt 13} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  {\tt 14} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  \\
-  {\tt 8} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{10}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{\#6}
-    \MemoryLabel{-7em}{0.75em}{10}
-    \vspace{.5em}
-  } \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-
-\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
-    the figure (at least in the latex build that I'm looking at...)}\fi}
-\label{fig:passingsimple}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-int *stash = 0;
-int main() {
-  int x, y;
-  f(x);
-  x = 0;
-  f(y);
-  return x;
-}
-
-void f(int &a) {
-  if(!stash)
-    stash = &a;
-  *stash = 5;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.49\textwidth}
-{\small
-\begin{verbatim}
-   stash: .word 0
-1  main:  add $2,%sp      ; allocate frame
-2         mov %sp,%ra
-3         add -2,%ra      ; ra = &x
-4         call f          ; SHARE: %sp-2
-5         mov $0,-2(%sp)  ; x = 0
-6         mov %sp,%ra
-7         add -1,%ra      ; ra = &y
-8         call f          ; SHARE: %sp-1
-9         mov -2(%sp),%ra ; set return value
-10        sub $2,%sp      ; deallocate frame
-11        ret
-
-20 f:     cmp stash, $0   ; stash = 0?
-21        breq #13        ; branch if so
-22        mov %ra,stash   ; stash = &a
-23        jmp #15
-24        mov $5, (stash) ; *stash = 5
-25        ret
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.29\textwidth}
-    \begin{center}
-    \begin{tabular}{l l}
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{S(0)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$ \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}
-      \memory{1}{\mainsealc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{5}
-      \\
-      {\tt 8} &
-      \memoryaddrs[(0)]{12em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 20} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{S(0)}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
-      \\
-      {\tt 25} &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\mainpassc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{\bf 5}
-\end{tabular}
-\end{center}
-\vspace{\abovedisplayskip}
-\end{subfigure}
-\caption{A violation with pass-by-reference}
-\label{fig:passing}
-\end{figure}
-
-To this point we have ignored function arguments, since our simplest model places them in
-registers. We now show how to implement two kinds of protected
-sharing: passing by value on the stack, and passing by reference. We first explore
-some examples at the same level of detail as our earlier ones, and then move on to
-extending our formalization.
-
-Example~\ref{fig:passingsimple} shows a sample trace of
-a program that passes a value on the stack, without an attack.
-We extend our call annotation with an additional parameter identifying a memory location to be
-passed (the diagram displays it in magenta). Our notion of domains is extended as well,
-to include \(\passed{\depth,\depth'}\), recording two depths. We note the
-current depth of the active function next to the stack pointer in the diagram, for reference.
-The argument {\tt a} has
-its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
-as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
-modified, and naturally this is permitted as well.
-
-The same annotation mechanism can be used to pass
-variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference.
-{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
-{\tt f} again passing a reference to {\tt y}.
-We would expect that the first call to {\tt f} might modify
-{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
-
-To avoid cluttering the picture at the assembly level, we revert to assuming that the
-argument {\tt a} is passed in a register, namely {\tt \%ra}.
-The question is how the local variables {\tt x} and {\tt y} themselves
-should be labelled in order for the policy to detect an integrity violation.
-At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
-return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity.
-%For that matter
-%a future call with no arguments passed by reference would also violate integrity if it wrote
-%to either variable.
-
-When passing by reference, it is possible to repeatedly pass the same reference further
-and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices for the domain to track two depths for each passed
-cell: the original owner, and the highest depth it has been passed to. The policy considers
-it to be accessible anywhere between the two.
-
-Our model provides only limited support for handling fully first-class pointers into
-stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike the case for pass-by-reference, there is no easy way to model the intended
-sharing behavior of such pointers in terms of stack frame activation patterns.
-The best approximation we can make is to treat an address-taken variable in the
-same way as a passed argument, with the additional property
-that it remains accessible at \emph{all} stack depths above the caller.
-This prevents most violations of temporal safety, although stale pointers can still be
-used in certain corner cases.
-But more seriously, callees are permitted to access the location without having
-been explicitly given a pointer to it.  Preventing the latter kind of abuse
-is the provence of a fine-grained memory safety policy, e.g. based on capabilities
-(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
-\emph{per se}.
-
-\paragraph{Formalization of variable passing.}
-We add extra information to our call
-annotations to reflect that they may now pass variables on the stack, and that the locations
-of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the components
-that should be passed, expressed as offsets from the stack pointer.
-%
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \WORDS)\]
-
-We extend the set of domains to include passed components,
-which are labeled with the range of depths of the caller that has passed them.
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
-The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-\times (\listT (\target)) \]
-The initial context again maps all components to \(\unsealed\).
-We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map at a call from depth \(\depth\).
-\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
-\begin{cases}
-  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and }
-                                 \component - (\mach ~ \rsp) \not \in P \\
-  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and }
-                                 \component - (\mach ~ \rsp) \in P \\
-  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \\
-                               & \text{ and } (\mach ~ \rsp) \in P, \text{ for any } \depth' \\
-  \domMap~\component & \text{else} \\
-\end{cases}\]
-
-At the matching return we just unseal the components at depth \(\depth\):
-%
-\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
-\begin{cases}
-  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
-  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
-                              \text{ for any } \depth' < \depth \\
-  \domMap~\component        & \text{else} \\
-\end{cases}\]
-
-Now we can inductively define the context update function; these rules again apply in the order
-presented. \ifspace\bcp{Too much space between the second and third.}\fi
-
-\judgmenttwo[ Call]
-            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
-            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmenttwo[ ReturnFound]
-            {\(\mach \stepsto \mach' \in \target\)}
-            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmentthree[ ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-5ex}
-\judgment[ Default]
-         {}
-         {\(\underscore,\context \stepsto_C \context\)}
-
-\definition Let the protected set of a context, \(\prot_P(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
-or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
-\(|\rts| > \depth'\).
-
-\definition Let the \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\) or
-\(\domMap ~ \component = \passed{\depth,\depth'}\)
-where \(\depth \leq |\rts| \leq \depth'\).
-
-\definition
-For each of the stepwise, deferred, and observational versions of
-stack integrity in the simple model, the equivalent property in the
-passing model substitutes the above \(\CONTEXTS\) type and update
-function, and \(\prot_P\), for their equivalents in the simple model.
-
-For both the stepwise and observational versions of stack confidentiality in
-the original model, the passing model equivalents additionally substitute \(\secret_P\).
-
-\subsection{Coroutines}
-
-Our final extension to the model adds coroutines. We assume a very simple model with
-a fixed number of coroutines, each with its own stack. 
-Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids.  Annotations and code maps are unchanged from
-the subroutine model.
-%\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-%\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
-
-The first major difference in this model is the domains. Stack domains are now nested inside
-top-level domains, and associated with stack identifiers.
-Our context now consists of a domain map, a map from stack identifiers to return target stacks,
-an additional map of targets for yields, and a stack identifier of the active stack.
-A yield target map \(\yts\) is a map from stack identifiers to targets.
-In a standard system an initial yield target map will map each stack identifier to the
-set of all states with their stack pointer at the stack base, and include any restrictions
-on entry points. Just as a return target tells us that we have really returned, a yield
-target tells us that we have really yielded. When yielding from state \(\mach\),
-\(\ycon ~ \mach \in \MACHS\) gives us the yield target for coming back to the yielding
-coroutine.
-%
-\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
-\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
-\[(\domMap,\rtm,\yts,\stackid) = \context \in \CONTEXTS ::= 
- (\COMPONENTS \rightarrow \STKDOMS) \times
-(\STACKS \rightarrow \listT(\target)) \times (\STACKS \rightarrow \target) \times \STACKS\]
-
-The initial domain map maps each stack address in stack \(\stackid\) to
-\(\instack{\stackid}{\unsealed}\), the initial return map maps all stack identifiers to empty
-return stacks, the initial yield targets are defined above, and the initial stack id matches the initial
-coroutine. Together these form the initial context.
-
-In our context update rules, calls and returns work the same as in the sharing model,
-with domain updates restricted to the top-level domain corresponding to the active stack,
-and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
-
-\judgmentfour[ Yield]
-             {\(\mach \stepsto \mach'\) ~ ~}
-             {~ ~ \(\stackid' = \stackof ~ (\mach'~ \rsp) \not = \stackid\) ~ ~}
-             {~ ~ \(\yts ~ \stackid'\) ~ ~}
-             {~ ~ \(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
-             {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
-
-This calls for some explanation. First, we determine which coroutine we are yielding to by
-kchecking which stack the stack pointer of the next state is pointing to. Then we can only yield
-if the next state is a valid yield target for that stack, and we update the yield target for
-the current stack, ensuring that when the system yields back we expect the current coroutine
-to continue from where it left off. So, it is possible for a coroutine to claim to yield
-but not actually change the active stack, with similar effects to when a callee fails
-to return properly. Now we can define both stack safety and coroutine safety under this model.
-
-Stack integrity and confidentiality apply in the coroutine model, based on straightforward
-projections of each stack's domains, etc. and the machinery of the passing model. We will give
-the coroutine-specific properties.
-
-%\definition
-%The protected and secret sets of a context as used for calls,
-%\(\prot_C(\domMap,\rtm,\yts,\stackid)\) and \(\secret_C(\domMap,\rtm,\yts,\stackid)\),
-%are equivalent to their counterparts \(\prot_P\) and \(\secret_P\) applied to a domain map
-%that projects out the stack domains of components in stack \(\stackid\) and a
-%return target stack \(\rts = \rtm ~ \stackid\). Intuitively, within the active stack we protect
-%the same components we would have in the passing model.
-
-%\definition
-%For each of the stepwise, deferred, and observational versions of
-%stack integrity in the simple model, the equivalent property in the
-%passing model substitutes the above definitions of \(\context\), initial contexts and update
-%function, and \(\prot_C\) for their simple equivalents, and is otherwise identical.
-
-%\definition \(\MPCT^*\) is now defined as
-%\[\begin{aligned}
-%\bigcup_{\stackid \in \STACKS,\depth \in \mathbb{Z}} \hspace{-1em} \{\MPCT | \MPCT\langle f_\depth \rangle
-%\text{ is a reachable segment}\} \hspace{1em} &
-%\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rtm,\yts,\stackid')) \triangleq |\rtm ~ \stackid| \geq \depth \\
-%\end{aligned}\]
-%and \(\brk{\MPCT}\) similarly projects the depth of the active stack.
-
-%\definition
-%For either of the stepwise and observational versions of
-%stack confidentiality in the passing model, the equivalent property in the
-%coroutine model substitutes the above definitions.
-
-\definition \(\MPCT^\wedge\) is the \emph{yield set}, defined as
-\[\begin{aligned}
-\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{\MPCT, \neg f_\stackid | \MPCT\langle f_\stackid \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\stackid ~ (\underscore,\underscore,(\underscore,\underscore,\underscore,\stackid')) \triangleq \stackid \not = \stackid' \\
-\end{aligned}\]
-
-Here, we are capturing the entire span that the system spends {\em away} from a given
-coroutine.
-
-\definition
-Let the protected set of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
-
-\definition
-Let the secret set of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
-be the set of components \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
-
-\definition
-We define \textbf{\em stepwise coroutine integrity} in the same way as stepwise stack integrity,
-substituting the above \(\CONTEXTS\) type, initial context and update
-function, and \(\prot_Y\) for their equivalents. Likewise
-\textbf{\em deferred coroutine integrity} and \textbf{\em observational coroutine integrity} are defined
-in terms of their simple equivalents using \(\MPCT^\wedge\) in place of \(\MPCT^*\).
-
-\definition
-We define \textbf{\em stepwise coroutine confidentiality} in the same way as stepwise stack
-integrity, substituting the above definitions. Likewise
-\textbf{\em observational coroutine confidentiality} is defined in terms of observational
-stack confidentiality.
-
-\section{Related Work}
-\label{sec:relwork}
-
-\paragraph{Formal Stack Safety on Capability Machines}
-%
-To our knowledge, the only line of work to date that has attempted
-a positive and formal characterization of
-stack safety begins with \citet{SkorstengaardLocal}, who introduce a calling
-convention that uses local capabilities to preserve local state
-encapsulation and well-bracketed control flow
-using a logical relation to reason about the stack safety of concrete
-programs.
-% Although the required hardware support is readily available
-%in capability machines like Cheri,
-%% this technique incurs significant
-%% costs, because
-%it requires the entire unused part of the stack to be
-%cleared whenever a security boundary is crossed.
-Their logical
-relation captures capability safety without ``externally observable
-side-effects (like console output or memory access traces)''
-and can be used to reason about individual programs.
-\citet{SkorstengaardSTK} acknowledge the difficulty of defining
-these concepts precisely:
-``while [they] claim that [their] calling
-convention enforces control-flow correctness, [they] do not prove a
-general theorem that shows this, because it is not clear what such a
-theorem should look like,'' noting that the correctness property
-enjoyed by their technique ``is not made very explicit.''
-
-StkTokens \citep{SkorstengaardSTK} continues this work by defining a more efficient
-calling convention that makes use of linear capabilities for stack and return
-pointers.
-%% The convention operates on a single shared stack and requires that
-%% protected components avoid compromising their own security by following certain
-%% simple rules---like not leaking their private capability seals. More
-%% precisely but
-Still informally, local state encapsulation is defined as
-restricting accesses to the range of memory allocated to the current stack
-frame, and well-bracketed control flow as only allowing returns from the topmost
-frame to the immediately adjacent frame below. Formally, it improves on
-\citet{SkorstengaardLocal} by building those properties into the semantics of a
-capability machine with a built-in call stack and call and return instructions,
-which is proven fully abstract with respect to a more concrete capability
-machine.
-%% that replaces those pseudo-instructions with their calling convention.
-Their proof of full abstraction uses a standard notion of components, which
-import and export functions through their interfaces.
-%% ; their model of
-%% observations is limited to cotermination.
-
-\citet{SkorstengaardSTK} enumerate a useful list of criteria for judging
-proposed stack safety properties: they should be (1) intuitive; (2)
-useful for reasoning; (3) reusable in secure compiler chains; (4)
-arguably ``complete;'' and (5) potentially scalable.  We believe our
-approach competes well with theirs on criteria (1), (4), and (5).
-Their approach is specifically set up to address (3), while
-ours is not. Addressing (2) remains as future work for us.
-
-%% Although there are some important differences between
-%% \citeauthor{SkorstengaardSTK}'s properties and our own, the new
-%% properties arguably satisfy their list of desiderata for such
-%% formulations: notably, they reflect (and refine) programmer intuitions
-%% about stack safety, furnish a comprehensive picture of the meaning of
-%% stack safety, can be used to reason about program behavior, and are
-%% flexible enough to be allow extensions to richer settings.
-%% %
-%% \sna{Some of this seems like more detail than we need, and we need to
-%%   compare ourselves to their list of criteria.}
-%% \rb{Trimmed this a bit. Are you referring to their list of attack
-%%   examples, or to something else?\apt{Their list at start of section 4.
-%%   \rb{Had a go at this.}}}
-
-\paragraph{Protecting the Stack with Micro-Policies}
-%
-\citet{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
-attacks against stack data are in scope (but not side channels or hardware
-attacks), and study the protection of stack data through three families of
-micro-policies that tag stack objects with a pair of frame and object identifiers
-used to validate accesses to the stack: {\em Return Address Protection} (which
-prevents an adversary from overwriting designated return addresses), {\em Static
-Authorities} (which only allows the code of a function to access the stack frames
-of its own dynamic instances), and {\em Depth Isolation} (described in
-\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
-temporal locality of stack memory and local call graphs, as well as information
-generated by the compilation toolchain, to strike various balances between
-precision and cacheability, evaluated through benchmarks that demonstrate
-limited performance overhead. In addition to the baseline ``eager'' policies,
-they propose a number of lazy optimizations, notably {\em Lazy Tagging} and
-{\em Lazy
-Clearing}, discussed in \cref{sec:lazy}, with associated improvements in
-performance.
-
-We have evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
-  Clearing} against our properties, finding that {\em Depth Isolation} does
-indeed enforce stepwise stack safety and therefore also observable stack
-safety. We also found that neither lazy micro-policy actually does enforce
-stack safety, and proposed a fix.
-
-\paragraph{Heap Safety as a Security Property}
-
-Heap safety, like stack safety, may be framed as a security property
-in the form of
-noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
-be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
-(or ask arthur how!).}\fi. Just
-as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
-characterization of the meaning of (heap) memory safety, in this paper,
-inspired by their work, we aim to do the same for the stack. They are a
-significant inspiration for this work.
-
-\section{Future Work}
-\label{sec:future}
-
-We have introduced a characterization against which stack safety enforcement
-mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
-Our immediate next step is to implement testing of observational properties,
-both to validate them and to explore how to effectively test properties that do not
-fail immediately.
-From there there are two natural ways to extend the work. We should thoroughly evaluate
-other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
-the functionality of our definitions by showing that they can be used to prove
-the correctness of a programming logic with strengthened rules around calls and returns.
-
-\paragraph*{Stack Safety in the Cheri Ecosystem}
-
-There are several proposals around the use of Cheri capabilities to enforce stack safety,
-ranging from mechanisms that use the standard Cheri hardware (which includes local
-capabilities) \citep{SkorstengaardLocal},
-to those that propose entirely new types of capabilities, such as linear
-\citep{SkorstengaardSTK}, uninitialized \citep{Georges+21}, lifetime \citep{Tsampas+19}, and monotonic \citep{GeorgesComplete}.
-Of these, uninitialized capabilities seem particularly promising---they naturally enforce
-write-before-read behavior, and therefore will enable confidentiality without the need
-to clear large swaths of the stack. We plan to implement a model of the stack safety
-system described in \citet{Georges+21} and test it against our properties using
-QuickChick.
-
-\paragraph*{Program Logic for Stacks}
-
-Stack integrity suggests a reasoning principle in which knowledge about a caller's
-local data can be carried across a call site. We propose to develop this notion
-into a low-level separation logic with a ``call rule'' that preserves propositions
-whose domains are sealed across a call, while allowing those propositions proven
-about the behavior of the callee to apply to the accessible stack.
-
-It is straightforward to imagine that such a rule could be proven sound in a
-system with lockstep stack safety. But there is a real novelty in considering
-what kind of soundness can be proven from a system that obeys only our observational
-properties.
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Bibliography
 \bibliography{bcp.bib,local.bib}

### 2021-05-11 short/main.tex
Fixed build issue, added makefile
Polishing on short version

@@
 %% use defaults for review submission.
-%\acmJournal{PACMPL}
-%\acmVolume{1}
-%\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
-%\acmArticle{1}
-%\acmYear{2021}
-%\acmMonth{1}
-%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-%\startPage{1}
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
+\acmYear{2021}
+\acmMonth{1}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
 
@@
 
-We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
+Our testing supports the stack-safety claim of Roessler and Dehon's {\em Depth
+  Isolation} micro-policy, in
 which memory cells within each stack frame are tagged with the identity of
 the function activation that owns the frame and access to those locations is
-then permitted only when that activation is currently executing, validates our
-stepwise properties. On the other hand, we reason that \emph{Lazy Tagging and Clearing}
+then permitted only when that activation is currently executing. On the other hand, we find that their \emph{Lazy
+  Tagging and Clearing} policy
 violates the temporal aspect of confidentiality in
@@
 and also violates integrity if the leak happens to use the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
-confidentiality, albeit at some performance cost.
+propose a variant of {\em Lazy Tagging and Clearing} that does enforce
+confidentiality (and test that it does), albeit at some performance cost.
 

### 2021-05-13 short/main.tex
Working on Andrew's feedback

--- 2021-05-10 short/main.tex
@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmtog,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
+\citestyle{acmnumeric}   %% For author/year citations
 
@@
 We propose such a characterization,
-using the technical framework of language-based security: a
-combination of an {\em integrity property},
-a {\em confidentiality property}, and a {\em well-bracketedness property}.
-%
-In a setting containing unstructured machine code, true
-control-flow attacks are only meaningful when they bypass a caller that was
-waiting for its return, so well-bracketness guarantees that callees return
-only to their callers. Other control-flow attacks, like
-return-oriented-programming (ROP), appear to return but keep
-the callee in control while accessing the caller's data.
-%
-Our properties treat this scenario as the callee not returning at all,
-so if the ROP attack touches the caller's data, it violates integrity
-or confidentiality.
+using the technical framework of language-based security.
+Stack safety protects a caller from its callee, guaranteeing
+the {\em integrity} and {\em confidentiality} of the caller's local state
+until it regains control. This formulation not only captures the intuition
+that the callee cannot directly access the caller's state, but gives a novel
+way of looking at control-flow attacks such as return-oriented-programming (ROP).
+In an ROP attack, the callee pretends to return to its caller, thus accessing the
+caller's privileges, but does so in such a way that it is still in control of
+execution. Our model allows the caller to dictate the terms of a valid return --
+typically, that the stack pointer is restored and execution proceeds from the
+return address. An ROP attack that ``returns'' and reads the caller's state is
+therefore treated precisely as if the callee had never returned, but attempted
+to read the caller's state directly. It is a confidentiality violation. Likewise
+an attack that would corrupt the caller's state is an integrity violation.
 
-Integrity and confidentiality are also novel in that they
-are ``nested'': {\em each} caller is protected from its
-callees. They come in {\em stepwise} variants, in
-which a caller's data is {\em never}
+We formalize integrity and confidentiality as trace properties, in two different
+variants: {\em stepwise} variants, in which a caller's data is {\em never}
 read or modified during a call, and {\em observational} ones, in which
@@
 even if it does not prevent every single dangerous read or write.
+Confidentiality is especially interesting, as it is based on the traditional
+notion of noninterference. But where noninterference is normally presented as an
+end-to-end hyperproperty, stack confidentiality is noninterference applied over
+multiple nested subtraces -- one for every call.
 

### 2021-05-13 short/main.tex
Polishing / comments

@@
 consequences ranging from leakage or corruption of private stack data to
-control-flow hijacking. To prevent or detect such attacks, a profusion of
-software and hardware protections have been proposed,
+control-flow hijacking. A profusion of
+software and hardware protections have been proposed for detecting or
+preventing such attacks,
 %
@@
 the {\em integrity} and {\em confidentiality} of the caller's local state
-until it regains control. This formulation not only captures the intuition
+until control returns to it. This formulation not only captures the intuition
 that the callee cannot directly access the caller's state, but gives a novel
-way of looking at control-flow attacks such as return-oriented-programming (ROP).
+perspective on control-flow attacks such as return-oriented-programming (ROP).
 In an ROP attack, the callee pretends to return to its caller, thus accessing the
 caller's privileges, but does so in such a way that it is still in control of
-execution. Our model allows the caller to dictate the terms of a valid return --
+execution\bcp{that's not very clear to me}. Our model allows the caller to
+dictate the terms of a valid return -- 
 typically, that the stack pointer is restored and execution proceeds from the
@@
 therefore treated precisely as if the callee had never returned, but attempted
-to read the caller's state directly. It is a confidentiality violation. Likewise
-an attack that would corrupt the caller's state is an integrity violation.
+to read the caller's state directly: it is a confidentiality
+violation. Likewise, an attack that would corrupt the caller's state is an
+integrity violation.
 
-We formalize integrity and confidentiality as trace properties, in two different
-variants: {\em stepwise} variants, in which a caller's data is {\em never}
+We formalize integrity and confidentiality as trace properties\bcp{hyperproperties??}, in two different
+variants: {\em stepwise} variants, in which a caller's data must {\em never} be
 read or modified during a call, and {\em observational} ones, in which
@@
 tool for Coq, to generate random programs and check
-that these micro-policies correctly abort programs that
+that Rossleer and Dehon's micro-policies correctly abort programs that
 would violate stack safety.

### 2021-05-13 short/main.tex
Nit

@@
 To demonstrate the utility of our properties, we use them
-to validate an existing mechanism, the
+to evaluate an existing mechanism, the
 {\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented

### 2021-05-14 short/main.tex
Discussion from today and some additional detail

--- 2021-05-13 short/main.tex
@@
 the {\em integrity} and {\em confidentiality} of the caller's local state
-until it regains control. This formulation not only captures the intuition
-that the callee cannot directly access the caller's state, but gives a novel
-way of looking at control-flow attacks such as return-oriented-programming (ROP).
-In an ROP attack, the callee pretends to return to its caller, thus accessing the
-caller's privileges, but does so in such a way that it is still in control of
-execution. Our model allows the caller to dictate the terms of a valid return --
-typically, that the stack pointer is restored and execution proceeds from the
-return address. An ROP attack that ``returns'' and reads the caller's state is
-therefore treated precisely as if the callee had never returned, but attempted
-to read the caller's state directly. It is a confidentiality violation. Likewise
-an attack that would corrupt the caller's state is an integrity violation.
+until it regains control.
 
@@
 even if it does not prevent every single dangerous read or write.
+
+%SNA: will this audience know "hyperproperty?"
 Confidentiality is especially interesting, as it is based on the traditional
@@
 end-to-end hyperproperty, stack confidentiality is noninterference applied over
-multiple nested subtraces -- one for every call.
+multiple nested subtraces -- each individual call's behavior must be invariant
+regardless of the state of the stack at its entry.
+
+This formulation not only captures the intuition
+that the callee cannot directly access the caller's state, but gives a novel
+way of looking at control-flow attacks. We are not interested in attacks that
+merely aim to execute arbitrary code -- in fact, our attacker model assumes
+that the (attacking) callee may already execute arbitrary code.
+%
+By extension, we do not place any particular restrictions on control flow.
+Perhaps suprisingly, we can still meaningfully distinguish the caller from
+its callee, and talk about protecting the caller's data. After a call, control
+is considered to belong to the callee until the trace reaches a valid
+``return target,'' typically a state with the program counter at the next instruction
+and the stack pointer restored. If the caller's data are accessed before then,
+it is a violation of integrity or confidentiality.
+
+These properties can optionally be extended with a notion of {\em well-bracketed control flow}
+as in \citet{SkorstengaardSTK}, the system-wide control-flow property that callees always return
+to their immediate caller, if they return at all. This property is largely orthogonal to our
+data-protection properties.
 
@@
 that these micro-policies correctly abort programs that
-would violate stack safety.
+would violate stack safety. Furthermore, we check incorrect
+enforcement variants (both variants that we accidentally created
+during our re-implementation of the micro-policy and ones that we
+intentionally crafted to be broken in order to increase our confidence
+in testing and the enforcement mechanism itself). The testing framework
+is able to generate counterexamples that violate our properties but are not halted by
+incorrect enforcement mechanisms, increasing our confidence when it finds
+no counterexamples under the real enforcement mechanisms.
 
@@
   Isolation} micro-policy, in
-which memory cells within each stack frame are tagged with the identity of
+which memory cells within each stack frame are tagged with the depth of
 the function activation that owns the frame and access to those locations is
-then permitted only when that activation is currently executing. On the other hand, we find that their \emph{Lazy
-  Tagging and Clearing} policy
+then permitted only when that activation is currently executing.
+On the other hand, we find that their \emph{Lazy Tagging and Clearing} policy
 violates the temporal aspect of confidentiality in
@@
 and also violates integrity if the leak happens to use the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that does enforce
-confidentiality (and test that it does), albeit at some performance cost.
+propose a variant of {\em Lazy Tagging and Clearing} that should enforce observational
+integrity and confidentiality, albeit at some performance cost. The next step
+is to expand testing to these properties and validate the variant; efficiently
+testing observational properties presents unique challenges because a property
+violation can be arbitrarily far removed from the initial ``risky'' read or write.
 
-Finally, we demonstrate our model's flexibility by extending it with the passing
-of variables on the stack and to a simple coroutine model.
+Finally, we demonstrate our model's flexibility by extending it to different settings.
+First, we allow variables to be passed on the stack, allowing some of the state of
+the caller to be shared with the callee (and potentially with a further
+callee, and so on.) Then we expand the number of stacks as we move to a simple
+coroutine model, albiet one with statically bounded regions for each stack.
+
+In addition to testing lazy tag-based policies against our observational properties,
+we are adapting the uninitialized capability model of \citet{Georges+21} to work with
+our testing framework for validation against our properties. Of the several Cheri-based
+stack protection schemes, its treatment of unitialized frames seems most compatible
+with our model of confidentiality, without the need for expensive stack clearing
+between calls. We expect that Cheri-based techniques support our properties if we
+assume extra cooperation from callers, namely that they do not leak their capabilities
+to the attacker. There is no equivalent dynamic requirement in a tag-based system because
+the policy can dynamically enforce cooperation.
 

### 2021-05-14 short/main.tex

@@
 
-We formalize integrity and confidentiality as trace properties, in two different
-variants: {\em stepwise} variants, in which a caller's data is {\em never}
+We formalize integrity and confidentiality as trace (hyper-)properties, in two different
+variants: {\em stepwise} variants, in which a caller's data must {\em never} be
 read or modified during a call, and {\em observational} ones, in which
@@
 
-%SNA: will this audience know "hyperproperty?"
 Confidentiality is especially interesting, as it is based on the traditional
 notion of noninterference. But where noninterference is normally presented as an
-end-to-end hyperproperty, stack confidentiality is noninterference applied over
+end-to-end hyper-property, stack confidentiality is noninterference applied over
 multiple nested subtraces -- each individual call's behavior must be invariant
@@
 tool for Coq, to generate random programs and check
-that these micro-policies correctly abort programs that
-would violate stack safety. Furthermore, we check incorrect
+that Roessler and Dehon's micro-policies correctly abort programs that
+would violate stack safety.
+Furthermore, we check incorrect
 enforcement variants (both variants that we accidentally created

### 2021-05-14 short/main.tex
Polishing on the first part

@@
 The call stack is a perennial target for low-level attacks, leading to
-consequences ranging from leakage or corruption of private stack data to
+consequences  from leakage or corruption of private stack data to
 control-flow hijacking. A profusion of
@@
 
-We propose such a characterization,
-using the technical framework of language-based security.
-Stack safety protects a caller from its callee, guaranteeing
+We propose such a characterization
+using the tools of language-based security: The informal claim that
+``stack safety protects a caller from its callee'' amounts to saying
+that it guarantees
 the {\em integrity} and {\em confidentiality} of the caller's local state
@@
 
-We formalize integrity and confidentiality as trace (hyper-)properties, in two different
+We formalize these integrity and confidentiality requirements as trace
+(hyper-)properties, with two different
 variants: {\em stepwise} variants, in which a caller's data must {\em never} be
@@
 long as these ``risky'' behaviors do not affect the system's observable
-behavior. The observational properties are more extensional, and
-any reasonable protection mechanism ought to enforce them,
-even if it does not prevent every single dangerous read or write.
+behavior. The observational properties are more extensional and thus
+allow ``lazier'' protection mechanism that permit ``dangerous'' reads or
+writes as long as they do not affect the program's visible outcomes.
 
-Confidentiality is especially interesting, as it is based on the traditional
-notion of noninterference. But where noninterference is normally presented as an
-end-to-end hyper-property, stack confidentiality is noninterference applied over
-multiple nested subtraces -- each individual call's behavior must be invariant
-regardless of the state of the stack at its entry.
+Confidentiality is especially interesting.  It is based on a traditional
+notion of noninterference, but where ordinary noninterference is 
+an end-to-end hyper-property on whole program runs, stack confidentiality is
+a ``nested'' form of 
+noninterference applied to
+subtraces of the whole program trace, requiring that, between each call
+instruction and the corresponding return, the
+caller's stack frame is protected from the callee.
 

### 2021-05-14 short/main.tex
some edits

@@
 The call stack is a perennial target for low-level attacks, leading to
-consequences  from leakage or corruption of private stack data to
+consequences that range from leakage or corruption of private stack data to
 control-flow hijacking. A profusion of
@@
 of concrete examples of attacks that they can prevent.
-At best, they define stack safety by reference to an idealized machine that
+At best, stack safety is defined by reference to an idealized machine that
 is arguably stack safe by construction \citep{SkorstengaardSTK}.
@@
 We propose such a characterization
-using the tools of language-based security: The informal claim that
+using the tools of language-based security. The informal claim that
 ``stack safety protects a caller from its callee'' amounts to saying
@@
 until it regains control.
-
+%
 We formalize these integrity and confidentiality requirements as trace
@@
 behavior. The observational properties are more extensional and thus
-allow ``lazier'' protection mechanism that permit ``dangerous'' reads or
+allow ``lazier'' protection mechanism that permit risky reads or
 writes as long as they do not affect the program's visible outcomes.
@@
 Confidentiality is especially interesting.  It is based on a traditional
-notion of noninterference, but where ordinary noninterference is 
+notion of noninterference, but whereas ordinary noninterference is 
 an end-to-end hyper-property on whole program runs, stack confidentiality is
@@
 noninterference applied to
-subtraces of the whole program trace, requiring that, between each call
-instruction and the corresponding return, the
-caller's stack frame is protected from the callee.
+subtraces of the whole program trace delimited by call instructions and
+corresponding returns, requiring that the callee's behavior is invariant
+under hypothetical changes to the caller's stack frame.
+%behavior of the callee is
+%independent of the caller's stack frame.
+%, between each call
+%instruction and the corresponding return, the
+%caller's stack frame is protected from the callee.
 
@@
 
-Finally, we demonstrate our model's flexibility by extending it to different settings.
-First, we allow variables to be passed on the stack, allowing some of the state of
+We have demonstrated our model's flexibility by extending it to different settings.
+First, we can support passing arguments on the stack, allowing some of the state of
 the caller to be shared with the callee (and potentially with a further
-callee, and so on.) Then we expand the number of stacks as we move to a simple
-coroutine model, albiet one with statically bounded regions for each stack.
+callee, and so on.)  We can also describe a simple coroutine model, with one
+statically bounded stack per coroutine, each protected from the others.
 
-In addition to testing lazy tag-based policies against our observational properties,
-we are adapting the uninitialized capability model of \citet{Georges+21} to work with
+In ongoing work, we are investigating whether stack protection schemes
+based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
+can enforce our properites.  In particular, we are adapting 
+the uninitialized capability model of \citet{Georges+21} to work with
 our testing framework for validation against our properties. Of the several Cheri-based
@@
 assume extra cooperation from callers, namely that they do not leak their capabilities
-to the attacker. There is no equivalent dynamic requirement in a tag-based system because
-the policy can dynamically enforce cooperation.
+to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
+%the policy can dynamically enforce cooperation.
 

### 2021-05-14 short/main.tex
typo

@@
 based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
-can enforce our properites.  In particular, we are adapting 
+can enforce our properties.  In particular, we are adapting 
 the uninitialized capability model of \citet{Georges+21} to work with

### 2021-05-14 short/main.tex
Polish the rest

@@
 behavior. The observational properties are more extensional and thus
-allow ``lazier'' protection mechanism that permit risky reads or
-writes as long as they do not affect the program's visible outcomes.
+allow ``lazier'' protection mechanisms that permit risky reads or
+writes as long as they do not affect visible outcomes.
 
@@
 corresponding returns, requiring that the callee's behavior is invariant
-under hypothetical changes to the caller's stack frame.
+under hypothetical changes to\bcp{scrambling of?} the caller's stack frame.
 %behavior of the callee is
@@
 This formulation not only captures the intuition
-that the callee cannot directly access the caller's state, but gives a novel
-way of looking at control-flow attacks. We are not interested in attacks that
-merely aim to execute arbitrary code -- in fact, our attacker model assumes
-that the (attacking) callee may already execute arbitrary code.
+that the callee cannot directly access the caller's state; it also gives a novel
+way of looking at control-flow attacks. We are not interested in\bcp{I find
+  ``We are not interested in'' confusing here---the previous sentence
+  promises a novel way of looking, and this opening feels irrelevant to
+  that.} attacks that
+merely aim to execute arbitrary code -- indeed, our model assumes
+that attackers (i.e., callees) may already execute arbitrary code.
 %
@@
 and the stack pointer restored. If the caller's data are accessed before then,
-it is a violation of integrity or confidentiality.
+it is a violation of integrity or confidentiality.  \bcp{Actually, the whole
+paragraph doesn't make very much sense to me.}
 
 These properties can optionally be extended with a notion of {\em well-bracketed control flow}
-as in \citet{SkorstengaardSTK}, the system-wide control-flow property that callees always return
-to their immediate caller, if they return at all. This property is largely orthogonal to our
+as in \citet{SkorstengaardSTK}---a global requirement that callees
+must always return
+to their immediate caller, if they return at all. This property is
+largely\bcp{``largely'' feels like a weasel word...} orthogonal to our
 data-protection properties.
@@
 {\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a RISC-V specification. We
+in the Coq proof assistant on top of a RISC-V ISA model\bcp{instead of
+  ``spec''.  OK?}. We
 use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
@@
 would violate stack safety.
-Furthermore, we check incorrect
-enforcement variants (both variants that we accidentally created
+Our testing regime is also able to detect incorrect variants of the
+enforcement mechanisms---both variants that we accidentally created
 during our re-implementation of the micro-policy and ones that we
 intentionally crafted to be broken in order to increase our confidence
-in testing and the enforcement mechanism itself). The testing framework
-is able to generate counterexamples that violate our properties but are not halted by
-incorrect enforcement mechanisms, increasing our confidence when it finds
-no counterexamples under the real enforcement mechanisms.
+in testing and the enforcement mechanism itself. For these broken
+mechanisms, our tester
+generates counterexamples that violate our properties but escape the
+enforcement.  This increases our confidence in enforcement
+mechanisms for which the tester is unable to find  counterexamples.
 
@@
 and also violates integrity if the leak happens to use the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that should enforce observational
-integrity and confidentiality, albeit at some performance cost. The next step
-is to expand testing to these properties and validate the variant; efficiently
-testing observational properties presents unique challenges because a property
-violation can be arbitrarily far removed from the initial ``risky'' read or write.
+propose a variant of {\em Lazy Tagging and Clearing} that should\bcp{``that
+  we believe will''?} enforce observational
+integrity and confidentiality, albeit at some performance cost. (An obvious next step
+is to expand testing to these properties and validate the variant, but
+efficiently
+testing observational properties is more challenging because a property
+violation can be arbitrarily far removed from the initial ``risky'' read or
+write.)
 
-We have demonstrated our model's flexibility by extending it to different settings.
+We have demonstrated\bcp{let's use the same verb tense consistently---''we
+  do X'' [my preference] or ``we have done X'' or ``we did X''---all the way
+through} our model's flexibility by extending it to different settings.
 First, we can support passing arguments on the stack, allowing some of the state of
-the caller to be shared with the callee (and potentially with a further
-callee, and so on.)  We can also describe a simple coroutine model, with one
+the caller to be shared with the callee (and potentially with a sub-callee,
+and so on.)  We can also describe a simple coroutine model, with one
 statically bounded stack per coroutine, each protected from the others.
@@
 can enforce our properties.  In particular, we are adapting
-the uninitialized capability model of \citet{Georges+21} to work with
-our testing framework for validation against our properties. Of the several Cheri-based
-stack protection schemes, its treatment of unitialized frames seems most compatible
+the ``uninitialized capability model'' of \citet{Georges+21} to work with
+our testing framework in order to validate it against our properties. Out of several Cheri-based
+stack protection schemes, this one's treatment of unitialized frames seems most compatible
 with our model of confidentiality, without the need for expensive stack clearing
-between calls. We expect that Cheri-based techniques support our properties if we
+between calls. We expect that Cheri-based techniques can support our properties if we
 assume extra cooperation from callers, namely that they do not leak their capabilities

### 2021-05-14 short/main.tex
Hopefully small changes, I don't remember

--- 2021-05-14 short/main.tex
@@
 The call stack is a perennial target for low-level attacks, leading to
-consequences that range from leakage or corruption of private stack data to
+consequences ranging from leakage or corruption of private stack data to
 control-flow hijacking. A profusion of
@@
 of concrete examples of attacks that they can prevent.
-At best, stack safety is defined by reference to an idealized machine that
+At best, they define stack safety by reference to an idealized machine that
 is arguably stack safe by construction \citep{SkorstengaardSTK}.
@@
 
-We propose such a characterization
-using the tools of language-based security. The informal claim that
-``stack safety protects a caller from its callee'' amounts to saying
-that it guarantees
+We propose such a characterization,
+using the technical framework of language-based security.
+Stack safety protects a caller from its callee, guaranteeing
 the {\em integrity} and {\em confidentiality} of the caller's local state
 until it regains control.
-%
-We formalize these integrity and confidentiality requirements as trace
-(hyper-)properties, with two different
+
+We formalize integrity and confidentiality as trace properties, in two different
 variants: {\em stepwise} variants, in which a caller's data must {\em never} be
@@
 long as these ``risky'' behaviors do not affect the system's observable
-behavior. The observational properties are more extensional and thus
-allow ``lazier'' protection mechanism that permit risky reads or
-writes as long as they do not affect the program's visible outcomes.
+behavior. The observational properties are more extensional, and
+any reasonable protection mechanism ought to enforce them,
+even if it does not prevent every single dangerous read or write.
 
-Confidentiality is especially interesting.  It is based on a traditional
-notion of noninterference, but whereas ordinary noninterference is 
-an end-to-end hyper-property on whole program runs, stack confidentiality is
-a ``nested'' form of 
-noninterference applied to
-subtraces of the whole program trace delimited by call instructions and
-corresponding returns, requiring that the callee's behavior is invariant
-under hypothetical changes to the caller's stack frame.
-%behavior of the callee is
-%independent of the caller's stack frame.
-%, between each call
-%instruction and the corresponding return, the
-%caller's stack frame is protected from the callee.
+Confidentiality is especially interesting, as it is based on the traditional
+notion of noninterference. But where noninterference is normally presented as an
+end-to-end hyper-property, stack confidentiality is noninterference applied over
+multiple nested subtraces -- each individual call's behavior must be invariant
+regardless of the state of the stack at its entry.
 
@@
 
-We have demonstrated our model's flexibility by extending it to different settings.
-First, we can support passing arguments on the stack, allowing some of the state of
+Finally, we demonstrate our model's flexibility by extending it to different settings.
+First, we allow variables to be passed on the stack, allowing some of the state of
 the caller to be shared with the callee (and potentially with a further
-callee, and so on.)  We can also describe a simple coroutine model, with one
-statically bounded stack per coroutine, each protected from the others.
+callee, and so on.) Then we expand the number of stacks as we move to a simple
+coroutine model, albiet one with statically bounded regions for each stack.
 
-In ongoing work, we are investigating whether stack protection schemes
-based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
-can enforce our properties.  In particular, we are adapting 
-the uninitialized capability model of \citet{Georges+21} to work with
+In addition to testing lazy tag-based policies against our observational properties,
+we are adapting the uninitialized capability model of \citet{Georges+21} to work with
 our testing framework for validation against our properties. Of the several Cheri-based
@@
 assume extra cooperation from callers, namely that they do not leak their capabilities
-to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
-%the policy can dynamically enforce cooperation.
+to the attacker. There is no equivalent dynamic requirement in a tag-based system because
+the policy can dynamically enforce cooperation.
 

### 2021-05-14 short/main.tex

--- 2021-05-14 short/main.tex
@@
 corresponding returns, requiring that the callee's behavior is invariant
-under hypothetical changes to\bcp{scrambling of?} the caller's stack frame.
+under hypothetical scrambling of the caller's stack frame.
 %behavior of the callee is
@@
 that the callee cannot directly access the caller's state; it also gives a novel
-way of looking at control-flow attacks. We are not interested in\bcp{I find
-  ``We are not interested in'' confusing here---the previous sentence
-  promises a novel way of looking, and this opening feels irrelevant to
-  that.} attacks that
-merely aim to execute arbitrary code -- indeed, our model assumes
-that attackers (i.e., callees) may already execute arbitrary code.
-%
-By extension, we do not place any particular restrictions on control flow.
-Perhaps suprisingly, we can still meaningfully distinguish the caller from
-its callee, and talk about protecting the caller's data. After a call, control
-is considered to belong to the callee until the trace reaches a valid
-``return target,'' typically a state with the program counter at the next instruction
-and the stack pointer restored. If the caller's data are accessed before then,
-it is a violation of integrity or confidentiality.  \bcp{Actually, the whole
-paragraph doesn't make very much sense to me.}
+interpretation of integrity and confidentiality in the presence of control-flow
+attacks. The key is that every call has a corresponding predicate on states
+designated a ``return target''
+that defines what it means for that call to truly return. The program counter
+must be at the next instruction, and the stack pointer restored to its original
+value. A callee may execute the instructions of a return,
+but if it does not reach a return target, we simply do not recognize it as a return.
+So the callee can be ``active'', and the caller protected, even if the callee is
+executing instructions that look like they belong to the caller.
+
+Our attacker model is strong, allowing the attacker to execute arbitrary code, so
+the callee has no reason to execute the caller's instructions -- any instruction it could
+jump to, it could already execute. The only reason to perform a ``fake'' return in this setting
+would be to access the caller's data, which our properties will correctly identify as a violation.
 
@@
 must always return
-to their immediate caller, if they return at all. This property is
-largely\bcp{``largely'' feels like a weasel word...} orthogonal to our
-data-protection properties.
+to their immediate caller, if they return at all. This property is orthogonal to our
+data-protection properties. It remains meaningful in our strong attacker model, because
+it supports reasoning about the ordering of valid returns.
 
@@
 {\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a RISC-V ISA model\bcp{instead of
-  ``spec''.  OK?}. We
+in the Coq proof assistant on top of a RISC-V ISA model. We
 use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
@@
 and also violates integrity if the leak happens to use the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that should\bcp{``that
-  we believe will''?} enforce observational
+propose a variant of {\em Lazy Tagging and Clearing} that
+  we believe will enforce observational
 integrity and confidentiality, albeit at some performance cost. (An obvious next step
@@
 
-We have demonstrated\bcp{let's use the same verb tense consistently---''we
-  do X'' [my preference] or ``we have done X'' or ``we did X''---all the way
-through} our model's flexibility by extending it to different settings.
+We demonstrate our model's flexibility by extending it to different settings.
 First, we can support passing arguments on the stack, allowing some of the state of

### 2021-05-14 short/main.tex
another stab at problematic paras

@@
 
-This formulation not only captures the intuition
-that the callee cannot directly access the caller's state; it also gives a novel
-interpretation of integrity and confidentiality in the presence of control-flow
-attacks. The key is that every call has a corresponding predicate on states
+\apt{Here's another stab:}
+Our attacker model is very strong, allowing the attacker to execute arbitrary code.
+In particular, we do not attempt to prevent \apt{most forms of?} control-flow attacks
+\apt{such as ROP?}. The notions of integrity and confidentiality remain useful even
+in the absence of normal control flow.
+%This formulation not only captures the intuition
+%that the callee cannot directly access the caller's state; it also gives a novel
+%interpretation of integrity and confidentiality in the presence of control-flow
+%attacks.
+The key is that every call has a corresponding predicate on states
 designated a ``return target''
-that defines what it means for that call to truly return. The program counter
+that defines what it means for that call to truly return: the program counter
 must be at the next instruction, and the stack pointer restored to its original
@@
 
-Our attacker model is strong, allowing the attacker to execute arbitrary code, so
-the callee has no reason to execute the caller's instructions -- any instruction it could
-jump to, it could already execute. The only reason to perform a ``fake'' return in this setting
-would be to access the caller's data, which our properties will correctly identify as a violation.
+%Our attacker model is strong, allowing the attacker to execute arbitrary code, so
+%the callee has no reason to execute the caller's instructions -- any instruction it could
+%jump to, it could already execute. The only reason to perform a ``fake'' return in this setting
+%would be to access the caller's data, which our properties will correctly identify as a violation.
 

### 2021-05-14 short/main.tex
Tweaked paragraph four

@@
 \apt{Here's another stab:}
+\sna{Tweaked slightly. Trying to make it clear that control flow attacks don't change what we're protecting.}
 Our attacker model is very strong, allowing the attacker to execute arbitrary code.
-In particular, we do not attempt to prevent \apt{most forms of?} control-flow attacks
-\apt{such as ROP?}. The notions of integrity and confidentiality remain useful even
-in the absence of normal control flow.
+In particular, we do not attempt to prevent control-flow attacks
+such as ROP. Even in the absence of normal control flow, confidentiality and integrity
+still apply to all of the data they would under more idealized conditions.
 %This formulation not only captures the intuition

### 2021-05-14 short/main.tex
more tweaking on para 4

@@
 \sna{Tweaked slightly. Trying to make it clear that control flow attacks don't change what we're protecting.}
+\apt{Yet more tweaking.}
 Our attacker model is very strong, allowing the attacker to execute arbitrary code.
@@
 such as ROP. Even in the absence of normal control flow, confidentiality and integrity
-still apply to all of the data they would under more idealized conditions.
+still apply to data in suspended stack frames.
+% all of the data they would under more idealized conditions.
 %This formulation not only captures the intuition
@@
 %attacks.
-The key is that every call has a corresponding predicate on states
-designated a ``return target''
-that defines what it means for that call to truly return: the program counter
+The key is that every call has a corresponding predicate on states,
+designated as the ``return target,''
+which defines what it means for that call to truly return: the program counter
 must be at the next instruction, and the stack pointer restored to its original
-value. A callee may execute the instructions of a return,
-but if it does not reach a return target, we simply do not recognize it as a return.
+value. A callee may execute a return instruction,
+but if it does not reach a return target we simply do not recognize it as a return.
 So the callee can be ``active'', and the caller protected, even if the callee is
-executing instructions that look like they belong to the caller.
+executing instructions outside of its normal footprint.
+% that don't look like they belong to the caller.
 

### 2021-05-14 preprint/main.tex

--- 2021-04-17 tex/main.tex
@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% Author with single affiliation.
-%\author{First1 Last1}
+\author{Sean Noble Anderson}
 %\authornote{with author1 note}          %% \authornote is optional;
@@
 %\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-%\affiliation{
+\affiliation{
 %  \position{Position1}
-%  \department{Department1}              %% \department is recommended
-%  \institution{Institution1}            %% \institution is required
+%  \department{Computer Science}              %% \department is recommended
+  \institution{Portland State University}            %% \institution is required
 %  \streetaddress{Street1 Address1}
@@
 %  \country{Country1}                    %% \country is recommended
-%}
-%\email{first1.last1@inst1.edu}          %% \email is recommended
+}
+\email{ander28@pdx.edu}          %% \email is recommended
+
+\author{Leonidas Lampropoulos}
+\affiliation{
+  \institution{University of Maryland, College Park}
+}
+\email{leonidas@umd.edu}
+
+\author{Roberto Blanco}
+\affiliation{
+  \institution{Max Planck Institute for Security and Privacy}
+}
+\email{roberto.blanco@mpi-sp.org}
+
+\author{Benjamin C. Pierce}
+\affiliation{
+  \institution{University of Pennsylvania}
+}
+\email{bcpierce@cis.upenn.edu}
+
+\author{Andrew Tolmach}
+\affiliation{
+  \institution{Portland State University}
+}
+\email{tolmach@pdx.edu}
+
 %
@@
 \section{Introduction}
-
-\newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}
-\newcommand{\memoryaddrs}[2][]
-  {
-    \MemoryLabel{#2}{2em}{$\downarrow$\SP #1}
-  }
 
@@
 capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\bcp{Yes, going back to MIT days---we should
-  include several more of these, if only to give readers the impression that
-this is a well-studied mechanism (so formalizing its protections is
-useful).}
-\fi
+and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}.
 %
@@
 \emph{explicit} definition using the technical framework of language-based
-security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
-  survey?}\fi, phrased as the conjunction of
+security, phrased as the conjunction of
 \emph{stack integrity} and \emph{stack confidentiality}.
@@
 %confidence that it can also catch unknown ones, strengthening our confidence
-%in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
-%  sentence belongs in the testing section?}\fi)
+%in the enforcement mechanism itself.)
 
@@
 %
-\ifaftersubmission
-\bcp{Did Nick and Andre know about this deficiency?  Did they mention
-  it in their paper?  (If neither, we should make a bigger deal out of the
-  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
-    They refer to ``temporal security properties'' as something to do with dangling
-    pointers, but they don't define it.
-    Then they say that Depth Isolation has weaker temporal security
-    guarantees than Static Authorities (tag=function name), but they don't explain why
-    (and I don't see why). They don't say anything explicit about the temporal
-    security properties of Lazy Tagging and Clearing, but they consider it to enforce
-    a completely different (and generally weaker) policy anyway (``data flow
-    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
-  leave this issue to deal with later (in consultation with them, which we
-  don't have time to do a good job of today).  At any rate, by my reading we
-are not saying anything incorrect at the moment.}
-\fi
-
 % A second challenge is that enforcement mechanisms may weaken their
@@
 properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy
-\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
+machine-code level for a variety of ISAs, including RISCy ones.  In
 particular, there may be no single ``call'' instruction that does all the
@@
 \end{subfigure}
-\caption{An integrity violation
-\ifaftersubmission\bcp{I wonder if we
-    could make the figures easier to read by putting very light, differently
-  colored backgrounds under the different parts}\fi}
+\caption{An integrity violation}
 \label{fig:int1}
@@
 shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
-state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
-understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
+state with arbitrary values~\(v_0\), etc., in place of the values that should
 be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
@@
 
-\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
-    the figure (at least in the latex build that I'm looking at...)}\fi}
+\caption{Sharing on the stack}
 \label{fig:passingsimple}
@@
 in the form of
-noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
-be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
-(or ask arthur how!).}\fi. Just
+noninterference~\citep{DBLP:conf/post/AmorimHP18}. Just
 as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
@@
 %% Acknowledgments
-\begin{acks}                            %% acks environment is optional
+%\begin{acks}                            %% acks environment is optional
                                         %% contents suppressed with 'anonymous'
@@
   %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
+%  This material is based upon work supported by the
+%  \grantsponsor{GS100000001}{National Science
+%    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+%  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+%  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+%  conclusions or recommendations expressed in this material are those
+%  of the author and do not necessarily reflect the views of the
+%  National Science Foundation.
+%\end{acks}
 

### 2021-05-14 short/main.tex

@@
 
-\apt{Here's another stab:}
-\sna{Tweaked slightly. Trying to make it clear that control flow attacks don't change what we're protecting.}
-\apt{Yet more tweaking.}
-Our attacker model is very strong, allowing the attacker to execute arbitrary code.
-In particular, we do not attempt to prevent control-flow attacks
-such as ROP. Even in the absence of normal control flow, confidentiality and integrity
+Our threat model is very strong, allowing the attacker (i.e., the callee) to
+execute arbitrary code, including code that might attempt to smash
+the stack to disrupt the program's control flow.
+Our stack safety properties demand that, even in the presence of such
+attackers, confidentiality and integrity must
 still apply to data in suspended stack frames.
@@
 %attacks.
+\bcp{I tried rewriting the first part of the paragraph, and I like it
+  better, but now there's a big gap between that and what comes next.  I
+  wonder whether we need to turn the next bit into a separate discussion of
+  ``what it means to return''...  In any case, I felt like the previous
+  version of the whole paragraph didn't quite nail what we needed to say
+  about all this.}
 The key is that every call has a corresponding predicate on states,

### 2021-05-14 short/main.tex
Attempted better transition in para. 4

@@
   about all this.}
-The key is that every call has a corresponding predicate on states,
-designated as the ``return target,''
+\sna{Better transition.}
+The question then arises: when are suspended frames reactivated, or equivalently,
+what constitutes a return in this setting? We define returns according to the behavior
+we would expect in a more structured environment.
+Every call has a corresponding predicate on states,
+its ``return target,''
 which defines what it means for that call to truly return: the program counter

### 2021-05-15 short/main.tex
Another try

--- 2021-05-14 short/main.tex
@@
 attackers, confidentiality and integrity must
-still apply to data in suspended stack frames.
+still apply to data in suspended stack frames until the callee returns.
 % all of the data they would under more idealized conditions.
@@
 %attacks.
-\bcp{I tried rewriting the first part of the paragraph, and I like it
-  better, but now there's a big gap between that and what comes next.  I
-  wonder whether we need to turn the next bit into a separate discussion of
-  ``what it means to return''...  In any case, I felt like the previous
-  version of the whole paragraph didn't quite nail what we needed to say
-  about all this.}
-The key is that every call has a corresponding predicate on states,
-designated as the ``return target,''
+% \bcp{I tried rewriting the first part of the paragraph, and I like it
+%   better, but now there's a big gap between that and what comes next.  I
+%   wonder whether we need to turn the next bit into a separate discussion of
+%   ``what it means to return''...  In any case, I felt like the previous
+%   version of the whole paragraph didn't quite nail what we needed to say
+%   about all this.}
+% \sna{Better transition.}
+The question then arises: when, exactly, are suspended frames
+reactivated---or equivalently,
+what constitutes a return in this setting? It is not obvious, since our
+setting is a concrete, low-level ISA where the same opcode might sometimes
+be used to return from a call and sometimes for other purposes.
+Our solution is to associate with each call a corresponding predicate on states,
+its ``return target,''
 which defines what it means for that call to truly return: the program counter
@@
 value. A callee may execute a return instruction,
-but if it does not reach a return target we simply do not recognize it as a return.
+but if it does not reach a return target we simply do not recognize it as a
+return.
 So the callee can be ``active'', and the caller protected, even if the callee is
-executing instructions outside of its normal footprint.
-% that don't look like they belong to the caller.
+executing instructions outside of its normal footprint.  \bcp{I tried
+  another rewrite of the end of the paragraph.  I think it's getting there,
+  but it's probably now a bit too long as well as too detailed.}
 

### 2021-05-15 short/main.tex
Cleaned up some spot, tried to make para. 4 better

@@
 still apply to data in suspended stack frames until the callee returns.
-% all of the data they would under more idealized conditions.
-%This formulation not only captures the intuition
-%that the callee cannot directly access the caller's state; it also gives a novel
-%interpretation of integrity and confidentiality in the presence of control-flow
-%attacks.
-% \bcp{I tried rewriting the first part of the paragraph, and I like it
-%   better, but now there's a big gap between that and what comes next.  I
-%   wonder whether we need to turn the next bit into a separate discussion of
-%   ``what it means to return''...  In any case, I felt like the previous
-%   version of the whole paragraph didn't quite nail what we needed to say
-%   about all this.}
-% \sna{Better transition.}
-The question then arises: when, exactly, are suspended frames
-reactivated---or equivalently,
+The question then arises:
 what constitutes a return in this setting? It is not obvious, since our
@@
 its ``return target,''
-which defines what it means for that call to truly return: the program counter
-must be at the next instruction, and the stack pointer restored to its original
-value. A callee may execute a return instruction,
-but if it does not reach a return target we simply do not recognize it as a
-return.
-So the callee can be ``active'', and the caller protected, even if the callee is
-executing instructions outside of its normal footprint.  \bcp{I tried
-  another rewrite of the end of the paragraph.  I think it's getting there,
-  but it's probably now a bit too long as well as too detailed.}
-
-%Our attacker model is strong, allowing the attacker to execute arbitrary code, so
-%the callee has no reason to execute the caller's instructions -- any instruction it could
-%jump to, it could already execute. The only reason to perform a ``fake'' return in this setting
-%would be to access the caller's data, which our properties will correctly identify as a violation.
+which holds on the states that would follow a well-behaved return.
+Whether or not a callee executes a return instruction,
+we only recognize it returning only when it reaches a return target.
+So the caller is protected until then, even if the callee is
+executing instructions outside of its normal footprint.
 
@@
 to their immediate caller, if they return at all. This property is orthogonal to our
-data-protection properties. It remains meaningful in our strong attacker model, because
-it supports reasoning about the ordering of valid returns.
+data-protection properties. It remains meaningful in our strong attacker model,
+to support reasoning about the ordering of valid returns.
 
@@
 would violate stack safety.
-Our testing regime is also able to detect incorrect variants of the
-enforcement mechanisms---both variants that we accidentally created
-during our re-implementation of the micro-policy and ones that we
-intentionally crafted to be broken in order to increase our confidence
+Our testing regime also detects incorrect variants of the
+enforcement mechanisms---both errors
+during our re-implementation of the micro-policy and
+intentionally crafted test, meant to be broken in order to increase our confidence
 in testing and the enforcement mechanism itself. For these broken
@@
 generates counterexamples that violate our properties but escape the
-enforcement.  This increases our confidence in enforcement
-mechanisms for which the tester is unable to find  counterexamples.
+enforcement.
 
@@
 
-In ongoing work, we are investigating whether stack protection schemes
+In ongoing work, we investigate whether stack protection schemes
 based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
-can enforce our properties.  In particular, we are adapting
-the ``uninitialized capability model'' of \citet{Georges+21} to work with
-our testing framework in order to validate it against our properties. Out of several Cheri-based
-stack protection schemes, this one's treatment of unitialized frames seems most compatible
+can enforce our properties.  In particular, we aim to validate
+the ``uninitialized capability model'' of \citet{Georges+21} using
+our testing framework. Out of several Cheri-based
+stack protection schemes, their treatment of unitialized frames seems most compatible
 with our model of confidentiality, without the need for expensive stack clearing

### 2021-05-15 short/main.tex
para 4

@@
 its ``return target,''
-which holds on the states that would follow a well-behaved return.
+which holds on the states that would follow a well-behaved return.\apt{I think this
+  was better before when the definition of return target was more concrete.}
 Whether or not a callee executes a return instruction,
-we only recognize it returning only when it reaches a return target.
+we consider it to be returning only when it reaches a return target.\apt{Reworded
+last sentence slightly; seems repetitive in anyc ase.}
 So the caller is protected until then, even if the callee is

### 2021-05-16 short/main.tex
Paragraph 4 tweaked

--- 2021-05-15 short/main.tex
@@
 its ``return target,''
-which holds on the states that would follow a well-behaved return.
-Whether or not a callee executes a return instruction,
-we only recognize it returning only when it reaches a return target.
-So the caller is protected until then, even if the callee is
+which defines what it means for that call to truly return: the program counter
+must be at the next instruction, and the stack pointer restored to its original
+value.
+So the caller is protected until its return target is reached, even if the callee is
 executing instructions outside of its normal footprint.

### 2021-05-16 short/main.tex
More work on four

@@
 Our stack safety properties demand that, even in the presence of such
-attackers, confidentiality and integrity must
+attackers, confidentiality and integrity
 still apply to data in suspended stack frames until the callee returns.
@@
 be used to return from a call and sometimes for other purposes.
-Our solution is to associate with each call a corresponding predicate on states,
+Our solution is to associate with each call a predicate on states,
 its ``return target,''
-which defines what it means for that call to truly return: the program counter
+which defines a correct return: the program counter
 must be at the next instruction, and the stack pointer restored to its original
-value.
-So the caller is protected until its return target is reached, even if the callee is
-executing instructions outside of its normal footprint.
+value. So the callee may
+execute instructions outside of its normal footprint, perhaps due to
+smashing the stack, but
+the caller is protected until the first state that satisfies its return target.
 

### 2021-05-17 short/main.tex
Nits

@@
 during our re-implementation of the micro-policy and
-intentionally crafted test, meant to be broken in order to increase our confidence
+intentionally crafted tests, meant to be broken in order to increase our confidence
 in testing and the enforcement mechanism itself. For these broken
@@
 the caller to be shared with the callee (and potentially with a sub-callee,
-and so on.)  We can also describe a simple coroutine model, with one
+and so on).  We can also describe a simple coroutine model, with one
 statically bounded stack per coroutine, each protected from the others.

### 2021-05-17 short/main.tex
Polish

@@
 Our threat model is very strong, allowing the attacker (i.e., the callee) to
-execute arbitrary code, including code that might attempt to smash
+execute arbitrary code, including attempting to smash
 the stack to disrupt the program's control flow.
@@
 The question then arises:
-what constitutes a return in this setting? It is not obvious, since our
-setting is a concrete, low-level ISA where the same opcode might sometimes
+what constitutes a return in this setting? It is not obvious, since
+our machine model is a low-level ISA where the same opcode might
+sometimes
 be used to return from a call and sometimes for other purposes.
-Our solution is to associate with each call a predicate on states,
-its ``return target,''
-which defines a correct return: the program counter
-must be at the next instruction, and the stack pointer restored to its original
-value. So the callee may
+Our solution is to associate, with each dynamic call in an execution trace, a
+``return target'' that records the address of the instruction following the
+call and the stack pointer when the call was made. Thus, so the callee may
 execute instructions outside of its normal footprint, perhaps due to
 smashing the stack, but
-the caller is protected until the first state that satisfies its return target.
+the caller is protected until the PC and SP simultaneously match its return
+target.
 
-These properties can optionally be extended with a notion of {\em well-bracketed control flow}
+We can also, optionally, include a notion of {\em well-bracketed control flow}
 as in \citet{SkorstengaardSTK}---a global requirement that callees
@@
 
-We demonstrate our model's flexibility by extending it to different settings.
+We demonstrate our model's flexibility by extending it in two significant ways.
 First, we can support passing arguments on the stack, allowing some of the state of
@@
 
-In ongoing work, we investigate whether stack protection schemes
+In ongoing work, we are investigating whether stack protection schemes
 based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}

### 2021-05-17 short/main.tex
wibbles

@@
 ``return target'' that records the address of the instruction following the
-call and the stack pointer when the call was made. Thus, so the callee may
+call and the stack pointer when the call was made. Thus, the callee may
 execute instructions outside of its normal footprint, perhaps due to
@@
 the function activation that owns the frame and access to those locations is
-then permitted only when that activation is currently executing.
+then permitted only when an activation at that depth is currently executing.
 On the other hand, we find that their \emph{Lazy Tagging and Clearing} policy

### 2021-06-14 submission2/response.txt

> OOPSLA 2021 Paper #16 Reviews and Comments
> ===========================================================================
> Paper #16 Stack Safety is a Security Property
> 
> 
> Review #16A
> ===========================================================================
> 
> Overall merit
> -------------
> 3. Weak accept
> 
> Review confidence
> -----------------
> 3. High
> 
> Paper summary
> -------------
> The paper gives an elegant, careful, formal treatment on what it means to ensure stack safety -- that callers and callees do not read or write each other's stacks, at least not in a way that has an observable impact, and that functions never return to the wrong place.   This is formalized in a suitably abstract fashion with a range of more eager to more lazy definitions of correctness such that particular enforcement mechanisms can be analyzed for correctness.  Property testing validates a particular eager algorithm from the literature and reveals problems, then corrected, with a lazier algorithm.  Extensions toward allowing unobservable stack corruptions and supporting a simple form of coroutines are then developed.
> 
> Strengths and Weaknesses
> ------------------------
> I really enjoyed this paper and would highly recommend it to anyone working on low-level security mechanisms.  The exposition is clear, precise, well-motivated, and in many ways synthesizing many years of insights into what it really means to protect the call stack.  The examples are well-thought out, the definitions unimpeachable, and the clarity of thought impressive.  Throughout reading the paper, I anticipated giving a 'strong accept' rating.
> 
> What gives me pause is the relative lack of validation or even discussion of why there isn't more validation.  I don't know if the authors simply haven't had time yet, think it is unnecessary, or ran into challenges or complexities that remain unexplained.  I do consider significant the property-based testing that was done both to validate the simpler tagging approach and identify (obvious in hindsight, but that's not a criticism) problems with the lazier tagging.  But that still leaves:
> 
> * Any discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.  After all, the authors already have a fully formal definition and an implementation in Coq.  While this sort of work does not necessarily need a correctness proof in Coq -- I still lean slightly toward accept after all -- it seems quite strange not to discuss even the possibility of validation including the ability to pursue a correctness argument for an enforcement mechanism beyond property testing.
> 
> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
> 
> * Any validation for the lazier observability definition (this is listed as future work.)
> 
> * Any validation for the coroutine work.
> 
> So: I believe this paper represents an important, sound conceptual advance that can inform the research community, but I was disappointed by the gaps in evaluation -- these gaps are acknowledged but not discussed (do they matter? how might they be closed? ...).
> 
> Comments for author
> -------------------
> There is one topic beyond my evaluation above worth considering: callee-save registers.  I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save registers hold data private to the caller that may be passed arbitrarily deeper into the call stack, then saved into arbitrary stack locations, but then must be restored before return.  Have you or the closely related prior work on enforcement mechanisms considered this complexity?  The invariants are clear enough informally, but the issue seems to add substantial difficulty to tracking that a function faithfully implements the calling convention.
> 
> As a minor question you don't need to answer, just take as low-level feedback: Do you /need/ to separate contexts (section 5.2) from policies, or is this just a matter of convenience since in practice various policies benefit from similar context information?
> 
> Another minor question: The difference between [nested] call-by-reference and passing address-of-local was interesting, but I'm not sure how much it makes sense to argue the latter is "beyond the scope of stack safety per se" [line 939] without arguing the same of the former.  Perhaps the argument can be made that address-of-local can be put in the heap while a by-reference parameter cannot and since you don't guarantee heap safety, you shouldn't try to do anything useful with address-of-local?
> 
> typos:
> * line 189
> * line 1020
> 
> Questions for authors’ response
> ---------------------------------
> See prior sections:
> 
> * Can you explain the gaps in validation/evaluation and how a revised paper can address them?
> 
> * Any thoughts on callee-save registers?
> 
> 
> 
> Review #16B
> ===========================================================================
> 
> Overall merit
> -------------
> 2. Weak reject
> 
> Review confidence
> -----------------
> 2. Medium
> 
> Paper summary
> -------------
> Low-level attacks exploit the unprotected access to the stack in languages like C. By writing outside the current stack frame, attackers may manipulate sensitive data and alter the control flow of the program. The paper formalizes the notion of _stack safety_, which protects against this type of attacks. Stack safety consists of two parts: Stack confidentiality and integrity, which say that protected regions on the stack cannot be read and altered. The definitions of stack safety target a language semantics in form of an abstract machine. The paper evaluates the approach by implementing an existing stack-safety enforcement mechanism and uncovering potential stack-safety violations.
> 
> Strengths and Weaknesses
> ------------------------
> Strengths:
> - The problem that this paper solves is relevant. Countless attacks that exploit unprotected stack access and protection methods against these attacks emphasize the relevance of this problem.
> - I was convinced by the evaluation of this paper. Their approach uncovered potential stack-safety violations in an existing enforcement mechanism by Reossler and DeHon.
> 
> Weaknesses:
> - I question the novelty of this work. In particular, it is unclear how this paper improves upon Skorstengaard et al (question 1).
> - It is unclear how this approach generalizes to other languages (question 2).
> - The presentation of the paper can be improved. In particular, section 2 and 3 are very abstract because of the lack of examples and the readability of definitions in section 5 and 6 can be improved (see detailed comments).
> 
> Because of the questionable novelty and a lacking presentation, I think this paper is not ready for publication yet.
> 
> Comments for author
> -------------------
> Detailed Comments
> =====================
> 
> The following comments are organised in the order of reading the paper front to back. On first read, I had some misunderstandings that got cleared up later, however, they may indicate places where the paper can be improved.
> 
> ### Section 1
> - L32-40: It is unclear how this paper improves over Skorstengaard et al. Why isn't their definition of stack safety not already good enough?
> 
> ### Section 2
> - L124: You did not introduce the acronyms "ISA" and "RISC"
> - I found it quite hard to understand section 2 because the text is very abstract. I would have appreciated a few examples, like those you provide in section 4.
> - L148: Why do you need these policies?
> - L154: In this paragraph you motivate why you need policies. I would have appreciated this motivation a paragraph earlier.
> 
> ### Section 3
> - The text of section 3 is equally hard to understand as in section 2 because of the lack of examples.
> 
> ### Section 4
> - L214: Remind the reader that the register `%sp` refers to the stack pointer.
> - I like the example in figure 1. The figure would be easier to understand if you would write "unsealed" and "sealed" instead of "U" and "S".
> - Figure 2: Why are there two stacks side by side?
> 
> ### Section 5
> - L389: What are "w" and "a"?
> - I find the terms "MPC-state" and "MP trace" confusing.
> 
> ### Section 6
> - L499-509: There are so many different variable names that make it difficult to read the rules ("m", "cm", "dm", "rts", "tar", "ctx"). These variable names are explained somewhere in the text, but their explanation is scatter across multiple pages. Consider declaring these variables in a table in a central place.
> - Definition 6.2: Explain what the definition means instead of just stating it mathematically. What is pi_m?
> - Consider applying these rules and definitions to an example of section 4 and explain where stack integrity and confidentiality is violated.
> 
> ### Section 10
> - The related work section seems a bit superficial since it only discusses 4 related works. Consider adding a discussion of the software and hardware protection techniques mentioned in the introduction.
> 
> Questions for authors’ response
> ---------------------------------
> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.
> 
> Q2. In L117-121 you state certain assumptions about the language semantics. It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply. For example, does a language with a heap allow you to break stack safety? What about a language with delimited continuations that manipulate the stack?
> 
> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?
> 
> 
> 
> Review #16C
> ===========================================================================
> 
> Overall merit
> -------------
> 1. Reject
> 
> Review confidence
> -----------------
> 2. Medium
> 
> Paper summary
> -------------
> The paper formalizes a definition of stack safety and presents extensions of it. The definition considers three properties: integrity, confidentiality, and well-bracketed control flow.
> The authors also use these properties to validate the stack-safety “micro-policies” proposed by earlier work.
> 
> Strengths and Weaknesses
> ------------------------
> - It is a bit difficult to assess the merits of this work. Yes, the stack safety property is important for security, but what new results or new capabilities does the proposed definition bring in? If it is positioned as a security solution,  does it lead to more efficient defenses of control flow attacks? Does it prevent any new attacks that existing defenses fail? The paper does not talk about data flow across heap and stack, which is an important element for many real-world memory corruption attacks.
> 
> - The formalization seems intuitively correct, however there is no formal proof for its overall soundness properties and there is no real-world evaluation of its applications.
> 
> - The paper is titled "stack safety is a security property". I don't see why this is surprising, or what's the novelty here. Memory safety is almost always connected with security properties.
> 
> Comments for author
> -------------------
> I might have missed something important. Are there any applications of the proposed formalization?

### 2021-06-15 submission2/response.txt
Filled in some of response.txt

@@
 While this sort of work does not necessarily need a correctness proof in Coq -- I still lean slightly toward accept after all -- it seems quite strange not to discuss even the possibility of validation including the ability to pursue a correctness argument for an enforcement mechanism beyond property testing.
+
+These properties would absolutely be amenable to a correctness proof. Such a proof would be a
+lot of work in a setting with a full ISA, but there is not reason it should be impossible. The
+primary challenge we would predict in such an effort, beyond raw effort, is in determining for a given
+enforcement mechanism what invariants it needs to maintain in order to work correctly, and proving that
+it maintains them. This is very feasible for a proof of an individual, well-understood enforcement
+mechanism, but not for our goal of evaluating several such mechanisms.
+
 > 
@@
 > * Any validation for the coroutine work.
+
+Here we should discuss our roadmap for testing in more detail. At present we are working to implement
+testing of observable properties alongside the fixed version of the lazy tagging/clearing micro-policy.
+Observational integrity is implemented, but confidentiality is not -- since the lazy policy enforced
+normal, stepwise confidentiality this is sufficient to test it. But we did not have this ready before
+the deadline.
+
+Our other major goal is to implement testing of a secondary architecture. We will evaluate a Cheri
+implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
+enforcement.
+
+We consider validation of the passing system to be a secondary priority and coroutines tertiary.
+The passing system has existing enforcement mechanisms and serves an important role in common
+stack use-cases. Coroutines will require the development of an enforcement mechanism (likely
+tag-based) to validate.
+
 > 
@@
 The invariants are clear enough informally, but the issue seems to add substantial difficulty to tracking that a function faithfully implements the calling convention.
-> 
+
+Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
+They actually provide another use-case for observational properties! If we focus on integrity
+the reason becomes clear. Non-argument registers will naturally be overwritten (violating
+stepwise integrity) but should always be restored before a return. This behavior corresponds
+to marking such registers as permanently sealed (never unsealing them) in the context of
+observational integrity. (Note that while our specific properties leave registers unsealed,
+the formalization permits the same set of domains to apply to them as to addresses.
+
+Confidentiality can be applied similarly. We would permanently treat the stack pointer and
+argument registers as passed, and so exempt from confidentiality. Then, when a non-argument
+register is copied to the stack and later restored before return, the register value does
+not represent a visible difference in state between the entry and the return.
+
+We had briefly considered this complexity and chosen to avoid it, because it is an extra layer
+of detail to apply domains to registers.
+
 > As a minor question you don't need to answer, just take as low-level feedback: Do you /need/ to separate contexts (section 5.2) from policies, or is this just a matter of convenience since in practice various policies benefit from similar context information?
-> 
+
+The distinction is moral: policies are part of the system being evaluated, while contexts
+inform the property.
+
 > Another minor question: The difference between [nested] call-by-reference and passing address-of-local was interesting, but I'm not sure how much it makes sense to argue the latter is "beyond the scope of stack safety per se" [line 939] without arguing the same of the former.
 Perhaps the argument can be made that address-of-local can be put in the heap while a by-reference parameter cannot and since you don't guarantee heap safety, you shouldn't try to do anything useful with address-of-local?
-> 
+
+Yes, this is the sort of issue that makes address-of-local hard to handle without heap safety.
+The restrictions that we might place on such a pointer are typically restrictions that also
+apply to heap pointers, such as treating them like capabilities. In the longer term we would
+like to extend our model with a heap, and then the interplay between heap-derived and stack-derived
+pointers (with similar behaviors but different lifespans) will be much more interesting.
+
 > typos:
@@
 > * Any thoughts on callee-save registers?
-> 
-> 
+
 > 
@@
 > Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.
-> 
+
+We improve upon Skorstengaard et. al. in the following ways.
+1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning
+2. We therefore do not assume a static division of trusted and untrusted code
+
+Taken together, these distinctions give the practical advantage that we can reason about stack safety
+in contexts where a buggy function is compromised and used to attack other instances of the same function.
+
+3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
+understanding of a specific abstract machine
+4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
+machine being architecturally similar to the concrete one (beyond the specific call and return semantics
+that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
+abstract machine. Our properties apply directly to a wide range of architectures.
+
+We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
+to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
+safety is expressible as a formal security property.
+
 > Q2.
@@
 What about a language with delimited continuations that manipulate the stack?
-> 
+
+The section should be broken into assumptions that are necessarily for our properties, and assumptions
+that we explicitly do not make. Our stack safety model applies regardless of how a language uses non-stack,
+non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
+known in advance.
+
+Delimited continuations are not supported by our base model, so a language that has them would violate
+stack safety. We extend the base model with argument passing on the stack to demonstrate how it can
+accommodate a programming language feature that shares stack data between callers and callees. A
+similar extension to delimited continuations would be more complex.
+
 > Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?
-> 
-> 
-> 
+
+We want our properties to be applicable even to low-level code that may not keep function
+footprints isolated from one another. Consider a program that mixes C code with hand-written
+assembly -- the assembly may make calls and returns, which should be protected by stack safety,
+as well as arbitrary jumps, which should not.
+
+In Skorstengaard, an untrusted adversary can exhibit such behavior, but only trusted code can
+expect its data to be protected. We extend this protection to all callers, even those that
+exhibit unusual control flow.
+
 > Review #16C
@@
 > - The paper is titled "stack safety is a security property". I don't see why this is surprising, or what's the novelty here. Memory safety is almost always connected with security properties.
-> 
+
+The title refers to a formal notion of security property as a mathematical description of
+the possible behaviors of a system. This may be unclear to a reader unfamiliar with the concept.
+
 > Comments for author
@@
 > I might have missed something important. Are there any applications of the proposed formalization?
+
+A formal definition of a type of security is valuable in its own right. The main application
+is identifying bugs in real world enforcement mechanisms -- without such a definition, we
+can't necessarily agree that a given behavior is a bug. It gives us the ability to reason
+clearly about what is gained from applying a given enforcement mechanism to a system. Since
+many enforcement mechanisms are quite costly in speed, dollars, and silicon, it is important
+to understand precisely what they buy us.

### 2021-06-15 submission2/response.txt
Note on callee-save in StkTokens

@@
 > -------------
-> The paper gives an elegant, careful, formal treatment on what it means to ensure stack safety -- that callers and callees do not read or write each other's stacks, at least not in a way that has an observable impact, and that functions never return to the wrong place.
-This is formalized in a suitably abstract fashion with a range of more eager to more lazy definitions of correctness such that particular enforcement mechanisms can be analyzed for correctness.
-Property testing validates a particular eager algorithm from the literature and reveals problems, then corrected, with a lazier algorithm.
-Extensions toward allowing unobservable stack corruptions and supporting a simple form of coroutines are then developed.
+> The paper gives an elegant, careful, formal treatment
+> on what it means to ensure stack safety -- that callers and callees
+> do not read or write each other's stacks, at least not in a way that
+> has an observable impact, and that functions never return to the
+> wrong place.  This is formalized in a suitably abstract fashion with
+> a range of more eager to more lazy definitions of correctness such
+> that particular enforcement mechanisms can be analyzed for
+> correctness.  Property testing validates a particular eager
+> algorithm from the literature and reveals problems, then corrected,
+> with a lazier algorithm.  Extensions toward allowing unobservable
+> stack corruptions and supporting a simple form of coroutines are
+> then developed.
+
 > 
@@
 > ------------------------
-> I really enjoyed this paper and would highly recommend it to anyone working on low-level security mechanisms.
-The exposition is clear, precise, well-motivated, and in many ways synthesizing many years of insights into what it really means to protect the call stack.
-The examples are well-thought out, the definitions unimpeachable, and the clarity of thought impressive.
-Throughout reading the paper, I anticipated giving a 'strong accept' rating.
+> I really enjoyed this paper and would
+> highly recommend it to anyone working on low-level security
+> mechanisms.  The exposition is clear, precise, well-motivated, and
+> in many ways synthesizing many years of insights into what it really
+> means to protect the call stack.  The examples are well-thought out,
+> the definitions unimpeachable, and the clarity of thought
+> impressive.  Throughout reading the paper, I anticipated giving a
+> 'strong accept' rating.
 >
-> What gives me pause is the relative lack of validation or even discussion of why there isn't more validation.
-I don't know if the authors simply haven't had time yet, think it is unnecessary, or ran into challenges or complexities that remain unexplained.
-I do consider significant the property-based testing that was done both to validate the simpler tagging approach and identify (obvious in hindsight, but that's not a criticism) problems with the lazier tagging.
-But that still leaves:
+>  What gives me pause is the relative lack of validation or even
+> discussion of why there isn't more validation.  I don't know if the
+> authors simply haven't had time yet, think it is unnecessary, or ran
+> into challenges or complexities that remain unexplained.  I do
+> consider significant the property-based testing that was done both
+> to validate the simpler tagging approach and identify (obvious in
+> hindsight, but that's not a criticism) problems with the lazier
+> tagging.  But that still leaves:
 >
-> * Any discussion on whether it is feasible/desirable/in-progress/etc.
-to prove enforcement mechanisms correct.
-After all, the authors already have a fully formal definition and an implementation in Coq.
-While this sort of work does not necessarily need a correctness proof in Coq -- I still lean slightly toward accept after all -- it seems quite strange not to discuss even the possibility of validation including the ability to pursue a correctness argument for an enforcement mechanism beyond property testing.
+> * Any discussion on whether it is
+> feasible/desirable/in-progress/etc. to prove enforcement mechanisms
+> correct.  After all, the authors already have a fully formal
+> definition and an implementation in Coq.  While this sort of work
+> does not necessarily need a correctness proof in Coq -- I still lean
+> slightly toward accept after all -- it seems quite strange not to
+> discuss even the possibility of validation including the ability to
+> pursue a correctness argument for an enforcement mechanism beyond
+> property testing.
+
+We will extend the paper with a more detailed discussion of our roadmap for verification and 
+testing, and the issues that we expect to arise.
 
 These properties would absolutely be amenable to a correctness proof. Such a proof would be a
-lot of work in a setting with a full ISA, but there is not reason it should be impossible. The
-primary challenge we would predict in such an effort, beyond raw effort, is in determining for a given
+lot of work in a setting with a full ISA, but there is no reason it should be impossible. The
+primary challenge we would predict in such an effort is in determining for a given
 enforcement mechanism what invariants it needs to maintain in order to work correctly, and proving that
-it maintains them. This is very feasible for a proof of an individual, well-understood enforcement
-mechanism, but not for our goal of evaluating several such mechanisms.
+it maintains them. [APT: That last sentence is not a very surprising observation; suggest omitting it.]
+This is very feasible for a proof of an individual, well-understood enforcement
+mechanism, but not for our goal of evaluating several such mechanisms.  [APT: surely
+that depends on our resources.]
 
 > 
-> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
+> * The testing for the lazy tagging/clearing found mistakes but the
+>   "potential fix [is] itself a prime candidate for future testing" [line 868]
 > 
@@
 
-Here we should discuss our roadmap for testing in more detail. At present we are working to implement
-testing of observable properties alongside the fixed version of the lazy tagging/clearing micro-policy.
-Observational integrity is implemented, but confidentiality is not -- since the lazy policy enforced
-normal, stepwise confidentiality this is sufficient to test it. But we did not have this ready before
+At present we are working to implement
+testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
+Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
+normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
 the deadline.
 
-Our other major goal is to implement testing of a secondary architecture. We will evaluate a Cheri
+Our other major priority is to implement testing of a second enforecment architecture. We will evaluate a Cheri
 implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
-enforcement.
+enforcement. [APT: Current status of this?]
 
-We consider validation of the passing system to be a secondary priority and coroutines tertiary.
+Validation of the argument-passing and coroutining models are longer-term priorities.
 The passing system has existing enforcement mechanisms and serves an important role in common
@@
 
-> 
-> So: I believe this paper represents an important, sound conceptual advance that can inform the research community, but I was disappointed by the gaps in evaluation -- these gaps are acknowledged but not discussed (do they matter?
-how might they be closed?
-...).
-> 
+>  So: I believe this paper represents an important, sound conceptual
+> advance that can inform the research community, but I was
+> disappointed by the gaps in evaluation -- these gaps are
+> acknowledged but not discussed (do they matter? how might they be
+> closed? ...).
+
 > Comments for author
 > -------------------
-> There is one topic beyond my evaluation above worth considering: callee-save registers.
-I fear these may be fairly tricky to support in your set-up.
-Upon function entry, callee-save registers hold data private to the caller that may be passed arbitrarily deeper into the call stack, then saved into arbitrary stack locations, but then must be restored before return.
-Have you or the closely related prior work on enforcement mechanisms considered this complexity?
-The invariants are clear enough informally, but the issue seems to add substantial difficulty to tracking that a function faithfully implements the calling convention.
+> There is one topic beyond my evaluation above
+> worth considering: callee-save registers.  I fear these may be
+> fairly tricky to support in your set-up.  Upon function entry,
+> callee-save registers hold data private to the caller that may be
+> passed arbitrarily deeper into the call stack, then saved into
+> arbitrary stack locations, but then must be restored before return.
+> Have you or the closely related prior work on enforcement mechanisms
+> considered this complexity?  The invariants are clear enough
+> informally, but the issue seems to add substantial difficulty to
+> tracking that a function faithfully implements the calling
+> convention.
 
-Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
-They actually provide another use-case for observational properties! If we focus on integrity
-the reason becomes clear. Non-argument registers will naturally be overwritten (violating
+We believe that callee-save registers can be described in our model, although that would require tweaks to the domain system.
+[APT: for my info, what tweaks?]
+They actually provide another use case for observational properties! If we focus on integrity
+the reason becomes clear. Non-argument [APT: "callee-save" ??] registers will naturally be overwritten (violating
 stepwise integrity) but should always be restored before a return. This behavior corresponds
-to marking such registers as permanently sealed (never unsealing them) in the context of
+to marking such registers as permanently sealed (never unsealing them) [APT: ?? don't they get
+unsealed after return?] in the context of [APT:??]
 observational integrity. (Note that while our specific properties leave registers unsealed,
-the formalization permits the same set of domains to apply to them as to addresses.
+the formalization permits the same set of domains to apply to them [APT: ??] as to addresses.)
 
 Confidentiality can be applied similarly. We would permanently treat the stack pointer and
-argument registers as passed, and so exempt from confidentiality. Then, when a non-argument
+argument registers as passed, and so exempt from confidentiality. Then, when a non-argument [APT: "callee-save"??]
 register is copied to the stack and later restored before return, the register value does
-not represent a visible difference in state between the entry and the return.
+not represent a visible difference in state between the entry and the return. [APT: I didn't
+understand this para]
 
@@
 
-> As a minor question you don't need to answer, just take as low-level feedback: Do you /need/ to separate contexts (section 5.2) from policies, or is this just a matter of convenience since in practice various policies benefit from similar context information?
+It is worth pointing out that the prior work of Skorstengaard et al. also relies
+on a calling convention which clears all non-argument registers before calls and
+all non-return registers before returns.
 
-The distinction is moral: policies are part of the system being evaluated, while contexts
+> As a minor question you don't need to answer, just take as low-level
+> feedback: Do you /need/ to separate contexts (section 5.2) from
+> policies, or is this just a matter of convenience since in practice
+> various policies benefit from similar context information?
+
+The distinction is moral [APT:??]: policies are part of the system being evaluated, while contexts
 inform the property.
 
-> Another minor question: The difference between [nested] call-by-reference and passing address-of-local was interesting, but I'm not sure how much it makes sense to argue the latter is "beyond the scope of stack safety per se" [line 939] without arguing the same of the former.
-Perhaps the argument can be made that address-of-local can be put in the heap while a by-reference parameter cannot and since you don't guarantee heap safety, you shouldn't try to do anything useful with address-of-local?
+> Another minor question: The difference between [nested]
+> call-by-reference and passing address-of-local was interesting, but
+> I'm not sure how much it makes sense to argue the latter is "beyond
+> the scope of stack safety per se" [line 939] without arguing the
+> same of the former.  Perhaps the argument can be made that
+> address-of-local can be put in the heap while a by-reference
+> parameter cannot and since you don't guarantee heap safety, you
+> shouldn't try to do anything useful with address-of-local?
 
-Yes, this is the sort of issue that makes address-of-local hard to handle without heap safety.
+Yes, this is exactly the sort of issue that makes address-of-local hard to handle without heap safety.
 The restrictions that we might place on such a pointer are typically restrictions that also
@@
 > Paper summary
+
 > -------------
-> Low-level attacks exploit the unprotected access to the stack in languages like C.
-By writing outside the current stack frame, attackers may manipulate sensitive data and alter the control flow of the program.
-The paper formalizes the notion of _stack safety_, which protects against this type of attacks.
-Stack safety consists of two parts: Stack confidentiality and integrity, which say that protected regions on the stack cannot be read and altered.
-The definitions of stack safety target a language semantics in form of an abstract machine.
-The paper evaluates the approach by implementing an existing stack-safety enforcement mechanism and uncovering potential stack-safety violations.
+
+> Low-level attacks exploit the unprotected access to the stack in
+> languages like C. By writing outside the current stack frame,
+> attackers may manipulate sensitive data and alter the control flow
+> of the program. The paper formalizes the notion of _stack safety_,
+> which protects against this type of attacks. Stack safety consists
+> of two parts: Stack confidentiality and integrity, which say that
+> protected regions on the stack cannot be read and altered. The
+> definitions of stack safety target a language semantics in form of
+> an abstract machine. The paper evaluates the approach by
+> implementing an existing stack-safety enforcement mechanism and
+> uncovering potential stack-safety violations.
 > 
@@
 > Strengths:
-> - The problem that this paper solves is relevant. Countless attacks that exploit unprotected stack access and protection methods against these attacks emphasize the relevance of this problem.
-> - I was convinced by the evaluation of this paper. Their approach uncovered potential stack-safety violations in an existing enforcement mechanism by Reossler and DeHon.
+
+> - The problem that this paper solves is relevant. Countless attacks
+>   that exploit unprotected stack access and protection methods
+>   against these attacks emphasize the relevance of this problem.
+> - I was convinced by the evaluation of this paper. Their approach
+>   uncovered potential stack-safety violations in an existing
+>   enforcement mechanism by Reossler and DeHon.
+
+> Weaknesses:
+
+> - I question the novelty of this work. In particular, it is unclear
+>   how this paper improves upon Skorstengaard et al (question 1).
 >
-> Weaknesses:
-> - I question the novelty of this work. In particular, it is unclear how this paper improves upon Skorstengaard et al (question 1).
 > - It is unclear how this approach generalizes to other languages (question 2).
-> - The presentation of the paper can be improved.
-In particular, section 2 and 3 are very abstract because of the lack of examples and the readability of definitions in section 5 and 6 can be improved (see detailed comments).
 >
-> Because of the questionable novelty and a lacking presentation, I think this paper is not ready for publication yet.
+> - The presentation of the paper can be improved. In particular,
+> section 2 and 3 are very abstract because of the lack of examples
+> and the readability of definitions in section 5 and 6 can be
+> improved (see detailed comments).
+>
+> Because of the questionable novelty and a lacking presentation, I
+> think this paper is not ready for publication yet.
 > 
@@
 > =====================
-> 
-> The following comments are organised in the order of reading the paper front to back.
-On first read, I had some misunderstandings that got cleared up later, however, they may indicate places where the paper can be improved.
+> The following comments are organised in the order of reading the
+> paper front to back. On first read, I had some misunderstandings
+> that got cleared up later, however, they may indicate places where
+> the paper can be improved.
+
 > 
 > ### Section 1
-> - L32-40: It is unclear how this paper improves over Skorstengaard et al. Why isn't their definition of stack safety not already good enough?
+> - L32-40: It is unclear how this paper improves over Skorstengaard et al.
+>   Why isn't their definition of stack safety not already good enough?
 > 
@@
 > - L124: You did not introduce the acronyms "ISA" and "RISC"
-> - I found it quite hard to understand section 2 because the text is very abstract. I would have appreciated a few examples, like those you provide in section 4.
+> - I found it quite hard to understand section 2 because the text is very abstract.
+>   I would have appreciated a few examples, like those you provide in section 4.
 > - L148: Why do you need these policies?
-> - L154: In this paragraph you motivate why you need policies. I would have appreciated this motivation a paragraph earlier.
+> - L154: In this paragraph you motivate why you need policies.
+>   I would have appreciated this motivation a paragraph earlier.
 > 
 > ### Section 3
-> - The text of section 3 is equally hard to understand as in section 2 because of the lack of examples.
+> - The text of section 3 is equally hard to understand as in section 2
+>   because of the lack of examples.
 > 
@@
 > - L214: Remind the reader that the register `%sp` refers to the stack pointer.
-> - I like the example in figure 1. The figure would be easier to understand if you would write "unsealed" and "sealed" instead of "U" and "S".
+> - I like the example in figure 1. The figure would be easier to understand
+>   if you would write "unsealed" and "sealed" instead of "U" and "S".
 > - Figure 2: Why are there two stacks side by side?
@@
 > ### Section 6
-> - L499-509: There are so many different variable names that make it difficult to read the rules ("m", "cm", "dm", "rts", "tar", "ctx").
-These variable names are explained somewhere in the text, but their explanation is scatter across multiple pages.
-Consider declaring these variables in a table in a central place.
-> - Definition 6.2: Explain what the definition means instead of just stating it mathematically. What is pi_m?
-> - Consider applying these rules and definitions to an example of section 4 and explain where stack integrity and confidentiality is violated.
+> - L499-509: There are so many different variable names that make it
+>   difficult to read the rules ("m", "cm", "dm", "rts", "tar",
+>   "ctx"). These variable names are explained somewhere in the text,
+>   but their explanation is scatter across multiple pages.
+>   Consider declaring these variables in a table in a central place.
+> - Definition 6.2: Explain what the definition means instead of just
+>   stating it mathematically. What is pi_m?
+> - Consider applying these rules and definitions to an example of
+>   section 4 and explain where stack integrity and confidentiality is
+>   violated.
 >
 > ### Section 10
-> - The related work section seems a bit superficial since it only discusses 4 related works.
-Consider adding a discussion of the software and hardware protection techniques mentioned in the introduction.
+> - The related work section seems a bit superficial since it only
+>   discusses 4 related works. Consider adding a discussion of the
+>   software and hardware protection techniques mentioned in the introduction.
 > 
@@
 > ---------------------------------
-> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.
+> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction
+> nor related work section give a satisfying answer to this.
 
 We improve upon Skorstengaard et. al. in the following ways.
-1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning
-2. We therefore do not assume a static division of trusted and untrusted code
+1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning [APT: what does this mean?]
+2. We therefore do not assume a static division between trusted and untrusted code
 
@@
 3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
-understanding of a specific abstract machine
+understanding of a specific abstract machine [APT: I don't understand this distinction.]
 4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
 machine being architecturally similar to the concrete one (beyond the specific call and return semantics
-that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
+that distinguish them). Therefore it is less portable: each architecture it is applied to requires its own
 abstract machine. Our properties apply directly to a wide range of architectures.
@@
 
-> Q2.
-In L117-121 you state certain assumptions about the language semantics.
-It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply.
-For example, does a language with a heap allow you to break stack safety?
-What about a language with delimited continuations that manipulate the stack?
+> Q2. In L117-121 you state certain assumptions about the language
+> semantics. It is unclear to me if these assumptions restrict the
+> languages to which your stack safety definition apply. For example,
+> does a language with a heap allow you to break stack safety? What
+> about a language with delimited continuations that manipulate the
+> stack?
 
 The section should be broken into assumptions that are necessarily for our properties, and assumptions
-that we explicitly do not make. Our stack safety model applies regardless of how a language uses non-stack,
-non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
-known in advance.
+that we explicitly do not make. [APT: ??]  Our stack safety model applies regardless of how a language uses non-stack,
+non-code memory, including to languages with a heap. It applies to any layout of memory as long as it is
+known in advance.[ APT: "statically"? ]
 
@@
 
-> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?
+> Q3. Why don't you assume well-bracketed control flow like
+> Skorstengaard et al (L289-291)? What are the implications of that?
 
@@
 expect its data to be protected. We extend this protection to all callers, even those that
-exhibit unusual control flow.
+exhibit unusual control flow.  [APT: Don't think I get this.]
 
@@
 > -------------
-> The paper formalizes a definition of stack safety and presents extensions of it. The definition considers three properties: integrity, confidentiality, and well-bracketed control flow.
-> The authors also use these properties to validate the stack-safety “micro-policies” proposed by earlier work.
+> The paper formalizes a definition of stack safety and
+> presents extensions of it. The definition considers three properties:
+> integrity, confidentiality, and well-bracketed control flow.  The
+> authors also use these properties to validate the stack-safety
+> “micro-policies” proposed by earlier work.
 > 
@@
 > ------------------------
-> - It is a bit difficult to assess the merits of this work.
-Yes, the stack safety property is important for security, but what new results or new capabilities does the proposed definition bring in?
-If it is positioned as a security solution,  does it lead to more efficient defenses of control flow attacks?
-Does it prevent any new attacks that existing defenses fail?
-The paper does not talk about data flow across heap and stack, which is an important element for many real-world memory corruption attacks.
+
+> - It is a bit difficult to assess the merits of this work. Yes, the
+>   stack safety property is important for security, but what new results
+>   or new capabilities does the proposed definition bring in? If it is
+>   positioned as a security solution, does it lead to more efficient
+>   defenses of control flow attacks? Does it prevent any new attacks
+>   that existing defenses fail? The paper does not talk about data flow
+>   across heap and stack, which is an important element for many
+>   real-world memory corruption attacks.
 >
-> - The formalization seems intuitively correct, however there is no formal proof for its overall soundness properties and there is no real-world evaluation of its applications.
 >
-> - The paper is titled "stack safety is a security property". I don't see why this is surprising, or what's the novelty here. Memory safety is almost always connected with security properties.
+
+> - The formalization seems intuitively correct, however there is no
+>   formal proof for its overall soundness properties and there is no
+>   real-world evaluation of its applications.
+
+> - The paper is titled "stack safety is a security property". I don't
+>   see why this is surprising, or what's the novelty here. Memory safety
+>   is almost always connected with security properties.
 
 The title refers to a formal notion of security property as a mathematical description of
-the possible behaviors of a system. This may be unclear to a reader unfamiliar with the concept.
+the possible behaviors of a system. This may be unclear to a reader unfamiliar with the concept,
+and we will consider choosing a clearer title.
 

### 2021-06-15 submission2/response.txt
reformatted review text

@@
 > Strengths and Weaknesses
-> ------------------------
-> I really enjoyed this paper and would
+
+> ------------------------ I really enjoyed this paper and would
 > highly recommend it to anyone working on low-level security
@@
 > hindsight, but that's not a criticism) problems with the lazier
-> tagging.  But that still leaves:
->
-> * Any discussion on whether it is
+> tagging.  But that still leaves: * Any discussion on whether it is
 > feasible/desirable/in-progress/etc. to prove enforcement mechanisms
@@
 
-We will extend the paper with a more detailed discussion of our roadmap for verification and 
-testing, and the issues that we expect to arise.
-
 These properties would absolutely be amenable to a correctness proof. Such a proof would be a
-lot of work in a setting with a full ISA, but there is no reason it should be impossible. The
-primary challenge we would predict in such an effort is in determining for a given
+lot of work in a setting with a full ISA, but there is not reason it should be impossible. The
+primary challenge we would predict in such an effort, beyond raw effort, is in determining for a given
 enforcement mechanism what invariants it needs to maintain in order to work correctly, and proving that
-it maintains them. [APT: That last sentence is not a very surprising observation; suggest omitting it.]
-This is very feasible for a proof of an individual, well-understood enforcement
-mechanism, but not for our goal of evaluating several such mechanisms.  [APT: surely
-that depends on our resources.]
+it maintains them. This is very feasible for a proof of an individual, well-understood enforcement
+mechanism, but not for our goal of evaluating several such mechanisms.
 
@@
 
-At present we are working to implement
-testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
-Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
-normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
+Here we should discuss our roadmap for testing in more detail. At present we are working to implement
+testing of observable properties alongside the fixed version of the lazy tagging/clearing micro-policy.
+Observational integrity is implemented, but confidentiality is not -- since the lazy policy enforced
+normal, stepwise confidentiality this is sufficient to test it. But we did not have this ready before
 the deadline.
 
-Our other major priority is to implement testing of a second enforecment architecture. We will evaluate a Cheri
+Our other major goal is to implement testing of a secondary architecture. We will evaluate a Cheri
 implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
-enforcement. [APT: Current status of this?]
+enforcement.
 
-Validation of the argument-passing and coroutining models are longer-term priorities.
+We consider validation of the passing system to be a secondary priority and coroutines tertiary.
 The passing system has existing enforcement mechanisms and serves an important role in common
@@
 
-We believe that callee-save registers can be described in our model, although that would require tweaks to the domain system.
-[APT: for my info, what tweaks?]
-They actually provide another use case for observational properties! If we focus on integrity
-the reason becomes clear. Non-argument [APT: "callee-save" ??] registers will naturally be overwritten (violating
+Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
+They actually provide another use-case for observational properties! If we focus on integrity
+the reason becomes clear. Non-argument registers will naturally be overwritten (violating
 stepwise integrity) but should always be restored before a return. This behavior corresponds
-to marking such registers as permanently sealed (never unsealing them) [APT: ?? don't they get
-unsealed after return?] in the context of [APT:??]
+to marking such registers as permanently sealed (never unsealing them) in the context of
 observational integrity. (Note that while our specific properties leave registers unsealed,
-the formalization permits the same set of domains to apply to them [APT: ??] as to addresses.)
+the formalization permits the same set of domains to apply to them as to addresses.
 
 Confidentiality can be applied similarly. We would permanently treat the stack pointer and
-argument registers as passed, and so exempt from confidentiality. Then, when a non-argument [APT: "callee-save"??]
+argument registers as passed, and so exempt from confidentiality. Then, when a non-argument
 register is copied to the stack and later restored before return, the register value does
-not represent a visible difference in state between the entry and the return. [APT: I didn't
-understand this para]
+not represent a visible difference in state between the entry and the return.
 
@@
 of detail to apply domains to registers.
-
-It is worth pointing out that the prior work of Skorstengaard et al. also relies
-on a calling convention which clears all non-argument registers before calls and
-all non-return registers before returns.
 
@@
 
-The distinction is moral [APT:??]: policies are part of the system being evaluated, while contexts
+The distinction is moral: policies are part of the system being evaluated, while contexts
 inform the property.
@@
 
-Yes, this is exactly the sort of issue that makes address-of-local hard to handle without heap safety.
+Yes, this is the sort of issue that makes address-of-local hard to handle without heap safety.
 The restrictions that we might place on such a pointer are typically restrictions that also
@@
 We improve upon Skorstengaard et. al. in the following ways.
-1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning [APT: what does this mean?]
-2. We therefore do not assume a static division between trusted and untrusted code
+1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning
+2. We therefore do not assume a static division of trusted and untrusted code
 
@@
 3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
-understanding of a specific abstract machine [APT: I don't understand this distinction.]
+understanding of a specific abstract machine
 4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
 machine being architecturally similar to the concrete one (beyond the specific call and return semantics
-that distinguish them). Therefore it is less portable: each architecture it is applied to requires its own
+that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
 abstract machine. Our properties apply directly to a wide range of architectures.
@@
 The section should be broken into assumptions that are necessarily for our properties, and assumptions
-that we explicitly do not make. [APT: ??]  Our stack safety model applies regardless of how a language uses non-stack,
-non-code memory, including to languages with a heap. It applies to any layout of memory as long as it is
-known in advance.[ APT: "statically"? ]
+that we explicitly do not make. Our stack safety model applies regardless of how a language uses non-stack,
+non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
+known in advance.
 
@@
 expect its data to be protected. We extend this protection to all callers, even those that
-exhibit unusual control flow.  [APT: Don't think I get this.]
+exhibit unusual control flow.
 
@@
 The title refers to a formal notion of security property as a mathematical description of
-the possible behaviors of a system. This may be unclear to a reader unfamiliar with the concept,
-and we will consider choosing a clearer title.
+the possible behaviors of a system. This may be unclear to a reader unfamiliar with the concept.
 

### 2021-06-15 submission2/response.txt
a few fixes, comments

@@
 > Strengths and Weaknesses
-
-> ------------------------ I really enjoyed this paper and would
+> ------------------------
+> I really enjoyed this paper and would
 > highly recommend it to anyone working on low-level security
@@
 > hindsight, but that's not a criticism) problems with the lazier
-> tagging.  But that still leaves: * Any discussion on whether it is
+> tagging.  But that still leaves:
+>
+> * Any discussion on whether it is
 > feasible/desirable/in-progress/etc. to prove enforcement mechanisms
@@
 These properties would absolutely be amenable to a correctness proof. Such a proof would be a
-lot of work in a setting with a full ISA, but there is not reason it should be impossible. The
+lot of work in a setting with a full ISA, but there is no reason it should be impossible. The
 primary challenge we would predict in such an effort, beyond raw effort, is in determining for a given
@@
 
-Our other major goal is to implement testing of a secondary architecture. We will evaluate a Cheri
+Our other major goal is to implement testing of a second enforecment architecture. We will evaluate a Cheri
 implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
@@
 
-We consider validation of the passing system to be a secondary priority and coroutines tertiary.
+Validation of the argument-passing and coroutining models are longer-term priorities.
 The passing system has existing enforcement mechanisms and serves an important role in common
@@
 3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
-understanding of a specific abstract machine
+understanding of a specific abstract machine [APT: I don't understand this distinction.]
 4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract

### 2021-06-15 src/response_formal.txt
Added formal response

** Overview **

** Changes **

1. Give a more thorough discussion of validation. Discuss the feasibility of proofs
   and of randomized testing of more complex properties. Detail our roadmap for
   further validation of different versions of the property.

2. Give an explicit, direct comparison to the work of Skorstengaard et. al. Currently
   we discuss them briefly in our related work section, but our novelty has not been
   obvious, so we can be more thorough in explaining the differences.

3. There are a number of improvements we can make to the presentation. We need to
   gradually introduce the examples and associated intuition (including our visual
   metaphors) as early as possible, starting in the introduction.

4. Draw a clearer distinction between language features that are compatible with our
   model but not assumed, those that are required for the model to function,
   and those that are incompatible.

** Detailed Comments **

* Reviewer A *

We propose above to add further discussion of validation. Here is a rough
outline of how it would cover your concerns.

> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.

These properties would absolutely be amenable to a correctness proof. Such a proof would be a
lot of work in a setting with a full ISA, but there is not reason it should be impossible. The
primary challenge we would predict in such an effort, beyond raw effort, is in determining for a given
enforcement mechanism what invariants it needs to maintain in order to work correctly, and proving that
it maintains them. This is very feasible for a proof of an individual, well-understood enforcement
mechanism, but not for our goal of evaluating several such mechanisms.

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
> 
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

Here we should discuss our roadmap for testing in more detail. At present we are working to implement
testing of observable properties alongside the fixed version of the lazy tagging/clearing micro-policy.
Observational integrity is implemented, but confidentiality is not -- since the lazy policy enforced
normal, stepwise confidentiality this is sufficient to test it. But we did not have this ready before
the deadline. Note that this has been implemented since submission, so we will need to take care not
to describe it as a contribution.

Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement.

We consider validation of the passing system to be a secondary priority and coroutines tertiary.
The passing system has existing enforcement mechanisms and serves an important role in common
stack use-cases. Coroutines will require the development of an enforcement mechanism (likely
tag-based) to validate.

> There is one topic beyond my evaluation above worth considering: callee-save registers.
> I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save
> registers hold data private to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored before return.
> Have you or the closely related prior work on enforcement mechanisms considered this complexity?
> The invariants are clear enough informally, but the issue seems to add substantial difficulty to
> tracking that a function faithfully implements the calling convention.

Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
They actually provide another use-case for observational properties! Focusing on integrity:
callee-save registers may be overwritten (violating stepwise integrity) but should always be
restored before a return. This behavior corresponds to marking such registers as permanently sealed
in the context of observational integrity -- recall that observational integrity compares sealed
components at the start and end of a callee, so the interim changes are not violations.
(Note as well that the formalization permits domains such as "sealed" to apply to registers as well
as to addresses, we just haven't taken advantage of this ability.)

Confidentiality can be applied similarly. We would permanently treat the stack pointer and
argument registers as "passed", and so exempt from confidentiality. Then, when a callee-save
register is copied to the stack and later restored before return, the register value does
not represent a visible difference in state between the entry and the return.

We had briefly considered this complexity and chosen to avoid it, because it is an extra layer
of detail to apply domains to registers.

> Another minor question: The difference between [nested] call-by-reference and passing
> address-of-local was interesting, but I'm not sure how much it makes sense to argue the
> latter is "beyond the scope of stack safety per se" [line 939] without arguing the same
> of the former.  Perhaps the argument can be made that address-of-local can be put in the
> heap while a by-reference parameter cannot and since you don't guarantee heap safety,
> you shouldn't try to do anything useful with address-of-local?

Yes, this is the sort of issue that makes address-of-local hard to handle without heap safety.
More broadly the issue is one of assuming intent on the part of the code. We tend to
assume that an address-of-local should be treated similarly to the address of a heap object,
but with a different lifetime. For instance, if a heap model treats heap pointers like capabilities,
we would treat stack pointers the same. In heap model that permits pointer forging (restricted to
within compartments, say,) we would expect to be able to forge pointers to address-taken locals
in the same circumstances. By contrast, the expected behavior of stack allocated arguments is
closely tied to that of the stack itself.

In the longer term we would like to extend our model with a heap, and then the interplay between
heap-derived and stack-derived pointers (with similar behaviors but different lifespans) will
be much more interesting.

* Reviewer B *

> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M

We improve upon Skorstengaard et. al. in the following ways.
1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning
2. We therefore do not assume a static division of trusted and untrusted code

Taken together, these distinctions give the practical advantage that we can reason about stack safety
in contexts where a buggy function is compromised and used to attack other instances of the same function.

3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
understanding of a specific abstract machine
4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties apply directly to a wide range of architectures.

We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
safety is expressible as a formal security property.

> Q2. In L117-121 you state certain assumptions about the language semantics. It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply. For example, does a language with a heap allow you to break stack safety? What about a language with delimited continuations that manipulate the stack?^M

The section should be broken into assumptions that are necessarily for our properties, and assumptions
that we explicitly do not make. Our stack safety model applies regardless of how a language uses non-stack,
non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
known in advance.

Delimited continuations are not supported by our base model, so a language that has them would violate
stack safety. We extend the base model with argument passing on the stack to demonstrate how it can
accommodate a programming language feature that shares stack data between callers and callees. A
similar extension to delimited continuations would be more complex.

> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?

We want our properties to be applicable even to low-level code that may not keep function
footprints isolated from one another. Consider a program that mixes C code with hand-written
assembly -- the assembly may make calls and returns, which should be protected by stack safety,
as well as arbitrary jumps, which should not.

In Skorstengaard, an untrusted adversary can exhibit such behavior, but only trusted code can
expect its data to be protected. We extend this protection to all callers, even those that
exhibit unusual control flow.

* Reviewer C *

> I might have missed something important. Are there any applications of the proposed formalization?

A formal definition of a type of security is valuable in its own right. The main application
is identifying bugs in real world enforcement mechanisms -- without such a definition, we
can't necessarily agree that a given behavior is a bug. It gives us the ability to reason
clearly about what is gained from applying a given enforcement mechanism to a system. Since
many enforcement mechanisms are quite costly in speed, dollars, and silicon, it is important
to understand precisely what they buy us.

### 2021-06-15 submission2/response.txt
more comments; done for now

--- 2021-06-15 submission2/response.txt
@@
 that depends on our resources.]
+[RB: Agreed, it could also be framed in terms of testing as a potential
+low effort, high reward strategy.]
 
@@
 
-It is worth pointing out that the prior work of Skorstengaard et al. also relies
-on a calling convention which clears all non-argument registers before calls and
-all non-return registers before returns.
+[RB: It is worth pointing out that the prior work of Skorstengaard et al. also
+relies on a calling convention which clears all non-argument registers before
+calls and all non-return registers before returns.]
 

### 2021-06-15 submission2/response_formal.txt
Added formal response (properly)

--- 2021-06-15 src/response_formal.txt
@@
 
-We propose above to add further discussion of validation. Here is a rough
+We will extend the paper with a more detailed discussion of our roadmap for verification and
+testing, and the issues that we expect to arise. Here is a rough
 outline of how it would cover your concerns.
@@
 These properties would absolutely be amenable to a correctness proof. Such a proof would be a
-lot of work in a setting with a full ISA, but there is not reason it should be impossible. The
-primary challenge we would predict in such an effort, beyond raw effort, is in determining for a given
-enforcement mechanism what invariants it needs to maintain in order to work correctly, and proving that
-it maintains them. This is very feasible for a proof of an individual, well-understood enforcement
-mechanism, but not for our goal of evaluating several such mechanisms.
+lot of work in a setting with a full ISA, but there is not reason it should be impossible.
+Testing simply makes more sense for our goal of evaluating multiple enforcement mechanisms
+to uncover possible bugs -- it is lower effort, but if bugs are found, high reward.
 
@@
 
-Here we should discuss our roadmap for testing in more detail. At present we are working to implement
-testing of observable properties alongside the fixed version of the lazy tagging/clearing micro-policy.
-Observational integrity is implemented, but confidentiality is not -- since the lazy policy enforced
-normal, stepwise confidentiality this is sufficient to test it. But we did not have this ready before
-the deadline. Note that this has been implemented since submission, so we will need to take care not
+At present we are working to implement
+testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
+Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
+normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
+the deadline. We will therefore take care not
 to describe it as a contribution.
@@
 implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
-enforcement.
+enforcement. [APT: Current status of this?]
 
@@
 Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
+[APT: for my info, what tweaks?] [SNA: I should be clearer -- it needs its own initCtx and updateCtx functions.]
 They actually provide another use-case for observational properties! Focusing on integrity:
@@
 restored before a return. This behavior corresponds to marking such registers as permanently sealed
-in the context of observational integrity -- recall that observational integrity compares sealed
+while enforcing observational integrity -- recall that observational integrity compares sealed
 components at the start and end of a callee, so the interim changes are not violations.
@@
 register is copied to the stack and later restored before return, the register value does
-not represent a visible difference in state between the entry and the return.
+not represent a visible difference in state between the entry and the return. [APT: I didn't
+understand this para]
 
@@
 of detail to apply domains to registers.
+
+[RB: It is worth pointing out that the prior work of Skorstengaard et al. also
+relies on a calling convention which clears all non-argument registers before
+calls and all non-return registers before returns.]
 
@@
 We improve upon Skorstengaard et. al. in the following ways.
-1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning
+1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning [APT: what does this mean?]
 2. We therefore do not assume a static division of trusted and untrusted code
@@
 The section should be broken into assumptions that are necessarily for our properties, and assumptions
-that we explicitly do not make. Our stack safety model applies regardless of how a language uses non-stack,
+that we explicitly do not make. [APT: ??] Our stack safety model applies regardless of how a language uses non-stack,
 non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
-known in advance.
+known statically.
 
@@
 expect its data to be protected. We extend this protection to all callers, even those that
-exhibit unusual control flow.
+exhibit unusual control flow. [APT: Don't think I get this.]
 

### 2021-06-15 submission2/response_formal.txt
Cerise fit

@@
 enforcement. [APT: Current status of this?]
+[RB: SNA refactored the models and Cerise (the capability machine development)
+fits quite well in it, so a lot of the machinery for generation and testing is
+easily reusable. The tricky part will probably be the generation of blessed
+sequences corresponding to the StkTokens-like calling convention, as this is
+quite involved.]
 

### 2021-06-16 submission2/response_formal.txt
Polishing and comments

@@
 
-These properties would absolutely be amenable to a correctness proof. Such a proof would be a
-lot of work in a setting with a full ISA, but there is not reason it should be impossible.
-Testing simply makes more sense for our goal of evaluating multiple enforcement mechanisms
-to uncover possible bugs -- it is lower effort, but if bugs are found, high reward.
+These properties would absolutely be amenable to a correctness proof. Such a
+proof would be a lot of work in a setting with a full ISA, but there is no
+reason it should be impossible.  We concentrated here on testing because it
+is better aligned with our immediate goal of evaluating multiple enforcement
+mechanisms to uncover possible bugs -- it is lower effort, but if bugs are
+found, high reward.
+
+[BCP: Might be a little dangerous to say "multiple" here because we didn't
+actually do very many yet.]
 
@@
 the deadline. We will therefore take care not
-to describe it as a contribution.
+to describe it as a contribution.  [BCP: Last sentence not needed.]
 
-Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
-implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
-enforcement. [APT: Current status of this?]
-[RB: SNA refactored the models and Cerise (the capability machine development)
-fits quite well in it, so a lot of the machinery for generation and testing is
-easily reusable. The tricky part will probably be the generation of blessed
-sequences corresponding to the StkTokens-like calling convention, as this is
-quite involved.]
+Our other immediate goal is to implement testing of a second, rather
+different architecture. We plan to evaluate a Cheri implementation of stack
+safety to demonstrate that our properties are not strictly tied to tag-based
+enforcement. [APT: Current status of this?]  [RB: SNA refactored the models
+and Cerise (the capability machine development) fits quite well in it, so a
+lot of the machinery for generation and testing is easily reusable. The
+tricky part will probably be the generation of blessed sequences
+corresponding to the StkTokens-like calling convention, as this is quite
+involved.] [BCP: And thus likely to be wrong. :-) ]
 
-We consider validation of the passing system to be a secondary priority and coroutines tertiary.
-The passing system has existing enforcement mechanisms and serves an important role in common
-stack use-cases. Coroutines will require the development of an enforcement mechanism (likely
-tag-based) to validate.
+We consider validation of the passing system [BCP: ??] to be a secondary
+priority and coroutines tertiary.  The passing system has existing
+enforcement mechanisms and serves an important role in common stack
+use-cases. Coroutines will require the development of an enforcement
+mechanism (e.g., one based on tags) to validate.
 
@@
 
-We had briefly considered this complexity and chosen to avoid it, because it is an extra layer
+We had briefly considered adding this complexity and chosen to avoid it, because it is an extra layer
 of detail to apply domains to registers.
@@
 relies on a calling convention which clears all non-argument registers before
-calls and all non-return registers before returns.]
+calls and all non-return registers before returns.] [BCP: Yes!]
 
@@
 
-In the longer term we would like to extend our model with a heap, and then the interplay between
-heap-derived and stack-derived pointers (with similar behaviors but different lifespans) will
+In the longer term we would like to extend our model with a heap,[BCP: Cite
+Arthur's paper and point out that it is phrased in similar terms and should
+thus be compatible!] and then the interplay between heap-derived and
+stack-derived pointers (with similar behaviors but different lifespans) will
 be much more interesting.
@@
 Taken together, these distinctions give the practical advantage that we can reason about stack safety
-in contexts where a buggy function is compromised and used to attack other instances of the same function.
+in contexts where a buggy function is compromised and used to attack other
+instances of the same function.  [BCP: Example? Plus a clear explanation of
+why S. et al cannot handle it while we can.]
 
@@
 that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
-abstract machine. Our properties apply directly to a wide range of architectures.
+abstract machine. Our properties should apply directly to a wide range of
+architectures.
 
@@
 
+> - It is a bit difficult to assess the merits of this work. Yes, the
+>   stack safety property is important for security, but what new results
+>   or new capabilities does the proposed definition bring in? If it is
+>   positioned as a security solution, does it lead to more efficient
+>   defenses of control flow attacks? Does it prevent any new attacks
+>   that existing defenses fail? The paper does not talk about data flow
+>   across heap and stack, which is an important element for many
+>   real-world memory corruption attacks.
+
+[BCP: We definitely need to respond directly to this.  First, we should
+agree that the title is confusing (of course it is a security property) and
+needs changing, and then we should say clearly what we meant by it.]
+
 > I might have missed something important. Are there any applications of the proposed formalization?
@@
 to understand precisely what they buy us.
+
+[BCP: Also, we intended to use it for bug finding, and the case study
+confirms that it is suitable for this purpose.]

### 2021-06-16 submission2/response_formal.txt
More detail on response

--- 2021-06-15 submission2/response_formal.txt
@@
 implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
-enforcement. [APT: Current status of this?]
+enforcement. We are currently working on the QuickChick generators for Cheri programs, which have a
+different (more complicated) calling convention.
 
@@
 
-Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
-[APT: for my info, what tweaks?] [SNA: I should be clearer -- it needs its own initCtx and updateCtx functions.]
+Callee-save registers are plausible in our setup, using a version of the sharing model
+with its context initialization and updates modified to apply appropriate domains to registers.
 They actually provide another use-case for observational properties! Focusing on integrity:
 callee-save registers may be overwritten (violating stepwise integrity) but should always be
-restored before a return. This behavior corresponds to marking such registers as permanently sealed
-while enforcing observational integrity -- recall that observational integrity compares sealed
+restored before a return. This requirement corresponds to observational integrity with the
+callee-save registers sealed -- recall that observational integrity compares sealed
 components at the start and end of a callee, so the interim changes are not violations.
-(Note as well that the formalization permits domains such as "sealed" to apply to registers as well
-as to addresses, we just haven't taken advantage of this ability.)
 
-Confidentiality can be applied similarly. We would permanently treat the stack pointer and
-argument registers as "passed", and so exempt from confidentiality. Then, when a callee-save
-register is copied to the stack and later restored before return, the register value does
-not represent a visible difference in state between the entry and the return. [APT: I didn't
-understand this para]
+Confidentiality can be applied similarly. A callee-save register should be secret, but
+in observational confidentiality, if it is copied to the stack and then restored before
+return, the property will allow the original register to be read normally after return.
+But the property will not allow it to be leaked in any other way.
 
@@
 We improve upon Skorstengaard et. al. in the following ways.
-1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning [APT: what does this mean?]
+1. We capture a desireable security characteristic of dynamically nested functions, that every caller is safe from its callee
 2. We therefore do not assume a static division of trusted and untrusted code
@@
 
-3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
-understanding of a specific abstract machine [APT: I don't understand this distinction.]
-4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
+3. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
 machine being architecturally similar to the concrete one (beyond the specific call and return semantics
@@
 
-The section should be broken into assumptions that are necessarily for our properties, and assumptions
-that we explicitly do not make. [APT: ??] Our stack safety model applies regardless of how a language uses non-stack,
+We will edit section 2 to draw a clearer distinction between language features
+that are compatible with our model but not assumed, those that are required for
+the model to function, and those that are incompatible.
+
+Our stack safety model applies regardless of how a language uses non-stack,
 non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is

### 2021-06-16 submission2/response_formal.txt

@@
 normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
-the deadline. We will therefore take care not
-to describe it as a contribution.
+the deadline.
 
@@
 enforcement. We are currently working on the QuickChick generators for Cheri programs, which have a
-different (more complicated) calling convention.
+different calling convention -- a more complicated one, which is easier to get wrong.
 
-We consider validation of the passing system to be a secondary priority and coroutines tertiary.
-The passing system has existing enforcement mechanisms and serves an important role in common
-stack use-cases. Coroutines will require the development of an enforcement mechanism (likely
-tag-based) to validate.
+We consider validation of the passing system [BCP: ??] to be a secondary
+priority and coroutines tertiary.  The passing system has existing
+enforcement mechanisms and serves an important role in common stack
+use-cases. Coroutines will require the development of an enforcement
+mechanism (e.g., one based on tags) to validate.
 

### 2021-06-16 submission2/response_formal.txt
Merged some of Andrew's changes I'd missed into response_formal

@@
 
-Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
+Our other immediate priority is to implement testing of a secondary enforcement mechanism. We will evaluate a Cheri
 implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
@@
 
-We consider validation of the passing system [BCP: ??] to be a secondary
-priority and coroutines tertiary.  The passing system has existing
-enforcement mechanisms and serves an important role in common stack
+Validation of the argument-passing and coroutining models are longer-term priorities.
+The passing system has existing enforcement mechanisms and serves an important role in common stack
 use-cases. Coroutines will require the development of an enforcement
@@
 
-Yes, this is the sort of issue that makes address-of-local hard to handle without heap safety.
+Yes, this is exactly the sort of issue that makes address-of-local hard to handle without heap safety.
 More broadly the issue is one of assuming intent on the part of the code. We tend to

### 2021-06-16 submission2/response_formal.txt
Taking a stab at overview

@@
 ** Overview **
+
+These reviews have identified some areas that the paper failed to explain adequately.
+Fortunately none of these issues are in the formalism itself, but meta questions: does
+the work improve on the competition? Do we have a plan to validate the model further?
+Can it be extended to other topics? We have answers to these questions, and we can
+address them explicitly in the paper.
 

### 2021-06-16 submission2/response_formal.txt
comments and polishing

--- 2021-06-16 submission2/response_formal.txt
@@
 ** Overview **
+
+We thank the reviewers for their comments. The reviews identify 
+some areas that the paper fails to explain adequately.
+Fortunately [APT: hmm] none of these issues are in the formalism itself, but meta questions: How does
+the work improve on existing formalizations of stack safety? Do we have a plan to validate the model further?
+Can our work be extended to cover more complex uses of the stack? We have answers to these questions, and we can
+address them explicitly in the paper.
 
@@
 
-1. Give a more thorough discussion of validation. Discuss the feasibility of proofs
-   and of randomized testing of more complex properties. Detail our roadmap for
+1. We will give a more thorough discussion of validation, discuss the feasibility of proofs
+   and of randomized testing of more complex properties, and detail our roadmap for
    further validation of different versions of the property.
 
-2. Give an explicit, direct comparison to the work of Skorstengaard et. al. Currently
-   we discuss them briefly in our related work section, but our novelty has not been
-   obvious, so we can be more thorough in explaining the differences.
+2. We will give an explicit, direct comparison to the work of Skorstengaard et. al. Currently
+   we discuss it briefly in our related work section, but our novelty has not been
+   obvious, so we will be more thorough in explaining the differences.
 
-3. There are a number of improvements we can make to the presentation. We need to
-   gradually introduce the examples and associated intuition (including our visual
-   metaphors) as early as possible, starting in the introduction.
+3. We will improve the presentation as suggested. We need to gradually
+   introduce the examples and associated intuition (including our
+   visual metaphors) as early as possible, starting in the
+   introduction.
 
-4. Draw a clearer distinction between language features that are compatible with our
-   model but not assumed, those that are required for the model to function,
+4. We will more clearly distinguish language features that are compatible with our
+   model but not assumed [APT: don't understand this phrasing], those that are required for the model to function,
    and those that are incompatible.
@@
 These properties would absolutely be amenable to a correctness proof. Such a
-proof would be a lot of work in a setting with a full ISA, but there is no
-reason it should be impossible.  We concentrated here on testing because it
+proof would be a lot of work in a setting with a full ISA, but we see no
+fundamental barriers.  We chose to concentrate on testing because it
 is better aligned with our immediate goal of evaluating multiple enforcement
@@
 
-Callee-save registers are plausible in our setup, using a version of the sharing model
+Callee-save registers are compatible with our approach, using a version of the sharing model
 with its context initialization and updates modified to apply appropriate domains to registers.
@@
 
-Yes, this is exactly the sort of issue that makes address-of-local hard to handle without heap safety.
-More broadly the issue is one of assuming intent on the part of the code. We tend to
-assume that an address-of-local should be treated similarly to the address of a heap object,
-but with a different lifetime. For instance, if a heap model treats heap pointers like capabilities,
-we would treat stack pointers the same. In heap model that permits pointer forging (restricted to
-within compartments, say,) we would expect to be able to forge pointers to address-taken locals
-in the same circumstances. By contrast, the expected behavior of stack allocated arguments is
-closely tied to that of the stack itself.
+Yes, this was exactly our thinking.  
+We know what constitutes legal access to stack-allocated arguments, but for explicit pointers
+this is much less clear, and depends on the intent of the programmer. We believe
+an address-of-local should be treated similarly to the address of a heap object,
+but with a different lifetime. For instance, if a heap safety model treats heap pointers like capabilities
+(possession implies access rights), we would treat stack pointers the same.
+Similarly, if the heap model permits pointer forging (restricted to
+within compartments, say) we would expect to be able to forge pointers to address-taken locals
+in the same circumstances.
 
@@
 We will edit section 2 to draw a clearer distinction between language features
-that are compatible with our model but not assumed, those that are required for
+that are compatible with our model but not assumed [APT: again, don't understand this language], those that are required for
 the model to function, and those that are incompatible.
@@
 Our stack safety model applies regardless of how a language uses non-stack,
-non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
+non-code memory, so it works for languages with a heap. It applies to any layout of memory as long as that is
 known statically.

### 2021-06-16 submission2/response_formal.txt
More work

@@
 some areas that the paper fails to explain adequately.
-Fortunately [APT: hmm] none of these issues are in the formalism itself, but meta questions: How does
+None of these issues are in the formalism itself, but meta questions: How does
 the work improve on existing formalizations of stack safety? Do we have a plan to validate the model further?
@@
 
-4. We will more clearly distinguish language features that are compatible with our
-   model but not assumed [APT: don't understand this phrasing], those that are required for the model to function,
-   and those that are incompatible.
+4. Our model avoids talking about language features other than the stack, but it is unclear at times
+   whether those features could be added without compromising stack safety. We will more clearly call out
+   common language features, such as heaps, that are compatible with stack safety.
+
+[SNA: Do we want to promise to add callee-saves registers? I wasn't going to at first, but actually there's
+      a good story here. Something like:]
+5. We will add a discussion of how to extend the model to a callee-saves registers convention, which
+   demonstrates a concrete use for observational properties and a clear advance over the state-of-the-art.
 
@@
 fundamental barriers.  We chose to concentrate on testing because it
-is better aligned with our immediate goal of evaluating multiple enforcement
+is better aligned with our immediate goal of evaluating enforcement
 mechanisms to uncover possible bugs -- it is lower effort, but if bugs are
-found, high reward.
-
-[BCP: Might be a little dangerous to say "multiple" here because we didn't
-actually do very many yet.]
+found, still high reward.
 
@@
 We had briefly considered adding this complexity and chosen to avoid it, because it is an extra layer
-of detail to apply domains to registers.
-
-[RB: It is worth pointing out that the prior work of Skorstengaard et al. also
-relies on a calling convention which clears all non-argument registers before
-calls and all non-return registers before returns.] [BCP: Yes!]
+of detail to apply domains to registers. It is worth noting that the primary existing work in this
+area, Skorstengaard et al., relies on a calling convention which clears all non-argument registers before
+calls and all non-return registers before returns.
 
@@
 
-In the longer term we would like to extend our model with a heap,[BCP: Cite
-Arthur's paper and point out that it is phrased in similar terms and should
-thus be compatible!] and then the interplay between heap-derived and
+In the longer term we would like to extend our model with a heap, in the style of
+Azevedo de Amorim et al.'s "The Meaning of Memory Safety," which is phrased in similar
+terms to our model and should be compatible. Then the interplay between heap-derived and
 stack-derived pointers (with similar behaviors but different lifespans) will
@@
 
-We will edit section 2 to draw a clearer distinction between language features
-that are compatible with our model but not assumed [APT: again, don't understand this language], those that are required for
-the model to function, and those that are incompatible.
-
 Our stack safety model applies regardless of how a language uses non-stack,
@@
 
+This distinction, between a feature that could be added without changing the properties and one that
+would require enhanced properties, is important, and we will clarify in section 2 which features fall
+into which category.
+
 > Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?
@@
 as well as arbitrary jumps, which should not.
-
-In Skorstengaard, an untrusted adversary can exhibit such behavior, but only trusted code can
-expect its data to be protected. We extend this protection to all callers, even those that
-exhibit unusual control flow. [APT: Don't think I get this.]
 
@@
 
+> - The paper is titled "stack safety is a security property". I don't
+>   see why this is surprising, or what's the novelty here. Memory safety
+>   is almost always connected with security properties.
+
+We agree that the title is confusing and should be changed. It refers to a concept in 
+theoretical security: a formal notion of security property as a mathematical description of
+the possible behaviors of a system. In that context, it was not actually clear that stack
+safety could be defined in these terms.
+
 [BCP: We definitely need to respond directly to this.  First, we should
@@
 
-A formal definition of a type of security is valuable in its own right. The main application
-is identifying bugs in real world enforcement mechanisms -- without such a definition, we
-can't necessarily agree that a given behavior is a bug. It gives us the ability to reason
+A formal definition of a type of security is valuable in its own right. It tells us whether
+any given behavior is a violation, and we can therefore use it to identify bugs. In our case
+study we identify a previously unknown bug in Roessler and Dehon's proposed "lazy tagging" micro-policy,
+and also use randomized testing to increase confidence that their "depth isolation" micro-policy
+correctly implements stack safety.
+
+More broadly, it gives us the ability to reason
 clearly about what is gained from applying a given enforcement mechanism to a system. Since
@@
 to understand precisely what they buy us.
-
-[BCP: Also, we intended to use it for bug finding, and the case study
-confirms that it is suitable for this purpose.]

### 2021-06-16 submission2/response_formal.txt
Handling comments

@@
 We improve upon Skorstengaard et. al. in the following ways.
-1. We capture a desireable security characteristic of dynamically nested functions, that every caller is safe from its callee
+1. We capture a desireable security characteristic of dynamically nested calls, that every caller is safe from its callee
 2. We therefore do not assume a static division of trusted and untrusted code
 
-Taken together, these distinctions give the practical advantage that we can reason about stack safety
-in contexts where a buggy function is compromised and used to attack other
-instances of the same function.  [BCP: Example? Plus a clear explanation of
-why S. et al cannot handle it while we can.]
+Specifically, Skorstengaard et. al. divide the world into trusted functions, which feature
+additional dynamic constraints on their behavior, and for which their properties hold,
+and adversaries, which may violate the dynamic "reasonableness" constraints. Suppose we
+have a single function that recursively traverses a linked list, loading data into local
+memory. What if it has a bug, and certain data cause it to violate the reasonableness condition?
+In Skorstengaard's model, it must be an adversary, and activations lower in the stack are not
+guaranteed protection. In our model, only the badly behaved activation is an attacker, and its
+caller(s) are guaranteed protection.
 
@@
 >   stack safety property is important for security, but what new results
->   or new capabilities does the proposed definition bring in? If it is
->   positioned as a security solution, does it lead to more efficient
->   defenses of control flow attacks? Does it prevent any new attacks
->   that existing defenses fail? The paper does not talk about data flow
->   across heap and stack, which is an important element for many
->   real-world memory corruption attacks.
-
-> - The paper is titled "stack safety is a security property". I don't
->   see why this is surprising, or what's the novelty here. Memory safety
->   is almost always connected with security properties.
-
-We agree that the title is confusing and should be changed. It refers to a concept in 
-theoretical security: a formal notion of security property as a mathematical description of
-the possible behaviors of a system. In that context, it was not actually clear that stack
-safety could be defined in these terms.
-
-[BCP: We definitely need to respond directly to this.  First, we should
-agree that the title is confusing (of course it is a security property) and
-needs changing, and then we should say clearly what we meant by it.]
-
-> I might have missed something important. Are there any applications of the proposed formalization?
+>   or new capabilities does the proposed definition bring in?
 
@@
 to understand precisely what they buy us.
+
+>   The paper does not talk about data flow
+>   across heap and stack, which is an important element for many
+>   real-world memory corruption attacks.
+
+We plan to combine our stack model with a heap model in the style of Acevedo de Amorim et al.
+as future work.
+
+> - The paper is titled "stack safety is a security property". I don't
+>   see why this is surprising, or what's the novelty here. Memory safety
+>   is almost always connected with security properties.
+
+We agree that the title is confusing and should be changed. It refers to a concept in 
+theoretical security: a formal notion of security property as a mathematical description of
+the possible behaviors of a system. In that context, it was not actually obvious that stack
+safety could be defined in these terms, but now we have done so. That said, the title should
+not require readers to understand this specific terminology to be comprehensible.

### 2021-06-16 submission2/response_formal.txt
Polishing on first half of response

@@
-** Overview **
+# Overview
 
-We thank the reviewers for their comments. The reviews identify 
-some areas that the paper fails to explain adequately.
-None of these issues are in the formalism itself, but meta questions: How does
-the work improve on existing formalizations of stack safety? Do we have a plan to validate the model further?
-Can our work be extended to cover more complex uses of the stack? We have answers to these questions, and we can
-address them explicitly in the paper.
+We thank the reviewers for their comments.
 
-** Changes **
+The reviews identify some areas that the paper fails to explain
+adequately -- in particular, "framing" questions: How does the work improve
+on existing formalizations of stack safety? Do we have a plan to validate
+the model further?  Can our work be extended to cover more complex uses of
+the stack?
 
-1. We will give a more thorough discussion of validation, discuss the feasibility of proofs
-   and of randomized testing of more complex properties, and detail our roadmap for
-   further validation of different versions of the property.
+In the next section, we outline the changes we plan to make in the paper to
+clarify these questions.  We discuss the issues in more detail in the third
+section.
 
-2. We will give an explicit, direct comparison to the work of Skorstengaard et. al. Currently
-   we discuss it briefly in our related work section, but our novelty has not been
-   obvious, so we will be more thorough in explaining the differences.
+## Changes
 
-3. We will improve the presentation as suggested. We need to gradually
-   introduce the examples and associated intuition (including our
-   visual metaphors) as early as possible, starting in the
-   introduction.
+1. We will give a more thorough discussion of validation, discuss the
+   feasibility of proofs and of randomized testing of more complex
+   properties, and detail our roadmap for further validation of different
+   versions of the property.
 
-4. Our model avoids talking about language features other than the stack, but it is unclear at times
-   whether those features could be added without compromising stack safety. We will more clearly call out
-   common language features, such as heaps, that are compatible with stack safety.
+2. We will give a more explicit and direct comparison to the work of
+   Skorstengaard et. al. Currently we discuss it briefly in our related work
+   section, but the novelty of our work does not come through clearly
+   enough; we will explain the differences more thoroughly.
 
-[SNA: Do we want to promise to add callee-saves registers? I wasn't going to at first, but actually there's
-      a good story here. Something like:]
-5. We will add a discussion of how to extend the model to a callee-saves registers convention, which
-   demonstrates a concrete use for observational properties and a clear advance over the state-of-the-art.
+3. We will improve the presentation as suggested. In particular, we will
+   gradually introduce the examples and associated intuitions (including our
+   visual metaphors) as early as possible, starting in the introduction.
 
-** Detailed Comments **
+4. We will more clearly call out common language features, such as the heap,
+   that are compatible with our treatment of stack safety.
 
-* Reviewer A *
+5. [SNA: Do we want to promise to add callee-saves registers? I wasn't going to
+   at first, but actually there's a good story here. Something like:] [BCP: If
+   we think we can really do this, we should say so!]
+   We will add a discussion of how to extend the model to a callee-saves
+   registers convention, which demonstrates a concrete use for the observational
+   properties.
+   [BCP: Doesn't the current draft also demonstrate a use for the observational
+   properties?]
 
-We will extend the paper with a more detailed discussion of our roadmap for verification and
-testing, and the issues that we expect to arise. Here is a rough
-outline of how it would cover your concerns.
+# Detailed Comments
+
+## Reviewer A
+
+We will extend the paper with a more detailed discussion of our roadmap for
+verification and testing and the issues that we expect to arise. Here is a
+rough outline of how it would cover the concerns raised in the review.
 
@@
 
-These properties would absolutely be amenable to a correctness proof. Such a
-proof would be a lot of work in a setting with a full ISA, but we see no
-fundamental barriers.  We chose to concentrate on testing because it
-is better aligned with our immediate goal of evaluating enforcement
-mechanisms to uncover possible bugs -- it is lower effort, but if bugs are
-found, still high reward.
+The formal security properties we define would absolutely be amenable to a
+correctness proof. Such a proof would be considerable work to carry out for
+a full ISA like RISC-V, but we see no fundamental barriers.  We chose to
+concentrate on testing because it is better aligned with our immediate goal
+of evaluating real enforcement mechanisms for real machine architectures.
 
@@
 
-At present we are working to implement
-testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
-Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
-normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
-the deadline.
+We are working to implement testing of observable properties and to apply
+this to the putatively fixed version of the lazy tagging/clearing
+micro-policy.  At the moment, testing of observational integrity is
+implemented; observational confidentiality testing is not, but -- since the
+lazy policy enforces normal, stepwise confidentiality -- this is actually
+sufficient to test it. [APT: Do we have the results of these tests?][BCP: +1
+-- if we could actually say something about results it would be fantastic!]
+But we did not have this ready before the deadline. [BCP: IIRC, we have not
+been forbidden from mentioning new results that we would include in the
+final paper if accepted, and IMO there is little risk of annoying the
+reviewers if we mention them in response to direct criticisms or questions.]
 
-Our other immediate priority is to implement testing of a secondary enforcement mechanism. We will evaluate a Cheri
-implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
-enforcement. We are currently working on the QuickChick generators for Cheri programs, which have a
-different calling convention -- a more complicated one, which is easier to get wrong.
+Our other short-term plan is to implement testing of a second enforcement
+mechanism -- a Cheri implementation of stack safety -- to demonstrate that
+our properties are not strictly tied to tag-based enforcement. We are
+currently working on the QuickChick generators for Cheri programs, which
+have a different calling convention (a more complicated one, which is easier
+to get wrong!).  [BCP: We should probably add something like: "But this
+effort falls outside the scope of the current paper" to make clear that we
+are not promising it.]
 
-Validation of the argument-passing and coroutining models are longer-term priorities.
-The passing system has existing enforcement mechanisms and serves an important role in common stack
-use-cases. Coroutines will require the development of an enforcement
-mechanism (e.g., one based on tags) to validate.
+Validation of the argument-passing and coroutine models are longer-term
+projects.  The passing system has existing enforcement mechanisms and serves
+an important role in common stack use-cases. Coroutines will require the
+development of an enforcement mechanism (e.g., one based on tags) to
+validate.
 
-> There is one topic beyond my evaluation above worth considering: callee-save registers.
-> I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save
-> registers hold data private to the caller that may be passed arbitrarily deeper into the call stack,
-> then saved into arbitrary stack locations, but then must be restored before return.
-> Have you or the closely related prior work on enforcement mechanisms considered this complexity?
-> The invariants are clear enough informally, but the issue seems to add substantial difficulty to
-> tracking that a function faithfully implements the calling convention.
+> There is one topic beyond my evaluation above worth considering:
+> callee-save registers.  I fear these may be fairly tricky to support in
+> your set-up.  Upon function entry, callee-save registers hold data private
+> to the caller that may be passed arbitrarily deeper into the call stack,
+> then saved into arbitrary stack locations, but then must be restored
+> before return.  Have you or the closely related prior work on enforcement
+> mechanisms considered this complexity?  The invariants are clear enough
+> informally, but the issue seems to add substantial difficulty to tracking
+> that a function faithfully implements the calling convention.
 
-Callee-save registers are compatible with our approach, using a version of the sharing model
-with its context initialization and updates modified to apply appropriate domains to registers.
-They actually provide another use-case for observational properties! Focusing on integrity:
-callee-save registers may be overwritten (violating stepwise integrity) but should always be
-restored before a return. This requirement corresponds to observational integrity with the
-callee-save registers sealed -- recall that observational integrity compares sealed
-components at the start and end of a callee, so the interim changes are not violations.
+Callee-save registers are compatible with our approach, using a version of
+the sharing model with its context initialization and updates modified to
+apply appropriate domains to registers.  Indeed, they actually provide
+another interesting use-case for the "observational" variants of our
+properties. Focusing on integrity: callee-save registers may be overwritten
+(violating stepwise integrity) but should always be restored before a
+return. This requirement corresponds to observational integrity with the
+callee-save registers sealed. (Recall that observational integrity compares
+sealed components at the start and end of a callee's execution, so the
+interim changes are not violations.)
 
-Confidentiality can be applied similarly. A callee-save register should be secret, but
-in observational confidentiality, if it is copied to the stack and then restored before
-return, the property will allow the original register to be read normally after return.
-But the property will not allow it to be leaked in any other way.
+Confidentiality for callee-saves is similar. A callee-save register should
+be secret, but in observational confidentiality, if it is copied to the
+stack and then restored before return, the property will allow the original
+register to be read normally after return.  But the property will not allow
+it to be leaked in any other way.
 
-We had briefly considered adding this complexity and chosen to avoid it, because it is an extra layer
-of detail to apply domains to registers. It is worth noting that the primary existing work in this
-area, Skorstengaard et al., relies on a calling convention which clears all non-argument registers before
-calls and all non-return registers before returns.
+We had considered adding this complexity and chosen to avoid it, because it
+is an extra layer of detail to apply domains to registers.  [BCP: Don't
+understand this.  Maybe just drop it and say we will add this discussion to
+the paper?  Or do we even want to say that we think we can extend the model
+and our testing framework to callee-saves before the camera-ready?... My
+sense is that this is too much work, but either way we should say (again)
+what we plan to do.]
 
-> Another minor question: The difference between [nested] call-by-reference and passing
-> address-of-local was interesting, but I'm not sure how much it makes sense to argue the
-> latter is "beyond the scope of stack safety per se" [line 939] without arguing the same
-> of the former.  Perhaps the argument can be made that address-of-local can be put in the
-> heap while a by-reference parameter cannot and since you don't guarantee heap safety,
+It is worth noting that the main prior work in this area by Skorstengaard et
+al. relies [BCP: fundamentally?] on a calling convention which clears all
+non-argument registers before calls and all non-return registers before
+returns.
+
+> Another minor question: The difference between [nested] call-by-reference
+> and passing address-of-local was interesting, but I'm not sure how much it
+> makes sense to argue the latter is "beyond the scope of stack safety per
+> se" [line 939] without arguing the same of the former.  Perhaps the
+> argument can be made that address-of-local can be put in the heap while a
+> by-reference parameter cannot and since you don't guarantee heap safety,
 > you shouldn't try to do anything useful with address-of-local?

### 2021-06-16 submission2/response_formal.txt
More polishing

@@
 
-Yes, this was exactly our thinking.
-We know what constitutes legal access to stack-allocated arguments, but for explicit pointers
-this is much less clear, and depends on the intent of the programmer. We believe
-an address-of-local should be treated similarly to the address of a heap object,
-but with a different lifetime. For instance, if a heap safety model treats heap pointers like capabilities
-(possession implies access rights), we would treat stack pointers the same.
-Similarly, if the heap model permits pointer forging (restricted to
-within compartments, say) we would expect to be able to forge pointers to address-taken locals
-in the same circumstances.
+Yes, this was exactly our thinking.  We know what constitutes legal access
+to stack-allocated arguments, but for explicit pointers this is much less
+clear, and depends on the intent of the programmer. We believe an
+address-of-local should be treated similarly to the address of a heap
+object, but with a different lifetime. For instance, if a heap safety model
+treats heap pointers like capabilities (possession implies access rights),
+we would treat stack pointers the same.  Similarly, if the heap model
+permits pointer forging (restricted to within compartments, say) we would
+expect to be able to forge pointers to address-taken locals in the same
+circumstances.
 
-In the longer term we would like to extend our model with a heap, in the style of
-Azevedo de Amorim et al.'s "The Meaning of Memory Safety," which is phrased in similar
-terms to our model and should be compatible. Then the interplay between heap-derived and
-stack-derived pointers (with similar behaviors but different lifespans) will
-be much more interesting.
+In the longer term we would like to extend our model with a heap, in the
+style of Azevedo de Amorim et al.'s "The Meaning of Memory Safety," which is
+phrased in similar terms to our model and should be compatible. In this
+setting the interplay between heap-derived and stack-derived pointers (with
+similar behaviors but different lifespans) will be much more interesting.
 
-* Reviewer B *
+## Reviewer B
 
-> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M
+> Q1. How does this work improve upon Skorstengaard et al? Neither the
+> introduction nor related work section give a satisfying answer to this.
 
 We improve upon Skorstengaard et. al. in the following ways.
-1. We capture a desireable security characteristic of dynamically nested calls, that every caller is safe from its callee
+
+1. We capture a desireable security characteristic of dynamically nested
+   calls, that every caller is safe from its callee.
+
 2. We therefore do not assume a static division of trusted and untrusted code
 
-Specifically, Skorstengaard et. al. divide the world into trusted functions, which feature
-additional dynamic constraints on their behavior, and for which their properties hold,
-and adversaries, which may violate the dynamic "reasonableness" constraints. Suppose we
-have a single function that recursively traverses a linked list, loading data into local
-memory. What if it has a bug, and certain data cause it to violate the reasonableness condition?
-In Skorstengaard's model, it must be an adversary, and activations lower in the stack are not
-guaranteed protection. In our model, only the badly behaved activation is an attacker, and its
-caller(s) are guaranteed protection.
+   Specifically, Skorstengaard et. al. divide the world into trusted
+   functions, which feature additional dynamic constraints on their
+   behavior, and for which their properties hold, and adversaries, which may
+   violate the dynamic "reasonableness" constraints. Suppose we have a
+   single function that recursively traverses a linked list, loading data
+   into local memory. What if it has a bug, and certain data cause it to
+   violate the reasonableness condition?  In Skorstengaard's model, it must
+   be an adversary, and activations lower in the stack are not guaranteed
+   protection. In our model, only the badly behaved activation is an
+   attacker, and its caller(s) are guaranteed protection.
 
-3. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
-machine being architecturally similar to the concrete one (beyond the specific call and return semantics
-that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
-abstract machine. Our properties should apply directly to a wide range of
-architectures.
+3. Skorstengaard's approach to proving safety via fully-abstract overlay
+   semantics relies on the abstract machine being architecturally similar to
+   the concrete one (beyond the specific call and return semantics that
+   distinguish them). Therefore it is less portable: each architecture it
+   applies to wants its own abstract machine. Our properties should apply
+   directly to a wide range of architectures.
 
-We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
-to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
-safety is expressible as a formal security property.
+   We also differ from Skorstengaard et. al. in our focus. We use the
+   language of formal security properties to apply the established
+   intuitions and formal techniques of non-interference. This demonstrates
+   that stack safety is expressible as a formal security property.
 
-> Q2.
-In L117-121 you state certain assumptions about the language semantics.
-It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply.
-For example, does a language with a heap allow you to break stack safety?
-What about a language with delimited continuations that manipulate the stack?^M
+> Q2. In L117-121 you state certain assumptions about the language
+> semantics. It is unclear to me if these assumptions restrict the languages
+> to which your stack safety definition apply. For example, does a language
+> with a heap allow you to break stack safety? What about a language with
+> delimited continuations that manipulate the stack?^M
 
@@
 
-* Reviewer C *
+## Reviewer C
 
@@
 
-A formal definition of a type of security is valuable in its own right. It tells us whether
-any given behavior is a violation, and we can therefore use it to identify bugs. In our case
-study we identify a previously unknown bug in Roessler and Dehon's proposed "lazy tagging" micro-policy,
-and also use randomized testing to increase confidence that their "depth isolation" micro-policy
-correctly implements stack safety.
+First, we hope (and believe) that a formal definition of stack safety is
+valuable in its own right, as a way of shifting the discussion away from
+specific examples of bad behavior that must be prevented and toward general
+mechanisms that ensure good behavior.  A formal definition gives us the
+ability to reason clearly about what is gained from applying a given
+enforcement mechanism to a system. Since many enforcement mechanisms are
+quite costly in speed, dollars, and/or silicon, it is important to
+understand precisely what they buy us.
 
-More broadly, it gives us the ability to reason
-clearly about what is gained from applying a given enforcement mechanism to a system. Since
-many enforcement mechanisms are quite costly in speed, dollars, and silicon, it is important
-to understand precisely what they buy us.
+More immediately, we can use a formal definition like this to test proposed
+enforecement mechanisms for violations of the policy.  In our case study, we
+identify a previously unknown bug in Roessler and Dehon's proposed "lazy
+tagging" micro-policy, and we use randomized testing to increase confidence
+that their "depth isolation" micro-policy correctly implements stack safety.
 
@@
 
-We plan to combine our stack model with a heap model in the style of Acevedo de Amorim et al.
-as future work.
+Agreed!  We plan to combine our stack model with a heap model in the style
+of Acevedo de Amorim et al. as future work.  Both models are phrased in the
+conceptual terminology of language-based security, so we have reason to
+believe that they will fit together cleanly.
 
@@
 
-We agree that the title is confusing and should be changed. It refers to a concept in
-theoretical security: a formal notion of security property as a mathematical description of
-the possible behaviors of a system. In that context, it was not actually obvious that stack
-safety could be defined in these terms, but now we have done so. That said, the title should
-not require readers to understand this specific terminology to be comprehensible.
+In hindsight, our choice of title was confusing: We did not just mean that
+stack safety impacts security (which is obvious), but that it can usefully
+be formulated using (a refinement of) the same formal terminology and
+mathematical definitions -- confidentiality, integrity, noninterference,
+etc. -- that have been used for decades to reason about security.
+
+We'll think of a more helpful title.
+[BCP: Any ideas?]

### 2021-06-16 submission2/response_formal.txt
Finish polishing pass -- lots of comments

@@
 1. We capture a desireable security characteristic of dynamically nested
-   calls, that every caller is safe from its callee.
+   calls: that EVERY caller should be protected from its callee.
 
-2. We therefore do not assume a static division of trusted and untrusted code
+   By contrast, Skorstengaard et. al. statically divide the world into
+   trusted functions, which feature additional dynamic constraints on their
+   behavior and for which their properties hold, and adversary functions,
+   which may violate the dynamic "reasonableness" constraints.
 
-   Specifically, Skorstengaard et. al. divide the world into trusted
-   functions, which feature additional dynamic constraints on their
-   behavior, and for which their properties hold, and adversaries, which may
-   violate the dynamic "reasonableness" constraints. Suppose we have a
-   single function that recursively traverses a linked list, loading data
-   into local memory. What if it has a bug, and certain data cause it to
-   violate the reasonableness condition?  In Skorstengaard's model, it must
-   be an adversary, and activations lower in the stack are not guaranteed
-   protection. In our model, only the badly behaved activation is an
-   attacker, and its caller(s) are guaranteed protection.
+   For example, suppose we have a single function that recursively traverses
+   a linked list, loading data into local memory. What if it has a bug, and
+   certain data cause it to violate the reasonableness condition?  In
+   Skorstengaard's model, it must be an adversary, and activations lower in
+   the stack are not guaranteed protection. In our model, only the badly
+   behaved activation is an attacker, and its caller(s) are guaranteed
+   protection.  [BCP: Not clear enough -- Indeed, I think the example might
+   be too subtle.]
 
-3. Skorstengaard's approach to proving safety via fully-abstract overlay
+2. Skorstengaard's approach to proving safety via fully abstract overlay
    semantics relies on the abstract machine being architecturally similar to
    the concrete one (beyond the specific call and return semantics that
-   distinguish them). Therefore it is less portable: each architecture it
+   distinguish them). This makes it less portable: each architecture it
    applies to wants its own abstract machine. Our properties should apply
@@
    We also differ from Skorstengaard et. al. in our focus. We use the
-   language of formal security properties to apply the established
-   intuitions and formal techniques of non-interference. This demonstrates
-   that stack safety is expressible as a formal security property.
+   language of formal security properties to apply established intuitions
+   and formal techniques of non-interference.
 
@@
 > with a heap allow you to break stack safety? What about a language with
-> delimited continuations that manipulate the stack?^M
+> delimited continuations that manipulate the stack?
 
 Our stack safety model applies regardless of how a language uses non-stack,
-non-code memory, so it works for languages with a heap. It applies to any layout of memory as long as that is
-known statically.
+non-code memory, so it works for languages with a heap.  It also applies to
+any layout of memory (that is known statically).
 
-Delimited continuations are not supported by our base model, so a language that has them would violate
-stack safety. We extend the base model with argument passing on the stack to demonstrate how it can
-accommodate a programming language feature that shares stack data between callers and callees. A
+Delimited continuations are not supported by our base model, so a language
+that has them would violate stack safety. We do extend the base model with
+argument passing on the stack to demonstrate how it can accommodate a
+language feature that shares stack data between callers and callees, but a
 similar extension to delimited continuations would be more complex.
 
-This distinction, between a feature that could be added without changing the properties and one that
-would require enhanced properties, is important, and we will clarify in section 2 which features fall
-into which category.
+This distinction -- between features that could be added without changing
+the properties and ones that would require enhanced properties -- is
+important, and we will clarify in section 2 which features fall into which
+category.  [BCP: Can we say anything here about what that categorization
+would look like?]
 
-> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?
+> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et
+> al (L289-291)? What are the implications of that?
 
-We want our properties to be applicable even to low-level code that may not keep function
-footprints isolated from one another. Consider a program that mixes C code with hand-written
-assembly -- the assembly may make calls and returns, which should be protected by stack safety,
-as well as arbitrary jumps, which should not.
+We want our properties to be applicable even to low-level code that may not
+keep function footprints isolated from one another. [BCP: Not clear what
+footprints have to do with well-bracketing...] For example, consider a
+program that mixes C code with hand-written assembly: the assembly may make
+calls and returns, which should be protected by stack safety, as well as
+arbitrary jumps, which should not.
 
@@
 
-First, we hope (and believe) that a formal definition of stack safety is
-valuable in its own right, as a way of shifting the discussion away from
-specific examples of bad behavior that must be prevented and toward general
+First, we contend that a formal definition of stack safety is valuable in
+its own right, as a way of shifting the discussion away from specific
+examples of bad behavior that must be prevented and toward general
 mechanisms that ensure good behavior.  A formal definition gives us the

### 2021-06-16 submission2/response_formal.txt
Some comments

@@
    section, but the novelty of our work does not come through clearly
-   enough; we will explain the differences more thoroughly.
+   enough; we will explain the differences more thoroughly. [LEO: I think
+   we need a forward pointer here to make sure reviewers B+C see this.]
 
@@
    [BCP: Doesn't the current draft also demonstrate a use for the observational
-   properties?]
+   properties?][LEO: I think the current motivation is efficiency in the
+   paper - this would tie things together beautifully IMO - and we can say
+   we left it out due to space/complexity concerns but that in retrospect
+   putting it in paints a more complete picture.]
 
@@
 our properties are not strictly tied to tag-based enforcement. We are
-currently working on the QuickChick generators for Cheri programs, which
+currently working on QuickChick generators for Cheri programs, which
 have a different calling convention (a more complicated one, which is easier
@@
 effort falls outside the scope of the current paper" to make clear that we
-are not promising it.]
+are not promising it.][LEO: Agreed.]
 
@@
 sense is that this is too much work, but either way we should say (again)
-what we plan to do.]
+what we plan to do.][LEO: The point I guess is that now we're only applying
+domains on stack addresses, while this would require domains over arbitrary
+components (including registers). I don't think it's particularly hard to
+do, just an extra layer of complexity.]
 
@@
    protection.  [BCP: Not clear enough -- Indeed, I think the example might
-   be too subtle.]
+   be too subtle.] [LEO: "We will expand on such an example in the revised
+   version"?]
 
@@
 calls and returns, which should be protected by stack safety, as well as
-arbitrary jumps, which should not.
+arbitrary jumps, which should not. [LEO: I also don't fully follow this
+answer]
 
@@
 We'll think of a more helpful title.
-[BCP: Any ideas?]
+[BCP: Any ideas?][LEO: s/is/as]

### 2021-06-16 submission2/response_formal.txt
Couple of comments

@@
    version"?]
+   [RB: This part also read confusingly to me. What needs to be made clear, I
+   think, is that in the world of StkTokens one needs to establish a boundary
+   separating trusted and untrusted components, but the reasonability conditions
+   on which trusted components rely is dynamic -- it might be approximated
+   statically, but we could shoot ourselves in the foot in a way the model can
+   do nothing about.]
 
@@
 Agreed!  We plan to combine our stack model with a heap model in the style
-of Acevedo de Amorim et al. as future work.  Both models are phrased in the
+of Azevedo de Amorim et al. as future work.  Both models are phrased in the
 conceptual terminology of language-based security, so we have reason to

### 2021-06-16 submission2/response_formal.txt
wibbles

@@
 Validation of the argument-passing and coroutine models are longer-term
-projects.  The passing system has existing enforcement mechanisms and serves
-an important role in common stack use-cases. Coroutines will require the
+projects.  The argument-passing model has existing enforcement mechanisms and serves
+an important role in common stack use cases. Coroutines will require the
 development of an enforcement mechanism (e.g., one based on tags) to
@@
 apply appropriate domains to registers.  Indeed, they actually provide
-another interesting use-case for the "observational" variants of our
+another interesting use case for the "observational" variants of our
 properties. Focusing on integrity: callee-save registers may be overwritten
@@
 It is worth noting that the main prior work in this area by Skorstengaard et
-al. relies [BCP: fundamentally?] on a calling convention which clears all
+al. relies [BCP: fundamentally?] on a calling convention that clears all
 non-argument registers before calls and all non-return registers before
@@
 In the longer term we would like to extend our model with a heap, in the
-style of Azevedo de Amorim et al.'s "The Meaning of Memory Safety," which is
-phrased in similar terms to our model and should be compatible. In this
+style of Azevedo de Amorim et al.'s "The Meaning of Memory Safety," POST 2018,
+which is phrased in similar terms to our model and should be compatible. In this
 setting the interplay between heap-derived and stack-derived pointers (with

### 2021-06-16 submission2/response_formal.txt
comments, polishing

@@
    we left it out due to space/complexity concerns but that in retrospect
-   putting it in paints a more complete picture.]
+   putting it in paints a more complete picture.][APT: Agreed, but I'm worried
+   about over-promising. However, see my note in the callee-save discussion
+   below: I'm not clear just how much we need to add.]
 
@@
 
-Callee-save registers are compatible with our approach, using a version of
-the sharing model with its context initialization and updates modified to
-apply appropriate domains to registers.  Indeed, they actually provide
-another interesting use case for the "observational" variants of our
-properties. Focusing on integrity: callee-save registers may be overwritten
-(violating stepwise integrity) but should always be restored before a
-return. This requirement corresponds to observational integrity with the
-callee-save registers sealed. (Recall that observational integrity compares
-sealed components at the start and end of a callee's execution, so the
-interim changes are not violations.)
+Callee-save registers are compatible with our approach, using a
+version of the sharing model with its context initialization and
+updates modified to apply appropriate domains to registers as well as
+stack locations.  Indeed, they actually provide another interesting
+use case for the "observational" variants of our properties. Focusing
+on integrity: callee-save registers may be overwritten (violating
+stepwise integrity) but should always be restored before a
+return. This requirement corresponds to observational integrity with
+the callee-save registers sealed. (Recall that observational integrity
+compares sealed components at the start and end of a callee's
+execution, so the interim changes are not violations.)
 
@@
 do, just an extra layer of complexity.]
+[APT: I'm a little confused here. Formally, doesn't the
+paper already state everything in terms of components, which include registers
+as well as stack locations? Does anything need to be added except a tweak to
+the sealing convention (which, by the way, seems to be a little underspecified anyhow) ?]
 
@@
 
-In the longer term we would like to extend our model with a heap, in the
-style of Azevedo de Amorim et al.'s "The Meaning of Memory Safety," POST 2018,
-which is phrased in similar terms to our model and should be compatible. In this
-setting the interplay between heap-derived and stack-derived pointers (with
-similar behaviors but different lifespans) will be much more interesting.
+In the longer term we would like to extend our model with a heap, in
+the style of Azevedo de Amorim et al. [1] which is phrased in similar
+terms to our model and should be compatible. In this setting the
+interplay between heap-derived and stack-derived pointers (with
+similar behaviors but different lifespans) will be much more
+interesting.
 
@@
 1. We capture a desireable security characteristic of dynamically nested
-   calls: that EVERY caller should be protected from its callee.
+   calls: that EVERY caller should be protected from its callees.
 
@@
    protection.  [BCP: Not clear enough -- Indeed, I think the example might
-   be too subtle.] [LEO: "We will expand on such an example in the revised
+   be too subtle.] [APT:+ 1] [LEO: "We will expand on such an example in the revised
    version"?]
@@
    distinguish them). This makes it less portable: each architecture it
-   applies to wants its own abstract machine. Our properties should apply
+   applies to needs its own abstract machine. Our properties should apply
    directly to a wide range of architectures.
@@
    and formal techniques of non-interference.
+   [APT: Can we distinguish the fact that we define things in terms of traces of behavior
+   rather than in terms of "correct by construction" machines?]
 
@@
 non-code memory, so it works for languages with a heap.  It also applies to
-any layout of memory (that is known statically).
+any (statically known) memory layout.
 
@@
 arbitrary jumps, which should not. [LEO: I also don't fully follow this
-answer]
+answer][APT: +1. Can we sketch a concrete example -- maybe a variant of
+one in the paper -- that we allow by SK does not?]
 
@@
 Agreed!  We plan to combine our stack model with a heap model in the style
-of Azevedo de Amorim et al. as future work.  Both models are phrased in the
+of Azevedo de Amorim et al. [1] as future work.  Both models are phrased in the
 conceptual terminology of language-based security, so we have reason to
@@
 We'll think of a more helpful title.
-[BCP: Any ideas?][LEO: s/is/as]
+[BCP: Any ideas?][LEO: s/is/as][APT: "Formalizing Stack Safety as a Security Property"]
+
+----------------
+
+[1] A. Azevedo de Amorim, C. Hritcu, and B. Pierce, "The Meaning of Memory Safety,"
+Proc. Principles of Security and Trust, 2018.

### 2021-06-16 submission2/response_formal.txt
Responding to things.

@@
 
-5. [SNA: Do we want to promise to add callee-saves registers? I wasn't going to
-   at first, but actually there's a good story here. Something like:] [BCP: If
-   we think we can really do this, we should say so!]
-   We will add a discussion of how to extend the model to a callee-saves
-   registers convention, which demonstrates a concrete use for the observational
-   properties.
-   [BCP: Doesn't the current draft also demonstrate a use for the observational
-   properties?][LEO: I think the current motivation is efficiency in the
-   paper - this would tie things together beautifully IMO - and we can say
-   we left it out due to space/complexity concerns but that in retrospect
-   putting it in paints a more complete picture.][APT: Agreed, but I'm worried
-   about over-promising. However, see my note in the callee-save discussion
-   below: I'm not clear just how much we need to add.]
+5. We will add a discussion of how to extend the model to a callee-saves
+   registers convention, which demonstrates a second concrete use for 
+   the observational properties. Previously we were concerned about space
+   and explanatory complexity, but in retrospect its presence would
+   paint a more complete picture.
 
@@
 -- if we could actually say something about results it would be fantastic!]
+[SNA: We are fighting with Quickchick a bit, trying to get these.]
 But we did not have this ready before the deadline. [BCP: IIRC, we have not
@@
 have a different calling convention (a more complicated one, which is easier
-to get wrong!).  [BCP: We should probably add something like: "But this
-effort falls outside the scope of the current paper" to make clear that we
-are not promising it.][LEO: Agreed.]
+to get wrong!). This work is for a future paper, but its plausibility is
+relevant to the discussion here.
 
@@
 
-We had considered adding this complexity and chosen to avoid it, because it
-is an extra layer of detail to apply domains to registers.  [BCP: Don't
-understand this.  Maybe just drop it and say we will add this discussion to
-the paper?  Or do we even want to say that we think we can extend the model
-and our testing framework to callee-saves before the camera-ready?... My
-sense is that this is too much work, but either way we should say (again)
-what we plan to do.][LEO: The point I guess is that now we're only applying
-domains on stack addresses, while this would require domains over arbitrary
-components (including registers). I don't think it's particularly hard to
-do, just an extra layer of complexity.]
+We will add this discussion to the paper, where previously we had ommitted
+it due to concerns for space and for the complexity of the explanations.
+But it both demonstrates the model's flexibility and ties in neatly to
+observational properties, so it is worth including.
+
 [APT: I'm a little confused here. Formally, doesn't the
@@
 the sealing convention (which, by the way, seems to be a little underspecified anyhow) ?]
+[SNA: initCtx needs to put the registers in Unsealed instead of Outside,
+as well.]
 
 It is worth noting that the main prior work in this area by Skorstengaard et
-al. relies [BCP: fundamentally?] on a calling convention that clears all
+al. relies fundamentally on a calling convention that clears all
 non-argument registers before calls and all non-return registers before
@@
    language of formal security properties to apply established intuitions
-   and formal techniques of non-interference.
-   [APT: Can we distinguish the fact that we define things in terms of traces of behavior
-   rather than in terms of "correct by construction" machines?]
+   and formal techniques of non-interference. This entails defining
+   our properties in terms of traces of the machine's behavior, rather
+   than in terms of a machine that is "correct by construction."
 
@@
 
-We want our properties to be applicable even to low-level code that may not
-keep function footprints isolated from one another. [BCP: Not clear what
-footprints have to do with well-bracketing...] For example, consider a
-program that mixes C code with hand-written assembly: the assembly may make
-calls and returns, which should be protected by stack safety, as well as
-arbitrary jumps, which should not. [LEO: I also don't fully follow this
-answer][APT: +1. Can we sketch a concrete example -- maybe a variant of
+Skorstengaard et al. don't assume well-bracketed control flow, but the
+terms in which they define differ from ours. Their version assumes
+a code layout where each function has an isolated footprint and their
+property doesn't permit unrestricted jumps between footprints. In that
+model it is always obvious when a function returns.
+
+We want our properties to be applicable even to low-level code 
+that may not keep function footprints isolated from one another, so our
+definition must recognize the returns...
+[APT: +1. Can we sketch a concrete example -- maybe a variant of
 one in the paper -- that we allow by SK does not?]
+[SNA: Still figuring out how to explain this... An example would be
+good, yes.]
 
@@
 
-We'll think of a more helpful title.
-[BCP: Any ideas?][LEO: s/is/as][APT: "Formalizing Stack Safety as a Security Property"]
+We will change the title to:
+"Formalizing Stack Safety as a Security Property"
 

### 2021-06-17 submission2/response_formal.txt
clear some comments; a little polishing

@@
    the observational properties. Previously we were concerned about space
-   and explanatory complexity, but in retrospect its presence would
-   paint a more complete picture.
+   and explanatory complexity, but in retrospect we see that discussing
+   register protection would paint a more complete picture.
 
@@
 
-The formal security properties we define would absolutely be amenable to a
-correctness proof. Such a proof would be considerable work to carry out for
-a full ISA like RISC-V, but we see no fundamental barriers.  We chose to
-concentrate on testing because it is better aligned with our immediate goal
-of evaluating real enforcement mechanisms for real machine architectures.
+Correctness of an enforcement mechanism with respect to our properties
+is certainly amenable to formal proof. Such a proof would be
+considerable work to carry out for a full ISA like RISC-V, but we see
+no fundamental barriers.  We chose to concentrate on testing because
+it is better aligned with our immediate goal of evaluating real
+enforcement mechanisms for real machine architectures.
 
@@
 have a different calling convention (a more complicated one, which is easier
-to get wrong!). This work is for a future paper, but its plausibility is
-relevant to the discussion here.
+to get wrong!). This effort is beyond the scope of this paper, but we will
+sketch it as future work.
 
@@
 
-Callee-save registers are compatible with our approach, using a
+Callee-save registers are supported by our approach, using a
 version of the sharing model with its context initialization and
@@
 
-We will add this discussion to the paper, where previously we had ommitted
-it due to concerns for space and for the complexity of the explanations.
-But it both demonstrates the model's flexibility and ties in neatly to
-observational properties, so it is worth including.
-
-[APT: I'm a little confused here. Formally, doesn't the
-paper already state everything in terms of components, which include registers
-as well as stack locations? Does anything need to be added except a tweak to
-the sealing convention (which, by the way, seems to be a little underspecified anyhow) ?]
-[SNA: initCtx needs to put the registers in Unsealed instead of Outside,
-as well.]
-
 It is worth noting that the main prior work in this area by Skorstengaard et
@@
 returns.
+
+We will add this discussion to the paper, where previously we had omitted
+it due to concerns for space and for the complexity of the explanations.
+But it both demonstrates the model's flexibility and ties in neatly to
+observational properties, so it is worth including.
 
@@
 
-We will change the title to:
+We plan to change our title to:
 "Formalizing Stack Safety as a Security Property"

### 2021-06-17 submission2/response_formal.txt
Tried for a simpler contrast with StkTokens

--- 2021-06-16 submission2/response_formal.txt
@@
    By contrast, Skorstengaard et. al. statically divide the world into
-   trusted functions, which feature additional dynamic constraints on their
-   behavior and for which their properties hold, and adversary functions,
-   which may violate the dynamic "reasonableness" constraints.
+   trusted functions, which feature additional dynamic "reasonableness"
+   constraints on their behavior and for which their properties hold,
+   and adversary functions, which may violate those constraints but
+   are not guaranteed safety.
 
-   For example, suppose we have a single function that recursively traverses
-   a linked list, loading data into local memory. What if it has a bug, and
-   certain data cause it to violate the reasonableness condition?  In
-   Skorstengaard's model, it must be an adversary, and activations lower in
-   the stack are not guaranteed protection. In our model, only the badly
-   behaved activation is an attacker, and its caller(s) are guaranteed
-   protection.  [BCP: Not clear enough -- Indeed, I think the example might
-   be too subtle.] [APT:+ 1] [LEO: "We will expand on such an example in the revised
-   version"?]
-   [RB: This part also read confusingly to me. What needs to be made clear, I
-   think, is that in the world of StkTokens one needs to establish a boundary
-   separating trusted and untrusted components, but the reasonability conditions
-   on which trusted components rely is dynamic -- it might be approximated
-   statically, but we could shoot ourselves in the foot in a way the model can
-   do nothing about.]
+   The dynamic nature of this bounary makes it an issue with 
+   functions that cannot be statically proven safe, but whose data
+   still need to be protected when they are behaving safely.
+   For a contrived example, consider a function that stores some
+   private state, but takes an argument "unsafe" and, if called with
+   "unsafe" set to 1, violates the reasonableness constraints. In
+   the StkTokens model, we can never guarantee the safety of the
+   data even if called with "unsafe" set to 0.
+
+   We will expand on such an example in the revised version.
 

### 2021-06-17 submission2/response_formal.txt
more polishing

--- 2021-06-17 submission2/response_formal.txt
@@
 
-   We also differ from Skorstengaard et. al. in our focus. We use the
-   language of formal security properties to apply established intuitions
-   and formal techniques of non-interference. This entails defining
-   our properties in terms of traces of the machine's behavior, rather
-   than in terms of a machine that is "correct by construction."
+3. We also differ from Skorstengaard et. al. in the nature of our
+   safety definition.  We use the language of formal security
+   properties to apply well-established concepts and formal techniques
+   of non-interference. Thus, we define our properties in terms of
+   traces of memory and register states, rather than in terms of an
+   ad-hoc abstract machine that is asserted to be "stack safe by construction."
+   Which approach is better may be a matter of taste, but we believe our
+   alternative merits consideration.
 
@@
 a code layout where each function has an isolated footprint and their
-property doesn't permit unrestricted jumps between footprints. In that
+property prohibits unrestricted jumps between footprints. In that
 model it is always obvious when a function returns.

### 2021-06-17 submission2/response_formal.txt
typo

@@
 
-   The dynamic nature of this bounary makes it an issue with 
+   The dynamic nature of this boundary makes it an issue with 
    functions that cannot be statically proven safe, but whose data

### 2021-06-17 submission2/response_formal.txt
Some more updates

--- 2021-06-17 submission2/response_formal.txt
@@
    section, but the novelty of our work does not come through clearly
-   enough; we will explain the differences more thoroughly. [LEO: I think
+   enough; we will explain the differences more thoroughly, all in one place.
+ [LEO: I think
    we need a forward pointer here to make sure reviewers B+C see this.]
@@
 
-3. We also differ from Skorstengaard et. al. in the nature of our
-   safety definition.  We use the language of formal security
-   properties to apply well-established concepts and formal techniques
-   of non-interference. Thus, we define our properties in terms of
-   traces of memory and register states, rather than in terms of an
-   ad-hoc abstract machine that is asserted to be "stack safe by construction."
-   Which approach is better may be a matter of taste, but we believe our
-   alternative merits consideration.
+   We also differ from Skorstengaard et. al. in our focus. We use the
+   language of formal security properties to apply established intuitions
+   and formal techniques of non-interference. This entails defining
+   our properties in terms of traces of the machine's behavior, rather
+   than in terms of a machine that is "correct by construction."
 
@@
 important, and we will clarify in section 2 which features fall into which
-category.  [BCP: Can we say anything here about what that categorization
-would look like?]
+category.
+
+Some features that can be added/changed without changing the properties:
+- a heap, as stated
+- architectural choices like a frame pointer
+- the machine's ISA
+
+Features that cannot be supported with changes to the model:
+- runtime code generation (needs dynamic code annotations)
+- continuations (needs its own model in the style of the stack-passing one)
+- parallelism (needs fundamental changes to step, trace models)
 
@@
 
-Skorstengaard et al. don't assume well-bracketed control flow, but the
-terms in which they define differ from ours. Their version assumes
-a code layout where each function has an isolated footprint and their
-property prohibits unrestricted jumps between footprints. In that
-model it is always obvious when a function returns.
+Skorstengaard et al. don't directly assume well-bracketed control flow,
+they enforce it. But the terms in which they define it differ somewhat
+from ours. Their version assumes a code layout where each function has
+an isolated footprint and they don't permit unrestricted jumps between
+footprints. In that model it is always obvious when a function returns.
 
 We want our properties to be applicable even to low-level code 
-that may not keep function footprints isolated from one another, so our
-definition must recognize the returns...
+that may not keep function footprints isolated from one another.
+Suppose we have a piece of hand-written assembly code that sometimes
+makes explicit calls, and sometimes computed jumps...
 [APT: +1. Can we sketch a concrete example -- maybe a variant of

### 2021-06-17 submission2/response_formal.txt

--- 2021-06-17 submission2/response_formal.txt
@@
 
-Skorstengaard et al. don't assume well-bracketed control flow, but the
-terms in which they define differ from ours. Their version assumes
+korstengaard et al. don't assume well-bracketed control flow, but the
+terms in which they define it differ from ours. Their version assumes
 a code layout where each function has an isolated footprint and their
@@
 We want our properties to be applicable even to low-level code 
-that may not keep function footprints isolated from one another, so our
-definition must recognize the returns...
+that may not keep function footprints isolated from one another.
+Suppose we have a piece of hand-written assembly code that sometimes
+makes explicit calls, and sometimes computed jumps...
 [APT: +1. Can we sketch a concrete example -- maybe a variant of
 one in the paper -- that we allow by SK does not?]
-[SNA: Still figuring out how to explain this... An example would be
-good, yes.]
+[SNA: Still brainstorming on this.]
 

### 2021-06-17 submission2/response_formal.txt
Dropped an S
Fixed merge trouble

@@
 
-korstengaard et al. don't assume well-bracketed control flow, but the
-terms in which they define it differ from ours. Their version assumes
-a code layout where each function has an isolated footprint and their
-property prohibits unrestricted jumps between footprints. In that
+Skorstengaard et al. don't directly assume well-bracketed control flow,
+they enforce it. But the terms in which they define it differ
+from ours. Their version assumes
+a code layout where each function has an isolated footprint and they
+don't permit unrestricted jumps between footprints. In that
 model it is always obvious when a function returns.

### 2021-06-17 submission2/response_formal.txt
a few more little comments

@@
 - a heap, as stated
-- architectural choices like a frame pointer
+- calling convention choices, like whether or not to have a frame pointer
 - the machine's ISA
 
-Features that cannot be supported with changes to the model:
+Features that cannot be supported without changes to the model:
 - runtime code generation (needs dynamic code annotations)
-- continuations (needs its own model in the style of the stack-passing one)
+- continuations (needs its own model in the style of the stack-passing [APT:??]  one)
 - parallelism (needs fundamental changes to step, trace models)
+[APT: exceptions?]
+[APT: address-taken locals?-- seems a bit ambiguous]
 

### 2021-06-17 submission2/response_formal.txt
More things we don't support

@@
 - runtime code generation (needs dynamic code annotations)
-- continuations (needs its own model in the style of the stack-passing [APT:??]  one)
+- continuations (would be similar to argument passing)
+- exceptions (needs annotations of throw/catch points)
 - parallelism (needs fundamental changes to step, trace models)
-[APT: exceptions?]
-[APT: address-taken locals?-- seems a bit ambiguous]
+- address-taken locals (would be similar to argument passing, but
+  benefits from more sophisticated, heap-like memory models)
 

### 2021-06-17 submission2/response_formal.txt
Forward pointers, optional content for if testing works.

@@
    properties, and detail our roadmap for further validation of different
-   versions of the property.
+   versions of the property. For details, see the comments responding
+   to reviewer A.
 
@@
    enough; we will explain the differences more thoroughly, all in one place.
- [LEO: I think
-   we need a forward pointer here to make sure reviewers B+C see this.]
+   For details, see the comments responding to reviewer B.
 
@@
 4. We will more clearly call out common language features, such as the heap,
-   that are compatible with our treatment of stack safety.
+   that are compatible with our treatment of stack safety. For details,
+   see the comments responding to reviewer B.
 
@@
    register protection would paint a more complete picture.
+   For details, see the comments responding to reviewer A.
 
@@
 
-We are working to implement testing of observable properties and to apply
+[ We have since tested lazy tagging/clearing and found it to pass the
+  newly implemented observational integrity property as well as normal,
+  stepwise confidentiality. But we did not have this ready before the
+  deadline.
+| We are working to implement testing of observational properties and to apply
   this to the putatively fixed version of the lazy tagging/clearing
-micro-policy.  At the moment, testing of observational integrity is
-implemented; observational confidentiality testing is not, but -- since the
-lazy policy enforces normal, stepwise confidentiality -- this is actually
-sufficient to test it. [APT: Do we have the results of these tests?][BCP: +1
--- if we could actually say something about results it would be fantastic!]
-[SNA: We are fighting with Quickchick a bit, trying to get these.]
-But we did not have this ready before the deadline. [BCP: IIRC, we have not
-been forbidden from mentioning new results that we would include in the
-final paper if accepted, and IMO there is little risk of annoying the
-reviewers if we mention them in response to direct criticisms or questions.]
+  micro-policy. ]
 
@@
 a code layout where each function has an isolated footprint and they
-don't permit unrestricted jumps between footprints. In that
-model it is always obvious when a function returns.
-
-We want our properties to be applicable even to low-level code 
-that may not keep function footprints isolated from one another.
-Suppose we have a piece of hand-written assembly code that sometimes
-makes explicit calls, and sometimes computed jumps...
-[APT: +1. Can we sketch a concrete example -- maybe a variant of
-one in the paper -- that we allow by SK does not?]
-[SNA: Still brainstorming on this.]
+don't permit unrestricted jumps between footprints.
+Our definitions are sensible without that assumption.
 

### 2021-06-17 submission2/response_formal.txt
one last bit of rewording and reordering a few bullets

@@
    ad-hoc abstract machine that is asserted to be "stack safe by construction."
-   Which approach is better may be a matter of taste, but we believe our
-   alternative merits consideration.
+   Which approach is better may be a matter of taste, but we believe that our
+   alternative has the merit of explicitly relating stack safety to existing
+   notions of security.
 
@@
 Features that cannot be supported without changes to the model:
+- address-taken locals (would be similar to argument passing, but
+  would benefit from more sophisticated, heap-like memory models)
+- exceptions (needs annotations of throw/catch points)
+- continuations (would be similar to argument passing)
 - runtime code generation (needs dynamic code annotations)
-- continuations (would be similar to argument passing)
-- exceptions (needs annotations of throw/catch points)
 - parallelism (needs fundamental changes to step, trace models)
-- address-taken locals (would be similar to argument passing, but
-  benefits from more sophisticated, heap-like memory models)
 
@@
 
-Skorstengaard et al. don't directly assume well-bracketed control flow,
-they enforce it. But the terms in which they define it differ
-from ours. Their version assumes
-a code layout where each function has an isolated footprint and they
-don't permit unrestricted jumps between footprints.
-Our definitions are sensible without that assumption.
+Skorstengaard et al. don't directly assume well-bracketed control
+flow, they enforce it. But the terms in which they define it differ
+from ours. Their version assumes a code layout where each function has
+an isolated footprint and they don't permit unrestricted jumps between
+footprints.  Our definitions are sensible without that assumption.
 

### 2021-06-26 submission3/main.tex
Adding submission3

--- 2021-04-17 tex/main.tex
@@
 Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the concrete machine that instantiates it.  We assume a register
+assumptions about the concrete machine that instantiates it. Without changing the
+properties, we can support any register
 machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
-details of the machine's ISA are unimportant. We assume that we know ahead of time
-the layout of memory, including where code lives and which addresses are within
-the bounds of the stack. We do not model a heap; programs might choose to use memory
-that is neither code nor stack to store global variables, as a heap, etc., but our
-properties do not protect it or assume any particular behavior. We also do not
-model run-time code generation.
-
-% \footnote{\bcp{Some possibly useful text: A major challenge in
-%     characterizing stack safety formally is that, in a low-level machine
-%     without built-in stacks, we are at the mercy\bcp{?} of
-%     architectural\bcp{?}  choices that do not relate to the abstract notion
-%     of stack safety. Some implementations may have only a stack pointer
-%     while others add a frame pointer, in some systems return addresses and
-%     data live in separate stacks, and so on. We abstract over these choices
-%     as much as possible.  }}%
+details of the machine's ISA are unimportant. We can support any static memory,
+including a heap -- our properties define stack protection regardless of how
+non-stack memory is used. Likewise, we can support a range of calling convention
+choices, like whether or not to have a frame pointer.
+
+Some calling convention choices, such as callee-save registers and argument
+passing on the stack, require enhancements to our base properties. We present
+an enhancement that supports argument passing. Other plausible enhancements
+include:
+
+\begin{itemize}
+\item address-taken locals -- similar to argument passing, but
+      would benefit from more sophisticated, heap-like memory models
+\item exceptions -- needs annotations of throw/catch points
+\item continuations -- would be similar to argument passing
+\end{itemize}
+
+Runtime code generation would be much more challenging to support, as
+it requires dynamic code annotations. Parallelism would require
+fundamental changes to the system model. So, we consider these features
+fully out of scope.
 
@@
 
+We improve upon \citet{SkorstengaardSTK} in two ways.
+
+\begin{enumerate}
+\item We capture a desireable security characteristic of dynamically nested
+      calls: that {\em every} caller should be protected from its callees.
+      By contrast, they statically divide the world into
+      trusted functions, which feature additional dynamic "reasonableness"
+      constraints on their behavior and for which their properties hold,
+      and adversary functions, which may violate those constraints but
+      are not guaranteed safety.
+
+      The dynamic nature of this boundary makes it an issue with 
+      functions that cannot be statically proven safe, but whose data
+      still need to be protected when they are behaving safely.
+      For a contrived example, consider a function that stores some
+      private state, but takes an argument ``unsafe'' and, if called with
+      ``unsafe'' set to 1, violates the reasonableness constraints. In
+      the StkTokens model, we can never guarantee the safety of the
+      data even if called with "unsafe" set to 0.
+
+      \sna{TODO: expand}
+
+\item Their approach to proving safety via fully abstract overlay
+      semantics relies on the abstract machine being architecturally similar to
+      the concrete one (beyond the specific call and return semantics that
+      distinguish them). This makes it less portable: each architecture it
+      applies to needs its own abstract machine. Our properties should apply
+      directly to a wide range of architectures.
+\end{enumerate}
+
 \citet{SkorstengaardSTK} enumerate a useful list of criteria for judging

### 2021-06-28 submission3/main.tex
Little calle-save registers writing

@@
 
+\subsection{Callee-saves Registers}
+
+In some systems, rather than the caller saving the non-argument data from its registers
+and clearing them (and therefore implicitly taking responsibility for protecting those data),
+the callee is responsible for this. Nevertheless we may wish for those registers to be
+protected, just as they would be in a caller-save convention.
+
+Intuitively, a caller expects a callee-saved register to be unchanged on return -- but it
+may change during the call, as long as it is restored. We similarly expect a callee-saved
+register to be secret from the callee, except when it is copied to the stack! Both scenarios
+reflect not stepwise but observational stack safety.
+
+We begin by adding information to Call annotations defining which registers will be saved
+and which are arguments. The stack pointer and the register containing the return address
+should never be saved. We then extend the set of domains with ``saved'' and ``unsaved.''
+%
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \REGS)\]
+%
+\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \unsaved + \saved\]
+
 \subsection{Coroutines}

### 2021-06-28 submission3/main.tex
Callee-saves registers discussion

@@
 be the set of components \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\) or
+\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
 \(\domMap ~ \component = \passed{\depth,\depth'}\)
-where \(\depth \leq |\rts| \leq \depth'\).
+where \(|\rts| < \depth\) or \(|\rts| > \depth'\).
 
@@
 
-We begin by adding information to Call annotations defining which registers will be saved
-and which are arguments. The stack pointer and the register containing the return address
-should never be saved. We then extend the set of domains with ``saved'' and ``unsaved.''
-%
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \REGS)\]
+We begin by adding information to Call annotations defining which registers will be {\em saved}
+and which are {\em unsaved}. The stack pointer and the register containing the return address
+should never be saved. We then extend the set of domains with these values.
+%
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(S \in \REGS \rightarrow (\unsaved + \saved))\]
 %
 \[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \unsaved + \saved\]
+
+\judgmentthree[ Call]
+              {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
+%              {\(\components = \sealcon ~ \mach\)}
+              {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
+              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \in \REGS \mapsto S \reg]\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+The remaining rules are as in the simple model.
+
+\definition Let the protected set of a context, \(\prot_S(\domMap,\rts)\),
+be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+or \(\domMap ~ \component = \saved\).
+
+\definition Let the secret set of a context, written \(\secret_S(\domMap,\rts)\),
+be the set of components \(\component\) such that
+\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
+\(\domMap ~ \component = \saved\).
+
+\definition
+For each of the stepwise, deferred, and observational versions of
+stack integrity in the simple model, the equivalent property in the
+passing model
+
+{\bf Observational stack integrity} with callee-saves registers substitutes the above
+\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
+
+{\bf Observational stack confidentiality} additionally substitutes \(\secret_S\).
+
+\paragraph{Discussion}
+
+Observational properties are not only for lazy enforcement mechanisms -- they reflect a useful
+loosening of the stepwise properties. Consider a single register \(\reg\) that is saved.
+It is copied to some stack address \(\addr\), violating stepwise confidentiality, and then
+overwritten, violating stepwise integrity. Then before the return, it is restored to its
+original value. In observational integrity terms, \(\reg\) is unchanged between call and return,
+and in observational confidentiality terms, both the original and variant values are unchanged
+from their originals, so that \(\reg\) is not in the \(\delta\) set. But \(\addr\) is in the
+\(\delta\) set, so confidentiality additionally requires that the value of \(\reg\) does not
+leak through the callee-saves mechanism, for instance if the caller allocates but fails to
+initialize more memory.
 

### 2021-07-01 submission3/response.txt
Response/todo

--- 2021-06-17 submission2/response_formal.txt
@@
 
+  *Status* - not incorporated
+
 2. We will give a more explicit and direct comparison to the work of
@@
 
+  *Status* - incorporated, but needs example/more discussion
+
 3. We will improve the presentation as suggested. In particular, we will
@@
 
+  *Status* - not incorporated
+
 4. We will more clearly call out common language features, such as the heap,
@@
    see the comments responding to reviewer B.
+
+  *Status* incorporated
 
@@
    For details, see the comments responding to reviewer A.
+
+  *Status* incorporated
 
@@
 
-> There is one topic beyond my evaluation above worth considering:
-> callee-save registers.  I fear these may be fairly tricky to support in
-> your set-up.  Upon function entry, callee-save registers hold data private
-> to the caller that may be passed arbitrarily deeper into the call stack,
-> then saved into arbitrary stack locations, but then must be restored
-> before return.  Have you or the closely related prior work on enforcement
-> mechanisms considered this complexity?  The invariants are clear enough
-> informally, but the issue seems to add substantial difficulty to tracking
-> that a function faithfully implements the calling convention.
-
-Callee-save registers are supported by our approach, using a
-version of the sharing model with its context initialization and
-updates modified to apply appropriate domains to registers as well as
-stack locations.  Indeed, they actually provide another interesting
-use case for the "observational" variants of our properties. Focusing
-on integrity: callee-save registers may be overwritten (violating
-stepwise integrity) but should always be restored before a
-return. This requirement corresponds to observational integrity with
-the callee-save registers sealed. (Recall that observational integrity
-compares sealed components at the start and end of a callee's
-execution, so the interim changes are not violations.)
-
-Confidentiality for callee-saves is similar. A callee-save register should
-be secret, but in observational confidentiality, if it is copied to the
-stack and then restored before return, the property will allow the original
-register to be read normally after return.  But the property will not allow
-it to be leaked in any other way.
-
-It is worth noting that the main prior work in this area by Skorstengaard et
-al. relies fundamentally on a calling convention that clears all
-non-argument registers before calls and all non-return registers before
-returns.
-
-We will add this discussion to the paper, where previously we had omitted
-it due to concerns for space and for the complexity of the explanations.
-But it both demonstrates the model's flexibility and ties in neatly to
-observational properties, so it is worth including.
-
 > Another minor question: The difference between [nested] call-by-reference
@@
 
-> Q1. How does this work improve upon Skorstengaard et al? Neither the
-> introduction nor related work section give a satisfying answer to this.
-
-We improve upon Skorstengaard et. al. in the following ways.
-
-1. We capture a desireable security characteristic of dynamically nested
-   calls: that EVERY caller should be protected from its callees.
-
-   By contrast, Skorstengaard et. al. statically divide the world into
-   trusted functions, which feature additional dynamic "reasonableness"
-   constraints on their behavior and for which their properties hold,
-   and adversary functions, which may violate those constraints but
-   are not guaranteed safety.
-
-   The dynamic nature of this boundary makes it an issue with 
-   functions that cannot be statically proven safe, but whose data
-   still need to be protected when they are behaving safely.
-   For a contrived example, consider a function that stores some
-   private state, but takes an argument "unsafe" and, if called with
-   "unsafe" set to 1, violates the reasonableness constraints. In
-   the StkTokens model, we can never guarantee the safety of the
-   data even if called with "unsafe" set to 0.
-
-   We will expand on such an example in the revised version.
-
-2. Skorstengaard's approach to proving safety via fully abstract overlay
-   semantics relies on the abstract machine being architecturally similar to
-   the concrete one (beyond the specific call and return semantics that
-   distinguish them). This makes it less portable: each architecture it
-   applies to needs its own abstract machine. Our properties should apply
-   directly to a wide range of architectures.
-
-3. We also differ from Skorstengaard et. al. in the nature of our
-   safety definition.  We use the language of formal security
-   properties to apply well-established concepts and formal techniques
-   of non-interference. Thus, we define our properties in terms of
-   traces of memory and register states, rather than in terms of an
-   ad-hoc abstract machine that is asserted to be "stack safe by construction."
-   Which approach is better may be a matter of taste, but we believe our
-   alternative merits consideration.
-
-> Q2. In L117-121 you state certain assumptions about the language
-> semantics. It is unclear to me if these assumptions restrict the languages
-> to which your stack safety definition apply. For example, does a language
-> with a heap allow you to break stack safety? What about a language with
-> delimited continuations that manipulate the stack?
-
-Our stack safety model applies regardless of how a language uses non-stack,
-non-code memory, so it works for languages with a heap.  It also applies to
-any (statically known) memory layout.
-
-Delimited continuations are not supported by our base model, so a language
-that has them would violate stack safety. We do extend the base model with
-argument passing on the stack to demonstrate how it can accommodate a
-language feature that shares stack data between callers and callees, but a
-similar extension to delimited continuations would be more complex.
-
-This distinction -- between features that could be added without changing
-the properties and ones that would require enhanced properties -- is
-important, and we will clarify in section 2 which features fall into which
-category.
-
-Some features that can be added/changed without changing the properties:
-- a heap, as stated
-- calling convention choices, like whether or not to have a frame pointer
-- the machine's ISA
-
-Features that cannot be supported without changes to the model:
-- runtime code generation (needs dynamic code annotations)
-- continuations (would be similar to argument passing)
-- exceptions (needs annotations of throw/catch points)
-- parallelism (needs fundamental changes to step, trace models)
-- address-taken locals (would be similar to argument passing, but
-  benefits from more sophisticated, heap-like memory models)
-
 > Q3. Why don't you assume well-bracketed control flow like Skorstengaard et

### 2021-07-06 submission3/response.txt
Changed title

@@
 
-> - The paper is titled "stack safety is a security property". I don't
->   see why this is surprising, or what's the novelty here. Memory safety
->   is almost always connected with security properties.
-
-In hindsight, our choice of title was confusing: We did not just mean that
-stack safety impacts security (which is obvious), but that it can usefully
-be formulated using (a refinement of) the same formal terminology and
-mathematical definitions -- confidentiality, integrity, noninterference,
-etc. -- that have been used for decades to reason about security.
-
-We plan to change our title to:
-"Formalizing Stack Safety as a Security Property"
-
 ----------------

### 2021-07-06 submission3/main.tex
Changed title
some work on stktokens comparison

@@
 %% Title information
-\title{Security Properties for Stack Safety}         %% [Short Title] is optional;
+\title{Formalizing Stack Safety as a Security Property}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 shadow stacks~\citep{Dang+15,Shanbhogue+19},
-capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
+capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
 and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
@@
 The most abstract characterization of stack safety that we are aware of is
-recent work by \citet{SkorstengaardSTK}, which defines stack safety as the conjunction of two
+recent work by \citet{SkorstengaardSTKJFP}, which defines stack safety as the conjunction of two
 properties:
@@
 condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $B$.  This is what \citet{SkorstengaardSTK} describe
+only return to $B$.  This is what \citet{SkorstengaardSTKJFP} describe
 as \emph{well-bracketed control flow (WBCF)}.
@@
 Finally, we formalize a version of WBCF, which captures the same informal idea
-as that of \citet{SkorstengaardSTK} using our context machinery.
+as that of \citet{SkorstengaardSTKJFP} using our context machinery.
 \definition
@@
 and can be used to reason about individual programs.
-\citet{SkorstengaardSTK} acknowledge the difficulty of defining
-these concepts precisely:
-``while [they] claim that [their] calling
-convention enforces control-flow correctness, [they] do not prove a
+\citet{SkorstengaardLocal} acknowledge the difficulty of defining
+these concepts precisely; they say
+``while we claim that our calling
+convention enforces control-flow correctness, we do not prove a
 general theorem that shows this, because it is not clear what such a
@@
 
-StkTokens \citep{SkorstengaardSTK} continues this work by defining a more efficient
+StkTokens \citep{SkorstengaardSTKJFP} continues this work by defining a more efficient
 calling convention that makes use of linear capabilities for stack and return
@@
 %% precisely but
-Still informally, local state encapsulation is defined as
+Still informally, local state encapsulation (LSE) is defined as
 restricting accesses to the range of memory allocated to the current stack
-frame, and well-bracketed control flow as only allowing returns from the topmost
+frame, and well-bracketed control flow (WBCF) as only allowing returns from the topmost
 frame to the immediately adjacent frame below. Formally, it improves on
@@
 
-We improve upon \citet{SkorstengaardSTK} in two ways.
-
-\begin{enumerate}
-\item We capture a desireable security characteristic of dynamically nested
+We improve upon \citet{SkorstengaardSTKJFP} in two ways.
+
+\begin{enumerate}[label=(\roman*)]
+\item We capture a desirable security characteristic of dynamically nested
       calls: that {\em every} caller should be protected from its callees.
@@
 
-      The dynamic nature of this boundary makes it an issue with 
+      Supporting a dynamic notion of adversary is important for 
       functions that cannot be statically proven safe, but whose data
       still need to be protected when they are behaving safely.
-      For a contrived example, consider a function that stores some
-      private state, but takes an argument ``unsafe'' and, if called with
-      ``unsafe'' set to 1, violates the reasonableness constraints. In
-      the StkTokens model, we can never guarantee the safety of the
-      data even if called with "unsafe" set to 0.
-
-      \sna{TODO: expand}
+      For example, consider a function that obeys the reasonableness
+      constraints on most inputs, but can be subverted by malicious inputs
+      in such a way that it violates the reasonableness constraints. In
+      the StkTokens model, we cannot guarantee to protect this function's
+      data from its callees even when it called with non-malicious inputs. 
+      \apt{too abstract now?}
+%      \sna{TODO: expand}
 
 \item Their approach to proving safety via fully abstract overlay
-      semantics relies on the abstract machine being architecturally similar to
-      the concrete one (beyond the specific call and return semantics that
-      distinguish them). This makes it less portable: each architecture it
-      applies to needs its own abstract machine. Our properties should apply
-      directly to a wide range of architectures.
+  semantics relies on the definition of a ``stack-safe-by-construction''
+  abstract machine that is closely tied to the underlying actual
+  machine architecture and and enforcement mechanism.
+  %the the abstract machine being architecturally similar to
+  %      the concrete one (beyond the specific call and return semantics that
+  %      distinguish them).
+  While the fundamental approach can no doubt be ported to other architectures and enforcement
+  schemes, each such port requires a new abstract machine definition whose
+  safe-by-construction character must be (informally) verified afresh.
+  Our properties are more generic, and apply directly to a wide range of architectures,
+  code features, and enforcement mechanisms.
+  \apt{too strong? perhaps should (re-)enumerate what we need to adjust in each setting.}
+
 \end{enumerate}
 
-\citet{SkorstengaardSTK} enumerate a useful list of criteria for judging
-proposed stack safety properties: they should be (1) intuitive; (2)
+\citet{SkorstengaardSTKJFP} propose a useful list of criteria for judging
+candidate stack safety properties: they should be (1) intuitive; (2)
 useful for reasoning; (3) reusable in secure compiler chains; (4)
-arguably ``complete;'' and (5) potentially scalable.  We believe our
-approach competes well with theirs on criteria (1), (4), and (5).
-Their approach is specifically set up to address (3), while
-ours is not. Addressing (2) remains as future work for us.
+arguably ``complete'' with respect to informal notions of stack
+safety; and (5) potentially scalable to richer settings such as
+dynamic code generation or multithreading.
+
+For criterion (1), we believe our property definitions are at least as intuitive
+as theirs. 
+We use the language of formal security
+properties to apply well-established concepts and formal techniques
+of non-interference. Thus, we define our properties in terms of
+traces of memory and register states, rather than in terms of an
+ad-hoc abstract machine. 
+Which approach is better may be a matter of taste, but we believe that our
+alternative has the merit of explicitly relating stack safety to existing
+notions of security.
+
+With respect to criterion (4), we believe that our support
+for dynamically changing notion of adversaries, as outlined in point (i) above,
+captures a better definition of safety than LSE; moreover, our observational
+properties capture a richer and more intensional notion of stack safety
+that can be applied to a broader range of enforcement mechanisms and code features.
+As to criterion (5), we have demonstrated that our approach extends
+to coroutines, which is a first step towards multithreading.
+Dynamic code generation is more challenging, but StkTokens does not yet
+handle this either.
+StkTokens is specifically set up to address criterion (3), while this is not
+a particular goal of our work. We do agree that criterion (2) is important,
+and this remains as future work for us.
+Finally, we believe that having a generic definition of stack
+safety that abstracts over machine details is also
+an important criterion, and here we do better than StkTokens, as described
+in point (ii) above.
 

### 2021-07-07 submission3/main.tex
Move section 4 before section 2

@@
 %  to leave it out of the title but hit it strongly in the abstract.}%
-What exactly does ``stack safety'' mean? The phrase is associated with a
+What does ``stack safety'' mean, exactly? The phrase is associated with a
 variety of compiler,
 run-time, and hardware mechanisms for protecting stack
-memory.  But these mechanisms typically lack precise specifications,
+memory, but these mechanisms typically lack precise specifications,
 relying instead on informal descriptions and examples of bad
@@
 % typical pattern of matched calls and returns.
-%
-We use these properties to validate the stack-safety ``micro-policies''
+
+We use these properties to validate the stack-safety {\em micro-policies}
 proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
@@
 ``lazy'' micro-policy corresponds to a slightly weaker and more extensional
-``observational'' variant of our properties.
+observational variant of our properties.
 %
@@
 
-The call stack is a perennial target for low-level attacks, leading to
-consequences ranging from leakage or corruption of private stack data to
-control-flow hijacking. To prevent or detect such attacks, a profusion of
+The call stack is a perennial target for low-level attacks, leading to a
+range of dire consequences, from leakage or corruption of private stack data
+to control-flow hijacking. To prevent or detect such attacks, a profusion of
 software and hardware protections have been proposed,
@@
 \fi
-%
+
 The protections offered by such mechanisms are commonly described in terms
@@
 But these mechanisms can be quite intricate,
-and they often require trading off protection and efficiency.
+and they often involve trading off protection and performance.\bcp{I'm a
+  little uncomfortable about that point: to the extent that applying these
+  mechanisms involves trading off performance and protection}
 Thus, it would be useful to have a precise, generic, and formal
 specification for stack
-safety, both as a basis for comparing the security claims of different
+safety, as a basis both for comparing the security claims of different
 enforcement techniques and for validating more rigorously that these claims
@@
 caller's data. Throughout, we note how we expect elements of the formalism
-to interact with enforcement mechanisms beyond micro-policies.
+to interact with\bcp{``interact with''??} enforcement mechanisms beyond micro-policies.
 
@@
 %% %efficient, but our variant is {\em not} efficient!}
-In summary, we offer the following contributions:
+In the rest of the paper, we begin with an informal overview of our key ideas via
+examples (\cref{sec:running-example}), followed by definitions of some key
+concepts (\cref{sec:setup}) and a discussion of the threat
+model we consider (\cref{sec:threat}). We then offer the following contributions:
 
@@
 
-We begin with a description of our setting (\cref{sec:setup}) and threat
-model (\cref{sec:threat}), and an informal overview of our key ideas via
-examples (\cref{sec:running-example}). \Cref{sec:relwork,sec:future} discuss related and future work.
-
-\section{Setting}
-\label{sec:setup}
-
-\paragraph*{Machine model}
-Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the concrete machine that instantiates it. Without changing the
-properties, we can support any register
-machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
-details of the machine's ISA are unimportant. We can support any static memory,
-including a heap -- our properties define stack protection regardless of how
-non-stack memory is used. Likewise, we can support a range of calling convention
-choices, like whether or not to have a frame pointer.
-
-Some calling convention choices, such as callee-save registers and argument
-passing on the stack, require enhancements to our base properties. We present
-an enhancement that supports argument passing. Other plausible enhancements
-include:
-
-\begin{itemize}
-\item address-taken locals -- similar to argument passing, but
-      would benefit from more sophisticated, heap-like memory models
-\item exceptions -- needs annotations of throw/catch points
-\item continuations -- would be similar to argument passing
-\end{itemize}
-
-Runtime code generation would be much more challenging to support, as
-it requires dynamic code annotations. Parallelism would require
-fundamental changes to the system model. So, we consider these features
-fully out of scope.
-
-\paragraph*{What is a ``call''?}  One significant challenge in formulating our
-properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy
-\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
-particular, there may be no single ``call'' instruction that does all the
-required manipulation of the \(\PCname\) and stack pointer. Instead, there
-might be just a simple branch instruction that stores the old \(\PCname\) in
-a register, with the rest of the call sequence performed by other
-instructions either before or after the jump.  Similarly, there may be no
-``return'' macro-instruction, but rather a ``return sequence'' involving an
-indirect branch at some point.  Thus, the same instruction opcode may be
-used in some places as part of a call or return sequence and in other places
-for other purposes.
-
-Since our stack-safety properties are all about protecting ``callers'' from
-their ``callees,'' we need some clear-cut way to recognize when the program
-makes or returns from a call.  Accordingly, we assume we are given a machine
-language program {\em annotated} with the locations of instructions
-representing calls.  We declare that a call has occurred when the machine
-has just executed an instruction marked as a call.
-%
-These call annotations would typically come from a compiler together with
-the machine code, but we do not assume anything about their provenance.
-Annotations connect the enforcement mechanism to the property: we don't
-protect unmarked calls, because a mechanism's starting conditions might
-depend on knowing where calls are.
-
-Returns, on the other hand, need not be annotated: we can tell when a callee
-has returned control to the caller when... we see control return to the
-caller! We don't require the callee's cooperation at all.
-For example, under a standard calling convention, we can see that a
-return has occurred occurred when state of the machine reaches a state where
-the stack pointer has the same value as it did at the point of some earlier
-call and the $\PCname$ points to the instruction following that call.  We
-abstract away from the details of specific calling conventions by assuming
-that we are given some way of telling when a given machine state is a
-{\em return target} with respect to some set of pending calls.
-
-%, returns, and function entry points\BCP{I
-%  still want to talk about whether returns and entry points need to be
-%  annotated.  To me, these annotations do not make intuitive sense: it is
-%  the caller's responsibility to know whether the address that it is calling
-%is really an entry point, and we can trust it to check this.  Similarly, it
-%is not the callee's role to say when it is returning; it is the caller's
-%prerogative to decide when it has been returned to!  I.e., from the caller's
-%point of view, it has been returned to when control resumes at the
-%instruction following the call with the SP reset to where it was at the
-%point of the call. (This is the point where it wants to be able to access
-%its own local state again and expected to find it unmolested.)}.
-
-\paragraph*{Policies}
-We model the base machine's behavior as a step function over machine states,
-and a machine
-enhanced with an enforcement mechanism, or {\em policy} for short, as a
-partial step function over machine states extended with some
-auxiliary state, refining the base machine. When the policy-enhanced machine
-does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
-action that would violate the security property;
-our properties are therefore naturally \emph{termination insensitive.}
-This model will be made more precise in \cref{sec:prelim}.
-
-Dividing policy state from machine state helps us reason about
-enforcement mechanisms that keep a significant amount of data separate from the
-primary execution data to inform security. Obviously tag-based systems do this,
-with tags and extra policy state separated from and untouchable by application code.
-Another significant example is the shadow stack, where a second stack
-mirrors the real one, and a return may only occur if both stacks agree on the
-target address. Typically the shadow stack is isolated using a separate protection
-mechanism.
-
-The separation is smaller in protection mechanisms that interact more directly
-with application code. Some mechanisms might exhibit failstop behavior, but have
-no protected policy state; others might never even failstop. Then the policy-enhanced
-machine is simply identical to the original. This is the case for software-based
-techniques.
-
-\section{Threat Model}
-\label{sec:threat}
-
-Consider a common scenario: a developer wants to link against an existing
-code-base, which will then occupy the same address space as their own code.
-But perhaps the external code is buggy, or even malicious, and contains
-exploits that can be used to cause it to modify the developer's data or
-exfiltrate secrets. For that matter, the developer may not be a particularly
-secure programmer, and may not be able to rely on their own defensive programming
-to keep their secrets safe.
-
-The developer likely relies instead on simple reasoning principles about how
-their code should work. Local variables do not change across function calls,
-and the callees generally do the same thing regardless of when and where they
-are called (up to a reasonable interface, such as their parameters and global
-variables.) If these intuitions prove false, then the programmer will reason
-incorrectly about their program and especially about its security, and
-be exposed to attack.
-
-Intuitively, stack safety means that the stack frame of each caller is protected
-from attempts by a confused or malicious callee (including all its
-descendant callees) to read or write its private data.
-That is, the attacker model treats every caller as a defender and its callee as
-an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways. We do rule out some
-scenarios with a ``well-formedness'' criterion on initial states, ensuring that
-they are set up properly for the enforcement mechanism in question.
-When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
-well-behaved calls and returns are implemented by fixed sequences of instructions
-which the tag policy will force to be executed from start to finish,
-termed {\em blessed sequences}. Well-formedness requires that all annotated
-instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
-This is all we need for the policy to protect itself: it can then failstop if it would
-execute the blessed sequence out of order.
-
-We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. Hardware and timing attacks are out of scope.
-So are attacks on the heap -- we believe that the work of
-\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with our
-to protect both stack and heap.
-
-%We model ordinary program state and policy-enforcement state separately to make it
-%easier to use well-established concepts from language-based security.
-%In particular, confidentiality has an elegant extensional characterization in terms of
-%\emph{noninterference}: a program preserves the confidentiality of certain data if
-%\emph{varying} that data would not change the program's observable
-%behavior~\citep{6234468}.
-%In applying this idea to a system that incorporates policy enforcement, only
-%ordinary program data should be varied, while any associated policy
-%state should be left alone; our model makes this distinction obvious.
-%That is, we compare the actual execution of the program with a \emph{hypothetical}
-%execution where data has been varied.
-%We will also introduce a notion of {\em observational integrity} that relies on the idea of
-%\emph{rolling back} program state; again, our model makes it easy to describe
-%rollback of ordinary program data while leaving policy state alone.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
-%We believe that it is flexible enough to capture code-altering approaches, such as code
-%rewriting or capability-based techniques, though perhaps with less elegance.
-
-%\paragraph*{Micro-Policies}
-%
-%Micro-Policies are a flexible tag-based, hardware-accelerated
-%reference monitoring mechanism; they been applied to stack safety
-%enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
-%Here the policy state consists of metadata tags (e.g., identifying stack frames)
-%attached to each value in memory or registers, and the policy step
-%function checks that each machine operation obeys a set of rules on tags
-%(e.g., that the current $\PCname$ tag matches the tag on the stack
-%location being accessed), and halts the machine if not.  Distinguishing
-%ordinary machine state and policy state is very natural in this setting.
-%% Here the machine state is extended with tags paired with all components of the
-%% system, and a monitor checks for each operation that the tags on the operands
-%% obey a set of rules. One could treat tags directly as part of the machine
-%% state, and build the rules into the step function. But local state
-%% encapsulation properties as we frame them are not intuitive in their
-%% interaction with tags or other additional enforcement state, so it is helpful
-%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
-%% a software fail state.
-%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-%%   policy? }
-
-%\iftrue\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
-%These approaches are represented as policies that
-%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
-%being produced by the compiler.) The set of policy states is again the unit type
-%and the step function constant. Property enforcement comes from the fact that
-%only ill-formed states can induce a trace that violates the property.
-
-%% \Paragraph*{Null policy}
-%% %
-%% The simplest possible policy is the neutral policy that carries no information
-%% and does nothing. Its set of policy states is the unit type, and its policy step
-%% function is the constant unit function, thereby allowing all calls to the machine
-%% step function. Its initialization function is defined on all states. This
-%% corresponds to the lifting of our machine model to the policy setting. Machines
-%% that fulfill a property by construction will also fulfill that property with the
-%% null policy.
-
-%
-%\leo{TODO: Add citations of mitigation techniques}
-%\rb{And for those techniques, explain how they reflect their motivating
-%examples. What does each technique do for those examples? How are they
-%connected to this work?}
-%
-%\paragraph*{Bad Write}
-%
-%\[
-%44 : [\rsp - 5] \leftarrow r_4
-%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
-%\]
-%
-%Instruction 44 currently is
-%$ [\rsp - 2] \leftarrow r_4$,
-%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
-%with the
-%effect of storing the contents of $r_4$ as the result of {\tt g}. If
-%instead it wrote to location $\rsp - 5$ then it would be overwriting
-%private data (the return address) belonging to {\tt f}.
-%
-%\paragraph*{Bad Read}
-%
-%\[
-%30 : r_5  \leftarrow [\rsp - 6]
-%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-%\]
-%
-%Instruction 30 currently is
-%$ r_5 \leftarrow [\rsp - 4]$,
-%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
-%with the
-%effect of loading the argument of {\tt f} to $r_5$. If instead it
-%loaded the contents of location $\rsp - 6$, it would obtain access
-%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
-%outputs the contents of $r_5$, that value would immediately be made public.
-%
-%\paragraph*{Bad Control Flow}
-%
-%\[
-%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
-%%5 : \jalr ~ r_4 ~ \rra ~ 0
-%\]
-%If a {\jalr} instruction was added to {\tt main} with its target being
-%the middle of the code block of {\tt g}.\apt{describe result of this?}
+\Cref{sec:relwork,sec:future} discuss related and future work.
 
@@
 
-This section introduces the intuition behind the properties through a series
-of examples. We show our examples both in C syntax and in corresponding assembly code for a
-simple two-address machine. The address of each instruction is shown to its left.
+This section introduces the intuition behind our formal stack safety properties through a series
+of examples, each presented both in C syntax and in corresponding assembly code for a
+simple two-address machine (see Figure~\ref{fig:int1}). The address of each
+instruction is shown to its left.
 The assembly code reflects a simple compilation model in which
@@
 as \emph{well-bracketed control flow (WBCF)}.
+
+\section{Key Concepts}
+\label{sec:setup}
+
+\paragraph*{Machine model}
+Our properties are defined in terms of an abstract machine model that makes very few
+assumptions about the concrete machine that instantiates it. Without changing the
+properties, we can support any register
+machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
+details of the machine's ISA are unimportant. We can support any static memory,
+including a heap -- our properties define stack protection regardless of how
+non-stack memory is used. Likewise, we can support a range of calling convention
+choices, like whether or not to have a frame pointer.
+
+Some calling convention choices, such as callee-save registers and argument
+passing on the stack, require enhancements to our base properties. We present
+an enhancement that supports argument passing. Other plausible enhancements
+include:
+
+\begin{itemize}
+\item address-taken locals -- similar to argument passing, but
+      would benefit from more sophisticated, heap-like memory models
+\item exceptions -- needs annotations of throw/catch points
+\item continuations -- would be similar to argument passing
+\end{itemize}
+
+Runtime code generation would be much more challenging to support, as
+it requires dynamic code annotations. Parallelism would require
+fundamental changes to the system model. So, we consider these features
+fully out of scope.
+
+\paragraph*{What is a ``call''?}  One significant challenge in formulating our
+properties arises from the fact that we want to define stack safety at the
+machine-code level for a variety of ISAs, including RISCy
+\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
+particular, there may be no single ``call'' instruction that does all the
+required manipulation of the \(\PCname\) and stack pointer. Instead, there
+might be just a simple branch instruction that stores the old \(\PCname\) in
+a register, with the rest of the call sequence performed by other
+instructions either before or after the jump.  Similarly, there may be no
+``return'' macro-instruction, but rather a ``return sequence'' involving an
+indirect branch at some point.  Thus, the same instruction opcode may be
+used in some places as part of a call or return sequence and in other places
+for other purposes.
+
+Since our stack-safety properties are all about protecting ``callers'' from
+their ``callees,'' we need some clear-cut way to recognize when the program
+makes or returns from a call.  Accordingly, we assume we are given a machine
+language program {\em annotated} with the locations of instructions
+representing calls.  We declare that a call has occurred when the machine
+has just executed an instruction marked as a call.
+%
+These call annotations would typically come from a compiler together with
+the machine code, but we do not assume anything about their provenance.
+Annotations connect the enforcement mechanism to the property: we don't
+protect unmarked calls, because a mechanism's starting conditions might
+depend on knowing where calls are.
+
+Returns, on the other hand, need not be annotated: we can tell when a callee
+has returned control to the caller when... we see control return to the
+caller! We don't require the callee's cooperation at all.
+For example, under a standard calling convention, we can see that a
+return has occurred occurred when state of the machine reaches a state where
+the stack pointer has the same value as it did at the point of some earlier
+call and the $\PCname$ points to the instruction following that call.  We
+abstract away from the details of specific calling conventions by assuming
+that we are given some way of telling when a given machine state is a
+{\em return target} with respect to some set of pending calls.
+
+%, returns, and function entry points\BCP{I
+%  still want to talk about whether returns and entry points need to be
+%  annotated.  To me, these annotations do not make intuitive sense: it is
+%  the caller's responsibility to know whether the address that it is calling
+%is really an entry point, and we can trust it to check this.  Similarly, it
+%is not the callee's role to say when it is returning; it is the caller's
+%prerogative to decide when it has been returned to!  I.e., from the caller's
+%point of view, it has been returned to when control resumes at the
+%instruction following the call with the SP reset to where it was at the
+%point of the call. (This is the point where it wants to be able to access
+%its own local state again and expected to find it unmolested.)}.
+
+\paragraph*{Policies}
+We model the base machine's behavior as a step function over machine states,
+and a machine
+enhanced with an enforcement mechanism, or {\em policy} for short, as a
+partial step function over machine states extended with some
+auxiliary state, refining the base machine. When the policy-enhanced machine
+does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
+action that would violate the security property;
+our properties are therefore naturally \emph{termination insensitive.}
+This model will be made more precise in \cref{sec:prelim}.
+
+Dividing policy state from machine state helps us reason about
+enforcement mechanisms that keep a significant amount of data separate from the
+primary execution data to inform security. Obviously tag-based systems do this,
+with tags and extra policy state separated from and untouchable by application code.
+Another significant example is the shadow stack, where a second stack
+mirrors the real one, and a return may only occur if both stacks agree on the
+target address. Typically the shadow stack is isolated using a separate protection
+mechanism.
+
+The separation is smaller in protection mechanisms that interact more directly
+with application code. Some mechanisms might exhibit failstop behavior, but have
+no protected policy state; others might never even failstop. Then the policy-enhanced
+machine is simply identical to the original. This is the case for software-based
+techniques.
+
+\section{Threat Model}
+\label{sec:threat}
+
+Consider a common scenario: a developer wants to link against an existing
+code-base, which will then occupy the same address space as their own code.
+But perhaps the external code is buggy, or even malicious, and contains
+exploits that can be used to cause it to modify the developer's data or
+exfiltrate secrets. For that matter, the developer may not be a particularly
+secure programmer, and may not be able to rely on their own defensive programming
+to keep their secrets safe.
+
+The developer likely relies instead on simple reasoning principles about how
+their code should work. Local variables do not change across function calls,
+and the callees generally do the same thing regardless of when and where they
+are called (up to a reasonable interface, such as their parameters and global
+variables.) If these intuitions prove false, then the programmer will reason
+incorrectly about their program and especially about its security, and
+be exposed to attack.
+
+Intuitively, stack safety means that the stack frame of each caller is protected
+from attempts by a confused or malicious callee (including all its
+descendant callees) to read or write its private data.
+That is, the attacker model treats every caller as a defender and its callee as
+an attacker, meaning that with nested calls, a single function invocation
+may be viewed in both ways. We do rule out some
+scenarios with a ``well-formedness'' criterion on initial states, ensuring that
+they are set up properly for the enforcement mechanism in question.
+When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
+well-behaved calls and returns are implemented by fixed sequences of instructions
+which the tag policy will force to be executed from start to finish,
+termed {\em blessed sequences}. Well-formedness requires that all annotated
+instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
+This is all we need for the policy to protect itself: it can then failstop if it would
+execute the blessed sequence out of order.
+
+We adopt a strong threat model in which both caller and callee may execute
+arbitrary machine code. Hardware and timing attacks are out of scope.
+So are attacks on the heap -- we believe that the work of
+\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with our
+to protect both stack and heap.
+
+%We model ordinary program state and policy-enforcement state separately to make it
+%easier to use well-established concepts from language-based security.
+%In particular, confidentiality has an elegant extensional characterization in terms of
+%\emph{noninterference}: a program preserves the confidentiality of certain data if
+%\emph{varying} that data would not change the program's observable
+%behavior~\citep{6234468}.
+%In applying this idea to a system that incorporates policy enforcement, only
+%ordinary program data should be varied, while any associated policy
+%state should be left alone; our model makes this distinction obvious.
+%That is, we compare the actual execution of the program with a \emph{hypothetical}
+%execution where data has been varied.
+%We will also introduce a notion of {\em observational integrity} that relies on the idea of
+%\emph{rolling back} program state; again, our model makes it easy to describe
+%rollback of ordinary program data while leaving policy state alone.
+
+%% \leo{Policies of interest: null, micro, cherri, software-only}
+%% \leo{TODO: Write down carefully}
+
+%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
+%We believe that it is flexible enough to capture code-altering approaches, such as code
+%rewriting or capability-based techniques, though perhaps with less elegance.
+
+%\paragraph*{Micro-Policies}
+%
+%Micro-Policies are a flexible tag-based, hardware-accelerated
+%reference monitoring mechanism; they been applied to stack safety
+%enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
+%Here the policy state consists of metadata tags (e.g., identifying stack frames)
+%attached to each value in memory or registers, and the policy step
+%function checks that each machine operation obeys a set of rules on tags
+%(e.g., that the current $\PCname$ tag matches the tag on the stack
+%location being accessed), and halts the machine if not.  Distinguishing
+%ordinary machine state and policy state is very natural in this setting.
+%% Here the machine state is extended with tags paired with all components of the
+%% system, and a monitor checks for each operation that the tags on the operands
+%% obey a set of rules. One could treat tags directly as part of the machine
+%% state, and build the rules into the step function. But local state
+%% encapsulation properties as we frame them are not intuitive in their
+%% interaction with tags or other additional enforcement state, so it is helpful
+%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
+%% a software fail state.
+%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
+%%   policy? }
+
+%\iftrue\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
+%These approaches are represented as policies that
+%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
+%being produced by the compiler.) The set of policy states is again the unit type
+%and the step function constant. Property enforcement comes from the fact that
+%only ill-formed states can induce a trace that violates the property.
+
+%% \Paragraph*{Null policy}
+%% %
+%% The simplest possible policy is the neutral policy that carries no information
+%% and does nothing. Its set of policy states is the unit type, and its policy step
+%% function is the constant unit function, thereby allowing all calls to the machine
+%% step function. Its initialization function is defined on all states. This
+%% corresponds to the lifting of our machine model to the policy setting. Machines
+%% that fulfill a property by construction will also fulfill that property with the
+%% null policy.
+
+%
+%\leo{TODO: Add citations of mitigation techniques}
+%\rb{And for those techniques, explain how they reflect their motivating
+%examples. What does each technique do for those examples? How are they
+%connected to this work?}
+%
+%\paragraph*{Bad Write}
+%
+%\[
+%44 : [\rsp - 5] \leftarrow r_4
+%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
+%\]
+%
+%Instruction 44 currently is
+%$ [\rsp - 2] \leftarrow r_4$,
+%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
+%with the
+%effect of storing the contents of $r_4$ as the result of {\tt g}. If
+%instead it wrote to location $\rsp - 5$ then it would be overwriting
+%private data (the return address) belonging to {\tt f}.
+%
+%\paragraph*{Bad Read}
+%
+%\[
+%30 : r_5  \leftarrow [\rsp - 6]
+%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
+%\]
+%
+%Instruction 30 currently is
+%$ r_5 \leftarrow [\rsp - 4]$,
+%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
+%with the
+%effect of loading the argument of {\tt f} to $r_5$. If instead it
+%loaded the contents of location $\rsp - 6$, it would obtain access
+%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
+%outputs the contents of $r_5$, that value would immediately be made public.
+%
+%\paragraph*{Bad Control Flow}
+%
+%\[
+%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
+%%5 : \jalr ~ r_4 ~ \rra ~ 0
+%\]
+%If a {\jalr} instruction was added to {\tt main} with its target being
+%the middle of the code block of {\tt g}.\apt{describe result of this?}
 

### 2021-07-07 submission3/main.tex
Polishing

@@
 For the moment, arguments and return values are passed in registers.
-The machine is word-addressed. The stack grows upward in
+The stack grows upward in
 memory and {\tt \%sp} points to the first unused word above the top of the stack.
@@
 At the right of each example, we show the layout of stack memory just before
-the specified instruction is executed; each box represents a memory word, and
-addresses increase to the right.
-
+the specified instruction is executed; each box represents a memory word
+(the machine is word addressed), and addresses increase to the right.
+
+\bcp{Have we made the diagrams comprehensible when printed in
+  B/W or viewed by a colorblind person?  If so we should say that the colors
+are just extra hints.}%
 Colored boxes indicate an assignment of the address to a
-{\em domain}---a label reflecting a particular set of access rights. Domains
+{\em domain}---a label reflecting a particular set of intended access
+rights.\bcp{should we say here that it is the job of an enforcement mechanism to
+  manage and enforce these access permissions?} Domains
 are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
@@
 figures. For our purposes, \(\outside\) will apply to global
-variables and code. Domains are part of the {\em context} of a state.
-We will introduce the remaining context as needed.
-
-Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
+variables and code. Domains are part of the {\em context} of a machine state.
+(We will introduce some other parts of the context as needed.)
+
+Figure~\ref{fig:int1} shows a program with a simple integrity violation. It consists of a {\tt main}
 function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
@@
   \centering
+\hspace*{-6em}
   \begin{subfigure}[t]{.2\textwidth}
@@
 \label{sec:setup}
+
+\bcp{I am having trouble deciding what this section should be called.
+  Probably this means I don't really understand its purpose...}
 

### 2021-07-07 submission3/main.tex
Answered BCP and APT as much as I could, and added forward pointer to callee-saves

@@
 
-\section{Key Concepts}
+\section{Setting}
 \label{sec:setup}
@@
   Probably this means I don't really understand its purpose...}
+\sna{Propose merging it with next section like so.}
 
@@
 
-\section{Threat Model}
+\paragraph*{Threat Model}
 \label{sec:threat}
@@
 
+Lazy policies are not the only scenario to call for a weaker property of this kind.
+Some code features may involve moving data, overwriting its original location,
+and ultimately restoring it before the change becomes visible; for example, the
+``callee-saves'' registers in some calling conventions will be stashed and restored.
+Stepwise stack safety can never support such a feature, but in section \ref{calleesave}
+we show how {\em observational properties} can.
+
 \paragraph*{Deferring Integrity Checks}
@@
       \apt{too abstract now?}
-%      \sna{TODO: expand}
+      \sna{Maybe, but I'm not sure how much being concrete about it serves us.}
 
@@
   safe-by-construction character must be (informally) verified afresh.
-  Our properties are more generic, and apply directly to a wide range of architectures,
-  code features, and enforcement mechanisms.
-  \apt{too strong? perhaps should (re-)enumerate what we need to adjust in each setting.}
+  Our properties are more generic: since we abstract over most of the machine
+  state as well as the sealing and return conventions, they apply directly
+  to a range of architectures, code features, and enforcement mechanisms.
+
+  Likewise, it would be possible to extend their abstract machine to pass arguments
+  on the stack, or to support callee-save registers. It might even viable to
+  construct a safe-by-construction abstract coroutine machine, but so far this is
+  hypothetical. We implement these scenarios in fact, with simple modifications to
+  the domain model of each one.
 

### 2021-07-07 submission3/main.tex
Intro pass

@@
 use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate many random programs and check
+tool for Coq, to generate random programs and check
 that Roessler and Dehon's micro-policies correctly abort the ones that
@@
 caller's data. Throughout, we note how we expect elements of the formalism
-to interact with\bcp{``interact with''??} enforcement mechanisms beyond micro-policies.
+to apply to enforcement mechanisms beyond micro-policies.
 
@@
 safety~(\cref{sec:prelim,sec:lse}).
-\item We use property-based randomized testing to check that this stepwise
+\item We use property-based random testing to check that this stepwise
 property is actually enforced by the {\em Depth Isolation} policy

### 2021-07-07 submission3/main.tex
Nit

@@
 rights.\bcp{should we say here that it is the job of an enforcement mechanism to
-  manage and enforce these access permissions?} Domains
-are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
-(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
+  manage and enforce these access permissions?} Such a domain 
+can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
+(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
 (outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
-figures. For our purposes, \(\outside\) will apply to global
+figures\leo{No outside?}. For our purposes, \(\outside\) will apply to global
 variables and code. Domains are part of the {\em context} of a machine state.

### 2021-07-07 submission3/main.tex
Nit in 2

@@
 A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, here represented by sharing a color with the sealed data.
+when it is reached, represented here by sharing a color with the sealed data.
 

### 2021-07-07 submission3/main.tex
title for 3
WIP

@@
 
-\bcp{I am having trouble deciding what this section should be called.
-  Probably this means I don't really understand its purpose...}
-\sna{Propose merging it with next section like so.}
+\bcp{Needs an intro sentence.}
 
@@
 exploits that can be used to cause it to modify the developer's data or
-exfiltrate secrets. For that matter, the developer may not be a particularly
-secure programmer, and may not be able to rely on their own defensive programming
+exfiltrate secrets. For that matter, the developer may not experienced at
+secure programming and so may not be able to rely on their own skill
 to keep their secrets safe.
@@
 We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. Hardware and timing attacks are out of scope.
-So are attacks on the heap -- we believe that the work of
-\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with our
+arbitrary machine code. Hardware and timing attacks are out of scope, as are
+are attacks on the heap---though we believe that the work of
+\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
 to protect both stack and heap.
@@
 
+\vspace*{-\medskipamount}
+
 \paragraph*{Context Segments}
@@
 \end{center}
+
+\vspace*{-\bigskipamount}
 
@@
 When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We now proceed to state our properties.
+call stack we are and which components are sealed. We now state our properties.
 The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
@@
 While deferred integrity meaningfully captures the intuition
-behind lazy tagging, we seek properties that can be used to evaluate the broadest
-possible range of reasonable
+behind lazy tagging, we are interested in properties that can be used to
+evaluate the broadest possible range of reasonable
 enforcement mechanisms. Consider a hypothetical
@@
 the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
-initial observation, a machine-trace, and an observation trace:
+initial observation, a machine-trace, and an observation trace:\bcp{This
+  would read better and take less space if all three rules were squished
+  onto one line.}
 

### 2021-07-07 submission3/main.tex
Machine model

--- 2021-07-07 submission3/main.tex
@@
 \sna{Propose merging it with next section like so.}
+\leo{Machine, Enforcement, and Threat Model?}
 
@@
 Some calling convention choices, such as callee-save registers and argument
-passing on the stack, require enhancements to our base properties. We present
-an enhancement that supports argument passing. Other plausible enhancements
+passing on the stack, require natural enhancements to our base properties. We
+present such enhancements in Section~\ref{sec:ext}, after we formalize the
+essence of stack safety. Other plausible enhancements
 include:
@@
 fundamental changes to the system model. So, we consider these features
-fully out of scope.
+fully out of scope.\leo{Shouldn't coroutines be mentioned here?}
 

### 2021-07-07 submission3/main.tex

--- 2021-07-07 submission3/main.tex
@@
 
-\bcp{Needs an intro sentence.}
+\bcp{I am having trouble deciding what this section should be called.
+  Probably this means I don't really understand its purpose...}
+\sna{Propose merging it with next section like so.}
+\leo{Machine, Enforcement, and Threat Model?}
+
+\bcp{And: Needs an intro sentence.}
 

### 2021-07-07 submission3/main.tex
Nit

@@
 For example, under a standard calling convention, we can see that a
-return has occurred occurred when state of the machine reaches a state where
+return has occurred when state of the machine reaches a state where
 the stack pointer has the same value as it did at the point of some earlier
@@
 
-\paragraph*{Policies}
+\paragraph*{Policies}\leo{Change this to ``enforcement model'' if we go with that title}
 We model the base machine's behavior as a step function over machine states,

### 2021-07-07 submission3/main.tex
Callee saves

--- 2021-07-07 submission3/main.tex
@@
 %  approaches}\bcp{Yes, we have to be very careful about this}.
-%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
+%\leo{Examples at a high level}: such as---does not overwrite return addresses, or ...
 The most abstract characterization of stack safety that we are aware of is
@@
 refers to the fact that the property must hold on every step. %Later we will
-%loosen this condition in the interest of extensionality -- the caller only
+%loosen this condition in the interest of extensionality---the caller only
 %cares about changes it can see after the return, not changes during execution.
@@
 details of the machine's ISA are unimportant. We can support any static memory,
-including a heap -- our properties define stack protection regardless of how
+including a heap---our properties define stack protection regardless of how
 non-stack memory is used. Likewise, we can support a range of calling convention
@@
 Some calling convention choices, such as callee-save registers and argument
-passing on the stack, require natural enhancements to our base properties. We
-present such enhancements in Section~\ref{sec:ext}, after we formalize the
-essence of stack safety. Other plausible enhancements
+passing on the stack, require enhancements to our base properties. We present
+an enhancement that supports argument passing. Other plausible enhancements
 include:
@@
 \begin{itemize}
-\item address-taken locals -- similar to argument passing, but
+\item address-taken locals---similar to argument passing, but
       would benefit from more sophisticated, heap-like memory models
-\item exceptions -- needs annotations of throw/catch points
-\item continuations -- would be similar to argument passing
+\item exceptions---needs annotations of throw/catch points
+\item continuations---would be similar to argument passing
 \end{itemize}
@@
 fundamental changes to the system model. So, we consider these features
-fully out of scope.\leo{Shouldn't coroutines be mentioned here?}
+fully out of scope.
 
@@
 %A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-%\Cref{sec:enforcement}. -- APT: We just said this.
+%\Cref{sec:enforcement}.---APT: We just said this.
 
@@
 
-We now can sketch observational versions of both integrity and confidentiality.
+We now sketch observational versions of both integrity and confidentiality.
 Integrity is already straightforward, simply weakening the condition that must hold
@@
 
-We have shown how our properties apply to an extremely simple
-subroutine model and validated them through randomized testing.
-But what about more realistic systems? We now demonstrate the
+We have seen how our properties apply to an extremely simple
+subroutine model.
+What about more realistic uses of the stack? We now demonstrate the
 flexibility of our core concepts by extending
 them to a system that models the passing of
-parameters on the stack, and then to a very simple coroutine model.
+parameters on the stack (Section~\ref{sec:passing}), to callee-saves
+registers  (\ref{sec:callee}), and to a very simple coroutine model
+(\ref{sec:coroutines}).
 
@@
 sharing: passing by value on the stack, and passing by reference. We first explore
-some examples at the same level of detail as our earlier ones, and then move on to
+some examples and then move on to
 extending our formalization.
@@
 its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
-as passed between depths 0 and 1, it is not varied. Then at instruction 13 it is even
-modified, and naturally this is permitted as well.
+as passed between depths 0 and 1, it is not varied. At instruction 13 it is
+modified---naturally this is permitted as well.
 
@@
 as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference.
-{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
-{\tt f} again passing a reference to {\tt y}.
-We would expect that the first call to {\tt f} might modify
+{\tt f} takes a single argument by reference.  The function
+{\tt main} calls {\tt f}, passing a reference to {\tt x}, then clears {\tt x} and calls
+{\tt f} again, passing a reference to {\tt y}.
+We would expect that the first call to {\tt f} can modify
 {\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++, although its behavior is undefined.)
+{\tt x} in the second call. (Incidentally, this is perfectly compilable C++,
+although its behavior is officially undefined.)
 
@@
 stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike the case for pass-by-reference, there is no easy way to model the intended
+Unlike pass-by-reference, there is no easy way to model the intended
 sharing behavior of such pointers in terms of stack frame activation patterns.
@@
 that it remains accessible at \emph{all} stack depths above the caller.
-This prevents most violations of temporal safety, although stale pointers can still be
-used in certain corner cases.
-But more seriously, callees are permitted to access the location without having
+This prevents most violations of temporal safety, although stale pointers
+can still be used in certain corner cases.
+More seriously, callees are permitted to access the location without having
 been explicitly given a pointer to it.  Preventing the latter kind of abuse
-is the provence of a fine-grained memory safety policy, e.g. based on capabilities
-(implemented by tags or otherwise); we argue it is beyond the scope of stack safety
-\emph{per se}.
+is the provence of a fine-grained memory safety policy, e.g., based on
+capabilities
+(implemented by tags or otherwise); we would argue it is beyond the scope of
+stack safety \emph{per se}.
 
@@
 \subsection{Callee-saves Registers}
-
-In some systems, rather than the caller saving the non-argument data from its registers
-and clearing them (and therefore implicitly taking responsibility for protecting those data),
-the callee is responsible for this. Nevertheless we may wish for those registers to be
-protected, just as they would be in a caller-save convention.
-
-Intuitively, a caller expects a callee-saved register to be unchanged on return -- but it
+\label{sec:callee}
+
+Under some calling conventions, rather than the caller saving the non-argument
+data from its registers
+and clearing them (and therefore implicitly taking responsibility for
+protecting this data),
+the callee is responsible for this. But of course we may still wish for
+those registers to be
+protected, just as they would be in a caller-saves convention.
+
+Intuitively, a caller expects a callee-saved register to be unchanged on
+return: it
 may change during the call, as long as it is restored. We similarly expect a callee-saved
@@
 
-The remaining rules are as in the simple model.
+The remaining rules are the same as in the simple model.
 
@@
 stack integrity in the simple model, the equivalent property in the
-passing model
-
+passing model\bcp{Is something missing here?}
+%
 {\bf Observational stack integrity} with callee-saves registers substitutes the above
 \(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
-
+%
 {\bf Observational stack confidentiality} additionally substitutes \(\secret_S\).
@@
 
-Observational properties are not only for lazy enforcement mechanisms -- they reflect a useful
+Observational properties are not only for lazy enforcement mechanisms---they reflect a useful
 loosening of the stepwise properties. Consider a single register \(\reg\) that is saved.
@@
 \subsection{Coroutines}
+\label{sec:coroutines}
 

### 2021-07-07 submission3/main.tex
Comment

@@
 a positive and formal characterization of
-stack safety begins with \citet{SkorstengaardLocal}, who introduce a calling
+stack safety begins with \citet{SkorstengaardLocal}, who introduced a calling
 convention that uses local capabilities to preserve local state
@@
 \citet{SkorstengaardLocal} acknowledge the difficulty of defining
-these concepts precisely; they say
+these concepts precisely\bcp{since  their later papers do include precise
+  definitions, maybe there's no point in emphasizing what they said at the
+  beginning?}; they say
 ``while we claim that our calling

### 2021-07-07 submission3/main.tex
CP

@@
 
-StkTokens \citep{SkorstengaardSTKJFP} continues this work by defining a more efficient
-calling convention that makes use of linear capabilities for stack and return
+StkTokens \citep{SkorstengaardSTKJFP} continues this line of work by
+defining a more efficient
+calling convention that uses linear capabilities for stack and return
 pointers.
@@
 %% precisely but
-Still informally, local state encapsulation (LSE) is defined as
+Informally, local state encapsulation (LSE) is defined as
 restricting accesses to the range of memory allocated to the current stack
-frame, and well-bracketed control flow (WBCF) as only allowing returns from the topmost
-frame to the immediately adjacent frame below. Formally, it improves on
+frame, and well-bracketed control flow (WBCF) is defined as only allowing
+returns from the topmost
+frame to the immediately adjacent frame below. Formally, this work improves on
 \citet{SkorstengaardLocal} by building those properties into the semantics of a
-capability machine with a built-in call stack and call and return instructions,
-which is proven fully abstract with respect to a more concrete capability
-machine.
+capability machine with a built-in call stack and call and return
+instructions; this machine is proven fully abstract with respect to a more
+concrete capability machine.
 %% that replaces those pseudo-instructions with their calling convention.
-Their proof of full abstraction uses a standard notion of components, which
+The proof of full abstraction uses a standard notion of components, which
 import and export functions through their interfaces.
-%% ; their model of
-%% observations is limited to cotermination.
-
-We improve upon \citet{SkorstengaardSTKJFP} in two ways.
-
+
+We improve upon \citet{SkorstengaardSTKJFP} in two significant ways:
 \begin{enumerate}[label=(\roman*)]
-\item We capture a desirable security characteristic of dynamically nested
-      calls: that {\em every} caller should be protected from its callees.
-      By contrast, they statically divide the world into
-      trusted functions, which feature additional dynamic "reasonableness"
-      constraints on their behavior and for which their properties hold,
-      and adversary functions, which may violate those constraints but
-      are not guaranteed safety.
-
-      Supporting a dynamic notion of adversary is important for
-      functions that cannot be statically proven safe, but whose data
-      still need to be protected when they are behaving safely.
-      For example, consider a function that obeys the reasonableness
-      constraints on most inputs, but can be subverted by malicious inputs
-      in such a way that it violates the reasonableness constraints. In
-      the StkTokens model, we cannot guarantee to protect this function's
-      data from its callees even when it called with non-malicious inputs.
-      \apt{too abstract now?}
-      \sna{Maybe, but I'm not sure how much being concrete about it serves us.}
-
-\item Their approach to proving safety via fully abstract overlay
+\item Most importantly, our definition captures the intuition that {\em
+  every} caller should be protected from its callees.  By contrast,
+\citet{SkorstengaardSTKJFP}
+statically divide the world into trusted functions, which feature additional
+dynamic ``reasonableness'' constraints on their behavior and for which their
+properties hold, and adversary functions, which may violate those
+constraints but are not guaranteed safety.  \bcp{Readers may wonder: why
+  isn't this reasonable?  after all, if a
+  function is adversarial, why do we care whether it gets any safety
+  guarantees?  This is an opportunity to (re)explain that functions can be
+  wrong without being malicious.  The next paragraph is a step in that
+  direction, but I think the emphasis on proof gives the wrong idea.}
+
+\vspace{.7\smallskipamount}
+
+Supporting a dynamic notion of adversary is important for functions that
+cannot be statically proven safe, but whose data still needs to be protected
+when they are behaving safely.  For example, consider a function that obeys
+the reasonableness constraints on most inputs, but can be subverted by
+malicious inputs in such a way that it violates the reasonableness
+constraints. In the StkTokens model, we cannot guarantee to protect this
+function's data from its callees even when it called with non-malicious
+inputs.
+
+      % \apt{too abstract now?}
+      % \sna{Maybe, but I'm not sure how much being concrete about it serves us.}
+      % BCP: reads OK to me
+
+\item \citet{SkorstengaardSTKJFP}'s approach to proving safety via fully
+abstract overlay
   semantics relies on the definition of a ``stack-safe-by-construction''
-  abstract machine that is closely tied to the underlying actual
-  machine architecture and and enforcement mechanism.
+  abstract machine that is closely tied to an underlying enforcement
+  mechanism and actual machine architecture.
   %the the abstract machine being architecturally similar to
@@
   While the fundamental approach can no doubt be ported to other architectures and enforcement
-  schemes, each such port requires a new abstract machine definition whose
-  safe-by-construction character must be (informally) verified afresh.
+  schemes, each such port will require a new abstract machine definition whose
+  safety-by-construction property must be checked afresh.
   Our properties are more generic: since we abstract over most of the machine
@@
 
-  Likewise, it would be possible to extend their abstract machine to pass arguments
-  on the stack, or to support callee-save registers. It might even viable to
-  construct a safe-by-construction abstract coroutine machine, but so far this is
-  hypothetical. We implement these scenarios in fact, with simple modifications to
-  the domain model of each one.
-
+\vspace{.7\smallskipamount}
+
+  Conversely, it seems possible to extend \citet{SkorstengaardSTKJFP}'s
+  abstract machine to pass
+  arguments on the stack, or to support callee-save registers. It might even
+  viable to construct a safe-by-construction abstract coroutine machine, but
+  so far this is hypothetical. We implement these scenarios in fact, with
+  simple modifications to the domain model of each one.  \bcp{Don't
+    understand the last sentence}
+
+\ifcomments \item \bcp{Should we also mention our testing stuff?} \fi
 \end{enumerate}
@@
 
-For criterion (1), we believe our property definitions are at least as intuitive
-as theirs.
+For criterion (1), we contend that our property definitions are at least as
+intuitive as theirs, while being phrased in interestingly different terms.
 We use the language of formal security
 properties to apply well-established concepts and formal techniques
-of non-interference. Thus, we define our properties in terms of
+of non-interference, and we define our properties in terms of
 traces of memory and register states, rather than in terms of an
 ad-hoc abstract machine.
-Which approach is better may be a matter of taste, but we believe that our
-alternative has the merit of explicitly relating stack safety to existing
+Which approach is ``better'' is probably a matter of taste, but ours
+has at least the merit of explicitly relating stack safety to existing
 notions of security.
@@
 captures a better definition of safety than LSE; moreover, our observational
-properties capture a richer and more intensional notion of stack safety
+properties capture an even richer and more intensional notion of stack safety
 that can be applied to a broader range of enforcement mechanisms and code features.
-As to criterion (5), we have demonstrated that our approach extends
-to coroutines, which is a first step towards multithreading.
+
+As for criterion (5), we have demonstrated that our approach extends
+to coroutines, a first step towards multithreading.
 Dynamic code generation is more challenging, but StkTokens does not yet
@@
 StkTokens is specifically set up to address criterion (3), while this is not
-a particular goal of our work. We do agree that criterion (2) is important,
-and this remains as future work for us.
-Finally, we believe that having a generic definition of stack
-safety that abstracts over machine details is also
-an important criterion, and here we do better than StkTokens, as described
-in point (ii) above.
+a goal of our work. We do agree that criterion (2) is important;
+demonstrating it for our proposal remains future work.
+
+ Finally, we would propose one more criterion: having a generic definition
+ of stack safety that abstracts over machine details.  Here we do better
+ than StkTokens, as described in point (ii) above.
 
@@
 
-We have evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
+We evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
   Clearing} against our properties, finding that {\em Depth Isolation} does
 indeed enforce stepwise stack safety and therefore also observable stack
-safety. We also found that neither lazy micro-policy actually does enforce
-stack safety, and proposed a fix.
+safety. We also found that neither lazy micro-policy actually enforces
+stack safety, and we proposed a fix.
 
@@
 
-Heap safety, like stack safety, may be framed as a security property
-in the form of
-noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
-be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
-(or ask arthur how!).}\fi. Just
-as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
-characterization of the meaning of (heap) memory safety, in this paper,
-inspired by their work, we aim to do the same for the stack. They are a
-significant inspiration for this work.
+Heap safety, like stack safety, may be framed as a security property using
+the language of
+noninterference~\citep{DBLP:conf/post/AmorimHP18}\iftrue\bcp{Should be
+  spelled ``Azedevo et al.''  We should figure out how to make bibtex do
+  that (or ask arthur how!).}\fi.  Their paper was a significant inspiration
+for the present work, and we believe that the two safety conditions should
+combine smoothly.
 
@@
 
-We have introduced a characterization against which stack safety enforcement
+We introduced a characterization against which stack safety enforcement
 mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
-Our immediate next step is to implement testing of observational properties,
+
+\bcp{Out of date?}Our immediate next step is to implement testing of observational properties,
 both to validate them and to explore how to effectively test properties that do not

### 2021-07-07 submission3/main.tex
cp

@@
 
-\ifcomments \item \bcp{Should we also mention our testing stuff?} \fi
+\ifdraft \item \bcp{Should we also mention our testing stuff?} \fi
 \end{enumerate}
@@
 fail immediately.
-From there there are two natural ways to extend the work. We should thoroughly evaluate
-other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
+
+From there, we see two natural ways to extend the work. First, we should thoroughly evaluate
+other enforcement mechanisms, especially Cheri-based capability
+systems. Second, we can further demonstrate
 the functionality of our definitions by showing that they can be used to prove
@@
 There are several proposals around the use of Cheri capabilities to enforce stack safety,
-ranging from mechanisms that use the standard Cheri hardware (which includes local
+including mechanisms that use the standard Cheri hardware (which includes local
 capabilities) \citep{SkorstengaardLocal},
-to those that propose entirely new types of capabilities, such as linear
-\citep{SkorstengaardSTK}, uninitialized \citep{Georges+21}, lifetime \citep{Tsampas+19}, and monotonic \citep{GeorgesComplete}.
+and others that propose entirely new types of capabilities, such as linear
+\citep{SkorstengaardSTK}, uninitialized \citep{Georges+21}, lifetime
+\citep{Tsampas+19}, and monotonic \citep{GeorgesComplete} capabilities.
 Of these, uninitialized capabilities seem particularly promising---they naturally enforce
-write-before-read behavior, and therefore will enable confidentiality without the need
+write-before-read behavior and therefore will enable confidentiality without the need
 to clear large swaths of the stack. We plan to implement a model of the stack safety
@@
 
-\paragraph*{Program Logic for Stacks}
-
-Stack integrity suggests a reasoning principle in which knowledge about a caller's
-local data can be carried across a call site. We propose to develop this notion
-into a low-level separation logic with a ``call rule'' that preserves propositions
-whose domains are sealed across a call, while allowing those propositions proven
-about the behavior of the callee to apply to the accessible stack.
-
-It is straightforward to imagine that such a rule could be proven sound in a
-system with lockstep stack safety. But there is a real novelty in considering
-what kind of soundness can be proven from a system that obeys only our observational
-properties.
+\paragraph*{A Program Logic for Stacks}
+
+Stack integrity suggests a reasoning principle in which knowledge about a
+caller's local data can be carried across a call site. We aim to develop
+this idea into a low-level separation logic with a ``call rule'' that
+preserves propositions whose domains are sealed across a call, while
+allowing those propositions proven about the behavior of the callee to apply
+to the accessible stack.
+
+It is easy to imagine that such a rule could be proven sound in a system
+with lockstep stack safety. More challenging would be to ask what kind of
+proof rule would be appropriate for a system that obeys only our
+observational properties.
 

### 2021-07-07 submission3/main.tex
Some more polishing

@@
   little uncomfortable about that point: to the extent that applying these
-  mechanisms involves trading off performance and protection}
+  mechanisms involves trading off performance and protection, our properties
+will not apply at all!}
 Thus, it would be useful to have a precise, generic, and formal
@@
 
-In this section we explore how an existing enforcement mechanism, {\em
-  Depth Isolation}~\citep{DBLP:conf/sp/RoesslerD18}, implements
+In this section we explore how an existing enforcement mechanism called {\em
+  Depth Isolation}~\citep{DBLP:conf/sp/RoesslerD18} implements
 the formal stack-safety property described in the previous
-section. In fact, it enforces a slightly stronger, inductive version
+section. In fact, this mechanism enforces a slightly stronger, inductive version
 of that property which checks for integrity and confidentiality violations at
 every step of the callee's execution trace. We validate this claim
-with a property-based testing framework that can quickly uncover
+using a property-based random testing framework to uncover
 inconsistencies between the enforcement mechanism and the stack-safety
@@
 runtime monitor implementing the stack-safety
-micro-policy.\footnote{This code is {\em redacted for review}.} We
+micro-policy.\footnote{This code is {\em redacted for review}.\bcp{Why??}} We
   chose the Coq proof assistant as the setting for our implementation

### 2021-07-08 submission3/main.tex
Arthur's last name

@@
 the language of
-noninterference~\citep{DBLP:conf/post/AmorimHP18}\iftrue\bcp{Should be
-  spelled ``Azedevo et al.''  We should figure out how to make bibtex do
-  that (or ask arthur how!).}\fi.  Their paper was a significant inspiration
+noninterference~\citep{DBLP:conf/post/AmorimHP18}. Their paper was a significant inspiration
 for the present work, and we believe that the two safety conditions should

### 2021-07-08 submission3/main.tex
Response to comments

@@
 etc.---leaving a more abstract characterization to the reader's intuition.
-But these mechanisms can be quite intricate,
-and they often involve trading off protection and performance.\bcp{I'm a
-  little uncomfortable about that point: to the extent that applying these
-  mechanisms involves trading off performance and protection, our properties
-will not apply at all!}
-Thus, it would be useful to have a precise, generic, and formal
-specification for stack
+But these mechanisms can be quite intricate, and the behaviors they prevent
+can vary. Perhaps one conservatively overapproximates the ideal stack safety,
+while another fails to measure up. In order to say for certain,
+we need a precise, generic, and formal specification for stack
 safety, as a basis both for comparing the security claims of different
@@
 
-\bcp{Have we made the diagrams comprehensible when printed in
-  B/W or viewed by a colorblind person?  If so we should say that the colors
-are just extra hints.}%
-Colored boxes indicate an assignment of the address to a
+Labeled boxes indicate an assignment of the address to a
 {\em domain}---a label reflecting a particular set of intended access
-rights.\bcp{should we say here that it is the job of an enforcement mechanism to
-  manage and enforce these access permissions?} Such a domain
+rights. Unless otherwise marked, sequential addresses retain the previous domain,
+as hinted by their colors. Domains are descriptive; it is up to the enforcement
+mechanism to ensure that the system adheres to these permissions. Such a domain
 can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
@@
 (outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
-figures\leo{No outside?}. For our purposes, \(\outside\) will apply to global
-variables and code. Domains are part of the {\em context} of a machine state.
+figures. Not shown are global variables and code, which would be labeled \(\outside\).
+Domains are part of the {\em context} of a machine state.
 (We will introduce some other parts of the context as needed.)
@@
 
-\section{Setting}
+\section{Machines, Enforcement, and Threat Model}
 \label{sec:setup}
 
-\bcp{I am having trouble deciding what this section should be called.
-  Probably this means I don't really understand its purpose...}
-\sna{Propose merging it with next section like so.}
-\leo{Machine, Enforcement, and Threat Model?}
-
-\bcp{And: Needs an intro sentence.}
+Before introducing the model itself, we must situate it in terms of our
+assumptions about machines, enforcement mechanisms, and threats.
 
@@
 passing on the stack, require natural enhancements to our base properties. We
-present such enhancements in Section~\ref{sec:ext}, after we formalize the
-essence of stack safety. Other plausible enhancements
-include:
+present these enhancements, as well as a simple coroutine model,
+in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
+Other plausible enhancements include:
 
@@
 Runtime code generation would be much more challenging to support, as
-it requires dynamic code annotations. Parallelism would require
-fundamental changes to the system model. So, we consider these features
-fully out of scope.\leo{Shouldn't coroutines be mentioned here?}
+it requires dynamic code annotations. Parallelism beyond the current
+coroutine model would require much more substantial changes to the machine
+model. So, we consider these features out of scope.
 
@@
 
-\paragraph*{Policies}\leo{Change this to ``enforcement model'' if we go with that title}
+\paragraph*{Enforcement Model}
 We model the base machine's behavior as a step function over machine states,
@@
 the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
-initial observation, a machine-trace, and an observation trace:\bcp{This
-  would read better and take less space if all three rules were squished
-  onto one line.}
-
-\begin{minipage}[b]{.6\textwidth}
+initial observation, a machine-trace, and an observation trace; the helper is
+then invoked with an initial \(\tau\).
+
+\begin{minipage}[b]{.55\textwidth}
   \judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
@@
 \end{minipage}
-\begin{minipage}[b]{.35\textwidth}
+\begin{minipage}[b]{.22\textwidth}
   \judgment{}
@@
 \end{minipage}
-
-\begin{minipage}{.35\textwidth}
-  Then we start with a \(\tau\) observation:
-\end{minipage}
-\begin{minipage}{.6\textwidth}
+\begin{minipage}[b]{.22\textwidth}
 \judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
@@
 \definition
-For each of the stepwise, deferred, and observational versions of
-stack integrity in the simple model, the equivalent property in the
-passing model\bcp{Is something missing here?}
 %

### 2021-07-08 submission3/main.tex
Filled in more on testing

@@
 
-%% APT: removed because the tag part is too mysterious and the instruction part
-%% is standard.
-%% This is the entry
-%% sequence:
-%% %
-%% \setcounter{pcctr}{1}
-%% \[
-%%   \begin{array}{l|l|l|l}
-%%      & \mathit{Instruction} & \mathit{Tags} & \\
-%%     \hline
-%%     \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
-%%         {Store return address in stack}
-%%     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
-%%         {Increment stack pointer by frame size}
-%%   \end{array}
-%% \]
-%% %
-%% And the exit sequence:
-%% %
-%% \setcounter{pcctr}{1}
-%% \[
-%%   \begin{array}{l|l|l|l}
-%%      & \mathit{Instruction} & \mathit{Tags} & \\
-%%     \hline
-%%     \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
-%%         {Load return address}
-%%     \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
-%%         {Decrement stack pointer by frame size}
-%%     \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
-%%         {Return to caller ($\jalr$)}
-%%   \end{array}
-%% \]
-
-%% \rb{Positioning of the above sequences, relation to running example. The tagging
-%%   of stack frames would be considered part of these sequences.}
-
-%% These sequences also enforce the $\SP$ discipline. When stack frames
-%% are of fixed size, it suffices to increment $\SP$ and tag it with a
-%% dedicated tag in the header sequence, which is later checked after
-%% decrementing $\SP$ during the exit sequence to authorize the
-%% return. The well-formedness of the blessed sequences can be checked
-%% statically.
-
-%% Only programs whose sequences are well-formed are
-%% protected by the micro-policy, although programs are still protected
-%% (by fail-stopping) even if the intended sequences are missing. By
-%% tracking the execution of the blessed sequences in the tag state of
-%% the {\PCname} register, the micro-policy ensures that the entry
-%% sequences of calls declared in the call map are executed correctly.
-
-% \leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
-%   the weeds.  I don't think this level of detail is needed, in the context
-%   of a paper that is mostly about our definitions, not theirs.}
-% Note that in this policy both the entry and exit sequences
-% must write to each element of
-% the frame, which can be quite expensive, especially for programs that
-% allocate large but sparsely populated frames; we return to this point
-% in \cref{sec:lazy}.
-% %
-% For example, the program in \cref{fig:passingsimple} manages the stack
-% in this way. Here, the call pseudo-instruction to {\tt f} in line 5
-% hides a blessed sequence of instructions that allocate space to save
-% its return address to {\tt main} and seal the protected part of the
-% stack before {\tt f}'s body starts executing at 10. Conversely, the
-% return instruction in line 15 restores and jumps to the return address
-% and deallocates its frame space, and unseals the corresponding parts
-% of the stack. The function itself allocates and deallocates its own
-% private frame space in instructions 10 and 14. Additionally, unlike in
-% the simplified presentation of the policy, the caller shares the
-% arguments passed to the callee through the stack, and for this creates
-% and initializes (3--4) and destructs (6) the argument part of the
-% callee's stack frame.
-
-%\rb{Note that we have also considered and implemented more flexible schemes
-%  based on allowing operations above the current activation depth. How do we put
-%  these together?}
-%\rb{Do we want to discuss variations involving frame pointers or more implicit
-%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
-%\rb{Where to talk about the role of the compiler?}
-
-%% Because the access rights of callers subsume those of callees, it is essential
-%% that the micro-policy change tags when the property thinks that a call occurs,
-%% but it is safe for the micro-policy to fail to reset to the caller's tag when the
-%% property thinks a return occurs. (Dually, it is safe for the micro-policy to
-%% change tags even without executing a call point, but not for it to reset tags to
-%% the caller when the property thinks execution is still in the caller).
-
-%% Because the property does not identify return sites, we have no way to require
-%% that ``intended returns'' are tagged correspondingly. Some unintended results
-%% might also be so tagged, but as long as the jump has the effect of a legal
-%% return, it is OK.
-
 \subsection{Validation through Random Testing}
@@
 
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c}
-  Bug & Property Violated & MTTF (ms) \\
-  \hline
-      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
-      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
-      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:bug-table}
-\end{table}
-
-The mean-time-to-failure (MTTF) for various bugs can be found in
-Table~\ref{tab:bug-table}. All experiments were run in a desktop
-machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
-RAM.
+\sna{I am not sure about the MTTF table. Is it as
+  valuable at POPL? Do we need to go into more detail
+  about metholodology if we use it?}
+
+%\begin{table}[]
+%\centering
+%\begin{tabular}{c|c|c}
+%  Bug & Property Violated & MTTF (ms) \\
+%  \hline
+%      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
+%      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
+%      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
+%\end{tabular}
+%\vspace*{1em}
+%\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+%\vspace*{-2em}
+%\label{tab:bug-table}
+%\end{table}
+
+%The mean-time-to-failure (MTTF) for various bugs can be found in
+%Table~\ref{tab:bug-table}. All experiments were run in a desktop
+%machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
+%RAM.
 % \leo{Something like: There is really no baseline to compare this
@@
 %
-The artificially injected bugs focus on weakening the tag propagation
+The artificially injected bugs weaken the tag propagation
 and checking mechanism.
@@
 
+\subsection{Why not Proof?}
+
+There are several ways to evaluate whether an enforcement mechanism enforces
+our properties. Ideally such validation would be done through formal proof.
+Our model is certainly amenable to such a proof, with no fundamental barriers,
+but it would be considerable work to carry out for a full ISA like RISC-V.
+We concentrate instead on testing because it is better aligned with our immediate
+goal of evaluating real enforcement mechanisms for real machine architectures.
+\sna{More to say here?}
+
 \section{Stack Safety for Lazy Enforcement}
@@
 
-We believe that {\em Lazy Per-activation Tagging and Clearing}, in which
-each activation is tagged uniquely, ends up enforcing the combination
-of deferred stepwise stack integrity and stepwise confidentiality.
-This is because tags that persist between calls will not become active again.
+We test {\em Lazy Per-activation Tagging and Clearing}, in which
+each activation is tagged uniquely, and find that it obeys stepwise
+confidentiality and observational integrity. Our testing catches the
+violation in the original {\em Lazy Tagging and Clearing}, as well as
+an equivalent to {\em LOAD\_NO\_CHECK}, which now also violates observational
+integrity.
+
+Per-activation tagging resolves the issue because tags that persist between
+calls will not become active again.
 Our properties have allowed us to reject an enforcement mechanism as
@@
 and can be used to reason about individual programs.
-\citet{SkorstengaardLocal} acknowledge the difficulty of defining
-these concepts precisely\bcp{since  their later papers do include precise
-  definitions, maybe there's no point in emphasizing what they said at the
-  beginning?}; they say
-``while we claim that our calling
-convention enforces control-flow correctness, we do not prove a
-general theorem that shows this, because it is not clear what such a
-theorem should look like,'' noting that the correctness property
-enjoyed by their technique ``is not made very explicit.''
 
@@
 pointers.
-%% The convention operates on a single shared stack and requires that
-%% protected components avoid compromising their own security by following certain
-%% simple rules---like not leaking their private capability seals. More
-%% precisely but
 Informally, local state encapsulation (LSE) is defined as
@@
 properties hold, and adversary functions, which may violate those
-constraints but are not guaranteed safety.  \bcp{Readers may wonder: why
-  isn't this reasonable?  after all, if a
-  function is adversarial, why do we care whether it gets any safety
-  guarantees?  This is an opportunity to (re)explain that functions can be
-  wrong without being malicious.  The next paragraph is a step in that
-  direction, but I think the emphasis on proof gives the wrong idea.}
-
-\vspace{.7\smallskipamount}
-
-Supporting a dynamic notion of adversary is important for functions that
-cannot be statically proven safe, but whose data still needs to be protected
-when they are behaving safely.  For example, consider a function that obeys
-the reasonableness constraints on most inputs, but can be subverted by
-malicious inputs in such a way that it violates the reasonableness
-constraints. In the StkTokens model, we cannot guarantee to protect this
-function's data from its callees even when it called with non-malicious
-inputs.
-
-      % \apt{too abstract now?}
-      % \sna{Maybe, but I'm not sure how much being concrete about it serves us.}
-      % BCP: reads OK to me
+constraints but are not guaranteed safety. This static division
+is insufficient when we consider buggy, but not inherently adversarial,
+code --- one can imagine a function that, when called with the wrong
+inputs, behaves unreasonably, but under most circumstances can be
+considered reasonable. \citet{SkirstengaardSTKJFP} would be forced
+to classify such a function as an adversary, and would not guarantee it
+protection. We argue that if individual function activations are well-behaved,
+they should be protected.
 
@@
   Conversely, it seems possible to extend \citet{SkorstengaardSTKJFP}'s
-  abstract machine to pass
-  arguments on the stack, or to support callee-save registers. It might even
-  viable to construct a safe-by-construction abstract coroutine machine, but
-  so far this is hypothetical. We implement these scenarios in fact, with
-  simple modifications to the domain model of each one.  \bcp{Don't
-    understand the last sentence}
+  abstract machine to pass arguments on the stack, or to support callee-save
+  registers. It might even be viable to construct a safe-by-construction abstract
+  coroutine machine in the same style, but this is hypothetical.
+  We have implemented these scenarios with
+  simple modifications to the domain model.
 
@@
 
-\bcp{Out of date?}Our immediate next step is to implement testing of observational properties,
-both to validate them and to explore how to effectively test properties that do not
-fail immediately.
-
-From there, we see two natural ways to extend the work. First, we should thoroughly evaluate
+From here, we see two natural ways to extend the work. First, we should thoroughly evaluate
 other enforcement mechanisms, especially Cheri-based capability

### 2021-07-08 submission3/main.tex
Some little things

@@
 The protections offered by such mechanisms are commonly described in terms
-of concrete examples of attacks that they can prevent---corruption
-of return addresses, buffer overflows, use of uninitialized variables,
-etc.---leaving a more abstract characterization to the reader's intuition.
-But these mechanisms can be quite intricate, and the behaviors they prevent
-can vary. Perhaps one conservatively overapproximates the ideal stack safety,
-while another fails to measure up. In order to say for certain,
+of concrete examples of attacks that they can prevent---corruption of return
+addresses, buffer overflows, use of uninitialized variables, etc.---leaving
+a more abstract characterization to the reader's intuition.  But these
+mechanisms can be quite intricate, and the behaviors they aim to prevent are
+subtle and varied.  It can be hard to guess, based only on informal
+intuitions, whether a given mechanism actually blocks all potential
+attacks---or, conversely, whether it is overly conservative and disallows
+system behaviors that are actually safe.
+To settle such questions with confidence,
 we need a precise, generic, and formal specification for stack
 safety, as a basis both for comparing the security claims of different
-enforcement techniques and for validating more rigorously that these claims
+enforcement techniques and for validating that these claims
 are met by particular implementations.
@@
 rights. Unless otherwise marked, sequential addresses retain the previous domain,
-as hinted by their colors. Domains are descriptive; it is up to the enforcement
+as hinted by their colors. Domains are descriptive\bcp{??}; it is up to the enforcement
 mechanism to ensure that the system adheres to these permissions. Such a domain
@@
 figures. Not shown are global variables and code, which would be labeled \(\outside\).
-Domains are part of the {\em context} of a machine state.
+Domains form part of the {\em context} associated with a machine state.
 (We will introduce some other parts of the context as needed.)
@@
 
-Before introducing the model itself, we must situate it in terms of our
-assumptions about machines, enforcement mechanisms, and threats.
+Before introducing the model itself, let us set the context by stating some
+assumptions we are making about machines, enforcement mechanisms, and
+potential threats.
 
@@
 Runtime code generation would be much more challenging to support, as
-it requires dynamic code annotations. Parallelism beyond the current
-coroutine model would require much more substantial changes to the machine
-model. So, we consider these features out of scope.
+it requires dynamic code annotations. Similarly, parallelism beyond the current
+coroutine model would require more substantial changes to the machine
+model. We consider these features out of scope for now.
 

### 2021-07-08 submission3/main.tex
citation typo
pass on abstract and intro

@@
 memory, but these mechanisms typically lack precise specifications,
-relying instead on informal descriptions and examples of bad
+relying instead on informal descriptions and examples of the bad
 behaviors that they prevent.
 
-We propose a formal characterization
-of stack safety, formulated with concepts from language-based security: a
+We propose a \apt{``generic''?} formal characterization
+of stack safety based on concepts from language-based security: a
 combination of an integrity property (``the private
@@
 caller's private state''), and a well-bracketedness property (``each
-callee returns control to its immediate caller'').
+callee returns control to its immediate caller''). \apt{Do we really
+  want to include WBCF in the abstract?} 
 % These properties are sensible even in the presence of
@@
 
-We use these properties to validate the stack-safety {\em micro-policies}
+We use these properties to validate \apt{``evaluate'' ?} the stack-safety {\em micro-policies}
 proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
@@
 simple ``stepwise'' variant of our properties and correctly detects several
-broken variants, and that (a repaired version of) their more performant
-``lazy'' micro-policy corresponds to a slightly weaker and more extensional
+broken variants\apt{last phrase is grammatically disconnected}, and that (a repaired version of) their more performant
+``lazy'' micro-policy corresponds to\apt{``enforces''?} a slightly weaker and more extensional
 observational variant of our properties.
@@
 %\leo{Examples at a high level}: such as---does not overwrite return addresses, or ...
+\apt{This makes us seem very reactive to Skst.  Can we reorder things to give our story first, and
+  compare to Skst afterwards?}
 The most abstract characterization of stack safety that we are aware of is
@@
 In the present work, we also describe the protection of both stack data (like LSE) and
-control flow (like WBCF).  Our technical formulation of WBCF is similar to
-theirs, but our treatment of stack data protection is rather different,
+control flow (like WBCF).  Our technical formulation of WBCF is similar to that of
+Skorstengaard et al., but our treatment of stack data protection is rather different,
 as is our approach to the definition itself. Instead of defining stack safety implicitly,
@@
 \emph{explicit} definition using the technical framework of language-based
-security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
-  survey?}\fi, phrased as the conjunction of
+security~\citep{sabelfeld2003language}, phrased as the conjunction of
 \emph{stack integrity} and \emph{stack confidentiality}.
@@
 that Roessler and Dehon's micro-policies correctly abort the ones that
-attempt to violate one of our properties; furthermore, we
+attempt to violate one of our properties. Furthermore, we
 %
 check that the testing framework is able to generate counterexamples
-that violate our properties but are not halted by incorrect
-enforcement variants (both variants that we accidentally created
+that violate our properties but are \emph{not} halted by incorrect
+enforcement variants---both variants that we accidentally created
 during our re-implementation of the micro-policy and ones that we
 intentionally crafted to be broken in order to increase our confidence
-in testing and the enforcement mechanism itself).
+in testing and the enforcement mechanism itself.
 %introduce several
@@
 We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
-which memory cells within each stack frame are tagged with the identity of
+which memory cells within each stack frame are tagged with the identity\apt{``depth''?} of
 the function activation that owns the frame and access to those locations is
-then permitted only when that activation is currently executing, validates our
-stepwise properties. On the other hand, we reason that \emph{Lazy Tagging and Clearing}
+then permitted only when that activation\apt{``an actiation at that depth''?} is currently executing, validates our
+stepwise properties. On the other hand, we reason\apt{but now we test?}  that \emph{Lazy Tagging and Clearing}
 violates the temporal aspect of confidentiality in
@@
 propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
-confidentiality, albeit at some performance cost.
+confidentiality, albeit at some performance cost.\apt{Do we test?}
 %
@@
 to a fixed region of memory.  Though simplistic, this extension is a step
-toward full-blown concurrency and demonstrates that our approach is flexible
+toward full-blown multithreading and demonstrates that our approach is flexible
 enough to handle more sophisticated styles of control flow.
@@
 In the rest of the paper, we begin with an informal overview of our key ideas via
-examples (\cref{sec:running-example}), followed by definitions of some key
-concepts (\cref{sec:setup}) and a discussion of the threat
-model we consider (\cref{sec:threat}). We then offer the following contributions:
+examples (\cref{sec:running-example}), followed by a description of our key
+assumptions about the machine environment and threat model (\cref{sec:setup}). 
+We then offer the following contributions:
 
@@
 property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18}, and extend the generation by
+of~\citet{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
 execution techniques of \citet{TestingNI:ICFP} to handle nested
@@
 property that characterizes a refinement of Roessler and DeHon's more
-performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
+performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).\apt{can this be more precise?}
 \item We show that the definition can be
-extended to argument passing on the stack and to a simple coroutine system (\cref{sec:ext}).
+extended to argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
 \end{itemize}
@@
 
-Some calling convention choices, such as callee-save registers and argument
+Some calling convention choices, such as callee-saves registers and argument
 passing on the stack, require natural enhancements to our base properties. We
@@
 
-Intuitively, a caller expects a callee-saved register to be unchanged on
+Intuitively, a caller expects a callee-saves register to be unchanged on
 return: it
-may change during the call, as long as it is restored. We similarly expect a callee-saved
+may change during the call, as long as it is restored. We similarly expect a callee-saves
 register to be secret from the callee, except when it is copied to the stack! Both scenarios
@@
 inputs, behaves unreasonably, but under most circumstances can be
-considered reasonable. \citet{SkirstengaardSTKJFP} would be forced
+considered reasonable. \citet{SkorstengaardSTKJFP} would be forced
 to classify such a function as an adversary, and would not guarantee it
@@
   Conversely, it seems possible to extend \citet{SkorstengaardSTKJFP}'s
-  abstract machine to pass arguments on the stack, or to support callee-save
+  abstract machine to pass arguments on the stack, or to support callee-saves
   registers. It might even be viable to construct a safe-by-construction abstract

### 2021-07-08 submission3/main.tex
pass through 3

@@
 function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
-variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
+variable {\tt x}. (Of course, this is illegal C, because the write in {\tt f} triggers
+undefined behavior, but it is characteristic of what buggy or malicious code might do.) 
+In the initial memory layout the entire stack is {\em unsealed},
 meaning that it is eligible to be written to. At the call from {\tt main} to
@@
 
-Meanwhile, confidentiality intuitively states that a callee is insensitive to
+Dually, confidentiality intuitively states that a callee is insensitive to
 the context of its call (except for its arguments and any global
@@
 understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
-be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
+be hidden from {\tt f}. (We show the evolving stack memory  of the variant in a second parallel column.) At instruction 12, we note that the callee's first variable, {\tt y}, has
 changed to 5 in both the primary trace and the variant. The primary and variant agree on
@@
 return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
-up being called a third time, at which point it finally returns to instruction 5.
+up being called a third time, at which point it finally returns to instruction 5.\apt{The line numbers are wrong.}
 
@@
 Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.
+violates the integrity property.\apt{Ditto}
 
@@
 
-Before introducing the model itself, let us set the context by stating some
+Before introducing the model\apt{what does this really mean?}  itself, we set the context by stating some
 assumptions we are making about machines, enforcement mechanisms, and
@@
 assumptions about the concrete machine that instantiates it. Without changing the
-properties, we can support any register
+properties, we can support\apt{what does this really mean?} any register
 machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
@@
 Some calling convention choices, such as callee-saves registers and argument
-passing on the stack, require natural enhancements to our base properties. We
+passing on the stack, require natural enhancements to our base properties.\apt{
+  At this point it is unclear what these ``enhancements'' might be.}
+We
 present these enhancements, as well as a simple coroutine model,
 in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
-Other plausible enhancements include:
+Other plausible enhancements include:\apt{Better to move the ``what would
+  be needed'' parts to the future work section -- they are mysterious here.}
 
@@
 Runtime code generation would be much more challenging to support, as
-it requires dynamic code annotations. Similarly, parallelism beyond the current
-coroutine model would require more substantial changes to the machine
+it requires dynamic generation of code annotations. Similarly, true concurrency
+would require more substantial changes to the machine
 model. We consider these features out of scope for now.
@@
 instructions either before or after the jump.  Similarly, there may be no
-``return'' macro-instruction, but rather a ``return sequence'' involving an
+``return'' instruction, but rather a ``return sequence'' involving an
 indirect branch at some point.  Thus, the same instruction opcode may be
@@
 Annotations connect the enforcement mechanism to the property: we don't
-protect unmarked calls, because a mechanism's starting conditions might
+protect unmarked calls, because a mechanism's correct functioning might
 depend on knowing where calls are.
@@
 Returns, on the other hand, need not be annotated: we can tell when a callee
-has returned control to the caller when... we see control return to the
+has returned control to the caller when...we see control return to the
 caller! We don't require the callee's cooperation at all.
@@
 \paragraph*{Enforcement Model}
+\apt{What are our ambititions for genericity over enforcement?}
+
 We model the base machine's behavior as a step function over machine states,
@@
 enforcement mechanisms that keep a significant amount of data separate from the
-primary execution data to inform security. Obviously tag-based systems do this,
-with tags and extra policy state separated from and untouchable by application code.
-Another significant example is the shadow stack, where a second stack
+primary execution data to inform security. Obviously tag-based systems do this\apt{At this
+  point, we barely know what tags are},
+with tags and extra policy state separated from, and untouchable by, application code.
+Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
 mirrors the real one, and a return may only occur if both stacks agree on the
@@
 The separation is smaller in protection mechanisms that interact more directly
-with application code. Some mechanisms might exhibit failstop behavior, but have
+with application code.\apt{Vague. Mention capabilities?}  Some mechanisms might exhibit failstop behavior, but have
 no protected policy state; others might never even failstop. Then the policy-enhanced
 machine is simply identical to the original. This is the case for software-based
-techniques.
+techniques.\apt{Even if they use a shadow stack?}
 
@@
 an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways. We do rule out some
+may be viewed in both ways.
+\apt{everything down to here seems like it should go earlier -- in informal section
+  or even intro.}
+We do rule out some
 scenarios with a ``well-formedness'' criterion on initial states, ensuring that
@@
 We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. Hardware and timing attacks are out of scope, as are
+arbitrary machine code. \apt{This should be the lead sentence. But how does is
+  it consistent with the previous para on well-formedness?}
+  Hardware and timing attacks are out of scope, as are
 are attacks on the heap---though we believe that the work of

### 2021-07-08 submission3/main.tex
pass up through 5

@@
 
-We now define our machine model precisely. To make our definition of stack safety
+We now define our machine model \apt{and enforcement model?} precisely. To make our definition of stack safety
 as generic as possible, this section proposes an abstract interface to a
@@
 individual calls from a trace by taking subtraces in which the stack is
-at or below a given depth. If \(f\) is still a predicate on states, then
+at or below\apt{``above''?}  a given depth. If \(f\) is still a predicate on states, then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
@@
 
-Each system component is assigned to a {\em domain}, which may be {\em Outside}
-of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
+Each system component is assigned to a {\em domain}, which may be {\em outside}
+of the stack entirely, {\em unsealed} and therefore accessible, or {\em sealed}
 at a particular depth. Each caller registers a {\em target}
@@
          {\(\underscore,\context \stepsto_C \context\)}
-
+\apt{typesetting problem above}
+
+\noindent
 When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed. We now state our properties.
+call stack we are and which components are sealed.
+
+We now state our properties.
 The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
@@
 \definition
-Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
+The \emph{protected set} of a context, \(\prot(\domMap,\_)\), is the
 set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
@@
 
-\definition \(\MPCT^*\) is the \emph{call set}, defined as
+\definition The \emph{call set} \(\MPCT^*\) of a trace \(\MPCT\) \apt{right?} is
 \[\begin{aligned}
@@
 
+\apt{intuition for what the call set is?}
 We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
@@
 \definition
-Let the secret set of a context, \(\secret(\domMap,\rts)\), be the
+The \emph{secret set} of a context, written \(\secret(\domMap,\rts)\), is the
 set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).

### 2021-07-08 submission3/main.tex
More readable domain labels
Uniform, longer descriptions. These work well everywhere except in
Figure 5, where there is now a bit of whitespace between memory cells.
How to fix this?

@@
 
-We propose a \apt{``generic''?} formal characterization
+We propose a generic formal characterization
 of stack safety based on concepts from language-based security: a
@@
 callee returns control to its immediate caller''). \apt{Do we really
-  want to include WBCF in the abstract?} 
+  want to include WBCF in the abstract?}\bcp{I would delete.}
 % These properties are sensible even in the presence of
@@
 
-We use these properties to validate \apt{``evaluate'' ?} the stack-safety {\em micro-policies}
-proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
-property-based random testing) that Roessler and Dehon's ``eager''
+We use these properties to validate \apt{``evaluate'' ?}\bcp{I like
+  ``validate'' here.} the stack-safety {\em micro-policies} proposed
+by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we use property-based
+random testing to check (1) that Roessler and Dehon's ``eager''
 micro-policy, which catches violations as early as possible, enforces a
 simple ``stepwise'' variant of our properties and correctly detects several
-broken variants\apt{last phrase is grammatically disconnected}, and that (a repaired version of) their more performant
-``lazy'' micro-policy corresponds to\apt{``enforces''?} a slightly weaker and more extensional
+broken variants, and (2) that (a repaired version of) their more performant
+``lazy'' micro-policy enforces a slightly weaker and more extensional
 observational variant of our properties.
@@
 \apt{This makes us seem very reactive to Skst.  Can we reorder things to give our story first, and
-  compare to Skst afterwards?}
+  compare to Skst afterwards?}\bcp{+1}
 The most abstract characterization of stack safety that we are aware of is
@@
 \newcommand{\retptrc}{black}
+\newcommand{\sealdesc}[1]{Seal(#1)}
+\newcommand{\unsealdesc}{Unseal}
+\newcommand{\retptrdesc}{RetPtr}
+\newcommand{\passdesc}[2]{Pass(#1,#2)}
 
@@
 \memoryaddrs{8em}
-\memory{3}{\unsealc}[U]
+\memory{3}{\unsealc}[\unsealdesc]
 ~$\cdots$
@@
 \memoryaddrs{16em}
-\memory{1}{\mainsealc}[S(0)]%
-\memory{1}{\retptrc}[RetPtr]%
-\memory{1}{\unsealc}[U]
+\memory{1}{\mainsealc}[\sealdesc{0}]%
+\memory{1}{\retptrc}[\retptrdesc]%
+\memory{1}{\unsealc}[\unsealdesc]
 ~$\cdots$
@@
       \memoryaddrs{12em}
-      \memory{1}{\mainsealc}[S(0)]%
-      \memory{1}{\unsealc}[RetPtr]%
+      \memory{1}{\mainsealc}[\sealdesc{0}]%
+      \memory{1}{\unsealc}[\retptrdesc]%
       \memory{1}{\unsealc}%
@@
   \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[S(0)]%
-  \memory{1}{\mainpassc}[P(0,1)]%
-  \memory{1}{\retptrc}[RetPtr]%
+  \memory{1}{\mainsealc}[\sealdesc{0}]%
+  \memory{1}{\mainpassc}[\passdesc{0}{1}]%
+  \memory{1}{\retptrc}[\retptrdesc]%
   \memory{1}{\unsealc}
@@
       \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{S(0)}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
       \memory{1}{\retptrc}
@@
       \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{S(0)}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
+      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
+      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
       \memory{1}{\retptrc}
@@
   simple modifications to the domain model.
-
-\ifdraft \item \bcp{Should we also mention our testing stuff?} \fi
 \end{enumerate}
@@
 
-For criterion (1), we contend that our property definitions are at least as
+\bcp{We could squish all these little paragraphs into one if we are short of
+space.}
+
+For criterion (1), intuitiveness, we contend that our property definitions are at least as
 intuitive as theirs, while being phrased in interestingly different terms.
@@
 
-With respect to criterion (4), we believe that our support
+With respect to criterion (4), capturing programmers' informal
+understanding of stack safety, we believe that our support
 for dynamically changing notion of adversaries, as outlined in point (i) above,
@@
 
-As for criterion (5), we have demonstrated that our approach extends
-to coroutines, a first step towards multithreading.
-Dynamic code generation is more challenging, but StkTokens does not yet
-handle this either.
-StkTokens is specifically set up to address criterion (3), while this is not
-a goal of our work. We do agree that criterion (2) is important;
-demonstrating it for our proposal remains future work.
+As for criterion (5), broad applicability, we have demonstrated that our approach extends to
+coroutines, a first step towards multithreading.  Dynamic code generation is
+more challenging, but StkTokens does not yet handle this either.
+
+StkTokens
+is specifically set up to address criterion (3), reusability in secure
+compilation chains, while this is not a goal of
+our work.
+
+We agree that criterion (2), usefulness of the safety criterion
+for carrying out formal proofs (about individual programs), is important;
+demonstrating it for our proposal remains future work.  Instead, we have
+chosen to focus on property-based random testing as a methodology for
+validating enforcement mechanisms.
 

### 2021-07-08 submission3/main.tex
Fix Azevedo again; add testing to comparison with StkTokens

--- 2021-07-08 submission3/main.tex
@@
 memory, but these mechanisms typically lack precise specifications,
-relying instead on informal descriptions and examples of the bad
+relying instead on informal descriptions and examples of bad
 behaviors that they prevent.
 
-We propose a \apt{``generic''?} formal characterization
-of stack safety based on concepts from language-based security: a
+We propose a formal characterization
+of stack safety, formulated with concepts from language-based security: a
 combination of an integrity property (``the private
@@
 caller's private state''), and a well-bracketedness property (``each
-callee returns control to its immediate caller''). \apt{Do we really
-  want to include WBCF in the abstract?} 
+callee returns control to its immediate caller'').
 % These properties are sensible even in the presence of
@@
 
-We use these properties to validate \apt{``evaluate'' ?} the stack-safety {\em micro-policies}
+We use these properties to validate the stack-safety {\em micro-policies}
 proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
@@
 simple ``stepwise'' variant of our properties and correctly detects several
-broken variants\apt{last phrase is grammatically disconnected}, and that (a repaired version of) their more performant
-``lazy'' micro-policy corresponds to\apt{``enforces''?} a slightly weaker and more extensional
+broken variants, and that (a repaired version of) their more performant
+``lazy'' micro-policy corresponds to a slightly weaker and more extensional
 observational variant of our properties.
@@
 %\leo{Examples at a high level}: such as---does not overwrite return addresses, or ...
-\apt{This makes us seem very reactive to Skst.  Can we reorder things to give our story first, and
-  compare to Skst afterwards?}
 The most abstract characterization of stack safety that we are aware of is
@@
 In the present work, we also describe the protection of both stack data (like LSE) and
-control flow (like WBCF).  Our technical formulation of WBCF is similar to that of
-Skorstengaard et al., but our treatment of stack data protection is rather different,
+control flow (like WBCF).  Our technical formulation of WBCF is similar to
+theirs, but our treatment of stack data protection is rather different,
 as is our approach to the definition itself. Instead of defining stack safety implicitly,
@@
 \emph{explicit} definition using the technical framework of language-based
-security~\citep{sabelfeld2003language}, phrased as the conjunction of
+security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
+  survey?}\fi, phrased as the conjunction of
 \emph{stack integrity} and \emph{stack confidentiality}.
@@
 that Roessler and Dehon's micro-policies correctly abort the ones that
-attempt to violate one of our properties. Furthermore, we
+attempt to violate one of our properties; furthermore, we
 %
 check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-enforcement variants---both variants that we accidentally created
+that violate our properties but are not halted by incorrect
+enforcement variants (both variants that we accidentally created
 during our re-implementation of the micro-policy and ones that we
 intentionally crafted to be broken in order to increase our confidence
-in testing and the enforcement mechanism itself.
+in testing and the enforcement mechanism itself).
 %introduce several
@@
 We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
-which memory cells within each stack frame are tagged with the identity\apt{``depth''?} of
+which memory cells within each stack frame are tagged with the identity of
 the function activation that owns the frame and access to those locations is
-then permitted only when that activation\apt{``an actiation at that depth''?} is currently executing, validates our
-stepwise properties. On the other hand, we reason\apt{but now we test?}  that \emph{Lazy Tagging and Clearing}
+then permitted only when that activation is currently executing, validates our
+stepwise properties. On the other hand, we reason that \emph{Lazy Tagging and Clearing}
 violates the temporal aspect of confidentiality in
@@
 propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
-confidentiality, albeit at some performance cost.\apt{Do we test?}
+confidentiality, albeit at some performance cost.
 %
@@
 to a fixed region of memory.  Though simplistic, this extension is a step
-toward full-blown multithreading and demonstrates that our approach is flexible
+toward full-blown concurrency and demonstrates that our approach is flexible
 enough to handle more sophisticated styles of control flow.
@@
 In the rest of the paper, we begin with an informal overview of our key ideas via
-examples (\cref{sec:running-example}), followed by a description of our key
-assumptions about the machine environment and threat model (\cref{sec:setup}). 
-We then offer the following contributions:
+examples (\cref{sec:running-example}), followed by definitions of some key
+concepts (\cref{sec:setup}) and a discussion of the threat
+model we consider (\cref{sec:threat}). We then offer the following contributions:
 
@@
 property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
+of~\citet{DBLP:conf/sp/RoesslerD18}, and extend the generation by
 execution techniques of \citet{TestingNI:ICFP} to handle nested
@@
 property that characterizes a refinement of Roessler and DeHon's more
-performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).\apt{can this be more precise?}
+performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
 \item We show that the definition can be
-extended to argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
+extended to argument passing on the stack and to a simple coroutine system (\cref{sec:ext}).
 \end{itemize}
@@
 function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
-variable {\tt x}. (Of course, this is illegal C, because the write in {\tt f} triggers
-undefined behavior, but it is characteristic of what buggy or malicious code might do.) 
-In the initial memory layout the entire stack is {\em unsealed},
+variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
 meaning that it is eligible to be written to. At the call from {\tt main} to
@@
 
-Dually, confidentiality intuitively states that a callee is insensitive to
+Meanwhile, confidentiality intuitively states that a callee is insensitive to
 the context of its call (except for its arguments and any global
@@
 understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
-be hidden from {\tt f}. (We show the evolving stack memory  of the variant in a second parallel column.) At instruction 12, we note that the callee's first variable, {\tt y}, has
+be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
 changed to 5 in both the primary trace and the variant. The primary and variant agree on
@@
 return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
-up being called a third time, at which point it finally returns to instruction 5.\apt{The line numbers are wrong.}
+up being called a third time, at which point it finally returns to instruction 5.
 
@@
 Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.\apt{Ditto}
+violates the integrity property.
 
@@
 
-Before introducing the model\apt{what does this really mean?}  itself, we set the context by stating some
+Before introducing the model itself, let us set the context by stating some
 assumptions we are making about machines, enforcement mechanisms, and
@@
 assumptions about the concrete machine that instantiates it. Without changing the
-properties, we can support\apt{what does this really mean?} any register
+properties, we can support any register
 machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
@@
 
-Some calling convention choices, such as callee-saves registers and argument
-passing on the stack, require natural enhancements to our base properties.\apt{
-  At this point it is unclear what these ``enhancements'' might be.}
-We
+Some calling convention choices, such as callee-save registers and argument
+passing on the stack, require natural enhancements to our base properties. We
 present these enhancements, as well as a simple coroutine model,
 in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
-Other plausible enhancements include:\apt{Better to move the ``what would
-  be needed'' parts to the future work section -- they are mysterious here.}
+Other plausible enhancements include:
 
@@
 Runtime code generation would be much more challenging to support, as
-it requires dynamic generation of code annotations. Similarly, true concurrency
-would require more substantial changes to the machine
+it requires dynamic code annotations. Similarly, parallelism beyond the current
+coroutine model would require more substantial changes to the machine
 model. We consider these features out of scope for now.
@@
 instructions either before or after the jump.  Similarly, there may be no
-``return'' instruction, but rather a ``return sequence'' involving an
+``return'' macro-instruction, but rather a ``return sequence'' involving an
 indirect branch at some point.  Thus, the same instruction opcode may be
@@
 Annotations connect the enforcement mechanism to the property: we don't
-protect unmarked calls, because a mechanism's correct functioning might
+protect unmarked calls, because a mechanism's starting conditions might
 depend on knowing where calls are.
@@
 Returns, on the other hand, need not be annotated: we can tell when a callee
-has returned control to the caller when...we see control return to the
+has returned control to the caller when... we see control return to the
 caller! We don't require the callee's cooperation at all.
@@
 \paragraph*{Enforcement Model}
-\apt{What are our ambititions for genericity over enforcement?}
-
 We model the base machine's behavior as a step function over machine states,
@@
 enforcement mechanisms that keep a significant amount of data separate from the
-primary execution data to inform security. Obviously tag-based systems do this\apt{At this
-  point, we barely know what tags are},
-with tags and extra policy state separated from, and untouchable by, application code.
-Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
+primary execution data to inform security. Obviously tag-based systems do this,
+with tags and extra policy state separated from and untouchable by application code.
+Another significant example is the shadow stack, where a second stack
 mirrors the real one, and a return may only occur if both stacks agree on the
@@
 The separation is smaller in protection mechanisms that interact more directly
-with application code.\apt{Vague. Mention capabilities?}  Some mechanisms might exhibit failstop behavior, but have
+with application code. Some mechanisms might exhibit failstop behavior, but have
 no protected policy state; others might never even failstop. Then the policy-enhanced
 machine is simply identical to the original. This is the case for software-based
-techniques.\apt{Even if they use a shadow stack?}
+techniques.
 
@@
 an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways.
-\apt{everything down to here seems like it should go earlier -- in informal section
-  or even intro.}
-We do rule out some
+may be viewed in both ways. We do rule out some
 scenarios with a ``well-formedness'' criterion on initial states, ensuring that
@@
 We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. \apt{This should be the lead sentence. But how does is
-  it consistent with the previous para on well-formedness?}
-  Hardware and timing attacks are out of scope, as are
+arbitrary machine code. Hardware and timing attacks are out of scope, as are
 are attacks on the heap---though we believe that the work of
@@
 
-Intuitively, a caller expects a callee-saves register to be unchanged on
+Intuitively, a caller expects a callee-saved register to be unchanged on
 return: it
-may change during the call, as long as it is restored. We similarly expect a callee-saves
+may change during the call, as long as it is restored. We similarly expect a callee-saved
 register to be secret from the callee, except when it is copied to the stack! Both scenarios
@@
   Conversely, it seems possible to extend \citet{SkorstengaardSTKJFP}'s
-  abstract machine to pass arguments on the stack, or to support callee-saves
+  abstract machine to pass arguments on the stack, or to support callee-save
   registers. It might even be viable to construct a safe-by-construction abstract
@@
 
-As for criterion (5), we have demonstrated that our approach extends
-to coroutines, a first step towards multithreading.
-Dynamic code generation is more challenging, but StkTokens does not yet
-handle this either.
-StkTokens is specifically set up to address criterion (3), while this is not
-a goal of our work. We do agree that criterion (2) is important;
-demonstrating it for our proposal remains future work.
+As for criterion (5), we have demonstrated that our approach extends to
+coroutines, a first step towards multithreading.  Dynamic code generation is
+more challenging, but StkTokens does not yet handle this either.  StkTokens
+is specifically set up to address criterion (3), while this is not a goal of
+our work. We agree that criterion (2), usefulness of the safety criterion
+for carrying out formal proofs (about individual programs), is important;
+demonstrating it for our proposal remains future work.  Instead, we have
+chosen to focus on property-based random testing as a methodology for
+validating enforcement mechanisms.
 

### 2021-07-08 submission3/main.tex
Gloss and reformat criteria

@@
 
-For criterion (1), we contend that our property definitions are at least as
+For criterion (1), intuitiveness, we contend that our property definitions are at least as
 intuitive as theirs, while being phrased in interestingly different terms.
@@
 
-As for criterion (5), we have demonstrated that our approach extends to
+As for criterion (5), broad applicability, we have demonstrated that our approach extends to
 coroutines, a first step towards multithreading.  Dynamic code generation is
-more challenging, but StkTokens does not yet handle this either.  StkTokens
-is specifically set up to address criterion (3), while this is not a goal of
-our work. We agree that criterion (2), usefulness of the safety criterion
+more challenging, but StkTokens does not yet handle this either.
+
+StkTokens
+is specifically set up to address criterion (3), reusability in secure
+compilation chains, while this is not a goal of
+our work.
+
+We agree that criterion (2), usefulness of the safety criterion
 for carrying out formal proofs (about individual programs), is important;

### 2021-07-08 submission3/main.tex
Abstract

@@
 callee returns control to its immediate caller''). \apt{Do we really
-  want to include WBCF in the abstract?}
+  want to include WBCF in the abstract?}\bcp{I would delete.}
 % These properties are sensible even in the presence of
@@
 
+\bcp{We could squish all these little paragraphs into one if we are short of
+space.}
+
 For criterion (1), intuitiveness, we contend that our property definitions are at least as

### 2021-07-08 submission3/main.tex
starting on 6

--- 2021-07-08 submission3/main.tex
@@
 of that property which checks for integrity and confidentiality violations at
-every step of the callee's execution trace. We validate this claim
+every step of the callee's execution trace. We validate this claim \apt{which claim: that it enforces
+  a \emph{stronger} version?}
 using a property-based random testing framework to uncover
-inconsistencies between the enforcement mechanism and the stack-safety
+inconsistencies \apt{but are there any?} between the enforcement mechanism and the stack-safety
 property.
@@
 \citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
-with ownership information associated to each stack frame (and to each
+with ownership information \apt{based on depth!} associated to each stack frame (and to each
 separate object inside that frame, a more finely grained access
@@
   simple modifications to the domain model.
+
+\ifdraft \item \bcp{Should we also mention our testing stuff?} \fi
 \end{enumerate}
@@
 
-As for criterion (5), we have demonstrated that our approach extends to
-coroutines, a first step towards multithreading.  Dynamic code generation is
-more challenging, but StkTokens does not yet handle this either.  StkTokens
-is specifically set up to address criterion (3), while this is not a goal of
-our work. We agree that criterion (2), usefulness of the safety criterion
-for carrying out formal proofs (about individual programs), is important;
-demonstrating it for our proposal remains future work.  Instead, we have
-chosen to focus on property-based random testing as a methodology for
-validating enforcement mechanisms.
+As for criterion (5), we have demonstrated that our approach extends
+to coroutines, a first step towards multithreading.
+Dynamic code generation is more challenging, but StkTokens does not yet
+handle this either.
+StkTokens is specifically set up to address criterion (3), while this is not
+a goal of our work. We do agree that criterion (2) is important;
+demonstrating it for our proposal remains future work.
 

### 2021-07-08 submission3/main.tex
Adjusted section 1

--- 2021-07-08 submission3/main.tex
@@
 \begin{abstract}
-%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
-%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
-%  to leave it out of the title but hit it strongly in the abstract.}%
+
 What does ``stack safety'' mean, exactly? The phrase is associated with a
@@
 
-We propose a \apt{``generic''?} formal characterization
+We propose a generic, formal characterization
 of stack safety based on concepts from language-based security: a
@@
 state in each caller's stack frame is held invariant by the callee''),
-a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state''), and a well-bracketedness property (``each
-callee returns control to its immediate caller''). \apt{Do we really
-  want to include WBCF in the abstract?} 
-% These properties are sensible even in the presence of
-% abnormal control flow.\apt{Is this important enough to mention here?}
-% BCP: I think not in the abstract, but maybe in the intro.
-% We further apply\apt{what does this mean? we didn't
-%  invent it} a straightforward notion of
-% ``well-bracketed control flow'' that restricts control flow to the
-% typical pattern of matched calls and returns.
-
-We use these properties to validate \apt{``evaluate'' ?} the stack-safety {\em micro-policies}
+and a confidentiality property (``the callee's behavior is insensitive to the
+caller's private state''), which can optionally be extended with a control
+flow property.
+
+We use these properties to evaluate the stack-safety {\em micro-policies}
 proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
@@
 micro-policy, which catches violations as early as possible, enforces a
-simple ``stepwise'' variant of our properties and correctly detects several
-broken variants\apt{last phrase is grammatically disconnected}, and that (a repaired version of) their more performant
-``lazy'' micro-policy corresponds to\apt{``enforces''?} a slightly weaker and more extensional
-observational variant of our properties.
-%
-% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
-%   still not really clear yet on whether our definition applies / might apply
-%   to other enforcement mechanisms or not.  Saying that our attacker model is
-%   stronger does not help---that should make it easier for us to describe
-%   their protections, not harder.  So I don't understand what we're saying
-%   here.}%
-% %
-% While we do not rule out other enforcement mechanisms, our
-% definitions are strongly informed by micropolicy systems, especially
-% our attacker model, which is stronger than other enforcement mechanisms
-% can enable.
-
-% For the sake of concreteness, most of our presentation assumes a standard RISC
-% machine architecture with a basic subroutine calling convention (parameters
-% passed in registers) and a micropolicy-like enforcement mechanism. We
-% also show how to accomodate stack-allocated parameter passing and
-% coroutines.
-
-% \bcp{this is also a
-%   good place to admit that our definitions are tuned for micropolicies, if
-%   we can find a succinct way to say it.}
-% \sna{See above.}
-% \BCP{I added it to the last paragraph (which is pretty close to the preceding
-%   paragraph :-)  We should follow this up with some discussion in the (new)
-%   Assumptions section about how different enforcement mechanisms (us,
-%   StkTokens, maybe others) adopt different perspectives on what the stack is
-% for and how it is organized and so lead to somewhat different notions of
-% low-level safety.}
-
-% We argue \apt{do we?}\bcp{I guess we will not have time to.  Propose
-%   deleting this para.} that our properties are suited to a range of other
-% enforcement mechanisms. The strict stepwise properties can describe
-% mechanisms that catch errors immediately, as is typical. While the
-% observational properties are closely linked to lazy micro-policies they are
-% weak enough to be a lower bound that captures stack safety for any
-% reasonable enforcement mechanism.
-
-% \apt{This is a good story, but does it describe the paper we are able to produce this week?
-%   We were rejected before exactly because we claimed to have
-%   a general criterion, but did not in fact demonstrate its generality (which is not
-%   obvious). What has changed?}\bcp{I agree with these concerns. Deleted the
-%   para about the general story.  Does the rest now avoid making (or
-%   implying) claims that we can't defend?}
-% \bcp{Also, I think we should save
-%   WBCF for the intro---I can't find a way to discuss it here that is brief
-%   and not confusing/distracting.}
-
-  % \iftrue\bcp{should we mention stktokens here?}
-  % \leo{Not sure. We almost never cite stuff in the abstract, do we?}
-  % \bcp{Some people say not to, but I think their argument is weak: I've
-  %   added Nick/Andre above.  But I'm not sure we want to get into stkTokens
-  %   in the abstract.}
-  % \fi
+simple ``stepwise'' variant of our properties and correctly, and that
+(a repaired version of) their more performant
+``lazy'' micro-policy enforces a slightly weaker and more extensional
+observational property. Meanwhile our testing successfully detects violations
+in several broken variants, including Roessler and Dehon's original lazy policy.
+
 \end{abstract}
@@
 
-%\begin{figure}
-%  \begin{minipage}{\textwidth}
-%    \begin{center}
-%\MemoryLabel{10em}{1.5em}{40}
-%\MemoryLabel{16em}{1.5em}{60}
-%\MemoryLabel{22em}{1.5em}{80}
-%$\cdots$
-%\memory{4}{gray}%
-%\memory{5}{green}[{\makebox[0pt]{Caller}}]%
-%\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
-%\memory{8}{red}[{\makebox[0pt]{Callee}}]%
-%\memory{4}{\unusedc}%
-%~$\cdots$\\
-%  \end{center}
-%  \end{minipage}\\
-%~\\
-%~\\
-%  \begin{minipage}{0.45\textwidth}
-%    \begin{center}
-%\begin{verbatim}
-%jal ...
-%lw r1 r0 40
-%\end{verbatim}
-%    \end{center}
-%  \end{minipage}~
-%  \begin{minipage}{0.45\textwidth}
-%    \begin{center}
-%\begin{verbatim}
-%jal ...
-%sw r0 r0 40
-%\end{verbatim}
-%    \end{center}
-%  \end{minipage}
-%  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
-%      the boxes; the program fragments need comments explaining what each
-%      line does.}}
-%  \label{fig:stackunsafety}
-%\end{figure}
-%
-%\leo{I hate the writing here, but I wanted to make a start.}
-
-%\bcp{We don't actually say very clearly anywhere that we are interested in
-%  low-level enforcement mechanisms for machine-code programs (because this
-%  is where stack attacks occur, and because at higher levels of abstraction
-%  the stack is a built-in concept, not something that can be attacked even
-%  in principle).  This could cause confusion.\apt{not so worried}}
-%\leo{I agree. I've clarified the writing in the top of the running example
-%  a bit to make sure though}
-
 The call stack is a perennial target for low-level attacks, leading to a
@@
 
-%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
-%with the notion usually being
-%defined negatively: through examples of stack-based exploits
-%\rb{usually though not always; as noted below there are other
-%  approaches}\bcp{Yes, we have to be very careful about this}.
-%\leo{Examples at a high level}: such as---does not overwrite return addresses, or ...
-\apt{This makes us seem very reactive to Skst.  Can we reorder things to give our story first, and
-  compare to Skst afterwards?}
-The most abstract characterization of stack safety that we are aware of is
-recent work by \citet{SkorstengaardSTKJFP}, which defines stack safety as the conjunction of two
-properties:
+We propose such a characterization using the tools of language-
+based security~\citep{sabelfeld2003language}. The informal claim that
+“stack safety protects a caller
+from its callee” amounts to saying that it guarantees the {\em integrity} and
+{\em confidentiality} of the caller’s local state until it regains control.
+
+A key technical novelty in these definitions,
+compared to standard formulations of confidentiality and integrity from the
+security literature, is that they
+are ``nested'': {\em each} caller is guaranteed protection from its
+immediate callees (which, in turn, need protection from their immediate
+callees, etc.).
+Confidentiality is especially interesting. It is based on a traditional
+notion of noninterference, but whereas ordinary noninterference
+is an end-to-end hyper-property on whole program runs, stack
+confidentiality is a nested form of noninterference applied to
+subtraces of the whole program trace delimited by call instructions
+and corresponding returns, requiring that the callee’s behavior is
+invariant under hypothetical scrambling of the caller’s stack frame.
+
+Our threat model is very strong, allowing the attacker (i.e., the
+callee) to execute arbitrary code, including attempting to smash
+the stack to disrupt the program’s control flow. Our stack safety
+properties demand that, even in the presence of such attackers, confi-
+dentiality and integrity still apply to data in suspended stack frames
+until the callee returns.
+
+We can also, optionally, include a notion of well-bracketed con-
+trol flow as in \citet{SkorstengaardSTKJFP} --- a global requirement that
+callees must always return to their immediate caller, if they return
+at all. Skostengaard et al. define stack safety as the conjunction of two properties:
 {\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
@@
 return to the instruction following their call (if they return at all).
-Skorstengaard et al.{} formalize these properties by defining an idealized
-machine in which they hold {\em by construction};
-they then show that an enforcement
-mechanism (in their case, hardware capabilities) guarantees these properties, by
-proving that an ordinary machine with enforcement simulates the idealized
-machine.
-%an enforcement mechanism (in their case, hardware
-%capabilities)
-%can then be shown to guarantee stack safety by exhibiting a fully abstract
-%embedding function
-%from the idealized machine to an ordinary machine equipped with this
-%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
-%  understand that sentence.}
-%%
-%% While this work represents an important step in the right direction, we believe
-%% there is still room to craft an operational characterization
-%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
-%% In other words, we seek formal criteria that can be applied
-%% to a mechanism for protecting stack-based low-level programs to judge
-%% whether or not it actually ``guarantees stack safety.''
-%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
-%% confused by the references to these types of properties, yet to be introduced?}
-%% We begin with a strong
-%% local state encapsulation property that closely models the behavior of eager
-%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
-%% is too strong to characterize stack safety; reasonable enforcement mechanisms
-%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
-%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
-%% to support this technique and argue that the new property, defined relative to an
-%% arbitrary notion of observation. This property is still strong enough to prevent
-%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
-%% other reasonable enforcement mechanisms can enforce it.
-
-In the present work, we also describe the protection of both stack data (like LSE) and
-control flow (like WBCF).  Our technical formulation of WBCF is similar to that of
-Skorstengaard et al., but our treatment of stack data protection is rather different,
-as is our approach to the definition itself. Instead of defining stack safety implicitly,
-as ``what a certain idealized machine does,'' we give an
-\emph{explicit} definition using the technical framework of language-based
-security~\citep{sabelfeld2003language}, phrased as the conjunction of
-\emph{stack integrity} and \emph{stack confidentiality}.
-%
-A key technical novelty in these definitions,
-compared to standard formulations of confidentiality and integrity from the
-security literature, is that they
-are ``nested'': {\em each} caller is guaranteed protection from its
-immediate callees (which, in turn, need protection from their immediate
-callees, etc.). They are also stronger than LSE; even in the absence of the
-control flow guarantees of WBCF, they guarantee callers'
-data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
-%\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
-This reflects an arguably cleaner division between data protection and
-control-flow protection than LSE / WBCF.
+Skorstengaard et al. formalize these properties by defining an idealized
+machine in which they hold {\em by construction}. Our confidentiality and
+integrity properties are together stronger than LSE, guaranteeing protection even
+in the presence of control-flow violations, and can be seen as orthogonal to WBCF.
 
@@
 The observational properties are more extensional, and represent a baseline
-% meant to capture the same sense of security, extensionally enough
 that any reasonable protection mechanism ought to enforce,
 even if it does not prevent every single dangerous read or write.
-
-% {\em Confidentiality} says that a caller's
-% data is not leaked into its callee, and  that the callee cannot see uninitialized data.
-% As a reasoning principle, confidentiality guarantees that the callee's
-% behavior is determined only by a predictable interface (function arguments and
-% register contents) that does not include the caller's
-% local state or the uninitialized portion of the stack. This is formalized as a
-% {\em noninterference}-style property that, rather than comparing pairs of whole program traces,
-% compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
-% again
-% {\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
-% changes in the state are independent of the caller's data.
-
-% Together, these properties capture roughly the same intuition as LSE, but
-% they are somewhat stronger.
-% First, we model returns as returns to a valid return target\bcp{that will
-%   make no sense to the reader at this point}; an invalid return
-% is treated as a mere jump within the callee, and does not grant access to the caller's
-% data. So, even without WBCF, our stack safety properties rule out many control-flow attacks.
-% The attacker in such an instance is permitted to ``return'' so long as it avoids
-% violating the integrity and confidentiality of the caller.\bcp{ditto} Second, confidentiality
-% contains\bcp{?} temporal protections that LSE does not necessarily address\bcp{?}. The intuition
-% for LSE is that it protects the caller's stack frame, not necessarily memory that was
-% previously allocated but now outside of the caller's frame. (Cheri-capability-based
-% implementations of LSE {\em do} protect previously allocated data, typically by clearing
-% it, but because it is a potential vector for leaking capabilities into the caller's frame,
-% not as a goal of the property.\bcp{Oof.  We need to do better than that!})
 
@@
 in testing and the enforcement mechanism itself.
-%introduce several
-%intentionally broken micro-policies and check that the testing framework is
-%able to generate counterexamples that violate our properties but are not
-%halted by the incorrect enforcement variants.  (Showing that the testing
-%framework detects intentional bugs in enforcement mechanisms increases our
-%confidence that it can also catch unknown ones, strengthening our confidence
-%in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
-%  sentence belongs in the testing section?}\fi)
 
 We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
-which memory cells within each stack frame are tagged with the identity\apt{``depth''?} of
+which memory cells within each stack frame are tagged with the depth of
 the function activation that owns the frame and access to those locations is
-then permitted only when that activation\apt{``an actiation at that depth''?} is currently executing, validates our
-stepwise properties. On the other hand, we reason\apt{but now we test?}  that \emph{Lazy Tagging and Clearing}
+then permitted only when that an actiation at that depth is currently executing, validates our
+stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
 violates the temporal aspect of confidentiality in
@@
 and also violates integrity if the leak happens to use the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
-confidentiality, albeit at some performance cost.\apt{Do we test?}
+propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
+confidentiality, albeit at some performance cost.
 %
@@
 
-% A second challenge is that enforcement mechanisms may weaken their
-% protections\bcp{``weaken their protections'' doesn't convey what we mean
-%   very well}
-% in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
-% \citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
-% ways of identifying functions and deciding exactly which accesses are allowed.
-% These vary in their efficiency and in the spatial and temporal precision of their
-% security guarantees.
-% The Depth Isolation policy prevents a callee from writing outside its frame
-% at all, which is simple and intuitively correct, but it must manually clear the tags
-% on the frame on each return, which is very expensive. Conversely, their
-% \emph{Lazy Tagging and Clearing} policy improves
-% performance by assigning locations to the active function when they are written, which
-% permits  ``illicit'' writes to inaccessible stack frames so long as the
-% written locations are not subsequently read by their ``real'' owners. This violates our
-% straightforward properties, but ntuitively,
-% this form of laziness is justified if any dangerous writes will
-% ultimately be detected before they can affect program behavior.
-
-% \bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
-% What does ``affecting program behavior'' really mean? This micro-policy
-% calls for a second characterization of stack safety in terms of the externally
-% observable behavior of a stack-safe system. Such a characterization will have the
-% additional benefit of being more {\em extensional}, concerned primarily with the
-% external behavior of the system. In contract, our stepwise properties are {\em intensional},
-% focusing on internal details that may not actually matter to overall security.
-
-% \bcp{Ditto, even moreso.}
-% To formalize this intuition, we develop an \emph{observational integrity} property which
-% characterizes enforcement mechanisms that detect violations between the point when
-% they occur and the point when they make an observable impact. Intuitively this property
-% can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
-% integrity implies observational integrity. \sna{TODO: check this.}
-
-% We additionally sketch an associated {\em observational confidentiality} property
-% makes confidentiality extensional in the same way, in order to form an
-% {\em observational stack safety} that is highly extensional.
-% We argue that, despite its somewhat greater complexity,
-% it may be a better specification for
-% stack safety enforcement mechanisms in general.
-
-%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
-%  significant contribution, and partly to give the reader some context for
-%  why this is part of our story.  E.g., ``One nice thing we can do with our
-%  spec is to use it to evaluate the claims of existing or future proposals
-%  for enforcement mechanisms.  One relatively inexpensive way to carry out
-%  such an evaluation is to see whether property-based testing against this
-%  spec can distinguish correct implementations of the mechanism from buggy
-%  ones...''}
-
 For ease of exposition, our development initially assumes a single simple
@@
 
-%% \paragraph*{The Use of Stack Safety}
-
-%% How does the programmer rely on stack safety? The stack makes implicit promises about
-%% the behavior of function calls. First, that when a call is finished, the caller can continue
-%% from where it left off with its local state unchanged. And second, that the call's
-%% behavior is determined only by a predictable interface that does not include the caller's
-%% local state.
-
-%% %The most natural example is the first
-%% %in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
-%% %calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
-%% %{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
-%% %overwrite {\tt main}'s stack data.
-
-%% Both expectations imply predictable control flow, termed {\em well-bracketed
-%% control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls always jump to valid entry
-%% points and return to their call sites. Then the data portion of the first expectation is for
-%% {\em integrity}: that a caller's local data is protected from modification by its callee.
-%% This is vital for local reasoning about the behavior of the caller. The security implications
-%% are significant: without integrity, a caller that branches on a local variable could be
-%% manipulated by its callee into unexpected behavior.
-
-%% Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
-%% from leakage by its callee. As a reasoning principle, confidentiality tells us that to
-%% predict the behavior of a call we need only consider a subset of the machine state.
-%% And in security terms, we can guarantee the secrecy of local data that we haven't passed
-%% even when calling a badly behaved function. Together, integrity and confidentiality
-%% comprise \citeauthor{Skorstengaard+19}'s {\em local state encapsulation} (LSE) property.
-
-%% It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
-%% could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
-%% accessing the caller's state, but the execution would remain under the callee's control.
-
-%% To phrase LSE and WBCF as security properties requires an explicit attacker model:
-%% a clear understanding of who is being protected and from what. Most existing
-%% stack protection mechanisms envisage scenarios where a callee is
-%% influenced by external inputs to corrupt the stack in some way. To
-%% avoid getting into the details of exactly how this happens when
-%% formalizing our properties, we use an even stronger model: when we
-%% state the protections that a function can rely on when it makes a
-%% call, we assume that the callee can execute arbitrary instructions
-%% until it returns. Thus we have no need to differentiate safe code
-%% from adversarial --- all functions, and indeed all activations of the same function,
-%% are mutually distrusting.
-%% %
-%% For this model to make sense, we need to know when a call or a return is taking
-%% place. We assume that calls and returns are annotated in the code, so that
-%% we can track the point at which control transfers from the protected caller
-%% to the attacking callee --- and when that callee in turn becomes a protected
-%% caller in its own right.
-
-%% This ``nesting'' of properties is a key challenge in adapting the concepts
-%% of language-based security to the stack setting. We address this issue by
-%% turning any set of annotations into a dynamic model of {\em domains},
-%% in which some stack locations are sealed by the caller that needs to preserve
-%% them while the rest are unsealed and available for use by the active, attacking
-%% callee.
-
-%% First we introduce this model for a single simple stack with no sharing between callers and
-%% callees. Then we expand it to allow passing of stack data, granting a callee one-time
-%% access to the caller's data. We do not model more sophisticated sharing, because to do
-%% so realistically would require a separate protection scheme, such as a capability
-%% model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
-%% a separate policy outside the scope of stack safety. (A tag-based approach can easily
-%% compose multiple such policies.)
-
-%% Finally, we extend the model to a coroutine system featuring finitely-bounded stacks.
-%% Though simplistic, this extension is a step toward concurrency and demonstrates that
-%% the model is flexible enough to handle more sophisticated styles of control-flow.
-
-%% After formally defining stack safety, we show that the definition is
-%% enforceable using an existing tag-based enforcement
-%% mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
-%% \citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy.
-%% This policy testably enforces our {\em eager} properties, in which integrity is respected
-%% at every step of a program, and each call respects integrity at each of its internal steps.
-%% These properties are very strong, and our policies enforce them; this also makes them very
-%% intensional, in that they do not distinguish internal behavior from external.
-%% They also have the advantage that they are suitable for efficient testing, as they
-%% fail immediately when a violation occurs.
-
-%% The downside of the Depth Isolation policy is that it is rather expensive to implement.
-%% In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
-%% that is more efficient by virtue of being lazy in its enforcement.
-%% Instead of enforcing the integrity and confidentiality of the stack at the
-%% exact moment when a callee returns, this policy signals violations when the
-%% caller actually accesses data that the callee has written into its stack
-%% frame. \iftrue\bcp{We could also hint at the lazy confidentiality
-%%   property in the same way...}\fi
-%% The lazy policy does not enforce our eager integrity property, and to our
-%% knowledge there is no formal statement of the properties that it ought to enforce.
-%% We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
-%% policy that detects integrity violations between when they occur and when they make
-%% a visible impact. Intuitively this can be phrased as ``no visible reads after dangerous
-%% writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
-%% ``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
-%% testing, but formalizing the ideal property allows us to see how it falls short.
-%% We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
-%% at some performance cost.
-%% %We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
-%% %  and is it so interesting anyhow?}
-%% We argue that any system purporting to enforce stack safety ought to obey at least lazy
-%% stack safety, making it a plausible extensional characterization.
-
-%% %\bcp{I worry that we are being a little
-%% %  dishonest here: the stated advantage of the lazy policy is that it is more
-%% %efficient, but our variant is {\em not} efficient!}
 In the rest of the paper, we begin with an informal overview of our key ideas via
@@
 property, and argue that this property captures the intuitive concept of
-strict stack
+strict stack.
 safety~(\cref{sec:prelim,sec:lse}).
@@
 properties (\cref{sec:enforcement}).
-%% \bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
-\item We formalize a
-weaker but more extensional \emph{observational} variant of the stack safety
-property that characterizes a refinement of Roessler and DeHon's more
-performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).\apt{can this be more precise?}
+\item We formalize weaker but more extensional \emph{observational} variants of
+  confidentiality and integrity. Both are parameterized over a notion of external
+  observation, and are violated only if the accessing of secrets or overwriting of
+  data later causes a visible change in the system behavior (\cref{sec:lazy}).
 \item We show that the definition can be
@@
 
-We now define our machine model \apt{and enforcement model?} precisely. To make our definition of stack safety
+We now define our machine model precisely. To make our definition of stack safety
 as generic as possible, this section proposes an abstract interface to a
@@
 individual calls from a trace by taking subtraces in which the stack is
-at or below\apt{``above''?}  a given depth. If \(f\) is still a predicate on states, then
+at or below a given depth. If \(f\) is still a predicate on states, then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
@@
 
-Each system component is assigned to a {\em domain}, which may be {\em outside}
-of the stack entirely, {\em unsealed} and therefore accessible, or {\em sealed}
+Each system component is assigned to a {\em domain}, which may be {\em Outside}
+of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
 at a particular depth. Each caller registers a {\em target}
@@
          {\(\underscore,\context \stepsto_C \context\)}
-\apt{typesetting problem above}
-
-\noindent
+
 When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed.
-
-We now state our properties.
+call stack we are and which components are sealed. We now state our properties.
 The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
@@
 \definition
-The \emph{protected set} of a context, \(\prot(\domMap,\_)\), is the
+Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
 set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
@@
 
-\definition The \emph{call set} \(\MPCT^*\) of a trace \(\MPCT\) \apt{right?} is
+\definition \(\MPCT^*\) is the \emph{call set}, defined as
 \[\begin{aligned}
@@
 
-\apt{intuition for what the call set is?}
 We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
@@
 \definition
-The \emph{secret set} of a context, written \(\secret(\domMap,\rts)\), is the
+Let the secret set of a context, \(\secret(\domMap,\rts)\), be the
 set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
@@
 of that property which checks for integrity and confidentiality violations at
-every step of the callee's execution trace. We validate this claim \apt{which claim: that it enforces
-  a \emph{stronger} version?}
+every step of the callee's execution trace. We validate this claim
 using a property-based random testing framework to uncover
-inconsistencies \apt{but are there any?} between the enforcement mechanism and the stack-safety
+inconsistencies between the enforcement mechanism and the stack-safety
 property.
@@
 \citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
-with ownership information \apt{based on depth!} associated to each stack frame (and to each
+with ownership information associated to each stack frame (and to each
 separate object inside that frame, a more finely grained access

### 2021-07-08 submission3/main.tex
pass through 6

--- 2021-07-08 submission3/main.tex
@@
 on the {\em Depth Isolation} policy presented by
-\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
-with ownership information \apt{based on depth!} associated to each stack frame (and to each
+\citet{DBLP:conf/sp/RoesslerD18}. In their policies,  the stack memory is tagged
+with ownership information associated to each stack frame (and to each
 separate object inside that frame, a more finely grained access
@@
 
-The micro-policy maintains tags of the form $\tagStackDepth{n}$ for
+The Depth Isolation micro-policy maintains tags of the form $\tagStackDepth{n}$ for
 each stack location belonging to the stack frame at activation depth
@@
 
-More interesting changes to the tags (from a stack safety perspective)
-happen at exactly the points in the program where we switch from a
-caller to a callee or vice-versa:
+Interesting changes to the tags occur where the program switches from 
+a caller to a callee or vice-versa:
 
@@
 
-\item From caller to callee, when the machine executes an instruction marked as a call in
-  the call map.
-  At this point, the current
+\item From caller to callee: when the machine executes a call,
+  %, when the machine executes an instruction marked as a call in
+  % the call map.
+  the current
   $\tagPCDepth{n}$ tag is incremented to $\tagPCDepth{(n + 1)}$, and the function
@@
 
-\item From callee back to caller, when execution is about to reach a
-  return point.
+\item From callee back to caller: %when execution is about to reach a
+%  return point.
   %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
-  At the actual return instruction,
-  the $\PCname$ tag is decremented.
+  at the actual return instruction,
+  the $\PCname$ tag is decremented. \apt{Doesn't the exit sequence retag to $\tagNoDepth$?}
 %  \leo{Not sure the actual policy
@@
 
-This discipline suffices to enforce local stack encapsulation.
+This discipline suffices to enforce stack safety. 
 To implement it,  we define blessed instruction sequences
@@
 (with respect to a particular call map), all we need require is that
-the entry sequence be initiated at any instruction marked as a call
+the entry sequence be initiated at \apt{immediately following?}  any instruction marked as a call
 in the call map. This is achieved by giving these instructions (another) special tag,
@@
 \label{sec:testing}
+
+There are several ways to evaluate whether an enforcement mechanism enforces
+stack safety properties. Ideally such validation would be done through formal proof over
+the semantics of the enforcement-augmented machine. 
+However, while there are no fundamental barriers to producing such a proof, 
+it would be considerable work to carry out for a full ISA like RISC-V and
+complex enforcement mechanism like the Depth Isolation micro-policy.
+We therefore choose to validate the micro-policy of the previous section by
+systematically \emph{testing} that it satisfies our properties. 
+This focus on testing is better aligned with our immediate
+goal of evaluating real enforcement mechanisms for real machine architectures.
+Formal proof remains as possible future wrok.
 
@@
   chose the Coq proof assistant as the setting for our implementation
-  for two reasons: first, to ensure that our coinductive trace
+  to ensure that our coinductive trace
   definitions are well-formed, and reason about them, and to leverage the power of the
@@
 (3)
- a code map and a stack map, accurately characterizing the behavior
+ a code map and a stack map \apt{what's that ??}, accurately characterizing the behavior
   of callers and callees.
@@
 To write such generators we build on the work of
-\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
-introduced {\em generation by execution} to produce progams leading
-abstract machines towards interesting behaviors. Generation by
+\citet{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
+introduced {\em generation by execution} to produce progams that lead
+machines towards interesting behaviors. Generation by
 execution receives as an input a partially instantiated machine state
@@
 such as a blessed sequence) that makes sense locally (e.g., jumps go
-to a potentially valid code location and loads read from from a
+to a potentially valid code location and loads read from a
 potentially valid stack location). Then we step the machine and repeat
 the process until we generate or execute some target number of
-instructions, or we reach a point where the machine can't step
+instructions, or reach a point where the machine can't step
 anymore.
@@
 to failstop when executed in conjunction with the policy monitor,
-but also allows for revealing errors in our setup.
+but also allows for revealing errors in our setup.\apt{??}
 
@@
 micro-policy needs to enforce that the tag on the location being read
-is $\tagStackDepth{n}$ for some number $n$ and that the one of the
-current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
+is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
+current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
 that restriction by not checking the depth equality (row {\em
@@
 %  and the write from {\tt f} was already a violation.}
-
-\subsection{Why not Proof?}
-
-There are several ways to evaluate whether an enforcement mechanism enforces
-our properties. Ideally such validation would be done through formal proof.
-Our model is certainly amenable to such a proof, with no fundamental barriers,
-but it would be considerable work to carry out for a full ISA like RISC-V.
-We concentrate instead on testing because it is better aligned with our immediate
-goal of evaluating real enforcement mechanisms for real machine architectures.
-\sna{More to say here?}
 

### 2021-07-08 submission3/main.tex
Fixed merge error

--- 2021-07-08 submission3/main.tex
@@
 
-We use these properties to evaluate the stack-safety {\em micro-policies}
+We use these properties to validate the stack-safety {\em micro-policies}
 proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
@@
 on the {\em Depth Isolation} policy presented by
-\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
+\citet{DBLP:conf/sp/RoesslerD18}. In their policies,  the stack memory is tagged
 with ownership information associated to each stack frame (and to each
@@
 
-\subsection{Why not Proof?}
-
-There are several ways to evaluate whether an enforcement mechanism enforces
-our properties. Ideally such validation would be done through formal proof.
-Our model is certainly amenable to such a proof, with no fundamental barriers,
-but it would be considerable work to carry out for a full ISA like RISC-V.
-We concentrate instead on testing because it is better aligned with our immediate
-goal of evaluating real enforcement mechanisms for real machine architectures.
-\sna{More to say here?}
-
 \section{Stack Safety for Lazy Enforcement}
@@
 
-%\definition
-%The protected and secret sets of a context as used for calls,
-%\(\prot_C(\domMap,\rtm,\yts,\stackid)\) and \(\secret_C(\domMap,\rtm,\yts,\stackid)\),
-%are equivalent to their counterparts \(\prot_P\) and \(\secret_P\) applied to a domain map
-%that projects out the stack domains of components in stack \(\stackid\) and a
-%return target stack \(\rts = \rtm ~ \stackid\). Intuitively, within the active stack we protect
-%the same components we would have in the passing model.
-
-%\definition
-%For each of the stepwise, deferred, and observational versions of
-%stack integrity in the simple model, the equivalent property in the
-%passing model substitutes the above definitions of \(\context\), initial contexts and update
-%function, and \(\prot_C\) for their simple equivalents, and is otherwise identical.
-
-%\definition \(\MPCT^*\) is now defined as
-%\[\begin{aligned}
-%\bigcup_{\stackid \in \STACKS,\depth \in \mathbb{Z}} \hspace{-1em} \{\MPCT | \MPCT\langle f_\depth \rangle
-%\text{ is a reachable segment}\} \hspace{1em} &
-%\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rtm,\yts,\stackid')) \triangleq |\rtm ~ \stackid| \geq \depth \\
-%\end{aligned}\]
-%and \(\brk{\MPCT}\) similarly projects the depth of the active stack.
-
-%\definition
-%For either of the stepwise and observational versions of
-%stack confidentiality in the passing model, the equivalent property in the
-%coroutine model substitutes the above definitions.
-
 \definition \(\MPCT^\wedge\) is the \emph{yield set}, defined as
@@
 concrete capability machine.
-%% that replaces those pseudo-instructions with their calling convention.
 The proof of full abstraction uses a standard notion of components, which
@@
 
-With respect to criterion (4), we believe that our support
+We agree that criterion (2), usefulness of the safety criterion
+for carrying out formal proofs (about individual programs), is important;
+demonstrating it for our proposal remains future work, in which we
+plan to apply our properties to justifying a low-level program logic.
+
+StkTokens
+is specifically set up to address criterion (3), reusability in secure
+compilation chains, while this is not a goal of
+our work.
+
+With respect to criterion (4), capturing programmers' informal
+understanding of stack safety, we believe that our support
 for dynamically changing notion of adversaries, as outlined in point (i) above,
@@
 
-As for criterion (5), we have demonstrated that our approach extends
-to coroutines, a first step towards multithreading.
-Dynamic code generation is more challenging, but StkTokens does not yet
-handle this either.
-StkTokens is specifically set up to address criterion (3), while this is not
-a goal of our work. We do agree that criterion (2) is important;
-demonstrating it for our proposal remains future work.
+As for criterion (5), broad applicability, we have demonstrated that our approach extends to
+coroutines, a first step towards multithreading. Dynamic code generation is
+more challenging, but StkTokens does not yet handle this either.
 
 Finally, we would propose one more criterion: having a generic definition
-of stack safety that abstracts over machine details.  Here we do better
-than StkTokens, as described in point (ii) above.
-
-%% Although there are some important differences between
-%% \citeauthor{SkorstengaardSTK}'s properties and our own, the new
-%% properties arguably satisfy their list of desiderata for such
-%% formulations: notably, they reflect (and refine) programmer intuitions
-%% about stack safety, furnish a comprehensive picture of the meaning of
-%% stack safety, can be used to reason about program behavior, and are
-%% flexible enough to be allow extensions to richer settings.
-%% %
-%% \sna{Some of this seems like more detail than we need, and we need to
-%%   compare ourselves to their list of criteria.}
-%% \rb{Trimmed this a bit. Are you referring to their list of attack
-%%   examples, or to something else?\apt{Their list at start of section 4.
-%%   \rb{Had a go at this.}}}
+of stack safety that is directly reusable to evaluate multiple machines and
+enforcement mechanisms. They have used their approach to give strong validation
+to one machine/enforcement mechanism, a linear-capability machine,
+via formal proof. We have used ours to obtain a weaker, testing-based validation of
+two different mechanisms on a single machine, and will target a capability machine
+next as future work.
 

### 2021-07-08 submission3/main.tex

@@
 The conservative policy described in the previous section, while
-testable and enforceable, is rather slow to be of practical use. What
-we want, instead, is a somewhat more permissive policy that can also
-be efficiently implemented. Most of the performance overhead incurred
+testable\apt{is this important?} and enforceable\apt{?? it is an enforcement
+  mechanism so obviously it is enforceable -- whatever it may enforce!}, is rather slow in
+some important practical cases. \citet{DBLP:conf/sp/RoesslerD18} observe
+that most of the performance overhead incurred
 stems from the need to set stack activation tags as a frame is created
-and reset them as it is destructed, as
-\citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
-mitigate those costs, they propose a number of optimizations.
-
-First, they eliminate the initialization pass. In the {\em Lazy Tagging}
+and reset them as it is destructed;  this
+is particularly costly for frames that contain large uninitialized arrays.
+To mitigate those costs, they propose a number of optimizations to the policy
+that make it more permissive but also more efficient to implement.
+
+First, they eliminate the costly frame initialization pass. In the {\em Lazy Tagging}
 system, the stack is again initially tagged as \(\tagNoDepth\).
@@
 to overwrite \(\tagNoDepth\) memory, and when it does so its depth tag overwrites
-the clear tag, claiming the memory; the remainder of the policy functions
-as {\em Depth Isolation}. On return, the callee must clear its frame, restoring
-it to \(\tagNoDepth\), which is still a heavy cost.
+the \(\tagNoDepth\) tag, claiming the memory; the remainder of the policy functions
+as in Depth Isolation. However, on return, the callee must clear its frame, restoring
+each tag to \(\tagNoDepth\), which is still a heavy cost.
 
@@
 
-Lazy policies are not the only scenario to call for a weaker property of this kind.
-Some code features may involve moving data, overwriting its original location,
-and ultimately restoring it before the change becomes visible; for example, the
-``callee-saves'' registers in some calling conventions will be stashed and restored.
-Stepwise stack safety can never support such a feature, but in section \ref{calleesave}
-we show how {\em observational properties} can.
-
 \paragraph*{Deferring Integrity Checks}
@@
 \((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
+\apt{Intuitive description of what this means?}
 
@@
 
-%\apt{Given that the deferred property suffices to specifcy the lazy policies,
-%  the case for observable properties is pretty weak.  Should we just cut them?}
-
 While deferred integrity meaningfully captures the intuition
@@
 evaluate the broadest possible range of reasonable
-enforcement mechanisms. Consider a hypothetical
-taint-tracking policy that allows high-security to be copied as long as it is
-tagged as such and not exfiltrated, or that allows dangerous writes to be
-copied as long as they don't affect critical operations. This calls for
-an abstract mechanism to characterize behaviors that are externally relevant.
+enforcement mechanisms.
+For example, reading after a dangerous write may be acceptable if the value
+read does not affect the ultimate output of the program; this looser scheme
+could be enforced by tainting the dangerous writes and detecting taint
+at the point of output.
+%For example, consider a hypothetical
+%taint-tracking policy that allows high-security data to be copied as long as it is
+%tagged as such and not exfiltrated, or that allows dangerous writes to be
+%copied as long as they don't affect critical operations. \apt{Not quite sure
+%  how these last connect to stack safety.}
+And some code features may involve moving data, overwriting its original location,
+and ultimately restoring it before the change becomes visible; for example, the
+``callee-saves'' registers in some calling conventions will be stashed and restored.
+These examples call for an abstract mechanism to characterize behaviors that are externally
+\emph{observable}.
 
@@
 We say that two observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
-as far as an external observer is concerned, written \(\obsT_1 \eqsim
+from the perspetive of an external observer, written \(\obsT_1 \eqsim
 \obsT_2\), if the sequence of non-silent observations is the same. That is, we
@@
 This require us to be able to quantify over components that fail to preserve
-our \(\diamond\) relation.
+the change-together relation \(\diamond\).
 
@@
 
-Together with return-after-call, these two properties give us
+Together with return-after-call\apt{??}, these two properties give us
 {\em Observable Stack Safety}. This is a weaker property
-than the stepwise one, and also much more extensional.
+than the stepwise one, and also much more extensional.\apt{intensional?}
 We can think of stepwise, deferred, and observational as
@@
 but if we had to choose a single property to represent stack safety,
-it would be observational state safety.
+it would be the observational one.
 
@@
 violation would become visible. Now we can identify when lazy policies
-miss violations that can become visible. The {\em Lazy Tagging and Clearing}
+miss such violations. The Lazy Tagging and Clearing
 micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
@@
 for a future callee at the same depth to read. This violates confidentiality and,
-if the writes were in the caller's frame, integrity.
-
-{\em Lazy Tagging and Clearing} can be repaired if, instead of tagging each function
+if the writes were in the caller's frame, integrity. \apt{A more detailed example
+  would be great here, and should not take too much space.}
+\apt{Mention here that this was found by testing.}
+
+Lazy Tagging and Clearing can be repaired if, instead of tagging each function
 activation with its depth in the stack, we generate a fresh activation
 identifier on each call, which prevents a program from exploiting stale
-tags from previous activations. A related mechanism was explored in the
+tags from previous activations.
+(A related mechanism was explored in the
 Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
-associates a unique activation identifier to each function, which is
-however shared by all activations of the same function in order to obtain
-good cacheability of tag rules---but that policy allows leaks between
-instances of the same function.
+associates a unique \emph{static} identifier to each function, which is shared
+by all activations---but that policy allows leaks between
+instances of the same function.)
 
@@
 an equivalent to {\em LOAD\_NO\_CHECK}, which now also violates observational
-integrity.
-
-Per-activation tagging resolves the issue because tags that persist between
-calls will not become active again.
-Our properties have allowed us to reject an enforcement mechanism as
-not really stack safe, and describe a potential fix that itself is a
-prime candidate for future testing.
+integrity.\apt{Move latter to above.}
+
+%Per-activation tagging resolves the issue because tags that persist between
+%calls will not become active again.
+Our abstract properties have allowed us to identify an enforcement mechanism as
+not really stack safe, and to validate a possible fix. 
+Unfortunately, Lazy Per-activation Tagging and Clearing
+puts severe pressure on the tag cache, so it may not be usable in practice.
+However, we conjecture that no lazy scheme can be fully stack safe
+without using unique per-activation identifiers.
 

### 2021-07-08 submission3/main.tex
remove bad comment

@@
 {\em Observable Stack Safety}. This is a weaker property
-than the stepwise one, and also much more extensional.\apt{intensional?}
+than the stepwise one, and also much more extensional.
 We can think of stepwise, deferred, and observational as
@@
 captures a better definition of safety than LSE; moreover, our observational
-properties capture an even richer and more intensional notion of stack safety
+properties capture an even richer and more extensional notion of stack safety
 that can be applied to a broader range of enforcement mechanisms and code features.

### 2021-07-08 submission3/main.tex
Section 3 work

--- 2021-07-08 submission3/main.tex
@@
 
+Intuitively, stack safety means that the stack frame of each caller is protected
+from attempts by a confused or malicious callee (including all its
+descendant callees) to read or write its private data.
+That is, the attacker model treats every caller as a defender and its callee as
+an attacker, meaning that with nested calls, a single function invocation
+may be viewed in both ways.
+
 The protections offered by such mechanisms are commonly described in terms
@@
 
-\section{Machines, Enforcement, and Threat Model}
+\section{Threat Model, Machines, and Enforcement }
 \label{sec:setup}
 
-Before introducing the model\apt{what does this really mean?}  itself, we set the context by stating some
-assumptions we are making about machines, enforcement mechanisms, and
-potential threats.
+We will introduce a formal model of stack safety that abstracts over
+machine details. But first we must discuss the bounds of our assumptions ---
+how broadly we can abstract over machines, how we represent enforcement, and
+our threat model.
+
+We adopt a strong threat model in which both caller and callee may execute
+arbitrary machine code. But hardware and timing attacks are out of scope, as are
+are attacks on the heap---though we believe that the work of
+\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
+to protect both stack and heap.
 
@@
 Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the concrete machine that instantiates it. Without changing the
-properties, we can support\apt{what does this really mean?} any register
-machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
-details of the machine's ISA are unimportant. We can support any static memory,
-including a heap---our properties define stack protection regardless of how
-non-stack memory is used. Likewise, we can support a range of calling convention
-choices, like whether or not to have a frame pointer.
-
+assumptions about the concrete machine that instantiates it. Each formal property
+may be instantiated with any machine that meets the following criteria, and it
+will be a meaningful defition of stack safety for that machine. It must at least be a
+register machine with a program counter (\(\PCname\)), and there must be a well-defined
+mechanism for determining the bounds of a stack frame, and for recognizing when a call
+has returned. In this paper we discuss those machines with a stack pointer (\(\rsp\)),
+in which the current stack frame is everything below the stack pointer and above the
+previous frame, and a call from instruction \(\addr_\PCname\) with stack pointer \(\addr_\rsp\)
+is said to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
+is again at \(\addr_\rsp\). But, other configurations, such as those involving frame pointers,
+are equally feasible. A machine may use its non-stack memory in any way, including
+as a heap, as long as the stack has static bounds.
+
+Our core property assumes that all register values passed to the callee are public,
+and that all data in a stack frame is private.
 Some calling convention choices, such as callee-saves registers and argument
-passing on the stack, require natural enhancements to our base properties.\apt{
-  At this point it is unclear what these ``enhancements'' might be.}
-We
-present these enhancements, as well as a simple coroutine model,
-in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
-Other plausible enhancements include:\apt{Better to move the ``what would
-  be needed'' parts to the future work section -- they are mysterious here.}
-
-\begin{itemize}
-\item address-taken locals---similar to argument passing, but
-      would benefit from more sophisticated, heap-like memory models
-\item exceptions---needs annotations of throw/catch points
-\item continuations---would be similar to argument passing
-\end{itemize}
-
-Runtime code generation would be much more challenging to support, as
-it requires dynamic generation of code annotations. Similarly, true concurrency
-would require more substantial changes to the machine
-model. We consider these features out of scope for now.
+passing on the stack, are not supported directly by this property; we introduce
+secondary properties that use the same underlying formalism to handle these cases,
+as well as a simple coroutine model, in Section~\ref{sec:ext}.
+
+Other enhancements, such as runtime code generation and true concurrency,
+would be much more challenging to support, and we consider these features
+out of scope for now.
 
@@
 enforcement mechanisms that keep a significant amount of data separate from the
-primary execution data to inform security. Obviously tag-based systems do this\apt{At this
-  point, we barely know what tags are},
-with tags and extra policy state separated from, and untouchable by, application code.
-Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
-mirrors the real one, and a return may only occur if both stacks agree on the
-target address. Typically the shadow stack is isolated using a separate protection
-mechanism.
-
-The separation is smaller in protection mechanisms that interact more directly
-with application code.\apt{Vague. Mention capabilities?}  Some mechanisms might exhibit failstop behavior, but have
-no protected policy state; others might never even failstop. Then the policy-enhanced
-machine is simply identical to the original. This is the case for software-based
-techniques.\apt{Even if they use a shadow stack?}
-
-\paragraph*{Threat Model}
-\label{sec:threat}
-
-Consider a common scenario: a developer wants to link against an existing
-code-base, which will then occupy the same address space as their own code.
-But perhaps the external code is buggy, or even malicious, and contains
-exploits that can be used to cause it to modify the developer's data or
-exfiltrate secrets. For that matter, the developer may not experienced at
-secure programming and so may not be able to rely on their own skill
-to keep their secrets safe.
-
-The developer likely relies instead on simple reasoning principles about how
-their code should work. Local variables do not change across function calls,
-and the callees generally do the same thing regardless of when and where they
-are called (up to a reasonable interface, such as their parameters and global
-variables.) If these intuitions prove false, then the programmer will reason
-incorrectly about their program and especially about its security, and
-be exposed to attack.
-
-Intuitively, stack safety means that the stack frame of each caller is protected
-from attempts by a confused or malicious callee (including all its
-descendant callees) to read or write its private data.
-That is, the attacker model treats every caller as a defender and its callee as
-an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways.
-\apt{everything down to here seems like it should go earlier -- in informal section
-  or even intro.}
-We do rule out some
-scenarios with a ``well-formedness'' criterion on initial states, ensuring that
+primary execution data to inform security. Tag-based systems, in which a hardware
+monitor tracks metadata tags associated with a software-defined ``micro-policy,'' are
+the motivating application. The tags, and any extra policy state, are separated from
+and untouchable by application code.
+%Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
+%mirrors the real one, and a return may only occur if both stacks agree on the
+%target address. Typically the shadow stack is isolated using a separate protection
+%mechanism.
+\sna{Commented out shadow stacks until I can find a citation. May just skip it.}
+
+That said, our model is meant to apply generically to any enforcement mechanism,
+including those that do not take advantage of this separate.
+Some mechanisms, like hardware capability systems, might exhibit failstop behavior,
+but have no protected policy state --- this is represented by a ``unit'' policy
+that failstops when the hardware fault. Others might never even failstop, such as
+code rewriting techniques that turn stack safety violations into other (erroneous, but not
+stack-unsafe) behaviors. Then the policy-enhanced machine is identical to the original.
+\sna{I'm thinking of SFI here, but that's not stacks, right? Does Softbound work like
+  this.}
+
+An enforcement model may additionally
+apply a ``well-formedness'' criterion on initial states, ensuring that
 they are set up properly for the enforcement mechanism in question.
@@
 execute the blessed sequence out of order.
-
-We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. \apt{This should be the lead sentence. But how does is
-  it consistent with the previous para on well-formedness?}
-  Hardware and timing attacks are out of scope, as are
-are attacks on the heap---though we believe that the work of
-\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
-to protect both stack and heap.
 

### 2021-07-08 submission3/main.tex
starting on 8

--- 2021-07-08 submission3/main.tex
@@
 
-To this point we have ignored function arguments, since our simplest model places them in
-registers. We now show how to implement two kinds of protected
-sharing: passing by value on the stack, and passing by reference. We first explore
-some examples and then move on to
-extending our formalization.
+To this point we have ignored function arguments, by assuming that they are placed
+in registers (which are implicitly shared by all functions). 
+We now show how to implement two kinds of protected
+sharing of stack contents: passing by value on the stack, and passing by reference. We first explore
+some examples and then extend our formalization.
 
@@
 
-Our model provides only limited support for handling fully first-class pointers into
+Our approach provides only limited support for handling fully first-class pointers into
 stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike pass-by-reference, there is no easy way to model the intended
+Unlike with pass-by-reference, there is no easy way to model the intended
 sharing behavior of such pointers in terms of stack frame activation patterns.
@@
 (implemented by tags or otherwise); we would argue it is beyond the scope of
-stack safety \emph{per se}.
+stack safety \emph{per se}.\apt{Could try to incorporate some of 
+  our response to OOPSLA reviewer 1 about this, but I don't see an easy way.}
 

### 2021-07-08 submission3/main.tex
Nits

--- 2021-07-08 submission3/main.tex
@@
 based security~\citep{sabelfeld2003language}. The informal claim that
-“stack safety protects a caller
-from its callee” amounts to saying that it guarantees the {\em integrity} and
+``stack safety protects a caller
+from its callee'' amounts to saying that it guarantees the {\em integrity} and
 {\em confidentiality} of the caller’s local state until it regains control.
@@
 property, and argue that this property captures the intuitive concept of
-strict stack.
+strict stack
 safety~(\cref{sec:prelim,sec:lse}).
@@
 We say that two observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
-from the perspetive of an external observer, written \(\obsT_1 \eqsim
+from the perspective of an external observer, written \(\obsT_1 \eqsim
 \obsT_2\), if the sequence of non-silent observations is the same. That is, we

### 2021-07-08 submission3/main.tex
pass through 7

--- 2021-07-08 submission3/main.tex
@@
 \label{sec:setup}
+\label{sec:threat}
 
@@
 
-We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
+We now formalize the model of enforcement that we introduced in Section~\ref{sec:threat}.
 A {\em policy} consists of a set of policy
@@
 Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
-as a predicate on addresses.
+as a predicate on addresses.\apt{We never explain what $\ANNS$ is for the code maps in
+  this section.}
 
@@
 \definition
-{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
+\textbf{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
 where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
@@
 \definition
-A system enjoys {\em observational confidentiality} if for all
+A system enjoys \textbf{\em observational confidentiality} if for all
 \(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\pol,\context\) and
@@
 that should be passed, expressed as offsets from the stack pointer.
-%
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \WORDS)\]
-
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P)~\mbox{\rm where}~P \subseteq \WORDS\]
 We extend the set of domains to include passed components,
@@
 
-\definition Let the protected set of a context, \(\prot_P(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+\definition The \emph{protected set} of a context, \(\prot_P(\domMap,\rts)\),
+is the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
 or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
@@
 
-\definition Let the \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
-be the set of components \(\component\) such that
+\definition The \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
+is the set of components \(\component\) such that
 \(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
@@
 \definition
-For each of the stepwise, deferred, and observational versions of
-stack integrity in the simple model, the equivalent property in the
+For each of the \textbf{\em stepwise}, \textbf{\em deferred}, and \textbf{\em observational} versions of
+\textbf{\em stack integrity} in the simple model, the equivalent property in the
 passing model substitutes the above \(\CONTEXTS\) type and update
@@
 
-For both the stepwise and observational versions of stack confidentiality in
+For both the \textbf{\em stepwise} and \textbf{\em observational} versions of \textbf{\em stack confidentiality} in
 the original model, the passing model equivalents additionally substitute \(\secret_P\).
@@
 register to be secret from the callee, except when it is copied to the stack! Both scenarios
-reflect not stepwise but observational stack safety.
-
+reflect observational rather than stepwise stack safety.
+
+\apt{Mention that now for the first time we are putting register components into something other than the outside domain.}
 We begin by adding information to Call annotations defining which registers will be {\em saved}
@@
 %
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(S \in \REGS \rightarrow (\unsaved + \saved))\]
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(S) ~ \mbox{\rm{where}} ~ S \in \REGS \rightarrow (\unsaved + \saved)\]
 %
@@
               {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \in \REGS \mapsto S \reg]\)}
+              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
               {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
@@
 %
-{\bf Observational stack integrity} with callee-saves registers substitutes the above
+\textbf{\em Observational stack integrity} with callee-saves registers substitutes the above
 \(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
 %
-{\bf Observational stack confidentiality} additionally substitutes \(\secret_S\).
+\textbf{\em Observational stack confidentiality} additionally substitutes \(\secret_S\).
 
@@
 and in observational confidentiality terms, both the original and variant values are unchanged
-from their originals, so that \(\reg\) is not in the \(\delta\) set. But \(\addr\) is in the
+from their originals, so that \(\reg\) is not in the \(\delta\) set.\apt{?? what's that?}  But \(\addr\) is in the
 \(\delta\) set, so confidentiality additionally requires that the value of \(\reg\) does not
@@
 
+\apt{Should that be $\passed{\depth,\depth'}$ ?.  And why do we need to be precise about $\STKDOMS$ anyhow?
+  And shouldn't the first component of $\context$ be of type $\COMPONENTS \rightarrow \TOPDOMS$? }
+
 The initial domain map maps each stack address in stack \(\stackid\) to
@@
 This calls for some explanation. First, we determine which coroutine we are yielding to by
-kchecking which stack the stack pointer of the next state is pointing to. Then we can only yield
+checking which stack the stack pointer of the next state is pointing to. Then we can only yield
 if the next state is a valid yield target for that stack, and we update the yield target for
@@
 
-\definition \(\MPCT^\wedge\) is the \emph{yield set}, defined as
+\definition The \emph{yield set} \(\MPCT^\wedge\) of a trace \(\MPCT\) \apt{right?} is
 \[\begin{aligned}
@@
 
+\noindent
 Here, we are capturing the entire span that the system spends {\em away} from a given
@@
 \definition
-Let the protected set of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
-be the set of components \(\component\) such that
+The \emph{protected set} of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
+is the set of components \(\component\) such that
 \(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
@@
 \definition
-Let the secret set of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
-be the set of components \(\component\) such that
+The \emph{secret set} of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
+is the set of components \(\component\) such that
 \(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).

### 2021-07-09 submission3/main.tex
Responding to notes

--- 2021-07-08 submission3/main.tex
@@
 based security~\citep{sabelfeld2003language}. The informal claim that
-``stack safety protects a caller
-from its callee'' amounts to saying that it guarantees the {\em integrity} and
+“stack safety protects a caller
+from its callee” amounts to saying that it guarantees the {\em integrity} and
 {\em confidentiality} of the caller’s local state until it regains control.
@@
 rights. Unless otherwise marked, sequential addresses retain the previous domain,
-as hinted by their colors. Domains are descriptive\bcp{??}; it is up to the enforcement
-mechanism to ensure that the system adheres to these permissions. Such a domain
+as hinted by their colors. Domains describe the intended security role of the address
+that the enforcement mechanism must adhere to. Such a domain
 can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
@@
 
-We now define our machine model \apt{and enforcement model?} precisely. To make our definition of stack safety
+We now define our machine and enforcement model precisely. To make our definition of stack safety
 as generic as possible, this section proposes an abstract interface to a
@@
 individual calls from a trace by taking subtraces in which the stack is
-at or below\apt{``above''?}  a given depth. If \(f\) is still a predicate on states, then
+at or above a given depth. If \(f\) is still a predicate on states, then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
@@
 
-\definition The \emph{call set} \(\MPCT^*\) of a trace \(\MPCT\) \apt{right?} is
+\definition The \emph{call set} \(\MPCT^*\) of a machine is
 \[\begin{aligned}
@@
 
-\apt{intuition for what the call set is?}
+This is the set of all possible calls that be reached from any well-formed initial state.
+
 We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
@@
 until either both return together or one failstops. This implies that
-if the callee reads a secret location, it must failstop. %\apt{right?}\leo{yes}
+if the callee reads a secret location, it must failstop.
 %
@@
 the formal stack-safety property described in the previous
-section. In fact, this mechanism enforces a slightly stronger, inductive version
-of that property which checks for integrity and confidentiality violations at
-every step of the callee's execution trace. We validate this claim \apt{which claim: that it enforces
-  a \emph{stronger} version?}
-using a property-based random testing framework to uncover
-inconsistencies \apt{but are there any?} between the enforcement mechanism and the stack-safety
-property.
+section. We validate this claim 
+using a property-based random testing framework.
 
@@
   at the actual return instruction,
-  the $\PCname$ tag is decremented. \apt{Doesn't the exit sequence retag to $\tagNoDepth$?}
+  the $\PCname$ tag is decremented.
+		\apt{Doesn't the exit sequence retag to $\tagNoDepth$?}
+		\sna{On inspection, what's actually going on here is a sort of per-activation eager tagging.
+		Need to fix that to actually match the paper.}
 %  \leo{Not sure the actual policy
@@
 (with respect to a particular call map), all we need require is that
-the entry sequence be initiated at \apt{immediately following?}  any instruction marked as a call
+the entry sequence immediately follow  any instruction marked as a call
 in the call map. This is achieved by giving these instructions (another) special tag,
@@
 runtime monitor implementing the stack-safety
-micro-policy.\footnote{This code is {\em redacted for review}.\bcp{Why??}} We
+micro-policy. We
   chose the Coq proof assistant as the setting for our implementation
@@
 (3)
- a code map and a stack map \apt{what's that ??}, accurately characterizing the behavior
-  of callers and callees.
+ a code map identifying the locations of calls
 % \end{enumerate}
@@
 generated, we ensure that the appropriate policy tags and code
-annotations are set for the entry or return points. At the same time,
-we allow the generation to sometimes relax those constraints,
-introducing potentially ill-formed flows: this causes our programs
-to failstop when executed in conjunction with the policy monitor,
-but also allows for revealing errors in our setup.\apt{??}
+annotations are set for the entry or return points.% At the same time,
+%we allow the generation to sometimes relax those constraints,
+%introducing potentially ill-formed flows: this causes our programs
+%to failstop when executed in conjunction with the policy monitor,
+%but also allows for revealing errors in our setup.\apt{??}
+\sna{Good idea, but we don't actually do it.}
 
@@
 The conservative policy described in the previous section, while
-testable\apt{is this important?} and enforceable\apt{?? it is an enforcement
-  mechanism so obviously it is enforceable -- whatever it may enforce!}, is rather slow in
+testably enforcing stack safety, is rather slow in
 some important practical cases. \citet{DBLP:conf/sp/RoesslerD18} observe
@@
 \((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
-\apt{Intuitive description of what this means?}
+
+Intuitively, deferred stepwise integrity means that if any components change from
+the a call to its matching return, they will afterward be treated as confidential
+(in the sense of stepwise confidentiality; the return state and a variant would
+execute in lockstep.)
 
@@
 We say that two observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
-from the perspective of an external observer, written \(\obsT_1 \eqsim
+from the perspetive of an external observer, written \(\obsT_1 \eqsim
 \obsT_2\), if the sequence of non-silent observations is the same. That is, we
@@
 
-Together with return-after-call\apt{??}, these two properties give us
+These two properties give us
 {\em Observable Stack Safety}. This is a weaker property

### 2021-07-09 submission3/main.tex
Some work on section 2

--- 2021-07-09 submission3/main.tex
@@
 \end{itemize}
-
 \Cref{sec:relwork,sec:future} discuss related and future work.
@@
 
-This section introduces the intuition behind our formal stack safety properties through a series
+This section explains our formal stack safety properties through a series
 of examples, each presented both in C syntax and in corresponding assembly code for a
-simple two-address machine (see Figure~\ref{fig:int1}). The address of each
-instruction is shown to its left.
+simple two-address machine (see Figure~\ref{fig:int1}).
 The assembly code reflects a simple compilation model in which
@@
 (the machine is word addressed), and addresses increase to the right.
-
-Labeled boxes indicate an assignment of the address to a
-{\em domain}---a label reflecting a particular set of intended access
-rights. Unless otherwise marked, sequential addresses retain the previous domain,
-as hinted by their colors. Domains describe the intended security role of the address
-that the enforcement mechanism must adhere to. Such a domain
-can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
+The address of each instruction is shown to its left.  \bcp{Most of the
+  figure is kind of self-explanatory with this help, but the rightmost part
+  is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
+  numbers in the boxes?  What are these brackets under the boxes, ...}
+
+Boxes labeled $\sealdesc{0}$, $\retptrdesc$, etc., indicate an assignment of
+the corresponding address to the indicated
+{\em domain}, reflecting a particular set of intended access
+rights. Unless otherwise marked, further addresses retain the previous domain,
+as hinted by their colors\bcp{where do we see this happening?  Is it
+  important?}. \proposecut{Domains are descriptive\bcp{??}; it is up to the enforcement
+mechanism to ensure that the system adheres to these
+permissions.}\proposeadd{These domain annotations are not part of the
+machine state; they are used to define which program behaviors are to be
+considered safe.}  The domain of a given address
+can be\bcp{awkward: we say it can be one of three things, but in the picture
+it can also be RetPtr.} \(\unsealed\) (meaning: part of the stack and
+accessible to writes), \(\sealed{\depth}\)
 (in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
-(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
-figures. Not shown are global variables and code, which would be labeled \(\outside\).
+(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the
+figures. Global variables and code, which are not shown in the figures, would
+also be labeled \(\outside\).
 Domains form part of the {\em context} associated with a machine state.
@@
 {\tt main} expects that {\tt x} will be unchanged when it gets control back.
-This gives us our simple, intuitive statement of {\em stepwise integrity}:
+This leads to a simple, intuitive statement of {\em stepwise integrity}:
 if a component is sealed before a step, it will be unchanged afterward. Stepwise
@@
 
-Dually, confidentiality intuitively states that a callee is insensitive to
+Dually, confidentiality states intuitively that a callee is insensitive to
 the context of its call (except for its arguments and any global
 variables).
-We state this as a kind of {\em noninterference} property~\citep{Goguen1982SecurityPA}.
+We state this as a form of {\em noninterference}~\citep{Goguen1982SecurityPA}.
 Noninterference describes knowledge: it states that for system components
@@
 
-We now define our machine and enforcement model precisely. To make our definition of stack safety
+We now define our machine model \apt{and enforcement model?} precisely. To make our definition of stack safety
 as generic as possible, this section proposes an abstract interface to a
@@
 individual calls from a trace by taking subtraces in which the stack is
-at or above a given depth. If \(f\) is still a predicate on states, then
+at or below\apt{``above''?}  a given depth. If \(f\) is still a predicate on states, then
 we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
@@
 
-\definition The \emph{call set} \(\MPCT^*\) of a machine is
+\definition The \emph{call set} \(\MPCT^*\) of a trace \(\MPCT\) \apt{right?} is
 \[\begin{aligned}
@@
 
-This is the set of all possible calls that be reached from any well-formed initial state.
-
+\apt{intuition for what the call set is?}
 We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
@@
 until either both return together or one failstops. This implies that
-if the callee reads a secret location, it must failstop.
+if the callee reads a secret location, it must failstop. %\apt{right?}\leo{yes}
 %
@@
 the formal stack-safety property described in the previous
-section. We validate this claim 
-using a property-based random testing framework.
+section. In fact, this mechanism enforces a slightly stronger, inductive version
+of that property which checks for integrity and confidentiality violations at
+every step of the callee's execution trace. We validate this claim \apt{which claim: that it enforces
+  a \emph{stronger} version?}
+using a property-based random testing framework to uncover
+inconsistencies \apt{but are there any?} between the enforcement mechanism and the stack-safety
+property.
 
@@
   at the actual return instruction,
-  the $\PCname$ tag is decremented.
-		\apt{Doesn't the exit sequence retag to $\tagNoDepth$?}
-		\sna{On inspection, what's actually going on here is a sort of per-activation eager tagging.
-		Need to fix that to actually match the paper.}
+  the $\PCname$ tag is decremented. \apt{Doesn't the exit sequence retag to $\tagNoDepth$?}
 %  \leo{Not sure the actual policy
@@
 (with respect to a particular call map), all we need require is that
-the entry sequence immediately follow  any instruction marked as a call
+the entry sequence be initiated at \apt{immediately following?}  any instruction marked as a call
 in the call map. This is achieved by giving these instructions (another) special tag,
@@
 runtime monitor implementing the stack-safety
-micro-policy. We
+micro-policy.\footnote{This code is {\em redacted for review}.\bcp{Why??}} We
   chose the Coq proof assistant as the setting for our implementation
@@
 (3)
- a code map identifying the locations of calls
+ a code map and a stack map \apt{what's that ??}, accurately characterizing the behavior
+  of callers and callees.
 % \end{enumerate}
@@
 generated, we ensure that the appropriate policy tags and code
-annotations are set for the entry or return points.% At the same time,
-%we allow the generation to sometimes relax those constraints,
-%introducing potentially ill-formed flows: this causes our programs
-%to failstop when executed in conjunction with the policy monitor,
-%but also allows for revealing errors in our setup.\apt{??}
-\sna{Good idea, but we don't actually do it.}
+annotations are set for the entry or return points. At the same time,
+we allow the generation to sometimes relax those constraints,
+introducing potentially ill-formed flows: this causes our programs
+to failstop when executed in conjunction with the policy monitor,
+but also allows for revealing errors in our setup.\apt{??}
 
@@
 The conservative policy described in the previous section, while
-testably enforcing stack safety, is rather slow in
+testable\apt{is this important?} and enforceable\apt{?? it is an enforcement
+  mechanism so obviously it is enforceable -- whatever it may enforce!}, is rather slow in
 some important practical cases. \citet{DBLP:conf/sp/RoesslerD18} observe
@@
 \((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
-
-Intuitively, deferred stepwise integrity means that if any components change from
-the a call to its matching return, they will afterward be treated as confidential
-(in the sense of stepwise confidentiality; the return state and a variant would
-execute in lockstep.)
+\apt{Intuitive description of what this means?}
 
@@
 
-These two properties give us
+Together with return-after-call\apt{??}, these two properties give us
 {\em Observable Stack Safety}. This is a weaker property

### 2021-07-09 submission3/main.tex
Test policy descriptions

--- 2021-07-09 submission3/main.tex
@@
 
-\sna{I am not sure about the MTTF table. Is it as
-  valuable at POPL? Do we need to go into more detail
-  about metholodology if we use it?}
-
-%\begin{table}[]
-%\centering
-%\begin{tabular}{c|c|c}
-%  Bug & Property Violated & MTTF (ms) \\
-%  \hline
-%      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
-%      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
-%      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
-%\end{tabular}
-%\vspace*{1em}
-%\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
-%\vspace*{-2em}
-%\label{tab:bug-table}
-%\end{table}
-
-%The mean-time-to-failure (MTTF) for various bugs can be found in
-%Table~\ref{tab:bug-table}. All experiments were run in a desktop
-%machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
-%RAM.
+\begin{table}[]
+\centering
+\begin{tabular}{c|c|c}
+  Bug & Property Violated & MTTF (ms) \\
+  \hline
+      {\em LOAD\_NO\_CHECK}  & Confidentiality &  \\
+      {\em STORE\_NO\_CHECK} & Integrity &  \\
+      {\em HEADER\_NO\_INIT} & Integrity &  \\
+\end{tabular}
+\vspace*{1em}
+\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+\vspace*{-2em}
+\label{tab:bug-table}
+\end{table}
+
+The mean-time-to-failure (MTTF) for various bugs can be found in
+Table~\ref{tab:bug-table}. All experiments were run in a desktop
+machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
+RAM.
 % \leo{Something like: There is really no baseline to compare this
@@
 same depth. However, this allows violations of observable stack safety!
-Specifically, a callee might write into the stack and leave tags behind
-for a future callee at the same depth to read. This violates confidentiality and,
-if the writes were in the caller's frame, integrity. \apt{A more detailed example
-  would be great here, and should not take too much space.}
-\apt{Mention here that this was found by testing.}
+Our implementation of this policy, {\em PER\_DEPTH\_TAG}, has an error that we find
+in testing. If, at depth \(d\), a callee \(A\) writes into its caller's
+frame, then returns, and later another callee \(B\) at depth \(d\) reads the same
+address and prints it, \(A\) has violated integrity in a way that later becomes visible,
+and \(B\) has likewise violated confidentiality!
 
@@
 each activation is tagged uniquely, and find that it obeys stepwise
-confidentiality and observational integrity. Our testing catches the
-violation in the original {\em Lazy Tagging and Clearing}, as well as
-an equivalent to {\em LOAD\_NO\_CHECK}, which now also violates observational
-integrity.\apt{Move latter to above.}
+confidentiality and observational integrity. To round out our mutations testing 
+we also check an equivalent to {\em LOAD\_NO\_CHECK} and a version where
+stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
+
+\begin{table}[]
+\centering
+\begin{tabular}{c|c|c|c}
+	Bug & Property Violated & MTTF (ms) & Tests \\
+  \hline
+	{\em PER\_DEPTH\_TAG} & Observational Integrity &  & \\
+	{\em LOAD\_NO\_CHECK}  & Observational Integrity &  & \\
+	{\em STORE\_NO\_UPDATE} & Observational Integrity &  & \\
+\end{tabular}
+\vspace*{1em}
+\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+\vspace*{-2em}
+\label{tab:lazy-bug-table}
+\end{table}
 

### 2021-07-09 submission3/main.tex

--- 2021-07-09 submission3/main.tex
@@
 (the machine is word addressed), and addresses increase to the right.
+The address of each instruction is shown to its left.  \bcp{Most of the
+  figure is kind of self-explanatory with this help, but the rightmost part
+  is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
+  numbers in the boxes?  What are these brackets under the boxes, ...}
 
@@
 rights. Unless otherwise marked, sequential addresses retain the previous domain,
-as hinted by their colors. Domains describe the intended security role of the address
+as hinted by their colors \bcp{where do we see this happening?  Is it
+  important?}. Domains are not part of the machine state; rather, they
+describe the intended security role of the address 
 that the enforcement mechanism must adhere to. Such a domain
@@
 (in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
-(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
-figures. Not shown are global variables and code, which would be labeled \(\outside\).
+(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the
+figures. Global variables and code, which are not shown in the figures, would
+also be labeled \(\outside\).  \bcp{what about RetPtr?}
 Domains form part of the {\em context} associated with a machine state.

### 2021-07-09 submission3/main.tex
done with pass

--- 2021-07-09 submission3/main.tex
@@
 rights. Unless otherwise marked, sequential addresses retain the previous domain,
-as hinted by their colors. Domains describe the intended security role of the address
-that the enforcement mechanism must adhere to. Such a domain
+as hinted by their colors. Domains are descriptive\bcp{??}; it is up to the enforcement
+mechanism to ensure that the system adheres to these permissions. Such a domain
 can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
@@
 
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c}
-  Bug & Property Violated & MTTF (ms) \\
-  \hline
-      {\em LOAD\_NO\_CHECK}  & Confidentiality &  \\
-      {\em STORE\_NO\_CHECK} & Integrity &  \\
-      {\em HEADER\_NO\_INIT} & Integrity &  \\
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:bug-table}
-\end{table}
-
-The mean-time-to-failure (MTTF) for various bugs can be found in
-Table~\ref{tab:bug-table}. All experiments were run in a desktop
-machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
-RAM.
+\sna{I am not sure about the MTTF table. Is it as
+  valuable at POPL? Do we need to go into more detail
+  about metholodology if we use it?}
+
+%\begin{table}[]
+%\centering
+%\begin{tabular}{c|c|c}
+%  Bug & Property Violated & MTTF (ms) \\
+%  \hline
+%      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
+%      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
+%      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
+%\end{tabular}
+%\vspace*{1em}
+%\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+%\vspace*{-2em}
+%\label{tab:bug-table}
+%\end{table}
+
+%The mean-time-to-failure (MTTF) for various bugs can be found in
+%Table~\ref{tab:bug-table}. All experiments were run in a desktop
+%machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
+%RAM.
 % \leo{Something like: There is really no baseline to compare this
@@
 same depth. However, this allows violations of observable stack safety!
-Our implementation of this policy, {\em PER\_DEPTH\_TAG}, has an error that we find
-in testing. If, at depth \(d\), a callee \(A\) writes into its caller's
-frame, then returns, and later another callee \(B\) at depth \(d\) reads the same
-address and prints it, \(A\) has violated integrity in a way that later becomes visible,
-and \(B\) has likewise violated confidentiality!
+Specifically, a callee might write into the stack and leave tags behind
+for a future callee at the same depth to read. This violates confidentiality and,
+if the writes were in the caller's frame, integrity. \apt{A more detailed example
+  would be great here, and should not take too much space.}
+\apt{Mention here that this was found by testing.}
 
@@
 each activation is tagged uniquely, and find that it obeys stepwise
-confidentiality and observational integrity. To round out our mutations testing 
-we also check an equivalent to {\em LOAD\_NO\_CHECK} and a version where
-stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
-
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c|c}
-	Bug & Property Violated & MTTF (ms) & Tests \\
-  \hline
-	{\em PER\_DEPTH\_TAG} & Observational Integrity &  & \\
-	{\em LOAD\_NO\_CHECK}  & Observational Integrity &  & \\
-	{\em STORE\_NO\_UPDATE} & Observational Integrity &  & \\
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:lazy-bug-table}
-\end{table}
+confidentiality and observational integrity. Our testing catches the
+violation in the original {\em Lazy Tagging and Clearing}, as well as
+an equivalent to {\em LOAD\_NO\_CHECK}, which now also violates observational
+integrity.\apt{Move latter to above.}
 
@@
 To our knowledge, the only line of work to date that has attempted
-a positive and formal characterization of
-stack safety begins with \citet{SkorstengaardLocal}, who introduced a calling
-convention that uses local capabilities to preserve local state
-encapsulation and well-bracketed control flow
-using a logical relation to reason about the stack safety of concrete
-programs.
+a positive and formal characterization of stack safety
+%begins with \citet{SkorstengaardLocal}, who introduced a calling
+%convention that uses local capabilities to preserve local state
+%encapsulation and well-bracketed control flow
+%using a logical relation to reason about the stack safety of concrete
+%programs.
 % Although the required hardware support is readily available
@@
 %cleared whenever a security boundary is crossed.
-Their logical
-relation captures capability safety without ``externally observable
-side-effects (like console output or memory access traces)''
-and can be used to reason about individual programs.
-
-StkTokens \citep{SkorstengaardSTKJFP} continues this line of work by
-defining a more efficient
-calling convention that uses linear capabilities for stack and return
-pointers.
-Informally, local state encapsulation (LSE) is defined as
-restricting accesses to the range of memory allocated to the current stack
-frame, and well-bracketed control flow (WBCF) is defined as only allowing
-returns from the topmost
-frame to the immediately adjacent frame below. Formally, this work improves on
-\citet{SkorstengaardLocal} by building those properties into the semantics of a
+%Their logical
+%relation captures capability safety without ``externally observable
+%side-effects (like console output or memory access traces)''
+%and can be used to reason about individual programs.
+is that of \citet{SkorstengaardSTKJFP}. Their 
+StkTokens system enforces stack safety on hardware equipped with
+memory capabilities, similar to those used in the CHERI system~\citep{Woodruff+14,Chisnall+15}, but
+with additional support for linear capabilities.
+Stack safety is defined as the conjuncton of local state encapsulation (LSE) and
+well-bracketed control flow (WBCF). 
+Informally, LSE means that the current activiation can only access its own frame, 
+and WBCF means that the topmost frame can only return to 
+the immediately adjacent frame below. Formally, these properties are defined
+with reference to the semantics of a 
 capability machine with a built-in call stack and call and return
 instructions; this machine is proven fully abstract with respect to a more
-concrete capability machine.
+concrete capability machine.  They call this approach a ``fully abstract overlay semantics.''
 The proof of full abstraction uses a standard notion of components, which
@@
 \begin{enumerate}[label=(\roman*)]
-\item Most importantly, our definition captures the intuition that {\em
+\item Our definition captures the intuition that {\em
   every} caller should be protected from its callees.  By contrast,
@@
 to classify such a function as an adversary, and would not guarantee it
-protection. We argue that if individual function activations are well-behaved,
+protection.
+\apt{But see p. 17 para 2 of the JFP version of their paper, where they explicitly
+  dispute this characterization!}
+We argue that if individual function activations are well-behaved,
 they should be protected.
 
-\item \citet{SkorstengaardSTKJFP}'s approach to proving safety via fully
+\item \citet{SkorstengaardSTKJFP}'s approach to proving safety via a fully
   abstract overlay
@@
   to a range of architectures, code features, and enforcement mechanisms.
-
-\vspace{.7\smallskipamount}
-
-  Conversely, it seems possible to extend \citet{SkorstengaardSTKJFP}'s
-  abstract machine to pass arguments on the stack, or to support callee-saves
-  registers. It might even be viable to construct a safe-by-construction abstract
-  coroutine machine in the same style, but this is hypothetical.
-  We have implemented these scenarios with
-  simple modifications to the domain model.
+  Moreover, we have implemented extensions such as passing arguments on the stack,
+  describing callee-saves registers, and coroutines by making simple modifications
+  to our base domain model.
+  %% While it may well be possible to extend 
+  %% \citet{SkorstengaardSTKJFP}'s approach to cover these features.  
+  %% abstract machine to pass arguments on the stack, or to support callee-saves
+  %% registers. It might even be viable to construct a safe-by-construction abstract
+  %% coroutine machine in the same style, but this is hypothetical.
+
 \end{enumerate}

### 2021-07-09 submission3/main.tex
More polishing

--- 2021-07-09 submission3/main.tex
@@
   is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
-  numbers in the boxes?  What are these brackets under the boxes, ...}
+  numbers in the boxes?  What are these brackets under the boxes, ...  One
+  simple thing we could do to help might be to label the right-hand column
+  something like ``Snapshots,'' or at least explain early on in the text
+  that this is what they are.}
 
@@
 This leads to a simple, intuitive statement of {\em stepwise integrity}:
-if a component is sealed before a step, it will be unchanged afterward. Stepwise
-refers to the fact that the property must hold on every step. %Later we will
+if a component is sealed before a step, it will be unchanged afterward.
+% Stepwise
+% refers to the fact that the property must hold on every step.
+ %Later we will
 %loosen this condition in the interest of extensionality---the caller only
@@
 \end{subfigure}
-\caption{A confidentiality violation\ifspace\apt{Could save space by not repeating C or assemly code for {\tt main}.}\fi}
+\caption{A confidentiality violation\ifspace\apt{Could save space by not
+    repeating C or assemly code for {\tt main}.}\fi\bcp{Can we put some more
+  space above the stack trace?}}
 \label{fig:conf1}
@@
 containing secret values, we can substitute any values whatsoever without changing the results
-of our computation. If there exists some value for which we see different results, then
+of our computation.\bcp{That's too loose.  Even for standard NI, high inputs
+are allowed to change high results.  And for us it's even more subtle, since
+a given address can switch from high to low when a function returns.} If there exists some value for which we see different results, then
 in fact we can learn something about the initial state by observing those results, violating
@@
 
-In Figure~\ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
+In Figure~\ref{fig:conf1} we see the callee, {\tt f}, violating confidentiality
 by reading {\tt main}'s variable {\tt x}. The trace
-shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
-state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
-understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
-be hidden from {\tt f}. (We show the evolving stack memory  of the variant in a second parallel column.) At instruction 12, we note that the callee's first variable, {\tt y}, has
+shows how noninterference applies to this scenario. At instruction 10, we
+create an imaginary {\em variant}
+state with arbitrary values\bcp{IMO, it would be easier to
+understand if we chose concrete values for the examples}~(\(v_0\), etc.),
+in place of the actual values ($42$, ...) that should
+be hidden from {\tt f}; we show the evolving stack memory of the variant in a second, parallel column. At instruction 12, we note that the callee's first variable, {\tt y}, has
 changed to 5 in both the primary trace and the variant. The primary and variant agree on
@@
 
-We will term this form of confidentiality as {\em stepwise confidentiality}. Informally,
-for any state at the entry to a function, for any variant over the values in the stack,
-the primary state and its variant will step in lockstep until they return. To step in lockstep
+We can this form of confidentiality {\em stepwise confidentiality}. Informally,
+for any state at the entry to a function, and for any variant over the values in the stack,
+the primary state and the variant will step in lockstep until they
+return. To step in lockstep
 means that if both states step---that is, neither the primary nor the variant failstops---any
 {\em system component} (such as memory cell, register, program counter) that changes
-in one step has a matching value at the end of the other step.
-When we later extend the model with a notion of observable output (\cref{sec:lazy}),  we will instead apply the matching
-criteria to the outputs.
+in the original will change in the same way in the variant.
+When we later extend the model with a notion of observable output
+(\cref{sec:lazy}), we will apply this condition only to the outputs\bcp{what
+outputs?}.
 

### 2021-07-09 submission3/main.tex
Finished pass over 2

@@
 and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
-temporal stack safety.
+{\em temporal stack safety}.
 In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
@@
 In the primary trace, {\tt z} will have the value 5, but in the variant it can have
-an arbitrary value $v_4$; since the variant can behave different from the primary,
-the policy again signals a confidentiality violation.
+an arbitrary value $v_4$.  Since the variant can behave different from the primary,
+this again represents a violation of confidentiality.
 
@@
 Our stack safety properties do not assume that control flow follows a typical structure
-of matching sets of calls and returns nested within on another.
+of matching sets of calls and returns nested within one another.
 Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
 Consider Figure~\ref{fig:controlflow}.
-(Here the diagram shows return addresses explicitly as line numbers, in the form $\#n$.)
+(Here the diagram shows return addresses explicitly as line numbers, in the
+form $\#n$.\bcp{But the return address boxes are not black... should they
+  be, really?})
 Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
@@
 $-1$. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
-This has the appearance of an integrity violation.  How do we capture this in our
+This has the appearance of an integrity violation, so how do we capture it in our
 integrity property?
 
-The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
+The key intuition is that, to be considered as having returned to {\tt main}, {\tt f} needs
 to do more than just execute a return-labeled instruction: it must actually reach the
 appropriate return target in {\tt main}.
-We add a new piece of context to our diagram: we keep a stack of return targets, and a callee is
+To capture this intuition, we add a new piece of context to our diagram:
+a stack of return targets.  Now a callee is
 only considered to have returned when it reaches a state that matches the caller's target.
 A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, represented here by sharing a color with the sealed data.
+when it is reached, represented here by sharing a color with the sealed
+data.  \bcp{Don't understand that last sentence.}
 
@@
 Thus, a formal definition of safe use of the stack should include the
-condition that calls and returns are {\em well bracketed}, e.g., $C$ should
-only return to $B$.  This is what \citet{SkorstengaardSTKJFP} describe
-as \emph{well-bracketed control flow (WBCF)}.
+condition that calls and returns should be correctly nested---e.g., $C$ should
+only return to $B$.  This is what \citet{SkorstengaardSTKJFP}
+call \emph{well-bracketed control flow (WBCF)}.
 

### 2021-07-09 submission3/main.tex
Comment bashing

@@
 Noninterference describes knowledge: it states that for system components
-containing secret values, we can substitute any values whatsoever without changing the results
-of our computation.\bcp{That's too loose.  Even for standard NI, high inputs
-are allowed to change high results.  And for us it's even more subtle, since
-a given address can switch from high to low when a function returns.} If there exists some value for which we see different results, then
+containing secret values, we can substitute any values whatsoever for the secrets without changing the effect
+of our computation on non-secret values.
+% \bcp{That's too loose.  Even for standard NI, high inputs
+% are allowed to change high results.  And for us it's even more subtle, since
+% a given address can switch from high to low when a function returns.}
+If there exists some value for which we see different results, then
 in fact we can learn something about the initial state by observing those results, violating
@@
 When we later extend the model with a notion of observable output
-(\cref{sec:lazy}), we will apply this condition only to the outputs\bcp{what
-outputs?}.
+(\cref{sec:lazy}), we will apply this condition only to those.
 

### 2021-07-09 submission3/main.tex
More bashing

@@
 (in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
-(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the
-figures. Global variables and code, which are not shown in the figures, would
+(outside the stack entirely.)
+% These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
+Global variables and code, which are not shown in the figures, would
 also be labeled \(\outside\).  \bcp{what about RetPtr?}

### 2021-07-09 submission3/main.tex
pass on 1

@@
 proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
-property-based random testing) that Roessler and Dehon's ``eager''
+property-based random testing) that Roessler and DeHon's ``eager''
 micro-policy, which catches violations as early as possible, enforces a
@@
 observational property. Meanwhile our testing successfully detects violations
-in several broken variants, including Roessler and Dehon's original lazy policy.
+in several broken variants, including Roessler and DeHon's original lazy policy.
 
@@
 
-We propose such a characterization using the tools of language-
-based security~\citep{sabelfeld2003language}. The informal claim that
+We propose such a characterization using the tools of language-based
+security~\citep{sabelfeld2003language}. The informal claim that
 ``stack safety protects a caller
@@
 
-We can also, optionally, include a notion of well-bracketed con-
-trol flow as in \citet{SkorstengaardSTKJFP} --- a global requirement that
+We can also, optionally, include a notion of {\em well-bracketed control
+flow (WBCF)} as in \citet{SkorstengaardSTKJFP}---a global requirement that
 callees must always return to their immediate caller, if they return
-at all. Skostengaard et al. define stack safety as the conjunction of two properties:
-{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
+at all. Skostengaard et al. define stack safety as the conjunction of WBCF with
+{\em local state encapsulation (LSE)}.
 Informally, LSE says that the contents of
 a caller's stack frame are not read or written while its callees (and their callees, etc.) are
-executing, and WBCF says that callees always
-return to the instruction following their call (if they return at all).
+executing.
 Skorstengaard et al. formalize these properties by defining an idealized
@@
 integrity properties are together stronger than LSE, guaranteeing protection even
-in the presence of control-flow violations, and can be seen as orthogonal to WBCF.
+in the presence of control-flow violations, and are essentially orthogonal to WBCF.
 
@@
 tool for Coq, to generate random programs and check
-that Roessler and Dehon's micro-policies correctly abort the ones that
+that Roessler and DeHon's micro-policies correctly abort the ones that
 attempt to violate one of our properties. Furthermore, we
@@
 
-We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
+We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
 which memory cells within each stack frame are tagged with the depth of

### 2021-07-09 submission3/main.tex
Skorst nit

--- 2021-07-09 submission3/main.tex
@@
 considered reasonable. \citet{SkorstengaardSTKJFP} would be forced
-to classify such a function as an adversary, and would not guarantee it
+to classify such a function as an adversary, and would not guarantee its
 protection.
@@
   dispute this characterization!}
-We argue that if individual function activations are well-behaved,
-they should be protected.
+In contrast in our setting, individual function well-behaved activations are protected.
 
@@
 \bcp{We could squish all these little paragraphs into one if we are short of
-space.}
+space.}\leo{+1 for squishing even if we're not lacking the space}
 

### 2021-07-09 submission3/main.tex
Answered comment

--- 2021-07-09 submission3/main.tex
@@
 
-%We model ordinary program state and policy-enforcement state separately to make it
-%easier to use well-established concepts from language-based security.
-%In particular, confidentiality has an elegant extensional characterization in terms of
-%\emph{noninterference}: a program preserves the confidentiality of certain data if
-%\emph{varying} that data would not change the program's observable
-%behavior~\citep{6234468}.
-%In applying this idea to a system that incorporates policy enforcement, only
-%ordinary program data should be varied, while any associated policy
-%state should be left alone; our model makes this distinction obvious.
-%That is, we compare the actual execution of the program with a \emph{hypothetical}
-%execution where data has been varied.
-%We will also introduce a notion of {\em observational integrity} that relies on the idea of
-%\emph{rolling back} program state; again, our model makes it easy to describe
-%rollback of ordinary program data while leaving policy state alone.
-
-%% \leo{Policies of interest: null, micro, cherri, software-only}
-%% \leo{TODO: Write down carefully}
-
-%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
-%We believe that it is flexible enough to capture code-altering approaches, such as code
-%rewriting or capability-based techniques, though perhaps with less elegance.
-
-%\paragraph*{Micro-Policies}
-%
-%Micro-Policies are a flexible tag-based, hardware-accelerated
-%reference monitoring mechanism; they been applied to stack safety
-%enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
-%Here the policy state consists of metadata tags (e.g., identifying stack frames)
-%attached to each value in memory or registers, and the policy step
-%function checks that each machine operation obeys a set of rules on tags
-%(e.g., that the current $\PCname$ tag matches the tag on the stack
-%location being accessed), and halts the machine if not.  Distinguishing
-%ordinary machine state and policy state is very natural in this setting.
-%% Here the machine state is extended with tags paired with all components of the
-%% system, and a monitor checks for each operation that the tags on the operands
-%% obey a set of rules. One could treat tags directly as part of the machine
-%% state, and build the rules into the step function. But local state
-%% encapsulation properties as we frame them are not intuitive in their
-%% interaction with tags or other additional enforcement state, so it is helpful
-%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
-%% a software fail state.
-%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
-%%   policy? }
-
-%\iftrue\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
-%These approaches are represented as policies that
-%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
-%being produced by the compiler.) The set of policy states is again the unit type
-%and the step function constant. Property enforcement comes from the fact that
-%only ill-formed states can induce a trace that violates the property.
-
-%% \Paragraph*{Null policy}
-%% %
-%% The simplest possible policy is the neutral policy that carries no information
-%% and does nothing. Its set of policy states is the unit type, and its policy step
-%% function is the constant unit function, thereby allowing all calls to the machine
-%% step function. Its initialization function is defined on all states. This
-%% corresponds to the lifting of our machine model to the policy setting. Machines
-%% that fulfill a property by construction will also fulfill that property with the
-%% null policy.
-
-%
-%\leo{TODO: Add citations of mitigation techniques}
-%\rb{And for those techniques, explain how they reflect their motivating
-%examples. What does each technique do for those examples? How are they
-%connected to this work?}
-%
-%\paragraph*{Bad Write}
-%
-%\[
-%44 : [\rsp - 5] \leftarrow r_4
-%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
-%\]
-%
-%Instruction 44 currently is
-%$ [\rsp - 2] \leftarrow r_4$,
-%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
-%with the
-%effect of storing the contents of $r_4$ as the result of {\tt g}. If
-%instead it wrote to location $\rsp - 5$ then it would be overwriting
-%private data (the return address) belonging to {\tt f}.
-%
-%\paragraph*{Bad Read}
-%
-%\[
-%30 : r_5  \leftarrow [\rsp - 6]
-%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
-%\]
-%
-%Instruction 30 currently is
-%$ r_5 \leftarrow [\rsp - 4]$,
-%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
-%with the
-%effect of loading the argument of {\tt f} to $r_5$. If instead it
-%loaded the contents of location $\rsp - 6$, it would obtain access
-%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
-%outputs the contents of $r_5$, that value would immediately be made public.
-%
-%\paragraph*{Bad Control Flow}
-%
-%\[
-%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
-%%5 : \jalr ~ r_4 ~ \rra ~ 0
-%\]
-%If a {\jalr} instruction was added to {\tt main} with its target being
-%the middle of the code block of {\tt g}.\apt{describe result of this?}
-
 \section{Machines and Traces}
@@
 
-%\subsection{Well-bracketed Control Flow}
-%\label{sec:wbcf}
-
-%Both the integrity and confidentiality components of stack safety
-%are reasonable even in programs that violate expected
-%control flow. But even though stack safety prevents control-flow violations
-%from being used to gain privileges, it is worth protecting control flow as well.
-%Indeed, all of the enforcement mechanisms we discuss do preserve well-bracketed
-%control flow.
-
-%When we focus on subroutines, the essence of well-bracketed control flow
-%is that a caller can expect that if its callee ever returns, it returns
-%to the caller at the appropriate return target. This is tantamount to saying
-%that we never ``skip'' a return, or perform a ``false'' one. To formalize this in
-%the simple model:
-
-%\definition A system enjoys {\em well-bracketed control flow} if for all
-%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-%\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\underscore, \underscore::\rts)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore, \rts)\).
-
-%Extending this idea to coroutines, we use the same notion that there are
-%no ``false'' yields.
-
-%\definition A system enjoys {\em well-structured control flow} if for all
-%reachable states \(\mpcstatename\) and \(\mpcstatename'\) such that
-%\(\mpcstatename \stepstocon \mpcstatename'\), if
-%\(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid)\) and \(\pi_m(\mpcstatename)\)
-%is annotated as a return, then \(\pi_c(\mpcstatename) = (\underscore,\underscore,\underscore,\stackid')\)
-%where \(\stackid \not = \stackid'\).
-
 \section{Case Study: Do Micro-Policies Enforce Stack Safety?}
 \label{sec:enforcement}
-
-% \bcp{``Enforcement'' seems like too broad a title.  What about calling it
-%   ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
-%   Safety?''? And what about folding the testing section into this one?  That
-% would keep related things together, would avoid appearing to make overly
-% grand claims in either section, and would give us an excuse to write less
-% about the mechanism here---we only need to write enough about the mechanism
-% itself that readers can understand our validation case study.}
-% \leo{Agreed. Took a crack at it.}
-
-% \leo{I am not sure we've hit the correct level of detail in this
-%   section.}\bcp{Agreed!}
 
@@
 considered reasonable. \citet{SkorstengaardSTKJFP} would be forced
-to classify such a function as an adversary, and would not guarantee its
+to classify such a function as an adversary, and would not guarantee it
 protection.
@@
   dispute this characterization!}
-In contrast in our setting, individual function well-behaved activations are protected.
+\sna{But on p. 30, Definition 5, the JFP version still clearly defines that to be reasonable,
+  a component must {\em always} behave reasonably. And immediately after states that the
+  adversarial code can {\em never} be given LSE and WBCF guarantees. I suppose I'll cite that
+  definition directly.}
+We argue that if individual function activations are well-behaved,
+they should be protected.
 
@@
 
-\bcp{We could squish all these little paragraphs into one if we are short of
-space.}\leo{+1 for squishing even if we're not lacking the space}
-
 For criterion (1), intuitiveness, we contend that our property definitions are at least as

### 2021-07-09 submission3/main.tex
Actually fixed merge

@@
 they should be protected.
+In contrast in our setting, individual function well-behaved activations are protected.
 

### 2021-07-09 submission3/main.tex
pass on 2

--- 2021-07-09 submission3/main.tex
@@
 simple two-address machine (see Figure~\ref{fig:int1}).
+The address of each instruction is shown to its left.
 The assembly code reflects a simple compilation model in which
@@
 implicitly adjusting the stack pointer.
-At the right of each example, we show the layout of stack memory just before
-the specified instruction is executed; each box represents a memory word
+At the right of each example, we show snapshots of the layout of stack memory just before
+selected instructions (labeled by their addresses) are executed; each box represents a memory word
 (the machine is word addressed), and addresses increase to the right.
-The address of each instruction is shown to its left.  \bcp{Most of the
-  figure is kind of self-explanatory with this help, but the rightmost part
-  is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
-  numbers in the boxes?  What are these brackets under the boxes, ...  One
-  simple thing we could do to help might be to label the right-hand column
-  something like ``Snapshots,'' or at least explain early on in the text
-  that this is what they are.}
-
-Labeled boxes indicate an assignment of the address to a
-{\em domain}---a label reflecting a particular set of intended access
-rights. Unless otherwise marked, sequential addresses retain the previous domain,
-as hinted by their colors \bcp{where do we see this happening?  Is it
-  important?}. Domains are not part of the machine state; rather, they
+  %% \bcp{Most of the
+  %% figure is kind of self-explanatory with this help, but the rightmost part
+  %% is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
+  %% numbers in the boxes?  What are these brackets under the boxes, ...  One
+  %% simple thing we could do to help might be to label the right-hand column
+  %% something like ``Snapshots,'' or at least explain early on in the text
+  %% that this is what they are.}
+
+Labels under the boxes indicate an assignment of the address to a
+{\em domain}, which is a particular set of intended access
+rights. Unless otherwise marked, each address retains its domain as
+sequential instructions execute. %, as hinted by their colors.
+%\bcp{where do we see this happening?  Is it important?}.
+Domains are not part of the machine state; rather, they
 describe the intended security role of the address
@@
 \end{subfigure}
+\\
+\vspace*{0.3in}
 \begin{subfigure}{.59\textwidth}
@@
 \caption{A confidentiality violation\ifspace\apt{Could save space by not
-    repeating C or assemly code for {\tt main}.}\fi\bcp{Can we put some more
-  space above the stack trace?}}
+    repeating C or assemly code for {\tt main}.}\fi}
 \label{fig:conf1}
@@
   be, really?})
-Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
-3 and 5, respectively. But the first time, before returning, it stashes its
-return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
-up being called a third time, at which point it finally returns to instruction 5.\apt{The line numbers are wrong.}
+Function {\tt f} is called twice, at instructions 3 and 5; it should therefore return to instructions
+4 and 6, respectively. But the first time, before returning, it stashes its
+return pointer away, and the second time it returns to instruction 4 instead of 6; then it ends
+up being called a third time, at which point it finally returns to instruction 6.
 
@@
 A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, represented here by sharing a color with the sealed
-data.  \bcp{Don't understand that last sentence.}
+when it is reached, represented here by coloring both the target and its sealed data with the
+same color (blue).
 
 In the example, on its first return, {\tt f} reaches its return target, because the
-\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
-(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
+\(\PCname\) is 4 after its return. But {\tt f}'s second return is to the wrong instruction
+(4 rather than 6) because the stack pointer has been overwritten (shown in light red).
 Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.\apt{Ditto}
+violates the integrity property.
 
@@
   dispute this characterization!}
-In contrast in our setting, individual function well-behaved activations are protected.
+We argue that if individual function activations are well-behaved,
+they should be protected.
 
@@
 \bcp{We could squish all these little paragraphs into one if we are short of
-space.}\leo{+1 for squishing even if we're not lacking the space}
+space.}
 

### 2021-07-09 submission3/main.tex
wibble

@@
 the stack to disrupt the program’s control flow. Our stack safety
-properties demand that, even in the presence of such attackers, confi-
-dentiality and integrity still apply to data in suspended stack frames
+properties demand that, even in the presence of such attackers, confidentiality
+and integrity still apply to data in suspended stack frames
 until the callee returns.

### 2021-07-09 submission3/main.tex
pass through 3

@@
 
-We will introduce a formal model of stack safety that abstracts over
-machine details. But first we must discuss the bounds of our assumptions ---
+Before introducing our formal model of stack safety, we describe our assumptions:
 how broadly we can abstract over machines, how we represent enforcement, and
@@
 We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. But hardware and timing attacks are out of scope, as are
+arbitrary machine code. \apt{But don't they need to be well-formed?? In which case,
+  this should go at the end of the section, after that has been explained. [Earlier
+    I was only arguing for the explanatory background to be placed earlier, not
+    the actual statement of the threat model.]} But hardware and timing attacks are out of scope, as are
 are attacks on the heap---though we believe that the work of
@@
 That said, our model is meant to apply generically to any enforcement mechanism,
-including those that do not take advantage of this separate.
+including those that do not take advantage of this separation.
 Some mechanisms, like hardware capability systems, might exhibit failstop behavior,
-but have no protected policy state --- this is represented by a ``unit'' policy
-that failstops when the hardware fault. Others might never even failstop, such as
+but have no protected policy state---this is represented by a trivial policy
+that failstops when the hardware faults. Others might never even failstop, such as
 code rewriting techniques that turn stack safety violations into other (erroneous, but not
 stack-unsafe) behaviors. Then the policy-enhanced machine is identical to the original.
-\sna{I'm thinking of SFI here, but that's not stacks, right? Does Softbound work like
-  this.}
+\sna{I'm thinking of SFI here\apt{yes}, but that's not stacks, right? Does Softbound work like
+  this.\apt{dunno}}
 

### 2021-07-09 submission3/main.tex
Skirstengaard talk

--- 2021-07-09 submission3/main.tex
@@
 simple two-address machine (see Figure~\ref{fig:int1}).
-The address of each instruction is shown to its left.
 The assembly code reflects a simple compilation model in which
@@
 implicitly adjusting the stack pointer.
-At the right of each example, we show snapshots of the layout of stack memory just before
-selected instructions (labeled by their addresses) are executed; each box represents a memory word
+At the right of each example, we show the layout of stack memory just before
+the specified instruction is executed; each box represents a memory word
 (the machine is word addressed), and addresses increase to the right.
-  %% \bcp{Most of the
-  %% figure is kind of self-explanatory with this help, but the rightmost part
-  %% is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
-  %% numbers in the boxes?  What are these brackets under the boxes, ...  One
-  %% simple thing we could do to help might be to label the right-hand column
-  %% something like ``Snapshots,'' or at least explain early on in the text
-  %% that this is what they are.}
-
-Labels under the boxes indicate an assignment of the address to a
-{\em domain}, which is a particular set of intended access
-rights. Unless otherwise marked, each address retains its domain as
-sequential instructions execute. %, as hinted by their colors.
-%\bcp{where do we see this happening?  Is it important?}.
-Domains are not part of the machine state; rather, they
+The address of each instruction is shown to its left.  \bcp{Most of the
+  figure is kind of self-explanatory with this help, but the rightmost part
+  is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
+  numbers in the boxes?  What are these brackets under the boxes, ...  One
+  simple thing we could do to help might be to label the right-hand column
+  something like ``Snapshots,'' or at least explain early on in the text
+  that this is what they are.}
+
+Labeled boxes indicate an assignment of the address to a
+{\em domain}---a label reflecting a particular set of intended access
+rights. Unless otherwise marked, sequential addresses retain the previous domain,
+as hinted by their colors \bcp{where do we see this happening?  Is it
+  important?}. Domains are not part of the machine state; rather, they
 describe the intended security role of the address
@@
 \end{subfigure}
-\\
-\vspace*{0.3in}
 \begin{subfigure}{.59\textwidth}
@@
 \caption{A confidentiality violation\ifspace\apt{Could save space by not
-    repeating C or assemly code for {\tt main}.}\fi}
+    repeating C or assemly code for {\tt main}.}\fi\bcp{Can we put some more
+  space above the stack trace?}}
 \label{fig:conf1}
@@
   be, really?})
-Function {\tt f} is called twice, at instructions 3 and 5; it should therefore return to instructions
-4 and 6, respectively. But the first time, before returning, it stashes its
-return pointer away, and the second time it returns to instruction 4 instead of 6; then it ends
-up being called a third time, at which point it finally returns to instruction 6.
+Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
+3 and 5, respectively. But the first time, before returning, it stashes its
+return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
+up being called a third time, at which point it finally returns to instruction 5.\apt{The line numbers are wrong.}
 
@@
 A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, represented here by coloring both the target and its sealed data with the
-same color (blue).
+when it is reached, represented here by sharing a color with the sealed
+data.  \bcp{Don't understand that last sentence.}
 
 In the example, on its first return, {\tt f} reaches its return target, because the
-\(\PCname\) is 4 after its return. But {\tt f}'s second return is to the wrong instruction
-(4 rather than 6) because the stack pointer has been overwritten (shown in light red).
+\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
+(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
 Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.
+violates the integrity property.\apt{Ditto}
 
@@
   every} caller should be protected from its callees.  By contrast,
-\citet{SkorstengaardSTKJFP}
-statically divide the world into trusted functions, which feature additional
-dynamic ``reasonableness'' constraints on their behavior and for which their
-properties hold, and adversary functions, which may violate those
-constraints but are not guaranteed safety. This static division
-is insufficient when we consider buggy, but not inherently adversarial,
-code --- one can imagine a function that, when called with the wrong
+  \citet{SkorstengaardSTKJFP} describe a class of ``unreasonable'' code components
+  that are not protected, because the are capable of dynamic behavior that can compromise
+  their own security. Unreasonable code can be viewed as adversarial, but it is equally
+  likely to simply be buggy. One can imagine a function that, when called with the wrong
   inputs, behaves unreasonably, but under most circumstances can be
-considered reasonable. \citet{SkorstengaardSTKJFP} would be forced
-to classify such a function as an adversary, and would not guarantee its
-protection.
-\apt{But see p. 17 para 2 of the JFP version of their paper, where they explicitly
-  dispute this characterization!}
-\sna{But on p. 30, Definition 5, the JFP version still clearly defines that to be reasonable,
-  a component must {\em always} behave reasonably. And immediately after states that the
-  adversarial code can {\em never} be given LSE and WBCF guarantees. I suppose I'll cite that
-  definition directly.}
-We argue that if individual function activations are well-behaved,
-they should be protected.
-In contrast in our setting, individual function well-behaved activations are protected.
+  considered reasonable. \citet{SkorstengaardSTKJFP} would not consider this a
+  {\em reasonable component} as defined in their section 4.3.3, and so
+  would not guarantee its protection. In contrast, in our setting, individual
+  well-behaved activations of the buggy function would be protected.
+  \citet{SkorstengaardSTKJFP} note that they support mutually distrustful
+  components, but we go a step further. One can think of our hypothetical
+  function as ``self-distrustful'' in that the safety of a well-behaved activation
+  is not compromised by another, ill-behaved one.
 

### 2021-07-09 submission3/main.tex
wibbles

@@
               {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
+\apt{Typesetting problem above}
 The remaining rules are the same as in the simple model.
@@
   considered reasonable. \citet{SkorstengaardSTKJFP} would not consider this a
-  {\em reasonable component} as defined in their section 4.3.3, and so
+  {\em reasonable component} (as defined in their section 4.3.3), and so
   would not guarantee its protection. In contrast, in our setting, individual
@@
   function as ``self-distrustful'' in that the safety of a well-behaved activation
-  is not compromised by another, ill-behaved one.
+  is not compromised by another, ill-behaved activiation of the same function.
 

### 2021-07-09 submission3/main.tex
Addess taking

--- 2021-07-09 submission3/main.tex
@@
 stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike with pass-by-reference, there is no easy way to model the intended
+Unlike with pass-by-reference, there is no obvious way to model the intended
 sharing behavior of such pointers in terms of stack frame activation patterns.
-The best approximation we can make is to treat an address-taken variable in the
+In general we expect such a pointer to behave similarly to other pointers in
+the system, except that its lifespan is connected to that of the frame that contains
+it. In order to give a satisfying property, then, we would tie it to an existing (e.g.)
+heap safety property.
+
+The default otherwise would be to treat an address-taken variable in the
 same way as a passed argument, with the additional property
@@
 This prevents most violations of temporal safety, although stale pointers
-can still be used in certain corner cases.
-More seriously, callees are permitted to access the location without having
-been explicitly given a pointer to it.  Preventing the latter kind of abuse
-is the provence of a fine-grained memory safety policy, e.g., based on
-capabilities
-(implemented by tags or otherwise); we would argue it is beyond the scope of
-stack safety \emph{per se}.\apt{Could try to incorporate some of
-  our response to OOPSLA reviewer 1 about this, but I don't see an easy way.}
+can still be used in certain corner cases. Callees would be permitted
+to access the location without having been explicitly given a pointer
+to it.
+\sna{This is trying to get toward what we were saying at OOPSLA: it's not
+  a technical obstacle, it's that the behavior is tied to things we don't model.}
 

### 2021-07-09 submission3/main.tex
pass through 5

--- 2021-07-09 submission3/main.tex
@@
 that doesn't share between caller and callee: arguments and results are passed
-in registers. To accommodate different machine architectures and different calling
+in registers. The code map just marks the locations of calls.
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap\]
+To accommodate different machine architectures and different calling
 conventions, we parameterize over the structure of the stack, particularly
@@
 Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
-as a predicate on addresses.\apt{We never explain what $\ANNS$ is for the code maps in
-  this section.}
+as a predicate on addresses.
 
@@
               {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
-
 \vspace*{-3.5ex}
@@
          {\(\underscore,\context \stepsto_C \context\)}
-\apt{typesetting problem above}
+\apt{typesetting problem above. Also, I don't understand what ReturnRec is for!}
 

### 2021-07-09 submission3/main.tex
Policy description

--- 2021-07-09 submission3/main.tex
@@
 
-\item From callee back to caller: %when execution is about to reach a
-%  return point.
-  %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
-  at the actual return instruction,
+\item From callee back to caller: the callee clears its frame, and at
+  the actual return instruction,
   the $\PCname$ tag is decremented.
-                \apt{Doesn't the exit sequence retag to $\tagNoDepth$?}
-                \sna{On inspection, what's actually going on here is a sort of per-activation eager tagging.
-                Need to fix that to actually match the paper.}
+
 %  \leo{Not sure the actual policy

### 2021-07-09 submission3/main.tex
Small changes

@@
 
-%  \leo{Not sure the actual policy
-%    does that now (retag unused). But I don't think it's a
-%    confidentiality or integrity violation.}
-%  \rb{Correct, and it doesn't. Some versions of the policy do clear
-%    the stack, retagging would be easily added.}
-
 \end{itemize}
@@
 %but also allows for revealing errors in our setup.\apt{??}
-\sna{Good idea, but we don't actually do it.}
 
@@
 
-\apt{Mention that now for the first time we are putting register components into something other than the outside domain.}
 We begin by adding information to Call annotations defining which registers will be {\em saved}
 and which are {\em unsaved}. The stack pointer and the register containing the return address
-should never be saved. We then extend the set of domains with these values.
+should never be saved. We then extend the set of domains with these values, for the first time applying
+(non-{\em outside}) domains to registers as well as addresses.
 %

### 2021-07-09 submission3/main.tex
clear comment

--- 2021-07-09 submission3/main.tex
@@
 
-\item From callee back to caller: the callee clears its frame, and at
-  the actual return instruction,
+\item From callee back to caller: %when execution is about to reach a
+%  return point.
+  %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
+  at the actual return instruction,
   the $\PCname$ tag is decremented.
-
+                \apt{Doesn't the exit sequence retag to $\tagNoDepth$?}
+                \sna{On inspection, what's actually going on here is a sort of per-activation eager tagging.
+                Need to fix that to actually match the paper.}
 %  \leo{Not sure the actual policy
@@
 the system, except that its lifespan is connected to that of the frame that contains
-it. In order to give a satisfying property, then, we would tie it to an existing (e.g.)
+it. In order to give a satisfactory property, then, we would tie it to an existing (e.g.)
 heap safety property.
@@
 to it.
-\sna{This is trying to get toward what we were saying at OOPSLA: it's not
-  a technical obstacle, it's that the behavior is tied to things we don't model.}
-
+%\sna{This is trying to get toward what we were saying at OOPSLA: it's not
+%  a technical obstacle, it's that the behavior is tied to things we don't model.}
 \paragraph{Formalization of variable passing.}

### 2021-07-09 submission3/main.tex
Another comment

@@
 
-We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code. \apt{But don't they need to be well-formed?? In which case,
-  this should go at the end of the section, after that has been explained. [Earlier
-    I was only arguing for the explanatory background to be placed earlier, not
-    the actual statement of the threat model.]} But hardware and timing attacks are out of scope, as are
-are attacks on the heap---though we believe that the work of
-\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
-to protect both stack and heap.
-
 \paragraph*{Machine model}
@@
 execute the blessed sequence out of order.
+
+\paragraph*{Threat Model}
+
+We adopt a strong threat model in which both caller and callee may execute
+arbitrary machine code, subject to whatever well-formedness conditions the
+enforcement mechanism demands. But hardware and timing attacks are out of scope, as are
+are attacks on the heap---though we believe that the work of
+\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
+to protect both stack and heap.
 
@@
 \begin{tabular}{c|c|c}
-  Bug & Property Violated & MTTF (ms) \\
+  Bug & Property Violated & MTTF (ms) & Tests \\
   \hline
@@
 The mean-time-to-failure (MTTF) for various bugs can be found in
-Table~\ref{tab:bug-table}. All experiments were run in a desktop
-machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
-RAM.
+Table~\ref{tab:bug-table}, along with the average number of tests
+it took to find the failure. Experiments were run in a laptop
+machine equipped with i7-8550U CPU @ 1.8GHz with 32GB RAM.
 % \leo{Something like: There is really no baseline to compare this
@@
 the system, except that its lifespan is connected to that of the frame that contains
-it. In order to give a satisfactory property, then, we would tie it to an existing (e.g.)
+it. In order to give a satisfying property, then, we would tie it to an existing (e.g.)
 heap safety property.
@@
 to it.
-%\sna{This is trying to get toward what we were saying at OOPSLA: it's not
-%  a technical obstacle, it's that the behavior is tied to things we don't model.}
+\sna{This is trying to get toward what we were saying at OOPSLA: it's not
+  a technical obstacle, it's that the behavior is tied to things we don't model.}
+
 \paragraph{Formalization of variable passing.}
@@
 
-\definition The \emph{yield set} \(\MPCT^\wedge\) of a trace \(\MPCT\) \apt{right?} is
+\definition The \emph{yield set} \(\MPCT^\wedge\) of a system is
 \[\begin{aligned}

### 2021-07-09 submission3/main.tex
Fixed typesetting
fix build
More fixes

@@
 Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
-as a predicate on addresses.
+as a predicate on addresses. If a target is reached anywhere in the return stack, {\sc ReturnRec}
+unseals each frame down to the level that was returned to.
 
@@
 \vspace*{-1ex}
-\judgmentthree[ ReturnFound]
+\judgmenttwo[ ReturnFound]
               {\(\mach \stepsto \mach' \in \target\)}
               {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap' = \domMap[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap[\components \mapsto \unsealed],\rts\)}
 
 \vspace*{-1ex}
-\judgmentthree[ ReturnRec]
+\judgmenttwo[ ReturnRec]
             {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
             {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
+            {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'[\components \mapsto \unsealed],\rts'\)}
 \vspace*{-3.5ex}
@@
          {\(\underscore,\context \stepsto_C \context\)}
-\apt{typesetting problem above. Also, I don't understand what ReturnRec is for!}
 
@@
 \centering
-\begin{tabular}{c|c|c}
+\begin{tabular}{c|c|c|c}
   Bug & Property Violated & MTTF (ms) & Tests \\
@@
 
-\judgmentthree[ Call]
+\judgmenttwobrlong[ Call]
                   {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
@@
                   {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-\apt{Typesetting problem above}
 The remaining rules are the same as in the simple model.
@@
 and in observational confidentiality terms, both the original and variant values are unchanged
-from their originals, so that \(\reg\) is not in the \(\delta\) set.\apt{?? what's that?}  But \(\addr\) is in the
-\(\delta\) set, so confidentiality additionally requires that the value of \(\reg\) does not
+from their originals, so that \(\reg\) is not in the \(\Delta\) set.  But \(\addr\) is in the
+\(\Delta\) set, so confidentiality additionally requires that the value of \(\reg\) does not
 leak through the callee-saves mechanism, for instance if the caller allocates but fails to
@@
 %
-\[\stackDom \in \STKDOMS ::= \unsealed + \sealed{\depth} + \passed{\depth-\depth'}\]
 \[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
 \[(\domMap,\rtm,\yts,\stackid) = \context \in \CONTEXTS ::=
- (\COMPONENTS \rightarrow \STKDOMS) \times
+ (\COMPONENTS \rightarrow \TOPDOMS) \times
 (\STACKS \rightarrow \listT(\target)) \times (\STACKS \rightarrow \target) \times \STACKS\]
-
-\apt{Should that be $\passed{\depth,\depth'}$ ?.  And why do we need to be precise about $\STKDOMS$ anyhow?
-  And shouldn't the first component of $\context$ be of type $\COMPONENTS \rightarrow \TOPDOMS$? }
 

### 2021-07-09 submission3/main.tex
wordsmithing in 5

@@
 A call annotation tells us to push the return target onto the return stack and
-seal all components with the depth of the previous return stack. If we aren't making a call,
+seal all components with the depth of the previous return stack.
+(Recall that the code map \(\codemap\) is a partial function, and we can use its domain
+as a predicate on addresses.) If we aren't making a call,
 we identify a return by looking ahead and checking if we're about to step to a state
-that matches the topmost return target. If neither is true, the context remains unchanged.
-Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
-as a predicate on addresses. If a target is reached anywhere in the return stack, {\sc ReturnRec}
+that matches the topmost return target ({\sc ReturnFound}), and unsealing the caller's frame if so. 
+If a target is reached anywhere in the return stack, {\sc ReturnRec}
 unseals each frame down to the level that was returned to.
+If none of these situations hold, the context remains unchanged.
 
@@
             {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'[\components \mapsto \unsealed],\rts'\)}
-\vspace*{-3.5ex}
+\vspace*{-2.5ex}
 \judgment[ Default]

### 2021-07-09 submission3/main.tex
Some test numbers and some comments fixed

--- 2021-07-09 submission3/main.tex
@@
 Global variables and code, which are not shown in the figures, would
-also be labeled \(\outside\).  \bcp{what about RetPtr?}
+also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
+are shaded black and ignored for simplicity for now --- they will
+be treated properly later.
 Domains form part of the {\em context} associated with a machine state.
@@
 create an imaginary {\em variant}
-state with arbitrary values\bcp{IMO, it would be easier to
-understand if we chose concrete values for the examples}~(\(v_0\), etc.),
+state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
+  understand if we chose concrete values for the examples}\sna{FWIW I agree in
+  theory, but it looks bad every time I try it.\fi~(\(v_0\), etc.),
 in place of the actual values ($42$, ...) that should
@@
 Consider Figure~\ref{fig:controlflow}.
-(Here the diagram shows return addresses explicitly as line numbers, in the
-form $\#n$.\bcp{But the return address boxes are not black... should they
-  be, really?})
+(Now we treat return addresses properly, but write them as line numbers, in the
+form $\#n$.)
 Function {\tt f} is called twice, at instructions 3 and 5; it should therefore return to instructions
@@
 \paragraph*{Enforcement Model}
-\apt{What are our ambititions for genericity over enforcement?}
 
@@
 \begin{tabular}{c|c|c|c}
-  Bug & Property Violated & MTTF (ms) & Tests \\
+  Bug & Property Violated & MTTF (s) & Tests \\
   \hline
-      {\em LOAD\_NO\_CHECK}  & Confidentiality &  \\
-      {\em STORE\_NO\_CHECK} & Integrity &  \\
-      {\em HEADER\_NO\_INIT} & Integrity &  \\
+      {\em LOAD\_NO\_CHECK}  & Confidentiality & 24.2 & 13.3 \\
+      {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
+      {\em HEADER\_NO\_INIT} & Integrity &  & \\
 \end{tabular}
@@
 Table~\ref{tab:bug-table}, along with the average number of tests
-it took to find the failure. Experiments were run in a laptop
-machine equipped with i7-8550U CPU @ 1.8GHz with 32GB RAM.
+it took to find the failure. Experiments were run in a desktop
+machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
 % \leo{Something like: There is really no baseline to compare this

### 2021-07-09 submission3/main.tex
wibble

--- 2021-07-09 submission3/main.tex
@@
 
-In our context update rules, calls and returns work the same as in the sharing model,
+In our context update rules, calls and returns work the same as in the subroutine models,
 with domain updates restricted to the top-level domain corresponding to the active stack,

### 2021-07-09 submission3/main.tex
squeezing a bit

--- 2021-07-09 submission3/main.tex
@@
 \maketitle
+
+\newcommand{\paragraphx}[1]{\emph{#1.}}
 
@@
 
-\paragraph{Formal Stack Safety on Capability Machines}
+\paragraphx{Formal Stack Safety on Capability Machines}
 %
@@
 dynamic code generation or multithreading.
-
+%
 For criterion (1), intuitiveness, we contend that our property definitions are at least as
@@
 notions of security.
-
+%
 We agree that criterion (2), usefulness of the safety criterion
@@
 plan to apply our properties to justifying a low-level program logic.
-
+%
 StkTokens
@@
 our work.
-
+%
 With respect to criterion (4), capturing programmers' informal
@@
 that can be applied to a broader range of enforcement mechanisms and code features.
-
+%
 As for criterion (5), broad applicability, we have demonstrated that our approach extends to
@@
 more challenging, but StkTokens does not yet handle this either.
-
+%
 Finally, we would propose one more criterion: having a generic definition
@@
 
-\paragraph{Protecting the Stack with Micro-Policies}
+\paragraphx{Protecting the Stack with Micro-Policies}
 %
@@
 
-\paragraph{Heap Safety as a Security Property}
-
+\paragraphx{Heap Safety as a Security Property}
+%
 Heap safety, like stack safety, may be framed as a security property using
@@
 
-\paragraph*{Stack Safety in the Cheri Ecosystem}
-
+\paragraphx{Stack Safety in the Cheri Ecosystem}
+%
 There are several proposals around the use of Cheri capabilities to enforce stack safety,
@@
 
-\paragraph*{A Program Logic for Stacks}
-
+\paragraphx{A Program Logic for Stacks}
+%
 Stack integrity suggests a reasoning principle in which knowledge about a

### 2021-07-09 submission3/main.tex
Most of tests

--- 2021-07-09 submission3/main.tex
@@
       {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
-      {\em HEADER\_NO\_INIT} & Integrity &  & \\
+      {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
 \end{tabular}
@@
 \begin{tabular}{c|c|c|c}
-        Bug & Property Violated & MTTF (ms) & Tests \\
+        Bug & Property Violated & MTTF (s) & Tests \\
   \hline
-        {\em PER\_DEPTH\_TAG} & Observational Integrity &  & \\
-        {\em LOAD\_NO\_CHECK}  & Observational Integrity &  & \\
-        {\em STORE\_NO\_UPDATE} & Observational Integrity &  & \\
+        {\em PER\_DEPTH\_TAG} & Observational Integrity & 70 & \\
+        {\em LOAD\_NO\_CHECK}  & Observational Integrity & 23.5 & 12 \\
+        {\em LOAD\_NO\_CHECK}  & Confidentiality & 19.2 & 695.5 \\
+        {\em STORE\_NO\_UPDATE} & Observational Integrity & 80.6 &  \\
+        {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
 \end{tabular}

### 2021-07-09 submission3/main.tex
Filled in testing

@@
 stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
+It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
+taking twice as long to catch as the next longest.
 
@@
   \hline
-        {\em PER\_DEPTH\_TAG} & Observational Integrity & 70 & \\
-        {\em LOAD\_NO\_CHECK}  & Observational Integrity & 23.5 & 12 \\
+        {\em PER\_DEPTH\_TAG} & Observational Integrity & 189.7 & 8342.5  \\
+        {\em LOAD\_NO\_CHECK}  & Observational Integrity & 23.5 & 12.0 \\
         {\em LOAD\_NO\_CHECK}  & Confidentiality & 19.2 & 695.5 \\
-        {\em STORE\_NO\_UPDATE} & Observational Integrity & 80.6 &  \\
+        {\em STORE\_NO\_UPDATE} & Observational Integrity & 70 & 80.6  \\
         {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\

### 2021-07-09 submission3/main.tex
Space saving
Ditched para

@@
 
-Intuitively, stack safety means that the stack frame of each caller is protected
-from attempts by a confused or malicious callee (including all its
-descendant callees) to read or write its private data.
-That is, the attacker model treats every caller as a defender and its callee as
-an attacker, meaning that with nested calls, a single function invocation
-may be viewed in both ways.
-
 The protections offered by such mechanisms are commonly described in terms
@@
 and untouchable by application code.
-%Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
-%mirrors the real one, and a return may only occur if both stacks agree on the
-%target address. Typically the shadow stack is isolated using a separate protection
-%mechanism.
-\sna{Commented out shadow stacks until I can find a citation. May just skip it.}
 
@@
 stack-unsafe) behaviors. Then the policy-enhanced machine is identical to the original.
-\sna{I'm thinking of SFI here\apt{yes}, but that's not stacks, right? Does Softbound work like
-  this.\apt{dunno}}
 
@@
 %
+\vspace{-3.5ex}
 \[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \unsaved + \saved\]
-
+%
 \judgmenttwobrlong[ Call]

### 2021-07-09 submission3/main.tex
More space

@@
 
-\subsection{Values and States}
+\paragraph*{Values and States}
 
@@
 
-\subsection{Contexts}
+\paragraph*{Contexts}
 
@@
 
-\subsection{Machine Traces}
+\paragraph*{Machine Traces}
 \label{sec:traces}
@@
 Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
-nested: we will quantify over subtraces that represent whole calls. For each such
-subtrace, we will quantify over {\em variants} of its initial state. Then we require
+nested: we will quantify over subtraces that represent whole calls. For each
+subtrace, we quantify over {\em variants} of its initial state. Then we require
 that the original, {\em primary} traces move in {\em lockstep} with the variant.
@@
 \end{minipage}
-
-%\leo{I still don't know how to do this transition. Maybe contrast
-%with the ``perhaps surprisingly'' later?} Note that
-%similarity of observation traces could alternatively be defined coinductively,
-%just like the prefix relation, but by dropping the asymmetric rules that allow
-%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}
 
@@
 
-Now we can inductively define the context update function; these rules again apply in the order
-presented. \ifspace\bcp{Too much space between the second and third.}\fi
+Now we can inductively define the context update function, as before.
 
@@
 \vspace*{-1ex}
-\judgmentthree[ ReturnRec]
+\judgmenttwo[ ReturnRec]
               {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}

### 2021-07-09 submission3/main.tex
Some savings in Section 9

@@
 StkTokens system enforces stack safety on hardware equipped with
-memory capabilities, similar to those used in the CHERI system~\citep{Woodruff+14,Chisnall+15}, but
+memory capabilities, similar to those used in the Cheri system~\citep{Woodruff+14,Chisnall+15}, but
 with additional support for linear capabilities.
@@
 micro-policies that tag stack objects with a pair of frame and object identifiers
-used to validate accesses to the stack: {\em Return Address Protection} (which
+used to validate accesses: {\em Return Address Protection} (which
 prevents an adversary from overwriting designated return addresses), {\em Static
 Authorities} (which only allows the code of a function to access the stack frames
-of its own dynamic instances), and {\em Depth Isolation} (described in
-\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
+of its own dynamic instances), and {\em Depth Isolation}
+(\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
 temporal locality of stack memory and local call graphs, as well as information
@@
 {\em Lazy
-Clearing}, discussed in \cref{sec:lazy}, with associated improvements in
-performance.
+Clearing} (\cref{sec:lazy}).
 
@@
   Clearing} against our properties, finding that {\em Depth Isolation} does
-indeed enforce stepwise stack safety and therefore also observable stack
+indeed enforce stepwise stack safety and therefore observable stack
 safety. We also found that neither lazy micro-policy actually enforces
@@
 to the accessible stack.
-
+%
 It is easy to imagine that such a rule could be proven sound in a system

### 2021-07-09 submission3/main.tex
Inlining StkTokens criteria in Section 9

@@
 
-\citet{SkorstengaardSTKJFP} propose a useful list of criteria for judging
-candidate stack safety properties: they should be (1) intuitive; (2)
-useful for reasoning; (3) reusable in secure compiler chains; (4)
-arguably ``complete'' with respect to informal notions of stack
-safety; and (5) potentially scalable to richer settings such as
-dynamic code generation or multithreading.
-%
-For criterion (1), intuitiveness, we contend that our property definitions are at least as
+\citet{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
+candidate stack safety properties.
+%% : they should be (1) intuitive; (2)
+%% useful for reasoning; (3) reusable in secure compiler chains; (4)
+%% arguably ``complete'' with respect to informal notions of stack
+%% safety; and (5) potentially scalable to richer settings such as
+%% dynamic code generation or multithreading.
+%
+(1) {\em Intuitiveness}. We contend that our property definitions are at least as
 intuitive as theirs, while being phrased in interestingly different terms.
@@
 %
-We agree that criterion (2), usefulness of the safety criterion
-for carrying out formal proofs (about individual programs), is important;
+(2) {\em Usefulness} of the safety criterion
+for carrying out formal proofs (about individual programs). We agree that this is an important consideration;
 demonstrating it for our proposal remains future work, in which we
@@
 %
+(3) {\em Reusability} in secure compilation chains.
 StkTokens
-is specifically set up to address criterion (3), reusability in secure
-compilation chains, while this is not a goal of
+is specifically set up to address this,
+while it is not a goal of
 our work.
 %
-With respect to criterion (4), capturing programmers' informal
-understanding of stack safety, we believe that our support
+(4) {\em Comprehensiveness} in capturing programmers' informal
+understanding of stack safety. We believe that our support
 for dynamically changing notion of adversaries, as outlined in point (i) above,
@@
 %
-As for criterion (5), broad applicability, we have demonstrated that our approach extends to
+(5) {\em Broad applicability} to richer settings. We have demonstrated that our approach extends to
 coroutines, a first step towards multithreading. Dynamic code generation is

### 2021-07-09 submission3/main.tex
Under limit!

@@
 that failstops when the hardware faults. Others might never even failstop, such as
-code rewriting techniques that turn stack safety violations into other (erroneous, but not
-stack-unsafe) behaviors. Then the policy-enhanced machine is identical to the original.
+code rewriting techniques that prevent stack safety violations directly, in which
+case their policy state does nothing at all.
 
@@
 
-We now define our machine and enforcement model precisely. To make our definition of stack safety
-as generic as possible, this section proposes an abstract interface to a
-machine model plus some kind of (software- or hardware-enforced) policy monitor.
-In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
-policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.
+We now define our machine and enforcement model precisely.
 
@@
 a positive and formal characterization of stack safety
-%begins with \citet{SkorstengaardLocal}, who introduced a calling
-%convention that uses local capabilities to preserve local state
-%encapsulation and well-bracketed control flow
-%using a logical relation to reason about the stack safety of concrete
-%programs.
-% Although the required hardware support is readily available
-%in capability machines like Cheri,
-%% this technique incurs significant
-%% costs, because
-%it requires the entire unused part of the stack to be
-%cleared whenever a security boundary is crossed.
-%Their logical
-%relation captures capability safety without ``externally observable
-%side-effects (like console output or memory access traces)''
-%and can be used to reason about individual programs.
 is that of \citet{SkorstengaardSTKJFP}. Their
@@
 the immediately adjacent frame below. Formally, these properties are defined
-with reference to the semantics of a
-capability machine with a built-in call stack and call and return
+with reference to a capability machine with a built-in call stack and call and return
 instructions; this machine is proven fully abstract with respect to a more
-concrete capability machine.  They call this approach a ``fully abstract overlay semantics.''
+concrete capability machine, an approach termed ``fully abstract overlay semantics.''
 The proof of full abstraction uses a standard notion of components, which
@@
   every} caller should be protected from its callees.  By contrast,
-  \citet{SkorstengaardSTKJFP} describe a class of ``unreasonable'' code components
-  that are not protected, because the are capable of dynamic behavior that can compromise
-  their own security. Unreasonable code can be viewed as adversarial, but it is equally
-  likely to simply be buggy. One can imagine a function that, when called with the wrong
+  \citet{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
+  that are not protected, because they are capable of dynamic behavior that can compromise
+  their own security. Unreasonable code can be viewed as adversarial, or merely as
+  buggy. One can imagine a function that, when called with the wrong
   inputs, behaves unreasonably, but under most circumstances can be
   considered reasonable. \citet{SkorstengaardSTKJFP} would not consider this a
-  {\em reasonable component} (as defined in their section 4.3.3), and so
-  would not guarantee its protection. In contrast, in our setting, individual
+  {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
+  component is always reasonable, and so they
+  would not guarantee its protection. In our setting, individual
   well-behaved activations of the buggy function would be protected.
-  \citet{SkorstengaardSTKJFP} note that they support mutually distrustful
-  components, but we go a step further. One can think of our hypothetical
+  \citet{SkorstengaardSTKJFP} support mutually distrustful
+  components; one can think of our hypothetical
   function as ``self-distrustful'' in that the safety of a well-behaved activation

### 2021-07-09 submission3/main.tex
Some nits in Section 7

@@
 
-To eliminate this cost, \citet{DBLP:conf/sp/RoesslerD18} augment Lazy Tagging
+To eliminate this cost, they augment Lazy Tagging
 with {\em Lazy Clearing}. Under this combination of optimizations,
@@
 For example, reading after a dangerous write may be acceptable if the value
-read does not affect the ultimate output of the program; this looser scheme
+read does not affect the output of the program; this looser scheme
 could be enforced by tainting the dangerous writes and detecting taint
@@
 
-The ``observations-of'' operator, written \(\obsof\), takes a machine trace and
+This operator, written \(\obsof\), takes a machine trace and
 gives the trace of observations of each step in
@@
 contain leaked information on either side of the first pair of variant traces.
-This becomes the basis for the second variation, because all of these values
+This becomes the basis for the second variation, as all of these values
 must be protected from leaking for the remainder of the program.

### 2021-07-09 submission3/main.tex
Table footers and descriptions

@@
 the process until we generate or execute some target number of
-instructions, or reach a point where the machine can't step
-anymore.
+instructions, or reach a point where the machine cannot step
+any more.
 
@@
 
-In addition, we need to extend this technique to handle the nested
+In addition, we need to further extend this technique to handle the nested
 nature of our confidentiality property: rather than just generating two
@@
 \vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+\caption{MTTF for finding bugs in erroneous eager policy enforcement mechanisms}
 \vspace*{-2em}
@@
 
-The mean-time-to-failure (MTTF) for various bugs can be found in
+The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
 Table~\ref{tab:bug-table}, along with the average number of tests
@@
 violation would become visible. Now we can identify when lazy policies
-miss such violations. The Lazy Tagging and Clearing
+miss such violations, which are again susceptible to testing;
+Table~\ref{tab:lazy-bug-table} summarizes the results of our evaluation.
+The Lazy Tagging and Clearing
 micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
@@
 \vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+\caption{MTTF for finding bugs in erroneous lazy policy enforcement mechanisms}
 \vspace*{-2em}

### 2021-09-04 submission3/response.txt
Added response file for submission3

@@
-# Overview
+Review #416A
+Overall merit
 
-We thank the reviewers for their comments.
+B. Weak Accept
+Reviewer expertise
 
-The reviews identify some areas that the paper fails to explain
-adequately -- in particular, "framing" questions: How does the work improve
-on existing formalizations of stack safety? Do we have a plan to validate
-the model further?  Can our work be extended to cover more complex uses of
-the stack?
+X. Expert
+Paper summary
 
-In the next section, we outline the changes we plan to make in the paper to
-clarify these questions.  We discuss the issues in more detail in the third
-section.
+This submission formalizes the general notion of stack safety (roughly speaking, every caller should be protected from its callees during program execution) using well-known notions of security (integrity, confidentiality and non interference), and defines several security policies for protecting the stack.
+The formalization is validated on existing (and less general) micro-policies for stack safety (depth isolation, lazy tagging and lazy clearing) using property-based random testing.
+This evaluation of former policies detected some security violations in former and less precise works.
 
-## Changes
+First, this submission introduces the main required security policies, in terms of traces of memory, register states and stepwise stack-safety: stepwise integrity (an execution step does not modify the memory slots that are reserved for a caller at a given depth), confidentiality (a callee is insensitive to the context of its call), stepwise confidentiality, temporal stack safety, and well-bracketed control-flow (calls and returns should be correctly nested).
+Then, the submission describes the requirements on the machine model (presence of a program counter, used to distinguish between the different stack-frames, and a mechanism for recognizing the return from a call) and the threat model.
+The machine model is abstract enough to represent different architectures.
 
-1. We will give a more thorough discussion of validation, discuss the
-   feasibility of proofs and of randomized testing of more complex
-   properties, and detail our roadmap for further validation of different
-   versions of the property. For details, see the comments responding
-   to reviewer A.
+Next, the submission details in a progressive way the formalization of machine behaviors and stack safety using semantic rules.
+The first formalization of the execution stack is simple, with a single stack and no sharing between callers and callees.
+Then the model is enriched with function parameters passed on the stack (by value or by reference), with callee-saves registers and with coroutines.
+Moreover, the first model is implemented in Coq and this implementation reuses an existing RISCV specification.
+The QuickChick tool is used to test the stepwise property and validate an existing enforcement mechanism (called depth isolation) of the literature.
+Furthermore, as the previous policy is conservative and rather slow to be practical, more permissive variants of confidentiality and integrity properties are formalized, resulting in an observational property that is again validated by testing.
+Strengths
 
-  *Status* - not incorporated
+    well-written submission, well-explained ideas
 
-2. We will give a more explicit and direct comparison to the work of
-   Skorstengaard et. al. Currently we discuss it briefly in our related work
-   section, but the novelty of our work does not come through clearly
-   enough; we will explain the differences more thoroughly, all in one place.
-   For details, see the comments responding to reviewer B.
+    general formalization that addresses a real challenge and that is validated by testing
 
-  *Status* - incorporated, but needs example/more discussion
+Weaknesses
 
-3. We will improve the presentation as suggested. In particular, we will
-   gradually introduce the examples and associated intuitions (including our
-   visual metaphors) as early as possible, starting in the introduction.
+    lack of discussion about the limitations of the work
 
-  *Status* - not incorporated
+    medium delta with respect to previous works
 
-4. We will more clearly call out common language features, such as the heap,
-   that are compatible with our treatment of stack safety. For details,
-   see the comments responding to reviewer B.
+Comments for author
 
-  *Status* incorporated
+This submission is well-written, with well-explained ideas.
+Examples of C and assembly programs are used throughout the submission to illustrate the various définitions.
+The formalization is provided as a supplementary material together with the test cases.
+The techniques and results of the submission are interesting and general enough to be reused in different settings.
 
-5. We will add a discussion of how to extend the model to a callee-saves
-   registers convention, which demonstrates a second concrete use for 
-   the observational properties. Previously we were concerned about space
-   and explanatory complexity, but in retrospect we see that discussing
-   register protection would paint a more complete picture.
-   For details, see the comments responding to reviewer A.
+I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.
 
-  *Status* incorporated
+Why did you choose this RISCV specification?
+Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?
+Would it be easy to adapt your work to another ISA specification?
+How realistic is the RISCV specification?
+In other words, are there any RISCV vulnerabilities that are not taken into account in your model?
+More generally, what are the limitations of the presented work?
+For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+Another example is the outside qualifier that characterizes only a whole component.
+Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?
 
-# Detailed Comments
+Could you comment on the Coq formalization that is provided as a supplementary material?
+Was it straightforward to write a Coq program from a C++ program?
+Moreover, I looked at the Coq development and found some admitted lemmas in it.
+Other comments
 
-## Reviewer A
+The supplementary material is not mentioned in the submission.
+Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').
 
-We will extend the paper with a more detailed discussion of our roadmap for
-verification and testing and the issues that we expect to arise. Here is a
-rough outline of how it would cover the concerns raised in the review.
+The first example programs (Fig.
+1 to 4) are C programs.
+The programs in Fig.
+5 and 6 are qualified as C++ programs, but they could be C programs and they are close to the first programs.
+Moreover, the C syntax for passing parameters by reference is surprising.
+I would have written f(&x)(line 934) and void f(int *a)(line 944).
 
-> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.
+    (83) when THAT an
+    (83) actiVation
+    (210) We CAN this form
+    (310, 886,895,947) The return instruction is confusing as the return type of fis void.
+    (357) when THE state
+    (389) are ARE
+    (568) two many closing parentheses (twice)
+    (626) we need REQUIRE
+    (626) followS
+    (652) calls. That
+    (796) I was expecting m' in (m,p',ctx').
+    (1005) lack of space between the last two rules
+    (1118) conjunctIon
 
-Correctness of an enforcement mechanism with respect to our properties
-is certainly amenable to formal proof. Such a proof would be
-considerable work to carry out for a full ISA like RISC-V, but we see
-no fundamental barriers.  We chose to concentrate on testing because
-it is better aligned with our immediate goal of evaluating real
-enforcement mechanisms for real machine architectures.
+Questions for the response period
 
-> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
->
-> * Any validation for the lazier observability definition (this is listed as future work.)^M
->
-> * Any validation for the coroutine work.
+See my former questions related to 1) the Coq formalization, 2) the choice of a RISCV model and 3) the limitations of the approach
 
-[ We have since tested lazy tagging/clearing and found it to pass the
-  newly implemented observational integrity property as well as normal,
-  stepwise confidentiality. But we did not have this ready before the
-  deadline.
-| We are working to implement testing of observational properties and to apply
-  this to the putatively fixed version of the lazy tagging/clearing
-  micro-policy. ]
+====================================================================================================================================
 
-Our other short-term plan is to implement testing of a second enforcement
-mechanism -- a Cheri implementation of stack safety -- to demonstrate that
-our properties are not strictly tied to tag-based enforcement. We are
-currently working on QuickChick generators for Cheri programs, which
-have a different calling convention (a more complicated one, which is easier
-to get wrong!). This effort is beyond the scope of this paper, but we will
-sketch it as future work.
+Review #416B
+Overall merit
 
-Validation of the argument-passing and coroutine models are longer-term
-projects.  The argument-passing model has existing enforcement mechanisms and serves
-an important role in common stack use cases. Coroutines will require the
-development of an enforcement mechanism (e.g., one based on tags) to
-validate.
+C. Weak Reject
+Reviewer expertise
 
-> Another minor question: The difference between [nested] call-by-reference
-> and passing address-of-local was interesting, but I'm not sure how much it
-> makes sense to argue the latter is "beyond the scope of stack safety per
-> se" [line 939] without arguing the same of the former.  Perhaps the
-> argument can be made that address-of-local can be put in the heap while a
-> by-reference parameter cannot and since you don't guarantee heap safety,
-> you shouldn't try to do anything useful with address-of-local?
+Y. Knowledgeable
+Paper summary
 
-Yes, this was exactly our thinking.  We know what constitutes legal access
-to stack-allocated arguments, but for explicit pointers this is much less
-clear, and depends on the intent of the programmer. We believe an
-address-of-local should be treated similarly to the address of a heap
-object, but with a different lifetime. For instance, if a heap safety model
-treats heap pointers like capabilities (possession implies access rights),
-we would treat stack pointers the same.  Similarly, if the heap model
-permits pointer forging (restricted to within compartments, say) we would
-expect to be able to forge pointers to address-taken locals in the same
-circumstances.
+This paper presents a semantics for formalising safety properties of stack safety enforcement systems.
+The paper uses a map to say which parts of the stack can currently be accessed and which can't, similar to the semantics of separation logic.
+It then connects this to a notion of call and return, that can be configured to enable different machine's implementations of functions.
 
-In the longer term we would like to extend our model with a heap, in
-the style of Azevedo de Amorim et al. [1] which is phrased in similar
-terms to our model and should be compatible. In this setting the
-interplay between heap-derived and stack-derived pointers (with
-similar behaviors but different lifespans) will be much more
-interesting.
+The paper uses a systematic testing approach with QuickChick to show that a hardware enforcement mechanism actually provides stack safety, and find bugs that are injected into the approach.
+It is great that you found the bug in lazy approach, however, it is fairly obvious type of security issue where uninitialised data leads to something escaping.
 
-## Reviewer B
+The paper discusses extension, where parameter passing can use the stack, and passing references to structures.
 
-> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et
-> al (L289-291)? What are the implications of that?
+The paper is primarily a semantics paper, though that seems fairly straightforward.
+The more interesting aspect of the paper is the more elaborate examples in Section 6 and 7, and the more elaborate extensions in 8.
+But I had not followed the earlier aspects well enough to appreciate those.
 
-Skorstengaard et al. don't directly assume well-bracketed control
-flow, they enforce it. But the terms in which they define it differ
-from ours. Their version assumes a code layout where each function has
-an isolated footprint and they don't permit unrestricted jumps between
-footprints.  Our definitions are sensible without that assumption.
+Overall, I didn't find the paper particularly compelling in its current form. It seems complex for what it achieves earlier on, and doesn't have clear enough exposition for the later parts.
+Strengths
 
-## Reviewer C
+    Checks properties of stack safety using capabilities
+    Interesting topic
 
-> - It is a bit difficult to assess the merits of this work. Yes, the
->   stack safety property is important for security, but what new results
->   or new capabilities does the proposed definition bring in?
+Weaknesses
 
-First, we contend that a formal definition of stack safety is valuable in
-its own right, as a way of shifting the discussion away from specific
-examples of bad behavior that must be prevented and toward general
-mechanisms that ensure good behavior.  A formal definition gives us the
-ability to reason clearly about what is gained from applying a given
-enforcement mechanism to a system. Since many enforcement mechanisms are
-quite costly in speed, dollars, and/or silicon, it is important to
-understand precisely what they buy us.
+    Exposition doesn't connect the different aspects of the paper well.
 
-More immediately, we can use a formal definition like this to test proposed
-enforecement mechanisms for violations of the policy.  In our case study, we
-identify a previously unknown bug in Roessler and Dehon's proposed "lazy
-tagging" micro-policy, and we use randomized testing to increase confidence
-that their "depth isolation" micro-policy correctly implements stack safety.
+    Seems overly complex in comparison to other relational reasoning papers.
 
->   The paper does not talk about data flow
->   across heap and stack, which is an important element for many
->   real-world memory corruption attacks.
+Comments for author
 
-Agreed!  We plan to combine our stack model with a heap model in the style
-of Azevedo de Amorim et al. [1] as future work.  Both models are phrased in the
-conceptual terminology of language-based security, so we have reason to
-believe that they will fit together cleanly.
+I found the paper did not connect the section well.
+Section 4 presents a lot of details for a framework, which is then used in Section 5.
+Seem very disconnected?
+For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section.
+How does it relate?
+Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them.
+Is this because they are different or just missing exposition.
 
-----------------
+Line 320: defition
 
-[1] A. Azevedo de Amorim, C. Hritcu, and B. Pierce, "The Meaning of Memory Safety,"
-Proc. Principles of Security and Trust, 2018.
+Line 329: I was really interested in address taken at this point, and whether the model could handle it.
+I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier
+
+Line 349: How does this related to control flow integrity/guard, and Intel's CET?
+
+Line 448: "If the predicate always holds" never holds?
+
+Line 510: "isRet" is a relation, but you are using as a function.
+
+Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?
+
+Line 642: wrok
+
+====================================================================================================================================
+
+Review #416C
+Overall merit
+
+B. Weak Accept
+Reviewer expertise
+
+Z. No familiarity
+Strengths
+
+The work presents a compelling model for stack security, framed in terms of existing notions of integrity and confidentiality.
+
+Models are sufficiently abstract and general to be transportable to a variety of mechanisms and machines.
+
+Notions of observable stack safety capture a more idealized notion that can reflect realistic violations, rather than "false positives" due to the limitations of particular mechanisms.
+This is my favorite contribution of this work.
+
+Some evidence that the approach can be tailored to complex control mechanisms, by showing how it applies to a simple fixed coroutine model.
+Weaknesses
+
+Validation is performed using random testing rather than formal proof.
+However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
+Comments for author
+
+line 13: variant of our properties and /does so/ correctly, and
+
+line 94: what about tail calls?
+
+line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.
+
+line 123: "we extend the generation ..." I'm not sure what that means.
+
+line 210: We /call/ this form...
+
+line 293: What about exceptions?
+
+line 563: all possible calls that /can/ be reached

### 2021-09-05 submission3/response.txt
Answers to some questions in response

@@
-Review #416A
-Overall merit
+> Review #416A
+> Overall merit
 
-B. Weak Accept
-Reviewer expertise
+> B. Weak Accept
+> Reviewer expertise
 
-X. Expert
-Paper summary
+> X. Expert
+> Paper summary
 
-This submission formalizes the general notion of stack safety (roughly speaking, every caller should be protected from its callees during program execution) using well-known notions of security (integrity, confidentiality and non interference), and defines several security policies for protecting the stack.
+> This submission formalizes the general notion of stack safety (roughly speaking, every caller should be protected from its callees during program execution) using well-known notions of security (integrity, confidentiality and non interference), and defines several security policies for protecting the stack.
 The formalization is validated on existing (and less general) micro-policies for stack safety (depth isolation, lazy tagging and lazy clearing) using property-based random testing.
@@
 
-First, this submission introduces the main required security policies, in terms of traces of memory, register states and stepwise stack-safety: stepwise integrity (an execution step does not modify the memory slots that are reserved for a caller at a given depth), confidentiality (a callee is insensitive to the context of its call), stepwise confidentiality, temporal stack safety, and well-bracketed control-flow (calls and returns should be correctly nested).
+> First, this submission introduces the main required security policies, in terms of traces of memory, register states and stepwise stack-safety: stepwise integrity (an execution step does not modify the memory slots that are reserved for a caller at a given depth), confidentiality (a callee is insensitive to the context of its call), stepwise confidentiality, temporal stack safety, and well-bracketed control-flow (calls and returns should be correctly nested).
 Then, the submission describes the requirements on the machine model (presence of a program counter, used to distinguish between the different stack-frames, and a mechanism for recognizing the return from a call) and the threat model.
@@
 
-Next, the submission details in a progressive way the formalization of machine behaviors and stack safety using semantic rules.
+> Next, the submission details in a progressive way the formalization of machine behaviors and stack safety using semantic rules.
 The first formalization of the execution stack is simple, with a single stack and no sharing between callers and callees.
@@
 Furthermore, as the previous policy is conservative and rather slow to be practical, more permissive variants of confidentiality and integrity properties are formalized, resulting in an observational property that is again validated by testing.
-Strengths
+> Strengths
 
-    well-written submission, well-explained ideas
+>    well-written submission, well-explained ideas
 
-    general formalization that addresses a real challenge and that is validated by testing
+>    general formalization that addresses a real challenge and that is validated by testing
 
-Weaknesses
+> Weaknesses
 
-    lack of discussion about the limitations of the work
+>    lack of discussion about the limitations of the work
 
-    medium delta with respect to previous works
+>    medium delta with respect to previous works
 
-Comments for author
+> Comments for author
 
-This submission is well-written, with well-explained ideas.
+> This submission is well-written, with well-explained ideas.
 Examples of C and assembly programs are used throughout the submission to illustrate the various définitions.
@@
 
-I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.
+> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.
 
-Why did you choose this RISCV specification?
-Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?
-Would it be easy to adapt your work to another ISA specification?
-How realistic is the RISCV specification?
-In other words, are there any RISCV vulnerabilities that are not taken into account in your model?
-More generally, what are the limitations of the presented work?
-For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
-Another example is the outside qualifier that characterizes only a whole component.
-Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?
+*** Questions Begin ***
 
-Could you comment on the Coq formalization that is provided as a supplementary material?
+> Why did you choose this RISCV specification?
+> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?
+
+Because the enforcement mechanisms we test are existing micropolicies defined in terms of a reference monitor attached to a RISCV machine.
+Note that the formalism itself is not dependent on the details of the architecture, only the testing framework to test that architecture is.
+There is little in the tested micropolicies that depends on RISCV, either, but moving them to another architecture such as ARM would nevertheless
+require translating the micropolicies rules to apply to that architecture.
+
+> Would it be easy to adapt your work to another ISA specification?
+
+Fairly easy. The formalism itself translates naturally to most ISA specifications. The largest amount of work
+is in the testing framework - modifying the code generation system to match the new ISA.
+We are currently working on extending the work to a model of a Cheri-style hardware capability machine, Cerise,
+and have not encountered major obstacles.
+
+> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?
+
+Timing channels; anything else?
+
+> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+
+> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?
+
+I don't understand this question, any insight would be helpful.
+
+> Could you comment on the Coq formalization that is provided as a supplementary material?
 Was it straightforward to write a Coq program from a C++ program?
 Moreover, I looked at the Coq development and found some admitted lemmas in it.
-Other comments
 
-The supplementary material is not mentioned in the submission.
+Writing Coq program from C++: I believe this one is for Rob
+Admitted Lemmas:
+
+> Other comments
+
+> The supplementary material is not mentioned in the submission.
 Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').
 
-The first example programs (Fig.
+The supplementary material should indeed be mentioned in the submission.
+Explanation of definition 7.4:
+
+> The first example programs (Fig.
 1 to 4) are C programs.
@@
 
-    (83) when THAT an
-    (83) actiVation
-    (210) We CAN this form
-    (310, 886,895,947) The return instruction is confusing as the return type of fis void.
-    (357) when THE state
-    (389) are ARE
-    (568) two many closing parentheses (twice)
-    (626) we need REQUIRE
-    (626) followS
-    (652) calls. That
-    (796) I was expecting m' in (m,p',ctx').
-    (1005) lack of space between the last two rules
-    (1118) conjunctIon
+>    (83) when THAT an
+>    (83) actiVation
+>    (210) We CAN this form
+>    (310, 886,895,947) The return instruction is confusing as the return type of fis void.
+>    (357) when THE state
+>    (389) are ARE
+>    (568) two many closing parentheses (twice)
+>    (626) we need REQUIRE
+>    (626) followS
+>    (652) calls. That
+>    (796) I was expecting m' in (m,p',ctx').
+>    (1005) lack of space between the last two rules
+>    (1118) conjunctIon
 
-Questions for the response period
+> Questions for the response period
 
-See my former questions related to 1) the Coq formalization, 2) the choice of a RISCV model and 3) the limitations of the approach
+> See my former questions related to 1) the Coq formalization, 2) the choice of a RISCV model and 3) the limitations of the approach
 
@@
 
-Review #416B
-Overall merit
+> Review #416B
+> Overall merit
 
-C. Weak Reject
-Reviewer expertise
+> C. Weak Reject
+> Reviewer expertise
 
-Y. Knowledgeable
-Paper summary
+> Y. Knowledgeable
+> Paper summary
 
-This paper presents a semantics for formalising safety properties of stack safety enforcement systems.
+> This paper presents a semantics for formalising safety properties of stack safety enforcement systems.
 The paper uses a map to say which parts of the stack can currently be accessed and which can't, similar to the semantics of separation logic.
@@
 
-The paper uses a systematic testing approach with QuickChick to show that a hardware enforcement mechanism actually provides stack safety, and find bugs that are injected into the approach.
+> The paper uses a systematic testing approach with QuickChick to show that a hardware enforcement mechanism actually provides stack safety, and find bugs that are injected into the approach.
 It is great that you found the bug in lazy approach, however, it is fairly obvious type of security issue where uninitialised data leads to something escaping.
 
-The paper discusses extension, where parameter passing can use the stack, and passing references to structures.
+> The paper discusses extension, where parameter passing can use the stack, and passing references to structures.
 
-The paper is primarily a semantics paper, though that seems fairly straightforward.
+> The paper is primarily a semantics paper, though that seems fairly straightforward.
 The more interesting aspect of the paper is the more elaborate examples in Section 6 and 7, and the more elaborate extensions in 8.
@@
 
-Overall, I didn't find the paper particularly compelling in its current form. It seems complex for what it achieves earlier on, and doesn't have clear enough exposition for the later parts.
-Strengths
+> Overall, I didn't find the paper particularly compelling in its current form. It seems complex for what it achieves earlier on, and doesn't have clear enough exposition for the later parts.
+> Strengths
 
-    Checks properties of stack safety using capabilities
-    Interesting topic
+>    Checks properties of stack safety using capabilities
+>    Interesting topic
 
-Weaknesses
+> Weaknesses
 
-    Exposition doesn't connect the different aspects of the paper well.
+>    Exposition doesn't connect the different aspects of the paper well.
 
-    Seems overly complex in comparison to other relational reasoning papers.
+>    Seems overly complex in comparison to other relational reasoning papers.
 
-Comments for author
+> Comments for author
 
-I found the paper did not connect the section well.
+> I found the paper did not connect the section well.
 Section 4 presents a lot of details for a framework, which is then used in Section 5.
@@
 
-Line 320: defition
+We should think if there is anything we can suggest here to improve organization, but it's unclear to me just what the problem was
 
-Line 329: I was really interested in address taken at this point, and whether the model could handle it.
+> Line 320: defition
+
+> Line 329: I was really interested in address taken at this point, and whether the model could handle it.
 I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier
 
-Line 349: How does this related to control flow integrity/guard, and Intel's CET?
+> Line 349: How does this related to control flow integrity/guard, and Intel's CET?
 
-Line 448: "If the predicate always holds" never holds?
+> Line 448: "If the predicate always holds" never holds?
 
-Line 510: "isRet" is a relation, but you are using as a function.
+> Line 510: "isRet" is a relation, but you are using as a function.
 
-Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?
+> Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?
 
-Line 642: wrok
+> Line 642: wrok
 
@@
 
-Review #416C
-Overall merit
+> Review #416C
+> Overall merit
 
-B. Weak Accept
-Reviewer expertise
+> B. Weak Accept
+> Reviewer expertise
 
-Z. No familiarity
-Strengths
+> Z. No familiarity
+> Strengths
 
-The work presents a compelling model for stack security, framed in terms of existing notions of integrity and confidentiality.
+> The work presents a compelling model for stack security, framed in terms of existing notions of integrity and confidentiality.
 
-Models are sufficiently abstract and general to be transportable to a variety of mechanisms and machines.
+> Models are sufficiently abstract and general to be transportable to a variety of mechanisms and machines.
 
-Notions of observable stack safety capture a more idealized notion that can reflect realistic violations, rather than "false positives" due to the limitations of particular mechanisms.
+> Notions of observable stack safety capture a more idealized notion that can reflect realistic violations, rather than "false positives" due to the limitations of particular mechanisms.
 This is my favorite contribution of this work.
 
-Some evidence that the approach can be tailored to complex control mechanisms, by showing how it applies to a simple fixed coroutine model.
-Weaknesses
+> Some evidence that the approach can be tailored to complex control mechanisms, by showing how it applies to a simple fixed coroutine model.
+> Weaknesses
 
-Validation is performed using random testing rather than formal proof.
+> Validation is performed using random testing rather than formal proof.
 However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
-Comments for author
+> Comments for author
 
-line 13: variant of our properties and /does so/ correctly, and
+> line 13: variant of our properties and /does so/ correctly, and
 
-line 94: what about tail calls?
+> line 94: what about tail calls?
 
-line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.
+Copy tail-call discussion from previous response
 
-line 123: "we extend the generation ..." I'm not sure what that means.
+> line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.
 
-line 210: We /call/ this form...
+> line 123: "we extend the generation ..." I'm not sure what that means.
 
-line 293: What about exceptions?
+> line 210: We /call/ this form...
 
-line 563: all possible calls that /can/ be reached
+> line 293: What about exceptions?
+
+Copy exception discussion from previous response
+
+> line 563: all possible calls that /can/ be reached

### 2021-09-06 submission3/response.txt
Added discussion of exceptions and tail-calls

@@
 
-Copy tail-call discussion from previous response
+Tail calls are a plausible extension of our framework, but not supported as written.
+Since a tail call is both a call and a return in a single step, our normal handling of
+returns will not work -- it would interpret the tail call as a call to and immediate return from
+a new function activation. We would need to distinguish tail calls from normal calls and essentially
+process the call and the return in the reverse order.
 
@@
 
-Copy exception discussion from previous response
+Exceptions are another plausible extension. They would involve interleaving the return
+target stack with entries that are not return targets, but "catch targets." Reaching a catch target
+would pop all return targets above it, much like returns that violate well-bracketedness already do.
 

### 2021-09-06 submission3/response.txt
Polishing and further comments

--- 2021-09-05 submission3/response.txt
@@
 
-Because the enforcement mechanisms we test are existing micropolicies defined in terms of a reference monitor attached to a RISCV machine.
-Note that the formalism itself is not dependent on the details of the architecture, only the testing framework to test that architecture is.
-There is little in the tested micropolicies that depends on RISCV, either, but moving them to another architecture such as ARM would nevertheless
-require translating the micropolicies rules to apply to that architecture.
+We chose RISCV for the sake of the evaluation: the enforcement mechanisms we
+test are existing micropolicies defined in terms of a reference monitor
+attached to a RISCV machine.
+
+The underlying formalism itself is not dependent on the details of the
+architecture, only the testing framework to test that architecture is.
+Indeed, even wrt. testing, there is little in the tested micropolicies that
+depends on RISCV either, but moving to another architecture such as ARM
+would require translating the micropolicy rules to apply to that
+architecture and its calling conventions.
 
@@
 
-Fairly easy. The formalism itself translates naturally to most ISA specifications. The largest amount of work
-is in the testing framework - modifying the code generation system to match the new ISA.
-We are currently working on extending the work to a model of a Cheri-style hardware capability machine, Cerise,
-and have not encountered major obstacles.
+Fairly easy. The formalism itself translates naturally to most ISA
+specifications. The largest amount of work is in the testing framework -
+modifying the example generator to match the new ISA.
+
+We are currently working on extending the work to a model of a Cheri-style
+hardware capability machine, Cerise, and have not encountered major
+obstacles.
 
@@
 
-Timing channels; anything else?
+[Timing channels; anything else?]  [BCP: Even if the RISCV model is fairly
+complete (though it doesn't include stuff like VM management, hypervisor
+instructions, ..., right?), we don't generate test cases for most
+instructions -- we're really just testing the protection mechanism for
+function calls.]
 
@@
 
+[BCP: ???]
+
 > Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?
 
-I don't understand this question, any insight would be helpful.
+[I don't understand this question, any insight would be helpful.]
+[BCP: Me either.  I wonder if we can ask for clarification...]
 
@@
 
+[BCP: I wonder whether they are asking about the compilation from our C/C++
+pseudo-source-code to RISCV...?]
+
 > Other comments
@@
 The supplementary material should indeed be mentioned in the submission.
-Explanation of definition 7.4:
+[Explanation of definition 7.4:]
 
@@
 I would have written f(&x)(line 934) and void f(int *a)(line 944).
+
+[BCP: We should check our C / C++ syntax.]
 
@@
 
-We should think if there is anything we can suggest here to improve organization, but it's unclear to me just what the problem was
+[We should think if there is anything we can suggest here to improve
+organization, but it's unclear to me just what the problem was
+
+[BCP: The review here gives one specific example -- what can we say about it?]
 
@@
 
+[BCP: We should just fix these sorts of things and then not mention them in
+the response.]
+
 > Line 329: I was really interested in address taken at this point, and whether the model could handle it.
 I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier
+
+[BCP: This is a good expositional idea.]
 
@@
 
-Copy tail-call discussion from previous response
+We believe tail calls are a straightforward extension of our framework,
+though they are not supported at present.  Since a tail call is both a call
+and a return in a single step, our normal handling of returns doesn't
+work -- it would interpret the tail call as a call to and immediate return
+from a new function activation. We would need to distinguish tail calls from
+normal calls and essentially process the call and the return in the reverse
+order.
 
@@
 
-Copy exception discussion from previous response
+Exceptions are another plausible extension. Adding them would involve
+interleaving the return target stack with entries that are not return
+targets, but "catch targets." Reaching a catch target would pop all return
+targets above it, much like returns that violate well-bracketedness already
+do.
 

### 2021-09-06 submission3/response.txt
comments

@@
 attached to a RISCV machine.
+[APT: Is this really true? Nick and Andre's paper uses Alpha.
+More accurate to say that we (and Draper, where much of the micropolicies work as been done)
+chose RISCV for its simplicity.]
 
@@
 architecture, only the testing framework to test that architecture is.
+[APT: Is it relevant to point out that we have generic Coq interfaces to
+describe the machine and policy?]
 Indeed, even wrt. testing, there is little in the tested micropolicies that
@@
 Admitted Lemmas:
+[APT: The admits don't seem to be important...]
 
@@
 pseudo-source-code to RISCV...?]
+[APT: ...perhaps under the misconception that we are showing RISCV code in
+the examples?]
 
@@
 [Explanation of definition 7.4:]
+[APT: Can we include a list of important links here in the response?]
 
@@
 [BCP: We should check our C / C++ syntax.]
+[APT: I think the reviewer is confused. C doesn't really have call-by-reference: (explicitly) passing a pointer
+is _not_ the same thing as (implicitly) passing a reference, as the latter cannot be used in a first-class way 
+in the callee: that's the whole distinction we're trying to make starting at line 954 of the text.
+Evidently we need to clarify this.]
 
@@
 [BCP: This is a good expositional idea.]
+[APT: Right, so we should acknowledge it politely -- but we did have reasons for ordering
+things the way we did....]
 

### 2021-09-06 submission3/response.txt
one more comment
Tiny

@@
 chose RISCV for its simplicity.]
+[BCP: es.]
 
@@
 [BCP: Me either.  I wonder if we can ask for clarification...]
+[APT: Seems like reviewer is thinking of components as having internal structure and
+wondering if we need to label different parts of the structure in different ways...]
 

### 2021-09-07 submission3/response.txt
Progress on response.txt

--- 2021-09-06 submission3/response.txt
@@
 
-We chose RISCV for the sake of the evaluation: the enforcement mechanisms we
-test are existing micropolicies defined in terms of a reference monitor
-attached to a RISCV machine.
-[APT: Is this really true? Nick and Andre's paper uses Alpha.
-More accurate to say that we (and Draper, where much of the micropolicies work as been done)
-chose RISCV for its simplicity.]
+We chose RISCV for the simplicity. It is a prominent base ISA for
+the PIPE extensions that implement the tag monitor for the same reason.
 
 The underlying formalism itself is not dependent on the details of the
-architecture, only the testing framework to test that architecture is.
-[APT: Is it relevant to point out that we have generic Coq interfaces to
-describe the machine and policy?]
-Indeed, even wrt. testing, there is little in the tested micropolicies that
+architecture, as in fact the properties are defined in terms of a generic
+interface over machine and micropolicy.
+
+Even wrt. testing, there is little in the tested micropolicies that
 depends on RISCV either, but moving to another architecture such as ARM
@@
 
-[Timing channels; anything else?]  [BCP: Even if the RISCV model is fairly
-complete (though it doesn't include stuff like VM management, hypervisor
-instructions, ..., right?), we don't generate test cases for most
-instructions -- we're really just testing the protection mechanism for
-function calls.]
+We use the RISCV specification to demonstrate that our properties are testable in
+a realistic setting. We do not exercize the entire RISCV specification, and the validity
+of the model is not dependent on the absence of vulnerabilities in the RISCV implementation.
+That said, the RISCV specification we used is fairly realistic, although lacking some features
+such as VM management instructions.
 
@@
 
-[BCP: ???]
+Currently, some discussions of our model's limitations is scattered throughout the paper.
+We will expand and organize it into an explicit discussion, as follows.
+Our goal is to specifically characterize stack safety.
+So we are particularly concerned with limitations that prevent the model from applying to realistic
+uses of the stack, and not with other structures such as the heap, which we intentionally make no
+guarantees about.
+
+Features that we do not currently support:
+- Dynamic code generation: this would be rather challenging to integrate into
+  our framework.
+- Address-taken locals: we have discussed why we don't consider our model suitable
+  for handling.
+- Exceptions, though we can see a way to extend the model to include them:
+  essentially, we would model a "setjump" operation by inserting a special
+  target in the return target stack. We would weaken well-bracketedness to
+  allow reach such a target from any depth. Since integrity and confidentiality
+  are distinct from well-bracketedness and already permit the popping of multiple
+  functions from the stack, they can simply treat the exception's target as another
+  return target.
+- Tail calls, but again we can see how we might extend the model for them:
+  though they are not supported at present.  Since a tail call is both a call
+  and a return in a single step, our normal handling of returns doesn't
+  work -- it would interpret the tail call as a call to and immediate return
+  from a new function activation. We would need to distinguish tail calls from
+  normal calls and essentially process the call and the return in the reverse
+  order.
+
+As for the distinction between values and addresses, conventional ISAs do not make this distinction.
+Ongoing work on the Cerise capability machine formalization has required some modification to distinguish
+values from addresses, as they are not the same in a capability machine, but it has little effect on the 
+model itself. Further, distinguishing between values and addresses is a higher-level abstraction that can
+"hide" potential bugs resulting from implicit conversions between them, so it may even be preferable to model
+them as the same when possible.
 
@@
 
-[I don't understand this question, any insight would be helpful.]
-[BCP: Me either.  I wonder if we can ask for clarification...]
+There appears to be some confusion arising from the term "components."
+In our model, components are already fine-grained: individual registers
+and memory addresses. We recognize that this is a poor choice of term
+given the common usage of component to refer to a larger module in a
+compartmentalization context, and will change the term to [SNA: still
+looking for a word. Atom?]
+Since components/atoms are so fine-grained, there is no need for finer
+access control.
 
@@
 
-Writing Coq program from C++: I believe this one is for Rob
-Admitted Lemmas:
-[APT: The admits don't seem to be important...]
-
-[BCP: I wonder whether they are asking about the compilation from our C/C++
-pseudo-source-code to RISCV...?]
-[APT: ...perhaps under the misconception that we are showing RISCV code in
-the examples?]
+See below.
 
@@
 The supplementary material should indeed be mentioned in the submission.
-[Explanation of definition 7.4:]
+The Coq formalization provided contains two parts: the formalization of our properties,
+which was a conceptual aid in their development and may be useful to readers' intuitions,
+and the testing framework, which is used to validate our properties by showing that
+computable versions of them are testable.
+
+Admitted lemmas are the beginnings of an effort to formalize [details]
+
+In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
+last instance of n' should be n''.
+
 [APT: Can we include a list of important links here in the response?]
@@
 
-[We should think if there is anything we can suggest here to improve
-organization, but it's unclear to me just what the problem was
+[We should think if there is anything we can suggest here to improve flow]
 
@@
 
-We believe tail calls are a straightforward extension of our framework,
-though they are not supported at present.  Since a tail call is both a call
-and a return in a single step, our normal handling of returns doesn't
-work -- it would interpret the tail call as a call to and immediate return
-from a new function activation. We would need to distinguish tail calls from
-normal calls and essentially process the call and the return in the reverse
-order.
-
 > line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.
@@
 
-Exceptions are another plausible extension. Adding them would involve
-interleaving the return target stack with entries that are not return
-targets, but "catch targets." Reaching a catch target would pop all return
-targets above it, much like returns that violate well-bracketedness already
-do.
-
 > line 563: all possible calls that /can/ be reached

### 2021-09-07 submission3/response.txt
More response.txt

@@
 
-[We should think if there is anything we can suggest here to improve flow]
+Many of the content that seems disconnected takes the form of a parameterizable relation and its various
+instatiations in different versions of the property. \rightarrow_C parameterized by the type ctx, which is section 5
+is instantiated as (K -> SD)x(list(target)). In later sections it will be instantiated by other types.
+Then for any \rightarrow_C, we can extend \rightharpoonup to form \rightharpoonup_C.
 
-[BCP: The review here gives one specific example -- what can we say about it?]
+We can do more to call out which parts of the formalization are parameterized and when they are being
+instantiated, so that the connections are more clear.
 
@@
 
+This is a good point, and we will examine ways to introduce that information sooner. However, we are
+hesitant to spend too much time on variants before introducing the basic formalism, because it can
+be a lot to keep track of.
+
 [BCP: This is a good expositional idea.]
@@
 things the way we did....]
+[SNA: That was roughly the reason, as I recall? I may take a stab tomorrow at sketching what
+moving them up would look like.]
 

### 2021-09-07 submission3/response.txt
Added coq map

@@
 
-[APT: Can we include a list of important links here in the response?]
+Map of the supplementary material:
+Core contains the abstract machine and policy models, as well as the RISCV instantiations with different policies.
+Properties contains Coq definitions corresponding to the different Properties:
+- Trace.v and ObsTrace.v define the trace model and observation model
+- TraceProperties.v defines arbitrary stack properties given a call trace and context type
+- SubroutineSimple.v defines the basic stack safety properties
+- SubroutineShare.v defines the version enhanced with argument passing on the stack
+- Coroutine.v defines the coroutine version
+- Lazy.v defines the lazy properties
+- CalleeSave.v defines stack safety with callee-save registers
+Testing contains the test system, with generators for different machines and the final test run
+via extraction in TestProperties.v
 

### 2021-09-07 submission3/response.txt
some polishing

@@
 > Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?
-
-We chose RISCV for the simplicity. It is a prominent base ISA for
-the PIPE extensions that implement the tag monitor for the same reason.
-
-The underlying formalism itself is not dependent on the details of the
-architecture, as in fact the properties are defined in terms of a generic
-interface over machine and micropolicy.
-
-Even wrt. testing, there is little in the tested micropolicies that
-depends on RISCV either, but moving to another architecture such as ARM
-would require translating the micropolicy rules to apply to that
-architecture and its calling conventions.
-
 > Would it be easy to adapt your work to another ISA specification?
 
-Fairly easy. The formalism itself translates naturally to most ISA
-specifications. The largest amount of work is in the testing framework -
-modifying the example generator to match the new ISA.
+We chose RISCV for its simplicity, and because it has been used as the
+base ISA for much of the existing work (by others) to design
+hardware that supports tag-based micro-policies.
+
+To clarify a possible misunderstanding: the Coq specification of RISCV
+that we use was developed by others (see footnote p. 14) and we were
+able to adapt it to our testing framework with relatively little
+effort.
+
+Our stack safety theory itself is not dependent on the details of the
+architecture, as in fact the properties are defined in terms of a
+generic interface over machine and enforcement policy; [see ... in the
+Coq development]
+
+Performing testing on another architecture such as ARM would require
+modifying the testing framework by example generator for the new ISA,
+and translating the micropolicy rules to apply to that architecture
+and its calling conventions.  This would take some work, but should be
+straightforward.
 
@@
 We use the RISCV specification to demonstrate that our properties are testable in
-a realistic setting. We do not exercize the entire RISCV specification, and the validity
+a realistic setting. We do not exercise the entire RISCV specification, and the validity
 of the model is not dependent on the absence of vulnerabilities in the RISCV implementation.
-That said, the RISCV specification we used is fairly realistic, although lacking some features
-such as VM management instructions.
+That said, the RISCV specification we use is complete for user mode integer instructions
+(IV32IM).
 

### 2021-09-07 submission3/response.txt
more polishing

@@
 
-Currently, some discussions of our model's limitations is scattered throughout the paper.
-We will expand and organize it into an explicit discussion, as follows.
-Our goal is to specifically characterize stack safety.
-So we are particularly concerned with limitations that prevent the model from applying to realistic
-uses of the stack, and not with other structures such as the heap, which we intentionally make no
-guarantees about.
+Currently, discussion of our model's limitations is scattered
+throughout the paper.  We will expand and organize it into an explicit
+section, as follows.
+
+Our goal is to specifically characterize stack safety.  So we are
+particularly concerned with limitations that prevent the model from
+applying to realistic uses of the stack, and not with other structures
+such as the heap, which we intentionally make no guarantees about.
 
 Features that we do not currently support:
-- Dynamic code generation: this would be rather challenging to integrate into
-  our framework.
-- Address-taken locals: we have discussed why we don't consider our model suitable
-  for handling.
-- Exceptions, though we can see a way to extend the model to include them:
-  essentially, we would model a "setjump" operation by inserting a special
-  target in the return target stack. We would weaken well-bracketedness to
-  allow reach such a target from any depth. Since integrity and confidentiality
-  are distinct from well-bracketedness and already permit the popping of multiple
-  functions from the stack, they can simply treat the exception's target as another
-  return target.
-- Tail calls, but again we can see how we might extend the model for them:
-  though they are not supported at present.  Since a tail call is both a call
-  and a return in a single step, our normal handling of returns doesn't
-  work -- it would interpret the tail call as a call to and immediate return
-  from a new function activation. We would need to distinguish tail calls from
-  normal calls and essentially process the call and the return in the reverse
+
+- Address-taken locals. As we discuss at line 954ff, a satisfactory
+  treatment of these seems to require integration with a more general
+  model of memory safety that covers the heap and static globals too.
+
+- Exceptions. We can see a way to extend the model to include them:
+  essentially, we would model a "setjump" operation by inserting a
+  special target in the return target stack. We would weaken
+  well-bracketedness to allow reaching such a target from any
+  depth. Since integrity and confidentiality are distinct from
+  well-bracketedness and already permit the popping of multiple
+  functions from the stack, they can simply treat the exception's
+  target as another return target.
+
+- Tail calls. Again, we can see how we might extend the model for them
+  Since a tail call is both a call and a return in a single step, our
+  normal handling of returns doesn't work -- it would interpret the
+  tail call as a call to and immediate return from a new function
+  activation. We would need to distinguish tail calls from normal
+  calls and essentially process the call and the return in the reverse
   order.
 
-As for the distinction between values and addresses, conventional ISAs do not make this distinction.
-Ongoing work on the Cerise capability machine formalization has required some modification to distinguish
-values from addresses, as they are not the same in a capability machine, but it has little effect on the 
-model itself. Further, distinguishing between values and addresses is a higher-level abstraction that can
-"hide" potential bugs resulting from implicit conversions between them, so it may even be preferable to model
-them as the same when possible.
+- Dynamic code generation. This would be rather challenging to
+  integrate into our framework, because we rely on annotations in the
+  code.
+[APT: Right? Can we imagine how this might be supported in future?]
+
+- Concurrency. We can handle a limited form of coroutines, as
+  described in 8.3. But handling true concurrency would require
+  significant changes in our underlying machine model.
+
+As for the distinction between values and addresses, conventional ISAs
+do not make this distinction.  Ongoing work on the Cerise capability
+machine formalization has required some modification to distinguish
+values from addresses, as they are not the same in a capability
+machine, but it has little effect on the model itself. Further,
+distinguishing between values and addresses is a higher-level
+abstraction that can "hide" potential bugs resulting from implicit
+conversions between them, so it may even be preferable to model them
+as the same when possible.
 
@@
 
-There appears to be some confusion arising from the term "components."
-In our model, components are already fine-grained: individual registers
-and memory addresses. We recognize that this is a poor choice of term
-given the common usage of component to refer to a larger module in a
-compartmentalization context, and will change the term to [SNA: still
-looking for a word. Atom?]
-Since components/atoms are so fine-grained, there is no need for finer
-access control.
+There may be some confusion arising from our use of the term
+"components."  In our model (see line 398), components are already
+fine-grained: individual registers and memory addresses. We recognize
+that this is a poor choice of term given the common usage of component
+to refer to a larger module in a compartmentalization context, and
+will change the term to [SNA: still looking for a word. Atom?
+APT: "state element" ?]  Since components/atoms are so fine-grained, there
+is no need for finer access control.
 

### 2021-09-07 submission3/response.txt
done with pass

@@
 via extraction in TestProperties.v
+[APT: Can we give pointers from each paper definition to corresponding Coq definition?]
 
@@
 
-[BCP: We should check our C / C++ syntax.]
-[APT: I think the reviewer is confused. C doesn't really have call-by-reference: (explicitly) passing a pointer
-is _not_ the same thing as (implicitly) passing a reference, as the latter cannot be used in a first-class way
-in the callee: that's the whole distinction we're trying to make starting at line 954 of the text.
-Evidently we need to clarify this.]
+Only Fig. 6 uses C++ features.  Although C programmers might indeed
+try to simulate call-by-reference by passing a pointer as suggested
+here, this is subtly different from a genuine by-reference argument in
+C++, because the latter cannot be directly used in a first-class way
+in the callee (without using the & operator, as e.g. at line 945).
+Thus, legal access to the value of a by-reference argument is
+restricted to the immediate callee (and any chain of sub-callees to
+which the argument is handed down by reference), and we can capture
+this access pattern in our policies.  On the other hand, pointer
+arguments might escape into arbitrary sub-callees, and there is no
+obvious way for policies in our style to specify when access through
+these pointers should be permitted. We will improve the discussion at
+line 954ff to clarify this point.
 
@@
 Many of the content that seems disconnected takes the form of a parameterizable relation and its various
-instatiations in different versions of the property. \rightarrow_C parameterized by the type ctx, which is section 5
+instantiations in different versions of the property. \rightarrow_C parameterized by the type ctx, which is section 5
 is instantiated as (K -> SD)x(list(target)). In later sections it will be instantiated by other types.
@@
 
-We can do more to call out which parts of the formalization are parameterized and when they are being
+We will do more to call out which parts of the formalization are parameterized and when they are being
 instantiated, so that the connections are more clear.
@@
 
+[APT: Add pointer]
+
 > line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.
@@
 > line 123: "we extend the generation ..." I'm not sure what that means.
+
+[APT: Add a clarification that is is "generation-by-execution" ?]
 
@@
 
+[APT: Add pointer]
+
 > line 563: all possible calls that /can/ be reached

### 2021-09-07 submission3/response.txt
Answered some edits

@@
   integrate into our framework, because we rely on annotations in the
-  code.
-[APT: Right? Can we imagine how this might be supported in future?]
+  code. We can easily envision supporting dynamic code generation
+  where the generated code is guaranteed never to contain a call
+  or other annotated instruction. We could also support annotated-code
+  generation in which existing code is copied along with its annotations.
 
@@
 to refer to a larger module in a compartmentalization context, and
-will change the term to [SNA: still looking for a word. Atom?
-APT: "state element" ?]  Since components/atoms are so fine-grained, there
-is no need for finer access control.
+will change the term to "state elements." Since components/elements
+are so fine-grained, there is no need for finer access control.
 

### 2021-09-08 submission3/response.txt
More responses

@@
 
-[APT: Add pointer]
+Tail calls are discussed above in our proposed "limitations" section.
+They are not currently supported, but we describe how they might be
+implemented.
 
@@
 
-[APT: Add a clarification that is is "generation-by-execution" ?]
+Line 123 should read "... we extend the generation-by-execution techniques...",
+referring to the method of randomly generating valid low-level programs by lazily adding
+instructions while simulating execution.
 
@@
 
-[APT: Add pointer]
+Exceptions are also discussed about in our "limitations" section,
+along with a description of what it would take to implement them.
 

### 2021-09-08 submission3/response.txt
Fixed a bunch of minor things

@@
 
->    (83) when THAT an
->    (83) actiVation
->    (210) We CAN this form
->    (310, 886,895,947) The return instruction is confusing as the return type of fis void.
->    (357) when THE state
->    (389) are ARE
+>    (310, 886,895,947) The return instruction is confusing as the return type of f is void.
 >    (568) two many closing parentheses (twice)
->    (626) we need REQUIRE
->    (626) followS
->    (652) calls. That
 >    (796) I was expecting m' in (m,p',ctx').
 >    (1005) lack of space between the last two rules
->    (1118) conjunctIon
 
@@
 
-> Line 448: "If the predicate always holds" never holds?
-
-> Line 510: "isRet" is a relation, but you are using as a function.
+> Line 508: "isRet" is a relation, but you are using as a function.
 
 > Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?
-
-> Line 642: wrok
 

### 2021-09-08 submission3/main.tex
Split StkTokens criteria and new criterion
Fixed a bunch of minor things

@@
 the function activation that owns the frame and access to those locations is
-then permitted only when that an actiation at that depth is currently executing, validates our
+then permitted only when an activation at that depth is currently executing, validates our
 stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
@@
 
-We can this form of confidentiality {\em stepwise confidentiality}. Informally,
+We call this form of confidentiality {\em stepwise confidentiality}. Informally,
 for any state at the entry to a function, and for any variant over the values in the stack,
@@
 For example, under a standard calling convention, we can see that a
-return has occurred when state of the machine reaches a state where
+return has occurred when the state of the machine reaches a state where
 the stack pointer has the same value as it did at the point of some earlier
@@
 arbitrary machine code, subject to whatever well-formedness conditions the
-enforcement mechanism demands. But hardware and timing attacks are out of scope, as are
+enforcement mechanism demands. But hardware and timing attacks are out of scope, as
 are attacks on the heap---though we believe that the work of
@@
 If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate always holds, then
+predicate must hold on its final element. If the predicate never holds, then
 \(\PUT\) is the identity function.
@@
 property. For a program to enjoy stack safety
-(with respect to a particular call map), all we need require is that
-the entry sequence immediately follow  any instruction marked as a call
+(with respect to a particular call map), all we require is that
+the entry sequence immediately follows any instruction marked as a call
 in the call map. This is achieved by giving these instructions (another) special tag,
@@
 goal of evaluating real enforcement mechanisms for real machine architectures.
-Formal proof remains as possible future wrok.
+Formal proof remains as possible future work.
 
@@
 (3)
- a code map identifying the locations of calls
+ a code map identifying the locations of calls.
 % \end{enumerate}
-%
+
 That is, our generators need to produce not only a low-level program
@@
 with additional support for linear capabilities.
-Stack safety is defined as the conjuncton of local state encapsulation (LSE) and
+Stack safety is defined as the conjunction of local state encapsulation (LSE) and
 well-bracketed control flow (WBCF).
@@
 more challenging, but StkTokens does not yet handle this either.
-%
+
 Finally, we would propose one more criterion: having a generic definition

### 2021-09-08 submission3/response.txt
More small fixes

@@
 
-> line 13: variant of our properties and /does so/ correctly, and
-
 > line 94: what about tail calls?
@@
 implemented.
-
-> line 118: please briefly summarize what "stepwise" means here. Doing so should not need the full formalism, just a notion of steps. Simply hoisting the discussion from line 521 would likely suffice.
 
@@
 
-> line 210: We /call/ this form...
-
 > line 293: What about exceptions?
@@
 along with a description of what it would take to implement them.
-
-> line 563: all possible calls that /can/ be reached

### 2021-09-08 submission3/main.tex
More small fixes

@@
 micro-policy, which catches violations as early as possible, enforces a
-simple ``stepwise'' variant of our properties and correctly, and that
+simple ``stepwise'' variant of our properties, and that
 (a repaired version of) their more performant
@@
 control flow to formalize a {\em stepwise stack-safety}
-property, and argue that this property captures the intuitive concept of
-strict stack
-safety~(\cref{sec:prelim,sec:lse}).
+property, and which captures the intuitive concept of
+strict stack safety enforced over each step of execution~(\cref{sec:prelim,sec:lse}).
 \item We use property-based random testing to check that this stepwise
@@
 
-This is the set of all possible calls that be reached from any well-formed initial state.
+This is the set of all possible calls that can be reached from any well-formed initial state.
 

### 2021-09-08 submission3/response.txt
small fixes

@@
 Performing testing on another architecture such as ARM would require
-modifying the testing framework by example generator for the new ISA,
-and translating the micropolicy rules to apply to that architecture
-and its calling conventions.  This would take some work, but should be
-straightforward.
+modifying the testing framework by adding an example generator for the
+new ISA, and translating the micropolicy rules to apply to that
+architecture and its calling conventions.  This would take some work,
+but should be straightforward.
 
@@
 That said, the RISCV specification we use is complete for user mode integer instructions
-(IV32IM).
+(RV32IM).
 

### 2021-09-08 submission3/response.txt
Nit + comment

@@
 
-- Address-taken locals. As we discuss at line 954ff, a satisfactory
+- Address-taken locals. As we discuss at line 954, a satisfactory
   treatment of these seems to require integration with a more general
   model of memory safety that covers the heap and static globals too.
+  [LEO: such as "The Meaning of Memory Safety" by Amorim et al.?]
 

### 2021-09-08 submission3/main.tex
Yet more small changes: more mapping of coq content coming soon

@@
 may be instantiated with any machine that meets the following criteria, and it
-will be a meaningful defition of stack safety for that machine. It must at least be a
+will be a meaningful definition of stack safety for that machine. It must at least be a
 register machine with a program counter (\(\PCname\)), and there must be a well-defined
@@
 A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
-\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts))\), and
-\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts'))\),
+\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
+\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
 if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).
@@
 the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
-and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
+and let \((\mach',\pol',\context') \hookrightarrow \MPCT'\).
 Then for any \(\nach\) such that
@@
 
-\vspace*{-5ex}
+\vspace*{-3ex}
 \judgment[ Default]

### 2021-09-09 submission3/response.txt
Yet more small changes: more mapping of coq content coming soon
Response

@@
-> Review #416A
-> Overall merit
+We first address for the most important high-level points, then go back to
+fill in detailed responses to all points made in the reviews.
 
-> B. Weak Accept
-> Reviewer expertise
+_[BCP: Fill this part in!!]_
 
-> X. Expert
+___________________________________________________________________
+
+_[BCP: This could all use some more attention to markdown formatting so that
+the output looks readable.]_
+
+_[BCP: Still some comments below marked BCP/LEO...]_
+
+# Reviewer A
+
 > Paper summary
 
-> This submission formalizes the general notion of stack safety (roughly speaking, every caller should be protected from its callees during program execution) using well-known notions of security (integrity, confidentiality and non interference), and defines several security policies for protecting the stack.
-The formalization is validated on existing (and less general) micro-policies for stack safety (depth isolation, lazy tagging and lazy clearing) using property-based random testing.
-This evaluation of former policies detected some security violations in former and less precise works.
+> This submission formalizes the general notion of stack safety (roughly
+  speaking, every caller should be protected from its callees during program
+  execution) using well-known notions of security (integrity,
+  confidentiality and non interference), and defines several security
+  policies for protecting the stack. The formalization is validated on
+  existing (and less general) micro-policies for stack safety (depth
+  isolation, lazy tagging and lazy clearing) using property-based random
+  testing. This evaluation of former policies detected some security
+  violations in former and less precise works.
 
-> First, this submission introduces the main required security policies, in terms of traces of memory, register states and stepwise stack-safety: stepwise integrity (an execution step does not modify the memory slots that are reserved for a caller at a given depth), confidentiality (a callee is insensitive to the context of its call), stepwise confidentiality, temporal stack safety, and well-bracketed control-flow (calls and returns should be correctly nested).
-Then, the submission describes the requirements on the machine model (presence of a program counter, used to distinguish between the different stack-frames, and a mechanism for recognizing the return from a call) and the threat model.
-The machine model is abstract enough to represent different architectures.
+> First, this submission introduces the main required security policies, in
+  terms of traces of memory, register states and stepwise stack-safety:
+  stepwise integrity (an execution step does not modify the memory slots
+  that are reserved for a caller at a given depth), confidentiality (a
+  callee is insensitive to the context of its call), stepwise
+  confidentiality, temporal stack safety, and well-bracketed control-flow
+  (calls and returns should be correctly nested). Then, the submission
+  describes the requirements on the machine model (presence of a program
+  counter, used to distinguish between the different stack-frames, and a
+  mechanism for recognizing the return from a call) and the threat
+  model. The machine model is abstract enough to represent different
+  architectures.
 
-> Next, the submission details in a progressive way the formalization of machine behaviors and stack safety using semantic rules.
-The first formalization of the execution stack is simple, with a single stack and no sharing between callers and callees.
-Then the model is enriched with function parameters passed on the stack (by value or by reference), with callee-saves registers and with coroutines.
-Moreover, the first model is implemented in Coq and this implementation reuses an existing RISCV specification.
-The QuickChick tool is used to test the stepwise property and validate an existing enforcement mechanism (called depth isolation) of the literature.
-Furthermore, as the previous policy is conservative and rather slow to be practical, more permissive variants of confidentiality and integrity properties are formalized, resulting in an observational property that is again validated by testing.
+> Next, the submission details in a progressive way the formalization of
+  machine behaviors and stack safety using semantic rules. The first
+  formalization of the execution stack is simple, with a single stack and no
+  sharing between callers and callees. Then the model is enriched with
+  function parameters passed on the stack (by value or by reference), with
+  callee-saves registers and with coroutines. Moreover, the first model is
+  implemented in Coq and this implementation reuses an existing RISCV
+  specification. The QuickChick tool is used to test the stepwise property
+  and validate an existing enforcement mechanism (called depth isolation) of
+  the literature. Furthermore, as the previous policy is conservative and
+  rather slow to be practical, more permissive variants of confidentiality
+  and integrity properties are formalized, resulting in an observational
+  property that is again validated by testing.
+
 > Strengths
@@
 > Why did you choose this RISCV specification?
-> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?
+> Why didn't you choose to reuse an ARM specification (as there are some of
+> them written in Coq in the literature)? 
 > Would it be easy to adapt your work to another ISA specification?
 
-We chose RISCV for its simplicity, and because it has been used as the
-base ISA for much of the existing work (by others) to design
-hardware that supports tag-based micro-policies.
+We chose RISCV for its simplicity, and because it has been used as the base
+ISA for much of the existing work (by others) on designing hardware that
+supports tag-based micro-policies.
 
-To clarify a possible misunderstanding: the Coq specification of RISCV
-that we use was developed by others (see footnote p. 14) and we were
-able to adapt it to our testing framework with relatively little
-effort.
+To clarify a possible misunderstanding: the Coq specification of RISCV that
+we use was developed by others (see footnote p. 14); we were able to adapt
+it to our testing framework with relatively little effort.
 
 Our stack safety theory itself is not dependent on the details of the
-architecture, as in fact the properties are defined in terms of a
-generic interface over machine and enforcement policy; [see ... in the
-Coq development]
+architecture, since the properties are defined in terms of a generic
+interface over machine and enforcement policy; [see ... in the Coq
+development] [BCP: Fill in!]
 
-Performing testing on another architecture such as ARM would require
-modifying the testing framework by adding an example generator for the
-new ISA, and translating the micropolicy rules to apply to that
-architecture and its calling conventions.  This would take some work,
-but should be straightforward.
+Porting the testing framework to another architecture such as ARM would
+require adding a test case generator for the new ISA (including its specific
+calling conventions) and translating the micropolicy rules to apply to that
+architecture / calling conventions.  This would take some work, but it
+should be straightforward.
 
-We are currently working on extending the work to a model of a Cheri-style
-hardware capability machine, Cerise, and have not encountered major
-obstacles.
+For example, we are currently working on extending the work to a model of a
+Cheri-style hardware capability machine, Cerise, and have not encountered
+major obstacles.
 
-> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?
+> How realistic is the RISCV specification? In other words, are there any
+  RISCV vulnerabilities that are not taken into account in your model?
 
-We use the RISCV specification to demonstrate that our properties are testable in
-a realistic setting. We do not exercise the entire RISCV specification, and the validity
-of the model is not dependent on the absence of vulnerabilities in the RISCV implementation.
-That said, the RISCV specification we use is complete for user mode integer instructions
-(RV32IM).
+We use the RISCV specification to demonstrate that our properties are
+testable in a realistic setting. We do not aim to exercise the entire RISCV
+specification, and the validity of the model is not dependent on the absence
+of vulnerabilities in the RISCV implementation. [BCP: It's not??] That said,
+the RISCV specification we use is complete for user mode integer
+instructions (RV32IM).
 
@@
 throughout the paper.  We will expand and organize it into an explicit
-section, as follows.
+section, along the following lines...
 
-Our goal is to specifically characterize stack safety.  So we are
-particularly concerned with limitations that prevent the model from
-applying to realistic uses of the stack, and not with other structures
-such as the heap, which we intentionally make no guarantees about.
+Our focus is on characterizing stack safety.  So we are particularly
+concerned with limitations that prevent the model from applying to realistic
+uses of the stack, as opposed to modeling other structures such as the heap,
+which we intentionally say nothing about.  (Heap safety has already been
+studied, in isolation, by Azevedo et al, and we expect that their
+correctness condition and ours can be combined cleanly to characterize
+correct protection of stack + heap.)
 
@@
   [LEO: such as "The Meaning of Memory Safety" by Amorim et al.?]
+  [BCP: Note that this is mentioned now, just above.]
 
@@
 
-- Concurrency. We can handle a limited form of coroutines, as
-  described in 8.3. But handling true concurrency would require
-  significant changes in our underlying machine model.
+- Concurrency. We can handle a limited form of coroutines, as described in
+  8.3. Handling full-blown preemptive concurrency would require significant
+  changes in our underlying machine model.
 
-As for the distinction between values and addresses, conventional ISAs
-do not make this distinction.  Ongoing work on the Cerise capability
-machine formalization has required some modification to distinguish
-values from addresses, as they are not the same in a capability
-machine, but it has little effect on the model itself. Further,
-distinguishing between values and addresses is a higher-level
-abstraction that can "hide" potential bugs resulting from implicit
-conversions between them, so it may even be preferable to model them
-as the same when possible.
+As for the distinction between values and addresses, conventional ISAs do
+not make this distinction.  Ongoing work on the Cerise capability machine
+formalization has required some modifications to our model to distinguish
+values from addresses, as they are _not_ the same in a capability machine,
+but it has little global effect on the model itself. Indeed, distinguishing
+between values and addresses is a higher-level abstraction that can "hide"
+potential bugs resulting from implicit conversions between them, so we feel
+it is preferable to model them as the same when possible.
 
-> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?
+> Another example is the outside qualifier that characterizes only a whole
+  component. Wouldn't you need a finer-grain access control policy (and a
+  qualifier for partially outside components)?
 
-There may be some confusion arising from our use of the term
-"components."  In our model (see line 398), components are already
-fine-grained: individual registers and memory addresses. We recognize
-that this is a poor choice of term given the common usage of component
-to refer to a larger module in a compartmentalization context, and
-will change the term to "state elements." Since components/elements
-are so fine-grained, there is no need for finer access control.
+We don't entirely understand this question, but we wonder whether there may
+be some confusion arising from our use of the term "components."  In our
+model (see line 398), components are already fine-grained: individual
+registers and memory addresses. We recognize that this is a poor choice of
+term given the common usage of component to refer to a larger module in a
+compartmentalization context, and we plan to change the term to "state
+elements."  Since components/elements are so fine-grained, there is no need
+for finer access control.
 
-> Could you comment on the Coq formalization that is provided as a supplementary material?
-Was it straightforward to write a Coq program from a C++ program?
-Moreover, I looked at the Coq development and found some admitted lemmas in it.
+> Could you comment on the Coq formalization that is provided as a
+  supplementary material? Was it straightforward to write a Coq program from
+  a C++ program? Moreover, I looked at the Coq development and found some
+  admitted lemmas in it.
 
@@
 
-Admitted lemmas are the beginnings of an effort to formalize [details]
+Admitted lemmas are the beginnings of an effort to formalize [details] [BCP:
+N.b.]
 
@@
 >    (310, 886,895,947) The return instruction is confusing as the return type of f is void.
->    (568) two many closing parentheses (twice)
->    (796) I was expecting m' in (m,p',ctx').
->    (1005) lack of space between the last two rules
+
+[SNA: should we do jlr?]
 
@@
 
-====================================================================================================================================
+___________________________________________________________________________
+# Reviewer B
 
-> Review #416B
-> Overall merit
-
-> C. Weak Reject
-> Reviewer expertise
-
-> Y. Knowledgeable
 > Paper summary
@@
 
-Many of the content that seems disconnected takes the form of a parameterizable relation and its various
-instantiations in different versions of the property. \rightarrow_C parameterized by the type ctx, which is section 5
-is instantiated as (K -> SD)x(list(target)). In later sections it will be instantiated by other types.
-Then for any \rightarrow_C, we can extend \rightharpoonup to form \rightharpoonup_C.
+Much of the content that may seem disconnected takes the form of a
+parameterizable relation and its various instantiations in different
+versions of the property. \rightarrow_C parameterized by the type ctx, which
+is section 5 is instantiated as (K -> SD)x(list(target)). In later sections
+it will be instantiated by other types.  Then for any \rightarrow_C, we can
+extend \rightharpoonup to form \rightharpoonup_C.
+[BCP: The formatting here makes it hard to read.  Can we improve it?]
 
-We will do more to call out which parts of the formalization are parameterized and when they are being
-instantiated, so that the connections are more clear.
+We will do more to call out which parts of the formalization are
+parameterized and when they are being instantiated, so that the connections
+are more clear.
 
-> Line 320: defition
+> Line 329: I was really interested in address taken at this point, and
+  whether the model could handle it. I really think the informal aspects of
+  the extensions, should be pulled forward, so these questions are answered
+  earlier
 
-[BCP: We should just fix these sorts of things and then not mention them in
-the response.]
+This is a good point, and we will examine ways to introduce that information
+sooner. But we are hesitant to go into too much detail on variants while
+introducing the basic formalism, because it can be a lot to keep track of.
 
-> Line 329: I was really interested in address taken at this point, and whether the model could handle it.
-I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier
+> Line 349: How does this related to control flow integrity/guard, and
+  Intel's CET?
 
-This is a good point, and we will examine ways to introduce that information sooner. However, we are
-hesitant to spend too much time on variants before introducing the basic formalism, because it can
-be a lot to keep track of.
-
-[BCP: This is a good expositional idea.]
-[APT: Right, so we should acknowledge it politely -- but we did have reasons for ordering
-things the way we did....]
-[SNA: That was roughly the reason, as I recall? I may take a stab tomorrow at sketching what
-moving them up would look like.]
-
-> Line 349: How does this related to control flow integrity/guard, and Intel's CET?
+[BCP: ???]
 
@@
 
-> Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?
+> Line 515: ReturnRec: This seems like it would allow you to access outer
+  frames as long as you didn't return?
 
-====================================================================================================================================
+[BCP: ???]
 
-> Review #416C
-> Overall merit
+___________________________________________________________________________
+# Reviewer C
 
-> B. Weak Accept
-> Reviewer expertise
-
-> Z. No familiarity
 > Strengths
 
-> The work presents a compelling model for stack security, framed in terms of existing notions of integrity and confidentiality.
+> The work presents a compelling model for stack security, framed in terms
+  of existing notions of integrity and confidentiality.
 
-> Models are sufficiently abstract and general to be transportable to a variety of mechanisms and machines.
+> Models are sufficiently abstract and general to be transportable to a
+  variety of mechanisms and machines.
 
-> Notions of observable stack safety capture a more idealized notion that can reflect realistic violations, rather than "false positives" due to the limitations of particular mechanisms.
-This is my favorite contribution of this work.
+> Notions of observable stack safety capture a more idealized notion that
+  can reflect realistic violations, rather than "false positives" due to the
+  limitations of particular mechanisms. This is my favorite contribution of
+  this work.
 
-> Some evidence that the approach can be tailored to complex control mechanisms, by showing how it applies to a simple fixed coroutine model.
+> Some evidence that the approach can be tailored to complex control
+  mechanisms, by showing how it applies to a simple fixed coroutine model.
+
 > Weaknesses
 
-> Validation is performed using random testing rather than formal proof.
-However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
+> Validation is performed using random testing rather than formal
+  proof. However, testing proved sufficient to uncover shortcomings in prior
+  lazy models (which could then be fixed), and mutation testing was used to
+  ensure that the testing approach would catch errors.
+
 > Comments for author
@@
 
-Tail calls are discussed above in our proposed "limitations" section.
-They are not currently supported, but we describe how they might be
-implemented.
+Tail calls are discussed above in our proposed "limitations" section.  They
+are not currently supported, but we describe there how they might be.
 
@@
 
-Line 123 should read "... we extend the generation-by-execution techniques...",
-referring to the method of randomly generating valid low-level programs by lazily adding
-instructions while simulating execution.
+Line 123 should read "... we extend the generation-by-execution
+techniques...", referring to the method of randomly generating valid
+low-level programs by lazily adding instructions while simulating execution.
 
@@
 
-Exceptions are also discussed about in our "limitations" section,
-along with a description of what it would take to implement them.
+Exceptions are also discussed in our "limitations" section, along with a
+description of what it would take to implement them.

### 2021-09-09 submission3/response.txt
Comment bash 1

@@
 architecture, since the properties are defined in terms of a generic
-interface over machine and enforcement policy; [see ... in the Coq
-development] [BCP: Fill in!]
+interface over machine and enforcement policy; see the `Properties`
+directory in the Coq development as explained below for concrete
+modelling details.
 

### 2021-09-09 submission3/response.txt
Comment bashing

@@
 specification, and the validity of the model is not dependent on the absence
-of vulnerabilities in the RISCV implementation. [BCP: It's not??] That said,
+of (not stack-related) vulnerabilities in the RISCV implementation. That said,
 the RISCV specification we use is complete for user mode integer

### 2021-09-09 submission3/response.txt
Comment bashing cont.

@@
   treatment of these seems to require integration with a more general
-  model of memory safety that covers the heap and static globals too.
-  [LEO: such as "The Meaning of Memory Safety" by Amorim et al.?]
-  [BCP: Note that this is mentioned now, just above.]
+  model of memory safety that covers the heap and static globals too,
+  such as the aforementioned model of Azevedo et al.
 

### 2021-09-09 submission3/response.txt
Map of coq code

--- 2021-09-09 submission3/response.txt
@@
-We first address for the most important high-level points, then go back to
-fill in detailed responses to all points made in the reviews.
+> Review #416A
+> Overall merit
 
-_[BCP: Fill this part in!!]_
+> B. Weak Accept
+> Reviewer expertise
 
-___________________________________________________________________
-
-_[BCP: This could all use some more attention to markdown formatting so that
-the output looks readable.]_
-
-_[BCP: Still some comments below marked BCP/LEO...]_
-
-# Reviewer A
-
+> X. Expert
 > Paper summary
 
-> This submission formalizes the general notion of stack safety (roughly
-  speaking, every caller should be protected from its callees during program
-  execution) using well-known notions of security (integrity,
-  confidentiality and non interference), and defines several security
-  policies for protecting the stack. The formalization is validated on
-  existing (and less general) micro-policies for stack safety (depth
-  isolation, lazy tagging and lazy clearing) using property-based random
-  testing. This evaluation of former policies detected some security
-  violations in former and less precise works.
+> This submission formalizes the general notion of stack safety (roughly speaking, every caller should be protected from its callees during program execution) using well-known notions of security (integrity, confidentiality and non interference), and defines several security policies for protecting the stack.
+The formalization is validated on existing (and less general) micro-policies for stack safety (depth isolation, lazy tagging and lazy clearing) using property-based random testing.
+This evaluation of former policies detected some security violations in former and less precise works.
 
-> First, this submission introduces the main required security policies, in
-  terms of traces of memory, register states and stepwise stack-safety:
-  stepwise integrity (an execution step does not modify the memory slots
-  that are reserved for a caller at a given depth), confidentiality (a
-  callee is insensitive to the context of its call), stepwise
-  confidentiality, temporal stack safety, and well-bracketed control-flow
-  (calls and returns should be correctly nested). Then, the submission
-  describes the requirements on the machine model (presence of a program
-  counter, used to distinguish between the different stack-frames, and a
-  mechanism for recognizing the return from a call) and the threat
-  model. The machine model is abstract enough to represent different
-  architectures.
+> First, this submission introduces the main required security policies, in terms of traces of memory, register states and stepwise stack-safety: stepwise integrity (an execution step does not modify the memory slots that are reserved for a caller at a given depth), confidentiality (a callee is insensitive to the context of its call), stepwise confidentiality, temporal stack safety, and well-bracketed control-flow (calls and returns should be correctly nested).
+Then, the submission describes the requirements on the machine model (presence of a program counter, used to distinguish between the different stack-frames, and a mechanism for recognizing the return from a call) and the threat model.
+The machine model is abstract enough to represent different architectures.
 
-> Next, the submission details in a progressive way the formalization of
-  machine behaviors and stack safety using semantic rules. The first
-  formalization of the execution stack is simple, with a single stack and no
-  sharing between callers and callees. Then the model is enriched with
-  function parameters passed on the stack (by value or by reference), with
-  callee-saves registers and with coroutines. Moreover, the first model is
-  implemented in Coq and this implementation reuses an existing RISCV
-  specification. The QuickChick tool is used to test the stepwise property
-  and validate an existing enforcement mechanism (called depth isolation) of
-  the literature. Furthermore, as the previous policy is conservative and
-  rather slow to be practical, more permissive variants of confidentiality
-  and integrity properties are formalized, resulting in an observational
-  property that is again validated by testing.
-
+> Next, the submission details in a progressive way the formalization of machine behaviors and stack safety using semantic rules.
+The first formalization of the execution stack is simple, with a single stack and no sharing between callers and callees.
+Then the model is enriched with function parameters passed on the stack (by value or by reference), with callee-saves registers and with coroutines.
+Moreover, the first model is implemented in Coq and this implementation reuses an existing RISCV specification.
+The QuickChick tool is used to test the stepwise property and validate an existing enforcement mechanism (called depth isolation) of the literature.
+Furthermore, as the previous policy is conservative and rather slow to be practical, more permissive variants of confidentiality and integrity properties are formalized, resulting in an observational property that is again validated by testing.
 > Strengths
@@
 > Why did you choose this RISCV specification?
-> Why didn't you choose to reuse an ARM specification (as there are some of
-> them written in Coq in the literature)? 
+> Why didn't you choose to reuse an ARM specification (as there are some of them written in Coq in the literature)?
 > Would it be easy to adapt your work to another ISA specification?
 
-We chose RISCV for its simplicity, and because it has been used as the base
-ISA for much of the existing work (by others) on designing hardware that
-supports tag-based micro-policies.
+We chose RISCV for its simplicity, and because it has been used as the
+base ISA for much of the existing work (by others) to design
+hardware that supports tag-based micro-policies.
 
-To clarify a possible misunderstanding: the Coq specification of RISCV that
-we use was developed by others (see footnote p. 14); we were able to adapt
-it to our testing framework with relatively little effort.
+To clarify a possible misunderstanding: the Coq specification of RISCV
+that we use was developed by others (see footnote p. 14) and we were
+able to adapt it to our testing framework with relatively little
+effort.
 
 Our stack safety theory itself is not dependent on the details of the
-architecture, since the properties are defined in terms of a generic
-interface over machine and enforcement policy; see the `Properties`
-directory in the Coq development as explained below for concrete
-modelling details.
+architecture, as in fact the properties are defined in terms of a
+generic interface over machine and enforcement policy; [see ... in the
+Coq development]
 
-Porting the testing framework to another architecture such as ARM would
-require adding a test case generator for the new ISA (including its specific
-calling conventions) and translating the micropolicy rules to apply to that
-architecture / calling conventions.  This would take some work, but it
-should be straightforward.
+Performing testing on another architecture such as ARM would require
+modifying the testing framework by adding an example generator for the
+new ISA, and translating the micropolicy rules to apply to that
+architecture and its calling conventions.  This would take some work,
+but should be straightforward.
 
-For example, we are currently working on extending the work to a model of a
-Cheri-style hardware capability machine, Cerise, and have not encountered
-major obstacles.
+We are currently working on extending the work to a model of a Cheri-style
+hardware capability machine, Cerise, and have not encountered major
+obstacles.
 
-> How realistic is the RISCV specification? In other words, are there any
-  RISCV vulnerabilities that are not taken into account in your model?
+> How realistic is the RISCV specification? In other words, are there any RISCV vulnerabilities that are not taken into account in your model?
 
-We use the RISCV specification to demonstrate that our properties are
-testable in a realistic setting. We do not aim to exercise the entire RISCV
-specification, and the validity of the model is not dependent on the absence
-of (not stack-related) vulnerabilities in the RISCV implementation. That said,
-the RISCV specification we use is complete for user mode integer
-instructions (RV32IM).
+We use the RISCV specification to demonstrate that our properties are testable in
+a realistic setting. We do not exercise the entire RISCV specification, and the validity
+of the model is not dependent on the absence of vulnerabilities in the RISCV implementation.
+That said, the RISCV specification we use is complete for user mode integer instructions
+(RV32IM).
 
@@
 throughout the paper.  We will expand and organize it into an explicit
-section, along the following lines...
+section, as follows.
 
-Our focus is on characterizing stack safety.  So we are particularly
-concerned with limitations that prevent the model from applying to realistic
-uses of the stack, as opposed to modeling other structures such as the heap,
-which we intentionally say nothing about.  (Heap safety has already been
-studied, in isolation, by Azevedo et al, and we expect that their
-correctness condition and ours can be combined cleanly to characterize
-correct protection of stack + heap.)
+Our goal is to specifically characterize stack safety.  So we are
+particularly concerned with limitations that prevent the model from
+applying to realistic uses of the stack, and not with other structures
+such as the heap, which we intentionally make no guarantees about.
 
@@
   [LEO: such as "The Meaning of Memory Safety" by Amorim et al.?]
-  [BCP: Note that this is mentioned now, just above.]
 
@@
 
-- Concurrency. We can handle a limited form of coroutines, as described in
-  8.3. Handling full-blown preemptive concurrency would require significant
-  changes in our underlying machine model.
+- Concurrency. We can handle a limited form of coroutines, as
+  described in 8.3. But handling true concurrency would require
+  significant changes in our underlying machine model.
 
-As for the distinction between values and addresses, conventional ISAs do
-not make this distinction.  Ongoing work on the Cerise capability machine
-formalization has required some modifications to our model to distinguish
-values from addresses, as they are _not_ the same in a capability machine,
-but it has little global effect on the model itself. Indeed, distinguishing
-between values and addresses is a higher-level abstraction that can "hide"
-potential bugs resulting from implicit conversions between them, so we feel
-it is preferable to model them as the same when possible.
+As for the distinction between values and addresses, conventional ISAs
+do not make this distinction.  Ongoing work on the Cerise capability
+machine formalization has required some modification to distinguish
+values from addresses, as they are not the same in a capability
+machine, but it has little effect on the model itself. Further,
+distinguishing between values and addresses is a higher-level
+abstraction that can "hide" potential bugs resulting from implicit
+conversions between them, so it may even be preferable to model them
+as the same when possible.
 
-> Another example is the outside qualifier that characterizes only a whole
-  component. Wouldn't you need a finer-grain access control policy (and a
-  qualifier for partially outside components)?
+> Another example is the outside qualifier that characterizes only a whole component. Wouldn't you need a finer-grain access control policy (and a qualifier for partially outside components)?
 
-We don't entirely understand this question, but we wonder whether there may
-be some confusion arising from our use of the term "components."  In our
-model (see line 398), components are already fine-grained: individual
-registers and memory addresses. We recognize that this is a poor choice of
-term given the common usage of component to refer to a larger module in a
-compartmentalization context, and we plan to change the term to "state
-elements."  Since components/elements are so fine-grained, there is no need
-for finer access control.
+There may be some confusion arising from our use of the term
+"components."  In our model (see line 398), components are already
+fine-grained: individual registers and memory addresses. We recognize
+that this is a poor choice of term given the common usage of component
+to refer to a larger module in a compartmentalization context, and
+will change the term to "state elements." Since components/elements
+are so fine-grained, there is no need for finer access control.
 
-> Could you comment on the Coq formalization that is provided as a
-  supplementary material? Was it straightforward to write a Coq program from
-  a C++ program? Moreover, I looked at the Coq development and found some
-  admitted lemmas in it.
+> Could you comment on the Coq formalization that is provided as a supplementary material?
+Was it straightforward to write a Coq program from a C++ program?
+Moreover, I looked at the Coq development and found some admitted lemmas in it.
 
@@
 
-Admitted lemmas are the beginnings of an effort to formalize [details] [BCP:
-N.b.]
+Admitted lemmas are the beginnings of an effort to formalize [details]
 
@@
 - Trace.v and ObsTrace.v define the trace model and observation model
-- TraceProperties.v defines arbitrary stack properties given a call trace and context type
-- SubroutineSimple.v defines the basic stack safety properties
-- SubroutineShare.v defines the version enhanced with argument passing on the stack
-- Coroutine.v defines the coroutine version
-- Lazy.v defines the lazy properties
-- CalleeSave.v defines stack safety with callee-save registers
+
+- SubroutineSimple.v defines the basic stack safety properties.
+  Some comments are outdated, and will be updated or removed for the final submission.
+
+  \rightarrow_C => Line 126 (disregard comment regarding returns)
+  Definition 5.2 => Line 184
+  Definition 5.9 => Line 255
+  Definition 5.10 => Line 277
+
+- CalleeSave.v defines stack safety with callee-save registers.
+  \rightarrow_C => Line 62
+  Definition 8.6 => Line 123 & Line 130
+  These definitions refer to general integrity and confidentiality properties for arbitrary
+  traces defined in TraceProperties.v at Lines 51 & 60.
+
+- SubroutineShare.v defines the version enhanced with argument passing on the stack.
+  Note that a preliminary treatment of address-taken locals is present; we found this
+  insufficient for inclusion in the paper.
+
+  \rightarrow_C => Line 82
+  Definition 8.3 (Stepwise only) => Line 152
+  Note that this definition refers to a general confidentiality property of traces defined in 
+  TraceProperties.v at Line 35.
+
+- Coroutine.v defines the coroutine version. Unlike in the paper, passing on the stack is included.
+  \rightarrow_C => Line 107
+  Definition 8.10 (Stepwise only) => Line 192 & Line 199
+  Definition 8.11 (Stepwise only) => Line 212 & Line 221
+  This also refers to the general properties presented in TraceProperties.v
+
+  Lazy.v is an outdated file that was included by mistake.
+  [SNA: I'm not sure how strongly we should call this out, but it's pretty obvious that it doesn't match the paper.]
+
 Testing contains the test system, with generators for different machines and the final test run
 via extraction in TestProperties.v
-[APT: Can we give pointers from each paper definition to corresponding Coq definition?]
 
@@
 
-___________________________________________________________________________
-# Reviewer B
+====================================================================================================================================
 
+> Review #416B
+> Overall merit
+
+> C. Weak Reject
+> Reviewer expertise
+
+> Y. Knowledgeable
 > Paper summary
@@
 
-Much of the content that may seem disconnected takes the form of a
-parameterizable relation and its various instantiations in different
-versions of the property. \rightarrow_C parameterized by the type ctx, which
-is section 5 is instantiated as (K -> SD)x(list(target)). In later sections
-it will be instantiated by other types.  Then for any \rightarrow_C, we can
-extend \rightharpoonup to form \rightharpoonup_C.
-[BCP: The formatting here makes it hard to read.  Can we improve it?]
+Many of the content that seems disconnected takes the form of a parameterizable relation and its various
+instantiations in different versions of the property. \rightarrow_C parameterized by the type ctx, which is section 5
+is instantiated as (K -> SD)x(list(target)). In later sections it will be instantiated by other types.
+Then for any \rightarrow_C, we can extend \rightharpoonup to form \rightharpoonup_C.
 
-We will do more to call out which parts of the formalization are
-parameterized and when they are being instantiated, so that the connections
-are more clear.
+We will do more to call out which parts of the formalization are parameterized and when they are being
+instantiated, so that the connections are more clear.
 
-> Line 329: I was really interested in address taken at this point, and
-  whether the model could handle it. I really think the informal aspects of
-  the extensions, should be pulled forward, so these questions are answered
-  earlier
+> Line 329: I was really interested in address taken at this point, and whether the model could handle it.
+I really think the informal aspects of the extensions, should be pulled forward, so these questions are answered earlier
 
-This is a good point, and we will examine ways to introduce that information
-sooner. But we are hesitant to go into too much detail on variants while
-introducing the basic formalism, because it can be a lot to keep track of.
+This is a good point, and we will examine ways to introduce that information sooner. However, we are
+hesitant to spend too much time on variants before introducing the basic formalism, because it can
+be a lot to keep track of.
 
-> Line 349: How does this related to control flow integrity/guard, and
-  Intel's CET?
+[BCP: This is a good expositional idea.]
+[APT: Right, so we should acknowledge it politely -- but we did have reasons for ordering
+things the way we did....]
+[SNA: That was roughly the reason, as I recall? I may take a stab tomorrow at sketching what
+moving them up would look like.]
 
-[BCP: ???]
+> Line 349: How does this related to control flow integrity/guard, and Intel's CET?
 
@@
 
-> Line 515: ReturnRec: This seems like it would allow you to access outer
-  frames as long as you didn't return?
+> Line 515: ReturnRec: This seems like it would allow you to access outer frames as long as you didn't return?
 
-[BCP: ???]
+====================================================================================================================================
 
-___________________________________________________________________________
-# Reviewer C
+> Review #416C
+> Overall merit
 
+> B. Weak Accept
+> Reviewer expertise
+
+> Z. No familiarity
 > Strengths
 
-> The work presents a compelling model for stack security, framed in terms
-  of existing notions of integrity and confidentiality.
+> The work presents a compelling model for stack security, framed in terms of existing notions of integrity and confidentiality.
 
-> Models are sufficiently abstract and general to be transportable to a
-  variety of mechanisms and machines.
+> Models are sufficiently abstract and general to be transportable to a variety of mechanisms and machines.
 
-> Notions of observable stack safety capture a more idealized notion that
-  can reflect realistic violations, rather than "false positives" due to the
-  limitations of particular mechanisms. This is my favorite contribution of
-  this work.
+> Notions of observable stack safety capture a more idealized notion that can reflect realistic violations, rather than "false positives" due to the limitations of particular mechanisms.
+This is my favorite contribution of this work.
 
-> Some evidence that the approach can be tailored to complex control
-  mechanisms, by showing how it applies to a simple fixed coroutine model.
-
+> Some evidence that the approach can be tailored to complex control mechanisms, by showing how it applies to a simple fixed coroutine model.
 > Weaknesses
 
-> Validation is performed using random testing rather than formal
-  proof. However, testing proved sufficient to uncover shortcomings in prior
-  lazy models (which could then be fixed), and mutation testing was used to
-  ensure that the testing approach would catch errors.
-
+> Validation is performed using random testing rather than formal proof.
+However, testing proved sufficient to uncover shortcomings in prior lazy models (which could then be fixed), and mutation testing was used to ensure that the testing approach would catch errors.
 > Comments for author
@@
 
-Tail calls are discussed above in our proposed "limitations" section.  They
-are not currently supported, but we describe there how they might be.
+Tail calls are discussed above in our proposed "limitations" section.
+They are not currently supported, but we describe how they might be
+implemented.
 
@@
 
-Line 123 should read "... we extend the generation-by-execution
-techniques...", referring to the method of randomly generating valid
-low-level programs by lazily adding instructions while simulating execution.
+Line 123 should read "... we extend the generation-by-execution techniques...",
+referring to the method of randomly generating valid low-level programs by lazily adding
+instructions while simulating execution.
 
@@
 
-Exceptions are also discussed in our "limitations" section, along with a
-description of what it would take to implement them.
+Exceptions are also discussed about in our "limitations" section,
+along with a description of what it would take to implement them.

### 2021-09-09 submission3/response.txt
Comments
Answer to Sean

@@
 The supplementary material should indeed be mentioned in the submission.
-The Coq formalization provided contains two parts: the formalization of our properties,
-which was a conceptual aid in their development and may be useful to readers' intuitions,
-and the testing framework, which is used to validate our properties by showing that
-computable versions of them are testable.
+
+The Coq formalization provided contains two parts: the formalization of our
+properties, which was a conceptual aid in their development and may be
+useful to readers' intuitions, and the testing framework, which is used to
+validate our properties by showing that computable versions of them are
+testable.
 
 Admitted lemmas are the beginnings of an effort to formalize [details] [BCP:
-N.b.]
+N.b. missing]
 
@@
 Map of the supplementary material:
-Core contains the abstract machine and policy models, as well as the RISCV instantiations with different policies.
+
+Core contains the abstract machine and policy models, as well as the RISCV
+instantiations with different policies.
+
 Properties contains Coq definitions corresponding to the different Properties:
@@
 - SubroutineSimple.v defines the basic stack safety properties.
-  Some comments are outdated, and will be updated or removed for the final submission.
+  Some comments are outdated, and will be updated or removed for the final
+  submission.
 
@@
   [SNA: I'm not sure how strongly we should call this out, but it's pretty obvious that it doesn't match the paper.]
+  [BCP: This is fine, IMO.]
 

### 2021-09-09 submission3/response.txt
Answer to question

--- 2021-09-09 submission3/response.txt
@@
 Admitted lemmas are the beginnings of an effort to formalize [details] [BCP:
-N.b.]
+N.b.][LEO: Not sure what we can say here.]
 
@@
 
-[BCP: ???]
+[LEO: Sean, check this.]
+All ReturnRec does is unseal the frames between the levels before and
+after the return. Frames marked "outside" are left unaffected, while
+out frames that had been sealed by earlier calls can only be accessed
+after returning to the outer level.
 

### 2021-09-09 submission3/response.txt

--- 2021-09-09 submission3/response.txt
@@
 Admitted lemmas are the beginnings of an effort to formalize [details] [BCP:
-N.b. missing]
+N.b. missing][LEO: Not sure what we can say here.]
 

### 2021-09-09 submission3/response.txt
address a few comments

@@
 
-Admitted lemmas are the beginnings of an effort to formalize [details] [BCP:
-N.b. missing][LEO: Not sure what we can say here.]
+The theorems and lemmas in the formalization files serve only as
+sanity checks on our definitions.  The Admitted and commented-out ones
+are not essential, and were included by mistake.
+[BCP: N.b. missing][LEO: Not sure what we can say here.][APT: tried to complete.]
 
@@
   [SNA: I'm not sure how strongly we should call this out, but it's pretty obvious that it doesn't match the paper.]
-  [BCP: This is fine, IMO.]
+  [BCP: This is fine, IMO.][APT:+1]
 
@@
 
-This is a good point, and we will examine ways to introduce that information
-sooner. But we are hesitant to go into too much detail on variants while
-introducing the basic formalism, because it can be a lot to keep track of.
+This is a good point, and we will examine ways to introduce that
+information sooner. But we are hesitant to go into too much detail on
+variants while introducing the basic formalism, because it can be a
+lot for the reader to keep track of.
 
@@
 
-[BCP: ???]
+[BCP: ???]  [APT: made a stab; others please check.]
+The identification of call points, as described in this paragraph, is
+required for all our properties: integrity, confidentiality, and
+well-bracketed control flow (WBCF). Software or hardware support for
+control flow integrity (e.g. the CET shadow stack) could be used to
+enforce WBCF, but not the other two properties.
 
@@
 
-Exceptions are also discussed in our "limitations" section, along with a
-description of what it would take to implement them.
+Exceptions are also discussed above in our "limitations" section,
+along with a description of what it would take to implement them.

### 2021-09-09 submission3/response.txt
wibbles

@@
-We first address for the most important high-level points, then go back to
+We first address the most important high-level points, then go back to
 fill in detailed responses to all points made in the reviews.
@@
 
-> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+> More generally, what are the limitations of the presented work?
 
@@
 throughout the paper.  We will expand and organize it into an explicit
-section, along the following lines...
+section, along the following lines:
 
@@
 
-As for the distinction between values and addresses, conventional ISAs do
-not make this distinction.  Ongoing work on the Cerise capability machine
-formalization has required some modifications to our model to distinguish
-values from addresses, as they are _not_ the same in a capability machine,
-but it has little global effect on the model itself. Indeed, distinguishing
-between values and addresses is a higher-level abstraction that can "hide"
-potential bugs resulting from implicit conversions between them, so we feel
-it is preferable to model them as the same when possible.
+> For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+
+Conventional ISAs do not make this distinction.  Our ongoing work on the
+Cerise capability machine formalization has required some
+modifications to our model to distinguish values from addresses, as
+they are _not_ the same in a capability machine, but it has little
+global effect on the model itself. Indeed, distinguishing between
+values and addresses is a higher-level abstraction that can "hide"
+potential bugs resulting from implicit conversions between them, so we
+feel it is preferable to model them as the same when possible.
 
@@
 [SNA: should we do jlr?]
+[APT: The reviewer is uncomfortable seeing a C `return` with no argument. Not sure why.
+I don't think this needs a response at all.]
 

### 2021-09-09 submission3/response.txt
Rearrannged content down through reviewer 1

--- 2021-09-09 submission3/response.txt
@@
-We first address for the most important high-level points, then go back to
+We first address the most important high-level points, then go back to
 fill in detailed responses to all points made in the reviews.
 
-_[BCP: Fill this part in!!]_
-
-___________________________________________________________________
-
-_[BCP: This could all use some more attention to markdown formatting so that
-the output looks readable.]_
-
-_[BCP: Still some comments below marked BCP/LEO...]_
-
-# Reviewer A
-
-> Paper summary
-
-> This submission formalizes the general notion of stack safety (roughly
-  speaking, every caller should be protected from its callees during program
-  execution) using well-known notions of security (integrity,
-  confidentiality and non interference), and defines several security
-  policies for protecting the stack. The formalization is validated on
-  existing (and less general) micro-policies for stack safety (depth
-  isolation, lazy tagging and lazy clearing) using property-based random
-  testing. This evaluation of former policies detected some security
-  violations in former and less precise works.
-
-> First, this submission introduces the main required security policies, in
-  terms of traces of memory, register states and stepwise stack-safety:
-  stepwise integrity (an execution step does not modify the memory slots
-  that are reserved for a caller at a given depth), confidentiality (a
-  callee is insensitive to the context of its call), stepwise
-  confidentiality, temporal stack safety, and well-bracketed control-flow
-  (calls and returns should be correctly nested). Then, the submission
-  describes the requirements on the machine model (presence of a program
-  counter, used to distinguish between the different stack-frames, and a
-  mechanism for recognizing the return from a call) and the threat
-  model. The machine model is abstract enough to represent different
-  architectures.
-
-> Next, the submission details in a progressive way the formalization of
-  machine behaviors and stack safety using semantic rules. The first
-  formalization of the execution stack is simple, with a single stack and no
-  sharing between callers and callees. Then the model is enriched with
-  function parameters passed on the stack (by value or by reference), with
-  callee-saves registers and with coroutines. Moreover, the first model is
-  implemented in Coq and this implementation reuses an existing RISCV
-  specification. The QuickChick tool is used to test the stepwise property
-  and validate an existing enforcement mechanism (called depth isolation) of
-  the literature. Furthermore, as the previous policy is conservative and
-  rather slow to be practical, more permissive variants of confidentiality
-  and integrity properties are formalized, resulting in an observational
-  property that is again validated by testing.
-
-> Strengths
-
->    well-written submission, well-explained ideas
-
->    general formalization that addresses a real challenge and that is validated by testing
-
-> Weaknesses
-
->    lack of discussion about the limitations of the work
-
->    medium delta with respect to previous works
-
-> Comments for author
-
-> This submission is well-written, with well-explained ideas.
-Examples of C and assembly programs are used throughout the submission to illustrate the various définitions.
-The formalization is provided as a supplementary material together with the test cases.
-The techniques and results of the submission are interesting and general enough to be reused in different settings.
-
-> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.
-
-*** Questions Begin ***
-
-> Why did you choose this RISCV specification?
-> Why didn't you choose to reuse an ARM specification (as there are some of
-> them written in Coq in the literature)? 
-> Would it be easy to adapt your work to another ISA specification?
-
-We chose RISCV for its simplicity, and because it has been used as the base
-ISA for much of the existing work (by others) on designing hardware that
-supports tag-based micro-policies.
-
-To clarify a possible misunderstanding: the Coq specification of RISCV that
-we use was developed by others (see footnote p. 14); we were able to adapt
-it to our testing framework with relatively little effort.
-
-Our stack safety theory itself is not dependent on the details of the
-architecture, since the properties are defined in terms of a generic
-interface over machine and enforcement policy; see the `Properties`
-directory in the Coq development as explained below for concrete
-modelling details.
-
-Porting the testing framework to another architecture such as ARM would
-require adding a test case generator for the new ISA (including its specific
-calling conventions) and translating the micropolicy rules to apply to that
-architecture / calling conventions.  This would take some work, but it
-should be straightforward.
-
-For example, we are currently working on extending the work to a model of a
-Cheri-style hardware capability machine, Cerise, and have not encountered
-major obstacles.
-
-> How realistic is the RISCV specification? In other words, are there any
-  RISCV vulnerabilities that are not taken into account in your model?
-
-We use the RISCV specification to demonstrate that our properties are
-testable in a realistic setting. We do not aim to exercise the entire RISCV
-specification, and the validity of the model is not dependent on the absence
-of (not stack-related) vulnerabilities in the RISCV implementation. That said,
-the RISCV specification we use is complete for user mode integer
-instructions (RV32IM).
-
-> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+# Limitations #
 
@@
 
-As for the distinction between values and addresses, conventional ISAs do
-not make this distinction.  Ongoing work on the Cerise capability machine
-formalization has required some modifications to our model to distinguish
-values from addresses, as they are _not_ the same in a capability machine,
-but it has little global effect on the model itself. Indeed, distinguishing
-between values and addresses is a higher-level abstraction that can "hide"
-potential bugs resulting from implicit conversions between them, so we feel
-it is preferable to model them as the same when possible.
+# Supplementary Coq Material #
 
-> Another example is the outside qualifier that characterizes only a whole
-  component. Wouldn't you need a finer-grain access control policy (and a
-  qualifier for partially outside components)?
-
-We don't entirely understand this question, but we wonder whether there may
-be some confusion arising from our use of the term "components."  In our
-model (see line 398), components are already fine-grained: individual
-registers and memory addresses. We recognize that this is a poor choice of
-term given the common usage of component to refer to a larger module in a
-compartmentalization context, and we plan to change the term to "state
-elements."  Since components/elements are so fine-grained, there is no need
-for finer access control.
-
-> Could you comment on the Coq formalization that is provided as a
-  supplementary material? Was it straightforward to write a Coq program from
-  a C++ program? Moreover, I looked at the Coq development and found some
-  admitted lemmas in it.
-
-See below.
-
-> Other comments
-
-> The supplementary material is not mentioned in the submission.
-Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').
-
-The supplementary material should indeed be mentioned in the submission.
+The supplementary material should be mentioned in the submission, and its
+purpose described.
 
@@
 
-Admitted lemmas are the beginnings of an effort to formalize [details] [BCP:
-N.b. missing][LEO: Not sure what we can say here.]
-
-In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
-last instance of n' should be n''.
-
 Map of the supplementary material:
 
-Core contains the abstract machine and policy models, as well as the RISCV
+Core/ contains the abstract machine and policy models, as well as the RISCV
 instantiations with different policies.
 
-Properties contains Coq definitions corresponding to the different Properties:
+Properties/ contains Coq definitions corresponding to the different Properties:
 - Trace.v and ObsTrace.v define the trace model and observation model
@@
   Lazy.v is an outdated file that was included by mistake.
-  [SNA: I'm not sure how strongly we should call this out, but it's pretty obvious that it doesn't match the paper.]
-  [BCP: This is fine, IMO.]
 
-Testing contains the test system, with generators for different machines and the final test run
+Testing/ contains the test system, with generators for different machines and the final test run
 via extraction in TestProperties.v
+
+___________________________________________________________________
+
+_[BCP: This could all use some more attention to markdown formatting so that
+the output looks readable.]_
+
+_[BCP: Still some comments below marked BCP/LEO...]_
+
+# Reviewer A
+
+> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.
+
+> Why did you choose this RISCV specification?
+> Why didn't you choose to reuse an ARM specification (as there are some of
+> them written in Coq in the literature)? 
+> Would it be easy to adapt your work to another ISA specification?
+
+We chose RISCV for its simplicity, and because it has been used as the base
+ISA for much of the existing work (by others) on designing hardware that
+supports tag-based micro-policies.
+
+To clarify a possible misunderstanding: the Coq specification of RISCV that
+we use was developed by others (see footnote p. 14); we were able to adapt
+it to our testing framework with relatively little effort.
+
+Our stack safety theory itself is not dependent on the details of the
+architecture, since the properties are defined in terms of a generic
+interface over machine and enforcement policy; see the `Properties`
+directory in the Coq development as explained below for concrete
+modelling details.
+
+Porting the testing framework to another architecture such as ARM would
+require adding a test case generator for the new ISA (including its specific
+calling conventions) and translating the micropolicy rules to apply to that
+architecture / calling conventions.  This would take some work, but it
+should be straightforward.
+
+For example, we are currently working on extending the work to a model of a
+Cheri-style hardware capability machine, Cerise, and have not encountered
+major obstacles.
+
+> How realistic is the RISCV specification? In other words, are there any
+> RISCV vulnerabilities that are not taken into account in your model?
+
+We use the RISCV specification to demonstrate that our properties are
+testable in a realistic setting. We do not aim to exercise the entire RISCV
+specification, and the validity of the model is not dependent on the absence
+of (not stack-related) vulnerabilities in the RISCV implementation. That said,
+the RISCV specification we use is complete for user mode integer
+instructions (RV32IM).
+
+> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+
+As for the distinction between values and addresses, conventional ISAs do
+not make this distinction.  Ongoing work on the Cerise capability machine
+formalization has required some modifications to our model to distinguish
+values from addresses, as they are _not_ the same in a capability machine,
+but it has little global effect on the model itself. Indeed, distinguishing
+between values and addresses is a higher-level abstraction that can "hide"
+potential bugs resulting from implicit conversions between them, so we feel
+it is preferable to model them as the same when possible.
+
+> Another example is the outside qualifier that characterizes only a whole
+> component. Wouldn't you need a finer-grain access control policy (and a
+> qualifier for partially outside components)?
+
+We don't entirely understand this question, but we wonder whether there may
+be some confusion arising from our use of the term "components."  In our
+model (see line 398), components are already fine-grained: individual
+registers and memory addresses. We recognize that this is a poor choice of
+term given the common usage of component to refer to a larger module in a
+compartmentalization context, and we plan to change the term to "state
+elements."  Since components/elements are so fine-grained, there is no need
+for finer access control.
+
+> Could you comment on the Coq formalization that is provided as a
+> supplementary material? Was it straightforward to write a Coq program from
+> a C++ program? Moreover, I looked at the Coq development and found some
+> admitted lemmas in it.
+
+See above for a discussion of the coq formalism in general.
+
+_[SNA: not sure what to say about writing Coq programs from C++ programs... really we wrote assembly programs anyway.]_
+
+Admitted lemmas are the beginnings of an effort to formalize the relationships
+between our various properties, especially that Stepwise properties are stronger
+than Observational.
+
+> Other comments
+
+> The supplementary material is not mentioned in the submission.
+Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').
+
+We agree, the Coq formalization should be better integrated into the submission. We give a map of
+the files above. In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
+last instance of n' should be n''.
 

### 2021-09-09 submission3/response.txt
add comment

--- 2021-09-09 submission3/response.txt
@@
 parameterizable relation and its various instantiations in different
-versions of the property. \rightarrow_C parameterized by the type ctx, which
+versions of the property. \rightarrow_C is parameterized by the type ctx, which
 is section 5 is instantiated as (K -> SD)x(list(target)). In later sections
@@
 parameterized and when they are being instantiated, so that the connections
-are more clear.
+are clearer.
 
@@
 > Line 508: "isRet" is a relation, but you are using as a function.
+
+[APT: Complaint is correct, no? Can admit to it and say we'll avoid
+this abuse of notation. Otherwise, just don't include this point
+in reply.]
 

### 2021-09-09 submission3/response.txt
add comment

--- 2021-09-09 submission3/response.txt
@@
 
-___________________________________________________________________
-
-_[BCP: This could all use some more attention to markdown formatting so that
-the output looks readable.]_
-
-_[BCP: Still some comments below marked BCP/LEO...]_
-
 # Reviewer A
-
-> I am in favor of accepting this paper; however there are a few questions that I would like to be addressed.
 
@@
 
-> More generally, what are the limitations of the presented work? For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+> More generally, what are the limitations of the presented work?
 
-As for the distinction between values and addresses, conventional ISAs do
-not make this distinction.  Ongoing work on the Cerise capability machine
-formalization has required some modifications to our model to distinguish
-values from addresses, as they are _not_ the same in a capability machine,
-but it has little global effect on the model itself. Indeed, distinguishing
-between values and addresses is a higher-level abstraction that can "hide"
-potential bugs resulting from implicit conversions between them, so we feel
-it is preferable to model them as the same when possible.
+> For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+
+Conventional ISAs do not make this distinction.  Our ongoing work on the
+Cerise capability machine formalization has required some
+modifications to our model to distinguish values from addresses, as
+they are _not_ the same in a capability machine, but it has little
+global effect on the model itself. Indeed, distinguishing between
+values and addresses is a higher-level abstraction that can "hide"
+potential bugs resulting from implicit conversions between them, so we
+feel it is preferable to model them as the same when possible.
 
@@
 
-See above for a discussion of the coq formalism in general.
+The theorems and lemmas in the formalization files serve only as
+sanity checks on our definitions. The Admitted and commented-out ones
+are mostly part of an early effort to formalize the relationships
+between our various properties, especially that Stepwise properties are stronger
+than Observational. They were included by mistake.
 
@@
 
-Admitted lemmas are the beginnings of an effort to formalize the relationships
-between our various properties, especially that Stepwise properties are stronger
-than Observational.
+In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
+last instance of n' should be n''.
 

### 2021-09-09 submission3/response.txt

@@
   insufficient for inclusion in the paper.
+  [APT: Does it differ from the (confusing) discussion at line 960?]
 

### 2021-09-09 submission3/response.txt
Further edits

@@
 
->    (310, 886,895,947) The return instruction is confusing as the return type of f is void.
-
-[SNA: should we do jlr?]
-[APT: The reviewer is uncomfortable seeing a C `return` with no argument. Not sure why.
-I don't think this needs a response at all.]
-
-> Questions for the response period
-
-> See my former questions related to 1) the Coq formalization, 2) the choice of a RISCV model and 3) the limitations of the approach
-
 ___________________________________________________________________________
 # Reviewer B
-
-> Paper summary
-
-> This paper presents a semantics for formalising safety properties of stack safety enforcement systems.
-The paper uses a map to say which parts of the stack can currently be accessed and which can't, similar to the semantics of separation logic.
-It then connects this to a notion of call and return, that can be configured to enable different machine's implementations of functions.
-
-> The paper uses a systematic testing approach with QuickChick to show that a hardware enforcement mechanism actually provides stack safety, and find bugs that are injected into the approach.
-It is great that you found the bug in lazy approach, however, it is fairly obvious type of security issue where uninitialised data leads to something escaping.
-
-> The paper discusses extension, where parameter passing can use the stack, and passing references to structures.
-
-> The paper is primarily a semantics paper, though that seems fairly straightforward.
-The more interesting aspect of the paper is the more elaborate examples in Section 6 and 7, and the more elaborate extensions in 8.
-But I had not followed the earlier aspects well enough to appreciate those.
-
-> Overall, I didn't find the paper particularly compelling in its current form. It seems complex for what it achieves earlier on, and doesn't have clear enough exposition for the later parts.
-> Strengths
-
->    Checks properties of stack safety using capabilities
->    Interesting topic
-
-> Weaknesses
-
->    Exposition doesn't connect the different aspects of the paper well.
-
->    Seems overly complex in comparison to other relational reasoning papers.
-
-> Comments for author
 
@@
 > Line 329: I was really interested in address taken at this point, and
-  whether the model could handle it. I really think the informal aspects of
-  the extensions, should be pulled forward, so these questions are answered
-  earlier
+>  whether the model could handle it. I really think the informal aspects of
+>  the extensions, should be pulled forward, so these questions are answered
+>  earlier
 
@@
 > Line 349: How does this related to control flow integrity/guard, and
-  Intel's CET?
+>  Intel's CET?
 
@@
 
-[APT: Complaint is correct, no? Can admit to it and say we'll avoid
-this abuse of notation. Otherwise, just don't include this point
-in reply.]
+Yes, we will avoid this abuse of notation.
 
@@
 
-[LEO: Sean, check this.]
-All ReturnRec does is unseal the frames between the levels before and
-after the return. Frames marked "outside" are left unaffected, while
-out frames that had been sealed by earlier calls can only be accessed
-after returning to the outer level.
+[LEO: Sean, check this.][SNA: I think they're confused about us allowing long returns.]
+ReturnRec defines a return in the case that well-bracketedness does
+not hold and a caller is returned to out of order. This has the cost of
+additional complexity, but allows our integrity and confidentiality properties
+to be orthogonal to well-bracketed control flow. It also opens the possibility
+of implementing exceptions, as described in the Limitations section above.
+So, accessing outer frames through ReturnRec is valid because if ReturnRec
+applies, there has been a return.
 

### 2021-09-09 submission3/response.txt
Another stab at the returnrec question

@@
 > Line 515: ReturnRec: This seems like it would allow you to access outer
-  frames as long as you didn't return?
+> frames as long as you didn't return?
 
-[LEO: Sean, check this.][SNA: I think they're confused about us allowing long returns.]
-ReturnRec defines a return in the case that well-bracketedness does
-not hold and a caller is returned to out of order. This has the cost of
-additional complexity, but allows our integrity and confidentiality properties
-to be orthogonal to well-bracketed control flow. It also opens the possibility
+ReturnRec allows us to handle the case that well-bracketedness is violated and
+a caller is returned to before its callee. We choose to define this as a return,
+so that integrity and confidentiality are meaningful even without 
+well-bracketed control flow. This also opens the possibility
 of implementing exceptions, as described in the Limitations section above.
-So, accessing outer frames through ReturnRec is valid because if ReturnRec
+In short, accessing outer frames through ReturnRec is valid because if ReturnRec
 applies, there has been a return.
@@
 # Reviewer C
-
-> Strengths
-
-> The work presents a compelling model for stack security, framed in terms
-  of existing notions of integrity and confidentiality.
-
-> Models are sufficiently abstract and general to be transportable to a
-  variety of mechanisms and machines.
-
-> Notions of observable stack safety capture a more idealized notion that
-  can reflect realistic violations, rather than "false positives" due to the
-  limitations of particular mechanisms. This is my favorite contribution of
-  this work.
-
-> Some evidence that the approach can be tailored to complex control
-  mechanisms, by showing how it applies to a simple fixed coroutine model.
-
-> Weaknesses
-
-> Validation is performed using random testing rather than formal
-  proof. However, testing proved sufficient to uncover shortcomings in prior
-  lazy models (which could then be fixed), and mutation testing was used to
-  ensure that the testing approach would catch errors.
-
-> Comments for author
 

### 2021-09-09 submission3/response.txt
Symbols

@@
 
-  \rightarrow_C => Line 126 (disregard comment regarding returns)
+  &#8594;<sub>C</sub> => Line 126 (disregard comment regarding returns)
   Definition 5.2 => Line 184
@@
 - CalleeSave.v defines stack safety with callee-save registers.
-  \rightarrow_C => Line 62
+  &#8594;<sub>C</sub> => Line 62
   Definition 8.6 => Line 123 & Line 130
@@
 
-  \rightarrow_C => Line 82
+  &#8594;<sub>C</sub> => Line 82
   Definition 8.3 (Stepwise only) => Line 152
@@
 - Coroutine.v defines the coroutine version. Unlike in the paper, passing on the stack is included.
-  \rightarrow_C => Line 107
+  &#8594;<sub>C</sub> => Line 107
   Definition 8.10 (Stepwise only) => Line 192 & Line 199
@@
 Seem very disconnected?
-For instance, Line 501: ←C\leftarrow_C←C​ this has a different type to the semantics in the previous section.
+For instance, Line 501: ←C\leftarrow<sub>C</sub>←C​ this has a different type to the semantics in the previous section.
 How does it relate?
@@
 parameterizable relation and its various instantiations in different
-versions of the property. \rightarrow_C is parameterized by the type ctx, which
+versions of the property. &#8594;<sub>C</sub> is parameterized by the type ctx, which
 is section 5 is instantiated as (K -> SD)x(list(target)). In later sections
-it will be instantiated by other types.  Then for any \rightarrow_C, we can
-extend \rightharpoonup to form \rightharpoonup_C.
-[BCP: The formatting here makes it hard to read.  Can we improve it?]
+it will be instantiated by other types.  Then for any &#8594;<sub>C</sub>, we can
+extend &#8640;  to form &#8640;<sub>C</sub>.
 

### 2021-09-09 submission3/response.txt
some minor fixes

--- 2021-09-09 submission3/response.txt
@@
+We thank the reviewers for their comments and questions.
 We first address the most important high-level points, then go back to
@@
 Our focus is on characterizing stack safety.  So we are particularly
-concerned with limitations that prevent the model from applying to realistic
-uses of the stack, as opposed to modeling other structures such as the heap,
-which we intentionally say nothing about.  (Heap safety has already been
-studied, in isolation, by Azevedo et al, and we expect that their
-correctness condition and ours can be combined cleanly to characterize
-correct protection of stack + heap.)
+concerned with limitations that prevent the model from applying to
+realistic uses of the stack, as opposed to modeling other structures
+such as the heap, which we intentionally say nothing about.  (Heap
+safety has already been studied, in isolation, by [Azevedo et al 2018],
+and we expect that their correctness condition and ours can be
+combined cleanly to characterize correct protection of stack + heap.)
 
@@
 
-The supplementary material should be mentioned in the submission, and its
-purpose described.
+We should have referenced the supplementary material from the submission, and
+described its purpose.
 
@@
 
-Map of the supplementary material:
+Here is a brief map of the supplementary material:
 
@@
   insufficient for inclusion in the paper.
-  [APT: Does it differ from the (confusing) discussion at line 960?]
+_[APT: Does it differ from the (confusing) discussion at line 960?]_
 
@@
 
+_[APT: Omit, or give back pointer.]_
+
 > For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
@@
 
+_[APT: Include backward pointer re formalization.]_
+
 The theorems and lemmas in the formalization files serve only as
@@
 between our various properties, especially that Stepwise properties are stronger
-than Observational. They were included by mistake.
+than Observational ones. They were included by mistake.
+_[APT: Maybe this should move up to the top-level discussion of the formalization? Not sure.]_
 
 _[SNA: not sure what to say about writing Coq programs from C++ programs... really we wrote assembly programs anyway.]_
-
-In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
-last instance of n' should be n''.
+_[APT: here's a stab.]_
+We don't quite understand this question; none of the Coq code was derived from C++.
 
@@
 
-We agree, the Coq formalization should be better integrated into the submission. We give a map of
-the files above. In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
+We agree, the Coq formalization should be better integrated into the submission, as described at the head of
+this response. In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
 last instance of n' should be n''.
@@
 
+>    (310, 886,895,947) The return instruction is confusing as the return type of f is void.
+
+[SNA: should we do jlr?]
+[APT: The reviewer is uncomfortable seeing a C `return` with no argument. Not sure why.
+I don't think this needs a response at all.]
+
+> Questions for the response period
+
+> See my former questions related to 1) the Coq formalization, 2) the choice of a RISCV model and 3) the limitations of the approach
+
 ___________________________________________________________________________
 # Reviewer B
+
+> Paper summary
+
+> This paper presents a semantics for formalising safety properties of stack safety enforcement systems.
+The paper uses a map to say which parts of the stack can currently be accessed and which can't, similar to the semantics of separation logic.
+It then connects this to a notion of call and return, that can be configured to enable different machine's implementations of functions.
+
+> The paper uses a systematic testing approach with QuickChick to show that a hardware enforcement mechanism actually provides stack safety, and find bugs that are injected into the approach.
+It is great that you found the bug in lazy approach, however, it is fairly obvious type of security issue where uninitialised data leads to something escaping.
+
+> The paper discusses extension, where parameter passing can use the stack, and passing references to structures.
+
+> The paper is primarily a semantics paper, though that seems fairly straightforward.
+The more interesting aspect of the paper is the more elaborate examples in Section 6 and 7, and the more elaborate extensions in 8.
+But I had not followed the earlier aspects well enough to appreciate those.
+
+> Overall, I didn't find the paper particularly compelling in its current form. It seems complex for what it achieves earlier on, and doesn't have clear enough exposition for the later parts.
+> Strengths
+
+>    Checks properties of stack safety using capabilities
+>    Interesting topic
+
+> Weaknesses
+
+>    Exposition doesn't connect the different aspects of the paper well.
+
+>    Seems overly complex in comparison to other relational reasoning papers.
+
+> Comments for author
 
@@
 > Line 329: I was really interested in address taken at this point, and
->  whether the model could handle it. I really think the informal aspects of
->  the extensions, should be pulled forward, so these questions are answered
->  earlier
+  whether the model could handle it. I really think the informal aspects of
+  the extensions, should be pulled forward, so these questions are answered
+  earlier
 
@@
 > Line 349: How does this related to control flow integrity/guard, and
->  Intel's CET?
+  Intel's CET?
 
@@
 
-Yes, we will avoid this abuse of notation.
+[APT: Complaint is correct, no? Can admit to it and say we'll avoid
+this abuse of notation. Otherwise, just don't include this point
+in reply.]
 
@@
 
-[LEO: Sean, check this.][SNA: I think they're confused about us allowing long returns.]
-ReturnRec defines a return in the case that well-bracketedness does
-not hold and a caller is returned to out of order. This has the cost of
-additional complexity, but allows our integrity and confidentiality properties
-to be orthogonal to well-bracketed control flow. It also opens the possibility
-of implementing exceptions, as described in the Limitations section above.
-So, accessing outer frames through ReturnRec is valid because if ReturnRec
-applies, there has been a return.
+[LEO: Sean, check this.]
+All ReturnRec does is unseal the frames between the levels before and
+after the return. Frames marked "outside" are left unaffected, while
+out frames that had been sealed by earlier calls can only be accessed
+after returning to the outer level.
 

### 2021-09-09 submission3/response.txt
Small tweaks:wq :wq

@@
 - SubroutineShare.v defines the version enhanced with argument passing on the stack.
-  Note that a preliminary treatment of address-taken locals is present; we found this
-  insufficient for inclusion in the paper.
-_[APT: Does it differ from the (confusing) discussion at line 960?]_
+  Note that a preliminary treatment of address-taken locals is present; we discuss this briefly
+  in the submission at line 960, but find it uninteresting in a model that lacks a heap.
 
@@
 
-_[APT: Omit, or give back pointer.]_
+We discuss limitations broadly in our top-level response.
 
@@
 > Could you comment on the Coq formalization that is provided as a
->  supplementary material? Was it straightforward to write a Coq program from
->  a C++ program? Moreover, I looked at the Coq development and found some
+>  supplementary material? Moreover, I looked at the Coq development and found some
 >  admitted lemmas in it.
 
-_[APT: Include backward pointer re formalization.]_
-
+We have discussed the Coq formalization broadly in our general response.
 The theorems and lemmas in the formalization files serve only as
@@
 _[APT: Maybe this should move up to the top-level discussion of the formalization? Not sure.]_
+_[SNA: We probably wouldn't bring them up if not asked explicitly.]_
 
-_[SNA: not sure what to say about writing Coq programs from C++ programs... really we wrote assembly programs anyway.]_
-_[APT: here's a stab.]_
-We don't quite understand this question; none of the Coq code was derived from C++.
+> Was it straightforward to write a Coq program from a C++ program?
+
+_[SNA: I think this is what they meant.]_
+If this question is referring to the example RISCV programs in the Coq
+development, they do not necessarily correspond to C++ programs. They served as
+intended as early test-cases for the property checker. More broadly, we do not
+generate specifically C or C++ code for the testing, but arbitrary low-level code
+that contains C-like call and return conventions.
 

### 2021-09-09 submission3/response.txt
Removed comment

@@
 
-[BCP: ???]  [APT: made a stab; others please check.]
 The identification of call points, as described in this paragraph, is

### 2021-09-09 submission3/response.txt
a word
clean comments; reformatting

@@
 We first address the most important high-level points, then go back to
-fill in detailed responses to all points made in the reviews.
+fill in detailed responses to all questions posed in the reviews.
 
@@
 
-> For example, in the model, there seems to be no distinction between values and addresses; is it a limitation?
+> For example, in the model, there seems to be no distinction between
+> values and addresses; is it a limitation?
 
@@
 than Observational ones. They were included by mistake.
-_[APT: Maybe this should move up to the top-level discussion of the formalization? Not sure.]_
-_[SNA: We probably wouldn't bring them up if not asked explicitly.]_
 
@@
 
-_[SNA: I think this is what they meant.]_
 If this question is referring to the example RISCV programs in the Coq
@@
 
-> The supplementary material is not mentioned in the submission.
-Links to the Coq formalization would have been useful (e.g., for definition 7.4 that I did not fully understand in the submission, as K' is defined but not used, and so is n'').
+> The supplementary material is not mentioned in the submission. Links
+> to the Coq formalization would have been useful (e.g., for
+> definition 7.4 that I did not fully understand in the submission, as
+> K' is defined but not used, and so is n'').
 
-We agree, the Coq formalization should be better integrated into the submission, as described at the head of
-this response. In the case of definition 7.4, this is a typo. The K in the third bullet should be K', and the
-last instance of n' should be n''.
+We agree, the Coq formalization should be better integrated into the
+submission, as described at the head of this response. In the case of
+definition 7.4, this is a typo. The K in the third bullet should be
+K', and the last instance of n' should be n''.
 
-> The first example programs (Fig.
-1 to 4) are C programs.
-The programs in Fig.
-5 and 6 are qualified as C++ programs, but they could be C programs and they are close to the first programs.
-Moreover, the C syntax for passing parameters by reference is surprising.
-I would have written f(&x)(line 934) and void f(int *a)(line 944).
+> The first example programs (Fig. 1 to 4) are C programs. The
+> programs in Fig. 5 and 6 are qualified as C++ programs, but they
+> could be C programs and they are close to the first
+> programs. Moreover, the C syntax for passing parameters by reference
+> is surprising. I would have written f(&x)(line 934) and void f(int > *a)
+> (line 944).
 
@@
 
-> I found the paper did not connect the section well.
-Section 4 presents a lot of details for a framework, which is then used in Section 5.
-Seem very disconnected?
-For instance, Line 501: ←C\leftarrow<sub>C</sub>←C​ this has a different type to the semantics in the previous section.
-How does it relate?
-Generally, there are a lot of things that are similar to the previous section, but you are not actually connecting them.
-Is this because they are different or just missing exposition.
+> I found the paper did not connect the section well. Section 4
+> presents a lot of details for a framework, which is then used in
+> Section 5. Seem very disconnected? For instance, Line 501:
+> ←C\leftarrow<sub>C</sub>←C​ this has a different type to the
+> semantics in the previous section. How does it relate? Generally,
+> there are a lot of things that are similar to the previous section,
+> but you are not actually connecting them. Is this because they are
+> different or just missing exposition.
 
@@
 parameterizable relation and its various instantiations in different
-versions of the property. &#8594;<sub>C</sub> is parameterized by the type ctx, which
-is section 5 is instantiated as (K -> SD)x(list(target)). In later sections
-it will be instantiated by other types.  Then for any &#8594;<sub>C</sub>, we can
-extend &#8640;  to form &#8640;<sub>C</sub>.
+versions of the property. &#8594;<sub>C</sub> is parameterized by the
+type ctx, which is section 5 is instantiated as (K -> SD)x(list(target)).
+In later sections it will be instantiated by other
+types.  Then for any &#8594;<sub>C</sub>, we can extend &#8640; to
+form &#8640;<sub>C</sub>.
 
 We will do more to call out which parts of the formalization are
-parameterized and when they are being instantiated, so that the connections
-are clearer.
+parameterized and when they are being instantiated, so that the
+connections are clearer.
 
-> Line 329: I was really interested in address taken at this point, and
->  whether the model could handle it. I really think the informal aspects of
->  the extensions, should be pulled forward, so these questions are answered
->  earlier
+> Line 329: I was really interested in address taken at this point,
+>  and whether the model could handle it. I really think the informal
+>  aspects of the extensions, should be pulled forward, so these
+>  questions are answered earlier
 
@@
 
-Yes, we will avoid this abuse of notation.
+Yes; we will avoid this abuse of notation.
 
@@
 
-ReturnRec allows us to handle the case that well-bracketedness is violated and
-a caller is returned to before its callee. We choose to define this as a return,
-so that integrity and confidentiality are meaningful even without 
-well-bracketed control flow. This also opens the possibility
-of implementing exceptions, as described in the Limitations section above.
-In short, accessing outer frames through ReturnRec is valid because if ReturnRec
-applies, there has been a return.
+ReturnRec allows us to handle the case that well-bracketedness is
+violated and a caller is returned to before its callee. We choose to
+define this as a return, so that integrity and confidentiality are
+meaningful even without well-bracketed control flow. This also opens
+the possibility of implementing exceptions, as described in the
+Limitations section above.  In short, accessing outer frames through
+ReturnRec is valid because if ReturnRec applies, there has been a
+return.
 

### 2021-09-09 submission3/response.txt
Moved coq map
Shrunk suggestions in limitations

@@
 - Exceptions. We can see a way to extend the model to include them:
-  essentially, we would model a "setjump" operation by inserting a
+  we would model a "setjump" operation by inserting a
   special target in the return target stack. We would weaken
@@
 
-- Tail calls. Again, we can see how we might extend the model for them
-  Since a tail call is both a call and a return in a single step, our
-  normal handling of returns doesn't work -- it would interpret the
-  tail call as a call to and immediate return from a new function
-  activation. We would need to distinguish tail calls from normal
-  calls and essentially process the call and the return in the reverse
-  order.
+- Tail calls. To implement these, we would need to distinguish them
+  from normal calls and extend &#8594;<sub>C</sub> to process them
+  as a simultaneous return and call.
 
-- Dynamic code generation. This would be rather challenging to
-  integrate into our framework, because we rely on annotations in the
-  code. We can easily envision supporting dynamic code generation
-  where the generated code is guaranteed never to contain a call
-  or other annotated instruction. We could also support annotated-code
-  generation in which existing code is copied along with its annotations.
+- Dynamic code generation. This would be challenging to
+  integrate into our framework, because we rely on code annotations.
+  But we can easily envision supporting limited cases in which
+  the generated code never contains a call or other annotated instruction,
+  or in which existing code is copied along with its annotations.
 
@@
 validate our properties by showing that computable versions of them are
-testable.
-
-Here is a brief map of the supplementary material:
-
-Core/ contains the abstract machine and policy models, as well as the RISCV
-instantiations with different policies.
-
-Properties/ contains Coq definitions corresponding to the different Properties:
-- Trace.v and ObsTrace.v define the trace model and observation model
-
-- SubroutineSimple.v defines the basic stack safety properties.
-  Some comments are outdated, and will be updated or removed for the final
-  submission.
-
-  &#8594;<sub>C</sub> => Line 126 (disregard comment regarding returns)
-  Definition 5.2 => Line 184
-  Definition 5.9 => Line 255
-  Definition 5.10 => Line 277
-
-- CalleeSave.v defines stack safety with callee-save registers.
-  &#8594;<sub>C</sub> => Line 62
-  Definition 8.6 => Line 123 & Line 130
-  These definitions refer to general integrity and confidentiality properties for arbitrary
-  traces defined in TraceProperties.v at Lines 51 & 60.
-
-- SubroutineShare.v defines the version enhanced with argument passing on the stack.
-  Note that a preliminary treatment of address-taken locals is present; we discuss this briefly
-  in the submission at line 960, but find it uninteresting in a model that lacks a heap.
-
-  &#8594;<sub>C</sub> => Line 82
-  Definition 8.3 (Stepwise only) => Line 152
-  Note that this definition refers to a general confidentiality property of traces defined in 
-  TraceProperties.v at Line 35.
-
-- Coroutine.v defines the coroutine version. Unlike in the paper, passing on the stack is included.
-  &#8594;<sub>C</sub> => Line 107
-  Definition 8.10 (Stepwise only) => Line 192 & Line 199
-  Definition 8.11 (Stepwise only) => Line 212 & Line 221
-  This also refers to the general properties presented in TraceProperties.v
-
-  Lazy.v is an outdated file that was included by mistake.
-
-Testing/ contains the test system, with generators for different machines and the final test run
-via extraction in TestProperties.v
+testable. The contents are laid out in detail at the end of our response,
+with definitions mapped from the paper to the coq where relevant.
 
@@
 along with a description of what it would take to implement them.
+
+# Map of Supplementary Material
+
+Here is a brief map of the supplementary material:
+
+Core/ contains the abstract machine and policy models, as well as the RISCV
+instantiations with different policies.
+
+Properties/ contains Coq definitions corresponding to the different Properties:
+- Trace.v and ObsTrace.v define the trace model and observation model
+
+- SubroutineSimple.v defines the basic stack safety properties.
+  Some comments are outdated, and will be updated or removed for the final
+  submission.
+
+  &#8594;<sub>C</sub> => Line 126 (disregard comment regarding returns)
+  Definition 5.2 => Line 184
+  Definition 5.9 => Line 255
+  Definition 5.10 => Line 277
+
+- CalleeSave.v defines stack safety with callee-save registers.
+  &#8594;<sub>C</sub> => Line 62
+  Definition 8.6 => Line 123 & Line 130
+  These definitions refer to general integrity and confidentiality properties for arbitrary
+  traces defined in TraceProperties.v at Lines 51 & 60.
+
+- SubroutineShare.v defines the version enhanced with argument passing on the stack.
+  Note that a preliminary treatment of address-taken locals is present; we discuss this briefly
+  in the submission at line 960, but find it uninteresting in a model that lacks a heap.
+
+  &#8594;<sub>C</sub> => Line 82
+  Definition 8.3 (Stepwise only) => Line 152
+  Note that this definition refers to a general confidentiality property of traces defined in 
+  TraceProperties.v at Line 35.
+
+- Coroutine.v defines the coroutine version. Unlike in the paper, passing on the stack is included.
+  &#8594;<sub>C</sub> => Line 107
+  Definition 8.10 (Stepwise only) => Line 192 & Line 199
+  Definition 8.11 (Stepwise only) => Line 212 & Line 221
+  This also refers to the general properties presented in TraceProperties.v
+
+  Lazy.v is an outdated file that was included by mistake.
+
+Testing/ contains the test system, with generators for different machines and the final test run
+via extraction in TestProperties.v

### 2021-09-12 submission4/main.tex
Added limitations subsection and started submission4

--- 2021-09-08 submission3/main.tex
@@
 
-Other enhancements, such as runtime code generation and true concurrency,
-would be much more challenging to support, and we consider these features
-out of scope for now.
-
 \paragraph*{What is a ``call''?}  One significant challenge in formulating our
@@
 
-%, returns, and function entry points\BCP{I
-%  still want to talk about whether returns and entry points need to be
-%  annotated.  To me, these annotations do not make intuitive sense: it is
-%  the caller's responsibility to know whether the address that it is calling
-%is really an entry point, and we can trust it to check this.  Similarly, it
-%is not the callee's role to say when it is returning; it is the caller's
-%prerogative to decide when it has been returned to!  I.e., from the caller's
-%point of view, it has been returned to when control resumes at the
-%instruction following the call with the SP reset to where it was at the
-%point of the call. (This is the point where it wants to be able to access
-%its own local state again and expected to find it unmolested.)}.
-
 \paragraph*{Enforcement Model}
@@
 to protect both stack and heap.
+
+\paragraph{Limitations}
+
+Our model is flexible enough to support many realistic uses of the stack.
+We intentionall say little or nothing about other structures
+such as the heap.  (Heap safety has already been studied, in isolation,
+by [Azevedo et al 2018], and we expect that their correctness condition and ours can be
+combined cleanly to characterize correct protection of stack + heap.)
+
+Features that we do not currently support:
+
+\begin{itemize}
+\item Address-taken locals. A satisfactory
+  treatment of these seems to require integration with a more general
+  model of memory safety that covers the heap and static globals too,
+  such as the aforementioned model of Azevedo et al.
+
+\item Exceptions. We can see a way to extend the model to include them:
+  we would model a "setjump" operation by inserting a
+  special target in the return target stack. We would weaken
+  well-bracketedness to allow reaching such a target from any
+  depth. Since integrity and confidentiality are distinct from
+  well-bracketedness and already permit the popping of multiple
+  functions from the stack, they can simply treat the exception's
+  target as another return target.
+
+\item Tail calls. To implement these, we would need to distinguish them
+  from normal calls and slightly adjust the mechanics of the return stack.
+
+\item Dynamic code generation. This would be challenging to
+  integrate into our framework, because we rely on code annotations.
+  But we can easily envision supporting limited cases in which
+  the generated code never contains a call or other annotated instruction,
+  or in which existing code is copied along with its annotations.
+
+\item Concurrency. We can handle a limited form of coroutines, as described in
+  8.3. Handling full-blown preemptive concurrency would require significant
+  changes in our underlying machine model.
+\end{itemize}
 

### 2021-09-13 submission3/response.txt
add a couple of divider lines
Last-minute changes

@@
 - Tail calls. To implement these, we would need to distinguish them
-  from normal calls and extend &#8594;<sub>C</sub> to process them
+  from normal calls and extend $\rightarrow_C$ to process them
   as a simultaneous return and call.
@@
 parameterizable relation and its various instantiations in different
-versions of the property. &#8594;<sub>C</sub> is parameterized by the
+versions of the property. $\rightarrow_C$ is parameterized by the
 type ctx, which is section 5 is instantiated as (K -> SD)x(list(target)).
 In later sections it will be instantiated by other
-types.  Then for any &#8594;<sub>C</sub>, we can extend &#8640; to
-form &#8640;<sub>C</sub>.
+types.  Then for any $\rightarrow_C$, we can extend $\rightharpoonup$; to
+form $\rightharpoonup_C$.
 
@@
 
-  &#8594;<sub>C</sub> => Line 126 (disregard comment regarding returns)
+  $\rightarrow_C$ => Line 126 (disregard comment regarding returns)
   Definition 5.2 => Line 184
@@
 - CalleeSave.v defines stack safety with callee-save registers.
-  &#8594;<sub>C</sub> => Line 62
+  $\rightarrow_C$ => Line 62
   Definition 8.6 => Line 123 & Line 130
@@
 
-  &#8594;<sub>C</sub> => Line 82
+  $\rightarrow_C$ => Line 82
   Definition 8.3 (Stepwise only) => Line 152
@@
 - Coroutine.v defines the coroutine version. Unlike in the paper, passing on the stack is included.
-  &#8594;<sub>C</sub> => Line 107
+  $\rightarrow_C$ => Line 107
   Definition 8.10 (Stepwise only) => Line 192 & Line 199

### 2021-09-27 submission4/main.tex
Initial switch to IEEEtran and cleanup

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-%\acmJournal{PACMPL}
-%\acmVolume{1}
-%\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
-%\acmArticle{1}
-%\acmYear{2021}
-%\acmMonth{1}
-%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-%\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
+\documentclass[12pt,conference]{ieeetran}%\settopmatter{printfolios=true,printccs=false,printacmref=false}
+
 \usepackage{booktabs}   %% For formal tables:
@@
                         %% http://ctan.org/pkg/subcaption
+\usepackage{amsmath,amsthm,amssymb}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
 
@@
 %% Title information
-\title{Formalizing Stack Safety as a Security Property}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%\titlenote{with title note}             %% \titlenote is optional;
-%                                        %% can be repeated if necessary;
-%                                        %% contents suppressed with 'anonymous'
-%\subtitle{Subtitle}                     %% \subtitle is optional
-%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-%                                        %% can be repeated if necessary;
-%                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{Sean Noble Anderson}
-%\authornote{with author1 note}          %% \authornote is optional;
-%                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-%  \position{Position1}
-%  \department{Computer Science}              %% \department is recommended
-  \institution{Portland State University}            %% \institution is required
-%  \streetaddress{Street1 Address1}
-%  \city{City1}
-%  \state{State1}
-%  \postcode{Post-Code1}
-%  \country{Country1}                    %% \country is recommended
-}
-\email{ander28@pdx.edu}          %% \email is recommended
-
-\author{Leonidas Lampropoulos}
-\affiliation{
-  \institution{University of Maryland, College Park}
-}
-\email{leonidas@umd.edu}
-
-\author{Roberto Blanco}
-\affiliation{
-  \institution{Max Planck Institute for Security and Privacy}
-}
-\email{roberto.blanco@mpi-sp.org}
-
-\author{Benjamin C. Pierce}
-\affiliation{
-  \institution{University of Pennsylvania}
-}
-\email{bcpierce@cis.upenn.edu}
-
-\author{Andrew Tolmach}
-\affiliation{
-  \institution{Portland State University}
-}
-\email{tolmach@pdx.edu}
-
-%
-%%% Author with two affiliations and emails.
-%\author{First2 Last2}
-%\authornote{with author2 note}          %% \authornote is optional;
-%                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-%\affiliation{
-%  \position{Position2a}
-%  \department{Department2a}             %% \department is recommended
-%  \institution{Institution2a}           %% \institution is required
-%  \streetaddress{Street2a Address2a}
-%  \city{City2a}
-%  \state{State2a}
-%  \postcode{Post-Code2a}
-%  \country{Country2a}                   %% \country is recommended
+\title{Formalizing Stack Safety as a Security Property}
+
+%\author{
+%  \IEEEauthorblockN{
+%    Sean Noble Anderson
+%  }
+%  \IEEEauthorblockA{
+%    Portland State University\\
+%    ander28@pdx.edu
+%  }
+%  \and
+%  \IEEEauthorblockN{
+%    Leonidas Lampropoulos,
+%  }
+%  \IEEEauthorblockA{
+%    University of Maryland, College Park\\
+%    leonidas@umd.edu
+%  }
+%  \and
+%  \IEEEauthorblockN{
+%    Roberto Blanco,
+%  }
+%  \IEEEauthorblockA{
+%    Max Planck Institute for Security and Privacy\\
+%    roberto.blanco@mpi-sp.org
+%  }
+%  \and
+%  \IEEEauthorblockN{
+%    Benjamin C. Pierce,
+%  }
+%  \IEEEauthorblockA{
+%    University of Pennsylvania\\
+%    bcpierce@cis.upenn.edu
+%  }
+%  \and
+%  \IEEEauthorblockN{
+%    Andrew Tolmach
+%  }
+%  \IEEEauthorblockA{
+%    Portland State University\\
+%    tolmach@pdx.edu
+%  }
 %}
-%\email{first2.last2@inst2a.com}         %% \email is recommended
-%\affiliation{
-%  \position{Position2b}
-%  \department{Department2b}             %% \department is recommended
-%  \institution{Institution2b}           %% \institution is required
-%  \streetaddress{Street3b Address2b}
-%  \city{City2b}
-%  \state{State2b}
-%  \postcode{Post-Code2b}
-%  \country{Country2b}                   %% \country is recommended
-%}
-%\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
+
+%% Keywords
+%% comma separated list
+\ifcameraready
+\keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission
+\fi
+
+\maketitle
+
 \begin{abstract}
@@
 We use these properties to validate the stack-safety {\em micro-policies}
-proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
-property-based random testing) that Roessler and DeHon's ``eager''
+proposed by Roessler and DeHon%~\cite{DBLP:conf/sp/RoesslerD18}.
+Specifically, we check (with
+property-based random testing) that their ``eager''
 micro-policy, which catches violations as early as possible, enforces a
@@
 
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\ifcameraready
-\keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission
-\fi
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
 \newcommand{\paragraphx}[1]{\emph{#1.}}
@@
 %
-including stack canaries~\citep{Cowan+98},
-bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
-split stacks~\citep{Kuznetsov+14},
-shadow stacks~\citep{Dang+15,Shanbhogue+19},
-capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
+including stack canaries~\cite{Cowan+98},
+bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
+split stacks~\cite{Kuznetsov+14},
+shadow stacks~\cite{Dang+15,Shanbhogue+19},
+capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
   nick; there could be more}\bcp{Yes, going back to MIT days---we should
@@
 We propose such a characterization using the tools of language-based
-security~\citep{sabelfeld2003language}. The informal claim that
+security~\cite{sabelfeld2003language}. The informal claim that
 ``stack safety protects a caller
@@
 We can also, optionally, include a notion of {\em well-bracketed control
-flow (WBCF)} as in \citet{SkorstengaardSTKJFP}---a global requirement that
+flow (WBCF)} as in Skorstengaard et al. \cite{SkorstengaardSTKJFP}---a global requirement that
 callees must always return to their immediate caller, if they return
@@
 properties to validate and improve an existing enforcement mechanism, the
-{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
+{\em stack-safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
 in the Coq proof assistant on top of a RISC-V specification.  We
-use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
+use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
 tool for Coq, to generate random programs and check
@@
 property is actually enforced by the {\em Depth Isolation} policy
-of~\citet{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
-execution techniques of \citet{TestingNI:ICFP} to handle nested
+of Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
+execution techniques of Hri\c{t}cu et al.\cite{TestingNI:ICFP} to handle nested
 properties (\cref{sec:enforcement}).
@@
 
-\begin{figure}
+\begin{figure*}
 
@@
 \label{fig:int1}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{subfigure}[t]{.25\textwidth}
@@
 \label{fig:conf1}
-\end{figure}
+\end{figure*}
 
@@
 variables).
-We state this as a form of {\em noninterference}~\citep{Goguen1982SecurityPA}.
+We state this as a form of {\em noninterference}~\cite{Goguen1982SecurityPA}.
 Noninterference describes knowledge: it states that for system components
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
@@
 \label{fig:conf2}
-\end{figure}
+\end{figure*}
 
@@
 \paragraph*{Control-Flow Attacks}
-\begin{figure}
+\begin{figure*}
   \centering
@@
   \label{fig:controlflow}
-\end{figure}
+\end{figure*}
 
@@
 condition that calls and returns should be correctly nested---e.g., $C$ should
-only return to $B$.  This is what \citet{SkorstengaardSTKJFP}
+only return to $B$.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
 call \emph{well-bracketed control flow (WBCF)}.
@@
 they are set up properly for the enforcement mechanism in question.
-When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
+When validating Roessler and DeHon's \cite{DBLP:conf/sp/RoesslerD18}'s micro-policies,
 well-behaved calls and returns are implemented by fixed sequences of instructions
@@
 are attacks on the heap---though we believe that the work of
-\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
+Azevedo de Amorim et al.\cite{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
 to protect both stack and heap.
@@
 such as the heap.  (Heap safety has already been studied, in isolation,
-by [Azevedo et al 2018], and we expect that their correctness condition and ours can be
+by Azevedo de Amorim et al., and we expect that their correctness condition and ours can be
 combined cleanly to characterize correct protection of stack + heap.)
@@
   model of memory safety that covers the heap and static globals too,
-  such as the aforementioned model of Azevedo et al.
+  such as the aforementioned model of Azevedo de Amorim et al.
 
@@
 
-\paragraph*{Values and States}
-
 The building blocks of the machine are {\em values} and {\em addresses}.
@@
 A {\em machine state} is a map from components to values.
-%
-    \[\component \in \COMPONENTS ::= \WORDS + \REGS  \hspace{1in}
-    \mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-%
+
+\begin{minipage}{.45\columnwidth}
+  \[\component \in \COMPONENTS ::= \WORDS + \REGS\]
+\vspace{\afterdisplayskip}
+\end{minipage}
+\begin{minipage}{.45\columnwidth}
+  \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
+\vspace{\afterdisplayskip}
+\end{minipage}
+
 We have a total step function between machine states, written \(\mach \stepsto \mach'\).
@@
 \begin{center}
+\vspace*{-\medskipamount}
 \judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
@@
 \begin{center}
-\begin{minipage}{.38\textwidth}
-  \judgmenttwo[ Now]
+\vspace*{-\medskipamount}
+  \judgmenttwo[]
               {\(f ~ \mpcstatename\)}
@@
               {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
-\end{minipage}
-%
-\begin{minipage}{.58\textwidth}
-  \judgmentthree[ Skip]
+\vspace*{-\bigskipamount}
+  \judgmentthree[]
                 {\(f ~ \mpcstatename\)}
@@
                 {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{minipage}
-%
-  \judgmentthree[ Search]
+\vspace*{-\bigskipamount}
+  \judgmentthree[]
                 {\(\neg f ~ \mpcstatename\)}
@@
 
-\vspace*{-\bigskipamount}
+\vspace*{-\medskipamount}
 
@@
 Finally, we formalize a version of WBCF, which captures the same informal idea
-as that of \citet{SkorstengaardSTKJFP} using our context machinery.
+as that of Skorstengaard et al.\cite{SkorstengaardSTKJFP} using our context machinery.
 \definition
@@
 In this section we explore how an existing enforcement mechanism called {\em
-  Depth Isolation}~\citep{DBLP:conf/sp/RoesslerD18} implements
+  Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements
 the formal stack-safety property described in the previous
@@
 %
-The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
+The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
 on a programmable, tag-based reference monitor that runs alongside the
@@
 allowing fine-grained manipulation of metadata tags to encode so-called
-\emph{micro-policies}~\citep{pump_oakland2015}.
+\emph{micro-policies}~\cite{pump_oakland2015}.
 Every value in
@@
 the machine fail-stops with a policy violation.
-Existing work \citep{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
+Existing work \cite{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
 %\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
@@
 Efficient execution of these micro-policies relies on hardware acceleration,
-such as the PUMP architecture~\citep{pump:asplos2015}.
+such as the PUMP architecture~\cite{pump:asplos2015}.
 The hardware incorporates a rule cache
@@
 on the {\em Depth Isolation} policy presented by
-\citet{DBLP:conf/sp/RoesslerD18}. In their policies,  the stack memory is tagged
+Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}. In their policies,  the stack memory is tagged
 with ownership information associated to each stack frame (and to each
@@
   definitions are well-formed, and reason about them, and to leverage the power of the
-  QuickChick property-based testing framework~\citep{Pierce:SF4}.
+  QuickChick property-based testing framework~\cite{Pierce:SF4}.
 
@@
 To write such generators we build on the work of
-\citet{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
+Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
 introduced {\em generation by execution} to produce progams that lead
@@
 To ensure the effectiveness of testing against our formal properties, we
-use {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
+use {\em mutation testing}~\cite{JiaH11}. In mutation testing, we inject errors
 (mutations) in a program that should cause the property of interest (here,
@@
 testably enforcing stack safety, is rather slow in
-some important practical cases. \citet{DBLP:conf/sp/RoesslerD18} observe
+some important practical cases. Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18} observe
 that most of the performance overhead incurred
@@
 The Lazy Tagging and Clearing
-micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
+micro-policy from Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
 at which it was written, and enforces that it must be read from the
@@
 (A related mechanism was explored in the
-Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
+Static Authorities policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}, which
 associates a unique \emph{static} identifier to each function, which is shared
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
@@
 \label{fig:passingsimple}
-\end{figure}
-
-\begin{figure}
+\end{figure*}
+
+\begin{figure*}
   \centering
@@
 \label{fig:passing}
-\end{figure}
+\end{figure*}
 
@@
 a positive and formal characterization of stack safety
-is that of \citet{SkorstengaardSTKJFP}. Their
+is that of Skorstengaard et al. \cite{SkorstengaardSTKJFP}. Their
 StkTokens system enforces stack safety on hardware equipped with
-memory capabilities, similar to those used in the Cheri system~\citep{Woodruff+14,Chisnall+15}, but
+memory capabilities, similar to those used in the Cheri system~\cite{Woodruff+14,Chisnall+15}, but
 with additional support for linear capabilities.
@@
 
-We improve upon \citet{SkorstengaardSTKJFP} in two significant ways:
+We improve upon Skorstengaard et al. \cite{SkorstengaardSTKJFP} in two significant ways:
 \begin{enumerate}[label=(\roman*)]
@@
   every} caller should be protected from its callees.  By contrast,
-  \citet{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
+  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
   that are not protected, because they are capable of dynamic behavior that can compromise
@@
   inputs, behaves unreasonably, but under most circumstances can be
-  considered reasonable. \citet{SkorstengaardSTKJFP} would not consider this a
+  considered reasonable. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
   {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
@@
   well-behaved activations of the buggy function would be protected.
-  \citet{SkorstengaardSTKJFP} support mutually distrustful
+  Skorstengaard et al. \cite{SkorstengaardSTKJFP} support mutually distrustful
   components; one can think of our hypothetical
@@
 
-\item \citet{SkorstengaardSTKJFP}'s approach to proving safety via a fully
+\item Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s approach to proving safety via a fully
   abstract overlay
@@
 
-\citet{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
+Skorstengaard et al. \cite{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
 candidate stack safety properties.
@@
 %
-\citet{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
 attacks against stack data are in scope (but not side channels or hardware
@@
 the language of
-noninterference~\citep{DBLP:conf/post/AmorimHP18}. Their paper was a significant inspiration
+noninterference~\cite{DBLP:conf/post/AmorimHP18}. Their paper was a significant inspiration
 for the present work, and we believe that the two safety conditions should
@@
 including mechanisms that use the standard Cheri hardware (which includes local
-capabilities) \citep{SkorstengaardLocal},
+capabilities) \cite{SkorstengaardLocal},
 and others that propose entirely new types of capabilities, such as linear
-\citep{SkorstengaardSTK}, uninitialized \citep{Georges+21}, lifetime
-\citep{Tsampas+19}, and monotonic \citep{GeorgesComplete} capabilities.
+\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
+\cite{Tsampas+19}, and monotonic \cite{GeorgesComplete} capabilities.
 Of these, uninitialized capabilities seem particularly promising---they naturally enforce
@@
 to clear large swaths of the stack. We plan to implement a model of the stack safety
-system described in \citet{Georges+21} and test it against our properties using
+system described in Georges et al. \cite{Georges+21} and test it against our properties using
 QuickChick.
@@
 %% Bibliography
+\bibliographystyle{ieeetran}
 \bibliography{bcp.bib,local.bib}

### 2021-09-27 submission4/main.tex
Removed policy state

@@
 
-We model the base machine's behavior as a step function over machine states,
-and a machine
-enhanced with an enforcement mechanism, or {\em policy} for short, as a
-partial step function over machine states extended with some
-auxiliary state, refining the base machine. When the policy-enhanced machine
-does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
+We model the machine's behavior as a partial step function over machine states.
+Machine states may contain arbitrary data, including data used exclusively by the
+enforcement mechanism, while our properties contend only with the data that are
+relevant to normal execution: addresses and registers.
+When the machine does not step from a given state, this
+is a {\em failstop}, halting the program before it can perform an
 action that would violate the security property;
 our properties are therefore naturally \emph{termination insensitive.}
-This model will be made more precise in \cref{sec:prelim}.
-
-Dividing policy state from machine state helps us reason about
-enforcement mechanisms that keep a significant amount of data separate from the
-primary execution data to inform security. Tag-based systems, in which a hardware
-monitor tracks metadata tags associated with a software-defined ``micro-policy,'' are
-the motivating application. The tags, and any extra policy state, are separated from
-and untouchable by application code.
 
 That said, our model is meant to apply generically to any enforcement mechanism,
-including those that do not take advantage of this separation.
-Some mechanisms, like hardware capability systems, might exhibit failstop behavior,
-but have no protected policy state---this is represented by a trivial policy
-that failstops when the hardware faults. Others might never even failstop, such as
-code rewriting techniques that prevent stack safety violations directly, in which
-case their policy state does nothing at all.
+including those that do not failstop, such as
+code rewriting techniques that prevent stack safety violations directly.
 
@@
 
-The building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively.
-%
-Our machine states are composed of {\em components} \(\component\),
-which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special-purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-A {\em machine state} is a map from components to values.
-
-\begin{minipage}{.45\columnwidth}
-  \[\component \in \COMPONENTS ::= \WORDS + \REGS\]
-\vspace{\afterdisplayskip}
-\end{minipage}
-\begin{minipage}{.45\columnwidth}
-  \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-\vspace{\afterdisplayskip}
-\end{minipage}
-
-We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-
-We now formalize the model of enforcement that we introduced in Section~\ref{sec:threat}.
-A {\em policy} consists of a set of policy
-states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. The policy also provides a
-well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
-initial states that are properly configured.
-%
-%A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-%\Cref{sec:enforcement}.---APT: We just said this.
-
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
-               \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol
-               \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-Note that the base machine has a total step function, while the policy-enhanced machine
-refines it to a partial function. The extra policy state never changes behavior,
-but it can cause
-premature termination, representing failstop behavior.
+The building blocks of the machine are {\em values}, {\em addresses}, and {\em registers}.
+Values and addresses are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
+\(\addr\), respectively. Register names (\(\reg\)) drawn from some set
+\(\REGS\), which is assumed to include a special-purpose register, the
+program counter {\PCname}. Collectively addresses and registers are {\em components} \(\component\)
+in the set \(\COMPONENTS ::= \WORDS + \REGS\).
+%
+A {\em machine state} \(\mach\) is drawn from a set \(\MACHS\) and supports a mapping
+from components to values, written
+\(\mach[\component] ::= \MACHS \rightarrow \COMPONENTS \rightarrow \WORDS \).
+
+We have a partial step function between machine states, written \(\mach \stepsto \mach'\), which
+is undefined on input states that correspond to a policy fault. So, a security mechanism is
+represented by the capacity of the machine to ``fail-stop,'' as well as by whatever initial
+well-formedness conditions are required for the mechanism to function properly.
 
@@
 that relates each code address to its annotation, if any.
-We lift contexts into an MP trace in much the same way as we lifted policy states to get
-a triple state \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times
-\CONTEXTS\).
-
-\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol \mpstate{\mach'}{\pol'}\)}
+We pair a machine state with a context state to get a combined ``MC-state''
+\(\mpcstatename \in \MPCS = \MACHS \times \CONTEXTS\), and lift the context-step
+function into the machine step like so:
+
+\judgmenttwo{\(\mach \stepsto \mach'\)}
               {\(\mach, \context \constep \context'\)}
-              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon
-                \mpcstate{\mach'}{\pol'}{\context'}\)}
+              {\(\mach,\context \stepstocon
+                 \mach',\context'\)}
 
@@
 
-A {\em machine trace} is a nonempty, finite or infinite sequence of elements
-of MPC-states, ranged over by \(\MPCT\) and \(\NPCT\).
+A {\em machine trace} is a nonempty, finite or infinite sequence
+of MC-states, ranged over by \(\MPCT\) and \(\NPCT\).
 We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces (ordinary lists will use
@@
 
-We project out the machine state of an MPC-state with \(\pi_\mach\), the
-policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
+We project out the machine state of an MPC-state with \(\pi_\mach\)
+and the context with \(\pi_\context\).
 We take the first element of a trace with \(\head(\MPCT)\), which is a total
@@
 properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
-where \((\mach,\pol) \in \WF\), \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
+is {\em reachable} if for some initial state \(\mpcstatename = \mach,\context_0\)
+where \(\mach\) is well-formed, \(\mpcstatename \hookrightarrow \MPCT\)
+and \(\MPCT\) contains \(\mpcstatename'\).
+A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
 such that from some similarly well-formed initial state \(\mpcstatename\),
@@
 A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
-\(\mach,\pol,\context\) such that \(\mach,\pol,\context \stepstocon \mpcstatename'\),
+\(\mach,\context\) such that \(\mach,\context \stepstocon \mach',\context'\),
 and any component \(\component \in \prot(\context)\),
-\(\pi_\mach(\mpcstatename') ~ \component = \mach ~ \component\).
+\(\mach'[\component] = \mach[\component]\).
 
@@
 \text{ is a reachable segment}\} \hspace{1em} &
-\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
+\text{where } f_\depth ~ (\mach,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
 \end{aligned}\]
@@
 A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
+for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\context\),
 any \(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\), and any trace
 %\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
-\(\NPCT\) such that \(\nach,\pol,\context \hookrightarrow \NPCT | \brk\),
+\(\NPCT\) such that \(\nach,\context \hookrightarrow \NPCT | \brk\),
 \(\MPCT \doteq \NPCT\) holds.
@@
 A system enjoys \textbf{\textit{deferred stepwise integrity}} if,
-for any \(\MPCT,\brk \in \MPCT^*\) where \(\head(\MPCT) = \mach,\pol,\context\) and
-\(\last(\MPCT) = \mach',\pol',\context'\), the following holds. Let
-\((\mach',\pol',\context') \hookrightarrow \MPCT'\) and
+for any \(\MPCT,\brk \in \MPCT^*\) where \(\head(\MPCT) = \mach,\context\) and
+\(\last(\MPCT) = \mach',\context'\), the following holds. Let
+\((\mach',\context') \hookrightarrow \MPCT'\) and
 \(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
@@
 For any \(\nach'\) and \(\NPCT'\) such that \(\mach' \approx_{\components} \nach'\) and
-\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
+\((\nach',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
 
@@
 \textbf{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
-where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
+where \(\head(\MPCT) = \mach\context\) and \(\last(\MPCT) = \mach',\context'\),
 the following holds. Let \(\components\) be the set of protected components that changed during
 the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
-and let \((\mach',\pol',\context') \hookrightarrow \MPCT'\).
+and let \((\mach',\context') \hookrightarrow \MPCT'\).
 Then for any \(\nach\) such that
 \(\mach' \approx_{\components'} \nach\), and its induced trace
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
+\((\nach,\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
 
@@
 A system enjoys \textbf{\em observational confidentiality} if for all
-\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\pol,\context\) and
+\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\context\) and
 \(\components = \secret(\context)\), the following holds.
@@
 (\last(\NPCT))\), i.e. if one call returns, so does the other
-\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
+\item Assuming both calls return, let \(\last(\MPCT) = \mach',\context'\)
   and \(\pi_\mach(\last(\NPCT)) = \nach'\),
@@
   and let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
-  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
-  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
+  Then if \(\mach',\context' \hookrightarrow \MPCT'\) and
+  \(\nach',\context' \hookrightarrow \NPCT'\), it must be
   that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)

### 2021-09-27 submission4/main.tex
Fixed displayskip name
Fixed definitions to fit in two column mode

--- 2021-09-27 submission4/main.tex
@@
   \MemoryLabel{-6em}{0.75em}{\(v_4\)}
-%  \MemoryLabel{-2em}{0.75em}{90}
+  \MemoryLabel{-2em}{0.75em}{90}
   \\
@@
   \[\component \in \COMPONENTS ::= \WORDS + \REGS\]
-\vspace{\afterdisplayskip}
+\vspace{\belowdisplayskip}
 \end{minipage}
@@
   \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-\vspace{\afterdisplayskip}
+\vspace{\belowdisplayskip}
 \end{minipage}
@@
 
-\judgmentthree[ Call]
+\judgmenttwobrlong[ Call]
               {\(\codemap ~ (\mach ~ \PCname)\)}
@@
 \vspace*{-1ex}
-\judgmenttwo[ ReturnRec]
+\judgmentbr[ ReturnRec]
             {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
@@
 \bigcup_{\depth \in \mathbb{Z}} \{\MPCT, \neg f_\depth | \MPCT\langle f_\depth \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
+\text{ is a reachable segment}\} \hspace{1em} \\
 \text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
@@
 
-\begin{minipage}{.3\textwidth}
 \judgment{}
          {\(\mpcstatename \doteq \npcstatename\)}
-\end{minipage}
-\begin{minipage}{.6\textwidth}
 \judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
@@
             {\(\notfinished{(\mach,\underscore,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore,\underscore)}{\nachT}\)}
-\end{minipage}
 
@@
 
-\begin{minipage}[b]{.55\textwidth}
   \judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
@@
                 {\(\mathit{obsfor} ~ \obs' ~ \MPCT = \obsT\)}
-                {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}%
-\end{minipage}
-\begin{minipage}[b]{.22\textwidth}
+                {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}
+
+\begin{minipage}[b]{.4\columnwidth}
   \judgment{}
@@
 \end{minipage}
-\begin{minipage}[b]{.22\textwidth}
+\begin{minipage}[b]{.4\columnwidth}
 \judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
@@
 
-\begin{minipage}{.2\textwidth}
+\begin{minipage}{.4\columnwidth}
   \judgment{}{\(\obsT \eqsim \obsT\)}
 \end{minipage}
-\begin{minipage}{.2\textwidth}
+\begin{minipage}{.4\columnwidth}
   \judgment{\(\obsT_1 \eqsim \obsT_2\)}
@@
 \end{minipage}
-\begin{minipage}{.2\textwidth}
+
+\begin{minipage}{.4\columnwidth}
   \judgment{\(\obsT_1 \eqsim \obsT_2\)}
@@
 \end{minipage}
-\begin{minipage}{.2\textwidth}
+\begin{minipage}{.4\columnwidth}
   \judgment{\(\obsT_1 \eqsim \obsT_2\)}
@@
   \hline
-        {\em PER\_DEPTH\_TAG} & Observational Integrity & 189.7 & 8342.5  \\
-        {\em LOAD\_NO\_CHECK}  & Observational Integrity & 23.5 & 12.0 \\
+        {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
+        {\em LOAD\_NO\_CHECK}  & Obs. Integrity & 23.5 & 12.0 \\
         {\em LOAD\_NO\_CHECK}  & Confidentiality & 19.2 & 695.5 \\
-        {\em STORE\_NO\_UPDATE} & Observational Integrity & 70 & 80.6  \\
+        {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
         {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
@@
 which are labeled with the range of depths of the caller that has passed them.
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
+\[\begin{split}
+\stackDom \in \STKDOMS ::= & \outside + \unsealed\\
+& + \sealed{\depth} + \passed{\depth,\depth'}
+\end{split}\]
 The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-\times (\listT (\target)) \]
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+\times (\rts \in \listT (\target)) \]
 The initial context again maps all components to \(\unsealed\).
@@
 update a domain map at a call from depth \(\depth\).
-\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
-\begin{cases}
-  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and }
-                                 \component - (\mach ~ \rsp) \not \in P \\
-  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
-                                 \sealcon ~ \mach ~ \component \text{ and }
-                                 \component - (\mach ~ \rsp) \in P \\
-  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
-                                 \sealcon ~ \mach ~ \component \\
-                               & \text{ and } (\mach ~ \rsp) \in P, \text{ for any } \depth' \\
-  \domMap~\component & \text{else} \\
-\end{cases}\]
+
+\judgmentthree[]
+              {\(\domMap ~ \component = \unsealed\)}
+              {\(\sealcon ~ \mach ~ \component\)}
+              {\(\component - (\mach ~ \rsp) \not \in P\)}
+              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \sealed{\depth}\)}
+\judgmentthree[]
+              {\(\domMap ~ \component = \unsealed\)}
+              {\(\sealcon ~ \mach ~ \component\)}
+              {\(\component - (\mach ~ \rsp) \in P\)}
+              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth,\depth+1}\)}
+\judgmentthree[]
+              {\(\domMap ~ \component = \passed{\depth',\depth}\)}
+              {\(\sealcon ~ \mach ~ \component\)}
+              {\(\component - (\mach ~ \rsp) \in P\)}
+              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth',\depth+1}\)}
+\judgment[]
+         {}
+         {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \domMap ~ \component\)}
 
          At the matching return we just unseal the components at depth \(\depth\):
-%
-\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
-\begin{cases}
-  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
-  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
-  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
-                              \text{ for any } \depth' < \depth \\
-  \domMap~\component        & \text{else} \\
-\end{cases}\]
+
+\judgment[]
+         {\(\domMap ~ \component = \sealed{\depth}\)}
+         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
+\judgment[]
+         {\(\domMap ~ \component = \passed{\depth-1,\depth}\)}
+         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
+\judgmenttwo[]
+           {\(\domMap ~ \component = \passed{\depth',\depth}\)}
+           {\(\depth' < \depth\)}
+           {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
+\judgment[]
+         {}
+         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
 
@@
 
-\judgmenttwo[ Call]
+\judgmenttwo[ CallSeal]
             {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
@@
 %
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(S) ~ \mbox{\rm{where}} ~ S \in \REGS \rightarrow (\unsaved + \saved)\]
-%
-\vspace{-3.5ex}
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \unsaved + \saved\]
+\[\begin{split}
+\codemap \in \CODEMAPS ::= & \WORDS \rightharpoonup \callmap(S)\\
+\mbox{\rm{where}} ~ & S \in \REGS \rightarrow (\unsaved + \saved) \\
+\end{split}\]
+
+\[\begin{split}
+\stackDom \in \STKDOMS ::= & \outside + \unsealed + \sealed{\depth} \\
+& + \unsaved + \saved
+\end{split}\]
 %
@@
 
-\paragraph{Discussion}
+\paragraph*{Discussion}
 
@@
 \[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
-\[(\domMap,\rtm,\yts,\stackid) = \context \in \CONTEXTS ::=
- (\COMPONENTS \rightarrow \TOPDOMS) \times
-(\STACKS \rightarrow \listT(\target)) \times (\STACKS \rightarrow \target) \times \STACKS\]
+\[\begin{split}
+\context \in \CONTEXTS ::= &
+ (\domMap \in \COMPONENTS \rightarrow \TOPDOMS) \times \\
+& (\rtm \in \STACKS \rightarrow \listT(\target)) \times \\
+& (\yts \in \STACKS \rightarrow \target) \times \\
+& \stackid \in \STACKS
+\end{split}\]
 
@@
 
-\judgmentfour[ Yield]
-             {\(\mach \stepsto \mach'\) ~ ~}
-             {~ ~ \(\stackid' = \stackof ~ (\mach'~ \rsp) \not = \stackid\) ~ ~}
-             {~ ~ \(\yts ~ \stackid'\) ~ ~}
-             {~ ~ \(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
+\judgmenttwobr[ Yld]
+             {\(\mach \stepsto \mach'\)}
+             {\(\stackid' = \stackof ~ (\mach'~ \rsp) \not = \stackid\)}
+             {\(\yts ~ \stackid'\)}
+             {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
              {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
@@
 \bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{\MPCT, \neg f_\stackid | \MPCT\langle f_\stackid \rangle
-\text{ is a reachable segment}\} \hspace{1em} &
+\text{ is a reachable segment}\} \hspace{1em} \\
 \text{where } f_\stackid ~ (\underscore,\underscore,(\underscore,\underscore,\underscore,\stackid')) \triangleq \stackid \not = \stackid' \\

### 2021-09-27 submission4/main.tex
Shrunk text in various ways; removed deferred integrity

@@
 \begin{figure*}
-\begin{subfigure}[t]{.25\textwidth}
+\begin{subfigure}[t]{.16\textwidth}
 {\small
 \begin{verbatim}
-int main() {
-  int x = 42;
-  return f()+x;
-}
-
 int f() {
@@
 \end{subfigure}
-\begin{subfigure}[t]{.5\textwidth}
+\begin{subfigure}[t]{.3\textwidth}
 {\small
 \begin{verbatim}
-1 main: add $1,%sp      ; allocate frame
-2       mov $42,-1(%sp) ; initialize local
-3       call f
-4       add -1(%sp),%ra ; set return value
-5       sub $1,%sp      ; deallocate frame
-6       ret
-
-10 f:   add $2,%sp      ; allocate frame
+   f:
+10 add $2,%sp      ; allocate frame
 11 mov $5,-2(%sp)  ; assign into y
@@
 
-\begin{figure*}
-\begin{center}
-\begin{subfigure}[t]{.2\textwidth}
+\begin{figure}
+%\begin{center}
+\begin{subfigure}[t]{.32\columnwidth}
   {\small
@@
 }
-
+\end{verbatim}
+  }
+\end{subfigure}
+\begin{subfigure}[t]{.32\columnwidth}
+  {\small
+\begin{verbatim}
 int f() {
@@
 }
-
+\end{verbatim}
+  }
+\end{subfigure}
+\begin{subfigure}[t]{.32\columnwidth}
+  {\small
+\begin{verbatim}
 int g() {
@@
 \end{subfigure}
-\begin{subfigure}[t]{.68\textwidth}
-  {\small
-\begin{verbatim}
-1 main:  add $1,%sp      ; allocate frame
-2        call f
-3        mov %ra,-1(%sp) ; store x
-4        call g
-5        sub $1,%sp      ; deallocate frame
-6        ret
-
-10 f:    add $1,%sp      ; allocate frame
-11       mov $5,-1(%sp)  ; initialize y
-12       mov -1(%sp),%ra ; set return value
-13       sub $1,%sp      ; deallocate frame
-14       ret
-
-20 g:    add $1,%sp      ; allocate frame
-21       mov -1(%sp),%ra ; violation!
-22       sub $1,%sp      ; deallocate frame
-23       ret
-}
-\end{verbatim}
-}
-\end{subfigure}
-\end{center}
-\begin{subfigure}{.65\textwidth}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 2} &
-  \multicolumn{2}{c}{
-    \memoryaddrs{8em}
-    \memory{3}{\unsealc}
-    ~$\cdots$
-%    \MemoryLabel{-18em}{0.75em}{0}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
- % \MemoryLabel{-19em}{0.75em}{0}
-%  \MemoryLabel{-10em}{0.75em}{0}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
-  \\
-  {\tt 12} &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
- % \MemoryLabel{-19em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(5\)}
-  \\
-  {\tt 4} &
-  \multicolumn{2}{c}{
-    \memoryaddrs{8em}
-%    \memory{4}{\unsealc}
-    \memory{1}{\unsealc}
-    \memory{1}{\retptrc}
-    \memory{1}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-14em}{0.75em}{5}
-%    \MemoryLabel{-10em}{0.75em}{\(v_1\)}
-    \MemoryLabel{-6em}{0.75em}{5}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  }
-  \\
-  {\tt 22} &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\bf 5}
- % \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs{16em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-15em}{0.75em}{\(v_3\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_4\)}
-  \MemoryLabel{-2em}{0.75em}{90}
-  \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
+%\begin{subfigure}[t]{.68\textwidth}
+%  {\small
+%\begin{verbatim}
+%1 main:  add $1,%sp      ; allocate frame
+%2        call f
+%3        mov %ra,-1(%sp) ; store x
+%4        call g
+%5        sub $1,%sp      ; deallocate frame
+%6        ret
+%
+%10 f:    add $1,%sp      ; allocate frame
+%11       mov $5,-1(%sp)  ; initialize y
+%12       mov -1(%sp),%ra ; set return value
+%13       sub $1,%sp      ; deallocate frame
+%14       ret
+%
+%20 g:    add $1,%sp      ; allocate frame
+%21       mov -1(%sp),%ra ; violation!
+%22       sub $1,%sp      ; deallocate frame
+%23       ret
+%}
+%\end{verbatim}
+%}
+%\end{subfigure}
+%\end{center}
+%\begin{subfigure}{.65\textwidth}
+%\begin{center}
+%\begin{tabular}{l r | l}
+%  {\tt 2} &
+%  \multicolumn{2}{c}{
+%    \memoryaddrs{8em}
+%    \memory{3}{\unsealc}
+%    ~$\cdots$
+%    \vspace{.5em}
+%  } \\
+%  {\tt 11} &
+%  \memoryaddrs{16em}
+%  \memory{1}{\unsealc}
+%  \memory{1}{\retptrc}
+%  \memory{1}{\unsealc}
+%  ~$\cdots$
+%  &
+%  \memoryaddrs{16em}
+%  \memory{1}{\unsealc}
+%  \memory{1}{\retptrc}
+%  \memory{1}{\unsealc}
+%  ~$\cdots$
+%  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
+%  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
+%  \\
+%  {\tt 12} &
+%  \memoryaddrs{16em}
+%  \memory{1}{\unsealc}
+%  \memory{1}{\retptrc}
+%  \memory{1}{\unsealc}
+%  ~$\cdots$
+%  \MemoryLabel{-6em}{0.75em}{5}
+%  &
+%  \memoryaddrs{16em}
+%  \memory{1}{\unsealc}
+%  \memory{1}{\retptrc}
+%  \memory{1}{\unsealc}
+%  ~$\cdots$
+%  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
+%  \MemoryLabel{-6em}{0.75em}{\(5\)}
+%  \\
+%  {\tt 4} &
+%  \multicolumn{2}{c}{
+%    \memoryaddrs{8em}
+%    \memory{1}{\unsealc}
+%    \memory{1}{\retptrc}
+%    \memory{1}{\unsealc}
+%    ~$\cdots$
+%    \MemoryLabel{-14em}{0.75em}{5}
+%    \MemoryLabel{-6em}{0.75em}{5}
+%    \vspace{.5em}
+%  }
+%  \\
+%  {\tt 22} &
+%  \memoryaddrs{16em}
+%  \memory{1}{\unsealc}
+%  \memory{1}{\retptrc}
+%  \memory{1}{\unsealc}
+%  ~$\cdots$
+%  \MemoryLabel{-15em}{0.75em}{5}
+%  \MemoryLabel{-6em}{0.75em}{\bf 5}
+%  &
+%  \memoryaddrs{16em}
+%  \memory{1}{\unsealc}
+%  \memory{1}{\retptrc}
+%  \memory{1}{\unsealc}
+%  ~$\cdots$
+%  \MemoryLabel{-15em}{0.75em}{\(v_3\)}
+%  \MemoryLabel{-6em}{0.75em}{\(v_4\)}
+%  \MemoryLabel{-2em}{0.75em}{90}
+%  \\
+%\end{tabular}
+%\end{center}
+%
+%\vspace{\abovedisplayskip}
+%
+%\end{subfigure}
 \caption{Another confidentiality violation}
 \label{fig:conf2}
-\end{figure*}
+\end{figure}
 
 It may be unclear in Figure~\ref{fig:conf1} why we vary the initial values of {\tt y}
-and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
-{\em temporal stack safety}.
-In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
+and {\tt z} as well as that of {\tt x}. This is to detect some violations of so-called
+{\em temporal stack safety}. Suppose, in a 
+In Figure~\ref{fig:conf2}, {\tt f} leaves behind a value in {\tt y}
 that later becomes the value of {\tt main}'s variable {\tt x}. Then {\tt g} is able to access
-{\tt y} and hence {\tt x} without even directly reading the stack pointer.
-In the primary trace, {\tt z} will have the value 5, but in the variant it can have
-an arbitrary value $v_4$.  Since the variant can behave different from the primary,
-this again represents a violation of confidentiality.
+{\tt y} and hence {\tt x} without even directly reading the stack pointer. In terms of noninterference,
+in the primary trace, {\tt z} will have the value 5, but in the variant it can have
+an arbitrary value. We regard this as a violation of confidentiality as well.
 
@@
 
-\label{sec:controlflow} % RB: ???
-
 %From the classic stack smashing attack to return oriented programming,
@@
 \begin{center}
-\vspace*{-\medskipamount}
 \judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
@@
 \end{center}
-
-\vspace*{-\medskipamount}
 
@@
 \begin{center}
-\vspace*{-\medskipamount}
   \judgmenttwo[]
@@
               {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
-\vspace*{-\bigskipamount}
   \judgmentthree[]
@@
                 {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\vspace*{-\bigskipamount}
   \judgmentthree[]
@@
 \(\MPCT \doteq \NPCT\) holds.
-
-%\medskip
 
@@
 
-\paragraph*{Deferring Integrity Checks}
-
-We first notice that these properties prevent invalid reads, as well as reads
-that follow invalid writes, but the writes themselves are permitted.
-So we wish in particular to find an integrity property that characterizes them.
-The intuition is that sealed memory can change, but it is
-then tainted, and its original owner cannot actually read it. This
-leads us to a hybrid of integrity and confidentiality.
-
-\definition
-A system enjoys \textbf{\textit{deferred stepwise integrity}} if,
-for any \(\MPCT,\brk \in \MPCT^*\) where \(\head(\MPCT) = \mach,\pol,\context\) and
-\(\last(\MPCT) = \mach',\pol',\context'\), the following holds. Let
-\((\mach',\pol',\context') \hookrightarrow \MPCT'\) and
-\(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
-set of sealed components that changed during the call.
-For any \(\nach'\) and \(\NPCT'\) such that \(\mach' \approx_{\components} \nach'\) and
-\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\).
-
-Intuitively, deferred stepwise integrity means that if any components change from
-the a call to its matching return, they will afterward be treated as confidential
-(in the sense of stepwise confidentiality; the return state and a variant would
-execute in lockstep.)
-
-For example, consider again the integrity violation in
-Figure~\ref{fig:int1}. Under a deferred discipline, the write to {\tt
-  x} in instruction 11 taints the variable but does not in itself
-constitute a violation, which happens only when the tainted value is
-read in instruction 4. However, if this instruction added the constant
-that was originally stored in {\tt x} without reading the variable,
-i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
-Instead of strictly forbidding such dangerous writes, we say,
-``No reads after dangerous writes.''
-
 \subsection{Observability}
 
-While deferred integrity meaningfully captures the intuition
-behind lazy tagging, we are interested in properties that can be used to
-evaluate the broadest possible range of reasonable
+In addition to capturing the intuition behind lazy tagging, we are interested
+in properties that can be used to evaluate the broadest possible range of reasonable
 enforcement mechanisms.
-For example, reading after a dangerous write may be acceptable if the value
-read does not affect the output of the program; this looser scheme
-could be enforced by tainting the dangerous writes and detecting taint
-at the point of output.
-%For example, consider a hypothetical
-%taint-tracking policy that allows high-security data to be copied as long as it is
-%tagged as such and not exfiltrated, or that allows dangerous writes to be
-%copied as long as they don't affect critical operations. \apt{Not quite sure
-%  how these last connect to stack safety.}
+
 And some code features may involve moving data, overwriting its original location,
@@
 which is carried over into the policy and context step functions:
-%
-$\mach \stepsto[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS $.
-%
 
@@
 %  a technical obstacle, it's that the behavior is tied to things we don't model.}
-\paragraph{Formalization of variable passing.}
+\paragraph*{Formalization of variable passing.}
 We add extra information to our call

### 2021-09-27 submission4/main.tex
Removed later references to deferred integrity

@@
 than the stepwise one, and also much more extensional.
-We can think of stepwise, deferred, and observational as
-three points along a spectrum from eager to lazy. The first two
-are useful for understanding particular enforcement mechanisms,
-but if we had to choose a single property to represent stack safety,
-it would be the observational one.
+We can think of the stepwise properties as instantiating
+the observation model with a view of the entire state, and
+thus as one end of a spectrum that can become arbitrarily permissive.
+Therefore, if we had to choose a single property to represent stack safety,
+it would be the observational formalization.
 
@@
 \definition
-For each of the \textbf{\em stepwise}, \textbf{\em deferred}, and \textbf{\em observational} versions of
+For each of the \textbf{\em stepwise} and \textbf{\em observational} versions of
 \textbf{\em stack integrity} in the simple model, the equivalent property in the
@@
 function, and \(\prot_Y\) for their equivalents. Likewise
-\textbf{\em deferred coroutine integrity} and \textbf{\em observational coroutine integrity} are defined
+\textbf{\em observational coroutine integrity} is defined
 in terms of their simple equivalents using \(\MPCT^\wedge\) in place of \(\MPCT^*\).

### 2021-09-27 submission4/main.tex

@@
 
-The building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively.
-%
-Our machine states are composed of {\em components} \(\component\),
-which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special-purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-A {\em machine state} is a map from components to values.
-
-\begin{minipage}{.45\columnwidth}
-  \[\component \in \COMPONENTS ::= \WORDS + \REGS\]
-\vspace{\belowdisplayskip}
-\end{minipage}
-\begin{minipage}{.45\columnwidth}
-  \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-\vspace{\belowdisplayskip}
-\end{minipage}
-
-We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-
-We now formalize the model of enforcement that we introduced in Section~\ref{sec:threat}.
-A {\em policy} consists of a set of policy
-states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. The policy also provides a
-well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
-initial states that are properly configured.
-%
-%A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-%\Cref{sec:enforcement}.---APT: We just said this.
-
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
-               \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol
-               \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-Note that the base machine has a total step function, while the policy-enhanced machine
-refines it to a partial function. The extra policy state never changes behavior,
-but it can cause
-premature termination, representing failstop behavior.
+The building blocks of the machine are {\em values}, {\em addresses}, and {\em registers}.
+Values and addresses are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
+\(\addr\), respectively. Register names (\(\reg\)) drawn from some set
+\(\REGS\), which is assumed to include a special-purpose register, the
+program counter {\PCname}. Collectively addresses and registers are {\em components} \(\component\)
+in the set \(\COMPONENTS ::= \WORDS + \REGS\).
+%
+A {\em machine state} \(\mach\) is drawn from a set \(\MACHS\) and supports a mapping
+from components to values, written
+\(\mach[\component] ::= \MACHS \rightarrow \COMPONENTS \rightarrow \WORDS \).
+
+We have a partial step function between machine states, written \(\mach \stepsto \mach'\), which
+is undefined on input states that correspond to a policy fault. So, a security mechanism is
+represented by the capacity of the machine to ``fail-stop,'' as well as by whatever initial
+well-formedness conditions are required for the mechanism to function properly.
 
@@
 \text{ is a reachable segment}\} \hspace{1em} \\
-\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
+\text{where } f_\depth ~ (\mach,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
 \end{aligned}\]

### 2021-09-28 submission4/main.tex
Shortened micropolicy descriptions

--- 2021-09-27 submission4/main.tex
@@
 
-\paragraph*{Generators}
-
-To write such generators we build on the work of
-Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
-introduced {\em generation by execution} to produce progams that lead
-machines towards interesting behaviors. Generation by
-execution receives as an input a partially instantiated machine state
-and attempts to generate an instruction (or a sequence of instructions
-such as a blessed sequence) that makes sense locally (e.g., jumps go
-to a potentially valid code location and loads read from a
-potentially valid stack location). Then we step the machine and repeat
-the process until we generate or execute some target number of
-instructions, or reach a point where the machine cannot step
-any more.
-
-We extend this technique to keep track of the control flow behavior of
-the program being generated: each time a call or return sequence is
-generated, we ensure that the appropriate policy tags and code
-annotations are set for the entry or return points.% At the same time,
-%we allow the generation to sometimes relax those constraints,
-%introducing potentially ill-formed flows: this causes our programs
-%to failstop when executed in conjunction with the policy monitor,
-%but also allows for revealing errors in our setup.\apt{??}
-
-In addition, we need to further extend this technique to handle the nested
-nature of our confidentiality property: rather than just generating two
-initial machines that are variants of one another and letting them
-execute to test for noninterference, we generate a new variant
-{\em every time a call is made} and check confidentiality for the
-subtrace produced from that variant state until its corresponding
-return. As a result, a ``single'' confidentiality test compactly
-checks the confidentiality of multiple nested calls.
+%\paragraph*{Generators}
+%
+%To write such generators we build on the work of
+%Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
+%introduced {\em generation by execution} to produce progams that lead
+%machines towards interesting behaviors. Generation by
+%execution receives as an input a partially instantiated machine state
+%and attempts to generate an instruction (or a sequence of instructions
+%such as a blessed sequence) that makes sense locally (e.g., jumps go
+%to a potentially valid code location and loads read from a
+%potentially valid stack location). Then we step the machine and repeat
+%the process until we generate or execute some target number of
+%instructions, or reach a point where the machine cannot step
+%any more.
+%
+%We extend this technique to keep track of the control flow behavior of
+%the program being generated: each time a call or return sequence is
+%generated, we ensure that the appropriate policy tags and code
+%annotations are set for the entry or return points.% At the same time,
+%%we allow the generation to sometimes relax those constraints,
+%%introducing potentially ill-formed flows: this causes our programs
+%%to failstop when executed in conjunction with the policy monitor,
+%%but also allows for revealing errors in our setup.\apt{??}
+%
+%In addition, we need to further extend this technique to handle the nested
+%nature of our confidentiality property: rather than just generating two
+%initial machines that are variants of one another and letting them
+%execute to test for noninterference, we generate a new variant
+%{\em every time a call is made} and check confidentiality for the
+%subtrace produced from that variant state until its corresponding
+%return. As a result, a ``single'' confidentiality test compactly
+%checks the confidentiality of multiple nested calls.
 
@@
 
-\section{Stack Safety for Lazy Enforcement}
+\section{Observable Stack Safety}
 \label{sec:lazy}
@@
 is particularly costly for frames that contain large uninitialized arrays.
-To mitigate those costs, they propose a number of optimizations to the policy
-that make it more permissive but also more efficient to implement.
-
-First, they eliminate the costly frame initialization pass. In the {\em Lazy Tagging}
-system, the stack is again initially tagged as \(\tagNoDepth\).
-A function is permitted
-to overwrite \(\tagNoDepth\) memory, and when it does so its depth tag overwrites
-the \(\tagNoDepth\) tag, claiming the memory; the remainder of the policy functions
-as in Depth Isolation. However, on return, the callee must clear its frame, restoring
-each tag to \(\tagNoDepth\), which is still a heavy cost.
-
-To eliminate this cost, they augment Lazy Tagging
-with {\em Lazy Clearing}. Under this combination of optimizations,
-the policy does not initialize stack frames on entry \emph{or} clear them on exit,
+To mitigate those costs, they eliminate the costly frame initialization pass.
+
+Under the most optimized version of the micro-policy, {\em Lazy Tagging and Clearing},
+it does not initialize stack frames on entry \emph{or} clear them on exit,
 and it permits all writes to the stack, even when the $\PCname$ tag
@@
 if it ever tries to read from that location.
-%
-These \emph{lazy policies} admit more efficient implementations, but
-they deliberately allow temporary violations of stack integrity,
-with the checks deferred until the point a violation truly becomes
-harmful. The natural question then is: how do we characterize the
+
+This \emph{lazy policy} admits more a efficient implementations, but
+they deliberately allow temporary violations of stack integrity.
+The natural question then is: how do we characterize the
 protections provided by these policies once stepwise stack safety is broken?
@@
 
-\subsection{Observability}
-
-In addition to capturing the intuition behind lazy tagging, we are interested
-in properties that can be used to evaluate the broadest possible range of reasonable
-enforcement mechanisms.
-
-And some code features may involve moving data, overwriting its original location,
+Beyond this one enforcement mechanism, some code features may involve
+moving data, overwriting its original location,
 and ultimately restoring it before the change becomes visible; for example, the
@@
 
-\subsection{Observational Stack Safety}
+\paragraph*{Observational Stack Safety}
 
@@
 \begin{verbatim}
-void main() {
-  int x = 5;
-  x = f(x);
-  return;
-}
-
-void f(int a) {
-  int y = 2*a;
-  a = 0;
-  return;
-}
-\end{verbatim}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.6\textwidth}
-{\small
-\begin{verbatim}
-1  main: add $1,%sp          ; allocate frame
-2        mov $5,-1(%sp)      ; initalize x
-3        add $1,%sp          ; allocate arg
-4        mov -2(%sp),-1(%sp) ; set arg
-5        call f              ; SHARE: %sp-1
-6        sub $1,%sp          ; deallocate arg
-7        mov %ra,-1(%sp)     ; store x
-8        sub $1,%sp          ; deallocate frame
-9        ret
-
-10  f:   add $1,%sp          ; allocate frame
-11       mov -3(%sp),-1(%sp) ; y = a
-12       mul $2,-1(%sp)      ; y = a*2
-13       mov $0,-3(%sp)      ; a = 0
-14       sub $1,%sp          ; deallocate frame
-15       ret
-\end{verbatim}
-}
-  \end{subfigure}
-\begin{subfigure}{.65\textwidth}
-\vspace*{0.2in}
-\begin{center}
-\begin{tabular}{l r | l}
-  {\tt 3} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{5}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
-    \vspace{.5em}
-  } \\
-  {\tt 11} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}[\sealdesc{0}]%
-  \memory{1}{\mainpassc}[\passdesc{0}{1}]%
-  \memory{1}{\retptrc}[\retptrdesc]%
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{0}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  {\tt 13} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{5}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  {\tt 14} &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{5}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  &
-  \memoryaddrs[(1)]{21.5em}
-  \memory{1}{\mainsealc}
-  \memory{1}{\mainpassc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-14em}{0.75em}{0}
-  \MemoryLabel{-6em}{0.75em}{10}
-  \\
-  \\
-  {\tt 8} &
-  \multicolumn{2}{c}{
-    \memoryaddrs[(0)]{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \MemoryLabel{-19em}{0.75em}{10}
-    \MemoryLabel{-14em}{0.75em}{0}
-    \MemoryLabel{-10em}{0.75em}{\#6}
-    \MemoryLabel{-7em}{0.75em}{10}
-    \vspace{.5em}
-  } \\
-\end{tabular}
-\end{center}
-
-\vspace{\abovedisplayskip}
-
-\end{subfigure}
-
-\caption{Sharing on the stack \ifaftersubmission\bcp{Not enough space above the bottom part of
-    the figure (at least in the latex build that I'm looking at...)}\fi}
-\label{fig:passingsimple}
-\end{figure*}
-
-\begin{figure*}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
 int *stash = 0;
@@
   \end{subfigure}
-  \begin{subfigure}[t]{.49\textwidth}
+  \begin{subfigure}[t]{.47\textwidth}
 {\small
@@
    stash: .word 0
-1  main:  add $2,%sp      ; allocate frame
+1  main:  add $2,%sp      ; alloc frame
 2         mov %sp,%ra
@@
 8         call f          ; SHARE: %sp-1
-9         mov -2(%sp),%ra ; set return value
-10        sub $2,%sp      ; deallocate frame
+9         mov -2(%sp),%ra ; set return
+10        sub $2,%sp      ; dealloc frame
 11        ret
@@
   \end{subfigure}
-  \begin{subfigure}[t]{.29\textwidth}
+  \begin{subfigure}[t]{.27\textwidth}
     \begin{center}
@@
 \end{center}
-\vspace{\abovedisplayskip}
 \end{subfigure}
@@
 
-Example~\ref{fig:passingsimple} shows a sample trace of
-a program that passes a value on the stack, without an attack.
-We extend our call annotation with an additional parameter identifying a memory location to be
-passed (the diagram displays it in magenta). Our notion of domains is extended as well,
-to include \(\passed{\depth,\depth'}\), recording two depths. We note the
-current depth of the active function next to the stack pointer in the diagram, for reference.
-The argument {\tt a} has
-its value copied from {\tt x}, and we note that {\tt f} can read it---because it is marked
-as passed between depths 0 and 1, it is not varied. At instruction 13 it is
-modified---naturally this is permitted as well.
-
-The same annotation mechanism can be used to pass
-variables by reference, and here there are important stack safety concerns,
-as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference.  The function
+Example~\ref{fig:passingsimple} shows a sample trace of a program that passes
+variables by reference. In this example, written in C++ notation,
+{\tt f} takes a single argument by reference (the diagram displays it in magenta). The function
 {\tt main} calls {\tt f}, passing a reference to {\tt x}, then clears {\tt x} and calls
@@
 The question is how the local variables {\tt x} and {\tt y} themselves
-should be labelled in order for the policy to detect an integrity violation.
+should be labelled in order for the property to correctly classify an integrity violation.
+Our notion of domains is extended to include \(\passed{\depth,\depth'}\),
+recording two depths for reasons explained below.
 At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
@@
                   {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
-%              {\(\components = \sealcon ~ \mach\)}
                   {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}

### 2021-09-28 submission4/main.tex
Updated state access notation

@@
 %              {\(\components = \sealcon ~ \mach\)}
-              {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
+              {\(\components = \{\component | \sealcon ~ \mach[\component]\}\)}
               {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
@@
 written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\).
+all \(\component \not \in \components\), \(\mach[\component] = \nach ~ \component\).
 
@@
 is the set of components \(\component\)
-such that \(\mach ~ \component \not = \mach' ~ \component\).
+such that \(\mach[\component] \not = \mach' ~ \component\).
 
@@
               {\(\domMap ~ \component = \unsealed\)}
-              {\(\sealcon ~ \mach ~ \component\)}
+              {\(\sealcon ~ \mach[\component]\)}
               {\(\component - (\mach ~ \rsp) \not \in P\)}
@@
               {\(\domMap ~ \component = \unsealed\)}
-              {\(\sealcon ~ \mach ~ \component\)}
+              {\(\sealcon ~ \mach[\component]\)}
               {\(\component - (\mach ~ \rsp) \in P\)}
@@
               {\(\domMap ~ \component = \passed{\depth',\depth}\)}
-              {\(\sealcon ~ \mach ~ \component\)}
+              {\(\sealcon ~ \mach[\component]\)}
               {\(\component - (\mach ~ \rsp) \in P\)}
@@
                   {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
-                  {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
+                  {\(\components = \{\component | \sealcon ~ \mach[\component]\}\)}
                   {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}

### 2021-09-28 submission4/main.tex
s/component/element

--- 2021-09-28 submission4/main.tex
@@
 (the machine is word addressed), and addresses increase to the right.
+We refer to registers and addresses collectively as {\em state elements} or just elements.
   %% \bcp{Most of the
@@
 %\bcp{where do we see this happening?  Is it important?}.
-Domains are not part of the machine state; rather, they
-describe the intended security role of the address
-that the enforcement mechanism must adhere to. Such a domain
+Domains are not part of the machine state; rather, they are part
+of the {\em context} of a state that describes the security restrictions
+that the enforcement mechanism must adhere to.
+(We will introduce some other parts of the context as needed.)
+Such a domain
 can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
@@
 be treated properly later.
-Domains form part of the {\em context} associated with a machine state.
-(We will introduce some other parts of the context as needed.)
 
@@
 function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
-variable {\tt x}. (Of course, this is illegal C, because the write in {\tt f} triggers
-undefined behavior, but it is characteristic of what buggy or malicious code might do.)
+variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
 In the initial memory layout the entire stack is {\em unsealed},
@@
 This leads to a simple, intuitive statement of {\em stepwise integrity}:
-if a component is sealed before a step, it will be unchanged afterward.
+if a element is sealed before a step, it will be unchanged afterward.
 % Stepwise
@@
 \begin{figure*}
-\begin{subfigure}[t]{.16\textwidth}
+\begin{subfigure}[t]{.2\textwidth}
 {\small
@@
 \begin{verbatim}
-   f:
-10 add $2,%sp      ; allocate frame
+10 f: add $2,%sp      ; allocate frame
 11     mov $5,-2(%sp)  ; assign into y
@@
 
-Dually, confidentiality states intuitively that a callee is insensitive to
+Confidentiality states intuitively that a callee is insensitive to
 the context of its call (except for its arguments and any global
@@
 We state this as a form of {\em noninterference}~\cite{Goguen1982SecurityPA}.
-Noninterference describes knowledge: it states that for system components
-containing secret values, we can substitute any values whatsoever for the secrets without changing the effect
-of our computation on non-secret values.
+Noninterference describes knowledge: its standard form states that for state elements
+containing secret values, we can substitute any values whatsoever for the secrets
+without changing the effect of our computation on non-secret values.
 % \bcp{That's too loose.  Even for standard NI, high inputs
@@
 If there exists some value for which we see different results, then
-in fact we can learn something about the initial state by observing those results, violating
-confidentiality.
+in fact we can learn something about the initial state by observing those results.
+To adapt this to the stack, the notion of ``secret'' must be dynamic with function
+calls and returns.
 
@@
 be hidden from {\tt f}; we show the evolving stack memory of the variant in a second, parallel column. At instruction 12, we note that the callee's first variable, {\tt y}, has
-changed to 5 in both the primary trace and the variant. The primary and variant agree on
-the value in the changed location, so confidentiality was not violated. But at instruction 13,
+changed to 5 in both the primary trace and the variant. The primary and variant correctly agree on
+the value in the changed location. But at instruction 13,
 the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
@@
 means that if both states step---that is, neither the primary nor the variant failstops---any
-{\em system component} (such as memory cell, register, program counter) that changes
-in the original will change in the same way in the variant.
-When we later extend the model with a notion of observable output
-(\cref{sec:lazy}), we will apply this condition only to those.
-
-%\bcp{Remove this para?}We alternatively could imagine a version of stepwise confidentiality in which we explicitly
-%track which addresses may be read (i.e., which have been initialized), and at each step
-%guarantee that the step alters state identically to a step from a variant state. Then we
-%wouldn't need to divide up the execution by calls. But the approach we take is more
-%consistent with a typical noninterference approach, in that it varies the memory once
-%at the beginning and trace changes through the execution.
+state element that changes in the original will change in the same way in the variant.
 
@@
 %
-Our machine states are composed of {\em components} \(\component\),
+Our machine states are composed of {\em elements} \(\component\),
 which are either addresses or register names (\(\reg\)) drawn from some set
@@
 program counter {\PCname} and the stack pointer \(\SP\).
-A {\em machine state} is a map from components to values.
+A {\em machine state} is a map from elements to values.
 
@@
 
-Each system component is assigned to a {\em domain}, which may be {\em outside}
+Each state element is assigned to a {\em domain}, which may be {\em outside}
 of the stack entirely, {\em unsealed} and therefore accessible, or {\em sealed}
@@
 has been returned to.
-A context is a pair of a ``domain map'' from components to domains, and a
+A context is a pair of a ``domain map'' from elements to domains, and a
 ``return target list,'' which is a stack of targets, one for each caller awaiting
@@
 A call annotation tells us to push the return target onto the return stack and
-seal all components with the depth of the previous return stack.
+seal all elements with the depth of the previous return stack.
 (Recall that the code map \(\codemap\) is a partial function, and we can use its domain
@@
 When we build machine traces with this step function, we always know how deep in the
-call stack we are and which components are sealed.
+call stack we are and which elements are sealed.
 
@@
 quantified over each step of execution from any initial state. The criterion for integrity
-is simple: sealed components do not change.
+is simple: sealed elements do not change.
 
@@
 The \emph{protected set} of a context, \(\prot(\domMap,\_)\), is the
-set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
+set of elements \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
 for some depth \(\depth\).
@@
 \(\mach,\pol,\context\) such that \(\mach,\pol,\context \stepstocon \mpcstatename'\),
-and any component \(\component \in \prot(\context)\),
+and any element \(\component \in \prot(\context)\),
 \(\pi_\mach(\mpcstatename') ~ \component = \mach ~ \component\).
@@
 \definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
-is the set of components \(\component\)
+is the set of elements \(\component\)
 such that \(\mach ~ \component \not = \mach' ~ \component\).
@@
 \definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
- {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all components \(\component\) in
+ {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all elements \(\component\) in
 \(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
@@
 The \emph{secret set} of a context, written \(\secret(\domMap,\rts)\), is the
-set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
+set of elements \(\component\) such that \(\domMap ~ \component \not = \outside\).
 
@@
 Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
-of the components in \(\components\), and in addition their subtraces are also in
+of the elements in \(\components\), and in addition their subtraces are also in
 \(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
@@
 where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
-the following holds. Let \(\components\) be the set of protected components that changed during
+the following holds. Let \(\components\) be the set of protected elements that changed during
 the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
@@
 secrets do not escape during a call {\em and} are not leaked during a return.
-This require us to be able to quantify over components that fail to preserve
+This require us to be able to quantify over elements that fail to preserve
 the change-together relation \(\diamond\).
@@
 be pairs of states. Their {\em unsafe set}, written
-\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all components
+\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all elements
 \(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
@@
 
-This last condition is the most interesting: we identify the components that
+This last condition is the most interesting: we identify the elements that
 contain leaked information on either side of the first pair of variant traces.
@@
 of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the components
+A call annotation now includes a predicate relating machine states to the elements
 that should be passed, expressed as offsets from the stack pointer.
 \[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P)~\mbox{\rm where}~P \subseteq \WORDS\]
-We extend the set of domains to include passed components,
+We extend the set of domains to include passed elements,
 which are labeled with the range of depths of the caller that has passed them.
@@
 \times (\rts \in \listT (\target)) \]
-The initial context again maps all components to \(\unsealed\).
+The initial context again maps all elements to \(\unsealed\).
 We straightforwardly describe how the sealing convention and passing predicate combine to
@@
 
-         At the matching return we just unseal the components at depth \(\depth\):
+         At the matching return we just unseal the elements at depth \(\depth\):
 
@@
 \definition The \emph{protected set} of a context, \(\prot_P(\domMap,\rts)\),
-is the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+is the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
 or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
@@
 \definition The \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
-is the set of components \(\component\) such that
+is the set of elements \(\component\) such that
 \(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
@@
 \definition Let the protected set of a context, \(\prot_S(\domMap,\rts)\),
-be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+be the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
 or \(\domMap ~ \component = \saved\).
@@
 \definition Let the secret set of a context, written \(\secret_S(\domMap,\rts)\),
-be the set of components \(\component\) such that
+be the set of elements \(\component\) such that
 \(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
@@
 The \emph{protected set} of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
-is the set of components \(\component\) such that
+is the set of elements \(\component\) such that
 \(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
@@
 The \emph{secret set} of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
-is the set of components \(\component\) such that
+is the set of elements \(\component\) such that
 \(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
@@
 concrete capability machine, an approach termed ``fully abstract overlay semantics.''
-The proof of full abstraction uses a standard notion of components, which
+The proof of full abstraction uses a standard notion of elements, which
 import and export functions through their interfaces.
@@
   every} caller should be protected from its callees.  By contrast,
-  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
+  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' elements
   that are not protected, because they are capable of dynamic behavior that can compromise
@@
   considered reasonable. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
-  {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
-  component is always reasonable, and so they
+  {\em reasonable element} (as defined in their section 4.3.3), as a reasonable
+  element is always reasonable, and so they
   would not guarantee its protection. In our setting, individual
@@
   Skorstengaard et al. \cite{SkorstengaardSTKJFP} support mutually distrustful
-  components; one can think of our hypothetical
+  elements; one can think of our hypothetical
   function as ``self-distrustful'' in that the safety of a well-behaved activation

### 2021-09-28 submission4/main.tex
Moved callee-saves

@@
 
-\paragraph{Limitations}
+\paragraph*{Limitations}
 
@@
 The Lazy Tagging and Clearing
-micro-policy from Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
-at which it was written, and enforces that it must be read from the
-same depth. However, this allows violations of observable stack safety!
+micro-policy from Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} in fact
+allows violations of observable stack safety!
 Our implementation of this policy, {\em PER\_DEPTH\_TAG}, has an error that we find
@@
 
-%Per-activation tagging resolves the issue because tags that persist between
-%calls will not become active again.
 Our abstract properties have allowed us to identify an enforcement mechanism as
@@
 without using unique per-activation identifiers.
+
+\subsection{Callee-saves Registers}
+\label{sec:callee}
+
+Another application of observational properties is characterizing a ``callee-saves''
+calling convention.
+Intuitively, a caller expects a callee-saves register to be unchanged on
+return, but it
+may change during the call, as long as it is restored. We similarly expect a callee-saves
+register to be secret from the callee, except when it is copied to the stack. Both scenarios
+reflect observational rather than stepwise stack safety.
+
+We begin by adding information to Call annotations defining which registers will be {\em saved}
+and which are {\em unsaved}. The stack pointer and the register containing the return address
+should never be saved. We then extend the set of domains with these values, for the first time applying
+(non-{\em outside}) domains to registers as well as addresses.
+%
+\[\begin{split}
+\codemap \in \CODEMAPS ::= & \WORDS \rightharpoonup \callmap(S)\\
+\mbox{\rm{where}} ~ & S \in \REGS \rightarrow (\unsaved + \saved) \\
+\end{split}\]
+
+\[\begin{split}
+\stackDom \in \STKDOMS ::= & \outside + \unsealed + \sealed{\depth} \\
+& + \unsaved + \saved
+\end{split}\]
+%
+\judgmenttwobrlong[ Call]
+                  {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
+                  {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
+                  {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
+                  {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+The remaining rules are the same as in the simple model.
+
+\definition Let the protected set of a context, \(\prot_S(\domMap,\rts)\),
+be the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+or \(\domMap ~ \component = \saved\).
+
+\definition Let the secret set of a context, written \(\secret_S(\domMap,\rts)\),
+be the set of elements \(\component\) such that
+\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
+\(\domMap ~ \component = \saved\).
+
+\definition
+%
+\textbf{\em Observational stack integrity} with callee-saves registers substitutes the above
+\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
+%
+\textbf{\em Observational stack confidentiality} additionally substitutes \(\secret_S\).
 
@@
 
-\subsection{Callee-saves Registers}
-\label{sec:callee}
-
-Under some calling conventions, rather than the caller saving the non-argument
-data from its registers
-and clearing them (and therefore implicitly taking responsibility for
-protecting this data),
-the callee is responsible for this. But of course we may still wish for
-those registers to be
-protected, just as they would be in a caller-saves convention.
-
-Intuitively, a caller expects a callee-saves register to be unchanged on
-return: it
-may change during the call, as long as it is restored. We similarly expect a callee-saves
-register to be secret from the callee, except when it is copied to the stack! Both scenarios
-reflect observational rather than stepwise stack safety.
-
-We begin by adding information to Call annotations defining which registers will be {\em saved}
-and which are {\em unsaved}. The stack pointer and the register containing the return address
-should never be saved. We then extend the set of domains with these values, for the first time applying
-(non-{\em outside}) domains to registers as well as addresses.
-%
-\[\begin{split}
-\codemap \in \CODEMAPS ::= & \WORDS \rightharpoonup \callmap(S)\\
-\mbox{\rm{where}} ~ & S \in \REGS \rightarrow (\unsaved + \saved) \\
-\end{split}\]
-
-\[\begin{split}
-\stackDom \in \STKDOMS ::= & \outside + \unsealed + \sealed{\depth} \\
-& + \unsaved + \saved
-\end{split}\]
-%
-\judgmenttwobrlong[ Call]
-                  {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
-                  {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
-                  {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
-                  {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-The remaining rules are the same as in the simple model.
-
-\definition Let the protected set of a context, \(\prot_S(\domMap,\rts)\),
-be the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
-or \(\domMap ~ \component = \saved\).
-
-\definition Let the secret set of a context, written \(\secret_S(\domMap,\rts)\),
-be the set of elements \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
-\(\domMap ~ \component = \saved\).
-
-\definition
-%
-\textbf{\em Observational stack integrity} with callee-saves registers substitutes the above
-\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
-%
-\textbf{\em Observational stack confidentiality} additionally substitutes \(\secret_S\).
-
-\paragraph*{Discussion}
-
-Observational properties are not only for lazy enforcement mechanisms---they reflect a useful
-loosening of the stepwise properties. Consider a single register \(\reg\) that is saved.
-It is copied to some stack address \(\addr\), violating stepwise confidentiality, and then
-overwritten, violating stepwise integrity. Then before the return, it is restored to its
-original value. In observational integrity terms, \(\reg\) is unchanged between call and return,
-and in observational confidentiality terms, both the original and variant values are unchanged
-from their originals, so that \(\reg\) is not in the \(\Delta\) set.  But \(\addr\) is in the
-\(\Delta\) set, so confidentiality additionally requires that the value of \(\reg\) does not
-leak through the callee-saves mechanism, for instance if the caller allocates but fails to
-initialize more memory.
-
 \subsection{Coroutines}

### 2021-09-28 submission4/main.tex
Reorganizing example 2 into example 1's code to save space

@@
 int f() {
-  int y;
+  int y, z;
+  z = *(&y-2);
   *(&y-2) = 0;
@@
 
-10 f:   add $1,%sp      ; allocate frame
-11      mov $0,-3(%sp)  ; violation!
-12      mov $5, %ra     ; set return value
-13      sub $1,%sp      ; deallocate frame
-14      ret
+10 f:   add $2,%sp      ; allocate frame
+11      mov -4(%sp),-1(%sp) ; violation!
+12      mov $0,-4(%sp)      ; violation!
+13      mov $5, %ra     ; set return value
+14      sub $1,%sp      ; deallocate frame
+15      ret
 \end{verbatim}
@@
 \\
-{\tt 11} &
+{\tt 12} &
 \memoryaddrs{16em}
@@
 \\
-{\tt 12} &
+{\tt 13} &
 \memoryaddrs{16em}
@@
 Figure~\ref{fig:int1} shows a program with a simple integrity violation. It consists of a {\tt main}
-function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
+function that performs a call to a function {\tt f}, which reads and then overwrites {\tt main}'s
 variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
@@
 {\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
-in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
+in this case 0. (Note that this does not include the return address, which is saved
+by the call instruction.)
+Sealing is the fundamental integrity contract between the caller and callee:
 {\tt main} expects that {\tt x} will be unchanged when it gets control back.
@@
 %cares about changes it can see after the return, not changes during execution.
-In this case, after instruction 11, {\tt f} has violated integrity by writing to {\tt x},
+In this case, after instruction 12, {\tt f} has violated integrity by writing to {\tt x},
 so the final return value produced by {\tt main} is wrong.
-%
-(The black rectangle represents the return address, which is saved by the call
-instruction itself, so it is not sealed.)
 
 \begin{figure*}
-\begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{verbatim}
-int f() {
-  int y,z;
-  y = 5;
-  z = *(&y-2);
-  return y+z;
-}
-\end{verbatim}
-}
-\end{subfigure}
-\begin{subfigure}[t]{.3\textwidth}
-{\small
-\begin{verbatim}
-10 f: add $2,%sp      ; allocate frame
-11     mov $5,-2(%sp)  ; assign into y
-12     mov -4(%sp),-1(%sp)  ; violation!
-13     mov -2(%sp),%ra ; calculate and
-14     add -1(%sp),%ra ; set return value
-15     sub $2,%sp      ; deallocate frame
-16     ret
-\end{verbatim}
-}
-\end{subfigure}
-\\
-\vspace*{0.3in}
-\begin{subfigure}{.59\textwidth}
 \centering
@@
 
-\end{subfigure}
 \caption{A confidentiality violation\ifspace\apt{Could save space by not
@@
 
+But before violating integrity, {\tt f} violates confidentiality by
+reading {\tt x}.
 Confidentiality states intuitively that a callee is insensitive to
@@
 
-In Figure~\ref{fig:conf1} we see the callee, {\tt f}, violating confidentiality
-by reading {\tt main}'s variable {\tt x}. The trace
-shows how noninterference applies to this scenario. At instruction 10, we
+In Figure~\ref{fig:conf1} we see how noninterference applies to this scenario.
+At instruction 10, we
 create an imaginary {\em variant}
@@
 in place of the actual values ($42$, ...) that should
-be hidden from {\tt f}; we show the evolving stack memory of the variant in a second, parallel column. At instruction 12, we note that the callee's first variable, {\tt y}, has
-changed to 5 in both the primary trace and the variant. The primary and variant correctly agree on
-the value in the changed location. But at instruction 13,
-the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
-in the variant trace it might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
+be hidden from {\tt f}; we show the evolving stack memory of the variant in a second, parallel column.
+At instruction 11, we note that the callee's variable, {\tt z}, has
+changed to 42 in the primary trace, but
+in the variant trace it instead might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
 has behaved differently than the primary, violating confidentiality.
@@
 with additional support for linear capabilities.
-Stack safety is defined as the conjunction of local state encapsulation (LSE) and
+They define Stack safety as the conjunction of local state encapsulation (LSE) and
 well-bracketed control flow (WBCF).
@@
 concrete capability machine, an approach termed ``fully abstract overlay semantics.''
-The proof of full abstraction uses a standard notion of elements, which
-import and export functions through their interfaces.
-
 We improve upon Skorstengaard et al. \cite{SkorstengaardSTKJFP} in two significant ways:
@@
   that are not protected, because they are capable of dynamic behavior that can compromise
-  their own security. Unreasonable code can be viewed as adversarial, or merely as
-  buggy. One can imagine a function that, when called with the wrong
-  inputs, behaves unreasonably, but under most circumstances can be
-  considered reasonable. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
+  their own security. Consider a function that, when called with the wrong
+  inputs, behaves unreasonably. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
   {\em reasonable element} (as defined in their section 4.3.3), as a reasonable
-  element is always reasonable, and so they
-  would not guarantee its protection. In our setting, individual
+  element is always reasonable. In our setting, individual
   well-behaved activations of the buggy function would be protected.
-  Skorstengaard et al. \cite{SkorstengaardSTKJFP} support mutually distrustful
-  elements; one can think of our hypothetical
-  function as ``self-distrustful'' in that the safety of a well-behaved activation
-  is not compromised by another, ill-behaved activiation of the same function.
-
-\item Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s approach to proving safety via a fully
+
+\item Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s fully
   abstract overlay
-  semantics relies on the definition of a ``stack-safe-by-construction''
-  abstract machine that is closely tied to an underlying enforcement
-  mechanism and actual machine architecture.
-  %the the abstract machine being architecturally similar to
-  %      the concrete one (beyond the specific call and return semantics that
-  %      distinguish them).
+  semantics rely on the definition of a ``stack-safe-by-construction''
+  abstract machine that is closely tied to the underlying machine architecture.
   While the fundamental approach can no doubt be ported to other architectures and enforcement
@@
   safety-by-construction property must be checked afresh.
-  Our properties are more generic: since we abstract over most of the machine
-  state as well as the sealing and return conventions, they apply directly
+  Since we abstract over most of the machine
+  state as well as the sealing and return conventions, our properties apply directly
   to a range of architectures, code features, and enforcement mechanisms.
@@
   to our base domain model.
-  %% While it may well be possible to extend
-  %% \citet{SkorstengaardSTKJFP}'s approach to cover these features.
-  %% abstract machine to pass arguments on the stack, or to support callee-saves
-  %% registers. It might even be viable to construct a safe-by-construction abstract
-  %% coroutine machine in the same style, but this is hypothetical.
 
@@
 %
-(1) {\em Intuitiveness}. We contend that our property definitions are at least as
+(1) {\em Intuitiveness}. Our property definitions are at least as
 intuitive as theirs, while being phrased in interestingly different terms.
@@
 
-Finally, we would propose one more criterion: having a generic definition
-of stack safety that is directly reusable to evaluate multiple machines and
+Finally, we would propose one more criterion: reusability in evaluating multiple machines and
 enforcement mechanisms. They have used their approach to give strong validation
@@
 %
-Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
-attacks against stack data are in scope (but not side channels or hardware
-attacks), and study the protection of stack data through three families of
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}
+study the protection of stack data through three families of
 micro-policies that tag stack objects with a pair of frame and object identifiers
@@
 
-We evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
+We evaluate {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
   Clearing} against our properties, finding that {\em Depth Isolation} does
 indeed enforce stepwise stack safety and therefore observable stack
-safety. We also found that neither lazy micro-policy actually enforces
+safety. We also find that neither lazy micro-policy actually enforces
 stack safety, and we proposed a fix.
@@
 \label{sec:future}
-
-We introduced a characterization against which stack safety enforcement
-mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
 

### 2021-09-29 submission4/main.tex
font = 10
Trim space

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[12pt,conference]{ieeetran}%\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[10pt,conference]{ieeetran}%\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
@@
 program counter {\PCname} and the stack pointer \(\SP\).
+$\component \in \COMPONENTS ::= \WORDS + \REGS$.
 A {\em machine state} is a map from elements to values.
-
-\begin{minipage}{.45\columnwidth}
-  \[\component \in \COMPONENTS ::= \WORDS + \REGS\]
-\vspace{\belowdisplayskip}
-\end{minipage}
-\begin{minipage}{.45\columnwidth}
-  \[\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\]
-\vspace{\belowdisplayskip}
-\end{minipage}
+$\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS$.
 
@@
 
-\section{Case Study: Do Micro-Policies Enforce Stack Safety?}
+\section{Case Study: Micro-Policies}
 \label{sec:enforcement}
@@
 confidentiality. We use a Coq specification of the RISC-V
-architecture\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
+architecture\bcp{change from footnote to citation :-)}\footnote{Available at
+  \url{https://github.com/mit-plv/riscv-coq}.}
 %\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}

### 2021-09-29 submission4/main.tex
Space hacks

@@
 We use these properties to validate the stack-safety {\em micro-policies}
-proposed by Roessler and DeHon%~\cite{DBLP:conf/sp/RoesslerD18}.
+proposed by Roessler and DeHon. %~\cite{DBLP:conf/sp/RoesslerD18}.
 Specifically, we check (with
@@
 
-We now define our machine and enforcement model precisely.
+% We now define our machine and enforcement model.
 
@@
 and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first element on which \(f\) holds.
+\(\MPCT\) ending with the first element on which \(f\) holds.  \bcp{Any
+  chance of getting two of these on the same line?}
 
@@
 that doesn't share between caller and callee: arguments and results are passed
-in registers. The code map just marks the locations of calls.
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap\]
+in registers. The code map just marks the locations of calls:
+$\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap$.
+
 To accommodate different machine architectures and different calling
@@
 ``return target list,'' which is a stack of targets, one for each caller awaiting
-return.
+return.  \bcp{The first displayed line can be inlined in the text. (I've
+  done a few others like this already.  There are probably more.)}
 %
@@
 section. We validate this claim
-using a property-based random testing framework.
+using property-based random testing.
 
@@
 $n$, or the tag $\tagNoDepth$ for stack locations that are currently
-unused (but could potentially be used in the future).
+unused (but could be used in the future).
 %
@@
 goal of evaluating real enforcement mechanisms for real machine architectures.
-Formal proof remains as possible future work.
+Formal proof remains future work.
 

### 2021-09-29 submission4/main.tex
Figure 2 to one column

@@
 10 f:   add $2,%sp      ; allocate frame
-11      mov -4(%sp),-1(%sp) ; violation!
-12      mov $0,-4(%sp)      ; violation!
+11      mov -4(%sp),-1(%sp) ; violation 1!
+12      mov $0,-4(%sp)      ; violation 2!
 13      mov $5, %ra     ; set return value
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
@@
 \label{fig:conf1}
-\end{figure*}
+\end{figure}
 
@@
 \end{subfigure}
-%\begin{subfigure}[t]{.68\textwidth}
-%  {\small
-%\begin{verbatim}
-%1 main:  add $1,%sp      ; allocate frame
-%2        call f
-%3        mov %ra,-1(%sp) ; store x
-%4        call g
-%5        sub $1,%sp      ; deallocate frame
-%6        ret
-%
-%10 f:    add $1,%sp      ; allocate frame
-%11       mov $5,-1(%sp)  ; initialize y
-%12       mov -1(%sp),%ra ; set return value
-%13       sub $1,%sp      ; deallocate frame
-%14       ret
-%
-%20 g:    add $1,%sp      ; allocate frame
-%21       mov -1(%sp),%ra ; violation!
-%22       sub $1,%sp      ; deallocate frame
-%23       ret
-%}
-%\end{verbatim}
-%}
-%\end{subfigure}
-%\end{center}
-%\begin{subfigure}{.65\textwidth}
-%\begin{center}
-%\begin{tabular}{l r | l}
-%  {\tt 2} &
-%  \multicolumn{2}{c}{
-%    \memoryaddrs{8em}
-%    \memory{3}{\unsealc}
-%    ~$\cdots$
-%    \vspace{.5em}
-%  } \\
-%  {\tt 11} &
-%  \memoryaddrs{16em}
-%  \memory{1}{\unsealc}
-%  \memory{1}{\retptrc}
-%  \memory{1}{\unsealc}
-%  ~$\cdots$
-%  &
-%  \memoryaddrs{16em}
-%  \memory{1}{\unsealc}
-%  \memory{1}{\retptrc}
-%  \memory{1}{\unsealc}
-%  ~$\cdots$
-%  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-%  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
-%  \\
-%  {\tt 12} &
-%  \memoryaddrs{16em}
-%  \memory{1}{\unsealc}
-%  \memory{1}{\retptrc}
-%  \memory{1}{\unsealc}
-%  ~$\cdots$
-%  \MemoryLabel{-6em}{0.75em}{5}
-%  &
-%  \memoryaddrs{16em}
-%  \memory{1}{\unsealc}
-%  \memory{1}{\retptrc}
-%  \memory{1}{\unsealc}
-%  ~$\cdots$
-%  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
-%  \MemoryLabel{-6em}{0.75em}{\(5\)}
-%  \\
-%  {\tt 4} &
-%  \multicolumn{2}{c}{
-%    \memoryaddrs{8em}
-%    \memory{1}{\unsealc}
-%    \memory{1}{\retptrc}
-%    \memory{1}{\unsealc}
-%    ~$\cdots$
-%    \MemoryLabel{-14em}{0.75em}{5}
-%    \MemoryLabel{-6em}{0.75em}{5}
-%    \vspace{.5em}
-%  }
-%  \\
-%  {\tt 22} &
-%  \memoryaddrs{16em}
-%  \memory{1}{\unsealc}
-%  \memory{1}{\retptrc}
-%  \memory{1}{\unsealc}
-%  ~$\cdots$
-%  \MemoryLabel{-15em}{0.75em}{5}
-%  \MemoryLabel{-6em}{0.75em}{\bf 5}
-%  &
-%  \memoryaddrs{16em}
-%  \memory{1}{\unsealc}
-%  \memory{1}{\retptrc}
-%  \memory{1}{\unsealc}
-%  ~$\cdots$
-%  \MemoryLabel{-15em}{0.75em}{\(v_3\)}
-%  \MemoryLabel{-6em}{0.75em}{\(v_4\)}
-%  \MemoryLabel{-2em}{0.75em}{90}
-%  \\
-%\end{tabular}
-%\end{center}
-%
-%\vspace{\abovedisplayskip}
-%
-%\end{subfigure}
 \caption{Another confidentiality violation}
@@
 
-The building blocks of the machine are {\em values} and {\em addresses}.
-Both are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively.
-%
-Our machine states are composed of {\em elements} \(\component\),
-which are either addresses or register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include two special-purpose registers: the
-program counter {\PCname} and the stack pointer \(\SP\).
-$\component \in \COMPONENTS ::= \WORDS + \REGS$.
-A {\em machine state} is a map from elements to values.
-$\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS$.
-
-We have a total step function between machine states, written \(\mach \stepsto \mach'\).
-
-We now formalize the model of enforcement that we introduced in Section~\ref{sec:threat}.
-A {\em policy} consists of a set of policy
-states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
-\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
-it is undefined on input configurations that correspond to a policy fault. The policy
-can see the machine state, but cannot modify it. The policy also provides a
-well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
-initial states that are properly configured.
-%
-%A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
-%\Cref{sec:enforcement}.---APT: We just said this.
-
-We lift the policy step function to operate on such pairs
-by combining it with the regular step function for machine states.
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
-               \polstep \pol'\)}
-            {\(\mpstate{\mach}{\pol} \stepstopol
-               \mpstate{\mach'}{\pol'}\)}
-
-\noindent
-Note that the base machine has a total step function, while the policy-enhanced machine
-refines it to a partial function. The extra policy state never changes behavior,
-but it can cause
-premature termination, representing failstop behavior.
+The building blocks of the machine are {\em values}, {\em addresses}, and {\em registers}.
+Values and addresses are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
+\(\addr\), respectively. Register names (\(\reg\)) drawn from some set
+\(\REGS\), which is assumed to include a special-purpose register, the
+program counter {\PCname}. Collectively addresses and registers are {\em components} \(\component\)
+in the set \(\COMPONENTS ::= \WORDS + \REGS\).
+%
+A {\em machine state} \(\mach\) is drawn from a set \(\MACHS\) and supports a mapping
+from components to values, written
+\(\mach[\component] ::= \MACHS \rightarrow \COMPONENTS \rightarrow \WORDS \).
+
+We have a partial step function between machine states, written \(\mach \stepsto \mach'\), which
+is undefined on input states that correspond to a policy fault. So, a security mechanism is
+represented by the capacity of the machine to ``fail-stop,'' as well as by whatever initial
+well-formedness conditions are required for the mechanism to function properly.
 
@@
 A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
-\(\mach,\pol,\context\) such that \(\mach,\pol,\context \stepstocon \mpcstatename'\),
+\(\mach,\context\) such that \(\mach,\context \stepstocon \mach',\context'\),
 and any element \(\component \in \prot(\context)\),
-\(\pi_\mach(\mpcstatename') ~ \component = \mach ~ \component\).
+\(\mach'[\component] = \mach[\component]\).
 
@@
 is the set of elements \(\component\)
-such that \(\mach ~ \component \not = \mach' ~ \component\).
+such that \(\mach[\component] \not = \mach' ~ \component\).
 
@@
 \textbf{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
-where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
+where \(\head(\MPCT) = \mach\context\) and \(\last(\MPCT) = \mach',\context'\),
 the following holds. Let \(\components\) be the set of protected elements that changed during
 the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
-and let \((\mach',\pol',\context') \hookrightarrow \MPCT'\).
+and let \((\mach',\context') \hookrightarrow \MPCT'\).
 Then for any \(\nach\) such that
 \(\mach' \approx_{\components'} \nach\), and its induced trace
-\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
+\((\nach,\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
 

### 2021-09-29 submission4/main.tex
Moved initial testing

@@
 
-\paragraph{Well-bracketed control flow.} Together,
+\paragraph*{Well-bracketed control flow.} Together,
 integrity and confidentiality protect a caller's data until its callee
@@
 $B$'s stack frames are improperly read or written by $C$.  But it is still
-arguably unsafe in general.  For example, suppose that $B$'s continuation
-after the call to $C$ takes $C$'s result, doubles it, and returns this to
-$A$.  If $A$'s correct behavior after the call to $B$ returns depends on the
-result it receives being a even number---i.e., $A$ is trusting that $B$'s
-continuation will execute after $C$ returns, perhaps because $A$ and $B$ are
-written by the same person while $C$ is an untrusted library---then $C$ can
-cause it to behave incorrectly, possibly even insecurely, by returning an
-odd result directly to $A$.
+arguably unsafe if $A$'s programmer depends on $B$ completing its execution before
+it returns.
 %
@@
 will be a meaningful definition of stack safety for that machine. It must at least be a
-register machine with a program counter (\(\PCname\)), and there must be a well-defined
+register machine with a program counter (\(\PCname\)) and there must be a well-defined
 mechanism for determining the bounds of a stack frame, and for recognizing when a call
@@
 are equally feasible. A machine may use its non-stack memory in any way, including
-as a heap, as long as the stack has static bounds.
+as a heap, as long as the stack has static bounds. It might also possess additional
+state beyond its registers and memory addresses, such as the state of a hardware
+enforcement mechanism.
 
@@
 including those that do not failstop, such as
-code rewriting techniques that prevent stack safety violations directly.
-
-An enforcement model may additionally
-apply a ``well-formedness'' criterion on initial states, ensuring that
-they are set up properly for the enforcement mechanism in question.
-When validating Roessler and DeHon's \cite{DBLP:conf/sp/RoesslerD18}'s micro-policies,
-well-behaved calls and returns are implemented by fixed sequences of instructions
-which the tag policy will force to be executed from start to finish,
-termed {\em blessed sequences}. Well-formedness requires that all annotated
-instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
-This is all we need for the policy to protect itself: it can then failstop if it would
+code rewriting techniques that prevent stack safety violations directly by
+restricting the set of possible initial states.
+
+Most enforcement mechanisms make assumptions about their initial conditions,
+so we assume that any given machine will restrict its conditions a ``well-formedness''
+criterion on initial states, ensuring that they are set up properly. In the case of
+Roessler and DeHon's \cite{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-formedness
+requires that all annotated instructions will be situated at the appropriate instruction
+of a {\em blessed sequence} -- a series of instructions that are tagged to enforce correct
+behavior. This is all we need for the policy to protect itself: it can then failstop if it would
 execute the blessed sequence out of order.
@@
 We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code, subject to whatever well-formedness conditions the
-enforcement mechanism demands. But hardware and timing attacks are out of scope, as
-are attacks on the heap---though we believe that the work of
-Azevedo de Amorim et al.\cite{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
-to protect both stack and heap.
+arbitrary machine code, subject to whatever ``well-formedness'' conditions the
+enforcement mechanism demands. But hardware and timing attacks are out of scope,
+and our properties are termination insensitive as described above.
 
@@
 
-We project out the machine state of an MPC-state with \(\pi_\mach\)
+We project out the machine state of an MC-state with \(\pi_\mach\)
 and the context with \(\pi_\context\).
@@
 
-\section{Case Study: Micro-Policies}
+\subsection{Case Study: Micro-Policies}
 \label{sec:enforcement}
@@
 the formal stack-safety property described in the previous
-section. We validate this claim
-using property-based random testing.
+section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not..
+We later validate this claim using property-based random testing.
 
@@
 
+Tags in the PIPE system are not accessible to application code, being separated
+from their associated values at the hardware level. So they are not modeled
+as part of the values, and therefore our properties make no claims about
+their behavior.
+
 Efficient execution of these micro-policies relies on hardware acceleration,
@@
 
-%\paragraph{A Conservative Tag Policy}
+%\paragraph*{A Conservative Tag Policy}
 %\label{sec:conservative}
@@
 
-\subsection{Validation through Random Testing}
-\label{sec:testing}
-
-There are several ways to evaluate whether an enforcement mechanism enforces
-stack safety properties. Ideally such validation would be done through formal proof over
-the semantics of the enforcement-augmented machine.
-However, while there are no fundamental barriers to producing such a proof,
-it would be considerable work to carry out for a full ISA like RISC-V and
-complex enforcement mechanism like the Depth Isolation micro-policy.
-We therefore choose to validate the micro-policy of the previous section by
-systematically \emph{testing} that it satisfies our properties.
-This focus on testing is better aligned with our immediate
-goal of evaluating real enforcement mechanisms for real machine architectures.
-Formal proof remains future work.
-
-We validate the micro-policy of the previous section by
-systematically testing that it satisfies eager stack integrity and
-confidentiality. We use a Coq specification of the RISC-V
-architecture\bcp{change from footnote to citation :-)}\footnote{Available at
-  \url{https://github.com/mit-plv/riscv-coq}.}
-%\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}
-and extend it with a
-runtime monitor implementing the stack-safety
-micro-policy. We
-  chose the Coq proof assistant as the setting for our implementation
-  to ensure that our coinductive trace
-  definitions are well-formed, and reason about them, and to leverage the power of the
-  QuickChick property-based testing framework~\cite{Pierce:SF4}.
-
-To use QuickChick, we build random test-case generators that produce
-% \begin{enumerate}
-% \item
-(1) an initial RISC-V machine state, including most notably
-  the program to be executed;
-  % \item
-(2)
-  an initial policy state, tagging instructions corresponding to
-  blessed call or return sequences appropriately, while marking
-  all potential stack locations as $\tagNoDepth$; and
-% \item
-(3)
- a code map identifying the locations of calls.
-% \end{enumerate}
-
-That is, our generators need to produce not only a low-level program
-that exhibits interesting control-flow patterns, but also the policy
-meta-information and code annotations that would normally be produced
-by the compiler.
-
-%\paragraph*{Generators}
-%
-%To write such generators we build on the work of
-%Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
-%introduced {\em generation by execution} to produce progams that lead
-%machines towards interesting behaviors. Generation by
-%execution receives as an input a partially instantiated machine state
-%and attempts to generate an instruction (or a sequence of instructions
-%such as a blessed sequence) that makes sense locally (e.g., jumps go
-%to a potentially valid code location and loads read from a
-%potentially valid stack location). Then we step the machine and repeat
-%the process until we generate or execute some target number of
-%instructions, or reach a point where the machine cannot step
-%any more.
-%
-%We extend this technique to keep track of the control flow behavior of
-%the program being generated: each time a call or return sequence is
-%generated, we ensure that the appropriate policy tags and code
-%annotations are set for the entry or return points.% At the same time,
-%%we allow the generation to sometimes relax those constraints,
-%%introducing potentially ill-formed flows: this causes our programs
-%%to failstop when executed in conjunction with the policy monitor,
-%%but also allows for revealing errors in our setup.\apt{??}
-%
-%In addition, we need to further extend this technique to handle the nested
-%nature of our confidentiality property: rather than just generating two
-%initial machines that are variants of one another and letting them
-%execute to test for noninterference, we generate a new variant
-%{\em every time a call is made} and check confidentiality for the
-%subtrace produced from that variant state until its corresponding
-%return. As a result, a ``single'' confidentiality test compactly
-%checks the confidentiality of multiple nested calls.
-
-\paragraph*{Findings}
-
-To ensure the effectiveness of testing against our formal properties, we
-use {\em mutation testing}~\cite{JiaH11}. In mutation testing, we inject errors
-(mutations) in a program that should cause the property of interest (here,
-stack safety) to fail, and ensure that the testing framework can find
-them. The bugs we use for our evaluation are either artificially generated
-by us (deliberately weakening the micro-policy in ways that we expect
-should break its guarantees), or actual bugs that we discovered through
-testing our implementation. We elaborate on some such bugs below.
-
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c|c}
-  Bug & Property Violated & MTTF (s) & Tests \\
-  \hline
-      {\em LOAD\_NO\_CHECK}  & Confidentiality & 24.2 & 13.3 \\
-      {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
-      {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous eager policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:bug-table}
-\end{table}
-
-The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
-Table~\ref{tab:bug-table}, along with the average number of tests
-it took to find the failure. Experiments were run in a desktop
-machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
-% \leo{Something like: There is really no baseline to compare this
-  %against, but include it here for the future/reasons?}
-%
-The artificially injected bugs weaken the tag propagation
-and checking mechanism.
-%
-For example, when loading from a stack location, the correct
-micro-policy needs to enforce that the tag on the location being read
-is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
-current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
-that restriction by not checking the depth equality (row {\em
-  LOAD\_NO\_CHECK}).
-%
-Similarly, when storing to a stack location, the correct micro-policy
-needs to ensure that the tag on the memory location is either
-$\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing that constraint causes violations to the integrity
-property (row {\em STORE\_NO\_CHECK}).
-
-Naturally, testing also revealed a number of errors in our
-implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code;
-%\bcp{right?}\leo{yeah}
-we re-implemented it in Coq targeting RISC-V).  These errors range
-from trivial typos to ones that require an intriguingly complex setup
-to appear.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
-sequence %/policy combination\apt{??}
-did not initialize all locations for the
-newly allocated stack frame correctly, but left some of them as
-$\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured.
-The smallest counterexample requires calling a function {\tt f},
-which fails to initialize some of its frame during the blessed sequence,
-but writes into an uninitialized location $l$ later, treating \(l\) as outside
-the stack. Then {\tt f} calls a further function {\tt g} (which should have
-the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
-which is allowed because the enforcement mechanism still has
-$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
-%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
-%  and the write from {\tt f} was already a violation.}
-
-\section{Observable Stack Safety}
-\label{sec:lazy}
-
-The conservative policy described in the previous section, while
-testably enforcing stack safety, is rather slow in
+\paragraph*{A Lazy Alternative}
+
+The conservative policy described above is rather slow in
 some important practical cases. Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18} observe
@@
 This \emph{lazy policy} admits more a efficient implementations, but
-they deliberately allow temporary violations of stack integrity.
+deliberately allows temporary violations of stack integrity.
 The natural question then is: how do we characterize the
@@
 
-Beyond this one enforcement mechanism, some code features may involve
+\section{Observable Stack Safety}
+\label{sec:lazy}
+
+We answer the above questions by introducing a notion of observations to our model.
+This has uses beyond this one enforcement mechanism, as some code features may involve
 moving data, overwriting its original location,
@@
 
-So we extend our model with a notion of observations. We abstract over an
+We abstract over an
 observation type \(\obs \in \OBSS\), containing at minimum the silent observation
@@
 
+\section{Validation through Random Testing}
+\label{sec:testing}
+
+There are several ways to evaluate whether an enforcement mechanism enforces
+stack safety properties. Ideally such validation would be done through formal proof over
+the semantics of the enforcement-augmented machine.
+However, while there are no fundamental barriers to producing such a proof,
+it would be considerable work to carry out for a full ISA like RISC-V and
+complex enforcement mechanism like the Depth Isolation micro-policy.
+We therefore choose to validate the micro-policy of the previous section by
+systematically \emph{testing} that it satisfies our properties.
+This focus on testing is better aligned with our immediate
+goal of evaluating real enforcement mechanisms for real machine architectures.
+Formal proof remains future work.
+
+We validate the micro-policy of the previous section by
+systematically testing that it satisfies eager stack integrity and
+confidentiality. We use a Coq specification of the RISC-V
+architecture\bcp{change from footnote to citation :-)}\footnote{Available at
+  \url{https://github.com/mit-plv/riscv-coq}.}
+%\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}
+and extend it with a
+runtime monitor implementing the stack-safety
+micro-policy. We
+  chose the Coq proof assistant as the setting for our implementation
+  to ensure that our coinductive trace
+  definitions are well-formed, and reason about them, and to leverage the power of the
+  QuickChick property-based testing framework~\cite{Pierce:SF4}.
+
+To use QuickChick, we build random test-case generators that produce
+% \begin{enumerate}
+% \item
+(1) an initial RISC-V machine state, including most notably
+  the program to be executed;
+  % \item
+(2)
+  an initial policy state, tagging instructions corresponding to
+  blessed call or return sequences appropriately, while marking
+  all potential stack locations as $\tagNoDepth$; and
+% \item
+(3)
+ a code map identifying the locations of calls.
+% \end{enumerate}
+
+That is, our generators need to produce not only a low-level program
+that exhibits interesting control-flow patterns, but also the policy
+meta-information and code annotations that would normally be produced
+by the compiler.
+
+%\paragraph*{Generators}
+%
+%To write such generators we build on the work of
+%Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
+%introduced {\em generation by execution} to produce progams that lead
+%machines towards interesting behaviors. Generation by
+%execution receives as an input a partially instantiated machine state
+%and attempts to generate an instruction (or a sequence of instructions
+%such as a blessed sequence) that makes sense locally (e.g., jumps go
+%to a potentially valid code location and loads read from a
+%potentially valid stack location). Then we step the machine and repeat
+%the process until we generate or execute some target number of
+%instructions, or reach a point where the machine cannot step
+%any more.
+%
+%We extend this technique to keep track of the control flow behavior of
+%the program being generated: each time a call or return sequence is
+%generated, we ensure that the appropriate policy tags and code
+%annotations are set for the entry or return points.% At the same time,
+%%we allow the generation to sometimes relax those constraints,
+%%introducing potentially ill-formed flows: this causes our programs
+%%to failstop when executed in conjunction with the policy monitor,
+%%but also allows for revealing errors in our setup.\apt{??}
+%
+%In addition, we need to further extend this technique to handle the nested
+%nature of our confidentiality property: rather than just generating two
+%initial machines that are variants of one another and letting them
+%execute to test for noninterference, we generate a new variant
+%{\em every time a call is made} and check confidentiality for the
+%subtrace produced from that variant state until its corresponding
+%return. As a result, a ``single'' confidentiality test compactly
+%checks the confidentiality of multiple nested calls.
+
+\paragraph*{Findings}
+
+To ensure the effectiveness of testing against our formal properties, we
+use {\em mutation testing}~\cite{JiaH11}. In mutation testing, we inject errors
+(mutations) in a program that should cause the property of interest (here,
+stack safety) to fail, and ensure that the testing framework can find
+them. The bugs we use for our evaluation are either artificially generated
+by us (deliberately weakening the micro-policy in ways that we expect
+should break its guarantees), or actual bugs that we discovered through
+testing our implementation. We elaborate on some such bugs below.
+
+\begin{table}[]
+\centering
+\begin{tabular}{c|c|c|c}
+  Bug & Property Violated & MTTF (s) & Tests \\
+  \hline
+      {\em LOAD\_NO\_CHECK}  & Confidentiality & 24.2 & 13.3 \\
+      {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
+      {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
+\end{tabular}
+\vspace*{1em}
+\caption{MTTF for finding bugs in erroneous eager policy enforcement mechanisms}
+\vspace*{-2em}
+\label{tab:bug-table}
+\end{table}
+
+The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
+Table~\ref{tab:bug-table}, along with the average number of tests
+it took to find the failure. Experiments were run in a desktop
+machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
+% \leo{Something like: There is really no baseline to compare this
+  %against, but include it here for the future/reasons?}
+%
+The artificially injected bugs weaken the tag propagation
+and checking mechanism.
+%
+For example, when loading from a stack location, the correct
+micro-policy needs to enforce that the tag on the location being read
+is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
+current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
+that restriction by not checking the depth equality (row {\em
+  LOAD\_NO\_CHECK}).
+%
+Similarly, when storing to a stack location, the correct micro-policy
+needs to ensure that the tag on the memory location is either
+$\tagNoDepth$ or has again the same depth as the current $\PCname$
+tag. Relaxing that constraint causes violations to the integrity
+property (row {\em STORE\_NO\_CHECK}).
+
+Naturally, testing also revealed a number of errors in our
+implementation of the enforcement mechanism (the original was written in C++
+and targeted ARM machine code;
+%\bcp{right?}\leo{yeah}
+we re-implemented it in Coq targeting RISC-V).  These errors range
+from trivial typos to ones that require an intriguingly complex setup
+to appear.  The most interesting bug (included in the table as row
+{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
+sequence %/policy combination\apt{??}
+did not initialize all locations for the
+newly allocated stack frame correctly, but left some of them as
+$\tagNoDepth$. This allowed for a potential integrity violation, but
+only if a rather complicated sequence of events occured.
+The smallest counterexample requires calling a function {\tt f},
+which fails to initialize some of its frame during the blessed sequence,
+but writes into an uninitialized location $l$ later, treating \(l\) as outside
+the stack. Then {\tt f} calls a further function {\tt g} (which should have
+the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
+which is allowed because the enforcement mechanism still has
+$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
+%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
+%  and the write from {\tt f} was already a violation.}
+
 \section{Related Work}

### 2021-09-29 submission4/main.tex
Appendicitis

@@
 
-Our abstract properties have allowed us to identify an enforcement mechanism as
+Our properties have allowed us to identify an enforcement mechanism as
 not really stack safe, and to validate a possible fix.
@@
 Another application of observational properties is characterizing a ``callee-saves''
-calling convention.
-Intuitively, a caller expects a callee-saves register to be unchanged on
-return, but it
-may change during the call, as long as it is restored. We similarly expect a callee-saves
+calling convention. Intuitively, a caller expects a callee-saves register to be unchanged on
+return, but it may change during the call, as long as it is restored. We similarly expect a callee-saves
 register to be secret from the callee, except when it is copied to the stack. Both scenarios
@@
 should never be saved. We then extend the set of domains with these values, for the first time applying
-(non-{\em outside}) domains to registers as well as addresses.
-%
-\[\begin{split}
-\codemap \in \CODEMAPS ::= & \WORDS \rightharpoonup \callmap(S)\\
-\mbox{\rm{where}} ~ & S \in \REGS \rightarrow (\unsaved + \saved) \\
-\end{split}\]
-
-\[\begin{split}
-\stackDom \in \STKDOMS ::= & \outside + \unsealed + \sealed{\depth} \\
-& + \unsaved + \saved
-\end{split}\]
-%
-\judgmenttwobrlong[ Call]
-                  {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
-                  {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
-                  {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
-                  {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-The remaining rules are the same as in the simple model.
-
-\definition Let the protected set of a context, \(\prot_S(\domMap,\rts)\),
-be the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
-or \(\domMap ~ \component = \saved\).
-
-\definition Let the secret set of a context, written \(\secret_S(\domMap,\rts)\),
-be the set of elements \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
-\(\domMap ~ \component = \saved\).
-
-\definition
-%
-\textbf{\em Observational stack integrity} with callee-saves registers substitutes the above
-\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
-%
-\textbf{\em Observational stack confidentiality} additionally substitutes \(\secret_S\).
+(non-{\em outside}) domains to registers as well as addresses. On each call, we save precisely those
+registers given by the annotation, and the rest are unsaved; {\em saved} registers are treated as part
+of the protected and secret sets for purposes of \textbf{\em Observational stack integrity} and
+\textbf{\em Observational stack confidentiality}, respectively. For full formal definitions, see
+Appendix \ref{app:calleesave}.
 
@@
 %  a technical obstacle, it's that the behavior is tied to things we don't model.}
-\paragraph*{Formalization of variable passing.}
-We add extra information to our call
-annotations to reflect that they may now pass variables on the stack, and that the locations
-of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the elements
-that should be passed, expressed as offsets from the stack pointer.
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P)~\mbox{\rm where}~P \subseteq \WORDS\]
-We extend the set of domains to include passed elements,
-which are labeled with the range of depths of the caller that has passed them.
-\[\begin{split}
-\stackDom \in \STKDOMS ::= & \outside + \unsealed\\
-& + \sealed{\depth} + \passed{\depth,\depth'}
-\end{split}\]
-The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\rts \in \listT (\target)) \]
-The initial context again maps all elements to \(\unsealed\).
-We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map at a call from depth \(\depth\).
-
-\judgmentthree[]
-              {\(\domMap ~ \component = \unsealed\)}
-              {\(\sealcon ~ \mach[\component]\)}
-              {\(\component - (\mach ~ \rsp) \not \in P\)}
-              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \sealed{\depth}\)}
-\judgmentthree[]
-              {\(\domMap ~ \component = \unsealed\)}
-              {\(\sealcon ~ \mach[\component]\)}
-              {\(\component - (\mach ~ \rsp) \in P\)}
-              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth,\depth+1}\)}
-\judgmentthree[]
-              {\(\domMap ~ \component = \passed{\depth',\depth}\)}
-              {\(\sealcon ~ \mach[\component]\)}
-              {\(\component - (\mach ~ \rsp) \in P\)}
-              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth',\depth+1}\)}
-\judgment[]
-         {}
-         {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \domMap ~ \component\)}
-
-         At the matching return we just unseal the elements at depth \(\depth\):
-
-\judgment[]
-         {\(\domMap ~ \component = \sealed{\depth}\)}
-         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-\judgment[]
-         {\(\domMap ~ \component = \passed{\depth-1,\depth}\)}
-         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-\judgmenttwo[]
-           {\(\domMap ~ \component = \passed{\depth',\depth}\)}
-           {\(\depth' < \depth\)}
-           {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-\judgment[]
-         {}
-         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-
-Now we can inductively define the context update function, as before.
-
-\judgmenttwo[ CallSeal]
-            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
-            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmenttwo[ ReturnFound]
-            {\(\mach \stepsto \mach' \in \target\)}
-            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
-
-\vspace*{-1ex}
-\judgmenttwo[ ReturnRec]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3ex}
-\judgment[ Default]
-         {}
-         {\(\underscore,\context \stepsto_C \context\)}
-
-\definition The \emph{protected set} of a context, \(\prot_P(\domMap,\rts)\),
-is the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
-or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
-\(|\rts| > \depth'\).
-
-\definition The \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
-is the set of elements \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
-\(\domMap ~ \component = \passed{\depth,\depth'}\)
-where \(|\rts| < \depth\) or \(|\rts| > \depth'\).
-
-\definition
-For each of the \textbf{\em stepwise} and \textbf{\em observational} versions of
-\textbf{\em stack integrity} in the simple model, the equivalent property in the
-passing model substitutes the above \(\CONTEXTS\) type and update
-function, and \(\prot_P\), for their equivalents in the simple model.
-
-For both the \textbf{\em stepwise} and \textbf{\em observational} versions of \textbf{\em stack confidentiality} in
-the original model, the passing model equivalents additionally substitute \(\secret_P\).
 
@@
 %% Appendix
-%\appendix
-%\section{Appendix}
+\appendix
+\paragraph*{Formalization of Callee-saves Registers}
+\label{app:calleesave}
+%
+\[\begin{split}
+\codemap \in \CODEMAPS ::= & \WORDS \rightharpoonup \callmap(S)\\
+\mbox{\rm{where}} ~ & S \in \REGS \rightarrow (\unsaved + \saved) \\
+\end{split}\]
+
+\[\begin{split}
+\stackDom \in \STKDOMS ::= & \outside + \unsealed + \sealed{\depth} \\
+& + \unsaved + \saved
+\end{split}\]
+%
+\judgmenttwobrlong[ Call]
+                  {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
+                  {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
+                  {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
+                  {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+The remaining rules are the same as in the simple model.
+
+\definition Let the protected set of a context, \(\prot_S(\domMap,\rts)\),
+be the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+or \(\domMap ~ \component = \saved\).
+
+\definition Let the secret set of a context, written \(\secret_S(\domMap,\rts)\),
+be the set of elements \(\component\) such that
+\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
+\(\domMap ~ \component = \saved\).
+
+\definition
+%
+\textbf{\em Observational stack integrity} with callee-saves registers substitutes the above
+\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
+%
+\textbf{\em Observational stack confidentiality} additionally substitutes \(\secret_S\).
+
+\paragraph*{Formalization of variable passing.}
+We add extra information to our call
+annotations to reflect that they may now pass variables on the stack, and that the locations
+of those variables may be dynamic (as in the case of pass-by-reference).
+A call annotation now includes a predicate relating machine states to the elements
+that should be passed, expressed as offsets from the stack pointer.
+\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P)~\mbox{\rm where}~P \subseteq \WORDS\]
+We extend the set of domains to include passed elements,
+which are labeled with the range of depths of the caller that has passed them.
+\[\begin{split}
+\stackDom \in \STKDOMS ::= & \outside + \unsealed\\
+& + \sealed{\depth} + \passed{\depth,\depth'}
+\end{split}\]
+The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
+\times (\rts \in \listT (\target)) \]
+The initial context again maps all elements to \(\unsealed\).
+We straightforwardly describe how the sealing convention and passing predicate combine to
+update a domain map at a call from depth \(\depth\).
+
+\judgmentthree[]
+              {\(\domMap ~ \component = \unsealed\)}
+              {\(\sealcon ~ \mach[\component]\)}
+              {\(\component - (\mach ~ \rsp) \not \in P\)}
+              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \sealed{\depth}\)}
+\judgmentthree[]
+              {\(\domMap ~ \component = \unsealed\)}
+              {\(\sealcon ~ \mach[\component]\)}
+              {\(\component - (\mach ~ \rsp) \in P\)}
+              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth,\depth+1}\)}
+\judgmentthree[]
+              {\(\domMap ~ \component = \passed{\depth',\depth}\)}
+              {\(\sealcon ~ \mach[\component]\)}
+              {\(\component - (\mach ~ \rsp) \in P\)}
+              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth',\depth+1}\)}
+\judgment[]
+         {}
+         {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \domMap ~ \component\)}
+
+         At the matching return we just unseal the elements at depth \(\depth\):
+
+\judgment[]
+         {\(\domMap ~ \component = \sealed{\depth}\)}
+         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
+\judgment[]
+         {\(\domMap ~ \component = \passed{\depth-1,\depth}\)}
+         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
+\judgmenttwo[]
+           {\(\domMap ~ \component = \passed{\depth',\depth}\)}
+           {\(\depth' < \depth\)}
+           {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
+\judgment[]
+         {}
+         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
+
+Now we can inductively define the context update function, as before.
+
+\judgmenttwo[ CallSeal]
+            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
+            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
+            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+
+\vspace*{-1ex}
+\judgmenttwo[ ReturnFound]
+            {\(\mach \stepsto \mach' \in \target\)}
+            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
+            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+
+\vspace*{-1ex}
+\judgmenttwo[ ReturnRec]
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
+              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
+              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
+
+\vspace*{-3ex}
+\judgment[ Default]
+         {}
+         {\(\underscore,\context \stepsto_C \context\)}
+
+\definition The \emph{protected set} of a context, \(\prot_P(\domMap,\rts)\),
+is the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
+or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
+\(|\rts| > \depth'\).
+
+\definition The \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
+is the set of elements \(\component\) such that
+\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
+\(\domMap ~ \component = \passed{\depth,\depth'}\)
+where \(|\rts| < \depth\) or \(|\rts| > \depth'\).
+
+\definition
+For each of the \textbf{\em stepwise} and \textbf{\em observational} versions of
+\textbf{\em stack integrity} in the simple model, the equivalent property in the
+passing model substitutes the above \(\CONTEXTS\) type and update
+function, and \(\prot_P\), for their equivalents in the simple model.
+
+For both the \textbf{\em stepwise} and \textbf{\em observational} versions of \textbf{\em stack confidentiality} in
+the original model, the passing model equivalents additionally substitute \(\secret_P\).
+
 %Text of appendix \ldots

### 2021-09-29 submission4/main.tex
More appendicitis

@@
 and when {\tt f} instead tries to write to {\tt x}, it violates integrity.
-%For that matter
-%a future call with no arguments passed by reference would also violate integrity if it wrote
-%to either variable.
+
+This extension to our model is formalized in Appendix \ref{app:argpass}.
 
@@
 \paragraph*{Formalization of variable passing.}
+\ref{app:argpass}
+
 We add extra information to our call

### 2021-09-29 submission4/main.tex
Moved testing to the end

@@
 violation would become visible. Now we can identify when lazy policies
-miss such violations, which are again susceptible to testing;
-Table~\ref{tab:lazy-bug-table} summarizes the results of our evaluation.
-The Lazy Tagging and Clearing
-micro-policy from Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} in fact
-allows violations of observable stack safety!
-Our implementation of this policy, {\em PER\_DEPTH\_TAG}, has an error that we find
-in testing. If, at depth \(d\), a callee \(A\) writes into its caller's
+miss such violations. It turns out that {\em Lazy Tagging and Clearing},
+as written, falls short of observational stack safety.
+If, at depth \(d\), a callee \(A\) writes into its caller's
 frame, then returns, and later another callee \(B\) at depth \(d\) reads the same
@@
 by all activations---but that policy allows leaks between
-instances of the same function.)
-
-We test {\em Lazy Per-activation Tagging and Clearing}, in which
-each activation is tagged uniquely, and find that it obeys stepwise
-confidentiality and observational integrity. To round out our mutations testing
-we also check an equivalent to {\em LOAD\_NO\_CHECK} and a version where
-stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
-It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
-taking twice as long to catch as the next longest.
-
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c|c}
-        Bug & Property Violated & MTTF (s) & Tests \\
-  \hline
-        {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
-        {\em LOAD\_NO\_CHECK}  & Obs. Integrity & 23.5 & 12.0 \\
-        {\em LOAD\_NO\_CHECK}  & Confidentiality & 19.2 & 695.5 \\
-        {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
-        {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous lazy policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:lazy-bug-table}
-\end{table}
-
-Our properties have allowed us to identify an enforcement mechanism as
-not really stack safe, and to validate a possible fix.
-Unfortunately, Lazy Per-activation Tagging and Clearing
-puts severe pressure on the tag cache, so it may not be usable in practice.
-However, we conjecture that no lazy scheme can be fully stack safe
-without using unique per-activation identifiers.
+instances of the same function.) We validate this repaired micro-policy in
+\ref{sec:testing}.
 
@@
   \end{subfigure}
-  \begin{subfigure}[t]{.47\textwidth}
+  \begin{subfigure}[t]{.4\textwidth}
 {\small
@@
   \end{subfigure}
-  \begin{subfigure}[t]{.27\textwidth}
+  \begin{subfigure}[t]{.25\textwidth}
     \begin{center}
@@
   \url{https://github.com/mit-plv/riscv-coq}.}
-%\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}
-and extend it with a
-runtime monitor implementing the stack-safety
-micro-policy. We
-  chose the Coq proof assistant as the setting for our implementation
-  to ensure that our coinductive trace
-  definitions are well-formed, and reason about them, and to leverage the power of the
-  QuickChick property-based testing framework~\cite{Pierce:SF4}.
+and extend it with a runtime monitor implementing a stack-safety
+micro-policy. We chose the Coq proof assistant as the setting for our implementation
+to ensure that our coinductive trace definitions are well-formed, and reason about
+them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.
 
@@
 
+To write such generators we build on the work of
+Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
+introduced {\em generation by execution} to produce progams that lead
+machines towards interesting behaviors. We extend the technique by tracking
+control flow information so that programs will usually --- but not always ---
+return to their call sites.
+
 %\paragraph*{Generators}
 %
-%To write such generators we build on the work of
-%Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
-%introduced {\em generation by execution} to produce progams that lead
-%machines towards interesting behaviors. Generation by
+% Generation by
 %execution receives as an input a partially instantiated machine state
@@
 %
-%In addition, we need to further extend this technique to handle the nested
-%nature of our confidentiality property: rather than just generating two
-%initial machines that are variants of one another and letting them
-%execute to test for noninterference, we generate a new variant
-%{\em every time a call is made} and check confidentiality for the
-%subtrace produced from that variant state until its corresponding
-%return. As a result, a ``single'' confidentiality test compactly
-%checks the confidentiality of multiple nested calls.
-
-\paragraph*{Findings}
+We need to further extend typical testing to handle the nested
+nature of confidentiality: rather than just generating two
+initial machines that are variants of one another and letting them
+execute to test for noninterference, we generate a new variant
+{\em every time a call is made} and check confidentiality for the
+subtrace produced from that variant state until its corresponding
+return. As a result, a ``single'' confidentiality test compactly
+checks the confidentiality of multiple nested calls.
+
+Our primary testing targets are the eager {\em Depth Isolation}
+and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
+We test the former against stepwise integrity and confidentiality, and
+the latter against their observational counterparts.
 
@@
 
+For example, when loading from a stack location, {\em Depth Isolation}
+needs to enforce that the tag on the location being read
+is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
+current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
+that restriction by not checking the depth equality (row {\em
+  LOAD\_NO\_CHECK\_DI}).
+
+Similarly, when storing to a stack location, the correct micro-policy
+needs to ensure that the tag on the memory location is either
+$\tagNoDepth$ or has again the same depth as the current $\PCname$
+tag. Relaxing that constraint causes violations to the integrity
+property (row {\em STORE\_NO\_CHECK}).
+
 \begin{table}[]
@@
   \hline
-      {\em LOAD\_NO\_CHECK}  & Confidentiality & 24.2 & 13.3 \\
+      {\em LOAD\_NO\_CHECK\_DI}  & Confidentiality & 24.2 & 13.3 \\
       {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
       {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
+  \hline
+  \hline
+      {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
+      {\em LOAD\_NO\_CHECK\_LT}  & Obs. Integrity & 23.5 & 12.0 \\
+      {\em LOAD\_NO\_CHECK\_LT}  & Confidentiality & 19.2 & 695.5 \\
+      {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
+      {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
+  \hline
 \end{tabular}
 \vspace*{1em}
-\caption{MTTF for finding bugs in erroneous eager policy enforcement mechanisms}
+\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
 \vspace*{-2em}
@@
 machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
-% \leo{Something like: There is really no baseline to compare this
-  %against, but include it here for the future/reasons?}
-%
-The artificially injected bugs weaken the tag propagation
-and checking mechanism.
-%
-For example, when loading from a stack location, the correct
-micro-policy needs to enforce that the tag on the location being read
-is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
-current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
-that restriction by not checking the depth equality (row {\em
-  LOAD\_NO\_CHECK}).
-%
-Similarly, when storing to a stack location, the correct micro-policy
-needs to ensure that the tag on the memory location is either
-$\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing that constraint causes violations to the integrity
-property (row {\em STORE\_NO\_CHECK}).
 
@@
 %  and the write from {\tt f} was already a violation.}
+
+Table~\ref{tab:lazy-bug-table} summarizes the results of our evaluation on lazy micro-policies.
+The {\em Lazy Tagging and Clearing} micro-policy implemented as {\em PER\_DEPTH\_TAG},
+fails in testing, in cases where data is leaked between sequential calls.
+To round out our mutations testing
+we also check an equivalent to {\em LOAD\_NO\_CHECK\_DI}, {\em LOAD\_NO\_CHECK\_LT}
+and a version where stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
+It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
+taking twice as long to catch as the next longest.
+
+Our properties have allowed us to identify an enforcement mechanism as
+not really stack safe, and to validate a possible fix.
+Unfortunately, Lazy Per-activation Tagging and Clearing
+puts severe pressure on the tag cache, so it may not be usable in practice.
+However, we conjecture that no lazy scheme can be fully stack safe
+without using unique per-activation identifiers.
 

### 2021-09-29 submission4/main.tex
Equation shrinking

@@
 \(\mpcstatename \in \MPCS = \MACHS \times \CONTEXTS\), and lift the context-step
-function into the machine step like so:
+function into the machine step:
 
@@
 coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\)
-
+produced by repeated application of \(\stepstocon\):
+%
+\vspace{-\bigskipamount}
 \begin{center}
-\begin{minipage}{.4\textwidth}
+\begin{minipage}[t]{.6\columnwidth}
 \judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
@@
 \end{minipage}
-\begin{minipage}{.4\textwidth}
+\begin{minipage}[t]{.35\columnwidth}
 \judgment%[Default]
@@
 \end{center}
-%
+\vspace{-\smallskipamount}
 
@@
 and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first element on which \(f\) holds.  \bcp{Any
-  chance of getting two of these on the same line?}
-
+\(\MPCT\) ending with the first element on which \(f\) holds.
+
+\vspace{-\medskipamount}
 \begin{center}
-  \begin{minipage}{.3\textwidth}
+  \begin{minipage}[b]{.5\columnwidth}
     \judgment{\(f ~ \mpcstatename\)}
@@
 %
-  \begin{minipage}{.3\textwidth}
+  \begin{minipage}[b]{.3\columnwidth}
     \judgment{}
@@
   \end{minipage}
-  \begin{minipage}{.3\textwidth}
+\end{center}
     \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
                 {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \notfinished{\mpcstatename}{\MPCT'}\)}
-  \end{minipage}
-\end{center}
-%
+
 If the resulting trace is a strict prefix of the parameter, then the
@@
 
-Each state element is assigned to a {\em domain}, which may be {\em outside}
-of the stack entirely, {\em unsealed} and therefore accessible, or {\em sealed}
-at a particular depth. Each caller registers a {\em target}
+Each state element is assigned to a {\em domain} \(\stackDom \in \STKDOMS\), which may be \(\outside\)
+of the stack entirely, \(\unsealed\) and therefore accessible, or \(\sealed{\depth}\)
+at a particular depth \(\depth\). Each caller registers a {\em target}
 condition, \(\target \subseteq \MACHS\), which is a predicate on machine states that means it
@@
 ``return target list,'' which is a stack of targets, one for each caller awaiting
-return.  \bcp{The first displayed line can be inlined in the text. (I've
-  done a few others like this already.  There are probably more.)}
-%
-\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
+return.
 \[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
@@
 
-\subsection{A Conservative Stack-Safety Enforcement Mechanism}
+\paragraph*{A Conservative Stack-Safety Enforcement Mechanism}
 %

### 2021-09-29 submission4/main.tex
Fixed some state accesses

@@
 written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach ~ \component\).
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
 
@@
 is the set of elements \(\component\)
-such that \(\mach[\component] \not = \mach' ~ \component\).
+such that \(\mach[\component] \not = \mach'[\component]\).
 
@@
 \(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
-\(\mach' ~ \component = \nach' ~ \component\).
+\(\mach'[\component] = \nach'[\component]\).
 

### 2021-09-29 submission4/main.tex
Some tightening

@@
 
-What does ``stack safety'' mean, exactly? The phrase is associated with a
+What does ``stack safety'' mean, precisely? The phrase is associated with a
 variety of compiler,
@@
 and a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state''), which can optionally be extended with a control
-flow property.
+caller's private state''), plus an optional control-flow property.
 
@@
 property-based random testing) that their ``eager''
-micro-policy, which catches violations as early as possible, enforces a
+micro-policy, which catches violations early, enforces a
 simple ``stepwise'' variant of our properties, and that
@@
 range of dire consequences, from leakage or corruption of private stack data
-to control-flow hijacking. To prevent or detect such attacks, a profusion of
+to control-flow hijacking. To foil such attacks, a profusion of
 software and hardware protections have been proposed,
@@
 We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
-which memory cells within each stack frame are tagged with the depth of
-the function activation that owns the frame and access to those locations is
+which memory within each stack frame is tagged with the depth of
+the function activation that owns the frame and access is
 then permitted only when an activation at that depth is currently executing, validates our
@@
 
-In the rest of the paper, we begin with an informal overview of our key ideas via
+\ifspace
+In the rest of the paper, we
+\else
+We
+\fi
+begin with an informal overview of our key ideas via
 examples (\cref{sec:running-example}), followed by a description of our key
@@
 \item We use property-based random testing to check that this stepwise
-property is actually enforced by the {\em Depth Isolation} policy
+property is  enforced by the {\em Depth Isolation} policy
 of Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
@@
   confidentiality and integrity. Both are parameterized over a notion of external
-  observation, and are violated only if the accessing of secrets or overwriting of
-  data later causes a visible change in the system behavior (\cref{sec:lazy}).
+  observation, and are violated only if accessing secrets or overwriting
+  data causes a visible change in the system behavior (\cref{sec:lazy}).
 \item We show that the definition can be
@@
 \centering
+\hspace*{-1cm}
 \begin{tabular}{l l | l}
@@
 in fact we can learn something about the initial state by observing those results.
-To adapt this to the stack, the notion of ``secret'' must be dynamic with function
+To adapt this to the stack, the notion of ``secret'' must vary with function
 calls and returns.
@@
   understand if we chose concrete values for the examples}\sna{FWIW I agree in
-  theory, but it looks bad every time I try it.\fi~(\(v_0\), etc.),
+  theory, but it looks bad every time I try it.}\fi~(\(v_0\), etc.),
 in place of the actual values ($42$, ...) that should
-be hidden from {\tt f}; we show the evolving stack memory of the variant in a second, parallel column.
-At instruction 11, we note that the callee's variable, {\tt z}, has
-changed to 42 in the primary trace, but
-in the variant trace it instead might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
+be hidden from {\tt f}; we show the evolving stack memory of the variant in the second column.
+At instruction 11, the callee's variable, {\tt z},
+changes to 42 in the primary trace, but
+in the variant trace it instead might have any value \(v_0\). If \(v_0 \neq 42\), then the variant
 has behaved differently than the primary, violating confidentiality.
@@
 means that if both states step---that is, neither the primary nor the variant failstops---any
-state element that changes in the original will change in the same way in the variant.
+state element that changes in the original will do the same in the variant.
 
@@
   \end{subfigure}
-  \caption{A control-flow attack.}
+  \caption{A control-flow attack. \bcp{Move the caption up a bit}}
   \label{fig:controlflow}
@@
 $-1$. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
-This has the appearance of an integrity violation, so how do we capture it in our
-integrity property?
+This seems like an integrity violation; how do we capture it in our
+\ifspace integrity \fi property?
 
@@
 integrity and confidentiality protect a caller's data until its callee
-returns to it. However, it is possible that an intermediate caller is not
+returns to it. However, it is possible that the caller is not
 returned {\em to}, but rather returned {\em past}, to some ancestor caller!
@@
 %
-Thus, a formal definition of safe use of the stack should include the
-condition that calls and returns should be correctly nested---e.g., $C$ should
+Thus, a formal definition of safe use of the stack should require that calls
+and returns be correctly nested---e.g., $C$ should
 only return to $B$.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
@@
 Before introducing our formal model of stack safety, we describe our assumptions:
-how broadly we can abstract over machines, how we represent enforcement, and
+how we abstract over machines, how we represent enforcement, and
 our threat model.

### 2021-09-29 submission4/main.tex

@@
 
-\caption{A confidentiality violation\ifspace\apt{Could save space by not
+\caption{A confidentiality violation\bcp{tighten so it doesn't spill into margins}\ifspace\apt{Could save space by not
     repeating C or assemly code for {\tt main}.}\fi}
@@
 create an imaginary {\em variant}
-state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
-  understand if we chose concrete values for the examples}\sna{FWIW I agree in
-  theory, but it looks bad every time I try it.}\fi~(\(v_0\), etc.),
+state with arbitrary values~%
+%\ifaftersubmission\bcp{IMO, it would be easier to
+%  understand if we chose concrete values for the examples}\sna{FWIW I agree in
+%  theory, but it looks bad every time I try it.}\fi
+(\(v_0\), etc.),
 in place of the actual values ($42$, ...) that should

### 2021-09-29 submission4/main.tex
Layout of named inference rules
beginning a pass

@@
 capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}. \apt{Can we choose a bibstyle that
+  doesn't bracket each reference separately?}
+  \ifaftersubmission\apt{Mostly from
   nick; there could be more}\bcp{Yes, going back to MIT days---we should
@@
 callees must always return to their immediate caller, if they return
-at all. Skostengaard et al. define stack safety as the conjunction of WBCF with
+at all.
+\apt{Rest of this para could be delayed or perhaps removed altogether.}
+Skorstengaard et al. define stack safety as the conjunction of WBCF with
 {\em local state encapsulation (LSE)}.
@@
 
+\apt{Next two paras could be condensed.}
 For ease of exposition, our development initially assumes a single simple
@@
   data causes a visible change in the system behavior (\cref{sec:lazy}).
-\item We show that the definition can be
-extended to argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
+\item We show that our definitions can be
+extended to describe argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
 \end{itemize}
@@
 \end{subfigure}
-\caption{An integrity violation
+\caption{Integrity and confidentiality violations
 \ifaftersubmission\bcp{I wonder if we
     could make the figures easier to read by putting very light, differently
-  colored backgrounds under the different parts}\fi}
+  colored backgrounds under the different parts}\fi}\apt{Picture is now wrong at step 12 (missing 42 at [sp-1]).}
 \label{fig:int1}
@@
 of examples, each presented both in C syntax and in corresponding assembly code for a
-simple two-address machine (see Figure~\ref{fig:int1}).
+simple imaginary two-address machine (see Figure~\ref{fig:int1}).
 The address of each instruction is shown to its left.
@@
 (We will introduce some other parts of the context as needed.)
-Such a domain
-can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
+The possible domains
+are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
 (in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
@@
 also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
-are shaded black and ignored for simplicity for now --- they will
+are shaded black and ignored for simplicity for now---they will
 be treated properly later.

### 2021-09-30 submission4/main.tex
pass up to start of IV-A

@@
 The address of each instruction is shown to its left.
-The assembly code reflects a simple compilation model in which
-local variables are stored in the stack frame.
+The assembly code reflects a simple compilation strategy in which
+local variables are stored in consecutive slots of the stack frame.
 For the moment, arguments and return values are passed in registers.
@@
 
-Figure~\ref{fig:int1} shows a program with a simple integrity violation. It consists of a {\tt main}
+Figure~\ref{fig:int1} shows a program with a simple integrity violation.\apt{Mention confidentiality here too.} It consists of a {\tt main}
 function that performs a call to a function {\tt f}, which reads and then overwrites {\tt main}'s
@@
 and {\tt z} as well as that of {\tt x}. This is to detect some violations of so-called
-{\em temporal stack safety}. Suppose, in a
-In Figure~\ref{fig:conf2}, {\tt f} leaves behind a value in {\tt y}
-that later becomes the value of {\tt main}'s variable {\tt x}. Then {\tt g} is able to access
-{\tt y} and hence {\tt x} without even directly reading the stack pointer. In terms of noninterference,
+{\em temporal stack safety}. 
+Consider the code in Figure~\ref{fig:conf2}.
+Using the same compilation strategy as before, {\tt y} and {\tt z} will end up occupying the
+same memory location in the stack,  
+so the value {\tt f} leaves behind in {\tt y}, which later becomes the value of {\tt x}, also
+becomes the value of {\tt z}. Thus, {\tt g} is able to learn the value of its caller's
+variable {\tt x}.  We regard this as a violation of confidentiality as well. In terms of noninterference,
 in the primary trace, {\tt z} will have the value 5, but in the variant it can have
-an arbitrary value. We regard this as a violation of confidentiality as well.
+an arbitrary value.
 
@@
 
-In the example, on its first return, {\tt f} reaches its return target, because the
-\(\PCname\) is 4 after its return. But {\tt f}'s second return is to the wrong instruction
+In the example, {\tt f} reaches its return target on its first return, to instruction 4. 
+But {\tt f}'s second return is to the wrong instruction
 (4 rather than 6) because the stack pointer has been overwritten (shown in light red).
@@
 call \emph{well-bracketed control flow (WBCF)}.
+\apt{We've lost the story about how we treat WBCF, namely as an optional additional property.}
 
@@
 Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the concrete machine that instantiates it. Each formal property
+assumptions about the concrete machine that instantiates it. \apt{This sentence adds little:} Each formal property
 may be instantiated with any machine that meets the following criteria, and it
-will be a meaningful definition of stack safety for that machine. It must at least be a
+will be a meaningful definition of stack safety for that machine. We require a 
 register machine with a program counter (\(\PCname\)) and there must be a well-defined
 mechanism for determining the bounds of a stack frame, and for recognizing when a call
-has returned. In this paper we discuss those machines with a stack pointer (\(\rsp\)),
+has returned. In this paper we discuss machines with a stack pointer (\(\rsp\)),
 in which the current stack frame is everything below the stack pointer and above the
 previous frame, and a call from instruction \(\addr_\PCname\) with stack pointer \(\addr_\rsp\)
-is said to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
-is again at \(\addr_\rsp\). But, other configurations, such as those involving frame pointers,
-are equally feasible. A machine may use its non-stack memory in any way, including
+is considered to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
+is again at \(\addr_\rsp\). But our properties apply equally to machines with
+other architectural conventions, such as those involving frame pointers.
+A machine may use its non-stack memory in any way, including
 as a heap, as long as the stack has static bounds. It might also possess additional
@@
 secondary properties that use the same underlying formalism to handle these cases,
-as well as a simple coroutine model, in Section~\ref{sec:ext}.
+\apt{omit this clause?: ``as well as a simple coroutine model,''} in Section~\ref{sec:ext}.
 
@@
 Our model is flexible enough to support many realistic uses of the stack.
-We intentionall say little or nothing about other structures
+We intentionally say little or nothing about other structures
 such as the heap.  (Heap safety has already been studied, in isolation,
@@
 
+\apt{Suggest just listing these, with the briefest of descriptions, and omitting
+ the mitigations/excuses/etc. And turn this into an in-paragraph itemization.}
 Features that we do not currently support:
@@
 
-\judgmenttwobrlong[ Call]
+\apt{splitting these into separated tabulars would allow them to flow over the column break, presumably.}
+\begin{tabular}{l}
+{\sc Call} \\
+\judgmenttwobrlong[]
               {\(\codemap ~ (\mach ~ \PCname)\)}
@@
               {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
-
-\vspace*{-1ex}
-\judgmenttwo[ ReturnFound]
+                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)} \\
+{\sc ReturnFound} \\
+\judgmenttwo[]
               {\(\mach \stepsto \mach' \in \target\)}
               {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap[\components \mapsto \unsealed],\rts\)}
-
-\vspace*{-1ex}
-\judgmentbr[ ReturnRec]
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap[\components \mapsto \unsealed],\rts\)} \\
+{\sc ReturnRec} \\
+\judgmentbr[]
             {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
             {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-            {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'[\components \mapsto \unsealed],\rts'\)}
-\vspace*{-2.5ex}
-\judgment[ Default]
+            {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'[\components \mapsto \unsealed],\rts'\)} \\
+{\sc Default} \\
+\judgment[]
          {}
-         {\(\underscore,\context \stepsto_C \context\)}
+         {\(\underscore,\context \stepsto_C \context\)} \\
+\end{tabular}
 
@@
 
+\apt{This doesn't want to be a lettered subsection, but it wants to be somehow
+  a bigger division than the paragraphs below.}
 \subsection{Case Study: Micro-Policies}

### 2021-09-30 submission4/main.tex
Tweaking

--- 2021-09-29 submission4/main.tex
@@
 them to a system that models the passing of
-parameters on the stack (Section~\ref{sec:passing}), to callee-saves
-registers  (\ref{sec:callee}), and to a very simple coroutine model
-(\ref{sec:coroutines}).
-
-\subsection{Parameters Passed on the Stack}
-\label{sec:passing}
+parameters on the stack, to callee-saves
+registers\bcp{No longer here!}, and to a simple coroutine model.
 
@@
 
-To this point we have ignored function arguments, by assuming that they are placed
+\medskip
+\paragraph*{Parameters Passed on the Stack}
+%
+Up to this point we have ignored function arguments, by assuming that they are placed
 in registers (which are implicitly shared by all functions).
@@
 
-\subsection{Coroutines}
-\label{sec:coroutines}
+\medskip
+
+\paragraph*{Coroutines}
 

### 2021-09-30 submission4/main.tex
Tightening

@@
 including those that do not failstop, such as
-code rewriting techniques that prevent stack safety violations directly by
+code rewriting techniques that protect stack safety directly by
 restricting the set of possible initial states.
@@
 of MC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces (ordinary lists will use
-``::'').
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces and ``::'' for list-cons.
 
@@
 
-\apt{This doesn't want to be a lettered subsection, but it wants to be somehow
-  a bigger division than the paragraphs below.}
-\subsection{Case Study: Micro-Policies}
+\medskip
+\paragraph*{Case Study: Micro-Policies}
 \label{sec:enforcement}
@@
 There remains the question of how to ensure that the sequences are
-invoked at the right places from the perspective of the stack-safety
-property. For a program to enjoy stack safety
+invoked at the right places\ifspace from the perspective of the stack-safety
+property\fi. For a program to enjoy stack safety
 (with respect to a particular call map), all we require is that
@@
 
+\medskip
 \paragraph*{A Lazy Alternative}
@@
 
-We answer the above questions by introducing a notion of observations to our model.
-This has uses beyond this one enforcement mechanism, as some code features may involve
-moving data, overwriting its original location,
-and ultimately restoring it before the change becomes visible; for example, the
-``callee-saves'' registers in some calling conventions will be stashed and restored.
+We answer these questions by introducing a notion of observations to our model.
+This will also be useful when, for example,
+a compiler wants to move data, overwrite its original location,
+and ultimately restore it before the change becomes visible, as in
+``callee-saves'' calling conventioons.
 These examples call for an abstract mechanism to characterize behaviors that are externally
@@
 
-Confidentiality is more sophisticated. We must capture the intuition that
+For confidentiality, we must capture the intuition that
 secrets do not escape during a call {\em and} are not leaked during a return.
 This require us to be able to quantify over elements that fail to preserve
-the change-together relation \(\diamond\).
+the change-together relation~\(\diamond\).
 
@@
 
-\subsection{Connection to Policies}
+\medskip
+\paragraph*{Connection to Policies}
 
@@
 
-\subsection{Callee-saves Registers}
+\medskip
+\paragraph*{Callee-saves Registers}
 \label{sec:callee}

### 2021-09-30 submission4/main.tex
Answered various comments and shrunk things

@@
 at all.
-\apt{Rest of this para could be delayed or perhaps removed altogether.}
-Skorstengaard et al. define stack safety as the conjunction of WBCF with
-{\em local state encapsulation (LSE)}.
-Informally, LSE says that the contents of
-a caller's stack frame are not read or written while its callees (and their callees, etc.) are
-executing.
-Skorstengaard et al. formalize these properties by defining an idealized
-machine in which they hold {\em by construction}. Our confidentiality and
-integrity properties are together stronger than LSE, guaranteeing protection even
-in the presence of control-flow violations, and are essentially orthogonal to WBCF.
 
@@
 
-\apt{Next two paras could be condensed.}
-For ease of exposition, our development initially assumes a single simple
+For ease of exposition, we initially assume a single simple
 stack with no sharing between callers and callees: all parameters and return
-values are passed in registers.  Later, we show how to refine the properties
-to allow passing of scalar stack data, granting a callee access to the
-caller's data. Throughout, we note how we expect elements of the formalism
-to apply to enforcement mechanisms beyond micro-policies.
-
-Lastly, we define integrity and confidentiality properties for a coroutine
-system with a static layout, in which each coroutine's stack is constrained
-to a fixed region of memory.  Though simplistic, this extension is a step
-toward full-blown multithreading and demonstrates that our approach is flexible
-enough to handle more sophisticated styles of control flow.
-
-\ifspace
-In the rest of the paper, we
-\else
-We
-\fi
-begin with an informal overview of our key ideas via
+values are passed in registers.  Later, we will introduce enhanced versions
+supporting
+1) passing of scalar stack data,
+2) callee-saves registers calling conventions, and
+3) a coroutine system with a static layout.
+
+We begin with an informal overview of our key ideas via
 examples (\cref{sec:running-example}), followed by a description of our key
@@
 
-Figure~\ref{fig:int1} shows a program with a simple integrity violation.\apt{Mention confidentiality here too.} It consists of a {\tt main}
+Figure~\ref{fig:int1} shows a program with simple integrity and confidentiality violations. It consists of a {\tt main}
 function that performs a call to a function {\tt f}, which reads and then overwrites {\tt main}'s
@@
 
-\caption{A confidentiality violation\bcp{tighten so it doesn't spill into margins}\ifspace\apt{Could save space by not
-    repeating C or assemly code for {\tt main}.}\fi}
+\caption{A confidentiality violation\bcp{tighten so it doesn't spill into margins}}
 \label{fig:conf1}
@@
 %
-Thus, a formal definition of safe use of the stack should require that calls
-and returns be correctly nested---e.g., $C$ should
+Thus, another aspect of safe use of the stack is that calls
+and returns should be correctly nested---e.g., $C$ should
 only return to $B$.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
-call \emph{well-bracketed control flow (WBCF)}.
-\apt{We've lost the story about how we treat WBCF, namely as an optional additional property.}
+call \emph{well-bracketed control flow (WBCF)}, and is orthogonal to
+confidentiality and integrity, so we treat is as an optional additional property.
 
@@
 Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the concrete machine that instantiates it. \apt{This sentence adds little:} Each formal property
-may be instantiated with any machine that meets the following criteria, and it
-will be a meaningful definition of stack safety for that machine. We require a
+assumptions about the concrete machine that instantiates it. We require a
 register machine with a program counter (\(\PCname\)) and there must be a well-defined
@@
 secondary properties that use the same underlying formalism to handle these cases,
-\apt{omit this clause?: ``as well as a simple coroutine model,''} in Section~\ref{sec:ext}.
+in Section~\ref{sec:ext}.
 
@@
 
-\apt{Suggest just listing these, with the briefest of descriptions, and omitting
- the mitigations/excuses/etc. And turn this into an in-paragraph itemization.}
-Features that we do not currently support:
-
-\begin{itemize}
-\item Address-taken locals. A satisfactory
-  treatment of these seems to require integration with a more general
-  model of memory safety that covers the heap and static globals too,
-  such as the aforementioned model of Azevedo de Amorim et al.
-
-\item Exceptions. We can see a way to extend the model to include them:
-  we would model a "setjump" operation by inserting a
-  special target in the return target stack. We would weaken
-  well-bracketedness to allow reaching such a target from any
-  depth. Since integrity and confidentiality are distinct from
-  well-bracketedness and already permit the popping of multiple
-  functions from the stack, they can simply treat the exception's
-  target as another return target.
-
-\item Tail calls. To implement these, we would need to distinguish them
-  from normal calls and slightly adjust the mechanics of the return stack.
-
-\item Dynamic code generation. This would be challenging to
-  integrate into our framework, because we rely on code annotations.
-  But we can easily envision supporting limited cases in which
-  the generated code never contains a call or other annotated instruction,
-  or in which existing code is copied along with its annotations.
-
-\item Concurrency. We can handle a limited form of coroutines, as described in
-  8.3. Handling full-blown preemptive concurrency would require significant
-  changes in our underlying machine model.
-\end{itemize}
+We do not currently support:
+1) address-taken locals, as a satisfactory treatment of these seems to require
+integration with a more general model of memory safety,
+2) exceptions, \sna{I could just write up how to do these...}
+3) tail calls,
+4) dynamic code generation, which is difficult to integrate with code annotations in general, and
+5) concurrency --- we can handle a limited form of coroutines, but
+  not full-blown preemptive concurrency.
 
@@
 from components to values, written
-\(\mach[\component] ::= \MACHS \rightarrow \COMPONENTS \rightarrow \WORDS \).
+\(\mach[\component] \in \MACHS \rightarrow \COMPONENTS \rightarrow \WORDS \).
 
@@
 
-\apt{splitting these into separated tabulars would allow them to flow over the column break, presumably.}
 \begin{tabular}{l}
@@
                 {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)} \\
+\end{tabular}
+\begin{tabular}{l}
 {\sc ReturnFound} \\
@@
 {\sc ReturnRec} \\
+\end{tabular}
+\begin{tabular}{l}
 \judgmentbr[]
@@
             {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'[\components \mapsto \unsealed],\rts'\)} \\
+\end{tabular}
+\begin{tabular}{l}
 {\sc Default} \\
@@
 
+\smallskip
+\begin{minipage}[b]{.2\columnwidth}
 \judgment{}
          {\(\mpcstatename \doteq \npcstatename\)}
+\end{minipage}
+\begin{minipage}[b]{.75\columnwidth}
 \judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
             {\(\machT \doteq \nachT\)}
-            {\(\notfinished{(\mach,\underscore,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore,\underscore)}{\nachT}\)}
+            {\(\notfinished{(\mach,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore)}{\nachT}\)}
+\end{minipage}
 
@@
 \(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
-\(\mach' ~ \component \not = \nach' ~ \component\).
+\(\mach'[\component] \not = \nach'[\component]\).
 
@@
 them to a system that models the passing of
-parameters on the stack, to callee-saves
-registers\bcp{No longer here!}, and to a simple coroutine model.
+parameters on the stack and to a simple coroutine model.
 
@@
 to it.
-%\sna{This is trying to get toward what we were saying at OOPSLA: it's not
-%  a technical obstacle, it's that the behavior is tied to things we don't model.}
 
@@
              {\(\mach \stepsto \mach'\)}
-             {\(\stackid' = \stackof ~ (\mach'~ \rsp) \not = \stackid\)}
+             {\(\stackid \not = \stackid' = \stackof ~ (\mach'[\rsp])\)}
              {\(\yts ~ \stackid'\)}

### 2021-09-30 submission4/main.tex
Work on figures

@@
 \usepackage{amsmath,amsthm,amssymb}
+
+\usepackage{xcolor,listings}
 
@@
 \begin{figure*}
-
 \begin{subfigure}[t]{.2\textwidth}
 {\small
-\begin{verbatim}
+  \begin{lstlisting}%[backgroundcolor=\color{gray!20}]
 int main() {
@@
 }
-\end{verbatim}
+  \end{lstlisting}
 }
 \end{subfigure}
-\begin{subfigure}[t]{.5\textwidth}
+\begin{subfigure}[t]{.45\textwidth}
   {\small
-\begin{verbatim}
+    \begin{lstlisting}[backgroundcolor=\color{lgray}]
 1 main: add $1,%sp      ; allocate frame
@@
 12      mov $0,-4(%sp)      ; violation 2!
-13      mov $5, %ra     ; set return value
+13      mov $5,%ra      ; set return value
 14      sub $1,%sp      ; deallocate frame
 15      ret
-\end{verbatim}
+    \end{lstlisting}    
 }
@@
 \begin{tabular}{l l}
-{\tt 2} &
+2 &
 \memoryaddrs{8em}
 \memory{3}{\unsealc}[\unsealdesc]
+
 ~$\cdots$
@@
 \\
-{\tt 3} &
+3 &
 \memoryaddrs{8em}
@@
 \\
-{\tt 12} &
+12 &
 \memoryaddrs{16em}
@@
 ~$\cdots$
-\MemoryLabel{-15em}{0.75em}{42}
+\MemoryLabel{-16em}{0.75em}{42}
+\MemoryLabel{-7.5em}{0.75em}{42}
 \vspace{.5em}
 \\
-{\tt 13} &
+13 &
 \memoryaddrs{16em}
@@
     could make the figures easier to read by putting very light, differently
-  colored backgrounds under the different parts}\fi}\apt{Picture is now wrong at step 12 (missing 42 at [sp-1]).}
+  colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
 \label{fig:int1}
 \end{figure*}
+
+\addtolength{\abovecaptionskip}{-1em}
 
@@
 \centering
-\hspace*{-1cm}
-\begin{tabular}{l l | l}
-  {\tt 2} &
+\addtolength{\tabcolsep}{-1em}    
+\begin{tabular}{l | l}
   \multicolumn{2}{c}{
+    2 \hspace{-1.5em}
     \memoryaddrs{8em}
@@
     ~$\cdots$
-%    \MemoryLabel{-18em}{0.75em}{0}
-%    \MemoryLabel{-14em}{0.75em}{0}
-%    \MemoryLabel{-10em}{0.75em}{0}
-%    \MemoryLabel{-7em}{0.75em}{0}
     \vspace{.5em}
   } \\
-  {\tt 11} &
+  11 \hspace{-1.5em}
   \memoryaddrs{21em}
@@
   \MemoryLabel{-19em}{0.75em}{42}
-%  \MemoryLabel{-10em}{0.75em}{0}
-%  \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   \\
-  {\tt 12} &
+  12 \hspace{-1.5em}
   \memoryaddrs{21em}
@@
   \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-10em}{0.75em}{5}
- % \MemoryLabel{-6em}{0.75em}{0}
   &
@@
   \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-10em}{0.75em}{5}
   \MemoryLabel{-6em}{0.75em}{\(v_2\)}
   \\
-  {\tt 13} &
+  13 \hspace{-1.5em}
   \memoryaddrs{21em}
@@
   \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-10em}{0.75em}{5}
   \MemoryLabel{-6em}{0.75em}{42}
@@
   \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-10em}{0.75em}{5}
   \MemoryLabel{-6em}{0.75em}{\(v_0\)}
@@
 \end{tabular}
+\addtolength{\tabcolsep}{1em}
 
@@
 
-\caption{A confidentiality violation\bcp{tighten so it doesn't spill into margins}}
+\caption{Noninterference, violated}
 \label{fig:conf1}
@@
 %\begin{center}
-\begin{subfigure}[t]{.32\columnwidth}
+\begin{subfigure}[t]{.28\columnwidth}
   {\small
-\begin{verbatim}
+\begin{lstlisting}
 int main() {
@@
 }
-\end{verbatim}
+\end{lstlisting}
+  }
+\end{subfigure}
+\begin{subfigure}[t]{.27\columnwidth}
+  {\small
+\begin{lstlisting}[backgroundcolor=\color{lgray}]
+int f() {
+  int y = 5;
+  return y;
+}
+\end{lstlisting}
   }
@@
   {\small
-\begin{verbatim}
-int f() {
-  int y = 5;
-  return y;
-}
-\end{verbatim}
-  }
-\end{subfigure}
-\begin{subfigure}[t]{.32\columnwidth}
-  {\small
-\begin{verbatim}
+\begin{lstlisting}
 int g() {
@@
 }
-\end{verbatim}
+\end{lstlisting}
   }
@@
 {\footnotesize
-\begin{verbatim}
+\begin{lstlisting}
 int *stash = 0;
@@
 }
-\end{verbatim}
+\end{lstlisting}
 }
@@
 {\footnotesize
-\begin{verbatim}
+\begin{lstlisting}[backgroundcolor=\color{lgray}]
    stash: .word 0
@@
 18        ret
-\end{verbatim}
+\end{lstlisting}
 }
@@
     \begin{tabular}{l l l}
-      {\tt 3} &
+      3 &
       \memoryaddrs{8em}
@@
       \\
-      {\tt 4} &
+      4 &
       \memoryaddrs{8em}
@@
       \MemoryLabel{-15em}{0.75em}{1} \\
-      {\tt 11} &
+      11 &
       \memoryaddrs{16em}
@@
       \\
-      {\tt 18} &
+      18 &
       \memoryaddrs{12em}
@@
       \\
-      {\tt 5} &
+      5 &
       \memoryaddrs{8em}
@@
   \end{subfigure}
-  \caption{A control-flow attack. \bcp{Move the caption up a bit}}
+  \caption{A control-flow attack.}
   \label{fig:controlflow}
@@
 {\small
-\begin{verbatim}
+\begin{lstlisting}
 int *stash = 0;
@@
 }
-\end{verbatim}
+\end{lstlisting}
 }
   \end{subfigure}
-  \begin{subfigure}[t]{.4\textwidth}
+  \begin{subfigure}[t]{.5\textwidth}
 {\small
-\begin{verbatim}
+  \begin{lstlisting}[backgroundcolor=\color{lgray}]
    stash: .word 0
-1  main:  add $2,%sp      ; alloc frame
+1  main:  add $2,%sp      ; allocate frame
 2         mov %sp,%ra
@@
 9         mov -2(%sp),%ra ; set return
-10        sub $2,%sp      ; dealloc frame
+10        sub $2,%sp      ; deallocate frame
 11        ret
@@
     \begin{center}
+      \vspace{1em}
     \begin{tabular}{l l}
-      {\tt 20} &
+      20 &
       \memoryaddrs[(1)]{17em}
@@
       ~$\cdots$ \\
-      {\tt 25} &
+      25 &
       \memoryaddrs[(1)]{17em}
@@
       ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{5}
+      \MemoryLabel{-14em}{0.75em}{5}
       \\
-      {\tt 8} &
+      8 &
       \memoryaddrs[(0)]{12em}
@@
       ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
+      \MemoryLabel{-14em}{0.75em}{0}
       \\
-      {\tt 20} &
+      20 &
       \memoryaddrs[(1)]{17em}
@@
       ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{0}
+      \MemoryLabel{-14em}{0.75em}{0}
       \\
-      {\tt 25} &
+      25 &
       \memoryaddrs[(1)]{17em}
@@
       ~$\cdots$
-      \MemoryLabel{-18em}{0.75em}{\bf 5}
+      \MemoryLabel{-14em}{0.75em}{\bf 5}
 \end{tabular}
@@
 The first major difference in this model is the domains. Stack domains are now nested inside
-top-level domains, and associated with stack identifiers.
+top-level domains, \(\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\),
+ssociated with stack identifiers.
 Our context now consists of a domain map, a map from stack identifiers to return target stacks,
@@
 %
-\[\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\]
 \[\begin{split}
-\context \in \CONTEXTS ::= &
- (\domMap \in \COMPONENTS \rightarrow \TOPDOMS) \times \\
+\context \in \CONTEXTS ::= & (\domMap \in \COMPONENTS \rightarrow \TOPDOMS) \times \\
 & (\rtm \in \STACKS \rightarrow \listT(\target)) \times \\
-& (\yts \in \STACKS \rightarrow \target) \times \\
-& \stackid \in \STACKS
+& (\yts \in \STACKS \rightarrow \target) \times \stackid \in \STACKS
 \end{split}\]

### 2021-09-30 submission4/main.tex
More small tweaks

@@
 unseals each frame down to the level that was returned to.
-If none of these situations hold, the context remains unchanged.
+If none of these rules apply, the default rule leaves context unchanged.
 
@@
 \begin{tabular}{l}
-{\sc Default} \\
-\judgment[]
-         {}
-         {\(\underscore,\context \stepsto_C \context\)} \\
 \end{tabular}
@@
 
-This calls for some explanation. First, we determine which coroutine we are yielding to by
+First, we determine which coroutine we are yielding to by
 checking which stack the stack pointer of the next state is pointing to. Then we can only yield
@@
 
-Table~\ref{tab:lazy-bug-table} summarizes the results of our evaluation on lazy micro-policies.
-The {\em Lazy Tagging and Clearing} micro-policy implemented as {\em PER\_DEPTH\_TAG},
+As for lazy micro-policies and observational properties,
+the original {\em Lazy Tagging and Clearing} micro-policy, implemented as {\em PER\_DEPTH\_TAG},
 fails in testing, in cases where data is leaked between sequential calls.
@@
 We improve upon Skorstengaard et al. \cite{SkorstengaardSTKJFP} in two significant ways:
-\begin{enumerate}[label=(\roman*)]
-\item Our definition captures the intuition that {\em
+
+i) Our definition captures the intuition that {\em
   every} caller should be protected from its callees.  By contrast,
@@
 
-\item Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s fully
+ii) Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s fully
   abstract overlay
@@
   to our base domain model.
-
-\end{enumerate}
 

### 2021-09-30 submission4/main.tex
Cut WBCF down more

@@
 appropriate return target in {\tt main}.
-To capture this intuition, we add a new piece of context to our diagram:
+To capture this intuition, we add a new piece of context:
 a stack of return targets.  Now a callee is
@@
 In the example, {\tt f} reaches its return target on its first return, to instruction 4.
-But {\tt f}'s second return is to the wrong instruction
-(4 rather than 6) because the stack pointer has been overwritten (shown in light red).
+But {\tt f}'s second return is to 4 rather than 6
+because the stack pointer has been overwritten (shown in light red).
 Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
@@
 
-\paragraph*{Well-bracketed control flow.} Together,
+\paragraph*{Well-bracketed control flow} Together,
 integrity and confidentiality protect a caller's data until its callee
@@
 returned {\em to}, but rather returned {\em past}, to some ancestor caller!
-For example, function $A$ calls function $B$, which in turn calls $C$, which
-returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
-and branching to the address after the call from $A$ to $B$.  Strictly
-speaking, this is not an information-flow violation, since neither $A$'s nor
-$B$'s stack frames are improperly read or written by $C$.  But it is still
-arguably unsafe if $A$'s programmer depends on $B$ completing its execution before
+Strictly speaking, this is not an information-flow violation. But it is still
+arguably unsafe if the ancestor depends on the skipped function completing its execution before
 it returns.
@@
 Thus, another aspect of safe use of the stack is that calls
-and returns should be correctly nested---e.g., $C$ should
-only return to $B$.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
+and returns should be correctly nested---a caller should
+only return to its own callee.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
 call \emph{well-bracketed control flow (WBCF)}, and is orthogonal to
@@
 properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs, including RISCy ones.  In
-particular, there may be no single ``call'' instruction that does all the
+machine-code level for a variety of ISAs.  There may be no single ``call''
+instruction that does all the
 required manipulation of the \(\PCname\) and stack pointer. Instead, there
@@
 
-In other words, stepwise stack confidentiality states that if, at the start of the call,
-we vary all stack locations that are supposed to be secret from the callee
-(which in our simple setting is the entirety of the stack),
+In short, if we consider a variant of the stack at the start of a call, 
 the resulting trace remains in lockstep with the original trace
-until either both return together or one failstops. This implies that
-if the callee reads a secret location, it must failstop.
-%
+until either both return together or one failstops.
 Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
 of the elements in \(\components\), and in addition their subtraces are also in
-\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
-will contain these calls and returns as well, so its return stack pushes and pops
-in lockstep until it breaks at the final return.
-
-Finally, we formalize a version of WBCF, which captures the same informal idea
-as that of Skorstengaard et al.\cite{SkorstengaardSTKJFP} using our context machinery.
+\(\MPCT^*\), and therefore confidentiality must hold on them separately.
+
+Finally, we formalize WBCF as a simple property of the behavior of returns.
 \definition
@@
 from their associated values at the hardware level. So they are not modeled
-as part of the values, and therefore our properties make no claims about
+as values, and therefore our properties make no claims about
 their behavior.

### 2021-09-30 submission4/main.tex
Save some space in related work

@@
 well-bracketed control flow (WBCF).
-Informally, LSE means that the current activiation can only access its own frame,
+Informally, LSE means that the current activation can only access its own frame,
 and WBCF means that the topmost frame can only return to
@@
 ii) Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s fully
-  abstract overlay
-  semantics rely on the definition of a ``stack-safe-by-construction''
+  abstract overlay semantics rely on the definition of a ``stack-safe-by-construction''
   abstract machine that is closely tied to the underlying machine architecture.
-  While the fundamental approach can no doubt be ported to other architectures and enforcement
+  While the approach can no doubt be ported to other architectures and enforcement
   schemes, each such port will require a new abstract machine definition whose
-  safety-by-construction property must be checked afresh.
-  Since we abstract over most of the machine
-  state as well as the sealing and return conventions, our properties apply directly
-  to a range of architectures, code features, and enforcement mechanisms.
-  Moreover, we have implemented extensions such as passing arguments on the stack,
-  describing callee-saves registers, and coroutines by making simple modifications
-  to our base domain model.
+  safety-by-construction must be checked afresh.
+  Since we abstract over most of the machine state as well as the sealing and return conventions, and
+  can support further enhancements such as coroutines,
+  our properties apply directly to a range of architectures, code features, and enforcement mechanisms.
 

### 2021-09-30 submission4/main.tex
Assorted edits

@@
 \usepackage{xcolor,listings}
+
+\usepackage[noadjust]{cite}
 
@@
 Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt x} will be unchanged when it gets control back.
+{\tt main} expects that {\tt x} will be unchanged when it gets control back,
+and usually at every step in between.
 This leads to a simple, intuitive statement of {\em stepwise integrity}:
 if a element is sealed before a step, it will be unchanged afterward.
-% Stepwise
-% refers to the fact that the property must hold on every step.
- %Later we will
-%loosen this condition in the interest of extensionality---the caller only
-%cares about changes it can see after the return, not changes during execution.
-In this case, after instruction 12, {\tt f} has violated integrity by writing to {\tt x},
-so the final return value produced by {\tt main} is wrong.
 
@@
 of a {\em blessed sequence} -- a series of instructions that are tagged to enforce correct
-behavior. This is all we need for the policy to protect itself: it can then failstop if it would
+behavior. It can then failstop if it would
 execute the blessed sequence out of order.
@@
 
+\sna{Necessary? }
 Note that an infinite silent trace is a
@@
 
-That is, our generators need to produce not only a low-level program
-that exhibits interesting control-flow patterns, but also the policy
-meta-information and code annotations that would normally be produced
-by the compiler.
-
 To write such generators we build on the work of
@@
 return. As a result, a ``single'' confidentiality test compactly
-checks the confidentiality of multiple nested calls.
+checks multiple nested calls.
 
@@
 fails in testing, in cases where data is leaked between sequential calls.
-To round out our mutations testing
-we also check an equivalent to {\em LOAD\_NO\_CHECK\_DI}, {\em LOAD\_NO\_CHECK\_LT}
+To round out our mutations testing we also check {\em LOAD\_NO\_CHECK\_LT},
+equivalent to its counterpart in depth isolation,
 and a version where stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
@@
 %% Bibliography
-\bibliographystyle{ieeetran}
+\bibliographystyle{IEEEtran}
 \bibliography{bcp.bib,local.bib}
@@
 
+\paragraph*{Exceptions}
+
 %Text of appendix \ldots

### 2021-09-30 submission4/main.tex
Added mit-riscv citation, other small changes

@@
 top-level domains, \(\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\),
-ssociated with stack identifiers.
+associated with stack identifiers.
 Our context now consists of a domain map, a map from stack identifiers to return target stacks,
@@
 & (\rtm \in \STACKS \rightarrow \listT(\target)) \times \\
-& (\yts \in \STACKS \rightarrow \target) \times \stackid \in \STACKS
+& (\yts \in \STACKS \rightarrow \target) \times (\stackid \in \STACKS)
 \end{split}\]
@@
 \definition
-The \emph{protected set} of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
-is the set of elements \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
-
-\definition
-The \emph{secret set} of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
-is the set of elements \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
+Let \(\prot_Y(\domMap,\rtm,\yts,\stackid)\) be the set of elements \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\),
+and \(\secret_Y(\domMap,\rtm,\yts,\stackid)\) the same.
 
@@
 confidentiality. We use a Coq specification of the RISC-V
-architecture\bcp{change from footnote to citation :-)}\footnote{Available at
-  \url{https://github.com/mit-plv/riscv-coq}.}
+architecture~\cite{Bourgeat2021AMF}.
 and extend it with a runtime monitor implementing a stack-safety
@@
 properties to apply well-established concepts and formal techniques
-of non-interference, and we define our properties in terms of
+of noninterference, and we define our properties in terms of
 traces of memory and register states, rather than in terms of an
@@
 (3) {\em Reusability} in secure compilation chains.
-StkTokens
-is specifically set up to address this,
-while it is not a goal of
-our work.
+StkTokens is specifically set up to address this,
+while it is not a goal of our work.
 %
@@
 
-We evaluate {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
-  Clearing} against our properties, finding that {\em Depth Isolation} does
+We evaluate {\em Depth Isolation} and {\em Lazy Tagging and Clearing}
+against our properties, finding that the former does
 indeed enforce stepwise stack safety and therefore observable stack
-safety. We also find that neither lazy micro-policy actually enforces
+safety. We find that the latter does not actually enforce
 stack safety, and we proposed a fix.

### 2021-09-30 submission4/main.tex
re-added section VI; adjusted contributions

@@
 until the callee returns.
-
-We can also, optionally, include a notion of {\em well-bracketed control
-flow (WBCF)} as in Skorstengaard et al. \cite{SkorstengaardSTKJFP}---a global requirement that
+%
+We can also, optionally, enforce {\em well-bracketed control
+flow (WBCF)}~\cite{SkorstengaardSTKJFP}---a global requirement that
 callees must always return to their immediate caller, if they return
 at all.
+\apt{Originally, we were using this spot to contrast with Skorstengaard et al. in the intro.
+  Do we really need to do that? Reads better without!}
 
@@
 strict stack safety enforced over each step of execution~(\cref{sec:prelim,sec:lse}).
-\item We use property-based random testing to check that this stepwise
-property is  enforced by the {\em Depth Isolation} policy
-of Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
-execution techniques of Hri\c{t}cu et al.\cite{TestingNI:ICFP} to handle nested
-properties (\cref{sec:enforcement}).
+As a case study, we explore how our properties are enforced 
+by the {\em Depth Isolation} micro-policy
+of Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}, but not by some of its optimized variants~(\cref{sec:enforcement}).
 \item We formalize weaker but more extensional \emph{observational} variants of
@@
 extended to describe argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
+\item We use property-based random testing to validate the relationship between
+  our properties and micro-policies; in the process, we extend the generation by
+execution techniques of Hri\c{t}cu et al.\cite{TestingNI:ICFP} to handle nested
+properties (\cref{sec:testing})\apt{Seems like some of the relevant support for this contribution is now cut?}
 \end{itemize}
@@
 We intentionally say little or nothing about other structures
-such as the heap.  (Heap safety has already been studied, in isolation,
-by Azevedo de Amorim et al., and we expect that their correctness condition and ours can be
-combined cleanly to characterize correct protection of stack + heap.)
+such as the heap.  Heap safety has already been studied, in isolation,
+by Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18}, and we expect that their correctness condition and ours can be
+combined cleanly to characterize correct protection of stack + heap.
 
@@
 3) tail calls,
-4) dynamic code generation, which is difficult to integrate with code annotations in general, and
+4) dynamic code generation, which is difficult to integrate with code annotations in general, or
 5) concurrency --- we can handle a limited form of coroutines, but
@@
 \medskip
-\paragraph*{Case Study: Micro-Policies}
+\section{Case Study: Micro-Policies}
 \label{sec:enforcement}
@@
 the formal stack-safety property described in the previous
-section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not..
-We later validate this claim using property-based random testing.
-
-\paragraph*{A Conservative Stack-Safety Enforcement Mechanism}
+section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not.
+We validate this claim using property-based random testing in \cref{sec:testing}.
+
+\paragraph*{Conservative Stack-Safety Enforcement Mechanism}
 %
@@
 \medskip
-\paragraph*{A Lazy Alternative}
+\paragraph*{Lazy Alternative}
 

### 2021-09-30 submission4/main.tex
a few small things

@@
   colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
-\label{fig:int1}
+\apt{We've lost the first line of main!}\label{fig:int1}
 \end{figure*}
@@
 %
+\apt{This whole section is indeed a bit verbose; could probably be condensed.}
 The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
@@
 on the {\em Depth Isolation} policy presented by
-Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}. In their policies,  the stack memory is tagged
-with ownership information associated to each stack frame (and to each
+Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}. In their policies, each stack location is tagged
+with ownership information associated to a stack frame (and to each
 separate object inside that frame, a more finely grained access
-control that we do not need to consider here), and tags registers containing
-stack pointers with access permission information. The policy also
+control that we do not need to consider here), and each register containing
+a stack pointer is tagged with access permission information. The policy also
 uses tags on instructions to identify the code
@@
 property\fi. For a program to enjoy stack safety
-(with respect to a particular call map), all we require is that
+(with respect to a particular call map\apt{?? what's that?}), all we require is that
 the entry sequence immediately follows any instruction marked as a call
@@
 
-Under the most optimized version of the micro-policy, {\em Lazy Tagging and Clearing},
-it does not initialize stack frames on entry \emph{or} clear them on exit,
+The most optimized version of the micro-policy, {\em Lazy Tagging and Clearing},
+does not initialize stack frames on entry \emph{or} clear them on exit,
 and it permits all writes to the stack, even when the $\PCname$ tag

### 2021-09-30 submission4/main.tex
Added exceptions to appendix

--- 2021-09-30 submission4/main.tex
@@
 \judgmenttwobrlong[]
-              {\(\codemap ~ (\mach ~ \PCname)\)}
+              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
 %             {\(\components = \sealcon ~ \mach\)}
@@
 systems. Second, we can further demonstrate
-the functionality of our definitions by showing that they can be used to prove
+the functionality of our definitions by using them to prove
 the correctness of a programming logic with strengthened rules around calls and returns.
@@
 
-%Text of appendix \ldots
+Our core formalism already models control flow patterns in which a return skips over one or more callers to
+reach a point further down the call stack. This is essentially the behavior of an exception; we need only add an
+annotation representing the {\tt setjump} operation or equivalent to enable a ``return'' to an arbitrary point
+not associated with a call. We add an annotation, {\sc SetJump}, and define a target label as either a call-return
+or an exception, \(\mathit{TL} ::= \text{\sc CR} + \text{\sc EX}\), which annotates targets in the return stack.
+\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times (\rts \in \listT (\mathit{TL} \times \target)) \]
+
+We add a rule to handle {\sc SetJump} annotations, and modify the {\sc Call} rule to insert {\sc CR} labels into the
+return stack.
+\judgment[]
+         {\(\codemap ~ (\mach ~ \PCname) = {\sc SetJump}\)}
+         {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc EX}, \retcon ~ \mach)::\rts\)}
+\judgmenttwobrlong[]
+              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
+              {\(\components = \{\component | \sealcon ~ \mach[\component]\}\)}
+              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc CR}, \retcon ~ \mach)::\rts\)}
+The return rules operate identically regardless of the target labels.
+
+\definition
+We define \textbf{\em stepwise stack integrity}, \textbf{\em stepwise stack confidentiality},
+\textbf{\em observational stack integrity}, \textbf{\em observational stack confidentiality},
+in the same way as in the core model, substituting the above \(\CONTEXTS\) type and update
+function for their original.
+
+\definition
+A system enjoys \textit{\textbf{well-bracketed control flow}} with exceptions when, for every reachable state
+\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
+\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
+if \(|\rts'| < |\rts|\), then either \(|\rts'| = |\rts|-1\) or \(\rts = \underscore +\!\!+ (\text{\sc EX},\underscore) +\!\!+ \rts'\).
 

### 2021-09-30 submission4/main.tex
Getting closer to 12 pages :-)

--- 2021-09-30 submission4/main.tex
@@
 
-What does ``stack safety'' mean, precisely? The phrase is associated with a
+What does ``stack safety'' mean? The phrase is associated with a
 variety of compiler,
@@
 Labels under the boxes indicate an assignment of the address to a
-{\em domain}, which is a particular set of intended access
+{\em domain}---a set of intended access
 rights. Unless otherwise marked, each address retains its domain as
-sequential instructions execute. %, as hinted by their colors.
+instructions execute. %, as hinted by their colors.
 %\bcp{where do we see this happening?  Is it important?}.
@@
 that the enforcement mechanism must adhere to.
-(We will introduce some other parts of the context as needed.)
 The possible domains
@@
 be treated properly later.
+(We will introduce some other parts of the context as needed.)
 
@@
 \paragraph*{Machine model}
-Our properties are defined in terms of an abstract machine model that makes very few
-assumptions about the concrete machine that instantiates it. We require a
-register machine with a program counter (\(\PCname\)) and there must be a well-defined
-mechanism for determining the bounds of a stack frame, and for recognizing when a call
-has returned. In this paper we discuss machines with a stack pointer (\(\rsp\)),
+Our properties are defined in terms of an abstract model that makes few
+assumptions about the concrete machine that instantiates it. We assume a
+register machine with a program counter (\(\PCname\)) and well-defined
+mechanisms for determining the bounds of a stack frame and recognizing when a call
+has returned. Our examples consider machines with a stack pointer (\(\rsp\)),
 in which the current stack frame is everything below the stack pointer and above the
-previous frame, and a call from instruction \(\addr_\PCname\) with stack pointer \(\addr_\rsp\)
+previous frame, and where a call from instruction \(\addr_\PCname\) with
+stack pointer \(\addr_\rsp\)
 is considered to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
-is again at \(\addr_\rsp\). But our properties apply equally to machines with
-other architectural conventions, such as those involving frame pointers.
-A machine may use its non-stack memory in any way, including
-as a heap, as long as the stack has static bounds. It might also possess additional
-state beyond its registers and memory addresses, such as the state of a hardware
+is again at \(\addr_\rsp\). However, our properties apply equally to
+other architectural conventions, such as frame pointers.
+The non-stack memory can be used in any way, including
+as a heap, as long as the stack has static bounds. There can be
+state other than registers and memory, such as the internal state of a hardware
 enforcement mechanism.
@@
 instruction that does all the
-required manipulation of the \(\PCname\) and stack pointer. Instead, there
+required manipulation of the \(\PCname\) and stack pointer. Instead, the
+same instruction opcode may be
+used in some places in a call or return sequence and in other places
+for other purposes.
+\ifspace
+For example, there
 might be just a simple branch instruction that stores the old \(\PCname\) in
 a register, with the rest of the call sequence performed by other
-instructions either before or after the jump.  Similarly, there may be no
-``return'' instruction, but rather a ``return sequence'' involving an
-indirect branch at some point.  Thus, the same instruction opcode may be
-used in some places as part of a call or return sequence and in other places
-for other purposes.
+instructions either before or after the jump.  Similarly, instead of a single
+``return'' instruction,  there may be a ``return sequence'' involving an
+indirect branch at some point.
+\fi
 
@@
 
-In this section we explore how an existing enforcement mechanism called {\em
+We next explore how an \ifspace existing \fi enforcement mechanism called {\em
   Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements

### 2021-09-30 submission4/main.tex
Answer some comments and describe exceptions

--- 2021-09-30 submission4/main.tex
@@
 \item We show that our definitions can be
-extended to describe argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
+extended to describe argument passing on the stack, callee-saves registers, exceptions, and a simple coroutine system (\cref{sec:ext}).
 \item We use property-based random testing to validate the relationship between
-  our properties and micro-policies; in the process, we extend the generation by
-execution techniques of Hri\c{t}cu et al.\cite{TestingNI:ICFP} to handle nested
-properties (\cref{sec:testing})\apt{Seems like some of the relevant support for this contribution is now cut?}
+  our properties and micro-policies (\cref{sec:testing})
 \end{itemize}
@@
 {\small
-  \begin{lstlisting}%[backgroundcolor=\color{gray!20}]
+  \begin{lstlisting}
 int main() {
@@
   colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
-\apt{We've lost the first line of main!}\label{fig:int1}
+\label{fig:int1}
 \end{figure*}
@@
 integration with a more general model of memory safety,
-2) exceptions, \sna{I could just write up how to do these...}
-3) tail calls,
-4) dynamic code generation, which is difficult to integrate with code annotations in general, or
-5) concurrency --- we can handle a limited form of coroutines, but
+2) tail calls,
+3) dynamic code generation, which is difficult to integrate with code annotations in general, or
+4) concurrency --- we can handle a limited form of coroutines, but
   not full-blown preemptive concurrency.
@@
 %
-\apt{This whole section is indeed a bit verbose; could probably be condensed.}
 The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
@@
 micro-policies can be defined using this scheme.
+Efficient execution of these micro-policies relies on hardware acceleration,
+such as the PUMP architecture~\cite{pump:asplos2015}.
 
@@
 
-Efficient execution of these micro-policies relies on hardware acceleration,
-such as the PUMP architecture~\cite{pump:asplos2015}.
-The hardware incorporates a rule cache
-to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
-the hardware traps to a software handler (running in a privileged context or
-on a co-processor) to compute the rule result. To obtain adequate performance,
-it is important to design micro-policies so that they hit in the cache as
-much as possible. Thus, practical policies need to maintain a small working
-set of distinct tags.
-%% %
-%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
-%% policies}
-
-%\paragraph*{A Conservative Tag Policy}
-%\label{sec:conservative}
-
-The micro-policy we present here to illustrate enforcement of stack
-safety as defined in \cref{sec:lse} is a small variation
+The micro-policy we reference here is a variation
 on the {\em Depth Isolation} policy presented by
-Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}. In their policies, each stack location is tagged
-with ownership information associated to a stack frame (and to each
-separate object inside that frame, a more finely grained access
-control that we do not need to consider here), and each register containing
-a stack pointer is tagged with access permission information. The policy also
-uses tags on instructions to identify the code
-sequences that have permission to manage the stack, say during calls and returns.
-%\leo{If we have the simple exposition so far, no need to mention the following sentence:}
-%In the following, for ease of exposition, we will assume that no
-%arguments are passed on the stack.
-%% \rb{Conceivably on alloc
-%%   operations, etc., which we have not yet mentioned.}
-
+Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}.
 The Depth Isolation micro-policy maintains tags of the form $\tagStackDepth{n}$ for
 each stack location belonging to the stack frame at activation depth
-$n$, or the tag $\tagNoDepth$ for stack locations that are currently
+$\depth$, or the tag $\tagNoDepth$ for stack locations that are currently
 unused (but could be used in the future).
 %
-The $\PCname$ is tagged similarly as $\tagPCDepth{n}$.
+The $\PCname$ is tagged similarly as $\tagPCDepth{\depth}$.
 %
@@
   the current
-  $\tagPCDepth{n}$ tag is incremented to $\tagPCDepth{(n + 1)}$, and the function
+  $\tagPCDepth{\depth}$ tag is incremented to $\tagPCDepth{(\depth + 1)}$, and the function
   entry sequence initializes  all locations of the new function frame with
-  tag $\tagStackDepth{(n + 1)}$.
+  tag $\tagStackDepth{(\depth + 1)}$.
 
@@
   the $\PCname$ tag is decremented.
-
 \end{itemize}
 
-This discipline suffices to enforce stack safety.
-To implement it,  we define blessed instruction sequences
-intended to appear at the entry and exit of each function,
+To implement this discipline, blessed instruction sequences are designated
+to appear at the entry and exit of each function,
 which manipulate tags as just described in addition to performing the
@@
 property\fi. For a program to enjoy stack safety
-(with respect to a particular call map\apt{?? what's that?}), all we require is that
+(with respect to a particular set of call annotations), all we require is that
 the entry sequence immediately follows any instruction marked as a call
@@
 
+\medskip
+
+\paragraph*{Exceptions}
+
+Exceptions are another useful form of control flow that disobeys the typical call and return pattern.
+We can extend our model easily to handle integrity and confidentiality in the presence of exceptions.
+In a typical exception model, the code point that ``catches'' a thrown exception is indicated by a {\tt setjump}
+instruction or equivalent, which is later jumped to by a {\tt longjump}. We treat {\tt setjump} as placing
+a special return target on the target stack, which then behaves like a normal return in terms of our properties.
+A full treatment is given in Appendix \ref{app:exceptions}.
+
 \section{Validation through Random Testing}

### 2021-09-30 submission4/main.tex
a couple of comments

@@
 %
+\apt{Could probably condense/omit this.}
 Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}
@@
 
-From here, we see two natural ways to extend the work. First, we should thoroughly evaluate
+From here, we see two natural ways to extend the work. \apt{what about addressing some of the limitations?}
+First, we should thoroughly evaluate
 other enforcement mechanisms, especially Cheri-based capability

### 2021-09-30 submission4/main.tex
Lots of gains with space hacking
wibble
12 pages (and some comments)

@@
     colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
+\vspace*{-1em}
 \label{fig:int1}
@@
 \caption{Noninterference, violated}
+\vspace*{-1em}
 \label{fig:conf1}
@@
 \caption{Another confidentiality violation}
+\vspace*{-1em}
 \label{fig:conf2}
@@
 
-Finally, we formalize WBCF as a simple property of the behavior of returns.
+Finally, WBCF is a \ifspace simple \fi property of the behavior of returns.
 \definition
@@
 collection of software-defined rules.
-At each step of the machine, the relevant rule is applied to the tags on
+On each machine step, the relevant rule is applied to the tags on
 the instruction's inputs ($\PCname$, registers, memory) and on the
-instruction opcode itself, producing
+\ifspace instruction \fi opcode \ifspace itself\fi, producing
 one of two outcomes: either the instruction is permitted to execute (and
@@
 We can think of the stepwise properties as instantiating
-the observation model with a view of the entire state, and
-thus as one end of a spectrum that can become arbitrarily permissive.
-Therefore, if we had to choose a single property to represent stack safety,
-it would be the observational formalization.
+the observation model with a view of the entire state---one end of a
+spectrum that can become arbitrarily permissive.
+Thus, if we had to choose a single property to represent stack safety,
+it would be the observational one.
 
@@
 We now show how to implement two kinds of protected
-sharing of stack contents: passing by value on the stack, and passing by reference. We first explore
+sharing of stack contents: passing by value on the stack, and passing by
+reference.
+\ifspace
+We first explore
 some examples and then extend our formalization.
+\fi
 
@@
   Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' elements
-  that are not protected, because they are capable of dynamic behavior that can compromise
+  that are not protected, because they are capable of compromising
   their own security. Consider a function that, when called with the wrong
@@
 %
-(1) {\em Intuitiveness}. Our property definitions are at least as
-intuitive as theirs, while being phrased in interestingly different terms.
+In terms of (1) {\em intuitiveness}, we contend that our property
+definitions are at least as
+intuitive as theirs, though phrased in rather different terms.
 We use the language of formal security
-properties to apply well-established concepts and formal techniques
-of noninterference, and we define our properties in terms of
-traces of memory and register states, rather than in terms of an
+properties to apply well-established concepts \ifspace and formal techniques \fi
+of noninterference, and we phrase our properties in terms of
+traces of memory and register states, rather than an
 ad-hoc abstract machine.
 Which approach is ``better'' is probably a matter of taste, but ours
-has at least the merit of explicitly relating stack safety to existing
+has \ifspace at least \fi the merit of explicitly relating stack safety to existing
 notions of security.
 %
-(2) {\em Usefulness} of the safety criterion
-for carrying out formal proofs (about individual programs). We agree that this is an important consideration;
-demonstrating it for our proposal remains future work, in which we
-plan to apply our properties to justifying a low-level program logic.
-%
-(3) {\em Reusability} in secure compilation chains.
-StkTokens is specifically set up to address this,
-while it is not a goal of our work.
-%
-(4) {\em Comprehensiveness} in capturing programmers' informal
-understanding of stack safety. We believe that our support
-for dynamically changing notion of adversaries, as outlined in point (i) above,
-captures a better definition of safety than LSE; moreover, our observational
-properties capture an even richer and more extensional notion of stack safety
-that can be applied to a broader range of enforcement mechanisms and code features.
-%
-(5) {\em Broad applicability} to richer settings. We have demonstrated that our approach extends to
+Demonstrating (2) {\em usefulness} of the safety criterion
+for carrying out formal proofs (about individual programs) for our proposal
+remains future work; concretely, we
+hope to apply our properties to justifying a low-level program logic.
+%
+(3) StkTokens is specifically set up to address {\em reusability} in secure
+compilation chains, while this is not a goal of our work.\bcp{Can we explain
+what it means??}
+%
+(4) Regarding {\em comprehensiveness} in capturing programmers' informal
+understanding of stack safety, we believe that our support
+for a dynamically changing notion of ``adversary,'' as outlined in point (i) above,
+captures an even better definition of safety than LSE; moreover, our observational
+properties are richer and\bcp{cut "richar and"??} more extensional, and thus
+applicable to a broader range of enforcement mechanisms and code features.
+%
+(5) Finally, regarding {\em broad applicability}, we have demonstrated that
+\bcp{unlike StkTokens}
+our approach extends to
 coroutines, a first step towards multithreading. Dynamic code generation is
@@
 
-Finally, we would propose one more criterion: reusability in evaluating multiple machines and
-enforcement mechanisms. They have used their approach to give strong validation
-to one machine/enforcement mechanism, a linear-capability machine,
-via formal proof. We have used ours to obtain a weaker, testing-based validation of
-two different mechanisms on a single machine, and will target a capability machine
+We would add one more criterion: reusability across machine models and
+enforcement mechanisms. StkTokens has been used to validatate
+one machine/enforcement mechanism---a linear-capability machine---via formal
+proof. We have used ours to obtain a weaker (testing-based) validation of
+two different mechanisms on a single machine; we hope to target a capability machine
 next as future work.
@@
 return stack.
+
 \judgment[]

### 2021-10-01 submission4/main.tex
Paragraph 2 skorst

--- 2021-09-30 submission4/main.tex
@@
 enforcement techniques and for validating that these claims
-are met by particular implementations.
+are met by particular implementations. With the exception of the model used
+in Skorstengaard et al.'s StkTokens work~\cite{SkorstengaardSTKJFP}, which we
+will discuss in Section \ref{sec:relwork}, there is a dearth of such properties.
 
@@
 at all.
-\apt{Originally, we were using this spot to contrast with Skorstengaard et al. in the intro.
-  Do we really need to do that? Reads better without!}
 
@@
 it means to return with a {\em return convention},
-%\(\retcon ~ (\mach \in \MACHS) ~ (\mach' \in \MACHS)\),
-\(\retcon \subseteq (\MACHS \times \MACHS)\),
-which relates call states to the states that can be considered valid returns to that call.
+\(\retcon ~ (\mach \in \MACHS) \subseteq \MACHS\), a function from states to sets of states,
+which maps call states to the states that can be considered valid returns to that call.
 The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
@@
 Up to this point we have ignored function arguments, by assuming that they are placed
-in registers (which are implicitly shared by all functions).
-We now show how to implement two kinds of protected
-sharing of stack contents: passing by value on the stack, and passing by reference. We first explore
-some examples and then extend our formalization.
-
-Example~\ref{fig:passingsimple} shows a sample trace of a program that passes
+in registers (which are shared by all functions).
+We now implement protected
+sharing of stack contents, as would be used for passing by value on the stack and for passing by reference.
+Here we give an example and high level discussion; the formalization can be found in Appendix \ref{app:argpass}.
+
+Example~\ref{fig:passing} shows a sample trace of a program that passes
 variables by reference. In this example, written in C++ notation,
@@
 and when {\tt f} instead tries to write to {\tt x}, it violates integrity.
-
-This extension to our model is formalized in Appendix \ref{app:argpass}.
 

### 2021-10-01 submission4/main.tex

--- 2021-09-30 submission4/main.tex
@@
 Function {\tt f} is called twice, at instructions 3 and 5; it should therefore return to instructions
-4 and 6, respectively. But the first time, before returning, it stashes its
-return pointer away, and the second time it returns to instruction 4 instead of 6; then it ends
-up being called a third time, at which point it finally returns to instruction 6.
+4 and 6, respectively. But during the first call it stashes its
+return pointer, and the second time it returns to instruction 4 instead of 6; it thus ends
+up being called a third time before finally returning to instruction 6.
 
@@
 We now show how to implement two kinds of protected
-sharing of stack contents: passing by value on the stack, and passing by
-reference.
-\ifspace
-We first explore
-some examples and then extend our formalization.
-\fi
-
-Example~\ref{fig:passingsimple} shows a sample trace of a program that passes
+sharing of stack contents: passing by value on the stack and passing by reference.
+Here we give an example and high level discussion; the formalization can be found in Appendix \ref{app:argpass}.
+
+Example~\ref{fig:passing} shows a sample trace of a program that passes
 variables by reference. In this example, written in C++ notation,

### 2021-10-01 submission4/main.tex
Fixed inference rule labels

@@
 
-\judgmenttwobr[ Yld]
+\begin{tabular}{l}
+  {\sc Yield} \\
+  \judgmenttwobr[]
                 {\(\mach \stepsto \mach'\)}
@@
                 {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
-             {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)}
+                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)} \\
+\end{tabular}
 
@@
 %
-\judgmenttwobrlong[ Call]
+\begin{tabular}{l}
+  {\sc Call} \\
+  \judgmenttwobrlong[]
                     {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
@@
                     {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
-                  {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+                    {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)} \\
+\end{tabular}
 The remaining rules are the same as in the simple model.
@@
 
-\judgmenttwo[ CallSeal]
+\begin{tabular}{l}
+  {\sc Call} \\
+  \judgmentbr[]
               {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
               {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}
+              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)} \\
+\end{tabular}
 
 \vspace*{-1ex}
-\judgmenttwo[ ReturnFound]
+\begin{tabular}{l}
+  {\sc ReturnFound} \\
+  \judgmenttwo[]
               {\(\mach \stepsto \mach' \in \target\)}
               {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}
+              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)} \\
+\end{tabular}
 
 \vspace*{-1ex}
-\judgmenttwo[ ReturnRec]
+\begin{tabular}{l}
+  {\sc ReturnRec} \\
+  \judgmenttwo[]
               {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
               {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)}
-
-\vspace*{-3ex}
-\judgment[ Default]
-         {}
-         {\(\underscore,\context \stepsto_C \context\)}
+              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)} \\
+\end{tabular}
 
@@
 
+\begin{tabular}{l}
+  {\sc SetJump} \\
 \judgment[]
@@
          {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc EX}, \retcon ~ \mach)::\rts\)}
+\end{tabular}
+
+\begin{tabular}{l}
+  {\sc Call} \\
   \judgmenttwobrlong[]
@@
                     {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc CR}, \retcon ~ \mach)::\rts\)}
-The return rules operate identically regardless of the target labels.
+                    {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc CR}, \retcon ~ \mach)::\rts\)} \\
+\end{tabular}
+
+The return rules operate identically to the core model, regardless of the target labels.
 

### 2021-10-01 submission4/main.tex
Tighten up stuff

@@
 This section explains our formal stack safety properties through a series
-of examples, each presented both in C syntax and in corresponding assembly code for a
-simple imaginary two-address machine (see Figure~\ref{fig:int1}).
-The address of each instruction is shown to its left.
+of examples, each presented both in C syntax and in corresponding assembly code for an
+imaginary two-address machine (see Figure~\ref{fig:int1}).
+Each instruction's address is shown to its left.
 The assembly code reflects a simple compilation strategy in which
@@
 At the right of each example, we show snapshots of the layout of stack memory just before
-selected instructions (labeled by their addresses) are executed; each box represents a memory word
-(the machine is word addressed), and addresses increase to the right.
+selected instructions (labeled by address) are executed; each box represents a memory word
+(the machine is word addressed). Addresses increase to the right.
 We refer to registers and addresses collectively as {\em state elements} or just elements.
@@
 \end{tabular}
+
 \begin{tabular}{l}
@@
              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'[\components \mapsto \unsealed],\rts'\)} \\
-\end{tabular}
-\begin{tabular}{l}
 \end{tabular}

### 2021-10-01 submission4/main.tex
Appendix labels

@@
 requires that all annotated instructions will be situated at the appropriate instruction
-of a {\em blessed sequence} -- a series of instructions that are tagged to enforce correct
+of a {\em blessed sequence} --- a series of instructions that are tagged to enforce correct
 behavior. It can then failstop if it would
@@
 We intentionally say little or nothing about other structures
-such as the heap.  Heap safety has already been studied, in isolation,
+such as the heap.  Heap safety has been studied in isolation
 by Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18}, and we expect that their correctness condition and ours can be
@@
 of its own dynamic instances), and {\em Depth Isolation}
-(\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
-temporal locality of stack memory and local call graphs, as well as information
-generated by the compilation toolchain, to strike various balances between
-precision and cacheability, evaluated through benchmarks that demonstrate
+(\cref{sec:enforcement}). All these policies strike various balances between
+precision and performance, evaluated through benchmarks that demonstrate
 limited performance overhead. In addition to the baseline ``eager'' policies,
@@
 
+\pagebreak
+
 %% Appendix
-\appendix
-\paragraph*{Formalization of Callee-saves Registers}
+\appendices
+\section{Formalization of Callee-saves Registers}
 \label{app:calleesave}
@@
 
-\paragraph*{Formalization of variable passing.}
-\ref{app:argpass}
+\section{Formalization of variable passing.}
+\label{app:argpass}
 
@@
 
-\paragraph*{Exceptions}
+\section{Formalization of Exceptions}
 

### 2021-10-01 submission4/main.tex
Written content for appendices

@@
 \label{app:calleesave}
-%
+
+In order to support callee-saves registers, we extend {\sc Call} annotations to
+carry information about which registers are to be saved and which are used for arguments,
+etc. This takes the form of a ``saving function'' \(S\) that maps registers to domains \(\saved\) and
+\(\unsaved\). These also extend \(\STKDOMS\).
+
 \[\begin{split}
@@
 \stackDom \in \STKDOMS ::= & \outside + \unsealed + \sealed{\depth} \\
-& + \unsaved + \saved
+& + \unsaved + \saved \\
 \end{split}\]
-%
+
+This property reuses most of the structure of the core property. Here we will highlight
+the definitions that have changed. Namely, the {\sc Call} rule for \(\stepsto_C\) now
+applies the saving function to determine which registers are saved. The remaining rules
+remain unchanged. Note that this means that at any given time, a register will be marked
+\(\saved\) or \(\unsaved\) depending on the most recent call chronologically (i.e., it is
+not nested like \(\sealed{\underscore}\)). This is because the protected and secret sets will be based
+only on the entry state of the call, and those are all that affect observational properties.
+
 \begin{tabular}{l}
@@
 \end{tabular}
-The remaining rules are the same as in the simple model.
 
@@
 \end{split}\]
-The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
-\times (\rts \in \listT (\target)) \]
-The initial context again maps all elements to \(\unsealed\).
+
 We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map at a call from depth \(\depth\).
+update a domain map at a call from depth \(\depth\), written as functions \(\mathit{push}\)
+and \(\mathit{pop}\) and written here as inference rules.
 
@@
 
-         At the matching return we just unseal the elements at depth \(\depth\):
-
 \judgment[]
@@
 
-Now we can inductively define the context update function, as before.
+These are applied to update the context at calls and returns, respectively.
 

### 2021-10-01 submission4/main.tex
Fix thingy
more reformatting of ctx definitions
pass through III and fix to code in fig. 5

@@
 
-We use these properties to validate the stack-safety {\em micro-policies}
-proposed by Roessler and DeHon. %~\cite{DBLP:conf/sp/RoesslerD18}.
+We use these properties to validate the stack-safety micro-policies
+proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
 Specifically, we check (with
@@
 capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}. \apt{Can we choose a bibstyle that
-  doesn't bracket each reference separately?}
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
   \ifaftersubmission\apt{Mostly from
@@
 enforcement techniques and for validating that these claims
-are met by particular implementations. With the exception of the model used
+are met by particular implementations. With the notable exception of the model used
 in Skorstengaard et al.'s StkTokens work~\cite{SkorstengaardSTKJFP}, which we
-will discuss in Section \ref{sec:relwork}, there is a dearth of such properties.
-
-We propose such a characterization using the tools of language-based
+discuss in Section \ref{sec:relwork}, there is a dearth of such properties in
+the literature.
+
+We propose a novel characterization of stack safety using the tools of language-based
 security~\cite{sabelfeld2003language}. The informal claim that
@@
 from its callee'' amounts to saying that it guarantees the {\em integrity} and
-{\em confidentiality} of the caller’s local state until it regains control.
+{\em confidentiality} of the caller’s local state until it regains control after the call.
 
@@
 flow (WBCF)}~\cite{SkorstengaardSTKJFP}---a global requirement that
-callees must always return to their immediate caller, if they return
+each callee must always return to its immediate caller, if it returns
 at all.
@@
 read or modified during a call. The observational variants
-allow a callee to read from and write to their caller's stack frame, as
-long as nothing that they read affects their observable behavior and nothing
-they write affects the observable behavior of the caller after they return.
+allow a callee to read from and write to its caller's stack frame, as
+long as nothing that it reads affects its observable behavior and nothing
+that it writes affects the observable behavior of the caller after it returns.
 The observational properties are more extensional, and represent a baseline
@@
 the function activation that owns the frame and access is
-then permitted only when an activation at that depth is currently executing, validates our
+permitted only when an activation at that depth is currently executing, validates our
 stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
@@
 supporting
-1) passing of scalar stack data,
-2) callee-saves registers calling conventions, and
-3) a coroutine system with a static layout.
+(1) passing of scalar stack data,
+(2) callee-saves registers calling conventions, and
+(3) a coroutine system with a static layout.
 
 We begin with an informal overview of our key ideas via
-examples (\cref{sec:running-example}), followed by a description of our key
+examples (\cref{sec:running-example}), followed by a description of our
 assumptions about the machine environment and threat model (\cref{sec:setup}).
@@
 \item We adapt concepts of integrity and confidentiality from language-based
-security, combine these with an existing notion of well-bracketed
+security and combine these with an existing notion of well-bracketed
 control flow to formalize a {\em stepwise stack-safety}
-property, and which captures the intuitive concept of
+property that captures the intuitive concept of
 strict stack safety enforced over each step of execution~(\cref{sec:prelim,sec:lse}).
@@
 \Cref{sec:relwork,sec:future} discuss related and future work.
+\apt{Say something about the Coq artifact?}
 
@@
 
-Labels under the boxes indicate an assignment of the address to a
+Labels under the boxes indicate an association of the address to a
 {\em domain}---a set of intended access
@@
 (in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
-(outside the stack entirely.)
+(outside the stack entirely).
 % These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
@@
 
-Figure~\ref{fig:int1} shows a program with simple integrity and confidentiality violations. It consists of a {\tt main}
-function that performs a call to a function {\tt f}, which reads and then overwrites {\tt main}'s
+Figure~\ref{fig:int1} shows a program with simple integrity and confidentiality violations. Function {\tt main}
+calls function {\tt f}, which reads and then overwrites {\tt main}'s
 variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
@@
 if a element is sealed before a step, it will be unchanged afterward.
+The write at line 12 violates this property.
 
@@
   int y;
-  if (!stash) {
+  if (stash == 0) {
     stash = *(&y-1);
@@
 11        cmp stash, $0  ; stash = 0?
-12        breq #15       ; branch if so
+12        brneq #15      ; branch if not
 13        mov -2(%sp),stash ; set stash
@@
 Consider Figure~\ref{fig:controlflow}.
-(Now we treat return addresses properly, but write them as line numbers, in the
+(Here we treat return addresses properly, but write them as line numbers, in the
 form $\#n$.)
@@
 Thus, another aspect of safe use of the stack is that calls
-and returns should be correctly nested---a caller should
-only return to its own callee.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
+and returns should be correctly nested---a callee should
+only return to its own caller.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
 call \emph{well-bracketed control flow (WBCF)}, and is orthogonal to
@@
 stack pointer \(\addr_\rsp\)
-is considered to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
+is considered to have returned when the \(\PCname\) is at \(\addr_\PCname + 4\) and the stack pointer
 is again at \(\addr_\rsp\). However, our properties apply equally to
@@
 Some calling convention choices, such as callee-saves registers and argument
-passing on the stack, are not supported directly by this property; we introduce
-secondary properties that use the same underlying formalism to handle these cases,
-in Section~\ref{sec:ext}.
+passing on the stack, are not supported directly by this property;
+in Section~\ref{sec:ext},
+we introduce
+secondary properties that use the same underlying formalism to handle these cases.
 
@@
 caller! We don't require the callee's cooperation at all.
+\ifaftersubmission\apt{Rest of this is repetitious with machine model section above.}\fi
 For example, under a standard calling convention, we can see that a
@@
 Machine states may contain arbitrary data, including data used exclusively by the
-enforcement mechanism, while our properties contend only with the data that are
+enforcement mechanism, while our properties refer only to the data that are
 relevant to normal execution: addresses and registers.
@@
 our properties are therefore naturally \emph{termination insensitive.}
-
+%
 That said, our model is meant to apply generically to any enforcement mechanism,
@@
 so we assume that any given machine will restrict its conditions a ``well-formedness''
-criterion on initial states, ensuring that they are set up properly. In the case of
+criterion on initial states, ensuring that they are set up properly. For eample, in the case of
 Roessler and DeHon's \cite{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-formedness
 requires that all annotated instructions will be situated at the appropriate instruction
-of a {\em blessed sequence} --- a series of instructions that are tagged to enforce correct
+of a {\em blessed sequence}---a series of instructions that are tagged to enforce correct
 behavior. It can then failstop if it would
@@
 We do not currently support:
-1) address-taken locals, as a satisfactory treatment of these seems to require
-integration with a more general model of memory safety,
-2) tail calls,
-3) dynamic code generation, which is difficult to integrate with code annotations in general, or
-4) concurrency --- we can handle a limited form of coroutines, but
+address-taken locals, as a satisfactory treatment of these seems to require
+integration with a more general model of memory safety;
+tail calls;
+dynamic code generation, which is difficult to integrate with code annotations in general; or
+concurrency---we can handle a limited form of coroutines, but
   not full-blown preemptive concurrency.
@@
 \[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-  \times (\listT (\target)) \]
+  \times \listT (\target) \]
 %
@@
 void f(int &a) {
-  if(!stash)
+  if (stash == 0)
     stash = &a;
@@
 20 f:     cmp stash, $0   ; stash = 0?
-21        breq #13        ; branch if so
+21        brneq #23       ; branch if not
 22        mov %ra,stash   ; stash = &a
-23        jmp #15
+23        jmp #25
 24        mov $5, (stash) ; *stash = 5
@@
 %
-\[\begin{split}
-\context \in \CONTEXTS ::= & (\domMap \in \COMPONENTS \rightarrow \TOPDOMS) \times \\
-& (\rtm \in \STACKS \rightarrow \listT(\target)) \times \\
-& (\yts \in \STACKS \rightarrow \target) \times (\stackid \in \STACKS)
-\end{split}\]
+\[\begin{array}{lr}
+\multicolumn{2}{l}{(\domMap, \rtm, \yts, \stackid) \in \CONTEXTS ::=} \\
+& (\COMPONENTS \rightarrow \TOPDOMS) \times
+(\STACKS \rightarrow \listT(\target)) \times 
+(\STACKS \rightarrow \target) \times \STACKS
+\end{array}\]
 
@@
 \section{Formalization of Exceptions}
+\label{app:exceptions}
 
@@
 or an exception, \(\mathit{TL} ::= \text{\sc CR} + \text{\sc EX}\), which annotates targets in the return stack.
-\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS) \times (\rts \in \listT (\mathit{TL} \times \target)) \]
+\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS) \times \listT (\mathit{TL} \times \target) \]
 

### 2021-10-01 submission4/main.tex
Fixed example 1

@@
 int f() {
-  int y, z;
-  z = *(&y-2);
+  int y;
+  y = *(&y-2);
   *(&y-2) = 0;
@@
 variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
-In the initial memory layout the entire stack is {\em unsealed},
-meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
-in this case 0. (Note that this does not include the return address, which is saved
-by the call instruction.)
-Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt x} will be unchanged when it gets control back,
-and usually at every step in between.
-This leads to a simple, intuitive statement of {\em stepwise integrity}:
-if a element is sealed before a step, it will be unchanged afterward.
-The write at line 12 violates this property.
-
-\begin{figure}
-\centering
-\addtolength{\tabcolsep}{-1em}
-\begin{tabular}{l | l}
-  \multicolumn{2}{c}{
-    2 \hspace{-1.5em}
-    \memoryaddrs{8em}
-    \memory{4}{\unsealc}
-    ~$\cdots$
-    \vspace{.5em}
-  } \\
-  11 \hspace{-1.5em}
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-  &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
-  \\
-  12 \hspace{-1.5em}
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-  &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{2}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
-  \\
-  13 \hspace{-1.5em}
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{42}
-  \MemoryLabel{-6em}{0.75em}{42}
-  \vspace{.5em}
-  &
-  \memoryaddrs{21em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
-  \vspace{.5em}
-\end{tabular}
-\addtolength{\tabcolsep}{1em}
-
-\vspace{\abovedisplayskip}
-
-\caption{Noninterference, violated}
-\vspace*{-1em}
-\label{fig:conf1}
-\end{figure}
-
-But before violating integrity, {\tt f} violates confidentiality by
-reading {\tt x}.
+
+First, {\tt f} violates confidentiality by reading {\tt x}.
 Confidentiality states intuitively that a callee is insensitive to
@@
 without changing the effect of our computation on non-secret values.
-% \bcp{That's too loose.  Even for standard NI, high inputs
-% are allowed to change high results.  And for us it's even more subtle, since
-% a given address can switch from high to low when a function returns.}
+
 If there exists some value for which we see different results, then
@@
 state element that changes in the original will do the same in the variant.
+
+Then, line 12 violates integrity.
+In the initial memory layout the entire stack is {\em unsealed},
+meaning that it is eligible to be written to. At the call from {\tt main} to
+{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
+in this case 0. (Note that this does not include the return address, which is saved
+by the call instruction.)
+Sealing is the fundamental integrity contract between the caller and callee:
+{\tt main} expects that {\tt f} will not change {\tt x}.
+This leads to a simple, intuitive statement of {\em stepwise integrity}:
+if a element is sealed before a step, it will be unchanged afterward.
+Overwriting {\tt x} violates this property.
+
+\begin{figure}
+\centering
+\addtolength{\tabcolsep}{-1em}
+\begin{tabular}{l | l}
+  \multicolumn{2}{c}{
+    2 \hspace{-1.5em}
+    \memoryaddrs{8em}
+    \memory{3}{\unsealc}
+    ~$\cdots$
+    \vspace{.5em}
+  } \\
+  11 \hspace{-1.5em}
+  \memoryaddrs{17em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-16em}{0.75em}{42}
+  &
+  \memoryaddrs{17em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{1}{\unsealc}
+  ~$\cdots$
+  \MemoryLabel{-16em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
+  \\
+  12 \hspace{-1.5em}
+  \memoryaddrs{17em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{1}{\badc}
+  ~$\cdots$
+  \MemoryLabel{-16em}{0.75em}{42}
+  \MemoryLabel{-6em}{0.75em}{42}
+  \vspace{.5em}
+  &
+  \memoryaddrs{17em}
+  \memory{1}{\unsealc}
+  \memory{1}{\retptrc}
+  \memory{1}{\badc}
+  ~$\cdots$
+  \MemoryLabel{-16em}{0.75em}{\(v_0\)}
+  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
+  \vspace{.5em}
+\end{tabular}
+\addtolength{\tabcolsep}{1em}
+
+\vspace{\abovedisplayskip}
+
+\caption{Noninterference, violated}
+\vspace*{-1em}
+\label{fig:conf1}
+\end{figure}
 
@@
 
-It may be unclear in Figure~\ref{fig:conf1} why we vary the initial values of {\tt y}
-and {\tt z} as well as that of {\tt x}. This is to detect some violations of so-called
+It may be unclear in Figure~\ref{fig:conf1} why we vary the initial value of {\tt y}
+as well as that of {\tt x}. This is to detect some violations of so-called
 {\em temporal stack safety}.

### 2021-10-01 submission4/main.tex
Messing with figures
Example reorder in text
Merged figures
Fixed alignment

@@
 \begin{figure*}
-\begin{subfigure}[t]{.2\textwidth}
+\begin{subfigure}[]{.16\textwidth}
 {\small
@@
 \end{subfigure}
-\begin{subfigure}[t]{.45\textwidth}
+\begin{subfigure}[]{.44\textwidth}
   {\small
     \begin{lstlisting}[backgroundcolor=\color{lgray}]
-1 main: add $1,%sp      ; allocate frame
-2       mov $42,-1(%sp) ; initialize local
+1 main: add $1,%sp      ;allocate frame
+2       mov $42,-1(%sp) ;initialize local
 3       call f
-4       add -1(%sp),%ra ; set return value
-5       sub $1,%sp      ; deallocate frame
+4       add -1(%sp),%ra ;set return value
+5       sub $1,%sp      ;deallocate frame
 6       ret
 
-10 f:   add $2,%sp      ; allocate frame
-11      mov -4(%sp),-1(%sp) ; violation 1!
-12      mov $0,-4(%sp)      ; violation 2!
-13      mov $5,%ra      ; set return value
-14      sub $1,%sp      ; deallocate frame
+10 f:   add $1,%sp      ;allocate frame
+11      mov -4(%sp),-1(%sp) ;violation 1!
+12      mov $0,-4(%sp)      ;violation 2!
+13      mov $5,%ra      ;set return value
+14      sub $1,%sp      ;deallocate frame
 15      ret
@@
 \end{subfigure}
-\begin{subfigure}[t]{.25\textwidth}
-\begin{center}
-\begin{tabular}{l l}
-2 &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}[\unsealdesc]
-
-~$\cdots$
-\vspace{.5em}
-\\
-3 &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}
-~$\cdots$
-    \MemoryLabel{-15em}{0.75em}{42}
-    \vspace{.5em}
-\\
-12 &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}[\sealdesc{0}]%
-\memory{1}{\retptrc}[\retptrdesc]%
-\memory{1}{\unsealc}[\unsealdesc]
-~$\cdots$
-\MemoryLabel{-16em}{0.75em}{42}
-\MemoryLabel{-7.5em}{0.75em}{42}
-\vspace{.5em}
-\\
-13 &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}
-\memory{1}{\retptrc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-15em}{0.75em}{\bf 0}
-\vspace{.5em}
-\end{tabular}
-\end{center}
-\vspace{-4ex}
-\end{subfigure}
-\caption{Integrity and confidentiality violations
-\ifaftersubmission\bcp{I wonder if we
-    could make the figures easier to read by putting very light, differently
-    colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
-\vspace*{-1em}
-\label{fig:int1}
-\end{figure*}
-
-\addtolength{\abovecaptionskip}{-1em}
-
-This section explains our formal stack safety properties through a series
-of examples, each presented both in C syntax and in corresponding assembly code for an
-imaginary two-address machine (see Figure~\ref{fig:int1}).
-Each instruction's address is shown to its left.
-The assembly code reflects a simple compilation strategy in which
-local variables are stored in consecutive slots of the stack frame.
-For the moment, arguments and return values are passed in registers.
-The stack grows upward in
-memory and {\tt \%sp} points to the first unused word above the top of the stack.
-The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
-implicitly adjusting the stack pointer.
-At the right of each example, we show snapshots of the layout of stack memory just before
-selected instructions (labeled by address) are executed; each box represents a memory word
-(the machine is word addressed). Addresses increase to the right.
-We refer to registers and addresses collectively as {\em state elements} or just elements.
-  %% \bcp{Most of the
-  %% figure is kind of self-explanatory with this help, but the rightmost part
-  %% is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
-  %% numbers in the boxes?  What are these brackets under the boxes, ...  One
-  %% simple thing we could do to help might be to label the right-hand column
-  %% something like ``Snapshots,'' or at least explain early on in the text
-  %% that this is what they are.}
-
-Labels under the boxes indicate an association of the address to a
-{\em domain}---a set of intended access
-rights. Unless otherwise marked, each address retains its domain as
-instructions execute. %, as hinted by their colors.
-%\bcp{where do we see this happening?  Is it important?}.
-Domains are not part of the machine state; rather, they are part
-of the {\em context} of a state that describes the security restrictions
-that the enforcement mechanism must adhere to.
-The possible domains
-are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
-(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
-(outside the stack entirely).
-% These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
-Global variables and code, which are not shown in the figures, would
-also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
-are shaded black and ignored for simplicity for now---they will
-be treated properly later.
-(We will introduce some other parts of the context as needed.)
-
-Figure~\ref{fig:int1} shows a program with simple integrity and confidentiality violations. Function {\tt main}
-calls function {\tt f}, which reads and then overwrites {\tt main}'s
-variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
-
-First, {\tt f} violates confidentiality by reading {\tt x}.
-Confidentiality states intuitively that a callee is insensitive to
-the context of its call (except for its arguments and any global
-variables).
-We state this as a form of {\em noninterference}~\cite{Goguen1982SecurityPA}.
-Noninterference describes knowledge: its standard form states that for state elements
-containing secret values, we can substitute any values whatsoever for the secrets
-without changing the effect of our computation on non-secret values.
-
-If there exists some value for which we see different results, then
-in fact we can learn something about the initial state by observing those results.
-To adapt this to the stack, the notion of ``secret'' must vary with function
-calls and returns.
-
-In Figure~\ref{fig:conf1} we see how noninterference applies to this scenario.
-At instruction 10, we
-create an imaginary {\em variant}
-state with arbitrary values~%
-%\ifaftersubmission\bcp{IMO, it would be easier to
-%  understand if we chose concrete values for the examples}\sna{FWIW I agree in
-%  theory, but it looks bad every time I try it.}\fi
-(\(v_0\), etc.),
-in place of the actual values ($42$, ...) that should
-be hidden from {\tt f}; we show the evolving stack memory of the variant in the second column.
-At instruction 11, the callee's variable, {\tt z},
-changes to 42 in the primary trace, but
-in the variant trace it instead might have any value \(v_0\). If \(v_0 \neq 42\), then the variant
-has behaved differently than the primary, violating confidentiality.
-
-We call this form of confidentiality {\em stepwise confidentiality}. Informally,
-for any state at the entry to a function, and for any variant over the values in the stack,
-the primary state and the variant will step in lockstep until they
-return. To step in lockstep
-means that if both states step---that is, neither the primary nor the variant failstops---any
-state element that changes in the original will do the same in the variant.
-
-Then, line 12 violates integrity.
-In the initial memory layout the entire stack is {\em unsealed},
-meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
-in this case 0. (Note that this does not include the return address, which is saved
-by the call instruction.)
-Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt f} will not change {\tt x}.
-This leads to a simple, intuitive statement of {\em stepwise integrity}:
-if a element is sealed before a step, it will be unchanged afterward.
-Overwriting {\tt x} violates this property.
-
-\begin{figure}
+\begin{subfigure}[]{.43\textwidth}
 \centering
@@
 \addtolength{\tabcolsep}{1em}
-
-\vspace{\abovedisplayskip}
-
-\caption{Noninterference, violated}
+\end{subfigure}
+\caption{Integrity and confidentiality violations
+\ifaftersubmission\bcp{I wonder if we
+    could make the figures easier to read by putting very light, differently
+    colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
 \vspace*{-1em}
 \label{fig:conf1}
-\end{figure}
+\end{figure*}
+
+\addtolength{\abovecaptionskip}{-1em}
 
 \begin{figure}
-%\begin{center}
-\begin{subfigure}[t]{.28\columnwidth}
+\begin{subfigure}{.4\columnwidth}
   {\small
@@
 }
-\end{lstlisting}
-  }
-\end{subfigure}
-\begin{subfigure}[t]{.27\columnwidth}
-  {\small
-\begin{lstlisting}[backgroundcolor=\color{lgray}]
+
 int f() {
@@
 }
-\end{lstlisting}
-  }
-\end{subfigure}
-\begin{subfigure}[t]{.32\columnwidth}
-  {\small
-\begin{lstlisting}
+
 int g() {
@@
   }
-\end{subfigure}
-\caption{Another confidentiality violation}
+\caption{Leaking through uninitialized data}
 \vspace*{-1em}
 \label{fig:conf2}
+\end{subfigure}
+\begin{subfigure}{.5\columnwidth}
+  \begin{center}
+\begin{tabular}{l l}
+2 &
+\memoryaddrs{8em}
+\memory{3}{\unsealc}[\unsealdesc]
+
+~$\cdots$
+\vspace{.5em}
+\\
+3 &
+\memoryaddrs{8em}
+\memory{3}{\unsealc}
+~$\cdots$
+    \MemoryLabel{-15em}{0.75em}{42}
+    \vspace{.5em}
+\\
+12 &
+\memoryaddrs{16em}
+\memory{1}{\mainsealc}[\sealdesc{0}]%
+\memory{1}{\retptrc}[\retptrdesc]%
+\memory{1}{\unsealc}[\unsealdesc]
+~$\cdots$
+\MemoryLabel{-16em}{0.75em}{42}
+\MemoryLabel{-7.5em}{0.75em}{42}
+\vspace{.5em}
+\\
+13 &
+\memoryaddrs{16em}
+\memory{1}{\mainsealc}
+\memory{1}{\retptrc}
+\memory{1}{\unsealc}
+~$\cdots$
+\MemoryLabel{-15em}{0.75em}{\bf 0}
+\vspace{.5em}
+\end{tabular}
+\end{center}
+\caption{Integrity Trace}
+\label{fig:int1}
+\end{subfigure}
 \end{figure}
+
+This section explains our formal stack safety properties through a series
+of examples, each presented both in C syntax and in corresponding assembly code for an
+imaginary two-address machine (see Figure~\ref{fig:int1}).
+Each instruction's address is shown to its left.
+The assembly code reflects a simple compilation strategy in which
+local variables are stored in consecutive slots of the stack frame.
+For the moment, arguments and return values are passed in registers.
+The stack grows upward in
+memory and {\tt \%sp} points to the first unused word above the top of the stack.
+The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
+implicitly adjusting the stack pointer.
+At the right of each example, we show snapshots of the layout of stack memory just before
+selected instructions (labeled by address) are executed; each box represents a memory word
+(the machine is word addressed). Addresses increase to the right.
+We refer to registers and addresses collectively as {\em state elements} or just elements.
+  %% \bcp{Most of the
+  %% figure is kind of self-explanatory with this help, but the rightmost part
+  %% is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
+  %% numbers in the boxes?  What are these brackets under the boxes, ...  One
+  %% simple thing we could do to help might be to label the right-hand column
+  %% something like ``Snapshots,'' or at least explain early on in the text
+  %% that this is what they are.}
+
+Labels under the boxes indicate an association of the address to a
+{\em domain}---a set of intended access
+rights. Unless otherwise marked, each address retains its domain as
+instructions execute. %, as hinted by their colors.
+%\bcp{where do we see this happening?  Is it important?}.
+Domains are not part of the machine state; rather, they are part
+of the {\em context} of a state that describes the security restrictions
+that the enforcement mechanism must adhere to.
+The possible domains
+are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
+(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
+(outside the stack entirely).
+% These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
+Global variables and code, which are not shown in the figures, would
+also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
+are shaded black and ignored for simplicity for now---they will
+be treated properly later.
+(We will introduce some other parts of the context as needed.)
+
+Figure~\ref{fig:int1} shows a program with simple integrity and confidentiality violations. Function {\tt main}
+calls function {\tt f}, which reads and then overwrites {\tt main}'s
+variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
+
+First, {\tt f} violates confidentiality by reading {\tt x}.
+Confidentiality states intuitively that a callee is insensitive to
+the context of its call (except for its arguments and any global
+variables).
+We state this as a form of {\em noninterference}~\cite{Goguen1982SecurityPA}.
+Noninterference describes knowledge: its standard form states that for state elements
+containing secret values, we can substitute any values whatsoever for the secrets
+without changing the effect of our computation on non-secret values.
+
+If there exists some value for which we see different results, then
+in fact we can learn something about the initial state by observing those results.
+To adapt this to the stack, the notion of ``secret'' must vary with function
+calls and returns.
+
+In Figure~\ref{fig:conf1} we see how noninterference applies to this scenario.
+At instruction 10, we
+create an imaginary {\em variant}
+state with arbitrary values~%
+%\ifaftersubmission\bcp{IMO, it would be easier to
+%  understand if we chose concrete values for the examples}\sna{FWIW I agree in
+%  theory, but it looks bad every time I try it.}\fi
+(\(v_0\), etc.),
+in place of the actual values ($42$, ...) that should
+be hidden from {\tt f}; we show the evolving stack memory of the variant in the second column.
+At instruction 11, the callee's variable, {\tt z},
+changes to 42 in the primary trace, but
+in the variant trace it instead might have any value \(v_0\). If \(v_0 \neq 42\), then the variant
+has behaved differently than the primary, violating confidentiality.
+
+We call this form of confidentiality {\em stepwise confidentiality}. Informally,
+for any state at the entry to a function, and for any variant over the values in the stack,
+the primary state and the variant will step in lockstep until they
+return. To step in lockstep
+means that if both states step---that is, neither the primary nor the variant failstops---any
+state element that changes in the original will do the same in the variant.
 
@@
 an arbitrary value.
+
+Then, line 12 violates integrity.
+In the initial memory layout the entire stack is {\em unsealed},
+meaning that it is eligible to be written to. At the call from {\tt main} to
+{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
+in this case 0. (Note that this does not include the return address, which is saved
+by the call instruction.)
+Sealing is the fundamental integrity contract between the caller and callee:
+{\tt main} expects that {\tt f} will not change {\tt x}.
+This leads to a simple, intuitive statement of {\em stepwise integrity}:
+if a element is sealed before a step, it will be unchanged afterward.
+Overwriting {\tt x} violates this property.
 

### 2021-10-02 submission4/main.tex
fix figure refs in II

@@
 \end{subfigure}
-\caption{Integrity and confidentiality violations
+\caption{Integrity and confidentiality violations with confidentiality traces
 \ifaftersubmission\bcp{I wonder if we
@@
   }
-\caption{Leaking through uninitialized data}
+\vspace*{-1em}
+\caption{Leaking via uninitialized data}
 \vspace*{-1em}
@@
 \end{subfigure}
+\caption{}
 \end{figure}
@@
 of examples, each presented both in C syntax and in corresponding assembly code for an
-imaginary two-address machine (see Figure~\ref{fig:int1}).
+imaginary two-address machine (e.g., see Figure~\ref{fig:conf1}).
 Each instruction's address is shown to its left.
@@
 
-Figure~\ref{fig:int1} shows a program with simple integrity and confidentiality violations. Function {\tt main}
+Figure~\ref{fig:conf1} shows a program with simple integrity and confidentiality violations. Function {\tt main}
 calls function {\tt f}, which reads and then overwrites {\tt main}'s
@@
 without changing the effect of our computation on non-secret values.
-
+%
 If there exists some value for which we see different results, then
@@
 
-In Figure~\ref{fig:conf1} we see how noninterference applies to this scenario.
+At the right of Figure~\ref{fig:conf1}, we see how noninterference applies to this scenario.
 At instruction 10, we
@@
 
-Then, line 12 violates integrity.
+Returning to the code in Figure~\ref{fig:conf1}, line 12 violates integrity.
 In the initial memory layout the entire stack is {\em unsealed},

### 2021-10-02 submission4/main.tex
add reference to Fig 2b

@@
 \end{subfigure}
-\caption{Integrity and confidentiality violations with confidentiality traces
+\caption{Integrity and confidentiality violations with confidentiality violation traces
 \ifaftersubmission\bcp{I wonder if we
@@
 
-\addtolength{\abovecaptionskip}{-1em}
+%\addtolength{\abovecaptionskip}{-1em}
 
@@
 }
-
 int f() {
@@
 }
-
 int g() {
@@
 \end{center}
-\caption{Integrity Trace}
+\caption{Integrity Violation Trace}
 \label{fig:int1}
@@
 if a element is sealed before a step, it will be unchanged afterward.
-Overwriting {\tt x} violates this property.
+Overwriting {\tt x} violates this property, as shown in Figure~\ref{fig:int1}.
 

### 2021-10-02 submission4/main.tex
oops; revert change and get back to 12 pages
not all components should have become elements, and some elements should have become something else

@@
 
-%\addtolength{\abovecaptionskip}{-1em}
+\addtolength{\abovecaptionskip}{-1em}
 
@@
 \end{subfigure}
+\vspace*{1em}
 \caption{}
@@
 This leads to a simple, intuitive statement of {\em stepwise integrity}:
-if a element is sealed before a step, it will be unchanged afterward.
+if an element is sealed before a step, it will be unchanged afterward.
 Overwriting {\tt x} violates this property, as shown in Figure~\ref{fig:int1}.
@@
 and the context with \(\pi_\context\).
-We take the first element of a trace with \(\head(\MPCT)\), which is a total
-function since traces are non-empty, and the final element (if one exists) with
+We extract the first entry from a trace with \(\head(\MPCT)\), which is a total
+function since traces are non-empty, and the final entry (if one exists) with
 \(\last(\MPCT)\), which is partial.
@@
 and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first element on which \(f\) holds.
+\(\MPCT\) ending with the first trace entry on which \(f\) holds.
 
@@
 If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final element. If the predicate never holds, then
+predicate must hold on its final entry. If the predicate never holds, then
 \(\PUT\) is the identity function.
@@
   every} caller should be protected from its callees.  By contrast,
-  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' elements
+  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
   that are not protected, because they are capable of compromising
@@
   inputs, behaves unreasonably. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
-  {\em reasonable element} (as defined in their section 4.3.3), as a reasonable
-  element is always reasonable. In our setting, individual
+  {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
+  component is always reasonable. In our setting, individual
   well-behaved activations of the buggy function would be protected.

### 2021-10-02 submission4/main.tex
fix cross ref from IX to VI

@@
 complex enforcement mechanism like the Depth Isolation micro-policy.
-We therefore choose to validate the micro-policy of the previous section by
+We therefore choose to validate the micro-policy of \cref{sec:enforcement} by 
 systematically \emph{testing} that it satisfies our properties.
@@
   schemes, each such port will require a new abstract machine definition whose
-  safety-by-construction must be checked afresh.
+  safety-by-construction must be argued afresh.
   Since we abstract over most of the machine state as well as the sealing and return conventions, and
@@
 %
-Heap safety, like stack safety, may be framed as a security property using
-the language of
-noninterference~\cite{DBLP:conf/post/AmorimHP18}. Their paper was a significant inspiration
+Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18} show that 
+heap safety may be framed as a security property using
+the language of noninterference. Their paper was a significant inspiration
 for the present work, and we believe that the two safety conditions should
@@
 to clear large swaths of the stack. We plan to implement a model of the stack safety
-system described in Georges et al. \cite{Georges+21} and test it against our properties using
+system described by Georges et al. \cite{Georges+21} and test it against our properties using
 QuickChick.

### 2021-10-02 submission4/main.tex
pass through V

@@
 \(\REGS\), which is assumed to include a special-purpose register, the
-program counter {\PCname}. Collectively addresses and registers are {\em components} \(\component\)
+program counter {\PCname}. Collectively addresses and registers are {\em state elements} \(\component\)
 in the set \(\COMPONENTS ::= \WORDS + \REGS\).
 %
-A {\em machine state} \(\mach\) is drawn from a set \(\MACHS\) and supports a mapping
-from components to values, written
-\(\mach[\component] \in \MACHS \rightarrow \COMPONENTS \rightarrow \WORDS \).
+A {\em machine state} \(\mach\) is drawn from a set \(\MACHS\) and defines a mapping
+from state elements to values, written
+\(\mach[.] \in \COMPONENTS \rightarrow \WORDS \).
 
@@
 of MC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces and ``::'' for list-cons.
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::'' for list-cons.
 
@@
 function since traces are non-empty, and the final entry (if one exists) with
-\(\last(\MPCT)\), which is partial.
+\(\last(\MPCT)\), which is partial, since traces may be infinite.
 
@@
 If \(f\) is a predicate on machine states, we define \(\mpcstatename
-\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)''):
+\hookrightarrow \MPCT | f\) (pronounced ``\(\MPCT\) is the prefix of
+running \(\mpcstatename\) up to \(f\)'') thus:
 %
@@
 individual calls from a trace by taking subtraces in which the stack is
-at or above a given depth. If \(f\) is still a predicate on states, then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
+at or above a given depth. If \(f\) is again a predicate on states, then
+we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\) (pronounced
+``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\)'') thus:
 \ifspace{\bcp{Throughout, we could save space and improve beauty by removing some of the
@@
 properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if for some initial state \(\mpcstatename = \mach,\context_0\)
+is {\em reachable} if, for some initial state \(\mpcstatename = \mach,\context_0\)
 where \(\mach\) is well-formed, \(\mpcstatename \hookrightarrow \MPCT\)
@@
 
-We are finally ready for our definition of stack safety. We begin
+We are finally ready to give our definition of stack safety. We begin
 by describing how we model knowledge about the program structure,
@@
 which addresses must be protected until the return. Similarly, we abstract over what
-it means to return with a {\em return convention},
-\(\retcon ~ (\mach \in \MACHS) \subseteq \MACHS\), a function from states to sets of states,
-which maps call states to the states that can be considered valid returns to that call.
+it means to return using a {\em return convention},
+\(\retcon : \MACHS \rightarrow 2^{\MACHS}\), which maps each call state 
+to the set of states that can be considered valid returns to that call.
 The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
@@
 %
-The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
-the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
+The initial context \(\context_0 = (\domMap_0, [])\), where \(\domMap_0\) maps
+the stack to \(\unsealed\) and all other elements to \(\outside\).
 
@@
 unseals each frame down to the level that was returned to.
-If none of these rules apply, the default rule leaves context unchanged.
+If none of these rules apply, the context remains unchanged.
 
@@
 
+\noindent
 This is the set of all possible calls that can be reached from any well-formed initial state.
@@
 A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\context\),
+for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = (\mach,\context\)),
 any \(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\), and any trace

### 2021-10-02 submission4/main.tex
pass through VII

@@
 \[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{\MPCT, \neg f_\depth | \MPCT\langle f_\depth \rangle
+\bigcup_{\depth \in \mathbb{Z}} \{(\MPCT, \neg f_\depth) | \MPCT\langle f_\depth \rangle
 \text{ is a reachable segment}\} \hspace{1em} \\
@@
 A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = (\mach,\context\)),
+for any \((\MPCT,\brk) \in \MPCT^*\) with \(\head(\MPCT) = (\mach,\context\)),
 any \(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\), and any trace
 %\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
-\(\NPCT\) such that \(\nach,\context \hookrightarrow \NPCT | \brk\),
+\(\NPCT\) such that \((\nach,\context) \hookrightarrow \NPCT | \brk\),
 \(\MPCT \doteq \NPCT\) holds.
@@
 section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not.
-We validate this claim using property-based random testing in \cref{sec:testing}.
+We validate these claims using property-based random testing in \cref{sec:testing}.
 
@@
 the instruction's inputs ($\PCname$, registers, memory) and on the
-\ifspace instruction \fi opcode \ifspace itself\fi, producing
+\ifspace instruction \fi opcode\ifspace itself\fi, producing
 one of two outcomes: either the instruction is permitted to execute (and
@@
 from their associated values at the hardware level. So they are not modeled
-as values, and therefore our properties make no claims about
-their behavior.
-
-The micro-policy we reference here is a variation
+as state elements, and therefore our properties make no claims about
+their values.
+
+The micro-policy we describe here is a variation
 on the {\em Depth Isolation} policy presented by
@@
 
-This \emph{lazy policy} admits more a efficient implementations, but
+This \emph{lazy policy} admits a more efficient implementation, but
 deliberately allows temporary violations of stack integrity.
@@
 \(\tau\). Our machine step function now takes a state and returns a state and an observation,
-which is carried over into the policy and context step functions:
+which is carried over into the policy and context step functions.
 
@@
 \definition
-\textbf{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
-where \(\head(\MPCT) = \mach\context\) and \(\last(\MPCT) = \mach',\context'\),
-the following holds. Let \(\components\) be the set of protected elements that changed during
-the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
+\textbf{\em Observational integrity} means that for any \((\MPCT,\brk) \in \MPCT^*\)
+where \(\head(\MPCT) = (\mach,\context\)) and \(\last(\MPCT) = (\mach',\context')\),
+the following holds. Let \(\components = \prot(\context) \cap \Delta(\mach, \mach')\)
+be the set of protected elements that changed during
+the call,
 and let \((\mach',\context') \hookrightarrow \MPCT'\).
@@
 \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all elements
-\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
+\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) such that
 \(\mach'[\component] \not = \nach'[\component]\).
@@
 \definition
-A system enjoys \textbf{\em observational confidentiality} if for all
-\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\context\) and
+A system enjoys \textbf{\em observational confidentiality} if, for all
+\((\MPCT,\brk) \in \MPCT^*\) with \(\head(\MPCT) = (\mach,\context)\) and
 \(\components = \secret(\context)\), the following holds.
@@
 (\last(\NPCT))\), i.e. if one call returns, so does the other
-\item Assuming both calls return, let \(\last(\MPCT) = \mach',\context'\)
+\item Assuming both calls return, let \(\last(\MPCT) = (\mach',\context')\)
   and \(\pi_\mach(\last(\NPCT)) = \nach'\),
@@
   and let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
-  Then if \(\mach',\context' \hookrightarrow \MPCT'\) and
-  \(\nach',\context' \hookrightarrow \NPCT'\), it must be
+  Then if \((\mach',\context') \hookrightarrow \MPCT'\) and
+  \((\nach',\context') \hookrightarrow \NPCT'\), it must be
   that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
@@
 instances of the same function.) We validate this repaired micro-policy in
-\ref{sec:testing}.
+\cref{sec:testing}.
 
@@
 
-We begin by adding information to Call annotations defining which registers will be {\em saved}
+We model this by adding information to Call annotations defining which registers will be {\em saved}
 and which are {\em unsaved}. The stack pointer and the register containing the return address
-should never be saved. We then extend the set of domains with these values, for the first time applying
+should never be saved. We then extend the set of domains with these values, for the first time associating
 (non-{\em outside}) domains to registers as well as addresses. On each call, we save precisely those

### 2021-10-02 submission4/main.tex
pass through XIII

@@
 
-Example~\ref{fig:passing} shows a sample trace of a program that passes
+Figure~\ref{fig:passing} shows a sample trace of a program that passes
 variables by reference. In this example, written in C++ notation,
@@
 sharing behavior of such pointers in terms of stack frame activation patterns.
-In general we expect such a pointer to behave similarly to other pointers in
-the system, except that its lifespan is connected to that of the frame that contains
-it. In order to give a satisfactory property, then, we would tie it to an existing (e.g.)
+In general we expect such a pointer to behave similarly to heap pointers, 
+except that its lifespan is connected to that of the frame that contains
+it. In order to give a satisfactory property, then, we should tie it to a broader
 heap safety property.
 
-The default otherwise would be to treat an address-taken variable in the
-same way as a passed argument, with the additional property
+An alternative is to treat every address-taken variable in the
+same way as a passed argument, with the modification 
 that it remains accessible at \emph{all} stack depths above the caller.
-This prevents most violations of temporal safety, although stale pointers
-can still be used in certain corner cases. Callees would be permitted
+Callees would be permitted
 to access the location without having been explicitly given a pointer
-to it.
+to it. This would prevent most violations of temporal safety, but stale pointers
+could still be used in certain corner cases.
 
@@
 \(\instack{\stackid}{\unsealed}\), the initial return map maps all stack identifiers to empty
-return stacks, the initial yield targets are defined above, and the initial stack id matches the initial
+return stacks, the initial yield targets are defined as above, and the initial stack id matches the initial
 coroutine. Together these form the initial context.
@@
 \[\begin{aligned}
-\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{\MPCT, \neg f_\stackid | \MPCT\langle f_\stackid \rangle
+\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{(\MPCT, \neg f_\stackid) | \MPCT\langle f_\stackid \rangle
 \text{ is a reachable segment}\} \hspace{1em} \\
@@
 \definition
-Let \(\prot_Y(\domMap,\rtm,\yts,\stackid)\) be the set of elements \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\),
-and \(\secret_Y(\domMap,\rtm,\yts,\stackid)\) the same.
+Let both \(\prot_Y(\domMap,\rtm,\yts,\stackid)\) and \(\secret_Y(\domMap,\rtm,\yts,\stackid)\) be the set of elements \(\component\) such that
+\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
 

### 2021-10-02 submission4/main.tex
no more hyphenated stack-safety

@@
 
-We use these properties to validate the stack-safety micro-policies
+We use these properties to validate the stack safety micro-policies
 proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
@@
 properties to validate and improve an existing enforcement mechanism, the
-{\em stack-safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
+{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
 in the Coq proof assistant on top of a RISC-V specification.  We
@@
 security and combine these with an existing notion of well-bracketed
-control flow to formalize a {\em stepwise stack-safety}
+control flow to formalize a {\em stepwise stack safety}
 property that captures the intuitive concept of
@@
 
-Since our stack-safety properties are all about protecting ``callers'' from
+Since our stack safety properties are all about protecting ``callers'' from
 their ``callees,'' we need some clear-cut way to recognize when the program
@@
   Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements
-the formal stack-safety property described in the previous
+the formal stack safety property described in the previous
 section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not.
@@
 
-\paragraph*{Conservative Stack-Safety Enforcement Mechanism}
+\paragraph*{Conservative Stack Safety Enforcement Mechanism}
 %
@@
 There remains the question of how to ensure that the sequences are
-invoked at the right places\ifspace from the perspective of the stack-safety
+invoked at the right places\ifspace from the perspective of the stack safety
 property\fi. For a program to enjoy stack safety
@@
 If the code fails to initiate an exit
-sequence at a point where the stack-safety property expects a return,
+sequence at a point where the stack safety property expects a return,
 the micro-policy will incorrectly behave as if execution
@@
 confidentiality. We use a Coq specification of the RISC-V
-architecture~\cite{Bourgeat2021AMF}.
-and extend it with a runtime monitor implementing a stack-safety
+architecture~\cite{Bourgeat2021AMF},
+and extend it with a runtime monitor implementing a stack safety
 micro-policy. We chose the Coq proof assistant as the setting for our implementation

### 2021-10-02 submission4/main.tex

@@
 micro-policy. We chose the Coq proof assistant as the setting for our implementation
-to ensure that our coinductive trace definitions are well-formed, and reason about
+to ensure that our coinductive trace definitions are well-formed, to reason about
 them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.
@@
 machines towards interesting behaviors. We extend the technique by tracking
-control flow information so that programs will usually --- but not always ---
-return to their call sites.
+control flow information so that programs will usually---but not always---return to their call sites.
 
@@
 %
-We need to further extend typical testing to handle the nested
+We need to extend typical testing schemes further to handle the nested
 nature of confidentiality: rather than just generating two
@@
 from trivial typos to ones that require an intriguingly complex setup
-to appear.  The most interesting bug (included in the table as row
+to provoke.  The most interesting bug (included in the table as row
 {\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
@@
 fails in testing, in cases where data is leaked between sequential calls.
-To round out our mutations testing we also check {\em LOAD\_NO\_CHECK\_LT},
+To round out our mutation testing we also check {\em LOAD\_NO\_CHECK\_LT},
 equivalent to its counterpart in depth isolation,

### 2021-10-02 submission4/main.tex
finished full pass

@@
 
+\vspace*{1em}
 \begin{tabular}{l}
@@
 \(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
-if \(|\rts'| < |\rts|\), then either \(|\rts'| = |\rts|-1\) or \(\rts = \underscore +\!\!+ (\text{\sc EX},\underscore) +\!\!+ \rts'\).
+if \(|\rts'| < |\rts|\), then either \(|\rts'| = |\rts|-1\) or \(\rts = \underscore +\!\!+ (\text{\sc EX},\underscore)::\rts'\).
 

### 2021-10-02 submission4/main.tex
Referenced artifact and fixed figure numbers

--- 2021-10-02 submission4/main.tex
@@
 violates the temporal aspect of confidentiality in
-corner cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak happens to use the caller's frame. We
+cases where data can leak across repeated calls to the same callee,
+and also violates integrity if the leak uses the caller's frame. We
 propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
@@
 \Cref{sec:relwork,sec:future} discuss related and future work.
-\apt{Say something about the Coq artifact?}
+
+Our artifact contains Coq formalizations of our properties for illustrative purposes
+as well as the testing implementation.
 
@@
 \vspace*{-1em}
-\caption{Leaking via uninitialized data}
+\caption*{Fig. \arabic{figure}: Leaking via\\ uninitialized data}
+\addtocounter{figure}{1}
 \vspace*{-1em}
@@
 \end{center}
-\caption{Integrity Violation Trace}
+\caption*{Fig. \arabic{figure}: Integrity Violation Trace}
 \label{fig:int1}
@@
 \vspace*{1em}
-\caption{}
+\vspace*{-2em}
 \end{figure}

### 2021-10-02 submission4/main.tex
Fixed figure labels in text

@@
 {\em temporal stack safety}.
-Consider the code in Figure~\ref{fig:conf2}.
+Consider the code in Figure~2.
 Using the same compilation strategy as before, {\tt y} and {\tt z} will end up occupying the
@@
 if an element is sealed before a step, it will be unchanged afterward.
-Overwriting {\tt x} violates this property, as shown in Figure~\ref{fig:int1}.
+Overwriting {\tt x} violates this property, as shown in Figure~3.
 

### 2021-12-21 submission4/main.tex
Last submission

@@
 
+\makeatletter
+\newcommand{\linebreakand}{%
+  \end{@IEEEauthorhalign}
+  \hfill\mbox{}\par
+  \mbox{}\hfill\begin{@IEEEauthorhalign}
+}
+\makeatother
+
 \begin{document}
@@
 
-%\author{
-%  \IEEEauthorblockN{
-%    Sean Noble Anderson
-%  }
-%  \IEEEauthorblockA{
-%    Portland State University\\
-%    ander28@pdx.edu
-%  }
-%  \and
-%  \IEEEauthorblockN{
-%    Leonidas Lampropoulos,
-%  }
-%  \IEEEauthorblockA{
-%    University of Maryland, College Park\\
-%    leonidas@umd.edu
-%  }
-%  \and
-%  \IEEEauthorblockN{
-%    Roberto Blanco,
-%  }
-%  \IEEEauthorblockA{
-%    Max Planck Institute for Security and Privacy\\
-%    roberto.blanco@mpi-sp.org
-%  }
-%  \and
-%  \IEEEauthorblockN{
-%    Benjamin C. Pierce,
-%  }
-%  \IEEEauthorblockA{
-%    University of Pennsylvania\\
-%    bcpierce@cis.upenn.edu
-%  }
-%  \and
-%  \IEEEauthorblockN{
-%    Andrew Tolmach
-%  }
-%  \IEEEauthorblockA{
-%    Portland State University\\
-%    tolmach@pdx.edu
-%  }
-%}
+\author{
+  \IEEEauthorblockN{
+    Sean Noble Anderson
+  }
+  \IEEEauthorblockA{
+    Portland State University\\
+    ander28@pdx.edu\\
+  }
+  \and
+  \IEEEauthorblockN{
+    Leonidas Lampropoulos
+  }
+  \IEEEauthorblockA{
+    University of Maryland, College Park\\
+    leonidas@umd.edu\\
+  }
+  \and
+  \IEEEauthorblockN{
+    Roberto Blanco
+  }
+  \IEEEauthorblockA{
+    Max Planck Institute for Security and Privacy\\
+    roberto.blanco@mpi-sp.org\\
+  }
+  \linebreakand
+  \IEEEauthorblockN{
+    Benjamin C. Pierce,
+  }
+  \IEEEauthorblockA{
+    University of Pennsylvania\\
+    bcpierce@cis.upenn.edu\\
+  }
+  \and
+  \IEEEauthorblockN{
+    Andrew Tolmach
+  }
+  \IEEEauthorblockA{
+    Portland State University\\
+    tolmach@pdx.edu\\
+  }
+}
 

### 2022-08-03 submission5/main.tex
Attempt 5

--- 2021-12-21 submission4/main.tex
@@
 
-What does ``stack safety'' mean? The phrase is associated with a
-variety of compiler,
-run-time, and hardware mechanisms for protecting stack
-memory, but these mechanisms typically lack precise specifications,
+``Stack safety'' is a term associated with a
+variety of compiler, run-time, and hardware mechanisms for protecting stack
+memory. But unlike heap memory, the stack does not correspond to any single
+high-level language concept. The result is that
+these mechanisms typically lack precise specifications,
 relying instead on informal descriptions and examples of the bad
-behaviors that they prevent.
-
-We propose a generic, formal characterization
-of stack safety based on concepts from language-based security: a
+behaviors that they prevent. Such a specification is especially desirable
+for an unusual enforcement mechanism, like the ``lazy'' stack safety
+micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
+
+We propose a formal characterization of ``observable'' stack safety based on
+concepts from language-based security. Stack safety is a
 combination of an integrity property (``the private
-state in each caller's stack frame is held invariant by the callee''),
+state in each caller's stack frame is held invariant by the callee'')
 and a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state''), plus an optional control-flow property.
-
-We use these properties to validate the stack safety micro-policies
-proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
-Specifically, we check (with
-property-based random testing) that their ``eager''
-micro-policy, which catches violations early, enforces a
-simple ``stepwise'' variant of our properties, and that
-(a repaired version of) their more performant
-``lazy'' micro-policy enforces a slightly weaker and more extensional
-observational property. Meanwhile our testing successfully detects violations
+caller's private state''), both compatible with lazy tagging, plus a control-flow property.
+
+The stack governs a large number of language features that are often excluded
+from discussion of stack safety. Our stack safety property targets a system with
+both caller- and callee-saved registers, coroutines, arguments passed on the stack,
+exceptions, [tail call elimination, heaps, and concurrency.]
+
+We use these properties to validate Roessler and DeHon's micro-policies
+via property-based random testing. Our testing successfully detects violations
 in several broken variants, including Roessler and DeHon's original lazy policy.

### 2022-08-17 submission5/onlyformalism.tex
Working on macro-instruction formalism

--- 2021-12-21 submission4/main.tex
@@
 \usepackage{xcolor,listings}
+
+\usepackage{multirow}
+
+\usepackage{stmaryrd}
 
@@
 
-\begin{abstract}
-
-What does ``stack safety'' mean? The phrase is associated with a
-variety of compiler,
-run-time, and hardware mechanisms for protecting stack
-memory, but these mechanisms typically lack precise specifications,
-relying instead on informal descriptions and examples of the bad
-behaviors that they prevent.
-
-We propose a generic, formal characterization
-of stack safety based on concepts from language-based security: a
-combination of an integrity property (``the private
-state in each caller's stack frame is held invariant by the callee''),
-and a confidentiality property (``the callee's behavior is insensitive to the
-caller's private state''), plus an optional control-flow property.
-
-We use these properties to validate the stack safety micro-policies
-proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
-Specifically, we check (with
-property-based random testing) that their ``eager''
-micro-policy, which catches violations early, enforces a
-simple ``stepwise'' variant of our properties, and that
-(a repaired version of) their more performant
-``lazy'' micro-policy enforces a slightly weaker and more extensional
-observational property. Meanwhile our testing successfully detects violations
-in several broken variants, including Roessler and DeHon's original lazy policy.
-
-\end{abstract}
-
-\newcommand{\paragraphx}[1]{\emph{#1.}}
-
-\section{Introduction}
-
-The call stack is a perennial target for low-level attacks, leading to a
-range of dire consequences, from leakage or corruption of private stack data
-to control-flow hijacking. To foil such attacks, a profusion of
-software and hardware protections have been proposed,
-%
-including stack canaries~\cite{Cowan+98},
-bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
-split stacks~\cite{Kuznetsov+14},
-shadow stacks~\cite{Dang+15,Shanbhogue+19},
-capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
-  \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\bcp{Yes, going back to MIT days---we should
-  include several more of these, if only to give readers the impression that
-this is a well-studied mechanism (so formalizing its protections is
-useful).}
-\fi
-
-The protections offered by such mechanisms are commonly described in terms
-of concrete examples of attacks that they can prevent---corruption of return
-addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more abstract characterization to the reader's intuition.  But these
-mechanisms can be quite intricate, and the behaviors they aim to prevent are
-subtle and varied. It can be hard to guess, based only on informal
-intuitions, whether a given mechanism actually blocks all potential
-attacks---or, conversely, whether it is overly conservative and disallows
-system behaviors that are actually safe.
-To settle such questions with confidence,
-we need a precise, generic, and formal specification for stack
-safety, as a basis both for comparing the security claims of different
-enforcement techniques and for validating that these claims
-are met by particular implementations. With the notable exception of the model used
-in Skorstengaard et al.'s StkTokens work~\cite{SkorstengaardSTKJFP}, which we
-discuss in Section \ref{sec:relwork}, there is a dearth of such properties in
-the literature.
-
-We propose a novel characterization of stack safety using the tools of language-based
-security~\cite{sabelfeld2003language}. The informal claim that
-``stack safety protects a caller
-from its callee'' amounts to saying that it guarantees the {\em integrity} and
-{\em confidentiality} of the caller’s local state until it regains control after the call.
-
-A key technical novelty in these definitions,
-compared to standard formulations of confidentiality and integrity from the
-security literature, is that they
-are ``nested'': {\em each} caller is guaranteed protection from its
-immediate callees (which, in turn, need protection from their immediate
-callees, etc.).
-Confidentiality is especially interesting. It is based on a traditional
-notion of noninterference, but whereas ordinary noninterference
-is an end-to-end hyper-property on whole program runs, stack
-confidentiality is a nested form of noninterference applied to
-subtraces of the whole program trace delimited by call instructions
-and corresponding returns, requiring that the callee’s behavior is
-invariant under hypothetical scrambling of the caller’s stack frame.
-
-Our threat model is very strong, allowing the attacker (i.e., the
-callee) to execute arbitrary code, including attempting to smash
-the stack to disrupt the program’s control flow. Our stack safety
-properties demand that, even in the presence of such attackers, confidentiality
-and integrity still apply to data in suspended stack frames
-until the callee returns.
-%
-We can also, optionally, enforce {\em well-bracketed control
-flow (WBCF)}~\cite{SkorstengaardSTKJFP}---a global requirement that
-each callee must always return to its immediate caller, if it returns
-at all.
-
-For confidentiality and integrity, we consider both {\em stepwise}
-and {\em observational} variants.  The simpler stepwise
-variants formalize the intuition that a caller's local data is {\em never}
-read or modified during a call. The observational variants
-allow a callee to read from and write to its caller's stack frame, as
-long as nothing that it reads affects its observable behavior and nothing
-that it writes affects the observable behavior of the caller after it returns.
-The observational properties are more extensional, and represent a baseline
-that any reasonable protection mechanism ought to enforce,
-even if it does not prevent every single dangerous read or write.
-
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly abort the ones that
-attempt to violate one of our properties. Furthermore, we
-%
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-enforcement variants---both variants that we accidentally created
-during our re-implementation of the micro-policy and ones that we
-intentionally crafted to be broken in order to increase our confidence
-in testing and the enforcement mechanism itself.
-
-We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
-which memory within each stack frame is tagged with the depth of
-the function activation that owns the frame and access is
-permitted only when an activation at that depth is currently executing, validates our
-stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-%
-\ifaftersubmission
-\bcp{Did Nick and Andre know about this deficiency?  Did they mention
-  it in their paper?  (If neither, we should make a bigger deal out of the
-  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
-    They refer to ``temporal security properties'' as something to do with dangling
-    pointers, but they don't define it.
-    Then they say that Depth Isolation has weaker temporal security
-    guarantees than Static Authorities (tag=function name), but they don't explain why
-    (and I don't see why). They don't say anything explicit about the temporal
-    security properties of Lazy Tagging and Clearing, but they consider it to enforce
-    a completely different (and generally weaker) policy anyway (``data flow
-    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
-  leave this issue to deal with later (in consultation with them, which we
-  don't have time to do a good job of today).  At any rate, by my reading we
-are not saying anything incorrect at the moment.}
-\fi
-
-For ease of exposition, we initially assume a single simple
-stack with no sharing between callers and callees: all parameters and return
-values are passed in registers.  Later, we will introduce enhanced versions
-supporting
-(1) passing of scalar stack data,
-(2) callee-saves registers calling conventions, and
-(3) a coroutine system with a static layout.
-
-We begin with an informal overview of our key ideas via
-examples (\cref{sec:running-example}), followed by a description of our
-assumptions about the machine environment and threat model (\cref{sec:setup}).
-We then offer the following contributions:
-
+\section{Threat Model, Machines, and Enforcement }
+
+Before introducing our formal model of stack safety, we describe
+our machine exemplar, enforcement mechanism, and threat model.
+
+We begin with a RISC-V-like machine enhanced with PIPE, a tag-based reference
+monitory. We extend this machine with a {\it security overlay semantics}:
+a semantics in which calls, returns, and other operations of interest are condensed
+into atomic {\it macro-instructions}, which update an additional security context.
+Our properties are defined in terms of the security context at the end of an appropriate
+macro-instruction.
+
+The underlying machine is a standard RISC-V ISA as modeled in [MIT-RISCV].
+Our overlay semantics provide macro-instructions for the following code features:
 \begin{itemize}
-\item We adapt concepts of integrity and confidentiality from language-based
-security and combine these with an existing notion of well-bracketed
-control flow to formalize a {\em stepwise stack safety}
-property that captures the intuitive concept of
-strict stack safety enforced over each step of execution~(\cref{sec:prelim,sec:lse}).
-As a case study, we explore how our properties are enforced
-by the {\em Depth Isolation} micro-policy
-of Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}, but not by some of its optimized variants~(\cref{sec:enforcement}).
-\item We formalize weaker but more extensional \emph{observational} variants of
-  confidentiality and integrity. Both are parameterized over a notion of external
-  observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in the system behavior (\cref{sec:lazy}).
-\item We show that our definitions can be
-extended to describe argument passing on the stack, callee-saves registers, exceptions, and a simple coroutine system (\cref{sec:ext}).
-\item We use property-based random testing to validate the relationship between
-  our properties and micro-policies (\cref{sec:testing})
+\item Function calls and returns, with caller- and callee-saved registers according
+  to the RISC-V ABI
+\item Arguments passed by reference
+\item Exceptions
+\item Coroutines and cooperative threading
+\item Tail-call Elimination
 \end{itemize}
-\Cref{sec:relwork,sec:future} discuss related and future work.
-
-Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as the testing implementation.
-
-\section{Key Ideas by Example}
-\label{sec:running-example}
-\colorlet{lgray}{gray!40}
-\colorlet{lred}{red!40}
-\colorlet{lblue}{blue!20}
-
-\newcommand{\mainsealc}{cyan}
-\newcommand{\fsealc}{green}
-\newcommand{\unsealc}{lgray}
-\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
-\newcommand{\fulloutc}{white}
-\newcommand{\badc}{lred}
-\newcommand{\goodc}{lblue}
-\newcommand{\retptrc}{black}
-\newcommand{\sealdesc}[1]{Seal(#1)}
-\newcommand{\unsealdesc}{Unseal}
-\newcommand{\retptrdesc}{RetPtr}
-\newcommand{\passdesc}[2]{Pass(#1,#2)}
-
-\begin{figure*}
-\begin{subfigure}[]{.16\textwidth}
-{\small
-  \begin{lstlisting}
-int main() {
-  int x = 42;
-  return f()+x;
-}
-
-int f() {
-  int y;
-  y = *(&y-2);
-  *(&y-2) = 0;
-  return 5;
-}
-  \end{lstlisting}
-}
-\end{subfigure}
-\begin{subfigure}[]{.44\textwidth}
-  {\small
-    \begin{lstlisting}[backgroundcolor=\color{lgray}]
-1 main: add $1,%sp      ;allocate frame
-2       mov $42,-1(%sp) ;initialize local
-3       call f
-4       add -1(%sp),%ra ;set return value
-5       sub $1,%sp      ;deallocate frame
-6       ret
-
-10 f:   add $1,%sp      ;allocate frame
-11      mov -4(%sp),-1(%sp) ;violation 1!
-12      mov $0,-4(%sp)      ;violation 2!
-13      mov $5,%ra      ;set return value
-14      sub $1,%sp      ;deallocate frame
-15      ret
-    \end{lstlisting}
-}
-\end{subfigure}
-\begin{subfigure}[]{.43\textwidth}
-\centering
-\addtolength{\tabcolsep}{-1em}
-\begin{tabular}{l | l}
-  \multicolumn{2}{c}{
-    2 \hspace{-1.5em}
-    \memoryaddrs{8em}
-    \memory{3}{\unsealc}
-    ~$\cdots$
-    \vspace{.5em}
-  } \\
-  11 \hspace{-1.5em}
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{42}
-  &
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  12 \hspace{-1.5em}
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{42}
-  \MemoryLabel{-6em}{0.75em}{42}
-  \vspace{.5em}
-  &
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
-  \vspace{.5em}
+
+\paragraph*{What is a ``call''?} Some ISAs have a single ``call''
+instruction that does all the required manipulation of the program counter
+(\(\PCname\)) and stack pointer; others perform a series of instructions
+whose constituent opcodes may be used in some places for other purposes.
+Our RISC-V machine is the latter, so calls must be explicitly designated
+as such by the compiler, by identifying them as macro-instructions.
+The same goes for other operations, such as exceptions and coroutine yields.
+Macro-instructions connect the enforcement mechanism to the property: we don't
+expect the enforcement mechanism to provide any guarantees for code sequences
+that resemble calls or returns, but which are not macro-instructions.
+
+Because macro-instructions must be generated by the compiler, they only
+carry data that can be known by the compiler at compile-time. For example,
+the \(\mathbf{call}\) macro-instruction identifies the register \(r\) containing
+the address of the function to be called, as well as a {\it type record} identifying
+the registers and addresses in which the callee expects to find arguments,
+and a {\it public set}.
+Let \(tr \in TR\) be a type record; then \(tr.args\) is a set of those registers
+that will be used for arguments, and \(tr.stack\_args\) defines a set of
+offsets from the stack pointer that are expected to contain additional
+arguments.The public set \(pubs\) defines a similar set of offsets in
+the caller that are declared public and not to be protected.
+
+%\paragraph*{Enforcement Model and Properties}
+
+%A machine state consists of the standard addresses and registers, as well as
+%additional policy state used by the enforcement mechanism. Our properties
+%refer only to addresses and registers; they do not constrain the behavior
+%of policy state. We model an execution step as a partial function on states:
+%when the machine does not step from a given state, this
+%is a {\em failstop}, halting the program before it can perform an
+%action that would violate the security property.
+%our properties are therefore naturally \emph{termination insensitive.}
+
+\paragraph*{Threat Model}
+
+We trust the compiler has correctly placed macro-instructions according to
+the semantics of the source language, but we do not assume that the code
+is otherwise reasonable. In particular, while we are agnostic as to the source
+language, C is very plausible, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code.
+
+In general, it is impossible to distinguish buggy source code from an attacker;
+we assume nothing about the intent of the source code, except that it makes calls
+and returns, and guarantee that the caller and callee are protected from one
+another.
+
+This is a strong threat model, but hardware and timing attacks are out of scope,
+and our properties are termination insensitive as a result of the enforcement mechanism
+(below).
+
+\paragraph*{Limitations}
+
+We do not support dynamic code generation or concurrency---we can handle a
+limited form of coroutines, but not full-blown preemptive concurrency.
+[Say something about Heap Safety]
+
+\subsection{The Machine in Detail}
+\label{sec:prelim}
+
+The building blocks of the machine are {\em words} and {\em registers}.
+Words are 64-bit integers ranged over by \(w\) and, when used as addresses,
+\(a\). There are 32 registers, ranged over in general by \(r\),
+plus a program counter (\(\PCname\)). The following table divides
+the registers into metavariables, some belonging to caller- or callee- sets:
+
+\vspace{\abovedisplayskip}
+\begin{tabular}{| l | l | l |}
+  \hline
+  Set & Names & Purpose \\
+  \hline
+  CLR\_SV & \(rt_0 - rt_6\) & Caller-saved temps \\
+  & \(ra_0 - ra_1\) & Caller-saved args / return vals \\
+  & \(ra_2 - ra_7\) & Caller-saved args \\
+  \hline
+  CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
+  \hline
+  PUBLIC & \(rra\) & Return Address \\
+  & \(rsp\) & Stack Pointer \\
+  & \(pc\) & Program Counter \\
+  \hline
 \end{tabular}
-\addtolength{\tabcolsep}{1em}
-\end{subfigure}
-\caption{Integrity and confidentiality violations with confidentiality violation traces
-\ifaftersubmission\bcp{I wonder if we
-    could make the figures easier to read by putting very light, differently
-    colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
-\vspace*{-1em}
-\label{fig:conf1}
-\end{figure*}
-
-\addtolength{\abovecaptionskip}{-1em}
+\vspace{\abovedisplayskip}
+
+Collectively addresses and registers are {\em state elements} \(\component\)
+in the set \(\COMPONENTS ::= \WORDS + \REGS\).
+%
+A {\em machine state} \(\mach \in \COMPONENTS \rightarrow \WORDS \)
+is a mapping from state elements to values. For convenience, we abbreviate
+multiple updates \(m[\component_0 \mapsto w][\component_1 \mapsto w]\dots\)
+as \(m \llbracket \component \mapsto w | P ~ \component\rrbracket\), where
+\(P\) is some predicate that defines the set of elements to update.
+
+Separate from the machine state, we keep track of additional security context
+information over the course of the run. This data is not relevant to execution,
+it is only used to define the stack safety property. We define a set of
+{\it security classes}:
+\[sc \in SEC ::= \sealed | \passed | \mathit{public} | \unsealed\]
+And of {\it targets}:
+\[\begin{split}
+tar \in TAR ::= & \mathit{Return} ~ a_1 ~ a_2 \\
+| & \mathit{Catch} ~ a_1 ~ a_2 \\
+\end{split}\]
+A {\it security view} \(V \in \COMPONENTS \rightarrow SEC\) divides the machine
+elements in terms of their purpose relative to the active function.
+And a security context is a stack of pairs of targets and views:
+\[\context \in \CONTEXTS ::= \mathit{list} ~ (TAR \times SEC)\]
+
+The set of instructions, \(I\), contains the RISC-V ISA. There is additionally a set
+of macro-instructions \(J\), described in detail in Figure \ref{fig:macros}.
+Every macro-instruction \(j \in J\) has a
+size \(\mathit{size}(j)\) and can be decomposed into a sequence of fragments
+\(j_0 \dots j_{\mathit{size}(J)}\), where \(j_n \in J^*\).
+The partial function \(\mathit{decode} : \WORDS \rightharpoonup I + J^*\) may decode a value
+into either an instruction or a macro-instruction fragment.
 
 \begin{figure}
-\begin{subfigure}{.4\columnwidth}
-  {\small
-\begin{lstlisting}
-int main() {
-  int x;
-  x = f();
-  return g();
-}
-int f() {
-  int y = 5;
-  return y;
-}
-int g() {
-  int z;
-  return z;
-}
-\end{lstlisting}
-  }
-\vspace*{-1em}
-\caption*{Fig. \arabic{figure}: Leaking via\\ uninitialized data}
-\addtocounter{figure}{1}
-\vspace*{-1em}
-\label{fig:conf2}
-\end{subfigure}
-\begin{subfigure}{.5\columnwidth}
-  \begin{center}
-\begin{tabular}{l l}
-2 &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}[\unsealdesc]
-
-~$\cdots$
-\vspace{.5em}
-\\
-3 &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}
-~$\cdots$
-    \MemoryLabel{-15em}{0.75em}{42}
-    \vspace{.5em}
-\\
-12 &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}[\sealdesc{0}]%
-\memory{1}{\retptrc}[\retptrdesc]%
-\memory{1}{\unsealc}[\unsealdesc]
-~$\cdots$
-\MemoryLabel{-16em}{0.75em}{42}
-\MemoryLabel{-7.5em}{0.75em}{42}
-\vspace{.5em}
-\\
-13 &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}
-\memory{1}{\retptrc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-15em}{0.75em}{\bf 0}
-\vspace{.5em}
+  \begin{tabular}{| l | l l |}
+    \hline
+    Macro & \multicolumn{2}{|l|}{Parameters} \\
+    \hline
+    \multirow{3}{*}{\(\mathbf{call}\)}
+    & \(\reg\) & Register containing destination \\
+    & \(tr\) & Type record \\
+    & \(pubs\) & Public set \\
+    \hline
+    \multirow{3}{*}{\(\mathbf{tailcall}\)}
+    & \(\reg\) & Register containing destination \\
+    & \(tr\) & Type record \\
+    & \(pubs\) & Public set \\
+    \hline
+    \multirow{2}{*}{\(\mathbf{enter}\)}
+    & \(tr\) & Type record \\
+    & \(n\) & Size of locals \\
+    \hline
+    \(\mathbf{return}\)
+    & \(pubs\) & Public set \\
+    \hline
+    \(\mathbf{yield}\) & & \\
+    \(\mathbf{setex}\) & & \\
+    \(\mathbf{throwex}\) & & \\
+    \hline
   \end{tabular}
-\end{center}
-\caption*{Fig. \arabic{figure}: Integrity Violation Trace}
-\label{fig:int1}
-\end{subfigure}
-\vspace*{1em}
-\vspace*{-2em}
+  \caption{Macro-instructions}
+  \label{fig:macros}
 \end{figure}
 
-This section explains our formal stack safety properties through a series
-of examples, each presented both in C syntax and in corresponding assembly code for an
-imaginary two-address machine (e.g., see Figure~\ref{fig:conf1}).
-Each instruction's address is shown to its left.
-The assembly code reflects a simple compilation strategy in which
-local variables are stored in consecutive slots of the stack frame.
-For the moment, arguments and return values are passed in registers.
-The stack grows upward in
-memory and {\tt \%sp} points to the first unused word above the top of the stack.
-The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
-implicitly adjusting the stack pointer.
-At the right of each example, we show snapshots of the layout of stack memory just before
-selected instructions (labeled by address) are executed; each box represents a memory word
-(the machine is word addressed). Addresses increase to the right.
-We refer to registers and addresses collectively as {\em state elements} or just elements.
-  %% \bcp{Most of the
-  %% figure is kind of self-explanatory with this help, but the rightmost part
-  %% is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
-  %% numbers in the boxes?  What are these brackets under the boxes, ...  One
-  %% simple thing we could do to help might be to label the right-hand column
-  %% something like ``Snapshots,'' or at least explain early on in the text
-  %% that this is what they are.}
-
-Labels under the boxes indicate an association of the address to a
-{\em domain}---a set of intended access
-rights. Unless otherwise marked, each address retains its domain as
-instructions execute. %, as hinted by their colors.
-%\bcp{where do we see this happening?  Is it important?}.
-Domains are not part of the machine state; rather, they are part
-of the {\em context} of a state that describes the security restrictions
-that the enforcement mechanism must adhere to.
-The possible domains
-are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
-(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
-(outside the stack entirely).
-% These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
-Global variables and code, which are not shown in the figures, would
-also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
-are shaded black and ignored for simplicity for now---they will
-be treated properly later.
-(We will introduce some other parts of the context as needed.)
-
-Figure~\ref{fig:conf1} shows a program with simple integrity and confidentiality violations. Function {\tt main}
-calls function {\tt f}, which reads and then overwrites {\tt main}'s
-variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
-
-First, {\tt f} violates confidentiality by reading {\tt x}.
-Confidentiality states intuitively that a callee is insensitive to
-the context of its call (except for its arguments and any global
-variables).
-We state this as a form of {\em noninterference}~\cite{Goguen1982SecurityPA}.
-Noninterference describes knowledge: its standard form states that for state elements
-containing secret values, we can substitute any values whatsoever for the secrets
-without changing the effect of our computation on non-secret values.
-%
-If there exists some value for which we see different results, then
-in fact we can learn something about the initial state by observing those results.
-To adapt this to the stack, the notion of ``secret'' must vary with function
-calls and returns.
-
-At the right of Figure~\ref{fig:conf1}, we see how noninterference applies to this scenario.
-At instruction 10, we
-create an imaginary {\em variant}
-state with arbitrary values~%
-%\ifaftersubmission\bcp{IMO, it would be easier to
-%  understand if we chose concrete values for the examples}\sna{FWIW I agree in
-%  theory, but it looks bad every time I try it.}\fi
-(\(v_0\), etc.),
-in place of the actual values ($42$, ...) that should
-be hidden from {\tt f}; we show the evolving stack memory of the variant in the second column.
-At instruction 11, the callee's variable, {\tt z},
-changes to 42 in the primary trace, but
-in the variant trace it instead might have any value \(v_0\). If \(v_0 \neq 42\), then the variant
-has behaved differently than the primary, violating confidentiality.
-
-We call this form of confidentiality {\em stepwise confidentiality}. Informally,
-for any state at the entry to a function, and for any variant over the values in the stack,
-the primary state and the variant will step in lockstep until they
-return. To step in lockstep
-means that if both states step---that is, neither the primary nor the variant failstops---any
-state element that changes in the original will do the same in the variant.
-
-It may be unclear in Figure~\ref{fig:conf1} why we vary the initial value of {\tt y}
-as well as that of {\tt x}. This is to detect some violations of so-called
-{\em temporal stack safety}.
-Consider the code in Figure~2.
-Using the same compilation strategy as before, {\tt y} and {\tt z} will end up occupying the
-same memory location in the stack,
-so the value {\tt f} leaves behind in {\tt y}, which later becomes the value of {\tt x}, also
-becomes the value of {\tt z}. Thus, {\tt g} is able to learn the value of its caller's
-variable {\tt x}.  We regard this as a violation of confidentiality as well. In terms of noninterference,
-in the primary trace, {\tt z} will have the value 5, but in the variant it can have
-an arbitrary value.
-
-Returning to the code in Figure~\ref{fig:conf1}, line 12 violates integrity.
-In the initial memory layout the entire stack is {\em unsealed},
-meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
-in this case 0. (Note that this does not include the return address, which is saved
-by the call instruction.)
-Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt f} will not change {\tt x}.
-This leads to a simple, intuitive statement of {\em stepwise integrity}:
-if an element is sealed before a step, it will be unchanged afterward.
-Overwriting {\tt x} violates this property, as shown in Figure~3.
-
-\paragraph*{Control-Flow Attacks}
-\begin{figure*}
-  \centering
-\hspace*{-6em}
-  \begin{subfigure}[t]{.2\textwidth}
-{\footnotesize
-\begin{lstlisting}
-int *stash = 0;
-int main() {
-  int x = 1;
-  f();
-  x = -x;
-  f();
-  return x;
-}
-
-void f() {
-  int y;
-  if (stash == 0) {
-    stash = *(&y-1);
-  } else {
-    *(&y-1) = stash;
-    stash = 0;
-  }
-  return;
-}
-\end{lstlisting}
-}
-\end{subfigure}
-  \begin{subfigure}[t]{.4\textwidth}
-{\footnotesize
-\begin{lstlisting}[backgroundcolor=\color{lgray}]
-   stash: .word 0
-1  main:  add $1,%sp     ; alloc frame
-2         mov $1,-1(%sp) ; init x
-3         call f
-4         neg -1(%sp)    ; x = -x
-5         call f
-6         mov -1(%sp),%ra ; set ret val
-7         sub $1,%sp     ; dealloc frame
-6         ret
-
-10 f:     add $1,%sp     ; alloc frame
-11        cmp stash, $0  ; stash = 0?
-12        brneq #15      ; branch if not
-13        mov -2(%sp),stash ; set stash
-14        jmp #17
-15        mov stash,-2(%sp) ; get stash
-16        mov $0,stash
-17        sub $1,%sp     ; dealloc frame
-18        ret
-\end{lstlisting}
-}
-  \end{subfigure}
-%
-  \begin{subfigure}[t]{.2\textwidth}
-    \begin{center}
-    \begin{tabular}{l l l}
-      3 &
-      \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      {\tt 18} &
-      \memoryaddrs{12em}
-      \memory{1}{\mainsealc}[\sealdesc{0}]%
-      \memory{1}{\unsealc}[\retptrdesc]%
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1}
-      \MemoryLabel{-11em}{0.75em}{\#4}
-      \vspace{.5em} &
-      \hspace*{-1.5em}
-      \memory[1.2em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-2.2em}{0.75em}{\#4}
-      \\
-      4 &
-      \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      11 &
-      \memoryaddrs{16em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#6} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      18 &
-      \memoryaddrs{12em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\badc}
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#4} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      4 &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      5 &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{\bf 1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
-    \end{tabular}
-    \end{center}
+Execution is modeled by a partial step function between machine states,
+written \(\mach, \context \stepsto \mach', \context'\).
+If \(\mach[\PCname]\) decodes to an instruction, then
+\(\mach, \context \stepsto \mach', \context\) if \(\mach\) steps to \(\mach'\)
+under the RISC-V semantics.
+If, for some \(j\), \(\mach[\PCname + i]\) decodes to \(j_i\) for all \(i\) between 0 and
+\(\mathit{size}(j)\), then the step is via a special multi-step. We present these steps
+as separate step relations: \(\cdot \harpoonunder{j}_M \cdot \in \MACHS \times J \times \MACHS\)
+to step the machine state,
+and \(\cdot, \cdot \stepstounder{j}_C \cdot \in \MACHS \times \CONTEXTS \times J \times \CONTEXTS\)
+for the security context. We then combine them:
+
+\judgmenttwo[]
+            {\(m \harpoonunder{j}_M m'\)}
+            {\(m',c \stepstounder{j}_C c'\)}
+            {\(m,c \stepsto m',c'\)}
+
+\paragraph*{Macro-step machine steps}
+
+Here we give the semantics of the machine side of the macro step for calls and returns.
+The machine multi-step for a call (\ref{fig:callmacro}) updates the stack pointer
+and program counter, and zeros non-argument, caller-save registers. (This means that
+the caller should have saved them before initiating the call.)
+
 \vspace{\abovedisplayskip}
-  \end{subfigure}
-  \caption{A control-flow attack.}
-  \label{fig:controlflow}
-\end{figure*}
-
-%From the classic stack smashing attack to return oriented programming,
-One of the most pernicious ways to attack the stack is by hijacking control flow.
-Our stack safety properties do not assume that control flow follows a typical structure
-of matching sets of calls and returns nested within one another.
-Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
-Consider Figure~\ref{fig:controlflow}.
-(Here we treat return addresses properly, but write them as line numbers, in the
-form $\#n$.)
-Function {\tt f} is called twice, at instructions 3 and 5; it should therefore return to instructions
-4 and 6, respectively. But during the first call it stashes its
-return pointer, and the second time it returns to instruction 4 instead of 6; it thus ends
-up being called a third time before finally returning to instruction 6.
-
-Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that {\tt main} returns
-$-1$. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
-This seems like an integrity violation; how do we capture it in our
-\ifspace integrity \fi property?
-
-The key intuition is that, to be considered as having returned to {\tt main}, {\tt f} needs
-to do more than just execute a return-labeled instruction: it must actually reach the
-appropriate return target in {\tt main}.
-To capture this intuition, we add a new piece of context:
-a stack of return targets.  Now a callee is
-only considered to have returned when it reaches a state that matches the caller's target.
-A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, represented here by coloring both the target and its sealed data with the
-same color (blue).
-
-In the example, {\tt f} reaches its return target on its first return, to instruction 4.
-But {\tt f}'s second return is to 4 rather than 6
-because the stack pointer has been overwritten (shown in light red).
-Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.
-
-\paragraph*{Well-bracketed control flow} Together,
-integrity and confidentiality protect a caller's data until its callee
-returns to it. However, it is possible that the caller is not
-returned {\em to}, but rather returned {\em past}, to some ancestor caller!
-Strictly speaking, this is not an information-flow violation. But it is still
-arguably unsafe if the ancestor depends on the skipped function completing its execution before
-it returns.
-%
-Thus, another aspect of safe use of the stack is that calls
-and returns should be correctly nested---a callee should
-only return to its own caller.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
-call \emph{well-bracketed control flow (WBCF)}, and is orthogonal to
-confidentiality and integrity, so we treat is as an optional additional property.
-
-\section{Threat Model, Machines, and Enforcement }
-\label{sec:setup}
-\label{sec:threat}
-
-Before introducing our formal model of stack safety, we describe our assumptions:
-how we abstract over machines, how we represent enforcement, and
-our threat model.
-
-\paragraph*{Machine model}
-Our properties are defined in terms of an abstract model that makes few
-assumptions about the concrete machine that instantiates it. We assume a
-register machine with a program counter (\(\PCname\)) and well-defined
-mechanisms for determining the bounds of a stack frame and recognizing when a call
-has returned. Our examples consider machines with a stack pointer (\(\rsp\)),
-in which the current stack frame is everything below the stack pointer and above the
-previous frame, and where a call from instruction \(\addr_\PCname\) with
-stack pointer \(\addr_\rsp\)
-is considered to have returned when the \(\PCname\) is at \(\addr_\PCname + 4\) and the stack pointer
-is again at \(\addr_\rsp\). However, our properties apply equally to
-other architectural conventions, such as frame pointers.
-The non-stack memory can be used in any way, including
-as a heap, as long as the stack has static bounds. There can be
-state other than registers and memory, such as the internal state of a hardware
-enforcement mechanism.
-
-Our core property assumes that all register values passed to the callee are public,
-and that all data in a stack frame is private.
-Some calling convention choices, such as callee-saves registers and argument
-passing on the stack, are not supported directly by this property;
-in Section~\ref{sec:ext},
-we introduce
-secondary properties that use the same underlying formalism to handle these cases.
-
-\paragraph*{What is a ``call''?}  One significant challenge in formulating our
-properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs.  There may be no single ``call''
-instruction that does all the
-required manipulation of the \(\PCname\) and stack pointer. Instead, the
-same instruction opcode may be
-used in some places in a call or return sequence and in other places
-for other purposes.
-\ifspace
-For example, there
-might be just a simple branch instruction that stores the old \(\PCname\) in
-a register, with the rest of the call sequence performed by other
-instructions either before or after the jump.  Similarly, instead of a single
-``return'' instruction,  there may be a ``return sequence'' involving an
-indirect branch at some point.
-\fi
-
-Since our stack safety properties are all about protecting ``callers'' from
-their ``callees,'' we need some clear-cut way to recognize when the program
-makes or returns from a call.  Accordingly, we assume we are given a machine
-language program {\em annotated} with the locations of instructions
-representing calls.  We declare that a call has occurred when the machine
-has just executed an instruction marked as a call.
-%
-These call annotations would typically come from a compiler together with
-the machine code, but we do not assume anything about their provenance.
-Annotations connect the enforcement mechanism to the property: we don't
-protect unmarked calls, because a mechanism's correct functioning might
-depend on knowing where calls are.
-
-Returns, on the other hand, need not be annotated: we can tell when a callee
-has returned control to the caller when...we see control return to the
-caller! We don't require the callee's cooperation at all.
-\ifaftersubmission\apt{Rest of this is repetitious with machine model section above.}\fi
-For example, under a standard calling convention, we can see that a
-return has occurred when the state of the machine reaches a state where
-the stack pointer has the same value as it did at the point of some earlier
-call and the $\PCname$ points to the instruction following that call.  We
-abstract away from the details of specific calling conventions by assuming
-that we are given some way of telling when a given machine state is a
-{\em return target} with respect to some set of pending calls.
-
-\paragraph*{Enforcement Model}
-
-We model the machine's behavior as a partial step function over machine states.
-Machine states may contain arbitrary data, including data used exclusively by the
-enforcement mechanism, while our properties refer only to the data that are
-relevant to normal execution: addresses and registers.
-When the machine does not step from a given state, this
-is a {\em failstop}, halting the program before it can perform an
-action that would violate the security property;
-our properties are therefore naturally \emph{termination insensitive.}
-%
-That said, our model is meant to apply generically to any enforcement mechanism,
-including those that do not failstop, such as
-code rewriting techniques that protect stack safety directly by
-restricting the set of possible initial states.
-
-Most enforcement mechanisms make assumptions about their initial conditions,
-so we assume that any given machine will restrict its conditions a ``well-formedness''
-criterion on initial states, ensuring that they are set up properly. For eample, in the case of
-Roessler and DeHon's \cite{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-formedness
-requires that all annotated instructions will be situated at the appropriate instruction
-of a {\em blessed sequence}---a series of instructions that are tagged to enforce correct
-behavior. It can then failstop if it would
-execute the blessed sequence out of order.
-
-\paragraph*{Threat Model}
-
-We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code, subject to whatever ``well-formedness'' conditions the
-enforcement mechanism demands. But hardware and timing attacks are out of scope,
-and our properties are termination insensitive as described above.
-
-\paragraph*{Limitations}
-
-Our model is flexible enough to support many realistic uses of the stack.
-We intentionally say little or nothing about other structures
-such as the heap.  Heap safety has been studied in isolation
-by Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18}, and we expect that their correctness condition and ours can be
-combined cleanly to characterize correct protection of stack + heap.
-
-We do not currently support:
-address-taken locals, as a satisfactory treatment of these seems to require
-integration with a more general model of memory safety;
-tail calls;
-dynamic code generation, which is difficult to integrate with code annotations in general; or
-concurrency---we can handle a limited form of coroutines, but
-  not full-blown preemptive concurrency.
-
-\section{Machines and Traces}
-\label{sec:prelim}
-
-% We now define our machine and enforcement model.
-
-The building blocks of the machine are {\em values}, {\em addresses}, and {\em registers}.
-Values and addresses are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively. Register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include a special-purpose register, the
-program counter {\PCname}. Collectively addresses and registers are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \WORDS + \REGS\).
-%
-A {\em machine state} \(\mach\) is drawn from a set \(\MACHS\) and defines a mapping
-from state elements to values, written
-\(\mach[.] \in \COMPONENTS \rightarrow \WORDS \).
-
-We have a partial step function between machine states, written \(\mach \stepsto \mach'\), which
-is undefined on input states that correspond to a policy fault. So, a security mechanism is
-represented by the capacity of the machine to ``fail-stop,'' as well as by whatever initial
-well-formedness conditions are required for the mechanism to function properly.
-
-\paragraph*{Contexts}
-
-Separate from the machine state, we keep track of additional context
-information representing the history of the run. This has no influence on execution, it merely
-describes information relevant to the particular property. Different information is needed
-for different versions of the model, so the definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
-an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
-\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
-\(\codemap \in \WORDS \rightharpoonup \ANNS\)
-that relates each code address to its annotation, if any.
-We pair a machine state with a context state to get a combined ``MC-state''
-\(\mpcstatename \in \MPCS = \MACHS \times \CONTEXTS\), and lift the context-step
-function into the machine step:
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}
-              {\(\mach, \context \constep \context'\)}
-              {\(\mach,\context \stepstocon
-                \mach',\context'\)}
+\judgmentbrbr[CallM]
+             {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathbf{call} ~ r ~ tr ~ pubs\)}
+             {\(m' = m[rra \mapsto m[pc] + 4]\)}
+             {\(m'' = m' \llbracket r \mapsto 0 | r \in CLR\_SV - tr.args \rrbracket\)}
+             {\(m \harpoonunder{j}_M m''[pc \mapsto m[r]]\)}
+\vspace{\belowdisplayskip}
+
+Then, on enter, the callee stores the stack pointer and allocates new stack space.
+[TODO: also have it do the saving of other callee-save, little complicated.]
+
+\vspace{\abovedisplayskip}
+\judgmentbrbr[EnterM]
+           {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathit{enter} ~ tr\)}
+           {\(m' = m[m[sp] \mapsto m[rra]]\)}
+           {\(m'' = m'[sp \mapsto m[sp]+n+1]]\)}
+           {\(m \harpoonunder{j}_M m''[pc \mapsto m[pc]+4*\mathit{size}(j)]\)}
+\vspace{\belowdisplayskip}
+
+Finally, on return, the program counter is restored from the stack, and the callee's
+frame is deallocated.
+
+\vspace{\abovedisplayskip}
+\judgmentbr[ReturnM]
+             {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathit{return} ~ n\)}
+             {\(m' = m[sp \mapsto m[sp]-n-1]\)}
+             {\(m \harpoonunder{j}_M m'[pc \mapsto m'[sp]]\)}
+\vspace{\belowdisplayskip}
+
+%For the context step, the macro-instruction additionally identifies which objects
+%(contiguous regions defined relative to the stack pointer) have been placed there
+%as: arguments that overflowed to the stack (\(sargs\)), which are existing objects
+%passed by reference (\(pbrs\)), and which are marked public (\(pubs\)). We use this
+%information to construct the callee's security view.
+%\[obj \in OBJ ::= \mathbb{Z} \times \mathbb{Z}\]
+%\[in\_obj ~ a_{sp} ~ (lo, hi) ~ a \triangleq a_{sp} - lo \leq a < a_{sp} - hi\]
+
+%\judgmenttwobrlong[Return]
+%                    {\(m' = m[sp \mapsto m[sp] - \mathit{size}(frsize)]\)}
+%                    {\(m'' = m \llbracket r \mapsto 0 | r \not \in args \land r \in caller \rrbracket\)}
+%                    {\(\mathit{decode}^* ~ m ~ pc = \mathit{return} ~ frsize\)}
+%                    {\(m \stepsto_M m'[pc \mapsto f]\)}
+
+%If \(\mathit{decode}\) is undefined on that value, \(\mach \stepsto \mach\). The
+%only times \(\stepsto\) is undefined is when there is a policy fault,
+%(and the machine ``failstops'') or when it attempts to begin executing a
+%macro-instruction out of order. The latter represents a failstop in a lower-level
+%policy in the underlying ISA, which guarantees that the macro-instruction
+%executes in order.
+
+\paragraph*{Compiling Macro-instructions}
+
+[TODO]
 
@@
 A {\em machine trace} is a nonempty, finite or infinite sequence
-of MC-states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::'' for list-cons.
+of states, ranged over by \(\MPCT\) and \(\NPCT\).
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
+for list-cons.
 
@@
 
-We are finally ready to give our definition of stack safety. We begin
-by describing how we model knowledge about the program structure,
-then dive into the formal definitions of integrity, confidentiality, and
-well-bracketed control flow, which together constitute stack safety.
-
-We will focus on a simple stack model, with a single stack
-that doesn't share between caller and callee: arguments and results are passed
-in registers. The code map just marks the locations of calls:
-$\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap$.
-
-To accommodate different machine architectures and different calling
-conventions, we parameterize over the structure of the stack, particularly
-the means by
-which a caller identifies the data that should be protected during a call. We term
-this the {\em sealing convention},
-%\(\sealcon ~ (\mach \in \MACHS) ~ (\component \in \COMPONENTS)\),
-\(\sealcon \subseteq (\MACHS \times \COMPONENTS)\),
-a relation between machine states and the addresses that should be sealed.
-If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
-which addresses must be protected until the return. Similarly, we abstract over what
-it means to return using a {\em return convention},
-\(\retcon : \MACHS \rightarrow 2^{\MACHS}\), which maps each call state 
-to the set of states that can be considered valid returns to that call.
-The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
-and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).
-
-Each state element is assigned to a {\em domain} \(\stackDom \in \STKDOMS\), which may be \(\outside\)
-of the stack entirely, \(\unsealed\) and therefore accessible, or \(\sealed{\depth}\)
-at a particular depth \(\depth\). Each caller registers a {\em target}
-condition, \(\target \subseteq \MACHS\), which is a predicate on machine states that means it
-has been returned to.
-A context is a pair of a ``domain map'' from elements to domains, and a
-``return target list,'' which is a stack of targets, one for each caller awaiting
-return.
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
-  \times \listT (\target) \]
-%
-The initial context \(\context_0 = (\domMap_0, [])\), where \(\domMap_0\) maps
-the stack to \(\unsealed\) and all other elements to \(\outside\).
-
-The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
-A call annotation tells us to push the return target onto the return stack and
-seal all elements with the depth of the previous return stack.
-(Recall that the code map \(\codemap\) is a partial function, and we can use its domain
-as a predicate on addresses.) If we aren't making a call,
-we identify a return by looking ahead and checking if we're about to step to a state
-that matches the topmost return target ({\sc ReturnFound}), and unsealing the caller's frame if so.
-If a target is reached anywhere in the return stack, {\sc ReturnRec}
-unseals each frame down to the level that was returned to.
-If none of these rules apply, the context remains unchanged.
-
-\begin{tabular}{l}
-  {\sc Call} \\
-  \judgmenttwobrlong[]
-                    {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
-                    %             {\(\components = \sealcon ~ \mach\)}
-                    {\(\components = \{\component | \sealcon ~ \mach[\component]\}\)}
-                    {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                    {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)} \\
-\end{tabular}
-
-\begin{tabular}{l}
-  {\sc ReturnFound} \\
-  \judgmenttwo[]
-              {\(\mach \stepsto \mach' \in \target\)}
-              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap[\components \mapsto \unsealed],\rts\)} \\
-              {\sc ReturnRec} \\
-\end{tabular}
-
-\begin{tabular}{l}
-  \judgmentbr[]
-             {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-             {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
-             {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'[\components \mapsto \unsealed],\rts'\)} \\
-\end{tabular}
-
-\noindent
-When we build machine traces with this step function, we always know how deep in the
-call stack we are and which elements are sealed.
+Now we can give our definition of stack safety. Each macro-instruction defines a
+context step; in turn, the context at the point of a call or return defines a trace
+property each for integrity, confidentiality, and well-bracketed control flow.
+The context updates involved in calls and returns are given in \ref{fig:callcontexts}.
+First, we define a function, {\it update view}, that defines the view of a callee.
+Addresses that, offset from stack pointer, are in public regions are marked
+\(\mathit{public}\). Those that are being passed on the stack are marked
+\(\mathit{passed}\). [When we add pass-by-reference arguments, we wil use
+\(V\)]
+
+To update the context, we pair the updated view with the return target of the
+call: the
+
+\begin{figure*}
+\[\begin{split}
+& \mathit{update\_view} ~ V ~ a_{sp} ~ tr ~ pubs \triangleq
+ \lambda \component .
+\begin{cases}
+  \mathit{public} & \textnormal{if } \component \in \WORDS \textnormal{ and }
+  (a_{sp} - \component) \in pubs \\
+%  \passed & \textnormal{else if } \component \in \WORDS \textnormal{ and }
+%  in\_obj ~ a_{sp} ~ pbr ~ \component \textnormal{ for some } pbr \textnormal{ in } pbrs \\
+  \passed & \textnormal{else if } \component \in \WORDS \textnormal{ and }
+  (a_{sp} - \component) \in tr.sargs \\
+  \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } \component < a_{sp} \\
+  \mathit{public} & \textnormal{else if } \component \in \REGS \textnormal{ and }
+  \component \in tr.args \textnormal{ or } \component \in \mathit{PUBLIC} \\
+  \unsealed & \textnormal{otherwise} \\
+\end{cases} \\
+\end{split}\]
+
+\judgmenttwobrlong[CallCCons]
+           {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ tr ~ pubs\)}
+           {\(j = \mathbf{call} ~ r ~ tr\)}
+           {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V')::(tar,V)::\context\)}
+           {\(m, (tar,V)::\context \stepstounder{j}_C \context'\)}
+
+\judgmenttwobrlong[CallCNil]
+           {\(V' = \mathit{update\_view} ~ (\lambda \component . \unsealed) ~ m[sp] ~ tr ~ pubs\)}
+           {\(j = \mathbf{call} ~ r ~ tr ~ pubs\)}
+           {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V')::[]\)}
+           {\(m, [] \stepstounder{j}_C \context'\)}
+
+\judgment[]
+         {\(j = \mathbf{enter} ~ tr ~ n\)}
+         {\(m,c \stepstounder{j}_M c\)}
+
+\judgmentbr[ReturnCCons]
+           {\(\)}
+           {\(j = \mathbf{return} ~ pubs\)}
+           {\(m, V::tars \stepstounder{j}_C ((m[pc] + 4, m[sp]), V')::tars\)}
+
+\judgmentbr[ReturnCNil]
+           {\(V' = \lambda \component . \textnormal{if } \component = \PCname \textnormal{ then }
+             \unsealed)\)}
+           {\(\mathit{decode}^* ~ m ~ pc = \mathit{call} ~ f ~ args ~ sargs ~ pubs\)}
+           {\(m, [] \stepstounder{j}_C ((m[pc] + 4, m[sp]),V')::[]\)}
+
+\caption{Call, enter, and return macro-instructions}
+\label{fig:callcontexts}
+\end{figure*}
 
@@
 
-\section{Extensions}
-\label{sec:ext}
-
-We have seen how our properties apply to an extremely simple
-subroutine model.
-What about more realistic uses of the stack? We now demonstrate the
-flexibility of our core concepts by extending
-them to a system that models the passing of
-parameters on the stack and to a simple coroutine model.
-
-\newcommand{\mainpassc}{magenta}
-
-\begin{figure*}
-  \centering
-  \begin{subfigure}[t]{.2\textwidth}
-{\small
-\begin{lstlisting}
-int *stash = 0;
-int main() {
-  int x, y;
-  f(x);
-  x = 0;
-  f(y);
-  return x;
-}
-
-void f(int &a) {
-  if (stash == 0)
-    stash = &a;
-  *stash = 5;
-  return;
-}
-\end{lstlisting}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.5\textwidth}
-{\small
-  \begin{lstlisting}[backgroundcolor=\color{lgray}]
-   stash: .word 0
-1  main:  add $2,%sp      ; allocate frame
-2         mov %sp,%ra
-3         add -2,%ra      ; ra = &x
-4         call f          ; SHARE: %sp-2
-5         mov $0,-2(%sp)  ; x = 0
-6         mov %sp,%ra
-7         add -1,%ra      ; ra = &y
-8         call f          ; SHARE: %sp-1
-9         mov -2(%sp),%ra ; set return
-10        sub $2,%sp      ; deallocate frame
-11        ret
-
-20 f:     cmp stash, $0   ; stash = 0?
-21        brneq #23       ; branch if not
-22        mov %ra,stash   ; stash = &a
-23        jmp #25
-24        mov $5, (stash) ; *stash = 5
-25        ret
-}
-\end{lstlisting}
-}
-  \end{subfigure}
-  \begin{subfigure}[t]{.25\textwidth}
-    \begin{center}
-      \vspace{1em}
-    \begin{tabular}{l l}
-      20 &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
-      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$ \\
-      25 &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainpassc}
-      \memory{1}{\mainsealc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-14em}{0.75em}{5}
-      \\
-      8 &
-      \memoryaddrs[(0)]{12em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-14em}{0.75em}{0}
-      \\
-      20 &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
-      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-14em}{0.75em}{0}
-      \\
-      25 &
-      \memoryaddrs[(1)]{17em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\mainpassc}
-      \memory{1}{\retptrc}
-%      \memory{1}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-14em}{0.75em}{\bf 5}
-\end{tabular}
-\end{center}
-\end{subfigure}
-\caption{A violation with pass-by-reference}
-\label{fig:passing}
-\end{figure*}
-
 \medskip

### 2022-08-21 submission5/onlyformalism.tex
Working on pseudoinstruction formalism

@@
 into atomic {\it macro-instructions}, which update an additional security context.
-Our properties are defined in terms of the security context at the end of an appropriate
-macro-instruction.
+As opposed to a fully-abstract overlay semantics, in which the semantics is safe
+by construction and compilation to an enforcement mechanism must be proved fully abstract,
+our machine's semantics are not safe unless the tag policy is correctly
+instantiated, but the security context informs trace properties that
+must hold under a policy for that policy to be considered valid.
 
@@
 %
-A {\em machine state} \(\mach \in \COMPONENTS \rightarrow \WORDS \)
+A {\em machine state} \(\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\)
 is a mapping from state elements to values. For convenience, we abbreviate
-multiple updates \(m[\component_0 \mapsto w][\component_1 \mapsto w]\dots\)
-as \(m \llbracket \component \mapsto w | P ~ \component\rrbracket\), where
-\(P\) is some predicate that defines the set of elements to update.
+multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
+as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
+are fomulae with free variables bound to sets in \(C\).
+
+The machine is equipped with additional state representing the tag-based hardware.
+We represent this as a policy state, \(\pol \in \POLS ::= \COMPONENTS \rightarrow \mathcal{T}\),
+where \(\mathcal{T}\) is the set of tags, discussed [in the section on security.]
+We represent a {\it policy} via the partial function
+\(\cdot, \cdot \rightharpoonup_P \cdot \in \MACHS \rightarrow \POLS \rightharpoonup \POLS\),
+where the function being undefined represents a {\it failstop}.
 
@@
 
-Execution is modeled by a partial step function between machine states,
-written \(\mach, \context \stepsto \mach', \context'\).
-If \(\mach[\PCname]\) decodes to an instruction, then
-\(\mach, \context \stepsto \mach', \context\) if \(\mach\) steps to \(\mach'\)
-under the RISC-V semantics.
+Each of these types of state is combined in to a triple called an
+MPC-state, \(\mpcstatename \in \MPCS ::= \MACHS \times \POLS \times \CONTEXTS\).
+Execution is modeled by a partial step function between MPC-states,
+written \(\mach, \pol, \context \harpoonunder{} \mach', \pol', \context'\),
+optionally annotated with a macro-instruction \(\harpoonunder{j}\).
+If \(\mach[\PCname]\) decodes to an instruction, then execution follows
+the RISC-V semantics and the policy, without touching the security context.
+
+\judgmentthree[]
+              {\(\mathit{decode} ~ \mach ~ \mach[\PCname] \in I\)}
+              {\(\mach \stepstounderfull \mach'\)}
+              {\(\mach,\pol \rightharpoonup_P \pol'\)}
+              {\(\mach,\pol,\context \harpoonunder{} \mach',\pol',\context\)}
+
 If, for some \(j\), \(\mach[\PCname + i]\) decodes to \(j_i\) for all \(i\) between 0 and
-\(\mathit{size}(j)\), then the step is via a special multi-step. We present these steps
-as separate step relations: \(\cdot \harpoonunder{j}_M \cdot \in \MACHS \times J \times \MACHS\)
-to step the machine state,
-and \(\cdot, \cdot \stepstounder{j}_C \cdot \in \MACHS \times \CONTEXTS \times J \times \CONTEXTS\)
+\(\mathit{size}(j)\), written \(\mathit{decode*} = j\), then the step is via a special
+multi-step. We present these steps as separate step relations:
+\(\cdot \harpoonunder{\cdot}_M \cdot \in \MACHS \times J \times \MACHS\) to step
+the machine state,
+\(\cdot, \cdot \harpoonunder{\cdot}_P \cdot \in \MACHS \times \POLS \times J \times \POLS\)
+for the policy state, and
+ \(\cdot, \cdot \stepstounder{j}_C \cdot \in \MACHS \times \CONTEXTS \times J \times \CONTEXTS\)
 for the security context. We then combine them:
 
-\judgmenttwo[]
-            {\(m \harpoonunder{j}_M m'\)}
-            {\(m',c \stepstounder{j}_C c'\)}
-            {\(m,c \stepsto m',c'\)}
+\judgmenttwobr[]
+               {\(\mathit{decode*} = j\)}
+               {\(\mach \harpoonunder{j}_M \mach'\)}
+               {\(\mach,\pol \rightharpoonup_P \pol'\)}
+               {\(\mach',\context \stepstounder{j}_C \context'\)}
+               {\(\mach,\pol,\context \harpoonunder{j} \mach',\pol',\context'\)}
+
+In this section we will define \(\harpoonunder{j}_M\) for our machine; \(\harpoonunder{j}_P\)
+for the relevant policies will be discussed in [enforcement section], and \(\stepstounder{j}_C\)
+in Section \ref{sec:properties}.
 
@@
 
-Then, on enter, the callee stores the stack pointer and allocates new stack space.
-[TODO: also have it do the saving of other callee-save, little complicated.]
+Then, on enter, the callee stores the stack pointer and callee-save registers,
+and allocates new stack space.
 
 \vspace{\abovedisplayskip}
-\judgmentbrbr[EnterM]
+\judgmentbrbrbr[EnterM]
                {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathit{enter} ~ tr\)}
                {\(m' = m[m[sp] \mapsto m[rra]]\)}
-           {\(m'' = m'[sp \mapsto m[sp]+n+1]]\)}
-           {\(m \harpoonunder{j}_M m''[pc \mapsto m[pc]+4*\mathit{size}(j)]\)}
+               {\(m'' = m\llbracket m[m[sp]+i+1 \mapsto m[rs_i] | 0 \leq i < 8 \rrbracket\)}
+               {\(m''' = m''[sp \mapsto m[sp]+n+9]]\)}
+               {\(m \harpoonunder{j}_M m'''[pc \mapsto m[pc]+4*\mathit{size}(j)]\)}
 \vspace{\belowdisplayskip}
 
-Finally, on return, the program counter is restored from the stack, and the callee's
-frame is deallocated.
+Finally, on return, the program counter and register contents are restored from the stack,
+and the callee's frame is deallocated.
 
 \vspace{\abovedisplayskip}
-\judgmentbr[ReturnM]
+\judgmentbrbr[ReturnM]
              {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathit{return} ~ n\)}
-             {\(m' = m[sp \mapsto m[sp]-n-1]\)}
-             {\(m \harpoonunder{j}_M m'[pc \mapsto m'[sp]]\)}
+             {\(m' = m[sp \mapsto m[sp]-n-9]\)}
+             {\(m'' = m\llbracket m[rs_i] \mapsto m[m[sp]+i+1 | 0 \leq i < 8 \rrbracket\)}
+             {\(m \harpoonunder{j}_M m''[pc \mapsto m''[sp]]\)}
 \vspace{\belowdisplayskip}
@@
 
-[TODO]
-
-\paragraph*{Machine Traces}
+Every macro-instruction is associated with a sequence of normal RISC-V instructions
+of the appropriate length, so compiling from the macro-instruction machine to normal
+RISC-V is as simple as replacing \(j_i\) with its corresponding instruction, tagged
+as a ``blessed sequence'' -- a standard tag-based enforcement mechanism that guarantees
+that instructions may {\it only} step in the appropriate order.
+
+\paragraph*{Machine and Observation Traces}
 \label{sec:traces}
@@
 We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
-for list-cons.
-
-The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
+for list-cons. An {\em event trace} \(\obsT\) is a similar sequence of
+events \(\obs \in \OBSS\), which are defined as follows:
+
+\[\begin{split}
+\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{dst} ~ \overline{w_{args}} \\
+| & ~ \tau \\
+\end{split}\]
+
+%This system of observations resembles that used by CompCert [cite], in that
+%the observable behaviors of the system are calls and returns (in our case, not
+%distinguishing between internal and external, and including exceptions), as well
+%as a special output akin to storing to volatile memory. We add the silent observation
+%\(\tau\).
+
+We convert macro-instructions into events via the \(\mathit{observe}\)
+function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.
+
+\begin{figure*}
+\[\mathit{observe} ~ j ~ m =
+\begin{cases}
+  \mathbf{callE} ~ m[r] ~ m \llbracket k | k \in tr.args \lor m[rsp] - k \in tr.sargs \rrbracket
+  & \textnormal{when } ~ j = \mathbf{call} ~ r ~ tr ~ pubs \\
+  \tau & \textnormal{otherwise} \\
+\end{cases}\]
+\label{fig:observe}
+\end{figure*}
+
+The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT \),
 coinductively relates an initial state with the trace of states
-produced by repeated application of \(\stepstocon\):
-%
-\vspace{-\bigskipamount}
-\begin{center}
-\begin{minipage}[t]{.6\columnwidth}
-\judgmenttwo{\(\mpcstatename \stepstocon \mpcstatename'\)}
-            {\(\mpcstatename' \hookrightarrow \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \notfinished{\mpcstatename'}{\MPCT}\)}%
-\end{minipage}
-\begin{minipage}[t]{.35\columnwidth}
+produced by repeated application of \(\rightharpoonup\):
+
+\judgmenttwo{\(\mpcstatename \rightharpoonup \mpcstatename'\)}
+            {\(\mpcstatename' \hookrightarrow \MPCT \)}
+            {\(\mpcstatename \hookrightarrow \notfinished{\mpcstatename'}{\MPCT} \)}%
 \judgment%[Default]
@@
          {\(\mpcstatename \hookrightarrow \mpcstatename\)}
-\end{minipage}
-\end{center}
-\vspace{-\smallskipamount}
 
@@
 
-\paragraph*{Until}
-The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
-and a predicate on states \(f \subseteq \MPCS\) and gives the prefix of
-\(\MPCT\) ending with the first trace entry on which \(f\) holds.
-
-\vspace{-\medskipamount}
-\begin{center}
-  \begin{minipage}[b]{.5\columnwidth}
-    \judgment{\(f ~ \mpcstatename\)}
-             {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \mpcstatename\)}
+\paragraph*{Observations-Of}
+
+This operator, written \(\obsof\), takes a machine trace and
+gives the trace of events of each step in
+the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
+initial event, a machine-trace, and an event trace; the helper is
+then invoked with an initial \(\tau\).
+
+\judgmenttwobrlong{\(\mpcstatename \harpoonunder{j} \mpcstatename'\)}
+                  {\(\mpcstatename' \hookrightarrow \MPCT\)}
+                  {\(\mathit{obsfor} ~ (\mathit{observe} ~ \pi_m(\mpcstatename) ~ j) ~ \MPCT = \obsT\)}
+                  {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}
+
+\begin{minipage}[b]{.4\columnwidth}
+  \judgment{}
+           {\(\mathit{obsfor} ~ \obs ~ \mpcstatename = \obs\)}
 \end{minipage}
-%
-  \begin{minipage}[b]{.3\columnwidth}
-    \judgment{}
-             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
-%
+\begin{minipage}[b]{.4\columnwidth}
+\judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
+         {\(\obsof(\MPCT) = \obsT\)}
 \end{minipage}
-\end{center}
-    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
-                {\(\PUT ~ f ~ (\notfinished{\mpcstatename}{\MPCT}) = \notfinished{\mpcstatename}{\MPCT'}\)}
-
-If the resulting trace is a strict prefix of the parameter, then the
-predicate must hold on its final entry. If the predicate never holds, then
-\(\PUT\) is the identity function.
-
-We will frequently take a prefix of the trace from an initial state up
-through the first state where some condition holds on the machine state.
-This can easily be implemented using \(\PUT\), and we provide special notation
-for convenience.
-%
-If \(f\) is a predicate on machine states, we define \(\mpcstatename
-\hookrightarrow \MPCT | f\) (pronounced ``\(\MPCT\) is the prefix of
-running \(\mpcstatename\) up to \(f\)'') thus:
-%
-\begin{center}
-\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
-            {\(\PUT ~ f ~ \MPCT' = \MPCT\)}
-            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-\end{center}
-
-\paragraph*{Context Segments}
-
-In particular, we wish to formulate properties around consecutive subtraces
-in which the context obeys a given predicate. This will commonly be used to extract
-individual calls from a trace by taking subtraces in which the stack is
-at or above a given depth. If \(f\) is again a predicate on states, then
-we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\) (pronounced
-``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\)'') thus:
-\ifspace{\bcp{Throughout, we could save space and improve beauty by removing some of the
-vertical space between inference rules.}\fi
-
-\begin{center}
-  \judgmenttwo[]
-              {\(f ~ \mpcstatename\)}
-              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
-  \judgmentthree[]
-                {\(f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-  \judgmentthree[]
-                {\(\neg f ~ \mpcstatename\)}
-                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
-                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
-                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
-\end{center}
+
+\paragraph*{Observational Similarity}
+
+We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar}
+from the perspective of an external observer, written \(\obsT_1 \eqsim
+\obsT_2\), if the sequence of non-silent events is the same. That is, we
+compare up to deletion of \(\tau\) events, coinductively:
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{}{\(\obsT \eqsim \obsT\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
+\end{minipage}
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
+\end{minipage}
+
+Note that an infinite silent trace is a
+prefix of (and similar to) any other trace. While this might seem
+surprising at first, it makes sense in a timing-insensitive context:
+an external observer looking at two machine runs cannot (computably)
+distinguish between a machine that steps forever and a machine that
+steps for a long time before producing some output.
 
@@
 properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if, for some initial state \(\mpcstatename = \mach,\context_0\)
+is {\em reachable} if, for some initial state \(\mpcstatename = \mach,\pol,[]\)
 where \(\mach\) is well-formed, \(\mpcstatename \hookrightarrow \MPCT\)
 and \(\MPCT\) contains \(\mpcstatename'\).
-A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
-such that from some similarly well-formed initial state \(\mpcstatename\),
-\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
 
 \section{Stack Safety, Formally}
-\label{sec:lse}
+\label{sec:properties}
 
@@
 property each for integrity, confidentiality, and well-bracketed control flow.
-The context updates involved in calls and returns are given in \ref{fig:callcontexts}.
+The context updates involved in calls and returns are given in Figure \ref{fig:callcontexts}.
 First, we define a function, {\it update view}, that defines the view of a callee.
@@
 \(\mathit{public}\). Those that are being passed on the stack are marked
-\(\mathit{passed}\). [When we add pass-by-reference arguments, we wil use
-\(V\)]
+\(\mathit{passed}\). [When I add pass-by-reference arguments, we will use
+the recursive \(V\)]
 
 To update the context, we pair the updated view with the return target of the
-call: the
+call: the return target tells us whether a return is to the proper location
+which the stack pointer restored, and the view tells us how to treat the
+security of the callee's state.
 
@@
 
-\judgment[]
+\judgment[EnterC]
          {\(j = \mathbf{enter} ~ tr ~ n\)}
@@
 
-\judgmentbr[ReturnCCons]
-           {\(\)}
+\judgmentthree[ReturnCValid]
+              {\(m[pc] = a_{pc}\)}
+              {\(m[rsp] = a_{sp}\)}
               {\(j = \mathbf{return} ~ pubs\)}
-           {\(m, V::tars \stepstounder{j}_C ((m[pc] + 4, m[sp]), V')::tars\)}
-
-\judgmentbr[ReturnCNil]
-           {\(V' = \lambda \component . \textnormal{if } \component = \PCname \textnormal{ then }
-             \unsealed)\)}
-           {\(\mathit{decode}^* ~ m ~ pc = \mathit{call} ~ f ~ args ~ sargs ~ pubs\)}
-           {\(m, [] \stepstounder{j}_C ((m[pc] + 4, m[sp]),V')::[]\)}
+              {\(m, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \stepstounder{j}_C \context\)}
+
+\judgmenttwobrlong[ReturnCInvalid]
+                  {\(m[pc] \not = a_{pc} \lor m[rsp] \not = a_{sp}\)}
+                  {\(j = \mathbf{return} ~ pubs\)}
+                  {\(V' = V \llbracket r \mapsto \unsealed | r \in \{pc, rsp\} \rrbracket\)}
+                  {\(m, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \stepstounder{j}_C (\mathit{return} ~ a_{pc} ~ a_{sp}, V')::\context\)}
+
+\judgmenttwo[ReturnCNil]
+            {\(V' = \lambda \component . \unsealed\)}
+            {\(j = \mathbf{return} ~ pubs\)}
+            {\(m, [] \stepstounder{j}_C []\)}
 
@@
 
-We now state our properties.
-The first is {\em stepwise stack integrity}, where ``stepwise'' means that the property is
-quantified over each step of execution from any initial state. The criterion for integrity
-is simple: sealed elements do not change.
-
-\definition
-The \emph{protected set} of a context, \(\prot(\domMap,\_)\), is the
-set of elements \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
-for some depth \(\depth\).
-
-\definition
-A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
-\(\mach,\context\) such that \(\mach,\context \stepstocon \mach',\context'\),
-and any element \(\component \in \prot(\context)\),
-\(\mach'[\component] = \mach[\component]\).
-
-Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
-nested: we will quantify over subtraces that represent whole calls. For each
-subtrace, we quantify over {\em variants} of its initial state. Then we require
-that the original, {\em primary} traces move in {\em lockstep} with the variant.
-
-\definition The \emph{call set} \(\MPCT^*\) of a machine is
-\[\begin{aligned}
-\bigcup_{\depth \in \mathbb{Z}} \{(\MPCT, \neg f_\depth) | \MPCT\langle f_\depth \rangle
-\text{ is a reachable segment}\} \hspace{1em} \\
-\text{where } f_\depth ~ (\mach,(\domMap,\rts)) \triangleq |\rts| \geq \depth \\
-\end{aligned}\]
-
-\noindent
-This is the set of all possible calls that can be reached from any well-formed initial state.
-
-We term \(\neg f_\depth\) the {\em break condition} \(\brk\), a predicate on
-states that holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.
+\paragraph*{Eventual Integrity}
+
+The intuition behind eventual integrity is that, when control is passed from the
+caller to the callee, the caller doesn't necessarily need its data to remain intact --
+as long as they are restored by the time control returns to it. We define this as a more
+general notion of {\it eventual integrity.}
+
+Let \(R \subseteq \CONTEXTS\) be a predicate on contexts, representing the condition at
+which integrity must be restored, and \(I\) be a predicate on states defining what integrity
+means. Then we find the first state on which \(R\) holds on the context and assert that
+\(I\) also holds on the machine state, with the \(\mathit{when}\) relation.
+
+\judgmenttwo{\(R ~ \context\)}
+            {\(I ~ \mach\)}
+            {\(\mathit{when} ~ R ~ I ~ (\mach,\pol,\context)\)}
+
+\judgmentthree{\(\neg R ~ \context\)}
+              {\((\mach,\pol,\context) \rightharpoonup \mpcstatename\)}
+              {\(\mathit{when} ~ R ~ I ~ \mpcstatename\)}
+              {\(\mathit{when} ~ R ~ I ~ \mpcstatename\)}
+
+\definition Let \(\components\) be a set of elements and \((\mach,\pol,\context)\)
+be the first state of a new call. We define \(R\) to hold on those states
+\((\mach',\pol',\context')\) where \(|\context'| < |\context|\). And we define
+\(I\) to hold when \(\mach[\component] = \mach'[\component]\) for all elements
+\(\component \in \COMPONENTS\). Then \((\mach,\pol,\context)\) enjoys eventual
+integrity of \(\components\) if \(\mathit{when} ~ R ~ I ~ (\mach,\pol,\context)\).
+
+Howver, eventual integrity is not sufficient to express the integrity protected by
+Lazy Tagging, because it is possible that values in memory are overwritten -- but
+are also guaranteed never to be accessible. We need to define a slightly weaker property.
+
+\paragraph*{Variants and Transient Elements}
+
+One or more elements of a given state are {\it transient} if their values
+in that state have no observable influence on future execution. We define this
+formally as a sort of non-interference: for a state \(\mpcstatename\) and
+a set of elements \(\components\), \(\mathit{transient} ~ \mpcstatename ~ \components\)
+holds if we can replace the contents of \(\components\) with arbitrary values
+and have the resulting state produce an equivalent trace. The state created
+by modifying \(\components\) in this way is termed a variant.
 
@@
 
-\definition The \emph{difference} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
-is the set of elements \(\component\)
+\definition An elements set \(\components\) is transient in state \((\mach,\pol,\context)\),
+written \(\mathit{transient} ~ (\mach,\pol,\context) ~ \components\), if for all
+\(\nach\) such that \(\mach \approx_{\components'} \nach\), if 
+\((\mach,\pol,\context) \hookrightarrow \MPCT\) and
+\((\nach,\pol,\context) \hookrightarrow \NPCT\), then
+\(\obsof(\MPCT) \eqsim \obsof(\NPCT)\).
+
+\definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
 such that \(\mach[\component] \not = \mach'[\component]\).
 
-\definition Two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\)
- {\em change together}, written \((\mach,\mach') \diamond (\nach,\nach')\), if, for all elements \(\component\) in
-\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
-\(\mach'[\component] = \nach'[\component]\).
-
-\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep},
-written \(\machT \doteq \nachT\),
-if their adjacent pairs of states change together.  We define this as a coinductive relation:
-
-\smallskip
-\begin{minipage}[b]{.2\columnwidth}
-\judgment{}
-         {\(\mpcstatename \doteq \npcstatename\)}
-\end{minipage}
-\begin{minipage}[b]{.75\columnwidth}
-\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
-            {\(\machT \doteq \nachT\)}
-            {\(\notfinished{(\mach,\underscore)}{\machT} \doteq \notfinished{(\nach,\underscore)}{\nachT}\)}
-\end{minipage}
-
-\definition
-The \emph{secret set} of a context, written \(\secret(\domMap,\rts)\), is the
-set of elements \(\component\) such that \(\domMap ~ \component \not = \outside\).
-
-\definition
-A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
-for any \((\MPCT,\brk) \in \MPCT^*\) with \(\head(\MPCT) = (\mach,\context\)),
-any \(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\), and any trace
-%\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
-\(\NPCT\) such that \((\nach,\context) \hookrightarrow \NPCT | \brk\),
-\(\MPCT \doteq \NPCT\) holds.
-
-In short, if we consider a variant of the stack at the start of a call,
-the resulting trace remains in lockstep with the original trace
-until either both return together or one failstops.
-Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
-of the elements in \(\components\), and in addition their subtraces are also in
-\(\MPCT^*\), and therefore confidentiality must hold on them separately.
-
-Finally, WBCF is a \ifspace simple \fi property of the behavior of returns.
-\definition
-A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
-\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
-\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
-if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).
-
-\medskip
-\section{Case Study: Micro-Policies}
-\label{sec:enforcement}
-
-We next explore how an \ifspace existing \fi enforcement mechanism called {\em
-  Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements
-the formal stack safety property described in the previous
-section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not.
-We validate these claims using property-based random testing in \cref{sec:testing}.
-
-\paragraph*{Conservative Stack Safety Enforcement Mechanism}
-%
-The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
-on a programmable, tag-based reference monitor that runs alongside the
-program,
-allowing fine-grained manipulation of metadata tags to encode so-called
-\emph{micro-policies}~\cite{pump_oakland2015}.
-Every value in
-memory and registers (including the $\PCname$)
-is enriched with an abstract metadata tag, which can represent
-arbitrary information about the value. A micro-policy is then defined as a
-set of tags and a
-collection of software-defined rules.
-On each machine step, the relevant rule is applied to the tags on
-the instruction's inputs ($\PCname$, registers, memory) and on the
-\ifspace instruction \fi opcode\ifspace itself\fi, producing
-one of two outcomes: either the instruction is permitted to execute (and
-generates tags for the result of the operation and the new $\PCname$), or
-the machine fail-stops with a policy violation.
-Existing work \cite{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
-%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
-has shown that a wide range of
-micro-policies can be defined using this scheme.
-Efficient execution of these micro-policies relies on hardware acceleration,
-such as the PUMP architecture~\cite{pump:asplos2015}.
-
-Tags in the PIPE system are not accessible to application code, being separated
-from their associated values at the hardware level. So they are not modeled
-as state elements, and therefore our properties make no claims about
-their values.
-
-The micro-policy we describe here is a variation
-on the {\em Depth Isolation} policy presented by
-Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}.
-The Depth Isolation micro-policy maintains tags of the form $\tagStackDepth{n}$ for
-each stack location belonging to the stack frame at activation depth
-$\depth$, or the tag $\tagNoDepth$ for stack locations that are currently
-unused (but could be used in the future).
-%
-The $\PCname$ is tagged similarly as $\tagPCDepth{\depth}$.
-%
-Initially, the entire stack is considered unused (and therefore tagged
-$\tagNoDepth$), and the {\PCname} has tag $\tagPCDepth{0}$.
-%
-During normal execution, the micro-policy rules only permit load and
-store operations when the target memory is tagged {\em with the same
- depth} as the current {\PCname} tag, and store operations also when the
-target memory is $\tagNoDepth$.
-
-Interesting changes to the tags occur where the program switches from
-a caller to a callee or vice-versa:
-
+\definition Let \(\components\) be a set of elements and \((\mach,\pol,\context)\)
+be the first state of a new call. We define \(R\) to hold on those states
+\((\mach',\pol',\context')\) where \(|\context'| < |\context|\). And we define
+\(I\) to hold on \(\mach'\) when \(\Delta(\mach,\mach') \cap \components\) is transient
+with respect to \(\mach'\). Then \((\mach,\pol,\context)\) enjoys
+{\it eventual intransient integrity} of \(\components\) if
+\(\mathit{when} ~ R ~ I ~ (\mach,\pol,\context)\).
+
+\definition A system enjoys {\it observational stack integrity} if, for all
+reachable states \(\mpcstatename\) such that
+\(\mpcstatename \harpoonunder{\mathbf{Call}\dots} (\mach,\pol,(tar, V)::\context)\),
+\((\mach, \pol, (tar, V)::\context)\) enjoys eventual intransient integrity with
+respect to the set of components that \(V\) maps to \(\sealed\).
+
+\paragraph*{Temporary Confidentiality}
+
+Confidentiality properties are also modeled as non-interference, but with a twist:
+the caller's confidential data should be kept secret from the callee, but after return,
+will naturally be accessible to the caller once again. So, we must model confidentiality
+in terms of events inside the callee. But confidentiality violations might not become
+visible during the callee's lifetime -- the callee might read a secret and stash it somewhere
+that it will be accessed later. So we must also consider the machine state at the end
+of the call. First we'll define a useful general property that implements this structure,
+similar to \(\mathit{when}\), called \(\mathit{during\_both}\).
+
+\(\mathit{during\_both}\) relates two initial states in their traces up until they
+both reach an end condition, again defines by a predicate \(R\). Instead of \(I\)
+we have an event-trace relation \(E\) that must relate the event traces up through
+\(R\), and a final relation \(F\) that must relate the final states.
+
+\definition Let \(\mpcstatename\) and \(\npcstatename\) be initial states
+and \(R\) be a predicate on states such that
+\(\mpcstatename \hookrightarrow \MPCT | R\) and
+\(\npcstatename \hookrightarrow \NPCT | R\).
+Let \(E\) be a relation on event traces, and \(F\) be a relation on machine states.
+Then \(\mathit{during\_both} ~ \mpcstatename ~ \npcstatename ~ R ~ E ~ F\) holds if both:
 \begin{itemize}
-
-\item From caller to callee: when the machine executes a call,
-  %, when the machine executes an instruction marked as a call in
-  % the call map.
-  the current
-  $\tagPCDepth{\depth}$ tag is incremented to $\tagPCDepth{(\depth + 1)}$, and the function
-  entry sequence initializes  all locations of the new function frame with
-  tag $\tagStackDepth{(\depth + 1)}$.
-
-\item From callee back to caller: the callee clears its frame, and at
-  the actual return instruction,
-  the $\PCname$ tag is decremented.
+\item \(E ~ (\obsof \MPCT) ~ (\obsof \NPCT)\) holds
+\item And, if \(\last ~ \MPCT = (\mach,\_,\_)\) and \(\last ~ \NPCT = (\nach,\_,\_)\),
+  then \(F ~ \mach ~ \nach\) holds.
 \end{itemize}
 
-To implement this discipline, blessed instruction sequences are designated
-to appear at the entry and exit of each function,
-which manipulate tags as just described in addition to performing the
-usual calling convention tasks of saving/restoring the return address to/from
-the stack and adjusting the stack pointer. With the aid of an additional tag on
-the $\rsp$ register, these sequences also serve to enforce
-well-bracketed control flow.
-The micro-policy guarantees atomic execution of these sequences
-using a combination of tags on the instructions
-and an additional tag on the $\PCname$.
-
-There remains the question of how to ensure that the sequences are
-invoked at the right places\ifspace from the perspective of the stack safety
-property\fi. For a program to enjoy stack safety
-(with respect to a particular set of call annotations), all we require is that
-the entry sequence immediately follows any instruction marked as a call
-in the call map. This is achieved by giving these instructions (another) special tag,
-and it is easy to check statically that this has been done correctly.
-If the code fails to initiate an exit
-sequence at a point where the stack safety property expects a return,
-the micro-policy will incorrectly behave as if execution
-is continuing in the callee, but since the callee never has stronger access
-rights than the caller, this is harmless.
-
-\medskip
-\paragraph*{Lazy Alternative}
-
-The conservative policy described above is rather slow in
-some important practical cases. Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18} observe
-that most of the performance overhead incurred
-stems from the need to set stack activation tags as a frame is created
-and reset them as it is destructed;  this
-is particularly costly for frames that contain large uninitialized arrays.
-To mitigate those costs, they eliminate the costly frame initialization pass.
-
-The most optimized version of the micro-policy, {\em Lazy Tagging and Clearing},
-does not initialize stack frames on entry \emph{or} clear them on exit,
-and it permits all writes to the stack, even when the $\PCname$ tag
-does not match the memory tag; the \(\PCname\) tag is always propagated
-to the written location. Reads from the stack \emph{do} require $\PCname$
-tag and memory tag to match. Thus, even if a callee illicitly writes to
-a private location in its caller, the caller will eventually detect this
-if it ever tries to read from that location.
-
-This \emph{lazy policy} admits a more efficient implementation, but
-deliberately allows temporary violations of stack integrity.
-The natural question then is: how do we characterize the
-protections provided by these policies once stepwise stack safety is broken?
-How does a harmful violation that must be caught later differ from a
-harmless one?
-
-\section{Observable Stack Safety}
-\label{sec:lazy}
-
-We answer these questions by introducing a notion of observations to our model.
-This will also be useful when, for example,
-a compiler wants to move data, overwrite its original location,
-and ultimately restore it before the change becomes visible, as in
-``callee-saves'' calling conventioons.
-These examples call for an abstract mechanism to characterize behaviors that are externally
-\emph{observable}.
-
-We abstract over an
-observation type \(\obs \in \OBSS\), containing at minimum the silent observation
-\(\tau\). Our machine step function now takes a state and returns a state and an observation,
-which is carried over into the policy and context step functions.
-
-\paragraph*{Observations-Of}
-
-This operator, written \(\obsof\), takes a machine trace and
-gives the trace of observations of each step in
-the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
-initial observation, a machine-trace, and an observation trace; the helper is
-then invoked with an initial \(\tau\).
-
-  \judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
-                {\(\mpcstatename' \hookrightarrow \MPCT\)}
-                {\(\mathit{obsfor} ~ \obs' ~ \MPCT = \obsT\)}
-                {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}
-
-\begin{minipage}[b]{.4\columnwidth}
-  \judgment{}
-           {\(\mathit{obsfor} ~ \obs ~ \mpcstatename = \obs\)}
-\end{minipage}
-\begin{minipage}[b]{.4\columnwidth}
-\judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
-         {\(\obsof(\MPCT) = \obsT\)}
-\end{minipage}
-
-\paragraph*{Observational Similarity}
-
-We say that two observation traces $\obsT_1$ and $\obsT_2$ are {\em similar}
-from the perspective of an external observer, written \(\obsT_1 \eqsim
-\obsT_2\), if the sequence of non-silent observations is the same. That is, we
-compare up to deletion of \(\tau\) observations, coinductively:
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{}{\(\obsT \eqsim \obsT\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
-\end{minipage}
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
-\end{minipage}
-
-\sna{Necessary? }
-Note that an infinite silent trace is a
-prefix of (and similar to) any other trace. While this might seem
-surprising at first, it makes sense in a timing-insensitive context:
-an external observer looking at two machine runs cannot (computably)
-distinguish between a machine that steps forever and a machine that
-steps for a long time before producing some output.
-
-\paragraph*{Observational Stack Safety}
-
-We now sketch observational versions of both integrity and confidentiality.
-Integrity is already straightforward, simply weakening the condition that must hold
-after the return---instead of preventing dangerous writes from being read, we
-prevent them from being leaked.
-
-\definition
-\textbf{\em Observational integrity} means that for any \((\MPCT,\brk) \in \MPCT^*\)
-where \(\head(\MPCT) = (\mach,\context\)) and \(\last(\MPCT) = (\mach',\context')\),
-the following holds. Let \(\components = \prot(\context) \cap \Delta(\mach, \mach')\)
-be the set of protected elements that changed during
-the call,
-and let \((\mach',\context') \hookrightarrow \MPCT'\).
-Then for any \(\nach\) such that
-\(\mach' \approx_{\components'} \nach\), and its induced trace
-\((\nach,\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
-
-\medskip
-
-For confidentiality, we must capture the intuition that
-secrets do not escape during a call {\em and} are not leaked during a return.
-This require us to be able to quantify over elements that fail to preserve
-the change-together relation~\(\diamond\).
+We could define a natural, strict form of confidentiality that requires that the final
+states be identical. However, as with integrity, this is stricter than we need -- there
+may be some conditions in which a secret is copied, but never used again. So, we once
+again wish for our final condition to make the values {\it transient}. Which values?
 
@@
 
-\definition
-A system enjoys \textbf{\em observational confidentiality} if, for all
-\((\MPCT,\brk) \in \MPCT^*\) with \(\head(\MPCT) = (\mach,\context)\) and
-\(\components = \secret(\context)\), the following holds.
-%
-For any \(\nach\) such that \(\mach \approx_\components \nach\) and
-\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
-\(\MPCT\) and \(\NPCT\) respect three conditions:
-
-\begin{itemize}
-\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
-  meaning that no secrets leak during the call
-\item \(\brk ~ (\last(\MPCT)) \leftrightarrow \brk ~
-(\last(\NPCT))\), i.e. if one call returns, so does the other
-\item Assuming both calls return, let \(\last(\MPCT) = (\mach',\context')\)
-  and \(\pi_\mach(\last(\NPCT)) = \nach'\),
-  let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\),
-  and let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
-  Then if \((\mach',\context') \hookrightarrow \MPCT'\) and
-  \((\nach',\context') \hookrightarrow \NPCT'\), it must be
-  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
-\end{itemize}
-
-This last condition is the most interesting: we identify the elements that
-contain leaked information on either side of the first pair of variant traces.
-This becomes the basis for the second variation, as all of these values
-must be protected from leaking for the remainder of the program.
-
-These two properties give us
-{\em Observable Stack Safety}. This is a weaker property
-than the stepwise one, and also much more extensional.
-We can think of the stepwise properties as instantiating
-the observation model with a view of the entire state---one end of a
-spectrum that can become arbitrarily permissive.
-Thus, if we had to choose a single property to represent stack safety,
-it would be the observational one.
-
-\medskip
-\paragraph*{Connection to Policies}
-
-Observable properties allow us to defer enforcement until a property
-violation would become visible. Now we can identify when lazy policies
-miss such violations. It turns out that {\em Lazy Tagging and Clearing},
-as written, falls short of observational stack safety.
-If, at depth \(d\), a callee \(A\) writes into its caller's
-frame, then returns, and later another callee \(B\) at depth \(d\) reads the same
-address and prints it, \(A\) has violated integrity in a way that later becomes visible,
-and \(B\) has likewise violated confidentiality!
-
-{\em Lazy Tagging and Clearing} can be repaired if, instead of tagging each function
-activation with its depth in the stack, we generate a fresh activation
-identifier on each call, which prevents a program from exploiting stale
-tags from previous activations.
-(A related mechanism was explored in the
-Static Authorities policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}, which
-associates a unique \emph{static} identifier to each function, which is shared
-by all activations---but that policy allows leaks between
-instances of the same function.) We validate this repaired micro-policy in
-\cref{sec:testing}.
-
-\medskip
-\paragraph*{Callee-saves Registers}
-\label{sec:callee}
-
-Another application of observational properties is characterizing a ``callee-saves''
-calling convention. Intuitively, a caller expects a callee-saves register to be unchanged on
-return, but it may change during the call, as long as it is restored. We similarly expect a callee-saves
-register to be secret from the callee, except when it is copied to the stack. Both scenarios
-reflect observational rather than stepwise stack safety.
-
-We model this by adding information to Call annotations defining which registers will be {\em saved}
-and which are {\em unsaved}. The stack pointer and the register containing the return address
-should never be saved. We then extend the set of domains with these values, for the first time associating
-(non-{\em outside}) domains to registers as well as addresses. On each call, we save precisely those
-registers given by the annotation, and the rest are unsaved; {\em saved} registers are treated as part
-of the protected and secret sets for purposes of \textbf{\em Observational stack integrity} and
-\textbf{\em Observational stack confidentiality}, respectively. For full formal definitions, see
-Appendix \ref{app:calleesave}.
-
-\medskip
-\paragraph*{Parameters Passed on the Stack}
-%
-Up to this point we have ignored function arguments, by assuming that they are placed
-in registers (which are implicitly shared by all functions).
-We now show how to implement two kinds of protected
-sharing of stack contents: passing by value on the stack and passing by reference.
-Here we give an example and high level discussion; the formalization can be found in Appendix \ref{app:argpass}.
-
-Figure~\ref{fig:passing} shows a sample trace of a program that passes
-variables by reference. In this example, written in C++ notation,
-{\tt f} takes a single argument by reference (the diagram displays it in magenta). The function
-{\tt main} calls {\tt f}, passing a reference to {\tt x}, then clears {\tt x} and calls
-{\tt f} again, passing a reference to {\tt y}.
-We would expect that the first call to {\tt f} can modify
-{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
-{\tt x} in the second call. (Incidentally, this is perfectly compilable C++,
-although its behavior is officially undefined.)
-
-To avoid cluttering the picture at the assembly level, we revert to assuming that the
-argument {\tt a} is passed in a register, namely {\tt \%ra}.
-The question is how the local variables {\tt x} and {\tt y} themselves
-should be labelled in order for the property to correctly classify an integrity violation.
-Our notion of domains is extended to include \(\passed{\depth,\depth'}\),
-recording two depths for reasons explained below.
-At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
-return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
-and when {\tt f} instead tries to write to {\tt x}, it violates integrity.
-
-When passing by reference, it is possible to repeatedly pass the same reference further
-and further down the stack. Note, however, that this sharing of the reference must be over
-an unbroken chain of calls. So it suffices for the domain to track two depths for each passed
-cell: the original owner, and the highest depth it has been passed to. The policy considers
-it to be accessible anywhere between the two.
-
-Our approach provides only limited support for handling fully first-class pointers into
-stack frames, e.g., pointers obtained by taking the address of a local variable.
-Unlike with pass-by-reference, there is no obvious way to model the intended
-sharing behavior of such pointers in terms of stack frame activation patterns.
-In general we expect such a pointer to behave similarly to heap pointers, 
-except that its lifespan is connected to that of the frame that contains
-it. In order to give a satisfactory property, then, we should tie it to a broader
-heap safety property.
-
-An alternative is to treat every address-taken variable in the
-same way as a passed argument, with the modification 
-that it remains accessible at \emph{all} stack depths above the caller.
-Callees would be permitted
-to access the location without having been explicitly given a pointer
-to it. This would prevent most violations of temporal safety, but stale pointers
-could still be used in certain corner cases.
-
-\medskip
-
-\paragraph*{Coroutines}
-
-Our final extension to the model adds coroutines. We assume a very simple model with
-a fixed number of coroutines, each with its own stack.
-Stack identifiers are drawn from a set \(\STACKS\), and there is a static partial map
-\(\stackof\) from addresses to stack ids.  Annotations and code maps are unchanged from
-the subroutine model.
-%\[\ann \in \ANNS ::= \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
-%\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \ann \in \ANNS\]
-
-The first major difference in this model is the domains. Stack domains are now nested inside
-top-level domains, \(\topDom \in \TOPDOMS ::= \outside + \instack{\stackid}{\stackDom}\),
-associated with stack identifiers.
-Our context now consists of a domain map, a map from stack identifiers to return target stacks,
-an additional map of targets for yields, and a stack identifier of the active stack.
-A yield target map \(\yts\) is a map from stack identifiers to targets.
-In a standard system an initial yield target map will map each stack identifier to the
-set of all states with their stack pointer at the stack base, and include any restrictions
-on entry points. Just as a return target tells us that we have really returned, a yield
-target tells us that we have really yielded. When yielding from state \(\mach\),
-\(\ycon ~ \mach \in \MACHS\) gives us the yield target for coming back to the yielding
-coroutine.
-%
-\[\begin{array}{lr}
-\multicolumn{2}{l}{(\domMap, \rtm, \yts, \stackid) \in \CONTEXTS ::=} \\
-& (\COMPONENTS \rightarrow \TOPDOMS) \times
-(\STACKS \rightarrow \listT(\target)) \times 
-(\STACKS \rightarrow \target) \times \STACKS
-\end{array}\]
-
-The initial domain map maps each stack address in stack \(\stackid\) to
-\(\instack{\stackid}{\unsealed}\), the initial return map maps all stack identifiers to empty
-return stacks, the initial yield targets are defined as above, and the initial stack id matches the initial
-coroutine. Together these form the initial context.
-
-In our context update rules, calls and returns work the same as in the subroutine models,
-with domain updates restricted to the top-level domain corresponding to the active stack,
-and likewise using the return target stacks. We will omit these rules and focus on the new yield rule.
-
-\begin{tabular}{l}
-  {\sc Yield} \\
-  \judgmenttwobr[]
-                {\(\mach \stepsto \mach'\)}
-                {\(\stackid \not = \stackid' = \stackof ~ (\mach'[\rsp])\)}
-                {\(\yts ~ \stackid'\)}
-                {\(\yts' = \yts[\stackid \mapsto \ycon ~ \mach]\)}
-                {\(\mach,(\domMap,\rts,\yts,\stackid) \stepsto_C \domMap,\rts,\yts',\stackid'\)} \\
-\end{tabular}
-
-First, we determine which coroutine we are yielding to by
-checking which stack the stack pointer of the next state is pointing to. Then we can only yield
-if the next state is a valid yield target for that stack, and we update the yield target for
-the current stack, ensuring that when the system yields back we expect the current coroutine
-to continue from where it left off. So, it is possible for a coroutine to claim to yield
-but not actually change the active stack, with similar effects to when a callee fails
-to return properly. Now we can define both stack safety and coroutine safety under this model.
-
-Stack integrity and confidentiality apply in the coroutine model, based on straightforward
-projections of each stack's domains, etc. and the machinery of the passing model. We will give
-the coroutine-specific properties.
-
-\definition The \emph{yield set} \(\MPCT^\wedge\) of a system is
-\[\begin{aligned}
-\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{(\MPCT, \neg f_\stackid) | \MPCT\langle f_\stackid \rangle
-\text{ is a reachable segment}\} \hspace{1em} \\
-\text{where } f_\stackid ~ (\underscore,\underscore,(\underscore,\underscore,\underscore,\stackid')) \triangleq \stackid \not = \stackid' \\
-\end{aligned}\]
-
-\noindent
-Here, we are capturing the entire span that the system spends {\em away} from a given
-coroutine.
-
-\definition
-Let both \(\prot_Y(\domMap,\rtm,\yts,\stackid)\) and \(\secret_Y(\domMap,\rtm,\yts,\stackid)\) be the set of elements \(\component\) such that
-\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
-
-\definition
-We define \textbf{\em stepwise coroutine integrity} in the same way as stepwise stack integrity,
-substituting the above \(\CONTEXTS\) type, initial context and update
-function, and \(\prot_Y\) for their equivalents. Likewise
-\textbf{\em observational coroutine integrity} is defined
-in terms of their simple equivalents using \(\MPCT^\wedge\) in place of \(\MPCT^*\).
-
-\definition
-We define \textbf{\em stepwise coroutine confidentiality} in the same way as stepwise stack
-integrity, substituting the above definitions. Likewise
-\textbf{\em observational coroutine confidentiality} is defined in terms of observational
-stack confidentiality.
-
-\medskip
-
-\paragraph*{Exceptions}
-
-Exceptions are another useful form of control flow that disobeys the typical call and return pattern.
-We can extend our model easily to handle integrity and confidentiality in the presence of exceptions.
-In a typical exception model, the code point that ``catches'' a thrown exception is indicated by a {\tt setjump}
-instruction or equivalent, which is later jumped to by a {\tt longjump}. We treat {\tt setjump} as placing
-a special return target on the target stack, which then behaves like a normal return in terms of our properties.
-A full treatment is given in Appendix \ref{app:exceptions}.
-
-\section{Validation through Random Testing}
-\label{sec:testing}
-
-There are several ways to evaluate whether an enforcement mechanism enforces
-stack safety properties. Ideally such validation would be done through formal proof over
-the semantics of the enforcement-augmented machine.
-However, while there are no fundamental barriers to producing such a proof,
-it would be considerable work to carry out for a full ISA like RISC-V and
-complex enforcement mechanism like the Depth Isolation micro-policy.
-We therefore choose to validate the micro-policy of \cref{sec:enforcement} by 
-systematically \emph{testing} that it satisfies our properties.
-This focus on testing is better aligned with our immediate
-goal of evaluating real enforcement mechanisms for real machine architectures.
-Formal proof remains future work.
-
-We validate the micro-policy of the previous section by
-systematically testing that it satisfies eager stack integrity and
-confidentiality. We use a Coq specification of the RISC-V
-architecture~\cite{Bourgeat2021AMF},
-and extend it with a runtime monitor implementing a stack safety
-micro-policy. We chose the Coq proof assistant as the setting for our implementation
-to ensure that our coinductive trace definitions are well-formed, to reason about
-them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.
-
-To use QuickChick, we build random test-case generators that produce
-% \begin{enumerate}
-% \item
-(1) an initial RISC-V machine state, including most notably
-  the program to be executed;
-  % \item
-(2)
-  an initial policy state, tagging instructions corresponding to
-  blessed call or return sequences appropriately, while marking
-  all potential stack locations as $\tagNoDepth$; and
-% \item
-(3)
- a code map identifying the locations of calls.
-% \end{enumerate}
-
-To write such generators we build on the work of
-Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
-introduced {\em generation by execution} to produce progams that lead
-machines towards interesting behaviors. We extend the technique by tracking
-control flow information so that programs will usually---but not always---return to their call sites.
-
-%\paragraph*{Generators}
-%
-% Generation by
-%execution receives as an input a partially instantiated machine state
-%and attempts to generate an instruction (or a sequence of instructions
-%such as a blessed sequence) that makes sense locally (e.g., jumps go
-%to a potentially valid code location and loads read from a
-%potentially valid stack location). Then we step the machine and repeat
-%the process until we generate or execute some target number of
-%instructions, or reach a point where the machine cannot step
-%any more.
-%
-%We extend this technique to keep track of the control flow behavior of
-%the program being generated: each time a call or return sequence is
-%generated, we ensure that the appropriate policy tags and code
-%annotations are set for the entry or return points.% At the same time,
-%%we allow the generation to sometimes relax those constraints,
-%%introducing potentially ill-formed flows: this causes our programs
-%%to failstop when executed in conjunction with the policy monitor,
-%%but also allows for revealing errors in our setup.\apt{??}
-%
-We need to extend typical testing schemes further to handle the nested
-nature of confidentiality: rather than just generating two
-initial machines that are variants of one another and letting them
-execute to test for noninterference, we generate a new variant
-{\em every time a call is made} and check confidentiality for the
-subtrace produced from that variant state until its corresponding
-return. As a result, a ``single'' confidentiality test compactly
-checks multiple nested calls.
-
-Our primary testing targets are the eager {\em Depth Isolation}
-and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
-We test the former against stepwise integrity and confidentiality, and
-the latter against their observational counterparts.
-
-To ensure the effectiveness of testing against our formal properties, we
-use {\em mutation testing}~\cite{JiaH11}. In mutation testing, we inject errors
-(mutations) in a program that should cause the property of interest (here,
-stack safety) to fail, and ensure that the testing framework can find
-them. The bugs we use for our evaluation are either artificially generated
-by us (deliberately weakening the micro-policy in ways that we expect
-should break its guarantees), or actual bugs that we discovered through
-testing our implementation. We elaborate on some such bugs below.
-
-For example, when loading from a stack location, {\em Depth Isolation}
-needs to enforce that the tag on the location being read
-is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
-current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
-that restriction by not checking the depth equality (row {\em
-  LOAD\_NO\_CHECK\_DI}).
-
-Similarly, when storing to a stack location, the correct micro-policy
-needs to ensure that the tag on the memory location is either
-$\tagNoDepth$ or has again the same depth as the current $\PCname$
-tag. Relaxing that constraint causes violations to the integrity
-property (row {\em STORE\_NO\_CHECK}).
-
-\begin{table}[]
-\centering
-\begin{tabular}{c|c|c|c}
-  Bug & Property Violated & MTTF (s) & Tests \\
-  \hline
-      {\em LOAD\_NO\_CHECK\_DI}  & Confidentiality & 24.2 & 13.3 \\
-      {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
-      {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
-  \hline
-  \hline
-      {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
-      {\em LOAD\_NO\_CHECK\_LT}  & Obs. Integrity & 23.5 & 12.0 \\
-      {\em LOAD\_NO\_CHECK\_LT}  & Confidentiality & 19.2 & 695.5 \\
-      {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
-      {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
-  \hline
-\end{tabular}
-\vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
-\vspace*{-2em}
-\label{tab:bug-table}
-\end{table}
-
-The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
-Table~\ref{tab:bug-table}, along with the average number of tests
-it took to find the failure. Experiments were run in a desktop
-machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
-
-Naturally, testing also revealed a number of errors in our
-implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code;
-%\bcp{right?}\leo{yeah}
-we re-implemented it in Coq targeting RISC-V).  These errors range
-from trivial typos to ones that require an intriguingly complex setup
-to provoke.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
-sequence %/policy combination\apt{??}
-did not initialize all locations for the
-newly allocated stack frame correctly, but left some of them as
-$\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured.
-The smallest counterexample requires calling a function {\tt f},
-which fails to initialize some of its frame during the blessed sequence,
-but writes into an uninitialized location $l$ later, treating \(l\) as outside
-the stack. Then {\tt f} calls a further function {\tt g} (which should have
-the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
-which is allowed because the enforcement mechanism still has
-$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
-%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
-%  and the write from {\tt f} was already a violation.}
-
-As for lazy micro-policies and observational properties,
-the original {\em Lazy Tagging and Clearing} micro-policy, implemented as {\em PER\_DEPTH\_TAG},
-fails in testing, in cases where data is leaked between sequential calls.
-To round out our mutation testing we also check {\em LOAD\_NO\_CHECK\_LT},
-equivalent to its counterpart in depth isolation,
-and a version where stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
-It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
-taking twice as long to catch as the next longest.
-
-Our properties have allowed us to identify an enforcement mechanism as
-not really stack safe, and to validate a possible fix.
-Unfortunately, Lazy Per-activation Tagging and Clearing
-puts severe pressure on the tag cache, so it may not be usable in practice.
-However, we conjecture that no lazy scheme can be fully stack safe
-without using unique per-activation identifiers.
-
-\section{Related Work}
-\label{sec:relwork}
-
-\paragraphx{Formal Stack Safety on Capability Machines}
-%
-To our knowledge, the only line of work to date that has attempted
-a positive and formal characterization of stack safety
-is that of Skorstengaard et al. \cite{SkorstengaardSTKJFP}. Their
-StkTokens system enforces stack safety on hardware equipped with
-memory capabilities, similar to those used in the Cheri system~\cite{Woodruff+14,Chisnall+15}, but
-with additional support for linear capabilities.
-They define Stack safety as the conjunction of local state encapsulation (LSE) and
-well-bracketed control flow (WBCF).
-Informally, LSE means that the current activation can only access its own frame,
-and WBCF means that the topmost frame can only return to
-the immediately adjacent frame below. Formally, these properties are defined
-with reference to a capability machine with a built-in call stack and call and return
-instructions; this machine is proven fully abstract with respect to a more
-concrete capability machine, an approach termed ``fully abstract overlay semantics.''
-We improve upon Skorstengaard et al. \cite{SkorstengaardSTKJFP} in two significant ways:
-
-i) Our definition captures the intuition that {\em
-  every} caller should be protected from its callees.  By contrast,
-  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
-  that are not protected, because they are capable of compromising
-  their own security. Consider a function that, when called with the wrong
-  inputs, behaves unreasonably. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
-  {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
-  component is always reasonable. In our setting, individual
-  well-behaved activations of the buggy function would be protected.
-
-ii) Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s fully
-  abstract overlay semantics rely on the definition of a ``stack-safe-by-construction''
-  abstract machine that is closely tied to the underlying machine architecture.
-  While the approach can no doubt be ported to other architectures and enforcement
-  schemes, each such port will require a new abstract machine definition whose
-  safety-by-construction must be argued afresh.
-  Since we abstract over most of the machine state as well as the sealing and return conventions, and
-  can support further enhancements such as coroutines,
-  our properties apply directly to a range of architectures, code features, and enforcement mechanisms.
-
-Skorstengaard et al. \cite{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
-candidate stack safety properties.
-%% : they should be (1) intuitive; (2)
-%% useful for reasoning; (3) reusable in secure compiler chains; (4)
-%% arguably ``complete'' with respect to informal notions of stack
-%% safety; and (5) potentially scalable to richer settings such as
-%% dynamic code generation or multithreading.
-%
-In terms of (1) {\em intuitiveness}, we contend that our property
-definitions are at least as
-intuitive as theirs, though phrased in rather different terms.
-We use the language of formal security
-properties to apply well-established concepts \ifspace and formal techniques \fi
-of noninterference, and we phrase our properties in terms of
-traces of memory and register states, rather than an
-ad-hoc abstract machine.
-Which approach is ``better'' is probably a matter of taste, but ours
-has \ifspace at least \fi the merit of explicitly relating stack safety to existing
-notions of security.
-%
-Demonstrating (2) {\em usefulness} of the safety criterion
-for carrying out formal proofs (about individual programs) for our proposal
-remains future work; concretely, we
-hope to apply our properties to justifying a low-level program logic.
-%
-(3) StkTokens is specifically set up to address {\em reusability} in secure
-compilation chains, while this is not a goal of our work.\bcp{Can we explain
-what it means??}
-%
-(4) Regarding {\em comprehensiveness} in capturing programmers' informal
-understanding of stack safety, we believe that our support
-for a dynamically changing notion of ``adversary,'' as outlined in point (i) above,
-captures an even better definition of safety than LSE; moreover, our observational
-properties are richer and\bcp{cut "richar and"??} more extensional, and thus
-applicable to a broader range of enforcement mechanisms and code features.
-%
-(5) Finally, regarding {\em broad applicability}, we have demonstrated that
-\bcp{unlike StkTokens}
-our approach extends to
-coroutines, a first step towards multithreading. Dynamic code generation is
-more challenging, but StkTokens does not yet handle this either.
-
-We would add one more criterion: reusability across machine models and
-enforcement mechanisms. StkTokens has been used to validatate
-one machine/enforcement mechanism---a linear-capability machine---via formal
-proof. We have used ours to obtain a weaker (testing-based) validation of
-two different mechanisms on a single machine; we hope to target a capability machine
-next as future work.
-
-\paragraphx{Protecting the Stack with Micro-Policies}
-%
-\apt{Could probably condense/omit this.}
-Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}
-study the protection of stack data through three families of
-micro-policies that tag stack objects with a pair of frame and object identifiers
-used to validate accesses: {\em Return Address Protection} (which
-prevents an adversary from overwriting designated return addresses), {\em Static
-Authorities} (which only allows the code of a function to access the stack frames
-of its own dynamic instances), and {\em Depth Isolation}
-(\cref{sec:enforcement}). All these policies strike various balances between
-precision and performance, evaluated through benchmarks that demonstrate
-limited performance overhead. In addition to the baseline ``eager'' policies,
-they propose a number of lazy optimizations, notably {\em Lazy Tagging} and
-{\em Lazy
-Clearing} (\cref{sec:lazy}).
-
-We evaluate {\em Depth Isolation} and {\em Lazy Tagging and Clearing}
-against our properties, finding that the former does
-indeed enforce stepwise stack safety and therefore observable stack
-safety. We find that the latter does not actually enforce
-stack safety, and we proposed a fix.
-
-\paragraphx{Heap Safety as a Security Property}
-%
-Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18} show that 
-heap safety may be framed as a security property using
-the language of noninterference. Their paper was a significant inspiration
-for the present work, and we believe that the two safety conditions should
-combine smoothly.
-
-\section{Future Work}
-\label{sec:future}
-
-From here, we see two natural ways to extend the work. \apt{what about addressing some of the limitations?}
-First, we should thoroughly evaluate
-other enforcement mechanisms, especially Cheri-based capability
-systems. Second, we can further demonstrate
-the functionality of our definitions by using them to prove
-the correctness of a programming logic with strengthened rules around calls and returns.
-
-\paragraphx{Stack Safety in the Cheri Ecosystem}
-%
-There are several proposals around the use of Cheri capabilities to enforce stack safety,
-including mechanisms that use the standard Cheri hardware (which includes local
-capabilities) \cite{SkorstengaardLocal},
-and others that propose entirely new types of capabilities, such as linear
-\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
-\cite{Tsampas+19}, and monotonic \cite{GeorgesComplete} capabilities.
-Of these, uninitialized capabilities seem particularly promising---they naturally enforce
-write-before-read behavior and therefore will enable confidentiality without the need
-to clear large swaths of the stack. We plan to implement a model of the stack safety
-system described by Georges et al. \cite{Georges+21} and test it against our properties using
-QuickChick.
-
-\paragraphx{A Program Logic for Stacks}
-%
-Stack integrity suggests a reasoning principle in which knowledge about a
-caller's local data can be carried across a call site. We aim to develop
-this idea into a low-level separation logic with a ``call rule'' that
-preserves propositions whose domains are sealed across a call, while
-allowing those propositions proven about the behavior of the callee to apply
-to the accessible stack.
-%
-It is easy to imagine that such a rule could be proven sound in a system
-with lockstep stack safety. More challenging would be to ask what kind of
-proof rule would be appropriate for a system that obeys only our
-observational properties.
-
-%% Acknowledgments
-%\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-%  This material is based upon work supported by the
-%  \grantsponsor{GS100000001}{National Science
-%    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%  conclusions or recommendations expressed in this material are those
-%  of the author and do not necessarily reflect the views of the
-%  National Science Foundation.
-%\end{acks}
-
-%% Bibliography
+\definition Let \((\mach,\pol,\context)\) be a state and \(\components\) be a set of
+elements. Let \(R\) hold on contexts \(\context'\) such that \(|\context'| < |\context|\).
+Let \(E \triangleq ~ \simeq\).
+Let \(F ~ \mpcstatename ~ \npcstatename \triangleq
+\mathit{transient} ~ \mpcstatename ~
+\bar{\Diamond}(\mach,\pi_m(\mpcstatename),\nach,\pi_m(\npcstatename))\).
+
+\((\mach,\pol,\context)\) enjoys {\it temporary confidentiality} of \(\components\)
+if, for all \(\nach\) such that \(\mach \approx_{\components} \nach\),
+\(\mathit{during\_both} ~ R ~ E ~ F\).
+
+\definition A system enjoys {\it observational caller confidentiality} if, for all
+reachable states \(\mpcstatename\) such that
+\(\mpcstatename \harpoonunder{\mathbf{Call}\dots} (\mach,\pol,(tar, V)::\context)\),
+\((\mach, \pol, (tar, V)::\context)\) enjoys temporary confidentiality with
+respect to the set of components that \(V\) maps to \(\unsealed\) or \(\sealed\).
+
 \bibliographystyle{IEEEtran}
@@
 
-\pagebreak
-
-%% Appendix
-\appendices
-\section{Formalization of Callee-saves Registers}
-\label{app:calleesave}
-
-In order to support callee-saves registers, we extend {\sc Call} annotations to
-carry information about which registers are to be saved and which are used for arguments,
-etc. This takes the form of a ``saving function'' \(S\) that maps registers to domains \(\saved\) and
-\(\unsaved\). These also extend \(\STKDOMS\).
-
-\[\begin{split}
-\codemap \in \CODEMAPS ::= & \WORDS \rightharpoonup \callmap(S)\\
-\mbox{\rm{where}} ~ & S \in \REGS \rightarrow (\unsaved + \saved) \\
-\end{split}\]
-
-\[\begin{split}
-\stackDom \in \STKDOMS ::= & \outside + \unsealed + \sealed{\depth} \\
-& + \unsaved + \saved \\
-\end{split}\]
-
-This property reuses most of the structure of the core property. Here we will highlight
-the definitions that have changed. Namely, the {\sc Call} rule for \(\stepsto_C\) now
-applies the saving function to determine which registers are saved. The remaining rules
-remain unchanged. Note that this means that at any given time, a register will be marked
-\(\saved\) or \(\unsaved\) depending on the most recent call chronologically (i.e., it is
-not nested like \(\sealed{\underscore}\)). This is because the protected and secret sets will be based
-only on the entry state of the call, and those are all that affect observational properties.
-
-\begin{tabular}{l}
-  {\sc Call} \\
-  \judgmenttwobrlong[]
-                    {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
-                    {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
-                    {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
-                    {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)} \\
-\end{tabular}
-
-\definition Let the protected set of a context, \(\prot_S(\domMap,\rts)\),
-be the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
-or \(\domMap ~ \component = \saved\).
-
-\definition Let the secret set of a context, written \(\secret_S(\domMap,\rts)\),
-be the set of elements \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
-\(\domMap ~ \component = \saved\).
-
-\definition
-%
-\textbf{\em Observational stack integrity} with callee-saves registers substitutes the above
-\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
-%
-\textbf{\em Observational stack confidentiality} additionally substitutes \(\secret_S\).
-
-\section{Formalization of variable passing.}
-\label{app:argpass}
-
-We add extra information to our call
-annotations to reflect that they may now pass variables on the stack, and that the locations
-of those variables may be dynamic (as in the case of pass-by-reference).
-A call annotation now includes a predicate relating machine states to the elements
-that should be passed, expressed as offsets from the stack pointer.
-\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P)~\mbox{\rm where}~P \subseteq \WORDS\]
-We extend the set of domains to include passed elements,
-which are labeled with the range of depths of the caller that has passed them.
-\[\begin{split}
-\stackDom \in \STKDOMS ::= & \outside + \unsealed\\
-& + \sealed{\depth} + \passed{\depth,\depth'}
-\end{split}\]
-
-We straightforwardly describe how the sealing convention and passing predicate combine to
-update a domain map at a call from depth \(\depth\), written as functions \(\mathit{push}\)
-and \(\mathit{pop}\) and written here as inference rules.
-
-\judgmentthree[]
-              {\(\domMap ~ \component = \unsealed\)}
-              {\(\sealcon ~ \mach[\component]\)}
-              {\(\component - (\mach ~ \rsp) \not \in P\)}
-              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \sealed{\depth}\)}
-\judgmentthree[]
-              {\(\domMap ~ \component = \unsealed\)}
-              {\(\sealcon ~ \mach[\component]\)}
-              {\(\component - (\mach ~ \rsp) \in P\)}
-              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth,\depth+1}\)}
-\judgmentthree[]
-              {\(\domMap ~ \component = \passed{\depth',\depth}\)}
-              {\(\sealcon ~ \mach[\component]\)}
-              {\(\component - (\mach ~ \rsp) \in P\)}
-              {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \passed{\depth',\depth+1}\)}
-\judgment[]
-         {}
-         {\((\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) \component = \domMap ~ \component\)}
-
-\judgment[]
-         {\(\domMap ~ \component = \sealed{\depth}\)}
-         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-\judgment[]
-         {\(\domMap ~ \component = \passed{\depth-1,\depth}\)}
-         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-\judgmenttwo[]
-           {\(\domMap ~ \component = \passed{\depth',\depth}\)}
-           {\(\depth' < \depth\)}
-           {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-\judgment[]
-         {}
-         {\((\mathit{pop} ~ \domMap ~ \depth) \component = \domMap ~ \component\)}
-
-These are applied to update the context at calls and returns, respectively.
-
-\begin{tabular}{l}
-  {\sc Call} \\
-  \judgmentbr[]
-              {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
-              {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)} \\
-\end{tabular}
-
-\vspace*{-1ex}
-\begin{tabular}{l}
-  {\sc ReturnFound} \\
-  \judgmenttwo[]
-              {\(\mach \stepsto \mach' \in \target\)}
-              {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
-              {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)} \\
-\end{tabular}
-
-\vspace*{-1ex}
-\begin{tabular}{l}
-  {\sc ReturnRec} \\
-  \judgmenttwo[]
-              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
-              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
-              {\(\mach,(\domMap,\underscore::\rts) \stepsto_C \domMap'',\rts'\)} \\
-\end{tabular}
-
-\definition The \emph{protected set} of a context, \(\prot_P(\domMap,\rts)\),
-is the set of elements \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
-or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
-\(|\rts| > \depth'\).
-
-\definition The \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
-is the set of elements \(\component\) such that
-\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
-\(\domMap ~ \component = \passed{\depth,\depth'}\)
-where \(|\rts| < \depth\) or \(|\rts| > \depth'\).
-
-\definition
-For each of the \textbf{\em stepwise} and \textbf{\em observational} versions of
-\textbf{\em stack integrity} in the simple model, the equivalent property in the
-passing model substitutes the above \(\CONTEXTS\) type and update
-function, and \(\prot_P\), for their equivalents in the simple model.
-
-For both the \textbf{\em stepwise} and \textbf{\em observational} versions of \textbf{\em stack confidentiality} in
-the original model, the passing model equivalents additionally substitute \(\secret_P\).
-
-\section{Formalization of Exceptions}
-\label{app:exceptions}
-
-Our core formalism already models control flow patterns in which a return skips over one or more callers to
-reach a point further down the call stack. This is essentially the behavior of an exception; we need only add an
-annotation representing the {\tt setjump} operation or equivalent to enable a ``return'' to an arbitrary point
-not associated with a call. We add an annotation, {\sc SetJump}, and define a target label as either a call-return
-or an exception, \(\mathit{TL} ::= \text{\sc CR} + \text{\sc EX}\), which annotates targets in the return stack.
-\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS) \times \listT (\mathit{TL} \times \target) \]
-
-We add a rule to handle {\sc SetJump} annotations, and modify the {\sc Call} rule to insert {\sc CR} labels into the
-return stack.
-
-\vspace*{1em}
-\begin{tabular}{l}
-  {\sc SetJump} \\
-\judgment[]
-         {\(\codemap ~ (\mach ~ \PCname) = {\sc SetJump}\)}
-         {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc EX}, \retcon ~ \mach)::\rts\)}
-\end{tabular}
-
-\begin{tabular}{l}
-  {\sc Call} \\
-  \judgmenttwobrlong[]
-                    {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
-                    {\(\components = \{\component | \sealcon ~ \mach[\component]\}\)}
-                    {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
-                    {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc CR}, \retcon ~ \mach)::\rts\)} \\
-\end{tabular}
-
-The return rules operate identically to the core model, regardless of the target labels.
-
-\definition
-We define \textbf{\em stepwise stack integrity}, \textbf{\em stepwise stack confidentiality},
-\textbf{\em observational stack integrity}, \textbf{\em observational stack confidentiality},
-in the same way as in the core model, substituting the above \(\CONTEXTS\) type and update
-function for their original.
-
-\definition
-A system enjoys \textit{\textbf{well-bracketed control flow}} with exceptions when, for every reachable state
-\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
-\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
-if \(|\rts'| < |\rts|\), then either \(|\rts'| = |\rts|-1\) or \(\rts = \underscore +\!\!+ (\text{\sc EX},\underscore)::\rts'\).
-
 \end{document}

### 2022-08-23 submission5/onlyformalism.tex
Working on formalism still

@@
 where \(\mathcal{T}\) is the set of tags, discussed [in the section on security.]
-We represent a {\it policy} via the partial function
-\(\cdot, \cdot \rightharpoonup_P \cdot \in \MACHS \rightarrow \POLS \rightharpoonup \POLS\),
-where the function being undefined represents a {\it failstop}.
+We represent a {\it policy} via the step relation
+\(\cdot, \cdot \rightharpoonup_P \cdot \subseteq \MACHS \times \POLS \times \POLS\).
+Steps are deterministic and  undefined represents a {\it failstop}.
 
@@
 \[sc \in SEC ::= \sealed | \passed | \mathit{public} | \unsealed\]
-And of {\it targets}:
+And of {\it targets}, which identify the addresses of the program counter or
+stack pointer at a return or catch point:
 \[\begin{split}
-tar \in TAR ::= & \mathit{Return} ~ a_1 ~ a_2 \\
-| & \mathit{Catch} ~ a_1 ~ a_2 \\
+tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
+| & \mathit{Catch} ~ a_{pc} ~ a_{sp} \\
 \end{split}\]
-A {\it security view} \(V \in \COMPONENTS \rightarrow SEC\) divides the machine
+A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\) divides the machine
 elements in terms of their purpose relative to the active function.
 And a security context is a stack of pairs of targets and views:
-\[\context \in \CONTEXTS ::= \mathit{list} ~ (TAR \times SEC)\]
+\[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]
 
-The set of instructions, \(I\), contains the RISC-V ISA. There is additionally a set
+The set of instructions, \(I\), contains the RISC-V ISA. We further introduce a set
+a set \(J\) of {\it overlay sequences}: sequences of instructions that operate as if
+they were single, atomic pseudoinstructions in a larger machine.
+
+execute atomically
+via a special {\it overlay step} and which carry additional source-level information
+that designates their purpose, and which therefore comes with security obligations.
+An overlay sequence has two parts, a {\it pseudo-instruction} drawn from those
+
+There is additionally a set
 of macro-instructions \(J\), described in detail in Figure \ref{fig:macros}.

### 2022-08-24 submission5/onlyformalism.tex
Emergency push as laptop makes nasty fan noises

@@
 a set \(J\) of {\it overlay sequences}: sequences of instructions that operate as if
-they were single, atomic pseudoinstructions in a larger machine.
-
-execute atomically
-via a special {\it overlay step} and which carry additional source-level information
-that designates their purpose, and which therefore comes with security obligations.
-An overlay sequence has two parts, a {\it pseudo-instruction} drawn from those
-
-There is additionally a set
-of macro-instructions \(J\), described in detail in Figure \ref{fig:macros}.
-Every macro-instruction \(j \in J\) has a
-size \(\mathit{size}(j)\) and can be decomposed into a sequence of fragments
-\(j_0 \dots j_{\mathit{size}(J)}\), where \(j_n \in J^*\).
-The partial function \(\mathit{decode} : \WORDS \rightharpoonup I + J^*\) may decode a value
-into either an instruction or a macro-instruction fragment.
+they were single, atomic pseudoinstructions. Overlay sequences are {\it calls},
+{\it returns}, {\it tailcalls}, exception {\it catches} and {\it throws},
+and object {\it allocs}, and each carries a sequence of instructions
+that implements the operation. When
 
@@
     & \(tr\) & Type record \\
-    & \(pubs\) & Public set \\
     \hline
@@
     & \(tr\) & Type record \\
-    & \(n\) & Size of locals \\
     \hline

### 2022-08-25 submission5/onlyformalism.tex
Through multiprocessor formalism

@@
 Words are 64-bit integers ranged over by \(w\) and, when used as addresses,
-\(a\). There are 32 registers, ranged over in general by \(r\),
+\(a\). Each processor has 32 registers, ranged over in general by \(r\),
 plus a program counter (\(\PCname\)). The following table divides
@@
 
+We specify that a register belongs to processor \(\rho \in \mathcal{R}\) by writing
+it \(r^\rho\), but will omit this notation when the active
+processor is unambiguous from context.
+
 Collectively addresses and registers are {\em state elements} \(\component\)
@@
 are fomulae with free variables bound to sets in \(C\).
+The base machine step relation 
+\(\mach \harpoonunder{\rho}_M \mach' \subseteq \MACHS \times \mathcal{R} \times \MACHS\)
+represents a single step of the processor \(\rho\).
 
@@
 We represent a {\it policy} via the step relation
-\(\cdot, \cdot \rightharpoonup_P \cdot \subseteq \MACHS \times \POLS \times \POLS\).
+\(\mach, \pol \harpoonunder{\rho}_P \pol' \subseteq \MACHS \times \POLS \times \mathcal{R} \times \POLS\).
 Steps are deterministic and  undefined represents a {\it failstop}.
@@
 {\it security classes}:
-\[sc \in SEC ::= \sealed | \passed | \mathit{public} | \unsealed\]
+\[sc \in SEC ::= \sealed | \mathit{object}(\phi) | \unsealed\]
 And of {\it targets}, which identify the addresses of the program counter or
@@
 elements in terms of their purpose relative to the active function.
-And a security context is a stack of pairs of targets and views:
+And a {\it stack context} is a stack of pairs of targets and views:
 \[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]
 
-The set of instructions, \(I\), contains the RISC-V ISA. We further introduce a set
-a set \(J\) of {\it overlay sequences}: sequences of instructions that operate as if
-they were single, atomic pseudoinstructions. Overlay sequences are {\it calls},
-{\it returns}, {\it tailcalls}, exception {\it catches} and {\it throws},
-and object {\it allocs}, and each carries a sequence of instructions
-that implements the operation. When
+Our properties additionally depend on a rudimentary model of memory-safe
+objects and pointer provenance. A provenance \(\phi \in \mathbb{N}\)
+identifies a value derived from the \(\phi\)th object allocated since
+the beginning of execution. A provenance map
+\(\prov \in \PROVS ::= \COMPONENTS \rightarrow 2^{\mathbb{N}}\)
+designates, for every machine element, which provenances its current
+value contains. We define a step relation on provenance maps
+\(\mach, \prov \harpoonunder{\rho}_{pm} \prov' \subseteq \MACHS \times
+\PROVS \times \mathcal{R} \times \PROVS\),
+exerpted in Figure \ref{fig:provenance} [TODO].
 
-\begin{figure}
-  \begin{tabular}{| l | l l |}
-    \hline
-    Macro & \multicolumn{2}{|l|}{Parameters} \\
-    \hline
-    \multirow{3}{*}{\(\mathbf{call}\)}
-    & \(\reg\) & Register containing destination \\
-    & \(tr\) & Type record \\
-    \hline
-    \multirow{3}{*}{\(\mathbf{tailcall}\)}
-    & \(\reg\) & Register containing destination \\
-    & \(tr\) & Type record \\
-    \hline
-    \multirow{2}{*}{\(\mathbf{enter}\)}
-    & \(tr\) & Type record \\
-    \hline
-    \(\mathbf{return}\)
-    & \(pubs\) & Public set \\
-    \hline
-    \(\mathbf{yield}\) & & \\
-    \(\mathbf{setex}\) & & \\
-    \(\mathbf{throwex}\) & & \\
-    \hline
-  \end{tabular}
-  \caption{Macro-instructions}
-  \label{fig:macros}
-\end{figure}
+The set of instructions, \(I \subseteq \WORDS\), corresponds to the RISC-V ISA.
+We further introduce {\it overlay sequences}: sequences of
+instructions that operate as if they were single, atomic pseudoinstructions.
+An overlay sequence consists of both a list
+of instructions and a pseudoinstruction that identifies its purpose in the
+property. The pseudoinstructions, and the data that they carry, are as follows.
+[TODO: still handle stack args by offset.]
 
-Each of these types of state is combined in to a triple called an
-MPC-state, \(\mpcstatename \in \MPCS ::= \MACHS \times \POLS \times \CONTEXTS\).
-Execution is modeled by a partial step function between MPC-states,
-written \(\mach, \pol, \context \harpoonunder{} \mach', \pol', \context'\),
-optionally annotated with a macro-instruction \(\harpoonunder{j}\).
-If \(\mach[\PCname]\) decodes to an instruction, then execution follows
-the RISC-V semantics and the policy, without touching the security context.
+\[\begin{align*}
+\psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+| & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+| & \mathbf{return} & \\
+| & \mathbf{alloc} ~ size ~ public & size \in \mathbb{N}, public \in \mathbb{B} \\
+| & \mathbf{setex} \\
+| & \mathbf{throwex} \\
+\end{align*}\]
 
-\judgmentthree[]
-              {\(\mathit{decode} ~ \mach ~ \mach[\PCname] \in I\)}
-              {\(\mach \stepstounderfull \mach'\)}
-              {\(\mach,\pol \rightharpoonup_P \pol'\)}
-              {\(\mach,\pol,\context \harpoonunder{} \mach',\pol',\context\)}
+An overlay sequence \(j \in J\) is a pair of a pseudoinstruction and
+a non-empty list of instructions.
 
-If, for some \(j\), \(\mach[\PCname + i]\) decodes to \(j_i\) for all \(i\) between 0 and
-\(\mathit{size}(j)\), written \(\mathit{decode*} = j\), then the step is via a special
-multi-step. We present these steps as separate step relations:
-\(\cdot \harpoonunder{\cdot}_M \cdot \in \MACHS \times J \times \MACHS\) to step
-the machine state,
-\(\cdot, \cdot \harpoonunder{\cdot}_P \cdot \in \MACHS \times \POLS \times J \times \POLS\)
-for the policy state, and
- \(\cdot, \cdot \stepstounder{j}_C \cdot \in \MACHS \times \CONTEXTS \times J \times \CONTEXTS\)
-for the security context. We then combine them:
+We define a step relation over triples of a machine state,
+a policy state, and a provenance map:
 
-\judgmenttwobr[]
-               {\(\mathit{decode*} = j\)}
-               {\(\mach \harpoonunder{j}_M \mach'\)}
-               {\(\mach,\pol \rightharpoonup_P \pol'\)}
-               {\(\mach',\context \stepstounder{j}_C \context'\)}
-               {\(\mach,\pol,\context \harpoonunder{j} \mach',\pol',\context'\)}
+\judgmentthree{\(\mach \harpoonunder{\rho}_M \mach'\)}
+              {\(\mach, \pol \harpoonunder{\rho}_P \pol'\)}
+              {\(\mach, \prov \harpoonunder{\rho}_{pm} \prov'\)}
+              {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
 
-In this section we will define \(\harpoonunder{j}_M\) for our machine; \(\harpoonunder{j}_P\)
-for the relevant policies will be discussed in [enforcement section], and \(\stepstounder{j}_C\)
-in Section \ref{sec:properties}.
+And we define a {\it ranged-step} relation that, given a pair of addresses
+\(a_{low}\) and \(a_{high}\), steps as long the program counter is between them.
 
-\paragraph*{Macro-step machine steps}
+\judgment[Done]
+         {\(\mach[\PCname] < a_{low} \lor \mach[\PCname] > a_{high}\)}
+         {\(a_{low},a_{high} \vdash (\mach,\pol,\prov) \manystepstounder{\rho} (\mach,\pol,\prov)\)}
 
-Here we give the semantics of the machine side of the macro step for calls and returns.
-The machine multi-step for a call (\ref{fig:callmacro}) updates the stack pointer
-and program counter, and zeros non-argument, caller-save registers. (This means that
-the caller should have saved them before initiating the call.)
+\judgmenttwobrlong[Step]
+                  {\(a_{low} \leq \mach[\PCname] < a_{high}\)}
+                  {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
+                  {\(a_{low},a_{high} \vdash (\mach',\pol',\prov') \manystepstounder{\rho} (\mach'',\pol'',\prov'')\)}
+                  {\(a_{low},a_{high} \vdash (\mach,\pol,\prov) \manystepstounder{\rho} (\mach'',\pol'',\prov'')\)}
 
-\vspace{\abovedisplayskip}
-\judgmentbrbr[CallM]
-             {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathbf{call} ~ r ~ tr ~ pubs\)}
-             {\(m' = m[rra \mapsto m[pc] + 4]\)}
-             {\(m'' = m' \llbracket r \mapsto 0 | r \in CLR\_SV - tr.args \rrbracket\)}
-             {\(m \harpoonunder{j}_M m''[pc \mapsto m[r]]\)}
-\vspace{\belowdisplayskip}
+Now we're finally ready to introduce the overlay semantics.
+A {\it full state} \(\sigma\) consists of a 4-tuple of a machine state,
+a policy state, a provenance map, and a stack context:
+\[\sigma \in \mathcal{S} ::=
+\MACHS \times \POLS \times \PROVS \times \CONTEXTS\]
 
-Then, on enter, the callee stores the stack pointer and callee-save registers,
-and allocates new stack space.
+The execution relation for full states
+\(\sigma \stepstounder{\rho;\psi?} \sigma' \subseteq
+\mathcal{S} \times \mathcal{R} \times (\Psi + \bot) \times \mathcal{S}\)
+proceeds by an {\it overlay step}
+if the program counter points to the start of an overlay sequence.
+Otherwise, it proceeds by the default step. The definitions of context steps
+\(\harpoonunder{\rho;\psi}_C\) will be defined as we go.
 
-\vspace{\abovedisplayskip}
-\judgmentbrbrbr[EnterM]
-               {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathit{enter} ~ tr\)}
-               {\(m' = m[m[sp] \mapsto m[rra]]\)}
-               {\(m'' = m\llbracket m[m[sp]+i+1 \mapsto m[rs_i] | 0 \leq i < 8 \rrbracket\)}
-               {\(m''' = m''[sp \mapsto m[sp]+n+9]]\)}
-               {\(m \harpoonunder{j}_M m'''[pc \mapsto m[pc]+4*\mathit{size}(j)]\)}
-\vspace{\belowdisplayskip}
+\judgmenttwobrlongbrlong[Overlay]
+                        {\(j = (\psi, \bar{i})\)}
+                        {\(\mach',\prov',\context \harpoonunder{\rho;\psi}_C \context'\)}
+                        {\((\mach,\pol,\prov) \manystepstounder{\rho} (\mach',\pol',\prov')\)}
+                        {\(\mach[\PCname+n] = \bar{i}[n]\) for each \(n\) where \(0 \leq n < |\bar{i}|\)}
+                        {\((\mach,\pol,\prov,\context) \stepstounder{\rho;j} (\mach',\pol',\prov',\context')\)}
 
-Finally, on return, the program counter and register contents are restored from the stack,
-and the callee's frame is deallocated.
+\judgment[Default]
+         {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
+         {\((\mach,\pol,\prov,\context) \stepstounder{\rho; \bot} (\mach',\pol',\prov',\context)\)}
 
-\vspace{\abovedisplayskip}
-\judgmentbrbr[ReturnM]
-             {\(j = \mathit{decode}^* ~ m ~ m[pc] = \mathit{return} ~ n\)}
-             {\(m' = m[sp \mapsto m[sp]-n-9]\)}
-             {\(m'' = m\llbracket m[rs_i] \mapsto m[m[sp]+i+1 | 0 \leq i < 8 \rrbracket\)}
-             {\(m \harpoonunder{j}_M m''[pc \mapsto m''[sp]]\)}
-\vspace{\belowdisplayskip}
+\paragraph*{Concurrency}
 
-%For the context step, the macro-instruction additionally identifies which objects
-%(contiguous regions defined relative to the stack pointer) have been placed there
-%as: arguments that overflowed to the stack (\(sargs\)), which are existing objects
-%passed by reference (\(pbrs\)), and which are marked public (\(pubs\)). We use this
-%information to construct the callee's security view.
-%\[obj \in OBJ ::= \mathbb{Z} \times \mathbb{Z}\]
-%\[in\_obj ~ a_{sp} ~ (lo, hi) ~ a \triangleq a_{sp} - lo \leq a < a_{sp} - hi\]
+Thus far, we have defined the step relation of a single processor.
+We model multiple processors able to step independently by a global step relation,
+\(\stepsto\), which simply steps an arbitrary processor:
 
-%\judgmenttwobrlong[Return]
-%                    {\(m' = m[sp \mapsto m[sp] - \mathit{size}(frsize)]\)}
-%                    {\(m'' = m \llbracket r \mapsto 0 | r \not \in args \land r \in caller \rrbracket\)}
-%                    {\(\mathit{decode}^* ~ m ~ pc = \mathit{return} ~ frsize\)}
-%                    {\(m \stepsto_M m'[pc \mapsto f]\)}
-
-%If \(\mathit{decode}\) is undefined on that value, \(\mach \stepsto \mach\). The
-%only times \(\stepsto\) is undefined is when there is a policy fault,
-%(and the machine ``failstops'') or when it attempts to begin executing a
-%macro-instruction out of order. The latter represents a failstop in a lower-level
-%policy in the underlying ISA, which guarantees that the macro-instruction
-%executes in order.
-
-\paragraph*{Compiling Macro-instructions}
-
-Every macro-instruction is associated with a sequence of normal RISC-V instructions
-of the appropriate length, so compiling from the macro-instruction machine to normal
-RISC-V is as simple as replacing \(j_i\) with its corresponding instruction, tagged
-as a ``blessed sequence'' -- a standard tag-based enforcement mechanism that guarantees
-that instructions may {\it only} step in the appropriate order.
+\judgment[]
+         {\(\sigma \stepstounder{\rho} \sigma'\)}
+         {\(\sigma \stepsto \sigma'\)}
 

### 2022-08-26 submission5/onlyformalism.tex
Did a pass of formalism with concurrency and provenance

@@
 
-\[\begin{align*}
+\begin{align*}
 \psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
@@
 | & \mathbf{throwex} \\
-\end{align*}\]
+\end{align*}
 
@@
 a policy state, a provenance map, and a stack context:
-\[\sigma \in \mathcal{S} ::=
-\MACHS \times \POLS \times \PROVS \times \CONTEXTS\]
+\[\sigma = (\mach, \pol, \prov, ctxs) \in \mathcal{S} ::=
+\MACHS \times \POLS \times \PROVS \times (\mathcal{R} \rightarrow \CONTEXTS)\]
 
@@
 Otherwise, it proceeds by the default step. The definitions of context steps
-\(\harpoonunder{\rho;\psi}_C\) will be defined as we go.
+\(\harpoonunder{\rho;\psi}_C\) will be defined for each pseudoinstruction as we go.
 
@@
                         {\(j = (\psi, \bar{i})\)}
-                        {\(\mach',\prov',\context \harpoonunder{\rho;\psi}_C \context'\)}
+                        {\(\mach',\prov',ctxs ~ \rho \harpoonunder{\rho;\psi}_C \context\)}
                         {\((\mach,\pol,\prov) \manystepstounder{\rho} (\mach',\pol',\prov')\)}
                         {\(\mach[\PCname+n] = \bar{i}[n]\) for each \(n\) where \(0 \leq n < |\bar{i}|\)}
-                        {\((\mach,\pol,\prov,\context) \stepstounder{\rho;j} (\mach',\pol',\prov',\context')\)}
+                        {\((\mach,\pol,\prov,ctxs) \stepstounder{\rho;\psi} (\mach',\pol',\prov',ctxs[\rho \mapsto \context])\)}
 
@@
          {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
-         {\((\mach,\pol,\prov,\context) \stepstounder{\rho; \bot} (\mach',\pol',\prov',\context)\)}
+         {\((\mach,\pol,\prov,ctxs) \stepstounder{\rho; \bot} (\mach',\pol',\prov',ctxs)\)}
+
+\paragraph*{Overlay sequence definitions}
+
+[TBD]
+
+Note that there can be multiple overlay sequences with different instructions
+that represent the same pseudoinstruction.
+[This is less likely when we have smaller pseudoinstructions.]
 
@@
 We model multiple processors able to step independently by a global step relation,
-\(\stepsto\), which simply steps an arbitrary processor:
+\(\stepstounder{}\), which simply steps an arbitrary processor:
 
@@
          {\(\sigma \stepstounder{\rho} \sigma'\)}
-         {\(\sigma \stepsto \sigma'\)}
+         {\(\sigma \stepstounder{} \sigma'\)}
 
-\paragraph*{Machine and Observation Traces}
-\label{sec:traces}
+\section{Stack Safety With Calls and Returns}
 
-A {\em machine trace} is a nonempty, finite or infinite sequence
-of states, ranged over by \(\MPCT\) and \(\NPCT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
-for list-cons. An {\em event trace} \(\obsT\) is a similar sequence of
-events \(\obs \in \OBSS\), which are defined as follows:
+In this section we will present the context steps for allocations,
+calls, and returns, and show how they
+translate into properties. They are given formally in Figure \ref{fig:callcontexts}.
+Allocations are simple: from the current stack pointer, a range up to the size of
+the allocation is marked \(\mathit{object}(\phi)\) for a fresh identifier \(\phi\).
 
+First, we define a function, {\it update view}, that defines the view of a callee
+in terms of that of the callee, together with the provenance map.
+First, all stack locations below the stack pointer
+are \(\unsealed\) -- fair game for the callee to allocate or write to, and
+to read after writing. Above the stack pointer, anything that is not an object
+becomes \(\sealed\). Argument and system registers are automatically marked
+\(\mathit{public}\), and all others are \(\sealed\).
+Finally, we must address stack objects: we recursively identify all objects
+that are reachable via pointer provenances found in \(\mathit{public}\)
+registers and in reachable objects, and these retain their view, while
+all others become \(\sealed\). [TODO: implement \(\mathit{reach}\) to do this.]
+
+To update the context, we pair the updated view with the return target of the
+call: the return target tells us whether a return is to the proper location
+which the stack pointer restored, and the view tells us how to treat the
+security of the callee's state. We define \(\mathit{hd\_def}\) as taking
+a context and returning its head, or \(((0,0),\lambda \component.\unsealed)\)
+if the context is empty. [TODO: structure it so that we don't need to return
+  a dummy target.]
+
+\begin{figure*}
 \[\begin{split}
-\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{dst} ~ \overline{w_{args}} \\
+& \mathit{update\_view} ~ V ~ \prov ~ a_{sp} ~ \overline{r_{args}} \triangleq
+\lambda \component .
+\begin{cases}
+  \mathit{public} & \textnormal{else if } \component \in \REGS \textnormal{ and }
+  \component \in \overline{r_{args}} \textnormal{ or } \component \in \mathit{PUBLIC} \\
+  \mathit{object}(\phi) & \textnormal{else if } (V ~ \component) = \mathit{object}(\phi)
+  \textnormal{ and } \phi \in \mathit{reach} ~ \prov ~ V ~ \overline{r_{args}} \\
+  \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } a_{sp} < \component \leq a_{sbase} \\
+  \unsealed & \textnormal{otherwise} \\
+\end{cases} \\
+\end{split}\]
+
+\judgmenttwobrlong[CallC]
+                  {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}}\)}
+                  {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V)::\context\)}
+                  {\(V = \mathit{update\_view} ~ (snd ~ (hd\_def ~ \context)) ~ m[sp] ~ \overline{r_{args}}\)}
+                  {\(\mach, \prov, \context \constep{\rho;\psi} \context'\)}
+
+\judgment[ReturnCCons]
+         {\(j = \mathbf{return}\)}
+         {\(m, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\rho;\psi} \context\)}
+
+\judgment[ReturnCNil]
+         {\(j = \mathbf{return}\)}
+         {\(m, [] \constep{\rho;\psi} []\)}
+
+\caption{Call, enter, and return macro-instructions}
+\label{fig:callcontexts}
+\end{figure*}
+
+\section{Stack Safety, Formally}
+\label{sec:properties}
+
+Now we can give our definition of stack safety. Each overlay sequence defines a
+context step; in turn, the context at the point of a call or return defines a
+property each for integrity, confidentiality, and well-bracketed control flow.
+
+\subsection{Facts Abouts Calls and Returns}
+
+Here we define some logical operations to reason about the behavior of the
+system over time. These have a temporal-logic flavor, as they reflect
+the expected behavior of the system in the future, after a possible return.
+
+\paragraph*{Event Traces}
+
+Some steps of execution can be interpreted as {\it events} that can be observed
+from outside the system. We define events \(\obs \in \OBSS\) as calls, labeled
+with the target address and the values of the arguments, or else as the silent
+event \(\tau\):
+\[\begin{split}
+\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
 | & ~ \tau \\
@@
 
-%This system of observations resembles that used by CompCert [cite], in that
-%the observable behaviors of the system are calls and returns (in our case, not
-%distinguishing between internal and external, and including exceptions), as well
-%as a special output akin to storing to volatile memory. We add the silent observation
-%\(\tau\).
+A {\em trace} is a nonempty, finite or infinite sequence
+of pairs of processors and events, ranged over by \(\obsT\).
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
+for list-cons.
 
@@
 \begin{figure*}
-\[\mathit{observe} ~ j ~ m =
+\[\mathit{observe} ~ \rho ~ \psi? ~ m =
 \begin{cases}
-  \mathbf{callE} ~ m[r] ~ m \llbracket k | k \in tr.args \lor m[rsp] - k \in tr.sargs \rrbracket
-  & \textnormal{when } ~ j = \mathbf{call} ~ r ~ tr ~ pubs \\
-  \tau & \textnormal{otherwise} \\
+  (\rho, \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket)
+  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
+  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
+  (\rho, \tau) & \textnormal{otherwise} \\
 \end{cases}\]
+\caption{Converting overlay steps to events}
 \label{fig:observe}
@@
 
-The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT \),
-coinductively relates an initial state with the trace of states
-produced by repeated application of \(\rightharpoonup\):
+We write that execution from a state produces an observation trace \(\sigma \hookrightarrow \obsT\)
+as follows, coinductively:
 
-\judgmenttwo{\(\mpcstatename \rightharpoonup \mpcstatename'\)}
-            {\(\mpcstatename' \hookrightarrow \MPCT \)}
-            {\(\mpcstatename \hookrightarrow \notfinished{\mpcstatename'}{\MPCT} \)}%
-\judgment%[Default]
-         {\(\not\exists \mpcstatename'. \mpcstatename \stepstocon \mpcstatename'\)}
-         {\(\mpcstatename \hookrightarrow \mpcstatename\)}
+\judgment{\(\not \exists \sigma' . \sigma \stepstounder{} \sigma'\)}
+         {\(\sigma \hookrightarrow \tau\)}
 
-We project out the machine state of an MC-state with \(\pi_\mach\)
-and the context with \(\pi_\context\).
-We extract the first entry from a trace with \(\head(\MPCT)\), which is a total
-function since traces are non-empty, and the final entry (if one exists) with
-\(\last(\MPCT)\), which is partial, since traces may be infinite.
+\judgmenttwo{\(\sigma \stepstounder{\rho; \psi?} \sigma'\)}
+            {\(\sigma' \hookrightarrow \obsT\)}
+            {\(\sigma \hookrightarrow \notfinished{(\mathit{observe} ~ \rho ~ \psi? (\pi_m{\sigma'}))}{\obsT}\)}
 
-\paragraph*{Observations-Of}
+\paragraph*{Trace Until Return}
 
-This operator, written \(\obsof\), takes a machine trace and
-gives the trace of events of each step in
-the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
-initial event, a machine-trace, and an event trace; the helper is
-then invoked with an initial \(\tau\).
+We define another relation that takes a trace of a callee up to its return, if one occurs.
+We write this \(\context \downarrow \sigma \hookrightarrow \obsT\), where \(\context\)
+is the context of the caller.
 
-\judgmenttwobrlong{\(\mpcstatename \harpoonunder{j} \mpcstatename'\)}
-                  {\(\mpcstatename' \hookrightarrow \MPCT\)}
-                  {\(\mathit{obsfor} ~ (\mathit{observe} ~ \pi_m(\mpcstatename) ~ j) ~ \MPCT = \obsT\)}
-                  {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}
+\judgment{\(\context = \pi_c(\sigma)\)}
+         {\(\context \downarrow \sigma \hookrightarrow \tau\)}
 
-\begin{minipage}[b]{.4\columnwidth}
-  \judgment{}
-           {\(\mathit{obsfor} ~ \obs ~ \mpcstatename = \obs\)}
-\end{minipage}
-\begin{minipage}[b]{.4\columnwidth}
-\judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
-         {\(\obsof(\MPCT) = \obsT\)}
-\end{minipage}
+\judgmentthree{\(\sigma \stepstounder{\rho,\psi?} \sigma'\)}
+              {\(\context \not = \pi_c(\sigma')\)}
+              {\(\context \downarrow \sigma' \hookrightarrow \obsT\)}
+              {\(\context \downarrow \sigma \hookrightarrow \notfinished{(\mathit{observe} ~ \rho ~ \psi? (\pi_m{\sigma'}))}{\obsT}\)}
 
@@
 
-Note that an infinite silent trace is a
-prefix of (and similar to) any other trace. While this might seem
-surprising at first, it makes sense in a timing-insensitive context:
-an external observer looking at two machine runs cannot (computably)
-distinguish between a machine that steps forever and a machine that
-steps for a long time before producing some output.
+\paragraph*{On-return}
 
-\paragraph*{Reachability}
+The intuition behind {\it return-time integrity} (below) is that a caller may expect its
+sealed data to be unchanged when control returns to it. In fact, the callee
+may overwrite such data -- when the data are found in callee-saved registers
+this is perfectly legal -- as long as it either restores it, or has some guarantee
+that its changes will impact the caller.
 
-We often want to quantify over all traces from any initial state so that our
-properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
-is {\em reachable} if, for some initial state \(\mpcstatename = \mach,\pol,[]\)
-where \(\mach\) is well-formed, \(\mpcstatename \hookrightarrow \MPCT\)
-and \(\MPCT\) contains \(\mpcstatename'\).
+We start by defining a second-order logical operator
+\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
+where \(P\) is a predicate on machine states. This is a coinductive relation
+similar to ``weak until'' in temporal logic -- it holds if the condition
+is never reached.
 
-\section{Stack Safety, Formally}
-\label{sec:properties}
+\judgment{\(P ~ \mach\)}
+         {\((\context \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
 
-Now we can give our definition of stack safety. Each macro-instruction defines a
-context step; in turn, the context at the point of a call or return defines a trace
-property each for integrity, confidentiality, and well-bracketed control flow.
-The context updates involved in calls and returns are given in Figure \ref{fig:callcontexts}.
-First, we define a function, {\it update view}, that defines the view of a callee.
-Addresses that, offset from stack pointer, are in public regions are marked
-\(\mathit{public}\). Those that are being passed on the stack are marked
-\(\mathit{passed}\). [When I add pass-by-reference arguments, we will use
-the recursive \(V\)]
+\judgmenttwobrlong{\(\context_0 \not = \context\)}
+                  {\((\context_0 \uparrow P) ~ (\mach', \pol', \prov', \context')\)}
+                  {\((\mach, \pol, \prov, \context) \stepstounder{\rho,\psi?}
+                    (\mach', \pol', \prov', \context')\)}
+                  {\((\context_0 \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
 
-To update the context, we pair the updated view with the return target of the
-call: the return target tells us whether a return is to the proper location
-which the stack pointer restored, and the view tells us how to treat the
-security of the callee's state.
+Similarly, in {\it caller confidentiality}, we will want to compare future states,
+so we give a binary equivalent, \(\context \Uparrow R\), so that
+\(\sigma (\context \Uparrow R) \sigma'\) holds if \(R\) holds on the
+first return states of after \(\sigma\) and \(\sigma'\).
 
-\begin{figure*}
-\[\begin{split}
-& \mathit{update\_view} ~ V ~ a_{sp} ~ tr ~ pubs \triangleq
- \lambda \component .
-\begin{cases}
-  \mathit{public} & \textnormal{if } \component \in \WORDS \textnormal{ and }
-  (a_{sp} - \component) \in pubs \\
-%  \passed & \textnormal{else if } \component \in \WORDS \textnormal{ and }
-%  in\_obj ~ a_{sp} ~ pbr ~ \component \textnormal{ for some } pbr \textnormal{ in } pbrs \\
-  \passed & \textnormal{else if } \component \in \WORDS \textnormal{ and }
-  (a_{sp} - \component) \in tr.sargs \\
-  \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } \component < a_{sp} \\
-  \mathit{public} & \textnormal{else if } \component \in \REGS \textnormal{ and }
-  \component \in tr.args \textnormal{ or } \component \in \mathit{PUBLIC} \\
-  \unsealed & \textnormal{otherwise} \\
-\end{cases} \\
-\end{split}\]
+\judgment[Returned]
+         {\(\mach ~ R ~ \mach'\)}
+         {\((\mach,\pol,\prov,\context) ~ (\context \Uparrow R) ~ (\mach',\pol',\prov',\context)\)}
 
-\judgmenttwobrlong[CallCCons]
-           {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ tr ~ pubs\)}
-           {\(j = \mathbf{call} ~ r ~ tr\)}
-           {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V')::(tar,V)::\context\)}
-           {\(m, (tar,V)::\context \stepstounder{j}_C \context'\)}
+\judgmentthree[Left]
+              {\(\context \not = \pi_c(\sigma_1)\)}
+              {\(\sigma_1 \stepstounder{\rho;\psi?} \sigma_1'\)}
+              {\(\sigma_1' ~ (\context \Uparrow R) ~ \sigma_2\)}     
+              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
 
-\judgmenttwobrlong[CallCNil]
-           {\(V' = \mathit{update\_view} ~ (\lambda \component . \unsealed) ~ m[sp] ~ tr ~ pubs\)}
-           {\(j = \mathbf{call} ~ r ~ tr ~ pubs\)}
-           {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V')::[]\)}
-           {\(m, [] \stepstounder{j}_C \context'\)}
+\judgmentthree[Right]
+              {\(\context \not = \pi_c(\sigma_2)\)}
+              {\(\sigma_2 \stepstounder{\rho;\psi?} \sigma_2'\)}
+              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2'\)}     
+              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
 
-\judgment[EnterC]
-         {\(j = \mathbf{enter} ~ tr ~ n\)}
-         {\(m,c \stepstounder{j}_M c\)}
+\paragraph*{Variants and Safe Elements}
 
-\judgmentthree[ReturnCValid]
-              {\(m[pc] = a_{pc}\)}
-              {\(m[rsp] = a_{sp}\)}
-              {\(j = \mathbf{return} ~ pubs\)}
-              {\(m, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \stepstounder{j}_C \context\)}
-
-\judgmenttwobrlong[ReturnCInvalid]
-                  {\(m[pc] \not = a_{pc} \lor m[rsp] \not = a_{sp}\)}
-                  {\(j = \mathbf{return} ~ pubs\)}
-                  {\(V' = V \llbracket r \mapsto \unsealed | r \in \{pc, rsp\} \rrbracket\)}
-                  {\(m, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \stepstounder{j}_C (\mathit{return} ~ a_{pc} ~ a_{sp}, V')::\context\)}
-
-\judgmenttwo[ReturnCNil]
-            {\(V' = \lambda \component . \unsealed\)}
-            {\(j = \mathbf{return} ~ pubs\)}
-            {\(m, [] \stepstounder{j}_C []\)}
-
-\caption{Call, enter, and return macro-instructions}
-\label{fig:callcontexts}
-\end{figure*}
-
-\paragraph*{Eventual Integrity}
-
-The intuition behind eventual integrity is that, when control is passed from the
-caller to the callee, the caller doesn't necessarily need its data to remain intact --
-as long as they are restored by the time control returns to it. We define this as a more
-general notion of {\it eventual integrity.}
-
-Let \(R \subseteq \CONTEXTS\) be a predicate on contexts, representing the condition at
-which integrity must be restored, and \(I\) be a predicate on states defining what integrity
-means. Then we find the first state on which \(R\) holds on the context and assert that
-\(I\) also holds on the machine state, with the \(\mathit{when}\) relation.
-
-\judgmenttwo{\(R ~ \context\)}
-            {\(I ~ \mach\)}
-            {\(\mathit{when} ~ R ~ I ~ (\mach,\pol,\context)\)}
-
-\judgmentthree{\(\neg R ~ \context\)}
-              {\((\mach,\pol,\context) \rightharpoonup \mpcstatename\)}
-              {\(\mathit{when} ~ R ~ I ~ \mpcstatename\)}
-              {\(\mathit{when} ~ R ~ I ~ \mpcstatename\)}
-
-\definition Let \(\components\) be a set of elements and \((\mach,\pol,\context)\)
-be the first state of a new call. We define \(R\) to hold on those states
-\((\mach',\pol',\context')\) where \(|\context'| < |\context|\). And we define
-\(I\) to hold when \(\mach[\component] = \mach'[\component]\) for all elements
-\(\component \in \COMPONENTS\). Then \((\mach,\pol,\context)\) enjoys eventual
-integrity of \(\components\) if \(\mathit{when} ~ R ~ I ~ (\mach,\pol,\context)\).
-
-Howver, eventual integrity is not sufficient to express the integrity protected by
-Lazy Tagging, because it is possible that values in memory are overwritten -- but
-are also guaranteed never to be accessible. We need to define a slightly weaker property.
-
-\paragraph*{Variants and Transient Elements}
-
-One or more elements of a given state are {\it transient} if their values
+One or more elements of a given state are {\it safe} if their values
 in that state have no observable influence on future execution. We define this
-formally as a sort of non-interference: for a state \(\mpcstatename\) and
-a set of elements \(\components\), \(\mathit{transient} ~ \mpcstatename ~ \components\)
+formally as a sort of non-interference: for a state \(\sigma\) and
+a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
 holds if we can replace the contents of \(\components\) with arbitrary values
@@
 
-\definition An elements set \(\components\) is transient in state \((\mach,\pol,\context)\),
-written \(\mathit{transient} ~ (\mach,\pol,\context) ~ \components\), if for all
-\(\nach\) such that \(\mach \approx_{\components'} \nach\), if 
-\((\mach,\pol,\context) \hookrightarrow \MPCT\) and
-\((\nach,\pol,\context) \hookrightarrow \NPCT\), then
-\(\obsof(\MPCT) \eqsim \obsof(\NPCT)\).
+\definition Full states \((\mach,\pol,\prov,\context)\) and \((\nach,\pol,\prov,\context)\)
+are also {\em \(\components\)-variants} if \(\mach\) and \(\nach\) are.
+
+\definition An element set \(\components\) is safe in state \(\sigma)\),
+written \(\components ~ \mathit{SAFE} ~ \sigma\), if for all
+\(\sigma'\) such that \(\sigma \approx_{\components'} \sigma'\), if 
+\(\sigma \hookrightarrow \obsT\) and
+\(\sigma' \hookrightarrow \obsT'\), then
+\(\obsT \eqsim \obsT'\).
 
@@
 
-\definition Let \(\components\) be a set of elements and \((\mach,\pol,\context)\)
-be the first state of a new call. We define \(R\) to hold on those states
-\((\mach',\pol',\context')\) where \(|\context'| < |\context|\). And we define
-\(I\) to hold on \(\mach'\) when \(\Delta(\mach,\mach') \cap \components\) is transient
-with respect to \(\mach'\). Then \((\mach,\pol,\context)\) enjoys
-{\it eventual intransient integrity} of \(\components\) if
-\(\mathit{when} ~ R ~ I ~ (\mach,\pol,\context)\).
+\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) and
+\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component = \sealed\}\).
+Let \(P\) be a predicate on states \(\sigma'\) that holds if
+\(\Delta(\mach, \pi_m(\sigma')) \cap \components ~ \mathit{SAFE} ~ \sigma'\).
+Then \(\sigma\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.
 
-\definition A system enjoys {\it observational stack integrity} if, for all
-reachable states \(\mpcstatename\) such that
-\(\mpcstatename \harpoonunder{\mathbf{Call}\dots} (\mach,\pol,(tar, V)::\context)\),
-\((\mach, \pol, (tar, V)::\context)\) enjoys eventual intransient integrity with
-respect to the set of components that \(V\) maps to \(\sealed\).
+\definition A system enjoys {\it stack integrity} if, for all reachable states \(\sigma\) such that
+\(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
+\(\sigma'\) enjoys return-time integrity.
 
-\paragraph*{Temporary Confidentiality}
+\paragraph*{Caller Confidentiality}
 
@@
 that it will be accessed later. So we must also consider the machine state at the end
-of the call. First we'll define a useful general property that implements this structure,
-similar to \(\mathit{when}\), called \(\mathit{during\_both}\).
-
-\(\mathit{during\_both}\) relates two initial states in their traces up until they
-both reach an end condition, again defines by a predicate \(R\). Instead of \(I\)
-we have an event-trace relation \(E\) that must relate the event traces up through
-\(R\), and a final relation \(F\) that must relate the final states.
-
-\definition Let \(\mpcstatename\) and \(\npcstatename\) be initial states
-and \(R\) be a predicate on states such that
-\(\mpcstatename \hookrightarrow \MPCT | R\) and
-\(\npcstatename \hookrightarrow \NPCT | R\).
-Let \(E\) be a relation on event traces, and \(F\) be a relation on machine states.
-Then \(\mathit{during\_both} ~ \mpcstatename ~ \npcstatename ~ R ~ E ~ F\) holds if both:
-\begin{itemize}
-\item \(E ~ (\obsof \MPCT) ~ (\obsof \NPCT)\) holds
-\item And, if \(\last ~ \MPCT = (\mach,\_,\_)\) and \(\last ~ \NPCT = (\nach,\_,\_)\),
-  then \(F ~ \mach ~ \nach\) holds.
-\end{itemize}
+of the call.
 
@@
 may be some conditions in which a secret is copied, but never used again. So, we once
-again wish for our final condition to make the values {\it transient}. Which values?
+again wish for our final condition to make the values {\it safe}. Which values?
 
 \definition Let \(\mach,\mach'\) and \(\nach,\nach'\)
-be pairs of states. Their {\em unsafe set}, written
+be pairs of machine states. Their {\em corrupted set}, written
 \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all elements
@@
 
-\definition Let \((\mach,\pol,\context)\) be a state and \(\components\) be a set of
-elements. Let \(R\) hold on contexts \(\context'\) such that \(|\context'| < |\context|\).
-Let \(E \triangleq ~ \simeq\).
-Let \(F ~ \mpcstatename ~ \npcstatename \triangleq
-\mathit{transient} ~ \mpcstatename ~
-\bar{\Diamond}(\mach,\pi_m(\mpcstatename),\nach,\pi_m(\npcstatename))\).
+\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) be a state and
+\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component \in
+\{\sealed, \unsealed\}\}\).
 
-\((\mach,\pol,\context)\) enjoys {\it temporary confidentiality} of \(\components\)
-if, for all \(\nach\) such that \(\mach \approx_{\components} \nach\),
-\(\mathit{during\_both} ~ R ~ E ~ F\).
+Then \(\sigma\) enjoys {\it caller confidentiality} if, for any \(\sigma'\)
+that is a \(\components\)-variant of \(\sigma\), we can take
+\(\context \uparrow \sigma \hookrightarrow \obsT\) and
+\(\context \uparrow \sigma' \hookrightarrow \obsT'\) and have that
+\(\obsT \simeq \obsT'\). And if, additionally, \(\sigma ~ (\context \Uparrow R) ~ \sigma'\),
+where \(\sigma_1 ~ R ~ \sigma_2\) if
+\(\bar{\Diamond}(\mach,\pi_m(\sigma_1),\pi_m(\sigma'),\pi_m(\sigma_2)) ~ \mathit{SAFE} ~ \sigma_1\).
 
-\definition A system enjoys {\it observational caller confidentiality} if, for all
-reachable states \(\mpcstatename\) such that
-\(\mpcstatename \harpoonunder{\mathbf{Call}\dots} (\mach,\pol,(tar, V)::\context)\),
-\((\mach, \pol, (tar, V)::\context)\) enjoys temporary confidentiality with
-respect to the set of components that \(V\) maps to \(\unsealed\) or \(\sealed\).
+\definition A system enjoys {\it universal caller confidentiality} if, for all reachable states
+\(\sigma\) such that \(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
+\(\sigma'\) enjoys caller confidentiality.
 

### 2022-08-26 submission5/onlyformalism.tex
Update

@@
 \item Exceptions
-\item Coroutines and cooperative threading
 \item Tail-call Elimination
@@
 as such by the compiler, by identifying them as macro-instructions.
-The same goes for other operations, such as exceptions and coroutine yields.
+The same goes for other operations, such as exceptions.
 Macro-instructions connect the enforcement mechanism to the property: we don't
@@
 carry data that can be known by the compiler at compile-time. For example,
-the \(\mathbf{call}\) macro-instruction identifies the register \(r\) containing
-the address of the function to be called, as well as a {\it type record} identifying
-the registers and addresses in which the callee expects to find arguments,
-and a {\it public set}.
-Let \(tr \in TR\) be a type record; then \(tr.args\) is a set of those registers
-that will be used for arguments, and \(tr.stack\_args\) defines a set of
-offsets from the stack pointer that are expected to contain additional
-arguments.The public set \(pubs\) defines a similar set of offsets in
-the caller that are declared public and not to be protected.
+the \(\mathbf{call}\) macro-instruction identifies the register \(r_{target}\) containing
+the address of the function to be called, as well as the registers that hold arguments.
 
@@
 we assume nothing about the intent of the source code, except that it makes calls
-and returns, and guarantee that the caller and callee are protected from one
+and returns, and we must guarantee that the caller and callee are protected from one
 another.
@@
 
-We do not support dynamic code generation or concurrency---we can handle a
-limited form of coroutines, but not full-blown preemptive concurrency.
-[Say something about Heap Safety]
+We do not support dynamic code generation, and our concurrency model is fairly
+simplistic, assuming a fixed number of threads each with its own dedicated processor.
+We model memory safe stack objects, but not a heap. Regions outside of
+stacks can be used however the compiler likes, including as a heap, but no protection is
+built in and our properties assume that if a pointer to a stack object is stored there,
+it is permanently compromised.
 
@@
 The building blocks of the machine are {\em words} and {\em registers}.
-Words are 64-bit integers ranged over by \(w\) and, when used as addresses,
+Words (\(\WORDS\)) are 64-bit integers ranged over by \(w\) and, when used as addresses,
 \(a\). Each processor has 32 registers, ranged over in general by \(r\),
@@
 
-We specify that a register belongs to processor \(\rho \in \mathcal{R}\) by writing
+We specify that a register belongs to processor \(\rho \in \PROCS\) by writing
 it \(r^\rho\), but will omit this notation when the active
-processor is unambiguous from context.
+processor is unambiguous from context. The set of all registers across all processors
+is \(\REGS\).
 
@@
 The base machine step relation 
-\(\mach \harpoonunder{\rho}_M \mach' \subseteq \MACHS \times \mathcal{R} \times \MACHS\)
+\(\mach \machstep{\rho} \mach' \subseteq \MACHS \times \PROCS \times \MACHS\)
 represents a single step of the processor \(\rho\).
@@
 We represent a {\it policy} via the step relation
-\(\mach, \pol \harpoonunder{\rho}_P \pol' \subseteq \MACHS \times \POLS \times \mathcal{R} \times \POLS\).
+\(\mach, \pol \harpoonunder{\rho}_P \pol' \subseteq \MACHS \times \POLS \times \PROCS \times \POLS\).
 Steps are deterministic and  undefined represents a {\it failstop}.
 
-Separate from the machine state, we keep track of additional security context
-information over the course of the run. This data is not relevant to execution,
+Our properties additionally depend on a rudimentary model of memory-safe
+objects and pointer provenance. A provenance \(\phi \in \mathbb{N}\)
+identifies a value derived from the \(\phi\)th object allocated since
+the beginning of execution. A provenance map
+\(\prov \in \PROVS ::= \COMPONENTS \rightarrow 2^{\mathbb{N}}\)
+designates, for every machine element, which provenances its current
+value contains. We define a step relation on provenance maps
+\(\mach, \prov \harpoonunder{\rho}_{pm} \prov' \subseteq \MACHS \times
+\PROVS \times \PROCS \times \PROVS\),
+exerpted in Figure \ref{fig:provenance} [TODO].
+
+We define a step relation over triples of a machine state,
+a policy state, and a provenance map:
+
+\judgmentthree{\(\mach \machstep{\rho} \mach'\)}
+              {\(\mach, \pol \polstep{\rho} \pol'\)}
+              {\(\mach, \prov \provstep{\rho} \prov'\)}
+              {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
+
+As the last piece of state, we keep track of security context information
+specific to each stack
+over the course of the run. This data is not relevant to execution,
 it is only used to define the stack safety property. We define a set of
 {\it security classes}:
-\[sc \in SEC ::= \sealed | \mathit{object}(\phi) | \unsealed\]
+\[sc \in SEC ::= \public | \sealed | \mathit{object}(\phi) | \unsealed\]
 And of {\it targets}, which identify the addresses of the program counter or
@@
 
-Our properties additionally depend on a rudimentary model of memory-safe
-objects and pointer provenance. A provenance \(\phi \in \mathbb{N}\)
-identifies a value derived from the \(\phi\)th object allocated since
-the beginning of execution. A provenance map
-\(\prov \in \PROVS ::= \COMPONENTS \rightarrow 2^{\mathbb{N}}\)
-designates, for every machine element, which provenances its current
-value contains. We define a step relation on provenance maps
-\(\mach, \prov \harpoonunder{\rho}_{pm} \prov' \subseteq \MACHS \times
-\PROVS \times \mathcal{R} \times \PROVS\),
-exerpted in Figure \ref{fig:provenance} [TODO].
-
 The set of instructions, \(I \subseteq \WORDS\), corresponds to the RISC-V ISA.
-We further introduce {\it overlay sequences}: sequences of
+We further introduce {\it macro-instructions}: sequences of
 instructions that operate as if they were single, atomic pseudoinstructions.
-An overlay sequence consists of both a list
+A macro-instruction consists of both a list
 of instructions and a pseudoinstruction that identifies its purpose in the
@@
 
-An overlay sequence \(j \in J\) is a pair of a pseudoinstruction and
-a non-empty list of instructions.
-
-We define a step relation over triples of a machine state,
-a policy state, and a provenance map:
-
-\judgmentthree{\(\mach \harpoonunder{\rho}_M \mach'\)}
-              {\(\mach, \pol \harpoonunder{\rho}_P \pol'\)}
-              {\(\mach, \prov \harpoonunder{\rho}_{pm} \prov'\)}
-              {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
-
-And we define a {\it ranged-step} relation that, given a pair of addresses
+An macro-instruction \((\bar{i},\psi) \in J\) is a pair of a pseudoinstruction and
+a non-empty list of instructions. In order to implement the atomic nature of
+executing macro-instructions,
+we define a {\it ranged-step} relation that, given a pair of addresses
 \(a_{low}\) and \(a_{high}\), steps as long the program counter is between them.
@@
 A {\it full state} \(\sigma\) consists of a 4-tuple of a machine state,
-a policy state, a provenance map, and a stack context:
+a policy state, a provenance map, and a map from processors to stack contexts:
 \[\sigma = (\mach, \pol, \prov, ctxs) \in \mathcal{S} ::=
-\MACHS \times \POLS \times \PROVS \times (\mathcal{R} \rightarrow \CONTEXTS)\]
+\MACHS \times \POLS \times \PROVS \times (\PROCS \rightarrow \CONTEXTS)\]
 
@@
 \(\sigma \stepstounder{\rho;\psi?} \sigma' \subseteq
-\mathcal{S} \times \mathcal{R} \times (\Psi + \bot) \times \mathcal{S}\)
+\mathcal{S} \times \PROCS \times (\Psi + \bot) \times \mathcal{S}\)
 proceeds by an {\it overlay step}
-if the program counter points to the start of an overlay sequence.
+if the program counter points to the start of a macro-instruction.
 Otherwise, it proceeds by the default step. The definitions of context steps
-\(\harpoonunder{\rho;\psi}_C\) will be defined for each pseudoinstruction as we go.
+\(\constep{\rho;\psi}\) will be defined for each pseudoinstruction as we introduce
+topics.
 
 \judgmenttwobrlongbrlong[Overlay]
-                        {\(j = (\psi, \bar{i})\)}
-                        {\(\mach',\prov',ctxs ~ \rho \harpoonunder{\rho;\psi}_C \context\)}
+                        {\((\bar{i}, \psi) \in J\)}
+                        {\(\mach',\prov',(ctxs ~ \rho) \constep{\rho;\psi} \context\)}
                         {\((\mach,\pol,\prov) \manystepstounder{\rho} (\mach',\pol',\prov')\)}
@@
 
-\paragraph*{Overlay sequence definitions}
+\paragraph*{Macro-instruction definitions}
 
@@
 
-Note that there can be multiple overlay sequences with different instructions
+Note that there can be multiple macro-instructions with different instructions
 that represent the same pseudoinstruction.
@@
 
-\paragraph*{Concurrency}
+\paragraph*{Concurrency and Traces}
 
-Thus far, we have defined the step relation of a single processor.
-We model multiple processors able to step independently by a global step relation,
-\(\stepstounder{}\), which simply steps an arbitrary processor:
+So far we have defined the execution steps of single processors. Processors
+have their own register banks, but share memory, and may step independently
+in any order. We can represent sequences of these steps as traces.
+Some steps of execution can be interpreted as {\it events} that can be observed
+from outside the system. We define events \(\obs \in \OBSS\) as calls, labeled
+with the target address and the values of the arguments, or else as the silent
+event \(\tau\):
+\[\begin{split}
+\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
+| & ~ \tau \\
+\end{split}\]
 
-\judgment[]
-         {\(\sigma \stepstounder{\rho} \sigma'\)}
-         {\(\sigma \stepstounder{} \sigma'\)}
+A {\em trace} is a nonempty, finite or infinite sequence
+of pairs of processors and events, ranged over by \(\obsT\).
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
+for list-cons.
+
+We convert macro-instructions into events via the \(\mathit{observe}\)
+function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.
+
+\begin{figure*}
+\[\mathit{observe} ~ \rho ~ \psi? ~ m =
+\begin{cases}
+  (\rho, \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket)
+  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
+%  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
+  (\rho, \tau) & \textnormal{otherwise} \\
+\end{cases}\]
+\caption{Converting overlay steps to events}
+\label{fig:observe}
+\end{figure*}
+
+We write that execution from a state produces an observation trace \(\sigma \hookrightarrow \obsT\)
+as follows, coinductively:
+
+\judgment{\(\not \exists \sigma' . \sigma \stepstounder{} \sigma'\)}
+         {\(\sigma \hookrightarrow \tau\)}
+
+\judgmenttwo{\(\sigma \stepstounder{\rho; \psi?} \sigma'\)}
+            {\(\sigma' \hookrightarrow \obsT\)}
+            {\(\sigma \hookrightarrow \notfinished{(\rho,\mathit{observe} ~ \rho ~ \psi? ~ \pi_m(\sigma'))}{\obsT}\)}
+
+Naturally, most states will admit many possible traces, depending on the order in which
+processors step.
+
+We define another relation that takes a trace of a callee up to its return, if one occurs.
+In this case we must specify that we are considering a single processor.
+We write this \(\rho,\context \downarrow \sigma \hookrightarrow \obsT\), where \(\context\)
+is the context of the caller.
+
+\judgment{\(\context = \pi_c(\sigma) ~ \rho\)}
+         {\(\rho, \context \downarrow \sigma \hookrightarrow \tau\)}
+
+\judgmentthree{\(\sigma \stepstounder{\rho',\psi?} \sigma'\)}
+              {\(\rho \not = \rho'\)}
+              {\(\rho, \context \downarrow \sigma' \hookrightarrow \obsT\)}
+              {\(\rho, \context \downarrow \sigma \hookrightarrow \obsT\)}
+
+\judgmentthree{\(\sigma \stepstounder{\rho,\psi?} \sigma'\)}
+              {\(\context \not = \pi_c(\sigma') ~ \rho\)}
+              {\(\context \downarrow \sigma' \hookrightarrow \obsT\)}
+              {\(\context \downarrow \sigma \hookrightarrow \notfinished{(\mathit{observe} ~ \rho ~ \psi? (\pi_m{\sigma'}))}{\obsT}\)}
+
+\paragraph*{Observational Similarity}
+
+We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar}
+from the perspective of processor \(\rho\), written \(\obsT_1 \eqsim_\rho
+\obsT_2\), if the sequence of non-silent \(\rho\) events is the same. That is, we
+compare up to deletion of \(\tau\) events and events from other processors, coinductively:
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{}{\(\obsT \eqsim_\rho \obsT\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
+           {\(\notfinished{(\rho, \obs)}{\obsT_1} \eqsim_\rho \notfinished{(\rho, \obs)}{\obsT_2}\)}
+\end{minipage}
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
+           {\(\notfinished{(\rho,\tau)}{\obsT_1} \eqsim_\rho \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
+           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho,\tau)}{\obsT_2}\)}
+\end{minipage}
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
+           {\(\notfinished{(\rho',\obs)}{\obsT_1} \eqsim_\rho \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
+           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho',\obs)}{\obsT_2}\)}
+\end{minipage}
 
@@
 the allocation is marked \(\mathit{object}(\phi)\) for a fresh identifier \(\phi\).
+[TODO: actually generate fresh \(\phi\) by keeping counter...]
+[TODO: need to put the original \(\phi\) somewhere. Maybe macro-instruction
+  for offsetting from stack pointer? Ugh.]
 
@@
 \begin{figure*}
+\judgmenttwobrlong[AllocC]
+                  {\(\psi = \mathbf{alloc} ~ size\)}
+                  {\((T,V) = (hd\_def ~ \context)\)}
+                  {\(V' = V\llbracket m[sp]+i \mapsto \mathit{object}(\phi) | 0 < i \leq size \rrbracket\)}
+                  {\(\mach, \prov, \context \constep{\rho;\psi} ctxs[\rho \mapsto (T,V')::(tl\_def ~ \context]\)}
+
 \[\begin{split}
@@
                   {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V)::\context\)}
-                  {\(V = \mathit{update\_view} ~ (snd ~ (hd\_def ~ \context)) ~ m[sp] ~ \overline{r_{args}}\)}
+                  {\(V = \mathit{update\_view} ~ (snd ~ \context) ~ m[sp] ~ \overline{r_{args}}\)}
                   {\(\mach, \prov, \context \constep{\rho;\psi} \context'\)}
@@
 
-\section{Stack Safety, Formally}
-\label{sec:properties}
-
-Now we can give our definition of stack safety. Each overlay sequence defines a
-context step; in turn, the context at the point of a call or return defines a
-property each for integrity, confidentiality, and well-bracketed control flow.
+Now we can give our definition of stack safety in terms of the context at the point of a call.
 
@@
 the expected behavior of the system in the future, after a possible return.
-
-\paragraph*{Event Traces}
-
-Some steps of execution can be interpreted as {\it events} that can be observed
-from outside the system. We define events \(\obs \in \OBSS\) as calls, labeled
-with the target address and the values of the arguments, or else as the silent
-event \(\tau\):
-\[\begin{split}
-\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
-| & ~ \tau \\
-\end{split}\]
-
-A {\em trace} is a nonempty, finite or infinite sequence
-of pairs of processors and events, ranged over by \(\obsT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
-for list-cons.
-
-We convert macro-instructions into events via the \(\mathit{observe}\)
-function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.
-
-\begin{figure*}
-\[\mathit{observe} ~ \rho ~ \psi? ~ m =
-\begin{cases}
-  (\rho, \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket)
-  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
-  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
-  (\rho, \tau) & \textnormal{otherwise} \\
-\end{cases}\]
-\caption{Converting overlay steps to events}
-\label{fig:observe}
-\end{figure*}
-
-We write that execution from a state produces an observation trace \(\sigma \hookrightarrow \obsT\)
-as follows, coinductively:
-
-\judgment{\(\not \exists \sigma' . \sigma \stepstounder{} \sigma'\)}
-         {\(\sigma \hookrightarrow \tau\)}
-
-\judgmenttwo{\(\sigma \stepstounder{\rho; \psi?} \sigma'\)}
-            {\(\sigma' \hookrightarrow \obsT\)}
-            {\(\sigma \hookrightarrow \notfinished{(\mathit{observe} ~ \rho ~ \psi? (\pi_m{\sigma'}))}{\obsT}\)}
-
-\paragraph*{Trace Until Return}
-
-We define another relation that takes a trace of a callee up to its return, if one occurs.
-We write this \(\context \downarrow \sigma \hookrightarrow \obsT\), where \(\context\)
-is the context of the caller.
-
-\judgment{\(\context = \pi_c(\sigma)\)}
-         {\(\context \downarrow \sigma \hookrightarrow \tau\)}
-
-\judgmentthree{\(\sigma \stepstounder{\rho,\psi?} \sigma'\)}
-              {\(\context \not = \pi_c(\sigma')\)}
-              {\(\context \downarrow \sigma' \hookrightarrow \obsT\)}
-              {\(\context \downarrow \sigma \hookrightarrow \notfinished{(\mathit{observe} ~ \rho ~ \psi? (\pi_m{\sigma'}))}{\obsT}\)}
-
-\paragraph*{Observational Similarity}
-
-We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar}
-from the perspective of an external observer, written \(\obsT_1 \eqsim
-\obsT_2\), if the sequence of non-silent events is the same. That is, we
-compare up to deletion of \(\tau\) events, coinductively:
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{}{\(\obsT \eqsim \obsT\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
-\end{minipage}
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
-\end{minipage}
 
@@
 
+\paragraph*{Property Definitions}
+
+And with the preliminary definitions out the way, we can define integrity.
+
 \definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)

### 2022-08-26 submission5/onlyformalism.tex
Last night's changes

@@
 | & \mathbf{return} & \\
-| & \mathbf{alloc} ~ size ~ public & size \in \mathbb{N}, public \in \mathbb{B} \\
+| & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
 | & \mathbf{setex} \\
@@
   for offsetting from stack pointer? Ugh.]
+
+Alloc macro-instructions model allocations that might, or might not, be
+shared. A function can always allocate more opjects by adjusting its
+stack pointer outside of a macro-instruction, but this will not be reflected
+in the context.
 

### 2022-09-01 submission5/onlyformalism.tex
Going to progressive introduction of concepts

@@
 it is permanently compromised.
+
+\section{The Basic Machine}
+
+The building blocks of the machine are {\em words} and {\em registers}.
+Words (\(\WORDS\)) are 64-bit integers ranged over by \(\word\) and, when used as addresses,
+\(\addr\). The processor has 32 registers, ranged over in general by \(\reg \in \REGS\),
+plus a program counter (\(\PCname\)). The following table divides
+the registers into metavariables, some belonging to caller- or callee- sets:
+
+\vspace{\abovedisplayskip}
+\begin{tabular}{| l | l | l |}
+  \hline
+  Set & Names & Purpose \\
+  \hline
+  CLR\_SV & \(rt_0 - rt_6\) & Caller-saved temps \\
+  & \(ra_0 - ra_1\) & Caller-saved args / return vals \\
+  & \(ra_2 - ra_7\) & Caller-saved args \\
+  \hline
+  CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
+  \hline
+  PUBLIC & \(rra\) & Return Address \\
+  & \(rsp\) & Stack Pointer \\
+  & \(pc\) & Program Counter \\
+  \hline
+\end{tabular}
+\vspace{\abovedisplayskip}
+
+Collectively addresses and registers are {\em state elements} \(\component\)
+in the set \(\COMPONENTS ::= \WORDS + \REGS\).
+%
+A {\em machine state} \(\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\)
+is a mapping from state elements to values. For convenience, we abbreviate
+multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
+as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
+are fomulae with free variables bound to sets in \(C\).
+The base machine step function
+\(\mach \stepstounder{} \mach' \in \MACHS \rightarrow \MACHS\) represents a single step.
+
+The set of instructions, \(I \subseteq \WORDS\), corresponds to the RISC-V ISA.
+We further introduce {\it macro-instructions}: sequences of
+instructions that operate as if they were single, atomic pseudoinstructions.
+A macro-instruction consists of both a list
+of instructions and a pseudoinstruction that identifies its purpose in the
+property. The pseudoinstructions, and the data that they carry, are as follows.
+
+\begin{align*}
+\psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+| & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+| & \mathbf{return} & \\
+| & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
+| & \mathbf{setex} \\
+| & \mathbf{throwex} \\
+\end{align*}
+
+An macro-instruction \((\bar{i},\psi) \in J\) is a pair of a pseudoinstruction and
+a non-empty list of instructions. In order to implement the atomic nature of
+executing macro-instructions,
+we define a {\it ranged-step} relation that, given a pair of addresses
+\(a_{low}\) and \(a_{high}\), steps as long the program counter is between them.
+
+\judgment[Done]
+         {\(\mach[\PCname] < a_{low} \lor \mach[\PCname] > a_{high}\)}
+         {\(a_{low},a_{high} \vdash \mach \manystepstounder{} \mach\)}
+
+\judgmenttwobrlong[Step]
+                  {\(a_{low} \leq \mach[\PCname] < a_{high}\)}
+                  {\(\mach \stepstounder{} \mach'\)}
+                  {\(a_{low},a_{high} \vdash \mach' \manystepstounder{} \mach''\)}
+                  {\(a_{low},a_{high} \vdash \mach \manystepstounder{} \mach''\)}
+
+Now we're ready to introduce the overlay semantics.
+Additionally, we keep track of security context information
+specific to each stack over the course of the run.
+This data is not relevant to execution,
+it is only used to define the stack safety property. We define a set of
+{\it security classes}:
+\[sc \in SEC ::= \public | \sealed | \unsealed\]
+A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\) divides the machine
+elements in terms of their purpose relative to the active function.
+The {\it initial view} \(V_0\) maps all stack locations and non-public registers
+to \(\unsealed\) and all other locations and public registers to \(\public\).
+A set of {\it targets} identify the addresses of the program counter or
+stack pointer at a return point:
+\[\begin{split}
+tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
+& \mathit{Catch} ~ a_{pc} ~ a_{sp} \\
+& \mathit{Done} \\
+\end{split}\]
+And a {\it stack context} is a non-empty stack of pairs of targets and views:
+\[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]
+
+The initial stack context \(\context_0\) is \(\lbracket (\mathit{Done}, V_0) \rbracket\).
+
+An {\it overlay state} \(\sigma\) consists of a pair of a machine state
+and a stack contexts.
+\[\sigma = (\mach, \context) \in \mathcal{S} ::= \MACHS \times \CONTEXTS\]
+
+The labeled step function for overlay states
+\(\sigma \stepstounder{\psi?} \sigma' \in
+\mathcal{S} \rightarrow  (\Psi + \bot) \times \mathcal{S}\)
+proceeds by an {\it overlay step}
+if the program counter points to the start of a macro-instruction's instruction sequence.
+Otherwise, it proceeds by the default step. The definitions of context steps
+\(\constep{\psi} \in \CONTEXTS \times \Psi \rightarrow \CONTEXTS\)
+will be defined for each pseudoinstruction as we introduce new topics.
+
+\judgmentthreebrlong[Overlay]
+                    {\((\bar{i}, \psi) \in J\)}
+                    {\(\mach',\context \constep{\psi} \context'\)}
+                    {\(\mach \manystepstounder{} \mach'\)}
+                    {\(\mach[\PCname+n] = \bar{i}[n]\) for each \(n\) where \(0 \leq n < |\bar{i}|\)}
+                    {\((\mach,\context)) \stepstounder{\psi} (\mach',\context')\)}
+
+\judgment[Default]
+         {\(\mach \stepstounder{} \mach'\)}
+         {\((\mach,\context) \stepstounder{\bot} (\mach',\context)\)}
+
+\paragraph*{Macro-instruction definitions}
+
+[TBD]
+
+Note that there can be multiple macro-instructions with different instructions
+that represent the same pseudoinstruction.
+[This is less likely when we have smaller pseudoinstructions.]
+
+\paragraph*{Events and Traces}
+
+Some macro-instructions may be interpreted as {\it events} that can be observed
+from outside the system. For now we define events \(\obs \in \OBSS\) as calls, labeled
+with the target address and the values of the arguments, or else as the silent
+event \(\tau\):
+\[\begin{split}
+\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
+| & ~ \tau \\
+\end{split}\]
+
+We convert macro-instructions into events via the \(\mathit{observe}\)
+function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.
+
+A {\em trace} is a nonempty, finite or infinite sequence
+of pairs of processors and events, ranged over by \(\obsT\).
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
+for list-cons.
+
+\begin{figure*}
+\[\mathit{observe} ~ \psi? ~ m =
+\begin{cases}
+  \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
+  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
+  \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
+  & \textnormal{when } ~ \psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}} \\
+%  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
+  \tau & \textnormal{otherwise} \\
+\end{cases}\]
+\caption{Converting overlay steps to events}
+\label{fig:observe}
+\end{figure*}
+
+We write that execution from a state produces an observation trace \(\sigma \hookrightarrow \obsT\)
+as follows, coinductively:
+
+\judgmenttwo{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
+            {\((\mach',\context') \hookrightarrow \obsT\)}
+            {\((\mach,\context) \hookrightarrow \notfinished{\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}
+
+Naturally, most states will admit many possible traces, depending on the order in which
+processors step.
+
+We define another relation that takes a trace of a callee up to its return, or until
+an exception is thrown, if either occurs.
+We write this \(\context' \downarrow (\mach,\context) \hookrightarrow \obsT\), where \(\context\)
+is the context of the caller.
+
+\judgment{\(|\context| \leq |\context'|\)}
+         {\(\context' \downarrow (\mach,\context) \hookrightarrow \tau\)}
+
+\judgmenttwobrlong{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
+                  {\(|\context'| > |\context''|\)}
+                  {\(\context'' \downarrow (\mach',\context') \hookrightarrow \obsT\)}
+                  {\(\context'' \downarrow (\mach,\context) \hookrightarrow \notfinished{(\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}
+
+\paragraph*{Observational Similarity}
+
+We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
+written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
+is the same. That is, we compare up to deletion of \(\tau\) events.
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{}{\(\obsT \eqsim \obsT\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
+\end{minipage}
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
+\end{minipage}
+
+%\begin{minipage}{.4\columnwidth}
+%  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
+%           {\(\notfinished{(\rho',\obs)}{\obsT_1} \eqsim_\rho \obsT_2\)}
+%\end{minipage}
+%\begin{minipage}{.4\columnwidth}
+%  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
+%           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho',\obs)}{\obsT_2}\)}
+%\end{minipage}
+
+\section{Stack Safety With Calls and Returns}
+
+In this section we will present the context steps for calls and returns,
+and show how they translate into properties. They are given formally in
+Figure \ref{fig:callcontexts}.
+
+For calls, we define a function, {\it update view}, that defines the view
+of a caller in terms of that of the callee.
+First, all stack locations below the stack pointer
+are \(\unsealed\) -- fair game for the callee to allocate or write to, and
+to read after writing. All \(\unsealed\) locations above the stack pointer
+become \(\sealed\) -- their contents are both secret to and protected
+from the callee. Argument and system registers are automatically marked
+\(\mathit{public}\). All other locations above the stack pointer
+retain their prior status (these will only be \(\sealed\) for now,
+until we introduce shareable allocations shortly.)
+
+To update the context for a call, we pair the updated view with the return target of the
+call and push it to the stack. The return target will tell us whether a future
+return is to the proper location with the stack pointer restored.
+The view tells us how to treat the security of the callee's state, and as
+it becomes more complicated, will benefit from the history that we maintain.
+We take the head of a list, \(\mathit{hd}\), in the usual way, here a
+total function because the context will never be empty.
+
+On a return, we pop the topmost pair from the context that has
+a \(\mathit{return}\), along with any other pairs we pass.
+
+\begin{figure*}
+\[\begin{split}
+& \mathit{update\_view} ~ V ~ a_{sp} ~ \overline{r_{args}} \triangleq
+\lambda \component .
+\begin{cases}
+  \public & \textnormal{else if } \component \in \REGS \textnormal{ and }
+  \component \in \overline{r_{args}} \textnormal{ or } \component \in \mathit{PUBLIC} \\
+  \unsealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } \component \leq a_{sp} \\
+  \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } a_{sp} < \component \textnormal { and } V ~ \component = \unsealed \\
+  V ~ \component & \textnormal{otherwise} \\
+\end{cases} \\
+\end{split}\]
+
+\judgmenttwobrlong[CallC]
+                  {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}}\)}
+                  {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V)::\context\)}
+                  {\(V = \mathit{update\_view} ~ (snd ~ (hd ~ \context)) ~ m[sp] ~ \overline{r_{args}}\)}
+                  {\(\mach, \context \constep{\psi} \context'\)}
+
+\judgment[ReturnCRet]
+         {\(\psi = \mathbf{return}\)}
+         {\(\mach, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\psi} \context\)}
+
+\judgment[ReturnCDone]
+         {\(\psi = \mathbf{return}\)}
+         {\(m, [\mathit{Done},V] \constep{\psi} [\mathit{Done},V]\)}
+
+\judgmenttwo[ReturnCSkip]
+            {\(\psi = \mathbf{return}\)}
+            {\(\mach, \context \constep{\psi} \context'\)}
+            {\(\mach, (\_::\context) \constep{\psi} \context'\)}
+
+\caption{Call, enter, and return macro-instructions}
+\label{fig:callcontexts}
+\end{figure*}
+
+Now we can give our definition of stack safety in terms of the context at the point of a call.
+
+\subsection{Facts Abouts Calls and Returns}
+
+Here we define some logical operations to reason about the behavior of the
+system over time. These have a temporal-logic flavor, as they reflect
+the expected behavior of the system in the future, after a possible return.
+
+\paragraph*{On-return}
+
+The intuition behind {\it return-time integrity} (below) is that a caller may expect its
+sealed data to be unchanged when control returns to it. In fact, the callee
+may overwrite such data -- when the data are found in callee-saved registers
+this is perfectly legal -- as long as it either restores it, or has some guarantee
+that its changes will impact the caller.
+
+We start by defining a second-order logical operator
+\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
+where \(P\) is a predicate on machine states. This is a coinductive relation
+similar to ``weak until'' in temporal logic -- it holds if the condition
+is never reached.
+
+\judgment{\(P ~ \mach\)}
+         {\((\context \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
+
+\judgmenttwobrlong{\(\context_0 \not = \context\)}
+                  {\((\context_0 \uparrow P) ~ (\mach', \pol', \prov', \context')\)}
+                  {\((\mach, \pol, \prov, \context) \stepstounder{\rho,\psi?}
+                    (\mach', \pol', \prov', \context')\)}
+                  {\((\context_0 \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
+
+Similarly, in {\it caller confidentiality}, we will want to compare future states,
+so we give a binary equivalent, \(\context \Uparrow R\), so that
+\(\sigma (\context \Uparrow R) \sigma'\) holds if \(R\) holds on the
+first return states of after \(\sigma\) and \(\sigma'\).
+
+\judgment[Returned]
+         {\(\mach ~ R ~ \mach'\)}
+         {\((\mach,\pol,\prov,\context) ~ (\context \Uparrow R) ~ (\mach',\pol',\prov',\context)\)}
+
+\judgmentthree[Left]
+              {\(\context \not = \pi_c(\sigma_1)\)}
+              {\(\sigma_1 \stepstounder{\rho;\psi?} \sigma_1'\)}
+              {\(\sigma_1' ~ (\context \Uparrow R) ~ \sigma_2\)}     
+              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
+
+\judgmentthree[Right]
+              {\(\context \not = \pi_c(\sigma_2)\)}
+              {\(\sigma_2 \stepstounder{\rho;\psi?} \sigma_2'\)}
+              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2'\)}     
+              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
+
+\paragraph*{Variants and Safe Elements}
+
+One or more elements of a given state are {\it safe} if their values
+in that state have no observable influence on future execution. We define this
+formally as a sort of non-interference: for a state \(\sigma\) and
+a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
+holds if we can replace the contents of \(\components\) with arbitrary values
+and have the resulting state produce an equivalent trace. The state created
+by modifying \(\components\) in this way is termed a variant.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+written \(\mach \approx_\components \nach\), if, for
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
+
+\definition Full states \((\mach,\pol,\prov,\context)\) and \((\nach,\pol,\prov,\context)\)
+are also {\em \(\components\)-variants} if \(\mach\) and \(\nach\) are.
+
+\definition An element set \(\components\) is safe in state \(\sigma)\),
+written \(\components ~ \mathit{SAFE} ~ \sigma\), if for all
+\(\sigma'\) such that \(\sigma \approx_{\components'} \sigma'\), if 
+\(\sigma \hookrightarrow \obsT\) and
+\(\sigma' \hookrightarrow \obsT'\), then
+\(\obsT \eqsim \obsT'\).
+
+\paragraph*{Property Definitions}
+
+And with the preliminary definitions out the way, we can define integrity.
+
+\definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
+such that \(\mach[\component] \not = \mach'[\component]\).
+
+\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) and
+\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component = \sealed\}\).
+Let \(P\) be a predicate on states \(\sigma'\) that holds if
+\(\Delta(\mach, \pi_m(\sigma')) \cap \components ~ \mathit{SAFE} ~ \sigma'\).
+Then \(\sigma\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.
+
+\definition A system enjoys {\it stack integrity} if, for all reachable states \(\sigma\) such that
+\(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
+\(\sigma'\) enjoys return-time integrity.
+
+\paragraph*{Caller Confidentiality}
+
+Confidentiality properties are also modeled as non-interference, but with a twist:
+the caller's confidential data should be kept secret from the callee, but after return,
+will naturally be accessible to the caller once again. So, we must model confidentiality
+in terms of events inside the callee. But confidentiality violations might not become
+visible during the callee's lifetime -- the callee might read a secret and stash it somewhere
+that it will be accessed later. So we must also consider the machine state at the end
+of the call.
+
+We could define a natural, strict form of confidentiality that requires that the final
+states be identical. However, as with integrity, this is stricter than we need -- there
+may be some conditions in which a secret is copied, but never used again. So, we once
+again wish for our final condition to make the values {\it safe}. Which values?
+
+\definition Let \(\mach,\mach'\) and \(\nach,\nach'\)
+be pairs of machine states. Their {\em corrupted set}, written
+\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all elements
+\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) such that
+\(\mach'[\component] \not = \nach'[\component]\).
+
+\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) be a state and
+\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component \in
+\{\sealed, \unsealed\}\}\).
+
+Then \(\sigma\) enjoys {\it caller confidentiality} if, for any \(\sigma'\)
+that is a \(\components\)-variant of \(\sigma\), we can take
+\(\context \uparrow \sigma \hookrightarrow \obsT\) and
+\(\context \uparrow \sigma' \hookrightarrow \obsT'\) and have that
+\(\obsT \simeq \obsT'\). And if, additionally, \(\sigma ~ (\context \Uparrow R) ~ \sigma'\),
+where \(\sigma_1 ~ R ~ \sigma_2\) if
+\(\bar{\Diamond}(\mach,\pi_m(\sigma_1),\pi_m(\sigma'),\pi_m(\sigma_2)) ~ \mathit{SAFE} ~ \sigma_1\).
+
+\definition A system enjoys {\it universal caller confidentiality} if, for all reachable states
+\(\sigma\) such that \(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
+\(\sigma'\) enjoys caller confidentiality.
 

### 2022-09-01 submission5/onlyformalism.tex
Pretty well done

@@
 %
-A {\em machine state} \(\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\)
-is a mapping from state elements to values. For convenience, we abbreviate
+A {\em machine state} \(\mach \in \MACHS\) is a type that exposes a mapping
+\(\mach[\component]\) from state elements to values, and may also contain additional
+security state. For convenience, we abbreviate
 multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
@@
 The base machine step function
-\(\mach \stepstounder{} \mach' \in \MACHS \rightarrow \MACHS\) represents a single step.
+\(\mach \stepstounder{} \mach' \in \MACHS \rightharpoonup \MACHS\) represents a single step.
+It is partial to represent the capacity of the machine to failstop.
+In order to accommodate failstop behavior, our properties will be {\it termination insensitive}.
 
@@
 | & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
-| & \mathbf{setex} \\
-| & \mathbf{throwex} \\
+| & \mathbf{setex} ~ \addr_{buf} \\
+| & \mathbf{throwex} ~ \addr_{buf} \\
 \end{align*}
@@
 tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
-& \mathit{Catch} ~ a_{pc} ~ a_{sp} \\
+& \mathit{Catch} ~ a_{buf} ~ a_{pc} ~ a_{sp} \\
 & \mathit{Done} \\
@@
 
-The initial stack context \(\context_0\) is \(\lbracket (\mathit{Done}, V_0) \rbracket\).
+The initial stack context \(\context_0\) is \(\left[ (\mathit{Done}, V_0) \right]\).
 
@@
 
+\judgment{\(\not \exists \mach' ~ \context'. (\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
+         {\((\mach,\context) \hookrightarrow \tau\)}
+
 \judgmenttwo{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
@@
                   {\(\context'' \downarrow (\mach',\context') \hookrightarrow \obsT\)}
-                  {\(\context'' \downarrow (\mach,\context) \hookrightarrow \notfinished{(\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}
+                  {\(\context'' \downarrow (\mach,\context) \hookrightarrow \notfinished{\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}
 
@@
 it becomes more complicated, will benefit from the history that we maintain.
-We take the head of a list, \(\mathit{hd}\), in the usual way, here a
-total function because the context will never be empty.
 
 On a return, we pop the topmost pair from the context that has
-a \(\mathit{return}\), along with any other pairs we pass.
+a \(\mathit{return}\), along with any other pairs (catch points,
+for instance) that we pass.
 
@@
                   {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}}\)}
-                  {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V)::\context\)}
-                  {\(V = \mathit{update\_view} ~ (snd ~ (hd ~ \context)) ~ m[sp] ~ \overline{r_{args}}\)}
-                  {\(\mach, \context \constep{\psi} \context'\)}
+                  {\(T' = \mathit{Return} ~ (m[pc] + 4) ~ m[sp]\)}
+                  {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ \overline{r_{args}}\)}
+                  {\(\mach, (T,V)::\context \constep{\psi} (T',V')::\context\)}
 
@@
 
-\caption{Call, enter, and return macro-instructions}
+\caption{Call and return context updates}
 \label{fig:callcontexts}
@@
 We start by defining a second-order logical operator
-\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
+\(\context_{ret} \uparrow P\), read ``\(P\) holds on return to \(\context_{ret}\),''
 where \(P\) is a predicate on machine states. This is a coinductive relation
@@
 
-\judgment{\(P ~ \mach\)}
-         {\((\context \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
+\judgmenttwo[Returned]
+            {\(P ~ (\mach,\context)\)}
+            {\(|\context| \leq |\context_{ret}|\)}
+            {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
 
-\judgmenttwobrlong{\(\context_0 \not = \context\)}
-                  {\((\context_0 \uparrow P) ~ (\mach', \pol', \prov', \context')\)}
-                  {\((\mach, \pol, \prov, \context) \stepstounder{\rho,\psi?}
-                    (\mach', \pol', \prov', \context')\)}
-                  {\((\context_0 \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
+\judgmenttwobrlong[Step]
+                  {\(|\context| > |\context_{ret}|\)}
+                  {\((\mach, \context) \stepstounder{\psi?} (\mach', \context')\)}
+                  {\((\context_{ret} \uparrow P) ~ (\mach', \context')\)}
+                  {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
 
@@
 so we give a binary equivalent, \(\context \Uparrow R\), so that
-\(\sigma (\context \Uparrow R) \sigma'\) holds if \(R\) holds on the
-first return states of after \(\sigma\) and \(\sigma'\).
+\((\mach,\context) ~ (\context'' \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
+first return states after \((\mach,\context)\) and \((\mach',\context')\).
 
-\judgment[Returned]
-         {\(\mach ~ R ~ \mach'\)}
-         {\((\mach,\pol,\prov,\context) ~ (\context \Uparrow R) ~ (\mach',\pol',\prov',\context)\)}
+\judgmentthree[Returned]
+              {\((\mach,\context) ~ R ~ (\mach',\context')\)}
+              {\(|\context| \leq |\context_{ret}|\)}
+              {\(|\context'| \leq |\context_{ret}|\)}
+              {\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\)}
 
-\judgmentthree[Left]
-              {\(\context \not = \pi_c(\sigma_1)\)}
-              {\(\sigma_1 \stepstounder{\rho;\psi?} \sigma_1'\)}
-              {\(\sigma_1' ~ (\context \Uparrow R) ~ \sigma_2\)}     
-              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
+\judgmenttwobr[Left]
+              {\(\context_1 > \context_{ret}\)}
+              {\((\mach_1,\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
+              {\(\context_2 > \context_{ret}\)}
+              {\((\mach_1',\context_1') ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
+              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
 
-\judgmentthree[Right]
-              {\(\context \not = \pi_c(\sigma_2)\)}
-              {\(\sigma_2 \stepstounder{\rho;\psi?} \sigma_2'\)}
-              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2'\)}     
-              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
+\judgmenttwobr[Right]
+              {\(\context_1 > \context_{ret}\)}
+              {\((\mach_2,\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
+              {\(\context_2 > \context_{ret}\)}
+              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2',\context_2')\)}
+              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
 
-\paragraph*{Variants and Safe Elements}
+\paragraph*{Variants and Vestigial Elements}
 
-One or more elements of a given state are {\it safe} if their values
+One or more elements of a given state are {\it vestigial} if their values
 in that state have no observable influence on future execution. We define this
-formally as a sort of non-interference: for a state \(\sigma\) and
-a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
+formally as a sort of non-interference: for a state \((\mach,\context)\) and
+a set of elements \(\components\), \((\mach,\context) \parallel \components\)
 holds if we can replace the contents of \(\components\) with arbitrary values
@@
 
-\definition Full states \((\mach,\pol,\prov,\context)\) and \((\nach,\pol,\prov,\context)\)
-are also {\em \(\components\)-variants} if \(\mach\) and \(\nach\) are.
-
-\definition An element set \(\components\) is safe in state \(\sigma)\),
-written \(\components ~ \mathit{SAFE} ~ \sigma\), if for all
-\(\sigma'\) such that \(\sigma \approx_{\components'} \sigma'\), if 
-\(\sigma \hookrightarrow \obsT\) and
-\(\sigma' \hookrightarrow \obsT'\), then
+\definition An element set \(\components\) is vestigial in state \(\sigma)\),
+written \((\mach,\context) \parallel \components\), if for all
+\(\nach\) such that \(\mach \approx_{\components'} \mach'\), if 
+\((\mach,\context) \hookrightarrow \obsT\) and
+\((\nach,\context) \hookrightarrow \obsT'\), then
 \(\obsT \eqsim \obsT'\).
@@
 
-\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) and
-\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component = \sealed\}\).
-Let \(P\) be a predicate on states \(\sigma'\) that holds if
-\(\Delta(\mach, \pi_m(\sigma')) \cap \components ~ \mathit{SAFE} ~ \sigma'\).
-Then \(\sigma\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.
+\definition Let \((\mach,(T,V)::\context)\) be a state and
+\(\components = \{\component | V ~ \component = \sealed\}\).
+Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
+\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
+Then \(\mach\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.
 
-\definition A system enjoys {\it stack integrity} if, for all reachable states \(\sigma\) such that
-\(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
-\(\sigma'\) enjoys return-time integrity.
+\definition A system enjoys {\it universal stack integrity} if, for all reachable states
+\((\mach,\context)\) such that
+\((\mach,\context) \stepstounder{\psi} (\mach',\context')\) with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
+\((\mach',\context')\) enjoys return-time integrity.
 
@@
 may be some conditions in which a secret is copied, but never used again. So, we once
-again wish for our final condition to make the values {\it safe}. Which values?
+again wish for our final condition to allow values to be vestigial. Which values?
 
@@
 
-\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) be a state and
-\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component \in
-\{\sealed, \unsealed\}\}\).
+\definition Let \((\mach,(T,V)::\context)\) be a state and
+\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
 
-Then \(\sigma\) enjoys {\it caller confidentiality} if, for any \(\sigma'\)
-that is a \(\components\)-variant of \(\sigma\), we can take
-\(\context \uparrow \sigma \hookrightarrow \obsT\) and
-\(\context \uparrow \sigma' \hookrightarrow \obsT'\) and have that
-\(\obsT \simeq \obsT'\). And if, additionally, \(\sigma ~ (\context \Uparrow R) ~ \sigma'\),
-where \(\sigma_1 ~ R ~ \sigma_2\) if
-\(\bar{\Diamond}(\mach,\pi_m(\sigma_1),\pi_m(\sigma'),\pi_m(\sigma_2)) ~ \mathit{SAFE} ~ \sigma_1\).
+Then \((\mach,\context)\) enjoys {\it caller confidentiality} with respect to
+some \(\context_{ret}\) if, for any \(\nach\)
+that is a \(\components\)-variant of \(\mach\), we can take
+\(\context \uparrow (\mach,\context) \hookrightarrow \obsT\) and
+\(\context \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
+\(\obsT \simeq \obsT'\);
+and additionally if \((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\nach,\context)\),
+where \((\mach_1,\context_1) ~ R ~ (\nach_2,\context_2)\) if
+\((\mach_1 \parallel \bar{\Diamond}(\mach,\mach_1,\nach,\nach_2)\).
 
 \definition A system enjoys {\it universal caller confidentiality} if, for all reachable states
-\(\sigma\) such that \(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
-\(\sigma'\) enjoys caller confidentiality.
+\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
+with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
+\((\mach',\context')\) enjoys caller confidentiality with respect to \(\context\).
 
-\subsection{The Machine in Detail}
-\label{sec:prelim}
+\section{Allocating Memory}
 
-The building blocks of the machine are {\em words} and {\em registers}.
-Words (\(\WORDS\)) are 64-bit integers ranged over by \(w\) and, when used as addresses,
-\(a\). Each processor has 32 registers, ranged over in general by \(r\),
-plus a program counter (\(\PCname\)). The following table divides
-the registers into metavariables, some belonging to caller- or callee- sets:
+We now introduce the context step for the \(\mathbf{alloc} ~ \mathit{size}\)
+macro-instruction. Without \(\mathbf{alloc}\), our system already allowed a function
+to allocate as much private memory as it liked by moving the stack pointer. With
+\(\mathbf{alloc}\) it can now also allocate public objects that can be accessed
+freely by its callees. Later we will refine this system to restrict public objects
+to be accessed only via capabilities.
 
-\vspace{\abovedisplayskip}
-\begin{tabular}{| l | l | l |}
-  \hline
-  Set & Names & Purpose \\
-  \hline
-  CLR\_SV & \(rt_0 - rt_6\) & Caller-saved temps \\
-  & \(ra_0 - ra_1\) & Caller-saved args / return vals \\
-  & \(ra_2 - ra_7\) & Caller-saved args \\
-  \hline
-  CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
-  \hline
-  PUBLIC & \(rra\) & Return Address \\
-  & \(rsp\) & Stack Pointer \\
-  & \(pc\) & Program Counter \\
-  \hline
-\end{tabular}
-\vspace{\abovedisplayskip}
+\judgmentbr[AllocC]
+           {\(\psi = \mathbf{alloc} ~ size\)}
+           {\(V' = V\llbracket m[sp]+i \mapsto \public | 0 < i \leq size \rrbracket\)}
+           {\((\mach, (T,V)::\context) \constep{\psi} (T,V')::\context\)}
 
-We specify that a register belongs to processor \(\rho \in \PROCS\) by writing
-it \(r^\rho\), but will omit this notation when the active
-processor is unambiguous from context. The set of all registers across all processors
-is \(\REGS\).
+The remainder of our property definitions are unchanged. The \(\mathit{update\_view}\)
+function preserves the public status of allocated objects, and they are exempt from
+integrity and confidentiality properties.
 
-Collectively addresses and registers are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \WORDS + \REGS\).
-%
-A {\em machine state} \(\mach \in \MACHS ::= \COMPONENTS \rightarrow \WORDS\)
-is a mapping from state elements to values. For convenience, we abbreviate
-multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
-as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
-are fomulae with free variables bound to sets in \(C\).
-The base machine step relation 
-\(\mach \machstep{\rho} \mach' \subseteq \MACHS \times \PROCS \times \MACHS\)
-represents a single step of the processor \(\rho\).
+\section{Exceptions}
 
-The machine is equipped with additional state representing the tag-based hardware.
-We represent this as a policy state, \(\pol \in \POLS ::= \COMPONENTS \rightarrow \mathcal{T}\),
-where \(\mathcal{T}\) is the set of tags, discussed [in the section on security.]
-We represent a {\it policy} via the step relation
-\(\mach, \pol \harpoonunder{\rho}_P \pol' \subseteq \MACHS \times \POLS \times \PROCS \times \POLS\).
-Steps are deterministic and  undefined represents a {\it failstop}.
-
-Our properties additionally depend on a rudimentary model of memory-safe
-objects and pointer provenance. A provenance \(\phi \in \mathbb{N}\)
-identifies a value derived from the \(\phi\)th object allocated since
-the beginning of execution. A provenance map
-\(\prov \in \PROVS ::= \COMPONENTS \rightarrow 2^{\mathbb{N}}\)
-designates, for every machine element, which provenances its current
-value contains. We define a step relation on provenance maps
-\(\mach, \prov \harpoonunder{\rho}_{pm} \prov' \subseteq \MACHS \times
-\PROVS \times \PROCS \times \PROVS\),
-exerpted in Figure \ref{fig:provenance} [TODO].
-
-We define a step relation over triples of a machine state,
-a policy state, and a provenance map:
-
-\judgmentthree{\(\mach \machstep{\rho} \mach'\)}
-              {\(\mach, \pol \polstep{\rho} \pol'\)}
-              {\(\mach, \prov \provstep{\rho} \prov'\)}
-              {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
-
-As the last piece of state, we keep track of security context information
-specific to each stack
-over the course of the run. This data is not relevant to execution,
-it is only used to define the stack safety property. We define a set of
-{\it security classes}:
-\[sc \in SEC ::= \public | \sealed | \mathit{object}(\phi) | \unsealed\]
-And of {\it targets}, which identify the addresses of the program counter or
-stack pointer at a return or catch point:
-\[\begin{split}
-tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
-| & \mathit{Catch} ~ a_{pc} ~ a_{sp} \\
-\end{split}\]
-A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\) divides the machine
-elements in terms of their purpose relative to the active function.
-And a {\it stack context} is a stack of pairs of targets and views:
-\[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]
-
-The set of instructions, \(I \subseteq \WORDS\), corresponds to the RISC-V ISA.
-We further introduce {\it macro-instructions}: sequences of
-instructions that operate as if they were single, atomic pseudoinstructions.
-A macro-instruction consists of both a list
-of instructions and a pseudoinstruction that identifies its purpose in the
-property. The pseudoinstructions, and the data that they carry, are as follows.
-[TODO: still handle stack args by offset.]
-
-\begin{align*}
-\psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
-| & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
-| & \mathbf{return} & \\
-| & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
-| & \mathbf{setex} \\
-| & \mathbf{throwex} \\
-\end{align*}
-
-An macro-instruction \((\bar{i},\psi) \in J\) is a pair of a pseudoinstruction and
-a non-empty list of instructions. In order to implement the atomic nature of
-executing macro-instructions,
-we define a {\it ranged-step} relation that, given a pair of addresses
-\(a_{low}\) and \(a_{high}\), steps as long the program counter is between them.
-
-\judgment[Done]
-         {\(\mach[\PCname] < a_{low} \lor \mach[\PCname] > a_{high}\)}
-         {\(a_{low},a_{high} \vdash (\mach,\pol,\prov) \manystepstounder{\rho} (\mach,\pol,\prov)\)}
-
-\judgmenttwobrlong[Step]
-                  {\(a_{low} \leq \mach[\PCname] < a_{high}\)}
-                  {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
-                  {\(a_{low},a_{high} \vdash (\mach',\pol',\prov') \manystepstounder{\rho} (\mach'',\pol'',\prov'')\)}
-                  {\(a_{low},a_{high} \vdash (\mach,\pol,\prov) \manystepstounder{\rho} (\mach'',\pol'',\prov'')\)}
-
-Now we're finally ready to introduce the overlay semantics.
-A {\it full state} \(\sigma\) consists of a 4-tuple of a machine state,
-a policy state, a provenance map, and a map from processors to stack contexts:
-\[\sigma = (\mach, \pol, \prov, ctxs) \in \mathcal{S} ::=
-\MACHS \times \POLS \times \PROVS \times (\PROCS \rightarrow \CONTEXTS)\]
-
-The execution relation for full states
-\(\sigma \stepstounder{\rho;\psi?} \sigma' \subseteq
-\mathcal{S} \times \PROCS \times (\Psi + \bot) \times \mathcal{S}\)
-proceeds by an {\it overlay step}
-if the program counter points to the start of a macro-instruction.
-Otherwise, it proceeds by the default step. The definitions of context steps
-\(\constep{\rho;\psi}\) will be defined for each pseudoinstruction as we introduce
-topics.
-
-\judgmenttwobrlongbrlong[Overlay]
-                        {\((\bar{i}, \psi) \in J\)}
-                        {\(\mach',\prov',(ctxs ~ \rho) \constep{\rho;\psi} \context\)}
-                        {\((\mach,\pol,\prov) \manystepstounder{\rho} (\mach',\pol',\prov')\)}
-                        {\(\mach[\PCname+n] = \bar{i}[n]\) for each \(n\) where \(0 \leq n < |\bar{i}|\)}
-                        {\((\mach,\pol,\prov,ctxs) \stepstounder{\rho;\psi} (\mach',\pol',\prov',ctxs[\rho \mapsto \context])\)}
-
-\judgment[Default]
-         {\((\mach,\pol,\prov) \stepstounder{\rho} (\mach',\pol',\prov')\)}
-         {\((\mach,\pol,\prov,ctxs) \stepstounder{\rho; \bot} (\mach',\pol',\prov',ctxs)\)}
-
-\paragraph*{Macro-instruction definitions}
-
-[TBD]
-
-Note that there can be multiple macro-instructions with different instructions
-that represent the same pseudoinstruction.
-[This is less likely when we have smaller pseudoinstructions.]
-
-\paragraph*{Concurrency and Traces}
-
-So far we have defined the execution steps of single processors. Processors
-have their own register banks, but share memory, and may step independently
-in any order. We can represent sequences of these steps as traces.
-Some steps of execution can be interpreted as {\it events} that can be observed
-from outside the system. We define events \(\obs \in \OBSS\) as calls, labeled
-with the target address and the values of the arguments, or else as the silent
-event \(\tau\):
-\[\begin{split}
-\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
-| & ~ \tau \\
-\end{split}\]
-
-A {\em trace} is a nonempty, finite or infinite sequence
-of pairs of processors and events, ranged over by \(\obsT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
-for list-cons.
-
-We convert macro-instructions into events via the \(\mathit{observe}\)
-function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.
+We implement exceptions via another pair of macro-instructions:
+\(\mathbf{setex} ~ a_{buf}\), akin to \(\mathit{setjmp}\), with the
+jump buffer stored at \(a_{buf}\), and \(\mathbf{throwex} ~ a_{buf}\),
+akin to \(\mathit{longjmp}\). The context updates for exceptions are given
+in Figure \ref{fig:excontexts}. The update for \(\mathbf{setex}\) pushes
+a \(\mathit{Catch}\) target onto the stack, identified by \(\addr_{buf}\),
+with the same view as the current function. For \(\mathbf{throwex}\),
+we pop the stack until we find a \(\mathit{Catch} ~ a_{buf} ~ \_ ~ \_\),
+popping it as well. Rules apply in order.
 
 \begin{figure*}
-\[\mathit{observe} ~ \rho ~ \psi? ~ m =
-\begin{cases}
-  (\rho, \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket)
-  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
-%  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
-  (\rho, \tau) & \textnormal{otherwise} \\
-\end{cases}\]
-\caption{Converting overlay steps to events}
-\label{fig:observe}
+
+\judgmenttwo[SetExC]
+            {\(\psi = \mathbf{setex} ~ \addr_{buf}\)}
+            {\(T' = \mathit{Catch} ~ \addr_{buf} ~ (\mach[pc] + 4) ~ \mach[sp]\)}
+            {\(\mach, (T,V)::\context \constep{\psi} (T',V)::(T,V)::\context\)}
+
+\judgment[ThrowExCCatch]
+         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
+         {\(\mach, (\mathit{Catch} \addr_{buf} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\psi} \context\)}
+
+\judgment[ThrowExCDone]
+         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
+         {\(m, [\mathit{Done},V] \constep{\psi} [\mathit{Done},V]\)}
+
+\judgmenttwo[ThrowExCSkip]
+         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
+         {\(\mach, \context \constep{\psi} \context'\)}
+         {\(\mach, (\_::\context) \constep{\psi} \context'\)}
+
+\caption{Exception context updates}
+\label{fig:excontexts}
 \end{figure*}
 
-We write that execution from a state produces an observation trace \(\sigma \hookrightarrow \obsT\)
-as follows, coinductively:
+\section{Tail Call Elimination}
 
-\judgment{\(\not \exists \sigma' . \sigma \stepstounder{} \sigma'\)}
-         {\(\sigma \hookrightarrow \tau\)}
-
-\judgmenttwo{\(\sigma \stepstounder{\rho; \psi?} \sigma'\)}
-            {\(\sigma' \hookrightarrow \obsT\)}
-            {\(\sigma \hookrightarrow \notfinished{(\rho,\mathit{observe} ~ \rho ~ \psi? ~ \pi_m(\sigma'))}{\obsT}\)}
-
-Naturally, most states will admit many possible traces, depending on the order in which
-processors step.
-
-We define another relation that takes a trace of a callee up to its return, if one occurs.
-In this case we must specify that we are considering a single processor.
-We write this \(\rho,\context \downarrow \sigma \hookrightarrow \obsT\), where \(\context\)
-is the context of the caller.
-
-\judgment{\(\context = \pi_c(\sigma) ~ \rho\)}
-         {\(\rho, \context \downarrow \sigma \hookrightarrow \tau\)}
-
-\judgmentthree{\(\sigma \stepstounder{\rho',\psi?} \sigma'\)}
-              {\(\rho \not = \rho'\)}
-              {\(\rho, \context \downarrow \sigma' \hookrightarrow \obsT\)}
-              {\(\rho, \context \downarrow \sigma \hookrightarrow \obsT\)}
-
-\judgmentthree{\(\sigma \stepstounder{\rho,\psi?} \sigma'\)}
-              {\(\context \not = \pi_c(\sigma') ~ \rho\)}
-              {\(\context \downarrow \sigma' \hookrightarrow \obsT\)}
-              {\(\context \downarrow \sigma \hookrightarrow \notfinished{(\mathit{observe} ~ \rho ~ \psi? (\pi_m{\sigma'}))}{\obsT}\)}
-
-\paragraph*{Observational Similarity}
-
-We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar}
-from the perspective of processor \(\rho\), written \(\obsT_1 \eqsim_\rho
-\obsT_2\), if the sequence of non-silent \(\rho\) events is the same. That is, we
-compare up to deletion of \(\tau\) events and events from other processors, coinductively:
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{}{\(\obsT \eqsim_\rho \obsT\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
-           {\(\notfinished{(\rho, \obs)}{\obsT_1} \eqsim_\rho \notfinished{(\rho, \obs)}{\obsT_2}\)}
-\end{minipage}
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
-           {\(\notfinished{(\rho,\tau)}{\obsT_1} \eqsim_\rho \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
-           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho,\tau)}{\obsT_2}\)}
-\end{minipage}
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
-           {\(\notfinished{(\rho',\obs)}{\obsT_1} \eqsim_\rho \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
-           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho',\obs)}{\obsT_2}\)}
-\end{minipage}
-
-\section{Stack Safety With Calls and Returns}
-
-In this section we will present the context steps for allocations,
-calls, and returns, and show how they
-translate into properties. They are given formally in Figure \ref{fig:callcontexts}.
-Allocations are simple: from the current stack pointer, a range up to the size of
-the allocation is marked \(\mathit{object}(\phi)\) for a fresh identifier \(\phi\).
-[TODO: actually generate fresh \(\phi\) by keeping counter...]
-[TODO: need to put the original \(\phi\) somewhere. Maybe macro-instruction
-  for offsetting from stack pointer? Ugh.]
-
-Alloc macro-instructions model allocations that might, or might not, be
-shared. A function can always allocate more opjects by adjusting its
-stack pointer outside of a macro-instruction, but this will not be reflected
-in the context.
-
-First, we define a function, {\it update view}, that defines the view of a callee
-in terms of that of the callee, together with the provenance map.
-First, all stack locations below the stack pointer
-are \(\unsealed\) -- fair game for the callee to allocate or write to, and
-to read after writing. Above the stack pointer, anything that is not an object
-becomes \(\sealed\). Argument and system registers are automatically marked
-\(\mathit{public}\), and all others are \(\sealed\).
-Finally, we must address stack objects: we recursively identify all objects
-that are reachable via pointer provenances found in \(\mathit{public}\)
-registers and in reachable objects, and these retain their view, while
-all others become \(\sealed\). [TODO: implement \(\mathit{reach}\) to do this.]
-
-To update the context, we pair the updated view with the return target of the
-call: the return target tells us whether a return is to the proper location
-which the stack pointer restored, and the view tells us how to treat the
-security of the callee's state. We define \(\mathit{hd\_def}\) as taking
-a context and returning its head, or \(((0,0),\lambda \component.\unsealed)\)
-if the context is empty. [TODO: structure it so that we don't need to return
-  a dummy target.]
+I submit that tail calls will work perfectly fine with all of our definitions
+using the context update in Figure \ref{fig:tailcallcontexts}. All of the
+``until return'' stuff works fine, because all the calls return at once.
+We need to define callee confidentiality still.
 
 \begin{figure*}
-\judgmenttwobrlong[AllocC]
-                  {\(\psi = \mathbf{alloc} ~ size\)}
-                  {\((T,V) = (hd\_def ~ \context)\)}
-                  {\(V' = V\llbracket m[sp]+i \mapsto \mathit{object}(\phi) | 0 < i \leq size \rrbracket\)}
-                  {\(\mach, \prov, \context \constep{\rho;\psi} ctxs[\rho \mapsto (T,V')::(tl\_def ~ \context]\)}
 
-\[\begin{split}
-& \mathit{update\_view} ~ V ~ \prov ~ a_{sp} ~ \overline{r_{args}} \triangleq
-\lambda \component .
-\begin{cases}
-  \mathit{public} & \textnormal{else if } \component \in \REGS \textnormal{ and }
-  \component \in \overline{r_{args}} \textnormal{ or } \component \in \mathit{PUBLIC} \\
-  \mathit{object}(\phi) & \textnormal{else if } (V ~ \component) = \mathit{object}(\phi)
-  \textnormal{ and } \phi \in \mathit{reach} ~ \prov ~ V ~ \overline{r_{args}} \\
-  \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } a_{sp} < \component \leq a_{sbase} \\
-  \unsealed & \textnormal{otherwise} \\
-\end{cases} \\
-\end{split}\]
+\judgmenttwo[TailCallC]
+            {\(\psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
+            {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ \overline{r_{args}}\)}
+            {\(\mach, (T,V)::\context \constep{\psi} (T,V')::\context\)}
 
-\judgmenttwobrlong[CallC]
-                  {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}}\)}
-                  {\(\context' = (\mathit{Return} ~ (m[pc] + 4) ~ m[sp], V)::\context\)}
-                  {\(V = \mathit{update\_view} ~ (snd ~ \context) ~ m[sp] ~ \overline{r_{args}}\)}
-                  {\(\mach, \prov, \context \constep{\rho;\psi} \context'\)}
-
-\judgment[ReturnCCons]
-         {\(j = \mathbf{return}\)}
-         {\(m, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\rho;\psi} \context\)}
-
-\judgment[ReturnCNil]
-         {\(j = \mathbf{return}\)}
-         {\(m, [] \constep{\rho;\psi} []\)}
-
-\caption{Call, enter, and return macro-instructions}
-\label{fig:callcontexts}
+\caption{Tailcall context updates}
+\label{fig:tailcallcontexts}
 \end{figure*}
 
-Now we can give our definition of stack safety in terms of the context at the point of a call.
+\section{Provenance, Capabilities, and Protecting Objects}
 
-\subsection{Facts Abouts Calls and Returns}
+So far, we have given functions the ability to allocate completely unprotected
+objects. What if we want to express a finer-grained notion of safety, in which
+such objects are protected unless the function that owns them intentionally
+passes a pointer to them?
 
-Here we define some logical operations to reason about the behavior of the
-system over time. These have a temporal-logic flavor, as they reflect
-the expected behavior of the system in the future, after a possible return.
+In order to express such a property, we need our machine to carry some notion
+of {\it pointer provenance} -- a distinction between a pointer that is intended to
+point to a given object, and non-pointer integers as well as pointers to other objects.
 
-\paragraph*{On-return}
-
-The intuition behind {\it return-time integrity} (below) is that a caller may expect its
-sealed data to be unchanged when control returns to it. In fact, the callee
-may overwrite such data -- when the data are found in callee-saved registers
-this is perfectly legal -- as long as it either restores it, or has some guarantee
-that its changes will impact the caller.
-
-We start by defining a second-order logical operator
-\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
-where \(P\) is a predicate on machine states. This is a coinductive relation
-similar to ``weak until'' in temporal logic -- it holds if the condition
-is never reached.
-
-\judgment{\(P ~ \mach\)}
-         {\((\context \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
-
-\judgmenttwobrlong{\(\context_0 \not = \context\)}
-                  {\((\context_0 \uparrow P) ~ (\mach', \pol', \prov', \context')\)}
-                  {\((\mach, \pol, \prov, \context) \stepstounder{\rho,\psi?}
-                    (\mach', \pol', \prov', \context')\)}
-                  {\((\context_0 \uparrow P) ~ (\mach, \pol, \prov, \context)\)}
-
-Similarly, in {\it caller confidentiality}, we will want to compare future states,
-so we give a binary equivalent, \(\context \Uparrow R\), so that
-\(\sigma (\context \Uparrow R) \sigma'\) holds if \(R\) holds on the
-first return states of after \(\sigma\) and \(\sigma'\).
-
-\judgment[Returned]
-         {\(\mach ~ R ~ \mach'\)}
-         {\((\mach,\pol,\prov,\context) ~ (\context \Uparrow R) ~ (\mach',\pol',\prov',\context)\)}
-
-\judgmentthree[Left]
-              {\(\context \not = \pi_c(\sigma_1)\)}
-              {\(\sigma_1 \stepstounder{\rho;\psi?} \sigma_1'\)}
-              {\(\sigma_1' ~ (\context \Uparrow R) ~ \sigma_2\)}     
-              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
-
-\judgmentthree[Right]
-              {\(\context \not = \pi_c(\sigma_2)\)}
-              {\(\sigma_2 \stepstounder{\rho;\psi?} \sigma_2'\)}
-              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2'\)}     
-              {\(\sigma_1 ~ (\context \Uparrow R) ~ \sigma_2\)}
-
-\paragraph*{Variants and Safe Elements}
-
-One or more elements of a given state are {\it safe} if their values
-in that state have no observable influence on future execution. We define this
-formally as a sort of non-interference: for a state \(\sigma\) and
-a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
-holds if we can replace the contents of \(\components\) with arbitrary values
-and have the resulting state produce an equivalent trace. The state created
-by modifying \(\components\) in this way is termed a variant.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
-
-\definition Full states \((\mach,\pol,\prov,\context)\) and \((\nach,\pol,\prov,\context)\)
-are also {\em \(\components\)-variants} if \(\mach\) and \(\nach\) are.
-
-\definition An element set \(\components\) is safe in state \(\sigma)\),
-written \(\components ~ \mathit{SAFE} ~ \sigma\), if for all
-\(\sigma'\) such that \(\sigma \approx_{\components'} \sigma'\), if 
-\(\sigma \hookrightarrow \obsT\) and
-\(\sigma' \hookrightarrow \obsT'\), then
-\(\obsT \eqsim \obsT'\).
-
-\paragraph*{Property Definitions}
-
-And with the preliminary definitions out the way, we can define integrity.
-
-\definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
-such that \(\mach[\component] \not = \mach'[\component]\).
-
-\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) and
-\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component = \sealed\}\).
-Let \(P\) be a predicate on states \(\sigma'\) that holds if
-\(\Delta(\mach, \pi_m(\sigma')) \cap \components ~ \mathit{SAFE} ~ \sigma'\).
-Then \(\sigma\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.
-
-\definition A system enjoys {\it stack integrity} if, for all reachable states \(\sigma\) such that
-\(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
-\(\sigma'\) enjoys return-time integrity.
-
-\paragraph*{Caller Confidentiality}
-
-Confidentiality properties are also modeled as non-interference, but with a twist:
-the caller's confidential data should be kept secret from the callee, but after return,
-will naturally be accessible to the caller once again. So, we must model confidentiality
-in terms of events inside the callee. But confidentiality violations might not become
-visible during the callee's lifetime -- the callee might read a secret and stash it somewhere
-that it will be accessed later. So we must also consider the machine state at the end
-of the call.
-
-We could define a natural, strict form of confidentiality that requires that the final
-states be identical. However, as with integrity, this is stricter than we need -- there
-may be some conditions in which a secret is copied, but never used again. So, we once
-again wish for our final condition to make the values {\it safe}. Which values?
-
-\definition Let \(\mach,\mach'\) and \(\nach,\nach'\)
-be pairs of machine states. Their {\em corrupted set}, written
-\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all elements
-\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) such that
-\(\mach'[\component] \not = \nach'[\component]\).
-
-\definition Let \(\sigma = (\mach,\pol,\prov,\context)\) be a state and
-\(\components = \{\component | \mathit{hd\_def} ~ \context ~ \component \in
-\{\sealed, \unsealed\}\}\).
-
-Then \(\sigma\) enjoys {\it caller confidentiality} if, for any \(\sigma'\)
-that is a \(\components\)-variant of \(\sigma\), we can take
-\(\context \uparrow \sigma \hookrightarrow \obsT\) and
-\(\context \uparrow \sigma' \hookrightarrow \obsT'\) and have that
-\(\obsT \simeq \obsT'\). And if, additionally, \(\sigma ~ (\context \Uparrow R) ~ \sigma'\),
-where \(\sigma_1 ~ R ~ \sigma_2\) if
-\(\bar{\Diamond}(\mach,\pi_m(\sigma_1),\pi_m(\sigma'),\pi_m(\sigma_2)) ~ \mathit{SAFE} ~ \sigma_1\).
-
-\definition A system enjoys {\it universal caller confidentiality} if, for all reachable states
-\(\sigma\) such that \(\sigma \stepstounder{\rho;\mathbf{Call}\dots} \sigma'\),
-\(\sigma'\) enjoys caller confidentiality.
+\section{Simple Concurrency}
 

### 2022-09-05 submission5/securityoverlays.tex
Playing with security overlay concept

\documentclass{article}

\usepackage{geometry}
\usepackage{amsmath, amssymb}
\usepackage{mathtools}

\begin{document}

\section{Security Overlay Semantics (Concrete)}

A {\it security overlay semantics} divides a program's execution into
{\it principals}, and specifies what it means for such principals to be
protected from one another by a high-level abstraction implemented
in a concrete machine. Specifications may take the form of predicates on
future states: statements of the form, ``When control returns to me...'',
or of relations on traces of future execution (hyper-properties.)

What this means in practice is that we augment a concrete machine with
additional {\it overlay state}, which tracks information related to the
principals and their privileges, as well as which is active.
We identify some sequences of instructions
as implementing {\it overlay operations}. An overlay operation either
transfers control between principals, or modifies the active principal's
privileges. And we enrich the underlying machine with an {\it overlay step}
that modifies the overlay state accordingly.

Let \(\Psi\) be the set of overlay operations ranged over by \(\psi\).

We begin with the base machine, whose states are drawn from a set \(\mathcal{M}\),
and which transitions via partial step function \(\rightharpoonup\).
We define several operations on states. \(m[k]\) accesses the word in register
or address \(k\). \(m\langle PC \rangle\) maps a state to a pair of a word
and an optional indexed transition \((\psi,i)?\) found at the address pointed to by the
program counter.

If \(m\langle PC \rangle = (w, \bot)\) or \((w, (\psi,i))\) for \(i > 0),
then we step by \(\rightharpoonup\).
If \(m\langle PC \rangle = (w, (\psi,0))\), we instead step by \(\xRightarrow{\psi}\),
which executes until it reaches a state \(m'\) such that \(m\langle PC \rangle = (w',\psi?)\)
where \(\psi? = \bot\) or \((\psi',i)\) for \(\psi \not = \psi'\).

\subsection{Starting Overlay Operations in the Middle}

An overlay operation can contain multiple instructions, and its security guarantees
may not hold if they are executed out of order. In fact, they almost certainly will not!
So we do not treat such executions as overlay operations, and our properties can be
statisfied even if such executions behave insecurely. It is the responsibility of the
enforcement mechanism to guarantee that this does not happen, for instance by tagging
the operation as a blessed sequence.

\subsection{The Compiler's Responsibility}

The code that is acted on by a security overlay semantics is precisely that code
that the compiler provides, down to any instruction tags that are needed to enforce
the associated micro-policy. The compiler is also responsible for labeling overlay
operations as such. As noted above, this puts the burden on the compiler to be aware
of the available overlay operations and use them intentionally.

In fact, in the strictest sense, the overlay semantics has {\it nothing} to say
about which sequences make up an operation. It is for the compiler to determine that
when it constructs the sequence. So, the compiler can make the sequence as short or
as long as it wants; there ends up not being much reason for a sequence longer than
a single instruction, except in the very strange case where the same sequence might be
entered from the start {\it or} in the middle, and the compiler wants to assert protections
only in the former case.

\subsection{Further Compilation}

The overlay semantics makes absolutely no changes to the behavior of the system,
so further compilation is not necessary.

\section{Security Overlay Semantics (Abstract)}

A {\it security overlay semantics} implements a high-level abstraction
using the instructions available to a concrete machine, and makes assertions
about its security. These assertions divide a program's execution into
{\it principals} and specify what it means for such principals to be
protected from one another by the high-level abstraction.
Specifications may take the form of predicates on
future states: statements of the form, ``When control returns to me...'',
or of relations on traces of future execution (hyper-properties.)

What this means in practice is that we augment a concrete machine with
additional {\it overlay state}, which tracks information related to the
principals and their privileges, as well as which is active.
We identify some sequences of instructions
as implementing {\it overlay transitions}. An overlay transition either
transfers control between principals, or modifies the active principal's
privileges. And we enrich the underlying machine with an {\it overlay step}
that executes the instructions of an overlay transition atomically
and modifies the overlay state accordingly.

Let \(\Psi\) be the set of overlay transitions ranged over by \(\psi\).

We begin with the base machine, whose states are drawn from a set \(\mathcal{M}\),
and which transitions via partial step function \(\rightharpoonup\).
We define several operations on states. \(m[k]\) accesses the word in register
or address \(k\). \(m\langle PC \rangle\) maps a state to a pair of a word
and an optional transition \(\psi?\) found at the address pointed to by the
program counter.

If \(m\langle PC \rangle = (w, \bot)\), then we step by \(\rightharpoonup\).
If \(m\langle PC \rangle = (w, (\psi,0))\), we instead step by \(\xRightarrow{\psi}\),
which executes until it reaches a state \(m'\) such that \(m\langle PC \rangle = (w',\psi?)\)
where \(\psi? = \bot\) or \((\psi',i)\) for \(\psi \not = \psi'\). We do not step
from any state where \(m\langle PC \rangle = (w, (\psi,i))\) for \(i > 0\).

\subsection{Starting Overlay Operations in the Middle}

An overlay operation can contain multiple instructions, and its security guarantees
may not hold if they are executed out of order. In fact, they almost certainly will not!
So we do not treat such executions as overlay operations, and our properties can be
statisfied even if such executions behave insecurely. It is the responsibility of the
enforcement mechanism to guarantee that this does not happen, for instance by tagging
the operation as a blessed sequence.

\subsection{The Compiler's Responsibility}

\subsection{Further Compilation / Protection}

A security overlay semantics uses the same instructions as the base machine,
with the same semantics, and protected by the same security mechanism. But it
is a more abstract machine in that it is also

\section{A Barely Abstract Stack Machine}

\subsection{The Compiler's Responsibility}

\subsection{Further Compilation / Protection}

\section{A Slightly More Abstract Stack Machine}

\subsection{The Compiler's Responsibility}

\subsection{Further Compilation / Protection}

\end{document}

### 2022-09-05 submission5/onlyformalism.tex

@@
 Otherwise, it proceeds by the default step. The definitions of context steps
-\(\constep{\psi} \in \CONTEXTS \times \Psi \rightarrow \CONTEXTS\)
+\(\constep{\psi} \in \MACHS \times \CONTEXTS \times \Psi \rightarrow \CONTEXTS\)
 will be defined for each pseudoinstruction as we introduce new topics.
@@
                     {\(\mach[\PCname+n] = \bar{i}[n]\) for each \(n\) where \(0 \leq n < |\bar{i}|\)}
-                    {\((\mach,\context)) \stepstounder{\psi} (\mach',\context')\)}
+                    {\((\mach,\context) \stepstounder{\psi} (\mach',\context')\)}
 
@@
   \unsealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } \component \leq a_{sp} \\
+  \unsealed & \textnormal{else if } \component \in \REGS \textnormal{ and } \component \in \mathit{CLR\_SV} \\
   \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } a_{sp} < \component \textnormal { and } V ~ \component = \unsealed \\
+  \sealed & \textnormal{else if } \component \in \REGS  \\
   V ~ \component & \textnormal{otherwise} \\
@@
 
-\definition An element set \(\components\) is vestigial in state \(\sigma)\),
+\definition An element set \(\components\) is vestigial in state \((\mach,\context)\),
 written \((\mach,\context) \parallel \components\), if for all
@@
 that is a \(\components\)-variant of \(\mach\), we can take
-\(\context \uparrow (\mach,\context) \hookrightarrow \obsT\) and
-\(\context \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
+\(\context_{ret} \uparrow (\mach,\context) \hookrightarrow \obsT\) and
+\(\context_{ret} \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
 \(\obsT \simeq \obsT'\);

### 2022-09-06 submission5/onlyformalism.tex
Activation ids

@@
 We begin with a RISC-V-like machine enhanced with PIPE, a tag-based reference
-monitory. We extend this machine with a {\it security overlay semantics}:
-a semantics in which calls, returns, and other operations of interest are condensed
-into atomic {\it macro-instructions}, which update an additional security context.
-As opposed to a fully-abstract overlay semantics, in which the semantics is safe
-by construction and compilation to an enforcement mechanism must be proved fully abstract,
-our machine's semantics are not safe unless the tag policy is correctly
-instantiated, but the security context informs trace properties that
-must hold under a policy for that policy to be considered valid.
+monitor. We extend this machine with a {\it security overlay semantics},
+which divides a program into dynamic {\it principals} that ought to be
+protected from one another, and specifies that protection over a set
+of abstract {\it overlay operations}. Specifications may take the form of
+predicates on future states: statements of the form,
+``When control returns to me...'', or of relations on traces of future execution
+(hyper-properties.)
+
+As opposed to a fully-abstract overlay semantics, in which overlay operations
+are given safe-by-construction semantics then proven fully abstract,
+our machine is not safe unless the monitor is correctly instantiated;
+the overlay gives us the means to distinguish a monitor that implements
+stack safety from one that does not.
 
 The underlying machine is a standard RISC-V ISA as modeled in [MIT-RISCV].
-Our overlay semantics provide macro-instructions for the following code features:
+Our overlay semantics provide overlay operations for the following code features:
 \begin{itemize}
@@
 \end{itemize}
-
-\paragraph*{What is a ``call''?} Some ISAs have a single ``call''
-instruction that does all the required manipulation of the program counter
-(\(\PCname\)) and stack pointer; others perform a series of instructions
-whose constituent opcodes may be used in some places for other purposes.
-Our RISC-V machine is the latter, so calls must be explicitly designated
-as such by the compiler, by identifying them as macro-instructions.
-The same goes for other operations, such as exceptions.
-Macro-instructions connect the enforcement mechanism to the property: we don't
-expect the enforcement mechanism to provide any guarantees for code sequences
-that resemble calls or returns, but which are not macro-instructions.
-
-Because macro-instructions must be generated by the compiler, they only
-carry data that can be known by the compiler at compile-time. For example,
-the \(\mathbf{call}\) macro-instruction identifies the register \(r_{target}\) containing
-the address of the function to be called, as well as the registers that hold arguments.
-
-%\paragraph*{Enforcement Model and Properties}
-
-%A machine state consists of the standard addresses and registers, as well as
-%additional policy state used by the enforcement mechanism. Our properties
-%refer only to addresses and registers; they do not constrain the behavior
-%of policy state. We model an execution step as a partial function on states:
-%when the machine does not step from a given state, this
-%is a {\em failstop}, halting the program before it can perform an
-%action that would violate the security property.
-%our properties are therefore naturally \emph{termination insensitive.}
 
@@
 
-\section{The Basic Machine}
+\subsection{Security Overlay Semantics}
+
+A security overlay semantics divides a program's execution into principals,
+in this case function activations, and identifies which principal is active
+in a given state. Overlay operations can generate new principals, transition
+which principal is active, and allocate stack objects belonging to the active
+principal. From this information, we derive the specifications that make up
+stack safety. Specifications may take the form of predicates on
+future states: statements of the form, ``When control returns to me...'',
+or of relations on traces of future execution (hyper-properties.)
+
+What this means in practice is that the compiler that generates a piece of
+code must identify some instructions as implementing the overlay operations.
+In the likely event that the operation requires multiple instructions,
+typically the {\it last} instruction will be labeled.
+
+Let \(\Psi\) be the set of overlay operations ranged over by \(\psi\).
+We begin with the base machine, whose states are drawn from a set \(\mathcal{M}\),
+and which transitions via partial step function \(\rightharpoonup\).
+We define the operation \(m[k]\) as accessing the word in register
+or at address \(k\) in state \(m\). The compiler provides a partial map
+\(\mathit{over} \in \mathcal{W} \rightharpoonup \Psi\).
+We lift \(\rightharpoonup\) into a labeled transition
+\(\harpoonunder{\psi?} \in \mathcal{M} \rightharpoonup
+\mathcal{M} \times (\Psi + \bot)\) by labeling each
+transition \(\mach \harpoonunder{\psi?} \mach'\) where
+\(\psi = \mathit{over} ~ \mach[\PCname]\) if it is defined and \(\bot\) if not.
+
+\paragraph{Starting Overlay Operations in the Middle}
+
+The concrete implementation of an overlay operation might contain multiple instructions,
+and its security guarantees may not hold if they are executed out of order.
+In fact, they almost certainly will not! But only one instruction is labeled.
+It is the responsibility of the monitor to guarantee that this instruction is
+executed at the appropriate place in its sequence.
+
+\paragraph{The Compiler's Responsibility}
+
+The code that is acted on by a security overlay semantics is precisely that code
+that the compiler provides, down to any instruction tags that are needed to enforce
+the associated micro-policy. The compiler is also responsible for labeling overlay
+operations as such. This puts the burden on the compiler to be aware
+of the available overlay operations and use them intentionally. We give a
+concrete example of one such labeling scheme in [Testing Section.]
+
+\subsection{The Basic Machine}
 
@@
 The base machine step function
-\(\mach \stepstounder{} \mach' \in \MACHS \rightharpoonup \MACHS\) represents a single step.
-It is partial to represent the capacity of the machine to failstop.
+\(\mach \stepstounder{\psi?} \mach' \in \MACHS \rightharpoonup \MACHS\) is partial to
+represent the capacity of the machine to failstop, and may be labeled with an overlay
+operation.
 In order to accommodate failstop behavior, our properties will be {\it termination insensitive}.
 
-The set of instructions, \(I \subseteq \WORDS\), corresponds to the RISC-V ISA.
-We further introduce {\it macro-instructions}: sequences of
-instructions that operate as if they were single, atomic pseudoinstructions.
-A macro-instruction consists of both a list
-of instructions and a pseudoinstruction that identifies its purpose in the
-property. The pseudoinstructions, and the data that they carry, are as follows.
+\subsection{Overlay Operations, Principals, and Overlay State}
+
+Overlay operations are drawn from a set \(\Psi\), as follows:
 
@@
 | & \mathbf{return} & \\
-| & \mathbf{alloc} ~ size & size \in \mathbb{N} \\
 | & \mathbf{setex} ~ \addr_{buf} \\
 | & \mathbf{throwex} ~ \addr_{buf} \\
+| & \mathbf{alloc} ~ sizes & sizes \in \mathit{list} ~ \mathbb{N} \\
 \end{align*}
 
-An macro-instruction \((\bar{i},\psi) \in J\) is a pair of a pseudoinstruction and
-a non-empty list of instructions. In order to implement the atomic nature of
-executing macro-instructions,
-we define a {\it ranged-step} relation that, given a pair of addresses
-\(a_{low}\) and \(a_{high}\), steps as long the program counter is between them.
+Calls and tailcalls identify the target of the call, and the argument registers.
+[Still a todo: stack arguments.] Return needs no further information. The operations
+that setup and then throw to an exception handler take the address at which it is
+stored. Alloc takes a list of sizes of objects to be allocated, because if allocating
+objects means simply moving the function pointer, multiple
+objects may be allocated at once.
 
-\judgment[Done]
-         {\(\mach[\PCname] < a_{low} \lor \mach[\PCname] > a_{high}\)}
-         {\(a_{low},a_{high} \vdash \mach \manystepstounder{} \mach\)}
+Principals are function activations \(\alpha \in A\), with an operation
+\(\mathit{fresh}\) that always generates a unique \(\alpha\).
+Overlay state consists of a context stack, which contains the security-relevant
+information associated with the active principal, and with inactive principals
+waiting for control to return to them.
 
-\judgmenttwobrlong[Step]
-                  {\(a_{low} \leq \mach[\PCname] < a_{high}\)}
-                  {\(\mach \stepstounder{} \mach'\)}
-                  {\(a_{low},a_{high} \vdash \mach' \manystepstounder{} \mach''\)}
-                  {\(a_{low},a_{high} \vdash \mach \manystepstounder{} \mach''\)}
+We define a set of {\it security classes}:
+\[sc \in SEC ::= \public | \sealed | \unsealed | \mathit{object}\]
 
-Now we're ready to introduce the overlay semantics.
-Additionally, we keep track of security context information
-specific to each stack over the course of the run.
-This data is not relevant to execution,
-it is only used to define the stack safety property. We define a set of
-{\it security classes}:
-\[sc \in SEC ::= \public | \sealed | \unsealed\]
-A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\) divides the machine
-elements in terms of their purpose relative to the active function.
+These indicate, from the perspective of the active principal, whether an element is
+always accessible (\(\public\)), claimed by an inactive principal (\(\sealed\)),
+available to be allocated and later sealed (\(\unsealed\)), or allocated by
+the active principal (\(\mathit{object}\)).
+
+A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\)
+assigns each element a security class.
+
 The {\it initial view} \(V_0\) maps all stack locations and non-public registers
 to \(\unsealed\) and all other locations and public registers to \(\public\).
-A set of {\it targets} identify the addresses of the program counter or
-stack pointer at a return point:
+
+We additionally define a set of {\it targets}, which identify how a
+principal expects to have control returned to it. This could be via a normal
+return to the correct instruction \(\addr_{pc}\), or by catching the exception
+whose state is saved at \(\addr_{buf}\). The \(\mathit{Active}\) target indicates
+the active principal.
+
 \[\begin{split}
-tar \in TAR ::= & \mathit{Return} ~ a_{pc} ~ a_{sp} \\
-& \mathit{Catch} ~ a_{buf} ~ a_{pc} ~ a_{sp} \\
-& \mathit{Done} \\
+tar \in TAR ::= & \mathit{Return} ~ a_{pc} \\
+| & \mathit{Catch} ~ a_{buf} ~ a_{pc} \\
+| & \mathit{Active} \\
 \end{split}\]
-And a {\it stack context} is a non-empty stack of pairs of targets and views:
-\[\context \in \CONTEXTS ::= \mathit{list} ~ (\mathit{TAR} \times \mathit{VIEW})\]
 
-The initial stack context \(\context_0\) is \(\left[ (\mathit{Done}, V_0) \right]\).
+Finally, a {\it context stack} is a non-empty stack of triples of an activation,
+a target, and a view. The initial context stack \(\context_0 = (\mathit{fresh},\mathit{Active},V_0)\),
+and the top target on the stack will always be \(\mathit{Active}\).
+\[\context \in \CONTEXTS ::= \mathit{list} ~ (A \times \mathit{TAR} \times \mathit{VIEW})\]
 
-An {\it overlay state} \(\sigma\) consists of a pair of a machine state
-and a stack contexts.
-\[\sigma = (\mach, \context) \in \mathcal{S} ::= \MACHS \times \CONTEXTS\]
+We define an update function for context stacks indexed by an overlay operation.
+This will be defined for each operation as we introduce them in detail.
+\[\constep{\psi} \in \MACHS \times \CONTEXTS \times \Psi \rightarrow \CONTEXTS\]
 
-The labeled step function for overlay states
-\(\sigma \stepstounder{\psi?} \sigma' \in
-\mathcal{S} \rightarrow  (\Psi + \bot) \times \mathcal{S}\)
-proceeds by an {\it overlay step}
-if the program counter points to the start of a macro-instruction's instruction sequence.
-Otherwise, it proceeds by the default step. The definitions of context steps
-\(\constep{\psi} \in \MACHS \times \CONTEXTS \times \Psi \rightarrow \CONTEXTS\)
-will be defined for each pseudoinstruction as we introduce new topics.
+From a machine state and a context stack, we create a {\it combined state} \((\mach,\context)\)
+and a transition that steps the context according to the label on the transition,
+if there is one, and otherwise maintains it.
 
-\judgmentthreebrlong[Overlay]
-                    {\((\bar{i}, \psi) \in J\)}
+\judgmenttwo[Overlay]
+            {\(\mach \harpoonunder{\psi} \mach'\)}
             {\(\mach',\context \constep{\psi} \context'\)}
-                    {\(\mach \manystepstounder{} \mach'\)}
-                    {\(\mach[\PCname+n] = \bar{i}[n]\) for each \(n\) where \(0 \leq n < |\bar{i}|\)}
             {\((\mach,\context) \stepstounder{\psi} (\mach',\context')\)}
@@
 \judgment[Default]
-         {\(\mach \stepstounder{} \mach'\)}
-         {\((\mach,\context) \stepstounder{\bot} (\mach',\context)\)}
+         {\(\mach \harpoonunder{\bot} \mach'\)}
+         {\((\mach,\context) \stepstounder{} (\mach',\context)\)}
 
-\paragraph*{Macro-instruction definitions}
+\section{Context Updates With Calls and Returns}
 
-[TBD]
+In this section we will present the overlay steps for calls and returns.
+They are given formally in Figure \ref{fig:callcontexts}.
 
-Note that there can be multiple macro-instructions with different instructions
-that represent the same pseudoinstruction.
-[This is less likely when we have smaller pseudoinstructions.]
+In order to understand the motivation behind the call rule, though, we must first
+describe how a function allocates objects. An allocation operation is parameterized
+by a list of sizes; for each entry in the list, it marks an appropriate range
+(based on offset from the stack pointer) as \(\mathit{object}\). This has no
+direct bearing on the active principal, which may continue using the memory,
+initialize it or not, and so on.
 
-\paragraph*{Events and Traces}
+For calls, we define a function, {\it update view}, that defines the view
+of a caller in terms of that of the callee. All argument registers are marked
+\(\public\), along with all actual public registers. Callee-save registers,
+including the stack pointer, are \(\sealed\); the callee is obligated to restore
+them to their original value. The remaining caller-save registers are \(\unsealed\).
+
+Then, all private objects are \(\sealed\), while the remainder of the stack retains
+its old security context.
+
+The formerly active principal is given a \(\mathit{Return}\) target, and on top of
+it in the context stack we push a principal with a fresh identifier, the \(\mathit{Active}\)
+target, and the updated view. On return, we search the stack for the first
+\(\mathit{Return}\) target, and restore that principal to active status with its
+original view. The rules given in Figure \ref{fig:callcontexts} take precedence
+in order.
+
+\begin{figure*}
+  \judgment[AllocCNil]
+           {\(\psi = \mathbf{alloc} ~ [ ~ ]\)}
+           {\(\mach, \context \constep{\psi} \context\)}
+
+  \judgmentthreebrtwo[AllocCCons]
+                     {\(\psi = \mathbf{alloc} ~ s::sizes\)}
+                     {\(\mach, (\alpha,T,V')::\context \constep{\psi'} \context'\)}
+                     {\(b = \mach[sp] + sum(sizes)\)}
+                     {\(\psi' = \mathbf{alloc} ~ sizes\)}
+                     {\(V' = V\llbracket \addr \mapsto \mathit{object} | b \leq \addr < b+s \land V ~ \addr = \unsealed \rrbracket\)}
+                     {\(\mach, (\alpha,T,V)::\context \constep{\psi} \context'\)}
+
+  \[\mathit{reg\_view} ~ \overline{r_{args}} \triangleq \lambda \reg .
+  \begin{cases}
+    \public & \textnormal{if } \reg \in \overline{r_{args}} \textnormal{ or } \reg \in \mathit{PUBLIC} \\
+    \unsealed & \textnormal{else if } \reg \in \mathit{CLR\_SV} \\
+    \sealed & \textnormal{else} \\
+  \end{cases}\]
+
+  \[\mathit{update\_view} ~ V ~ \overline{r_{args}} \triangleq \lambda \component
+  \begin{cases}
+    \mathit{reg\_view} ~ \overline{r_{args}} ~ \component & \textnormal{if } \component \in \REGS \\
+    \sealed & \textnormal{else if } V ~ \component = \mathit{object} ~ \mathbf{F} \\
+
+      V ~ \component & \textnormal{otherwise} \\
+  \end{cases}\]
+
+  \judgmenttwobr[CallC]
+              {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{\reg_{args}}\)}
+              {\(T = \mathit{Return} ~ (\mach[pc] + 4)\)}
+              {\(V' = \mathit{update\_view} ~ V ~ \overline{\reg_{args}}\)}
+              {\(\alpha' = \mathit{fresh}\)}
+              {\(\mach, (\alpha,\mathit{Active},V)::\context \constep{\psi}
+                (\alpha',\mathit{Active},V')::(\alpha,T,V)::\context\)}
+
+  \judgmenttwo[ReturnCRet]
+              {\(\psi = \mathbf{return}\)}
+              {\(T = \mathit{Return} ~ \_\)}
+              {\(\mach, (\_, \mathit{Active}, \_)::(\alpha, T, V)::\context
+                \constep{\psi} (\alpha, \mathit{Active}, V)::\context\)}
+
+  \judgmenttwo[ReturnCSkip]
+              {\(\psi = \mathbf{return}\)}
+              {\(\mach, \context \constep{\psi} \context'\)}
+              {\(\mach, (\alpha, \mathit{Active}, V)::\_::\context
+                \constep{\psi} (\alpha, \mathit{Active}, V)::\context\)}
+
+  \judgment[ReturnCNoRet]
+           {\(\psi = \mathbf{return}\)}
+           {\(\mach, [(\alpha,\mathit{Active},V)] \constep{\psi}
+             [(\alpha,\mathit{Active},V)]\)}
+
+\caption{Call and return context updates}
+\label{fig:callcontexts}
+\end{figure*}
+
+Now we can give our definition of stack safety in terms of the context at the point of a call.
+
+\subsection{Events and Traces}
 
@@
 A {\em trace} is a nonempty, finite or infinite sequence
-of pairs of processors and events, ranged over by \(\obsT\).
+of events, ranged over by \(\obsT\).
 We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
@@
 
-We write that execution from a state produces an observation trace \(\sigma \hookrightarrow \obsT\)
+We write that execution from a state produces an observation trace \(\mach,\context \hookrightarrow \obsT\)
 as follows, coinductively:
@@
 
-We define another relation that takes a trace of a callee up to its return, or until
-an exception is thrown, if either occurs.
-We write this \(\context' \downarrow (\mach,\context) \hookrightarrow \obsT\), where \(\context\)
+We define another relation that takes a trace of an activation until it is dead -- neither
+active, nor inactive and waiting for return.
+We write this \(\context' \downarrow (\mach,\context) \hookrightarrow \obsT\), where \(\context'\)
 is the context of the caller.
@@
 
-%\begin{minipage}{.4\columnwidth}
-%  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
-%           {\(\notfinished{(\rho',\obs)}{\obsT_1} \eqsim_\rho \obsT_2\)}
-%\end{minipage}
-%\begin{minipage}{.4\columnwidth}
-%  \judgment{\(\obsT_1 \eqsim_\rho \obsT_2\)}
-%           {\(\obsT_1 \eqsim_\rho \notfinished{(\rho',\obs)}{\obsT_2}\)}
-%\end{minipage}
-
-\section{Stack Safety With Calls and Returns}
-
-In this section we will present the context steps for calls and returns,
-and show how they translate into properties. They are given formally in
-Figure \ref{fig:callcontexts}.
-
-For calls, we define a function, {\it update view}, that defines the view
-of a caller in terms of that of the callee.
-First, all stack locations below the stack pointer
-are \(\unsealed\) -- fair game for the callee to allocate or write to, and
-to read after writing. All \(\unsealed\) locations above the stack pointer
-become \(\sealed\) -- their contents are both secret to and protected
-from the callee. Argument and system registers are automatically marked
-\(\mathit{public}\). All other locations above the stack pointer
-retain their prior status (these will only be \(\sealed\) for now,
-until we introduce shareable allocations shortly.)
-
-To update the context for a call, we pair the updated view with the return target of the
-call and push it to the stack. The return target will tell us whether a future
-return is to the proper location with the stack pointer restored.
-The view tells us how to treat the security of the callee's state, and as
-it becomes more complicated, will benefit from the history that we maintain.
-
-On a return, we pop the topmost pair from the context that has
-a \(\mathit{return}\), along with any other pairs (catch points,
-for instance) that we pass.
-
-\begin{figure*}
-\[\begin{split}
-& \mathit{update\_view} ~ V ~ a_{sp} ~ \overline{r_{args}} \triangleq
-\lambda \component .
-\begin{cases}
-  \public & \textnormal{else if } \component \in \REGS \textnormal{ and }
-  \component \in \overline{r_{args}} \textnormal{ or } \component \in \mathit{PUBLIC} \\
-  \unsealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } \component \leq a_{sp} \\
-  \unsealed & \textnormal{else if } \component \in \REGS \textnormal{ and } \component \in \mathit{CLR\_SV} \\
-  \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } a_{sp} < \component \textnormal { and } V ~ \component = \unsealed \\
-  \sealed & \textnormal{else if } \component \in \REGS  \\
-  V ~ \component & \textnormal{otherwise} \\
-\end{cases} \\
-\end{split}\]
-
-\judgmenttwobrlong[CallC]
-                  {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}}\)}
-                  {\(T' = \mathit{Return} ~ (m[pc] + 4) ~ m[sp]\)}
-                  {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ \overline{r_{args}}\)}
-                  {\(\mach, (T,V)::\context \constep{\psi} (T',V')::\context\)}
-
-\judgment[ReturnCRet]
-         {\(\psi = \mathbf{return}\)}
-         {\(\mach, (\mathit{return} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\psi} \context\)}
-
-\judgment[ReturnCDone]
-         {\(\psi = \mathbf{return}\)}
-         {\(m, [\mathit{Done},V] \constep{\psi} [\mathit{Done},V]\)}
-
-\judgmenttwo[ReturnCSkip]
-            {\(\psi = \mathbf{return}\)}
-            {\(\mach, \context \constep{\psi} \context'\)}
-            {\(\mach, (\_::\context) \constep{\psi} \context'\)}
-
-\caption{Call and return context updates}
-\label{fig:callcontexts}
-\end{figure*}
-
-Now we can give our definition of stack safety in terms of the context at the point of a call.
-
 \subsection{Facts Abouts Calls and Returns}
@@
 We start by defining a second-order logical operator
-\(\context_{ret} \uparrow P\), read ``\(P\) holds on return to \(\context_{ret}\),''
+\(\alpha \uparrow P\), read ``\(P\) holds on return to \(\alpha\),''
 where \(P\) is a predicate on machine states. This is a coinductive relation
@@
 
-\judgmenttwo[Returned]
-            {\(P ~ (\mach,\context)\)}
-            {\(|\context| \leq |\context_{ret}|\)}
-            {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
+\judgment[Returned]
+         {\(P ~ (\mach,(\alpha,T,V)::\context)\)}
+         {\((\alpha \uparrow P) ~ (\mach, (\alpha,T,V)::\context)\)}
 
 \judgmenttwobrlong[Step]
-                  {\(|\context| > |\context_{ret}|\)}
-                  {\((\mach, \context) \stepstounder{\psi?} (\mach', \context')\)}
-                  {\((\context_{ret} \uparrow P) ~ (\mach', \context')\)}
-                  {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
+                  {\(\alpha \not = \alpha'\)}
+                  {\((\alpha \uparrow P) ~ (\mach', \context')\)}
+                  {\((\mach, (\alpha',T,V)::\context) \stepstounder{\psi?} (\mach', \context')\)}
+                  {\((\alpha \uparrow P) ~ (\mach, (\alpha',T,V)::\context)\)}
 
 Similarly, in {\it caller confidentiality}, we will want to compare future states,
-so we give a binary equivalent, \(\context \Uparrow R\), so that
-\((\mach,\context) ~ (\context'' \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
+so we give a binary equivalent, \(\alpha \Uparrow R\), so that
+\((\mach,\context) ~ (\alpha \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
 first return states after \((\mach,\context)\) and \((\mach',\context')\).
 
-\judgmentthree[Returned]
-              {\((\mach,\context) ~ R ~ (\mach',\context')\)}
-              {\(|\context| \leq |\context_{ret}|\)}
-              {\(|\context'| \leq |\context_{ret}|\)}
-              {\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\)}
+\judgment[Returned]
+         {\((\mach,(\alpha,T,V)::\context) ~ R ~ (\mach',(\alpha,T',V')::\context')\)}
+         {\((\mach,(\alpha,T,V)::\context) ~ (\alpha \Uparrow R) ~ (\mach',(\alpha,T',V')::\context')\)}
 
-\judgmenttwobr[Left]
-              {\(\context_1 > \context_{ret}\)}
-              {\((\mach_1,\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
-              {\(\context_2 > \context_{ret}\)}
-              {\((\mach_1',\context_1') ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
-              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
+\judgmenttwobrlong[Left]
+              {\(\alpha \not = \alpha'\)}
+              {\((\mach_1',\context_1') ~ (\alpha \Uparrow R) ~ (\mach_2,\context_2)\)}
+              {\((\mach_1,(\alpha',T,V)::\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
+              {\((\mach_1,(\alpha',T,V)::\context_1) ~ (\alpha \Uparrow R) ~ (\mach_2,\context_2)\)}
 
-\judgmenttwobr[Right]
-              {\(\context_1 > \context_{ret}\)}
-              {\((\mach_2,\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
-              {\(\context_2 > \context_{ret}\)}
-              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2',\context_2')\)}
-              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
+\judgmenttwobrlong[Right]
+              {\(\alpha \not = \alpha'\)}
+              {\((\mach_1,\context_1) ~ (\alpha \Uparrow R) ~ (\mach_2',\context_2')\)}
+              {\((\mach_2,(\alpha',T,V)::\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
+              {\((\mach_1,\context_1) ~ (\alpha \Uparrow R) ~ (\mach_2,(\alpha',T,V)::\context_2)\)}
 
@@
 
-\judgmentbr[AllocC]
-           {\(\psi = \mathbf{alloc} ~ size\)}
-           {\(V' = V\llbracket m[sp]+i \mapsto \public | 0 < i \leq size \rrbracket\)}
-           {\((\mach, (T,V)::\context) \constep{\psi} (T,V')::\context\)}
-
 The remainder of our property definitions are unchanged. The \(\mathit{update\_view}\)

### 2022-09-11 submission5/onlyformalism.tex
Pass through with identifiers

@@
 
-\section{Context Updates With Calls and Returns}
+\section{Context Updates With Calls, Returns, and Private Allocations}
 
@@
     \mathit{reg\_view} ~ \overline{r_{args}} ~ \component & \textnormal{if } \component \in \REGS \\
-    \sealed & \textnormal{else if } V ~ \component = \mathit{object} ~ \mathbf{F} \\
+    \sealed & \textnormal{else if } V ~ \component = \mathit{object} \\
 
@@
 
-Now we can give our definition of stack safety in terms of the context at the point of a call.
-
-\subsection{Events and Traces}
+\section{Events and Traces}
 
@@
 
-We define another relation that takes a trace of an activation until it is dead -- neither
-active, nor inactive and waiting for return.
+We define another relation that takes a trace until we have returned to a particular
+activation, or past it (as might occur when we add exceptions.)
 We write this \(\context' \downarrow (\mach,\context) \hookrightarrow \obsT\), where \(\context'\)
@@
 
-\subsection{Facts Abouts Calls and Returns}
+\section{Facts Abouts Calls and Returns}
 
@@
 We start by defining a second-order logical operator
-\(\alpha \uparrow P\), read ``\(P\) holds on return to \(\alpha\),''
+\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
 where \(P\) is a predicate on machine states. This is a coinductive relation
@@
 
-\judgment[Returned]
-         {\(P ~ (\mach,(\alpha,T,V)::\context)\)}
-         {\((\alpha \uparrow P) ~ (\mach, (\alpha,T,V)::\context)\)}
+\judgmenttwo[Returned]
+            {\(|\context| \leq |\context_{ret}| }
+            {\(P ~ (\mach,\context)\)}
+            {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
 
 \judgmenttwobrlong[Step]
-                  {\(\alpha \not = \alpha'\)}
-                  {\((\alpha \uparrow P) ~ (\mach', \context')\)}
-                  {\((\mach, (\alpha',T,V)::\context) \stepstounder{\psi?} (\mach', \context')\)}
-                  {\((\alpha \uparrow P) ~ (\mach, (\alpha',T,V)::\context)\)}
+                  {\(|\context| \leq |\context_{tar}|\)}
+                  {\((\context_{ret} \uparrow P) ~ (\mach', \context')\)}
+                  {\((\mach, \context) \stepstounder{\psi?} (\mach', \context')\)}
+                  {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
 
 Similarly, in {\it caller confidentiality}, we will want to compare future states,
-so we give a binary equivalent, \(\alpha \Uparrow R\), so that
-\((\mach,\context) ~ (\alpha \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
+so we give a binary equivalent, \(\context \Uparrow R\), so that
+\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
 first return states after \((\mach,\context)\) and \((\mach',\context')\).
 
-\judgment[Returned]
-         {\((\mach,(\alpha,T,V)::\context) ~ R ~ (\mach',(\alpha,T',V')::\context')\)}
-         {\((\mach,(\alpha,T,V)::\context) ~ (\alpha \Uparrow R) ~ (\mach',(\alpha,T',V')::\context')\)}
+\judgmenttwobrlong[Returned]
+            {\(|\context| \leq |\context_{ret}|\)}
+            {\(|\context'| \leq |\context_{ret}|\)}
+            {\((\mach,\context) ~ R ~ (\mach',\context')\)}
+            {\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\)}
 
 \judgmenttwobrlong[Left]
-              {\(\alpha \not = \alpha'\)}
-              {\((\mach_1',\context_1') ~ (\alpha \Uparrow R) ~ (\mach_2,\context_2)\)}
-              {\((\mach_1,(\alpha',T,V)::\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
-              {\((\mach_1,(\alpha',T,V)::\context_1) ~ (\alpha \Uparrow R) ~ (\mach_2,\context_2)\)}
+              {\(|\context_1| \leq |\context_{ret}|\)}
+              {\((\mach_1,\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
+              {\((\mach_1',\context_1') ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
+              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
 
 \judgmenttwobrlong[Right]
-              {\(\alpha \not = \alpha'\)}
-              {\((\mach_1,\context_1) ~ (\alpha \Uparrow R) ~ (\mach_2',\context_2')\)}
-              {\((\mach_2,(\alpha',T,V)::\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
-              {\((\mach_1,\context_1) ~ (\alpha \Uparrow R) ~ (\mach_2,(\alpha',T,V)::\context_2)\)}
+              {\(|\context_2| \leq |\context_{ret}|\)}
+              {\((\mach_2,\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
+              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2',\context_2')\)}
+              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
 
-\paragraph*{Variants and Vestigial Elements}
+\section{Properties}
 
+We now finally have everything we need to offer our definitions for stack safety.
+We need one preliminary definition first.
 One or more elements of a given state are {\it vestigial} if their values
@@
 written \((\mach,\context) \parallel \components\), if for all
-\(\nach\) such that \(\mach \approx_{\components'} \mach'\), if 
+\(\nach\) such that \(\mach \approx_{\components} \mach'\), if 
 \((\mach,\context) \hookrightarrow \obsT\) and
@@
 
-\paragraph*{Property Definitions}
+\subsection{Integrity}
 
@@
 
-\definition Let \((\mach,(T,V)::\context)\) be a state and
+\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
 \(\components = \{\component | V ~ \component = \sealed\}\).
@@
 
-\definition A system enjoys {\it universal stack integrity} if, for all reachable states
+\definition A system enjoys {\it stack integrity} if, for all reachable states
 \((\mach,\context)\) such that
@@
 
-\paragraph*{Caller Confidentiality}
+\subsection{Caller Confidentiality}
 
@@
 
-\definition Let \((\mach,(T,V)::\context)\) be a state and
+\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
 \(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
 
-Then \((\mach,\context)\) enjoys {\it caller confidentiality} with respect to
+Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to
 some \(\context_{ret}\) if, for any \(\nach\)
@@
 \(\context_{ret} \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
-\(\obsT \simeq \obsT'\);
-and additionally if \((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\nach,\context)\),
+\(\obsT \simeq \obsT'\).
+
+\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
+\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
+
+Then \((\mach,\context)\) enjoys {\it return-time confidentiality} with respect to
+some \(\context_{ret}\) if \((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\nach,\context)\),
 where \((\mach_1,\context_1) ~ R ~ (\nach_2,\context_2)\) if
@@
 
-\definition A system enjoys {\it universal caller confidentiality} if, for all reachable states
+\definition A system enjoys {\it caller confidentiality} if, for all reachable states
 \((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
 with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
-\((\mach',\context')\) enjoys caller confidentiality with respect to \(\context\).
+\((\mach',\context')\) enjoys both internal and return-time confidentiality
+with respect to \(\context\).
 
-\section{Allocating Memory}
+\subsection{Callee Confidentiality}
 
-We now introduce the context step for the \(\mathbf{alloc} ~ \mathit{size}\)
-macro-instruction. Without \(\mathbf{alloc}\), our system already allowed a function
-to allocate as much private memory as it liked by moving the stack pointer. With
-\(\mathbf{alloc}\) it can now also allocate public objects that can be accessed
-freely by its callees. Later we will refine this system to restrict public objects
-to be accessed only via capabilities.
+Callees can have secrets too. The question is, what do we consider a callee's secrets?
+Anything in its frame? Anything that {\it could} be in its frame (i.e. above the stack pointer)?
+These are plausible options, but they will also capture data that has been written to those
+locations before the call. To get a narrower property specific to the data that is actually
+derived from the callee, we focus instead on those values that the callee has written.
 
-The remainder of our property definitions are unchanged. The \(\mathit{update\_view}\)
-function preserves the public status of allocated objects, and they are exempt from
-integrity and confidentiality properties.
+\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
+\(\components = \{\component | V ~ \component \in \{\sealed,\unsealed\}\}\).
+Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
+\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
+Then \(\mach\) enjoys {\it post-return confidentiality} if \(\context \downarrow P\) holds.
+
+\definition A system enjoys {\it callee confidentiality} if, for all reachable states
+\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
+with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
+\((\mach',\context')\) enjoys post-return confidentiality
+with respect to \(\context\).
+
+Interestingly, callee confidentiality subsumes caller integrity. This makes sense:
+one way for the callee's secrets to leak is by overwriting the caller's sealed data
+with them.
+
+\section{Public Memory}
+
+We can extend our basic machine to support public memory allocations that can
+be accessed by anyone until they are deallocated on their owner's return.
+We replace \(\mathbf{alloc} ~ \mathit{sizes}\) with \(\mathbf{alloc} ~ \mathit{objs}\),
+where \(\mathit{objs}\) is a list of pairs of sizes and booleans, representing whether
+the object in question should be public. The \(\mathit{update\_view}\)
+function only seals private objects, and the remained maintain their \(\mathit{object}\)
+class, making them exempt from both integrity and confidentiality properties.
 
@@
 popping it as well. Rules apply in order.
+[TODO: update these for the new \(\alpha\) stuff.]
 
@@
             {\(\psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
-            {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ \overline{r_{args}}\)}
-            {\(\mach, (T,V)::\context \constep{\psi} (T,V')::\context\)}
+            {\(V' = \mathit{update\_view} ~ V ~ \overline{r_{args}}\)}
+            {\(\mach, (\alpha,\mathit{Active},V)::\context \constep{\psi} (\alpha,\mathit{Active},V')::\context\)}
 
@@
 point to a given object, and non-pointer integers as well as pointers to other objects.
+Then, in \(\mathit{update\_view}\), we recursively identify all public objects that
+are reachable transitively from capabilities in registers; these keep their
+\(\mathit{object}\) status, while those that are not reachable become \(\sealed\).
 

### 2022-09-11 submission5/onlyformalism.tex
Changing frmaing

--- 2022-09-06 submission5/onlyformalism.tex
@@
 We begin with a RISC-V-like machine enhanced with PIPE, a tag-based reference
-monitor. We extend this machine with a {\it security overlay semantics},
+monitor. We extend this machine with a {\it security semantics},
 which divides a program into dynamic {\it principals} that ought to be
 protected from one another, and specifies that protection over a set
-of abstract {\it overlay operations}. Specifications may take the form of
+of abstract {\it security operations}. Specifications may take the form of
 predicates on future states: statements of the form,
@@
 our machine is not safe unless the monitor is correctly instantiated;
-the overlay gives us the means to distinguish a monitor that implements
+the security semantics gives us the means to distinguish a monitor that implements
 stack safety from one that does not.
@@
 
-\subsection{Security Overlay Semantics}
+\subsection{Security Semantics}
 
-A security overlay semantics divides a program's execution into principals,
-in this case function activations, and identifies which principal is active
-in a given state. Overlay operations can generate new principals, transition
+We begin with a machine that supports a number of security operations,
+defined above. For purposes of specifying security, we do not need to
+know the semantics of these operations; we only need the machine to
+distinguish them from normal execution. We model the machine as
+a transition system in which some transitions are marked with
+security operations. We then define a security semantics for stack safety
+in terms of those steps.
+
+The security semantics divides a program's execution into principals,
+namely function activations, and identifies which principal is active
+in a given state. Security operations can generate new principals, transition
 which principal is active, and allocate stack objects belonging to the active
@@
 or of relations on traces of future execution (hyper-properties.)
-
-What this means in practice is that the compiler that generates a piece of
-code must identify some instructions as implementing the overlay operations.
-In the likely event that the operation requires multiple instructions,
-typically the {\it last} instruction will be labeled.
-
-Let \(\Psi\) be the set of overlay operations ranged over by \(\psi\).
-We begin with the base machine, whose states are drawn from a set \(\mathcal{M}\),
-and which transitions via partial step function \(\rightharpoonup\).
-We define the operation \(m[k]\) as accessing the word in register
-or at address \(k\) in state \(m\). The compiler provides a partial map
-\(\mathit{over} \in \mathcal{W} \rightharpoonup \Psi\).
-We lift \(\rightharpoonup\) into a labeled transition
-\(\harpoonunder{\psi?} \in \mathcal{M} \rightharpoonup
-\mathcal{M} \times (\Psi + \bot)\) by labeling each
-transition \(\mach \harpoonunder{\psi?} \mach'\) where
-\(\psi = \mathit{over} ~ \mach[\PCname]\) if it is defined and \(\bot\) if not.
-
-\paragraph{Starting Overlay Operations in the Middle}
-
-The concrete implementation of an overlay operation might contain multiple instructions,
-and its security guarantees may not hold if they are executed out of order.
-In fact, they almost certainly will not! But only one instruction is labeled.
-It is the responsibility of the monitor to guarantee that this instruction is
-executed at the appropriate place in its sequence.
-
-\paragraph{The Compiler's Responsibility}
-
-The code that is acted on by a security overlay semantics is precisely that code
-that the compiler provides, down to any instruction tags that are needed to enforce
-the associated micro-policy. The compiler is also responsible for labeling overlay
-operations as such. This puts the burden on the compiler to be aware
-of the available overlay operations and use them intentionally. We give a
-concrete example of one such labeling scheme in [Testing Section.]
 
@@
 %
-A {\em machine state} \(\mach \in \MACHS\) is a type that exposes a mapping
-\(\mach[\component]\) from state elements to values, and may also contain additional
-security state. For convenience, we abbreviate
-multiple updates \(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
+{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
+that defines a mapping \(\mach[\component]\) from state elements to values.
+The machine step relation
+\(\mach \harpoonunder{\psi?} \mach' \in \MACHS \times (\psi+\bot) \times \MACHS\)
+is left-total and may be labeled by a security operation, defined below.
+While the tag-based enforcement mechanism may cause the machine to
+failstop, we model this as the machine silently diverging by returning to the
+same state. Since failstops are not distinguished from normal execution,
+our properties will necessarily be {\it termination insensitive}.
+
+For convenience, we will abbreviate multiple updates of a mapping
+\(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
 as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
 are fomulae with free variables bound to sets in \(C\).
-The base machine step function
-\(\mach \stepstounder{\psi?} \mach' \in \MACHS \rightharpoonup \MACHS\) is partial to
-represent the capacity of the machine to failstop, and may be labeled with an overlay
-operation.
-In order to accommodate failstop behavior, our properties will be {\it termination insensitive}.
 
-\subsection{Overlay Operations, Principals, and Overlay State}
+\subsection{Security Operations, Principals, and Security State}
 
-Overlay operations are drawn from a set \(\Psi\), as follows:
+Security operations are split into two categories: {\it privilege operations}
+and {\it control operations}. Privilege operations are those that adjust the
+privileges of future callees by allocating private or public data and by saving
+catch points for possible exceptions. Allocations specify the offset from
+the stack pointer and size of the object. Exceptions are tied to the address
+of the buffer that will contain saved state.
+%
+\begin{align*}
+  \mathit{pop} \in \mathit{POPS} ::= & \mathbf{alloc} ~ \mathit{off} ~ sz & \mathit{off}, sz \in \mathbb{N} \\
+  | & \mathbf{alloc\_pub} ~ \mathit{off} ~ sz & \mathit{off}, sz \in \mathbb{N} \\
+  | & \mathbf{setex} ~ \addr_{buf} \\
+\end{align*}
 
+Control operations are those that exchange
+control between principals: calls (including tailcalls), returns, and thrown
+exceptions. Calls and tailcalls identify the target of the call, and the argument registers.
+[Still a todo: stack arguments.] Return needs no further information. Throwing
+an exception specifies the address of the saved buffer.
+%
 \begin{align*}
-\psi \in \Psi ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+  \mathit{cop} \in \mathit{COPS} ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
   | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
   | & \mathbf{return} & \\
-| & \mathbf{setex} ~ \addr_{buf} \\
-| & \mathbf{throwex} ~ \addr_{buf} \\
-| & \mathbf{alloc} ~ sizes & sizes \in \mathit{list} ~ \mathbb{N} \\
+  | & \mathbf{throw} ~ \addr_{buf} \\
 \end{align*}
+%
+Security principals are function activations, which we model as an {\em activation record}.
+Such a record will track the status of the principals data and the ways in which its callees
+may return to it. To track the former, we define a set of {\it security classes}:
+\[sc \in SEC ::= \sealed | \mathit{object} | \unsealed | \public\]
 
-Calls and tailcalls identify the target of the call, and the argument registers.
-[Still a todo: stack arguments.] Return needs no further information. The operations
-that setup and then throw to an exception handler take the address at which it is
-stored. Alloc takes a list of sizes of objects to be allocated, because if allocating
-objects means simply moving the function pointer, multiple
-objects may be allocated at once.
-
-Principals are function activations \(\alpha \in A\), with an operation
-\(\mathit{fresh}\) that always generates a unique \(\alpha\).
-Overlay state consists of a context stack, which contains the security-relevant
-information associated with the active principal, and with inactive principals
-waiting for control to return to them.
-
-We define a set of {\it security classes}:
-\[sc \in SEC ::= \public | \sealed | \unsealed | \mathit{object}\]
-
-These indicate, from the perspective of the active principal, whether an element is
-always accessible (\(\public\)), claimed by an inactive principal (\(\sealed\)),
-available to be allocated and later sealed (\(\unsealed\)), or allocated by
-the active principal (\(\mathit{object}\)).
+These indicate, from the perspective of a given principal, whether an element is
+private to an inactive principal (\(\sealed\)),
+in an allocated object that is accessible to the current principal (\(\mathit{object}\)),
+available to be allocated (\(\unsealed\)), or
+always accessible (\(\public\)).
 
@@
 assigns each element a security class.
-
 The {\it initial view} \(V_0\) maps all stack locations and non-public registers
@@
     \sealed & \textnormal{else if } V ~ \component = \mathit{object} ~ \mathbf{F} \\
-
     V ~ \component & \textnormal{otherwise} \\

### 2022-09-12 submission5/onlyformalism.tex
New model through IV

@@
 \begin{align*}
-  \mathit{pop} \in \mathit{POPS} ::= & \mathbf{alloc} ~ \mathit{off} ~ sz & \mathit{off}, sz \in \mathbb{N} \\
-  | & \mathbf{alloc\_pub} ~ \mathit{off} ~ sz & \mathit{off}, sz \in \mathbb{N} \\
+  \phi \in \Phi ::= & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+  | & \mathbf{alloc\_pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
   | & \mathbf{setex} ~ \addr_{buf} \\
@@
 \begin{align*}
-  \mathit{cop} \in \mathit{COPS} ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+  \chi \in X ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
   | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
@@
 %
+A security operation may consist of any number of privilege operations and at most
+one control operation.
+%
+\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (X \times \mathit{list ~ \Phi})\]
+
 Security principals are function activations, which we model as an {\em activation record}.
-Such a record will track the status of the principals data and the ways in which its callees
+Such a record will track the status of the principal's data and the ways in which its callees
 may return to it. To track the former, we define a set of {\it security classes}:
@@
 
-We additionally define a set of {\it targets}, which identify how a
-principal expects to have control returned to it. This could be via a normal
-return to the correct instruction \(\addr_{pc}\), or by catching the exception
-whose state is saved at \(\addr_{buf}\). The \(\mathit{Active}\) target indicates
-the active principal.
+An single principal might register multiple exception handlers over the
+course of its execution, each of which will identify the value that the
+program counter ought to have when an exception is thrown to it. We model
+this as a partial map from addresses to addresses.
+So an the activation record for the current activation is a pair of a view
+and such a map:
+\[(V,ex) \in \mathit{ACT} ::= \mathit{VIEW \times (\WORDS \rightharpoonup \WORDS)}\]
 
-\[\begin{split}
-tar \in TAR ::= & \mathit{Return} ~ a_{pc} \\
-| & \mathit{Catch} ~ a_{buf} ~ a_{pc} \\
-| & \mathit{Active} \\
-\end{split}\]
+An inactive principal additional carries the address at which the program counter
+should point in the event of a normal return.
+\[(V, \addr_{ret}, ex) \in \mathit{INACT} ::= \mathit{VIEW \times \WORDS \times (\WORDS \rightharpoonup \WORDS)}\]
 
-Finally, a {\it context stack} is a non-empty stack of triples of an activation,
-a target, and a view. The initial context stack \(\context_0 = (\mathit{fresh},\mathit{Active},V_0)\),
-and the top target on the stack will always be \(\mathit{Active}\).
-\[\context \in \CONTEXTS ::= \mathit{list} ~ (A \times \mathit{TAR} \times \mathit{VIEW})\]
+We represent a {\it call stack} as a pair of one active record and a list of inactive ones.
+The initial call stack is \(\context_0 = ((V_0,[ ~ ]), [ ~ ])\).
+\[(\alpha, \sigma) = \context \in \CONTEXTS ::= \mathit{ACT \times list ~ INACT}\]
 
-We define an update function for context stacks indexed by an overlay operation.
-This will be defined for each operation as we introduce them in detail.
-\[\constep{\psi} \in \MACHS \times \CONTEXTS \times \Psi \rightarrow \CONTEXTS\]
+Every security operation manipulates the call stack: privilege operations act on
+the active principal, and control operations push or pop the active principal.
+We define the security effects of each operation as a function \(Pop\) or \(Cop\),
+from an operation, a machine state, and a call stack to a new call stack.
+As we introduce more complex operations, we will separately define
+the curried functions \(Pop ~ \phi\) or \(Cop ~ \chi\) for each operation.
+They are combined in \(Op\), which acts on full operations as follows.
+\[Pop \in \Phi \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
+\[Cop \in X \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
+\[Op ~ \psi =
+\begin{cases}
+  \lambda \context ~ \mach . \context & \textnormal{when } \psi = [~] \\
+  Pop ~ \phi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = \phi::\overline{\phi} \\
+  Cop ~ \chi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = (\chi,\overline{\phi}) \\
+\end{cases}\]
 
-From a machine state and a context stack, we create a {\it combined state} \((\mach,\context)\)
-and a transition that steps the context according to the label on the transition,
-if there is one, and otherwise maintains it.
+A control operation may additionally produce an observable event, drawn
+from a set \(\mathit{EVENTS}\) and ranged over by \(e\).
+A function \(Ev\) maps operations and machine states to events. The
+silent event \(\tau\) is relevant to our definitions here, and the
+remainder will be described in Section \ref{sec:events}. 
+%
+\[Ev \in X \rightarrow \MACHS \rightarrow \mathit{EVENTS}\]
+%
+From a machine state and a context stack, we create a {\it combined state}
+\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
+labeled with an event. If the underlying machine transition is labeled by
+some operation \(\psi\), the combined step applies \(Op\) to the contents
+of \(\psi\) to determine the new call stack; otherwise the call stack is
+unchanged. If there is a control operation,
+the transition is labeled with its event, otherwise it is silent.
+Rules apply in order.
 
-\judgmenttwo[Overlay]
+\judgmentthree[Control]
               {\(\mach \harpoonunder{\psi} \mach'\)}
-            {\(\mach',\context \constep{\psi} \context'\)}
-            {\((\mach,\context) \stepstounder{\psi} (\mach',\context')\)}
+              {\(\psi = (\chi, \overline{\phi})\)}
+              {\(e = Ev ~ \chi ~ \mach\)}
+              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \mach ~ \context)\)}
+
+\judgment[NonControl]
+         {\(\mach \harpoonunder{\psi} \mach'\)}
+         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \mach ~ \context)\)}
 
@@
          {\(\mach \harpoonunder{\bot} \mach'\)}
-         {\((\mach,\context) \stepstounder{} (\mach',\context)\)}
+         {\((\mach,\context) \stepstounder{\tau} (\mach',\context)\)}
 
@@
 describe how a function allocates objects. An allocation operation is parameterized
-by a list of sizes; for each entry in the list, it marks an appropriate range
-(based on offset from the stack pointer) as \(\mathit{object}\). This has no
-direct bearing on the active principal, which may continue using the memory,
-initialize it or not, and so on.
+by an offset \(\mathit{off}\) and a size \(\mathit{sz}\). It will attempt to claim
+a region of size \(\mathit{sz}\) starting at an offset of \(\mathit{off}\).
+Claiming the region in this case means marking it as \(\mathit{object} ~ \mathbf{F}\),
+where the boolean indicates whether the region is public.
+
+\judgmenttwobrlong{\(b = \mach[sp] - \mathit{off}\)}
+                  {\(\context = ((V, ex), \sigma)\)}
+                  {\(V' = V \llbracket \addr \mapsto \mathit{object} ~ \mathbf{F} |
+                    b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+                  {\(Pop ~ (\mathbf{alloc} ~ \mathit{off ~ sz}) ~ \context ~ \mach \triangleq
+                    ((V',ex), \sigma)\)}
 
@@
 them to their original value. The remaining caller-save registers are \(\unsealed\).
-
 Then, all private objects are \(\sealed\), while the remainder of the stack retains
@@
 
-The formerly active principal is given a \(\mathit{Return}\) target, and on top of
-it in the context stack we push a principal with a fresh identifier, the \(\mathit{Active}\)
-target, and the updated view. On return, we search the stack for the first
-\(\mathit{Return}\) target, and restore that principal to active status with its
-original view. The rules given in Figure \ref{fig:callcontexts} take precedence
-in order.
+The formerly active principal's record is pushed onto the inactive list,
+and replaced by a new record for the callee. Its return target is the intruction
+immediately following that of the program counter before the call.
+The callee's view is updated from the caller's with {\it update\_view}.
+
+\judgmenttwo{\(\chi = \mathbf{call} ~ r_{target} ~ \overline{\reg_{args}}\)}
+            {\(V' = \mathit{update\_view} ~ V ~ \overline{\reg_{args}}\)}
+            {\(Cop ~ \chi ~ ((V,ex),\sigma) ~ \mach \triangleq
+                ((V',[~]),(V, \mach[\PCname]+4, ex)::\sigma)\)}
+
+On return, we restore the topmost inactive principal, if there is one.
+Else, we default to retaining the current principal, as we do not model
+the main function returning.
+
+\judgmenttwo[Ret]
+            {\(\chi = \mathbf{return}\)}
+            {\(\sigma = (V,\addr_{ret},ex)::\sigma'\)}
+            {\(Cop ~ \chi ~ (\_, \sigma) ~ \mach \triangleq
+              ((V,ex), \sigma)\)}
+
+\judgment[RetDef]
+         {\(\chi = \mathbf{return}\)}
+         {\(Cop ~ \chi ~ (\alpha, [~]) ~ \mach \triangleq
+           (\alpha, [~])\)}
 
 \begin{figure*}
-  \judgment[AllocCNil]
-           {\(\psi = \mathbf{alloc} ~ [ ~ ]\)}
-           {\(\mach, \context \constep{\psi} \context\)}
-
-  \judgmentthreebrtwo[AllocCCons]
-                     {\(\psi = \mathbf{alloc} ~ s::sizes\)}
-                     {\(\mach, (\alpha,T,V')::\context \constep{\psi'} \context'\)}
-                     {\(b = \mach[sp] + sum(sizes)\)}
-                     {\(\psi' = \mathbf{alloc} ~ sizes\)}
-                     {\(V' = V\llbracket \addr \mapsto \mathit{object} | b \leq \addr < b+s \land V ~ \addr = \unsealed \rrbracket\)}
-                     {\(\mach, (\alpha,T,V)::\context \constep{\psi} \context'\)}
-
   \[\mathit{reg\_view} ~ \overline{r_{args}} \triangleq \lambda \reg .
@@
 
-  \judgmenttwobr[CallC]
-              {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{\reg_{args}}\)}
-              {\(T = \mathit{Return} ~ (\mach[pc] + 4)\)}
-              {\(V' = \mathit{update\_view} ~ V ~ \overline{\reg_{args}}\)}
-              {\(\alpha' = \mathit{fresh}\)}
-              {\(\mach, (\alpha,\mathit{Active},V)::\context \constep{\psi}
-                (\alpha',\mathit{Active},V')::(\alpha,T,V)::\context\)}
-
-  \judgmenttwo[ReturnCRet]
-              {\(\psi = \mathbf{return}\)}
-              {\(T = \mathit{Return} ~ \_\)}
-              {\(\mach, (\_, \mathit{Active}, \_)::(\alpha, T, V)::\context
-                \constep{\psi} (\alpha, \mathit{Active}, V)::\context\)}
-
-  \judgmenttwo[ReturnCSkip]
-              {\(\psi = \mathbf{return}\)}
-              {\(\mach, \context \constep{\psi} \context'\)}
-              {\(\mach, (\alpha, \mathit{Active}, V)::\_::\context
-                \constep{\psi} (\alpha, \mathit{Active}, V)::\context\)}
-
-  \judgment[ReturnCNoRet]
-           {\(\psi = \mathbf{return}\)}
-           {\(\mach, [(\alpha,\mathit{Active},V)] \constep{\psi}
-             [(\alpha,\mathit{Active},V)]\)}
-
-\caption{Call and return context updates}
+\caption{View updates for calls}
 \label{fig:callcontexts}
@@
 
-We convert macro-instructions into events via the \(\mathit{observe}\)
+We convert control operations into events via the \(Ev\)
 function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.
@@
 \begin{figure*}
-\[\mathit{observe} ~ \psi? ~ m =
+\[\mathit{Ev} ~ \chi ~ m =
 \begin{cases}
-  \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
-  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
-  \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
-  & \textnormal{when } ~ \psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}} \\
-%  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
+  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
+  & \textnormal{when } ~ \chi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
+  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
+  & \textnormal{when } ~ \chi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}} \\
   \tau & \textnormal{otherwise} \\
@@
 
-We write that execution from a state produces an observation trace \(\mach,\context \hookrightarrow \obsT\)
-as follows, coinductively:
+We write that execution from a state produces an observation trace
+\(\mach,\context \hookrightarrow \obsT\) as follows, coinductively:
 
-\judgment{\(\not \exists \mach' ~ \context'. (\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
-         {\((\mach,\context) \hookrightarrow \tau\)}
-
-\judgmenttwo{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
+\judgmenttwo{\((\mach,\context) \stepstounder{e} (\mach',\context')\)}
             {\((\mach',\context') \hookrightarrow \obsT\)}
-            {\((\mach,\context) \hookrightarrow \notfinished{\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}
-
-Naturally, most states will admit many possible traces, depending on the order in which
-processors step.
+            {\((\mach,\context) \hookrightarrow \notfinished{e}{\obsT}\)}
 
@@
 activation, or past it (as might occur when we add exceptions.)
-We write this \(\context' \downarrow (\mach,\context) \hookrightarrow \obsT\), where \(\context'\)
-is the context of the caller.
+We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
+\(d\) is the depth of the original call.
 
-\judgment{\(|\context| \leq |\context'|\)}
-         {\(\context' \downarrow (\mach,\context) \hookrightarrow \tau\)}
+\judgment{\(|\sigma| \leq d\)}
+         {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \tau\)}
 
-\judgmenttwobrlong{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
-                  {\(|\context'| > |\context''|\)}
-                  {\(\context'' \downarrow (\mach',\context') \hookrightarrow \obsT\)}
-                  {\(\context'' \downarrow (\mach,\context) \hookrightarrow \notfinished{\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}
+\judgmenttwobrlong{\((\mach,(\alpha,\sigma)) \stepstounder{e} (\mach',\context')\)}
+                  {\(|\sigma| > d\)}
+                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
+                  {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \notfinished{e}{\obsT}\)}
 
@@
 We start by defining a second-order logical operator
-\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
+\(d \uparrow P\), read ``\(P\) holds on return to depth \(d\),''
 where \(P\) is a predicate on machine states. This is a coinductive relation
-similar to ``weak until'' in temporal logic -- it holds if the condition
-is never reached.
+similar to ``weak until'' in temporal logic -- it holds if the program never
+returns to depth \(d\).
 
 \judgmenttwo[Returned]
-            {\(|\context| \leq |\context_{ret}| }
-            {\(P ~ (\mach,\context)\)}
-            {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
+            {\(|\sigma| \leq d }
+            {\(P ~ (\mach,(\alpha,\sigma))\)}
+            {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}
 
 \judgmenttwobrlong[Step]
-                  {\(|\context| \leq |\context_{tar}|\)}
-                  {\((\context_{ret} \uparrow P) ~ (\mach', \context')\)}
-                  {\((\mach, \context) \stepstounder{\psi?} (\mach', \context')\)}
-                  {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}
+                  {\(|\sigma| > d\)}
+                  {\((d \uparrow P) ~ (\mach', \context')\)}
+                  {\((\mach, (\alpha,\sigma)) \stepstounder{e} (\mach', \context')\)}
+                  {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}
 
 Similarly, in {\it caller confidentiality}, we will want to compare future states,
-so we give a binary equivalent, \(\context \Uparrow R\), so that
-\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
-first return states after \((\mach,\context)\) and \((\mach',\context')\).
+so we give a binary equivalent, \(d \Uparrow R\), so that
+\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
+first states that return to depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).
 
 \judgmenttwobrlong[Returned]
-            {\(|\context| \leq |\context_{ret}|\)}
-            {\(|\context'| \leq |\context_{ret}|\)}
-            {\((\mach,\context) ~ R ~ (\mach',\context')\)}
-            {\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\)}
+            {\(|\sigma_1| \leq d\)}
+            {\(|\sigma_2| \leq d\)}
+            {\((\mach_1,(\alpha_1,\sigma_1)) ~ R ~ (\mach_2,(\alpha_2,\sigma_2))\)}
+            {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
 
 \judgmenttwobrlong[Left]
-              {\(|\context_1| \leq |\context_{ret}|\)}
-              {\((\mach_1,\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
-              {\((\mach_1',\context_1') ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
-              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
+              {\(|\sigma_1| > d\)}
+              {\((\mach_1,(\alpha_1,\sigma_1)) \stepstounder{e} (\mach_1',\context_1')\)}
+              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
+              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
 
 \judgmenttwobrlong[Right]
-              {\(|\context_2| \leq |\context_{ret}|\)}
-              {\((\mach_2,\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
-              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2',\context_2')\)}
-              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
+              {\(|\sigma_2| > d\)}
+              {\((\mach_2,(\alpha_2,\sigma_2)) \stepstounder{e} (\mach_2',\context_2')\)}
+              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
+              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
 

### 2022-09-15 submission5/noexceptions.tex
For meeting

--- 2022-09-12 submission5/onlyformalism.tex
@@
   CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
+  & \(rra\) & Return Address \\
+  & \(rsp\) & Stack Pointer \\  
   \hline
-  PUBLIC & \(rra\) & Return Address \\
-  & \(rsp\) & Stack Pointer \\
-  & \(pc\) & Program Counter \\
+  PUBLIC & \(pc\) & Program Counter \\
   \hline
@@
 that defines a mapping \(\mach[\component]\) from state elements to values.
-The machine step relation
-\(\mach \harpoonunder{\psi?} \mach' \in \MACHS \times (\psi+\bot) \times \MACHS\)
-is left-total and may be labeled by a security operation, defined below.
+The machine step function
+\(\mach \xrightarrow{\psi} \mach' \in \MACHS \rightarrow \MACHS \times \Psi\)
+may be labeled by a security operation, defined below.
 While the tag-based enforcement mechanism may cause the machine to
@@
 For convenience, we will abbreviate multiple updates of a mapping
-\(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
-as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
+\(\mach[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
+as \(\mach \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
 are fomulae with free variables bound to sets in \(C\).
@@
   | & \mathbf{alloc\_pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
-  | & \mathbf{setex} ~ \addr_{buf} \\
 \end{align*}
@@
 \begin{align*}
-  \chi \in X ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+  \chi \in \mathcal{X} ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
   | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
   | & \mathbf{return} & \\
-  | & \mathbf{throw} ~ \addr_{buf} \\
 \end{align*}
@@
 %
-\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (X \times \mathit{list ~ \Phi})\]
+\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (\mathcal{X} \times \mathit{list ~ \Phi})\]
 
-Security principals are function activations, which we model as an {\em activation record}.
-Such a record will track the status of the principal's data and the ways in which its callees
-may return to it. To track the former, we define a set of {\it security classes}:
-\[sc \in SEC ::= \sealed | \mathit{object} | \unsealed | \public\]
+Security principals are function activations: at any given time, the current
+activation, and some number of pending ones. Each activation will have a {\it view}
+of the system that maps each element to one of a set of {\it security classes}:
+\[sc \in SEC ::= \sealed | \unsealed | \public\]
+\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]
 
@@
 private to an inactive principal (\(\sealed\)),
-in an allocated object that is accessible to the current principal (\(\mathit{object}\)),
+%in an allocated object that is accessible to the current principal (\(\mathit{object}\)),
 available to be allocated (\(\unsealed\)), or
 always accessible (\(\public\)).
-
-A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\)
-assigns each element a security class.
 The {\it initial view} \(V_0\) maps all stack locations and non-public registers
@@
 
-An single principal might register multiple exception handlers over the
-course of its execution, each of which will identify the value that the
-program counter ought to have when an exception is thrown to it. We model
-this as a partial map from addresses to addresses.
-So an the activation record for the current activation is a pair of a view
-and such a map:
-\[(V,ex) \in \mathit{ACT} ::= \mathit{VIEW \times (\WORDS \rightharpoonup \WORDS)}\]
-
-An inactive principal additional carries the address at which the program counter
-should point in the event of a normal return.
-\[(V, \addr_{ret}, ex) \in \mathit{INACT} ::= \mathit{VIEW \times \WORDS \times (\WORDS \rightharpoonup \WORDS)}\]
-
-We represent a {\it call stack} as a pair of one active record and a list of inactive ones.
-The initial call stack is \(\context_0 = ((V_0,[ ~ ]), [ ~ ])\).
-\[(\alpha, \sigma) = \context \in \CONTEXTS ::= \mathit{ACT \times list ~ INACT}\]
+A pending activation additional carries the addresses at which the program counter
+and stack pointer should point in the event of a return.
+We represent a {\it call stack} as a pair of the current activation's view
+and a list of pending ones.
+The initial call stack is \(\context_0 = (V_0, [ ~ ])\).
+\[(V, \addr_{ret}, \addr_{sp}) \in \mathit{PEND} ::= \mathit{VIEW \times \WORDS \times \WORDS}\]
+\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ PEND}\]
 
@@
 \judgmentthree[Control]
-              {\(\mach \harpoonunder{\psi} \mach'\)}
+              {\(\mach \xrightarrow{\psi} \mach'\)}
               {\(\psi = (\chi, \overline{\phi})\)}
               {\(e = Ev ~ \chi ~ \mach\)}
-              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \mach ~ \context)\)}
+              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \context ~ \mach)\)}
 
 \judgment[NonControl]
-         {\(\mach \harpoonunder{\psi} \mach'\)}
-         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \mach ~ \context)\)}
+         {\(\mach \xrightarrow{\psi} \mach'\)}
+         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \context ~ \mach)\)}
 
-\judgment[Default]
-         {\(\mach \harpoonunder{\bot} \mach'\)}
-         {\((\mach,\context) \stepstounder{\tau} (\mach',\context)\)}
+\section{Calls, Returns, and Private Allocations}
 
-\section{Context Updates With Calls, Returns, and Private Allocations}
+In this section we consider a simple setting with calls, returns, and private
+allocations. Assume that \(\MACHS\) describe a RISC-V machine, where a transition
+\(\mach \xrightarrow{\Psi?} \mach'\) is labeled as follows.
+If \(\mach[\PCname]\) points to {\tt addi sp,sp,N}, \(\Psi? = \mathbf{alloc} ~ (0,N)\);
+since we are not yet allocating public objects,
+we don't need to distinguish objects from one another if they are allocated
+simultaneously. Similarly, if \(\mach[\PCname]\) points to {\tt jal off ra}
+and \(\mach'[\PCname] = a_f\) where \(a_f\) is the entry of a function
+\(f\) that uses argument registers \(\overline{r}\), 
+\(\Psi? = \mathbf{call} ~ f ~ \overline{r}\). And a step via the
+instruction {\tt jalr ra} is labeled \(\mathbf{return}\).
+This simplistic system serves to demonstrate our core properties.
 
-In this section we will present the overlay steps for calls and returns.
-They are given formally in Figure \ref{fig:callcontexts}.
+Figure \ref{fig:example} gives a running example
+of a function {\tt main} that takes a secret in its arguments and allocates
+an array of potentially sensitive data. It then calls another function {\tt f},
+and afterward may decide to {\tt publish} its secret based on the contents of
+its sensitive data.
 
-In order to understand the motivation behind the call rule, though, we must first
-describe how a function allocates objects. An allocation operation is parameterized
-by an offset \(\mathit{off}\) and a size \(\mathit{sz}\). It will attempt to claim
-a region of size \(\mathit{sz}\) starting at an offset of \(\mathit{off}\).
-Claiming the region in this case means marking it as \(\mathit{object} ~ \mathbf{F}\),
-where the boolean indicates whether the region is public.
+\begin{figure*}
+  \begin{subfigure}[t]{.5\textwidth}
+    {\tt
+      void main(int argc, char*argv) \{
 
-\judgmenttwobrlong{\(b = \mach[sp] - \mathit{off}\)}
-                  {\(\context = ((V, ex), \sigma)\)}
-                  {\(V' = V \llbracket \addr \mapsto \mathit{object} ~ \mathbf{F} |
+      ~ ~ int secret = argv[1];
+
+      ~ ~ int sensitive[4] = \{0\};
+
+      ~ ~ f();
+
+      ~ ~ if (sensitive == 42) \{
+
+      ~ ~ ~ ~ publish(secret);
+
+      ~ ~ \}
+
+      \}
+    }
+  \end{subfigure}
+  \begin{subfigure}[t]{.5\textwidth}
+    {\tt
+      0: addi sp,sp,-20
+
+      4: sw a1,16(sp)
+
+      8: sd zero,8(sp)
+
+      12: sd zero,0(sp)
+
+      16: jal 84 ra
+
+      20: lw a4,12(sp)
+
+      24: li a5,42
+
+      28: bne a4,a5,L
+
+      32: ld a0,16(sp)
+
+      36: jal 164 ra
+
+      L(40): nop
+
+      44: addi sp,sp,20
+
+      48: return
+
+    }
+  \end{subfigure}
+
+  \caption{Example: Main}
+  \label{fig:example}
+\end{figure*}
+
+At the start of execution (PC 0) the initial call stack is \((V_0, [~])\).
+Instruction 0 allocates 20 bytes starting at the stack pointer. This will have the
+effect of marking those twenty bytes \(\sealed\), assuming they were previously
+\(\unsealed\):
+
+\judgmentbr{\(b = \mach[\SP] - \mathit{off}\)}
+           {\(V' = V \llbracket \addr \mapsto \sealed |
              b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
-                  {\(Pop ~ (\mathbf{alloc} ~ \mathit{off ~ sz}) ~ \context ~ \mach \triangleq
-                    ((V',ex), \sigma)\)}
+           {\(Pop ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach \triangleq
+             (V',\sigma)\)}
 
-For calls, we define a function, {\it update view}, that defines the view
-of a caller in terms of that of the callee. All argument registers are marked
-\(\public\), along with all actual public registers. Callee-save registers,
-including the stack pointer, are \(\sealed\); the callee is obligated to restore
-them to their original value. The remaining caller-save registers are \(\unsealed\).
-Then, all private objects are \(\sealed\), while the remainder of the stack retains
-its old security context.
+This memory being \(\sealed\) has no effect on the current activation --
+the security guarantees apply at the start of the call. Our next security operation
+occurs when we step from the call at instruction 16.
 
-The formerly active principal's record is pushed onto the inactive list,
-and replaced by a new record for the callee. Its return target is the intruction
+On a call, the formerly active principal's record is pushed onto the inactive list,
+and replaced by a new record for the callee. Its return target is the instruction
 immediately following that of the program counter before the call.
-The callee's view is updated from the caller's with {\it update\_view}.
+The callee's view is updated from the caller's such that the argument
+registers are \(\public\) and any non-argument, caller-saved registers
+are \(\unsealed\). All callee-save registers are always \(\sealed\), and
+the program counter is always public, but this is already the case.
 
-\judgmenttwo{\(\chi = \mathbf{call} ~ r_{target} ~ \overline{\reg_{args}}\)}
-            {\(V' = \mathit{update\_view} ~ V ~ \overline{\reg_{args}}\)}
-            {\(Cop ~ \chi ~ ((V,ex),\sigma) ~ \mach \triangleq
-                ((V',[~]),(V, \mach[\PCname]+4, ex)::\sigma)\)}
+\[\begin{split}
+\mathit{reg\_view} ~ V ~ \overline{r_{args}} \triangleq
+V & \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR\_SV} \rrbracket \\
+& \llbracket \reg \mapsto \public | \reg \in \overline{r_{args}} \rrbracket
+\end{split}\]
+
+\judgmenttwo{\(\chi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}}\)}
+            {\(V' = \mathit{reg\_view} ~ V ~ \overline{\reg_{args}}\)}
+            {\(Cop ~ \chi ~ (V,\sigma) ~ \mach \triangleq
+              (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}
+
+After the call step, we have a new activation, and {\tt main} is now pending.
+{\tt main}'s pending record tracks its return address.
+Everything that {\tt main} allocated, both {\tt secret} and {\tt sensitive},
+{\tt f} views as \(\sealed\), along with its callee-saved registers.
+Informally, the meaning of \(\sealed\) is: ``when control returns to
+{\tt main}, this elements value should be the same as it is now.''
+This is termed {\it return-time integrity}, and is defined below.
+The remainder of the stack, and the caller-save registers, are \(\unsealed\).
+The informal meaning of \(\unsealed\) is: ``{\tt f}'s behavior should not
+depend on this element's value, but {\tt f} is free to overwrite it.''
+This becomes relevant in out {\it confidentiality} properties.
+Finally, \(\public\) elements are those that {\tt f} is allowed to
+depend on, and may also overwrite.
 
@@
             {\(\chi = \mathbf{return}\)}
-            {\(\sigma = (V,\addr_{ret},ex)::\sigma'\)}
+            {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
             {\(Cop ~ \chi ~ (\_, \sigma) ~ \mach \triangleq
-              ((V,ex), \sigma)\)}
+              (V, \sigma)\)}
 
@@
          {\(\chi = \mathbf{return}\)}
-         {\(Cop ~ \chi ~ (\alpha, [~]) ~ \mach \triangleq
-           (\alpha, [~])\)}
+         {\(Cop ~ \chi ~ (V, [~]) ~ \mach \triangleq
+           (V, [~])\)}
 
-\begin{figure*}
-  \[\mathit{reg\_view} ~ \overline{r_{args}} \triangleq \lambda \reg .
-  \begin{cases}
-    \public & \textnormal{if } \reg \in \overline{r_{args}} \textnormal{ or } \reg \in \mathit{PUBLIC} \\
-    \unsealed & \textnormal{else if } \reg \in \mathit{CLR\_SV} \\
-    \sealed & \textnormal{else} \\
-  \end{cases}\]
+We are ready to define our first stack safety property for this system:
+{\it well-bracketed control flow}.
 
-  \[\mathit{update\_view} ~ V ~ \overline{r_{args}} \triangleq \lambda \component
-  \begin{cases}
-    \mathit{reg\_view} ~ \overline{r_{args}} ~ \component & \textnormal{if } \component \in \REGS \\
-    \sealed & \textnormal{else if } V ~ \component = \mathit{object} \\
-
-      V ~ \component & \textnormal{otherwise} \\
-  \end{cases}\]
-
-\caption{View updates for calls}
-\label{fig:callcontexts}
-\end{figure*}
+\definition A system enjoys well-bracketed control flow if, for all states
+\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context)\),
+\(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).
 
@@
 
-We define another relation that takes a trace until we have returned to a particular
-activation, or past it (as might occur when we add exceptions.)
+We define another relation that takes a trace until we have returned from the
+active principal.
 We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
-\(d\) is the depth of the original call.
+\(d\) is the depth of the current call.
 
-\judgment{\(|\sigma| \leq d\)}
-         {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \tau\)}
+\judgment{\(|\sigma| < d\)}
+         {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \tau\)}
 
-\judgmenttwobrlong{\((\mach,(\alpha,\sigma)) \stepstounder{e} (\mach',\context')\)}
-                  {\(|\sigma| > d\)}
+\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{e} (\mach',\context')\)}
+                  {\(|\sigma| \geq d\)}
                   {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
-                  {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \notfinished{e}{\obsT}\)}
+                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{e}{\obsT}\)}
 
@@
 We start by defining a second-order logical operator
-\(d \uparrow P\), read ``\(P\) holds on return to depth \(d\),''
+\(d \uparrow P\), read ``\(P\) holds on return from depth \(d\),''
 where \(P\) is a predicate on machine states. This is a coinductive relation
 similar to ``weak until'' in temporal logic -- it holds if the program never
-returns to depth \(d\).
+returns from depth \(d\).
 
 \judgmenttwo[Returned]
-            {\(|\sigma| \leq d }
-            {\(P ~ (\mach,(\alpha,\sigma))\)}
-            {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}
+            {\(|\sigma| < d\)}
+            {\(P ~ (\mach,(V,\sigma))\)}
+            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
 
 \judgmenttwobrlong[Step]
-                  {\(|\sigma| > d\)}
+                  {\(|\sigma| \geq d\)}
                   {\((d \uparrow P) ~ (\mach', \context')\)}
-                  {\((\mach, (\alpha,\sigma)) \stepstounder{e} (\mach', \context')\)}
-                  {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}
+                  {\((\mach, (V,\sigma)) \stepstounder{e} (\mach', \context')\)}
+                  {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
 
@@
 \((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
-first states that return to depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).
+first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).
 
 \judgmenttwobrlong[Returned]
-            {\(|\sigma_1| \leq d\)}
-            {\(|\sigma_2| \leq d\)}
-            {\((\mach_1,(\alpha_1,\sigma_1)) ~ R ~ (\mach_2,(\alpha_2,\sigma_2))\)}
-            {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
+            {\(|\sigma_1| < d\)}
+            {\(|\sigma_2| < d\)}
+            {\((\mach_1,(V_1,\sigma_1)) ~ R ~ (\mach_2,(V_2,\sigma_2))\)}
+            {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
 
 \judgmenttwobrlong[Left]
-              {\(|\sigma_1| > d\)}
-              {\((\mach_1,(\alpha_1,\sigma_1)) \stepstounder{e} (\mach_1',\context_1')\)}
-              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
-              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
+              {\(|\sigma_1| \geq d\)}
+              {\((\mach_1,(V_1,\sigma_1)) \stepstounder{e} (\mach_1',\context_1')\)}
+              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
 
 \judgmenttwobrlong[Right]
-              {\(|\sigma_2| > d\)}
-              {\((\mach_2,(\alpha_2,\sigma_2)) \stepstounder{e} (\mach_2',\context_2')\)}
-              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
-              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
+              {\(|\sigma_2| \geq d\)}
+              {\((\mach_2,(V_2,\sigma_2)) \stepstounder{e} (\mach_2',\context_2')\)}
+              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
+              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
 
@@
 
-\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
+\definition Let \((\mach,\context)\) be a combined state with
+\(\context = (V,\sigma)\), and
 \(\components = \{\component | V ~ \component = \sealed\}\).
@@
 \(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
-Then \(\mach\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.
+Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \downarrow P\) holds
+on \((\mach,\context)\).
 
@@
 \((\mach,\context)\) such that
-\((\mach,\context) \stepstounder{\psi} (\mach',\context')\) with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
+\((\mach,\context) \stepstounder{e} (\mach',\context')\) with
+\(e = \mathbf{callE} ~ \addr ~ \overline{\word}\),
 \((\mach',\context')\) enjoys return-time integrity.
@@
 
-\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
+\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
 \(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
@@
 Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to
-some \(\context_{ret}\) if, for any \(\nach\)
-that is a \(\components\)-variant of \(\mach\), we can take
-\(\context_{ret} \uparrow (\mach,\context) \hookrightarrow \obsT\) and
-\(\context_{ret} \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
+if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
+\(|\sigma| \uparrow (\mach,\context) \hookrightarrow \obsT\) and
+\(|\sigma| \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
 \(\obsT \simeq \obsT'\).
 
-\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
-\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
-
-Then \((\mach,\context)\) enjoys {\it return-time confidentiality} with respect to
-some \(\context_{ret}\) if \((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\nach,\context)\),
+\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\),
+and \(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
+Then \((\mach,\context)\) enjoys {\it return-time confidentiality}
+if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\),
+\((\mach,\context) ~ (|\sigma| \Uparrow R) ~ (\nach,\context)\),
 where \((\mach_1,\context_1) ~ R ~ (\nach_2,\context_2)\) if
-\((\mach_1 \parallel \bar{\Diamond}(\mach,\mach_1,\nach,\nach_2)\).
+\((\mach_1,\context_1) \parallel \bar{\Diamond}(\mach,\mach_1,\nach,\nach_2)\).
 
 \definition A system enjoys {\it caller confidentiality} if, for all reachable states
-\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
-with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
-\((\mach',\context')\) enjoys both internal and return-time confidentiality
-with respect to \(\context\).
+\((\mach,\context)\) such that \((\mach,\context) \stepstounder{e} (\mach',\context')\)
+with \(e = \mathbf{callE} ~ \addr ~ \overline{\word}\),
+\((\mach',\context')\) enjoys both internal and return-time confidentiality.
 
@@
 
-\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
+\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
 \(\components = \{\component | V ~ \component \in \{\sealed,\unsealed\}\}\).
@@
 \(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
-Then \(\mach\) enjoys {\it post-return confidentiality} if \(\context \downarrow P\) holds.
+Then \(\mach\) enjoys {\it post-return confidentiality} if \(|\sigma| \downarrow P\) holds.
 
 \definition A system enjoys {\it callee confidentiality} if, for all reachable states
-\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
-with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
+\((\mach,\context)\) such that \((\mach,\context) \stepstounder{e} (\mach',\context')\)
+with \(\psi = \mathbf{callE} ~ \addr ~ \overline{\word}\),
 \((\mach',\context')\) enjoys post-return confidentiality
@@
 We can extend our basic machine to support public memory allocations that can
-be accessed by anyone until they are deallocated on their owner's return.
-We replace \(\mathbf{alloc} ~ \mathit{sizes}\) with \(\mathbf{alloc} ~ \mathit{objs}\),
-where \(\mathit{objs}\) is a list of pairs of sizes and booleans, representing whether
-the object in question should be public. The \(\mathit{update\_view}\)
-function only seals private objects, and the remained maintain their \(\mathit{object}\)
-class, making them exempt from both integrity and confidentiality properties.
-
-\section{Exceptions}
-
-We implement exceptions via another pair of macro-instructions:
-\(\mathbf{setex} ~ a_{buf}\), akin to \(\mathit{setjmp}\), with the
-jump buffer stored at \(a_{buf}\), and \(\mathbf{throwex} ~ a_{buf}\),
-akin to \(\mathit{longjmp}\). The context updates for exceptions are given
-in Figure \ref{fig:excontexts}. The update for \(\mathbf{setex}\) pushes
-a \(\mathit{Catch}\) target onto the stack, identified by \(\addr_{buf}\),
-with the same view as the current function. For \(\mathbf{throwex}\),
-we pop the stack until we find a \(\mathit{Catch} ~ a_{buf} ~ \_ ~ \_\),
-popping it as well. Rules apply in order.
-[TODO: update these for the new \(\alpha\) stuff.]
-
-\begin{figure*}
-
-\judgmenttwo[SetExC]
-            {\(\psi = \mathbf{setex} ~ \addr_{buf}\)}
-            {\(T' = \mathit{Catch} ~ \addr_{buf} ~ (\mach[pc] + 4) ~ \mach[sp]\)}
-            {\(\mach, (T,V)::\context \constep{\psi} (T',V)::(T,V)::\context\)}
-
-\judgment[ThrowExCCatch]
-         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
-         {\(\mach, (\mathit{Catch} \addr_{buf} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\psi} \context\)}
-
-\judgment[ThrowExCDone]
-         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
-         {\(m, [\mathit{Done},V] \constep{\psi} [\mathit{Done},V]\)}
-
-\judgmenttwo[ThrowExCSkip]
-         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
-         {\(\mach, \context \constep{\psi} \context'\)}
-         {\(\mach, (\_::\context) \constep{\psi} \context'\)}
-
-\caption{Exception context updates}
-\label{fig:excontexts}
-\end{figure*}
+be accessed by anyone until they are deallocated on their owner's return,
+using the \(\mathbf{alloc\_pub}\) operation. This behaves identically to
+\(\mathbf{alloc}\) except that it marks the locations \(\public\).
+We should give an example of this, as it may also require us to use multiple
+allocations simultaneously to interleave public and private allocations with a
+single increase of the stack pointer.
 
@@
 \judgmenttwo[TailCallC]
-            {\(\psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
+            {\(\chi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
             {\(V' = \mathit{update\_view} ~ V ~ \overline{r_{args}}\)}
-            {\(\mach, (\alpha,\mathit{Active},V)::\context \constep{\psi} (\alpha,\mathit{Active},V')::\context\)}
+            {\(Cop ~ \chi ~ (V,\sigma) ~ (V',\sigma)\)}
 

### 2022-10-05 submission5/noexceptions.tex
Bunch of stuff

@@
 
-\section{Threat Model, Machines, and Enforcement }
+\section{The Setting}
 
-Before introducing our formal model of stack safety, we describe
-our machine exemplar, enforcement mechanism, and threat model.
+We state our properties in terms of a generic machine subject to a few constraints,
+but our examples and tests are focused on a RISC-V-like machine enhanced with PIPE,
+a tag-based reference monitor. Importantly, such a machine has no explicit ``call''
+instruction, only {\tt jal} instructions which may be used for that purpose or potentially
+other purposes. Similarly, there is no ``return'', only {\tt jalr}, and so on.
+We therefore define our properties under the assumption that we can distinguish the
+steps that represent these abstract operations from those that do not -- an assumption
+that is valid in our testing environment (Section [TBD]) and that is already necessary
+to correctly apply tag-based enforcement in practice (cite something about this.)
 
-We begin with a RISC-V-like machine enhanced with PIPE, a tag-based reference
-monitor. We extend this machine with a {\it security semantics},
-which divides a program into dynamic {\it principals} that ought to be
-protected from one another, and specifies that protection over a set
-of abstract {\it security operations}. Specifications may take the form of
+The set of special operations that are recognized in this way are termed
+{\it security operations}. Our properties are phrased as a {\it security semantics},
+which extends the underlying machine with additional information about its
+security principals and which registers and memory they require to be secure, which evolves
+dynamically through the execution of security operations.
+Specifications may take the form of
 predicates on future states: statements of the form,
 ``When control returns to me...'', or of relations on traces of future execution
-(hyper-properties.)
+(hyper-properties).
 
-As opposed to a fully-abstract overlay semantics, in which overlay operations
-are given safe-by-construction semantics then proven fully abstract,
-our machine is not safe unless the monitor is correctly instantiated;
-the security semantics gives us the means to distinguish a monitor that implements
-stack safety from one that does not.
+This approach is in contrast to approaches that give a safe-by-construction semantics
+for various security operations and then prove secure compilation to the underlying
+enforcement mechanism, in that our machine is explicitly not safe unless the enforcement
+is correctly instantiated; the security semantics gives us the means to distinguish an
+enforcement scheme that implements stack safety from one that does not.
 
-The underlying machine is a standard RISC-V ISA as modeled in [MIT-RISCV].
-Our overlay semantics provide overlay operations for the following code features:
+We will first give a detailed example of a security semantics for a simple setting
+in which our security operations are restricted to making calls and returns, and allocating
+private memory within the current function activation. We then extend this model
+to the full system that we test, which features:
 \begin{itemize}
-\item Function calls and returns, with caller- and callee-saved registers according
-  to the RISC-V ABI
-\item Arguments passed by reference
+\item Function calls and returns, with caller- and callee-saved registers
+\item Allocation of strictly private and strictly public regions on the stack
+\item Arguments spilled onto the stack and/or passed by reference
 \item Exceptions
@@
 
+Finally, we give a property definition for stack safety in which (some) stack-allocated
+objects are governed by a provenance-based capability system, accessible only to those
+functions which have recieved a valid pointer to them.
+
+In section (TBD) we discuss the extension of this model into a simple concurrency setup.
+
 \paragraph*{Threat Model}
 
-We trust the compiler has correctly placed macro-instructions according to
-the semantics of the source language, but we do not assume that the code
-is otherwise reasonable. In particular, while we are agnostic as to the source
-language, C is very plausible, and so any source function might contain undefined
-behavior resulting in its compilation to arbitrary machine code.
+We must trust that our method of distinguishing security operations is reasonable; if it
+involves labels placed on code by a compiler, that means trusting that the compiler placed
+those labels correctly. If it recognizes the operations based on the dynamic behavior of
+the system, similarly, those definitions must be realistic.
 
-In general, it is impossible to distinguish buggy source code from an attacker;
-we assume nothing about the intent of the source code, except that it makes calls
-and returns, and we must guarantee that the caller and callee are protected from one
-another.
+Otherwise, we do not assume that the code is reasonable or corresponds to any particular
+calling convention or source construct. In particular, while we are agnostic as to the source
+language, we certainly aim to support C, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code. A given enforcement
+mechanism may place additional constraints, of course, particularly on the behavior of
+call and return sequences.
+
+In general, it is impossible to distinguish buggy source code from an attacker; in
+our examples we will identify one function or another as an attacker, but we do not
+require any static division between trusted and untrusted code, and aim to protect
+even buggy code.
 
@@
 
-We do not support dynamic code generation, and our concurrency model is fairly
+Our concurrency model is fairly
 simplistic, assuming a fixed number of threads each with its own dedicated processor.
@@
 
-\subsection{Security Semantics}
-
-We begin with a machine that supports a number of security operations,
-defined above. For purposes of specifying security, we do not need to
-know the semantics of these operations; we only need the machine to
-distinguish them from normal execution. We model the machine as
-a transition system in which some transitions are marked with
-security operations. We then define a security semantics for stack safety
-in terms of those steps.
-
-The security semantics divides a program's execution into principals,
-namely function activations, and identifies which principal is active
-in a given state. Security operations can generate new principals, transition
-which principal is active, and allocate stack objects belonging to the active
-principal. From this information, we derive the specifications that make up
-stack safety. Specifications may take the form of predicates on
-future states: statements of the form, ``When control returns to me...'',
-or of relations on traces of future execution (hyper-properties.)
-
 \subsection{The Basic Machine}
@@
 The building blocks of the machine are {\em words} and {\em registers}.
-Words (\(\WORDS\)) are 64-bit integers ranged over by \(\word\) and, when used as addresses,
-\(\addr\). The processor has 32 registers, ranged over in general by \(\reg \in \REGS\),
-plus a program counter (\(\PCname\)). The following table divides
-the registers into metavariables, some belonging to caller- or callee- sets:
+Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
+and are drawn from the set (\(\WORDS\)) (by default, the set of 64-bit integers).
+Registers are ranged over in general by \(\reg \in \REGS\), other than the
+program counter \(\PCname\) and stack pointer \(\SP\).
+They are divided into sets of caller-saved and callee-saved registers;
+Figure \ref{fig:RISCVregs} gives an example division for a RISC-V machine as will
+appear in our examples and testing, along with the register names that will appear in
+example code, and their default security class (explained below).
 
-\vspace{\abovedisplayskip}
+\begin{figure}
   \begin{tabular}{| l | l | l |}
     \hline
-  Set & Names & Purpose \\
+    Set / & Names & Purpose \\
+    Class & & \\
     \hline
-  CLR\_SV & \(rt_0 - rt_6\) & Caller-saved temps \\
-  & \(ra_0 - ra_1\) & Caller-saved args / return vals \\
-  & \(ra_2 - ra_7\) & Caller-saved args \\
+    \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
+    \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
+    & {\tt a2} -- {\tt a7} & Caller-saved args \\
     \hline
-  CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
-  & \(rra\) & Return Address \\
-  & \(rsp\) & Stack Pointer \\  
+    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s7} & Callee-saved \\
+    \(\sealed\) & {\tt ra} & Return Address \\
+    & {\tt sp} & Stack Pointer \\  
     \hline
-  PUBLIC & \(pc\) & Program Counter \\
+    \(\mathit{PUBLIC}\) / & {\tt pc} & Program Counter \\
+    \(\public\) & & \\
     \hline
   \end{tabular}
-\vspace{\abovedisplayskip}
+  \caption{RISC-V register set}
+  \label{fig:RISCVregs}
+\end{figure}
 
@@
 that defines a mapping \(\mach[\component]\) from state elements to values.
+{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\).
 The machine step function
-\(\mach \xrightarrow{\psi} \mach' \in \MACHS \rightarrow \MACHS \times \Psi\)
-may be labeled by a security operation, defined below.
+\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \in \MACHS \rightarrow
+\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
+is labeled by a list of security operations (order matters), defined below,
+and an event.
 While the tag-based enforcement mechanism may cause the machine to
-failstop, we model this as the machine silently diverging by returning to the
+failstop, we model this as the machine silently diverging by remaining in the
 same state. Since failstops are not distinguished from normal execution,
-our properties will necessarily be {\it termination insensitive}.
+our properties will suffer the limitations typical of
+{\it termination insensitive} systems \cite{}.
 
@@
 
-Security operations are split into two categories: {\it privilege operations}
-and {\it control operations}. Privilege operations are those that adjust the
-privileges of future callees by allocating private or public data and by saving
-catch points for possible exceptions. Allocations specify the offset from
-the stack pointer and size of the object. Exceptions are tied to the address
-of the buffer that will contain saved state.
+Security operations consist of calls (including tailcalls), returns,
+allocations and deallocations. Calls and tailcalls identify the address of the call's target,
+and the argument registers. Allocations and deallocations are phrased in terms of a pair
+of an offset and a size. Returns don't require any additional information.
 %
 \begin{align*}
-  \phi \in \Phi ::= & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
-  | & \mathbf{alloc\_pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
-\end{align*}
-
-Control operations are those that exchange
-control between principals: calls (including tailcalls), returns, and thrown
-exceptions. Calls and tailcalls identify the target of the call, and the argument registers.
-[Still a todo: stack arguments.] Return needs no further information. Throwing
-an exception specifies the address of the saved buffer.
-%
-\begin{align*}
-  \chi \in \mathcal{X} ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
-  | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
+  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} \\
+  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} \\
   | & \mathbf{return} & \\
+  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
 \end{align*}
 %
-A security operation may consist of any number of privilege operations and at most
-one control operation.
-%
-\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (\mathcal{X} \times \mathit{list ~ \Phi})\]
-
 Security principals are function activations: at any given time, the current
@@
 of the system that maps each element to one of a set of {\it security classes}:
-\[sc \in SEC ::= \sealed | \unsealed | \public\]
+\[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
 \[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]
@@
 private to an inactive principal (\(\sealed\)),
-%in an allocated object that is accessible to the current principal (\(\mathit{object}\)),
+in an allocated object that is accessible to the current principal (\(\object\)),
 available to be allocated (\(\unsealed\)), or
 always accessible (\(\public\)).
-The {\it initial view} \(V_0\) maps all stack locations and non-public registers
-to \(\unsealed\) and all other locations and public registers to \(\public\).
+The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
+all other locations to \(\public\), and registers based on which set they
+belong to (Figure \ref{fig:RISCVregs}): \(\sealed\) for callee-saved,
+\(\unsealed\) for caller-saved, and \(\public\) otherwise.
 
@@
 
-Every security operation manipulates the call stack: privilege operations act on
-the active principal, and control operations push or pop the active principal.
-We define the security effects of each operation as a function \(Pop\) or \(Cop\),
-from an operation, a machine state, and a call stack to a new call stack.
-As we introduce more complex operations, we will separately define
-the curried functions \(Pop ~ \phi\) or \(Cop ~ \chi\) for each operation.
-They are combined in \(Op\), which acts on full operations as follows.
-\[Pop \in \Phi \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
-\[Cop \in X \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
-\[Op ~ \psi =
-\begin{cases}
-  \lambda \context ~ \mach . \context & \textnormal{when } \psi = [~] \\
-  Pop ~ \phi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = \phi::\overline{\phi} \\
-  Cop ~ \chi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = (\chi,\overline{\phi}) \\
-\end{cases}\]
+Every security operation manipulates the call stack via a function
+\(Op : \Psi \rightarrow \MACHS \rightarrow \CONTEXTS \rightarrow \CONTEXTS\). We will introduce the
+curried application of \(Op\) to each operation as we introduce them in detail.
 
-A control operation may additionally produce an observable event, drawn
-from a set \(\mathit{EVENTS}\) and ranged over by \(e\).
-A function \(Ev\) maps operations and machine states to events. The
-silent event \(\tau\) is relevant to our definitions here, and the
-remainder will be described in Section \ref{sec:events}. 
-%
-\[Ev \in X \rightarrow \MACHS \rightarrow \mathit{EVENTS}\]
-%
 From a machine state and a context stack, we create a {\it combined state}
 \((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
-labeled with an event. If the underlying machine transition is labeled by
-some operation \(\psi\), the combined step applies \(Op\) to the contents
-of \(\psi\) to determine the new call stack; otherwise the call stack is
-unchanged. If there is a control operation,
-the transition is labeled with its event, otherwise it is silent.
-Rules apply in order.
+labeled with the same operations. Note that \(Op\) takes as its state argument
+\(\mach\), the state before the step.
 
-\judgmentthree[Control]
-              {\(\mach \xrightarrow{\psi} \mach'\)}
-              {\(\psi = (\chi, \overline{\phi})\)}
-              {\(e = Ev ~ \chi ~ \mach\)}
-              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \context ~ \mach)\)}
-
-\judgment[NonControl]
-         {\(\mach \xrightarrow{\psi} \mach'\)}
-         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \context ~ \mach)\)}
+\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
+            {\(\mathit{foldl} ~ (Op ~ \psi ~ \mach) ~ \context = \context'\)}
+            {\((\mach,\context) \stepstounder{\psi :: \overline{\psi}} (\mach', \context'', \obs)\)}
 
@@
 
-In this section we consider a simple setting with calls, returns, and private
-allocations. Assume that \(\MACHS\) describe a RISC-V machine, where a transition
-\(\mach \xrightarrow{\Psi?} \mach'\) is labeled as follows.
-If \(\mach[\PCname]\) points to {\tt addi sp,sp,N}, \(\Psi? = \mathbf{alloc} ~ (0,N)\);
-since we are not yet allocating public objects,
-we don't need to distinguish objects from one another if they are allocated
-simultaneously. Similarly, if \(\mach[\PCname]\) points to {\tt jal off ra}
-and \(\mach'[\PCname] = a_f\) where \(a_f\) is the entry of a function
-\(f\) that uses argument registers \(\overline{r}\), 
-\(\Psi? = \mathbf{call} ~ f ~ \overline{r}\). And a step via the
-instruction {\tt jalr ra} is labeled \(\mathbf{return}\).
-This simplistic system serves to demonstrate our core properties.
-
-Figure \ref{fig:example} gives a running example
-of a function {\tt main} that takes a secret in its arguments and allocates
+In this section we consider a simple setting with calls, returns, and private allocations.
+Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
+that takes a secret in its arguments and allocates
 an array of potentially sensitive data. It then calls another function {\tt f},
@@
 
-\begin{figure*}
-  \begin{subfigure}[t]{.5\textwidth}
+\begin{figure}
+  \begin{subfigure}[b]{\columnwidth}
     {\tt
-      void main(int argc, char*argv) \{
+      void main(int argc, int* argv) \{
 
@@
 
-      ~ ~ int sensitive[4] = \{0\};
+      ~ ~ int sensitive[4] = \{0,0,0,0\};
 
-      ~ ~ f();
+      ~ ~ int res = f();
 
-      ~ ~ if (sensitive == 42) \{
+      ~ ~ if (sensitive[0] == 42) \{
 
@@
 
+      ~ ~ \} else \{
+
+      ~ ~ ~ ~ publish(res);
+
       ~ ~ \}
 
-      \}
-    }
+      \}}
   \end{subfigure}
-  \begin{subfigure}[t]{.5\textwidth}
+  \begin{subfigure}[b]{.45\columnwidth}
+    \vspace{2\abovedisplayskip}
+
     {\tt
-      0: addi sp,sp,-20
+      0: addi sp,sp,-24
 
-      4: sw a1,16(sp)
+      4: sw a1,24(sp)
 
-      8: sd zero,8(sp)
+      8: sd zero,16(sp)
 
-      12: sd zero,0(sp)
+      12: sd zero,8(sp)
 
-      16: jal 84 ra
+      16: jal 84,ra
 
-      20: lw a4,12(sp)
+      20: lw a4,4(sp)
 
@@
 
-      28: bne a4,a5,L
+      28: bne a4,a5,L1
 
-      32: ld a0,16(sp)
+      32: lw a0,16(sp)
 
-      36: jal 164 ra
+      36: jal 164,ra
 
-      L(40): nop
+      40: j L2:
 
-      44: addi sp,sp,20
+      L1:
 
-      48: return
+      44: jal 164,ra
 
+      L2:
+
+      48: addi sp,sp,24
     }
+  \end{subfigure}
+  \begin{subfigure}[b]{.5\columnwidth}
+    \includegraphics[width=\columnwidth]{stacklayout.png}
   \end{subfigure}
@@
 \caption{Example: Main}
-  \label{fig:example}
+\label{fig:main}
+\end{figure}
+
+Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
+to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
+assembly code in Figure \ref{fig:f}. It might take an offset from the stack
+pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
+subtly, even if somehow prevented from outputting {\tt secret} directly, it might
+instead return that value so that {\tt main} makes the call to {\tt publish}.
+
+Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive[0]}
+with 42, guaranteeing that {\tt main} publishes its own secret unintentionally!
+Attacks of this kind do not violate {\tt main}'s confidentiality, but its
+{\it integrity}.
+And last, but certainly not least, the attacker might attempt to return to the
+wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
+violating the program's {\it well-bracketed control flow} (WBCF.)
+
+\begin{figure}
+  \begin{subfigure}[b]{.5\columnwidth}
+    {\tt
+      100: addi sp,sp,-4
+
+      104: sw ra,4(sp)
+
+      108: lw a0,20(sp)
+
+      112: jal 88,ra
+
+      116: lw ra,4(sp)
+
+      120: jalr ra
+    }
+    \caption{Leaking {\tt secret} directly}
+  \end{subfigure}  
+  \begin{subfigure}[b]{.45\columnwidth}
+    {\tt
+      100: lw a4,16(sp)
+
+      104: sw a4,12(sp)
+
+      108: jalr ra
+    }
+    \caption{Leaking {\tt secret} indirectly}
+  \end{subfigure}
+
+  \vspace{\belowdisplayskip}
+
+  \begin{subfigure}[b]{.5\columnwidth}
+    {\tt
+      100: li a5,42
+
+      104: sw a5,12(sp)
+
+      108: jalr ra
+    }
+    \subcaption{Attacking {\tt sensitive}}
+  \end{subfigure}
+  \begin{subfigure}[b]{.45\columnwidth}
+    {\tt
+      100: addi ra,ra,20
+
+      104: jalr ra
+    }
+    \subcaption{Attacking control flow}
+  \end{subfigure}
+
+  \caption{Assembly code for {\tt f} as an attacker}
+  \label{fig:f}
+\end{figure}
+
+At the start of execution the program counter is 0 and the stack pointer -4, with code
+proceeding upward and the stack growing down. The functions {\tt f} and {\tt publish}
+are at addresses 100 and 200, respectively.
+
+Figure \ref{fig:exec1} shows how the program counter, stack pointer, and call stack
+develop over several steps, with labels on the steps that allocate space and then make
+a call. \(\downarrow \overline{\psi}\) in this diagram is a sideways
+\(\xrightarrow{\overline{\psi}}\). Events are omitted.
+
+The first step allocates a word for {\tt secret} and four for {\tt sensitive}. To
+maintain a two-word alignment it leaves a gap. This will have the
+effect of marking those bytes \(\object\), assuming they were previously
+\(\unsealed\).
+
+\begin{figure*}
+  \begin{subfigure}[t]{.5\textwidth}
+    \vskip 0pt
+    \begin{tabular}{| l | l | l | l |}
+      \hline
+      Step & \(\PCname\) & \(\SP\) & \(\context\) \\
+      \hline
+      0 & 0 & 0 & \(V_0, \emplist\) \\
+      \hline
+      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-4,4);\mathbf{alloc} ~ (-24,16)]\)} \\
+      \hline
+      1 & 4 & -24 & \(V_1 = V_0 \llbracket -24 \ldots -9, \mapsto \object \rrbracket\) \\
+          & & & \(\llbracket -4 \ldots -1, \mapsto \object \rrbracket, \emplist\)\\
+      \hline
+      \multicolumn{4}{l}{\(\vdots\)} \\
+      \hline
+      4 & 16 & -24 & \(V_1, \emplist\) \\
+      \hline
+      \multicolumn{4}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
+      \hline
+      5 & 100 & -24 & \(V_2 = V_1 \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket,\) \\
+      & & & \([(V_1, 20, -24)]\) \\
+      \hline
+      \multicolumn{4}{l}{\(\vdots\)} \\
+      \hline
+      \(n\) & ? & ? & \(V_2,[(V_1, 20, -24)]\) \\
+      \hline
+      \multicolumn{4}{l}{\(\downarrow [\mathbf{return}]\)} \\
+      \hline
+      \(n+1\) & 20 & -24 & \(V_1, \emplist\) \\
+      \hline
+    \end{tabular}
+  \end{subfigure}
+  \begin{subfigure}[t]{.4\textwidth}
+    \vskip 0px
+    \includegraphics[width=\columnwidth]{stack1.png}
+    \includegraphics[width=\columnwidth]{stack2.png}
+    \includegraphics[width=\columnwidth]{stack3.png}
+    \includegraphics[width=\columnwidth]{stack4.png}
+  \end{subfigure}
+
+\caption{Execution through return from {\tt f}}
+\label{fig:exec1}
 \end{figure*}
 
-At the start of execution (PC 0) the initial call stack is \((V_0, [~])\).
-Instruction 0 allocates 20 bytes starting at the stack pointer. This will have the
-effect of marking those twenty bytes \(\sealed\), assuming they were previously
-\(\unsealed\):
-
-\judgmentbr{\(b = \mach[\SP] - \mathit{off}\)}
-           {\(V' = V \llbracket \addr \mapsto \sealed |
-             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
-           {\(Pop ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach \triangleq
-             (V',\sigma)\)}
-
-This memory being \(\sealed\) has no effect on the current activation --
-the security guarantees apply at the start of the call. Our next security operation
-occurs when we step from the call at instruction 16.
-
-On a call, the formerly active principal's record is pushed onto the inactive list,
-and replaced by a new record for the callee. Its return target is the instruction
-immediately following that of the program counter before the call.
+On a call, the formerly active principal's record is pushed onto the inactive list.
+Its return target is the instruction immediately following that of the program
+counter before the call, and the stack pointer target is the current stack pointer.
 The callee's view is updated from the caller's such that the argument
 registers are \(\public\) and any non-argument, caller-saved registers
-are \(\unsealed\). All callee-save registers are always \(\sealed\), and
-the program counter is always public, but this is already the case.
+are \(\unsealed\). All callee-save registers are always \(\sealed\) and
+the program counter \(\public\), as they were initialized.
 
-\[\begin{split}
-\mathit{reg\_view} ~ V ~ \overline{r_{args}} \triangleq
-V & \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR\_SV} \rrbracket \\
-& \llbracket \reg \mapsto \public | \reg \in \overline{r_{args}} \rrbracket
-\end{split}\]
+We suppose that, on step \(n+1\), {\tt f} returns; if it does not, then
+{\tt main}'s secrets are safe as long as {\tt f} does not leak them directly.
+On that return, we restore the topmost inactive principal, {\tt main}'s principal.
 
-\judgmenttwo{\(\chi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}}\)}
-            {\(V' = \mathit{reg\_view} ~ V ~ \overline{\reg_{args}}\)}
-            {\(Cop ~ \chi ~ (V,\sigma) ~ \mach \triangleq
-              (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}
+\paragraph*{Well-bracketed Control Flow}
 
-After the call step, we have a new activation, and {\tt main} is now pending.
-{\tt main}'s pending record tracks its return address.
-Everything that {\tt main} allocated, both {\tt secret} and {\tt sensitive},
-{\tt f} views as \(\sealed\), along with its callee-saved registers.
-Informally, the meaning of \(\sealed\) is: ``when control returns to
-{\tt main}, this elements value should be the same as it is now.''
-This is termed {\it return-time integrity}, and is defined below.
-The remainder of the stack, and the caller-save registers, are \(\unsealed\).
-The informal meaning of \(\unsealed\) is: ``{\tt f}'s behavior should not
-depend on this element's value, but {\tt f} is free to overwrite it.''
-This becomes relevant in out {\it confidentiality} properties.
-Finally, \(\public\) elements are those that {\tt f} is allowed to
-depend on, and may also overwrite.
+But what if {\tt f}, does not return with \(\PCname = 20\) and \(\SP = -24\),
+as would be the case in the fourth example attack? Then it has violated WBCF.
+This property can be phrased in terms of a single step: for every transition
+labeled as a return, the program counter and stack pointer match those recorded
+in the top pending activation.
 
-On return, we restore the topmost inactive principal, if there is one.
-Else, we default to retaining the current principal, as we do not model
-the main function returning.
+\paragraph*{Stack Integrity}
 
-\judgmenttwo[Ret]
-            {\(\chi = \mathbf{return}\)}
-            {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-            {\(Cop ~ \chi ~ (\_, \sigma) ~ \mach \triangleq
-              (V, \sigma)\)}
+Integrity is more complicated, and requires us to compare the state at the start of
+the call to {\tt f} to that on return. Intuitively, the integrity of {\tt main}
+is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
+to be identical to when it made the call.
 
-\judgment[RetDef]
-         {\(\chi = \mathbf{return}\)}
-         {\(Cop ~ \chi ~ (V, [~]) ~ \mach \triangleq
-           (V, [~])\)}
+In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
+addresses -24 through -9 and -4 through -1 (allocated as private data) and the callee-saved registers.
+Note that the callee-saved registers may well have changed during the call -- but if
+so, the callee is obligated to restore them to their prior value.
 
-We are ready to define our first stack safety property for this system:
-{\it well-bracketed control flow}.
+One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
+in a way that is harmless -- or that is rendered harmless by the enforcement mechanism!
+This is the case in a lazy micro-policy: the monitor permits the caller to write into the
+callee's frame, but taints that write so that the callee cannot read it later without
+triggering a failstop. In such a scenario we will consider a value {\it harmless}
+if it could be replaced with any other value without changing the observable behavior
+of the machine. That is, for a given state \(\mach\) and set of elements \(\components\),
+we define a {\it variant} state that agrees with \(\mach\) on every element except those
+in \(\components\). Those elements are said to be harmless in \(\mach\) if the variant
+state produces an equivalent trace (see Section \ref{sec:events}.)
 
-\definition A system enjoys well-bracketed control flow if, for all states
-\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context)\),
-\(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).
+So, to define integrity, we need to know when a caller has been returned to;
+this is the first state after the call in which the depth of the call stack
+matches the state before the call. Integrity means that, at that return state,
+any element that is \(\sealed\) under the callee's view is either restored
+to its original value or is harmless.
+
+\paragraph*{Caller Confidentiality}
+
+We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
+means that its callee's behavior is dependent only on publicly visible data,
+not the caller's state. As we saw in our examples, the callee's behavior is defined
+both in terms of events that the callee produces during the call, but also in terms of
+the changes it makes to the state.
+
+Consider the state \((\mach, (V_2,\sigma))\) at step 5. If we take a variant state over
+the set of element(s) that are \(\sealed\) or \(\unsealed\) in \(V_2\),
+{\it internal confidentiality} means
+that, if we execute both states until they return, they will produce equivalent sequences
+of events.
+
+But, it is also possible that the callee could extract data from the caller that it does
+not use to produce a visible event, but instead stashes for later or uses to influence
+its return value -- which in turn can lead to future changes in behavior that do not
+even appear to be the fault of the callee!
+{\it Return-time confidentiality} means that any state element that changes in one variant
+must either change in both, or be harmless.
+
+Together, these definitions encompass the confidentiality of the caller from the callee.
+
+\paragraph*{Callee Confidentiality}
+
+Although we presented our initial example from the perspective of the caller, a callee
+may also have data that must be kept secret from its caller. Consider a callee that makes
+a privileged system call to obtain a secret key, and uses that key to perform a specific
+task. An untrustworthy or erroneous caller might attempt to read the key out of the callee's
+memory after return.
+
+We model this as nearly dual to caller integrity: the callee's secrets are those
+values that it has stored in \(\unsealed\) elements. So, callee confidentiality means that
+at the state following the callee's return, any element that was \(\unsealed\) under the
+callee's view either retains its original value or is harmless.
+
+\paragraph*{The Hierarchy of Stack Safety}
+
+The forms of stack safety described above are largely independent of one another, and
+can potentially be enforced separately in various combinations. This is important, because
+many enforcement mechanisms already in existence only enforce some of the stack safety
+properties. Stricter enforcement may have greater performance impacts, leading to decisions
+about trade-offs between different points in a hierarchy of properties.
+
+At the most basic, if we protect nothing else, it is vital to protect the control flow of the
+system by enforcing WBCF. None of the data-protection properties are meaningful if a
+callee can simply ``return'' to an arbitrary instruction and proceed with the privileges
+of its caller! And from a practical standpoint, control-flow attacks such as return-oriented
+programming are among the most severe attacks involving the stack and the core motivation
+for much existing work.
+
+\paragraph*{Moving Forward}
+
+In the next section we will define traces and their (hyper-)properties.
+Then in Section \ref{sec:facts} we will define a mechanism to make assertions
+about what will happen on a future return, and Section \ref{sec:props} will use these
+to give the formal definitions of integrity and confidentiality.
 
 \section{Events and Traces}
+\label{sec:events}
 
-Some macro-instructions may be interpreted as {\it events} that can be observed
-from outside the system. For now we define events \(\obs \in \OBSS\) as calls, labeled
-with the target address and the values of the arguments, or else as the silent
-event \(\tau\):
-\[\begin{split}
-\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
-| & ~ \tau \\
-\end{split}\]
-
-We convert control operations into events via the \(Ev\)
-function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.
-
+We abstract over the events that can be observed in the system, defining them
+only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
+event. Other events might represent certain function calls (i.e., system calls)
+or writes to special addresses representing mmapped regions.
 A {\em trace} is a nonempty, finite or infinite sequence
@@
 
-\begin{figure*}
-\[\mathit{Ev} ~ \chi ~ m =
-\begin{cases}
-  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
-  & \textnormal{when } ~ \chi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
-  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
-  & \textnormal{when } ~ \chi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}} \\
-  \tau & \textnormal{otherwise} \\
-\end{cases}\]
-\caption{Converting overlay steps to events}
-\label{fig:observe}
-\end{figure*}
-
 We write that execution from a state produces an observation trace
@@
 
-\judgmenttwo{\((\mach,\context) \stepstounder{e} (\mach',\context')\)}
+\judgmenttwo{\((\mach,\context) \stepstounder{} (\mach',\context',\obs)\)}
             {\((\mach',\context') \hookrightarrow \obsT\)}
-            {\((\mach,\context) \hookrightarrow \notfinished{e}{\obsT}\)}
+            {\((\mach,\context) \hookrightarrow \notfinished{\obs}{\obsT}\)}
 
@@
 
-\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{e} (\mach',\context')\)}
+\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
                   {\(|\sigma| \geq d\)}
                   {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
-                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{e}{\obsT}\)}
+                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{\obs}{\obsT}\)}
 
@@
 
+\paragraph*{Variants and Harmless Values}
+
+Two states are variants with respect to a set of elements, \(\components\),
+if they agree on the value of every element not in \(\components\). Our
+notion of non-interference involves comparing the traces of such
+\(\components\)-variants. We use this to define sets of vestigial elements.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+written \(\mach \approx_\components \nach\), if, for
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
+
+\definition An element set \(\components\) in state \((\mach,\context)\) contains
+harmless values, written \((\mach,\context) \parallel \components\), if for all
+\(\nach\) such that \(\mach \approx_{\components} \mach'\), if 
+\((\mach,\context) \hookrightarrow \obsT\) and
+\((\nach,\context) \hookrightarrow \obsT'\), then
+\(\obsT \eqsim \obsT'\).
+
 \section{Facts Abouts Calls and Returns}
+\label{sec:facts}
 
@@
 
-The intuition behind {\it return-time integrity} (below) is that a caller may expect its
+The intuition behind integrity (below) is that a caller may expect its
 sealed data to be unchanged when control returns to it. In fact, the callee
@@
 this is perfectly legal -- as long as it either restores it, or has some guarantee
-that its changes will impact the caller.
+that its changes will not impact the caller.
 
@@
                   {\((d \uparrow P) ~ (\mach', \context')\)}
-                  {\((\mach, (V,\sigma)) \stepstounder{e} (\mach', \context')\)}
+                  {\((\mach, (V,\sigma)) \stepstounder{\overline{\psi}} (\mach', \context',\obs)\)}
                   {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
 
-Similarly, in {\it caller confidentiality}, we will want to compare future states,
+Similarly, in confidentiality, we will want to compare pairs of future states,
 so we give a binary equivalent, \(d \Uparrow R\), so that
@@
 first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).
+Once again, \(\Uparrow\) is coinductive.
 
@@
               {\(|\sigma_1| \geq d\)}
-              {\((\mach_1,(V_1,\sigma_1)) \stepstounder{e} (\mach_1',\context_1')\)}
+              {\((\mach_1,(V_1,\sigma_1)) \stepstounder{\overline{\psi}} (\mach_1',\context_1',\obs)\)}
               {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
@@
               {\(|\sigma_2| \geq d\)}
-              {\((\mach_2,(V_2,\sigma_2)) \stepstounder{e} (\mach_2',\context_2')\)}
+              {\((\mach_2,(V_2,\sigma_2)) \stepstounder{\overline{\psi}} (\mach_2',\context_2',\obs)\)}
               {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
@@
 \section{Properties}
+\label{sec:props}
 
 We now finally have everything we need to offer our definitions for stack safety.
-We need one preliminary definition first.
-One or more elements of a given state are {\it vestigial} if their values
-in that state have no observable influence on future execution. We define this
-formally as a sort of non-interference: for a state \((\mach,\context)\) and
-a set of elements \(\components\), \((\mach,\context) \parallel \components\)
-holds if we can replace the contents of \(\components\) with arbitrary values
-and have the resulting state produce an equivalent trace. The state created
-by modifying \(\components\) in this way is termed a variant.
+Figure \ref{fig:oprules} gives the operation rules for all of the operations we've described
+so far.
 
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
+\begin{figure}
 
-\definition An element set \(\components\) is vestigial in state \((\mach,\context)\),
-written \((\mach,\context) \parallel \components\), if for all
-\(\nach\) such that \(\mach \approx_{\components} \mach'\), if 
-\((\mach,\context) \hookrightarrow \obsT\) and
-\((\nach,\context) \hookrightarrow \obsT'\), then
-\(\obsT \eqsim \obsT'\).
+  \judgmentbr[Alloc]
+             {\(b = \mach[\SP] + \mathit{off}\)}
+             {\(V' = V \llbracket \addr \mapsto \sealed |
+               b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+             {\(Op ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach = (V',\sigma)\)}
+
+  \judgmentbr[Call]
+             {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
+             {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+               \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+             {\(Op ~ \psi ~ (V,\sigma) ~ \mach =
+               (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}
+
+  \judgment[Ret]
+           {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
+           {\(Op ~ \mathbf{return} ~ (\_, \sigma) ~ \mach = (V, \sigma)\)}
+
+%\judgment[RetDef]
+%         {\(\chi = \mathbf{return}\)}
+%         {\(Cop ~ \chi ~ (V, \emplist) ~ \mach \triangleq
+%           (V, \emplist)\)}
+
+  \caption{Core operation rules for the simple system}
+  \label{fig:oprules}
+\end{figure}
+
+\subsection{Well-bracketed Control Flow}
+
+\definition A system enjoys well-bracketed control flow if, for all states
+\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context)\),
+\(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).
 
 \subsection{Integrity}
-
-And with the preliminary definitions out the way, we can define integrity.
 
@@
 \(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
-Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \downarrow P\) holds
+Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \uparrow P\) holds
 on \((\mach,\context)\).
@@
 \((\mach,\context)\) such that
-\((\mach,\context) \stepstounder{e} (\mach',\context')\) with
-\(e = \mathbf{callE} ~ \addr ~ \overline{\word}\),
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\) with
+\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\) or
+\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \psi\),
 \((\mach',\context')\) enjoys return-time integrity.
@@
 
-Confidentiality properties are also modeled as non-interference, but with a twist:
-the caller's confidential data should be kept secret from the callee, but after return,
-will naturally be accessible to the caller once again. So, we must model confidentiality
-in terms of events inside the callee. But confidentiality violations might not become
-visible during the callee's lifetime -- the callee might read a secret and stash it somewhere
-that it will be accessed later. So we must also consider the machine state at the end
-of the call.
-
-We could define a natural, strict form of confidentiality that requires that the final
-states be identical. However, as with integrity, this is stricter than we need -- there
-may be some conditions in which a secret is copied, but never used again. So, we once
-again wish for our final condition to allow values to be vestigial. Which values?
-
-\definition Let \(\mach,\mach'\) and \(\nach,\nach'\)
-be pairs of machine states. Their {\em corrupted set}, written
-\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all elements
+\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
+of four states treats \(\mach\) and \(\nach\) as the initial states and
+\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
+set of elements that were corrupted between them. That is, the set of all elements
 \(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) such that
@@
 and \(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
+Let \(\nach\) be a \(\components\)-variant of \(\mach\).
+Let \(R\) be a relation such that \((\mach_1,\context_1) ~ R ~ (\mach_2,\context_2)\)
+if \(\bar{\Diamond}(\mach,\mach_1,\nach,\mach_2)\) is harmless in \((\mach_1,\context_1)\).
 Then \((\mach,\context)\) enjoys {\it return-time confidentiality}
-if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\),
-\((\mach,\context) ~ (|\sigma| \Uparrow R) ~ (\nach,\context)\),
-where \((\mach_1,\context_1) ~ R ~ (\nach_2,\context_2)\) if
-\((\mach_1,\context_1) \parallel \bar{\Diamond}(\mach,\mach_1,\nach,\nach_2)\).
+if, for any such \(\nach\), \((\mach,\context) ~ (|\sigma| \Uparrow R) ~ (\nach,\context)\).
 
 \definition A system enjoys {\it caller confidentiality} if, for all reachable states
-\((\mach,\context)\) such that \((\mach,\context) \stepstounder{e} (\mach',\context')\)
-with \(e = \mathbf{callE} ~ \addr ~ \overline{\word}\),
+\((\mach,\context)\) such that
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\) or
+\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \psi\),
 \((\mach',\context')\) enjoys both internal and return-time confidentiality.
@@
 
-Callees can have secrets too. The question is, what do we consider a callee's secrets?
-Anything in its frame? Anything that {\it could} be in its frame (i.e. above the stack pointer)?
-These are plausible options, but they will also capture data that has been written to those
-locations before the call. To get a narrower property specific to the data that is actually
-derived from the callee, we focus instead on those values that the callee has written.
-
 \definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
-\(\components = \{\component | V ~ \component \in \{\sealed,\unsealed\}\}\).
+\(\components = \{\component | V ~ \component = \unsealed\}\).
 Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
@@
 \definition A system enjoys {\it callee confidentiality} if, for all reachable states
-\((\mach,\context)\) such that \((\mach,\context) \stepstounder{e} (\mach',\context')\)
-with \(\psi = \mathbf{callE} ~ \addr ~ \overline{\word}\),
+\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\) or
+\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\),
 \((\mach',\context')\) enjoys post-return confidentiality
@@
 
-Interestingly, callee confidentiality subsumes caller integrity. This makes sense:
-one way for the callee's secrets to leak is by overwriting the caller's sealed data
-with them.
+\section{Shared Memory}
 
-\section{Public Memory}
+%We can extend our basic machine to support public memory allocations that can
+%be accessed by anyone until they are deallocated on their owner's return,
+%using the \(\mathbf{alloc\_pub}\) operation. This behaves identically to
+%\(\mathbf{alloc\_priv}\) except that it marks the locations \(\public\).
+%For example, consider the following code:
 
-We can extend our basic machine to support public memory allocations that can
-be accessed by anyone until they are deallocated on their owner's return,
-using the \(\mathbf{alloc\_pub}\) operation. This behaves identically to
-\(\mathbf{alloc}\) except that it marks the locations \(\public\).
-We should give an example of this, as it may also require us to use multiple
-allocations simultaneously to interleave public and private allocations with a
-single increase of the stack pointer.
+%{\tt
+%  void g() \{
 
-\section{Tail Call Elimination}
+%  ~ ~ int x[3];
 
-I submit that tail calls will work perfectly fine with all of our definitions
-using the context update in Figure \ref{fig:tailcallcontexts}. All of the
-``until return'' stuff works fine, because all the calls return at once.
-We need to define callee confidentiality still.
+%  ~ ~ int y;
 
-\begin{figure*}
+%  ~ ~ h(\&x);
 
-\judgmenttwo[TailCallC]
-            {\(\chi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
-            {\(V' = \mathit{update\_view} ~ V ~ \overline{r_{args}}\)}
-            {\(Cop ~ \chi ~ (V,\sigma) ~ (V',\sigma)\)}
+%  \}
+%}
 
-\caption{Tailcall context updates}
-\label{fig:tailcallcontexts}
-\end{figure*}
+%Here we have allocated several arrays and variables; {\tt y}
+%will behave as normal with \(\mathit{alloc\_priv}\), but {\tt x} has its
+%address taken and passed to another function. That function might do anything at
+%all to {\tt x}, and we have to decide whether it can be said to violate stack
+%safety. A reasonable default is to assume that any address-taken local should be
+%public, while {\tt y} is still private.
 
-\section{Provenance, Capabilities, and Protecting Objects}
+%In the likely event that the space for the return address, {\tt x}, and {\tt y}
+%are all allocated in a single step, that step needs multiple labels to capture
+%that they are not all private. Its label would instead be:
+%\[\begin{split}
+%& \psi = [\mathbf{alloc\_priv} ~ (20,4); \\
+%  & \mathbf{alloc\_pub} ~ (16,12) ; \mathbf{alloc\_priv} ~ (4,4)]
+%\end{split}\]
 
-So far, we have given functions the ability to allocate completely unprotected
-objects. What if we want to express a finer-grained notion of safety, in which
+%Allocating public objects affects the call stack very similarly to private ones,
+%but instead of treating them as \(\sealed\) they become \(\public\). This makes these
+%regions of memory able to transmit information between caller and callee without
+%violating the stack safety properties. Note that there is no way to seal a location
+%once it has become \(\public\); only when {\tt g} returns will the memory allocated
+%to {\tt x} be available for private use.
+
+%\judgmentbr{\(b = \mach[\SP] - \mathit{off}\)}
+%           {\(V' = V \llbracket \addr \mapsto \public |
+%             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+%           {\(Pop ~ (\mathbf{alloc\_pub} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach \triangleq
+%             (V',\sigma)\)}
+
+\subsection{Stack Arguments}
+
+So far, we have not dealt with arguments spilled onto the stack, variadic arguments,
+or arguments that are passed by reference without their addresses being taken
+explicitly (i.e., structs and arrays passed directly.) Consider the simplest case,
+where a function has more than 8 arguments and therefore must (in RISC-V) spill one
+to the stack.
+
+\begin{minipage}[t]{.6\columnwidth}
+{\tt
+  void main() \{
+
+  ~ f(1,2,3,4,5,6,7,8,9);
+
+  ~ g(42);
+
+  \}
+}
+\end{minipage}
+\begin{minipage}[t]{.35\columnwidth}
+{\tt
+0: addi sp,sp,-24
+
+4: sd ra,16(sp)
+
+8: li a5,9
+
+12: sd a5,0(sp)
+
+16: li a7,8
+
+...
+
+48: jal 52,ra
+
+52: li a0,42
+
+56: jal 144,ra
+}
+\end{minipage}
+
+Under a typical calling convention, the callee expects the ninth argument to
+be passed to it at the stack pointer, so the caller places it there
+{\it without moving the stack pointer}. We model this by labeling the transition
+from instruction 48 as a call {\it followed by an allocation}.
+\([\mathbf{call} ~ 100 ~ [\mathtt{ra0-ra7}]; \mathbf{alloc} ~ (0,4)]\).
+
+This has the effect of first pushing a new view onto the call stack, then
+updating the callee's view so that the word following the stack pointer is
+already marked \(\object\).
+
+So, we first apply \(Op ~ (\mathbf{alloc} ~ (0,4))\) and mark that region as
+\(\object\). But we do not want to seal it when we make the call! So we must
+refine our call operation to make use of the information that we have about
+which stack offsets contain arguments, \(\overline{sa}\). We first define the
+helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\) based on those offsets,
+then extend the call operation to keep all objects in \(\mathit{passed}\) marked
+as \(\object\) and seal everything else.
+%
+\[\mathit{passed} ~ \overline{sa} ~ \mach \triangleq
+\bigcup_{(b,o) \in \overline{sa}} \{\mach[\SP]+i | b \leq i < b+o\}\]
+%
+\judgmentbrbr[]
+             {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
+             {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+               \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+             {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+             {\(Op ~ \psi ~ (V,\sigma) ~ \mach =
+               (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}
+
+The remaining definitions are unchanged. Now a stack-spilled argument can be allocated
+and remain accessible in the callee, while its memory is sealed for calls that do not
+use that memory as an argument.
+
+\subsection{Pass-by-reference}
+
+[Still to-do]
+
+\subsection{Provenance, Capabilities, and Protecting Objects}
+
+What if we want to express a finer-grained notion of safety, in which
 such objects are protected unless the function that owns them intentionally
@@
 point to a given object, and non-pointer integers as well as pointers to other objects.
-Then, in \(\mathit{update\_view}\), we recursively identify all public objects that
-are reachable transitively from capabilities in registers; these keep their
-\(\mathit{object}\) status, while those that are not reachable become \(\sealed\).
-
-\section{Simple Concurrency}
+Then \(\mathit{passed}\) should recursively identify all objects that
+are reachable transitively from valid pointers in registers and accessible memory;
+these keep their \(\object\) status, while those that are not reachable become \(\sealed\).
 

### 2022-10-05 submission5/example.c

@@ -1,0 +1,18 @@
+int f() {
+	return 0;
+}
+
+void publish(int x) {
+	return;
+}
+
+void main(int argc, int* argv) {
+	int secret = argv[1];
+	int sensitive[4] = {0,0,0,0};
+	int res = f();
+	if (sensitive[0] == 42)
+		publish(secret);
+	else
+		publish(res);
+	return;
+}

### 2022-10-05 submission5/reference.c

@@ -1,0 +1,13 @@
+struct S {
+	int x;
+};
+
+int f(struct S s) {
+	return 0;
+}
+
+void main() {
+	struct S s;
+	s.x = 0;
+	f(s);
+}

### 2022-10-05 submission5/spill.c

@@ -1,0 +1,13 @@
+int id(int a, int b, int c, int d, int e, int f, int g, int h, int i) {
+	return i;
+}
+
+int id2(int x) {
+	return x;
+}
+
+void main() {
+	id(1,2,3,4,5,6,7,8,9);
+	id2(42);
+	id(1,2,3,4,5,6,7,8,9);
+}

### 2022-10-05 submission5/test.c

@@ -1,0 +1,15 @@
+void swap (int* x, int*y) {
+	int t;
+	t = *x;
+	*x = *y;
+	*y = t;
+	return;
+}
+
+int main() {
+	int a = 1;
+	int b = 2;
+
+	swap(&a, &b);
+	return a;
+}

### 2022-10-06 submission5/noexceptions.tex
comments through III

@@
 to correctly apply tag-based enforcement in practice (cite something about this.)
+\apt{This is too vague about how the  ``distinguishing'' can be performed.
+  Should be clear up front that this will require annotating the code (at least
+  in most cases).}
 
@@
 {\it security operations}. Our properties are phrased as a {\it security semantics},
-which extends the underlying machine with additional information about its
-security principals and which registers and memory they require to be secure, which evolves
+which extends the underlying machine with additional information\apt{``context''?}  about its
+security principals and which registers and memory they require to be secure. This information\apt{``context''?}  evolves
 dynamically through the execution of security operations.
-Specifications may take the form of
-predicates on future states: statements of the form,
-``When control returns to me...'', or of relations on traces of future execution
+Specifications \apt{of what?} may be
+predicates on future states, e.g. assertions of the form
+``When control returns to me...'', or relations on traces of future execution
 (hyper-properties).
 
-This approach is in contrast to approaches that give a safe-by-construction semantics
+Unlike approaches that give a safe-by-construction semantics
 for various security operations and then prove secure compilation to the underlying
-enforcement mechanism, in that our machine is explicitly not safe unless the enforcement
-is correctly instantiated; the security semantics gives us the means to distinguish an
-enforcement scheme that implements stack safety from one that does not.
+enforcement mechanism, our machine is explicitly not safe unless the enforcement
+is correctly instantiated; the security semantics gives us the means to judge whether
+an enforcement scheme implements stack safety or not.
 
@@
 objects are governed by a provenance-based capability system, accessible only to those
-functions which have recieved a valid pointer to them.
+functions which have recieved a valid pointer to them. \apt{ok for now, but may be more
+  than can fit in the final paper}.
 
@@
 those labels correctly. If it recognizes the operations based on the dynamic behavior of
-the system, similarly, those definitions must be realistic.
+the system, similarly, those definitions must be realistic.\apt{The substance of this paragraph
+  is unclear. Isn't it implicit that any scheme must be ``reasonable'' to be useful?
+  All the reader needs to know here is what assumptions we make; then they can decide whether they
+  think these are reasonable or not.}
 
-Otherwise, we do not assume that the code is reasonable or corresponds to any particular
+Otherwise, we do not assume that the code is reasonable or corresponds\apt{wrong word?} to any particular
 calling convention or source construct. In particular, while we are agnostic as to the source
@@
 even buggy code.
+
+\apt{Need to discuss somewhere (maybe not here) what are the consequences of breaking our
+  assumptions by mis-labeling code. Can we salvage anything? Do we care?}
 
@@
 Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
-and are drawn from the set (\(\WORDS\)) (by default, the set of 64-bit integers).
-Registers are ranged over in general by \(\reg \in \REGS\), other than the
+and are drawn from the set \(\WORDS\) (by default\apt{??}, the set of 64-bit integers).
+Registers are ranged over in general by \(\reg \in \REGS\), other than\apt{??} the
 program counter \(\PCname\) and stack pointer \(\SP\).
@@
 {\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
-that defines a mapping \(\mach[\component]\) from state elements to values.
+that defines a mapping \(\mach[\component]\) from state elements to values\apt{which are what? $\WORDS$, right?}.
 {\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\).
+\apt{Maybe want to mention $\tau$ now?}
 The machine step function
@@
 \MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
-is labeled by a list of security operations (order matters), defined below,
-and an event.
+is labeled by an ordered list of security operations, defined below,
+and an event.\apt{Maybe put $\obs$ above the arrow too, as that is traditional.}
 While the tag-based enforcement mechanism may cause the machine to
@@
 our properties will suffer the limitations typical of
-{\it termination insensitive} systems \cite{}.
+{\it termination insensitive} systems \cite{}.\apt{better, but will need fuller
+  explanation later}
 
@@
 and the argument registers. Allocations and deallocations are phrased in terms of a pair
-of an offset and a size. Returns don't require any additional information.
+of an offset \apt{from what?} and a size. Returns don't require any additional information.
 %
@@
 Security principals are function activations: at any given time, the current
-activation, and some number of pending ones. Each activation will have a {\it view}
-of the system that maps each element to one of a set of {\it security classes}:
+activation, and some number of pending ones.\apt{?? Is the principal an activation or
+  a call stack (as defined below)?} Each activation has a {\it view}
+of the system that maps each state element to a{\it security classe}:
 \[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
@@
 
-These indicate, from the perspective of a given principal, whether an element is
+A view indicates, from the perspective of a given principal, whether an element is
 private to an inactive principal (\(\sealed\)),
@@
 
-A pending activation additional carries the addresses at which the program counter
+A pending activation also carries the addresses to which the program counter
 and stack pointer should point in the event of a return.
-We represent a {\it call stack} as a pair of the current activation's view
+We represent a {\it call stack} \apt{``context'' ?} as a pair of the current activation's view
 and a list of pending ones.
@@
 
-From a machine state and a context stack, we create a {\it combined state}
+From a machine state and a context stack\apt{just ``context''?}, we create a {\it combined state}
 \((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
@@
 
+\apt{I don't understand the notation above. What is the fold over? What is $\context''$? Why is the first $\psi$ split out?}
+
 \section{Calls, Returns, and Private Allocations}
@@
 that takes a secret in its arguments and allocates
-an array of potentially sensitive data. It then calls another function {\tt f},
+an array of potentially sensitive data. \apt{why an array? why not just a single value?}
+It then calls another function {\tt f},
 and afterward may decide to {\tt publish} its secret based on the contents of
@@
 
+\apt{C code, assembly code, and picture don't agree!
+  Also, why aren't you showing the storing of ra on the stack as you do in 3(a)? (This is surely needed,
+  and I think knowledgeable readers would expect it; anyhow, should be consistent across figures.)
+  As a presentation thing: in the picture, offset labels on left should be positive, from sp, 0 through 24.}
+
 Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
 to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
-assembly code in Figure \ref{fig:f}. It might take an offset from the stack
+assembly code in Figure \ref{fig:f}.\apt{explicitly reference the subfigure labels.}
+It might take an offset from the stack
 pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
@@
 instead return that value so that {\tt main} makes the call to {\tt publish}.
+\apt{not what is happening in 3(b). (I'm not sure what is supposed to be happening there...)}
 
@@
 Attacks of this kind do not violate {\tt main}'s confidentiality, but its
-{\it integrity}.
+{\it integrity}.s
 And last, but certainly not least, the attacker might attempt to return to the
@@
 At the start of execution the program counter is 0 and the stack pointer -4, with code
-proceeding upward and the stack growing down. The functions {\tt f} and {\tt publish}
-are at addresses 100 and 200, respectively.
+proceeding upward\apt{don't need to say this}  and the stack growing down. The functions {\tt f} and {\tt publish}
+are at addresses 100 and 200, respectively.\apt{not in the assembly code, they aren't!}
 
-Figure \ref{fig:exec1} shows how the program counter, stack pointer, and call stack
+Figure \ref{fig:exec1} shows how the program counter, stack pointer, and call stack \apt{i.e. context?.
+Confusing mix of real machine registers and abstract machine context.}
 develop over several steps, with labels on the steps that allocate space and then make
@@
 
+\apt{It is difficult to track what is going on in thie figure with the original code; we have to keep
+inspecting the pc. It would help to actually annotate the code with relevant security events.}
 The first step  allocates a word for {\tt secret} and four for {\tt sensitive}. To
-maintain a two-word alignment it leaves a gap. This will have the
+maintain a two-word alignment it leaves a gap.\apt{Not sure this is really what is
+  going on in compiled code, but if so, rework to the example to avoid it, as we discussed!} This will have the
 effect of marking those bytes \(\object\), assuming they were previously
@@
 On a call, the formerly active principal's record is pushed onto the inactive list.
-Its return target is the instruction immediately following that of the program
-counter before the call, and the stack pointer target is the current stack pointer.
+Its return target is the return address of the call, 
+and the stack pointer target is the stack pointer at the moment of call.
 The callee's view is updated from the caller's such that the argument
@@
 are \(\unsealed\). All callee-save registers are always \(\sealed\) and
-the program counter \(\public\), as they were initialized.
+the program counter \(\public\), as they were initialized. \apt{Explain shat happens to
+  $\object$s.}
 
@@
 {\tt main}'s secrets are safe as long as {\tt f} does not leak them directly.
+\apt{This is obscure. Are you talking about the jalr instruction serving as
+  a return (and if not, why not?) or about the security label return? Why do
+  you need to ``suppose'' anything; why not just show the code and explain that
+it is annotated?}
 On that return, we restore the topmost inactive principal, {\tt main}'s principal.
@@
 
-But what if {\tt f}, does not return with \(\PCname = 20\) and \(\SP = -24\),
-as would be the case in the fourth example attack? Then it has violated WBCF.
+But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or \(\SP \neq -24\)),
+as would be the case in attack (d)? Then it has violated the WBCF property.
 This property can be phrased in terms of a single step: for every transition
-labeled as a return, the program counter and stack pointer match those recorded
-in the top pending activation.
+labeled as a return, the program counter and stack pointer at the next state
+match those recorded in the top pending activation.
 
@@
 addresses -24 through -9 and -4 through -1 (allocated as private data) and the callee-saved registers.
+\apt{But there are no callee-saved registers in the example, so this is confusing. It would
+  be easy enough to add one (e.g. to hold {\tt secret}), but probably better just to delay telling the
+story about them until later.}
 Note that the callee-saved registers may well have changed during the call -- but if
@@
 callee's frame, but taints that write so that the callee cannot read it later without
-triggering a failstop. In such a scenario we will consider a value {\it harmless}
+triggering a failstop. In such a scenario we will consider a value {\it harmless}\apt{Better, but I still
+  think ``irrelevant'' would be better still. It is the write that is harmless, not the value.}
 if it could be replaced with any other value without changing the observable behavior
-of the machine. That is, for a given state \(\mach\) and set of elements \(\components\),
-we define a {\it variant} state that agrees with \(\mach\) on every element except those
-in \(\components\). Those elements are said to be harmless in \(\mach\) if the variant
-state produces an equivalent trace (see Section \ref{sec:events}.)
+of the machine. More precisely, for a given state \(\mach\) and set of elements \(\components\),
+we define a {\em \(\components\)-variant} state to be one agrees with \(\mach\) on every element except those
+in \(\components\). The elements of \(\components\) are said to be harmless in \(\mach\) if every
+variant state produces an equivalent trace (see Section \ref{sec:events}.)
 
@@
 this is the first state after the call in which the depth of the call stack
-matches the state before the call. Integrity means that, at that return state,
+matches the state before the call. Integrity holds if, at that return state,
 any element that is \(\sealed\) under the callee's view is either restored
 to its original value or is harmless.
-
+\apt{Can this be mapped map to the example?}
 \paragraph*{Caller Confidentiality}
 
+\apt{Remainder of this section is on the right track, but needs some expansion,
+  and to be tied back more clearly and consistently to the examples. I started
+  word-smithing this first para...}
 We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
 means that its callee's behavior is dependent only on publicly visible data,
-not the caller's state. As we saw in our examples, the callee's behavior is defined
-both in terms of events that the callee produces during the call, but also in terms of
-the changes it makes to the state.
+not the caller's private state. As we saw in the examples, we must consider both the observable events
+that the callee produces during the call and the changes that the callee makes to the state that might
+affect the caller after the callee returns.
 
@@
 its return value -- which in turn can lead to future changes in behavior that do not
-even appear to be the fault of the callee!
+even appear to be the fault of the callee!\apt{a bit repeetitious with above}
 {\it Return-time confidentiality} means that any state element that changes in one variant
-must either change in both, or be harmless.
+must either change in both, or be harmless. \apt{Explain this more! and tie back to examples.}
 
@@
 for much existing work.
+
+\apt{Things sort of trail off here...Are there more interesting points on the hierarchy to discuss?
+(And by the way, even WBCF needs to be loosened for exceptions.)}
 
@@
 notion of non-interference involves comparing the traces of such
-\(\components\)-variants. We use this to define sets of vestigial elements.
+\(\components\)-variants. We use this to define sets of vestigial\apt{oops. a meta-vestigial element!}  elements.
 
@@
 harmless values, written \((\mach,\context) \parallel \components\), if for all
-\(\nach\) such that \(\mach \approx_{\components} \mach'\), if 
+\(\nach\) such that \(\mach \approx_{\components} \nach\), if 
 \((\mach,\context) \hookrightarrow \obsT\) and

### 2022-10-10 submission5/noexceptions.tex
Responded to notes

--- 2022-10-05 submission5/noexceptions.tex
@@
 We therefore define our properties under the assumption that we can distinguish the
-steps that represent these abstract operations from those that do not -- an assumption
-that is valid in our testing environment (Section [TBD]) and that is already necessary
+steps that represent these abstract operations from those that do not -- for instance,
+by labeling instructions as calls, returns, etc. Such a labeling scheme appears in
+our testing environment (Section [TBD]) and is already necessary
 to correctly apply tag-based enforcement in practice (cite something about this.)
@@
 {\it security operations}. Our properties are phrased as a {\it security semantics},
-which extends the underlying machine with additional information about its
-security principals and which registers and memory they require to be secure, which evolves
-dynamically through the execution of security operations.
-Specifications may take the form of
-predicates on future states: statements of the form,
-``When control returns to me...'', or of relations on traces of future execution
+which extends the underlying machine with additional context about its
+security principals and which registers and memory they require to be secure.
+This context evolves dynamically through the execution of security operations.
+This notion of security may be phrased as
+predicates on future states, e.g. assertions of the form
+``When control returns to me...'', or relations on traces of future execution
 (hyper-properties).
 
-This approach is in contrast to approaches that give a safe-by-construction semantics
+Unlike approaches that give a safe-by-construction semantics
 for various security operations and then prove secure compilation to the underlying
-enforcement mechanism, in that our machine is explicitly not safe unless the enforcement
-is correctly instantiated; the security semantics gives us the means to distinguish an
-enforcement scheme that implements stack safety from one that does not.
+enforcement mechanism, our machine is explicitly not safe unless the enforcement
+is correctly instantiated; the security semantics gives us the means to judge whether
+an enforcement scheme implements stack safety or not.
 
@@
 
-We must trust that our method of distinguishing security operations is reasonable; if it
+We must trust that our method of distinguishing security operations is accurate; if it
 involves labels placed on code by a compiler, that means trusting that the compiler placed
-those labels correctly. If it recognizes the operations based on the dynamic behavior of
-the system, similarly, those definitions must be realistic.
+those labels correctly. If operations occur that are not recognized, those operations
+might not be guaranteed to protect their principals -- for instance, an unlabeled call
+might not protect the caller's data from the callee. On the flip side, applying an incorrect
+label most likely means that the property becomes too strong to be enforced.
 
-Otherwise, we do not assume that the code is reasonable or corresponds to any particular
+Otherwise, we do not assume that the code is reasonable or derived from any particular
 calling convention or source construct. In particular, while we are agnostic as to the source
@@
 Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
-and are drawn from the set (\(\WORDS\)) (by default, the set of 64-bit integers).
-Registers are ranged over in general by \(\reg \in \REGS\), other than the
+and are drawn from the set \(\WORDS\).
+Registers in the set \(\REGS\) are ranged over by \(\reg\), with special names given to the
 program counter \(\PCname\) and stack pointer \(\SP\).
@@
 {\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
-that defines a mapping \(\mach[\component]\) from state elements to values.
-{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\).
+that defines a mapping \(\mach[\component]\) from state elements to words.
+{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\), with a
+special silent event \(\tau\).
 The machine step function
-\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \in \MACHS \rightarrow
+\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
 \MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
-is labeled by a list of security operations (order matters), defined below,
+is labeled by an ordered list of security operations, defined below,
 and an event.
@@
 our properties will suffer the limitations typical of
-{\it termination insensitive} systems \cite{}.
+{\it termination insensitive} systems \cite{}. TODO: more detailed explanation later.
 
@@
 and the argument registers. Allocations and deallocations are phrased in terms of a pair
-of an offset and a size. Returns don't require any additional information.
+of an offset from the stack pointer and a size. Returns don't require any additional information.
 %
 \begin{align*}
-  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} \\
-  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} \\
+  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
+  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
+  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
+  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
   | & \mathbf{return} & \\
@@
 %
-Security principals are function activations: at any given time, the current
-activation, and some number of pending ones. Each activation will have a {\it view}
-of the system that maps each element to one of a set of {\it security classes}:
+Security principals are function activations: at any given time, pending activations
+have security requirements that are recorded in a call stack, and the current
+activation tracks context information that will inform its requirements when
+it makes a call. Each activation has a {\it view}
+of the system that maps each state element to a{\it security classe}:
 \[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
@@
 
-A pending activation additional carries the addresses at which the program counter
-and stack pointer should point in the event of a return.
-We represent a {\it call stack} as a pair of the current activation's view
+A pending activation also carries the addresses to which the program counter
+and stack pointer should point in the event of a return. We represent the call
+stack with a list of such triples. A {\it context} is a pair of the current activation's view
 and a list of pending ones.
-The initial call stack is \(\context_0 = (V_0, [ ~ ])\).
+The initial context is \(\context_0 = (V_0, [ ~ ])\).
 \[(V, \addr_{ret}, \addr_{sp}) \in \mathit{PEND} ::= \mathit{VIEW \times \WORDS \times \WORDS}\]
@@
 Every security operation manipulates the call  stack via a function
-\(Op : \Psi \rightarrow \MACHS \rightarrow \CONTEXTS \rightarrow \CONTEXTS\). We will introduce the
-curried application of \(Op\) to each operation as we introduce them in detail.
+\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
 
-From a machine state and a context stack, we create a {\it combined state}
+From a machine state and a context, we create a {\it combined state}
 \((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
@@
 \judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
-            {\(\mathit{foldl} ~ (Op ~ \psi ~ \mach) ~ \context = \context'\)}
-            {\((\mach,\context) \stepstounder{\psi :: \overline{\psi}} (\mach', \context'', \obs)\)}
+            {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
+            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context'' \obs)\)}
 
@@
 Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
-that takes a secret in its arguments and allocates
-an array of potentially sensitive data. It then calls another function {\tt f},
+that takes a secret in its arguments and initializes a variable to contain
+potentially sensitive data.
+It then calls another function {\tt f},
 and afterward may decide to {\tt publish} its secret based on the contents of
-its sensitive data.
+its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
+it should not -- instead it will publish the results of {\tt f}.
 
 \begin{figure}
-  \begin{subfigure}[b]{\columnwidth}
+  \begin{subfigure}{\columnwidth}
     {\tt
@@
 
-      ~ ~ int sensitive[4] = \{0,0,0,0\};
+      ~ ~ int sensitive = 0;
 
@@
 
-      ~ ~ if (sensitive[0] == 42) \{
+      ~ ~ if (sensitive == 42) \{
 
@@
   \end{subfigure}
-  \begin{subfigure}[b]{.45\columnwidth}
-    \vspace{2\abovedisplayskip}
-
-    {\tt
-      0: addi sp,sp,-24
-
-      4: sw a1,24(sp)
-
-      8: sd zero,16(sp)
-
-      12: sd zero,8(sp)
-
-      16: jal 84,ra
-
-      20: lw a4,4(sp)
-
-      24: li a5,42
-
-      28: bne a4,a5,L1
-
-      32: lw a0,16(sp)
-
-      36: jal 164,ra
-
-      40: j L2:
-
-      L1:
-
-      44: jal 164,ra
-
-      L2:
-
-      48: addi sp,sp,24
-    }
+  \begin{subfigure}{\columnwidth}
+    \begin{tabular}{r l | l}
+      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
+      \labeledrow{4:}{sd ra,12(sp)}{}
+      \labeledrow{8:}{sw a1,8(sp)}{}
+      \labeledrow{12:}{sw zero,4(sp)}{}
+      \labeledrow{16:}{jal 100,ra}{\(\mathbf{call} ~ 100 ~ \emplist ~ \emplist\)}
+      \labeledrow{20:}{sw r0,0(sp)}{}
+      \labeledrow{24:}{lw a4,4(sp)}{}
+      \labeledrow{28:}{li a5,42}{}
+      \labeledrow{32:}{bne a4,a5,L1}{}
+      \labeledrow{36:}{lw a0,8(sp)}{}
+      \labeledrow{40:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
+      \labeledrow{44:}{j L2:}{}
+      \labeledrow{L1, 48:}{lw r0,0(sp)}{}
+      \labeledrow{52:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
+      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
+      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
+      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
+    \end{tabular}
   \end{subfigure}
-  \begin{subfigure}[b]{.5\columnwidth}
+  \begin{subfigure}{\columnwidth}
     \includegraphics[width=\columnwidth]{stacklayout.png}
@@
 to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
-assembly code in Figure \ref{fig:f}. It might take an offset from the stack
+assembly code in Figure \ref{fig:f}.
+In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
 pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
-subtly, even if somehow prevented from outputting {\tt secret} directly, it might
-instead return that value so that {\tt main} makes the call to {\tt publish}.
+subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
+instead return that value so that {\tt main} makes the call to {\tt publish},
+as in Figure \ref{subfig:indirect}.
 
-Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive[0]}
-with 42, guaranteeing that {\tt main} publishes its own secret unintentionally!
+Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
+with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
+(Figure \ref{subfig:integrity})!
 Attacks of this kind do not violate {\tt main}'s confidentiality, but its
 {\it integrity}.
-And last, but certainly not least, the attacker might attempt to return to the
+And in Figure \ref{subfig:WBCF}, the attacker attempts to return to the
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
@@
 \begin{figure}
-  \begin{subfigure}[b]{.5\columnwidth}
-    {\tt
-      100: addi sp,sp,-4
-
-      104: sw ra,4(sp)
-
-      108: lw a0,20(sp)
-
-      112: jal 88,ra
-
-      116: lw ra,4(sp)
-
-      120: jalr ra
-    }
+  \begin{subfigure}[b]{\columnwidth}
+    \vspace{\abovedisplayskip}
+    \begin{tabular}{r l | l}
+      \labeledrow{100:}{addi sp,sp,-8}{\(\mathbf{alloc} ~ (-8,8)\)}
+      \labeledrow{104:}{sd ra,0(sp)}{}
+      \labeledrow{108:}{lw a0,12(sp)}{}
+      \labeledrow{112:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0} ~ \emplist\}\)}
+      \labeledrow{116:}{lw ra,0(sp)}{}
+      \labeledrow{120:}{addi sp,sp,8}{\(\mathbf{dealloc} ~ (0,8)\)}
+      \labeledrow{124:}{jalr ra}{\(\mathbf{return}\)}
+    \end{tabular}
     \caption{Leaking {\tt secret} directly}
+    \label{subfig:direct}
   \end{subfigure}  
-  \begin{subfigure}[b]{.45\columnwidth}
-    {\tt
-      100: lw a4,16(sp)
-
-      104: sw a4,12(sp)
-
-      108: jalr ra
-    }
+  \begin{subfigure}[b]{\columnwidth}
+    \vspace{\abovedisplayskip}
+    \begin{tabular}{r l | l}
+      \labeledrow{100:}{lw a0,12(sp)}{}
+      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
+    \end{tabular}
     \caption{Leaking {\tt secret} indirectly}
+    \label{subfig:indirect}
   \end{subfigure}  
-
-  \vspace{\belowdisplayskip}
-
-  \begin{subfigure}[b]{.5\columnwidth}
-    {\tt
-      100: li a5,42
-
-      104: sw a5,12(sp)
-
-      108: jalr ra
-    }
+  \begin{subfigure}[b]{\columnwidth}
+    \vspace{\abovedisplayskip}
+    \begin{tabular}{r l | l}
+      \labeledrow{100:}{li a5,42}{}
+      \labeledrow{104:}{sw a5,4(sp)}{}
+      \labeledrow{108:}{jalr ra}{\(\mathbf{return}\)}
+    \end{tabular}
     \subcaption{Attacking {\tt sensitive}}
+    \label{subfig:integrity}
   \end{subfigure}
-  \begin{subfigure}[b]{.45\columnwidth}
-    {\tt
-      100: addi ra,ra,20
-
-      104: jalr ra
-    }
+  \begin{subfigure}[b]{\columnwidth}
+    \vspace{\abovedisplayskip}
+    \begin{tabular}{r l | l}
+      \labeledrow{100:}{addi ra,ra,20}{}
+      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
+    \end{tabular}
     \subcaption{Attacking control flow}
+    \label{subfig:WBCF}
   \end{subfigure}
@@
 
-At the start of execution the program counter is 0 and the stack pointer -4, with code
-proceeding upward and the stack growing down. The functions {\tt f} and {\tt publish}
+At the start of execution the program counter is 0 and the stack pointer -4, with the
+stack growing down. The functions {\tt f} and {\tt publish}
 are at addresses 100 and 200, respectively.
 
-Figure \ref{fig:exec1} shows how the program counter, stack pointer, and call stack
+Figure \ref{fig:exec1} shows how the program counter, stack pointer, and context
 develop over several steps, with labels on the steps that allocate space and then make
@@
 
-The first step allocates a word for {\tt secret} and four for {\tt sensitive}. To
-maintain a two-word alignment it leaves a gap. This will have the
+The first step  allocates words for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
+as two words for the return address. This will have the
 effect of marking those bytes \(\object\), assuming they were previously
@@
     \vskip 0pt
-    \begin{tabular}{| l | l | l | l |}
+    \begin{tabular}{| l | l | l || l |}
       \hline
@@
       \hline
-      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-4,4);\mathbf{alloc} ~ (-24,16)]\)} \\
+      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
       \hline
-      1 & 4 & -24 & \(V_1 = V_0 \llbracket -24 \ldots -9, \mapsto \object \rrbracket\) \\
-          & & & \(\llbracket -4 \ldots -1, \mapsto \object \rrbracket, \emplist\)\\
+      1 & 4 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket\) \\
       \hline
@@
       \hline
-      4 & 16 & -24 & \(V_1, \emplist\) \\
+      4 & 16 & -20 & \(V_1, \emplist\) \\
       \hline
@@
       \hline
-      5 & 100 & -24 & \(V_2 = V_1 \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket,\) \\
-      & & & \([(V_1, 20, -24)]\) \\
+      5 & 100 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[(V_1, 20, -20)]\) \\
       \hline
@@
       \hline
-      \(n\) & ? & ? & \(V_2,[(V_1, 20, -24)]\) \\
+      \(n\) & ? & ? & \(V_2,[(V_1, 20, -20)]\) \\
       \hline
@@
       \hline
-      \(n+1\) & 20 & -24 & \(V_1, \emplist\) \\
+      \(n+1\) & 20 & -20 & \(V_1, \emplist\) \\
       \hline
@@
 On a call, the formerly active principal's record is pushed onto the inactive list.
-Its return target is the instruction immediately following that of the program
-counter before the call, and the stack pointer target is the current stack pointer.
-The callee's view is updated from the caller's such that the argument
-registers are \(\public\) and any non-argument, caller-saved registers
-are \(\unsealed\). All callee-save registers are always \(\sealed\) and
+Its return target is the return address of the call, 
+and the stack pointer target is the stack pointer at the moment of call.
+The callee's view is updated from the caller's such that all \(\object\) locations
+become \(\sealed\). (In more complex settings, objects that are intentionally passed
+to the callee will not get sealed, but for now we assume no sharing of memory.)
+For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments,
+in this case {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
+registers are mapped to \(\unsealed\). All callee-save registers are always \(\sealed\) and
 the program counter \(\public\), as they were initialized.
 
-We suppose that, on step \(n+1\), {\tt f} returns; if it does not, then
-{\tt main}'s secrets are safe as long as {\tt f} does not leak them directly.
-On that return, we restore the topmost inactive principal, {\tt main}'s principal.
+On step \(n\), {\tt f} returns, and we restore the topmost inactive principal,
+{\tt main}'s principal.
 
@@
 
-But what if {\tt f}, does not return with \(\PCname = 20\) and \(\SP = -24\),
-as would be the case in the fourth example attack? Then it has violated WBCF.
+But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or \(\SP \neq -24\)),
+as would be the case in attack (d)? Then it has violated the WBCF property.
 This property can be phrased in terms of a single step: for every transition
-labeled as a return, the program counter and stack pointer match those recorded
-in the top pending activation.
+labeled as a return, the program counter and stack pointer at the next state
+match those recorded in the top pending activation. The table in Figure \ref{fig:exec1}
+assumes that WBCF is followed.
 
@@
 In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
-addresses -24 through -9 and -4 through -1 (allocated as private data) and the callee-saved registers.
+addresses -20 through -1 (allocated as private data) and callee-saved registers such as
+the stack pointer and return address.
 Note that the callee-saved registers may well have changed during the call -- but if
@@
 callee's frame, but taints that write so that the callee cannot read it later without
-triggering a failstop. In such a scenario we will consider a value {\it harmless}
+triggering a failstop. In such a scenario we will consider a value {\it irrelevant}
 if it could be replaced with any other value without changing the observable behavior
-of the machine. That is, for a given state \(\mach\) and set of elements \(\components\),
-we define a {\it variant} state that agrees with \(\mach\) on every element except those
-in \(\components\). Those elements are said to be harmless in \(\mach\) if the variant
-state produces an equivalent trace (see Section \ref{sec:events}.)
+of the machine. More precisely, for a given state \(\mach\) and set of elements \(\components\),
+we define a {\em \(\components\)-variant} state to be one agrees with \(\mach\) on every element except those
+in \(\components\). The elements of \(\components\) are said to be irrelevant in \(\mach\) if every
+variant state produces an equivalent trace (see Section \ref{sec:events}.)
 
@@
 this is the first state after the call in which the depth of the call stack
-matches the state before the call. Integrity means that, at that return state,
+matches the state before the call. Integrity holds if, at that return state,
 any element that is \(\sealed\) under the callee's view is either restored
-to its original value or is harmless.
+to its original value or is irrelevant.
+
+Returning to our example: at step 4, when {\tt main} makes its call to {\tt f},
+the call stack in the context is empty. It makes its call, and at step 5, {\tt main}'s
+frame, including {\tt secret} and {\tt sensitive}, is mapped to \(\sealed\) from the
+perspective of {\tt f}. So, the next time the call stack is empty (implicitly, when
+we have returned to {\tt main}), elements of that frame must either match their
+values at step 4, or be irrelevant to future computation.
 
@@
 
+\apt{Remainder of this section is on the right track, but needs some expansion,
+  and to be tied back more clearly and consistently to the examples. I started
+  word-smithing this first para...}
 We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
 means that its callee's behavior is dependent only on publicly visible data,
-not the caller's state. As we saw in our examples, the callee's behavior is defined
-both in terms of events that the callee produces during the call, but also in terms of
-the changes it makes to the state.
+not the caller's private state. As we saw in the examples, we must consider both the observable events
+that the callee produces during the call and the changes that the callee makes to the state that might
+affect the caller after the callee returns.
 
@@
 
-But, it is also possible that the callee could extract data from the caller that it does
-not use to produce a visible event, but instead stashes for later or uses to influence
-its return value -- which in turn can lead to future changes in behavior that do not
-even appear to be the fault of the callee!
-{\it Return-time confidentiality} means that any state element that changes in one variant
-must either change in both, or be harmless.
+But, in Figure \ref{subfig:indirect}, we also saw an attacker that exfiltrated the secret
+by reading it and then returning it, in a context where the caller would output the returned
+value. In situations like this, we care about {\it return-time confidentiality}.
+In that case, we would identify that in our original execution, {\tt a0} changed between
+the call and return, being set to the value of {\tt secret}. But in a variant state,
+{\tt secret} might have a different value, and so will {\tt a0} at the return.
+This difference, found by comparing the initial and final states of both executions,
+indicates that {\tt a0} is corrupted. Return-time confidentiality means that
+an element that is corrupted in this way must be irrelevant.
 
@@
 at the state following the callee's return, any element that was \(\unsealed\) under the
-callee's view either retains its original value or is harmless.
+callee's view either retains its original value or is irrelevant.
 
@@
 At the most basic, if we protect nothing else, it is vital to protect the control flow of the
-system by enforcing WBCF. None of the data-protection properties are meaningful if a
-callee can simply ``return'' to an arbitrary instruction and proceed with the privileges
-of its caller! And from a practical standpoint, control-flow attacks such as return-oriented
-programming are among the most severe attacks involving the stack and the core motivation
-for much existing work.
+system by enforcing WBCF or a similar property. None of the data-protection properties are
+meaningful if a callee can simply ``return'' to an arbitrary instruction and proceed with
+the privileges of its caller! And from a practical standpoint, control-flow attacks such
+as return-oriented programming are among the most severe attacks involving the stack and
+the core motivation for much existing work.
+
+That said, many systems will not enforce WBCF, as it is insufficient to capture the
+control flow of features such as exceptions. So, the data-protection properties are
+technically orthogonal to WBCF. It remains the case that an appropriate control-flow
+property (extended with relevant control-flow features) is a vital base to any stack safety
+enforcement.
+
+Once control-flow is protected, we can think about integrity and confidentiality.
+Integrity is perhaps a more obviously useful property: major proposed formalizations
+of stack safety have omitted confidentiality in favor of combining WBCF and integrity
+\cite{SkorstengaardSTKJFP}. This is because integrity is directly tied to compiler correctness
+in a way that confidentiality is not: compilers may optimize based on assumptions
+about the invariance of a caller's frame over a call, and without integrity such optimizations
+are in fact unsound.
+
+Finally, at the strongest end of the hierarchy, we might enforce safe control flow,
+integrity, and confidentiality of both kinds.
 
@@
 
-\paragraph*{Variants and Harmless Values}
+\paragraph*{Variants and Irrelevant Values}
 
@@
 notion of non-interference involves comparing the traces of such
-\(\components\)-variants. We use this to define sets of vestigial elements.
+\(\components\)-variants. We use this to define sets of irrelevant elements.
 
@@
 \definition An element set \(\components\) in state \((\mach,\context)\) contains
-harmless values, written \((\mach,\context) \parallel \components\), if for all
-\(\nach\) such that \(\mach \approx_{\components} \mach'\), if 
+irrelevant values, written \((\mach,\context) \parallel \components\), if for all
+\(\nach\) such that \(\mach \approx_{\components} \nach\), if 
 \((\mach,\context) \hookrightarrow \obsT\) and
@@
                b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
-             {\(Op ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach = (V',\sigma)\)}
+             {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
 
@@
                \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-             {\(Op ~ \psi ~ (V,\sigma) ~ \mach =
+             {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
                (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}
@@
            {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-           {\(Op ~ \mathbf{return} ~ (\_, \sigma) ~ \mach = (V, \sigma)\)}
+           {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
 
@@
 \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\) with
-\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\) or
-\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \psi\),
+\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
+\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
 \((\mach',\context')\) enjoys return-time integrity.
@@
 Let \(R\) be a relation such that \((\mach_1,\context_1) ~ R ~ (\mach_2,\context_2)\)
-if \(\bar{\Diamond}(\mach,\mach_1,\nach,\mach_2)\) is harmless in \((\mach_1,\context_1)\).
+if \(\bar{\Diamond}(\mach,\mach_1,\nach,\mach_2)\) is irrelevant in \((\mach_1,\context_1)\).
 Then \((\mach,\context)\) enjoys {\it return-time confidentiality}
@@
 \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\) or
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
 \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \psi\),
@@
 \((\mach,\context)\) such that \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\) or
-\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \psi\),
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
+\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
 \((\mach',\context')\) enjoys post-return confidentiality
@@
 
-\begin{minipage}[t]{.6\columnwidth}
 {\tt
@@
 }
-\end{minipage}
-\begin{minipage}[t]{.35\columnwidth}
-{\tt
-0: addi sp,sp,-24
 
-4: sd ra,16(sp)
+\vspace{\abovedisplayskip}
 
-8: li a5,9
-
-12: sd a5,0(sp)
-
-16: li a7,8
-
-...
-
-48: jal 52,ra
-
-52: li a0,42
-
-56: jal 144,ra
-}
-\end{minipage}
+\begin{tabular}{r l | l}
+  \labeledrow{0:}{addi sp,sp,-12}{\(\mathbf{alloc} ~ (-12,12)\)}
+  \labeledrow{4:}{sd ra,4(sp)}{}
+  \labeledrow{8:}{li a5,9}{}
+  \labeledrow{12:}{sd a5,0(sp)}{}
+  \labeledrow{16:}{li a7,8}{}
+  \dots \\
+  \labeledrow{48:}{jal 100,ra}{\(\mathbf{call} ~ \{\mathtt{a0-a7}\} ~ \{(\SP,0,4)\}\)}
+  \labeledrow{52:}{li a0,42}{}
+  \labeledrow{56:}{jal 200,ra}{\(\mathbf{call} ~ \{\mathtt{a0}\} ~ \emplist\)}
+  \dots \\
+\end{tabular}
 
 Under a typical calling convention, the callee expects the ninth argument to
-be passed to it at the stack pointer, so the caller places it there
-{\it without moving the stack pointer}. We model this by labeling the transition
-from instruction 48 as a call {\it followed by an allocation}.
-\([\mathbf{call} ~ 100 ~ [\mathtt{ra0-ra7}]; \mathbf{alloc} ~ (0,4)]\).
+be passed to it at bottom of the caller's frame, so the caller allocates extra
+space for it. Then the call to {\tt f} identifies that we expect a range of four
+bytes from the stack pointer to be passed as an object instead of sealed.
 
-This has the effect of first pushing a new view onto the call stack, then
-updating the callee's view so that the word following the stack pointer is
-already marked \(\object\).
-
-So, we first apply \(Op ~ (\mathbf{alloc} ~ (0,4))\) and mark that region as
-\(\object\). But we do not want to seal it when we make the call! So we must
-refine our call operation to make use of the information that we have about
-which stack offsets contain arguments, \(\overline{sa}\). We first define the
-helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\) based on those offsets,
+So we must refine our call operation to make use of the information that we have about
+which memory contain arguments, \(\overline{sa}\). \(\overline{sa}\) is a set of
+triples of a register, an offset from the value of that register, and a size.
+We first define the helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\),
 then extend the call operation to keep all objects in \(\mathit{passed}\) marked
@@
 \[\mathit{passed} ~ \overline{sa} ~ \mach \triangleq
-\bigcup_{(b,o) \in \overline{sa}} \{\mach[\SP]+i | b \leq i < b+o\}\]
+\bigcup_{(\reg,b,o) \in \overline{sa}} \{\mach[\reg]+i | b \leq i < b+o\}\]
 %
@@
              {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-             {\(Op ~ \psi ~ (V,\sigma) ~ \mach =
+             {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
                (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}
@@
 
-\subsection{Pass-by-reference}
+This same structure works for pass-by-reference arguments: we know from the type of
+the called function which register it expects to hold the address of a given object,
+and the size of that object. In this case the offset will always be 0.
 
-[Still to-do]
+When the callee in such a scenario makes a call of its own, the objects passed to it
+will in turn be sealed -- unless it passes them on, which can happen with pass-by-reference
+arguments!
 

### 2022-10-10 submission5/noexceptions.tex
Delete random comment

@@
 
-\apt{Remainder of this section is on the right track, but needs some expansion,
-  and to be tied back more clearly and consistently to the examples. I started
-  word-smithing this first para...}
 We treat confidentiality as a form of non-interference as well: the confidentiality of a caller

### 2022-10-19 submission5/noexceptions.tex
More hierarchy

@@
 properties. Stricter enforcement may have greater performance impacts, leading to decisions
-about trade-offs between different points in a hierarchy of properties.
+about trade-offs between different degrees of stack safety.
+
+In principal, these properties can exist in any combination, but in practice we see examples
+of certain combinations that form a hierarchy. Many enforcement techniques focus purely on
+well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
+address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
+to enforce it completely.
+
+Others combine this protection with some degree of memory protection,
+chiefly focusing on integrity. Skorstengaard et al. \cite{SkorstengaardSTK} characterize stack
+safety as the combination of WBCF and {\it local state encapsulation} (LSE), and prove that their
+capability-based enforcement satisfies both. LSE corresponds to integrity, plus some degree of
+confidentiality. They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
+together as a safe-by-construction abstract machine. That machine does not allow a callee to
+read or write its caller's frame, but might allow it to read data that is left behind below
+the caller's frame, which is weaker than our full caller confidentiality and rules out
+callee confidentiality entirely. Georges et al. \cite{Georges22:TempsDesCerises} correct
+both limitations.
 

### 2022-10-19 submission5/noexceptions.tex
Small changes
More example discussion in definitions

@@
 of the machine. More precisely, for a given state \(\mach\) and set of elements \(\components\),
-we define a {\em \(\components\)-variant} state to be one agrees with \(\mach\) on every element except those
-in \(\components\). The elements of \(\components\) are said to be irrelevant in \(\mach\) if every
-variant state produces an equivalent trace (see Section \ref{sec:events}.)
+we define a {\em \(\components\)-variant} state to be one agrees with \(\mach\) on every
+element except those in \(\components\). The elements of \(\components\) are said to be
+irrelevant in \(\mach\) if every variant state produces an equivalent trace
+(see Section \ref{sec:events}.)
 
@@
 \definition A system enjoys well-bracketed control flow if, for all states
-\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context)\),
+\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context,\obs)\),
 \(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).
+
+\vspace{\abovedisplayskip}
+
+Consider again the attack in Figure \ref{subfig:WBCF}: the attacker adds
+20 to the return address and then returns, thus bypassing the check and publishing
+{\tt secret}. During the call step, we record in the context that the return to the
+caller should have the program counter at the next instruction (address 20) and the
+stack pointer restored to address -20. When the return occurs, we see that in the
+state following it, the program counter is 40 -- a violation of WBCF.
+
+To demonstrate why WBCF also cares about the stack pointer, consider an attack which
+returns with the stack pointer at -12. Now, compare the layout of the stack relative
+to the stack pointer as it should be (top) to the one that results by increasing the
+stack pointer before return (bottom):
+
+\begin{tabular}{| l | l | l | l | l | l | l |}
+  \multicolumn{3}{r}{\(\SP \downarrow\)} \\
+  \hline
+  & & res & sens & secr & ra & ra \\
+  \hline
+  \hline
+  res & sens & secr & ra & ra & & \\
+  \hline
+\end{tabular}
+
+\vspace{\abovedisplayskip}
+
+In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
+read part of the return address, and then it will attempt to publish
+{\tt res}, but instead will publish {\tt secret}!
+
+Alternatively, we can formalize well-bracketed control flow using \(\uparrow\).
+
+\definition For a machine state \(\mach\), let
+\(ret ~ (\mach,\context) = \{\mach' | \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]\}\).
+A system enjoys well-bracketed control flow if, for all states 
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
+\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
+\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+\(|\sigma| \uparrow ret(\mach')\).
+
+\vspace{\abovedisplayskip}
+
+Once again we can think about our example attack. When the call occurs,
+we define a predicate \(ret ~ (\mach,\context)\) on states that holds when the program counter
+and stack pointer have the appropriate values. Then we use apply the \(\uparrow\)
+operator to this predicate with the depth of the call stack {\it after the call} and
+the machine state {\it before the call}.
+
+After the call, the call stack is one higher than it was before the call,
+and if the callee makes further calls, it will continue to be incremented.
+Each return will similarly decrement it, until it is finally lower than the
+initial state. This is the return that matches the original call, and so
+the resulting state is when we require \(ret(\mach)\) to hold.
 
@@
 \(\components = \{\component | V ~ \component = \sealed\}\).
-Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
-\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
+Let \(P ~ (\mach',\context')\) be a predicate that holds on a state
+\(\Delta(\mach,\mach')\) is irrelevant in \((\mach', \context')\).
 Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \uparrow P\) holds
@@
 \((\mach',\context')\) enjoys return-time integrity.
+
+\vspace{\abovedisplayskip}
+
+The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
+which is \(\sealed\). At the call, we define the predicate \(P\), which holds on states
+where any \(\sealed\) elements that changed since the call are all irrelevant.
+As {\tt sensitive} is such an element (conveniently, the only one), the property
+has not yet been violated at the returned state \((\mach,\context)\). To determine whether this
+attack has violated stack integrity, we must consider an arbitrary variant state
+\(\nach = \mach[\addr_{\mathtt{sensitive}} \mapsto \word]\) for some \(\word\).
+The property is violated if the traces of \((\mach,\context)\) and \((\nach,\context)\)
+differ.
+
+In our case, {\tt publish} produces an observation, and if \(\word \not = 42\),
+that observation in the trace from \((\nach,\context)\) will be {\tt res}, which is 0.
+On the other hand, \((\mach,\context)\) will produce an event from {\tt secret},
+which may differ. So, integrity is indeed violated.
 
@@
 \((\mach',\context')\) enjoys both internal and return-time confidentiality.
+
+\vspace{\abovedisplayskip}
+
+Confidentiality is more complicated. We begin with a state immediately following
+a call. Let that be \((\mach,\context)\). We suppose an arbitrary variant state,
+\((\nach,\context)\), which may vary any element that is \(\sealed\) or \(\unsealed\).
+That is, any element that is not used legitimately to pass arguments. Caller confidentiality
+therefore can be thought of as the callee's insensitivity to parts of its initial states
+that are not part of the caller-callee interface.
+
+In the example in Figure \ref{subfig:direct}, this variation will quickly result
+in differing events, as {\tt publish} is called with arguments that may have different
+values. This is clearly a violation of confidentiality, in the form of internal
+confidentiality.
+
+When the variants return, however, the situation becomes more complicated.
+Consider Figure \ref{subfig:indirect}. The value of {\tt secret} is been extracted
+and placed in {\tt a0}, the return value register. Suppose that we have \(\mach\)
+and \(\nach\) as above, and they step, respectively, to \(\mach'\) and \(\nach'\),
+which each follow a return. We must identify which elements of the state
+contain leaked confidential data.
+
+An element contains confidential data, according to the definition of non-interference, 
+if it differs between the states. But some of those data belong to the original
+elements which are accessible to the caller and may now influence future execution
+legitimately. We are looking specifically for those elements that have {\it changed}
+from call to return, and that vary between the variants. As with integrity, we then
+must determine whether those leaks actually translate into events. If the set of
+such elements is not irrelevant, confidentiality is violated.
 

### 2022-10-19 submission5/noexceptions.tex
Fix hierarchy section

@@
 
-At the most basic, if we protect nothing else, it is vital to protect the control flow of the
-system by enforcing WBCF or a similar property. None of the data-protection properties are
-meaningful if a callee can simply ``return'' to an arbitrary instruction and proceed with
-the privileges of its caller! And from a practical standpoint, control-flow attacks such
-as return-oriented programming are among the most severe attacks involving the stack and
-the core motivation for much existing work.
-
-That said, many systems will not enforce WBCF, as it is insufficient to capture the
-control flow of features such as exceptions. So, the data-protection properties are
-technically orthogonal to WBCF. It remains the case that an appropriate control-flow
-property (extended with relevant control-flow features) is a vital base to any stack safety
-enforcement.
-
-Once control-flow is protected, we can think about integrity and confidentiality.
-Integrity is perhaps a more obviously useful property: major proposed formalizations
-of stack safety have omitted confidentiality in favor of combining WBCF and integrity
-\cite{SkorstengaardSTKJFP}. This is because integrity is directly tied to compiler correctness
-in a way that confidentiality is not: compilers may optimize based on assumptions
-about the invariance of a caller's frame over a call, and without integrity such optimizations
-are in fact unsound.
-
-Finally, at the strongest end of the hierarchy, we might enforce safe control flow,
-integrity, and confidentiality of both kinds.
-
 \paragraph*{Moving Forward}

### 2022-10-20 submission5/noexceptions.tex
more comments on I-V

@@
   \IEEEauthorblockN{
-    Benjamin C. Pierce,
+    Benjamin C. Pierce
   }
@@
 We therefore define our properties under the assumption that we can distinguish the
-steps that represent these abstract operations from those that do not -- for instance,
-by labeling instructions as calls, returns, etc. Such a labeling scheme appears in
+steps that represent these abstract operations from those that do not---for instance,
+by labeling instructions as calls, returns, etc.\apt{note proper way to do punctuating dashes}
+Such a labeling scheme appears in
 our testing environment (Section [TBD]) and is already necessary
@@
 those labels correctly. If operations occur that are not recognized, those operations
-might not be guaranteed to protect their principals -- for instance, an unlabeled call
+might not be guaranteed to protect their principals---for instance, an unlabeled call
 might not protect the caller's data from the callee. On the flip side, applying an incorrect
@@
 
-Otherwise, we do not assume that the code is reasonable or derived from any particular
+Otherwise, we do not assume that the code is reasonable\apt{still vague} or adheres to any particular
 calling convention or source construct. In particular, while we are agnostic as to the source
@@
 it makes a call. Each activation has a {\it view}
-of the system that maps each state element to a{\it security classe}:
+of the system that maps each state element to a {\it security class}:
 \[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
@@
             {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
-            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context'' \obs)\)}
-
-\section{Calls, Returns, and Private Allocations}
+            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context', \obs)\)}
+
+%\section{Calls, Returns, and Private Allocations}
+\section{Examples}
 
@@
 that takes a secret in its arguments and initializes a variable to contain
-potentially sensitive data.
+potentially sensitive data.\apt{Arguments to {\tt main} are overly complicated.
+  There's no need to match reality (and in fact {\tt argv} is really a {\tt char*}
+  anyway).}\apt{As we discussed, it would be simpler to represent ``publish'' by
+  writing to a specified global rather than with a call.}
 It then calls another function {\tt f},
@@
 its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
-it should not -- instead it will publish the results of {\tt f}.
+it should not---instead it will publish the results of {\tt f}.
+\apt{Would be better to use symbolic names instead of inetegers for {\tt jal}
+  targets. Figure at bottom is good, but also needs additional labelling to
+  show addresses relative to the new sp within {\tt f}.}
 
@@
 assembly code in Figure \ref{fig:f}.
+\apt{In (a,b), {\tt 12(sp)} maps to {\tt sensitive}, not {\tt secret} (I think!).
+  It would be nice if (a) didn't have to push a frame (i.e. again by making publication
+  just a global write) and if all the examples of {\tt f} were the same length, e.g.
+  by adding {\tt nop}s.}
+
 In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
@@
 
-At the start of execution the program counter is 0 and the stack pointer -4, with the
-stack growing down. The functions {\tt f} and {\tt publish}
+\apt{This para is a bit of an orphan.}
+At the start of execution the program counter is 0 and the stack pointer -4 \apt{or is it 0?},
+with the stack growing down. The functions {\tt f} and {\tt publish}
 are at addresses 100 and 200, respectively.
+
+\apt{Need a smoother transition here. Be explicit that we are now going to start considering the security
+  state.}
 
@@
 develop over several steps, with labels on the steps that allocate space and then make
-a call. \(\downarrow \overline{\psi}\) in this diagram is a sideways
-\(\xrightarrow{\overline{\psi}}\). Events are omitted.
+a call. \(\xrightarrow{\overline{\psi}}\) transitions are written \(\downarrow \overline{\psi}\) in this diagram;
+events are omitted.\apt{graphics on right of figure need to be keyed to corresponding steps.}
 
@@
 become \(\sealed\). (In more complex settings, objects that are intentionally passed
-to the callee will not get sealed, but for now we assume no sharing of memory.)
+to the callee will not get sealed, but for now we assume no sharing of memory between activations.)
 For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments,
@@
 On step \(n\), {\tt f} returns, and we restore the topmost inactive principal,
-{\tt main}'s principal.
+{\tt main}'s principal. \apt{This $n$ is a bit mysterious. Again, if all the snippets
+  were the same length, you could use a concrete number here.}
+
+\apt{Need a bridging sentence here, along the lines of: ``Now we consider how to define desirable
+  stack safety properties that rule out various kinds of bad behavior.''}
 
@@
 
-But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or \(\SP \neq -24\)),
+But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or \(\SP \neq -24\))\apt{-20?},
 as would be the case in attack (d)? Then it has violated the WBCF property.
@@
 the stack pointer and return address.
-Note that the callee-saved registers may well have changed during the call -- but if
+Note that the callee-saved registers may well have changed during the call---but if
 so, the callee is obligated to restore them to their prior value.
@@
 One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
-in a way that is harmless -- or that is rendered harmless by the enforcement mechanism!
-This is the case in a lazy micro-policy: the monitor permits the caller to write into the
+in a way that is harmless---or that is rendered harmless by the enforcement mechanism!
+\apt{This is still confusing, because the ``rendered harmless'' part applies just as much
+  to the strict integrity we've been discussing up to now.}
+This is the case in a lazy micro-policy\apt{presumes we've described these a bit in the introduction
+  (which we should)}: the monitor permits the caller to write into the
 callee's frame, but taints that write so that the callee cannot read it later without
@@
 we have returned to {\tt main}), elements of that frame must either match their
-values at step 4, or be irrelevant to future computation.
+values at step 4, or be irrelevant to future computation.\apt{picture?}
 
@@
 means that its callee's behavior is dependent only on publicly visible data,
-not the caller's private state. As we saw in the examples, we must consider both the observable events
+not the caller's private state. \apt{Spell out the implications of this for initialization.}
+As we saw in the examples, we must consider both the observable events
 that the callee produces during the call and the changes that the callee makes to the state that might
@@
 
-Together, these definitions encompass the confidentiality of the caller from the callee.
+Together, these definitions encompass the confidentiality of the caller from the callee.\apt{picture?}
 
@@
 can potentially be enforced separately in various combinations. This is important, because
-many enforcement mechanisms already in existence only enforce some of the stack safety
+many existing enforcement mechanisms enforce only some of the stack safety
 properties. Stricter enforcement may have greater performance impacts, leading to decisions
@@
 capability-based enforcement satisfies both. LSE corresponds to integrity, plus some degree of
-confidentiality. They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
+confidentiality.\apt{Be more explicit.}  They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
 together as a safe-by-construction abstract machine. That machine does not allow a callee to
@@
 sealed data to be unchanged when control returns to it. In fact, the callee
-may overwrite such data -- when the data are found in callee-saved registers
-this is perfectly legal -- as long as it either restores it, or has some guarantee
+may overwrite such data---when the data are found in callee-saved registers
+this is perfectly legal---as long as it either restores it, or has some guarantee
 that its changes will not impact the caller.
@@
 We start by defining a second-order logical operator
-\(d \uparrow P\), read ``\(P\) holds on return from depth \(d\),''
+\(d \uparrow P\), pronounced ``\(P\) holds on return from depth \(d\),''
 where \(P\) is a predicate on machine states. This is a coinductive relation
-similar to ``weak until'' in temporal logic -- it holds if the program never
+similar to ``weak until'' in temporal logic---it also holds if the program never
 returns from depth \(d\).
@@
 
-Similarly, in confidentiality, we will want to compare pairs of future states,
-so we give a binary equivalent, \(d \Uparrow R\), so that
+Similarly, for confidentiality, we will want to compare pairs of future states,
+so we give a binary equivalent, \(d \Uparrow R\), where
 \((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
-first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).
-Once again, \(\Uparrow\) is coinductive.
+first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\),
+respectively. Once again, \(\Uparrow\) is coinductive.
 
@@
 
-We now finally have everything we need to offer our definitions for stack safety.
+We now finally have everything we need to formalize our definitions for stack safety.
 Figure \ref{fig:oprules} gives the operation rules for all of the operations we've described
@@
 
-\definition A system enjoys well-bracketed control flow if, for all states
-\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context,\obs)\),
-\(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).
+\definition A system enjoys {\it well-bracketed control flow} if 
+\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context,\obs)\)
+implies \(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).
 
 \vspace{\abovedisplayskip}
+
+\apt{These explanations that relate the definitions back to the example snippets are
+  great. They would be even better if you can find a way to put them in Section II,
+  based on (less formal) definitions of the terms introduced here.}
 
@@
 stack pointer restored to address -20. When the return occurs, we see that in the
-state following it, the program counter is 40 -- a violation of WBCF.
+state following it, the program counter is 40---a violation of WBCF.
 
 To demonstrate why WBCF also cares about the stack pointer, consider an attack which
-returns with the stack pointer at -12. Now, compare the layout of the stack relative
+returns with the stack pointer at -12 instead of -20. Now, compare the layout of the stack relative
 to the stack pointer as it should be (top) to the one that results by increasing the
@@
 \end{tabular}
+
+\apt{I think it might be clearer just to have one row with the two different SP pointers
+  (the proper one and the bad one) marked.}
 
@@
 
-Alternatively, we can formalize well-bracketed control flow using \(\uparrow\).
+There is an alternative foramlization of well-bracketed control flow using \(\uparrow\).
 
 \definition For a machine state \(\mach\), let
-\(ret ~ (\mach,\context) = \{\mach' | \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]\}\).
-A system enjoys well-bracketed control flow if, for all states 
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
+\(ret ~ (\mach,\context) = \{\mach' \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]\}\).
+\apt{Note: you'll get better spacing by using $\mid$ rather than $|$ in set comprehensions.}
+A system enjoys  well-bracketed control flow if
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) where 
 \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
-\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-\(|\sigma| \uparrow ret(\mach')\).
+\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\)
+implies \(|\sigma| \uparrow ret(\mach')\).
 
@@
 
+\apt{Sudden appearance of $\mathbf{tailcall}$??}
 Once again we can think about our example attack. When the call occurs,
@@
 
-After the call, the call stack is one higher than it was before the call,
+After the call, the size of the call stack, \(|\sigma|\), is one greater than it was before the call,
 and if the callee makes further calls, it will continue to be incremented.
-Each return will similarly decrement it, until it is finally lower than the
+Similarly, each return will decrement it, until it is finally lower than in the
 initial state. This is the return that matches the original call, and so
@@
 \(\context = (V,\sigma)\), and
-\(\components = \{\component | V ~ \component = \sealed\}\).
+\(\components = \{\component \mid V ~ \component = \sealed\}\).
 Let \(P ~ (\mach',\context')\) be a predicate that holds on a state
-\(\Delta(\mach,\mach')\) is irrelevant in \((\mach', \context')\).
+\(\Delta(\mach,\mach')\) is irrelevant in \((\mach', \context')\).\apt{Something is broken here!}
 Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \uparrow P\) holds
@@
 \(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) such that
-\(\mach'[\component] \not = \nach'[\component]\).
+\(\mach'[\component] \not = \nach'[\component]\). ]\apt{Define first, explain after.}
 
 \definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
-\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
-
-Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to
+\(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
+Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to \apt{??}
 if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
@@
 \definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\),
-and \(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
+and \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
 Let \(\nach\) be a \(\components\)-variant of \(\mach\).
@@
 \definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
-\(\components = \{\component | V ~ \component = \unsealed\}\).
+\(\components = \{\component \mid V ~ \component = \unsealed\}\).
 Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
@@
 When the callee in such a scenario makes a call of its own, the objects passed to it
-will in turn be sealed -- unless it passes them on, which can happen with pass-by-reference
+will in turn be sealed---unless it passes them on, which can happen with pass-by-reference
 arguments!
@@
 In order to express such a property, we need our machine to carry some notion
-of {\it pointer provenance} -- a distinction between a pointer that is intended to
+of {\it pointer provenance}---a distinction between a pointer that is intended to
 point to a given object, and non-pointer integers as well as pointers to other objects.

### 2022-10-25 submission5/noexceptions.tex
New property versions in MachineModule (for now)

@@
 steps that represent these abstract operations from those that do not---for instance,
-by labeling instructions as calls, returns, etc.\apt{note proper way to do punctuating dashes}
+by labeling instructions as calls, returns, etc.
 Such a labeling scheme appears in

### 2022-11-01 submission5/noexceptions.tex
Moved more detailed explanation into earlier section

@@
 
+\begin{abstract}
+
+``Stack safety'' is a term associated with a
+variety of compiler, run-time, and hardware mechanisms for protecting stack
+memory. But unlike heap memory, the stack does not correspond to any single
+high-level language concept. The result is that
+these mechanisms typically lack precise specifications,
+relying instead on informal descriptions and examples of the bad
+behaviors that they prevent. Such a specification is especially desirable
+for an unusual enforcement mechanism, like the ``lazy'' stack safety
+micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
+
+We propose a formal characterization of ``observable'' stack safety based on
+concepts from language-based security. Stack safety is a
+combination of an integrity property (``the private
+state in each caller's stack frame is held invariant by the callee'')
+and confidentiality properties (``the caller's and callee's behavior are
+insensitive to one another's private state,''), both compatible with lazy tagging, as well as
+control-flow property.
+
+The stack governs a large number of language features that are often excluded
+from discussion of stack safety. Our stack safety property targets a system with
+both caller- and callee-saved registers, coroutines, arguments passed on the stack,
+exceptions, [tail call elimination, heaps, and concurrency.]
+
+We use these properties to validate Roessler and DeHon's micro-policies
+via property-based random testing. Our testing successfully detects violations
+in several broken variants, including Roessler and DeHon's original lazy policy.
+
+\end{abstract}
+
+\newcommand{\paragraphx}[1]{\emph{#1.}}
+
+\section{Introduction}
+
+The call stack is a perennial target for low-level attacks, leading to a
+range of dire consequences, from leakage or corruption of private stack data
+to control-flow hijacking. To foil such attacks, a profusion of
+software and hardware protections have been proposed,
+%
+including stack canaries~\cite{Cowan+98},
+bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
+split stacks~\cite{Kuznetsov+14},
+shadow stacks~\cite{Dang+15,Shanbhogue+19},
+capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
+  \ifaftersubmission\apt{Mostly from
+  nick; there could be more}\bcp{Yes, going back to MIT days---we should
+  include several more of these, if only to give readers the impression that
+this is a well-studied mechanism (so formalizing its protections is
+useful).}
+\fi
+
+The protections offered by such mechanisms are commonly described in terms
+of concrete examples of attacks that they can prevent---corruption of return
+addresses, buffer overflows, use of uninitialized variables, etc.---leaving
+a more abstract characterization to the reader's intuition.  But these
+mechanisms can be quite intricate, and the behaviors they aim to prevent are
+subtle and varied. It can be hard to guess, based only on informal
+intuitions, whether a given mechanism actually blocks all potential
+attacks---or, conversely, whether it is overly conservative and disallows
+system behaviors that are actually safe.
+To settle such questions with confidence,
+we need a precise, generic, and formal specification for stack
+safety, as a basis both for comparing the security claims of different
+enforcement techniques and for validating that these claims
+are met by particular implementations. With the notable exception of the model used
+in Skorstengaard et al.'s StkTokens work~\cite{SkorstengaardSTKJFP}, which we
+discuss in Section \ref{sec:relwork}, there is a dearth of such properties in
+the literature.
+
+We propose a novel characterization of stack safety using the tools of language-based
+security~\cite{sabelfeld2003language}, decomposing stack safety into
+the {\em integrity} and {\em confidentiality} of the caller’s local state
+during the callee's execution, and the confidentiality of the calle's local state
+afterward, as well as the control flow protection of {\em well-bracketed control
+  flow (WBCF)}~\cite{SkorstengaardSTKJFP}. We extend WBCF to a setting with exceptions,
+in which there may be multiple points at which a callee can ``return.''
+
+A key technical novelty in these definitions,
+compared to standard formulations of confidentiality and integrity from the
+security literature, is that they
+are ``nested'': {\em each} caller is guaranteed protection from its
+immediate callees (which, in turn, need protection from their immediate
+callees, etc.).
+Caller confidentiality is especially interesting. It is based on a traditional
+notion of noninterference, but whereas ordinary noninterference
+is an end-to-end hyper-property on whole program runs, caller
+confidentiality is a nested form of noninterference that applies to
+part of a run, requiring that the callee’s behavior is
+invariant under hypothetical scrambling of the caller’s stack
+but allowing the caller to access its own data upon return.
+
+Our threat model is very strong, allowing the attacker
+to execute arbitrary code, including attempting to smash
+the stack to disrupt the program’s control flow. Our stack safety
+properties demand that, even in the presence of such attackers, confidentiality
+and integrity still apply.
+%
+
+For confidentiality and integrity, our properties are {\em observational},
+meaning that they allow a function activation to read from and write to other activations'
+data, as long as such reads and writes do not affect the observable behavior of the
+system. This permissiveness enables our properties to handle lazy enforcement mechanisms.
+
+To demonstrate the utility of our formal characterization, we use these
+properties to validate and improve an existing enforcement mechanism, the
+{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
+in the Coq proof assistant on top of a RISC-V specification.  We
+use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for Coq, to generate random programs and check
+that Roessler and DeHon's micro-policies correctly abort the ones that
+attempt to violate one of our properties. Furthermore, we
+%
+check that the testing framework is able to generate counterexamples
+that violate our properties but are \emph{not} halted by incorrect
+enforcement variants---both variants that we accidentally created
+during our re-implementation of the micro-policy and ones that we
+intentionally crafted to be broken in order to increase our confidence
+in testing and the enforcement mechanism itself.
+
+We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
+which memory within each stack frame is tagged with the depth of
+the function activation that owns the frame and access is
+permitted only when an activation at that depth is currently executing, validates our
+stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
+violates the temporal aspect of confidentiality in
+cases where data can leak across repeated calls to the same callee,
+and also violates integrity if the leak uses the caller's frame. We
+propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
+confidentiality, albeit at some performance cost.
+%
+
+%For ease of exposition, we initially assume a single simple
+%stack with no sharing between callers and callees: all parameters and return
+%values are passed in registers.  Later, we will introduce enhanced versions
+%supporting
+%(1) passing of scalar stack data,
+%(2) callee-saves registers calling conventions, and
+%(3) a coroutine system with a static layout.
+
+In sum, we offer the following contributions:
+
+\begin{itemize}
+\item We formalize permissive stack confidentiality and integrity and integrity properties.
+  Both are parameterized over a notion of external
+  observation, and are violated only if accessing secrets or overwriting
+  data causes a visible change in the system behavior %(\cref{sec:lazy}).
+\item We apply these definitions to a complex system with features such as argument passing
+  on the stack, callee-saves registers, exceptions, and a simple coroutine system% (\cref{sec:ext}).
+\item We use property-based random testing to validate the relationship between
+  our properties and micro-policies %(\cref{sec:testing})
+\end{itemize}
+%\Cref{sec:relwork,sec:future} discuss related and future work.
+
+Our artifact contains Coq formalizations of our properties for illustrative purposes
+as well as the testing implementation.
+
 \section{The Setting}
@@
 
-Otherwise, we do not assume that the code is reasonable\apt{still vague} or adheres to any particular
-calling convention or source construct. In particular, while we are agnostic as to the source
+Otherwise, we do not assume that the code adheres to any particular
+calling convention or implements a source language construct.
+In particular, while we are agnostic as to the source
 language, we certainly aim to support C, and so any source function might contain undefined
@@
 This is a strong threat model, but hardware and timing attacks are out of scope,
-and our properties are termination insensitive as a result of the enforcement mechanism
-(below).
+and our properties are termination insensitive as a result of the enforcement mechanism.
 
@@
 {\it termination insensitive} systems \cite{}. TODO: more detailed explanation later.
-
-For convenience, we will abbreviate multiple updates of a mapping
-\(\mach[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
-as \(\mach \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
-are fomulae with free variables bound to sets in \(C\).
 
@@
 
-A pending activation also carries the addresses to which the program counter
-and stack pointer should point in the event of a return. We represent the call
-stack with a list of such triples. A {\it context} is a pair of the current activation's view
-and a list of pending ones.
-The initial context is \(\context_0 = (V_0, [ ~ ])\).
-\[(V, \addr_{ret}, \addr_{sp}) \in \mathit{PEND} ::= \mathit{VIEW \times \WORDS \times \WORDS}\]
-\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ PEND}\]
+For convenience, we will abbreviate multiple updates of a mapping
+\(V[\component_0 \mapsto sc_0][\component_1 \mapsto sc_1]\dots\)
+as \(V \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
+are fomulae with free variables bound to sets in \(C\).
+
+We represent the call stack with a list of views representing pending inactive
+principals, ranged over by \(\sigma\). A {\it context} is a pair of the current
+activation's view and that list. The initial context is \(\context_0 = (V_0, \emplist)\).
+\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
 
@@
 From a machine state and a context, we create a {\it combined state}
-\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
+\(s = (\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
 labeled with the same operations. Note that \(Op\) takes as its state argument
@@
 
+\subsection{Policy State}
+
+In addition to its elements, a system may contain additional ``policy'' state that carries
+information relevant to a hardware enforcement mechanism. We assume that policy state
+does not influence execution, except to cause the system to failstop, which is modeled
+as a state stepping to itself. A failstop is therefore indistinguishable from silent
+divergence.
+
+Formally, we can treat our machine state as
+\(\mach = (\mach[\cdot],\langle) \mach \rangle \in \MACHS
+::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
+Then \(\langle \mach \rangle\) is the security state. We assume that if
+\(\mach[\component] = \nach[\component]\) for all \(\component\),
+\(\mach \stepstounder{\bar{\psi_1}} \mach'\), and \(\nach \stepstounder{\bar{\psi_2}} \nach'\),
+then \(\bar{\psi_1} = \bar{\psi_2}\) and either
+\(\mach = \mach'\), \(\nach = \nach'\), or \(\mach'[\component] = \nach'[\component]\)
+for all \(\component\).
+
 %\section{Calls, Returns, and Private Allocations}
@@
 Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
-that takes a secret in its arguments and initializes a variable to contain
-potentially sensitive data.\apt{Arguments to {\tt main} are overly complicated.
-  There's no need to match reality (and in fact {\tt argv} is really a {\tt char*}
-  anyway).}\apt{As we discussed, it would be simpler to represent ``publish'' by
-  writing to a specified global rather than with a call.}
-It then calls another function {\tt f},
-and afterward may decide to {\tt publish} its secret based on the contents of
+that takes a secret as its argument and initializes a variable to contain
+potentially sensitive data. We will assume for this system that events are stores
+to a special address, called {\tt out}.
+
+{\tt main} then calls another function {\tt f},
+and afterward may decide to output its secret based on the contents of
 its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
 it should not---instead it will publish the results of {\tt f}.
-\apt{Would be better to use symbolic names instead of inetegers for {\tt jal}
-  targets. Figure at bottom is good, but also needs additional labelling to
-  show addresses relative to the new sp within {\tt f}.}
 
@@
     {\tt
-      void main(int argc, int* argv) \{
-
-      ~ ~ int secret = argv[1];
+      void main(int secret) \{
 
@@
 
-      ~ ~ ~ ~ publish(secret);
+      ~ ~ ~ ~ *out = secret;
 
@@
 
-      ~ ~ ~ ~ publish(res);
+      ~ ~ ~ ~ *out = res;
 
@@
       \labeledrow{12:}{sw zero,4(sp)}{}
-      \labeledrow{16:}{jal 100,ra}{\(\mathbf{call} ~ 100 ~ \emplist ~ \emplist\)}
+      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \mathtt{f} ~ \emplist ~ \emplist\)}
       \labeledrow{20:}{sw r0,0(sp)}{}
@@
       \labeledrow{36:}{lw a0,8(sp)}{}
-      \labeledrow{40:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
+      \labeledrow{40:}{sw a0,out}{}
       \labeledrow{44:}{j L2:}{}
       \labeledrow{L1, 48:}{lw r0,0(sp)}{}
-      \labeledrow{52:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
+      \labeledrow{52:}{sw a0,out}{}
       \labeledrow{L2, 56:}{ld ra,12(sp)}{}
@@
 assembly code in Figure \ref{fig:f}.
-\apt{In (a,b), {\tt 12(sp)} maps to {\tt sensitive}, not {\tt secret} (I think!).
-  It would be nice if (a) didn't have to push a frame (i.e. again by making publication
-  just a global write) and if all the examples of {\tt f} were the same length, e.g.
-  by adding {\tt nop}s.}
 
 In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
-pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
+pointer, accesses {\tt secret}, and directly outputs it. But more
 subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
-instead return that value so that {\tt main} makes the call to {\tt publish},
+can instead return that value so that {\tt main} stores it to {\tt out},
 as in Figure \ref{subfig:indirect}.
@@
     \begin{tabular}{r l | l}
-      \labeledrow{100:}{addi sp,sp,-8}{\(\mathbf{alloc} ~ (-8,8)\)}
-      \labeledrow{104:}{sd ra,0(sp)}{}
-      \labeledrow{108:}{lw a0,12(sp)}{}
-      \labeledrow{112:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0} ~ \emplist\}\)}
-      \labeledrow{116:}{lw ra,0(sp)}{}
-      \labeledrow{120:}{addi sp,sp,8}{\(\mathbf{dealloc} ~ (0,8)\)}
-      \labeledrow{124:}{jalr ra}{\(\mathbf{return}\)}
+      \labeledrow{100:}{lw a4,8(sp)}{}
+      \labeledrow{104:}{sw a4,out}{}
+      \labeledrow{108:}{li a0,1}{}
+      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
     \end{tabular}
@@
     \begin{tabular}{r l | l}
-      \labeledrow{100:}{lw a0,12(sp)}{}
-      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
+      \labeledrow{100:}{lw a4,8(sp)}{}
+      \labeledrow{104:}{mov a4,a0}{}
+      \labeledrow{108:}{sw zero,-4(sp)}{}
+      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
     \end{tabular}
@@
       \labeledrow{104:}{sw a5,4(sp)}{}
-      \labeledrow{108:}{jalr ra}{\(\mathbf{return}\)}
+      \labeledrow{108:}{li a0,1}{}
+      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
     \end{tabular}
@@
       \labeledrow{100:}{addi ra,ra,20}{}
-      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
+      \labeledrow{104:}{nop}{}
+      \labeledrow{108:}{nop}{}
+      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
     \end{tabular}
@@
 
-\apt{This para is a bit of an orphan.}
-At the start of execution the program counter is 0 and the stack pointer -4 \apt{or is it 0?},
-with the stack growing down. The functions {\tt f} and {\tt publish}
-are at addresses 100 and 200, respectively.
-
-\apt{Need a smoother transition here. Be explicit that we are now going to start considering the security
-  state.}
-
-Figure \ref{fig:exec1} shows how the program counter, stack pointer, and context
-develop over several steps, with labels on the steps that allocate space and then make
-a call. \(\xrightarrow{\overline{\psi}}\) transitions are written \(\downarrow \overline{\psi}\) in this diagram;
-events are omitted.\apt{graphics on right of figure need to be keyed to corresponding steps.}
+Figure \ref{fig:exec1} shows how the security context
+develops over several steps, with labels on the steps that allocate space and then make
+a call. Both program counter and stack pointer are initially 0, and the stack pointer
+is shown for convenience.
+\(\xrightarrow{\overline{\psi}}\) transitions are written \(\downarrow \overline{\psi}\)
+in this diagram; events are omitted.
 
@@
     \vskip 0pt
-    \begin{tabular}{| l | l | l || l |}
-      \hline
-      Step & \(\PCname\) & \(\SP\) & \(\context\) \\
-      \hline
-      0 & 0 & 0 & \(V_0, \emplist\) \\
-      \hline
-      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
-      \hline
-      1 & 4 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket\) \\
-      \hline
-      \multicolumn{4}{l}{\(\vdots\)} \\
-      \hline
-      4 & 16 & -20 & \(V_1, \emplist\) \\
-      \hline
-      \multicolumn{4}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
-      \hline
-      5 & 100 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[(V_1, 20, -20)]\) \\
-      \hline
-      \multicolumn{4}{l}{\(\vdots\)} \\
-      \hline
-      \(n\) & ? & ? & \(V_2,[(V_1, 20, -20)]\) \\
-      \hline
-      \multicolumn{4}{l}{\(\downarrow [\mathbf{return}]\)} \\
-      \hline
-      \(n+1\) & 20 & -20 & \(V_1, \emplist\) \\
+    \begin{tabular}{| l | l || l |}
+      \hline
+      Step & \(\SP\) & \(\context\) \\
+      \hline
+      0 & 0 & \(V_0, \emplist\) \\
+      \hline
+      \multicolumn{3}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
+      \hline
+      1 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket, \emplist\) \\
+      \hline
+      \multicolumn{3}{l}{\(\vdots\)} \\
+      \hline
+      4 & -20 & \(V_1, \emplist\) \\
+      \hline
+      \multicolumn{3}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
+      \hline
+      5 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[V_1]\) \\
+      \hline
+      \multicolumn{3}{l}{\(\vdots\)} \\
+      \hline
+      8 & -20 & \(V_2,[V_1]\) \\
+      \hline
+      \multicolumn{3}{l}{\(\downarrow [\mathbf{return}]\)} \\
+      \hline
+      9 & -20 & \(V_1, \emplist\) \\
       \hline
@@
 
-On step \(n\), {\tt f} returns, and we restore the topmost inactive principal,
-{\tt main}'s principal. \apt{This $n$ is a bit mysterious. Again, if all the snippets
-  were the same length, you could use a concrete number here.}
-
-\apt{Need a bridging sentence here, along the lines of: ``Now we consider how to define desirable
-  stack safety properties that rule out various kinds of bad behavior.''}
+On step 8, {\tt f} returns, and we restore the topmost inactive principal,
+{\tt main}'s principal. Now we will discuss the intuition behind how this
+structure translates into stack safety properties that rule out different
+classes of bad behavior.
 
@@
 
-But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or \(\SP \neq -24\))\apt{-20?},
-as would be the case in attack (d)? Then it has violated the WBCF property.
-This property can be phrased in terms of a single step: for every transition
-labeled as a return, the program counter and stack pointer at the next state
-match those recorded in the top pending activation. The table in Figure \ref{fig:exec1}
-assumes that WBCF is followed.
+But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
+\(\SP \neq -20\)), as would be the case in attack (d)? We consider this to
+violate well-bracketed control flow (WBCF). WBCF is a relationship between
+call steps and their corresponding return steps: after the return, the program
+counter should be at the instruction following the state before the call,
+and the stack pointer should be the same as it was before the call.
+
+Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
+20 to the return address and then returns, thus bypassing the check and outputting
+{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 20,
+so we define a predicate on states \(ret\) that holds when the
+program counter and stack pointer are both 20.
+
+Then we identify the state immediately following the return, if one occurs ---
+this will be the first state in which the pending call stack is smaller than it was
+inside of the call.
+WBCF requires that if \(s\) after that return, \(ret ~ s\) will hold.
+For nested calls, where the pending stack is initial larger, the same principal
+applies: \(ret ~ s\) must hold the next time the pending stack is the same size or smaller.
+
+To demonstrate why WBCF also cares about the stack pointer, consider an attack which
+returns with the stack pointer at -12 instead of -20. Now, compare the layout of the
+stack relative to the stack pointer as it should be (\(\SP\)) to the one that results by
+increasing the stack pointer before return (\(\SP'\)):
+
+\begin{tabular}{| l | l | l | l | l |}
+  \multicolumn{1}{r}{\(\SP \downarrow\)} &
+  \multicolumn{2}{r}{\(\SP' \downarrow\)} \\
+  \hline
+  res & sens & sec & ra & ra \\
+  \hline
+\end{tabular}
+
+\vspace{\abovedisplayskip}
+
+In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
+read part of the return address, and then it will attempt to output
+{\tt res}, but instead output {\tt secret}! Even absent other
+kinds of data protection, the stack pointer {\it must} be restored
+for the program to behave predictably.
 
@@
 
-Integrity is more complicated, and requires us to compare the state at the start of
-the call to {\tt f} to that on return. Intuitively, the integrity of {\tt main}
+Stack integrity applies a similar notion to all of the memory that a function has
+allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
+upon allocation, the view labeled is \(\object\), and when a call is made, it instead
+becomes \(\sealed\). In more sophisticated systems discussed later, we will distinguish
+sealed objects from those that are intentionally passed to the callee.
+Intuitively, the integrity of {\tt main}
 is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
@@
 addresses -20 through -1 (allocated as private data) and callee-saved registers such as
-the stack pointer and return address.
-Note that the callee-saved registers may well have changed during the call---but if
-so, the callee is obligated to restore them to their prior value.
+the stack pointer and return address. Note that callee-saved registers often change
+during the call---but if so, the callee is obligated to restore them to their prior value.
 
 One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
-in a way that is harmless---or that is rendered harmless by the enforcement mechanism!
-\apt{This is still confusing, because the ``rendered harmless'' part applies just as much
-  to the strict integrity we've been discussing up to now.}
-This is the case in a lazy micro-policy\apt{presumes we've described these a bit in the introduction
-  (which we should)}: the monitor permits the caller to write into the
+in a way that is harmless. Or, an enforcement mechanism might defer the checking
+for illegal writes until later in the execution, entering an error state at some
+point after return but before the illegal write is able to damage the system.
+This is the case in a lazy micro-policy: the monitor permits the caller to write into the
 callee's frame, but taints that write so that the callee cannot read it later without
@@
 if it could be replaced with any other value without changing the observable behavior
-of the machine. More precisely, for a given state \(\mach\) and set of elements \(\components\),
-we define a {\em \(\components\)-variant} state to be one agrees with \(\mach\) on every
-element except those in \(\components\). The elements of \(\components\) are said to be
+of the machine. More precisely, for a set of elements \(\components\),
+a pair of states \(\mach\) and \(\nach\)
+are {\em \(\components\)-variants} if they agree on the value of every
+element except those in \(\components\), and share the same policy state.
+The elements of \(\components\) are said to be
 irrelevant in \(\mach\) if every variant state produces an equivalent trace
@@
 
-So, to define integrity, we need to know when a caller has been returned to;
-this is the first state after the call in which the depth of the call stack
-matches the state before the call. Integrity holds if, at that return state,
+So, to define integrity, we again need to know when a caller has been returned to,
+and we use the same mechanism of checking the depth of the call stack.
+Integrity holds if, at that returned state,
 any element that is \(\sealed\) under the callee's view is either restored
@@
 
-Returning to our example: at step 4, when {\tt main} makes its call to {\tt f},
-the call stack in the context is empty. It makes its call, and at step 5, {\tt main}'s
-frame, including {\tt secret} and {\tt sensitive}, is mapped to \(\sealed\) from the
-perspective of {\tt f}. So, the next time the call stack is empty (implicitly, when
-we have returned to {\tt main}), elements of that frame must either match their
-values at step 4, or be irrelevant to future computation.\apt{picture?}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{variants.png}
+  \caption{Integrity Violation}
+  \label{fig:variant}
+\end{figure}
+
+The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
+which is \(\sealed\). Figure \ref{fig:variant} shows the state at step 5, just after the call,
+assuming that {\tt sec} is 5. Similar to WBCF, we will define
+\(int ~ s\) as a predicate on states that holds if, of the sealed
+addresses in state 5, all values that differ in \(s\) are irrelevant. We once again
+require that \(int\) hold on the state following the matching return.
+
+That state is reached at step 9. Since {\tt sensitive} has changed, we must determine whether
+it is irrelevant. We consider a variant state in which it has any other value, arbitrarily
+choosing 43. As execution continues after the return from the original state, it
+passes the check on {\tt sensitive}, while the execution from the variant does not, resulting
+in differing outputs. Therefore {\tt sensitive} is not irrelevant, so \(int\) does not hold,
+and integrity has been violated.
 
@@
 means that its callee's behavior is dependent only on publicly visible data,
-not the caller's private state. \apt{Spell out the implications of this for initialization.}
+not the caller's private state. This also requires that the callee initialize
+memory before reading it.
 As we saw in the examples, we must consider both the observable events
@@
 
-Consider the state \((\mach, (V_2,\sigma))\) at step 5. If we take a variant state over
-the set of element(s) that are \(\sealed\) or \(\unsealed\) in \(V_2\),
-{\it internal confidentiality} means
-that, if we execute both states until they return, they will produce equivalent sequences
-of events.
+Consider again the state \((\mach, (V_2,\sigma))\) at step 5, this time with the attacker from
+Figure \ref{subfig:direct}. We take a variant state over the set of element(s) that are
+\(\sealed\) or \(\unsealed\) in \(V_2\), which happens to be the entire stack (see
+Figure \ref{fig:variant2}.) If we take a trace of execution from each state until it returns,
+the traces may differ, in this case outputting 5 (the original value of {\tt sec}) and
+4 (its value in the variant) respectively. This is a violation of
+{\it internal confidentiality}.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{variants2.png}
+  \caption{Internal Confidentiality Violation}
+  \label{fig:variant2}
+\end{figure}
 
@@
 by reading it and then returning it, in a context where the caller would output the returned
-value. In situations like this, we care about {\it return-time confidentiality}.
-In that case, we would identify that in our original execution, {\tt a0} changed between
-the call and return, being set to the value of {\tt secret}. But in a variant state,
-{\tt secret} might have a different value, and so will {\tt a0} at the return.
-This difference, found by comparing the initial and final states of both executions,
-indicates that {\tt a0} is corrupted. Return-time confidentiality means that
-an element that is corrupted in this way must be irrelevant.
-
-Together, these definitions encompass the confidentiality of the caller from the callee.\apt{picture?}
+value. Figure \ref{subfig:variant3} shows the behavior of the same variants under this attacker,
+but in this case, there is no output during the call. Instead the value of {\tt secret} is
+extracted and placed in {\tt a0}, the return value register. To illustrate the subtlety
+of identifying a leak, {\tt f} has also stored a 0 below the stack pointer.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{variants3.png}
+  \caption{Return-time Confidentiality Violation}
+  \label{fig:variant3}
+\end{figure}
+
+Consider three elements: the address \(\SP - 4\) (blue), the address \(\SP + 12\) (yellow),
+and the register {\tt a0} (red). In the first case, the execution from the variant state
+has changed the value at \(\SP - 4\). But both return states agree on that value. Therefore,
+while the address was updated, it does not depend on a secret value.
+
+In the second case, the return states disagree on the value of \(\SP + 12\). But in neither
+case has that value changed since the original variants. So the difference is inherited from
+the original variation, and does not depend on a secret value either. This matters because,
+following the return, the caller is allowed to access its stack frame, so those addresses
+may influence future execution legitimately.
+
+But in the case of {\tt a0}, the value has changed during the call (on both sides,
+although only one would be necessary.) And its final value differs between the variants.
+Therefore, it must depend on a secret. Indeed, we happen to know that it depends on {\tt secret}
+itself. We must therefore insist that {\tt a0} be irrelevant in the original trace.
+We term this property {\it return-time confidentiality}.
+
+Structurally, return-time confidentiality resembles integrity, but now dealing with
+variants. We begin with a state immediately following
+a call. Let that be \((\mach,\context)\). We suppose an arbitrary variant state,
+\((\nach,\context)\), which may vary any element that is \(\sealed\) or \(\unsealed\).
+That is, any element that is not used legitimately to pass arguments. Caller confidentiality
+therefore can be thought of as the callee's insensitivity to parts of its initial states
+that are not part of the caller-callee interface.
+
+In place of a predicate we define \(\mathit{conf}\) as a binary relation on states.
+\(\mathit{conf}\) will hold on eventual return states \((\mach',\context')\) and
+\((\nach',\context'')\)
+if all elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\),
+and which differ between \(\mach\) and \(\nach\), are irrelevant. We then define
+return-time confidentiality as holding if \(conf\) holds on the matching return states
+from \(\mach\) and \(\nach\), respectively.
+
+Together, these definitions encompass the confidentiality of the caller from the callee.
 
@@
 safety as the combination of WBCF and {\it local state encapsulation} (LSE), and prove that their
-capability-based enforcement satisfies both. LSE corresponds to integrity, plus some degree of
-confidentiality.\apt{Be more explicit.}  They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
+capability-based enforcement satisfies both. LSE corresponds to integrity, plus the
+confidentiality of the caller's frame, but neither the unallocated stack nor data left
+behind in it by previous callees.
+They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
 together as a safe-by-construction abstract machine. That machine does not allow a callee to
@@
 written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\),
+and \(\langle\mach\rangle = \langle\nach\rangle\).
 
@@
              {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
-               (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}
+               (V',V::\sigma)\)}
 
@@
 
-\definition A system enjoys {\it well-bracketed control flow} if 
-\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context,\obs)\)
-implies \(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).
-
-\vspace{\abovedisplayskip}
-
-\apt{These explanations that relate the definitions back to the example snippets are
-  great. They would be even better if you can find a way to put them in Section II,
-  based on (less formal) definitions of the terms introduced here.}
-
-Consider again the attack in Figure \ref{subfig:WBCF}: the attacker adds
-20 to the return address and then returns, thus bypassing the check and publishing
-{\tt secret}. During the call step, we record in the context that the return to the
-caller should have the program counter at the next instruction (address 20) and the
-stack pointer restored to address -20. When the return occurs, we see that in the
-state following it, the program counter is 40---a violation of WBCF.
-
-To demonstrate why WBCF also cares about the stack pointer, consider an attack which
-returns with the stack pointer at -12 instead of -20. Now, compare the layout of the stack relative
-to the stack pointer as it should be (top) to the one that results by increasing the
-stack pointer before return (bottom):
-
-\begin{tabular}{| l | l | l | l | l | l | l |}
-  \multicolumn{3}{r}{\(\SP \downarrow\)} \\
-  \hline
-  & & res & sens & secr & ra & ra \\
-  \hline
-  \hline
-  res & sens & secr & ra & ra & & \\
-  \hline
-\end{tabular}
-
-\apt{I think it might be clearer just to have one row with the two different SP pointers
-  (the proper one and the bad one) marked.}
-
-\vspace{\abovedisplayskip}
-
-In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
-read part of the return address, and then it will attempt to publish
-{\tt res}, but instead will publish {\tt secret}!
-
-There is an alternative foramlization of well-bracketed control flow using \(\uparrow\).
-
-\definition For a machine state \(\mach\), let
-\(ret ~ (\mach,\context) = \{\mach' \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]\}\).
-\apt{Note: you'll get better spacing by using $\mid$ rather than $|$ in set comprehensions.}
-A system enjoys  well-bracketed control flow if
+\definition We define a correct return from \(\mach\) as:
+%
+\[ret(\mach) = \{(\mach',\context') \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}\]
+
+\definition
+A system enjoys  well-bracketed control flow if, whenever
 \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) where 
-\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
-\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\)
-implies \(|\sigma| \uparrow ret(\mach')\).
-
-\vspace{\abovedisplayskip}
-
-\apt{Sudden appearance of $\mathbf{tailcall}$??}
-Once again we can think about our example attack. When the call occurs,
-we define a predicate \(ret ~ (\mach,\context)\) on states that holds when the program counter
-and stack pointer have the appropriate values. Then we use apply the \(\uparrow\)
-operator to this predicate with the depth of the call stack {\it after the call} and
-the machine state {\it before the call}.
-
-After the call, the size of the call stack, \(|\sigma|\), is one greater than it was before the call,
-and if the callee makes further calls, it will continue to be incremented.
-Similarly, each return will decrement it, until it is finally lower than in the
-initial state. This is the return that matches the original call, and so
-the resulting state is when we require \(ret(\mach)\) to hold.
+\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
+%or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\)
+\(|\sigma| \uparrow ret(\mach)\) holds on \((\mach',(V,\sigma))\).
 
@@
 
-\definition Let \((\mach,\context)\) be a combined state with
-\(\context = (V,\sigma)\), and
-\(\components = \{\component \mid V ~ \component = \sealed\}\).
-Let \(P ~ (\mach',\context')\) be a predicate that holds on a state
-\(\Delta(\mach,\mach')\) is irrelevant in \((\mach', \context')\).\apt{Something is broken here!}
-Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \uparrow P\) holds
-on \((\mach,\context)\).
-
-\definition A system enjoys {\it stack integrity} if, for all reachable states
-\((\mach,\context)\) such that
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\) with
-\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
-\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-\((\mach',\context')\) enjoys return-time integrity.
-
-\vspace{\abovedisplayskip}
-
-The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
-which is \(\sealed\). At the call, we define the predicate \(P\), which holds on states
-where any \(\sealed\) elements that changed since the call are all irrelevant.
-As {\tt sensitive} is such an element (conveniently, the only one), the property
-has not yet been violated at the returned state \((\mach,\context)\). To determine whether this
-attack has violated stack integrity, we must consider an arbitrary variant state
-\(\nach = \mach[\addr_{\mathtt{sensitive}} \mapsto \word]\) for some \(\word\).
-The property is violated if the traces of \((\mach,\context)\) and \((\nach,\context)\)
-differ.
-
-In our case, {\tt publish} produces an observation, and if \(\word \not = 42\),
-that observation in the trace from \((\nach,\context)\) will be {\tt res}, which is 0.
-On the other hand, \((\mach,\context)\) will produce an event from {\tt secret},
-which may differ. So, integrity is indeed violated.
+\definition We define \(int(\mach,V)\), to hold on a state \((\mach',\context')\) when
+\(\{\component \mid V ~ \component = \sealed\} \cap \Delta(\mach,\mach')\)
+is irrelevant in \((\mach', \context')\).
+
+\definition A system enjoys {\it stack integrity} if, whenever
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
+\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
+% or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+we have:
+\[(|\sigma'| \uparrow int(\mach',V)) ~ (\mach',(V,\sigma))\]
 
 \subsection{Caller Confidentiality}
-
-\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
-of four states treats \(\mach\) and \(\nach\) as the initial states and
-\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
-set of elements that were corrupted between them. That is, the set of all elements
-\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) such that
-\(\mach'[\component] \not = \nach'[\component]\). ]\apt{Define first, explain after.}
 
@@
 \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
-Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to \apt{??}
+Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to \(\components\)
 if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
@@
 
-\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\),
-and \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
-Let \(\nach\) be a \(\components\)-variant of \(\mach\).
-Let \(R\) be a relation such that \((\mach_1,\context_1) ~ R ~ (\mach_2,\context_2)\)
-if \(\bar{\Diamond}(\mach,\mach_1,\nach,\mach_2)\) is irrelevant in \((\mach_1,\context_1)\).
-Then \((\mach,\context)\) enjoys {\it return-time confidentiality}
-if, for any such \(\nach\), \((\mach,\context) ~ (|\sigma| \Uparrow R) ~ (\nach,\context)\).
-
-\definition A system enjoys {\it caller confidentiality} if, for all reachable states
-\((\mach,\context)\) such that
+\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
+is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
+This treats
+\(\mach\) and \(\nach\) as the initial states and
+\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
+set of elements that were corrupted between them.
+
+\definition We define the relation \(conf(\mach,\nach)\) to hold on a pair of states
+\((\mach',\context)\) and \((\nach',\context')\) when
+\(\bar{\Diamond}(\mach,\nach,\mach',\nach')\) is irrelevant in \((\mach',\context)\).
+
+\definition A state \((\mach,(V,\sigma))\) enjoys {\it return-time confidentiality}
+if, for \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\)
+and any \(\nach\) that is a \(\components\)-variant of \(\mach\), we have:
+\[(\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \mathit{conf}(\mach,\nach) ~ (\nach,(V,\sigma))\]
+
+\definition A system enjoys {\it caller confidentiality} if, whenever
 \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
-\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \psi\),
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+%or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \psi\),
 \((\mach',\context')\) enjoys both internal and return-time confidentiality.
 
-\vspace{\abovedisplayskip}
-
-Confidentiality is more complicated. We begin with a state immediately following
-a call. Let that be \((\mach,\context)\). We suppose an arbitrary variant state,
-\((\nach,\context)\), which may vary any element that is \(\sealed\) or \(\unsealed\).
-That is, any element that is not used legitimately to pass arguments. Caller confidentiality
-therefore can be thought of as the callee's insensitivity to parts of its initial states
-that are not part of the caller-callee interface.
-
-In the example in Figure \ref{subfig:direct}, this variation will quickly result
-in differing events, as {\tt publish} is called with arguments that may have different
-values. This is clearly a violation of confidentiality, in the form of internal
-confidentiality.
-
-When the variants return, however, the situation becomes more complicated.
-Consider Figure \ref{subfig:indirect}. The value of {\tt secret} is been extracted
-and placed in {\tt a0}, the return value register. Suppose that we have \(\mach\)
-and \(\nach\) as above, and they step, respectively, to \(\mach'\) and \(\nach'\),
-which each follow a return. We must identify which elements of the state
-contain leaked confidential data.
-
-An element contains confidential data, according to the definition of non-interference, 
-if it differs between the states. But some of those data belong to the original
-elements which are accessible to the caller and may now influence future execution
-legitimately. We are looking specifically for those elements that have {\it changed}
-from call to return, and that vary between the variants. As with integrity, we then
-must determine whether those leaks actually translate into events. If the set of
-such elements is not irrelevant, confidentiality is violated.
-
 \subsection{Callee Confidentiality}
 
-\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
-\(\components = \{\component \mid V ~ \component = \unsealed\}\).
-Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
-\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
-Then \(\mach\) enjoys {\it post-return confidentiality} if \(|\sigma| \downarrow P\) holds.
-
-\definition A system enjoys {\it callee confidentiality} if, for all reachable states
-\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
-\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-\((\mach',\context')\) enjoys post-return confidentiality
-with respect to \(\context\).
+\definition We define \(\mathit{cconf}(\mach,V)\), to hold on a state \((\mach',\context')\) when
+\(\{\component \mid V ~ \component = \unsealed\} \cap \Delta(\mach,\mach')\)
+is irrelevant in \((\mach', \context')\).
+
+\definition A system enjoys {\it callee confidentiality} if, whenever
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\)
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+% or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+we have:
+\[(|\sigma| \uparrow \mathit{cconf}(\mach',V)) ~ (\mach',\context')\)
 

### 2022-11-06 submission5/noexceptions.tex
fix minor latex issue
edits and comments

@@
 The set of special operations that are recognized in this way are termed
-{\it security operations}. Our properties are phrased as a {\it security semantics},
+{\it security-relevant operations}. Our properties are phrased as a {\it security semantics},
 which extends the underlying machine with additional context about its
 security principals and which registers and memory they require to be secure.
-This context evolves dynamically through the execution of security operations.
+This context evolves dynamically through the execution of security-relevant operations.
 This notion of security may be phrased as
@@
 Unlike approaches that give a safe-by-construction semantics
-for various security operations and then prove secure compilation to the underlying
+for various security-relevant operations and then prove secure compilation to the underlying
 enforcement mechanism, our machine is explicitly not safe unless the enforcement
@@
 We will first give a detailed example of a security semantics for a simple setting
-in which our security operations are restricted to making calls and returns, and allocating
+in which our security-relevant operations are restricted to making calls and returns, and allocating
 private memory within the current function activation. We then extend this model
@@
 
-We must trust that our method of distinguishing security operations is accurate; if it
+We must trust that our method of distinguishing security-relevant operations is accurate; if it
 involves labels placed on code by a compiler, that means trusting that the compiler placed
@@
     & {\tt a2} -- {\tt a7} & Caller-saved args \\
-    \hline
-    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s7} & Callee-saved \\
-    \(\sealed\) & {\tt ra} & Return Address \\
-    & {\tt sp} & Stack Pointer \\  
-    \hline
-    \(\mathit{PUBLIC}\) / & {\tt pc} & Program Counter \\
-    \(\public\) & & \\
+    & {\tt ra} & Return Address \\
+    \hline
+    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
+    \(\sealed\) & {\tt sp} & Stack Pointer \\  
+    \hline
+    \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
+    \(\public\) & {\tt tp} & Thread Pointer \\
     \hline
   \end{tabular}
-  \caption{RISC-V register set}
+  \caption{RISC-V integer register set}
   \label{fig:RISCVregs}
@@
 \MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
-is labeled by an ordered list of security operations, defined below,
+is labeled by an ordered list of security-relevant operations, defined below,
 and an event.
@@
 
-\subsection{Security Operations, Principals, and Security State}
-
-Security operations consist of calls (including tailcalls), returns,
-allocations and deallocations. Calls and tailcalls identify the address of the call's target,
-and the argument registers. Allocations and deallocations are phrased in terms of a pair
+\subsection{Security-Relevant Operations, Principals, and Security State}
+
+\apt{Maybe should drop the math displays from this section, and keep to words.}
+Security-relevant operations consist of calls (including tailcalls), returns,
+allocations and deallocations. \apt{No tailscalls for now?} Calls and tailcalls identify the address of the call's target,
+and the argument registers. \apt{Either explain $\overline{sa}$ or omit it.} Allocations and deallocations are phrased in terms of a pair
 of an offset from the stack pointer and a size. Returns don't require any additional information.
@@
 are fomulae with free variables bound to sets in \(C\).
-
-We represent the call stack with a list of views representing pending inactive
-principals, ranged over by \(\sigma\). A {\it context} is a pair of the current
-activation's view and that list. The initial context is \(\context_0 = (V_0, \emplist)\).
+\apt{can omit this para}
+
+A {\it context} is a pair of the current activation's view and 
+a list of views representing the call stack (pending inactive
+principals), ranged over by \(\sigma\).  
 \[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
-
-Every security operation manipulates the call  stack via a function
+The initial context is \(\context_0 = (V_0, \emplist)\).
+
+\apt{Maybe don't need to describe $Op$ or $\stepstounder{\overline{\psi}}$ here, 
+  and maybe not even combined states, although they are used at the moment.}
+  Every security-relevant operation manipulates the call stack via a function
 \(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
@@
 
+\apt{As we discussed at length, it would certainly be nice to avoid separate policy state if possible.}
 In addition to its elements, a system may contain additional ``policy'' state that carries
 information relevant to a hardware enforcement mechanism. We assume that policy state
-does not influence execution, except to cause the system to failstop, which is modeled
+cannot influence execution, except possibly to cause the system to failstop, which is modeled
 as a state stepping to itself. A failstop is therefore indistinguishable from silent
@@
 Formally, we can treat our machine state as
-\(\mach = (\mach[\cdot],\langle) \mach \rangle \in \MACHS
+\(\mach = (\mach[\cdot],\langle \mach \rangle) \in \MACHS
 ::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
+\apt{This overloaded notation doesn't work. $\mach$ can be either a pair or a map, not both.
+Note that earlier text strongly implies that it is a map.}
 Then \(\langle \mach \rangle\) is the security state. We assume that if
@@
 In this section we consider a simple setting with calls, returns, and private allocations.
-Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
-that takes a secret as its argument and initializes a variable to contain
-potentially sensitive data. We will assume for this system that events are stores
+Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
+for a function {\tt main}. (Code is compiled using the standard RISC-V
+calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
+{\tt main} begins at address 0 and its callee {\tt f} at address 100.)
+
+{\tt main} takes a secret as its argument and initializes a variable to contain
+potentially sensitive data. We assume that observable events are stores
 to a special address, called {\tt out}. 
-
-{\tt main} then calls another function {\tt f},
+{\tt main} calls another function {\tt f},
 and afterward may decide to output its secret based on the contents of
@@
     {\tt
+      volatile int out;
+
       void main(int secret) \{
@@
 
-      ~ ~ if (sensitive == 42) \{
-
-      ~ ~ ~ ~ *out = secret;
-
-      ~ ~ \} else \{
-
-      ~ ~ ~ ~ *out = res;
-
-      ~ ~ \}
+      ~ ~ if (sensitive == 42)
+
+      ~ ~ ~ ~ out = secret;
+
+      ~ ~ else
+
+      ~ ~ ~ ~ out = res;
 
@@
       \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \mathtt{f} ~ \emplist ~ \emplist\)}
-      \labeledrow{20:}{sw r0,0(sp)}{}
+      \labeledrow{20:}{sw a0,0(sp)}{}
       \labeledrow{24:}{lw a4,4(sp)}{}
@@
       \labeledrow{44:}{j L2:}{}
-      \labeledrow{L1, 48:}{lw r0,0(sp)}{}
+      \labeledrow{L1, 48:}{lw a0,0(sp)}{}
       \labeledrow{52:}{sw a0,out}{}
@@
 to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
-assembly code in Figure \ref{fig:f}.
-
+assembly code in Figure \ref{fig:f}.\footnote{We arrange that all the snippets
+have the same length just to simplify some later explanations.}
+%
 In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
@@
 as in Figure \ref{subfig:indirect}.
-
+%
 Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
@@
 {\it integrity}.
-And in Figure \ref{subfig:WBCF}, the attacker attempts to return to the
+And in Figure \ref{subfig:WBCF}, the attacker arranges to return to the
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
@@
     \begin{tabular}{r l | l}
-      \labeledrow{100:}{addi ra,ra,20}{}
+      \labeledrow{100:}{addi ra,ra,16}{}
       \labeledrow{104:}{nop}{}
@@
 
-Figure \ref{fig:exec1} shows how the security context
-develops over several steps, with labels on the steps that allocate space and then make
-a call. Both program counter and stack pointer are initially 0, and the stack pointer
-is shown for convenience.
-\(\xrightarrow{\overline{\psi}}\) transitions are written \(\downarrow \overline{\psi}\)
-in this diagram; events are omitted.
-
-The first step  allocates words for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
+The security semantics for this program is based
+on the security-relevant events noted in the right columns of Figures~\ref{fig:main}
+and~\ref{fig:f}, namely execution of instructions that allocate or deallocate space,
+make a call, or make a return.
+Figure \ref{fig:exec1} shows how the corresponding security context
+develops over the first few steps of the program. We assume execution begins
+at the start of {\tt main} (\(\PCname = 0\)) and that the stack pointer is initially at
+address 1000. %  the \(PC\) and \(SP\) at each step are shown for clarity.
+\(\xrightarrow{\overline{\psi}}\) transitions are numbered and
+written \(\downarrow \overline{\psi}\) in this diagram; events are omitted.
+
+Step 1 allocates a word each for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
 as two words for the return address. This will have the
 effect of marking those bytes \(\object\), assuming they were previously
-\(\unsealed\).
-
+\(\unsealed\). (We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).\apt{To avoid confusion with list brackets.} )
+%
 \begin{figure*}
-  \begin{subfigure}[t]{.5\textwidth}
+  \begin{subfigure}[t]{.55\textwidth}
     \vskip 0pt
-    \begin{tabular}{| l | l || l |}
-      \hline
-      Step & \(\SP\) & \(\context\) \\
-      \hline
-      0 & 0 & \(V_0, \emplist\) \\
-      \hline
-      \multicolumn{3}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
-      \hline
-      1 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket, \emplist\) \\
-      \hline
-      \multicolumn{3}{l}{\(\vdots\)} \\
-      \hline
-      4 & -20 & \(V_1, \emplist\) \\
-      \hline
-      \multicolumn{3}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
-      \hline
-      5 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[V_1]\) \\
-      \hline
-      \multicolumn{3}{l}{\(\vdots\)} \\
-      \hline
-      8 & -20 & \(V_2,[V_1]\) \\
-      \hline
-      \multicolumn{3}{l}{\(\downarrow [\mathbf{return}]\)} \\
-      \hline
-      9 & -20 & \(V_1, \emplist\) \\
+    \begin{tabular}{|r|r||l|}
+      \hline
+      \(\PCname\) & \(\SP\) & \(\context\) \\
+      \hline
+      0 & 1000 & \(V_0, \emplist\) \\
+      \hline
+      \multicolumn{3}{l}{\(1 \downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
+      \hline
+      4 & 980 & \(V_1 = V_0 \llbracket 980..999 \mapsto \object\rrbracket, \emplist\) \\
+      \hline
+      \multicolumn{3}{l}{\(2 \downarrow []\)} \\
+      \multicolumn{3}{l}{\(\ \ \vdots\)} \\
+      \multicolumn{3}{l}{\(4 \downarrow []\)} \\
+      \hline
+      16 & 980 & \(V_1, \emplist\) \\
+      \hline
+      \multicolumn{3}{l}{\(5 \downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
+      \hline
+      100 & 980 & \(V_2 = V_1 \llbracket 980..999 \mapsto \sealed, \mathtt{a0} \mapsto \unsealed\rrbracket,[V_1]\) \\
+      \hline
+      \multicolumn{3}{l}{\(6 \downarrow []\)} \\
+      \multicolumn{3}{l}{\(\ \ \vdots\)} \\
+      \multicolumn{3}{l}{\(8 \downarrow []\)} \\
+      \hline
+      112 & 980 & \(V_2,[V_1]\) \\
+      \hline
+      \multicolumn{3}{l}{\(9 \downarrow [\mathbf{return}]\)} \\
+      \hline
+      20(?) & 980  & \(V_1, \emplist\) \\
       \hline
@@
 
-\caption{Execution through return from {\tt f}}
+\caption{Execution up through the return from {\tt f}\apt{Label rows on right. Also show registers?}}
 \label{fig:exec1}
 \end{figure*}
-
-On a call, the formerly active principal's record is pushed onto the inactive list.
+%
+At step 5, the formerly active principal's record is pushed onto the inactive list.
 Its return target is the return address of the call, 
@@
 For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments,
-in this case {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
-registers are mapped to \(\unsealed\). All callee-save registers are always \(\sealed\) and
-the program counter \(\public\), as they were initialized.
-
-On step 8, {\tt f} returns, and we restore the topmost inactive principal,
-{\tt main}'s principal. Now we will discuss the intuition behind how this
-structure translates into stack safety properties that rule out different
-classes of bad behavior.
+in this case just {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
+registers remain \apt{right?} mapped to \(\unsealed\). All callee-save registers are always \(\sealed\)
+as they were initialized\apt{???}.
+At step 9, {\tt f} returns, and the topmost inactive principal, for {\tt main}, is restored.
+
+We now show how this security semantiy can be used define notions of confidentiality,
+integrity, and correct control flow in such a way that many classes of
+bad behavior, including the attacks in Figure~\ref{fig:f}, are 
+are detected as security violations.
 
@@
 
-But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
-\(\SP \neq -20\)), as would be the case in attack (d)? We consider this to
+To begin with, what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
+\(\SP \neq 980\)). We consider this to
 violate well-bracketed control flow (WBCF). WBCF is a relationship between
-call steps and their corresponding return steps: after the return, the program
-counter should be at the instruction following the state before the call,
+call steps and their corresponding return steps: just after the return, the program
+counter should be at the next instruction following the call,
 and the stack pointer should be the same as it was before the call.
@@
 Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
-20 to the return address and then returns, thus bypassing the check and outputting
-{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 20,
-so we define a predicate on states \(ret\) that holds when the
-program counter and stack pointer are both 20.
-
-Then we identify the state immediately following the return, if one occurs ---
-this will be the first state in which the pending call stack is smaller than it was
-inside of the call.
-WBCF requires that if \(s\) after that return, \(ret ~ s\) will hold.
-For nested calls, where the pending stack is initial larger, the same principal
+16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
+{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 980,
+so we define a predicate \(ret\) \apt{suggest capitalizing these predicates}
+on states that should hold just after the return, namely, the program counter
+should be 20 and the stack pointer should be 980. \apt{Maybe write this symbolically.}
+
+We can identify the point just after the return (if a return occurs)
+as the first state in which the pending call stack is smaller than it was
+just after the call.
+WBCF requires that if \(s\) is the state at that point, then \(ret ~ s\) holds.
+For nested calls, where the pending stack is initially larger, the same principle
 applies: \(ret ~ s\) must hold the next time the pending stack is the same size or smaller.
 
-To demonstrate why WBCF also cares about the stack pointer, consider an attack which
-returns with the stack pointer at -12 instead of -20. Now, compare the layout of the
-stack relative to the stack pointer as it should be (\(\SP\)) to the one that results by
-increasing the stack pointer before return (\(\SP'\)):
-
+To see why WBCF also cares about the stack pointer, consider an attack \apt{should this
+  be shown as a fifth attack in the figure?} which
+returns with \(\SP' = 998\) instead of the correct \(\SP = 980\), and recall
+the layout of {\tt main}'s frame: 
 \begin{tabular}{| l | l | l | l | l |}
@@
 
+\noindent
 In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
@@
 
-Stack integrity applies a similar notion to all of the memory that a function has
+Stack integrity applies a similar \apt{to what?} notion to all of the memory that a function has
 allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
@@
 becomes \(\sealed\). In more sophisticated systems discussed later, we will distinguish
-sealed objects from those that are intentionally passed to the callee.
+sealed objects from those that are intentionally passed to the callee. \apt{already said this}
 Intuitively, the integrity of {\tt main}
@@
 to be identical to when it made the call.
-
+%
+Again, we need to know when a caller has been returned to,
+and we use the same mechanism of checking the depth of the call stack.
+%
 In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
-addresses -20 through -1 (allocated as private data) and callee-saved registers such as
-the stack pointer and return address. Note that callee-saved registers often change
+addresses 980 through 999 (allocated as private\apt{??} data) and callee-saved registers such as
+the stack pointer. Note that callee-saved registers often change
 during the call---but if so, the callee is obligated to restore them to their prior value.
 
-One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
-in a way that is harmless. Or, an enforcement mechanism might defer the checking
-for illegal writes until later in the execution, entering an error state at some
-point after return but before the illegal write is able to damage the system.
-This is the case in a lazy micro-policy: the monitor permits the caller to write into the
+While this notion of integrity is simple, it is stricter than necessary for many purposes.
+For example, it arguably doesn't matter if {\tt f} overwrites {\tt main}'s data provided that {\tt main} never
+reads that corrupted data.  This is the justification behind the lazy micro-policy~\cite{??}:
+the monitor permits the caller to write into the
 callee's frame, but taints that write so that the callee cannot read it later without
-triggering a failstop. In such a scenario we will consider a value {\it irrelevant}
-if it could be replaced with any other value without changing the observable behavior
+triggering a failstop. More generally, the corruption can be considered acceptable
+as long as it does not affect {\tt main}'s observable behavior. \apt{Maybe justify this
+  point of view by appeal to standard security notions?}
+To characterize this situation, we define a state element to be \emph{irrelevant}
+if its value can be replaced with any other value without changing the observable behavior
 of the machine. More precisely, for a set of elements \(\components\),
 a pair of states \(\mach\) and \(\nach\)
-are {\em \(\components\)-variants} if they agree on the value of every
-element except those in \(\components\), and share the same policy state.
-The elements of \(\components\) are said to be
-irrelevant in \(\mach\) if every variant state produces an equivalent trace
-(see Section \ref{sec:events}.)
-
-So, to define integrity, we again need to know when a caller has been returned to,
-and we use the same mechanism of checking the depth of the call stack.
-Integrity holds if, at that returned state,
-any element that is \(\sealed\) under the callee's view is either restored
-to its original value or is irrelevant.
+are {\em \(\components\)-variants} if t nhey agree on the value of every
+element except those in \(\components\), and share the same policy state\apt{hopefully can remove this last}.
+The elements of \(\components\) are \emph{irrelevant} in \(\mach\) if every variant state produces an equivalent trace
+(see Section \ref{sec:events}); all other elements are \emph{relevant}.
+We characterize \emph{observational integrity} as the property that
+every \emph{relevant} element that is \(\sealed\) under the callee's view is restored
+to its original value at the return point.
 
@@
 The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
-which is \(\sealed\). Figure \ref{fig:variant} shows the state at step 5, just after the call,
+which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the all at step 5,
 assuming that {\tt sec} is 5. Similar to WBCF, we will define
-\(int ~ s\) as a predicate on states that holds if, of the sealed
-addresses in state 5, all values that differ in \(s\) are irrelevant. We once again
-require that \(int\) hold on the state following the matching return.
-
-That state is reached at step 9. Since {\tt sensitive} has changed, we must determine whether
-it is irrelevant. We consider a variant state in which it has any other value, arbitrarily
+\(int\) \apt{suggest $Int$} as a predicate on states that holds if
+all relevant sealed addresses in \(s\) are the same as after step 5.
+We once again require that \(int\) hold on the state following the matching return.
+That state is reached by step 9. Although {\tt sensitive} has changed, it might be
+irrelevant. Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
 choosing 43. As execution continues after the return from the original state, it
-passes the check on {\tt sensitive}, while the execution from the variant does not, resulting
-in differing outputs. Therefore {\tt sensitive} is not irrelevant, so \(int\) does not hold,
-and integrity has been violated.
+passes the {\tt if}-test on {\tt sensitive}, whereas the execution from the variant does not, resulting
+in differing outputs. Therefore {\tt sensitive} is relevant, so \(int\) does not hold,
+and integrity has indeed been violated.
 
@@
 
-Consider again the state \((\mach, (V_2,\sigma))\) at step 5, this time with the attacker from
+Consider again the state \((\mach, (V_2,\sigma))\) after step 5, this time with the attacker from
 Figure \ref{subfig:direct}. We take a variant state over the set of element(s) that are
@@
   \includegraphics[width=\columnwidth]{variants2.png}
-  \caption{Internal Confidentiality Violation}
+  \caption{Internal Confidentiality Violation\apt{needs key cells to be colored}}
   \label{fig:variant2}
@@
 by reading it and then returning it, in a context where the caller would output the returned
-value. Figure \ref{subfig:variant3} shows the behavior of the same variants under this attacker,
+value. Figure \ref{fig:variant3} shows the behavior of the same variants under this attacker,
 but in this case, there is no output during the call. Instead the value of {\tt secret} is
-extracted and placed in {\tt a0}, the return value register. To illustrate the subtlety
-of identifying a leak, {\tt f} has also stored a 0 below the stack pointer.
-
+extracted and placed in {\tt a0}, the return value register. We wish to identify this as
+a confidentiality violation, again by considering variants of the \(\sealed\) and \(\unsealed\)
+elements in \(V_2\), but capturing the required property is subtle.
+
+To illustrate the issues, note that {\tt f} has also stores a 0 below the stack pointer.
+%
 \begin{figure}
@@
 \end{figure}
-
-Consider three elements: the address \(\SP - 4\) (blue), the address \(\SP + 12\) (yellow),
+%
+Now consider three elements: the address \(\SP - 4\) (blue), the address \(\SP + 12\) (yellow),
 and the register {\tt a0} (red). In the first case, the execution from the variant state
@@
 
-But in the case of {\tt a0}, the value has changed during the call (on both sides,
-although only one would be necessary.) And its final value differs between the variants.
-Therefore, it must depend on a secret. Indeed, we happen to know that it depends on {\tt secret}
-itself. We must therefore insist that {\tt a0} be irrelevant in the original trace.
-We term this property {\it return-time confidentiality}.
+But in the case of {\tt a0}, the value has changed during the call (in both the original
+and the variant, although only one of these would be necessary), and its final value differs between the variants.
+Therefore, it must depend on a secret (in fact, the variable {\tt secret}).
+Unless {\tt a0} happens to be irrelevant in the original trace, this is a violation of what
+we term {\it return-time confidentiality}.
 
@@
 variants. We begin with a state immediately following
-a call. Let that be \((\mach,\context)\). We suppose an arbitrary variant state,
-\((\nach,\context)\), which may vary any element that is \(\sealed\) or \(\unsealed\).
-That is, any element that is not used legitimately to pass arguments. Caller confidentiality
-therefore can be thought of as the callee's insensitivity to parts of its initial states
+a call. Let that be \((\mach,\context)\). \apt{do we need the $\context$ part??} We consider an arbitrary variant state,
+\((\nach,\context)\), which may vary any element that is \(\sealed\) or \(\unsealed\), 
+i.e., any element that is not used legitimately to pass arguments. Caller confidentiality
+therefore can be thought of as the callee's insensitivity to elements in its initial state
 that are not part of the caller-callee interface.
 
-In place of a predicate we define \(\mathit{conf}\) as a binary relation on states.
-\(\mathit{conf}\) will hold on eventual return states \((\mach',\context')\) and
+In place of a predicate we define \(\mathit{conf}\) as a binary relation on states, 
+which holds on eventual return states \((\mach',\context')\) and
 \((\nach',\context'')\)
-if all elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\),
-and which differ between \(\mach\) and \(\nach\), are irrelevant. We then define
-return-time confidentiality as holding if \(conf\) holds on the matching return states
+iff all relevant elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\)
+are the same in \(\mach\) and \(\nach\). We then define
+return-time confidentiality as holding if \(\mathit{conf}\) holds on the matching return states
 from \(\mach\) and \(\nach\), respectively.
 
-Together, these definitions encompass the confidentiality of the caller from the callee.
+Finally, we define caller confidentiality as the 
+combination of internal and return-time confidentiality.
 
@@
 we have:
-\[(|\sigma| \uparrow \mathit{cconf}(\mach',V)) ~ (\mach',\context')\)
+\[(|\sigma| \uparrow \mathit{cconf}(\mach',V)) ~ (\mach',\context')\]
 

### 2022-11-06 submission5/noexceptions.tex
typo

@@
 
-We now show how this security semantiy can be used define notions of confidentiality,
+We now show how this security semantics can be used define notions of confidentiality,
 integrity, and correct control flow in such a way that many classes of

### 2022-11-06 submission5/noexceptions.tex
more typos

@@
 callee's frame, but taints that write so that the callee cannot read it later without
-triggering a failstop. More generally, the corruption can be considered acceptable
+triggering a failstop. More generally, corruption by {\tt f} can be considered acceptable
 as long as it does not affect {\tt main}'s observable behavior. \apt{Maybe justify this
@@
 a pair of states \(\mach\) and \(\nach\)
-are {\em \(\components\)-variants} if t nhey agree on the value of every
+are {\em \(\components\)-variants} if they agree on the value of every
 element except those in \(\components\), and share the same policy state\apt{hopefully can remove this last}.
@@
 We characterize \emph{observational integrity} as the property that
-every \emph{relevant} element that is \(\sealed\) under the callee's view is restored
+every relevant element that is \(\sealed\) under the callee's view is restored
 to its original value at the return point.
@@
 The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
-which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the all at step 5,
+which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
 assuming that {\tt sec} is 5. Similar to WBCF, we will define

### 2022-11-14 submission5/noexceptions.tex
Progress

@@
 and are drawn from the set \(\WORDS\).
-Registers in the set \(\REGS\) are ranged over by \(\reg\), with special names given to the
-program counter \(\PCname\) and stack pointer \(\SP\).
+Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
+given the special name \(\SP\).
 They are divided into sets of caller-saved and callee-saved registers;
@@
 
+In addition to memory and registers, a machine has a program counter \(\PCname\)
+and may have additional architecture-specific state; particularly relevant is state
+that is used for security enforcement. We treat this as a set \(\mathcal{P}\) with
+\(\PCname \in \mathcal{P}\).
+
 \begin{figure}
@@
 Collectively addresses and registers are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \WORDS + \REGS\).
-%
-{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
-that defines a mapping \(\mach[\component]\) from state elements to words.
+in the set \(\COMPONENTS ::= \WORDS + \REGS + \mathcal{P}\).
+%
+The set of {\em machine states} \(\MACHS\) (ranged over by \(\mach\) and \(\nach\))
+defines the mapping \(\mach[\component]\) from state elements to words defined.
+
 {\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\), with a
@@
 and an event.
-While the tag-based enforcement mechanism may cause the machine to
+
+While some enforcement mechanisms may cause a machine to
 failstop, we model this as the machine silently diverging by remaining in the
@@
 
-\apt{Maybe should drop the math displays from this section, and keep to words.}
-Security-relevant operations consist of calls (including tailcalls), returns,
-allocations and deallocations. \apt{No tailscalls for now?} Calls and tailcalls identify the address of the call's target,
-and the argument registers. \apt{Either explain $\overline{sa}$ or omit it.} Allocations and deallocations are phrased in terms of a pair
+Security-relevant operations consist of calls, returns, and allocations.
+Calls identify the address of the call's target,
+and the argument registers. Allocations and deallocations are phrased in terms of a pair
 of an offset from the stack pointer and a size. Returns don't require any additional information.
@@
 \begin{align*}
-  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
-  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
-  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
-  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
+  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} & \\
   | & \mathbf{return} & \\
   | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
-  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
 \end{align*}
@@
 of the system that maps each state element to a {\it security class}:
-\[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
-\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]
-
-A view indicates, from the perspective of a given principal, whether an element is
-private to an inactive principal (\(\sealed\)),
+it may be private to an inactive principal (\(\sealed\)),
 in an allocated object that is accessible to the current principal (\(\object\)),
 available to be allocated (\(\unsealed\)), or
-always accessible (\(\public\)).
-The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
+always accessible (\(\public\)). The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
 all other locations to \(\public\), and registers based on which set they
 belong to (Figure \ref{fig:RISCVregs}): \(\sealed\) for callee-saved,
-\(\unsealed\) for caller-saved, and \(\public\) otherwise.
-
-For convenience, we will abbreviate multiple updates of a mapping
-\(V[\component_0 \mapsto sc_0][\component_1 \mapsto sc_1]\dots\)
-as \(V \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
-are fomulae with free variables bound to sets in \(C\).
-\apt{can omit this para}
+\(\unsealed\) for caller-saved, and \(\public\) otherwise. Elements of \(\mathcal{P}\),
+such as the program counter, are always \(\public\).
 
@@
 principals), ranged over by \(\sigma\).  
-\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
+%\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
 The initial context is \(\context_0 = (V_0, \emplist)\).
 
-\apt{Maybe don't need to describe $Op$ or $\stepstounder{\overline{\psi}}$ here, 
-  and maybe not even combined states, although they are used at the moment.}
-  Every security-relevant operation manipulates the call stack via a function
-\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
-
-From a machine state and a context, we create a {\it combined state}
-\(s = (\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
-labeled with the same operations. Note that \(Op\) takes as its state argument
-\(\mach\), the state before the step.
-
-\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
-            {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
-            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context', \obs)\)}
-
-\subsection{Policy State}
-
-\apt{As we discussed at length, it would certainly be nice to avoid separate policy state if possible.}
-In addition to its elements, a system may contain additional ``policy'' state that carries
-information relevant to a hardware enforcement mechanism. We assume that policy state
-cannot influence execution, except possibly to cause the system to failstop, which is modeled
-as a state stepping to itself. A failstop is therefore indistinguishable from silent
-divergence.
-
-Formally, we can treat our machine state as
-\(\mach = (\mach[\cdot],\langle \mach \rangle) \in \MACHS
-::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
-\apt{This overloaded notation doesn't work. $\mach$ can be either a pair or a map, not both.
-Note that earlier text strongly implies that it is a map.}
-Then \(\langle \mach \rangle\) is the security state. We assume that if
-\(\mach[\component] = \nach[\component]\) for all \(\component\),
-\(\mach \stepstounder{\bar{\psi_1}} \mach'\), and \(\nach \stepstounder{\bar{\psi_2}} \nach'\),
-then \(\bar{\psi_1} = \bar{\psi_2}\) and either
-\(\mach = \mach'\), \(\nach = \nach'\), or \(\mach'[\component] = \nach'[\component]\)
-for all \(\component\).
+As the machine takes a step that executes a particular operation, it updates the
+corresponding context, as described informally in the following section, and formally
+in Section \ref{sec:props}.
+
+%  Every security-relevant operation manipulates the call stack via a function
+%\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
+
+%From a machine state and a context, we create a {\it combined state}
+%\(s = (\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
+%labeled with the same operations. Note that \(Op\) takes as its state argument
+%\(\mach\), the state before the step.
+
+%\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
+%            {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
+%            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context', \obs)\)}
+
+%\subsection{Policy State}
+
+%\apt{As we discussed at length, it would certainly be nice to avoid separate policy state if possible.}
+%In addition to its elements, a system may contain additional ``policy'' state that carries
+%information relevant to a hardware enforcement mechanism. We assume that policy state
+%cannot influence execution, except possibly to cause the system to failstop, which is modeled
+%as a state stepping to itself. A failstop is therefore indistinguishable from silent
+%divergence.
+
+%Formally, we can treat our machine state as
+%\(\mach = (\mach[\cdot],\langle \mach \rangle) \in \MACHS
+%::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
+%\apt{This overloaded notation doesn't work. $\mach$ can be either a pair or a map, not both.
+%Note that earlier text strongly implies that it is a map.}
+%Then \(\langle \mach \rangle\) is the security state. We assume that if
+%\(\mach[\component] = \nach[\component]\) for all \(\component\),
+%%\(\mach \stepstounder{\bar{\psi_1}} \mach'\), and \(\nach \stepstounder{\bar{\psi_2}} \nach'\),
+%then \(\bar{\psi_1} = \bar{\psi_2}\) and either
+%\(\mach = \mach'\), \(\nach = \nach'\), or \(\mach'[\component] = \nach'[\component]\)
+%for all \(\component\).
 
@@
 its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
-it should not---instead it will publish the results of {\tt f}.
+it should not---instead it will output the results of {\tt f}.
 
@@
 at the start of {\tt main} (\(\PCname = 0\)) and that the stack pointer is initially at
-address 1000. %  the \(PC\) and \(SP\) at each step are shown for clarity.
+address 1000.
 \(\xrightarrow{\overline{\psi}}\) transitions are numbered and
@@
 effect of marking those bytes \(\object\), assuming they were previously
-\(\unsealed\). (We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).\apt{To avoid confusion with list brackets.} )
-%
+\(\unsealed\). (We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).)
+%
+
 \begin{figure*}
@@
       \hline
-      \(\PCname\) & \(\SP\) & \(\context\) \\
+      \(\PCname\) & \(\SP\) & Context \\
       \hline
@@
       \hline
-      20(?) & 980  & \(V_1, \emplist\) \\
+      20 & 980  & \(V_1, \emplist\) \\
       \hline
@@
     \vskip 0px
+
     \includegraphics[width=\columnwidth]{stack1.png}
@@
 in this case just {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
-registers remain \apt{right?} mapped to \(\unsealed\). All callee-save registers are always \(\sealed\)
-as they were initialized\apt{???}.
-At step 9, {\tt f} returns, and the topmost inactive principal, for {\tt main}, is restored.
+registers remain mapped to \(\unsealed\). All callee-save registers remain \(\sealed\) for all calls.
+At step 9, {\tt f} returns, and the topmost inactive view, that of {\tt main}, is restored.
 
@@
 {\tt secret}. Before the call, the program counter is 16 and the stack pointer is 980,
-so we define a predicate \(ret\) \apt{suggest capitalizing these predicates}
-on states that should hold just after the return, namely, the program counter
-should be 20 and the stack pointer should be 980. \apt{Maybe write this symbolically.}
+so we define a predicate \(\ret\)
+on states that should hold just after the return, namely, on states \(\mach\)
+such that \(\mach[\PCname] = 20\) and \(\mach[\SP] = 980\).
 
@@
 just after the call.
-WBCF requires that if \(s\) is the state at that point, then \(ret ~ s\) holds.
+WBCF requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
 For nested calls, where the pending stack is initially larger, the same principle
-applies: \(ret ~ s\) must hold the next time the pending stack is the same size or smaller.
+applies: \(\ret ~ \mach\) must hold the next time the pending stack is the same size or smaller.
 
@@
 
-Stack integrity applies a similar \apt{to what?} notion to all of the memory that a function has
+Like WBCF, stack integrity defines at the call a condition that must hold upon
+return. This time the condition applies to all of the memory that a function has
 allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
 upon allocation, the view labeled is \(\object\), and when a call is made, it instead
-becomes \(\sealed\). In more sophisticated systems discussed later, we will distinguish
-sealed objects from those that are intentionally passed to the callee. \apt{already said this}
-Intuitively, the integrity of {\tt main}
+becomes \(\sealed\). Intuitively, the integrity of {\tt main}
 is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
@@
 In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
-addresses 980 through 999 (allocated as private\apt{??} data) and callee-saved registers such as
+addresses 980 through 999 and callee-saved registers such as
 the stack pointer. Note that callee-saved registers often change
@@
 assuming that {\tt sec} is 5. Similar to WBCF, we will define
-\(int\) \apt{suggest $Int$} as a predicate on states that holds if
-all relevant sealed addresses in \(s\) are the same as after step 5.
-We once again require that \(int\) hold on the state following the matching return.
+\(\intProp\) as a predicate on states that holds if
+all relevant sealed addresses in \(\mach\) are the same as after step 5.
+We once again require that \(\intProp\) hold on the state following the matching return.
 That state is reached by step 9. Although {\tt sensitive} has changed, it might be
@@
 passes the {\tt if}-test on {\tt sensitive}, whereas the execution from the variant does not, resulting
-in differing outputs. Therefore {\tt sensitive} is relevant, so \(int\) does not hold,
+in differing outputs. Therefore {\tt sensitive} is relevant, so \(\intProp\) does not hold,
 and integrity has indeed been violated.
@@
 
-Consider again the state \((\mach, (V_2,\sigma))\) after step 5, this time with the attacker from
+Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\), with the attacker from
 Figure \ref{subfig:direct}. We take a variant state over the set of element(s) that are
@@
 Figure \ref{fig:variant2}.) If we take a trace of execution from each state until it returns,
-the traces may differ, in this case outputting 5 (the original value of {\tt sec}) and
+the traces may differ, in this case outputting 5 (the original value of {\tt secret}) and
 4 (its value in the variant) respectively. This is a violation of
@@
 variants. We begin with a state immediately following
-a call. Let that be \((\mach,\context)\). \apt{do we need the $\context$ part??} We consider an arbitrary variant state,
-\((\nach,\context)\), which may vary any element that is \(\sealed\) or \(\unsealed\), 
+a call, \(\mach\). We consider an arbitrary variant state,
+\(\nach\), which may vary any element that is \(\sealed\) or \(\unsealed\), 
 i.e., any element that is not used legitimately to pass arguments. Caller confidentiality
@@
 
-In place of a predicate we define \(\mathit{conf}\) as a binary relation on states, 
+In place of a predicate we define \(\conf\) as a binary relation on states, 
 which holds on eventual return states \((\mach',\context')\) and
@@
 are the same in \(\mach\) and \(\nach\). We then define
-return-time confidentiality as holding if \(\mathit{conf}\) holds on the matching return states
+return-time confidentiality as holding if \(\conf\) holds on the matching return states
 from \(\mach\) and \(\nach\), respectively.
@@
 written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\),
-and \(\langle\mach\rangle = \langle\nach\rangle\).
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
 
@@
 %
-\[ret(\mach) = \{(\mach',\context') \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}\]
+\[\ret(\mach) = \{(\mach',\context') \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}\]
 
@@
 \((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) where 
-\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
-%or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\)
-\(|\sigma| \uparrow ret(\mach)\) holds on \((\mach',(V,\sigma))\).
+\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\),
+\(|\sigma| \uparrow \ret(\mach)\) holds on \((\mach',(V,\sigma))\).
 
@@
 
-\definition We define \(int(\mach,V)\), to hold on a state \((\mach',\context')\) when
+\definition We define \(\intProp(\mach,V)\), to hold on a state \((\mach',\context')\) when
 \(\{\component \mid V ~ \component = \sealed\} \cap \Delta(\mach,\mach')\)
@@
 \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
-% or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
 we have:
-\[(|\sigma'| \uparrow int(\mach',V)) ~ (\mach',(V,\sigma))\]
+\[(|\sigma'| \uparrow \intProp(\mach',V)) ~ (\mach',(V,\sigma))\]
 
@@
 
-\definition We define the relation \(conf(\mach,\nach)\) to hold on a pair of states
+\definition We define the relation \(\conf(\mach,\nach)\) to hold on a pair of states
 \((\mach',\context)\) and \((\nach',\context')\) when
@@
 and any \(\nach\) that is a \(\components\)-variant of \(\mach\), we have:
-\[(\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \mathit{conf}(\mach,\nach) ~ (\nach,(V,\sigma))\]
+\[(\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \conf(\mach,\nach) ~ (\nach,(V,\sigma))\]
 
@@
 with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-%or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \psi\),
 \((\mach',\context')\) enjoys both internal and return-time confidentiality.
@@
 
-\definition We define \(\mathit{cconf}(\mach,V)\), to hold on a state \((\mach',\context')\) when
+\definition We define \(\cconf(\mach,V)\), to hold on a state \((\mach',\context')\) when
 \(\{\component \mid V ~ \component = \unsealed\} \cap \Delta(\mach,\mach')\)
@@
 we have:
-\[(|\sigma| \uparrow \mathit{cconf}(\mach',V)) ~ (\mach',\context')\]
-
-\section{Shared Memory}
+\[(|\sigma| \uparrow \cconf(\mach',V)) ~ (\mach',\context')\]
+
+\section{Advanced Machine}
+
+The advanced machine defines additional operations, and adds a field \(\overline{sa}\) to
+the \(\mathbf{call}\) operation, which is a set of triples of a register, a base offset, and a size,
+and denotes that the object at the offset from the given register is to be passed as an argument.
+
+\begin{align*}
+  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
+  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
+  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
+  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
+  | & \mathbf{return} & \\
+  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+\end{align*}
 
@@
 
+\section{Enforcement}
+
+\paragraph{Protecting Enforcement State}
+
+State beyond registers and memory is always considered \(\public\), and in a PIPE machine this
+includes tags. That does not mean that functions should be modifying them! It just means that we
+do not consider them to be part of the state being protected. It is less obvious that a function
+should be prevented reading a tag, and our properties do not rule it out, although in practice
+PIPE does not allow such reads.
+
 \bibliographystyle{IEEEtran}

### 2022-11-15 submission5/noexceptions.tex
Most updates

@@
 Collectively addresses and registers are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \WORDS + \REGS + \mathcal{P}\).
-%
-The set of {\em machine states} \(\MACHS\) (ranged over by \(\mach\) and \(\nach\))
-defines the mapping \(\mach[\component]\) from state elements to words defined.
+in the set \(\COMPONENTS ::= \WORDS + \REGS + \mathcal{P}\). A {\em machine state}
+\(\mach\) is a map from components to words, written \(\mach[\component]\).
+
+\sna{It occurs to me, based on some of your comments further down, that the old machine
+  states were unclear. My intention was not for \(\mach\) to be a map, but for
+  it to be an arbitrary type on which \(\cdot[\cdot]\) was defined. (Akin to
+  the coq idiom of a record with a field t:Type and proj:t\(\rightarrow\)k\(\rightarrow\)w.)
+  Therefore, implicitly, there could be additional state for enforcement and other
+  purposes. The new presentation with P makes that explicit but is less elegant.
+}
 
@@
 failstop, we model this as the machine silently diverging by remaining in the
-same state. Since failstops are not distinguished from normal execution,
+same state. Since failstops are not distinguished from normal divergence,
 our properties will suffer the limitations typical of
@@
 \(\unsealed\) for caller-saved, and \(\public\) otherwise. Elements of \(\mathcal{P}\),
-such as the program counter, are always \(\public\).
+such as the program counter, are always \(\public\). Even if they might deserve their
+own protection, that protection is orthogonal to stack safety.
+
+\sna{And the payoff of \(\mathcal{P}\) being explicit: we can just state this.}
 
@@
   \end{subfigure}
+  \begin{subfigure}[b]{\columnwidth}
+    \vspace{\abovedisplayskip}
+    \begin{tabular}{r l | l}
+      \labeledrow{100:}{addi sp,sp,8}{}
+      \labeledrow{104:}{nop}{}
+      \labeledrow{108:}{nop}{}
+      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
+    \end{tabular}
+    \subcaption{Relative address attack}
+    \label{subfig:WBCF2}
+  \end{subfigure}
 
@@
 
-    \includegraphics[width=\columnwidth]{stack1.png}
+    \(\underbrace{~ ~ ~}_\unsealed \uparrow \SP \)
+
     \includegraphics[width=\columnwidth]{stack2.png}
@@
 
-To see why WBCF also cares about the stack pointer, consider an attack \apt{should this
-  be shown as a fifth attack in the figure?} which
-returns with \(\SP' = 998\) instead of the correct \(\SP = 980\), and recall
-the layout of {\tt main}'s frame: 
+To see why WBCF also cares about the stack pointer, consider the attack in
+Figure \ref{subfig:WBCF2}. Here the attacker returns with \(\SP' = 998\) instead of the
+correct \(\SP = 980\). Given the layout of {\tt main}'s frame: 
 \begin{tabular}{| l | l | l | l | l |}
@@
 the stack pointer. Note that callee-saved registers often change
-during the call---but if so, the callee is obligated to restore them to their prior value.
-
-While this notion of integrity is simple, it is stricter than necessary for many purposes.
-For example, it arguably doesn't matter if {\tt f} overwrites {\tt main}'s data provided that {\tt main} never
-reads that corrupted data.  This is the justification behind the lazy micro-policy~\cite{??}:
+during the call---but if the caller accesses them after the call, it should find them
+restored to their prior value.
+
+While it would be simple to define integrity as ``all sealed elements retain their
+values after the call,'' this would be stricter than necessary for many purposes.
+For example, it arguably doesn't matter if {\tt f} overwrites {\tt main}'s data provided
+that {\tt main} never reads that corrupted data.
+This is the justification behind the lazy micro-policy~\cite{??}:
 the monitor permits the caller to write into the
 callee's frame, but taints that write so that the callee cannot read it later without
-triggering a failstop. More generally, corruption by {\tt f} can be considered acceptable
-as long as it does not affect {\tt main}'s observable behavior. \apt{Maybe justify this
-  point of view by appeal to standard security notions?}
+triggering a failstop. More generally, it is common to specify the behavior of a system,
+including security properties, in terms of traces of observable events rather than details
+of the machine state. For any such notion of observable behavior, we can consider
+corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.
+
 To characterize this situation, we define a state element to be \emph{irrelevant}
@@
 are {\em \(\components\)-variants} if they agree on the value of every
-element except those in \(\components\), and share the same policy state\apt{hopefully can remove this last}.
-The elements of \(\components\) are \emph{irrelevant} in \(\mach\) if every variant state produces an equivalent trace
+element except those in \(\components\). The elements of \(\components\) are \emph{irrelevant}
+in \(\mach\) if every variant state produces an equivalent trace
 (see Section \ref{sec:events}); all other elements are \emph{relevant}.

### 2022-11-15 submission5/noexceptions.tex
Fixing diagrams
Update to comments

@@
 
-\usepackage{multirow}
+\usepackage{array,multirow}
 
@@
 in Section \ref{sec:props}.
-
-%  Every security-relevant operation manipulates the call stack via a function
-%\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
-
-%From a machine state and a context, we create a {\it combined state}
-%\(s = (\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
-%labeled with the same operations. Note that \(Op\) takes as its state argument
-%\(\mach\), the state before the step.
-
-%\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
-%            {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
-%            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context', \obs)\)}
-
-%\subsection{Policy State}
-
-%\apt{As we discussed at length, it would certainly be nice to avoid separate policy state if possible.}
-%In addition to its elements, a system may contain additional ``policy'' state that carries
-%information relevant to a hardware enforcement mechanism. We assume that policy state
-%cannot influence execution, except possibly to cause the system to failstop, which is modeled
-%as a state stepping to itself. A failstop is therefore indistinguishable from silent
-%divergence.
-
-%Formally, we can treat our machine state as
-%\(\mach = (\mach[\cdot],\langle \mach \rangle) \in \MACHS
-%::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
-%\apt{This overloaded notation doesn't work. $\mach$ can be either a pair or a map, not both.
-%Note that earlier text strongly implies that it is a map.}
-%Then \(\langle \mach \rangle\) is the security state. We assume that if
-%\(\mach[\component] = \nach[\component]\) for all \(\component\),
-%%\(\mach \stepstounder{\bar{\psi_1}} \mach'\), and \(\nach \stepstounder{\bar{\psi_2}} \nach'\),
-%then \(\bar{\psi_1} = \bar{\psi_2}\) and either
-%\(\mach = \mach'\), \(\nach = \nach'\), or \(\mach'[\component] = \nach'[\component]\)
-%for all \(\component\).
 
@@
 
+\newcommand{\freebox}{\includegraphics[height=10px]{free.png}}
+\newcommand{\pubbox}{\includegraphics[height=10px]{public.png}}
+\newcommand{\objbox}{\includegraphics[height=10px]{object.png}}
+\newcommand{\sealbox}{\includegraphics[height=10px]{sealed.png}}
+
 \begin{figure*}
-  \begin{subfigure}[t]{.55\textwidth}
-    \vskip 0pt
-    \begin{tabular}{|r|r||l|}
-      \hline
-      \(\PCname\) & \(\SP\) & Context \\
-      \hline
-      0 & 1000 & \(V_0, \emplist\) \\
-      \hline
-      \multicolumn{3}{l}{\(1 \downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
-      \hline
-      4 & 980 & \(V_1 = V_0 \llbracket 980..999 \mapsto \object\rrbracket, \emplist\) \\
-      \hline
-      \multicolumn{3}{l}{\(2 \downarrow []\)} \\
-      \multicolumn{3}{l}{\(\ \ \vdots\)} \\
-      \multicolumn{3}{l}{\(4 \downarrow []\)} \\
-      \hline
-      16 & 980 & \(V_1, \emplist\) \\
-      \hline
-      \multicolumn{3}{l}{\(5 \downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
-      \hline
-      100 & 980 & \(V_2 = V_1 \llbracket 980..999 \mapsto \sealed, \mathtt{a0} \mapsto \unsealed\rrbracket,[V_1]\) \\
-      \hline
-      \multicolumn{3}{l}{\(6 \downarrow []\)} \\
-      \multicolumn{3}{l}{\(\ \ \vdots\)} \\
-      \multicolumn{3}{l}{\(8 \downarrow []\)} \\
-      \hline
+  \begin{tabular}{|r|r||l|r}
+    \cline{1-3}
+    \(\PCname\) & \(\SP\) & Context &
+    \multirow{3}{*}{\(\underbrace{\dots \freebox \freebox \freebox \freebox \freebox
+        \freebox \freebox \freebox \freebox \freebox}_\unsealed
+      \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\pubbox} \!\! \pubbox \pubbox \dots}_\public
+      ~ \stackrel{\mathtt{a1}}{\pubbox} ~ \stackrel{\mathtt{a2}}{\freebox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      \)} \\
+    \cline{1-3}
+    0 & 1000 & \(V_0, \emplist\)
+    \\
+    \cline{1-3}
+    \multicolumn{3}{l}{\multirow{2}{*}{\(1 \Big\downarrow [\mathbf{alloc} ~ (-20,20)]\)}} & \\
+    \multicolumn{3}{l}{} &
+    \multirow{3}{*}{\(\underbrace{\dots \freebox \freebox \freebox \freebox \freebox}_\unsealed
+      \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\objbox} \!\! \objbox \objbox \objbox \objbox}_\object
+      \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
+      ~ \stackrel{\mathtt{a1}}{\pubbox} ~ \stackrel{\mathtt{a2}}{\freebox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      \)}
+    \\
+    \cline{1-3}
+    4 & 980 & \(V_1 = V_0 \llbracket 980..999 \mapsto \object\rrbracket, \emplist\) &
+    \\
+    \cline{1-3}
+    \multicolumn{3}{l}{\multirow{2}{*}{2-4 \(\Big\downarrow \emplist\)}} \\ \multicolumn{3}{l}{} \\
+    \cline{1-3}
+    16 & 980 & \(V_1, \emplist\) & \\
+    \cline{1-3}
+    \multicolumn{3}{l}{\multirow{2}{*}{\(5 \Big\downarrow [\mathbf{call} ~ 100 ~ \emplist]\)}} & \\
+    \multicolumn{3}{l}{} &
+    \multirow{3}{*}{\(\underbrace{\dots \freebox \freebox \freebox \freebox \freebox}_\unsealed
+      \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\sealbox} \!\! \sealbox \sealbox \sealbox \sealbox}_\sealed
+      \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
+      ~ \stackrel{\mathtt{a1}}{\freebox} ~ \stackrel{\mathtt{a2}}{\freebox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      \)}
+    \\
+    \cline{1-3}
+    100 & 980 & \(V_2 = V_1 \llbracket 980..999 \mapsto \sealed, \mathtt{a0} \mapsto \unsealed\rrbracket,[V_1]\) & \\
+    \cline{1-3}
+    \multicolumn{3}{l}{\multirow{2}{*}{6-8 \(\Big\downarrow \emplist\)}} \\ \multicolumn{3}{l}{} \\
+    \cline{1-3}
     112 & 980 & \(V_2,[V_1]\) \\
-      \hline
-      \multicolumn{3}{l}{\(9 \downarrow [\mathbf{return}]\)} \\
-      \hline
-      20 & 980  & \(V_1, \emplist\) \\
-      \hline
+    \cline{1-3}
+    \multicolumn{3}{l}{\multirow{2}{*}{\(9 \Big\downarrow [\mathbf{return}]\)}} & \\
+    \multicolumn{3}{l}{} & \multirow{3}{*}{\(\underbrace{\dots \freebox \freebox \freebox \freebox \freebox}_\unsealed
+      \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\objbox} \!\! \objbox \objbox \objbox \objbox}_\object
+      \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
+      ~ \stackrel{\mathtt{a1}}{\pubbox} ~ \stackrel{\mathtt{a2}}{\freebox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      \)}
+    \\
+    \cline{1-3}
+    20 & 980  & \(V_1, \emplist\) &
+    \\
+    \cline{1-3}
+    \multicolumn{2}{l}{} \\
   \end{tabular}
-  \end{subfigure}
-  \begin{subfigure}[t]{.4\textwidth}
-    \vskip 0px
-
-    \(\underbrace{~ ~ ~}_\unsealed \uparrow \SP \)
-
-    \includegraphics[width=\columnwidth]{stack2.png}
-    \includegraphics[width=\columnwidth]{stack3.png}
-    \includegraphics[width=\columnwidth]{stack4.png}
-  \end{subfigure}
-
-\caption{Execution up through the return from {\tt f}\apt{Label rows on right. Also show registers?}}
+\caption{Execution up through the return from {\tt f}}
 \label{fig:exec1}
@@
   \includegraphics[width=\columnwidth]{variants2.png}
-  \caption{Internal Confidentiality Violation\apt{needs key cells to be colored}}
+  \caption{Internal Confidentiality Violation}
   \label{fig:variant2}
@@
 But in the case of {\tt a0}, the value has changed during the call (in both the original
-and the variant, although only one of these would be necessary), and its final value differs between the variants.
+and the variant, although only one of these would be necessary), and its final value
+differs between the variants.
 Therefore, it must depend on a secret (in fact, the variable {\tt secret}).
@@
 In place of a predicate we define \(\conf\) as a binary relation on states, 
-which holds on eventual return states \((\mach',\context')\) and
-\((\nach',\context'')\)
+which holds on eventual return states \(\mach'\) and \(\nach'\)
 iff all relevant elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\)
@@
 
-\section{Events and Traces}
+\section{Property Formalisms}
+
+We have described informally how the security context develops as the system performs
+security-relevant operations. Formally, we combine each machine state with a context
+to create a {\it combined state} \(s = (\mach,\context)\) and a lift the transition
+transition \(\stepstounder{}\) on combined states, labeled with the same operations.
+with each step, the context updates based on the function
+\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
+Note that \(Op\) takes as its first argument the state {\it before} the step.
+Since a single step might correspond to multiple operations, we apply
+\(Op\) as many times as is needed.
+
+\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
+            {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
+            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context', \obs)\)}
+
+The definition of \(Op\) is most convenient to present decomposed into
+rules for each operation. In the simple system, the operations affect the
+context as follows:
+
+\judgmentbr[Alloc]
+           {\(b = \mach[\SP] + \mathit{off}\)}
+           {\(V' = V \llbracket \addr \mapsto \sealed |
+             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+           {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+
+\judgmentbr[Call]
+           {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}}\)}
+           {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+             \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+           {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
+             (V',V::\sigma)\)}
+
+\judgment[Ret]
+         {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
+         {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
+
+\subsection{Events and Traces}
 \label{sec:events}
@@
 
-\section{Facts Abouts Calls and Returns}
+\subsection{Facts Abouts Calls and Returns}
 \label{sec:facts}
@@
 
-\section{Properties}
+\subsection{Properties}
 \label{sec:props}
 
-We now finally have everything we need to formalize our definitions for stack safety.
-Figure \ref{fig:oprules} gives the operation rules for all of the operations we've described
-so far.
-
-\begin{figure}
-
-  \judgmentbr[Alloc]
-             {\(b = \mach[\SP] + \mathit{off}\)}
-             {\(V' = V \llbracket \addr \mapsto \sealed |
-               b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
-             {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-
-  \judgmentbr[Call]
-             {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
-             {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-               \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-             {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
-               (V',V::\sigma)\)}
-
-  \judgment[Ret]
-           {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-           {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
-
-%\judgment[RetDef]
-%         {\(\chi = \mathbf{return}\)}
-%         {\(Cop ~ \chi ~ (V, \emplist) ~ \mach \triangleq
-%           (V, \emplist)\)}
-
-  \caption{Core operation rules for the simple system}
-  \label{fig:oprules}
-\end{figure}
-
-\subsection{Well-bracketed Control Flow}
+\paragraph*{Well-bracketed Control Flow}
 
@@
 
-\subsection{Integrity}
+\paragraph*{Integrity}
 
@@
 
-\subsection{Caller Confidentiality}
+\paragraph*{Caller Confidentiality}
 
@@
 
-\subsection{Callee Confidentiality}
+\paragraph*{Callee Confidentiality}
 
@@
 
-The advanced machine defines additional operations, and adds a field \(\overline{sa}\) to
-the \(\mathbf{call}\) operation, which is a set of triples of a register, a base offset, and a size,
-and denotes that the object at the offset from the given register is to be passed as an argument.
+The advanced machine defines additional operations, and extends some of the old ones.
+It adds a field \(\overline{sa}\) to the \(\mathbf{call}\) operation, which is a set of
+triples of a register, a base offset, and a size, and denotes that the object at the offset
+from the given register is to be passed as an argument. The new \(\mathbf{tailcall}\) operation
+has the same fields, but deals with tail call optimizations in which the callee will reuse
+the caller's stack. The \(\mathbf{alloc}\) operation
+gains a boolen flag indicating whether the allocation is to be public, accessible to
+all functions until deallocated. And, we add the corresponding \(\mathbf{dealloc}\)
+operation.
 
@@
   | & \mathbf{return} & \\
-  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+  | & \mathbf{alloc} ~ \mathit{pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N}, \mathit{pub} \in \mathbb{B} \\
   | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
@@
 
-%We can extend our basic machine to support public memory allocations that can
-%be accessed by anyone until they are deallocated on their owner's return,
-%using the \(\mathbf{alloc\_pub}\) operation. This behaves identically to
-%\(\mathbf{alloc\_priv}\) except that it marks the locations \(\public\).
-%For example, consider the following code:
-
-%{\tt
-%  void g() \{
-
-%  ~ ~ int x[3];
-
-%  ~ ~ int y;
-
-%  ~ ~ h(\&x);
-
-%  \}
-%}
-
-%Here we have allocated several arrays and variables; {\tt y}
-%will behave as normal with \(\mathit{alloc\_priv}\), but {\tt x} has its
-%address taken and passed to another function. That function might do anything at
-%all to {\tt x}, and we have to decide whether it can be said to violate stack
-%safety. A reasonable default is to assume that any address-taken local should be
-%public, while {\tt y} is still private.
-
-%In the likely event that the space for the return address, {\tt x}, and {\tt y}
-%are all allocated in a single step, that step needs multiple labels to capture
-%that they are not all private. Its label would instead be:
-%\[\begin{split}
-%& \psi = [\mathbf{alloc\_priv} ~ (20,4); \\
-%  & \mathbf{alloc\_pub} ~ (16,12) ; \mathbf{alloc\_priv} ~ (4,4)]
-%\end{split}\]
-
-%Allocating public objects affects the call stack very similarly to private ones,
-%but instead of treating them as \(\sealed\) they become \(\public\). This makes these
-%regions of memory able to transmit information between caller and callee without
-%violating the stack safety properties. Note that there is no way to seal a location
-%once it has become \(\public\); only when {\tt g} returns will the memory allocated
-%to {\tt x} be available for private use.
-
-%\judgmentbr{\(b = \mach[\SP] - \mathit{off}\)}
-%           {\(V' = V \llbracket \addr \mapsto \public |
-%             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
-%           {\(Pop ~ (\mathbf{alloc\_pub} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach \triangleq
-%             (V',\sigma)\)}
+\subsection{Public Allocations}
+
+The advanced machine supports public stack allocations that can
+be accessed by anyone until they are deallocated on their owner's return,
+represented by \(\mathbf{alloc} ~ \mathbf{t} \dots\).
+For example, consider the following code:
+
+{\tt
+  void g() \{
+
+  ~ ~ int x[3];
+
+  ~ ~ int y;
+
+  ~ ~ h(\&x);
+
+  \}
+}
+
+Here we have allocated several arrays and variables; {\tt y}
+will behave as normal with \(\mathbf{alloc} ~ \mathbf{f} \dots\), but
+{\tt x} has its address taken and passed to another function. That
+function might do anything at all to {\tt x}, and might store its pointer
+elsewhere (in the heap, for instance) so that the scope of legitimate accesses
+to {\tt x} are uncertain. We default to assuming that any address-taken local
+should be public. Meanwhile {\tt y} is still private.
+
+In the likely event that the space for the return address, {\tt x}, and {\tt y}
+are all allocated in a single step, that step makes multiple allocation operations:
+\[\bar{\psi} = [\mathbf{alloc} ~ \mathbf{f} ~ (-20,4);
+ \mathbf{alloc} ~ \mathbf{t} ~ (-16,12) ; \mathbf{alloc} ~ \mathbf{f} ~ (-4,4)]\]
+
+Allocating public objects affects the call stack very similarly to private ones,
+but instead of labeling them \(\object\) they become \(\public\), so they are
+not sealed at a call. This makes these
+regions of memory able to transmit information between caller and callee without
+violating the stack safety properties. Note that there is no way to seal a location
+once it has become \(\public\); only when {\tt x} is deallocated will its region
+be available to allocate privately.
+
+\judgmentbr[AllocPriv]
+           {\(b = \mach[\SP] + \mathit{off}\)}
+           {\(V' = V \llbracket \addr \mapsto \sealed |
+             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+           {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+
+\judgmentbr[AllocPub]
+           {\(b = \mach[\SP] + \mathit{off}\)}
+           {\(V' = V \llbracket \addr \mapsto \public |
+             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+           {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+
+\judgmentbr[Dealloc]
+           {\(b = \mach[\SP] + \mathit{off}\)}
+           {\(V' = V \llbracket \addr \mapsto \unsealed |
+             b \leq a < b+\mathit{sz} \land V ~ \addr \in \{\object, \public\} \rrbracket\)}
+           {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
 
@@
              {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
-               (V',V::\sigma)\)}
+               (V'',V::\sigma)\)}
 
@@
 
-This same structure works for pass-by-reference arguments: we know from the type of
-the called function which register it expects to hold the address of a given object,
-and the size of that object. In this case the offset will always be 0.
-
-When the callee in such a scenario makes a call of its own, the objects passed to it
-will in turn be sealed---unless it passes them on, which can happen with pass-by-reference
-arguments!
+This same structure works for pass-by-reference arguments. In this case {\tt main}
+passes an array {\tt x} directly, which is compiled as a pass-by-reference.
+
+{\tt
+  int main() \{
+
+  ~ int x[1];
+
+  ~ f(x);
+
+  ~ g();
+
+  ~ return x[0];
+
+  \}
+}
+
+Unlike the address-taken array, we can clearly delineate the scope in which {\tt x} may be
+accessed without violating stack safety: {\tt f} should be allowed to access it, but further
+calls from {\tt f} should not (unless {\tt f} passes it to them) and {\tt g} certainly
+should not. Say the compiler opts to pass the pointer to {\tt x} in register {\tt a0}.
+The {\tt f} is specifically being given access to the four bytes starting at that address.
+The call operation is \(\mathbf{call} ~ \emplist ~ \{(\mathtt{a0},0,4)\}\).
+
+\subsection{Tailcalls}
+
+The operation rule for a tailcall is similar to that for a normal call.
+We do not push the caller's view onto the stack,
+but replace it outright. This means that a tailcall does not increase the size of
+the call stack, and therefore for purposes of our properties, all tailcalls will
+be considered to return simultaneously when the eventual {\bf return} operation
+pops the top of the stack.
+
+Since the caller will not be returned to, it does not need integrity, though
+it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
+than \(\sealed\) to express this. Of course, some objects may be passed in place,
+easily modeled by including them in \(\overline{sa}\).
+
+\judgmentbrbr[]
+             {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
+             {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+               \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+             {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+             {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
+               (V',\sigma)\)}
 
@@
 should be prevented reading a tag, and our properties do not rule it out, although in practice
-PIPE does not allow such reads.
+PIPE does not allow such reads. It would be quite possible to implement
+the properties so that they would forbid these activities, for instance by structuring \(\mathcal{P}\)
+to treat tags separately and apply stack safety to the tags on the stack, but there is no obvious
+application for such a stringent property---the enforcement state must be protected in order that
+enforcement not be subverted, not for its own sake.
+
+I have a sketch of an argument that we would not gain anything meaningful from confidentiality
+of tags, but it still needs some work. Still, I think it's correct not to vary the policy state.
 

### 2022-11-17 submission5/noexceptions.tex
typos and a couple of comments

@@
 \(\mach\) is a map from components to words, written \(\mach[\component]\).
+\apt{As we discussed, this seems too restrictive a definition for $\mathcal{P}$.
+  For example, for tagged machines it might contain a mapping from $\WORDS + \REGS$ to tags.}
 
@@
   | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
 \end{align*}
@@
 violating the program's {\it well-bracketed control flow} (WBCF.)
+\apt{Add brief description of \ref{subfig:WBCF2}.}
 
@@
 To begin with, what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
-\(\SP \neq 980\)). We consider this to
+\(\SP \neq 980\))? We consider this to
 violate well-bracketed control flow (WBCF). WBCF is a relationship between
@@
 Figure \ref{subfig:WBCF2}. Here the attacker returns with \(\SP' = 998\) instead of the
-correct \(\SP = 980\). Given the layout of {\tt main}'s frame: 
+correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame 
 \begin{tabular}{| l | l | l | l | l |}
@@
 \noindent
-In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
-read part of the return address, and then it will attempt to output
-{\tt res}, but instead output {\tt secret}! Even absent other
+{\tt main}'s attempt to read {\tt sensitive} will instead 
+read part of the return address, and its attempt to output
+{\tt res} will instead output {\tt secret}! Even absent other
 kinds of data protection, the stack pointer {\it must} be restored
@@
 including security properties, in terms of traces of observable events rather than details
-of the machine state. For any such notion of observable behavior, we can consider
+of the machine state. Given such a specification, we can consider
 corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.
@@
 are {\em \(\components\)-variants} if they agree on the value of every
-element except those in \(\components\). The elements of \(\components\) are \emph{irrelevant}
-in \(\mach\) if every variant state produces an equivalent trace
+element except those in \(\components\). \apt{Explicitly note that this excludes elements of
+  of $\mathcal{P}$ (assuming that is indeed the case).} The elements of \(\components\) are \emph{irrelevant}
+in \(\mach\) if every variant state produces an equivalent observation trace
 (see Section \ref{sec:events}); all other elements are \emph{relevant}.
@@
 
-The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
+In our example setting, the observation trace consists of the sequence of values written to {\tt out}. The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
 which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
@@
 values that it has stored in \(\unsealed\) elements. So, callee confidentiality means that
-at the state following the callee's return, any element that was \(\unsealed\) under the
-callee's view either retains its original value or is irrelevant.
+at the state following the callee's return, any relevant element that was \(\unsealed\) under the callee's view retains its original value.
 
@@
 
-In principal, these properties can exist in any combination, but in practice we see examples
+In principle, these properties can exist in any combination, but in practice we see examples
 of certain combinations that form a hierarchy. Many enforcement techniques focus purely on
@@
 
-\paragraph*{Moving Forward}
+\paragraph*{Moving Forward}\apt{Suggest omit this para.}
 
@@
 
-\section{Property Formalisms}
+\section{Formalizing Properties}
 

### 2022-11-28 submission5/noexceptions.tex
Almost there

@@
 that is used for security enforcement. We treat this as a set \(\mathcal{P}\) with
-\(\PCname \in \mathcal{P}\).
+\(\PCname \in \mathcal{P}\). Collectively these are {\em state elements} \(\component\)
+in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS | \mathcal{P}\).
 
@@
 
-Collectively addresses and registers are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \WORDS + \REGS + \mathcal{P}\). A {\em machine state}
-\(\mach\) is a map from components to words, written \(\mach[\component]\).
-\apt{As we discussed, this seems too restrictive a definition for $\mathcal{P}$.
-  For example, for tagged machines it might contain a mapping from $\WORDS + \REGS$ to tags.}
-
-\sna{It occurs to me, based on some of your comments further down, that the old machine
-  states were unclear. My intention was not for \(\mach\) to be a map, but for
-  it to be an arbitrary type on which \(\cdot[\cdot]\) was defined. (Akin to
-  the coq idiom of a record with a field t:Type and proj:t\(\rightarrow\)k\(\rightarrow\)w.)
-  Therefore, implicitly, there could be additional state for enforcement and other
-  purposes. The new presentation with P makes that explicit but is less elegant.
-}
+A machine has a set \(\mathcal{V}\) of values such that \(\WORDS \subseteq \mathcal{V}\).
+A {\em machine state} \(\mach\) is a map from components to values, written \(\mach[\component]\).
 
@@
 own protection, that protection is orthogonal to stack safety.
-
-\sna{And the payoff of \(\mathcal{P}\) being explicit: we can just state this.}
 
@@
 of the machine. More precisely, for a set of elements \(\components\),
-a pair of states \(\mach\) and \(\nach\)
-are {\em \(\components\)-variants} if they agree on the value of every
-element except those in \(\components\). \apt{Explicitly note that this excludes elements of
-  of $\mathcal{P}$ (assuming that is indeed the case).} The elements of \(\components\) are \emph{irrelevant}
+a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
+agree on the value of every element except those in \(\components\). This includes
+public elements such as those in \(\mathcal{P}\).
+The elements of \(\components\) are \emph{irrelevant}
 in \(\mach\) if every variant state produces an equivalent observation trace
@@
 
-\paragraph*{Moving Forward}\apt{Suggest omit this para.}
-
-In the next section we will define traces and their (hyper-)properties.
-Then in Section \ref{sec:facts} we will define a mechanism to make assertions
-about what will happen on a future return, and Section \ref{sec:props} will use these
-to give the formal definitions of integrity and confidentiality.
-
 \section{Formalizing Properties}
@@
 
-\section{Enforcement}
-
-\paragraph{Protecting Enforcement State}
+\section{Enforcement and Testing}
+
+We use our properties to test the stack safety micro-policies of Roessler and DeHon,
+particularly their Lazy Tagging and Clearing. Our underlying machine is a RISC-V
+based on the MIT RISC-V formalism \cite{} and extended with {\it metadata tags},
+represented as additional state in \(\mathcal{P}\), one for each register, address,
+and the program counter:
+\[\mathcal{P} ::= t_\reg | t_\addr | t_\PCname\]
+%
+We divide values into normal memory words and tags drawn from a policy-specific
+set of tags \(\mathcal{T}\), \(\mathcal{V} ::= \mathcal{W} | \mathcal{T}\).
+The machine's step function corresponds to that of the basic RISC-V machine,
+with the behavior of the tags at each step parameterized by a {\it micro-policy},
+which may turn a normal step into an error (represented by returning to the same state.)
+
+\paragraph*{Protecting Enforcement State}
 

### 2022-11-29 submission5/noexceptions.tex
Ready for sending

@@
 For confidentiality and integrity, our properties are {\em observational},
-meaning that they allow a function activation to read from and write to other activations'
-data, as long as such reads and writes do not affect the observable behavior of the
+meaning that they are only concerned with the observable behavior of the system.
+A function activation might read from and write to other activations' data
+without violating an observational property, as long as such reads and writes do not
+affect the observable behavior of the
 system. This permissiveness enables our properties to handle lazy enforcement mechanisms.
@@
 by labeling instructions as calls, returns, etc.
-Such a labeling scheme appears in
-our testing environment (Section [TBD]) and is already necessary
-to correctly apply tag-based enforcement in practice (cite something about this.)
+Such a labeling scheme appears in our testing environment (Section [TBD]).% and is already necessary
+%to correctly apply tag-based enforcement in practice (cite something about this.)
 
@@
 
-In section (TBD) we discuss the extension of this model into a simple concurrency setup.
+%In section (TBD) we discuss the extension of this model into a simple concurrency setup.
 
@@
 in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS | \mathcal{P}\).
+One instance of \(\mathcal{P}\) is discussed in Section \ref{sec:enforcement}.
 
@@
 failstop, we model this as the machine silently diverging by remaining in the
-same state. Since failstops are not distinguished from normal divergence,
-our properties will suffer the limitations typical of
-{\it termination insensitive} systems \cite{}. TODO: more detailed explanation later.
+same state. We consider silently diverging traces to be indistinguishable
+(from the point of view of an attacker) from traces that might eventually
+produce output, an assumption that gives our properties the flavor of
+{\it termination insensitivity}.
 
@@
 {\it integrity}.
-And in Figure \ref{subfig:WBCF}, the attacker arranges to return to the
+In Figure \ref{subfig:WBCF}, the attacker arranges to return to the
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
-violating the program's {\it well-bracketed control flow} (WBCF.)
-\apt{Add brief description of \ref{subfig:WBCF2}.}
+violating the program's {\it well-bracketed control flow} (WBCF.) Figure \ref{subfig:WBCF2}
+also shows an attack that violates WBCF, this time by returning to the correct
+program counter but with the stack pointer offset.
 
@@
 What if we want to express a finer-grained notion of safety, in which
-such objects are protected unless the function that owns them intentionally
-passes a pointer to them?
-
-In order to express such a property, we need our machine to carry some notion
+stack objects are protected unless the function that owns them intentionally
+passes a pointer to them? This can be thought of as a {\it capability}-based
+notion of security. Capabilities are unforgeable tokens that grant access to
+a region of memory, typically corresponding to valid pointers to that region.
+So, in order to express such a property, we need our machine to carry some notion
 of {\it pointer provenance}---a distinction between a pointer that is intended to
 point to a given object, and non-pointer integers as well as pointers to other objects.
-Then \(\mathit{passed}\) should recursively identify all objects that
-are reachable transitively from valid pointers in registers and accessible memory;
-these keep their \(\object\) status, while those that are not reachable become \(\sealed\).
-
-\section{Enforcement and Testing}
+
+One such provenance model is Memarian et al.'s PVI \cite{}, in which pointers are
+annotated with the identity of the object they first pointed to. This annotation is
+propagated when the pointer is copied and when operations are performed on it---even
+integer-only operations.
+
+We can model this as a pair of additional security-relevant operations: one which
+declares a register to contain a valid pointer, and one which transmits the provenance
+of a pointer from one element to another.
+
+\begin{align*}
+  \psi \in \Psi ::= & \dots \\
+  & \mathbf{promote} ~ \reg ~ sa & sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
+  & \mathbf{propagate} ~ \component_{src} ~ \component_{dst} \\
+\end{align*}
+
+In addition to the normal call stack, our security context will carry a map \(\rho\) from
+elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
+Existing operations are extended to propagate \(\rho\), and the new operations work as follows:
+
+\judgmenttwobrlong[]
+                  {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},sz)\)}
+                  {\(b = \mach[\reg_{base}]+\mathit{off}\)}
+                  {\(\rho' = \rho[\reg \mapsto (b,b+sz)]\)}
+                  {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+
+\judgmenttwo[]
+            {\(\psi = \mathbf{propagate} ~ \component_{src} ~ \component_{dst}\)}
+            {\(\rho' = \rho[\component_{dst} \mapsto \rho[\component_{src}]]\)}
+            {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+
+We now have a notion of provenance, and must integrate it into the definition of
+stack safety. We essentially generalize the above notion of passing: we will consider
+a caller to have intentionally passed an object if that object is reachable by
+a capability that has been passed to the callee. This includes capabilities passed
+indirectly, by being stored in an object that is in turn passed. Formally, we call
+this set \(\mathit{capped}\), and define it recursively:
+
+\[\mathit{capped} ~ \components ~ \rho \triangleq
+\bigcup_{\component \in \components} \{\component' |
+\rho[\component] = (\mathit{base},\mathit{bound})
+\land \mathit{base} \leq \component' < \mathit{bound}\}\]
+
+We then tweak the call operation to seal only objects that are in \(\mathit{capped}\), or
+the previously defined \(\mathit{passed}\).
+
+\judgmentbrbrbr[]
+               {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
+               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(\components = \{\addr | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \cup (\mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \}\)}
+               {\(V'' = V'\llbracket \components \mapsto \sealed \rrbracket\)}
+               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) =
+                 (V'',V::\sigma,\rho)\)}
+
+Note that we have a degree of monotonicity here. Once an object is sealed (because its
+capability has not been passed to a callee), subsequent nested calls can never unseal it.
+On the other hand, an object that is passed via a pointer may be passed on indefinitely.
+
+\section{Enforcement}
+\label{sec:enforcement}
 
@@
 
-I have a sketch of an argument that we would not gain anything meaningful from confidentiality
-of tags, but it still needs some work. Still, I think it's correct not to vary the policy state.
+\section*{Remainder}
+
+These document sections lay out a way to present our stack safety properties to build intuition and
+then introduce the formalism. Next the full paper will need the testing section and an expanded
+enforcement section, but these can largely be borrowed from previous submissions.
 

### 2022-12-02 submission5/noexceptions.tex
Fixed recursion
Done integrating tags into examples

@@
 confidentiality, albeit at some performance cost.
-%
-
-%For ease of exposition, we initially assume a single simple
-%stack with no sharing between callers and callees: all parameters and return
-%values are passed in registers.  Later, we will introduce enhanced versions
-%supporting
-%(1) passing of scalar stack data,
-%(2) callee-saves registers calling conventions, and
-%(3) a coroutine system with a static layout.
 
@@
 
-In addition to memory and registers, a machine has a program counter \(\PCname\)
-and may have additional architecture-specific state; particularly relevant is state
-that is used for security enforcement. We treat this as a set \(\mathcal{P}\) with
-\(\PCname \in \mathcal{P}\). Collectively these are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS | \mathcal{P}\).
-One instance of \(\mathcal{P}\) is discussed in Section \ref{sec:enforcement}.
+Along with the program counter, \(\PCname\), these are {\em state elements} \(\component\)
+in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).
 
@@
 
-A machine has a set \(\mathcal{V}\) of values such that \(\WORDS \subseteq \mathcal{V}\).
-A {\em machine state} \(\mach\) is a map from components to values, written \(\mach[\component]\).
+A {\em machine state} \(\mach\) is a map from components to a set \(\mathcal{V}\) of
+values. A value \(v\) may be convertible to an address, written \(|v|\) (\(|\cdot|\)
+being a partial function.) We write \(\mach[\component]\) to denote the value at
+\(\component\) of \(\mach\) and \(\mach[v]\) to denote the value at \(|v|\) if
+defined.
 
@@
 
-In this section we consider a simple setting with calls, returns, and private allocations.
+In this section we consider a simple setting with calls, returns, and private allocations,
+on a RISC-V machine with a PIPE tagging system. For some type \(\mathcal{T}\) of tags,
+\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\), with \(|(w,t)| = w\).
 Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
@@
 
+In typical compiled code, WBCF is maintained by the compiler's call and return sequences.
+As long as a function deallocates the same-sized frame that it allocates, and properly
+saves and restores the return address, it should have WBCF---unless an error in the middle
+subverts those sequences by overwriting one of the relevant elements, or by jumping into
+the middle of the call or return sequence. We can therefore enforce WBCF by (1) ensuring that
+the call and return sequences execute in order and (2) permitting only call and return
+sequences to write to \(\SP\), {\tt ra}, and the address in memory where {\tt ra} is stored.
+
+Let \(\mathcal{T} ::= \mathit{gen} | \mathit{seq} ~ n ~ m | \mathit{priv}\), named for general
+execution, sequence, and privileged, respectively. Call and return sequences of length \(m\) are
+tagged \(\mathit{seq} ~ 1 ~ m\) through \(\mathit{seq} ~ m ~ l\), and the micro-policy enforces
+that they execute in order. The stack pointer and {\tt ra}
+are tagged \(\mathit{priv}\). Only instructions tagged \(\mathit{seq} \dots\) may write to memory
+or registers tagged \(\mathit{priv}\). Store instructions tagged \(\mathit{seq} \dots\) tag their
+targets \(\mathit{priv}\).
+
 \paragraph*{Stack Integrity}
@@
 that {\tt main} never reads that corrupted data.
-This is the justification behind the lazy micro-policy~\cite{??}:
-the monitor permits the caller to write into the
-callee's frame, but taints that write so that the callee cannot read it later without
-triggering a failstop. More generally, it is common to specify the behavior of a system,
-including security properties, in terms of traces of observable events rather than details
-of the machine state. Given such a specification, we can consider
+To illustrate, we can consider two integrity micro-policies: ``Depth Isolation'' and
+``Lazy Tagging and Clearing.''
+
+Depth Isolation tracks, via a tag on the program counter,
+the depth of the current activation. Stack frames are initialized and cleared, by writing
+to every address in them, so that during a call at depth \(d\), the current activation's
+frame is tagged with \(d\). Loads and stores targeting addresses so tagged failstop
+if the program counter does not match. In this setting, at depth \(d\), all memory tagged
+with a depth less than \(d\) corresponds to \(\sealed\) memory, and is inaccessible.
+When its depth is reached via a return, it will have the same value, so integrity is
+enforced.
+
+On the other hand, Lazy Tagging and Clearing skips the (problematically linear time)
+initialization and clearing steps, and only updates tags when memory is written to.
+It allows a callee to overwrite sealed memory, but that write tags the corrupted memory
+with the callee's identity. If the corrupted memory's owner (or any other activation) tries
+to read it, only then does the micro-policy failstop.
+
+Lazy Tagging and Clearing illustrates the value of ``extensional'' properties.
+It is common to specify the behavior of a system, including security properties, in
+terms of traces of observable events rather than details of the machine state.
+Lazy enforcement works because the deferred check catches errors before they appear
+in the trace. The nature of traces varies based on the properties of interest, but
+given a particular model of observation, we can consider
 corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.
@@
 a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
-agree on the value of every element except those in \(\components\). This includes
-public elements such as those in \(\mathcal{P}\).
+agree on the value of every element except those in \(\components\).
+In the tagged environment, we additionally need the values in \(\components\)
+to agree on tags, but not necessarily their data.
+
 The elements of \(\components\) are \emph{irrelevant}
@@
 
-In our example setting, the observation trace consists of the sequence of values written to {\tt out}. The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
+In our example setting, the observation trace consists of the sequence of values written to {\tt out}.
+The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
 which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
@@
 irrelevant. Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
-choosing 43. As execution continues after the return from the original state, it
-passes the {\tt if}-test on {\tt sensitive}, whereas the execution from the variant does not, resulting
-in differing outputs. Therefore {\tt sensitive} is relevant, so \(\intProp\) does not hold,
-and integrity has indeed been violated.
+choosing 43 (and keeping its tag, \(t_2'\), unchanged.) As execution continues after the return
+from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
+from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} is relevant,
+so \(\intProp\) does not hold, and integrity has indeed been violated.
 
@@
 combination of internal and return-time confidentiality.
+While we have separated the definitions of integrity and confidentiality,
+in terms of tag-based enforcement, both Depth Isolation and Lazy Tagging and Clearing
+enforce them.
 
@@
 values that it has stored in \(\unsealed\) elements. So, callee confidentiality means that
-at the state following the callee's return, any relevant element that was \(\unsealed\) under the callee's view retains its original value.
+at the state following the callee's return, any relevant element that was \(\unsealed\) under
+the callee's view retains its original value.
+
+Once again, both tag-based approaches do enforce callee confidentiality, but it is
+worth keeping separate because some enforcement mechanisms might not, and it has not
+historically been acknowledged by all formal treatments of stack safety \cite{Georges22:TempsDesCerise}.
 
@@
 
-\subsection{Tailcalls}
+\subsection{Tail Calls}
 
@@
 elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
-Existing operations are extended to propagate \(\rho\), and the new operations work as follows:
+Existing operations are extended to keep the value of \(\rho\) the same, and the new operations
+work as follows:
 
@@
 
-\[\mathit{capped} ~ \components ~ \rho \triangleq
-\bigcup_{\component \in \components} \{\component' |
-\rho[\component] = (\mathit{base},\mathit{bound})
-\land \mathit{base} \leq \component' < \mathit{bound}\}\]
+\[\begin{split}
+& \mathit{capped} ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{\component\} \cup \mathit{capped} ~ \components' ~ \rho \textnormal{ where} \\
+& \components' = \{\component' |\rho[\component] = (\mathit{base},\mathit{bound})
+\land \mathit{base} \leq \component' < \mathit{bound}\} \\
+\end{split}\]
 
@@
 
-\section{Enforcement}
-\label{sec:enforcement}
-
-We use our properties to test the stack safety micro-policies of Roessler and DeHon,
-particularly their Lazy Tagging and Clearing. Our underlying machine is a RISC-V
-based on the MIT RISC-V formalism \cite{} and extended with {\it metadata tags},
-represented as additional state in \(\mathcal{P}\), one for each register, address,
-and the program counter:
-\[\mathcal{P} ::= t_\reg | t_\addr | t_\PCname\]
-%
-We divide values into normal memory words and tags drawn from a policy-specific
-set of tags \(\mathcal{T}\), \(\mathcal{V} ::= \mathcal{W} | \mathcal{T}\).
-The machine's step function corresponds to that of the basic RISC-V machine,
-with the behavior of the tags at each step parameterized by a {\it micro-policy},
-which may turn a normal step into an error (represented by returning to the same state.)
-
-\paragraph*{Protecting Enforcement State}
-
-State beyond registers and memory is always considered \(\public\), and in a PIPE machine this
-includes tags. That does not mean that functions should be modifying them! It just means that we
-do not consider them to be part of the state being protected. It is less obvious that a function
-should be prevented reading a tag, and our properties do not rule it out, although in practice
-PIPE does not allow such reads. It would be quite possible to implement
-the properties so that they would forbid these activities, for instance by structuring \(\mathcal{P}\)
-to treat tags separately and apply stack safety to the tags on the stack, but there is no obvious
-application for such a stringent property---the enforcement state must be protected in order that
-enforcement not be subverted, not for its own sake.
-
-\section*{Remainder}
+%\section{Enforcement}
+%\label{sec:enforcement}
+
+%We use our properties to test the stack safety micro-policies of Roessler and DeHon,
+%particularly their Lazy Tagging and Clearing. Our underlying machine is a RISC-V
+%based on the MIT RISC-V formalism \cite{} and extended with {\it metadata tags},
+%represented as additional state in \(\mathcal{P}\), one for each register, address,
+%and the program counter:
+%\[\mathcal{P} ::= t_\reg | t_\addr | t_\PCname\]
+%
+%We divide values into normal memory words and tags drawn from a policy-specific
+%set of tags \(\mathcal{T}\), \(\mathcal{V} ::= \mathcal{W} | \mathcal{T}\).
+%The machine's step function corresponds to that of the basic RISC-V machine,
+%with the behavior of the tags at each step parameterized by a {\it micro-policy},
+%which may turn a normal step into an error (represented by returning to the same state.)
+
+%\paragraph*{Protecting Enforcement State}
+
+%State beyond registers and memory is always considered \(\public\), and in a PIPE machine this
+%includes tags. That does not mean that functions should be modifying them! It just means that we
+%do not consider them to be part of the state being protected. It is less obvious that a function
+%should be prevented reading a tag, and our properties do not rule it out, although in practice
+%PIPE does not allow such reads. It would be quite possible to implement
+%the properties so that they would forbid these activities, for instance by structuring \(\mathcal{P}\)
+%to treat tags separately and apply stack safety to the tags on the stack, but there is no obvious
+%application for such a stringent property---the enforcement state must be protected in order that
+%enforcement not be subverted, not for its own sake.
+
+\section*{[Meta] Remainder}
 

### 2022-12-03 submission5/noexceptions.tex

@@
 
-\section{The Setting}
-
-We state our properties in terms of a generic machine subject to a few constraints,
-but our examples and tests are focused on a RISC-V-like machine enhanced with PIPE,
-a tag-based reference monitor. Importantly, such a machine has no explicit ``call''
-instruction, only {\tt jal} instructions which may be used for that purpose or potentially
-other purposes. Similarly, there is no ``return'', only {\tt jalr}, and so on.
-We therefore define our properties under the assumption that we can distinguish the
-steps that represent these abstract operations from those that do not---for instance,
-by labeling instructions as calls, returns, etc.
-Such a labeling scheme appears in our testing environment (Section [TBD]).% and is already necessary
-%to correctly apply tag-based enforcement in practice (cite something about this.)
-
-The set of special operations that are recognized in this way are termed
-{\it security-relevant operations}. Our properties are phrased as a {\it security semantics},
-which extends the underlying machine with additional context about its
-security principals and which registers and memory they require to be secure.
-This context evolves dynamically through the execution of security-relevant operations.
-This notion of security may be phrased as
-predicates on future states, e.g. assertions of the form
-``When control returns to me...'', or relations on traces of future execution
-(hyper-properties).
-
-Unlike approaches that give a safe-by-construction semantics
-for various security-relevant operations and then prove secure compilation to the underlying
-enforcement mechanism, our machine is explicitly not safe unless the enforcement
-is correctly instantiated; the security semantics gives us the means to judge whether
-an enforcement scheme implements stack safety or not.
-
-We will first give a detailed example of a security semantics for a simple setting
+\section{Machine Model and Security Semantics}
+
+  We state our properties in terms of a generic machine, making minimal assumptions
+  on the machine architecture and the nature of its built-in safety enforcement mechanisms
+  (if any).  Our machine model is meant to be a faithful description of the hardware,
+  with a semantics that covers both ordinary instruction behavior and the behavior of
+  any tags, capabilities, etc. that the machine might support. 
+  This semantics is intentionally \emph{not} ``safe-by-construction;'' it gives
+  no guarantee that the enforcement hardware is being used correctly to achieve desired
+  security properties.
+
+  On top of this machine model, we layer a {\it security semantics}, which extends the
+  underlying machine with additional context about the identity of current and pending
+  functions (which act as security principals) and about the registers and memory they require
+  to be secure. This added context is purely notional;
+  it does not affect the real machine. The security context
+  evolves dynamically through the execution of {\it security-relevant operations},
+  which include events like calls, returns, and frame manipulations.
+  Our security properties are phrased in terms of this context, often as predicates
+  on future states, e.g. of the form ``when control returns to the current function...'',
+  or relations on traces of future execution (hyper-properties).
+
+  The security-relevant operations typically correspond to underling machine instructions,
+  but the correspondence may not be obvious just be inspecting the machine code.
+  For example, in the tagged RISC-V machine we use in our examples and tests,
+  calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
+  and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
+  instructions might also be used for other things. We therefore assume that there
+  is some external mechanism for distinguishing which instructions are security-relevant;
+  the most obvious technique is to add optional labels to instructions, and that is 
+  what we do in this paper, but the theory does not depend on this choice.
+
+  Our security semantics has similarities to the overlay semantics that have been proposed
+  to characterize stack safety in the past~\cite{??}, but, like the basic machine
+  semantics, it is not safe-by-construction. Rather, it gives us a means to \emph{define}
+  the security properties that a correct enforcement scheme should implement, using
+  some combination of hardware and software.
+
+In the next section, we give a detailed example of a security semantics for a simple setting
 in which our security-relevant operations are restricted to making calls and returns, and allocating
-private memory within the current function activation. We then extend this model
+private memory within the current function activation. In later sections, we extend this model
 to the full system that we test, which features:
@@
 objects are governed by a provenance-based capability system, accessible only to those
-functions which have recieved a valid pointer to them.
+functions which have received a valid pointer to them.
 
@@
 behavior resulting in its compilation to arbitrary machine code. A given enforcement
-mechanism may place additional constraints, of course, particularly on the behavior of
+mechanism may place additional constraints, particularly on the behavior of
 call and return sequences.
@@
 our examples we will identify one function or another as an attacker, but we do not
-require any static division between trusted and untrusted code, and aim to protect
+require any static division between trusted and untrusted code, and we aim to protect
 even buggy code.
@@
 This is a strong threat model, but hardware and timing attacks are out of scope,
-and our properties are termination insensitive as a result of the enforcement mechanism.
+and our properties are termination insensitive as a result of how failstops are
+modelled (see ~\ref{??}).
 
@@
 simplistic, assuming a fixed number of threads each with its own dedicated processor.
-We model memory safe stack objects, but not a heap. Regions outside of
+\apt{Is concurrency still alive?}
+We model memory-safe stack objects, but not a heap. Regions outside of
 stacks can be used however the compiler likes, including as a heap, but no protection is
@@
 Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
-given the special name \(\SP\).
-They are divided into sets of caller-saved and callee-saved registers;
-Figure \ref{fig:RISCVregs} gives an example division for a RISC-V machine as will
-appear in our examples and testing, along with the register names that will appear in
-example code, and their default security class (explained below).
-
+given the special name \(\SP\); 
+some registers may be classified as caller-saved or callee-saved.
+%
 Along with the program counter, \(\PCname\), these are {\em state elements} \(\component\)
@@
 
-A {\em machine state} \(\mach\) is a map from components to a set \(\mathcal{V}\) of
-values. A value \(v\) may be convertible to an address, written \(|v|\) (\(|\cdot|\)
-being a partial function.) We write \(\mach[\component]\) to denote the value at
+A {\em machine state} \(\mach\) is a map from elements to a set \(\mathcal{V}\) of
+\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\) (\(|\cdot|\).
+The details of value structure depend on the specific machine being modelled;  
+intuitively, the payload represents the part of the value that is relevant to
+the behavior of the basic machine, while the rest of the value may contain
+information relevant to a hardware enforcement mechanism (such as a tag).
+We write \(\mach[\component]\) to denote the value at
 \(\component\) of \(\mach\) and \(\mach[v]\) to denote the value at \(|v|\) if
@@
 all other locations to \(\public\), and registers based on which set they
-belong to (Figure \ref{fig:RISCVregs}): \(\sealed\) for callee-saved,
-\(\unsealed\) for caller-saved, and \(\public\) otherwise. Elements of \(\mathcal{P}\),
-such as the program counter, are always \(\public\). Even if they might deserve their
-own protection, that protection is orthogonal to stack safety.
+belong to: % (Figure \ref{fig:RISCVregs}):
+\(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
+%Elements of \(\mathcal{P}\),
+%such as the program counter, are always \(\public\). Even if they might deserve their
+%own protection, that protection is orthogonal to stack safety.
 
@@
 %\section{Calls, Returns, and Private Allocations}
-\section{Examples}
+\section{Properties by Example}
 
 In this section we consider a simple setting with calls, returns, and private allocations,
-on a RISC-V machine with a PIPE tagging system. For some type \(\mathcal{T}\) of tags,
-\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\), with \(|(w,t)| = w\).
+on a RISC-V machine with a PIPE tagging system. Here 
+\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
+for some type \(\mathcal{T}\) of tags to be elaborated later, with \(|(w,t)| = w\).
+Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
+security classes. 
 Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
@@
 {\tt main} takes a secret as its argument and initializes a variable to contain
-potentially sensitive data. We assume that observable events are stores
+potentially sensitive data. We assume that the only observable events are stores
 to a special address, called {\tt out}. 
@@
 violating the program's {\it well-bracketed control flow} (WBCF.) Figure \ref{subfig:WBCF2}
-also shows an attack that violates WBCF, this time by returning to the correct
+shows another attack that violates WBCF, this time by returning to the correct
 program counter but with the stack pointer offset.
@@
 {\tt secret}. Before the call, the program counter is 16 and the stack pointer is 980,
-so we define a predicate \(\ret\)
-on states that should hold just after the return, namely, on states \(\mach\)
-such that \(\mach[\PCname] = 20\) and \(\mach[\SP] = 980\).
-
+so we define a predicate on states that should hold just after the return:
+\(\ret\ \mach \triangleq \mach[\PCname] = 20 \wedge \mach[\SP] = 980\).
+%
 We can identify the point just after the return (if a return occurs)
@@
 WBCF requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
-For nested calls, where the pending stack is initially larger, the same principle
-applies: \(\ret ~ \mach\) must hold the next time the pending stack is the same size or smaller.
+%For nested calls, where the pending stack is initially larger, the same principle
+%applies: \(\ret ~ \mach\) must hold the next time the pending stack is the same size or smaller.
 
@@
 the middle of the call or return sequence. We can therefore enforce WBCF by (1) ensuring that
-the call and return sequences execute in order and (2) permitting only call and return
+these ``blessed'' call and return sequences execute in order and (2) permitting only blessed
 sequences to write to \(\SP\), {\tt ra}, and the address in memory where {\tt ra} is stored.
 
-Let \(\mathcal{T} ::= \mathit{gen} | \mathit{seq} ~ n ~ m | \mathit{priv}\), named for general
-execution, sequence, and privileged, respectively. Call and return sequences of length \(m\) are
-tagged \(\mathit{seq} ~ 1 ~ m\) through \(\mathit{seq} ~ m ~ l\), and the micro-policy enforces
+To perform this enforcement using tags, we can take
+\(\mathcal{T} ::= \mathit{gen} | \mathit{bseq} ~ n ~ m | \mathit{prot}\), named for general
+execution, blessed sequence, and protected, respectively. Blessed sequences of length \(m\) are
+tagged \(\mathit{bseq} ~ 1 ~ m\) through \(\mathit{bseq} ~ m ~ l\), and the micro-policy enforces
 that they execute in order. The stack pointer and {\tt ra}
-are tagged \(\mathit{priv}\). Only instructions tagged \(\mathit{seq} \dots\) may write to memory
-or registers tagged \(\mathit{priv}\). Store instructions tagged \(\mathit{seq} \dots\) tag their
-targets \(\mathit{priv}\).
+are tagged \(\mathit{prot}\). Only instructions tagged \(\mathit{bseq}\) may write to memory
+or registers tagged \(\mathit{prot}\). Store instructions tagged \(\mathit{bseq}\) tag their
+targets \(\mathit{prot}\).
 
@@
 
-Like WBCF, stack integrity defines at the call a condition that must hold upon
+Like WBCF, stack integrity defines a condition at the call that must hold upon
 return. This time the condition applies to all of the memory that a function has
 allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
-upon allocation, the view labeled is \(\object\), and when a call is made, it instead
+upon allocation, the view labels it \(\object\), and when a call is made, it instead
 becomes \(\sealed\). Intuitively, the integrity of {\tt main}
@@
 
-Depth Isolation tracks, via a tag on the program counter,
-the depth of the current activation. Stack frames are initialized and cleared, by writing
+Depth Isolation tracks the depth of the current activation via a tag on the program counter.
+Stack frames are initialized and cleared, by writing
 to every address in them, so that during a call at depth \(d\), the current activation's
-frame is tagged with \(d\). Loads and stores targeting addresses so tagged failstop
-if the program counter does not match. In this setting, at depth \(d\), all memory tagged
-with a depth less than \(d\) corresponds to \(\sealed\) memory, and is inaccessible.
-When its depth is reached via a return, it will have the same value, so integrity is
-enforced.
-
-On the other hand, Lazy Tagging and Clearing skips the (problematically linear time)
+frame is tagged with \(d\). Loads and stores targeting addresses in frames failstop
+if the program counter tag does not match the frame tag. In this setting, at depth \(d\),
+all memory tagged with a depth less than \(d\) corresponds to \(\sealed\) memory,
+and should be indeed be inaccessible.
+
+On the other hand, Lazy Tagging and Clearing skips the (inefficient)
 initialization and clearing steps, and only updates tags when memory is written to.
 It allows a callee to overwrite sealed memory, but that write tags the corrupted memory
-with the callee's identity. If the corrupted memory's owner (or any other activation) tries
-to read it, only then does the micro-policy failstop.
+with the callee's identity. Only if the corrupted memory's owner (or any other activation) tries
+to read it will the micro-policy failstop.
 
@@
 
-To characterize this situation, we define a state element to be \emph{irrelevant}
+To characterize this situation in a property, we define a state element to be \emph{irrelevant}
 if its value can be replaced with any other value without changing the observable behavior
@@
 a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
-agree on the value of every element except those in \(\components\).
-In the tagged environment, we additionally need the values in \(\components\)
-to agree on tags, but not necessarily their data.
+agree on the value of every element not in \(\components\).
+In the tagged environment, we additionally require that the values of
+\emph{all} elements in \(\mach\) and \(\nach\) agree on their tags. 
+\apt{Right?  This does not generalize so nicely, since we have not defined the
+  non-payload part of values, but it will do ok here, I think.}
 
@@
   \includegraphics[width=\columnwidth]{variants.png}
-  \caption{Integrity Violation}
+  \caption{Integrity Violation\apt{Tags are too small to read (also in next figures)}}
   \label{fig:variant}
@@
 all relevant sealed addresses in \(\mach\) are the same as after step 5.
-We once again require that \(\intProp\) hold on the state following the matching return.
-That state is reached by step 9. Although {\tt sensitive} has changed, it might be
-irrelevant. Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
+We once again require that \(\intProp\) hold on the state following the matching return, 
+which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
+Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
 choosing 43 (and keeping its tag, \(t_2'\), unchanged.) As execution continues after the return
 from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
-from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} is relevant,
+from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} \emph{is} relevant,
 so \(\intProp\) does not hold, and integrity has indeed been violated.
@@
 following the return, the caller is allowed to access its stack frame, so those addresses
-may influence future execution legitimately.
+may legitmately influence future execution.
 
@@
 While we have separated the definitions of integrity and confidentiality,
-in terms of tag-based enforcement, both Depth Isolation and Lazy Tagging and Clearing
-enforce them.
+in terms of tag-based enforcement, Depth Isolation and Lazy Tagging and Clearing
+both enforce both properties.
 
@@
 capability-based enforcement satisfies both. LSE corresponds to integrity, plus the
-confidentiality of the caller's frame, but neither the unallocated stack nor data left
+confidentiality of the caller's frame, but not of the unallocated stack or data left
 behind in it by previous callees.
@@
 all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
+\apt{Needs revision to discuss payloads/tags.}
 

### 2022-12-04 submission5/noexceptions.tex
Fixed tag text size

@@
 \usepackage{xcolor,listings}
+
+\usepackage{tikz}
 
@@
 simplistic, assuming a fixed number of threads each with its own dedicated processor.
-\apt{Is concurrency still alive?}
+\sna{There is no concurrency in this document. Attempting to build it in from the beginning
+  was rather complicated. The question is where to put that extra complexity? It's not difficult
+  from a technical standpoint, subject to the one-thread-per-processor constraint.}
 We model memory-safe stack objects, but not a heap. Regions outside of
@@
 
-\newcommand{\freebox}{\includegraphics[height=10px]{free.png}}
-\newcommand{\pubbox}{\includegraphics[height=10px]{public.png}}
-\newcommand{\objbox}{\includegraphics[height=10px]{object.png}}
-\newcommand{\sealbox}{\includegraphics[height=10px]{sealed.png}}
+\newcommand{\freebox}{\tikz \filldraw[fill=blue] (0,0) rectangle (10px,10px);}
+\newcommand{\pubbox}{\tikz \filldraw[fill=lightgray] (0,0) rectangle (10px,10px);}
+\newcommand{\objbox}{\tikz \filldraw[fill=yellow] (0,0) rectangle (10px,10px);}
+\newcommand{\sealbox}{\tikz \filldraw[fill=red] (0,0) rectangle (10px,10px);}
 
@@
 To characterize this situation in a property, we define a state element to be \emph{irrelevant}
-if its value can be replaced with any other value without changing the observable behavior
-of the machine. More precisely, for a set of elements \(\components\),
-a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
-agree on the value of every element not in \(\components\).
-In the tagged environment, we additionally require that the values of
-\emph{all} elements in \(\mach\) and \(\nach\) agree on their tags. 
-\apt{Right?  This does not generalize so nicely, since we have not defined the
-  non-payload part of values, but it will do ok here, I think.}
+if its value can be replaced with another value without changing the observable behavior
+of the machine. This replacement value must respect any enforcement-relevant constraints on
+values; in the tagged setting, we define two values as variants if they share the same tag
+component. Then, for a set of elements \(\components\),
+a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
+they agree on the value of every element not in \(\components\), and the values
+in all elements are variants.
 
 The elements of \(\components\) are \emph{irrelevant}
-in \(\mach\) if every variant state produces an equivalent observation trace
+in \(\mach\) if every \(\components\)-variant state produces an equivalent observation trace
 (see Section \ref{sec:events}); all other elements are \emph{relevant}.
@@
   \includegraphics[width=\columnwidth]{variants.png}
-  \caption{Integrity Violation\apt{Tags are too small to read (also in next figures)}}
+  \caption{Integrity Violation}
   \label{fig:variant}
@@
 %
-Now consider three elements: the address \(\SP - 4\) (blue), the address \(\SP + 12\) (yellow),
-and the register {\tt a0} (red). In the first case, the execution from the variant state
-has changed the value at \(\SP - 4\). But both return states agree on that value. Therefore,
-while the address was updated, it does not depend on a secret value.
-
-In the second case, the return states disagree on the value of \(\SP + 12\). But in neither
+Now consider three elements: the address \(\SP - 4\), the address \(\SP + 12\),
+and the register {\tt a0}. The execution from the right-hand state to its return
+has changed the value at \(\SP - 4\), but that value matches that of the
+left-hand return. Therefore, the change does not represent a leak.
+
+The return states disagree on the value of \(\SP + 12\). But in neither
 case has that value changed since the original variants. So the difference is inherited from
-the original variation, and does not depend on a secret value either. This matters because,
-following the return, the caller is allowed to access its stack frame, so those addresses
-may legitmately influence future execution.
+the original variation, and does not represent a leak either. We do not continue executing the
+variant state after return, so these values will not cause the caller to behave differently.
 
@@
 Therefore, it must depend on a secret (in fact, the variable {\tt secret}).
-Unless {\tt a0} happens to be irrelevant in the original trace, this is a violation of what
+Unless {\tt a0} happens to be irrelevant, this is a violation of what
 we term {\it return-time confidentiality}.
@@
 
+For a given system, assume an equivalence class \(\sim\) on values.
 Two states are variants with respect to a set of elements, \(\components\),
-if they agree on the value of every element not in \(\components\). Our
-notion of non-interference involves comparing the traces of such
-\(\components\)-variants. We use this to define sets of irrelevant elements.
+if they agree on the value of every element not in \(\components\) and
+all of their values are equivalent under \(\sim\).
+Our notion of non-interference involves comparing the traces of such
+\(\components\)-variants. We use this to define sets of irrelevant elements,
+and secrets in general.
 
@@
 written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
-\apt{Needs revision to discuss payloads/tags.}
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
+and for all \(\component\), \(\mach[\component] \sim \nach[\component]\).
 

### 2022-12-04 submission5/noexceptions.tex
Cited Skorstengaard

@@
   Our security semantics has similarities to the overlay semantics that have been proposed
-  to characterize stack safety in the past~\cite{??}, but, like the basic machine
+  to characterize stack safety in the past~\cite{SkorstengaardSTK}, but, like the basic machine
   semantics, it is not safe-by-construction. Rather, it gives us a means to \emph{define}

### 2022-12-04 submission5/noexceptions.tex
More detail at the end, and fixed formatting

@@
 once it has become \(\public\); only when {\tt x} is deallocated will its region
-be available to allocate privately.
-
+be available to allocate privately. We define a helper function, \(\mathit{range}\),
+to map an offset from a register and a size into a set of addresses.
+%
+\[\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach \triangleq
+\{\mach[\reg]+i | \mathit{off} \leq i < \mathit{off+sz}\}\]
+%
 \judgmentbr[AllocPriv]
-           {\(b = \mach[\SP] + \mathit{off}\)}
-           {\(V' = V \llbracket \addr \mapsto \sealed |
-             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+           {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+             \{\addr \mid V ~ \addr = \unsealed\}\)}
+           {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
            {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
@@
 \judgmentbr[AllocPub]
-           {\(b = \mach[\SP] + \mathit{off}\)}
-           {\(V' = V \llbracket \addr \mapsto \public |
-             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
+           {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+             \{\addr \mid V ~ \addr = \unsealed\}\)}
+           {\(V' = V \llbracket \addr \mapsto \public \mid \addr \in \components \rrbracket\)}
            {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
@@
 \judgmentbr[Dealloc]
-           {\(b = \mach[\SP] + \mathit{off}\)}
-           {\(V' = V \llbracket \addr \mapsto \unsealed |
-             b \leq a < b+\mathit{sz} \land V ~ \addr \in \{\object, \public\} \rrbracket\)}
+           {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+             \{\addr \mid V ~ \addr = \unsealed\}\)}
+           {\(V' = V \llbracket \addr \mapsto \unsealed \mid \addr \in \components \rrbracket\)}
            {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
@@
 %
-\[\mathit{passed} ~ \overline{sa} ~ \mach \triangleq
-\bigcup_{(\reg,b,o) \in \overline{sa}} \{\mach[\reg]+i | b \leq i < b+o\}\]
+\[\mathit{passed} ~ \overline{sa} ~ \mach = \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
+\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
 %
@@
 it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
-than \(\sealed\) to express this. Of course, some objects may be passed in place,
-easily modeled by including them in \(\overline{sa}\).
+than \(\sealed\) to express this. Of course, some objects may be passed as arguments
+in place, easily modeled by including them in \(\overline{sa}\). More sophisticated
+behaviors may require sequences of {\bf dealloc}, {\bf tailcall}, then {\bf alloc}.
 
@@
 
-We can model this as a pair of additional security-relevant operations: one which
-declares a register to contain a valid pointer, and one which transmits the provenance
-of a pointer from one element to another.
+We can model this as a trio of additional security-relevant operations: one which
+declares a register to contain a valid pointer, one which transmits the provenance
+of a pointer from one element to another, and one which clears the provenance
+(for instance, when a pointer is modified in place in a way that makes it invalid.)
 
@@
   & \mathbf{propagate} ~ \component_{src} ~ \component_{dst} \\
+  & \mathbf{clear} ~ \component \\
 \end{align*}
@@
 
-\judgmenttwobrlong[]
-                  {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},sz)\)}
-                  {\(b = \mach[\reg_{base}]+\mathit{off}\)}
-                  {\(\rho' = \rho[\reg \mapsto (b,b+sz)]\)}
+\judgmentbr[Promote]
+           {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})\)}
+           {\(\rho' = \rho[\reg_{dst} \mapsto \mathit{range} ~ \reg_{base} ~ \mathit{off} ~ \mathit{sz}]\)}
            {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
 
-\judgmenttwo[]
+\judgmentbr[Propagate]
            {\(\psi = \mathbf{propagate} ~ \component_{src} ~ \component_{dst}\)}
@@
            {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+
+\judgment[Clear]
+         {\(\psi = \mathbf{clear} ~ \component\)}
+         {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
 
@@
 this set \(\mathit{capped}\), and define it recursively:
-
+%
 \[\begin{split}
@@
 
+There are some advanced features that would be nice to work in. Any of these that we support should
+all be folded into the ``advanced machine,'' so that the full formalism is only presented once.
+They can still be explained progressively. First, concurrency: the coroutine properties from earlier
+drafts should translate just fine into this formalism, but the challenge is that concurrency introduces
+non-determinism. I think it's feasible to formalize. I do wonder what level of testing is appropriate;
+two threads switching at intervals and mostly minding their own business seems doable, but the interesting
+parts happen when they interact through memory.
+
+It would also be nice to have a story about exceptions. I've done some thinking about how to model them,
+and we'll want to choose one particular style (e.g., C++-scoping with or without stack unwinding).
+
 \bibliographystyle{IEEEtran}

### 2022-12-07 submission5/core.tex
fix typo

--- 2022-12-04 submission5/noexceptions.tex
@@
 A {\em machine state} \(\mach\) is a map from elements to a set \(\mathcal{V}\) of
-\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\) (\(|\cdot|\).
+\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
 The details of value structure depend on the specific machine being modelled;  
@@
 information relevant to a hardware enforcement mechanism (such as a tag).
-We write \(\mach[\component]\) to denote the value at
-\(\component\) of \(\mach\) and \(\mach[v]\) to denote the value at \(|v|\) if
-defined.
+We write \(\mach[\component]\) to denote the value of \(\mach\) at
+\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
 

### 2022-12-07 submission5/core.tex
Wordsmithing

@@
 
-To characterize this situation in a property, we define a state element to be \emph{irrelevant}
-if its value can be replaced with another value without changing the observable behavior
-of the machine. This replacement value must respect any enforcement-relevant constraints on
-values; in the tagged setting, we define two values as variants if they share the same tag
-component. Then, for a set of elements \(\components\),
+To characterize this situation in a property, we use the concept of {\it non-interference}
+based on variant states. Specifically, we consider two values to be variants, even if their
+payload values differ, as long as the remainder of their state respects other system-specific
+constraints; in the case of the tagged system, the tag component must be identical.
+Then, for a set of elements \(\components\),
 a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
 they agree on the value of every element not in \(\components\), and the values
-in all elements are variants.
-
-The elements of \(\components\) are \emph{irrelevant}
+in \(\components\) are variants.
+
+We say that the elements of \(\components\) are \emph{irrelevant}
 in \(\mach\) if every \(\components\)-variant state produces an equivalent observation trace

### 2022-12-07 submission5/core.tex
Variants in formalism

@@
 
-For a given system, assume an equivalence class \(\sim\) on values.
-Two states are variants with respect to a set of elements, \(\components\),
-if they agree on the value of every element not in \(\components\) and
-all of their values are equivalent under \(\sim\).
-Our notion of non-interference involves comparing the traces of such
-\(\components\)-variants. We use this to define sets of irrelevant elements,
-and secrets in general.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+\definition For a given system, assume an equivalence relation \(\sim\) on values.
+Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
 written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
-and for all \(\component\), \(\mach[\component] \sim \nach[\component]\).
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\),
+and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
 

### 2022-12-07 submission5/core.tex
wordsmith; add compatiblity terminology

@@
 
-To characterize this situation in a property, we use the concept of {\it non-interference}
-based on variant states. Specifically, we consider two values to be variants, even if their
-payload values differ, as long as the remainder of their state respects other system-specific
-constraints; in the case of the tagged system, the tag component must be identical.
-Then, for a set of elements \(\components\),
+To characterize this situation in a property, we borrow the idea of \emph{variant} states
+from the theory of non-interference. 
+For a set of elements \(\components\),
 a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
-they agree on the value of every element not in \(\components\), and the values
-in \(\components\) are variants.
+their values an only disagree on elements in \(\components\). We further require that
+the respective values in \(\components\) be \emph{compatible}, meaning that even though
+their payloads may differ, their tags (or whatever other data is important to
+the machine's security enforcement mechanism) must be identical.
 
 We say that the elements of \(\components\) are \emph{irrelevant}
-in \(\mach\) if every \(\components\)-variant state produces an equivalent observation trace
-(see Section \ref{sec:events}); all other elements are \emph{relevant}.
+in \(\mach\) if they can be replaced by compatible values without changing the
+observable behavior of the machine. (We will formalize this in Section~\ref{sec:events}
+by saying that every \(\components\)-variant state produces
+an equivalent observation trace.) All other elements are \emph{relevant}.
 We characterize \emph{observational integrity} as the property that
@@
 
-\definition For a given system, assume an equivalence relation \(\sim\) on values.
-Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+For a given system, assume an compatibility equivalence relation \(\sim\) on values.
+Two states are variants with respect to a set of elements \(\components\)
+if they agree on the value of every element not in \(\components\) and
+and have compatible values for every element in \(\components\).
+Our notion of non-interference involves comparing the traces of such
+\(\components\)-variants. We use this to define sets of irrelevant elements,
+and secrets in general.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
 written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\),
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
 and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).

### 2022-12-29 submission5/core.tex
Comments on the current draft, and some polishing here and there

@@
 
-``Stack safety'' is a term associated with a
+The term {\em stack safety} is associated with a
 variety of compiler, run-time, and hardware mechanisms for protecting stack
-memory. But unlike heap memory, the stack does not correspond to any single
-high-level language concept. The result is that
-these mechanisms typically lack precise specifications,
-relying instead on informal descriptions and examples of the bad
-behaviors that they prevent. Such a specification is especially desirable
-for an unusual enforcement mechanism, like the ``lazy'' stack safety
+memory. But, unlike ``the heap,'' the stack does not correspond to a single
+high-level language concept, and mechanisms for protecting it
+typically lack precise specifications,
+relying instead on informal descriptions and examples of bad
+behaviors that they prevent. A more rigorous specification is especially desirable
+for unusual enforcement mechanisms\bcp{why especially for this?}
+like the ``lazy'' stack safety
 micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
+% The term {\em Stack safety} is associated with a
+% variety of compiler, run-time, and hardware mechanisms for protecting stack
+% memory. But unlike heap memory, the stack does not correspond to any single
+% high-level language concept. The result is that
+% these mechanisms typically lack precise specifications,
+% relying instead on informal descriptions and examples of the bad
+% behaviors that they prevent. Such a specification is especially desirable
+% for an unusual enforcement mechanism, like the ``lazy'' stack safety
+% micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
 
@@
 state in each caller's stack frame is held invariant by the callee'')
-and confidentiality properties (``the caller's and callee's behavior are
-insensitive to one another's private state,''), both compatible with lazy tagging, as well as
-control-flow property.
+and\bcp{two?} confidentiality properties\bcp{plural?} (``the caller's
+and callee's behavior are insensitive to one another's private
+state,''), both compatible with lazy tagging\bcp{undefined term; also,
+the fact that we've mentioned tagged enforcement twice already makes
+it really feel like we are proposing an abstract theory of exactly one
+thing}, plus a control-flow property.
 
@@
 
-The call stack is a perennial target for low-level attacks, leading to a
-range of dire consequences, from leakage or corruption of private stack data
-to control-flow hijacking. To foil such attacks, a profusion of
+The call stack is a perennial target for low-level attacks, leading to
+dire consequences from leakage and corruption of private stack data
+to control-flow hijacking. To foil such attacks, a menagerie of
 software and hardware protections have been proposed,
@@
 
-The protections offered by such mechanisms are commonly described in terms
-of concrete examples of attacks that they can prevent---corruption of return
+The protections offered by such mechanisms are commonly described by giving
+concrete examples of attacks that they can prevent---corruption of return
 addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more abstract characterization to the reader's intuition.  But these
+a more generic characterization to the reader's intuition.  But these
 mechanisms can be quite intricate, and the behaviors they aim to prevent are
@@
 intuitions, whether a given mechanism actually blocks all potential
-attacks---or, conversely, whether it is overly conservative and disallows
+attacks in a given class---or, conversely, whether it is overly conservative and disallows
 system behaviors that are actually safe.
@@
 discuss in Section \ref{sec:relwork}, there is a dearth of such properties in
-the literature.
-
-We propose a novel characterization of stack safety using the tools of language-based
+the literature. \bcp{can we give a very quick preview here of the comparison?}
+
+We propose a novel characterization of stack safety using the formal tools of language-based
 security~\cite{sabelfeld2003language}, decomposing stack safety into
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-during the callee's execution, and the confidentiality of the calle's local state
-afterward, as well as the control flow protection of {\em well-bracketed control
-  flow (WBCF)}~\cite{SkorstengaardSTKJFP}. We extend WBCF to a setting with exceptions,
+during the callee's execution, the confidentiality of the callee's local state
+afterward, and the control flow protection of {\em well-bracketed control
+  flow (WBCF)}~\cite{SkorstengaardSTKJFP}. We also extend WBCF to a setting with exceptions,
 in which there may be multiple points at which a callee can ``return.''
@@
 are ``nested'': {\em each} caller is guaranteed protection from its
-immediate callees (which, in turn, need protection from their immediate
+immediate callees (which, in turn, are protected from their immediate
 callees, etc.).
-Caller confidentiality is especially interesting. It is based on a traditional
-notion of noninterference, but whereas ordinary noninterference
+Caller confidentiality is especially interesting: It is based on a traditional
+notion of noninterference, but, whereas ordinary noninterference
 is an end-to-end hyper-property on whole program runs, caller
 confidentiality is a nested form of noninterference that applies to
-part of a run, requiring that the callee’s behavior is
+part of a run, requiring that the callee’s behavior be
 invariant under hypothetical scrambling of the caller’s stack
@@
 
-Our threat model is very strong, allowing the attacker
+Our threat model is very strong: it allows the attacker
 to execute arbitrary code, including attempting to smash
 the stack to disrupt the program’s control flow. Our stack safety
-properties demand that, even in the presence of such attackers, confidentiality
+properties promise that, even in the presence of such attackers, confidentiality
 and integrity still apply.
@@
 
-For confidentiality and integrity, our properties are {\em observational},
+For confidentiality and integrity, our properties\bcp{by this point,
+  the reader may be getting confused about why we are talking about
+  multiple properties, rather than just one (with multiple clauses)} are {\em observational},
 meaning that they are only concerned with the observable behavior of the system.
@@
 affect the observable behavior of the
-system. This permissiveness enables our properties to handle lazy enforcement mechanisms.
+system. This permissiveness enables our properties to handle lazy
+enforcement mechanisms.  \bcp{do we have any in mind aside from Nick's
+tagging-based ones?}
 
@@
 tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly abort the ones that
+that Roessler and DeHon's micro-policies correctly detect the ones that
 attempt to violate one of our properties. Furthermore, we
@@
 that violate our properties but are \emph{not} halted by incorrect
-enforcement variants---both variants that we accidentally created
-during our re-implementation of the micro-policy and ones that we
-intentionally crafted to be broken in order to increase our confidence
-in testing and the enforcement mechanism itself.
+variants of the enforcement mechanisms, including ones that we accidentally created
+while implementating the micro-policy and ones that we
+intentionally crafted to increase our confidence in the effectiveness
+of our testing.
 
@@
 \begin{itemize}
-\item We formalize permissive stack confidentiality and integrity and integrity properties.
+\item We formalize permissive stack confidentiality and integrity properties.
   Both are parameterized over a notion of external
   observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in the system behavior %(\cref{sec:lazy}).
-\item We apply these definitions to a complex system with features such as argument passing
-  on the stack, callee-saves registers, exceptions, and a simple coroutine system% (\cref{sec:ext}).
+  data causes a visible change in system behavior %(\cref{sec:lazy}).
+\item We apply\bcp{instantiate?} these definitions to a complex\bcp{is
+  its complexity a selling point?} system
+  with features such as argument passing
+  on the stack, callee-saves registers, exceptions, and a simple
+  coroutine system% (\cref{sec:ext}).
+  \bcp{Are these features really independent of the generic model?  If
+  so, that's a nice point that should be brought out clearly.}
 \item We use property-based random testing to validate the relationship between
-  our properties and micro-policies %(\cref{sec:testing})
+  our properties and micro-policies. %(\cref{sec:testing})
+  \bcp{Micro-policies have not been explained, and we haven't
+    motivated why this would be an important or useful thing to do.}
 \end{itemize}
 %\Cref{sec:relwork,sec:future} discuss related and future work.
+\bcp{Why did the section references get taken away from all that?  And
+why delete the last sentence?}
 
@@
 as well as the testing implementation.
+\bcp{Make clear that there are no proofs!}
 
@@
   on the machine architecture and the nature of its built-in safety enforcement mechanisms
-  (if any).  Our machine model is meant to be a faithful description of the hardware,
+  (if any).  Our machine model is meant to be a faithful description of the\bcp{??} hardware,
   with a semantics that covers both ordinary instruction behavior and the behavior of
@@
   The security-relevant operations typically correspond to underling machine instructions,
-  but the correspondence may not be obvious just be inspecting the machine code.
+  but the correspondence may not be obvious just by inspecting the machine code.
   For example, in the tagged RISC-V machine we use in our examples and tests,
@@
 In the next section, we give a detailed example of a security semantics for a simple setting
-in which our security-relevant operations are restricted to making calls and returns, and allocating
+in which our security-relevant operations are restricted to calls,
+returns, and allocation of
 private memory within the current function activation. In later sections, we extend this model
-to the full system that we test, which features:
-\begin{itemize}
-\item Function calls and returns, with caller- and callee-saved registers
-\item Allocation of strictly private and strictly public regions on the stack
-\item Arguments spilled onto the stack and/or passed by reference
-\item Exceptions
-\item Tail-call Elimination
-\end{itemize}
+to the full system that we test, which features
+% \begin{itemize}
+% \item
+function calls and returns, with caller- and callee-saved registers,
+% \item
+allocation of strictly private and strictly public regions on the stack,
+% \item
+arguments spilled onto the stack and/or passed by reference,
+% \item
+exceptions,
+and
+% \item
+tail-call elimination.
+% \end{itemize}
 
@@
 objects are governed by a provenance-based capability system, accessible only to those
-functions which have received a valid pointer to them.
+functions which have received a valid pointer to them.  \bcp{???}
 
@@
 
-In general, it is impossible to distinguish buggy source code from an attacker; in
+In general, it is impossible to distinguish buggy machine code from an attacker; in
 our examples we will identify one function or another as an attacker, but we do not
@@
 
-This is a strong threat model, but hardware and timing attacks are out of scope,
+This is a strong threat model, but it does omit some important aspects of
+security enforcement: In particular, hardware and timing attacks are out of scope,
 and our properties are termination insensitive as a result of how failstops are
-modelled (see ~\ref{??}).
+modelled (see~\ref{??}).
 
@@
   was rather complicated. The question is where to put that extra complexity? It's not difficult
-  from a technical standpoint, subject to the one-thread-per-processor constraint.}
+  from a technical standpoint, subject to the one-thread-per-processor constraint.}\bcp{Appendix?}
 We model memory-safe stack objects, but not a heap. Regions outside of
@@
 
-A {\em machine state} \(\mach\) is a map from elements to a set \(\mathcal{V}\) of
+A {\em machine state} \(\mach\) is a map from state elements to a set \(\mathcal{V}\) of
 \emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
@@
 
-{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\), with a
+{\em Events} are drawn from a set \(\OBSS\)\bcp{lowercase?} and ranged over by \(\obs\), with a
 special silent event \(\tau\).
@@
 (from the point of view of an attacker) from traces that might eventually
-produce output, an assumption that gives our properties the flavor of
+produce output\bcp{why?}, an assumption that gives our properties the flavor of
 {\it termination insensitivity}.
 
-\subsection{Security-Relevant Operations, Principals, and Security State}
+\subsection{Security-Relevant Operations, Principals, and State}
 
@@
 Calls identify the address of the call's target,
-and the argument registers. Allocations and deallocations are phrased in terms of a pair
+plus the argument registers. Allocations and deallocations are phrased in terms of a pair
 of an offset from the stack pointer and a size. Returns don't require any additional information.
@@
 \begin{align*}
-  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} & \\
-  | & \mathbf{return} & \\
-  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
-  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+  \psi \in \Psi \ \ ::= \ \ & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} & \\
+  | \ & \mathbf{return} & \\
+  | \ & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
+  | \ & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N}
 \end{align*}
 %
-Security principals are function activations: at any given time, pending activations
+The security principals in our account of stack safety are function
+activations: at any given time, pending activations
 have security requirements that are recorded in a call stack, and the current
@@
 In this section we consider a simple setting with calls, returns, and private allocations,
-on a RISC-V machine with a PIPE tagging system. Here 
+on a RISC-V machine with a PIPE tagging system\bcp{undefined}. Here
 \(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
@@
 Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
-for a function {\tt main}. (Code is compiled using the standard RISC-V
+for a function {\tt main}. Code is compiled using the standard RISC-V
 calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
-{\tt main} begins at address 0 and its callee {\tt f} at address 100.)
+{\tt main} begins at address 0 and its callee {\tt f} at address 100.
 
@@
 
-Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
-to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
+Suppose that {\tt f} is actually an attacker seeking
+to leak {\tt secret}. It might do so in a number of ways, shown as snippets of
 assembly code in Figure \ref{fig:f}.\footnote{We arrange that all the snippets
@@
 with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
-(Figure \ref{subfig:integrity})!
-Attacks of this kind do not violate {\tt main}'s confidentiality, but its
-{\it integrity}.
+(Figure \ref{subfig:integrity}).
+Attacks of this kind do not violate {\tt main}'s confidentiality, but
+rather its {\it integrity}.
 In Figure \ref{subfig:WBCF}, the attacker arranges to return to the
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
-violating the program's {\it well-bracketed control flow} (WBCF.) Figure \ref{subfig:WBCF2}
-shows another attack that violates WBCF, this time by returning to the correct
-program counter but with the stack pointer offset.
+violating the program's {\it well-bracketed control flow} (WBCF.) In
+Figure \ref{subfig:WBCF2}, a different attack violates WBCF, this time
+by returning to the correct program counter but with the wrong stack pointer.
 
@@
 Figure \ref{subfig:WBCF2}. Here the attacker returns with \(\SP' = 998\) instead of the
-correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame 
+correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame,
+\begin{center}
 \begin{tabular}{| l | l | l | l | l |}
@@
 \end{tabular}
+\end{center}
 
@@
 read part of the return address, and its attempt to output
-{\tt res} will instead output {\tt secret}! Even absent other
-kinds of data protection, the stack pointer {\it must} be restored
-for the program to behave predictably.
+{\tt res} will instead output {\tt secret}!
+% Even absent other
+% kinds of data protection, the stack pointer {\it must} be restored
+% for the program to behave predictably.
 
@@
 To perform this enforcement using tags, we can take
-\(\mathcal{T} ::= \mathit{gen} | \mathit{bseq} ~ n ~ m | \mathit{prot}\), named for general
+\(\mathcal{T} ::= \mathit{gen} \mathrel{|} \mathit{bseq} ~ n ~ m \mathrel{|} \mathit{prot}\), named for general
 execution, blessed sequence, and protected, respectively. Blessed sequences of length \(m\) are
@@
 that {\tt main} never reads that corrupted data.
-To illustrate, we can consider two integrity micro-policies: ``Depth Isolation'' and
+To illustrate why this matters, let's consider two integrity
+micro-policies: ``Depth Isolation'' and
 ``Lazy Tagging and Clearing.''
@@
 Depth Isolation tracks the depth of the current activation via a tag on the program counter.
-Stack frames are initialized and cleared, by writing
+Stack frames are initialized and cleared by writing
 to every address in them, so that during a call at depth \(d\), the current activation's
@@
 
-In our example setting, the observation trace consists of the sequence of values written to {\tt out}.
+In our example setting, the observation trace consists of the sequence
+of values written to {\tt out}.\bcp{Is this the first place we've
+  discussed what the observations actually are?  Is this the only
+  thing they could be, or does the model support other options?}
 The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
@@
 callee confidentiality entirely. Georges et al. \cite{Georges22:TempsDesCerises} correct
-both limitations.
+both limitations.\bcp{Then why are we not comparing to them instead??}
 
@@
 
-\subsection{Facts Abouts Calls and Returns}
+\subsection{Facts About Calls and Returns}
 \label{sec:facts}
@@
 two threads switching at intervals and mostly minding their own business seems doable, but the interesting
-parts happen when they interact through memory.
+parts happen when they interact through memory. \bcp{This seems
+  infeasibly hard to address at this point.}
 

### 2022-12-29 submission5/main.tex
Comments on the current draft, and some polishing here and there

@@
 
-``Stack safety'' is a term associated with a
+The term {\em stack safety} is associated with a
 variety of compiler, run-time, and hardware mechanisms for protecting stack
-memory. But unlike heap memory, the stack does not correspond to any single
-high-level language concept. The result is that
-these mechanisms typically lack precise specifications,
-relying instead on informal descriptions and examples of the bad
-behaviors that they prevent. Such a specification is especially desirable
-for an unusual enforcement mechanism, like the ``lazy'' stack safety
+memory. But, unlike ``the heap,'' the stack does not correspond to a single
+high-level language concept, and mechanisms for protecting it
+typically lack precise specifications,
+relying instead on informal descriptions and examples of bad
+behaviors that they prevent. A more rigorous specification is especially desirable
+for unusual enforcement mechanisms\bcp{why especially for this?}
+like the ``lazy'' stack safety
 micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
@@
 
-\section{Key Ideas by Example}
-\label{sec:running-example}
-\colorlet{lgray}{gray!40}
-\colorlet{lred}{red!40}
-\colorlet{lblue}{blue!20}
-
-\newcommand{\mainsealc}{cyan}
-\newcommand{\fsealc}{green}
-\newcommand{\unsealc}{lgray}
-\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
-\newcommand{\fulloutc}{white}
-\newcommand{\badc}{lred}
-\newcommand{\goodc}{lblue}
-\newcommand{\retptrc}{black}
-\newcommand{\sealdesc}[1]{Seal(#1)}
-\newcommand{\unsealdesc}{Unseal}
-\newcommand{\retptrdesc}{RetPtr}
-\newcommand{\passdesc}[2]{Pass(#1,#2)}
+\section{Threat Model, Machines, and Enforcement }
+
+Before introducing our formal model of stack safety, we describe
+our machine exemplar, enforcement mechanism, and threat model.
+
+We begin with a RISC-V-like machine enhanced with PIPE, a tag-based reference
+monitory. We extend this machine with a {\it security overlay semantics}:
+a semantics in which calls, returns, and other operations of interest are condensed
+into atomic {\it macro-instructions}, which update an additional security context.
+Our properties are defined in terms of the security context at the end of an appropriate
+macro-instruction.
+
+The underlying machine is a standard RISC-V ISA as modeled in [MIT-RISCV].
+Our overlay semantics provide macro-instructions for the following code features:
+\begin{itemize}
+\item Function calls and returns, with caller- and callee-saved registers according
+  to the RISC-V ABI
+\item Arguments passed by reference
+\item Exceptions
+\item Coroutines and cooperative threading
+\item Tail-call Elimination
+\end{itemize}
+
+\paragraph*{What is a ``call''?} Some ISAs have a single ``call''
+instruction that does all the required manipulation of the program counter
+(\(\PCname\)) and stack pointer; others perform a series of instructions
+whose constituent opcodes may be used in some places for other purposes.
+Our RISC-V machine is the latter, so calls must be explicitly designated
+as such by the compiler, by identifying them as macro-instructions.
+Macro-instructions connect the enforcement mechanism to the property: we don't
+expect the enforcement mechanism to provide any guarantees for code sequences
+that resemble calls or returns, but which are not macro-instructions.
+
+\paragraph*{Enforcement Model and Properties}
+
+A machine state consists of the standard addresses and registers, as well as
+additional policy state used by the enforcement mechanism. Our properties
+refer only to addresses and registers; they do not constrain the behavior
+of policy state. We model an execution step as a partial function on states:
+when the machine does not step from a given state, this
+is a {\em failstop}, halting the program before it can perform an
+action that would violate the security property.
+our properties are therefore naturally \emph{termination insensitive.}
+
+\paragraph*{Threat Model}
+
+We trust the compiler has correctly placed macro-instructions according to
+the semantics of the source language, but we do not assume that the compiler
+is otherwise reasonable. In particular, while we are agnostic as to the source
+language, C is very plausible, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code.
+
+In general, it is impossible to distinguish buggy source code from an attacker;
+we assume nothing about the intent of the source code, except that it makes calls
+and returns, and guarantee that the caller and callee are protected from one
+another.
+
+This is quite a strong threat model, but hardware and timing attacks are out of scope,
+and our properties are termination insensitive as described above.
+
+\paragraph*{Limitations}
+
+We do not support dynamic code generation or concurrency---we can handle a
+limited form of coroutines, but not full-blown preemptive concurrency.
+
+\paragraph*{Heap Safety}
+
+\subsection{The Machine in Detail}
+\label{sec:prelim}
+
+The building blocks of the machine are {\em words} and {\em registers}.
+Words are 64-bit integers ranged over by \(w\) and, when used as addresses,
+\(a\). Registers are written \(r_i\) for \(0 \leq i < 32\), and generally
+ranged over by \(r\). The program counter \(pc\) is a special register,
+while the stack pointer is \(sp = r_1\), as in the RISC-V ABI.
+
+Collectively addresses and registers are {\em state elements} \(\component\)
+in the set \(\COMPONENTS ::= \WORDS + \REGS\).
+%
+A {\em machine state} \(\mach \in \COMPONENTS \rightarrow \WORDS \)
+is a mapping from state elements to values. For convenience, we abbreviate
+multiple updates \(m[\component_0 \mapsto w][\component_1 \mapsto w]\dots\)
+as \(m \llbracket \component \mapsto w | P ~ \component\rrbracket\), where
+\(P\) is some predicate that defines the set of elements to update.
+
+Separate from the machine state, we keep track of additional security context
+information over the course of the run. This data is not relevant to execution,
+it is only used to define the stack safety property. We define a set of
+{\it security classes}:
+\[sc \in SEC ::= \sealed | \passed | \mathit{public} | \unsealed\]
+And of {\it targets}:
+\[\begin{split}
+tar \in TAR ::= & \mathit{Return} ~ a_1 ~ a_2 \\
+| & \mathit{Catch} ~ a_1 ~ a_2 \\
+\end{split}\]
+A {\it security view} \(V \in \COMPONENTS \rightarrow SEC\) divides the machine
+elements in terms of their purpose relative to the active function.
+And a security context is a stack of pairs of targets and views:
+\[\context \in \CONTEXTS ::= \mathit{list} ~ (TAR \times SEC)\]
+
+The set of instructions, \(I\), contains the RISC-V ISA. There is additionally a set
+of macro-instructions \(J\), defined in detail in Figure \ref{fig:macros}.
+Every macro-instruction \(j \in J\) has a
+size \(\mathit{size}(j)\) and can be decomposed into a sequence of fragmanets
+\(j_0 \dots j_{\mathit{size}(J)}\), where \(j_n \in J^*\).
+The partial function \(\mathit{decode} : \WORDS \rightharpoonup I + J^*\) may decode a value
+into either an instruction or a macro-instruction fragment.
+
+\begin{figure}
+  \begin{tabular}{| l | l l |}
+    \hline
+    Macro & \multicolumn{2}{|l|}{Parameters} \\
+    \hline
+    \multirow{4}{*}{\(\mathbf{call}\)}
+    & \(\reg\) & Register containing destination \\
+    & \(args\) & Set of argument registers \\
+    & \(sarg\) & Set of stack object bounds \\
+    & \(pub\) & Set of public object bounds \\
+    \hline
+    \multirow{3}{*}{\(\mathbf{tailcall}\)}
+    & \(\reg\) & Register containing destination \\
+    & \(args\) & Set of argument registers \\
+    \hline
+    \multirow{2}{*}{\(\mathbf{enter}\)}
+    & \(\funid\) & Function identifier \\
+    & \(n\) & Size of locals \\
+    \hline
+    \(\mathbf{return}\)
+    & \(\overline{obj}\) & Public objects \\
+    \hline
+    \(\mathbf{yield}\) & & \\
+    \(\mathbf{setex}\) & & \\
+    \(\mathbf{throwex}\) & & \\
+    \hline
+  \end{tabular}
+  \caption{Macro-instructions}
+  \label{fig:macros}
+\end{figure}
+
+Execution is modeled by a partial step function between machine states,
+written \(\mach, \context \stepsto \mach', \context'\).
+If \(\mach[\PCname]\) decodes to an instruction, then
+\(\mach, \context \stepsto \mach', \context\) if \(\mach \stepsto_{\mathit{RISC-V}} \mach'\).
+If, for some \(j\), \(\mach[\PCname + i]\) decodes to \(j_i\) for all \(i\) between 0 and
+\(\mathit{size}(j)\), then the step is via a special multi-step. We present these steps
+as separate step relations: \(\cdot \harpoonunder{j}_M \cdot \in \MACHS \times J \times \MACHS\)
+to step the machine state,
+and \(\cdot, \cdot \stepstounder{j}_C \cdot \in \MACHS \times \CONTEXTS \times J \times \CONTEXTS\)
+for the security context. We then combine them:
+
+\judgmenttwo[]
+            {\(m \harpoonunder{j}_M m'\)}
+            {\(m',c \stepstounder{j}_C c'\)}
+            {\(m,c \stepsto m',c'\)}
+
+The machine multi-step for a call (\ref{fig:callmacro}) updates the stack pointer
+and program counter, and zeros non-argument, caller-save registers. (This means that
+the caller should have saved them before initiating the call.)
+
+For the context step, the macro-instruction additionally identifies which objects
+(contiguous regions defined relative to the stack pointer) have been placed there
+as: arguments that overflowed to the stack (\(sargs\)), which are existing objects
+passed by reference (\(pbrs\)), and which are marked public (\(pubs\)). We use this
+information to construct the callee's security view.
+\[obj \in OBJ ::= \mathbb{Z} \times \mathbb{Z}\]
+\[in\_obj ~ a_{sp} ~ (lo, hi) ~ a \triangleq a_{sp} - lo \leq a < a_{sp} - hi\]
 
 \begin{figure*}
-\begin{subfigure}[]{.16\textwidth}
-{\small
-  \begin{lstlisting}
-int main() {
-  int x = 42;
-  return f()+x;
-}
-
-int f() {
-  int y;
-  y = *(&y-2);
-  *(&y-2) = 0;
-  return 5;
-}
-  \end{lstlisting}
-}
-\end{subfigure}
-\begin{subfigure}[]{.44\textwidth}
-  {\small
-    \begin{lstlisting}[backgroundcolor=\color{lgray}]
-1 main: add $1,%sp      ;allocate frame
-2       mov $42,-1(%sp) ;initialize local
-3       call f
-4       add -1(%sp),%ra ;set return value
-5       sub $1,%sp      ;deallocate frame
-6       ret
-
-10 f:   add $1,%sp      ;allocate frame
-11      mov -4(%sp),-1(%sp) ;violation 1!
-12      mov $0,-4(%sp)      ;violation 2!
-13      mov $5,%ra      ;set return value
-14      sub $1,%sp      ;deallocate frame
-15      ret
-    \end{lstlisting}
-}
-\end{subfigure}
-\begin{subfigure}[]{.43\textwidth}
-\centering
-\addtolength{\tabcolsep}{-1em}
-\begin{tabular}{l | l}
-  \multicolumn{2}{c}{
-    2 \hspace{-1.5em}
-    \memoryaddrs{8em}
-    \memory{3}{\unsealc}
-    ~$\cdots$
-    \vspace{.5em}
-  } \\
-  11 \hspace{-1.5em}
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{42}
-  &
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\unsealc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
-  \\
-  12 \hspace{-1.5em}
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{42}
-  \MemoryLabel{-6em}{0.75em}{42}
-  \vspace{.5em}
-  &
-  \memoryaddrs{17em}
-  \memory{1}{\unsealc}
-  \memory{1}{\retptrc}
-  \memory{1}{\badc}
-  ~$\cdots$
-  \MemoryLabel{-16em}{0.75em}{\(v_0\)}
-  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
-  \vspace{.5em}
-\end{tabular}
-\addtolength{\tabcolsep}{1em}
-\end{subfigure}
-\caption{Integrity and confidentiality violations with confidentiality violation traces
-\ifaftersubmission\bcp{I wonder if we
-    could make the figures easier to read by putting very light, differently
-    colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
-\vspace*{-1em}
-\label{fig:conf1}
+\judgmenttwobrlong[]
+                  {\(m' = m[r_0 \mapsto m[pc] + 4]\)}
+                  {\(m'' = m' \llbracket r_i \mapsto 0 | i \geq 5 \land r \not \in args \land r \in caller \rrbracket\)}
+                  {\(j = \mathit{decode}^* ~ m ~ pc = \mathit{call} ~ f ~ args ~ sargs ~ pubs\)}
+                  {\(m \harpoonunder{j}_M m''[pc \mapsto f]\)}
+\[\begin{split}
+& \mathit{update\_view} ~ V ~ a_{sp} ~ args ~ sargs ~ pubs \triangleq
+ \lambda \component .
+\begin{cases}
+  \mathit{public} & \textnormal{if } \component \in \WORDS \textnormal{ and }
+  in\_obj ~ a_{sp} ~ pub ~ \component \textnormal{ for some } pub \textnormal{ in } pubs \\
+%  \passed & \textnormal{else if } \component \in \WORDS \textnormal{ and }
+%  in\_obj ~ a_{sp} ~ pbr ~ \component \textnormal{ for some } pbr \textnormal{ in } pbrs \\
+  V ~ \component & \textnormal{else if } \component \in \WORDS \textnormal{ and }
+  in\_obj ~ a_{sp} ~ sarg ~ \component \textnormal{ for some } sarg \textnormal{ in } sargs \\
+  \sealed & \textnormal{else if } \component \in \WORDS \textnormal{ and } \component < a_{sp} \\
+  \unsealed & \textnormal{else if } \component \in \REGS \textnormal{ and }
+  \component \textnormal{ is in } args \textnormal{ or is } \PCname \textnormal{ or } sp \\
+  \sealed & \textnormal{otherwise} \\
+\end{cases} \\
+\end{split}\]
+
+\begin{minipage}{.49\textwidth}
+\judgmentbr[]
+           {\(V' = \mathit{update\_view} ~ V ~ m[sp] ~ args ~ sargs ~ pubs\)}
+           {\(j = \mathit{call} ~ f ~ args ~ sargs ~ pubs\)}
+           {\(m, V::tars \stepstounder{j}_C ((m[pc] + 4, m[sp]), V')::V::tars\)}
+\end{minipage}
+\begin{minipage}{.49\textwidth}
+\judgmentbr[]
+           {\(V' = \mathit{update\_view} ~ (\lambda \component . \unsealed) ~ m[sp] ~ args ~ sargs ~ pubs\)}
+           {\(j = \mathit{call} ~ f ~ args ~ sargs ~ pubs\)}
+           {\(m, [] \stepstounder{j}_C ((m[pc] + 4, m[sp]),V')::[]\)}
+\end{minipage}
+
+\judgmenttwobrlong[]
+                  {\(m' = m[m[sp] \mapsto m[r_0]]\)}
+                  {\(m'' = m' \llbracket a \mapsto 0 | m[sp+1] \leq a < m[sp+n+1] \rrbracket\)}
+                  {\(\mathit{decode}^* ~ m ~ pc = \mathit{enter} ~ f ~ n\)}
+                  {\(m \harpoonunder{Enter}_M m''[sp \mapsto m[sp]+n+1]\)}
+
+\judgment[]
+         {}
+         {\(m,c \stepstounder{Enter}_M c\)}
+
+\judgmenttwobrlong[]
+                  {\(m' = m[sp \mapsto m[sp]-n-1]\)}
+                  {\(m'' = m' \llbracket r_i \mapsto 0 | 5 \leq r < 10 \lor 12 \leq r < 32 \rrbracket\)}
+                  {\(\mathit{decode}^* ~ m ~ pc = \mathit{return} ~ n\)}
+                  {\(m \harpoonunder{Return}_M m''[pc \mapsto m[sp]]\)}
+
+\begin{minipage}{.49\textwidth}
+\judgmentbr[]
+           {\(\)}
+           {\(\mathit{decode}^* ~ m ~ pc = \mathit{return} ~ f ~ args ~ sargs ~ pubs\)}
+           {\(m, V::tars \stepstounder{Return}_C ((m[pc] + 4, m[sp]), V')::tars\)}
+\end{minipage}
+\begin{minipage}{.49\textwidth}
+\judgmentbr[]
+           {\(V' = \mathit{update\_view} ~ (\lambda \component . \unsealed) ~ m[sp] ~ args ~ sargs ~ pubs\)}
+           {\(\mathit{decode}^* ~ m ~ pc = \mathit{call} ~ f ~ args ~ sargs ~ pubs\)}
+           {\(m, [] \stepstounder{Return}_C ((m[pc] + 4, m[sp]),V')::[]\)}
+\end{minipage}
+
+\caption{Call, enter, and return macro-instructions}
+\label{fig:callmacro}
 \end{figure*}
 
-\addtolength{\abovecaptionskip}{-1em}
-
-\begin{figure}
-\begin{subfigure}{.4\columnwidth}
-  {\small
-\begin{lstlisting}
-int main() {
-  int x;
-  x = f();
-  return g();
-}
-int f() {
-  int y = 5;
-  return y;
-}
-int g() {
-  int z;
-  return z;
-}
-\end{lstlisting}
-  }
-\vspace*{-1em}
-\caption*{Fig. \arabic{figure}: Leaking via\\ uninitialized data}
-\addtocounter{figure}{1}
-\vspace*{-1em}
-\label{fig:conf2}
-\end{subfigure}
-\begin{subfigure}{.5\columnwidth}
-  \begin{center}
-\begin{tabular}{l l}
-2 &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}[\unsealdesc]
-
-~$\cdots$
-\vspace{.5em}
-\\
-3 &
-\memoryaddrs{8em}
-\memory{3}{\unsealc}
-~$\cdots$
-    \MemoryLabel{-15em}{0.75em}{42}
-    \vspace{.5em}
-\\
-12 &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}[\sealdesc{0}]%
-\memory{1}{\retptrc}[\retptrdesc]%
-\memory{1}{\unsealc}[\unsealdesc]
-~$\cdots$
-\MemoryLabel{-16em}{0.75em}{42}
-\MemoryLabel{-7.5em}{0.75em}{42}
-\vspace{.5em}
-\\
-13 &
-\memoryaddrs{16em}
-\memory{1}{\mainsealc}
-\memory{1}{\retptrc}
-\memory{1}{\unsealc}
-~$\cdots$
-\MemoryLabel{-15em}{0.75em}{\bf 0}
-\vspace{.5em}
-\end{tabular}
-\end{center}
-\caption*{Fig. \arabic{figure}: Integrity Violation Trace}
-\label{fig:int1}
-\end{subfigure}
-\vspace*{1em}
-\vspace*{-2em}
-\end{figure}
-
-This section explains our formal stack safety properties through a series
-of examples, each presented both in C syntax and in corresponding assembly code for an
-imaginary two-address machine (e.g., see Figure~\ref{fig:conf1}).
-Each instruction's address is shown to its left.
-The assembly code reflects a simple compilation strategy in which
-local variables are stored in consecutive slots of the stack frame.
-For the moment, arguments and return values are passed in registers.
-The stack grows upward in
-memory and {\tt \%sp} points to the first unused word above the top of the stack.
-The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
-implicitly adjusting the stack pointer.
-At the right of each example, we show snapshots of the layout of stack memory just before
-selected instructions (labeled by address) are executed; each box represents a memory word
-(the machine is word addressed). Addresses increase to the right.
-We refer to registers and addresses collectively as {\em state elements} or just elements.
-  %% \bcp{Most of the
-  %% figure is kind of self-explanatory with this help, but the rightmost part
-  %% is a little opaque: what do 2, 3, 11, 12 mean?  What is SP?  What are the
-  %% numbers in the boxes?  What are these brackets under the boxes, ...  One
-  %% simple thing we could do to help might be to label the right-hand column
-  %% something like ``Snapshots,'' or at least explain early on in the text
-  %% that this is what they are.}
-
-Labels under the boxes indicate an association of the address to a
-{\em domain}---a set of intended access
-rights. Unless otherwise marked, each address retains its domain as
-instructions execute. %, as hinted by their colors.
-%\bcp{where do we see this happening?  Is it important?}.
-Domains are not part of the machine state; rather, they are part
-of the {\em context} of a state that describes the security restrictions
-that the enforcement mechanism must adhere to.
-The possible domains
-are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
-(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
-(outside the stack entirely).
-% These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
-Global variables and code, which are not shown in the figures, would
-also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
-are shaded black and ignored for simplicity for now---they will
-be treated properly later.
-(We will introduce some other parts of the context as needed.)
-
-Figure~\ref{fig:conf1} shows a program with simple integrity and confidentiality violations. Function {\tt main}
-calls function {\tt f}, which reads and then overwrites {\tt main}'s
-variable {\tt x}. (Of course, this is illegal C, but it is characteristic of what buggy or malicious code might do.)
-
-First, {\tt f} violates confidentiality by reading {\tt x}.
-Confidentiality states intuitively that a callee is insensitive to
-the context of its call (except for its arguments and any global
-variables).
-We state this as a form of {\em noninterference}~\cite{Goguen1982SecurityPA}.
-Noninterference describes knowledge: its standard form states that for state elements
-containing secret values, we can substitute any values whatsoever for the secrets
-without changing the effect of our computation on non-secret values.
-%
-If there exists some value for which we see different results, then
-in fact we can learn something about the initial state by observing those results.
-To adapt this to the stack, the notion of ``secret'' must vary with function
-calls and returns.
-
-At the right of Figure~\ref{fig:conf1}, we see how noninterference applies to this scenario.
-At instruction 10, we
-create an imaginary {\em variant}
-state with arbitrary values~%
-%\ifaftersubmission\bcp{IMO, it would be easier to
-%  understand if we chose concrete values for the examples}\sna{FWIW I agree in
-%  theory, but it looks bad every time I try it.}\fi
-(\(v_0\), etc.),
-in place of the actual values ($42$, ...) that should
-be hidden from {\tt f}; we show the evolving stack memory of the variant in the second column.
-At instruction 11, the callee's variable, {\tt z},
-changes to 42 in the primary trace, but
-in the variant trace it instead might have any value \(v_0\). If \(v_0 \neq 42\), then the variant
-has behaved differently than the primary, violating confidentiality.
-
-We call this form of confidentiality {\em stepwise confidentiality}. Informally,
-for any state at the entry to a function, and for any variant over the values in the stack,
-the primary state and the variant will step in lockstep until they
-return. To step in lockstep
-means that if both states step---that is, neither the primary nor the variant failstops---any
-state element that changes in the original will do the same in the variant.
-
-It may be unclear in Figure~\ref{fig:conf1} why we vary the initial value of {\tt y}
-as well as that of {\tt x}. This is to detect some violations of so-called
-{\em temporal stack safety}.
-Consider the code in Figure~2.
-Using the same compilation strategy as before, {\tt y} and {\tt z} will end up occupying the
-same memory location in the stack,
-so the value {\tt f} leaves behind in {\tt y}, which later becomes the value of {\tt x}, also
-becomes the value of {\tt z}. Thus, {\tt g} is able to learn the value of its caller's
-variable {\tt x}.  We regard this as a violation of confidentiality as well. In terms of noninterference,
-in the primary trace, {\tt z} will have the value 5, but in the variant it can have
-an arbitrary value.
-
-Returning to the code in Figure~\ref{fig:conf1}, line 12 violates integrity.
-In the initial memory layout the entire stack is {\em unsealed},
-meaning that it is eligible to be written to. At the call from {\tt main} to
-{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
-in this case 0. (Note that this does not include the return address, which is saved
-by the call instruction.)
-Sealing is the fundamental integrity contract between the caller and callee:
-{\tt main} expects that {\tt f} will not change {\tt x}.
-This leads to a simple, intuitive statement of {\em stepwise integrity}:
-if an element is sealed before a step, it will be unchanged afterward.
-Overwriting {\tt x} violates this property, as shown in Figure~3.
-
-\paragraph*{Control-Flow Attacks}
-\begin{figure*}
-  \centering
-\hspace*{-6em}
-  \begin{subfigure}[t]{.2\textwidth}
-{\footnotesize
-\begin{lstlisting}
-int *stash = 0;
-int main() {
-  int x = 1;
-  f();
-  x = -x;
-  f();
-  return x;
-}
-
-void f() {
-  int y;
-  if (stash == 0) {
-    stash = *(&y-1);
-  } else {
-    *(&y-1) = stash;
-    stash = 0;
-  }
-  return;
-}
-\end{lstlisting}
-}
-\end{subfigure}
-  \begin{subfigure}[t]{.4\textwidth}
-{\footnotesize
-\begin{lstlisting}[backgroundcolor=\color{lgray}]
-   stash: .word 0
-1  main:  add $1,%sp     ; alloc frame
-2         mov $1,-1(%sp) ; init x
-3         call f
-4         neg -1(%sp)    ; x = -x
-5         call f
-6         mov -1(%sp),%ra ; set ret val
-7         sub $1,%sp     ; dealloc frame
-6         ret
-
-10 f:     add $1,%sp     ; alloc frame
-11        cmp stash, $0  ; stash = 0?
-12        brneq #15      ; branch if not
-13        mov -2(%sp),stash ; set stash
-14        jmp #17
-15        mov stash,-2(%sp) ; get stash
-16        mov $0,stash
-17        sub $1,%sp     ; dealloc frame
-18        ret
-\end{lstlisting}
-}
-  \end{subfigure}
-%
-  \begin{subfigure}[t]{.2\textwidth}
-    \begin{center}
-    \begin{tabular}{l l l}
-      3 &
-      \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      {\tt 18} &
-      \memoryaddrs{12em}
-      \memory{1}{\mainsealc}[\sealdesc{0}]%
-      \memory{1}{\unsealc}[\retptrdesc]%
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1}
-      \MemoryLabel{-11em}{0.75em}{\#4}
-      \vspace{.5em} &
-      \hspace*{-1.5em}
-      \memory[1.2em]{1}{\mainsealc}[Target]%
-      \MemoryLabel{-2.2em}{0.75em}{\#4}
-      \\
-      4 &
-      \memoryaddrs{8em}
-      \memory{3}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{1} \\
-      11 &
-      \memoryaddrs{16em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#6} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      18 &
-      \memoryaddrs{12em}
-      \memory{1}{\mainsealc}
-      \memory{1}{\badc}
-      \memory{1}{\unsealc}%
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1}
-      \MemoryLabel{-11em}{0.75em}{\#4} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      4 &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{-1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6}
-      \\
-      5 &
-      \memoryaddrs{8em}
-      \memory{1}{\mainsealc}
-      \memory{2}{\unsealc}
-      ~$\cdots$
-      \MemoryLabel{-15em}{0.75em}{\bf 1} &
-      \hspace*{-1em}
-      \memory[1.2em]{1}{\mainsealc}
-      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
-    \end{tabular}
-    \end{center}
-    \vspace{\abovedisplayskip}
-  \end{subfigure}
-  \caption{A control-flow attack.}
-  \label{fig:controlflow}
-\end{figure*}
-
-%From the classic stack smashing attack to return oriented programming,
-One of the most pernicious ways to attack the stack is by hijacking control flow.
-Our stack safety properties do not assume that control flow follows a typical structure
-of matching sets of calls and returns nested within one another.
-Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
-Consider Figure~\ref{fig:controlflow}.
-(Here we treat return addresses properly, but write them as line numbers, in the
-form $\#n$.)
-Function {\tt f} is called twice, at instructions 3 and 5; it should therefore return to instructions
-4 and 6, respectively. But during the first call it stashes its
-return pointer, and the second time it returns to instruction 4 instead of 6; it thus ends
-up being called a third time before finally returning to instruction 6.
-
-Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that {\tt main} returns
-$-1$. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
-This seems like an integrity violation; how do we capture it in our
-\ifspace integrity \fi property?
-
-The key intuition is that, to be considered as having returned to {\tt main}, {\tt f} needs
-to do more than just execute a return-labeled instruction: it must actually reach the
-appropriate return target in {\tt main}.
-To capture this intuition, we add a new piece of context:
-a stack of return targets.  Now a callee is
-only considered to have returned when it reaches a state that matches the caller's target.
-A return target is associated with a set of sealed locations that will become unsealed
-when it is reached, represented here by coloring both the target and its sealed data with the
-same color (blue).
-
-In the example, {\tt f} reaches its return target on its first return, to instruction 4.
-But {\tt f}'s second return is to 4 rather than 6
-because the stack pointer has been overwritten (shown in light red).
-Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
-violates the integrity property.
-
-\paragraph*{Well-bracketed control flow} Together,
-integrity and confidentiality protect a caller's data until its callee
-returns to it. However, it is possible that the caller is not
-returned {\em to}, but rather returned {\em past}, to some ancestor caller!
-Strictly speaking, this is not an information-flow violation. But it is still
-arguably unsafe if the ancestor depends on the skipped function completing its execution before
-it returns.
-%
-Thus, another aspect of safe use of the stack is that calls
-and returns should be correctly nested---a callee should
-only return to its own caller.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
-call \emph{well-bracketed control flow (WBCF)}, and is orthogonal to
-confidentiality and integrity, so we treat is as an optional additional property.
-
-\section{Threat Model, Machines, and Enforcement }
-\label{sec:setup}
-\label{sec:threat}
-
-Before introducing our formal model of stack safety, we describe our assumptions:
-how we abstract over machines, how we represent enforcement, and
-our threat model.
-
-\paragraph*{Machine model}
-Our properties are defined in terms of an abstract model that makes few
-assumptions about the concrete machine that instantiates it. We assume a
-register machine with a program counter (\(\PCname\)) and well-defined
-mechanisms for determining the bounds of a stack frame and recognizing when a call
-has returned. Our examples consider machines with a stack pointer (\(\rsp\)),
-in which the current stack frame is everything below the stack pointer and above the
-previous frame, and where a call from instruction \(\addr_\PCname\) with
-stack pointer \(\addr_\rsp\)
-is considered to have returned when the \(\PCname\) is at \(\addr_\PCname + 4\) and the stack pointer
-is again at \(\addr_\rsp\). However, our properties apply equally to
-other architectural conventions, such as frame pointers.
-The non-stack memory can be used in any way, including
-as a heap, as long as the stack has static bounds. There can be
-state other than registers and memory, such as the internal state of a hardware
-enforcement mechanism.
-
-Our core property assumes that all register values passed to the callee are public,
-and that all data in a stack frame is private.
-Some calling convention choices, such as callee-saves registers and argument
-passing on the stack, are not supported directly by this property;
-in Section~\ref{sec:ext},
-we introduce
-secondary properties that use the same underlying formalism to handle these cases.
-
-\paragraph*{What is a ``call''?}  One significant challenge in formulating our
-properties arises from the fact that we want to define stack safety at the
-machine-code level for a variety of ISAs.  There may be no single ``call''
-instruction that does all the
-required manipulation of the \(\PCname\) and stack pointer. Instead, the
-same instruction opcode may be
-used in some places in a call or return sequence and in other places
-for other purposes.
-\ifspace
-For example, there
-might be just a simple branch instruction that stores the old \(\PCname\) in
-a register, with the rest of the call sequence performed by other
-instructions either before or after the jump.  Similarly, instead of a single
-``return'' instruction,  there may be a ``return sequence'' involving an
-indirect branch at some point.
-\fi
-
-Since our stack safety properties are all about protecting ``callers'' from
-their ``callees,'' we need some clear-cut way to recognize when the program
-makes or returns from a call.  Accordingly, we assume we are given a machine
-language program {\em annotated} with the locations of instructions
-representing calls.  We declare that a call has occurred when the machine
-has just executed an instruction marked as a call.
-%
-These call annotations would typically come from a compiler together with
-the machine code, but we do not assume anything about their provenance.
-Annotations connect the enforcement mechanism to the property: we don't
-protect unmarked calls, because a mechanism's correct functioning might
-depend on knowing where calls are.
-
-Returns, on the other hand, need not be annotated: we can tell when a callee
-has returned control to the caller when...we see control return to the
-caller! We don't require the callee's cooperation at all.
-\ifaftersubmission\apt{Rest of this is repetitious with machine model section above.}\fi
-For example, under a standard calling convention, we can see that a
-return has occurred when the state of the machine reaches a state where
-the stack pointer has the same value as it did at the point of some earlier
-call and the $\PCname$ points to the instruction following that call.  We
-abstract away from the details of specific calling conventions by assuming
-that we are given some way of telling when a given machine state is a
-{\em return target} with respect to some set of pending calls.
-
-\paragraph*{Enforcement Model}
-
-We model the machine's behavior as a partial step function over machine states.
-Machine states may contain arbitrary data, including data used exclusively by the
-enforcement mechanism, while our properties refer only to the data that are
-relevant to normal execution: addresses and registers.
-When the machine does not step from a given state, this
-is a {\em failstop}, halting the program before it can perform an
-action that would violate the security property;
-our properties are therefore naturally \emph{termination insensitive.}
-%
-That said, our model is meant to apply generically to any enforcement mechanism,
-including those that do not failstop, such as
-code rewriting techniques that protect stack safety directly by
-restricting the set of possible initial states.
-
-Most enforcement mechanisms make assumptions about their initial conditions,
-so we assume that any given machine will restrict its conditions a ``well-formedness''
-criterion on initial states, ensuring that they are set up properly. For eample, in the case of
-Roessler and DeHon's \cite{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-formedness
-requires that all annotated instructions will be situated at the appropriate instruction
-of a {\em blessed sequence}---a series of instructions that are tagged to enforce correct
-behavior. It can then failstop if it would
-execute the blessed sequence out of order.
-
-\paragraph*{Threat Model}
-
-We adopt a strong threat model in which both caller and callee may execute
-arbitrary machine code, subject to whatever ``well-formedness'' conditions the
-enforcement mechanism demands. But hardware and timing attacks are out of scope,
-and our properties are termination insensitive as described above.
-
-\paragraph*{Limitations}
-
-Our model is flexible enough to support many realistic uses of the stack.
-We intentionally say little or nothing about other structures
-such as the heap.  Heap safety has been studied in isolation
-by Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18}, and we expect that their correctness condition and ours can be
-combined cleanly to characterize correct protection of stack + heap.
-
-We do not currently support:
-address-taken locals, as a satisfactory treatment of these seems to require
-integration with a more general model of memory safety;
-tail calls;
-dynamic code generation, which is difficult to integrate with code annotations in general; or
-concurrency---we can handle a limited form of coroutines, but
-  not full-blown preemptive concurrency.
-
-\section{Machines and Traces}
-\label{sec:prelim}
-
-% We now define our machine and enforcement model.
-
-The building blocks of the machine are {\em values}, {\em addresses}, and {\em registers}.
-Values and addresses are drawn from some set of {\em words} \(\WORDS\), ranged over by \(\word\) and
-\(\addr\), respectively. Register names (\(\reg\)) drawn from some set
-\(\REGS\), which is assumed to include a special-purpose register, the
-program counter {\PCname}. Collectively addresses and registers are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \WORDS + \REGS\).
-%
-A {\em machine state} \(\mach\) is drawn from a set \(\MACHS\) and defines a mapping
-from state elements to values, written
-\(\mach[.] \in \COMPONENTS \rightarrow \WORDS \).
-
-We have a partial step function between machine states, written \(\mach \stepsto \mach'\), which
-is undefined on input states that correspond to a policy fault. So, a security mechanism is
-represented by the capacity of the machine to ``fail-stop,'' as well as by whatever initial
-well-formedness conditions are required for the mechanism to function properly.
-
-\paragraph*{Contexts}
-
-Separate from the machine state, we keep track of additional context
-information representing the history of the run. This has no influence on execution, it merely
-describes information relevant to the particular property. Different information is needed
-for different versions of the model, so the definitions are parameterized
-over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
-an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
-\(\mach, \context \constep \context' \in \MACHS \times \CONTEXTS
-\rightarrow \CONTEXTS\). We also assume the existence of a \emph{code map},
-\(\codemap \in \WORDS \rightharpoonup \ANNS\)
-that relates each code address to its annotation, if any.
-We pair a machine state with a context state to get a combined ``MC-state''
-\(\mpcstatename \in \MPCS = \MACHS \times \CONTEXTS\), and lift the context-step
-function into the machine step:
-
-\judgmenttwo{\(\mach \stepsto \mach'\)}
-              {\(\mach, \context \constep \context'\)}
-              {\(\mach,\context \stepstocon
-                \mach',\context'\)}
+%\judgmenttwobrlong[Return]
+%                    {\(m' = m[sp \mapsto m[sp] - \mathit{size}(frsize)]\)}
+%                    {\(m'' = m \llbracket r \mapsto 0 | r \not \in args \land r \in caller \rrbracket\)}
+%                    {\(\mathit{decode}^* ~ m ~ pc = \mathit{return} ~ frsize\)}
+%                    {\(m \stepsto_M m'[pc \mapsto f]\)}
+
+If \(\mathit{decode}\) is undefined on that value, \(\mach \stepsto \mach\). The
+only times \(\stepsto\) is undefined is when there is a policy fault,
+(and the machine ``failstops'') or when it attempts to begin executing a
+macro-instruction out of order. The latter represents a failstop in a lower-level
+policy in the underlying ISA, which guarantees that the macro-instruction
+executes in order.
 

### 2023-01-05 submission5/core.tex
Added that paragraph about the factorization

@@
 in which there may be multiple points at which a callee can ``return.''
+
+This approach has several benefits: first, not all state-safety enforcement mechanisms
+aim to protect all aspects of stack safety. Most protect control flow, but relatively
+few care about the confidentiality of the callee after return. Second, our formalism
+factors stack safety into three questions: (1) ``when does this machine call, return, etc.?''
+(2) ``during any given call, which elements of the state are in need of protection?'' and
+(3) ``when an element needs protection, how can we tell that it is protected?''
+Different machines, including those with different enforcement mechanisms, may answer question
+(1) differently, without needing new definitions for (2) or (3). In order to adjust our notion
+of permissible behavior to correspond to features such as pass-by-reference behavior, which
+alter its security requirements (typically by relaxing them), we need only add nuance to (2).
+Regardless of the assumptions that we make for either (1) or (2), the security definitions that
+answer (3) need not change.
 

### 2023-01-05 submission5/core.tex
Tweaking the connection to lazy tagged enforcement, as we discuessed in the meeting Tuesday

--- 2022-12-29 submission5/core.tex
@@
 addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more generic characterization to the reader's intuition.  But these
-mechanisms can be quite intricate, and the behaviors they aim to prevent are
-subtle and varied. It can be hard to guess, based only on informal
-intuitions, whether a given mechanism actually blocks all potential
-attacks in a given class---or, conversely, whether it is overly conservative and disallows
-system behaviors that are actually safe.
-To settle such questions with confidence,
-we need a precise, generic, and formal specification for stack
-safety, as a basis both for comparing the security claims of different
-enforcement techniques and for validating that these claims
-are met by particular implementations. With the notable exception of the model used
-in Skorstengaard et al.'s StkTokens work~\cite{SkorstengaardSTKJFP}, which we
-discuss in Section \ref{sec:relwork}, there is a dearth of such properties in
-the literature. \bcp{can we give a very quick preview here of the comparison?}
+a more formal characterization to the reader's intuition. Previous attempts
+to produce a formal characterization do so using a {\it fully abstract overlay semantics},
+a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
+The latest stack-safety overlay semantics \cite{Georges22:TempsDesCerises} capture
+both control-flow and data protection (confidentiality and integrity).
+But, this approach is not suitable for {\em lazy} enforcement mechanisms like those
+of Roessler and DeHon, in which stores are not checked, and illegal stores are
+only caught when their location is later read. We aim to provide a stack-safety
+property suitable for testing lazy enforcement.
 
@@
 in which there may be multiple points at which a callee can ``return.''
+
+Stack safety is challenging to specify in full generality, because the stack is fundamentally
+an architecture-specific construct. Its desired behavior varies based on ISA, calling convention,
+and the features of whichever programming language originated the code that must be protected.
+Our approach centers around a tagged architecture, like PIPE \cite{???} or STAR \cite{???},
+particularly using a lazy micro-policy. However, our design is factored into three pieces:
+(1) the definition of {\em security relevant operations} in a given machine,
+(2) the definition of a {\em security context} which tracks the security needs of different
+elements of the state over the course of execution, and
+(3) the definition, for a given security context, of how we distinguish secure behavior
+from insecure. This separation of concerns allows (3) to be relatively general, while
+architecture-specifics are isolated in (1) and some special features require extending
+(2).
 
@@
 
+\sna{Version that is less tied to lazy enforcement. It's still rather tied, actually,
+  more than I remembered.}
+
+One or more elements of a given state are {\it safe} if their values
+in that state have no observable influence on future execution. We define this
+formally as a sort of non-interference: for a state \(\sigma\) and
+a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
+holds if we can replace the contents of \(\components\) with arbitrary values
+and have the resulting state produce an equivalent trace. The state created
+by modifying \(\components\) in this way is termed a variant.}
+
 \begin{figure}

### 2023-01-17 submission5/core.tex
In the processing of rearranging the paper

@@
 
-\usepackage{xcolor,listings}
+\usepackage[table]{xcolor}
+
+\usepackage{listings}
 
@@
 memory. But, unlike ``the heap,'' the stack does not correspond to a single
-high-level language concept, and mechanisms for protecting it
-typically lack precise specifications,
-relying instead on informal descriptions and examples of bad
-behaviors that they prevent. A more rigorous specification is especially desirable
-for unusual enforcement mechanisms\bcp{why especially for this?}
-like the ``lazy'' stack safety
-micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
-% The term {\em Stack safety} is associated with a
-% variety of compiler, run-time, and hardware mechanisms for protecting stack
-% memory. But unlike heap memory, the stack does not correspond to any single
-% high-level language concept. The result is that
-% these mechanisms typically lack precise specifications,
-% relying instead on informal descriptions and examples of the bad
-% behaviors that they prevent. Such a specification is especially desirable
-% for an unusual enforcement mechanism, like the ``lazy'' stack safety
-% micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
+high-level language concept, making a formal specification elusive.
+Protection mechanisms typically lack precise specifications, relying instead
+on informal descriptions and examples of bad behaviors that they prevent.
+
+When the prevented behaviors are straightforward, describing them may be sufficient
+to justify an enforcement mechanism's effectiveness. Less intuitive enforcement
+mechanisms instead leave us to ask: what security do we actually hope to gain,
+and are we successful?
+
+In this paper we focus on one such non-intuitive enforcement mechanism,
+the ``lazy'' stack safety micro-policies proposed by
+Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}. These permit functions
+to write into one another's frames, but taint the result so that the frame's
+owner cannot access it. No existent characterization of stack safety
+captures this style of safety.
 
@@
 combination of an integrity property (``the private
-state in each caller's stack frame is held invariant by the callee'')
-and\bcp{two?} confidentiality properties\bcp{plural?} (``the caller's
+state in each caller's stack frame is held invariant by the callee''),
+two confidentiality properties (``the caller's
 and callee's behavior are insensitive to one another's private
-state,''), both compatible with lazy tagging\bcp{undefined term; also,
-the fact that we've mentioned tagged enforcement twice already makes
-it really feel like we are proposing an abstract theory of exactly one
-thing}, plus a control-flow property.
+state,''), and a control flow property. By defining the properties
+in terms of the observable behavior of the system, we expand the
+space of supported enforcement mechanisms to include those that
+allow certain (intuitively incorrect) behaviors but prevent them
+from causing harm.
 
 The stack governs a large number of language features that are often excluded
-from discussion of stack safety. Our stack safety property targets a system with
-both caller- and callee-saved registers, coroutines, arguments passed on the stack,
-exceptions, [tail call elimination, heaps, and concurrency.]
+from discussion of stack safety. Our properties support a system with
+both caller- and callee-saved registers, arguments passed on the stack,
+exceptions, [tail call elimination, and concurrency.] They are modular by
+design, in order to be further extensible to other features and architectures.
 
@@
 a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
-The latest stack-safety overlay semantics \cite{Georges22:TempsDesCerises} capture
-both control-flow and data protection (confidentiality and integrity).
-But, this approach is not suitable for {\em lazy} enforcement mechanisms like those
+But this approach is not suitable for the {\em lazy} enforcement mechanism
 of Roessler and DeHon, in which stores are not checked, and illegal stores are
@@
 in which there may be multiple points at which a callee can ``return.''
-
-Stack safety is challenging to specify in full generality, because the stack is fundamentally
-an architecture-specific construct. Its desired behavior varies based on ISA, calling convention,
-and the features of whichever programming language originated the code that must be protected.
-Our approach centers around a tagged architecture, like PIPE \cite{???} or STAR \cite{???},
-particularly using a lazy micro-policy. However, our design is factored into three pieces:
-(1) the definition of {\em security relevant operations} in a given machine,
-(2) the definition of a {\em security context} which tracks the security needs of different
-elements of the state over the course of execution, and
-(3) the definition, for a given security context, of how we distinguish secure behavior
-from insecure. This separation of concerns allows (3) to be relatively general, while
-architecture-specifics are isolated in (1) and some special features require extending
-(2).
-
-A key technical novelty in these definitions,
-compared to standard formulations of confidentiality and integrity from the
-security literature, is that they
-are ``nested'': {\em each} caller is guaranteed protection from its
-immediate callees (which, in turn, are protected from their immediate
-callees, etc.).
-Caller confidentiality is especially interesting: It is based on a traditional
-notion of noninterference, but, whereas ordinary noninterference
-is an end-to-end hyper-property on whole program runs, caller
-confidentiality is a nested form of noninterference that applies to
-part of a run, requiring that the callee’s behavior be
-invariant under hypothetical scrambling of the caller’s stack
-but allowing the caller to access its own data upon return.
-
-Our threat model is very strong: it allows the attacker
-to execute arbitrary code, including attempting to smash
-the stack to disrupt the program’s control flow. Our stack safety
-properties promise that, even in the presence of such attackers, confidentiality
-and integrity still apply.
-%
-
-For confidentiality and integrity, our properties\bcp{by this point,
-  the reader may be getting confused about why we are talking about
-  multiple properties, rather than just one (with multiple clauses)} are {\em observational},
-meaning that they are only concerned with the observable behavior of the system.
-A function activation might read from and write to other activations' data
-without violating an observational property, as long as such reads and writes do not
-affect the observable behavior of the
-system. This permissiveness enables our properties to handle lazy
-enforcement mechanisms.  \bcp{do we have any in mind aside from Nick's
-tagging-based ones?}
 
@@
 attempt to violate one of our properties. Furthermore, we
-%
 check that the testing framework is able to generate counterexamples
@@
 
-We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
-which memory within each stack frame is tagged with the depth of
-the function activation that owns the frame and access is
-permitted only when an activation at that depth is currently executing, validates our
-stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
+We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
 violates the temporal aspect of confidentiality in
@@
 
+\subsection{Motivation}
+
+Security mechanisms designed to prevent a certain class of attacks can be brittle---successfully
+eliminating the attack as it exists at the time of publication, but leaving room for attackers
+to find new, similar attacks. To mitigate this risk, mechanisms should aim to enforce formal
+properties that can be proven or tested. Such properties become the specification against which
+enforcement can be validated; even enforcement mechanisms that do not fulfill the property
+benefit from the ability to articulate {\it why}.
+
+Stack safety is challenging to specify in full generality, because the stack is fundamentally
+an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
+which is broadly consistent across architectures and therefore amenable to a very general approach.)
+The desired behavior of the stack varies based on ISA, calling convention,
+and the features of whichever programming language originated the code.
+For instance, the current state-of-the-art stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable. Subtle tweaks to the definition
+accompany any addition to the model.
+
+We are not attempting to create a universal definition of stack safety for all systems.
+Rather, we define the stack safety of a single, fairly realistic system, decomposed into
+(1) a lightweight model of the system and its security-relevant features, from which we
+derive the integrity, confidentiality, and control-flow requirements of each call and
+return, (2) the application of that model to a particular target system, and (3)
+the formal criteria for integrity, confidentiality, and control flow
+to obey the requirements.
+
+We treat stack safety as multiple properties, rather than one multi-faceted property, because
+not all enforcement mechanisms attempt to implement all of them. Some focus exclusively on
+control flow, others protect the confidentiality of the caller but not the callee, etc.
+
+\paragraph*{Permissiveness, Laziness, and Tagged Architectures}
+
+Security properties tend to be conservative, identifying violations that would be
+harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
+and of compatibility with existing codebases. In the motivating case for this work,
+Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
+using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
+STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
+A set of software-defined rules, termed a micro-policy, determines the tags for operation
+results based on the tags of their inputs---or, if the input tags are in a dangerous
+combination, it may halt the machine (a ``failstop.'')
+
+Roessler and DeHon find that the obvious way to enforce stack safety in this setting
+is inefficient when allocating and clearing stack frames, because each word must be initialized
+with a tag corresponding to the (depth of the) current function activation. They propose
+{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
+when a function writes to memory. This means that such writes cannot be protected. Instead
+the micro-policy enforces that memory be read by the same fuction that wrote it. This
+lazy enforcement is incompatible with existing models of stack safety, particularly the
+integrity side, because an attacker can modify another function's data. But the owner of
+the data will never see the changes, so it may still be secure in some sense. We aim to elaborate
+that sense, and to test whether the micro-policy fulfills it.
+This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
+machine states that make claims about the future observable behavior of execution from those states.
+
+\subsection{Contributions}
+
 In sum, we offer the following contributions:
@@
   observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in system behavior %(\cref{sec:lazy}).
-\item We apply\bcp{instantiate?} these definitions to a complex\bcp{is
-  its complexity a selling point?} system
-  with features such as argument passing
-  on the stack, callee-saves registers, exceptions, and a simple
-  coroutine system% (\cref{sec:ext}).
-  \bcp{Are these features really independent of the generic model?  If
-  so, that's a nice point that should be brought out clearly.}
+  data causes a visible change in system behavior.
+\item We instantiate these definitions with a realistic
+  system with features such as argument passing
+  on the stack, callee-saves registers, and tail-call elimination.
+  Our model is modular enough that adding these features is straightforward.
+\item Our formalization of well-bracketed control flow in the presence of tail-call
+  elimination is a slight novelty over existing models.
 \item We use property-based random testing to validate the relationship between
-  our properties and micro-policies. %(\cref{sec:testing})
-  \bcp{Micro-policies have not been explained, and we haven't
-    motivated why this would be an important or useful thing to do.}
+  our properties and micro-policies.
 \end{itemize}
-%\Cref{sec:relwork,sec:future} discuss related and future work.
-\bcp{Why did the section references get taken away from all that?  And
-why delete the last sentence?}
 
 Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as the testing implementation.
-\bcp{Make clear that there are no proofs!}
-
-\section{Machine Model and Security Semantics}
-
-  We state our properties in terms of a generic machine, making minimal assumptions
-  on the machine architecture and the nature of its built-in safety enforcement mechanisms
-  (if any).  Our machine model is meant to be a faithful description of the\bcp{??} hardware,
-  with a semantics that covers both ordinary instruction behavior and the behavior of
-  any tags, capabilities, etc. that the machine might support.
-  This semantics is intentionally \emph{not} ``safe-by-construction;'' it gives
-  no guarantee that the enforcement hardware is being used correctly to achieve desired
-  security properties.
-
-  On top of this machine model, we layer a {\it security semantics}, which extends the
-  underlying machine with additional context about the identity of current and pending
+as well as their testing implementation. There are no proofs---we use primarily Coq for the
+QuickChick library and to ensure that the formalizations are unambiguous.
+
+\section{Key Ideas}
+
+The paper is structured as follows. In \cref{sec:example}, we will walk through a function
+call in a simple example machine and discuss how each of our properties applies to it,
+informally. Then in \cref{sec:machine} we will introduce the formal machine model and give
+its {\em security semantics}, described below. In \cref{sec:props} we give abbreviated
+formal statements of the properties, while the full formalism is found in \cref{app:props}.
+\Cref{sec:ext} gives the security semantics for the full, tested machine, and
+\cref{sec:testing} describes the testing framework used to test it.
+\Cref{sec:relwork,sec:future} discuss related and future work.
+The remainder of this section introduces ideas that will be important throughout.
+
+\paragraph*{Security Semantics}
+
+A {\it security semantics} extends a machine
+with additional context about the identity of current and pending
 functions (which act as security principals) and about the registers and memory they require
@@
 and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
-  instructions might also be used for other things. We therefore assume that there
-  is some external mechanism for distinguishing which instructions are security-relevant;
-  the most obvious technique is to add optional labels to instructions, and that is
-  what we do in this paper, but the theory does not depend on this choice.
-
-  Our security semantics has similarities to the overlay semantics that have been proposed
-  to characterize stack safety in the past~\cite{SkorstengaardSTK}, but, like the basic machine
-  semantics, it is not safe-by-construction. Rather, it gives us a means to \emph{define}
-  the security properties that a correct enforcement scheme should implement, using
-  some combination of hardware and software.
-
-In the next section, we give a detailed example of a security semantics for a simple setting
-in which our security-relevant operations are restricted to calls,
-returns, and allocation of
-private memory within the current function activation. In later sections, we extend this model
-to the full system that we test, which features
-% \begin{itemize}
-% \item
-function calls and returns, with caller- and callee-saved registers,
-% \item
-allocation of strictly private and strictly public regions on the stack,
-% \item
-arguments spilled onto the stack and/or passed by reference,
-% \item
-exceptions,
-and
-% \item
-tail-call elimination.
-% \end{itemize}
-
-Finally, we give a property definition for stack safety in which (some) stack-allocated
-objects are governed by a provenance-based capability system, accessible only to those
-functions which have received a valid pointer to them.  \bcp{???}
-
-%In section (TBD) we discuss the extension of this model into a simple concurrency setup.
-
-\paragraph*{Threat Model}
-
-We must trust that our method of distinguishing security-relevant operations is accurate; if it
-involves labels placed on code by a compiler, that means trusting that the compiler placed
-those labels correctly. If operations occur that are not recognized, those operations
-might not be guaranteed to protect their principals---for instance, an unlabeled call
-might not protect the caller's data from the callee. On the flip side, applying an incorrect
-label most likely means that the property becomes too strong to be enforced.
-
-Otherwise, we do not assume that the code adheres to any particular
-calling convention or implements a source language construct.
-In particular, while we are agnostic as to the source
-language, we certainly aim to support C, and so any source function might contain undefined
-behavior resulting in its compilation to arbitrary machine code. A given enforcement
-mechanism may place additional constraints, particularly on the behavior of
-call and return sequences.
-
-In general, it is impossible to distinguish buggy machine code from an attacker; in
-our examples we will identify one function or another as an attacker, but we do not
-require any static division between trusted and untrusted code, and we aim to protect
-even buggy code.
-
-This is a strong threat model, but it does omit some important aspects of
-security enforcement: In particular, hardware and timing attacks are out of scope,
-and our properties are termination insensitive as a result of how failstops are
-modelled (see~\ref{??}).
-
-\paragraph*{Limitations}
-
-Our concurrency model is fairly
-simplistic, assuming a fixed number of threads each with its own dedicated processor.
-\sna{There is no concurrency in this document. Attempting to build it in from the beginning
-  was rather complicated. The question is where to put that extra complexity? It's not difficult
-  from a technical standpoint, subject to the one-thread-per-processor constraint.}\bcp{Appendix?}
-We model memory-safe stack objects, but not a heap. Regions outside of
-stacks can be used however the compiler likes, including as a heap, but no protection is
-built in and our properties assume that if a pointer to a stack object is stored there,
-it is permanently compromised.
-
-\subsection{The Basic Machine}
-
-The building blocks of the machine are {\em words} and {\em registers}.
-Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
-and are drawn from the set \(\WORDS\).
-Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
-given the special name \(\SP\);
-some registers may be classified as caller-saved or callee-saved.
-%
-Along with the program counter, \(\PCname\), these are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).
-
-\begin{figure}
-  \begin{tabular}{| l | l | l |}
+instructions might also be used for other things. An instantiation of an operation
+might require some external mechanism for distinguishing which instructions are
+security-relevant; the most obvious technique is to add optional labels to instructions,
+but the theory does not depend on this choice.
+
+The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
+the following table. Those that we explore in detail in Section \ref{sec:examples}
+are shown in white: calls, returns, and the allocation and deallocation of local memory.
+In the full formalism (Appendix \ref{app:full}) and the machine that we test,
+we cover additional operations add parameters to our existing ones, to support
+various forms of memory sharing and tail-call elimination. These are shown in light gray.
+We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
+based on pointer provenance in Appendix \ref{app:ptr}, represented by the dark grey
+operations, but we do not test this.
+
+\newcommand{\example}{\rowcolor{black!0}}
+\newcommand{\testing}{\rowcolor{black!10}}
+\newcommand{\theory}{\rowcolor{black!25}}
+
+\vspace{\abovedisplayskip}
+
+\begin{tabular}{| l | l |}
   \hline
-    Set / & Names & Purpose \\
-    Class & & \\
+  \(\psi \in \Psi\) & Parameters \\
   \hline
-    \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
-    \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
-    & {\tt a2} -- {\tt a7} & Caller-saved args \\
-    & {\tt ra} & Return Address \\
+  \example \(\mathbf{call}\) & Target address \\
+  \example & Argument registers \\
+  \testing & Stack argument offset \& size \\
+  \example \(\mathbf{return}\) & \\
+  \example \(\mathbf{alloc}\) & Offset \& size \\
+  \testing & Public flag \\
+  \example \(\mathbf{dealloc}\) & Offset \& size \\
+  \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
   \hline
-    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
-    \(\sealed\) & {\tt sp} & Stack Pointer \\
+  \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
   \hline
-    \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
-    \(\public\) & {\tt tp} & Thread Pointer \\
+  \theory \(\mathbf{promote}\) & Register, offset \& size \\
+  \theory \(\mathbf{propagate}\) & Source register/address \\
+  \theory & Destination register/address \\
+  \theory \(\mathbf{clear}\) & Target register/address \\
   \hline
 \end{tabular}
-  \caption{RISC-V integer register set}
-  \label{fig:RISCVregs}
-\end{figure}
-
-A {\em machine state} \(\mach\) is a map from state elements to a set \(\mathcal{V}\) of
-\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
-The details of value structure depend on the specific machine being modelled;
-intuitively, the payload represents the part of the value that is relevant to
-the behavior of the basic machine, while the rest of the value may contain
-information relevant to a hardware enforcement mechanism (such as a tag).
-We write \(\mach[\component]\) to denote the value of \(\mach\) at
-\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
-
-{\em Events} are drawn from a set \(\OBSS\)\bcp{lowercase?} and ranged over by \(\obs\), with a
-special silent event \(\tau\).
-The machine step function
-\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
-\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
-is labeled by an ordered list of security-relevant operations, defined below,
-and an event.
-
-While some enforcement mechanisms may cause a machine to
-failstop, we model this as the machine silently diverging by remaining in the
-same state. We consider silently diverging traces to be indistinguishable
-(from the point of view of an attacker) from traces that might eventually
-produce output\bcp{why?}, an assumption that gives our properties the flavor of
-{\it termination insensitivity}.
-
-\subsection{Security-Relevant Operations, Principals, and State}
-
-Security-relevant operations consist of calls, returns, and allocations.
-Calls identify the address of the call's target,
-plus the argument registers. Allocations and deallocations are phrased in terms of a pair
-of an offset from the stack pointer and a size. Returns don't require any additional information.
-%
-\begin{align*}
-  \psi \in \Psi \ \ ::= \ \ & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} & \\
-  | \ & \mathbf{return} & \\
-  | \ & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
-  | \ & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N}
-\end{align*}
-%
+
+\vspace{\belowdisplayskip}
+
+Our security semantics has similarities to the overlay semantics that have been proposed
+to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
+the behavior of the underlying machine in any way. Rather, it tracks additional context
+about the history of security-relevant operations, which inform the criterion a machine
+must satisfy in order to correctly obey the properties.
+
+\paragraph*{Security Classes}
+
 The security principals in our account of stack safety are function
@@
 all other locations to \(\public\), and registers based on which set they
-belong to: % (Figure \ref{fig:RISCVregs}):
-\(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
-%Elements of \(\mathcal{P}\),
-%such as the program counter, are always \(\public\). Even if they might deserve their
-%own protection, that protection is orthogonal to stack safety.
+belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
 
@@
 principals), ranged over by \(\sigma\).
-%\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
-The initial context is \(\context_0 = (V_0, \emplist)\).
-
-As the machine takes a step that executes a particular operation, it updates the
-corresponding context, as described informally in the following section, and formally
-in Section \ref{sec:props}.
+
+\paragraph*{Variant States}
+
+\paragraph*{Events and Traces}
+
+It is common to specify the behavior of a system, including security properties, in
+terms of traces of observable events rather than details of the machine state.
+Lazy enforcement works because the deferred check catches errors before they appear
+in the trace. The nature of traces varies based on the properties of interest, but
+given a particular model of observation, we can consider
+corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.
+
+\paragraph*{Threat Model}
+
+We must trust that our method of distinguishing security-relevant operations is accurate; if it
+involves labels placed on code by a compiler, that means trusting that the compiler placed
+those labels correctly. If operations occur that are not recognized, those operations
+might not be guaranteed to protect their principals---for instance, an unlabeled call
+might not protect the caller's data from the callee. On the flip side, applying an incorrect
+label most likely means that the property becomes too strong to be enforced.
+
+Otherwise, we do not assume that the code adheres to any particular
+calling convention or implements a source language construct.
+In particular, while we are agnostic as to the source
+language, we certainly aim to support C, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code. A given enforcement
+mechanism may place additional constraints, particularly on the behavior of
+call and return sequences.
+
+In general, it is impossible to distinguish buggy machine code from an attacker; in
+our examples we will identify one function or another as an attacker, but we do not
+require any static division between trusted and untrusted code, and we aim to protect
+even buggy code.
+
+This is a strong threat model, but it does omit some important aspects of
+security enforcement: In particular, hardware and timing attacks are out of scope,
+and our properties are termination insensitive as a result of how failstops are
+modelled (see~\ref{??}).
+
+\paragraph*{Limitations}
+
+%Our concurrency model is fairly
+%simplistic, assuming a fixed number of threads each with its own dedicated processor.
+We model memory-safe stack objects, but not a heap. Regions outside of
+stacks can be used however the compiler likes, including as a heap, but no protection is
+built in and our properties assume that if a pointer to a stack object is stored there,
+it is permanently compromised.
 
@@
 \section{Properties by Example}
+\label{sec:example}
 
 In this section we consider a simple setting with calls, returns, and private allocations,
-on a RISC-V machine with a PIPE tagging system\bcp{undefined}. Here
-\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
-for some type \(\mathcal{T}\) of tags to be elaborated later, with \(|(w,t)| = w\).
+on a RISC-V machine with a PIPE tagging system. In this system a value is a pair of a
+word (the ``payload'') and a tag drawn from some set \(\mathcal{T}\).
+%\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
+%for some type  of tags to be elaborated later, with \(|(w,t)| = w\).
 Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
@@
 While it would be simple to define integrity as ``all sealed elements retain their
-values after the call,'' this would be stricter than necessary for many purposes.
-For example, it arguably doesn't matter if {\tt f} overwrites {\tt main}'s data provided
-that {\tt main} never reads that corrupted data.
-To illustrate why this matters, let's consider two integrity
-micro-policies: ``Depth Isolation'' and
-``Lazy Tagging and Clearing.''
-
-Depth Isolation tracks the depth of the current activation via a tag on the program counter.
-Stack frames are initialized and cleared by writing
-to every address in them, so that during a call at depth \(d\), the current activation's
-frame is tagged with \(d\). Loads and stores targeting addresses in frames failstop
-if the program counter tag does not match the frame tag. In this setting, at depth \(d\),
-all memory tagged with a depth less than \(d\) corresponds to \(\sealed\) memory,
-and should be indeed be inaccessible.
-
-On the other hand, Lazy Tagging and Clearing skips the (inefficient)
-initialization and clearing steps, and only updates tags when memory is written to.
-It allows a callee to overwrite sealed memory, but that write tags the corrupted memory
-with the callee's identity. Only if the corrupted memory's owner (or any other activation) tries
-to read it will the micro-policy failstop.
-
-Lazy Tagging and Clearing illustrates the value of ``extensional'' properties.
-It is common to specify the behavior of a system, including security properties, in
-terms of traces of observable events rather than details of the machine state.
-Lazy enforcement works because the deferred check catches errors before they appear
-in the trace. The nature of traces varies based on the properties of interest, but
-given a particular model of observation, we can consider
-corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.
+values after the call,'' this would be stricter than necessary. Suppose that
+a callee overwrites some data of its caller, but the caller never accesses that data.
+By sheer luck, the callee has managed not to impact the observable execution beyond that one write.
+Luck, or enforcement: Lazy Tagging and Clearing, from Roessler and Dehon, works by
+failstopping any time a caller accesses callee-tainted data, so that out-of-frame
+writes should never impact execution.
 
@@
 
-\sna{Version that is less tied to lazy enforcement. It's still rather tied, actually,
-  more than I remembered.}
-
-One or more elements of a given state are {\it safe} if their values
-in that state have no observable influence on future execution. We define this
-formally as a sort of non-interference: for a state \(\sigma\) and
-a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
-holds if we can replace the contents of \(\components\) with arbitrary values
-and have the resulting state produce an equivalent trace. The state created
-by modifying \(\components\) in this way is termed a variant.}
-
 \begin{figure}
@@
 In our example setting, the observation trace consists of the sequence
-of values written to {\tt out}.\bcp{Is this the first place we've
-  discussed what the observations actually are?  Is this the only
-  thing they could be, or does the model support other options?}
+of values written to {\tt out}.
 The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
@@
 
-\section{Formalizing Properties}
+\section{Machine Model and Security Semantics}
+
+\subsection{The Basic Machine}
+
+The building blocks of the machine are {\em words} and {\em registers}.
+Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
+and are drawn from the set \(\WORDS\).
+Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
+given the special name \(\SP\);
+some registers may be classified as caller-saved or callee-saved.
+%
+Along with the program counter, \(\PCname\), these are {\em state elements} \(\component\)
+in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).
+
+\begin{figure}
+  \begin{tabular}{| l | l | l |}
+    \hline
+    Set / & Names & Purpose \\
+    Class & & \\
+    \hline
+    \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
+    \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
+    & {\tt a2} -- {\tt a7} & Caller-saved args \\
+    & {\tt ra} & Return Address \\
+    \hline
+    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
+    \(\sealed\) & {\tt sp} & Stack Pointer \\
+    \hline
+    \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
+    \(\public\) & {\tt tp} & Thread Pointer \\
+    \hline
+  \end{tabular}
+  \caption{RISC-V integer register set}
+  \label{fig:RISCVregs}
+\end{figure}
+
+A {\em machine state} \(\mach\) is a map from state elements to a set \(\mathcal{V}\) of
+\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
+The details of value structure depend on the specific machine being modelled;
+intuitively, the payload represents the part of the value that is relevant to
+the behavior of the basic machine, while the rest of the value may contain
+information relevant to a hardware enforcement mechanism (such as a tag).
+We write \(\mach[\component]\) to denote the value of \(\mach\) at
+\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
+
+{\em Events} are drawn from a set \(\OBSS\)\bcp{lowercase?} and ranged over by \(\obs\), with a
+special silent event \(\tau\).
+The machine step function
+\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
+\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
+is labeled by an ordered list of security-relevant operations, defined below,
+and an event.
+
+While some enforcement mechanisms may cause a machine to
+failstop, we model this as the machine silently diverging by remaining in the
+same state. We consider silently diverging traces to be indistinguishable
+(from the point of view of an attacker) from traces that might eventually
+produce output\bcp{why?}, an assumption that gives our properties the flavor of
+{\it termination insensitivity}.
+
+\subsection{Security-Relevant Operations, Principals, and State}
+
+Security-relevant operations consist of calls, returns, and allocations.
+Calls identify the address of the call's target,
+plus the argument registers. Allocations and deallocations are phrased in terms of a pair
+of an offset from the stack pointer and a size. Returns don't require any additional information.
+%
+
+%\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
+The initial context is \(\context_0 = (V_0, \emplist)\).
 
@@
 
-\subsection{Events and Traces}
-\label{sec:events}
-
-We abstract over the events that can be observed in the system, defining them
-only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
-event. Other events might represent certain function calls (i.e., system calls)
-or writes to special addresses representing mmapped regions.
-A {\em trace} is a nonempty, finite or infinite sequence
-of events, ranged over by \(\obsT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
-for list-cons.
-
-We write that execution from a state produces an observation trace
-\(\mach,\context \hookrightarrow \obsT\) as follows, coinductively:
-
-\judgmenttwo{\((\mach,\context) \stepstounder{} (\mach',\context',\obs)\)}
-            {\((\mach',\context') \hookrightarrow \obsT\)}
-            {\((\mach,\context) \hookrightarrow \notfinished{\obs}{\obsT}\)}
-
-We define another relation that takes a trace until we have returned from the
-active principal.
-We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
-\(d\) is the depth of the current call.
-
-\judgment{\(|\sigma| < d\)}
-         {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \tau\)}
-
-\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
-                  {\(|\sigma| \geq d\)}
-                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
-                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{\obs}{\obsT}\)}
-
-\paragraph*{Observational Similarity}
-
-We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
-written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
-is the same. That is, we compare up to deletion of \(\tau\) events.
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{}{\(\obsT \eqsim \obsT\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
-\end{minipage}
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
-\end{minipage}
-
-\paragraph*{Variants and Irrelevant Values}
-
-For a given system, assume an compatibility equivalence relation \(\sim\) on values.
-Two states are variants with respect to a set of elements \(\components\)
-if they agree on the value of every element not in \(\components\) and
-and have compatible values for every element in \(\components\).
-Our notion of non-interference involves comparing the traces of such
-\(\components\)-variants. We use this to define sets of irrelevant elements,
-and secrets in general.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
-and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
-
-\definition An element set \(\components\) in state \((\mach,\context)\) contains
-irrelevant values, written \((\mach,\context) \parallel \components\), if for all
-\(\nach\) such that \(\mach \approx_{\components} \nach\), if
-\((\mach,\context) \hookrightarrow \obsT\) and
-\((\nach,\context) \hookrightarrow \obsT'\), then
-\(\obsT \eqsim \obsT'\).
-
-\subsection{Facts About Calls and Returns}
-\label{sec:facts}
-
-Here we define some logical operations to reason about the behavior of the
-system over time. These have a temporal-logic flavor, as they reflect
-the expected behavior of the system in the future, after a possible return.
-
-\paragraph*{On-return}
-
-The intuition behind integrity (below) is that a caller may expect its
-sealed data to be unchanged when control returns to it. In fact, the callee
-may overwrite such data---when the data are found in callee-saved registers
-this is perfectly legal---as long as it either restores it, or has some guarantee
-that its changes will not impact the caller.
-
-We start by defining a second-order logical operator
-\(d \uparrow P\), pronounced ``\(P\) holds on return from depth \(d\),''
-where \(P\) is a predicate on machine states. This is a coinductive relation
-similar to ``weak until'' in temporal logic---it also holds if the program never
-returns from depth \(d\).
-
-\judgmenttwo[Returned]
-            {\(|\sigma| < d\)}
-            {\(P ~ (\mach,(V,\sigma))\)}
-            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
-
-\judgmenttwobrlong[Step]
-                  {\(|\sigma| \geq d\)}
-                  {\((d \uparrow P) ~ (\mach', \context')\)}
-                  {\((\mach, (V,\sigma)) \stepstounder{\overline{\psi}} (\mach', \context',\obs)\)}
-                  {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
-
-Similarly, for confidentiality, we will want to compare pairs of future states,
-so we give a binary equivalent, \(d \Uparrow R\), where
-\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
-first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\),
-respectively. Once again, \(\Uparrow\) is coinductive.
-
-\judgmenttwobrlong[Returned]
-            {\(|\sigma_1| < d\)}
-            {\(|\sigma_2| < d\)}
-            {\((\mach_1,(V_1,\sigma_1)) ~ R ~ (\mach_2,(V_2,\sigma_2))\)}
-            {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-
-\judgmenttwobrlong[Left]
-              {\(|\sigma_1| \geq d\)}
-              {\((\mach_1,(V_1,\sigma_1)) \stepstounder{\overline{\psi}} (\mach_1',\context_1',\obs)\)}
-              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-
-\judgmenttwobrlong[Right]
-              {\(|\sigma_2| \geq d\)}
-              {\((\mach_2,(V_2,\sigma_2)) \stepstounder{\overline{\psi}} (\mach_2',\context_2',\obs)\)}
-              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
-              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-
-\subsection{Properties}
-\label{sec:props}
-
-\paragraph*{Well-bracketed Control Flow}
-
-\definition We define a correct return from \(\mach\) as:
-%
-\[\ret(\mach) = \{(\mach',\context') \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}\]
-
-\definition
-A system enjoys  well-bracketed control flow if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) where
-\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\),
-\(|\sigma| \uparrow \ret(\mach)\) holds on \((\mach',(V,\sigma))\).
-
-\paragraph*{Integrity}
-
-\definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
-such that \(\mach[\component] \not = \mach'[\component]\).
-
-\definition We define \(\intProp(\mach,V)\), to hold on a state \((\mach',\context')\) when
-\(\{\component \mid V ~ \component = \sealed\} \cap \Delta(\mach,\mach')\)
-is irrelevant in \((\mach', \context')\).
-
-\definition A system enjoys {\it stack integrity} if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
-\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
-we have:
-\[(|\sigma'| \uparrow \intProp(\mach',V)) ~ (\mach',(V,\sigma))\]
-
-\paragraph*{Caller Confidentiality}
-
-\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
-\(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
-Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to \(\components\)
-if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
-\(|\sigma| \uparrow (\mach,\context) \hookrightarrow \obsT\) and
-\(|\sigma| \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
-\(\obsT \simeq \obsT'\).
-
-\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
-is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
-This treats
-\(\mach\) and \(\nach\) as the initial states and
-\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
-set of elements that were corrupted between them.
-
-\definition We define the relation \(\conf(\mach,\nach)\) to hold on a pair of states
-\((\mach',\context)\) and \((\nach',\context')\) when
-\(\bar{\Diamond}(\mach,\nach,\mach',\nach')\) is irrelevant in \((\mach',\context)\).
-
-\definition A state \((\mach,(V,\sigma))\) enjoys {\it return-time confidentiality}
-if, for \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\)
-and any \(\nach\) that is a \(\components\)-variant of \(\mach\), we have:
-\[(\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \conf(\mach,\nach) ~ (\nach,(V,\sigma))\]
-
-\definition A system enjoys {\it caller confidentiality} if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-\((\mach',\context')\) enjoys both internal and return-time confidentiality.
-
-\paragraph*{Callee Confidentiality}
-
-\definition We define \(\cconf(\mach,V)\), to hold on a state \((\mach',\context')\) when
-\(\{\component \mid V ~ \component = \unsealed\} \cap \Delta(\mach,\mach')\)
-is irrelevant in \((\mach', \context')\).
-
-\definition A system enjoys {\it callee confidentiality} if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-% or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-we have:
-\[(|\sigma| \uparrow \cconf(\mach',V)) ~ (\mach',\context')\]
+\section{Formalizing Properties}
 
@@
 operation.
-
-\begin{align*}
-  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
-  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
-  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
-  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
-  | & \mathbf{return} & \\
-  | & \mathbf{alloc} ~ \mathit{pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N}, \mathit{pub} \in \mathbb{B} \\
-  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
-\end{align*}
 
@@
 
-\begin{align*}
-  \psi \in \Psi ::= & \dots \\
-  & \mathbf{promote} ~ \reg ~ sa & sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
-  & \mathbf{propagate} ~ \component_{src} ~ \component_{dst} \\
-  & \mathbf{clear} ~ \component \\
-\end{align*}
-
 In addition to the normal call stack, our security context will carry a map \(\rho\) from
@@
 
-%\section{Enforcement}
-%\label{sec:enforcement}
-
-%We use our properties to test the stack safety micro-policies of Roessler and DeHon,
-%particularly their Lazy Tagging and Clearing. Our underlying machine is a RISC-V
-%based on the MIT RISC-V formalism \cite{} and extended with {\it metadata tags},
-%represented as additional state in \(\mathcal{P}\), one for each register, address,
-%and the program counter:
-%\[\mathcal{P} ::= t_\reg | t_\addr | t_\PCname\]
-%
-%We divide values into normal memory words and tags drawn from a policy-specific
-%set of tags \(\mathcal{T}\), \(\mathcal{V} ::= \mathcal{W} | \mathcal{T}\).
-%The machine's step function corresponds to that of the basic RISC-V machine,
-%with the behavior of the tags at each step parameterized by a {\it micro-policy},
-%which may turn a normal step into an error (represented by returning to the same state.)
-
-%\paragraph*{Protecting Enforcement State}
-
-%State beyond registers and memory is always considered \(\public\), and in a PIPE machine this
-%includes tags. That does not mean that functions should be modifying them! It just means that we
-%do not consider them to be part of the state being protected. It is less obvious that a function
-%should be prevented reading a tag, and our properties do not rule it out, although in practice
-%PIPE does not allow such reads. It would be quite possible to implement
-%the properties so that they would forbid these activities, for instance by structuring \(\mathcal{P}\)
-%to treat tags separately and apply stack safety to the tags on the stack, but there is no obvious
-%application for such a stringent property---the enforcement state must be protected in order that
-%enforcement not be subverted, not for its own sake.
-
-\section*{[Meta] Remainder}
-
-These document sections lay out a way to present our stack safety properties to build intuition and
-then introduce the formalism. Next the full paper will need the testing section and an expanded
-enforcement section, but these can largely be borrowed from previous submissions.
-
-There are some advanced features that would be nice to work in. Any of these that we support should
-all be folded into the ``advanced machine,'' so that the full formalism is only presented once.
-They can still be explained progressively. First, concurrency: the coroutine properties from earlier
-drafts should translate just fine into this formalism, but the challenge is that concurrency introduces
-non-determinism. I think it's feasible to formalize. I do wonder what level of testing is appropriate;
-two threads switching at intervals and mostly minding their own business seems doable, but the interesting
-parts happen when they interact through memory. \bcp{This seems
-  infeasibly hard to address at this point.}
-
-It would also be nice to have a story about exceptions. I've done some thinking about how to model them,
-and we'll want to choose one particular style (e.g., C++-scoping with or without stack unwinding).
+\section{Validation through Random Testing}
+\label{sec:testing}
+
+There are several ways to evaluate whether an enforcement mechanism enforces
+stack safety properties. Ideally such validation would be done through formal proof over
+the semantics of the enforcement-augmented machine.
+However, while there are no fundamental barriers to producing such a proof,
+it would be considerable work to carry out for a full ISA like RISC-V and
+complex enforcement mechanism like the Depth Isolation micro-policy.
+We therefore choose to validate the micro-policy of \cref{sec:enforcement} by
+systematically \emph{testing} that it satisfies our properties.
+This focus on testing is better aligned with our immediate
+goal of evaluating real enforcement mechanisms for real machine architectures.
+Formal proof remains future work.
+
+We validate the micro-policy of the previous section by
+systematically testing that it satisfies eager stack integrity and
+confidentiality. We use a Coq specification of the RISC-V
+architecture~\cite{Bourgeat2021AMF},
+and extend it with a runtime monitor implementing a stack safety
+micro-policy. We chose the Coq proof assistant as the setting for our implementation
+to ensure that our coinductive trace definitions are well-formed, to reason about
+them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.
+
+To use QuickChick, we build random test-case generators that produce
+% \begin{enumerate}
+% \item
+(1) an initial RISC-V machine state, including most notably
+  the program to be executed;
+  % \item
+(2)
+  an initial policy state, tagging instructions corresponding to
+  blessed call or return sequences appropriately, while marking
+  all potential stack locations as $\tagNoDepth$; and
+% \item
+(3)
+ a code map identifying the locations of calls.
+% \end{enumerate}
+
+To write such generators we build on the work of
+Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
+introduced {\em generation by execution} to produce progams that lead
+machines towards interesting behaviors. We extend the technique by tracking
+control flow information so that programs will usually---but not always---return to their call sites.
+
+%\paragraph*{Generators}
+%
+% Generation by
+%execution receives as an input a partially instantiated machine state
+%and attempts to generate an instruction (or a sequence of instructions
+%such as a blessed sequence) that makes sense locally (e.g., jumps go
+%to a potentially valid code location and loads read from a
+%potentially valid stack location). Then we step the machine and repeat
+%the process until we generate or execute some target number of
+%instructions, or reach a point where the machine cannot step
+%any more.
+%
+%We extend this technique to keep track of the control flow behavior of
+%the program being generated: each time a call or return sequence is
+%generated, we ensure that the appropriate policy tags and code
+%annotations are set for the entry or return points.% At the same time,
+%%we allow the generation to sometimes relax those constraints,
+%%introducing potentially ill-formed flows: this causes our programs
+%%to failstop when executed in conjunction with the policy monitor,
+%%but also allows for revealing errors in our setup.\apt{??}
+%
+We need to extend typical testing schemes further to handle the nested
+nature of confidentiality: rather than just generating two
+initial machines that are variants of one another and letting them
+execute to test for noninterference, we generate a new variant
+{\em every time a call is made} and check confidentiality for the
+subtrace produced from that variant state until its corresponding
+return. As a result, a ``single'' confidentiality test compactly
+checks multiple nested calls.
+
+Our primary testing targets are the eager {\em Depth Isolation}
+and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
+We test the former against stepwise integrity and confidentiality, and
+the latter against their observational counterparts.
+
+To ensure the effectiveness of testing against our formal properties, we
+use {\em mutation testing}~\cite{JiaH11}. In mutation testing, we inject errors
+(mutations) in a program that should cause the property of interest (here,
+stack safety) to fail, and ensure that the testing framework can find
+them. The bugs we use for our evaluation are either artificially generated
+by us (deliberately weakening the micro-policy in ways that we expect
+should break its guarantees), or actual bugs that we discovered through
+testing our implementation. We elaborate on some such bugs below.
+
+For example, when loading from a stack location, {\em Depth Isolation}
+needs to enforce that the tag on the location being read
+is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
+current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
+that restriction by not checking the depth equality (row {\em
+  LOAD\_NO\_CHECK\_DI}).
+
+Similarly, when storing to a stack location, the correct micro-policy
+needs to ensure that the tag on the memory location is either
+$\tagNoDepth$ or has again the same depth as the current $\PCname$
+tag. Relaxing that constraint causes violations to the integrity
+property (row {\em STORE\_NO\_CHECK}).
+
+\begin{table}[]
+\centering
+\begin{tabular}{c|c|c|c}
+  Bug & Property Violated & MTTF (s) & Tests \\
+  \hline
+      {\em LOAD\_NO\_CHECK\_DI}  & Confidentiality & 24.2 & 13.3 \\
+      {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
+      {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
+  \hline
+  \hline
+      {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
+      {\em LOAD\_NO\_CHECK\_LT}  & Obs. Integrity & 23.5 & 12.0 \\
+      {\em LOAD\_NO\_CHECK\_LT}  & Confidentiality & 19.2 & 695.5 \\
+      {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
+      {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
+  \hline
+\end{tabular}
+\vspace*{1em}
+\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+\vspace*{-2em}
+\label{tab:bug-table}
+\end{table}
+
+The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
+Table~\ref{tab:bug-table}, along with the average number of tests
+it took to find the failure. Experiments were run in a desktop
+machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
+
+Naturally, testing also revealed a number of errors in our
+implementation of the enforcement mechanism (the original was written in C++
+and targeted ARM machine code;
+%\bcp{right?}\leo{yeah}
+we re-implemented it in Coq targeting RISC-V).  These errors range
+from trivial typos to ones that require an intriguingly complex setup
+to provoke.  The most interesting bug (included in the table as row
+{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
+sequence %/policy combination\apt{??}
+did not initialize all locations for the
+newly allocated stack frame correctly, but left some of them as
+$\tagNoDepth$. This allowed for a potential integrity violation, but
+only if a rather complicated sequence of events occured.
+The smallest counterexample requires calling a function {\tt f},
+which fails to initialize some of its frame during the blessed sequence,
+but writes into an uninitialized location $l$ later, treating \(l\) as outside
+the stack. Then {\tt f} calls a further function {\tt g} (which should have
+the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
+which is allowed because the enforcement mechanism still has
+$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
+%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
+%  and the write from {\tt f} was already a violation.}
+
+As for lazy micro-policies and observational properties,
+the original {\em Lazy Tagging and Clearing} micro-policy, implemented as {\em PER\_DEPTH\_TAG},
+fails in testing, in cases where data is leaked between sequential calls.
+To round out our mutation testing we also check {\em LOAD\_NO\_CHECK\_LT},
+equivalent to its counterpart in depth isolation,
+and a version where stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
+It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
+taking twice as long to catch as the next longest.
+
+Our properties have allowed us to identify an enforcement mechanism as
+not really stack safe, and to validate a possible fix.
+Unfortunately, Lazy Per-activation Tagging and Clearing
+puts severe pressure on the tag cache, so it may not be usable in practice.
+However, we conjecture that no lazy scheme can be fully stack safe
+without using unique per-activation identifiers.
+
+\section{Related Work}
+\label{sec:relwork}
+
+\paragraphx{Formal Stack Safety on Capability Machines}
+%
+To our knowledge, the only line of work to date that has attempted
+a positive and formal characterization of stack safety
+is that of Skorstengaard et al. \cite{SkorstengaardSTKJFP}. Their
+StkTokens system enforces stack safety on hardware equipped with
+memory capabilities, similar to those used in the Cheri system~\cite{Woodruff+14,Chisnall+15}, but
+with additional support for linear capabilities.
+They define Stack safety as the conjunction of local state encapsulation (LSE) and
+well-bracketed control flow (WBCF).
+Informally, LSE means that the current activation can only access its own frame,
+and WBCF means that the topmost frame can only return to
+the immediately adjacent frame below. Formally, these properties are defined
+with reference to a capability machine with a built-in call stack and call and return
+instructions; this machine is proven fully abstract with respect to a more
+concrete capability machine, an approach termed ``fully abstract overlay semantics.''
+We improve upon Skorstengaard et al. \cite{SkorstengaardSTKJFP} in two significant ways:
+
+i) Our definition captures the intuition that {\em
+  every} caller should be protected from its callees.  By contrast,
+  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
+  that are not protected, because they are capable of compromising
+  their own security. Consider a function that, when called with the wrong
+  inputs, behaves unreasonably. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
+  {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
+  component is always reasonable. In our setting, individual
+  well-behaved activations of the buggy function would be protected.
+
+ii) Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s fully
+  abstract overlay semantics rely on the definition of a ``stack-safe-by-construction''
+  abstract machine that is closely tied to the underlying machine architecture.
+  While the approach can no doubt be ported to other architectures and enforcement
+  schemes, each such port will require a new abstract machine definition whose
+  safety-by-construction must be argued afresh.
+  Since we abstract over most of the machine state as well as the sealing and return conventions, and
+  can support further enhancements such as coroutines,
+  our properties apply directly to a range of architectures, code features, and enforcement mechanisms.
+
+Skorstengaard et al. \cite{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
+candidate stack safety properties.
+%% : they should be (1) intuitive; (2)
+%% useful for reasoning; (3) reusable in secure compiler chains; (4)
+%% arguably ``complete'' with respect to informal notions of stack
+%% safety; and (5) potentially scalable to richer settings such as
+%% dynamic code generation or multithreading.
+%
+In terms of (1) {\em intuitiveness}, we contend that our property
+definitions are at least as
+intuitive as theirs, though phrased in rather different terms.
+We use the language of formal security
+properties to apply well-established concepts \ifspace and formal techniques \fi
+of noninterference, and we phrase our properties in terms of
+traces of memory and register states, rather than an
+ad-hoc abstract machine.
+Which approach is ``better'' is probably a matter of taste, but ours
+has \ifspace at least \fi the merit of explicitly relating stack safety to existing
+notions of security.
+%
+Demonstrating (2) {\em usefulness} of the safety criterion
+for carrying out formal proofs (about individual programs) for our proposal
+remains future work; concretely, we
+hope to apply our properties to justifying a low-level program logic.
+%
+(3) StkTokens is specifically set up to address {\em reusability} in secure
+compilation chains, while this is not a goal of our work.\bcp{Can we explain
+what it means??}
+%
+(4) Regarding {\em comprehensiveness} in capturing programmers' informal
+understanding of stack safety, we believe that our support
+for a dynamically changing notion of ``adversary,'' as outlined in point (i) above,
+captures an even better definition of safety than LSE; moreover, our observational
+properties are richer and\bcp{cut "richar and"??} more extensional, and thus
+applicable to a broader range of enforcement mechanisms and code features.
+%
+(5) Finally, regarding {\em broad applicability}, we have demonstrated that
+\bcp{unlike StkTokens}
+our approach extends to
+coroutines, a first step towards multithreading. Dynamic code generation is
+more challenging, but StkTokens does not yet handle this either.
+
+We would add one more criterion: reusability across machine models and
+enforcement mechanisms. StkTokens has been used to validatate
+one machine/enforcement mechanism---a linear-capability machine---via formal
+proof. We have used ours to obtain a weaker (testing-based) validation of
+two different mechanisms on a single machine; we hope to target a capability machine
+next as future work.
+
+\paragraphx{Protecting the Stack with Micro-Policies}
+%
+\apt{Could probably condense/omit this.}
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}
+study the protection of stack data through three families of
+micro-policies that tag stack objects with a pair of frame and object identifiers
+used to validate accesses: {\em Return Address Protection} (which
+prevents an adversary from overwriting designated return addresses), {\em Static
+Authorities} (which only allows the code of a function to access the stack frames
+of its own dynamic instances), and {\em Depth Isolation}
+(\cref{sec:enforcement}). All these policies strike various balances between
+precision and performance, evaluated through benchmarks that demonstrate
+limited performance overhead. In addition to the baseline ``eager'' policies,
+they propose a number of lazy optimizations, notably {\em Lazy Tagging} and
+{\em Lazy
+Clearing} (\cref{sec:lazy}).
+
+We evaluate {\em Depth Isolation} and {\em Lazy Tagging and Clearing}
+against our properties, finding that the former does
+indeed enforce stepwise stack safety and therefore observable stack
+safety. We find that the latter does not actually enforce
+stack safety, and we proposed a fix.
+
+\paragraphx{Heap Safety as a Security Property}
+%
+Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18} show that
+heap safety may be framed as a security property using
+the language of noninterference. Their paper was a significant inspiration
+for the present work, and we believe that the two safety conditions should
+combine smoothly.
+
+\section{Future Work}
+\label{sec:future}
+
+From here, we see two natural ways to extend the work. \apt{what about addressing some of the limitations?}
+First, we should thoroughly evaluate
+other enforcement mechanisms, especially Cheri-based capability
+systems. Second, we can further demonstrate
+the functionality of our definitions by using them to prove
+the correctness of a programming logic with strengthened rules around calls and returns.
+
+\paragraphx{Stack Safety in the Cheri Ecosystem}
+%
+There are several proposals around the use of Cheri capabilities to enforce stack safety,
+including mechanisms that use the standard Cheri hardware (which includes local
+capabilities) \cite{SkorstengaardLocal},
+and others that propose entirely new types of capabilities, such as linear
+\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
+\cite{Tsampas+19}, and monotonic \cite{GeorgesComplete} capabilities.
+Of these, uninitialized capabilities seem particularly promising---they naturally enforce
+write-before-read behavior and therefore will enable confidentiality without the need
+to clear large swaths of the stack. We plan to implement a model of the stack safety
+system described by Georges et al. \cite{Georges+21} and test it against our properties using
+QuickChick.
 
@@
 
+\appendix
+
+\subsection{Events and Traces}
+\label{sec:events}
+
+We abstract over the events that can be observed in the system, defining them
+only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
+event. Other events might represent certain function calls (i.e., system calls)
+or writes to special addresses representing mmapped regions.
+A {\em trace} is a nonempty, finite or infinite sequence
+of events, ranged over by \(\obsT\).
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
+for list-cons.
+
+We write that execution from a state produces an observation trace
+\(\mach,\context \hookrightarrow \obsT\) as follows, coinductively:
+
+\judgmenttwo{\((\mach,\context) \stepstounder{} (\mach',\context',\obs)\)}
+            {\((\mach',\context') \hookrightarrow \obsT\)}
+            {\((\mach,\context) \hookrightarrow \notfinished{\obs}{\obsT}\)}
+
+We define another relation that takes a trace until we have returned from the
+active principal.
+We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
+\(d\) is the depth of the current call.
+
+\judgment{\(|\sigma| < d\)}
+         {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \tau\)}
+
+\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
+                  {\(|\sigma| \geq d\)}
+                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
+                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{\obs}{\obsT}\)}
+
+\paragraph*{Observational Similarity}
+
+We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
+written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
+is the same. That is, we compare up to deletion of \(\tau\) events.
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{}{\(\obsT \eqsim \obsT\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
+\end{minipage}
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
+\end{minipage}
+
+\paragraph*{Variants and Irrelevant Values}
+
+For a given system, assume an compatibility equivalence relation \(\sim\) on values.
+Two states are variants with respect to a set of elements \(\components\)
+if they agree on the value of every element not in \(\components\) and
+and have compatible values for every element in \(\components\).
+Our notion of non-interference involves comparing the traces of such
+\(\components\)-variants. We use this to define sets of irrelevant elements,
+and secrets in general.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+written \(\mach \approx_\components \nach\), if, for
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
+and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
+
+\definition An element set \(\components\) in state \((\mach,\context)\) contains
+irrelevant values, written \((\mach,\context) \parallel \components\), if for all
+\(\nach\) such that \(\mach \approx_{\components} \nach\), if
+\((\mach,\context) \hookrightarrow \obsT\) and
+\((\nach,\context) \hookrightarrow \obsT'\), then
+\(\obsT \eqsim \obsT'\).
+
+\subsection{Properties at Return}
+\label{sec:facts}
+
+Here we define some logical operations to reason about the behavior of the
+system over time. These have a temporal-logic flavor, as they reflect
+the expected behavior of the system in the future, after a possible return.
+
+\paragraph*{On-return}
+
+The intuition behind integrity (below) is that a caller may expect its
+sealed data to be unchanged when control returns to it. In fact, the callee
+may overwrite such data---when the data are found in callee-saved registers
+this is perfectly legal---as long as it either restores it, or has some guarantee
+that its changes will not impact the caller.
+
+We start by defining a second-order logical operator
+\(d \uparrow P\), pronounced ``\(P\) holds on return from depth \(d\),''
+where \(P\) is a predicate on machine states. This is a coinductive relation
+similar to ``weak until'' in temporal logic---it also holds if the program never
+returns from depth \(d\).
+
+\judgmenttwo[Returned]
+            {\(|\sigma| < d\)}
+            {\(P ~ (\mach,(V,\sigma))\)}
+            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
+
+\judgmenttwobrlong[Step]
+                  {\(|\sigma| \geq d\)}
+                  {\((d \uparrow P) ~ (\mach', \context')\)}
+                  {\((\mach, (V,\sigma)) \stepstounder{\overline{\psi}} (\mach', \context',\obs)\)}
+                  {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
+
+Similarly, for confidentiality, we will want to compare pairs of future states,
+so we give a binary equivalent, \(d \Uparrow R\), where
+\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
+first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\),
+respectively. Once again, \(\Uparrow\) is coinductive.
+
+\judgmenttwobrlong[Returned]
+            {\(|\sigma_1| < d\)}
+            {\(|\sigma_2| < d\)}
+            {\((\mach_1,(V_1,\sigma_1)) ~ R ~ (\mach_2,(V_2,\sigma_2))\)}
+            {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+
+\judgmenttwobrlong[Left]
+              {\(|\sigma_1| \geq d\)}
+              {\((\mach_1,(V_1,\sigma_1)) \stepstounder{\overline{\psi}} (\mach_1',\context_1',\obs)\)}
+              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+
+\judgmenttwobrlong[Right]
+              {\(|\sigma_2| \geq d\)}
+              {\((\mach_2,(V_2,\sigma_2)) \stepstounder{\overline{\psi}} (\mach_2',\context_2',\obs)\)}
+              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
+              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+
+\subsection{Properties}
+\label{sec:props}
+
+\paragraph*{Well-bracketed Control Flow}
+
+\definition We define a correct return from \(\mach\) as:
+%
+\[\ret(\mach) = \{(\mach',\context') \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}\]
+
+\definition
+A system enjoys  well-bracketed control flow if, whenever
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) where
+\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\),
+\(|\sigma| \uparrow \ret(\mach)\) holds on \((\mach',(V,\sigma))\).
+
+\paragraph*{Integrity}
+
+\definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
+such that \(\mach[\component] \not = \mach'[\component]\).
+
+\definition We define \(\intProp(\mach,V)\), to hold on a state \((\mach',\context')\) when
+\(\{\component \mid V ~ \component = \sealed\} \cap \Delta(\mach,\mach')\)
+is irrelevant in \((\mach', \context')\).
+
+\definition A system enjoys {\it stack integrity} if, whenever
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
+\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
+we have:
+\[(|\sigma'| \uparrow \intProp(\mach',V)) ~ (\mach',(V,\sigma))\]
+
+\paragraph*{Caller Confidentiality}
+
+\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
+\(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
+Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to \(\components\)
+if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
+\(|\sigma| \uparrow (\mach,\context) \hookrightarrow \obsT\) and
+\(|\sigma| \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
+\(\obsT \simeq \obsT'\).
+
+\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
+is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
+This treats
+\(\mach\) and \(\nach\) as the initial states and
+\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
+set of elements that were corrupted between them.
+
+\definition We define the relation \(\conf(\mach,\nach)\) to hold on a pair of states
+\((\mach',\context)\) and \((\nach',\context')\) when
+\(\bar{\Diamond}(\mach,\nach,\mach',\nach')\) is irrelevant in \((\mach',\context)\).
+
+\definition A state \((\mach,(V,\sigma))\) enjoys {\it return-time confidentiality}
+if, for \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\)
+and any \(\nach\) that is a \(\components\)-variant of \(\mach\), we have:
+\[(\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \conf(\mach,\nach) ~ (\nach,(V,\sigma))\]
+
+\definition A system enjoys {\it caller confidentiality} if, whenever
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+\((\mach',\context')\) enjoys both internal and return-time confidentiality.
+
+\paragraph*{Callee Confidentiality}
+
+\definition We define \(\cconf(\mach,V)\), to hold on a state \((\mach',\context')\) when
+\(\{\component \mid V ~ \component = \unsealed\} \cap \Delta(\mach,\mach')\)
+is irrelevant in \((\mach', \context')\).
+
+\definition A system enjoys {\it callee confidentiality} if, whenever
+\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\)
+with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+% or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
+we have:
+\[(|\sigma| \uparrow \cconf(\mach',V)) ~ (\mach',\context')\]
+
 \end{document}

### 2023-01-17 submission5/core.tex
Finished reorg, largely

@@
 both caller- and callee-saved registers, arguments passed on the stack,
-exceptions, [tail call elimination, and concurrency.] They are modular by
+tail-call elimination, [exceptions, and concurrency.] They are modular by
 design, in order to be further extensible to other features and architectures.
@@
 
-Security mechanisms designed to prevent a certain class of attacks can be brittle---successfully
+Security mechanisms designed to prevent a certain attack can be brittle---successfully
 eliminating the attack as it exists at the time of publication, but leaving room for attackers
 to find new, similar attacks. To mitigate this risk, mechanisms should aim to enforce formal
-properties that can be proven or tested. Such properties become the specification against which
-enforcement can be validated; even enforcement mechanisms that do not fulfill the property
-benefit from the ability to articulate {\it why}.
+properties of correct behavior that can be proven or tested. Such properties become the
+specification against which enforcement can be validated; even enforcement mechanisms that do
+not fulfill the property benefit from the ability to articulate {\it why} they fail.
 
@@
 a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Subtle tweaks to the definition
-accompany any addition to the model.
+return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
+must be accompanied by tweaks to the model.
 
@@
 the formal criteria for integrity, confidentiality, and control flow
-to obey the requirements.
+to obey the requirements. We consider the criteria in (3) to represent universal definitions
+that are parameterized by (1).
 
 We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all of them. Some focus exclusively on
-control flow, others protect the confidentiality of the caller but not the callee, etc.
+not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
+exclusively on control flow, others protect the confidentiality of the caller but not the
+callee, etc.
 
@@
 
+\paragraph*{Limitations}
+
+%Our concurrency model is fairly
+%simplistic, assuming a fixed number of threads each with its own dedicated processor.
+We model memory-safe stack objects, but not a heap. Regions outside of
+stacks can be used however the compiler likes, including as a heap, but no protection is
+built in and our properties assume that if a pointer to a stack object is stored there,
+it is permanently compromised.
+
+Our properties are termination insensitive: in order to support enforcement mechanisms
+that failstop rather than execute dangerous code, the properties must treat failstops
+as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
+This is consistent with our treatment of properties that specify the state after a return,
+in that a callee that diverges for any reason will never reach the return, and will fulfill
+the property vacuously.
+
 \section{Key Ideas}
+\label{sec:ideas}
 
@@
 The remainder of this section introduces ideas that will be important throughout.
+
+\paragraph*{Terminology and Notation}
+
+The building blocks of a machine are {\em words} and {\em registers}.
+Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
+and are drawn from the set \(\WORDS\).
+Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
+given the special name \(\SP\);
+some registers may be classified as caller-saved or callee-saved.
+Along with the program counter, \(\PCname\), these are referred to as
+{\em state elements} \(\component\) in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).
+
+A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
+\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
+The details of value structure depend on the specific machine being modelled;
+intuitively, the payload represents the part of the value that is relevant to
+the behavior of the basic machine, while the rest of the value may contain
+information relevant to a hardware enforcement mechanism (such as a tag).
+We write \(\mach[\component]\) to denote the value of \(\mach\) at
+\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
 
@@
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
-the following table. Those that we explore in detail in Section \ref{sec:examples}
+\cref{tab:psi}. Those that we explore in detail in \cref{sec:examples}
 are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In the full formalism (Appendix \ref{app:full}) and the machine that we test,
+In the full formalism (\cref{app:full}) and the machine that we test,
 we cover additional operations add parameters to our existing ones, to support
@@
 We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
-based on pointer provenance in Appendix \ref{app:ptr}, represented by the dark grey
+based on pointer provenance in \cref{app:ptr}, represented by the dark grey
 operations, but we do not test this.
@@
 
+\begin{table}
   \begin{tabular}{| l | l |}
@@
   \end{tabular}
+  \label{tab:psi}
+\end{table}
 
@@
 
-\paragraph*{Variant States}
-
-\paragraph*{Events and Traces}
+\paragraph*{Events, Steps, and Traces}
 
 It is common to specify the behavior of a system, including security properties, in
-terms of traces of observable events rather than details of the machine state.
+terms of traces of observable {\em events} rather than details of the machine state.
 Lazy enforcement works because the deferred check catches errors before they appear
-in the trace. The nature of traces varies based on the properties of interest, but
-given a particular model of observation, we can consider
-corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.
+in the trace. The nature of events is a parameter of the machine, and can be specialized
+to any notion of observable behavior.
+
+Events are ranged over by \(\obs\), and include the special silent event
+\(\tau\). Traces are non-empty, potentially infinite (coinductive) series of
+events, and traces are {\em similar} \(\obsT_1 \simeq \obsT_2\) if they are
+equal up to deletion of \(\tau\) events. This means that an infinite silent trace is
+similar to any trace, consistent with termination insensitivity.
+
+A machine step function
+\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
+\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
+is labeled by an ordered list of security-relevant operations and an event.
+We lift it into the security semantics by adding a security context to the
+state, giving us a combined step function
+\(\mach,\context \stepstounder{\overline{\psi},\obs} \mach',\context' \in
+\MACHS \times \CONTEXTS \rightarrow \MACHS \times \CONTEXT \times \mathit{list} ~ \Psi \times \OBSS\).
+We will define \(\stepstounder{}\) informally in the next section, and formally
+in \cref{sec:props}.
 
@@
 modelled (see~\ref{??}).
-
-\paragraph*{Limitations}
-
-%Our concurrency model is fairly
-%simplistic, assuming a fixed number of threads each with its own dedicated processor.
-We model memory-safe stack objects, but not a heap. Regions outside of
-stacks can be used however the compiler likes, including as a heap, but no protection is
-built in and our properties assume that if a pointer to a stack object is stored there,
-it is permanently compromised.
 
@@
 
-\section{Machine Model and Security Semantics}
-
-\subsection{The Basic Machine}
-
-The building blocks of the machine are {\em words} and {\em registers}.
-Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
-and are drawn from the set \(\WORDS\).
-Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
-given the special name \(\SP\);
-some registers may be classified as caller-saved or callee-saved.
-%
-Along with the program counter, \(\PCname\), these are {\em state elements} \(\component\)
-in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).
+\section{Machine and Security Semantics, Formally}
+\label{sec:props}
 
@@
 
-A {\em machine state} \(\mach\) is a map from state elements to a set \(\mathcal{V}\) of
-\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
-The details of value structure depend on the specific machine being modelled;
-intuitively, the payload represents the part of the value that is relevant to
-the behavior of the basic machine, while the rest of the value may contain
-information relevant to a hardware enforcement mechanism (such as a tag).
-We write \(\mach[\component]\) to denote the value of \(\mach\) at
-\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
-
-{\em Events} are drawn from a set \(\OBSS\)\bcp{lowercase?} and ranged over by \(\obs\), with a
-special silent event \(\tau\).
-The machine step function
-\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
-\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
-is labeled by an ordered list of security-relevant operations, defined below,
-and an event.
-
-While some enforcement mechanisms may cause a machine to
-failstop, we model this as the machine silently diverging by remaining in the
-same state. We consider silently diverging traces to be indistinguishable
-(from the point of view of an attacker) from traces that might eventually
-produce output\bcp{why?}, an assumption that gives our properties the flavor of
-{\it termination insensitivity}.
-
-\subsection{Security-Relevant Operations, Principals, and State}
-
-Security-relevant operations consist of calls, returns, and allocations.
-Calls identify the address of the call's target,
-plus the argument registers. Allocations and deallocations are phrased in terms of a pair
-of an offset from the stack pointer and a size. Returns don't require any additional information.
-%
-
-%\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
+Recall from \cref{sec:ideas} that a machine is a map from state elements
+to values, with a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
+The step function is total---while some enforcement mechanisms may cause a machine
+to failstop, we model this as the machine silently diverging by remaining in the
+same state.
+
+A view \(V\) maps elements to security classes, and a context consists of the
+view of the active function together with a stack of inactive views.
 The initial context is \(\context_0 = (V_0, \emplist)\).
-
+%
+\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
+%
 We have described informally how the security context develops as the system performs
@@
 The definition of \(Op\) is most convenient to present decomposed into
-rules for each operation. In the simple system, the operations affect the
-context as follows:
-
+rules for each operation. We have already seen the intuition behind the rules for
+\(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\). Their precise behavior
+is found in \cref{fig:basicops}.
+
+\begin{figure}
 \judgmentbr[Alloc]
            {\(b = \mach[\SP] + \mathit{off}\)}
-           {\(V' = V \llbracket \addr \mapsto \sealed |
+           {\(V' = V \llbracket \addr \mapsto \object |
              b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
            {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+
+\judgmentbr[Dealloc]
+           {\(b = \mach[\SP] + \mathit{off}\)}
+           {\(V' = V \llbracket \addr \mapsto \unsealed |
+             b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
+           {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
 
@@
 
-\section{Formalizing Properties}
-
-\section{Advanced Machine}
+         \caption{The basic operations}
+         \label{fig:basicops}
+\end{figure}
+
+\subsection{Advanced Operations}
 
 The advanced machine defines additional operations, and extends some of the old ones.
-It adds a field \(\overline{sa}\) to the \(\mathbf{call}\) operation, which is a set of
-triples of a register, a base offset, and a size, and denotes that the object at the offset
-from the given register is to be passed as an argument. The new \(\mathbf{tailcall}\) operation
-has the same fields, but deals with tail call optimizations in which the callee will reuse
-the caller's stack. The \(\mathbf{alloc}\) operation
+We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
+triples of a register \(\reg\), a base offset \(off\), and a size \(sz\), denoting
+that the object at the offset \(off\) from \(\reg\) is to be passed as an argument.
+The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
+optimizations in which the callee will reuse the caller's stack. The \(\mathbf{alloc}\) operation
 gains a boolen flag indicating whether the allocation is to be public, accessible to
-all functions until deallocated. And, we add the corresponding \(\mathbf{dealloc}\)
-operation.
+all functions until deallocated.
 
@@
 be available to allocate privately. We define a helper function, \(\mathit{range}\),
-to map an offset from a register and a size into a set of addresses.
-%
+to map an offset from a register and a size into a set of addresses, and update
+the operation's security semantics as shown in \cref{sfig:publicalloc}.
+
+\begin{figure*}
+  \begin{subfigure}{0.4\textwidth}
     \[\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach \triangleq
     \{\mach[\reg]+i | \mathit{off} \leq i < \mathit{off+sz}\}\]
-%
-\judgmentbr[AllocPriv]
+
+    \judgmentbr[]
                {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
@@
 
-\judgmentbr[AllocPub]
+    \judgmentbr[]
                {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
@@
                {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-
-\judgmentbr[Dealloc]
-           {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
-             \{\addr \mid V ~ \addr = \unsealed\}\)}
-           {\(V' = V \llbracket \addr \mapsto \unsealed \mid \addr \in \components \rrbracket\)}
-           {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+    \caption{Allocating Public Regions}
+    \label{sfig:publicalloc}
+  \end{subfigure}
+  \begin{subfigure}{0.6\textwidth}
+    %
+    \[\mathit{passed} ~ \overline{sa} ~ \mach = \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
+    \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
+    %
+    \judgmentbr[]
+               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+                 ~ (V,\sigma) = (V'',V::\sigma)\)}
+    %
+    \judgmentbr[]
+               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{tailcall} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+                 ~ (V,\sigma) = (V',\sigma)\)}
+    \caption{Argument Passing on Stack}
+    \label{sfig:stkargs}
+  \end{subfigure}
+  \caption{Advanced Operations}
+  \label{fig:advops}
+\end{figure*}
 
@@
 then extend the call operation to keep all objects in \(\mathit{passed}\) marked
-as \(\object\) and seal everything else.
-%
-\[\mathit{passed} ~ \overline{sa} ~ \mach = \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
-\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
-%
-\judgmentbrbr[]
-             {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
-             {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-               \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-             {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-             {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
-               (V'',V::\sigma)\)}
+as \(\object\) and seal everything else (\cref{sfig:stkargs}).
 
@@
 behaviors may require sequences of {\bf dealloc}, {\bf tailcall}, then {\bf alloc}.
-
-\judgmentbrbr[]
-             {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
-             {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-               \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-             {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-             {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
-               (V',\sigma)\)}
 
@@
 \subsection{Properties}
-\label{sec:props}
+\label{app:props}
 

### 2023-01-17 submission5/core.tex
Tweak to layout.

@@
 call in a simple example machine and discuss how each of our properties applies to it,
-informally. Then in \cref{sec:machine} we will introduce the formal machine model and give
-its {\em security semantics}, described below. In \cref{sec:props} we give abbreviated
-formal statements of the properties, while the full formalism is found in \cref{app:props}.
-\Cref{sec:ext} gives the security semantics for the full, tested machine, and
-\cref{sec:testing} describes the testing framework used to test it.
+informally. Then in \cref{sec:props} we will introduce the machine model that we test, and give
+its {\em security semantics}, described in brief below.
+%In \cref{sec:props} we give abbreviated
+%formal statements of the properties, while
+The full formalism is found in \cref{app:props}.
+\Cref{sec:testing} describes the testing framework used to test it.
 \Cref{sec:relwork,sec:future} discuss related and future work.

### 2023-01-20 submission5/core.tex
Stray S restored
Editing through key ideas

@@
 an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
-which is broadly consistent across architectures and therefore amenable to a very general approach.)
+which is broadly consistent across architectures and language semantics, and therefore amenable
+to a very general approach.)
 The desired behavior of the stack varies based on ISA, calling convention,
@@
 harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
-and of compatibility with existing codebases. In the motivating case for this work,
+and of compatibility with existing codebases. In a crucial motivating example for this work,
 Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
@@
 
-Roessler and DeHon find that the obvious way to enforce stack safety in this setting
+Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
 is inefficient when allocating and clearing stack frames, because each word must be initialized
@@
 the micro-policy enforces that memory be read by the same fuction that wrote it. This
-lazy enforcement is incompatible with existing models of stack safety, particularly the
-integrity side, because an attacker can modify another function's data. But the owner of
-the data will never see the changes, so it may still be secure in some sense. We aim to elaborate
+lazy enforcement is incompatible with existing models of stack safety,
+because an attacker can modify another function's data. But the owner of
+the data will never see the changes, so it is secure in the sense that the attacker is
+unsuccessful at changing program behavior (except, perhaps, by causing a fault.) We aim to elaborate
 that sense, and to test whether the micro-policy fulfills it.
+
 This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
 machine states that make claims about the future observable behavior of execution from those states.
+That is not to say that our properties are specialized for lazy micro-policies. It also supports
+normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, our
+observational properties are even weaker than necessary to support Roessler and Dehon's micro-policy.
 
@@
 call in a simple example machine and discuss how each of our properties applies to it,
-informally. Then in \cref{sec:props} we will introduce the machine model that we test, and give
-its {\em security semantics}, described in brief below.
-%In \cref{sec:props} we give abbreviated
-%formal statements of the properties, while
-The full formalism is found in \cref{app:props}.
-\Cref{sec:testing} describes the testing framework used to test it.
-\Cref{sec:relwork,sec:future} discuss related and future work.
+informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
+its {\em security semantics}, described in brief below. We formalize the definitions of
+stack safety that derive from the security semantics in \cref{sec:props}.
+\Cref{sec:testing} describes the testing framework, and
+\cref{sec:relwork,sec:future} discuss related and future work.
 The remainder of this section introduces ideas that will be important throughout.
 
-\paragraph*{Terminology and Notation}
-
-The building blocks of a machine are {\em words} and {\em registers}.
-Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
-and are drawn from the set \(\WORDS\).
-Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
-given the special name \(\SP\);
-some registers may be classified as caller-saved or callee-saved.
-Along with the program counter, \(\PCname\), these are referred to as
-{\em state elements} \(\component\) in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).
-
-A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
-\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
-The details of value structure depend on the specific machine being modelled;
-intuitively, the payload represents the part of the value that is relevant to
-the behavior of the basic machine, while the rest of the value may contain
-information relevant to a hardware enforcement mechanism (such as a tag).
-We write \(\mach[\component]\) to denote the value of \(\mach\) at
-\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
-
-\paragraph*{Security Semantics}
+\paragraph*{Security Semantics and Property Structure}
 
@@
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
-\cref{tab:psi}. Those that we explore in detail in \cref{sec:examples}
+\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
 are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In the full formalism (\cref{app:full}) and the machine that we test,
-we cover additional operations add parameters to our existing ones, to support
+In \cref{sec:machine}), we cover additional operations add parameters to the examples', to support
 various forms of memory sharing and tail-call elimination. These are shown in light gray.
@@
 \newcommand{\theory}{\rowcolor{black!25}}
-
-\vspace{\abovedisplayskip}
 
@@
     \hline
-    \example \(\mathbf{call}\) & Target address \\
-    \example & Argument registers \\
+    \example \(\mathbf{call}\) & Target address, argument registers \\
     \testing & Stack argument offset \& size \\
@@
   \end{tabular}
+  \caption{Security-relevant Operations and their Parameters}
   \label{tab:psi}
 \end{table}
-
-\vspace{\belowdisplayskip}
 
@@
 must satisfy in order to correctly obey the properties.
-
-\paragraph*{Security Classes}
 
@@
 available to be allocated (\(\unsealed\)), or
-always accessible (\(\public\)). The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
-all other locations to \(\public\), and registers based on which set they
-belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
-
-A {\it context} is a pair of the current activation's view and
-a list of views representing the call stack (pending inactive
-principals), ranged over by \(\sigma\).
-
-\paragraph*{Events, Steps, and Traces}
+always accessible (\(\public\)).
+
+\paragraph*{The Properties and their Hierarchy}
+
+We have five different properties to introduce:
+
+\begin{itemize}
+\item Well=bracketed Control Flow (\(WBCF\))
+\item Caller Integrity (\(CLRI\))
+\item Caller Confidentiality (\(CLRC\))
+\item Callee Confidentiality (\(CLEC\))
+\item Callee Integrity (\(CLEI\))
+\end{itemize}
+
+The properties form a rough hierarchy; although they are technically separate and need
+not coexist in any particular combination, some are likely to take priority in terms of their
+importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
+as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
+Many also cover some degree of (caller) integrity. [TODO: more hierarchy].
+
+\paragraph*{Events}
 
@@
 to any notion of observable behavior.
-
-Events are ranged over by \(\obs\), and include the special silent event
-\(\tau\). Traces are non-empty, potentially infinite (coinductive) series of
-events, and traces are {\em similar} \(\obsT_1 \simeq \obsT_2\) if they are
-equal up to deletion of \(\tau\) events. This means that an infinite silent trace is
-similar to any trace, consistent with termination insensitivity.
-
-A machine step function
-\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
-\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
-is labeled by an ordered list of security-relevant operations and an event.
-We lift it into the security semantics by adding a security context to the
-state, giving us a combined step function
-\(\mach,\context \stepstounder{\overline{\psi},\obs} \mach',\context' \in
-\MACHS \times \CONTEXTS \rightarrow \MACHS \times \CONTEXT \times \mathit{list} ~ \Psi \times \OBSS\).
-We will define \(\stepstounder{}\) informally in the next section, and formally
-in \cref{sec:props}.
 
@@
 This is a strong threat model, but it does omit some important aspects of
-security enforcement: In particular, hardware and timing attacks are out of scope,
-and our properties are termination insensitive as a result of how failstops are
-modelled (see~\ref{??}).
-
-%\section{Calls, Returns, and Private Allocations}
+security enforcement: In particular, hardware and timing attacks are out of scope.
+
 \section{Properties by Example}
@@
 {\tt main} takes a secret as its argument and initializes a variable to contain
-potentially sensitive data. We assume that the only observable events are stores
+potentially sensitive data. We define an observable event to be a store
 to a special address, called {\tt out}.
@@
 Figure \ref{fig:exec1} shows how the corresponding security context
-develops over the first few steps of the program. We assume execution begins
-at the start of {\tt main} (\(\PCname = 0\)) and that the stack pointer is initially at
-address 1000.
-\(\xrightarrow{\overline{\psi}}\) transitions are numbered and
-written \(\downarrow \overline{\psi}\) in this diagram; events are omitted.
+develops over the first few steps of the program, beginning with
+the start of {\tt main} (\(\PCname = 0\)) and with the stack pointer at address 1000.
+State transitions are numbered and labeled with a list of security operations, written
+\(\downarrow \overline{\psi}\) between steps.
 
@@
 values after the call,'' this would be stricter than necessary. Suppose that
-a callee overwrites some data of its caller, but the caller never accesses that data.
-By sheer luck, the callee has managed not to impact the observable execution beyond that one write.
-Luck, or enforcement: Lazy Tagging and Clearing, from Roessler and Dehon, works by
-failstopping any time a caller accesses callee-tainted data, so that out-of-frame
-writes should never impact execution.
+a callee overwrites some data of its caller, but the caller never accesses that data
+(or only does so after re-initializing it.) This would be harmless, with the callee
+essentially using the caller's memory as scratch space, but the caller never seeing any change.
 
@@
 
-In place of a predicate we define \(\conf\) as a binary relation on states,
+In place of a predicate we define \(\confProp\) as a binary relation on states,
 which holds on eventual return states \(\mach'\) and \(\nach'\)
@@
 are the same in \(\mach\) and \(\nach\). We then define
-return-time confidentiality as holding if \(\conf\) holds on the matching return states
+return-time confidentiality as holding if \(\confProp\) holds on the matching return states
 from \(\mach\) and \(\nach\), respectively.
@@
 \section{Machine and Security Semantics, Formally}
-\label{sec:props}
+\label{sec:machine}
 
@@
 
-Recall from \cref{sec:ideas} that a machine is a map from state elements
-to values, with a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
-The step function is total---while some enforcement mechanisms may cause a machine
-to failstop, we model this as the machine silently diverging by remaining in the
-same state.
+\paragraph*{Terminology and Notation}
+
+The building blocks of a machine are {\em words} and {\em registers}.
+Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
+and are drawn from the set \(\WORDS\).
+Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
+given the special name \(\SP\);
+some registers may be classified as caller-saved or callee-saved.
+Along with the program counter, \(\PCname\), these are referred to as
+{\em state elements} \(\component\) in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).
+
+A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
+\emph{values},  with a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
+Each value \(v\) contains a \emph{payload} word, written \(|v|\).
+The details of value structure depend on the specific machine being modelled;
+intuitively, the payload represents the part of the value that is relevant to
+the behavior of the basic machine, while the rest of the value may contain
+information relevant to a hardware enforcement mechanism (such as a tag).
+We write \(\mach[\component]\) to denote the value of \(\mach\) at
+\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
+
+Security classes are ranged over by \(C\).
+For some security class \(C\), we write \(C(V)\)
+to denote the set of elements \(\component\) such that \(V ~ \component = C\).
+The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
+all other locations to \(\public\), and registers based on which set they
+belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
+
+A {\it context} is a pair of the current activation's view and
+a list of views representing the call stack (pending inactive
+principals), ranged over by \(\sigma\).
 
@@
 
-\subsection{Provenance, Capabilities, and Protecting Objects}
-
-What if we want to express a finer-grained notion of safety, in which
-stack objects are protected unless the function that owns them intentionally
-passes a pointer to them? This can be thought of as a {\it capability}-based
-notion of security. Capabilities are unforgeable tokens that grant access to
-a region of memory, typically corresponding to valid pointers to that region.
-So, in order to express such a property, we need our machine to carry some notion
-of {\it pointer provenance}---a distinction between a pointer that is intended to
-point to a given object, and non-pointer integers as well as pointers to other objects.
-
-One such provenance model is Memarian et al.'s PVI \cite{}, in which pointers are
-annotated with the identity of the object they first pointed to. This annotation is
-propagated when the pointer is copied and when operations are performed on it---even
-integer-only operations.
-
-We can model this as a trio of additional security-relevant operations: one which
-declares a register to contain a valid pointer, one which transmits the provenance
-of a pointer from one element to another, and one which clears the provenance
-(for instance, when a pointer is modified in place in a way that makes it invalid.)
-
-In addition to the normal call stack, our security context will carry a map \(\rho\) from
-elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
-Existing operations are extended to keep the value of \(\rho\) the same, and the new operations
-work as follows:
-
-\judgmentbr[Promote]
-           {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})\)}
-           {\(\rho' = \rho[\reg_{dst} \mapsto \mathit{range} ~ \reg_{base} ~ \mathit{off} ~ \mathit{sz}]\)}
-           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-
-\judgmentbr[Propagate]
-           {\(\psi = \mathbf{propagate} ~ \component_{src} ~ \component_{dst}\)}
-           {\(\rho' = \rho[\component_{dst} \mapsto \rho[\component_{src}]]\)}
-           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-
-\judgment[Clear]
-         {\(\psi = \mathbf{clear} ~ \component\)}
-         {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
-
-We now have a notion of provenance, and must integrate it into the definition of
-stack safety. We essentially generalize the above notion of passing: we will consider
-a caller to have intentionally passed an object if that object is reachable by
-a capability that has been passed to the callee. This includes capabilities passed
-indirectly, by being stored in an object that is in turn passed. Formally, we call
-this set \(\mathit{capped}\), and define it recursively:
-%
-\[\begin{split}
-& \mathit{capped} ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{\component\} \cup \mathit{capped} ~ \components' ~ \rho \textnormal{ where} \\
-& \components' = \{\component' |\rho[\component] = (\mathit{base},\mathit{bound})
-\land \mathit{base} \leq \component' < \mathit{bound}\} \\
-\end{split}\]
-
-We then tweak the call operation to seal only objects that are in \(\mathit{capped}\), or
-the previously defined \(\mathit{passed}\).
-
-\judgmentbrbrbr[]
-               {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
-               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(\components = \{\addr | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \cup (\mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \}\)}
-               {\(V'' = V'\llbracket \components \mapsto \sealed \rrbracket\)}
-               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) =
-                 (V'',V::\sigma,\rho)\)}
-
-Note that we have a degree of monotonicity here. Once an object is sealed (because its
-capability has not been passed to a callee), subsequent nested calls can never unseal it.
-On the other hand, an object that is passed via a pointer may be passed on indefinitely.
+\section{Formal properties}
+\label{sec:props}
+
+Now we are ready to introduce the formal properties that define correctly enforced
+security semantics. First, we must give a few prelimaries: the generation and comparison
+of traces, variant states, and ``on-return'' assertions.
+
+\subsection{Events and Traces}
+\label{sec:events}
+
+We abstract over the events that can be observed in the system, defining them
+only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
+event. Other events might represent certain function calls (i.e., system calls)
+or writes to special addresses representing mmapped regions.
+A {\em trace} is a nonempty, finite or infinite sequence
+of events, ranged over by \(\obsT\).
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
+for list-cons.
+
+We write the trace of execution from a state up to the first
+state below depth \(d\) via the \(\hookrightarrow\) operator, defined coinductively.
+When \(d = 0\), the trace will always be infinite. In this case we
+omit \(d\) and just write \(\hookrightarrow s\).
+
+\judgment{\(|\sigma| < d\)}
+         {\(d \hookrightarrow (\mach,(V,\sigma)) = \tau\)}
+
+\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
+              {\(|\sigma| \geq d\)}
+              {\(d \hookrightarrow (\mach',\context') = \obsT\)}
+              {\(d \hookrightarrow (\mach,(V,\sigma)) = \notfinished{\obs}{\obsT}\)}
+
+Two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
+written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
+is the same. That is, we compare up to deletion of \(\tau\) events.
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{}{\(\obsT \eqsim \obsT\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
+\end{minipage}
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
+\end{minipage}
+
+\paragraph*{Variants and Irrelevant Values}
+
+For a given system, assume an compatibility equivalence relation \(\sim\) on values.
+Two states are variants with respect to a set of elements \(\components\)
+if they agree on the value of every element not in \(\components\) and
+and have compatible values for every element in \(\components\).
+Our notion of non-interference involves comparing the traces of such
+\(\components\)-variants. We use this to define sets of irrelevant elements,
+and secrets in general.
+
+\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
+written \(\mach \approx_\components \nach\), if, for
+all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
+and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
+
+\definition An element set \(\components\) in state \((\mach,\context)\) contains
+irrelevant values, written \((\mach,\context) \parallel \components\), if for all
+\(\nach\) such that \(\mach \approx_{\components} \nach\),
+\(\hookrightarrow (\mach,\context)  \eqsim \hookrightarrow (\nach,\context)\).
+
+\subsection{Properties at Return}
+\label{sec:facts}
+
+Here we define some logical operations to reason about the behavior of the
+system over time. These have a temporal-logic flavor, as they reflect
+the expected behavior of the system in the future, after a possible return.
+
+\paragraph*{On-return}
+
+The intuition behind integrity (below) is that a caller may expect its
+sealed data to be unchanged when control returns to it. In fact, the callee
+may overwrite such data---when the data are found in callee-saved registers
+this is perfectly legal---as long as it either restores it, or has some guarantee
+that its changes will not impact the caller.
+
+We start by defining a second-order logical operator
+\(d \uparrow P\), pronounced ``\(P\) holds on return from depth \(d\),''
+where \(P\) is a predicate on machine states. This is a coinductive relation
+similar to ``weak until'' in temporal logic---it also holds if the program never
+returns from depth \(d\).
+
+\judgmenttwo[Returned]
+            {\(|\sigma| < d\)}
+            {\(P ~ \mach\)}
+            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
+
+\judgmenttwobrlong[Step]
+                  {\(|\sigma| \geq d\)}
+                  {\((d \uparrow P) ~ (\mach', \context')\)}
+                  {\((\mach, (V,\sigma)) \stepstounder{\overline{\psi}} (\mach', \context',\obs)\)}
+                  {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
+
+Similarly, for confidentiality, we will want to compare pairs of future states,
+so we give a binary equivalent, \(d \Uparrow R\), where
+\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
+first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\),
+respectively. Once again, \(\Uparrow\) is coinductive.
+
+\judgmentthree[Returned]
+              {\(|\sigma_1| < d\)}
+              {\(|\sigma_2| < d\)}
+              {\(\mach_1 ~ R ~ \mach_2\)}
+              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+
+\judgmenttwobrlong[Left]
+                  {\(|\sigma_1| \geq d\)}
+                  {\((\mach_1,(V_1,\sigma_1)) \stepstounder{\overline{\psi}} (\mach_1',\context_1',\obs)\)}
+                  {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+                  {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+
+\judgmenttwobrlong[Right]
+                  {\(|\sigma_2| \geq d\)}
+                  {\((\mach_2,(V_2,\sigma_2)) \stepstounder{\overline{\psi}} (\mach_2',\context_2',\obs)\)}
+                  {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
+                  {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
+
+\subsection{Properties}
+\label{app:props}
+
+[TODO: back-propagate abbreviations used here]
+
+Finally, with the preliminaries out of the way, we can define our core properties. We write
+``when \(s\) is called'' as a shorthand for quantifying over all \(s\) reachable in any execution
+such that \(s \stepstounder{\overline{\psi}} s'\) and
+\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\).
+
+Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
+such that \(\mach[\component] \not = \mach'[\component]\).
+The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
+is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
+This treats \(\mach\) and \(\nach\) as the initial states and
+\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
+set of elements that were corrupted between them.
+
+\begin{table*}
+  \begin{align*}
+    \mathit{WBCF} \triangleq & |\sigma| \uparrow \ret
+    && \textnormal{ where } \ret ~ \mach' \textnormal{ iff }
+    \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}
+    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called} \\
+    %
+    \mathit{ClrI} \triangleq & (|\sigma| \uparrow \intProp) ~ (\mach,(V,\sigma))
+    && \textnormal{ where } \intProp ~ \mach' \textnormal{ iff }
+    \mach' \parallel \sealed(V) \cap \Delta(\mach,\mach')
+    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called} \\
+    %
+    \mathit{ClrC} \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\sealed(V)} \nach, &&
+    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
+    & |\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma)) \\
+    & \textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))
+    && \textnormal{ where } (\mach' ~ \confProp ~ \nach') \textnormal{ iff }
+    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach') \\
+    %
+    \mathit{CleC} \triangleq & (|\sigma| \uparrow \cconfProp) ~ (\mach,(V,\sigma))
+    && \textnormal{ where } \cconfProp ~ \mach' \textnormal{ iff }
+    \mach' \parallel \unsealed(V) \cap \Delta(\mach,\mach')
+    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
+    %
+    \mathit{CleI} \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\unsealed(V)} \nach, &&
+    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
+    & |\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma)) \\
+    & \textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))
+    && \textnormal{ where } (\mach' ~ \cintProp ~ \nach') \textnormal{ iff }
+    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach') \\
+  \end{align*}
+  \caption{Properties}
+  \label{tbl:props}
+\end{table*}
 
@@
 %annotations are set for the entry or return points.% At the same time,
-%%we allow the generation to sometimes relax those constraints,
+%%we allow the generation to sometimes relax constraints,
 %%introducing potentially ill-formed flows: this causes our programs
@@
 
-\subsection{Events and Traces}
-\label{sec:events}
-
-We abstract over the events that can be observed in the system, defining them
-only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
-event. Other events might represent certain function calls (i.e., system calls)
-or writes to special addresses representing mmapped regions.
-A {\em trace} is a nonempty, finite or infinite sequence
-of events, ranged over by \(\obsT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
-for list-cons.
-
-We write that execution from a state produces an observation trace
-\(\mach,\context \hookrightarrow \obsT\) as follows, coinductively:
-
-\judgmenttwo{\((\mach,\context) \stepstounder{} (\mach',\context',\obs)\)}
-            {\((\mach',\context') \hookrightarrow \obsT\)}
-            {\((\mach,\context) \hookrightarrow \notfinished{\obs}{\obsT}\)}
-
-We define another relation that takes a trace until we have returned from the
-active principal.
-We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
-\(d\) is the depth of the current call.
-
-\judgment{\(|\sigma| < d\)}
-         {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \tau\)}
-
-\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
-                  {\(|\sigma| \geq d\)}
-                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
-                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{\obs}{\obsT}\)}
-
-\paragraph*{Observational Similarity}
-
-We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
-written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
-is the same. That is, we compare up to deletion of \(\tau\) events.
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{}{\(\obsT \eqsim \obsT\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
-\end{minipage}
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
-\end{minipage}
-
-\paragraph*{Variants and Irrelevant Values}
-
-For a given system, assume an compatibility equivalence relation \(\sim\) on values.
-Two states are variants with respect to a set of elements \(\components\)
-if they agree on the value of every element not in \(\components\) and
-and have compatible values for every element in \(\components\).
-Our notion of non-interference involves comparing the traces of such
-\(\components\)-variants. We use this to define sets of irrelevant elements,
-and secrets in general.
-
-\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
-and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
-
-\definition An element set \(\components\) in state \((\mach,\context)\) contains
-irrelevant values, written \((\mach,\context) \parallel \components\), if for all
-\(\nach\) such that \(\mach \approx_{\components} \nach\), if
-\((\mach,\context) \hookrightarrow \obsT\) and
-\((\nach,\context) \hookrightarrow \obsT'\), then
-\(\obsT \eqsim \obsT'\).
-
-\subsection{Properties at Return}
-\label{sec:facts}
-
-Here we define some logical operations to reason about the behavior of the
-system over time. These have a temporal-logic flavor, as they reflect
-the expected behavior of the system in the future, after a possible return.
-
-\paragraph*{On-return}
-
-The intuition behind integrity (below) is that a caller may expect its
-sealed data to be unchanged when control returns to it. In fact, the callee
-may overwrite such data---when the data are found in callee-saved registers
-this is perfectly legal---as long as it either restores it, or has some guarantee
-that its changes will not impact the caller.
-
-We start by defining a second-order logical operator
-\(d \uparrow P\), pronounced ``\(P\) holds on return from depth \(d\),''
-where \(P\) is a predicate on machine states. This is a coinductive relation
-similar to ``weak until'' in temporal logic---it also holds if the program never
-returns from depth \(d\).
-
-\judgmenttwo[Returned]
-            {\(|\sigma| < d\)}
-            {\(P ~ (\mach,(V,\sigma))\)}
-            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
-
-\judgmenttwobrlong[Step]
-                  {\(|\sigma| \geq d\)}
-                  {\((d \uparrow P) ~ (\mach', \context')\)}
-                  {\((\mach, (V,\sigma)) \stepstounder{\overline{\psi}} (\mach', \context',\obs)\)}
-                  {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
-
-Similarly, for confidentiality, we will want to compare pairs of future states,
-so we give a binary equivalent, \(d \Uparrow R\), where
-\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
-first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\),
-respectively. Once again, \(\Uparrow\) is coinductive.
-
-\judgmenttwobrlong[Returned]
-            {\(|\sigma_1| < d\)}
-            {\(|\sigma_2| < d\)}
-            {\((\mach_1,(V_1,\sigma_1)) ~ R ~ (\mach_2,(V_2,\sigma_2))\)}
-            {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-
-\judgmenttwobrlong[Left]
-              {\(|\sigma_1| \geq d\)}
-              {\((\mach_1,(V_1,\sigma_1)) \stepstounder{\overline{\psi}} (\mach_1',\context_1',\obs)\)}
-              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-
-\judgmenttwobrlong[Right]
-              {\(|\sigma_2| \geq d\)}
-              {\((\mach_2,(V_2,\sigma_2)) \stepstounder{\overline{\psi}} (\mach_2',\context_2',\obs)\)}
-              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
-              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
-
-\subsection{Properties}
-\label{app:props}
-
-\paragraph*{Well-bracketed Control Flow}
-
-\definition We define a correct return from \(\mach\) as:
-%
-\[\ret(\mach) = \{(\mach',\context') \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}\]
-
-\definition
-A system enjoys  well-bracketed control flow if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) where
-\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\),
-\(|\sigma| \uparrow \ret(\mach)\) holds on \((\mach',(V,\sigma))\).
-
-\paragraph*{Integrity}
-
-\definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
-such that \(\mach[\component] \not = \mach'[\component]\).
-
-\definition We define \(\intProp(\mach,V)\), to hold on a state \((\mach',\context')\) when
-\(\{\component \mid V ~ \component = \sealed\} \cap \Delta(\mach,\mach')\)
-is irrelevant in \((\mach', \context')\).
-
-\definition A system enjoys {\it stack integrity} if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
-\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
-we have:
-\[(|\sigma'| \uparrow \intProp(\mach',V)) ~ (\mach',(V,\sigma))\]
-
-\paragraph*{Caller Confidentiality}
-
-\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
-\(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
-Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to \(\components\)
-if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
-\(|\sigma| \uparrow (\mach,\context) \hookrightarrow \obsT\) and
-\(|\sigma| \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
-\(\obsT \simeq \obsT'\).
-
-\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
-is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
-This treats
-\(\mach\) and \(\nach\) as the initial states and
-\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
-set of elements that were corrupted between them.
-
-\definition We define the relation \(\conf(\mach,\nach)\) to hold on a pair of states
-\((\mach',\context)\) and \((\nach',\context')\) when
-\(\bar{\Diamond}(\mach,\nach,\mach',\nach')\) is irrelevant in \((\mach',\context)\).
-
-\definition A state \((\mach,(V,\sigma))\) enjoys {\it return-time confidentiality}
-if, for \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\)
-and any \(\nach\) that is a \(\components\)-variant of \(\mach\), we have:
-\[(\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \conf(\mach,\nach) ~ (\nach,(V,\sigma))\]
-
-\definition A system enjoys {\it caller confidentiality} if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-\((\mach',\context')\) enjoys both internal and return-time confidentiality.
-
-\paragraph*{Callee Confidentiality}
-
-\definition We define \(\cconf(\mach,V)\), to hold on a state \((\mach',\context')\) when
-\(\{\component \mid V ~ \component = \unsealed\} \cap \Delta(\mach,\mach')\)
-is irrelevant in \((\mach', \context')\).
-
-\definition A system enjoys {\it callee confidentiality} if, whenever
-\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\)
-with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-% or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
-we have:
-\[(|\sigma| \uparrow \cconf(\mach',V)) ~ (\mach',\context')\]
+\subsection{Provenance, Capabilities, and Protecting Objects}
+\label{app:ptr}
+
+What if we want to express a finer-grained notion of safety, in which
+stack objects are protected unless the function that owns them intentionally
+passes a pointer to them? This can be thought of as a {\it capability}-based
+notion of security. Capabilities are unforgeable tokens that grant access to
+a region of memory, typically corresponding to valid pointers to that region.
+So, in order to express such a property, we need our machine to carry some notion
+of {\it pointer provenance}---a distinction between a pointer that is intended to
+point to a given object, and non-pointer integers as well as pointers to other objects.
+
+One such provenance model is Memarian et al.'s PVI \cite{}, in which pointers are
+annotated with the identity of the object they first pointed to. This annotation is
+propagated when the pointer is copied and when operations are performed on it---even
+integer-only operations.
+
+We can model this as a trio of additional security-relevant operations: one which
+declares a register to contain a valid pointer, one which transmits the provenance
+of a pointer from one element to another, and one which clears the provenance
+(for instance, when a pointer is modified in place in a way that makes it invalid.)
+
+In addition to the normal call stack, our security context will carry a map \(\rho\) from
+elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
+Existing operations are extended to keep the value of \(\rho\) the same, and the new operations
+work as follows:
+
+\judgmentbr[Promote]
+           {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})\)}
+           {\(\rho' = \rho[\reg_{dst} \mapsto \mathit{range} ~ \reg_{base} ~ \mathit{off} ~ \mathit{sz}]\)}
+           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+
+\judgmentbr[Propagate]
+           {\(\psi = \mathbf{propagate} ~ \component_{src} ~ \component_{dst}\)}
+           {\(\rho' = \rho[\component_{dst} \mapsto \rho[\component_{src}]]\)}
+           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+
+\judgment[Clear]
+         {\(\psi = \mathbf{clear} ~ \component\)}
+         {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
+
+We now have a notion of provenance, and must integrate it into the definition of
+stack safety. We essentially generalize the above notion of passing: we will consider
+a caller to have intentionally passed an object if that object is reachable by
+a capability that has been passed to the callee. This includes capabilities passed
+indirectly, by being stored in an object that is in turn passed. Formally, we call
+this set \(\mathit{capped}\), and define it recursively:
+%
+\[\begin{split}
+& \mathit{capped} ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{\component\} \cup \mathit{capped} ~ \components' ~ \rho \textnormal{ where} \\
+& \components' = \{\component' |\rho[\component] = (\mathit{base},\mathit{bound})
+\land \mathit{base} \leq \component' < \mathit{bound}\} \\
+\end{split}\]
+
+We then tweak the call operation to seal only objects that are in \(\mathit{capped}\), or
+the previously defined \(\mathit{passed}\).
+
+\judgmentbrbrbr[]
+               {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
+               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(\components = \{\addr | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \cup (\mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \}\)}
+               {\(V'' = V'\llbracket \components \mapsto \sealed \rrbracket\)}
+               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) =
+                 (V'',V::\sigma,\rho)\)}
+
+Note that we have a degree of monotonicity here. Once an object is sealed (because its
+capability has not been passed to a callee), subsequent nested calls can never unseal it.
+On the other hand, an object that is passed via a pointer may be passed on indefinitely.
 

### 2023-01-21 submission5/core.tex
Next iteration of the reorg

@@
 
-Security classes are ranged over by \(C\).
-For some security class \(C\), we write \(C(V)\)
-to denote the set of elements \(\component\) such that \(V ~ \component = C\).
+Security classes are ranged over by \(l \in \{\public, \object, \sealed, \unsealed\}\).
+For some security class \(l\), we write \(l(V)\)
+to denote the set of elements \(\component\) such that \(V ~ \component = l\).
 The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
@@
 rules for each operation. We have already seen the intuition behind the rules for
-\(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\). Their precise behavior
-is found in \cref{fig:basicops}.
-
-\begin{figure}
-\judgmentbr[Alloc]
-           {\(b = \mach[\SP] + \mathit{off}\)}
-           {\(V' = V \llbracket \addr \mapsto \object |
-             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
-           {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-
-\judgmentbr[Dealloc]
-           {\(b = \mach[\SP] + \mathit{off}\)}
-           {\(V' = V \llbracket \addr \mapsto \unsealed |
-             b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
-           {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-
-\judgmentbr[Call]
-           {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}}\)}
-           {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-             \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-           {\(Op ~ \mach ~ \psi ~ (V,\sigma) =
-             (V',V::\sigma)\)}
-
-\judgment[Ret]
-         {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-         {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
-
-         \caption{The basic operations}
-         \label{fig:basicops}
-\end{figure}
-
-\subsection{Advanced Operations}
-
-The advanced machine defines additional operations, and extends some of the old ones.
+\(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\).
+The full machine defines additional operations, and extends some of the old ones.
 We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
@@
 gains a boolen flag indicating whether the allocation is to be public, accessible to
-all functions until deallocated.
+all functions until deallocated. The rules are given in \cref{fig:advops}; the
+example is equivalent to these rules with \(\mathbf{call}\) always instantiated with
+\(\overline{sa}\) as the empty set, and \(\mathbf{alloc}\) with \(\mathbf{f}\).
+
+\subsection{Events and Traces}
+\label{sec:events}
+
+We abstract over the events that can be observed in the system, defining them
+only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
+event. Other events might represent certain function calls (i.e., system calls)
+or writes to special addresses representing mmapped regions.
+A {\em trace} is a nonempty, finite or infinite sequence
+of events, ranged over by \(\obsT\).
+We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
+for list-cons.
+
+We write the trace of execution from a state up to the first
+state below depth \(d\) via the \(\hookrightarrow\) operator, defined coinductively.
+When \(d = 0\), the trace will always be infinite. In this case we
+omit \(d\) and just write \(\hookrightarrow s\).
+
+\judgment{\(|\sigma| < d\)}
+         {\(d \hookrightarrow (\mach,(V,\sigma)) = \tau\)}
+
+\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
+              {\(|\sigma| \geq d\)}
+              {\(d \hookrightarrow (\mach',\context') = \obsT\)}
+              {\(d \hookrightarrow (\mach,(V,\sigma)) = \notfinished{\obs}{\obsT}\)}
+
+Two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
+written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
+is the same. That is, we compare up to deletion of \(\tau\) events.
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{}{\(\obsT \eqsim \obsT\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
+\end{minipage}
+
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
+\end{minipage}
 
@@
 
-The advanced machine supports public stack allocations that can
+The full machine supports public stack allocations that can
 be accessed by anyone until they are deallocated on their owner's return,
@@
 
-    \judgmentbr[]
-               {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
                  \{\addr \mid V ~ \addr = \unsealed\}\)}
@@
                {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-    \caption{Allocating Public Regions}
+    %
+    \judgmentbr{\(b = \mach[\SP] + \mathit{off}\)}
+               {\(V' = V \llbracket \addr \mapsto \unsealed |
+                 b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+
+    \caption{Memory Allocation}
     \label{sfig:publicalloc}
@@
     %
-    \judgmentbr[]
-               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
                  \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
@@
                  ~ (V,\sigma) = (V',\sigma)\)}
-    \caption{Argument Passing on Stack}
+    %
+    \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
+             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
+
+    \caption{Argument Passing on }
     \label{sfig:stkargs}
   \end{subfigure}
-  \caption{Advanced Operations}
+  \caption{Operations}
   \label{fig:advops}
@@
 
-\subsection{Events and Traces}
-\label{sec:events}
-
-We abstract over the events that can be observed in the system, defining them
-only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
-event. Other events might represent certain function calls (i.e., system calls)
-or writes to special addresses representing mmapped regions.
-A {\em trace} is a nonempty, finite or infinite sequence
-of events, ranged over by \(\obsT\).
-We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
-for list-cons.
-
-We write the trace of execution from a state up to the first
-state below depth \(d\) via the \(\hookrightarrow\) operator, defined coinductively.
-When \(d = 0\), the trace will always be infinite. In this case we
-omit \(d\) and just write \(\hookrightarrow s\).
-
-\judgment{\(|\sigma| < d\)}
-         {\(d \hookrightarrow (\mach,(V,\sigma)) = \tau\)}
-
-\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
-              {\(|\sigma| \geq d\)}
-              {\(d \hookrightarrow (\mach',\context') = \obsT\)}
-              {\(d \hookrightarrow (\mach,(V,\sigma)) = \notfinished{\obs}{\obsT}\)}
-
-Two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
-written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
-is the same. That is, we compare up to deletion of \(\tau\) events.
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{}{\(\obsT \eqsim \obsT\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
-\end{minipage}
-
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
-           {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
-\end{minipage}
-
-\paragraph*{Variants and Irrelevant Values}
-
-For a given system, assume an compatibility equivalence relation \(\sim\) on values.
+For a given system, we assume an compatibility equivalence relation \(\sim\) on values.
 Two states are variants with respect to a set of elements \(\components\)
@@
 
-\subsection{Properties at Return}
-\label{sec:facts}
-
-Here we define some logical operations to reason about the behavior of the
-system over time. These have a temporal-logic flavor, as they reflect
-the expected behavior of the system in the future, after a possible return.
-
-\paragraph*{On-return}
-
-The intuition behind integrity (below) is that a caller may expect its
-sealed data to be unchanged when control returns to it. In fact, the callee
-may overwrite such data---when the data are found in callee-saved registers
-this is perfectly legal---as long as it either restores it, or has some guarantee
-that its changes will not impact the caller.
-
-We start by defining a second-order logical operator
+\definition \(\Delta(\mach,\mach')\) is the set of elements \(\component\)
+such that \(\mach[\component] \not = \mach'[\component]\).
+
+\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
+is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
+
+Conceptually, the corrupted set treats \(\mach\) and \(\nach\) as initial states whose
+and \(\mach'\) and \(\nach'\) as their corresponding final states, such that any element
+that has changed from its initial value but does not match between the final states
+is considered to have propagated some varied data from the original pair.
+
+Our ``on-return'' assertions are defined using a second-order logical operator
 \(d \uparrow P\), pronounced ``\(P\) holds on return from depth \(d\),''
@@
 
-Similarly, for confidentiality, we will want to compare pairs of future states,
-so we give a binary equivalent, \(d \Uparrow R\), where
+Similarly, we give a binary equivalent for use in confidentiality. We define \(\Uparrow\) so that
 \((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
@@
 
-\subsection{Properties}
-\label{app:props}
-
 [TODO: back-propagate abbreviations used here]
 
-Finally, with the preliminaries out of the way, we can define our core properties. We write
-``when \(s\) is called'' as a shorthand for quantifying over all \(s\) reachable in any execution
-such that \(s \stepstounder{\overline{\psi}} s'\) and
-\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\).
-
-Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
-such that \(\mach[\component] \not = \mach'[\component]\).
-The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
-is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
-This treats \(\mach\) and \(\nach\) as the initial states and
-\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
-set of elements that were corrupted between them.
-
-\begin{table*}
+Finally, we can define our core properties. They can be found in \cref{tbl:props},
+arranged to show their commonalities and distinctions. Each definition gives a criterion
+quantified over states \(s\) that immediately follow call (or tail-call) steps.
+If an execution can reach a state \(s\) such that \(s \stepstounder{\overline{\psi}} s'\)
+where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s'\) is the target
+of a call. Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
+\(s'\) is the target of a tailcall. As a shorthand, we write that each property is defined
+by a criterion that must hold ``when \(s'\) is called,'' i.e., for every such \(s'\).
+
+\(\wbcf\), \(\clri\), and \(\clec\) are each straightforward: they define predicates that must
+hold on return from the call. \(\clrc\) and \(\clei\) both compare the execution of the
+call state with that of a variant state, and are concerned with both their internal and
+their return-time behavior. So, for some \(\nach\) that is an appropriate variant,
+all states must fulfill an ``internal'' clause that their traces are similar,
+and a ``return-time'' clause, that any potentially corrupted memory at their returns
+is irrelevant.
+
+\begin{table*}[h]
   \begin{align*}
-    \mathit{WBCF} \triangleq & |\sigma| \uparrow \ret
+    \wbcf \triangleq & |\sigma| \uparrow \ret
     && \textnormal{ where } \ret ~ \mach' \textnormal{ iff }
@@
     %
-    \mathit{ClrI} \triangleq & (|\sigma| \uparrow \intProp) ~ (\mach,(V,\sigma))
+    \clri \triangleq & (|\sigma| \uparrow \intProp) ~ (\mach,(V,\sigma))
     && \textnormal{ where } \intProp ~ \mach' \textnormal{ iff }
@@
     %
-    \mathit{ClrC} \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\sealed(V)} \nach, &&
+    \clrc \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\sealed(V)} \nach, &&
     && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
@@
     %
-    \mathit{CleC} \triangleq & (|\sigma| \uparrow \cconfProp) ~ (\mach,(V,\sigma))
+    \clec \triangleq & (|\sigma| \uparrow \cconfProp) ~ (\mach,(V,\sigma))
     && \textnormal{ where } \cconfProp ~ \mach' \textnormal{ iff }
@@
     %
-    \mathit{CleI} \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\unsealed(V)} \nach, &&
+    \clei \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\unsealed(V)} \nach, &&
     && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
@@
 \end{table*}
+
+[TODO: words about the complementary properties.]
 
@@
 To use QuickChick, we build random test-case generators that produce
-% \begin{enumerate}
-% \item
 (1) an initial RISC-V machine state, including most notably
   the program to be executed;
-  % \item
 (2)
@@
   all potential stack locations as $\tagNoDepth$; and
-% \item
 (3)
- a code map identifying the locations of calls.
-% \end{enumerate}
+  a map from instructions to lists of their operations in the security semantics.
 
@@
 a positive and formal characterization of stack safety
-is that of Skorstengaard et al. \cite{SkorstengaardSTKJFP}. Their
+begins with Skorstengaard et al. \cite{SkorstengaardSTKJFP} and is improved
+by Georges et al. \cite{Georges22:TempsDesCerises}. The former's
 StkTokens system enforces stack safety on hardware equipped with
 memory capabilities, similar to those used in the Cheri system~\cite{Woodruff+14,Chisnall+15}, but
-with additional support for linear capabilities.
-They define Stack safety as the conjunction of local state encapsulation (LSE) and
+with additional support for linear capabilities, while the latter offer a calling
+convention that can replace the expensive and likely-impractical linear capabilities
+with more efficient ``directed'' capabilities.
+
+Both works define stack safety as the conjunction of local state encapsulation (LSE) and
 well-bracketed control flow (WBCF).
@@
 concrete capability machine, an approach termed ``fully abstract overlay semantics.''
-We improve upon Skorstengaard et al. \cite{SkorstengaardSTKJFP} in two significant ways:
-
-i) Our definition captures the intuition that {\em
-  every} caller should be protected from its callees.  By contrast,
-  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
-  that are not protected, because they are capable of compromising
-  their own security. Consider a function that, when called with the wrong
-  inputs, behaves unreasonably. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
-  {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
-  component is always reasonable. In our setting, individual
-  well-behaved activations of the buggy function would be protected.
-
-ii) Skorstengaard et al. \cite{SkorstengaardSTKJFP}'s fully
-  abstract overlay semantics rely on the definition of a ``stack-safe-by-construction''
-  abstract machine that is closely tied to the underlying machine architecture.
-  While the approach can no doubt be ported to other architectures and enforcement
-  schemes, each such port will require a new abstract machine definition whose
-  safety-by-construction must be argued afresh.
-  Since we abstract over most of the machine state as well as the sealing and return conventions, and
-  can support further enhancements such as coroutines,
-  our properties apply directly to a range of architectures, code features, and enforcement mechanisms.
-
-Skorstengaard et al. \cite{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
-candidate stack safety properties.
-%% : they should be (1) intuitive; (2)
-%% useful for reasoning; (3) reusable in secure compiler chains; (4)
-%% arguably ``complete'' with respect to informal notions of stack
-%% safety; and (5) potentially scalable to richer settings such as
-%% dynamic code generation or multithreading.
-%
-In terms of (1) {\em intuitiveness}, we contend that our property
-definitions are at least as
-intuitive as theirs, though phrased in rather different terms.
-We use the language of formal security
-properties to apply well-established concepts \ifspace and formal techniques \fi
-of noninterference, and we phrase our properties in terms of
-traces of memory and register states, rather than an
-ad-hoc abstract machine.
-Which approach is ``better'' is probably a matter of taste, but ours
-has \ifspace at least \fi the merit of explicitly relating stack safety to existing
-notions of security.
-%
-Demonstrating (2) {\em usefulness} of the safety criterion
-for carrying out formal proofs (about individual programs) for our proposal
-remains future work; concretely, we
-hope to apply our properties to justifying a low-level program logic.
-%
-(3) StkTokens is specifically set up to address {\em reusability} in secure
-compilation chains, while this is not a goal of our work.\bcp{Can we explain
-what it means??}
-%
-(4) Regarding {\em comprehensiveness} in capturing programmers' informal
-understanding of stack safety, we believe that our support
-for a dynamically changing notion of ``adversary,'' as outlined in point (i) above,
-captures an even better definition of safety than LSE; moreover, our observational
-properties are richer and\bcp{cut "richar and"??} more extensional, and thus
-applicable to a broader range of enforcement mechanisms and code features.
-%
-(5) Finally, regarding {\em broad applicability}, we have demonstrated that
-\bcp{unlike StkTokens}
-our approach extends to
-coroutines, a first step towards multithreading. Dynamic code generation is
-more challenging, but StkTokens does not yet handle this either.
-
-We would add one more criterion: reusability across machine models and
-enforcement mechanisms. StkTokens has been used to validatate
-one machine/enforcement mechanism---a linear-capability machine---via formal
-proof. We have used ours to obtain a weaker (testing-based) validation of
-two different mechanisms on a single machine; we hope to target a capability machine
-next as future work.
-
-\paragraphx{Protecting the Stack with Micro-Policies}
-%
-\apt{Could probably condense/omit this.}
-Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}
-study the protection of stack data through three families of
-micro-policies that tag stack objects with a pair of frame and object identifiers
-used to validate accesses: {\em Return Address Protection} (which
-prevents an adversary from overwriting designated return addresses), {\em Static
-Authorities} (which only allows the code of a function to access the stack frames
-of its own dynamic instances), and {\em Depth Isolation}
-(\cref{sec:enforcement}). All these policies strike various balances between
-precision and performance, evaluated through benchmarks that demonstrate
-limited performance overhead. In addition to the baseline ``eager'' policies,
-they propose a number of lazy optimizations, notably {\em Lazy Tagging} and
-{\em Lazy
-Clearing} (\cref{sec:lazy}).
-
-We evaluate {\em Depth Isolation} and {\em Lazy Tagging and Clearing}
-against our properties, finding that the former does
-indeed enforce stepwise stack safety and therefore observable stack
-safety. We find that the latter does not actually enforce
-stack safety, and we proposed a fix.
+
+[TODO: more detailed comparison with Aina's work primarily.]
 
@@
 the language of noninterference. Their paper was a significant inspiration
-for the present work, and we believe that the two safety conditions should
-combine smoothly.
+for the present work, and we believe that their safety conditions should
+combine smoothly into ours.
 
@@
 
-From here, we see two natural ways to extend the work. \apt{what about addressing some of the limitations?}
+From here, we see two natural ways to extend the work.
 First, we should thoroughly evaluate
@@
 \cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
-\cite{Tsampas+19}, and monotonic \cite{GeorgesComplete} capabilities.
-Of these, uninitialized capabilities seem particularly promising---they naturally enforce
-write-before-read behavior and therefore will enable confidentiality without the need
-to clear large swaths of the stack. We plan to implement a model of the stack safety
-system described by Georges et al. \cite{Georges+21} and test it against our properties using
-QuickChick.
+\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
+Of these, the last seems particularly promising for testing.
+[TODO: more about this. (Rob's suggestion of citing Aina could go here.)]
 

### 2023-01-21 submission5/core.tex
Checkpoint: intro comments

@@
 
+\rb{\textbf{The call stack and its security.} Maybe a few opening words to set
+  the stage and bring two things to the front: (1) functions and subroutines as
+  fundamental building blocks of any kind of structured programming; and (2) the
+  stack as the natural abstraction to manage these in a computer. And also maybe
+  a little bit more about the width, the gravity and the \emph{foundational
+  nature} of attacks on the stack, which the sentence below touches on
+  summarily.}
+%
 The call stack is a perennial target for low-level attacks, leading to
 dire consequences from leakage and corruption of private stack data
-to control-flow hijacking. To foil such attacks, a menagerie of
+to control-flow hijacking.
+%
+\rb{We can mention things like e.g. the most recent CWE Top 25 Most Dangerous
+    Software Weaknesses from MITRE, cf.
+    \url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}, where
+    e.g. \#1 (Out-of-bounds Write), \#5 (Out-of-bounds Read) and \#7 (Use After
+    Free), that is a good number of the most important and prevalent
+    vulnerability types are closely related to the stack. These will show up in
+    other popular rankings as well.}
+%
+To foil such attacks, a menagerie of
 software and hardware protections have been proposed,
@@
 \fi
+%
+\rb{Did we want to add even more techniques? Is this selection broad enough?
+  Anyway reference that we'll be looking at these more closely later.}
 
@@
 addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more formal characterization to the reader's intuition. Previous attempts
+a more formal characterization to the reader's intuition.
+%
+\rb{Begins to anticipate subsequent discussion. This is even something of an
+  understatement: some papers (esp. newer ones?), make concrete claims, though
+  they remain informal. Arguably, none have a truly formal characterization in
+  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
+  of the novelties of this work!}
+%
+Previous attempts
 to produce a formal characterization do so using a {\it fully abstract overlay semantics},
@@
 property suitable for testing lazy enforcement.
+%
+\rb{Need to discuss overlay semantics, but also note that it is very intensional
+  and very specific, as the authors acknowledge. I'd introduce this a bit more
+  gradually. Also, if phrased this way, it sounds like its main drawback is a
+  lack of generality, to which we should contrast a more general approach, not
+  just one that is specialized to a specific policy, i.e. lazy micro-policies
+  (I'd try to present these more gradually as well).}
 
@@
 in which there may be multiple points at which a callee can ``return.''
+%
+\rb{Connect this to the shortcomings of previous work. I'd give at least a quick
+  explanation of what the various properties roughly entail in practice (e.g.,
+  secrecy of private data, etc.). The comment about the single extension of
+  exceptions seems disconnected from the rest.}
 
@@
 {\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a RISC-V specification.  We
+in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
 use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
@@
 of our testing.
+%
+\rb{This and the following paragraph could start more generally (we need to
+  answer the question of why these experiments, what is their interest), and
+  probably motivate the use of PBT (experiment results below show there's
+  benefit in them!) and, somewhere at least, talk about other points in the
+  spectrum.}
 
@@
 confidentiality, albeit at some performance cost.
+%
+\rb{Anything to add from personal communications? More details later? Shows
+  these things are very hard to get right -- like the latest work on Cerise,
+  also motivated by this very work.}
 
@@
 not fulfill the property benefit from the ability to articulate {\it why} they fail.
+%
+\rb{Good stuff! There is some overlap between this discussion and the opening
+  motivation, maybe we could restructure things a bit so that those threads are
+  unified?}
 
@@
 must be accompanied by tweaks to the model.
+%
+\rb{This point is extremely important!}
 
@@
 that are parameterized by (1).
+%
+\rb{Sharpen, contrast the universal definitions (or a framework of parametric
+  definitions, say) with the concrete instance we study, again with its interest
+  properly motivated.}
 
@@
 callee, etc.
+%
+\rb{Also sharpen after extended discussion?}
 
@@
 observational properties are even weaker than necessary to support Roessler and Dehon's micro-policy.
+%
+\rb{Does this discussion (at least in part) belong in the general background
+  section, some even in key ideas? It still feels a little out of order to me,
+  and occasionally too detailed for the introduction?}
 
@@
 QuickChick library and to ensure that the formalizations are unambiguous.
+%
+\rb{Will have another look after the full pass.}
 
@@
 the property vacuously.
+%
+\rb{Limitations in the introduction or later? Orthogonality of the heap, morally
+  similar interesting setups on the stack? Termination sensitivity and
+  capability machine semantics?}
 

### 2023-01-21 submission5/core.tex
An attempt to classify stack security policies

@@
   and occasionally too detailed for the introduction?}
+
+\rb{\textbf{A menagerie of protection mechanisms.}
+%
+We have made explicit the key point that there is \emph{not} a single stack
+abstraction. This paves the way for a \emph{general} idea of stack protection
+that is widely applicable with high-level concepts that remain unchanged (and I
+think it's particularly important to at least introduce these more immutable
+parts early on). The fact that this general ``framework'' requires low-level or
+specific specializations is a predictable consequence of the protean nature of
+the call stack.
+%
+\\
+%
+Let us return to a selection of representative policies in our menagerie and
+consider how they fit into the framework. At the highest level, we can classify
+all those techniques into two large groups, software-based and hardware-based.
+Traditionally, most practical protections are implemented in software, so we
+start by considering this first group.
+%
+\\
+%
+We can make two general observations about these software based techniques.
+Firstly, (as we will see: discussion? technical details?) these techniques are
+\emph{less modular} than their hardware counterparts. That is, the
+countermeasures are tightly coupled with the program code they protect, and
+consequently more effort is required to disentangle the program from the
+protections, and to understand the impact of the latter on the former. This is
+true of any systematic attempt to study the security of stack protection
+policies, including their adaptation to our framework. (Owing to this added
+complexity, software policies will not be the cleanest or most elegant exemplars
+of use of the framework, its salient features and practical utility; this partly
+motivates the focus on hardware policies.)
+%
+\\
+%
+(The above remark holds \emph{at least in a ``single-level model''} that
+considers assembly language/machine code programs. In general, (low-level)
+protections are introduced by a compiler or similar tool. One could consider a
+multi-level extension of the framework pairing models of high-level programs and
+the security transformations performed by a (stack-secure) compiler.)
+%
+\\
+%
+Secondly, the guarantees generally offered by software-based policies are
+\emph{less absolute}, being defined more or less informally and usually
+providing few definite guarantees of absence of \emph{general} classes of
+errors. (The following discussion elaborates on this. It is conceivable that an
+extended and more sophisticated framework could add \emph{probabilistic
+reasoning} capabilities to capture and reason about the effects of less strict
+protection mechanisms. This would be another orthogonal development to what is
+presented here. In this paper, we are interested in assessing the complete
+\emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
+or defects. In passing, this would extend the security properties to e.g. larger
+system verification efforts.)
+%
+\\
+%
+One by one, consider the following specimens of software protection mechanisms
+(quick explanations inlined?):
+%
+\\
+%
+\begin{itemize}
+%
+\item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
+  difficulty of some attacks on the stack, usually (always?) buffer overflows.
+  They can never provide any absolute security guarantees and therefore do not
+  satisfy any of our properties (some of the properties, e.g. confidentiality,
+  are completely out of scope.) Although tedious, we could we could model this
+  mechanism in our framework with the appropriate adaptations --- and
+  unsurprisingly, testing would find property violations, which we already know
+  are there: the results of this effort would be unsurprising and not as useful
+  or instructive as other experiments.
+%
+\item \emph{Bounds checking}. This is fundamentally a high-level technique, even
+  more obviously so than stack canaries (which could still detect violations
+  caused by external code). Bounds checking only applies to code that has been
+  instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
+  impose any constraints on adversarial code. The attacker model is therefore
+  rather weak. (Certain variants, like SoftBound, are good targets for lifting
+  our framework to the source level by e.g. using a C-language semantics and
+  using Tagged C for security modeling; the lifted versions of our properties
+  would likely hold under the expected assumptions. HardBound is probably an
+  interesting transitional case because it uses capability-like hardware
+  primitives for enforcement; it could be close to ongoing and future work to
+  verify stack safety on capability machines, e.g. based on Cerise -- discuss
+  together?)
+%
+\item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
+  ``probabilistic techniques'' in that they ``restrict the flexibility available
+  in creating gadget chains'' (Shanbhogue et al.). This technique is solely
+  concerned with control-flow integrity, and so closely related to our
+  well-bracketed control flow. Nevertheless, and again, it is only an
+  approximate technique, and our testing framework would be able to find the
+  expected attacks against our security property.
+%
+\item \emph{Split stacks}. The security claims of split stacks are stronger than
+  those of most software countermeasures, i.e., they are meant to guarantee
+  their titular code-pointer integrity. They explicitly offer no protections on
+  stack data (i.e., integrity and confidentiality), and so restrict themselves
+  to well-bracketed control flow. (This is probably closer to our framework than
+  most other software techniques, and it seems like it would be easier to model
+  as well. However, this is again work based on high-level languages and would
+  benefit from a higher level of abstraction.)
+%
+\item \ldots
+%
+\end{itemize}
+%
+Opposed to the above group of software techniques we have hardware-based
+protection policies (more in background/technical preliminaries?). Although
+programs must effectively leverage the protection mechanisms offered by the
+hardware, the distinction between those protections and the program logic itself
+is much cleaner: it is fairly simple to separate the use of the security
+features in the hardware (and their effects in program code) from regular
+program code. They also tend to aspire to offer complete protection against
+entire classes of attacks. Both aspects make this family of techniques a better
+fit for illustrating and showcasing our techniques.
+%
+\\
+%
+Why favor micro-policies? Their design is more flexible and arguably more
+general than capability machines (of course, they pay a performance price for
+this). This generality and ease of programming make them good candidates for an
+exploration the design space of stack protection policies and proof of concept
+implementations. On the practical side, there is a number of existing policies
+available to examine (not just lazy policies!) --- and we find how they satisfy
+our properties, how they do not, and errors in some of them, and see how to
+correct them. Lazy enforcement is particularly novel, but not the only
+possibility. This additional breadth is beneficial for the proof of concept
+stage.
+%
+\\
+%
+Regarding capability machines, we know that most designs based on capability
+machines do not satisfy our properties (or do so incurring significant costs),
+and that fixes are not trivial and involve new types of capabilities (in a
+sense, are they any more ``realistic?'' Only the latest Cerise paper would seem
+to satisfy our properties, is directly influenced by previous versions of this
+work, and is a reasonable target for ongoing/further study -- but this is not
+easy, and even the author(s) estimate it would be a significant amount of work,
+especially full proofs (references, communications, etc.).
+%
+\\
+%
+\ldots
+%
+}
 

### 2023-01-23 submission5/core.tex
Reorg and closer edits

--- 2023-01-21 submission5/core.tex
@@
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-during the callee's execution, the confidentiality of the callee's local state
-afterward, and the control flow protection of {\em well-bracketed control
-  flow (WBCF)}~\cite{SkorstengaardSTKJFP}. We also extend WBCF to a setting with exceptions,
+and the callee's behavior during the callee's execution, plus the control flow protection
+of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
+\ifexceptions
+We also extend WBCF to a setting with exceptions,
 in which there may be multiple points at which a callee can ``return.''
+\fi
+
+This gives us a total of five core properties: \(\wbcf\)
+caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
+callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
 
@@
 not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
-exclusively on control flow, others protect the confidentiality of the caller but not the
-callee, etc.
+exclusively on control flow, some enforce integrity but not confidentiality, and many
+focus on the caller over the callee.
 
@@
 {\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
-when a function writes to memory. This means that such writes cannot be protected. Instead
-the micro-policy enforces that memory be read by the same fuction that wrote it. This
-lazy enforcement is incompatible with existing models of stack safety,
-because an attacker can modify another function's data. But the owner of
-the data will never see the changes, so it is secure in the sense that the attacker is
-unsuccessful at changing program behavior (except, perhaps, by causing a fault.) We aim to elaborate
-that sense, and to test whether the micro-policy fulfills it.
+when a function writes to memory. This means that writes cannot be prevented, necessarily
+violating existing models of stack safety. Instead the micro-policy enforces that memory be
+read by the same fuction that wrote it. The owner of the data will never see the changes,
+so it is secure in the sense that the attacker is unsuccessful at changing program behavior
+(except, perhaps, by causing a fault.) We elaborate that sense, and test whether the
+micro-policy fulfills it.
 
@@
 That is not to say that our properties are specialized for lazy micro-policies. It also supports
-normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, our
-observational properties are even weaker than necessary to support Roessler and Dehon's micro-policy.
+normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
+and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
+while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
 
@@
 \begin{itemize}
-\item We formalize permissive stack confidentiality and integrity properties.
-  Both are parameterized over a notion of external
+\item We formalize permissive stack confidentiality and integrity properties,
+  which are parameterized over a notion of external
   observation, and are violated only if accessing secrets or overwriting
@@
 \item Our formalization of well-bracketed control flow in the presence of tail-call
-  elimination is a slight novelty over existing models.
+  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
 \item We use property-based random testing to validate the relationship between
@@
 are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In \cref{sec:machine}), we cover additional operations add parameters to the examples', to support
-various forms of memory sharing and tail-call elimination. These are shown in light gray.
+In \cref{sec:machine}), we cover additional operations that add parameters to the examples',
+to support various forms of memory sharing \ifexceptions , exceptions, \fi
+and tail-call elimination. These are shown in light gray.
 We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
@@
 
+\sna{Might like a better name than object, because morally it ought to apply to arguments as well.}
+
 \paragraph*{The Properties and their Hierarchy}
 
-We have five different properties to introduce:
-
-\begin{itemize}
-\item Well=bracketed Control Flow (\(WBCF\))
-\item Caller Integrity (\(CLRI\))
-\item Caller Confidentiality (\(CLRC\))
-\item Callee Confidentiality (\(CLEC\))
-\item Callee Integrity (\(CLEI\))
-\end{itemize}
-
-The properties form a rough hierarchy; although they are technically separate and need
-not coexist in any particular combination, some are likely to take priority in terms of their
+The properties form a hierarchy, shown in [TODO: diagram].
+Within properties that are orthogonal
+some are likely to take priority in terms of their
 importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
 as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
-Many also cover some degree of (caller) integrity. [TODO: more hierarchy].
+Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
+and callee-oriented properties even moreso.
 
@@
 mechanism may place additional constraints, particularly on the behavior of
-call and return sequences.
+call and return sequences. For instance, extant implementations tend to assume
+implicitly that callee-saved registers are treated appropriately by whichever compiler
+generated their code. Our properties explicitly require them to be treated as belonging
+to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
 
@@
 make a call, or make a return.
-Figure \ref{fig:exec1} shows how the corresponding security context
-develops over the first few steps of the program, beginning with
-the start of {\tt main} (\(\PCname = 0\)) and with the stack pointer at address 1000.
+
+Recall that our security semantics attach a security context to the machine state,
+which consists of a view \(V\) and a stack \(\sigma\) of pending activations' views.
+(See \cref{sec:machine} for details.)
+Figure \ref{fig:exec1} shows how the security context develops over the first few
+steps of the program, according to the rules in \cref{fig:advops}, slightly simplified.
+It begins at the start of {\tt main}, where the program counter (\(\PCname\)) is zero,
+and with the stack pointer (\(\SP\)) at address 1000.
 State transitions are numbered and labeled with a list of security operations, written
@@
 To begin with, what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
-\(\SP \neq 980\))? We consider this to
-violate well-bracketed control flow (WBCF). WBCF is a relationship between
+\(\SP \neq 980\))? We consider this to violate \(\wbcf\). \(\wbcf\) is a relationship between
 call steps and their corresponding return steps: just after the return, the program
@@
 
-To see why WBCF also cares about the stack pointer, consider the attack in
+To see why \(\wbcf\) also cares about the stack pointer, consider the attack in
 Figure \ref{subfig:WBCF2}. Here the attacker returns with \(\SP' = 998\) instead of the
@@
 
-In typical compiled code, WBCF is maintained by the compiler's call and return sequences.
-As long as a function deallocates the same-sized frame that it allocates, and properly
-saves and restores the return address, it should have WBCF---unless an error in the middle
-subverts those sequences by overwriting one of the relevant elements, or by jumping into
-the middle of the call or return sequence. We can therefore enforce WBCF by (1) ensuring that
-these ``blessed'' call and return sequences execute in order and (2) permitting only blessed
-sequences to write to \(\SP\), {\tt ra}, and the address in memory where {\tt ra} is stored.
-
-To perform this enforcement using tags, we can take
-\(\mathcal{T} ::= \mathit{gen} \mathrel{|} \mathit{bseq} ~ n ~ m \mathrel{|} \mathit{prot}\), named for general
-execution, blessed sequence, and protected, respectively. Blessed sequences of length \(m\) are
-tagged \(\mathit{bseq} ~ 1 ~ m\) through \(\mathit{bseq} ~ m ~ l\), and the micro-policy enforces
-that they execute in order. The stack pointer and {\tt ra}
-are tagged \(\mathit{prot}\). Only instructions tagged \(\mathit{bseq}\) may write to memory
-or registers tagged \(\mathit{prot}\). Store instructions tagged \(\mathit{bseq}\) tag their
-targets \(\mathit{prot}\).
-
 \paragraph*{Stack Integrity}
@@
 To characterize this situation in a property, we borrow the idea of \emph{variant} states
-from the theory of non-interference.
-For a set of elements \(\components\),
+from the theory of non-interference. For a set of elements \(\components\),
 a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
@@
 the machine's security enforcement mechanism) must be identical.
-
 We say that the elements of \(\components\) are \emph{irrelevant}
 in \(\mach\) if they can be replaced by compatible values without changing the
-observable behavior of the machine. (We will formalize this in Section~\ref{sec:events}
-by saying that every \(\components\)-variant state produces
-an equivalent observation trace.) All other elements are \emph{relevant}.
-We characterize \emph{observational integrity} as the property that
+observable behavior of the machine. All other elements are \emph{relevant}.
+Variants and relevance are defined in \cref{sec:props}.
+
+We characterize \(\clri\) as the property that
 every relevant element that is \(\sealed\) under the callee's view is restored
@@
 Figure \ref{subfig:direct}. We take a variant state over the set of element(s) that are
-\(\sealed\) or \(\unsealed\) in \(V_2\), which happens to be the entire stack (see
-Figure \ref{fig:variant2}.) If we take a trace of execution from each state until it returns,
+\(\sealed\) in \(V_2\) (see Figure \ref{fig:variant2}.)
+            [TODO: fix figure to avoid varying the entire stack.]
+If we take a trace of execution from each state until it returns,
 the traces may differ, in this case outputting 5 (the original value of {\tt secret}) and
 4 (its value in the variant) respectively. This is a violation of
-{\it internal confidentiality}.
+{\it internal confidentiality} (\cref{tab:props}, line 3a.)
 
@@
 extracted and placed in {\tt a0}, the return value register. We wish to identify this as
-a confidentiality violation, again by considering variants of the \(\sealed\) and \(\unsealed\)
+a confidentiality violation, again by considering variants of the \(\sealed\)
 elements in \(V_2\), but capturing the required property is subtle.
@@
 Unless {\tt a0} happens to be irrelevant, this is a violation of what
-we term {\it return-time confidentiality}.
+we term {\it return-time confidentiality} (\cref{tab:props}, line 3b.)
 
@@
 
-Finally, we define caller confidentiality as the
+Finally, we define \(\clrc\) as the
 combination of internal and return-time confidentiality.
-While we have separated the definitions of integrity and confidentiality,
-in terms of tag-based enforcement, Depth Isolation and Lazy Tagging and Clearing
-both enforce both properties.
-
-\paragraph*{Callee Confidentiality}
-
-Although we presented our initial example from the perspective of the caller, a callee
-may also have data that must be kept secret from its caller. Consider a callee that makes
-a privileged system call to obtain a secret key, and uses that key to perform a specific
-task. An untrustworthy or erroneous caller might attempt to read the key out of the callee's
-memory after return.
-
-We model this as nearly dual to caller integrity: the callee's secrets are those
-values that it has stored in \(\unsealed\) elements. So, callee confidentiality means that
-at the state following the callee's return, any relevant element that was \(\unsealed\) under
-the callee's view retains its original value.
-
-Once again, both tag-based approaches do enforce callee confidentiality, but it is
-worth keeping separate because some enforcement mechanisms might not, and it has not
-historically been acknowledged by all formal treatments of stack safety \cite{Georges22:TempsDesCerise}.
+
+\paragraph*{The Callee's Perspective}
+
+We presented our initial example from the perspective of the caller, a callee
+may also have privilege that its caller lacks, and which must be protected from the
+caller. Consider a function that makes a privileged system call to obtain a secret key,
+and uses that key to perform a specific task. An untrustworthy or erroneous caller might
+attempt to read the key out of the callee's memory after return, or to influence the callee
+to cause it to misuse the key itself!
+
+Where the caller's confidentiality and integrity are concerned with protecting specific,
+identifiable state---the caller's stack frame---their callee equivalents are concerned
+with enforcing the expected interface between caller and callee. Communication between
+the principals should occur only through the state elements that are designated for the
+purpose: those which we label \(\public\) and (in the presence of memory sharing) \(\object\).
+
+The properties themselves correspond to their caller equivalents, inverted.
 
@@
 
-The forms of stack safety described above are largely independent of one another, and
-can potentially be enforced separately in various combinations. This is important, because
+The forms of stack safety described above can potentially be enforced separately
+in various combinations. This is important, because
 many existing enforcement mechanisms enforce only some of the stack safety
@@
 
-In principle, these properties can exist in any combination, but in practice we see examples
-of certain combinations that form a hierarchy. Many enforcement techniques focus purely on
+ \(\clec\) subsumes \(\clri\), and \(\clei\) subsumes
+\(\clrc\), but integrity and confidentiality are in general orthogonal to each other
+and to \(\wbcf\). Despite this, many enforcement techniques focus purely on
 well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
 address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
-to enforce it completely.
-
-Others combine this protection with some degree of memory protection,
-chiefly focusing on integrity. Skorstengaard et al. \cite{SkorstengaardSTK} characterize stack
-safety as the combination of WBCF and {\it local state encapsulation} (LSE), and prove that their
-capability-based enforcement satisfies both. LSE corresponds to integrity, plus the
-confidentiality of the caller's frame, but not of the unallocated stack or data left
-behind in it by previous callees.
-They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
-together as a safe-by-construction abstract machine. That machine does not allow a callee to
-read or write its caller's frame, but might allow it to read data that is left behind below
-the caller's frame, which is weaker than our full caller confidentiality and rules out
-callee confidentiality entirely. Georges et al. \cite{Georges22:TempsDesCerises} correct
-both limitations.\bcp{Then why are we not comparing to them instead??}
+to enforce it completely. Others combine this protection with some degree of memory protection,
+chiefly focusing on integrity.
 
@@
 
-\paragraph*{Terminology and Notation}
-
 The building blocks of a machine are {\em words} and {\em registers}.
@@
 
+[TODO: this section could use some trimming. I think we have room for this
+discussion, it's just more verbose than everything else and therefore feels weird.]
+
 The full machine supports public stack allocations that can
@@
 
-[TODO: back-propagate abbreviations used here]
-
-Finally, we can define our core properties. They can be found in \cref{tbl:props},
+Finally, we can define our core properties. They can be found in \cref{tab:props},
 arranged to show their commonalities and distinctions. Each definition gives a criterion
@@
 \begin{table*}[h]
-  \begin{align*}
-    \wbcf \triangleq & |\sigma| \uparrow \ret
-    && \textnormal{ where } \ret ~ \mach' \textnormal{ iff }
-    \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}
-    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called} \\
-    %
-    \clri \triangleq & (|\sigma| \uparrow \intProp) ~ (\mach,(V,\sigma))
-    && \textnormal{ where } \intProp ~ \mach' \textnormal{ iff }
-    \mach' \parallel \sealed(V) \cap \Delta(\mach,\mach')
-    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called} \\
-    %
-    \clrc \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\sealed(V)} \nach, &&
-    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
-    & |\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma)) \\
-    & \textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))
-    && \textnormal{ where } (\mach' ~ \confProp ~ \nach') \textnormal{ iff }
-    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach') \\
-    %
-    \clec \triangleq & (|\sigma| \uparrow \cconfProp) ~ (\mach,(V,\sigma))
-    && \textnormal{ where } \cconfProp ~ \mach' \textnormal{ iff }
-    \mach' \parallel \unsealed(V) \cap \Delta(\mach,\mach')
-    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
-    %
-    \clei \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\unsealed(V)} \nach, &&
-    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
-    & |\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma)) \\
-    & \textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))
-    && \textnormal{ where } (\mach' ~ \cintProp ~ \nach') \textnormal{ iff }
-    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach') \\
-  \end{align*}
+  \setlength{\tabcolsep}{1pt}
+  \center
+  \begin{tabular}{l r l l l}
+    \rowcolor{black!20}
+    1
+    & \(\wbcf \triangleq\) & \(|\sigma| \uparrow \ret\)
+    & \(\textnormal{ where } \ret ~ \mach' \textnormal{ iff }\)
+    \(\mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\)
+    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
+    %
+    \rowcolor{black!10}
+    2
+    & \(\clri \triangleq\) & \((|\sigma| \uparrow \intProp) ~ (\mach,(V,\sigma))\)
+    & \(\textnormal{ where } \intProp ~ \mach' \textnormal{ iff }
+    \mach' \parallel \sealed(V) \cap \Delta(\mach,\mach')\)
+    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
+    %
+    \rowcolor{black!20}
+    3
+    & \(\clrc \triangleq\) & \(\forall \nach \textnormal{ s.t. } \mach \approx_{\components} \nach,\)
+    & \(\textnormal{ where } \components = \sealed(V)\)
+    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+    \rowcolor{black!20}
+    3a & & \(|\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma))\) & & \\
+    \rowcolor{black!20}
+    3b & & \(\textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))\)
+    & \(\textnormal{ where } (\mach' ~ \confProp ~ \nach') \textnormal{ iff }
+    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach')\) & \\
+    %
+    \rowcolor{black!10}
+    4
+    & \(\clec \triangleq\) & \((|\sigma| \uparrow \cconfProp) ~ (\mach,(V,\sigma))\)
+    & \(\textnormal{ where } \cconfProp ~ \mach' \textnormal{ iff }
+    \mach' \parallel \Delta(\mach,\mach') - \components\)
+    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+    \rowcolor{black!10}
+    & & & \(\textnormal{ where } \components = \public(V) \cup \object(V)\) & \\
+    %
+    \rowcolor{black!20}
+    5
+    & \(\clei \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\components} \nach,\)
+    & \(\textnormal{ where } \components = \public(V) \cup \object(V)\)
+    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+    \rowcolor{black!20}
+    5a & & \(|\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma))\) & & \\
+    \rowcolor{black!20}
+    5b & & \(\textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))\)
+    & \(\textnormal{ where } (\mach' ~ \cintProp ~ \nach') \textnormal{ iff }
+    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach')\) & \\
+  \end{tabular}
   \caption{Properties}
-  \label{tbl:props}
+  \label{tab:props}
 \end{table*}
-
-[TODO: words about the complementary properties.]
 
@@
 \label{sec:testing}
+
+[TODO: this is still last submission's test. Kept here to get a sense of length.]
 

### 2023-01-23 submission5/core.tex
Fix mathmode errror

@@
 \fi
+%
+\rb{Connect this to the shortcomings of previous work. I'd give at least a quick
+  explanation of what the various properties roughly entail in practice (e.g.,
+  secrecy of private data, etc.). The comment about the single extension of
+  exceptions seems disconnected from the rest.}
 
@@
 focus on the caller over the callee.
+%
+\rb{Also sharpen after extended discussion?}
 
@@
 while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
+%
+\rb{Does this discussion (at least in part) belong in the general background
+  section, some even in key ideas? It still feels a little out of order to me,
+  and occasionally too detailed for the introduction?}
+
+\rb{\textbf{A menagerie of protection mechanisms.}
+%
+We have made explicit the key point that there is \emph{not} a single stack
+abstraction. This paves the way for a \emph{general} idea of stack protection
+that is widely applicable with high-level concepts that remain unchanged (and I
+think it's particularly important to at least introduce these more immutable
+parts early on). The fact that this general ``framework'' requires low-level or
+specific specializations is a predictable consequence of the protean nature of
+the call stack.
+%
+\\
+%
+Let us return to a selection of representative policies in our menagerie and
+consider how they fit into the framework. At the highest level, we can classify
+all those techniques into two large groups, software-based and hardware-based.
+Traditionally, most practical protections are implemented in software, so we
+start by considering this first group.
+%
+\\
+%
+We can make two general observations about these software based techniques.
+Firstly, (as we will see: discussion? technical details?) these techniques are
+\emph{less modular} than their hardware counterparts. That is, the
+countermeasures are tightly coupled with the program code they protect, and
+consequently more effort is required to disentangle the program from the
+protections, and to understand the impact of the latter on the former. This is
+true of any systematic attempt to study the security of stack protection
+policies, including their adaptation to our framework. (Owing to this added
+complexity, software policies will not be the cleanest or most elegant exemplars
+of use of the framework, its salient features and practical utility; this partly
+motivates the focus on hardware policies.)
+%
+\\
+%
+(The above remark holds \emph{at least in a ``single-level model''} that
+considers assembly language/machine code programs. In general, (low-level)
+protections are introduced by a compiler or similar tool. One could consider a
+multi-level extension of the framework pairing models of high-level programs and
+the security transformations performed by a (stack-secure) compiler.)
+%
+\\
+%
+Secondly, the guarantees generally offered by software-based policies are
+\emph{less absolute}, being defined more or less informally and usually
+providing few definite guarantees of absence of \emph{general} classes of
+errors. (The following discussion elaborates on this. It is conceivable that an
+extended and more sophisticated framework could add \emph{probabilistic
+reasoning} capabilities to capture and reason about the effects of less strict
+protection mechanisms. This would be another orthogonal development to what is
+presented here. In this paper, we are interested in assessing the complete
+\emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
+or defects. In passing, this would extend the security properties to e.g. larger
+system verification efforts.)
+%
+\\
+%
+One by one, consider the following specimens of software protection mechanisms
+(quick explanations inlined?):
+%
+\\
+%
+\begin{itemize}
+%
+\item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
+  difficulty of some attacks on the stack, usually (always?) buffer overflows.
+  They can never provide any absolute security guarantees and therefore do not
+  satisfy any of our properties (some of the properties, e.g. confidentiality,
+  are completely out of scope.) Although tedious, we could we could model this
+  mechanism in our framework with the appropriate adaptations --- and
+  unsurprisingly, testing would find property violations, which we already know
+  are there: the results of this effort would be unsurprising and not as useful
+  or instructive as other experiments.
+%
+\item \emph{Bounds checking}. This is fundamentally a high-level technique, even
+  more obviously so than stack canaries (which could still detect violations
+  caused by external code). Bounds checking only applies to code that has been
+  instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
+  impose any constraints on adversarial code. The attacker model is therefore
+  rather weak. (Certain variants, like SoftBound, are good targets for lifting
+  our framework to the source level by e.g. using a C-language semantics and
+  using Tagged C for security modeling; the lifted versions of our properties
+  would likely hold under the expected assumptions. HardBound is probably an
+  interesting transitional case because it uses capability-like hardware
+  primitives for enforcement; it could be close to ongoing and future work to
+  verify stack safety on capability machines, e.g. based on Cerise -- discuss
+  together?)
+%
+\item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
+  ``probabilistic techniques'' in that they ``restrict the flexibility available
+  in creating gadget chains'' (Shanbhogue et al.). This technique is solely
+  concerned with control-flow integrity, and so closely related to our
+  well-bracketed control flow. Nevertheless, and again, it is only an
+  approximate technique, and our testing framework would be able to find the
+  expected attacks against our security property.
+%
+\item \emph{Split stacks}. The security claims of split stacks are stronger than
+  those of most software countermeasures, i.e., they are meant to guarantee
+  their titular code-pointer integrity. They explicitly offer no protections on
+  stack data (i.e., integrity and confidentiality), and so restrict themselves
+  to well-bracketed control flow. (This is probably closer to our framework than
+  most other software techniques, and it seems like it would be easier to model
+  as well. However, this is again work based on high-level languages and would
+  benefit from a higher level of abstraction.)
+%
+\item \ldots
+%
+\end{itemize}
+%
+Opposed to the above group of software techniques we have hardware-based
+protection policies (more in background/technical preliminaries?). Although
+programs must effectively leverage the protection mechanisms offered by the
+hardware, the distinction between those protections and the program logic itself
+is much cleaner: it is fairly simple to separate the use of the security
+features in the hardware (and their effects in program code) from regular
+program code. They also tend to aspire to offer complete protection against
+entire classes of attacks. Both aspects make this family of techniques a better
+fit for illustrating and showcasing our techniques.
+%
+\\
+%
+Why favor micro-policies? Their design is more flexible and arguably more
+general than capability machines (of course, they pay a performance price for
+this). This generality and ease of programming make them good candidates for an
+exploration the design space of stack protection policies and proof of concept
+implementations. On the practical side, there is a number of existing policies
+available to examine (not just lazy policies!) --- and we find how they satisfy
+our properties, how they do not, and errors in some of them, and see how to
+correct them. Lazy enforcement is particularly novel, but not the only
+possibility. This additional breadth is beneficial for the proof of concept
+stage.
+%
+\\
+%
+Regarding capability machines, we know that most designs based on capability
+machines do not satisfy our properties (or do so incurring significant costs),
+and that fixes are not trivial and involve new types of capabilities (in a
+sense, are they any more ``realistic?'' Only the latest Cerise paper would seem
+to satisfy our properties, is directly influenced by previous versions of this
+work, and is a reasonable target for ongoing/further study -- but this is not
+easy, and even the author(s) estimate it would be a significant amount of work,
+especially full proofs (references, communications, etc.).
+%
+\\
+%
+\ldots
+%
+}
 
@@
     5
-    & \(\clei \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\components} \nach,\)
+    & \(\clei \triangleq\) & \(\forall \nach \textnormal{ s.t. } \mach \approx_{\components} \nach,\)
     & \(\textnormal{ where } \components = \public(V) \cup \object(V)\)

### 2023-01-24 submission5/core.tex
Relocate intro text to separate file
Temporary move, should go away after the rewrite attempt.

--- 2023-01-23 submission5/core.tex
@@
 
-The call stack is a perennial target for low-level attacks, leading to
-dire consequences from leakage and corruption of private stack data
-to control-flow hijacking. To foil such attacks, a menagerie of
-software and hardware protections have been proposed,
-%
-including stack canaries~\cite{Cowan+98},
-bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
-split stacks~\cite{Kuznetsov+14},
-shadow stacks~\cite{Dang+15,Shanbhogue+19},
-capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
-  \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\bcp{Yes, going back to MIT days---we should
-  include several more of these, if only to give readers the impression that
-this is a well-studied mechanism (so formalizing its protections is
-useful).}
-\fi
-
-The protections offered by such mechanisms are commonly described by giving
-concrete examples of attacks that they can prevent---corruption of return
-addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more formal characterization to the reader's intuition. Previous attempts
-to produce a formal characterization do so using a {\it fully abstract overlay semantics},
-a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
-But this approach is not suitable for the {\em lazy} enforcement mechanism
-of Roessler and DeHon, in which stores are not checked, and illegal stores are
-only caught when their location is later read. We aim to provide a stack-safety
-property suitable for testing lazy enforcement.
-
-We propose a novel characterization of stack safety using the formal tools of language-based
-security~\cite{sabelfeld2003language}, decomposing stack safety into
-the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the control flow protection
-of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
-\ifexceptions
-We also extend WBCF to a setting with exceptions,
-in which there may be multiple points at which a callee can ``return.''
-\fi
-
-This gives us a total of five core properties: \(\wbcf\)
-caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
-
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly detect the ones that
-attempt to violate one of our properties. Furthermore, we
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms, including ones that we accidentally created
-while implementating the micro-policy and ones that we
-intentionally crafted to increase our confidence in the effectiveness
-of our testing.
-
-We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-
-\subsection{Motivation}
-
-Security mechanisms designed to prevent a certain attack can be brittle---successfully
-eliminating the attack as it exists at the time of publication, but leaving room for attackers
-to find new, similar attacks. To mitigate this risk, mechanisms should aim to enforce formal
-properties of correct behavior that can be proven or tested. Such properties become the
-specification against which enforcement can be validated; even enforcement mechanisms that do
-not fulfill the property benefit from the ability to articulate {\it why} they fail.
-
-Stack safety is challenging to specify in full generality, because the stack is fundamentally
-an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
-which is broadly consistent across architectures and language semantics, and therefore amenable
-to a very general approach.)
-The desired behavior of the stack varies based on ISA, calling convention,
-and the features of whichever programming language originated the code.
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
-must be accompanied by tweaks to the model.
-
-We are not attempting to create a universal definition of stack safety for all systems.
-Rather, we define the stack safety of a single, fairly realistic system, decomposed into
-(1) a lightweight model of the system and its security-relevant features, from which we
-derive the integrity, confidentiality, and control-flow requirements of each call and
-return, (2) the application of that model to a particular target system, and (3)
-the formal criteria for integrity, confidentiality, and control flow
-to obey the requirements. We consider the criteria in (3) to represent universal definitions
-that are parameterized by (1).
-
-We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
-exclusively on control flow, some enforce integrity but not confidentiality, and many
-focus on the caller over the callee.
-
-\paragraph*{Permissiveness, Laziness, and Tagged Architectures}
-
-Security properties tend to be conservative, identifying violations that would be
-harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
-and of compatibility with existing codebases. In a crucial motivating example for this work,
-Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
-using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
-STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
-A set of software-defined rules, termed a micro-policy, determines the tags for operation
-results based on the tags of their inputs---or, if the input tags are in a dangerous
-combination, it may halt the machine (a ``failstop.'')
-
-Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
-is inefficient when allocating and clearing stack frames, because each word must be initialized
-with a tag corresponding to the (depth of the) current function activation. They propose
-{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
-when a function writes to memory. This means that writes cannot be prevented, necessarily
-violating existing models of stack safety. Instead the micro-policy enforces that memory be
-read by the same fuction that wrote it. The owner of the data will never see the changes,
-so it is secure in the sense that the attacker is unsuccessful at changing program behavior
-(except, perhaps, by causing a fault.) We elaborate that sense, and test whether the
-micro-policy fulfills it.
-
-This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
-machine states that make claims about the future observable behavior of execution from those states.
-That is not to say that our properties are specialized for lazy micro-policies. It also supports
-normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
-and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
-while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
-
-\subsection{Contributions}
-
-In sum, we offer the following contributions:
-
-\begin{itemize}
-\item We formalize permissive stack confidentiality and integrity properties,
-  which are parameterized over a notion of external
-  observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in system behavior.
-\item We instantiate these definitions with a realistic
-  system with features such as argument passing
-  on the stack, callee-saves registers, and tail-call elimination.
-  Our model is modular enough that adding these features is straightforward.
-\item Our formalization of well-bracketed control flow in the presence of tail-call
-  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
-\item We use property-based random testing to validate the relationship between
-  our properties and micro-policies.
-\end{itemize}
-
-Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as their testing implementation. There are no proofs---we use primarily Coq for the
-QuickChick library and to ensure that the formalizations are unambiguous.
-
-\paragraph*{Limitations}
-
-%Our concurrency model is fairly
-%simplistic, assuming a fixed number of threads each with its own dedicated processor.
-We model memory-safe stack objects, but not a heap. Regions outside of
-stacks can be used however the compiler likes, including as a heap, but no protection is
-built in and our properties assume that if a pointer to a stack object is stored there,
-it is permanently compromised.
-
-Our properties are termination insensitive: in order to support enforcement mechanisms
-that failstop rather than execute dangerous code, the properties must treat failstops
-as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
-This is consistent with our treatment of properties that specify the state after a return,
-in that a callee that diverges for any reason will never reach the return, and will fulfill
-the property vacuously.
+\input{intro-v1}
 

### 2023-01-24 submission5/intro-v2.tex
Start revised intro from original text

\rb{\textbf{The call stack and its security.} Maybe a few opening words to set
  the stage and bring two things to the front: (1) functions and subroutines as
  fundamental building blocks of any kind of structured programming; and (2) the
  stack as the natural abstraction to manage these in a computer. And also maybe
  a little bit more about the width, the gravity and the \emph{foundational
  nature} of attacks on the stack, which the sentence below touches on
  summarily.}
%
The call stack is a perennial target for low-level attacks, leading to
dire consequences from leakage and corruption of private stack data
to control-flow hijacking.
%
\rb{We can mention things like e.g. the most recent CWE Top 25 Most Dangerous
    Software Weaknesses from MITRE, cf.
    \url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}, where
    e.g. \#1 (Out-of-bounds Write), \#5 (Out-of-bounds Read) and \#7 (Use After
    Free), that is a good number of the most important and prevalent
    vulnerability types are closely related to the stack. These will show up in
    other popular rankings as well.}
%
To foil such attacks, a menagerie of
software and hardware protections have been proposed,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
  \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
\rb{Did we want to add even more techniques? Is this selection broad enough?
  Anyway reference that we'll be looking at these more closely later.}

The protections offered by such mechanisms are commonly described by giving
concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more formal characterization to the reader's intuition.
%
\rb{Begins to anticipate subsequent discussion. This is even something of an
  understatement: some papers (esp. newer ones?), make concrete claims, though
  they remain informal. Arguably, none have a truly formal characterization in
  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
  of the novelties of this work!}
%
Previous attempts
to produce a formal characterization do so using a {\it fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
But this approach is not suitable for the {\em lazy} enforcement mechanism
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read. We aim to provide a stack-safety
property suitable for testing lazy enforcement.
%
\rb{Need to discuss overlay semantics, but also note that it is very intensional
  and very specific, as the authors acknowledge. I'd introduce this a bit more
  gradually. Also, if phrased this way, it sounds like its main drawback is a
  lack of generality, to which we should contrast a more general approach, not
  just one that is specialized to a specific policy, i.e. lazy micro-policies
  (I'd try to present these more gradually as well).}

We propose a novel characterization of stack safety using the formal tools of language-based
security~\cite{sabelfeld2003language}, decomposing stack safety into
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the control flow protection
of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
\ifexceptions
We also extend WBCF to a setting with exceptions,
in which there may be multiple points at which a callee can ``return.''
\fi
%
\rb{Connect this to the shortcomings of previous work. I'd give at least a quick
  explanation of what the various properties roughly entail in practice (e.g.,
  secrecy of private data, etc.). The comment about the single extension of
  exceptions seems disconnected from the rest.}

This gives us a total of five core properties: \(\wbcf\)
caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and DeHon's micro-policies correctly detect the ones that
attempt to violate one of our properties. Furthermore, we
check that the testing framework is able to generate counterexamples
that violate our properties but are \emph{not} halted by incorrect
variants of the enforcement mechanisms, including ones that we accidentally created
while implementating the micro-policy and ones that we
intentionally crafted to increase our confidence in the effectiveness
of our testing.
%
\rb{This and the following paragraph could start more generally (we need to
  answer the question of why these experiments, what is their interest), and
  probably motivate the use of PBT (experiment results below show there's
  benefit in them!) and, somewhere at least, talk about other points in the
  spectrum.}

We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak uses the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.
%
\rb{Anything to add from personal communications? More details later? Shows
  these things are very hard to get right -- like the latest work on Cerise,
  also motivated by this very work.}

\subsection{Motivation}

Security mechanisms designed to prevent a certain attack can be brittle---successfully
eliminating the attack as it exists at the time of publication, but leaving room for attackers
to find new, similar attacks. To mitigate this risk, mechanisms should aim to enforce formal
properties of correct behavior that can be proven or tested. Such properties become the
specification against which enforcement can be validated; even enforcement mechanisms that do
not fulfill the property benefit from the ability to articulate {\it why} they fail.
%
\rb{Good stuff! There is some overlap between this discussion and the opening
  motivation, maybe we could restructure things a bit so that those threads are
  unified?}

Stack safety is challenging to specify in full generality, because the stack is fundamentally
an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
which is broadly consistent across architectures and language semantics, and therefore amenable
to a very general approach.)
The desired behavior of the stack varies based on ISA, calling convention,
and the features of whichever programming language originated the code.
For instance, the current state-of-the-art stack-safety specification from Georges et al.
\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
a callee should always return to its caller. But, in the presence of tail-call elimination, a
return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
must be accompanied by tweaks to the model.
%
\rb{This point is extremely important!}

We are not attempting to create a universal definition of stack safety for all systems.
Rather, we define the stack safety of a single, fairly realistic system, decomposed into
(1) a lightweight model of the system and its security-relevant features, from which we
derive the integrity, confidentiality, and control-flow requirements of each call and
return, (2) the application of that model to a particular target system, and (3)
the formal criteria for integrity, confidentiality, and control flow
to obey the requirements. We consider the criteria in (3) to represent universal definitions
that are parameterized by (1).
%
\rb{Sharpen, contrast the universal definitions (or a framework of parametric
  definitions, say) with the concrete instance we study, again with its interest
  properly motivated.}

We treat stack safety as multiple properties, rather than one multi-faceted property, because
not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
exclusively on control flow, some enforce integrity but not confidentiality, and many
focus on the caller over the callee.
%
\rb{Also sharpen after extended discussion?}

\paragraph*{Permissiveness, Laziness, and Tagged Architectures}

Security properties tend to be conservative, identifying violations that would be
harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
and of compatibility with existing codebases. In a crucial motivating example for this work,
Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
A set of software-defined rules, termed a micro-policy, determines the tags for operation
results based on the tags of their inputs---or, if the input tags are in a dangerous
combination, it may halt the machine (a ``failstop.'')

Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
is inefficient when allocating and clearing stack frames, because each word must be initialized
with a tag corresponding to the (depth of the) current function activation. They propose
{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
when a function writes to memory. This means that writes cannot be prevented, necessarily
violating existing models of stack safety. Instead the micro-policy enforces that memory be
read by the same fuction that wrote it. The owner of the data will never see the changes,
so it is secure in the sense that the attacker is unsuccessful at changing program behavior
(except, perhaps, by causing a fault.) We elaborate that sense, and test whether the
micro-policy fulfills it.

This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
machine states that make claims about the future observable behavior of execution from those states.
That is not to say that our properties are specialized for lazy micro-policies. It also supports
normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
%
\rb{Does this discussion (at least in part) belong in the general background
  section, some even in key ideas? It still feels a little out of order to me,
  and occasionally too detailed for the introduction?}

\rb{\textbf{A menagerie of protection mechanisms.}
%
We have made explicit the key point that there is \emph{not} a single stack
abstraction. This paves the way for a \emph{general} idea of stack protection
that is widely applicable with high-level concepts that remain unchanged (and I
think it's particularly important to at least introduce these more immutable
parts early on). The fact that this general ``framework'' requires low-level or
specific specializations is a predictable consequence of the protean nature of
the call stack.
%
\\
%
Let us return to a selection of representative policies in our menagerie and
consider how they fit into the framework. At the highest level, we can classify
all those techniques into two large groups, software-based and hardware-based.
Traditionally, most practical protections are implemented in software, so we
start by considering this first group.
%
\\
%
We can make two general observations about these software based techniques.
Firstly, (as we will see: discussion? technical details?) these techniques are
\emph{less modular} than their hardware counterparts. That is, the
countermeasures are tightly coupled with the program code they protect, and
consequently more effort is required to disentangle the program from the
protections, and to understand the impact of the latter on the former. This is
true of any systematic attempt to study the security of stack protection
policies, including their adaptation to our framework. (Owing to this added
complexity, software policies will not be the cleanest or most elegant exemplars
of use of the framework, its salient features and practical utility; this partly
motivates the focus on hardware policies.)
%
\\
%
(The above remark holds \emph{at least in a ``single-level model''} that
considers assembly language/machine code programs. In general, (low-level)
protections are introduced by a compiler or similar tool. One could consider a
multi-level extension of the framework pairing models of high-level programs and
the security transformations performed by a (stack-secure) compiler.)
%
\\
%
Secondly, the guarantees generally offered by software-based policies are
\emph{less absolute}, being defined more or less informally and usually
providing few definite guarantees of absence of \emph{general} classes of
errors. (The following discussion elaborates on this. It is conceivable that an
extended and more sophisticated framework could add \emph{probabilistic
reasoning} capabilities to capture and reason about the effects of less strict
protection mechanisms. This would be another orthogonal development to what is
presented here. In this paper, we are interested in assessing the complete
\emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
or defects. In passing, this would extend the security properties to e.g. larger
system verification efforts.)
%
\\
%
One by one, consider the following specimens of software protection mechanisms
(quick explanations inlined?):
%
\\
%
\begin{itemize}
%
\item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
  difficulty of some attacks on the stack, usually (always?) buffer overflows.
  They can never provide any absolute security guarantees and therefore do not
  satisfy any of our properties (some of the properties, e.g. confidentiality,
  are completely out of scope.) Although tedious, we could we could model this
  mechanism in our framework with the appropriate adaptations --- and
  unsurprisingly, testing would find property violations, which we already know
  are there: the results of this effort would be unsurprising and not as useful
  or instructive as other experiments.
%
\item \emph{Bounds checking}. This is fundamentally a high-level technique, even
  more obviously so than stack canaries (which could still detect violations
  caused by external code). Bounds checking only applies to code that has been
  instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
  impose any constraints on adversarial code. The attacker model is therefore
  rather weak. (Certain variants, like SoftBound, are good targets for lifting
  our framework to the source level by e.g. using a C-language semantics and
  using Tagged C for security modeling; the lifted versions of our properties
  would likely hold under the expected assumptions. HardBound is probably an
  interesting transitional case because it uses capability-like hardware
  primitives for enforcement; it could be close to ongoing and future work to
  verify stack safety on capability machines, e.g. based on Cerise -- discuss
  together?)
%
\item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
  ``probabilistic techniques'' in that they ``restrict the flexibility available
  in creating gadget chains'' (Shanbhogue et al.). This technique is solely
  concerned with control-flow integrity, and so closely related to our
  well-bracketed control flow. Nevertheless, and again, it is only an
  approximate technique, and our testing framework would be able to find the
  expected attacks against our security property.
%
\item \emph{Split stacks}. The security claims of split stacks are stronger than
  those of most software countermeasures, i.e., they are meant to guarantee
  their titular code-pointer integrity. They explicitly offer no protections on
  stack data (i.e., integrity and confidentiality), and so restrict themselves
  to well-bracketed control flow. (This is probably closer to our framework than
  most other software techniques, and it seems like it would be easier to model
  as well. However, this is again work based on high-level languages and would
  benefit from a higher level of abstraction.)
%
\item \ldots
%
\end{itemize}
%
Opposed to the above group of software techniques we have hardware-based
protection policies (more in background/technical preliminaries?). Although
programs must effectively leverage the protection mechanisms offered by the
hardware, the distinction between those protections and the program logic itself
is much cleaner: it is fairly simple to separate the use of the security
features in the hardware (and their effects in program code) from regular
program code. They also tend to aspire to offer complete protection against
entire classes of attacks. Both aspects make this family of techniques a better
fit for illustrating and showcasing our techniques.
%
\\
%
Why favor micro-policies? Their design is more flexible and arguably more
general than capability machines (of course, they pay a performance price for
this). This generality and ease of programming make them good candidates for an
exploration the design space of stack protection policies and proof of concept
implementations. On the practical side, there is a number of existing policies
available to examine (not just lazy policies!) --- and we find how they satisfy
our properties, how they do not, and errors in some of them, and see how to
correct them. Lazy enforcement is particularly novel, but not the only
possibility. This additional breadth is beneficial for the proof of concept
stage.
%
\\
%
Regarding capability machines, we know that most designs based on capability
machines do not satisfy our properties (or do so incurring significant costs),
and that fixes are not trivial and involve new types of capabilities (in a
sense, are they any more ``realistic?'' Only the latest Cerise paper would seem
to satisfy our properties, is directly influenced by previous versions of this
work, and is a reasonable target for ongoing/further study -- but this is not
easy, and even the author(s) estimate it would be a significant amount of work,
especially full proofs (references, communications, etc.).
%
\\
%
\ldots
%
}

\subsection{Contributions}

In sum, we offer the following contributions:

\begin{itemize}
\item We formalize permissive stack confidentiality and integrity properties,
  which are parameterized over a notion of external
  observation, and are violated only if accessing secrets or overwriting
  data causes a visible change in system behavior.
\item We instantiate these definitions with a realistic
  system with features such as argument passing
  on the stack, callee-saves registers, and tail-call elimination.
  Our model is modular enough that adding these features is straightforward.
\item Our formalization of well-bracketed control flow in the presence of tail-call
  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
\item We use property-based random testing to validate the relationship between
  our properties and micro-policies.
\end{itemize}

Our artifact contains Coq formalizations of our properties for illustrative purposes
as well as their testing implementation. There are no proofs---we use primarily Coq for the
QuickChick library and to ensure that the formalizations are unambiguous.
%
\rb{Will have another look after the full pass.}

\paragraph*{Limitations}

%Our concurrency model is fairly
%simplistic, assuming a fixed number of threads each with its own dedicated processor.
We model memory-safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

Our properties are termination insensitive: in order to support enforcement mechanisms
that failstop rather than execute dangerous code, the properties must treat failstops
as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
This is consistent with our treatment of properties that specify the state after a return,
in that a callee that diverges for any reason will never reach the return, and will fulfill
the property vacuously.
%
\rb{Limitations in the introduction or later? Orthogonality of the heap, morally
  similar interesting setups on the stack? Termination sensitivity and
  capability machine semantics?}

### 2023-01-24 submission5/intro-v2.tex

@@
   nature} of attacks on the stack, which the sentence below touches on
-  summarily.}
+  summarily.
+\rb{(+SNA) Connection to functions (and adversarial code?). Attacks on the stack
+  are also attacks on this fundamental abstraction (functions, subroutines,
+  procedures, methods, however you call them)}
+}
 %
@@
     vulnerability types are closely related to the stack. These will show up in
-    other popular rankings as well.}
+    other popular rankings as well.
+\rb{(+SNA) List major CWEs? No matter the class of weakness considered, many
+  involve a \emph{flow} going through the stack (cf. Eternal war on memory).}
+}
 %
@@
   mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
-  of the novelties of this work!}
+  of the novelties of this work!
+\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
+  than previous work, without showing a fully general framework (which would be
+  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
+  start because of their flexibility, elegance, enabling a wide of interesting
+  experiments, etc.}
+}
 %
@@
   these things are very hard to get right -- like the latest work on Cerise,
-  also motivated by this very work.}
+  also motivated by this very work.
+\rb{(+SNA) At the very least, these models are subtle and getting them right,
+  and describing them unambiguously, is hard without a framework like this.}
+}
 
@@
   or instructive as other experiments.
+  \rb{(+SNA) They are kind of lazy, aren't they?}
 %
@@
 especially full proofs (references, communications, etc.).
+\rb{(+SNA) Discuss the degree in which the various papers satisfy the
+  properties\ldots and the degrees to which they \emph{aim to}. E.g. originally
+  they don't talk about confidentiality. Do they have callee integrity? Maybe
+  they aren't thinking about it? (This kind of discussion is a good motivation
+  for the formal properties. Also related to the hierarchy. And: how to convey
+  that the latest paper draws inspiration directly from this work?}
 %
@@
 }
+\rb{(+SNA) Any interesting tables comparing specimens in the menagerie?}
+\rb{(+SNA) A general multi-level extension of the framework would need to model
+  multiple source languages. Among other things, the stack is the meeting place
+  of all sorts of different language abstractions! One of the advantages of the
+  single-level model is that it lets us be agnostic about the language. (On the
+  other hand, it is likely that a single source and target mediated by the
+  compiler suffice to model many existing techniques, coming e.g. from systems
+  programming and C-like languages.}
 

### 2023-01-25 submission5/intro-v2.tex
Checkpoint on intro

@@
-\rb{\textbf{The call stack and its security.} Maybe a few opening words to set
-  the stage and bring two things to the front: (1) functions and subroutines as
-  fundamental building blocks of any kind of structured programming; and (2) the
-  stack as the natural abstraction to manage these in a computer. And also maybe
-  a little bit more about the width, the gravity and the \emph{foundational
-  nature} of attacks on the stack, which the sentence below touches on
-  summarily.
-\rb{(+SNA) Connection to functions (and adversarial code?). Attacks on the stack
-  are also attacks on this fundamental abstraction (functions, subroutines,
-  procedures, methods, however you call them)}
-}
+\subsection{The Call Stack and Its Security}
+
+Computer programming at any scale relies on \emph{functions} (variously known as
+subroutines, procedures, methods, etc.) as a \emph{foundational abstraction} to
+structure programs into building blocks representing units of computation that
+can \emph{call} one another to create larger computations.
 %
-The call stack is a perennial target for low-level attacks, leading to
-dire consequences from leakage and corruption of private stack data
+At a low level, a function call comprises its own data (private or shared), as
+well as information about its \emph{caller} and \emph{callee} functions. The
+runtime must keep track of the local contexts of all active function activations
+at any given time; the \emph{fundamental data structure} charged with realizing
+this abstraction is the \emph{(call) stack}.
+%
+\rb{This should be well known, so no need for illustration?}
+
+The correct execution on a program---and its security---is contingent on all of
+its functions properly implementing this abstraction by means of a careful use
+of the call stack. Be it accidental or intentional, incorrect management of the
+stack is at the root of a myriad of faults whose effects range from (rarely)
+innocuous to (usually) catastrophic.
+%
+From a security perspective, attacks on the stack (say, from a malicious
+function that chooses to abuse its access to the contents of this data
+structure) are as fundamental, varied and pervasive as the functional
+abstraction itself.
+%
+\rb{Begins to prefigure the idea of multiple stack implementations.}
+
+%% \rb{\textbf{The call stack and its security.} Maybe a few opening words to set
+%%   the stage and bring two things to the front: (1) functions and subroutines as
+%%   fundamental building blocks of any kind of structured programming; and (2) the
+%%   stack as the natural abstraction to manage these in a computer. And also maybe
+%%   a little bit more about the width, the gravity and the \emph{foundational
+%%   nature} of attacks on the stack, which the sentence below touches on
+%%   summarily.
+%% \rb{(+SNA) Connection to functions (and adversarial code?). Attacks on the stack
+%%   are also attacks on this fundamental abstraction (functions, subroutines,
+%%   procedures, methods, however you call them)}
+%% }
+
+Indeed, the call stack is a perennial target for low-level attacks, leading to
+dire consequences, from leakage and corruption of private stack data
 to control-flow hijacking.
 %
-\rb{We can mention things like e.g. the most recent CWE Top 25 Most Dangerous
-    Software Weaknesses from MITRE, cf.
-    \url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}, where
-    e.g. \#1 (Out-of-bounds Write), \#5 (Out-of-bounds Read) and \#7 (Use After
-    Free), that is a good number of the most important and prevalent
-    vulnerability types are closely related to the stack. These will show up in
-    other popular rankings as well.
-\rb{(+SNA) List major CWEs? No matter the class of weakness considered, many
-  involve a \emph{flow} going through the stack (cf. Eternal war on memory).}
-}
+Even a cursory glance at the most recent release of the authoritative CWE Top 25
+Most Dangerous Software Weaknesses
+list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
+shows general classes of vulnerabilities closely related to the stack \rb{(though not exclusively)}
+consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read), \#7 (CWE-416, Use
+After Free) or \#11 (CWE-476, NULL Pointer Dereference).
 %
+Moreover, many other types of faults involve some kind of control or data flow
+using data stored in the stack \rb{cf. Eternal war in memory, Google/Microsoft
+memory safety statistics, etc.}.
+
 To foil such attacks, a menagerie of
@@
 capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}---we discuss these in more
+detail in \cref{XXX}.
   \ifaftersubmission\apt{Mostly from
@@
 %
-\rb{Did we want to add even more techniques? Is this selection broad enough?
-  Anyway reference that we'll be looking at these more closely later.}
+\rb{Broad enough? Do we want to add more techniques?}
+%
+Yet despite this wealth of potential solutions, the stack remains a central
+security concern---why?
+%
+Security mechanisms designed to prevent a certain attack can be brittle---successfully
+eliminating the attack as it exists at the time of publication, but leaving room for attackers
+to find new, similar attacks.
+%
+%% \rb{And so the cat-and-mouse game continues.}
 
-The protections offered by such mechanisms are commonly described by giving
+To mitigate this risk, \rb{we claim that} mechanisms should aim to enforce \emph{formal
+properties} of correct \rb{safe?} behavior that can be proven or tested. Such properties become the
+\emph{specification} against which enforcement can be validated; even enforcement mechanisms that do
+not fulfill a property benefit from the ability to articulate \emph{why} they fail.
+%
+%% \rb{Good stuff! There is some overlap between this discussion and the opening
+%%   motivation, maybe we could restructure things a bit so that those threads are
+%%   unified?}
+%
+This is the problem we begin to address in this paper \rb{or some such
+  connecting words}.
+
+\subsection{Security Properties for Stack Safety}
+
+The protections offered by stack protection mechanisms are commonly described by giving
 concrete examples of attacks that they can prevent---corruption of return
@@
 
-Security mechanisms designed to prevent a certain attack can be brittle---successfully
-eliminating the attack as it exists at the time of publication, but leaving room for attackers
-to find new, similar attacks. To mitigate this risk, mechanisms should aim to enforce formal
-properties of correct behavior that can be proven or tested. Such properties become the
-specification against which enforcement can be validated; even enforcement mechanisms that do
-not fulfill the property benefit from the ability to articulate {\it why} they fail.
-%
-\rb{Good stuff! There is some overlap between this discussion and the opening
-  motivation, maybe we could restructure things a bit so that those threads are
-  unified?}
-
 Stack safety is challenging to specify in full generality, because the stack is fundamentally
-an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
+an architecture-specific construct. (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
 which is broadly consistent across architectures and language semantics, and therefore amenable

### 2023-01-26 submission5/intro-v2.tex
Some more work on intro

@@
 \subsection{The Call Stack and Its Security}
+
+\rb{A.K.A. motivation}
 
@@
 memory safety statistics, etc.}.
-
+%
 To foil such attacks, a menagerie of
@@
 security concern---why?
-%
+
 Security mechanisms designed to prevent a certain attack can be brittle---successfully
@@
 %% \rb{And so the cat-and-mouse game continues.}
-
+%
 To mitigate this risk, \rb{we claim that} mechanisms should aim to enforce \emph{formal
@@
 %
-This is the problem we begin to address in this paper \rb{or some such
+This is the problem we begin to address in this paper \rb{or a few more such
   connecting words}.
 
-\subsection{Security Properties for Stack Safety}
+\subsection{A Menagerie of Stack Protection Mechanisms}
+
+\rb{A.K.A. related work (part 1?). I wonder if this goes before or after a first
+  look at the idea of the properties.}
 
@@
 %
-\rb{Begins to anticipate subsequent discussion. This is even something of an
-  understatement: some papers (esp. newer ones?), make concrete claims, though
-  they remain informal. Arguably, none have a truly formal characterization in
-  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
-  of the novelties of this work!
-\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
-  than previous work, without showing a fully general framework (which would be
-  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
-  start because of their flexibility, elegance, enabling a wide of interesting
-  experiments, etc.}
-}
+To make these notions more precise, we review a selection of specimens from the
+menagerie and begin to consider the difficulties involved in specifying and
+verifying the protection they offer.
+
+\subsubsection{Software Techniques}
+
+At the highest level, we can classify techniques into two large groups depending
+on whether their protection is furnished at the software or the hardware level.
+Up to date, most proposals, including all commonly deployed practical
+protections, are implemented in software, so we start by considering exemplars
+from this first group.
+
+\paragraph{Stack Canaries}
 %
-Previous attempts
-to produce a formal characterization do so using a {\it fully abstract overlay semantics},
-a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
-But this approach is not suitable for the {\em lazy} enforcement mechanism
-of Roessler and DeHon, in which stores are not checked, and illegal stores are
-only caught when their location is later read. We aim to provide a stack-safety
-property suitable for testing lazy enforcement.
+Canaries are special values inserted by the caller at certain points of the
+stack and checked for corruption on return, primarily for protection against
+some classes of control-flow hijacking (i.e., seeking to overwrite the caller's
+return address through a buffer overflow that would also overwrite the canary
+detectably).
 %
-\rb{Need to discuss overlay semantics, but also note that it is very intensional
-  and very specific, as the authors acknowledge. I'd introduce this a bit more
-  gradually. Also, if phrased this way, it sounds like its main drawback is a
-  lack of generality, to which we should contrast a more general approach, not
-  just one that is specialized to a specific policy, i.e. lazy micro-policies
-  (I'd try to present these more gradually as well).}
+Observe that this is a \emph{hardening} technique that increases the difficulty
+of some control-flow attacks on the stack, but can never provide any absolute
+guarantees on control flow integrity; neither does it do anything to protect
+function data stored in the stack.
+%
+Also interestingly, a successful attack is not detected the moment it takes
+place, instead action is \emph{deferred} until the attacker returns control to
+the caller.
+
+\paragraph{Shadow Stacks}
+%
+A shadow stack is a mirror image of the control-relevant parts of the stack,
+maintained by the program to attempt to detect attempts to hijack its control
+flow. They are similar in purpose and can be seen as related to stack canaries.
+Like canaries, they offer hardened but not guaranteed, deferred protection
+against control-flow attacks, ``restrict the flexibility available in creating
+gadget chains'' \cite{Shanbhogue+19}, and do not cover the data-relevant regions
+of the stack.
+%
+\rb{Could probably find other variants (e.g. SmashGuard, on hardware?), but this
+  seems to be the basic idea. Some of those require OS support, etc.}
+
+\paragraph{Bounds Checking}
+%
+Under a bounds checking discipline, all the pointers in a program are extended
+with some disjoint metadata, and these are combined to gate memory accesses,
+usually with the objective of enforcing some form of \emph{memory
+safety} \rb{(relate this to the basic high-level concepts and types of attacks
+we want to consider: control flow, some integrity and confidentiality?)}.
+%
+Compared to stack canaries and shadow stacks, bounds checking is eminently
+a \emph{high-level} technique, designed with a particular programming language
+in mind and implemented through a \emph{secure compiler} \cite{} that performs
+the necessary code transformations. Also unlike those techniques (which make few
+assumptions on callee code), it offers little to no protection
+against \emph{adversarial low-level} code that is not constrained to obey the
+bounds checking discipline.
+%
+\rb{Some related work, e.g. HardBound, is inspired by this but more closely
+  related to hardware capabilities.}
+
+\paragraph{Split Stacks}
+%
+Code-pointer integrity seeks to prevent control-flow hijacking attacks by
+splitting the program memory, including the stack, into a regular region and a
+safe region; those objects that require protection are located in the latter and
+all accesses protected by static or dynamic checks.
+%
+Like shadow stacks, bounds checking is a high-level technique designed to
+protect code written in a high-level language, but, unlike them, it is meant to
+offer protection under a stronger threat model that also considers the presence
+of untrusted code. Similar to most software-based methods, it purports to
+protect exclusively the control-relevant parts of the stack, making no claims
+about its data-relevant parts.
+
+\subsubsection{Hardware Techniques}
+
+In addition to many the software-based techniques developed to provide greater
+degrees of protection to the stack, often at a significant cost in performance,
+some proposals have turned to specialized hardware features to offset those
+penalties and attempt to offer greater levels of protection.
+%
+\rb{Something about how realistic/practical these alternatives are? Are we
+  selecting partly based on prototypes, etc.?}
+
+\paragraph{Capabilities}
+%
+\rb{(There is plenty of variety here, should we focus on the Cerise-and-friends
+  (CHERI-adjacent) work exclusively?)}
+%
+Capability machines are able to represent and manipulate unforgeable tokens of
+authority over specific memory regions using a combination of dedicated
+out-of-band memory and a collection of specialized ISA instructions.
+%
+These hardware capabilities can be used to implement \emph{secure calling
+conventions} \ldots
+
+\paragraph{Micro-Policies}
+%
+\rb{(This partly overlaps with the technical preliminaries)}
+%
+\ldots
+
+\subsection{A Language-Based View of Stack Safety}
+
+Stack safety is challenging to specify in full generality, because the stack is fundamentally
+an architecture-specific construct. (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
+which is broadly consistent across architectures and language semantics, and therefore amenable
+to a very general approach.)
+The desired behavior of the stack varies based on ISA, calling convention,
+and the features of whichever programming language originated the code.
+For instance, the current state-of-the-art stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
+must be accompanied by tweaks to the model.
+%
+\rb{This point is extremely important!}
 
@@
 callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
-
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly detect the ones that
-attempt to violate one of our properties. Furthermore, we
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms, including ones that we accidentally created
-while implementating the micro-policy and ones that we
-intentionally crafted to increase our confidence in the effectiveness
-of our testing.
-%
-\rb{This and the following paragraph could start more generally (we need to
-  answer the question of why these experiments, what is their interest), and
-  probably motivate the use of PBT (experiment results below show there's
-  benefit in them!) and, somewhere at least, talk about other points in the
-  spectrum.}
-
-We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-%
-\rb{Anything to add from personal communications? More details later? Shows
-  these things are very hard to get right -- like the latest work on Cerise,
-  also motivated by this very work.
-\rb{(+SNA) At the very least, these models are subtle and getting them right,
-  and describing them unambiguously, is hard without a framework like this.}
-}
-
-\subsection{Motivation}
-
-Stack safety is challenging to specify in full generality, because the stack is fundamentally
-an architecture-specific construct. (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across architectures and language semantics, and therefore amenable
-to a very general approach.)
-The desired behavior of the stack varies based on ISA, calling convention,
-and the features of whichever programming language originated the code.
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
-must be accompanied by tweaks to the model.
-%
-\rb{This point is extremely important!}
 
@@
   and occasionally too detailed for the introduction?}
+
+\rb{Begins to anticipate subsequent discussion. This is even something of an
+  understatement: some papers (esp. newer ones?), make concrete claims, though
+  they remain informal. Arguably, none have a truly formal characterization in
+  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
+  of the novelties of this work!
+\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
+  than previous work, without showing a fully general framework (which would be
+  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
+  start because of their flexibility, elegance, enabling a wide of interesting
+  experiments, etc.
+\rb{Basically, none reach the point of being an independent (and executable) specification}
+}
+}
+%
+Previous attempts
+to produce a formal characterization do so using a {\it fully abstract overlay semantics},
+a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
+But this approach is not suitable for the {\em lazy} enforcement mechanism
+of Roessler and DeHon, in which stores are not checked, and illegal stores are
+only caught when their location is later read. We aim to provide a stack-safety
+property suitable for testing lazy enforcement.
+%
+\rb{Need to discuss overlay semantics, but also note that it is very intensional
+  and very specific, as the authors acknowledge. I'd introduce this a bit more
+  gradually. Also, if phrased this way, it sounds like its main drawback is a
+  lack of generality, to which we should contrast a more general approach, not
+  just one that is specialized to a specific policy, i.e. lazy micro-policies
+  (I'd try to present these more gradually as well).}
+
+\begin{table*}
+  \begin{tabular}{| l | l | l | l | l | l | l | l | l |}
+    \hline
+    Technique       & Approximate & Deferred & High-level & Partial & Well-bracketed & Integrity & Confidentiality & Formal \\
+    \hline
+    Stack canaries  & Yes         & Yes      & No         & Yes     & Yes            & No        & No              & XXX    \\
+    Shadow stacks   & Yes         & Yes      & No         & Yes     & Yes            & No        & No              & XXX    \\
+    Bounds checking & No          & No       & Yes        & No      & Yes            & XXX       & XXX             & XXX    \\
+    Split stacks    & No          & No       & Yes        & Yes     & Yes            & No        & No              & XXX    \\
+    Capabilities    &             &          &            &         &                &           &                 & XXX    \\
+    Micro-policies  &             & Yes      &            & Yes     & Yes            & Yes       & Yes             & XXX    \\
+    \hline
+  \end{tabular}
+  \caption{Comparison of Stack Protection Mechanisms}
+  \label{tab:relwork}
+\end{table*}
+
+To demonstrate the utility of our formal characterization, we use these
+properties to validate and improve an existing enforcement mechanism, the
+{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
+in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
+use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for Coq, to generate random programs and check
+that Roessler and DeHon's micro-policies correctly detect the ones that
+attempt to violate one of our properties. Furthermore, we
+check that the testing framework is able to generate counterexamples
+that violate our properties but are \emph{not} halted by incorrect
+variants of the enforcement mechanisms, including ones that we accidentally created
+while implementating the micro-policy and ones that we
+intentionally crafted to increase our confidence in the effectiveness
+of our testing.
+%
+\rb{This and the following paragraph could start more generally (we need to
+  answer the question of why these experiments, what is their interest), and
+  probably motivate the use of PBT (experiment results below show there's
+  benefit in them!) and, somewhere at least, talk about other points in the
+  spectrum.}
+
+We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
+violates the temporal aspect of confidentiality in
+cases where data can leak across repeated calls to the same callee,
+and also violates integrity if the leak uses the caller's frame. We
+propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
+confidentiality, albeit at some performance cost.
+%
+\rb{Anything to add from personal communications? More details later? Shows
+  these things are very hard to get right -- like the latest work on Cerise,
+  also motivated by this very work.
+\rb{(+SNA) At the very least, these models are subtle and getting them right,
+  and describing them unambiguously, is hard without a framework like this.}
+}
 

### 2023-01-27 submission5/intro-v2.tex
Draft rest of intro, checkpoint on new key ideas

@@
 verifying the protection they offer.
+%
+For an assessment of these methods in the context of our stack security
+framework, see \cref{XXX} below.
 
@@
 %
-Capability machines are able to represent and manipulate unforgeable tokens of
-authority over specific memory regions using a combination of dedicated
-out-of-band memory and a collection of specialized ISA instructions.
+Capability machines can represent and manipulate their homonymous
+capabilities---unforgeable tokens of authority over specific regions of
+memory---using a combination of dedicated out-of-band memory and a collection of
+specialized ISA instructions.
 %
-These hardware capabilities can be used to implement \emph{secure calling
-conventions} \ldots
+The most prominent contemporary representative of this long line of work is
+CHERI \cite{}, a modern architecture designed to provide efficient fine-grained
+memory protection and compartmentalization.
+%
+Previous work has demonstrated the use of these general principles \rb{in
+  simplified models} to implement \emph{secure calling conventions} for the call
+stack \cite{}. Those conventions can be seen as combining the advantages and
+protections of various families of software-based methods.
+%
+They are closely related \rb{too strong?} to this work; we discuss them in more
+detail in \cref{XXX}.
+%
+\rb{HardBound is also related to this}
 
@@
 %
-\rb{(This partly overlaps with the technical preliminaries)}
+\rb{(This partly overlaps with the technical preliminaries, try to offer a quick
+  synthesis here.)}
 %
-\ldots
+Micro-policy machines are a family of tagged architectures that decorate
+addressable memory and registers with arbitrary metadata tags, and extend the
+processor with a programmable hardware monitor that uses custom rules on
+metadata tags to implement security policies of various types.
+%
+This flexibility can be harnessed to implement a wide array of stack protection policies with relative ease, covering and going beyond the capabilities \rb{word?} of the various software-based techniques \cite{} \rb{as demonstrated by}.
+%
+They comprise very general forms of \emph{lazy enforcement}, which can tolerate
+potential but harmless violations \rb{the system programmer's motto: ``trust me,
+  I know what I'm doing''} while still detecting them if and when their effects
+become tangible.
+%
+Micro-policies are one of the centerpieces of our technical development; we
+present them in more detail in \cref{XXX}.
 
@@
 
-Stack safety is challenging to specify in full generality, because the stack is fundamentally
-an architecture-specific construct. (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across architectures and language semantics, and therefore amenable
-to a very general approach.)
-The desired behavior of the stack varies based on ISA, calling convention,
-and the features of whichever programming language originated the code.
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
-must be accompanied by tweaks to the model.
+As we begin to intuit, the security of the call stack and its contents can be
+viewed through many different lenses. Each of these points of view results in a
+\emph{unique specification} of the meaning of \emph{stack safety} for a
+particular protection mechanism---mostly an informal specification if one is
+produced, and almost always left implicit to some degree.
 %
-\rb{This point is extremely important!}
+Moreover, different protections operate under different threat models, further
+complicating their comparison and the assessment of the safeguards they
+effectively offer their clients---the programs themselves.
+%
+On the face of it, it seems perplexing that there is not a \emph{standard
+specification} (or set of specifications) for such a fundamental problem.
+Instead of reverse engineering an specific enforcement mechanism obtain a
+(necessarily approximate) specification, we turn the problem around on its head
+and seek more general solutions from first principles.
+%
+\rb{(Problems/notes: ``backward specs'' not the most useful to their clients,
+  not designed with their security needs in mind---goes in the direction of
+  program logics---; to reason about security in complex systems an explicit and
+  clear spec is preferable; many of the informal specs are at least
+  superficially similar and overlapping; hopefully all of them can be understood
+  from a small number of fundamental concepts)}
 
@@
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the control flow protection
+and the callee's behavior during the callee's execution, plus the control-flow protection
 of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
@@
 %
-\rb{Connect this to the shortcomings of previous work. I'd give at least a quick
-  explanation of what the various properties roughly entail in practice (e.g.,
-  secrecy of private data, etc.). The comment about the single extension of
-  exceptions seems disconnected from the rest.}
-
-This gives us a total of five core properties: \(\wbcf\)
+We treat stack safety as multiple properties, rather than one multi-faceted property, because
+not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
+exclusively on control flow, some enforce integrity but not confidentiality, and many
+focus on the caller over the callee.
+%
+\rb{Return to this after extended discussion?}
+%
+This gives us a total of five core properties: \(\wbcf\),
 caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
 callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
+%
+\rb{Forward references?}
 
@@
 %
-\rb{Sharpen, contrast the universal definitions (or a framework of parametric
-  definitions, say) with the concrete instance we study, again with its interest
-  properly motivated.}
+\rb{FIXME: Contrast some universal definitions (or a parametric framework) with
+  the concrete instance we study, again with its interest properly motivated.
+  (Coming in key ideas?) No ``gold standard''?}
 
-We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
-exclusively on control flow, some enforce integrity but not confidentiality, and many
-focus on the caller over the callee.
+Stack safety is challenging to specify in full generality, because the stack is fundamentally
+an architecture-specific construct. (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
+which is broadly consistent across architectures and language semantics, and therefore amenable
+to a very general approach.)
+The desired behavior of the stack varies based on ISA, calling convention,
+and the features of whichever programming language originated the code.
 %
-\rb{Also sharpen after extended discussion?}
+\rb{(Extremely important!)}
+%
+For instance, the current state-of-the-art stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
+must be accompanied by tweaks to the model.
+%
+\rb{(Make this second point later?)}
+
+To demonstrate the utility of our formal characterization, we use these
+properties to validate and improve an existing enforcement mechanism, the
+{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
+in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
+use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for Coq, to generate random programs and check
+that Roessler and DeHon's micro-policies correctly detect the ones that
+attempt to violate one of our properties. Furthermore, we
+check that the testing framework is able to generate counterexamples
+that violate our properties but are \emph{not} halted by incorrect
+variants of the enforcement mechanisms, including ones that we accidentally created
+while implementating the micro-policy and ones that we
+intentionally crafted to increase our confidence in the effectiveness
+of our testing.
+%
+\rb{This and the following paragraph could start more generally (we need to
+  answer the question of why these experiments, what is their interest), and
+  probably motivate the use of PBT (experiment results below show there's
+  benefit in them!) and, somewhere at least, talk about other points in the
+  spectrum.}
+
+We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
+violates the temporal aspect of confidentiality in
+cases where data can leak across repeated calls to the same callee,
+and also violates integrity if the leak uses the caller's frame. We
+propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
+confidentiality, albeit at some performance cost.
+%
+\rb{Anything to add from personal communications? More details later? Shows
+  these things are very hard to get right -- like the latest work on Cerise,
+  also motivated by this very work.
+\rb{(+SNA) At the very least, these models are subtle and getting them right,
+  and describing them unambiguously, is hard without a framework like this.}
+}
+
+\subsection{Contributions}
+
+In sum, we offer the following contributions:
+
+\begin{itemize}
+\item We formalize permissive stack confidentiality and integrity properties,
+  which are parameterized over a notion of external
+  observation, and are violated only if accessing secrets or overwriting
+  data causes a visible change in system behavior.
+\item We instantiate these definitions with a realistic
+  system with features such as argument passing
+  on the stack, callee-saves registers, and tail-call elimination.
+  Our model is modular enough that adding these features is straightforward.
+\item Our formalization of well-bracketed control flow in the presence of tail-call
+  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
+\item We use property-based random testing to validate the relationship between
+  our properties and micro-policies.
+\end{itemize}
+
+Our artifact contains Coq formalizations of our properties for illustrative purposes
+as well as their testing implementation. There are no proofs---we use primarily Coq for the
+QuickChick library and to ensure that the formalizations are unambiguous.
+%
+\rb{Will have another look after the full pass.}
+
+\rb{*** KEY IDEAS FROM HERE? ***}
 
@@
 
-\begin{table*}
-  \begin{tabular}{| l | l | l | l | l | l | l | l | l |}
-    \hline
-    Technique       & Approximate & Deferred & High-level & Partial & Well-bracketed & Integrity & Confidentiality & Formal \\
-    \hline
-    Stack canaries  & Yes         & Yes      & No         & Yes     & Yes            & No        & No              & XXX    \\
-    Shadow stacks   & Yes         & Yes      & No         & Yes     & Yes            & No        & No              & XXX    \\
-    Bounds checking & No          & No       & Yes        & No      & Yes            & XXX       & XXX             & XXX    \\
-    Split stacks    & No          & No       & Yes        & Yes     & Yes            & No        & No              & XXX    \\
-    Capabilities    &             &          &            &         &                &           &                 & XXX    \\
-    Micro-policies  &             & Yes      &            & Yes     & Yes            & Yes       & Yes             & XXX    \\
-    \hline
+\newcommand*{\rot}[1]{\hbox to1em{\hss\rotatebox[origin=br]{-60}{#1}}}
+\newcommand*{\tabyes}{\CIRCLE\xspace}
+\newcommand*{\tabno}{\Circle\xspace}
+\newcommand*{\tabjein}{\RIGHTcircle\xspace}
+
+\begin{table}
+\begin{threeparttable}
+  \begin{tabular}{l | c c c c | c c c | c}
+    \toprule
+    \multicolumn{1}{l |}{Mechanism}
+    & \multicolumn{4}{l |}{Features}
+    & \multicolumn{3}{l |}{Properties}
+    & %% Formal criteria
+    \\
+    \midrule
+    %% Technique
+    \multicolumn{1}{c}{}
+    & \rot{Strict enforcement}
+    & \rot{Deferred checking}
+    & \rot{Low-level code support}
+    & \multicolumn{1}{c}{\rot{Partial program support}}
+    & \rot{Well-bracketed control flow}
+    & \rot{Data integrity}
+    & \multicolumn{1}{c}{\rot{Data confidentiality}}
+    & \rot{Formal criteria}
+    \\
+    \midrule
+    Stack canaries  & \tabno   & \tabjein & \tabyes  & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
+    Shadow stacks   & \tabno   & \tabjein &          & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
+    Bounds checking & \tabyes  & \tabno   & \tabno   & \tabno   & \tabyes  &          &          &          \\
+    Split stacks    & \tabyes  & \tabno   & \tabno   & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
+    \midrule
+    Capabilities    &          & \tabno   &          & \tabyes  &          &          &          & \tabjein \\
+    Micro-policies  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  \\
+    \bottomrule
   \end{tabular}
   \caption{Comparison of Stack Protection Mechanisms}
+  \begin{tablenotes}
+  \item \tabyes indicates a protection mechanism possesses the feature
+  \item \tabno indicates the mechanism lacks the feature
+  \item \tabjein indicates partial or incomplete support
+  \item \rb{Complete and check columns, describe. Capitalize?}
+  \item \rb{Order? Split int/conf? Deferred/lazy at the end of features}
+  \item \rb{2+ lines for capabilities, before and after using our draft?}
+  \end{tablenotes}
   \label{tab:relwork}
-\end{table*}
-
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly detect the ones that
-attempt to violate one of our properties. Furthermore, we
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms, including ones that we accidentally created
-while implementating the micro-policy and ones that we
-intentionally crafted to increase our confidence in the effectiveness
-of our testing.
-%
-\rb{This and the following paragraph could start more generally (we need to
-  answer the question of why these experiments, what is their interest), and
-  probably motivate the use of PBT (experiment results below show there's
-  benefit in them!) and, somewhere at least, talk about other points in the
-  spectrum.}
-
-We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-%
-\rb{Anything to add from personal communications? More details later? Shows
-  these things are very hard to get right -- like the latest work on Cerise,
-  also motivated by this very work.
-\rb{(+SNA) At the very least, these models are subtle and getting them right,
-  and describing them unambiguously, is hard without a framework like this.}
-}
+\end{threeparttable}
+\end{table}
 
@@
 %
+CHERI and lightweight formal methods?
+%
+Questionably that they are strictly more realistic in the sense that it is very
+costly or requires special sophisticated/experimental types of capability not
+available/perhaps not very feasible to implement.
+%
+Problems of adapting this work to capability machines: security semantics tightly
+coupled to regular machine semantics and to a complex calling convention, with
+more complex side conditions...
+%
+\\
+%
 Why favor micro-policies? Their design is more flexible and arguably more
@@
 
-\subsection{Contributions}
-
-In sum, we offer the following contributions:
-
-\begin{itemize}
-\item We formalize permissive stack confidentiality and integrity properties,
-  which are parameterized over a notion of external
-  observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in system behavior.
-\item We instantiate these definitions with a realistic
-  system with features such as argument passing
-  on the stack, callee-saves registers, and tail-call elimination.
-  Our model is modular enough that adding these features is straightforward.
-\item Our formalization of well-bracketed control flow in the presence of tail-call
-  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
-\item We use property-based random testing to validate the relationship between
-  our properties and micro-policies.
-\end{itemize}
-
-Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as their testing implementation. There are no proofs---we use primarily Coq for the
-QuickChick library and to ensure that the formalizations are unambiguous.
-%
-\rb{Will have another look after the full pass.}
-
 \paragraph*{Limitations}
@@
   capability machine semantics?}
+
+%%%
+
+\rb{Connect this (novel characterization?) to the shortcomings of previous work (later!). I'd give at least a quick
+  explanation of what the various properties roughly entail in practice (e.g.,
+  secrecy of private data, etc.). The comment about the single extension of
+  exceptions (still in intro) seems disconnected from the rest.}

### 2023-01-28 submission5/core.tex
Swap introductions for build
Temporarily move key ideas to a separate file

@@
 \usepackage[table]{xcolor}
+
+\usepackage{threeparttable}
+
+\usepackage{wasysym}
 
@@
 
-\input{intro-v1}
+\input{intro-v2}
 
@@
 
-The paper is structured as follows. In \cref{sec:example}, we will walk through a function
-call in a simple example machine and discuss how each of our properties applies to it,
-informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
-its {\em security semantics}, described in brief below. We formalize the definitions of
-stack safety that derive from the security semantics in \cref{sec:props}.
-\Cref{sec:testing} describes the testing framework, and
-\cref{sec:relwork,sec:future} discuss related and future work.
-The remainder of this section introduces ideas that will be important throughout.
-
-\paragraph*{Security Semantics and Property Structure}
-
-A {\it security semantics} extends a machine
-with additional context about the identity of current and pending
-functions (which act as security principals) and about the registers and memory they require
-to be secure. This added context is purely notional;
-it does not affect the real machine. The security context
-evolves dynamically through the execution of {\it security-relevant operations},
-which include events like calls, returns, and frame manipulations.
-Our security properties are phrased in terms of this context, often as predicates
-on future states, e.g. of the form ``when control returns to the current function...'',
-or relations on traces of future execution (hyper-properties).
-
-The security-relevant operations typically correspond to underling machine instructions,
-but the correspondence may not be obvious just by inspecting the machine code.
-For example, in the tagged RISC-V machine we use in our examples and tests,
-calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
-and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
-instructions might also be used for other things. An instantiation of an operation
-might require some external mechanism for distinguishing which instructions are
-security-relevant; the most obvious technique is to add optional labels to instructions,
-but the theory does not depend on this choice.
-
-The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
-\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
-are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In \cref{sec:machine}), we cover additional operations that add parameters to the examples',
-to support various forms of memory sharing \ifexceptions , exceptions, \fi
-and tail-call elimination. These are shown in light gray.
-We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
-based on pointer provenance in \cref{app:ptr}, represented by the dark grey
-operations, but we do not test this.
-
-\newcommand{\example}{\rowcolor{black!0}}
-\newcommand{\testing}{\rowcolor{black!10}}
-\newcommand{\theory}{\rowcolor{black!25}}
-
-\begin{table}
-  \begin{tabular}{| l | l |}
-    \hline
-    \(\psi \in \Psi\) & Parameters \\
-    \hline
-    \example \(\mathbf{call}\) & Target address, argument registers \\
-    \testing & Stack argument offset \& size \\
-    \example \(\mathbf{return}\) & \\
-    \example \(\mathbf{alloc}\) & Offset \& size \\
-    \testing & Public flag \\
-    \example \(\mathbf{dealloc}\) & Offset \& size \\
-    \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
-    \hline
-    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
-    \hline
-    \theory \(\mathbf{promote}\) & Register, offset \& size \\
-    \theory \(\mathbf{propagate}\) & Source register/address \\
-    \theory & Destination register/address \\
-    \theory \(\mathbf{clear}\) & Target register/address \\
-    \hline
-  \end{tabular}
-  \caption{Security-relevant Operations and their Parameters}
-  \label{tab:psi}
-\end{table}
-
-Our security semantics has similarities to the overlay semantics that have been proposed
-to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
-the behavior of the underlying machine in any way. Rather, it tracks additional context
-about the history of security-relevant operations, which inform the criterion a machine
-must satisfy in order to correctly obey the properties.
-
-The security principals in our account of stack safety are function
-activations: at any given time, pending activations
-have security requirements that are recorded in a call stack, and the current
-activation tracks context information that will inform its requirements when
-it makes a call. Each activation has a {\it view}
-of the system that maps each state element to a {\it security class}:
-it may be private to an inactive principal (\(\sealed\)),
-in an allocated object that is accessible to the current principal (\(\object\)),
-available to be allocated (\(\unsealed\)), or
-always accessible (\(\public\)).
-
-\sna{Might like a better name than object, because morally it ought to apply to arguments as well.}
-
-\paragraph*{The Properties and their Hierarchy}
-
-The properties form a hierarchy, shown in [TODO: diagram].
-Within properties that are orthogonal
-some are likely to take priority in terms of their
-importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
-as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
-Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
-and callee-oriented properties even moreso.
-
-\paragraph*{Events}
-
-It is common to specify the behavior of a system, including security properties, in
-terms of traces of observable {\em events} rather than details of the machine state.
-Lazy enforcement works because the deferred check catches errors before they appear
-in the trace. The nature of events is a parameter of the machine, and can be specialized
-to any notion of observable behavior.
-
-\paragraph*{Threat Model}
-
-We must trust that our method of distinguishing security-relevant operations is accurate; if it
-involves labels placed on code by a compiler, that means trusting that the compiler placed
-those labels correctly. If operations occur that are not recognized, those operations
-might not be guaranteed to protect their principals---for instance, an unlabeled call
-might not protect the caller's data from the callee. On the flip side, applying an incorrect
-label most likely means that the property becomes too strong to be enforced.
-
-Otherwise, we do not assume that the code adheres to any particular
-calling convention or implements a source language construct.
-In particular, while we are agnostic as to the source
-language, we certainly aim to support C, and so any source function might contain undefined
-behavior resulting in its compilation to arbitrary machine code. A given enforcement
-mechanism may place additional constraints, particularly on the behavior of
-call and return sequences. For instance, extant implementations tend to assume
-implicitly that callee-saved registers are treated appropriately by whichever compiler
-generated their code. Our properties explicitly require them to be treated as belonging
-to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
-
-In general, it is impossible to distinguish buggy machine code from an attacker; in
-our examples we will identify one function or another as an attacker, but we do not
-require any static division between trusted and untrusted code, and we aim to protect
-even buggy code.
-
-This is a strong threat model, but it does omit some important aspects of
-security enforcement: In particular, hardware and timing attacks are out of scope.
+\include{ideas-v1}
 

### 2023-01-28 submission5/ideas-v1.tex
Temporarily move key ideas to a separate file

The paper is structured as follows. In \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
its {\em security semantics}, described in brief below. We formalize the definitions of
stack safety that derive from the security semantics in \cref{sec:props}.
\Cref{sec:testing} describes the testing framework, and
\cref{sec:relwork,sec:future} discuss related and future work.
The remainder of this section introduces ideas that will be important throughout.

\paragraph*{Security Semantics and Property Structure}

A {\it security semantics} extends a machine
with additional context about the identity of current and pending
functions (which act as security principals) and about the registers and memory they require
to be secure. This added context is purely notional;
it does not affect the real machine. The security context
evolves dynamically through the execution of {\it security-relevant operations},
which include events like calls, returns, and frame manipulations.
Our security properties are phrased in terms of this context, often as predicates
on future states, e.g. of the form ``when control returns to the current function...'',
or relations on traces of future execution (hyper-properties).

The security-relevant operations typically correspond to underling machine instructions,
but the correspondence may not be obvious just by inspecting the machine code.
For example, in the tagged RISC-V machine we use in our examples and tests,
calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
instructions might also be used for other things. An instantiation of an operation
might require some external mechanism for distinguishing which instructions are
security-relevant; the most obvious technique is to add optional labels to instructions,
but the theory does not depend on this choice.

The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
are shown in white: calls, returns, and the allocation and deallocation of local memory.
In \cref{sec:machine}), we cover additional operations that add parameters to the examples',
to support various forms of memory sharing \ifexceptions , exceptions, \fi
and tail-call elimination. These are shown in light gray.
We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
based on pointer provenance in \cref{app:ptr}, represented by the dark grey
operations, but we do not test this.

\newcommand{\example}{\rowcolor{black!0}}
\newcommand{\testing}{\rowcolor{black!10}}
\newcommand{\theory}{\rowcolor{black!25}}

\begin{table}
  \begin{tabular}{| l | l |}
    \hline
    \(\psi \in \Psi\) & Parameters \\
    \hline
    \example \(\mathbf{call}\) & Target address, argument registers \\
    \testing & Stack argument offset \& size \\
    \example \(\mathbf{return}\) & \\
    \example \(\mathbf{alloc}\) & Offset \& size \\
    \testing & Public flag \\
    \example \(\mathbf{dealloc}\) & Offset \& size \\
    \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
    \hline
    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
    \hline
    \theory \(\mathbf{promote}\) & Register, offset \& size \\
    \theory \(\mathbf{propagate}\) & Source register/address \\
    \theory & Destination register/address \\
    \theory \(\mathbf{clear}\) & Target register/address \\
    \hline
  \end{tabular}
  \caption{Security-relevant Operations and their Parameters}
  \label{tab:psi}
\end{table}

Our security semantics has similarities to the overlay semantics that have been proposed
to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
the behavior of the underlying machine in any way. Rather, it tracks additional context
about the history of security-relevant operations, which inform the criterion a machine
must satisfy in order to correctly obey the properties.

The security principals in our account of stack safety are function
activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
it may be private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).

\sna{Might like a better name than object, because morally it ought to apply to arguments as well.}

\paragraph*{The Properties and their Hierarchy}

The properties form a hierarchy, shown in [TODO: diagram].
Within properties that are orthogonal
some are likely to take priority in terms of their
importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
and callee-oriented properties even moreso.

\paragraph*{Events}

It is common to specify the behavior of a system, including security properties, in
terms of traces of observable {\em events} rather than details of the machine state.
Lazy enforcement works because the deferred check catches errors before they appear
in the trace. The nature of events is a parameter of the machine, and can be specialized
to any notion of observable behavior.

\paragraph*{Threat Model}

We must trust that our method of distinguishing security-relevant operations is accurate; if it
involves labels placed on code by a compiler, that means trusting that the compiler placed
those labels correctly. If operations occur that are not recognized, those operations
might not be guaranteed to protect their principals---for instance, an unlabeled call
might not protect the caller's data from the callee. On the flip side, applying an incorrect
label most likely means that the property becomes too strong to be enforced.

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, particularly on the behavior of
call and return sequences. For instance, extant implementations tend to assume
implicitly that callee-saved registers are treated appropriately by whichever compiler
generated their code. Our properties explicitly require them to be treated as belonging
to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.

In general, it is impossible to distinguish buggy machine code from an attacker; in
our examples we will identify one function or another as an attacker, but we do not
require any static division between trusted and untrusted code, and we aim to protect
even buggy code.

This is a strong threat model, but it does omit some important aspects of
security enforcement: In particular, hardware and timing attacks are out of scope.

### 2023-01-28 submission5/ideas-v2.tex
Move leftover notes to key ideas file

--- 2023-01-25 submission5/intro-v2.tex
@@
-\subsection{The Call Stack and Its Security}
+The paper is structured as follows. In \cref{sec:example}, we will walk through a function
+call in a simple example machine and discuss how each of our properties applies to it,
+informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
+its {\em security semantics}, described in brief below. We formalize the definitions of
+stack safety that derive from the security semantics in \cref{sec:props}.
+\Cref{sec:testing} describes the testing framework, and
+\cref{sec:relwork,sec:future} discuss related and future work.
+The remainder of this section introduces ideas that will be important throughout.
 
-Computer programming at any scale relies on \emph{functions} (variously known as
-subroutines, procedures, methods, etc.) as a \emph{foundational abstraction} to
-structure programs into building blocks representing units of computation that
-can \emph{call} one another to create larger computations.
-%
-At a low level, a function call comprises its own data (private or shared), as
-well as information about its \emph{caller} and \emph{callee} functions. The
-runtime must keep track of the local contexts of all active function activations
-at any given time; the \emph{fundamental data structure} charged with realizing
-this abstraction is the \emph{(call) stack}.
-%
-\rb{This should be well known, so no need for illustration?}
+\paragraph*{Security Semantics and Property Structure}
 
-The correct execution on a program---and its security---is contingent on all of
-its functions properly implementing this abstraction by means of a careful use
-of the call stack. Be it accidental or intentional, incorrect management of the
-stack is at the root of a myriad of faults whose effects range from (rarely)
-innocuous to (usually) catastrophic.
-%
-From a security perspective, attacks on the stack (say, from a malicious
-function that chooses to abuse its access to the contents of this data
-structure) are as fundamental, varied and pervasive as the functional
-abstraction itself.
-%
-\rb{Begins to prefigure the idea of multiple stack implementations.}
+A {\it security semantics} extends a machine
+with additional context about the identity of current and pending
+functions (which act as security principals) and about the registers and memory they require
+to be secure. This added context is purely notional;
+it does not affect the real machine. The security context
+evolves dynamically through the execution of {\it security-relevant operations},
+which include events like calls, returns, and frame manipulations.
+Our security properties are phrased in terms of this context, often as predicates
+on future states, e.g. of the form ``when control returns to the current function...'',
+or relations on traces of future execution (hyper-properties).
 
-%% \rb{\textbf{The call stack and its security.} Maybe a few opening words to set
-%%   the stage and bring two things to the front: (1) functions and subroutines as
-%%   fundamental building blocks of any kind of structured programming; and (2) the
-%%   stack as the natural abstraction to manage these in a computer. And also maybe
-%%   a little bit more about the width, the gravity and the \emph{foundational
-%%   nature} of attacks on the stack, which the sentence below touches on
-%%   summarily.
-%% \rb{(+SNA) Connection to functions (and adversarial code?). Attacks on the stack
-%%   are also attacks on this fundamental abstraction (functions, subroutines,
-%%   procedures, methods, however you call them)}
-%% }
+The security-relevant operations typically correspond to underling machine instructions,
+but the correspondence may not be obvious just by inspecting the machine code.
+For example, in the tagged RISC-V machine we use in our examples and tests,
+calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
+and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
+instructions might also be used for other things. An instantiation of an operation
+might require some external mechanism for distinguishing which instructions are
+security-relevant; the most obvious technique is to add optional labels to instructions,
+but the theory does not depend on this choice.
 
-Indeed, the call stack is a perennial target for low-level attacks, leading to
-dire consequences, from leakage and corruption of private stack data
-to control-flow hijacking.
-%
-Even a cursory glance at the most recent release of the authoritative CWE Top 25
-Most Dangerous Software Weaknesses
-list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
-shows general classes of vulnerabilities closely related to the stack \rb{(though not exclusively)}
-consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read), \#7 (CWE-416, Use
-After Free) or \#11 (CWE-476, NULL Pointer Dereference).
-%
-Moreover, many other types of faults involve some kind of control or data flow
-using data stored in the stack \rb{cf. Eternal war in memory, Google/Microsoft
-memory safety statistics, etc.}.
+The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
+\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
+are shown in white: calls, returns, and the allocation and deallocation of local memory.
+In \cref{sec:machine}), we cover additional operations that add parameters to the examples',
+to support various forms of memory sharing \ifexceptions , exceptions, \fi
+and tail-call elimination. These are shown in light gray.
+We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
+based on pointer provenance in \cref{app:ptr}, represented by the dark grey
+operations, but we do not test this.
 
-To foil such attacks, a menagerie of
-software and hardware protections have been proposed,
-%
-including stack canaries~\cite{Cowan+98},
-bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
-split stacks~\cite{Kuznetsov+14},
-shadow stacks~\cite{Dang+15,Shanbhogue+19},
-capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}---we discuss these in more
-detail in \cref{XXX}.
-  \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\bcp{Yes, going back to MIT days---we should
-  include several more of these, if only to give readers the impression that
-this is a well-studied mechanism (so formalizing its protections is
-useful).}
-\fi
-%
-\rb{Broad enough? Do we want to add more techniques?}
-%
-Yet despite this wealth of potential solutions, the stack remains a central
-security concern---why?
-%
-Security mechanisms designed to prevent a certain attack can be brittle---successfully
-eliminating the attack as it exists at the time of publication, but leaving room for attackers
-to find new, similar attacks.
-%
-%% \rb{And so the cat-and-mouse game continues.}
+\newcommand{\example}{\rowcolor{black!0}}
+\newcommand{\testing}{\rowcolor{black!10}}
+\newcommand{\theory}{\rowcolor{black!25}}
 
-To mitigate this risk, \rb{we claim that} mechanisms should aim to enforce \emph{formal
-properties} of correct \rb{safe?} behavior that can be proven or tested. Such properties become the
-\emph{specification} against which enforcement can be validated; even enforcement mechanisms that do
-not fulfill a property benefit from the ability to articulate \emph{why} they fail.
-%
-%% \rb{Good stuff! There is some overlap between this discussion and the opening
-%%   motivation, maybe we could restructure things a bit so that those threads are
-%%   unified?}
-%
-This is the problem we begin to address in this paper \rb{or some such
-  connecting words}.
+\begin{table}
+  \begin{tabular}{| l | l |}
+    \hline
+    \(\psi \in \Psi\) & Parameters \\
+    \hline
+    \example \(\mathbf{call}\) & Target address, argument registers \\
+    \testing & Stack argument offset \& size \\
+    \example \(\mathbf{return}\) & \\
+    \example \(\mathbf{alloc}\) & Offset \& size \\
+    \testing & Public flag \\
+    \example \(\mathbf{dealloc}\) & Offset \& size \\
+    \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
+    \hline
+    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
+    \hline
+    \theory \(\mathbf{promote}\) & Register, offset \& size \\
+    \theory \(\mathbf{propagate}\) & Source register/address \\
+    \theory & Destination register/address \\
+    \theory \(\mathbf{clear}\) & Target register/address \\
+    \hline
+  \end{tabular}
+  \caption{Security-relevant Operations and their Parameters}
+  \label{tab:psi}
+\end{table}
 
-\subsection{Security Properties for Stack Safety}
+Our security semantics has similarities to the overlay semantics that have been proposed
+to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
+the behavior of the underlying machine in any way. Rather, it tracks additional context
+about the history of security-relevant operations, which inform the criterion a machine
+must satisfy in order to correctly obey the properties.
 
-The protections offered by stack protection mechanisms are commonly described by giving
-concrete examples of attacks that they can prevent---corruption of return
-addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more formal characterization to the reader's intuition.
-%
-\rb{Begins to anticipate subsequent discussion. This is even something of an
-  understatement: some papers (esp. newer ones?), make concrete claims, though
-  they remain informal. Arguably, none have a truly formal characterization in
-  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
-  of the novelties of this work!
-\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
-  than previous work, without showing a fully general framework (which would be
-  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
-  start because of their flexibility, elegance, enabling a wide of interesting
-  experiments, etc.}
-}
-%
-Previous attempts
-to produce a formal characterization do so using a {\it fully abstract overlay semantics},
-a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
-But this approach is not suitable for the {\em lazy} enforcement mechanism
-of Roessler and DeHon, in which stores are not checked, and illegal stores are
-only caught when their location is later read. We aim to provide a stack-safety
-property suitable for testing lazy enforcement.
-%
-\rb{Need to discuss overlay semantics, but also note that it is very intensional
-  and very specific, as the authors acknowledge. I'd introduce this a bit more
-  gradually. Also, if phrased this way, it sounds like its main drawback is a
-  lack of generality, to which we should contrast a more general approach, not
-  just one that is specialized to a specific policy, i.e. lazy micro-policies
-  (I'd try to present these more gradually as well).}
+The security principals in our account of stack safety are function
+activations: at any given time, pending activations
+have security requirements that are recorded in a call stack, and the current
+activation tracks context information that will inform its requirements when
+it makes a call. Each activation has a {\it view}
+of the system that maps each state element to a {\it security class}:
+it may be private to an inactive principal (\(\sealed\)),
+in an allocated object that is accessible to the current principal (\(\object\)),
+available to be allocated (\(\unsealed\)), or
+always accessible (\(\public\)).
 
-We propose a novel characterization of stack safety using the formal tools of language-based
-security~\cite{sabelfeld2003language}, decomposing stack safety into
-the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the control flow protection
-of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
-\ifexceptions
-We also extend WBCF to a setting with exceptions,
-in which there may be multiple points at which a callee can ``return.''
-\fi
-%
-\rb{Connect this to the shortcomings of previous work. I'd give at least a quick
-  explanation of what the various properties roughly entail in practice (e.g.,
-  secrecy of private data, etc.). The comment about the single extension of
-  exceptions seems disconnected from the rest.}
+\sna{Might like a better name than object, because morally it ought to apply to arguments as well.}
 
-This gives us a total of five core properties: \(\wbcf\)
-caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
+\paragraph*{The Properties and their Hierarchy}
 
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly detect the ones that
-attempt to violate one of our properties. Furthermore, we
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms, including ones that we accidentally created
-while implementating the micro-policy and ones that we
-intentionally crafted to increase our confidence in the effectiveness
-of our testing.
-%
-\rb{This and the following paragraph could start more generally (we need to
-  answer the question of why these experiments, what is their interest), and
-  probably motivate the use of PBT (experiment results below show there's
-  benefit in them!) and, somewhere at least, talk about other points in the
-  spectrum.}
+The properties form a hierarchy, shown in [TODO: diagram].
+Within properties that are orthogonal
+some are likely to take priority in terms of their
+importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
+as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
+Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
+and callee-oriented properties even moreso.
 
-We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-%
-\rb{Anything to add from personal communications? More details later? Shows
-  these things are very hard to get right -- like the latest work on Cerise,
-  also motivated by this very work.
-\rb{(+SNA) At the very least, these models are subtle and getting them right,
-  and describing them unambiguously, is hard without a framework like this.}
-}
+\paragraph*{Events}
 
-\subsection{Motivation}
+It is common to specify the behavior of a system, including security properties, in
+terms of traces of observable {\em events} rather than details of the machine state.
+Lazy enforcement works because the deferred check catches errors before they appear
+in the trace. The nature of events is a parameter of the machine, and can be specialized
+to any notion of observable behavior.
 
-Stack safety is challenging to specify in full generality, because the stack is fundamentally
-an architecture-specific construct. (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across architectures and language semantics, and therefore amenable
-to a very general approach.)
-The desired behavior of the stack varies based on ISA, calling convention,
-and the features of whichever programming language originated the code.
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
-must be accompanied by tweaks to the model.
-%
-\rb{This point is extremely important!}
+\paragraph*{Threat Model}
 
-We are not attempting to create a universal definition of stack safety for all systems.
-Rather, we define the stack safety of a single, fairly realistic system, decomposed into
-(1) a lightweight model of the system and its security-relevant features, from which we
-derive the integrity, confidentiality, and control-flow requirements of each call and
-return, (2) the application of that model to a particular target system, and (3)
-the formal criteria for integrity, confidentiality, and control flow
-to obey the requirements. We consider the criteria in (3) to represent universal definitions
-that are parameterized by (1).
-%
-\rb{Sharpen, contrast the universal definitions (or a framework of parametric
-  definitions, say) with the concrete instance we study, again with its interest
-  properly motivated.}
+We must trust that our method of distinguishing security-relevant operations is accurate; if it
+involves labels placed on code by a compiler, that means trusting that the compiler placed
+those labels correctly. If operations occur that are not recognized, those operations
+might not be guaranteed to protect their principals---for instance, an unlabeled call
+might not protect the caller's data from the callee. On the flip side, applying an incorrect
+label most likely means that the property becomes too strong to be enforced.
 
-We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
-exclusively on control flow, some enforce integrity but not confidentiality, and many
-focus on the caller over the callee.
-%
-\rb{Also sharpen after extended discussion?}
+Otherwise, we do not assume that the code adheres to any particular
+calling convention or implements a source language construct.
+In particular, while we are agnostic as to the source
+language, we certainly aim to support C, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code. A given enforcement
+mechanism may place additional constraints, particularly on the behavior of
+call and return sequences. For instance, extant implementations tend to assume
+implicitly that callee-saved registers are treated appropriately by whichever compiler
+generated their code. Our properties explicitly require them to be treated as belonging
+to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
+
+In general, it is impossible to distinguish buggy machine code from an attacker; in
+our examples we will identify one function or another as an attacker, but we do not
+require any static division between trusted and untrusted code, and we aim to protect
+even buggy code.
+
+This is a strong threat model, but it does omit some important aspects of
+security enforcement: In particular, hardware and timing attacks are out of scope.
+
+%%%
 
@@
   and occasionally too detailed for the introduction?}
+
+\rb{Begins to anticipate subsequent discussion. This is even something of an
+  understatement: some papers (esp. newer ones?), make concrete claims, though
+  they remain informal. Arguably, none have a truly formal characterization in
+  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
+  of the novelties of this work!
+\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
+  than previous work, without showing a fully general framework (which would be
+  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
+  start because of their flexibility, elegance, enabling a wide of interesting
+  experiments, etc.
+\rb{Basically, none reach the point of being an independent (and executable) specification}
+}
+}
+%
+Previous attempts
+to produce a formal characterization do so using a {\it fully abstract overlay semantics},
+a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
+But this approach is not suitable for the {\em lazy} enforcement mechanism
+of Roessler and DeHon, in which stores are not checked, and illegal stores are
+only caught when their location is later read. We aim to provide a stack-safety
+property suitable for testing lazy enforcement.
+%
+\rb{Need to discuss overlay semantics, but also note that it is very intensional
+  and very specific, as the authors acknowledge. I'd introduce this a bit more
+  gradually. Also, if phrased this way, it sounds like its main drawback is a
+  lack of generality, to which we should contrast a more general approach, not
+  just one that is specialized to a specific policy, i.e. lazy micro-policies
+  (I'd try to present these more gradually as well).}
+
+\newcommand*{\rot}[1]{\hbox to1em{\hss\rotatebox[origin=br]{-60}{#1}}}
+\newcommand*{\tabyes}{\CIRCLE\xspace}
+\newcommand*{\tabno}{\Circle\xspace}
+\newcommand*{\tabjein}{\RIGHTcircle\xspace}
+
+\begin{table}
+\begin{threeparttable}
+  \begin{tabular}{l | c c c c | c c c | c}
+    \toprule
+    \multicolumn{1}{l |}{Mechanism}
+    & \multicolumn{4}{l |}{Features}
+    & \multicolumn{3}{l |}{Properties}
+    & %% Formal criteria
+    \\
+    \midrule
+    %% Technique
+    \multicolumn{1}{c}{}
+    & \rot{Strict enforcement}
+    & \rot{Deferred checking}
+    & \rot{Low-level code support}
+    & \multicolumn{1}{c}{\rot{Partial program support}}
+    & \rot{Well-bracketed control flow}
+    & \rot{Data integrity}
+    & \multicolumn{1}{c}{\rot{Data confidentiality}}
+    & \rot{Formal criteria}
+    \\
+    \midrule
+    Stack canaries  & \tabno   & \tabjein & \tabyes  & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
+    Shadow stacks   & \tabno   & \tabjein &          & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
+    Bounds checking & \tabyes  & \tabno   & \tabno   & \tabno   & \tabyes  &          &          &          \\
+    Split stacks    & \tabyes  & \tabno   & \tabno   & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
+    \midrule
+    Capabilities    &          & \tabno   &          & \tabyes  &          &          &          & \tabjein \\
+    Micro-policies  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  \\
+    \bottomrule
+  \end{tabular}
+  \caption{Comparison of Stack Protection Mechanisms}
+  \begin{tablenotes}
+  \item \tabyes indicates a protection mechanism possesses the feature
+  \item \tabno indicates the mechanism lacks the feature
+  \item \tabjein indicates partial or incomplete support
+  \item \rb{Complete and check columns, describe. Capitalize?}
+  \item \rb{Order? Split int/conf? Deferred/lazy at the end of features}
+  \item \rb{2+ lines for capabilities, before and after using our draft?}
+  \end{tablenotes}
+  \label{tab:relwork}
+\end{threeparttable}
+\end{table}
 

### 2023-01-28 submission5/intro-v2.tex
Move leftover notes to key ideas file

@@
 \rb{Will have another look after the full pass.}
-
-\rb{*** KEY IDEAS FROM HERE? ***}
-
-\paragraph*{Permissiveness, Laziness, and Tagged Architectures}
-
-Security properties tend to be conservative, identifying violations that would be
-harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
-and of compatibility with existing codebases. In a crucial motivating example for this work,
-Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
-using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
-STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
-A set of software-defined rules, termed a micro-policy, determines the tags for operation
-results based on the tags of their inputs---or, if the input tags are in a dangerous
-combination, it may halt the machine (a ``failstop.'')
-
-Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
-is inefficient when allocating and clearing stack frames, because each word must be initialized
-with a tag corresponding to the (depth of the) current function activation. They propose
-{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
-when a function writes to memory. This means that writes cannot be prevented, necessarily
-violating existing models of stack safety. Instead the micro-policy enforces that memory be
-read by the same fuction that wrote it. The owner of the data will never see the changes,
-so it is secure in the sense that the attacker is unsuccessful at changing program behavior
-(except, perhaps, by causing a fault.) We elaborate that sense, and test whether the
-micro-policy fulfills it.
-
-This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
-machine states that make claims about the future observable behavior of execution from those states.
-That is not to say that our properties are specialized for lazy micro-policies. It also supports
-normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
-and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
-while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
-%
-\rb{Does this discussion (at least in part) belong in the general background
-  section, some even in key ideas? It still feels a little out of order to me,
-  and occasionally too detailed for the introduction?}
-
-\rb{Begins to anticipate subsequent discussion. This is even something of an
-  understatement: some papers (esp. newer ones?), make concrete claims, though
-  they remain informal. Arguably, none have a truly formal characterization in
-  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
-  of the novelties of this work!
-\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
-  than previous work, without showing a fully general framework (which would be
-  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
-  start because of their flexibility, elegance, enabling a wide of interesting
-  experiments, etc.
-\rb{Basically, none reach the point of being an independent (and executable) specification}
-}
-}
-%
-Previous attempts
-to produce a formal characterization do so using a {\it fully abstract overlay semantics},
-a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
-But this approach is not suitable for the {\em lazy} enforcement mechanism
-of Roessler and DeHon, in which stores are not checked, and illegal stores are
-only caught when their location is later read. We aim to provide a stack-safety
-property suitable for testing lazy enforcement.
-%
-\rb{Need to discuss overlay semantics, but also note that it is very intensional
-  and very specific, as the authors acknowledge. I'd introduce this a bit more
-  gradually. Also, if phrased this way, it sounds like its main drawback is a
-  lack of generality, to which we should contrast a more general approach, not
-  just one that is specialized to a specific policy, i.e. lazy micro-policies
-  (I'd try to present these more gradually as well).}
-
-\newcommand*{\rot}[1]{\hbox to1em{\hss\rotatebox[origin=br]{-60}{#1}}}
-\newcommand*{\tabyes}{\CIRCLE\xspace}
-\newcommand*{\tabno}{\Circle\xspace}
-\newcommand*{\tabjein}{\RIGHTcircle\xspace}
-
-\begin{table}
-\begin{threeparttable}
-  \begin{tabular}{l | c c c c | c c c | c}
-    \toprule
-    \multicolumn{1}{l |}{Mechanism}
-    & \multicolumn{4}{l |}{Features}
-    & \multicolumn{3}{l |}{Properties}
-    & %% Formal criteria
-    \\
-    \midrule
-    %% Technique
-    \multicolumn{1}{c}{}
-    & \rot{Strict enforcement}
-    & \rot{Deferred checking}
-    & \rot{Low-level code support}
-    & \multicolumn{1}{c}{\rot{Partial program support}}
-    & \rot{Well-bracketed control flow}
-    & \rot{Data integrity}
-    & \multicolumn{1}{c}{\rot{Data confidentiality}}
-    & \rot{Formal criteria}
-    \\
-    \midrule
-    Stack canaries  & \tabno   & \tabjein & \tabyes  & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
-    Shadow stacks   & \tabno   & \tabjein &          & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
-    Bounds checking & \tabyes  & \tabno   & \tabno   & \tabno   & \tabyes  &          &          &          \\
-    Split stacks    & \tabyes  & \tabno   & \tabno   & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
-    \midrule
-    Capabilities    &          & \tabno   &          & \tabyes  &          &          &          & \tabjein \\
-    Micro-policies  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  \\
-    \bottomrule
-  \end{tabular}
-  \caption{Comparison of Stack Protection Mechanisms}
-  \begin{tablenotes}
-  \item \tabyes indicates a protection mechanism possesses the feature
-  \item \tabno indicates the mechanism lacks the feature
-  \item \tabjein indicates partial or incomplete support
-  \item \rb{Complete and check columns, describe. Capitalize?}
-  \item \rb{Order? Split int/conf? Deferred/lazy at the end of features}
-  \item \rb{2+ lines for capabilities, before and after using our draft?}
-  \end{tablenotes}
-  \label{tab:relwork}
-\end{threeparttable}
-\end{table}
-
-\rb{\textbf{A menagerie of protection mechanisms.}
-%
-We have made explicit the key point that there is \emph{not} a single stack
-abstraction. This paves the way for a \emph{general} idea of stack protection
-that is widely applicable with high-level concepts that remain unchanged (and I
-think it's particularly important to at least introduce these more immutable
-parts early on). The fact that this general ``framework'' requires low-level or
-specific specializations is a predictable consequence of the protean nature of
-the call stack.
-%
-\\
-%
-Let us return to a selection of representative policies in our menagerie and
-consider how they fit into the framework. At the highest level, we can classify
-all those techniques into two large groups, software-based and hardware-based.
-Traditionally, most practical protections are implemented in software, so we
-start by considering this first group.
-%
-\\
-%
-We can make two general observations about these software based techniques.
-Firstly, (as we will see: discussion? technical details?) these techniques are
-\emph{less modular} than their hardware counterparts. That is, the
-countermeasures are tightly coupled with the program code they protect, and
-consequently more effort is required to disentangle the program from the
-protections, and to understand the impact of the latter on the former. This is
-true of any systematic attempt to study the security of stack protection
-policies, including their adaptation to our framework. (Owing to this added
-complexity, software policies will not be the cleanest or most elegant exemplars
-of use of the framework, its salient features and practical utility; this partly
-motivates the focus on hardware policies.)
-%
-\\
-%
-(The above remark holds \emph{at least in a ``single-level model''} that
-considers assembly language/machine code programs. In general, (low-level)
-protections are introduced by a compiler or similar tool. One could consider a
-multi-level extension of the framework pairing models of high-level programs and
-the security transformations performed by a (stack-secure) compiler.)
-%
-\\
-%
-Secondly, the guarantees generally offered by software-based policies are
-\emph{less absolute}, being defined more or less informally and usually
-providing few definite guarantees of absence of \emph{general} classes of
-errors. (The following discussion elaborates on this. It is conceivable that an
-extended and more sophisticated framework could add \emph{probabilistic
-reasoning} capabilities to capture and reason about the effects of less strict
-protection mechanisms. This would be another orthogonal development to what is
-presented here. In this paper, we are interested in assessing the complete
-\emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
-or defects. In passing, this would extend the security properties to e.g. larger
-system verification efforts.)
-%
-\\
-%
-One by one, consider the following specimens of software protection mechanisms
-(quick explanations inlined?):
-%
-\\
-%
-\begin{itemize}
-%
-\item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
-  difficulty of some attacks on the stack, usually (always?) buffer overflows.
-  They can never provide any absolute security guarantees and therefore do not
-  satisfy any of our properties (some of the properties, e.g. confidentiality,
-  are completely out of scope.) Although tedious, we could we could model this
-  mechanism in our framework with the appropriate adaptations --- and
-  unsurprisingly, testing would find property violations, which we already know
-  are there: the results of this effort would be unsurprising and not as useful
-  or instructive as other experiments.
-  \rb{(+SNA) They are kind of lazy, aren't they?}
-%
-\item \emph{Bounds checking}. This is fundamentally a high-level technique, even
-  more obviously so than stack canaries (which could still detect violations
-  caused by external code). Bounds checking only applies to code that has been
-  instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
-  impose any constraints on adversarial code. The attacker model is therefore
-  rather weak. (Certain variants, like SoftBound, are good targets for lifting
-  our framework to the source level by e.g. using a C-language semantics and
-  using Tagged C for security modeling; the lifted versions of our properties
-  would likely hold under the expected assumptions. HardBound is probably an
-  interesting transitional case because it uses capability-like hardware
-  primitives for enforcement; it could be close to ongoing and future work to
-  verify stack safety on capability machines, e.g. based on Cerise -- discuss
-  together?)
-%
-\item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
-  ``probabilistic techniques'' in that they ``restrict the flexibility available
-  in creating gadget chains'' (Shanbhogue et al.). This technique is solely
-  concerned with control-flow integrity, and so closely related to our
-  well-bracketed control flow. Nevertheless, and again, it is only an
-  approximate technique, and our testing framework would be able to find the
-  expected attacks against our security property.
-%
-\item \emph{Split stacks}. The security claims of split stacks are stronger than
-  those of most software countermeasures, i.e., they are meant to guarantee
-  their titular code-pointer integrity. They explicitly offer no protections on
-  stack data (i.e., integrity and confidentiality), and so restrict themselves
-  to well-bracketed control flow. (This is probably closer to our framework than
-  most other software techniques, and it seems like it would be easier to model
-  as well. However, this is again work based on high-level languages and would
-  benefit from a higher level of abstraction.)
-%
-\item \ldots
-%
-\end{itemize}
-%
-Opposed to the above group of software techniques we have hardware-based
-protection policies (more in background/technical preliminaries?). Although
-programs must effectively leverage the protection mechanisms offered by the
-hardware, the distinction between those protections and the program logic itself
-is much cleaner: it is fairly simple to separate the use of the security
-features in the hardware (and their effects in program code) from regular
-program code. They also tend to aspire to offer complete protection against
-entire classes of attacks. Both aspects make this family of techniques a better
-fit for illustrating and showcasing our techniques.
-%
-\\
-%
-CHERI and lightweight formal methods?
-%
-Questionably that they are strictly more realistic in the sense that it is very
-costly or requires special sophisticated/experimental types of capability not
-available/perhaps not very feasible to implement.
-%
-Problems of adapting this work to capability machines: security semantics tightly
-coupled to regular machine semantics and to a complex calling convention, with
-more complex side conditions...
-%
-\\
-%
-Why favor micro-policies? Their design is more flexible and arguably more
-general than capability machines (of course, they pay a performance price for
-this). This generality and ease of programming make them good candidates for an
-exploration the design space of stack protection policies and proof of concept
-implementations. On the practical side, there is a number of existing policies
-available to examine (not just lazy policies!) --- and we find how they satisfy
-our properties, how they do not, and errors in some of them, and see how to
-correct them. Lazy enforcement is particularly novel, but not the only
-possibility. This additional breadth is beneficial for the proof of concept
-stage.
-%
-\\
-%
-Regarding capability machines, we know that most designs based on capability
-machines do not satisfy our properties (or do so incurring significant costs),
-and that fixes are not trivial and involve new types of capabilities (in a
-sense, are they any more ``realistic?'' Only the latest Cerise paper would seem
-to satisfy our properties, is directly influenced by previous versions of this
-work, and is a reasonable target for ongoing/further study -- but this is not
-easy, and even the author(s) estimate it would be a significant amount of work,
-especially full proofs (references, communications, etc.).
-\rb{(+SNA) Discuss the degree in which the various papers satisfy the
-  properties\ldots and the degrees to which they \emph{aim to}. E.g. originally
-  they don't talk about confidentiality. Do they have callee integrity? Maybe
-  they aren't thinking about it? (This kind of discussion is a good motivation
-  for the formal properties. Also related to the hierarchy. And: how to convey
-  that the latest paper draws inspiration directly from this work?}
-%
-\\
-%
-\ldots
-%
-}
-\rb{(+SNA) Any interesting tables comparing specimens in the menagerie?}
-\rb{(+SNA) A general multi-level extension of the framework would need to model
-  multiple source languages. Among other things, the stack is the meeting place
-  of all sorts of different language abstractions! One of the advantages of the
-  single-level model is that it lets us be agnostic about the language. (On the
-  other hand, it is likely that a single source and target mediated by the
-  compiler suffice to model many existing techniques, coming e.g. from systems
-  programming and C-like languages.}
-
-\paragraph*{Limitations}
-
-%Our concurrency model is fairly
-%simplistic, assuming a fixed number of threads each with its own dedicated processor.
-We model memory-safe stack objects, but not a heap. Regions outside of
-stacks can be used however the compiler likes, including as a heap, but no protection is
-built in and our properties assume that if a pointer to a stack object is stored there,
-it is permanently compromised.
-
-Our properties are termination insensitive: in order to support enforcement mechanisms
-that failstop rather than execute dangerous code, the properties must treat failstops
-as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
-This is consistent with our treatment of properties that specify the state after a return,
-in that a callee that diverges for any reason will never reach the return, and will fulfill
-the property vacuously.
-%
-\rb{Limitations in the introduction or later? Orthogonality of the heap, morally
-  similar interesting setups on the stack? Termination sensitivity and
-  capability machine semantics?}
-
-%%%
-
-\rb{Connect this (novel characterization?) to the shortcomings of previous work (later!). I'd give at least a quick
-  explanation of what the various properties roughly entail in practice (e.g.,
-  secrecy of private data, etc.). The comment about the single extension of
-  exceptions (still in intro) seems disconnected from the rest.}

### 2023-01-29 submission5/ideas-v2.tex
Work on key ideas

@@
 The remainder of this section introduces ideas that will be important throughout.
+%
+%% \rb{Or shorter: references in contributions + final sentence at end of intro}
 
-\paragraph*{Security Semantics and Property Structure}
+%% \rb{Turn these back into paragraphs later?}
+\subsection{Permissiveness, Laziness, and Tagged Architectures}
+
+Security properties tend to be conservative, identifying violations that would be
+harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
+and of compatibility with existing codebases. In a crucial motivating example for this work,
+Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
+using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
+STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
+A set of software-defined rules, termed a micro-policy, determines the tags for operation
+results based on the tags of their inputs---or, if the input tags are in a dangerous
+combination, it may halt the machine (a ``failstop.'')
+
+Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
+is inefficient when allocating and clearing stack frames, because each word must be initialized
+with a tag corresponding to the (depth of the) current function activation. They propose
+{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
+when a function writes to memory. This means that writes cannot be prevented, necessarily
+violating existing models of stack safety. Instead the micro-policy enforces that memory be
+read by the same fuction that wrote it. The owner of the data will never see the changes,
+so it is secure in the sense that the attacker is unsuccessful at changing program behavior
+(except, perhaps, by causing a fault.) We elaborate that sense, and test whether the
+micro-policy fulfills it.
+
+This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
+machine states that make claims about the future observable behavior of execution from those states.
+That is not to say that our properties are specialized for lazy micro-policies. It also supports
+normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
+and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
+while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
+%
+\rb{Emphasize the point about eager properties, here, when discussing other
+  enforcement types?}
+%
+\rb{(Mentioned properties are still ``forward references'' of a sort as they
+  haven't been formally introduced yet)}
+%% \rb{Does this discussion (at least in part) belong in the general background
+%%   section, some even in key ideas? It still feels a little out of order to me,
+%%   and occasionally too detailed for the introduction?}
+
+\rb{Moved last bit to the framework discussion below, does this work?}
+
+\rb{Test: the comparison of enforcement mechanisms \emph{and} the hierarchy of
+  properties \emph{before} the security semantics. Does this work? Easy to move
+  around/consolidate. Some things still feel a bit out of order. The comparison
+  motivates the hierarchy, and also parts of the security semantics -- too much
+  at once? Also: bad subsection/paragraph title ahead.}
+
+\subsection{A Formal and Flexible Security Framework}
+
+While a concrete aim is to provide a stack-safety property suitable for testing
+lazy enforcement for a micro-policy architecture, the formal framework we
+develop could be adapted to other protection mechanisms.
+%
+The need for these policy-specific adaptations is a natural and predictable
+consequence of the protean nature of the call stack (discussed in \cref{XXX}),
+and does not detract from the generality of our properties and their
+underlying \rb{high-level?} design principles.
+%
+Nevertheless, some enforcement methods lends themselves to modeling more easily
+than others. Micro-policies in particular stand out for their elegance, which
+results in particularly clean instances of our properties, and their
+flexibility, which enables them to implement very general forms of stack
+protection \rb{(and so we get to exercise our properties fully)}.
+
+Let us revisit our menagerie of stack protections from \cref{XXX} to motivate
+our choices. \cref{tab:relwork} summarizes and compares those techniques along
+several representative axes.
+
+\newcommand*{\rot}[1]{\hbox to1em{\hss\rotatebox[origin=br]{-60}{#1}}}
+\newcommand*{\tabyes}{\CIRCLE\xspace}
+\newcommand*{\tabno}{\Circle\xspace}
+\newcommand*{\tabjein}{\RIGHTcircle\xspace}
+
+\begin{table}
+\begin{threeparttable}
+  \begin{tabular}{l | c c c c | c c c | c}
+    \toprule
+    \multicolumn{1}{l |}{Mechanism}
+    & \multicolumn{4}{l |}{Features}
+    & \multicolumn{3}{l |}{Properties}
+    & %% Formal criteria
+    \\
+    \midrule
+    %% Technique
+    \multicolumn{1}{c}{}
+    & \rot{Strict enforcement}
+    & \rot{Deferred checking}
+    & \rot{Low-level code support}
+    & \multicolumn{1}{c}{\rot{Partial program support}}
+    & \rot{Well-bracketed control flow}
+    & \rot{Data integrity}
+    & \multicolumn{1}{c}{\rot{Data confidentiality}}
+    & \rot{Formal criteria}
+    \\
+    \midrule
+    Stack canaries  & \tabno   & \tabjein & \tabyes  & \tabyes  & \tabjein & \tabno   & \tabno   &          \\
+    Shadow stacks   & \tabno   & \tabjein &          & \tabyes  & \tabjein & \tabno   & \tabno   &          \\
+    Bounds checking & \tabyes  & \tabno   & \tabno   & \tabno   & \tabyes  &          &          &          \\
+    Split stacks    & \tabyes  & \tabno   & \tabno   & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
+    \midrule
+    Capabilities    &          & \tabno   &          & \tabyes  &          &          &          & \tabjein \\
+    Micro-policies  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  \\
+    \bottomrule
+  \end{tabular}
+  \caption{Comparison of Stack Protection Mechanisms}
+  \begin{tablenotes}
+  \item \tabyes indicates a protection mechanism possesses the feature
+  \item \tabno indicates the mechanism lacks the feature
+  \item \tabjein indicates partial or incomplete support
+  \item \rb{Complete and check columns, describe. Capitalize?}
+  \item \rb{Order? Split int/conf? Deferred/lazy at the end of features}
+  \item \rb{2+ lines for capabilities, before and after using our draft?}
+  \end{tablenotes}
+  \label{tab:relwork}
+\end{threeparttable}
+\end{table}
+
+\subsubsection{Strict Property Enforcement}
+%
+We are primarily interested in the \emph{strict} enforcement of a set of
+security properties of choice, i.e., no violations are allowed. Some \rb{most?}
+of the most commonly deployed countermeasures, like stack canaries and shadow
+stacks, offer only \emph{approximate} enforcement. As such, they will not
+satisfy any of our properties per se. We could use our framework to model them
+with some effort, but the interest of this exercise would be limited: testing
+would easily find the expected, unavoidable property violations. A more complex,
+\emph{probabilistic} version of the framework could offer adequate reasoning
+capabilities for these defenses.
+
+\subsubsection{Deferred and Lazy Properties}
+%
+We seek to support both \emph{eager} property enforcement (where potential
+violations are detected immediately) as well as more permissive \emph{lazy}
+properties (as motivated in \cref{XXX}).
+%
+Micro-policies are well suited to both types of properties. There is an
+interesting connection between lazy enforcement and the deferred checks of stack
+canaries and shadow stacks. Other techniques, like split stacks and bounds
+checking---and the closely related hardware capabilities---are fundamentally
+eager techniques.
+
+\subsubsection{Low-Level Component Support}
+%
+Our framework considers a \emph{single-, low-level view} of all program code and
+data; a key advantage of this approach is that it allows us to remain
+\emph{language-agnostic}.
+%
+Some software-based protection mechanisms are designed specifically with
+a \emph{high-level language} in mind, e.g. bounds checking and split stacks.
+They tend to offer stronger security guarantees than other software techniques.
+In exchange, their threat model becomes weaker, as they make additional
+assumptions about the admissible behavior of program components
+%
+(for example, bounds checking only applies to code that has been instrumented by
+a \emph{secure compiler} \ref{}).
+%
+In general, a multi-level extension of the framework offering adequate reasoning
+capabilities for this class of policies would need to model multiple source
+language and a notion of code provenance---a reflection of the call stack as the
+melting pot of language abstractions.
+%
+The common case of one source and one target language mediated by a compiler may
+be solved by a simpler lifting of our framework from the machine to the source
+level \rb{e.g., SoftBound could use a C language semantics and Tagged C for
+security modeling (lifted properties would likely hold under the expected
+assumptions)}.
+
+\subsubsection{Partial Program Protection}
+%
+A consequence of the use of high-level software techniques is the difficulty of
+keeping their shields up against a \emph{low-level attacker} that is not bound
+to play by their rules.
+%
+Insofar as their protections continue to work against untrusted code, like split
+stacks, they do so by relying on additional support, e.g. from the operating
+system.
+%
+\rb{(The same remarks about the language-agnostic framework apply here. Split
+  stacks are probably relatively easy to model without big changes, but would
+  still benefit from a higher level of abstraction.)}
+%
+\rb{(partial programs and split stacks? jein?)}
+
+\subsubsection{Enforced Properties}
+%
+\rb{(FIXME: Present with the hierarchy of properties? Basically all policies try
+  to protect {\wbcf}. Without it things are quite strange. Some of the more
+  restrictive software methods and the hardware methods can cover additional
+  properties.)}
+
+\subsubsection{Software vs. Hardware}
+%
+One first general observation about software-based techniques is that they
+are \emph{less modular} than their hardware counterparts, that is, the
+countermeasures are more tightly coupled with the program code they protect.
+%
+By contrast, although programs must effectively leverage the protection
+mechanisms offered by the hardware, the distinction between those protections
+and the program logic itself is much cleaner.
+%
+Our framework relies on a \emph{separation} of security-relevant operations and
+regular program code \rb{stress that any scalable, composable solution must do
+this!}.
+%
+While the various extensions suggested in this section could help abstract the
+complexities of software safeguards, hardware techniques remain the most elegant
+and general exemplars for the methods we seek to demonstrate.
+%
+\rb{(This is also related to our single-level model)}
+
+\subsubsection{Formal Models}
+%
+A second general observation about software-based techniques is that they tend
+to be \emph{less absolute}: they enforce weaker properties or under stronger
+assumptions, and their threat models are usually fairly informal.
+%
+%
+The closest previous attempt
+to produce a formal characterization of stack safety does so using a {\em fully abstract overlay semantics},
+a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
+But this approach
+builds the security properties into the semantics of the machine without externalizing them or their meaning.
+It is also not suitable for the {\em lazy} enforcement mechanism
+of Roessler and DeHon, in which stores are not checked, and illegal stores are
+only caught when their location is later read.
+%
+\rb{Need to discuss overlay semantics, but also note that it is very intensional
+  and very specific, as the authors acknowledge. I'd introduce this a bit more
+  gradually. Also, if phrased this way, it sounds like its main drawback is a
+  lack of generality, to which we should contrast a more general approach, not
+  just one that is specialized to a specific policy, i.e. lazy micro-policies
+  (I'd try to present these more gradually as well).}
+%
+\rb{Also, testing certainly (and motivate), but not only?}
+
+Overall, the generality and flexibility of micro-policies, arising from an
+elegant design that clearly separates the program from its security, are the
+most appropriate choice for an exploration of this new framework.
+%
+\rb{FIXME: Finish motivation. We focus on micro-policies, but\ldots Something
+  about existing policies, exercising the framework most generally,
+  representative experiments?}
+
+\rb{Begins to anticipate subsequent discussion. This is even something of an
+  understatement: some papers (esp. newer ones?), make concrete claims, though
+  they remain informal. Arguably, none have a truly formal characterization in
+  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
+  of the novelties of this work!
+\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
+  than previous work, without showing a fully general framework (which would be
+  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
+  start because of their flexibility, elegance, enabling a wide of interesting
+  experiments, etc.
+\rb{Basically, none reach the point of being an independent (and executable) specification}
+}
+}
+
+\subsection{The Properties and Their Hierarchy}
+
+The properties form a hierarchy, shown in [TODO: diagram].
+Within properties that are orthogonal
+some are likely to take priority in terms of their
+importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
+as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
+Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
+and callee-oriented properties even moreso.
+
+\subsection{Security Semantics and Property Structure}
 
@@
 
-\paragraph*{The Properties and their Hierarchy}
-
-The properties form a hierarchy, shown in [TODO: diagram].
-Within properties that are orthogonal
-some are likely to take priority in terms of their
-importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
-as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
-Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
-and callee-oriented properties even moreso.
-
-\paragraph*{Events}
+\subsection{Events}
 
@@
 
-\paragraph*{Threat Model}
+\subsection{Threat Model}
 
@@
 
+\subsection{Limitations}
+
+%Our concurrency model is fairly
+%simplistic, assuming a fixed number of threads each with its own dedicated processor.
+We model memory-safe stack objects, but not a heap. Regions outside of
+stacks can be used however the compiler likes, including as a heap, but no protection is
+built in and our properties assume that if a pointer to a stack object is stored there,
+it is permanently compromised.
+
+Our properties are termination insensitive: in order to support enforcement mechanisms
+that failstop rather than execute dangerous code, the properties must treat failstops
+as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
+This is consistent with our treatment of properties that specify the state after a return,
+in that a callee that diverges for any reason will never reach the return, and will fulfill
+the property vacuously.
+%
+\rb{Merge with threat model? Orthogonality of the heap, morally
+  similar interesting setups on the stack? Termination sensitivity and
+  capability machine semantics?}
+
 %%%
-
-\paragraph*{Permissiveness, Laziness, and Tagged Architectures}
-
-Security properties tend to be conservative, identifying violations that would be
-harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
-and of compatibility with existing codebases. In a crucial motivating example for this work,
-Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
-using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
-STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
-A set of software-defined rules, termed a micro-policy, determines the tags for operation
-results based on the tags of their inputs---or, if the input tags are in a dangerous
-combination, it may halt the machine (a ``failstop.'')
-
-Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
-is inefficient when allocating and clearing stack frames, because each word must be initialized
-with a tag corresponding to the (depth of the) current function activation. They propose
-{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
-when a function writes to memory. This means that writes cannot be prevented, necessarily
-violating existing models of stack safety. Instead the micro-policy enforces that memory be
-read by the same fuction that wrote it. The owner of the data will never see the changes,
-so it is secure in the sense that the attacker is unsuccessful at changing program behavior
-(except, perhaps, by causing a fault.) We elaborate that sense, and test whether the
-micro-policy fulfills it.
-
-This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
-machine states that make claims about the future observable behavior of execution from those states.
-That is not to say that our properties are specialized for lazy micro-policies. It also supports
-normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
-and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
-while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
-%
-\rb{Does this discussion (at least in part) belong in the general background
-  section, some even in key ideas? It still feels a little out of order to me,
-  and occasionally too detailed for the introduction?}
-
-\rb{Begins to anticipate subsequent discussion. This is even something of an
-  understatement: some papers (esp. newer ones?), make concrete claims, though
-  they remain informal. Arguably, none have a truly formal characterization in
-  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
-  of the novelties of this work!
-\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
-  than previous work, without showing a fully general framework (which would be
-  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
-  start because of their flexibility, elegance, enabling a wide of interesting
-  experiments, etc.
-\rb{Basically, none reach the point of being an independent (and executable) specification}
-}
-}
-%
-Previous attempts
-to produce a formal characterization do so using a {\it fully abstract overlay semantics},
-a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
-But this approach is not suitable for the {\em lazy} enforcement mechanism
-of Roessler and DeHon, in which stores are not checked, and illegal stores are
-only caught when their location is later read. We aim to provide a stack-safety
-property suitable for testing lazy enforcement.
-%
-\rb{Need to discuss overlay semantics, but also note that it is very intensional
-  and very specific, as the authors acknowledge. I'd introduce this a bit more
-  gradually. Also, if phrased this way, it sounds like its main drawback is a
-  lack of generality, to which we should contrast a more general approach, not
-  just one that is specialized to a specific policy, i.e. lazy micro-policies
-  (I'd try to present these more gradually as well).}
-
-\newcommand*{\rot}[1]{\hbox to1em{\hss\rotatebox[origin=br]{-60}{#1}}}
-\newcommand*{\tabyes}{\CIRCLE\xspace}
-\newcommand*{\tabno}{\Circle\xspace}
-\newcommand*{\tabjein}{\RIGHTcircle\xspace}
-
-\begin{table}
-\begin{threeparttable}
-  \begin{tabular}{l | c c c c | c c c | c}
-    \toprule
-    \multicolumn{1}{l |}{Mechanism}
-    & \multicolumn{4}{l |}{Features}
-    & \multicolumn{3}{l |}{Properties}
-    & %% Formal criteria
-    \\
-    \midrule
-    %% Technique
-    \multicolumn{1}{c}{}
-    & \rot{Strict enforcement}
-    & \rot{Deferred checking}
-    & \rot{Low-level code support}
-    & \multicolumn{1}{c}{\rot{Partial program support}}
-    & \rot{Well-bracketed control flow}
-    & \rot{Data integrity}
-    & \multicolumn{1}{c}{\rot{Data confidentiality}}
-    & \rot{Formal criteria}
-    \\
-    \midrule
-    Stack canaries  & \tabno   & \tabjein & \tabyes  & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
-    Shadow stacks   & \tabno   & \tabjein &          & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
-    Bounds checking & \tabyes  & \tabno   & \tabno   & \tabno   & \tabyes  &          &          &          \\
-    Split stacks    & \tabyes  & \tabno   & \tabno   & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
-    \midrule
-    Capabilities    &          & \tabno   &          & \tabyes  &          &          &          & \tabjein \\
-    Micro-policies  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  \\
-    \bottomrule
-  \end{tabular}
-  \caption{Comparison of Stack Protection Mechanisms}
-  \begin{tablenotes}
-  \item \tabyes indicates a protection mechanism possesses the feature
-  \item \tabno indicates the mechanism lacks the feature
-  \item \tabjein indicates partial or incomplete support
-  \item \rb{Complete and check columns, describe. Capitalize?}
-  \item \rb{Order? Split int/conf? Deferred/lazy at the end of features}
-  \item \rb{2+ lines for capabilities, before and after using our draft?}
-  \end{tablenotes}
-  \label{tab:relwork}
-\end{threeparttable}
-\end{table}
 
@@
 %
-We have made explicit the key point that there is \emph{not} a single stack
-abstraction. This paves the way for a \emph{general} idea of stack protection
-that is widely applicable with high-level concepts that remain unchanged (and I
-think it's particularly important to at least introduce these more immutable
-parts early on). The fact that this general ``framework'' requires low-level or
-specific specializations is a predictable consequence of the protean nature of
-the call stack.
-%
-\\
-%
-Let us return to a selection of representative policies in our menagerie and
-consider how they fit into the framework. At the highest level, we can classify
-all those techniques into two large groups, software-based and hardware-based.
-Traditionally, most practical protections are implemented in software, so we
-start by considering this first group.
-%
-\\
-%
-We can make two general observations about these software based techniques.
-Firstly, (as we will see: discussion? technical details?) these techniques are
-\emph{less modular} than their hardware counterparts. That is, the
-countermeasures are tightly coupled with the program code they protect, and
-consequently more effort is required to disentangle the program from the
-protections, and to understand the impact of the latter on the former. This is
-true of any systematic attempt to study the security of stack protection
-policies, including their adaptation to our framework. (Owing to this added
-complexity, software policies will not be the cleanest or most elegant exemplars
-of use of the framework, its salient features and practical utility; this partly
-motivates the focus on hardware policies.)
-%
-\\
-%
-(The above remark holds \emph{at least in a ``single-level model''} that
-considers assembly language/machine code programs. In general, (low-level)
-protections are introduced by a compiler or similar tool. One could consider a
-multi-level extension of the framework pairing models of high-level programs and
-the security transformations performed by a (stack-secure) compiler.)
-%
-\\
-%
-Secondly, the guarantees generally offered by software-based policies are
-\emph{less absolute}, being defined more or less informally and usually
-providing few definite guarantees of absence of \emph{general} classes of
-errors. (The following discussion elaborates on this. It is conceivable that an
-extended and more sophisticated framework could add \emph{probabilistic
-reasoning} capabilities to capture and reason about the effects of less strict
-protection mechanisms. This would be another orthogonal development to what is
-presented here. In this paper, we are interested in assessing the complete
-\emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
-or defects. In passing, this would extend the security properties to e.g. larger
-system verification efforts.)
-%
-\\
-%
-One by one, consider the following specimens of software protection mechanisms
-(quick explanations inlined?):
-%
-\\
-%
-\begin{itemize}
-%
-\item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
-  difficulty of some attacks on the stack, usually (always?) buffer overflows.
-  They can never provide any absolute security guarantees and therefore do not
-  satisfy any of our properties (some of the properties, e.g. confidentiality,
-  are completely out of scope.) Although tedious, we could we could model this
-  mechanism in our framework with the appropriate adaptations --- and
-  unsurprisingly, testing would find property violations, which we already know
-  are there: the results of this effort would be unsurprising and not as useful
-  or instructive as other experiments.
-  \rb{(+SNA) They are kind of lazy, aren't they?}
-%
-\item \emph{Bounds checking}. This is fundamentally a high-level technique, even
-  more obviously so than stack canaries (which could still detect violations
-  caused by external code). Bounds checking only applies to code that has been
-  instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
-  impose any constraints on adversarial code. The attacker model is therefore
-  rather weak. (Certain variants, like SoftBound, are good targets for lifting
-  our framework to the source level by e.g. using a C-language semantics and
-  using Tagged C for security modeling; the lifted versions of our properties
-  would likely hold under the expected assumptions. HardBound is probably an
-  interesting transitional case because it uses capability-like hardware
-  primitives for enforcement; it could be close to ongoing and future work to
-  verify stack safety on capability machines, e.g. based on Cerise -- discuss
-  together?)
-%
-\item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
-  ``probabilistic techniques'' in that they ``restrict the flexibility available
-  in creating gadget chains'' (Shanbhogue et al.). This technique is solely
-  concerned with control-flow integrity, and so closely related to our
-  well-bracketed control flow. Nevertheless, and again, it is only an
-  approximate technique, and our testing framework would be able to find the
-  expected attacks against our security property.
-%
-\item \emph{Split stacks}. The security claims of split stacks are stronger than
-  those of most software countermeasures, i.e., they are meant to guarantee
-  their titular code-pointer integrity. They explicitly offer no protections on
-  stack data (i.e., integrity and confidentiality), and so restrict themselves
-  to well-bracketed control flow. (This is probably closer to our framework than
-  most other software techniques, and it seems like it would be easier to model
-  as well. However, this is again work based on high-level languages and would
-  benefit from a higher level of abstraction.)
-%
-\item \ldots
-%
-\end{itemize}
-%
-Opposed to the above group of software techniques we have hardware-based
-protection policies (more in background/technical preliminaries?). Although
-programs must effectively leverage the protection mechanisms offered by the
-hardware, the distinction between those protections and the program logic itself
-is much cleaner: it is fairly simple to separate the use of the security
-features in the hardware (and their effects in program code) from regular
-program code. They also tend to aspire to offer complete protection against
-entire classes of attacks. Both aspects make this family of techniques a better
-fit for illustrating and showcasing our techniques.
-%
-\\
-%
+%% We have made explicit the key point that there is \emph{not} a single stack
+%% abstraction. This paves the way for a \emph{general} idea of stack protection
+%% that is widely applicable with high-level concepts that remain unchanged (and I
+%% think it's particularly important to at least introduce these more immutable
+%% parts early on). The fact that this general ``framework'' requires low-level or
+%% specific specializations is a predictable consequence of the protean nature of
+%% the call stack.
+%% %
+%% \\
+%% %
+%% Let us return to a selection of representative policies in our menagerie and
+%% consider how they fit into the framework. At the highest level, we can classify
+%% all those techniques into two large groups, software-based and hardware-based.
+%% Traditionally, most practical protections are implemented in software, so we
+%% start by considering this first group.
+%% %
+%% \\
+%% %
+%% We can make two general observations about these software based techniques.
+%% Firstly, (as we will see: discussion? technical details?) these techniques are
+%% \emph{less modular} than their hardware counterparts. That is, the
+%% countermeasures are tightly coupled with the program code they protect, and
+%% consequently more effort is required to disentangle the program from the
+%% protections, and to understand the impact of the latter on the former. This is
+%% true of any systematic attempt to study the security of stack protection
+%% policies, including their adaptation to our framework. (Owing to this added
+%% complexity, software policies will not be the cleanest or most elegant exemplars
+%% of use of the framework, its salient features and practical utility; this partly
+%% motivates the focus on hardware policies.)
+%% %
+%% \\
+%% %
+%% (The above remark holds \emph{at least in a ``single-level model''} that
+%% considers assembly language/machine code programs. In general, (low-level)
+%% protections are introduced by a compiler or similar tool. One could consider a
+%% multi-level extension of the framework pairing models of high-level programs and
+%% the security transformations performed by a (stack-secure) compiler.)
+%% %
+%% \\
+%% %
+%% Secondly, the guarantees generally offered by software-based policies are
+%% \emph{less absolute}, being defined more or less informally and usually
+%% providing few definite guarantees of absence of \emph{general} classes of
+%% errors. (The following discussion elaborates on this. It is conceivable that an
+%% extended and more sophisticated framework could add \emph{probabilistic
+%% reasoning} capabilities to capture and reason about the effects of less strict
+%% protection mechanisms. This would be another orthogonal development to what is
+%% presented here. In this paper, we are interested in assessing the complete
+%% \emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
+%% or defects. In passing, this would extend the security properties to e.g. larger
+%% system verification efforts.)
+%% %
+%% \\
+%% %
+%% One by one, consider the following specimens of software protection mechanisms
+%% (quick explanations inlined?):
+%% %
+%% \\
+%% %
+%% \begin{itemize}
+%% %
+%% \item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
+%%   difficulty of some attacks on the stack, usually (always?) buffer overflows.
+%%   They can never provide any absolute security guarantees and therefore do not
+%%   satisfy any of our properties (some of the properties, e.g. confidentiality,
+%%   are completely out of scope.) Although tedious, we could we could model this
+%%   mechanism in our framework with the appropriate adaptations --- and
+%%   unsurprisingly, testing would find property violations, which we already know
+%%   are there: the results of this effort would be unsurprising and not as useful
+%%   or instructive as other experiments.
+%%   \rb{(+SNA) They are kind of lazy, aren't they?}
+%% %
+%% \item \emph{Bounds checking}. This is fundamentally a high-level technique, even
+%%   more obviously so than stack canaries (which could still detect violations
+%%   caused by external code). Bounds checking only applies to code that has been
+%%   instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
+%%   impose any constraints on adversarial code. The attacker model is therefore
+%%   rather weak. (Certain variants, like SoftBound, are good targets for lifting
+%%   our framework to the source level by e.g. using a C-language semantics and
+%%   using Tagged C for security modeling; the lifted versions of our properties
+%%   would likely hold under the expected assumptions. HardBound is probably an
+%%   interesting transitional case because it uses capability-like hardware
+%%   primitives for enforcement; it could be close to ongoing and future work to
+%%   verify stack safety on capability machines, e.g. based on Cerise -- discuss
+%%   together?)
+%% %
+%% \item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
+%%   ``probabilistic techniques'' in that they ``restrict the flexibility available
+%%   in creating gadget chains'' (Shanbhogue et al.). This technique is solely
+%%   concerned with control-flow integrity, and so closely related to our
+%%   well-bracketed control flow. Nevertheless, and again, it is only an
+%%   approximate technique, and our testing framework would be able to find the
+%%   expected attacks against our security property.
+%% %
+%% \item \emph{Split stacks}. The security claims of split stacks are stronger than
+%%   those of most software countermeasures, i.e., they are meant to guarantee
+%%   their titular code-pointer integrity. They explicitly offer no protections on
+%%   stack data (i.e., integrity and confidentiality), and so restrict themselves
+%%   to well-bracketed control flow. (This is probably closer to our framework than
+%%   most other software techniques, and it seems like it would be easier to model
+%%   as well. However, this is again work based on high-level languages and would
+%%   benefit from a higher level of abstraction.)
+%% %
+%% \item \ldots
+%% %
+%% \end{itemize}
+%% %
+%% Opposed to the above group of software techniques we have hardware-based
+%% protection policies (more in background/technical preliminaries?). Although
+%% programs must effectively leverage the protection mechanisms offered by the
+%% hardware, the distinction between those protections and the program logic itself
+%% is much cleaner: it is fairly simple to separate the use of the security
+%% features in the hardware (and their effects in program code) from regular
+%% program code. They also tend to aspire to offer complete protection against
+%% entire classes of attacks. Both aspects make this family of techniques a better
+%% fit for illustrating and showcasing our techniques.
+%% %
+%% \\
+%% %
 CHERI and lightweight formal methods?
@@
 }
-\rb{(+SNA) Any interesting tables comparing specimens in the menagerie?}
-\rb{(+SNA) A general multi-level extension of the framework would need to model
-  multiple source languages. Among other things, the stack is the meeting place
-  of all sorts of different language abstractions! One of the advantages of the
-  single-level model is that it lets us be agnostic about the language. (On the
-  other hand, it is likely that a single source and target mediated by the
-  compiler suffice to model many existing techniques, coming e.g. from systems
-  programming and C-like languages.}
-
-\paragraph*{Limitations}
-
-%Our concurrency model is fairly
-%simplistic, assuming a fixed number of threads each with its own dedicated processor.
-We model memory-safe stack objects, but not a heap. Regions outside of
-stacks can be used however the compiler likes, including as a heap, but no protection is
-built in and our properties assume that if a pointer to a stack object is stored there,
-it is permanently compromised.
-
-Our properties are termination insensitive: in order to support enforcement mechanisms
-that failstop rather than execute dangerous code, the properties must treat failstops
-as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
-This is consistent with our treatment of properties that specify the state after a return,
-in that a callee that diverges for any reason will never reach the return, and will fulfill
-the property vacuously.
-%
-\rb{Limitations in the introduction or later? Orthogonality of the heap, morally
-  similar interesting setups on the stack? Termination sensitivity and
-  capability machine semantics?}
-
-%%%
+%% \rb{(+SNA) Any interesting tables comparing specimens in the menagerie?}
+%% \rb{(+SNA) A general multi-level extension of the framework would need to model
+%%   multiple source languages. Among other things, the stack is the meeting place
+%%   of all sorts of different language abstractions! One of the advantages of the
+%%   single-level model is that it lets us be agnostic about the language. (On the
+%%   other hand, it is likely that a single source and target mediated by the
+%%   compiler suffice to model many existing techniques, coming e.g. from systems
+%%   programming and C-like languages.}
 
@@
   exceptions (still in intro) seems disconnected from the rest.}
+
+\rb{Deferred/lazy connection of canaries and shadow stacks -- arguably these helps them support things other software policies need more help with?}
+
+\rb{Capabilities and HardBound?}

### 2023-01-29 submission5/ideas-v2.tex
Nits

@@
 %
+\rb{(Mention complexity/relation between eager/lazy properties somewhere? Later?)}
+%
 Micro-policies are well suited to both types of properties. There is an
@@
 capabilities for this class of policies would need to model multiple source
-language and a notion of code provenance---a reflection of the call stack as the
+languages and a notion of code provenance---a reflection of the call stack as the
 melting pot of language abstractions.
@@
 By contrast, although programs must effectively leverage the protection
-mechanisms offered by the hardware, the distinction between those protections
+mechanisms offered by a hardware-based mechanism, the distinction between those protections
 and the program logic itself is much cleaner.
@@
 regular program code \rb{stress that any scalable, composable solution must do
-this!}.
+this!, add fwd ref?}.
 %
@@
 A second general observation about software-based techniques is that they tend
-to be \emph{less absolute}: they enforce weaker properties or under stronger
+to be \emph{less absolute}: they enforce weaker properties or rely on stronger
 assumptions, and their threat models are usually fairly informal.
 %
-%
-The closest previous attempt
+The closest attempt
 to produce a formal characterization of stack safety does so using a {\em fully abstract overlay semantics},
@@
 only caught when their location is later read.
+%
+\rb{Move the bit about laziness to the corresponding point above? Also emphasize
+  (and contributions?) that we go further by having an extensional, executable
+  specification, which we can test or verify.}
 %
@@
   about existing policies, exercising the framework most generally,
-  representative experiments?}
+  representative experiments? This kind of repeats the opening of the section.}
 
@@
 
-A {\it security semantics} extends a machine
+A {\em security semantics} extends a machine
 with additional context about the identity of current and pending
@@
 it does not affect the real machine. The security context
-evolves dynamically through the execution of {\it security-relevant operations},
+evolves dynamically through the execution of {\em security-relevant operations},
 which include events like calls, returns, and frame manipulations.
@@
 %
-Questionably that they are strictly more realistic in the sense that it is very
-costly or requires special sophisticated/experimental types of capability not
-available/perhaps not very feasible to implement.
+\\
 %
-Problems of adapting this work to capability machines: security semantics tightly
+One may wonder whether capability methods are strictly more realistic in the sense they are very costly
+costly or require sophisticated/experimental types of capability not
+available/perhaps not very feasible to implement---it's an open question.
+%
+\\
+%
+Challenges of adapting this work to capability machines: security semantics tightly
 coupled to regular machine semantics and to a complex calling convention, with

### 2023-01-29 submission5/intro-v2.tex
Nits

@@
 well as information about its \emph{caller} and \emph{callee} functions. The
-runtime must keep track of the local contexts of all active function activations
+runtime must keep track of the local contexts of all function activations
 at any given time; the \emph{fundamental data structure} charged with realizing
@@
 
-The correct execution on a program---and its security---is contingent on all of
+The correct execution on a program---and its security!---is contingent on all of
 its functions properly implementing this abstraction by means of a careful use
@@
 menagerie and begin to consider the difficulties involved in specifying and
-verifying the protection they offer.
+verifying the protections they offer.
 %
 For an assessment of these methods in the context of our stack security
-framework, see \cref{XXX} below.
+framework, see \cref{XXX}.
 
@@
 splitting the program memory, including the stack, into a regular region and a
-safe region; those objects that require protection are located in the latter and
-all accesses protected by static or dynamic checks.
+safe region; objects that require protection are placed in the latter and
+all accesses to them protected by static or dynamic checks.
 %
 Like shadow stacks, bounds checking is a high-level technique designed to
-protect code written in a high-level language, but, unlike them, it is meant to
+protect code written in a high-level language, but it is designed to
 offer protection under a stronger threat model that also considers the presence
-of untrusted code. Similar to most software-based methods, it purports to
+of untrusted code. As with most other software-based methods, it purports to
 protect exclusively the control-relevant parts of the stack, making no claims
@@
 
-In addition to many the software-based techniques developed to provide greater
-degrees of protection to the stack, often at a significant cost in performance,
+In addition to the many software-based techniques developed to provide greater
+degrees of protection to the stack, often incurring a significant performance penalty,
 some proposals have turned to specialized hardware features to offset those
-penalties and attempt to offer greater levels of protection.
+costs and attempt to offer greater levels of protection.
 %
@@
 %
-Capability machines can represent and manipulate their homonymous
-capabilities---unforgeable tokens of authority over specific regions of
-memory---using a combination of dedicated out-of-band memory and a collection of
+Capability machines can represent and manipulate
+unforgeable tokens of authority over specific regions of
+memory using a combination of dedicated out-of-band memory and
 specialized ISA instructions.
@@
 processor with a programmable hardware monitor that uses custom rules on
-metadata tags to implement security policies of various types.
+metadata tags to implement many different kinds of security policies.
 %
-This flexibility can be harnessed to implement a wide array of stack protection policies with relative ease, covering and going beyond the capabilities \rb{word?} of the various software-based techniques \cite{} \rb{as demonstrated by}.
+This flexibility can be harnessed to implement a wide array of stack protection policies with relative ease, covering and going beyond the capabilities \rb{word choice?} of the various software-based techniques \cite{} \rb{as demonstrated by}.
 %
-They comprise very general forms of \emph{lazy enforcement}, which can tolerate
+These policies comprise very general forms of \emph{lazy enforcement}, which can tolerate
 potential but harmless violations \rb{the system programmer's motto: ``trust me,
@@
 \emph{unique specification} of the meaning of \emph{stack safety} for a
-particular protection mechanism---mostly an informal specification if one is
+particular protection mechanism---usually an informal specification if one is
 produced, and almost always left implicit to some degree.
@@
 specification} (or set of specifications) for such a fundamental problem.
-Instead of reverse engineering an specific enforcement mechanism obtain a
+Instead of reverse-engineering a specific enforcement mechanism to obtain a
 (necessarily approximate) specification, we turn the problem around on its head
@@
 %
-\rb{Return to this after extended discussion?}
+\rb{Return to this after extended discussion? Forward ref?}
 %

### 2023-01-30 submission5/intro-v2.tex
Comments on the new intro section

@@
 \rb{This should be well known, so no need for illustration?}
+
+\sna{I agree, but: a discussion of the low-level abstraction should discuss
+  the ABI and the fact that things like private vs. shared data are fundamentally
+  not specified.}
 
@@
 To mitigate this risk, \rb{we claim that} mechanisms should aim to enforce \emph{formal
-properties} of correct \rb{safe?} behavior that can be proven or tested. Such properties become the
+properties} of safe behavior that can be proven or tested. Such properties become the
 \emph{specification} against which enforcement can be validated; even enforcement mechanisms that do
@@
 
-\subsubsection{Software Techniques}
-
 At the highest level, we can classify techniques into two large groups depending
@@
 from this first group.
+
+\subsubsection{Software Techniques}
 
@@
 
+\sna{Would HardBound warrant a section in the hardware techniques section,
+  for hardware-accelerated bounds checking (that is not capability-based)?}
+
 \paragraph{Split Stacks}
@@
   (CHERI-adjacent) work exclusively?)}
+
+\sna{If we have good examples of other capability systems, we should absolutely
+  talk about them. No need to privilege CHERI.}
 %
@@
 %
-Previous work has demonstrated the use of these general principles \rb{in
-  simplified models} to implement \emph{secure calling conventions} for the call
-stack \cite{}. Those conventions can be seen as combining the advantages and
-protections of various families of software-based methods.
+Previous work has uses these principles in simplified models to implement
+\emph{secure calling conventions} for the call
+stack \cite{}. Those conventions can be seen as combining the
+protections of bounds-checking with software-based methods of control-flow protection.
 %
@@
   from a small number of fundamental concepts)}
+
+\sna{Still very worried about whether this is plausible to claim given our focus
+  on one enforcement mechanism. There is probably a way to thread the needle, though.
+  Something like: we begin with a mechanism that we know is not supported by any
+  other model, and seek to build a model that is not specific to it by appealing to
+  first principles.}
 

### 2023-01-30 submission5/core.tex
Input instead of include
Comments on the new intro section
Detailed readthrough and notes on intro

@@
 
-\include{ideas-v1}
+\input{ideas-v2}
 
@@
 
-\subsection{Public Allocations}
-
-[TODO: this section could use some trimming. I think we have room for this
-discussion, it's just more verbose than everything else and therefore feels weird.]
-
-The full machine supports public stack allocations that can
-be accessed by anyone until they are deallocated on their owner's return,
-represented by \(\mathbf{alloc} ~ \mathbf{t} \dots\).
-For example, consider the following code:
-
-{\tt
-  void g() \{
-
-  ~ ~ int x[3];
-
-  ~ ~ int y;
-
-  ~ ~ h(\&x);
-
-  \}
-}
-
-Here we have allocated several arrays and variables; {\tt y}
-will behave as normal with \(\mathbf{alloc} ~ \mathbf{f} \dots\), but
-{\tt x} has its address taken and passed to another function. That
-function might do anything at all to {\tt x}, and might store its pointer
-elsewhere (in the heap, for instance) so that the scope of legitimate accesses
-to {\tt x} are uncertain. We default to assuming that any address-taken local
-should be public. Meanwhile {\tt y} is still private.
-
-In the likely event that the space for the return address, {\tt x}, and {\tt y}
-are all allocated in a single step, that step makes multiple allocation operations:
-\[\bar{\psi} = [\mathbf{alloc} ~ \mathbf{f} ~ (-20,4);
- \mathbf{alloc} ~ \mathbf{t} ~ (-16,12) ; \mathbf{alloc} ~ \mathbf{f} ~ (-4,4)]\]
-
+\subsection{Sharing Stack Memory}
+
+In our examples so far, we have presented a vision of stack safety in which
+the interface between caller and callee is in the registers that pass
+arguments and return values. This is frequently not the case in a realistic
+setting. Arguments may be passed on the stack due to being spilled, as an implementation
+of variadic arguments, or because they are reference objects that inherently have
+pass-by-reference semantics. The caller may also take the address of a local object and
+pass that as a pointer to its callee.
+
+We refine our call operation to make use of the information that we have about
+which memory contain arguments, \(\overline{sa}\). \(\overline{sa}\) is a set of
+triples of a register, an offset from the value of that register, and a size.
+We first define the helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\),
+then extend the call operation to keep all objects in \(\mathit{passed}\) marked
+as \(\object\) and seal everything else (\cref{sfig:stkargs}).
+
+In this mechanism, we can describe an argument being passed by their relative address
+with the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
+In the case of pass-by-reference semantics, if the compiler designates register
+\(\reg\) as holding the reference, then the operations is
+\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
+
+In both cases, the immediate callee retains access to the passed object, but future
+callees will not---unless the object is passed-by-reference further down the stack.
+On the other hand, in the case of a caller explicitly taking the address of an
+object and passing it as a pointer, we have no way obvious way to limit the scope
+of such sharing.
+
+Our primary machine treats this scenario as the caller allocating a ``public''
+object that can be accessed by anyone until it's deallocated. We extend the
+\(\mathbf{alloc}\) operation with a boolean flag, where {\bf t} indicates
+that the allocation is public, and {\bf f} is private.
+In the likely event that the space for multiple objects is allocated in a single step,
+that step can make multiple allocation operations, each labeled appropriately.
 Allocating public objects affects the call stack very similarly to private ones,
 but instead of labeling them \(\object\) they become \(\public\), so they are
-not sealed at a call. This makes these
-regions of memory able to transmit information between caller and callee without
-violating the stack safety properties. Note that there is no way to seal a location
-once it has become \(\public\); only when {\tt x} is deallocated will its region
-be available to allocate privately. We define a helper function, \(\mathit{range}\),
-to map an offset from a register and a size into a set of addresses, and update
-the operation's security semantics as shown in \cref{sfig:publicalloc}.
+never sealed at a call (\cref{sfig:publicalloc}).
 
@@
 
-    \caption{Argument Passing on }
+    \caption{Calls with Argument Passing on the Stack}
     \label{sfig:stkargs}
@@
 \end{figure*}
-
-\subsection{Stack Arguments}
-
-So far, we have not dealt with arguments spilled onto the stack, variadic arguments,
-or arguments that are passed by reference without their addresses being taken
-explicitly (i.e., structs and arrays passed directly.) Consider the simplest case,
-where a function has more than 8 arguments and therefore must (in RISC-V) spill one
-to the stack.
-
-{\tt
-  void main() \{
-
-  ~ f(1,2,3,4,5,6,7,8,9);
-
-  ~ g(42);
-
-  \}
-}
-
-\vspace{\abovedisplayskip}
-
-\begin{tabular}{r l | l}
-  \labeledrow{0:}{addi sp,sp,-12}{\(\mathbf{alloc} ~ (-12,12)\)}
-  \labeledrow{4:}{sd ra,4(sp)}{}
-  \labeledrow{8:}{li a5,9}{}
-  \labeledrow{12:}{sd a5,0(sp)}{}
-  \labeledrow{16:}{li a7,8}{}
-  \dots \\
-  \labeledrow{48:}{jal 100,ra}{\(\mathbf{call} ~ \{\mathtt{a0-a7}\} ~ \{(\SP,0,4)\}\)}
-  \labeledrow{52:}{li a0,42}{}
-  \labeledrow{56:}{jal 200,ra}{\(\mathbf{call} ~ \{\mathtt{a0}\} ~ \emplist\)}
-  \dots \\
-\end{tabular}
-
-Under a typical calling convention, the callee expects the ninth argument to
-be passed to it at bottom of the caller's frame, so the caller allocates extra
-space for it. Then the call to {\tt f} identifies that we expect a range of four
-bytes from the stack pointer to be passed as an object instead of sealed.
-
-So we must refine our call operation to make use of the information that we have about
-which memory contain arguments, \(\overline{sa}\). \(\overline{sa}\) is a set of
-triples of a register, an offset from the value of that register, and a size.
-We first define the helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\),
-then extend the call operation to keep all objects in \(\mathit{passed}\) marked
-as \(\object\) and seal everything else (\cref{sfig:stkargs}).
-
-The remaining definitions are unchanged. Now a stack-spilled argument can be allocated
-and remain accessible in the callee, while its memory is sealed for calls that do not
-use that memory as an argument.
-
-This same structure works for pass-by-reference arguments. In this case {\tt main}
-passes an array {\tt x} directly, which is compiled as a pass-by-reference.
-
-{\tt
-  int main() \{
-
-  ~ int x[1];
-
-  ~ f(x);
-
-  ~ g();
-
-  ~ return x[0];
-
-  \}
-}
-
-Unlike the address-taken array, we can clearly delineate the scope in which {\tt x} may be
-accessed without violating stack safety: {\tt f} should be allowed to access it, but further
-calls from {\tt f} should not (unless {\tt f} passes it to them) and {\tt g} certainly
-should not. Say the compiler opts to pass the pointer to {\tt x} in register {\tt a0}.
-The {\tt f} is specifically being given access to the four bytes starting at that address.
-The call operation is \(\mathbf{call} ~ \emplist ~ \{(\mathtt{a0},0,4)\}\).
 
@@
 
-\section{Related Work}
-\label{sec:relwork}
-
-\paragraphx{Formal Stack Safety on Capability Machines}
-%
-To our knowledge, the only line of work to date that has attempted
-a positive and formal characterization of stack safety
-begins with Skorstengaard et al. \cite{SkorstengaardSTKJFP} and is improved
-by Georges et al. \cite{Georges22:TempsDesCerises}. The former's
-StkTokens system enforces stack safety on hardware equipped with
-memory capabilities, similar to those used in the Cheri system~\cite{Woodruff+14,Chisnall+15}, but
-with additional support for linear capabilities, while the latter offer a calling
-convention that can replace the expensive and likely-impractical linear capabilities
-with more efficient ``directed'' capabilities.
-
-Both works define stack safety as the conjunction of local state encapsulation (LSE) and
-well-bracketed control flow (WBCF).
-Informally, LSE means that the current activation can only access its own frame,
-and WBCF means that the topmost frame can only return to
-the immediately adjacent frame below. Formally, these properties are defined
-with reference to a capability machine with a built-in call stack and call and return
-instructions; this machine is proven fully abstract with respect to a more
-concrete capability machine, an approach termed ``fully abstract overlay semantics.''
-
-[TODO: more detailed comparison with Aina's work primarily.]
-
-\paragraphx{Heap Safety as a Security Property}
-%
-Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18} show that
-heap safety may be framed as a security property using
-the language of noninterference. Their paper was a significant inspiration
-for the present work, and we believe that their safety conditions should
-combine smoothly into ours.
-
 \section{Future Work}

### 2023-01-30 submission5/intro-v2.tex
Detailed readthrough and notes on intro

@@
 at any given time; the \emph{fundamental data structure} charged with realizing
-this abstraction is the \emph{(call) stack}.
-%
-\rb{This should be well known, so no need for illustration?}
+this abstraction is the \emph{(call) stack}. But at the level of the ABI
+\sna{Do we need to unpack the acronym ABI?}, the specification of a well-behaved
+stack is minimal. It is not concerned with such questions as whether a caller's
+frame should be readable or writable to its callee, leaving those concerns
+to be determined by a high-level language built atop the stack.
 
-\sna{I agree, but: a discussion of the low-level abstraction should discuss
-  the ABI and the fact that things like private vs. shared data are fundamentally
-  not specified.}
-
-The correct execution on a program---and its security!---is contingent on all of
-its functions properly implementing this abstraction by means of a careful use
+The correct execution---and the security!---of a program is contingent on all of
+its functions properly implementing the intended abstraction by means of a careful use
 of the call stack. Be it accidental or intentional, incorrect management of the
-stack is at the root of a myriad of faults whose effects range from (rarely)
+stack is at the root of many faults whose effects range from (rarely)
 innocuous to (usually) catastrophic.
@@
 %
-\rb{Begins to prefigure the idea of multiple stack implementations.}
 
@@
 %
-Even a cursory glance at the most recent release of the authoritative CWE Top 25
+The most recent release of the authoritative CWE Top 25
 Most Dangerous Software Weaknesses
@@
 shows general classes of vulnerabilities closely related to the stack \rb{(though not exclusively)}
-consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read), \#7 (CWE-416, Use
-After Free) or \#11 (CWE-476, NULL Pointer Dereference).
+consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read), \#7 (CWE-416, Use After Free) or \#11 (CWE-476, NULL Pointer Dereference).
 %
@@
   memory safety statistics, etc.}.
+\sna{Trying to think how to make the case that these CWEs are about the stack, since they are
+  primarily described as memory attacks.}
 %
@@
 \rb{Broad enough? Do we want to add more techniques?}
+\sna{I think this is plenty unless there's a really major thing we're forgetting.}
 %
 Yet despite this wealth of potential solutions, the stack remains a central
-security concern---why?
+security concern.
+
+\paragraph{Formal Properties}
 
@@
 %
-To mitigate this risk, \rb{we claim that} mechanisms should aim to enforce \emph{formal
+To mitigate this risk, we claim that mechanisms should aim to enforce \emph{formal
 properties} of safe behavior that can be proven or tested. Such properties become the
@@
 %
-This is the problem we begin to address in this paper \rb{or a few more such
-  connecting words}.
+Our work is focused on the creation of such properties for stack safety.
 
-\subsection{A Menagerie of Stack Protection Mechanisms}
-
-\rb{A.K.A. related work (part 1?). I wonder if this goes before or after a first
-  look at the idea of the properties.}
+\subsection{Related Work}
 
@@
 costs and attempt to offer greater levels of protection.
+
+We focus in particular on two families of hardware mechanisms: capability machines
+and tag-based reference monitors. Both types of system move beyond simply
+offering hardware acceleration for software-based techniques, offering hardware
+primitives for both memory and control-flow enforcement that can offer a degree
+of protection even with minimal cooperation from the code.
 %
@@
 %
-\rb{HardBound is also related to this}
 
@@
 
-As we begin to intuit, the security of the call stack and its contents can be
-viewed through many different lenses. Each of these points of view results in a
-\emph{unique specification} of the meaning of \emph{stack safety} for a
-particular protection mechanism---usually an informal specification if one is
-produced, and almost always left implicit to some degree.
-%
+We begin to see that the security of the call stack and its contents is not
+clearly defined. Different protection mechanisms aim to protect the stack
+in fundamentally different ways, each offering a \emph{unique}, \emph{implicit}
+specification of the meaning of ``stack safety.'' These specifications
+may be more or less compatible with the needs of a given high-level language.
 Moreover, different protections operate under different threat models, further
@@
 effectively offer their clients---the programs themselves.
+It would be far preferable to have a \emph{standard}, \emph{explicit} specification.
+Only a handful of works have offered any explicit specification.
+
+Stack safety is challenging to specify in full generality, because the stack is
+fundamentally an architecture-specific construct.
+(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
+which is broadly consistent across architectures and language semantics, and therefore
+amenable to a very general approach.)
+The desired behavior of the stack varies based on ISA, calling convention,
+and the features of whichever programming language originated the code.
 %
-On the face of it, it seems perplexing that there is not a \emph{standard
-specification} (or set of specifications) for such a fundamental problem.
-Instead of reverse-engineering a specific enforcement mechanism to obtain a
-(necessarily approximate) specification, we turn the problem around on its head
-and seek more general solutions from first principles.
-%
-\rb{(Problems/notes: ``backward specs'' not the most useful to their clients,
-  not designed with their security needs in mind---goes in the direction of
-  program logics---; to reason about security in complex systems an explicit and
-  clear spec is preferable; many of the informal specs are at least
-  superficially similar and overlapping; hopefully all of them can be understood
-  from a small number of fundamental concepts)}
+For instance, the current state-of-the-art stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable.
 
-\sna{Still very worried about whether this is plausible to claim given our focus
-  on one enforcement mechanism. There is probably a way to thread the needle, though.
-  Something like: we begin with a mechanism that we know is not supported by any
-  other model, and seek to build a model that is not specific to it by appealing to
-  first principles.}
-
-We propose a novel characterization of stack safety using the formal tools of language-based
+In this work, we propose an explicit specification, built from first principles
+in the hopes that it will be widely applicable, and demonstrate that it supports
+an enforcement mechanism that does not yet have one.
+We use the formal tools of language-based
 security~\cite{sabelfeld2003language}, decomposing stack safety into
@@
 caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
+callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)), defined formally
+in \cref{sec:props}.
 %
-\rb{Forward references?}
 
@@
 the formal criteria for integrity, confidentiality, and control flow
-to obey the requirements. We consider the criteria in (3) to represent universal definitions
-that are parameterized by (1).
+to obey the requirements. This decomposition can be thought of as a step toward
+a universal definition, in that the definitions in (3) capture fundamental
+notions of security that can then be applied flexibly to other systems by varying
+the model in (1) and (2).
 %
@@
 
-Stack safety is challenging to specify in full generality, because the stack is fundamentally
-an architecture-specific construct. (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across architectures and language semantics, and therefore amenable
-to a very general approach.)
-The desired behavior of the stack varies based on ISA, calling convention,
-and the features of whichever programming language originated the code.
-%
-\rb{(Extremely important!)}
-%
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
-must be accompanied by tweaks to the model.
-%
-\rb{(Make this second point later?)}
+\sna{Not sure what your FIXME means exactly. Can you elaborate?}
 

### 2023-01-31 submission5/core.tex
comments

--- 2023-01-28 submission5/core.tex
@@
 on informal descriptions and examples of bad behaviors that they prevent.
+\apt{Can we connect stack safety more explicitly to a broader idea of ``subroutine safety''
+  as discussed in the intro? Note that this also includes registers.}
 
@@
 mechanisms instead leave us to ask: what security do we actually hope to gain,
-and are we successful?
-
-In this paper we focus on one such non-intuitive enforcement mechanism,
+and are we successful?\apt{For this audience, the need for a \emph{formal}
+  characterization doesn't need to be belabored.}
+
+In this paper we focus \apt{Is this still accurate?} on one such non-intuitive enforcement mechanism,
 the ``lazy'' stack safety micro-policies proposed by
@@
 to write into one another's frames, but taint the result so that the frame's
-owner cannot access it. No existent characterization of stack safety
+owner cannot access it. No existing characterization of stack safety
 captures this style of safety.
@@
 
-The stack governs a large number of language features that are often excluded
+The stack interacts with a large number of language features that are often excluded
 from discussion of stack safety. Our properties support a system with
@@
 to leak {\tt secret}. It might do so in a number of ways, shown as snippets of
-assembly code in Figure \ref{fig:f}.\footnote{We arrange that all the snippets
-have the same length just to simplify some later explanations.}
+assembly code in Figure \ref{fig:f}.
 %
@@
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
-violating the program's {\it well-bracketed control flow} (WBCF.) In
-Figure \ref{subfig:WBCF2}, a different attack violates WBCF, this time
+violating the program's {\it well-bracketed control flow} (WBCF.)
+%
+In Figure \ref{subfig:WBCF2}, a different attack violates WBCF, this time
 by returning to the correct program counter but with the wrong stack pointer.
+\footnote{We pad the last two variants with {\tt nop}s just so that all the
+snippets have the same length, which simplifies some later explanations.}
 
@@
 
-  \caption{Assembly code for {\tt f} as an attacker}
+  \caption{Assembly code alternatives for {\tt f} as an attacker.
+  }
   \label{fig:f}
@@
 
-Recall that our security semantics attach a security context to the machine state,
+Recall\apt{can't; we've lost these} that our security semantics attach a security context to the machine state,
 which consists of a view \(V\) and a stack \(\sigma\) of pending activations' views.
 (See \cref{sec:machine} for details.)
-Figure \ref{fig:exec1} shows how the security context develops over the first few
-steps of the program, according to the rules in \cref{fig:advops}, slightly simplified.
+Figure \ref{fig:exec1} shows how the security context evolves over the first few
+steps of the program.  (The rules governing this evolution are formalized in \cref{fig:advops}.)
 It begins at the start of {\tt main}, where the program counter (\(\PCname\)) is zero,
@@
 %
-At step 5, the formerly active principal's record is pushed onto the inactive list.
+At step 5, the current principal's record is pushed onto the inactive list.
 Its return target is the return address of the call,
@@
 For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments, 
-in this case just {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
+in this case just {\tt a0}. \apt{We've lost the explanation of the arguments to call, and they are used inconsistent now too, so this is utterly mysterious.}
+These are mapped to \(\public\), while any non-argument, caller-saved
 registers remain mapped to \(\unsealed\). All callee-save registers remain \(\sealed\) for all calls.
@@
 observable behavior of the machine. All other elements are \emph{relevant}.
-Variants and relevance are defined in \cref{sec:props}.
-
-We characterize \(\clri\) as the property that
+Variants and relevance are defined formally in \cref{sec:props}.
+
+We define \emph{caller integrity} (\(\clri\))  as the property that
 every relevant element that is \(\sealed\) under the callee's view is restored
@@
   \includegraphics[width=\columnwidth]{variants.png}
-  \caption{Integrity Violation}
+  \caption{Integrity Violation\apt{Needs a machine-drawn version, as do the rest.}}
   \label{fig:variant}
@@
 4 (its value in the variant) respectively. This is a violation of
-{\it internal confidentiality} (\cref{tab:props}, line 3a.)
+{\it internal confidentiality} (formalized in \cref{tab:props}, line 3a).
 
@@
 Unless {\tt a0} happens to be irrelevant, this is a violation of what
-we term {\it return-time confidentiality} (\cref{tab:props}, line 3b.)
+we term {\it return-time confidentiality} (formalized in \cref{tab:props}, line 3b).
 
@@
 
-In place of a predicate we define \(\confProp\) as a binary relation on states,
+In place of a predicate, we define \(\confProp\) as a binary relation on states,
 which holds on eventual return states \(\mach'\) and \(\nach'\)
@@
 
-Finally, we define \(\clrc\) as the
-combination of internal and return-time confidentiality.
+Finally, we define \emph{caller confidentiality} (\(\clrc\)) as the
+combination of internal and return-time confidentiality (\cref{tab:props}, line 3).
 
@@
 
-We presented our initial example from the perspective of the caller, a callee
+We presented our initial example from the perspective of the caller, but a callee
 may also have privilege that its caller lacks, and which must be protected from the
@@
 to cause it to misuse the key itself!
+\apt{I believe this is the first point at which we suggest that our properties might be
+  useful for enforcing (conventional) security goals --- which are definitely \emph{not}
+  part of the usual targets for ``stack safety'' mechanisms. This needs a lot more foreshadowing!}
 
@@
 
-The properties themselves correspond to their caller equivalents, inverted.
+The properties themselves correspond to their caller equivalents, inverted. \apt{Need at least their
+  names, with references to formalizations in the table. And for the discussion in the next paragraph
+  to make sense, we will need more than that. Especially because the formalization section now has
+nothing to say!}
 
@@
 
+\apt{This discussion should be merged with others on same topic, and maybe moved.}
 The forms of stack safety described above can potentially be enforced separately
@@
 We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
-triples of a register \(\reg\), a base offset \(off\), and a size \(sz\), denoting
-that the object at the offset \(off\) from \(\reg\) is to be passed as an argument.
+triples of a register \(\reg\), a base offset \(\mathit{off}\), and a size \(sz\), denoting
+that the object at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
 The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
@@
 
-\subsection{Public Allocations}
+\apt{Remainder of this section should be moved to AFTER the rest of the formalization.}
+
+\subsection{Memory Sharing}
 
@@
 
+\apt{Duplicates title of last section.}
+
 Now we are ready to introduce the formal properties that define correctly enforced
@@
 quantified over states \(s\) that immediately follow call (or tail-call) steps.
-If an execution can reach a state \(s\) such that \(s \stepstounder{\overline{\psi}} s'\)
-where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s'\) is the target
+If an execution can reach a state \(s'\) such that \(s' \stepstounder{\overline{\psi}} s\)
+where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
 of a call. Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
-\(s'\) is the target of a tailcall. As a shorthand, we write that each property is defined
-by a criterion that must hold ``when \(s'\) is called,'' i.e., for every such \(s'\).
+\(s\) is the target of a tailcall. As a shorthand, we write that each property is defined
+by a criterion that must hold ``for all (tail)call targets \(s\).'
 
@@
     \(\mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\)
-    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
+%    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
+    & \(\textnormal{ for all call targets } (\mach,(V,\sigma))\) \\
     %
@@
     \mach' \parallel \sealed(V) \cap \Delta(\mach,\mach')\)
-    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
+%    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
+    & \(\textnormal{ for all call targets } (\mach,(V,\sigma))\) \\
     %
@@
     & \(\textnormal{ where } \components = \sealed(V)\)
-    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+%    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+    & \(\textnormal{ for all (tail)call targets } (\mach,(V,\sigma))\) \\
     \rowcolor{black!20}
@@
     \mach' \parallel \Delta(\mach,\mach') - \components\)
-    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+%    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+    & \(\textnormal{ for all (tail)call targets } (\mach,(V,\sigma))\) \\
     \rowcolor{black!10}
@@
     & \(\textnormal{ where } \components = \public(V) \cup \object(V)\)
-    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+%    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
+    & \(\textnormal{ for all (tail)call targets } (\mach,(V,\sigma))\) \\
    \rowcolor{black!20}

### 2023-01-31 submission5/ideas-v2.tex
comments

@@
 
-Security properties tend to be conservative, identifying violations that would be
+Security properties tend to be conservative, identifying violations that may be
 harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
@@
 using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
-STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
+STAR \cite{???}, \apt{double-check this} every value in memory and registers is paired with a {\em metadata tag}.
 A set of software-defined rules, termed a micro-policy, determines the tags for operation
@@
 
-Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
-is inefficient when allocating and clearing stack frames, because each word must be initialized
-with a tag corresponding to the (depth of the) current function activation. They propose
+The most obvious way to enforce stack safety in this setting is to tag each word
+in the stack frame with the (depth of) its owning function activation: tags are
+set when the frame is allocated, checked against the current activation tag at each read or write,
+and cleared when the frame is deallocated. 
+However, Roessler and DeHon find that this approach is inefficient, because tag setting
+and clearing must be performed even on elements of the frame that are never accessed,
+which is quite common, e.g. in code that allocates unnecessarily large local arrays.
+They therefore propose
 {\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
 when a function writes to memory. This means that writes cannot be prevented, necessarily
-violating existing models of stack safety. Instead the micro-policy enforces that memory be
-read by the same fuction that wrote it. The owner of the data will never see the changes,
+violating existing models of stack safety.\apt{Like what??}  Instead the micro-policy enforces that memory be
+read by the same function activation that wrote it. The owner of the data will never see the changes,
 so it is secure in the sense that the attacker is unsuccessful at changing program behavior
-(except, perhaps, by causing a fault.) We elaborate that sense, and test whether the
+(except, perhaps, by causing a fault). We elaborate this notion of observational safety, and test whether the
 micro-policy fulfills it.
@@
 \subsection{A Formal and Flexible Security Framework}
+
+\apt{Sorry, but I don't understand what this section is trying to achieve.
+What is the ``comparison'' in Table~\cite{tab:relwork} for?
+It would be great if we could describe how our properties are suitable
+for describing the kinds of things various protection mechanisms are trying
+to acheive, and I can see a sketch of this going in related/future work, but 
+in fact we DON'T have any evidence for anything except micropolicies.}
 
@@
 
-\subsubsection{Low-Level Component Support}
+\subsubsection{Low-Level Component\apt{``Code''?} Support}
 %
@@
 some are likely to take priority in terms of their
-importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
+importance to security. Almost all stack-safety enforcement mechanisms will try to enforce {\wbcf},
 as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
 Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
-and callee-oriented properties even moreso.
+and callee-oriented properties even more so.
 
@@
 are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In \cref{sec:machine}), we cover additional operations that add parameters to the examples',
+In \cref{sec:machine}), we cover additional operations that add parameters to the examples,
 to support various forms of memory sharing \ifexceptions , exceptions, \fi
@@
 
-\sna{Might like a better name than object, because morally it ought to apply to arguments as well.}
+\sna{Might like a better name than object, because morally it ought to apply to arguments as well.\apt{Maybe ``active'' ?}}
 

### 2023-01-31 submission5/intro-v2.tex
comments

--- 2023-01-30 submission5/intro-v2.tex
@@
 
-Computer programming at any scale relies on \emph{functions} (variously known as
-subroutines, procedures, methods, etc.) as a \emph{foundational abstraction} to
-structure programs into building blocks representing units of computation that
-can \emph{call} one another to create larger computations.
+Computer programming % at any scale
+relies on \emph{functions} (variously known as
+subroutines, procedures, methods, etc.) as a foundational abstraction.
+\apt{General comment: I think it's better to reserve italics for terms being defined, and avoid using them just for rhetorical emphasis.}
+Functions are units of computation that
+can \emph{call} one another to realize larger computations.
 %
-At a low level, a function call comprises its own data (private or shared), as
+At a low level, a function call \apt{``activation''?} comprises its own data (private or shared), as
 well as information about its \emph{caller} and \emph{callee} functions. The
-runtime must keep track of the local contexts of all function activations
-at any given time; the \emph{fundamental data structure} charged with realizing
+runtime must keep track of the local contexts \apt{not defined} of all function activations
+at any given time. The fundamental data structure charged with realizing
 this abstraction is the \emph{(call) stack}.
@@
 
-The correct execution on a program---and its security!---is contingent on all of
+The correct execution of a program---and its security!---is contingent on all of
 its functions properly implementing this abstraction by means of a careful use
 of the call stack. Be it accidental or intentional, incorrect management of the
-stack is at the root of a myriad of faults whose effects range from (rarely)
-innocuous to (usually) catastrophic.
+stack is at the root of a myriad of faults whose effects range from % (rarely)
+innocuous to % (usually)
+catastrophic.
 %
@@
 structure) are as fundamental, varied and pervasive as the functional
-abstraction itself.
+abstraction itself
 %
 \rb{Begins to prefigure the idea of multiple stack implementations.}
+\apt{I think this is all well-known.}
 
@@
 consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read), \#7 (CWE-416, Use
-After Free) or \#11 (CWE-476, NULL Pointer Dereference).
+After Free) or \#11 (CWE-476, NULL Pointer Dereference).\apt{That last one is not usually a stack issue.}
 %
@@
 memory safety statistics, etc.}.
+\apt{OK to say, but this is well-known.}
 %
@@
 security concern---why?
+\apt{Mainly because preventing common stack problems generally causes a performance hit.
+  And, to a (much) lesser extent, because C/C++ don't carry length information with
+  arrays.  The sequencing here seems to suggest that we are going to \emph{solve} the
+  problem of enforcing stack safety by defining it better---which I don't buy at all!}
 
@@
 not fulfill a property benefit from the ability to articulate \emph{why} they fail.
+
 %
@@
 %
-This is the problem we begin to address in this paper \rb{or a few more such
-  connecting words}.
 
@@
   look at the idea of the properties.}
+
+\apt{This section was a nice summary of enforcement approaches, but I am not
+  yet seeing why it justifies so much paper real estate, especially if we only 
+  actually apply our property definitions to (lazy) micropolicies. And in any case,
+  it makes for a huge delay before getting to our work and contributions.}
 
@@
   I know what I'm doing''} while still detecting them if and when their effects
-become tangible.
+become observable. 
 %
@@
   (Coming in key ideas?) No ``gold standard''?}
+\apt{Yes, there are at least two different kinds of parameterization going on.
+  One has to do with which properties we want to, or are able to, enforce; e.g.,
+  for a language with tail calls or exceptions, we need a more complex definition of the WBCF property itself.
+  Another has to do with how standard features, like argument passing, are implemented
+  by particular hardware/ABIs -- which shouldn't show up in the property definition.
+  We need to be clearer about this.}
 

### 2023-01-31 submission5/core.tex

@@
 
-\subsection{Memory Sharing}
-
-[TODO: this section could use some trimming. I think we have room for this
-discussion, it's just more verbose than everything else and therefore feels weird.]
-
-The full machine supports public stack allocations that can
-be accessed by anyone until they are deallocated on their owner's return,
-represented by \(\mathbf{alloc} ~ \mathbf{t} \dots\).
-For example, consider the following code:
-
-{\tt
-  void g() \{
-
-  ~ ~ int x[3];
-
-  ~ ~ int y;
-
-  ~ ~ h(\&x);
-
-  \}
-}
-
-Here we have allocated several arrays and variables; {\tt y}
-will behave as normal with \(\mathbf{alloc} ~ \mathbf{f} \dots\), but
-{\tt x} has its address taken and passed to another function. That
-function might do anything at all to {\tt x}, and might store its pointer
-elsewhere (in the heap, for instance) so that the scope of legitimate accesses
-to {\tt x} are uncertain. We default to assuming that any address-taken local
-should be public. Meanwhile {\tt y} is still private.
-
-In the likely event that the space for the return address, {\tt x}, and {\tt y}
-are all allocated in a single step, that step makes multiple allocation operations:
-\[\bar{\psi} = [\mathbf{alloc} ~ \mathbf{f} ~ (-20,4);
- \mathbf{alloc} ~ \mathbf{t} ~ (-16,12) ; \mathbf{alloc} ~ \mathbf{f} ~ (-4,4)]\]
-
+\subsection{Sharing Stack Memory}
+In our examples so far, we have presented a vision of stack safety in which
+the interface between caller and callee is in the registers that pass
+arguments and return values. This is frequently not the case in a realistic
+setting. Arguments may be passed on the stack due to being spilled, as an implementation
+of variadic arguments, or because they are reference objects that inherently have
+pass-by-reference semantics. The caller may also take the address of a local object and
+pass that as a pointer to its callee.
+
+We refine our call operation to make use of the information that we have about
+which memory contain arguments, \(\overline{sa}\). \(\overline{sa}\) is a set of
+triples of a register, an offset from the value of that register, and a size.
+We first define the helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\),
+then extend the call operation to keep all objects in \(\mathit{passed}\) marked
+as \(\object\) and seal everything else (\cref{sfig:stkargs}).
+
+In this mechanism, we can describe an argument being passed by their relative address
+with the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
+In the case of pass-by-reference semantics, if the compiler designates register
+\(\reg\) as holding the reference, then the operations is
+\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
+
+In both cases, the immediate callee retains access to the passed object, but future
+callees will not---unless the object is passed-by-reference further down the stack.
+On the other hand, in the case of a caller explicitly taking the address of an
+object and passing it as a pointer, we have no way obvious way to limit the scope
+of such sharing.
+
+Our primary machine treats this scenario as the caller allocating a ``public''
+object that can be accessed by anyone until it's deallocated. We extend the
+\(\mathbf{alloc}\) operation with a boolean flag, where {\bf t} indicates
+that the allocation is public, and {\bf f} is private.
+In the likely event that the space for multiple objects is allocated in a single step,
+that step can make multiple allocation operations, each labeled appropriately.
 Allocating public objects affects the call stack very similarly to private ones,
 but instead of labeling them \(\object\) they become \(\public\), so they are
-not sealed at a call. This makes these
-regions of memory able to transmit information between caller and callee without
-violating the stack safety properties. Note that there is no way to seal a location
-once it has become \(\public\); only when {\tt x} is deallocated will its region
-be available to allocate privately. We define a helper function, \(\mathit{range}\),
-to map an offset from a register and a size into a set of addresses, and update
-the operation's security semantics as shown in \cref{sfig:publicalloc}.
+never sealed at a call (\cref{sfig:publicalloc}).
 

### 2023-01-31 submission5/intro-v2.tex

--- 2023-01-30 submission5/intro-v2.tex
@@
 
-Computer programming at any scale relies on \emph{functions} (variously known as
-subroutines, procedures, methods, etc.) as a \emph{foundational abstraction} to
-structure programs into building blocks representing units of computation that
-can \emph{call} one another to create larger computations.
+Computer programming % at any scale
+relies on \emph{functions} (variously known as
+subroutines, procedures, methods, etc.) as a foundational abstraction.
+\apt{General comment: I think it's better to reserve italics for terms being defined, and avoid using them just for rhetorical emphasis.}
+Functions are units of computation that
+can \emph{call} one another to realize larger computations.
 %
-At a low level, a function call comprises its own data (private or shared), as
+At a low level, a function call \apt{``activation''?} comprises its own data (private or shared), as
 well as information about its \emph{caller} and \emph{callee} functions. The
-runtime must keep track of the local contexts of all function activations
-at any given time; the \emph{fundamental data structure} charged with realizing
+runtime must keep track of the local contexts \apt{not defined} of all function activations
+at any given time; the fundamental data structure charged with realizing
 this abstraction is the \emph{(call) stack}. But at the level of the ABI
@@
 stack is at the root of many faults whose effects range from (rarely)
-innocuous to (usually) catastrophic.
+innocuous to (usually) catastrophic.\apt{Long-winded.}
 %
@@
 structure) are as fundamental, varied and pervasive as the functional
-abstraction itself.
+abstraction itself
 %
+
+\apt{I think this is all well-known.}
 
@@
 \sna{Trying to think how to make the case that these CWEs are about the stack, since they are
-  primarily described as memory attacks.}
+  primarily described as memory attacks.}\apt{Agree: the out-of-bounds ones are often on the stack, but the NULL pointer deref is not.
+Also, this story is very well known, and doesn't need belaboring.}
 %
@@
 Yet despite this wealth of potential solutions, the stack remains a central
-security concern.
+security concern---why?
+\apt{Mainly because preventing common stack problems generally causes a performance hit.
+  And, to a (much) lesser extent, because C/C++ don't carry length information with
+  arrays.  The sequencing here seems to suggest that we are going to \emph{solve} the
+  problem of enforcing stack safety by defining it better---which I don't buy at all!}
 
@@
 %
-Our work is focused on the creation of such properties for stack safety.
 
 \subsection{Related Work}
+
+\apt{This section was a nice summary of enforcement approaches, but I am not
+  yet seeing why it justifies so much paper real estate, especially if we only 
+  actually apply our property definitions to (lazy) micropolicies. And in any case,
+  it makes for a huge delay before getting to our work and contributions.}
 

### 2023-01-31 submission5/ideas-v1.tex
Moved things around in the introduction and went back to v1 of key ideas. Also trimmed down v1 even further.

@@
 
-\sna{Might like a better name than object, because morally it ought to apply to arguments as well.}
-
-\paragraph*{The Properties and their Hierarchy}
-
-The properties form a hierarchy, shown in [TODO: diagram].
-Within properties that are orthogonal
-some are likely to take priority in terms of their
-importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
-as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
-Many also cover some degree of (caller) integrity. Caller confidentiality has been fairly neglected,
-and callee-oriented properties even moreso.
-
 \paragraph*{Events}
@@
 to any notion of observable behavior.
+
+\paragraph*{Variants and Non-interference}
+
+We model confidentiality by drawing on
 
@@
 This is a strong threat model, but it does omit some important aspects of
-security enforcement: In particular, hardware and timing attacks are out of scope.
+security enforcement: in particular, hardware and timing attacks are out of scope.
+
+\paragraph*{Tagged Architectures and Laziness}
+
+In \cref{sec:testing}, we apply our properties to test the existing stack-safety
+mechanism of Roessler and DeHon \cite{}, who present a collection of tag-based enforcement
+mechanisms using PIPE \cite{???}, a tagged hardware reference monitor.
+In monitors such as PIPE and STAR \cite{???}, \apt{double-check this} every value in memory
+and registers is paired with a {\em metadata tag}.
+A set of software-defined rules, termed a micro-policy, determines the tags for operation
+results based on the tags of their inputs---or, if the input tags are in a dangerous
+combination, it may halt the machine (a ``failstop.'')
+
+We are particularly interested in their {\em Lazy Tagging and Clearing} micro-policy.
+In their conventional micro-policies, they tag each word
+in a stack frame with the (depth of) its owning function activation.
+Roessler and DeHon find that it is inefficient, because tag setting
+and clearing must be performed at every allocation, even on elements that are never accessed,
+or that are initialized again and accessed long afterward. This is quite common, e.g. in code
+that allocates large local arrays.
+
+Instead they offer Lazy Tagging and Clearing, in which tags are not initialized at allocation, but only
+when a function writes to memory. This means that writes cannot be prevented, which in turn means
+that the existing line of stack safety specifications from Georges et al. will deem it unsafe.
+Instead the micro-policy enforces that memory be read by the same function activation that wrote it.
+The owner of the data will never see the changes,
+so it is secure in the sense that the attacker is unsuccessful at changing program behavior
+(except, perhaps, by causing a fault). We elaborate this notion of observational safety, and test whether the
+micro-policy fulfills it.
+
+\paragraph*{Randomized Property-based Testing}
+
+[TODO: Brief discussion of PBT and why we use it, explain the lack of proofs.]

### 2023-01-31 submission5/intro-v2.tex
Moved things around in the introduction and went back to v1 of key ideas. Also trimmed down v1 even further.

@@
 
-\rb{A.K.A. motivation}
-
-Computer programming % at any scale
-relies on \emph{functions} (variously known as
-subroutines, procedures, methods, etc.) as a foundational abstraction.
-\apt{General comment: I think it's better to reserve italics for terms being defined, and avoid using them just for rhetorical emphasis.}
-Functions are units of computation that
-can \emph{call} one another to realize larger computations.
+Computer programming relies on functions (subroutines, procedures, methods, etc.)
+as a foundational abstraction. Functions are units of computation that
+can call one another to realize larger computations.
 %
-At a low level, a function call \apt{``activation''?} comprises its own data (private or shared), as
-well as information about its \emph{caller} and \emph{callee} functions. The
-runtime must keep track of the local contexts \apt{not defined} of all function activations
-at any given time; the fundamental data structure charged with realizing
-this abstraction is the \emph{(call) stack}. But at the level of the ABI
-\sna{Do we need to unpack the acronym ABI?}, the specification of a well-behaved
-stack is minimal. It is not concerned with such questions as whether a caller's
-frame should be readable or writable to its callee, leaving those concerns
-to be determined by a high-level language built atop the stack.
-
-The correct execution---and the security!---of a program is contingent on all of
-its functions properly implementing the intended abstraction by means of a careful use
-of the call stack. Be it accidental or intentional, incorrect management of the
-stack is at the root of many faults whose effects range from (rarely)
-innocuous to (usually) catastrophic.\apt{Long-winded.}
-%
-From a security perspective, attacks on the stack (say, from a malicious
-function that chooses to abuse its access to the contents of this data
-structure) are as fundamental, varied and pervasive as the functional
-abstraction itself
-%
-
-\apt{I think this is all well-known.}
-
-%% \rb{\textbf{The call stack and its security.} Maybe a few opening words to set
-%%   the stage and bring two things to the front: (1) functions and subroutines as
-%%   fundamental building blocks of any kind of structured programming; and (2) the
-%%   stack as the natural abstraction to manage these in a computer. And also maybe
-%%   a little bit more about the width, the gravity and the \emph{foundational
-%%   nature} of attacks on the stack, which the sentence below touches on
-%%   summarily.
-%% \rb{(+SNA) Connection to functions (and adversarial code?). Attacks on the stack
-%%   are also attacks on this fundamental abstraction (functions, subroutines,
-%%   procedures, methods, however you call them)}
-%% }
+At a low level, a function activation comprises its own data, as
+well as information about its \emph{caller} function, to which it must return.
+The \emph{(call) stack} is the fundamental data structure charged with realizing
+this abstraction. From a security perspective, attacks on the stack are attacks
+on the fundamental functional abstraction itself, and are therefore every bit as
+pervasive.
 
@@
 %
-Yet despite this wealth of potential solutions, the stack remains a central
-security concern---why?
-\apt{Mainly because preventing common stack problems generally causes a performance hit.
-  And, to a (much) lesser extent, because C/C++ don't carry length information with
-  arrays.  The sequencing here seems to suggest that we are going to \emph{solve} the
-  problem of enforcing stack safety by defining it better---which I don't buy at all!}
 
-\paragraph{Formal Properties}
+\paragraph{Towards a Formal Property for Stack Safety}
 
@@
 to find new, similar attacks.
-%
-%% \rb{And so the cat-and-mouse game continues.}
-%
-To mitigate this risk, we claim that mechanisms should aim to enforce \emph{formal
-properties} of safe behavior that can be proven or tested. Such properties become the
+
+To mitigate this risk, mechanisms should aim to enforce \emph{formal properties}
+of safe behavior that can be proven or tested. Such properties become the
 \emph{specification} against which enforcement can be validated; even enforcement mechanisms that do
@@
 
+But most such properties are described at the level of a high-level programming language.
+The stack as an abstraction exists below the level of
+any language that uses it, and at the level of the ABI
+\sna{Do we need to unpack the acronym ABI?}, the specification of a well-behaved
+stack is minimal. It is not concerned with such questions as whether a caller's
+frame should be readable or writable to its callee, leaving those concerns
+to be determined by a high-level language built atop the stack.
+
+Specifying stack safety in full generality would not merely be a matter of handling
+a wide range of architectural variations. The basic security behavior that we expect
+of the stack varies based on calling convention, compiler behavior, and the features of
+whichever programming language originated the code.
+(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
+which is broadly consistent across language semantics, and therefore
+amenable to a very general approach.)
+
+For instance, the current state-of-the-art stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable.
+
+In this work, we propose an explicit specification, built from first principles
+in the hopes that it will be widely applicable, and demonstrate that it supports
+an enforcement mechanism that does not yet have one.
+We use the formal tools of language-based
+security~\cite{sabelfeld2003language}, decomposing stack safety into
+the {\em integrity} and {\em confidentiality} of the caller’s local state
+and the callee's behavior during the callee's execution, plus the control-flow protection
+of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
+\ifexceptions
+We also extend WBCF to a setting with exceptions,
+in which there may be multiple points at which a callee can ``return.''
+\fi
 %
-%% \rb{Good stuff! There is some overlap between this discussion and the opening
-%%   motivation, maybe we could restructure things a bit so that those threads are
-%%   unified?}
+We treat stack safety as multiple properties, rather than one multi-faceted property, because
+not all enforcement mechanisms attempt to implement all aspects of stack safety (see \cref{???}.)
 %
+This gives us a total of five core properties: \(\wbcf\),
+caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
+callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)), defined formally
+in \cref{sec:props}.
 
-\subsection{Related Work}
+We are not attempting to create a universal definition of stack safety for all systems.
+Rather, we define the stack safety of a single, fairly realistic system, decomposed into
+(1) a lightweight model of the system and its security-relevant features, from which we
+derive the integrity, confidentiality, and control-flow requirements of each call and
+return, (2) the application of that model to a particular target system, and (3)
+the formal criteria for integrity, confidentiality, and control flow
+to obey the requirements. This decomposition can be thought of as a step toward
+a universal definition, in that the definitions in (3) capture fundamental
+notions of security that can then be applied flexibly to other systems by varying
+the model in (1) and (2).
+%
+\rb{FIXME: Contrast some universal definitions (or a parametric framework) with
+  the concrete instance we study, again with its interest properly motivated.
+  (Coming in key ideas?) No ``gold standard''?}
+\apt{Yes, there are at least two different kinds of parameterization going on.
+  One has to do with which properties we want to, or are able to, enforce; e.g.,
+  for a language with tail calls or exceptions, we need a more complex definition of the WBCF property itself.
+  Another has to do with how standard features, like argument passing, are implemented
+  by particular hardware/ABIs -- which shouldn't show up in the property definition.
+  We need to be clearer about this.}
+\sna{In discussion above, I tried to carve out some separation there, but more may be needed.}
+
+To demonstrate the utility of our formal characterization, we use these
+properties to validate and improve an existing enforcement mechanism, the
+{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
+in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
+use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for Coq, to generate random programs and check
+that Roessler and DeHon's micro-policies correctly detect the ones that
+attempt to violate one of our properties. Furthermore, we
+check that the testing framework is able to generate counterexamples
+that violate our properties but are \emph{not} halted by incorrect
+variants of the enforcement mechanisms, including ones that we accidentally created
+while implementating the micro-policy and ones that we
+intentionally crafted to increase our confidence in the effectiveness
+of our testing.
+%
+\rb{This and the following paragraph could start more generally (we need to
+  answer the question of why these experiments, what is their interest), and
+  probably motivate the use of PBT (experiment results below show there's
+  benefit in them!) and, somewhere at least, talk about other points in the
+  spectrum.}
+
+We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
+violates the temporal aspect of confidentiality in
+cases where data can leak across repeated calls to the same callee,
+and also violates integrity if the leak uses the caller's frame. We
+propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
+confidentiality, albeit at some performance cost.
+%
+\rb{Anything to add from personal communications? More details later? Shows
+  these things are very hard to get right -- like the latest work on Cerise,
+  also motivated by this very work.
+\rb{(+SNA) At the very least, these models are subtle and getting them right,
+  and describing them unambiguously, is hard without a framework like this.}
+}
+
+\subsection{Contributions}
+
+In sum, we offer the following contributions:
+
+\begin{itemize}
+\item We formalize permissive stack confidentiality and integrity properties,
+  which are parameterized over a notion of external
+  observation, and are violated only if accessing secrets or overwriting
+  data causes a visible change in system behavior.
+\item We instantiate these definitions with a realistic
+  system with features such as argument passing
+  on the stack, callee-saves registers, and tail-call elimination.
+  Our model is modular enough that adding these features is straightforward.
+\item Our formalization of well-bracketed control flow in the presence of tail-call
+  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
+\item We use property-based random testing to validate the relationship between
+  our properties and micro-policies.
+\end{itemize}
+
+Our artifact contains Coq formalizations of our properties for illustrative purposes
+as well as their testing implementation. There are no proofs---we use primarily Coq for the
+QuickChick library and to ensure that the formalizations are unambiguous.
+
+\subsection{Related Work: The Stack Safety Menagerie}
 
@@
   it makes for a huge delay before getting to our work and contributions.}
+\sna{I've tried out moving it here. To my mind, its main purpose is to justify
+  our claims about the wide variation in stack-safety enforcement, and therefore
+  to help argue that our properties not necessarily being universal is
+  not a flaw.}
 
@@
 
-\subsection{A Language-Based View of Stack Safety}
+\paragraph{Summary}
 
@@
 Only a handful of works have offered any explicit specification.
-
-Stack safety is challenging to specify in full generality, because the stack is
-fundamentally an architecture-specific construct.
-(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across architectures and language semantics, and therefore
-amenable to a very general approach.)
-The desired behavior of the stack varies based on ISA, calling convention,
-and the features of whichever programming language originated the code.
-%
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable.
-
-In this work, we propose an explicit specification, built from first principles
-in the hopes that it will be widely applicable, and demonstrate that it supports
-an enforcement mechanism that does not yet have one.
-We use the formal tools of language-based
-security~\cite{sabelfeld2003language}, decomposing stack safety into
-the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the control-flow protection
-of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
-\ifexceptions
-We also extend WBCF to a setting with exceptions,
-in which there may be multiple points at which a callee can ``return.''
-\fi
-%
-We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all aspects of stack safety. Some focus
-exclusively on control flow, some enforce integrity but not confidentiality, and many
-focus on the caller over the callee.
-%
-\rb{Return to this after extended discussion? Forward ref?}
-%
-This gives us a total of five core properties: \(\wbcf\),
-caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)), defined formally
-in \cref{sec:props}.
-%
-
-We are not attempting to create a universal definition of stack safety for all systems.
-Rather, we define the stack safety of a single, fairly realistic system, decomposed into
-(1) a lightweight model of the system and its security-relevant features, from which we
-derive the integrity, confidentiality, and control-flow requirements of each call and
-return, (2) the application of that model to a particular target system, and (3)
-the formal criteria for integrity, confidentiality, and control flow
-to obey the requirements. This decomposition can be thought of as a step toward
-a universal definition, in that the definitions in (3) capture fundamental
-notions of security that can then be applied flexibly to other systems by varying
-the model in (1) and (2).
-%
-\rb{FIXME: Contrast some universal definitions (or a parametric framework) with
-  the concrete instance we study, again with its interest properly motivated.
-  (Coming in key ideas?) No ``gold standard''?}
-\apt{Yes, there are at least two different kinds of parameterization going on.
-  One has to do with which properties we want to, or are able to, enforce; e.g.,
-  for a language with tail calls or exceptions, we need a more complex definition of the WBCF property itself.
-  Another has to do with how standard features, like argument passing, are implemented
-  by particular hardware/ABIs -- which shouldn't show up in the property definition.
-  We need to be clearer about this.}
-
-\sna{Not sure what your FIXME means exactly. Can you elaborate?}
-
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly detect the ones that
-attempt to violate one of our properties. Furthermore, we
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms, including ones that we accidentally created
-while implementating the micro-policy and ones that we
-intentionally crafted to increase our confidence in the effectiveness
-of our testing.
-%
-\rb{This and the following paragraph could start more generally (we need to
-  answer the question of why these experiments, what is their interest), and
-  probably motivate the use of PBT (experiment results below show there's
-  benefit in them!) and, somewhere at least, talk about other points in the
-  spectrum.}
-
-We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-%
-\rb{Anything to add from personal communications? More details later? Shows
-  these things are very hard to get right -- like the latest work on Cerise,
-  also motivated by this very work.
-\rb{(+SNA) At the very least, these models are subtle and getting them right,
-  and describing them unambiguously, is hard without a framework like this.}
-}
-
-\subsection{Contributions}
-
-In sum, we offer the following contributions:
-
-\begin{itemize}
-\item We formalize permissive stack confidentiality and integrity properties,
-  which are parameterized over a notion of external
-  observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in system behavior.
-\item We instantiate these definitions with a realistic
-  system with features such as argument passing
-  on the stack, callee-saves registers, and tail-call elimination.
-  Our model is modular enough that adding these features is straightforward.
-\item Our formalization of well-bracketed control flow in the presence of tail-call
-  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
-\item We use property-based random testing to validate the relationship between
-  our properties and micro-policies.
-\end{itemize}
-
-Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as their testing implementation. There are no proofs---we use primarily Coq for the
-QuickChick library and to ensure that the formalizations are unambiguous.
-%
-\rb{Will have another look after the full pass.}

### 2023-01-31 submission5/core.tex
Moved things around in the introduction and went back to v1 of key ideas. Also trimmed down v1 even further.
Moved advanced machine later again.

@@
 
-\input{ideas-v2}
+\input{ideas-v1}
 
@@
 \(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\).
-The full machine defines additional operations, and extends some of the old ones.
-We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
-triples of a register \(\reg\), a base offset \(\mathit{off}\), and a size \(sz\), denoting
-that the object at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
-The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
-optimizations in which the callee will reuse the caller's stack. The \(\mathbf{alloc}\) operation
-gains a boolen flag indicating whether the allocation is to be public, accessible to
-all functions until deallocated. The rules are given in \cref{fig:advops}; the
-example is equivalent to these rules with \(\mathbf{call}\) always instantiated with
-\(\overline{sa}\) as the empty set, and \(\mathbf{alloc}\) with \(\mathbf{f}\).
+For the machine described in the example, the \(Op\) rules would be those
+found in \cref{fig:basicops}.
+
+\begin{figure}
+  \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+    \{\addr \mid V ~ \addr = \unsealed\}\)}
+             {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
+             {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+    %
+    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+                 ~ (V,\sigma) = (V'',V::\sigma)\)}
+    %
+    \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
+             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
+  \caption{Basic Operations}
+  \label{fig:basicops}
+\end{figure}
 
@@
 
-\apt{Remainder of this section should be moved to AFTER the rest of the formalization.}
-
-\subsection{Sharing Stack Memory}
-In our examples so far, we have presented a vision of stack safety in which
-the interface between caller and callee is in the registers that pass
-arguments and return values. This is frequently not the case in a realistic
-setting. Arguments may be passed on the stack due to being spilled, as an implementation
-of variadic arguments, or because they are reference objects that inherently have
-pass-by-reference semantics. The caller may also take the address of a local object and
-pass that as a pointer to its callee.
-
-We refine our call operation to make use of the information that we have about
-which memory contain arguments, \(\overline{sa}\). \(\overline{sa}\) is a set of
-triples of a register, an offset from the value of that register, and a size.
-We first define the helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\),
-then extend the call operation to keep all objects in \(\mathit{passed}\) marked
-as \(\object\) and seal everything else (\cref{sfig:stkargs}).
-
-In this mechanism, we can describe an argument being passed by their relative address
-with the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
-In the case of pass-by-reference semantics, if the compiler designates register
-\(\reg\) as holding the reference, then the operations is
-\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
-
-In both cases, the immediate callee retains access to the passed object, but future
-callees will not---unless the object is passed-by-reference further down the stack.
-On the other hand, in the case of a caller explicitly taking the address of an
-object and passing it as a pointer, we have no way obvious way to limit the scope
-of such sharing.
-
-Our primary machine treats this scenario as the caller allocating a ``public''
-object that can be accessed by anyone until it's deallocated. We extend the
-\(\mathbf{alloc}\) operation with a boolean flag, where {\bf t} indicates
-that the allocation is public, and {\bf f} is private.
-In the likely event that the space for multiple objects is allocated in a single step,
-that step can make multiple allocation operations, each labeled appropriately.
-Allocating public objects affects the call stack very similarly to private ones,
-but instead of labeling them \(\object\) they become \(\public\), so they are
-never sealed at a call (\cref{sfig:publicalloc}).
-
-\begin{figure*}
-  \begin{subfigure}{0.4\textwidth}
-    \[\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach \triangleq
-    \{\mach[\reg]+i | \mathit{off} \leq i < \mathit{off+sz}\}\]
-
-    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
-                 \{\addr \mid V ~ \addr = \unsealed\}\)}
-               {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-
-    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
-                 \{\addr \mid V ~ \addr = \unsealed\}\)}
-               {\(V' = V \llbracket \addr \mapsto \public \mid \addr \in \components \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-    %
-    \judgmentbr{\(b = \mach[\SP] + \mathit{off}\)}
-               {\(V' = V \llbracket \addr \mapsto \unsealed |
-                 b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-
-    \caption{Memory Allocation}
-    \label{sfig:publicalloc}
-  \end{subfigure}
-  \begin{subfigure}{0.6\textwidth}
-    %
-    \[\mathit{passed} ~ \overline{sa} ~ \mach = \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
-    \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
-    %
-    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
-                 ~ (V,\sigma) = (V'',V::\sigma)\)}
-    %
-    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{tailcall} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
-                 ~ (V,\sigma) = (V',\sigma)\)}
-    %
-    \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
-
-    \caption{Calls with Argument Passing on the Stack}
-    \label{sfig:stkargs}
-  \end{subfigure}
-  \caption{Operations}
-  \label{fig:advops}
-\end{figure*}
-
-\subsection{Tail Calls}
-
-The operation rule for a tailcall is similar to that for a normal call.
-We do not push the caller's view onto the stack,
-but replace it outright. This means that a tailcall does not increase the size of
-the call stack, and therefore for purposes of our properties, all tailcalls will
-be considered to return simultaneously when the eventual {\bf return} operation
-pops the top of the stack.
-
-Since the caller will not be returned to, it does not need integrity, though
-it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
-than \(\sealed\) to express this. Of course, some objects may be passed as arguments
-in place, easily modeled by including them in \(\overline{sa}\). More sophisticated
-behaviors may require sequences of {\bf dealloc}, {\bf tailcall}, then {\bf alloc}.
-
 \section{Formal properties}
 \label{sec:props}
-
-\apt{Duplicates title of last section.}
 
@@
 \end{table*}
+
+\section{Expanding the Machine}
+
+The system we've modeled so far has been very simple, but our properties
+support an important dimension of generality: adding features. We will now
+briefly introduce the features of our primary testing machine.
+
+The full machine defines additional operations, and extends some of the old ones.
+We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
+triples of a register \(\reg\), a base offset \(\mathit{off}\), and a size \(sz\), denoting
+that the object at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
+The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
+optimizations in which the callee will reuse the caller's stack. The \(\mathbf{alloc}\) operation
+gains a boolen flag indicating whether the allocation is to be public, accessible to
+all functions until deallocated. The rules are given in \cref{fig:advops}; the
+rules in \cref{fig:basicops} can be recaptured by instantiating
+\(\mathbf{call}\) with \(\overline{sa}\) as the empty set, and \(\mathbf{alloc}\)
+with \(\mathbf{f}\).
+
+\subsection{Sharing Stack Memory}
+In our examples, we have presented a vision of stack safety in which
+the interface between caller and callee is in the registers that pass
+arguments and return values. This is frequently not the case in a realistic
+setting. Arguments may be passed on the stack due to being spilled, as an implementation
+of variadic arguments, or because they are reference objects that inherently have
+pass-by-reference semantics. The caller may also take the address of a local object and
+pass that as a pointer to its callee.
+
+We refine our call operation to make use of the information that we have about
+which memory contain arguments, \(\overline{sa}\). \(\overline{sa}\) is a set of
+triples of a register, an offset from the value of that register, and a size.
+We first define the helpful set \(\mathit{passed} ~ \overline{sa} ~ \mach\),
+then extend the call operation to keep all objects in \(\mathit{passed}\) marked
+as \(\object\) and seal everything else (\cref{sfig:stkargs}).
+
+In this mechanism, we can describe an argument being passed by their relative address
+with the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
+In the case of pass-by-reference semantics, if the compiler designates register
+\(\reg\) as holding the reference, then the operations is
+\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
+
+In both cases, the immediate callee retains access to the passed object, but future
+callees will not---unless the object is passed-by-reference further down the stack.
+On the other hand, in the case of a caller explicitly taking the address of an
+object and passing it as a pointer, we have no way obvious way to limit the scope
+of such sharing.
+
+Our primary machine treats this scenario as the caller allocating a ``public''
+object that can be accessed by anyone until it's deallocated. We extend the
+\(\mathbf{alloc}\) operation with a boolean flag, where {\bf t} indicates
+that the allocation is public, and {\bf f} is private.
+In the likely event that the space for multiple objects is allocated in a single step,
+that step can make multiple allocation operations, each labeled appropriately.
+Allocating public objects affects the call stack very similarly to private ones,
+but instead of labeling them \(\object\) they become \(\public\), so they are
+never sealed at a call (\cref{sfig:publicalloc}).
+
+\begin{figure*}
+  \begin{subfigure}{0.4\textwidth}
+    \[\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach \triangleq
+    \{\mach[\reg]+i | \mathit{off} \leq i < \mathit{off+sz}\}\]
+
+    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+                 \{\addr \mid V ~ \addr = \unsealed\}\)}
+               {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+
+    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+                 \{\addr \mid V ~ \addr = \unsealed\}\)}
+               {\(V' = V \llbracket \addr \mapsto \public \mid \addr \in \components \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+    %
+    \judgmentbr{\(b = \mach[\SP] + \mathit{off}\)}
+               {\(V' = V \llbracket \addr \mapsto \unsealed |
+                 b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+
+    \caption{Memory Allocation}
+    \label{sfig:publicalloc}
+  \end{subfigure}
+  \begin{subfigure}{0.6\textwidth}
+    %
+    \[\mathit{passed} ~ \overline{sa} ~ \mach = \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
+    \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
+    %
+    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+                 ~ (V,\sigma) = (V'',V::\sigma)\)}
+    %
+    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{tailcall} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+                 ~ (V,\sigma) = (V',\sigma)\)}
+    %
+    \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
+             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
+
+    \caption{Calls with Argument Passing on the Stack}
+    \label{sfig:stkargs}
+  \end{subfigure}
+  \caption{Operations}
+  \label{fig:advops}
+\end{figure*}
+
+\subsection{Tail Calls}
+
+The operation rule for a tailcall is similar to that for a normal call.
+We do not push the caller's view onto the stack,
+but replace it outright. This means that a tailcall does not increase the size of
+the call stack, and therefore for purposes of our properties, all tailcalls will
+be considered to return simultaneously when the eventual {\bf return} operation
+pops the top of the stack.
+
+Since the caller will not be returned to, it does not need integrity, though
+it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
+than \(\sealed\) to express this. Of course, some objects may be passed as arguments
+in place, easily modeled by including them in \(\overline{sa}\). More sophisticated
+behaviors may require sequences of {\bf dealloc}, {\bf tailcall}, then {\bf alloc}.
 

### 2023-02-01 submission5/intro-v2.tex

@@
 shows general classes of vulnerabilities closely related to the stack \rb{(though not exclusively)}
-consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read), \#7 (CWE-416, Use After Free) or \#11 (CWE-476, NULL Pointer Dereference).
+consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read) or \#7 (CWE-416, Use After Free).
 %
@@
 \sna{Trying to think how to make the case that these CWEs are about the stack, since they are
-  primarily described as memory attacks.}\apt{Agree: the out-of-bounds ones are often on the stack, but the NULL pointer deref is not.
+  primarily described as memory attacks.}\apt{Agree: the out-of-bounds ones are often on the stack, but the NULL pointer deref is not \rb{Indeed, removed the last one.}
 Also, this story is very well known, and doesn't need belaboring.}
@@
 \fi
-%
-\rb{Broad enough? Do we want to add more techniques?}
-\sna{I think this is plenty unless there's a really major thing we're forgetting.}
-%
 

### 2023-02-01 submission5/core.tex
Reorganizing the introduction to try to thread the needle right away.

@@
 memory. But, unlike ``the heap,'' the stack does not correspond to a single
-high-level language concept, making a formal specification elusive.
-Protection mechanisms typically lack precise specifications, relying instead
-on informal descriptions and examples of bad behaviors that they prevent.
-\apt{Can we connect stack safety more explicitly to a broader idea of ``subroutine safety''
-  as discussed in the intro? Note that this also includes registers.}
-
-When the prevented behaviors are straightforward, describing them may be sufficient
-to justify an enforcement mechanism's effectiveness. Less intuitive enforcement
-mechanisms instead leave us to ask: what security do we actually hope to gain,
-and are we successful?\apt{For this audience, the need for a \emph{formal}
-  characterization doesn't need to be belabored.}
-
-In this paper we focus \apt{Is this still accurate?} on one such non-intuitive enforcement mechanism,
+high-level language concept. Rather, it corresponds to the fundamental
+abstraction of functions, in all of the forms that they can take.
+The protean nature of the functional abstraction makes stack safety
+difficult to specify.
+
+We propose a formal characterization of stack safety based on
+concepts from language-based security. Stack safety is decomposed
+into an integrity property and a confidentiality property for each
+of the caller and the callee, as well as a control-flow property.
+
+Our motivating enforcement mechanism,
 the ``lazy'' stack safety micro-policies proposed by
-Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}. These permit functions
+Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, permit functions
 to write into one another's frames, but taint the result so that the frame's
 owner cannot access it. No existing characterization of stack safety
-captures this style of safety.
-
-We propose a formal characterization of ``observable'' stack safety based on
-concepts from language-based security. Stack safety is a
-combination of an integrity property (``the private
-state in each caller's stack frame is held invariant by the callee''),
-two confidentiality properties (``the caller's
-and callee's behavior are insensitive to one another's private
-state,''), and a control flow property. By defining the properties
-in terms of the observable behavior of the system, we expand the
-space of supported enforcement mechanisms to include those that
-allow certain (intuitively incorrect) behaviors but prevent them
-from causing harm.
+captures this style of safety. We capture it by defining the properties in
+terms of the observable behavior of the system.
 
@@
 both caller- and callee-saved registers, arguments passed on the stack,
-tail-call elimination, [exceptions, and concurrency.] They are modular by
-design, in order to be further extensible to other features and architectures.
-
-We use these properties to validate Roessler and DeHon's micro-policies
+tail-call elimination, \ifexceptions exceptions \fi. They are modular by
+design, in order to be further extensible to other features.
+
+We validate our properties by using them to distinguish correct implementations
+of Roessler and DeHon's micro-policies from incorrect ones
 via property-based random testing. Our testing successfully detects violations
 in several broken variants, including Roessler and DeHon's original lazy policy.
+A fixed version of the policy does pass our tests.
 

### 2023-02-01 submission5/intro-v2.tex
Reorganizing the introduction to try to thread the needle right away.

@@
 
-Indeed, the call stack is a perennial target for low-level attacks, leading to
-dire consequences, from leakage and corruption of private stack data
-to control-flow hijacking.
+Indeed, the call stack is a perennial target for low-level attacks.
 %
@@
   memory safety statistics, etc.}.
-\sna{Trying to think how to make the case that these CWEs are about the stack, since they are
-  primarily described as memory attacks.}\apt{Agree: the out-of-bounds ones are often on the stack, but the NULL pointer deref is not \rb{Indeed, removed the last one.}
-Also, this story is very well known, and doesn't need belaboring.}
-%
-To foil such attacks, a menagerie of
-software and hardware protections have been proposed,
+
+The variety in attacks on the stack is mirrored by the 
+software and hardware protections that aim to prevent them,
 %
@@
 and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}---we discuss these in more
-detail in \cref{XXX}.
-  \ifaftersubmission\apt{Mostly from
-  nick; there could be more}\bcp{Yes, going back to MIT days---we should
-  include several more of these, if only to give readers the impression that
-this is a well-studied mechanism (so formalizing its protections is
-useful).}
-\fi
+detail in \cref{XXX}. Few of these articulate any formal property that they aim to enforce,
+with the exception of Skorstengaard and Georges. Some are fundamentally ill-suited for
+formalization, as they aim to impede attackers but could not provide universal protection.
 
-\paragraph{Towards a Formal Property for Stack Safety}
-
-Security mechanisms designed to prevent a certain attack can be brittle---successfully
-eliminating the attack as it exists at the time of publication, but leaving room for attackers
-to find new, similar attacks.
-
-To mitigate this risk, mechanisms should aim to enforce \emph{formal properties}
-of safe behavior that can be proven or tested. Such properties become the
-\emph{specification} against which enforcement can be validated; even enforcement mechanisms that do
-not fulfill a property benefit from the ability to articulate \emph{why} they fail.
-
-But most such properties are described at the level of a high-level programming language.
-The stack as an abstraction exists below the level of
-any language that uses it, and at the level of the ABI
-\sna{Do we need to unpack the acronym ABI?}, the specification of a well-behaved
-stack is minimal. It is not concerned with such questions as whether a caller's
-frame should be readable or writable to its callee, leaving those concerns
-to be determined by a high-level language built atop the stack.
-
-Specifying stack safety in full generality would not merely be a matter of handling
-a wide range of architectural variations. The basic security behavior that we expect
-of the stack varies based on calling convention, compiler behavior, and the features of
-whichever programming language originated the code.
-(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across language semantics, and therefore
-amenable to a very general approach.)
-
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable.
-
-In this work, we propose an explicit specification, built from first principles
+In this work, we propose such a property, built from first principles
 in the hopes that it will be widely applicable, and demonstrate that it supports
@@
 \fi
-%
-We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all aspects of stack safety (see \cref{???}.)
-%
-This gives us a total of five core properties: \(\wbcf\),
-caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)), defined formally
-in \cref{sec:props}.
+
+Our target for validation is the ``Lazy Tagging and Clearing'' \(LTC\) micropolicy of Roessler
+and DeHon \cite{DBLP:conf/sp/RoesslerD18}. \(LTC\) is peculiar in that function activations
+may write into one another's stack framess, as long as the data cannot be accessed afterward.
+It is therefore too weak to be characterized by the state-of-the-art properties from
+Georges et al. We give a characterization that does apply to it, and expand the features that
+it supports.
+
+\paragraph{Towards a Formal Property for Stack Safety}
+
+Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
+other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
+or tested. Such properties become the specification against which enforcement can be validated;
+even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
+\emph{why} they fail.
+
+But most such properties are described at the level of a high-level programming language.
+The stack as an abstraction exists below the level of
+any language that uses it, and at the level of the ABI, the specification of a well-behaved
+stack is minimal. It is not concerned with such questions as whether a caller's
+frame should be readable or writable to its callee, leaving those concerns
+to be determined by a high-level language built atop the stack.
+(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
+which is broadly consistent across language semantics, and therefore
+amenable to a very general approach.)
+
+For instance, the current state-of-the-art stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable.
 
@@
 %
-\rb{FIXME: Contrast some universal definitions (or a parametric framework) with
-  the concrete instance we study, again with its interest properly motivated.
-  (Coming in key ideas?) No ``gold standard''?}
-\apt{Yes, there are at least two different kinds of parameterization going on.
-  One has to do with which properties we want to, or are able to, enforce; e.g.,
-  for a language with tail calls or exceptions, we need a more complex definition of the WBCF property itself.
-  Another has to do with how standard features, like argument passing, are implemented
-  by particular hardware/ABIs -- which shouldn't show up in the property definition.
-  We need to be clearer about this.}
-\sna{In discussion above, I tried to carve out some separation there, but more may be needed.}
+We treat stack safety as multiple properties, rather than one multi-faceted property, because
+not all enforcement mechanisms attempt to implement all aspects of stack safety (see \cref{???}.)
+%
+This gives us a total of five core properties: \(\wbcf\),
+caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
+callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)), defined formally
+in \cref{sec:props}.
 
-To demonstrate the utility of our formal characterization, we use these
-properties to validate and improve an existing enforcement mechanism, the
+We demonstrate that our properties are general in that they apply easily to systems with various
+features, and that they support ``al la carte'' enforcement based on the security needs of the
+system.
+
+To validate our formal characterization, we use these
+properties to test and improve an existing enforcement mechanism, the
 {\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented

### 2023-02-01 submission5/intro-v2.tex
Checkpoint, missing things to intro next

@@
 
-Computer programming relies on functions (subroutines, procedures, methods, etc.)
+Computer programming relies on functions (or subroutines, procedures, methods, etc.)
 as a foundational abstraction. Functions are units of computation that
@@
 The \emph{(call) stack} is the fundamental data structure charged with realizing
-this abstraction. From a security perspective, attacks on the stack are attacks
+this abstraction. From a security perspective, attacks on the stack are really attacks
 on the fundamental functional abstraction itself, and are therefore every bit as
@@
 list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
-shows general classes of vulnerabilities closely related to the stack \rb{(though not exclusively)}
+shows general classes of vulnerabilities closely related to the stack
+%% \rb{(though not exclusively)}
 consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read) or \#7 (CWE-416, Use After Free).
@@
 Moreover, many other types of faults involve some kind of control or data flow
-using data stored in the stack \rb{cf. Eternal war in memory, Google/Microsoft
-  memory safety statistics, etc.}.
+using data stored in the stack \cite{DBLP:conf/raid/VeendCB12,
+  DBLP:conf/sp/SzekeresPWS13,
+  DBLP:conf/sp/HuSACSL16}.\footnote{\url{https://github.com/Microsoft/MSRC-Security-Research/blob/master/presentations/2019_02_BlueHatIL/}}\footnote{\url{https://www.chromium.org/Home/chromium-security/memory-safety/}}
 
@@
 detail in \cref{XXX}. Few of these articulate any formal property that they aim to enforce,
-with the exception of Skorstengaard and Georges. Some are fundamentally ill-suited for
-formalization, as they aim to impede attackers but could not provide universal protection.
+with the exception of Skorstengaard and Georges \rb{(refs, and add a ``but'' here that motivates our development!)}. Others are fundamentally ill-suited for
+formalization, as they aim to impede attackers but could never provide universal protection
+against whole classes of attacks.
 
@@
 in the hopes that it will be widely applicable, and demonstrate that it supports
-an enforcement mechanism that does not yet have one.
+an \rb{(flexible? ``general''?)} enforcement mechanism that does not yet have one.
 We use the formal tools of language-based
@@
 and the callee's behavior during the callee's execution, plus the control-flow protection
-of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
+of {\em well-bracketed control flow} (\(\wbcf\))~\cite{SkorstengaardSTKJFP}.
 \ifexceptions
@@
 
-Our target for validation is the ``Lazy Tagging and Clearing'' \(LTC\) micropolicy of Roessler
+Our target for validation is the {\em Lazy Tagging and Clearing} (\(LTC\)) micro-policy of Roessler
 and DeHon \cite{DBLP:conf/sp/RoesslerD18}. \(LTC\) is peculiar in that function activations
-may write into one another's stack framess, as long as the data cannot be accessed afterward.
+may write into one another's stack frames, as long as the data cannot be accessed afterward.
 It is therefore too weak to be characterized by the state-of-the-art properties from
-Georges et al. We give a characterization that does apply to it, and expand the features that
-it supports.
+Georges et al. \rb{(Do we explain why anywhere?)} We give a characterization that does apply to it, and expand the features that
+it supports \rb{(try to be a bit more explicit?)}.
 
@@
 amenable to a very general approach.)
+\rb{(Really, the abstraction makes the problem largely if not completely transparent at the language level: these are low-level attacks that break that abstraction)}
 
@@
 \cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
+control flow, and therefore defines {\em Well-bracketed Control Flow} (WBCF) under the assumption that
 a callee should always return to its caller. But, in the presence of tail-call elimination, a
@@
 
-We are not attempting to create a universal definition of stack safety for all systems.
-Rather, we define the stack safety of a single, fairly realistic system, decomposed into
+We are not attempting to create \rb{(``to present''?)} a universal definition of stack safety for all systems.
+Rather, we define the stack safety of a single, fairly realistic system \rb{(if exotic capability machines count as realistic, this is no less so than those)}, decomposed into
 (1) a lightweight model of the system and its security-relevant features, from which we
@@
 We demonstrate that our properties are general in that they apply easily to systems with various
-features, and that they support ``al la carte'' enforcement based on the security needs of the
+features, and that they support {\it \`a la carte} enforcement based on the security needs of the
 system.

### 2023-02-01 submission5/intro-v2.tex
When in doubt, go bold

--- 2023-02-01 submission5/intro-v2.tex
@@
 shadow stacks~\cite{Dang+15,Shanbhogue+19},
-capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}---we discuss these in more
-detail in \cref{XXX}. Few of these articulate any formal property that they aim to enforce,
-with the exception of Skorstengaard and Georges. Some are fundamentally ill-suited for
-formalization, as they aim to impede attackers but could not provide universal protection.
+capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
 
-In this work, we propose such a property, built from first principles
-in the hopes that it will be widely applicable, and demonstrate that it supports
-an enforcement mechanism that does not yet have one.
-We use the formal tools of language-based
-security~\cite{sabelfeld2003language}, decomposing stack safety into
+Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
+as they aim to impede attackers but cannot provide universal protection. Stack canaries,
+for instance, aim to reduce the ability of attackers to find gadgets, not to categorically
+rule out attacks. Others may be amenable to formalization but do not offer it.
+The Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
+Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
+to offer (various) stack safety formalizations.
+
+Meanwhile, Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
+micro-policies for stack safety, at least some of which should be amenable to making
+formal guarantees. Among these is their ``Lazy Tagging and Clearing'' \(LTC\) micropolicy,
+which is is peculiar in that function activations may write into one another's stack frames,
+as long as the data cannot be accessed afterward. As a result, \(LTC\) definitely would not
+fulfill a tag-based adaptation Georges et al.'s state-of-the-art property.
+Roessler and DeHon themselves do not claim that it intends to enforce stack safety,
+only an approximation.
+
+We disagree. \(LTC\) does not always protect individual addresses on the stack, but it aims
+to prevent one function activation from interfering with another's behavior---to protect
+the functional abstraction. We argue that a formal characterization of \(LTC\)'s protections
+should fall under the umbrella of stack safety.
+
+In this work, we propose such a characterization, built from first principles
+in the hopes that it will be broadly applicable, and demonstrate
+that it can distinguish between correct and flawed versions of Roessler and DeHon's
+micro-policies. We use the formal tools of language-based
+security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
+properties describing
 the {\em integrity} and {\em confidentiality} of the caller’s local state
@@
 of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
+
+The structure of our properties enables them to be easily extended to systems with
+sophisticated features such as callee-save registers, argument-passing on the stack,
+and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
+test the extended micro-policy.
 \ifexceptions
-We also extend WBCF to a setting with exceptions,
+We extend WBCF to a setting with exceptions,
 in which there may be multiple points at which a callee can ``return.''
 \fi
-
-Our target for validation is the ``Lazy Tagging and Clearing'' \(LTC\) micropolicy of Roessler
-and DeHon \cite{DBLP:conf/sp/RoesslerD18}. \(LTC\) is peculiar in that function activations
-may write into one another's stack framess, as long as the data cannot be accessed afterward.
-It is therefore too weak to be characterized by the state-of-the-art properties from
-Georges et al. We give a characterization that does apply to it, and expand the features that
-it supports.
+It also enables us to separate out only those properties that an enforcement mechanism
+actually attempts to enforce.
 

### 2023-02-01 submission5/intro-v2.tex

@@
 fulfill a tag-based adaptation Georges et al.'s state-of-the-art property.
+\rb{(Do we explain why anywhere?)}
 Roessler and DeHon themselves do not claim that it intends to enforce stack safety,
@@
 \fi
-It also enables us to separate out only those properties that an enforcement mechanism
-actually attempts to enforce.
 

### 2023-02-01 submission5/intro-v2.tex

@@
 and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
-test the extended micro-policy.
-\ifexceptions
-We extend WBCF to a setting with exceptions,
-in which there may be multiple points at which a callee can ``return.''
-\fi
+test the extended micro-policy. We find that, apart from the extensions, \(LTC\)
+is flawed in a way that undermines integrity and confidentiality, and we correct
+this flaw.
 
@@
 
-Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
-other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
-or tested. Such properties become the specification against which enforcement can be validated;
-even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
-\emph{why} they fail.
+%Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
+%other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
+%or tested. Such properties become the specification against which enforcement can be validated;
+%even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
+%\emph{why} they fail.
 
-But most such properties are described at the level of a high-level programming language.
-The stack as an abstraction exists below the level of
-any language that uses it, and at the level of the ABI, the specification of a well-behaved
-stack is minimal. It is not concerned with such questions as whether a caller's
-frame should be readable or writable to its callee, leaving those concerns
-to be determined by a high-level language built atop the stack.
+Most security properties are described at the level of a high-level programming language,
+and translated to the target-level via a secure compiler. Stack safety cannot
+be defined in this way, as the abstraction is completely transparent in most source languages.
 (Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
 which is broadly consistent across language semantics, and therefore
-amenable to a very general approach.)
-\rb{(Really, the abstraction makes the problem largely if not completely transparent at the language level: these are low-level attacks that break that abstraction)}
+amenable to a very general high-level approach.)
+But it is not satisfying as a low-level property; at the level of the ABI, the
+specification of a well-behaved stack is minimal. It is not concerned with such
+questions as whether a caller's frame should be readable or writable to its callee,
+leaving those concerns to be determined by a high-level language built atop the stack.
 
@@
 a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable.
+return from a callee to a distant ancestor is perfectly reasonable. Related, their predecessor
+Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
+their enforcement mechanism did protect the confidentiality of the caller (not the callee.)
 
-We are not attempting to create \rb{(``to present''?)} a universal definition of stack safety for all systems.
-Rather, we define the stack safety of a single, fairly realistic system \rb{(if exotic capability machines count as realistic, this is no less so than those)}, decomposed into
+These two cases represent two dimensions of generality that make stack safety difficult to
+define as a single, universal property: what constitutes safe behavior may differ between
+systems that support different features, and different types of security may be more or less
+desirable depending on context.
+
+We are not attempting to present a universal definition of stack safety for all systems.
+Rather, we define the stack safety of a single, fairly realistic system (one with common
+features such as memory arguments and tailcall elimination), decomposed into
 (1) a lightweight model of the system and its security-relevant features, from which we
@@
 the model in (1) and (2).
-%
-We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all aspects of stack safety (see \cref{???}.)
-%
-This gives us a total of five core properties: \(\wbcf\),
-caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)), defined formally
-in \cref{sec:props}.
 
-We demonstrate that our properties are general in that they apply easily to systems with various
-features, and that they support {\it \`a la carte} enforcement based on the security needs of the
-system.
+To address the second class of generality, we present a total of five core properties:
+\(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
+callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)). These are defined
+formally in \cref{sec:props}. They are related to one another as seen in
+\cref{fig:propertydiagram}: \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
+\(\clri\), but otherwise the properties are orthogonal. We further define two
+important combinations of properties: full stack safety \(\textnormal{\sc FSS}\),
+and caller safety \(\textnormal{\sc ClrS}\). Of these, most existing concepts of
+stack safety correspond to either \(\wbcf\) or \(\textnormal{\sc ClrS}\), which are
+concerned with the basic guarantees that the caller recieves during the call, while
+\(\textnormal{\sc FSS}\) is a natural extension that enforces a higher-level concept
+of a correct interface between caller and callee. \sna{Blah. Needs work.}
 

### 2023-02-02 submission5/core.tex
Checkpoint, missing things to intro next
typo
Move menagerie to future work

@@
 \usepackage[noadjust]{cite}
+
+\usepackage[multiple]{footmisc}
+
+\usepackage[hyphens]{url}
 
@@
 \(s\) is the target of a tailcall. As a shorthand, we write that each property is defined
-by a criterion that must hold ``for all (tail)call targets \(s\).'
+by a criterion that must hold ``for all (tail)call targets \(s\).''
 
@@
 
-From here, we see two natural ways to extend the work.
-First, we should thoroughly evaluate
-other enforcement mechanisms, especially Cheri-based capability
-systems. Second, we can further demonstrate
-the functionality of our definitions by using them to prove
-the correctness of a programming logic with strengthened rules around calls and returns.
-
-\paragraphx{Stack Safety in the Cheri Ecosystem}
-%
-There are several proposals around the use of Cheri capabilities to enforce stack safety,
-including mechanisms that use the standard Cheri hardware (which includes local
-capabilities) \cite{SkorstengaardLocal},
-and others that propose entirely new types of capabilities, such as linear
-\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
-\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
-Of these, the last seems particularly promising for testing.
-[TODO: more about this. (Rob's suggestion of citing Aina could go here.)]
+\input{future_work.tex}
 

### 2023-02-02 submission5/future_work.tex
Move menagerie to future work

--- 2023-02-01 submission5/intro-v2.tex
@@
-\subsection{The Call Stack and Its Security}
+\subsection{Testing The Stack Safety Menagerie}
 
-Computer programming relies on functions (subroutines, procedures, methods, etc.)
-as a foundational abstraction. Functions are units of computation that
-can call one another to realize larger computations.
-%
-At a low level, a function activation comprises its own data, as
-well as information about its \emph{caller} function, to which it must return.
-The \emph{(call) stack} is the fundamental data structure charged with realizing
-this abstraction. From a security perspective, attacks on the stack are attacks
-on the fundamental functional abstraction itself, and are therefore every bit as
-pervasive.
-
-Indeed, the call stack is a perennial target for low-level attacks.
-%
-The most recent release of the authoritative CWE Top 25
-Most Dangerous Software Weaknesses
-list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
-shows general classes of vulnerabilities closely related to the stack \rb{(though not exclusively)}
-consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read) or \#7 (CWE-416, Use After Free).
-%
-Moreover, many other types of faults involve some kind of control or data flow
-using data stored in the stack \rb{cf. Eternal war in memory, Google/Microsoft
-  memory safety statistics, etc.}.
-
-The variety in attacks on the stack is mirrored by the 
-software and hardware protections that aim to prevent them,
-%
-including stack canaries~\cite{Cowan+98},
-bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
-split stacks~\cite{Kuznetsov+14},
-shadow stacks~\cite{Dang+15,Shanbhogue+19},
-capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
-
-Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
-as they aim to impede attackers but cannot provide universal protection. Stack canaries,
-for instance, aim to reduce the ability of attackers to find gadgets, not to categorically
-rule out attacks. Others may be amenable to formalization but do not offer it.
-The Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
-Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
-to offer (various) stack safety formalizations.
-
-Meanwhile, Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
-micro-policies for stack safety, at least some of which should be amenable to making
-formal guarantees. Among these is their ``Lazy Tagging and Clearing'' \(LTC\) micropolicy,
-which is is peculiar in that function activations may write into one another's stack frames,
-as long as the data cannot be accessed afterward. As a result, \(LTC\) definitely would not
-fulfill a tag-based adaptation Georges et al.'s state-of-the-art property.
-Roessler and DeHon themselves do not claim that it intends to enforce stack safety,
-only an approximation.
-
-We disagree. \(LTC\) does not always protect individual addresses on the stack, but it aims
-to prevent one function activation from interfering with another's behavior---to protect
-the functional abstraction. We argue that a formal characterization of \(LTC\)'s protections
-should fall under the umbrella of stack safety.
-
-In this work, we propose such a characterization, built from first principles
-in the hopes that it will be broadly applicable, and demonstrate
-that it can distinguish between correct and flawed versions of Roessler and DeHon's
-micro-policies. We use the formal tools of language-based
-security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
-properties describing
-the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the control-flow protection
-of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
-
-The structure of our properties enables them to be easily extended to systems with
-sophisticated features such as callee-save registers, argument-passing on the stack,
-and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
-test the extended micro-policy.
-\ifexceptions
-We extend WBCF to a setting with exceptions,
-in which there may be multiple points at which a callee can ``return.''
-\fi
-It also enables us to separate out only those properties that an enforcement mechanism
-actually attempts to enforce.
-
-\paragraph{Towards a Formal Property for Stack Safety}
-
-Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
-other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
-or tested. Such properties become the specification against which enforcement can be validated;
-even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
-\emph{why} they fail.
-
-But most such properties are described at the level of a high-level programming language.
-The stack as an abstraction exists below the level of
-any language that uses it, and at the level of the ABI, the specification of a well-behaved
-stack is minimal. It is not concerned with such questions as whether a caller's
-frame should be readable or writable to its callee, leaving those concerns
-to be determined by a high-level language built atop the stack.
-(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across language semantics, and therefore
-amenable to a very general approach.)
-
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable.
-
-We are not attempting to create a universal definition of stack safety for all systems.
-Rather, we define the stack safety of a single, fairly realistic system, decomposed into
-(1) a lightweight model of the system and its security-relevant features, from which we
-derive the integrity, confidentiality, and control-flow requirements of each call and
-return, (2) the application of that model to a particular target system, and (3)
-the formal criteria for integrity, confidentiality, and control flow
-to obey the requirements. This decomposition can be thought of as a step toward
-a universal definition, in that the definitions in (3) capture fundamental
-notions of security that can then be applied flexibly to other systems by varying
-the model in (1) and (2).
-%
-We treat stack safety as multiple properties, rather than one multi-faceted property, because
-not all enforcement mechanisms attempt to implement all aspects of stack safety (see \cref{???}.)
-%
-This gives us a total of five core properties: \(\wbcf\),
-caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)), defined formally
-in \cref{sec:props}.
-
-We demonstrate that our properties are general in that they apply easily to systems with various
-features, and that they support ``al la carte'' enforcement based on the security needs of the
-system.
-
-To validate our formal characterization, we use these
-properties to test and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly detect the ones that
-attempt to violate one of our properties. Furthermore, we
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms, including ones that we accidentally created
-while implementating the micro-policy and ones that we
-intentionally crafted to increase our confidence in the effectiveness
-of our testing.
-%
-\rb{This and the following paragraph could start more generally (we need to
-  answer the question of why these experiments, what is their interest), and
-  probably motivate the use of PBT (experiment results below show there's
-  benefit in them!) and, somewhere at least, talk about other points in the
-  spectrum.}
-
-We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-%
-\rb{Anything to add from personal communications? More details later? Shows
-  these things are very hard to get right -- like the latest work on Cerise,
-  also motivated by this very work.
-\rb{(+SNA) At the very least, these models are subtle and getting them right,
-  and describing them unambiguously, is hard without a framework like this.}
-}
-
-\subsection{Contributions}
-
-In sum, we offer the following contributions:
-
-\begin{itemize}
-\item We formalize permissive stack confidentiality and integrity properties,
-  which are parameterized over a notion of external
-  observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in system behavior.
-\item We instantiate these definitions with a realistic
-  system with features such as argument passing
-  on the stack, callee-saves registers, and tail-call elimination.
-  Our model is modular enough that adding these features is straightforward.
-\item Our formalization of well-bracketed control flow in the presence of tail-call
-  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
-\item We use property-based random testing to validate the relationship between
-  our properties and micro-policies.
-\end{itemize}
-
-Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as their testing implementation. There are no proofs---we use primarily Coq for the
-QuickChick library and to ensure that the formalizations are unambiguous.
-
-\subsection{Related Work: The Stack Safety Menagerie}
-
-\apt{This section was a nice summary of enforcement approaches, but I am not
-  yet seeing why it justifies so much paper real estate, especially if we only 
-  actually apply our property definitions to (lazy) micropolicies. And in any case,
-  it makes for a huge delay before getting to our work and contributions.}
-\sna{I've tried out moving it here. To my mind, its main purpose is to justify
-  our claims about the wide variation in stack-safety enforcement, and therefore
-  to help argue that our properties not necessarily being universal is
-  not a flaw.}
-
-The protections offered by stack protection mechanisms are commonly described by giving
-concrete examples of attacks that they can prevent---corruption of return
-addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more formal characterization to the reader's intuition.
-%
-To make these notions more precise, we review a selection of specimens from the
-menagerie and begin to consider the difficulties involved in specifying and
-verifying the protections they offer.
-%
-For an assessment of these methods in the context of our stack security
-framework, see \cref{XXX}.
-
-At the highest level, we can classify techniques into two large groups depending
-on whether their protection is furnished at the software or the hardware level.
-Up to date, most proposals, including all commonly deployed practical
-protections, are implemented in software, so we start by considering exemplars
-from this first group.
+In \cref{sec:intro}, we discussed the four dimensions of extensibility
+that our properties should exhibit: architectures, enforcement mechanisms,
+features, and requirements. We have focused on extending our basic system
+with advanced features; a logical next step is extending to other enforcement
+mechanisms. In the following, we discuss how different existing mechanisms
+might fit into our model.
 
@@
 
-\paragraph{Stack Canaries}
+\paragraph{Stack Canaries and Shadow Stacks}
 %
 Canaries are special values inserted by the caller at certain points of the
-stack and checked for corruption on return, primarily for protection against
-some classes of control-flow hijacking (i.e., seeking to overwrite the caller's
+stack and checked for corruption on return, for protection against
+some classes of control-flow hijacking (i.e., overwriting the caller's
 return address through a buffer overflow that would also overwrite the canary
-detectably).
+detectably). A shadow stack is a mirror image of the control-relevant parts of the stack,
+maintained by the program to attempt to detect attempts to hijack its control
+flow and ``restrict the flexibility available in creating gadget chains''
+\cite{Shanbhogue+19}.
 %
-Observe that this is a \emph{hardening} technique that increases the difficulty
-of some control-flow attacks on the stack, but can never provide any absolute
-guarantees on control flow integrity; neither does it do anything to protect
-function data stored in the stack.
-%
-Also interestingly, a successful attack is not detected the moment it takes
-place, instead action is \emph{deferred} until the attacker returns control to
-the caller.
-
-\paragraph{Shadow Stacks}
-%
-A shadow stack is a mirror image of the control-relevant parts of the stack,
-maintained by the program to attempt to detect attempts to hijack its control
-flow. They are similar in purpose and can be seen as related to stack canaries.
-Like canaries, they offer hardened but not guaranteed, deferred protection
-against control-flow attacks, ``restrict the flexibility available in creating
-gadget chains'' \cite{Shanbhogue+19}, and do not cover the data-relevant regions
-of the stack.
-%
-\rb{Could probably find other variants (e.g. SmashGuard, on hardware?), but this
-  seems to be the basic idea. Some of those require OS support, etc.}
+These are hardening techniques that increase the difficulty
+of some control-flow attacks on the stack, but cannot provide absolute
+guarantees on \(\wbcf\) under a normal attacker model.
+Interestingly, these are also lazy enforcement mechanisms, in that
+the attack may occur and be detected some time later, before taking the
 
@@
 bounds checking discipline.
-%
-\rb{Some related work, e.g. HardBound, is inspired by this but more closely
-  related to hardware capabilities.}
-
-\sna{Would HardBound warrant a section in the hardware techniques section,
-  for hardware-accelerated bounds checking (that is not capability-based)?}
 

### 2023-02-02 submission5/intro-v2.tex
Move menagerie to future work

@@
 QuickChick library and to ensure that the formalizations are unambiguous.
-
-\subsection{Related Work: The Stack Safety Menagerie}
-
-\apt{This section was a nice summary of enforcement approaches, but I am not
-  yet seeing why it justifies so much paper real estate, especially if we only 
-  actually apply our property definitions to (lazy) micropolicies. And in any case,
-  it makes for a huge delay before getting to our work and contributions.}
-\sna{I've tried out moving it here. To my mind, its main purpose is to justify
-  our claims about the wide variation in stack-safety enforcement, and therefore
-  to help argue that our properties not necessarily being universal is
-  not a flaw.}
-
-The protections offered by stack protection mechanisms are commonly described by giving
-concrete examples of attacks that they can prevent---corruption of return
-addresses, buffer overflows, use of uninitialized variables, etc.---leaving
-a more formal characterization to the reader's intuition.
-%
-To make these notions more precise, we review a selection of specimens from the
-menagerie and begin to consider the difficulties involved in specifying and
-verifying the protections they offer.
-%
-For an assessment of these methods in the context of our stack security
-framework, see \cref{XXX}.
-
-At the highest level, we can classify techniques into two large groups depending
-on whether their protection is furnished at the software or the hardware level.
-Up to date, most proposals, including all commonly deployed practical
-protections, are implemented in software, so we start by considering exemplars
-from this first group.
-
-\subsubsection{Software Techniques}
-
-\paragraph{Stack Canaries}
-%
-Canaries are special values inserted by the caller at certain points of the
-stack and checked for corruption on return, primarily for protection against
-some classes of control-flow hijacking (i.e., seeking to overwrite the caller's
-return address through a buffer overflow that would also overwrite the canary
-detectably).
-%
-Observe that this is a \emph{hardening} technique that increases the difficulty
-of some control-flow attacks on the stack, but can never provide any absolute
-guarantees on control flow integrity; neither does it do anything to protect
-function data stored in the stack.
-%
-Also interestingly, a successful attack is not detected the moment it takes
-place, instead action is \emph{deferred} until the attacker returns control to
-the caller.
-
-\paragraph{Shadow Stacks}
-%
-A shadow stack is a mirror image of the control-relevant parts of the stack,
-maintained by the program to attempt to detect attempts to hijack its control
-flow. They are similar in purpose and can be seen as related to stack canaries.
-Like canaries, they offer hardened but not guaranteed, deferred protection
-against control-flow attacks, ``restrict the flexibility available in creating
-gadget chains'' \cite{Shanbhogue+19}, and do not cover the data-relevant regions
-of the stack.
-%
-\rb{Could probably find other variants (e.g. SmashGuard, on hardware?), but this
-  seems to be the basic idea. Some of those require OS support, etc.}
-
-\paragraph{Bounds Checking}
-%
-Under a bounds checking discipline, all the pointers in a program are extended
-with some disjoint metadata, and these are combined to gate memory accesses,
-usually with the objective of enforcing some form of \emph{memory
-safety} \rb{(relate this to the basic high-level concepts and types of attacks
-we want to consider: control flow, some integrity and confidentiality?)}.
-%
-Compared to stack canaries and shadow stacks, bounds checking is eminently
-a \emph{high-level} technique, designed with a particular programming language
-in mind and implemented through a \emph{secure compiler} \cite{} that performs
-the necessary code transformations. Also unlike those techniques (which make few
-assumptions on callee code), it offers little to no protection
-against \emph{adversarial low-level} code that is not constrained to obey the
-bounds checking discipline.
-%
-\rb{Some related work, e.g. HardBound, is inspired by this but more closely
-  related to hardware capabilities.}
-
-\sna{Would HardBound warrant a section in the hardware techniques section,
-  for hardware-accelerated bounds checking (that is not capability-based)?}
-
-\paragraph{Split Stacks}
-%
-Code-pointer integrity seeks to prevent control-flow hijacking attacks by
-splitting the program memory, including the stack, into a regular region and a
-safe region; objects that require protection are placed in the latter and
-all accesses to them protected by static or dynamic checks.
-%
-Like shadow stacks, bounds checking is a high-level technique designed to
-protect code written in a high-level language, but it is designed to
-offer protection under a stronger threat model that also considers the presence
-of untrusted code. As with most other software-based methods, it purports to
-protect exclusively the control-relevant parts of the stack, making no claims
-about its data-relevant parts.
-
-\subsubsection{Hardware Techniques}
-
-In addition to the many software-based techniques developed to provide greater
-degrees of protection to the stack, often incurring a significant performance penalty,
-some proposals have turned to specialized hardware features to offset those
-costs and attempt to offer greater levels of protection.
-
-We focus in particular on two families of hardware mechanisms: capability machines
-and tag-based reference monitors. Both types of system move beyond simply
-offering hardware acceleration for software-based techniques, offering hardware
-primitives for both memory and control-flow enforcement that can offer a degree
-of protection even with minimal cooperation from the code.
-%
-\rb{Something about how realistic/practical these alternatives are? Are we
-  selecting partly based on prototypes, etc.?}
-
-\paragraph{Capabilities}
-%
-\rb{(There is plenty of variety here, should we focus on the Cerise-and-friends
-  (CHERI-adjacent) work exclusively?)}
-
-\sna{If we have good examples of other capability systems, we should absolutely
-  talk about them. No need to privilege CHERI.}
-%
-Capability machines can represent and manipulate
-unforgeable tokens of authority over specific regions of
-memory using a combination of dedicated out-of-band memory and
-specialized ISA instructions.
-%
-The most prominent contemporary representative of this long line of work is
-CHERI \cite{}, a modern architecture designed to provide efficient fine-grained
-memory protection and compartmentalization.
-%
-Previous work has uses these principles in simplified models to implement
-\emph{secure calling conventions} for the call
-stack \cite{}. Those conventions can be seen as combining the
-protections of bounds-checking with software-based methods of control-flow protection.
-%
-They are closely related \rb{too strong?} to this work; we discuss them in more
-detail in \cref{XXX}.
-%
-
-\paragraph{Micro-Policies}
-%
-\rb{(This partly overlaps with the technical preliminaries, try to offer a quick
-  synthesis here.)}
-%
-Micro-policy machines are a family of tagged architectures that decorate
-addressable memory and registers with arbitrary metadata tags, and extend the
-processor with a programmable hardware monitor that uses custom rules on
-metadata tags to implement many different kinds of security policies.
-%
-This flexibility can be harnessed to implement a wide array of stack protection policies with relative ease, covering and going beyond the capabilities \rb{word choice?} of the various software-based techniques \cite{} \rb{as demonstrated by}.
-%
-These policies comprise very general forms of \emph{lazy enforcement}, which can tolerate
-potential but harmless violations \rb{the system programmer's motto: ``trust me,
-  I know what I'm doing''} while still detecting them if and when their effects
-become observable. 
-%
-Micro-policies are one of the centerpieces of our technical development; we
-present them in more detail in \cref{XXX}.
-
-\paragraph{Summary}
-
-We begin to see that the security of the call stack and its contents is not
-clearly defined. Different protection mechanisms aim to protect the stack
-in fundamentally different ways, each offering a \emph{unique}, \emph{implicit}
-specification of the meaning of ``stack safety.'' These specifications
-may be more or less compatible with the needs of a given high-level language.
-Moreover, different protections operate under different threat models, further
-complicating their comparison and the assessment of the safeguards they
-effectively offer their clients---the programs themselves.
-It would be far preferable to have a \emph{standard}, \emph{explicit} specification.
-Only a handful of works have offered any explicit specification.

### 2023-02-02 submission5/core.tex
work on 3 and a few small things in 4

@@
 
-In this section we consider a simple setting with calls, returns, and private allocations,
-on a RISC-V machine with a PIPE tagging system. In this system a value is a pair of a
-word (the ``payload'') and a tag drawn from some set \(\mathcal{T}\).
+In this section we introduce our security property definitions by means
+of small code examples, using a simplified set of security-relevent operations for
+or calls, returns, and private allocations.
+% In this system a value is a pair of a
+%word (the ``payload'') and a tag drawn from some set \(\mathcal{T}\).
 %\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
 %for some type  of tags to be elaborated later, with \(|(w,t)| = w\).
-Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
-security classes.
-Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
-for a function {\tt main}. Code is compiled using the standard RISC-V
-calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
-{\tt main} begins at address 0 and its callee {\tt f} at address 100.
-
-{\tt main} takes a secret as its argument and initializes a variable to contain
-potentially sensitive data. We define an observable event to be a store
-to a special address, called {\tt out}.
-{\tt main} calls another function {\tt f},
-and afterward may decide to output its secret based on the contents of
-its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
-it should not---instead it will output the results of {\tt f}.
+Figure \ref{fig:main} gives C code and corresponding compiled 32-bit RISC-V code
+for a function {\tt main}, which  
+takes an argument {\tt secret} and initializes a local variable {\tt sensitive} to contain
+potentially sensitive data. 
+Then {\tt main} calls another function {\tt f},
+and afterward performs a test on {\tt sensitive} to decide whether
+to output {\tt secret}.  Since {\tt sensitive} is initialized to 0,
+the test should fail, and {\tt main} should instead output the return value of {\tt f}.
+Output is performed by writing to the special global {\tt out},
+and we assume that such writes are the only observable events in the system.
+
+The C code is compiled using the standard RISC-V calling conventions~\cite{RISC-V-CC}.
+In particular, the first function argument and the function
+return value are both passed in {\tt a0}.  
+Memory is byte-addressed and the stack grows towards
+lower addresses. We assume that {\tt main} begins at address 0 and its callee {\tt f} at address 100.
 
@@
       \labeledrow{4:}{sd ra,12(sp)}{}
-      \labeledrow{8:}{sw a1,8(sp)}{}
+      \labeledrow{8:}{sw a0,8(sp)}{}
       \labeledrow{12:}{sw zero,4(sp)}{}
-      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \mathtt{f} ~ \emplist ~ \emplist\)}
+      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \emplist \)}
       \labeledrow{20:}{sw a0,0(sp)}{}
@@
 
-\caption{Example: Main}
+\caption{Example: {\tt main}\apt{Needs machine-set diagram.}}
 \label{fig:main}
@@
 
-Suppose that {\tt f} is actually an attacker seeking
+%Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
+%security classes.  APT: Don't think we need this figure.
+
+We now consider how {\tt f} might misbehave and violate desirable
+stack safety properties associated with {\tt main}. To put the violations in a
+security framework, suppose that {\tt f} is actually an attacker seeking
 to leak {\tt secret}. It might do so in a number of ways, shown as snippets of
@@
 %
+Leakage is most obviously viewed as a violation of {\tt main}'s {\it confidentiality}.
 In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
@@
     \end{tabular}
-    \subcaption{Relative address attack}
+    \subcaption{Attacking stack pointer integrity}
     \label{subfig:WBCF2}
@@
 
-  \caption{Assembly code alternatives for {\tt f} as an attacker.
+  \caption{Example: assembly code alternatives for {\tt f} as an attacker.
   }
@@
 
-Recall\apt{can't; we've lost these} that our security semantics attach a security context to the machine state,
+Recall\apt{make sure this was said in previous section!} that our security semantics attach a security context to the machine state,
 which consists of a view \(V\) and a stack \(\sigma\) of pending activations' views.
-(See \cref{sec:machine} for details.)
 Figure \ref{fig:exec1} shows how the security context evolves over the first few
-steps of the program.  (The rules governing this evolution are formalized in \cref{fig:advops}.)
-It begins at the start of {\tt main}, where the program counter (\(\PCname\)) is zero,
+steps of the program.  (The formal details of the security semantics are described in
+\cref{sec:machine}, and the context evolution rules are formalized in \cref{fig:basicops}.)
+Execution begins at the start of {\tt main}, where the program counter (\(\PCname\)) is zero,
 and with the stack pointer (\(\SP\)) at address 1000.
@@
 
+The initial view \(V_0\) maps all stack addresses below \(\SP\) to \(\unsealed\) and the remainder of
+memory to \(\public\). The sole used argument register, {\tt a0}, is mapped to \(public\) \apt{why not \(unsealed\)?};
+other caller-save registers are mapped to \(\unsealed\) and callee-save registers to \(\sealed\). 
 Step 1 allocates a word each for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
-as two words for the return address. This will have the
-effect of marking those bytes \(\object\), assuming they were previously
-\(\unsealed\). (We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).)
+as two words for the return address. This has the
+effect of marking those bytes \(\object\). 
+(We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).)
 %
@@
       \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\pubbox} \!\! \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a1}}{\pubbox} ~ \stackrel{\mathtt{a2}}{\freebox}
-      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox}
+%      ~ \stackrel{\mathtt{s0}}{\sealbox}
       \)} \\
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a1}}{\pubbox} ~ \stackrel{\mathtt{a2}}{\freebox}
-      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox}
+%      ~ \stackrel{\mathtt{s0}}{\sealbox}
       \)}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a1}}{\freebox} ~ \stackrel{\mathtt{a2}}{\freebox}
-      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{a0}}{\freebox} ~ \stackrel{\mathtt{a1}}{\freebox}
+%      ~ \stackrel{\mathtt{s0}}{\sealbox}
       \)}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a1}}{\pubbox} ~ \stackrel{\mathtt{a2}}{\freebox}
-      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox}
+ %     ~ \stackrel{\mathtt{s0}}{\sealbox}
       \)}
@@
   \end{tabular}
-\caption{Execution up through the return from {\tt f}}
+  \caption{Execution of example up through the return from {\tt f}. In right-hand diagram, addresses increase to the right and
+  boxes represent 4-byte words.}
 \label{fig:exec1}
@@
 to the callee will not get sealed, but for now we assume no sharing of memory between activations.)
-For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments, 
-in this case just {\tt a0}. \apt{We've lost the explanation of the arguments to call, and they are used inconsistent now too, so this is utterly mysterious.}
-These are mapped to \(\public\), while any non-argument, caller-saved
-registers remain mapped to \(\unsealed\). All callee-save registers remain \(\sealed\) for all calls.
+%For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments, 
+% in this case just {\tt a0}. \apt{We've lost the explanation of the arguments to call, and they are used inconsistent now too, so this is utterly mysterious.}
+\apt{needs review}
+{\tt f} does not take any arguments; if it did, any registers containing them would be 
+mapped to \(\public\), while any non-argument, caller-saved
+registers are mapped to \(\unsealed\). All callee-save registers remain \(\sealed\) for all calls.
 At step 9, {\tt f} returns, and the topmost inactive view, that of {\tt main}, is restored.
@@
 bad behavior, including the attacks in Figure~\ref{fig:f}, are
-are detected as security violations.
+detected as security violations.
 
@@
 and the stack pointer should be the same as it was before the call.
-
-Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
+Both of these are essential. In Figure \ref{subfig:WBCF}, the attacker adds
 16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
-{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 980,
-so we define a predicate on states that should hold just after the return:
-\(\ret\ \mach \triangleq \mach[\PCname] = 20 \wedge \mach[\SP] = 980\).
-%
-We can identify the point just after the return (if a return occurs)
-as the first state in which the pending call stack is smaller than it was
-just after the call.
-WBCF requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
-%For nested calls, where the pending stack is initially larger, the same principle
-%applies: \(\ret ~ \mach\) must hold the next time the pending stack is the same size or smaller.
-
-To see why \(\wbcf\) also cares about the stack pointer, consider the attack in
-Figure \ref{subfig:WBCF2}. Here the attacker returns with \(\SP' = 998\) instead of the
+{\tt secret}.
+In Figure \ref{subfig:WBCF2}, the attacker returns with \(\SP' = 998\) instead of the
 correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame,
@@
 read part of the return address, and its attempt to output
-{\tt res} will instead output {\tt secret}!
+{\tt res} will instead output {\tt secret}.
+
+Before the call, the program counter is 16 and the stack pointer is 980.
+So we define a predicate on states that should hold just after the return:
+\(\ret\ \mach \triangleq \mach[\PCname] = 20 \wedge \mach[\SP] = 980\).
+%
+We can identify the point just after the return (if a return occurs)
+as the first state in which the pending call stack is smaller than it was
+just after the call.
+WBCF requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
+%For nested calls, where the pending stack is initially larger, the same principle
+%applies: \(\ret ~ \mach\) must hold the next time the pending stack is the same size or smaller.
+
 % Even absent other
@@
 Like WBCF, stack integrity defines a condition at the call that must hold upon
-return. This time the condition applies to all of the memory that a function has
-allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
+return. This time the condition applies to all of the memory in the function's 
+frame. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
 upon allocation, the view labels it \(\object\), and when a call is made, it instead
 becomes \(\sealed\). Intuitively, the integrity of {\tt main}
-is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
-to be identical to when it made the call.
+is preserved if, when control returns to it, any \(\sealed\) elements
+are identical to when it made the call.
 %
@@
 a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
-their values an only disagree on elements in \(\components\). We further require that
-the respective values in \(\components\) be \emph{compatible}, meaning that even though
-their payloads may differ, their tags (or whatever other data is important to
-the machine's security enforcement mechanism) must be identical.
+their values an only disagree on elements in \(\components\).
 We say that the elements of \(\components\) are \emph{irrelevant}
-in \(\mach\) if they can be replaced by compatible values without changing the
+in \(\mach\) if they can be replaced by arbitrary other values without changing the
 observable behavior of the machine. All other elements are \emph{relevant}.
-Variants and relevance are defined formally in \cref{sec:props}.
+
+(This story is slightly over-simplified. If an enforcement mechanism maintains
+additional state associated with elements, such as tags, we don't want that
+state to vary; only the payloads should. Formal definitions of variants and
+relevance that incorporate this wrinkle are given in \cref{sec:props}.)
 
@@
   \includegraphics[width=\columnwidth]{variants.png}
-  \caption{Integrity Violation\apt{Needs a machine-drawn version, as do the rest.}}
+  \caption{Integrity Violation\apt{Needs a machine-drawn version, as do the rest. Use colors consistently. No tags!}}
   \label{fig:variant}
@@
 which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
-assuming that {\tt sec} is 5. Similar to WBCF, we will define
+assuming that {\tt sec} is 5. Similar to WBCF, we define
 \(\intProp\) as a predicate on states that holds if
 all relevant sealed addresses in \(\mach\) are the same as after step 5.
-We once again require that \(\intProp\) hold on the state following the matching return,
+We require that \(\intProp\) hold on the state following the matching return,
 which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
 Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
-choosing 43 (and keeping its tag, \(t_2'\), unchanged.) As execution continues after the return
+choosing 43. As execution continues after the return
 from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
@@
 has changed the value at \(\SP - 4\), but that value matches that of the
-left-hand return. Therefore, the change does not represent a leak.
+left-hand variant at its return. Therefore, the change does not represent a leak.
 
@@
 Therefore, it must depend on a secret (in fact, the variable {\tt secret}).
-Unless {\tt a0} happens to be irrelevant, this is a violation of what
+Unless {\tt a0} happens to be irrelevant to the caller, this is a violation of what
 we term {\it return-time confidentiality} (formalized in \cref{tab:props}, line 3b).
@@
 
-In place of a predicate, we define \(\confProp\) as a binary relation on states,
+We define a binary relation \(\confProp\) on pairs of states,
 which holds on eventual return states \(\mach'\) and \(\nach'\)
 iff all relevant elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\)
-are the same in \(\mach\) and \(\nach\). We then define
-return-time confidentiality as holding if \(\confProp\) holds on the matching return states
+are the same in \(\mach\) and \(\nach\). \apt{This is a confusing formulation! Introduce ``corrupted'' here?}
+We then define return-time confidentiality as holding if \(\confProp\) holds on the matching return states
 from \(\mach\) and \(\nach\), respectively.
@@
 to cause it to misuse the key itself!
-\apt{I believe this is the first point at which we suggest that our properties might be
-  useful for enforcing (conventional) security goals --- which are definitely \emph{not}
-  part of the usual targets for ``stack safety'' mechanisms. This needs a lot more foreshadowing!}
 
@@
 
+\apt{We may need a new section that describes how a tag-based system would enforce the properties
+  described so far.}
+
 \section{Machine and Security Semantics, Formally}
@@
 
-\begin{figure}
-  \begin{tabular}{| l | l | l |}
-    \hline
-    Set / & Names & Purpose \\
-    Class & & \\
-    \hline
-    \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
-    \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
-    & {\tt a2} -- {\tt a7} & Caller-saved args \\
-    & {\tt ra} & Return Address \\
-    \hline
-    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
-    \(\sealed\) & {\tt sp} & Stack Pointer \\
-    \hline
-    \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
-    \(\public\) & {\tt tp} & Thread Pointer \\
-    \hline
-  \end{tabular}
-  \caption{RISC-V integer register set}
-  \label{fig:RISCVregs}
-\end{figure}
+% APT: I don't think we need this.
+%% \begin{figure}
+%%   \begin{tabular}{| l | l | l |}
+%%     \hline
+%%     Set / & Names & Purpose \\
+%%     Class & & \\
+%%     \hline
+%%     \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
+%%     \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
+%%     & {\tt a2} -- {\tt a7} & Caller-saved args \\
+%%     & {\tt ra} & Return Address \\
+%%     \hline
+%%     \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
+%%     \(\sealed\) & {\tt sp} & Stack Pointer \\
+%%     \hline
+%%     \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
+%%     \(\public\) & {\tt tp} & Thread Pointer \\
+%%     \hline
+%%   \end{tabular}
+%%   \caption{RISC-V integer register set}
+%%   \label{fig:RISCVregs}
+%% \end{figure}
 
@@
 \begin{figure}
+    \[\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach \triangleq
+    \{\mach[\reg]+i | \mathit{off} \leq i < \mathit{off+sz}\}\]
+
   \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
@@
              {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
-             {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+             {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+    %
+    \judgmentbr{\(b = \mach[\SP] + \mathit{off}\)}
+               {\(V' = V \llbracket \addr \mapsto \unsealed |
+                 b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
     %
@@
                  \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \rrbracket\)}
+               {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}})
                  ~ (V,\sigma) = (V'',V::\sigma)\)}
@@
     \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
+             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma')\)}
   \caption{Basic Operations}
@@
 
-For a given system, we assume an compatibility equivalence relation \(\sim\) on values.
+For a given system, we assume a \emph{compatibility} equivalence relation \(\sim\) on values.
+Values are compatible if their tags (or whatever other data is important to
+the machine's security enforcement mechanism) are identical,
+even though their payloads may differ.
 Two states are variants with respect to a set of elements \(\components\)
@@
     & \(\wbcf \triangleq\) & \(|\sigma| \uparrow \ret\)
-    & \(\textnormal{ where } \ret ~ \mach' \textnormal{ iff }\)
+    & \(\textnormal{ where } \ret ~ \mach' \triangleq \)
     \(\mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\)
@@
     & \(\clri \triangleq\) & \((|\sigma| \uparrow \intProp) ~ (\mach,(V,\sigma))\)
-    & \(\textnormal{ where } \intProp ~ \mach' \textnormal{ iff }
+    & \(\textnormal{ where } \intProp ~ \mach' \triangleq 
     \mach' \parallel \sealed(V) \cap \Delta(\mach,\mach')\)
@@
     3b & & \(\textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))\)
-    & \(\textnormal{ where } (\mach' ~ \confProp ~ \nach') \textnormal{ iff }
+    & \(\textnormal{ where } (\mach' ~ \confProp ~ \nach') \triangleq 
     \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach')\) & \\
@@
     & \(\clec \triangleq\) & \((|\sigma| \uparrow \cconfProp) ~ (\mach,(V,\sigma))\)
-    & \(\textnormal{ where } \cconfProp ~ \mach' \textnormal{ iff }
+    & \(\textnormal{ where } \cconfProp ~ \mach' \triangleq
     \mach' \parallel \Delta(\mach,\mach') - \components\)
@@
     5b & & \(\textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))\)
-    & \(\textnormal{ where } (\mach' ~ \cintProp ~ \nach') \textnormal{ iff }
+    & \(\textnormal{ where } (\mach' ~ \cintProp ~ \nach') \triangleq
     \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach')\) & \\
   \end{tabular}
-  \caption{Properties}
+  \caption{Properties. \apt{Something wrong with last one. $\cintProp$ vs. $\confProp$? }}
   \label{tab:props}
@@
 The system we've modeled so far has been very simple, but our properties
-support an important dimension of generality: adding features. We will now
+support an important dimension of generality: adding features. We now
 briefly introduce the features of our primary testing machine.
@@
 triples of a register \(\reg\), a base offset \(\mathit{off}\), and a size \(sz\), denoting
-that the object at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
+that the value at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
 The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
@@
     \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
+             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma')\)}
 
@@
   \end{subfigure}
-  \caption{Operations}
+  \caption{Operations supporting tailcalls and argument passing on stack.}
   \label{fig:advops}

### 2023-02-02 submission5/core.tex
Reworked ideas, fixed various comments

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 
-Recall\apt{make sure this was said in previous section!} that our security semantics attach a security context to the machine state,
+Our security semantics attach a security context to the machine state,
 which consists of a view \(V\) and a stack \(\sigma\) of pending activations' views.
@@
 The initial view \(V_0\) maps all stack addresses below \(\SP\) to \(\unsealed\) and the remainder of
-memory to \(\public\). The sole used argument register, {\tt a0}, is mapped to \(public\) \apt{why not \(unsealed\)?};
+memory to \(\public\). The sole used argument register, {\tt a0}, is mapped to \(\object\);
 other caller-save registers are mapped to \(\unsealed\) and callee-save registers to \(\sealed\). 
@@
 and the stack pointer target is the stack pointer at the moment of call.
-The callee's view is updated from the caller's such that all \(\object\) locations
-become \(\sealed\). (In more complex settings, objects that are intentionally passed
-to the callee will not get sealed, but for now we assume no sharing of memory between activations.)
-%For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments, 
-% in this case just {\tt a0}. \apt{We've lost the explanation of the arguments to call, and they are used inconsistent now too, so this is utterly mysterious.}
+The callee's view is updated from the caller's such that all \(\object\) memory locations
+become \(\sealed\). (For now we assume no sharing of memory between activations; data is
+passed only through argument registers, which remain active. In the presence of memory
+sharing, some memory would remain active, too.)
+
 \apt{needs review}
 {\tt f} does not take any arguments; if it did, any registers containing them would be 
-mapped to \(\public\), while any non-argument, caller-saved
+mapped to \(\object\), while any non-argument, caller-saved
 registers are mapped to \(\unsealed\). All callee-save registers remain \(\sealed\) for all calls.
@@
 
-To characterize this situation in a property, we borrow the idea of \emph{variant} states
-from the theory of non-interference. For a set of elements \(\components\),
+For a set of elements \(\components\),
 a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
@@
 which holds on eventual return states \(\mach'\) and \(\nach'\)
-iff all relevant elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\)
-are the same in \(\mach\) and \(\nach\). \apt{This is a confusing formulation! Introduce ``corrupted'' here?}
-We then define return-time confidentiality as holding if \(\confProp\) holds on the matching return states
-from \(\mach\) and \(\nach\), respectively.
+if all relevant elements are {\em uncorrupted} relative to \(\mach\) and \(\nach\).
+An element is {\em corrupted} if it differs between \(\mach'\) and \(\nach'\),
+and it either changed between \(\mach\) and \(\mach'\) or between \(\nach\) and \(\nach'\).
 
@@
 
-The properties themselves correspond to their caller equivalents, inverted. \apt{Need at least their
-  names, with references to formalizations in the table. And for the discussion in the next paragraph
-  to make sense, we will need more than that. Especially because the formalization section now has
-nothing to say!}
-
-\paragraph*{The Hierarchy of Stack Safety}
-
-\apt{This discussion should be merged with others on same topic, and maybe moved.}
-The forms of stack safety described above can potentially be enforced separately
-in various combinations. This is important, because
-many existing enforcement mechanisms enforce only some of the stack safety
-properties. Stricter enforcement may have greater performance impacts, leading to decisions
-about trade-offs between different degrees of stack safety.
-
- \(\clec\) subsumes \(\clri\), and \(\clei\) subsumes
-\(\clrc\), but integrity and confidentiality are in general orthogonal to each other
-and to \(\wbcf\). Despite this, many enforcement techniques focus purely on
-well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
-address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
-to enforce it completely. Others combine this protection with some degree of memory protection,
-chiefly focusing on integrity.
-
-\apt{We may need a new section that describes how a tag-based system would enforce the properties
-  described so far.}
+Applying this intuition using our framework, \(\clec\) turns out to resemble
+\(\clri\), extended to every element that is not marked \(\object\) at call-time. The callee's
+internal behavior is represented by those elements that change over the course of its
+execution, and which are not part of the interface with the caller. At return, those
+elements should become irrelevant.
+
+Similarly, in \(\clei\), only elements marked \(\object\) at the call should influence the
+behavior of the callee. This is the same principle as \(\clrc\), extended from \(\sealed\)
+elements to all non-\(\object\) ones.
 
@@
 to create a {\it combined state} \(s = (\mach,\context)\) and a lift the transition
-transition \(\stepstounder{}\) on combined states, labeled with the same operations.
+to \(\stepstounder{}\) on combined states, labeled with the same operations.
 with each step, the context updates based on the function
@@
 
-\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
+\judgmenttwo{\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \)}
             {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
-            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context', \obs)\)}
+            {\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\)}
 
@@
     \rowcolor{black!20}
-    5b & & \(\textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))\)
+    5b & & \(\textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \cintProp) ~ (\nach,(V,\sigma))\)
     & \(\textnormal{ where } (\mach' ~ \cintProp ~ \nach') \triangleq
@@
   \end{tabular}
-  \caption{Properties. \apt{Something wrong with last one. $\cintProp$ vs. $\confProp$? }}
+  \caption{Properties}
   \label{tab:props}
@@
 
+\section{Enforcement}
+
+\paragraph{Depth Isolation}
+
+\paragraph{Lazy Tagging and Clearing}
+
+\subsection{Enforcing \(\clei\) and \(\clec\)}
+
 \section{Validation through Random Testing}
@@
 
-[TODO: this is still last submission's test. Kept here to get a sense of length.]
-
-There are several ways to evaluate whether an enforcement mechanism enforces
+There are several ways to evaluate whether an enforcement mechanism enforces the above
 stack safety properties. Ideally such validation would be done through formal proof over
@@
 it would be considerable work to carry out for a full ISA like RISC-V and
-complex enforcement mechanism like the Depth Isolation micro-policy.
-We therefore choose to validate the micro-policy of \cref{sec:enforcement} by
-systematically \emph{testing} that it satisfies our properties.
-This focus on testing is better aligned with our immediate
-goal of evaluating real enforcement mechanisms for real machine architectures.
-Formal proof remains future work.
-
-We validate the micro-policy of the previous section by
-systematically testing that it satisfies eager stack integrity and
-confidentiality. We use a Coq specification of the RISC-V
-architecture~\cite{Bourgeat2021AMF},
+complex enforcement mechanisms like Roessler and DeHon's micro-policies.
+We therefore choose to systematically \emph{test} their micro-policies.
+
+We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
 and extend it with a runtime monitor implementing a stack safety
-micro-policy. We chose the Coq proof assistant as the setting for our implementation
+micro-policy. We choose the Coq proof assistant as the setting for our implementation
 to ensure that our coinductive trace definitions are well-formed, to reason about
@@
 To use QuickChick, we build random test-case generators that produce
-(1) an initial RISC-V machine state, including most notably
-  the program to be executed;
+(1) an initial RISC-V machine state, including the program to be executed;
 (2)
-  an initial policy state, tagging instructions corresponding to
-  blessed call or return sequences appropriately, while marking
-  all potential stack locations as $\tagNoDepth$; and
+  an initial policy state, tagging instructions and the stack regions appropriately
+  for the micro-policy (see \cref{subsec:policies})
 (3)
@@
 introduced {\em generation by execution} to produce progams that lead
-machines towards interesting behaviors. We extend the technique by tracking
-control flow information so that programs will usually---but not always---return to their call sites.
-
-%\paragraph*{Generators}
-%
-% Generation by
-%execution receives as an input a partially instantiated machine state
-%and attempts to generate an instruction (or a sequence of instructions
-%such as a blessed sequence) that makes sense locally (e.g., jumps go
-%to a potentially valid code location and loads read from a
-%potentially valid stack location). Then we step the machine and repeat
-%the process until we generate or execute some target number of
-%instructions, or reach a point where the machine cannot step
-%any more.
-%
-%We extend this technique to keep track of the control flow behavior of
-%the program being generated: each time a call or return sequence is
-%generated, we ensure that the appropriate policy tags and code
-%annotations are set for the entry or return points.% At the same time,
-%%we allow the generation to sometimes relax constraints,
-%%introducing potentially ill-formed flows: this causes our programs
-%%to failstop when executed in conjunction with the policy monitor,
-%%but also allows for revealing errors in our setup.\apt{??}
-%
+machines towards interesting behaviors. We extend the technique with a degree
+of statefulness, to nudge the generator toward executions that will carry
+on without early failstops and thus have a better chance of reaching a
+state that presents opportunities for an attack.
+
 We need to extend typical testing schemes further to handle the nested
@@
 and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
-We test the former against stepwise integrity and confidentiality, and
-the latter against their observational counterparts.
+
+\subsection{Mutation Testing}
 
@@
 
+\subsection{Results}
+
 The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
@@
 The smallest counterexample requires calling a function {\tt f},
-which fails to initialize some of its frame during the blessed sequence,
+which fails to initialize some of its frame,
 but writes into an uninitialized location $l$ later, treating \(l\) as outside
@@
 
-As for lazy micro-policies and observational properties,
-the original {\em Lazy Tagging and Clearing} micro-policy, implemented as {\em PER\_DEPTH\_TAG},
+As for \(LTC\), the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
 fails in testing, in cases where data is leaked between sequential calls.
@@
 not really stack safe, and to validate a possible fix.
-Unfortunately, Lazy Per-activation Tagging and Clearing
-puts severe pressure on the tag cache, so it may not be usable in practice.
-However, we conjecture that no lazy scheme can be fully stack safe
-without using unique per-activation identifiers.
+
+\section{Related Work}
+
+\sna{Not really sure what goes here anymore.
+  We've discussed Cerise and Nick and Andre quite a bit throughout.}
 

### 2023-02-02 submission5/future_work.tex
Reworked ideas, fixed various comments

@@
 
-In \cref{sec:intro}, we discussed the four dimensions of extensibility
-that our properties should exhibit: architectures, enforcement mechanisms,
-features, and requirements. We have focused on extending our basic system
-with advanced features; a logical next step is extending to other enforcement
-mechanisms. In the following, we discuss how different existing mechanisms
+In \cref{sec:intro}, we discussed several dimensions of flexibility
+that our properties should exhibit: applicability to different system
+features and different security requirements. Another important dimension
+is applicability to multiple enforcement mechanisms.
+In the following, we discuss how different existing mechanisms
 might fit into our model.
-
-\subsubsection{Software Techniques}
 
@@
 %
-These are hardening techniques that increase the difficulty
+Interestingly, these are lazy enforcement mechanisms, in that
+the attack may occur and be detected some time later, as long as
+it is detected before it can become dangerous. That would make our
+formalism a good fit for defining their security, except that
+they are chiefly hardening techniques: they increase the difficulty
 of some control-flow attacks on the stack, but cannot provide absolute
 guarantees on \(\wbcf\) under a normal attacker model.
-Interestingly, these are also lazy enforcement mechanisms, in that
-the attack may occur and be detected some time later, before taking the
-
-\paragraph{Bounds Checking}
-%
-Under a bounds checking discipline, all the pointers in a program are extended
-with some disjoint metadata, and these are combined to gate memory accesses,
-usually with the objective of enforcing some form of \emph{memory
-safety} \rb{(relate this to the basic high-level concepts and types of attacks
-we want to consider: control flow, some integrity and confidentiality?)}.
-%
-Compared to stack canaries and shadow stacks, bounds checking is eminently
-a \emph{high-level} technique, designed with a particular programming language
-in mind and implemented through a \emph{secure compiler} \cite{} that performs
-the necessary code transformations. Also unlike those techniques (which make few
-assumptions on callee code), it offers little to no protection
-against \emph{adversarial low-level} code that is not constrained to obey the
-bounds checking discipline.
 
@@
 protect exclusively the control-relevant parts of the stack, making no claims
-about its data-relevant parts.
+about its data-relevant parts---we would solely be testing \(\wbcf\).
 
-\subsubsection{Hardware Techniques}
+\paragraph{Bounds Checking}
+%
+Under a bounds checking discipline such as Softbound \cite{}, all the pointers
+in a program are extended with some disjoint metadata, and these are combined
+to gate memory accesses. These approaches enforce a form of \emph{memory safety},
+and we would therefore expect them to enforce \(\clri\) and \(\clrc\). They aim
+to enforce \(\wbcf\) by cutting off attacks that involve memory-safety violations,
+but that may not be sufficient.
 
-In addition to the many software-based techniques developed to provide greater
-degrees of protection to the stack, often incurring a significant performance penalty,
-some proposals have turned to specialized hardware features to offset those
-costs and attempt to offer greater levels of protection.
-
-We focus in particular on two families of hardware mechanisms: capability machines
-and tag-based reference monitors. Both types of system move beyond simply
-offering hardware acceleration for software-based techniques, offering hardware
-primitives for both memory and control-flow enforcement that can offer a degree
-of protection even with minimal cooperation from the code.
-%
-\rb{Something about how realistic/practical these alternatives are? Are we
-  selecting partly based on prototypes, etc.?}
+Bounds checking approaches require substantial compiler cooperation. This is not a
+problem for our properties in general, but it is not very compatible with
+generation-by-execution of low-level code. A better choice would be to generate
+high-level code using a tool like CSmith \cite{}, or prove the properties instead.
 
 \paragraph{Capabilities}
-%
-\rb{(There is plenty of variety here, should we focus on the Cerise-and-friends
-  (CHERI-adjacent) work exclusively?)}
 
-\sna{If we have good examples of other capability systems, we should absolutely
-  talk about them. No need to privilege CHERI.}
-%
 Capability machines can represent and manipulate
@@
 protections of bounds-checking with software-based methods of control-flow protection.
-%
-They are closely related \rb{too strong?} to this work; we discuss them in more
-detail in \cref{XXX}.
-%
 
-\paragraph{Micro-Policies}
-%
-\rb{(This partly overlaps with the technical preliminaries, try to offer a quick
-  synthesis here.)}
-%
-Micro-policy machines are a family of tagged architectures that decorate
-addressable memory and registers with arbitrary metadata tags, and extend the
-processor with a programmable hardware monitor that uses custom rules on
-metadata tags to implement many different kinds of security policies.
-%
-This flexibility can be harnessed to implement a wide array of stack protection policies with relative ease, covering and going beyond the capabilities \rb{word choice?} of the various software-based techniques \cite{} \rb{as demonstrated by}.
-%
-These policies comprise very general forms of \emph{lazy enforcement}, which can tolerate
-potential but harmless violations \rb{the system programmer's motto: ``trust me,
-  I know what I'm doing''} while still detecting them if and when their effects
-become observable. 
-%
-Micro-policies are one of the centerpieces of our technical development; we
-present them in more detail in \cref{XXX}.
+There are several proposals around the use of Cheri capabilities to enforce stack safety,
+including mechanisms that use the standard Cheri hardware (which includes local
+capabilities) \cite{SkorstengaardLocal},
+and others that propose entirely new types of capabilities, such as linear
+\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
+\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
+It is a high priority to test this most recent work by Georges et al., as it is designed
+to be similar to all of our properties except for \(\clei\).
 
-\paragraph{Summary}
+\subsection{More Features}
 
-We begin to see that the security of the call stack and its contents is not
-clearly defined. Different protection mechanisms aim to protect the stack
-in fundamentally different ways, each offering a \emph{unique}, \emph{implicit}
-specification of the meaning of ``stack safety.'' These specifications
-may be more or less compatible with the needs of a given high-level language.
-Moreover, different protections operate under different threat models, further
-complicating their comparison and the assessment of the safeguards they
-effectively offer their clients---the programs themselves.
-It would be far preferable to have a \emph{standard}, \emph{explicit} specification.
-Only a handful of works have offered any explicit specification.
+Several common features did not make it into this work: most importantly concurrency and
+exceptions. We plan to extend the model to include them, and then test them. We would also
+like to test the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.

### 2023-02-02 submission5/ideas-v1.tex
Reworked ideas, fixed various comments

--- 2023-01-28 submission5/ideas-v1.tex
@@
 call in a simple example machine and discuss how each of our properties applies to it,
-informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
+informally. In the process we discuss some motivation for the properties, from a security
+perspective: we consider a simple attacker, and discuss the harm that it can do to
+its caller. This perspective is essential to understanding how lazy enforcement can
+be safe: even if an attacker can overwrite some memory, if it cannot accomplish
+its bigger-picture goals, the system is safe.
+
+Then in \cref{sec:machine} we will introduce the machine model that we test, and give
 its {\em security semantics}, described in brief below. We formalize the definitions of
@@
 but the theory does not depend on this choice.
+
+Whatever the mechanism for labeling transitions, we get a base
+machine transition \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
+an event (see below) and \(\bar{\psi}\) is a list of security-relevant operations.
+We then lift this into a transition between pairs of machine states and contexts,
+\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
 
@@
 
-\paragraph*{Threat Model}
+\paragraph*{Variants and Non-interference}
 
-We must trust that our method of distinguishing security-relevant operations is accurate; if it
-involves labels placed on code by a compiler, that means trusting that the compiler placed
-those labels correctly. If operations occur that are not recognized, those operations
-might not be guaranteed to protect their principals---for instance, an unlabeled call
-might not protect the caller's data from the callee. On the flip side, applying an incorrect
-label most likely means that the property becomes too strong to be enforced.
+To characterize confidentiality, we borrow the idea of \emph{variant} states
+from the theory of non-interference. Non-interference is a way of expressing
+the flow of information through the system. If we begin with a state \(\mach\)
+and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
+then execute both \(\mach\) and \(\nach\), any differences in the outputs they
+produce must be the result of accessing \(\component\).
 
-Otherwise, we do not assume that the code adheres to any particular
-calling convention or implements a source language construct.
-In particular, while we are agnostic as to the source
-language, we certainly aim to support C, and so any source function might contain undefined
-behavior resulting in its compilation to arbitrary machine code. A given enforcement
-mechanism may place additional constraints, particularly on the behavior of
-call and return sequences. For instance, extant implementations tend to assume
-implicitly that callee-saved registers are treated appropriately by whichever compiler
-generated their code. Our properties explicitly require them to be treated as belonging
-to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
+\paragraph*{Putting It Together}
 
-In general, it is impossible to distinguish buggy machine code from an attacker; in
-our examples we will identify one function or another as an attacker, but we do not
-require any static division between trusted and untrusted code, and we aim to protect
-even buggy code.
-
-This is a strong threat model, but it does omit some important aspects of
-security enforcement: In particular, hardware and timing attacks are out of scope.
+Putting these ideas together, we get the basic structure of a property in our model.
+First we define the security-relevant operations that we support, a security context,
+and the function according to which the security context updates with each step.
+Then, we define a predicate on states and contexts that expresses the requirements
+for our security property, using variants, assertions about future states, and comparison
+of event traces. Finally, in order to apply the property to a specific machine, we
+must determine how the machine's execution translates into security relevant operations.

### 2023-02-02 submission5/intro-v2.tex
Reworked ideas, fixed various comments

@@
 Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
-as they aim to impede attackers but cannot provide universal protection. Stack canaries,
-for instance, aim to reduce the ability of attackers to find gadgets, not to categorically
-rule out attacks. Others may be amenable to formalization but do not offer it.
+as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
+for instance, aim to ``restrict the flexibility available in creating gadget chains''
+\cite{Shanbhogue+19} not to categorically
+rule out attacks. Other mechanisms are stronger, but relatively few formally state their guarantees.
 The Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
@@
 fulfill a tag-based adaptation Georges et al.'s state-of-the-art property.
-\rb{(Do we explain why anywhere?)}
-Roessler and DeHon themselves do not claim that it intends to enforce stack safety,
-only an approximation.
 
-We disagree. \(LTC\) does not always protect individual addresses on the stack, but it aims
+We believe that, nevertheless, \(LTC\) enforces stack safety.
+It does not always protect individual addresses on the stack, but it aims
 to prevent one function activation from interfering with another's behavior---to protect
@@
 
-%Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
-%other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
-%or tested. Such properties become the specification against which enforcement can be validated;
-%even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
-%\emph{why} they fail.
+Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
+other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
+or tested. Such properties become the specification against which enforcement can be validated;
+even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
+\emph{why} they fail.
 
+We take the view that security properties should be motivated by security requirements,
+and flexible enough to support various sets of security requirements depending on the
+goals of the system that they validate.
+Following this reasoning, our vision of stack safety is as a collection of properties
+describing the security needs of a system with regard to its functional abstractions.
+While the stack still plays a key role, our stack safety is about functions. The most
+dramatic consequence of this is that we treat registers as part of a function activation's
+state, and hold them to the same protection requirements as stack memory.
+
+Flexibility is especially important for stack safety.
+%
 Most security properties are described at the level of a high-level programming language,
@@
 amenable to a very general high-level approach.)
+%
 But it is not satisfying as a low-level property; at the level of the ABI, the
-specification of a well-behaved stack is minimal. It is not concerned with such
+specification of a well-behaved stack is minimal. The ABI is not concerned with such
 questions as whether a caller's frame should be readable or writable to its callee,
 leaving those concerns to be determined by a high-level language built atop the stack.
+These are high-level security questions.
 
@@
 
-These two cases represent two dimensions of generality that make stack safety difficult to
-define as a single, universal property: what constitutes safe behavior may differ between
-systems that support different features, and different types of security may be more or less
-desirable depending on context.
+These two cases represent two dimensions in which a stack safety definition requires
+flexibility: what constitutes safe behavior differs between systems that support different
+features, and different contexts have different top-level security requirements.
 
-We are not attempting to present a universal definition of stack safety for all systems.
-Rather, we define the stack safety of a single, fairly realistic system (one with common
-features such as memory arguments and tailcall elimination), decomposed into
+We are not attempting to present a universal definition of stack safety for all systems,
+but we do construct our definition with these dimensions of flexibility in mind.
+Our model is decomposed into:
 (1) a lightweight model of the system and its security-relevant features, from which we
@@
 return, (2) the application of that model to a particular target system, and (3)
-the formal criteria for integrity, confidentiality, and control flow
-to obey the requirements. This decomposition can be thought of as a step toward
-a universal definition, in that the definitions in (3) capture fundamental
-notions of security that can then be applied flexibly to other systems by varying
-the model in (1) and (2).
+the formal criteria for the integrity, confidentiality, and control flow
+requirements.
 
-To address the second class of generality, we present a total of five core properties:
+We show that we can extend our system to new features by first introducing a simple
+example model, and then moving to a more realistic model with multiple new features:
+callee-save registers, argument passing on the stack, and tailcall elimination, simply
+by refining a few definitions in (1). This becomes the model that we validate.
+In \cref{???}, we extend the model further with arbitrary sharing of memory-safe
+pointers. Our criteria in (3) cover a total of five core requirements:
 \(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)). These are defined
-formally in \cref{sec:props}. They are related to one another as seen in
+callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
+They are related to one another as seen in
 \cref{fig:propertydiagram}: \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
-\(\clri\), but otherwise the properties are orthogonal. We further define two
-important combinations of properties: full stack safety \(\textnormal{\sc FSS}\),
-and caller safety \(\textnormal{\sc ClrS}\). Of these, most existing concepts of
-stack safety correspond to either \(\wbcf\) or \(\textnormal{\sc ClrS}\), which are
-concerned with the basic guarantees that the caller recieves during the call, while
-\(\textnormal{\sc FSS}\) is a natural extension that enforces a higher-level concept
-of a correct interface between caller and callee. \sna{Blah. Needs work.}
+\(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
+according to the needs of a system.
 
-To validate our formal characterization, we use these
-properties to test and improve an existing enforcement mechanism, the
-{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
-in the Coq proof assistant on top of a \rb{formal} RISC-V specification.  We
-use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check
-that Roessler and DeHon's micro-policies correctly detect the ones that
-attempt to violate one of our properties. Furthermore, we
+In particular, many enforcement techniques focus purely on
+well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
+address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
+to enforce it completely. Others combine this protection with some degree of memory protection,
+chiefly focusing on integrity. Interestingly, Skorstengaard et al. \cite{SkorstengaardSTKJFP}
+describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
+to the combination of \(\clri\) and \(\clrc\). It is not until Georges et al. \cite{Georges22:TempsDesCerises}
+that the Cerise line of work considers confidentiality; they also consider \(\clec\), but not
+\(\clei\). \sna{It is really tempting to find a way to take credit for this.}
+
+%These are defined
+%formally in \cref{sec:props}. . We further define two
+%important combinations of properties: full stack safety \(\textnormal{\sc FSS}\),
+%and caller safety \(\textnormal{\sc ClrS}\). Of these, most existing concepts of
+%stack safety correspond to either \(\wbcf\) or \(\textnormal{\sc ClrS}\), which are
+%concerned with the basic guarantees that the caller recieves during the call, while
+%\(\textnormal{\sc FSS}\) is a natural extension that enforces a higher-level concept
+%of a correct interface between caller and callee. \sna{Blah. Needs work.}
+
+Flexibility aside, we show that our properties are useful for their purpose:
+distinguishing correct enforcement from incorrect. We use
+QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for Coq, to generate random programs and check that the
+stack safety {\em micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
+correctly detect the ones that attempt to violate one of our properties. Furthermore, we
 check that the testing framework is able to generate counterexamples
 that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms, including ones that we accidentally created
-while implementating the micro-policy and ones that we
-intentionally crafted to increase our confidence in the effectiveness
-of our testing.
-%
-\rb{This and the following paragraph could start more generally (we need to
-  answer the question of why these experiments, what is their interest), and
-  probably motivate the use of PBT (experiment results below show there's
-  benefit in them!) and, somewhere at least, talk about other points in the
-  spectrum.}
-
-We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
-violates the temporal aspect of confidentiality in
-cases where data can leak across repeated calls to the same callee,
-and also violates integrity if the leak uses the caller's frame. We
-propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
-confidentiality, albeit at some performance cost.
-%
-\rb{Anything to add from personal communications? More details later? Shows
-  these things are very hard to get right -- like the latest work on Cerise,
-  also motivated by this very work.
-\rb{(+SNA) At the very least, these models are subtle and getting them right,
-  and describing them unambiguously, is hard without a framework like this.}
-}
+variants of the enforcement mechanisms---including the original \(LTC\)!
 
@@
 \begin{itemize}
-\item We formalize permissive stack confidentiality and integrity properties,
-  which are parameterized over a notion of external
-  observation, and are violated only if accessing secrets or overwriting
-  data causes a visible change in system behavior.
-\item We instantiate these definitions with a realistic
-  system with features such as argument passing
+\item A novel characterization of stack safety as a collection of properties:
+  confidentiality and integrity for callee and caller, and well-bracketed
+  control-flow. These can characterize the security requirements of systems
+  that support only some portion of stack safety, and are
+  parameterized over a notion of external observation, so that they can characterize
+  lazy enforcement.
+\item An instantiation of these definitions in a realistic
+  setting with features such as argument passing
   on the stack, callee-saves registers, and tail-call elimination.
   Our model is modular enough that adding these features is straightforward.
-\item Our formalization of well-bracketed control flow in the presence of tail-call
-  elimination \ifexceptions and exceptions \fi is a novelty over existing models.
-\item We use property-based random testing to validate the relationship between
-  our properties and micro-policies.
+\item The validation, via property-based randomize testing, of tag-based
+  stack-safety policies. We find that \emph{Lazy Tagging and Clearing}
+  falls short of stack safety, and propose a fix.
 \end{itemize}
@@
 QuickChick library and to ensure that the formalizations are unambiguous.
+
+\subsection{Threat Model and Limitations}
+
+When our properties are used to evaluate a system, the threat model will depend on the
+details of that system. However, there are some constraints that our design puts on
+any system. We must trust that the method of distinguishing security-relevant operations is accurate; if it
+involves labels placed on code by a compiler, that means trusting that the compiler placed
+those labels correctly. If operations occur that are not recognized, those operations
+might not be guaranteed to protect their principals---for instance, an unlabeled call
+might not protect the caller's data from the callee. On the flip side, applying an incorrect
+label most likely means that the property becomes too strong to be enforced.
+
+Otherwise, we do not assume that the code adheres to any particular
+calling convention or implements a source language construct.
+In particular, while we are agnostic as to the source
+language, we certainly aim to support C, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code. A given enforcement
+mechanism may place additional constraints, particularly on the behavior of
+call and return sequences. For instance, extant implementations tend to assume
+implicitly that callee-saved registers are treated appropriately by whichever compiler
+generated their code. Our properties explicitly require them to be treated as belonging
+to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
+
+In general, it is impossible to distinguish buggy machine code from an attacker; in
+our examples we will identify one function or another as an attacker, but we do not
+require any static division between trusted and untrusted code, and we aim to protect
+even buggy code.
+
+This is a strong threat model, but it does omit some important aspects of
+security enforcement: in particular, hardware and timing attacks are out of scope, and
+we do not address concurrency.

### 2023-02-02 submission5/intro-v2.tex
Quick pass on I+II

@@
 for instance, aim to ``restrict the flexibility available in creating gadget chains''
-\cite{Shanbhogue+19} not to categorically
+\cite{Shanbhogue+19}, not to categorically
 rule out attacks. Other mechanisms are stronger, but relatively few formally state their guarantees.
@@
 \(\clei\). \sna{It is really tempting to find a way to take credit for this.}
+%
+\rb{This is the direct support I can find: ``Following Anderson et al. [2021],
+we must distinguish between {\em local state integrity} and {\em local state
+confidentiality}.'' + similar discussion in their related work + their novel
+claim of adding support for confidentiality properties, thus ``following'' us
+\ldots + personal communication?}
 
@@
 QuickChick library and to ensure that the formalizations are unambiguous.
+%
+\rb{Mention proofs in future work?}
 
@@
 to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
+%
+\rb{This last bit about callee-saved registers belonging to the caller is subtle, I think it risks the reader missing the point.}
 
@@
 we do not address concurrency.
+%
+\rb{Anything about concurrency in future work, though?}

### 2023-02-02 submission5/core.tex
Add missing references

@@
 
-One such provenance model is Memarian et al.'s PVI \cite{}, in which pointers are
+One such provenance model is Memarian et al.'s PVI \cite{provenance}, in which pointers are
 annotated with the identity of the object they first pointed to. This annotation is

### 2023-02-02 submission5/future_work.tex
Add missing references

@@
 %
-Under a bounds checking discipline such as Softbound \cite{}, all the pointers
+Under a bounds checking discipline such as SoftBound \cite{NagarakatteZMZ09}, all the pointers
 in a program are extended with some disjoint metadata, and these are combined
@@
 generation-by-execution of low-level code. A better choice would be to generate
-high-level code using a tool like CSmith \cite{}, or prove the properties instead.
+high-level code using a tool like CSmith \cite{DBLP:conf/pldi/YangCER11}, or prove the properties instead.
 
@@
 The most prominent contemporary representative of this long line of work is
-CHERI \cite{}, a modern architecture designed to provide efficient fine-grained
+CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
 memory protection and compartmentalization.
@@
 \emph{secure calling conventions} for the call
-stack \cite{}. Those conventions can be seen as combining the
+stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
 protections of bounds-checking with software-based methods of control-flow protection.
 
-There are several proposals around the use of Cheri capabilities to enforce stack safety,
-including mechanisms that use the standard Cheri hardware (which includes local
+There are several proposals around the use of CHERI capabilities to enforce stack safety,
+including mechanisms that use the standard CHERI hardware (which includes local
 capabilities) \cite{SkorstengaardLocal},

### 2023-02-02 submission5/core.tex
section 3: 64-bit RISC-V

@@
 %for some type  of tags to be elaborated later, with \(|(w,t)| = w\).
-Figure \ref{fig:main} gives C code and corresponding compiled 32-bit RISC-V code
+Figure \ref{fig:main} gives C code and corresponding compiled 64-bit RISC-V code
 for a function {\tt main}, which  
@@
       \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\pubbox} \!\! \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
 %      ~ \stackrel{\mathtt{s0}}{\sealbox}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
 %      ~ \stackrel{\mathtt{s0}}{\sealbox}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\freebox} ~ \stackrel{\mathtt{a1}}{\freebox}
+      ~ \stackrel{\mathtt{a0}}{\freebox\freebox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
 %      ~ \stackrel{\mathtt{s0}}{\sealbox}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
  %     ~ \stackrel{\mathtt{s0}}{\sealbox}

### 2023-02-02 submission5/core.tex
Nits

@@
 %for some type  of tags to be elaborated later, with \(|(w,t)| = w\).
-Figure \ref{fig:main} gives C code and corresponding compiled 64-bit RISC-V code
+Figure \ref{fig:main} gives C code and corresponding compiled 32-bit RISC-V code
 for a function {\tt main}, which
@@
 
+\bcp{Bits missing here!}
+
 \paragraph{Depth Isolation}

### 2023-02-02 submission5/core.tex
add callee-save description to example

--- 2023-02-02 submission5/core.tex
@@
       ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
-%      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
       \)} \\
@@
       ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
-%      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
       \)}
@@
       ~ \stackrel{\mathtt{a0}}{\freebox\freebox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
-%      ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
       \)}
@@
       ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
- %     ~ \stackrel{\mathtt{s0}}{\sealbox}
+      ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
       \)}
@@
 sharing, some memory would remain active, too.)
-
-\apt{needs review}
-{\tt f} does not take any arguments; if it did, any registers containing them would be 
+Function {\tt f} does not take any arguments; if it did, any registers containing them would be 
 mapped to \(\object\), while any non-argument, caller-saved 
-registers are mapped to \(\unsealed\). All callee-save registers remain \(\sealed\) for all calls.
+registers are mapped to \(\unsealed\). In the current example, only register {\tt a0} has a change in
+security class. All callee-save registers remain \(\sealed\) for all calls.
+Thus, if we varied the assembly code for the example slightly so that {\tt sensitive} was stored
+in a callee-save register (e.g. {\tt s0}) rather than in memory, its security class would still be \(\sealed\)
+at the entry to {\tt f}.
+%
 At step 9, {\tt f} returns, and the topmost inactive view, that of {\tt main}, is restored.
 
-We now show how this security semantics can be used define notions of confidentiality,
+We now show how this security semantics can be used to define notions of confidentiality,
 integrity, and correct control flow in such a way that many classes of

### 2023-02-02 submission5/intro-v2.tex

@@
 \cite{Shanbhogue+19}, not to categorically
-rule out attacks. Other mechanisms are stronger, but relatively few formally state their guarantees.
-The Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
+rule out attacks. Other mechanisms are stronger, but formal
+guarantees are rarely discussed.
+Indeed, the Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
 Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
-to offer (various) stack safety formalizations.
+to offer (various)\bcp{Maybe we can omit that word here and spell it
+  out in the related work.} stack safety formalizations.
 

### 2023-02-02 submission5/ideas-v1.tex
Added security classes to ideas

@@
 The security principals in our account of stack safety are function
-activations: at any given time, pending activations
-have security requirements that are recorded in a call stack, and the current
-activation tracks context information that will inform its requirements when
-it makes a call. Each activation has a {\it view}
-of the system that maps each state element to a {\it security class}:
-it may be private to an inactive principal (\(\sealed\)),
-in an allocated object that is accessible to the current principal (\(\object\)),
-available to be allocated (\(\unsealed\)), or
-always accessible (\(\public\)).
+activations: the context tracks, for the active function and all pending
+functions, their {\em view} of the security state of the system.
+
+\paragraph*{Views}
+
+The security context consists of a stack of views: functions that map
+each state element to a {\it security class}. The security classes are
+\(\public\), \(\unsealed\), \(\object\), and \(\sealed\).
+
+State elements that are outside of the stack---general-purpose memory used for
+globals and the heap, as well as the code region and globally-relevant
+registers---are always labeled \(\public\). We place security requirements on some
+\(\public\) elements for purposes of \(\wbcf\), and a given enforcement mechanism
+might restrict their access (e.g., by rendering code immutable) but for integrity
+and confidentiality purposes they are considered accessible at all times.
+
+For a newly active function, every element that is available for use but uninitialized
+is seen as \(\unsealed\). From the perspective of the caller, the callee has no obligations
+regarding its use of \(\unsealed\) elements, but the callee itself
+
+Arguments are seen as \(\object\), meaning that their contents may be safely used.
+When the active function allocates memory, that memory will also be \(\object\).
+Then, on a call, \(\object\) elements that are not being used to communicate with
+the new callee will become \(\sealed\)---reserved for an inactive principal,
+and expected to be unchanged when it becomes active again.
 
@@
 and the function according to which the security context updates with each step.
-Then, we define a predicate on states and contexts that expresses the requirements
-for our security property, using variants, assertions about future states, and comparison
-of event traces. Finally, in order to apply the property to a specific machine, we
-must determine how the machine's execution translates into security relevant operations.
+The security context gives each function activation a view, mapping each element to a
+security class.
+
+Then, we define our security requirements in terms of the security classes of state
+elements. These take the form of predicates on states and contexts that use variants,
+assertions about future return states, and comparison of event traces to capture the
+desired behavior without reference to implementation details.
+
+Separately, in order to apply a property to a specific machine, we do need to consider
+implementation details, by determining how the machine's execution translates into
+security relevant operations. But this step is transparent to the properties themselves.

### 2023-02-02 submission5/intro-v2.tex
Added security classes to ideas
Words about Aina's work citing ours

@@
 describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
-to the combination of \(\clri\) and \(\clrc\). It is not until Georges et al. \cite{Georges22:TempsDesCerises}
-that the Cerise line of work considers confidentiality; they also consider \(\clec\), but not
-\(\clei\). \sna{It is really tempting to find a way to take credit for this.}
-%
-\rb{This is the direct support I can find: ``Following Anderson et al. [2021],
-we must distinguish between {\em local state integrity} and {\em local state
-confidentiality}.'' + similar discussion in their related work + their novel
-claim of adding support for confidentiality properties, thus ``following'' us
-\ldots + personal communication?}
+to the combination of \(\clri\) and \(\clrc\). Their state-of-the-art successor,
+Georges et al. \cite{Georges22:TempsDesCerises}, cite an early preprint of our work
+as inspiration for extending their formal discussion to include confidentiality.
+When checking if a preliminary version of these properties applied to
+the Cerise calling convention, they noted that the Skostengaard et al. calling convention
+did not enforce callee confidentiality, and made sure that their version would
+\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
+This demonstrates the benefit of our choice to explicitly state properties in security
+terms: specifying security mechanisms is hard, and when the specification takes the
+form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
+security requirement.
 
@@
   on the stack, callee-saves registers, and tail-call elimination.
-  Our model is modular enough that adding these features is straightforward.
+  Our model is modular enough that handling these features is straightforward.
 \item The validation, via property-based randomize testing, of tag-based

### 2023-02-02 submission5/core.tex
pass through formalism

@@
   \IEEEauthorblockN{
+    Roberto Blanco
+  }
+  \IEEEauthorblockA{
+    Max Planck Institute for Security and Privacy\\
+    roberto.blanco@mpi-sp.org\\
+  }
+  \and
+  \IEEEauthorblockN{
     Leonidas Lampropoulos
@@
     leonidas@umd.edu\\
-  }
-  \and
-  \IEEEauthorblockN{
-    Roberto Blanco
-  }
-  \IEEEauthorblockA{
-    Max Planck Institute for Security and Privacy\\
-    roberto.blanco@mpi-sp.org\\
   }
@@
 
-\section{Machine and Security Semantics, Formally}
+\section{Formalization}
 \label{sec:machine}
@@
 
+We now give a formal description of our machine model, security semantics,
+and properties. Our definitions abstract over the details of the target machine
+architecture and ABI, the set of security-relevant operations, the rules that
+govern evolution of the security context, the set of observable events, and
+a notion of value compatibility.
+\apt{Not quite true: the property definitions require a notion of
+  call entry.}
+\apt{Any more? Give a more listlike presentation?
+  Need to clarify exactly what the parameters are, both in the initial
+  presentation and in the various extensions, including the appendix.}
+
+\subsection{Machine}
 The building blocks of a machine are {\em words} and {\em registers}.
@@
 A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
-\emph{values},  with a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
+\emph{values}.
 Each value \(v\) contains a \emph{payload} word, written \(|v|\).
+We write \(\mach[\component]\) to denote the value of \(\mach\) at
+\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
 The details of value structure depend on the specific machine being modelled;
@@
 information relevant to a hardware enforcement mechanism (such as a tag).
-We write \(\mach[\component]\) to denote the value of \(\mach\) at
-\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
-
-Security classes are ranged over by \(l \in \{\public, \object, \sealed, \unsealed\}\).
-For some security class \(l\), we write \(l(V)\)
+We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
+and lift it element-wise to states.
+Values are compatible if their non-payload information is identical, 
+even though their payloads may differ.
+\apt{This is still not very convincing!
+  empting to say that only the payloads can affect the step function, but that is
+  too strong, right? (OK for tags, but not, e.g. capabilities?}
+
+The machine has a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
+Except for the annotations over the arrow, this function just encodes the usual
+ISA description of the machine's instruction set. The annotations serve to connect
+the machine's operation to our security setting: 
+\(\bar{\psi}\) is a list of security-relevant operations drawn from an assumed set \(\Psi\),
+and \(\obs\) is an (potentially silent) observable event; these are described further below.
+
+\subsection{Security semantics}
+
+The security semantics operates in parallel with the 
+Each state element (memory word or register) is given a \emph{security class} 
+\(l \in \{\public, \object, \sealed, \unsealed\}\).
+A \emph{view} \(V \in \mathit{VIEW}\) maps elements to security classes.
+For any security class \(l\), we write \(l(V)\)
 to denote the set of elements \(\component\) such that \(V ~ \component = l\).
@@
 belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
-
-A {\it context} is a pair of the current activation's view and
+\apt{This doesn't quite match example, where there is magically an initial argument register labeled \(\active\).}
+
+A (security) \emph{context} is
+a pair of the current activation's view and
 a list of views representing the call stack (pending inactive
 principals), ranged over by \(\sigma\).
-
-A view \(V\) maps elements to security classes, and a context consists of the
-view of the active function together with a stack of inactive views.
+%
+\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
+%
 The initial context is \(\context_0 = (V_0, \emplist)\).
-%
-\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
-%
-We have described informally how the security context develops as the system performs
+
+\cref{sec:example} describes informally how the security context evolves as the system performs
 security-relevant operations. Formally, we combine each machine state with a context
 to create a {\it combined state} \(s = (\mach,\context)\) and a lift the transition
-to \(\stepstounder{}\) on combined states, labeled with the same operations.
-with each step, the context updates based on the function
+to \(\stepstounder{}\) on combined states. 
+At each step, the context updates based on the function
 \(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
-Note that \(Op\) takes as its first argument the state {\it before} the step.
+(Note that \(Op\) takes as its first argument the state {\it before} the step.)
 Since a single step might correspond to multiple operations, we apply
-\(Op\) as many times as is needed.
+\(Op\) as many times as needed, using \(\mathit{foldl}\).
 
@@
 
-The definition of \(Op\) is most convenient to present decomposed into
+A definition of \(Op\) is most convenient to present decomposed into
 rules for each operation. We have already seen the intuition behind the rules for
@@
 
-We write the trace of execution from a state up to the first
-state below depth \(d\) via the \(\hookrightarrow\) operator, defined coinductively.
-When \(d = 0\), the trace will always be infinite. In this case we
-omit \(d\) and just write \(\hookrightarrow s\).
+We are particular interested in traces that end just after a function returns.
+We define these in terms of the depth \(d\) of the security context's call stack \(\sigma\).
+We write \(d \hookrightarrow s\) for the trace of execution from a state \(s\)
+up to the first point where the stack depth is smaller than \(d\), defined
+coinductively by these rules:
 
@@
 
+\noindent
+When \(d = 0\), the trace will always be infinite; in this case we
+omit \(d\) and just write \(\hookrightarrow s\).
+
 Two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
@@
 is the same. That is, we compare up to deletion of \(\tau\) events.
+\apt{Didn't we use to say something about the (arguably) odd nature of this definition?}
 
@@
 
-\section{Formal properties}
+\subsection{Variants, corrupted sets, and ``on-return'' assertions}
 \label{sec:props}
 
-Now we are ready to introduce the formal properties that define correctly enforced
-security semantics. First, we must give a few prelimaries: the generation and comparison
-of traces, variant states, and ``on-return'' assertions.
-
-For a given system, we assume a \emph{compatibility} equivalence relation \(\sim\) on values.
-Values are compatible if their tags (or whatever other data is important to
-the machine's security enforcement mechanism) are identical,
-even though their payloads may differ.
-Two states are variants with respect to a set of elements \(\components\)
-if they agree on the value of every element not in \(\components\) and
-and have compatible values for every element in \(\components\).
+Two (compatible) states are variants with respect to a set of elements \(\components\)
+if they agree on the value of every element not in \(\components\). 
 Our notion of non-interference involves comparing the traces of such
 \(\components\)-variants. We use this to define sets of irrelevant elements,
-and secrets in general.
+and secrets in general.\apt{last phrase is cryptic}
+
+\definition The \emph{difference set} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
+is the set of elements \(\component\) such that \(\mach[\component] \not = \mach'[\component]\).
 
 \definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
-written \(\mach \approx_\components \nach\), if, for
-all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
-and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
-
-\definition An element set \(\components\) in state \((\mach,\context)\) contains
-irrelevant values, written \((\mach,\context) \parallel \components\), if for all
+written \(\mach \approx_\components \nach\), if \(\mach\sim\nach\) and
+\(\Delta(\mach,\nach) \subseteq \components\). 
+%if all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
+%and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
+
+\definition An element set \(\components\) is \emph{irrelevant} to state \((\mach,\context)\),
+written \((\mach,\context) \parallel \components\), if for all
 \(\nach\) such that \(\mach \approx_{\components} \nach\),
@@
 
-\definition \(\Delta(\mach,\mach')\) is the set of elements \(\component\)
-such that \(\mach[\component] \not = \mach'[\component]\).
-
 \definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
@@
 
-Conceptually, the corrupted set treats \(\mach\) and \(\nach\) as initial states whose
-and \(\mach'\) and \(\nach'\) as their corresponding final states, such that any element
-that has changed from its initial value but does not match between the final states
-is considered to have propagated some varied data from the original pair.
+If we consider two execution sequences, one from \(\mach\) to \(\mach'\)
+and the other from \(\nach\) to \(\nach'\),
+then \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\) is the set of elements that
+change in one or both executions and end up with different values. Intuitively,
+this captures the effect of any differences between \(\mach\) and \(\nach\), i.e.,
+the set of values that are ``corrupted'' by those differences.
 
@@
                   {\((d \uparrow P) ~ (\mach', \context')\)}
-                  {\((\mach, (V,\sigma)) \stepstounder{\overline{\psi}} (\mach', \context',\obs)\)}
+                  {\((\mach, (V,\sigma)) \stepstounder{\overline{\psi},\obs} (\mach', \context')\)}
                   {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}
@@
                   {\(|\sigma_1| \geq d\)}
-                  {\((\mach_1,(V_1,\sigma_1)) \stepstounder{\overline{\psi}} (\mach_1',\context_1',\obs)\)}
+                  {\((\mach_1,(V_1,\sigma_1)) \stepstounder{\overline{\psi},\obs} (\mach_1',\context_1')\)}
                   {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
@@
                   {\(|\sigma_2| \geq d\)}
-                  {\((\mach_2,(V_2,\sigma_2)) \stepstounder{\overline{\psi}} (\mach_2',\context_2',\obs)\)}
-                  {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
+                  {\((\mach_2,(V_2,\sigma_2)) \stepstounder{\overline{\psi},\obs} (\mach_2',\context_2')\)}
+                  {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2',\context_2')\)}
                   {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
@@
 where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
-of a call. Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
-\(s\) is the target of a tailcall. As a shorthand, we write that each property is defined
-by a criterion that must hold ``for all (tail)call targets \(s\).''
+of a call.
+%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
+%\(s\) is the target of a tailcall.
+As a shorthand, we write that each property is defined
+by a criterion that must hold ``for all call targets \(s\).''
 
@@
 is irrelevant.
+\apt{This paragraph needs major expansion, to paraphrase the definitions informally,
+  and TO SAY MUCH MORE ABOUT THE CALLEE PROPERTIES, which are not intuitive at all (to me)!}
 
@@
     1
-    & \(\wbcf \triangleq\) & \(|\sigma| \uparrow \ret\)
+    & \(\wbcf \triangleq\) & \((|\sigma| \uparrow \ret) ~ ????\)
     & \(\textnormal{ where } \ret ~ \mach' \triangleq \)
@@
     & \(\textnormal{ where } \intProp ~ \mach' \triangleq
-    \mach' \parallel \sealed(V) \cap \Delta(\mach,\mach')\)
+    \mach' \parallel (\sealed(V) \cap \Delta(\mach,\mach'))\)
 %    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
@@
 %    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
-    & \(\textnormal{ for all (tail)call targets } (\mach,(V,\sigma))\) \\
+    & \(\textnormal{ for all call targets } (\mach,(V,\sigma))\) \\
     \rowcolor{black!20}
@@
     & \(\textnormal{ where } \cconfProp ~ \mach' \triangleq
-    \mach' \parallel \Delta(\mach,\mach') - \components\)
+    \mach' \parallel (\Delta(\mach,\mach') - \components)\)
 %    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
-    & \(\textnormal{ for all (tail)call targets } (\mach,(V,\sigma))\) \\
+    & \(\textnormal{ for all call targets } (\mach,(V,\sigma))\) \\
     \rowcolor{black!10}
@@
 %    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
-    & \(\textnormal{ for all (tail)call targets } (\mach,(V,\sigma))\) \\
+    & \(\textnormal{ for all call targets } (\mach,(V,\sigma))\) \\
    \rowcolor{black!20}
@@
   \end{tabular}
-  \caption{Properties}
+  \caption{Properties\apt{In WBCF, surely we need a state arg? But it needs to be the state \emph{before} the call,
+      doesn't it?}}
   \label{tab:props}
@@
 The system we've modeled so far has been very simple, but our properties
-support an important dimension of generality: adding features. We now
-briefly introduce the features of our primary testing machine.
+support an important dimension of generality: adding features.\apt{Just what needs to be changed
+to add these features?}
+We now briefly introduce the features of our primary testing machine.\apt{Make sure
+testing has been introduced properly before now. And why ``primary''?}
 
@@
 \(\mathbf{call}\) with \(\overline{sa}\) as the empty set, and \(\mathbf{alloc}\)
-with \(\mathbf{f}\).
+with flag \(\mathbf{f}\).
 
@@
 arguments and return values. This is frequently not the case in a realistic
-setting. Arguments may be passed on the stack due to being spilled, as an implementation
-of variadic arguments, or because they are reference objects that inherently have
+setting. Arguments may be passed on the stack (e.g. because there are too many
+to pass in registers), as an implementation
+of variadic arguments, or because they are reference objects \apt{??} that inherently have
 pass-by-reference semantics. The caller may also take the address of a local object and
@@
 
-In this mechanism, we can describe an argument being passed by their relative address
+Using this mechanism, we can describe an argument being passed on the stack
+by giving its \(\SP\)-relative offset,
 with the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
@@
 \(\reg\) as holding the reference, then the operations is
-\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
-
-In both cases, the immediate callee retains access to the passed object, but future
+\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).\apt{This is unclear to me. 
+  The compiler knows where the cbr parameter lives in the caller's stack
+  frame, so why wouldn't it use a sp-relative spec for that too? Does it have
+  to do with passing a CBR parameter further down to another callee as you allude
+  to in the next para?}
+
+In both cases, the immediate callee retains access to the passed object, but deeper
 callees will not---unless the object is passed-by-reference further down the stack.
@@
 
-Our primary machine treats this scenario as the caller allocating a ``public''
+Our primary \apt{again...} machine treats this scenario as the caller allocating a ``public''
 object that can be accessed by anyone until it's deallocated. We extend the
@@
 than \(\sealed\) to express this. Of course, some objects may be passed as arguments
-in place, easily modeled by including them in \(\overline{sa}\). More sophisticated
+in place\apt{??}, easily modeled by including them in \(\overline{sa}\). More sophisticated
 behaviors may require sequences of {\bf dealloc}, {\bf tailcall}, then {\bf alloc}.

### 2023-02-02 submission5/ideas-v1.tex
pass through formalism

@@
 an event (see below) and \(\bar{\psi}\) is a list of security-relevant operations.
+\apt{Example doesn't include the event annotation. Should remove here, or add a note there.}
 We then lift this into a transition between pairs of machine states and contexts,
-\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
+\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).\apt{We don't need this notation for the example.}
 

### 2023-02-02 submission5/core.tex
Merging in the separate files

@@
 
-\input{intro-v2}
+\subsection{The Call Stack and Its Security}
+
+Computer programming relies on functions (or subroutines, procedures, methods, etc.)
+as a foundational abstraction. Functions are units of computation that
+can call one another to realize larger computations.
+%
+At a low level, a function activation comprises its own data, as
+well as information about its \emph{caller} function, to which it must return.
+The \emph{(call) stack} is the fundamental data structure charged with realizing
+this abstraction. From a security perspective, attacks on the stack are really attacks
+on the fundamental functional abstraction itself, and are therefore every bit as
+pervasive.
+
+Indeed, the call stack is a perennial target for low-level attacks.
+%
+The most recent release of the authoritative CWE Top 25
+Most Dangerous Software Weaknesses
+list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
+shows general classes of vulnerabilities closely related to the stack
+%% \rb{(though not exclusively)}
+consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read) or \#7 (CWE-416, Use After Free).
+%
+Moreover, many other types of faults involve some kind of control or data flow
+using data stored in the stack \cite{DBLP:conf/raid/VeendCB12,
+  DBLP:conf/sp/SzekeresPWS13,
+  DBLP:conf/sp/HuSACSL16}.\footnote{\url{https://github.com/Microsoft/MSRC-Security-Research/blob/master/presentations/2019_02_BlueHatIL/}}\footnote{\url{https://www.chromium.org/Home/chromium-security/memory-safety/}}
+
+The variety in attacks on the stack is mirrored by the
+software and hardware protections that aim to prevent them,
+%
+including stack canaries~\cite{Cowan+98},
+bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
+split stacks~\cite{Kuznetsov+14},
+shadow stacks~\cite{Dang+15,Shanbhogue+19},
+capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
+
+Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
+as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
+for instance, aim to ``restrict the flexibility available in creating gadget chains''
+\cite{Shanbhogue+19}, not to categorically
+rule out attacks. Other mechanisms are stronger, but formal
+guarantees are rarely discussed.
+Indeed, the Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
+Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
+to offer (various)\bcp{Maybe we can omit that word here and spell it
+  out in the related work.} stack safety formalizations.
+
+Meanwhile, Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
+micro-policies for stack safety, at least some of which should be amenable to making
+formal guarantees. Among these is their ``Lazy Tagging and Clearing'' \(LTC\) micropolicy,
+which is is peculiar in that function activations may write into one another's stack frames,
+as long as the data cannot be accessed afterward. As a result, \(LTC\) definitely would not
+fulfill a tag-based adaptation Georges et al.'s state-of-the-art property.
+
+We believe that, nevertheless, \(LTC\) enforces stack safety.
+It does not always protect individual addresses on the stack, but it aims
+to prevent one function activation from interfering with another's behavior---to protect
+the functional abstraction. We argue that a formal characterization of \(LTC\)'s protections
+should fall under the umbrella of stack safety.
+
+In this work, we propose such a characterization, built from first principles
+in the hopes that it will be broadly applicable, and demonstrate
+that it can distinguish between correct and flawed versions of Roessler and DeHon's
+micro-policies. We use the formal tools of language-based
+security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
+properties describing
+the {\em integrity} and {\em confidentiality} of the caller’s local state
+and the callee's behavior during the callee's execution, plus the control-flow protection
+of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
+
+The structure of our properties enables them to be easily extended to systems with
+sophisticated features such as callee-save registers, argument-passing on the stack,
+and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
+test the extended micro-policy. We find that, apart from the extensions, \(LTC\)
+is flawed in a way that undermines integrity and confidentiality, and we correct
+this flaw.
+
+\paragraph{Towards a Formal Property for Stack Safety}
+
+Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
+other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
+or tested. Such properties become the specification against which enforcement can be validated;
+even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
+\emph{why} they fail.
+
+We take the view that security properties should be motivated by security requirements,
+and flexible enough to support various sets of security requirements depending on the
+goals of the system that they validate.
+Following this reasoning, our vision of stack safety is as a collection of properties
+describing the security needs of a system with regard to its functional abstractions.
+While the stack still plays a key role, our stack safety is about functions. The most
+dramatic consequence of this is that we treat registers as part of a function activation's
+state, and hold them to the same protection requirements as stack memory.
+
+Flexibility is especially important for stack safety.
+%
+Most security properties are described at the level of a high-level programming language,
+and translated to the target-level via a secure compiler. Stack safety cannot
+be defined in this way, as the abstraction is completely transparent in most source languages.
+(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
+which is broadly consistent across language semantics, and therefore
+amenable to a very general high-level approach.)
+%
+But it is not satisfying as a low-level property; at the level of the ABI, the
+specification of a well-behaved stack is minimal. The ABI is not concerned with such
+questions as whether a caller's frame should be readable or writable to its callee,
+leaving those concerns to be determined by a high-level language built atop the stack.
+These are high-level security questions.
+
+For instance, the current state-of-the-art stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\em Well-bracketed Control Flow} (WBCF) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable. Related, their predecessor
+Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
+their enforcement mechanism did protect the confidentiality of the caller (not the callee.)
+
+These two cases represent two dimensions in which a stack safety definition requires
+flexibility: what constitutes safe behavior differs between systems that support different
+features, and different contexts have different top-level security requirements.
+
+We are not attempting to present a universal definition of stack safety for all systems,
+but we do construct our definition with these dimensions of flexibility in mind.
+Our model is decomposed into:
+(1) a lightweight model of the system and its security-relevant features, from which we
+derive the integrity, confidentiality, and control-flow requirements of each call and
+return, (2) the application of that model to a particular target system, and (3)
+the formal criteria for the integrity, confidentiality, and control flow
+requirements.
+
+We show that we can extend our system to new features by first introducing a simple
+example model, and then moving to a more realistic model with multiple new features:
+callee-save registers, argument passing on the stack, and tailcall elimination, simply
+by refining a few definitions in (1). This becomes the model that we validate.
+In \cref{???}, we extend the model further with arbitrary sharing of memory-safe
+pointers. Our criteria in (3) cover a total of five core requirements:
+\(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
+callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
+They are related to one another as seen in
+\cref{fig:propertydiagram}: \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
+\(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
+according to the needs of a system.
+
+In particular, many enforcement techniques focus purely on
+well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
+address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
+to enforce it completely. Others combine this protection with some degree of memory protection,
+chiefly focusing on integrity. Interestingly, Skorstengaard et al. \cite{SkorstengaardSTKJFP}
+describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
+to the combination of \(\clri\) and \(\clrc\). Their state-of-the-art successor,
+Georges et al. \cite{Georges22:TempsDesCerises}, cite an early preprint of our work
+as inspiration for extending their formal discussion to include confidentiality.
+When checking if a preliminary version of these properties applied to
+the Cerise calling convention, they noted that the Skostengaard et al. calling convention
+did not enforce callee confidentiality, and made sure that their version would
+\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
+This demonstrates the benefit of our choice to explicitly state properties in security
+terms: specifying security mechanisms is hard, and when the specification takes the
+form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
+security requirement.
+
+%These are defined
+%formally in \cref{sec:props}. . We further define two
+%important combinations of properties: full stack safety \(\textnormal{\sc FSS}\),
+%and caller safety \(\textnormal{\sc ClrS}\). Of these, most existing concepts of
+%stack safety correspond to either \(\wbcf\) or \(\textnormal{\sc ClrS}\), which are
+%concerned with the basic guarantees that the caller recieves during the call, while
+%\(\textnormal{\sc FSS}\) is a natural extension that enforces a higher-level concept
+%of a correct interface between caller and callee. \sna{Blah. Needs work.}
+
+Flexibility aside, we show that our properties are useful for their purpose:
+distinguishing correct enforcement from incorrect. We use
+QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for Coq, to generate random programs and check that the
+stack safety {\em micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
+correctly detect the ones that attempt to violate one of our properties. Furthermore, we
+check that the testing framework is able to generate counterexamples
+that violate our properties but are \emph{not} halted by incorrect
+variants of the enforcement mechanisms---including the original \(LTC\)!
+
+\subsection{Contributions}
+
+In sum, we offer the following contributions:
+
+\begin{itemize}
+\item A novel characterization of stack safety as a collection of properties:
+  confidentiality and integrity for callee and caller, and well-bracketed
+  control-flow. These can characterize the security requirements of systems
+  that support only some portion of stack safety, and are
+  parameterized over a notion of external observation, so that they can characterize
+  lazy enforcement.
+\item An instantiation of these definitions in a realistic
+  setting with features such as argument passing
+  on the stack, callee-saves registers, and tail-call elimination.
+  Our model is modular enough that handling these features is straightforward.
+\item The validation, via property-based randomize testing, of tag-based
+  stack-safety policies. We find that \emph{Lazy Tagging and Clearing}
+  falls short of stack safety, and propose a fix.
+\end{itemize}
+
+Our artifact contains Coq formalizations of our properties for illustrative purposes
+as well as their testing implementation. There are no proofs---we use primarily Coq for the
+QuickChick library and to ensure that the formalizations are unambiguous.
+%
+\rb{Mention proofs in future work?}
+
+\subsection{Threat Model and Limitations}
+
+When our properties are used to evaluate a system, the threat model will depend on the
+details of that system. However, there are some constraints that our design puts on
+any system. We must trust that the method of distinguishing security-relevant operations is accurate; if it
+involves labels placed on code by a compiler, that means trusting that the compiler placed
+those labels correctly. If operations occur that are not recognized, those operations
+might not be guaranteed to protect their principals---for instance, an unlabeled call
+might not protect the caller's data from the callee. On the flip side, applying an incorrect
+label most likely means that the property becomes too strong to be enforced.
+
+Otherwise, we do not assume that the code adheres to any particular
+calling convention or implements a source language construct.
+In particular, while we are agnostic as to the source
+language, we certainly aim to support C, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code. A given enforcement
+mechanism may place additional constraints, particularly on the behavior of
+call and return sequences. For instance, extant implementations tend to assume
+implicitly that callee-saved registers are treated appropriately by whichever compiler
+generated their code. Our properties explicitly require them to be treated as belonging
+to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
+%
+\rb{This last bit about callee-saved registers belonging to the caller is subtle, I think it risks the reader missing the point.}
+
+In general, it is impossible to distinguish buggy machine code from an attacker; in
+our examples we will identify one function or another as an attacker, but we do not
+require any static division between trusted and untrusted code, and we aim to protect
+even buggy code.
+
+This is a strong threat model, but it does omit some important aspects of
+security enforcement: in particular, hardware and timing attacks are out of scope, and
+we do not address concurrency.
+%
+\rb{Anything about concurrency in future work, though?}
 
@@
 
-\input{ideas-v1}
+The paper is structured as follows. In \cref{sec:example}, we will walk through a function
+call in a simple example machine and discuss how each of our properties applies to it,
+informally. In the process we discuss some motivation for the properties, from a security
+perspective: we consider a simple attacker, and discuss the harm that it can do to
+its caller. This perspective is essential to understanding how lazy enforcement can
+be safe: even if an attacker can overwrite some memory, if it cannot accomplish
+its bigger-picture goals, the system is safe.
+
+Then in \cref{sec:machine} we will introduce the machine model that we test, and give
+its {\em security semantics}, described in brief below. We formalize the definitions of
+stack safety that derive from the security semantics in \cref{sec:props}.
+\Cref{sec:testing} describes the testing framework, and
+\cref{sec:relwork,sec:future} discuss related and future work.
+The remainder of this section introduces ideas that will be important throughout.
+
+\paragraph*{Security Semantics and Property Structure}
+
+A {\it security semantics} extends a machine
+with additional context about the identity of current and pending
+functions (which act as security principals) and about the registers and memory they require
+to be secure. This added context is purely notional;
+it does not affect the real machine. The security context
+evolves dynamically through the execution of {\it security-relevant operations},
+which include events like calls, returns, and frame manipulations.
+Our security properties are phrased in terms of this context, often as predicates
+on future states, e.g. of the form ``when control returns to the current function...'',
+or relations on traces of future execution (hyper-properties).
+
+The security-relevant operations typically correspond to underling machine instructions,
+but the correspondence may not be obvious just by inspecting the machine code.
+For example, in the tagged RISC-V machine we use in our examples and tests,
+calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
+and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
+instructions might also be used for other things. An instantiation of an operation
+might require some external mechanism for distinguishing which instructions are
+security-relevant; the most obvious technique is to add optional labels to instructions,
+but the theory does not depend on this choice.
+
+Whatever the mechanism for labeling transitions, we get a base
+machine transition \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
+an event (see below) and \(\bar{\psi}\) is a list of security-relevant operations.
+We then lift this into a transition between pairs of machine states and contexts,
+\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
+
+The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
+\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
+are shown in white: calls, returns, and the allocation and deallocation of local memory.
+In \cref{sec:machine}), we cover additional operations that add parameters to the examples',
+to support various forms of memory sharing \ifexceptions , exceptions, \fi
+and tail-call elimination. These are shown in light gray.
+We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
+based on pointer provenance in \cref{app:ptr}, represented by the dark grey
+operations, but we do not test this.
+
+\newcommand{\example}{\rowcolor{black!0}}
+\newcommand{\testing}{\rowcolor{black!10}}
+\newcommand{\theory}{\rowcolor{black!25}}
+
+\begin{table}
+  \begin{tabular}{| l | l |}
+    \hline
+    \(\psi \in \Psi\) & Parameters \\
+    \hline
+    \example \(\mathbf{call}\) & Target address, argument registers \\
+    \testing & Stack argument offset \& size \\
+    \example \(\mathbf{return}\) & \\
+    \example \(\mathbf{alloc}\) & Offset \& size \\
+    \testing & Public flag \\
+    \example \(\mathbf{dealloc}\) & Offset \& size \\
+    \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
+    \hline
+    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
+    \hline
+    \theory \(\mathbf{promote}\) & Register, offset \& size \\
+    \theory \(\mathbf{propagate}\) & Source register/address \\
+    \theory & Destination register/address \\
+    \theory \(\mathbf{clear}\) & Target register/address \\
+    \hline
+  \end{tabular}
+  \caption{Security-relevant Operations and their Parameters}
+  \label{tab:psi}
+\end{table}
+
+Our security semantics has similarities to the overlay semantics that have been proposed
+to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
+the behavior of the underlying machine in any way. Rather, it tracks additional context
+about the history of security-relevant operations, which inform the criterion a machine
+must satisfy in order to correctly obey the properties.
+
+The security principals in our account of stack safety are function
+activations: the context tracks, for the active function and all pending
+functions, their {\em view} of the security state of the system.
+
+\paragraph*{Views}
+
+The security context consists of a stack of views: functions that map
+each state element to a {\it security class}. The security classes are
+\(\public\), \(\unsealed\), \(\object\), and \(\sealed\).
+
+State elements that are outside of the stack---general-purpose memory used for
+globals and the heap, as well as the code region and globally-relevant
+registers---are always labeled \(\public\). We place security requirements on some
+\(\public\) elements for purposes of \(\wbcf\), and a given enforcement mechanism
+might restrict their access (e.g., by rendering code immutable) but for integrity
+and confidentiality purposes they are considered accessible at all times.
+
+For a newly active function, every element that is available for use but uninitialized
+is seen as \(\unsealed\). From the perspective of the caller, the callee has no obligations
+regarding its use of \(\unsealed\) elements, but the callee itself
+
+Arguments are seen as \(\object\), meaning that their contents may be safely used.
+When the active function allocates memory, that memory will also be \(\object\).
+Then, on a call, \(\object\) elements that are not being used to communicate with
+the new callee will become \(\sealed\)---reserved for an inactive principal,
+and expected to be unchanged when it becomes active again.
+
+\paragraph*{Events}
+
+It is common to specify the behavior of a system, including security properties, in
+terms of traces of observable {\em events} rather than details of the machine state.
+Lazy enforcement works because the deferred check catches errors before they appear
+in the trace. The nature of events is a parameter of the machine, and can be specialized
+to any notion of observable behavior.
+
+\paragraph*{Variants and Non-interference}
+
+To characterize confidentiality, we borrow the idea of \emph{variant} states
+from the theory of non-interference. Non-interference is a way of expressing
+the flow of information through the system. If we begin with a state \(\mach\)
+and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
+then execute both \(\mach\) and \(\nach\), any differences in the outputs they
+produce must be the result of accessing \(\component\).
+
+\paragraph*{Putting It Together}
+
+Putting these ideas together, we get the basic structure of a property in our model.
+First we define the security-relevant operations that we support, a security context,
+and the function according to which the security context updates with each step.
+The security context gives each function activation a view, mapping each element to a
+security class.
+
+Then, we define our security requirements in terms of the security classes of state
+elements. These take the form of predicates on states and contexts that use variants,
+assertions about future return states, and comparison of event traces to capture the
+desired behavior without reference to implementation details.
+
+Separately, in order to apply a property to a specific machine, we do need to consider
+implementation details, by determining how the machine's execution translates into
+security relevant operations. But this step is transparent to the properties themselves.
 

### 2023-02-02 submission5/core.tex
Also merging in future work

@@
 
-\input{future_work.tex}
+\subsection{Testing The Stack Safety Menagerie}
+
+In \cref{sec:intro}, we discussed several dimensions of flexibility
+that our properties should exhibit: applicability to different system
+features and different security requirements. Another important dimension
+is applicability to multiple enforcement mechanisms.
+In the following, we discuss how different existing mechanisms
+might fit into our model.
+
+\paragraph{Stack Canaries and Shadow Stacks}
+%
+Canaries are special values inserted by the caller at certain points of the
+stack and checked for corruption on return, for protection against
+some classes of control-flow hijacking (i.e., overwriting the caller's
+return address through a buffer overflow that would also overwrite the canary
+detectably). A shadow stack is a mirror image of the control-relevant parts of the stack,
+maintained by the program to attempt to detect attempts to hijack its control
+flow and ``restrict the flexibility available in creating gadget chains''
+\cite{Shanbhogue+19}.
+%
+Interestingly, these are lazy enforcement mechanisms, in that
+the attack may occur and be detected some time later, as long as
+it is detected before it can become dangerous. That would make our
+formalism a good fit for defining their security, except that
+they are chiefly hardening techniques: they increase the difficulty
+of some control-flow attacks on the stack, but cannot provide absolute
+guarantees on \(\wbcf\) under a normal attacker model.
+
+\paragraph{Split Stacks}
+%
+Code-pointer integrity seeks to prevent control-flow hijacking attacks by
+splitting the program memory, including the stack, into a regular region and a
+safe region; objects that require protection are placed in the latter and
+all accesses to them protected by static or dynamic checks.
+%
+Like shadow stacks, bounds checking is a high-level technique designed to
+protect code written in a high-level language, but it is designed to
+offer protection under a stronger threat model that also considers the presence
+of untrusted code. As with most other software-based methods, it purports to
+protect exclusively the control-relevant parts of the stack, making no claims
+about its data-relevant parts---we would solely be testing \(\wbcf\).
+
+\paragraph{Bounds Checking}
+%
+Under a bounds checking discipline such as SoftBound \cite{NagarakatteZMZ09}, all the pointers
+in a program are extended with some disjoint metadata, and these are combined
+to gate memory accesses. These approaches enforce a form of \emph{memory safety},
+and we would therefore expect them to enforce \(\clri\) and \(\clrc\). They aim
+to enforce \(\wbcf\) by cutting off attacks that involve memory-safety violations,
+but that may not be sufficient.
+
+Bounds checking approaches require substantial compiler cooperation. This is not a
+problem for our properties in general, but it is not very compatible with
+generation-by-execution of low-level code. A better choice would be to generate
+high-level code using a tool like CSmith \cite{DBLP:conf/pldi/YangCER11}, or prove the properties instead.
+
+\paragraph{Capabilities}
+
+Capability machines can represent and manipulate
+unforgeable tokens of authority over specific regions of
+memory using a combination of dedicated out-of-band memory and
+specialized ISA instructions.
+%
+The most prominent contemporary representative of this long line of work is
+CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
+memory protection and compartmentalization.
+%
+Previous work has uses these principles in simplified models to implement
+\emph{secure calling conventions} for the call
+stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
+protections of bounds-checking with software-based methods of control-flow protection.
+
+There are several proposals around the use of CHERI capabilities to enforce stack safety,
+including mechanisms that use the standard CHERI hardware (which includes local
+capabilities) \cite{SkorstengaardLocal},
+and others that propose entirely new types of capabilities, such as linear
+\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
+\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
+It is a high priority to test this most recent work by Georges et al., as it is designed
+to be similar to all of our properties except for \(\clei\).
+
+\subsection{More Features}
+
+Several common features did not make it into this work: most importantly concurrency and
+exceptions. We plan to extend the model to include them, and then test them. We would also
+like to test the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
 

### 2023-02-02 submission5/core.tex
add direction info to captions

@@
 
-\caption{Example: {\tt main}\apt{Needs machine-set diagram.}}
+\caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; addresses increase to the right and the stadk grows to the left.\apt{Needs machine-set diagram.}}
 \label{fig:main}
@@
   \end{tabular}
-  \caption{Execution of example up through the return from {\tt f}. In right-hand diagram, addresses increase to the right and
-  boxes represent 4-byte words.}
+  \caption{Execution of example up through the return from {\tt f}. In stack diagram, addresses increase to the right, stack grows to the left, and boxes represent 4-byte words.}
 \label{fig:exec1}

### 2023-02-03 submission5/core.tex
some work on 1

@@
 as a foundational abstraction. Functions are units of computation that
-can call one another to realize larger computations.
-%
-At a low level, a function activation comprises its own data, as
+can call one another to realize larger computations in a modular fashion.
+%
+At a low level, a function activation manages its own data, such as local variables, as
 well as information about its \emph{caller} function, to which it must return.
 The \emph{(call) stack} is the fundamental data structure charged with realizing
-this abstraction. From a security perspective, attacks on the stack are really attacks
-on the fundamental functional abstraction itself, and are therefore every bit as
-pervasive.
-
-Indeed, the call stack is a perennial target for low-level attacks.
-%
-The most recent release of the authoritative CWE Top 25
-Most Dangerous Software Weaknesses
-list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
-shows general classes of vulnerabilities closely related to the stack
+this abstraction in memory. \an{registers?} From a security perspective, attacks on the stack are really attacks
+on the fundamental functional abstraction itself. %, and are therefore every bit as
+%pervasive.
+
+Indeed, the call stack is a ancient~\cite{one06phrack} and perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
+  DBLP:conf/sp/SzekeresPWS13,
+  DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
+  target for low-level attacks. \apt{characterize the variety of attacks?}
+%
+%The most recent release of the authoritative CWE Top 25
+%Most Dangerous Software Weaknesses~\cite{
+%list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
+%shows that general classes of vulnerabilities closely related to the stack
 %% \rb{(though not exclusively)}
-consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read) or \#7 (CWE-416, Use After Free).
-%
-Moreover, many other types of faults involve some kind of control or data flow
-using data stored in the stack \cite{DBLP:conf/raid/VeendCB12,
-  DBLP:conf/sp/SzekeresPWS13,
-  DBLP:conf/sp/HuSACSL16}.\footnote{\url{https://github.com/Microsoft/MSRC-Security-Research/blob/master/presentations/2019_02_BlueHatIL/}}\footnote{\url{https://www.chromium.org/Home/chromium-security/memory-safety/}}
-
+%consistently rank at and near the top of the list, e.g., \#1 Out-of-bounds Write, \#5 Out-of-bounds Read, and \#7 Use After Free.
+%
+%Moreover, many other types of faults involve some kind of control or data flow
+%using data stored in the stack 
 The variety in attacks on the stack is mirrored by the
@@
 capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
-and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
-
+and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18,Gollapudi+23}.
+
+\apt{Add brief case for having formal guarantees.}
 Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
@@
 guarantees are rarely discussed.
-Indeed, the Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
+Indeed, the Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP}\apt{that's not Cerise} and
 Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
@@
 
-Meanwhile, Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
-micro-policies for stack safety, at least some of which should be amenable to making
-formal guarantees. Among these is their ``Lazy Tagging and Clearing'' \(LTC\) micropolicy,
+\apt{Transition}
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
+micro-policies~\cite{pump_oakland2015} for stack safety, at least some of which should be amenable to making
+formal guarantees. Among these is their ``Lazy Tagging and Clearing'' (\(LTC\)) micro-policy,
 which is is peculiar in that function activations may write into one another's stack frames,
-as long as the data cannot be accessed afterward. As a result, \(LTC\) definitely would not
-fulfill a tag-based adaptation Georges et al.'s state-of-the-art property.
-
-We believe that, nevertheless, \(LTC\) enforces stack safety.
-It does not always protect individual addresses on the stack, but it aims
-to prevent one function activation from interfering with another's behavior---to protect
-the functional abstraction. We argue that a formal characterization of \(LTC\)'s protections
-should fall under the umbrella of stack safety.
-
-In this work, we propose such a characterization, built from first principles
-in the hopes that it will be broadly applicable, and demonstrate
-that it can distinguish between correct and flawed versions of Roessler and DeHon's
-micro-policies. We use the formal tools of language-based
+as long as the data cannot be accessed afterward. Roessler and DeHon themselves say that
+\(LTC\) does not enforce memory safety, but rather a form of data-flow integrity~\cite{castro+06}.
+And \(LTC\) definitely would not enforce a tag-based adaptation of Georges et al.'s property.
+
+Neverthess, we believe that there is a useful sense in which \(LTC\) does (attempt to) enforce stack safety. 
+It does not always protect individual addresses on the stack, but it tries to 
+prevent one function activation from interfering with another's behavior---protecting 
+the functional abstraction, which we argue is the real essence of stack safety.
+%We argue that a formal characterization of \(LTC\)'s protections
+%should fall under the umbrella of stack safety.
+
+In this work, we propose a formal characterization of stack safety based on the idea
+of protecting each function activation from other activations.
+We use the formal tools of language-based
 security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
@@
 and the callee's behavior during the callee's execution, plus the control-flow protection
-of {\em well-bracketed control flow}~\cite{SkorstengaardSTKJFP} (\(\wbcf\)).
+of {\em well-bracketed control flow} (\(\wbcf\))~\cite{SkorstengaardSTKJFP}.
+While our properties are motivated by a desire to specify \(LTC\), 
+they are built from first principles in the hopes that they will be broadly applicable
+to a range of enforcement mechanisms.
 
@@
 and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
-test the extended micro-policy. We find that, apart from the extensions, \(LTC\)
+test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
 is flawed in a way that undermines integrity and confidentiality, and we correct
@@
 
-\paragraph{Towards a Formal Property for Stack Safety}
+\paragraph*{Towards Formal Properties for Stack Safety}
 
@@
 
-For instance, the current state-of-the-art stack-safety specification from Georges et al.
+For instance, the stack-safety specification from Georges et al.
 \cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\em Well-bracketed Control Flow} (WBCF) under the assumption that
+control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
 a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Related, their predecessor
+return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
 Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
-their enforcement mechanism did protect the confidentiality of the caller (not the callee.)
+their enforcement mechanism did protect the confidentiality of the caller (though not the callee).
 
@@
 describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
-to the combination of \(\clri\) and \(\clrc\). Their state-of-the-art successor,
-Georges et al. \cite{Georges22:TempsDesCerises}, cite an early preprint of our work
+to the combination of \(\clri\) and \(\clrc\). The follow-up work by 
+Georges et al. \cite{Georges22:TempsDesCerises} cites an early preprint of our work
 as inspiration for extending their formal discussion to include confidentiality.
@@
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
-violating the program's {\it well-bracketed control flow} (WBCF.)
-%
-In Figure \ref{subfig:WBCF2}, a different attack violates WBCF, this time
+violating the program's {\it well-bracketed control flow} (\(\wbcf\).)
+%
+In Figure \ref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
 by returning to the correct program counter but with the wrong stack pointer.
@@
 just after the call.
-WBCF requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
+\(\wbcf\) requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
 %For nested calls, where the pending stack is initially larger, the same principle
@@
 
-Like WBCF, stack integrity defines a condition at the call that must hold upon
+Like \(\wbcf\), stack integrity defines a condition at the call that must hold upon
 return. This time the condition applies to all of the memory in the function's
@@
 which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
-assuming that {\tt sec} is 5. Similar to WBCF, we define
+assuming that {\tt sec} is 5. Similar to \(\wbcf\), we define
 \(\intProp\) as a predicate on states that holds if

### 2023-02-03 submission5/core.tex
Worked through sections 1 and 2

--- 2023-02-02 submission5/core.tex
@@
 The \emph{(call) stack} is the fundamental data structure charged with realizing
-this abstraction. From a security perspective, attacks on the stack are really attacks
+this abstraction, aided by an ABI that defines how registers are shared between activations.
+From a security perspective, attacks on the stack are really attacks
 on the fundamental functional abstraction itself, and are therefore every bit as
@@
 
-Indeed, the call stack is a perennial target for low-level attacks.
-%
-The most recent release of the authoritative CWE Top 25
-Most Dangerous Software Weaknesses
-list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
-shows general classes of vulnerabilities closely related to the stack
-%% \rb{(though not exclusively)}
-consistently at and near the top, e.g., \#1 (CWE-787, Out-of-bounds Write), \#5 (CWE-125, Out-of-bounds Read) or \#7 (CWE-416, Use After Free).
-%
-Moreover, many other types of faults involve some kind of control or data flow
-using data stored in the stack \cite{DBLP:conf/raid/VeendCB12,
+Indeed, the call stack is a perennial target for low-level attacks \cite{phrack96:smashingthestack}.
+%
+The CWE Top 25 Most Dangerous Software Weaknesses list
+\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
+is consistently topped by classes of vulnerabilities closely related to the stack
+e.g., \#1 Out-of-bounds Write, \#5 Out-of-bounds Read, or \#7 Use After Free.
+%
+And many other types of attack involve compromising the stack as a step in a larger workfloaw
+\cite{DBLP:conf/raid/VeendCB12,
   DBLP:conf/sp/SzekeresPWS13,
-  DBLP:conf/sp/HuSACSL16}.\footnote{\url{https://github.com/Microsoft/MSRC-Security-Research/blob/master/presentations/2019_02_BlueHatIL/}}\footnote{\url{https://www.chromium.org/Home/chromium-security/memory-safety/}}
+  DBLP:conf/sp/HuSACSL16}.\footnote{\url{https://github.com/Microsoft/MSRC-Security-Research/blob/master/presentations/2019_02_BlueHatIL/}}\footnote{\url{https://www.chromium.org/Home/chromium-security/memory-safety/}}.
 
@@
 \cite{Shanbhogue+19}, not to categorically
-rule out attacks. Other mechanisms are stronger, but formal
-guarantees are rarely discussed.
+rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
+integrity\cite{Kuznetsov+14}
+are stronger, but most do not give formal guarantees.
 Indeed, the Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
 Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
-to offer (various)\bcp{Maybe we can omit that word here and spell it
-  out in the related work.} stack safety formalizations.
-
-Meanwhile, Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
-micro-policies for stack safety, at least some of which should be amenable to making
-formal guarantees. Among these is their ``Lazy Tagging and Clearing'' \(LTC\) micropolicy,
-which is is peculiar in that function activations may write into one another's stack frames,
-as long as the data cannot be accessed afterward. As a result, \(LTC\) definitely would not
-fulfill a tag-based adaptation Georges et al.'s state-of-the-art property.
+to offer stack safety formalizations.
+
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
+micro-policies for stack safety, at least some of which should be amenable to making strong
+formal guarantees. Their most realistic micro-policy, ``Lazy Tagging and Clearing'' (\(LTC\)),
+which makes an interesting trade-off for performance: it allows function activations to write
+into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
+it afterward. Because the memory does get corrupted, \(LTC\) would not fulfill an
+adaptation Georges et al.'s property to the tagged setting.
 
@@
 
-The structure of our properties enables them to be easily extended to systems with
+The structure of our properties enables them to be easily expanded to characterize systems with
 sophisticated features such as callee-save registers, argument-passing on the stack,
@@
 
-\paragraph{Towards a Formal Property for Stack Safety}
+\paragraph{Towards Formal Properties for Stack Safety}
 
@@
 
-We take the view that security properties should be motivated by security requirements,
+Security properties should be motivated by security requirements
 and flexible enough to support various sets of security requirements depending on the
@@
 by refining a few definitions in (1). This becomes the model that we validate.
-In \cref{???}, we extend the model further with arbitrary sharing of memory-safe
+In \cref{app:ptr}, we extend the model further with arbitrary sharing of memory-safe
 pointers. Our criteria in (3) cover a total of five core requirements:
@@
 stack safety {\em micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
-correctly detect the ones that attempt to violate one of our properties. Furthermore, we
-check that the testing framework is able to generate counterexamples
-that violate our properties but are \emph{not} halted by incorrect
-variants of the enforcement mechanisms---including the original \(LTC\)!
+correctly detect the ones that attempt to violate one of our properties.
+
+%Furthermore, we
+%check that the testing framework is able to generate counterexamples
+%that violate our properties but are \emph{not} halted by incorrect
+%variants of the enforcement mechanisms---including the original \(LTC\)!
+% SNA: we talk about this in the testing section, and it's a "best practice" for
+% PBT, but not really a contribution in and of itself
 
@@
 Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as their testing implementation. There are no proofs---we use primarily Coq for the
-QuickChick library and to ensure that the formalizations are unambiguous.
-%
-\rb{Mention proofs in future work?}
+as well as their testing implementation. There are no proofs---we use Coq primarily
+for the QuickChick library and to ensure that the formalizations are unambiguous.
+Proofs are future work.
 
@@
 informally. In the process we discuss some motivation for the properties, from a security
-perspective: we consider a simple attacker, and discuss the harm that it can do to
+perspective. We consider a simple attacker, and discuss the harm that it can do to
 its caller. This perspective is essential to understanding how lazy enforcement can
@@
 
+\paragraph*{Building a Property}
+
+Our properties follow a three-part structure that we term a {\em security semantics}.
+First we define the {\em security-relevant operations} that we support, and an extra
+piece of state called a {\em security context}, which updates alongside each step
+of execution. The security context gives each function activation a {\em view},
+mapping each element to a {\em security class}.
+
+Then, we define our high-level security requirements---integrity, confidentiality,
+and well-bracketed control flow---in terms of the security classes of state
+elements. These take the form of predicates on states and contexts that use variants,
+assertions about future return states, and comparison of event traces to capture the
+desired behavior without reference to implementation details.
+
+Separately, in order to apply a property to a specific machine, we do need to consider
+implementation details, by determining how the machine's execution translates into
+security relevant operations. But this step is transparent to the properties themselves.
+
 \paragraph*{Security Semantics and Property Structure}
 
-A {\it security semantics} extends a machine
+A security extends a machine
 with additional context about the identity of current and pending
@@
 it does not affect the real machine. The security context
-evolves dynamically through the execution of {\it security-relevant operations},
+evolves dynamically through the execution of security-relevant operations,
 which include events like calls, returns, and frame manipulations.
@@
 
-\paragraph*{Putting It Together}
-
-Putting these ideas together, we get the basic structure of a property in our model.
-First we define the security-relevant operations that we support, a security context,
-and the function according to which the security context updates with each step.
-The security context gives each function activation a view, mapping each element to a
-security class.
-
-Then, we define our security requirements in terms of the security classes of state
-elements. These take the form of predicates on states and contexts that use variants,
-assertions about future return states, and comparison of event traces to capture the
-desired behavior without reference to implementation details.
-
-Separately, in order to apply a property to a specific machine, we do need to consider
-implementation details, by determining how the machine's execution translates into
-security relevant operations. But this step is transparent to the properties themselves.
-
 \section{Properties by Example}
@@
 or calls, returns, and private allocations.
-% In this system a value is a pair of a
-%word (the ``payload'') and a tag drawn from some set \(\mathcal{T}\).
-%\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
-%for some type  of tags to be elaborated later, with \(|(w,t)| = w\).
 Figure \ref{fig:main} gives C code and corresponding compiled 32-bit RISC-V code
@@
 
+\subsection{Proofs}
+
 \bibliographystyle{IEEEtran}

### 2023-02-03 submission5/core.tex
a few things in 2

--- 2023-02-03 submission5/core.tex
@@
 
-The paper is structured as follows. In \cref{sec:example}, we will walk through a function
-call in a simple example machine and discuss how each of our properties applies to it,
-informally. In the process we discuss some motivation for the properties, from a security
+\apt{Not a good place for this.}
+The paper is structured as follows. In \cref{sec:example}, we walk through a function
+call in a simple example machine and discuss informally how each of our properties applies to it.
+In the process we discuss some motivation for the properties, from a security
 perspective: we consider a simple attacker, and discuss the harm that it can do to
@@
 be safe: even if an attacker can overwrite some memory, if it cannot accomplish
-its bigger-picture goals, the system is safe.
-
-Then in \cref{sec:machine} we will introduce the machine model that we test, and give
-its {\em security semantics}, described in brief below. We formalize the definitions of
-stack safety that derive from the security semantics in \cref{sec:props}.
+subvert the observable behavior of the caller, the system is safe.
+
+In \cref{sec:machine} we formalize the machine model, 
+its {\em security semantics}, and the stack safety properties built on these. 
 \Cref{sec:testing} describes the testing framework, and
@@
 The remainder of this section introduces ideas that will be important throughout.
+
+\apt{Something about how the machine can incorporate enforcement, with examples
+(tags, capabilities, SW only). }
 
@@
 might require some external mechanism for distinguishing which instructions are
-security-relevant; the most obvious technique is to add optional labels to instructions,
-but the theory does not depend on this choice.
+security-relevant. The most obvious technique, which we use in our examples, is to add labels to
+selected instructions, but the theory does not depend on this choice.
 
@@
 an event (see below) and \(\bar{\psi}\) is a list of security-relevant operations.
-We then lift this into a transition between pairs of machine states and contexts,
-\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
+%We then lift this into a transition between pairs of machine states and contexts,
+%\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
 
@@
 \begin{table}
+\begin{center}
   \begin{tabular}{| l | l |}
@@
     \hline
-    \example \(\mathbf{call}\) & Target address, argument registers \\
-    \testing & Stack argument offset \& size \\
+    \example \(\mathbf{call}\) & target address, argument registers \\
+    \testing & stack arguments (base, offset \& size) \\
     \example \(\mathbf{return}\) & \\
-    \example \(\mathbf{alloc}\) & Offset \& size \\
-    \testing & Public flag \\
-    \example \(\mathbf{dealloc}\) & Offset \& size \\
-    \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
+    \example \(\mathbf{alloc}\) & offset \& size \\
+    \testing & public flag \\
+    \example \(\mathbf{dealloc}\) & offset \& size \\
+    \testing \(\mathbf{tailcall}\) & (same as for \(\mathbf{call}\)) \\
     \hline
@@
     \hline
-    \theory \(\mathbf{promote}\) & Register, offset \& size \\
-    \theory \(\mathbf{propagate}\) & Source register/address \\
-    \theory & Destination register/address \\
-    \theory \(\mathbf{clear}\) & Target register/address \\
+    \theory \(\mathbf{promote}\) & register, offset \& size \\
+    \theory \(\mathbf{propagate}\) & source register/address \\
+    \theory & destination register/address \\
+    \theory \(\mathbf{clear}\) & target register/address \\
     \hline
   \end{tabular}
+\end{center}
   \caption{Security-relevant Operations and their Parameters}
@@
 is seen as \(\unsealed\). From the perspective of the caller, the callee has no obligations
-regarding its use of \(\unsealed\) elements, but the callee itself
+regarding its use of \(\unsealed\) elements, but the callee itself \apt{???}
 
@@
 \paragraph*{Putting It Together}
+\apt{Agree with AN that this should come first instead of last.}
 

### 2023-02-03 submission5/core.tex

--- 2023-02-03 submission5/core.tex
@@
 The \emph{(call) stack} is the fundamental data structure charged with realizing
-this abstraction in memory. \an{registers?} From a security perspective, attacks on the stack are really attacks
-on the fundamental functional abstraction itself. %, and are therefore every bit as
-%pervasive.
-
-Indeed, the call stack is a ancient~\cite{one06phrack} and perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
+this abstraction, aided by an ABI that defines how registers are shared between activations.
+From a security perspective, attacks on the stack are really attacks
+on the fundamental functional abstraction itself, and are therefore every bit as
+pervasive.
+
+Indeed, the call stack is a ancient~\cite{one06phrack} and
+perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
   DBLP:conf/sp/SzekeresPWS13,
   DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
-  target for low-level attacks. \apt{characterize the variety of attacks?}
+target for low-level attacks, frequently involving control-flow hijacking via corrupting
+the return address, and memory corruption more generally.
+
 %
@@
 \cite{Shanbhogue+19}, not to categorically
-rule out attacks. Other mechanisms are stronger, but formal
-guarantees are rarely discussed.
-Indeed, the Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP}\apt{that's not Cerise} and
+rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
+integrity\cite{Kuznetsov+14} are stronger, but most do not give formal guarantees.
+Indeed, the secure calling conventions of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
 Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
-to offer (various)\bcp{Maybe we can omit that word here and spell it
-  out in the related work.} stack safety formalizations.
-
-\apt{Transition}
+to offer stack safety formalizations.
+
+Among the mechanisms that should be amenable to making strong formal guarantees,
 Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
-micro-policies~\cite{pump_oakland2015} for stack safety, at least some of which should be amenable to making
-formal guarantees. Among these is their ``Lazy Tagging and Clearing'' (\(LTC\)) micro-policy,
-which is is peculiar in that function activations may write into one another's stack frames,
-as long as the data cannot be accessed afterward. Roessler and DeHon themselves say that
-\(LTC\) does not enforce memory safety, but rather a form of data-flow integrity~\cite{castro+06}.
-And \(LTC\) definitely would not enforce a tag-based adaptation of Georges et al.'s property.
+micro-policies~\cite{pump_oakland2015} for stack safety.
+Their most realistic micro-policy, ``Lazy Tagging and Clearing'' (\(LTC\)),
+makes an interesting trade-off for performance: it allows function activations to write
+into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
+it afterward. Because the memory does get corrupted, \(LTC\) would not fulfill an
+adaptation Georges et al.'s property to the tagged setting.
 
@@
 the functional abstraction, which we argue is the real essence of stack safety.
-%We argue that a formal characterization of \(LTC\)'s protections
-%should fall under the umbrella of stack safety.
 
@@
 behavior resulting in its compilation to arbitrary machine code. A given enforcement
-mechanism may place additional constraints, particularly on the behavior of
+mechanism or target architecture might place additional constraints, particularly on the behavior of
 call and return sequences. For instance, extant implementations tend to assume
-implicitly that callee-saved registers are treated appropriately by whichever compiler
-generated their code. Our properties explicitly require them to be treated as belonging
-to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
-%
-\rb{This last bit about callee-saved registers belonging to the caller is subtle, I think it risks the reader missing the point.}
+implicitly that callee-saved registers have their values maintained by whichever compiler
+generated their code. Our properties explicitly state this as a requirement,
+which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
 
@@
 security enforcement: in particular, hardware and timing attacks are out of scope, and
-we do not address concurrency.
-%
-\rb{Anything about concurrency in future work, though?}
+we do not address concurrency---that is a high-priority future work.
+%
 

### 2023-02-03 submission5/core.tex

@@
 
+\apt{Not a good place for this.}
 The paper is structured as follows. In \cref{sec:example}, we will walk through a function
@@
 security relevant operations. But this step is transparent to the properties themselves.
+
+\apt{Something about how the machine can incorporate enforcement, with examples
+(tags, capabilities, SW only). }
 

### 2023-02-03 submission5/core.tex
Testing updates

@@
 
-Indeed, the call stack is a ancient~\cite{one06phrack} and
+Indeed, the call stack is an ancient~\cite{one06phrack} and
 perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
@@
 We therefore choose to systematically \emph{test} their micro-policies.
+\leo{moved the following here, I think it makes more sense here.}
+Our primary testing targets are the eager {\em Depth Isolation}
+and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
 
@@
 to ensure that our coinductive trace definitions are well-formed, to reason about
-them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.
-
-To use QuickChick, we build random test-case generators that produce
+them\leo{where do we mention reasoning about them?}, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.
+
+\subsection{Test Generation}
+To use QuickChick, we developed random test-case generators that produce
 (1) an initial RISC-V machine state, including the program to be executed;
@@
 
-To write such generators we build on the work of
-Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
-introduced {\em generation by execution} to produce progams that lead
-machines towards interesting behaviors. We extend the technique with a degree
-of statefulness, to nudge the generator toward executions that will carry
-on without early failstops and thus have a better chance of reaching a
-state that presents opportunities for an attack.
-
-We need to extend typical testing schemes further to handle the nested
-nature of confidentiality: rather than just generating two
-initial machines that are variants of one another and letting them
-execute to test for noninterference, we generate a new variant
-{\em every time a call is made} and check confidentiality for the
-subtrace produced from that variant state until its corresponding
-return. As a result, a ``single'' confidentiality test compactly
-checks multiple nested calls.
-
-Our primary testing targets are the eager {\em Depth Isolation}
-and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
+To write such generators we build on the work of Hri\c{t}cu et
+al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
+introduced {\em generation by execution}, a technique that produces
+programs that lead to longer executions---and hopefully towards more
+interesting behaviors as a result. We extend their technique with an additional
+degree of statefulness: when complex calling conventions need to be
+observed to avoid early failstops, extra care must be taken to ensure
+that any such calls are well behaved. For example, immediately
+following a call we need to initialize any stack-allocated variables
+with an increased probability. However, at the same time, we need to
+allow for the possibility of {\em not} performing such an initialization
+to allow for potential attack vectors to manifest.
+
+Moreover, we need to further extend typical hyperproperty testing
+schemes to handle the nested nature of confidentiality: rather than
+just generating two initial machines that are variants of one another
+and letting them execute to test for noninterference, we generate a
+new variant {\em every time a call is made} and check confidentiality
+for the subtrace produced from that variant state until its
+corresponding return. As a result, a ``single'' confidentiality test
+compactly checks multiple nested calls.
 

### 2023-02-03 submission5/core.tex
First diagram, and macros for it

--- 2023-02-03 submission5/core.tex
@@
 is seen as \(\unsealed\). From the perspective of the caller, the callee has no obligations
-regarding its use of \(\unsealed\) elements, but the callee itself \apt{???}
+regarding its use of \(\unsealed\) elements.
 
@@
 lower addresses. We assume that {\tt main} begins at address 0 and its callee {\tt f} at address 100.
+
+\newcommand{\freebox}[1][10px]{\tikz \filldraw[fill=blue] (0,0) rectangle (#1,#1);}
+\newcommand{\pubbox}[1][10px]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (#1,#1);}
+\newcommand{\objbox}[1][10px]{\tikz \filldraw[fill=yellow] (0,0) rectangle (#1,#1);}
+\newcommand{\sealbox}[1][10px]{\tikz \filldraw[fill=red] (0,0) rectangle (#1,#1);}
+
+\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30px,30px) node[pos=.5]{#1};}
 
@@
   \begin{subfigure}{\columnwidth}
-    \includegraphics[width=\columnwidth]{stacklayout.png}
+    \center
+    \vspace{\abovedisplayskip}
+    \(\textnormal{\large \dots} \pubbox[30px] \!\! \stackrel{\textsc{\large sp}}{\textbox[\tt res]}
+    \stackrel{\textsc{\large 4(sp)}}{\textbox[\tt sens]}
+    \stackrel{\textsc{\large 8(sp)}}{\textbox[\tt sec]}
+    \stackrel{\textsc{\large 12(sp)} \hfill}{\textbox[\(\mathtt{ra_1}\)]\textbox[\(\mathtt{ra_2}\)]}
+    \textnormal{\large \dots}\)
   \end{subfigure}
 
-\caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; addresses increase to the right and the stadk grows to the left.\apt{Needs machine-set diagram.}}
+  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; addresses increase to the right and the stack grows to the left.}
   \label{fig:main}
 \end{figure}
-
-%Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
-%security classes.  APT: Don't think we need this figure.
 
@@
 
-  \caption{Example: assembly code alternatives for {\tt f} as an attacker.
-  }
+  \caption{Example: assembly code alternatives for {\tt f} as an attacker.}
   \label{fig:f}
@@
 (We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).)
-%
-
-\newcommand{\freebox}{\tikz \filldraw[fill=blue] (0,0) rectangle (10px,10px);}
-\newcommand{\pubbox}{\tikz \filldraw[fill=lightgray] (0,0) rectangle (10px,10px);}
-\newcommand{\objbox}{\tikz \filldraw[fill=yellow] (0,0) rectangle (10px,10px);}
-\newcommand{\sealbox}{\tikz \filldraw[fill=red] (0,0) rectangle (10px,10px);}
 
@@
   \end{tabular}
-  \caption{Execution of example up through the return from {\tt f}. In stack diagram, addresses increase to the right, stack grows to the left, and boxes represent 4-byte words.}
+  \caption{Execution of example up through the return from {\tt f}. In stack diagrams, addresses increase to the right, stack grows to the left, and boxes represent 4-byte words.}
 \label{fig:exec1}

### 2023-02-03 submission5/core.tex
Added micropolicy discussion

@@
 
-\bcp{Bits missing here!}
-
-\paragraph{Depth Isolation}
-
-\paragraph{Lazy Tagging and Clearing}
-
-\subsection{Enforcing \(\clei\) and \(\clec\)}
+We implement and test two micro-policies from Roessler and DeHon \cite{}:
+{\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
+They share a common structure: each function activation is assigned a ``color'' \(n\)
+representing its identity. Stack locations belonging to that activation are
+tagged \(\tagStackDepth{n}\), and when it is active, the tag on the
+program counter (PC tag) is \(\tagPCDepth{n}\). The tag \(\tagNoDepth\)
+is for stack locations that are currently unused.
+
+In \(DI\), \(n\) always corresponds to the depth of the stack when
+the function is called. A functions must initialize its entire frame on a call
+in order to tag it, and then clear it on a return.
+During normal execution, the micro-policy rules only permit load and
+store operations when the target memory is tagged {\em with the same depth}
+as the current {\PCname} tag, and store operations also when the
+target memory is \(\tagNoDepth\).
+
+In \(LTC\), rather than initialize the entire frame,
+a function will initialize each location's tag when it writes to it. It does
+not check if those writes are legal! If the PC tag is \(\tagPCDepth{n}\),
+then any stack location that recieves a store will be tagged \(\tagStackDepth{n}\).
+On a load, the micro-policy failstops if the value that it would load
+is tagged \(\tagStackDepth{n}\) or \(\tagNoDepth\) and the PC tag doesn't
+match \(n\).
+
+\paragraph{Blessed Sequences}
+
+To implement this discipline, {\em blessed instruction sequences} appear at
+the entry and exit of each function, which manipulate tags as just described
+while performing the usual tasks of saving/restoring the return address to/from
+the stack and adjusting the stack pointer. A blessed sequence also guarantees
+that the full sequence executes from the beginning---no jumping into the middle.
+
+\paragraph{Protecting Registers}
+
+Roessler and DeHon do not protect registers, but we consider registers to be
+an extension of the stack. In particular, \(\clri\) requires callee-saved
+registers to be saved and restored properly. We extend \(DI\) and \(LTC\)
+so that callee-saved registers, too, are tagged with the color of the 
+function that is using them. In \(DI\) they are tagged as part of the entry
+sequence, while in \(LTC\) they are tagged when a value is placed in them.
 
@@
 equivalent to its counterpart in depth isolation,
-and a version where stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
+and a version where stores succeed but fails to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
 It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
@@
 \section{Related Work}
-
-\sna{Not really sure what goes here anymore.
-  We've discussed Cerise and Nick and Andre quite a bit throughout.}
 

### 2023-02-03 submission5/core.tex
Halfway through figures

@@
 
-\apt{Add brief case for having formal guarantees.}
-Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
+Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
+other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
+or tested. Such properties become the specification against which enforcement can be validated;
+even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
+\emph{why} they fail.
+
+Of the mechanisms above, many are fundamentally ill-suited for offering formal guarantees,
 as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
@@
 \paragraph*{Towards Formal Properties for Stack Safety}
-
-Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
-other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
-or tested. Such properties become the specification against which enforcement can be validated;
-even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
-\emph{why} they fail.
 
@@
 
-\section{Key Ideas}
-\label{sec:ideas}
-
-\apt{Not a good place for this.}
-The paper is structured as follows. In \cref{sec:example}, we will walk through a function
+\paragraph{Organization}
+
+In the next section, we will give a brief overview of the structure of properties.
+Then in \cref{sec:example}, we will walk through a function
 call in a simple example machine and discuss how each of our properties applies to it,
 informally. In the process we discuss some motivation for the properties, from a security
-perspective. We consider a simple attacker, and discuss the harm that it can do to
-its caller. This perspective is essential to understanding how lazy enforcement can
-be safe: even if an attacker can overwrite some memory, if it cannot accomplish
-subvert the observable behavior of the caller, the system is safe.
+perspective.
 
@@
 its {\em security semantics}, and the stack safety properties built on these. 
-\Cref{sec:testing} describes the testing framework, and
+\Cref{sec:enforcement} describes the micro-policies that we test,
+\cref{sec:test} the testing framework itself, and
 \cref{sec:relwork,sec:future} discuss related and future work.
-The remainder of this section introduces ideas that will be important throughout.
-
-\paragraph*{Building a Property}
+
+%We consider a simple attacker, and discuss the harm that it can do to
+%its caller. This perspective is essential to understanding how lazy enforcement can
+%be safe: even if an attacker can overwrite some memory, if it cannot accomplish
+%subvert the observable behavior of the caller, the system is safe.
+
+\section{Building a Property}
+\label{sec:ideas}
 
@@
 \newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30px,30px) node[pos=.5]{#1};}
+\newcommand{\numbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
+\newcommand{\leftbox}[1][]{\tikz \filldraw[fill=blue!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
+\newcommand{\rightbox}[1][]{\tikz \filldraw[fill=orange!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{variants.png}
-  \caption{Integrity Violation\apt{Needs a machine-drawn version, as do the rest. Use colors consistently. No tags!}}
+  \centering
+  \[
+  \stackrel{\texttt{res}}{\numbox[0]}
+  \stackrel{\texttt{sens}}{\numbox[0]}
+  \stackrel{\texttt{sec}}{\numbox[5]}
+  \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+  %
+  \[\big\Downarrow\]
+  %
+  \[
+  \stackrel{\texttt{res}}{\numbox[0]}
+  \stackrel{\texttt{sens}}{\numbox[\bf 42]}
+  \stackrel{\texttt{sec}}{\numbox[5]}
+  \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+  %
+  \[\overbrace{
+    \stackrel{\texttt{res}}{\numbox[0]}
+    \stackrel{\texttt{sens}}{\numbox[\bf 42]}
+    \stackrel{\texttt{sec}}{\numbox[5]}
+    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}
+    \hspace{1cm}
+    \stackrel{\texttt{res}}{\numbox[0]}
+    \stackrel{\texttt{sens}}{\numbox[\bf 0]}
+    \stackrel{\texttt{sec}}{\numbox[5]}
+    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}}\]
+  \[\large\hookrightarrow 5 \hspace{1cm} \large\hookrightarrow 1\]
+
+  \caption{Integrity Violation}
   \label{fig:variant}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{variants2.png}
+    \centering
+    \[
+    \stackrel{\texttt{res}}{\numbox[0]}
+    \stackrel{\texttt{sens}}{\numbox[0]}
+    \stackrel{\texttt{sec}}{\numbox[5]}
+    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+    %
+    \[\overbrace{
+    \stackrel{\texttt{res}}{\leftbox[0]}
+    \stackrel{\texttt{sens}}{\leftbox[0]}
+    \stackrel{\texttt{sec}}{\leftbox[5]}
+    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
+    \hspace{1cm}
+    %
+    \stackrel{\texttt{res}}{\rightbox[2]}
+    \stackrel{\texttt{sens}}{\rightbox[3]}
+    \stackrel{\texttt{sec}}{\rightbox[4]}
+    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
+    \]
+    \[\large\Downarrow \hspace{1cm} \large\hookrightarrow 5 \hspace{0.5cm} \not \approx \hspace{0.5cm}
+    \large\hookrightarrow 4 \hspace{1cm} \large\Downarrow\]
+    \[
+    \stackrel{\texttt{res}}{\leftbox[\textbf{0}]}
+    \stackrel{\texttt{sens}}{\leftbox[0]}
+    \stackrel{\texttt{sec}}{\leftbox[5]}
+    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
+    \hspace{1cm}
+    %
+    \stackrel{\texttt{res}}{\rightbox[\textbf{2}]}
+    \stackrel{\texttt{sens}}{\rightbox[3]}
+    \stackrel{\texttt{sec}}{\rightbox[4]}
+    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
+    \]
+
   \caption{Internal Confidentiality Violation}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{variants3.png}
+    \centering
+    \[
+    \stackrel{\texttt{res}}{\numbox[0]}
+    \stackrel{\texttt{sens}}{\numbox[0]}
+    \stackrel{\texttt{sec}}{\numbox[5]}
+    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+    %
+    \[\overbrace{
+    \stackrel{\texttt{res}}{\leftbox[0]}
+    \stackrel{\texttt{sens}}{\leftbox[0]}
+    \stackrel{\texttt{sec}}{\leftbox[5]}
+    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
+    \hspace{1cm}
+    %
+    \stackrel{\texttt{res}}{\rightbox[2]}
+    \stackrel{\texttt{sens}}{\rightbox[3]}
+    \stackrel{\texttt{sec}}{\rightbox[4]}
+    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
+    \]
+    \[\large\Downarrow \hspace{1cm} \mathtt{a0} \leftbox[0] \hspace{1cm}
+    \rightbox[7] \mathtt{a0} \hspace{1cm} \large\Downarrow\]
+    \[
+    \stackrel{\texttt{res}}{\leftbox[0]}
+    \stackrel{\texttt{sens}}{\leftbox[0]}
+    \stackrel{\texttt{sec}}{\leftbox[5]}
+    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
+    \hspace{1cm}
+    %
+    \stackrel{\texttt{res}}{\rightbox[2]}
+    \stackrel{\texttt{sens}}{\rightbox[3]}
+    \stackrel{\texttt{sec}}{\rightbox[4]}
+    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
+    \]
+    \[\mathtt{a0} \leftbox[0] \hspace{1cm}
+    \mathtt{a0} \rightbox[0]\]
   \caption{Return-time Confidentiality Violation}

### 2023-02-03 submission5/core.tex
More diagrams, working through comments

@@
 
-% APT: I don't think we need this.
-%% \begin{figure}
-%%   \begin{tabular}{| l | l | l |}
-%%     \hline
-%%     Set / & Names & Purpose \\
-%%     Class & & \\
-%%     \hline
-%%     \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
-%%     \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
-%%     & {\tt a2} -- {\tt a7} & Caller-saved args \\
-%%     & {\tt ra} & Return Address \\
-%%     \hline
-%%     \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
-%%     \(\sealed\) & {\tt sp} & Stack Pointer \\
-%%     \hline
-%%     \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
-%%     \(\public\) & {\tt tp} & Thread Pointer \\
-%%     \hline
-%%   \end{tabular}
-%%   \caption{RISC-V integer register set}
-%%   \label{fig:RISCVregs}
-%% \end{figure}
-
 We now give a formal description of our machine model, security semantics,
-and properties. Our definitions abstract over the details of the target machine
-architecture and ABI, the set of security-relevant operations, the rules that
-govern evolution of the security context, the set of observable events, and
-a notion of value compatibility.
-\apt{Not quite true: the property definitions require a notion of
-  call entry.}
+and properties. Our definitions abstract over: (1) the details of  the target machine
+architecture and ABI, (2) the set of observable events, and (3) a notion of value compatibility.
+They can be extended to a more sophisticated machine model by expanding the set of
+security-relevant events and defining
+
 \apt{Any more? Give a more listlike presentation?
@@
   presentation and in the various extensions, including the appendix.}
+\sna{Need to come back to this.}
 
@@
 and lift it element-wise to states.
-Values are compatible if their non-payload information is identical, 
-even though their payloads may differ.
+In the tagged setting, values are compatible if their tags are identical.
 \apt{This is still not very convincing!
@@
   too strong, right? (OK for tags, but not, e.g. capabilities?}
+\sna{Yeah, I don't think we can say something universal about when values should be
+  considered compatible.}
 
@@
 all other locations to \(\public\), and registers based on which set they
-belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
-\apt{This doesn't quite match example, where there is magically an initial argument register labeled \(\active\).}
+belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved except for those
+that contain arguments at the start of execution, which are \(\object\), and \(\public\) otherwise.
 
@@
 is the same. That is, we compare up to deletion of \(\tau\) events.
-\apt{Didn't we use to say something about the (arguably) odd nature of this definition?}
+Note that this results in an infinite silent trace being similar to
+any trace. So, a trace that silently diverges due to a failstop will
+fulfill this property vacuously.
 

### 2023-02-03 submission5/core.tex
Quick pass over III

@@
 of small code examples, using a simplified set of security-relevent operations for
-or calls, returns, and private allocations.
+calls, returns, and private allocations.
 Figure \ref{fig:main} gives C code and corresponding compiled 32-bit RISC-V code
@@
 In Figure \ref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
-by returning to the correct program counter but with the wrong stack pointer.
+by returning to the correct program counter but with the wrong stack pointer.%
 \footnote{We pad the last two variants with {\tt nop}s just so that all the
@@
 essentially using the caller's memory as scratch space, but the caller never seeing any change.
+\rb{(Indeed, this even applies to callee-saved registers, if it is worth mentioning)}
 
@@
 Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\), with the attacker from
-Figure \ref{subfig:direct}. We take a variant state over the set of element(s) that are
+Figure \ref{subfig:direct}. We take a variant state over the set of elements that are
 \(\sealed\) in \(V_2\) (see Figure \ref{fig:variant2}.)
@@
 
-To illustrate the issues, note that {\tt f} has also stores a 0 below the stack pointer.
+To illustrate the issues, note that {\tt f} has also stored a 0 below the stack pointer.
 %

### 2023-02-03 submission5/core.tex
Quick pass over IV

@@
 
-The security semantics operates in parallel with the 
+The security semantics operates in parallel with the machine.
 Each state element (memory word or register) is given a \emph{security class} 
@@
 security-relevant operations. Formally, we combine each machine state with a context
-to create a {\it combined state} \(s = (\mach,\context)\) and a lift the transition
+to create a {\it combined state} \(s = (\mach,\context)\) and lift the transition
 to \(\stepstounder{}\) on combined states. 
@@
 event. Other events might represent certain function calls (i.e., system calls)
-or writes to special addresses representing mmapped regions.
+or writes to special addresses representing memory-mapped regions.
 A {\em trace} is a nonempty, finite or infinite sequence
@@
 
-We are particular interested in traces that end just after a function returns.
+We are particularly interested in traces that end just after a function returns.
 We define these in terms of the depth \(d\) of the security context's call stack \(\sigma\).

### 2023-02-03 submission5/core.tex
ref to cref

@@
 calls, returns, and private allocations.
-Figure \ref{fig:main} gives C code and corresponding compiled 32-bit RISC-V code
+\Cref{fig:main} gives C code and possible corresponding compiled 32-bit RISC-V code
 for a function {\tt main}, which
@@
 to leak {\tt secret}. It might do so in a number of ways, shown as snippets of
-assembly code in Figure \ref{fig:f}.
+assembly code in \cref{fig:f}.
 %
 Leakage is most obviously viewed as a violation of {\tt main}'s {\it confidentiality}.
-In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
+In \cref{subfig:direct}, {\tt f} takes an offset from the stack
 pointer, accesses {\tt secret}, and directly outputs it. But more
@@
 can instead return that value so that {\tt main} stores it to {\tt out},
-as in Figure \ref{subfig:indirect}.
+as in \cref{subfig:indirect}.
 %
@@
 with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
-(Figure \ref{subfig:integrity}).
+(\cref{subfig:integrity}).
 Attacks of this kind do not violate {\tt main}'s confidentiality, but
 rather its {\it integrity}.
-In Figure \ref{subfig:WBCF}, the attacker arranges to return to the
+In \cref{subfig:WBCF}, the attacker arranges to return to the
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
@@
 %
-In Figure \ref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
+In \cref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
 by returning to the correct program counter but with the wrong stack pointer.%
 \footnote{We pad the last two variants with {\tt nop}s just so that all the
-snippets have the same length, which simplifies some later explanations.}
+snippets have the same length, which keeps the step numbering uniform in~\cref{fig:exec1}.}
 
@@
 The security semantics for this program is based
-on the security-relevant events noted in the right columns of Figures~\ref{fig:main}
-and~\ref{fig:f}, namely execution of instructions that allocate or deallocate space,
+on the security-relevant events noted in the right columns of \cref{fig:main,fig:f},
+namely execution of instructions that allocate or deallocate space,
 make a call, or make a return.
@@
 which consists of a view \(V\) and a stack \(\sigma\) of pending activations' views.
-Figure \ref{fig:exec1} shows how the security context evolves over the first few
+\Cref{fig:exec1} shows how the security context evolves over the first few
 steps of the program.  (The formal details of the security semantics are described in
@@
 integrity, and correct control flow in such a way that many classes of
-bad behavior, including the attacks in Figure~\ref{fig:f}, are
+bad behavior, including the attacks in \cref{fig:f}, are
 detected as security violations.
@@
 and the stack pointer should be the same as it was before the call.
-Both of these are essential. In Figure \ref{subfig:WBCF}, the attacker adds
+Both of these are essential. In \cref{subfig:WBCF}, the attacker adds
 16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
 {\tt secret}.
-In Figure \ref{subfig:WBCF2}, the attacker returns with \(\SP' = 998\) instead of the
+In \cref{subfig:WBCF2}, the attacker returns with \(\SP' = 998\) instead of the
 correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame,
@@
 return. This time the condition applies to all of the memory in the function's
-frame. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
+frame. In \cref{fig:exec1} we see the lifecycle of an allocated frame:
 upon allocation, the view labels it \(\object\), and when a call is made, it instead
@@
 of values written to {\tt out}.
-The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
-which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
+The example in \cref{subfig:integrity} modifies the value of {\tt sensitive},
+which is \(\sealed\). \Cref{fig:variant} shows the state just after the call at step 5,
 assuming that {\tt sec} is 5. Similar to \(\wbcf\), we define
@@
 Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\), with the attacker from
-Figure \ref{subfig:direct}. We take a variant state over the set of elements that are
-\(\sealed\) in \(V_2\) (see Figure \ref{fig:variant2}.)
+\cref{subfig:direct}. We take a variant state over the set of elements that are
+\(\sealed\) in \(V_2\) (see \cref{fig:variant2}.)
             [TODO: fix figure to avoid varying the entire stack.]
@@
 
-But, in Figure \ref{subfig:indirect}, we also saw an attacker that exfiltrated the secret
+But, in \cref{subfig:indirect}, we also saw an attacker that exfiltrated the secret
 by reading it and then returning it, in a context where the caller would output the returned
-value. Figure \ref{fig:variant3} shows the behavior of the same variants under this attacker,
+value. \Cref{fig:variant3} shows the behavior of the same variants under this attacker,
 but in this case, there is no output during the call. Instead the value of {\tt secret} is
@@
 
-\cref{sec:example} describes informally how the security context evolves as the system performs
+\Cref{sec:example} describes informally how the security context evolves as the system performs
 security-relevant operations. Formally, we combine each machine state with a context
@@
 The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
-Table~\ref{tab:bug-table}, along with the average number of tests
+\cref{tab:bug-table}, along with the average number of tests
 it took to find the failure. Experiments were run in a desktop

### 2023-02-03 submission5/core.tex
Fixed some stack passing discussion

--- 2023-02-03 submission5/core.tex
@@
 of small code examples, using a simplified set of security-relevent operations for
-calls, returns, and private allocations.
+or calls, returns, and private allocations.
 Figure \ref{fig:main} gives C code and corresponding compiled 32-bit RISC-V code
@@
 In Figure \ref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
-by returning to the correct program counter but with the wrong stack pointer.%
+by returning to the correct program counter but with the wrong stack pointer.
 \footnote{We pad the last two variants with {\tt nop}s just so that all the
@@
 essentially using the caller's memory as scratch space, but the caller never seeing any change.
-\rb{(Indeed, this even applies to callee-saved registers, if it is worth mentioning)}
 
@@
 Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\), with the attacker from
-Figure \ref{subfig:direct}. We take a variant state over the set of elements that are
+Figure \ref{subfig:direct}. We take a variant state over the set of element(s) that are
 \(\sealed\) in \(V_2\) (see Figure \ref{fig:variant2}.)
@@
 
-To illustrate the issues, note that {\tt f} has also stored a 0 below the stack pointer.
+To illustrate the issues, note that {\tt f} has also stores a 0 below the stack pointer.
 %
@@
 Our notion of non-interference involves comparing the traces of such
-\(\components\)-variants. We use this to define sets of irrelevant elements,
-and secrets in general.\apt{last phrase is cryptic}
-
-\definition The \emph{difference set} of two machine states \(\mach\) and \(\mach'\), written \(\Delta(\mach,\mach')\),
+\(\components\)-variants. We use this to define sets of irrelevant elements.
+
+\definition The \emph{difference set} of two machine states \(\mach\) and \(\mach'\),
+written \(\Delta(\mach,\mach')\),
 is the set of elements \(\component\) such that \(\mach[\component] \not = \mach'[\component]\).
@@
 arguments and return values. This is frequently not the case in a realistic
-setting. Arguments may be passed on the stack (e.g. because there are too many
-to pass in registers), as an implementation
-of variadic arguments, or because they are reference objects \apt{??} that inherently have
-pass-by-reference semantics. The caller may also take the address of a local object and
-pass that as a pointer to its callee.
+setting. Arguments may be passed on the stack because there are too many
+to pass in registers, as an implementation of variadic arguments, or
+because they are composite types that inherently have
+pass-by-reference semantics. The caller may also pass an object by reference
+in the C++ style, or take its address and pass it as a pointer.
 
@@
 
-Using this mechanism, we can describe an argument being passed on the stack
-by giving its \(\SP\)-relative offset,
-with the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
-In the case of pass-by-reference semantics, if the compiler designates register
-\(\reg\) as holding the reference, then the operations is
-\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).\apt{This is unclear to me. 
-  The compiler knows where the cbr parameter lives in the caller's stack
-  frame, so why wouldn't it use a sp-relative spec for that too? Does it have
-  to do with passing a CBR parameter further down to another callee as you allude
-  to in the next para?}
-
-In both cases, the immediate callee retains access to the passed object, but deeper
-callees will not---unless the object is passed-by-reference further down the stack.
-On the other hand, in the case of a caller explicitly taking the address of an
-object and passing it as a pointer, we have no way obvious way to limit the scope
-of such sharing.
-
-Our primary \apt{again...} machine treats this scenario as the caller allocating a ``public''
-object that can be accessed by anyone until it's deallocated. We extend the
-\(\mathbf{alloc}\) operation with a boolean flag, where {\bf t} indicates
-that the allocation is public, and {\bf f} is private.
+Using this mechanism, an argument passed on the stack via its \(\SP\)-relative offset
+uses the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
+C++-style reference passing that places the reference in \(\reg\) uses
+\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
+In the second case, the reference can be passed further down the stack, but otherwise
+only the immediate callee gains access.
+
+On the other hand, if the address of an object is taken directly and passed it as a
+pointer, we have no way obvious way to limit the scope of such sharing.
+We treat this scenario as the caller allocating a ``public''
+object that recieves no protection. We extend the \(\mathbf{alloc}\) operation
+with a boolean flag, where {\bf t} indicates
+that the allocation is public, and {\bf f} that it is private.
 In the likely event that the space for multiple objects is allocated in a single step,
 that step can make multiple allocation operations, each labeled appropriately.
-Allocating public objects affects the call stack very similarly to private ones,
-but instead of labeling them \(\object\) they become \(\public\), so they are
+Public objects are labeled \(\public\) rather than \(\object\), so they are
 never sealed at a call (\cref{sfig:publicalloc}).
@@
 
-Since the caller will not be returned to, it does not need integrity, though
+Since the caller will not be returned to, it does not need integrity, but
 it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
-than \(\sealed\) to express this. Of course, some objects may be passed as arguments
-in place\apt{??}, easily modeled by including them in \(\overline{sa}\). More sophisticated
-behaviors may require sequences of {\bf dealloc}, {\bf tailcall}, then {\bf alloc}.
+than \(\sealed\) to express this.
 

### 2023-02-03 submission5/core.tex
wibbles in 3
edits, comments in 5

@@
     \vspace{\abovedisplayskip}
-    \(\textnormal{\large \dots} \pubbox[30px] \!\! \stackrel{\textsc{\large sp}}{\textbox[\tt res]}
-    \stackrel{\textsc{\large 4(sp)}}{\textbox[\tt sens]}
-    \stackrel{\textsc{\large 8(sp)}}{\textbox[\tt sec]}
-    \stackrel{\textsc{\large 12(sp)} \hfill}{\textbox[\(\mathtt{ra_1}\)]\textbox[\(\mathtt{ra_2}\)]}
+    \(\textnormal{\large \dots} \pubbox[30px] \!\! \stackrel{\textsc{\normalsize sp}}{\textbox[\tt res]}
+    \stackrel{\textsc{\normalsize 4(sp)}}{\textbox[\tt sens]}
+    \stackrel{\textsc{\normalsize 8(sp)}}{\textbox[\tt sec]}
+    \stackrel{\textsc{\normalsize 12(sp)} \hfill}{\textbox[\(\mathtt{ra_1}\)]\textbox[\(\mathtt{ra_2}\)]}
     \textnormal{\large \dots}\)
@@
 {\tt secret}.
-In \cref{subfig:WBCF2}, the attacker returns with \(\SP' = 998\) instead of the
+In \cref{subfig:WBCF2}, the attacker returns with \(\SP' = 988\) instead of the
 correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame,
@@
   \hline
-  res & sens & sec & ra & ra \\
+  {\tt res} & {\tt sens} & {\tt sec} & \({\mathtt ra_1}\) & \({\mathtt ra_2}\) \\
   \hline
@@
 the principals should occur only through the state elements that are designated for the
-purpose: those which we label \(\public\) and (in the presence of memory sharing) \(\object\).
-
-Applying this intuition using our framework, \(\clec\) turns out to resemble
+purpose: those which we label \(\public\) and (in the presence of memory sharing) \(\object\).\apt{Are these backwards?}
+
+Applying this intuition using our framework, \emph{callee confidentiality} (\(\clec\)) turns out to resemble
 \(\clri\), extended to every element that is not marked \(\object\) at call-time. The callee's
@@
 execution, and which are not part of the interface with the caller. At return, those
-elements should become irrelevant.
-
-Similarly, in \(\clei\), only elements marked \(\object\) at the call should influence the
+elements should become irrelevant to the subsequent behavior of the caller.
+
+Similarly, in \emph{callee integrity} (\(\clei\)), only elements marked \(\object\) at the call should influence the
 behavior of the callee. This is the same principle as \(\clrc\), extended from \(\sealed\)
-elements to all non-\(\object\) ones.
+elements to all non-\(\object\) ones. \apt{Still not obvious why we're calling this an integrity property.}
 
@@
 arranged to show their commonalities and distinctions. Each definition gives a criterion
-quantified over states \(s\) that immediately follow call (or tail-call) steps.
+quantified over states \(s\) that immediately follow call (or tail-call) steps.\apt{No, we don't
+  want to mention tail-calls here. But we do need to explain how the interpretation of the
+  property needs to change when we add tail-calls later.}
 If an execution can reach a state \(s'\) such that \(s' \stepstounder{\overline{\psi}} s\)
@@
 
-Using this mechanism, an argument passed on the stack via its \(\SP\)-relative offset
-uses the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
-C++-style reference passing that places the reference in \(\reg\) uses
-\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
-In the second case, the reference can be passed further down the stack, but otherwise
-only the immediate callee gains access.
+Using this mechanism, an call-by-value argument passed on the stack at an \(\SP\)-relative offset
+is specified by the triple \((\SP, \mathit{off}, \mathit{sz})\).
+In this case, only the immediate callee gains access to the argument location.
+A C++-style call-by-reference argument where the reference is passed in \(\reg\)
+is instead specified by \((\reg, 0, \mathit{sz})\), which allows
+the reference to be passed further down the stack.
+\apt{What if the reference is itself passed on the stack?}
+%In this second case, the reference can be passed further down the stack, but otherwise
+%only the immediate callee gains access. APT:
 
@@
 We treat this scenario as the caller allocating a ``public''
-object that recieves no protection. We extend the \(\mathbf{alloc}\) operation
+object that receives no protection. We extend the \(\mathbf{alloc}\) operation
 with a boolean flag, where {\bf t} indicates
 that the allocation is public, and {\bf f} that it is private.
-In the likely event that the space for multiple objects is allocated in a single step,
+If space for multiple objects is allocated in a single step,
 that step can make multiple allocation operations, each labeled appropriately.

### 2023-02-03 submission5/core.tex
pass on 6

@@
 
-We implement and test two micro-policies from Roessler and DeHon \cite{}:
+We implement and test two micro-policies from Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
 {\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
@@
 representing its identity. Stack locations belonging to that activation are
-tagged \(\tagStackDepth{n}\), and when it is active, the tag on the
-program counter (PC tag) is \(\tagPCDepth{n}\). The tag \(\tagNoDepth\)
-is for stack locations that are currently unused.
+tagged \(\tagStackDepth{n}\), and while the activation is running, the tag on the
+program counter (PC tag) is \(\tagPCDepth{n}\). Stack locations not part of
+any activation are tagged \(\tagNoDepth\).
 
 In \(DI\), \(n\) always corresponds to the depth of the stack when
-the function is called. A functions must initialize its entire frame on a call
-in order to tag it, and then clear it on a return.
+the function is called. A function must initialize its entire frame upon entry
+in order to tag it, and then clear the frame before returning.
 During normal execution, the micro-policy rules only permit load and
 store operations when the target memory is tagged {\em with the same depth}
-as the current {\PCname} tag, and store operations also when the
-target memory is \(\tagNoDepth\).
-
-In \(LTC\), rather than initialize the entire frame,
-a function will initialize each location's tag when it writes to it. It does
+as the current {\PCname} tag, or, for store operations, if the target memory
+is tagged \(\tagNoDepth\).
+
+In \(LTC\), rather than initialize the entire frame at entry,
+a function initializes each location's tag when it writes to it. It does
 not check if those writes are legal! If the PC tag is \(\tagPCDepth{n}\),
 then any stack location that recieves a store will be tagged \(\tagStackDepth{n}\).
-On a load, the micro-policy failstops if the value that it would load
-is tagged \(\tagStackDepth{n}\) or \(\tagNoDepth\) and the PC tag doesn't
-match \(n\).
-
-\paragraph{Blessed Sequences}
+On a load, the micro-policy failstops if the source memory location 
+is tagged \(\tagNoDepth\) or \(\tagStackDepth{n}\) for some \(n\) that
+doesn't match the PC tag.
+
+\paragraph*{Blessed Sequences}
 
@@
 while performing the usual tasks of saving/restoring the return address to/from
-the stack and adjusting the stack pointer. A blessed sequence also guarantees
-that the full sequence executes from the beginning---no jumping into the middle.
-
-\paragraph{Protecting Registers}
+the stack and adjusting the stack pointer. A blessed sequence uses further tags
+to guarantee that the full sequence executes from the beginning---no jumping into the middle.
+
+\paragraph*{Protecting Registers}
 
@@
 sequence, while in \(LTC\) they are tagged when a value is placed in them.
+\apt{Presumably R\&dH don't care about registers because they already have
+  the compiler in the TCB.  We should not (perhaps earlier than here!) that
+  we have a stronger attacker model since we don't require the callee code
+  to come from a trusted  compiler.}
 

### 2023-02-03 submission5/core.tex
abstract + intro

@@
 abstraction of functions, in all of the forms that they can take.
-The protean nature of the functional abstraction makes stack safety
+This protean nature of the functional abstraction makes stack safety
 difficult to specify.
@@
 both caller- and callee-saved registers, arguments passed on the stack,
-tail-call elimination, \ifexceptions exceptions \fi. They are modular by
+\ifexceptions tail-call elimination, and exceptions.
+\else and tail-call elimination. \fi They are modular by
 design, in order to be further extensible to other features.
@@
 
-Indeed, the call stack is an ancient~\cite{one06phrack} and
+Indeed, the call stack is an ancient~\cite{phrack96:smashingthestack} and
 perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
@@
 the return address, and memory corruption more generally.
-
 %
@@
 Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
-other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
+other, similar attacks. To avoid this \leo{game of whack-a-mole}, we seek formal properties of safe behavior that can be proven
 or tested. Such properties become the specification against which enforcement can be validated;
@@
 They are related to one another as seen in
-\cref{fig:propertydiagram}: \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
+\cref{fig:propertydiagram}:\leo{is this supposed to be \cref{tab:props}?} \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
 \(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
@@
 describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
-to the combination of \(\clri\) and \(\clrc\). The follow-up work by 
-Georges et al. \cite{Georges22:TempsDesCerises} cites an early preprint of our work
-as inspiration for extending their formal discussion to include confidentiality.
+to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by 
+Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
 When checking if a preliminary version of these properties applied to
@@
 for the QuickChick library and to ensure that the formalizations are unambiguous.
-Proofs are future work.
+Formal proofs are left as future work.
 
@@
 
-\paragraph{Organization}
+\paragraph*{Organization}
 
@@
 
-\newcommand{\freebox}[1][10px]{\tikz \filldraw[fill=blue] (0,0) rectangle (#1,#1);}
-\newcommand{\pubbox}[1][10px]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (#1,#1);}
-\newcommand{\objbox}[1][10px]{\tikz \filldraw[fill=yellow] (0,0) rectangle (#1,#1);}
-\newcommand{\sealbox}[1][10px]{\tikz \filldraw[fill=red] (0,0) rectangle (#1,#1);}
-
-\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30px,30px) node[pos=.5]{#1};}
-\newcommand{\numbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
-\newcommand{\leftbox}[1][]{\tikz \filldraw[fill=blue!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
-\newcommand{\rightbox}[1][]{\tikz \filldraw[fill=orange!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
+\newcommand{\freebox}[1][10pt]{\tikz \filldraw[fill=blue] (0,0) rectangle (#1,#1);}
+\newcommand{\pubbox}[1][10pt]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (#1,#1);}
+\newcommand{\objbox}[1][10pt]{\tikz \filldraw[fill=yellow] (0,0) rectangle (#1,#1);}
+\newcommand{\sealbox}[1][10pt]{\tikz \filldraw[fill=red] (0,0) rectangle (#1,#1);}
+
+\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30pt,30pt) node[pos=.5]{#1};}
+\newcommand{\numbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
+\newcommand{\leftbox}[1][]{\tikz \filldraw[fill=blue!20] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
+\newcommand{\rightbox}[1][]{\tikz \filldraw[fill=orange!20] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
 
@@
     \vspace{\abovedisplayskip}
-    \(\textnormal{\large \dots} \pubbox[30px] \!\! \stackrel{\textsc{\normalsize sp}}{\textbox[\tt res]}
+    \(\textnormal{\large \dots} \pubbox[30pt] \!\! \stackrel{\textsc{\normalsize sp}}{\textbox[\tt res]}
     \stackrel{\textsc{\normalsize 4(sp)}}{\textbox[\tt sens]}

### 2023-02-03 submission5/core.tex
fix some unresolved refs, a few notes

--- 2023-02-03 submission5/core.tex
@@
 
-In \cref{sec:machine} we formalize the machine model, 
+In \cref{sec:formal} we formalize the machine model, 
 its {\em security semantics}, and the stack safety properties built on these. 
 \Cref{sec:enforcement} describes the micro-policies that we test,
-\cref{sec:test} the testing framework itself, and
+\cref{sec:testing} the testing framework itself, and
 \cref{sec:relwork,sec:future} discuss related and future work.
@@
 
-A security extends a machine
+A security semantics extends a machine
 with additional context about the identity of current and pending
@@
 on future states, e.g. of the form ``when control returns to the current function...'',
-or relations on traces of future execution (hyper-properties).
+or as relations on traces of future execution (hyper-properties).
 
@@
 %\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
-
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
-\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
+\cref{tab:psi}. Those that we explore in detail in \cref{sec:example,sec:formal}
 are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In \cref{sec:machine}), we cover additional operations that add parameters to the examples,
+In \cref{sec:extensions}, we cover additional operations and and parameters 
 to support various forms of memory sharing \ifexceptions , exceptions, \fi
-and tail-call elimination. These are shown in light gray.
+and tail-call elimination; these are shown in light gray. These are also
+the operations we test in \cref{sec:testing}.
 We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
-based on pointer provenance in \cref{app:ptr}, represented by the dark grey
+based on pointer provenance in \cref{app:ptr}, represented by the dark gray
 operations, but we do not test this.
@@
 \end{center}
-  \caption{Security-relevant Operations and their Parameters}
+\caption{Security-relevant operations and their parameters.}
   \label{tab:psi}
@@
 
-To characterize confidentiality, we borrow the idea of \emph{variant} states
+To characterize confidentiality\apt{and sometimes integrity}, we borrow the idea of \emph{variant} states
 from the theory of non-interference. Non-interference is a way of expressing
@@
 In this section we introduce our security property definitions by means
-of small code examples, using a simplified set of security-relevent operations for
+of small code examples, using a simplified set of security-relevant operations for
 calls, returns, and private allocations.
@@
 steps of the program.  (The formal details of the security semantics are described in
-\cref{sec:machine}, and the context evolution rules are formalized in \cref{fig:basicops}.)
+\cref{sec:formal}, and the context evolution rules are formalized in \cref{fig:basicops}.)
 Execution begins at the start of {\tt main}, where the program counter (\(\PCname\)) is zero,
@@
   \hline
-  {\tt res} & {\tt sens} & {\tt sec} & \({\mathtt ra_1}\) & \({\mathtt ra_2}\) \\
+  {\tt res} & {\tt sens} & {\tt sec} & \(\mbox{\tt ra}_1\) & \(\mbox{\tt ra}_2\) \\
   \hline
@@
 \section{Formalization}
-\label{sec:machine}
+\label{sec:formal}
 
@@
 \section{Expanding the Machine}
+\label{sec:extensions}
 
@@
 \section{Enforcement}
+\label{sec:enforcement}
 
@@
   an initial policy state, tagging instructions and the stack regions appropriately
-  for the micro-policy (see \cref{subsec:policies})
+  for the micro-policy (see \cref{sec:enforcement})
 (3)
@@
 \section{Related Work}
+\label{sec:relwork}
 

### 2023-02-03 submission5/core.tex
wibbles

@@
 
-Our motivating enforcement mechanism,
+Our motivating enforcement mechanisms,
 the ``lazy'' stack safety micro-policies proposed by
@@
 Note that this results in an infinite silent trace being similar to
-any trace. So, a trace that silently diverges due to a failstop will
+any trace. So, a trace that silently diverges due to a failstop\apt{?? haven't explained this!}  will
 fulfill this property vacuously.
@@
 \(\hookrightarrow (\mach,\context)  \eqsim \hookrightarrow (\nach,\context)\).
+
+\apt{needs small transition}
 
@@
 
-The system we've modeled so far has been very simple, but our properties
+The system we model in \cref{sec:example,sec:formal} is very simple, but our properties
 support an important dimension of generality: adding features.\apt{Just what needs to be changed
@@
 \apt{Presumably R\&dH don't care about registers because they already have
-  the compiler in the TCB.  We should not (perhaps earlier than here!) that
+  the compiler in the TCB.  We should note (perhaps earlier than here!) that
   we have a stronger attacker model since we don't require the callee code

### 2023-02-03 submission5/core.tex
edits in 5

@@
 
-On the other hand, if the address of an object is taken directly and passed it as a
-pointer, we have no way obvious way to limit the scope of such sharing.
-We treat this scenario as the caller allocating a ``public''
-object that receives no protection. We extend the \(\mathbf{alloc}\) operation
+If the address of an object is taken directly and passed as a
+pointer, we simply classify the object as ``public'' and give it no protection against
+access by other functions. We extend the \(\mathbf{alloc}\) operation
 with a boolean flag, where {\bf t} indicates
@@
 never sealed at a call (\cref{sfig:publicalloc}).
+Providing more fine-grained control over sharing is desirable, but
+seems to require a considerably more complex model; \cref{app:ptr} describes one
+approach, based on capabilities.
 
@@
 than \(\sealed\) to express this.
+
+\apt{good place to detail the needed changes in interpretation of \cref{tab:props}.}
 

### 2023-02-03 submission5/core.tex
Related work

@@
 the principals should occur only through the state elements that are designated for the
-purpose: those which we label \(\public\) and (in the presence of memory sharing) \(\object\).\apt{Are these backwards?}
-
-Applying this intuition using our framework, \emph{callee confidentiality} (\(\clec\)) turns out to resemble
-\(\clri\), extended to every element that is not marked \(\object\) at call-time. The callee's
-internal behavior is represented by those elements that change over the course of its
-execution, and which are not part of the interface with the caller. At return, those
-elements should become irrelevant to the subsequent behavior of the caller.
-
-Similarly, in \emph{callee integrity} (\(\clei\)), only elements marked \(\object\) at the call should influence the
-behavior of the callee. This is the same principle as \(\clrc\), extended from \(\sealed\)
-elements to all non-\(\object\) ones. \apt{Still not obvious why we're calling this an integrity property.}
+purpose: those labeled \(\public\) and \(\object\).
+
+Applying this intuition using our framework, \emph{callee confidentiality} (\(\clec\))
+turns out to resemble \(\clri\), extended to every element that is not marked \(\object\)
+or \(\public\) at call-time. The callee's internal behavior is represented by those
+elements that change over the course of its execution, and which are not part of the
+interface with the caller. At return, those elements should become irrelevant to the
+subsequent behavior of the caller.
+
+Similarly, in \emph{callee integrity} (\(\clei\)), only elements marked \(\object\)
+or \(\public\) at the call should influence the behavior of the callee. It may seem
+odd to call this integrity, as the callee does not have a private state. But
+an erroneous callee that performs a read-before-write within its stack
+frame, or which uses a non-argument register without initializing it, is vulnerable
+to its caller seeding those elements with values that will change its behavior.
+The fact that well-behaved callees have integrity by definition is probably why
+callee integrity is not typically discussed.
 
@@
 arranged to show their commonalities and distinctions. Each definition gives a criterion
-quantified over states \(s\) that immediately follow call (or tail-call) steps.\apt{No, we don't
-  want to mention tail-calls here. But we do need to explain how the interpretation of the
-  property needs to change when we add tail-calls later.}
+quantified over states \(s\) that immediately follow call steps.
 If an execution can reach a state \(s'\) such that \(s' \stepstounder{\overline{\psi}} s\)
@@
 As a shorthand, we write that each property is defined
-by a criterion that must hold ``for all call targets \(s\).''
-
-\(\wbcf\), \(\clri\), and \(\clec\) are each straightforward: they define predicates that must
-hold on return from the call. \(\clrc\) and \(\clei\) both compare the execution of the
-call state with that of a variant state, and are concerned with both their internal and
-their return-time behavior. So, for some \(\nach\) that is an appropriate variant,
-all states must fulfill an ``internal'' clause that their traces are similar,
-and a ``return-time'' clause, that any potentially corrupted memory at their returns
-is irrelevant.
-\apt{This paragraph needs major expansion, to paraphrase the definitions informally,
-  and TO SAY MUCH MORE ABOUT THE CALLEE PROPERTIES, which are not intuitive at all (to me)!}
+by a criterion that must hold ``for all call targets \(s\).'' In the case of \(\wbcf\),
+we also need to know about the callee, so we refer to the ``call step \(s \stepstounder{} s'\).
+
+\paragraph*{\(\wbcf\)}
+Given a call step \((\mach,(V,\sigma)) \stepstounder{} (\mach',(V',\sigma'))\),
+we define the predicate \(\ret\) to hold on states \(\mach''\)
+whose stack pointer matches that of \(\mach\)
+and whose program counter is at the next instruction. We enjoy \(\wbcf\) if
+\(\ret\) holds on return from \(\mach'\).
+
+\paragraph*{\(\clri\)}
+When our call target is \((\mach,(V,\sigma))\), we define the predicate \(\intProp\) to hold
+on states \(\mach'\) if any elements that are both sealed in \(V\) and in the difference
+set between \(\mach\) and \(\mach'\) are irrelevant. We enjoy \(\clri\) if \(\intProp\)
+holds on return from \(\mach\).
+
+\paragraph*{\(\clrc\)}
+When our call target \((\mach,(V,\sigma))\), we begin by taking an arbitrary \(\nach\)
+that is a \(\components\)-variant of \(\mach\), where \(\components\) is the set of sealed elements
+in \(V\). We require that two clauses hold. On line a, the behavior of a trace from
+\((\mach,(V,\sigma))\) up to its return must match that of \((\nach,(V,\sigma))\).
+On line b, we define a relation \(\confProp\) that relates states \(\mach'\) and \(\nach'\)
+if their corrupted set (relative to \(\mach\) and \(\nach\)) is irrelevant, and require
+that it hold on the returns from \((\mach,(V,\sigma))\) and \((\nach,(V,\sigma))\).
+We enjoy \(\clrc\) if both clauses hold.
+
+\paragraph*{\(\clec\)}
+We consider the callee's private behavior to be any changes that it makes to the state
+outside of legitimate channels---elements marked \(\object\) or \(\public\). The remainder
+should be kept secret, which is to say, irrelevant. Similar to \(\clri\), given a call target
+\((\mach,(V,\sigma))\), we define a predicate \(\cconfProp\) to hold
+on states \(\mach'\) if the difference set between \(\mach\) and \(\mach'\), that is
+not \(\object\) or \(\public\), is irrelevant. If \(\cconfProp\) holds on return,
+we enjoy \(\clec\).
+
+\paragraph*{\(\clei\)}
+Callee integrity means that the caller does not influence it outside of legitimate
+channels. The caller's influence can be seen internally, or in corrupted data on return,
+just like the caller's secrets would under \(\clrc\). So, for a call target
+\((\mach,(V,\sigma))\), we take an arbitrary \(\nach\) that is a \(\components\)-variant
+of \(\mach\), where \(\components\) is the set all elements that are not \(\object\)
+or \(\public\). The remainder of the property is identical to \(\clrc\).
 
@@
     1
-    & \(\wbcf \triangleq\) & \((|\sigma| \uparrow \ret) ~ ????\)
-    & \(\textnormal{ where } \ret ~ \mach' \triangleq \)
-    \(\mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\)
+    & \(\wbcf \triangleq\) & \((|\sigma'| \uparrow \ret) ~ (\mach', (V',\sigma'))\)
+    & \(\textnormal{ where } \ret ~ \mach'' \triangleq \)
+    \(\mach''[\SP] = \mach[\SP]\)
 %    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called}\) \\
-    & \(\textnormal{ for all call targets } (\mach,(V,\sigma))\) \\
+    & \(\textnormal{ for all calls } (\mach,(V,\sigma)) \stepstounder{} (\mach',(V',\sigma'))\) \\
+    %
+    \rowcolor{black!20}
+    & & & \(\textnormal{ and } \mach''[\PCname] = \mach[\PCname]+4\) & \\
     %
@@
     & \(\clei \triangleq\) & \(\forall \nach \textnormal{ s.t. } \mach \approx_{\components} \nach,\)
-    & \(\textnormal{ where } \components = \public(V) \cup \object(V)\)
+    & \(\textnormal{ where } \components = \COMPONENTS - (\public(V) \cup \object(V))\)
 %    & \(\textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called}\) \\
@@
   \end{tabular}
-  \caption{Properties\apt{In WBCF, surely we need a state arg? But it needs to be the state \emph{before} the call,
-      doesn't it?}}
+  \caption{Properties}
   \label{tab:props}
@@
 
+The centrality of the functional abstraction and its security are behind the
+many software and hardware mechanisms proposed for its protection
+\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
+Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
+SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
+DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey briefly in
+\cref{sec:intro}. Among these techniques, the micro-policies of Roessler and
+DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
+range of safeguards including lazy enforcement, and suitable for formulating
+and validating strong security guarantees.
+
+The only other line of work that sets out to rigorously characterize the
+security of the stack is the Cerise family of program logics
+\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
+These works define secure calling conventions for capability machines and
+overlay semantics that internalize those conventions and their related stack
+safety properties, phrased in terms of logical relations. Most flavors of
+Cerise define experimental types of hardware capabilities in order to obtain
+efficient calling conventions. Their latest result
+\cite{Georges22:TempsDesCerises} uses our work to extend its notion of stack
+safety to cover caller confidentiality (\(\clrc\)).
+
+HardBound \cite{DeviettiBMZ08} is another hardware extension related to
+capabilities, which closely follows SoftBound \cite{NagarakatteZMZ09} and
+simply aims to accelerate its enforcement of memory safety for compiled code.
+
 \section{Future Work}

### 2023-02-03 submission5/core.tex
Expanded property discussion and definition

@@
 abstraction of functions, in all of the forms that they can take.
-This protean nature of the functional abstraction makes stack safety
+The protean nature of the functional abstraction makes stack safety
 difficult to specify.
@@
 
-Our motivating enforcement mechanisms,
+Our motivating enforcement mechanism,
 the ``lazy'' stack safety micro-policies proposed by
@@
 both caller- and callee-saved registers, arguments passed on the stack,
-\ifexceptions tail-call elimination, and exceptions.
-\else and tail-call elimination. \fi They are modular by
+tail-call elimination, \ifexceptions exceptions \fi. They are modular by
 design, in order to be further extensible to other features.
@@
 
-Indeed, the call stack is an ancient~\cite{phrack96:smashingthestack} and
+Indeed, the call stack is an ancient~\cite{one06phrack} and
 perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
@@
 the return address, and memory corruption more generally.
+
 %
@@
 Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
-other, similar attacks. To avoid this \leo{game of whack-a-mole}, we seek formal properties of safe behavior that can be proven
+other, similar attacks. To avoid this, we seek formal properties of safe behavior that can be proven
 or tested. Such properties become the specification against which enforcement can be validated;
@@
 They are related to one another as seen in
-\cref{fig:propertydiagram}:\leo{is this supposed to be \cref{tab:props}?} \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
+\cref{fig:propertydiagram}: \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
 \(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
@@
 describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
-to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by 
-Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
+to the combination of \(\clri\) and \(\clrc\). The follow-up work by 
+Georges et al. \cite{Georges22:TempsDesCerises} cites an early preprint of our work
+as inspiration for extending their formal discussion to include confidentiality.
 When checking if a preliminary version of these properties applied to
@@
 for the QuickChick library and to ensure that the formalizations are unambiguous.
-Formal proofs are left as future work.
+Proofs are future work.
 
@@
 
-\paragraph*{Organization}
+\paragraph{Organization}
 
@@
 
-In \cref{sec:formal} we formalize the machine model, 
+In \cref{sec:machine} we formalize the machine model, 
 its {\em security semantics}, and the stack safety properties built on these. 
 \Cref{sec:enforcement} describes the micro-policies that we test,
-\cref{sec:testing} the testing framework itself, and
+\cref{sec:test} the testing framework itself, and
 \cref{sec:relwork,sec:future} discuss related and future work.
@@
 
-A security semantics extends a machine
+A security extends a machine
 with additional context about the identity of current and pending
@@
 on future states, e.g. of the form ``when control returns to the current function...'',
-or as relations on traces of future execution (hyper-properties).
+or relations on traces of future execution (hyper-properties).
 
@@
 %\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
+
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
-\cref{tab:psi}. Those that we explore in detail in \cref{sec:example,sec:formal}
+\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
 are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In \cref{sec:extensions}, we cover additional operations and and parameters 
+In \cref{sec:machine}), we cover additional operations that add parameters to the examples,
 to support various forms of memory sharing \ifexceptions , exceptions, \fi
-and tail-call elimination; these are shown in light gray. These are also
-the operations we test in \cref{sec:testing}.
+and tail-call elimination. These are shown in light gray.
 We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
-based on pointer provenance in \cref{app:ptr}, represented by the dark gray
+based on pointer provenance in \cref{app:ptr}, represented by the dark grey
 operations, but we do not test this.
@@
 \end{center}
-\caption{Security-relevant operations and their parameters.}
+  \caption{Security-relevant Operations and their Parameters}
   \label{tab:psi}
@@
 
-To characterize confidentiality\apt{and sometimes integrity}, we borrow the idea of \emph{variant} states
+To characterize confidentiality, we borrow the idea of \emph{variant} states
 from the theory of non-interference. Non-interference is a way of expressing
@@
 In this section we introduce our security property definitions by means
-of small code examples, using a simplified set of security-relevant operations for
+of small code examples, using a simplified set of security-relevent operations for
 calls, returns, and private allocations.
@@
 
-\newcommand{\freebox}[1][10pt]{\tikz \filldraw[fill=blue] (0,0) rectangle (#1,#1);}
-\newcommand{\pubbox}[1][10pt]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (#1,#1);}
-\newcommand{\objbox}[1][10pt]{\tikz \filldraw[fill=yellow] (0,0) rectangle (#1,#1);}
-\newcommand{\sealbox}[1][10pt]{\tikz \filldraw[fill=red] (0,0) rectangle (#1,#1);}
-
-\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30pt,30pt) node[pos=.5]{#1};}
-\newcommand{\numbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
-\newcommand{\leftbox}[1][]{\tikz \filldraw[fill=blue!20] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
-\newcommand{\rightbox}[1][]{\tikz \filldraw[fill=orange!20] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
+\newcommand{\freebox}[1][10px]{\tikz \filldraw[fill=blue] (0,0) rectangle (#1,#1);}
+\newcommand{\pubbox}[1][10px]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (#1,#1);}
+\newcommand{\objbox}[1][10px]{\tikz \filldraw[fill=yellow] (0,0) rectangle (#1,#1);}
+\newcommand{\sealbox}[1][10px]{\tikz \filldraw[fill=red] (0,0) rectangle (#1,#1);}
+
+\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30px,30px) node[pos=.5]{#1};}
+\newcommand{\numbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
+\newcommand{\leftbox}[1][]{\tikz \filldraw[fill=blue!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
+\newcommand{\rightbox}[1][]{\tikz \filldraw[fill=orange!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
 
@@
     \vspace{\abovedisplayskip}
-    \(\textnormal{\large \dots} \pubbox[30pt] \!\! \stackrel{\textsc{\normalsize sp}}{\textbox[\tt res]}
+    \(\textnormal{\large \dots} \pubbox[30px] \!\! \stackrel{\textsc{\normalsize sp}}{\textbox[\tt res]}
     \stackrel{\textsc{\normalsize 4(sp)}}{\textbox[\tt sens]}
@@
 steps of the program.  (The formal details of the security semantics are described in
-\cref{sec:formal}, and the context evolution rules are formalized in \cref{fig:basicops}.)
+\cref{sec:machine}, and the context evolution rules are formalized in \cref{fig:basicops}.)
 Execution begins at the start of {\tt main}, where the program counter (\(\PCname\)) is zero,
@@
   \hline
-  {\tt res} & {\tt sens} & {\tt sec} & \(\mbox{\tt ra}_1\) & \(\mbox{\tt ra}_2\) \\
+  {\tt res} & {\tt sens} & {\tt sec} & \({\mathtt ra_1}\) & \({\mathtt ra_2}\) \\
   \hline
@@
 \section{Formalization}
-\label{sec:formal}
+\label{sec:machine}
 
@@
 Note that this results in an infinite silent trace being similar to
-any trace. So, a trace that silently diverges due to a failstop\apt{?? haven't explained this!}  will
+any trace. So, a trace that silently diverges due to a failstop will
 fulfill this property vacuously.
@@
 \(\hookrightarrow (\mach,\context)  \eqsim \hookrightarrow (\nach,\context)\).
-
-\apt{needs small transition}
 
@@
 \section{Expanding the Machine}
-\label{sec:extensions}
-
-The system we model in \cref{sec:example,sec:formal} is very simple, but our properties
+
+The system we've modeled so far has been very simple, but our properties
 support an important dimension of generality: adding features.\apt{Just what needs to be changed
@@
 
-If the address of an object is taken directly and passed as a
-pointer, we simply classify the object as ``public'' and give it no protection against
-access by other functions. We extend the \(\mathbf{alloc}\) operation
+On the other hand, if the address of an object is taken directly and passed it as a
+pointer, we have no way obvious way to limit the scope of such sharing.
+We treat this scenario as the caller allocating a ``public''
+object that receives no protection. We extend the \(\mathbf{alloc}\) operation
 with a boolean flag, where {\bf t} indicates
@@
 never sealed at a call (\cref{sfig:publicalloc}).
-Providing more fine-grained control over sharing is desirable, but
-seems to require a considerably more complex model; \cref{app:ptr} describes one
-approach, based on capabilities.
 
@@
 
-\apt{good place to detail the needed changes in interpretation of \cref{tab:props}.}
-
 \section{Enforcement}
-\label{sec:enforcement}
-
-We implement and test two micro-policies from Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
+
+We implement and test two micro-policies from Roessler and DeHon \cite{}:
 {\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
@@
 representing its identity. Stack locations belonging to that activation are
-tagged \(\tagStackDepth{n}\), and while the activation is running, the tag on the
-program counter (PC tag) is \(\tagPCDepth{n}\). Stack locations not part of
-any activation are tagged \(\tagNoDepth\).
+tagged \(\tagStackDepth{n}\), and when it is active, the tag on the
+program counter (PC tag) is \(\tagPCDepth{n}\). The tag \(\tagNoDepth\)
+is for stack locations that are currently unused.
 
 In \(DI\), \(n\) always corresponds to the depth of the stack when
-the function is called. A function must initialize its entire frame upon entry
-in order to tag it, and then clear the frame before returning.
+the function is called. A functions must initialize its entire frame on a call
+in order to tag it, and then clear it on a return.
 During normal execution, the micro-policy rules only permit load and
 store operations when the target memory is tagged {\em with the same depth}
-as the current {\PCname} tag, or, for store operations, if the target memory
-is tagged \(\tagNoDepth\).
-
-In \(LTC\), rather than initialize the entire frame at entry,
-a function initializes each location's tag when it writes to it. It does
+as the current {\PCname} tag, and store operations also when the
+target memory is \(\tagNoDepth\).
+
+In \(LTC\), rather than initialize the entire frame,
+a function will initialize each location's tag when it writes to it. It does
 not check if those writes are legal! If the PC tag is \(\tagPCDepth{n}\),
 then any stack location that recieves a store will be tagged \(\tagStackDepth{n}\).
-On a load, the micro-policy failstops if the source memory location 
-is tagged \(\tagNoDepth\) or \(\tagStackDepth{n}\) for some \(n\) that
-doesn't match the PC tag.
-
-\paragraph*{Blessed Sequences}
+On a load, the micro-policy failstops if the value that it would load
+is tagged \(\tagStackDepth{n}\) or \(\tagNoDepth\) and the PC tag doesn't
+match \(n\).
+
+\paragraph{Blessed Sequences}
 
@@
 while performing the usual tasks of saving/restoring the return address to/from
-the stack and adjusting the stack pointer. A blessed sequence uses further tags
-to guarantee that the full sequence executes from the beginning---no jumping into the middle.
-
-\paragraph*{Protecting Registers}
+the stack and adjusting the stack pointer. A blessed sequence also guarantees
+that the full sequence executes from the beginning---no jumping into the middle.
+
+\paragraph{Protecting Registers}
 
@@
 sequence, while in \(LTC\) they are tagged when a value is placed in them.
-\apt{Presumably R\&dH don't care about registers because they already have
-  the compiler in the TCB.  We should note (perhaps earlier than here!) that
-  we have a stronger attacker model since we don't require the callee code
-  to come from a trusted  compiler.}
 
@@
   an initial policy state, tagging instructions and the stack regions appropriately
-  for the micro-policy (see \cref{sec:enforcement})
+  for the micro-policy (see \cref{subsec:policies})
 (3)
@@
 \section{Related Work}
-\label{sec:relwork}
-
-The centrality of the functional abstraction and its security are behind the
-many software and hardware mechanisms proposed for its protection
-\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
-Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
-SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
-DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey briefly in
-\cref{sec:intro}. Among these techniques, the micro-policies of Roessler and
-DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
-range of safeguards including lazy enforcement, and suitable for formulating
-and validating strong security guarantees.
-
-The only other line of work that sets out to rigorously characterize the
-security of the stack is the Cerise family of program logics
-\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
-These works define secure calling conventions for capability machines and
-overlay semantics that internalize those conventions and their related stack
-safety properties, phrased in terms of logical relations. Most flavors of
-Cerise define experimental types of hardware capabilities in order to obtain
-efficient calling conventions. Their latest result
-\cite{Georges22:TempsDesCerises} uses our work to extend its notion of stack
-safety to cover caller confidentiality (\(\clrc\)).
-
-HardBound \cite{DeviettiBMZ08} is another hardware extension related to
-capabilities, which closely follows SoftBound \cite{NagarakatteZMZ09} and
-simply aims to accelerate its enforcement of memory safety for compiled code.
 

### 2023-02-03 submission5/core.tex

--- 2023-02-03 submission5/core.tex
@@
 
-The centrality of the functional abstraction and its security are behind the
-many software and hardware mechanisms proposed for its protection
-\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
-Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
-SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
-DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey briefly in
-\cref{sec:intro}. Among these techniques, the micro-policies of Roessler and
-DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
-range of safeguards including lazy enforcement, and suitable for formulating
-and validating strong security guarantees.
-
-The only other line of work that sets out to rigorously characterize the
-security of the stack is the Cerise family of program logics
-\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
-These works define secure calling conventions for capability machines and
-overlay semantics that internalize those conventions and their related stack
-safety properties, phrased in terms of logical relations. Most flavors of
-Cerise define experimental types of hardware capabilities in order to obtain
-efficient calling conventions. Their latest result
-\cite{Georges22:TempsDesCerises} uses our work to extend its notion of stack
-safety to cover caller confidentiality (\(\clrc\)).
-
-HardBound \cite{DeviettiBMZ08} is another hardware extension related to
-capabilities, which closely follows SoftBound \cite{NagarakatteZMZ09} and
-simply aims to accelerate its enforcement of memory safety for compiled code.
-
 \section{Future Work}

### 2023-02-03 submission5/core.tex
More explanations

@@
 which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
+Roessler and DeHon have the compiler in their trusted computing base, so our properties
+are stronger than necessary to characterize their micro-policies.
 
@@
 
-To characterize confidentiality\apt{and sometimes integrity}, we borrow the idea of \emph{variant} states
+To characterize situations in which values should be kept secret or prevented
+from influencing future execution, we borrow the idea of \emph{variant} states
 from the theory of non-interference. Non-interference is a way of expressing
@@
 then execute both \(\mach\) and \(\nach\), any differences in the outputs they
-produce must be the result of accessing \(\component\).
+produce leaks information about \(\component\).
 
@@
 
-\apt{needs small transition}
+When comparing the behavior of variant states, we need a notion of how their
+differences have influenced them.
 
@@
 The system we model in \cref{sec:example,sec:formal} is very simple, but our properties
-support an important dimension of generality: adding features.\apt{Just what needs to be changed
-to add these features?}
-We now briefly introduce the features of our primary testing machine.\apt{Make sure
-testing has been introduced properly before now. And why ``primary''?}
-
-The full machine defines additional operations, and extends some of the old ones.
-We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
-triples of a register \(\reg\), a base offset \(\mathit{off}\), and a size \(sz\), denoting
-that the value at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
-The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
-optimizations in which the callee will reuse the caller's stack. The \(\mathbf{alloc}\) operation
-gains a boolen flag indicating whether the allocation is to be public, accessible to
-all functions until deallocated. The rules are given in \cref{fig:advops}; the
+are designed to make it easy to add features. To support argument passing on the stack,
+we just new parameters to the existing security-relevant operations, and refine how they
+update the security context. The remainder of the properties do not change at all.
+To add tail-calls, we add and define a new operation, and since it is a kind of call,
+we add it to the definition of call targets again.
+The rules for the expanded machine are given in \cref{fig:advops}; the
 rules in \cref{fig:basicops} can be recaptured by instantiating
@@
 \apt{What if the reference is itself passed on the stack?}
-%In this second case, the reference can be passed further down the stack, but otherwise
-%only the immediate callee gains access. APT:
+\sna{Ugh. We could make \(\reg\) an arbitrary element, but that seems like it's getting
+  dangerously close to indirection.}
 
@@
 it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
-than \(\sealed\) to express this.
-
-\apt{good place to detail the needed changes in interpretation of \cref{tab:props}.}
+than \(\sealed\) to express this. In terms of \cref{tab:props}, we replace
+``call'' with ``call or tailcall'' in \(\clrc\), \(\clec\), and \(\clei\).
 
@@
 
-Roessler and DeHon do not protect registers, but we consider registers to be
-an extension of the stack. In particular, \(\clri\) requires callee-saved
+Roessler and DeHon do not protect registers, since the compiler is in the
+trusted computing base, but we support threat models that do not.
+In particular, \(\clri\) requires callee-saved
 registers to be saved and restored properly. We extend \(DI\) and \(LTC\)

### 2023-02-03 submission5/core.tex
wordsmithing in 5

@@
 
-Finally, we can define our core properties. They can be found in \cref{tab:props},
+\subsection{Properties}
+
+Finally, the core property definitions are given in \cref{tab:props},
 arranged to show their commonalities and distinctions. Each definition gives a criterion
 quantified over states \(s\) that immediately follow call steps.
-If an execution can reach a state \(s'\) such that \(s' \stepstounder{\overline{\psi}} s\)
+If an execution includes a transition \(s' \stepstounder{\overline{\psi}} s\)
 where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
@@
 As a shorthand, we write that each property is defined
-by a criterion that must hold ``for all call targets \(s\).'' In the case of \(\wbcf\),
-we also need to know about the callee, so we refer to the ``call step \(s \stepstounder{} s'\).
-
-\paragraph*{\(\wbcf\)}
+by a criterion that must hold ``for all call targets \(s\),'' or, in the case of \(\wbcf\),
+``for all call steps \(s \stepstounder{} s'\).''
+
+\paragraph*{1.~\(\wbcf\)}
 Given a call step \((\mach,(V,\sigma)) \stepstounder{} (\mach',(V',\sigma'))\),
@@
 whose stack pointer matches that of \(\mach\)
-and whose program counter is at the next instruction. We enjoy \(\wbcf\) if
-\(\ret\) holds on return from \(\mach'\).
-
-\paragraph*{\(\clri\)}
-When our call target is \((\mach,(V,\sigma))\), we define the predicate \(\intProp\) to hold
+and whose program counter is at the next instruction. A system enjoys \(\wbcf\) if,
+for every call transition, \(\ret\) holds just after the callee returns (i.e.,
+the call stack shrinks).
+
+\paragraph*{2.~\(\clri\)}
+When the call target is \((\mach,(V,\sigma))\), we define the predicate \(\intProp\) to hold
 on states \(\mach'\) if any elements that are both sealed in \(V\) and in the difference
-set between \(\mach\) and \(\mach'\) are irrelevant. We enjoy \(\clri\) if \(\intProp\)
-holds on return from \(\mach\).
-
-\paragraph*{\(\clrc\)}
-When our call target \((\mach,(V,\sigma))\), we begin by taking an arbitrary \(\nach\)
+set between \(\mach\) and \(\mach'\) are irrelevant. A system enjoys \(\clri\) if, for every
+call, \(\intProp\) holds just after the corresponding return.
+
+\paragraph*{3.~\(\clrc\)}
+When the call target is \((\mach,(V,\sigma))\), we begin by taking an arbitrary \(\nach\)
 that is a \(\components\)-variant of \(\mach\), where \(\components\) is the set of sealed elements
-in \(V\). We require that two clauses hold. On line a, the behavior of a trace from
+in \(V\). We require that two clauses hold. On line 3a, the behavior of a trace from
 \((\mach,(V,\sigma))\) up to its return must match that of \((\nach,(V,\sigma))\).
-On line b, we define a relation \(\confProp\) that relates states \(\mach'\) and \(\nach'\)
+On line 3b, we define a relation \(\confProp\) that relates states \(\mach'\) and \(\nach'\)
 if their corrupted set (relative to \(\mach\) and \(\nach\)) is irrelevant, and require
-that it hold on the returns from \((\mach,(V,\sigma))\) and \((\nach,(V,\sigma))\).
-We enjoy \(\clrc\) if both clauses hold.
-
-\paragraph*{\(\clec\)}
+that it hold just after the returns from the callees that start at \((\mach,(V,\sigma))\) and \((\nach,(V,\sigma))\).
+A system enjoys \(\clrc\) if both clauses hold for every call.
+
+\paragraph*{4.~\(\clec\)}
 We consider the callee's private behavior to be any changes that it makes to the state
 outside of legitimate channels---elements marked \(\object\) or \(\public\). The remainder
-should be kept secret, which is to say, irrelevant. Similar to \(\clri\), given a call target
+should be kept secret, which is to say, irrelevant to future execution. Similar to \(\clri\), given a call target
 \((\mach,(V,\sigma))\), we define a predicate \(\cconfProp\) to hold
-on states \(\mach'\) if the difference set between \(\mach\) and \(\mach'\), that is
-not \(\object\) or \(\public\), is irrelevant. If \(\cconfProp\) holds on return,
-we enjoy \(\clec\).
-
-\paragraph*{\(\clei\)}
+on states \(\mach'\) if the difference set between \(\mach\) and \(\mach'\), excluding 
+\(\object\) or \(\public\) locations, is irrelevant.
+A system enjoys \(\clec\) if, for every call, \(\cconfProp\) holds just after the corresponding return.
+
+\paragraph*{5.~\(\clei\)}
 Callee integrity means that the caller does not influence it outside of legitimate
 channels. The caller's influence can be seen internally, or in corrupted data on return,
-just like the caller's secrets would under \(\clrc\). So, for a call target
+just like the caller's secrets would be under \(\clrc\). So, for a call target
 \((\mach,(V,\sigma))\), we take an arbitrary \(\nach\) that is a \(\components\)-variant
@@
 To add tail-calls, we add and define a new operation, and since it is a kind of call,
-we add it to the definition of call targets again.
+we add it to the definition of call targets.
 The rules for the expanded machine are given in \cref{fig:advops}; the

### 2023-02-03 submission5/core.tex
work in 4

@@
 and properties. Our definitions abstract over: (1) the details of  the target machine
-architecture and ABI, (2) the set of observable events, and (3) a notion of value compatibility.
-They can be extended to a more sophisticated machine model by expanding the set of
-security-relevant events and defining
-
+architecture and ABI, (2) the set of security-relevant operations and their effects on
+the security context, (3) the set of observable events, and (4) a notion of value compatibility.
+%They can be extended to a more sophisticated machine model by expanding the set of
+%security-relevant events and defining 
 \apt{Any more? Give a more listlike presentation?
@@
 the machine's operation to our security setting: 
-\(\bar{\psi}\) is a list of security-relevant operations drawn from an assumed set \(\Psi\),
+\(\bar{\psi}\) is a list of security-relevant operations drawn from an assumed given set \(\Psi\),
 and \(\obs\) is an (potentially silent) observable event; these are described further below.
@@
 to \(\stepstounder{}\) on combined states. 
-At each step, the context updates based on the function
+At each step, the context updates based on an assumed given function
 \(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
-(Note that \(Op\) takes as its first argument the state {\it before} the step.)
 Since a single step might correspond to multiple operations, we apply
@@
 found in \cref{fig:basicops}.
+Note that \(Op\) takes as its first argument the state {\it before} the step.
 
@@
 
-We abstract over the events that can be observed in the system, defining them
-only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
+We abstract over the events that can be observed in the system, assuming just 
+oa given set \(\OBSS\) that contains at least the element \(\tau\), the silent
 event. Other events might represent certain function calls (i.e., system calls)
@@
 differences have influenced them.
-
 \definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
@@
 \subsection{Properties}
-
-Finally, the core property definitions are given in \cref{tab:props},
-arranged to show their commonalities and distinctions. Each definition gives a criterion
-quantified over states \(s\) that immediately follow call steps.
-If an execution includes a transition \(s' \stepstounder{\overline{\psi}} s\)
-where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
-of a call.
-%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
-%\(s\) is the target of a tailcall.
-As a shorthand, we write that each property is defined
-by a criterion that must hold ``for all call targets \(s\),'' or, in the case of \(\wbcf\),
-``for all call steps \(s \stepstounder{} s'\).''
-
-\paragraph*{1.~\(\wbcf\)}
-Given a call step \((\mach,(V,\sigma)) \stepstounder{} (\mach',(V',\sigma'))\),
-we define the predicate \(\ret\) to hold on states \(\mach''\)
-whose stack pointer matches that of \(\mach\)
-and whose program counter is at the next instruction. A system enjoys \(\wbcf\) if,
-for every call transition, \(\ret\) holds just after the callee returns (i.e.,
-the call stack shrinks).
-
-\paragraph*{2.~\(\clri\)}
-When the call target is \((\mach,(V,\sigma))\), we define the predicate \(\intProp\) to hold
-on states \(\mach'\) if any elements that are both sealed in \(V\) and in the difference
-set between \(\mach\) and \(\mach'\) are irrelevant. A system enjoys \(\clri\) if, for every
-call, \(\intProp\) holds just after the corresponding return.
-
-\paragraph*{3.~\(\clrc\)}
-When the call target is \((\mach,(V,\sigma))\), we begin by taking an arbitrary \(\nach\)
-that is a \(\components\)-variant of \(\mach\), where \(\components\) is the set of sealed elements
-in \(V\). We require that two clauses hold. On line 3a, the behavior of a trace from
-\((\mach,(V,\sigma))\) up to its return must match that of \((\nach,(V,\sigma))\).
-On line 3b, we define a relation \(\confProp\) that relates states \(\mach'\) and \(\nach'\)
-if their corrupted set (relative to \(\mach\) and \(\nach\)) is irrelevant, and require
-that it hold just after the returns from the callees that start at \((\mach,(V,\sigma))\) and \((\nach,(V,\sigma))\).
-A system enjoys \(\clrc\) if both clauses hold for every call.
-
-\paragraph*{4.~\(\clec\)}
-We consider the callee's private behavior to be any changes that it makes to the state
-outside of legitimate channels---elements marked \(\object\) or \(\public\). The remainder
-should be kept secret, which is to say, irrelevant to future execution. Similar to \(\clri\), given a call target
-\((\mach,(V,\sigma))\), we define a predicate \(\cconfProp\) to hold
-on states \(\mach'\) if the difference set between \(\mach\) and \(\mach'\), excluding 
-\(\object\) or \(\public\) locations, is irrelevant.
-A system enjoys \(\clec\) if, for every call, \(\cconfProp\) holds just after the corresponding return.
-
-\paragraph*{5.~\(\clei\)}
-Callee integrity means that the caller does not influence it outside of legitimate
-channels. The caller's influence can be seen internally, or in corrupted data on return,
-just like the caller's secrets would be under \(\clrc\). So, for a call target
-\((\mach,(V,\sigma))\), we take an arbitrary \(\nach\) that is a \(\components\)-variant
-of \(\mach\), where \(\components\) is the set all elements that are not \(\object\)
-or \(\public\). The remainder of the property is identical to \(\clrc\).
 
@@
 \end{table*}
+
+Finally, the core property definitions are given in \cref{tab:props},
+arranged to show their commonalities and distinctions. Each definition gives a criterion
+quantified over states \(s\) that immediately follow call steps.
+If an execution includes a transition \(s' \stepstounder{\overline{\psi}} s\)
+where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
+of a call.
+%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
+%\(s\) is the target of a tailcall.
+As a shorthand, we write that each property is defined
+by a criterion that must hold ``for all call targets \(s\),'' or, in the case of \(\wbcf\),
+``for all call steps \(s \stepstounder{} s'\).''
+
+\paragraph*{1.~\(\wbcf\)}
+Given a call step \((\mach,(V,\sigma)) \stepstounder{} (\mach',(V',\sigma'))\),
+we define the predicate \(\ret\) to hold on states \(\mach''\)
+whose stack pointer matches that of \(\mach\)
+and whose program counter is at the next instruction. A system enjoys \(\wbcf\) if,
+for every call transition, \(\ret\) holds just after the callee returns (i.e.,
+the call stack shrinks).
+
+\paragraph*{2.~\(\clri\)}
+When the call target is \((\mach,(V,\sigma))\), we define the predicate \(\intProp\) to hold
+on states \(\mach'\) if any elements that are both sealed in \(V\) and in the difference
+set between \(\mach\) and \(\mach'\) are irrelevant. A system enjoys \(\clri\) if, for every
+call, \(\intProp\) holds just after the corresponding return.
+
+\paragraph*{3.~\(\clrc\)}
+When the call target is \((\mach,(V,\sigma))\), we begin by taking an arbitrary \(\nach\)
+that is a \(\components\)-variant of \(\mach\), where \(\components\) is the set of sealed elements
+in \(V\). We require that two clauses hold. On line 3a, the behavior of a trace from
+\((\mach,(V,\sigma))\) up to its return must match that of \((\nach,(V,\sigma))\).
+On line 3b, we define a relation \(\confProp\) that relates states \(\mach'\) and \(\nach'\)
+if their corrupted set (relative to \(\mach\) and \(\nach\)) is irrelevant, and require
+that it hold just after the returns from the callees that start at \((\mach,(V,\sigma))\) and \((\nach,(V,\sigma))\).
+A system enjoys \(\clrc\) if both clauses hold for every call.
+
+\paragraph*{4.~\(\clec\)}
+We consider the callee's private behavior to be any changes that it makes to the state
+outside of legitimate channels---elements marked \(\object\) or \(\public\). The remainder
+should be kept secret, which is to say, irrelevant to future execution. Similar to \(\clri\), given a call target
+\((\mach,(V,\sigma))\), we define a predicate \(\cconfProp\) to hold
+on states \(\mach'\) if the difference set between \(\mach\) and \(\mach'\), excluding 
+\(\object\) or \(\public\) locations, is irrelevant.
+A system enjoys \(\clec\) if, for every call, \(\cconfProp\) holds just after the corresponding return.
+
+\paragraph*{5.~\(\clei\)}
+Callee integrity means that the caller does not influence it outside of legitimate
+channels. The caller's influence can be seen internally, or in corrupted data on return,
+just like the caller's secrets would be under \(\clrc\). So, for a call target
+\((\mach,(V,\sigma))\), we take an arbitrary \(\nach\) that is a \(\components\)-variant
+of \(\mach\), where \(\components\) is the set all elements that are not \(\object\)
+or \(\public\). The remainder of the property is identical to \(\clrc\).
 

### 2023-02-03 submission5/core.tex
wibbles and comment in intro

@@
 it afterward. Because the memory does get corrupted, \(LTC\) would not fulfill an
-adaptation Georges et al.'s property to the tagged setting.
+adaptation of Georges et al.'s property to the tagged setting.
 
 Neverthess, we believe that there is a useful sense in which \(LTC\) does (attempt to) enforce stack safety. 
-It does not always protect individual addresses on the stack, but it tries to 
+While it does not always protect individual addresses on the stack, it tries to 
 prevent one function activation from interfering with another's behavior---protecting 
@@
 
-\paragraph*{Towards Formal Properties for Stack Safety}
+\subsection {Towards Formal Properties for Stack Safety}
+
+\apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
+  Again, we are not providing evidence that these attributes are true, although we can
+  and should conjecture claim them in passing.}
 

### 2023-02-03 submission5/core.tex
64 bit

@@
 calls, returns, and private allocations.
-\Cref{fig:main} gives C code and possible corresponding compiled 32-bit RISC-V code
+\Cref{fig:main} gives C code and possible corresponding compiled 64-bit RISC-V code
 for a function {\tt main}, which

### 2023-02-03 submission5/core.tex
Merge branch 'master' of https://github.com/bcpierce00/interfering
* 'master' of https://github.com/bcpierce00/interfering:
  Eager testing file
  work in 4
  Add copy of lazy properties file for eager DI
  Add eager default layout
  Corrections to eager files
  wordsmithing in 5
  Copy lazy files as baseline for revised eager DI
  Move outdated Eager modules

# Conflicts:
#	submission5/core.tex

--- 2023-02-03 submission5/core.tex
@@
 
-The term {\em stack safety} is associated with a
-variety of compiler, run-time, and hardware mechanisms for protecting stack
-memory. But, unlike ``the heap,'' the stack does not correspond to a single
-high-level language concept. Rather, it corresponds to the fundamental
-abstraction of functions, in all of the forms that they can take.
-This protean nature of the functional abstraction makes stack safety
-difficult to specify.
-
-We propose a formal characterization of stack safety based on
-concepts from language-based security. Stack safety is decomposed
-into an integrity property and a confidentiality property for each
-of the caller and the callee, as well as a control-flow property.
-
-Our motivating enforcement mechanisms,
-the ``lazy'' stack safety micro-policies proposed by
-Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, permit functions
-to write into one another's frames, but taint the result so that the frame's
-owner cannot access it. No existing characterization of stack safety
-captures this style of safety. We capture it by defining the properties in
-terms of the observable behavior of the system.
-
-The stack interacts with a large number of language features that are often excluded
-from discussion of stack safety. Our properties support a system with
-both caller- and callee-saved registers, arguments passed on the stack,
-\ifexceptions tail-call elimination, and exceptions.
-\else and tail-call elimination. \fi They are modular by
-design, in order to be further extensible to other features.
-
-We validate our properties by using them to distinguish correct implementations
-of Roessler and DeHon's micro-policies from incorrect ones
-via property-based random testing. Our testing successfully detects violations
-in several broken variants, including Roessler and DeHon's original lazy policy.
-A fixed version of the policy does pass our tests.
+The term {\em stack safety} is used for a variety of compiler,
+run-time, and hardware mechanisms for protecting stack memory. Unlike
+``the heap,'' the hardware stack does not correspond to a single
+high-level language concept: different compilers use it in different
+ways to support procedural and functional abstraction mechanisms from
+a wide range of languages.  This protean nature makes it difficult to
+nail down what it means to correctly enforce stack safety.
+
+We propose a formal characterization of stack safety using concepts
+from language-based security. Rather than packaging all aspects of
+stack safety into a monolithic property, we decompose it into an
+integrity property and a confidentiality property for each of the
+caller and the callee, plus a control-flow property---five properties
+in all.
+
+This presentation is motivated by a particular class of enforcement
+mechanisms, the ``lazy'' stack safety micro-policies studied by
+Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, which permit
+functions to write into one another's frames, but which taint the
+changed locations so that the frame's owner cannot access it. No
+existing characterization of stack safety captures this style of
+safety. We capture it here by stating our properties in terms of the
+observable behavior of the system.
+
+The hardware stack interacts with a large number of language features,
+some of which are typically excluded from discussions of stack
+safety. Our properties go further than previous proposals in
+supporting caller- and callee-saved registers, arguments passed on the
+stack, \ifexceptions tail-call elimination, and exceptions.  \else and
+tail-call elimination. \fi They are modular by design, in order to be
+further extensible to other features.
+
+We validate our properties by using them to distinguish between
+correct and incorrect implementations of Roessler and DeHon's
+micro-policies using property-based random testing. Our test harness
+successfully identifies several broken variants, including Roessler
+and DeHon's original lazy policy.  A fixed version of their policy
+does pass our tests.
 

### 2023-02-03 submission5/core.tex
a0 is active not public in fig 3
Abstract

@@
 some of which are typically excluded from discussions of stack
-safety. Our properties go further than previous proposals in
+safety. Our properties go further than previous proposals,
 supporting caller- and callee-saved registers, arguments passed on the
 stack, \ifexceptions tail-call elimination, and exceptions.  \else and
-tail-call elimination. \fi They are modular by design, in order to be
-further extensible to other features.
+tail-call elimination. \fi They are modular by design so that they
+can be further extended to other features.
 

### 2023-02-03 submission5/core.tex
Abstract

@@
 run-time, and hardware mechanisms for protecting stack memory. Unlike
-``the heap,'' the hardware stack does not correspond to a single
+``the heap,'' the ISA-level stack does not correspond to a single
 high-level language concept: different compilers use it in different
@@
 
-The hardware stack interacts with a large number of language features,
-some of which are typically excluded from discussions of stack
-safety. Our properties go further than previous proposals,
-supporting caller- and callee-saved registers, arguments passed on the
-stack, \ifexceptions tail-call elimination, and exceptions.  \else and
-tail-call elimination. \fi They are modular by design so that they
-can be further extended to other features.
+% The stack plays a central role in realizing large number of
+% high-level language features, some of which are typically excluded
+% from discussions of stack safety. 
+Our properties go further than previous formal definitions of stack
+safety, supporting caller- and callee-saved registers, arguments
+passed on the stack, \ifexceptions tail-call elimination, and
+exceptions.  \else and tail-call elimination. \fi 
+% \bcp{Out of all of
+%   these, only exceptions are high-level language features!}
+They are modular by design so that they can be further extended to
+other features.
 
@@
       \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\pubbox} \!\! \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
+      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
       ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
+      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
       ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}

### 2023-02-03 submission5/core.tex
wip

--- 2023-02-03 submission5/core.tex
@@
 \subsection{Threat Model and Limitations}
+\apt{This could probably go better at end of next section.}
 
@@
 
-\section{Building a Property}
+\section{Framework}
 \label{sec:ideas}
+
+Our properties are
+
+We assume a conventional hardware ISA, such as x86-64, ARM, or RISC-V, with a program counter and stack pointer,
+and a standard ABI with a contiguous stack that holds interleaved data and control information.
+The machine might be enhanced with additional monitoring or other security mechanisms,
+such as hardware tags or capabilities. We do not , but....
 

### 2023-02-03 submission5/core.tex
Related work also in IX

--- 2023-02-03 submission5/core.tex
@@
 SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
-DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey briefly in
-\cref{sec:intro}. Among these techniques, the micro-policies of Roessler and
+DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey in
+Sections~\ref{sec:intro} and~\ref{sec:future}. Among these techniques, the micro-policies of Roessler and
 DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
@@
 %
-Previous work has uses these principles in simplified models to implement
+Previous work has used these principles in simplified models to implement
 \emph{secure calling conventions} for the call

### 2023-02-03 submission5/core.tex
CP intro

--- 2023-02-03 submission5/core.tex
@@
 
-This presentation is motivated by a particular class of enforcement
+This formulation is motivated by a particular class of enforcement
 mechanisms, the ``lazy'' stack safety micro-policies studied by
@@
 successfully identifies several broken variants, including Roessler
-and DeHon's original lazy policy.  A fixed version of their policy
-does pass our tests.
+and DeHon's lazy policy; a repaired version of their policy does pass
+our tests.
 
@@
 
-Computer programming relies on functions (or subroutines, procedures, methods, etc.)
-as a foundational abstraction. Functions are units of computation that
-can call one another to realize larger computations in a modular fashion.
-%
-At a low level, a function activation manages its own data, such as local variables, as
-well as information about its \emph{caller} function, to which it must return.
-The \emph{(call) stack} is the fundamental data structure charged with realizing
-this abstraction, aided by an ABI that defines how registers are shared between activations.
-From a security perspective, attacks on the stack are really attacks
-on the fundamental functional abstraction itself, and are therefore every bit as
-pervasive.
-
-Indeed, the call stack is an ancient~\cite{phrack96:smashingthestack} and
+{\em Functions} in high-level languages, and related abstractions such
+as subroutines, procedures, methods, etc., are units of computation
+that call one another to modularly realize larger computations.
+%
+At a low level, each function activation manages its own data---local
+variables, etc.---as well as information about the \emph{caller} to
+which it must return.
+%
+The \emph{call stack} is the fundamental data structure used to
+implement functions, aided by an ABI that defines how registers are shared
+between activations.
+
+From a security perspective, attacks on the call stack
+are attacks on the function abstraction itself.
+%
+Indeed, the stack is an ancient~\cite{phrack96:smashingthestack} and
 perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
@@
   DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
-target for low-level attacks, frequently involving control-flow hijacking via corrupting
-the return address, and memory corruption more generally.
+target for low-level attacks, sometimes involving control-flow
+hijacking via corrupting the return address, sometimes memory corruption
+more generally.  
 %
@@
 %using data stored in the stack 
-The variety in attacks on the stack is mirrored by the
+The variety in attacks on the stack is mirrored in the range of
 software and hardware protections that aim to prevent them,
@@
 
-Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
-other, similar attacks. To avoid this \leo{game of whack-a-mole}, we seek formal properties of safe behavior that can be proven
-or tested. Such properties become the specification against which enforcement can be validated;
-even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
-\emph{why} they fail.
-
-Of the mechanisms above, many are fundamentally ill-suited for offering formal guarantees,
-as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
-for instance, aim to ``restrict the flexibility available in creating gadget chains''
-\cite{Shanbhogue+19}, not to categorically
-rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
-integrity\cite{Kuznetsov+14} are stronger, but most do not give formal guarantees.
-Indeed, the secure calling conventions of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
-Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
-to offer stack safety formalizations.
-
-Among the mechanisms that should be amenable to making strong formal guarantees,
-Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
-micro-policies~\cite{pump_oakland2015} for stack safety.
-Their most realistic micro-policy, ``Lazy Tagging and Clearing'' (\(LTC\)),
-makes an interesting trade-off for performance: it allows function activations to write
-into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
-it afterward. Because the memory does get corrupted, \(LTC\) would not fulfill an
-adaptation of Georges et al.'s property to the tagged setting.
-
-Neverthess, we believe that there is a useful sense in which \(LTC\) does (attempt to) enforce stack safety. 
-While it does not always protect individual addresses on the stack, it tries to 
-prevent one function activation from interfering with another's behavior---protecting 
-the functional abstraction, which we argue is the real essence of stack safety.
-
-In this work, we propose a formal characterization of stack safety based on the idea
+Enforcement mechanisms can be brittle, successfully eliminating one
+attack while leaving room for others. To avoid an endless game of
+whack-a-mole, we seek formal properties of safe behavior that can be
+proven or at least rigorously tested. Such properties can be used as
+the specification against which enforcement can be validated; even
+enforcement mechanisms that do not fulfill a property benefit from the
+ability to articulate why and when they may fail.
+
+Of the mechanisms listed above, many are fundamentally ill suited for
+offering formal guarantees: may impede attackers, but do not provide
+universal protection. Shadow stacks, for instance, aim to ``restrict
+the flexibility available in creating gadget chains''
+\cite{Shanbhogue+19}, not to categorically rule out attacks. Other
+mechanisms, such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
+integrity\cite{Kuznetsov+14}, do aim for stronger guarantees, but even
+these do not make formal promises.  Indeed, to our knowledge, the sole
+line of work making a formal claim to protect stack safety is the
+secure calling conventions of Skorstengaard et
+al.~\cite{SkorstengaardSTKJFP} and Georges et
+al.~\cite{Georges22:TempsDesCerises}.
+
+Fortunately, other mechanisms besides these should be
+amenable to strong 
+formal guarantees.  In particular, 
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} have presented an
+array of tag-based micro-policies~\cite{pump_oakland2015} for stack
+safety that give strong guarantees.
+Their most realistic micro-policy, called Lazy Tagging and Clearing (\(LTC\)),
+makes an interesting performance trade-off: it allows function
+activations to improperly write into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
+it afterward.
+
+Because memory does get corrupted, \(LTC\) would not fulfill an
+adaptation of Georges et al.'s property to the tagged setting.  But
+there is a useful sense in which \(LTC\) does enforce stack safety.
+While it does not necessarily protect individual addresses on the
+stack, it does prevent one function activation from interfering with
+another's observable behavior.
+
+We propose a formal characterization of stack safety based on the intuition
 of protecting each function activation from other activations.
@@
 and the callee's behavior during the callee's execution, plus the control-flow protection
-of {\em well-bracketed control flow} (\(\wbcf\))~\cite{SkorstengaardSTKJFP}.
-While our properties are motivated by a desire to specify \(LTC\), 
-they are built from first principles in the hopes that they will be broadly applicable
-to a range of enforcement mechanisms.
-
-The structure of our properties enables them to be easily expanded to characterize systems with
-sophisticated features such as callee-save registers, argument-passing on the stack,
-and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
+of {\em well-bracketed control flow} (\(\wbcf\)) articulated by
+Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
+While our properties are motivated by the desire to specify \(LTC\) precisely, 
+they are stated somewhat abstractly in hopes that they may be applicable
+to other enforcement mechanisms.
+
+The structure of our properties enables them to be easily extended\bcp{??} to characterize systems with
+features such as callee-save registers, argument-passing on the stack,
+and tail-call elimination. We likewise extend \(LTC\) to protect these features, and
 test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
@@
 example model, and then moving to a more realistic model with multiple new features:
-callee-save registers, argument passing on the stack, and tailcall elimination, simply
+callee-save registers, argument passing on the stack, and tail-call elimination, simply
 by refining a few definitions in (1). This becomes the model that we validate.
@@
     \example \(\mathbf{dealloc}\) & offset \& size \\
-    \testing \(\mathbf{tailcall}\) & (same as for \(\mathbf{call}\)) \\
+    \testing \(\mathbf{tail call}\) & (same as for \(\mathbf{call}\)) \\
     \hline
@@
 of a call.
-%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
-%\(s\) is the target of a tailcall.
+%Likewise, if \(\mathbf{tail call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
+%\(s\) is the target of a tail call.
 As a shorthand, we write that each property is defined
@@
                {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{tailcall} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+               {\(Op ~ \mach ~ (\mathbf{tail call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
                  ~ (V,\sigma) = (V',\sigma)\)}
@@
   \end{subfigure}
-  \caption{Operations supporting tailcalls and argument passing on stack.}
+  \caption{Operations supporting tail calls and argument passing on stack.}
   \label{fig:advops}
@@
 
-The operation rule for a tailcall is similar to that for a normal call.
+The operation rule for a tail call is similar to that for a normal call.
 We do not push the caller's view onto the stack,
-but replace it outright. This means that a tailcall does not increase the size of
-the call stack, and therefore for purposes of our properties, all tailcalls will
+but replace it outright. This means that a tail call does not increase the size of
+the call stack, and therefore for purposes of our properties, all tail
+calls will
 be considered to return simultaneously when the eventual {\bf return} operation
@@
 than \(\sealed\) to express this. In terms of \cref{tab:props}, we replace
-``call'' with ``call or tailcall'' in \(\clrc\), \(\clec\), and \(\clei\).
+``call'' with ``call or tail call'' in \(\clrc\), \(\clec\), and \(\clei\).
 

### 2023-02-04 submission5/core.tex
Clearer point in related work
CP intro -- rewritten contributions -- some notes and questions

@@
 
-Fortunately, other mechanisms besides these should be
-amenable to strong 
-formal guarantees.  In particular, 
-Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} have presented an
-array of tag-based micro-policies~\cite{pump_oakland2015} for stack
-safety that give strong guarantees.
-Their most realistic micro-policy, called Lazy Tagging and Clearing (\(LTC\)),
-makes an interesting performance trade-off: it allows function
-activations to improperly write into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
-it afterward.
-
-Because memory does get corrupted, \(LTC\) would not fulfill an
-adaptation of Georges et al.'s property to the tagged setting.  But
-there is a useful sense in which \(LTC\) does enforce stack safety.
-While it does not necessarily protect individual addresses on the
-stack, it does prevent one function activation from interfering with
-another's observable behavior.
-
-We propose a formal characterization of stack safety based on the intuition
-of protecting each function activation from other activations.
-We use the formal tools of language-based
+Other mechanisms besides these should also be amenable to strong
+formal guarantees.  In particular, Roessler and DeHon
+\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
+micro-policies~\cite{pump_oakland2015} for stack safety that offer
+strong guarantees.  Their most realistic micro-policy, called Lazy
+Tagging and Clearing (\(LTC\)), makes an interesting performance
+trade-off: it allows function activations to improperly write into one
+another's stack frames, but ensures that the owner of the corrupted
+memory cannot access it afterward.
+
+Under this policy, one function activation {\em can} corrupt another's
+memory---just not in ways that do not affect observable behavior.
+Therefore, \(LTC\) would not fulfill Georges et al.'s property
+(appropriately adapted to the tagged setting).  But \(LTC\) does
+arguably enforce stack safety.  A looser, more observational
+definition of stack safety is needed to fit this situation.
+
+We propose here a formal characterization of stack safety, based on
+the intuition of protecting function activations from each other.
+\bcp{Aren't other formal definitions based on the same intuition?}
+We use the tools of language-based
 security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
@@
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the control-flow protection
-of {\em well-bracketed control flow} (\(\wbcf\)) articulated by
+and the callee's behavior during the callee's execution, plus the
+{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
 Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
 While our properties are motivated by the desire to specify \(LTC\) precisely, 
-they are stated somewhat abstractly in hopes that they may be applicable
+they are stated abstractly in hopes that they may also apply
 to other enforcement mechanisms.
 
-The structure of our properties enables them to be easily extended\bcp{??} to characterize systems with
-features such as callee-save registers, argument-passing on the stack,
-and tail-call elimination. We likewise extend \(LTC\) to protect these features, and
-test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
-is flawed in a way that undermines integrity and confidentiality, and we correct
-this flaw.
+We find, via property-based random testing, that \(LTC\) is flawed in
+a way that undermines both integrity and confidentiality; after
+correcting this flaw, the repaired \(LTC\) satisfies all our
+properties.  Moreover, we show that our properties can be extended to
+describe calling conventions with features such as callee-save
+registers, argument-passing on the stack, and tail-call
+elimination. We modify \(LTC\) to protect these features and apply
+random testing to validate this extended protection mechanism against
+the extended properties.
 
@@
 \begin{itemize}
-\item A novel characterization of stack safety as a collection of properties:
-  confidentiality and integrity for callee and caller, and well-bracketed
-  control-flow. These can characterize the security requirements of systems
-  that support only some portion of stack safety, and are
-  parameterized over a notion of external observation, so that they can characterize
-  lazy enforcement.
-\item An instantiation of these definitions in a realistic
-  setting with features such as argument passing
-  on the stack, callee-saves registers, and tail-call elimination.
-  Our model is modular enough that handling these features is straightforward.
-\item The validation, via property-based randomize testing, of tag-based
-  stack-safety policies. We find that \emph{Lazy Tagging and Clearing}
-  falls short of stack safety, and propose a fix.
+\item We give a novel characterization of stack safety as a collection
+of properties: confidentiality and integrity for callee and caller,
+plus well-bracketed control-flow. Different subsets of these
+properties characterize the security guarantees offered by mechanisms
+that support only some aspects of stack safety\bcp{where do we show
+  this?}.  The properties they are parameterized over a notion of
+external observation, allowing them to characterize lazy enforcement
+mechanisms.
+\item We instantiate\bcp{extend??} these definitions in a realistic
+setting with features such as\bcp{``such as''? are there others?}
+argument passing on the stack, callee-saves registers, and tail-call
+elimination.\bcp{How about: We extend these core definitions to
+  describe a realistic setting with argument passing on the stack,
+  callee-saves registers, and tail-call elimination.}  Our model is
+modular enough that handling these features is straightforward.
+\bcp{Throughout, the way we talk about the ``core model'' and its
+  intantiation / extension / enrichment with realistic features feels
+  quite weasily.}
+\item We validate tag-based stack-safety enforcement via
+property-based randomize testing. We find that \emph{Lazy Tagging and
+  Clearing} falls short of ensuring stack safety, propose a fix, and
+validate it.
 \end{itemize}

### 2023-02-04 submission5/core.tex
CP

@@
 
-\apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
-  Again, we are not providing evidence that these attributes are true, although we can
-  and should conjecture claim them in passing.}
+% \apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
+%   Again, we are not providing evidence that these attributes are true, although we can
+%   and should conjecture claim them in passing.}
 
@@
   quite weasily.}
-\item We validate tag-based stack-safety enforcement via
-property-based randomize testing. We find that \emph{Lazy Tagging and
-  Clearing} falls short of ensuring stack safety, propose a fix, and
-validate it.
+\item We validate a tag-based stack-safety enforcement mechanism,
+\emph{Lazy Tagging and Clearing}, via property-based random testing,
+find that it falls short of ensuring stack safety, and propose and
+validate a fix.
 \end{itemize}
 
-Our artifact contains Coq formalizations of our properties for illustrative purposes
-as well as their testing implementation. There are no proofs---we use Coq primarily
-for the QuickChick library and to ensure that the formalizations are unambiguous.
-Formal proofs are left as future work.
+The accompanying artifact contains formal definitions (in Coq) of our
+properties, plus our testing framework.  It does not include proofs:
+we use Coq primarily for the QuickChick testing library and for
+ensuring that our definitions are unambiguous.  Formal proofs are left
+as future work.
 
@@
 Cerise define experimental types of hardware capabilities in order to obtain
-efficient calling conventions. Their latest result
-\cite{Georges22:TempsDesCerises} uses our work to extend its notion of stack
-safety to cover caller confidentiality (\(\clrc\)).
+efficient calling conventions. Their latest paper
+\cite{Georges22:TempsDesCerises} uses our work to extend their notion of stack
+safety to include caller confidentiality (\(\clrc\)).
 

### 2023-02-04 submission5/core.tex
CP

@@
 
-% \apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
-%   Again, we are not providing evidence that these attributes are true, although we can
-%   and should conjecture claim them in passing.}
+\apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
+  Again, we are not providing evidence that these attributes are true, although we can
+  and should conjecture claim them in passing.}
+
+\bcp{Agreed.  I think in particular the first paragraph can just go.
+  But this is not enough.  Also, the section is trying to do at least
+  two things---talk about earlier work and how we compare, and talk
+  about why we believe a flexible approach is good.  The whole
+  discussion is a bit wooly and confusing.}
 
@@
   quite weasily.}
-\item We validate a tag-based stack-safety enforcement mechanism,
-\emph{Lazy Tagging and Clearing}, via property-based random testing,
-find that it falls short of ensuring stack safety, and propose and
-validate a fix.
+\item We validate a tag-based enforcement mechanism, \emph{Lazy
+  Tagging and Clearing}, via property-based random testing, find that
+it falls short, and propose and validate a fix.
 \end{itemize}
+
+In the next section, we give a brief overview of the structure of
+properties.  In \cref{sec:example}, we walk through a function call in
+a simple example machine and discuss how each of our properties
+applies to it, informally. In the process we motivate the properties
+from a security perspective.
+%
+In \cref{sec:formal} we formalize the machine model,
+its {\em security semantics}, and the stack safety properties built on these.
+In \Cref{sec:enforcement} we describe the micro-policies that we test,
+in \cref{sec:testing} the testing framework itself, and in
+\cref{sec:relwork,sec:future} related and future work.
 
@@
 
-When our properties are used to evaluate a system, the threat model will depend on the
-details of that system. However, there are some constraints that our design puts on
-any system. We must trust that the method of distinguishing security-relevant operations is accurate; if it
-involves labels placed on code by a compiler, that means trusting that the compiler placed
-those labels correctly. If operations occur that are not recognized, those operations
-might not be guaranteed to protect their principals---for instance, an unlabeled call
-might not protect the caller's data from the callee. On the flip side, applying an incorrect
-label most likely means that the property becomes too strong to be enforced.
-
-Otherwise, we do not assume that the code adheres to any particular
-calling convention or implements a source language construct.
-In particular, while we are agnostic as to the source
-language, we certainly aim to support C, and so any source function might contain undefined
-behavior resulting in its compilation to arbitrary machine code. A given enforcement
+When our properties are used to evaluate a system, the threat model
+will depend on the details of that system. However, there are some
+constraints that our design puts on any system. In particular, we must
+trust that the method of distinguishing security-relevant operations
+is accurate.  For example, if the enforcement mnechanism involves
+labels placed on code by a compiler, we must trust that the compiler
+placed those labels correctly. For instance, if a compiled function
+call is not marked as such, then the caller's data might not be
+protected from the callee.  Conversely, marking too many operations as
+calls might mean that the property becomes too strong to be
+enforced.\bcp{I don't understand that---wouldn't it mean that
+  enforcing the property causes almost all programs to be rejected?
+  (I.e., isn't it a case of false negatives rather than false
+  positives?)}
+
+\bcp{Still working here:}
+
+We do not assume that low-level code adheres to any fixed calling
+convention or implements any particular source-language constructs.
+Indeed, if the source
+language is C, then high-level programs might contain undefined
+behavior, in which case they might be compiled to arbitrary machine
+code. A given enforcement 
 mechanism or target architecture might place additional constraints, particularly on the behavior of
@@
 which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
-Roessler and DeHon have the compiler in their trusted computing base, so our properties
+Roessler and DeHon include the compiler in their trusted computing
+base, so our properties 
 are stronger than necessary to characterize their micro-policies.
 
-In general, it is impossible to distinguish buggy machine code from an attacker; in
-our examples we will identify one function or another as an attacker, but we do not
-require any static division between trusted and untrusted code, and we aim to protect
-even buggy code.
-
-This is a strong threat model, but it does omit some important aspects of
-security enforcement: in particular, hardware and timing attacks are out of scope, and
-we do not address concurrency---that is a high-priority future work.
-%
-
-\paragraph*{Organization}
-
-In the next section, we will give a brief overview of the structure of properties.
-Then in \cref{sec:example}, we will walk through a function
-call in a simple example machine and discuss how each of our properties applies to it,
-informally. In the process we discuss some motivation for the properties, from a security
-perspective.
-
-In \cref{sec:formal} we formalize the machine model,
-its {\em security semantics}, and the stack safety properties built on these.
-\Cref{sec:enforcement} describes the micro-policies that we test,
-\cref{sec:testing} the testing framework itself, and
-\cref{sec:relwork,sec:future} discuss related and future work.
+In general, it is impossible to distinguish buggy machine code from an
+attacker.  In our examples we will identify one function or another as
+an attacker, but we do not require any static division between trusted
+and untrusted code, and we aim to protect even buggy code.
+
+This is a strong threat model, but it does omit some important aspects
+of stack safety in real systems: in particular, it does not address
+concurrency.  Hardware and timing attacks are also out of scope.
+%
 

### 2023-02-04 submission5/core.tex

--- 2023-02-04 submission5/core.tex
@@
 run-time, and hardware mechanisms for protecting stack memory. Unlike
-``the heap,'' the ISA-level stack does not correspond to a single
+``the heap,'' the hardware stack does not correspond to a single
 high-level language concept: different compilers use it in different
@@
 
-This formulation is motivated by a particular class of enforcement
+This presentation is motivated by a particular class of enforcement
 mechanisms, the ``lazy'' stack safety micro-policies studied by
@@
 
-% The stack plays a central role in realizing large number of
-% high-level language features, some of which are typically excluded
-% from discussions of stack safety. 
-Our properties go further than previous formal definitions of stack
-safety, supporting caller- and callee-saved registers, arguments
-passed on the stack, \ifexceptions tail-call elimination, and
-exceptions.  \else and tail-call elimination. \fi 
-% \bcp{Out of all of
-%   these, only exceptions are high-level language features!}
-They are modular by design so that they can be further extended to
-other features.
+The hardware stack interacts with a large number of language features,
+some of which are typically excluded from discussions of stack
+safety. Our properties go further than previous proposals,
+supporting caller- and callee-saved registers, arguments passed on the
+stack, \ifexceptions tail-call elimination, and exceptions.  \else and
+tail-call elimination. \fi They are modular by design so that they
+can be further extended to other features.
 
@@
 successfully identifies several broken variants, including Roessler
-and DeHon's lazy policy; a repaired version of their policy does pass
-our tests.
+and DeHon's original lazy policy.  A fixed version of their policy
+does pass our tests.
 
@@
 
-{\em Functions} in high-level languages, and related abstractions such
-as subroutines, procedures, methods, etc., are units of computation
-that call one another to modularly realize larger computations.
-%
-At a low level, each function activation manages its own data---local
-variables, etc.---as well as information about the \emph{caller} to
-which it must return.
-%
-The \emph{call stack} is the fundamental data structure used to
-implement functions, aided by an ABI that defines how registers are shared
-between activations.
-
-From a security perspective, attacks on the call stack
-are attacks on the function abstraction itself.
-%
-Indeed, the stack is an ancient~\cite{phrack96:smashingthestack} and
+Computer programming relies on functions (or subroutines, procedures, methods, etc.)
+as a foundational abstraction. Functions are units of computation that
+can call one another to realize larger computations in a modular fashion.
+%
+At a low level, a function activation manages its own data, such as local variables, as
+well as information about its \emph{caller} function, to which it must return.
+The \emph{(call) stack} is the fundamental data structure charged with realizing
+this abstraction, aided by an ABI that defines how registers are shared between activations.
+From a security perspective, attacks on the stack are really attacks
+on the fundamental functional abstraction itself, and are therefore every bit as
+pervasive.
+
+Indeed, the call stack is an ancient~\cite{phrack96:smashingthestack} and
 perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
@@
   DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
-target for low-level attacks, sometimes involving control-flow
-hijacking via corrupting the return address, sometimes memory corruption
-more generally.  
+target for low-level attacks, frequently involving control-flow hijacking via corrupting
+the return address, and memory corruption more generally.
 %
@@
 %using data stored in the stack 
-The variety in attacks on the stack is mirrored in the range of
+The variety in attacks on the stack is mirrored by the
 software and hardware protections that aim to prevent them,
@@
 
-Enforcement mechanisms can be brittle, successfully eliminating one
-attack while leaving room for others. To avoid an endless game of
-whack-a-mole, we seek formal properties of safe behavior that can be
-proven or at least rigorously tested. Such properties can be used as
-the specification against which enforcement can be validated; even
-enforcement mechanisms that do not fulfill a property benefit from the
-ability to articulate why and when they may fail.
-
-Of the mechanisms listed above, many are fundamentally ill suited for
-offering formal guarantees: may impede attackers, but do not provide
-universal protection. Shadow stacks, for instance, aim to ``restrict
-the flexibility available in creating gadget chains''
-\cite{Shanbhogue+19}, not to categorically rule out attacks. Other
-mechanisms, such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
-integrity\cite{Kuznetsov+14}, do aim for stronger guarantees, but even
-these do not make formal promises.  Indeed, to our knowledge, the sole
-line of work making a formal claim to protect stack safety is the
-secure calling conventions of Skorstengaard et
-al.~\cite{SkorstengaardSTKJFP} and Georges et
-al.~\cite{Georges22:TempsDesCerises}.
-
-Other mechanisms besides these should also be amenable to strong
-formal guarantees.  In particular, Roessler and DeHon
-\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
-micro-policies~\cite{pump_oakland2015} for stack safety that offer
-strong guarantees.  Their most realistic micro-policy, called Lazy
-Tagging and Clearing (\(LTC\)), makes an interesting performance
-trade-off: it allows function activations to improperly write into one
-another's stack frames, but ensures that the owner of the corrupted
-memory cannot access it afterward.
-
-Under this policy, one function activation {\em can} corrupt another's
-memory---just not in ways that do not affect observable behavior.
-Therefore, \(LTC\) would not fulfill Georges et al.'s property
-(appropriately adapted to the tagged setting).  But \(LTC\) does
-arguably enforce stack safety.  A looser, more observational
-definition of stack safety is needed to fit this situation.
-
-We propose here a formal characterization of stack safety, based on
-the intuition of protecting function activations from each other.
-\bcp{Aren't other formal definitions based on the same intuition?}
-We use the tools of language-based
+Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
+other, similar attacks. To avoid this \leo{game of whack-a-mole}, we seek formal properties of safe behavior that can be proven
+or tested. Such properties become the specification against which enforcement can be validated;
+even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
+\emph{why} they fail.
+
+Of the mechanisms above, many are fundamentally ill-suited for offering formal guarantees,
+as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
+for instance, aim to ``restrict the flexibility available in creating gadget chains''
+\cite{Shanbhogue+19}, not to categorically
+rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
+integrity\cite{Kuznetsov+14} are stronger, but most do not give formal guarantees.
+Indeed, the secure calling conventions of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
+Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
+to offer stack safety formalizations.
+
+Among the mechanisms that should be amenable to making strong formal guarantees,
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
+micro-policies~\cite{pump_oakland2015} for stack safety.
+Their most realistic micro-policy, ``Lazy Tagging and Clearing'' (\(LTC\)),
+makes an interesting trade-off for performance: it allows function activations to write
+into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
+it afterward. Because the memory does get corrupted, \(LTC\) would not fulfill an
+adaptation of Georges et al.'s property to the tagged setting.
+
+Neverthess, we believe that there is a useful sense in which \(LTC\) does (attempt to) enforce stack safety. 
+While it does not always protect individual addresses on the stack, it tries to 
+prevent one function activation from interfering with another's behavior---protecting 
+the functional abstraction, which we argue is the real essence of stack safety.
+
+In this work, we propose a formal characterization of stack safety based on the idea
+of protecting each function activation from other activations.
+We use the formal tools of language-based
 security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
@@
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the
-{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
-Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
-While our properties are motivated by the desire to specify \(LTC\) precisely, 
-they are stated abstractly in hopes that they may also apply
-to other enforcement mechanisms.
-
-We find, via property-based random testing, that \(LTC\) is flawed in
-a way that undermines both integrity and confidentiality; after
-correcting this flaw, the repaired \(LTC\) satisfies all our
-properties.  Moreover, we show that our properties can be extended to
-describe calling conventions with features such as callee-save
-registers, argument-passing on the stack, and tail-call
-elimination. We modify \(LTC\) to protect these features and apply
-random testing to validate this extended protection mechanism against
-the extended properties.
+and the callee's behavior during the callee's execution, plus the control-flow protection
+of {\em well-bracketed control flow} (\(\wbcf\))~\cite{SkorstengaardSTKJFP}.
+While our properties are motivated by a desire to specify \(LTC\), 
+they are built from first principles in the hopes that they will be broadly applicable
+to a range of enforcement mechanisms.
+
+The structure of our properties enables them to be easily expanded to characterize systems with
+sophisticated features such as callee-save registers, argument-passing on the stack,
+and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
+test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
+is flawed in a way that undermines integrity and confidentiality, and we correct
+this flaw.
 
@@
 example model, and then moving to a more realistic model with multiple new features:
-callee-save registers, argument passing on the stack, and tail-call elimination, simply
+callee-save registers, argument passing on the stack, and tailcall elimination, simply
 by refining a few definitions in (1). This becomes the model that we validate.
@@
 well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
-address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
+address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
 to enforce it completely. Others combine this protection with some degree of memory protection,
-chiefly focusing on integrity. Interestingly, Skorstengaard et al. \cite{SkorstengaardSTKJFP}
+chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
 describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
@@
 \begin{itemize}
-\item We give a novel characterization of stack safety as a collection
-of properties: confidentiality and integrity for callee and caller,
-plus well-bracketed control-flow. Different subsets of these
-properties characterize the security guarantees offered by mechanisms
-that support only some aspects of stack safety\bcp{where do we show
-  this?}.  The properties they are parameterized over a notion of
-external observation, allowing them to characterize lazy enforcement
-mechanisms.
-\item We instantiate\bcp{extend??} these definitions in a realistic
-setting with features such as\bcp{``such as''? are there others?}
-argument passing on the stack, callee-saves registers, and tail-call
-elimination.\bcp{How about: We extend these core definitions to
-  describe a realistic setting with argument passing on the stack,
-  callee-saves registers, and tail-call elimination.}  Our model is
-modular enough that handling these features is straightforward.
-\bcp{Throughout, the way we talk about the ``core model'' and its
-  intantiation / extension / enrichment with realistic features feels
-  quite weasily.}
-\item We validate tag-based stack-safety enforcement via
-property-based randomize testing. We find that \emph{Lazy Tagging and
-  Clearing} falls short of ensuring stack safety, propose a fix, and
-validate it.
+\item A novel characterization of stack safety as a collection of properties:
+  confidentiality and integrity for callee and caller, and well-bracketed
+  control-flow. These can characterize the security requirements of systems
+  that support only some portion of stack safety, and are
+  parameterized over a notion of external observation, so that they can characterize
+  lazy enforcement.
+\item An instantiation of these definitions in a realistic
+  setting with features such as argument passing
+  on the stack, callee-saves registers, and tail-call elimination.
+  Our model is modular enough that handling these features is straightforward.
+\item The validation, via property-based randomize testing, of tag-based
+  stack-safety policies. We find that \emph{Lazy Tagging and Clearing}
+  falls short of stack safety, and propose a fix.
 \end{itemize}
@@
 
-\section{Building a Property}
+\section{Framework and Assumptions}
 \label{sec:ideas}
 
-Our properties follow a three-part structure that we term a {\em security semantics}.
-First we define the {\em security-relevant operations} that we support, and an extra
-piece of state called a {\em security context}, which updates alongside each step
-of execution. The security context gives each function activation a {\em view},
-mapping each element to a {\em security class}.
-
-Then, we define our high-level security requirements---integrity, confidentiality,
+Stack safety properties need to describe the behavior of machine code. but they naturally 
+talk about function activations and stack contents---abstractions that
+are (typically) not visible at machine level. To bridge this gap, 
+our properties are defined in terms of a {\em security semantics} layered on top of
+the standard execution semantics of the machine.  This semantics identifies certain
+state transitions of the machine as {\em security-relevant operations}, which update
+a notional {\em security context}.  This context consists of
+an (abstract) stack of function activations, each associated with a {\em view} 
+that maps each element of the machine state (memory location or register)
+to a {\em security class} (active, sealed, etc.) which specifies how the activation
+can access the element.
+The action of the security-relevant operations on the context is defined by a set
+of rules that depend on how the machine code makes use of the stack and registers
+to implement the function abstraction.
+
+We define stack safety properties in
+terms of traces of {\em observable events} rather than internal machine states. 
+This is particularly important for
+specifying lazy enforcement, which works because the deferred check catches problems 
+before they can affect the trace.
+Events might include certain function calls (i.e., system calls performing I/O)
+or writes to special addresses representing memory-mapped regions, etc.
+%The nature of events is a parameter of the framework, and can be specialized
+%to any notion of observable behavior.
+
+Finally, we define our high-level security properties---integrity, confidentiality,
 and well-bracketed control flow---in terms of the security classes of state
@@
 assertions about future return states, and comparison of event traces to capture the
-desired behavior without reference to implementation details.
-
-Separately, in order to apply a property to a specific machine, we do need to consider
-implementation details, by determining how the machine's execution translates into
-security relevant operations. But this step is transparent to the properties themselves.
-
-\apt{Something about how the machine can incorporate enforcement, with examples
-(tags, capabilities, SW only). }
+desired behavior without reference to internal implementation details.
+We introduce these properties by example in \cref{sec:example}. In the remainder
+of this section we describe the underlying framework in more detail.
+
+Our security semantics has similarities to the overlay semantics proposed
+by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it does not restrict
+the behavior of the underlying machine in any way. Rather, it tracks additional context
+about the history of security-relevant operations, which inform the criterion a machine
+must satisfy in order to correctly obey the properties.\apt{Sharpen this comparison?}
+
+%The security principals in our account of stack safety are function
+%activations: the context tracks, for the active function and all pending
+%functions, their {\em view} of the security state of the system.
+
+\paragraph*{Machine model}
+We assume a conventional processor ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
+and stack pointer, and a standard ABI.
+We make no particular assumptions about the provenance of the machine code; in particular,
+we do not assume the use of any particular compiler.
+The machine may possibly be enhanced with security
+enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
+we assume that the behavior of these mechanisms is incorporated into the basic
+step semantics of the machine, but we distinguish the \emph{payload} part of
+each state element---the part relevant to the behavior of the basic machine---from
+any additional enforcement information (such as a tag).
 
 \paragraph*{Security Semantics and Property Structure}
-
 A security semantics extends a machine
@@
 selected instructions, but the theory does not depend on this choice.
-
-Whatever the mechanism for labeling transitions, we get a base
-machine transition \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
-an event (see below) and \(\bar{\psi}\) is a list of security-relevant operations.
-%We then lift this into a transition between pairs of machine states and contexts,
-%\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
+Whatever the mechanism for labeling transitions, we end up with an annotated version of the 
+machine transition function \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
+an event and \(\bar{\psi}\) is a list of security-relevant operations.
+
+We then lift this into a transition between pairs of machine states and contexts,
+by applying a set of operation-specific rules to transform each state and context to
+a new context, in parallel with the ordinary transition on states.
+The most important rules describe call and return operations.
+A call pushes a new activation view onto the context stack and changes the class of the
+caller's data to protect it from the new callee; a return reverses these steps.
+Other operations can serve to signal how parts of the stack frame are being used to store
+or share data, and their corresponding rules alter the classes of different
+state elements accordingly.
+
+Exactly which operations and rules are needed depends on
+what code features we wish to support. 
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
@@
     \example \(\mathbf{dealloc}\) & offset \& size \\
-    \testing \(\mathbf{tail call}\) & (same as for \(\mathbf{call}\)) \\
+    \testing \(\mathbf{tailcall}\) & (same as for \(\mathbf{call}\)) \\
     \hline
@@
 
-Our security semantics has similarities to the overlay semantics that have been proposed
-to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
-the behavior of the underlying machine in any way. Rather, it tracks additional context
-about the history of security-relevant operations, which inform the criterion a machine
-must satisfy in order to correctly obey the properties.
-
-The security principals in our account of stack safety are function
-activations: the context tracks, for the active function and all pending
-functions, their {\em view} of the security state of the system.
-
-\paragraph*{Views}
+\paragraph*{Views and security classes}
 
@@
 
-\paragraph*{Events}
-
-It is common to specify the behavior of a system, including security properties, in
-terms of traces of observable {\em events} rather than details of the machine state.
-Lazy enforcement works because the deferred check catches errors before they appear
-in the trace. The nature of events is a parameter of the machine, and can be specialized
-to any notion of observable behavior.
-
-\paragraph*{Variants and Non-interference}
-
-To characterize situations in which values should be kept secret or prevented
-from influencing future execution, we borrow the idea of \emph{variant} states
-from the theory of non-interference. Non-interference is a way of expressing
-the flow of information through the system. If we begin with a state \(\mach\)
-and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
-then execute both \(\mach\) and \(\nach\), any differences in the outputs they
-produce leaks information about \(\component\).
+%% \paragraph*{Variants and Non-interference}
+
+%% To characterize situations in which values should be kept secret or prevented
+%% from influencing future execution, we borrow the idea of \emph{variant} states
+%% from the theory of non-interference. Non-interference is a way of expressing
+%% the flow of information through the system. If we begin with a state \(\mach\)
+%% and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
+%% then execute both \(\mach\) and \(\nach\), any differences in the outputs they
+%% produce leaks information about \(\component\).
+
+\paragraph*{Instantiating the Framework}
+
+The property framework has many parameters to support broad applicability.
+Conceptually, the following steps are needed to instantiate the framework to a specific machine
+and code style: (i) define the base machine semantics; (ii) distinguish payload
+from enforcement parts (if any) of the machine state; (iii) identify the set of
+security-relevant operations and rules required by the code style; (iv) determine
+how to label (or otherwise associate) machine instruction with security-relevant
+operations as appropriate; (v) specify the form of observable events.
+
+In \cref{sec:example} we give such an instantiation, and use it to explain the
+security properties.  But we stress that the property definitions themselves
+are \emph{independent} of the parameter instantiation.
+
+\apt{Suggest moving threat model to here.}
 
@@
 of a call.
-%Likewise, if \(\mathbf{tail call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
-%\(s\) is the target of a tail call.
+%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
+%\(s\) is the target of a tailcall.
 As a shorthand, we write that each property is defined
@@
                {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{tail call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
+               {\(Op ~ \mach ~ (\mathbf{tailcall} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
                  ~ (V,\sigma) = (V',\sigma)\)}
@@
   \end{subfigure}
-  \caption{Operations supporting tail calls and argument passing on stack.}
+  \caption{Operations supporting tailcalls and argument passing on stack.}
   \label{fig:advops}
@@
 
-The operation rule for a tail call is similar to that for a normal call.
+The operation rule for a tailcall is similar to that for a normal call.
 We do not push the caller's view onto the stack,
-but replace it outright. This means that a tail call does not increase the size of
-the call stack, and therefore for purposes of our properties, all tail
-calls will
+but replace it outright. This means that a tailcall does not increase the size of
+the call stack, and therefore for purposes of our properties, all tailcalls will
 be considered to return simultaneously when the eventual {\bf return} operation
@@
 than \(\sealed\) to express this. In terms of \cref{tab:props}, we replace
-``call'' with ``call or tail call'' in \(\clrc\), \(\clec\), and \(\clei\).
+``call'' with ``call or tailcall'' in \(\clrc\), \(\clec\), and \(\clei\).
 
@@
 SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
-DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey in
-Sections~\ref{sec:intro} and~\ref{sec:future}. Among these techniques, the micro-policies of Roessler and
+DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey briefly in
+\cref{sec:intro}. Among these techniques, the micro-policies of Roessler and
 DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
@@
 %
-Previous work has used these principles in simplified models to implement
+Previous work has uses these principles in simplified models to implement
 \emph{secure calling conventions} for the call

### 2023-02-04 submission5/core.tex
Mostly figures

--- 2023-02-04 submission5/core.tex
@@
 
-Other mechanisms besides these should also be amenable to strong
-formal guarantees.  In particular, Roessler and DeHon
-\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
-micro-policies~\cite{pump_oakland2015} for stack safety that offer
-strong guarantees.  Their most realistic micro-policy, called Lazy
-Tagging and Clearing (\(LTC\)), makes an interesting performance
-trade-off: it allows function activations to improperly write into one
-another's stack frames, but ensures that the owner of the corrupted
-memory cannot access it afterward.
-
-Under this policy, one function activation {\em can} corrupt another's
-memory---just not in ways that do not affect observable behavior.
-Therefore, \(LTC\) would not fulfill Georges et al.'s property
-(appropriately adapted to the tagged setting).  But \(LTC\) does
-arguably enforce stack safety.  A looser, more observational
-definition of stack safety is needed to fit this situation.
-
-We propose here a formal characterization of stack safety, based on
-the intuition of protecting function activations from each other.
-\bcp{Aren't other formal definitions based on the same intuition?}
-We use the tools of language-based
+Fortunately, other mechanisms besides these should be
+amenable to strong 
+formal guarantees.  In particular, 
+Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} have presented an
+array of tag-based micro-policies~\cite{pump_oakland2015} for stack
+safety that give strong guarantees.
+Their most realistic micro-policy, called Lazy Tagging and Clearing (\(LTC\)),
+makes an interesting performance trade-off: it allows function
+activations to improperly write into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
+it afterward.
+
+Because memory does get corrupted, \(LTC\) would not fulfill an
+adaptation of Georges et al.'s property to the tagged setting.  But
+there is a useful sense in which \(LTC\) does enforce stack safety.
+While it does not necessarily protect individual addresses on the
+stack, it does prevent one function activation from interfering with
+another's observable behavior.
+
+We propose a formal characterization of stack safety based on the intuition
+of protecting each function activation from other activations.
+We use the formal tools of language-based
 security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
@@
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the
-{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
+and the callee's behavior during the callee's execution, plus the control-flow protection
+of {\em well-bracketed control flow} (\(\wbcf\)) articulated by
 Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
 While our properties are motivated by the desire to specify \(LTC\) precisely, 
-they are stated abstractly in hopes that they may also apply
+they are stated somewhat abstractly in hopes that they may be applicable
 to other enforcement mechanisms.
 
-We find, via property-based random testing, that \(LTC\) is flawed in
-a way that undermines both integrity and confidentiality; after
-correcting this flaw, the repaired \(LTC\) satisfies all our
-properties.  Moreover, we show that our properties can be extended to
-describe calling conventions with features such as callee-save
-registers, argument-passing on the stack, and tail-call
-elimination. We modify \(LTC\) to protect these features and apply
-random testing to validate this extended protection mechanism against
-the extended properties.
+The structure of our properties enables them to be easily extended\bcp{??} to characterize systems with
+features such as callee-save registers, argument-passing on the stack,
+and tail-call elimination. We likewise extend \(LTC\) to protect these features, and
+test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
+is flawed in a way that undermines integrity and confidentiality, and we correct
+this flaw.
 
@@
   and should conjecture claim them in passing.}
-
-\bcp{Agreed.  I think in particular the first paragraph can just go.
-  But this is not enough.  Also, the section is trying to do at least
-  two things---talk about earlier work and how we compare, and talk
-  about why we believe a flexible approach is good.  The whole
-  discussion is a bit wooly and confusing.}
 
@@
 \begin{itemize}
-\item We give a novel characterization of stack safety as a collection
-of properties: confidentiality and integrity for callee and caller,
-plus well-bracketed control-flow. Different subsets of these
-properties characterize the security guarantees offered by mechanisms
-that support only some aspects of stack safety\bcp{where do we show
-  this?}.  The properties they are parameterized over a notion of
-external observation, allowing them to characterize lazy enforcement
-mechanisms.
-\item We instantiate\bcp{extend??} these definitions in a realistic
-setting with features such as\bcp{``such as''? are there others?}
-argument passing on the stack, callee-saves registers, and tail-call
-elimination.\bcp{How about: We extend these core definitions to
-  describe a realistic setting with argument passing on the stack,
-  callee-saves registers, and tail-call elimination.}  Our model is
-modular enough that handling these features is straightforward.
-\bcp{Throughout, the way we talk about the ``core model'' and its
-  intantiation / extension / enrichment with realistic features feels
-  quite weasily.}
-\item We validate a tag-based enforcement mechanism, \emph{Lazy
-  Tagging and Clearing}, via property-based random testing, find that
-it falls short, and propose and validate a fix.
+\item A novel characterization of stack safety as a collection of properties:
+  confidentiality and integrity for callee and caller, and well-bracketed
+  control-flow. These can characterize the security requirements of systems
+  that support only some portion of stack safety, and are
+  parameterized over a notion of external observation, so that they can characterize
+  lazy enforcement.
+\item An instantiation of these definitions in a realistic
+  setting with features such as argument passing
+  on the stack, callee-saves registers, and tail-call elimination.
+  Our model is modular enough that handling these features is straightforward.
+\item The validation, via property-based randomize testing, of tag-based
+  stack-safety policies. We find that \emph{Lazy Tagging and Clearing}
+  falls short of stack safety, and propose a fix.
 \end{itemize}
 
-In the next section, we give a brief overview of the structure of
-properties.  In \cref{sec:example}, we walk through a function call in
-a simple example machine and discuss how each of our properties
-applies to it, informally. In the process we motivate the properties
-from a security perspective.
-%
-In \cref{sec:formal} we formalize the machine model,
-its {\em security semantics}, and the stack safety properties built on these.
-In \Cref{sec:enforcement} we describe the micro-policies that we test,
-in \cref{sec:testing} the testing framework itself, and in
-\cref{sec:relwork,sec:future} related and future work.
-
-The accompanying artifact contains formal definitions (in Coq) of our
-properties, plus our testing framework.  It does not include proofs:
-we use Coq primarily for the QuickChick testing library and for
-ensuring that our definitions are unambiguous.  Formal proofs are left
-as future work.
+Our artifact contains Coq formalizations of our properties for illustrative purposes
+as well as their testing implementation. There are no proofs---we use Coq primarily
+for the QuickChick library and to ensure that the formalizations are unambiguous.
+Formal proofs are left as future work.
 
@@
 
-When our properties are used to evaluate a system, the threat model
-will depend on the details of that system. However, there are some
-constraints that our design puts on any system. In particular, we must
-trust that the method of distinguishing security-relevant operations
-is accurate.  For example, if the enforcement mnechanism involves
-labels placed on code by a compiler, we must trust that the compiler
-placed those labels correctly. For instance, if a compiled function
-call is not marked as such, then the caller's data might not be
-protected from the callee.  Conversely, marking too many operations as
-calls might mean that the property becomes too strong to be
-enforced.\bcp{I don't understand that---wouldn't it mean that
-  enforcing the property causes almost all programs to be rejected?
-  (I.e., isn't it a case of false negatives rather than false
-  positives?)}
-
-\bcp{Still working here:}
-
-We do not assume that low-level code adheres to any fixed calling
-convention or implements any particular source-language constructs.
-Indeed, if the source
-language is C, then high-level programs might contain undefined
-behavior, in which case they might be compiled to arbitrary machine
-code. A given enforcement 
+When our properties are used to evaluate a system, the threat model will depend on the
+details of that system. However, there are some constraints that our design puts on
+any system. We must trust that the method of distinguishing security-relevant operations is accurate; if it
+involves labels placed on code by a compiler, that means trusting that the compiler placed
+those labels correctly. If operations occur that are not recognized, those operations
+might not be guaranteed to protect their principals---for instance, an unlabeled call
+might not protect the caller's data from the callee. On the flip side, applying an incorrect
+label most likely means that the property becomes too strong to be enforced.
+
+Otherwise, we do not assume that the code adheres to any particular
+calling convention or implements a source language construct.
+In particular, while we are agnostic as to the source
+language, we certainly aim to support C, and so any source function might contain undefined
+behavior resulting in its compilation to arbitrary machine code. A given enforcement
 mechanism or target architecture might place additional constraints, particularly on the behavior of
@@
 which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
-Roessler and DeHon include the compiler in their trusted computing
-base, so our properties 
+Roessler and DeHon have the compiler in their trusted computing base, so our properties
 are stronger than necessary to characterize their micro-policies.
 
-In general, it is impossible to distinguish buggy machine code from an
-attacker.  In our examples we will identify one function or another as
-an attacker, but we do not require any static division between trusted
-and untrusted code, and we aim to protect even buggy code.
-
-This is a strong threat model, but it does omit some important aspects
-of stack safety in real systems: in particular, it does not address
-concurrency.  Hardware and timing attacks are also out of scope.
-%
+In general, it is impossible to distinguish buggy machine code from an attacker; in
+our examples we will identify one function or another as an attacker, but we do not
+require any static division between trusted and untrusted code, and we aim to protect
+even buggy code.
+
+This is a strong threat model, but it does omit some important aspects of
+security enforcement: in particular, hardware and timing attacks are out of scope, and
+we do not address concurrency---that is a high-priority future work.
+%
+
+\paragraph*{Organization}
+
+In the next section, we will give a brief overview of the structure of properties.
+Then in \cref{sec:example}, we will walk through a function
+call in a simple example machine and discuss how each of our properties applies to it,
+informally. In the process we discuss some motivation for the properties, from a security
+perspective.
+
+In \cref{sec:formal} we formalize the machine model, 
+its {\em security semantics}, and the stack safety properties built on these. 
+\Cref{sec:enforcement} describes the micro-policies that we test,
+\cref{sec:testing} the testing framework itself, and
+\cref{sec:relwork,sec:future} discuss related and future work.
 
@@
 
-\newcommand{\freebox}[1][10pt]{\tikz \filldraw[fill=blue] (0,0) rectangle (#1,#1);}
-\newcommand{\pubbox}[1][10pt]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (#1,#1);}
-\newcommand{\objbox}[1][10pt]{\tikz \filldraw[fill=yellow] (0,0) rectangle (#1,#1);}
-\newcommand{\sealbox}[1][10pt]{\tikz \filldraw[fill=red] (0,0) rectangle (#1,#1);}
-
-\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30pt,30pt) node[pos=.5]{#1};}
-\newcommand{\numbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
-\newcommand{\leftbox}[1][]{\tikz \filldraw[fill=blue!20] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
-\newcommand{\rightbox}[1][]{\tikz \filldraw[fill=orange!20] (0,0) rectangle (20pt,20pt) node[pos=.5]{#1};}
-
 \begin{figure}
+  \newcommand{\figonebox}[1][]{\genbox{30px}{lightgray}{#1}}
   \begin{subfigure}{\columnwidth}
@@
     \vspace{\abovedisplayskip}
-    \(\textnormal{\large \dots} \pubbox[30pt] \!\! \stackrel{\textsc{\normalsize sp}}{\textbox[\tt res]}
-    \stackrel{\textsc{\normalsize 4(sp)}}{\textbox[\tt sens]}
-    \stackrel{\textsc{\normalsize 8(sp)}}{\textbox[\tt sec]}
-    \stackrel{\textsc{\normalsize 12(sp)} \hfill}{\textbox[\(\mathtt{ra_1}\)]\textbox[\(\mathtt{ra_2}\)]}
-    \textnormal{\large \dots}\)
+    \(\xleftarrow{\figonebox[\dots] \stackrel{\textsc{\normalsize sp}}{\figonebox[\tt res]}
+    \stackrel{\textsc{\normalsize 4(sp)}}{\figonebox[\tt sens]}
+    \stackrel{\textsc{\normalsize 8(sp)}}{\figonebox[\tt sec]}
+    \stackrel{\textsc{\normalsize 12(sp)} \hfill}{\figonebox[\(\mathtt{ra_1}\)]\figonebox[\(\mathtt{ra_2}\)]}
+    }\)
   \end{subfigure}
@@
 
+\newcommand{\figfourbox}[1][]{\genbox{20px}{red}{#1}}
+
 \begin{figure}
@@
   \[
-  \stackrel{\texttt{res}}{\numbox[0]}
-  \stackrel{\texttt{sens}}{\numbox[0]}
-  \stackrel{\texttt{sec}}{\numbox[5]}
-  \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+  \stackrel{\texttt{res}}{\figfourbox[0]}
+  \stackrel{\texttt{sens}}{\figfourbox[0]}
+  \stackrel{\texttt{sec}}{\figfourbox[5]}
+  \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}\]
   %
@@
   \[
-  \stackrel{\texttt{res}}{\numbox[0]}
-  \stackrel{\texttt{sens}}{\numbox[\bf 42]}
-  \stackrel{\texttt{sec}}{\numbox[5]}
-  \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+  \stackrel{\texttt{res}}{\figfourbox[0]}
+  \stackrel{\texttt{sens}}{\genbox{20px}{red!50}{\bf 42}}
+  \stackrel{\texttt{sec}}{\figfourbox[5]}
+  \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}\]
   %
   \[\overbrace{
-    \stackrel{\texttt{res}}{\numbox[0]}
-    \stackrel{\texttt{sens}}{\numbox[\bf 42]}
-    \stackrel{\texttt{sec}}{\numbox[5]}
-    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}
+    \stackrel{\texttt{res}}{\figfourbox[0]}
+    \stackrel{\texttt{sens}}{\genbox{20px}{\leftvariant}{\bf 42}}
+    \stackrel{\texttt{sec}}{\figfourbox[5]}
+    \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}
     \hspace{1cm}
-    \stackrel{\texttt{res}}{\numbox[0]}
-    \stackrel{\texttt{sens}}{\numbox[\bf 0]}
-    \stackrel{\texttt{sec}}{\numbox[5]}
-    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}}\]
-  \[\large\hookrightarrow 5 \hspace{1cm} \large\hookrightarrow 1\]
+    \stackrel{\texttt{res}}{\figfourbox[0]}
+    \stackrel{\texttt{sens}}{\genbox{20px}{\rightvariant}{\bf 0}}
+    \stackrel{\texttt{sec}}{\figfourbox[5]}
+    \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}}\]
+  \[\stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
+  \raisebox{\height}{\not\approx}
+  \hspace{0.5cm}
+  \stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{1}}\]
 
@@
 
+\newcommand{\leftbox}[1][]{\genbox{20px}{\leftvariant}{#1}}
+\newcommand{\rightbox}[1][]{\genbox{20px}{\rightvariant}{#1}}
+
 \begin{figure}
@@
     \[
-    \stackrel{\texttt{res}}{\numbox[0]}
-    \stackrel{\texttt{sens}}{\numbox[0]}
-    \stackrel{\texttt{sec}}{\numbox[5]}
-    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+    \stackrel{\texttt{res}}{\figfourbox[0]}
+    \stackrel{\texttt{sens}}{\figfourbox[0]}
+    \stackrel{\texttt{sec}}{\figfourbox[5]}
+    \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}\]
     %
@@
     \]
-    \[\large\Downarrow \hspace{1cm} \large\hookrightarrow 5 \hspace{0.5cm} \not \approx \hspace{0.5cm}
-    \large\hookrightarrow 4 \hspace{1cm} \large\Downarrow\]
+    \[\raisebox{.5\height}{\Bigg\Downarrow} \hspace{1cm}
+    \stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
+    \raisebox{\height}{\not\approx}
+    \hspace{0.5cm}
+    \stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{4}} \hspace{1cm}
+    \raisebox{.5\height}{\Bigg\Downarrow}\]
     \[
-    \stackrel{\texttt{res}}{\leftbox[\textbf{0}]}
+    \stackrel{\texttt{res}}{\leftbox[0]}
     \stackrel{\texttt{sens}}{\leftbox[0]}
@@
     %
-    \stackrel{\texttt{res}}{\rightbox[\textbf{2}]}
+    \stackrel{\texttt{res}}{\rightbox[2]}
     \stackrel{\texttt{sens}}{\rightbox[3]}

### 2023-02-04 submission5/core.tex
CP intro

--- 2023-02-04 submission5/core.tex
@@
 
-\bcp{Agreed.  I think in particular the first paragraph can just go.
+\bcp{Agreed.  I think in particular the first paragraph can just go
+  (I've deleted it).
   But this is not enough.  Also, the section is trying to do at least
@@
 
-Security properties should be motivated by security requirements
-and flexible enough to support various sets of security requirements depending on the
-goals of the system that they validate.
-Following this reasoning, our vision of stack safety is as a collection of properties
-describing the security needs of a system with regard to its functional abstractions.
-While the stack still plays a key role, our stack safety is about functions. The most
-dramatic consequence of this is that we treat registers as part of a function activation's
-state, and hold them to the same protection requirements as stack memory.
-
-Flexibility is especially important for stack safety.
-%
-Most security properties are described at the level of a high-level programming language,
-and translated to the target-level via a secure compiler. Stack safety cannot
-be defined in this way, as the abstraction is completely transparent in most source languages.
-(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
-which is broadly consistent across language semantics, and therefore
-amenable to a very general high-level approach.)
-%
-But it is not satisfying as a low-level property; at the level of the ABI, the
-specification of a well-behaved stack is minimal. The ABI is not concerned with such
-questions as whether a caller's frame should be readable or writable to its callee,
-leaving those concerns to be determined by a high-level language built atop the stack.
-These are high-level security questions.
+% Security properties should be motivated by security requirements
+% and flexible enough to support various sets of security requirements depending on the
+% goals of the system that they validate.
+% Following this reasoning, our vision of stack safety is as a collection of properties
+% describing the security needs of a system with regard to its functional abstractions.
+% While the stack still plays a key role, our stack safety is about functions. The most
+% dramatic consequence of this is that we treat registers as part of a function activation's
+% state, and hold them to the same protection requirements as stack memory.
+
+Flexibility is important for talking rigorously about stack safety.
+%
+The reason is that, while many security properties can be described at
+the level of a high-level programming language and translated to a
+target machine by a secure compiler, stack safety cannot be defined in
+this way, since ``the stack'' is completely transparent in the
+definitions of most source languages.\footnote{Contrast Azevedo de
+  Amorim et al.'s work on heap safety
+  \cite{DBLP:conf/post/AmorimHP18}: the concept of the heap figures
+  directly in high-level language semantics and its security is
+  therefore amenable to a high-level treatment.}
+%
+But neither is it satisfying as an entirely low-level property;
+indeed, at the lowest level, the specification of a ``well-behaved
+stack'' is almost vacuous. The ISA is not concerned with such
+questions as whether a caller's frame should be readable or writable
+to its callee: those are the purview of high-level languages built
+atop the hardware stack.
+
+\bcp{There's a broken transition here: the paragraph below is not an
+  instance of the paragraph above.  Indeed, I don't see how they
+  connect at all.}
 
@@
 In particular, many enforcement techniques focus purely on
-well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
-address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
+well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
+address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
 to enforce it completely. Others combine this protection with some degree of memory protection,
-chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
+chiefly focusing on integrity. Interestingly, Skorstengaard et al. \cite{SkorstengaardSTKJFP}
 describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
@@
 \subsection{Threat Model and Limitations}
-\apt{This could probably go better at end of next section.}
 
@@
 
-\section{Framework and Assumptions}
+\section{Building a Property}
 \label{sec:ideas}
 
-Stack safety properties need to describe the behavior of machine code. but they naturally 
-talk about function activations and stack contents---abstractions that
-are (typically) not visible at machine level. To bridge this gap, 
-our properties are defined in terms of a {\em security semantics} layered on top of
-the standard execution semantics of the machine.  This semantics identifies certain
-state transitions of the machine as {\em security-relevant operations}, which update
-a notional {\em security context}.  This context consists of
-an (abstract) stack of function activations, each associated with a {\em view} 
-that maps each element of the machine state (memory location or register)
-to a {\em security class} (active, sealed, etc.) which specifies how the activation
-can access the element.
-The action of the security-relevant operations on the context is defined by a set
-of rules that depend on how the machine code makes use of the stack and registers
-to implement the function abstraction.
-
-We define stack safety properties in
-terms of traces of {\em observable events} rather than internal machine states. 
-This is particularly important for
-specifying lazy enforcement, which works because the deferred check catches problems 
-before they can affect the trace.
-Events might include certain function calls (i.e., system calls performing I/O)
-or writes to special addresses representing memory-mapped regions, etc.
-%The nature of events is a parameter of the framework, and can be specialized
-%to any notion of observable behavior.
-
-Finally, we define our high-level security properties---integrity, confidentiality,
+Our properties follow a three-part structure that we term a {\em security semantics}.
+First we define the {\em security-relevant operations} that we support, and an extra
+piece of state called a {\em security context}, which updates alongside each step
+of execution. The security context gives each function activation a {\em view},
+mapping each element to a {\em security class}.
+
+Then, we define our high-level security requirements---integrity, confidentiality,
 and well-bracketed control flow---in terms of the security classes of state
@@
 assertions about future return states, and comparison of event traces to capture the
-desired behavior without reference to internal implementation details.
-We introduce these properties by example in \cref{sec:example}. In the remainder
-of this section we describe the underlying framework in more detail.
-
-Our security semantics has similarities to the overlay semantics proposed
-by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it does not restrict
-the behavior of the underlying machine in any way. Rather, it tracks additional context
-about the history of security-relevant operations, which inform the criterion a machine
-must satisfy in order to correctly obey the properties.\apt{Sharpen this comparison?}
-
-%The security principals in our account of stack safety are function
-%activations: the context tracks, for the active function and all pending
-%functions, their {\em view} of the security state of the system.
-
-\paragraph*{Machine model}
-We assume a conventional processor ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
-and stack pointer, and a standard ABI.
-We make no particular assumptions about the provenance of the machine code; in particular,
-we do not assume the use of any particular compiler.
-The machine may possibly be enhanced with security
-enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
-we assume that the behavior of these mechanisms is incorporated into the basic
-step semantics of the machine, but we distinguish the \emph{payload} part of
-each state element---the part relevant to the behavior of the basic machine---from
-any additional enforcement information (such as a tag).
+desired behavior without reference to implementation details.
+
+Separately, in order to apply a property to a specific machine, we do need to consider
+implementation details, by determining how the machine's execution translates into
+security relevant operations. But this step is transparent to the properties themselves.
+
+\apt{Something about how the machine can incorporate enforcement, with examples
+(tags, capabilities, SW only). }
 
 \paragraph*{Security Semantics and Property Structure}
+
 A security semantics extends a machine
@@
 selected instructions, but the theory does not depend on this choice.
-Whatever the mechanism for labeling transitions, we end up with an annotated version of the 
-machine transition function \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
-an event and \(\bar{\psi}\) is a list of security-relevant operations.
-
-We then lift this into a transition between pairs of machine states and contexts,
-by applying a set of operation-specific rules to transform each state and context to
-a new context, in parallel with the ordinary transition on states.
-The most important rules describe call and return operations.
-A call pushes a new activation view onto the context stack and changes the class of the
-caller's data to protect it from the new callee; a return reverses these steps.
-Other operations can serve to signal how parts of the stack frame are being used to store
-or share data, and their corresponding rules alter the classes of different
-state elements accordingly.
-
-Exactly which operations and rules are needed depends on
-what code features we wish to support. 
+
+Whatever the mechanism for labeling transitions, we get a base
+machine transition \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
+an event (see below) and \(\bar{\psi}\) is a list of security-relevant operations.
+%We then lift this into a transition between pairs of machine states and contexts,
+%\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
@@
 
-\paragraph*{Views and security classes}
+Our security semantics has similarities to the overlay semantics that have been proposed
+to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
+the behavior of the underlying machine in any way. Rather, it tracks additional context
+about the history of security-relevant operations, which inform the criterion a machine
+must satisfy in order to correctly obey the properties.
+
+The security principals in our account of stack safety are function
+activations: the context tracks, for the active function and all pending
+functions, their {\em view} of the security state of the system.
+
+\paragraph*{Views}
 
@@
 
-%% \paragraph*{Variants and Non-interference}
-
-%% To characterize situations in which values should be kept secret or prevented
-%% from influencing future execution, we borrow the idea of \emph{variant} states
-%% from the theory of non-interference. Non-interference is a way of expressing
-%% the flow of information through the system. If we begin with a state \(\mach\)
-%% and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
-%% then execute both \(\mach\) and \(\nach\), any differences in the outputs they
-%% produce leaks information about \(\component\).
-
-\paragraph*{Instantiating the Framework}
-
-The property framework has many parameters to support broad applicability.
-Conceptually, the following steps are needed to instantiate the framework to a specific machine
-and code style: (i) define the base machine semantics; (ii) distinguish payload
-from enforcement parts (if any) of the machine state; (iii) identify the set of
-security-relevant operations and rules required by the code style; (iv) determine
-how to label (or otherwise associate) machine instruction with security-relevant
-operations as appropriate; (v) specify the form of observable events.
-
-In \cref{sec:example} we give such an instantiation, and use it to explain the
-security properties.  But we stress that the property definitions themselves
-are \emph{independent} of the parameter instantiation.
-
-\apt{Suggest moving threat model to here.}
+\paragraph*{Events}
+
+It is common to specify the behavior of a system, including security properties, in
+terms of traces of observable {\em events} rather than details of the machine state.
+Lazy enforcement works because the deferred check catches errors before they appear
+in the trace. The nature of events is a parameter of the machine, and can be specialized
+to any notion of observable behavior.
+
+\paragraph*{Variants and Non-interference}
+
+To characterize situations in which values should be kept secret or prevented
+from influencing future execution, we borrow the idea of \emph{variant} states
+from the theory of non-interference. Non-interference is a way of expressing
+the flow of information through the system. If we begin with a state \(\mach\)
+and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
+then execute both \(\mach\) and \(\nach\), any differences in the outputs they
+produce leaks information about \(\component\).
 

### 2023-02-04 submission5/core.tex

--- 2023-02-04 submission5/core.tex
@@
 In particular, many enforcement techniques focus purely on
-well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
+well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
 address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})

### 2023-02-04 submission5/core.tex
Repair bad merge from earlier

@@
 \bcp{Aren't other formal definitions based on the same intuition?}
+\sna{Two differences we're trying to get across. First, protecting
+  functions vs. protecting stack frames: we bring registers into
+  the story because it isn't just about the stack as a data structure.
+  Second: describing security in terms of functions' high-level
+  security goals (f is trying to extract a secret) rather than
+  describing operationally what they can and can't do (as the CHERI
+  folks do.)
+}
 We use the tools of language-based
@@
 
-\bcp{Agreed.  I think in particular the first paragraph can just go
-  (I've deleted it).
+\bcp{Agreed.  I think in particular the first paragraph can just go.
   But this is not enough.  Also, the section is trying to do at least
@@
 
-% Security properties should be motivated by security requirements
-% and flexible enough to support various sets of security requirements depending on the
-% goals of the system that they validate.
-% Following this reasoning, our vision of stack safety is as a collection of properties
-% describing the security needs of a system with regard to its functional abstractions.
-% While the stack still plays a key role, our stack safety is about functions. The most
-% dramatic consequence of this is that we treat registers as part of a function activation's
-% state, and hold them to the same protection requirements as stack memory.
-
-Flexibility is important for talking rigorously about stack safety.
-%
-The reason is that, while many security properties can be described at
-the level of a high-level programming language and translated to a
-target machine by a secure compiler, stack safety cannot be defined in
-this way, since ``the stack'' is completely transparent in the
-definitions of most source languages.\footnote{Contrast Azevedo de
-  Amorim et al.'s work on heap safety
-  \cite{DBLP:conf/post/AmorimHP18}: the concept of the heap figures
-  directly in high-level language semantics and its security is
-  therefore amenable to a high-level treatment.}
-%
-But neither is it satisfying as an entirely low-level property;
-indeed, at the lowest level, the specification of a ``well-behaved
-stack'' is almost vacuous. The ISA is not concerned with such
-questions as whether a caller's frame should be readable or writable
-to its callee: those are the purview of high-level languages built
-atop the hardware stack.
-
-\bcp{There's a broken transition here: the paragraph below is not an
-  instance of the paragraph above.  Indeed, I don't see how they
-  connect at all.}
+\sna{Should probably move the discussion of the CHERI work to related work,
+  and focus on justifying our approach.}
+
+Security properties should be motivated by security requirements
+and flexible enough to support various sets of security requirements depending on the
+goals of the system that they validate.
+Following this reasoning, our vision of stack safety is as a collection of properties
+describing the security needs of a system with regard to its functional abstractions.
+While the stack still plays a key role, our stack safety is about functions. The most
+dramatic consequence of this is that we treat registers as part of a function activation's
+state, and hold them to the same protection requirements as stack memory.
+
+Flexibility is especially important for stack safety.
+%
+Most security properties are described at the level of a high-level programming language,
+and translated to the target-level via a secure compiler. Stack safety cannot
+be defined in this way, as the abstraction is completely transparent in most source languages.
+(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
+which is broadly consistent across language semantics, and therefore
+amenable to a very general high-level approach.)
+%
+But it is not satisfying as a low-level property; at the level of the ABI, the
+specification of a well-behaved stack is minimal. The ABI is not concerned with such
+questions as whether a caller's frame should be readable or writable to its callee,
+leaving those concerns to be determined by a high-level language built atop the stack.
+These are high-level security questions.
 
@@
 In particular, many enforcement techniques focus purely on
-well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
+well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
 address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
@@
 that support only some aspects of stack safety\bcp{where do we show
-  this?}.  The properties they are parameterized over a notion of
+  this?}. \sna{Nowhere clearly enough. We need to make it clear in section 6
+  that Nick and Andre are included there---only aiming at caller-side,
+  not callee-side. The discussion of CHERI work should touch on this as well.}
+The properties they are parameterized over a notion of
 external observation, allowing them to characterize lazy enforcement
@@
 protected from the callee.  Conversely, marking too many operations as
-calls might mean that the property becomes too strong to be
-enforced.\bcp{I don't understand that---wouldn't it mean that
-  enforcing the property causes almost all programs to be rejected?
-  (I.e., isn't it a case of false negatives rather than false
-  positives?)}
+calls will simply cause otherwise safe programs to be rejected.
 
@@
   \[
-  \stackrel{\texttt{res}}{\numbox[0]}
-  \stackrel{\texttt{sens}}{\numbox[\bf 42]}
-  \stackrel{\texttt{sec}}{\numbox[5]}
-  \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
+  \stackrel{\texttt{res}}{\figfourbox[0]}
+  \stackrel{\texttt{sens}}{\genbox{20px}{red!50}{\bf 42}}
+  \stackrel{\texttt{sec}}{\figfourbox[5]}
+  \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}\]
   %
   \[\overbrace{
-    \stackrel{\texttt{res}}{\numbox[0]}
-    \stackrel{\texttt{sens}}{\numbox[\bf 42]}
-    \stackrel{\texttt{sec}}{\numbox[5]}
-    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}
+    \stackrel{\texttt{res}}{\figfourbox[0]}
+    \stackrel{\texttt{sens}}{\genbox{20px}{\leftvariant}{\bf 42}}
+    \stackrel{\texttt{sec}}{\figfourbox[5]}
+    \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}
     \hspace{1cm}
-    \stackrel{\texttt{res}}{\numbox[0]}
-    \stackrel{\texttt{sens}}{\numbox[\bf 0]}
-    \stackrel{\texttt{sec}}{\numbox[5]}
-    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}}\]
-  \[\large\hookrightarrow 5 \hspace{1cm} \large\hookrightarrow 1\]
-
+    \stackrel{\texttt{res}}{\figfourbox[0]}
+    \stackrel{\texttt{sens}}{\genbox{20px}{\rightvariant}{\bf 0}}
+    \stackrel{\texttt{sec}}{\figfourbox[5]}
+    \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}}\]
+  \[\stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
+  \raisebox{\height}{\not\approx}
+  \hspace{0.5cm}
+  \stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{1}}\]
   \caption{Integrity Violation}

### 2023-02-04 submission5/core.tex

@@
       \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\pubbox} \!\! \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
       ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
+      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
       ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}

### 2023-02-04 submission5/core.tex
CP
Responses

@@
 in all.
-
+%
 This formulation is motivated by a particular class of enforcement
@@
 %   these, only exceptions are high-level language features!}
-They are modular by design so that they can be further extended to
-other features.
-
+\proposecut{They are modular by design so that they can be further extended to
+other features.}
+%
 We validate our properties by using them to distinguish between
@@
   describing operationally what they can and can't do (as the CHERI
-  folks do.)
-}
+  folks do.)}
+\bcp{Both useful!  But neither comes across, at least not very clearly.}
 We use the tools of language-based
@@
   and focus on justifying our approach.}
+\bcp{+1}
 

### 2023-02-04 submission5/core.tex
Fix raisebox

--- 2023-02-04 submission5/core.tex
@@
 \(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
-They are related to one another as seen in
-\cref{fig:propertydiagram}:\leo{is this supposed to be \cref{tab:props}?} \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
+\(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
 \(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
@@
     \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}}\]
-  \[\stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
-  \raisebox{\height}{\not\approx}
+  \[\stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
+  %\raisebox{\height}{\not\approx}
   \hspace{0.5cm}
-  \stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{1}}\]
+  \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{1}}\]
   \caption{Integrity Violation}
@@
     \]
-    \[\raisebox{.5\height}{\Bigg\Downarrow} \hspace{1cm}
-    \stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
-    \raisebox{\height}{\not\approx}
+    \[%\raisebox{.5\height}{\Bigg\Downarrow} \hspace{1cm}
+    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
+    %\raisebox{\height}{\not\approx}
     \hspace{0.5cm}
-    \stackrel{\(\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{4}} \hspace{1cm}
-    \raisebox{.5\height}{\Bigg\Downarrow}\]
+    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{4}} \hspace{1cm}
+    %    \raisebox{.5\height}{\Bigg\Downarrow}
+    \]
     \[

### 2023-02-04 submission5/core.tex
CP

--- 2023-02-04 submission5/core.tex
@@
 talk about function activations and stack contents---abstractions that
-are (typically) not visible at machine level. To bridge this gap,
+are typically not visible at machine level. To bridge this gap,
 our properties are defined in terms of a {\em security semantics} layered on top of
-the standard execution semantics of the machine.  This semantics identifies certain
-state transitions of the machine as {\em security-relevant operations}, which update
+the standard execution semantics of the machine.  The security semantics identifies certain
+state transitions of the machine as {\em security-relevant
+  operations}, which update
 a notional {\em security context}.  This context consists of
@@
 that maps each element of the machine state (memory location or register)
-to a {\em security class} (active, sealed, etc.) which specifies how the activation
+to a {\em security class} (active, sealed, etc.) specifying how the activation
 can access the element.
-The action of the security-relevant operations on the context is defined by a set
-of rules that depend on how the machine code makes use of the stack and registers
+The action of a security-relevant operation on the context is defined by a set
+of rules that describe how the machine code makes use of the stack and registers
 to implement the function abstraction.

### 2023-02-04 submission5/core.tex
no message

@@
 The \emph{call stack} is the fundamental data structure used to
-implement functions, aided by an ABI that defines how registers are shared
+implement functions, aided by an ABI\bcp{will all CSF readers know
+  what an ABI is?} that defines how registers are shared
 between activations.
@@
 This is particularly important for
-specifying lazy enforcement, which works because the deferred check catches problems
+specifying lazy enforcement, which catches bad behaviors after they occur but
 before they can affect the trace.
-Events might include certain function calls (i.e., system calls performing I/O)
-or writes to special addresses representing memory-mapped regions, etc.
+Observable events might include specific function calls (e.g., system
+calls that perform I/O), writes to special addresses representing
+memory-mapped regions, etc.
 %The nature of events is a parameter of the framework, and can be specialized
@@
 
-Finally, we define our high-level security properties---integrity, confidentiality,
+Finally, we define high-level security properties---integrity, confidentiality,
 and well-bracketed control flow---in terms of the security classes of state
-elements. These take the form of predicates on states and contexts that use variants,
+elements. These take the form of predicates on states and contexts
+that use variants\bcp{undefined},
 assertions about future return states, and comparison of event traces to capture the
 desired behavior without reference to internal implementation details.
-We introduce these properties by example in \cref{sec:example}. In the remainder
+We introduce these properties by example in \cref{sec:example} and
+formally in \cref{sec:formal}. In the remainder
 of this section we describe the underlying framework in more detail.
 
-Our security semantics has similarities to the overlay semantics proposed
-by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it does not restrict
-the behavior of the underlying machine in any way. Rather, it tracks additional context
-about the history of security-relevant operations, which inform the criterion a machine
-must satisfy in order to correctly obey the properties.\apt{Sharpen this comparison?}
+Our security semantics has similarities to the overlay semantics
+proposed by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it
+does not restrict the behavior of the underlying machine in any
+way\bcp{Will people not familiar with Skorstengaard understand
+  that?}. Rather, it tracks additional context about the history of
+security-relevant operations, which informs the criterion a machine
+must satisfy in order to correctly obey the properties.\apt{Sharpen
+  this comparison?}\bcp{+1.  Or maybe, in the interest of time and to
+  keeping this section flowing, move it to related work.}
 
@@
 \paragraph*{Machine model}
-We assume a conventional processor ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
+We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
 and stack pointer, and a standard ABI.
@@
 each state element---the part relevant to the behavior of the basic machine---from
-any additional enforcement information (such as a tag).
+any additional enforcement information (such as a tag).\bcp{Why?}
 
@@
 A security semantics extends a machine
-with additional context about the identity of current and pending
+with additional context about the identities of current and pending
 functions (which act as security principals) and about the registers and memory they require
 to be secure. This added context is purely notional;
-it does not affect the real machine. The security context
+it does not affect the behavior of the real machine. The security context
 evolves dynamically through the execution of security-relevant operations,
-which include events like calls, returns, and frame manipulations.
+which include events like calls, returns, and frame manipulation.
 Our security properties are phrased in terms of this context, often as predicates
-on future states, e.g. of the form ``when control returns to the current function...'',
-or as relations on traces of future execution (hyper-properties).
-
-The security-relevant operations typically correspond to underling machine instructions,
+on future states (``when control returns to the current function...'')
+or as relations on traces of future execution
+(hyper-properties).\bcp{This is the first we've mentioned
+  hyperproperties.  CSF readers may be surprised.}
+
+Security-relevant operations typically correspond to
+\proposeadd{single?} underlying machine instructions,
 but the correspondence may not be obvious just by inspecting the machine code.
-For example, in the tagged RISC-V machine we use in our examples and tests,
-calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
-and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
-instructions might also be used for other things. An instantiation of an operation
+For example, in the tagged RISC-V architecture that we use in our
+examples and tests,
+calls and returns are conventionally performed using the {\tt jal}
+(``jump-and-link'')
+and {\tt jalr} (``jump-and-link-register'') instructions, but these
+instructions might also be used for other things. An instantiation of
+an operation
 might require some external mechanism for distinguishing which instructions are
-security-relevant. The most obvious technique, which we use in our examples, is to add labels to
-selected instructions, but the theory does not depend on this choice.
+security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique, which we use in our examples, is to add labels to
+selected instructions, but the theory does not depend on this
+choice.\bcp{What could one do instead?}
+
 Whatever the mechanism for labeling transitions, we end up with an annotated version of the
-machine transition function \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
-an event and \(\bar{\psi}\) is a list of security-relevant operations.
-
+machine transition function: we write \(\mach \xrightarrow{\bar{\psi},
+  \obs} \mach'\), where \(\mach\) and \(\mach\) are machine states, \(\obs\) is
+an event, and \(\bar{\psi}\) is a list of security-relevant operations.
+%
 We then lift this into a transition between pairs of machine states and contexts,
@@
 The most important rules describe call and return operations.
-A call pushes a new activation view onto the context stack and changes the class of the
+A call pushes a new activation view\bcp{it was just called ``view'' above} onto the context stack and changes the class of the
 caller's data to protect it from the new callee; a return reverses these steps.
@@
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
-\cref{tab:psi}. Those that we explore in detail in \cref{sec:example,sec:formal}
+\cref{tab:psi}.\bcp{I was confused by the table because I looked at
+  the figure before I read that sentence!} Those that we explore in detail in \cref{sec:example,sec:formal}
 are shown in white: calls, returns, and the allocation and deallocation of local memory.
@@
 We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
-based on pointer provenance in \cref{app:ptr}, represented by the dark gray
+based on pointer provenance in \cref{app:ptr}\bcp{first
+  mention of this appendix; if it's important, mention it in the
+  outline in section I.  Otherwise drop it.}, represented by the dark gray
 operations, but we do not test this.
+
+\bcp{As a reader, I'm feeling a bit overwhelmed at this point.
+  Everything feels fluid and flexible---''It could be like this, but
+  it could also be another way if you want...''  I know we're past the
+  point of doing more than tweaking the way we talk about things, but
+  if there are any tweaks that reduce the sense of flexibility, let's
+  do them!}
 
@@
     \hline
-    \(\psi \in \Psi\) & Parameters \\
+    \(\psi \in \Psi\)\bcp{huh?} & Parameters \\
     \hline

### 2023-02-04 submission5/core.tex
Took a hatchet to 1B.

--- 2023-02-04 submission5/core.tex
@@
 
-\apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
-  Again, we are not providing evidence that these attributes are true, although we can
-  and should conjecture claim them in passing.}
-
-\bcp{Agreed.  I think in particular the first paragraph can just go
-  (I've deleted it).
-  But this is not enough.  Also, the section is trying to do at least
-  two things---talk about earlier work and how we compare, and talk
-  about why we believe a flexible approach is good.  The whole
-  discussion is a bit wooly and confusing.}
-
-\sna{Should probably move the discussion of the CHERI work to related work,
-  and focus on justifying our approach.}
-\bcp{+1}
-
-% Security properties should be motivated by security requirements
-% and flexible enough to support various sets of security requirements depending on the
-% goals of the system that they validate.
-% Following this reasoning, our vision of stack safety is as a collection of properties
-% describing the security needs of a system with regard to its functional abstractions.
-% While the stack still plays a key role, our stack safety is about functions. The most
-% dramatic consequence of this is that we treat registers as part of a function activation's
-% state, and hold them to the same protection requirements as stack memory.
-
-Flexibility is important for talking rigorously about stack safety.
-%
-The reason is that, while many security properties can be described at
+We are not attempting to present a universal definition of stack safety
+for all systems. While many security properties can be described at
 the level of a high-level programming language and translated to a
@@
 questions as whether a caller's frame should be readable or writable
-to its callee: those are the purview of high-level languages built
+to its callee. Those are the purview of high-level languages built
 atop the hardware stack.
 
-\bcp{There's a broken transition here: the paragraph below is not an
-  instance of the paragraph above.  Indeed, I don't see how they
-  connect at all.}
-
-For instance, the stack-safety specification from Georges et al.
-\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
-a callee should always return to its caller. But, in the presence of tail-call elimination, a
-return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
-Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
-their enforcement mechanism did protect the confidentiality of the caller (though not the callee).
-
-These two cases represent two dimensions in which a stack safety definition requires
-flexibility: what constitutes safe behavior differs between systems that support different
-features, and different contexts have different top-level security requirements.
-
-We are not attempting to present a universal definition of stack safety for all systems,
-but we do construct our definition with these dimensions of flexibility in mind.
-Our model is decomposed into:
-(1) a lightweight model of the system and its security-relevant features, from which we
-derive the integrity, confidentiality, and control-flow requirements of each call and
-return, (2) the application of that model to a particular target system, and (3)
-the formal criteria for the integrity, confidentiality, and control flow
-requirements.
-
-We show that we can extend our system to new features by first introducing a simple
-example model, and then moving to a more realistic model with multiple new features:
-callee-save registers, argument passing on the stack, and tail-call elimination, simply
-by refining a few definitions in (1). This becomes the model that we validate.
-In \cref{app:ptr}, we extend the model further with arbitrary sharing of memory-safe
-pointers. Our criteria in (3) cover a total of five core requirements:
-\(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
-\(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
-\(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
-according to the needs of a system.
-
-In particular, many enforcement techniques focus purely on
-well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
-address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
-to enforce it completely. Others combine this protection with some degree of memory protection,
-chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
-describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
-to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by
-Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
-When checking if a preliminary version of these properties applied to
-the Cerise calling convention, they noted that the Skostengaard et al. calling convention
-did not enforce callee confidentiality, and made sure that their version would
-\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
-This demonstrates the benefit of our choice to explicitly state properties in security
-terms: specifying security mechanisms is hard, and when the specification takes the
-form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
-security requirement.
-
-%These are defined
-%formally in \cref{sec:props}. . We further define two
-%important combinations of properties: full stack safety \(\textnormal{\sc FSS}\),
-%and caller safety \(\textnormal{\sc ClrS}\). Of these, most existing concepts of
-%stack safety correspond to either \(\wbcf\) or \(\textnormal{\sc ClrS}\), which are
-%concerned with the basic guarantees that the caller recieves during the call, while
-%\(\textnormal{\sc FSS}\) is a natural extension that enforces a higher-level concept
-%of a correct interface between caller and callee. \sna{Blah. Needs work.}
-
-Flexibility aside, we show that our properties are useful for their purpose:
-distinguishing correct enforcement from incorrect. We use
-QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
-tool for Coq, to generate random programs and check that the
-stack safety {\em micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
+Therefore, any low-level treatment of stack safety must begin by asking:
+which high-level features does our system support, and what are their
+security requirements? Different answers will beget different properties.
+We identify and define five security requirements that a system might try
+to enforce upon a function call: well-bracketed control flow, and the
+integrity and confidentiality of each of the caller and the callee.
+These are by no means exhaustive. We apply these first to a simple system
+with very few features, then to a more realistic one supporting
+tailcall-elimination, argument passing on the stack, and callee-saves
+registers.
+
+We show that our properties are useful for distinguishing correct enforcement
+from incorrect using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
+testing tool for Coq. We generate random programs, using the features of our
+more realistic system, and check that the stack safety {\em micro-policies}
+of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
 correctly detect the ones that attempt to violate one of our properties.
@@
 talk about function activations and stack contents---abstractions that
-are typically not visible at machine level. To bridge this gap,
+are (typically) not visible at machine level. To bridge this gap,
 our properties are defined in terms of a {\em security semantics} layered on top of
-the standard execution semantics of the machine.  The security semantics identifies certain
-state transitions of the machine as {\em security-relevant
-  operations}, which update
+the standard execution semantics of the machine.  This semantics identifies certain
+state transitions of the machine as {\em security-relevant operations}, which update
 a notional {\em security context}.  This context consists of
@@
 that maps each element of the machine state (memory location or register)
-to a {\em security class} (active, sealed, etc.) specifying how the activation
+to a {\em security class} (active, sealed, etc.) which specifies how the activation
 can access the element.
-The action of a security-relevant operation on the context is defined by a set
-of rules that describe how the machine code makes use of the stack and registers
+The action of the security-relevant operations on the context is defined by a set
+of rules that depend on how the machine code makes use of the stack and registers
 to implement the function abstraction.
@@
 
+\sna{See commented out tex for content that could be massaged to fit in here,
+  depending on whether we're interested in delving deeper.}
+%For instance, the stack-safety specification from Georges et al.
+%\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+%control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
+%a callee should always return to its caller. But, in the presence of tail-call elimination, a
+%return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
+%Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
+%their enforcement mechanism did protect the confidentiality of the caller (though not the callee).
+
+%In particular, many enforcement techniques focus purely on
+%well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
+%address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
+%to enforce it completely. Others combine this protection with some degree of memory protection,
+%chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
+%describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
+%to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by
+%Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
+%When checking if a preliminary version of these properties applied to
+%the Cerise calling convention, they noted that the Skostengaard et al. calling convention
+%did not enforce callee confidentiality, and made sure that their version would
+%\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
+%This demonstrates the benefit of our choice to explicitly state properties in security
+%terms: specifying security mechanisms is hard, and when the specification takes the
+%form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
+%security requirement.
+
 HardBound \cite{DeviettiBMZ08} is another hardware extension related to

### 2023-02-04 submission5/core.tex
squeeze on IX; didn't help as much as I hoped

--- 2023-02-04 submission5/core.tex
@@
 
-We are not attempting to present a universal definition of stack safety
-for all systems. While many security properties can be described at
+\apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
+  Again, we are not providing evidence that these attributes are true, although we can
+  and should conjecture claim them in passing.}
+
+\bcp{Agreed.  I think in particular the first paragraph can just go
+  (I've deleted it).
+  But this is not enough.  Also, the section is trying to do at least
+  two things---talk about earlier work and how we compare, and talk
+  about why we believe a flexible approach is good.  The whole
+  discussion is a bit wooly and confusing.}
+
+\sna{Should probably move the discussion of the CHERI work to related work,
+  and focus on justifying our approach.}
+\bcp{+1}
+
+% Security properties should be motivated by security requirements
+% and flexible enough to support various sets of security requirements depending on the
+% goals of the system that they validate.
+% Following this reasoning, our vision of stack safety is as a collection of properties
+% describing the security needs of a system with regard to its functional abstractions.
+% While the stack still plays a key role, our stack safety is about functions. The most
+% dramatic consequence of this is that we treat registers as part of a function activation's
+% state, and hold them to the same protection requirements as stack memory.
+
+Flexibility is important for talking rigorously about stack safety.
+%
+The reason is that, while many security properties can be described at
 the level of a high-level programming language and translated to a
@@
 questions as whether a caller's frame should be readable or writable
-to its callee. Those are the purview of high-level languages built
+to its callee: those are the purview of high-level languages built
 atop the hardware stack.
 
-Therefore, any low-level treatment of stack safety must begin by asking:
-which high-level features does our system support, and what are their
-security requirements? Different answers will beget different properties.
-We identify and define five security requirements that a system might try
-to enforce upon a function call: well-bracketed control flow, and the
-integrity and confidentiality of each of the caller and the callee.
-These are by no means exhaustive. We apply these first to a simple system
-with very few features, then to a more realistic one supporting
-tailcall-elimination, argument passing on the stack, and callee-saves
-registers.
-
-We show that our properties are useful for distinguishing correct enforcement
-from incorrect using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
-testing tool for Coq. We generate random programs, using the features of our
-more realistic system, and check that the stack safety {\em micro-policies}
-of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
+\bcp{There's a broken transition here: the paragraph below is not an
+  instance of the paragraph above.  Indeed, I don't see how they
+  connect at all.}
+
+For instance, the stack-safety specification from Georges et al.
+\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
+a callee should always return to its caller. But, in the presence of tail-call elimination, a
+return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
+Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
+their enforcement mechanism did protect the confidentiality of the caller (though not the callee).
+
+These two cases represent two dimensions in which a stack safety definition requires
+flexibility: what constitutes safe behavior differs between systems that support different
+features, and different contexts have different top-level security requirements.
+
+We are not attempting to present a universal definition of stack safety for all systems,
+but we do construct our definition with these dimensions of flexibility in mind.
+Our model is decomposed into:
+(1) a lightweight model of the system and its security-relevant features, from which we
+derive the integrity, confidentiality, and control-flow requirements of each call and
+return, (2) the application of that model to a particular target system, and (3)
+the formal criteria for the integrity, confidentiality, and control flow
+requirements.
+
+We show that we can extend our system to new features by first introducing a simple
+example model, and then moving to a more realistic model with multiple new features:
+callee-save registers, argument passing on the stack, and tail-call elimination, simply
+by refining a few definitions in (1). This becomes the model that we validate.
+In \cref{app:ptr}, we extend the model further with arbitrary sharing of memory-safe
+pointers. Our criteria in (3) cover a total of five core requirements:
+\(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
+\(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
+\(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
+according to the needs of a system.
+
+In particular, many enforcement techniques focus purely on
+well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
+address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
+to enforce it completely. Others combine this protection with some degree of memory protection,
+chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
+describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
+to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by
+Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
+When checking if a preliminary version of these properties applied to
+the Cerise calling convention, they noted that the Skostengaard et al. calling convention
+did not enforce callee confidentiality, and made sure that their version would
+\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
+This demonstrates the benefit of our choice to explicitly state properties in security
+terms: specifying security mechanisms is hard, and when the specification takes the
+form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
+security requirement.
+
+%These are defined
+%formally in \cref{sec:props}. . We further define two
+%important combinations of properties: full stack safety \(\textnormal{\sc FSS}\),
+%and caller safety \(\textnormal{\sc ClrS}\). Of these, most existing concepts of
+%stack safety correspond to either \(\wbcf\) or \(\textnormal{\sc ClrS}\), which are
+%concerned with the basic guarantees that the caller recieves during the call, while
+%\(\textnormal{\sc FSS}\) is a natural extension that enforces a higher-level concept
+%of a correct interface between caller and callee. \sna{Blah. Needs work.}
+
+Flexibility aside, we show that our properties are useful for their purpose:
+distinguishing correct enforcement from incorrect. We use
+QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
+tool for Coq, to generate random programs and check that the
+stack safety {\em micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
 correctly detect the ones that attempt to violate one of our properties.
@@
 
-\sna{See commented out tex for content that could be massaged to fit in here,
-  depending on whether we're interested in delving deeper.}
-%For instance, the stack-safety specification from Georges et al.
-%\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
-%control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
-%a callee should always return to its caller. But, in the presence of tail-call elimination, a
-%return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
-%Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
-%their enforcement mechanism did protect the confidentiality of the caller (though not the callee).
-
-%In particular, many enforcement techniques focus purely on
-%well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
-%address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
-%to enforce it completely. Others combine this protection with some degree of memory protection,
-%chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
-%describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
-%to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by
-%Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
-%When checking if a preliminary version of these properties applied to
-%the Cerise calling convention, they noted that the Skostengaard et al. calling convention
-%did not enforce callee confidentiality, and made sure that their version would
-%\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
-%This demonstrates the benefit of our choice to explicitly state properties in security
-%terms: specifying security mechanisms is hard, and when the specification takes the
-%form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
-%security requirement.
-
 HardBound \cite{DeviettiBMZ08} is another hardware extension related to
@@
 
-\subsection{Testing The Stack Safety Menagerie}
-
 In \cref{sec:intro}, we discussed several dimensions of flexibility
@@
 features and different security requirements. Another important dimension
-is applicability to multiple enforcement mechanisms.
-In the following, we discuss how different existing mechanisms
-might fit into our model.
-
-\paragraph{Stack Canaries and Shadow Stacks}
-%
-Canaries are special values inserted by the caller at certain points of the
-stack and checked for corruption on return, for protection against
-some classes of control-flow hijacking (i.e., overwriting the caller's
-return address through a buffer overflow that would also overwrite the canary
-detectably). A shadow stack is a mirror image of the control-relevant parts of the stack,
-maintained by the program to attempt to detect attempts to hijack its control
-flow and ``restrict the flexibility available in creating gadget chains''
-\cite{Shanbhogue+19}.
-%
-Interestingly, these are lazy enforcement mechanisms, in that
-the attack may occur and be detected some time later, as long as
-it is detected before it can become dangerous. That would make our
-formalism a good fit for defining their security, except that
-they are chiefly hardening techniques: they increase the difficulty
-of some control-flow attacks on the stack, but cannot provide absolute
-guarantees on \(\wbcf\) under a normal attacker model.
-
-\paragraph{Split Stacks}
-%
-Code-pointer integrity seeks to prevent control-flow hijacking attacks by
-splitting the program memory, including the stack, into a regular region and a
-safe region; objects that require protection are placed in the latter and
-all accesses to them protected by static or dynamic checks.
-%
-Like shadow stacks, bounds checking is a high-level technique designed to
-protect code written in a high-level language, but it is designed to
-offer protection under a stronger threat model that also considers the presence
-of untrusted code. As with most other software-based methods, it purports to
-protect exclusively the control-relevant parts of the stack, making no claims
-about its data-relevant parts---we would solely be testing \(\wbcf\).
-
-\paragraph{Bounds Checking}
+is applicability to multiple enforcement mechanisms. \apt{Could we mention that
+  in intro as well?}  There is much future work in each of these dimensions.
+
+% \subsection{Testing The Stack Safety Menagerie}
+
+\paragraph*{Enforcement with capabilities}
+Capability machines can represent and manipulate
+unforgeable tokens of authority over specific regions of
+memory using a combination of dedicated out-of-band memory and
+specialized ISA instructions.
+%
+The most prominent contemporary representative of this long line of work is
+CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
+memory protection and compartmentalization.
+%
+Previous work has used these principles in simplified models to implement
+\emph{secure calling conventions} for the call
+stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
+protections of bounds-checking with software-based methods of control-flow protection.
+%
+%There are several proposals around the use of CHERI capabilities to enforce stack safety,
+%including mechanisms that use the standard CHERI hardware (which includes local
+%capabilities) \cite{SkorstengaardLocal},
+%and others that propose entirely new types of capabilities, such as linear
+%\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
+%\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
+It is a high priority to test this most recent work by Georges et al.~\cite{Georges22:TempsDesCerises}, as it is designed
+to enforce analogues of all of our properties except for \(\clei\).
+
+\paragraph*{Enforcement in software}
 %
@@
 but that may not be sufficient.
-
 Bounds checking approaches require substantial compiler cooperation. This is not a
 problem for our properties in general, but it is not very compatible with
-generation-by-execution of low-level code. A better choice would be to generate
+generation-by-execution of low-level code. A better choice might be to generate
 high-level code using a tool like CSmith \cite{DBLP:conf/pldi/YangCER11}, or prove the properties instead.
 
-\paragraph{Capabilities}
-
-Capability machines can represent and manipulate
-unforgeable tokens of authority over specific regions of
-memory using a combination of dedicated out-of-band memory and
-specialized ISA instructions.
-%
-The most prominent contemporary representative of this long line of work is
-CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
-memory protection and compartmentalization.
-%
-Previous work has used these principles in simplified models to implement
-\emph{secure calling conventions} for the call
-stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
-protections of bounds-checking with software-based methods of control-flow protection.
-
-There are several proposals around the use of CHERI capabilities to enforce stack safety,
-including mechanisms that use the standard CHERI hardware (which includes local
-capabilities) \cite{SkorstengaardLocal},
-and others that propose entirely new types of capabilities, such as linear
-\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
-\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
-It is a high priority to test this most recent work by Georges et al., as it is designed
-to be similar to all of our properties except for \(\clei\).
-
-\subsection{More Features}
-
-Several common features did not make it into this work: most importantly concurrency and
-exceptions. We plan to extend the model to include them, and then test them. We would also
-like to test the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
-
-\subsection{Proofs}
+\paragraph*{More Features}
+
+We have preliminary work on extending our model to handle C++-style
+exceptions, which, like tailcalls, obey only a weakened version of \(\wbcf\). 
+We are also exploring extensions to concurrency, starting with a model of
+statically allocated co-routines.  These extensions will also require non-trivial
+testing effort.  We also plan to test
+the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
+
+\paragraph*{Weaker Requirements}
+
+Several popular enforcement mechanisms are not designed to provide
+absolute guarantees of security. For example, stack canaries~\cite{Cowan+98}
+and shadow stacks~\cite{Dang+15,Shanbhogue+19}
+are chiefly hardening techniques: they increase the difficulty
+of some control-flow attacks on the stack, but cannot provide absolute
+guarantees on \(\wbcf\) under a normal attacker model.
+%
+Interestingly, these are lazy enforcement mechanisms, in that
+the attack may occur and be detected some time later, as long as
+it is detected before it can become dangerous. That would make our
+observation-based formalism a good fit for defining their security,
+if we could find a formal characterization of what they do acheive
+(perhaps in terms of a base machine with restricted addressing power).
+
+%% \paragraph{Split Stacks}
+%% %
+%% Code-pointer integrity seeks to prevent control-flow hijacking attacks by
+%% splitting the program memory, including the stack, into a regular region and a
+%% safe region; objects that require protection are placed in the latter and
+%% all accesses to them protected by static or dynamic checks.
+%% %
+%% Like shadow stacks, bounds checking is a high-level technique designed to
+%% protect code written in a high-level language, but it is designed to
+%% offer protection under a stronger threat model that also considers the presence
+%% of untrusted code. As with most other software-based methods, it purports to
+%% protect exclusively the control-relevant parts of the stack, making no claims
+%% about its data-relevant parts---we would solely be testing \(\wbcf\).
+
+\paragraph*{Proofs}
+
+The present work relies on randomized testing to validate properties versus
+enforcement mechanisms, but full formal proofs are certainly possible in principle,
+and we would like to investigate whether they are feasible in practice.
+Having formalizing our definitions in Coq gives us a head start in this effort.
 

### 2023-02-04 submission5/core.tex
Comments and polishing on 2

@@
 The security context consists of a stack of views: functions that map
-each state element to a {\it security class}. The security classes are
-\(\public\), \(\unsealed\), \(\object\), and \(\sealed\).
+each state element to a {\it security class}: 
+\(\public\), \(\unsealed\), \(\object\), or \(\sealed\).
 
@@
 globals and the heap, as well as the code region and globally-relevant
-registers---are always labeled \(\public\). We place security requirements on some
-\(\public\) elements for purposes of \(\wbcf\), and a given enforcement mechanism
-might restrict their access (e.g., by rendering code immutable) but for integrity
+registers\bcp{what is a globally-relevant
+register?}---are always labeled \(\public\). We place security requirements on some
+\(\public\) elements for purposes of our \(\wbcf\) property, and a
+given enforcement mechanism
+might restrict their access (e.g., by rendering code immutable), but for integrity
 and confidentiality purposes they are considered accessible at all times.
 
-For a newly active function, every element that is available for use but uninitialized
+For a newly active function, every stack location that is available
+for use but not yet initialized
 is seen as \(\unsealed\). From the perspective of the caller, the callee has no obligations
-regarding its use of \(\unsealed\) elements.
-
-Arguments are seen as \(\object\), meaning that their contents may be safely used.
-When the active function allocates memory, that memory will also be \(\object\).
+regarding its use of free elements.
+
+Arguments are marked \(\object\), meaning that their contents may be safely used.
+When the active function\bcp{are functions active, or are their
+  arguments active?} allocates memory, that memory will also be \(\object\).
 Then, on a call, \(\object\) elements that are not being used to communicate with
@@
 
-The property framework has many parameters to support broad applicability.
+\proposecut{Our property framework has many parameters to support broad applicability.}
 Conceptually, the following steps are needed to instantiate the framework to a specific machine
@@
 
-In \cref{sec:example} we give such an instantiation, and use it to explain the
-security properties.  But we stress that the property definitions themselves
-are \emph{independent} of the parameter instantiation.
-
-\apt{Suggest moving threat model to here.}
+\proposecut{In \cref{sec:example} we give such an instantiation, and
+  use it to illustrate our security properties.  But we stress that
+  the property definitions themselves are independent of the
+  instantiation.}
+
+\apt{Suggest moving threat model to here.}\bcp{Fine.}
 
@@
 
+\subsection{Testing The Stack Safety Menagerie}
+
 In \cref{sec:intro}, we discussed several dimensions of flexibility
@@
 features and different security requirements. Another important dimension
-is applicability to multiple enforcement mechanisms. \apt{Could we mention that
-  in intro as well?}  There is much future work in each of these dimensions.
-
-% \subsection{Testing The Stack Safety Menagerie}
-
-\paragraph*{Enforcement with capabilities}
-Capability machines can represent and manipulate
-unforgeable tokens of authority over specific regions of
-memory using a combination of dedicated out-of-band memory and
-specialized ISA instructions.
-%
-The most prominent contemporary representative of this long line of work is
-CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
-memory protection and compartmentalization.
-%
-Previous work has used these principles in simplified models to implement
-\emph{secure calling conventions} for the call
-stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
-protections of bounds-checking with software-based methods of control-flow protection.
-%
-%There are several proposals around the use of CHERI capabilities to enforce stack safety,
-%including mechanisms that use the standard CHERI hardware (which includes local
-%capabilities) \cite{SkorstengaardLocal},
-%and others that propose entirely new types of capabilities, such as linear
-%\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
-%\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
-It is a high priority to test this most recent work by Georges et al.~\cite{Georges22:TempsDesCerises}, as it is designed
-to enforce analogues of all of our properties except for \(\clei\).
-
-\paragraph*{Enforcement in software}
+is applicability to multiple enforcement mechanisms.
+In the following, we discuss how different existing mechanisms
+might fit into our model.
+
+\paragraph{Stack Canaries and Shadow Stacks}
+%
+Canaries are special values inserted by the caller at certain points of the
+stack and checked for corruption on return, for protection against
+some classes of control-flow hijacking (i.e., overwriting the caller's
+return address through a buffer overflow that would also overwrite the canary
+detectably). A shadow stack is a mirror image of the control-relevant parts of the stack,
+maintained by the program to attempt to detect attempts to hijack its control
+flow and ``restrict the flexibility available in creating gadget chains''
+\cite{Shanbhogue+19}.
+%
+Interestingly, these are lazy enforcement mechanisms, in that
+the attack may occur and be detected some time later, as long as
+it is detected before it can become dangerous. That would make our
+formalism a good fit for defining their security, except that
+they are chiefly hardening techniques: they increase the difficulty
+of some control-flow attacks on the stack, but cannot provide absolute
+guarantees on \(\wbcf\) under a normal attacker model.
+
+\paragraph{Split Stacks}
+%
+Code-pointer integrity seeks to prevent control-flow hijacking attacks by
+splitting the program memory, including the stack, into a regular region and a
+safe region; objects that require protection are placed in the latter and
+all accesses to them protected by static or dynamic checks.
+%
+Like shadow stacks, bounds checking is a high-level technique designed to
+protect code written in a high-level language, but it is designed to
+offer protection under a stronger threat model that also considers the presence
+of untrusted code. As with most other software-based methods, it purports to
+protect exclusively the control-relevant parts of the stack, making no claims
+about its data-relevant parts---we would solely be testing \(\wbcf\).
+
+\paragraph{Bounds Checking}
 %
@@
 but that may not be sufficient.
+
 Bounds checking approaches require substantial compiler cooperation. This is not a
 problem for our properties in general, but it is not very compatible with
-generation-by-execution of low-level code. A better choice might be to generate
+generation-by-execution of low-level code. A better choice would be to generate
 high-level code using a tool like CSmith \cite{DBLP:conf/pldi/YangCER11}, or prove the properties instead.
 
-\paragraph*{More Features}
-
-We have preliminary work on extending our model to handle C++-style
-exceptions, which, like tailcalls, obey only a weakened version of \(\wbcf\). 
-We are also exploring extensions to concurrency, starting with a model of
-statically allocated co-routines.  These extensions will also require non-trivial
-testing effort.  We also plan to test
-the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
-
-\paragraph*{Weaker Requirements}
-
-Several popular enforcement mechanisms are not designed to provide
-absolute guarantees of security. For example, stack canaries~\cite{Cowan+98}
-and shadow stacks~\cite{Dang+15,Shanbhogue+19}
-are chiefly hardening techniques: they increase the difficulty
-of some control-flow attacks on the stack, but cannot provide absolute
-guarantees on \(\wbcf\) under a normal attacker model.
-%
-Interestingly, these are lazy enforcement mechanisms, in that
-the attack may occur and be detected some time later, as long as
-it is detected before it can become dangerous. That would make our
-observation-based formalism a good fit for defining their security,
-if we could find a formal characterization of what they do acheive
-(perhaps in terms of a base machine with restricted addressing power).
-
-%% \paragraph{Split Stacks}
-%% %
-%% Code-pointer integrity seeks to prevent control-flow hijacking attacks by
-%% splitting the program memory, including the stack, into a regular region and a
-%% safe region; objects that require protection are placed in the latter and
-%% all accesses to them protected by static or dynamic checks.
-%% %
-%% Like shadow stacks, bounds checking is a high-level technique designed to
-%% protect code written in a high-level language, but it is designed to
-%% offer protection under a stronger threat model that also considers the presence
-%% of untrusted code. As with most other software-based methods, it purports to
-%% protect exclusively the control-relevant parts of the stack, making no claims
-%% about its data-relevant parts---we would solely be testing \(\wbcf\).
-
-\paragraph*{Proofs}
-
-The present work relies on randomized testing to validate properties versus
-enforcement mechanisms, but full formal proofs are certainly possible in principle,
-and we would like to investigate whether they are feasible in practice.
-Having formalizing our definitions in Coq gives us a head start in this effort.
+\paragraph{Capabilities}
+
+Capability machines can represent and manipulate
+unforgeable tokens of authority over specific regions of
+memory using a combination of dedicated out-of-band memory and
+specialized ISA instructions.
+%
+The most prominent contemporary representative of this long line of work is
+CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
+memory protection and compartmentalization.
+%
+Previous work has used these principles in simplified models to implement
+\emph{secure calling conventions} for the call
+stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
+protections of bounds-checking with software-based methods of control-flow protection.
+
+There are several proposals around the use of CHERI capabilities to enforce stack safety,
+including mechanisms that use the standard CHERI hardware (which includes local
+capabilities) \cite{SkorstengaardLocal},
+and others that propose entirely new types of capabilities, such as linear
+\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
+\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
+It is a high priority to test this most recent work by Georges et al., as it is designed
+to be similar to all of our properties except for \(\clei\).
+
+\subsection{More Features}
+
+Several common features did not make it into this work: most importantly concurrency and
+exceptions. We plan to extend the model to include them, and then test them. We would also
+like to test the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
+
+\subsection{Proofs}
 

### 2023-02-04 submission5/core.tex
Fix

@@
 
-\proposecut{In \cref{sec:example} we give such an instantiation, and
+\bcp{cut this?}In \cref{sec:example} we give such an instantiation, and
   use it to illustrate our security properties.  But we stress that
   the property definitions themselves are independent of the
-  instantiation.}
+  instantiation.
 

### 2023-02-04 submission5/core.tex
Notes and edits in section 2

--- 2023-02-04 submission5/core.tex
@@
 
-Our security semantics has similarities to the overlay semantics
-proposed by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it
-does not restrict the behavior of the underlying machine in any
-way\bcp{Will people not familiar with Skorstengaard understand
-  that?}. Rather, it tracks additional context about the history of
-security-relevant operations, which informs the criterion a machine
-must satisfy in order to correctly obey the properties.\apt{Sharpen
-  this comparison?}\bcp{+1.  Or maybe, in the interest of time and to
-  keeping this section flowing, move it to related work.}
-
 %The security principals in our account of stack safety are function
@@
 and stack pointer, and a standard ABI.
+\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
+  This assumption is made in our example and testing.}
 We make no particular assumptions about the provenance of the machine code; in particular,
@@
 any additional enforcement information (such as a tag).\bcp{Why?}
+\sna{I think that the payload concept is too nitty-gritty for this section.
+  The chief use of the payload is that it allows every value to be interpreted
+  as an address. In the tag world, it happens to coincide with a natural instance
+  of our variant-compatibility relation.}
 
@@
   hyperproperties.  CSF readers may be surprised.}
+\sna{Surprised that we have hyperproperties,
+  or that we made it this far without mentioning them?}
 
@@
 might require some external mechanism for distinguishing which instructions are
-security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique, which we use in our examples, is to add labels to
-selected instructions, but the theory does not depend on this
-choice.\bcp{What could one do instead?}
+security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique,
+which we use in our examples, is to add labels to selected instructions, but the theory does not
+depend on this choice. One could instead recognize some operations dynamically. For instance,
+if we make a call to untrusted (and unlabeled) code, we might decide to recognize
+any jump back to trusted code as a return, to which \(\wbcf\) needs to apply.
 
@@
 The security context consists of a stack of views: functions that map
-each state element to a {\it security class}: 
-\(\public\), \(\unsealed\), \(\object\), or \(\sealed\).
+each state element to a {\it security class}. The security classes are
+\(\public\), \(\unsealed\), \(\object\), and \(\sealed\).
 
@@
 globals and the heap, as well as the code region and globally-relevant
-registers\bcp{what is a globally-relevant
-register?}---are always labeled \(\public\). We place security requirements on some
-\(\public\) elements for purposes of our \(\wbcf\) property, and a
-given enforcement mechanism
-might restrict their access (e.g., by rendering code immutable), but for integrity
+registers---are always labeled \(\public\). We place security requirements on some
+\(\public\) elements for purposes of \(\wbcf\), and a given enforcement mechanism
+might restrict their access (e.g., by rendering code immutable) but for integrity
 and confidentiality purposes they are considered accessible at all times.
 
-For a newly active function, every stack location that is available
-for use but not yet initialized
+For a newly active function, every element that is available for use but uninitialized
 is seen as \(\unsealed\). From the perspective of the caller, the callee has no obligations
-regarding its use of free elements.
-
-Arguments are marked \(\object\), meaning that their contents may be safely used.
-When the active function\bcp{are functions active, or are their
-  arguments active?} allocates memory, that memory will also be \(\object\).
+regarding its use of \(\unsealed\) elements.
+
+Arguments are seen as \(\object\), meaning that their contents may be safely used.
+When the active function allocates memory, that memory will also be \(\object\).
 Then, on a call, \(\object\) elements that are not being used to communicate with
@@
 
-\proposecut{Our property framework has many parameters to support broad applicability.}
+The property framework has many parameters to support broad applicability.
 Conceptually, the following steps are needed to instantiate the framework to a specific machine
@@
 
-\proposecut{In \cref{sec:example} we give such an instantiation, and
-  use it to illustrate our security properties.  But we stress that
-  the property definitions themselves are independent of the
-  instantiation.}
-
-\apt{Suggest moving threat model to here.}\bcp{Fine.}
+In \cref{sec:example} we give such an instantiation, and use it to explain the
+security properties.  But we stress that the property definitions themselves
+are \emph{independent} of the parameter instantiation.
+
+\apt{Suggest moving threat model to here.}
 
@@
 
+\sna{See commented out tex for content that could be massaged to fit in here,
+  depending on whether we're interested in delving deeper.}
+%For instance, the stack-safety specification from Georges et al.
+%\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
+%control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
+%a callee should always return to its caller. But, in the presence of tail-call elimination, a
+%return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
+%Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
+%their enforcement mechanism did protect the confidentiality of the caller (though not the callee).
+
+%In particular, many enforcement techniques focus purely on
+%well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
+%address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
+%to enforce it completely. Others combine this protection with some degree of memory protection,
+%chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
+%describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
+%to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by
+%Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
+%When checking if a preliminary version of these properties applied to
+%the Cerise calling convention, they noted that the Skostengaard et al. calling convention
+%did not enforce callee confidentiality, and made sure that their version would
+%\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
+%This demonstrates the benefit of our choice to explicitly state properties in security
+%terms: specifying security mechanisms is hard, and when the specification takes the
+%form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
+%security requirement.
+
+%Our security semantics has similarities to the overlay semantics
+%proposed by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it
+%does not restrict the behavior of the underlying machine in any
+%way\bcp{Will people not familiar with Skorstengaard understand
+%  that?}. Rather, it tracks additional context about the history of
+%security-relevant operations, which informs the criterion a machine
+%must satisfy in order to correctly obey the properties.\apt{Sharpen
+%  this comparison?}\bcp{+1.  Or maybe, in the interest of time and to
+%  keeping this section flowing, move it to related work.}
+
 HardBound \cite{DeviettiBMZ08} is another hardware extension related to

### 2023-02-04 submission5/core.tex

@@
 The security context consists of a stack of views: functions that map
-each state element to a {\it security class}. The security classes are
-\(\public\), \(\unsealed\), \(\object\), and \(\sealed\).
+each state element to a {\it security class}:
+\(\public\), \(\unsealed\), \(\object\), or \(\sealed\).
 
@@
 
-In \cref{sec:example} we give such an instantiation, and use it to explain the
-security properties.  But we stress that the property definitions themselves
-are \emph{independent} of the parameter instantiation.
-
-\apt{Suggest moving threat model to here.}
+\bcp{cut this?}In \cref{sec:example} we give such an instantiation, and
+  use it to illustrate our security properties.  But we stress that
+  the property definitions themselves are independent of the
+  instantiation.
+
+\apt{Suggest moving threat model to here.}\bcp{Fine.}
 
@@
 but that may not be sufficient.
-
 Bounds checking approaches require substantial compiler cooperation. This is not a
 problem for our properties in general, but it is not very compatible with
-generation-by-execution of low-level code. A better choice would be to generate
+generation-by-execution of low-level code. A better choice might be to generate
 high-level code using a tool like CSmith \cite{DBLP:conf/pldi/YangCER11}, or prove the properties instead.
 
-\paragraph{Capabilities}
-
-Capability machines can represent and manipulate
-unforgeable tokens of authority over specific regions of
-memory using a combination of dedicated out-of-band memory and
-specialized ISA instructions.
-%
-The most prominent contemporary representative of this long line of work is
-CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
-memory protection and compartmentalization.
-%
-Previous work has used these principles in simplified models to implement
-\emph{secure calling conventions} for the call
-stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
-protections of bounds-checking with software-based methods of control-flow protection.
-
-There are several proposals around the use of CHERI capabilities to enforce stack safety,
-including mechanisms that use the standard CHERI hardware (which includes local
-capabilities) \cite{SkorstengaardLocal},
-and others that propose entirely new types of capabilities, such as linear
-\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
-\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
-It is a high priority to test this most recent work by Georges et al., as it is designed
-to be similar to all of our properties except for \(\clei\).
-
-\subsection{More Features}
-
-Several common features did not make it into this work: most importantly concurrency and
-exceptions. We plan to extend the model to include them, and then test them. We would also
-like to test the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
-
-\subsection{Proofs}
+\paragraph*{More Features}
+
+We have preliminary work on extending our model to handle C++-style
+exceptions, which, like tailcalls, obey only a weakened version of \(\wbcf\).
+We are also exploring extensions to concurrency, starting with a model of
+statically allocated co-routines.  These extensions will also require non-trivial
+testing effort.  We also plan to test
+the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
+
+\paragraph*{Weaker Requirements}
+
+Several popular enforcement mechanisms are not designed to provide
+absolute guarantees of security. For example, stack canaries~\cite{Cowan+98}
+and shadow stacks~\cite{Dang+15,Shanbhogue+19}
+are chiefly hardening techniques: they increase the difficulty
+of some control-flow attacks on the stack, but cannot provide absolute
+guarantees on \(\wbcf\) under a normal attacker model.
+%
+Interestingly, these are lazy enforcement mechanisms, in that
+the attack may occur and be detected some time later, as long as
+it is detected before it can become dangerous. That would make our
+observation-based formalism a good fit for defining their security,
+if we could find a formal characterization of what they do acheive
+(perhaps in terms of a base machine with restricted addressing power).
+
+%% \paragraph{Split Stacks}
+%% %
+%% Code-pointer integrity seeks to prevent control-flow hijacking attacks by
+%% splitting the program memory, including the stack, into a regular region and a
+%% safe region; objects that require protection are placed in the latter and
+%% all accesses to them protected by static or dynamic checks.
+%% %
+%% Like shadow stacks, bounds checking is a high-level technique designed to
+%% protect code written in a high-level language, but it is designed to
+%% offer protection under a stronger threat model that also considers the presence
+%% of untrusted code. As with most other software-based methods, it purports to
+%% protect exclusively the control-relevant parts of the stack, making no claims
+%% about its data-relevant parts---we would solely be testing \(\wbcf\).
+
+\paragraph*{Proofs}
+
+The present work relies on randomized testing to validate properties versus
+enforcement mechanisms, but full formal proofs are certainly possible in principle,
+and we would like to investigate whether they are feasible in practice.
+Having formalizing our definitions in Coq gives us a head start in this effort.
 

### 2023-02-04 submission5/core.tex
A few quick space hacks

--- 2023-02-04 submission5/core.tex
@@
 {\em Functions} in high-level languages, and related abstractions such
-as subroutines, procedures, methods, etc., are units of computation
+as procedures, methods, etc., are units of computation
 that call one another to modularly realize larger computations.
@@
 Under this policy, one function activation {\em can} corrupt another's
-memory---just not in ways that do not affect observable behavior.
+memory---just not in ways that affect observable behavior.
 Therefore, \(LTC\) would not fulfill Georges et al.'s property
-(appropriately adapted to the tagged setting).  But \(LTC\) does
+(adapted to the tagged setting).  But \(LTC\) does
 arguably enforce stack safety.  A looser, more observational
@@
 the intuition of protecting function activations from each other.
-\bcp{Aren't other formal definitions based on the same intuition?}
+\bcp{Aren't other formal definitions based on the same intuition?}%
 \sna{Two differences we're trying to get across. First, protecting
@@
   describing operationally what they can and can't do (as the CHERI
-  folks do.)}
-\bcp{Both useful!  But neither comes across, at least not very clearly.}
+  folks do.)}%
+\bcp{Both useful!  But neither comes across, at least not very clearly.}%
 We use the tools of language-based
@@
 Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
-While our properties are motivated by the desire to specify \(LTC\) precisely,
-they are stated abstractly in hopes that they may also apply
-to other enforcement mechanisms.
+\proposecut{While our properties are motivated by the desire to specify \(LTC\) precisely,
+they are stated abstractly in hopes of applying
+to other enforcement mechanisms.}
 

### 2023-02-04 submission5/core.tex
Last responses

@@
 integrity and confidentiality of each of the caller and the callee.
-These are by no means exhaustive. We apply these first to a simple system
+These are by no means exhaustive\bcp{That comment belongs in the
+  future work
+  section, not here}. We apply these first to a simple system
 with very few features, then to a more realistic one supporting
@@
   or that we made it this far without mentioning them?}
+\bcp{The latter!}
 
@@
 \sna{See commented out tex for content that could be massaged to fit in here,
-  depending on whether we're interested in delving deeper.}
+  depending on whether we're interested in delving deeper.}\bcp{If
+  there's space, some of this could be resuscitated in the related
+  work section.  But removing it from here is a good call.}
 %For instance, the stack-safety specification from Georges et al.

### 2023-02-04 submission5/core.tex
Must stop space hacking...  Really stopping now...

@@
 schemes to handle the nested nature of confidentiality: rather than
-just generating two initial machines that are variants of one another
+just generating two variant initial machines
 and letting them execute to test for noninterference, we generate a
@@
 that our properties should exhibit: applicability to different system
-features and different security requirements. Another important dimension
+features and different security requirements. Another \proposecut{important dimension}
 is applicability to multiple enforcement mechanisms. \apt{Could we mention that
-  in intro as well?}  There is much future work in each of these dimensions.
+  in intro as well?}  \proposecut{There is much future work in each of these dimensions.}
 
@@
 %\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
-It is a high priority to test this most recent work by Georges et al.~\cite{Georges22:TempsDesCerises}, as it is designed
+We want to test the most recent work by Georges et
+al.~\cite{Georges22:TempsDesCerises}, which is designed
 to enforce analogues of all of our properties except for \(\clei\).
@@
 testing effort.  We also plan to test
-the model given in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
+the model in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
 
@@
 and we would like to investigate whether they are feasible in practice.
-Having formalizing our definitions in Coq gives us a head start in this effort.
+\proposecut{Having formalized our definitions in Coq gives us a head
+  start in this effort.} \bcp{Or just just the whole paragraph.}
 

### 2023-02-04 submission5/core.tex
Mostly just comments

--- 2023-02-04 submission5/core.tex
@@
 {\em Functions} in high-level languages, and related abstractions such
-as procedures, methods, etc., are units of computation
+as subroutines, procedures, methods, etc., are units of computation
 that call one another to modularly realize larger computations.
@@
 Under this policy, one function activation {\em can} corrupt another's
-memory---just not in ways that affect observable behavior.
+memory---just not in ways that do not affect observable behavior.
 Therefore, \(LTC\) would not fulfill Georges et al.'s property
-(adapted to the tagged setting).  But \(LTC\) does
+(appropriately adapted to the tagged setting).  But \(LTC\) does
 arguably enforce stack safety.  A looser, more observational
@@
 the intuition of protecting function activations from each other.
-\bcp{Aren't other formal definitions based on the same intuition?}%
+\bcp{Aren't other formal definitions based on the same intuition?}
 \sna{Two differences we're trying to get across. First, protecting
@@
   describing operationally what they can and can't do (as the CHERI
-  folks do.)}%
-\bcp{Both useful!  But neither comes across, at least not very clearly.}%
+  folks do.)}
+\bcp{Both useful!  But neither comes across, at least not very clearly.}
 We use the tools of language-based
@@
 Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
-\proposecut{While our properties are motivated by the desire to specify \(LTC\) precisely,
-they are stated abstractly in hopes of applying
-to other enforcement mechanisms.}
+While our properties are motivated by the desire to specify \(LTC\) precisely,
+they are stated abstractly in hopes that they may also apply
+to other enforcement mechanisms.
 
@@
   do them!}
+
+\sna{Agreed. The point of the operations was originally that we shouldn't
+  have to think about all of these things when presenting the formalism,
+  not that we should open with all the different ways we can think about them.}
 

### 2023-02-04 submission5/core.tex
colors in Fig. 3
spacing in fig. 6
work on 1,2

--- 2023-02-04 submission5/core.tex
@@
 
-\subsection{The Call Stack and Its Security}
+%\subsection{The Call Stack and Its Security}
 
@@
 The \emph{call stack} is the fundamental data structure used to
-implement functions, aided by an ABI\bcp{will all CSF readers know
-  what an ABI is?} that defines how registers are shared
-between activations.
+implement functions, aided by an Application Binary Interface (ABI)
+that defines how registers are shared between activations.
 
@@
 memory cannot access it afterward.
-
+%
 Under this policy, one function activation {\em can} corrupt another's
@@
 We propose here a formal characterization of stack safety, based on
-the intuition of protecting function activations from each other.
-\bcp{Aren't other formal definitions based on the same intuition?}%
-\sna{Two differences we're trying to get across. First, protecting
-  functions vs. protecting stack frames: we bring registers into
-  the story because it isn't just about the stack as a data structure.
-  Second: describing security in terms of functions' high-level
-  security goals (f is trying to extract a secret) rather than
-  describing operationally what they can and can't do (as the CHERI
-  folks do.)}%
-\bcp{Both useful!  But neither comes across, at least not very clearly.}%
+the intuition of protecting function activations (both registers and
+stack frames) from each other.
+%\bcp{Aren't other formal definitions based on the same intuition?}%
+%% \sna{Two differences we're trying to get across. First, protecting
+%%   functions vs. protecting stack frames: we bring registers into
+%%   the story because it isn't just about the stack as a data structure.
+%%   Second: describing security in terms of functions' high-level
+%%   security goals (f is trying to extract a secret) rather than
+%%   describing operationally what they can and can't do (as the CHERI
+%%   folks do.)}%
+%% \bcp{Both useful!  But neither comes across, at least not very clearly.}%
 We use the tools of language-based
-security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
+security~\cite{sabelfeld2003language}, treating function activations as
+security principals.  We decompose stack safety into a family of
 properties describing
@@
 
-We find, via property-based random testing, that \(LTC\) is flawed in
-a way that undermines both integrity and confidentiality; after
-correcting this flaw, the repaired \(LTC\) satisfies all our
-properties.  Moreover, we show that our properties can be extended to
-describe calling conventions with features such as callee-save
-registers, argument-passing on the stack, and tail-call
-elimination. We modify \(LTC\) to protect these features and apply
-random testing to validate this extended protection mechanism against
-the extended properties.
-
-\subsection {Towards Formal Properties for Stack Safety}
+%\subsection {Towards Formal Properties for Stack Safety}
 
@@
 integrity and confidentiality of each of the caller and the callee.
-These are by no means exhaustive\bcp{That comment belongs in the
-  future work
-  section, not here}. We apply these first to a simple system
+%These are by no means exhaustive\bcp{That comment belongs in the
+%  future work
+%  section, not here}.
+We apply these first to a simple system
 with very few features, then to a more realistic one supporting
-tailcall-elimination, argument passing on the stack, and callee-saves
+tail-call elimination, argument passing on the stack, and callee-save
 registers.
@@
 from incorrect using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
-testing tool for Coq. We generate random programs, using the features of our
-more realistic system, and check that the stack safety {\em micro-policies}
-of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
-correctly detect the ones that attempt to violate one of our properties.
+random testing tool for Coq.
+We find that \(LTC\) is flawed in
+a way that undermines both integrity and confidentiality; after
+correcting this flaw, the repaired \(LTC\) satisfies all our
+properties.  Moreover, we modify \(LTC\) to protect the features of
+our more realistic system, and apply
+random testing to validate this extended protection mechanism against
+the extended properties.
+
+%We generate random programs, using the features of our
+%more realistic system, and check that the stack safety {\em micro-policies}
+%of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
+%correctly detect the ones that attempt to violate one of our properties.
 
@@
 
-\subsection{Contributions}
+%\subsection{Contributions}
 
@@
 mechanisms.
-\item We instantiate\bcp{extend??} these definitions in a realistic
-setting with features such as\bcp{``such as''? are there others?}
-argument passing on the stack, callee-saves registers, and tail-call
-elimination.\bcp{How about: We extend these core definitions to
+\item
+  %We instantiate\bcp{extend??} these definitions in a realistic
+%setting with features such as\bcp{``such as''? are there others?}
+%argument passing on the stack, callee-saves registers, and tail-call
+  %elimination.\bcp{How about:
+ We extend these core definitions to
   describe a realistic setting with argument passing on the stack,
-  callee-saves registers, and tail-call elimination.}  Our model is
+  callee-saves registers, and tail-call elimination.  Our model is
 modular enough that handling these features is straightforward.
@@
 
-In the next section, we give a brief overview of the structure of
-properties.  In \cref{sec:example}, we walk through a function call in
+In the next section, we give a brief overview of our framework and
+assumptions. In \cref{sec:example}, we walk through a function call in
 a simple example machine and discuss how each of our properties
@@
 as future work.
-
-\subsection{Threat Model and Limitations}
-\apt{This could probably go better at end of next section.}
-
-When our properties are used to evaluate a system, the threat model
-will depend on the details of that system. However, there are some
-constraints that our design puts on any system. In particular, we must
-trust that the method of distinguishing security-relevant operations
-is accurate.  For example, if the enforcement mnechanism involves
-labels placed on code by a compiler, we must trust that the compiler
-placed those labels correctly. For instance, if a compiled function
-call is not marked as such, then the caller's data might not be
-protected from the callee.  Conversely, marking too many operations as
-calls will simply cause otherwise safe programs to be rejected.
-
-\bcp{Still working here:}
-
-We do not assume that low-level code adheres to any fixed calling
-convention or implements any particular source-language constructs.
-Indeed, if the source
-language is C, then high-level programs might contain undefined
-behavior, in which case they might be compiled to arbitrary machine
-code. A given enforcement
-mechanism or target architecture might place additional constraints, particularly on the behavior of
-call and return sequences. For instance, extant implementations tend to assume
-implicitly that callee-saved registers have their values maintained by whichever compiler
-generated their code. Our properties explicitly state this as a requirement,
-which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
-Roessler and DeHon include the compiler in their trusted computing
-base, so our properties
-are stronger than necessary to characterize their micro-policies.
-
-In general, it is impossible to distinguish buggy machine code from an
-attacker.  In our examples we will identify one function or another as
-an attacker, but we do not require any static division between trusted
-and untrusted code, and we aim to protect even buggy code.
-
-This is a strong threat model, but it does omit some important aspects
-of stack safety in real systems: in particular, it does not address
-concurrency.  Hardware and timing attacks are also out of scope.
-%
 
@@
 
+%% \paragraph*{Variants and Non-interference}
+
+To characterize situations in which values should be kept secret or prevented
+from influencing future execution, we borrow the idea of \emph{variant} states
+from the theory of non-interference.  For example, to show that certain locations
+are kept secret, it suffices to vary their contents them and check that the observable
+output does not change.
+
+%Non-interference is a way of expressing
+%the flow of information through the system. If we begin with a state \(\mach\)
+%and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
+%then execute both \(\mach\) and \(\nach\), any differences in the outputs they
+%produce leaks information about \(\component\).
+
 Finally, we define high-level security properties---integrity, confidentiality,
@@
 elements. These take the form of predicates on states and contexts
-that use variants\bcp{undefined},
+that use variants,
 assertions about future return states, and comparison of event traces to capture the
@@
 We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
-and stack pointer, and a standard ABI.
-\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
-  This assumption is made in our example and testing.}
+and stack pointer. 
+%, and a standard ABI.
+%\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
+%  This assumption is made in our example and testing.}
 We make no particular assumptions about the provenance of the machine code; in particular,
@@
 we assume that the behavior of these mechanisms is incorporated into the basic
-step semantics of the machine, but we distinguish the \emph{payload} part of
-each state element---the part relevant to the behavior of the basic machine---from
-any additional enforcement information (such as a tag).\bcp{Why?}
-\sna{I think that the payload concept is too nitty-gritty for this section.
-  The chief use of the payload is that it allows every value to be interpreted
-  as an address. In the tag world, it happens to coincide with a natural instance
-  of our variant-compatibility relation.}
+step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
+%% \sna{I think that the payload concept is too nitty-gritty for this section.
+%%   The chief use of the payload is that it allows every value to be interpreted
+%%   as an address. In the tag world, it happens to coincide with a natural instance
+%%   of our variant-compatibility relation.}
 
@@
 Security-relevant operations typically correspond to
-\proposeadd{single?} underlying machine instructions,
+single underlying machine instructions,
 but the correspondence may not be obvious just by inspecting the machine code.
@@
 and {\tt jalr} (``jump-and-link-register'') instructions, but these
-instructions might also be used for other things. An instantiation of
-an operation
-might require some external mechanism for distinguishing which instructions are
-security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique,
-which we use in our examples, is to add labels to selected instructions, but the theory does not
-depend on this choice. One could instead recognize some operations dynamically. For instance,
-if we make a call to untrusted (and unlabeled) code, we might decide to recognize
-any jump back to trusted code as a return, to which \(\wbcf\) needs to apply.
-
-Whatever the mechanism for labeling transitions, we end up with an annotated version of the
-machine transition function: we write \(\mach \xrightarrow{\bar{\psi},
+instructions might also be used for other things. Thus we identify
+security-relevant operations by labeling selected instructions. 
+%% An instantiation of
+%% an operation
+%% might require some external mechanism for distinguishing which instructions are
+%% security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique,
+%% which we use in our examples, is to add labels to selected instructions, but the theory does not
+%% depend on this choice. One could instead recognize some operations dynamically. For instance,
+%% if we make a call to untrusted (and unlabeled) code, we might decide to recognize
+%% any jump back to trusted code as a return, to which \(\wbcf\) needs to apply.
+
+We end up with an annotated version of the
+machine transition function written \(\mach \xrightarrow{\bar{\psi},
   \obs} \mach'\), where \(\mach\) and \(\mach\) are machine states, \(\obs\) is
@@
 The most important rules describe call and return operations.
-A call pushes a new activation view\bcp{it was just called ``view'' above} onto the context stack and changes the class of the
+A call pushes a new view onto the context stack and changes the class of the
 caller's data to protect it from the new callee; a return reverses these steps.
@@
 state elements accordingly.
+
+\apt{Still working from here down}
 
@@
 
-%% \paragraph*{Variants and Non-interference}
-
-%% To characterize situations in which values should be kept secret or prevented
-%% from influencing future execution, we borrow the idea of \emph{variant} states
-%% from the theory of non-interference. Non-interference is a way of expressing
-%% the flow of information through the system. If we begin with a state \(\mach\)
-%% and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
-%% then execute both \(\mach\) and \(\nach\), any differences in the outputs they
-%% produce leaks information about \(\component\).
-
 \paragraph*{Instantiating the Framework}
-
-\proposecut{Our property framework has many parameters to support broad applicability.}
 Conceptually, the following steps are needed to instantiate the framework to a specific machine
-and code style: (i) define the base machine semantics; (ii) distinguish payload
-from enforcement parts (if any) of the machine state; (iii) identify the set of
-security-relevant operations and rules required by the code style; (iv) determine
-how to label (or otherwise associate) machine instruction with security-relevant
-operations as appropriate; (v) specify the form of observable events.
+and code style: (i) define the base machine semantics, including any hardware
+security enforcement features; (ii) identify the set of
+security-relevant operations and rules required by the code style; (iii) determine
+how to label machine instruction with security-relevant
+operations as appropriate; (iv) specify the form of observable events.
 
@@
 
-\apt{Suggest moving threat model to here.}\bcp{Fine.}
+\paragraph*{Threat Model and Limitations}
+
+When our properties are used to evaluate a system, the threat model
+will depend on the details of that system. However, there are some
+constraints that our design puts on any system. In particular, we must
+trust that the method of distinguishing security-relevant operations
+is accurate.  For example, if the enforcement mnechanism involves
+labels placed on code by a compiler, we must trust that the compiler
+placed those labels correctly. For instance, if a compiled function
+call is not marked as such, then the caller's data might not be
+protected from the callee.  Conversely, marking too many operations as
+calls will simply cause otherwise safe programs to be rejected.
+
+\bcp{Still working here:}
+
+We do not assume that low-level code adheres to any fixed calling
+convention or implements any particular source-language constructs.
+Indeed, if the source
+language is C, then high-level programs might contain undefined
+behavior, in which case they might be compiled to arbitrary machine
+code. A given enforcement
+mechanism or target architecture might place additional constraints, particularly on the behavior of
+call and return sequences. For instance, extant implementations tend to assume
+implicitly that callee-saved registers have their values maintained by whichever compiler
+generated their code. Our properties explicitly state this as a requirement,
+which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
+Roessler and DeHon include the compiler in their trusted computing
+base, so our properties
+are stronger than necessary to characterize their micro-policies.
+
+In general, it is impossible to distinguish buggy machine code from an
+attacker.  In our examples we will identify one function or another as
+an attacker, but we do not require any static division between trusted
+and untrusted code, and we aim to protect even buggy code.
+
+This is a strong threat model, but it does omit some important aspects
+of stack safety in real systems: in particular, it does not address
+concurrency.  Hardware and timing attacks are also out of scope.
+%
 
@@
 in \(\mach\) if they can be replaced by arbitrary other values without changing the
-observable behavior of the machine. All other elements are \emph{relevant}.
-
-(This story is slightly over-simplified. If an enforcement mechanism maintains
+observable behavior of the machine. All other elements are \emph{relevant}.\footnote{
+This story is slightly over-simplified. If an enforcement mechanism maintains
 additional state associated with elements, such as tags, we don't want that
-state to vary; only the payloads should. Formal definitions of variants and
-relevance that incorporate this wrinkle are given in \cref{sec:props}.)
+state to vary. Formal definitions of variants and
+relevance that incorporate this wrinkle are given in \cref{sec:props}.}
 
@@
     \]
-    \[\large\Downarrow \hspace{1cm} \mathtt{a0} \leftbox[0] \hspace{1cm}
-    \rightbox[7] \mathtt{a0} \hspace{1cm} \large\Downarrow\]
+    \[\large\Downarrow \hspace{1cm} \mathtt{a0} ~ \leftbox[0] \hspace{1cm}
+    \rightbox[7] ~ \mathtt{a0} \hspace{1cm} \large\Downarrow\]
     \[
@@
     \]
-    \[\mathtt{a0} \leftbox[0] \hspace{1cm}
-    \mathtt{a0} \rightbox[0]\]
+    \[\mathtt{a0} ~ \leftbox[0] \hspace{1cm}
+    \rightbox[0] ~ \mathtt{a0}\]
   \caption{Return-time Confidentiality Violation}
@@
 \(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
-The details of value structure depend on the specific machine being modelled;
-intuitively, the payload represents the part of the value that is relevant to
-the behavior of the basic machine, while the rest of the value may contain
-information relevant to a hardware enforcement mechanism (such as a tag).
+The details of value structure depend on the specific machine being modelled.
+Intuitively, the payload represents the part of the value that is relevant to
+the behavior of the basic machine, and it is the part that should be changed
+when constructing variants (see below); the rest of the value may contain
+information relevant to a hardware enforcement mechanism (such as a tag), which
+should not be varied.
 We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
 and lift it element-wise to states.
-In the tagged setting, values are compatible if their tags are identical.
-\apt{This is still not very convincing!
-  empting to say that only the payloads can affect the step function, but that is
-  too strong, right? (OK for tags, but not, e.g. capabilities?}
-\sna{Yeah, I don't think we can say something universal about when values should be
-  considered compatible.}
+In a tag-based enforcement scheme, values are compatible if their tags are identical.
+%% \apt{This is still not very convincing!
+%%   empting to say that only the payloads can affect the step function, but that is
+%%   too strong, right? (OK for tags, but not, e.g. capabilities?}
+%% \sna{Yeah, I don't think we can say something universal about when values should be
+%%   considered compatible.}
 

### 2023-02-04 submission5/core.tex
done with 1,2 pass

@@
 \cref{sec:relwork,sec:future} related and future work.
+\Cref{app:ptr} gives an (untested) proof-of-concept formalization of a more sophisticated,
+heap-like sharing model based on pointer provenance.
 
@@
 
-\apt{Still working from here down}
-
 Exactly which operations and rules are needed depends on
@@
 The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
-\cref{tab:psi}.\bcp{I was confused by the table because I looked at
-  the figure before I read that sentence!} Those that we explore in detail in \cref{sec:example,sec:formal}
-are shown in white: calls, returns, and the allocation and deallocation of local memory.
-In \cref{sec:extensions}, we cover additional operations and and parameters
-to support various forms of memory sharing \ifexceptions , exceptions, \fi
-and tail-call elimination; these are shown in light gray. These are also
-the operations we test in \cref{sec:testing}.
-We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
-based on pointer provenance in \cref{app:ptr}\bcp{first
-  mention of this appendix; if it's important, mention it in the
-  outline in section I.  Otherwise drop it.}, represented by the dark gray
-operations, but we do not test this.
+\cref{tab:psi}. A core set of operations covering calls, returns, and local
+memory is introduced in the example in \cref{sec:example}
+and formalized in \cref{sec:formal}. An extended set covering simple memory sharing and
+tail-call elimination are described in \cref{sec:extensions} and tested in \cref{sec:testing}.
+The remaining operations are needed for the capability-based model in \cref{app:ptr}. 
+%\bcp{I was confused by the table because I looked at
+% the figure before I read that sentence!}
+%Those that we explore in detail in \cref{sec:example,sec:formal}
+%are shown in white: calls, returns, and the allocation and deallocation of local memory.
+%In \cref{sec:extensions}, we cover additional operations and and parameters
+%to support various forms of memory sharing \ifexceptions , exceptions, \fi
+%and tail-call elimination; these are shown in light gray. These are also
+%the operations we test in \cref{sec:testing}.
 
@@
 
+\apt{Hopefully better now.}
+
 \newcommand{\example}{\rowcolor{black!0}}
@@
 \begin{center}
-  \begin{tabular}{| l | l |}
+  \begin{tabular}{| l | l | l |}
     \hline
-    \(\psi \in \Psi\)\bcp{huh?} & Parameters \\
+    Operation \(\psi \in \Psi\) & Parameters & Sections\\
     \hline
-    \example \(\mathbf{call}\) & target address, argument registers \\
-    \testing & stack arguments (base, offset \& size) \\
-    \example \(\mathbf{return}\) & \\
-    \example \(\mathbf{alloc}\) & offset \& size \\
-    \testing & public flag \\
-    \example \(\mathbf{dealloc}\) & offset \& size \\
-    \testing \(\mathbf{tail call}\) & (same as for \(\mathbf{call}\)) \\
-    \hline
-    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
-    \hline
-    \theory \(\mathbf{promote}\) & register, offset \& size \\
-    \theory \(\mathbf{propagate}\) & source register/address \\
-    \theory & destination register/address \\
-    \theory \(\mathbf{clear}\) & target register/address \\
+    \example \(\mathbf{call}\) & target address, argument registers & \ref{sec:example},\ref{sec:formal}\\
+    \testing & stack arguments (base, offset \& size) & \ref{sec:extensions},\ref{sec:testing} \\
+    \example \(\mathbf{return}\) & & \ref{sec:example},\ref{sec:formal}\\
+    \example \(\mathbf{alloc}\) & offset \& size & \ref{sec:example},\ref{sec:formal}\\
+    \testing & public flag & \ref{sec:extensions},\ref{sec:testing} \\
+    \example \(\mathbf{dealloc}\) & offset \& size &  \ref{sec:example},\ref{sec:formal}\\
+    \testing \(\mathbf{tail call}\) & (same as for \(\mathbf{call}\)) & \ref{sec:extensions},\ref{sec:testing} \\
+%    \hline
+%    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
+%    \hline
+    \theory \(\mathbf{promote}\) & register, offset \& size & \ref{app:ptr} \\
+    \theory \(\mathbf{propagate}\) & source register/address & \ref{app:ptr} \\
+    \theory & destination register/address & \ref{app:ptr} \\
+    \theory \(\mathbf{clear}\) & target register/address & \ref{app:ptr} \\
     \hline
@@
 \end{center}
-\caption{Security-relevant operations and their parameters.}
+\caption{Security-relevant operations and their parameters, with the Sections in which
+  they are first used in the paper.}
   \label{tab:psi}
@@
 State elements that are outside of the stack---general-purpose memory used for
-globals and the heap, as well as the code region and globally-relevant
-registers\bcp{what is a globally-relevant
-register?}---are always labeled \(\public\). We place security requirements on some
-\(\public\) elements for purposes of our \(\wbcf\) property, and a
+globals and the heap, as well as the code region and globally shared
+registers---are always labeled \(\public\). We place security requirements on some
+\(\public\) elements for purposes of the \(\wbcf\) property, and a
 given enforcement mechanism
@@
 Arguments are marked \(\object\), meaning that their contents may be safely used.
-When the active function\bcp{are functions active, or are their
-  arguments active?} allocates memory, that memory will also be \(\object\).
+When a function allocates memory for its own stack frame, that memory will also be \(\object\).
 Then, on a call, \(\object\) elements that are not being used to communicate with
@@
 constraints that our design puts on any system. In particular, we must
-trust that the method of distinguishing security-relevant operations
-is accurate.  For example, if the enforcement mnechanism involves
-labels placed on code by a compiler, we must trust that the compiler
-placed those labels correctly. For instance, if a compiled function
+trust that the security-relevant operations have been correctly labeled.
+If a compiled function
 call is not marked as such, then the caller's data might not be
-protected from the callee.  Conversely, marking too many operations as
+protected from the callee; conversely, marking too many operations as
 calls will simply cause otherwise safe programs to be rejected.

### 2023-02-04 submission5/core.tex
misc. cleaning

@@
 its {\em security semantics}, and the stack safety properties built on these.
+\Cref{sec:extensions} describes how to support an extended set of code features.
 In \Cref{sec:enforcement} we describe the micro-policies that we test,
@@
 
-\section{Expanding the Machine}
+\section{Extended Code Features}
 \label{sec:extensions}
 
-The system we model in \cref{sec:example,sec:formal} is very simple, but our properties
-are designed to make it easy to add features. To support argument passing on the stack,
-we just new parameters to the existing security-relevant operations, and refine how they
+The system we model in \cref{sec:example,sec:formal} is very simple, but our framework
+is designed to make it easy to add support for additional code features. To support argument passing on the stack,
+we just add new parameters to the existing security-relevant operations, and refine how they
 update the security context. The remainder of the properties do not change at all.
@@
 we add it to the definition of call targets.
-The rules for the expanded machine are given in \cref{fig:advops}; the
+The rules for the extended security semantics are given in \cref{fig:advops}; the
 rules in \cref{fig:basicops} can be recaptured by instantiating

### 2023-02-04 submission5/core.tex
wibbles

@@
 they are stated abstractly in hopes of applying
-to other enforcement mechanisms.}
+to other enforcement mechanisms.}\apt{I think we should try to keep something
+  along these lines in the intro, if not this particular sentence.}
 
@@
 \(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
-The details of value structure depend on the specific machine being modelled.
-Intuitively, the payload represents the part of the value that is relevant to
-the behavior of the basic machine, and it is the part that should be changed
-when constructing variants (see below); the rest of the value may contain
-information relevant to a hardware enforcement mechanism (such as a tag), which
-should not be varied.
-We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
-and lift it element-wise to states.
-In a tag-based enforcement scheme, values are compatible if their tags are identical.
+Depending on the specific machine being modeled, values may also contain other
+information relevant to hardware enforcement (such as a tag). 
+%Intuitively, the payload represents the part of the value that is relevant to
+%the behavior of the basic machine, and it is the part that should be changed
+When constructing variants (see below), this additional information should
+not be varied. To capture this idea, we assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
+and lift it element-wise to states.  Two values should be compatible if their
+non-payload information (e.g. their tag) is identical.
 %% \apt{This is still not very convincing!
@@
 SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
-DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey in
-Sections~\ref{sec:intro} and~\ref{sec:future}. Among these techniques, the micro-policies of Roessler and
+DBLP:conf/sp/RoesslerD18, Gollapudi+23}.
+%, which we survey in Sections~\ref{sec:intro} and~\ref{sec:future}.
+Among these techniques, the micro-policies of Roessler and
 DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
@@
 
-The only other line of work that sets out to rigorously characterize the
+To our knowledge, the only other line of work that sets out to rigorously characterize the
 security of the stack is the Cerise family of program logics
@@
   depending on whether we're interested in delving deeper.}\bcp{If
-  there's space, some of this could be resuscitated in the related
-  work section.  But removing it from here is a good call.}
+  there's space, some of this could be resuscitated.}
+\apt{It would be very good to do this, because reviewers will want a careful comparison with this work,
+  but I don't feel quite confident enough to do it (especially after my unfortunate attempt last time)!}
 %For instance, the stack-safety specification from Georges et al.
@@
 
-HardBound \cite{DeviettiBMZ08} is another hardware extension related to
+\proposecut{HardBound \cite{DeviettiBMZ08} is another hardware extension related to
 capabilities, which closely follows SoftBound \cite{NagarakatteZMZ09} and
-simply aims to accelerate its enforcement of memory safety for compiled code.
+simply aims to accelerate its enforcement of memory safety for compiled code.}
 
@@
 that our properties should exhibit: applicability to different system
-features and different security requirements. Another \proposecut{important dimension}
-is applicability to multiple enforcement mechanisms. \apt{Could we mention that
-  in intro as well?}  \proposecut{There is much future work in each of these dimensions.}
+features and different security requirements. Another 
+is applicability to multiple enforcement mechanisms.
 

### 2023-02-04 submission5/core.tex
comments on appendix

@@
 %% Title information
-\title{Formalizing Stack Safety as a Security Property}
-
+% RB: same amount of whitespace between title and text, also when authors are anonymized
+\title{Formalizing Stack Safety as a Security Property\ifcameraready\else\vspace{-0.75cm}\fi}
+
+\ifcameraready
 \author{
@@
 }
+\fi
 
@@
 which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
-Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
-choosing 43. As execution continues after the return
+Consider a variant state in which {\tt sensitive} has any other value, % arbitrarily choosing
+say 43. As execution continues after the return
 from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
@@
 \((\mach,(V,\sigma))\), we take an arbitrary \(\nach\) that is a \(\components\)-variant
-of \(\mach\), where \(\components\) is the set all elements that are not \(\object\)
+of \(\mach\), where \(\components\) is the set elements that are not \(\object\)
 or \(\public\). The remainder of the property is identical to \(\clrc\).
@@
 setting. Arguments may be passed on the stack because there are too many
-to pass in registers, as an implementation of variadic arguments, or
+to pass in registers, as % an implementation of
+variadic arguments, or
 because they are composite types that inherently have
@@
 
-Using this mechanism, an call-by-value argument passed on the stack at an \(\SP\)-relative offset
+Using this mechanism, a call-by-value argument passed on the stack at an \(\SP\)-relative offset
 is specified by the triple \((\SP, \mathit{off}, \mathit{sz})\).
 In this case, only the immediate callee gains access to the argument location.
-A C++-style call-by-reference argument where the reference is passed in \(\reg\)
+A C++-style call-by-reference argument % where the reference is
+passed in \(\reg\)
 is instead specified by \((\reg, 0, \mathit{sz})\), which allows
@@
 
-The operation rule for a tail call is similar to that for a normal call.
+The rule for a tail call is similar to that for a normal call.
 We do not push the caller's view onto the stack,
@@
 than \(\sealed\) to express this. In terms of \cref{tab:props}, we replace
+% RB: could shorten "in terms of"
 ``call'' with ``call or tail call'' in \(\clrc\), \(\clec\), and \(\clei\).
@@
 to ensure that our coinductive trace definitions are well-formed, to reason about
-them\leo{where do we mention reasoning about them?}, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.
+them\leo{where do we mention reasoning about them?}, and to leverage the power of the QuickChick % property-based
+testing framework~\cite{Pierce:SF4}.
 
@@
 observed to avoid early failstops, extra care must be taken to ensure
-that any such calls are well behaved. For example, immediately
-following a call we need to initialize any stack-allocated variables
+that % any such
+calls are well behaved. For example, immediately
+after a call we need to initialize any stack-allocated variables
 with an increased probability. However, at the same time, we need to
@@
 To ensure the effectiveness of testing against our formal properties, we
-use {\em mutation testing}~\cite{JiaH11}. In mutation testing, we inject errors
+use {\em mutation testing}~\cite{JiaH11} to inject errors
 (mutations) in a program that should cause the property of interest (here,
@@
 current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
-that restriction by not checking the depth equality (row {\em
+that restriction by omitting the check (row {\em
   LOAD\_NO\_CHECK\_DI}).
-
+%
 Similarly, when storing to a stack location, the correct micro-policy
@@
 Naturally, testing also revealed a number of errors in our
-implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code;
+implementation of the enforcement mechanism (originally written in C++
+and targeting ARM machine code;
 %\bcp{right?}\leo{yeah}
-we re-implemented it in Coq targeting RISC-V).  These errors range
-from trivial typos to ones that require an intriguingly complex setup
+then re-implemented in Coq targeting RISC-V).  These errors range
+from trivial typos to ones that require a %n intriguingly
+complex setup
 to provoke.  The most interesting bug (included in the table as row
@@
 It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
-taking twice as long to catch as the next longest.
-
-Our properties have allowed us to identify an enforcement mechanism as
+%taking twice as long to catch as the next longest.
+taking longer to catch.
+
+Our properties allow us to identify an enforcement mechanism as
 not really stack safe, and to validate a possible fix.
@@
 %, which we survey in Sections~\ref{sec:intro} and~\ref{sec:future}.
-Among these techniques, the micro-policies of Roessler and
+% Among these techniques,
+The micro-policies of Roessler and
 DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
@@
 %
-Under a bounds checking discipline such as SoftBound \cite{NagarakatteZMZ09}, all the pointers
-in a program are extended with some disjoint metadata, and these are combined
+Under a bounds checking discipline~%such as SoftBound
+\cite{NagarakatteZMZ09}, all the pointers
+in a program are extended with some disjoint metadata %, and these are combined
+used
 to gate memory accesses. These approaches enforce a form of \emph{memory safety},
@@
 
+\apt{Sbould admit up front that this extension requires a fundamental change in the
+  security context.}
+
 What if we want to express a finer-grained notion of safety, in which
@@
 
-Note that we have a degree of monotonicity here. Once an object is sealed (because its
+Note that we have a degree of monotonicity \apt{??} here. Once an object is sealed (because its
 capability has not been passed to a callee), subsequent nested calls can never unseal it.

### 2023-02-04 submission5/core.tex
Done with figures, and some tweaks

@@
 
-% The stack plays a central role in realizing large number of
-% high-level language features, some of which are typically excluded
-% from discussions of stack safety.
 Our properties go further than previous formal definitions of stack
@@
 %   these, only exceptions are high-level language features!}
-\proposecut{They are modular by design so that they can be further extended to
-other features.}
 %
@@
 mechanisms.
-\item
-  %We instantiate\bcp{extend??} these definitions in a realistic
-%setting with features such as\bcp{``such as''? are there others?}
-%argument passing on the stack, callee-saves registers, and tail-call
-  %elimination.\bcp{How about:
- We extend these core definitions to
+\item We extend these core definitions to
   describe a realistic setting with argument passing on the stack,
@@
 security-relevant operations by labeling selected instructions.
-%% An instantiation of
-%% an operation
-%% might require some external mechanism for distinguishing which instructions are
-%% security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique,
-%% which we use in our examples, is to add labels to selected instructions, but the theory does not
-%% depend on this choice. One could instead recognize some operations dynamically. For instance,
-%% if we make a call to untrusted (and unlabeled) code, we might decide to recognize
-%% any jump back to trusted code as a return, to which \(\wbcf\) needs to apply.
 
@@
 \paragraph*{Instantiating the Framework}
+
 Conceptually, the following steps are needed to instantiate the framework to a specific machine
@@
 operations as appropriate; (iv) specify the form of observable events.
-
-\bcp{cut this?}In \cref{sec:example} we give such an instantiation, and
-  use it to illustrate our security properties.  But we stress that
-  the property definitions themselves are independent of the
-  instantiation.
 
@@
 calls will simply cause otherwise safe programs to be rejected.
-
-\bcp{Still working here:}
 
@@
 \(\sealed\) in \(V_2\) (see \cref{fig:variant2}.)
-            [TODO: fix figure to avoid varying the entire stack.]
 If we take a trace of execution from each state until it returns,
@@
     %
-    \stackrel{\texttt{res}}{\rightbox[2]}
-    \stackrel{\texttt{sens}}{\rightbox[3]}
-    \stackrel{\texttt{sec}}{\rightbox[4]}
-    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
+    \stackrel{\texttt{res}}{\rightbox[1]}
+    \stackrel{\texttt{sens}}{\rightbox[2]}
+    \stackrel{\texttt{sec}}{\rightbox[3]}
+    \stackrel{\texttt{ra}}{\rightbox[4]\rightbox[5]}}
     \]
-    \[%\raisebox{.5\height}{\Bigg\Downarrow} \hspace{1cm}
+    \[\raisebox{.5\height}{\Bigg\Downarrow} \hspace{1cm}
     \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
-    %\raisebox{\height}{\not\approx}
+    \raisebox{\height}{\(\not\approx\)}
     \hspace{0.5cm}
-    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{4}} \hspace{1cm}
-    %    \raisebox{.5\height}{\Bigg\Downarrow}
+    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{3}} \hspace{1cm}
+    \raisebox{.5\height}{\Bigg\Downarrow}
     \]
@@
     %
-    \stackrel{\texttt{res}}{\rightbox[2]}
-    \stackrel{\texttt{sens}}{\rightbox[3]}
-    \stackrel{\texttt{sec}}{\rightbox[4]}
-    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
+    \stackrel{\texttt{res}}{\rightbox[1]}
+    \stackrel{\texttt{sens}}{\rightbox[2]}
+    \stackrel{\texttt{sec}}{\rightbox[3]}
+    \stackrel{\texttt{ra}}{\rightbox[4]\rightbox[5]}
     \]
@@
 elements in \(V_2\), but capturing the required property is subtle.
-
-To illustrate the issues, note that {\tt f} has also stored a 0 below the stack pointer.
 %
@@
     %
-    \stackrel{\texttt{res}}{\rightbox[2]}
-    \stackrel{\texttt{sens}}{\rightbox[3]}
-    \stackrel{\texttt{sec}}{\rightbox[4]}
-    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
+    \stackrel{\texttt{res}}{\rightbox[1]}
+    \stackrel{\texttt{sens}}{\rightbox[2]}
+    \stackrel{\texttt{sec}}{\rightbox[3]}
+    \stackrel{\texttt{ra}}{\rightbox[4]\rightbox[5]}}
     \]
-    \[\large\Downarrow \hspace{1cm} \mathtt{a0} ~ \leftbox[0] \hspace{1cm}
-    \rightbox[7] ~ \mathtt{a0} \hspace{1cm} \large\Downarrow\]
+    \[\raisebox{.5\height}{\Bigg\Downarrow} \hspace{1cm} \stackrel{\mathtt{a0}}{\leftbox[\bf 0]} \hspace{1cm}
+    \stackrel{\mathtt{a0}}{\rightbox[\bf 6]} \hspace{1cm} \raisebox{.5\height}{\Bigg\Downarrow}\]
     \[
@@
     %
-    \stackrel{\texttt{res}}{\rightbox[2]}
-    \stackrel{\texttt{sens}}{\rightbox[3]}
-    \stackrel{\texttt{sec}}{\rightbox[4]}
-    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
+    \stackrel{\texttt{res}}{\rightbox[1]}
+    \stackrel{\texttt{sens}}{\rightbox[2]}
+    \stackrel{\texttt{sec}}{\rightbox[3]}
+    \stackrel{\texttt{ra}}{\rightbox[4]\rightbox[5]}
     \]
-    \[\mathtt{a0} ~ \leftbox[0] \hspace{1cm}
-    \rightbox[0] ~ \mathtt{a0}\]
+    \[\stackrel{\mathtt{a0}}{\leftbox[\bf 5]} \hspace{1cm}
+    \stackrel{\mathtt{a0}}{\rightbox[\bf 3]}\]
   \caption{Return-time Confidentiality Violation}
@@
 %
-Now consider three elements: the address \(\SP - 4\), the address \(\SP + 12\),
-and the register {\tt a0}. The execution from the right-hand state to its return
-has changed the value at \(\SP - 4\), but that value matches that of the
-left-hand variant at its return. Therefore, the change does not represent a leak.
-
-The return states disagree on the value of \(\SP + 12\). But in neither
-case has that value changed since the original variants. So the difference is inherited from
-the original variation, and does not represent a leak either. We do not continue executing the
-variant state after return, so these values will not cause the caller to behave differently.
-
-But in the case of {\tt a0}, the value has changed during the call (in both the original
-and the variant, although only one of these would be necessary), and its final value
-differs between the variants.
-Therefore, it must depend on a secret (in fact, the variable {\tt secret}).
+The return states do not agree on the value of {\tt a0}, but then, the disagree on
+all of the other states, too. The difference is that {\tt a0} has changed during the call
+(in both the original and the variant, although only one of these would be necessary).
+The disagreement between the return states must therefore depend on a secret
+(in fact, the variable {\tt secret}).
 Unless {\tt a0} happens to be irrelevant to the caller, this is a violation of what
@@
 We therefore choose to systematically \emph{test} their micro-policies.
-\leo{moved the following here, I think it makes more sense here.}
 Our primary testing targets are the eager {\em Depth Isolation}
@@
 micro-policy. We choose the Coq proof assistant as the setting for our implementation
-to ensure that our coinductive trace definitions are well-formed, to reason about
-them\leo{where do we mention reasoning about them?}, and to leverage the power of the QuickChick % property-based
+to ensure that our definitions are well-formed and to leverage the
+power of the QuickChick % property-based
 testing framework~\cite{Pierce:SF4}.
@@
 and targeting ARM machine code;
-%\bcp{right?}\leo{yeah}
 then re-implemented in Coq targeting RISC-V).  These errors range
@@
 
-\proposecut{HardBound \cite{DeviettiBMZ08} is another hardware extension related to
-capabilities, which closely follows SoftBound \cite{NagarakatteZMZ09} and
-simply aims to accelerate its enforcement of memory safety for compiled code.}
-
 \section{Future Work}
@@
 
-\paragraph*{Proofs}
-
-The present work relies on randomized testing to validate properties versus
-enforcement mechanisms, but full formal proofs are certainly possible in principle,
-and we would like to investigate whether they are feasible in practice.
-\proposecut{Having formalized our definitions in Coq gives us a head
-  start in this effort.} \bcp{Or just just the whole paragraph.}
-
 \bibliographystyle{IEEEtran}
@@
 
-Note that we have a degree of monotonicity \apt{??} here. Once an object is sealed (because its
+Int the resulting property, once an object is sealed (because its
 capability has not been passed to a callee), subsequent nested calls can never unseal it.

### 2023-02-04 submission5/core.tex
More Cerise related work

--- 2023-02-04 submission5/core.tex
@@
 its {\em security semantics}, and the stack safety properties built on these.
-\Cref{sec:extensions} describes how to support an extended set of code features.
+\Cref{sec:extensions} describes how to support an extended set of features.
 In \Cref{sec:enforcement} we describe the micro-policies that we test,
@@
 we do not assume the use of any particular compiler.
-The machine may possibly be enhanced with security
+The machine may possibly be enhanced with % security
 enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
@@
 % Among these techniques,
-The micro-policies of Roessler and
+%
+Many enforcement techniques focus purely on \(\wbcf\);
+others combine this with some degree of memory protection,
+chiefly focusing on integrity.
+%
+%The micro-policies of
+Roessler and
 DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
-range of safeguards including lazy enforcement, and suitable for formulating
-and validating strong security guarantees.
-
-To our knowledge, the only other line of work that sets out to rigorously characterize the
+range of safeguards including lazy enforcement, and suitable to formulate
+and validate strong security guarantees.
+
+To our knowledge, the only other line of work that aims to rigorously characterize the
 security of the stack is the Cerise family of program logics
 \cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
+Most of those define experimental types of hardware capabilities to achieve
+efficient implementations.
+%
 These works define secure calling conventions for capability machines and
 overlay semantics that internalize those conventions and their related stack
-safety properties, phrased in terms of logical relations. Most flavors of
-Cerise define experimental types of hardware capabilities in order to obtain
-efficient calling conventions. Their latest paper
-\cite{Georges22:TempsDesCerises} uses our work to extend their notion of stack
-safety to include caller confidentiality (\(\clrc\)).
+safety properties, phrased in terms of logical relations instead of trace
+properties.
+%
+Originally, they describe ``local state
+encapsulation''~\cite{SkorstengaardSTKJFP} in terms of integrity (but it is
+equivalent to % the combination of
+\(\clri\) {\em and} \(\clrc\)). Their latest
+paper \cite{Georges22:TempsDesCerises} was inspired by the properties presented
+in this paper to extend their formalism to include confidentiality. When
+checking if our properties applied to their old calling convention, they noted that
+it did not enforce \(\clec\), and made sure that their new version
+would in addition to building it into their formalism
+\cite{Georges22:personalcommunication}. This demonstrates the benefit of our
+choice to explicitly state properties in security terms: specifying security
+% mechanisms
+is hard, and when the spec takes the form of a ``correct by
+construction'' machine, it is easy to neglect a non-obvious security
+requirement.
 

### 2023-02-04 submission5/core.tex
Resolved more comments

@@
   of properties: confidentiality and integrity for callee and caller,
-plus well-bracketed control-flow. Different subsets of these
-properties characterize the security guarantees offered by mechanisms
-that support only some aspects of stack safety\bcp{where do we show
-  this?}. \sna{Nowhere clearly enough. We need to make it clear in section 6
-  that Nick and Andre are included there---only aiming at caller-side,
-  not callee-side. The discussion of CHERI work should touch on this as well.}
-The properties they are parameterized over a notion of
+  plus well-bracketed control-flow.
+  The properties are parameterized over a notion of
   external observation, allowing them to characterize lazy enforcement
@@
   modular enough that handling these features is straightforward.
-\bcp{Throughout, the way we talk about the ``core model'' and its
-  intantiation / extension / enrichment with realistic features feels
-  quite weasily.}
 \item We validate a tag-based enforcement mechanism, \emph{Lazy
@@
 memory-mapped regions, etc.
-%The nature of events is a parameter of the framework, and can be specialized
-%to any notion of observable behavior.
-
-%% \paragraph*{Variants and Non-interference}
 
@@
 from the theory of non-interference.  For example, to show that certain locations
-are kept secret, it suffices to vary their contents them and check that the observable
+are kept secret, it suffices to vary their contents and check that the observable
 output does not change.
-
-%Non-interference is a way of expressing
-%the flow of information through the system. If we begin with a state \(\mach\)
-%and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
-%then execute both \(\mach\) and \(\nach\), any differences in the outputs they
-%produce leaks information about \(\component\).
 
@@
 of this section we describe the underlying framework in more detail.
-
-%The security principals in our account of stack safety are function
-%activations: the context tracks, for the active function and all pending
-%functions, their {\em view} of the security state of the system.
 
@@
 evolves dynamically through the execution of security-relevant operations,
-which include events like calls, returns, and frame manipulation.
+which include calls, returns, and frame manipulation.
 Our security properties are phrased in terms of this context, often as predicates
@@
 or as relations on traces of future execution
-(hyper-properties).\bcp{This is the first we've mentioned
-  hyperproperties.  CSF readers may be surprised.}
-\sna{Surprised that we have hyperproperties,
-  or that we made it this far without mentioning them?}
-\bcp{The latter!}
-
-Security-relevant operations typically correspond to
-single underlying machine instructions,
-but the correspondence may not be obvious just by inspecting the machine code.
-For example, in the tagged RISC-V architecture that we use in our
-examples and tests,
+(hyper-properties).
+
+Security-relevant operations abstract over the implementation details of the
+actions they take. In this paper, each operation corresponds to a
+single underlying machine instruction. One instruction may perform multiple
+operations. We assume that a compiler or other trusted source has provided
+labels to disambiguate instructions that have multiple purposes. For instance,
+in the tagged RISC-V architecture
+that we use in our examples and tests,
 calls and returns are conventionally performed using the {\tt jal}
@@
 and {\tt jalr} (``jump-and-link-register'') instructions, but these
-instructions might also be used for other things. Thus we identify
-security-relevant operations by labeling selected instructions.
+instructions might also be used for other things.
 
@@
 The remaining operations are needed for the capability-based model in \cref{app:ptr}.
-%\bcp{I was confused by the table because I looked at
-% the figure before I read that sentence!}
-%Those that we explore in detail in \cref{sec:example,sec:formal}
-%are shown in white: calls, returns, and the allocation and deallocation of local memory.
-%In \cref{sec:extensions}, we cover additional operations and and parameters
-%to support various forms of memory sharing \ifexceptions , exceptions, \fi
-%and tail-call elimination; these are shown in light gray. These are also
-%the operations we test in \cref{sec:testing}.
-
-\bcp{As a reader, I'm feeling a bit overwhelmed at this point.
-  Everything feels fluid and flexible---''It could be like this, but
-  it could also be another way if you want...''  I know we're past the
-  point of doing more than tweaking the way we talk about things, but
-  if there are any tweaks that reduce the sense of flexibility, let's
-  do them!}
-
-\sna{Agreed. The point of the operations was originally that we shouldn't
-  have to think about all of these things when presenting the formalism,
-  not that we should open with all the different ways we can think about them.}
-
-\apt{Hopefully better now.}
 
@@
 We now give a formal description of our machine model, security semantics,
-and properties. Our definitions abstract over: (1) the details of  the target machine
-architecture and ABI, (2) the set of security-relevant operations and their effects on
-the security context, (3) the set of observable events, and (4) a notion of value compatibility.
-%They can be extended to a more sophisticated machine model by expanding the set of
-%security-relevant events and defining
-\apt{Any more? Give a more listlike presentation?
-  Need to clarify exactly what the parameters are, both in the initial
-  presentation and in the various extensions, including the appendix.}
-\sna{Need to come back to this.}
+and properties. Our definitions abstract over: (i) the details of  the target machine
+architecture and ABI, (ii) the set of security-relevant operations and their effects on
+the security context, (iii) the set of observable events, and (iv) a notion of value compatibility.
 
@@
 the reference to be passed further down the stack.
-\apt{What if the reference is itself passed on the stack?}
-\sna{Ugh. We could make \(\reg\) an arbitrary element, but that seems like it's getting
-  dangerously close to indirection.}
 
@@
 %
-%The micro-policies of
-Roessler and
-DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
-range of safeguards including lazy enforcement, and suitable to formulate
-and validate strong security guarantees.
+Roessler and DeHon's {\it Depth Isolation} and {\it Lazy Tagging and Clearing}
+\cite{DBLP:conf/sp/RoesslerD18} both offer protections corresponding to
+\(\wbcf\), \(\clri\), and \(\clrc\), though they do not give a formal description
+of this. They are generally not concerned with protecting callees.
 
@@
 requirement.
-
-\sna{See commented out tex for content that could be massaged to fit in here,
-  depending on whether we're interested in delving deeper.}\bcp{If
-  there's space, some of this could be resuscitated.}
-\apt{It would be very good to do this, because reviewers will want a careful comparison with this work,
-  but I don't feel quite confident enough to do it (especially after my unfortunate attempt last time)!}
+%
 %For instance, the stack-safety specification from Georges et al.
@@
 
-\apt{Sbould admit up front that this extension requires a fundamental change in the
-  security context.}
-
 What if we want to express a finer-grained notion of safety, in which
@@
 point to a given object, and non-pointer integers as well as pointers to other objects.
+This constitutes a substantial addition to the security context.
 
@@
 
-Int the resulting property, once an object is sealed (because its
+In the resulting property, once an object is sealed (because its
 capability has not been passed to a callee), subsequent nested calls can never unseal it.

### 2023-02-04 submission5/core.tex
Extended CHERI discussion further

@@
 Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
-\proposecut{While our properties are motivated by the desire to specify \(LTC\) precisely,
+While our properties are motivated by the desire to specify \(LTC\) precisely,
 they are stated abstractly in hopes of applying
-to other enforcement mechanisms.}\apt{I think we should try to keep something
-  along these lines in the intro, if not this particular sentence.}
+to other enforcement mechanisms.
 
@@
 step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
+Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
+(and thus silently diverging.)
 %% \sna{I think that the payload concept is too nitty-gritty for this section.
@@
 
-  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; addresses increase to the right and the stack grows to the left.}
+  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; the stack grows to the left.}
   \label{fig:main}
@@
 To our knowledge, the only other line of work that aims to rigorously characterize the
-security of the stack is the Cerise family of program logics
-\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
-Most of those define experimental types of hardware capabilities to achieve
-efficient implementations.
+security of the stack is the StkTokens-Cerise family of CHERI-enforced secure calling
+conventions \cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
 %
@@
 %
-Originally, they describe ``local state
-encapsulation''~\cite{SkorstengaardSTKJFP} in terms of integrity (but it is
-equivalent to % the combination of
-\(\clri\) {\em and} \(\clrc\)). Their latest
+Originally, they describe ``local state encapsulation''~\cite{SkorstengaardSTKJFP}
+in terms of integrity only (but it has confidentiality, equivalent
+to \(\clri\) {\em and} \(\clrc\)). Their latest
 paper \cite{Georges22:TempsDesCerises} was inspired by the properties presented
@@
 choice to explicitly state properties in security terms: specifying security
-% mechanisms
 is hard, and when the spec takes the form of a ``correct by
@@
 requirement.
+
+In terms of direct feature comparison with Georges et al. (being the most
+recent work in the line), with the addition of confidentiality to their formalism, we
+are roughly at parity in terms of the expressiveness of our properties.
+We have additionally proposed callee-integrity, but it is probably the least
+practical of our properties. We extend our model to tailcalls, which they do
+not, and to the passing of pointers to stack objects. They discuss stack objects
+and the interaction between stack and heap, but their calling convention does not
+guarantee safety in the presence of pointer passing without additional checks.
+We test a limited degree of pointer passing, which does not guarantee memory
+safety for the passed pointer but which does not undermine the security of its
+frame, and we offer an untested formalism for memory-safe passing of pointers.
+On the other hand, their properties are validated by proof, while ours are
+only tested.
+
 %
@@
 
-In \cref{sec:intro}, we discussed several dimensions of flexibility
-that our properties should exhibit: applicability to different system
-features and different security requirements. Another 
-is applicability to multiple enforcement mechanisms.
-
-% \subsection{Testing The Stack Safety Menagerie}
-
-\paragraph*{Enforcement with capabilities}
-Capability machines can represent and manipulate
-unforgeable tokens of authority over specific regions of
-memory using a combination of dedicated out-of-band memory and
-specialized ISA instructions.
-%
-The most prominent contemporary representative of this long line of work is
-CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed to provide efficient fine-grained
+We plan to test our properties against multiple enforcement mechanisms.
+The top priority is capability machines, namely 
+CHERI \cite{DBLP:conf/sp/WatsonWNMACDDGL15}, a modern architecture designed
+to provide efficient fine-grained
 memory protection and compartmentalization.
 %
-Previous work has used these principles in simplified models to implement
-\emph{secure calling conventions} for the call
-stack \cite{SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises}. Those conventions can be seen as combining the
-protections of bounds-checking with software-based methods of control-flow protection.
-%
-%There are several proposals around the use of CHERI capabilities to enforce stack safety,
-%including mechanisms that use the standard CHERI hardware (which includes local
-%capabilities) \cite{SkorstengaardLocal},
-%and others that propose entirely new types of capabilities, such as linear
-%\cite{SkorstengaardSTK}, uninitialized \cite{Georges+21}, lifetime
-%\cite{Tsampas+19}, and directed \cite{Georges22:TempsDesCerises} capabilities.
 We want to test the most recent work by Georges et
@@
 
-\paragraph*{Enforcement in software}
-%
+It would also be interesting to test a software approach.
 Under a bounds checking discipline~%such as SoftBound
@@
 
-\paragraph*{More Features}
+Several popular enforcement mechanisms are not designed to provide
+absolute guarantees of security. For example, stack canaries~\cite{Cowan+98}
+and shadow stacks~\cite{Dang+15,Shanbhogue+19}
+are chiefly hardening techniques: they increase the difficulty
+of some control-flow attacks on the stack, but cannot provide absolute
+guarantees on \(\wbcf\) under a normal attacker model.
+%
+Interestingly, these are lazy enforcement mechanisms, in that
+the attack may occur and be detected some time later, as long as
+it is detected before it can become dangerous. That would make our
+observation-based formalism a good fit for defining their security,
+if we could find a formal characterization of what they do acheive
+(perhaps in terms of a base machine with restricted addressing power).
 
@@
 the model in \cref{app:ptr} for arbitrary memory-safe pointer sharing.
-
-\paragraph*{Weaker Requirements}
-
-Several popular enforcement mechanisms are not designed to provide
-absolute guarantees of security. For example, stack canaries~\cite{Cowan+98}
-and shadow stacks~\cite{Dang+15,Shanbhogue+19}
-are chiefly hardening techniques: they increase the difficulty
-of some control-flow attacks on the stack, but cannot provide absolute
-guarantees on \(\wbcf\) under a normal attacker model.
-%
-Interestingly, these are lazy enforcement mechanisms, in that
-the attack may occur and be detected some time later, as long as
-it is detected before it can become dangerous. That would make our
-observation-based formalism a good fit for defining their security,
-if we could find a formal characterization of what they do acheive
-(perhaps in terms of a base machine with restricted addressing power).
-
-%% \paragraph{Split Stacks}
-%% %
-%% Code-pointer integrity seeks to prevent control-flow hijacking attacks by
-%% splitting the program memory, including the stack, into a regular region and a
-%% safe region; objects that require protection are placed in the latter and
-%% all accesses to them protected by static or dynamic checks.
-%% %
-%% Like shadow stacks, bounds checking is a high-level technique designed to
-%% protect code written in a high-level language, but it is designed to
-%% offer protection under a stronger threat model that also considers the presence
-%% of untrusted code. As with most other software-based methods, it purports to
-%% protect exclusively the control-relevant parts of the stack, making no claims
-%% about its data-relevant parts---we would solely be testing \(\wbcf\).
 

### 2023-02-04 submission5/core.tex
Small space tweaks

@@
 and extend it with a runtime monitor implementing a stack safety
-micro-policy. We choose the Coq proof assistant as the setting for our implementation
+micro-policy. We choose Coq for our implementation
 to ensure that our definitions are well-formed and to leverage the
-power of the QuickChick % property-based
+power of the QuickChick
 testing framework~\cite{Pierce:SF4}.

### 2023-02-04 submission5/core.tex
Space!

@@
 
-Naturally, testing also revealed a number of errors in our
-implementation of the enforcement mechanism (originally written in C++
-and targeting ARM machine code;
-then re-implemented in Coq targeting RISC-V).  These errors range
-from trivial typos to ones that require a %n intriguingly
-complex setup
-to provoke.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
-sequence %/policy combination\apt{??}
-did not initialize all locations for the
-newly allocated stack frame correctly, but left some of them as
-$\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured.
-The smallest counterexample requires calling a function {\tt f},
-which fails to initialize some of its frame,
-but writes into an uninitialized location $l$ later, treating \(l\) as outside
-the stack. Then {\tt f} calls a further function {\tt g} (which should have
-the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
-which is allowed because the enforcement mechanism still has
-$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
-%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
-%  and the write from {\tt f} was already a violation.}
-
-As for \(LTC\), the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
+For \(LTC\), the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
 fails in testing, in cases where data is leaked between sequential calls.
@@
 %
-These works define secure calling conventions for capability machines and
-overlay semantics that internalize those conventions and their related stack
+These works define stack safety as overlay semantics and related stack
 safety properties, phrased in terms of logical relations instead of trace

### 2023-02-04 submission5/core.tex
Intro nit

@@
 \cite{Shanbhogue+19}, not to categorically rule out attacks. Other
-mechanisms, such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
-integrity\cite{Kuznetsov+14}, do aim for stronger guarantees, but even
+mechanisms, such as SoftBound~\cite{NagarakatteZMZ09} and code-pointer
+integrity~\cite{Kuznetsov+14}, do aim for stronger guarantees, but even
 these do not make formal promises.  Indeed, to our knowledge, the sole

### 2023-02-04 submission5/core.tex
Nits

@@
 
-Stack safety properties need to describe the behavior of machine code. but they naturally
+Stack safety properties need to describe the behavior of machine code, but they naturally
 talk about function activations and stack contents---abstractions that
@@
 
-\paragraph*{Machine model}
+\paragraph*{Machine Model}
 We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
@@
 and formalized in \cref{sec:formal}. An extended set covering simple memory sharing and
-tail-call elimination are described in \cref{sec:extensions} and tested in \cref{sec:testing}.
+tail-call elimination is described in \cref{sec:extensions} and tested in \cref{sec:testing}.
 The remaining operations are needed for the capability-based model in \cref{app:ptr}.
@@
 
-\paragraph*{Views and security classes}
+\paragraph*{Views and Security Classes}
 
@@
 security-relevant operations and rules required by the code style; (iii) determine
-how to label machine instruction with security-relevant
+how to label machine instructions with security-relevant
 operations as appropriate; (iv) specify the form of observable events.

### 2023-05-10 submission5/core.tex
Beginning of camera ready edits

@@
 %% comma separated list
-\ifcameraready
-\keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission
-\fi
+%\ifcameraready
+%\keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission
+%\fi
 
@@
 target machine by a secure compiler, stack safety cannot be defined in
-this way, since ``the stack'' is completely transparent in the
+this way, since ``the stack'' is not explicitly present in the
 definitions of most source languages.\footnote{Contrast Azevedo de
@@
 atop the hardware stack.
+
+\sna{Clarify that we do not mean that the notion of a stack is absent, but
+  rather that it is precisely the expectations referenced by reviewer 2 that
+  we are trying to codify. Priority 2.}
 
@@
 \begin{figure}
-  \newcommand{\figonebox}[1][]{\genbox{30px}{lightgray}{#1}}
+  \newcommand{\figonebox}[1][]{\genbox{30pt}{lightgray}{#1}}
   \begin{subfigure}{\columnwidth}
@@
 
-Our security semantics attach a security context to the machine state,
+Our security semantics attaches a security context to the machine state,
 which consists of a view \(V\) and a stack \(\sigma\) of pending activations' views.
@@
   \end{tabular}
-  \caption{Execution of example up through the return from {\tt f}. In stack diagrams, addresses increase to the right, stack grows to the left, and boxes represent 4-byte words.}
+  \caption{Execution of example up through the return from {\tt f}.
+In stack diagrams, addresses increase to the right, stack grows to the left, and boxes represent 4-byte words.
+\sna{Reviewer 2 found s0 and a1 confusing, as well as lack of ra a4 and a5.
+Priority 3.}}
 \label{fig:exec1}
@@
 
-\newcommand{\figfourbox}[1][]{\genbox{20px}{red}{#1}}
+\newcommand{\figfourbox}[1][]{\genbox{20pt}{red}{#1}}
 
@@
   \stackrel{\texttt{res}}{\figfourbox[0]}
-  \stackrel{\texttt{sens}}{\genbox{20px}{red!50}{\bf 42}}
+  \stackrel{\texttt{sens}}{\genbox{20pt}{red!50}{\bf 42}}
   \stackrel{\texttt{sec}}{\figfourbox[5]}
@@
     \stackrel{\texttt{res}}{\figfourbox[0]}
-    \stackrel{\texttt{sens}}{\genbox{20px}{\leftvariant}{\bf 42}}
+    \stackrel{\texttt{sens}}{\genbox{20pt}{\leftvariant}{\bf 42}}
     \stackrel{\texttt{sec}}{\figfourbox[5]}
@@
     \stackrel{\texttt{res}}{\figfourbox[0]}
-    \stackrel{\texttt{sens}}{\genbox{20px}{\rightvariant}{\bf 0}}
+    \stackrel{\texttt{sens}}{\genbox{20pt}{\rightvariant}{\bf 0}}
     \stackrel{\texttt{sec}}{\figfourbox[5]}
     \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}}\]
-  \[\stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
+  \[\stackrel{\hookrightarrow \mathtt{out}}{\genbox{20pt}{white}{5}} \hspace{0.5cm}
   %\raisebox{\height}{\not\approx}
   \hspace{0.5cm}
-  \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{1}}\]
+  \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20pt}{white}{1}}\]
   \caption{Integrity Violation}
@@
 
-\newcommand{\leftbox}[1][]{\genbox{20px}{\leftvariant}{#1}}
-\newcommand{\rightbox}[1][]{\genbox{20px}{\rightvariant}{#1}}
+\newcommand{\leftbox}[1][]{\genbox{20pt}{\leftvariant}{#1}}
+\newcommand{\rightbox}[1][]{\genbox{20pt}{\rightvariant}{#1}}
 
@@
     \[\raisebox{.5\height}{\Bigg\Downarrow} \hspace{1cm}
-    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{5}} \hspace{0.5cm}
+    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20pt}{white}{5}} \hspace{0.5cm}
     \raisebox{\height}{\(\not\approx\)}
     \hspace{0.5cm}
-    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20px}{white}{3}} \hspace{1cm}
+    \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20pt}{white}{3}} \hspace{1cm}
     \raisebox{.5\height}{\Bigg\Downarrow}
@@
 (in both the original and the variant, although only one of these would be necessary).
-The disagreement between the return states must therefore depend on a secret
-(in fact, the variable {\tt secret}).
+The disagreement between the return states must therefore depend on a secret.
 Unless {\tt a0} happens to be irrelevant to the caller, this is a violation of what
 we term {\it return-time confidentiality} (formalized in \cref{tab:props}, line 3b).
+\sna{Clarify, per reviewer 2. Priority 2. Also consider connections to robust
+declassification. Priority 4.}
 
@@
 the principals should occur only through the state elements that are designated for the
-purpose: those labeled \(\public\) and \(\object\).
+purpose: those labeled \(\public\) and \(\object\). \sna{Define state elements before this. Priority 3.}
 
@@
 At each step, the context updates based on an assumed given function
-\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
+\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \Psi \rightarrow \CONTEXTS\).
 Since a single step might correspond to multiple operations, we apply
@@
 We abstract over the events that can be observed in the system, assuming just
-oa given set \(\OBSS\) that contains at least the element \(\tau\), the silent
+a given set \(\OBSS\) that contains at least the element \(\tau\), the silent
 event. Other events might represent certain function calls (i.e., system calls)
@@
 
-\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
+\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{\obs} (\mach',\context')\)}
               {\(|\sigma| \geq d\)}
@@
 written \(\mach \approx_\components \nach\), if \(\mach\sim\nach\) and
-\(\Delta(\mach,\nach) \subseteq \components\).
+\(\Delta(\mach,\nach) \subseteq \components\). \sna{define \(\sim\). Priority 2.}
 %if all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
@@
 \(\nach\) such that \(\mach \approx_{\components} \nach\),
-\(\hookrightarrow (\mach,\context)  \eqsim \hookrightarrow (\nach,\context)\).
+\(\hookrightarrow (\mach,\context) ~ \eqsim ~ \hookrightarrow (\nach,\context)\).
 
@@
 
+\sna{Discuss the distinction that they make between memory safety and
+  dataflow integrity; we have framed stack safety as a dataflow integrity
+  problem. They did not give a formal specification, and their implementation
+  does not enforce ours. Spell out the violation. Priority 1.}
+
 We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
@@
 
+\sna{Discuss false positives, false negatives, and the degree to which
+  testing can increase our confidence. Priority 2.}
+
 \subsection{Test Generation}
+\sna{Restore Leo's old test gen discussion, then ping Leo if it seems that more
+  is needed. Priority 2.}
+
 To use QuickChick, we developed random test-case generators that produce
@@
 compactly checks multiple nested calls.
+\sna{Justify this. Priority 3.}
 
@@
   \hline
-      {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
-      {\em LOAD\_NO\_CHECK\_LT}  & Obs. Integrity & 23.5 & 12.0 \\
+      {\em PER\_DEPTH\_TAG} & Integrity & 189.7 & 8342.5  \\
+      {\em LOAD\_NO\_CHECK\_LT}  & Integrity & 23.5 & 12.0 \\
       {\em LOAD\_NO\_CHECK\_LT}  & Confidentiality & 19.2 & 695.5 \\
-      {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
+      {\em STORE\_NO\_UPDATE} & Integrity & 70 & 80.6  \\
       {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
@@
 
+\sna{Look into wasm compilation per reviewer 1. Priority 4.}
+
 We plan to test our properties against multiple enforcement mechanisms.
@@
 
+\sna{Move to section 5, cleanup. Priority 3.}
+
 What if we want to express a finer-grained notion of safety, in which

### 2023-05-11 submission5/core.tex
Discussing Nick and Andre

@@
 
+\paragraph*{Lazy Applicability}
+
+Roessler and DeHon describe {\em Lazy Tagging and Clearing} as aiming to enforce
+a fundamentally different property than {\em Depth Isolation}. They consider \(DI\)
+a form of memory safety, while they claim that {\em dataflow integrity} enforces a
+form of {\em dataflow integrity}. Because our properties are phrased in
+terms of dataflow, they represent a plausible definition  against which \(LTC\) can be tested.
+
+\sna{Discuss the distinction that they make between memory safety and
+  dataflow integrity; we have framed stack safety as a dataflow integrity
+  problem. They did not give a formal specification, and their implementation
+  does not enforce ours. Spell out the violation. Priority 1.}
+
 \paragraph*{Blessed Sequences}
@@
 and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
-
-\sna{Discuss the distinction that they make between memory safety and
-  dataflow integrity; we have framed stack safety as a dataflow integrity
-  problem. They did not give a formal specification, and their implementation
-  does not enforce ours. Spell out the violation. Priority 1.}
 

### 2023-05-12 submission5/core.tex
Writeup about Nick and Andre.

--- 2023-05-10 submission5/core.tex
@@
 
-We implement and test two micro-policies from Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
+We implement and test two micro-policies inspired by
+Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
 {\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
+Their connections to Roessler and DeHon's work is discussed below.
 They share a common structure: each function activation is assigned a ``color'' \(n\)
@@
 
-\paragraph*{Blessed Sequences}
-
 To implement this discipline, {\em blessed instruction sequences} appear at
@@
 
+\paragraph*{Applicability to R\&D}
+
+In Roessler and DeHon, Lazy Tagging and Lazy Clearing are both optimizations that
+can be applied to their Depth Isolation policy. Our version of \(LTC\)
+corresponds to Depth Isolation with both optimizations applied. How closely
+do our properties correspond to the specifications of their policies?
+
+Roessler and DeHon differentiate between memory safety policies (without lazy optimization)
+and {\em data-flow integrity} policies (with lazy opimization.) Our properties
+are phrased in terms of data flow, which corresponds to the optimized policies,
+and which we argue is also sufficient to test non-optimized Depth Isolation.
+Explicit formal properties are out of their scope, but they do list the
+behaviors that they expect their data-flow integrity policies to prevent.
+Data-flow integrity policies should prevent reads from sealed objects
+(our caller confidentiality), should catch writes to sealed objects
+if they are later read (our caller integrity), and should prevent reads
+from deallocated objects (our callee confidentiality.)
+They also note that Lazy Clearing prevents uninitialized reads,
+which corresponds to our callee integrity.
+
+They note a flaw in Data-flow Integrity: because function activations
+are identified by depth, a dangling pointer into a stack frame might be usable
+when a new frame is allocated at the same depth. Our testing does not discover
+this flaw, because we do not test address-taken objects, but it discovers a
+related flaw under Lazy Tagging and Clearing that does not require
+an object's address to be taken. If an activation reads a location
+that was previously written by an activation at the same depth, it will
+violate callee confidentiality. If that location was in a caller's frame,
+it also violates caller integrity and confidentiality.
+
+Roessler and DeHon propose addressing the dangling-pointer issue by
+tracking both the depth of the current activation and the identity
+of the active function. This would not eliminate all instances, but it
+would require the confidentiality-violating activation to be of the same
+function that wrote the data in the first place, which is a significantly higher bar.
+We propose instead tracking every activation uniquely, which in our tests eliminates
+the issue entirely.
+
 \paragraph*{Protecting Registers}
 
-Roessler and DeHon do not protect registers, since the compiler is in the
-trusted computing base, but we support threat models that do not.
+Roessler and DeHon do not need to protect registers, since the compiler is in their
+trusted computing base, but we target threat models that do not.
 In particular, \(\clri\) requires callee-saved

### 2023-05-12 submission5/core.tex
Use our property nicknames

@@
 Data-flow integrity policies should prevent reads from sealed objects
-(our caller confidentiality), should catch writes to sealed objects
-if they are later read (our caller integrity), and should prevent reads
-from deallocated objects (our callee confidentiality.)
+(\(\clrc\)), should catch writes to sealed objects
+if they are later read (\(\clri\)), and should prevent reads
+from deallocated objects (\(\clec\).)
 They also note that Lazy Clearing prevents uninitialized reads,
-which corresponds to our callee integrity.
+which corresponds roughly to \(\clei\).
 

### 2023-05-12 submission5/core.tex
some clarification of new text and nits

--- 2023-05-11 submission5/core.tex
@@
 Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
-(and thus silently diverging.)
+(and thus silently diverging).
 %% \sna{I think that the payload concept is too nitty-gritty for this section.
@@
 wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
-violating the program's {\it well-bracketed control flow} (\(\wbcf\).)
+violating the program's {\it well-bracketed control flow} (\(\wbcf\)).
 %
@@
 a callee overwrites some data of its caller, but the caller never accesses that data
-(or only does so after re-initializing it.) This would be harmless, with the callee
+(or only does so after re-initializing it). This would be harmless, with the callee
 essentially using the caller's memory as scratch space, but the caller never seeing any change.
@@
 \cref{subfig:direct}. We take a variant state over the set of elements that are
-\(\sealed\) in \(V_2\) (see \cref{fig:variant2}.)
+\(\sealed\) in \(V_2\) (see \cref{fig:variant2}).
 If we take a trace of execution from each state until it returns,
@@
 
-We implement and test two micro-policies from Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
+We implement and test two micro-policies inspired by
+Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
 {\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
+Their precise connection to Roessler and DeHon's work is discussed below.
 They share a common structure: each function activation is assigned a ``color'' \(n\)
@@
 
-\paragraph*{Lazy Applicability}
-
-Roessler and DeHon describe {\em Lazy Tagging and Clearing} as aiming to enforce
-a fundamentally different property than {\em Depth Isolation}. They consider \(DI\)
-a form of memory safety, while they claim that {\em dataflow integrity} enforces a
-form of {\em dataflow integrity}. Because our properties are phrased in
-terms of dataflow, they represent a plausible definition  against which \(LTC\) can be tested.
-
-\sna{Discuss the distinction that they make between memory safety and
-  dataflow integrity; we have framed stack safety as a dataflow integrity
-  problem. They did not give a formal specification, and their implementation
-  does not enforce ours. Spell out the violation. Priority 1.}
-
-\paragraph*{Blessed Sequences}
-
 To implement this discipline, {\em blessed instruction sequences} appear at
@@
 
+\paragraph*{Applicability to Roessler \& DeHon~\cite{DBLP:conf/sp/RoesslerD18}:}
+
+For Roessler and DeHon (henceforward \emph{R\&D}), Lazy Tagging and Lazy Clearing are both optimizations that
+can be applied to their Depth Isolation policy. Our version of \(LTC\)
+corresponds to Depth Isolation with both optimizations applied. How closely
+do our properties correspond to the specifications of their policies?
+
+R\&D differentiate between memory safety policies (without lazy optimization)
+and {\em data-flow integrity} policies (with lazy optimization). Our properties
+are phrased in terms of data flow, and we apply them to both optimized and non-optimized
+Depth Isolation.
+R\&D do not attempt to define explicit formal properties, but they do list the
+behaviors that they expect their data-flow integrity policies to prevent, namely:
+reads from sealed objects
+(our \(\clrc\)), writes to sealed objects
+if they are later read (our \(\clri\)), and reads
+from deallocated objects (our \(\clec\)).
+They also note that Lazy Clearing prevents uninitialized reads,
+which corresponds roughly to our \(\clei\).
+
+R\&D note a flaw in Depth Isolation: because function activations
+are identified by depth, a dangling pointer into a stack frame might be usable
+when a new frame is allocated at the same depth. Our testing does not discover
+this flaw, because we do not test address-taken objects, but it discovers a
+related flaw under Lazy Tagging and Clearing that does not require
+an object's address to be taken. If an activation reads a location
+that was previously written by an earlier activation at the same depth, it will
+violate callee confidentiality. If that location was in a caller's frame,
+it also violates caller integrity and confidentiality.
+
+R\&D propose addressing the dangling-pointer issue by
+tracking both the depth of the current activation and the static identity
+of the active function. This would not eliminate all instances of this issue, but it
+would require the confidentiality-violating activation to be of the same
+function that wrote the data in the first place, which is a significantly higher bar.
+We propose instead tracking every activation uniquely, which should eliminate the
+issue entirely, as it does in our tests.
+
 \paragraph*{Protecting Registers}
 
-Roessler and DeHon do not protect registers, since the compiler is in the
-trusted computing base, but we support threat models that do not.
+R\&D do not need to protect registers, since they include the compiler in their
+trusted computing base, but we target threat models that do not.
 In particular, \(\clri\) requires callee-saved
@@
 of a pointer from one element to another, and one which clears the provenance
-(for instance, when a pointer is modified in place in a way that makes it invalid.)
+(for instance, when a pointer is modified in place in a way that makes it invalid).
 

### 2023-05-15 submission5/core.tex
Two priority 2 things

@@
 this way, since ``the stack'' is not explicitly present in the
-definitions of most source languages.\footnote{Contrast Azevedo de
+definitions of most source languages. It is implied by the semantics
+of features such as calls and return, but is not itself a feature.
+\footnote{Contrast Azevedo de
   Amorim et al.'s work on heap safety
@@
 to its callee. Those are the purview of high-level languages built
-atop the hardware stack.
-
-\sna{Clarify that we do not mean that the notion of a stack is absent, but
-  rather that it is precisely the expectations referenced by reviewer 2 that
-  we are trying to codify. Priority 2.}
+atop the hardware stack\
 
@@
 but in this case, there is no output during the call. Instead the value of {\tt secret} is
-extracted and placed in {\tt a0}, the return value register. We wish to identify this as
-a confidentiality violation, again by considering variants of the \(\sealed\)
-elements in \(V_2\), but capturing the required property is subtle.
+extracted and placed in {\tt a0}, the return value register.
 %
@@
 %
-The return states do not agree on the value of {\tt a0}, but then, the disagree on
-all of the other states, too. The difference is that {\tt a0} has changed during the call
-(in both the original and the variant, although only one of these would be necessary).
-The disagreement between the return states must therefore depend on a secret.
-Unless {\tt a0} happens to be irrelevant to the caller, this is a violation of what
-we term {\it return-time confidentiality} (formalized in \cref{tab:props}, line 3b).
-\sna{Clarify, per reviewer 2. Priority 2. Also consider connections to robust
-declassification. Priority 4.}
+In this case, {\tt a0} has changed during the call, {\em and}
+the return states do not agree on its value. We can therefore deduce that
+it carries some information derived from the original varied elements. By
+contrast, the sealed stack frame has not changed during the call, so the
+fact that it varies between the return states represents that the initial
+states disagreed, not a leak.
+Unless {\tt a0} happens to be irrelevant to the caller, this example
+is a violation of what we term {\it return-time confidentiality}
+(formalized in \cref{tab:props}, line 3b).
+\sna{Consider connections to robust declassification. Priority 4.}
 
@@
 
+\subsubsection*{Acknowledgements}
+
+We thank the reviewers for their comments, CHR Chhak and Allison Naaktgeboren for
+feedback during the writing process, and Aina Georges for her enthusiastic reception
+of our early work.
+
+This work was supported by the National Science Foundation under Grant No. 2048499, Specifying and Verifying Secure Compilation of C Code to Tagged Hardware.
+
 \bibliographystyle{IEEEtran}

### 2023-05-16 submission5/core.tex
Consolidate discussion on testing

@@
 micro-policy. We choose Coq for our implementation
-to ensure that our definitions are well-formed and to leverage the
+to ensure that our definitions are well-formed \rb{, to reason about them,} and to leverage the
 power of the QuickChick
@@
   an initial policy state, tagging instructions and the stack regions appropriately
-  for the micro-policy (see \cref{sec:enforcement})
+  for the micro-policy (see \cref{sec:enforcement}); and
 (3)
   a map from instructions to lists of their operations in the security semantics.
+%
+\rb{((say a few more words about 3?))}
+%
+\rb{
+That is, our generators need to produce not only a low-level program
+that exhibits interesting control-flow patterns, but also the policy
+meta-information and code annotations that would normally be produced
+by the compiler.
+}
 
@@
 programs that lead to longer executions---and hopefully towards more
-interesting behaviors as a result. We extend their technique with an additional
-degree of statefulness: when complex calling conventions need to be
+interesting behaviors as a result.
+%
+\rb{
+Generation by
+execution receives as an input a partially instantiated machine state
+and attempts to generate an instruction (or a sequence of instructions
+such as a blessed sequence) that makes sense locally (e.g., jumps go
+to a potentially valid code location and loads read from a
+potentially valid stack location). Then we step the machine and repeat
+the process until we generate or execute some target number of
+instructions, or reach a point where the machine cannot step
+any more.
+}
+
+We extend their technique with an additional
+degree of statefulness:
+\rb{
+we keep track of the control flow behavior of
+the program being generated. Each time a call or return sequence is
+generated, we ensure that the appropriate policy tags and code
+annotations are set for the entry or return points.
+}
+When complex calling conventions need to be
 observed to avoid early failstops, extra care must be taken to ensure
@@
 with an increased probability. However, at the same time, we need to
-allow for the possibility of {\em not} performing such an initialization
+\rb{allow the generation to sometimes relax those constraints,
+introducing potentially ill-formed flows:}
+% allow for the possibility of,
+for example, {\em not} performing such an initialization
+\rb{or \emph{not} returning to the proper call site, so as}
 to allow for potential attack vectors to manifest.
+\rb{
+This causes our programs
+to failstop when executed in conjunction with the policy monitor,
+but also allows for revealing errors in our setup.}
 
@@
 
+\rb{
+Naturally, testing also revealed a number of errors in our
+implementation of the enforcement mechanism (the original was written in C++
+and targeted ARM machine code;
+%\bcp{right?}\leo{yeah}
+we re-implemented it in Coq targeting RISC-V).  These errors range
+from trivial typos to ones that require an intriguingly complex setup
+to provoke.  The most interesting bug (included in the table as row
+{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
+sequence %/policy combination\apt{??}
+did not initialize all locations for the
+newly allocated stack frame correctly, but left some of them as
+$\tagNoDepth$. This allowed for a potential integrity violation, but
+only if a rather complicated sequence of events occured.
+The smallest counterexample requires calling a function {\tt f},
+which fails to initialize some of its frame during the blessed sequence,
+but writes into an uninitialized location $l$ later, treating \(l\) as outside
+the stack. Then {\tt f} calls a further function {\tt g} (which should have
+the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
+which is allowed because the enforcement mechanism still has
+$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
+%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
+%  and the write from {\tt f} was already a violation.}
+}
+
 For \(LTC\), the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
@@
 not really stack safe, and to validate a possible fix.
+\rb{
+((Cache pressure discussion relevant?))
+%% Unfortunately, Lazy Per-activation Tagging and Clearing
+%% puts severe pressure on the tag cache, so it may not be usable in practice.
+%% However, we conjecture that no lazy scheme can be fully stack safe
+%% without using unique per-activation identifiers.
+}
 

### 2023-05-16 submission5/core.tex
Acknowledgments plus

@@
 We thank the reviewers for their comments, CHR Chhak and Allison Naaktgeboren for
-feedback during the writing process, and Aina Georges for her enthusiastic reception
+feedback during the writing process, and A\"ina Linn Georges for her enthusiastic reception
 of our early work.
 
-This work was supported by the National Science Foundation under Grant No. 2048499, Specifying and Verifying Secure Compilation of C Code to Tagged Hardware.
+This work was supported by the National Science Foundation under Grant No. 2048499, Specifying and Verifying Secure Compilation of C Code to Tagged Hardware;
+and by Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972.
 

### 2023-05-17 submission5/core.tex
Priority twos done

@@
 \(\components\)-variants. We use this to define sets of irrelevant elements.
+Recall that \(\sim\) is a policy-specific compatibility relation.
 
@@
 written \(\mach \approx_\components \nach\), if \(\mach\sim\nach\) and
-\(\Delta(\mach,\nach) \subseteq \components\). \sna{define \(\sim\). Priority 2.}
+\(\Delta(\mach,\nach) \subseteq \components\).
 %if all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
@@
 testing framework~\cite{Pierce:SF4}.
-
-\sna{Discuss false positives, false negatives, and the degree to which
-  testing can increase our confidence. Priority 2.}
+%
+QuickChick is a randomized property-based testing framework. We use it to generate
+random programs, then execute them and test that they fulfill our criteria.
+
+Such testing is sound---it will not produce false positives---but necessarily
+necessarily incomplete. We might test a flawed policy but fail to find a
+program that exploits the flaw. Additionally, detecting violations of noninterference-style
+properties is dependent on choosing appropriate variant states, so it is possible
+to find a dangerous program but have it pass the test due to variant selection.
+We can increased our confidence in our test coverage by {\em mutation testing},
+in which we intentionally inject flaws into our policies and demonstrate that our testing
+can find them.
 
 \subsection{Test Generation}
-\sna{Restore Leo's old test gen discussion, then ping Leo if it seems that more
-  is needed. Priority 2.}
 
 To use QuickChick, we developed random test-case generators that produce
-(1) an initial RISC-V machine state, including the program to be executed;
-(2)
-  an initial policy state, tagging instructions and the stack regions appropriately
-  for the micro-policy (see \cref{sec:enforcement}); and
-(3)
-  a map from instructions to lists of their operations in the security semantics.
-%
-\rb{((say a few more words about 3?))}
-%
-\rb{
-That is, our generators need to produce not only a low-level program
-that exhibits interesting control-flow patterns, but also the policy
-meta-information and code annotations that would normally be produced
-by the compiler.
-}
+an initial RISC-V machine state tagged appropriately for the micro-policy
+(see \cref{sec:enforcement}), including a code region containing a
+low-level program. They must also produce the meta-information about
+how instructions in that program map to security-relevant operations,
+information that would normally be provided by the compiler.
 
@@
 interesting behaviors as a result.
-%
-\rb{
-Generation by
-execution receives as an input a partially instantiated machine state
-and attempts to generate an instruction (or a sequence of instructions
-such as a blessed sequence) that makes sense locally (e.g., jumps go
+
+Each step of generation by execution takes a partially instantiated
+machine state and attempts to generate an instruction
+that makes sense locally (e.g., jumps go
 to a potentially valid code location and loads read from a
-potentially valid stack location). Then we step the machine and repeat
-the process until we generate or execute some target number of
-instructions, or reach a point where the machine cannot step
-any more.
-}
-
-We extend their technique with an additional
-degree of statefulness:
-\rb{
-we keep track of the control flow behavior of
-the program being generated. Each time a call or return sequence is
-generated, we ensure that the appropriate policy tags and code
-annotations are set for the entry or return points.
-}
-When complex calling conventions need to be
-observed to avoid early failstops, extra care must be taken to ensure
-that % any such
-calls are well behaved. For example, immediately
-after a call we need to initialize any stack-allocated variables
-with an increased probability. However, at the same time, we need to
-\rb{allow the generation to sometimes relax those constraints,
-introducing potentially ill-formed flows:}
-% allow for the possibility of,
-for example, {\em not} performing such an initialization
-\rb{or \emph{not} returning to the proper call site, so as}
+potentially valid stack location). The generator repeats this process
+for an arbitrary number of steps, or until it reaches a point where
+the machine cannot step any more. Each time it generates a call or return,
+it places the appropriate policy tags on the relevant instruction(s)
+and records the operation.
+
+We extend Hrit\c{t}cu et al.'s technique with additional statefulness
+to avoid early failstops. For example, immediately
+after a call we initialize any stack-allocated variables
+with an increased probability. However, periodically, the generator
+will relax those constraints and generate potentially ill-formed
+code, such {\em not} performing such an initialization, writing outside
+of the current stack frame, or attempting an ill-formed return sequence,
 to allow for potential attack vectors to manifest.
-\rb{
-This causes our programs
-to failstop when executed in conjunction with the policy monitor,
-but also allows for revealing errors in our setup.}
 

### 2023-05-18 submission5/core.tex
Fix diagram

@@
       \! \underbrace{\stackrel{\stackrel{\SP}{\downarrow}}{\pubbox} \!\! \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
-      ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
+      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a4}}{\freebox\freebox}
+      ~ \stackrel{\mathtt{a5}}{\freebox\freebox}
       \)} \\
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
-      ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
+      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a4}}{\freebox\freebox}
+      ~ \stackrel{\mathtt{a5}}{\freebox\freebox}
       \)}
@@
       \! \underbrace{\pubbox \pubbox \pubbox \dots}_\public
-      ~ \stackrel{\mathtt{a0}}{\freebox\freebox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
-      ~ \stackrel{\mathtt{s0}}{\sealbox\sealbox}
+      ~ \stackrel{\mathtt{a0}}{\freebox\freebox} ~ \stackrel{\mathtt{a4}}{\freebox\freebox}
+      ~ \stackrel{\mathtt{a5}}{\freebox\freebox}
       \)}
@@
   \end{tabular}
-  \caption{Execution of example up through the return from {\tt f}.
-In stack diagrams, addresses increase to the right, stack grows to the left, and boxes represent 4-byte words.
-\sna{Reviewer 2 found s0 and a1 confusing, as well as lack of ra a4 and a5.
-Priority 3.}}
+  \caption{Execution of example up through the return from {\tt f}. In stack diagrams, addresses increase to the right, stack grows to the left, and boxes represent 4-byte words.}
 \label{fig:exec1}
@@
 registers are mapped to \(\unsealed\). In the current example, only register {\tt a0} has a change in
-security class. All callee-save registers remain \(\sealed\) for all calls.
-Thus, if we varied the assembly code for the example slightly so that {\tt sensitive} was stored
-in a callee-save register (e.g. {\tt s0}) rather than in memory, its security class would still be \(\sealed\)
-at the entry to {\tt f}.
+security class. All callee-save registers remain \(\sealed\) for all calls, so
+if we varied the assembly code for the example so that {\tt sensitive} was stored
+in a callee-save register (e.g. {\tt s0}) rather than in memory, its security class would still
+be \(\sealed\) at the entry to {\tt f}.
 %
@@
 ``call'' with ``call or tail call'' in \(\clrc\), \(\clec\), and \(\clei\).
+
+\subsection{Provenance, Capabilities, and Protecting Objects}
+\label{app:ptr}
+
+\sna{Cleanup. Priority 3.}
+
+What if we want to express a finer-grained notion of safety, in which
+stack objects are protected unless the function that owns them intentionally
+passes a pointer to them? This can be thought of as a {\it capability}-based
+notion of security. Capabilities are unforgeable tokens that grant access to
+a region of memory, typically corresponding to valid pointers to that region.
+So, in order to express such a property, we need our machine to carry some notion
+of {\it pointer provenance}---a distinction between a pointer that is intended to
+point to a given object, and non-pointer integers as well as pointers to other objects.
+This constitutes a substantial addition to the security context.
+
+One such provenance model is Memarian et al.'s PVI \cite{provenance}, in which pointers are
+annotated with the identity of the object they first pointed to. This annotation is
+propagated when the pointer is copied and when operations are performed on it---even
+integer-only operations.
+
+We can model this as a trio of additional security-relevant operations: one which
+declares a register to contain a valid pointer, one which transmits the provenance
+of a pointer from one element to another, and one which clears the provenance
+(for instance, when a pointer is modified in place in a way that makes it invalid).
+
+In addition to the normal call stack, our security context will carry a map \(\rho\) from
+elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
+Existing operations are extended to keep the value of \(\rho\) the same, and the new operations
+work as follows:
+
+\judgmentbr[Promote]
+           {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})\)}
+           {\(\rho' = \rho[\reg_{dst} \mapsto \mathit{range} ~ \reg_{base} ~ \mathit{off} ~ \mathit{sz}]\)}
+           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+
+\judgmentbr[Propagate]
+           {\(\psi = \mathbf{propagate} ~ \component_{src} ~ \component_{dst}\)}
+           {\(\rho' = \rho[\component_{dst} \mapsto \rho[\component_{src}]]\)}
+           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+
+\judgment[Clear]
+         {\(\psi = \mathbf{clear} ~ \component\)}
+         {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
+
+We now have a notion of provenance, and must integrate it into the definition of
+stack safety. We essentially generalize the above notion of passing: we will consider
+a caller to have intentionally passed an object if that object is reachable by
+a capability that has been passed to the callee. This includes capabilities passed
+indirectly, by being stored in an object that is in turn passed. Formally, we call
+this set \(\mathit{capped}\), and define it recursively:
+%
+\[\begin{split}
+& \mathit{capped} ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{\component\} \cup \mathit{capped} ~ \components' ~ \rho \textnormal{ where} \\
+& \components' = \{\component' |\rho[\component] = (\mathit{base},\mathit{bound})
+\land \mathit{base} \leq \component' < \mathit{bound}\} \\
+\end{split}\]
+
+We then tweak the call operation to seal only objects that are in \(\mathit{capped}\), or
+the previously defined \(\mathit{passed}\).
+
+\judgmentbrbrbr[]
+               {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
+               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
+                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
+               {\(\components = \{\addr | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \cup (\mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \}\)}
+               {\(V'' = V'\llbracket \components \mapsto \sealed \rrbracket\)}
+               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) =
+                 (V'',V::\sigma,\rho)\)}
+
+In the resulting property, once an object is sealed (because its
+capability has not been passed to a callee), subsequent nested calls can never unseal it.
+On the other hand, an object that is passed via a pointer may be passed on indefinitely.
 
@@
 It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
-%taking twice as long to catch as the next longest.
 taking longer to catch.
@@
 not really stack safe, and to validate a possible fix.
-\rb{
-((Cache pressure discussion relevant?))
+
+%% I don't think this discussion is relevant here
 %% Unfortunately, Lazy Per-activation Tagging and Clearing
@@
 %% without using unique per-activation identifiers.
-}
 
@@
 
-\appendix
-
-\subsection{Provenance, Capabilities, and Protecting Objects}
-\label{app:ptr}
-
-\sna{Move to section 5, cleanup. Priority 3.}
-
-What if we want to express a finer-grained notion of safety, in which
-stack objects are protected unless the function that owns them intentionally
-passes a pointer to them? This can be thought of as a {\it capability}-based
-notion of security. Capabilities are unforgeable tokens that grant access to
-a region of memory, typically corresponding to valid pointers to that region.
-So, in order to express such a property, we need our machine to carry some notion
-of {\it pointer provenance}---a distinction between a pointer that is intended to
-point to a given object, and non-pointer integers as well as pointers to other objects.
-This constitutes a substantial addition to the security context.
-
-One such provenance model is Memarian et al.'s PVI \cite{provenance}, in which pointers are
-annotated with the identity of the object they first pointed to. This annotation is
-propagated when the pointer is copied and when operations are performed on it---even
-integer-only operations.
-
-We can model this as a trio of additional security-relevant operations: one which
-declares a register to contain a valid pointer, one which transmits the provenance
-of a pointer from one element to another, and one which clears the provenance
-(for instance, when a pointer is modified in place in a way that makes it invalid).
-
-In addition to the normal call stack, our security context will carry a map \(\rho\) from
-elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
-Existing operations are extended to keep the value of \(\rho\) the same, and the new operations
-work as follows:
-
-\judgmentbr[Promote]
-           {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})\)}
-           {\(\rho' = \rho[\reg_{dst} \mapsto \mathit{range} ~ \reg_{base} ~ \mathit{off} ~ \mathit{sz}]\)}
-           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-
-\judgmentbr[Propagate]
-           {\(\psi = \mathbf{propagate} ~ \component_{src} ~ \component_{dst}\)}
-           {\(\rho' = \rho[\component_{dst} \mapsto \rho[\component_{src}]]\)}
-           {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-
-\judgment[Clear]
-         {\(\psi = \mathbf{clear} ~ \component\)}
-         {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
-
-We now have a notion of provenance, and must integrate it into the definition of
-stack safety. We essentially generalize the above notion of passing: we will consider
-a caller to have intentionally passed an object if that object is reachable by
-a capability that has been passed to the callee. This includes capabilities passed
-indirectly, by being stored in an object that is in turn passed. Formally, we call
-this set \(\mathit{capped}\), and define it recursively:
-%
-\[\begin{split}
-& \mathit{capped} ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{\component\} \cup \mathit{capped} ~ \components' ~ \rho \textnormal{ where} \\
-& \components' = \{\component' |\rho[\component] = (\mathit{base},\mathit{bound})
-\land \mathit{base} \leq \component' < \mathit{bound}\} \\
-\end{split}\]
-
-We then tweak the call operation to seal only objects that are in \(\mathit{capped}\), or
-the previously defined \(\mathit{passed}\).
-
-\judgmentbrbrbr[]
-               {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
-               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(\components = \{\addr | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \cup (\mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \}\)}
-               {\(V'' = V'\llbracket \components \mapsto \sealed \rrbracket\)}
-               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) =
-                 (V'',V::\sigma,\rho)\)}
-
-In the resulting property, once an object is sealed (because its
-capability has not been passed to a callee), subsequent nested calls can never unseal it.
-On the other hand, an object that is passed via a pointer may be passed on indefinitely.
-
 \end{document}

### 2023-05-18 submission5/core.tex
Introduced state elements earlier

@@
 Finally, we define high-level security properties---integrity, confidentiality,
-and well-bracketed control flow---in terms of the security classes of state
-elements. These take the form of predicates on states and contexts
-that use variants,
-assertions about future return states, and comparison of event traces to capture the
+and well-bracketed control flow---in terms of the security classes of {\em state elements}
+(memory addresses and registers). The properties are predicates use variants, assertions
+about future return states, and comparison of event traces to capture
 desired behavior without reference to internal implementation details.
+
 We introduce these properties by example in \cref{sec:example} and
@@
 the principals should occur only through the state elements that are designated for the
-purpose: those labeled \(\public\) and \(\object\). \sna{Define state elements before this. Priority 3.}
+purpose: those labeled \(\public\) and \(\object\).
 

### 2023-05-18 submission5/core.tex
Clean up in V.C

@@
 
-\sna{Cleanup. Priority 3.}
-
-What if we want to express a finer-grained notion of safety, in which
+Lastly, what if we want to express a finer-grained notion of safety, in which
 stack objects are protected unless the function that owns them intentionally
@@
 a region of memory, typically corresponding to valid pointers to that region.
-So, in order to express such a property, we need our machine to carry some notion
-of {\it pointer provenance}---a distinction between a pointer that is intended to
-point to a given object, and non-pointer integers as well as pointers to other objects.
-This constitutes a substantial addition to the security context.
-
-One such provenance model is Memarian et al.'s PVI \cite{provenance}, in which pointers are
-annotated with the identity of the object they first pointed to. This annotation is
-propagated when the pointer is copied and when operations are performed on it---even
-integer-only operations.
-
-We can model this as a trio of additional security-relevant operations: one which
+As such, this capability safety relies on some preexisting notion of pointer
+validity, i.e. {\it pointer provenance}.
+Memarian et al.'s PVI \cite{provenance} (provenance via integer)
+memory model is a good option: it annotates pointers with the identity of the
+object they first pointed to, and propagates the annotation
+pointer is copied and when operations are performed on it.
+This constitutes a substantial addition to the security context, which is why
+this enhancement is more speculative than the others, and we do not test it.
+
+We can model the provenance model as a trio of additional security-relevant operations: one which
 declares a register to contain a valid pointer, one which transmits the provenance
@@
 micro-policy. We choose Coq for our implementation
-to ensure that our definitions are well-formed \rb{, to reason about them,} and to leverage the
+to ensure that our definitions are well-formed, and to leverage the
 power of the QuickChick

### 2023-05-18 submission5/core.tex
Priority 3s done

@@
 Moreover, we need to further extend typical hyperproperty testing
-schemes to handle the nested nature of confidentiality: rather than
-just generating two variant initial machines
-and letting them execute to test for noninterference, we generate a
-new variant {\em every time a call is made} and check confidentiality
-for the subtrace produced from that variant state until its
-corresponding return. As a result, a ``single'' confidentiality test
-compactly checks multiple nested calls.
-\sna{Justify this. Priority 3.}
+schemes to handle the nested nature of confidentiality. In normal
+noninterference testing, we would generate a pair of initial
+variant states, execute them to completion, and compare the results.
+Instead, for our setup to test the confidentiality of every call,
+it would need to create a variant state at each call point, execute
+that until return, then generate a post-call variant based on any tainted
+values that would execute alongside the ``primary'' execution until
+completion. Tracking a number of variant executions that is linear
+in the total number of calls is rather inefficient! For performance,
+we instead track a single ``post-call taint'' execution. This machine
+executes in parallel with the original. Everytime a function returns
+with elements containing tainted data, those elements are varied in
+the tainted execution. So, at any given time, we need only simulate
+(1) the original execution, (2) the tainted execution, and (3) one
+variant execution for each call on the call stack. This approach
+makes testing longer runs substantially faster, at the cost of making it
+harder to identify which call is the source of a failure.
 
@@
 property (row {\em STORE\_NO\_CHECK}).
+
+In additional intentional mutations, our testing catches errors in our
+own implementation of the enforcement mechanism, including one interesting
+bug where the initial function's frame included space allocated for its
+return address, but this uninitialized (and therefore \(\tagNoDepth\)-tagged)
+space was treated as private data but left unprotected. We added this to
+our set of mutations as {\em HEADER\_NO\_INIT}.
+%(the original was written in C++
+%and targeted ARM machine code;
+%\bcp{right?}\leo{yeah}
+%we re-implemented it in Coq targeting RISC-V).  These errors range
+%from trivial typos to ones that require an intriguingly complex setup
+%to provoke.  The most interesting bug (included in the table as row
+%{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
+%sequence %/policy combination\apt{??}
+%did not initialize all locations for the
+%newly allocated stack frame correctly, but left some of them as
+%$\tagNoDepth$. This allowed for a potential integrity violation, but
+%only if a rather complicated sequence of events occured.
+%The smallest counterexample requires calling a function {\tt f},
+%which fails to initialize some of its frame during the blessed sequence,
+%but writes into an uninitialized location $l$ later, treating \(l\) as outside
+%the stack. Then {\tt f} calls a further function {\tt g} (which should have
+%the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
+%which is allowed because the enforcement mechanism still has
+%$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
+%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
+%  and the write from {\tt f} was already a violation.}
+%}
 
@@
 
-\rb{
-Naturally, testing also revealed a number of errors in our
-implementation of the enforcement mechanism (the original was written in C++
-and targeted ARM machine code;
-%\bcp{right?}\leo{yeah}
-we re-implemented it in Coq targeting RISC-V).  These errors range
-from trivial typos to ones that require an intriguingly complex setup
-to provoke.  The most interesting bug (included in the table as row
-{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
-sequence %/policy combination\apt{??}
-did not initialize all locations for the
-newly allocated stack frame correctly, but left some of them as
-$\tagNoDepth$. This allowed for a potential integrity violation, but
-only if a rather complicated sequence of events occured.
-The smallest counterexample requires calling a function {\tt f},
-which fails to initialize some of its frame during the blessed sequence,
-but writes into an uninitialized location $l$ later, treating \(l\) as outside
-the stack. Then {\tt f} calls a further function {\tt g} (which should have
-the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
-which is allowed because the enforcement mechanism still has
-$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
-%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
-%  and the write from {\tt f} was already a violation.}
-}
-
 For \(LTC\), the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},

### 2023-05-19 submission5/core.tex
Did a once-over

@@
 
-{\em Functions} in high-level languages, and related abstractions such
+Functions in high-level languages, and related abstractions such
 as procedures, methods, etc., are units of computation
@@
 Of the mechanisms listed above, many are fundamentally ill suited for
-offering formal guarantees: may impede attackers, but do not provide
+offering formal guarantees: they may impede attackers, but do not provide
 universal protection. Shadow stacks, for instance, aim to ``restrict
@@
 mechanisms, such as SoftBound~\cite{NagarakatteZMZ09} and code-pointer
-integrity~\cite{Kuznetsov+14}, do aim for stronger guarantees, but even
-these do not make formal promises.  Indeed, to our knowledge, the sole
+integrity~\cite{Kuznetsov+14}, do aim for stronger guarantees, but not
+typically formal ones.  To our knowledge, the sole
 line of work making a formal claim to protect stack safety is the
@@
 
-Other mechanisms besides these should also be amenable to strong
+Mechanisms besides this line of work should also be amenable to strong
 formal guarantees.  In particular, Roessler and DeHon
@@
 micro-policies~\cite{pump_oakland2015} for stack safety that offer
-strong guarantees.  Their most realistic micro-policy, called Lazy
+universal protection, but not yet tied to a particular formal guarantee.
+Their most realistic micro-policy, called Lazy
 Tagging and Clearing (\(LTC\)), makes an interesting performance
@@
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during the callee's execution, plus the
+and the callee's behavior during (and after) the callee's execution, plus the
 {\em well-bracketed control flow} (\(\wbcf\)) property articulated by
@@
 definitions of most source languages. It is implied by the semantics
-of features such as calls and return, but is not itself a feature.
+of features such as calls and returns, but is not itself a feature.
 \footnote{Contrast Azevedo de
@@
 %
-But neither is it satisfying as an entirely low-level property;
+But neither can stack safety be an entirely satisfying low-level property;
 indeed, at the lowest level, the specification of a ``well-behaved
@@
 to its callee. Those are the purview of high-level languages built
-atop the hardware stack\
+atop the hardware stack.
 
@@
 \Cref{sec:extensions} describes how to support an extended set of features.
+%\Cref{app:ptr} gives an (untested) proof-of-concept formalization of a more sophisticated,
+%heap-like sharing model based on pointer provenance.
 In \Cref{sec:enforcement} we describe the micro-policies that we test,
@@
 \cref{sec:relwork,sec:future} related and future work.
-\Cref{app:ptr} gives an (untested) proof-of-concept formalization of a more sophisticated,
-heap-like sharing model based on pointer provenance.
 
@@
 properties, plus our testing framework.  It does not include proofs:
-we use Coq primarily for the QuickChick testing library and for
-ensuring that our definitions are unambiguous.  Formal proofs are left
+we use Coq primarily for the QuickChick testing library and to
+ensure that our definitions are unambiguous.  Formal proofs are left
 as future work.
-
-%We consider a simple attacker, and discuss the harm that it can do to
-%its caller. This perspective is essential to understanding how lazy enforcement can
-%be safe: even if an attacker can overwrite some memory, if it cannot accomplish
-%subvert the observable behavior of the caller, the system is safe.
 
@@
 the standard execution semantics of the machine.  The security semantics identifies certain
-state transitions of the machine as {\em security-relevant
-  operations}, which update
+state transitions of the machine as {\em security-relevant operations}, which update
 a notional {\em security context}.  This context consists of
 an (abstract) stack of function activations, each associated with a {\em view}
-that maps each element of the machine state (memory location or register)
+that maps each {\em state element} of the machine (memory location or register)
 to a {\em security class} (active, sealed, etc.) specifying how the activation
@@
 
-We define stack safety properties in
-terms of traces of {\em observable events} rather than internal machine states.
-This is particularly important for
-specifying lazy enforcement, which catches bad behaviors after they occur but
-before they can affect the trace.
-Observable events might include specific function calls (e.g., system
+Given the security classes of the machine's state elements, we define high-level
+security properties---integrity, confidentiality, and well-bracketed control flow---
+as predicates that must hold on each call. These predicates draw on the idea of
+\emph{variant} states from the theory of non-interference, and a notion of
+{\em observable events}, which might include specific function calls (e.g., system
 calls that perform I/O), writes to special addresses representing
-memory-mapped regions, etc.
-
-To characterize situations in which values should be kept secret or prevented
-from influencing future execution, we borrow the idea of \emph{variant} states
-from the theory of non-interference.  For example, to show that certain locations
-are kept secret, it suffices to vary their contents and check that the observable
-output does not change.
-
-Finally, we define high-level security properties---integrity, confidentiality,
-and well-bracketed control flow---in terms of the security classes of {\em state elements}
-(memory addresses and registers). The properties are predicates use variants, assertions
-about future return states, and comparison of event traces to capture
-desired behavior without reference to internal implementation details.
+memory-mapped regions, etc. For example, to show that certain locations are kept
+secret, it suffices to compare the execution of states which vary at those locations
+and check that their traces of observable events are the same. This structure
+allows us to talk about the eventual implact of leaks or memory corruption without
+reference to internal implementation details, and to support lazy enforcement by
+delaying checks until a corrupted value would actually impact visible behavior.
 
@@
 convention or implements any particular source-language constructs.
-Indeed, if the source
-language is C, then high-level programs might contain undefined
-behavior, in which case they might be compiled to arbitrary machine
-code. A given enforcement
-mechanism or target architecture might place additional constraints, particularly on the behavior of
+Indeed, if the source language is C, then high-level programs might
+contain undefined behavior, in which case they might be compiled to
+arbitrary machine code. A given enforcement mechanism or target
+architecture might place additional constraints, particularly on the behavior of
 call and return sequences. For instance, extant implementations tend to assume
@@
 which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
-Roessler and DeHon include the compiler in their trusted computing
-base, so our properties
-are stronger than necessary to characterize their micro-policies.
 
@@
 state to vary. Formal definitions of variants and
-relevance that incorporate this wrinkle are given in \cref{sec:props}.}
+relevance that incorporate this wrinkle are given in \cref{sec:props}.
+\sna{Might need to make this concept a touch more explicit in \cref{sec:props}, or else
+promise a little less here.}
+}
 
@@
 (formalized in \cref{tab:props}, line 3b).
-\sna{Consider connections to robust declassification. Priority 4.}
+\sna{Consider connections to robust declassification. Low priority.}
 
@@
 We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
-and extend it with a runtime monitor implementing a stack safety
-micro-policy. We choose Coq for our implementation
-to ensure that our definitions are well-formed, and to leverage the
-power of the QuickChick
-testing framework~\cite{Pierce:SF4}.
-%
-QuickChick is a randomized property-based testing framework. We use it to generate
-random programs, then execute them and test that they fulfill our criteria.
+extend it with a runtime monitor implementing a stack safety micro-policy,
+and test it using QuickChick~\cite{Pierce:SF4}, a randomized property-based
+testing framework. QuickChick works by generating
+random programs, executing them, and checking that they fulfill our criteria.
 
@@
 to find a dangerous program but have it pass the test due to variant selection.
-We can increased our confidence in our test coverage by {\em mutation testing},
+We increase our confidence in our test coverage by {\em mutation testing},
 in which we intentionally inject flaws into our policies and demonstrate that our testing
@@
 
-To use QuickChick, we developed random test-case generators that produce
+To use QuickChick, we develope random test-case generators that produce
 an initial RISC-V machine state tagged appropriately for the micro-policy
 (see \cref{sec:enforcement}), including a code region containing a
-low-level program. They must also produce the meta-information about
+low-level program. They also produce the meta-information about
 how instructions in that program map to security-relevant operations,
@@
 
-To write such generators we build on the work of Hri\c{t}cu et
+Our generators build on the work of Hri\c{t}cu et
 al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
@@
 that makes sense locally (e.g., jumps go
-to a potentially valid code location and loads read from a
+to a potentially valid code location, loads read from a
 potentially valid stack location). The generator repeats this process
@@
 
-We extend Hrit\c{t}cu et al.'s technique with additional statefulness
-to avoid early failstops. For example, immediately
-after a call we initialize any stack-allocated variables
-with an increased probability. However, periodically, the generator
+We extend Hri\c{t}cu et al.'s technique with additional statefulness
+to avoid early failstops. For example, immediately after a call, we
+increase the probability of initializing any stack-allocated variables.
+However, periodically, the generator
 will relax those constraints and generate potentially ill-formed
-code, such {\em not} performing such an initialization, writing outside
+code, such as failing to initialize variables, writing outside
 of the current stack frame, or attempting an ill-formed return sequence,
@@
 
-Moreover, we need to further extend typical hyperproperty testing
-schemes to handle the nested nature of confidentiality. In normal
-noninterference testing, we would generate a pair of initial
-variant states, execute them to completion, and compare the results.
-Instead, for our setup to test the confidentiality of every call,
+\subsection{Property-based Testing}
+
+Once a test program is generated, QuickChick tests it against a
+property. A typical hyperproperty testing scheme might do so by
+generating a pair of initial variant states, executing them to completion,
+and comparing the results. We extend this procedure to handle the nested
+nature of confidentiality.
+
+For our setup to na\"{i}vely test the confidentiality of every call,
 it would need to create a variant state at each call point, execute
-that until return, then generate a post-call variant based on any tainted
-values that would execute alongside the ``primary'' execution until
-completion. Tracking a number of variant executions that is linear
-in the total number of calls is rather inefficient! For performance,
-we instead track a single ``post-call taint'' execution. This machine
-executes in parallel with the original. Everytime a function returns
-with elements containing tainted data, those elements are varied in
-the tainted execution. So, at any given time, we need only simulate
+it until return, then generate a post-call variant based on any tainted
+values. The post-call variant would execute alongside the ``primary''
+execution until the test is finished. This results in tracking a number
+of variant executions that is linear in the total number of calls!
+
+For performance, we instead maintain a single ``post-call taint''
+execution that executes in parallel with the original. Everytime a function returns
+with elements containing tainted data, the test process sets those elements
+to random values in the tainted execution. So, at any given time, we need only simulate
 (1) the original execution, (2) the tainted execution, and (3) one
 variant execution for each call on the call stack. This approach
-makes testing longer runs substantially faster, at the cost of making it
+makes testing longer executions substantially faster, at the cost of making it
 harder to identify which call is the source of a failure.
@@
 
-\sna{Look into wasm compilation per reviewer 1. Priority 4.}
+\sna{Look into wasm compilation per reviewer 1. Low priority.}
 

### 2023-05-23 submission5/core.tex
fixes and comments

@@
 
-The term {\em stack safety} is used for a variety of compiler,
+The term {\em stack safety} is used to describe a variety of compiler,
 run-time, and hardware mechanisms for protecting stack memory. Unlike
@@
 We propose a formal characterization of stack safety using concepts
-from language-based security. Rather than packaging all aspects of
-stack safety into a monolithic property, we decompose it into an
+from language-based security. Rather than treating 
+stack safety as a monolithic property, we decompose it into an
 integrity property and a confidentiality property for each of the
@@
 functions to write into one another's frames, but which taint the
-changed locations so that the frame's owner cannot access it. No
+changed locations so that the frame's owner cannot access them. No
 existing characterization of stack safety captures this style of
@@
 as procedures, methods, etc., are units of computation
-that call one another to modularly realize larger computations.
+that call one another to define larger computations in a modular way.
 %
 At a low level, each function activation manages its own data---local
-variables, etc.---as well as information about the \emph{caller} to
+variables, spilled temporaries, etc.---as well as information about the \emph{caller} to
 which it must return.
@@
 whack-a-mole, we seek formal properties of safe behavior that can be
-proven or at least rigorously tested. Such properties can be used as
+proven, or at least rigorously tested. Such properties can be used as
 the specification against which enforcement can be validated; even
@@
 
-Of the mechanisms listed above, many are fundamentally ill suited for
+Of the mechanisms listed above, many are fundamentally ill-suited for
 offering formal guarantees: they may impede attackers, but do not provide
@@
 line of work making a formal claim to protect stack safety is the
-secure calling conventions of Skorstengaard et
+study of secure calling conventions by Skorstengaard et
 al.~\cite{SkorstengaardSTKJFP} and Georges et
@@
 formal guarantees.  In particular, Roessler and DeHon
-\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
+\cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
 micro-policies~\cite{pump_oakland2015} for stack safety that offer
@@
 Tagging and Clearing (\(LTC\)), makes an interesting performance
-trade-off: it allows function activations to improperly write into one
+trade-off: it allows function activations to write improperly into one
 another's stack frames, but ensures that the owner of the corrupted
@@
 While our properties are motivated by the desire to specify \(LTC\) precisely,
-they are stated abstractly in hopes of applying
+they are stated abstractly in the hope that they can also be applied
 to other enforcement mechanisms.
@@
 definitions of most source languages. It is implied by the semantics
-of features such as calls and returns, but is not itself a feature.
-\footnote{Contrast Azevedo de
+of features such as calls and returns, but is not itself a feature.\footnote{
+Contrast Azevedo de
   Amorim et al.'s work on heap safety
@@
 
-We show that our properties are useful for distinguishing correct enforcement
-from incorrect using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
+We show that our properties are useful for distinguishing between correct and
+incorrect enforcement using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
 random testing tool for Coq.
@@
 assumptions. In \cref{sec:example}, we walk through a function call in
-a simple example machine and discuss how each of our properties
-applies to it, informally. In the process we motivate the properties
+a simple example machine and discuss informally how each of our properties
+applies to it. In the process we motivate the properties
 from a security perspective.
@@
 
-The accompanying artifact contains formal definitions (in Coq) of our
+The accompanying artifact \apt{what is the state of that? Depending on conference preferences,
+should give a URL or other pointer.} contains formal definitions (in Coq) of our
 properties, plus our testing framework.  It does not include proofs:
@@
 an (abstract) stack of function activations, each associated with a {\em view}
-that maps each {\em state element} of the machine (memory location or register)
+that maps each machine {\em state element} (memory location or register)
 to a {\em security class} (active, sealed, etc.) specifying how the activation
@@
 The action of a security-relevant operation on the context is defined by a set
-of rules that describe how the machine code makes use of the stack and registers
-to implement the function abstraction.
+of rules that describe how the machine code implements the function abstractoin
+using the stack and registers.
 
 Given the security classes of the machine's state elements, we define high-level
-security properties---integrity, confidentiality, and well-bracketed control flow---
-as predicates that must hold on each call. These predicates draw on the idea of
+security properties---integrity, confidentiality, and well-bracketed control flow---as
+predicates that must hold on each call. These predicates draw on the idea of
 \emph{variant} states from the theory of non-interference, and a notion of
@@
 reference to internal implementation details, and to support lazy enforcement by
-delaying checks until a corrupted value would actually impact visible behavior.
+flagging corruption of values only when it can actually impact visible behavior.
 
@@
   \obs} \mach'\), where \(\mach\) and \(\mach\) are machine states, \(\obs\) is
-an event, and \(\bar{\psi}\) is a list of security-relevant operations.
+an (optional) external event, and \(\bar{\psi}\) is a list of security-relevant operations.
 %
@@
 implicitly that callee-saved registers have their values maintained by whichever compiler
-generated their code. Our properties explicitly state this as a requirement,
+generated their code. \apt{I do not understand the last sentence. Implementations of what?}
+Our properties explicitly state this as a requirement,
 which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
@@
       \labeledrow{40:}{sw a0,out}{}
-      \labeledrow{44:}{j L2:}{}
+      \labeledrow{44:}{j L2}{}
       \labeledrow{L1, 48:}{lw a0,0(sp)}{}
@@
 
-  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; the stack grows to the left.}
+  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; the stack grows to the left.\apt{In this and later figures, can we lighten the blue backgrounds?}}
   \label{fig:main}
@@
 by returning to the correct program counter but with the wrong stack pointer.%
-\footnote{We pad the last two variants with {\tt nop}s just so that all the
+\footnote{We pad some of the variants with {\tt nop}s just so that all the
 snippets have the same length, which keeps the step numbering uniform in~\cref{fig:exec1}.}
@@
       \labeledrow{100:}{lw a4,8(sp)}{}
-      \labeledrow{104:}{mov a4,a0}{}
-      \labeledrow{108:}{sw zero,-4(sp)}{}
+      \labeledrow{104:}{mov a0,a4}{}
+      \labeledrow{108:}{nop}{}
       \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
@@
 on the security-relevant events noted in the right columns of \cref{fig:main,fig:f},
-namely execution of instructions that allocate or deallocate space,
-make a call, or make a return.
+namely execution of instructions that allocate or deallocate space (specified by
+an \(\SP\)-relative offset and size), make a call (with a specified list of argument registers),
+or make a return.
 
@@
 State transitions are numbered and labeled with a list of security operations, written
-\(\downarrow \overline{\psi}\) between steps.
+\(\downarrow \overline{\psi}\) between steps. We write \(\emplist\) for empty lists.
 
@@
 At step 5, the current principal's record is pushed onto the inactive list.
-Its return target is the return address of the call,
-and the stack pointer target is the stack pointer at the moment of call.
+% APT: I think this is dead:
+%Its return target is the return address of the call,
+%and the stack pointer target is the stack pointer at the moment of call.
 The callee's view is updated from the caller's such that all \(\object\) memory locations
@@
 security class. All callee-save registers remain \(\sealed\) for all calls, so
-if we varied the assembly code for the example so that {\tt sensitive} was stored
+if, in the example, we varied the assembly code for {\tt main} so that {\tt sensitive} were stored
 in a callee-save register (e.g. {\tt s0}) rather than in memory, its security class would still
@@
 and the stack pointer should be the same as it was before the call.
-Both of these are essential. In \cref{subfig:WBCF}, the attacker adds
-16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
+Both of these are essential for security. In \cref{subfig:WBCF}, the attacker adds
+16 to the return address and then returns; this bypasses the {\tt if}-test in the code and outputs
 {\tt secret}.
@@
 \(\wbcf\) requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
+(This property is formalized in \cref{tab:props}, line 1).
 %For nested calls, where the pending stack is initially larger, the same principle
@@
 Like \(\wbcf\), stack integrity defines a condition at the call that must hold upon
-return. This time the condition applies to all of the memory in the function's
+return. This time the condition applies to all of the memory in the caller's
 frame. In \cref{fig:exec1} we see the lifecycle of an allocated frame:
@@
 to its original value at the return point. 
+(This property is formalized in \cref{tab:props}, line 2).
 
@@
   \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20pt}{white}{1}}\]
-  \caption{Integrity Violation}
+  \caption{Integrity Violation\apt{Unclear how these states are related. Extend caption?  And what do the colors means?}}
   \label{fig:variant}
@@
 which is \(\sealed\). \Cref{fig:variant} shows the state just after the call at step 5,
-assuming that {\tt sec} is 5. Similar to \(\wbcf\), we define
+assuming that {\tt sec} is 5. \apt{But the figure seems to show several states. Unclear!}
+Similar to \(\wbcf\), we define
 \(\intProp\) as a predicate on states that holds if
@@
 which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
+\apt{Rest of this paragraph is still not very clear, especially in (non-?)relation to the figure.}
 Consider a variant state in which {\tt sensitive} has any other value, % arbitrarily choosing
@@
 
-Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\), with the attacker from
-\cref{subfig:direct}. We take a variant state over the set of elements that are
+Consider the state after step 5,
+% APT: trying to remove mysterious state notation:
+% Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\),
+5with the attacker code from \cref{subfig:direct}. We take a variant state over the set of elements that are
 \(\sealed\) in \(V_2\) (see \cref{fig:variant2}).
@@
 the traces may differ, in this case outputting 5 (the original value of {\tt secret}) and
-4 (its value in the variant) respectively. This is a violation of
+4 \apt{3???} (its value in the variant) respectively. This is a violation of
 {\it internal confidentiality} (formalized in \cref{tab:props}, line 3a).
@@
 states disagreed, not a leak.
+\apt{Last sentence still quite mysterious.}
 Unless {\tt a0} happens to be irrelevant to the caller, this example
@@
 given the special name \(\SP\);
-some registers may be classified as caller-saved or callee-saved.
+some registers may be classified as caller-saved (CLR) or callee-saved (CLE).
 Along with the program counter, \(\PCname\), these are referred to as
@@
 %the behavior of the basic machine, and it is the part that should be changed
-When constructing variants (see below), this additional information should
+When constructing variants (see~\cref{sec:props}, this additional information should
 not be varied. To capture this idea, we assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
@@
 For any security class \(l\), we write \(l(V)\)
-to denote the set of elements \(\component\) such that \(V ~ \component = l\).
+to denote the set of elements \(\component\) such that \(V ~ \component = l\).\apt{Did you ever use this?
+  I've edited some places to do do.}
 The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
@@
 
-  \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
-    \{\addr \mid V ~ \addr = \unsealed\}\)}
-             {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
+  \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
+             {\(V' = V \llbracket \addr \mapsto \object \mid \addr \in \components \rrbracket\)}
              {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
     %
-    \judgmentbr{\(b = \mach[\SP] + \mathit{off}\)}
-               {\(V' = V \llbracket \addr \mapsto \unsealed |
-                 b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
+    \judgmentbr{\(\components = \mathit{range}~ \SP ~ \mathit{off} ~ \mathit {sz} ~ \mach \cap \object(V)\)}
+               {\(V' = V \llbracket \addr \mapsto \unsealed \mid \addr \in \components \rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
@@
                  \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \rrbracket\)}
+              {\(V'' = V' \llbracket \addr \mapsto \sealed | \addr \in \object(V')\rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}})
@@
     %
-    \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma')\)}
-  \caption{Basic Operations}
+    \judgment{}
+             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, (V,\sigma')) = (V, \sigma')\)}
+             \caption{Basic Operations\apt{This was absolutely full of problems. Please triple-check my changes.
+             And do we really need two-stage calculation in the call rule, or was that just to avoid type-setting issues?}}
   \label{fig:basicops}
@@
 
-\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{\obs} (\mach',\context')\)}
+\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{\overline{\psi},\obs} (\mach',\context')\)}
               {\(|\sigma| \geq d\)}
@@
 \noindent
-When \(d = 0\), the trace will always be infinite; in this case we
+When \(d = 0\), the trace will always be infinite\apt{why? remind us about machine not halting (if that is the reason)?}; in this case we
 omit \(d\) and just write \(\hookrightarrow s\).
@@
 any trace. So, a trace that silently diverges due to a failstop will
-fulfill this property vacuously.
+fulfill this property \apt{what property?} vacuously.
 
@@
 
-Similarly, we give a binary equivalent for use in confidentiality. We define \(\Uparrow\) so that
+Similarly, we give a analogous binary relation for use in confidentiality. We define \(\Uparrow\) so that
 \((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
@@
     \rowcolor{black!20}
-    & & & \(\textnormal{ and } \mach''[\PCname] = \mach[\PCname]+4\) & \\
+    & & & \(\textnormal{ \hspace{0.8in}} \land \mach''[\PCname] = \mach[\PCname]+4\) & \\
     %
@@
   \end{tabular}
-  \caption{Properties}
+  \caption{Properties\apt{The hardwired +4 is a little sad, but oh well.}}
   \label{tab:props}
@@
 \paragraph*{5.~\(\clei\)}
-Callee integrity means that the caller does not influence it outside of legitimate
+Callee integrity means that the caller does not influence the callee outside of legitimate
 channels. The caller's influence can be seen internally, or in corrupted data on return,
@@
 \((\mach,(V,\sigma))\), we take an arbitrary \(\nach\) that is a \(\components\)-variant
-of \(\mach\), where \(\components\) is the set elements that are not \(\object\)
+of \(\mach\), where \(\components\) is the set of elements that are not \(\object\)
 or \(\public\). The remainder of the property is identical to \(\clrc\).
@@
 because they are composite types that inherently have
-pass-by-reference semantics. The caller may also pass an object by reference
+pass-by-reference semantics. The caller may also pass a stack-allocated  object by reference
 in the C++ style, or take its address and pass it as a pointer.
@@
 We refine our call operation to make use of the information that we have about
-which memory contain arguments, \(\overline{sa}\). \(\overline{sa}\) is a set of
+which stack memory locations contain arguments. The new annotation \(\overline{sa}\) is a set of
 triples of a register, an offset from the value of that register, and a size.
@@
 is instead specified by \((\reg, 0, \mathit{sz})\), which allows
-the reference to be passed further down the stack.
-
+the reference to be passed further down the stack.\apt{Last sentence is completely mysterious.}
+
+\apt{This needs light revision to reflect fact that Capabilities section is now on same page, not much later!}
 If the address of an object is taken directly and passed as a
@@
   \end{subfigure}
-  \caption{Operations supporting tail calls and argument passing on stack.}
+  \caption{Operations supporting tail calls and argument passing on stack.\apt{This needs similar edits to the previous operations figure.
+In particular, same question about the $V''$ in call and tailcall.}}
   \label{fig:advops}
@@
 it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
-than \(\sealed\) to express this. In terms of \cref{tab:props}, we replace
+than \(\sealed\) to express this. In \cref{tab:props}, we replace
 % RB: could shorten "in terms of"
-``call'' with ``call or tail call'' in \(\clrc\), \(\clec\), and \(\clei\).
+``call targets'' with ``call or tail call targets'' in \(\clrc\), \(\clec\), and \(\clei\).
 
@@
 memory model is a good option: it annotates pointers with the identity of the
-object they first pointed to, and propagates the annotation
+object they first pointed to, and propagates the annotation when the 
 pointer is copied and when operations are performed on it.
 This constitutes a substantial addition to the security context, which is why
-this enhancement is more speculative than the others, and we do not test it.
+this enhancement is more speculative than the others, and we have not tested it.
 
@@
 elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
-Existing operations are extended to keep the value of \(\rho\) the same, and the new operations
-work as follows:
+Existing operations are extended to preserve the value of \(\rho\), and the new operations
+work as follows\apt{Put these and tweaked call rule into a separate figure in the same style as the existing two}:
 
@@
 
-In \(LTC\), rather than initialize the entire frame at entry,
-a function initializes each location's tag when it writes to it. It does
+In \(LTC\), a function neither initializes the frame at entry nor clears it at exist;
+instead, it simply sets each location's tag to the PC tag when that location is written. It does
 not check if those writes are legal! If the PC tag is \(\tagPCDepth{n}\),
@@
 We propose instead tracking every activation uniquely, which should eliminate the
-issue entirely, as it does in our tests.
+issue entirely---and does in our tests.
 
@@
 registers to be saved and restored properly. We extend \(DI\) and \(LTC\)
-so that callee-saved registers, too, are tagged with the color of the
+so that callee-saved registers are also tagged with the color of the
 function that is using them. In \(DI\) they are tagged as part of the entry
@@
 Our primary testing targets are the eager {\em Depth Isolation}
-and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
+and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.\apt{Reads
+  awkwardly coming just after last section.}
 
@@
 
-Such testing is sound---it will not produce false positives---but necessarily
-necessarily incomplete. We might test a flawed policy but fail to find a
+Such testing is sound---it will not produce false positives---but 
+necessarily incomplete. We might test a flawed policy but fail to generate a
 program that exploits the flaw. Additionally, detecting violations of noninterference-style
 properties is dependent on choosing appropriate variant states, so it is possible
-to find a dangerous program but have it pass the test due to variant selection.
+to generate a dangerous program but have it pass the test due to variant selection.
 We increase our confidence in our test coverage by {\em mutation testing},
-in which we intentionally inject flaws into our policies and demonstrate that our testing
+in which we intentionally inject flaws into the policies and demonstrate that testing
 can find them.
@@
 
-To use QuickChick, we develope random test-case generators that produce
+To use QuickChick, we develop random test-case generators that produce
 an initial RISC-V machine state tagged appropriately for the micro-policy
@@
 how instructions in that program map to security-relevant operations,
-information that would normally be provided by the compiler.
+which would normally be provided by the compiler.
 
@@
 interesting behaviors as a result.
-
+%
 Each step of generation by execution takes a partially instantiated
@@
 to avoid early failstops. For example, immediately after a call, we
-increase the probability of initializing any stack-allocated variables.
-However, periodically, the generator
-will relax those constraints and generate potentially ill-formed
+increase the probability of generating code that initializes any stack-allocated variables.
+To allow for potential attack vectors to manifest,
+the generator periodically relaxes those constraints and generates potentially ill-formed
 code, such as failing to initialize variables, writing outside
 of the current stack frame, or attempting an ill-formed return sequence,
-to allow for potential attack vectors to manifest.
 
@@
 Once a test program is generated, QuickChick tests it against a
-property. A typical hyperproperty testing scheme might do so by
+property. A typical hyperproperty testing scheme might do this by
 generating a pair of initial variant states, executing them to completion,
@@
 
-For performance, we instead maintain a single ``post-call taint''
+For better performance, we instead maintain a single ``post-call taint''
 execution that executes in parallel with the original. Everytime a function returns
-with elements containing tainted data, the test process sets those elements
+with elements containing tainted data\apt{I don't understand what that is}, the test process sets those elements
 to random values in the tainted execution. So, at any given time, we need only simulate
@@
 current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
-that restriction by omitting the check (row {\em
+that restriction by omitting the check (bug {\em
   LOAD\_NO\_CHECK\_DI}).
@@
 tag. Relaxing that constraint causes violations to the integrity
-property (row {\em STORE\_NO\_CHECK}).
+property (bug {\em STORE\_NO\_CHECK}).
 
@@
 \vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
+\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms\apt{what does the horizontal line mean?}}
 \vspace*{-2em}
@@
 Our properties allow us to identify an enforcement mechanism as
-not really stack safe, and to validate a possible fix.
+not really stack safe, and to validate a possible fix.\apt{This paragraph
+  comes as a bit of a non-sequitur here.}
 
@@
 
-The centrality of the functional abstraction and its security are behind the
+The centrality of the function abstraction and its security are behind the
 many software and hardware mechanisms proposed for its protection
@@
 %
-These works define stack safety as overlay semantics and related stack
+The authors define stack safety as overlay semantics and related stack
 safety properties, phrased in terms of logical relations instead of trace
@@
 
-In terms of direct feature comparison with Georges et al. (being the most
+In terms of direct feature comparison with Georges et al.~\cite{Georges22:TempsDesCerises} (the most
 recent work in the line), with the addition of confidentiality to their formalism, we
@@
 
-It would also be interesting to test a software approach.
+It would also be interesting to test a software enforcement approach.
 Under a bounds checking discipline~%such as SoftBound
@@
 feedback during the writing process, and A\"ina Linn Georges for her enthusiastic reception
-of our early work.
+of our early work.\apt{would consider rephrasing that last.}
 
 This work was supported by the National Science Foundation under Grant No. 2048499, Specifying and Verifying Secure Compilation of C Code to Tagged Hardware;
-and by Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972.
+and by Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972.\apt{Benjamin?}{Leo?}
 

### 2023-05-23 submission5/core.tex
Responded to Andrew's notes

@@
 The accompanying artifact \apt{what is the state of that? Depending on conference preferences,
-should give a URL or other pointer.} contains formal definitions (in Coq) of our
+  should give a URL or other pointer.} \sna{Git repo is currently private, I think. I didn't see much
+  in the way of conference preferences on a quick glance, but I'll look again in the morning.}
+contains formal definitions (in Coq) of our
 properties, plus our testing framework.  It does not include proofs:
@@
 arbitrary machine code. A given enforcement mechanism or target
-architecture might place additional constraints, particularly on the behavior of
-call and return sequences. For instance, extant implementations tend to assume
-implicitly that callee-saved registers have their values maintained by whichever compiler
-generated their code. \apt{I do not understand the last sentence. Implementations of what?}
-Our properties explicitly state this as a requirement,
-which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
+architecture might place additional constraints, particularly on
+call and return sequences.
+
+%tend to assume
+%implicitly that callee-saved registers have their values maintained by whichever compiler
+%generated their code. \apt{I do not understand the last sentence. Implementations of what?}
+%Our properties explicitly state this as a requirement,
+%which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
 
@@
 \begin{figure}
-  \newcommand{\figonebox}[1][]{\genbox{30pt}{lightgray}{#1}}
+  \newcommand{\figonebox}[1][]{\genbox{30pt}{blue!20}{#1}}
   \begin{subfigure}{\columnwidth}
@@
 
-  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; the stack grows to the left.\apt{In this and later figures, can we lighten the blue backgrounds?}}
+  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; the stack grows to the left.}
   \label{fig:main}
@@
 At step 5, the current principal's record is pushed onto the inactive list.
-% APT: I think this is dead:
-%Its return target is the return address of the call,
-%and the stack pointer target is the stack pointer at the moment of call.
 The callee's view is updated from the caller's such that all \(\object\) memory locations
@@
 additional state associated with elements, such as tags, we don't want that
-state to vary. Formal definitions of variants and
-relevance that incorporate this wrinkle are given in \cref{sec:props}.
-\sna{Might need to make this concept a touch more explicit in \cref{sec:props}, or else
-promise a little less here.}
+state to vary. This is touched on in \cref{sec:props}.
 }
@@
   \stackrel{\hookrightarrow \mathtt{out}}{\genbox{20pt}{white}{1}}\]
-  \caption{Integrity Violation\apt{Unclear how these states are related. Extend caption?  And what do the colors means?}}
+  \caption{Integrity Violation: {\tt sensitive} changed, and if varied, changes future outputs}
   \label{fig:variant}
@@
 In our example setting, the observation trace consists of the sequence
-of values written to {\tt out}.
-The example in \cref{subfig:integrity} modifies the value of {\tt sensitive},
-which is \(\sealed\). \Cref{fig:variant} shows the state just after the call at step 5,
-assuming that {\tt sec} is 5. \apt{But the figure seems to show several states. Unclear!}
-Similar to \(\wbcf\), we define
+of values written to {\tt out}. In \cref{subfig:integrity} the states before
+and after the call differ in the value of {\tt sensitive}. \Cref{fig:variant}
+shows the states before and after the call, which disagree on the value at
+{\tt sensitive}. If we consider a
+variant of the original return state in which {\tt sensitive} is 0 (orange)
+as opposed to 42 (blue), that state will eventually output 1, while the actual
+execution outputs 5. This means that {\tt sensitive} is relevant.
+
+To be more explicit, similar to \(\wbcf\), we define
 \(\intProp\) as a predicate on states that holds if
@@
 We require that \(\intProp\) hold on the state following the matching return,
-which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
-\apt{Rest of this paragraph is still not very clear, especially in (non-?)relation to the figure.}
-Consider a variant state in which {\tt sensitive} has any other value, % arbitrarily choosing
-say 43. As execution continues after the return
-from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
-from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} \emph{is} relevant,
-so \(\intProp\) does not hold, and integrity has indeed been violated.
+which is reached by step 9. Here {\tt sensitive} has obviously changed, but we
+just saw that it is relevant.
 
@@
 
-Consider the state after step 5,
-% APT: trying to remove mysterious state notation:
-% Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\),
-5with the attacker code from \cref{subfig:direct}. We take a variant state over the set of elements that are
-\(\sealed\) in \(V_2\) (see \cref{fig:variant2}).
-If we take a trace of execution from each state until it returns,
-the traces may differ, in this case outputting 5 (the original value of {\tt secret}) and
-4 \apt{3???} (its value in the variant) respectively. This is a violation of
+Consider the state after step 5, shown at the top of \cref{fig:variant2},
+with the attacker code from \cref{subfig:direct} and the assumption that
+{\tt secret} has the value 5. We take a variant state over
+the set of elements that are \(\sealed\) in \(V_2\) (orange), and compare it to the
+original (blue). During the execution, the value of {\tt secret} is written
+to the output, and the information leak is evidenced by the fact that the
+outputs do not agree---the original outputs 5, while the variant outputs 3.
+This is a violation of
 {\it internal confidentiality} (formalized in \cref{tab:props}, line 3a).
@@
 \end{figure}
-%
-In this case, {\tt a0} has changed during the call, {\em and}
-the return states do not agree on its value. We can therefore deduce that
-it carries some information derived from the original varied elements. By
-contrast, the sealed stack frame has not changed during the call, so the
-fact that it varies between the return states represents that the initial
-states disagreed, not a leak.
-\apt{Last sentence still quite mysterious.}
+
+At the end of the call, we can deduce that every element on which
+the variant states disagree must carry some information derived from
+the original varied elements. In most cases, that is because the element
+is one of the original varied elements and has not changed during the
+call, which does not represent a leak. But in the case of {\tt a0}, it
+has changed during the call, {\em and} the return states do not agree
+on its value. This represents data that has been leaked, and should
+not be used to affect future execution.
 Unless {\tt a0} happens to be irrelevant to the caller, this example
@@
 (formalized in \cref{tab:props}, line 3b).
-\sna{Consider connections to robust declassification. Low priority.}
 
@@
 For any security class \(l\), we write \(l(V)\)
-to denote the set of elements \(\component\) such that \(V ~ \component = l\).\apt{Did you ever use this?
-  I've edited some places to do do.}
+to denote the set of elements \(\component\) such that \(V ~ \component = l\).
 The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
@@
     %
-    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-              {\(V'' = V' \llbracket \addr \mapsto \sealed | \addr \in \object(V')\rrbracket\)}
+    \judgment{\(V' = V \left\llbracket
+      k \mapsto
+      \begin{cases}
+        \unsealed & \textnormal{if } k \in CLR \\
+        \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
+        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } k \in \object(V') \\
+        V(k) & \textnormal{otherwise} \\
+      \end{cases}\right\rrbracket\)}
              {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}})
-                 ~ (V,\sigma) = (V'',V::\sigma)\)}
+               ~ (V,\sigma) = (V',V::\sigma)\)}
     %
@@
              {\(Op ~ \mach ~ \mathbf{return} ~ (\_, (V,\sigma')) = (V, \sigma')\)}
-             \caption{Basic Operations\apt{This was absolutely full of problems. Please triple-check my changes.
-             And do we really need two-stage calculation in the call rule, or was that just to avoid type-setting issues?}}
+             \caption{Basic Operations\apt{...Do we really need two-stage calculation in the call rule, or was that just to avoid type-setting issues?}\sna{The latter.
+I took a stab at trying to avoid it, but I'm ambivalent about  that ``cases'' style here.}}
   \label{fig:basicops}
@@
 \noindent
-When \(d = 0\), the trace will always be infinite\apt{why? remind us about machine not halting (if that is the reason)?}; in this case we
+When \(d = 0\), the trace will always be infinite because the machine never halts; in this case we
 omit \(d\) and just write \(\hookrightarrow s\).
@@
 any trace. So, a trace that silently diverges due to a failstop will
-fulfill this property \apt{what property?} vacuously.
+be vacuously similar to all other traces.
 
@@
     \rowcolor{black!20}
-    & & & \(\textnormal{ \hspace{0.8in}} \land \mach''[\PCname] = \mach[\PCname]+4\) & \\
+    & & & \(\textnormal{ \hspace{0.8in}} \land \mach''[\PCname] = \mach[\PCname]+\mathit{sz}\) & where \(\mathit{sz}\) is the size of one instruction \\
     %
@@
   \end{tabular}
-  \caption{Properties\apt{The hardwired +4 is a little sad, but oh well.}}
+  \caption{Properties}
   \label{tab:props}
@@
 In this case, only the immediate callee gains access to the argument location.
-A C++-style call-by-reference argument % where the reference is
-passed in \(\reg\)
-is instead specified by \((\reg, 0, \mathit{sz})\), which allows
-the reference to be passed further down the stack.\apt{Last sentence is completely mysterious.}
-
-\apt{This needs light revision to reflect fact that Capabilities section is now on same page, not much later!}
-If the address of an object is taken directly and passed as a
-pointer, we simply classify the object as ``public'' and give it no protection against
-access by other functions. We extend the \(\mathbf{alloc}\) operation
-with a boolean flag, where {\bf t} indicates
+A C++-style call-by-reference argument where the reference is passed in \(\reg\)
+is instead specified by the triple \((\reg, 0, \mathit{sz})\). Such a call-by-reference
+argument could be passed through multiple calls, provided that it is in \(\overline{sa}\)
+each time.
+
+Absent the more sophisticated capability model (below), if the address of an object
+is taken directly and passed as a pointer, we simply classify the object as ``public''
+and give it no protection against access by other functions.
+We extend the \(\mathbf{alloc}\) operation with a boolean flag, where {\bf t} indicates
 that the allocation is public, and {\bf f} that it is private.
@@
 never sealed at a call (\cref{sfig:publicalloc}).
-Providing more fine-grained control over sharing is desirable, but
-seems to require a considerably more complex model; \cref{app:ptr} describes one
-approach, based on capabilities.
+Providing more fine-grained control over sharing is desirable, but requires a considerably
+more complex model. This simple model is included in our testing; we describe
+an untested approach based on capabilities below.
 
@@
 
-    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
-                 \{\addr \mid V ~ \addr = \unsealed\}\)}
+    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
                {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
@@
     \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
-                 \{\addr \mid V ~ \addr = \unsealed\}\)}
+                 \unsealed(V)\)}
                {\(V' = V \llbracket \addr \mapsto \public \mid \addr \in \components \rrbracket\)}
@@
                {\(V' = V \llbracket \addr \mapsto \unsealed |
-                 b \leq a < b+\mathit{sz} \land V ~ \addr = \object \rrbracket\)}
+                 b \leq a < b+\mathit{sz} \land \addr \in \object(V) \rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
@@
     %
-    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+    \judgment{\(V' = V \left\llbracket
+      k \mapsto
+      \begin{cases}
+        \unsealed & \textnormal{if } k \in CLR \\
+        \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
+        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } \\
+        & k \in \object(V) - \mathit{passed} ~ \overline{sa} ~ \mach \\
+        V(k) & \textnormal{otherwise} \\
+      \end{cases}\right\rrbracket\)}
              {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
-                 ~ (V,\sigma) = (V'',V::\sigma)\)}
-    %
-    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(V'' = V'\llbracket \addr \mapsto \unsealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
+               ~ (V,\sigma) = (V',V::\sigma)\)}
+    %
+    \judgment{\(V' = V \left\llbracket
+      k \mapsto
+      \begin{cases}
+        \unsealed & \textnormal{if } k \in CLR \\
+        \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
+        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } \\
+        & k \in \object(V) - \mathit{passed} ~ \overline{sa} ~ \mach \\
+        V(k) & \textnormal{otherwise} \\
+      \end{cases}\right\rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{tail call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
@@
   \end{subfigure}
-  \caption{Operations supporting tail calls and argument passing on stack.\apt{This needs similar edits to the previous operations figure.
-In particular, same question about the $V''$ in call and tailcall.}}
+  \caption{Operations supporting tail calls and argument passing on stack.}
   \label{fig:advops}
@@
 elements to memory regions, represented as a base and a bound \(\context = (V, \sigma, \rho)\).
-Existing operations are extended to preserve the value of \(\rho\), and the new operations
-work as follows\apt{Put these and tweaked call rule into a separate figure in the same style as the existing two}:
-
+Most existing operations are extended to preserve the value of \(\rho\), while the new operations
+and the call operation work as seen in \cref{fig:capops}.
+
+\begin{figure*}
+  \begin{minipage}{0.5\textwidth}
     \judgmentbr[Promote]
@@
                {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-
+               %
+    \judgment[Clear]
+             {\(\psi = \mathbf{clear} ~ \component\)}
+             {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
+  \end{minipage}
+  \begin{minipage}{0.5\textwidth}
     \judgmentbr[Propagate]
@@
                {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-
-\judgment[Clear]
-         {\(\psi = \mathbf{clear} ~ \component\)}
-         {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
-
-We now have a notion of provenance, and must integrate it into the definition of
-stack safety. We essentially generalize the above notion of passing: we will consider
-a caller to have intentionally passed an object if that object is reachable by
-a capability that has been passed to the callee. This includes capabilities passed
-indirectly, by being stored in an object that is in turn passed. Formally, we call
-this set \(\mathit{capped}\), and define it recursively:
                %
@@
 
-We then tweak the call operation to seal only objects that are in \(\mathit{capped}\), or
-the previously defined \(\mathit{passed}\).
-
-\judgmentbrbrbr[]
-               {\(\psi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa}\)}
-               {\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
-                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
-               {\(\components = \{\addr | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \cup (\mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \}\)}
-               {\(V'' = V'\llbracket \components \mapsto \sealed \rrbracket\)}
-               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) =
-                 (V'',V::\sigma,\rho)\)}
+  \end{minipage}
+  \begin{minipage}{1\textwidth}
+    \vspace{\abovedisplayskip}
+    \judgment{\(V' = V \left\llbracket
+      k \mapsto
+      \begin{cases}
+        \unsealed & \textnormal{if } k \in CLR \\
+        \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
+        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } k \in \object(V) - (\mathit{passed} ~ \overline{sa} ~ \mach \cup \mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \\
+        V(k) & \textnormal{otherwise} \\
+      \end{cases}\right\rrbracket\)}
+             {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V',V::\sigma,\rho)\)}
+
+  \end{minipage}
+
+\caption{Operations supporting provenance-based protection of passed objects}
+\label{fig:capops}
+\end{figure*}
+
+This essentially generalizes the above notion of passing: we will consider
+a caller to have intentionally passed an object if that object is reachable by
+a capability that has been passed to the callee. Reachability includes capabilities passed
+indirectly, by being stored in an object that is in turn passed. We call
+the set of reachable addresses \(\mathit{capped}\). The call operation in this setting
+will seal only objects that are neither in \(\mathit{capped}\) nor the previously defined \(\mathit{passed}\).
 
@@
 complex enforcement mechanisms like Roessler and DeHon's micro-policies.
-We therefore choose to systematically \emph{test} their micro-policies.
-Our primary testing targets are the eager {\em Depth Isolation}
-and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.\apt{Reads
-  awkwardly coming just after last section.}
+We therefore choose to systematically \emph{test} their {\em Depth Isolation}
+and {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
 
@@
 
+\sna{Maybe a quick example or even two would be of benefit here.}
 For better performance, we instead maintain a single ``post-call taint''
@@
 \vspace*{1em}
-\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms\apt{what does the horizontal line mean?}}
+\caption{MTTF for finding bugs in erroneous \(DI\) (top) and \(LTC\) (bottom) micro-policies}
 \vspace*{-2em}
@@
 taking longer to catch.
-
-Our properties allow us to identify an enforcement mechanism as
-not really stack safe, and to validate a possible fix.\apt{This paragraph
-  comes as a bit of a non-sequitur here.}
 
@@
 
-\sna{Look into wasm compilation per reviewer 1. Low priority.}
-
 We plan to test our properties against multiple enforcement mechanisms.
@@
 We thank the reviewers for their comments, CHR Chhak and Allison Naaktgeboren for
-feedback during the writing process, and A\"ina Linn Georges for her enthusiastic reception
-of our early work.\apt{would consider rephrasing that last.}
+feedback during the writing process, and A\"ina Linn Georges for significant technical
+feedback and encouragement.
 
 This work was supported by the National Science Foundation under Grant No. 2048499, Specifying and Verifying Secure Compilation of C Code to Tagged Hardware;
-and by Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972.\apt{Benjamin?}{Leo?}
+by Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972;
+and by NSF award \#2145649, CAREER: Fuzzing Formal Specifications.
+\apt{Benjamin?}
 

### 2023-05-23 submission5/core.tex
BCP grants

@@
 by Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972;
-and by NSF award \#2145649, CAREER: Fuzzing Formal Specifications.
-\apt{Benjamin?}
+by NSF award \#2145649, CAREER: Fuzzing Formal Specifications, 
+by
+NSF award \#1955610, Bringing Python Up To Speed,
+and by %
+NSF award \#1521523, Expeditions in Computing: The Science of Deep
+  Specification.
 

### 2023-05-23 submission5/core.tex
A bit of polishing

@@
 integrity property and a confidentiality property for each of the
-caller and the callee, plus a control-flow property---five properties
+caller and the callee, plus a control-flow property: five properties
 in all.
@@
 Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, which permit
-functions to write into one another's frames, but which taint the
+functions to write into one another's frames but taint the
 changed locations so that the frame's owner cannot access them. No
 existing characterization of stack safety captures this style of
-safety. We capture it here by stating our properties in terms of the
+safety; we capture it here by stating our properties in terms of the
 observable behavior of the system.
@@
 %
-We validate our properties by using them to distinguish between
+We validate the properties by using them to distinguish between
 correct and incorrect implementations of Roessler and DeHon's
@@
 successfully identifies several broken variants, including Roessler
-and DeHon's lazy policy; a repaired version of their policy does pass
+and DeHon's lazy policy; a repaired version of their policy passes
 our tests.
@@
 
-Functions in high-level languages, and related abstractions such
-as procedures, methods, etc., are units of computation
-that call one another to define larger computations in a modular way.
-%
-At a low level, each function activation manages its own data---local
-variables, spilled temporaries, etc.---as well as information about the \emph{caller} to
-which it must return.
+Functions in high-level languages (and related abstractions such
+as procedures, methods, etc.) are units of computation
+that invoke one another to define larger computations in a modular way.
+%
+At a low level, each function activation manages its own local
+variables, spilled temporaries, etc., as well as information about the
+{caller} to which it will return.
 %
@@
 %using data stored in the stack
+
 The variety in attacks on the stack is mirrored in the range of
@@
 and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18,Gollapudi+23}.
-
-Enforcement mechanisms can be brittle, successfully eliminating one
+%
+But enforcement mechanisms can be brittle, successfully eliminating one
 attack while leaving room for others. To avoid an endless game of
@@
 proven, or at least rigorously tested. Such properties can be used as
-the specification against which enforcement can be validated; even
-enforcement mechanisms that do not fulfill a property benefit from the
+the specification against which enforcement can be validated---even
+enforcement mechanisms that do {\em not} fulfill a property can benefit from the
 ability to articulate why and when they may fail.
 
-Of the mechanisms listed above, many are fundamentally ill-suited for
-offering formal guarantees: they may impede attackers, but do not provide
+Many of the mechanisms listed above are fundamentally ill-suited for
+offering formal guarantees: they may impede attackers, but they do not provide
 universal protection. Shadow stacks, for instance, aim to ``restrict
@@
 integrity~\cite{Kuznetsov+14}, do aim for stronger guarantees, but not
-typically formal ones.  To our knowledge, the sole
+typically\bcp{do we need that word?} formal ones.  To our knowledge, the sole
 line of work making a formal claim to protect stack safety is the
@@
 
-Mechanisms besides this line of work should also be amenable to strong
+There are other mechanisms that should also be amenable to strong
 formal guarantees.  In particular, Roessler and DeHon
 \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
-micro-policies~\cite{pump_oakland2015} for stack safety that offer
-universal protection, but not yet tied to a particular formal guarantee.
+micro-policies~\cite{pump_oakland2015} for stack safety that aim to offer
+universal protection.  But the reasoning involved is somewhat subtle.
 Their most realistic micro-policy, called Lazy

### 2023-05-23 submission5/core.tex
Intro

@@
 
-We propose a formal characterization of stack safety using concepts
+We propose a new formal characterization of stack safety using concepts
 from language-based security. Rather than treating
@@
 
-There are other mechanisms that should also be amenable to strong
+Some of the other mechanisms listed above should also be amenable to strong
 formal guarantees.  In particular, Roessler and DeHon
@@
 Their most realistic micro-policy, called Lazy
-Tagging and Clearing (\(LTC\)), makes an interesting performance
+Tagging and Clearing ({\LTC}\bcp{does it have to be italic?}), makes an interesting performance
 trade-off: it allows function activations to write improperly into one
@@
 memory---just not in ways that affect observable behavior.
-Therefore, \(LTC\) would not fulfill Georges et al.'s property
-(adapted to the tagged setting).  But \(LTC\) does
+Therefore, {\LTC} would not fulfill Georges et al.'s property
+(adapted to the tagged setting).  But {\LTC} does
 arguably enforce stack safety.  A looser, more observational
@@
 
-We propose here a formal characterization of stack safety, based on
-the intuition of protecting function activations (both registers and
-stack frames) from each other.
+We propose here a formal characterization of stack safety based on
+the intuition of protecting function activations
+(both registers and stack frames)\bcp{Can we delete that parenthetical
+remark?  I think it distracts more than it clarifies.}
+from each other and
 %\bcp{Aren't other formal definitions based on the same intuition?}%
@@
 %% \bcp{Both useful!  But neither comes across, at least not very clearly.}%
-We use the tools of language-based
-security~\cite{sabelfeld2003language}, treating function activations as
+using the tools of language-based
+security~\cite{sabelfeld2003language} to treat function activations as
 security principals.  We decompose stack safety into a family of
@@
 the {\em integrity} and {\em confidentiality} of the caller’s local state
-and the callee's behavior during (and after) the callee's execution, plus the
+and the callee's behavior during (and after) the callee's execution,
+together with the
 {\em well-bracketed control flow} (\(\wbcf\)) property articulated by
 Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
-While our properties are motivated by the desire to specify \(LTC\) precisely,
-they are stated abstractly in the hope that they can also be applied
-to other enforcement mechanisms.
-
-%\subsection {Towards Formal Properties for Stack Safety}
-
-We are not attempting to present a universal definition of stack safety
-for all systems. While many security properties can be described at
+
+Our properties are stated abstractly in the hope that they can also be
+applied to other enforcement mechanisms besides \LTC.
+%
+However, it does not seem feasible to give
+a universal definition of stack safety
+that applies to all architectures and compilers.
+While many security properties can be described purely at
 the level of a high-level programming language and translated to a
@@
 this way, since ``the stack'' is not explicitly present in the
-definitions of most source languages. It is implied by the semantics
-of features such as calls and returns, but is not itself a feature.\footnote{
+definitions of most source languages but rather is implicit in the semantics
+of features such as calls and returns.\footnote{
 Contrast Azevedo de
   Amorim et al.'s work on heap safety
-  \cite{DBLP:conf/post/AmorimHP18}: the concept of the heap figures
+  \cite{DBLP:conf/post/AmorimHP18}, where the concept of the heap figures
   directly in high-level language semantics and its security is
@@
 %
-But neither can stack safety be an entirely satisfying low-level property;
+But neither can stack safety be described coherently as a purely
+low-level property;
 indeed, at the lowest level, the specification of a ``well-behaved
@@
 
-Therefore, any low-level treatment of stack safety must begin by asking:
-which high-level features does our system support, and what are their
+Thus, any low-level treatment of stack safety must begin by asking:
+which high-level features are supported in a given
+setting using the stack, and what are their
 security requirements? Different answers will beget different properties.
+\bcp{There's an awkward conceptual transition here: We pose a question
+(``which high-level features are supported in a given
+setting using the stack, and what are their
+security requirements'') but then we don't seem to be going ahead and
+answering it.  I suggest adding a sentence here that answers the
+question directly, then a paragraph break.}
 We identify and define five security requirements that a system might try
@@
 integrity and confidentiality of each of the caller and the callee.
-%These are by no means exhaustive\bcp{That comment belongs in the
-%  future work
-%  section, not here}.
+%
 We apply these first to a simple system
@@
 random testing tool for Coq.
-We find that \(LTC\) is flawed in
+We find that {\LTC} is flawed in
 a way that undermines both integrity and confidentiality; after
-correcting this flaw, the repaired \(LTC\) satisfies all our
-properties.  Moreover, we modify \(LTC\) to protect the features of
+correcting this flaw, the repaired {\LTC} satisfies all our
+properties.  Moreover, we modify {\LTC} to protect the features of
 our more realistic system, and apply
@@
 %that violate our properties but are \emph{not} halted by incorrect
-%variants of the enforcement mechanisms---including the original \(LTC\)!
+%variants of the enforcement mechanisms---including the original {\LTC}!
 % SNA: we talk about this in the testing section, and it's a "best practice" for
@@
 Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
-{\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
+{\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} ({\LTC}).
 Their precise connection to Roessler and DeHon's work is discussed below.
@@
 
-In \(LTC\), a function neither initializes the frame at entry nor clears it at exist;
+In {\LTC}, a function neither initializes the frame at entry nor clears it at exist;
 instead, it simply sets each location's tag to the PC tag when that location is written. It does
@@
 For Roessler and DeHon (henceforward \emph{R\&D}), Lazy Tagging and Lazy Clearing are both optimizations that
-can be applied to their Depth Isolation policy. Our version of \(LTC\)
+can be applied to their Depth Isolation policy. Our version of {\LTC}
 corresponds to Depth Isolation with both optimizations applied. How closely
@@
 In particular, \(\clri\) requires callee-saved
-registers to be saved and restored properly. We extend \(DI\) and \(LTC\)
+registers to be saved and restored properly. We extend \(DI\) and {\LTC}
 so that callee-saved registers are also tagged with the color of the
 function that is using them. In \(DI\) they are tagged as part of the entry
-sequence, while in \(LTC\) they are tagged when a value is placed in them.
+sequence, while in {\LTC} they are tagged when a value is placed in them.
 
@@
 \vspace*{1em}
-\caption{MTTF for finding bugs in erroneous \(DI\) (top) and \(LTC\) (bottom) micro-policies}
+\caption{MTTF for finding bugs in erroneous \(DI\) (top) and {\LTC} (bottom) micro-policies}
 \vspace*{-2em}
@@
 
-For \(LTC\), the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
+For {\LTC}, the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
 fails in testing, in cases where data is leaked between sequential calls.

### 2023-05-23 submission5/core.tex
Polishing and comments

@@
 answering it.  I suggest adding a sentence here that answers the
-question directly, then a paragraph break.}
+question directly, then a paragraph break.} \bcp{Though, actually, the
+next sentence repeats points that have already been made above.  Maybe
+better to just delete it and find a new transition to the sentence
+after.}
 We identify and define five security requirements that a system might try
@@
 
-We show that our properties are useful for distinguishing between correct and
-incorrect enforcement using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
+We demonstrate the usefulness of our properties for distinguishing
+between correct and
+incorrect enforcement using
+QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
 random testing tool for Coq.
-We find that {\LTC} is flawed in
+Indeed, we find that the published version of {\LTC} is flawed in
 a way that undermines both integrity and confidentiality; after
-correcting this flaw, the repaired {\LTC} satisfies all our
-properties.  Moreover, we modify {\LTC} to protect the features of
-our more realistic system, and apply
+correcting this flaw, {\LTC} satisfies all our
+properties.  Further, we modify {\LTC} to protect the features of
+our more realistic system and apply
 random testing to validate this extended protection mechanism against
@@
 
-In sum, we offer the following contributions:
+In summary, we offer the following contributions:
 
 \begin{itemize}
-\item We give a novel characterization of stack safety as a collection
-  of properties: confidentiality and integrity for callee and caller,
-  plus well-bracketed control-flow.
+\item We give a novel characterization of stack safety as a conjunction
+  of security properties---confidentiality and integrity for callee
+  and caller---plus well-bracketed control-flow.
   The properties are parameterized over a notion of
@@
   describe a realistic setting with argument passing on the stack,
-  callee-saves registers, and tail-call elimination. Our model is
-  modular enough that handling these features is straightforward.
-\item We validate a tag-based enforcement mechanism, \emph{Lazy
+  callee-saves registers, and tail-call elimination. The model is
+  modular enough that adding these features is straightforward.
+\item We validate a published enforcement mechanism, \emph{Lazy
   Tagging and Clearing}, via property-based random testing, find that
@@
 \end{itemize}
-
-In the next section, we give a brief overview of our framework and
-assumptions. In \cref{sec:example}, we walk through a function call in
-a simple example machine and discuss informally how each of our properties
-applies to it. In the process we motivate the properties
+%
+The following section offers a brief overview of our framework and
+assumptions. \cref{sec:example} walks through a function call in
+a simple example machine, discusses informally how each of our properties
+applies to it, and motivates the properties
 from a security perspective.
 %
-In \cref{sec:formal} we formalize the machine model,
+\cref{sec:formal} formalizes the machine model,
 its {\em security semantics}, and the stack safety properties built on these.
@@
 %heap-like sharing model based on pointer provenance.
-In \Cref{sec:enforcement} we describe the micro-policies that we test,
-in \cref{sec:testing} the testing framework itself, and in
+\Cref{sec:enforcement} describes the micro-policies that we test,
+\cref{sec:testing} the testing framework itself, and
 \cref{sec:relwork,sec:future} related and future work.
@@
 contains formal definitions (in Coq) of our
-properties, plus our testing framework.  It does not include proofs:
+properties, plus our testing framework.  It does not include proofs, since
 we use Coq primarily for the QuickChick testing library and to
@@
 The action of a security-relevant operation on the context is defined by a set
-of rules that describe how the machine code implements the function abstractoin
-using the stack and registers.
-
-Given the security classes of the machine's state elements, we define high-level
-security properties---integrity, confidentiality, and well-bracketed control flow---as
+of rules that characterize how particular machine code sequences are used
+to implement the function abstraction in terms of the stack and registers.\bcp{OK?}
+
+Given the security classes of the elements of the machine state, we
+define high-level
+security properties---integrity, confidentiality, and well-bracketed
+control flow---as
 predicates that must hold on each call. These predicates draw on the idea of
-\emph{variant} states from the theory of non-interference, and a notion of
+\emph{variant} states from the theory of non-interference, plus a notion of
 {\em observable events}, which might include specific function calls (e.g., system
@@
 memory-mapped regions, etc. For example, to show that certain locations are kept
-secret, it suffices to compare the execution of states which vary at those locations
+secret, it suffices to compare executions starting at machine states which vary at those locations
 and check that their traces of observable events are the same. This structure
-allows us to talk about the eventual implact of leaks or memory corruption without
-reference to internal implementation details, and to support lazy enforcement by
+allows us to talk about the eventual impact of leaks or memory corruption without
+reference to internal implementation details and, in particular, to support lazy enforcement by
 flagging corruption of values only when it can actually impact visible behavior.
@@
 formally in \cref{sec:formal}. In the remainder
-of this section we describe the underlying framework in more detail.
+of this section we introduce the underlying \bcp{semantic?} framework
+in more detail.
 
 \paragraph*{Machine Model}
-We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
+We assume a conventional ISA (e.g., RISC-V, x86-64, etc.), with registers including a program counter
 and stack pointer.
@@
 We make no particular assumptions about the provenance of the machine code; in particular,
-we do not assume the use of any particular compiler.
-The machine may possibly be enhanced with % security
-enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
-we assume that the behavior of these mechanisms is incorporated into the basic
-step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
+we do not assume any particular compiler.
+If the machine is enhanced with % security
+enforcement mechanisms such as hardware
+tags~\cite{pump_hasp2014,Gollapudi+23} or
+capabilities~\cite{Woodruff+14}, we 
+assume that the behavior of these mechanisms is incorporated into the basic
+step semantics of the machine, but we keep the enforcement state
+unchanged when constructing variant states\bcp{As a reader, I wonder
+  ``How would they do that?  Seems like they are saying the tag
+  enforcement or whatever is just part of the basic machine step
+  relation and we can't see it, but then going on to place a
+  requirement on it. How would such a requirement be phrased, formally?''}.
 Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
@@
 \end{center}
-\caption{Security-relevant operations and their parameters, with the Sections in which
-  they are first used in the paper.}
+\caption{Security-relevant operations and their parameters, with the
+  sections where
+  they are first used.}
   \label{tab:psi}
@@
 if, in the example, we varied the assembly code for {\tt main} so that {\tt sensitive} were stored
-in a callee-save register (e.g. {\tt s0}) rather than in memory, its security class would still
+in a callee-save register (e.g., {\tt s0}) rather than in memory, its security class would still
 be \(\sealed\) at the entry to {\tt f}.
@@
 
-To begin with, what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
+To begin with, what if {\tt f} returns to an unexpected place (i.e., \(\PCname \neq 20\) or
 \(\SP \neq 980\))? We consider this to violate \(\wbcf\). \(\wbcf\) is a relationship between
@@
 and lift it element-wise to states.  Two values should be compatible if their
-non-payload information (e.g. their tag) is identical.
+non-payload information (e.g., their tag) is identical.
 %% \apt{This is still not very convincing!
 %%   empting to say that only the payloads can affect the step function, but that is
-%%   too strong, right? (OK for tags, but not, e.g. capabilities?}
+%%   too strong, right? (OK for tags, but not, e.g., capabilities?}
 %% \sna{Yeah, I don't think we can say something universal about when values should be
@@
 As such, this capability safety relies on some preexisting notion of pointer
-validity, i.e. {\it pointer provenance}.
+validity, i.e., {\it pointer provenance}.
 Memarian et al.'s PVI \cite{provenance} (provenance via integer)
@@
 %well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
-%address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
+%address, and shadow stacks with protection (e.g., Return Address Defender~\cite{Chiueh2001RAD})
 %to enforce it completely. Others combine this protection with some degree of memory protection,

### 2023-05-23 submission5/core.tex
Polishing and comments

@@
 
-\paragraph*{Security Semantics and Property Structure}
-A security semantics extends a machine
+\paragraph*{Security Semantics and Property Structure}\bcp{Delete
+  ``and Property Structure''?}
+A security semantics extends the core machine model
 with additional context about the identities of current and pending
-functions (which act as security principals) and about the registers and memory they require
-to be secure. This added context is purely notional;
-it does not affect the behavior of the real machine. The security context
+functions (which act as security principals) and about their security
+requirements on registers and memory. This added context is purely notional;
+it does not affect the behavior of the core machine. The security context
 evolves dynamically through the execution of security-relevant operations,
@@
 Our security properties are phrased in terms of this context, often as predicates
-on future states (``when control returns to the current function...'')
+on future states (``when control returns to the current function, X
+must hold...'')
 or as relations on traces of future execution
@@
 actions they take. In this paper, each operation corresponds to a
-single underlying machine instruction. One instruction may perform multiple
-operations. We assume that a compiler or other trusted source has provided
-labels to disambiguate instructions that have multiple purposes. For instance,
+single underlying machine instruction\bcp{That sentence is opaque.}.
+Since the same machine instruction may be used by compilers for
+different purposes, we assume that the compiler or another trusted
+source has provided
+labels to identify the security-relevant purpose of each instruction,
+if any. For instance,
 in the tagged RISC-V architecture
@@
 
-We end up with an annotated version of the
-machine transition function written \(\mach \xrightarrow{\bar{\psi},
+These considerations lead to an annotated version of the
+machine transition function, written \(\mach \xrightarrow{\bar{\psi},
   \obs} \mach'\), where \(\mach\) and \(\mach\) are machine states, \(\obs\) is
-an (optional) external event, and \(\bar{\psi}\) is a list of security-relevant operations.
-%
-We then lift this into a transition between pairs of machine states and contexts,
-by applying a set of operation-specific rules to transform each state and context to
-a new context, in parallel with the ordinary transition on states.
-The most important rules describe call and return operations.
+an optional external\bcp{observable?} event, and \(\bar{\psi}\) is a
+list of security-relevant operations\bcp{that bit is not completely
+  clear---e.g., why a list?  because the same
+machine step can do several distinct security-revevant things}.
+%
+This is then lifted into a transition between pairs of machine states
+and contexts
+by applying a set of operation-specific rules\bcp{why is it a set of
+  rules rather than a single transition function?} that, given a
+state and associated state, yield a new context corresponding to
+the new state calculated by the ordinary transition relation on states.
+%
+The most important of these rules describe call and return operations.
 A call pushes a new view onto the context stack and changes the class of the
 caller's data to protect it from the new callee; a return reverses these steps.
-Other operations can serve to signal how parts of the stack frame are being used to store
+Other operations signal how parts of the stack frame are being used to store
 or share data, and their corresponding rules alter the classes of different
@@
 tail-call elimination is described in \cref{sec:extensions} and tested in \cref{sec:testing}.
-The remaining operations are needed for the capability-based model in \cref{app:ptr}.
+The remaining operations are needed for the capability-based model in
+\cref{app:ptr}.
 
@@
   sections where
-  they are first used.}
+  they are first used.  \bcp{Say what the gray levels mean.}}
   \label{tab:psi}
@@
 
-The security context consists of a stack of views: functions that map
-each state element to a {\it security class}:
+The security context consists of a stack of \emph{views}, where a view is a
+function mapping
+each state element to a {\it security class}---one of
 \(\public\), \(\unsealed\), \(\object\), or \(\sealed\).
@@
 registers---are always labeled \(\public\). We place security requirements on some
-\(\public\) elements for purposes of the \(\wbcf\) property, and a
+\(\public\) elements for purposes of the well-bracketed control flow \(\wbcf\) property, and a
 given enforcement mechanism
@@
 
-For a newly active function, every stack location that is available
+When a function is newly activated, every stack location that is available
 for use but not yet initialized
-is seen as \(\unsealed\). From the perspective of the caller, the callee has no obligations
+is \(\unsealed\). From the perspective of the caller, the callee has no obligations
 regarding its use of free elements.
 
-Arguments are marked \(\object\), meaning that their contents may be safely used.
+Arguments are marked \(\object\), meaning that their contents may be
+used safely.
 When a function allocates memory for its own stack frame, that memory will also be \(\object\).
 Then, on a call, \(\object\) elements that are not being used to communicate with
-the new callee will become \(\sealed\)---reserved for an inactive principal,
+the callee will become \(\sealed\)---i.e., reserved for an inactive principal
 and expected to be unchanged when it becomes active again.
@@
 Conceptually, the following steps are needed to instantiate the framework to a specific machine
-and code style: (i) define the base machine semantics, including any hardware
+and coding conventions: (i) define the base machine semantics, including any hardware
 security enforcement features; (ii) identify the set of
-security-relevant operations and rules required by the code style; (iii) determine
+security-relevant operations and rules required by the coding conventions; (iii) determine
 how to label machine instructions with security-relevant
@@
 protected from the callee; conversely, marking too many operations as
-calls will simply cause otherwise safe programs to be rejected.
-
-We do not assume that low-level code adheres to any fixed calling
-convention or implements any particular source-language constructs.
+calls may cause otherwise safe programs to be rejected.
+
+We do not assume that low-level code adheres to any fixed\bcp{single?
+(I think we do assume that the calling convention is fixed in any
+particular instance of our framework, no?)} calling
+convention or is being used to implement any particular
+source-language constructs.
 Indeed, if the source language is C, then high-level programs might
@@
 architecture might place additional constraints, particularly on
-call and return sequences.
+call and return sequences.\bcp{Not sure that last sentence is helpful.}
 
@@
 In general, it is impossible to distinguish buggy machine code from an
-attacker.  In our examples we will identify one function or another as
-an attacker, but we do not require any static division between trusted
+attacker.  In examples, we often identify one function or another as
+an attacker, but our framework does not require any static division between trusted
 and untrusted code, and we aim to protect even buggy code.
@@
 
-In this section we introduce our security property definitions by means
-of small code examples, using a simplified set of security-relevant operations for
+In this section, we introduce our security properties by means
+of small code examples, using a simple set of security-relevant operations for
 calls, returns, and private allocations.
+
 \Cref{fig:main} gives C code and possible corresponding compiled 64-bit RISC-V code
@@
 Then {\tt main} calls another function {\tt f},
-and afterward performs a test on {\tt sensitive} to decide whether
+and afterward it performs a test on {\tt sensitive} to decide whether
 to output {\tt secret}.  Since {\tt sensitive} is initialized to 0,
-the test should fail, and {\tt main} should instead output the return value of {\tt f}.
+the test should always fail, and {\tt main} should instead output the return value of {\tt f}.
 Output is performed by writing to the special global {\tt out},
@@
 The C code is compiled using the standard RISC-V calling conventions~\cite{RISC-V-CC}.
-In particular, the first function argument and the function
+In particular, the function's first argument and its
 return value are both passed in {\tt a0}.
-Memory is byte-addressed and the stack grows towards
-lower addresses. We assume that {\tt main} begins at address 0 and its callee {\tt f} at address 100.
+Memory is byte addressed, and the stack grows towards
+lower addresses. We assume that {\tt main} begins at address 0 and its
+callee {\tt f} at address 100.  \bcp{What are alloc(-20,20) and
+  dealloc(0,20)? What is $\varepsilon$ in call $\varepsilon$?  Also, readers who don't have the RISCV ISA at their
+fingertips may wonder what all the instructions actually do.  We
+should (if it's true) explicitly say that they don't need to worry
+about the details.}
 
@@
   \newcommand{\figonebox}[1][]{\genbox{30pt}{blue!20}{#1}}
-  \begin{subfigure}{\columnwidth}
+  ~~~~~~\begin{subfigure}{\columnwidth}
     {\tt
@@
 
-  \caption{Example: C and assembly code for {\tt main}, and layout of its stack frame; the stack grows to the left.}
+  \caption{Example: C and assembly code for {\tt main} and layout of its stack frame (the stack grows to the left).}
   \label{fig:main}
@@
 
-We now consider how {\tt f} might misbehave and violate desirable
-stack safety properties associated with {\tt main}. To put the violations in a
-security framework, suppose that {\tt f} is actually an attacker seeking
+Now, suppose that {\tt f} is actually an attacker seeking
 to leak {\tt secret}. It might do so in a number of ways, shown as snippets of
@@
 In \cref{subfig:direct}, {\tt f} takes an offset from the stack
-pointer, accesses {\tt secret}, and directly outputs it. But more
-subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
-can instead return that value so that {\tt main} stores it to {\tt out},
+pointer, accesses {\tt secret}, and directly outputs it.  More
+subtly, even if it is somehow prevented from outputting {\tt secret}
+directly, {\tt f}
+can instead return its value so that {\tt main} stores it to {\tt out},
 as in \cref{subfig:indirect}.
@@
 with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
-(\cref{subfig:integrity}).
-Attacks of this kind do not violate {\tt main}'s confidentiality, but
+(\cref{subfig:integrity}); this does not violate {\tt main}'s
+confidentiality, but
 rather its {\it integrity}.
 In \cref{subfig:WBCF}, the attacker arranges to return to the
-wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
-violating the program's {\it well-bracketed control flow} (\(\wbcf\)).
+wrong instruction, thereby bypassing the check and publishing {\tt
+  secret} regardless; this
+violates the program's {\it well-bracketed control flow} (\(\wbcf\)).
 %
 In \cref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
-by returning to the correct program counter but with the wrong stack pointer.%
-\footnote{We pad some of the variants with {\tt nop}s just so that all the
-snippets have the same length, which keeps the step numbering uniform in~\cref{fig:exec1}.}
+by returning to the correct program counter but with the wrong stack pointer.
+%
+(We pad some of these variants with {\tt nop}s just so that all the
+snippets have the same length, which keeps the step numbering uniform in~\cref{fig:exec1}.)
 
@@
 Our security semantics attaches a security context to the machine state,
-which consists of a view \(V\) and a stack \(\sigma\) of pending activations' views.
+consisting of a view \(V\) and a stack \(\sigma\) of pending activations' views.
 \Cref{fig:exec1} shows how the security context evolves over the first few
@@
 \cref{sec:formal}, and the context evolution rules are formalized in \cref{fig:basicops}.)
-Execution begins at the start of {\tt main}, where the program counter (\(\PCname\)) is zero,
-and with the stack pointer (\(\SP\)) at address 1000.
+Execution begins at the start of {\tt main}, with the program counter
+(\(\PCname\)) set to zero and the stack pointer (\(\SP\)) at address 1000.
 State transitions are numbered and labeled with a list of security operations, written
-\(\downarrow \overline{\psi}\) between steps. We write \(\emplist\) for empty lists.
+\(\downarrow \overline{\psi}\), between steps. We write \(\emplist\)
+for empty lists.  \bcp{Is it ever anything other than empty or a
+  singleton?}
 
@@
 Step 1 allocates a word each for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
-as two words for the return address. This has the
+as two words for the return address; this has the
 effect of marking those bytes \(\object\).
-(We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).)
+We use \(V\llbracket\ldots\rrbracket\) to denote updates to \(V\).
 
@@
 The callee's view is updated from the caller's such that all \(\object\) memory locations
-become \(\sealed\). (For now we assume no sharing of memory between activations; data is
+become \(\sealed\). (For now we assume no sharing of stack memory between activations; data is
 passed only through argument registers, which remain active. In the presence of memory
@@
 mapped to \(\object\), while any non-argument, caller-saved
-registers are mapped to \(\unsealed\). In the current example, only register {\tt a0} has a change in
+registers are mapped to \(\unsealed\). In the current example, only register {\tt a0}  changes
 security class. All callee-save registers remain \(\sealed\) for all calls, so
-if, in the example, we varied the assembly code for {\tt main} so that {\tt sensitive} were stored
+if, in the example, we varied the assembly code for {\tt main} so that {\tt sensitive} was stored
 in a callee-save register (e.g., {\tt s0}) rather than in memory, its security class would still
@@
 %
-At step 9, {\tt f} returns, and the topmost inactive view, that of {\tt main}, is restored.
+At step 9, {\tt f} returns and the topmost inactive view, that of {\tt main}, is restored.
 
@@
 
-\paragraph*{Well-bracketed Control Flow}
-
-To begin with, what if {\tt f} returns to an unexpected place (i.e., \(\PCname \neq 20\) or
-\(\SP \neq 980\))? We consider this to violate \(\wbcf\). \(\wbcf\) is a relationship between
+\paragraph*{Well-Bracketed Control Flow}
+
+To begin with, if {\tt f} returns to an unexpected place (i.e., \(\PCname \neq 20\) or
+\(\SP \neq 980\)), we say that it has violated \(\wbcf\). \(\wbcf\) is a relationship between
 call steps and their corresponding return steps: just after the return, the program
-counter should be at the next instruction following the call,
-and the stack pointer should be the same as it was before the call.
+counter should be at the next instruction below the call,
+and the stack pointer should have the same value that it had before the call.
 Both of these are essential for security. In \cref{subfig:WBCF}, the attacker adds
@@
 \noindent
-{\tt main}'s attempt to read {\tt sensitive} will instead
+{\tt main}'s attempt to read {\tt sensitive} will\bcp{may?} instead
 read part of the return address, and its attempt to output
@@
 just after the call.
-\(\wbcf\) requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
-(This property is formalized in \cref{tab:props}, line 1).
+\(\wbcf\) requires that, if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
+This property is formalized in \cref{tab:props}, line 1.
 %For nested calls, where the pending stack is initially larger, the same principle

### 2023-05-24 submission5/core.tex
Andrew's comments+many of Benjamin's

@@
 integrity~\cite{Kuznetsov+14}, do aim for stronger guarantees, but not
-typically\bcp{do we need that word?} formal ones.  To our knowledge, the sole
+formal ones.  To our knowledge, the sole
 line of work making a formal claim to protect stack safety is the
@@
 We propose here a formal characterization of stack safety based on
-the intuition of protecting function activations
-(both registers and stack frames)\bcp{Can we delete that parenthetical
-remark?  I think it distracts more than it clarifies.}
-from each other and
-%\bcp{Aren't other formal definitions based on the same intuition?}%
-%% \sna{Two differences we're trying to get across. First, protecting
-%%   functions vs. protecting stack frames: we bring registers into
-%%   the story because it isn't just about the stack as a data structure.
-%%   Second: describing security in terms of functions' high-level
-%%   security goals (f is trying to extract a secret) rather than
-%%   describing operationally what they can and can't do (as the CHERI
-%%   folks do.)}%
-%% \bcp{Both useful!  But neither comes across, at least not very clearly.}%
+the intuition of protecting function activations from each other and
 using the tools of language-based
@@
 formally in \cref{sec:formal}. In the remainder
-of this section we introduce the underlying \bcp{semantic?} framework
+of this section we introduce the underlying semantic framework
 in more detail.
@@
 
-\paragraph*{Security Semantics and Property Structure}\bcp{Delete
-  ``and Property Structure''?}
+\paragraph*{Security Semantics}
 A security semantics extends the core machine model
@@
 Security-relevant operations abstract over the implementation details of the
-actions they take. In this paper, each operation corresponds to a
-single underlying machine instruction\bcp{That sentence is opaque.}.
-Since the same machine instruction may be used by compilers for
+actions they take. Since the same machine instruction may be used by compilers for
 different purposes, we assume that the compiler or another trusted
@@
   \obs} \mach'\), where \(\mach\) and \(\mach\) are machine states, \(\obs\) is
-an optional external\bcp{observable?} event, and \(\bar{\psi}\) is a
-list of security-relevant operations\bcp{that bit is not completely
-  clear---e.g., why a list?  because the same
-machine step can do several distinct security-revevant things}.
+an optional externally observable event, and \(\overline{\psi}\) is a
+list of security-relevant operations---necessary because a single step might
+perform multiple simultaneous operations.
 %
 This is then lifted into a transition between pairs of machine states
-and contexts
-by applying a set of operation-specific rules\bcp{why is it a set of
-  rules rather than a single transition function?} that, given a
-state and associated state, yield a new context corresponding to
-the new state calculated by the ordinary transition relation on states.
+and contexts by applying a transition function parameterized by the operation.
+We will decompose this function into rules associated with each operation and introduce
+them as needed.
 %
@@
 (\(\PCname\)) set to zero and the stack pointer (\(\SP\)) at address 1000.
-State transitions are numbered and labeled with a list of security operations, written
-\(\downarrow \overline{\psi}\), between steps. We write \(\emplist\)
-for empty lists.  \bcp{Is it ever anything other than empty or a
-  singleton?}
+State transitions are numbered and may be labeled with a security operation, written
+\(\downarrow \psi\), between steps.
 
@@
     \cline{1-3}
-    \multicolumn{3}{l}{\multirow{2}{*}{\(1 \Big\downarrow [\mathbf{alloc} ~ (-20,20)]\)}} & \\
+    \multicolumn{3}{l}{\multirow{2}{*}{\(1 \Big\downarrow \mathbf{alloc} ~ (-20,20)\)}} & \\
     \multicolumn{3}{l}{} &
@@
     \cline{1-3}
-    \multicolumn{3}{l}{\multirow{2}{*}{2-4 \(\Big\downarrow \emplist\)}} \\ \multicolumn{3}{l}{} \\
+    \multicolumn{3}{l}{\multirow{2}{*}{2-4 \(\Big\downarrow\)}} \\ \multicolumn{3}{l}{} \\
     \cline{1-3}
@@
     \cline{1-3}
-    \multicolumn{3}{l}{\multirow{2}{*}{\(5 \Big\downarrow [\mathbf{call} ~ 100 ~ \emplist]\)}} & \\
+    \multicolumn{3}{l}{\multirow{2}{*}{\(5 \Big\downarrow \mathbf{call} ~ 100 ~ \emplist\)}} & \\
     \multicolumn{3}{l}{} &
@@
     \cline{1-3}
-    \multicolumn{3}{l}{\multirow{2}{*}{6-8 \(\Big\downarrow \emplist\)}} \\ \multicolumn{3}{l}{} \\
+    \multicolumn{3}{l}{\multirow{2}{*}{6-8 \(\Big\downarrow\)}} \\ \multicolumn{3}{l}{} \\
     \cline{1-3}
@@
     \cline{1-3}
-    \multicolumn{3}{l}{\multirow{2}{*}{\(9 \Big\downarrow [\mathbf{return}]\)}} & \\
+    \multicolumn{3}{l}{\multirow{2}{*}{\(9 \Big\downarrow \mathbf{return}\)}} & \\
     \multicolumn{3}{l}{} & \multirow{3}{*}{\(\underbrace{\dots \freebox \freebox \freebox \freebox \freebox}_\unsealed
@@
 \noindent
-{\tt main}'s attempt to read {\tt sensitive} will\bcp{may?} instead
+{\tt main}'s attempt to read {\tt sensitive} may instead
 read part of the return address, and its attempt to output
@@
     %
-    \judgment{\(V' = V \left\llbracket
-      k \mapsto
+    \judgment{\(V' = \lambda \component .
       \begin{cases}
@@
         V(k) & \textnormal{otherwise} \\
-      \end{cases}\right\rrbracket\)}
+      \end{cases}\)}
              {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}})
@@
              {\(Op ~ \mach ~ \mathbf{return} ~ (\_, (V,\sigma')) = (V, \sigma')\)}
-             \caption{Basic Operations\apt{...Do we really need two-stage calculation in the call rule, or was that just to avoid type-setting issues?}\sna{The latter.
-I took a stab at trying to avoid it, but I'm ambivalent about  that ``cases'' style here.}}
+             \caption{Basic Operations}
   \label{fig:basicops}
@@
     \rowcolor{black!20}
-    & & & \(\textnormal{ \hspace{0.8in}} \land \mach''[\PCname] = \mach[\PCname]+\mathit{sz}\) & where \(\mathit{sz}\) is the size of one instruction \\
+    & & & \(\textnormal{ \hspace{0.8in}} \land \mach''[\PCname] = \mach[\PCname]+\mathit{sz}\) & \textnormal{ where} \(\mathit{sz}\) is the size of instruction at \(\mach[\PCname]\) \\
     %
@@
     \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
-               {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
+               {\(V' = V \llbracket \addr \mapsto \object \mid \addr \in \components \rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
@@
     %
-    \judgmentbr{\(b = \mach[\SP] + \mathit{off}\)}
-               {\(V' = V \llbracket \addr \mapsto \unsealed |
-                 b \leq a < b+\mathit{sz} \land \addr \in \object(V) \rrbracket\)}
+    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \object(V)\)}
+               {\(V' = V \llbracket \addr \mapsto \unsealed | \addr \in \components \rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
@@
     %
-    \[\mathit{passed} ~ \overline{sa} ~ \mach = \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
-    \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
-    %
-    \judgment{\(V' = V \left\llbracket
-      k \mapsto
+    \[\mathit{push} ~ V ~ \overline{\reg} ~ K \triangleq \lambda k .
     \begin{cases}
@@
       \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } \\
-        & k \in \object(V) - \mathit{passed} ~ \overline{sa} ~ \mach \\
+      & k \in \object(V) - \components \\
       V(k) & \textnormal{otherwise} \\
-      \end{cases}\right\rrbracket\)}
+    \end{cases}      
+    \]
+    %
+    \[\mathit{passed} ~ \overline{sa} ~ \mach \triangleq \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
+    \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
+    %
+    \judgmenttwo[\sc Call]
+        {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
+        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ \components\)}
         {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
@@
     %
-    \judgment{\(V' = V \left\llbracket
-      k \mapsto
-      \begin{cases}
-        \unsealed & \textnormal{if } k \in CLR \\
-        \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
-        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } \\
-        & k \in \object(V) - \mathit{passed} ~ \overline{sa} ~ \mach \\
-        V(k) & \textnormal{otherwise} \\
-      \end{cases}\right\rrbracket\)}
+    \judgmenttwo[\sc Tailcall]
+        {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
+        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ \components\)}
         {\(Op ~ \mach ~ (\mathbf{tail call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
           ~ (V,\sigma) = (V',\sigma)\)}
-    %
-    \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
-             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma')\)}
 
@@
 
-\subsection{Provenance, Capabilities, and Protecting Objects}
+\section{Provenance, Capabilities, and Protecting Objects}
 \label{app:ptr}
@@
 \begin{figure*}
-  \begin{minipage}{0.5\textwidth}
-    \judgmentbr[Promote]
-               {\(\psi = \mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})\)}
+    \judgment[Promote]
              {\(\rho' = \rho[\reg_{dst} \mapsto \mathit{range} ~ \reg_{base} ~ \mathit{off} ~ \mathit{sz}]\)}
-               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+             {\(Op ~ \mach ~ (\mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})) ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
              %
     \judgment[Clear]
-             {\(\psi = \mathbf{clear} ~ \component\)}
-             {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho[\component \mapsto \emptyset])\)}
-  \end{minipage}
-  \begin{minipage}{0.5\textwidth}
-    \judgmentbr[Propagate]
-               {\(\psi = \mathbf{propagate} ~ \component_{src} ~ \component_{dst}\)}
+             {\(\rho' = \rho[\component \mapsto \emptyset]\)}
+             {\(Op ~ \mach ~ (\mathbf{clear} ~ \component) ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+             %
+    \judgment[Propagate]
              {\(\rho' = \rho[\component_{dst} \mapsto \rho[\component_{src}]]\)}
-               {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-               %
-    \[\begin{split}
-    & \mathit{capped} ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{\component\} \cup \mathit{capped} ~ \components' ~ \rho \textnormal{ where} \\
-    & \components' = \{\component' |\rho[\component] = (\mathit{base},\mathit{bound})
-    \land \mathit{base} \leq \component' < \mathit{bound}\} \\
-    \end{split}\]
-
-  \end{minipage}
-  \begin{minipage}{1\textwidth}
+             {\(Op ~ \mach ~ (\mathbf{propagate} ~ \component_{src} ~ \component_{dst}) ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
+               %
+
+    \[\mathit{reach} ~ \component ~ \rho \triangleq \{\component' | \mathit{base} \leq \component' < \mathit{bound}
+    \textnormal{ where } \rho[\component] = (\mathit{base},\mathit{bound})\}\]
+
+    \[\mathit{reach}^* ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{ \component \} \cup \mathit{reach}^* ~ (\mathit{reach} ~ \component ~ \rho) ~ \rho\]
+
     \vspace{\abovedisplayskip}
-    \judgment{\(V' = V \left\llbracket
-      k \mapsto
-      \begin{cases}
-        \unsealed & \textnormal{if } k \in CLR \\
-        \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
-        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } k \in \object(V) - (\mathit{passed} ~ \overline{sa} ~ \mach \cup \mathit{capped} ~ \overline{\reg_{args}} ~ \rho) \\
-        V(k) & \textnormal{otherwise} \\
-      \end{cases}\right\rrbracket\)}
-             {\(Op ~ \mach ~ \psi ~ (V,\sigma,\rho) = (V',V::\sigma,\rho)\)}
-
-  \end{minipage}
+    \judgmentthree[\sc Call]
+        {\(\components = \mathit{passed} ~ \overline{sa} ~ \cup ~ \overline{\reg_{args}}\)}
+        {\(\components' = \mathit{reach}^* ~ \components ~ \rho\)}
+        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ K'\)}
+        {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{\mathit{target}} ~ \overline{\reg_{args}} ~ \overline{sa}) ~ (V,\sigma,\rho) = (V',V::\sigma,\rho)\)}
+
+    \vspace{\abovedisplayskip}
+    \judgmentthree[\sc Tailcall]
+        {\(\components = \mathit{passed} ~ \overline{sa} ~ \cup ~ \overline{\reg_{args}}\)}
+        {\(\components' = \mathit{reach}^* ~ \components ~ \rho\)}
+        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ K'\)}
+        {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{\mathit{target}} ~ \overline{\reg_{args}} ~ \overline{sa}) ~ (V,\sigma,\rho) = (V',V::\sigma,\rho)\)}
 

### 2023-05-24 submission5/core.tex
Evolution of the Cerise line

@@
 %
-Originally, they describe ``local state encapsulation''~\cite{SkorstengaardSTKJFP}
+Originally, they define an informal notion of stack safety as the combination of WBCF
+and ``local state encapsulation''~\cite{SkorstengaardSTKJFP},
+and describe the latter
 in terms of integrity only (but it has confidentiality, equivalent
-to \(\clri\) {\em and} \(\clrc\)). Their latest
+to \(\clri\) {\em and} \(\clrc\)).
+%
+StkTokens \cite{SkorstengaardSTKJFP} makes this conception of stack safety
+explicit through an overlay semantics
+%(that internalizes the semantics of the calling convention and is proved fully
+%abstract w.r.t. the concrete machine executing the actual calling convention).
+%This overlay manipulates an abstract call stack whose elements are frames. A
+%function call creates a new stack frame off of a capability representing the
+%available stack space, and a function return merges the corresponding frame
+%back into the stack capability. It is assumed that components do not leave any
+%capabilities on the stack when they return.
+which (1) on call mints new a stack frame from a capability representing the
+available stack space, and (2) on return merges the current frame back into the
+stack capability, under the assumption that there are no capabilities left on
+the stack.
+
+Their latest
 paper \cite{Georges22:TempsDesCerises} was inspired by the properties presented
@@
 would in addition to building it into their formalism
-\cite{Georges22:personalcommunication}. This demonstrates the benefit of our
+\cite{Georges22:personalcommunication}.
+%
+To do so, they redesign the overlay semantics to actually pop stack frames on
+return and have them disappear from the stack.
+%
+This demonstrates the benefit of our
 choice to explicitly state properties in security terms: specifying security
@@
 requirement.
+
+% \cite{Georges22:TempsDesCerises} revisits this design and modifies the
+% overlay in an interesting way to actually pop stack frames on return and have
+% them disappear from the stack rather than be merged back into it. It thus
+% realizes callee confidentiality (and no dangling stack pointers).
 

### 2023-05-24 submission5/core.tex
Nit
Another grant

@@
 
-\paragraph*{Applicability to Roessler \& DeHon~\cite{DBLP:conf/sp/RoesslerD18}:}
+\paragraph*{Applicability to Roessler \& DeHon~\cite{DBLP:conf/sp/RoesslerD18}}
 
@@
 This work was supported by the National Science Foundation under Grant No. 2048499, Specifying and Verifying Secure Compilation of C Code to Tagged Hardware;
-by Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972;
+by ERC Starting Grant SECOMP (715753), Efficient Formally Secure Compilers to a Tagged Architecture;
+by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) as part of the Excellence Strategy of the German Federal and State Governments, EXC 2092 CASA -- 390781972;
 by NSF award \#2145649, CAREER: Fuzzing Formal Specifications,

### 2023-05-24 submission5/core.tex
Footnote personal communication

@@
 it did not enforce \(\clec\), and made sure that their new version
-would in addition to building it into their formalism
-\cite{Georges22:personalcommunication}.
+would in addition to building it into their formalism.\footnote{A. L. Georges,
+personal communication.}
+%\cite{Georges22:personalcommunication}.
 %

### 2023-05-24 submission5/core.tex
Benjamin's comments handled

--- 2023-05-24 submission5/core.tex
@@
 micro-policies~\cite{pump_oakland2015} for stack safety that aim to offer
-universal protection.  But the reasoning involved is somewhat subtle.
-Their most realistic micro-policy, called Lazy
-Tagging and Clearing ({\LTC}\bcp{does it have to be italic?}), makes an interesting performance
-trade-off: it allows function activations to write improperly into one
-another's stack frames, but ensures that the owner of the corrupted
-memory cannot access it afterward.
+universal protection.  But the reasoning involved can be subtle:
+they include micro-policy optimizations, Lazy Tagging and Lazy Clearing
+(likely to be deployed together, which we hereafter refer to as Lazy
+Tagging and Clearing, or LTC). LTC allows function activations to write
+improperly into one another's stack frames, but ensures that the owner of
+the corrupted memory cannot access it afterward, avoiding expensive
+clearings of the stack frame.
 %
@@
 memory---just not in ways that affect observable behavior.
-Therefore, {\LTC} would not fulfill Georges et al.'s property
-(adapted to the tagged setting).  But {\LTC} does
-arguably enforce stack safety.  A looser, more observational
+Therefore, LTC would not fulfill Georges et al.'s property
+(adapted to the tagged setting).  But LTC does arguably enforce stack safety,
+or as Roessler and DeHon describe it informally, a sort of data-flow
+integrity tied to the stack. A looser, more observational
 definition of stack safety is needed to fit this situation.
@@
 Our properties are stated abstractly in the hope that they can also be
-applied to other enforcement mechanisms besides \LTC.
+applied to other enforcement mechanisms besides LTC.
 %
@@
 which high-level features are supported in a given
-setting using the stack, and what are their
-security requirements? Different answers will beget different properties.
-\bcp{There's an awkward conceptual transition here: We pose a question
-(``which high-level features are supported in a given
-setting using the stack, and what are their
-security requirements'') but then we don't seem to be going ahead and
-answering it.  I suggest adding a sentence here that answers the
-question directly, then a paragraph break.} \bcp{Though, actually, the
-next sentence repeats points that have already been made above.  Maybe
-better to just delete it and find a new transition to the sentence
-after.}
-We identify and define five security requirements that a system might try
-to enforce upon a function call: well-bracketed control flow, and the
-integrity and confidentiality of each of the caller and the callee.
-%
-We apply these first to a simple system
-with very few features, then to a more realistic one supporting
+setting using the stack, and how does their presence influence
+the expectation of well-bracketed control flow, confidentiality,
+and integrity? We begin with a simple system
+with very few features, then move to a more realistic one supporting
 tail-call elimination, argument passing on the stack, and callee-save
-registers.
+registers. Our properties are factored so that the basic structure
+of each of our five properties remains constant while the presence or
+absence of different features leads to subtler differences in how
+they behave.
 
@@
 random testing tool for Coq.
-Indeed, we find that the published version of {\LTC} is flawed in
+Indeed, we find that the published version of LTC is flawed in
 a way that undermines both integrity and confidentiality; after
-correcting this flaw, {\LTC} satisfies all our
-properties.  Further, we modify {\LTC} to protect the features of
+correcting this flaw, LTC satisfies all of our
+properties.  Further, we modify LTC to protect the features of
 our more realistic system and apply
@@
 the extended properties.
-
-%We generate random programs, using the features of our
-%more realistic system, and check that the stack safety {\em micro-policies}
-%of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
-%correctly detect the ones that attempt to violate one of our properties.
-
-%Furthermore, we
-%check that the testing framework is able to generate counterexamples
-%that violate our properties but are \emph{not} halted by incorrect
-%variants of the enforcement mechanisms---including the original {\LTC}!
-% SNA: we talk about this in the testing section, and it's a "best practice" for
-% PBT, but not really a contribution in and of itself
-
-%\subsection{Contributions}
 
@@
 can access the element.
-The action of a security-relevant operation on the context is defined by a set
-of rules that characterize how particular machine code sequences are used
-to implement the function abstraction in terms of the stack and registers.\bcp{OK?}
+The action of a security-relevant operation on the context is defined by a function
+that characterizes how the operation's underling machine code ought to
+implement the function abstraction in terms of the stack and registers.
 
@@
 assume that the behavior of these mechanisms is incorporated into the basic
-step semantics of the machine, but we keep the enforcement state
-unchanged when constructing variant states\bcp{As a reader, I wonder
-  ``How would they do that?  Seems like they are saying the tag
-  enforcement or whatever is just part of the basic machine step
-  relation and we can't see it, but then going on to place a
-  requirement on it. How would such a requirement be phrased, formally?''}.
+step semantics of the machine, with a notion of ``compatible'' states that
+share security behavior that may be defined based on the enforcement mechanism.
 Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
@@
 \caption{Security-relevant operations and their parameters, with the
-  sections where
-  they are first used.  \bcp{Say what the gray levels mean.}}
+  sections where they are first defined or used. Entries in light grey do not appear in
+  our examples, but are part of our testing. Dark grey entries are not tested.}
 \label{tab:psi}
@@
 
-We do not assume that low-level code adheres to any fixed\bcp{single?
-(I think we do assume that the calling convention is fixed in any
-particular instance of our framework, no?)} calling
+We do not assume that low-level code adheres to any single calling
 convention or is being used to implement any particular
@@
 contain undefined behavior, in which case they might be compiled to
-arbitrary machine code. A given enforcement mechanism or target
-architecture might place additional constraints, particularly on
-call and return sequences.\bcp{Not sure that last sentence is helpful.}
+arbitrary machine code.
 
@@
 lower addresses. We assume that {\tt main} begins at address 0 and its
-callee {\tt f} at address 100.  \bcp{What are alloc(-20,20) and
-  dealloc(0,20)? What is $\varepsilon$ in call $\varepsilon$?  Also, readers who don't have the RISCV ISA at their
-fingertips may wonder what all the instructions actually do.  We
-should (if it's true) explicitly say that they don't need to worry
-about the details.}
+callee {\tt f} at address 100. The annotations in the right-hand column are
+security-relevant operations, described further below.
+The assembly is a simplified but otherwise typical compilation of the
+source code into RISC-V; its details are less important than the positions
+of the security-relevant operations.
 
@@
 Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
-{\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} ({\LTC}).
-Their precise connection to Roessler and DeHon's work is discussed below.
+{\em Depth Isolation} without lazy optimizations (DI) and with both
+Lazy Tagging and Lazy Clearing optimizations (LTC).
+(The connection between our properties and Roessler and DeHon's work is discussed below.)
 They share a common structure: each function activation is assigned a ``color'' \(n\)
@@
 
-In \(DI\), \(n\) always corresponds to the depth of the stack when
+In DI, \(n\) always corresponds to the depth of the stack when
 the function is called. A function must initialize its entire frame upon entry
@@
 
-In {\LTC}, a function neither initializes the frame at entry nor clears it at exist;
+In LTC, a function neither initializes the frame at entry nor clears it at exist;
 instead, it simply sets each location's tag to the PC tag when that location is written. It does
@@
 
-For Roessler and DeHon (henceforward \emph{R\&D}), Lazy Tagging and Lazy Clearing are both optimizations that
-can be applied to their Depth Isolation policy. Our version of {\LTC}
-corresponds to Depth Isolation with both optimizations applied. How closely
-do our properties correspond to the specifications of their policies?
-
+Roessler and DeHon (henceforward \emph{R\&D})
 R\&D differentiate between memory safety policies (without lazy optimization)
@@
 
-R\&D propose addressing the dangling-pointer issue by
+They propose addressing the dangling-pointer issue by
 tracking both the depth of the current activation and the static identity
@@
 In particular, \(\clri\) requires callee-saved
-registers to be saved and restored properly. We extend \(DI\) and {\LTC}
+registers to be saved and restored properly. We extend DI and LTC
 so that callee-saved registers are also tagged with the color of the
 function that is using them. In \(DI\) they are tagged as part of the entry
-sequence, while in {\LTC} they are tagged when a value is placed in them.
+sequence, while in LTC they are tagged when a value is placed in them.
 
@@
 \vspace*{1em}
-\caption{MTTF for finding bugs in erroneous \(DI\) (top) and {\LTC} (bottom) micro-policies}
+\caption{MTTF for finding bugs in erroneous micro-policies: DI (top) and LTC (bottom)}
 \vspace*{-2em}
@@
 
-For {\LTC}, the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
+For LTC, the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
 fails in testing, in cases where data is leaked between sequential calls.
@@
 %
-Originally, they define an informal notion of stack safety as the combination of WBCF
-and ``local state encapsulation''~\cite{SkorstengaardSTKJFP},
-and describe the latter
+Originally, they describe ``local state encapsulation''~\cite{SkorstengaardSTKJFP}
 in terms of integrity only (but it has confidentiality, equivalent
-to \(\clri\) {\em and} \(\clrc\)).
-%
-StkTokens \cite{SkorstengaardSTKJFP} makes this conception of stack safety
-explicit through an overlay semantics
-%(that internalizes the semantics of the calling convention and is proved fully
-%abstract w.r.t. the concrete machine executing the actual calling convention).
-%This overlay manipulates an abstract call stack whose elements are frames. A
-%function call creates a new stack frame off of a capability representing the
-%available stack space, and a function return merges the corresponding frame
-%back into the stack capability. It is assumed that components do not leave any
-%capabilities on the stack when they return.
-which (1) on call mints new a stack frame from a capability representing the
-available stack space, and (2) on return merges the current frame back into the
-stack capability, under the assumption that there are no capabilities left on
-the stack.
-
-Their latest
+to \(\clri\) {\em and} \(\clrc\)). Their latest
 paper \cite{Georges22:TempsDesCerises} was inspired by the properties presented
@@
 would in addition to building it into their formalism
-\cite{Georges22:personalcommunication}.
-%
-To do so, they redesign the overlay semantics to actually pop stack frames on
-return and have them disappear from the stack.
-%
-This demonstrates the benefit of our
+\cite{Georges22:personalcommunication}. This demonstrates the benefit of our
 choice to explicitly state properties in security terms: specifying security
@@
 requirement.
-
-% \cite{Georges22:TempsDesCerises} revisits this design and modifies the
-% overlay in an interesting way to actually pop stack frames on return and have
-% them disappear from the stack rather than be merged back into it. It thus
-% realizes callee confidentiality (and no dangling stack pointers).
 

### 2023-05-24 submission5/core.tex

@@
 it did not enforce \(\clec\), and made sure that their new version
-would in addition to building it into their formalism
-\cite{Georges22:personalcommunication}. This demonstrates the benefit of our
+would in addition to building it into their formalism.\footnote{A. L. Georges,
+personal communication.}
+%\cite{Georges22:personalcommunication}.
+%
+To do so, they redesign the overlay semantics to actually pop stack frames on
+return and have them disappear from the stack.
+%
+This demonstrates the benefit of our
 choice to explicitly state properties in security terms: specifying security

### 2023-05-24 submission5/core.tex
Added git repo

@@
 
-The accompanying artifact \apt{what is the state of that? Depending on conference preferences,
-  should give a URL or other pointer.} \sna{Git repo is currently private, I think. I didn't see much
-  in the way of conference preferences on a quick glance, but I'll look again in the morning.}
+The accompanying artifact \footnote{https://github.com/SNoAnd/stack-safety}
 contains formal definitions (in Coq) of our
@@
 
-\sna{Maybe a quick example or even two would be of benefit here.}
-For better performance, we instead maintain a single ``post-call taint''
-execution that executes in parallel with the original. Everytime a function returns
-with elements containing tainted data\apt{I don't understand what that is}, the test process sets those elements
-to random values in the tainted execution. So, at any given time, we need only simulate
+For better performance, we instead maintain a single execution that combines
+all of the variants that would be spawned at returns.
+So, at any given time, we need only simulate
 (1) the original execution, (2) the tainted execution, and (3) one

### 2023-05-24 submission5/core.tex
Remove "capped"

--- 2023-05-24 submission5/core.tex
@@
   \begin{subfigure}{0.4\textwidth}
-    \[\mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach \triangleq
-    \{\mach[\reg]+i | \mathit{off} \leq i < \mathit{off+sz}\}\]
-
     \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
@@
 a capability that has been passed to the callee. Reachability includes capabilities passed
-indirectly, by being stored in an object that is in turn passed. We call
-the set of reachable addresses \(\mathit{capped}\). The call operation in this setting
-will seal only objects that are neither in \(\mathit{capped}\) nor the previously defined \(\mathit{passed}\).
+indirectly, by being stored in an object that is in turn passed. We define
+the set of reachable addresses using \(\mathit{reach*}\), the transitive closure of elements
+that can be reached from the arguments of the call. The call operation in this setting
+will seal only objects that are not in \(\mathit{reach*}\) nor the previously defined \(\mathit{passed}\).
 

### 2023-05-24 submission5/core.tex
Small fixes

--- 2023-05-24 submission5/core.tex
@@
 When the call target is \((\mach,(V,\sigma))\), we define the predicate \(\intProp\) to hold
-on states \(\mach'\) if any elements that are both sealed in \(V\) and in the difference
+on states \(\mach'\) if all elements that are both sealed in \(V\) and in the difference
 set between \(\mach\) and \(\mach'\) are irrelevant. A system enjoys \(\clri\) if, for every
@@
                {\(V' = V \llbracket \addr \mapsto \object \mid \addr \in \components \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ (\mathit{off, sz})) ~ (V,\sigma) = (V',\sigma)\)}
 
@@
                {\(V' = V \llbracket \addr \mapsto \public \mid \addr \in \components \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-    %
+               {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{t} ~ (\mathit{off, sz})) ~ (V,\sigma) = (V',\sigma)\)}
+
     \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \object(V)\)}
                {\(V' = V \llbracket \addr \mapsto \unsealed | \addr \in \components \rrbracket\)}
-               {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
+               {\(Op ~ \mach ~ (\mathbf{dealloc} ~ (\mathit{off, sz})) ~ (V,\sigma) = (V',\sigma)\)}
 
@@
 so that callee-saved registers are also tagged with the color of the
-function that is using them. In \(DI\) they are tagged as part of the entry
+function that is using them. In DI they are tagged as part of the entry
 sequence, while in LTC they are tagged when a value is placed in them.

### 2023-05-24 submission5/core.tex
Clarification on capability calling conventions

--- 2023-05-24 submission5/core.tex
@@
 the stack.
+%
+The underlying unary logical relation does not capture confidentiality proper,
+although it does capture some of its facets.
 
@@
 paper \cite{Georges22:TempsDesCerises} was inspired by the properties presented
-in this paper to extend their formalism to include confidentiality. When
+in this paper to extend their formalism to include confidentiality through a binary logical relation. When
 checking if our properties applied to their old calling convention, they noted that

### 2023-05-24 submission5/core.tex
Proofreading

--- 2023-05-24 submission5/core.tex
@@
         \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
-        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } k \in \object(V') \\
+        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } k \in \object(V) \\
         V(k) & \textnormal{otherwise} \\
@@
         {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ K'\)}
-        {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{\mathit{target}} ~ \overline{\reg_{args}} ~ \overline{sa}) ~ (V,\sigma,\rho) = (V',V::\sigma,\rho)\)}
+        {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{\mathit{target}} ~ \overline{\reg_{args}} ~ \overline{sa}) ~ (V,\sigma,\rho) = (V',\sigma,\rho)\)}
 

### 2023-05-24 submission5/core.tex
MPI-SP branding policy
Updated test runs

@@
 \begin{tabular}{c|c|c|c}
-  Bug & Property Violated & MTTF (s) & Tests \\
+  Bug & Property Violated & Ave. MTTF (s) & Tests \\
   \hline
-      {\em LOAD\_NO\_CHECK\_DI}  & Confidentiality & 24.2 & 13.3 \\
+      {\em LOAD\_NO\_CHECK}  & Confidentiality & 24.2 & 13.3 \\
       {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
@@
   \hline
-      {\em PER\_DEPTH\_TAG} & Integrity & 189.7 & 8342.5  \\
-      {\em LOAD\_NO\_CHECK\_LT}  & Integrity & 23.5 & 12.0 \\
-      {\em LOAD\_NO\_CHECK\_LT}  & Confidentiality & 19.2 & 695.5 \\
-      {\em STORE\_NO\_UPDATE} & Integrity & 70 & 80.6  \\
-      {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
+      {\em PER\_DEPTH\_TAG} & Integrity & 10.5 & 82 \\
+      {\em PER\_DEPTH\_TAG} & Confidentiality & 16.85 & 88 \\
+      {\em LOAD\_NO\_CHECK}  & Integrity & 8.82 & 34.3 \\
+      {\em LOAD\_NO\_CHECK}  & Confidentiality & 22.55 & 127 \\
+      {\em STORE\_NO\_UPDATE} & Integrity & 6.96 & 101 \\
+      {\em STORE\_NO\_UPDATE} & Confidentiality & 17.34 & 11 \\
   \hline

### 2023-05-24 submission5/core.tex
Whitespace

@@
   \IEEEauthorblockA{
-    Max Planck Institute for Security and Privacy\\
+    MPI-SP\\
     roberto.blanco@mpi-sp.org\\
@@
 \begin{figure*}
-  \begin{subfigure}{0.4\textwidth}
-    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
+  \begin{subfigure}{0.48\textwidth}
+    \judgmentbr[~AllocF]
+        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
         {\(V' = V \llbracket \addr \mapsto \object \mid \addr \in \components \rrbracket\)}
@@
 
-    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
+    \vspace{\abovedisplayskip}
+
+    \judgmentbr[~AllocT]
+               {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
                  \unsealed(V)\)}
@@
 
-    \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \object(V)\)}
+    \vspace{\abovedisplayskip}
+
+    \judgmentbr[~Dealloc]
+        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \object(V)\)}
         {\(V' = V \llbracket \addr \mapsto \unsealed | \addr \in \components \rrbracket\)}
@@
   \end{subfigure}
-  \begin{subfigure}{0.6\textwidth}
+  \begin{subfigure}{0.52\textwidth}
     %
@@
     %
-    \judgmenttwo[\sc Call]
+    \judgmenttwo[~Call]
         {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
@@
     %
-    \judgmenttwo[\sc Tailcall]
+    \judgmenttwo[~Tailcall]
         {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
@@
 fails in testing, in cases where data is leaked between sequential calls.
-To round out our mutation testing we also check {\em LOAD\_NO\_CHECK\_LT},
+To round out our mutation testing we also check {\em LOAD\_NO\_CHECK},
 equivalent to its counterpart in depth isolation,
@@
 taking longer to catch.
-
-%% I don't think this discussion is relevant here
-%% Unfortunately, Lazy Per-activation Tagging and Clearing
-%% puts severe pressure on the tag cache, so it may not be usable in practice.
-%% However, we conjecture that no lazy scheme can be fully stack safe
-%% without using unique per-activation identifiers.
 

### 2023-05-24 submission5/core.tex
Ready to go!

@@
 
-  \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
+    \judgmentbr[~Alloc]
+               {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
                {\(V' = V \llbracket \addr \mapsto \object \mid \addr \in \components \rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-    %
-    \judgmentbr{\(\components = \mathit{range}~ \SP ~ \mathit{off} ~ \mathit {sz} ~ \mach \cap \object(V)\)}
+
+    \vspace{\abovedisplayskip}
+
+    \judgmentbr[~Dealloc]
+               {\(\components = \mathit{range}~ \SP ~ \mathit{off} ~ \mathit {sz} ~ \mach \cap \object(V)\)}
                {\(V' = V \llbracket \addr \mapsto \unsealed \mid \addr \in \components \rrbracket\)}
                {\(Op ~ \mach ~ (\mathbf{dealloc} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
-    %
-    \judgment{\(V' = \lambda \component .
+
+    \vspace{\abovedisplayskip}
+
+    \judgment[~Call]
+             {\(V' = \lambda \component .
                \begin{cases}
@@
                ~ (V,\sigma) = (V',V::\sigma)\)}
-    %
-    \judgment{}
+
+    \vspace{\abovedisplayskip}
+
+    \judgment[~Return]
+             {}
              {\(Op ~ \mach ~ \mathbf{return} ~ (\_, (V,\sigma')) = (V, \sigma')\)}
@@
 
-\judgment{\(|\sigma| < d\)}
+\judgment[~Done]
+         {\(|\sigma| < d\)}
          {\(d \hookrightarrow (\mach,(V,\sigma)) = \tau\)}
 
-\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{\overline{\psi},\obs} (\mach',\context')\)}
+\vspace{\abovedisplayskip}
+
+\judgmenttwobrlong[~Step]
                   {\(|\sigma| \geq d\)}
                   {\(d \hookrightarrow (\mach',\context') = \obsT\)}
+                  {\((\mach,(V,\sigma)) \stepstounder{\overline{\psi},\obs} (\mach',\context')\)}
                   {\(d \hookrightarrow (\mach,(V,\sigma)) = \notfinished{\obs}{\obsT}\)}
@@
 \begin{minipage}{.4\columnwidth}
-  \judgment{}{\(\obsT \eqsim \obsT\)}
+  \judgment[~SimRefl]{}{\(\obsT \eqsim \obsT\)}
 \end{minipage}
 \begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+  \judgment[~SimEvent]
+           {\(\obsT_1 \eqsim \obsT_2\)}
            {\(\notfinished{\obs}{\obsT_1} \eqsim \notfinished{\obs}{\obsT_2}\)}
@@
 \begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+  \judgment[~SimLeft]
+           {\(\obsT_1 \eqsim \obsT_2\)}
            {\(\notfinished{\tau}{\obsT_1} \eqsim \obsT_2\)}
 \end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \judgment{\(\obsT_1 \eqsim \obsT_2\)}
+\begin{minipage}{.5\columnwidth}
+  \judgment[~SimRight]
+           {\(\obsT_1 \eqsim \obsT_2\)}
            {\(\obsT_1 \eqsim \notfinished{\tau}{\obsT_2}\)}
@@
 
-\judgmenttwo[Returned]
+\vspace{\abovedisplayskip}
+
+\judgmenttwo[~Returned]
             {\(|\sigma| < d\)}
@@
 
-\judgmenttwobrlong[Step]
+\vspace{\abovedisplayskip}
+
+\judgmenttwobrlong[~Step]
                   {\(|\sigma| \geq d\)}
@@
 
-\judgmentthree[Returned]
+\vspace{\abovedisplayskip}
+\judgmentthree[~Returned]
               {\(|\sigma_1| < d\)}
@@
 
-\judgmenttwobrlong[Left]
+\judgmenttwobrlong[~Left]
                   {\(|\sigma_1| \geq d\)}
@@
 
-\judgmenttwobrlong[Right]
+\judgmenttwobrlong[~Right]
                   {\(|\sigma_2| \geq d\)}
@@
   \begin{subfigure}{0.52\textwidth}
-    %
+
     \[\mathit{push} ~ V ~ \overline{\reg} ~ K \triangleq \lambda k .
@@
     \]
-    %
+
     \[\mathit{passed} ~ \overline{sa} ~ \mach \triangleq \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
     \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
-    %
+
     \judgmenttwo[~Call]
@@
           ~ (V,\sigma) = (V',V::\sigma)\)}
-    %
+
+    \vspace{\abovedisplayskip}
+
     \judgmenttwo[~Tailcall]
@@
              {\(Op ~ \mach ~ (\mathbf{promote} ~ \reg_{dst} ~ (\reg_{base},\mathit{off},\mathit{sz})) ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-             %
+
+    \vspace{\abovedisplayskip}
+
     \judgment[Clear]
@@
              {\(Op ~ \mach ~ (\mathbf{clear} ~ \component) ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-             %
+
+    \vspace{\abovedisplayskip}
+
     \judgment[Propagate]
@@
              {\(Op ~ \mach ~ (\mathbf{propagate} ~ \component_{src} ~ \component_{dst}) ~ (V,\sigma,\rho) = (V,\sigma,\rho')\)}
-               %
+
+    \vspace{\abovedisplayskip}
 
@@
     \[\mathit{reach}^* ~ \components ~ \rho \triangleq \bigcup_{\component \in \components} \{ \component \} \cup \mathit{reach}^* ~ (\mathit{reach} ~ \component ~ \rho) ~ \rho\]
+    \vspace{\abovedisplayskip}
 

### 2023-05-25 submission5/core.tex
Last minute nits
Fixed typo

@@
 
-\begin{figure*}
-  \begin{subfigure}{0.48\textwidth}
+\begin{figure*}[h]
+  \begin{subfigure}{\textwidth}
     \judgmentbr[~AllocF]
@@
     \judgmentbr[~Dealloc]
-        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \object(V)\)}
+        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap (\object(V) \cup \public(V))\)}
         {\(V' = V \llbracket \addr \mapsto \unsealed | \addr \in \components \rrbracket\)}
@@
   \end{subfigure}
-  \begin{subfigure}{0.52\textwidth}
-
+  \begin{subfigure}{\textwidth}
+
+    \begin{minipage}{0.4\textwidth}
       \[\mathit{push} ~ V ~ \overline{\reg} ~ K \triangleq \lambda k .
@@
       \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
-
+    \end{minipage}
+    \begin{minipage}{0.6\textwidth}    
       \judgmenttwo[~Call]
@@
           ~ (V,\sigma) = (V',\sigma)\)}
+    \end{minipage}
 
@@
 that restriction by omitting the check (bug {\em
-  LOAD\_NO\_CHECK\_DI}).
+  LOAD\_NO\_CHECK}).
 %
@@
 our set of mutations as {\em HEADER\_NO\_INIT}.
+
+For LTC, the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
+fails in testing, in cases where data is leaked between sequential calls.
+To round out our mutation testing we also check {\em LOAD\_NO\_CHECK},
+equivalent to its counterpart in depth isolation,
+and a version where stores succeed but fails to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
+
+The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
+\cref{tab:bug-table}, along with the average number of tests
+it took to find the failure. Experiments were run in a desktop
+machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
 %(the original was written in C++
@@
 
-\subsection{Results}
-
-The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
-\cref{tab:bug-table}, along with the average number of tests
-it took to find the failure. Experiments were run in a desktop
-machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
-
-For LTC, the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
-fails in testing, in cases where data is leaked between sequential calls.
-To round out our mutation testing we also check {\em LOAD\_NO\_CHECK},
-equivalent to its counterpart in depth isolation,
-and a version where stores succeed but fails to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
-It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
-taking longer to catch.
-
 \section{Related Work}
